### 2024-01-09 appendix.qmd
Initial commit

---
title: "Appendix: Visualising How Non-linear Dimension Reduction Warps Your Data"
author:
  - name: Jayani P.G. Lakshika
    affiliations:
      - name: Monash University
        department: Econometrics & Business Statistics
        address: Clayton
        city: VIC 
        country: Austria
        postal-code: 3800
    orcid: 0000-0002-6265-6481
    email: jayani.piyadigamage@monash.edu
    url: https://jayanilakshika.netlify.app/
  - name: Dianne Cook
    affiliations:
      - name: Monash University
        department: Econometrics & Business Statistics
        address: Clayton
        city: VIC 
        country: Austria
        postal-code: 3800
    orcid: 0000-0002-3813-7155
    email: dicook@monash.edu 
    url: http://www.dicook.org/
  - name: Paul Harrison
    affiliations:
      - name: Monash University
        department: MGBP, BDInstitute
        address: Clayton
        city: VIC 
        country: Austria
        postal-code: 3800
    orcid: 0000-0002-3980-268X
    email: 	paul.harrison@monash.edu
    url: 
  - name: Michael Lydeamore
    affiliations:
      - name: Monash University
        department: Econometrics & Business Statistics
        address: Clayton
        city: VIC 
        country: Austria
        postal-code: 3800
    orcid: 0000-0001-6515-827X
    email: michael.lydeamore@monash.edu
    url: 
  - name: Thiyanga S. Talagala
    affiliations:
      - name: University of Sri Jayewardenepura
        department: Statistics
        address: Gangodawila
        city: Nugegoda 
        country: Sri Lanka
        postal-code: 10100
    orcid: 0000-0002-0656-9789
    email: ttalagala@sjp.ac.lk 
    url: https://thiyanga.netlify.app/
tbl-cap-location: bottom
---

```{r}
#| warning: false
#| echo: false
library(dplyr)
# remotes::install_github("jlmelville/snedata")
library(snedata)
library(ggflowchart)
library(purrr) ## map function
library(gridExtra) ## for grid.arrange
library(rsample)
library(DT)
library(ggbeeswarm)
library(ggplot2)
library(readr)

library(Rtsne)
library(umap)
library(phateR)
library(reticulate)
library(patchwork)

library(grid)

# install.packages("Seurat")
# remotes::install_github("satijalab/seurat-data")
#library(SeuratData) ## For the application

# use_python("~/miniforge3/envs/pcamp_env/bin/python")
# use_condaenv("pcamp_env")
# 
# reticulate::source_python(paste0(here::here(), "/examples_for_the_paper/function_scripts/Fit_PacMAP_code.py"))
# reticulate::source_python(paste0(here::here(), "/examples_for_the_paper/function_scripts/Fit_TriMAP_code.py"))

set.seed(20230531)

source("quollr_code.R", local = TRUE)
source("nldr_code.R", local = TRUE)
```

```{r}
#| warning: false
#| echo: false

data <- read_csv(paste0(here::here(), "/data/s_curve.csv"))
```

```{r}
#| warning: false
#| echo: false

data <- data |>
  mutate(ID = row_number())

data_split <- initial_split(data)
training_data <- training(data_split)
test_data <- testing(data_split)
```

```{r}
#| echo: false

UMAP_fit <- umap(training_data |> dplyr::select(-ID), n_neighbors = 50, n_components =  2)

UMAP_data <- UMAP_fit$layout |>
  as.data.frame()
names(UMAP_data)[1:(ncol(UMAP_data))] <- paste0(rep("UMAP",(ncol(UMAP_data))), 1:(ncol(UMAP_data)))

UMAP_data <- UMAP_data |>
  mutate(ID = training_data$ID)

num_bins_x <- calculate_effective_x_bins(.data = UMAP_data, x = UMAP1,
                           cell_area = 1)

shape_value <- calculate_effective_shape_value(.data = UMAP_data, 
                                             x = UMAP1, y = UMAP2)

## To extract bin centroids
hexbin_data_object <-extract_hexbin_centroids(nldr_df = UMAP_data, num_bins = num_bins_x, shape_val = shape_value)

df_bin_centroids <- hexbin_data_object$hexdf_data

##########

## Data set with all possible centroids in the hexagonal grid

full_centroid_df <- generate_full_grid_centroids(df_bin_centroids)

## To map hexID to hexbin centroids in the full grid

vec1 <- stats::setNames(rep("", 2), c("x", "y"))  ## Define column names

full_grid_with_hexbin_id <- dplyr::bind_rows(vec1)[0, ]
full_grid_with_hexbin_id <- full_grid_with_hexbin_id |>
  dplyr::mutate_if(is.character, as.numeric)

for(i in 1:length(sort(unique(full_centroid_df$y)))){

  ## Filter the data set with specific y value
  specific_y_val_df <- full_centroid_df |>
    dplyr::filter(y == sort(unique(full_centroid_df$y))[i])

  ordered_x_df <- specific_y_val_df |>
    dplyr::arrange(x)

  full_grid_with_hexbin_id <- dplyr::bind_rows(full_grid_with_hexbin_id, ordered_x_df)

}

full_grid_with_hexbin_id <- full_grid_with_hexbin_id |>
  dplyr::mutate(hexID = row_number())

full_grid_with_hexbin_id <- full_grid_with_hexbin_id |>
  dplyr::rename("c_x" = "x",
         "c_y" = "y")

full_grid_with_hexbin_id <- dplyr::full_join(full_grid_with_hexbin_id, df_bin_centroids, by = c("hexID" = "hexID")) |>
  dplyr::select(-c(x, y)) #|> 
  #dplyr::mutate(counts = tidyr::replace_na(counts, 0))

full_grid_with_hexbin_id <- full_grid_with_hexbin_id |>
    dplyr::mutate(std_counts = counts/max(counts, na.rm = TRUE))

## Generate all coordinates of hexagons
hex_grid <- full_hex_grid(full_centroid_df)

full_grid_with_polygon_id_df <- map_polygon_id(full_grid_with_hexbin_id, hex_grid)

full_grid_with_hexbin_id_rep <- full_grid_with_polygon_id_df |>
  dplyr::slice(rep(1:n(), each = 6)) |>
  dplyr::arrange(polygon_id)

hex_full_count_df <- dplyr::bind_cols(hex_grid, full_grid_with_hexbin_id_rep)

##########

min_std_cell_threshold <- 0.05

df_bin_centroids_all <- hexbin_data_object$hexdf_data ## All the centroids without removing low-density hexagons

df_bin_centroids <- df_bin_centroids |>
  dplyr::mutate(stand_cell_count = counts/max(counts)) |>
  dplyr::filter(stand_cell_count > min_std_cell_threshold)

UMAP_data_with_hb_id <- UMAP_data |> 
  dplyr::mutate(hb_id = hexbin_data_object$hb_data@cID)

## To generate a data set with high-D and 2D training data
df_all <- dplyr::bind_cols(training_data |> dplyr::select(-ID), UMAP_data_with_hb_id)

## Averaged on high-D
df_bin <- avg_highD_data(.data = df_all)

## Triangulate bin centroids
tr1_object <- triangulate_bin_centroids(df_bin_centroids, x, y)
tr_from_to_df <- generate_edge_info(triangular_object = tr1_object)

## Compute 2D distances
distance <- cal_2D_dist(.data = tr_from_to_df)

## To find the benchmark value
benchmark <- find_benchmark_value(.data = distance, distance_col = distance)

trimesh <- ggplot(df_bin_centroids, aes(x = x, y = y)) + 
    geom_segment(data = tr_from_to_df, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
    geom_point(size = 2, colour = "#33a02c") +
    coord_equal()

  # ggplot(df_bin_centroids, aes(x = x, y = y)) + 
  # geom_point(size = 1, colour = "#33a02c") + 
  # geom_trimesh() + 
  # coord_equal()

trimesh <- trimesh +
  #ggtitle("(a)") + 
  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  theme_light() +
    theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
          axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
    ) 
  # theme(axis.text = element_text(size = 5),
  #       axis.title = element_text(size = 7))

trimesh_gr <- colour_long_edges(.data = distance, benchmark_value = benchmark, 
                                triangular_object = tr1_object, distance_col = distance)

trimesh_gr <- trimesh_gr + 
    geom_point(size = 2, colour = "#33a02c") + 
  #ggtitle("(b)") + 
  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  theme_light() +
    #coord_equal() +
    theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
          axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
    )

trimesh_removed <- remove_long_edges(.data = distance, benchmark_value = benchmark, 
                                triangular_object = tr1_object, distance_col = distance)

trimesh_removed <- trimesh_removed + 
       geom_point(size = 2, colour = "#33a02c") + 
  #ggtitle("(b)") + 
  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  theme_light() +
    #coord_equal() +
    theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
          axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
    )

tour1 <- show_langevitour(df_all, df_bin, df_bin_centroids, benchmark_value = benchmark, distance = distance, distance_col = distance)
```

```{r}
#| echo: false

## To plot the distribution of distance
plot_dist <- function(distance_df){
  distance_df$group <- "1"
  dist_plot <- ggplot(distance_df, aes(x = group, y = distance)) +
    geom_quasirandom()+
    ylim(0, max(unlist(distance_df$distance))+ 0.5) + coord_flip()
  return(dist_plot)
}
```

```{r}
#| echo: false

distance_plot <- plot_dist(distance) +
  #ggtitle("(b)" ) + 
  ylab(expression(d^{(2)})) +
  theme(axis.text = element_text(size = 5),
        axis.title = element_text(size = 12))
```

## Paramter tunning

### Total number of bins

The total number of bins represents the overall count of hexagonal bins within the hexagonal grid. This count is determined by multiplying the number of bins along the x-axis ($b_1$) with the number of bins along the y-axis ($b_2$), according to the formula:

$$
 b = b_1 \times b_2
$$ {#eq-equationp1}

Here, $b$ denotes the total number of bins. By adjusting the parameter $b_1$, we have control over the total number of bins ($b$). Hence, fine-tuning $b_1$ enables us to customize and optimize the total bin count based on the desired configuration along the x-axis.

```{r}
#| echo: false
#| message: false
#| warning: false

hexgrid_with_points_hex <- ggplot(data = hex_full_count_df |> dplyr::filter(!is.na(std_counts)), aes(x = c_x, y = c_y, fill = std_counts, hexID = hexID)) + 
    geom_hex(stat = "identity", color = "#969696") +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
  theme_light() +
    #coord_equal() +
    theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
          axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
    )

```

```{r}
#| echo: false
#| message: false
#| warning: false

hexgrid_with_points_m <- ggplot(data = hex_full_count_df, aes(x = x, y = y)) + 
  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) + 
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
  theme_light() +
    #coord_equal() +
    theme(plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
          axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
    )

hexgrid_with_points <- hexgrid_with_points_m + theme(legend.position = "none")
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| out-width: 100%

hexgrid_with_points_hex + hexgrid_with_points +
  plot_layout(ncol = 3)

```
<!--
```{r}
#| echo: false
#| message: false
#| warning: false
 # Loop over loop.vector
  #i <- 1
  num_bins <- 3
  shape_val <- calculate_effective_shape_value(.data = UMAP_data, x = UMAP1, y = UMAP2)

  hexbin_data_object_loop <- extract_hexbin_centroids(UMAP_data, num_bins, shape_val)

  df_bin_centroids_loop <- hexbin_data_object_loop$hexdf_data

  ## Data set with all possible centroids in the hexagonal grid

  full_centroid_df_loop <- generate_full_grid_centroids(df_bin_centroids_loop)

  ## To map hexID to hexbin centroids in the full grid

  vec1 <- stats::setNames(rep("", 2), c("x", "y"))  ## Define column names

  full_grid_with_hexbin_id_loop <- dplyr::bind_rows(vec1)[0, ]
  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
    dplyr::mutate_if(is.character, as.numeric)

  for(i in 1:length(sort(unique(full_centroid_df_loop$y)))){

    ## Filter the data set with specific y value
    specific_y_val_df_loop <- full_centroid_df_loop |>
      dplyr::filter(y == sort(unique(full_centroid_df_loop$y))[i])

    ordered_x_df_loop <- specific_y_val_df_loop |>
      dplyr::arrange(x)

    full_grid_with_hexbin_id_loop <- dplyr::bind_rows(full_grid_with_hexbin_id_loop, ordered_x_df_loop)

  }

  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
    dplyr::mutate(hexID = row_number())

  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
    dplyr::rename("c_x" = "x",
           "c_y" = "y")

  full_grid_with_hexbin_id_loop <- dplyr::full_join(full_grid_with_hexbin_id_loop, df_bin_centroids_loop, by = c("hexID" = "hexID")) |>
    dplyr::select(-c(x, y))

  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
      dplyr::mutate(std_counts = counts/max(counts, na.rm = TRUE))

  ## Generate all coordinates of hexagons
  hex_grid_loop <- full_hex_grid(full_centroid_df_loop)

  full_grid_with_polygon_id_df_loop <- map_polygon_id(full_grid_with_hexbin_id_loop, hex_grid_loop)

  full_grid_with_hexbin_id_rep_loop <- full_grid_with_polygon_id_df_loop |>
  dplyr::slice(rep(1:n(), each = 6)) |>
  dplyr::arrange(polygon_id)

  hex_full_count_df_loop <- dplyr::bind_cols(hex_grid_loop, full_grid_with_hexbin_id_rep_loop)

  p1 <-  ggplot(data = hex_full_count_df_loop, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
        geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
    theme_void() +
    theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
          axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
          legend.title = element_text(size=8), #change legend title font size
          legend.text = element_text(size=6)) +
    guides(fill = guide_colourbar(title = "Standardized count")) +
    annotate(geom = 'text', label = "a", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)

```

```{r}
#| echo: false
#| message: false
#| warning: false
 # Loop over loop.vector
  #i <- 1
  num_bins <- 8
  shape_val <- calculate_effective_shape_value(.data = UMAP_data, x = UMAP1, y = UMAP2)

  hexbin_data_object_loop <- extract_hexbin_centroids(UMAP_data, num_bins, shape_val)

  df_bin_centroids_loop <- hexbin_data_object_loop$hexdf_data

  ## Data set with all possible centroids in the hexagonal grid

  full_centroid_df_loop <- generate_full_grid_centroids(df_bin_centroids_loop)

  ## To map hexID to hexbin centroids in the full grid

  vec1 <- stats::setNames(rep("", 2), c("x", "y"))  ## Define column names

  full_grid_with_hexbin_id_loop <- dplyr::bind_rows(vec1)[0, ]
  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
    dplyr::mutate_if(is.character, as.numeric)

  for(i in 1:length(sort(unique(full_centroid_df_loop$y)))){

    ## Filter the data set with specific y value
    specific_y_val_df_loop <- full_centroid_df_loop |>
      dplyr::filter(y == sort(unique(full_centroid_df_loop$y))[i])

    ordered_x_df_loop <- specific_y_val_df_loop |>
      dplyr::arrange(x)

    full_grid_with_hexbin_id_loop <- dplyr::bind_rows(full_grid_with_hexbin_id_loop, ordered_x_df_loop)

  }

  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
    dplyr::mutate(hexID = row_number())

  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
    dplyr::rename("c_x" = "x",
           "c_y" = "y")

  full_grid_with_hexbin_id_loop <- dplyr::full_join(full_grid_with_hexbin_id_loop, df_bin_centroids_loop, by = c("hexID" = "hexID")) |>
    dplyr::select(-c(x, y))

  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
      dplyr::mutate(std_counts = counts/max(counts, na.rm = TRUE))

  ## Generate all coordinates of hexagons
  hex_grid_loop <- full_hex_grid(full_centroid_df_loop)

  full_grid_with_polygon_id_df_loop <- map_polygon_id(full_grid_with_hexbin_id_loop, hex_grid_loop)

  full_grid_with_hexbin_id_rep_loop <- full_grid_with_polygon_id_df_loop |>
  dplyr::slice(rep(1:n(), each = 6)) |>
  dplyr::arrange(polygon_id)

  hex_full_count_df_loop <- dplyr::bind_cols(hex_grid_loop, full_grid_with_hexbin_id_rep_loop)

  p2 <-  ggplot(data = hex_full_count_df_loop, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
    geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
        theme_void() +
    theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
          axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
          legend.title = element_text(size=8), #change legend title font size
          legend.text = element_text(size=6)) +
    guides(fill = guide_colourbar(title = "Standardized count")) +
    annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)

```

```{r}
#| echo: false
#| message: false
#| warning: false
 # Loop over loop.vector
  #i <- 1
  num_bins <- 15
  shape_val <- calculate_effective_shape_value(.data = UMAP_data, x = UMAP1, y = UMAP2)

  hexbin_data_object_loop <- extract_hexbin_centroids(UMAP_data, num_bins, shape_val)

  df_bin_centroids_loop <- hexbin_data_object_loop$hexdf_data

  ## Data set with all possible centroids in the hexagonal grid

  full_centroid_df_loop <- generate_full_grid_centroids(df_bin_centroids_loop)

  ## To map hexID to hexbin centroids in the full grid

  vec1 <- stats::setNames(rep("", 2), c("x", "y"))  ## Define column names

  full_grid_with_hexbin_id_loop <- dplyr::bind_rows(vec1)[0, ]
  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
    dplyr::mutate_if(is.character, as.numeric)

  for(i in 1:length(sort(unique(full_centroid_df_loop$y)))){

    ## Filter the data set with specific y value
    specific_y_val_df_loop <- full_centroid_df_loop |>
      dplyr::filter(y == sort(unique(full_centroid_df_loop$y))[i])

    ordered_x_df_loop <- specific_y_val_df_loop |>
      dplyr::arrange(x)

    full_grid_with_hexbin_id_loop <- dplyr::bind_rows(full_grid_with_hexbin_id_loop, ordered_x_df_loop)

  }

  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
    dplyr::mutate(hexID = row_number())

  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
    dplyr::rename("c_x" = "x",
           "c_y" = "y")

  full_grid_with_hexbin_id_loop <- dplyr::full_join(full_grid_with_hexbin_id_loop, df_bin_centroids_loop, by = c("hexID" = "hexID")) |>
    dplyr::select(-c(x, y))

  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
      dplyr::mutate(std_counts = counts/max(counts, na.rm = TRUE))

  ## Generate all coordinates of hexagons
  hex_grid_loop <- full_hex_grid(full_centroid_df_loop)

  full_grid_with_polygon_id_df_loop <- map_polygon_id(full_grid_with_hexbin_id_loop, hex_grid_loop)

  full_grid_with_hexbin_id_rep_loop <- full_grid_with_polygon_id_df_loop |>
  dplyr::slice(rep(1:n(), each = 6)) |>
  dplyr::arrange(polygon_id)

  hex_full_count_df_loop <- dplyr::bind_cols(hex_grid_loop, full_grid_with_hexbin_id_rep_loop)

  p3 <-  ggplot(data = hex_full_count_df_loop, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
        geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
        theme_void() +
    theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
          axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
          legend.title = element_text(size=8), #change legend title font size
          legend.text = element_text(size=6)) +
    guides(fill = guide_colourbar(title = "Standardized count")) +
    annotate(geom = 'text', label = "c", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)

```

```{r}
#| echo: false
#| label: fig-binsize
#| fig-pos: H
#| fig-cap: "Hexbin plots from different number of bins for the same **s_curve_noise_umap** data: (a) b = (4, 8), s = 2.031, (b) b = (9, 20), s = 2.031, and (c) b = (16, 36), s = 2.031. The hexbins are colored based on the density of points, with yellow indicating higher point density and darker colors representing lower point density within each bin. Does  a value of number of bins exist to effectively represent the low-dimensional data?"

p1 + p2 + p3
```
-->

### Benchmark value to remove low-density hexagons

In the step of addressing low-density hexagons, which can arise from sparsely represented data in certain regions, we employ a systematic strategy. The goal is to ensure a more comprehensive coverage of the data by removing hexagons with low data density. To achieve this, we initiate the process by identifying, for each hex bin, the six nearest hex bins based on an equal 2D distance metric. Following this, we calculate the mean density (see @eq-equationp3), as outlined in the equation:

$$
 \text{standard count} = \frac{\text{count}}{\text{max count}}
$$ {#eq-equationp2}

$$
 \text{mean density} = \frac{\text{standard count}}{6}
$$ {#eq-equationp3}

The standard count is derived from the number of observations in the hex bins (see @eq-equationp2). Next, we examine the distribution of mean densities across all hex bins and designate the first quartile as the benchmark value for removing low-density hexagons. Finally, hex bins with mean densities below this benchmark value are removed from consideration. This meticulous procedure ensures the elimination of regions with inadequate data density, allowing the focus to shift to areas with more significant data representation. The result is the preservation of the overall structure of the data in the low-dimensional space, as illustrated in @fig-bintorm (c).

```{r}
#| echo: false

df_bin_centroids <- df_bin_centroids |>
  dplyr::mutate(ID = row_number())

# To store mean densities of hexagons
mean_density_vec <- c()

for (i in 1:length(df_bin_centroids$hexID)) {

  df_bin_centroids_coordinates_spec_bin <- df_bin_centroids |>
    filter(hexID == df_bin_centroids$hexID[i])

  available_near_check <- df_bin_centroids |>
    dplyr::filter((hexID == (df_bin_centroids$hexID[i] + 1)) | (hexID == (df_bin_centroids$hexID[i] - 1))) |>
    head(1)

  if (NROW(available_near_check) == 0) {

    df_bin_centroids_coordinates_spec_bin_near1 <- df_bin_centroids |>
      filter((hexID == (df_bin_centroids$hexID[i] + (num_bins_x + 1))) | (hexID == (df_bin_centroids$hexID[i] + num_bins_x)) | (hexID == (df_bin_centroids$hexID[i] - (num_bins_x + 1))) | (hexID == (df_bin_centroids$hexID[i] - num_bins_x))) |>
      head(1)

  } else {

    df_bin_centroids_coordinates_spec_bin_near1 <- df_bin_centroids |>
      filter((hexID == (df_bin_centroids$hexID[i] + 1)) | (hexID == (df_bin_centroids$hexID[i] - 1))) |>
      head(1)

  }

  near_df_1 <- dplyr::bind_rows(df_bin_centroids_coordinates_spec_bin, df_bin_centroids_coordinates_spec_bin_near1)

  start <- unlist(near_df_1[1, c("x","y")])
  end <- unlist(near_df_1[2, c("x","y")])
  nearest_dist <- sqrt(sum((start - end)^2)) # Distance to nearest centroid

  df_bin_centroids$distance <- lapply(seq(nrow(df_bin_centroids)), function(x) {
    start <- unlist(df_bin_centroids[(df_bin_centroids_coordinates_spec_bin |> pull(ID)), c("x","y")])
    end <- unlist(df_bin_centroids[x, c("x","y")])
    sqrt(sum((start - end)^2))})

  df_bin_centroids <- df_bin_centroids %>%
    dplyr::select(names(df_bin_centroids), "distance")

  df_bin_centroids$distance <- round(unlist(df_bin_centroids$distance), 7)

  neighbor_df <- df_bin_centroids |>
    filter(distance == round(nearest_dist, 7))

  mean_density <- neighbor_df |>
    pull(std_counts) |>
    sum()/6 ## The reason to take the mean is to check the density in a considerable amount

  mean_density_vec <- append(mean_density_vec, mean_density)

}

df_bin_centroids <- df_bin_centroids |>
  dplyr::mutate(mean_density = mean_density_vec)

cell_count_plot <- ggplot(df_bin_centroids, aes(x = reorder(as.factor(hexID), -mean_density), y = mean_density)) +
  geom_quasirandom() + xlab("hexagonal id") + ylab("Standardized cell count") +
  geom_hline(yintercept = 0.255, colour = "#de2d26") +
  theme(axis.text = element_text(size = 5),
        axis.title = element_text(size = 7),
        axis.text.x = element_text(angle = 90))

cell_count_plot
```

```{r}
#| echo: false

hexgrid_with_points <- hexgrid_with_points +
  coord_equal()

hex_full_count_df <- full_join(hex_full_count_df, df_bin_centroids |> dplyr::select(hexID, mean_density))

hex_full_count_df <- hex_full_count_df |>
  dplyr::mutate(type = if_else(mean_density < 0.255, "low", "high"))

hex_full_count_df_filtered <- hex_full_count_df |>
  dplyr::filter(type == "low")

hex_grid_centers_rm_l_n <- ggplot(data = hex_full_count_df, aes(x = x, y = y)) + 
  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts))   +
  # geom_polygon(data = hex_full_count_df_filtered, color = "black", aes(group = polygon_id), fill = "red") +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
  theme_light() +
    #coord_equal() +
    theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
          axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
    )

hex_grid_centers_rm_l_n <- hex_grid_centers_rm_l_n +
  geom_text(data = hex_full_count_df_filtered, color = "black", aes(x = c_x, y = c_y, group = polygon_id, label = "X", size = 0.5)) +
  coord_equal()

hexgrid_with_points + hex_grid_centers_rm_l_n +
  plot_layout(ncol = 2)

```

### Benchmark value to remove long edges

In this step, we aim to create a smoother surface in the low-dimensional space by removing long edges from the triangular mesh. This process helps eliminate outliers and noise while preserving important local relationships and intricate structures within the data. To achieve this, distances between vertices are sorted, and unique distance values are extracted. A data frame is created to calculate the differences between consecutive distance values, and the largest difference serves as a threshold to identify long edges. By removing edges that exceed this threshold, the algorithm refines the triangular mesh (see @fig-traingularmeshgr).

![A flow diagram detailing the steps taken to find the benchmark value to remove long edges.](figures/remove_long_edges_workflow.png){#fig-rmlgmeth fig-align="center" width="100%" height="100%"}

```{r}
#| echo: false

dist_p <- plot_dist(distance)  +
  geom_hline(yintercept = benchmark, colour = "red") + 
  ylab(expression(d^{(2)})) +
  theme(axis.text = element_text(size = 5),
        axis.title = element_text(size = 12))

dist_p
```

```{r}
#| echo: false

trimesh + trimesh_gr +
  plot_layout(ncol = 2)

```

### 2024-01-09 appendix.tex
Initial commit

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt]{article}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-1in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.7in}%
\addtolength{\topmargin}{-1in}%

\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{agsm}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Appendix: Visualising How Non-linear Dimension Reduction Warps Your Data},
  pdfauthor={Jayani P.G. Lakshika; Dianne Cook; Paul Harrison; Michael Lydeamore; Thiyanga S. Talagala},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\begin{document}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{\bf Appendix: Visualising How Non-linear Dimension Reduction
Warps Your Data}
\author{
Jayani P.G. Lakshika\\
Econometrics \& Business Statistics, Monash University\\
and\\Dianne Cook\\
Econometrics \& Business Statistics, Monash University\\
and\\Paul Harrison\\
MGBP, BDInstitute, Monash University\\
and\\Michael Lydeamore\\
Econometrics \& Business Statistics, Monash University\\
and\\Thiyanga S. Talagala\\
Statistics, University of Sri Jayewardenepura\\
}
\maketitle

\bigskip
\bigskip
\begin{abstract}

\end{abstract}

\newpage
\spacingset{1} % DON'T change the spacing! (Default 1.9)

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, boxrule=0pt, breakable, enhanced]}{\end{tcolorbox}}\fi

Paramter tunning

\hypertarget{total-number-of-bins}{%
\section{Total number of bins}\label{total-number-of-bins}}

The total number of bins represents the overall count of hexagonal bins
within the hexagonal grid. This count is determined by multiplying the
number of bins along the x-axis (\(b_1\)) with the number of bins along
the y-axis (\(b_2\)), according to the formula:

\begin{equation}\protect\hypertarget{eq-equationp1}{}{
 b = b_1 \times b_2
}\label{eq-equationp1}\end{equation}

Here, \(b\) denotes the total number of bins. By adjusting the parameter
\(b_1\), we have control over the total number of bins (\(b\)). Hence,
fine-tuning \(b_1\) enables us to customize and optimize the total bin
count based on the desired configuration along the x-axis.

\includegraphics[width=1\textwidth,height=\textheight]{appendix_files/figure-pdf/unnamed-chunk-9-1.pdf}

\hypertarget{benchmark-value-to-remove-low-density-hexagons}{%
\section{Benchmark value to remove low-density
hexagons}\label{benchmark-value-to-remove-low-density-hexagons}}

In the step of addressing low-density hexagons, which can arise from
sparsely represented data in certain regions, we employ a systematic
strategy. The goal is to ensure a more comprehensive coverage of the
data by removing hexagons with low data density. To achieve this, we
initiate the process by identifying, for each hex bin, the six nearest
hex bins based on an equal 2D distance metric. Following this, we
calculate the mean density (see Equation~\ref{eq-equationp3}), as
outlined in the equation:

\begin{equation}\protect\hypertarget{eq-equationp2}{}{
 \text{standard count} = \frac{\text{count}}{\text{max count}}
}\label{eq-equationp2}\end{equation}

\begin{equation}\protect\hypertarget{eq-equationp3}{}{
 \text{mean density} = \frac{\text{standard count}}{6}
}\label{eq-equationp3}\end{equation}

The standard count is derived from the number of observations in the hex
bins (see Equation~\ref{eq-equationp2}). Next, we examine the
distribution of mean densities across all hex bins and designate the
first quartile as the benchmark value for removing low-density hexagons.
Finally, hex bins with mean densities below this benchmark value are
removed from consideration. This meticulous procedure ensures the
elimination of regions with inadequate data density, allowing the focus
to shift to areas with more significant data representation. The result
is the preservation of the overall structure of the data in the
low-dimensional space, as illustrated in \textbf{?@fig-bintorm} (c).

\includegraphics{appendix_files/figure-pdf/unnamed-chunk-14-1.pdf}

\begin{verbatim}
Joining with `by = join_by(hexID)`
\end{verbatim}

\includegraphics{appendix_files/figure-pdf/unnamed-chunk-15-1.pdf}

\hypertarget{benchmark-value-to-remove-long-edges}{%
\section{Benchmark value to remove long
edges}\label{benchmark-value-to-remove-long-edges}}

In this step, we aim to create a smoother surface in the low-dimensional
space by removing long edges from the triangular mesh. This process
helps eliminate outliers and noise while preserving important local
relationships and intricate structures within the data. To achieve this,
distances between vertices are sorted, and unique distance values are
extracted. A data frame is created to calculate the differences between
consecutive distance values, and the largest difference serves as a
threshold to identify long edges. By removing edges that exceed this
threshold, the algorithm refines the triangular mesh (see
\textbf{?@fig-traingularmeshgr}).

\begin{figure}

{\centering \includegraphics[width=1\textwidth,height=1\textheight]{figures/remove_long_edges_workflow.png}

}

\caption{\label{fig-rmlgmeth}A flow diagram detailing the steps taken to
find the benchmark value to remove long edges.}

\end{figure}

\includegraphics{appendix_files/figure-pdf/unnamed-chunk-16-1.pdf}

\includegraphics{appendix_files/figure-pdf/unnamed-chunk-17-1.pdf}

\end{document}

### 2024-01-09 paper.qmd
Initial commit

---
title: "Visualising How Non-linear Dimension Reduction Warps Your Data"
format: 
    jasa-pdf:
        keep-tex: true
    jasa-html: default
author:
  - name: Jayani P.G. Lakshika
    affiliations:
      - name: Monash University
        department: Econometrics & Business Statistics
        address: Clayton
        city: VIC 
        country: Austria
        postal-code: 3800
    orcid: 0000-0002-6265-6481
    email: jayani.piyadigamage@monash.edu
    url: https://jayanilakshika.netlify.app/
  - name: Dianne Cook
    affiliations:
      - name: Monash University
        department: Econometrics & Business Statistics
        address: Clayton
        city: VIC 
        country: Austria
        postal-code: 3800
    orcid: 0000-0002-3813-7155
    email: dicook@monash.edu 
    url: http://www.dicook.org/
  - name: Paul Harrison
    affiliations:
      - name: Monash University
        department: MGBP, BDInstitute
        address: Clayton
        city: VIC 
        country: Austria
        postal-code: 3800
    orcid: 0000-0002-3980-268X
    email: 	paul.harrison@monash.edu
    url: 
  - name: Michael Lydeamore
    affiliations:
      - name: Monash University
        department: Econometrics & Business Statistics
        address: Clayton
        city: VIC 
        country: Austria
        postal-code: 3800
    orcid: 0000-0001-6515-827X
    email: michael.lydeamore@monash.edu
    url: 
  - name: Thiyanga S. Talagala
    affiliations:
      - name: University of Sri Jayewardenepura
        department: Statistics
        address: Gangodawila
        city: Nugegoda 
        country: Sri Lanka
        postal-code: 10100
    orcid: 0000-0002-0656-9789
    email: ttalagala@sjp.ac.lk 
    url: https://thiyanga.netlify.app/
tbl-cap-location: bottom
abstract: |
  Non-Linear Dimension Reduction (NLDR) techniques have emerged as powerful tools to visualize high-dimensional data. However, their complexity and parameter choices may lead to distrustful or misleading results. To address this challenge, we propose a novel approach that combines the tour technique with a low-dimensional manifold generated using NLDR techniques, hexagonal binning, and triangulation. This integration enables a clear examination of the low-dimensional representation in the original high-dimensional space. Our approach not only preserves the advantages of both tours and NLDR but also offers a more intuitive perception of complex data structures and facilitates accurate data transformation assessments. The method and example data sets are available in the **quollr** R package.

keywords: [high-dimensional data, dimension reduction, triangulation, hexagonal binning, low-dimensional manifold, manifold learning, tour, data vizualization]
keywords-formatted: [high-dimensional data, dimension reduction, triangulation, hexagonal binning, low-dimensional manifold, manifold learning, tour, data vizualization]

bibliography: bibliography.bib  
---

```{r}
#| warning: false
#| echo: false
library(dplyr)
# remotes::install_github("jlmelville/snedata")
library(snedata)
library(ggflowchart)
library(purrr) ## map function
library(gridExtra) ## for grid.arrange
library(rsample)
library(DT)
library(ggbeeswarm)
library(ggplot2)
library(readr)

library(Rtsne)
library(umap)
library(phateR)
library(reticulate)
library(patchwork)

library(grid)

# install.packages("Seurat")
# remotes::install_github("satijalab/seurat-data")
#library(SeuratData) ## For the application

# use_python("~/miniforge3/envs/pcamp_env/bin/python")
# use_condaenv("pcamp_env")
# 
# reticulate::source_python(paste0(here::here(), "/examples_for_the_paper/function_scripts/Fit_PacMAP_code.py"))
# reticulate::source_python(paste0(here::here(), "/examples_for_the_paper/function_scripts/Fit_TriMAP_code.py"))

set.seed(20230531)

source("quollr_code.R", local = TRUE)
source("nldr_code.R", local = TRUE)
```

```{=html}
<!-- 
Notes
* Use American spelling
-->
```

## Introduction {#sec-intro}

High-dimensional (high-D) data is widespread in many fields including ecology and bioinformatics [@Guo2023], in part because of new data collection technologies [@Johnstone2009; @ayesha2020overview]. Working with high-dimensional data poses considerable challenges due to the difficulty in visualizing beyond two dimensions [@Jia2022]. High-dimensional data also presents difficulties for model fitting [@Johnstone2009], both computationally and interpretation, each of which benefits from visualization.

To create visual representations of high-dimensional data, it is common to apply dimension reduction techniques. Linear methods such as principal component analysis (PCA) [@Karl1901] have been used for many years, and remain popular. Non-linear methods such as multi-dimensional scaling (MDS) [@Torgerson1967] have also been routinely used. In the past decade, there has merged many new techniques non-linear dimension reduction (NLDR), such as t-distributed stochastic neighbor embedding (tSNE) [@Laurens2008], uniform manifold approximation and projection (UMAP) [@Leland2018], designed to capture the complex and non-linear relationships present within high-dimensional data [@Johnstone2009].

However, projecting high-dimensional data has limitations, such as information loss and potential distortion of essential structures and patterns [@Jia2022, @article53]. The choice of technique and parameters further impacts the accuracy of the visualization, necessitating careful consideration for meaningful interpretation (see @fig-nldervis).

Interactive and dynamic graphics systems have also been developed over the years to enable visualizing high dimensions. One method, called a tour [@Asimov1985], shows a sequence of linear projections is shown as a movie, allowing exploration without warping the space [@lee2021review]. Interactive tools like **XGobi** and **GGobi** have been successful in incorporating tours for exploring high-dimensional data [@article60]. The R package **tourr** [@article61] further enhances tour visualization within R, although it may face limitations in frame rate and interactive features compared to **GGobi**.

To overcome these limitations, the R package **detourr** [@article22] has been developed, leveraging a Javascript widget via htmlwidgets [@Ramnath2023] to achieve higher frame rates and enhanced interactivity. Additionally, the R package **langevitour** [@article09] utilizes Langevin Dynamics to generate a continuous path of projections, eliminating the need for interpolation between projections for animation. The tour technique has proven valuable in exploring statistical model fits [@article58] and factorial experimental designs [@article59]. Augmenting the results of non-linear dimensional reduction methods with the tour, as demonstrated in the **liminal** R package [@article21], further enhances data exploration.

While tours [@Asimov1985] preserve space without warping [@lee2021review], they require integrating multiple low-dimensional views mentally to perceive high-dimensional structures. To address this challenge, we propose a novel approach by combining the tour technique with a low-dimensional manifold. This manifold is created through the synergistic use of Non-Linear Dimension Reduction (NLDR) techniques, hexagonal binning, and triangulation. By merging these techniques, our approach offers a comprehensive and efficient means to visualize and explore high-dimensional data while retaining the advantages of both tours and NLDR. This integration facilitates a more intuitive perception of complex data structures and empowers analysts with a robust tool for assessing the accuracy of data transformations. The implementation of our approach is available as an R package called **quollr**.

The outline of this paper is as follows. The @sec-background provides an detailed overview of dimension reduction methods, triangulation, and tours. Building upon this foundation, the @sec-methods delves into the proposed algorithm, **quollr**, and its implementation details. In @sec-prediction, discusses the effectiveness of the learned low-dimensional manifold in accurately representing the complex high-dimensional data. Following that, @sec-simpleex presents simple examples from simulations to illustrate the functionality of the algorithm. Subsequently, @sec-applications showcases real-world applications of **quollr** on different data sets, particularly in single-cell RNA-seq data. These applications reveal insights into the performance and trustworthiness of NLDR algorithms. We analyze the results to identify situations where NLDR techniques may lead to misleading interpretations. Finally, @sec-conclusions concludes by summarizing the findings and emphasizing the significance of the proposed approach in tackling the challenges of high-dimensional data visualization.

## Background {#sec-background}

### Dimension Reduction

Consider the high-dimensional data a rectangular matrix $X$, where $X = \begin{bmatrix} \textbf{x}_{1} & \textbf{x}_{2} & \cdots & \textbf{x}_{n}\\ \end{bmatrix}$, with $n$ observations in $p$ dimensions. The objective is to discover a low-dimensional projection $Y = \begin{bmatrix} \textbf{y}_{1} & \textbf{y}_{2} & \cdots & \textbf{y}_{n}\\ \end{bmatrix}$, represented as an $n$ × $d$ matrix, where $d \ll p$. The reduction process seeks to remove noise from the original data set while retaining essential information.

There are two main categories of dimension reduction techniques: linear and non-linear methods. Linear techniques involve a linear transformation of the data, with one popular example being PCA. PCA performs an eigen-decomposition of the sample covariance matrix to obtain orthogonal principal components that capture the variance of the data [@Karl1901]. However, linear methods may not fully capture complex non-linear relationships present in the data.

In contrast, NLDR techniques generate the low-dimensional representation $Y$ from the high-dimensional data $X$, often using pre-processing techniques like $k$-nearest neighbors graph or kernel transformations. Multidimensional Scaling (MDS) is a class of NLDR methods that aims to construct an embedding $Y$ in a low-dimensional space, approximating the pair-wise distances in $X$ [@Torgerson1967]. Variants of MDS include non-metric scaling [@article62] and Isomap, which estimate geodesic distances to create the low-dimensional representation [@article63]. Other approaches based on diffusion processes, like diffusion maps [@article64] and the PHATE algorithm [@article03], also fall under NLDR methods.

A challenge with NLDR methods is selecting and tuning appropriate parameters. One specific technique we focus on is Pairwise Controlled Manifold Approximation (PaCMAP). Similar considerations apply to related methods like tSNE [@Laurens2008], UMAP [@Leland2018], and TrMAP [@article02].

It is important to note that methods like PCA and auto-encoders [@article65] provide a reverse mapping from the low-dimensional space back to the high-dimensional space, enabling data reconstruction. However, many non-linear methods, including tSNE, prioritize visualization and exploration over reconstruction. Their focus is on capturing complex structures that may not be easily represented in the original space, making a straightforward reverse mapping challenging.

#### Introduction to UMAP

The UMAP algorithm constructs a fuzzy topological representation of the high-dimensional data and then optimizes the low-dimensional embedding to be as similar to this representation as possible [@Leland2018].

The UMAP algorithm constructs a weighted k-nearest neighbors graph based on the local similarity of data points in the original high-dimensional space. The weighted graph is defined by a conditional probability distribution $P_{ij}$, which represents the probability of observing data point $x_i$ given data point $x_j$ as a neighbor:

$$
P_{ij} = \frac{\exp(-\lVert x_i - x_j \rVert^2)}{\sum_{k \neq i} \exp(-\lVert x_i - x_k \rVert^2)}
$$

Similarly, a fuzzy topological representation of the data is constructed in the low-dimensional space with conditional probability distribution $Q_{ij}$:

$$
Q_{ij} = \frac{\exp(-\lVert y_i - y_j \rVert^2)}{\sum_{k \neq i} \exp(-\lVert y_i - y_k \rVert^2)}
$$

where $y_i$ and $y_j$ are the corresponding low-dimensional representations of data points $x_i$ and $x_j$.

UMAP minimizes the cross-entropy between the high-dimensional fuzzy topological representation $P_{ij}$ and the low-dimensional representation $Q_{ij}$ to find an optimal low-dimensional embedding. The loss function for UMAP is defined as:

$$
L = \sum_{i} \sum_{j} P_{ij} \log \frac{P_{ij}}{Q_{ij}}
$$

The algorithm uses stochastic gradient descent to optimize the positions of the data points in the low-dimensional space to minimize this loss function.

In R, the `umap` package is often used to implement UMAP and generate results.

The "n_neighbors" parameter in UMAP determines the number of nearest neighbors to consider for each data point during the construction of the fuzzy topological representation. Higher values of "n_neighbors" capture more global structure, while lower values emphasize local structure. Choosing an appropriate value for "n_neighbors" is crucial, as it affects the balance between local and global structure in the resulting low-dimensional representation.

```{r}
#| warning: false
#| echo: false

data <- read_csv(paste0(here::here(), "/data/s_curve.csv"))
```

```{r}
#| warning: false
#| echo: false

data <- data |>
  mutate(ID = row_number())

data_split <- initial_split(data)
training_data <- training(data_split) |>
  arrange(ID)
test_data <- testing(data_split) |>
  arrange(ID)
```

```{r}
#| warning: false
#| echo: false

#plot_list <- list()

tSNE_data <- Fit_tSNE(training_data, opt_perplexity = calculate_effective_perplexity(training_data), with_seed = 20230531)
#(perplexity: ", calculate_effective_perplexity(data), ")
plot_list1 <- plot_tSNE_2D(tSNE_data) + #ggtitle("(a)") + 
  theme_linedraw() +
    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

```

```{r}
#| warning: false
#| echo: false

UMAP_fit <- umap(training_data |> dplyr::select(-ID), n_neighbors = 50, n_components =  2)

UMAP_data <- UMAP_fit$layout |>
  as.data.frame()
names(UMAP_data)[1:(ncol(UMAP_data))] <- paste0(rep("UMAP",(ncol(UMAP_data))), 1:(ncol(UMAP_data)))

UMAP_data <- UMAP_data |>
  mutate(ID = training_data$ID)

#(n-neighbors: 50)
plot_list2 <- plot_UMAP_2D(UMAP_data) + #ggtitle("(b)") + 
  theme_linedraw() +
    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

```

```{r}
#| warning: false
#| echo: false

#(knn: 5)
# PHATE_data <- Fit_PHATE(training_data, knn = 5, with_seed = 20230531)
# write_csv(PHATE_data, paste0(here::here(), "/data/phate_data_s_curve.csv"))

PHATE_data <- read_csv(paste0(here::here(), "/data/phate_data_s_curve.csv"))

plot_list3 <- plot_PHATE_2D(PHATE_data) + #ggtitle("(c)") + 
  theme_linedraw() +
    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

```

```{r}
#| warning: false
#| echo: false

# tem_dir <- tempdir()
# 
# Fit_TriMAP_data(training_data, tem_dir)
# 
# path <- file.path(tem_dir, "df_2_without_class.csv")
# path2 <- file.path(tem_dir, "dataset_3_TriMAP_values.csv")
# 
# Fit_TriMAP(as.integer(2), as.integer(5), as.integer(4), as.integer(3), path, path2)
# 
# TriMAP_data <- read_csv(path2)
# write_csv(TriMAP_data, paste0(here::here(), "/data/trimap_data_s_curve.csv"))

TriMAP_data <- read_csv(paste0(here::here(), "/data/trimap_data_s_curve.csv"))

#(n-inliers: 5, \n n-outliers: 4, n-random: 3)
plot_list4 <- plot_TriMAP_2D(TriMAP_data) + #ggtitle("(d)") + 
  theme_linedraw() +
    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)

```

```{r}
#| warning: false
#| echo: false

# tem_dir <- tempdir()
# 
# Fit_PacMAP_data(training_data, tem_dir)
# 
# path <- file.path(tem_dir, "df_2_without_class.csv")
# path2 <- file.path(tem_dir, "dataset_3_PaCMAP_values.csv")
# 
# Fit_PaCMAP(as.integer(2), as.integer(10), "random", 0.9, as.integer(2), path, path2)
# 
# PacMAP_data <- read_csv(path2)
# write_csv(PacMAP_data, paste0(here::here(), "/data/pacmap_data_s_curve.csv"))

PaCMAP_data <- read_csv(paste0(here::here(), "/data/pacmap_data_s_curve.csv"))

#(knn: 10, init: random, \n MN-ratio: 0.9, FP-ratio: 2)
plot_list5 <- plot_PaCMAP_2D(PaCMAP_data) + #ggtitle("(e)") + 
  theme_linedraw() +
    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'e', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

```

```{r}
#| echo: false
#| fig-cap: "2D layouts from different NLDR techniques applied the same data: (a) tSNE (perplexity = 32), (b) UMAP (n_neighbors = 50), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2). Is there a best representation of the original data or are they all providing  equivalent information?"
#| label: fig-nldervis
#| out-width: 100%

plot_list1 + plot_list2 + plot_list3 + plot_list4 + plot_list5 +
  plot_layout(ncol=5, widths=c(0.35, 0.15, 0.30, 0.12, 0.11))
```

### Linear overviews using tours

A tour is a powerful visualization technique used to explore high-dimensional data by generating a sequence of projections, typically into two dimensions. There are two main types of tours: the Grand Tour and the Guided Tour. The Grand Tour explores the data's shape and global structure by using random projections [@Asimov1985]. In contrast, the Guided Tour focuses on specific patterns by moving towards interesting projections defined by a predefined index function [@article29].

The process begins with a real data matrix $X$ containing $n$ observations in $p$ dimensions. It generates a sequence of $p$ × $d$ orthonormal projection matrices (bases), usually 1 or 2 dimensions. For each pair of orthonormal bases $A_t$ and $A_{t+1}$, a geodesic path is interpolated to create smooth animation between projections.

In the Grand Tour, new orthonormal bases are randomly chosen to explore the $d$-dimensional subspace. The data is often sphered via principal components to reduce dimensionality. The Guided Tour uses a predefined index function to generate a sequence of 'interesting' projections. The resulting tour continuously visualizes the projected data $Y_t$ = $XA_t$ as it interpolates between successive bases.

While both tours can be used to visualize data, examples often focus on using the Grand Tour to observe global structures. However, software like **langevitour** can visualize both types of tours, providing flexibility for exploring high-dimensional data with various objectives.

### Non-parametric models

#### Hexagonal binning

Hexagonal binning is a data visualization technique that aggregates high-dimensional data into a two-dimensional representation using hexagonal regions called bins [@article66]. The hexagonal shape is preferred due to its circular-like appearance, providing smoother transitions between neighboring bins and efficient data aggregation. This technique is especially useful for handling large data sets as it reduces visual clutter while capturing the underlying data distribution effectively. By considering the data distribution, important features are captured, and gaps between bins are minimized, leading to a more accurate representation of the data. In our algorithm, hexagonal binning is utilized to create a low-dimensional manifold and facilitate data visualization in the original high-dimensional space.

#### Triangulation

Triangulation is a fundamental process in computational geometry and computer graphics that involves dividing a set of points in a given space into interconnected triangles [@article30] finding applications in mesh generation and representing complex geometric structures efficiently and discretely. One common algorithm used for triangulation is Delaunay triangulation [@article26], where points are connected in a way that maximizes the minimum angles of the resulting triangles, leading to a more regular and well-conditioned triangulation.

In our algorithm, we are interested in triangulation as it plays a key role in providing a clear and concise representation of the data. By converting the data points into interconnected triangles, we can visualize and analyze the data in a more structured and organized manner. Triangulation helps us identify geometric relationships and patterns in the data, enabling us to make informed decisions and gain valuable insights from the information presented in the triangular form.

## Methodology {#sec-methods}

<!--In this paper, we introduce a novel approach, which is designed to determine which method, and which parameter choice provide the most useful representation of high dimensional data. Our approach involves dividing the data set into two subsets: a training set, which serves to construct the low-dimensional manifold, and a test set, utilized to generate predictive values and residuals.

For the implementation of our approach, a 2D embedding data set serves as the initiation point for the subsequent analysis and visualization (see @fig-nldervis (b)).-->

<!--### Overview {#sec-algorithm}-->

Our algorithm comprises two main phases: (1) generate the model in the 2D space, and (2) generate the model in the high-D space. These two phases are described in details in this section.

![A flow diagram detailing the steps taken to create the low-dimensional manifold in the high dimensional space. There are two basic phases, one to generate the model in the 2D space, and other to generate the model in the high-D space.](figures/workflow.png){#fig-meth fig-align="center" width="100%" height="100%"}

### Preprocessing steps

To tackle the complexities and noise in high-dimensional data, we apply PCA as a pre-processing step [@article67, @article68, @article69]. This step helps in noise reduction by identifying principal components that represent directions of maximum variance, capturing essential patterns in the data.

### Constructing the 2D model

**Step 1: Computing the hexagonal grid configuration**

**(a) Determine the number of bins along the x-axis** ($b_1$)

The first step involves to determine the optimal number of bins ($b$) for creating regular hexagons in the hexagonal grid. To achieve this, we rely on the relationship between the diameter ($h$) and the area ($A$) of regular hexagons, as given by @eq-equation3.

$$
 \text{A} = \frac{\sqrt{3}}{2}h^2
$$ {#eq-equation3}

To capture the data's structure effectively, we consider the optimal $A = 1$ (see @fig-binsize). Using @eq-equation4, we then calculate $h$ of a regular hexagon.

$$
 \text{h} = \sqrt{\frac{2}{\sqrt{3}}A}
$$ {#eq-equation4}

Next, we refer the `hexbin` function in R, which provides a relationship between the diameter of the hexagon ($h$) and the number of bins along the x-axis ($b_1$) (@eq-equation6).

$$
 h = \frac{r_1}{b_1}
$$ {#eq-equation6}

By utilizing the calculated $h$ and the range of the non-linear projection component 1 ($r_1$), we determine the $b_1$ (@eq-equation5). The result is rounded up to the nearest whole number to ensure an appropriate number of bins that capture the variability of the data along the x-axis.

$$
 b_1 = \frac{r_1}{h}
$$ {#eq-equation5}

**(b) Determine the effective shape parameter** ($s$)

In this step, we determine the optimal shape parameter ($s$) for the hexagonal bins, which significantly influences their shape and arrangement within the grid. The shape parameter ($s$) in the hexagonal binning algorithm is defined as the ratio of the height ($y$) to the width ($x$) of the plotting region, referring to the `hexbin` function in R (see @eq-equation1). It determines the shape of the plotting regions and plays a vital role in generating an appropriate hexagonal grid for the data visualization.

$$
  s = \frac{y}{x}
$$ {#eq-equation1}

To calculate the effective shape parameter ($s$) for our algorithm, we consider the ranges of the non-linear projection components. Denoting the range of the non-linear projection component 1 as $r_1$ and the range of the non-linear projection component 2 as $r_2$, we find the shape parameter using @eq-equation2.

$$
s = \frac{r_2}{r_1}
$$ {#eq-equation2}

<!--first value-->

```{r}
#| echo: false
#| message: false
#| warning: false
 # Loop over loop.vector
  #i <- 1
  num_bins <- 3
  shape_val <- calculate_effective_shape_value(.data = UMAP_data, x = UMAP1, y = UMAP2)

  hexbin_data_object_loop <- extract_hexbin_centroids(UMAP_data, num_bins, shape_val)

  df_bin_centroids_loop <- hexbin_data_object_loop$hexdf_data

  ## Data set with all possible centroids in the hexagonal grid

  full_centroid_df_loop <- generate_full_grid_centroids(df_bin_centroids_loop)

  ## To map hexID to hexbin centroids in the full grid

  vec1 <- stats::setNames(rep("", 2), c("x", "y"))  ## Define column names

  full_grid_with_hexbin_id_loop <- dplyr::bind_rows(vec1)[0, ]
  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
    dplyr::mutate_if(is.character, as.numeric)

  for(i in 1:length(sort(unique(full_centroid_df_loop$y)))){

    ## Filter the data set with specific y value
    specific_y_val_df_loop <- full_centroid_df_loop |>
      dplyr::filter(y == sort(unique(full_centroid_df_loop$y))[i])

    ordered_x_df_loop <- specific_y_val_df_loop |>
      dplyr::arrange(x)

    full_grid_with_hexbin_id_loop <- dplyr::bind_rows(full_grid_with_hexbin_id_loop, ordered_x_df_loop)

  }

  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
    dplyr::mutate(hexID = row_number())

  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
    dplyr::rename("c_x" = "x",
           "c_y" = "y")

  full_grid_with_hexbin_id_loop <- dplyr::full_join(full_grid_with_hexbin_id_loop, df_bin_centroids_loop, by = c("hexID" = "hexID")) |>
    dplyr::select(-c(x, y))

  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
      dplyr::mutate(std_counts = counts/max(counts, na.rm = TRUE))

  ## Generate all coordinates of hexagons
  hex_grid_loop <- full_hex_grid(full_centroid_df_loop)

  full_grid_with_polygon_id_df_loop <- map_polygon_id(full_grid_with_hexbin_id_loop, hex_grid_loop)

  full_grid_with_hexbin_id_rep_loop <- full_grid_with_polygon_id_df_loop |>
  dplyr::slice(rep(1:n(), each = 6)) |>
  dplyr::arrange(polygon_id)

  hex_full_count_df_loop <- dplyr::bind_cols(hex_grid_loop, full_grid_with_hexbin_id_rep_loop)

  p1 <-  ggplot(data = hex_full_count_df_loop, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
        geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
    theme_void() +
    theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
          axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
          legend.title = element_text(size=8), #change legend title font size
          legend.text = element_text(size=6)) +
    guides(fill = guide_colourbar(title = "Standardized count")) +
    annotate(geom = 'text', label = "a", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)

```

```{r}
#| echo: false
#| message: false
#| warning: false
 # Loop over loop.vector
  #i <- 1
  num_bins <- 8
  shape_val <- calculate_effective_shape_value(.data = UMAP_data, x = UMAP1, y = UMAP2)

  hexbin_data_object_loop <- extract_hexbin_centroids(UMAP_data, num_bins, shape_val)

  df_bin_centroids_loop <- hexbin_data_object_loop$hexdf_data

  ## Data set with all possible centroids in the hexagonal grid

  full_centroid_df_loop <- generate_full_grid_centroids(df_bin_centroids_loop)

  ## To map hexID to hexbin centroids in the full grid

  vec1 <- stats::setNames(rep("", 2), c("x", "y"))  ## Define column names

  full_grid_with_hexbin_id_loop <- dplyr::bind_rows(vec1)[0, ]
  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
    dplyr::mutate_if(is.character, as.numeric)

  for(i in 1:length(sort(unique(full_centroid_df_loop$y)))){

    ## Filter the data set with specific y value
    specific_y_val_df_loop <- full_centroid_df_loop |>
      dplyr::filter(y == sort(unique(full_centroid_df_loop$y))[i])

    ordered_x_df_loop <- specific_y_val_df_loop |>
      dplyr::arrange(x)

    full_grid_with_hexbin_id_loop <- dplyr::bind_rows(full_grid_with_hexbin_id_loop, ordered_x_df_loop)

  }

  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
    dplyr::mutate(hexID = row_number())

  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
    dplyr::rename("c_x" = "x",
           "c_y" = "y")

  full_grid_with_hexbin_id_loop <- dplyr::full_join(full_grid_with_hexbin_id_loop, df_bin_centroids_loop, by = c("hexID" = "hexID")) |>
    dplyr::select(-c(x, y))

  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
      dplyr::mutate(std_counts = counts/max(counts, na.rm = TRUE))

  ## Generate all coordinates of hexagons
  hex_grid_loop <- full_hex_grid(full_centroid_df_loop)

  full_grid_with_polygon_id_df_loop <- map_polygon_id(full_grid_with_hexbin_id_loop, hex_grid_loop)

  full_grid_with_hexbin_id_rep_loop <- full_grid_with_polygon_id_df_loop |>
  dplyr::slice(rep(1:n(), each = 6)) |>
  dplyr::arrange(polygon_id)

  hex_full_count_df_loop <- dplyr::bind_cols(hex_grid_loop, full_grid_with_hexbin_id_rep_loop)

  p2 <-  ggplot(data = hex_full_count_df_loop, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
    geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
        theme_void() +
    theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
          axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
          legend.title = element_text(size=8), #change legend title font size
          legend.text = element_text(size=6)) +
    guides(fill = guide_colourbar(title = "Standardized count")) +
    annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)

```

```{r}
#| echo: false
#| message: false
#| warning: false
 # Loop over loop.vector
  #i <- 1
  num_bins <- 15
  shape_val <- calculate_effective_shape_value(.data = UMAP_data, x = UMAP1, y = UMAP2)

  hexbin_data_object_loop <- extract_hexbin_centroids(UMAP_data, num_bins, shape_val)

  df_bin_centroids_loop <- hexbin_data_object_loop$hexdf_data

  ## Data set with all possible centroids in the hexagonal grid

  full_centroid_df_loop <- generate_full_grid_centroids(df_bin_centroids_loop)

  ## To map hexID to hexbin centroids in the full grid

  vec1 <- stats::setNames(rep("", 2), c("x", "y"))  ## Define column names

  full_grid_with_hexbin_id_loop <- dplyr::bind_rows(vec1)[0, ]
  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
    dplyr::mutate_if(is.character, as.numeric)

  for(i in 1:length(sort(unique(full_centroid_df_loop$y)))){

    ## Filter the data set with specific y value
    specific_y_val_df_loop <- full_centroid_df_loop |>
      dplyr::filter(y == sort(unique(full_centroid_df_loop$y))[i])

    ordered_x_df_loop <- specific_y_val_df_loop |>
      dplyr::arrange(x)

    full_grid_with_hexbin_id_loop <- dplyr::bind_rows(full_grid_with_hexbin_id_loop, ordered_x_df_loop)

  }

  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
    dplyr::mutate(hexID = row_number())

  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
    dplyr::rename("c_x" = "x",
           "c_y" = "y")

  full_grid_with_hexbin_id_loop <- dplyr::full_join(full_grid_with_hexbin_id_loop, df_bin_centroids_loop, by = c("hexID" = "hexID")) |>
    dplyr::select(-c(x, y))

  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
      dplyr::mutate(std_counts = counts/max(counts, na.rm = TRUE))

  ## Generate all coordinates of hexagons
  hex_grid_loop <- full_hex_grid(full_centroid_df_loop)

  full_grid_with_polygon_id_df_loop <- map_polygon_id(full_grid_with_hexbin_id_loop, hex_grid_loop)

  full_grid_with_hexbin_id_rep_loop <- full_grid_with_polygon_id_df_loop |>
  dplyr::slice(rep(1:n(), each = 6)) |>
  dplyr::arrange(polygon_id)

  hex_full_count_df_loop <- dplyr::bind_cols(hex_grid_loop, full_grid_with_hexbin_id_rep_loop)

  p3 <-  ggplot(data = hex_full_count_df_loop, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
        geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
        theme_void() +
    theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
          axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
          legend.title = element_text(size=8), #change legend title font size
          legend.text = element_text(size=6)) +
    guides(fill = guide_colourbar(title = "Standardized count")) +
    annotate(geom = 'text', label = "c", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)

```

```{r}
#| echo: false
#| label: fig-binsize
#| fig-pos: H
#| fig-cap: "Hexbin plots from different number of bins for the same **s_curve_noise_umap** data: (a) b = (4, 8), s = 2.031, (b) b = (9, 20), s = 2.031, and (c) b = (16, 36), s = 2.031. The hexbins are colored based on the density of points, with yellow indicating higher point density and darker colors representing lower point density within each bin. Does  a value of number of bins exist to effectively represent the low-dimensional data?"

p1 + p2 + p3
```

**Step 2: Obtain bin centroids**

After computing hexagonal grid configurations, the 2D emdeddings are binned into hexagon cells. As a result of hexagonal binning , the bin centroids ($C_k^{(2)} \equiv (C_{kx}, C_{ky}$)) (see @fig-meth (Step 2)) are obtained.

**Step 3: Triangulate bin centroids**

In this step, the algorithm proceeds to triangulate the centroids (see @fig-meth (Step 3)) of the hex bins using the `tripack` package in R [@article54]. Since we are working with the centroids of regular hexagonal bins, the resulting mesh will predominantly comprise equal-sized regular triangles. However, the triangulation also helps span any gaps that may exist between clusters of points, allowing for a more complete and interconnected representation of the data.

### Lifting the model into high dimensions

Creating the model in the high-D space involves a two-step process. Initially, we identify observations within each hexagonal bin. After this, we calculate the average of the high-dimensional data within each hex bin. The resulting values represent the averaged high-dimensional data points ($C_k^{(p)} \equiv (C_{kx_1}, ..., C_{kx_p}$)). Essentially, these averaged high-dimensional data points serve as the representation of the hex bin centroids in the high-D space.

```{r}
#| echo: false

UMAP_fit <- umap(training_data |> dplyr::select(-ID), n_neighbors = 50, n_components =  2)

UMAP_data <- UMAP_fit$layout |>
  as.data.frame()
names(UMAP_data)[1:(ncol(UMAP_data))] <- paste0(rep("UMAP",(ncol(UMAP_data))), 1:(ncol(UMAP_data)))

UMAP_data <- UMAP_data |>
  mutate(ID = training_data$ID)

num_bins_x <- calculate_effective_x_bins(.data = UMAP_data, x = UMAP1,
                           cell_area = 1)

shape_value <- calculate_effective_shape_value(.data = UMAP_data, 
                                             x = UMAP1, y = UMAP2)

## To extract bin centroids
hexbin_data_object <-extract_hexbin_centroids(nldr_df = UMAP_data, num_bins = num_bins_x, shape_val = shape_value)

df_bin_centroids <- hexbin_data_object$hexdf_data

##########

## Data set with all possible centroids in the hexagonal grid

full_centroid_df <- generate_full_grid_centroids(df_bin_centroids)

## To map hexID to hexbin centroids in the full grid

vec1 <- stats::setNames(rep("", 2), c("x", "y"))  ## Define column names

full_grid_with_hexbin_id <- dplyr::bind_rows(vec1)[0, ]
full_grid_with_hexbin_id <- full_grid_with_hexbin_id |>
  dplyr::mutate_if(is.character, as.numeric)

for(i in 1:length(sort(unique(full_centroid_df$y)))){

  ## Filter the data set with specific y value
  specific_y_val_df <- full_centroid_df |>
    dplyr::filter(y == sort(unique(full_centroid_df$y))[i])

  ordered_x_df <- specific_y_val_df |>
    dplyr::arrange(x)

  full_grid_with_hexbin_id <- dplyr::bind_rows(full_grid_with_hexbin_id, ordered_x_df)

}

full_grid_with_hexbin_id <- full_grid_with_hexbin_id |>
  dplyr::mutate(hexID = row_number())

full_grid_with_hexbin_id <- full_grid_with_hexbin_id |>
  dplyr::rename("c_x" = "x",
         "c_y" = "y")

full_grid_with_hexbin_id <- dplyr::full_join(full_grid_with_hexbin_id, df_bin_centroids, by = c("hexID" = "hexID")) |>
  dplyr::select(-c(x, y)) #|> 
  #dplyr::mutate(counts = tidyr::replace_na(counts, 0))

full_grid_with_hexbin_id <- full_grid_with_hexbin_id |>
    dplyr::mutate(std_counts = counts/max(counts, na.rm = TRUE))

## Generate all coordinates of hexagons
hex_grid <- full_hex_grid(full_centroid_df)

full_grid_with_polygon_id_df <- map_polygon_id(full_grid_with_hexbin_id, hex_grid)

full_grid_with_hexbin_id_rep <- full_grid_with_polygon_id_df |>
  dplyr::slice(rep(1:n(), each = 6)) |>
  dplyr::arrange(polygon_id)

hex_full_count_df <- dplyr::bind_cols(hex_grid, full_grid_with_hexbin_id_rep)

##########

min_std_cell_threshold <- 0.05

df_bin_centroids_all <- hexbin_data_object$hexdf_data ## All the centroids without removing low-density hexagons

df_bin_centroids <- df_bin_centroids |>
  dplyr::mutate(stand_cell_count = counts/max(counts)) |>
  dplyr::filter(stand_cell_count > min_std_cell_threshold)

UMAP_data_with_hb_id <- UMAP_data |> 
  dplyr::mutate(hb_id = hexbin_data_object$hb_data@cID)

## To generate a data set with high-D and 2D training data
df_all <- dplyr::bind_cols(training_data |> dplyr::select(-ID), UMAP_data_with_hb_id)

## Averaged on high-D
df_bin <- avg_highD_data(.data = df_all)

## Triangulate bin centroids
tr1_object <- triangulate_bin_centroids(df_bin_centroids, x, y)
tr_from_to_df <- generate_edge_info(triangular_object = tr1_object)

## Compute 2D distances
distance <- cal_2D_dist(.data = tr_from_to_df)

## To find the benchmark value
benchmark <- find_benchmark_value(.data = distance, distance_col = distance)

trimesh <- ggplot(df_bin_centroids, aes(x = x, y = y)) + 
    geom_segment(data = tr_from_to_df, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
    geom_point(size = 2, colour = "#33a02c") +
    coord_equal()

  # ggplot(df_bin_centroids, aes(x = x, y = y)) + 
  # geom_point(size = 1, colour = "#33a02c") + 
  # geom_trimesh() + 
  # coord_equal()

trimesh <- trimesh +
  #ggtitle("(a)") + 
  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  theme_light() +
    theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
          axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
    ) 
  # theme(axis.text = element_text(size = 5),
  #       axis.title = element_text(size = 7))

trimesh_gr <- colour_long_edges(.data = distance, benchmark_value = benchmark, 
                                triangular_object = tr1_object, distance_col = distance)

trimesh_gr <- trimesh_gr + 
    geom_point(size = 2, colour = "#33a02c") + 
  #ggtitle("(b)") + 
  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  theme_light() +
    #coord_equal() +
    theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
          axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
    )

trimesh_removed <- remove_long_edges(.data = distance, benchmark_value = benchmark, 
                                triangular_object = tr1_object, distance_col = distance)

trimesh_removed <- trimesh_removed + 
       geom_point(size = 2, colour = "#33a02c") + 
  #ggtitle("(b)") + 
  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  theme_light() +
    #coord_equal() +
    theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
          axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
    )

tour1 <- show_langevitour(df_all, df_bin, df_bin_centroids, benchmark_value = benchmark, distance = distance, distance_col = distance)
```

```{r}
#| echo: false

## To plot the distribution of distance
plot_dist <- function(distance_df){
  distance_df$group <- "1"
  dist_plot <- ggplot(distance_df, aes(x = group, y = distance)) +
    geom_quasirandom()+
    ylim(0, max(unlist(distance_df$distance))+ 0.5) + coord_flip()
  return(dist_plot)
}
```

```{r}
#| echo: false

distance_plot <- plot_dist(distance) +
  #ggtitle("(b)" ) + 
  ylab(expression(d^{(2)})) +
  theme(axis.text = element_text(size = 5),
        axis.title = element_text(size = 12))
```

```{r}
#| echo: false

a1 <- colour_long_edges(.data = distance, benchmark_value = 13, 
                  triangular_object = tr1_object, distance_col = distance) + 
    geom_point(size = 2, colour = "#33a02c") + 
    #ggtitle("(b)") + 
    xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
    theme_light() +
    #coord_equal() +
    theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
          axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
    )

UMAP_data_with_hb_id_selected <- UMAP_data_with_hb_id |> dplyr::filter(hb_id %in% c(14, 177)) |>
  dplyr::mutate(select_point = if_else(hb_id == 14, "data1", "data2"))

a2 <- ggplot(data = hex_full_count_df, aes(x = x, y = y)) + 
  geom_polygon(color = "black", aes(group = polygon_id), fill = "#ffffff") +
  geom_point(data = UMAP_data_with_hb_id_selected, aes(x = UMAP1, y = UMAP2, colour = select_point), alpha = 0.5) +
  scale_color_manual(values=c("#e6550d", "#000000")) +
  geom_point(data = df_bin_centroids_all |> dplyr::filter(hexID %in% c(14, 177)), aes(x = x, y = y), colour = "#33a02c") +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
  theme_light() +
  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
          axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
    )

## Script to make coloring data points with a line

df1 <- df_all |> 
  dplyr::filter(hb_id == "14") |>
  dplyr::select(tidyselect::starts_with("x")) |>
  dplyr::mutate(type = "data1") ## original dataset

df2 <- df_all |> 
  dplyr::filter(hb_id == "177") |>
  dplyr::select(tidyselect::starts_with("x")) |>
  dplyr::mutate(type = "data2") ## original dataset

df3 <- df_all |> 
  dplyr::filter(!(hb_id %in% c("14", "177"))) |>
  dplyr::select(tidyselect::starts_with("x")) |>
  dplyr::mutate(type = "data") ## original dataset

df_b <- df_bin |> 
  dplyr::filter(hb_id %in% df_bin_centroids$hexID) |>
  dplyr::select(-hb_id) |>
  dplyr::mutate(type = "model") ## Data with summarized mean

df_exe <- dplyr::bind_rows(df_b, df1, df2, df3)

distance_df_small_edges <- distance |> 
  dplyr::filter(from == 3) |>
  dplyr::filter(to == 49) |>
  head(1)

lg_high <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)], group = df_exe$type, pointSize = 3, levelColors = c("#6a3d9a", "#e6550d", "#000000", "#33a02c"), lineFrom = distance_df_small_edges$from, lineTo = distance_df_small_edges$to, lineColors = "#e31a1c")

```

```{r}
#| warning: false
#| echo: false
#| label: fig-wkhighD
#| fig-cap: Visualization.

a1 + a2 +
  plot_annotation(tag_levels = 'a') +
  plot_layout(guides='collect', ncol = 3) &
  theme(legend.position='bottom', plot.tag = element_text(size = 8))
```
### Tunning the model

### Summaries of the model {#sec-diagnostics}

#### Predicted values and residuals

In our prediction process, we employ the K-nearest neighbors (KNN) algorithm to predict the hex bin IDs. Specifically, for each new high-dimensional data point, the algorithm assigns the first neighboring hex bin ID as the predicted hex bin ID.

<!-- move to goodness of fit stat-->
Moving on to the assessment of prediction accuracy, we calculate the Mean Squared Error (MSE). The MSE helps measure the average squared differences between the actual high-dimensional data ($x_{ij}$) and the predicted bin centroid data in high-D ($C_{x_ij}$) values across all bins. Mathematically, this is expressed as:

$$
\text{MSE} = \sum_{j = 1}^{n} \frac{\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2}{\text{total number of bins}}
$$ {#eq-equation9}

Here, $b$ signifies the total number of bins, $p$ denotes the number of dimensions in the high-dimensional (high-D) data, and $n$ represents the number of observations.

Additionally, we gauge the model's performance using the Akaike Information Criterion (AIC), calculated by the formula:

$$
\text{AIC} = 2bp + np * log(\text{MSE})
$$ {#eq-equation10}

These metrics, MSE and AIC, collectively offer valuable insights into the model's predictive performance, considering both accuracy and complexity in the predictions.

```{r}
#| echo: false
predict_UMAP_df <- predict(UMAP_fit, test_data |> dplyr::select(-ID)) |>
  as.data.frame()

names(predict_UMAP_df)[1:(ncol(predict_UMAP_df))] <- paste0(rep("UMAP",(ncol(predict_UMAP_df))), 1:(ncol(predict_UMAP_df)))

predict_UMAP_df <- predict_UMAP_df |>
  mutate(ID = test_data$ID)
```

```{r}
#| echo: false

predict_umap_plot <- plot_UMAP_2D(UMAP_data) +
  geom_point(data = predict_UMAP_df, aes(x = UMAP1, y = UMAP2), color = "red") + 
  theme_linedraw() +
    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
```

<!--predict hexid for the test data -->

```{r}
#| echo: false

pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = UMAP_data, num_bins = num_bins_x, shape_val = shape_value)
pred_df_training <- pred_df_training_object$pred_data
centroid_df_training <- pred_df_training_object$df_bin_centroids
avg_df_training <- pred_df_training_object$df_bin

eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_x)

pred_df_test_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = predict_UMAP_df, num_bins = num_bins_x, shape_val = shape_value)
pred_df_test <- pred_df_test_object$pred_data
centroid_df_test <- pred_df_test_object$df_bin_centroids
avg_df_test <- pred_df_test_object$df_bin

eval_df_test <- generate_eval_df(data = data, prediction_df = pred_df_test, df_bin_centroids = centroid_df_test, df_bin = avg_df_test, num_bins = num_bins_x)
```

```{r}
#| echo: false

## Data set with all possible centroids in the hexagonal grid
full_centroid_df <- generate_full_grid_centroids(df_bin_centroids)

## Generate all coordinates of hexagons
hex_grid <- full_hex_grid(full_centroid_df)

#hex_full_count_df <- generate_full_grid_info(df_bin_centroids)

predict_hex_ID_plot <- ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
  geom_point(data = df_bin_centroids, aes(x = x, y = y), color = "#33a02c") +
  geom_point(data = predict_UMAP_df, aes(x = UMAP1, y = UMAP2), color = "red") +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
    theme_void() +
    theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
          axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
          legend.title = element_text(size=8), #change legend title font size
          legend.text = element_text(size=6)) +
    guides(fill = guide_colourbar(title = "Standardized count")) +
    annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3) 
```

<!--To draw predict values in hexbin grid-->

```{r}
#| echo: false
#| fig-cap: The .
#| label: fig-predict

predict_umap_plot  + predict_hex_ID_plot +
  plot_layout(guides='collect', ncol = 3) &
  theme(legend.position='bottom', plot.tag = element_text(size = 8))
```

```{r}
#| echo: false
#num_bins_vec <- append(sample(1:sqrt(750), 10), num_bins_x)

#num_bins_vec <- sample(2:1000, 40)
num_bins_vec <- 1:10

vec <- stats::setNames(rep("", 6), c("number_of_bins", "number_of_observations", "total_error", "totol_error_method_2", "totol_error_method_3", "total_mse"))  ## Define column names

eval_data_test <- dplyr::bind_rows(vec)[0, ]
eval_data_test <- eval_data_test |>
  dplyr::mutate_if(is.character, as.numeric)

eval_data_training <- dplyr::bind_rows(vec)[0, ]
eval_data_training <- eval_data_training |>
  dplyr::mutate_if(is.character, as.numeric)

for (i in 1:length(num_bins_vec)) {

  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = UMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value)
  pred_df_training <- pred_df_training_object$pred_data
  centroid_df_training <- pred_df_training_object$df_bin_centroids
  avg_df_training <- pred_df_training_object$df_bin

  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i])

  pred_df_test_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = predict_UMAP_df, num_bins = num_bins_vec[i], shape_val = shape_value)
  pred_df_test <- pred_df_test_object$pred_data
  centroid_df_test <- pred_df_test_object$df_bin_centroids
  avg_df_test <- pred_df_test_object$df_bin

  eval_df_test <- generate_eval_df(data = data, prediction_df = pred_df_test, df_bin_centroids = centroid_df_test, df_bin = avg_df_test, num_bins = num_bins_vec[i])

  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training) 
  eval_data_test <- dplyr::bind_rows(eval_data_test, eval_df_test)

}
```

```{r}
#| echo: false

eval_data_training <- eval_data_training |>
  mutate(data_type = "training")

eval_data_test <- eval_data_test |>
  mutate(data_type = "test")

MSE_df <- dplyr::bind_rows(eval_data_training, eval_data_test)
```

```{r}
#| echo: false
aic_plot <- ggplot(MSE_df |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
                   y = total_error,
                   color = data_type
)) +  
  geom_point() +
  geom_line() +
  #geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-15000, label=paste0("effective number of bins = ",as.character(NROW(full_grid_with_hexbin_id))), angle=90, size = 2.5) + 
  theme(legend.position = "none", legend.title=element_blank(), plot.title=element_text(size=8), axis.text = element_text(size = 5),
          axis.title = element_text(size = 7),
        legend.text = element_text(size=7)) +
  scale_color_manual(values = c("#d95f02")) + 
  ylab("AIC") +
  xlab("Total number of bins")
```

```{r}
#| echo: false
mse_plot <- ggplot(MSE_df, aes(x = number_of_bins,
                   y = total_mse,
                   color = data_type
)) +  
  geom_point() +
  geom_line() +
  #geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.05, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90, size = 2.5) +
  scale_color_manual(values = c("#1b9e77", "#d95f02")) + 
  theme(legend.position = "bottom", legend.title=element_blank(), plot.title=element_text(size=8), axis.text = element_text(size = 5),
          axis.title = element_text(size = 7),
        legend.text = element_text(size=7)) + 
  ylab("MSE") +
  xlab("Total number of bins")
```

```{r}
#| echo: false
#| fig-cap: The .
#| label: fig-diagnosticplt
#| out-width: 100%

aic_plot + mse_plot +
  plot_annotation(tag_levels = 'a') +
  plot_layout(guides='collect') &
  theme(legend.position='bottom', plot.tag = element_text(size = 8))
```

#### Goodness of fit statistics

In this analysis, we're looking at how well our model is doing in two aspects. First, we're checking the distances in our lower-dimensional space (see @fig-diag (a)). We want to see if the distances between different types of edges are similar. If they are, it means our model is doing a good job of keeping the relationships consistent. If there are big differences, it could mean our model is struggling with certain types of edges.

Next, we're examining how our model handles the relationship between high-dimensional and low-dimensional distances for each type of edge (see @fig-diag (b)). If our data points nicely line up around the expected line, it tells us that our model is preserving the original structure well. If they're scattered, it could mean our model is having trouble with specific data points. This two-part analysis helps us really understand how our model fits the data.

```{r}
#| echo: false

sorted_distance_df <- distance[order(distance$distance),] ## Sort the distances

sorted_distance_df <- cbind(sorted_distance_df, rbind(NA, apply(sorted_distance_df, 2, diff))) ## Calculate differences between unique distance
sorted_distance_df <- sorted_distance_df[, c(3,6)]
sorted_distance_df[is.na(sorted_distance_df)] <- 0.0000 ## To replace missing values with zero
names(sorted_distance_df)[2] <- "difference"

sorted_distance_df <- sorted_distance_df %>% 
  mutate(Edge_type = ifelse(distance >= benchmark, "Long edge", "Short edge"))

diagnostic_plot1 <- ggplot(sorted_distance_df, aes(x = distance, y = difference, color = Edge_type)) +
  geom_point(alpha = 0.5) +
  geom_rug(col="#e31a1c",alpha=0.1, linewidth=1.5) +
  #ggtitle("Visualization of difference between \n 2D distances by edge type.") +
  scale_color_manual(values=c("#33a02c", "#000000")) + 
  theme(legend.position = "bottom", legend.title=element_blank(), plot.title=element_text(size=8), axis.text = element_text(size = 5),
          axis.title = element_text(size = 7),
        legend.text = element_text(size=7)) + 
  xlab(expression(d^{(2)})) 
```

```{r}
#| echo: false

distance <- distance %>% 
  rename(distance_2D = distance)

#num_pca <- 4

high_d <- df_bin[, 1:7]

data_matrix <- data.matrix(high_d, rownames.force = NA)

# calculate Euclidean distance between each row in matrix
d <- dist(data_matrix)

df <- reshape2::melt(as.matrix(d), varnames = c("Start_point", "End_point"))

df <- df[df$Start_point > df$End_point,]

rownames(df) <- 1:nrow(df)

df <- df %>% 
  rename(distance_high_D = value)

high_low_dist <- left_join(distance, df, by = c("from" = "End_point", "to" = "Start_point"))
high_low_dist <- high_low_dist %>% distinct()
high_low_dist <- high_low_dist[complete.cases(high_low_dist),]

lm_fit <- lm(distance_high_D ~ distance_2D, data = high_low_dist)

high_low_dist <- high_low_dist %>% 
  mutate(Edge_type = ifelse(distance_2D >= benchmark, "Long edge", "Short edge"))

diagnostic_plot2 <- ggplot(high_low_dist, aes(x = distance_2D, y = distance_high_D, color = Edge_type)) +
  geom_point(alpha = 0.5) +
  geom_line(data = broom::augment(lm_fit),
            aes(x = distance_2D, y = .fitted),
            color="blue",
            linewidth=1) +
  geom_rug(col="#e31a1c",alpha=0.1, linewidth=1.5) +
  #ggtitle("Visualization of 2D and high-D distances \n by edge type.") +
  scale_color_manual(values=c("#33a02c","#000000")) + 
  theme(legend.position = "bottom", legend.title=element_blank(), plot.title=element_text(size=8), axis.text = element_text(size = 5),
          axis.title = element_text(size = 7),
        legend.text = element_text(size=7)) + 
  xlab(expression(d^{(2)})) + 
  ylab(expression(d^{(p)}))
```

```{r}
#| echo: false
#| fig-cap: The two diagnostic plot assess the effectiveness of the high-D representation algorithm  performed for the **s_curve_noise_umap** data set$\colon$ (a) visualizing the differences between low-D distances and (b) visualizing the relationship between high-D ($d^{(p)}$) and low-D distances ($d^{(2)}$). 
#| label: fig-diag
#| out-width: 100%

diagnostic_plot1 + diagnostic_plot2 +
  plot_annotation(tag_levels = 'a') +
  plot_layout(guides='collect') &
  theme(legend.position='bottom', plot.tag = element_text(size = 8))
```

<!--### Software

Our algorithm for high-dimensional (high-D) representation is available as an open-source R package named quollr. This package provides analysts with the capability to construct a low-dimensional manifold and seamlessly visualize it in the high-D space alongside the original data.-->

### Simulated data examples {#sec-simpleex}

In this section, we demonstrate the effectiveness of our approach through the use of simulation data. These simulations are designed to have known cluster structures and underlying data geometries. We begin with a straightforward example where we generate spherical clusters with a doublet structure. This allows us to assess how well our algorithm can identify and represent these known structures in the low-dimensional space.

We then proceed to a more complex example where the data exhibits a more intricate geometry. This simulation provides a challenging test for our approach to capture and preserve the underlying data structure accurately.

We *strongly* recommend viewing the linked videos for each study while reading. Links to the videos are available in the figures for each example. The videos show the visual appearance of the **langevitour** interface with low-dimensional view and how we can interact with the tour via the controls.

**Example 1: Exploring spherical Gaussian clusters with doublet structure**

The next data set consists of two distinct types of clusters embedded in a 10-$d$ space. The first set of clusters includes three Gaussian clusters with equal variance. Each of these Gaussian clusters is equidistant from the others, forming a well-separated and symmetrical arrangement. The second set of clusters consists of three clusters with a doublet structure. These clusters are also embedded in the 10-$d$ space, but each is positioned in the middle of its parent Gaussian clusters. Moreover, each doublet cluster contains an average number of points from its corresponding parent Gaussian clusters.

For analysis, we run tSNE on the training data with a perplexity setting 18. @fig-example3 (a) displays the tSNE results, indicating that the sub-clusters have been correctly identified. However, the relative locations of certain clusters to each other appear distorted in the low-dimensional view, such as the red and pink clusters being far apart.

Nonetheless, we find no apparent impact on the perception of the data structure when we examine the tour view and the model (see @fig-example3) through the linked video (see @fig-exp1_sc). The model demonstrates accurate identification of clusters and captures the internal variations within each cluster, reflecting the inherent variance present in the data. The configurations used for the analysis are detailed in @tbl-table02. Additionally, the representation of the parent clusters in the low-dimensional space appears stretched, aligning with the variance of these clusters. On the other hand, the doublet clusters are squeezed, reflecting their specific characteristics and relationship to their parent Gaussian clusters.

```{r}
#| warning: false
#| echo: false
sample_size <- 450
cluster_size <- sample_size/4.5

df1 <- tibble::tibble(x1=rnorm(cluster_size, mean = 3, sd = 0.05), x2 = rnorm(cluster_size, mean = 1, sd = 0.05), x3=rnorm(cluster_size, mean = 1, sd = 0.05), x4=rnorm(cluster_size, mean = 1, sd = 0.05),
                      x5=rnorm(cluster_size, mean = 1, sd = 0.05),
                      x6=rnorm(cluster_size, mean = 1, sd = 0.05),
                      x7=rnorm(cluster_size, mean = 1, sd = 0.05),
                      x8=rnorm(cluster_size, mean = 1, sd = 0.05),
                      x9=rnorm(cluster_size, mean = 1, sd = 0.05),
                      x10=rnorm(cluster_size, mean = 1, sd = 0.05))

df1 <- df1 %>%
  mutate(cluster = 1)

df2 <- tibble::tibble(x1=rnorm(cluster_size, mean = 1, sd = 0.05), x2=rnorm(cluster_size, mean = 1, sd = 0.05), x3=rnorm(cluster_size, mean = 1, sd = 0.05), x4=rnorm(cluster_size, mean = 1, sd = 0.05),
                      x5=rnorm(cluster_size, mean = 1, sd = 0.05),
                      x6=rnorm(cluster_size, mean = 1, sd = 0.05),
                      x7=rnorm(cluster_size, mean = 1, sd = 0.05),
                      x8=rnorm(cluster_size, mean = 1, sd = 0.05),
                      x9=rnorm(cluster_size, mean = 1, sd = 0.05),
                      x10=rnorm(cluster_size, mean = 1, sd = 0.05))

df2 <- df2 %>%
  mutate(cluster = 2)

df3_new <- (df1 + df2) / 2

samp <- sample(nrow(df3_new), cluster_size * 0.50) ## 50% from the original dataset

# data in the sample
df3 <- df3_new[samp,]

df3 <- df3 %>%
  mutate(cluster = 3)

df4 <- tibble::tibble(x1=rnorm(cluster_size, mean = 1, sd = 0.05), x2=rnorm(cluster_size, mean = 1, sd = 0.05), x3=rnorm(cluster_size, mean = 1, sd = 0.05), x4=rnorm(cluster_size, mean = 3, sd = 0.05),
                      x5=rnorm(cluster_size, mean = 1, sd = 0.05),
                      x6=rnorm(cluster_size, mean = 1, sd = 0.05),
                      x7=rnorm(cluster_size, mean = 1, sd = 0.05),
                      x8=rnorm(cluster_size, mean = 1, sd = 0.05),
                      x9=rnorm(cluster_size, mean = 1, sd = 0.05),
                      x10=rnorm(cluster_size, mean = 1, sd = 0.05))

df4 <- df4 %>%
  mutate(cluster = 4)

df5_new <- (df2 + df4) / 2

samp1 <- sample(nrow(df5_new), cluster_size * 0.50) ## 50% from the original dataset

df5 <- df5_new[samp1,]

df5 <- df5 %>%
  mutate(cluster = 5)

df6_new <- (df1 + df4) / 2

samp2 <- sample(nrow(df6_new), cluster_size * 0.50) ## 50% from the original dataset

df6 <- df6_new[samp2,]

df6 <- df6 %>%
  mutate(cluster = 6)

df_3 <- bind_rows(df1, df2, df3, df4, df5, df6)
```

```{r}
#| warning: false
#| echo: false

# langevitour::langevitour(df_3[, 1:10], group = df_3$cluster, levelColors = c("#b15928", "#1f78b4", "#bc80bd", "#33a02c", "#fb9a99", "#e31a1c"))
```

```{r}
#| warning: false
#| echo: false

df_3 <- df_3 |>
  mutate(ID = row_number())

data_split_sp <- initial_split(df_3)
training_data_5 <- training(data_split_sp) |>
  arrange(ID)
test_data_5 <- testing(data_split_sp) |>
  arrange(ID)

clusters <- training_data_5$cluster
```

```{r}
#| warning: false
#| echo: false 
tSNE_data_gau <- Fit_tSNE(training_data_5 |> dplyr::select(-ID, -cluster), with_seed = 20230531)
```

```{r}
#| warning: false
#| echo: false

tSNE_cluster <- tSNE_data_gau %>%
  ggplot(aes(x = tSNE1,
             y = tSNE2, color = as.character(clusters))) +
  geom_point(alpha=0.5) +
  coord_equal() +
  theme_linedraw() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold"),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          axis.text = element_text(size = 5),
          axis.title = element_text(size = 7), #change legend key width
        legend.title = element_text(size=7), #change legend title font size
        legend.text = element_text(size=7)) +
  scale_color_manual(values=c("#b15928", "#1f78b4", "#bc80bd", "#33a02c", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#cab2d6", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
  guides(color = guide_legend(title = "Cluster"))

```

```{r}
#| warning: false
#| echo: false

cell_area <- 1

num_bins_tsne_gau <- calculate_effective_x_bins(.data = tSNE_data_gau, x = tSNE1,
                                                cell_area = 1)
shape_val_tsne_gau <- calculate_effective_shape_value(.data = tSNE_data_gau,
                                                      x = tSNE1, y = tSNE2)

## To extract bin centroids
hexbin_data_object_tsne_gau <- extract_hexbin_centroids(nldr_df = tSNE_data_gau, num_bins = num_bins_tsne_gau, shape_val = shape_val_tsne_gau, x = tSNE1, y = tSNE2)

df_bin_centroids_tsne_gau <- hexbin_data_object_tsne_gau$hexdf_data

# min_std_cell_threshold3 <- 0.01
# 
# df_bin_centroids_pbmc <- df_bin_centroids_pbmc |>
#   dplyr::mutate(stand_cell_count = counts/max(counts)) |>
#   dplyr::filter(stand_cell_count > min_std_cell_threshold3)

tSNE_data_with_hb_id_gau <- tSNE_data_gau |> 
  dplyr::mutate(hb_id = hexbin_data_object_tsne_gau$hb_data@cID)

## To generate a data set with high-D and 2D training data
df_all_tsne_gau <- dplyr::bind_cols(training_data_5 |> dplyr::select(-ID), tSNE_data_with_hb_id_gau)

## Averaged on high-D
df_bin_tsne_gau <- avg_highD_data(.data = df_all_tsne_gau)

## Triangulate bin centroids
tr1_object_tsne_gau <- triangulate_bin_centroids(df_bin_centroids_tsne_gau, x, y)
tr_from_to_df_tsne_gau <- generate_edge_info(triangular_object = tr1_object_tsne_gau)

## Compute 2D distances
distance_tsne_gau <- cal_2D_dist(.data = tr_from_to_df_tsne_gau)

## To find the benchmark value
benchmark_tsne_gau <- find_benchmark_value(.data = distance_tsne_gau, distance_col = distance)
#benchmark_tsne <- 5

tour_tsne_gau <- show_langevitour(df_all_tsne_gau, df_bin_tsne_gau, df_bin_centroids_tsne_gau, benchmark_value = benchmark_tsne_gau, distance = distance_tsne_gau, distance_col = distance)

trimesh_cluster <- ggplot(df_bin_centroids_tsne_gau, aes(x = x, y = y)) +
  geom_trimesh() +
  coord_equal()

trimesh_cluster <- trimesh_cluster +
  #ggtitle("(b)") +
  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  theme(axis.text = element_text(size = 5),
        axis.title = element_text(size = 7))

trimesh_gr_cluster <- colour_long_edges(.data = distance_tsne_gau, benchmark_value = benchmark_tsne_gau,
                                     triangular_object = tr1_object_tsne_gau, distance_col = distance)

trimesh_gr_cluster <- trimesh_gr_cluster +
  #ggtitle("(c)") +
  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  theme(axis.text = element_text(size = 5),
        axis.title = element_text(size = 7), #change legend key width
        legend.title = element_text(size=7), #change legend title font size
        legend.text = element_text(size=5))

tour_cluster <- show_langevitour(df_all_tsne_gau, df_bin_tsne_gau, df_bin_centroids_tsne_gau, benchmark_value = benchmark_tsne_gau, distance = distance_tsne_gau, distance_col = distance)
```

```{r}
#| warning: false
#| echo: false
#| out-width: 100%
#| label: fig-example3
#| fig-cap: Visualization and refinement of low-dimensional representation for sub-clustered data$\colon$ (a) 2D layout of tSNE (perplexity = 18), (b) triangular mesh, and (c) triangular mesh colored by edge type. Edges with lengths less than the benchmark value of $2$ are classifiedas small, while edges with lengths greater than or equal to this value are labeled as long. The combination of tSNE, hexagonal binning, and triangulation provides the effectiveness of our approach in visualizing and refining the low-dimensional representation of the sub-clustered data.

tSNE_cluster + trimesh_cluster + trimesh_gr_cluster +
  plot_annotation(tag_levels = 'a') +
  plot_layout(guides='collect') &
  theme(legend.position='bottom', plot.tag = element_text(size = 8))
```

::: {#fig-exp1_sc layout-ncol="3"}
![](figures/example_1a.png){#fig-exp1_sc1}

![](figures/example_1b.png){#fig-exp1_sc2}

![](figures/example_1c.png){#fig-exp1_sc3}

Screen shots of the **langevitour** with the low-dimensional view applied to sub-clustered data set, a video of the tour animation is available at (\url{https://drive.google.com/file/d/15nAbmc4UoVlUyFcGtnlnmyxrJEbK39nx/view}). Combination of the **langevitour** and low-dimensional representation of the sub-clustered data provides the effectiveness of our approach.
:::

**Example 2: Exploring data with piecewise linear structure**

In our next exploration, we delve into simulated noisy tree-structured data designed to mimic branching trajectories of cell differentiation. The simulation is crafted to resemble the hierarchical clustering pattern observed in cell differentiation trajectories, where mature cells are akin to the tips of branches.

To begin our analysis, we apply principal components to the data and focus on the first ten principal components, collectively accounting for approximately 68% of the variance explained in the data set. Then, we run TriMAP with the default parameters on the training data (n_inliers: 5, n_outliers: 4, and n_random: 3).  Through this process, we create a low-dimensional manifold using the configurations detailed in @tbl-table02. Finally, we present the model along with the data using the tour view (Figure @fig-exp2_sc), allowing us to visualize and explore the intricate relationships and structures present in the data's low-dimensional representation.

```{r}
#| warning: false
#| echo: false

data_phate <- read_csv(paste0(here::here(), "/data/data_phate.csv"))

data_phate <- data_phate |>
  dplyr::mutate(ID = row_number())
```

```{r}
#| warning: false
#| echo: false 
data_split <- initial_split(data_phate)
training_data <- training(data_split) |>
  arrange(ID)
test_data <- testing(data_split) |>
  arrange(ID)
```

```{r}
#| warning: false
#| echo: false 
## num_pcs: selected number of PCs to get results (deafukt is all)

calculate_pca <- function(data, num_pcs = NCOL(data), center = TRUE, scale = FALSE){
  pcaY_cal <- prcomp(data, center = TRUE, scale = FALSE)

  PCAresults <- tibble::as_tibble(pcaY_cal$x[, 1:num_pcs])
  names(PCAresults) <- paste0(rep("PC", num_pcs), 1:num_pcs)

  summary_pca <- summary(pcaY_cal)

  var_explained_df <- tibble::tibble(PC = paste0("PC",1:NCOL(data)),
                               var_explained = (pcaY_cal$sdev[1:NCOL(data)])^2/sum((pcaY_cal$sdev[1:NCOL(data)])^2))

  ## Scree plot
  plot <- var_explained_df %>%
    ggplot(aes(x = PC,y = var_explained, group = 1))+
    geom_point(size=1)+
    geom_line() +
    theme_minimal() +
    scale_x_discrete(limits = paste0(rep("PC", 20), 1:20)) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
          axis.text = element_text(size = 5),
          axis.title = element_text(size = 7)) +
    xlab("Principal Components (PC)") +
    ylab("Variance explained")

  return(list(prcomp_out = pcaY_cal, pca_components = PCAresults, summary = summary_pca, var_explained_pca  = var_explained_df, plot = plot))
}
```

```{r}
#| warning: false
#| echo: false 
pca_out <- calculate_pca(training_data, 10)
data_pca <- pca_out$pca_components
data_pca <- data_pca %>% 
  distinct()

var_explain_example2 <- pca_out$plot
```

```{r}
#| warning: false
#| echo: false 
TriMAP_data_tree <- read_csv(paste0(here::here(), "/data/trimap_data_linear_structure_pca.csv")) %>%
    mutate(ID = row_number())

```

```{r}
#| warning: false
#| echo: false 
TriMAP_tree <- plot_TriMAP_2D(TriMAP_data_tree) +
  theme_linedraw() +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

```{r}
#| warning: false
#| echo: false

cell_area <- 1

num_bins_tree <- calculate_effective_x_bins(.data = TriMAP_data_tree, x = TriMAP1,
                                            cell_area = 1)
shape_val_tree <- calculate_effective_shape_value(.data = TriMAP_data_tree,
                                                  x = TriMAP1, y = TriMAP2)

## To extract bin centroids
hexbin_data_object_tree <- extract_hexbin_centroids(nldr_df = TriMAP_data_tree, num_bins = num_bins_tree, shape_val = shape_val_tree, x = TriMAP1, y = TriMAP2)

df_bin_centroids_tree <- hexbin_data_object_tree$hexdf_data

# min_std_cell_threshold3 <- 0.01
# 
# df_bin_centroids_tree <- df_bin_centroids_tree |>
#   dplyr::mutate(stand_cell_count = counts/max(counts)) |>
#   dplyr::filter(stand_cell_count > min_std_cell_threshold3)

TriMAP_data_with_hb_id_tree <- TriMAP_data_tree |> 
  dplyr::mutate(hb_id = hexbin_data_object_tree$hb_data@cID)

## To generate a data set with high-D and 2D training data
data_pca_n <- data_pca
names(data_pca_n) <- paste0(rep("x", 7), 1:7)
df_all_tree <- dplyr::bind_cols(data_pca_n, TriMAP_data_with_hb_id_tree)

## Averaged on high-D
df_bin_tree <- avg_highD_data(.data = df_all_tree)

## Triangulate bin centroids
tr1_object_tree <- triangulate_bin_centroids(df_bin_centroids_tree, x, y)
tr_from_to_df_tree <- generate_edge_info(triangular_object = tr1_object_tree)

## Compute 2D distances
distance_tree <- cal_2D_dist(.data = tr_from_to_df_tree)

## To find the benchmark value
benchmark_tree <- find_benchmark_value(.data = distance_tree, distance_col = distance)

benchmark_tree <- 2

names(df_all_tree)[1:7] <- paste0(rep("PC_", 7), 1:7)
names(df_bin_tree) <- c("hb_id", paste0(rep("PC_", 7), 1:7))

# tour_tree <- show_langevitour(df_all_tree, df_bin_tree, df_bin_centroids_tree, benchmark_value = benchmark_tree, distance = distance_tree, distance_col = distance, col_start = "PC")

trimesh_tree <- ggplot(df_bin_centroids_tree, aes(x = x, y = y)) +
  geom_trimesh() +
  coord_equal()

trimesh_tree <- trimesh_tree +
  #ggtitle("(b)") +
  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  theme(axis.text = element_text(size = 5),
        axis.title = element_text(size = 7))

trimesh_gr_tree <- colour_long_edges(.data = distance_tree, benchmark_value = benchmark_tree,
                                     triangular_object = tr1_object_tree, distance_col = distance)

trimesh_gr_tree <- trimesh_gr_tree +
  #ggtitle("(c)") +
  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  theme(axis.text = element_text(size = 5),
        axis.title = element_text(size = 7), #change legend key width
        legend.title = element_text(size=7), #change legend title font size
        legend.text = element_text(size=5))

tour_tree <- show_langevitour(df_all_tree, df_bin_tree, df_bin_centroids_tree, benchmark_value = benchmark_tree, distance = distance_tree, distance_col = distance, col_start = "PC")

```

```{r}
#| warning: false
#| echo: false
#| out-width: 100%
#| label: fig-example2
#| fig-pos: H
#| fig-cap: Visualization and refinement of low-dimensional representation for tree structured data$\colon$ (a) 2D layout of TriMAP (n-inliers = 5, n-outliers = 4, n-random = 3), (b) triangular mesh, and (c) triangular mesh colored by edge type. Edges with lengths less than the benchmark value of $2$ are classified as small, while edges with lengths greater than or equal to this value are labeled as long. The combination of TriMAP, hexagonal binning, and triangulation provides the effectiveness of our approach in visualizing and refining the low-dimensional representation of the tree structured data.

TriMAP_tree + trimesh_tree + trimesh_gr_tree +
  plot_annotation(tag_levels = 'a') +
  plot_layout(guides='collect') &
  theme(legend.position='bottom', plot.tag = element_text(size = 8))
```

::: {#fig-exp2_sc layout-ncol="3"}
![](figures/example_2a.png)

![](figures/example_2b.png)

![](figures/example_2c.png)

Screen shots of the **langevitour** with the low-dimensional view applied to tree structured data set, a video of the tour animation is available at (\url{https://drive.google.com/file/d/1YQx4UGlzDify1mGqeDtlqTHncgqidDhq/view}). Combination of the **langevitour** and low-dimensional representation of the tree structured data provides the effectiveness of our approach.
:::

## Applications {#sec-applications}

### Single-cell RNA-seq data of human

In the field of single-cell studies, a common analysis task involves clustering to identify groups of cells with similar expression profiles. Analysts often turn to NLDR techniques to verify and identify these clusters and explore developmental trajectories (e.g., example 1). In clustering workflows, the main objective is to verify the existence of clusters and subsequently identify them as specific cell types by examining the expression of "known" marker genes. In this context, a "faithful" embedding should ideally preserve the topology of the data, ensuring that cells corresponding to the same cell type are situated close to the high-dimensional space.

To begin our analysis, we installed the Peripheral Blood Mononuclear Cells (pbmc) data set obtained from 10x Genomics using the `SeuratData` R package [@Rahul2019], which facilitates the distribution of data sets in the form of Seurat objects [@Yuhan2021]. This data set contains 13,714 features across 2,700 samples within a single assay. The active assay is RNA, with 13,714 features representing different gene expressions. After loading the data set, we obtained the principal components (PCs) and assessed the variance explained by each PC. Based on this evaluation, we selected seven PCs, representing approximately 50% of the variance in gene expression, for further analysis.

Next, we employed the UMAP technique with default parameter settings. As illustrated in @fig-pbmc, the cell types B and Platelet are well-separated in the UMAP layout. Moreover, CD14+ Mono, FCGR3A+ Mono, and DC form a distinct cluster, while Naive CD4 T, NK, Memory CD4 T, and CD8 T are grouped together in another cluster. The values utilized to construct the smooth low-dimensional manifold are presented in @tbl-table02. The linked video, demonstrating the tour with the model, showcases the generation of a smooth surface for this application, enabling a comprehensive exploration of the data's structure and relationships (see @fig-pbmc_sc).

```{r}
#| warning: false
#| echo: false
# InstallData("pbmc3k")
# data("pbmc3k.final")
# 
# data_pca <- as.data.frame(pbmc3k.final@reductions$pca@cell.embeddings)
# names(data_pca) <- paste0(rep("PC", ncol(data_pca)), 1:ncol(data_pca))
# 
# data_pca_with_label <- bind_cols(pbmc3k.final@reductions$pca@cell.embeddings, pbmc3k.final@meta.data$seurat_clusters)
# names(data_pca_with_label)[ncol(pbmc3k.final@reductions$pca@cell.embeddings) + 1] <- "cluster_label"
# 
# levels(data_pca_with_label$cluster_label) <- c("Naive CD4 T", "Memory CD4 T", "CD14+ Mono", "B", "CD8 T", "FCGR3A+ Mono", "NK", "DC", "Platelet")

data_pca_with_label <- readRDS(file = paste0(here::here(), "/data/pbmc_with_label.rds"))

data_pca <- data_pca_with_label %>%
  select(-cluster_label)
```

```{r}
#| warning: false
#| echo: false
# eigs <- pbmc3k.final@reductions$pca@stdev^2 ## To calculate eigen values
# 
# summary_pca <- rbind(
#   SD = sqrt(eigs),
#   Proportion = eigs/sum(eigs),
#   Cumulative = cumsum(eigs)/sum(eigs))
# 
# var_explained_df <- data.frame(PC = paste0("PC", 1:ncol(pbmc3k.final@reductions$pca@cell.embeddings)),
#                                var_explained = (pbmc3k.final@reductions$pca@stdev)^2/sum((pbmc3k.final@reductions$pca@stdev)^2))

var_explained_df <- readRDS(file = paste0(here::here(), "/data/pbmc_var_explained.rds"))

pbmc_var_explain <- var_explained_df %>%
    ggplot(aes(x = PC,y = var_explained, group = 1))+
    geom_point(size=1)+
    geom_line() +
    theme_minimal() +
    scale_x_discrete(limits = paste0(rep("PC", 20), 1:20)) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
          axis.text = element_text(size = 5),
          axis.title = element_text(size = 7)) +
    xlab("Principal Components (PC)") +
    ylab("Variance explained")
```

```{r}
#| warning: false
#| echo: false

data_pca <- data_pca[,1:7]
UMAP_fit_pbmc <- umap(data_pca, n_neighbors = 15, n_components =  2)

UMAP_data_pbmc <- UMAP_fit_pbmc$layout |>
  as.data.frame()

names(UMAP_data_pbmc)[1:(ncol(UMAP_data_pbmc))] <- paste0(rep("UMAP",(ncol(UMAP_data_pbmc))), 1:(ncol(UMAP_data_pbmc)))

UMAP_data_pbmc <- UMAP_data_pbmc |>
  mutate(ID = row_number())

plot_pbmc <- UMAP_data_pbmc %>%
    ggplot(aes(x = UMAP1,
               y = UMAP2, color = data_pca_with_label$cluster_label))+
    geom_point(alpha=0.5) +
    coord_equal() +
    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
  theme_linedraw() +
    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
         legend.key.height = unit(0.25, 'cm'),
         legend.key.width = unit(0.25, 'cm')) +
  scale_color_manual(values=c("#b15928", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#cab2d6", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#bc80bd", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
  guides(color = guide_legend(title = "Cell type"),
         color = guide_legend(nrow = 3))
```

```{r}
#| echo: false
cell_area <- 1

num_bins_pbmc <- calculate_effective_x_bins(.data = UMAP_data_pbmc, x = UMAP1,
                                        cell_area = 1)
shape_val_pbmc <- calculate_effective_shape_value(.data = UMAP_data_pbmc,
                                              x = UMAP1, y = UMAP2)

## To extract bin centroids
hexbin_data_object_pbmc <- extract_hexbin_centroids(nldr_df = UMAP_data_pbmc, num_bins = num_bins_pbmc, shape_val = shape_val_pbmc)

df_bin_centroids_pbmc <- hexbin_data_object_pbmc$hexdf_data

# min_std_cell_threshold3 <- 0.01
# 
# df_bin_centroids_pbmc <- df_bin_centroids_pbmc |>
#   dplyr::mutate(stand_cell_count = counts/max(counts)) |>
#   dplyr::filter(stand_cell_count > min_std_cell_threshold3)

UMAP_data_with_hb_id_pbmc <- UMAP_data_pbmc |> 
  dplyr::mutate(hb_id = hexbin_data_object_pbmc$hb_data@cID)

## To generate a data set with high-D and 2D training data
data_pca_n <- data_pca
names(data_pca_n) <- paste0(rep("x", 7), 1:7)
df_all_pbmc <- dplyr::bind_cols(data_pca_n, UMAP_data_with_hb_id_pbmc)

## Averaged on high-D
df_bin_pbmc <- avg_highD_data(.data = df_all_pbmc)

## Triangulate bin centroids
tr1_object_pbmc <- triangulate_bin_centroids(df_bin_centroids_pbmc, x, y)
tr_from_to_df_pbmc <- generate_edge_info(triangular_object = tr1_object_pbmc)

## Compute 2D distances
distance_pbmc <- cal_2D_dist(.data = tr_from_to_df_pbmc)

## To find the benchmark value
benchmark_pbmc <- find_benchmark_value(.data = distance_pbmc, distance_col = distance)

benchmark_pbmc <- 2

names(df_all_pbmc)[1:7] <- paste0(rep("PC_", 7), 1:7)
names(df_bin_pbmc) <- c("hb_id", paste0(rep("PC_", 7), 1:7))

# tour_pbmc <- show_langevitour(df_all_pbmc, df_bin_pbmc, df_bin_centroids_pbmc, benchmark_value = benchmark_pbmc, distance = distance_pbmc, distance_col = distance, col_start = "PC")

df_1 <- df_all_pbmc |> 
  dplyr::select(tidyselect::starts_with("PC")) |>
  dplyr::mutate(type = data_pca_with_label$cluster_label) ## original dataset

df_b_1 <- df_bin_pbmc |> 
  dplyr::filter(hb_id %in% df_bin_centroids_pbmc$hexID) |>
  dplyr::select(-hb_id) |>
  dplyr::mutate(type = "model") ## Data with summarized mean

df_exe <- dplyr::bind_rows(df_b_1, df_1)

## Set the maximum difference as the criteria
distance_df_small_edges <- distance_pbmc %>% 
  dplyr::filter(distance < benchmark_pbmc)
## Since erase brushing is considerd.

trimesh_pbmc <- ggplot(df_bin_centroids_pbmc, aes(x = x, y = y)) +
  geom_trimesh() +
  coord_equal()

trimesh_pbmc <- trimesh_pbmc +
  #ggtitle("(b)") +
  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  theme(axis.text = element_text(size = 5),
          axis.title = element_text(size = 7))

trimesh_gr_pbmc <- colour_long_edges(.data = distance_pbmc, benchmark_value = benchmark_pbmc,
                 triangular_object = tr1_object_pbmc, distance_col = distance)

trimesh_gr_pbmc <- trimesh_gr_pbmc +
  #ggtitle("(c)") +
  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  theme(axis.text = element_text(size = 5),
          axis.title = element_text(size = 7), #change legend key width
        legend.title = element_text(size=7), #change legend title font size
        legend.text = element_text(size=5))

tour_pbmc <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)], lineFrom = distance_df_small_edges$from, lineTo = distance_df_small_edges$to, group = df_exe$type, levelColors = c("#33a02c", "#b2df8a", "#fb9a99", "#ff7f00", "#e31a1c", "#1f78b4", "black", "#b15928", "#6a3d9a", "#cab2d6"), pointSize = 3)
```

```{r}
#| warning: false
#| echo: false
#| out-width: 100%
#| label: fig-pbmc
#| fig-pos: H
#| fig-cap: Visualization and refinement of low-dimensional representation for pbmc data$\colon$ (a) 2D layout of UMAP (n-neigbors = 15), (b) triangular mesh, and (c) triangular mesh colored by edge type. Edges with lengths less than the benchmark value of $2$ are classified as small, while edges with lengths greater than or equal to this value are labeled as long. The combination of UMAP, hexagonal binning, and triangulation provides the effectiveness of our approach in visualizing and refining the low-dimensional representation of the pbmc data.

plot_pbmc + trimesh_pbmc + trimesh_gr_pbmc + #A = 1, b = (19, 27), s = 1.255103
  plot_annotation(tag_levels = 'a') +
  plot_layout(guides='collect') &
  theme(legend.position='bottom', plot.tag = element_text(size = 8))
```

::: {#fig-pbmc_sc layout-ncol="3"}
![](figures/pbmc_a.png){#fig-pbmc_sc1}

![](figures/pbmc_b.png){#fig-pbmc_sc2}

![](figures/pbmc_c.png){#fig-pbmc_sc3}

Screen shots of the **langevitour** with the low-dimensional view applied to pbmc data set, a video of the tour animation is available at (\url{https://drive.google.com/file/d/1F7Xb2EIpfUruCxYihPw2rKro4FB5Y2uL/view}). Combination of the **langevitour** and low-dimensional representation of the pbmc data provides the effectiveness of our approach.
:::

### MNIST

The MNIST dataset is a widely recognized benchmark in the realm of computer vision and machine learning, which is a collection of 28x28 pixel grayscale images of handwritten digits (0 through 9), serving as a foundational resource for developing and evaluating various image classification algorithms. With its simplicity and clarity, MNIST provides an ideal platform for researchers and practitioners to experiment with and benchmark their models.

## Discussion {#sec-discussion}

Our research introduces a comprehensive framework that leverages tours for interactive exploration of high-dimensional data coupled with a low-dimensional manifold, facilitated by the `quollr` R package. Regardless of the Non-Linear Dimension Reduction (NLDR) technique in use, our approach demonstrates effectiveness through simulation examples, particularly in the iterative removal of long edges for a smoother representation and capturing cluster variance.

In the example with doublets, our method successfully captures the tweak within each cluster, indicating the variance present within them. However, the model may not appear smooth in high-dimensional space due to considerable noise when the data has a piecewise linear geometry, such as the tree simulation.

The practical application of our framework, as showcased with the UMAP view, enables visual inspection of well-separated clusters. Furthermore, the combined tour and model provide a robust assessment of whether UMAP preserves the data structure and accurately transforms the data.

The advantages of our approach include its versatility across various NLDR techniques and the ability to generate interactive visualizations for detailed exploration. The tour provides an intuitive way to navigate and comprehend high-dimensional data while assessing the accuracy of dimensionality reduction.

However, one limitation is that the approach may be less effective in cases with significant noise, as seen in the tree simulation example. Additionally, while our method aids in visual verification, quantifying the accuracy of embeddings might require further evaluation metrics.

In conclusion, our framework presents a powerful tool for researchers and analysts in single-cell studies to assess their embeddings by visually inspecting them alongside the original data. By leveraging the advantages of tours and low-dimensional manifolds, our approach offers valuable insights into the data transformation process, empowering users to make informed decisions in analyzing high-dimensional data. Future work could enhance the method's robustness in the presence of noise and explore additional evaluation metrics for quantifying embedding accuracy.

## References {.unnumbered}

::: {#refs}
:::

{{< pagebreak >}}

### 2024-01-09 paper.tex
Initial commit

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt]{article}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-1in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.7in}%
\addtolength{\topmargin}{-1in}%

\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{agsm}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Visualising How Non-linear Dimension Reduction Warps Your Data},
  pdfauthor={Jayani P.G. Lakshika; Dianne Cook; Paul Harrison; Michael Lydeamore; Thiyanga S. Talagala},
  pdfkeywords={high-dimensional data, dimension
reduction, triangulation, hexagonal binning, low-dimensional
manifold, manifold learning, tour, data vizualization},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\begin{document}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{\bf Visualising How Non-linear Dimension Reduction Warps Your
Data}
\author{
Jayani P.G. Lakshika\\
Econometrics \& Business Statistics, Monash University\\
and\\Dianne Cook\\
Econometrics \& Business Statistics, Monash University\\
and\\Paul Harrison\\
MGBP, BDInstitute, Monash University\\
and\\Michael Lydeamore\\
Econometrics \& Business Statistics, Monash University\\
and\\Thiyanga S. Talagala\\
Statistics, University of Sri Jayewardenepura\\
}
\maketitle

\bigskip
\bigskip
\begin{abstract}
Non-Linear Dimension Reduction (NLDR) techniques have emerged as
powerful tools to visualize high-dimensional data. However, their
complexity and parameter choices may lead to distrustful or misleading
results. To address this challenge, we propose a novel approach that
combines the tour technique with a low-dimensional manifold generated
using NLDR techniques, hexagonal binning, and triangulation. This
integration enables a clear examination of the low-dimensional
representation in the original high-dimensional space. Our approach not
only preserves the advantages of both tours and NLDR but also offers a
more intuitive perception of complex data structures and facilitates
accurate data transformation assessments. The method and example data
sets are available in the \textbf{quollr} R package.
\end{abstract}

\noindent%
{\it Keywords:} high-dimensional data, dimension
reduction, triangulation, hexagonal binning, low-dimensional
manifold, manifold learning, tour, data vizualization
\vfill

\newpage
\spacingset{1} % DON'T change the spacing! (Default 1.9)

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, sharp corners, boxrule=0pt, interior hidden, frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable]}{\end{tcolorbox}}\fi

\hypertarget{sec-intro}{%
\section{Introduction}\label{sec-intro}}

High-dimensional (high-D) data is widespread in many fields including
ecology and bioinformatics \citep{Guo2023}, in part because of new data
collection technologies \citep{Johnstone2009, ayesha2020overview}.
Working with high-dimensional data poses considerable challenges due to
the difficulty in visualizing beyond two dimensions \citep{Jia2022}.
High-dimensional data also presents difficulties for model fitting
\citep{Johnstone2009}, both computationally and interpretation, each of
which benefits from visualization.

To create visual representations of high-dimensional data, it is common
to apply dimension reduction techniques. Linear methods such as
principal component analysis (PCA) \citep{Karl1901} have been used for
many years, and remain popular. Non-linear methods such as
multi-dimensional scaling (MDS) \citep{Torgerson1967} have also been
routinely used. In the past decade, there has merged many new techniques
non-linear dimension reduction (NLDR), such as t-distributed stochastic
neighbor embedding (tSNE) \citep{Laurens2008}, uniform manifold
approximation and projection (UMAP) \citep{Leland2018}, designed to
capture the complex and non-linear relationships present within
high-dimensional data \citep{Johnstone2009}.

However, projecting high-dimensional data has limitations, such as
information loss and potential distortion of essential structures and
patterns \citep[\citet{article53}]{Jia2022}. The choice of technique and
parameters further impacts the accuracy of the visualization,
necessitating careful consideration for meaningful interpretation (see
Figure~\ref{fig-nldervis}).

Interactive and dynamic graphics systems have also been developed over
the years to enable visualizing high dimensions. One method, called a
tour \citep{Asimov1985}, shows a sequence of linear projections is shown
as a movie, allowing exploration without warping the space
\citep{lee2021review}. Interactive tools like \textbf{XGobi} and
\textbf{GGobi} have been successful in incorporating tours for exploring
high-dimensional data \citep{article60}. The R package \textbf{tourr}
\citep{article61} further enhances tour visualization within R, although
it may face limitations in frame rate and interactive features compared
to \textbf{GGobi}.

To overcome these limitations, the R package \textbf{detourr}
\citep{article22} has been developed, leveraging a Javascript widget via
htmlwidgets \citep{Ramnath2023} to achieve higher frame rates and
enhanced interactivity. Additionally, the R package \textbf{langevitour}
\citep{article09} utilizes Langevin Dynamics to generate a continuous
path of projections, eliminating the need for interpolation between
projections for animation. The tour technique has proven valuable in
exploring statistical model fits \citep{article58} and factorial
experimental designs \citep{article59}. Augmenting the results of
non-linear dimensional reduction methods with the tour, as demonstrated
in the \textbf{liminal} R package \citep{article21}, further enhances
data exploration.

While tours \citep{Asimov1985} preserve space without warping
\citep{lee2021review}, they require integrating multiple low-dimensional
views mentally to perceive high-dimensional structures. To address this
challenge, we propose a novel approach by combining the tour technique
with a low-dimensional manifold. This manifold is created through the
synergistic use of Non-Linear Dimension Reduction (NLDR) techniques,
hexagonal binning, and triangulation. By merging these techniques, our
approach offers a comprehensive and efficient means to visualize and
explore high-dimensional data while retaining the advantages of both
tours and NLDR. This integration facilitates a more intuitive perception
of complex data structures and empowers analysts with a robust tool for
assessing the accuracy of data transformations. The implementation of
our approach is available as an R package called \textbf{quollr}.

The outline of this paper is as follows. The
Section~\ref{sec-background} provides an detailed overview of dimension
reduction methods, triangulation, and tours. Building upon this
foundation, the Section~\ref{sec-methods} delves into the proposed
algorithm, \textbf{quollr}, and its implementation details. In
\textbf{?@sec-prediction}, discusses the effectiveness of the learned
low-dimensional manifold in accurately representing the complex
high-dimensional data. Following that, Section~\ref{sec-simpleex}
presents simple examples from simulations to illustrate the
functionality of the algorithm. Subsequently,
Section~\ref{sec-applications} showcases real-world applications of
\textbf{quollr} on different data sets, particularly in single-cell
RNA-seq data. These applications reveal insights into the performance
and trustworthiness of NLDR algorithms. We analyze the results to
identify situations where NLDR techniques may lead to misleading
interpretations. Finally, \textbf{?@sec-conclusions} concludes by
summarizing the findings and emphasizing the significance of the
proposed approach in tackling the challenges of high-dimensional data
visualization.

\hypertarget{sec-background}{%
\section{Background}\label{sec-background}}

\hypertarget{dimension-reduction}{%
\subsection{Dimension Reduction}\label{dimension-reduction}}

Consider the high-dimensional data a rectangular matrix \(X\), where
\(X = \begin{bmatrix} \textbf{x}_{1} & \textbf{x}_{2} & \cdots & \textbf{x}_{n}\\ \end{bmatrix}\),
with \(n\) observations in \(p\) dimensions. The objective is to
discover a low-dimensional projection
\(Y = \begin{bmatrix} \textbf{y}_{1} & \textbf{y}_{2} & \cdots & \textbf{y}_{n}\\ \end{bmatrix}\),
represented as an \(n\) × \(d\) matrix, where \(d \ll p\). The reduction
process seeks to remove noise from the original data set while retaining
essential information.

There are two main categories of dimension reduction techniques: linear
and non-linear methods. Linear techniques involve a linear
transformation of the data, with one popular example being PCA. PCA
performs an eigen-decomposition of the sample covariance matrix to
obtain orthogonal principal components that capture the variance of the
data \citep{Karl1901}. However, linear methods may not fully capture
complex non-linear relationships present in the data.

In contrast, NLDR techniques generate the low-dimensional representation
\(Y\) from the high-dimensional data \(X\), often using pre-processing
techniques like \(k\)-nearest neighbors graph or kernel transformations.
Multidimensional Scaling (MDS) is a class of NLDR methods that aims to
construct an embedding \(Y\) in a low-dimensional space, approximating
the pair-wise distances in \(X\) \citep{Torgerson1967}. Variants of MDS
include non-metric scaling \citep{article62} and Isomap, which estimate
geodesic distances to create the low-dimensional representation
\citep{article63}. Other approaches based on diffusion processes, like
diffusion maps \citep{article64} and the PHATE algorithm
\citep{article03}, also fall under NLDR methods.

A challenge with NLDR methods is selecting and tuning appropriate
parameters. One specific technique we focus on is Pairwise Controlled
Manifold Approximation (PaCMAP). Similar considerations apply to related
methods like tSNE \citep{Laurens2008}, UMAP \citep{Leland2018}, and
TrMAP \citep{article02}.

It is important to note that methods like PCA and auto-encoders
\citep{article65} provide a reverse mapping from the low-dimensional
space back to the high-dimensional space, enabling data reconstruction.
However, many non-linear methods, including tSNE, prioritize
visualization and exploration over reconstruction. Their focus is on
capturing complex structures that may not be easily represented in the
original space, making a straightforward reverse mapping challenging.

\hypertarget{introduction-to-umap}{%
\subsubsection{Introduction to UMAP}\label{introduction-to-umap}}

The UMAP algorithm constructs a fuzzy topological representation of the
high-dimensional data and then optimizes the low-dimensional embedding
to be as similar to this representation as possible \citep{Leland2018}.

The UMAP algorithm constructs a weighted k-nearest neighbors graph based
on the local similarity of data points in the original high-dimensional
space. The weighted graph is defined by a conditional probability
distribution \(P_{ij}\), which represents the probability of observing
data point \(x_i\) given data point \(x_j\) as a neighbor:

\[
P_{ij} = \frac{\exp(-\lVert x_i - x_j \rVert^2)}{\sum_{k \neq i} \exp(-\lVert x_i - x_k \rVert^2)}
\]

Similarly, a fuzzy topological representation of the data is constructed
in the low-dimensional space with conditional probability distribution
\(Q_{ij}\):

\[
Q_{ij} = \frac{\exp(-\lVert y_i - y_j \rVert^2)}{\sum_{k \neq i} \exp(-\lVert y_i - y_k \rVert^2)}
\]

where \(y_i\) and \(y_j\) are the corresponding low-dimensional
representations of data points \(x_i\) and \(x_j\).

UMAP minimizes the cross-entropy between the high-dimensional fuzzy
topological representation \(P_{ij}\) and the low-dimensional
representation \(Q_{ij}\) to find an optimal low-dimensional embedding.
The loss function for UMAP is defined as:

\[
L = \sum_{i} \sum_{j} P_{ij} \log \frac{P_{ij}}{Q_{ij}}
\]

The algorithm uses stochastic gradient descent to optimize the positions
of the data points in the low-dimensional space to minimize this loss
function.

In R, the \texttt{umap} package is often used to implement UMAP and
generate results.

The ``n\_neighbors'' parameter in UMAP determines the number of nearest
neighbors to consider for each data point during the construction of the
fuzzy topological representation. Higher values of ``n\_neighbors''
capture more global structure, while lower values emphasize local
structure. Choosing an appropriate value for ``n\_neighbors'' is
crucial, as it affects the balance between local and global structure in
the resulting low-dimensional representation.

\begin{figure}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervis-1.pdf}

}

\caption{\label{fig-nldervis}2D layouts from different NLDR techniques
applied the same data: (a) tSNE (perplexity = 32), (b) UMAP
(n\_neighbors = 50), (c) PHATE (knn = 5), (d) TriMAP (n\_inliers = 5,
n\_outliers = 4, n\_random = 3), and (e) PaCMAP (n\_neighbors = 10, init
= random, MN\_ratio = 0.9, FP\_ratio = 2). Is there a best
representation of the original data or are they all providing equivalent
information?}

\end{figure}

\hypertarget{linear-overviews-using-tours}{%
\subsection{Linear overviews using
tours}\label{linear-overviews-using-tours}}

A tour is a powerful visualization technique used to explore
high-dimensional data by generating a sequence of projections, typically
into two dimensions. There are two main types of tours: the Grand Tour
and the Guided Tour. The Grand Tour explores the data's shape and global
structure by using random projections \citep{Asimov1985}. In contrast,
the Guided Tour focuses on specific patterns by moving towards
interesting projections defined by a predefined index function
\citep{article29}.

The process begins with a real data matrix \(X\) containing \(n\)
observations in \(p\) dimensions. It generates a sequence of \(p\) ×
\(d\) orthonormal projection matrices (bases), usually 1 or 2
dimensions. For each pair of orthonormal bases \(A_t\) and \(A_{t+1}\),
a geodesic path is interpolated to create smooth animation between
projections.

In the Grand Tour, new orthonormal bases are randomly chosen to explore
the \(d\)-dimensional subspace. The data is often sphered via principal
components to reduce dimensionality. The Guided Tour uses a predefined
index function to generate a sequence of `interesting' projections. The
resulting tour continuously visualizes the projected data \(Y_t\) =
\(XA_t\) as it interpolates between successive bases.

While both tours can be used to visualize data, examples often focus on
using the Grand Tour to observe global structures. However, software
like \textbf{langevitour} can visualize both types of tours, providing
flexibility for exploring high-dimensional data with various objectives.

\hypertarget{non-parametric-models}{%
\subsection{Non-parametric models}\label{non-parametric-models}}

\hypertarget{hexagonal-binning}{%
\subsubsection{Hexagonal binning}\label{hexagonal-binning}}

Hexagonal binning is a data visualization technique that aggregates
high-dimensional data into a two-dimensional representation using
hexagonal regions called bins \citep{article66}. The hexagonal shape is
preferred due to its circular-like appearance, providing smoother
transitions between neighboring bins and efficient data aggregation.
This technique is especially useful for handling large data sets as it
reduces visual clutter while capturing the underlying data distribution
effectively. By considering the data distribution, important features
are captured, and gaps between bins are minimized, leading to a more
accurate representation of the data. In our algorithm, hexagonal binning
is utilized to create a low-dimensional manifold and facilitate data
visualization in the original high-dimensional space.

\hypertarget{triangulation}{%
\subsubsection{Triangulation}\label{triangulation}}

Triangulation is a fundamental process in computational geometry and
computer graphics that involves dividing a set of points in a given
space into interconnected triangles \citep{article30} finding
applications in mesh generation and representing complex geometric
structures efficiently and discretely. One common algorithm used for
triangulation is Delaunay triangulation \citep{article26}, where points
are connected in a way that maximizes the minimum angles of the
resulting triangles, leading to a more regular and well-conditioned
triangulation.

In our algorithm, we are interested in triangulation as it plays a key
role in providing a clear and concise representation of the data. By
converting the data points into interconnected triangles, we can
visualize and analyze the data in a more structured and organized
manner. Triangulation helps us identify geometric relationships and
patterns in the data, enabling us to make informed decisions and gain
valuable insights from the information presented in the triangular form.

\hypertarget{sec-methods}{%
\section{Methodology}\label{sec-methods}}

Our algorithm comprises two main phases: (1) generate the model in the
2D space, and (2) generate the model in the high-D space. These two
phases are described in details in this section.

\begin{figure}

{\centering \includegraphics[width=1\textwidth,height=1\textheight]{figures/workflow.png}

}

\caption{\label{fig-meth}A flow diagram detailing the steps taken to
create the low-dimensional manifold in the high dimensional space. There
are two basic phases, one to generate the model in the 2D space, and
other to generate the model in the high-D space.}

\end{figure}

\hypertarget{preprocessing-steps}{%
\subsection{Preprocessing steps}\label{preprocessing-steps}}

To tackle the complexities and noise in high-dimensional data, we apply
PCA as a pre-processing step \citep[\citet{article68},
\citet{article69}]{article67}. This step helps in noise reduction by
identifying principal components that represent directions of maximum
variance, capturing essential patterns in the data.

\hypertarget{constructing-the-2d-model}{%
\subsection{Constructing the 2D model}\label{constructing-the-2d-model}}

\textbf{Step 1: Computing the hexagonal grid configuration}

\textbf{(a) Determine the number of bins along the x-axis} (\(b_1\))

The first step involves to determine the optimal number of bins (\(b\))
for creating regular hexagons in the hexagonal grid. To achieve this, we
rely on the relationship between the diameter (\(h\)) and the area
(\(A\)) of regular hexagons, as given by Equation~\ref{eq-equation3}.

\begin{equation}\protect\hypertarget{eq-equation3}{}{
 \text{A} = \frac{\sqrt{3}}{2}h^2
}\label{eq-equation3}\end{equation}

To capture the data's structure effectively, we consider the optimal
\(A = 1\) (see Figure~\ref{fig-binsize}). Using
Equation~\ref{eq-equation4}, we then calculate \(h\) of a regular
hexagon.

\begin{equation}\protect\hypertarget{eq-equation4}{}{
 \text{h} = \sqrt{\frac{2}{\sqrt{3}}A}
}\label{eq-equation4}\end{equation}

Next, we refer the \texttt{hexbin} function in R, which provides a
relationship between the diameter of the hexagon (\(h\)) and the number
of bins along the x-axis (\(b_1\)) (Equation~\ref{eq-equation6}).

\begin{equation}\protect\hypertarget{eq-equation6}{}{
 h = \frac{r_1}{b_1}
}\label{eq-equation6}\end{equation}

By utilizing the calculated \(h\) and the range of the non-linear
projection component 1 (\(r_1\)), we determine the \(b_1\)
(Equation~\ref{eq-equation5}). The result is rounded up to the nearest
whole number to ensure an appropriate number of bins that capture the
variability of the data along the x-axis.

\begin{equation}\protect\hypertarget{eq-equation5}{}{
 b_1 = \frac{r_1}{h}
}\label{eq-equation5}\end{equation}

\textbf{(b) Determine the effective shape parameter} (\(s\))

In this step, we determine the optimal shape parameter (\(s\)) for the
hexagonal bins, which significantly influences their shape and
arrangement within the grid. The shape parameter (\(s\)) in the
hexagonal binning algorithm is defined as the ratio of the height
(\(y\)) to the width (\(x\)) of the plotting region, referring to the
\texttt{hexbin} function in R (see Equation~\ref{eq-equation1}). It
determines the shape of the plotting regions and plays a vital role in
generating an appropriate hexagonal grid for the data visualization.

\begin{equation}\protect\hypertarget{eq-equation1}{}{
  s = \frac{y}{x}
}\label{eq-equation1}\end{equation}

To calculate the effective shape parameter (\(s\)) for our algorithm, we
consider the ranges of the non-linear projection components. Denoting
the range of the non-linear projection component 1 as \(r_1\) and the
range of the non-linear projection component 2 as \(r_2\), we find the
shape parameter using Equation~\ref{eq-equation2}.

\begin{equation}\protect\hypertarget{eq-equation2}{}{
s = \frac{r_2}{r_1}
}\label{eq-equation2}\end{equation}

\begin{figure}[H]

{\centering \includegraphics{paper_files/figure-pdf/fig-binsize-1.pdf}

}

\caption{\label{fig-binsize}Hexbin plots from different number of bins
for the same \textbf{s\_curve\_noise\_umap} data: (a) b = (4, 8), s =
2.031, (b) b = (9, 20), s = 2.031, and (c) b = (16, 36), s = 2.031. The
hexbins are colored based on the density of points, with yellow
indicating higher point density and darker colors representing lower
point density within each bin. Does a value of number of bins exist to
effectively represent the low-dimensional data?}

\end{figure}

\textbf{Step 2: Obtain bin centroids}

After computing hexagonal grid configurations, the 2D emdeddings are
binned into hexagon cells. As a result of hexagonal binning , the bin
centroids (\(C_k^{(2)} \equiv (C_{kx}, C_{ky}\))) (see
Figure~\ref{fig-meth} (Step 2)) are obtained.

\textbf{Step 3: Triangulate bin centroids}

In this step, the algorithm proceeds to triangulate the centroids (see
Figure~\ref{fig-meth} (Step 3)) of the hex bins using the
\texttt{tripack} package in R \citep{article54}. Since we are working
with the centroids of regular hexagonal bins, the resulting mesh will
predominantly comprise equal-sized regular triangles. However, the
triangulation also helps span any gaps that may exist between clusters
of points, allowing for a more complete and interconnected
representation of the data.

\hypertarget{lifting-the-model-into-high-dimensions}{%
\subsection{Lifting the model into high
dimensions}\label{lifting-the-model-into-high-dimensions}}

Creating the model in the high-D space involves a two-step process.
Initially, we identify observations within each hexagonal bin. After
this, we calculate the average of the high-dimensional data within each
hex bin. The resulting values represent the averaged high-dimensional
data points (\(C_k^{(p)} \equiv (C_{kx_1}, ..., C_{kx_p}\))).
Essentially, these averaged high-dimensional data points serve as the
representation of the hex bin centroids in the high-D space.

\begin{figure}

{\centering \includegraphics{paper_files/figure-pdf/fig-wkhighD-1.pdf}

}

\caption{\label{fig-wkhighD}Visualization.}

\end{figure}

\hypertarget{tunning-the-model}{%
\subsection{Tunning the model}\label{tunning-the-model}}

\hypertarget{sec-diagnostics}{%
\subsection{Summaries of the model}\label{sec-diagnostics}}

\hypertarget{predicted-values-and-residuals}{%
\subsubsection{Predicted values and
residuals}\label{predicted-values-and-residuals}}

In our prediction process, we employ the K-nearest neighbors (KNN)
algorithm to predict the hex bin IDs. Specifically, for each new
high-dimensional data point, the algorithm assigns the first neighboring
hex bin ID as the predicted hex bin ID.

Moving on to the assessment of prediction accuracy, we calculate the
Mean Squared Error (MSE). The MSE helps measure the average squared
differences between the actual high-dimensional data (\(x_{ij}\)) and
the predicted bin centroid data in high-D (\(C_{x_ij}\)) values across
all bins. Mathematically, this is expressed as:

\begin{equation}\protect\hypertarget{eq-equation9}{}{
\text{MSE} = \sum_{j = 1}^{n} \frac{\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2}{\text{total number of bins}}
}\label{eq-equation9}\end{equation}

Here, \(b\) signifies the total number of bins, \(p\) denotes the number
of dimensions in the high-dimensional (high-D) data, and \(n\)
represents the number of observations.

Additionally, we gauge the model's performance using the Akaike
Information Criterion (AIC), calculated by the formula:

\begin{equation}\protect\hypertarget{eq-equation10}{}{
\text{AIC} = 2bp + np * log(\text{MSE})
}\label{eq-equation10}\end{equation}

These metrics, MSE and AIC, collectively offer valuable insights into
the model's predictive performance, considering both accuracy and
complexity in the predictions.

\begin{figure}

{\centering \includegraphics{paper_files/figure-pdf/fig-predict-1.pdf}

}

\caption{\label{fig-predict}The .}

\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticplt-1.pdf}

}

\caption{\label{fig-diagnosticplt}The .}

\end{figure}

\hypertarget{goodness-of-fit-statistics}{%
\subsubsection{Goodness of fit
statistics}\label{goodness-of-fit-statistics}}

In this analysis, we're looking at how well our model is doing in two
aspects. First, we're checking the distances in our lower-dimensional
space (see Figure~\ref{fig-diag} (a)). We want to see if the distances
between different types of edges are similar. If they are, it means our
model is doing a good job of keeping the relationships consistent. If
there are big differences, it could mean our model is struggling with
certain types of edges.

Next, we're examining how our model handles the relationship between
high-dimensional and low-dimensional distances for each type of edge
(see Figure~\ref{fig-diag} (b)). If our data points nicely line up
around the expected line, it tells us that our model is preserving the
original structure well. If they're scattered, it could mean our model
is having trouble with specific data points. This two-part analysis
helps us really understand how our model fits the data.

\begin{figure}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diag-1.pdf}

}

\caption{\label{fig-diag}The two diagnostic plot assess the
effectiveness of the high-D representation algorithm performed for the
\textbf{s\_curve\_noise\_umap} data set\(\colon\) (a) visualizing the
differences between low-D distances and (b) visualizing the relationship
between high-D (\(d^{(p)}\)) and low-D distances (\(d^{(2)}\)).}

\end{figure}

\hypertarget{sec-simpleex}{%
\subsection{Simulated data examples}\label{sec-simpleex}}

In this section, we demonstrate the effectiveness of our approach
through the use of simulation data. These simulations are designed to
have known cluster structures and underlying data geometries. We begin
with a straightforward example where we generate spherical clusters with
a doublet structure. This allows us to assess how well our algorithm can
identify and represent these known structures in the low-dimensional
space.

We then proceed to a more complex example where the data exhibits a more
intricate geometry. This simulation provides a challenging test for our
approach to capture and preserve the underlying data structure
accurately.

We \emph{strongly} recommend viewing the linked videos for each study
while reading. Links to the videos are available in the figures for each
example. The videos show the visual appearance of the
\textbf{langevitour} interface with low-dimensional view and how we can
interact with the tour via the controls.

\textbf{Example 1: Exploring spherical Gaussian clusters with doublet
structure}

The next data set consists of two distinct types of clusters embedded in
a 10-\(d\) space. The first set of clusters includes three Gaussian
clusters with equal variance. Each of these Gaussian clusters is
equidistant from the others, forming a well-separated and symmetrical
arrangement. The second set of clusters consists of three clusters with
a doublet structure. These clusters are also embedded in the 10-\(d\)
space, but each is positioned in the middle of its parent Gaussian
clusters. Moreover, each doublet cluster contains an average number of
points from its corresponding parent Gaussian clusters.

For analysis, we run tSNE on the training data with a perplexity setting
18. Figure~\ref{fig-example3} (a) displays the tSNE results, indicating
that the sub-clusters have been correctly identified. However, the
relative locations of certain clusters to each other appear distorted in
the low-dimensional view, such as the red and pink clusters being far
apart.

Nonetheless, we find no apparent impact on the perception of the data
structure when we examine the tour view and the model (see
Figure~\ref{fig-example3}) through the linked video (see
Figure~\ref{fig-exp1_sc}). The model demonstrates accurate
identification of clusters and captures the internal variations within
each cluster, reflecting the inherent variance present in the data. The
configurations used for the analysis are detailed in
\textbf{?@tbl-table02}. Additionally, the representation of the parent
clusters in the low-dimensional space appears stretched, aligning with
the variance of these clusters. On the other hand, the doublet clusters
are squeezed, reflecting their specific characteristics and relationship
to their parent Gaussian clusters.

\begin{figure}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-example3-1.pdf}

}

\caption{\label{fig-example3}Visualization and refinement of
low-dimensional representation for sub-clustered data\(\colon\) (a) 2D
layout of tSNE (perplexity = 18), (b) triangular mesh, and (c)
triangular mesh colored by edge type. Edges with lengths less than the
benchmark value of \(2\) are classifiedas small, while edges with
lengths greater than or equal to this value are labeled as long. The
combination of tSNE, hexagonal binning, and triangulation provides the
effectiveness of our approach in visualizing and refining the
low-dimensional representation of the sub-clustered data.}

\end{figure}

\begin{figure}

\begin{minipage}[t]{0.33\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{figures/example_1a.png}

}

}

\subcaption{\label{fig-exp1_sc1}}
\end{minipage}%
%
\begin{minipage}[t]{0.33\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{figures/example_1b.png}

}

}

\subcaption{\label{fig-exp1_sc2}}
\end{minipage}%
%
\begin{minipage}[t]{0.33\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{figures/example_1c.png}

}

}

\subcaption{\label{fig-exp1_sc3}}
\end{minipage}%

\caption{\label{fig-exp1_sc}Screen shots of the \textbf{langevitour}
with the low-dimensional view applied to sub-clustered data set, a video
of the tour animation is available at
(\url{https://drive.google.com/file/d/15nAbmc4UoVlUyFcGtnlnmyxrJEbK39nx/view}).
Combination of the \textbf{langevitour} and low-dimensional
representation of the sub-clustered data provides the effectiveness of
our approach.}

\end{figure}

\textbf{Example 2: Exploring data with piecewise linear structure}

In our next exploration, we delve into simulated noisy tree-structured
data designed to mimic branching trajectories of cell differentiation.
The simulation is crafted to resemble the hierarchical clustering
pattern observed in cell differentiation trajectories, where mature
cells are akin to the tips of branches.

To begin our analysis, we apply principal components to the data and
focus on the first ten principal components, collectively accounting for
approximately 68\% of the variance explained in the data set. Then, we
run TriMAP with the default parameters on the training data (n\_inliers:
5, n\_outliers: 4, and n\_random: 3). Through this process, we create a
low-dimensional manifold using the configurations detailed in
\textbf{?@tbl-table02}. Finally, we present the model along with the
data using the tour view (Figure Figure~\ref{fig-exp2_sc}), allowing us
to visualize and explore the intricate relationships and structures
present in the data's low-dimensional representation.

\begin{figure}[H]

{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-example2-1.pdf}

}

\caption{\label{fig-example2}Visualization and refinement of
low-dimensional representation for tree structured data\(\colon\) (a) 2D
layout of TriMAP (n-inliers = 5, n-outliers = 4, n-random = 3), (b)
triangular mesh, and (c) triangular mesh colored by edge type. Edges
with lengths less than the benchmark value of \(2\) are classified as
small, while edges with lengths greater than or equal to this value are
labeled as long. The combination of TriMAP, hexagonal binning, and
triangulation provides the effectiveness of our approach in visualizing
and refining the low-dimensional representation of the tree structured
data.}

\end{figure}

\begin{figure}

\begin{minipage}[t]{0.33\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{figures/example_2a.png}

}

}

\end{minipage}%
%
\begin{minipage}[t]{0.33\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{figures/example_2b.png}

}

}

\end{minipage}%
%
\begin{minipage}[t]{0.33\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{figures/example_2c.png}

}

}

\end{minipage}%

\caption{\label{fig-exp2_sc}Screen shots of the \textbf{langevitour}
with the low-dimensional view applied to tree structured data set, a
video of the tour animation is available at
(\url{https://drive.google.com/file/d/1YQx4UGlzDify1mGqeDtlqTHncgqidDhq/view}).
Combination of the \textbf{langevitour} and low-dimensional
representation of the tree structured data provides the effectiveness of
our approach.}

\end{figure}

\hypertarget{sec-applications}{%
\section{Examples/Applications}\label{sec-applications}}

\hypertarget{single-cell-rna-seq-data-of-human}{%
\subsection{Single-cell RNA-seq data of
human}\label{single-cell-rna-seq-data-of-human}}

In the field of single-cell studies, a common analysis task involves
clustering to identify groups of cells with similar expression profiles.
Analysts often turn to NLDR techniques to verify and identify these
clusters and explore developmental trajectories (e.g., example 1). In
clustering workflows, the main objective is to verify the existence of
clusters and subsequently identify them as specific cell types by
examining the expression of ``known'' marker genes. In this context, a
``faithful'' embedding should ideally preserve the topology of the data,
ensuring that cells corresponding to the same cell type are situated
close to the high-dimensional space.

To begin our analysis, we installed the Peripheral Blood Mononuclear
Cells (pbmc) data set obtained from 10x Genomics using the
\texttt{SeuratData} R package \citep{Rahul2019}, which facilitates the
distribution of data sets in the form of Seurat objects
\citep{Yuhan2021}. This data set contains 13,714 features across 2,700
samples within a single assay. The active assay is RNA, with 13,714
features representing different gene expressions. After loading the data
set, we obtained the principal components (PCs) and assessed the
variance explained by each PC. Based on this evaluation, we selected
seven PCs, representing approximately 50\% of the variance in gene
expression, for further analysis.

Next, we employed the UMAP technique with default parameter settings. As
illustrated in Figure~\ref{fig-pbmc}, the cell types B and Platelet are
well-separated in the UMAP layout. Moreover, CD14+ Mono, FCGR3A+ Mono,
and DC form a distinct cluster, while Naive CD4 T, NK, Memory CD4 T, and
CD8 T are grouped together in another cluster. The values utilized to
construct the smooth low-dimensional manifold are presented in
\textbf{?@tbl-table02}. The linked video, demonstrating the tour with
the model, showcases the generation of a smooth surface for this
application, enabling a comprehensive exploration of the data's
structure and relationships (see Figure~\ref{fig-pbmc_sc}).

\begin{figure}[H]

{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-pbmc-1.pdf}

}

\caption{\label{fig-pbmc}Visualization and refinement of low-dimensional
representation for pbmc data\(\colon\) (a) 2D layout of UMAP (n-neigbors
= 15), (b) triangular mesh, and (c) triangular mesh colored by edge
type. Edges with lengths less than the benchmark value of \(2\) are
classified as small, while edges with lengths greater than or equal to
this value are labeled as long. The combination of UMAP, hexagonal
binning, and triangulation provides the effectiveness of our approach in
visualizing and refining the low-dimensional representation of the pbmc
data.}

\end{figure}

\begin{figure}

\begin{minipage}[t]{0.33\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{figures/pbmc_a.png}

}

}

\subcaption{\label{fig-pbmc_sc1}}
\end{minipage}%
%
\begin{minipage}[t]{0.33\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{figures/pbmc_b.png}

}

}

\subcaption{\label{fig-pbmc_sc2}}
\end{minipage}%
%
\begin{minipage}[t]{0.33\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{figures/pbmc_c.png}

}

}

\subcaption{\label{fig-pbmc_sc3}}
\end{minipage}%

\caption{\label{fig-pbmc_sc}Screen shots of the \textbf{langevitour}
with the low-dimensional view applied to pbmc data set, a video of the
tour animation is available at
(\url{https://drive.google.com/file/d/1F7Xb2EIpfUruCxYihPw2rKro4FB5Y2uL/view}).
Combination of the \textbf{langevitour} and low-dimensional
representation of the pbmc data provides the effectiveness of our
approach.}

\end{figure}

\hypertarget{sec-discussion}{%
\section{Discussion}\label{sec-discussion}}

Our research introduces a comprehensive framework that leverages tours
for interactive exploration of high-dimensional data coupled with a
low-dimensional manifold, facilitated by the \texttt{quollr} R package.
Regardless of the Non-Linear Dimension Reduction (NLDR) technique in
use, our approach demonstrates effectiveness through simulation
examples, particularly in the iterative removal of long edges for a
smoother representation and capturing cluster variance.

In the example with doublets, our method successfully captures the tweak
within each cluster, indicating the variance present within them.
However, the model may not appear smooth in high-dimensional space due
to considerable noise when the data has a piecewise linear geometry,
such as the tree simulation.

The practical application of our framework, as showcased with the UMAP
view, enables visual inspection of well-separated clusters. Furthermore,
the combined tour and model provide a robust assessment of whether UMAP
preserves the data structure and accurately transforms the data.

The advantages of our approach include its versatility across various
NLDR techniques and the ability to generate interactive visualizations
for detailed exploration. The tour provides an intuitive way to navigate
and comprehend high-dimensional data while assessing the accuracy of
dimensionality reduction.

However, one limitation is that the approach may be less effective in
cases with significant noise, as seen in the tree simulation example.
Additionally, while our method aids in visual verification, quantifying
the accuracy of embeddings might require further evaluation metrics.

In conclusion, our framework presents a powerful tool for researchers
and analysts in single-cell studies to assess their embeddings by
visually inspecting them alongside the original data. By leveraging the
advantages of tours and low-dimensional manifolds, our approach offers
valuable insights into the data transformation process, empowering users
to make informed decisions in analyzing high-dimensional data. Future
work could enhance the method's robustness in the presence of noise and
explore additional evaluation metrics for quantifying embedding
accuracy.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\renewcommand{\bibsection}{}
\bibliography{bibliography.bib}

\newpage{}

\end{document}

### 2024-01-09 Supplementary/automation_of_benchmark_value_to_rm_lg.R

@@ -1,0 +1,330 @@
+library(dplyr)
+library(snedata)
+library(ggflowchart)
+library(purrr) ## map function
+library(gridExtra) ## for grid.arrange
+library(rsample)
+library(DT)
+library(ggbeeswarm)
+library(ggplot2)
+library(readr)
+
+library(Rtsne)
+library(umap)
+library(phateR)
+library(reticulate)
+library(patchwork)
+
+library(grid)
+set.seed(20230531)
+
+source(paste0(here::here(), "/paper-nldr-vis-algorithm/quollr_code.R"))
+source(paste0(here::here(), "/paper-nldr-vis-algorithm/nldr_code.R"))
+
+data <- read_csv(paste0(here::here(), "/paper-nldr-vis-algorithm/data/s_curve.csv"))
+
+data <- data |>
+  mutate(ID = row_number())
+
+data_split <- initial_split(data)
+training_data <- training(data_split) |>
+  arrange(ID)
+test_data <- testing(data_split) |>
+  arrange(ID)
+
+UMAP_fit <- umap(training_data |> dplyr::select(-ID), n_neighbors = 50, n_components =  2)
+
+UMAP_data <- UMAP_fit$layout |>
+  as.data.frame()
+names(UMAP_data)[1:(ncol(UMAP_data))] <- paste0(rep("UMAP",(ncol(UMAP_data))), 1:(ncol(UMAP_data)))
+
+UMAP_data <- UMAP_data |>
+  mutate(ID = training_data$ID)
+
+num_bins_x <- calculate_effective_x_bins(.data = UMAP_data, x = UMAP1,
+                                         cell_area = 1)
+
+shape_value <- calculate_effective_shape_value(.data = UMAP_data,
+                                               x = UMAP1, y = UMAP2)
+
+## To extract bin centroids
+hexbin_data_object <-extract_hexbin_centroids(nldr_df = UMAP_data, num_bins = num_bins_x, shape_val = shape_value)
+
+df_bin_centroids <- hexbin_data_object$hexdf_data
+
+UMAP_data_with_hb_id <- UMAP_data |>
+  dplyr::mutate(hb_id = hexbin_data_object$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all <- dplyr::bind_cols(training_data |> dplyr::select(-ID), UMAP_data_with_hb_id)
+
+## Averaged on high-D
+df_bin <- avg_highD_data(.data = df_all)
+
[…]

### 2024-01-09 Supplementary/automation_of_benchmark_value_to_rm_low_hex.R

@@ -1,0 +1,374 @@
+library(dplyr)
+library(snedata)
+library(ggflowchart)
+library(purrr) ## map function
+library(gridExtra) ## for grid.arrange
+library(rsample)
+library(DT)
+library(ggbeeswarm)
+library(ggplot2)
+library(readr)
+
+library(Rtsne)
+library(umap)
+library(phateR)
+library(reticulate)
+library(patchwork)
+
+library(grid)
+set.seed(20230531)
+
+source(paste0(here::here(), "/paper-nldr-vis-algorithm/quollr_code.R"))
+source(paste0(here::here(), "/paper-nldr-vis-algorithm/nldr_code.R"))
+
+data <- read_csv(paste0(here::here(), "/paper-nldr-vis-algorithm/data/s_curve.csv"))
+
+data <- data |>
+  mutate(ID = row_number())
+
+data_split <- initial_split(data)
+training_data <- training(data_split) |>
+  arrange(ID)
+test_data <- testing(data_split) |>
+  arrange(ID)
+
+UMAP_fit <- umap(training_data |> dplyr::select(-ID), n_neighbors = 50, n_components =  2)
+
+UMAP_data <- UMAP_fit$layout |>
+  as.data.frame()
+names(UMAP_data)[1:(ncol(UMAP_data))] <- paste0(rep("UMAP",(ncol(UMAP_data))), 1:(ncol(UMAP_data)))
+
+UMAP_data <- UMAP_data |>
+  mutate(ID = training_data$ID)
+
+num_bins_x <- calculate_effective_x_bins(.data = UMAP_data, x = UMAP1,
+                                         cell_area = 1)
+
+shape_value <- calculate_effective_shape_value(.data = UMAP_data,
+                                               x = UMAP1, y = UMAP2)
+
+## To extract bin centroids
+hexbin_data_object <- extract_hexbin_centroids(nldr_df = UMAP_data, num_bins = num_bins_x, shape_val = shape_value)
+
+df_bin_centroids <- hexbin_data_object$hexdf_data
+
+df_bin_centroids_all <- df_bin_centroids
+
+## Benchmark value to remove low-density hexagons is 0.255
+identify_rm_bins <- find_low_density_hexagons(df_bin_centroids, num_bins_x)
+
+df_bin_centroids <- df_bin_centroids |>
+  filter(!(hexID %in% identify_rm_bins))
+
[…]

### 2024-01-09 Supplementary/bin_centroids_weighted_mean_high_D.R

@@ -1,0 +1,143 @@
+library(dplyr)
+library(snedata)
+library(ggflowchart)
+library(purrr) ## map function
+library(gridExtra) ## for grid.arrange
+library(rsample)
+library(DT)
+library(ggbeeswarm)
+library(ggplot2)
+library(readr)
+
+library(Rtsne)
+library(umap)
+library(phateR)
+library(reticulate)
+library(patchwork)
+
+library(grid)
+set.seed(20230531)
+
+source(paste0(here::here(), "/paper-nldr-vis-algorithm/quollr_code.R"))
+source(paste0(here::here(), "/paper-nldr-vis-algorithm/nldr_code.R"))
+
+data <- read_csv(paste0(here::here(), "/paper-nldr-vis-algorithm/data/s_curve.csv"))
+
+data <- data |>
+  mutate(ID = row_number())
+
+data_split <- initial_split(data)
+training_data <- training(data_split) |>
+  arrange(ID)
+test_data <- testing(data_split) |>
+  arrange(ID)
+
+UMAP_fit <- umap(training_data |> dplyr::select(-ID), n_neighbors = 50, n_components =  2)
+
+UMAP_data <- UMAP_fit$layout |>
+  as.data.frame()
+names(UMAP_data)[1:(ncol(UMAP_data))] <- paste0(rep("UMAP",(ncol(UMAP_data))), 1:(ncol(UMAP_data)))
+
+UMAP_data <- UMAP_data |>
+  mutate(ID = training_data$ID)
+
+num_bins_x <- calculate_effective_x_bins(.data = UMAP_data, x = UMAP1,
+                                         cell_area = 1)
+
+shape_value <- calculate_effective_shape_value(.data = UMAP_data,
+                                               x = UMAP1, y = UMAP2)
+
+## To extract bin centroids
+hexbin_data_object <- extract_hexbin_centroids(nldr_df = UMAP_data, num_bins = num_bins_x, shape_val = shape_value)
+
+df_bin_centroids <- hexbin_data_object$hexdf_data
+
+## Data set with all possible centroids in the hexagonal grid
+full_centroid_df <- generate_full_grid_centroids(df_bin_centroids)
+## Generate all coordinates of hexagons
+hex_grid <- full_hex_grid(full_centroid_df)
+
+ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
+  geom_point(data = df_bin_centroids, aes(x = x, y = y), color = "#33a02c")
+
[…]

### 2024-01-09 paper.qmd
Selected another application with mouse barin data

@@
 
-### Non-parametric models
-
-#### Hexagonal binning
+## Methodology {#sec-methods}
+
+<!--In this paper, we introduce a novel approach, which is designed to determine which method, and which parameter choice provide the most useful representation of high dimensional data.
+Our approach involves dividing the data set into two subsets: a training set, which serves to construct the low-dimensional manifold, and a test set, utilized to generate predictive values and residuals.
+
+For the implementation of our approach, a 2D embedding data set serves as the initiation point for the subsequent analysis and visualization (see @fig-nldervis (b)).-->
+
+<!--### Overview {#sec-algorithm}-->
+
+Our algorithm comprises two main phases: (1) generate the model in the 2D space, and (2) generate the model in the high-D space. These two phases are described in details in this section.
+
+![A flow diagram detailing the steps taken to create the low-dimensional manifold in the high dimensional space.
+There are two basic phases, one to generate the model in the 2D space, and other to generate the model in the high-D space.](figures/workflow.png){#fig-meth fig-align="center" width="100%" height="100%"}
+
+### Preprocessing steps
+
+To tackle the complexities and noise in high-dimensional data, we apply PCA as a pre-processing step [@article67, @article68, @article69].
+This step helps in noise reduction by identifying principal components that represent directions of maximum variance, capturing essential patterns in the data.
+
+### Constructing the 2D model
+
+**Step 1: Computing the hexagonal grid configuration**
 
@@
 By considering the data distribution, important features are captured, and gaps between bins are minimized, leading to a more accurate representation of the data.
-In our algorithm, hexagonal binning is utilized to create a low-dimensional manifold and facilitate data visualization in the original high-dimensional space.
-
-#### Triangulation
-
-Triangulation is a fundamental process in computational geometry and computer graphics that involves dividing a set of points in a given space into interconnected triangles [@article30] finding applications in mesh generation and representing complex geometric structures efficiently and discretely.
-One common algorithm used for triangulation is Delaunay triangulation [@article26], where points are connected in a way that maximizes the minimum angles of the resulting triangles, leading to a more regular and well-conditioned triangulation.
-
-In our algorithm, we are interested in triangulation as it plays a key role in providing a clear and concise representation of the data.
-By converting the data points into interconnected triangles, we can visualize and analyze the data in a more structured and organized manner.
-Triangulation helps us identify geometric relationships and patterns in the data, enabling us to make informed decisions and gain valuable insights from the information presented in the triangular form.
-
-## Methodology {#sec-methods}
-
-<!--In this paper, we introduce a novel approach, which is designed to determine which method, and which parameter choice provide the most useful representation of high dimensional data.
-Our approach involves dividing the data set into two subsets: a training set, which serves to construct the low-dimensional manifold, and a test set, utilized to generate predictive values and residuals.
-
-For the implementation of our approach, a 2D embedding data set serves as the initiation point for the subsequent analysis and visualization (see @fig-nldervis (b)).-->
-
-<!--### Overview {#sec-algorithm}-->
-
-Our algorithm comprises two main phases: (1) generate the model in the 2D space, and (2) generate the model in the high-D space. These two phases are described in details in this section.
-
-![A flow diagram detailing the steps taken to create the low-dimensional manifold in the high dimensional space.
-There are two basic phases, one to generate the model in the 2D space, and other to generate the model in the high-D space.](figures/workflow.png){#fig-meth fig-align="center" width="100%" height="100%"}
-
-### Preprocessing steps
-
-To tackle the complexities and noise in high-dimensional data, we apply PCA as a pre-processing step [@article67, @article68, @article69].
-This step helps in noise reduction by identifying principal components that represent directions of maximum variance, capturing essential patterns in the data.
-
-### Constructing the 2D model
-
-**Step 1: Computing the hexagonal grid configuration**
+In our algorithm, hexagonal binning is utilized to create a low-dimensional manifold.
 
@@
 
-In this step, the algorithm proceeds to triangulate the centroids (see @fig-meth (Step 3)) of the hex bins using the `tripack` package in R [@article54].
+In this step, the algorithm proceeds to triangulate the centroids (see @fig-meth (Step 3)) of the hex bins.
+Triangulation is a fundamental process in computational geometry and computer graphics that involves dividing a set of points in a given space into interconnected triangles [@article30].
+One common algorithm used for triangulation is Delaunay triangulation [@article26], where points are connected in a way that maximizes the minimum angles of the resulting triangles, leading to a more regular and well-conditioned triangulation.
+In our algorithm, triangulation helps us identify geometric relationships and patterns in the data.
+
 Since we are working with the centroids of regular hexagonal bins, the resulting mesh will predominantly comprise equal-sized regular triangles.
@@
 ```
+
 ### Tunning the model
 
-### Summaries of the model {#sec-diagnostics}
+### Summaries of the model {#sec-summary}
 
@@
 
-In our prediction process, we employ the K-nearest neighbors (KNN) algorithm to predict the hex bin IDs.
-Specifically, for each new high-dimensional data point, the algorithm assigns the first neighboring hex bin ID as the predicted hex bin ID.
-
-<!-- move to goodness of fit stat-->
+In this context, the term "prediction values" refers to the 2D coordinates predicted for the NLDR technique.
+The approach involves employing the K-nearest neighbors (KNN) algorithm to identify the nearest hexagonal bin centroid in the 2D space.
+Subsequently, the coordinates of this centroid are assigned as the low-dimensional predicted values for the test data in 2D space.
+It is noteworthy that traditional NLDR methods, such as t-SNE, often lack a direct predict function, making our approach valuable for generating predicted values in the absence of such functionalities.
+
+The concept of "residuals" is pivotal in evaluating the accuracy of representing bin centroids in high dimensions.
+To quantify this accuracy, we introduce an error metric, which measures the sum of squared differences between the high-dimensional data (\(x_{ij}\)) and the predicted bin centroid data in high-dimensional space (\(C_{x_ij}\)) across all bins and dimensions.
+Mathematically, this error is expressed as:
+
+$$
+\text{Error} = \sum_{j = 1}^{n}\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2
+$$ {#eq-equation11}
+
+Here, \(n\) represents the number of bins, \(p\) represents the dimensions, \(x_{ij}\) is the actual high-dimensional data, and \(C_{x_ij}\) is the predicted bin centroid data in high dimensions.
+
+The error metric outlined above provides valuable insights into the overall accuracy of our predictive model.
+By quantifying the squared deviations between the actual and predicted values across all bins and dimensions, we gain a comprehensive understanding of how well our method captures and represents the underlying structure of the data in the reduced 2D space.
+This assessment is crucial for evaluating the efficacy of our NLDR technique in preserving the essential information present in the original high-dimensional data.
+
+#### Goodness of fit statistics
+
 Moving on to the assessment of prediction accuracy, we calculate the Mean Squared Error (MSE).
@@
 
-Here, $b$ signifies the total number of bins, $p$ denotes the number of dimensions in the high-dimensional (high-D) data, and $n$ represents the number of observations.
+Here, $b$ signifies the total number of bins, $p$ denotes the number of dimensions in the high-dimensional data, and $n$ represents the number of observations.
 
@@
 
-#### Goodness of fit statistics
-
-In this analysis, we're looking at how well our model is doing in two aspects.
+<!--In this analysis, we're looking at how well our model is doing in two aspects.
 First, we're checking the distances in our lower-dimensional space (see @fig-diag (a)).
@@
 
-<!--### Software
+### Software
 
@@
 
-### MNIST
-
-The MNIST dataset is a widely recognized benchmark in the realm of computer vision and machine learning, which is a collection of 28x28 pixel grayscale images of handwritten digits (0 through 9), serving as a foundational resource for developing and evaluating various image classification algorithms.
-With its simplicity and clarity, MNIST provides an ideal platform for researchers and practitioners to experiment with and benchmark their models.
+### Single-Cell Tagged Reverse Transcription sequencing data of mouse
+
+The Zeisel mouse brain dataset, obtained through Spatial Transcriptomics (STRT-Seq).
+Within this dataset, information is collected from a substantial 2,816 individual mouse brain cells.
+Each of these cells acts as a molecular snapshot, capturing the distinctive genetic activity within various cell types.
+This diversity spans neurons, glial cells, and other essential components of the brain, offering a comprehensive view of the cellular tapestry.
+
+What makes this dataset particularly valuable is its ability to shed light on the spatial distribution of cells.
+Researchers can explore how gene expression patterns vary across different regions of the mouse brain, unlocking insights into the functional specialization of these regions and the intricate networks that underpin neural processes.
 

### 2024-01-09 paper.tex
Selected another application with mouse barin data

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, sharp corners, boxrule=0pt, interior hidden, frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, sharp corners, breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, enhanced]}{\end{tcolorbox}}\fi
 
@@
 
-\hypertarget{non-parametric-models}{%
-\subsection{Non-parametric models}\label{non-parametric-models}}
-
-\hypertarget{hexagonal-binning}{%
-\subsubsection{Hexagonal binning}\label{hexagonal-binning}}
+\hypertarget{sec-methods}{%
+\section{Methodology}\label{sec-methods}}
+
+Our algorithm comprises two main phases: (1) generate the model in the
+2D space, and (2) generate the model in the high-D space. These two
+phases are described in details in this section.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\textwidth,height=1\textheight]{figures/workflow.png}
+
+}
+
+\caption{\label{fig-meth}A flow diagram detailing the steps taken to
+create the low-dimensional manifold in the high dimensional space. There
+are two basic phases, one to generate the model in the 2D space, and
+other to generate the model in the high-D space.}
+
+\end{figure}
+
+\hypertarget{preprocessing-steps}{%
+\subsection{Preprocessing steps}\label{preprocessing-steps}}
+
+To tackle the complexities and noise in high-dimensional data, we apply
+PCA as a pre-processing step \citep[\citet{article68},
+\citet{article69}]{article67}. This step helps in noise reduction by
+identifying principal components that represent directions of maximum
+variance, capturing essential patterns in the data.
+
+\hypertarget{constructing-the-2d-model}{%
+\subsection{Constructing the 2D model}\label{constructing-the-2d-model}}
+
+\textbf{Step 1: Computing the hexagonal grid configuration}
 
@@
 accurate representation of the data. In our algorithm, hexagonal binning
-is utilized to create a low-dimensional manifold and facilitate data
-visualization in the original high-dimensional space.
-
-\hypertarget{triangulation}{%
-\subsubsection{Triangulation}\label{triangulation}}
-
-Triangulation is a fundamental process in computational geometry and
-computer graphics that involves dividing a set of points in a given
-space into interconnected triangles \citep{article30} finding
-applications in mesh generation and representing complex geometric
-structures efficiently and discretely. One common algorithm used for
-triangulation is Delaunay triangulation \citep{article26}, where points
-are connected in a way that maximizes the minimum angles of the
-resulting triangles, leading to a more regular and well-conditioned
-triangulation.
-
-In our algorithm, we are interested in triangulation as it plays a key
-role in providing a clear and concise representation of the data. By
-converting the data points into interconnected triangles, we can
-visualize and analyze the data in a more structured and organized
-manner. Triangulation helps us identify geometric relationships and
-patterns in the data, enabling us to make informed decisions and gain
-valuable insights from the information presented in the triangular form.
-
-\hypertarget{sec-methods}{%
-\section{Methodology}\label{sec-methods}}
-
-Our algorithm comprises two main phases: (1) generate the model in the
-2D space, and (2) generate the model in the high-D space. These two
-phases are described in details in this section.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\textwidth,height=1\textheight]{figures/workflow.png}
-
-}
-
-\caption{\label{fig-meth}A flow diagram detailing the steps taken to
-create the low-dimensional manifold in the high dimensional space. There
-are two basic phases, one to generate the model in the 2D space, and
-other to generate the model in the high-D space.}
-
-\end{figure}
-
-\hypertarget{preprocessing-steps}{%
-\subsection{Preprocessing steps}\label{preprocessing-steps}}
-
-To tackle the complexities and noise in high-dimensional data, we apply
-PCA as a pre-processing step \citep[\citet{article68},
-\citet{article69}]{article67}. This step helps in noise reduction by
-identifying principal components that represent directions of maximum
-variance, capturing essential patterns in the data.
-
-\hypertarget{constructing-the-2d-model}{%
-\subsection{Constructing the 2D model}\label{constructing-the-2d-model}}
-
-\textbf{Step 1: Computing the hexagonal grid configuration}
+is utilized to create a low-dimensional manifold.
 
@@
 In this step, the algorithm proceeds to triangulate the centroids (see
-Figure~\ref{fig-meth} (Step 3)) of the hex bins using the
-\texttt{tripack} package in R \citep{article54}. Since we are working
-with the centroids of regular hexagonal bins, the resulting mesh will
-predominantly comprise equal-sized regular triangles. However, the
-triangulation also helps span any gaps that may exist between clusters
-of points, allowing for a more complete and interconnected
-representation of the data.
+Figure~\ref{fig-meth} (Step 3)) of the hex bins. Triangulation is a
+fundamental process in computational geometry and computer graphics that
+involves dividing a set of points in a given space into interconnected
+triangles \citep{article30}. One common algorithm used for triangulation
+is Delaunay triangulation \citep{article26}, where points are connected
+in a way that maximizes the minimum angles of the resulting triangles,
+leading to a more regular and well-conditioned triangulation. In our
+algorithm, triangulation helps us identify geometric relationships and
+patterns in the data.
+
+Since we are working with the centroids of regular hexagonal bins, the
+resulting mesh will predominantly comprise equal-sized regular
+triangles. However, the triangulation also helps span any gaps that may
+exist between clusters of points, allowing for a more complete and
+interconnected representation of the data.
 
@@
 
-\hypertarget{sec-diagnostics}{%
-\subsection{Summaries of the model}\label{sec-diagnostics}}
+\hypertarget{sec-summary}{%
+\subsection{Summaries of the model}\label{sec-summary}}
 
@@
 
-In our prediction process, we employ the K-nearest neighbors (KNN)
-algorithm to predict the hex bin IDs. Specifically, for each new
-high-dimensional data point, the algorithm assigns the first neighboring
-hex bin ID as the predicted hex bin ID.
+In this context, the term ``prediction values'' refers to the 2D
+coordinates predicted for the NLDR technique. The approach involves
+employing the K-nearest neighbors (KNN) algorithm to identify the
+nearest hexagonal bin centroid in the 2D space. Subsequently, the
+coordinates of this centroid are assigned as the low-dimensional
+predicted values for the test data in 2D space. It is noteworthy that
+traditional NLDR methods, such as t-SNE, often lack a direct predict
+function, making our approach valuable for generating predicted values
+in the absence of such functionalities.
+
+The concept of ``residuals'' is pivotal in evaluating the accuracy of
+representing bin centroids in high dimensions. To quantify this
+accuracy, we introduce an error metric, which measures the sum of
+squared differences between the high-dimensional data ((x\_\{ij\})) and
+the predicted bin centroid data in high-dimensional space
+((C\_\{x\_ij\})) across all bins and dimensions. Mathematically, this
+error is expressed as:
+
+\begin{equation}\protect\hypertarget{eq-equation11}{}{
+\text{Error} = \sum_{j = 1}^{n}\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2
+}\label{eq-equation11}\end{equation}
+
+Here, (n) represents the number of bins, (p) represents the dimensions,
+(x\_\{ij\}) is the actual high-dimensional data, and (C\_\{x\_ij\}) is
+the predicted bin centroid data in high dimensions.
+
+The error metric outlined above provides valuable insights into the
+overall accuracy of our predictive model. By quantifying the squared
+deviations between the actual and predicted values across all bins and
+dimensions, we gain a comprehensive understanding of how well our method
+captures and represents the underlying structure of the data in the
+reduced 2D space. This assessment is crucial for evaluating the efficacy
+of our NLDR technique in preserving the essential information present in
+the original high-dimensional data.
+
+\hypertarget{goodness-of-fit-statistics}{%
+\subsubsection{Goodness of fit
+statistics}\label{goodness-of-fit-statistics}}
 
@@
 Here, \(b\) signifies the total number of bins, \(p\) denotes the number
-of dimensions in the high-dimensional (high-D) data, and \(n\)
-represents the number of observations.
+of dimensions in the high-dimensional data, and \(n\) represents the
+number of observations.
 
@@
 \caption{\label{fig-diagnosticplt}The .}
-
-\end{figure}
-
-\hypertarget{goodness-of-fit-statistics}{%
-\subsubsection{Goodness of fit
-statistics}\label{goodness-of-fit-statistics}}
-
-In this analysis, we're looking at how well our model is doing in two
-aspects. First, we're checking the distances in our lower-dimensional
-space (see Figure~\ref{fig-diag} (a)). We want to see if the distances
-between different types of edges are similar. If they are, it means our
-model is doing a good job of keeping the relationships consistent. If
-there are big differences, it could mean our model is struggling with
-certain types of edges.
-
-Next, we're examining how our model handles the relationship between
-high-dimensional and low-dimensional distances for each type of edge
-(see Figure~\ref{fig-diag} (b)). If our data points nicely line up
-around the expected line, it tells us that our model is preserving the
-original structure well. If they're scattered, it could mean our model
-is having trouble with specific data points. This two-part analysis
-helps us really understand how our model fits the data.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diag-1.pdf}
-
-}
-
-\caption{\label{fig-diag}The two diagnostic plot assess the
-effectiveness of the high-D representation algorithm performed for the
-\textbf{s\_curve\_noise\_umap} data set\(\colon\) (a) visualizing the
-differences between low-D distances and (b) visualizing the relationship
-between high-D (\(d^{(p)}\)) and low-D distances (\(d^{(2)}\)).}
 
@@
 \hypertarget{sec-applications}{%
-\section{Examples/Applications}\label{sec-applications}}
+\section{Applications}\label{sec-applications}}
 
@@
 
+\hypertarget{zeisel-mouse-brain-strt-seq}{%
+\subsection{Zeisel mouse brain
+(STRT-Seq)}\label{zeisel-mouse-brain-strt-seq}}
+
+The Zeisel mouse brain dataset, obtained through Spatial Transcriptomics
+(STRT-Seq). Within this dataset, information is collected from a
+substantial 2,816 individual mouse brain cells. Each of these cells acts
+as a molecular snapshot, capturing the distinctive genetic activity
+within various cell types. This diversity spans neurons, glial cells,
+and other essential components of the brain, offering a comprehensive
+view of the cellular tapestry.
+
+What makes this dataset particularly valuable is its ability to shed
+light on the spatial distribution of cells. Researchers can explore how
+gene expression patterns vary across different regions of the mouse
+brain, unlocking insights into the functional specialization of these
+regions and the intricate networks that underpin neural processes.
+
 \hypertarget{sec-discussion}{%

### 2024-01-10 paper.qmd
Add hexbin grid generation with different bins into another script and generated the grids from the source file

@@
 library(patchwork)
+library(langevitour)
 
@@
 
-set.seed(20230531)
+set.seed(20240110)
 
@@
 
-tSNE_data <- Fit_tSNE(training_data, opt_perplexity = calculate_effective_perplexity(training_data), with_seed = 20230531)
+tSNE_data <- Fit_tSNE(training_data, opt_perplexity = calculate_effective_perplexity(training_data), with_seed = 20240110)
 #(perplexity: ", calculate_effective_perplexity(data), ")
@@
 
-UMAP_fit <- umap(training_data |> dplyr::select(-ID), n_neighbors = 50, n_components =  2)
-
-UMAP_data <- UMAP_fit$layout |>
-  as.data.frame()
-names(UMAP_data)[1:(ncol(UMAP_data))] <- paste0(rep("UMAP",(ncol(UMAP_data))), 1:(ncol(UMAP_data)))
-
-UMAP_data <- UMAP_data |>
-  mutate(ID = training_data$ID)
+UMAP_data <- read_rds(file = "data/s_curve/s_curve_umap.rds")
 
@@
 #(knn: 5)
-# PHATE_data <- Fit_PHATE(training_data, knn = 5, with_seed = 20230531)
+# PHATE_data <- Fit_PHATE(training_data, knn = 5, with_seed = 20240110)
 # write_csv(PHATE_data, paste0(here::here(), "/data/phate_data_s_curve.csv"))
@@
 #| warning: false
- # Loop over loop.vector
-  #i <- 1
-  num_bins <- 3
-  shape_val <- calculate_effective_shape_value(.data = UMAP_data, x = UMAP1, y = UMAP2)
-
-  hexbin_data_object_loop <- extract_hexbin_centroids(UMAP_data, num_bins, shape_val)
-
-  df_bin_centroids_loop <- hexbin_data_object_loop$hexdf_data
-
-  ## Data set with all possible centroids in the hexagonal grid
-
-  full_centroid_df_loop <- generate_full_grid_centroids(df_bin_centroids_loop)
-
-  ## To map hexID to hexbin centroids in the full grid
-
-  vec1 <- stats::setNames(rep("", 2), c("x", "y"))  ## Define column names
-
-  full_grid_with_hexbin_id_loop <- dplyr::bind_rows(vec1)[0, ]
-  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
-    dplyr::mutate_if(is.character, as.numeric)
-
-  for(i in 1:length(sort(unique(full_centroid_df_loop$y)))){
-
-    ## Filter the data set with specific y value
-    specific_y_val_df_loop <- full_centroid_df_loop |>
-      dplyr::filter(y == sort(unique(full_centroid_df_loop$y))[i])
-
-    ordered_x_df_loop <- specific_y_val_df_loop |>
-      dplyr::arrange(x)
-
-    full_grid_with_hexbin_id_loop <- dplyr::bind_rows(full_grid_with_hexbin_id_loop, ordered_x_df_loop)
-
-  }
-
-  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
-    dplyr::mutate(hexID = row_number())
-
-  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
-    dplyr::rename("c_x" = "x",
-           "c_y" = "y")
-
-  full_grid_with_hexbin_id_loop <- dplyr::full_join(full_grid_with_hexbin_id_loop, df_bin_centroids_loop, by = c("hexID" = "hexID")) |>
-    dplyr::select(-c(x, y))
-
-  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
-      dplyr::mutate(std_counts = counts/max(counts, na.rm = TRUE))
-
-  ## Generate all coordinates of hexagons
-  hex_grid_loop <- full_hex_grid(full_centroid_df_loop)
-
-  full_grid_with_polygon_id_df_loop <- map_polygon_id(full_grid_with_hexbin_id_loop, hex_grid_loop)
-
-  full_grid_with_hexbin_id_rep_loop <- full_grid_with_polygon_id_df_loop |>
-  dplyr::slice(rep(1:n(), each = 6)) |>
-  dplyr::arrange(polygon_id)
-
-  hex_full_count_df_loop <- dplyr::bind_cols(hex_grid_loop, full_grid_with_hexbin_id_rep_loop)
-
-  p1 <-  ggplot(data = hex_full_count_df_loop, aes(x = x, y = y)) +
+hex_full_count_df_loop1 <- read_rds("data/s_curve/s_curve_hex_3.rds")
+
+p1 <-  ggplot(data = hex_full_count_df_loop1, aes(x = x, y = y)) +
       geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
@@
 #| warning: false
- # Loop over loop.vector
-  #i <- 1
-  num_bins <- 8
-  shape_val <- calculate_effective_shape_value(.data = UMAP_data, x = UMAP1, y = UMAP2)
-
-  hexbin_data_object_loop <- extract_hexbin_centroids(UMAP_data, num_bins, shape_val)
-
-  df_bin_centroids_loop <- hexbin_data_object_loop$hexdf_data
-
-  ## Data set with all possible centroids in the hexagonal grid
-
-  full_centroid_df_loop <- generate_full_grid_centroids(df_bin_centroids_loop)
-
-  ## To map hexID to hexbin centroids in the full grid
-
-  vec1 <- stats::setNames(rep("", 2), c("x", "y"))  ## Define column names
-
-  full_grid_with_hexbin_id_loop <- dplyr::bind_rows(vec1)[0, ]
-  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
-    dplyr::mutate_if(is.character, as.numeric)
-
-  for(i in 1:length(sort(unique(full_centroid_df_loop$y)))){
-
-    ## Filter the data set with specific y value
-    specific_y_val_df_loop <- full_centroid_df_loop |>
-      dplyr::filter(y == sort(unique(full_centroid_df_loop$y))[i])
-
-    ordered_x_df_loop <- specific_y_val_df_loop |>
-      dplyr::arrange(x)
-
-    full_grid_with_hexbin_id_loop <- dplyr::bind_rows(full_grid_with_hexbin_id_loop, ordered_x_df_loop)
-
-  }
-
-  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
-    dplyr::mutate(hexID = row_number())
-
-  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
-    dplyr::rename("c_x" = "x",
-           "c_y" = "y")
-
-  full_grid_with_hexbin_id_loop <- dplyr::full_join(full_grid_with_hexbin_id_loop, df_bin_centroids_loop, by = c("hexID" = "hexID")) |>
-    dplyr::select(-c(x, y))
-
-  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
-      dplyr::mutate(std_counts = counts/max(counts, na.rm = TRUE))
-
-  ## Generate all coordinates of hexagons
-  hex_grid_loop <- full_hex_grid(full_centroid_df_loop)
-
-  full_grid_with_polygon_id_df_loop <- map_polygon_id(full_grid_with_hexbin_id_loop, hex_grid_loop)
-
-  full_grid_with_hexbin_id_rep_loop <- full_grid_with_polygon_id_df_loop |>
-  dplyr::slice(rep(1:n(), each = 6)) |>
-  dplyr::arrange(polygon_id)
-
-  hex_full_count_df_loop <- dplyr::bind_cols(hex_grid_loop, full_grid_with_hexbin_id_rep_loop)
-
-  p2 <-  ggplot(data = hex_full_count_df_loop, aes(x = x, y = y)) +
+hex_full_count_df_loop2 <- read_rds("data/s_curve/s_curve_hex_8.rds")
+
+p2 <-  ggplot(data = hex_full_count_df_loop2, aes(x = x, y = y)) +
   geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
@@
 #| warning: false
- # Loop over loop.vector
-  #i <- 1
-  num_bins <- 15
-  shape_val <- calculate_effective_shape_value(.data = UMAP_data, x = UMAP1, y = UMAP2)
-
-  hexbin_data_object_loop <- extract_hexbin_centroids(UMAP_data, num_bins, shape_val)
-
-  df_bin_centroids_loop <- hexbin_data_object_loop$hexdf_data
-
-  ## Data set with all possible centroids in the hexagonal grid
-
-  full_centroid_df_loop <- generate_full_grid_centroids(df_bin_centroids_loop)
-
-  ## To map hexID to hexbin centroids in the full grid
-
-  vec1 <- stats::setNames(rep("", 2), c("x", "y"))  ## Define column names
-
-  full_grid_with_hexbin_id_loop <- dplyr::bind_rows(vec1)[0, ]
-  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
-    dplyr::mutate_if(is.character, as.numeric)
-
-  for(i in 1:length(sort(unique(full_centroid_df_loop$y)))){
-
-    ## Filter the data set with specific y value
-    specific_y_val_df_loop <- full_centroid_df_loop |>
-      dplyr::filter(y == sort(unique(full_centroid_df_loop$y))[i])
-
-    ordered_x_df_loop <- specific_y_val_df_loop |>
-      dplyr::arrange(x)
-
-    full_grid_with_hexbin_id_loop <- dplyr::bind_rows(full_grid_with_hexbin_id_loop, ordered_x_df_loop)
-
-  }
-
-  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
-    dplyr::mutate(hexID = row_number())
-
-  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
-    dplyr::rename("c_x" = "x",
-           "c_y" = "y")
-
-  full_grid_with_hexbin_id_loop <- dplyr::full_join(full_grid_with_hexbin_id_loop, df_bin_centroids_loop, by = c("hexID" = "hexID")) |>
-    dplyr::select(-c(x, y))
-
-  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
-      dplyr::mutate(std_counts = counts/max(counts, na.rm = TRUE))
-
-  ## Generate all coordinates of hexagons
-  hex_grid_loop <- full_hex_grid(full_centroid_df_loop)
-
-  full_grid_with_polygon_id_df_loop <- map_polygon_id(full_grid_with_hexbin_id_loop, hex_grid_loop)
-
-  full_grid_with_hexbin_id_rep_loop <- full_grid_with_polygon_id_df_loop |>
-  dplyr::slice(rep(1:n(), each = 6)) |>
-  dplyr::arrange(polygon_id)
-
-  hex_full_count_df_loop <- dplyr::bind_cols(hex_grid_loop, full_grid_with_hexbin_id_rep_loop)
-
-  p3 <-  ggplot(data = hex_full_count_df_loop, aes(x = x, y = y)) +
+hex_full_count_df_loop3 <- read_rds("data/s_curve/s_curve_hex_15.rds")
+
+p3 <-  ggplot(data = hex_full_count_df_loop3, aes(x = x, y = y)) +
   geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
@@
 
-UMAP_fit <- umap(training_data |> dplyr::select(-ID), n_neighbors = 50, n_components =  2)
-
-UMAP_data <- UMAP_fit$layout |>
-  as.data.frame()
-names(UMAP_data)[1:(ncol(UMAP_data))] <- paste0(rep("UMAP",(ncol(UMAP_data))), 1:(ncol(UMAP_data)))
-
-UMAP_data <- UMAP_data |>
-  mutate(ID = training_data$ID)
+# UMAP_fit <- umap(training_data |> dplyr::select(-ID), n_neighbors = 50, n_components =  2)
+# 
+# UMAP_data <- UMAP_fit$layout |>
+#   as.data.frame()
+# names(UMAP_data)[1:(ncol(UMAP_data))] <- paste0(rep("UMAP",(ncol(UMAP_data))), 1:(ncol(UMAP_data)))
+# 
+# UMAP_data <- UMAP_data |>
+#   mutate(ID = training_data$ID)
 
@@
 #| echo: false
-
-a1 <- colour_long_edges(.data = distance, benchmark_value = 13, 
+# distance |> dplyr::filter(distance == max(distance))
+# df_bin_centroids |> dplyr::filter(row_number() %in% c(10, 65))
+
+a1 <- colour_long_edges(.data = distance, benchmark_value = 15, 
                   triangular_object = tr1_object, distance_col = distance) + 
@@
 
-UMAP_data_with_hb_id_selected <- UMAP_data_with_hb_id |> dplyr::filter(hb_id %in% c(14, 177)) |>
-  dplyr::mutate(select_point = if_else(hb_id == 14, "data1", "data2"))
+UMAP_data_with_hb_id_selected <- UMAP_data_with_hb_id |> dplyr::filter(hb_id %in% c(31, 239)) |>
+  dplyr::mutate(select_point = if_else(hb_id == 31, "data1", "data2"))
 
@@
   scale_color_manual(values=c("#e6550d", "#000000")) +
-  geom_point(data = df_bin_centroids_all |> dplyr::filter(hexID %in% c(14, 177)), aes(x = x, y = y), colour = "#33a02c") +
+  geom_point(data = df_bin_centroids_all |> dplyr::filter(hexID %in% c(31, 239)), aes(x = x, y = y), colour = "#33a02c") +
   scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
@@
 df1 <- df_all |> 
-  dplyr::filter(hb_id == "14") |>
+  dplyr::filter(hb_id == "31") |>
   dplyr::select(tidyselect::starts_with("x")) |>
@@
 df2 <- df_all |> 
-  dplyr::filter(hb_id == "177") |>
+  dplyr::filter(hb_id == "239") |>
   dplyr::select(tidyselect::starts_with("x")) |>
@@
 df3 <- df_all |> 
-  dplyr::filter(!(hb_id %in% c("14", "177"))) |>
+  dplyr::filter(!(hb_id %in% c("31", "239"))) |>
   dplyr::select(tidyselect::starts_with("x")) |>
@@
 distance_df_small_edges <- distance |> 
-  dplyr::filter(from == 3) |>
-  dplyr::filter(to == 49) |>
+  dplyr::filter(from == 10) |>
+  dplyr::filter(to == 65) |>
   head(1)
@@
 #| echo: false
-predict_UMAP_df <- predict(UMAP_fit, test_data |> dplyr::select(-ID)) |>
-  as.data.frame()
-
-names(predict_UMAP_df)[1:(ncol(predict_UMAP_df))] <- paste0(rep("UMAP",(ncol(predict_UMAP_df))), 1:(ncol(predict_UMAP_df)))
-
-predict_UMAP_df <- predict_UMAP_df |>
-  mutate(ID = test_data$ID)
+predict_UMAP_df <- read_rds("data/s_curve/s_curve_umap_predict.rds")
 ```
@@
 
-In this section, we demonstrate the effectiveness of our approach through the use of simulation data.
+<!--six different clusters-->
+
+<!--Gaussian globe or sphere-->
+
+<!--In this section, we demonstrate the effectiveness of our approach through the use of simulation data.
 These simulations are designed to have known cluster structures and underlying data geometries.
@@
 #| echo: false 
-tSNE_data_gau <- Fit_tSNE(training_data_5 |> dplyr::select(-ID, -cluster), with_seed = 20230531)
+tSNE_data_gau <- Fit_tSNE(training_data_5 |> dplyr::select(-ID, -cluster), with_seed = 20240110)
 ```
@@
 :::
+-->
 

### 2024-01-10 paper.tex
Add hexbin grid generation with different bins into another script and generated the grids from the source file

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, sharp corners, breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, interior hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}, frame hidden, boxrule=0pt, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 
-In this section, we demonstrate the effectiveness of our approach
-through the use of simulation data. These simulations are designed to
-have known cluster structures and underlying data geometries. We begin
-with a straightforward example where we generate spherical clusters with
-a doublet structure. This allows us to assess how well our algorithm can
-identify and represent these known structures in the low-dimensional
-space.
-
-We then proceed to a more complex example where the data exhibits a more
-intricate geometry. This simulation provides a challenging test for our
-approach to capture and preserve the underlying data structure
-accurately.
-
-We \emph{strongly} recommend viewing the linked videos for each study
-while reading. Links to the videos are available in the figures for each
-example. The videos show the visual appearance of the
-\textbf{langevitour} interface with low-dimensional view and how we can
-interact with the tour via the controls.
-
-\textbf{Example 1: Exploring spherical Gaussian clusters with doublet
-structure}
-
-The next data set consists of two distinct types of clusters embedded in
-a 10-\(d\) space. The first set of clusters includes three Gaussian
-clusters with equal variance. Each of these Gaussian clusters is
-equidistant from the others, forming a well-separated and symmetrical
-arrangement. The second set of clusters consists of three clusters with
-a doublet structure. These clusters are also embedded in the 10-\(d\)
-space, but each is positioned in the middle of its parent Gaussian
-clusters. Moreover, each doublet cluster contains an average number of
-points from its corresponding parent Gaussian clusters.
-
-For analysis, we run tSNE on the training data with a perplexity setting
-18. Figure~\ref{fig-example3} (a) displays the tSNE results, indicating
-that the sub-clusters have been correctly identified. However, the
-relative locations of certain clusters to each other appear distorted in
-the low-dimensional view, such as the red and pink clusters being far
-apart.
-
-Nonetheless, we find no apparent impact on the perception of the data
-structure when we examine the tour view and the model (see
-Figure~\ref{fig-example3}) through the linked video (see
-Figure~\ref{fig-exp1_sc}). The model demonstrates accurate
-identification of clusters and captures the internal variations within
-each cluster, reflecting the inherent variance present in the data. The
-configurations used for the analysis are detailed in
-\textbf{?@tbl-table02}. Additionally, the representation of the parent
-clusters in the low-dimensional space appears stretched, aligning with
-the variance of these clusters. On the other hand, the doublet clusters
-are squeezed, reflecting their specific characteristics and relationship
-to their parent Gaussian clusters.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-example3-1.pdf}
-
-}
-
-\caption{\label{fig-example3}Visualization and refinement of
-low-dimensional representation for sub-clustered data\(\colon\) (a) 2D
-layout of tSNE (perplexity = 18), (b) triangular mesh, and (c)
-triangular mesh colored by edge type. Edges with lengths less than the
-benchmark value of \(2\) are classifiedas small, while edges with
-lengths greater than or equal to this value are labeled as long. The
-combination of tSNE, hexagonal binning, and triangulation provides the
-effectiveness of our approach in visualizing and refining the
-low-dimensional representation of the sub-clustered data.}
-
-\end{figure}
-
-\begin{figure}
-
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/example_1a.png}
-
-}
-
-}
-
-\subcaption{\label{fig-exp1_sc1}}
-\end{minipage}%
-%
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/example_1b.png}
-
-}
-
-}
-
-\subcaption{\label{fig-exp1_sc2}}
-\end{minipage}%
-%
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/example_1c.png}
-
-}
-
-}
-
-\subcaption{\label{fig-exp1_sc3}}
-\end{minipage}%
-
-\caption{\label{fig-exp1_sc}Screen shots of the \textbf{langevitour}
-with the low-dimensional view applied to sub-clustered data set, a video
-of the tour animation is available at
-(\url{https://drive.google.com/file/d/15nAbmc4UoVlUyFcGtnlnmyxrJEbK39nx/view}).
-Combination of the \textbf{langevitour} and low-dimensional
-representation of the sub-clustered data provides the effectiveness of
-our approach.}
-
-\end{figure}
-
-\textbf{Example 2: Exploring data with piecewise linear structure}
-
-In our next exploration, we delve into simulated noisy tree-structured
-data designed to mimic branching trajectories of cell differentiation.
-The simulation is crafted to resemble the hierarchical clustering
-pattern observed in cell differentiation trajectories, where mature
-cells are akin to the tips of branches.
-
-To begin our analysis, we apply principal components to the data and
-focus on the first ten principal components, collectively accounting for
-approximately 68\% of the variance explained in the data set. Then, we
-run TriMAP with the default parameters on the training data (n\_inliers:
-5, n\_outliers: 4, and n\_random: 3). Through this process, we create a
-low-dimensional manifold using the configurations detailed in
-\textbf{?@tbl-table02}. Finally, we present the model along with the
-data using the tour view (Figure Figure~\ref{fig-exp2_sc}), allowing us
-to visualize and explore the intricate relationships and structures
-present in the data's low-dimensional representation.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-example2-1.pdf}
-
-}
-
-\caption{\label{fig-example2}Visualization and refinement of
-low-dimensional representation for tree structured data\(\colon\) (a) 2D
-layout of TriMAP (n-inliers = 5, n-outliers = 4, n-random = 3), (b)
-triangular mesh, and (c) triangular mesh colored by edge type. Edges
-with lengths less than the benchmark value of \(2\) are classified as
-small, while edges with lengths greater than or equal to this value are
-labeled as long. The combination of TriMAP, hexagonal binning, and
-triangulation provides the effectiveness of our approach in visualizing
-and refining the low-dimensional representation of the tree structured
-data.}
-
-\end{figure}
-
-\begin{figure}
-
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/example_2a.png}
-
-}
-
-}
-
-\end{minipage}%
-%
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/example_2b.png}
-
-}
-
-}
-
-\end{minipage}%
-%
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/example_2c.png}
-
-}
-
-}
-
-\end{minipage}%
-
-\caption{\label{fig-exp2_sc}Screen shots of the \textbf{langevitour}
-with the low-dimensional view applied to tree structured data set, a
-video of the tour animation is available at
-(\url{https://drive.google.com/file/d/1YQx4UGlzDify1mGqeDtlqTHncgqidDhq/view}).
-Combination of the \textbf{langevitour} and low-dimensional
-representation of the tree structured data provides the effectiveness of
-our approach.}
-
-\end{figure}
-
 \hypertarget{sec-applications}{%
@@
 
-\hypertarget{zeisel-mouse-brain-strt-seq}{%
-\subsection{Zeisel mouse brain
-(STRT-Seq)}\label{zeisel-mouse-brain-strt-seq}}
+\hypertarget{single-cell-tagged-reverse-transcription-sequencing-data-of-mouse}{%
+\subsection{Single-Cell Tagged Reverse Transcription sequencing data of
+mouse}\label{single-cell-tagged-reverse-transcription-sequencing-data-of-mouse}}
 

### 2024-01-10 paper.qmd
Update the paper with sumaary for the example, s-curve data set, for s-curve dataset added summary with different parameter setting in UMAP, generate required data sets beforehand and store in necessary folders, generate scripts to generate embeddings

@@
 library(phateR)
-library(reticulate)
 library(patchwork)
 library(langevitour)
+library(colorspace)
 
@@
 
-#### Introduction to UMAP
-
-The UMAP algorithm constructs a fuzzy topological representation of the high-dimensional data and then optimizes the low-dimensional embedding to be as similar to this representation as possible [@Leland2018].
-
-The UMAP algorithm constructs a weighted k-nearest neighbors graph based on the local similarity of data points in the original high-dimensional space.
-The weighted graph is defined by a conditional probability distribution $P_{ij}$, which represents the probability of observing data point $x_i$ given data point $x_j$ as a neighbor:
-
-$$
-P_{ij} = \frac{\exp(-\lVert x_i - x_j \rVert^2)}{\sum_{k \neq i} \exp(-\lVert x_i - x_k \rVert^2)}
-$$
-
-Similarly, a fuzzy topological representation of the data is constructed in the low-dimensional space with conditional probability distribution $Q_{ij}$:
-
-$$
-Q_{ij} = \frac{\exp(-\lVert y_i - y_j \rVert^2)}{\sum_{k \neq i} \exp(-\lVert y_i - y_k \rVert^2)}
-$$
-
-where $y_i$ and $y_j$ are the corresponding low-dimensional representations of data points $x_i$ and $x_j$.
-
-UMAP minimizes the cross-entropy between the high-dimensional fuzzy topological representation $P_{ij}$ and the low-dimensional representation $Q_{ij}$ to find an optimal low-dimensional embedding.
-The loss function for UMAP is defined as:
-
-$$
-L = \sum_{i} \sum_{j} P_{ij} \log \frac{P_{ij}}{Q_{ij}}
-$$
-
-The algorithm uses stochastic gradient descent to optimize the positions of the data points in the low-dimensional space to minimize this loss function.
-
-In R, the `umap` package is often used to implement UMAP and generate results.
-
-The "n_neighbors" parameter in UMAP determines the number of nearest neighbors to consider for each data point during the construction of the fuzzy topological representation.
-Higher values of "n_neighbors" capture more global structure, while lower values emphasize local structure.
-Choosing an appropriate value for "n_neighbors" is crucial, as it affects the balance between local and global structure in the resulting low-dimensional representation.
+#### Non-linear dimension reduction techniques
+
+Non-linear dimension reduction techniques play a crucial role in the analysis and visualization of high-dimensional data, where the complexities of relationships among variables may not be adequately captured by linear methods.
+Among these techniques, tSNE stands out for its emphasis on preserving pairwise distances.
+By minimizing the divergence between probability distributions in both the high and low-dimensional spaces, t-SNE effectively reveals intricate structures and patterns within the data.
+Its application is widespread in tasks requiring the visualization of clusters and local relationships, though it does require careful consideration of the perplexity parameter for optimal results.
+
+UMAP is another powerful non-linear technique that strikes a balance between preserving local and global structures.
+Constructing a fuzzy topological representation using a weighted k-nearest neighbors graph, UMAP optimizes the low-dimensional embedding to resemble this representation.
+Known for its efficiency and scalability, UMAP is versatile across various scales of relationships in the data, although parameter sensitivity, particularly concerning the choice of neighbors, must be taken into account.
+
+For trajectory data, PHATE provides specialized capabilities.
+It models the affinity between data points, simulating a heat diffusion process to capture developmental processes, particularly in single-cell genomics.
+While PHATE excels in revealing trajectory structures and offering insights into cellular development, it necessitates careful parameter tuning due to its specialized nature.
+
+TriMAP adopts a unique approach by approximating the data manifold through the construction of a triangulated graph representation.
+This technique efficiently captures both global and local structures by representing the data as a network of triangles.
+TriMAP's strength lies in its ability to efficiently capture complex structures, albeit with sensitivity to parameter choices, including the number of neighbors.
+
+In contrast, PaCMAP introduces supervised learning to create a low-dimensional representation while preserving pair-wise relationships.
+Constructing a graph based on pair-wise distances, PaCMAP optimizes an embedding using a customizable loss function.
+Particularly notable is PaCMAP's flexibility in incorporating class labels or additional information to guide the embedding process, offering users a means to customize its behavior and performance.
 
@@
 
-data <- data |>
-  mutate(ID = row_number())
-
-data_split <- initial_split(data)
-training_data <- training(data_split) |>
-  arrange(ID)
-test_data <- testing(data_split) |>
-  arrange(ID)
+# data <- data |>
+#   mutate(ID = row_number())
+# 
+# data_split <- initial_split(data)
+# training_data <- training(data_split) |>
+#   arrange(ID)
+# test_data <- testing(data_split) |>
+#   arrange(ID)
+training_data <- read_rds("data/s_curve/s_curve_training.rds")
+test_data <- read_rds("data/s_curve/s_curve_test.rds")
 ```
@@
 
-tSNE_data <- Fit_tSNE(training_data, opt_perplexity = calculate_effective_perplexity(training_data), with_seed = 20240110)
+tSNE_data <- read_rds("data/s_curve/s_curve_tsne_27.rds")
 #(perplexity: ", calculate_effective_perplexity(data), ")
@@
 
-PHATE_data <- read_csv(paste0(here::here(), "/data/phate_data_s_curve.csv"))
+PHATE_data <- read_rds(file = "data/s_curve/s_curve_phate.rds")
 
@@
 
-TriMAP_data <- read_csv(paste0(here::here(), "/data/trimap_data_s_curve.csv"))
+TriMAP_data <- read_rds(file = "data/s_curve/s_curve_trimap.rds")
 
@@
 
-PaCMAP_data <- read_csv(paste0(here::here(), "/data/pacmap_data_s_curve.csv"))
+PaCMAP_data <- read_rds(file = "data/s_curve/s_curve_pacmap.rds")
 
@@
 #| echo: false
-#| fig-cap: "2D layouts from different NLDR techniques applied the same data: (a) tSNE (perplexity = 32), (b) UMAP (n_neighbors = 50), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2).
+#| fig-cap: "2D layouts from different NLDR techniques applied the same data: (a) tSNE (perplexity = 27), (b) UMAP (n_neighbors = 50), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2).
 Is there a best representation of the original data or are they all providing  equivalent information?"
@@
 #| warning: false
-hex_full_count_df_loop2 <- read_rds("data/s_curve/s_curve_hex_8.rds")
+hex_full_count_df_loop2 <- read_rds("data/s_curve/s_curve_hex_11.rds")
 
@@
 #| warning: false
-hex_full_count_df_loop3 <- read_rds("data/s_curve/s_curve_hex_15.rds")
+hex_full_count_df_loop3 <- read_rds("data/s_curve/s_curve_hex_20.rds")
 
@@
 #| fig-pos: H
-#| fig-cap: "Hexbin plots from different number of bins for the same **s_curve_noise_umap** data: (a) b = (4, 8), s = 2.031, (b) b = (9, 20), s = 2.031, and (c) b = (16, 36), s = 2.031.
+#| fig-cap: "Hexbin plots from different number of bins for the same **s_curve_noise_umap** data: (a) b = (4, 8), s = 1.643542, (b) b = (12, 22), s = 1.643542, and (c) b = (21, 40), s = 1.643542.
 The hexbins are colored based on the density of points, with yellow indicating higher point density and darker colors representing lower point density within each bin.
@@
 
-##########
-
-## Data set with all possible centroids in the hexagonal grid
-
-full_centroid_df <- generate_full_grid_centroids(df_bin_centroids)
-
-## To map hexID to hexbin centroids in the full grid
-
-vec1 <- stats::setNames(rep("", 2), c("x", "y"))  ## Define column names
-
-full_grid_with_hexbin_id <- dplyr::bind_rows(vec1)[0, ]
-full_grid_with_hexbin_id <- full_grid_with_hexbin_id |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for(i in 1:length(sort(unique(full_centroid_df$y)))){
-
-  ## Filter the data set with specific y value
-  specific_y_val_df <- full_centroid_df |>
-    dplyr::filter(y == sort(unique(full_centroid_df$y))[i])
-
-  ordered_x_df <- specific_y_val_df |>
-    dplyr::arrange(x)
-
-  full_grid_with_hexbin_id <- dplyr::bind_rows(full_grid_with_hexbin_id, ordered_x_df)
-
-}
-
-full_grid_with_hexbin_id <- full_grid_with_hexbin_id |>
-  dplyr::mutate(hexID = row_number())
-
-full_grid_with_hexbin_id <- full_grid_with_hexbin_id |>
-  dplyr::rename("c_x" = "x",
-         "c_y" = "y")
-
-full_grid_with_hexbin_id <- dplyr::full_join(full_grid_with_hexbin_id, df_bin_centroids, by = c("hexID" = "hexID")) |>
-  dplyr::select(-c(x, y)) #|> 
-  #dplyr::mutate(counts = tidyr::replace_na(counts, 0))
-
-full_grid_with_hexbin_id <- full_grid_with_hexbin_id |>
-    dplyr::mutate(std_counts = counts/max(counts, na.rm = TRUE))
-
-## Generate all coordinates of hexagons
-hex_grid <- full_hex_grid(full_centroid_df)
-
-full_grid_with_polygon_id_df <- map_polygon_id(full_grid_with_hexbin_id, hex_grid)
-
-full_grid_with_hexbin_id_rep <- full_grid_with_polygon_id_df |>
-  dplyr::slice(rep(1:n(), each = 6)) |>
-  dplyr::arrange(polygon_id)
-
-hex_full_count_df <- dplyr::bind_cols(hex_grid, full_grid_with_hexbin_id_rep)
+# ##########
+# 
+# ## Data set with all possible centroids in the hexagonal grid
+# 
+# full_centroid_df <- generate_full_grid_centroids(df_bin_centroids)
+# 
+# ## To map hexID to hexbin centroids in the full grid
+# 
+# vec1 <- stats::setNames(rep("", 2), c("x", "y"))  ## Define column names
+# 
+# full_grid_with_hexbin_id <- dplyr::bind_rows(vec1)[0, ]
+# full_grid_with_hexbin_id <- full_grid_with_hexbin_id |>
+#   dplyr::mutate_if(is.character, as.numeric)
+# 
+# for(i in 1:length(sort(unique(full_centroid_df$y)))){
+#   
+#   ## Filter the data set with specific y value
+#   specific_y_val_df <- full_centroid_df |>
+#     dplyr::filter(y == sort(unique(full_centroid_df$y))[i])
+#   
+#   ordered_x_df <- specific_y_val_df |>
+#     dplyr::arrange(x) 
+#   
+#   full_grid_with_hexbin_id <- dplyr::bind_rows(full_grid_with_hexbin_id, ordered_x_df)
+#   
+# }
+# 
+# 
+# full_grid_with_hexbin_id <- full_grid_with_hexbin_id |>
+#   dplyr::mutate(hexID = row_number())
+# 
+# full_grid_with_hexbin_id <- full_grid_with_hexbin_id |>
+#   dplyr::rename("c_x" = "x",
+#          "c_y" = "y") 
+# 
+# full_grid_with_hexbin_id <- dplyr::full_join(full_grid_with_hexbin_id, df_bin_centroids, by = c("hexID" = "hexID")) |>
+#   dplyr::select(-c(x, y)) #|> 
+#   #dplyr::mutate(counts = tidyr::replace_na(counts, 0))
+# 
+# full_grid_with_hexbin_id <- full_grid_with_hexbin_id |>
+#     dplyr::mutate(std_counts = counts/max(counts, na.rm = TRUE))
+# 
+# ## Generate all coordinates of hexagons
+# hex_grid <- full_hex_grid(full_centroid_df)
+# 
+# full_grid_with_polygon_id_df <- map_polygon_id(full_grid_with_hexbin_id, hex_grid)
+# 
+# full_grid_with_hexbin_id_rep <- full_grid_with_polygon_id_df |>
+#   dplyr::slice(rep(1:n(), each = 6)) |>
+#   dplyr::arrange(polygon_id)
+
+hex_full_count_df <- read_rds("data/s_curve/s_curve_hex_11.rds")
 
@@
 
+In our model tuning process, we strategically adjust three key parameters to optimize the performance and accuracy of our approach.
+They are (i) the total number of bins ($b$), (ii) a benchmark value to remove low-density hexagons, and (iii) a benchmark value to remove long edges.
+
+### Total number of bins
+
+The total number of bins is a crucial parameter that defines the granularity of our hexagonal grid.
+This parameter particularly influential in capturing the intricacies of the data structure.
+It is determined by multiplying the number of bins along the x-axis (\(b_1\)) with the number of bins along the y-axis (\(b_2\)), as per the formula:
+
+$$
+b = b_1 \times b_2
+$${#eq-equationp1}
+
+Here, \(b\) represents the total number of bins.
+Adjusting the parameter \(b_1\) provides control over the total bin count (\(b\)), allowing us to fine-tune the grid's configuration along the x-axis based on specific requirements.
+
+### Benchmark value to remove low-density hexagons
+
+Addressing low-density hexagons is a systematic process to handle sparsely represented data in certain regions.
+For each hex bin, we identify the six nearest hex bins using an equal 2D distance metric.
+Then, we calculate the mean density, as outlined in the equations:
+
+$$
+\text{standard count} = \frac{\text{count}}{\text{max count}} 
+$${#eq-equationp2}
+
+$$
+\text{mean density} = \frac{\text{standard count}}{6} 
+$${#eq-equationp3}
+
+The standard count is derived from the number of observations in the hex bins.
+By examining the distribution of mean densities and designating the first quartile as the benchmark value, hex bins with mean densities below this benchmark are removed.
+This process ensures the elimination of regions with insufficient data density, focusing on areas with more significant data representation and preserving the overall structure in the low-dimensional space.
+
+### Benchmark value to remove long edges
+
+The removal of long edges is a critical step to create a smoother representation  by iteratively eliminating hexagons with excessive distances between centroids.
+This process eliminates outliers and noise while preserving essential local relationships within the data.
+To achieve this, distances between vertices are sorted, and unique distance values are extracted.
+By setting a threshold based on the largest difference between consecutive distance values, long edges are identified and removed.
+This refinement step contributes to enhancing the quality of the triangular mesh, ensuring a more accurate representation of the data structure.
+
 ### Summaries of the model {#sec-summary}
@@
 ```{r}
-#| echo: false
-predict_UMAP_df <- read_rds("data/s_curve/s_curve_umap_predict.rds")
-```
-
-```{r}
-#| echo: false
-
-predict_umap_plot <- plot_UMAP_2D(UMAP_data) +
-  geom_point(data = predict_UMAP_df, aes(x = UMAP1, y = UMAP2), color = "red") + 
-  theme_linedraw() +
-    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-```
-
-<!--predict hexid for the test data -->
-
-```{r}
-#| echo: false
-
-pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = UMAP_data, num_bins = num_bins_x, shape_val = shape_value)
+#| warning: false
+#| echo: false
+
+data <- read_csv("data/s_curve.csv")
+data <- data |>
+  mutate(ID = row_number())
+
+## tSNE
+
+shape_value_curve <- calculate_effective_shape_value(.data = tSNE_data,
+                                                   x = tSNE1, y = tSNE2)
+
+num_bins_vec <- 1:10 ## Number of bins along the x-axis
+
+vec <- stats::setNames(rep("", 6), c("number_of_bins", "number_of_observations", "total_error", "totol_error_method_2", "totol_error_method_3", "total_mse"))  ## Define column names
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(num_bins_vec)) {
+
+  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = tSNE_data, nldr_df_test = tSNE_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "tSNE1", y = "tSNE2", col_start = "x")
   pred_df_training <- pred_df_training_object$pred_data
@@
 
-eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_x)
-
-pred_df_test_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = predict_UMAP_df, num_bins = num_bins_x, shape_val = shape_value)
+  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+
+  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+
+}
+
+## Add new column with data types
+
+eval_data_training <- eval_data_training |>
+  mutate(data_type = "training")
+
+MSE_df_1 <- eval_data_training |>
+  dplyr::mutate(method = "tSNE")
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## UMAP
+## Prediction
+
+shape_value_curve <- calculate_effective_shape_value(.data = UMAP_data,
+                                                   x = UMAP1, y = UMAP2)
+
+num_bins_vec <- 1:10 ## Number of bins along the x-axis
+
+vec <- stats::setNames(rep("", 6), c("number_of_bins", "number_of_observations", "total_error", "totol_error_method_2", "totol_error_method_3", "total_mse"))  ## Define column names
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(num_bins_vec)) {
+
+  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = UMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
+  pred_df_training <- pred_df_training_object$pred_data
+  centroid_df_training <- pred_df_training_object$df_bin_centroids
+  avg_df_training <- pred_df_training_object$df_bin
+
+  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+
+  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+
+}
+
+## Add new column with data types
+
+eval_data_training <- eval_data_training |>
+  mutate(data_type = "training")
+
+MSE_df_2 <- eval_data_training |>
+  dplyr::mutate(method = "UMAP")
+```
+
+```{r}
+#| warning: false
+#| echo: false
+## PAHTE
+## Prediction
+
+shape_value_curve <- calculate_effective_shape_value(.data = PHATE_data,
+                                                   x = PHATE1, y = PHATE2)
+
+num_bins_vec <- 1:10 ## Number of bins along the x-axis
+
+vec <- stats::setNames(rep("", 6), c("number_of_bins", "number_of_observations", "total_error", "totol_error_method_2", "totol_error_method_3", "total_mse"))  ## Define column names
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(num_bins_vec)) {
+
+  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = PHATE_data, nldr_df_test = PHATE_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "PHATE1", y = "PHATE2", col_start = "x")
+  pred_df_training <- pred_df_training_object$pred_data
+  centroid_df_training <- pred_df_training_object$df_bin_centroids
+  avg_df_training <- pred_df_training_object$df_bin
+
+  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+
+  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+
+}
+
+## Add new column with data types
+
+eval_data_training <- eval_data_training |>
+  mutate(data_type = "training")
+
+MSE_df_3 <- eval_data_training |>
+  dplyr::mutate(method = "PHATE")
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## TriMAP
+
+## Prediction
+
+shape_value_curve <- calculate_effective_shape_value(.data = TriMAP_data,
+                                                   x = TriMAP1, y = TriMAP2)
+
+num_bins_vec <- 1:10 ## Number of bins along the x-axis
+
+vec <- stats::setNames(rep("", 6), c("number_of_bins", "number_of_observations", "total_error", "totol_error_method_2", "totol_error_method_3", "total_mse"))  ## Define column names
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(num_bins_vec)) {
+
+  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = TriMAP_data, nldr_df_test = TriMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "TriMAP1", y = "TriMAP2", col_start = "x")
+  pred_df_training <- pred_df_training_object$pred_data
+  centroid_df_training <- pred_df_training_object$df_bin_centroids
+  avg_df_training <- pred_df_training_object$df_bin
+
+  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+
+  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+
+}
+
+## Add new column with data types
+
+eval_data_training <- eval_data_training |>
+  mutate(data_type = "training")
+
+MSE_df_4 <- eval_data_training |>
+  dplyr::mutate(method = "TriMAP")
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## PaCMAP
+
+## Prediction
+
+shape_value_curve <- calculate_effective_shape_value(.data = PaCMAP_data,
+                                                   x = PaCMAP1, y = PaCMAP2)
+
+num_bins_vec <- 1:10 ## Number of bins along the x-axis
+
+vec <- stats::setNames(rep("", 6), c("number_of_bins", "number_of_observations", "total_error", "totol_error_method_2", "totol_error_method_3", "total_mse"))  ## Define column names
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(num_bins_vec)) {
+
+  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = PaCMAP_data, nldr_df_test = PaCMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "PaCMAP1", y = "PaCMAP2", col_start = "x")
+  pred_df_training <- pred_df_training_object$pred_data
+  centroid_df_training <- pred_df_training_object$df_bin_centroids
+  avg_df_training <- pred_df_training_object$df_bin
+
+  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+
+  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+
+}
+
+## Add new column with data types
+
+eval_data_training <- eval_data_training |>
+  mutate(data_type = "training")
+
+MSE_df_5 <- eval_data_training |>
+  dplyr::mutate(method = "PaCMAP")
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+MSE_df <- dplyr::bind_rows(MSE_df_1, MSE_df_2, MSE_df_3, MSE_df_4, MSE_df_5)
+
+## To draw with AIC
+aic_plot <- ggplot(MSE_df |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
+                                                                                 y = total_error,
+                                                                                 color = method
+)) +
+  geom_point() +
+  geom_line() +
+  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  theme_light() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  scale_color_discrete_qualitative() +
+  ylab("AIC") +
+  xlab("Total number of bins")
+## Effective number of bins along x-axis
+
+mse_plot <- ggplot(MSE_df, aes(x = number_of_bins,
+                                       y = total_mse,
+                                       color = method
+)) +
+  geom_point() +
+  geom_line() +
+  theme_light() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  scale_color_discrete_qualitative() +
+  ylab("MSE") +
+  xlab("Total number of bins")
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: The .
+#| label: fig-diagnosticplt
+#| out-width: 100%
+
+aic_plot + mse_plot +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='bottom', plot.tag = element_text(size = 8))
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## UMAP with n_neighbor: 7
+## Prediction
+
+UMAP_data <- read_rds("data/s_curve/s_curve_umap_7.rds")
+predict_UMAP_df <- read_rds("data/s_curve/s_curve_umap_7_predict.rds")
+
+shape_value_curve <- calculate_effective_shape_value(.data = UMAP_data,
+                                                     x = UMAP1, y = UMAP2)
+
+num_bins_vec <- 1:10 ## Number of bins along the x-axis
+
+vec <- stats::setNames(rep("", 6), c("number_of_bins", "number_of_observations", "total_error", "totol_error_method_2", "totol_error_method_3", "total_mse"))  ## Define column names
+
+eval_data_test <- dplyr::bind_rows(vec)[0, ]
+eval_data_test <- eval_data_test |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(num_bins_vec)) {
+
+  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = UMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
+  pred_df_training <- pred_df_training_object$pred_data
+  centroid_df_training <- pred_df_training_object$df_bin_centroids
+  avg_df_training <- pred_df_training_object$df_bin
+
+  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+
+  pred_df_test_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = predict_UMAP_df, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
   pred_df_test <- pred_df_test_object$pred_data
@@
 
-eval_df_test <- generate_eval_df(data = data, prediction_df = pred_df_test, df_bin_centroids = centroid_df_test, df_bin = avg_df_test, num_bins = num_bins_x)
-```
-
-```{r}
-#| echo: false
-
-## Data set with all possible centroids in the hexagonal grid
-full_centroid_df <- generate_full_grid_centroids(df_bin_centroids)
-
-## Generate all coordinates of hexagons
-hex_grid <- full_hex_grid(full_centroid_df)
-
-#hex_full_count_df <- generate_full_grid_info(df_bin_centroids)
-
-predict_hex_ID_plot <- ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
-  geom_point(data = df_bin_centroids, aes(x = x, y = y), color = "#33a02c") +
-  geom_point(data = predict_UMAP_df, aes(x = UMAP1, y = UMAP2), color = "red") +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-    theme_void() +
-    theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-          axis.title.x = element_blank(), axis.title.y = element_blank(),
-          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-          panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-          legend.title = element_text(size=8), #change legend title font size
-          legend.text = element_text(size=6)) +
-    guides(fill = guide_colourbar(title = "Standardized count")) +
-    annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3) 
-```
-
-<!--To draw predict values in hexbin grid-->
-
-```{r}
-#| echo: false
-#| fig-cap: The .
-#| label: fig-predict
-
-predict_umap_plot  + predict_hex_ID_plot +
-  plot_layout(guides='collect', ncol = 3) &
-  theme(legend.position='bottom', plot.tag = element_text(size = 8))
-```
-
-```{r}
-#| echo: false
-#num_bins_vec <- append(sample(1:sqrt(750), 10), num_bins_x)
-
-#num_bins_vec <- sample(2:1000, 40)
-num_bins_vec <- 1:10
+  eval_df_test <- generate_eval_df(data = data, prediction_df = pred_df_test, df_bin_centroids = centroid_df_test, df_bin = avg_df_test, num_bins = num_bins_vec[i], col_start = "x")
+
+  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+  eval_data_test <- dplyr::bind_rows(eval_data_test, eval_df_test)
+
+}
+
+## Add new column with data types
+
+eval_data_training <- eval_data_training |>
+  mutate(data_type = "training")
+
+eval_data_test <- eval_data_test |>
+  mutate(data_type = "test")
+
+MSE_df_umap_1 <- bind_rows(eval_data_training, eval_data_test) |>
+  dplyr::mutate(param = "n_neighbors: 7")
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## UMAP with n_neighbor: 15
+## Prediction
+
+UMAP_data <- read_rds("data/s_curve/s_curve_umap_15.rds")
+predict_UMAP_df <- read_rds("data/s_curve/s_curve_umap_15_predict.rds")
+
+shape_value_curve <- calculate_effective_shape_value(.data = UMAP_data,
+                                                     x = UMAP1, y = UMAP2)
+
+num_bins_vec <- 1:10 ## Number of bins along the x-axis
 
@@
 
-  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = UMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value)
+  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = UMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
   pred_df_training <- pred_df_training_object$pred_data
@@
 
-  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i])
-
-  pred_df_test_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = predict_UMAP_df, num_bins = num_bins_vec[i], shape_val = shape_value)
+  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+
+  pred_df_test_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = predict_UMAP_df, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
   pred_df_test <- pred_df_test_object$pred_data
@@
 
-  eval_df_test <- generate_eval_df(data = data, prediction_df = pred_df_test, df_bin_centroids = centroid_df_test, df_bin = avg_df_test, num_bins = num_bins_vec[i])
+  eval_df_test <- generate_eval_df(data = data, prediction_df = pred_df_test, df_bin_centroids = centroid_df_test, df_bin = avg_df_test, num_bins = num_bins_vec[i], col_start = "x")
 
@@
 }
-```
-
-```{r}
-#| echo: false
+
+## Add new column with data types
 
@@
 
-MSE_df <- dplyr::bind_rows(eval_data_training, eval_data_test)
-```
-
-```{r}
-#| echo: false
-aic_plot <- ggplot(MSE_df |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
+MSE_df_umap_2 <- bind_rows(eval_data_training, eval_data_test) |>
+  dplyr::mutate(param = "n_neighbors: 15")
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## UMAP with n_neighbor: 32
+## Prediction
+
+UMAP_data <- read_rds("data/s_curve/s_curve_umap_32.rds")
+predict_UMAP_df <- read_rds("data/s_curve/s_curve_umap_32_predict.rds")
+
+shape_value_curve <- calculate_effective_shape_value(.data = UMAP_data,
+                                                     x = UMAP1, y = UMAP2)
+
+num_bins_vec <- 1:10 ## Number of bins along the x-axis
+
+vec <- stats::setNames(rep("", 6), c("number_of_bins", "number_of_observations", "total_error", "totol_error_method_2", "totol_error_method_3", "total_mse"))  ## Define column names
+
+eval_data_test <- dplyr::bind_rows(vec)[0, ]
+eval_data_test <- eval_data_test |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(num_bins_vec)) {
+
+  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = UMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
+  pred_df_training <- pred_df_training_object$pred_data
+  centroid_df_training <- pred_df_training_object$df_bin_centroids
+  avg_df_training <- pred_df_training_object$df_bin
+
+  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+
+  pred_df_test_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = predict_UMAP_df, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
+  pred_df_test <- pred_df_test_object$pred_data
+  centroid_df_test <- pred_df_test_object$df_bin_centroids
+  avg_df_test <- pred_df_test_object$df_bin
+
+  eval_df_test <- generate_eval_df(data = data, prediction_df = pred_df_test, df_bin_centroids = centroid_df_test, df_bin = avg_df_test, num_bins = num_bins_vec[i], col_start = "x")
+
+  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+  eval_data_test <- dplyr::bind_rows(eval_data_test, eval_df_test)
+
+}
+
+## Add new column with data types
+
+eval_data_training <- eval_data_training |>
+  mutate(data_type = "training")
+
+eval_data_test <- eval_data_test |>
+  mutate(data_type = "test")
+
+MSE_df_umap_3 <- bind_rows(eval_data_training, eval_data_test) |>
+  dplyr::mutate(param = "n_neighbors: 32")
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## UMAP with n_neighbor: 50
+## Prediction
+
+UMAP_data <- read_rds("data/s_curve/s_curve_umap.rds")
+predict_UMAP_df <- read_rds("data/s_curve/s_curve_umap_predict.rds")
+
+shape_value_curve <- calculate_effective_shape_value(.data = UMAP_data,
+                                                     x = UMAP1, y = UMAP2)
+
+num_bins_vec <- 1:10 ## Number of bins along the x-axis
+
+vec <- stats::setNames(rep("", 6), c("number_of_bins", "number_of_observations", "total_error", "totol_error_method_2", "totol_error_method_3", "total_mse"))  ## Define column names
+
+eval_data_test <- dplyr::bind_rows(vec)[0, ]
+eval_data_test <- eval_data_test |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(num_bins_vec)) {
+
+  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = UMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
+  pred_df_training <- pred_df_training_object$pred_data
+  centroid_df_training <- pred_df_training_object$df_bin_centroids
+  avg_df_training <- pred_df_training_object$df_bin
+
+  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+
+  pred_df_test_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = predict_UMAP_df, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
+  pred_df_test <- pred_df_test_object$pred_data
+  centroid_df_test <- pred_df_test_object$df_bin_centroids
+  avg_df_test <- pred_df_test_object$df_bin
+
+  eval_df_test <- generate_eval_df(data = data, prediction_df = pred_df_test, df_bin_centroids = centroid_df_test, df_bin = avg_df_test, num_bins = num_bins_vec[i], col_start = "x")
+
+  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+  eval_data_test <- dplyr::bind_rows(eval_data_test, eval_df_test)
+
+}
+
+## Add new column with data types
+
+eval_data_training <- eval_data_training |>
+  mutate(data_type = "training")
+
+eval_data_test <- eval_data_test |>
+  mutate(data_type = "test")
+
+MSE_df_umap_4 <- bind_rows(eval_data_training, eval_data_test) |>
+  dplyr::mutate(param = "n_neighbors: 50")
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+MSE_df_umap <- dplyr::bind_rows(MSE_df_umap_1, MSE_df_umap_2, MSE_df_umap_3, MSE_df_umap_4)
+
+## To draw with AIC
+aic_plot_param <- ggplot(MSE_df_umap |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
                                                                          y = total_error,
-                   color = data_type
+                                                                         color = param
 )) +
@@
   geom_line() +
-  #geom_smooth(method = "loess", se = FALSE) +
-  geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-15000, label=paste0("effective number of bins = ",as.character(NROW(full_grid_with_hexbin_id))), angle=90, size = 2.5) + 
-  theme(legend.position = "none", legend.title=element_blank(), plot.title=element_text(size=8), axis.text = element_text(size = 5),
+  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  theme_light() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
         axis.title = element_text(size = 7),
-        legend.text = element_text(size=7)) +
-  scale_color_manual(values = c("#d95f02")) + 
+        axis.text = element_text(size = 7)) +
+  scale_color_discrete_qualitative() +
   ylab("AIC") +
   xlab("Total number of bins")
-```
-
-```{r}
-#| echo: false
-mse_plot <- ggplot(MSE_df, aes(x = number_of_bins,
+## Effective number of bins along x-axis
+
+mse_plot_param_training <- ggplot(MSE_df_umap |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
                                y = total_mse,
-                   color = data_type
+                               color = param
 )) +
@@
   geom_line() +
-  #geom_smooth(method = "loess", se = FALSE) +
-  geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.05, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90, size = 2.5) +
-  scale_color_manual(values = c("#1b9e77", "#d95f02")) + 
-  theme(legend.position = "bottom", legend.title=element_blank(), plot.title=element_text(size=8), axis.text = element_text(size = 5),
+  theme_light() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
         axis.title = element_text(size = 7),
-        legend.text = element_text(size=7)) + 
+        axis.text = element_text(size = 7)) +
+  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  scale_color_discrete_qualitative() +
   ylab("MSE") +
   xlab("Total number of bins")
+
+mse_plot_param_test <- ggplot(MSE_df_umap |> dplyr::filter(data_type == "test"), aes(x = number_of_bins,
+                                                                                             y = total_mse,
+                                                                                             color = param
+)) +
+  geom_point() +
+  geom_line() +
+  theme_light() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  scale_color_discrete_qualitative() +
+  ylab("MSE") +
+  xlab("Total number of bins")
+
 ```
@@
 #| fig-cap: The .
-#| label: fig-diagnosticplt
+#| label: fig-diagnosticpltDiffParam
 #| out-width: 100%
 
-aic_plot + mse_plot +
+aic_plot_param + mse_plot_param_training + mse_plot_param_test  +
   plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect') &
+  plot_layout(guides='collect', ncol = 3) &
   theme(legend.position='bottom', plot.tag = element_text(size = 8))
@@
 
-### Simulated data examples {#sec-simpleex}
+### Simulated data example {#sec-simpleex}
+
+In this section, we showcase the effectiveness of our methodology using simulated data.
+The dataset comprises five spherical Gaussian clusters in 4-$d$, with each cluster containing an equal number of points and consistent within variation.
+
+We *strongly* recommend viewing the linked videos for each study while reading.
+Links to the videos are available in the figures for each example.
+The videos show the visual appearance of the **langevitour** interface with low-dimensional view and how we can interact with the tour via the controls.
+
+```{r}
+#| warning: false
+#| echo: false
+
+## Import data
+df_2 <- read_rds("data/five_gau_clusters/data_five_gau.rds")
+training_data_1 <- read_rds("data/five_gau_clusters/data_five_gau_training.rds")
+test_1 <- read_rds("data/five_gau_clusters/data_five_gau_test.rds")
+
+tSNE_data_gau <- read_rds("data/five_gau_clusters/tsne_data_five_gau.rds")
+UMAP_data_gau <- read_rds("data/five_gau_clusters/umap_data_five_gau.rds")
+PHATE_data_gau <- read_rds("data/five_gau_clusters/phate_data_five_gau.rds")
+TriMAP_data_gau <- read_rds("data/five_gau_clusters/trimap_data_five_gau.rds")
+PaCMAP_data_gau <- read_rds("data/five_gau_clusters/pacmap_data_five_gau.rds")
+
+## Visualise embeddings
+
+plot_list1_gau <- plot_tSNE_2D(tSNE_data_gau) + #ggtitle("(a)") + 
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+plot_list2_gau <- plot_UMAP_2D(UMAP_data_gau) + #ggtitle("(b)") + 
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+plot_list3_gau <- plot_PHATE_2D(PHATE_data_gau) + #ggtitle("(c)") + 
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+plot_list4_gau <- plot_TriMAP_2D(TriMAP_data_gau) + #ggtitle("(d)") + 
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+plot_list5_gau <- plot_PaCMAP_2D(PaCMAP_data_gau) + #ggtitle("(e)") + 
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'e', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "2D layouts from different NLDR techniques applied the same data: (a) tSNE (perplexity = 61), (b) UMAP (n_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2).
+Is there a best representation of the original data or are they all providing  equivalent information?"
+#| label: fig-nldervis5Gau
+#| out-width: 100%
+
+plot_list1_gau + plot_list2_gau + plot_list3_gau + plot_list4_gau + plot_list5_gau +
+  plot_layout(ncol=5)
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## tSNE
+
+shape_value_gau <- calculate_effective_shape_value(.data = tSNE_data_gau,
+                                                   x = tSNE1, y = tSNE2)
+
+num_bins_vec <- 1:10 ## Number of bins along the x-axis
+
+vec <- stats::setNames(rep("", 6), c("number_of_bins", "number_of_observations", "total_error", "totol_error_method_2", "totol_error_method_3", "total_mse"))  ## Define column names
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(num_bins_vec)) {
+
+  pred_df_training_object <- predict_hex_id(training_data = training_data_1, nldr_df = tSNE_data_gau, nldr_df_test = tSNE_data_gau, num_bins = num_bins_vec[i], shape_val = shape_value_gau, x = "tSNE1", y = "tSNE2", col_start = "x")
+  pred_df_training <- pred_df_training_object$pred_data
+  centroid_df_training <- pred_df_training_object$df_bin_centroids
+  avg_df_training <- pred_df_training_object$df_bin
+
+  eval_df_training <- generate_eval_df(data = df_2, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+
+  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+
+}
+
+## Add new column with data types
+
+eval_data_training <- eval_data_training |>
+  mutate(data_type = "training")
+
+MSE_df_1_gau <- eval_data_training |>
+  dplyr::mutate(method = "tSNE")
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## UMAP
+## Prediction
+
+shape_value_gau <- calculate_effective_shape_value(.data = UMAP_data_gau,
+                                                   x = UMAP1, y = UMAP2)
+
+num_bins_vec <- 1:10 ## Number of bins along the x-axis
+
+vec <- stats::setNames(rep("", 6), c("number_of_bins", "number_of_observations", "total_error", "totol_error_method_2", "totol_error_method_3", "total_mse"))  ## Define column names
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(num_bins_vec)) {
+
+  pred_df_training_object <- predict_hex_id(training_data = training_data_1, nldr_df = UMAP_data_gau, nldr_df_test = UMAP_data_gau, num_bins = num_bins_vec[i], shape_val = shape_value_gau, x = "UMAP1", y = "UMAP2", col_start = "x")
+  pred_df_training <- pred_df_training_object$pred_data
+  centroid_df_training <- pred_df_training_object$df_bin_centroids
+  avg_df_training <- pred_df_training_object$df_bin
+
+  eval_df_training <- generate_eval_df(data = df_2, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+
+  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+
+}
+
+## Add new column with data types
+
+eval_data_training <- eval_data_training |>
+  mutate(data_type = "training")
+
+MSE_df_2_gau <- eval_data_training |>
+  dplyr::mutate(method = "UMAP")
+```
+
+```{r}
+#| warning: false
+#| echo: false
+## PAHTE
+## Prediction
+
+shape_value_gau <- calculate_effective_shape_value(.data = PHATE_data_gau,
+                                                   x = PHATE1, y = PHATE2)
+
+num_bins_vec <- 1:10 ## Number of bins along the x-axis
+
+vec <- stats::setNames(rep("", 6), c("number_of_bins", "number_of_observations", "total_error", "totol_error_method_2", "totol_error_method_3", "total_mse"))  ## Define column names
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(num_bins_vec)) {
+
+  pred_df_training_object <- predict_hex_id(training_data = training_data_1, nldr_df = PHATE_data_gau, nldr_df_test = PHATE_data_gau, num_bins = num_bins_vec[i], shape_val = shape_value_gau, x = "PHATE1", y = "PHATE2", col_start = "x")
+  pred_df_training <- pred_df_training_object$pred_data
+  centroid_df_training <- pred_df_training_object$df_bin_centroids
+  avg_df_training <- pred_df_training_object$df_bin
+
+  eval_df_training <- generate_eval_df(data = df_2, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+
+  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+
+}
+
+## Add new column with data types
+
+eval_data_training <- eval_data_training |>
+  mutate(data_type = "training")
+
+MSE_df_3_gau <- eval_data_training |>
+  dplyr::mutate(method = "PHATE")
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## TriMAP
+
+## Prediction
+
+shape_value_gau <- calculate_effective_shape_value(.data = TriMAP_data_gau,
+                                                   x = TriMAP1, y = TriMAP2)
+
+num_bins_vec <- 1:10 ## Number of bins along the x-axis
+
+vec <- stats::setNames(rep("", 6), c("number_of_bins", "number_of_observations", "total_error", "totol_error_method_2", "totol_error_method_3", "total_mse"))  ## Define column names
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(num_bins_vec)) {
+
+  pred_df_training_object <- predict_hex_id(training_data = training_data_1, nldr_df = TriMAP_data_gau, nldr_df_test = TriMAP_data_gau, num_bins = num_bins_vec[i], shape_val = shape_value_gau, x = "TriMAP1", y = "TriMAP2", col_start = "x")
+  pred_df_training <- pred_df_training_object$pred_data
+  centroid_df_training <- pred_df_training_object$df_bin_centroids
+  avg_df_training <- pred_df_training_object$df_bin
+
+  eval_df_training <- generate_eval_df(data = df_2, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+
+  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+
+}
+
+## Add new column with data types
+
+eval_data_training <- eval_data_training |>
+  mutate(data_type = "training")
+
+MSE_df_4_gau <- eval_data_training |>
+  dplyr::mutate(method = "TriMAP")
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## PaCMAP
+
+## Prediction
+
+shape_value_gau <- calculate_effective_shape_value(.data = PaCMAP_data_gau,
+                                                   x = PaCMAP1, y = PaCMAP2)
+
+num_bins_vec <- 1:10 ## Number of bins along the x-axis
+
+vec <- stats::setNames(rep("", 6), c("number_of_bins", "number_of_observations", "total_error", "totol_error_method_2", "totol_error_method_3", "total_mse"))  ## Define column names
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(num_bins_vec)) {
+
+  pred_df_training_object <- predict_hex_id(training_data = training_data_1, nldr_df = PaCMAP_data_gau, nldr_df_test = PaCMAP_data_gau, num_bins = num_bins_vec[i], shape_val = shape_value_gau, x = "PaCMAP1", y = "PaCMAP2", col_start = "x")
+  pred_df_training <- pred_df_training_object$pred_data
+  centroid_df_training <- pred_df_training_object$df_bin_centroids
+  avg_df_training <- pred_df_training_object$df_bin
+
+  eval_df_training <- generate_eval_df(data = df_2, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+
+  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+
+}
+
+## Add new column with data types
+
+eval_data_training <- eval_data_training |>
+  mutate(data_type = "training")
+
+MSE_df_5_gau <- eval_data_training |>
+  dplyr::mutate(method = "PaCMAP")
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+MSE_df_gau <- dplyr::bind_rows(MSE_df_1_gau, MSE_df_2_gau, MSE_df_3_gau, MSE_df_4_gau, MSE_df_5_gau)
+
+## To draw with AIC
+aic_gau_plot <- ggplot(MSE_df_gau |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
+                                                             y = total_error,
+                                                             color = method
+)) +
+  geom_point() +
+  geom_line() +
+  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  theme_light() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  scale_color_discrete_qualitative() +
+  ylab("AIC") +
+  xlab("Total number of bins")
+## Effective number of bins along x-axis
+
+mse_gau_plot <- ggplot(MSE_df_gau, aes(x = number_of_bins,
+                   y = total_mse,
+                   color = method
+)) +
+  geom_point() +
+  geom_line() +
+  theme_light() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  scale_color_discrete_qualitative() +
+  ylab("MSE") +
+  xlab("Total number of bins")
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+aic_gau_plot + mse_gau_plot +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='bottom', plot.tag = element_text(size = 8))
+```
 

### 2024-01-10 paper.tex
Update the paper with sumaary for the example, s-curve data set, for s-curve dataset added summary with different parameter setting in UMAP, generate required data sets beforehand and store in necessary folders, generate scripts to generate embeddings

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, interior hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}, frame hidden, boxrule=0pt, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, breakable, sharp corners, borderline west={3pt}{0pt}{shadecolor}, enhanced, boxrule=0pt, interior hidden]}{\end{tcolorbox}}\fi
 
@@
 
-\hypertarget{introduction-to-umap}{%
-\subsubsection{Introduction to UMAP}\label{introduction-to-umap}}
+\hypertarget{non-linear-dimension-reduction-techniques}{%
+\subsubsection{Non-linear dimension reduction
+techniques}\label{non-linear-dimension-reduction-techniques}}
 
-The UMAP algorithm constructs a fuzzy topological representation of the
-high-dimensional data and then optimizes the low-dimensional embedding
-to be as similar to this representation as possible \citep{Leland2018}.
+Non-linear dimension reduction techniques play a crucial role in the
+analysis and visualization of high-dimensional data, where the
+complexities of relationships among variables may not be adequately
+captured by linear methods. Among these techniques, tSNE stands out for
+its emphasis on preserving pairwise distances. By minimizing the
+divergence between probability distributions in both the high and
+low-dimensional spaces, t-SNE effectively reveals intricate structures
+and patterns within the data. Its application is widespread in tasks
+requiring the visualization of clusters and local relationships, though
+it does require careful consideration of the perplexity parameter for
+optimal results.
 
-The UMAP algorithm constructs a weighted k-nearest neighbors graph based
-on the local similarity of data points in the original high-dimensional
-space. The weighted graph is defined by a conditional probability
-distribution \(P_{ij}\), which represents the probability of observing
-data point \(x_i\) given data point \(x_j\) as a neighbor:
+UMAP is another powerful non-linear technique that strikes a balance
+between preserving local and global structures. Constructing a fuzzy
+topological representation using a weighted k-nearest neighbors graph,
+UMAP optimizes the low-dimensional embedding to resemble this
+representation. Known for its efficiency and scalability, UMAP is
+versatile across various scales of relationships in the data, although
+parameter sensitivity, particularly concerning the choice of neighbors,
+must be taken into account.
 
-\[
-P_{ij} = \frac{\exp(-\lVert x_i - x_j \rVert^2)}{\sum_{k \neq i} \exp(-\lVert x_i - x_k \rVert^2)}
-\]
+For trajectory data, PHATE provides specialized capabilities. It models
+the affinity between data points, simulating a heat diffusion process to
+capture developmental processes, particularly in single-cell genomics.
+While PHATE excels in revealing trajectory structures and offering
+insights into cellular development, it necessitates careful parameter
+tuning due to its specialized nature.
 
-Similarly, a fuzzy topological representation of the data is constructed
-in the low-dimensional space with conditional probability distribution
-\(Q_{ij}\):
+TriMAP adopts a unique approach by approximating the data manifold
+through the construction of a triangulated graph representation. This
+technique efficiently captures both global and local structures by
+representing the data as a network of triangles. TriMAP's strength lies
+in its ability to efficiently capture complex structures, albeit with
+sensitivity to parameter choices, including the number of neighbors.
 
-\[
-Q_{ij} = \frac{\exp(-\lVert y_i - y_j \rVert^2)}{\sum_{k \neq i} \exp(-\lVert y_i - y_k \rVert^2)}
-\]
-
-where \(y_i\) and \(y_j\) are the corresponding low-dimensional
-representations of data points \(x_i\) and \(x_j\).
-
-UMAP minimizes the cross-entropy between the high-dimensional fuzzy
-topological representation \(P_{ij}\) and the low-dimensional
-representation \(Q_{ij}\) to find an optimal low-dimensional embedding.
-The loss function for UMAP is defined as:
-
-\[
-L = \sum_{i} \sum_{j} P_{ij} \log \frac{P_{ij}}{Q_{ij}}
-\]
-
-The algorithm uses stochastic gradient descent to optimize the positions
-of the data points in the low-dimensional space to minimize this loss
-function.
-
-In R, the \texttt{umap} package is often used to implement UMAP and
-generate results.
-
-The ``n\_neighbors'' parameter in UMAP determines the number of nearest
-neighbors to consider for each data point during the construction of the
-fuzzy topological representation. Higher values of ``n\_neighbors''
-capture more global structure, while lower values emphasize local
-structure. Choosing an appropriate value for ``n\_neighbors'' is
-crucial, as it affects the balance between local and global structure in
-the resulting low-dimensional representation.
+In contrast, PaCMAP introduces supervised learning to create a
+low-dimensional representation while preserving pair-wise relationships.
+Constructing a graph based on pair-wise distances, PaCMAP optimizes an
+embedding using a customizable loss function. Particularly notable is
+PaCMAP's flexibility in incorporating class labels or additional
+information to guide the embedding process, offering users a means to
+customize its behavior and performance.
 
@@
 \caption{\label{fig-nldervis}2D layouts from different NLDR techniques
-applied the same data: (a) tSNE (perplexity = 32), (b) UMAP
+applied the same data: (a) tSNE (perplexity = 27), (b) UMAP
 (n\_neighbors = 50), (c) PHATE (knn = 5), (d) TriMAP (n\_inliers = 5,
@@
 for the same \textbf{s\_curve\_noise\_umap} data: (a) b = (4, 8), s =
-2.031, (b) b = (9, 20), s = 2.031, and (c) b = (16, 36), s = 2.031. The
-hexbins are colored based on the density of points, with yellow
-indicating higher point density and darker colors representing lower
-point density within each bin. Does a value of number of bins exist to
-effectively represent the low-dimensional data?}
+1.643542, (b) b = (12, 22), s = 1.643542, and (c) b = (21, 40), s =
+1.643542. The hexbins are colored based on the density of points, with
+yellow indicating higher point density and darker colors representing
+lower point density within each bin. Does a value of number of bins
+exist to effectively represent the low-dimensional data?}
 
@@
 \subsection{Tunning the model}\label{tunning-the-model}}
+
+In our model tuning process, we strategically adjust three key
+parameters to optimize the performance and accuracy of our approach.
+They are (i) the total number of bins (\(b\)), (ii) a benchmark value to
+remove low-density hexagons, and (iii) a benchmark value to remove long
+edges.
+
+\hypertarget{total-number-of-bins}{%
+\subsection{Total number of bins}\label{total-number-of-bins}}
+
+The total number of bins is a crucial parameter that defines the
+granularity of our hexagonal grid. This parameter particularly
+influential in capturing the intricacies of the data structure. It is
+determined by multiplying the number of bins along the x-axis ((b\_1))
+with the number of bins along the y-axis ((b\_2)), as per the formula:
+
+\begin{equation}\protect\hypertarget{eq-equationp1}{}{
+b = b_1 \times b_2
+}\label{eq-equationp1}\end{equation}
+
+Here, (b) represents the total number of bins. Adjusting the parameter
+(b\_1) provides control over the total bin count ((b)), allowing us to
+fine-tune the grid's configuration along the x-axis based on specific
+requirements.
+
+\hypertarget{benchmark-value-to-remove-low-density-hexagons}{%
+\subsection{Benchmark value to remove low-density
+hexagons}\label{benchmark-value-to-remove-low-density-hexagons}}
+
+Addressing low-density hexagons is a systematic process to handle
+sparsely represented data in certain regions. For each hex bin, we
+identify the six nearest hex bins using an equal 2D distance metric.
+Then, we calculate the mean density, as outlined in the equations:
+
+\begin{equation}\protect\hypertarget{eq-equationp2}{}{
+\text{standard count} = \frac{\text{count}}{\text{max count}} 
+}\label{eq-equationp2}\end{equation}
+
+\begin{equation}\protect\hypertarget{eq-equationp3}{}{
+\text{mean density} = \frac{\text{standard count}}{6} 
+}\label{eq-equationp3}\end{equation}
+
+The standard count is derived from the number of observations in the hex
+bins. By examining the distribution of mean densities and designating
+the first quartile as the benchmark value, hex bins with mean densities
+below this benchmark are removed. This process ensures the elimination
+of regions with insufficient data density, focusing on areas with more
+significant data representation and preserving the overall structure in
+the low-dimensional space.
+
+\hypertarget{benchmark-value-to-remove-long-edges}{%
+\subsection{Benchmark value to remove long
+edges}\label{benchmark-value-to-remove-long-edges}}
+
+The removal of long edges is a critical step to create a smoother
+representation by iteratively eliminating hexagons with excessive
+distances between centroids. This process eliminates outliers and noise
+while preserving essential local relationships within the data. To
+achieve this, distances between vertices are sorted, and unique distance
+values are extracted. By setting a threshold based on the largest
+difference between consecutive distance values, long edges are
+identified and removed. This refinement step contributes to enhancing
+the quality of the triangular mesh, ensuring a more accurate
+representation of the data structure.
 
@@
 
-{\centering \includegraphics{paper_files/figure-pdf/fig-predict-1.pdf}
-
-}
-
-\caption{\label{fig-predict}The .}
-
-\end{figure}
-
-\begin{figure}
-
 {\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticplt-1.pdf}
@@
 
+\begin{figure}
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticpltDiffParam-1.pdf}
+
+}
+
+\caption{\label{fig-diagnosticpltDiffParam}The .}
+
+\end{figure}
+
 \hypertarget{sec-simpleex}{%
-\subsection{Simulated data examples}\label{sec-simpleex}}
+\subsection{Simulated data example}\label{sec-simpleex}}
+
+In this section, we showcase the effectiveness of our methodology using
+simulated data. The dataset comprises five spherical Gaussian clusters
+in 4-\(d\), with each cluster containing an equal number of points and
+consistent within variation.
+
+We \emph{strongly} recommend viewing the linked videos for each study
+while reading. Links to the videos are available in the figures for each
+example. The videos show the visual appearance of the
+\textbf{langevitour} interface with low-dimensional view and how we can
+interact with the tour via the controls.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervis5Gau-1.pdf}
+
+}
+
+\caption{\label{fig-nldervis5Gau}2D layouts from different NLDR
+techniques applied the same data: (a) tSNE (perplexity = 61), (b) UMAP
+(n\_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n\_inliers = 5,
+n\_outliers = 4, n\_random = 3), and (e) PaCMAP (n\_neighbors = 10, init
+= random, MN\_ratio = 0.9, FP\_ratio = 2). Is there a best
+representation of the original data or are they all providing equivalent
+information?}
+
+\end{figure}
+
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-43-1.pdf}
 

### 2024-01-10 applications/pbmc/gen_pbmc_summary_df.R

@@ -1,0 +1,202 @@
+library(readr)
+library(dplyr)
+
+set.seed(20240110)
+
+source("quollr_code.R", local = TRUE)
+
+data <- read_csv("pbmc/pbmc.rds")
+training_data <- read_csv("pbmc/pbmc_training.rds")
+test_data <- read_csv("pbmc/pbmc_test.rds")
+
+tSNE_data <- read_csv("pbmc/pbmc_tsne_30.rds")
+UMAP_data <- read_csv("pbmc/pbmc_umap.rds")
+PHATE_data <- read_csv("pbmc/pbmc_phate.rds")
+TriMAP_data <- read_csv("pbmc/pbmc_trimap.rds")
+PaCMAP_data <- read_csv("pbmc/pbmc_pacmap.rds")
+
+
+## tSNE
+
+shape_value_curve <- calculate_effective_shape_value(.data = tSNE_data,
+                                                     x = tSNE1, y = tSNE2)
+
+num_bins_vec <- 1:10 ## Number of bins along the x-axis
+
+vec <- stats::setNames(rep("", 6), c("number_of_bins", "number_of_observations", "total_error", "totol_error_method_2", "totol_error_method_3", "total_mse"))  ## Define column names
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(num_bins_vec)) {
+
+  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = tSNE_data, nldr_df_test = tSNE_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "tSNE1", y = "tSNE2", col_start = "x")
+  pred_df_training <- pred_df_training_object$pred_data
+  centroid_df_training <- pred_df_training_object$df_bin_centroids
+  avg_df_training <- pred_df_training_object$df_bin
+
+  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+
+  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+
+
+}
+
+
+## Add new column with data types
+
+eval_data_training <- eval_data_training |>
+  mutate(data_type = "training")
+
+MSE_df_1 <- eval_data_training |>
+  dplyr::mutate(method = "tSNE")
+
+
[…]

### 2024-01-10 paper.qmd
Add summary for PBMC and Mouse data sets

@@
 
+```{r}
+#| warning: false
+#| echo: false
+
+MSE_df_pbmc <- read_rds("data/pbmc/summary_pbmc.rds")
+
+## To draw with AIC
+aic_plot_pbmc <- ggplot(MSE_df_pbmc |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
+                                                                                 y = total_error,
+                                                                                 color = method
+)) +
+  geom_point() +
+  geom_line() +
+  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  theme_light() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  scale_color_discrete_qualitative() +
+  ylab("AIC") +
+  xlab("Total number of bins")
+## Effective number of bins along x-axis
+
+mse_plot_pbmc <- ggplot(MSE_df_pbmc, aes(x = number_of_bins,
+                                       y = total_mse,
+                                       color = method
+)) +
+  geom_point() +
+  geom_line() +
+  theme_light() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  scale_color_discrete_qualitative() +
+  ylab("MSE") +
+  xlab("Total number of bins")
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: The .
+#| label: fig-diagnosticpltPBMC
+#| out-width: 100%
+
+aic_plot_pbmc + mse_plot_pbmc +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='bottom', plot.tag = element_text(size = 8))
+```
+
 ### Single-Cell Tagged Reverse Transcription sequencing data of mouse
@@
 
+```{r}
+#| warning: false
+#| echo: false
+
+MSE_df_zei <- read_rds("data/zeisel/summary_zei.rds")
+
+## To draw with AIC
+aic_plot_zei <- ggplot(MSE_df_zei |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
+                                                                                 y = total_error,
+                                                                                 color = method
+)) +
+  geom_point() +
+  geom_line() +
+  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  theme_light() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  scale_color_discrete_qualitative() +
+  ylab("AIC") +
+  xlab("Total number of bins")
+## Effective number of bins along x-axis
+
+mse_plot_zei <- ggplot(MSE_df_zei, aes(x = number_of_bins,
+                                       y = total_mse,
+                                       color = method
+)) +
+  geom_point() +
+  geom_line() +
+  theme_light() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  scale_color_discrete_qualitative() +
+  ylab("MSE") +
+  xlab("Total number of bins")
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: The .
+#| label: fig-diagnosticpltZEI
+#| out-width: 100%
+
+aic_plot_zei + mse_plot_zei +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='bottom', plot.tag = element_text(size = 8))
+```
+
 ## Discussion {#sec-discussion}

### 2024-01-10 paper.tex
Add summary for PBMC and Mouse data sets

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, breakable, sharp corners, borderline west={3pt}{0pt}{shadecolor}, enhanced, boxrule=0pt, interior hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, enhanced, breakable, interior hidden, frame hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 
+\begin{figure}
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticpltPBMC-1.pdf}
+
+}
+
+\caption{\label{fig-diagnosticpltPBMC}The .}
+
+\end{figure}
+
 \hypertarget{single-cell-tagged-reverse-transcription-sequencing-data-of-mouse}{%
@@
 regions and the intricate networks that underpin neural processes.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticpltZEI-1.pdf}
+
+}
+
+\caption{\label{fig-diagnosticpltZEI}The .}
+
+\end{figure}
 

### 2024-01-10 paper.qmd
Added figure titles

--- 2024-01-10 paper.qmd
@@
       scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  xlim(-5, 7) +
+  ylim(-10, 10) +
         theme_void() +
@@
   scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+    xlim(-5, 8) +
+  ylim(-10, 10) +
         theme_void() +
@@
 #| label: fig-wkhighD
-#| fig-cap: Visualization.
-
-a1 + a2 +
+#| fig-cap: How the 2D model lift into high dimensions?
+(a) visualize the points and the hexagonal bin centroids related 31th and 239th hexagons, (b) visualization of the edge connected the 31th and 239th hexagons (colored in red) in the triangular mesh.
+A video of tour animation is available at (need to add the link of lg_high).
+
+a2 + a1 +
   plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', ncol = 3) &
+  plot_layout(ncol = 3) &
   theme(legend.position='bottom', plot.tag = element_text(size = 8))
@@
   theme_light() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
         axis.title = element_text(size = 7),
@@
 #| echo: false
-#| fig-cap: The .
-#| label: fig-diagnosticplt
+#| fig-cap: Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset. What is the best NLDR technique to represent the original data in 2D?
+#| label: fig-diagnosticpltScurve
 #| out-width: 100%
@@
 #| echo: false
-#| fig-cap: The .
+#| fig-cap: Goodness of fit statistics from different n_neighbors parameter of UMAP applied to training S-curve dataset. What is the best parameter choice in UMAP to represent the original data in 2D?
 #| label: fig-diagnosticpltDiffParam
@@
 ```
-
-<!--In this analysis, we're looking at how well our model is doing in two aspects.
-First, we're checking the distances in our lower-dimensional space (see @fig-diag (a)).
-We want to see if the distances between different types of edges are similar.
-If they are, it means our model is doing a good job of keeping the relationships consistent.
-If there are big differences, it could mean our model is struggling with certain types of edges.
-
-Next, we're examining how our model handles the relationship between high-dimensional and low-dimensional distances for each type of edge (see @fig-diag (b)).
-If our data points nicely line up around the expected line, it tells us that our model is preserving the original structure well.
-If they're scattered, it could mean our model is having trouble with specific data points.
-This two-part analysis helps us really understand how our model fits the data.
-
-```{r}
-#| echo: false
-
-sorted_distance_df <- distance[order(distance$distance),] ## Sort the distances
-
-sorted_distance_df <- cbind(sorted_distance_df, rbind(NA, apply(sorted_distance_df, 2, diff))) ## Calculate differences between unique distance
-sorted_distance_df <- sorted_distance_df[, c(3,6)]
-sorted_distance_df[is.na(sorted_distance_df)] <- 0.0000 ## To replace missing values with zero
-names(sorted_distance_df)[2] <- "difference"
-
-sorted_distance_df <- sorted_distance_df %>% 
-  mutate(Edge_type = ifelse(distance >= benchmark, "Long edge", "Short edge"))
-
-diagnostic_plot1 <- ggplot(sorted_distance_df, aes(x = distance, y = difference, color = Edge_type)) +
-  geom_point(alpha = 0.5) +
-  geom_rug(col="#e31a1c",alpha=0.1, linewidth=1.5) +
-  #ggtitle("Visualization of difference between \n 2D distances by edge type.") +
-  scale_color_manual(values=c("#33a02c", "#000000")) + 
-  theme(legend.position = "bottom", legend.title=element_blank(), plot.title=element_text(size=8), axis.text = element_text(size = 5),
-          axis.title = element_text(size = 7),
-        legend.text = element_text(size=7)) + 
-  xlab(expression(d^{(2)})) 
-```
-
-```{r}
-#| echo: false
-
-distance <- distance %>% 
-  rename(distance_2D = distance)
-
-#num_pca <- 4
-
-high_d <- df_bin[, 1:7]
-
-data_matrix <- data.matrix(high_d, rownames.force = NA)
-
-# calculate Euclidean distance between each row in matrix
-d <- dist(data_matrix)
-
-df <- reshape2::melt(as.matrix(d), varnames = c("Start_point", "End_point"))
-
-df <- df[df$Start_point > df$End_point,]
-
-rownames(df) <- 1:nrow(df)
-
-df <- df %>% 
-  rename(distance_high_D = value)
-
-high_low_dist <- left_join(distance, df, by = c("from" = "End_point", "to" = "Start_point"))
-high_low_dist <- high_low_dist %>% distinct()
-high_low_dist <- high_low_dist[complete.cases(high_low_dist),]
-
-lm_fit <- lm(distance_high_D ~ distance_2D, data = high_low_dist)
-
-high_low_dist <- high_low_dist %>% 
-  mutate(Edge_type = ifelse(distance_2D >= benchmark, "Long edge", "Short edge"))
-
-diagnostic_plot2 <- ggplot(high_low_dist, aes(x = distance_2D, y = distance_high_D, color = Edge_type)) +
-  geom_point(alpha = 0.5) +
-  geom_line(data = broom::augment(lm_fit),
-            aes(x = distance_2D, y = .fitted),
-            color="blue",
-            linewidth=1) +
-  geom_rug(col="#e31a1c",alpha=0.1, linewidth=1.5) +
-  #ggtitle("Visualization of 2D and high-D distances \n by edge type.") +
-  scale_color_manual(values=c("#33a02c","#000000")) + 
-  theme(legend.position = "bottom", legend.title=element_blank(), plot.title=element_text(size=8), axis.text = element_text(size = 5),
-          axis.title = element_text(size = 7),
-        legend.text = element_text(size=7)) + 
-  xlab(expression(d^{(2)})) + 
-  ylab(expression(d^{(p)}))
-```
-
-```{r}
-#| echo: false
-#| fig-cap: The two diagnostic plot assess the effectiveness of the high-D representation algorithm  performed for the **s_curve_noise_umap** data set$\colon$ (a) visualizing the differences between low-D distances and (b) visualizing the relationship between high-D ($d^{(p)}$) and low-D distances ($d^{(2)}$).
-#| label: fig-diag
-#| out-width: 100%
-
-diagnostic_plot1 + diagnostic_plot2 +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect') &
-  theme(legend.position='bottom', plot.tag = element_text(size = 8))
-```
-
-### Software
-
-Our algorithm for high-dimensional (high-D) representation is available as an open-source R package named quollr.
-This package provides analysts with the capability to construct a low-dimensional manifold and seamlessly visualize it in the high-D space alongside the original data.-->
 
@@
 ```{r}
-#| warning: false
-#| echo: false
+#| echo: false
+#| fig-cap: Goodness of fit statistics from different NLDR techniques applied to training five spherical Gaussian cluster dataset. What is the best NLDR technique to represent the original data in 2D?
+#| label: fig-diagnosticpltGau
+#| out-width: 100%
 
@@
 
+<!--
 ::: {#fig-pbmc_sc layout-ncol="3"}
@@
 Combination of the **langevitour** and low-dimensional representation of the pbmc data provides the effectiveness of our approach.
-:::
+:::-->
+
+```{r}
+#| warning: false
+#| echo: false
+
+MSE_df_pbmc <- read_rds("data/pbmc/summary_pbmc.rds")
+
+## To draw with AIC
+aic_plot_pbmc <- ggplot(MSE_df_pbmc |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
+                                                                                 y = total_error,
+                                                                                 color = method
+)) +
+  geom_point() +
+  geom_line() +
+  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  theme_light() +
+  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  scale_color_discrete_qualitative() +
+  ylab("AIC") +
+  xlab("Total number of bins")
+## Effective number of bins along x-axis
+
+mse_plot_pbmc <- ggplot(MSE_df_pbmc, aes(x = number_of_bins,
+                                       y = total_mse,
+                                       color = method
+)) +
+  geom_point() +
+  geom_line() +
+  theme_light() +
+  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  scale_color_discrete_qualitative() +
+  ylab("MSE") +
+  xlab("Total number of bins")
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: Goodness of fit statistics from different NLDR techniques applied to training PBMC dataset. What is the best NLDR technique to represent the original data in 2D?
+#| label: fig-diagnosticpltPBMC
+#| out-width: 100%
+
+aic_plot_pbmc + mse_plot_pbmc +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='bottom', plot.tag = element_text(size = 8))
+```
 
@@
 
+```{r}
+#| warning: false
+#| echo: false
+
+MSE_df_zei <- read_rds("data/zeisel/summary_zei.rds")
+
+## To draw with AIC
+aic_plot_zei <- ggplot(MSE_df_zei |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
+                                                                                 y = total_error,
+                                                                                 color = method
+)) +
+  geom_point() +
+  geom_line() +
+  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  theme_light() +
+  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  scale_color_discrete_qualitative() +
+  ylab("AIC") +
+  xlab("Total number of bins")
+## Effective number of bins along x-axis
+
+mse_plot_zei <- ggplot(MSE_df_zei, aes(x = number_of_bins,
+                                       y = total_mse,
+                                       color = method
+)) +
+  geom_point() +
+  geom_line() +
+  theme_light() +
+  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  scale_color_discrete_qualitative() +
+  ylab("MSE") +
+  xlab("Total number of bins")
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: Goodness of fit statistics from different NLDR techniques applied to training Zeisel mouse brain dataset. What is the best NLDR technique to represent the original data in 2D?
+#| label: fig-diagnosticpltZEI
+#| out-width: 100%
+
+aic_plot_zei + mse_plot_zei +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='bottom', plot.tag = element_text(size = 8))
+```
+
 ## Discussion {#sec-discussion}

### 2024-01-10 paper.tex
Added figure titles

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, enhanced, breakable, interior hidden, frame hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, interior hidden, boxrule=0pt, frame hidden, enhanced, breakable, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 
-\caption{\label{fig-wkhighD}Visualization.}
+\caption{\label{fig-wkhighD}How the 2D model lift into high dimensions?
+(a) visualize the points and the hexagonal bin centroids related 31th
+and 239th hexagons, (b) visualization of the edge connected the 31th and
+239th hexagons (colored in red) in the triangular mesh. A video of tour
+animation is available at (need to add the link of lg\_high).}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticplt-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticpltScurve-1.pdf}
 
@@
 
-\caption{\label{fig-diagnosticplt}The .}
+\caption{\label{fig-diagnosticpltScurve}Goodness of fit statistics from
+different NLDR techniques applied to training S-curve dataset. What is
+the best NLDR technique to represent the original data in 2D?}
 
@@
 
-\caption{\label{fig-diagnosticpltDiffParam}The .}
+\caption{\label{fig-diagnosticpltDiffParam}Goodness of fit statistics
+from different n\_neighbors parameter of UMAP applied to training
+S-curve dataset. What is the best parameter choice in UMAP to represent
+the original data in 2D?}
 
@@
 
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-43-1.pdf}
+\begin{figure}
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticpltGau-1.pdf}
+
+}
+
+\caption{\label{fig-diagnosticpltGau}Goodness of fit statistics from
+different NLDR techniques applied to training five spherical Gaussian
+cluster dataset. What is the best NLDR technique to represent the
+original data in 2D?}
+
+\end{figure}
 
@@
 application, enabling a comprehensive exploration of the data's
-structure and relationships (see Figure~\ref{fig-pbmc_sc}).
+structure and relationships (see \textbf{?@fig-pbmc\_sc}).
 
@@
 
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/pbmc_a.png}
-
-}
-
-}
-
-\subcaption{\label{fig-pbmc_sc1}}
-\end{minipage}%
-%
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/pbmc_b.png}
-
-}
-
-}
-
-\subcaption{\label{fig-pbmc_sc2}}
-\end{minipage}%
-%
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/pbmc_c.png}
-
-}
-
-}
-
-\subcaption{\label{fig-pbmc_sc3}}
-\end{minipage}%
-
-\caption{\label{fig-pbmc_sc}Screen shots of the \textbf{langevitour}
-with the low-dimensional view applied to pbmc data set, a video of the
-tour animation is available at
-(\url{https://drive.google.com/file/d/1F7Xb2EIpfUruCxYihPw2rKro4FB5Y2uL/view}).
-Combination of the \textbf{langevitour} and low-dimensional
-representation of the pbmc data provides the effectiveness of our
-approach.}
-
-\end{figure}
-
-\begin{figure}
-
 {\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticpltPBMC-1.pdf}
@@
 
-\caption{\label{fig-diagnosticpltPBMC}The .}
+\caption{\label{fig-diagnosticpltPBMC}Goodness of fit statistics from
+different NLDR techniques applied to training PBMC dataset. What is the
+best NLDR technique to represent the original data in 2D?}
 
@@
 
-\caption{\label{fig-diagnosticpltZEI}The .}
+\caption{\label{fig-diagnosticpltZEI}Goodness of fit statistics from
+different NLDR techniques applied to training Zeisel mouse brain
+dataset. What is the best NLDR technique to represent the original data
+in 2D?}
 

### 2024-01-10 paper.qmd
Add 2D layouts of PBMC and Mouse data

@@
   plot_layout(ncol = 3) &
-  theme(legend.position='bottom', plot.tag = element_text(size = 8))
+  theme(legend.position='none', plot.tag = element_text(size = 8))
 ```
@@
   theme_light() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+  theme(legend.title = element_blank(), legend.text = element_text(size=7), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
         axis.title = element_text(size = 7),
@@
   theme_light() +
-  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+  theme(legend.title = element_blank(), legend.text = element_text(size=7), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
         axis.title = element_text(size = 7),
@@
 #| echo: false
-# InstallData("pbmc3k")
-# data("pbmc3k.final")
-# 
-# data_pca <- as.data.frame(pbmc3k.final@reductions$pca@cell.embeddings)
-# names(data_pca) <- paste0(rep("PC", ncol(data_pca)), 1:ncol(data_pca))
-# 
-# data_pca_with_label <- bind_cols(pbmc3k.final@reductions$pca@cell.embeddings, pbmc3k.final@meta.data$seurat_clusters)
-# names(data_pca_with_label)[ncol(pbmc3k.final@reductions$pca@cell.embeddings) + 1] <- "cluster_label"
-# 
-# levels(data_pca_with_label$cluster_label) <- c("Naive CD4 T", "Memory CD4 T", "CD14+ Mono", "B", "CD8 T", "FCGR3A+ Mono", "NK", "DC", "Platelet")
-
-data_pca_with_label <- readRDS(file = paste0(here::here(), "/data/pbmc_with_label.rds"))
-
-data_pca <- data_pca_with_label %>%
-  select(-cluster_label)
-```
-
-```{r}
-#| warning: false
-#| echo: false
-# eigs <- pbmc3k.final@reductions$pca@stdev^2 ## To calculate eigen values
-# 
-# summary_pca <- rbind(
-#   SD = sqrt(eigs),
-#   Proportion = eigs/sum(eigs),
-#   Cumulative = cumsum(eigs)/sum(eigs))
-# 
-# var_explained_df <- data.frame(PC = paste0("PC", 1:ncol(pbmc3k.final@reductions$pca@cell.embeddings)),
-#                                var_explained = (pbmc3k.final@reductions$pca@stdev)^2/sum((pbmc3k.final@reductions$pca@stdev)^2))
-
-var_explained_df <- readRDS(file = paste0(here::here(), "/data/pbmc_var_explained.rds"))
-
-pbmc_var_explain <- var_explained_df %>%
-    ggplot(aes(x = PC,y = var_explained, group = 1))+
-    geom_point(size=1)+
-    geom_line() +
-    theme_minimal() +
-    scale_x_discrete(limits = paste0(rep("PC", 20), 1:20)) +
-    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
-          axis.text = element_text(size = 5),
-          axis.title = element_text(size = 7)) +
-    xlab("Principal Components (PC)") +
-    ylab("Variance explained")
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-data_pca <- data_pca[,1:7]
-UMAP_fit_pbmc <- umap(data_pca, n_neighbors = 15, n_components =  2)
-
-UMAP_data_pbmc <- UMAP_fit_pbmc$layout |>
-  as.data.frame()
-
-names(UMAP_data_pbmc)[1:(ncol(UMAP_data_pbmc))] <- paste0(rep("UMAP",(ncol(UMAP_data_pbmc))), 1:(ncol(UMAP_data_pbmc)))
-
-UMAP_data_pbmc <- UMAP_data_pbmc |>
-  mutate(ID = row_number())
-
-plot_pbmc <- UMAP_data_pbmc %>%
-    ggplot(aes(x = UMAP1,
-               y = UMAP2, color = data_pca_with_label$cluster_label))+
-    geom_point(alpha=0.5) +
-    coord_equal() +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+
+## Import data
+df_2 <- read_rds("data/pbmc/pbmc.rds")
+training_data_1 <- read_rds("data/pbmc/pbmc_training.rds")
+test_1 <- read_rds("data/pbmc/pbmc_test.rds")
+
+tSNE_pbmc <- read_rds("data/pbmc/pbmc_tsne_30.rds")
+UMAP_pbmc <- read_rds("data/pbmc/pbmc_umap.rds")
+PHATE_pbmc <- read_rds("data/pbmc/pbmc_phate.rds")
+TriMAP_pbmc <- read_rds("data/pbmc/pbmc_trimap.rds")
+PaCMAP_pbmc <- read_rds("data/pbmc/pbmc_pacmap.rds")
+
+## Visualise embeddings
+
+plot_list1_pbmc <- plot_tSNE_2D(tSNE_pbmc) + #ggtitle("(a)") + 
   theme_linedraw() +
@@
         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#b15928", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#cab2d6", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#bc80bd", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
-  guides(color = guide_legend(title = "Cell type"),
-         color = guide_legend(nrow = 3))
-```
-
-```{r}
-#| echo: false
-cell_area <- 1
-
-num_bins_pbmc <- calculate_effective_x_bins(.data = UMAP_data_pbmc, x = UMAP1,
-                                        cell_area = 1)
-shape_val_pbmc <- calculate_effective_shape_value(.data = UMAP_data_pbmc,
-                                              x = UMAP1, y = UMAP2)
-
-## To extract bin centroids
-hexbin_data_object_pbmc <- extract_hexbin_centroids(nldr_df = UMAP_data_pbmc, num_bins = num_bins_pbmc, shape_val = shape_val_pbmc)
-
-df_bin_centroids_pbmc <- hexbin_data_object_pbmc$hexdf_data
-
-# min_std_cell_threshold3 <- 0.01
-# 
-# df_bin_centroids_pbmc <- df_bin_centroids_pbmc |>
-#   dplyr::mutate(stand_cell_count = counts/max(counts)) |>
-#   dplyr::filter(stand_cell_count > min_std_cell_threshold3)
-
-UMAP_data_with_hb_id_pbmc <- UMAP_data_pbmc |> 
-  dplyr::mutate(hb_id = hexbin_data_object_pbmc$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-data_pca_n <- data_pca
-names(data_pca_n) <- paste0(rep("x", 7), 1:7)
-df_all_pbmc <- dplyr::bind_cols(data_pca_n, UMAP_data_with_hb_id_pbmc)
-
-## Averaged on high-D
-df_bin_pbmc <- avg_highD_data(.data = df_all_pbmc)
-
-## Triangulate bin centroids
-tr1_object_pbmc <- triangulate_bin_centroids(df_bin_centroids_pbmc, x, y)
-tr_from_to_df_pbmc <- generate_edge_info(triangular_object = tr1_object_pbmc)
-
-## Compute 2D distances
-distance_pbmc <- cal_2D_dist(.data = tr_from_to_df_pbmc)
-
-## To find the benchmark value
-benchmark_pbmc <- find_benchmark_value(.data = distance_pbmc, distance_col = distance)
-
-benchmark_pbmc <- 2
-
-names(df_all_pbmc)[1:7] <- paste0(rep("PC_", 7), 1:7)
-names(df_bin_pbmc) <- c("hb_id", paste0(rep("PC_", 7), 1:7))
-
-# tour_pbmc <- show_langevitour(df_all_pbmc, df_bin_pbmc, df_bin_centroids_pbmc, benchmark_value = benchmark_pbmc, distance = distance_pbmc, distance_col = distance, col_start = "PC")
-
-df_1 <- df_all_pbmc |> 
-  dplyr::select(tidyselect::starts_with("PC")) |>
-  dplyr::mutate(type = data_pca_with_label$cluster_label) ## original dataset
-
-df_b_1 <- df_bin_pbmc |> 
-  dplyr::filter(hb_id %in% df_bin_centroids_pbmc$hexID) |>
-  dplyr::select(-hb_id) |>
-  dplyr::mutate(type = "model") ## Data with summarized mean
-
-df_exe <- dplyr::bind_rows(df_b_1, df_1)
-
-## Set the maximum difference as the criteria
-distance_df_small_edges <- distance_pbmc %>% 
-  dplyr::filter(distance < benchmark_pbmc)
-## Since erase brushing is considerd.
-
-trimesh_pbmc <- ggplot(df_bin_centroids_pbmc, aes(x = x, y = y)) +
-  geom_trimesh() +
-  coord_equal()
-
-trimesh_pbmc <- trimesh_pbmc +
-  #ggtitle("(b)") +
-  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  theme(axis.text = element_text(size = 5),
-          axis.title = element_text(size = 7))
-
-trimesh_gr_pbmc <- colour_long_edges(.data = distance_pbmc, benchmark_value = benchmark_pbmc,
-                 triangular_object = tr1_object_pbmc, distance_col = distance)
-
-trimesh_gr_pbmc <- trimesh_gr_pbmc +
-  #ggtitle("(c)") +
-  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  theme(axis.text = element_text(size = 5),
-          axis.title = element_text(size = 7), #change legend key width
-        legend.title = element_text(size=7), #change legend title font size
-        legend.text = element_text(size=5))
-
-tour_pbmc <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)], lineFrom = distance_df_small_edges$from, lineTo = distance_df_small_edges$to, group = df_exe$type, levelColors = c("#33a02c", "#b2df8a", "#fb9a99", "#ff7f00", "#e31a1c", "#1f78b4", "black", "#b15928", "#6a3d9a", "#cab2d6"), pointSize = 3)
-```
-
-```{r}
-#| warning: false
-#| echo: false
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+plot_list2_pbmc <- plot_UMAP_2D(UMAP_pbmc) + #ggtitle("(b)") + 
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+plot_list3_pbmc <- plot_PHATE_2D(PHATE_pbmc) + #ggtitle("(c)") + 
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+plot_list4_pbmc <- plot_TriMAP_2D(TriMAP_pbmc) + #ggtitle("(d)") + 
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+plot_list5_pbmc <- plot_PaCMAP_2D(PaCMAP_pbmc) + #ggtitle("(e)") + 
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'e', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "2D layouts from different NLDR techniques applied for the training PBMC dataset: (a) tSNE (perplexity = 30), (b) UMAP (n_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2).
+Is there a best representation of the original data or are they all providing  equivalent information?"
+#| label: fig-nldervis5PBMC
 #| out-width: 100%
-#| label: fig-pbmc
-#| fig-pos: H
-#| fig-cap: Visualization and refinement of low-dimensional representation for pbmc data$\colon$ (a) 2D layout of UMAP (n-neigbors = 15), (b) triangular mesh, and (c) triangular mesh colored by edge type.
-Edges with lengths less than the benchmark value of $2$ are classified as small, while edges with lengths greater than or equal to this value are labeled as long.
-The combination of UMAP, hexagonal binning, and triangulation provides the effectiveness of our approach in visualizing and refining the low-dimensional representation of the pbmc data.
-
-plot_pbmc + trimesh_pbmc + trimesh_gr_pbmc + #A = 1, b = (19, 27), s = 1.255103
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect') &
-  theme(legend.position='bottom', plot.tag = element_text(size = 8))
-```
-
-<!--
-::: {#fig-pbmc_sc layout-ncol="3"}
-![](figures/pbmc_a.png){#fig-pbmc_sc1}
-
-![](figures/pbmc_b.png){#fig-pbmc_sc2}
-
-![](figures/pbmc_c.png){#fig-pbmc_sc3}
-
-Screen shots of the **langevitour** with the low-dimensional view applied to pbmc data set, a video of the tour animation is available at (\url{https://drive.google.com/file/d/1F7Xb2EIpfUruCxYihPw2rKro4FB5Y2uL/view}).
-Combination of the **langevitour** and low-dimensional representation of the pbmc data provides the effectiveness of our approach.
-:::-->
+
+plot_list1_pbmc + plot_list2_pbmc + plot_list3_pbmc + plot_list4_pbmc + plot_list5_pbmc +
+  plot_layout(ncol=5)
+```
 
@@
 
+## Import data
+df_2 <- read_rds("data/zeisel/zeisel.rds")
+training_data_1 <- read_rds("data/zeisel/zeisel_training.rds")
+test_1 <- read_rds("data/zeisel/zeisel_test.rds")
+
+tSNE_zeisel <- read_rds("data/zeisel/zeisel_tsne_30.rds")
+UMAP_zeisel <- read_rds("data/zeisel/zeisel_umap.rds")
+PHATE_zeisel <- read_rds("data/zeisel/zeisel_phate.rds")
+TriMAP_zeisel <- read_rds("data/zeisel/zeisel_trimap.rds")
+PaCMAP_zeisel <- read_rds("data/zeisel/zeisel_pacmap.rds")
+
+## Visualise embeddings
+
+plot_list1_zeisel <- plot_tSNE_2D(tSNE_zeisel) + #ggtitle("(a)") + 
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+plot_list2_zeisel <- plot_UMAP_2D(UMAP_zeisel) + #ggtitle("(b)") + 
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+plot_list3_zeisel <- plot_PHATE_2D(PHATE_zeisel) + #ggtitle("(c)") + 
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+plot_list4_zeisel <- plot_TriMAP_2D(TriMAP_zeisel) + #ggtitle("(d)") + 
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+plot_list5_zeisel <- plot_PaCMAP_2D(PaCMAP_zeisel) + #ggtitle("(e)") + 
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'e', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "2D layouts from different NLDR techniques applied for the training Zeisel mouse brain dataset: (a) tSNE (perplexity = 30), (b) UMAP (n_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2).
+Is there a best representation of the original data or are they all providing  equivalent information?"
+#| label: fig-nldervis5Mouse
+#| out-width: 100%
+
+plot_list1_zeisel + plot_list2_zeisel + plot_list3_zeisel + plot_list4_zeisel + plot_list5_zeisel +
+  plot_layout(ncol=5)
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
 MSE_df_zei <- read_rds("data/zeisel/summary_zei.rds")

### 2024-01-10 paper.tex
Add 2D layouts of PBMC and Mouse data

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, interior hidden, boxrule=0pt, frame hidden, enhanced, breakable, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden, breakable, enhanced]}{\end{tcolorbox}}\fi
 
@@
 Next, we employed the UMAP technique with default parameter settings. As
-illustrated in Figure~\ref{fig-pbmc}, the cell types B and Platelet are
+illustrated in \textbf{?@fig-pbmc}, the cell types B and Platelet are
 well-separated in the UMAP layout. Moreover, CD14+ Mono, FCGR3A+ Mono,
@@
 
-\begin{figure}[H]
+\begin{figure}
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-pbmc-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervis5PBMC-1.pdf}
 
@@
 
-\caption{\label{fig-pbmc}Visualization and refinement of low-dimensional
-representation for pbmc data\(\colon\) (a) 2D layout of UMAP (n-neigbors
-= 15), (b) triangular mesh, and (c) triangular mesh colored by edge
-type. Edges with lengths less than the benchmark value of \(2\) are
-classified as small, while edges with lengths greater than or equal to
-this value are labeled as long. The combination of UMAP, hexagonal
-binning, and triangulation provides the effectiveness of our approach in
-visualizing and refining the low-dimensional representation of the pbmc
-data.}
+\caption{\label{fig-nldervis5PBMC}2D layouts from different NLDR
+techniques applied for the training PBMC dataset: (a) tSNE (perplexity =
+30), (b) UMAP (n\_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP
+(n\_inliers = 5, n\_outliers = 4, n\_random = 3), and (e) PaCMAP
+(n\_neighbors = 10, init = random, MN\_ratio = 0.9, FP\_ratio = 2). Is
+there a best representation of the original data or are they all
+providing equivalent information?}
 
@@
 regions and the intricate networks that underpin neural processes.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervis5Mouse-1.pdf}
+
+}
+
+\caption{\label{fig-nldervis5Mouse}2D layouts from different NLDR
+techniques applied for the training Zeisel mouse brain dataset: (a) tSNE
+(perplexity = 30), (b) UMAP (n\_neighbors = 15), (c) PHATE (knn = 5),
+(d) TriMAP (n\_inliers = 5, n\_outliers = 4, n\_random = 3), and (e)
+PaCMAP (n\_neighbors = 10, init = random, MN\_ratio = 0.9, FP\_ratio =
+2). Is there a best representation of the original data or are they all
+providing equivalent information?}
+
+\end{figure}
 

### 2024-01-11 paper.tex
fix gitignore so that questions of overwriting are not asked on git pull

--- 2024-01-10 paper.tex
@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, enhanced, breakable, interior hidden, frame hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, sharp corners, boxrule=0pt, breakable]}{\end{tcolorbox}}\fi
 
@@
 
-\hypertarget{non-linear-dimension-reduction-techniques}{%
-\subsubsection{Non-linear dimension reduction
-techniques}\label{non-linear-dimension-reduction-techniques}}
+\hypertarget{introduction-to-umap}{%
+\subsubsection{Introduction to UMAP}\label{introduction-to-umap}}
 
-Non-linear dimension reduction techniques play a crucial role in the
-analysis and visualization of high-dimensional data, where the
-complexities of relationships among variables may not be adequately
-captured by linear methods. Among these techniques, tSNE stands out for
-its emphasis on preserving pairwise distances. By minimizing the
-divergence between probability distributions in both the high and
-low-dimensional spaces, t-SNE effectively reveals intricate structures
-and patterns within the data. Its application is widespread in tasks
-requiring the visualization of clusters and local relationships, though
-it does require careful consideration of the perplexity parameter for
-optimal results.
+The UMAP algorithm constructs a fuzzy topological representation of the
+high-dimensional data and then optimizes the low-dimensional embedding
+to be as similar to this representation as possible \citep{Leland2018}.
 
-UMAP is another powerful non-linear technique that strikes a balance
-between preserving local and global structures. Constructing a fuzzy
-topological representation using a weighted k-nearest neighbors graph,
-UMAP optimizes the low-dimensional embedding to resemble this
-representation. Known for its efficiency and scalability, UMAP is
-versatile across various scales of relationships in the data, although
-parameter sensitivity, particularly concerning the choice of neighbors,
-must be taken into account.
+The UMAP algorithm constructs a weighted k-nearest neighbors graph based
+on the local similarity of data points in the original high-dimensional
+space. The weighted graph is defined by a conditional probability
+distribution \(P_{ij}\), which represents the probability of observing
+data point \(x_i\) given data point \(x_j\) as a neighbor:
 
-For trajectory data, PHATE provides specialized capabilities. It models
-the affinity between data points, simulating a heat diffusion process to
-capture developmental processes, particularly in single-cell genomics.
-While PHATE excels in revealing trajectory structures and offering
-insights into cellular development, it necessitates careful parameter
-tuning due to its specialized nature.
+\[
+P_{ij} = \frac{\exp(-\lVert x_i - x_j \rVert^2)}{\sum_{k \neq i} \exp(-\lVert x_i - x_k \rVert^2)}
+\]
 
-TriMAP adopts a unique approach by approximating the data manifold
-through the construction of a triangulated graph representation. This
-technique efficiently captures both global and local structures by
-representing the data as a network of triangles. TriMAP's strength lies
-in its ability to efficiently capture complex structures, albeit with
-sensitivity to parameter choices, including the number of neighbors.
+Similarly, a fuzzy topological representation of the data is constructed
+in the low-dimensional space with conditional probability distribution
+\(Q_{ij}\):
 
-In contrast, PaCMAP introduces supervised learning to create a
-low-dimensional representation while preserving pair-wise relationships.
-Constructing a graph based on pair-wise distances, PaCMAP optimizes an
-embedding using a customizable loss function. Particularly notable is
-PaCMAP's flexibility in incorporating class labels or additional
-information to guide the embedding process, offering users a means to
-customize its behavior and performance.
+\[
+Q_{ij} = \frac{\exp(-\lVert y_i - y_j \rVert^2)}{\sum_{k \neq i} \exp(-\lVert y_i - y_k \rVert^2)}
+\]
+
+where \(y_i\) and \(y_j\) are the corresponding low-dimensional
+representations of data points \(x_i\) and \(x_j\).
+
+UMAP minimizes the cross-entropy between the high-dimensional fuzzy
+topological representation \(P_{ij}\) and the low-dimensional
+representation \(Q_{ij}\) to find an optimal low-dimensional embedding.
+The loss function for UMAP is defined as:
+
+\[
+L = \sum_{i} \sum_{j} P_{ij} \log \frac{P_{ij}}{Q_{ij}}
+\]
+
+The algorithm uses stochastic gradient descent to optimize the positions
+of the data points in the low-dimensional space to minimize this loss
+function.
+
+In R, the \texttt{umap} package is often used to implement UMAP and
+generate results.
+
+The ``n\_neighbors'' parameter in UMAP determines the number of nearest
+neighbors to consider for each data point during the construction of the
+fuzzy topological representation. Higher values of ``n\_neighbors''
+capture more global structure, while lower values emphasize local
+structure. Choosing an appropriate value for ``n\_neighbors'' is
+crucial, as it affects the balance between local and global structure in
+the resulting low-dimensional representation.
 
@@
 \caption{\label{fig-nldervis}2D layouts from different NLDR techniques
-applied the same data: (a) tSNE (perplexity = 27), (b) UMAP
+applied the same data: (a) tSNE (perplexity = 32), (b) UMAP
 (n\_neighbors = 50), (c) PHATE (knn = 5), (d) TriMAP (n\_inliers = 5,
@@
 for the same \textbf{s\_curve\_noise\_umap} data: (a) b = (4, 8), s =
-1.643542, (b) b = (12, 22), s = 1.643542, and (c) b = (21, 40), s =
-1.643542. The hexbins are colored based on the density of points, with
-yellow indicating higher point density and darker colors representing
-lower point density within each bin. Does a value of number of bins
-exist to effectively represent the low-dimensional data?}
+2.031, (b) b = (9, 20), s = 2.031, and (c) b = (16, 36), s = 2.031. The
+hexbins are colored based on the density of points, with yellow
+indicating higher point density and darker colors representing lower
+point density within each bin. Does a value of number of bins exist to
+effectively represent the low-dimensional data?}
 
@@
 \subsection{Tunning the model}\label{tunning-the-model}}
-
-In our model tuning process, we strategically adjust three key
-parameters to optimize the performance and accuracy of our approach.
-They are (i) the total number of bins (\(b\)), (ii) a benchmark value to
-remove low-density hexagons, and (iii) a benchmark value to remove long
-edges.
-
-\hypertarget{total-number-of-bins}{%
-\subsection{Total number of bins}\label{total-number-of-bins}}
-
-The total number of bins is a crucial parameter that defines the
-granularity of our hexagonal grid. This parameter particularly
-influential in capturing the intricacies of the data structure. It is
-determined by multiplying the number of bins along the x-axis ((b\_1))
-with the number of bins along the y-axis ((b\_2)), as per the formula:
-
-\begin{equation}\protect\hypertarget{eq-equationp1}{}{
-b = b_1 \times b_2
-}\label{eq-equationp1}\end{equation}
-
-Here, (b) represents the total number of bins. Adjusting the parameter
-(b\_1) provides control over the total bin count ((b)), allowing us to
-fine-tune the grid's configuration along the x-axis based on specific
-requirements.
-
-\hypertarget{benchmark-value-to-remove-low-density-hexagons}{%
-\subsection{Benchmark value to remove low-density
-hexagons}\label{benchmark-value-to-remove-low-density-hexagons}}
-
-Addressing low-density hexagons is a systematic process to handle
-sparsely represented data in certain regions. For each hex bin, we
-identify the six nearest hex bins using an equal 2D distance metric.
-Then, we calculate the mean density, as outlined in the equations:
-
-\begin{equation}\protect\hypertarget{eq-equationp2}{}{
-\text{standard count} = \frac{\text{count}}{\text{max count}} 
-}\label{eq-equationp2}\end{equation}
-
-\begin{equation}\protect\hypertarget{eq-equationp3}{}{
-\text{mean density} = \frac{\text{standard count}}{6} 
-}\label{eq-equationp3}\end{equation}
-
-The standard count is derived from the number of observations in the hex
-bins. By examining the distribution of mean densities and designating
-the first quartile as the benchmark value, hex bins with mean densities
-below this benchmark are removed. This process ensures the elimination
-of regions with insufficient data density, focusing on areas with more
-significant data representation and preserving the overall structure in
-the low-dimensional space.
-
-\hypertarget{benchmark-value-to-remove-long-edges}{%
-\subsection{Benchmark value to remove long
-edges}\label{benchmark-value-to-remove-long-edges}}
-
-The removal of long edges is a critical step to create a smoother
-representation by iteratively eliminating hexagons with excessive
-distances between centroids. This process eliminates outliers and noise
-while preserving essential local relationships within the data. To
-achieve this, distances between vertices are sorted, and unique distance
-values are extracted. By setting a threshold based on the largest
-difference between consecutive distance values, long edges are
-identified and removed. This refinement step contributes to enhancing
-the quality of the triangular mesh, ensuring a more accurate
-representation of the data structure.
 
@@
 
+{\centering \includegraphics{paper_files/figure-pdf/fig-predict-1.pdf}
+
+}
+
+\caption{\label{fig-predict}The .}
+
+\end{figure}
+
+\begin{figure}
+
 {\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticplt-1.pdf}
@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticpltDiffParam-1.pdf}
-
-}
-
-\caption{\label{fig-diagnosticpltDiffParam}The .}
-
-\end{figure}
-
 \hypertarget{sec-simpleex}{%
-\subsection{Simulated data example}\label{sec-simpleex}}
-
-In this section, we showcase the effectiveness of our methodology using
-simulated data. The dataset comprises five spherical Gaussian clusters
-in 4-\(d\), with each cluster containing an equal number of points and
-consistent within variation.
-
-We \emph{strongly} recommend viewing the linked videos for each study
-while reading. Links to the videos are available in the figures for each
-example. The videos show the visual appearance of the
-\textbf{langevitour} interface with low-dimensional view and how we can
-interact with the tour via the controls.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervis5Gau-1.pdf}
-
-}
-
-\caption{\label{fig-nldervis5Gau}2D layouts from different NLDR
-techniques applied the same data: (a) tSNE (perplexity = 61), (b) UMAP
-(n\_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n\_inliers = 5,
-n\_outliers = 4, n\_random = 3), and (e) PaCMAP (n\_neighbors = 10, init
-= random, MN\_ratio = 0.9, FP\_ratio = 2). Is there a best
-representation of the original data or are they all providing equivalent
-information?}
-
-\end{figure}
-
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-43-1.pdf}
+\subsection{Simulated data examples}\label{sec-simpleex}}
 
@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticpltPBMC-1.pdf}
-
-}
-
-\caption{\label{fig-diagnosticpltPBMC}The .}
-
-\end{figure}
-
 \hypertarget{single-cell-tagged-reverse-transcription-sequencing-data-of-mouse}{%
@@
 regions and the intricate networks that underpin neural processes.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticpltZEI-1.pdf}
-
-}
-
-\caption{\label{fig-diagnosticpltZEI}The .}
-
-\end{figure}
 

### 2024-01-11 paper.tex

--- 2024-01-10 paper.tex
@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden, breakable, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, enhanced, sharp corners]}{\end{tcolorbox}}\fi
 

### 2024-01-11 paper.qmd
Update the color of the points and consistant with coloring

@@
 
-data <- read_csv(paste0(here::here(), "/data/s_curve.csv"))
+data <- read_rds("data/s_curve/s_curve.rds")
 ```
@@
 plot_list1 + plot_list2 + plot_list3 + plot_list4 + plot_list5 +
-  plot_layout(ncol=5, widths=c(0.35, 0.15, 0.30, 0.12, 0.11))
+  plot_layout(ncol=5)
 ```
@@
 
-### Total number of bins
+#### Total number of bins
 
@@
 
-### Benchmark value to remove low-density hexagons
+#### Benchmark value to remove low-density hexagons
 
@@
 
-### Benchmark value to remove long edges
+#### Benchmark value to remove long edges
 
@@
 
-### Summaries of the model {#sec-summary}
+### Model summaries {#sec-summary}
 
@@
 The concept of "residuals" is pivotal in evaluating the accuracy of representing bin centroids in high dimensions.
-To quantify this accuracy, we introduce an error metric, which measures the sum of squared differences between the high-dimensional data (\(x_{ij}\)) and the predicted bin centroid data in high-dimensional space (\(C_{x_ij}\)) across all bins and dimensions.
+To quantify this accuracy, we introduce an error metric, which measures the sum of squared differences between the high-dimensional data ($x_{ij}$) and the predicted bin centroid data in high-dimensional space ($C_{x_ij}$) across all bins and dimensions.
 Mathematically, this error is expressed as:
@@
 
-Here, \(n\) represents the number of bins, \(p\) represents the dimensions, \(x_{ij}\) is the actual high-dimensional data, and \(C_{x_ij}\) is the predicted bin centroid data in high dimensions.
+Here, $n$ represents the number of bins, $p$ represents the dimensions, $x_{ij}$ is the actual high-dimensional data, and $C_{x_ij}$ is the predicted bin centroid data in high dimensions.
 
@@
 
-data <- read_csv("data/s_curve.csv")
-data <- data |>
-  mutate(ID = row_number())
+data <- read_rds("data/s_curve/s_curve.rds")
 
@@
 MSE_df <- dplyr::bind_rows(MSE_df_1, MSE_df_2, MSE_df_3, MSE_df_4, MSE_df_5)
+
+MSE_df$method <- factor(MSE_df$method, levels = c("tSNE", "UMAP", "PHATE", "TriMAP", "PaCMAP"))
 
@@
         axis.text = element_text(size = 7)) +
-  scale_color_discrete_qualitative() +
+  scale_colour_manual(values = c("#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
   ylab("AIC") +
@@
   # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  scale_color_discrete_qualitative() +
+  scale_colour_manual(values = c("#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
   ylab("MSE") +
@@
 
+<!--UMAP with different param choices-->
+
+```{r}
+#| warning: false
+#| echo: false
+
+UMAP_data_7 <- read_rds(file = "data/s_curve/s_curve_umap_7.rds")
+
+#(n-neighbors: 50)
+plot_list1_umap <- plot_UMAP_2D(UMAP_data_7) + #ggtitle("(b)") +
+  geom_point(color = "#8dd3c7") +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+UMAP_data_15 <- read_rds(file = "data/s_curve/s_curve_umap_15.rds")
+
+#(n-neighbors: 50)
+plot_list2_umap <- plot_UMAP_2D(UMAP_data_15) + #ggtitle("(b)") + 
+  geom_point(color = "#a65628") +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+UMAP_data_32 <- read_rds(file = "data/s_curve/s_curve_umap_32.rds")
+
+#(n-neighbors: 50)
+plot_list3_umap <- plot_UMAP_2D(UMAP_data_32) + #ggtitle("(b)") +
+  geom_point(color = "#f781bf") +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+UMAP_data <- read_rds(file = "data/s_curve/s_curve_umap.rds")
+
+#(n-neighbors: 50)
+plot_list4_umap <- plot_UMAP_2D(UMAP_data) + #ggtitle("(b)") +
+  geom_point(color = "#999999") +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'd', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "2D layouts from UMAP applied for the S-curve data: (a) UMAP (n_neighbors = 7), (b) UMAP (n_neighbors = 15), (c) UMAP (n_neighbors = 32), (d) UMAP (n_neighbors = 50).
+Is there a best parameter choice in representing UMAP or are they all providing  equivalent information?"
+#| label: fig-nldervisUMAP
+#| out-width: 100%
+
+plot_list1_umap + plot_list2_umap + plot_list3_umap + plot_list4_umap +
+  plot_layout(ncol=4)
+```
+
 ```{r}
@@
 MSE_df_umap <- dplyr::bind_rows(MSE_df_umap_1, MSE_df_umap_2, MSE_df_umap_3, MSE_df_umap_4)
+MSE_df_umap$param <- factor(MSE_df_umap$param, levels = c("n_neighbors: 7", "n_neighbors: 15", "n_neighbors: 32", "n_neighbors: 50"))
 
@@
         axis.text = element_text(size = 7)) +
-  scale_color_discrete_qualitative() +
+  scale_color_manual(values = c("#8dd3c7", "#a65628", "#f781bf", "#999999")) +
   ylab("AIC") +
@@
   # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  scale_color_discrete_qualitative() +
+  scale_color_manual(values = c("#8dd3c7", "#a65628", "#f781bf", "#999999")) +
   ylab("MSE") +
@@
 MSE_df_gau <- dplyr::bind_rows(MSE_df_1_gau, MSE_df_2_gau, MSE_df_3_gau, MSE_df_4_gau, MSE_df_5_gau)
+
+MSE_df_gau$method <- factor(MSE_df_gau$method, levels = c("tSNE", "UMAP", "PHATE", "TriMAP", "PaCMAP"))
 
@@
 ```
-
-<!--six different clusters-->
-
-<!--Gaussian globe or sphere-->
-
-<!--In this section, we demonstrate the effectiveness of our approach through the use of simulation data.
-These simulations are designed to have known cluster structures and underlying data geometries.
-We begin with a straightforward example where we generate spherical clusters with a doublet structure.
-This allows us to assess how well our algorithm can identify and represent these known structures in the low-dimensional space.
-
-We then proceed to a more complex example where the data exhibits a more intricate geometry.
-This simulation provides a challenging test for our approach to capture and preserve the underlying data structure accurately.
-
-We *strongly* recommend viewing the linked videos for each study while reading.
-Links to the videos are available in the figures for each example.
-The videos show the visual appearance of the **langevitour** interface with low-dimensional view and how we can interact with the tour via the controls.
-
-**Example 1: Exploring spherical Gaussian clusters with doublet structure**
-
-The next data set consists of two distinct types of clusters embedded in a 10-$d$ space.
-The first set of clusters includes three Gaussian clusters with equal variance.
-Each of these Gaussian clusters is equidistant from the others, forming a well-separated and symmetrical arrangement.
-The second set of clusters consists of three clusters with a doublet structure.
-These clusters are also embedded in the 10-$d$ space, but each is positioned in the middle of its parent Gaussian clusters.
-Moreover, each doublet cluster contains an average number of points from its corresponding parent Gaussian clusters.
-
-For analysis, we run tSNE on the training data with a perplexity setting 18.
-@fig-example3 (a) displays the tSNE results, indicating that the sub-clusters have been correctly identified.
-However, the relative locations of certain clusters to each other appear distorted in the low-dimensional view, such as the red and pink clusters being far apart.
-
-Nonetheless, we find no apparent impact on the perception of the data structure when we examine the tour view and the model (see @fig-example3) through the linked video (see @fig-exp1_sc).
-The model demonstrates accurate identification of clusters and captures the internal variations within each cluster, reflecting the inherent variance present in the data.
-The configurations used for the analysis are detailed in @tbl-table02.
-Additionally, the representation of the parent clusters in the low-dimensional space appears stretched, aligning with the variance of these clusters.
-On the other hand, the doublet clusters are squeezed, reflecting their specific characteristics and relationship to their parent Gaussian clusters.
-
-```{r}
-#| warning: false
-#| echo: false
-sample_size <- 450
-cluster_size <- sample_size/4.5
-
-df1 <- tibble::tibble(x1=rnorm(cluster_size, mean = 3, sd = 0.05), x2 = rnorm(cluster_size, mean = 1, sd = 0.05), x3=rnorm(cluster_size, mean = 1, sd = 0.05), x4=rnorm(cluster_size, mean = 1, sd = 0.05),
-                      x5=rnorm(cluster_size, mean = 1, sd = 0.05),
-                      x6=rnorm(cluster_size, mean = 1, sd = 0.05),
-                      x7=rnorm(cluster_size, mean = 1, sd = 0.05),
-                      x8=rnorm(cluster_size, mean = 1, sd = 0.05),
-                      x9=rnorm(cluster_size, mean = 1, sd = 0.05),
-                      x10=rnorm(cluster_size, mean = 1, sd = 0.05))
-
-df1 <- df1 %>%
-  mutate(cluster = 1)
-
-df2 <- tibble::tibble(x1=rnorm(cluster_size, mean = 1, sd = 0.05), x2=rnorm(cluster_size, mean = 1, sd = 0.05), x3=rnorm(cluster_size, mean = 1, sd = 0.05), x4=rnorm(cluster_size, mean = 1, sd = 0.05),
-                      x5=rnorm(cluster_size, mean = 1, sd = 0.05),
-                      x6=rnorm(cluster_size, mean = 1, sd = 0.05),
-                      x7=rnorm(cluster_size, mean = 1, sd = 0.05),
-                      x8=rnorm(cluster_size, mean = 1, sd = 0.05),
-                      x9=rnorm(cluster_size, mean = 1, sd = 0.05),
-                      x10=rnorm(cluster_size, mean = 1, sd = 0.05))
-
-df2 <- df2 %>%
-  mutate(cluster = 2)
-
-df3_new <- (df1 + df2) / 2
-
-samp <- sample(nrow(df3_new), cluster_size * 0.50) ## 50% from the original dataset
-
-# data in the sample
-df3 <- df3_new[samp,]
-
-df3 <- df3 %>%
-  mutate(cluster = 3)
-
-df4 <- tibble::tibble(x1=rnorm(cluster_size, mean = 1, sd = 0.05), x2=rnorm(cluster_size, mean = 1, sd = 0.05), x3=rnorm(cluster_size, mean = 1, sd = 0.05), x4=rnorm(cluster_size, mean = 3, sd = 0.05),
-                      x5=rnorm(cluster_size, mean = 1, sd = 0.05),
-                      x6=rnorm(cluster_size, mean = 1, sd = 0.05),
-                      x7=rnorm(cluster_size, mean = 1, sd = 0.05),
-                      x8=rnorm(cluster_size, mean = 1, sd = 0.05),
-                      x9=rnorm(cluster_size, mean = 1, sd = 0.05),
-                      x10=rnorm(cluster_size, mean = 1, sd = 0.05))
-
-df4 <- df4 %>%
-  mutate(cluster = 4)
-
-df5_new <- (df2 + df4) / 2
-
-samp1 <- sample(nrow(df5_new), cluster_size * 0.50) ## 50% from the original dataset
-
-df5 <- df5_new[samp1,]
-
-df5 <- df5 %>%
-  mutate(cluster = 5)
-
-df6_new <- (df1 + df4) / 2
-
-samp2 <- sample(nrow(df6_new), cluster_size * 0.50) ## 50% from the original dataset
-
-df6 <- df6_new[samp2,]
-
-df6 <- df6 %>%
-  mutate(cluster = 6)
-
-df_3 <- bind_rows(df1, df2, df3, df4, df5, df6)
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-# langevitour::langevitour(df_3[, 1:10], group = df_3$cluster, levelColors = c("#b15928", "#1f78b4", "#bc80bd", "#33a02c", "#fb9a99", "#e31a1c"))
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-df_3 <- df_3 |>
-  mutate(ID = row_number())
-
-data_split_sp <- initial_split(df_3)
-training_data_5 <- training(data_split_sp) |>
-  arrange(ID)
-test_data_5 <- testing(data_split_sp) |>
-  arrange(ID)
-
-clusters <- training_data_5$cluster
-```
-
-```{r}
-#| warning: false
-#| echo: false 
-tSNE_data_gau <- Fit_tSNE(training_data_5 |> dplyr::select(-ID, -cluster), with_seed = 20240110)
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-tSNE_cluster <- tSNE_data_gau %>%
-  ggplot(aes(x = tSNE1,
-             y = tSNE2, color = as.character(clusters))) +
-  geom_point(alpha=0.5) +
-  coord_equal() +
-  theme_linedraw() +
-  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold"),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-          axis.text = element_text(size = 5),
-          axis.title = element_text(size = 7), #change legend key width
-        legend.title = element_text(size=7), #change legend title font size
-        legend.text = element_text(size=7)) +
-  scale_color_manual(values=c("#b15928", "#1f78b4", "#bc80bd", "#33a02c", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#cab2d6", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
-  guides(color = guide_legend(title = "Cluster"))
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-cell_area <- 1
-
-num_bins_tsne_gau <- calculate_effective_x_bins(.data = tSNE_data_gau, x = tSNE1,
-                                                cell_area = 1)
-shape_val_tsne_gau <- calculate_effective_shape_value(.data = tSNE_data_gau,
-                                                      x = tSNE1, y = tSNE2)
-
-## To extract bin centroids
-hexbin_data_object_tsne_gau <- extract_hexbin_centroids(nldr_df = tSNE_data_gau, num_bins = num_bins_tsne_gau, shape_val = shape_val_tsne_gau, x = tSNE1, y = tSNE2)
-
-df_bin_centroids_tsne_gau <- hexbin_data_object_tsne_gau$hexdf_data
-
-# min_std_cell_threshold3 <- 0.01
-# 
-# df_bin_centroids_pbmc <- df_bin_centroids_pbmc |>
-#   dplyr::mutate(stand_cell_count = counts/max(counts)) |>
-#   dplyr::filter(stand_cell_count > min_std_cell_threshold3)
-
-tSNE_data_with_hb_id_gau <- tSNE_data_gau |> 
-  dplyr::mutate(hb_id = hexbin_data_object_tsne_gau$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_tsne_gau <- dplyr::bind_cols(training_data_5 |> dplyr::select(-ID), tSNE_data_with_hb_id_gau)
-
-## Averaged on high-D
-df_bin_tsne_gau <- avg_highD_data(.data = df_all_tsne_gau)
-
-## Triangulate bin centroids
-tr1_object_tsne_gau <- triangulate_bin_centroids(df_bin_centroids_tsne_gau, x, y)
-tr_from_to_df_tsne_gau <- generate_edge_info(triangular_object = tr1_object_tsne_gau)
-
-## Compute 2D distances
-distance_tsne_gau <- cal_2D_dist(.data = tr_from_to_df_tsne_gau)
-
-## To find the benchmark value
-benchmark_tsne_gau <- find_benchmark_value(.data = distance_tsne_gau, distance_col = distance)
-#benchmark_tsne <- 5
-
-tour_tsne_gau <- show_langevitour(df_all_tsne_gau, df_bin_tsne_gau, df_bin_centroids_tsne_gau, benchmark_value = benchmark_tsne_gau, distance = distance_tsne_gau, distance_col = distance)
-
-trimesh_cluster <- ggplot(df_bin_centroids_tsne_gau, aes(x = x, y = y)) +
-  geom_trimesh() +
-  coord_equal()
-
-trimesh_cluster <- trimesh_cluster +
-  #ggtitle("(b)") +
-  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  theme(axis.text = element_text(size = 5),
-        axis.title = element_text(size = 7))
-
-trimesh_gr_cluster <- colour_long_edges(.data = distance_tsne_gau, benchmark_value = benchmark_tsne_gau,
-                                     triangular_object = tr1_object_tsne_gau, distance_col = distance)
-
-trimesh_gr_cluster <- trimesh_gr_cluster +
-  #ggtitle("(c)") +
-  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  theme(axis.text = element_text(size = 5),
-        axis.title = element_text(size = 7), #change legend key width
-        legend.title = element_text(size=7), #change legend title font size
-        legend.text = element_text(size=5))
-
-tour_cluster <- show_langevitour(df_all_tsne_gau, df_bin_tsne_gau, df_bin_centroids_tsne_gau, benchmark_value = benchmark_tsne_gau, distance = distance_tsne_gau, distance_col = distance)
-```
-
-```{r}
-#| warning: false
-#| echo: false
-#| out-width: 100%
-#| label: fig-example3
-#| fig-cap: Visualization and refinement of low-dimensional representation for sub-clustered data$\colon$ (a) 2D layout of tSNE (perplexity = 18), (b) triangular mesh, and (c) triangular mesh colored by edge type.
-Edges with lengths less than the benchmark value of $2$ are classifiedas small, while edges with lengths greater than or equal to this value are labeled as long.
-The combination of tSNE, hexagonal binning, and triangulation provides the effectiveness of our approach in visualizing and refining the low-dimensional representation of the sub-clustered data.
-
-tSNE_cluster + trimesh_cluster + trimesh_gr_cluster +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect') &
-  theme(legend.position='bottom', plot.tag = element_text(size = 8))
-```
-
-::: {#fig-exp1_sc layout-ncol="3"}
-![](figures/example_1a.png){#fig-exp1_sc1}
-
-![](figures/example_1b.png){#fig-exp1_sc2}
-
-![](figures/example_1c.png){#fig-exp1_sc3}
-
-Screen shots of the **langevitour** with the low-dimensional view applied to sub-clustered data set, a video of the tour animation is available at (\url{https://drive.google.com/file/d/15nAbmc4UoVlUyFcGtnlnmyxrJEbK39nx/view}).
-Combination of the **langevitour** and low-dimensional representation of the sub-clustered data provides the effectiveness of our approach.
-:::
-
-**Example 2: Exploring data with piecewise linear structure**
-
-In our next exploration, we delve into simulated noisy tree-structured data designed to mimic branching trajectories of cell differentiation.
-The simulation is crafted to resemble the hierarchical clustering pattern observed in cell differentiation trajectories, where mature cells are akin to the tips of branches.
-
-To begin our analysis, we apply principal components to the data and focus on the first ten principal components, collectively accounting for approximately 68% of the variance explained in the data set.
-Then, we run TriMAP with the default parameters on the training data (n_inliers: 5, n_outliers: 4, and n_random: 3).
-Through this process, we create a low-dimensional manifold using the configurations detailed in @tbl-table02.
-Finally, we present the model along with the data using the tour view (Figure @fig-exp2_sc), allowing us to visualize and explore the intricate relationships and structures present in the data's low-dimensional representation.
-
-```{r}
-#| warning: false
-#| echo: false
-
-data_phate <- read_csv(paste0(here::here(), "/data/data_phate.csv"))
-
-data_phate <- data_phate |>
-  dplyr::mutate(ID = row_number())
-```
-
-```{r}
-#| warning: false
-#| echo: false 
-data_split <- initial_split(data_phate)
-training_data <- training(data_split) |>
-  arrange(ID)
-test_data <- testing(data_split) |>
-  arrange(ID)
-```
-
-```{r}
-#| warning: false
-#| echo: false 
-## num_pcs: selected number of PCs to get results (deafukt is all)
-
-calculate_pca <- function(data, num_pcs = NCOL(data), center = TRUE, scale = FALSE){
-  pcaY_cal <- prcomp(data, center = TRUE, scale = FALSE)
-
-  PCAresults <- tibble::as_tibble(pcaY_cal$x[, 1:num_pcs])
-  names(PCAresults) <- paste0(rep("PC", num_pcs), 1:num_pcs)
-
-  summary_pca <- summary(pcaY_cal)
-
-  var_explained_df <- tibble::tibble(PC = paste0("PC",1:NCOL(data)),
-                               var_explained = (pcaY_cal$sdev[1:NCOL(data)])^2/sum((pcaY_cal$sdev[1:NCOL(data)])^2))
-
-  ## Scree plot
-  plot <- var_explained_df %>%
-    ggplot(aes(x = PC,y = var_explained, group = 1))+
-    geom_point(size=1)+
-    geom_line() +
-    theme_minimal() +
-    scale_x_discrete(limits = paste0(rep("PC", 20), 1:20)) +
-    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
-          axis.text = element_text(size = 5),
-          axis.title = element_text(size = 7)) +
-    xlab("Principal Components (PC)") +
-    ylab("Variance explained")
-
-  return(list(prcomp_out = pcaY_cal, pca_components = PCAresults, summary = summary_pca, var_explained_pca  = var_explained_df, plot = plot))
-}
-```
-
-```{r}
-#| warning: false
-#| echo: false 
-pca_out <- calculate_pca(training_data, 10)
-data_pca <- pca_out$pca_components
-data_pca <- data_pca %>% 
-  distinct()
-
-var_explain_example2 <- pca_out$plot
-```
-
-```{r}
-#| warning: false
-#| echo: false 
-TriMAP_data_tree <- read_csv(paste0(here::here(), "/data/trimap_data_linear_structure_pca.csv")) %>%
-    mutate(ID = row_number())
-
-```
-
-```{r}
-#| warning: false
-#| echo: false 
-TriMAP_tree <- plot_TriMAP_2D(TriMAP_data_tree) +
-  theme_linedraw() +
-  theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank())
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-cell_area <- 1
-
-num_bins_tree <- calculate_effective_x_bins(.data = TriMAP_data_tree, x = TriMAP1,
-                                            cell_area = 1)
-shape_val_tree <- calculate_effective_shape_value(.data = TriMAP_data_tree,
-                                                  x = TriMAP1, y = TriMAP2)
-
-## To extract bin centroids
-hexbin_data_object_tree <- extract_hexbin_centroids(nldr_df = TriMAP_data_tree, num_bins = num_bins_tree, shape_val = shape_val_tree, x = TriMAP1, y = TriMAP2)
-
-df_bin_centroids_tree <- hexbin_data_object_tree$hexdf_data
-
-# min_std_cell_threshold3 <- 0.01
-# 
-# df_bin_centroids_tree <- df_bin_centroids_tree |>
-#   dplyr::mutate(stand_cell_count = counts/max(counts)) |>
-#   dplyr::filter(stand_cell_count > min_std_cell_threshold3)
-
-TriMAP_data_with_hb_id_tree <- TriMAP_data_tree |> 
-  dplyr::mutate(hb_id = hexbin_data_object_tree$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-data_pca_n <- data_pca
-names(data_pca_n) <- paste0(rep("x", 7), 1:7)
-df_all_tree <- dplyr::bind_cols(data_pca_n, TriMAP_data_with_hb_id_tree)
-
-## Averaged on high-D
-df_bin_tree <- avg_highD_data(.data = df_all_tree)
-
-## Triangulate bin centroids
-tr1_object_tree <- triangulate_bin_centroids(df_bin_centroids_tree, x, y)
-tr_from_to_df_tree <- generate_edge_info(triangular_object = tr1_object_tree)
-
-## Compute 2D distances
-distance_tree <- cal_2D_dist(.data = tr_from_to_df_tree)
-
-## To find the benchmark value
-benchmark_tree <- find_benchmark_value(.data = distance_tree, distance_col = distance)
-
-benchmark_tree <- 2
-
-names(df_all_tree)[1:7] <- paste0(rep("PC_", 7), 1:7)
-names(df_bin_tree) <- c("hb_id", paste0(rep("PC_", 7), 1:7))
-
-# tour_tree <- show_langevitour(df_all_tree, df_bin_tree, df_bin_centroids_tree, benchmark_value = benchmark_tree, distance = distance_tree, distance_col = distance, col_start = "PC")
-
-trimesh_tree <- ggplot(df_bin_centroids_tree, aes(x = x, y = y)) +
-  geom_trimesh() +
-  coord_equal()
-
-trimesh_tree <- trimesh_tree +
-  #ggtitle("(b)") +
-  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  theme(axis.text = element_text(size = 5),
-        axis.title = element_text(size = 7))
-
-trimesh_gr_tree <- colour_long_edges(.data = distance_tree, benchmark_value = benchmark_tree,
-                                     triangular_object = tr1_object_tree, distance_col = distance)
-
-trimesh_gr_tree <- trimesh_gr_tree +
-  #ggtitle("(c)") +
-  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  theme(axis.text = element_text(size = 5),
-        axis.title = element_text(size = 7), #change legend key width
-        legend.title = element_text(size=7), #change legend title font size
-        legend.text = element_text(size=5))
-
-tour_tree <- show_langevitour(df_all_tree, df_bin_tree, df_bin_centroids_tree, benchmark_value = benchmark_tree, distance = distance_tree, distance_col = distance, col_start = "PC")
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-#| out-width: 100%
-#| label: fig-example2
-#| fig-pos: H
-#| fig-cap: Visualization and refinement of low-dimensional representation for tree structured data$\colon$ (a) 2D layout of TriMAP (n-inliers = 5, n-outliers = 4, n-random = 3), (b) triangular mesh, and (c) triangular mesh colored by edge type.
-Edges with lengths less than the benchmark value of $2$ are classified as small, while edges with lengths greater than or equal to this value are labeled as long.
-The combination of TriMAP, hexagonal binning, and triangulation provides the effectiveness of our approach in visualizing and refining the low-dimensional representation of the tree structured data.
-
-TriMAP_tree + trimesh_tree + trimesh_gr_tree +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect') &
-  theme(legend.position='bottom', plot.tag = element_text(size = 8))
-```
-
-::: {#fig-exp2_sc layout-ncol="3"}
-![](figures/example_2a.png)
-
-![](figures/example_2b.png)
-
-![](figures/example_2c.png)
-
-Screen shots of the **langevitour** with the low-dimensional view applied to tree structured data set, a video of the tour animation is available at (\url{https://drive.google.com/file/d/1YQx4UGlzDify1mGqeDtlqTHncgqidDhq/view}).
-Combination of the **langevitour** and low-dimensional representation of the tree structured data provides the effectiveness of our approach.
-:::
--->
 
@@
 MSE_df_pbmc <- read_rds("data/pbmc/summary_pbmc.rds")
+MSE_df_pbmc$method <- factor(MSE_df_pbmc$method, levels = c("tSNE", "UMAP", "PHATE", "TriMAP", "PaCMAP"))
 
@@
 MSE_df_zei <- read_rds("data/zeisel/summary_zei.rds")
+MSE_df_zei$method <- factor(MSE_df_zei$method, levels = c("tSNE", "UMAP", "PHATE", "TriMAP", "PaCMAP"))
 

### 2024-01-11 paper.tex
Update the color of the points and consistant with coloring

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, enhanced, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden, breakable, enhanced]}{\end{tcolorbox}}\fi
 
@@
 \hypertarget{total-number-of-bins}{%
-\subsection{Total number of bins}\label{total-number-of-bins}}
+\subsubsection{Total number of bins}\label{total-number-of-bins}}
 
@@
 \hypertarget{benchmark-value-to-remove-low-density-hexagons}{%
-\subsection{Benchmark value to remove low-density
+\subsubsection{Benchmark value to remove low-density
 hexagons}\label{benchmark-value-to-remove-low-density-hexagons}}
@@
 \hypertarget{benchmark-value-to-remove-long-edges}{%
-\subsection{Benchmark value to remove long
+\subsubsection{Benchmark value to remove long
 edges}\label{benchmark-value-to-remove-long-edges}}
@@
 \hypertarget{sec-summary}{%
-\subsection{Summaries of the model}\label{sec-summary}}
+\subsection{Model summaries}\label{sec-summary}}
 
@@
 accuracy, we introduce an error metric, which measures the sum of
-squared differences between the high-dimensional data ((x\_\{ij\})) and
-the predicted bin centroid data in high-dimensional space
-((C\_\{x\_ij\})) across all bins and dimensions. Mathematically, this
-error is expressed as:
+squared differences between the high-dimensional data (\(x_{ij}\)) and
+the predicted bin centroid data in high-dimensional space (\(C_{x_ij}\))
+across all bins and dimensions. Mathematically, this error is expressed
+as:
 
@@
 
-Here, (n) represents the number of bins, (p) represents the dimensions,
-(x\_\{ij\}) is the actual high-dimensional data, and (C\_\{x\_ij\}) is
-the predicted bin centroid data in high dimensions.
+Here, \(n\) represents the number of bins, \(p\) represents the
+dimensions, \(x_{ij}\) is the actual high-dimensional data, and
+\(C_{x_ij}\) is the predicted bin centroid data in high dimensions.
 
@@
 the best NLDR technique to represent the original data in 2D?}
+
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervisUMAP-1.pdf}
+
+}
+
+\caption{\label{fig-nldervisUMAP}2D layouts from UMAP applied for the
+S-curve data: (a) UMAP (n\_neighbors = 7), (b) UMAP (n\_neighbors = 15),
+(c) UMAP (n\_neighbors = 32), (d) UMAP (n\_neighbors = 50). Is there a
+best parameter choice in representing UMAP or are they all providing
+equivalent information?}
 

### 2024-01-11 scripts/gen_summary_umap_15.R

@@ -1,0 +1,82 @@
+library(readr)
+library(dplyr)
+library(ggplot2)
+library(colorspace)
+
+set.seed(20240110)
+
+source("quollr_code.R", local = TRUE)
+source("nldr_code.R", local = TRUE)
+
+data <- read_rds("data/s_curve/s_curve.rds")
+training_data <- read_rds("data/s_curve/s_curve_training.rds")
+UMAP_data <- read_rds("data/s_curve/s_curve_umap_15.rds")
+
+shape_value_curve <- calculate_effective_shape_value(.data = UMAP_data,
+                                                     x = UMAP1, y = UMAP2)
+
+num_bins_vec <- 1:10 ## Number of bins along the x-axis
+
+vec <- stats::setNames(rep("", 6), c("number_of_bins", "number_of_observations", "total_error", "totol_error_method_2", "totol_error_method_3", "total_mse"))  ## Define column names
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(num_bins_vec)) {
+
+  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = UMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
+  pred_df_training <- pred_df_training_object$pred_data
+  centroid_df_training <- pred_df_training_object$df_bin_centroids
+  avg_df_training <- pred_df_training_object$df_bin
+
+  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+
+  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+
+
+}
+
+
+## Add new column with data types
+
+eval_data_training <- eval_data_training |>
+  mutate(data_type = "training")
+
+MSE_df_2 <- eval_data_training |>
+  dplyr::mutate(method = "UMAP")
+
+
+## To draw with AIC
+ggplot(MSE_df_2 |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
+                                                               y = total_error,
+                                                               color = method
+)) +
[…]

### 2024-01-11 scripts/gen_tour_with_umap_diff_param_15.R

@@ -1,0 +1,150 @@
+library(readr)
+library(dplyr)
+library(ggplot2)
+
+set.seed(20240110)
+
+source("quollr_code.R", local = TRUE)
+source("nldr_code.R", local = TRUE)
+
+## Import data
+df_s_curve <- read_csv("data/s_curve/s_curve.rds")
+
+training_data_s_curve <- read_rds("data/s_curve/s_curve_training.rds")
+
+UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap_15.rds")
+
+plot_UMAP_2D(UMAP_s_curve) + #ggtitle("(b)") +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+num_bins_s_curve <- calculate_effective_x_bins(.data = UMAP_s_curve, x = UMAP1,
+                                               cell_area = 1) ##18
+
+
+shape_value_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
+                                                       x = UMAP1, y = UMAP2) ## 1.259938
+
+## To extract bin centroids
+hexbin_data_object_s_curve <- extract_hexbin_centroids(nldr_df = UMAP_s_curve,
+                                                       num_bins = num_bins_s_curve,
+                                                       shape_val = shape_value_s_curve, x = UMAP1, y = UMAP2)
+
+df_bin_centroids_s_curve <- hexbin_data_object_s_curve$hexdf_data
+
+##########
+
+## Data set with all possible centroids in the hexagonal grid
+
+full_centroid_df <- generate_full_grid_centroids(df_bin_centroids_s_curve)
+
+## Generate all coordinates of hexagons
+hex_grid <- full_hex_grid(full_centroid_df)
+
+hex_full_count_df <- generate_full_grid_info(df_bin_centroids_s_curve)
+
+write_rds(hex_full_count_df, "data/s_curve/s_curve_umap_hex_18.rds")
+
+##########
+
[…]

### 2024-01-11 paper.qmd
Change eval function

@@
 
-vec <- stats::setNames(rep("", 6), c("number_of_bins", "number_of_observations", "total_error", "totol_error_method_2", "totol_error_method_3", "total_mse"))  ## Define column names
+vec <- stats::setNames(rep("", 4), c("number_of_bins", "number_of_observations", "total_error", "total_mse"))  ## Define column names
 

### 2024-01-11 paper.tex
Change eval function

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden, breakable, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, sharp corners, frame hidden, interior hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, breakable]}{\end{tcolorbox}}\fi
 

### 2024-01-12 paper.qmd
Change the wording 31th to 31st, rescaled with coord_cartesian

@@
   scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  xlim(-5, 7) +
-  ylim(-10, 10) +
+  coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
   theme_void() +
@@
   scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  xlim(-5, 8) +
-  ylim(-10, 10) +
+  coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
   theme_void() +
@@
   scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
   theme_void() +
@@
 #| fig-cap: How the 2D model lift into high dimensions?
-(a) visualize the points and the hexagonal bin centroids related 31th and 239th hexagons, (b) visualization of the edge connected the 31th and 239th hexagons (colored in red) in the triangular mesh.
+(a) visualize the points and the hexagonal bin centroids related 31st and 239th hexagons, (b) visualization of the edge connected the 31th and 239th hexagons (colored in red) in the triangular mesh.
 A video of tour animation is available at (need to add the link of lg_high).

### 2024-01-12 paper.tex
Change the wording 31th to 31st, rescaled with coord_cartesian

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, sharp corners, frame hidden, interior hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, breakable]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 \caption{\label{fig-wkhighD}How the 2D model lift into high dimensions?
-(a) visualize the points and the hexagonal bin centroids related 31th
+(a) visualize the points and the hexagonal bin centroids related 31st
 and 239th hexagons, (b) visualization of the edge connected the 31th and

### 2024-01-13 paper.qmd
Corrected the issue with tSNE AIC calculation

@@
 
-tSNE_data_gau <- read_rds("data/five_gau_clusters/tsne_data_five_gau.rds")
+tSNE_data_gau <- read_rds("data/five_gau_clusters/tsne_data_five_gau_61.rds")
 UMAP_data_gau <- read_rds("data/five_gau_clusters/umap_data_five_gau.rds")

### 2024-01-13 paper.tex
Corrected the issue with tSNE AIC calculation

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, frame hidden, interior hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, breakable, sharp corners]}{\end{tcolorbox}}\fi
 

### 2024-01-13 examples/five_gaussian_cluster_data_emb.R

@@ -62,6 +62,10 @@ write_rds(test_data_5, file = "data/five_gau_clusters/data_five_gau_test.rds")
 ### tSNE
 tSNE_data_gau <- Fit_tSNE(training_data_5 |> dplyr::select(-ID), opt_perplexity = calculate_effective_perplexity(training_data_5), with_seed = 20240110)
 
+tSNE_data_gau <- tSNE_data_gau |>
+  select(-ID) |>
+  mutate(ID = training_data_5$ID)
+
 plot_tSNE_2D(tSNE_data_gau)
 
 write_rds(tSNE_data_gau, file = paste0("data/five_gau_clusters/tsne_data_five_gau_", calculate_effective_perplexity(training_data_5),".rds"))

### 2024-01-13 paper.tex
Regenerate AIC plot of 5-Gaussian clust with PHATE

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, frame hidden, interior hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, breakable, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners, interior hidden, enhanced, boxrule=0pt, frame hidden]}{\end{tcolorbox}}\fi
 

### 2024-01-13 paper.qmd
Regenerated Figure 4

@@
 #| echo: false
-# distance |> dplyr::filter(distance == max(distance))
-# df_bin_centroids |> dplyr::filter(row_number() %in% c(10, 65))
-
-a1 <- colour_long_edges(.data = distance, benchmark_value = 15, 
-                        triangular_object = tr1_object, distance_col = distance) + 
-  geom_point(size = 2, colour = "#33a02c") + 
+#| warning: false
+data <- read_rds("data/s_curve/s_curve.rds")
+training_data <- read_rds("data/s_curve/s_curve_training.rds")
+UMAP_data <- read_rds(file = "data/s_curve/s_curve_umap.rds")
+
+num_bins_x <- 2
+
+shape_value <- calculate_effective_shape_value(.data = UMAP_data, 
+                                               x = UMAP1, y = UMAP2)
+
+## To extract bin centroids
+hexbin_data_object <-extract_hexbin_centroids(nldr_df = UMAP_data, num_bins = num_bins_x, shape_val = shape_value)
+
+df_bin_centroids <- hexbin_data_object$hexdf_data
+
+# ## Data set with all possible centroids in the hexagonal grid
+# 
+# full_centroid_df <- generate_full_grid_centroids(df_bin_centroids)
+# 
+# ## Generate all coordinates of hexagons
+# hex_grid <- full_hex_grid(full_centroid_df)
+# 
+# hex_full_count_df <- generate_full_grid_info(df_bin_centroids)
+# 
+# write_rds(hex_full_count_df, "data/s_curve/s_curve_hex_2.rds")
+
+min_std_cell_threshold <- 0.05
+
+df_bin_centroids_all <- hexbin_data_object$hexdf_data ## All the centroids without removing low-density hexagons
+
+hex_full_count_df <- read_rds("data/s_curve/s_curve_hex_2.rds")
+
+df_bin_centroids <- df_bin_centroids |>
+  dplyr::mutate(stand_cell_count = counts/max(counts)) |>
+  dplyr::filter(stand_cell_count > min_std_cell_threshold)
+
+UMAP_data_with_hb_id <- UMAP_data |> 
+  dplyr::mutate(hb_id = hexbin_data_object$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all <- dplyr::bind_cols(training_data |> dplyr::select(-ID), UMAP_data_with_hb_id)
+
+## Averaged on high-D
+df_bin <- avg_highD_data(.data = df_all)
+
+## Triangulate bin centroids
+tr1_object <- triangulate_bin_centroids(df_bin_centroids, x, y)
+tr_from_to_df <- generate_edge_info(triangular_object = tr1_object)
+
+# Create the tibble with x and y coordinates
+tr_df <- tibble::tibble(x = tr1_object$x, y = tr1_object$y)
+
+## Compute 2D distances
+distance <- cal_2D_dist(.data = tr_from_to_df)
+
+benchmark_value <- 19.4
+
+# Filter and label small and long edges
+distance_df_small_edges <- distance |>
+  dplyr::filter(distance < benchmark_value) |>
+  dplyr::mutate(type = "small_edges")
+
+distance_df_long_edges <- distance |>
+  dplyr::filter(distance >= benchmark_value) |>
+  dplyr::mutate(type = "long_edges")
+
+# Combine small and long edges
+distance_edges <- dplyr::bind_rows(distance_df_small_edges, distance_df_long_edges)
+
+tr_from_to_df_coord_with_group <- merge(tr_from_to_df, distance_edges, by = c("from", "to"))
+
+UMAP_data_with_hb_id_selected <- UMAP_data_with_hb_id |> dplyr::filter(hb_id %in% c(2, 15)) |>
+  dplyr::mutate(select_point = if_else(hb_id == 2, "data1", "data2"))
+
+a1 <- ggplot(tr_df |> dplyr::filter((y == min(tr_df$y)) | (y == max(tr_df$y))), aes(x = x, y = y)) +
+  geom_segment(
+    aes(x = x_from, y = y_from, xend = x_to, yend = y_to, color = type),
+    data = tr_from_to_df_coord_with_group
+  ) +
+  geom_point() +
+  coord_equal() +
+  scale_colour_manual(values = c("#000000", "#f6e8c3"))  + 
+  geom_point(size = 3, colour = "#33a02c") + 
   #ggtitle("(b)") + 
@@
 
-UMAP_data_with_hb_id_selected <- UMAP_data_with_hb_id |> dplyr::filter(hb_id %in% c(31, 239)) |>
-  dplyr::mutate(select_point = if_else(hb_id == 31, "data1", "data2"))
-
 a2 <- ggplot(data = hex_full_count_df, aes(x = x, y = y)) + 
-  geom_polygon(color = "black", aes(group = polygon_id), fill = "#ffffff") +
+  geom_polygon(color = "#e7e1ef", aes(group = polygon_id), fill = "#ffffff") +
   geom_point(data = UMAP_data_with_hb_id_selected, aes(x = UMAP1, y = UMAP2, colour = select_point), alpha = 0.5) +
   scale_color_manual(values=c("#e6550d", "#000000")) +
-  geom_point(data = df_bin_centroids_all |> dplyr::filter(hexID %in% c(31, 239)), aes(x = x, y = y), colour = "#33a02c") +
+  geom_point(data = df_bin_centroids_all |> dplyr::filter(hexID %in% c(2, 15)), aes(x = x, y = y), colour = "#33a02c", size = 3) +
   scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
@@
 df1 <- df_all |> 
-  dplyr::filter(hb_id == "31") |>
+  dplyr::filter(hb_id == "2") |>
   dplyr::select(tidyselect::starts_with("x")) |>
@@
 df2 <- df_all |> 
-  dplyr::filter(hb_id == "239") |>
+  dplyr::filter(hb_id == "15") |>
   dplyr::select(tidyselect::starts_with("x")) |>
@@
 df3 <- df_all |> 
-  dplyr::filter(!(hb_id %in% c("31", "239"))) |>
+  dplyr::filter(!(hb_id %in% c("2", "15"))) |>
   dplyr::select(tidyselect::starts_with("x")) |>
@@
   dplyr::filter(hb_id %in% df_bin_centroids$hexID) |>
+  dplyr::filter(hb_id %in% c("2", "15")) |>
   dplyr::select(-hb_id) |>
@@
 
-distance_df_small_edges <- distance |> 
-  dplyr::filter(from == 10) |>
-  dplyr::filter(to == 65) |>
-  head(1)
-
-lg_high <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)], group = df_exe$type, pointSize = 3, levelColors = c("#6a3d9a", "#e6550d", "#000000", "#33a02c"), lineFrom = distance_df_small_edges$from, lineTo = distance_df_small_edges$to, lineColors = "#e31a1c")
+lg_high <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)], group = df_exe$type, pointSize = 3, levelColors = c("#6a3d9a", "#e6550d", "#000000", "#33a02c"), lineFrom = 1, lineTo = 2, lineColors = "#e31a1c")
 
@@
 #| fig-cap: How the 2D model lift into high dimensions?
-(a) visualize the points and the hexagonal bin centroids related 31st and 239th hexagons, (b) visualization of the edge connected the 31th and 239th hexagons (colored in red) in the triangular mesh.
-A video of tour animation is available at (need to add the link of lg_high).
+(a) visualize the points and the hexagonal bin centroids related 2nd and 15th hexagons, (b) visualization of the edge connected the 2nd and 15th hexagons (colored in red) in the triangular mesh.
+A video of tour animation is available at <https://www.youtube.com/watch?v=vBvM30Plt24>.
 

### 2024-01-13 paper.tex
Regenerated Figure 4

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners, interior hidden, enhanced, boxrule=0pt, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, breakable, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, enhanced, frame hidden, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 \caption{\label{fig-wkhighD}How the 2D model lift into high dimensions?
-(a) visualize the points and the hexagonal bin centroids related 31st
-and 239th hexagons, (b) visualization of the edge connected the 31th and
-239th hexagons (colored in red) in the triangular mesh. A video of tour
-animation is available at (need to add the link of lg\_high).}
+(a) visualize the points and the hexagonal bin centroids related 2nd and
+15th hexagons, (b) visualization of the edge connected the 2nd and 15th
+hexagons (colored in red) in the triangular mesh. A video of tour
+animation is available at
+\url{https://www.youtube.com/watch?v=vBvM30Plt24}.}
 

### 2024-01-13 paper.qmd
Add a figure with how summary works when we have different  benchmark values to remove hexagons with S-curve Conclusion: By comparing MSE values, only phate need to rm lwd hexgons, value 0.24

@@
 
-num_bins_vec <- 1:10 ## Number of bins along the x-axis
+num_bins_vec <- 1:20 ## Number of bins along the x-axis
 
@@
 
-num_bins_vec <- 1:10 ## Number of bins along the x-axis
+num_bins_vec <- 1:14 ## Number of bins along the x-axis
 
@@
 
-num_bins_vec <- 1:10 ## Number of bins along the x-axis
+num_bins_vec <- 1:18 ## Number of bins along the x-axis
 
@@
 
+<!--lwd smoothing criteria with s-curve-->
+
+```{r}
+#| warning: false
+#| echo: false
+
+MSE_df_1_lwd <- read_rds("data/s_curve/s_curve_summary_lwd_tsne.rds")
+MSE_df_2_lwd <- read_rds("data/s_curve/s_curve_summary_lwd_umap.rds")
+MSE_df_3_lwd <- read_rds("data/s_curve/s_curve_summary_lwd_phate.rds")
+MSE_df_4_lwd <- read_rds("data/s_curve/s_curve_summary_lwd_trimap.rds")
+MSE_df_5_lwd <- read_rds("data/s_curve/s_curve_summary_lwd_pacmap.rds")
+
+MSE_df_lwd <- dplyr::bind_rows(MSE_df_1_lwd, MSE_df_2_lwd, MSE_df_3_lwd, MSE_df_4_lwd, MSE_df_5_lwd)
+
+MSE_df_lwd$method <- factor(MSE_df_lwd$method, levels = c("tSNE", "UMAP", "PHATE", "TriMAP", "PaCMAP"))
+
+## To draw with AIC
+aic_plot_lwd <- ggplot(MSE_df_lwd |> dplyr::filter(data_type == "training"), aes(x = benchmark_rm_hex,
+                                                                                 y = total_error,
+                                                                                 color = method
+)) +
+  geom_point() +
+  geom_line() +
+  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  theme_light() +
+  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  scale_colour_manual(values = c("#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("AIC") +
+  xlab("Benchmark value")
+## Effective number of bins along x-axis
+
+mse_plot_lwd <- ggplot(MSE_df_lwd, aes(x = benchmark_rm_hex,
+                                       y = total_mse,
+                                       color = method
+)) +
+  geom_point() +
+  geom_line() +
+  theme_light() +
+  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  scale_colour_manual(values = c("#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("MSE") +
+  xlab("Benchmark value")
+
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| fig-cap: Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset with different benchmark values to remove the low-density heaxgons.
+What is the effective benchark valuw to remove the low-density heaxgons?
+#| label: fig-diagnosticpltScurvelwd
+#| out-width: 100%
+
+aic_plot_lwd + mse_plot_lwd +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='bottom', plot.tag = element_text(size = 8))
+```
+
 ### Simulated data example {#sec-simpleex}

### 2024-01-13 paper.tex
Add a figure with how summary works when we have different  benchmark values to remove hexagons with S-curve Conclusion: By comparing MSE values, only phate need to rm lwd hexgons, value 0.24

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, breakable, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, enhanced, frame hidden, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, interior hidden, frame hidden, boxrule=0pt, enhanced, sharp corners, breakable]}{\end{tcolorbox}}\fi
 
@@
 
+\begin{figure}
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticpltScurvelwd-1.pdf}
+
+}
+
+\caption{\label{fig-diagnosticpltScurvelwd}Goodness of fit statistics
+from different NLDR techniques applied to training S-curve dataset with
+different benchmark values to remove the low-density heaxgons. What is
+the effective benchark valuw to remove the low-density heaxgons?}
+
+\end{figure}
+
 \hypertarget{sec-simpleex}{%

### 2024-01-13 quollr_code.R

@@ -548,33 +548,43 @@ compute_aic <- function(p, total, num_bins, num_obs) {
 
 ## df_bin_centroids_all: all bin centroids without removing any
 ## df_bin_centroids: bin centroids after removing removing any
-generate_eval_df <- function(data, prediction_df, df_bin_centroids, df_bin, num_bins, col_start = "x") {
+generate_eval_df <- function(data, prediction_df, df_bin_centroids, df_bin, num_bins, col_start = "x", rm_lwd_hex = FALSE, df_bin_centroids_all = NA) {
 
-  ## Generate all possible bin centroids in the full grid
-  full_centroid_df <- generate_full_grid_centroids(df_bin_centroids)
+  if (rm_lwd_hex == TRUE) {
+
+    ## Generate all possible bin centroids in the full grid
+    full_centroid_df <- generate_full_grid_centroids(df_bin_centroids_all)
+
+  } else {
+    ## Generate all possible bin centroids in the full grid
+    full_centroid_df <- generate_full_grid_centroids(df_bin_centroids)
+
+  }
+
+
 
   df_bin_centroids_filtered <- df_bin_centroids |>
     dplyr::select(hexID, x, y)
 
   ## To map centroid coordinates to predicted hexID
-  prediction_df <- dplyr::inner_join(prediction_df, df_bin_centroids_filtered, by = c("pred_hb_id" = "hexID"))
+  prediction_df_join <- dplyr::inner_join(prediction_df, df_bin_centroids_filtered, by = c("pred_hb_id" = "hexID"))
 
   df_bin_train <- df_bin
   names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])
 
-  prediction_df <- prediction_df |>
+  prediction_df_join <- prediction_df_join |>
     dplyr::left_join(df_bin_train, by = c("pred_hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates
 
-  prediction_df <- prediction_df |>
+  prediction_df_join <- prediction_df_join |>
     dplyr::left_join(data, by = c("ID" = "ID")) ## Map high-D data
 
   for (i in 1:(NCOL(df_bin_train) - 1)) {
 
-    prediction_df[ , paste0("error_square_", col_start, i)] <- (prediction_df[ , paste0(col_start, i)] - prediction_df[ , paste0("avg_", col_start, i)])^2
[…]

### 2024-01-13 paper.tex
Reconstructed the section to lift the model into high dimensions

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, interior hidden, frame hidden, boxrule=0pt, enhanced, sharp corners, breakable]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, boxrule=0pt, sharp corners, interior hidden, breakable, borderline west={3pt}{0pt}{shadecolor}, enhanced]}{\end{tcolorbox}}\fi
 
@@
 Consider the high-dimensional data a rectangular matrix \(X\), where
-\(X = \begin{bmatrix} \textbf{x}_{1} & \textbf{x}_{2} & \cdots & \textbf{x}_{n}\\ \end{bmatrix}\),
+\(X = \begin{bmatrix} \textbf{x}_{1} & \textbf{x}_{2} & \cdots & \textbf{x}_{n}\\ \end{bmatrix}^\top\),
 with \(n\) observations in \(p\) dimensions. The objective is to
 discover a low-dimensional projection
-\(Y = \begin{bmatrix} \textbf{y}_{1} & \textbf{y}_{2} & \cdots & \textbf{y}_{n}\\ \end{bmatrix}\),
+\(Y = \begin{bmatrix} \textbf{y}_{1} & \textbf{y}_{2} & \cdots & \textbf{y}_{n}\\ \end{bmatrix}^\top\),
 represented as an \(n\) × \(d\) matrix, where \(d \ll p\). The reduction
@@
 
-\hypertarget{constructing-the-2d-model}{%
-\subsection{Constructing the 2D model}\label{constructing-the-2d-model}}
+\hypertarget{sec-construct2d}{%
+\subsection{Constructing the 2D model}\label{sec-construct2d}}
 
@@
 binned into hexagon cells. As a result of hexagonal binning , the bin
-centroids (\(C_k^{(2)} \equiv (C_{kx}, C_{ky}\))) (see
-Figure~\ref{fig-meth} (Step 2)) are obtained.
+centroids (\(C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2}\)) (see
+Figure~\ref{fig-meth} Step 2) are obtained.
 
@@
 
-Creating the model in the high-D space involves a two-step process.
-Initially, we identify observations within each hexagonal bin. After
-this, we calculate the average of the high-dimensional data within each
-hex bin. The resulting values represent the averaged high-dimensional
-data points (\(C_k^{(p)} \equiv (C_{kx_1}, ..., C_{kx_p}\))).
-Essentially, these averaged high-dimensional data points serve as the
-representation of the hex bin centroids in the high-D space.
+\hypertarget{lifting-the-triangular-mesh-points-into-high-dimensions}{%
+\subsubsection{Lifting the triangular mesh points into high
+dimensions}\label{lifting-the-triangular-mesh-points-into-high-dimensions}}
+
+\textbf{Step1: Cluster 2D points to hexagons}
+
+Expanding upon the information regarding hexagonal binning discussed in
+Step 1 of Section~\ref{sec-construct2d}, the primary objective in this
+step is to determine the 2D embedding points associated with each
+hexagon. As the initial process of hexagonal binning, the 2D points are
+clustered into their respective hexagonal bins. By mapping this
+information with the hexagonal bin centroids
+(\(C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2})\)) that obtained in Step 2 of
+the 2D model building (see Section~\ref{sec-construct2d}), we can find
+which 2D points are assigned to which data set. (see
+Figure~\ref{fig-wkhighD} (a)).
+
+\textbf{Step2: Cluster high dimensional points to hexagons}
+
+Following the step 1, the main focus in this step to determine the
+corresponding high dimensional points for each hexagon. Every 2D
+embedding point serves as a projection of a data point belonging to high
+dimensional space. By using this mapping between the high dimensional
+data and their corresponding projections, the high dimensional points
+allocated to each hexagons are determined (see video linked in
+Figure~\ref{fig-wkhighD}).
+
+\textbf{Step3: Compute the mean within hexagon}
+
+Having identified the high-dimensional points associated with hexagons,
+the final step involves computing the mean within each hexagonal bin.
+This implies calculating the average of the high-dimensional data points
+located within each hex bin. These averaged high-dimensional data
+points, denoted as \(C_k^{(p)} \equiv (C_{kx_1}, ..., C_{kx_p})\), serve
+as the representative coordinates for the hex bin centroids within the
+expansive high-dimensional space (see video linked in
+Figure~\ref{fig-wkhighD}).
+
+\hypertarget{lifting-the-2d-triangular-mesh-into-high-dimensions}{%
+\subsubsection{Lifting the 2D triangular mesh into high
+dimensions}\label{lifting-the-2d-triangular-mesh-into-high-dimensions}}
+
+Based on the insights gained in Step 3 of Section~\ref{sec-construct2d},
+where connected edges in the 2D triangular mesh were identified, this
+step involves lifting these connections into the high dimensions. Having
+the mappings of all 2D triangular mesh points into high dimensions, the
+points connected in 2D are also connected in high dimensional space (see
+video linked in Figure~\ref{fig-wkhighD}).
 

### 2024-01-13 paper.tex
Update paper

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, boxrule=0pt, sharp corners, interior hidden, breakable, borderline west={3pt}{0pt}{shadecolor}, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, sharp corners, breakable, boxrule=0pt, interior hidden, enhanced, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 the 2D model building (see Section~\ref{sec-construct2d}), we can find
-which 2D points are assigned to which data set. (see
+which 2D points are assigned to which data set (see
 Figure~\ref{fig-wkhighD} (a)).

### 2024-01-14 paper.qmd
Rewrite the section "Preprocessing steps"

@@
 
-Consider the high-dimensional data a rectangular matrix $X$, where $X = \begin{bmatrix} \textbf{x}_{1} & \textbf{x}_{2} & \cdots & \textbf{x}_{n}\\ \end{bmatrix}$, with $n$ observations in $p$ dimensions.
-The objective is to discover a low-dimensional projection $Y = \begin{bmatrix} \textbf{y}_{1} & \textbf{y}_{2} & \cdots & \textbf{y}_{n}\\ \end{bmatrix}$, represented as an $n$ × $d$ matrix, where $d \ll p$.
+Consider the high-dimensional data a rectangular matrix $X$, where $X = \begin{bmatrix} \textbf{x}_{1} & \textbf{x}_{2} & \cdots & \textbf{x}_{n}\\ \end{bmatrix}^\top$, with $n$ observations in $p$ dimensions.
+The objective is to discover a low-dimensional projection $Y = \begin{bmatrix} \textbf{y}_{1} & \textbf{y}_{2} & \cdots & \textbf{y}_{n}\\ \end{bmatrix}^\top$, represented as an $n$ × $d$ matrix, where $d \ll p$.
 The reduction process seeks to remove noise from the original data set while retaining essential information.
@@
 
-To tackle the complexities and noise in high-dimensional data, we apply PCA as a pre-processing step [@article67, @article68, @article69].
-This step helps in noise reduction by identifying principal components that represent directions of maximum variance, capturing essential patterns in the data.
-
-### Constructing the 2D model
+In order to reduce the computational complexity associated with performing NLDR techniques to high-D data, and as a initial step of noise reduction of high-D data, PCA [@article67, @article68, @article69] is applied as a preprocessing step.
+Subsequently, the identified principal components, representing directions of maximum variance, are used to construct the model.
+
+### Constructing the 2D model {#sec-construct2d}
 
@@
   After computing hexagonal grid configurations, the 2D emdeddings are binned into hexagon cells.
-As a result of hexagonal binning , the bin centroids ($C_k^{(2)} \equiv (C_{kx}, C_{ky}$)) (see @fig-meth (Step 2)) are obtained.
+As a result of hexagonal binning , the bin centroids ($C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2}$) (see @fig-meth Step 2) are obtained.
 
@@
 
-Creating the model in the high-D space involves a two-step process.
-Initially, we identify observations within each hexagonal bin.
-After this, we calculate the average of the high-dimensional data within each hex bin.
-The resulting values represent the averaged high-dimensional data points ($C_k^{(p)} \equiv (C_{kx_1}, ..., C_{kx_p}$)).
-Essentially, these averaged high-dimensional data points serve as the representation of the hex bin centroids in the high-D space.
+#### Lifting the triangular mesh points into high dimensions
+
+**Step1: Cluster 2D points to hexagons**
+
+Expanding upon the information regarding hexagonal binning discussed in Step 1 of @sec-construct2d, the primary objective in this step is to determine the 2D embedding points associated with each hexagon.
+As the initial process of hexagonal binning, the 2D points are clustered into their respective hexagonal bins.
+By mapping this information with the hexagonal bin centroids ($C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2})$) that obtained in Step 2 of the 2D model building (see @sec-construct2d), we can find which 2D points are assigned to which data set (see @fig-wkhighD (a)).
+
+**Step2: Cluster high dimensional points to hexagons**
+
+Following the step 1, the main focus in this step to determine the corresponding high dimensional points for each hexagon.
+Every 2D embedding point serves as a projection of a data point belonging to high dimensional space.
+By using this mapping between the high dimensional data and their corresponding projections, the high dimensional points allocated to each hexagons are determined (see video linked in @fig-wkhighD).
+
+**Step3: Compute the mean within hexagon**
+
+Having identified the high-dimensional points associated with hexagons, the final step involves computing the mean within each hexagonal bin.
+This implies calculating the average of the high-dimensional data points located within each hex bin.
+These averaged high-dimensional data points, denoted as $C_k^{(p)} \equiv (C_{kx_1}, ..., C_{kx_p})$, serve as the representative coordinates for the hex bin centroids within the expansive high-dimensional space (see video linked in @fig-wkhighD).
+
+#### Lifting the 2D triangular mesh into high dimensions
+
+Based on the insights gained in Step 3 of @sec-construct2d, where connected edges in the 2D triangular mesh were identified, this step involves lifting these connections into the high dimensions.
+Having the mappings of all 2D triangular mesh points into high dimensions, the points connected in 2D are also connected in high dimensional space (see video linked in @fig-wkhighD).
 

### 2024-01-14 paper.tex
Rewrite the section "Preprocessing steps"

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, sharp corners, breakable, boxrule=0pt, interior hidden, enhanced, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, interior hidden, frame hidden, boxrule=0pt, breakable, borderline west={3pt}{0pt}{shadecolor}, enhanced]}{\end{tcolorbox}}\fi
 
@@
 
-To tackle the complexities and noise in high-dimensional data, we apply
-PCA as a pre-processing step \citep[\citet{article68},
-\citet{article69}]{article67}. This step helps in noise reduction by
-identifying principal components that represent directions of maximum
-variance, capturing essential patterns in the data.
+In order to reduce the computational complexity associated with
+performing NLDR techniques to high-D data, and as a initial step of
+noise reduction of high-D data, PCA \citep[\citet{article68},
+\citet{article69}]{article67} is applied as a preprocessing step.
+Subsequently, the identified principal components, representing
+directions of maximum variance, are used to construct the model.
 

### 2024-01-14 paper.qmd
Update construct 2D model section

@@
 
-Hexagonal binning is a data visualization technique that aggregates high-dimensional data into a two-dimensional representation using hexagonal regions called bins [@article66].
-The hexagonal shape is preferred due to its circular-like appearance, providing smoother transitions between neighboring bins and efficient data aggregation.
-This technique is especially useful for handling large data sets as it reduces visual clutter while capturing the underlying data distribution effectively.
-By considering the data distribution, important features are captured, and gaps between bins are minimized, leading to a more accurate representation of the data.
-In our algorithm, hexagonal binning is utilized to create a low-dimensional manifold.
+The hexagonal grid, formed through hexagonal binning [@Carr1987, @article66], serves as a type of bivariate histogram employed to visualize the structure of high-D data.
+Hexagons, being one of only three regular polygons capable of tessellating a plane [@Carr2013], possess both symmetry of nearest neighbors and the maximum number of sides for a regular tessellation of the plane.
+This unique combination makes hexagons more efficient in covering the plane compared to other regular tessellations.
+Additionally, hexagons exhibit lower visual bias when displaying densities, setting them apart from other regular tessellations [@Dan2023].
+In our algorithm, hexagonal binning is used as the initial step of constructing the 2D model and the total number of bins ($b$) is the crucial parameter that defines the granularity of the hexagonal grid.
 
@@
 
-The first step involves to determine the optimal number of bins ($b$) for creating regular hexagons in the hexagonal grid.
-To achieve this, we rely on the relationship between the diameter ($h$) and the area ($A$) of regular hexagons, as given by @eq-equation3.
+First, the number of bins along the x-axis ($b_1$) is computed using the relationship between the diameter ($h$) and the area ($A$) of regular hexagons (see @eq-equation3).
 
@@
 
-To capture the data's structure effectively, we consider the optimal $A = 1$ (see @fig-binsize). Using @eq-equation4, we then calculate $h$ of a regular hexagon.
+To construct regular hexagons, $A = 1$ (see @fig-binsize) use as the default setting. Then, the diameter ($h$) of the regular hexagons is calculated (see @eq-equation4).
 
@@
 
-  Next, we refer the `hexbin` function in R, which provides a relationship between the diameter of the hexagon ($h$) and the number of bins along the x-axis ($b_1$) (@eq-equation6).
+@Carr2013 mentioned about the relationship between the diameter ($h$) of regular hexagons and the height ($y$) of the plotting region.
+According to our algorithm, the height ($y$) of the plotting region is the the range of 2D embedding component 1 ($r_1$) (see @eq-equation6).
 
@@
 
-By utilizing the calculated $h$ and the range of the non-linear projection component 1 ($r_1$), we determine the $b_1$ (@eq-equation5).
-The result is rounded up to the nearest whole number to ensure an appropriate number of bins that capture the variability of the data along the x-axis.
+After rearranging the @eq-equation6 as @eq-equation5, $b_1$ is computed. The $b_1$ value is rounded up to the nearest whole number to have an integer value.
 
@@
 
-**(b) Determine the effective shape parameter** ($s$)
-
-In this step, we determine the optimal shape parameter ($s$) for the hexagonal bins, which significantly influences their shape and arrangement within the grid.
-The shape parameter ($s$) in the hexagonal binning algorithm is defined as the ratio of the height ($y$) to the width ($x$) of the plotting region, referring to the `hexbin` function in R (see @eq-equation1).
-It determines the shape of the plotting regions and plays a vital role in generating an appropriate hexagonal grid for the data visualization.
+**(b) Determine the shape parameter** ($s$)
+
+In this step, we determine the shape parameter ($s$) for the hexagonal bins, which significantly influences their shape and arrangement within the grid.
+The $s$ in the hexagonal binning algorithm is defined as the ratio of the height ($y$) to the width ($x$) of the plotting region as defined in @eq-equation1 [@Carr2013].
 
@@
 
-To calculate the effective shape parameter ($s$) for our algorithm, we consider the ranges of the non-linear projection components.
-Denoting the range of the non-linear projection component 1 as $r_1$ and the range of the non-linear projection component 2 as $r_2$, we find the shape parameter using @eq-equation2.
+The shape parameter ($s$) of our algorithm is calculated as the ratio of the ranges of 2D embedding components, where $r_1$ and $r_2$ represent the ranges of 2D embedding components 1 and component 2, respectively (see @eq-equation2).
 
@@
 $$ {#eq-equation2}
+
+**(c) Determine the number of bins along the y-axis** ($b_2$)
+
+Next, the number of bins along the y-axis is computed based on the number of bins along the x-axis ($b_1$) and the shape parameter ($s$) (see @eq-equation12) [@Carr2013].
+
+$$
+b_2 = 2 * \left(\frac{(b_1 \times s)}{\sqrt(3)} + 1.5001 \right)
+$${#eq-equation12}
+
+**(d) Determine the total number of bins** ($b$)
+
+The total number of bins is determined by multiplying the number of bins along the x-axis ($b_1$) with the number of bins along the y-axis ($b_2$) (see @eq-equation13).
+
+$$
+b = b_1 \times b_2
+$${#eq-equation13}
 
@@
 #| fig-pos: H
-#| fig-cap: "Hexbin plots from different number of bins for the same **s_curve_noise_umap** data: (a) b = (4, 8), s = 1.643542, (b) b = (12, 22), s = 1.643542, and (c) b = (21, 40), s = 1.643542.
-The hexbins are colored based on the density of points, with yellow indicating higher point density and darker colors representing lower point density within each bin.
+#| fig-cap: "Hexbin plots from different number of bins for the **UMAP** projections of **S-curve** data: (a) b = 32 (4, 8), s = 1.643542, (b) b = 264 (12, 22), s = 1.643542, and (c) b = 840 (21, 40), s = 1.643542.
+The hexbins are colored based on the density of points, with darker colors indicating higher point density and yellow color representing lower point density within each bin.
 Does a value of number of bins exist to effectively represent the low-dimensional data?"
@@
 
-  After computing hexagonal grid configurations, the 2D emdeddings are binned into hexagon cells.
-As a result of hexagonal binning , the bin centroids ($C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2}$) (see @fig-meth Step 2) are obtained.
+As a result of hexagonal binning for high-D data, all the high-D data are clustered into hexagons.
+In this step, the bin centroids ($C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2}$) (see @fig-meth Step 2) are obtained [@Carr2013].
 
@@
 
-  In this step, the algorithm proceeds to triangulate the centroids (see @fig-meth (Step 3)) of the hex bins.
+In this step, the algorithm proceeds to triangulate the hexagonal bin centroids (see @fig-meth (Step 3)).
 Triangulation is a fundamental process in computational geometry and computer graphics that involves dividing a set of points in a given space into interconnected triangles [@article30].
-One common algorithm used for triangulation is Delaunay triangulation [@article26], where points are connected in a way that maximizes the minimum angles of the resulting triangles, leading to a more regular and well-conditioned triangulation.
-In our algorithm, triangulation helps us identify geometric relationships and patterns in the data.
+One common algorithm used for triangulation is Delaunay triangulation [@article26, @article54], where points are connected in a way that maximizes the minimum angles of the resulting triangles, leading to a more regular and well-conditioned triangulation.
 
@@
 
-The total number of bins is a crucial parameter that defines the granularity of our hexagonal grid.
-This parameter particularly influential in capturing the intricacies of the data structure.
-It is determined by multiplying the number of bins along the x-axis (\(b_1\)) with the number of bins along the y-axis (\(b_2\)), as per the formula:
-
-$$
-b = b_1 \times b_2
-$${#eq-equationp1}
-
-  Here, \(b\) represents the total number of bins.
-Adjusting the parameter \(b_1\) provides control over the total bin count (\(b\)), allowing us to fine-tune the grid's configuration along the x-axis based on specific requirements.
+Adjusting the parameter $b_1$ provides control over the total number of bins $b$, allowing us to fine-tune the hexagonal grid.
 

### 2024-01-14 paper.tex
Update construct 2D model section

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, interior hidden, frame hidden, boxrule=0pt, breakable, borderline west={3pt}{0pt}{shadecolor}, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, breakable, boxrule=0pt, interior hidden, enhanced]}{\end{tcolorbox}}\fi
 
@@
 
-Hexagonal binning is a data visualization technique that aggregates
-high-dimensional data into a two-dimensional representation using
-hexagonal regions called bins \citep{article66}. The hexagonal shape is
-preferred due to its circular-like appearance, providing smoother
-transitions between neighboring bins and efficient data aggregation.
-This technique is especially useful for handling large data sets as it
-reduces visual clutter while capturing the underlying data distribution
-effectively. By considering the data distribution, important features
-are captured, and gaps between bins are minimized, leading to a more
-accurate representation of the data. In our algorithm, hexagonal binning
-is utilized to create a low-dimensional manifold.
+The hexagonal grid, formed through hexagonal binning
+\citep[\citet{article66}]{Carr1987}, serves as a type of bivariate
+histogram employed to visualize the structure of high-D data. Hexagons,
+being one of only three regular polygons capable of tessellating a plane
+\citep{Carr2013}, possess both symmetry of nearest neighbors and the
+maximum number of sides for a regular tessellation of the plane. This
+unique combination makes hexagons more efficient in covering the plane
+compared to other regular tessellations. Additionally, hexagons exhibit
+lower visual bias when displaying densities, setting them apart from
+other regular tessellations \citep{Dan2023}. In our algorithm, hexagonal
+binning is used as the initial step of constructing the 2D model and the
+total number of bins (\(b\)) is the crucial parameter that defines the
+granularity of the hexagonal grid.
 
@@
 
-The first step involves to determine the optimal number of bins (\(b\))
-for creating regular hexagons in the hexagonal grid. To achieve this, we
-rely on the relationship between the diameter (\(h\)) and the area
-(\(A\)) of regular hexagons, as given by Equation~\ref{eq-equation3}.
+First, the number of bins along the x-axis (\(b_1\)) is computed using
+the relationship between the diameter (\(h\)) and the area (\(A\)) of
+regular hexagons (see Equation~\ref{eq-equation3}).
 
@@
 
-To capture the data's structure effectively, we consider the optimal
-\(A = 1\) (see Figure~\ref{fig-binsize}). Using
-Equation~\ref{eq-equation4}, we then calculate \(h\) of a regular
-hexagon.
+To construct regular hexagons, \(A = 1\) (see Figure~\ref{fig-binsize})
+use as the default setting. Then, the diameter (\(h\)) of the regular
+hexagons is calculated (see Equation~\ref{eq-equation4}).
 
@@
 
-Next, we refer the \texttt{hexbin} function in R, which provides a
-relationship between the diameter of the hexagon (\(h\)) and the number
-of bins along the x-axis (\(b_1\)) (Equation~\ref{eq-equation6}).
+\citet{Carr2013} mentioned about the relationship between the diameter
+(\(h\)) of regular hexagons and the height (\(y\)) of the plotting
+region. According to our algorithm, the height (\(y\)) of the plotting
+region is the the range of 2D embedding component 1 (\(r_1\)) (see
+Equation~\ref{eq-equation6}).
 
@@
 
-By utilizing the calculated \(h\) and the range of the non-linear
-projection component 1 (\(r_1\)), we determine the \(b_1\)
-(Equation~\ref{eq-equation5}). The result is rounded up to the nearest
-whole number to ensure an appropriate number of bins that capture the
-variability of the data along the x-axis.
+After rearranging the Equation~\ref{eq-equation6} as
+Equation~\ref{eq-equation5}, \(b_1\) is computed. The \(b_1\) value is
+rounded up to the nearest whole number to have an integer value.
 
@@
 
-\textbf{(b) Determine the effective shape parameter} (\(s\))
-
-In this step, we determine the optimal shape parameter (\(s\)) for the
-hexagonal bins, which significantly influences their shape and
-arrangement within the grid. The shape parameter (\(s\)) in the
-hexagonal binning algorithm is defined as the ratio of the height
-(\(y\)) to the width (\(x\)) of the plotting region, referring to the
-\texttt{hexbin} function in R (see Equation~\ref{eq-equation1}). It
-determines the shape of the plotting regions and plays a vital role in
-generating an appropriate hexagonal grid for the data visualization.
+\textbf{(b) Determine the shape parameter} (\(s\))
+
+In this step, we determine the shape parameter (\(s\)) for the hexagonal
+bins, which significantly influences their shape and arrangement within
+the grid. The \(s\) in the hexagonal binning algorithm is defined as the
+ratio of the height (\(y\)) to the width (\(x\)) of the plotting region
+as defined in Equation~\ref{eq-equation1}.
 
@@
 
-To calculate the effective shape parameter (\(s\)) for our algorithm, we
-consider the ranges of the non-linear projection components. Denoting
-the range of the non-linear projection component 1 as \(r_1\) and the
-range of the non-linear projection component 2 as \(r_2\), we find the
-shape parameter using Equation~\ref{eq-equation2}.
+The shape parameter (\(s\)) of our algorithm is calculated as the ratio
+of the ranges of 2D embedding components, where \(r_1\) and \(r_2\)
+represent the ranges of 2D embedding components 1 and component 2,
+respectively (see Equation~\ref{eq-equation2}).
 
@@
 
+\textbf{(c) Determine the number of bins along the y-axis} (\(b_2\))
+
+Next, the number of bins along the y-axis is computed based on the
+number of bins along the x-axis (\(b_1\)) and the shape parameter
+(\(s\)) (see Equation~\ref{eq-equation12}) \citep{Carr2013}.
+
+\begin{equation}\protect\hypertarget{eq-equation12}{}{
+b_2 = 2 * \left(\frac{(b_1 \times s)}{\sqrt(3)} + 1.5001 \right)
+}\label{eq-equation12}\end{equation}
+
+\textbf{(d) Determine the total number of bins} (\(b\))
+
+The total number of bins is determined by multiplying the number of bins
+along the x-axis (\(b_1\)) with the number of bins along the y-axis
+(\(b_2\)) (see Equation~\ref{eq-equation13}).
+
+\begin{equation}\protect\hypertarget{eq-equation13}{}{
+b = b_1 \times b_2
+}\label{eq-equation13}\end{equation}
+
 \begin{figure}[H]
@@
 \caption{\label{fig-binsize}Hexbin plots from different number of bins
-for the same \textbf{s\_curve\_noise\_umap} data: (a) b = (4, 8), s =
-1.643542, (b) b = (12, 22), s = 1.643542, and (c) b = (21, 40), s =
-1.643542. The hexbins are colored based on the density of points, with
-yellow indicating higher point density and darker colors representing
-lower point density within each bin. Does a value of number of bins
-exist to effectively represent the low-dimensional data?}
+for the \textbf{UMAP} projections of \textbf{S-curve} data: (a) b = 32
+(4, 8), s = 1.643542, (b) b = 264 (12, 22), s = 1.643542, and (c) b =
+840 (21, 40), s = 1.643542. The hexbins are colored based on the density
+of points, with darker colors indicating higher point density and yellow
+color representing lower point density within each bin. Does a value of
+number of bins exist to effectively represent the low-dimensional data?}
 
@@
 
-After computing hexagonal grid configurations, the 2D emdeddings are
-binned into hexagon cells. As a result of hexagonal binning , the bin
-centroids (\(C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2}\)) (see
-Figure~\ref{fig-meth} Step 2) are obtained.
+As a result of hexagonal binning for high-D data, all the high-D data
+are clustered into hexagons. In this step, the bin centroids
+(\(C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2}\)) (see Figure~\ref{fig-meth}
+Step 2) are obtained \citep{Carr2013}.
 
@@
 
-In this step, the algorithm proceeds to triangulate the centroids (see
-Figure~\ref{fig-meth} (Step 3)) of the hex bins. Triangulation is a
+In this step, the algorithm proceeds to triangulate the hexagonal bin
+centroids (see Figure~\ref{fig-meth} (Step 3)). Triangulation is a
 fundamental process in computational geometry and computer graphics that
@@
 triangles \citep{article30}. One common algorithm used for triangulation
-is Delaunay triangulation \citep{article26}, where points are connected
-in a way that maximizes the minimum angles of the resulting triangles,
-leading to a more regular and well-conditioned triangulation. In our
-algorithm, triangulation helps us identify geometric relationships and
-patterns in the data.
+is Delaunay triangulation \citep[\citet{article54}]{article26}, where
+points are connected in a way that maximizes the minimum angles of the
+resulting triangles, leading to a more regular and well-conditioned
+triangulation.
 
@@
 
-The total number of bins is a crucial parameter that defines the
-granularity of our hexagonal grid. This parameter particularly
-influential in capturing the intricacies of the data structure. It is
-determined by multiplying the number of bins along the x-axis ((b\_1))
-with the number of bins along the y-axis ((b\_2)), as per the formula:
-
-\begin{equation}\protect\hypertarget{eq-equationp1}{}{
-b = b_1 \times b_2
-}\label{eq-equationp1}\end{equation}
-
-Here, (b) represents the total number of bins. Adjusting the parameter
-(b\_1) provides control over the total bin count ((b)), allowing us to
-fine-tune the grid's configuration along the x-axis based on specific
-requirements.
+Adjusting the parameter \(b_1\) provides control over the total number
+of bins \(b\), allowing us to fine-tune the hexagonal grid.
 

### 2024-01-14 paper.qmd
Add the prediction section where 2D embeddings are predicted.

@@
         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'e', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+  annotate(geom = 'text', label = 'e', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
 
@@
 
+<!--model for S-curve with 5 methods-->
+
+```{r}
+#| warning: false
+#| echo: false
+
+## Import data
+
+tSNE_s_curve <- read_rds("data/s_curve/s_curve_tsne_27.rds")
+UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap.rds")
+PHATE_s_curve <- read_rds("data/s_curve/s_curve_phate.rds")
+TriMAP_s_curve <- read_rds("data/s_curve/s_curve_trimap.rds")
+PaCMAP_s_curve <- read_rds("data/s_curve/s_curve_pacmap.rds")
+
+## tSNE
+
+num_bins_tsne_s_curve <- 8
+shape_val_tsne_s_curve <- calculate_effective_shape_value(.data = tSNE_s_curve,
+                                                          x = tSNE1, y = tSNE2) ## 1.259938
+## To extract bin centroids
+hexbin_data_object_tsne_s_curve <- extract_hexbin_centroids(nldr_df = tSNE_s_curve, num_bins = num_bins_tsne_s_curve, shape_val = shape_val_tsne_s_curve, x = tSNE1, y = tSNE2)
+
+df_bin_centroids_tsne_s_curve <- hexbin_data_object_tsne_s_curve$hexdf_data
+
+tSNE_data_with_hb_id_s_curve <- tSNE_s_curve |>
+  dplyr::mutate(hb_id = hexbin_data_object_tsne_s_curve$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_tsne_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), tSNE_data_with_hb_id_s_curve)
+
+## Averaged on high-D
+df_bin_tsne_s_curve <- avg_highD_data(.data = df_all_tsne_s_curve)
+
+## Triangulate bin centroids
+tr1_object_tsne_s_curve <- triangulate_bin_centroids(df_bin_centroids_tsne_s_curve, x, y)
+tr_from_to_df_tsne_s_curve <- generate_edge_info(triangular_object = tr1_object_tsne_s_curve)
+
+# ggplot(df_bin_centroids_tsne_s_curve, aes(x = x, y = y)) +
+#   geom_segment(data = tr_from_to_df_tsne_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
+#   geom_point(size = 2, colour = "#33a02c") +
+#   coord_equal()
+
+## Compute 2D distances
+distance_tsne_s_curve <- cal_2D_dist(.data = tr_from_to_df_tsne_s_curve)
+
+## To find the benchmark value
+benchmark_tsne_s_curve <- find_benchmark_value(.data = distance_tsne_s_curve, distance_col = distance)
+
+# colour_long_edges(.data = distance_tsne_s_curve, benchmark_value = benchmark_tsne_s_curve,
+#                   triangular_object = tr1_object_tsne_s_curve, distance_col = distance)
+
+trimesh_removed_tsne_s_curve <- remove_long_edges(.data = distance_tsne_s_curve, benchmark_value = benchmark_tsne_s_curve,
+                                                  triangular_object = tr1_object_tsne_s_curve, distance_col = distance)
+
+trimesh_removed_tsne_s_curve <- trimesh_removed_tsne_s_curve +
+  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  # theme(axis.text = element_text(size = 5),
+  #       axis.title = element_text(size = 7)) +
+  geom_point(colour = "#33a02c", size = 0.5) +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+# tour_tsne_s_curve <- show_langevitour(df_all_tsne_s_curve, df_bin_tsne_s_curve, df_bin_centroids_tsne_s_curve, benchmark_value = benchmark_tsne_s_curve, distance = distance_tsne_s_curve, distance_col = distance)
+
+## UMAP
+
+num_bins_umap_s_curve <- 6
+shape_val_umap_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
+                                                          x = UMAP1, y = UMAP2) ## 1.259938
+## To extract bin centroids
+hexbin_data_object_umap_s_curve <- extract_hexbin_centroids(nldr_df = UMAP_s_curve, num_bins = num_bins_umap_s_curve, shape_val = shape_val_umap_s_curve, x = UMAP1, y = UMAP2)
+
+df_bin_centroids_umap_s_curve <- hexbin_data_object_umap_s_curve$hexdf_data
+
+UMAP_data_with_hb_id_s_curve <- UMAP_s_curve |>
+  dplyr::mutate(hb_id = hexbin_data_object_umap_s_curve$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_umap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), UMAP_data_with_hb_id_s_curve)
+
+## Averaged on high-D
+df_bin_umap_s_curve <- avg_highD_data(.data = df_all_umap_s_curve)
+
+## Triangulate bin centroids
+tr1_object_umap_s_curve <- triangulate_bin_centroids(df_bin_centroids_umap_s_curve, x, y)
+tr_from_to_df_umap_s_curve <- generate_edge_info(triangular_object = tr1_object_umap_s_curve)
+
+# ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
+#   geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
+#   geom_point(size = 2, colour = "#33a02c") +
+#   coord_equal()
+
+## Compute 2D distances
+distance_umap_s_curve <- cal_2D_dist(.data = tr_from_to_df_umap_s_curve)
+
+## To find the benchmark value
+benchmark_umap_s_curve <- find_benchmark_value(.data = distance_umap_s_curve, distance_col = distance)
+
+# colour_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve,
+#                   triangular_object = tr1_object_umap_s_curve, distance_col = distance)
+
+trimesh_removed_umap_s_curve <- remove_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve,
+                                     triangular_object = tr1_object_umap_s_curve, distance_col = distance)
+
+trimesh_removed_umap_s_curve <- trimesh_removed_umap_s_curve +
+  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  # theme(axis.text = element_text(size = 5),
+  #       axis.title = element_text(size = 7)) +
+  geom_point(colour = "#33a02c", size = 0.5) +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+# tour_umap_s_curve <- show_langevitour(df_all_umap_s_curve, df_bin_umap_s_curve, df_bin_centroids_umap_s_curve, benchmark_value = benchmark_umap_s_curve, distance = distance_umap_s_curve, distance_col = distance)
+
+## TriMAP
+
+num_bins_trimap_s_curve <- 6
+shape_val_trimap_s_curve <- calculate_effective_shape_value(.data = TriMAP_s_curve,
+                                                            x = TriMAP1, y = TriMAP2) ## 1.259938
+## To extract bin centroids
+hexbin_data_object_trimap_s_curve <- extract_hexbin_centroids(nldr_df = TriMAP_s_curve, num_bins = num_bins_trimap_s_curve, shape_val = shape_val_trimap_s_curve, x = TriMAP1, y = TriMAP2)
+
+df_bin_centroids_trimap_s_curve <- hexbin_data_object_trimap_s_curve$hexdf_data
+
+TriMAP_data_with_hb_id_s_curve <- TriMAP_s_curve |>
+  dplyr::mutate(hb_id = hexbin_data_object_trimap_s_curve$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_trimap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), TriMAP_data_with_hb_id_s_curve)
+
+## Averaged on high-D
+df_bin_trimap_s_curve <- avg_highD_data(.data = df_all_trimap_s_curve)
+
+## Triangulate bin centroids
+tr1_object_trimap_s_curve <- triangulate_bin_centroids(df_bin_centroids_trimap_s_curve, x, y)
+tr_from_to_df_trimap_s_curve <- generate_edge_info(triangular_object = tr1_object_trimap_s_curve)
+
+# ggplot(df_bin_centroids_trimap_s_curve, aes(x = x, y = y)) +
+#   geom_segment(data = tr_from_to_df_trimap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
+#   geom_point(size = 2, colour = "#33a02c") +
+#   coord_equal()
+
+## Compute 2D distances
+distance_trimap_s_curve <- cal_2D_dist(.data = tr_from_to_df_trimap_s_curve)
+
+## To find the benchmark value
+benchmark_trimap_s_curve <- find_benchmark_value(.data = distance_trimap_s_curve, distance_col = distance)
+
+# colour_long_edges(.data = distance_trimap_s_curve, benchmark_value = benchmark_trimap_s_curve,
+#                   triangular_object = tr1_object_trimap_s_curve, distance_col = distance)
+
+trimesh_removed_trimap_s_curve <- remove_long_edges(.data = distance_trimap_s_curve, benchmark_value = benchmark_trimap_s_curve,
+                                                    triangular_object = tr1_object_trimap_s_curve, distance_col = distance)
+
+trimesh_removed_trimap_s_curve <- trimesh_removed_trimap_s_curve +
+  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  # theme(axis.text = element_text(size = 5),
+  #       axis.title = element_text(size = 7)) +
+  geom_point(colour = "#33a02c", size = 0.25) +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+# tour_trimap_s_curve <- show_langevitour(df_all_trimap_s_curve, df_bin_trimap_s_curve, df_bin_centroids_trimap_s_curve, benchmark_value = benchmark_trimap_s_curve, distance = distance_trimap_s_curve, distance_col = distance)
+
+## PacMAP
+
+num_bins_pacmap_s_curve <- 10
+shape_val_pacmap_s_curve <- calculate_effective_shape_value(.data = PaCMAP_s_curve,
+                                                            x = PaCMAP1, y = PaCMAP2) ## 1.259938
+## To extract bin centroids
+hexbin_data_object_pacmap_s_curve <- extract_hexbin_centroids(nldr_df = PaCMAP_s_curve, num_bins = num_bins_pacmap_s_curve, shape_val = shape_val_pacmap_s_curve, x = PaCMAP1, y = PaCMAP2)
+
+df_bin_centroids_pacmap_s_curve <- hexbin_data_object_pacmap_s_curve$hexdf_data
+
+PaCMAP_data_with_hb_id_s_curve <- PaCMAP_s_curve |>
+  dplyr::mutate(hb_id = hexbin_data_object_pacmap_s_curve$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_pacmap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), PaCMAP_data_with_hb_id_s_curve)
+
+## Averaged on high-D
+df_bin_pacmap_s_curve <- avg_highD_data(.data = df_all_pacmap_s_curve)
+
+## Triangulate bin centroids
+tr1_object_pacmap_s_curve <- triangulate_bin_centroids(df_bin_centroids_pacmap_s_curve, x, y)
+tr_from_to_df_pacmap_s_curve <- generate_edge_info(triangular_object = tr1_object_pacmap_s_curve)
+
+# ggplot(df_bin_centroids_pacmap_s_curve, aes(x = x, y = y)) +
+#   geom_segment(data = tr_from_to_df_pacmap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
+#   geom_point(size = 2, colour = "#33a02c") +
+#   coord_equal()
+
+## Compute 2D distances
+distance_pacmap_s_curve <- cal_2D_dist(.data = tr_from_to_df_pacmap_s_curve)
+
+## To find the benchmark value
+benchmark_pacmap_s_curve <- find_benchmark_value(.data = distance_pacmap_s_curve, distance_col = distance)
+
+# colour_long_edges(.data = distance_pacmap_s_curve, benchmark_value = benchmark_pacmap_s_curve,
+#                   triangular_object = tr1_object_pacmap_s_curve, distance_col = distance)
+
+trimesh_removed_pacmap_s_curve <- remove_long_edges(.data = distance_pacmap_s_curve, benchmark_value = benchmark_pacmap_s_curve,
+                                     triangular_object = tr1_object_pacmap_s_curve, distance_col = distance)
+
+trimesh_removed_pacmap_s_curve <- trimesh_removed_pacmap_s_curve +
+  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  # theme(axis.text = element_text(size = 5),
+  #       axis.title = element_text(size = 7)) +
+  geom_point(colour = "#33a02c", size = 0.5) +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'e', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+# tour_pacmap_s_curve <- show_langevitour(df_all_pacmap_s_curve, df_bin_pacmap_s_curve, df_bin_centroids_pacmap_s_curve, benchmark_value = benchmark_pacmap_s_curve, distance = distance_pacmap_s_curve, distance_col = distance)
+
+## PHATE
+
+num_bins_phate_s_curve <- 8
+shape_val_phate_s_curve <- calculate_effective_shape_value(.data = PHATE_s_curve,
+                                                           x = PHATE1, y = PHATE2) ## 1.259938
+## To extract bin centroids
+hexbin_data_object_phate_s_curve <- extract_hexbin_centroids(nldr_df = PHATE_s_curve, num_bins = num_bins_phate_s_curve, shape_val = shape_val_phate_s_curve, x = PHATE1, y = PHATE2)
+
+df_bin_centroids_phate_s_curve <- hexbin_data_object_phate_s_curve$hexdf_data
+
+## Identify bins with low-density
+identify_rm_bins <- find_low_density_hexagons(df_bin_centroids_phate_s_curve, num_bins_phate_s_curve, benchmark_rm_hex = 0.06)
+
+df_bin_centroids_phate_s_curve <- df_bin_centroids_phate_s_curve |>
+  filter(!(hexID %in% identify_rm_bins))
+
+PHATE_data_with_hb_id_s_curve <- PHATE_s_curve |>
+  dplyr::mutate(hb_id = hexbin_data_object_phate_s_curve$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_phate_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), PHATE_data_with_hb_id_s_curve)
+
+## Averaged on high-D
+df_bin_phate_s_curve <- avg_highD_data(.data = df_all_phate_s_curve)
+
+## Triangulate bin centroids
+tr1_object_phate_s_curve <- triangulate_bin_centroids(df_bin_centroids_phate_s_curve, x, y)
+tr_from_to_df_phate_s_curve <- generate_edge_info(triangular_object = tr1_object_phate_s_curve)
+
+# ggplot(df_bin_centroids_phate_s_curve, aes(x = x, y = y)) +
+#   geom_segment(data = tr_from_to_df_phate_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
+#   geom_point(size = 2, colour = "#33a02c") +
+#   coord_equal()
+
+## Compute 2D distances
+distance_phate_s_curve <- cal_2D_dist(.data = tr_from_to_df_phate_s_curve)
+
+## To find the benchmark value
+benchmark_phate_s_curve <- find_benchmark_value(.data = distance_phate_s_curve, distance_col = distance)
+
+# colour_long_edges(.data = distance_phate_s_curve, benchmark_value = benchmark_phate_s_curve,
+#                   triangular_object = tr1_object_phate_s_curve, distance_col = distance)
+
+trimesh_removed_phate_s_curve <- remove_long_edges(.data = distance_phate_s_curve, benchmark_value = benchmark_phate_s_curve,
+                                                   triangular_object = tr1_object_phate_s_curve, distance_col = distance)
+
+trimesh_removed_phate_s_curve <- trimesh_removed_phate_s_curve +
+  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  # theme(axis.text = element_text(size = 5),
+  #       axis.title = element_text(size = 7)) +
+  geom_point(colour = "#33a02c", size = 0.5) +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+# tour_phate_s_curve <- show_langevitour(df_all_phate_s_curve, df_bin_phate_s_curve, df_bin_centroids_phate_s_curve, benchmark_value = benchmark_phate_s_curve, distance = distance_phate_s_curve, distance_col = distance)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+#| label: fig-modelScurve
+#| fig-cap: Model.
+
+trimesh_removed_tsne_s_curve + trimesh_removed_umap_s_curve + trimesh_removed_phate_s_curve + trimesh_removed_trimap_s_curve + trimesh_removed_pacmap_s_curve +
+  #plot_annotation(tag_levels = 'a') +
+  plot_layout(ncol = 5) &
+  theme(legend.position='none', plot.tag = element_text(size = 8))
+```
+
 ### Tunning the model
@@
 
-In this context, the term "prediction values" refers to the 2D coordinates predicted for the NLDR technique.
 The approach involves employing the K-nearest neighbors (KNN) algorithm to identify the nearest hexagonal bin centroid in the 2D space.
 Subsequently, the coordinates of this centroid are assigned as the low-dimensional predicted values for the test data in 2D space.
-It is noteworthy that traditional NLDR methods, such as t-SNE, often lack a direct predict function, making our approach valuable for generating predicted values in the absence of such functionalities.
+It is noteworthy that traditional NLDR methods, such as tSNE, often lack a direct predict function, making our approach valuable for generating predicted values in the absence of such functionalities.
 
@@
 
+### Prediction {#sec-prediction}
+
+In this context, "prediction" denotes the 2D embedding generated for the NLDR technique.
+The methodology encompasses identifying the nearest averaged high-D points in the high-D space for the test data, by computing high-D Euclidean distances.
+As the averaged high-D point corresponds to the lifting of the 2D model, determining its nearest counterpart allows us to map its hexagonal bin centroid coordinates.
+Consequently, these centroid coordinates serve as the assigned low-dimensional predicted values for the test data in the 2D space.
+
+Some NLDR techniques, such as tSNE, often lack a direct prediction, making our approach valuable for generating predicted values in the absence of such functionalities.
+
 ### Simulated data example {#sec-simpleex}

### 2024-01-14 paper.tex
Add the prediction section where 2D embeddings are predicted.

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, breakable, boxrule=0pt, interior hidden, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, interior hidden, breakable, enhanced, sharp corners, frame hidden, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
 
@@
 algorithm, \textbf{quollr}, and its implementation details. In
-\textbf{?@sec-prediction}, discusses the effectiveness of the learned
+Section~\ref{sec-prediction}, discusses the effectiveness of the learned
 low-dimensional manifold in accurately representing the complex
@@
 
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-modelScurve-1.pdf}
+
+}
+
+\caption{\label{fig-modelScurve}Model.}
+
+\end{figure}
+
 \hypertarget{tunning-the-model}{%
@@
 
-In this context, the term ``prediction values'' refers to the 2D
-coordinates predicted for the NLDR technique. The approach involves
-employing the K-nearest neighbors (KNN) algorithm to identify the
-nearest hexagonal bin centroid in the 2D space. Subsequently, the
-coordinates of this centroid are assigned as the low-dimensional
-predicted values for the test data in 2D space. It is noteworthy that
-traditional NLDR methods, such as t-SNE, often lack a direct predict
-function, making our approach valuable for generating predicted values
-in the absence of such functionalities.
+The approach involves employing the K-nearest neighbors (KNN) algorithm
+to identify the nearest hexagonal bin centroid in the 2D space.
+Subsequently, the coordinates of this centroid are assigned as the
+low-dimensional predicted values for the test data in 2D space. It is
+noteworthy that traditional NLDR methods, such as tSNE, often lack a
+direct predict function, making our approach valuable for generating
+predicted values in the absence of such functionalities.
 
@@
 \end{figure}
+
+\hypertarget{sec-prediction}{%
+\subsection{Prediction}\label{sec-prediction}}
+
+In this context, ``prediction'' denotes the 2D embedding generated for
+the NLDR technique. The methodology encompasses identifying the nearest
+averaged high-D points in the high-D space for the test data, by
+computing high-D Euclidean distances. As the averaged high-D point
+corresponds to the lifting of the 2D model, determining its nearest
+counterpart allows us to map its hexagonal bin centroid coordinates.
+Consequently, these centroid coordinates serve as the assigned
+low-dimensional predicted values for the test data in the 2D space.
+
+Some NLDR techniques, such as tSNE, often lack a direct prediction,
+making our approach valuable for generating predicted values in the
+absence of such functionalities.
 

### 2024-01-14 scripts/gen_tour_with_diff_emb.R

@@ -5,9 +5,9 @@ set.seed(20240110)
 source("quollr_code.R", local = TRUE)
 
 ## Import data
-data <- read_rds("data/five_gau_clusters/data_five_gau.rds")
-training_data <- read_rds("data/five_gau_clusters/data_five_gau_training.rds")
-test <- read_rds("data/five_gau_clusters/data_five_gau_test.rds")
+data <- read_rds("data/s_curve/data_s_curve.rds")
+training_data <- read_rds("data/s_curve/data_s_curve_training.rds")
+test_data <- read_rds("data/s_curve/data_s_curve_test.rds")
 
 tSNE_s_curve <- read_rds("data/s_curve/s_curve_tsne_27.rds")
 UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap.rds")
@@ -39,10 +39,10 @@ df_bin_tsne_s_curve <- avg_highD_data(.data = df_all_tsne_s_curve)
 tr1_object_tsne_s_curve <- triangulate_bin_centroids(df_bin_centroids_tsne_s_curve, x, y)
 tr_from_to_df_tsne_s_curve <- generate_edge_info(triangular_object = tr1_object_tsne_s_curve)
 
-ggplot(df_bin_centroids_tsne_s_curve, aes(x = x, y = y)) +
-  geom_segment(data = tr_from_to_df_tsne_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
-  geom_point(size = 2, colour = "#33a02c") +
-  coord_equal()
+# ggplot(df_bin_centroids_tsne_s_curve, aes(x = x, y = y)) +
+#   geom_segment(data = tr_from_to_df_tsne_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
+#   geom_point(size = 2, colour = "#33a02c") +
+#   coord_equal()
 
 
 ## Compute 2D distances
@@ -51,8 +51,16 @@ distance_tsne_s_curve <- cal_2D_dist(.data = tr_from_to_df_tsne_s_curve)
 ## To find the benchmark value
 benchmark_tsne_s_curve <- find_benchmark_value(.data = distance_tsne_s_curve, distance_col = distance)
 
-colour_long_edges(.data = distance_tsne_s_curve, benchmark_value = benchmark_tsne_s_curve,
-                  triangular_object = tr1_object_tsne_s_curve, distance_col = distance)
[…]

### 2024-01-14 paper.qmd
Add links to the model of S-curve, and What I lear from the models

@@
 
-<!--model for S-curve with 5 methods-->
-
-```{r}
-#| warning: false
-#| echo: false
-
-## Import data
-
-tSNE_s_curve <- read_rds("data/s_curve/s_curve_tsne_27.rds")
-UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap.rds")
-PHATE_s_curve <- read_rds("data/s_curve/s_curve_phate.rds")
-TriMAP_s_curve <- read_rds("data/s_curve/s_curve_trimap.rds")
-PaCMAP_s_curve <- read_rds("data/s_curve/s_curve_pacmap.rds")
-
-## tSNE
-
-num_bins_tsne_s_curve <- 8
-shape_val_tsne_s_curve <- calculate_effective_shape_value(.data = tSNE_s_curve,
-                                                          x = tSNE1, y = tSNE2) ## 1.259938
-## To extract bin centroids
-hexbin_data_object_tsne_s_curve <- extract_hexbin_centroids(nldr_df = tSNE_s_curve, num_bins = num_bins_tsne_s_curve, shape_val = shape_val_tsne_s_curve, x = tSNE1, y = tSNE2)
-
-df_bin_centroids_tsne_s_curve <- hexbin_data_object_tsne_s_curve$hexdf_data
-
-tSNE_data_with_hb_id_s_curve <- tSNE_s_curve |>
-  dplyr::mutate(hb_id = hexbin_data_object_tsne_s_curve$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_tsne_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), tSNE_data_with_hb_id_s_curve)
-
-## Averaged on high-D
-df_bin_tsne_s_curve <- avg_highD_data(.data = df_all_tsne_s_curve)
-
-## Triangulate bin centroids
-tr1_object_tsne_s_curve <- triangulate_bin_centroids(df_bin_centroids_tsne_s_curve, x, y)
-tr_from_to_df_tsne_s_curve <- generate_edge_info(triangular_object = tr1_object_tsne_s_curve)
-
-# ggplot(df_bin_centroids_tsne_s_curve, aes(x = x, y = y)) +
-#   geom_segment(data = tr_from_to_df_tsne_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
-#   geom_point(size = 2, colour = "#33a02c") +
-#   coord_equal()
-
-## Compute 2D distances
-distance_tsne_s_curve <- cal_2D_dist(.data = tr_from_to_df_tsne_s_curve)
-
-## To find the benchmark value
-benchmark_tsne_s_curve <- find_benchmark_value(.data = distance_tsne_s_curve, distance_col = distance)
-
-# colour_long_edges(.data = distance_tsne_s_curve, benchmark_value = benchmark_tsne_s_curve,
-#                   triangular_object = tr1_object_tsne_s_curve, distance_col = distance)
-
-trimesh_removed_tsne_s_curve <- remove_long_edges(.data = distance_tsne_s_curve, benchmark_value = benchmark_tsne_s_curve,
-                                                  triangular_object = tr1_object_tsne_s_curve, distance_col = distance)
-
-trimesh_removed_tsne_s_curve <- trimesh_removed_tsne_s_curve +
-  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  # theme(axis.text = element_text(size = 5),
-  #       axis.title = element_text(size = 7)) +
-  geom_point(colour = "#33a02c", size = 0.5) +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-# tour_tsne_s_curve <- show_langevitour(df_all_tsne_s_curve, df_bin_tsne_s_curve, df_bin_centroids_tsne_s_curve, benchmark_value = benchmark_tsne_s_curve, distance = distance_tsne_s_curve, distance_col = distance)
-
-## UMAP
-
-num_bins_umap_s_curve <- 6
-shape_val_umap_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
-                                                          x = UMAP1, y = UMAP2) ## 1.259938
-## To extract bin centroids
-hexbin_data_object_umap_s_curve <- extract_hexbin_centroids(nldr_df = UMAP_s_curve, num_bins = num_bins_umap_s_curve, shape_val = shape_val_umap_s_curve, x = UMAP1, y = UMAP2)
-
-df_bin_centroids_umap_s_curve <- hexbin_data_object_umap_s_curve$hexdf_data
-
-UMAP_data_with_hb_id_s_curve <- UMAP_s_curve |>
-  dplyr::mutate(hb_id = hexbin_data_object_umap_s_curve$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_umap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), UMAP_data_with_hb_id_s_curve)
-
-## Averaged on high-D
-df_bin_umap_s_curve <- avg_highD_data(.data = df_all_umap_s_curve)
-
-## Triangulate bin centroids
-tr1_object_umap_s_curve <- triangulate_bin_centroids(df_bin_centroids_umap_s_curve, x, y)
-tr_from_to_df_umap_s_curve <- generate_edge_info(triangular_object = tr1_object_umap_s_curve)
-
-# ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
-#   geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
-#   geom_point(size = 2, colour = "#33a02c") +
-#   coord_equal()
-
-## Compute 2D distances
-distance_umap_s_curve <- cal_2D_dist(.data = tr_from_to_df_umap_s_curve)
-
-## To find the benchmark value
-benchmark_umap_s_curve <- find_benchmark_value(.data = distance_umap_s_curve, distance_col = distance)
-
-# colour_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve,
-#                   triangular_object = tr1_object_umap_s_curve, distance_col = distance)
-
-trimesh_removed_umap_s_curve <- remove_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve,
-                                     triangular_object = tr1_object_umap_s_curve, distance_col = distance)
-
-trimesh_removed_umap_s_curve <- trimesh_removed_umap_s_curve +
-  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  # theme(axis.text = element_text(size = 5),
-  #       axis.title = element_text(size = 7)) +
-  geom_point(colour = "#33a02c", size = 0.5) +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-# tour_umap_s_curve <- show_langevitour(df_all_umap_s_curve, df_bin_umap_s_curve, df_bin_centroids_umap_s_curve, benchmark_value = benchmark_umap_s_curve, distance = distance_umap_s_curve, distance_col = distance)
-
-## TriMAP
-
-num_bins_trimap_s_curve <- 6
-shape_val_trimap_s_curve <- calculate_effective_shape_value(.data = TriMAP_s_curve,
-                                                            x = TriMAP1, y = TriMAP2) ## 1.259938
-## To extract bin centroids
-hexbin_data_object_trimap_s_curve <- extract_hexbin_centroids(nldr_df = TriMAP_s_curve, num_bins = num_bins_trimap_s_curve, shape_val = shape_val_trimap_s_curve, x = TriMAP1, y = TriMAP2)
-
-df_bin_centroids_trimap_s_curve <- hexbin_data_object_trimap_s_curve$hexdf_data
-
-TriMAP_data_with_hb_id_s_curve <- TriMAP_s_curve |>
-  dplyr::mutate(hb_id = hexbin_data_object_trimap_s_curve$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_trimap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), TriMAP_data_with_hb_id_s_curve)
-
-## Averaged on high-D
-df_bin_trimap_s_curve <- avg_highD_data(.data = df_all_trimap_s_curve)
-
-## Triangulate bin centroids
-tr1_object_trimap_s_curve <- triangulate_bin_centroids(df_bin_centroids_trimap_s_curve, x, y)
-tr_from_to_df_trimap_s_curve <- generate_edge_info(triangular_object = tr1_object_trimap_s_curve)
-
-# ggplot(df_bin_centroids_trimap_s_curve, aes(x = x, y = y)) +
-#   geom_segment(data = tr_from_to_df_trimap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
-#   geom_point(size = 2, colour = "#33a02c") +
-#   coord_equal()
-
-## Compute 2D distances
-distance_trimap_s_curve <- cal_2D_dist(.data = tr_from_to_df_trimap_s_curve)
-
-## To find the benchmark value
-benchmark_trimap_s_curve <- find_benchmark_value(.data = distance_trimap_s_curve, distance_col = distance)
-
-# colour_long_edges(.data = distance_trimap_s_curve, benchmark_value = benchmark_trimap_s_curve,
-#                   triangular_object = tr1_object_trimap_s_curve, distance_col = distance)
-
-trimesh_removed_trimap_s_curve <- remove_long_edges(.data = distance_trimap_s_curve, benchmark_value = benchmark_trimap_s_curve,
-                                                    triangular_object = tr1_object_trimap_s_curve, distance_col = distance)
-
-trimesh_removed_trimap_s_curve <- trimesh_removed_trimap_s_curve +
-  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  # theme(axis.text = element_text(size = 5),
-  #       axis.title = element_text(size = 7)) +
-  geom_point(colour = "#33a02c", size = 0.25) +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-# tour_trimap_s_curve <- show_langevitour(df_all_trimap_s_curve, df_bin_trimap_s_curve, df_bin_centroids_trimap_s_curve, benchmark_value = benchmark_trimap_s_curve, distance = distance_trimap_s_curve, distance_col = distance)
-
-## PacMAP
-
-num_bins_pacmap_s_curve <- 10
-shape_val_pacmap_s_curve <- calculate_effective_shape_value(.data = PaCMAP_s_curve,
-                                                            x = PaCMAP1, y = PaCMAP2) ## 1.259938
-## To extract bin centroids
-hexbin_data_object_pacmap_s_curve <- extract_hexbin_centroids(nldr_df = PaCMAP_s_curve, num_bins = num_bins_pacmap_s_curve, shape_val = shape_val_pacmap_s_curve, x = PaCMAP1, y = PaCMAP2)
-
-df_bin_centroids_pacmap_s_curve <- hexbin_data_object_pacmap_s_curve$hexdf_data
-
-PaCMAP_data_with_hb_id_s_curve <- PaCMAP_s_curve |>
-  dplyr::mutate(hb_id = hexbin_data_object_pacmap_s_curve$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_pacmap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), PaCMAP_data_with_hb_id_s_curve)
-
-## Averaged on high-D
-df_bin_pacmap_s_curve <- avg_highD_data(.data = df_all_pacmap_s_curve)
-
-## Triangulate bin centroids
-tr1_object_pacmap_s_curve <- triangulate_bin_centroids(df_bin_centroids_pacmap_s_curve, x, y)
-tr_from_to_df_pacmap_s_curve <- generate_edge_info(triangular_object = tr1_object_pacmap_s_curve)
-
-# ggplot(df_bin_centroids_pacmap_s_curve, aes(x = x, y = y)) +
-#   geom_segment(data = tr_from_to_df_pacmap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
-#   geom_point(size = 2, colour = "#33a02c") +
-#   coord_equal()
-
-## Compute 2D distances
-distance_pacmap_s_curve <- cal_2D_dist(.data = tr_from_to_df_pacmap_s_curve)
-
-## To find the benchmark value
-benchmark_pacmap_s_curve <- find_benchmark_value(.data = distance_pacmap_s_curve, distance_col = distance)
-
-# colour_long_edges(.data = distance_pacmap_s_curve, benchmark_value = benchmark_pacmap_s_curve,
-#                   triangular_object = tr1_object_pacmap_s_curve, distance_col = distance)
-
-trimesh_removed_pacmap_s_curve <- remove_long_edges(.data = distance_pacmap_s_curve, benchmark_value = benchmark_pacmap_s_curve,
-                                     triangular_object = tr1_object_pacmap_s_curve, distance_col = distance)
-
-trimesh_removed_pacmap_s_curve <- trimesh_removed_pacmap_s_curve +
-  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  # theme(axis.text = element_text(size = 5),
-  #       axis.title = element_text(size = 7)) +
-  geom_point(colour = "#33a02c", size = 0.5) +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'e', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-# tour_pacmap_s_curve <- show_langevitour(df_all_pacmap_s_curve, df_bin_pacmap_s_curve, df_bin_centroids_pacmap_s_curve, benchmark_value = benchmark_pacmap_s_curve, distance = distance_pacmap_s_curve, distance_col = distance)
-
-## PHATE
-
-num_bins_phate_s_curve <- 8
-shape_val_phate_s_curve <- calculate_effective_shape_value(.data = PHATE_s_curve,
-                                                           x = PHATE1, y = PHATE2) ## 1.259938
-## To extract bin centroids
-hexbin_data_object_phate_s_curve <- extract_hexbin_centroids(nldr_df = PHATE_s_curve, num_bins = num_bins_phate_s_curve, shape_val = shape_val_phate_s_curve, x = PHATE1, y = PHATE2)
-
-df_bin_centroids_phate_s_curve <- hexbin_data_object_phate_s_curve$hexdf_data
-
-## Identify bins with low-density
-identify_rm_bins <- find_low_density_hexagons(df_bin_centroids_phate_s_curve, num_bins_phate_s_curve, benchmark_rm_hex = 0.06)
-
-df_bin_centroids_phate_s_curve <- df_bin_centroids_phate_s_curve |>
-  filter(!(hexID %in% identify_rm_bins))
-
-PHATE_data_with_hb_id_s_curve <- PHATE_s_curve |>
-  dplyr::mutate(hb_id = hexbin_data_object_phate_s_curve$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_phate_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), PHATE_data_with_hb_id_s_curve)
-
-## Averaged on high-D
-df_bin_phate_s_curve <- avg_highD_data(.data = df_all_phate_s_curve)
-
-## Triangulate bin centroids
-tr1_object_phate_s_curve <- triangulate_bin_centroids(df_bin_centroids_phate_s_curve, x, y)
-tr_from_to_df_phate_s_curve <- generate_edge_info(triangular_object = tr1_object_phate_s_curve)
-
-# ggplot(df_bin_centroids_phate_s_curve, aes(x = x, y = y)) +
-#   geom_segment(data = tr_from_to_df_phate_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
-#   geom_point(size = 2, colour = "#33a02c") +
-#   coord_equal()
-
-## Compute 2D distances
-distance_phate_s_curve <- cal_2D_dist(.data = tr_from_to_df_phate_s_curve)
-
-## To find the benchmark value
-benchmark_phate_s_curve <- find_benchmark_value(.data = distance_phate_s_curve, distance_col = distance)
-
-# colour_long_edges(.data = distance_phate_s_curve, benchmark_value = benchmark_phate_s_curve,
-#                   triangular_object = tr1_object_phate_s_curve, distance_col = distance)
-
-trimesh_removed_phate_s_curve <- remove_long_edges(.data = distance_phate_s_curve, benchmark_value = benchmark_phate_s_curve,
-                                                   triangular_object = tr1_object_phate_s_curve, distance_col = distance)
-
-trimesh_removed_phate_s_curve <- trimesh_removed_phate_s_curve +
-  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  # theme(axis.text = element_text(size = 5),
-  #       axis.title = element_text(size = 7)) +
-  geom_point(colour = "#33a02c", size = 0.5) +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-# tour_phate_s_curve <- show_langevitour(df_all_phate_s_curve, df_bin_phate_s_curve, df_bin_centroids_phate_s_curve, benchmark_value = benchmark_phate_s_curve, distance = distance_phate_s_curve, distance_col = distance)
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-#| label: fig-modelScurve
-#| fig-cap: Model.
-
-trimesh_removed_tsne_s_curve + trimesh_removed_umap_s_curve + trimesh_removed_phate_s_curve + trimesh_removed_trimap_s_curve + trimesh_removed_pacmap_s_curve +
-  #plot_annotation(tag_levels = 'a') +
-  plot_layout(ncol = 5) &
-  theme(legend.position='none', plot.tag = element_text(size = 8))
-```
-
 ### Tunning the model
@@
 $$
-\text{MSE} = \sum_{j = 1}^{n} \frac{\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2}{\text{total number of bins}}
+\text{MSE} = \sum_{j = 1}^{n} \frac{\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2}{b}
 $$ {#eq-equation9}
@@
 
+### Model interpretation
+
+The visualizations of NLDR techniques for the S-curve dataset shows different representations in 2D, as showcased in Figure @fig-nldervis.
+The question at hand is determining the best NLDR technique to represent the S-curve data.
+Examining the 2D layouts in @fig-nldervis suggests that TriMAP or PaCMAP may be best representations.
+However, the challenge is to decide the best representation with considerable evidence.
+
+After considering the models constructed from the algorithm for each NLDR techniques and visualizing them in high-D space with the original data, shows that all the NLDR techniques capture the non-linearity of the S-curve (see videos of the models linked in @fig-modelScurve).
+But thorough examination of each model in high-D space shows some surprising findings.
+The model of PHATE is flattened and lack of capturing the width of S-curve (see videos of the model with PHATE linked in @fig-modelScurve).
+This nature of PHATE is also evicted by the considerably higher AIC and MSE values for PHATE as shown in @fig-diagnosticpltScurve.
+Furthermore, there are two edges that are not connected to any points in the model of TriMAP (see see videos of the model with TriMAP linked in @fig-modelScurve).
+This model shows how the patterns in 2D representation of TriMAP capture by the model (see @fig-nldervis (d)).
+As shown in @fig-nldervis (b), UMAP have four clusters in a curvilinear structure.
+However, the model reveals that even though UMAP shows a curvilinear in 2D, UMAP actually capture the structure of the S-curve without any disjoint in the model (see see videos of the model with UMAP linked in @fig-modelScurve).
+In the 2D layouts of tSNE and PaCMAP (see @fig-nldervis (a), (e)), there are some sparse areas that are captured by the models effectively (see see videos of the model with tSNE and PaCMAP linked in @fig-modelScurve).
+
+Different NLDR techniques perform differently on the S-curve data and tSNE provides the best representation in 2D according to AIC and MSE values (see @fig-diagnosticpltScurve).
+
+<!--model for S-curve with 5 methods-->
+
+```{r}
+#| warning: false
+#| echo: false
+
+## Import data
+
+tSNE_s_curve <- read_rds("data/s_curve/s_curve_tsne_27.rds")
+UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap.rds")
+PHATE_s_curve <- read_rds("data/s_curve/s_curve_phate.rds")
+TriMAP_s_curve <- read_rds("data/s_curve/s_curve_trimap.rds")
+PaCMAP_s_curve <- read_rds("data/s_curve/s_curve_pacmap.rds")
+
+## tSNE
+
+num_bins_tsne_s_curve <- 8
+shape_val_tsne_s_curve <- calculate_effective_shape_value(.data = tSNE_s_curve,
+                                                          x = tSNE1, y = tSNE2) ## 1.259938
+## To extract bin centroids
+hexbin_data_object_tsne_s_curve <- extract_hexbin_centroids(nldr_df = tSNE_s_curve, num_bins = num_bins_tsne_s_curve, shape_val = shape_val_tsne_s_curve, x = tSNE1, y = tSNE2)
+
+df_bin_centroids_tsne_s_curve <- hexbin_data_object_tsne_s_curve$hexdf_data
+
+tSNE_data_with_hb_id_s_curve <- tSNE_s_curve |>
+  dplyr::mutate(hb_id = hexbin_data_object_tsne_s_curve$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_tsne_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), tSNE_data_with_hb_id_s_curve)
+
+## Averaged on high-D
+df_bin_tsne_s_curve <- avg_highD_data(.data = df_all_tsne_s_curve)
+
+## Triangulate bin centroids
+tr1_object_tsne_s_curve <- triangulate_bin_centroids(df_bin_centroids_tsne_s_curve, x, y)
+tr_from_to_df_tsne_s_curve <- generate_edge_info(triangular_object = tr1_object_tsne_s_curve)
+
+# ggplot(df_bin_centroids_tsne_s_curve, aes(x = x, y = y)) +
+#   geom_segment(data = tr_from_to_df_tsne_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
+#   geom_point(size = 2, colour = "#33a02c") +
+#   coord_equal()
+
+## Compute 2D distances
+distance_tsne_s_curve <- cal_2D_dist(.data = tr_from_to_df_tsne_s_curve)
+
+## To find the benchmark value
+benchmark_tsne_s_curve <- find_benchmark_value(.data = distance_tsne_s_curve, distance_col = distance)
+
+# colour_long_edges(.data = distance_tsne_s_curve, benchmark_value = benchmark_tsne_s_curve,
+#                   triangular_object = tr1_object_tsne_s_curve, distance_col = distance)
+
+trimesh_removed_tsne_s_curve <- remove_long_edges(.data = distance_tsne_s_curve, benchmark_value = benchmark_tsne_s_curve,
+                                                  triangular_object = tr1_object_tsne_s_curve, distance_col = distance)
+
+trimesh_removed_tsne_s_curve <- trimesh_removed_tsne_s_curve +
+  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  # theme(axis.text = element_text(size = 5),
+  #       axis.title = element_text(size = 7)) +
+  geom_point(colour = "#33a02c", size = 0.5) +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+# tour_tsne_s_curve <- show_langevitour(df_all_tsne_s_curve, df_bin_tsne_s_curve, df_bin_centroids_tsne_s_curve, benchmark_value = benchmark_tsne_s_curve, distance = distance_tsne_s_curve, distance_col = distance)
+
+## UMAP
+
+num_bins_umap_s_curve <- 6
+shape_val_umap_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
+                                                          x = UMAP1, y = UMAP2) ## 1.259938
+## To extract bin centroids
+hexbin_data_object_umap_s_curve <- extract_hexbin_centroids(nldr_df = UMAP_s_curve, num_bins = num_bins_umap_s_curve, shape_val = shape_val_umap_s_curve, x = UMAP1, y = UMAP2)
+
+df_bin_centroids_umap_s_curve <- hexbin_data_object_umap_s_curve$hexdf_data
+
+UMAP_data_with_hb_id_s_curve <- UMAP_s_curve |>
+  dplyr::mutate(hb_id = hexbin_data_object_umap_s_curve$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_umap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), UMAP_data_with_hb_id_s_curve)
+
+## Averaged on high-D
+df_bin_umap_s_curve <- avg_highD_data(.data = df_all_umap_s_curve)
+
+## Triangulate bin centroids
+tr1_object_umap_s_curve <- triangulate_bin_centroids(df_bin_centroids_umap_s_curve, x, y)
+tr_from_to_df_umap_s_curve <- generate_edge_info(triangular_object = tr1_object_umap_s_curve)
+
+# ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
+#   geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
+#   geom_point(size = 2, colour = "#33a02c") +
+#   coord_equal()
+
+## Compute 2D distances
+distance_umap_s_curve <- cal_2D_dist(.data = tr_from_to_df_umap_s_curve)
+
+## To find the benchmark value
+benchmark_umap_s_curve <- find_benchmark_value(.data = distance_umap_s_curve, distance_col = distance)
+
+# colour_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve,
+#                   triangular_object = tr1_object_umap_s_curve, distance_col = distance)
+
+trimesh_removed_umap_s_curve <- remove_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve,
+                                     triangular_object = tr1_object_umap_s_curve, distance_col = distance)
+
+trimesh_removed_umap_s_curve <- trimesh_removed_umap_s_curve +
+  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  # theme(axis.text = element_text(size = 5),
+  #       axis.title = element_text(size = 7)) +
+  geom_point(colour = "#33a02c", size = 0.5) +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+# tour_umap_s_curve <- show_langevitour(df_all_umap_s_curve, df_bin_umap_s_curve, df_bin_centroids_umap_s_curve, benchmark_value = benchmark_umap_s_curve, distance = distance_umap_s_curve, distance_col = distance)
+
+## TriMAP
+
+num_bins_trimap_s_curve <- 6
+shape_val_trimap_s_curve <- calculate_effective_shape_value(.data = TriMAP_s_curve,
+                                                            x = TriMAP1, y = TriMAP2) ## 1.259938
+## To extract bin centroids
+hexbin_data_object_trimap_s_curve <- extract_hexbin_centroids(nldr_df = TriMAP_s_curve, num_bins = num_bins_trimap_s_curve, shape_val = shape_val_trimap_s_curve, x = TriMAP1, y = TriMAP2)
+
+df_bin_centroids_trimap_s_curve <- hexbin_data_object_trimap_s_curve$hexdf_data
+
+TriMAP_data_with_hb_id_s_curve <- TriMAP_s_curve |>
+  dplyr::mutate(hb_id = hexbin_data_object_trimap_s_curve$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_trimap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), TriMAP_data_with_hb_id_s_curve)
+
+## Averaged on high-D
+df_bin_trimap_s_curve <- avg_highD_data(.data = df_all_trimap_s_curve)
+
+## Triangulate bin centroids
+tr1_object_trimap_s_curve <- triangulate_bin_centroids(df_bin_centroids_trimap_s_curve, x, y)
+tr_from_to_df_trimap_s_curve <- generate_edge_info(triangular_object = tr1_object_trimap_s_curve)
+
+# ggplot(df_bin_centroids_trimap_s_curve, aes(x = x, y = y)) +
+#   geom_segment(data = tr_from_to_df_trimap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
+#   geom_point(size = 2, colour = "#33a02c") +
+#   coord_equal()
+
+## Compute 2D distances
+distance_trimap_s_curve <- cal_2D_dist(.data = tr_from_to_df_trimap_s_curve)
+
+## To find the benchmark value
+benchmark_trimap_s_curve <- find_benchmark_value(.data = distance_trimap_s_curve, distance_col = distance)
+
+# colour_long_edges(.data = distance_trimap_s_curve, benchmark_value = benchmark_trimap_s_curve,
+#                   triangular_object = tr1_object_trimap_s_curve, distance_col = distance)
+
+trimesh_removed_trimap_s_curve <- remove_long_edges(.data = distance_trimap_s_curve, benchmark_value = benchmark_trimap_s_curve,
+                                                    triangular_object = tr1_object_trimap_s_curve, distance_col = distance)
+
+trimesh_removed_trimap_s_curve <- trimesh_removed_trimap_s_curve +
+  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  # theme(axis.text = element_text(size = 5),
+  #       axis.title = element_text(size = 7)) +
+  geom_point(colour = "#33a02c", size = 0.25) +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+# tour_trimap_s_curve <- show_langevitour(df_all_trimap_s_curve, df_bin_trimap_s_curve, df_bin_centroids_trimap_s_curve, benchmark_value = benchmark_trimap_s_curve, distance = distance_trimap_s_curve, distance_col = distance)
+
+## PacMAP
+
+num_bins_pacmap_s_curve <- 10
+shape_val_pacmap_s_curve <- calculate_effective_shape_value(.data = PaCMAP_s_curve,
+                                                            x = PaCMAP1, y = PaCMAP2) ## 1.259938
+## To extract bin centroids
+hexbin_data_object_pacmap_s_curve <- extract_hexbin_centroids(nldr_df = PaCMAP_s_curve, num_bins = num_bins_pacmap_s_curve, shape_val = shape_val_pacmap_s_curve, x = PaCMAP1, y = PaCMAP2)
+
+df_bin_centroids_pacmap_s_curve <- hexbin_data_object_pacmap_s_curve$hexdf_data
+
+PaCMAP_data_with_hb_id_s_curve <- PaCMAP_s_curve |>
+  dplyr::mutate(hb_id = hexbin_data_object_pacmap_s_curve$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_pacmap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), PaCMAP_data_with_hb_id_s_curve)
+
+## Averaged on high-D
+df_bin_pacmap_s_curve <- avg_highD_data(.data = df_all_pacmap_s_curve)
+
+## Triangulate bin centroids
+tr1_object_pacmap_s_curve <- triangulate_bin_centroids(df_bin_centroids_pacmap_s_curve, x, y)
+tr_from_to_df_pacmap_s_curve <- generate_edge_info(triangular_object = tr1_object_pacmap_s_curve)
+
+# ggplot(df_bin_centroids_pacmap_s_curve, aes(x = x, y = y)) +
+#   geom_segment(data = tr_from_to_df_pacmap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
+#   geom_point(size = 2, colour = "#33a02c") +
+#   coord_equal()
+
+## Compute 2D distances
+distance_pacmap_s_curve <- cal_2D_dist(.data = tr_from_to_df_pacmap_s_curve)
+
+## To find the benchmark value
+benchmark_pacmap_s_curve <- find_benchmark_value(.data = distance_pacmap_s_curve, distance_col = distance)
+
+# colour_long_edges(.data = distance_pacmap_s_curve, benchmark_value = benchmark_pacmap_s_curve,
+#                   triangular_object = tr1_object_pacmap_s_curve, distance_col = distance)
+
+trimesh_removed_pacmap_s_curve <- remove_long_edges(.data = distance_pacmap_s_curve, benchmark_value = benchmark_pacmap_s_curve,
+                                     triangular_object = tr1_object_pacmap_s_curve, distance_col = distance)
+
+trimesh_removed_pacmap_s_curve <- trimesh_removed_pacmap_s_curve +
+  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  # theme(axis.text = element_text(size = 5),
+  #       axis.title = element_text(size = 7)) +
+  geom_point(colour = "#33a02c", size = 0.5) +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'e', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+# tour_pacmap_s_curve <- show_langevitour(df_all_pacmap_s_curve, df_bin_pacmap_s_curve, df_bin_centroids_pacmap_s_curve, benchmark_value = benchmark_pacmap_s_curve, distance = distance_pacmap_s_curve, distance_col = distance)
+
+## PHATE
+
+num_bins_phate_s_curve <- 8
+shape_val_phate_s_curve <- calculate_effective_shape_value(.data = PHATE_s_curve,
+                                                           x = PHATE1, y = PHATE2) ## 1.259938
+## To extract bin centroids
+hexbin_data_object_phate_s_curve <- extract_hexbin_centroids(nldr_df = PHATE_s_curve, num_bins = num_bins_phate_s_curve, shape_val = shape_val_phate_s_curve, x = PHATE1, y = PHATE2)
+
+df_bin_centroids_phate_s_curve <- hexbin_data_object_phate_s_curve$hexdf_data
+
+## Identify bins with low-density
+identify_rm_bins <- find_low_density_hexagons(df_bin_centroids_phate_s_curve, num_bins_phate_s_curve, benchmark_rm_hex = 0.06)
+
+df_bin_centroids_phate_s_curve <- df_bin_centroids_phate_s_curve |>
+  filter(!(hexID %in% identify_rm_bins))
+
+PHATE_data_with_hb_id_s_curve <- PHATE_s_curve |>
+  dplyr::mutate(hb_id = hexbin_data_object_phate_s_curve$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_phate_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), PHATE_data_with_hb_id_s_curve)
+
+## Averaged on high-D
+df_bin_phate_s_curve <- avg_highD_data(.data = df_all_phate_s_curve)
+
+## Triangulate bin centroids
+tr1_object_phate_s_curve <- triangulate_bin_centroids(df_bin_centroids_phate_s_curve, x, y)
+tr_from_to_df_phate_s_curve <- generate_edge_info(triangular_object = tr1_object_phate_s_curve)
+
+# ggplot(df_bin_centroids_phate_s_curve, aes(x = x, y = y)) +
+#   geom_segment(data = tr_from_to_df_phate_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
+#   geom_point(size = 2, colour = "#33a02c") +
+#   coord_equal()
+
+## Compute 2D distances
+distance_phate_s_curve <- cal_2D_dist(.data = tr_from_to_df_phate_s_curve)
+
+## To find the benchmark value
+benchmark_phate_s_curve <- find_benchmark_value(.data = distance_phate_s_curve, distance_col = distance)
+
+# colour_long_edges(.data = distance_phate_s_curve, benchmark_value = benchmark_phate_s_curve,
+#                   triangular_object = tr1_object_phate_s_curve, distance_col = distance)
+
+trimesh_removed_phate_s_curve <- remove_long_edges(.data = distance_phate_s_curve, benchmark_value = benchmark_phate_s_curve,
+                                                   triangular_object = tr1_object_phate_s_curve, distance_col = distance)
+
+trimesh_removed_phate_s_curve <- trimesh_removed_phate_s_curve +
+  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  # theme(axis.text = element_text(size = 5),
+  #       axis.title = element_text(size = 7)) +
+  geom_point(colour = "#33a02c", size = 0.5) +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+# tour_phate_s_curve <- show_langevitour(df_all_phate_s_curve, df_bin_phate_s_curve, df_bin_centroids_phate_s_curve, benchmark_value = benchmark_phate_s_curve, distance = distance_phate_s_curve, distance_col = distance)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+#| label: fig-modelScurve
+#| fig-cap: Is there a best model to represent the original data in 2D space or are they all providing equivalent information?, (a) Model in the 2D space with tSNE (<https://youtu.be/uy9SkAo6gAE>), (b) Model in the 2D space with UMAP (<https://youtu.be/0MJDhHrh_Ug>), (c) Model in the 2D space with PHATE (<https://youtu.be/HbVv0uy0QWk>), (d) Model in the 2D space with TriMAP (<https://youtu.be/2OGwipAjzc8>), and (e) Model in the 2D space with PaCMAP (<https://youtu.be/pkokI8d-cBk>).
+
+trimesh_removed_tsne_s_curve + trimesh_removed_umap_s_curve + trimesh_removed_phate_s_curve + trimesh_removed_trimap_s_curve + trimesh_removed_pacmap_s_curve +
+  #plot_annotation(tag_levels = 'a') +
+  plot_layout(ncol = 5) &
+  theme(legend.position='none', plot.tag = element_text(size = 8))
+```
+
 ### Simulated data example {#sec-simpleex}

### 2024-01-14 paper.tex
Add links to the model of S-curve, and What I lear from the models

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, interior hidden, breakable, enhanced, sharp corners, frame hidden, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, interior hidden, boxrule=0pt, sharp corners, frame hidden, breakable, enhanced]}{\end{tcolorbox}}\fi
 
@@
 
-\begin{figure}
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-modelScurve-1.pdf}
-
-}
-
-\caption{\label{fig-modelScurve}Model.}
-
-\end{figure}
-
 \hypertarget{tunning-the-model}{%
@@
 \begin{equation}\protect\hypertarget{eq-equation9}{}{
-\text{MSE} = \sum_{j = 1}^{n} \frac{\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2}{\text{total number of bins}}
+\text{MSE} = \sum_{j = 1}^{n} \frac{\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2}{b}
 }\label{eq-equation9}\end{equation}
@@
 absence of such functionalities.
+
+\hypertarget{model-interpretation}{%
+\subsection{Model interpretation}\label{model-interpretation}}
+
+The visualizations of NLDR techniques for the S-curve dataset shows
+different representations in 2D, as showcased in Figure
+Figure~\ref{fig-nldervis}. The question at hand is determining the best
+NLDR technique to represent the S-curve data. Examining the 2D layouts
+in Figure~\ref{fig-nldervis} suggests that TriMAP or PaCMAP may be best
+representations. However, the challenge is to decide the best
+representation with considerable evidence.
+
+After considering the models constructed from the algorithm for each
+NLDR techniques and visualizing them in high-D space with the original
+data, shows that all the NLDR techniques capture the non-linearity of
+the S-curve (see videos of the models linked in
+Figure~\ref{fig-modelScurve}). But thorough examination of each model in
+high-D space shows some surprising findings. The model of PHATE is
+flattened and lack of capturing the width of S-curve (see videos of the
+model with PHATE linked in Figure~\ref{fig-modelScurve}). This nature of
+PHATE is also evicted by the considerably higher AIC and MSE values for
+PHATE as shown in Figure~\ref{fig-diagnosticpltScurve}. Furthermore,
+there are two edges that are not connected to any points in the model of
+TriMAP (see see videos of the model with TriMAP linked in
+Figure~\ref{fig-modelScurve}). This model shows how the patterns in 2D
+representation of TriMAP capture by the model (see
+Figure~\ref{fig-nldervis} (d)). As shown in Figure~\ref{fig-nldervis}
+(b), UMAP have four clusters in a curvilinear structure. However, the
+model reveals that even though UMAP shows a curvilinear in 2D, UMAP
+actually capture the structure of the S-curve without any disjoint in
+the model (see see videos of the model with UMAP linked in
+Figure~\ref{fig-modelScurve}). In the 2D layouts of tSNE and PaCMAP (see
+Figure~\ref{fig-nldervis} (a), (e)), there are some sparse areas that
+are captured by the models effectively (see see videos of the model with
+tSNE and PaCMAP linked in Figure~\ref{fig-modelScurve}).
+
+Different NLDR techniques perform differently on the S-curve data and
+tSNE provides the best representation in 2D according to AIC and MSE
+values (see Figure~\ref{fig-diagnosticpltScurve}).
+
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-modelScurve-1.pdf}
+
+}
+
+\caption{\label{fig-modelScurve}Is there a best model to represent the
+original data in 2D space or are they all providing equivalent
+information?, (a) Model in the 2D space with tSNE
+(\url{https://youtu.be/uy9SkAo6gAE}), (b) Model in the 2D space with
+UMAP (\url{https://youtu.be/0MJDhHrh_Ug}), (c) Model in the 2D space
+with PHATE (\url{https://youtu.be/HbVv0uy0QWk}), (d) Model in the 2D
+space with TriMAP (\url{https://youtu.be/2OGwipAjzc8}), and (e) Model in
+the 2D space with PaCMAP (\url{https://youtu.be/pkokI8d-cBk}).}
+
+\end{figure}
 

### 2024-01-14 quollr_code.R

@@ -478,7 +478,7 @@ show_langevitour <- function(df, df_b, df_b_with_center_data, benchmark_value = 
       dplyr::filter({{ distance_col }} < benchmark_value)
     ## Since erase brushing is considerd.
 
-    langevitour::langevitour(df_exe[1:(length(df_exe)-1)], lineFrom = distance_df_small_edges$from, lineTo = distance_df_small_edges$to, group = df_exe$type)
+    langevitour::langevitour(df_exe[1:(length(df_exe)-1)], lineFrom = distance_df_small_edges$from, lineTo = distance_df_small_edges$to, group = df_exe$type, pointSize = 3, levelColors = c("#6a3d9a", "#33a02c"))
 
   } else if ((is.na(benchmark_value)) && (!(is.na(min_points_threshold)))) {
     df_bin_centroids_filterd <- df_bin_centroids %>%

### 2024-01-15 paper.qmd
Rewrite the motivation of our paper

@@
 
-High-dimensional (high-D) data is widespread in many fields including ecology and bioinformatics [@Guo2023], in part because of new data collection technologies [@Johnstone2009; @ayesha2020overview].
+High-dimensional (high-D) data is prevalent across various fields, such as ecology and bioinformatics [@Guo2023], due to advancements in data collection technologies [@Johnstone2009; @ayesha2020overview].
+However, visualization of high-D data introduces significant challenges, because the complexity of visualizing data beyond two dimensions [@Jia2022].
+In recent years, interactive and dynamic graphics systems like **liminal** [@article21] —which employs interactive tools like brushing and linking [@article58]—and software tools such as **XGobi**, **GGobi** [@article60], **tourr**  [@article61], **detourr** [@article22], and **langevitour**  [@article09], involving dynamic methods like tours [@Asimov1985], have played a key role in visualizing high-D data (data-vis).
+
+To create low-dimensional representations (typically in 2D) of high-D data, it is common to apply dimension reduction (DR) techniques.
+Approaches for DR involve linear methods such as principal component analysis (PCA) [@Karl1901], non-linear methods such as multi-dimensional scaling (MDS) [@Torgerson1967].
+In the past decade, many new non-linear dimension reduction (NLDR) techniques have emerged, such as t-distributed stochastic neighbor embedding (tSNE) [@Laurens2008] and uniform manifold approximation and projection (UMAP) [@Leland2018], designed to capture the complex and non-linear relationships present within high-D data [@Johnstone2009].
+NLDR techniques are the 2D models of high-D data in our context.
+
+Visualization of NLDR techniques (m-vis) [@article59] for the same high-D data is particularly important in understanding and finding the best representation.
+If we have done so, the 2D models can be quite different from the data structure shown in high-D space.
+Therefore, visualizing the 2D model in high-D space (m-in-dis) is more useful to answer different types of questions:
+
+- Is there a best representation of high-D data or are they all providing  equivalent information?
+Is there a best parameter choice to fit the 2D model?
+How does the model change when it's parameters change?
+
+- How well the does the 2D models capture the data structure? Is the model fitting able to capture different data structure like non-linear, clustering?
+
+If we can't easily ask and answer these questions, the ability of understanding the models are restricted.
+To find the best 2D model and the parameter choices, better understanding of the underlying science is really important.
+
+Also, the importance of m-vis along with data-vis has been recognized and incorporated into interactive software, **liminal** [@article21].
+But the 2D model and high-D visualize side by side and interactive like brushing and linking connect the data in the two panels.
+To address this challenge, we propose a novel approach by combining the tour technique with a low-dimensional manifold.
+This manifold is created through the synergistic use of NLDR techniques, hexagonal binning, and triangulation.
+This integration facilitates a more understanding of the data structure, how well (or how poorly) NLDR techniques perform.
+
+The outline of this paper is as follows.
+The @sec-background provides an detailed overview of dimension reduction methods, and tours.
+Building upon this foundation, the @sec-methods delves into the proposed algorithm, its implementation details, how to tune the model, model summaries, and a synthetic example to illustrate the functionality of the algorithm.
+Subsequently, @sec-applications showcases applications of the algorithm on different data sets, particularly in single-cell RNA-seq data.
+These applications reveal insights into the performance and trustworthiness of NLDR algorithms.
+We analyze the results to identify situations where NLDR techniques may lead to misleading interpretations.
+Finally, @sec-conclusions concludes by summarizing the findings and emphasizing the significance of the proposed approach in tackling the challenges of high-dimensional data visualization.
+
+<!--High-dimensional (high-D) data is widespread in many fields including ecology and bioinformatics [@Guo2023], in part because of new data collection technologies [@Johnstone2009; @ayesha2020overview].
 Working with high-dimensional data poses considerable challenges due to the difficulty in visualizing beyond two dimensions [@Jia2022].
@@
 To create visual representations of high-dimensional data, it is common to apply dimension reduction techniques.
-Linear methods such as principal component analysis (PCA) [@Karl1901] have been used for many years, and remain popular.
-Non-linear methods such as multi-dimensional scaling (MDS) [@Torgerson1967] have also been routinely used.
-In the past decade, there has merged many new techniques non-linear dimension reduction (NLDR), such as t-distributed stochastic neighbor embedding (tSNE) [@Laurens2008], uniform manifold approximation and projection (UMAP) [@Leland2018], designed to capture the complex and non-linear relationships present within high-dimensional data [@Johnstone2009].
 
@@
 Augmenting the results of non-linear dimensional reduction methods with the tour, as demonstrated in the **liminal** R package [@article21], further enhances data exploration.
+
+The earlier approaches for DR involve linear methods such as PCA [@Karl1901].
+PCA aims to maintain the second-order statistics of the data by projecting the points into the low dimensional space that preserves the maximum amount of variance among all such projections.
+As a result, PCA has been shown to be effective in preserving the global structure of the data (Silva & Tenenbaum, 2003).
+The global structure includes the overall shape of the dataset, placement of the clusters, and the existence of potential outliers.
+Unlike PCA, much of the focus of the more recent non-linear methods, including t-SNE (Maaten & Hinton,2008), LargeVis (Tang et al., 2016), and UMAP (McInnes et al., 2018) has been on preserving the local neighborhood structure of each individual point.
+Mainly,
 
@@
 Finally, @sec-conclusions concludes by summarizing the findings and emphasizing the significance of the proposed approach in tackling the challenges of high-dimensional data visualization.
+-->
 
@@
 As a result of hexagonal binning for high-D data, all the high-D data are clustered into hexagons.
-In this step, the bin centroids ($C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2}$) (see @fig-meth Step 2) are obtained [@Carr2013].
+In this step, the bin centroids ($C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2})$) (see @fig-meth Step 2) are obtained [@Carr2013].
 
@@
 
-In this step, the algorithm proceeds to triangulate the hexagonal bin centroids (see @fig-meth (Step 3)).
+In this step, the algorithm proceeds to triangulate the hexagonal bin centroids (see @fig-meth Step 3).
 Triangulation is a fundamental process in computational geometry and computer graphics that involves dividing a set of points in a given space into interconnected triangles [@article30].
@@
 
-## Discussion {#sec-discussion}
-
-Our research introduces a comprehensive framework that leverages tours for interactive exploration of high-dimensional data coupled with a low-dimensional manifold, facilitated by the `quollr` R package.
+## Conclusion {#sec-conclusion}
+
+<!--Our research introduces a comprehensive framework that leverages tours for interactive exploration of high-dimensional data coupled with a low-dimensional manifold, facilitated by the `quollr` R package.
 Regardless of the Non-Linear Dimension Reduction (NLDR) technique in use, our approach demonstrates effectiveness through simulation examples, particularly in the iterative removal of long edges for a smoother representation and capturing cluster variance.
@@
 By leveraging the advantages of tours and low-dimensional manifolds, our approach offers valuable insights into the data transformation process, empowering users to make informed decisions in analyzing high-dimensional data.
-Future work could enhance the method's robustness in the presence of noise and explore additional evaluation metrics for quantifying embedding accuracy.
+Future work could enhance the method's robustness in the presence of noise and explore additional evaluation metrics for quantifying embedding accuracy.-->
 

### 2024-01-15 paper.tex
Rewrite the motivation of our paper

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, interior hidden, boxrule=0pt, sharp corners, frame hidden, breakable, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, interior hidden, enhanced, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi
 
@@
 
-High-dimensional (high-D) data is widespread in many fields including
-ecology and bioinformatics \citep{Guo2023}, in part because of new data
-collection technologies \citep{Johnstone2009, ayesha2020overview}.
-Working with high-dimensional data poses considerable challenges due to
-the difficulty in visualizing beyond two dimensions \citep{Jia2022}.
-High-dimensional data also presents difficulties for model fitting
-\citep{Johnstone2009}, both computationally and interpretation, each of
-which benefits from visualization.
-
-To create visual representations of high-dimensional data, it is common
-to apply dimension reduction techniques. Linear methods such as
-principal component analysis (PCA) \citep{Karl1901} have been used for
-many years, and remain popular. Non-linear methods such as
-multi-dimensional scaling (MDS) \citep{Torgerson1967} have also been
-routinely used. In the past decade, there has merged many new techniques
-non-linear dimension reduction (NLDR), such as t-distributed stochastic
-neighbor embedding (tSNE) \citep{Laurens2008}, uniform manifold
-approximation and projection (UMAP) \citep{Leland2018}, designed to
-capture the complex and non-linear relationships present within
-high-dimensional data \citep{Johnstone2009}.
-
-However, projecting high-dimensional data has limitations, such as
-information loss and potential distortion of essential structures and
-patterns \citep[\citet{article53}]{Jia2022}. The choice of technique and
-parameters further impacts the accuracy of the visualization,
-necessitating careful consideration for meaningful interpretation (see
-Figure~\ref{fig-nldervis}).
-
-Interactive and dynamic graphics systems have also been developed over
-the years to enable visualizing high dimensions. One method, called a
-tour \citep{Asimov1985}, shows a sequence of linear projections is shown
-as a movie, allowing exploration without warping the space
-\citep{lee2021review}. Interactive tools like \textbf{XGobi} and
-\textbf{GGobi} have been successful in incorporating tours for exploring
-high-dimensional data \citep{article60}. The R package \textbf{tourr}
-\citep{article61} further enhances tour visualization within R, although
-it may face limitations in frame rate and interactive features compared
-to \textbf{GGobi}.
-
-To overcome these limitations, the R package \textbf{detourr}
-\citep{article22} has been developed, leveraging a Javascript widget via
-htmlwidgets \citep{Ramnath2023} to achieve higher frame rates and
-enhanced interactivity. Additionally, the R package \textbf{langevitour}
-\citep{article09} utilizes Langevin Dynamics to generate a continuous
-path of projections, eliminating the need for interpolation between
-projections for animation. The tour technique has proven valuable in
-exploring statistical model fits \citep{article58} and factorial
-experimental designs \citep{article59}. Augmenting the results of
-non-linear dimensional reduction methods with the tour, as demonstrated
-in the \textbf{liminal} R package \citep{article21}, further enhances
-data exploration.
-
-While tours \citep{Asimov1985} preserve space without warping
-\citep{lee2021review}, they require integrating multiple low-dimensional
-views mentally to perceive high-dimensional structures. To address this
-challenge, we propose a novel approach by combining the tour technique
-with a low-dimensional manifold. This manifold is created through the
-synergistic use of Non-Linear Dimension Reduction (NLDR) techniques,
-hexagonal binning, and triangulation. By merging these techniques, our
-approach offers a comprehensive and efficient means to visualize and
-explore high-dimensional data while retaining the advantages of both
-tours and NLDR. This integration facilitates a more intuitive perception
-of complex data structures and empowers analysts with a robust tool for
-assessing the accuracy of data transformations. The implementation of
-our approach is available as an R package called \textbf{quollr}.
+High-dimensional (high-D) data is prevalent across various fields, such
+as ecology and bioinformatics \citep{Guo2023}, due to advancements in
+data collection technologies \citep{Johnstone2009, ayesha2020overview}.
+However, visualization of high-D data introduces significant challenges,
+because the complexity of visualizing data beyond two dimensions
+\citep{Jia2022}. In recent years, interactive and dynamic graphics
+systems like \textbf{liminal} \citep{article21} ---which employs
+interactive tools like brushing and linking \citep{article58}---and
+software tools such as \textbf{XGobi}, \textbf{GGobi} \citep{article60},
+\textbf{tourr} \citep{article61}, \textbf{detourr} \citep{article22},
+and \textbf{langevitour} \citep{article09}, involving dynamic methods
+like tours \citep{Asimov1985}, have played a key role in visualizing
+high-D data (data-vis).
+
+To create low-dimensional representations (typically in 2D) of high-D
+data, it is common to apply dimension reduction (DR) techniques.
+Approaches for DR involve linear methods such as principal component
+analysis (PCA) \citep{Karl1901}, non-linear methods such as
+multi-dimensional scaling (MDS) \citep{Torgerson1967}. In the past
+decade, many new non-linear dimension reduction (NLDR) techniques have
+emerged, such as t-distributed stochastic neighbor embedding (tSNE)
+\citep{Laurens2008} and uniform manifold approximation and projection
+(UMAP) \citep{Leland2018}, designed to capture the complex and
+non-linear relationships present within high-D data
+\citep{Johnstone2009}. NLDR techniques are the 2D models of high-D data
+in our context.
+
+Visualization of NLDR techniques (m-vis) \citep{article59} for the same
+high-D data is particularly important in understanding and finding the
+best representation. If we have done so, the 2D models can be quite
+different from the data structure shown in high-D space. Therefore,
+visualizing the 2D model in high-D space (m-in-dis) is more useful to
+answer different types of questions:
+
+\begin{itemize}
+\item
+  Is there a best representation of high-D data or are they all
+  providing equivalent information? Is there a best parameter choice to
+  fit the 2D model? How does the model change when it's parameters
+  change?
+\item
+  How well the does the 2D models capture the data structure? Is the
+  model fitting able to capture different data structure like
+  non-linear, clustering?
+\end{itemize}
+
+If we can't easily ask and answer these questions, the ability of
+understanding the models are restricted. To find the best 2D model and
+the parameter choices, better understanding of the underlying science is
+really important.
+
+Also, the importance of m-vis along with data-vis has been recognized
+and incorporated into interactive software, \textbf{liminal}
+\citep{article21}. But the 2D model and high-D visualize side by side
+and interactive like brushing and linking connect the data in the two
+panels. To address this challenge, we propose a novel approach by
+combining the tour technique with a low-dimensional manifold. This
+manifold is created through the synergistic use of NLDR techniques,
+hexagonal binning, and triangulation. This integration facilitates a
+more understanding of the data structure, how well (or how poorly) NLDR
+techniques perform.
 
@@
 Section~\ref{sec-background} provides an detailed overview of dimension
-reduction methods, triangulation, and tours. Building upon this
-foundation, the Section~\ref{sec-methods} delves into the proposed
-algorithm, \textbf{quollr}, and its implementation details. In
-Section~\ref{sec-prediction}, discusses the effectiveness of the learned
-low-dimensional manifold in accurately representing the complex
-high-dimensional data. Following that, Section~\ref{sec-simpleex}
-presents simple examples from simulations to illustrate the
-functionality of the algorithm. Subsequently,
-Section~\ref{sec-applications} showcases real-world applications of
-\textbf{quollr} on different data sets, particularly in single-cell
+reduction methods, and tours. Building upon this foundation, the
+Section~\ref{sec-methods} delves into the proposed algorithm, its
+implementation details, how to tune the model, model summaries, and a
+synthetic example to illustrate the functionality of the algorithm.
+Subsequently, Section~\ref{sec-applications} showcases applications of
+the algorithm on different data sets, particularly in single-cell
 RNA-seq data. These applications reveal insights into the performance
@@
 are clustered into hexagons. In this step, the bin centroids
-(\(C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2}\)) (see Figure~\ref{fig-meth}
+(\(C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2})\)) (see Figure~\ref{fig-meth}
 Step 2) are obtained \citep{Carr2013}.
@@
 In this step, the algorithm proceeds to triangulate the hexagonal bin
-centroids (see Figure~\ref{fig-meth} (Step 3)). Triangulation is a
+centroids (see Figure~\ref{fig-meth} Step 3). Triangulation is a
 fundamental process in computational geometry and computer graphics that
@@
 
-\hypertarget{sec-discussion}{%
-\section{Discussion}\label{sec-discussion}}
-
-Our research introduces a comprehensive framework that leverages tours
-for interactive exploration of high-dimensional data coupled with a
-low-dimensional manifold, facilitated by the \texttt{quollr} R package.
-Regardless of the Non-Linear Dimension Reduction (NLDR) technique in
-use, our approach demonstrates effectiveness through simulation
-examples, particularly in the iterative removal of long edges for a
-smoother representation and capturing cluster variance.
-
-In the example with doublets, our method successfully captures the tweak
-within each cluster, indicating the variance present within them.
-However, the model may not appear smooth in high-dimensional space due
-to considerable noise when the data has a piecewise linear geometry,
-such as the tree simulation.
-
-The practical application of our framework, as showcased with the UMAP
-view, enables visual inspection of well-separated clusters. Furthermore,
-the combined tour and model provide a robust assessment of whether UMAP
-preserves the data structure and accurately transforms the data.
-
-The advantages of our approach include its versatility across various
-NLDR techniques and the ability to generate interactive visualizations
-for detailed exploration. The tour provides an intuitive way to navigate
-and comprehend high-dimensional data while assessing the accuracy of
-dimensionality reduction.
-
-However, one limitation is that the approach may be less effective in
-cases with significant noise, as seen in the tree simulation example.
-Additionally, while our method aids in visual verification, quantifying
-the accuracy of embeddings might require further evaluation metrics.
-
-In conclusion, our framework presents a powerful tool for researchers
-and analysts in single-cell studies to assess their embeddings by
-visually inspecting them alongside the original data. By leveraging the
-advantages of tours and low-dimensional manifolds, our approach offers
-valuable insights into the data transformation process, empowering users
-to make informed decisions in analyzing high-dimensional data. Future
-work could enhance the method's robustness in the presence of noise and
-explore additional evaluation metrics for quantifying embedding
-accuracy.
+\hypertarget{sec-conclusion}{%
+\section{Conclusion}\label{sec-conclusion}}
 

### 2024-01-15 paper.qmd
Add logic to remove long edges and add error calculation

@@
 
+- Compute high-D distances (using averaged high-D data)
+- Fit a linear model using response as distance_high_D and predictor as distance_2D
+- Filter the model residuals only for short edges
+- compute error for each edge connection as (residual for that point)/(total residual for all edges * number of short edges)
+
+```{r}
+#| echo: false
+#| warning: false
+
+error_df_lg_1 <- read_rds("data/s_curve/s_curve_summary_lg_tsne.rds")
+error_df_lg_2 <- read_rds("data/s_curve/s_curve_summary_lg_umap.rds")
+error_df_lg_3 <- read_rds("data/s_curve/s_curve_summary_lg_phate.rds")
+error_df_lg_4 <- read_rds("data/s_curve/s_curve_summary_lg_trimap.rds")
+error_df_lg_5 <- read_rds("data/s_curve/s_curve_summary_lg_pacmap.rds")
+
+error_df_lg <- dplyr::bind_rows(error_df_lg_1, error_df_lg_2, error_df_lg_3, error_df_lg_4, error_df_lg_5)
+
+error_df_lg$method <- factor(error_df_lg$method, levels = c("tSNE", "UMAP", "PHATE", "TriMAP", "PaCMAP"))
+
+error_plot_lg_1 <- ggplot(error_df_lg_1, aes(x = benchmark_rm_lg, y = total_error, color = method)) +
+  geom_point() +
+  geom_line() +
+  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  theme_light() +
+  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  scale_colour_manual(values = c("#e41a1c")) +
+  ylab("error") +
+  xlab("Benchmark value")
+
+error_plot_lg_2 <- ggplot(error_df_lg_2, aes(x = benchmark_rm_lg, y = total_error, color = method)) +
+  geom_point() +
+  geom_line() +
+  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  theme_light() +
+  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  scale_colour_manual(values = c("#377eb8")) +
+  ylab("error") +
+  xlab("Benchmark value")
+
+error_plot_lg_3 <- ggplot(error_df_lg_3, aes(x = benchmark_rm_lg, y = total_error, color = method)) +
+  geom_point() +
+  geom_line() +
+  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  theme_light() +
+  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  scale_colour_manual(values = c("#4daf4a")) +
+  ylab("error") +
+  xlab("Benchmark value")
+
+error_plot_lg_4 <- ggplot(error_df_lg_4, aes(x = benchmark_rm_lg, y = total_error, color = method)) +
+  geom_point() +
+  geom_line() +
+  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  theme_light() +
+  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  scale_colour_manual(values = c("#984ea3")) +
+  ylab("error") +
+  xlab("Benchmark value")
+
+error_plot_lg_5 <- ggplot(error_df_lg_5, aes(x = benchmark_rm_lg, y = total_error, color = method)) +
+  geom_point() +
+  geom_line() +
+  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  theme_light() +
+  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  scale_colour_manual(values = c("#ff7f00")) +
+  ylab("error") +
+  xlab("Benchmark value")
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| fig-cap: Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset with different benchmark values to remove the long edges.
+What is the effective benchmark value to remove the long edges?
+#| label: fig-diagnosticpltScurvelgrm
+#| out-width: 100%
+
+error_plot_lg_1 + error_plot_lg_2 + error_plot_lg_3 + error_plot_lg_4 + error_plot_lg_5 +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(guides='collect', nrow = 2) &
+  theme(legend.position='bottom', plot.tag = element_text(size = 8)) 
+```
+
 ### Model summaries {#sec-summary}
@@
 plot_list1_umap <- plot_UMAP_2D(UMAP_data_7) + #ggtitle("(b)") +
-  geom_point(color = "#8dd3c7") +
+  geom_point(alpha=0.5, size = 0.5, colour = "#8dd3c7", fill = "#8dd3c7") +
   theme_linedraw() +
@@
 plot_list2_umap <- plot_UMAP_2D(UMAP_data_15) + #ggtitle("(b)") + 
-  geom_point(color = "#a65628") +
+  geom_point(alpha=0.5, size = 0.5, colour = "#a65628", fill = "#a65628") +
   theme_linedraw() +
@@
 plot_list3_umap <- plot_UMAP_2D(UMAP_data_32) + #ggtitle("(b)") +
-  geom_point(color = "#f781bf") +
+  geom_point(alpha=0.5, size = 0.5, color = "#f781bf", fill = "#f781bf") +
   theme_linedraw() +
@@
 plot_list4_umap <- plot_UMAP_2D(UMAP_data) + #ggtitle("(b)") +
-  geom_point(color = "#999999") +
+  geom_point(alpha=0.5, size = 0.5, colour = "#999999", fill = "#999999") +
   theme_linedraw() +
@@
 
-<!--lwd smoothing criteria with s-curve-->
+<!--lwd smoothing criteria 1 with s-curve-->
 
@@
 #| fig-cap: Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset with different benchmark values to remove the low-density heaxgons.
-What is the effective benchark valuw to remove the low-density heaxgons?
+What is the effective benchmark value to remove the low-density heaxgons?
 #| label: fig-diagnosticpltScurvelwd
@@
 
+https://satijalab.org/seurat/articles/pbmc3k_tutorial.html
+https://www.cell.com/fulltext/S0092-8674(15)00549-8
+
 In the field of single-cell studies, a common analysis task involves clustering to identify groups of cells with similar expression profiles.
@@
 
+### Retina dataset
+
+https://www.cell.com/cms/10.1016/j.cell.2015.05.002/attachment/fd52950c-fd35-401a-9e60-ca7588ef9001/mmc1.pdf
+
+https://www.cell.com/fulltext/S0092-8674(15)00549-8#app3
+
+32 PCs
+tSNE (perplexity 30)
+
 ## Conclusion {#sec-conclusion}

### 2024-01-15 paper.tex
Add logic to remove long edges and add error calculation

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, interior hidden, enhanced, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, enhanced, interior hidden, breakable, sharp corners, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 
+\begin{itemize}
+\tightlist
+\item
+  Compute high-D distances (using averaged high-D data)
+\item
+  Fit a linear model using response as distance\_high\_D and predictor
+  as distance\_2D
+\item
+  Filter the model residuals only for short edges
+\item
+  compute error for each edge connection as (residual for that
+  point)/(total residual for all edges * number of short edges)
+\end{itemize}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticpltScurvelgrm-1.pdf}
+
+}
+
+\caption{\label{fig-diagnosticpltScurvelgrm}Goodness of fit statistics
+from different NLDR techniques applied to training S-curve dataset with
+different benchmark values to remove the long edges. What is the
+effective benchmark value to remove the long edges?}
+
+\end{figure}
+
 \hypertarget{sec-summary}{%
@@
 different benchmark values to remove the low-density heaxgons. What is
-the effective benchark valuw to remove the low-density heaxgons?}
+the effective benchmark value to remove the low-density heaxgons?}
 
@@
 
+https://satijalab.org/seurat/articles/pbmc3k\_tutorial.html
+https://www.cell.com/fulltext/S0092-8674(15)00549-8
+
 In the field of single-cell studies, a common analysis task involves
@@
 
+\hypertarget{retina-dataset}{%
+\subsection{Retina dataset}\label{retina-dataset}}
+
+https://www.cell.com/cms/10.1016/j.cell.2015.05.002/attachment/fd52950c-fd35-401a-9e60-ca7588ef9001/mmc1.pdf
+
+https://www.cell.com/fulltext/S0092-8674(15)00549-8\#app3
+
+32 PCs tSNE (perplexity 30)
+
 \hypertarget{sec-conclusion}{%

### 2024-01-15 scripts/gen_summary_trimap_lg_rm.R

@@ -10,25 +10,25 @@ source("nldr_code.R", local = TRUE)
 df_s_curve <- read_rds("data/s_curve/s_curve.rds")
 training_data_s_curve <- read_rds("data/s_curve/s_curve_training.rds")
 
-UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap.rds")
+TriMAP_s_curve <- read_rds("data/s_curve/s_curve_trimap.rds")
 
 num_bins_s_curve <- 6
 
-shape_value_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
-                                                      x = UMAP1, y = UMAP2) ## 1.259938
+shape_value_s_curve <- calculate_effective_shape_value(.data = TriMAP_s_curve,
+                                                       x = TriMAP1, y = TriMAP2) ## 1.259938
 
 ## To extract bin centroids
-hexbin_data_object_s_curve <- extract_hexbin_centroids(nldr_df = UMAP_s_curve,
-                                                      num_bins = num_bins_s_curve,
-                                                      shape_val = shape_value_s_curve, x = UMAP1, y = UMAP2)
+hexbin_data_object_s_curve <- extract_hexbin_centroids(nldr_df = TriMAP_s_curve,
+                                                       num_bins = num_bins_s_curve,
+                                                       shape_val = shape_value_s_curve, x = TriMAP1, y = TriMAP2)
 
 df_bin_centroids_s_curve <- hexbin_data_object_s_curve$hexdf_data
 
-UMAP_s_curve_with_hb_id <- UMAP_s_curve |>
+TriMAP_s_curve_with_hb_id <- TriMAP_s_curve |>
   dplyr::mutate(hb_id = hexbin_data_object_s_curve$hb_data@cID)
 
 ## To generate a data set with high-D and 2D training data
-df_all_s_curve <- dplyr::bind_cols(training_data_s_curve |> dplyr::select(-ID), UMAP_s_curve_with_hb_id)
+df_all_s_curve <- dplyr::bind_cols(training_data_s_curve |> dplyr::select(-ID), TriMAP_s_curve_with_hb_id)
 
 ## Averaged on high-D
 df_bin_s_curve <- avg_highD_data(.data = df_all_s_curve, column_start_text = "x")
@@ -42,7 +42,7 @@ distance_s_curve <- cal_2D_dist(.data = tr_from_to_df_s_curve)
 
 ## To find the benchmark value
[…]

### 2024-01-17 paper.qmd
Regenerate PBMC3K UMAP results by following the same steps as the author in Literature

@@
 #| fig-cap: Is there a best model to represent the original data in 2D space or are they all providing equivalent information?, (a) Model in the 2D space with tSNE (<https://youtu.be/uy9SkAo6gAE>), (b) Model in the 2D space with UMAP (<https://youtu.be/0MJDhHrh_Ug>), (c) Model in the 2D space with PHATE (<https://youtu.be/HbVv0uy0QWk>), (d) Model in the 2D space with TriMAP (<https://youtu.be/2OGwipAjzc8>), and (e) Model in the 2D space with PaCMAP (<https://youtu.be/pkokI8d-cBk>).
+#| out-width: 100%
 
@@
   plot_layout(ncol=5)
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+### tSNE
+
+# num_bins_tsne_gau <- calculate_effective_x_bins(.data = tSNE_data_gau, x = tSNE1,
+#                                                 cell_area = 1) #56
+num_bins_tsne_gau <- 16
+
+shape_val_tsne_gau <- calculate_effective_shape_value(.data = tSNE_data_gau,
+                                                      x = tSNE1, y = tSNE2)
+## To extract bin centroids
+hexbin_data_object_tsne_gau <- extract_hexbin_centroids(nldr_df = tSNE_data_gau, num_bins = num_bins_tsne_gau, shape_val = shape_val_tsne_gau, x = tSNE1, y = tSNE2)
+
+df_bin_centroids_tsne_gau <- hexbin_data_object_tsne_gau$hexdf_data
+
+tSNE_data_with_hb_id_gau <- tSNE_data_gau |>
+  dplyr::mutate(hb_id = hexbin_data_object_tsne_gau$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_tsne_gau <- dplyr::bind_cols(training_data_1 |> dplyr::select(-ID), tSNE_data_with_hb_id_gau)
+
+## Averaged on high-D
+df_bin_tsne_gau <- avg_highD_data(.data = df_all_tsne_gau)
+
+## Triangulate bin centroids
+tr1_object_tsne_gau <- triangulate_bin_centroids(df_bin_centroids_tsne_gau, x, y)
+tr_from_to_df_tsne_gau <- generate_edge_info(triangular_object = tr1_object_tsne_gau)
+
+## Compute 2D distances
+distance_tsne_gau <- cal_2D_dist(.data = tr_from_to_df_tsne_gau)
+
+## To find the benchmark value
+benchmark_tsne_gau <- find_benchmark_value(.data = distance_tsne_gau, distance_col = distance)
+
+trimesh_removed_tsne_gau <- remove_long_edges(.data = distance_tsne_gau, benchmark_value = benchmark_tsne_gau,
+                                                  triangular_object = tr1_object_tsne_gau, distance_col = distance)
+
+trimesh_removed_tsne_gau <- trimesh_removed_tsne_gau +
+  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  # theme(axis.text = element_text(size = 5),
+  #       axis.title = element_text(size = 7)) +
+  geom_point(colour = "#33a02c", size = 0.05) +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+# tour_tsne_gau <- show_langevitour(df_all_tsne_gau, df_bin_tsne_gau, df_bin_centroids_tsne_gau, benchmark_value = benchmark_tsne_gau, distance = distance_tsne_gau, distance_col = distance)
+
+### UMAP
+
+# num_bins_umap_gau <- calculate_effective_x_bins(.data = UMAP_data_gau, x = UMAP1,
+#                                                 cell_area = 1) #41
+num_bins_umap_gau <- 26
+
+shape_val_umap_gau <- calculate_effective_shape_value(.data = UMAP_data_gau,
+                                                      x = UMAP1, y = UMAP2)
+## To extract bin centroids
+hexbin_data_object_umap_gau <- extract_hexbin_centroids(nldr_df = UMAP_data_gau, num_bins = num_bins_umap_gau, shape_val = shape_val_umap_gau, x = UMAP1, y = UMAP2)
+
+df_bin_centroids_umap_gau <- hexbin_data_object_umap_gau$hexdf_data
+
+UMAP_data_with_hb_id_gau <- UMAP_data_gau |>
+  dplyr::mutate(hb_id = hexbin_data_object_umap_gau$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_umap_gau <- dplyr::bind_cols(training_data_1 |> dplyr::select(-ID), UMAP_data_with_hb_id_gau)
+
+## Averaged on high-D
+df_bin_umap_gau <- avg_highD_data(.data = df_all_umap_gau)
+
+## Triangulate bin centroids
+tr1_object_umap_gau <- triangulate_bin_centroids(df_bin_centroids_umap_gau, x, y)
+tr_from_to_df_umap_gau <- generate_edge_info(triangular_object = tr1_object_umap_gau)
+
+## Compute 2D distances
+distance_umap_gau <- cal_2D_dist(.data = tr_from_to_df_umap_gau)
+
+## To find the benchmark value
+benchmark_umap_gau <- find_benchmark_value(.data = distance_umap_gau, distance_col = distance)
+
+trimesh_removed_umap_gau <- remove_long_edges(.data = distance_umap_gau, benchmark_value = benchmark_umap_gau,
+                                              triangular_object = tr1_object_umap_gau, distance_col = distance)
+
+trimesh_removed_umap_gau <- trimesh_removed_umap_gau +
+  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  # theme(axis.text = element_text(size = 5),
+  #       axis.title = element_text(size = 7)) +
+  geom_point(colour = "#33a02c", size = 0.05) +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+# tour_umap_gau <- show_langevitour(df_all_umap_gau, df_bin_umap_gau, df_bin_centroids_umap_gau, benchmark_value = benchmark_umap_gau, distance = distance_umap_gau, distance_col = distance)
+
+## PAHTE
+
+# num_bins_phate_gau <- calculate_effective_x_bins(.data = PHATE_data_gau, x = PHATE1,
+#                                                  cell_area = 1)
+num_bins_phate_gau <- 70
+
+shape_val_phate_gau <- calculate_effective_shape_value(.data = PHATE_data_gau,
+                                                       x = PHATE1, y = PHATE2)
+## To extract bin centroids
+hexbin_data_object_phate_gau <- extract_hexbin_centroids(nldr_df = PHATE_data_gau, num_bins = num_bins_phate_gau, shape_val = shape_val_phate_gau, x = PHATE1, y = PHATE2)
+
+df_bin_centroids_phate_gau <- hexbin_data_object_phate_gau$hexdf_data
+
+PHATE_data_with_hb_id_gau <- PHATE_data_gau |>
+  dplyr::mutate(hb_id = hexbin_data_object_phate_gau$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_phate_gau <- dplyr::bind_cols(training_data_1 |> dplyr::select(-ID), PHATE_data_with_hb_id_gau)
+
+## Averaged on high-D
+df_bin_phate_gau <- avg_highD_data(.data = df_all_phate_gau)
+
+## Triangulate bin centroids
+tr1_object_phate_gau <- triangulate_bin_centroids(df_bin_centroids_phate_gau, x, y)
+tr_from_to_df_phate_gau <- generate_edge_info(triangular_object = tr1_object_phate_gau)
+
+## Compute 2D distances
+distance_phate_gau <- cal_2D_dist(.data = tr_from_to_df_phate_gau)
+
+## To find the benchmark value
+benchmark_phate_gau <- find_benchmark_value(.data = distance_phate_gau, distance_col = distance)
+benchmark_phate_gau <- 0.00078
+
+trimesh_removed_phate_gau <- remove_long_edges(.data = distance_phate_gau, benchmark_value = benchmark_phate_gau,
+                                               triangular_object = tr1_object_phate_gau, distance_col = distance)
+
+trimesh_removed_phate_gau <- trimesh_removed_phate_gau +
+  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  # theme(axis.text = element_text(size = 5),
+  #       axis.title = element_text(size = 7)) +
+  geom_point(colour = "#33a02c", size = 0.05) +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+# tour_phate_gau <- show_langevitour(df_all_phate_gau, df_bin_phate_gau, df_bin_centroids_phate_gau, benchmark_value = benchmark_phate_gau, distance = distance_phate_gau, distance_col = distance)
+
+## TriMAP
+
+# num_bins_trimap_gau <- calculate_effective_x_bins(.data = TriMAP_data_gau, x = TriMAP1,
+#                                                   cell_area = 1) #143
+
+num_bins_trimap_gau <- 21
+shape_val_trimap_gau <- calculate_effective_shape_value(.data = TriMAP_data_gau,
+                                                        x = TriMAP1, y = TriMAP2)
+## To extract bin centroids
+hexbin_data_object_trimap_gau <- extract_hexbin_centroids(nldr_df = TriMAP_data_gau, num_bins = num_bins_trimap_gau, shape_val = shape_val_trimap_gau, x = TriMAP1, y = TriMAP2)
+
+df_bin_centroids_trimap_gau <- hexbin_data_object_trimap_gau$hexdf_data
+
+TriMAP_data_with_hb_id_gau <- TriMAP_data_gau |>
+  dplyr::mutate(hb_id = hexbin_data_object_trimap_gau$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_trimap_gau <- dplyr::bind_cols(training_data_1 |> dplyr::select(-ID), TriMAP_data_with_hb_id_gau)
+
+## Averaged on high-D
+df_bin_trimap_gau <- avg_highD_data(.data = df_all_trimap_gau)
+
+## Triangulate bin centroids
+tr1_object_trimap_gau <- triangulate_bin_centroids(df_bin_centroids_trimap_gau, x, y)
+tr_from_to_df_trimap_gau <- generate_edge_info(triangular_object = tr1_object_trimap_gau)
+
+## Compute 2D distances
+distance_trimap_gau <- cal_2D_dist(.data = tr_from_to_df_trimap_gau)
+
+## To find the benchmark value
+benchmark_trimap_gau <- find_benchmark_value(.data = distance_trimap_gau, distance_col = distance)
+
+trimesh_removed_trimap_gau <- remove_long_edges(.data = distance_trimap_gau, benchmark_value = benchmark_trimap_gau,
+                                                triangular_object = tr1_object_trimap_gau, distance_col = distance)
+
+trimesh_removed_trimap_gau <- trimesh_removed_trimap_gau +
+  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  # theme(axis.text = element_text(size = 5),
+  #       axis.title = element_text(size = 7)) +
+  geom_point(colour = "#33a02c", size = 0.05) +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+# tour_trimap_gau <- show_langevitour(df_all_trimap_gau, df_bin_trimap_gau, df_bin_centroids_trimap_gau, benchmark_value = benchmark_trimap_gau, distance = distance_trimap_gau, distance_col = distance)
+
+## PaCMAP
+
+# num_bins_pacmap_gau <- calculate_effective_x_bins(.data = PaCMAP_data_gau, x = PaCMAP1,
+#                                                   cell_area = 1) #30
+num_bins_pacmap_gau <- 18
+
+shape_val_pacmap_gau <- calculate_effective_shape_value(.data = PaCMAP_data_gau,
+                                                        x = PaCMAP1, y = PaCMAP2)
+## To extract bin centroids
+hexbin_data_object_pacmap_gau <- extract_hexbin_centroids(nldr_df = PaCMAP_data_gau, num_bins = num_bins_pacmap_gau, shape_val = shape_val_pacmap_gau, x = PaCMAP1, y = PaCMAP2)
+
+df_bin_centroids_pacmap_gau <- hexbin_data_object_pacmap_gau$hexdf_data
+
+PaCMAP_data_with_hb_id_gau <- PaCMAP_data_gau |>
+  dplyr::mutate(hb_id = hexbin_data_object_pacmap_gau$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_pacmap_gau <- dplyr::bind_cols(training_data_1 |> dplyr::select(-ID), PaCMAP_data_with_hb_id_gau)
+
+## Averaged on high-D
+df_bin_pacmap_gau <- avg_highD_data(.data = df_all_pacmap_gau)
+
+## Triangulate bin centroids
+tr1_object_pacmap_gau <- triangulate_bin_centroids(df_bin_centroids_pacmap_gau, x, y)
+tr_from_to_df_pacmap_gau <- generate_edge_info(triangular_object = tr1_object_pacmap_gau)
+
+## Compute 2D distances
+distance_pacmap_gau <- cal_2D_dist(.data = tr_from_to_df_pacmap_gau)
+
+## To find the benchmark value
+benchmark_pacmap_gau <- find_benchmark_value(.data = distance_pacmap_gau, distance_col = distance)
+
+trimesh_removed_pacmap_gau <- remove_long_edges(.data = distance_pacmap_gau, benchmark_value = benchmark_pacmap_gau,
+                                                triangular_object = tr1_object_pacmap_gau, distance_col = distance)
+
+trimesh_removed_pacmap_gau <- trimesh_removed_pacmap_gau +
+  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  # theme(axis.text = element_text(size = 5),
+  #       axis.title = element_text(size = 7)) +
+  geom_point(colour = "#33a02c", size = 0.05) +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+# tour_pacmap_gau <- show_langevitour(df_all_pacmap_gau, df_bin_pacmap_gau, df_bin_centroids_pacmap_gau, benchmark_value = benchmark_pacmap_gau, distance = distance_pacmap_gau, distance_col = distance)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+#| label: fig-modelfiveGau
+#| fig-cap: Is there a best model to represent the original data in 2D space or are they all providing equivalent information?, (a) Model in the 2D space with tSNE (<https://youtu.be/uy9SkAo6gAE>), (b) Model in the 2D space with UMAP (<https://youtu.be/0MJDhHrh_Ug>), (c) Model in the 2D space with PHATE (<https://youtu.be/HbVv0uy0QWk>), (d) Model in the 2D space with TriMAP (<https://youtu.be/2OGwipAjzc8>), and (e) Model in the 2D space with PaCMAP (<https://youtu.be/pkokI8d-cBk>).
+#| out-width: 100%
+
+trimesh_removed_tsne_gau + trimesh_removed_umap_gau + trimesh_removed_phate_gau + trimesh_removed_trimap_gau + trimesh_removed_pacmap_gau +
+  #plot_annotation(tag_levels = 'a') +
+  plot_layout(ncol = 5) &
+  theme(legend.position='none', plot.tag = element_text(size = 8))
 ```
@@
 ## Import data
-df_2 <- read_rds("data/pbmc/pbmc.rds")
-training_data_1 <- read_rds("data/pbmc/pbmc_training.rds")
-test_1 <- read_rds("data/pbmc/pbmc_test.rds")
-
-tSNE_pbmc <- read_rds("data/pbmc/pbmc_tsne_30.rds")
-UMAP_pbmc <- read_rds("data/pbmc/pbmc_umap.rds")
-PHATE_pbmc <- read_rds("data/pbmc/pbmc_phate.rds")
-TriMAP_pbmc <- read_rds("data/pbmc/pbmc_trimap.rds")
-PaCMAP_pbmc <- read_rds("data/pbmc/pbmc_pacmap.rds")
+#df_2 <- read_rds("data/pbmc/pbmc.rds")
+training_data_1 <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_pca_50.rds")
+training_data_1 <- training_data_1[, 1:15]
+#test_1 <- read_rds("data/pbmc/pbmc_test.rds")
+
+#tSNE_pbmc <- read_rds("data/pbmc/pbmc_tsne_30.rds")
+UMAP_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_umap.rds")
+# PHATE_pbmc <- read_rds("data/pbmc/pbmc_phate.rds")
+# TriMAP_pbmc <- read_rds("data/pbmc/pbmc_trimap.rds")
+# PaCMAP_pbmc <- read_rds("data/pbmc/pbmc_pacmap.rds")
+
+load("data/pbmc/pbmc_3k_festem/pbmc3k_label.RData")
 
@@
 
-plot_list1_pbmc <- plot_tSNE_2D(tSNE_pbmc) + #ggtitle("(a)") + 
+# plot_list1_pbmc <- plot_tSNE_2D(tSNE_pbmc) + #ggtitle("(a)") + 
+#   theme_linedraw() +
+#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+#         axis.title.x = element_blank(), axis.title.y = element_blank(),
+#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+#   annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+class_avg <- UMAP_pbmc |>
+  group_by(cell_label) |>
+  summarise(
+    UMAP1 = median(UMAP1),
+    UMAP2 = median(UMAP2)
+  )
+
+class_avg$cell_label <- as.factor(class_avg$cell_label)
+
+levels(class_avg$cell_label) <- c("Memory \nCD4 T", "Naive CD4 T", "CD14+ Mono",  "B", "CD8 T", "FCGR3A+ \n Mono", "NK", "M-MDSC\n-like", "CD27-CD+ \n Memory T", "DC")
+
+plot_list2_pbmc <- UMAP_pbmc |>
+    ggplot(aes(x = UMAP1,
+               y = UMAP2, color = cell_label))+
+    geom_point(alpha=0.5) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
   theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-plot_list2_pbmc <- plot_UMAP_2D(UMAP_pbmc) + #ggtitle("(b)") + 
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-plot_list3_pbmc <- plot_PHATE_2D(PHATE_pbmc) + #ggtitle("(c)") + 
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-plot_list4_pbmc <- plot_TriMAP_2D(TriMAP_pbmc) + #ggtitle("(d)") + 
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-plot_list5_pbmc <- plot_PaCMAP_2D(PaCMAP_pbmc) + #ggtitle("(e)") + 
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'e', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "2D layouts from different NLDR techniques applied for the training PBMC dataset: (a) tSNE (perplexity = 30), (b) UMAP (n_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2).
-Is there a best representation of the original data or are they all providing  equivalent information?"
-#| label: fig-nldervis5PBMC
-#| out-width: 100%
-
-plot_list1_pbmc + plot_list2_pbmc + plot_list3_pbmc + plot_list4_pbmc + plot_list5_pbmc +
-  plot_layout(ncol=5)
-```
-
+    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#b15928", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#cab2d6", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#bc80bd", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
+  geom_text(aes(x=UMAP1,y = UMAP2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) #+
+  # guides(color = guide_legend(title = "Cell type"),
+  #        color = guide_legend(nrow = 3)) #+
+  # annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+# plot_list3_pbmc <- plot_PHATE_2D(PHATE_pbmc) + #ggtitle("(c)") + 
+#   theme_linedraw() +
+#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+#         axis.title.x = element_blank(), axis.title.y = element_blank(),
+#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+#   annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+# plot_list4_pbmc <- plot_TriMAP_2D(TriMAP_pbmc) + #ggtitle("(d)") + 
+#   theme_linedraw() +
+#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+#         axis.title.x = element_blank(), axis.title.y = element_blank(),
+#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+#   annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+# 
+# plot_list5_pbmc <- plot_PaCMAP_2D(PaCMAP_pbmc) + #ggtitle("(e)") + 
+#   theme_linedraw() +
+#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+#         axis.title.x = element_blank(), axis.title.y = element_blank(),
+#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+#   annotate(geom = 'text', label = 'e', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "2D layout from UMAP (n_neighbors = 30, min_dist = 0.3) applied for the PBMC3k dataset. Is this a best representation of the original data?"
+#| label: fig-nldervisPBMCUMAP
+
+plot_list2_pbmc
+
+#  plot_list1_pbmc + plot_list3_pbmc + plot_list4_pbmc + plot_list5_pbmc +
+#   plot_layout(ncol=5) #| out-width: 100%
+
+# 2D layouts from different NLDR techniques applied for the training PBMC dataset: (a) tSNE (perplexity = 30), (b) UMAP (n_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2).
+Is there a best representation of the original data or are they all providing  equivalent information?
+```
+
+<!--
 ```{r}
@@
 ```
+-->
 

### 2024-01-17 paper.tex
Regenerate PBMC3K UMAP results by following the same steps as the author in Literature

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, enhanced, interior hidden, breakable, sharp corners, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, frame hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden, enhanced]}{\end{tcolorbox}}\fi
 
@@
 
-{\centering \includegraphics{paper_files/figure-pdf/fig-modelScurve-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-modelScurve-1.pdf}
 
@@
 information?}
+
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-modelfiveGau-1.pdf}
+
+}
+
+\caption{\label{fig-modelfiveGau}Is there a best model to represent the
+original data in 2D space or are they all providing equivalent
+information?, (a) Model in the 2D space with tSNE
+(\url{https://youtu.be/uy9SkAo6gAE}), (b) Model in the 2D space with
+UMAP (\url{https://youtu.be/0MJDhHrh_Ug}), (c) Model in the 2D space
+with PHATE (\url{https://youtu.be/HbVv0uy0QWk}), (d) Model in the 2D
+space with TriMAP (\url{https://youtu.be/2OGwipAjzc8}), and (e) Model in
+the 2D space with PaCMAP (\url{https://youtu.be/pkokI8d-cBk}).}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervis5PBMC-1.pdf}
-
-}
-
-\caption{\label{fig-nldervis5PBMC}2D layouts from different NLDR
-techniques applied for the training PBMC dataset: (a) tSNE (perplexity =
-30), (b) UMAP (n\_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP
-(n\_inliers = 5, n\_outliers = 4, n\_random = 3), and (e) PaCMAP
-(n\_neighbors = 10, init = random, MN\_ratio = 0.9, FP\_ratio = 2). Is
-there a best representation of the original data or are they all
-providing equivalent information?}
-
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticpltPBMC-1.pdf}
-
-}
-
-\caption{\label{fig-diagnosticpltPBMC}Goodness of fit statistics from
-different NLDR techniques applied to training PBMC dataset. What is the
-best NLDR technique to represent the original data in 2D?}
+{\centering \includegraphics{paper_files/figure-pdf/fig-nldervisPBMCUMAP-1.pdf}
+
+}
+
+\caption{\label{fig-nldervisPBMCUMAP}2D layout from UMAP (n\_neighbors =
+30, min\_dist = 0.3) applied for the PBMC3k dataset. Is this a best
+representation of the original data?}
 

### 2024-01-17 paper.qmd
Added kang dataset example with tSNE, fix the error of getting PCAs in PBMC3k data set, complete the tour generation code for PBMC3k (need to record the video)

@@
 
+https://github.com/XiDsLab/Festem_paper/tree/main
+
+@Chen2023
+
 https://satijalab.org/seurat/articles/pbmc3k_tutorial.html
@@
 
-load("data/pbmc/pbmc_3k_festem/pbmc3k_label.RData")
+#load("data/pbmc/pbmc_3k_festem/pbmc3k_label.RData")
 
@@
     UMAP2 = median(UMAP2)
-  )
+  ) |>
+  mutate(cell_label = as.numeric(cell_label)) |>
+  arrange(cell_label)
 
@@
 
+### Kang dataset
+
+```{r}
+#| warning: false
+#| echo: false
+
+## Import data
+#df_2 <- read_rds("data/pbmc/pbmc.rds")
+# training_data_1 <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_pca_50.rds")
+# training_data_1 <- training_data_1[, 1:25]
+#test_1 <- read_rds("data/pbmc/pbmc_test.rds")
+
+#tSNE_pbmc <- read_rds("data/pbmc/pbmc_tsne_30.rds")
+tSNE_kang <- read_rds("data/kang/kang_tsne.rds") |>
+  mutate(cell_label = if_else(cell_label %in% c("0", "3", "5"), "0", cell_label))
+# PHATE_pbmc <- read_rds("data/pbmc/pbmc_phate.rds")
+# TriMAP_pbmc <- read_rds("data/pbmc/pbmc_trimap.rds")
+# PaCMAP_pbmc <- read_rds("data/pbmc/pbmc_pacmap.rds")
+
+#load("data/pbmc/pbmc_3k_festem/pbmc3k_label.RData")
+
+## Visualise embeddings
+
+# plot_list1_pbmc <- plot_tSNE_2D(tSNE_pbmc) + #ggtitle("(a)") + 
+#   theme_linedraw() +
+#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+#         axis.title.x = element_blank(), axis.title.y = element_blank(),
+#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+#   annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+class_avg <- tSNE_kang |>
+  group_by(cell_label) |>
+  summarise(
+    tSNE1 = median(tSNE1),
+    tSNE2 = median(tSNE2)
+  ) |>
+  mutate(cell_label = as.numeric(cell_label)) |>
+  arrange(cell_label)
+
+class_avg$cell_label <- as.factor(class_avg$cell_label)
+
+levels(class_avg$cell_label) <- c("CD14 Monocyte", "CD4 \n Memory T", "CD4 Naive T", "CD16 Monocyte", "B", "T Activated", "CD8 T", "NK", "DC", "B \nActivated", "T cell:\nmonocyte \nComplex", "HSP+ CD4 T", "IFN^hi CD14 \n Monocyte", "CD34 \n Progenitors", "pDC")
+
+plot_list2_kang <- tSNE_kang |>
+    ggplot(aes(x = tSNE1,
+               y = tSNE2, color = cell_label))+
+    geom_point(alpha=0.5) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#b15928", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#cab2d6", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#bc80bd", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
+  geom_text(aes(x=tSNE1,y = tSNE2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) #+
+  # guides(color = guide_legend(title = "Cell type"),
+  #        color = guide_legend(nrow = 3)) #+
+  # annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+# plot_list3_pbmc <- plot_PHATE_2D(PHATE_pbmc) + #ggtitle("(c)") + 
+#   theme_linedraw() +
+#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+#         axis.title.x = element_blank(), axis.title.y = element_blank(),
+#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+#   annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+# plot_list4_pbmc <- plot_TriMAP_2D(TriMAP_pbmc) + #ggtitle("(d)") + 
+#   theme_linedraw() +
+#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+#         axis.title.x = element_blank(), axis.title.y = element_blank(),
+#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+#   annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+# 
+# plot_list5_pbmc <- plot_PaCMAP_2D(PaCMAP_pbmc) + #ggtitle("(e)") + 
+#   theme_linedraw() +
+#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+#         axis.title.x = element_blank(), axis.title.y = element_blank(),
+#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+#   annotate(geom = 'text', label = 'e', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "2D layout from tSNE (perplexity = 30) applied for the Kang dataset. Is this a best representation of the original data?"
+#| label: fig-nlderviskangtSNE
+
+plot_list2_kang
+
+#  plot_list1_pbmc + plot_list3_pbmc + plot_list4_pbmc + plot_list5_pbmc +
+#   plot_layout(ncol=5) #| out-width: 100%
+
+# 2D layouts from different NLDR techniques applied for the training PBMC dataset: (a) tSNE (perplexity = 30), (b) UMAP (n_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2).
+Is there a best representation of the original data or are they all providing  equivalent information?
+```
+
 ## Conclusion {#sec-conclusion}

### 2024-01-17 paper.tex
Added kang dataset example with tSNE, fix the error of getting PCAs in PBMC3k data set, complete the tour generation code for PBMC3k (need to record the video)

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, frame hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden, breakable, sharp corners, enhanced]}{\end{tcolorbox}}\fi
 
@@
 
+https://github.com/XiDsLab/Festem\_paper/tree/main
+
+\citet{Chen2023}
+
 https://satijalab.org/seurat/articles/pbmc3k\_tutorial.html
@@
 
+\hypertarget{kang-dataset}{%
+\subsection{Kang dataset}\label{kang-dataset}}
+
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-nlderviskangtSNE-1.pdf}
+
+}
+
+\caption{\label{fig-nlderviskangtSNE}2D layout from tSNE (perplexity =
+30) applied for the Kang dataset. Is this a best representation of the
+original data?}
+
+\end{figure}
+
 \hypertarget{sec-conclusion}{%

### 2024-01-17 applications/pbmc_3K_Festem/1_obtain_pca_umap.R
Add tour implemenation for kang data sets with necessary parameters (need to record the video)

@@ -1,0 +1,94 @@
+library(Seurat)
+library(ggplot2)
+library(ggpubr)
+library(tidyverse)
+library(scales)
+
+#################################
+load("data/pbmc/pbmc_3k_festem/pbmc3k_clustering_UMAP.RData")
+pbmc_umap <- umap.list$Festem@cell.embeddings |>
+  tibble::as_tibble() |>
+  mutate(cell_label = as.character(label.list$Festem)) |>
+  rename(c("UMAP1" = "UMAP_1", "UMAP2" = "UMAP_2"))
+
+
+write_rds(pbmc_umap, "data/pbmc/pbmc_3k_festem/pbmc_umap.rds")
+
+rownames(umap.list$Festem@cell.embeddings)
+plots.list[[1]]
+
+## Obtain gene names
+rw_names <- rownames(umap.list$Festem@cell.embeddings) |> str_split("-")
+
+name_vec <- c()
+for (i in 1:length(rw_names)) {
+  name <- rw_names[[i]][1]
+  name_vec <- append(name_vec, name)
+
+}
+
+####################PCA
+# library(RColorBrewer)
+# my.color <- hue_pal()(13)
+# names(my.color) <- 1:13
+# umap.for.plot <- function(umap,cluster){
+#   umap <- umap@cell.embeddings
+#   umap <- as.data.frame(umap)
+#   cbind(umap,cluster = cluster)
+# }
+load("data/pbmc/pbmc_3k_festem/pbmc3k_hvggenes.RData")
+pbmc <- readRDS("data/pbmc/pbmc_3k_festem/pbmc3k_final.rds")
+pbmc <- UpdateSeuratObject(pbmc)
+# ref <- pbmc@active.ident
+# names_fac <- names(pbmc@active.ident)
+#
+# ref <- as.character(ref)
+# ref[c("CGGGCATGACCCAA","CTTGATTGATCTTC")] <- "Platelet"
+# ref <- as.factor(ref)
+# names(ref) <- names_fac
+
+#pbmc <- pbmc[,!colnames(pbmc) %in% c("CGGGCATGACCCAA","CTTGATTGATCTTC")]
+pbmc <- pbmc[,colnames(pbmc) %in% name_vec]
+
+#pbmc <- pbmc[,ref!="Platelet"] ## Didn't run
+gene.list <- list(EM[1:1000])
+# ,
+# hvgvst[1:1000],
+# hvgdisp[1:1000],
+# dub,
+# devianceFS[1:1000],
+# trendvar[1:1000]
+
[…]

### 2024-01-18 paper.qmd
Add tour implemenation for kang data sets with necessary parameters (need to record the video)
Added video links for PBMC3k and Five Guassian clusters

@@
 #| label: fig-modelfiveGau
-#| fig-cap: Is there a best model to represent the original data in 2D space or are they all providing equivalent information?, (a) Model in the 2D space with tSNE (<https://youtu.be/uy9SkAo6gAE>), (b) Model in the 2D space with UMAP (<https://youtu.be/0MJDhHrh_Ug>), (c) Model in the 2D space with PHATE (<https://youtu.be/HbVv0uy0QWk>), (d) Model in the 2D space with TriMAP (<https://youtu.be/2OGwipAjzc8>), and (e) Model in the 2D space with PaCMAP (<https://youtu.be/pkokI8d-cBk>).
+#| fig-cap: Is there a best model to represent the original data in 2D space or are they all providing equivalent information?, (a) Model in the 2D space with tSNE (<https://youtu.be/x5VPB-wOLm4>), (b) Model in the 2D space with UMAP (<https://youtu.be/Cm1dW6iLyHQ>), (c) Model in the 2D space with PHATE (<https://youtu.be/HB0Y3ilz6sI>), (d) Model in the 2D space with TriMAP (<https://youtu.be/iSVkXIrOrSA>), and (e) Model in the 2D space with PaCMAP (<https://youtu.be/O2QBctsc4uM>).
 #| out-width: 100%
@@
 
-### Single-cell RNA-seq data of human
-
-https://github.com/XiDsLab/Festem_paper/tree/main
-
-@Chen2023
-
-https://satijalab.org/seurat/articles/pbmc3k_tutorial.html
-https://www.cell.com/fulltext/S0092-8674(15)00549-8
-
-In the field of single-cell studies, a common analysis task involves clustering to identify groups of cells with similar expression profiles.
-Analysts often turn to NLDR techniques to verify and identify these clusters and explore developmental trajectories (e.g., example 1).
-In clustering workflows, the main objective is to verify the existence of clusters and subsequently identify them as specific cell types by examining the expression of "known" marker genes.
-In this context, a "faithful" embedding should ideally preserve the topology of the data, ensuring that cells corresponding to the same cell type are situated close to the high-dimensional space.
-
-To begin our analysis, we installed the Peripheral Blood Mononuclear Cells (pbmc) data set obtained from 10x Genomics using the `SeuratData` R package [@Rahul2019], which facilitates the distribution of data sets in the form of Seurat objects [@Yuhan2021].
-This data set contains 13,714 features across 2,700 samples within a single assay.
-The active assay is RNA, with 13,714 features representing different gene expressions.
-After loading the data set, we obtained the principal components (PCs) and assessed the variance explained by each PC.
-Based on this evaluation, we selected seven PCs, representing approximately 50% of the variance in gene expression, for further analysis.
-
-  Next, we employed the UMAP technique with default parameter settings.
-As illustrated in @fig-pbmc, the cell types B and Platelet are well-separated in the UMAP layout.
-Moreover, CD14+ Mono, FCGR3A+ Mono, and DC form a distinct cluster, while Naive CD4 T, NK, Memory CD4 T, and CD8 T are grouped together in another cluster.
-The values utilized to construct the smooth low-dimensional manifold are presented in @tbl-table02.
-The linked video, demonstrating the tour with the model, showcases the generation of a smooth surface for this application, enabling a comprehensive exploration of the data's structure and relationships (see @fig-pbmc_sc).
+### Human peripheral blood mononclear cells (PBMCs) data
+
+- Intro to the data set (This data set contains 13,714 features across 2,700 samples within a single assay. The active assay is RNA, with 13,714 features representing different gene expressions.)
+
+- 2622 cells
+
+- Evaluate Festem on DEG detection (Festem enabled identification of often-missed fine cell types)
+
+- Using the Festem-selected genes, identified 10 clusters in the PBMCsk data and annotated them based on the expression of canonical markers
+
+- The 10 clusters included immune cells such as naive CD4 T cells, memory CD4 T cells, CD8 T cells, CD14 monocytes, FCGR3A monocytes, Natural Killer (NK) cells, B cells and Dendritic Cells (DC).
+
+- In addition to these common cell types, Festem identified a fine cell type, CD27− CD4+ memory T cells, which were often missed by other methods.
+
+- The CD27. CD4+ memory T cells identified by Festem expressed common marker genes (IL7R
+and S100A4) of memory T cells, but did not express CD27. These cells also had downregulated
+expression of SELL, CCR7, MAL and LEF1, and upregulated expression of CCL5 (Fig. 4B) and
+thus were the CD27- CD4+ memory T cells in the literature (36). The CD27- CD4+ memory T
+cells were known to be at a more differentiated state and have stronger antigen-recall responses
+than their CD27+ counterparts.
 
@@
 #df_2 <- read_rds("data/pbmc/pbmc.rds")
-training_data_1 <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_pca_50.rds")
-training_data_1 <- training_data_1[, 1:15]
+training_data_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_pca_50.rds")
+training_data_pbmc <- training_data_pbmc[, 1:15] |>
+  mutate(ID = 1:NROW(training_data_pbmc))
 #test_1 <- read_rds("data/pbmc/pbmc_test.rds")
@@
          legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#b15928", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#cab2d6", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#bc80bd", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
-  geom_text(aes(x=UMAP1,y = UMAP2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) #+
+  scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
+  geom_text(aes(x=UMAP1,y = UMAP2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) #+
   # guides(color = guide_legend(title = "Cell type"),
@@
 ```{r}
-#| echo: false
-#| fig-cap: "2D layout from UMAP (n_neighbors = 30, min_dist = 0.3) applied for the PBMC3k dataset. Is this a best representation of the original data?"
+#| warning: false
+#| echo: false
+num_bins_pbmc <- calculate_effective_x_bins(.data = UMAP_pbmc, x = UMAP1,
+                                              cell_area = 1) ##23
+
+shape_value_pbmc <- calculate_effective_shape_value(.data = UMAP_pbmc,
+                                                      x = UMAP1, y = UMAP2) ## 0.8772751
+
+## To extract bin centroids
+hexbin_data_object_pbmc <- extract_hexbin_centroids(nldr_df = UMAP_pbmc,
+                                                      num_bins = num_bins_pbmc,
+                                                      shape_val = shape_value_pbmc, x = UMAP1, y = UMAP2)
+
+df_bin_centroids_pbmc <- hexbin_data_object_pbmc$hexdf_data
+
+UMAP_pbmc_with_hb_id <- UMAP_pbmc |>
+  dplyr::mutate(hb_id = hexbin_data_object_pbmc$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), UMAP_pbmc_with_hb_id)
+
+## Averaged on high-D
+df_bin_pbmc <- avg_highD_data(.data = df_all_pbmc, column_start_text = "PC")
+
+## Triangulate bin centroids
+tr1_object_pbmc <- triangulate_bin_centroids(df_bin_centroids_pbmc, x, y)
+tr_from_to_df_pbmc <- generate_edge_info(triangular_object = tr1_object_pbmc)
+
+## Compute 2D distances
+distance_pbmc <- cal_2D_dist(.data = tr_from_to_df_pbmc)
+
+## To find the benchmark value
+benchmark_pbmc <- find_benchmark_value(.data = distance_pbmc, distance_col = distance)
+
+trimesh_removed_pbmc_umap <- remove_long_edges(.data = distance_pbmc, benchmark_value = benchmark_pbmc,
+                                                 triangular_object = tr1_object_pbmc, distance_col = distance)
+
+trimesh_removed_pbmc_umap <- trimesh_removed_pbmc_umap +
+  geom_point(size = 2, colour = "#33a02c") +
+  #ggtitle("(b)") +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "(a) 2D layout from UMAP (n_neighbors = 30, min_dist = 0.3) applied for the PBMC3k dataset.
+Is this a best representation of the original data?, (b)  Model in the 2D space with UMAP (<https://youtu.be/Ia3aL3L-fCk>)"
 #| label: fig-nldervisPBMCUMAP
-
-plot_list2_pbmc
+ #| out-width: 100%
+
+plot_list2_pbmc + trimesh_removed_pbmc_umap + 
+plot_layout(ncol=2)
 
@@
 <!--
+### Single-cell RNA-seq data of human
+
+https://github.com/XiDsLab/Festem_paper/tree/main
+
+@Chen2023
+
+https://satijalab.org/seurat/articles/pbmc3k_tutorial.html
+https://www.cell.com/fulltext/S0092-8674(15)00549-8
+
+In the field of single-cell studies, a common analysis task involves clustering to identify groups of cells with similar expression profiles.
+Analysts often turn to NLDR techniques to verify and identify these clusters and explore developmental trajectories (e.g., example 1).
+In clustering workflows, the main objective is to verify the existence of clusters and subsequently identify them as specific cell types by examining the expression of "known" marker genes.
+In this context, a "faithful" embedding should ideally preserve the topology of the data, ensuring that cells corresponding to the same cell type are situated close to the high-dimensional space.
+
+To begin our analysis, we installed the Peripheral Blood Mononuclear Cells (pbmc) data set obtained from 10x Genomics using the `SeuratData` R package [@Rahul2019], which facilitates the distribution of data sets in the form of Seurat objects [@Yuhan2021].
+This data set contains 13,714 features across 2,700 samples within a single assay.
+The active assay is RNA, with 13,714 features representing different gene expressions.
+After loading the data set, we obtained the principal components (PCs) and assessed the variance explained by each PC.
+Based on this evaluation, we selected seven PCs, representing approximately 50% of the variance in gene expression, for further analysis.
+
+Next, we employed the UMAP technique with default parameter settings.
+As illustrated in @fig-pbmc, the cell types B and Platelet are well-separated in the UMAP layout.
+Moreover, CD14+ Mono, FCGR3A+ Mono, and DC form a distinct cluster, while Naive CD4 T, NK, Memory CD4 T, and CD8 T are grouped together in another cluster.
+The values utilized to construct the smooth low-dimensional manifold are presented in @tbl-table02.
+The linked video, demonstrating the tour with the model, showcases the generation of a smooth surface for this application, enabling a comprehensive exploration of the data's structure and relationships (see @fig-pbmc_sc).
+
 ```{r}
@@
 
+Human retina: https://www.ebi.ac.uk/biostudies/arrayexpress/studies/E-MTAB-7316
+
 https://www.cell.com/cms/10.1016/j.cell.2015.05.002/attachment/fd52950c-fd35-401a-9e60-ca7588ef9001/mmc1.pdf
@@
 ### Kang dataset
+
+For the Kang lupus dataset (29), we obtained 16 cell types based on clustering of the Festem-
+selected genes (Fig. 4E and Fig. S10, S11), including 2 fine cell types that were often missed by
+other methods. The 2 cell types were IFNhi CD14+ monocytes with high expression of type I
+interferon-related (IFN) genes, and HSP+ CD4 T cells with up-regulated heat-
+shock-protein-related (HSP) genes. The HSP+ CD4 T cells might result from
+cryopreservation and thawing of T cells (37). Other than devianceFS, other available methods
+could not identify these HSP+ CD4 T cells (Fig. S11).
 

### 2024-01-18 paper.tex
Added video links for PBMC3k and Five Guassian clusters

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden, breakable, sharp corners, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, borderline west={3pt}{0pt}{shadecolor}, breakable, enhanced, interior hidden, boxrule=0pt, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 information?, (a) Model in the 2D space with tSNE
-(\url{https://youtu.be/uy9SkAo6gAE}), (b) Model in the 2D space with
-UMAP (\url{https://youtu.be/0MJDhHrh_Ug}), (c) Model in the 2D space
-with PHATE (\url{https://youtu.be/HbVv0uy0QWk}), (d) Model in the 2D
-space with TriMAP (\url{https://youtu.be/2OGwipAjzc8}), and (e) Model in
-the 2D space with PaCMAP (\url{https://youtu.be/pkokI8d-cBk}).}
+(\url{https://youtu.be/x5VPB-wOLm4}), (b) Model in the 2D space with
+UMAP (\url{https://youtu.be/Cm1dW6iLyHQ}), (c) Model in the 2D space
+with PHATE (\url{https://youtu.be/HB0Y3ilz6sI}), (d) Model in the 2D
+space with TriMAP (\url{https://youtu.be/iSVkXIrOrSA}), and (e) Model in
+the 2D space with PaCMAP (\url{https://youtu.be/O2QBctsc4uM}).}
 
@@
 
-\hypertarget{single-cell-rna-seq-data-of-human}{%
-\subsection{Single-cell RNA-seq data of
-human}\label{single-cell-rna-seq-data-of-human}}
-
-https://github.com/XiDsLab/Festem\_paper/tree/main
-
-\citet{Chen2023}
-
-https://satijalab.org/seurat/articles/pbmc3k\_tutorial.html
-https://www.cell.com/fulltext/S0092-8674(15)00549-8
-
-In the field of single-cell studies, a common analysis task involves
-clustering to identify groups of cells with similar expression profiles.
-Analysts often turn to NLDR techniques to verify and identify these
-clusters and explore developmental trajectories (e.g., example 1). In
-clustering workflows, the main objective is to verify the existence of
-clusters and subsequently identify them as specific cell types by
-examining the expression of ``known'' marker genes. In this context, a
-``faithful'' embedding should ideally preserve the topology of the data,
-ensuring that cells corresponding to the same cell type are situated
-close to the high-dimensional space.
-
-To begin our analysis, we installed the Peripheral Blood Mononuclear
-Cells (pbmc) data set obtained from 10x Genomics using the
-\texttt{SeuratData} R package \citep{Rahul2019}, which facilitates the
-distribution of data sets in the form of Seurat objects
-\citep{Yuhan2021}. This data set contains 13,714 features across 2,700
-samples within a single assay. The active assay is RNA, with 13,714
-features representing different gene expressions. After loading the data
-set, we obtained the principal components (PCs) and assessed the
-variance explained by each PC. Based on this evaluation, we selected
-seven PCs, representing approximately 50\% of the variance in gene
-expression, for further analysis.
-
-Next, we employed the UMAP technique with default parameter settings. As
-illustrated in \textbf{?@fig-pbmc}, the cell types B and Platelet are
-well-separated in the UMAP layout. Moreover, CD14+ Mono, FCGR3A+ Mono,
-and DC form a distinct cluster, while Naive CD4 T, NK, Memory CD4 T, and
-CD8 T are grouped together in another cluster. The values utilized to
-construct the smooth low-dimensional manifold are presented in
-\textbf{?@tbl-table02}. The linked video, demonstrating the tour with
-the model, showcases the generation of a smooth surface for this
-application, enabling a comprehensive exploration of the data's
-structure and relationships (see \textbf{?@fig-pbmc\_sc}).
+\hypertarget{human-peripheral-blood-mononclear-cells-pbmcs-data}{%
+\subsection{Human peripheral blood mononclear cells (PBMCs)
+data}\label{human-peripheral-blood-mononclear-cells-pbmcs-data}}
+
+\begin{itemize}
+\item
+  Intro to the data set (This data set contains 13,714 features across
+  2,700 samples within a single assay. The active assay is RNA, with
+  13,714 features representing different gene expressions.)
+\item
+  2622 cells
+\item
+  Evaluate Festem on DEG detection (Festem enabled identification of
+  often-missed fine cell types)
+\item
+  Using the Festem-selected genes, identified 10 clusters in the PBMCsk
+  data and annotated them based on the expression of canonical markers
+\item
+  The 10 clusters included immune cells such as naive CD4 T cells,
+  memory CD4 T cells, CD8 T cells, CD14 monocytes, FCGR3A monocytes,
+  Natural Killer (NK) cells, B cells and Dendritic Cells (DC).
+\item
+  In addition to these common cell types, Festem identified a fine cell
+  type, CD27− CD4+ memory T cells, which were often missed by other
+  methods.
+\item
+  The CD27. CD4+ memory T cells identified by Festem expressed common
+  marker genes (IL7R and S100A4) of memory T cells, but did not express
+  CD27. These cells also had downregulated expression of SELL, CCR7, MAL
+  and LEF1, and upregulated expression of CCL5 (Fig. 4B) and thus were
+  the CD27- CD4+ memory T cells in the literature (36). The CD27- CD4+
+  memory T cells were known to be at a more differentiated state and
+  have stronger antigen-recall responses than their CD27+ counterparts.
+\end{itemize}
 
@@
 
-\caption{\label{fig-nldervisPBMCUMAP}2D layout from UMAP (n\_neighbors =
-30, min\_dist = 0.3) applied for the PBMC3k dataset. Is this a best
-representation of the original data?}
+\caption{\label{fig-nldervisPBMCUMAP}(a) 2D layout from UMAP
+(n\_neighbors = 30, min\_dist = 0.3) applied for the PBMC3k dataset. Is
+this a best representation of the original data?, (b) Model in the 2D
+space with UMAP (\url{https://youtu.be/Ia3aL3L-fCk})}
 
@@
 
+Human retina:
+https://www.ebi.ac.uk/biostudies/arrayexpress/studies/E-MTAB-7316
+
 https://www.cell.com/cms/10.1016/j.cell.2015.05.002/attachment/fd52950c-fd35-401a-9e60-ca7588ef9001/mmc1.pdf
@@
 \subsection{Kang dataset}\label{kang-dataset}}
+
+For the Kang lupus dataset (29), we obtained 16 cell types based on
+clustering of the Festem- selected genes (Fig. 4E and Fig. S10, S11),
+including 2 fine cell types that were often missed by other methods. The
+2 cell types were IFNhi CD14+ monocytes with high expression of type I
+interferon-related (IFN) genes, and HSP+ CD4 T cells with up-regulated
+heat- shock-protein-related (HSP) genes. The HSP+ CD4 T cells might
+result from cryopreservation and thawing of T cells (37). Other than
+devianceFS, other available methods could not identify these HSP+ CD4 T
+cells (Fig. S11).
 

### 2024-01-18 paper.qmd
Add tSNE view of HARPS star data and videos with absolute error values for PBMC3k and HARPS data

@@
 
-### Tunning the model
+### Tuning the model
 
@@
 
+- learned from the model: There are 3 well-separated clusters in 2D, but if look at the model in high-D space, the three clusters are much closer.
+Also, there is some continuity within the clustering structure and it didn't capture well.
+Therefore UMAP with n:neighbour 30 is not a good representation for this data set.
+
 ```{r}
@@
 #| fig-cap: "(a) 2D layout from UMAP (n_neighbors = 30, min_dist = 0.3) applied for the PBMC3k dataset.
-Is this a best representation of the original data?, (b)  Model in the 2D space with UMAP (<https://youtu.be/Ia3aL3L-fCk>)"
+Is this a best representation of the original data?, (b)  Model in the 2D space with UMAP (<https://youtu.be/lIA6ZPZAcLE>)"
 #| label: fig-nldervisPBMCUMAP
@@
 
+### HARPS GTO sample dataset
+
+- In 2D, there are two well-separated clusters
+
+- One cluster have a sparse data space within the cluster
+
+- There is a line in right corner of the structure that is not connected any other trimesh points. This line create a long edge in high-D space.
+
+- In high-D view, the clusters having some non-linearity patter, which is also not captured in tSNE. Therefore, this is not a good choice.
+
+```{r}
+#| warning: false
+#| echo: false
+
+tSNE_data_harps <- read_rds("data/harps/harps_tsne_40.rds")
+
+plot_list2_harps <- tSNE_data_harps |>
+    ggplot(aes(x = tSNE1,
+               y = tSNE2, color = cluster))+
+    geom_point(alpha=0.5) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#b15928", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#cab2d6", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#bc80bd", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
+  guides(color = guide_legend(title = "cluster"),
+         color = guide_legend(nrow = 3)) #+
+  # annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## Import data
+training_data_1 <- read_rds("data/harps/harps_data.rds")
+training_data_1 <- training_data_1 |>
+  mutate(ID = 1:NROW(training_data_1))
+
+clusters <- training_data_1$cluster
+select_names <-  names(training_data_1)[-c(18, length(training_data_1))]
+
+training_data_1 <- training_data_1 |> select(-cluster)
+names(training_data_1) <- append(paste0("x", 1:17), "ID")
+
+tSNE_data_harps <- tSNE_data_harps |>
+  select(tSNE1, tSNE2) |>
+  mutate(ID = 1:NROW(tSNE_data_harps))
+
+### tSNE
+
+num_bins_tsne_harps <- calculate_effective_x_bins(.data = tSNE_data_harps, x = tSNE1,
+                                                cell_area = 1) #43
+
+shape_val_tsne_harps <- calculate_effective_shape_value(.data = tSNE_data_harps,
+                                                      x = tSNE1, y = tSNE2)
+## To extract bin centroids
+hexbin_data_object_tsne_harps <- extract_hexbin_centroids(nldr_df = tSNE_data_harps, num_bins = num_bins_tsne_harps, shape_val = shape_val_tsne_harps, x = tSNE1, y = tSNE2)
+
+df_bin_centroids_tsne_harps <- hexbin_data_object_tsne_harps$hexdf_data
+
+## Identify bins with low-density
+# identify_rm_bins <- find_low_density_hexagons(df_bin_centroids_tsne_harps, num_bins_tsne_harps, benchmark_rm_hex = 0.01)
+#
+# df_bin_centroids_tsne_harps <- df_bin_centroids_tsne_harps |>
+#   filter(!(hexID %in% identify_rm_bins))
+
+tSNE_data_with_hb_id_harps <- tSNE_data_harps |>
+  dplyr::mutate(hb_id = hexbin_data_object_tsne_harps$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_tsne_harps <- dplyr::bind_cols(training_data_1 |> dplyr::select(-ID), tSNE_data_with_hb_id_harps)
+
+## Averaged on high-D
+df_bin_tsne_harps <- avg_highD_data(.data = df_all_tsne_harps)
+
+## Triangulate bin centroids
+tr1_object_tsne_harps <- triangulate_bin_centroids(df_bin_centroids_tsne_harps, x, y)
+tr_from_to_df_tsne_harps <- generate_edge_info(triangular_object = tr1_object_tsne_harps)
+
+## Compute 2D distances
+distance_tsne_harps <- cal_2D_dist(.data = tr_from_to_df_tsne_harps)
+
+## To find the benchmark value
+benchmark_tsne_harps <- find_benchmark_value(.data = distance_tsne_harps, distance_col = distance)
+
+trimesh_removed_tsne_harps <- remove_long_edges(.data = distance_tsne_harps, benchmark_value = benchmark_tsne_harps,
+                                              triangular_object = tr1_object_tsne_harps, distance_col = distance)
+
+trimesh_removed_tsne_harps <- trimesh_removed_tsne_harps +
+  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  # theme(axis.text = element_text(size = 5),
+  #       axis.title = element_text(size = 7)) +
+  geom_point(colour = "#33a02c", size = 0.5) +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "(a) 2D layout from tSNE (perplexity = 40) applied for the HARPS GTO dataset.
+Is this a best representation of the original data?, (b) Model in the 2D space with tSNE (<https://youtu.be/txTD7GTxt6Y>)"
+#| label: fig-nldervisharpstSNE
+
+plot_list2_harps + trimesh_removed_tsne_harps +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='bottom', plot.tag = element_text(size = 8))
+
+```
+
+<!--
 ### Single-Cell Tagged Reverse Transcription sequencing data of mouse

### 2024-01-18 paper.tex
Add tSNE view of HARPS star data and videos with absolute error values for PBMC3k and HARPS data

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, borderline west={3pt}{0pt}{shadecolor}, breakable, enhanced, interior hidden, boxrule=0pt, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, frame hidden, interior hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, sharp corners, enhanced]}{\end{tcolorbox}}\fi
 
@@
 
-\hypertarget{tunning-the-model}{%
-\subsection{Tunning the model}\label{tunning-the-model}}
+\hypertarget{tuning-the-model}{%
+\subsection{Tuning the model}\label{tuning-the-model}}
 
@@
   have stronger antigen-recall responses than their CD27+ counterparts.
+\item
+  learned from the model: There are 3 well-separated clusters in 2D, but
+  if look at the model in high-D space, the three clusters are much
+  closer. Also, there is some continuity within the clustering structure
+  and it didn't capture well. Therefore UMAP with n:neighbour 30 is not
+  a good representation for this data set.
 \end{itemize}
@@
 this a best representation of the original data?, (b) Model in the 2D
-space with UMAP (\url{https://youtu.be/Ia3aL3L-fCk})}
-
-\end{figure}
-
-\hypertarget{single-cell-tagged-reverse-transcription-sequencing-data-of-mouse}{%
-\subsection{Single-Cell Tagged Reverse Transcription sequencing data of
-mouse}\label{single-cell-tagged-reverse-transcription-sequencing-data-of-mouse}}
-
-The Zeisel mouse brain dataset, obtained through Spatial Transcriptomics
-(STRT-Seq). Within this dataset, information is collected from a
-substantial 2,816 individual mouse brain cells. Each of these cells acts
-as a molecular snapshot, capturing the distinctive genetic activity
-within various cell types. This diversity spans neurons, glial cells,
-and other essential components of the brain, offering a comprehensive
-view of the cellular tapestry.
-
-What makes this dataset particularly valuable is its ability to shed
-light on the spatial distribution of cells. Researchers can explore how
-gene expression patterns vary across different regions of the mouse
-brain, unlocking insights into the functional specialization of these
-regions and the intricate networks that underpin neural processes.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervis5Mouse-1.pdf}
-
-}
-
-\caption{\label{fig-nldervis5Mouse}2D layouts from different NLDR
-techniques applied for the training Zeisel mouse brain dataset: (a) tSNE
-(perplexity = 30), (b) UMAP (n\_neighbors = 15), (c) PHATE (knn = 5),
-(d) TriMAP (n\_inliers = 5, n\_outliers = 4, n\_random = 3), and (e)
-PaCMAP (n\_neighbors = 10, init = random, MN\_ratio = 0.9, FP\_ratio =
-2). Is there a best representation of the original data or are they all
-providing equivalent information?}
-
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticpltZEI-1.pdf}
-
-}
-
-\caption{\label{fig-diagnosticpltZEI}Goodness of fit statistics from
-different NLDR techniques applied to training Zeisel mouse brain
-dataset. What is the best NLDR technique to represent the original data
-in 2D?}
-
-\end{figure}
-
-\hypertarget{retina-dataset}{%
-\subsection{Retina dataset}\label{retina-dataset}}
-
-Human retina:
-https://www.ebi.ac.uk/biostudies/arrayexpress/studies/E-MTAB-7316
-
-https://www.cell.com/cms/10.1016/j.cell.2015.05.002/attachment/fd52950c-fd35-401a-9e60-ca7588ef9001/mmc1.pdf
-
-https://www.cell.com/fulltext/S0092-8674(15)00549-8\#app3
-
-32 PCs tSNE (perplexity 30)
-
-\hypertarget{kang-dataset}{%
-\subsection{Kang dataset}\label{kang-dataset}}
-
-For the Kang lupus dataset (29), we obtained 16 cell types based on
-clustering of the Festem- selected genes (Fig. 4E and Fig. S10, S11),
-including 2 fine cell types that were often missed by other methods. The
-2 cell types were IFNhi CD14+ monocytes with high expression of type I
-interferon-related (IFN) genes, and HSP+ CD4 T cells with up-regulated
-heat- shock-protein-related (HSP) genes. The HSP+ CD4 T cells might
-result from cryopreservation and thawing of T cells (37). Other than
-devianceFS, other available methods could not identify these HSP+ CD4 T
-cells (Fig. S11).
-
-\begin{figure}
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-nlderviskangtSNE-1.pdf}
-
-}
-
-\caption{\label{fig-nlderviskangtSNE}2D layout from tSNE (perplexity =
-30) applied for the Kang dataset. Is this a best representation of the
-original data?}
+space with UMAP (\url{https://youtu.be/lIA6ZPZAcLE})}
+
+\end{figure}
+
+\hypertarget{harps-gto-sample-dataset}{%
+\subsection{HARPS GTO sample dataset}\label{harps-gto-sample-dataset}}
+
+\begin{itemize}
+\item
+  In 2D, there are two well-separated clusters
+\item
+  One cluster have a sparse data space within the cluster
+\item
+  There is a line in right corner of the structure that is not connected
+  any other trimesh points. This line create a long edge in high-D
+  space.
+\item
+  In high-D view, the clusters having some non-linearity patter, which
+  is also not captured in tSNE. Therefore, this is not a good choice.
+\end{itemize}
+
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-nldervisharpstSNE-1.pdf}
+
+}
+
+\caption{\label{fig-nldervisharpstSNE}(a) 2D layout from tSNE
+(perplexity = 40) applied for the HARPS GTO dataset. Is this a best
+representation of the original data?, (b) Model in the 2D space with
+tSNE (\url{https://youtu.be/txTD7GTxt6Y})}
 

### 2024-01-19 paper.qmd
Regenerated Figure 4 by comments

@@
 
-benchmark_value <- 19.4
+## Map hb_id
+
+df_bin_n <- df_bin |>
+  mutate(ID = 1:NROW(df_bin))
+
+tr_from_to_df_n <- left_join(tr_from_to_df, df_bin_n, by = c("from" = "ID")) |>
+  rename("from_hb_id" = "hb_id")
+
+tr_from_to_df_n <- left_join(tr_from_to_df_n, df_bin_n, by = c("to" = "ID")) |>
+  rename("to_hb_id" = "hb_id")
+
+hb_id_selected <- tr_from_to_df_n |>
+  dplyr::filter((from == 2) & (to == 3)) |>
+  head(1)
+
+# benchmark_value <- 19.4
 
@@
 distance_df_small_edges <- distance |>
-  dplyr::filter(distance < benchmark_value) |>
+  dplyr::filter((from != 2) | (to != 3)) |>
   dplyr::mutate(type = "small_edges")
@@
 distance_df_long_edges <- distance |>
-  dplyr::filter(distance >= benchmark_value) |>
+  dplyr::filter((from == 2) & (to == 3)) |>
   dplyr::mutate(type = "long_edges")
@@
 
-UMAP_data_with_hb_id_selected <- UMAP_data_with_hb_id |> dplyr::filter(hb_id %in% c(2, 15)) |>
-  dplyr::mutate(select_point = if_else(hb_id == 2, "data1", "data2"))
-
-a1 <- ggplot(tr_df |> dplyr::filter((y == min(tr_df$y)) | (y == max(tr_df$y))), aes(x = x, y = y)) +
+UMAP_data_with_hb_id_selected <- UMAP_data_with_hb_id |> dplyr::filter(hb_id %in% c(hb_id_selected$from_hb_id, hb_id_selected$to_hb_id)) |>
+  dplyr::mutate(select_point = if_else(hb_id == hb_id_selected$from_hb_id, "data1", "data2"))
+
+UMAP_data_with_hb_id_selected_no <- UMAP_data_with_hb_id |> dplyr::filter(!(hb_id %in% c(hb_id_selected$from_hb_id, hb_id_selected$to_hb_id)))
+
+# bin_coord_1 <- tr_df |>
+#   dplyr::filter((x == hb_id_selected$x_from)) |>
+#   dplyr::filter((y == hb_id_selected$y_from))
+
+bin_coord_1 <- tr_df |>
+  head(1)
+
+# bin_coord_2 <- tr_df |>
+#   dplyr::filter((x == hb_id_selected$x_to)) |>
+#   dplyr::filter((y == hb_id_selected$y_to))
+
+bin_coord_2 <- tr_df |>
+  dplyr::filter(row_number() == 2)
+
+bin_coord <- dplyr::bind_rows(bin_coord_1, bin_coord_2)
+
+a1 <- ggplot(bin_coord, aes(x = x, y = y)) +
   geom_segment(
-    aes(x = x_from, y = y_from, xend = x_to, yend = y_to, color = type),
-    data = tr_from_to_df_coord_with_group
+    aes(x = x_from, y = y_from, xend = x_to, yend = y_to),
+    data = tr_from_to_df_coord_with_group, colour = "#f6e8c3"
   ) +
@@
   coord_equal() +
-  scale_colour_manual(values = c("#000000", "#f6e8c3"))  + 
+  geom_point(data = UMAP_data_with_hb_id_selected, aes(x = UMAP1, y = UMAP2, colour = select_point), alpha = 0.5, size = 1) +
+  scale_color_manual(values=c("#e6550d", "#000000")) +
   geom_point(size = 3, colour = "#33a02c") + 
@@
 a2 <- ggplot(data = hex_full_count_df, aes(x = x, y = y)) + 
-  geom_polygon(color = "#e7e1ef", aes(group = polygon_id), fill = "#ffffff") +
-  geom_point(data = UMAP_data_with_hb_id_selected, aes(x = UMAP1, y = UMAP2, colour = select_point), alpha = 0.5) +
+  geom_polygon(color = "#deebf7", aes(group = polygon_id), fill = "#ffffff") +
+  geom_segment(
+    aes(x = x_from, y = y_from, xend = x_to, yend = y_to), colour = "#f6e8c3",
+    data = tr_from_to_df_coord_with_group
+  ) +
+  geom_point(data = UMAP_data_with_hb_id_selected, aes(x = UMAP1, y = UMAP2, colour = select_point), size = 1, alpha = 0.5) +
   scale_color_manual(values=c("#e6550d", "#000000")) +
-  geom_point(data = df_bin_centroids_all |> dplyr::filter(hexID %in% c(2, 15)), aes(x = x, y = y), colour = "#33a02c", size = 3) +
+  geom_point(data = df_bin_centroids_all |> dplyr::filter(hexID %in% c(hb_id_selected$from_hb_id, hb_id_selected$to_hb_id)), aes(x = x, y = y), colour = "#33a02c", size = 3) +
+    geom_point(data = UMAP_data_with_hb_id_selected_no, aes(x = UMAP1, y = UMAP2), colour = "#6a3d9a", size = 1, alpha = 0.5) +
   scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  coord_cartesian(xlim =c(-5, 8), ylim = c(-10, 10)) +
   theme_light() +
@@
 df1 <- df_all |> 
-  dplyr::filter(hb_id == "2") |>
+  dplyr::filter(hb_id == as.character(hb_id_selected$from_hb_id)) |>
   dplyr::select(tidyselect::starts_with("x")) |>
@@
 df2 <- df_all |> 
-  dplyr::filter(hb_id == "15") |>
+  dplyr::filter(hb_id == as.character(hb_id_selected$to_hb_id)) |>
   dplyr::select(tidyselect::starts_with("x")) |>
@@
 df3 <- df_all |> 
-  dplyr::filter(!(hb_id %in% c("2", "15"))) |>
+  dplyr::filter(!(hb_id %in% c(as.character(hb_id_selected$from_hb_id), as.character(hb_id_selected$to_hb_id)))) |>
   dplyr::select(tidyselect::starts_with("x")) |>
@@
   dplyr::filter(hb_id %in% df_bin_centroids$hexID) |>
-  dplyr::filter(hb_id %in% c("2", "15")) |>
+  dplyr::filter(hb_id %in% c(as.character(hb_id_selected$from_hb_id), as.character(hb_id_selected$to_hb_id))) |>
   dplyr::select(-hb_id) |>
@@
 (a) visualize the points and the hexagonal bin centroids related 2nd and 15th hexagons, (b) visualization of the edge connected the 2nd and 15th hexagons (colored in red) in the triangular mesh.
-A video of tour animation is available at <https://www.youtube.com/watch?v=vBvM30Plt24>.
+A video of tour animation is available at <https://youtu.be/yN-RNPV_cHQ>.
 
@@
 
-#### Benchmark value to remove low-density hexagons
+#### Remove low-density hexagons
 
@@
 
-#### Benchmark value to remove long edges
+#### Remove long edges
 

### 2024-01-19 paper.tex
Regenerated Figure 4 by comments

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, frame hidden, interior hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, sharp corners, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden, boxrule=0pt, enhanced, sharp corners, breakable]}{\end{tcolorbox}}\fi
 
@@
 hexagons (colored in red) in the triangular mesh. A video of tour
-animation is available at
-\url{https://www.youtube.com/watch?v=vBvM30Plt24}.}
+animation is available at \url{https://youtu.be/yN-RNPV_cHQ}.}
 
@@
 
-\hypertarget{benchmark-value-to-remove-low-density-hexagons}{%
-\subsubsection{Benchmark value to remove low-density
-hexagons}\label{benchmark-value-to-remove-low-density-hexagons}}
+\hypertarget{remove-low-density-hexagons}{%
+\subsubsection{Remove low-density
+hexagons}\label{remove-low-density-hexagons}}
 
@@
 
-\hypertarget{benchmark-value-to-remove-long-edges}{%
-\subsubsection{Benchmark value to remove long
-edges}\label{benchmark-value-to-remove-long-edges}}
+\hypertarget{remove-long-edges}{%
+\subsubsection{Remove long edges}\label{remove-long-edges}}
 

### 2024-01-19 paper.qmd
Add remove low density hex, long edges MSE, AIC plots with different evaluations

@@
 
-df_bin_n <- df_bin |>
-  mutate(ID = 1:NROW(df_bin))
-
-tr_from_to_df_n <- left_join(tr_from_to_df, df_bin_n, by = c("from" = "ID")) |>
-  rename("from_hb_id" = "hb_id")
-
-tr_from_to_df_n <- left_join(tr_from_to_df_n, df_bin_n, by = c("to" = "ID")) |>
-  rename("to_hb_id" = "hb_id")
+df_bin_centroids_n <- df_bin_centroids |>
+  select(hexID) |>
+  mutate(ID = 1:NROW(df_bin_centroids))
+
+tr_from_to_df_n <- left_join(tr_from_to_df, df_bin_centroids_n, by = c("from" = "ID")) |>
+  rename("from_hb_id" = "hexID")
+
+tr_from_to_df_n <- left_join(tr_from_to_df_n, df_bin_centroids_n, by = c("to" = "ID")) |>
+  rename("to_hb_id" = "hexID")
 
 hb_id_selected <- tr_from_to_df_n |>
-  dplyr::filter((from == 2) & (to == 3)) |>
-  head(1)
+  dplyr::filter((from == 2) & (to == 3))
 
@@
 
-tr_from_to_df_coord_with_group <- merge(tr_from_to_df, distance_edges, by = c("from", "to"))
+tr_from_to_df_coord_with_group <- inner_join(tr_from_to_df, distance_edges, by = c("from" = "from", "to" = "to"))
 
@@
 
+bin_coord_1 <- tr_df |>
+  dplyr::filter((x == hb_id_selected$x_from)) |>
+  dplyr::filter((y == hb_id_selected$y_from))
+
 # bin_coord_1 <- tr_df |>
-#   dplyr::filter((x == hb_id_selected$x_from)) |>
-#   dplyr::filter((y == hb_id_selected$y_from))
-
-bin_coord_1 <- tr_df |>
-  head(1)
+#   head(1)
+
+bin_coord_2 <- tr_df |>
+  dplyr::filter((x == hb_id_selected$x_to)) |>
+  dplyr::filter((y == hb_id_selected$y_to))
 
 # bin_coord_2 <- tr_df |>
-#   dplyr::filter((x == hb_id_selected$x_to)) |>
-#   dplyr::filter((y == hb_id_selected$y_to))
-
-bin_coord_2 <- tr_df |>
-  dplyr::filter(row_number() == 2)
+#   dplyr::filter(row_number() == 2)
 
@@
   geom_segment(
-    aes(x = x_from, y = y_from, xend = x_to, yend = y_to),
-    data = tr_from_to_df_coord_with_group, colour = "#f6e8c3"
-  ) +
+    aes(x = x_from, y = y_from, xend = x_to, yend = y_to, colour = type),
+    data = tr_from_to_df_coord_with_group) +
   geom_point() +
@@
   geom_point(data = UMAP_data_with_hb_id_selected, aes(x = UMAP1, y = UMAP2, colour = select_point), alpha = 0.5, size = 1) +
-  scale_color_manual(values=c("#e6550d", "#000000")) +
+  scale_color_manual(values=c("#e6550d", "#000000", "#e31a1c", "#f6e8c3")) +
   geom_point(size = 3, colour = "#33a02c") + 
@@
 (a) visualize the points and the hexagonal bin centroids related 2nd and 15th hexagons, (b) visualization of the edge connected the 2nd and 15th hexagons (colored in red) in the triangular mesh.
-A video of tour animation is available at <https://youtu.be/yN-RNPV_cHQ>.
+A video of tour animation is available at <https://youtu.be/ffWFVK1UTIc>.
 
@@
 Adjusting the parameter $b_1$ provides control over the total number of bins $b$, allowing us to fine-tune the hexagonal grid.
-
-#### Remove low-density hexagons
-
-Addressing low-density hexagons is a systematic process to handle sparsely represented data in certain regions.
-For each hex bin, we identify the six nearest hex bins using an equal 2D distance metric.
-Then, we calculate the mean density, as outlined in the equations:
-
-$$
-\text{standard count} = \frac{\text{count}}{\text{max count}} 
-$${#eq-equationp2}
-
-$$
-\text{mean density} = \frac{\text{standard count}}{6} 
-$${#eq-equationp3}
-
-The standard count is derived from the number of observations in the hex bins.
-By examining the distribution of mean densities and designating the first quartile as the benchmark value, hex bins with mean densities below this benchmark are removed.
-This process ensures the elimination of regions with insufficient data density, focusing on areas with more significant data representation and preserving the overall structure in the low-dimensional space.
-
-#### Remove long edges
-
-The removal of long edges is a critical step to create a smoother representation  by iteratively eliminating hexagons with excessive distances between centroids.
-This process eliminates outliers and noise while preserving essential local relationships within the data.
-To achieve this, distances between vertices are sorted, and unique distance values are extracted.
-By setting a threshold based on the largest difference between consecutive distance values, long edges are identified and removed.
-This refinement step contributes to enhancing the quality of the triangular mesh, ensuring a more accurate representation of the data structure.
-
-- Compute high-D distances (using averaged high-D data)
-- Fit a linear model using response as distance_high_D and predictor as distance_2D
-- Filter the model residuals only for short edges
-- compute error for each edge connection as (residual for that point)/(total residual for all edges * number of short edges)
-
-```{r}
-#| echo: false
-#| warning: false
-
-error_df_lg_1 <- read_rds("data/s_curve/s_curve_summary_lg_tsne.rds")
-error_df_lg_2 <- read_rds("data/s_curve/s_curve_summary_lg_umap.rds")
-error_df_lg_3 <- read_rds("data/s_curve/s_curve_summary_lg_phate.rds")
-error_df_lg_4 <- read_rds("data/s_curve/s_curve_summary_lg_trimap.rds")
-error_df_lg_5 <- read_rds("data/s_curve/s_curve_summary_lg_pacmap.rds")
-
-error_df_lg <- dplyr::bind_rows(error_df_lg_1, error_df_lg_2, error_df_lg_3, error_df_lg_4, error_df_lg_5)
-
-error_df_lg$method <- factor(error_df_lg$method, levels = c("tSNE", "UMAP", "PHATE", "TriMAP", "PaCMAP"))
-
-error_plot_lg_1 <- ggplot(error_df_lg_1, aes(x = benchmark_rm_lg, y = total_error, color = method)) +
-  geom_point() +
-  geom_line() +
-  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  theme_light() +
-  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  scale_colour_manual(values = c("#e41a1c")) +
-  ylab("error") +
-  xlab("Benchmark value")
-
-error_plot_lg_2 <- ggplot(error_df_lg_2, aes(x = benchmark_rm_lg, y = total_error, color = method)) +
-  geom_point() +
-  geom_line() +
-  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  theme_light() +
-  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  scale_colour_manual(values = c("#377eb8")) +
-  ylab("error") +
-  xlab("Benchmark value")
-
-error_plot_lg_3 <- ggplot(error_df_lg_3, aes(x = benchmark_rm_lg, y = total_error, color = method)) +
-  geom_point() +
-  geom_line() +
-  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  theme_light() +
-  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  scale_colour_manual(values = c("#4daf4a")) +
-  ylab("error") +
-  xlab("Benchmark value")
-
-error_plot_lg_4 <- ggplot(error_df_lg_4, aes(x = benchmark_rm_lg, y = total_error, color = method)) +
-  geom_point() +
-  geom_line() +
-  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  theme_light() +
-  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  scale_colour_manual(values = c("#984ea3")) +
-  ylab("error") +
-  xlab("Benchmark value")
-
-error_plot_lg_5 <- ggplot(error_df_lg_5, aes(x = benchmark_rm_lg, y = total_error, color = method)) +
-  geom_point() +
-  geom_line() +
-  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  theme_light() +
-  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  scale_colour_manual(values = c("#ff7f00")) +
-  ylab("error") +
-  xlab("Benchmark value")
-```
-
-```{r}
-#| echo: false
-#| warning: false
-#| fig-cap: Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset with different benchmark values to remove the long edges.
-What is the effective benchmark value to remove the long edges?
-#| label: fig-diagnosticpltScurvelgrm
-#| out-width: 100%
-
-error_plot_lg_1 + error_plot_lg_2 + error_plot_lg_3 + error_plot_lg_4 + error_plot_lg_5 +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', nrow = 2) &
-  theme(legend.position='bottom', plot.tag = element_text(size = 8)) 
-```
-
-### Model summaries {#sec-summary}
-
-#### Predicted values and residuals
-
-The approach involves employing the K-nearest neighbors (KNN) algorithm to identify the nearest hexagonal bin centroid in the 2D space.
-Subsequently, the coordinates of this centroid are assigned as the low-dimensional predicted values for the test data in 2D space.
-It is noteworthy that traditional NLDR methods, such as tSNE, often lack a direct predict function, making our approach valuable for generating predicted values in the absence of such functionalities.
-
-The concept of "residuals" is pivotal in evaluating the accuracy of representing bin centroids in high dimensions.
-To quantify this accuracy, we introduce an error metric, which measures the sum of squared differences between the high-dimensional data ($x_{ij}$) and the predicted bin centroid data in high-dimensional space ($C_{x_ij}$) across all bins and dimensions.
-Mathematically, this error is expressed as:
-
-$$
-\text{Error} = \sum_{j = 1}^{n}\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2
-$$ {#eq-equation11}
-
-Here, $n$ represents the number of bins, $p$ represents the dimensions, $x_{ij}$ is the actual high-dimensional data, and $C_{x_ij}$ is the predicted bin centroid data in high dimensions.
-
-The error metric outlined above provides valuable insights into the overall accuracy of our predictive model.
-By quantifying the squared deviations between the actual and predicted values across all bins and dimensions, we gain a comprehensive understanding of how well our method captures and represents the underlying structure of the data in the reduced 2D space.
-This assessment is crucial for evaluating the efficacy of our NLDR technique in preserving the essential information present in the original high-dimensional data.
-
-#### Goodness of fit statistics
-
-Moving on to the assessment of prediction accuracy, we calculate the Mean Squared Error (MSE).
-The MSE helps measure the average squared differences between the actual high-dimensional data ($x_{ij}$) and the predicted bin centroid data in high-D ($C_{x_ij}$) values across all bins.
-Mathematically, this is expressed as:
-
-$$
-\text{MSE} = \sum_{j = 1}^{n} \frac{\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2}{b}
-$$ {#eq-equation9}
-
-Here, $b$ signifies the total number of bins, $p$ denotes the number of dimensions in the high-dimensional data, and $n$ represents the number of observations.
-
-Additionally, we gauge the model's performance using the Akaike Information Criterion (AIC), calculated by the formula:
-
-$$
-\text{AIC} = 2bp + np * log(\text{MSE})
-$$ {#eq-equation10}
-
-  These metrics, MSE and AIC, collectively offer valuable insights into the model's predictive performance, considering both accuracy and complexity in the predictions.
 
@@
 
+#### Remove low-density hexagons
+
+Addressing low-density hexagons is a systematic process to handle sparsely represented data in certain regions.
+For each hex bin, we identify the six nearest hex bins using an equal 2D distance metric.
+Then, we calculate the mean density, as outlined in the equations:
+
+$$
+\text{standard count} = \frac{\text{count}}{\text{max count}} 
+$${#eq-equationp2}
+
+$$
+\text{mean density} = \frac{\text{standard count}}{6} 
+$${#eq-equationp3}
+
+The standard count is derived from the number of observations in the hex bins.
+By examining the distribution of mean densities and designating the first quartile as the benchmark value, hex bins with mean densities below this benchmark are removed.
+This process ensures the elimination of regions with insufficient data density, focusing on areas with more significant data representation and preserving the overall structure in the low-dimensional space.
+
+<!--lwd smoothing criteria 1 with s-curve-->
+
+```{r}
+#| warning: false
+#| echo: false
+
+MSE_df_1_lwd <- read_rds("data/s_curve/s_curve_summary_lwd_tsne.rds")
+MSE_df_2_lwd <- read_rds("data/s_curve/s_curve_summary_lwd_umap.rds")
+MSE_df_3_lwd <- read_rds("data/s_curve/s_curve_summary_lwd_phate.rds")
+MSE_df_4_lwd <- read_rds("data/s_curve/s_curve_summary_lwd_trimap.rds")
+MSE_df_5_lwd <- read_rds("data/s_curve/s_curve_summary_lwd_pacmap.rds")
+
+MSE_df_lwd <- dplyr::bind_rows(MSE_df_1_lwd, MSE_df_2_lwd, MSE_df_3_lwd, MSE_df_4_lwd, MSE_df_5_lwd)
+
+MSE_df_lwd$method <- factor(MSE_df_lwd$method, levels = c("tSNE", "UMAP", "PHATE", "TriMAP", "PaCMAP"))
+
+## To draw with AIC
+aic_plot_lwd <- ggplot(MSE_df_lwd |> dplyr::filter(data_type == "training"), aes(x = benchmark_rm_hex,
+                                                                                 y = total_error,
+                                                                                 color = method
+)) +
+  geom_point() +
+  geom_line() +
+  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  theme_light() +
+  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  scale_colour_manual(values = c("#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("AIC") +
+  xlab("Benchmark value")
+## Effective number of bins along x-axis
+
+mse_plot_lwd <- ggplot(MSE_df_lwd, aes(x = benchmark_rm_hex,
+                                       y = total_mse,
+                                       color = method
+)) +
+  geom_point() +
+  geom_line() +
+  theme_light() +
+  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  scale_colour_manual(values = c("#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("MSE") +
+  xlab("Benchmark value")
+
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| fig-cap: Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset with different benchmark values to remove the low-density heaxgons.
+What is the effective benchmark value to remove the low-density heaxgons?
+#| label: fig-diagnosticpltScurvelwd
+#| out-width: 100%
+
+aic_plot_lwd + mse_plot_lwd +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='bottom', plot.tag = element_text(size = 8))
+```
+
+#### Remove long edges
+
+The removal of long edges is a critical step to create a smoother representation  by iteratively eliminating hexagons with excessive distances between centroids.
+This process eliminates outliers and noise while preserving essential local relationships within the data.
+To achieve this, distances between vertices are sorted, and unique distance values are extracted.
+By setting a threshold based on the largest difference between consecutive distance values, long edges are identified and removed.
+This refinement step contributes to enhancing the quality of the triangular mesh, ensuring a more accurate representation of the data structure.
+
+```{r}
+#| echo: false
+#| warning: false
+
+error_df_lg_1 <- read_rds("data/s_curve/s_curve_summary_lg_tsne.rds")
+error_df_lg_2 <- read_rds("data/s_curve/s_curve_summary_lg_umap.rds")
+error_df_lg_3 <- read_rds("data/s_curve/s_curve_summary_lg_phate.rds")
+error_df_lg_4 <- read_rds("data/s_curve/s_curve_summary_lg_trimap.rds")
+error_df_lg_5 <- read_rds("data/s_curve/s_curve_summary_lg_pacmap.rds")
+
+error_df_lg <- dplyr::bind_rows(error_df_lg_1, error_df_lg_2, error_df_lg_3, error_df_lg_4, error_df_lg_5)
+
+error_df_lg$method <- factor(error_df_lg$method, levels = c("tSNE", "UMAP", "PHATE", "TriMAP", "PaCMAP"))
+
+## To draw with AIC
+aic_plot_lg <- ggplot(error_df_lg, aes(x = benchmark_rm_lg,
+                                                                                 y = total_error,
+                                                                                 color = method
+)) +
+  geom_point() +
+  geom_line() +
+  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  theme_light() +
+  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  scale_colour_manual(values = c("#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("AIC") +
+  xlab("Benchmark value")
+## Effective number of bins along x-axis
+
+mse_plot_lg <- ggplot(error_df_lg, aes(x = benchmark_rm_lg,
+                                       y = total_mse,
+                                       color = method
+)) +
+  geom_point() +
+  geom_line() +
+  theme_light() +
+  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  scale_colour_manual(values = c("#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("MSE") +
+  xlab("Benchmark value")
+
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| fig-cap: Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset with different benchmark values to remove the long edges.
+What is the effective benchmark value to remove the long edges?
+#| label: fig-diagnosticpltScurvelgrm
+#| out-width: 100%
+
+aic_plot_lg + mse_plot_lg +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='bottom', plot.tag = element_text(size = 8)) 
+```
+
+### Model summaries {#sec-summary}
+
+#### Predicted values and residuals
+
+The approach involves employing the K-nearest neighbors (KNN) algorithm to identify the nearest hexagonal bin centroid in the 2D space.
+Subsequently, the coordinates of this centroid are assigned as the low-dimensional predicted values for the test data in 2D space.
+It is noteworthy that traditional NLDR methods, such as tSNE, often lack a direct predict function, making our approach valuable for generating predicted values in the absence of such functionalities.
+
+The concept of "residuals" is pivotal in evaluating the accuracy of representing bin centroids in high dimensions.
+To quantify this accuracy, we introduce an error metric, which measures the sum of squared differences between the high-dimensional data ($x_{ij}$) and the predicted bin centroid data in high-dimensional space ($C_{x_ij}$) across all bins and dimensions.
+Mathematically, this error is expressed as:
+
+$$
+\text{Error} = \sum_{j = 1}^{n}\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2
+$$ {#eq-equation11}
+
+Here, $n$ represents the number of bins, $p$ represents the dimensions, $x_{ij}$ is the actual high-dimensional data, and $C_{x_ij}$ is the predicted bin centroid data in high dimensions.
+
+The error metric outlined above provides valuable insights into the overall accuracy of our predictive model.
+By quantifying the squared deviations between the actual and predicted values across all bins and dimensions, we gain a comprehensive understanding of how well our method captures and represents the underlying structure of the data in the reduced 2D space.
+This assessment is crucial for evaluating the efficacy of our NLDR technique in preserving the essential information present in the original high-dimensional data.
+
+#### Goodness of fit statistics
+
+Moving on to the assessment of prediction accuracy, we calculate the Mean Squared Error (MSE).
+The MSE helps measure the average squared differences between the actual high-dimensional data ($x_{ij}$) and the predicted bin centroid data in high-D ($C_{x_ij}$) values across all bins.
+Mathematically, this is expressed as:
+
+$$
+\text{MSE} = \sum_{j = 1}^{n} \frac{\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2}{b}
+$$ {#eq-equation9}
+
+Here, $b$ signifies the total number of bins, $p$ denotes the number of dimensions in the high-dimensional data, and $n$ represents the number of observations.
+
+Additionally, we gauge the model's performance using the Akaike Information Criterion (AIC), calculated by the formula:
+
+$$
+\text{AIC} = 2bp + np * log(\text{MSE})
+$$ {#eq-equation10}
+
+  These metrics, MSE and AIC, collectively offer valuable insights into the model's predictive performance, considering both accuracy and complexity in the predictions.
+
 <!--UMAP with different param choices-->
@@
 
-<!--lwd smoothing criteria 1 with s-curve-->
-
-```{r}
-#| warning: false
-#| echo: false
-
-MSE_df_1_lwd <- read_rds("data/s_curve/s_curve_summary_lwd_tsne.rds")
-MSE_df_2_lwd <- read_rds("data/s_curve/s_curve_summary_lwd_umap.rds")
-MSE_df_3_lwd <- read_rds("data/s_curve/s_curve_summary_lwd_phate.rds")
-MSE_df_4_lwd <- read_rds("data/s_curve/s_curve_summary_lwd_trimap.rds")
-MSE_df_5_lwd <- read_rds("data/s_curve/s_curve_summary_lwd_pacmap.rds")
-
-MSE_df_lwd <- dplyr::bind_rows(MSE_df_1_lwd, MSE_df_2_lwd, MSE_df_3_lwd, MSE_df_4_lwd, MSE_df_5_lwd)
-
-MSE_df_lwd$method <- factor(MSE_df_lwd$method, levels = c("tSNE", "UMAP", "PHATE", "TriMAP", "PaCMAP"))
-
-## To draw with AIC
-aic_plot_lwd <- ggplot(MSE_df_lwd |> dplyr::filter(data_type == "training"), aes(x = benchmark_rm_hex,
-                                                                                 y = total_error,
-                                                                                 color = method
-)) +
-  geom_point() +
-  geom_line() +
-  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  theme_light() +
-  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  scale_colour_manual(values = c("#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("AIC") +
-  xlab("Benchmark value")
-## Effective number of bins along x-axis
-
-mse_plot_lwd <- ggplot(MSE_df_lwd, aes(x = benchmark_rm_hex,
-                                       y = total_mse,
-                                       color = method
-)) +
-  geom_point() +
-  geom_line() +
-  theme_light() +
-  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  scale_colour_manual(values = c("#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("MSE") +
-  xlab("Benchmark value")
-
-```
-
-```{r}
-#| echo: false
-#| warning: false
-#| fig-cap: Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset with different benchmark values to remove the low-density heaxgons.
-What is the effective benchmark value to remove the low-density heaxgons?
-#| label: fig-diagnosticpltScurvelwd
-#| out-width: 100%
-
-aic_plot_lwd + mse_plot_lwd +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', ncol = 2) &
-  theme(legend.position='bottom', plot.tag = element_text(size = 8))
-```
-
 ### Prediction {#sec-prediction}

### 2024-01-19 paper.tex
Add remove low density hex, long edges MSE, AIC plots with different evaluations

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden, boxrule=0pt, enhanced, sharp corners, breakable]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, breakable, enhanced, sharp corners, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 hexagons (colored in red) in the triangular mesh. A video of tour
-animation is available at \url{https://youtu.be/yN-RNPV_cHQ}.}
+animation is available at \url{https://youtu.be/ffWFVK1UTIc}.}
 
@@
 of bins \(b\), allowing us to fine-tune the hexagonal grid.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticpltScurve-1.pdf}
+
+}
+
+\caption{\label{fig-diagnosticpltScurve}Goodness of fit statistics from
+different NLDR techniques applied to training S-curve dataset. What is
+the best NLDR technique to represent the original data in 2D?}
+
+\end{figure}
 
@@
 the low-dimensional space.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticpltScurvelwd-1.pdf}
+
+}
+
+\caption{\label{fig-diagnosticpltScurvelwd}Goodness of fit statistics
+from different NLDR techniques applied to training S-curve dataset with
+different benchmark values to remove the low-density heaxgons. What is
+the effective benchmark value to remove the low-density heaxgons?}
+
+\end{figure}
 
@@
 
-\begin{itemize}
-\tightlist
-\item
-  Compute high-D distances (using averaged high-D data)
-\item
-  Fit a linear model using response as distance\_high\_D and predictor
-  as distance\_2D
-\item
-  Filter the model residuals only for short edges
-\item
-  compute error for each edge connection as (residual for that
-  point)/(total residual for all edges * number of short edges)
-\end{itemize}
-
 \begin{figure}
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticpltScurve-1.pdf}
-
-}
-
-\caption{\label{fig-diagnosticpltScurve}Goodness of fit statistics from
-different NLDR techniques applied to training S-curve dataset. What is
-the best NLDR technique to represent the original data in 2D?}
-
-\end{figure}
-
-\begin{figure}
-
 {\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervisUMAP-1.pdf}
@@
 the original data in 2D?}
-
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticpltScurvelwd-1.pdf}
-
-}
-
-\caption{\label{fig-diagnosticpltScurvelwd}Goodness of fit statistics
-from different NLDR techniques applied to training S-curve dataset with
-different benchmark values to remove the low-density heaxgons. What is
-the effective benchmark value to remove the low-density heaxgons?}
 

### 2024-01-19 quollr_code.R

@@ -159,6 +159,14 @@ generate_edge_info <- function(triangular_object) {
   tr_arcs_df2 <- tibble::tibble(from = trang$node1, to = trang$node3)
   tr_arcs_df3 <- tibble::tibble(from = trang$node2, to = trang$node3)
   tr_arcs_df <- dplyr::bind_rows(tr_arcs_df1, tr_arcs_df2, tr_arcs_df3)  ## Create dataframe with from and to edges
+  tr_arcs_df <- tr_arcs_df |> ## To remove duplicates
+    distinct()
+
+  ## To extract unique combinations
+  tr_arcs_df <- tr_arcs_df |>
+    mutate(x = pmin(from, to), y = pmax(from, to)) |>
+    distinct(x, y) |>
+    rename(c("from" = "x", "to" = "y"))
 
   ## To obtain x and values of from to in a dataframe
   vec <- stats::setNames(rep("", 6), c("from", "to", "x_from", "y_from",

### 2024-01-19 scripts/gen_summary_trimap_lg_rm_n.R

@@ -1,0 +1,83 @@
+library(readr)
+library(dplyr)
+
+set.seed(20240110)
+
+source("quollr_code.R", local = TRUE)
+source("nldr_code.R", local = TRUE)
+
+## Import data
+df_s_curve <- read_rds("data/s_curve/s_curve.rds")
+training_data_s_curve <- read_rds("data/s_curve/s_curve_training.rds")
+
+TriMAP_s_curve <- read_rds("data/s_curve/s_curve_trimap.rds")
+
+num_bins_s_curve <- 6
+
+shape_value_s_curve <- calculate_effective_shape_value(.data = TriMAP_s_curve,
+                                                       x = TriMAP1, y = TriMAP2) ## 1.259938
+
+## To extract bin centroids
+hexbin_data_object_s_curve <- extract_hexbin_centroids(nldr_df = TriMAP_s_curve,
+                                                       num_bins = num_bins_s_curve,
+                                                       shape_val = shape_value_s_curve, x = TriMAP1, y = TriMAP2)
+
+df_bin_centroids_s_curve <- hexbin_data_object_s_curve$hexdf_data
+
+TriMAP_s_curve_with_hb_id <- TriMAP_s_curve |>
+  dplyr::mutate(hb_id = hexbin_data_object_s_curve$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_s_curve <- dplyr::bind_cols(training_data_s_curve |> dplyr::select(-ID), TriMAP_s_curve_with_hb_id)
+
+## Averaged on high-D
+df_bin_s_curve <- avg_highD_data(.data = df_all_s_curve, column_start_text = "x")
+
+## Triangulate bin centroids
+tr1_object_s_curve <- triangulate_bin_centroids(df_bin_centroids_s_curve, x, y)
+tr_from_to_df_s_curve <- generate_edge_info(triangular_object = tr1_object_s_curve)
+
+## Compute 2D distances
+distance_s_curve <- cal_2D_dist(.data = tr_from_to_df_s_curve)
+
+## To find the benchmark value
+benchmark <- find_benchmark_value(.data = distance_s_curve, distance_col = distance)
+#benchmark <- 4.8
+
+benchmark_dist_vec <- distance_s_curve$distance |> round(3) |> unique() |> sort()
+benchmark_dist_vec <- benchmark_dist_vec[1:10]
[…]

### 2024-01-19 scripts/gen_summary_tsne_lg_rm_n.R

@@ -1,0 +1,85 @@
+library(readr)
+library(dplyr)
+
+set.seed(20240110)
+
+source("quollr_code.R", local = TRUE)
+source("nldr_code.R", local = TRUE)
+
+## Import data
+df_s_curve <- read_rds("data/s_curve/s_curve.rds")
+training_data_s_curve <- read_rds("data/s_curve/s_curve_training.rds")
+
+tSNE_s_curve <- read_rds("data/s_curve/s_curve_tsne_27.rds")
+
+num_bins_s_curve <- 8
+
+shape_value_s_curve <- calculate_effective_shape_value(.data = tSNE_s_curve,
+                                                       x = tSNE1, y = tSNE2) ## 1.259938
+
+## To extract bin centroids
+hexbin_data_object_s_curve <- extract_hexbin_centroids(nldr_df = tSNE_s_curve,
+                                                       num_bins = num_bins_s_curve,
+                                                       shape_val = shape_value_s_curve, x = tSNE1, y = tSNE2)
+
+df_bin_centroids_s_curve <- hexbin_data_object_s_curve$hexdf_data
+
+tSNE_s_curve_with_hb_id <- tSNE_s_curve |>
+  dplyr::mutate(hb_id = hexbin_data_object_s_curve$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_s_curve <- dplyr::bind_cols(training_data_s_curve |> dplyr::select(-ID), tSNE_s_curve_with_hb_id)
+
+## Averaged on high-D
+df_bin_s_curve <- avg_highD_data(.data = df_all_s_curve, column_start_text = "x")
+
+## Triangulate bin centroids
+tr1_object_s_curve <- triangulate_bin_centroids(df_bin_centroids_s_curve, x, y)
+tr_from_to_df_s_curve <- generate_edge_info(triangular_object = tr1_object_s_curve)
+
+## Compute 2D distances
+distance_s_curve <- cal_2D_dist(.data = tr_from_to_df_s_curve)
+
+## To find the benchmark value
+benchmark <- find_benchmark_value(.data = distance_s_curve, distance_col = distance)
+#benchmark <- 4.8
+
+
+distance <- distance_s_curve
+
+benchmark_dist_vec <- distance_s_curve$distance |> round(3) |> unique() |> sort()
[…]

### 2024-01-20 appendix.qmd
Add section for total number of bins tuning

@@
 
+Addressing low-density hexagons is a systematic process to handle sparsely represented data in certain regions.
+For each hex bin, we identify the six nearest hex bins using an equal 2D distance metric.
+Then, we calculate the mean density, as outlined in the equations:
+
+$$
+\text{standard count} = \frac{\text{count}}{\text{max count}} 
+$${#eq-equationp2}
+
+$$
+\text{mean density} = \frac{\text{standard count}}{6} 
+$${#eq-equationp3}
+
+The standard count is derived from the number of observations in the hex bins.
+By examining the distribution of mean densities and designating the first quartile as the benchmark value, hex bins with mean densities below this benchmark are removed.
+This process ensures the elimination of regions with insufficient data density, focusing on areas with more significant data representation and preserving the overall structure in the low-dimensional space.
+
+<!--
 In the step of addressing low-density hexagons, which can arise from sparsely represented data in certain regions, we employ a systematic strategy.
@@
 The result is the preservation of the overall structure of the data in the low-dimensional space, as illustrated in @fig-bintorm (c).
+-->
 
@@
 
+<!--The removal of long edges is a critical step to create a smoother representation  by iteratively eliminating hexagons with excessive distances between centroids.
+This process eliminates outliers and noise while preserving essential local relationships within the data.
+To achieve this, distances between vertices are sorted, and unique distance values are extracted.
+By setting a threshold based on the largest difference between consecutive distance values, long edges are identified and removed.
+This refinement step contributes to enhancing the quality of the triangular mesh, ensuring a more accurate representation of the data structure.
+-->
+
 In this step, we aim to create a smoother surface in the low-dimensional space by removing long edges from the triangular mesh.

### 2024-01-20 paper.qmd
Add section for total number of bins tuning

@@
 library(colorspace)
+library(kableExtra)
 
 library(grid)
+library(xtable)
+
+xtable2kable <- function(x) {
+  out <- capture.output(print(x, table.placement = NULL))[-(1:2)]
+  out <- paste(out, collapse = "\n")
+  structure(out, format = "latex", class = "knitr_kable")
+}
 
@@
 $${#eq-equation13}
-
-<!--first value-->
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-hex_full_count_df_loop1 <- read_rds("data/s_curve/s_curve_hex_3.rds")
-
-p1 <-  ggplot(data = hex_full_count_df_loop1, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
-  geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
-  theme_void() +
-  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=8), #change legend title font size
-        legend.text = element_text(size=6)) +
-  guides(fill = guide_colourbar(title = "Standardized count")) +
-  annotate(geom = 'text', label = "a", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
-
-```
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-hex_full_count_df_loop2 <- read_rds("data/s_curve/s_curve_hex_11.rds")
-
-p2 <-  ggplot(data = hex_full_count_df_loop2, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
-  geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
-  theme_void() +
-  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=8), #change legend title font size
-        legend.text = element_text(size=6)) +
-  guides(fill = guide_colourbar(title = "Standardized count")) +
-  annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
-
-```
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-hex_full_count_df_loop3 <- read_rds("data/s_curve/s_curve_hex_20.rds")
-
-p3 <-  ggplot(data = hex_full_count_df_loop3, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
-  geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
-  theme_void() +
-  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=8), #change legend title font size
-        legend.text = element_text(size=6)) +
-  guides(fill = guide_colourbar(title = "Standardized count")) +
-  annotate(geom = 'text', label = "c", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
-
-```
-
-```{r}
-#| echo: false
-#| label: fig-binsize
-#| fig-pos: H
-#| fig-cap: "Hexbin plots from different number of bins for the **UMAP** projections of **S-curve** data: (a) b = 32 (4, 8), s = 1.643542, (b) b = 264 (12, 22), s = 1.643542, and (c) b = 840 (21, 40), s = 1.643542.
-The hexbins are colored based on the density of points, with darker colors indicating higher point density and yellow color representing lower point density within each bin.
-Does a value of number of bins exist to effectively represent the low-dimensional data?"
-
-p1 + p2 + p3
-```
 
@@
 
-In our model tuning process, we strategically adjust three key parameters to optimize the performance and accuracy of our approach.
-They are (i) the total number of bins ($b$), (ii) a benchmark value to remove low-density hexagons, and (iii) a benchmark value to remove long edges.
+The performance and robustness of our model depend on three key parameters: (i) the total number of bins ($b$), (ii) a benchmark value used to remove low-density hexagons, and (iii) a benchmark value used to remove long edges.
+However, there is no analytical formula to calculate an appropriate value for these parameters (see Appendix for details on how default values are calculated).
+The selection of these parameter values depends on the model performance computed by Mean Squared Error (MSE) (see @sec-goodfit).
 
@@
 
-Adjusting the parameter $b_1$ provides control over the total number of bins $b$, allowing us to fine-tune the hexagonal grid.
+The number of hexagonal bins in the hexagonal grid has a considerable impact on the construction of the model.
+This is because it is the initial step in building the 2D model.
+The hexagonal grid with the chosen total number of bins must be able to capture the structure of the NLDR data.
+If the number of bins is too low, the model may not be able to capture the structure of the NLDR data effectively (see @fig-binsize (a)), while if there are too many bins, it may result in over-fitting the individual points of the NLDR data (see @fig-binsize (c)).
+Therefore, it is important to determine an appropriate number of bins to build an effective model.
+
+Typically, the MSE plot will have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
+Then, the slope will gradually decline or level off, indicating that a higher number of bins generates a smaller error.
+This MSE plot is useful in determining the effective number of bins required to construct the 2D model.
+@fig-diagnosticpltScurve (b) has created a graph that shows how the Mean Squared Error (MSE) changes depending on the number of bins assigned in the hexagonal grid for various NLDR techniques applied to the S-curve data set.
+As per the rule of thumb mentioned earlier, the effective number of bins for each NLDR method is shown in the @tbl-msebinsnldr.
+Furthermore, @fig-modelScurve shows the 2D models constructed for each NLDR technique with the selected total number of bins and how the models look in high-D with original data.
+
+```{r msebinsnldr}
+#| echo: false
+#| message: false
+#| warning: false
+
+num_bins_for_nldr_methods_df <- tibble::tibble(NLDR_technique = "tSNE", tot_bins = 8)
+
+num_bins_for_nldr_methods_df <- num_bins_for_nldr_methods_df |> 
+  tibble::add_row(NLDR_technique = "UMAP", tot_bins = 6)
+
+num_bins_for_nldr_methods_df <- num_bins_for_nldr_methods_df |> 
+  tibble::add_row(NLDR_technique = "PHATE", tot_bins = 8)
+
+num_bins_for_nldr_methods_df <- num_bins_for_nldr_methods_df |> 
+  tibble::add_row(NLDR_technique = "TriMAP", tot_bins = 6)
+
+num_bins_for_nldr_methods_df <- num_bins_for_nldr_methods_df |> 
+  tibble::add_row(NLDR_technique = "PaCMAP", tot_bins = 10)
+
+# knitr::kable(
+#   num_bins_for_nldr_methods_df,
+#   col.names = c('NLDR technique', 'total number of bins'), format = "latex", 
+#              caption = "Total number of bins suggested by rule of thumb for each NLDR technique applied to S-curve dataset."
+# ) |> 
+#   kable_styling(latex_options = c("striped", "HOLD_position"), position = "center", full_width = FALSE) |>
+#   row_spec(0, bold = TRUE)
+num_bins_for_nldr_methods_df$tot_bins <- as.character(num_bins_for_nldr_methods_df$tot_bins)
+colnames(num_bins_for_nldr_methods_df) <- c('NLDR technique', 'total number of bins')
+num_bins_for_nldr_methods_df <- tibble::remove_rownames(num_bins_for_nldr_methods_df)
+
+xtable(num_bins_for_nldr_methods_df, caption = "Total number of bins suggested by rule of thumb for each NLDR technique applied to S-curve dataset.", label = "tbl-msebinsnldr") |>
+  xtable2kable() |> 
+  kable_styling(latex_options = c("striped", "HOLD_position"), position = "center", full_width = FALSE) |>
+  row_spec(0, bold = TRUE)
+
+```
+
+<!--Adjusting the parameter $b_1$ provides control over the total number of bins $b$, allowing us to fine-tune the hexagonal grid.-->
+
+<!--first value-->
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+hex_full_count_df_loop1 <- read_rds("data/s_curve/s_curve_hex_3.rds")
+
+p1 <-  ggplot(data = hex_full_count_df_loop1, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
+  geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
+  theme_void() +
+  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6)) +
+  guides(fill = guide_colourbar(title = "Standardized count")) +
+  annotate(geom = 'text', label = "a", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
+
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+hex_full_count_df_loop2 <- read_rds("data/s_curve/s_curve_hex_11.rds")
+
+p2 <-  ggplot(data = hex_full_count_df_loop2, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
+  geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
+  theme_void() +
+  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6)) +
+  guides(fill = guide_colourbar(title = "Standardized count")) +
+  annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
+
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+hex_full_count_df_loop3 <- read_rds("data/s_curve/s_curve_hex_20.rds")
+
+p3 <-  ggplot(data = hex_full_count_df_loop3, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
+  geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
+  theme_void() +
+  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6)) +
+  guides(fill = guide_colourbar(title = "Standardized count")) +
+  annotate(geom = 'text', label = "c", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
+
+```
+
+```{r}
+#| echo: false
+#| label: fig-binsize
+#| fig-pos: H
+#| fig-cap: "Hexbin plots from different number of bins for the **UMAP** projections of **S-curve** data: (a) b = 32 (4, 8), s = 1.643542, (b) b = 264 (12, 22), s = 1.643542, and (c) b = 840 (21, 40), s = 1.643542.
+The hexbins are colored based on the density of points, with darker colors indicating higher point density and yellow color representing lower point density within each bin.
+What is the number of bins that would be effective in representing low-dimensional data?"
+
+p1 + p2 + p3
+```
 
@@
 
-num_bins_vec <- 1:20 ## Number of bins along the x-axis
+num_bins_vec <- 1:18 ## Number of bins along the x-axis
 
@@
 
-num_bins_vec <- 1:14 ## Number of bins along the x-axis
+num_bins_vec <- 1:13 ## Number of bins along the x-axis
 
@@
 #| echo: false
-#| fig-cap: Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset. What is the best NLDR technique to represent the original data in 2D?
+#| fig-cap: "Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset.
+What is the most effective NLDR technique to represent the original data in a 2D space?
+What is the effective number of bins in each NLDR technique to create a 2D model?"
 #| label: fig-diagnosticpltScurve
@@
 
-#### Remove low-density hexagons
-
-Addressing low-density hexagons is a systematic process to handle sparsely represented data in certain regions.
+#### Benchmark value to remove low-density hexagons
+
+<!--Addressing low-density hexagons is a systematic process to handle sparsely represented data in certain regions.
 For each hex bin, we identify the six nearest hex bins using an equal 2D distance metric.
@@
 By examining the distribution of mean densities and designating the first quartile as the benchmark value, hex bins with mean densities below this benchmark are removed.
-This process ensures the elimination of regions with insufficient data density, focusing on areas with more significant data representation and preserving the overall structure in the low-dimensional space.
+This process ensures the elimination of regions with insufficient data density, focusing on areas with more significant data representation and preserving the overall structure in the low-dimensional space.-->
 
@@
 
-#### Remove long edges
-
-The removal of long edges is a critical step to create a smoother representation  by iteratively eliminating hexagons with excessive distances between centroids.
+#### Benchmark value to remove long edges
+
+<!--The removal of long edges is a critical step to create a smoother representation  by iteratively eliminating hexagons with excessive distances between centroids.
 This process eliminates outliers and noise while preserving essential local relationships within the data.
@@
 By setting a threshold based on the largest difference between consecutive distance values, long edges are identified and removed.
-This refinement step contributes to enhancing the quality of the triangular mesh, ensuring a more accurate representation of the data structure.
+This refinement step contributes to enhancing the quality of the triangular mesh, ensuring a more accurate representation of the data structure.-->
 
@@
 
-#### Goodness of fit statistics
+#### Goodness of fit statistics {#sec-goodfit}
 
@@
 #| label: fig-modelScurve
-#| fig-cap: Is there a best model to represent the original data in 2D space or are they all providing equivalent information?, (a) Model in the 2D space with tSNE (<https://youtu.be/uy9SkAo6gAE>), (b) Model in the 2D space with UMAP (<https://youtu.be/0MJDhHrh_Ug>), (c) Model in the 2D space with PHATE (<https://youtu.be/HbVv0uy0QWk>), (d) Model in the 2D space with TriMAP (<https://youtu.be/2OGwipAjzc8>), and (e) Model in the 2D space with PaCMAP (<https://youtu.be/pkokI8d-cBk>).
+#| fig-cap: Is there a good model to represent the original data in 2D space or are they all providing equivalent information?, (a) Model in the 2D space with tSNE (<https://youtu.be/uy9SkAo6gAE>), (b) Model in the 2D space with UMAP (<https://youtu.be/0MJDhHrh_Ug>), (c) Model in the 2D space with PHATE (<https://youtu.be/HbVv0uy0QWk>), (d) Model in the 2D space with TriMAP (<https://youtu.be/2OGwipAjzc8>), and (e) Model in the 2D space with PaCMAP (<https://youtu.be/pkokI8d-cBk>).
 #| out-width: 100%
@@
 
+We introduce a new tool to help to determine which method, which parameter choice provide the most useful representation of high-D data.
+
 <!--Our research introduces a comprehensive framework that leverages tours for interactive exploration of high-dimensional data coupled with a low-dimensional manifold, facilitated by the `quollr` R package.

### 2024-01-20 paper.tex
Add section for total number of bins tuning

--- 2024-01-18 paper.tex
@@
 
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 \makeatletter
@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, frame hidden, interior hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, sharp corners, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, borderline west={3pt}{0pt}{shadecolor}, interior hidden, breakable, sharp corners, boxrule=0pt, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 
-\begin{figure}[H]
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-binsize-1.pdf}
-
-}
-
-\caption{\label{fig-binsize}Hexbin plots from different number of bins
-for the \textbf{UMAP} projections of \textbf{S-curve} data: (a) b = 32
-(4, 8), s = 1.643542, (b) b = 264 (12, 22), s = 1.643542, and (c) b =
-840 (21, 40), s = 1.643542. The hexbins are colored based on the density
-of points, with darker colors indicating higher point density and yellow
-color representing lower point density within each bin. Does a value of
-number of bins exist to effectively represent the low-dimensional data?}
-
-\end{figure}
-
 \textbf{Step 2: Obtain bin centroids}
@@
 hexagons (colored in red) in the triangular mesh. A video of tour
-animation is available at
-\url{https://www.youtube.com/watch?v=vBvM30Plt24}.}
+animation is available at \url{https://youtu.be/ffWFVK1UTIc}.}
 
@@
 
-In our model tuning process, we strategically adjust three key
-parameters to optimize the performance and accuracy of our approach.
-They are (i) the total number of bins (\(b\)), (ii) a benchmark value to
-remove low-density hexagons, and (iii) a benchmark value to remove long
-edges.
+The performance and robustness of our model depend on three key
+parameters: (i) the total number of bins (\(b\)), (ii) a benchmark value
+used to remove low-density hexagons, and (iii) a benchmark value used to
+remove long edges. However, there is no analytical formula to calculate
+an appropriate value for these parameters (see Appendix for details on
+how default values are calculated). The selection of these parameter
+values depends on the model performance computed by Mean Squared Error
+(MSE) (see Section~\ref{sec-goodfit}).
 
@@
 
-Adjusting the parameter \(b_1\) provides control over the total number
-of bins \(b\), allowing us to fine-tune the hexagonal grid.
+The number of hexagonal bins in the hexagonal grid has a considerable
+impact on the construction of the model. This is because it is the
+initial step in building the 2D model. The hexagonal grid with the
+chosen total number of bins must be able to capture the structure of the
+NLDR data. If the number of bins is too low, the model may not be able
+to capture the structure of the NLDR data effectively (see
+Figure~\ref{fig-binsize} (a)), while if there are too many bins, it may
+result in over-fitting the individual points of the NLDR data (see
+Figure~\ref{fig-binsize} (c)). Therefore, it is important to determine
+an appropriate number of bins to build an effective model.
+
+Typically, the MSE plot will have a steep slope at the beginning,
+indicating that a smaller number of bins causes a larger amount of
+error. Then, the slope will gradually decline or level off, indicating
+that a higher number of bins generates a smaller error. This MSE plot is
+useful in determining the effective number of bins required to construct
+the 2D model. Figure~\ref{fig-diagnosticpltScurve} (b) has created a
+graph that shows how the Mean Squared Error (MSE) changes depending on
+the number of bins assigned in the hexagonal grid for various NLDR
+techniques applied to the S-curve data set. As per the rule of thumb
+mentioned earlier, the effective number of bins for each NLDR method is
+shown in the \textbf{?@tbl-msebinsnldr}. Furthermore,
+Figure~\ref{fig-modelScurve} shows the 2D models constructed for each
+NLDR technique with the selected total number of bins and how the models
+look in high-D with original data.
+
+\begin{table}[H]
+\centering
+\begin{tabular}{rll}
+  \hline
+\textbf{} & \textbf{NLDR technique} & \textbf{total number of bins }\\ 
+  \hline
+\cellcolor{gray!6}{1} & \cellcolor{gray!6}{tSNE} & \cellcolor{gray!6}{8 }\\ 
+  2 & UMAP & 6 \\ 
+\cellcolor{gray!6}{  3} & \cellcolor{gray!6}{PHATE} & \cellcolor{gray!6}{8 }\\ 
+  4 & TriMAP & 6 \\ 
+\cellcolor{gray!6}{  5} & \cellcolor{gray!6}{PaCMAP} & \cellcolor{gray!6}{10 }\\ 
+   \hline
+\end{tabular}
+\caption{Total number of bins suggested by rule of thumb for each NLDR technique applied to S-curve dataset.} 
+\label{tbl-msebinsnldr}
+\end{table}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-binsize-1.pdf}
+
+}
+
+\caption{\label{fig-binsize}Hexbin plots from different number of bins
+for the \textbf{UMAP} projections of \textbf{S-curve} data: (a) b = 32
+(4, 8), s = 1.643542, (b) b = 264 (12, 22), s = 1.643542, and (c) b =
+840 (21, 40), s = 1.643542. The hexbins are colored based on the density
+of points, with darker colors indicating higher point density and yellow
+color representing lower point density within each bin. What is the
+number of bins that would be effective in representing low-dimensional
+data?}
+
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticpltScurve-1.pdf}
+
+}
+
+\caption{\label{fig-diagnosticpltScurve}Goodness of fit statistics from
+different NLDR techniques applied to training S-curve dataset. What is
+the most effective NLDR technique to represent the original data in a 2D
+space? What is the effective number of bins in each NLDR technique to
+create a 2D model?}
+
+\end{figure}
 
@@
 
-Addressing low-density hexagons is a systematic process to handle
-sparsely represented data in certain regions. For each hex bin, we
-identify the six nearest hex bins using an equal 2D distance metric.
-Then, we calculate the mean density, as outlined in the equations:
-
-\begin{equation}\protect\hypertarget{eq-equationp2}{}{
-\text{standard count} = \frac{\text{count}}{\text{max count}} 
-}\label{eq-equationp2}\end{equation}
-
-\begin{equation}\protect\hypertarget{eq-equationp3}{}{
-\text{mean density} = \frac{\text{standard count}}{6} 
-}\label{eq-equationp3}\end{equation}
-
-The standard count is derived from the number of observations in the hex
-bins. By examining the distribution of mean densities and designating
-the first quartile as the benchmark value, hex bins with mean densities
-below this benchmark are removed. This process ensures the elimination
-of regions with insufficient data density, focusing on areas with more
-significant data representation and preserving the overall structure in
-the low-dimensional space.
+\begin{figure}
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticpltScurvelwd-1.pdf}
+
+}
+
+\caption{\label{fig-diagnosticpltScurvelwd}Goodness of fit statistics
+from different NLDR techniques applied to training S-curve dataset with
+different benchmark values to remove the low-density heaxgons. What is
+the effective benchmark value to remove the low-density heaxgons?}
+
+\end{figure}
 
@@
 edges}\label{benchmark-value-to-remove-long-edges}}
-
-The removal of long edges is a critical step to create a smoother
-representation by iteratively eliminating hexagons with excessive
-distances between centroids. This process eliminates outliers and noise
-while preserving essential local relationships within the data. To
-achieve this, distances between vertices are sorted, and unique distance
-values are extracted. By setting a threshold based on the largest
-difference between consecutive distance values, long edges are
-identified and removed. This refinement step contributes to enhancing
-the quality of the triangular mesh, ensuring a more accurate
-representation of the data structure.
-
-\begin{itemize}
-\tightlist
-\item
-  Compute high-D distances (using averaged high-D data)
-\item
-  Fit a linear model using response as distance\_high\_D and predictor
-  as distance\_2D
-\item
-  Filter the model residuals only for short edges
-\item
-  compute error for each edge connection as (residual for that
-  point)/(total residual for all edges * number of short edges)
-\end{itemize}
 
@@
 
-\hypertarget{goodness-of-fit-statistics}{%
-\subsubsection{Goodness of fit
-statistics}\label{goodness-of-fit-statistics}}
+\hypertarget{sec-goodfit}{%
+\subsubsection{Goodness of fit statistics}\label{sec-goodfit}}
 
@@
 
-\caption{\label{fig-modelScurve}Is there a best model to represent the
+\caption{\label{fig-modelScurve}Is there a good model to represent the
 original data in 2D space or are they all providing equivalent
@@
 
+We introduce a new tool to help to determine which method, which
+parameter choice provide the most useful representation of high-D data.
+
 \hypertarget{references}{%

### 2024-01-20 paper.qmd
Add remove low-density hexagon tuning section

@@
 
-The number of hexagonal bins in the hexagonal grid has a considerable impact on the construction of the model.
+The number of hexagonal bins in the hexagonal grid has a considerable impact on the construction of the 2D model.
 This is because it is the initial step in building the 2D model.
@@
 Therefore, it is important to determine an appropriate number of bins to build an effective model.
+
+Furthermore, the total number of bins is determined by the number of bins along the x-axis and y-axis (see @eq-equation13).
+To calculate the effective total number of bins, the minimum and approximate maximum number of bins along the x-axis are used as candidate values.
+The minimum number of bins along the x-axis is $1$.
+The maximum number of bins along the x-axis can be estimated by taking the square root of the number of NLDR data points.
+To investigate the impact of the number of bins on MSE, the analysis is focused on the MSE for different total numbers of bins.
+This range encompasses the total number of bins calculated for the minimum and maximum values of the number of bins along the x-axis.
 
@@
 
+After setting up the hexagonal grid with an appropriate number of bins, it is possible that some hexagonal bins may have very few or no data points within them.
+To ensure comprehensive coverage of the NLDR data, it is necessary to select hexagonal bins that have a considerable number of data points within their hexagons.
+To achieve this, the standard number of points within each hexagon is first calculated.
+This standard count is obtained by dividing the number of points within each hexagon by the maximum number of points in the grid (see @eq-equationp2).
+Next, the bins that have a standard number of points less than a certain benchmark value are removed.
+After removing the hexagons that have insufficient data density, the hexagons with more substantial data representation are used to construct the 2D model.
+
+$$
+\text{standard count} = \frac{\text{count}}{\text{max count}} 
+$${#eq-equationp2}
+
+Furthermore, it is crucial to choose the benchmark value to remove low-density hexagons carefully.
+If we remove unnecessary bins, then it may result in long edges and an uneven 2D model.
+Therefore, instead of removing hexagons only identified by the benchmark value, we examine the standard number of points in the neighboring hexagons of the identified low-density hexagons.
+If the neighboring bins also have low counts, then only those bins will be removed.
+The other identified bins will be kept and used for constructing the 2D model along with the high-density bins (see Appendix for more details).
+
+The benchmark value for removing low-density hexagons has a range of $0$ and $1$.
+When analyzing how these benchmark values affect model performance, it's important to observe the change in Mean Squared Error (MSE) as the benchmark value increases.
+The MSE exhibits a gradual increase as the benchmark value progresses from $0$ to $1$.
+Assessing this rate of increase is crucial.
+If the increment is not substantial, the decision might lean towards retaining low-density hexagons.
+According to @fig-diagnosticpltScurvelwd, the change in MSE for NLDR techniques is relatively small across different benchmark values, except for the PHATE method.
+
+Consequently, it is not necessary to remove low-density hexagons when constructing the 2D model for the S-curve dataset in NLDR techniques, except for PHATE.
+While there may be some fluctuations in MSE between benchmark values of 0 to 1, the absence of significant changes suggests that maintaining low-density hexagons is preferable.
+However, to choose the best selection, it's recommended to examine the benchmark values around the first local minimum in the MSE curve.
+However, the MSE for PHATE fluctuates when benchmark values range between 0 and 1.
+Therefore, it is necessary to explore the benchmark values surrounding the first local minimum to determine an effective benchmark value.
+There is no universal rule for selecting a process.
+As shown in @fig-diagnosticpltScurvelwd, the MSE values for PHATE fluctuate, and the first local minimum is at $0$.
+Therefore, nearby benchmark values such as $0.01$, $0.04$, and $0.06$ were explored to identify the best value for preserving the data structure.
+In this case, a benchmark value of $0.06$ was chosen.
+The resulting 2D model is shown in @fig-modelScurve (c), while the video link provides a visual representation of how the model appears in high-D.
+
 <!--Addressing low-density hexagons is a systematic process to handle sparsely represented data in certain regions.
@@
 
+### Model selection {#sec-selection}
+
+- By comaring AIC what is the best NLDR technique to reprrsnt NLDR data
+
 ### Prediction {#sec-prediction}

### 2024-01-20 paper.tex
Add remove low-density hexagon tuning section

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, borderline west={3pt}{0pt}{shadecolor}, interior hidden, breakable, sharp corners, boxrule=0pt, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, enhanced, interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, frame hidden, breakable]}{\end{tcolorbox}}\fi
 
@@
 The number of hexagonal bins in the hexagonal grid has a considerable
-impact on the construction of the model. This is because it is the
+impact on the construction of the 2D model. This is because it is the
 initial step in building the 2D model. The hexagonal grid with the
@@
 an appropriate number of bins to build an effective model.
+
+Furthermore, the total number of bins is determined by the number of
+bins along the x-axis and y-axis (see Equation~\ref{eq-equation13}). To
+calculate the effective total number of bins, the minimum and
+approximate maximum number of bins along the x-axis are used as
+candidate values. The minimum number of bins along the x-axis is \(1\).
+The maximum number of bins along the x-axis can be estimated by taking
+the square root of the number of NLDR data points. To investigate the
+impact of the number of bins on MSE, the analysis is focused on the MSE
+for different total numbers of bins. This range encompasses the total
+number of bins calculated for the minimum and maximum values of the
+number of bins along the x-axis.
 
@@
 
+After setting up the hexagonal grid with an appropriate number of bins,
+it is possible that some hexagonal bins may have very few or no data
+points within them. To ensure comprehensive coverage of the NLDR data,
+it is necessary to select hexagonal bins that have a considerable number
+of data points within their hexagons. To achieve this, the standard
+number of points within each hexagon is first calculated. This standard
+count is obtained by dividing the number of points within each hexagon
+by the maximum number of points in the grid (see
+Equation~\ref{eq-equationp2}). Next, the bins that have a standard
+number of points less than a certain benchmark value are removed. After
+removing the hexagons that have insufficient data density, the hexagons
+with more substantial data representation are used to construct the 2D
+model.
+
+\begin{equation}\protect\hypertarget{eq-equationp2}{}{
+\text{standard count} = \frac{\text{count}}{\text{max count}} 
+}\label{eq-equationp2}\end{equation}
+
+Furthermore, it is crucial to choose the benchmark value to remove
+low-density hexagons carefully. If we remove unnecessary bins, then it
+may result in long edges and an uneven 2D model. Therefore, instead of
+removing hexagons only identified by the benchmark value, we examine the
+standard number of points in the neighboring hexagons of the identified
+low-density hexagons. If the neighboring bins also have low counts, then
+only those bins will be removed. The other identified bins will be kept
+and used for constructing the 2D model along with the high-density bins
+(see Appendix for more details).
+
+The benchmark value for removing low-density hexagons has a range of
+\(0\) and \(1\). When analyzing how these benchmark values affect model
+performance, it's important to observe the change in Mean Squared Error
+(MSE) as the benchmark value increases. The MSE exhibits a gradual
+increase as the benchmark value progresses from \(0\) to \(1\).
+Assessing this rate of increase is crucial. If the increment is not
+substantial, the decision might lean towards retaining low-density
+hexagons. According to Figure~\ref{fig-diagnosticpltScurvelwd}, the
+change in MSE for NLDR techniques is relatively small across different
+benchmark values, except for the PHATE method.
+
+Consequently, it is not necessary to remove low-density hexagons when
+constructing the 2D model for the S-curve dataset in NLDR techniques,
+except for PHATE. While there may be some fluctuations in MSE between
+benchmark values of 0 to 1, the absence of significant changes suggests
+that maintaining low-density hexagons is preferable. However, to choose
+the best selection, it's recommended to examine the benchmark values
+around the first local minimum in the MSE curve. However, the MSE for
+PHATE fluctuates when benchmark values range between 0 and 1. Therefore,
+it is necessary to explore the benchmark values surrounding the first
+local minimum to determine an effective benchmark value. There is no
+universal rule for selecting a process. As shown in
+Figure~\ref{fig-diagnosticpltScurvelwd}, the MSE values for PHATE
+fluctuate, and the first local minimum is at \(0\). Therefore, nearby
+benchmark values such as \(0.01\), \(0.04\), and \(0.06\) were explored
+to identify the best value for preserving the data structure. In this
+case, a benchmark value of \(0.06\) was chosen. The resulting 2D model
+is shown in Figure~\ref{fig-modelScurve} (c), while the video link
+provides a visual representation of how the model appears in high-D.
+
 \begin{figure}
@@
 \end{figure}
+
+\hypertarget{sec-selection}{%
+\subsection{Model selection}\label{sec-selection}}
+
+\begin{itemize}
+\tightlist
+\item
+  By comaring AIC what is the best NLDR technique to reprrsnt NLDR data
+\end{itemize}
 

### 2024-01-20 paper.qmd
Add benchmark value to remove long edges tunning section

@@
 #| label: fig-diagnosticpltScurvelwd
-#| out-width: 100%
-
-aic_plot_lwd + mse_plot_lwd +
-  plot_annotation(tag_levels = 'a') +
+##| out-width: 100%
+
+#aic_plot_lwd + 
+  mse_plot_lwd +
+  #plot_annotation(tag_levels = 'a') +
   plot_layout(guides='collect', ncol = 2) &
@@
 #### Benchmark value to remove long edges
+
+Achieving a smooth representation in 2D space is crucial, and one factor influencing this smoothness is the presence of long edges.
+These long edges occur when a line connects points that are distant in the triangular mesh, impacting the overall smoothness of the mesh.
+
+To investigate the effect of removing long edges on the MSE, we analyzed the MSE for various benchmark values.
+Surprisingly, the results indicated that the MSE remained consistent across different benchmark values, as shown in @fig-diagnosticpltScurvelgrm.
+It's important to note that edges are defined only in the 2D model and do not extend to higher dimensions.
+Consequently, removing edges does not affect the model in higher dimensions.
+
+There is no definitive rule for determining the benchmark value to remove long edges.
+However, we have a method to find a default value (see Appendix).
+Adjusting values around this default can help to find benchmark value to remove long edges, contributing to the construction of a smoother 2D representation.
 
@@
 #| label: fig-diagnosticpltScurvelgrm
-#| out-width: 100%
-
-aic_plot_lg + mse_plot_lg +
-  plot_annotation(tag_levels = 'a') +
+##| out-width: 100%
+
+#aic_plot_lg + 
+  mse_plot_lg +
+  #plot_annotation(tag_levels = 'a') +
   plot_layout(guides='collect', ncol = 2) &

### 2024-01-20 paper.tex
Add benchmark value to remove long edges tunning section

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, enhanced, interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, frame hidden, breakable]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, sharp corners, enhanced, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden, breakable]}{\end{tcolorbox}}\fi
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticpltScurvelwd-1.pdf}
+{\centering \includegraphics{paper_files/figure-pdf/fig-diagnosticpltScurvelwd-1.pdf}
 
@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticpltScurvelgrm-1.pdf}
+Achieving a smooth representation in 2D space is crucial, and one factor
+influencing this smoothness is the presence of long edges. These long
+edges occur when a line connects points that are distant in the
+triangular mesh, impacting the overall smoothness of the mesh.
+
+To investigate the effect of removing long edges on the MSE, we analyzed
+the MSE for various benchmark values. Surprisingly, the results
+indicated that the MSE remained consistent across different benchmark
+values, as shown in Figure~\ref{fig-diagnosticpltScurvelgrm}. It's
+important to note that edges are defined only in the 2D model and do not
+extend to higher dimensions. Consequently, removing edges does not
+affect the model in higher dimensions.
+
+There is no definitive rule for determining the benchmark value to
+remove long edges. However, we have a method to find a default value
+(see Appendix). Adjusting values around this default can help to find
+benchmark value to remove long edges, contributing to the construction
+of a smoother 2D representation.
+
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-diagnosticpltScurvelgrm-1.pdf}
 

### 2024-01-20 paper.qmd
Add model selection section

@@
 If we have done so, the 2D models can be quite different from the data structure shown in high-D space.
-Therefore, visualizing the 2D model in high-D space (m-in-dis) is more useful to answer different types of questions:
+Therefore, visualizing the 2D model in high-D space (m-in-ds) is more useful to answer different types of questions:
 
@@
 The process begins with a real data matrix $X$ containing $n$ observations in $p$ dimensions.
-It generates a sequence of $p$ × $d$ orthonormal projection matrices (bases), usually 1 or 2 dimensions.
+It generates a sequence of $p$ × $d$ orthonormal projection matrices (bases) $A_t$, usually $d$ is $1$ or $2$ dimensions.
 For each pair of orthonormal bases $A_t$ and $A_{t+1}$, a geodesic path is interpolated to create smooth animation between projections.
@@
 <!--### Overview {#sec-algorithm}-->
+
+In this paper, we introduce a novel approach, which is designed to determine which method, and which parameter choice provide the most useful representation of high dimensional data.
+Our approach involves dividing the data set into two subsets: a training set, which serves to construct the low-dimensional manifold, and a test set, utilized to generate predictive values and residuals.
+For the implementation of our approach, a 2D embedding data set serves as the initiation point for the subsequent analysis and visualization.
+The output
+
+write about the S-curve as well.
 
@@
 
-num_bins_for_nldr_methods_df <- tibble::tibble(NLDR_technique = "tSNE", tot_bins = 8)
+num_bins_for_nldr_methods_df <- tibble::tibble(NLDR_technique = "tSNE", tot_bins = 72, aic = -21882.462)
 
 num_bins_for_nldr_methods_df <- num_bins_for_nldr_methods_df |> 
-  tibble::add_row(NLDR_technique = "UMAP", tot_bins = 6)
+  tibble::add_row(NLDR_technique = "UMAP", tot_bins = 98, aic = -20013.266)
 
 num_bins_for_nldr_methods_df <- num_bins_for_nldr_methods_df |> 
-  tibble::add_row(NLDR_technique = "PHATE", tot_bins = 8)
+  tibble::add_row(NLDR_technique = "PHATE", tot_bins = 90, aic = -15036.91)
 
 num_bins_for_nldr_methods_df <- num_bins_for_nldr_methods_df |> 
-  tibble::add_row(NLDR_technique = "TriMAP", tot_bins = 6)
+  tibble::add_row(NLDR_technique = "TriMAP", tot_bins = 154, aic = -23001.90)
 
 num_bins_for_nldr_methods_df <- num_bins_for_nldr_methods_df |> 
-  tibble::add_row(NLDR_technique = "PaCMAP", tot_bins = 10)
+  tibble::add_row(NLDR_technique = "PaCMAP", tot_bins = 132, aic = -21330.059)
 
@@
 num_bins_for_nldr_methods_df$tot_bins <- as.character(num_bins_for_nldr_methods_df$tot_bins)
-colnames(num_bins_for_nldr_methods_df) <- c('NLDR technique', 'total number of bins')
+colnames(num_bins_for_nldr_methods_df) <- c('NLDR technique', 'total number of bins', "AIC")
 num_bins_for_nldr_methods_df <- tibble::remove_rownames(num_bins_for_nldr_methods_df)
 
-xtable(num_bins_for_nldr_methods_df, caption = "Total number of bins suggested by rule of thumb for each NLDR technique applied to S-curve dataset.", label = "tbl-msebinsnldr") |>
+xtable(num_bins_for_nldr_methods_df, caption = "Total number of bins suggested by rule of thumb and AIC values for each NLDR technique applied to S-curve dataset.", label = "tbl-msebinsnldr") |>
   xtable2kable() |> 
   kable_styling(latex_options = c("striped", "HOLD_position"), position = "center", full_width = FALSE) |>
-  row_spec(0, bold = TRUE)
+  row_spec(0, bold = TRUE) |>
+  row_spec(4, bold=T, color = "red")
 
@@
 
-vec <- stats::setNames(rep("", 4), c("number_of_bins", "number_of_observations", "total_error", "total_mse"))  ## Define column names
+vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names
 
@@
 
+  eval_df_training <- eval_df_training |>
+    mutate(num_bins_x = num_bins_vec[i])
+
   eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
@@
   eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+
+  eval_df_training <- eval_df_training |>
+    mutate(num_bins_x = num_bins_vec[i])
 
@@
 
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+
+hex_full_count_df_loop2 <- hex_full_count_df_loop2 |>
+  dplyr::mutate(type = if_else(std_counts < 0.12, "low", "high"))
+
+hex_full_count_df_filtered <- hex_full_count_df_loop2 |>
+  dplyr::filter(type == "low")
+
+p2_n <-  ggplot(data = hex_full_count_df_loop2, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
+  #geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  #coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
+  theme_void() +
+  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6)) +
+  guides(fill = guide_colourbar(title = "Standardized count")) #+
+  #annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
+
+p2_n_rm <-  ggplot(data = hex_full_count_df_loop2, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
+  geom_text(data = hex_full_count_df_filtered, color = "red", aes(x = c_x, y = c_y, group = polygon_id, label = "X", size = 0.5)) +
+  #geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  #coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
+  theme_void() +
+  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6)) +
+  guides(fill = guide_colourbar(title = "Standardized count")) #+
+  #annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
+
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| fig-cap: Goodness of f
+#| label: fig-rmlowdenshex
+##| out-width: 100%
+
+p2_n + p2_n_rm +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='bottom', plot.tag = element_text(size = 8))
+```
+
 <!--Addressing low-density hexagons is a systematic process to handle sparsely represented data in certain regions.
@@
 #| warning: false
-#| fig-cap: Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset with different benchmark values to remove the low-density heaxgons.
-What is the effective benchmark value to remove the low-density heaxgons?
+#| fig-cap: "Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset with different benchmark values to remove the low-density heaxgons.
+What is the effective benchmark value to remove the low-density heaxgons?"
 #| label: fig-diagnosticpltScurvelwd
@@
 However, we have a method to find a default value (see Appendix).
-Adjusting values around this default can help to find benchmark value to remove long edges, contributing to the construction of a smoother 2D representation.
+Adjusting values around this default can help to find benchmark value to remove long edges, contributing to the construction of a smoother 2D representation (see @fig-diagnosticpltScurvelgrm).
 
@@
 
-### Model selection {#sec-selection}
-
-- By comaring AIC what is the best NLDR technique to reprrsnt NLDR data
+#### Model selection {#sec-selection}
+
+In the process of selecting the most suitable NLDR technique to represent high-D data, the Akaike Information Criterion (AIC) stands as a crucial tool for insightful comparisons.
+Once the effective number of bins has been determined for each method, AIC values can be compared to identify the best NLDR technique with the lowest AIC value.
+For example, as shown in @fig-diagnosticpltScurve (a), TriMAP is the most suitable NLDR technique to represent the S-curve data (see table 1).
 

### 2024-01-20 paper.tex
Add model selection section

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, sharp corners, enhanced, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden, breakable]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, frame hidden, interior hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, breakable, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 different from the data structure shown in high-D space. Therefore,
-visualizing the 2D model in high-D space (m-in-dis) is more useful to
+visualizing the 2D model in high-D space (m-in-ds) is more useful to
 answer different types of questions:
@@
 observations in \(p\) dimensions. It generates a sequence of \(p\) ×
-\(d\) orthonormal projection matrices (bases), usually 1 or 2
-dimensions. For each pair of orthonormal bases \(A_t\) and \(A_{t+1}\),
-a geodesic path is interpolated to create smooth animation between
-projections.
+\(d\) orthonormal projection matrices (bases) \(A_t\), usually \(d\) is
+\(1\) or \(2\) dimensions. For each pair of orthonormal bases \(A_t\)
+and \(A_{t+1}\), a geodesic path is interpolated to create smooth
+animation between projections.
 
@@
 \section{Methodology}\label{sec-methods}}
+
+In this paper, we introduce a novel approach, which is designed to
+determine which method, and which parameter choice provide the most
+useful representation of high dimensional data. Our approach involves
+dividing the data set into two subsets: a training set, which serves to
+construct the low-dimensional manifold, and a test set, utilized to
+generate predictive values and residuals. For the implementation of our
+approach, a 2D embedding data set serves as the initiation point for the
+subsequent analysis and visualization. The output
+
+write about the S-curve as well.
 
@@
 \centering
-\begin{tabular}{rll}
+\begin{tabular}{rllr}
   \hline
-\textbf{} & \textbf{NLDR technique} & \textbf{total number of bins }\\ 
+\textbf{} & \textbf{NLDR technique} & \textbf{total number of bins} & \textbf{AIC }\\ 
   \hline
-\cellcolor{gray!6}{1} & \cellcolor{gray!6}{tSNE} & \cellcolor{gray!6}{8 }\\ 
-  2 & UMAP & 6 \\ 
-\cellcolor{gray!6}{  3} & \cellcolor{gray!6}{PHATE} & \cellcolor{gray!6}{8 }\\ 
-  4 & TriMAP & 6 \\ 
-\cellcolor{gray!6}{  5} & \cellcolor{gray!6}{PaCMAP} & \cellcolor{gray!6}{10 }\\ 
+\cellcolor{gray!6}{1} & \cellcolor{gray!6}{tSNE} & \cellcolor{gray!6}{72} & \cellcolor{gray!6}{-21882.46 }\\ 
+  2 & UMAP & 98 & -20013.27 \\ 
+\cellcolor{gray!6}{  3} & \cellcolor{gray!6}{PHATE} & \cellcolor{gray!6}{90} & \cellcolor{gray!6}{-15036.91 }\\ 
+\textcolor{red}{\textbf{  4}} & \textcolor{red}{\textbf{TriMAP}} & \textcolor{red}{\textbf{154}} & \textcolor{red}{\textbf{-23001.90 }}\\ 
+\cellcolor{gray!6}{  5} & \cellcolor{gray!6}{PaCMAP} & \cellcolor{gray!6}{132} & \cellcolor{gray!6}{-21330.06 }\\ 
    \hline
 \end{tabular}
-\caption{Total number of bins suggested by rule of thumb for each NLDR technique applied to S-curve dataset.} 
+\caption{Total number of bins suggested by rule of thumb and AIC values for each NLDR technique applied to S-curve dataset.} 
 \label{tbl-msebinsnldr}
@@
 
+{\centering \includegraphics{paper_files/figure-pdf/fig-rmlowdenshex-1.pdf}
+
+}
+
+\caption{\label{fig-rmlowdenshex}Goodness of f}
+
+\end{figure}
+
+\begin{figure}
+
 {\centering \includegraphics{paper_files/figure-pdf/fig-diagnosticpltScurvelwd-1.pdf}
@@
 benchmark value to remove long edges, contributing to the construction
-of a smoother 2D representation.
+of a smoother 2D representation (see
+Figure~\ref{fig-diagnosticpltScurvelgrm}).
 
@@
 \hypertarget{sec-selection}{%
-\subsection{Model selection}\label{sec-selection}}
-
-\begin{itemize}
-\tightlist
-\item
-  By comaring AIC what is the best NLDR technique to reprrsnt NLDR data
-\end{itemize}
+\subsubsection{Model selection}\label{sec-selection}}
+
+In the process of selecting the most suitable NLDR technique to
+represent high-D data, the Akaike Information Criterion (AIC) stands as
+a crucial tool for insightful comparisons. Once the effective number of
+bins has been determined for each method, AIC values can be compared to
+identify the best NLDR technique with the lowest AIC value. For example,
+as shown in Figure~\ref{fig-diagnosticpltScurve} (a), TriMAP is the most
+suitable NLDR technique to represent the S-curve data (see table 1).
 

### 2024-01-21 paper.qmd
Add intro for methodology section

@@
 
-In this paper, we introduce a novel approach, which is designed to determine which method, and which parameter choice provide the most useful representation of high dimensional data.
-Our approach involves dividing the data set into two subsets: a training set, which serves to construct the low-dimensional manifold, and a test set, utilized to generate predictive values and residuals.
-For the implementation of our approach, a 2D embedding data set serves as the initiation point for the subsequent analysis and visualization.
-The output
-
-write about the S-curve as well.
-
-Our algorithm comprises two main phases: (1) generate the model in the 2D space, and (2) generate the model in the high-D space. These two phases are described in details in this section.
+In this paper, we present a novel method that aims to determine the most effective way to represent high-D data by selecting the best method and hyperparameter choice.
+Our approach involves dividing the data set into two parts: a training set to construct the model and a test set to generate predictive values and residuals.
+To implement our approach, we first use a 2D embedding data set as the initial point.
+The output of our algorithm is a tour that displays the model and original data in the high-dimensional space.
+A flow chart of the proposed algorithm is shown in @fig-meth.
+Our algorithm comprises two main phases: (1) generate the model in the 2D space, and (2) generate the model in the high-D space.
+These two phases are described in detail in this section using UMAP 2D embedding of the S-curve data set, which has seven dimensions, including four noise dimensions.
 

### 2024-01-21 paper.tex
Add intro for methodology section

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, frame hidden, interior hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, breakable, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden, frame hidden, breakable, sharp corners, enhanced]}{\end{tcolorbox}}\fi
 
@@
 
-In this paper, we introduce a novel approach, which is designed to
-determine which method, and which parameter choice provide the most
-useful representation of high dimensional data. Our approach involves
-dividing the data set into two subsets: a training set, which serves to
-construct the low-dimensional manifold, and a test set, utilized to
-generate predictive values and residuals. For the implementation of our
-approach, a 2D embedding data set serves as the initiation point for the
-subsequent analysis and visualization. The output
-
-write about the S-curve as well.
-
-Our algorithm comprises two main phases: (1) generate the model in the
-2D space, and (2) generate the model in the high-D space. These two
-phases are described in details in this section.
+In this paper, we present a novel method that aims to determine the most
+effective way to represent high-D data by selecting the best method and
+hyperparameter choice. Our approach involves dividing the data set into
+two parts: a training set to construct the model and a test set to
+generate predictive values and residuals. To implement our approach, we
+first use a 2D embedding data set as the initial point. The output of
+our algorithm is a tour that displays the model and original data in the
+high-dimensional space. A flow chart of the proposed algorithm is shown
+in Figure~\ref{fig-meth}. Our algorithm comprises two main phases: (1)
+generate the model in the 2D space, and (2) generate the model in the
+high-D space. These two phases are described in detail in this section
+using UMAP 2D embedding of the S-curve data set, which has seven
+dimensions, including four noise dimensions.
 

### 2024-01-21 paper.qmd
Do the chnages in the introduction

@@
 
-To create low-dimensional representations (typically in 2D) of high-D data, it is common to apply dimension reduction (DR) techniques.
+To create low-dimensional representations (typically in 2D) (m-vis) [@article59] of high-D data, it is common to apply dimension reduction (DR) techniques.
 Approaches for DR involve linear methods such as principal component analysis (PCA) [@Karl1901], non-linear methods such as multi-dimensional scaling (MDS) [@Torgerson1967].
@@
 
-Visualization of NLDR techniques (m-vis) [@article59] for the same high-D data is particularly important in understanding and finding the best representation.
-If we have done so, the 2D models can be quite different from the data structure shown in high-D space.
+It is important to visualize various Non-Linear Dimensionality Reduction (NLDR) techniques for the same high-D data in order to understand and find the best representation.
+After doing so, the 2D models may differ considerably from each other and may also deviate from the original data structure in high-dimensional space.
 Therefore, visualizing the 2D model in high-D space (m-in-ds) is more useful to answer different types of questions:
 
-- Is there a best representation of high-D data or are they all providing  equivalent information?
+- Is there a best 2D representation of high-D data or are they all providing  equivalent information?
 Is there a best parameter choice to fit the 2D model?
@@
 
-If we can't easily ask and answer these questions, the ability of understanding the models are restricted.
-To find the best 2D model and the parameter choices, better understanding of the underlying science is really important.
+If we cannot easily ask and answer these questions, our ability to understand the models is limited.
+To find the best 2D model and parameter choices, a better understanding of the underlying science is important.
 
@@
 
-Consider the high-dimensional data a rectangular matrix $X$, where $X = \begin{bmatrix} \textbf{x}_{1} & \textbf{x}_{2} & \cdots & \textbf{x}_{n}\\ \end{bmatrix}^\top$, with $n$ observations in $p$ dimensions.
-The objective is to discover a low-dimensional projection $Y = \begin{bmatrix} \textbf{y}_{1} & \textbf{y}_{2} & \cdots & \textbf{y}_{n}\\ \end{bmatrix}^\top$, represented as an $n$ × $d$ matrix, where $d \ll p$.
+Consider the high-dimensional data a rectangular matrix $X_{n \times p}$, where $X_{n \times p} = \begin{bmatrix} \textbf{x}_{1} & \textbf{x}_{2} & \cdots & \textbf{x}_{n}\\ \end{bmatrix}^\top$, with $n$ observations in $p$ dimensions.
+The objective is to discover a low-dimensional projection $Y_{n \times d} = \begin{bmatrix} \textbf{y}_{1} & \textbf{y}_{2} & \cdots & \textbf{y}_{n}\\ \end{bmatrix}^\top$, represented as an $n$ × $d$ matrix, where $d \ll p$.
 The reduction process seeks to remove noise from the original data set while retaining essential information.
@@
 
-#### Model selection {#sec-selection}
-
-In the process of selecting the most suitable NLDR technique to represent high-D data, the Akaike Information Criterion (AIC) stands as a crucial tool for insightful comparisons.
-Once the effective number of bins has been determined for each method, AIC values can be compared to identify the best NLDR technique with the lowest AIC value.
-For example, as shown in @fig-diagnosticpltScurve (a), TriMAP is the most suitable NLDR technique to represent the S-curve data (see table 1).
-
-### Prediction {#sec-prediction}
-
-In this context, "prediction" denotes the 2D embedding generated for the NLDR technique.
-The methodology encompasses identifying the nearest averaged high-D points in the high-D space for the test data, by computing high-D Euclidean distances.
-As the averaged high-D point corresponds to the lifting of the 2D model, determining its nearest counterpart allows us to map its hexagonal bin centroid coordinates.
-Consequently, these centroid coordinates serve as the assigned low-dimensional predicted values for the test data in the 2D space.
-
-Some NLDR techniques, such as tSNE, often lack a direct prediction, making our approach valuable for generating predicted values in the absence of such functionalities.
-
 ### Model interpretation
@@
 
-Different NLDR techniques perform differently on the S-curve data and tSNE provides the best representation in 2D according to AIC and MSE values (see @fig-diagnosticpltScurve).
+<!--Different NLDR techniques perform differently on the S-curve data and tSNE provides the best representation in 2D according to AIC and MSE values (see @fig-diagnosticpltScurve). -->
 
@@
 ```
+
+### Model selection {#sec-selection}
+
+In the process of selecting the most suitable NLDR technique to represent high-D data, the Akaike Information Criterion (AIC) stands as a crucial tool for insightful comparisons.
+Once the effective number of bins has been determined for each method, AIC values can be compared to identify the best NLDR technique with the lowest AIC value.
+For example, as shown in @fig-diagnosticpltScurve (a), TriMAP is the most suitable NLDR technique to represent the S-curve data (see table 1).
+
+### Prediction {#sec-prediction}
+
+In this context, "prediction" denotes the 2D embedding generated for the NLDR technique.
+The methodology encompasses identifying the nearest averaged high-D points in the high-D space for the test data, by computing high-D Euclidean distances.
+As the averaged high-D point corresponds to the lifting of the 2D model, determining its nearest counterpart allows us to map its hexagonal bin centroid coordinates.
+Consequently, these centroid coordinates serve as the assigned low-dimensional predicted values for the test data in the 2D space.
+
+Some NLDR techniques, such as tSNE, often lack a direct prediction, making our approach valuable for generating predicted values in the absence of such functionalities.
 
@@
 
+- 15 PCs
+
 - learned from the model: There are 3 well-separated clusters in 2D, but if look at the model in high-D space, the three clusters are much closer.
@@
 benchmark_pbmc <- find_benchmark_value(.data = distance_pbmc, distance_col = distance)
+benchmark_pbmc <- 1.072513
 
@@
 trimesh_removed_pbmc_umap <- trimesh_removed_pbmc_umap +
-  geom_point(size = 2, colour = "#33a02c") +
+  geom_point(size = 1, colour = "#33a02c") +
   #ggtitle("(b)") +
@@
 #| fig-cap: "(a) 2D layout from UMAP (n_neighbors = 30, min_dist = 0.3) applied for the PBMC3k dataset.
-Is this a best representation of the original data?, (b)  Model in the 2D space with UMAP (<https://youtu.be/lIA6ZPZAcLE>)"
+Is this a best representation of the original data?, (b)  Model in the 2D space with UMAP (<https://youtu.be/CNXjWZMyh_M>)"
 #| label: fig-nldervisPBMCUMAP

### 2024-01-21 paper.tex
Do the chnages in the introduction

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden, frame hidden, breakable, sharp corners, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, enhanced, interior hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, sharp corners, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 
-To create low-dimensional representations (typically in 2D) of high-D
-data, it is common to apply dimension reduction (DR) techniques.
-Approaches for DR involve linear methods such as principal component
-analysis (PCA) \citep{Karl1901}, non-linear methods such as
-multi-dimensional scaling (MDS) \citep{Torgerson1967}. In the past
-decade, many new non-linear dimension reduction (NLDR) techniques have
-emerged, such as t-distributed stochastic neighbor embedding (tSNE)
-\citep{Laurens2008} and uniform manifold approximation and projection
-(UMAP) \citep{Leland2018}, designed to capture the complex and
-non-linear relationships present within high-D data
+To create low-dimensional representations (typically in 2D) (m-vis)
+\citep{article59} of high-D data, it is common to apply dimension
+reduction (DR) techniques. Approaches for DR involve linear methods such
+as principal component analysis (PCA) \citep{Karl1901}, non-linear
+methods such as multi-dimensional scaling (MDS) \citep{Torgerson1967}.
+In the past decade, many new non-linear dimension reduction (NLDR)
+techniques have emerged, such as t-distributed stochastic neighbor
+embedding (tSNE) \citep{Laurens2008} and uniform manifold approximation
+and projection (UMAP) \citep{Leland2018}, designed to capture the
+complex and non-linear relationships present within high-D data
 \citep{Johnstone2009}. NLDR techniques are the 2D models of high-D data
@@
 
-Visualization of NLDR techniques (m-vis) \citep{article59} for the same
-high-D data is particularly important in understanding and finding the
-best representation. If we have done so, the 2D models can be quite
-different from the data structure shown in high-D space. Therefore,
-visualizing the 2D model in high-D space (m-in-ds) is more useful to
-answer different types of questions:
+It is important to visualize various Non-Linear Dimensionality Reduction
+(NLDR) techniques for the same high-D data in order to understand and
+find the best representation. After doing so, the 2D models may differ
+considerably from each other and may also deviate from the original data
+structure in high-dimensional space. Therefore, visualizing the 2D model
+in high-D space (m-in-ds) is more useful to answer different types of
+questions:
 
@@
 \item
-  Is there a best representation of high-D data or are they all
+  Is there a best 2D representation of high-D data or are they all
   providing equivalent information? Is there a best parameter choice to
@@
 
-If we can't easily ask and answer these questions, the ability of
-understanding the models are restricted. To find the best 2D model and
-the parameter choices, better understanding of the underlying science is
-really important.
+If we cannot easily ask and answer these questions, our ability to
+understand the models is limited. To find the best 2D model and
+parameter choices, a better understanding of the underlying science is
+important.
 
@@
 
-Consider the high-dimensional data a rectangular matrix \(X\), where
-\(X = \begin{bmatrix} \textbf{x}_{1} & \textbf{x}_{2} & \cdots & \textbf{x}_{n}\\ \end{bmatrix}^\top\),
+Consider the high-dimensional data a rectangular matrix
+\(X_{n \times p}\), where
+\(X_{n \times p} = \begin{bmatrix} \textbf{x}_{1} & \textbf{x}_{2} & \cdots & \textbf{x}_{n}\\ \end{bmatrix}^\top\),
 with \(n\) observations in \(p\) dimensions. The objective is to
 discover a low-dimensional projection
-\(Y = \begin{bmatrix} \textbf{y}_{1} & \textbf{y}_{2} & \cdots & \textbf{y}_{n}\\ \end{bmatrix}^\top\),
+\(Y_{n \times d} = \begin{bmatrix} \textbf{y}_{1} & \textbf{y}_{2} & \cdots & \textbf{y}_{n}\\ \end{bmatrix}^\top\),
 represented as an \(n\) × \(d\) matrix, where \(d \ll p\). The reduction
@@
 \end{figure}
-
-\hypertarget{sec-selection}{%
-\subsubsection{Model selection}\label{sec-selection}}
-
-In the process of selecting the most suitable NLDR technique to
-represent high-D data, the Akaike Information Criterion (AIC) stands as
-a crucial tool for insightful comparisons. Once the effective number of
-bins has been determined for each method, AIC values can be compared to
-identify the best NLDR technique with the lowest AIC value. For example,
-as shown in Figure~\ref{fig-diagnosticpltScurve} (a), TriMAP is the most
-suitable NLDR technique to represent the S-curve data (see table 1).
-
-\hypertarget{sec-prediction}{%
-\subsection{Prediction}\label{sec-prediction}}
-
-In this context, ``prediction'' denotes the 2D embedding generated for
-the NLDR technique. The methodology encompasses identifying the nearest
-averaged high-D points in the high-D space for the test data, by
-computing high-D Euclidean distances. As the averaged high-D point
-corresponds to the lifting of the 2D model, determining its nearest
-counterpart allows us to map its hexagonal bin centroid coordinates.
-Consequently, these centroid coordinates serve as the assigned
-low-dimensional predicted values for the test data in the 2D space.
-
-Some NLDR techniques, such as tSNE, often lack a direct prediction,
-making our approach valuable for generating predicted values in the
-absence of such functionalities.
 
@@
 
-Different NLDR techniques perform differently on the S-curve data and
-tSNE provides the best representation in 2D according to AIC and MSE
-values (see Figure~\ref{fig-diagnosticpltScurve}).
-
 \begin{figure}
@@
 
+\hypertarget{sec-selection}{%
+\subsection{Model selection}\label{sec-selection}}
+
+In the process of selecting the most suitable NLDR technique to
+represent high-D data, the Akaike Information Criterion (AIC) stands as
+a crucial tool for insightful comparisons. Once the effective number of
+bins has been determined for each method, AIC values can be compared to
+identify the best NLDR technique with the lowest AIC value. For example,
+as shown in Figure~\ref{fig-diagnosticpltScurve} (a), TriMAP is the most
+suitable NLDR technique to represent the S-curve data (see table 1).
+
+\hypertarget{sec-prediction}{%
+\subsection{Prediction}\label{sec-prediction}}
+
+In this context, ``prediction'' denotes the 2D embedding generated for
+the NLDR technique. The methodology encompasses identifying the nearest
+averaged high-D points in the high-D space for the test data, by
+computing high-D Euclidean distances. As the averaged high-D point
+corresponds to the lifting of the 2D model, determining its nearest
+counterpart allows us to map its hexagonal bin centroid coordinates.
+Consequently, these centroid coordinates serve as the assigned
+low-dimensional predicted values for the test data in the 2D space.
+
+Some NLDR techniques, such as tSNE, often lack a direct prediction,
+making our approach valuable for generating predicted values in the
+absence of such functionalities.
+
 \hypertarget{sec-simpleex}{%
@@
 \item
+  15 PCs
+\item
   learned from the model: There are 3 well-separated clusters in 2D, but
@@
 this a best representation of the original data?, (b) Model in the 2D
-space with UMAP (\url{https://youtu.be/lIA6ZPZAcLE})}
+space with UMAP (\url{https://youtu.be/CNXjWZMyh_M})}
 

### 2024-01-22 paper.qmd
Add selected three parameter settings for PBMC example

@@
 
+- The most important parameter is n_neighbors - the number of approximate nearest neighbors used to construct the initial high-dimensional graph.
+It effectively controls how UMAP balances local versus global structure - low values will push UMAP to focus more on local structure by constraining the number of neighboring points considered when analyzing the data in high dimensions, while high values will push UMAP towards representing the big-picture structure while losing fine detail.
+
+- The second parameter we’ll investigate is min_dist, or the minimum distance between points in low-dimensional space.
+This parameter controls how tightly UMAP clumps points together, with low values leading to more tightly packed embeddings.
+Larger values of min_dist will make UMAP pack points together more loosely, focusing instead on the preservation of the broad topological structure.
+
 ```{r}
@@
 
+<!--UMAP parm 2-->
+
+```{r}
+#| warning: false
+#| echo: false
+
+## Import data
+#df_2 <- read_rds("data/pbmc/pbmc.rds")
+training_data_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_pca_50.rds")
+training_data_pbmc <- training_data_pbmc[, 1:15] |>
+  mutate(ID = 1:NROW(training_data_pbmc))
+#test_1 <- read_rds("data/pbmc/pbmc_test.rds")
+
+#tSNE_pbmc <- read_rds("data/pbmc/pbmc_tsne_30.rds")
+UMAP_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_umap_5_min_dist_0.1.rds")
+# PHATE_pbmc <- read_rds("data/pbmc/pbmc_phate.rds")
+# TriMAP_pbmc <- read_rds("data/pbmc/pbmc_trimap.rds")
+# PaCMAP_pbmc <- read_rds("data/pbmc/pbmc_pacmap.rds")
+
+#load("data/pbmc/pbmc_3k_festem/pbmc3k_label.RData")
+
+## Visualise embeddings
+
+# plot_list1_pbmc <- plot_tSNE_2D(tSNE_pbmc) + #ggtitle("(a)") + 
+#   theme_linedraw() +
+#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+#         axis.title.x = element_blank(), axis.title.y = element_blank(),
+#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+#   annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+class_avg <- UMAP_pbmc |>
+  group_by(cell_label) |>
+  summarise(
+    UMAP1 = median(UMAP1),
+    UMAP2 = median(UMAP2)
+  ) |>
+  mutate(cell_label = as.numeric(cell_label)) |>
+  arrange(cell_label)
+
+class_avg$cell_label <- as.factor(class_avg$cell_label)
+
+levels(class_avg$cell_label) <- c("Memory \nCD4 T", "Naive CD4 T", "CD14+ Mono",  "B", "CD8 T", "FCGR3A+ \n Mono", "NK", "M-MDSC\n-like", "CD27-CD+ \n Memory T", "DC")
+
+plot_list2_pbmc <- UMAP_pbmc |>
+    ggplot(aes(x = UMAP1,
+               y = UMAP2, color = cell_label))+
+    geom_point(alpha=0.5) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
+  geom_text(aes(x=UMAP1,y = UMAP2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) #+
+  # guides(color = guide_legend(title = "Cell type"),
+  #        color = guide_legend(nrow = 3)) #+
+  # annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+# plot_list3_pbmc <- plot_PHATE_2D(PHATE_pbmc) + #ggtitle("(c)") + 
+#   theme_linedraw() +
+#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+#         axis.title.x = element_blank(), axis.title.y = element_blank(),
+#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+#   annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+# plot_list4_pbmc <- plot_TriMAP_2D(TriMAP_pbmc) + #ggtitle("(d)") + 
+#   theme_linedraw() +
+#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+#         axis.title.x = element_blank(), axis.title.y = element_blank(),
+#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+#   annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+# 
+# plot_list5_pbmc <- plot_PaCMAP_2D(PaCMAP_pbmc) + #ggtitle("(e)") + 
+#   theme_linedraw() +
+#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+#         axis.title.x = element_blank(), axis.title.y = element_blank(),
+#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+#   annotate(geom = 'text', label = 'e', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+```
+
+```{r}
+#| warning: false
+#| echo: false
+num_bins_pbmc <- calculate_effective_x_bins(.data = UMAP_pbmc, x = UMAP1,
+                                              cell_area = 1) ##23
+
+shape_value_pbmc <- calculate_effective_shape_value(.data = UMAP_pbmc,
+                                                      x = UMAP1, y = UMAP2) ## 0.8772751
+
+## To extract bin centroids
+hexbin_data_object_pbmc <- extract_hexbin_centroids(nldr_df = UMAP_pbmc,
+                                                      num_bins = num_bins_pbmc,
+                                                      shape_val = shape_value_pbmc, x = UMAP1, y = UMAP2)
+
+df_bin_centroids_pbmc <- hexbin_data_object_pbmc$hexdf_data
+
+UMAP_pbmc_with_hb_id <- UMAP_pbmc |>
+  dplyr::mutate(hb_id = hexbin_data_object_pbmc$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), UMAP_pbmc_with_hb_id)
+
+## Averaged on high-D
+df_bin_pbmc <- avg_highD_data(.data = df_all_pbmc, column_start_text = "PC")
+
+## Triangulate bin centroids
+tr1_object_pbmc <- triangulate_bin_centroids(df_bin_centroids_pbmc, x, y)
+tr_from_to_df_pbmc <- generate_edge_info(triangular_object = tr1_object_pbmc)
+
+## Compute 2D distances
+distance_pbmc <- cal_2D_dist(.data = tr_from_to_df_pbmc)
+
+## To find the benchmark value
+benchmark_pbmc <- find_benchmark_value(.data = distance_pbmc, distance_col = distance)
+benchmark_pbmc <- 1.055271
+
+trimesh_removed_pbmc_umap <- remove_long_edges(.data = distance_pbmc, benchmark_value = benchmark_pbmc,
+                                                 triangular_object = tr1_object_pbmc, distance_col = distance)
+
+trimesh_removed_pbmc_umap <- trimesh_removed_pbmc_umap +
+  geom_point(size = 1, colour = "#33a02c") +
+  #ggtitle("(b)") +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "(a) 2D layout from UMAP (n_neighbors = 5, min_dist = 0.1) applied for the PBMC3k dataset.
+Is this a best representation of the original data?, (b)  Model in the 2D space with UMAP (<https://youtu.be/SO3thJ8PMi8>)"
+#| label: fig-nldervisPBMCUMAP2
+ #| out-width: 100%
+
+plot_list2_pbmc + trimesh_removed_pbmc_umap + 
+plot_layout(ncol=2)
+
+#  plot_list1_pbmc + plot_list3_pbmc + plot_list4_pbmc + plot_list5_pbmc +
+#   plot_layout(ncol=5) #| out-width: 100%
+
+# 2D layouts from different NLDR techniques applied for the training PBMC dataset: (a) tSNE (perplexity = 30), (b) UMAP (n_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2).
+Is there a best representation of the original data or are they all providing  equivalent information?
+```
+
+<!--UMAP parm 3-->
+
+```{r}
+#| warning: false
+#| echo: false
+
+## Import data
+#df_2 <- read_rds("data/pbmc/pbmc.rds")
+training_data_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_pca_50.rds")
+training_data_pbmc <- training_data_pbmc[, 1:15] |>
+  mutate(ID = 1:NROW(training_data_pbmc))
+#test_1 <- read_rds("data/pbmc/pbmc_test.rds")
+
+#tSNE_pbmc <- read_rds("data/pbmc/pbmc_tsne_30.rds")
+UMAP_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_umap_30_min_dist_0.01.rds")
+# PHATE_pbmc <- read_rds("data/pbmc/pbmc_phate.rds")
+# TriMAP_pbmc <- read_rds("data/pbmc/pbmc_trimap.rds")
+# PaCMAP_pbmc <- read_rds("data/pbmc/pbmc_pacmap.rds")
+
+#load("data/pbmc/pbmc_3k_festem/pbmc3k_label.RData")
+
+## Visualise embeddings
+
+# plot_list1_pbmc <- plot_tSNE_2D(tSNE_pbmc) + #ggtitle("(a)") + 
+#   theme_linedraw() +
+#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+#         axis.title.x = element_blank(), axis.title.y = element_blank(),
+#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+#   annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+class_avg <- UMAP_pbmc |>
+  group_by(cell_label) |>
+  summarise(
+    UMAP1 = median(UMAP1),
+    UMAP2 = median(UMAP2)
+  ) |>
+  mutate(cell_label = as.numeric(cell_label)) |>
+  arrange(cell_label)
+
+class_avg$cell_label <- as.factor(class_avg$cell_label)
+
+levels(class_avg$cell_label) <- c("Memory \nCD4 T", "Naive CD4 T", "CD14+ Mono",  "B", "CD8 T", "FCGR3A+ \n Mono", "NK", "M-MDSC\n-like", "CD27-CD+ \n Memory T", "DC")
+
+plot_list2_pbmc <- UMAP_pbmc |>
+    ggplot(aes(x = UMAP1,
+               y = UMAP2, color = cell_label))+
+    geom_point(alpha=0.5) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
+  geom_text(aes(x=UMAP1,y = UMAP2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) #+
+  # guides(color = guide_legend(title = "Cell type"),
+  #        color = guide_legend(nrow = 3)) #+
+  # annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+# plot_list3_pbmc <- plot_PHATE_2D(PHATE_pbmc) + #ggtitle("(c)") + 
+#   theme_linedraw() +
+#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+#         axis.title.x = element_blank(), axis.title.y = element_blank(),
+#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+#   annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+# plot_list4_pbmc <- plot_TriMAP_2D(TriMAP_pbmc) + #ggtitle("(d)") + 
+#   theme_linedraw() +
+#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+#         axis.title.x = element_blank(), axis.title.y = element_blank(),
+#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+#   annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+# 
+# plot_list5_pbmc <- plot_PaCMAP_2D(PaCMAP_pbmc) + #ggtitle("(e)") + 
+#   theme_linedraw() +
+#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+#         axis.title.x = element_blank(), axis.title.y = element_blank(),
+#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+#   annotate(geom = 'text', label = 'e', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+```
+
+```{r}
+#| warning: false
+#| echo: false
+num_bins_pbmc <- calculate_effective_x_bins(.data = UMAP_pbmc, x = UMAP1,
+                                              cell_area = 1) ##23
+
+shape_value_pbmc <- calculate_effective_shape_value(.data = UMAP_pbmc,
+                                                      x = UMAP1, y = UMAP2) ## 0.8772751
+
+## To extract bin centroids
+hexbin_data_object_pbmc <- extract_hexbin_centroids(nldr_df = UMAP_pbmc,
+                                                      num_bins = num_bins_pbmc,
+                                                      shape_val = shape_value_pbmc, x = UMAP1, y = UMAP2)
+
+df_bin_centroids_pbmc <- hexbin_data_object_pbmc$hexdf_data
+
+UMAP_pbmc_with_hb_id <- UMAP_pbmc |>
+  dplyr::mutate(hb_id = hexbin_data_object_pbmc$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), UMAP_pbmc_with_hb_id)
+
+## Averaged on high-D
+df_bin_pbmc <- avg_highD_data(.data = df_all_pbmc, column_start_text = "PC")
+
+## Triangulate bin centroids
+tr1_object_pbmc <- triangulate_bin_centroids(df_bin_centroids_pbmc, x, y)
+tr_from_to_df_pbmc <- generate_edge_info(triangular_object = tr1_object_pbmc)
+
+## Compute 2D distances
+distance_pbmc <- cal_2D_dist(.data = tr_from_to_df_pbmc)
+
+## To find the benchmark value
+benchmark_pbmc <- find_benchmark_value(.data = distance_pbmc, distance_col = distance)
+benchmark_pbmc <- 1.057549
+
+trimesh_removed_pbmc_umap <- remove_long_edges(.data = distance_pbmc, benchmark_value = benchmark_pbmc,
+                                                 triangular_object = tr1_object_pbmc, distance_col = distance)
+
+trimesh_removed_pbmc_umap <- trimesh_removed_pbmc_umap +
+  geom_point(size = 1, colour = "#33a02c") +
+  #ggtitle("(b)") +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "(a) 2D layout from UMAP (n_neighbors = 30, min_dist = 0.01) applied for the PBMC3k dataset.
+Is this a best representation of the original data?, (b)  Model in the 2D space with UMAP (<https://youtu.be/hP8ie_oMXw0>)"
+#| label: fig-nldervisPBMCUMAP3
+ #| out-width: 100%
+
+plot_list2_pbmc + trimesh_removed_pbmc_umap + 
+plot_layout(ncol=2)
+
+#  plot_list1_pbmc + plot_list3_pbmc + plot_list4_pbmc + plot_list5_pbmc +
+#   plot_layout(ncol=5) #| out-width: 100%
+
+# 2D layouts from different NLDR techniques applied for the training PBMC dataset: (a) tSNE (perplexity = 30), (b) UMAP (n_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2).
+Is there a best representation of the original data or are they all providing  equivalent information?
+```
+
 <!--
@@
 
+- The parameter is, in a sense, a guess about the number of close neighbors each point has.
+
 ```{r}

### 2024-01-22 paper.tex
Add selected three parameter settings for PBMC example

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, enhanced, interior hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, sharp corners, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden, enhanced, breakable, frame hidden]}{\end{tcolorbox}}\fi
 
@@
   a good representation for this data set.
+\item
+  The most important parameter is n\_neighbors - the number of
+  approximate nearest neighbors used to construct the initial
+  high-dimensional graph. It effectively controls how UMAP balances
+  local versus global structure - low values will push UMAP to focus
+  more on local structure by constraining the number of neighboring
+  points considered when analyzing the data in high dimensions, while
+  high values will push UMAP towards representing the big-picture
+  structure while losing fine detail.
+\item
+  The second parameter we'll investigate is min\_dist, or the minimum
+  distance between points in low-dimensional space. This parameter
+  controls how tightly UMAP clumps points together, with low values
+  leading to more tightly packed embeddings. Larger values of min\_dist
+  will make UMAP pack points together more loosely, focusing instead on
+  the preservation of the broad topological structure.
 \end{itemize}
@@
 
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-nldervisPBMCUMAP2-1.pdf}
+
+}
+
+\caption{\label{fig-nldervisPBMCUMAP2}(a) 2D layout from UMAP
+(n\_neighbors = 5, min\_dist = 0.1) applied for the PBMC3k dataset. Is
+this a best representation of the original data?, (b) Model in the 2D
+space with UMAP (\url{https://youtu.be/SO3thJ8PMi8})}
+
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-nldervisPBMCUMAP3-1.pdf}
+
+}
+
+\caption{\label{fig-nldervisPBMCUMAP3}(a) 2D layout from UMAP
+(n\_neighbors = 30, min\_dist = 0.01) applied for the PBMC3k dataset. Is
+this a best representation of the original data?, (b) Model in the 2D
+space with UMAP (\url{https://youtu.be/hP8ie_oMXw0})}
+
+\end{figure}
+
 \hypertarget{harps-gto-sample-dataset}{%
@@
   is also not captured in tSNE. Therefore, this is not a good choice.
+\item
+  The parameter is, in a sense, a guess about the number of close
+  neighbors each point has.
 \end{itemize}

### 2024-01-22 paper.qmd
Add picorial representation for remove lond edges

@@
 #| label: fig-diagnosticpltScurve
-#| out-width: 100%
+##| out-width: 100%
 
@@
 
-Achieving a smooth representation in 2D space is crucial, and one factor influencing this smoothness is the presence of long edges.
+Achieving a smooth representation in 2D space is crucial, and one factor influencing this smoothness is the presence of long edges (see @fig-modelScurvermlgimp).
 These long edges occur when a line connects points that are distant in the triangular mesh, impacting the overall smoothness of the mesh.
@@
 This refinement step contributes to enhancing the quality of the triangular mesh, ensuring a more accurate representation of the data structure.-->
+
+```{r}
+#| warning: false
+#| echo: false
+
+## Import data
+
+UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap.rds")
+
+## UMAP
+
+num_bins_umap_s_curve <- 6
+shape_val_umap_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
+                                                          x = UMAP1, y = UMAP2) ## 1.259938
+## To extract bin centroids
+hexbin_data_object_umap_s_curve <- extract_hexbin_centroids(nldr_df = UMAP_s_curve, num_bins = num_bins_umap_s_curve, shape_val = shape_val_umap_s_curve, x = UMAP1, y = UMAP2)
+
+df_bin_centroids_umap_s_curve <- hexbin_data_object_umap_s_curve$hexdf_data
+
+UMAP_data_with_hb_id_s_curve <- UMAP_s_curve |>
+  dplyr::mutate(hb_id = hexbin_data_object_umap_s_curve$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_umap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), UMAP_data_with_hb_id_s_curve)
+
+## Averaged on high-D
+df_bin_umap_s_curve <- avg_highD_data(.data = df_all_umap_s_curve)
+
+## Triangulate bin centroids
+tr1_object_umap_s_curve <- triangulate_bin_centroids(df_bin_centroids_umap_s_curve, x, y)
+tr_from_to_df_umap_s_curve <- generate_edge_info(triangular_object = tr1_object_umap_s_curve)
+
+# ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
+#   geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
+#   geom_point(size = 2, colour = "#33a02c") +
+#   coord_equal()
+
+## Compute 2D distances
+distance_umap_s_curve <- cal_2D_dist(.data = tr_from_to_df_umap_s_curve)
+
+## To find the benchmark value
+benchmark_umap_s_curve <- find_benchmark_value(.data = distance_umap_s_curve, distance_col = distance)
+
+# colour_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve,
+#                   triangular_object = tr1_object_umap_s_curve, distance_col = distance)
+
+trimesh_umap_s_curve <- ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
+  geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
+  geom_point(size = 2, colour = "#33a02c") +
+  coord_equal()
+
+trimesh_umap_s_curve <- trimesh_umap_s_curve +
+  #ggtitle("(a)") +
+  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  theme_light() +
+  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
+  ) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
+# theme(axis.text = element_text(size = 5),
+#       axis.title = element_text(size = 7))
+
+trimesh_gr_umap_s_curve <- colour_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve, triangular_object = tr1_object_umap_s_curve, distance_col = distance)
+
+trimesh_gr_umap_s_curve <- trimesh_gr_umap_s_curve +
+  geom_point(size = 2, colour = "#33a02c") +
+  #ggtitle("(b)") +
+  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  theme_light() +
+  #coord_equal() +
+  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
+  ) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+# ### Define type column
+# df <- df_all_umap_s_curve |>
+#   dplyr::select(tidyselect::starts_with("x")) |>
+#   dplyr::mutate(type = "data") ## original dataset
+# 
+# df_b <- df_bin_umap_s_curve |>
+#   dplyr::filter(hb_id %in% df_bin_centroids_umap_s_curve$hexID) |>
+#   dplyr::select(-hb_id) |>
+#   dplyr::mutate(type = "model") ## Data with summarized mean
+# 
+# df_exe <- dplyr::bind_rows(df_b, df)
+# 
+# distance_df_small_edges <- distance_umap_s_curve %>%
+#   dplyr::filter(distance < benchmark_umap_s_curve)
+# 
+# distance_df_long_edges <- distance_umap_s_curve %>%
+#   dplyr::filter(distance >= benchmark_umap_s_curve)
+# 
+# distance_df_edges <- dplyr::bind_rows(distance_df_small_edges, distance_df_long_edges)
+# 
+# langevitour::langevitour(df_exe[1:(length(df_exe)-1)], lineFrom = distance_df_edges$from , lineTo = distance_df_edges$to, group = df_exe$type, pointSize = 3, levelColors = c("#6a3d9a", "#33a02c"), lineColors = append(rep("black", length(distance_df_small_edges$from)), rep("red", length(distance_df_long_edges$from))))
+#
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+#| label: fig-modelScurvermlgimp
+#| fig-cap: "What does the 2D model, constructed using UMAP (applied to the S-curve dataset), look like in high dimensions with long edges?
+(<https://youtu.be/pSL_G_n7-fM>) (a) 2D model without removing the long edges, and (b) 2D model with long edges coloured by red.
+The default benchmark value is used to identify long edges."
+##| out-width: 100%
+
+trimesh_umap_s_curve + trimesh_gr_umap_s_curve +
+  #plot_annotation(tag_levels = 'a') +
+  plot_layout(ncol = 2) &
+  theme(legend.position='none', plot.tag = element_text(size = 8))
+```
 
@@
 #| warning: false
-#| fig-cap: Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset with different benchmark values to remove the long edges.
-What is the effective benchmark value to remove the long edges?
+#| fig-cap: "Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset with different benchmark values to remove the long edges.
+What is the effective benchmark value to remove the long edges?"
 #| label: fig-diagnosticpltScurvelgrm

### 2024-01-22 paper.tex
Add picorial representation for remove lond edges

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden, enhanced, breakable, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, interior hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden, enhanced]}{\end{tcolorbox}}\fi
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticpltScurve-1.pdf}
+{\centering \includegraphics{paper_files/figure-pdf/fig-diagnosticpltScurve-1.pdf}
 
@@
 Achieving a smooth representation in 2D space is crucial, and one factor
-influencing this smoothness is the presence of long edges. These long
-edges occur when a line connects points that are distant in the
-triangular mesh, impacting the overall smoothness of the mesh.
+influencing this smoothness is the presence of long edges (see
+Figure~\ref{fig-modelScurvermlgimp}). These long edges occur when a line
+connects points that are distant in the triangular mesh, impacting the
+overall smoothness of the mesh.
 
@@
 Figure~\ref{fig-diagnosticpltScurvelgrm}).
+
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-modelScurvermlgimp-1.pdf}
+
+}
+
+\caption{\label{fig-modelScurvermlgimp}What does the 2D model,
+constructed using UMAP (applied to the S-curve dataset), look like in
+high dimensions with long edges? (\url{https://youtu.be/pSL_G_n7-fM})
+(a) 2D model without removing the long edges, and (b) 2D model with long
+edges coloured by red. The default benchmark value is used to identify
+long edges.}
+
+\end{figure}
 

### 2024-01-23 paper.qmd
Rewrite the tour section

@@
 
-A tour is a powerful visualization technique used to explore high-dimensional data by generating a sequence of projections, typically into two dimensions.
-There are two main types of tours: the Grand Tour and the Guided Tour.
-The Grand Tour explores the data's shape and global structure by using random projections [@Asimov1985].
-In contrast, the Guided Tour focuses on specific patterns by moving towards interesting projections defined by a predefined index function [@article29].
-
-The process begins with a real data matrix $X$ containing $n$ observations in $p$ dimensions.
-It generates a sequence of $p$ × $d$ orthonormal projection matrices (bases) $A_t$, usually $d$ is $1$ or $2$ dimensions.
-For each pair of orthonormal bases $A_t$ and $A_{t+1}$, a geodesic path is interpolated to create smooth animation between projections.
-
-In the Grand Tour, new orthonormal bases are randomly chosen to explore the $d$-dimensional subspace.
-The data is often sphered via principal components to reduce dimensionality.
-The Guided Tour uses a predefined index function to generate a sequence of 'interesting' projections.
-The resulting tour continuously visualizes the projected data $Y_t$ = $XA_t$ as it interpolates between successive bases.
-
-While both tours can be used to visualize data, examples often focus on using the Grand Tour to observe global structures.
+A tour is a powerful visualization technique used to explore the shape and global structure of high-dimensional data by generating a sequence of projections, typically into two dimensions.
+There are two main types of tours: the grand tour [@Asimov1985] and the guided tour [@article29].
+A grand tour involves randomly selecting new orthonormal bases, enabling users to understand the structure by exploring the subspace of d-dimensional projections [@Asimov1985].
+In contrast, a guided tour can be employed to generate a sequence of 'interesting' projections based on an index function [@article29].
+
+The process begins with the data matrix $X$.
+It generates a sequence of $p$ × $d$ orthonormal projection matrices (bases) $P_t$, usually $d$ is one or two dimensions.
+For each pair of orthonormal bases $P_t$ and $P_{t+1}$, a geodesic path is interpolated to create smooth animation between projections.
+The resulting tour continuously visualizes the projected data $Y_t$ = $XP_t$ as it interpolates between successive bases.
+
+Furthermore, software like **langevitour** can visualize both types of tours, providing flexibility for exploring high-dimensional data with various objectives.
+In our context, use grand tour along with the model to observe how effectively the model captures the underlying structure of the data.
+
+<!--While both tours can be used to visualize data, examples often focus on using the grand tour to observe global structures.
 However, software like **langevitour** can visualize both types of tours, providing flexibility for exploring high-dimensional data with various objectives.
+-->
 
@@
 
+## Import data
+
+UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap.rds")
+
+## UMAP
+
+num_bins_umap_s_curve <- 11
+shape_val_umap_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
+                                                          x = UMAP1, y = UMAP2) ## 1.259938
+## To extract bin centroids
+hexbin_data_object_umap_s_curve <- extract_hexbin_centroids(nldr_df = UMAP_s_curve, num_bins = num_bins_umap_s_curve, shape_val = shape_val_umap_s_curve, x = UMAP1, y = UMAP2)
+
+df_bin_centroids_umap_s_curve <- hexbin_data_object_umap_s_curve$hexdf_data
+
+df_bin_centroids_umap_s_curve <- df_bin_centroids_umap_s_curve |>
+  dplyr::filter(std_counts > 0.12)
+
+## Triangulate bin centroids
+tr1_object_umap_s_curve <- triangulate_bin_centroids(df_bin_centroids_umap_s_curve, x, y)
+tr_from_to_df_umap_s_curve <- generate_edge_info(triangular_object = tr1_object_umap_s_curve)
+
+## Compute 2D distances
+distance_umap_s_curve <- cal_2D_dist(.data = tr_from_to_df_umap_s_curve)
+
+## To find the benchmark value
+benchmark_umap_s_curve <- find_benchmark_value(.data = distance_umap_s_curve, distance_col = distance)
+
+trimesh_umap_s_curve_rm_low <- ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
+  geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
+  geom_point(size = 2, colour = "#33a02c") +
+  coord_equal()
+
+trimesh_umap_s_curve_rm_low <- trimesh_umap_s_curve_rm_low +
+  coord_cartesian(xlim =c(-5, 7.5), ylim = c(-10, 10)) +
+  #ggtitle("(a)") +
+  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  theme_light() +
+  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
+  ) #+
+  #annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
 ```
@@
 
-p2_n + p2_n_rm +
+p2_n + p2_n_rm + trimesh_umap_s_curve_rm_low +
   plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', ncol = 2) &
+  plot_layout(guides='collect', ncol = 3) &
   theme(legend.position='bottom', plot.tag = element_text(size = 8))
@@
   guides(color = guide_legend(title = "cluster"),
-         color = guide_legend(nrow = 3)) #+
-  # annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+         color = guide_legend(nrow = 3)) +
+ annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
 
@@
         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
 

### 2024-01-23 paper.tex
Rewrite the tour section

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, interior hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, frame hidden, enhanced, breakable, boxrule=0pt, interior hidden, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 
-A tour is a powerful visualization technique used to explore
-high-dimensional data by generating a sequence of projections, typically
-into two dimensions. There are two main types of tours: the Grand Tour
-and the Guided Tour. The Grand Tour explores the data's shape and global
-structure by using random projections \citep{Asimov1985}. In contrast,
-the Guided Tour focuses on specific patterns by moving towards
-interesting projections defined by a predefined index function
-\citep{article29}.
-
-The process begins with a real data matrix \(X\) containing \(n\)
-observations in \(p\) dimensions. It generates a sequence of \(p\) ×
-\(d\) orthonormal projection matrices (bases) \(A_t\), usually \(d\) is
-\(1\) or \(2\) dimensions. For each pair of orthonormal bases \(A_t\)
-and \(A_{t+1}\), a geodesic path is interpolated to create smooth
-animation between projections.
-
-In the Grand Tour, new orthonormal bases are randomly chosen to explore
-the \(d\)-dimensional subspace. The data is often sphered via principal
-components to reduce dimensionality. The Guided Tour uses a predefined
-index function to generate a sequence of `interesting' projections. The
-resulting tour continuously visualizes the projected data \(Y_t\) =
-\(XA_t\) as it interpolates between successive bases.
-
-While both tours can be used to visualize data, examples often focus on
-using the Grand Tour to observe global structures. However, software
-like \textbf{langevitour} can visualize both types of tours, providing
-flexibility for exploring high-dimensional data with various objectives.
+A tour is a powerful visualization technique used to explore the shape
+and global structure of high-dimensional data by generating a sequence
+of projections, typically into two dimensions. There are two main types
+of tours: the grand tour \citep{Asimov1985} and the guided tour
+\citep{article29}. A grand tour involves randomly selecting new
+orthonormal bases, enabling users to understand the structure by
+exploring the subspace of d-dimensional projections \citep{Asimov1985}.
+In contrast, a guided tour can be employed to generate a sequence of
+`interesting' projections based on an index function \citep{article29}.
+
+The process begins with the data matrix \(X\). It generates a sequence
+of \(p\) × \(d\) orthonormal projection matrices (bases) \(P_t\),
+usually \(d\) is one or two dimensions. For each pair of orthonormal
+bases \(P_t\) and \(P_{t+1}\), a geodesic path is interpolated to create
+smooth animation between projections. The resulting tour continuously
+visualizes the projected data \(Y_t\) = \(XP_t\) as it interpolates
+between successive bases.
+
+Furthermore, software like \textbf{langevitour} can visualize both types
+of tours, providing flexibility for exploring high-dimensional data with
+various objectives. In our context, use grand tour along with the model
+to observe how effectively the model captures the underlying structure
+of the data.
 

### 2024-01-23 paper.qmd
Rewrite the DR section

@@
 Approaches for DR involve linear methods such as principal component analysis (PCA) [@Karl1901], non-linear methods such as multi-dimensional scaling (MDS) [@Torgerson1967].
-In the past decade, many new non-linear dimension reduction (NLDR) techniques have emerged, such as t-distributed stochastic neighbor embedding (tSNE) [@Laurens2008] and uniform manifold approximation and projection (UMAP) [@Leland2018], designed to capture the complex and non-linear relationships present within high-D data [@Johnstone2009].
+In the past decade, many new non-linear dimension reduction (NLDR) techniques have emerged, such as t-distributed stochastic neighbor embedding (tSNE) [@Laurens2008] and uniform manifold approximation and projection (UMAP) [@Leland2018].
 NLDR techniques are the 2D models of high-D data in our context.
 
-It is important to visualize various Non-Linear Dimensionality Reduction (NLDR) techniques for the same high-D data in order to understand and find the best representation.
+It is important to visualize various non-linear dimensionality reduction (NLDR) techniques for the same high-D data in order to understand and find the best representation.
 After doing so, the 2D models may differ considerably from each other and may also deviate from the original data structure in high-dimensional space.
@@
 
-Consider the high-dimensional data a rectangular matrix $X_{n \times p}$, where $X_{n \times p} = \begin{bmatrix} \textbf{x}_{1} & \textbf{x}_{2} & \cdots & \textbf{x}_{n}\\ \end{bmatrix}^\top$, with $n$ observations in $p$ dimensions.
+Consider the high-D data a rectangular matrix $X_{n \times p}$, where $X_{n \times p} = \begin{bmatrix} \textbf{x}_{1} & \textbf{x}_{2} & \cdots & \textbf{x}_{n}\\ \end{bmatrix}^\top$, with $n$ observations in $p$ dimensions.
 The objective is to discover a low-dimensional projection $Y_{n \times d} = \begin{bmatrix} \textbf{y}_{1} & \textbf{y}_{2} & \cdots & \textbf{y}_{n}\\ \end{bmatrix}^\top$, represented as an $n$ × $d$ matrix, where $d \ll p$.
@@
 PCA performs an eigen-decomposition of the sample covariance matrix to obtain orthogonal principal components that capture the variance of the data [@Karl1901].
-However, linear methods may not fully capture complex non-linear relationships present in the data.
 
@@
 Variants of MDS include non-metric scaling [@article62] and Isomap, which estimate geodesic distances to create the low-dimensional representation [@article63].
-Other approaches based on diffusion processes, like diffusion maps [@article64] and the PHATE algorithm [@article03], also fall under NLDR methods.
-
-A challenge with NLDR methods is selecting and tuning appropriate parameters.
+Other approaches based on diffusion processes, like diffusion maps [@article64] and the PHATE (Potential of Heat-diffusion for Affinity-based Trajectory Embedding) algorithm [@article03], also fall under NLDR methods.
+
+<!--Their focus is on capturing complex structures that may not be easily represented in the original space, making a straightforward reverse mapping challenging.
+
 One specific technique we focus on is Pairwise Controlled Manifold Approximation (PaCMAP).
@@
 
-It is important to note that methods like PCA and auto-encoders [@article65] provide a reverse mapping from the low-dimensional space back to the high-dimensional space, enabling data reconstruction.
-However, many non-linear methods, including tSNE, prioritize visualization and exploration over reconstruction.
-Their focus is on capturing complex structures that may not be easily represented in the original space, making a straightforward reverse mapping challenging.
+However, linear methods may not fully capture complex non-linear relationships present in the data.-->
 
@@
 
-Non-linear dimension reduction techniques play a crucial role in the analysis and visualization of high-dimensional data, where the complexities of relationships among variables may not be adequately captured by linear methods.
-Among these techniques, tSNE stands out for its emphasis on preserving pairwise distances.
+NLDR techniques are crucial for analyzing and displaying high-dimensional data, where linear approaches may not adequately capture complexities in relationships between variables [@Johnstone2009].
+One of the challenges with NLDR techniques is the selection and tuning of appropriate hyperparameters [@liao2023].
+This process involves finding the suitable combination of hyperparameters that enhances the performance of the NLDR technique, considering the characteristics of the dataset and the specific goals of the analysis.
+
+<!--This process requires a delicate balance to optimize the performance of NLDR techniques, taking into account the intricacies of the dataset and the specific goals of the analysis.-->
+
+Additionally, another challenge is lack of reverse mapping.
+Techniques like PCA and auto-encoders [@article65] provide a way to map back from the low-dimensional space to the high-D space, facilitating data reconstruction.
+However, some NLDR methods, such as tSNE, don't have a specific way to reconstruct the original data from the low-dimensional space.
+
+In this article, mainly focus on five NLDR techniques. They are tSNE, UMAP, PHATE, TriMAP [@article02], and Pairwise Controlled Manifold Approximation (PaCMAP) [@article01].
+
+Among these, tSNE [@Laurens2008] stands out for its ability to preserve pairwise distances.
+By minimizing the divergence between probability distributions in both high and low-dimensional spaces, tSNE effectively uncovers intricate structures and patterns within the data.
+Its application is widespread, particularly in tasks requiring the visualization of clusters and local relationships.
+However, achieving effective results requires careful consideration of hyperparameters, such as perplexity.
+
+UMAP [@Leland2018] is a useful technique for simplifying data while maintaining both local and overall structures.
+It builds a fuzzy topological view by considering nearby data points and then optimizes a simplified version to match that view.
+UMAP is known for working well with different scales of relationships in data and is efficient in handling large datasets.
+However, it's important to choose parameters like neighbors and minimum distance carefully, as they can affect the results.
+
+Furthermore, PHATE [@article03] is great for understanding how things develop, especially in single-cell genomics.
+It uses a heat diffusion process to capture relationships between data points, like points along a trajectory.
+While PHATE is excellent for revealing these developmental structures, it requires careful tuning of its parameters because of its specialized focus.
+
+Additionally, TriMAP [@article02] takes a special approach by creating a triangulated graph representation of the data.
+This method is good at understanding both local and global structures by treating the data as a network of triangles.
+TriMAP is powerful in capturing complicated structures, but it's important to choose parameters carefully, like deciding how many neighbors to consider.
+
+PaCMAP [@article01] is different because it adds supervised learning to make a 2D representation while keeping the relationships between pairs of points.
+It builds a graph using distances between pairs and then makes the 2D representation better using a customizable loss function.
+What's special about PaCMAP is that it can use class labels or extra information to guide how it makes the 2D representation.
+This gives users a way to change how PaCMAP works to fit their needs better.
+
+<!--Among these techniques, tSNE stands out for its emphasis on preserving pairwise distances.
 By minimizing the divergence between probability distributions in both the high and low-dimensional spaces, t-SNE effectively reveals intricate structures and patterns within the data.
@@
 Constructing a graph based on pair-wise distances, PaCMAP optimizes an embedding using a customizable loss function.
-Particularly notable is PaCMAP's flexibility in incorporating class labels or additional information to guide the embedding process, offering users a means to customize its behavior and performance.
+Particularly notable is PaCMAP's flexibility in incorporating class labels or additional information to guide the embedding process, offering users a means to customize its behavior and performance.-->
 

### 2024-01-23 paper.tex
Rewrite the DR section

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, frame hidden, enhanced, breakable, boxrule=0pt, interior hidden, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, breakable, interior hidden, enhanced, boxrule=0pt, frame hidden, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 embedding (tSNE) \citep{Laurens2008} and uniform manifold approximation
-and projection (UMAP) \citep{Leland2018}, designed to capture the
-complex and non-linear relationships present within high-D data
-\citep{Johnstone2009}. NLDR techniques are the 2D models of high-D data
-in our context.
-
-It is important to visualize various Non-Linear Dimensionality Reduction
+and projection (UMAP) \citep{Leland2018}. NLDR techniques are the 2D
+models of high-D data in our context.
+
+It is important to visualize various non-linear dimensionality reduction
 (NLDR) techniques for the same high-D data in order to understand and
@@
 
-Consider the high-dimensional data a rectangular matrix
-\(X_{n \times p}\), where
+Consider the high-D data a rectangular matrix \(X_{n \times p}\), where
 \(X_{n \times p} = \begin{bmatrix} \textbf{x}_{1} & \textbf{x}_{2} & \cdots & \textbf{x}_{n}\\ \end{bmatrix}^\top\),
@@
 obtain orthogonal principal components that capture the variance of the
-data \citep{Karl1901}. However, linear methods may not fully capture
-complex non-linear relationships present in the data.
+data \citep{Karl1901}.
 
@@
 \citep{article63}. Other approaches based on diffusion processes, like
-diffusion maps \citep{article64} and the PHATE algorithm
+diffusion maps \citep{article64} and the PHATE (Potential of
+Heat-diffusion for Affinity-based Trajectory Embedding) algorithm
 \citep{article03}, also fall under NLDR methods.
-
-A challenge with NLDR methods is selecting and tuning appropriate
-parameters. One specific technique we focus on is Pairwise Controlled
-Manifold Approximation (PaCMAP). Similar considerations apply to related
-methods like tSNE \citep{Laurens2008}, UMAP \citep{Leland2018}, and
-TrMAP \citep{article02}.
-
-It is important to note that methods like PCA and auto-encoders
-\citep{article65} provide a reverse mapping from the low-dimensional
-space back to the high-dimensional space, enabling data reconstruction.
-However, many non-linear methods, including tSNE, prioritize
-visualization and exploration over reconstruction. Their focus is on
-capturing complex structures that may not be easily represented in the
-original space, making a straightforward reverse mapping challenging.
 
@@
 
-Non-linear dimension reduction techniques play a crucial role in the
-analysis and visualization of high-dimensional data, where the
-complexities of relationships among variables may not be adequately
-captured by linear methods. Among these techniques, tSNE stands out for
-its emphasis on preserving pairwise distances. By minimizing the
-divergence between probability distributions in both the high and
-low-dimensional spaces, t-SNE effectively reveals intricate structures
-and patterns within the data. Its application is widespread in tasks
-requiring the visualization of clusters and local relationships, though
-it does require careful consideration of the perplexity parameter for
-optimal results.
-
-UMAP is another powerful non-linear technique that strikes a balance
-between preserving local and global structures. Constructing a fuzzy
-topological representation using a weighted k-nearest neighbors graph,
-UMAP optimizes the low-dimensional embedding to resemble this
-representation. Known for its efficiency and scalability, UMAP is
-versatile across various scales of relationships in the data, although
-parameter sensitivity, particularly concerning the choice of neighbors,
-must be taken into account.
-
-For trajectory data, PHATE provides specialized capabilities. It models
-the affinity between data points, simulating a heat diffusion process to
-capture developmental processes, particularly in single-cell genomics.
-While PHATE excels in revealing trajectory structures and offering
-insights into cellular development, it necessitates careful parameter
-tuning due to its specialized nature.
-
-TriMAP adopts a unique approach by approximating the data manifold
-through the construction of a triangulated graph representation. This
-technique efficiently captures both global and local structures by
-representing the data as a network of triangles. TriMAP's strength lies
-in its ability to efficiently capture complex structures, albeit with
-sensitivity to parameter choices, including the number of neighbors.
-
-In contrast, PaCMAP introduces supervised learning to create a
-low-dimensional representation while preserving pair-wise relationships.
-Constructing a graph based on pair-wise distances, PaCMAP optimizes an
-embedding using a customizable loss function. Particularly notable is
-PaCMAP's flexibility in incorporating class labels or additional
-information to guide the embedding process, offering users a means to
-customize its behavior and performance.
+NLDR techniques are crucial for analyzing and displaying
+high-dimensional data, where linear approaches may not adequately
+capture complexities in relationships between variables
+\citep{Johnstone2009}. One of the challenges with NLDR techniques is the
+selection and tuning of appropriate hyperparameters \citep{liao2023}.
+This process involves finding the suitable combination of
+hyperparameters that enhances the performance of the NLDR technique,
+considering the characteristics of the dataset and the specific goals of
+the analysis.
+
+Additionally, another challenge is lack of reverse mapping. Techniques
+like PCA and auto-encoders \citep{article65} provide a way to map back
+from the low-dimensional space to the high-D space, facilitating data
+reconstruction. However, some NLDR methods, such as tSNE, don't have a
+specific way to reconstruct the original data from the low-dimensional
+space.
+
+In this article, mainly focus on five NLDR techniques. They are tSNE,
+UMAP, PHATE, TriMAP \citep{article02}, and Pairwise Controlled Manifold
+Approximation (PaCMAP) \citep{article01}.
+
+Among these, tSNE \citep{Laurens2008} stands out for its ability to
+preserve pairwise distances. By minimizing the divergence between
+probability distributions in both high and low-dimensional spaces, tSNE
+effectively uncovers intricate structures and patterns within the data.
+Its application is widespread, particularly in tasks requiring the
+visualization of clusters and local relationships. However, achieving
+effective results requires careful consideration of hyperparameters,
+such as perplexity.
+
+UMAP \citep{Leland2018} is a useful technique for simplifying data while
+maintaining both local and overall structures. It builds a fuzzy
+topological view by considering nearby data points and then optimizes a
+simplified version to match that view. UMAP is known for working well
+with different scales of relationships in data and is efficient in
+handling large datasets. However, it's important to choose parameters
+like neighbors and minimum distance carefully, as they can affect the
+results.
+
+Furthermore, PHATE \citep{article03} is great for understanding how
+things develop, especially in single-cell genomics. It uses a heat
+diffusion process to capture relationships between data points, like
+points along a trajectory. While PHATE is excellent for revealing these
+developmental structures, it requires careful tuning of its parameters
+because of its specialized focus.
+
+Additionally, TriMAP \citep{article02} takes a special approach by
+creating a triangulated graph representation of the data. This method is
+good at understanding both local and global structures by treating the
+data as a network of triangles. TriMAP is powerful in capturing
+complicated structures, but it's important to choose parameters
+carefully, like deciding how many neighbors to consider.
+
+PaCMAP \citep{article01} is different because it adds supervised
+learning to make a 2D representation while keeping the relationships
+between pairs of points. It builds a graph using distances between pairs
+and then makes the 2D representation better using a customizable loss
+function. What's special about PaCMAP is that it can use class labels or
+extra information to guide how it makes the 2D representation. This
+gives users a way to change how PaCMAP works to fit their needs better.
 

### 2024-01-23 paper.qmd
Add what I learn from the models for simple example with Gaussian clusters

@@
 
-We *strongly* recommend viewing the linked videos for each study while reading.
+In the 2D representations created by all NLDR techniques, as shown in @fig-nldervis5Gau, except for PHATE, there are five distinct clusters.
+In tSNE, these five clusters are closely located to each other (see @fig-nldervis5Gau (a)).
+However, in PHATE, two clusters are closely positioned, while the other three are more distant (see @fig-nldervis5Gau (c)).
+In PaCMAP, one cluster is at the center, and the remaining four are positioned in four different directions (see @fig-nldervis5Gau (e)).
+In TriMAP, two clusters are close, although not as close as in PHATE, and the other three are well-separated (see @fig-nldervis5Gau (d)).
+In UMAP, all clusters are arranged in a parallel manner, with three in one line and the other two in a separate line (see @fig-nldervis5Gau (b)).
+
+Visualizing the models alongside the original high-D data provides insights into how different techniques capture the underlying clustering structure.
+The tSNE model exhibits five well-separated clusters, effectively preserving both local and global structures (see video link of @fig-modelfiveGau (a)).
+UMAP, on the other hand, also presents five distinct clusters but with a more flattened surface appearance (see video link of @fig-modelfiveGau (b)).
+The PHATE model shows five separated clusters resembling triangles or partial triangles but struggles to capture local structures (see video link of @fig-modelfiveGau (c)).
+TriMAP and PaCMAP both display five well-separated clusters with flat surfaces, each capturing within-cluster variation to varying extents.
+Despite the various 2D representations, all NLDR techniques preserve the global structure (see video link of @fig-modelfiveGau (d), (e) respectively).
+However, tSNE, effectively capture both local and global structures, as indicated by lower AIC values in @fig-diagnosticpltGau (a).
+
+<!--By visualizing the models with original data in high-D space, we are able to understand how various techniques capture the clustering data structure.
+The tSNE model exhibits five well-separated clusters, effectively preserving both local and global structures.
+UMAP, on the other hand, also presents five distinct clusters but with a more flattened surface appearance.
+The PHATE model shows five clusters resembling triangles but struggles to capture local structures.
+TriMAP and PaCMAP both display five well-separated clusters with a flat surface appearance, capturing within-cluster variation to varying extents.
+As we can see, even though all NLDR techniques have different representations in 2D, all the NLDR techniques preserve the global structure, but some NLDR techniques fail to capture the local structure.
+According to the visualizations, we can see that tSNE preserve both local and global structure than other NLDR techniques that is also evident by the AIC plot shown in Figure 4.-->
+
+<!--We *strongly* recommend viewing the linked videos for each study while reading.
 Links to the videos are available in the figures for each example.
-The videos show the visual appearance of the **langevitour** interface with low-dimensional view and how we can interact with the tour via the controls.
+The videos show the visual appearance of the **langevitour** interface with low-dimensional view and how we can interact with the tour via the controls.-->
+
+<!--In all NLDR techniques, except for PHATE, there are five well separated clusters.
+In tSNE, the five clusters are close to each other.
+However, in PHATE, two clusters are really close to each other and other three clusters are far away from them.
+In PaCMAP, one cluster is at the center and other four positioned in four directions.
+In TriMAP also, there are two clusters close to each other and other three are well-separated from each other.
+In UMAP, four clusters are positioned parallel to each and remaining one far away from them.-->
+
+<!--The model with tSNE shows five well-separated clusters. Also, within cluster variation is captured by giving proof that with clustering tSNE preserve local and global structures of high-D data.
+
+The model with UMAP also shows five well-separated clusters. But now, the clusters are more flat surfaces.
+
+The model with PHATE also shows five well-separated clusters. The five clusters are more like triangles and couldn't able to capture local structure at all.
+
+In the model with TriMAP shows five well-separated clusters. Also, within cluster variation is captured but much like flat surfaces.
+
+In the model with PaCMAP shows five well-separated clusters. Also, within cluster variation is captured but much like flat surfaces.-->
 
@@
 plot_list1_gau <- plot_tSNE_2D(tSNE_data_gau) + #ggtitle("(a)") + 
+  geom_point(size = 0.00001, colour = "#e41a1c") +
   theme_linedraw() +
@@
 plot_list2_gau <- plot_UMAP_2D(UMAP_data_gau) + #ggtitle("(b)") +
+  geom_point(size = 0.00001, colour = "#377eb8") +
   theme_linedraw() +
@@
 plot_list3_gau <- plot_PHATE_2D(PHATE_data_gau) + #ggtitle("(c)") +
+  geom_point(size = 0.00001, colour = "#4daf4a") +
   theme_linedraw() +
@@
 plot_list4_gau <- plot_TriMAP_2D(TriMAP_data_gau) + #ggtitle("(d)") +
+  geom_point(size = 0.00001, colour = "#984ea3") +
   theme_linedraw() +
@@
 plot_list5_gau <- plot_PaCMAP_2D(PaCMAP_data_gau) + #ggtitle("(e)") +
+  geom_point(size = 0.00001, colour = "#ff7f00") +
   theme_linedraw() +
@@
         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+  annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
 
@@
         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+  annotate(geom = 'text', label = 'c', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
 
@@
         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+  annotate(geom = 'text', label = 'd', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
 
@@
         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+  annotate(geom = 'text', label = 'e', x = -Inf, y = Inf, hjust = -0.5, vjust = 1, size = 3)
 

### 2024-01-23 paper.tex
Add what I learn from the models for simple example with Gaussian clusters

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, breakable, interior hidden, enhanced, boxrule=0pt, frame hidden, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, borderline west={3pt}{0pt}{shadecolor}, interior hidden, boxrule=0pt, breakable, sharp corners, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 
-We \emph{strongly} recommend viewing the linked videos for each study
-while reading. Links to the videos are available in the figures for each
-example. The videos show the visual appearance of the
-\textbf{langevitour} interface with low-dimensional view and how we can
-interact with the tour via the controls.
+In the 2D representations created by all NLDR techniques, as shown in
+Figure~\ref{fig-nldervis5Gau}, except for PHATE, there are five distinct
+clusters. In tSNE, these five clusters are closely located to each other
+(see Figure~\ref{fig-nldervis5Gau} (a)). However, in PHATE, two clusters
+are closely positioned, while the other three are more distant (see
+Figure~\ref{fig-nldervis5Gau} (c)). In PaCMAP, one cluster is at the
+center, and the remaining four are positioned in four different
+directions (see Figure~\ref{fig-nldervis5Gau} (e)). In TriMAP, two
+clusters are close, although not as close as in PHATE, and the other
+three are well-separated (see Figure~\ref{fig-nldervis5Gau} (d)). In
+UMAP, all clusters are arranged in a parallel manner, with three in one
+line and the other two in a separate line (see
+Figure~\ref{fig-nldervis5Gau} (b)).
+
+Visualizing the models alongside the original high-D data provides
+insights into how different techniques capture the underlying clustering
+structure. The tSNE model exhibits five well-separated clusters,
+effectively preserving both local and global structures (see video link
+of Figure~\ref{fig-modelfiveGau} (a)). UMAP, on the other hand, also
+presents five distinct clusters but with a more flattened surface
+appearance (see video link of Figure~\ref{fig-modelfiveGau} (b)). The
+PHATE model shows five separated clusters resembling triangles or
+partial triangles but struggles to capture local structures (see video
+link of Figure~\ref{fig-modelfiveGau} (c)). TriMAP and PaCMAP both
+display five well-separated clusters with flat surfaces, each capturing
+within-cluster variation to varying extents. Despite the various 2D
+representations, all NLDR techniques preserve the global structure (see
+video link of Figure~\ref{fig-modelfiveGau} (d), (e) respectively).
+However, tSNE, effectively capture both local and global structures, as
+indicated by lower AIC values in Figure~\ref{fig-diagnosticpltGau} (a).
 

### 2024-01-24 paper.qmd
Add three parameter choices can discuss with star data set

@@
 #| echo: false
-#| fig-cap: Goodness of fit statistics from different n_neighbors parameter of UMAP applied to training S-curve dataset. What is the best parameter choice in UMAP to represent the original data in 2D?
+#| fig-cap: "Goodness of fit statistics from different n_neighbors parameter of UMAP applied to training S-curve dataset.
+What is the best parameter choice in UMAP to represent the original data in 2D?"
 #| label: fig-diagnosticpltDiffParam
@@
 #| label: fig-modelfiveGau
-#| fig-cap: Is there a best model to represent the original data in 2D space or are they all providing equivalent information?, (a) Model in the 2D space with tSNE (<https://youtu.be/x5VPB-wOLm4>), (b) Model in the 2D space with UMAP (<https://youtu.be/Cm1dW6iLyHQ>), (c) Model in the 2D space with PHATE (<https://youtu.be/HB0Y3ilz6sI>), (d) Model in the 2D space with TriMAP (<https://youtu.be/iSVkXIrOrSA>), and (e) Model in the 2D space with PaCMAP (<https://youtu.be/O2QBctsc4uM>).
+#| fig-cap: "Is there a best model to represent the original data in 2D space or are they all providing equivalent information?, (a) Model in the 2D space with tSNE (<https://youtu.be/x5VPB-wOLm4>), (b) Model in the 2D space with UMAP (<https://youtu.be/Cm1dW6iLyHQ>), (c) Model in the 2D space with PHATE (<https://youtu.be/HB0Y3ilz6sI>), (d) Model in the 2D space with TriMAP (<https://youtu.be/iSVkXIrOrSA>), and (e) Model in the 2D space with PaCMAP (<https://youtu.be/O2QBctsc4uM>)."
 #| out-width: 100%
@@
 #| echo: false
-#| fig-cap: Goodness of fit statistics from different NLDR techniques applied to training five spherical Gaussian cluster dataset. What is the best NLDR technique to represent the original data in 2D?
+#| fig-cap: "Goodness of fit statistics from different NLDR techniques applied to training five spherical Gaussian cluster dataset.
+What is the best NLDR technique to represent the original data in 2D?"
 #| label: fig-diagnosticpltGau
@@
 #| label: fig-nldervisPBMCUMAP
- #| out-width: 100%
+##| out-width: 100%
+#| fig-pos: h
 
@@
 #| label: fig-nldervisPBMCUMAP2
- #| out-width: 100%
+##| out-width: 100%
+#| fig-pos: h
 
@@
 #| label: fig-nldervisPBMCUMAP3
- #| out-width: 100%
+##| out-width: 100%
+#| fig-pos: h
 
@@
 
-- In 2D, there are two well-separated clusters
+- In 2D, there are two well-separated clusters, but the clusters are really close
 
@@
 
-- In high-D view, the clusters having some non-linearity patter, which is also not captured in tSNE. Therefore, this is not a good choice.
+- In high-D view, the clusters having some non-linearity patter, which is also not captured in tSNE.
+Therefore, this is not a good choice.
+Also, the two clusters are really near to each other and can't see any separation.
+Also continuity of the structure.
+Long edges are presence also.
 
@@
 
+<!-- Author suggestion-->
 ```{r}
@@
     geom_point(alpha=0.5) +
-    coord_equal() +
+    coord_cartesian(xlim = c(-25, 25), ylim = c(-12, 12)) +
     theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
@@
   geom_point(colour = "#33a02c", size = 0.5) +
+  #coord_cartesian(xlim = c(-25, 25), ylim = c(-12, 12)) +
   theme_linedraw() +
@@
 #| fig-cap: "(a) 2D layout from tSNE (perplexity = 40) applied for the HARPS GTO dataset.
-Is this a best representation of the original data?, (b) Model in the 2D space with tSNE (<https://youtu.be/txTD7GTxt6Y>)"
+Is this a best representation of the original data?, (b) Model in the 2D space with tSNE (<https://youtu.be/jxo_rPyGv9k>)"
 #| label: fig-nldervisharpstSNE
+#| fig-pos: h
+
+plot_list2_harps + trimesh_removed_tsne_harps +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='bottom', plot.tag = element_text(size = 8))
+
+```
+
+<!-- param set 1-->
+
+- 5 clusters shows in 2D
+
+- 3 clusters which are really close shows in high_d data with the model
+
+```{r}
+#| warning: false
+#| echo: false
+
+tSNE_data_harps <- read_rds("data/harps/harps_tsne_23.rds")
+
+plot_list2_harps <- tSNE_data_harps |>
+    ggplot(aes(x = tSNE1,
+               y = tSNE2, color = clusters))+
+    geom_point(alpha=0.5) +
+    coord_cartesian(xlim = c(-22, 18), ylim = c(-27, 32)) +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#b15928", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#cab2d6", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#bc80bd", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
+  guides(color = guide_legend(title = "cluster"),
+         color = guide_legend(nrow = 3)) +
+ annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## Import data
+# training_data_1 <- read_rds("data/harps/harps_data.rds")
+# training_data_1 <- training_data_1 |>
+#   mutate(ID = 1:NROW(training_data_1))
+# 
+# clusters <- training_data_1$cluster
+# select_names <-  names(training_data_1)[-c(18, length(training_data_1))]
+# 
+# training_data_1 <- training_data_1 |> select(-cluster)
+# names(training_data_1) <- append(paste0("x", 1:17), "ID")
+# 
+# tSNE_data_harps <- tSNE_data_harps |>
+#   select(tSNE1, tSNE2) |>
+#   mutate(ID = 1:NROW(tSNE_data_harps))
+
+### tSNE
+
+num_bins_tsne_harps <- calculate_effective_x_bins(.data = tSNE_data_harps, x = tSNE1,
+                                                cell_area = 1) #43
+num_bins_tsne_harps <- 30
+
+shape_val_tsne_harps <- calculate_effective_shape_value(.data = tSNE_data_harps,
+                                                      x = tSNE1, y = tSNE2)
+## To extract bin centroids
+hexbin_data_object_tsne_harps <- extract_hexbin_centroids(nldr_df = tSNE_data_harps, num_bins = num_bins_tsne_harps, shape_val = shape_val_tsne_harps, x = tSNE1, y = tSNE2)
+
+df_bin_centroids_tsne_harps <- hexbin_data_object_tsne_harps$hexdf_data
+
+## Identify bins with low-density
+# identify_rm_bins <- find_low_density_hexagons(df_bin_centroids_tsne_harps, num_bins_tsne_harps, benchmark_rm_hex = 0.01)
+# 
+# df_bin_centroids_tsne_harps <- df_bin_centroids_tsne_harps |>
+#   filter(!(hexID %in% identify_rm_bins))
+
+tSNE_data_with_hb_id_harps <- tSNE_data_harps |>
+  dplyr::mutate(hb_id = hexbin_data_object_tsne_harps$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_tsne_harps <- dplyr::bind_cols(training_data_1 |> dplyr::select(-ID), tSNE_data_with_hb_id_harps)
+
+## Averaged on high-D
+df_bin_tsne_harps <- avg_highD_data(.data = df_all_tsne_harps)
+
+## Triangulate bin centroids
+tr1_object_tsne_harps <- triangulate_bin_centroids(df_bin_centroids_tsne_harps, x, y)
+tr_from_to_df_tsne_harps <- generate_edge_info(triangular_object = tr1_object_tsne_harps)
+
+## Compute 2D distances
+distance_tsne_harps <- cal_2D_dist(.data = tr_from_to_df_tsne_harps)
+
+## To find the benchmark value
+benchmark_tsne_harps <- find_benchmark_value(.data = distance_tsne_harps, distance_col = distance)
+
+trimesh_removed_tsne_harps <- remove_long_edges(.data = distance_tsne_harps, benchmark_value = benchmark_tsne_harps,
+                                              triangular_object = tr1_object_tsne_harps, distance_col = distance)
+
+trimesh_removed_tsne_harps <- trimesh_removed_tsne_harps +
+  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  # theme(axis.text = element_text(size = 5),
+  #       axis.title = element_text(size = 7)) +
+  geom_point(colour = "#33a02c", size = 0.5) +
+  #coord_cartesian(xlim = c(-22, 18), ylim = c(-27, 32)) +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "(a) 2D layout from tSNE (perplexity = 23) applied for the HARPS GTO dataset.
+Is this a best representation of the original data?, (b) Model in the 2D space with tSNE (<https://youtu.be/_is9qStggFE>)"
+#| label: fig-nldervisharpstSNE2
+#| fig-pos: h
+
+plot_list2_harps + trimesh_removed_tsne_harps +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='bottom', plot.tag = element_text(size = 8))
+
+```
+
+<!-- param set 2-->
+
+```{r}
+#| warning: false
+#| echo: false
+
+tSNE_data_harps <- read_rds("data/harps/harps_tsne_91.rds")
+
+plot_list2_harps <- tSNE_data_harps |>
+    ggplot(aes(x = tSNE1,
+               y = tSNE2, color = clusters))+
+    geom_point(alpha=0.5) +
+    coord_cartesian(xlim = c(-8, 7), ylim = c(-10, 16)) +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#b15928", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#cab2d6", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#bc80bd", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
+  guides(color = guide_legend(title = "cluster"),
+         color = guide_legend(nrow = 3)) +
+ annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## Import data
+# training_data_1 <- read_rds("data/harps/harps_data.rds")
+# training_data_1 <- training_data_1 |>
+#   mutate(ID = 1:NROW(training_data_1))
+# 
+# clusters <- training_data_1$cluster
+# select_names <-  names(training_data_1)[-c(18, length(training_data_1))]
+# 
+# training_data_1 <- training_data_1 |> select(-cluster)
+# names(training_data_1) <- append(paste0("x", 1:17), "ID")
+# 
+# tSNE_data_harps <- tSNE_data_harps |>
+#   select(tSNE1, tSNE2) |>
+#   mutate(ID = 1:NROW(tSNE_data_harps))
+
+### tSNE
+
+num_bins_tsne_harps <- calculate_effective_x_bins(.data = tSNE_data_harps, x = tSNE1,
+                                                cell_area = 1) #43
+#num_bins_tsne_harps <- 30
+
+shape_val_tsne_harps <- calculate_effective_shape_value(.data = tSNE_data_harps,
+                                                      x = tSNE1, y = tSNE2)
+## To extract bin centroids
+hexbin_data_object_tsne_harps <- extract_hexbin_centroids(nldr_df = tSNE_data_harps, num_bins = num_bins_tsne_harps, shape_val = shape_val_tsne_harps, x = tSNE1, y = tSNE2)
+
+df_bin_centroids_tsne_harps <- hexbin_data_object_tsne_harps$hexdf_data
+
+## Identify bins with low-density
+# identify_rm_bins <- find_low_density_hexagons(df_bin_centroids_tsne_harps, num_bins_tsne_harps, benchmark_rm_hex = 0.01)
+# 
+# df_bin_centroids_tsne_harps <- df_bin_centroids_tsne_harps |>
+#   filter(!(hexID %in% identify_rm_bins))
+
+tSNE_data_with_hb_id_harps <- tSNE_data_harps |>
+  dplyr::mutate(hb_id = hexbin_data_object_tsne_harps$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_tsne_harps <- dplyr::bind_cols(training_data_1 |> dplyr::select(-ID), tSNE_data_with_hb_id_harps)
+
+## Averaged on high-D
+df_bin_tsne_harps <- avg_highD_data(.data = df_all_tsne_harps)
+
+## Triangulate bin centroids
+tr1_object_tsne_harps <- triangulate_bin_centroids(df_bin_centroids_tsne_harps, x, y)
+tr_from_to_df_tsne_harps <- generate_edge_info(triangular_object = tr1_object_tsne_harps)
+
+## Compute 2D distances
+distance_tsne_harps <- cal_2D_dist(.data = tr_from_to_df_tsne_harps)
+
+## To find the benchmark value
+benchmark_tsne_harps <- find_benchmark_value(.data = distance_tsne_harps, distance_col = distance)
+
+trimesh_removed_tsne_harps <- remove_long_edges(.data = distance_tsne_harps, benchmark_value = benchmark_tsne_harps,
+                                              triangular_object = tr1_object_tsne_harps, distance_col = distance)
+
+trimesh_removed_tsne_harps <- trimesh_removed_tsne_harps +
+  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  # theme(axis.text = element_text(size = 5),
+  #       axis.title = element_text(size = 7)) +
+  geom_point(colour = "#33a02c", size = 0.5) +
+  #coord_cartesian(xlim = c(-8, 7), ylim = c(-10, 16)) +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "(a) 2D layout from tSNE (perplexity = 91) applied for the HARPS GTO dataset.
+Is this a best representation of the original data?, (b) Model in the 2D space with tSNE (<https://youtu.be/aGvHFhWw_Xs>)"
+#| label: fig-nldervisharpstSNE3
+#| fig-pos: h
 

### 2024-01-24 paper.tex
Add three parameter choices can discuss with star data set

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, borderline west={3pt}{0pt}{shadecolor}, interior hidden, boxrule=0pt, breakable, sharp corners, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, sharp corners, frame hidden, interior hidden, breakable, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
 
@@
 \item
-  In 2D, there are two well-separated clusters
+  In 2D, there are two well-separated clusters, but the clusters are
+  really close
 \item
@@
   is also not captured in tSNE. Therefore, this is not a good choice.
+  Also, the two clusters are really near to each other and can't see any
+  separation. Also continuity of the structure. Long edges are presence
+  also.
 \item
@@
 
-\begin{figure}
+\begin{figure}[h]
 
@@
 representation of the original data?, (b) Model in the 2D space with
-tSNE (\url{https://youtu.be/txTD7GTxt6Y})}
+tSNE (\url{https://youtu.be/jxo_rPyGv9k})}
+
+\end{figure}
+
+\begin{itemize}
+\item
+  5 clusters shows in 2D
+\item
+  3 clusters which are really close shows in high\_d data with the model
+\end{itemize}
+
+\begin{figure}[h]
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-nldervisharpstSNE2-1.pdf}
+
+}
+
+\caption{\label{fig-nldervisharpstSNE2}(a) 2D layout from tSNE
+(perplexity = 23) applied for the HARPS GTO dataset. Is this a best
+representation of the original data?, (b) Model in the 2D space with
+tSNE (\url{https://youtu.be/_is9qStggFE})}
+
+\end{figure}
+
+\begin{figure}[h]
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-nldervisharpstSNE3-1.pdf}
+
+}
+
+\caption{\label{fig-nldervisharpstSNE3}(a) 2D layout from tSNE
+(perplexity = 91) applied for the HARPS GTO dataset. Is this a best
+representation of the original data?, (b) Model in the 2D space with
+tSNE (\url{https://youtu.be/aGvHFhWw_Xs})}
 

### 2024-01-25 paper.qmd
Add 3 cluster examples with suitable choice to discuss for PBMC3k

@@
 #tSNE_pbmc <- read_rds("data/pbmc/pbmc_tsne_30.rds")
-UMAP_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_umap_5_min_dist_0.1.rds")
+UMAP_pbmc <- read_rds("data/pbmc/pbmc_umap_100_min_dist_0.3.rds")
 # PHATE_pbmc <- read_rds("data/pbmc/pbmc_phate.rds")
@@
 benchmark_pbmc <- find_benchmark_value(.data = distance_pbmc, distance_col = distance)
-benchmark_pbmc <- 1.055271
+benchmark_pbmc <- 1.049766
 
@@
 #| echo: false
-#| fig-cap: "(a) 2D layout from UMAP (n_neighbors = 5, min_dist = 0.1) applied for the PBMC3k dataset.
-Is this a best representation of the original data?, (b)  Model in the 2D space with UMAP (<https://youtu.be/SO3thJ8PMi8>)"
+#| fig-cap: "(a) 2D layout from UMAP (n_neighbors = 100, min_dist = 0.3) applied for the PBMC3k dataset.
+Is this a best representation of the original data?, (b)  Model in the 2D space with UMAP (<https://youtu.be/1jNVVlzI4tM>)"
 #| label: fig-nldervisPBMCUMAP2
@@
 #tSNE_pbmc <- read_rds("data/pbmc/pbmc_tsne_30.rds")
-UMAP_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_umap_30_min_dist_0.01.rds")
+UMAP_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_umap_15_min_dist_0.3.rds")
 # PHATE_pbmc <- read_rds("data/pbmc/pbmc_phate.rds")
@@
 benchmark_pbmc <- find_benchmark_value(.data = distance_pbmc, distance_col = distance)
-benchmark_pbmc <- 1.057549
+benchmark_pbmc <- 1.035666
 
@@
 #| echo: false
-#| fig-cap: "(a) 2D layout from UMAP (n_neighbors = 30, min_dist = 0.01) applied for the PBMC3k dataset.
-Is this a best representation of the original data?, (b)  Model in the 2D space with UMAP (<https://youtu.be/hP8ie_oMXw0>)"
+#| fig-cap: "(a) 2D layout from UMAP (n_neighbors = 15, min_dist = 0.3) applied for the PBMC3k dataset.
+Is this a best representation of the original data?, (b)  Model in the 2D space with UMAP (<https://youtu.be/2GM_SuM866M>)"
 #| label: fig-nldervisPBMCUMAP3
@@
 
+@Anders2018
+
 - In 2D, there are two well-separated clusters, but the clusters are really close
@@
     geom_point(alpha=0.5) +
-    coord_cartesian(xlim = c(-25, 25), ylim = c(-12, 12)) +
+    #coord_cartesian(xlim = c(-25, 25), ylim = c(-12, 12)) +
     theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
@@
     geom_point(alpha=0.5) +
-    coord_cartesian(xlim = c(-22, 18), ylim = c(-27, 32)) +
+    coord_cartesian(xlim = c(-25, 20), ylim = c(-35, 25)) +
     theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
@@
   geom_point(colour = "#33a02c", size = 0.5) +
-  #coord_cartesian(xlim = c(-22, 18), ylim = c(-27, 32)) +
+  coord_cartesian(xlim = c(-25, 20), ylim = c(-35, 25)) +
   theme_linedraw() +
@@
 #| fig-cap: "(a) 2D layout from tSNE (perplexity = 23) applied for the HARPS GTO dataset.
-Is this a best representation of the original data?, (b) Model in the 2D space with tSNE (<https://youtu.be/_is9qStggFE>)"
+Is this a best representation of the original data?, (b) Model in the 2D space with tSNE (<https://youtu.be/VO_4SJEoAUA>)"
 #| label: fig-nldervisharpstSNE2
@@
   geom_point(colour = "#33a02c", size = 0.5) +
-  #coord_cartesian(xlim = c(-8, 7), ylim = c(-10, 16)) +
+  coord_cartesian(xlim = c(-8, 7), ylim = c(-10, 16)) +
   theme_linedraw() +
@@
 #| fig-cap: "(a) 2D layout from tSNE (perplexity = 91) applied for the HARPS GTO dataset.
-Is this a best representation of the original data?, (b) Model in the 2D space with tSNE (<https://youtu.be/aGvHFhWw_Xs>)"
+Is this a best representation of the original data?, (b) Model in the 2D space with tSNE (<https://youtu.be/vLvRfcogft8>)"
 #| label: fig-nldervisharpstSNE3

### 2024-01-25 paper.tex
Add 3 cluster examples with suitable choice to discuss for PBMC3k

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, sharp corners, frame hidden, interior hidden, breakable, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, borderline west={3pt}{0pt}{shadecolor}, enhanced, sharp corners, frame hidden, boxrule=0pt, interior hidden]}{\end{tcolorbox}}\fi
 
@@
 \caption{\label{fig-nldervisPBMCUMAP2}(a) 2D layout from UMAP
-(n\_neighbors = 5, min\_dist = 0.1) applied for the PBMC3k dataset. Is
+(n\_neighbors = 100, min\_dist = 0.3) applied for the PBMC3k dataset. Is
 this a best representation of the original data?, (b) Model in the 2D
-space with UMAP (\url{https://youtu.be/SO3thJ8PMi8})}
+space with UMAP (\url{https://youtu.be/1jNVVlzI4tM})}
 
@@
 \caption{\label{fig-nldervisPBMCUMAP3}(a) 2D layout from UMAP
-(n\_neighbors = 30, min\_dist = 0.01) applied for the PBMC3k dataset. Is
+(n\_neighbors = 15, min\_dist = 0.3) applied for the PBMC3k dataset. Is
 this a best representation of the original data?, (b) Model in the 2D
-space with UMAP (\url{https://youtu.be/hP8ie_oMXw0})}
+space with UMAP (\url{https://youtu.be/2GM_SuM866M})}
 
@@
 \subsection{HARPS GTO sample dataset}\label{harps-gto-sample-dataset}}
+
+\citet{Anders2018}
 
@@
 representation of the original data?, (b) Model in the 2D space with
-tSNE (\url{https://youtu.be/_is9qStggFE})}
+tSNE (\url{https://youtu.be/VO_4SJEoAUA})}
 
@@
 representation of the original data?, (b) Model in the 2D space with
-tSNE (\url{https://youtu.be/aGvHFhWw_Xs})}
+tSNE (\url{https://youtu.be/vLvRfcogft8})}
 

### 2024-01-25 nldr_code.R

@@ -34,7 +34,7 @@ plot_tSNE_2D <- function(tSNE_df){
   tSNE_df_plot <- tSNE_df %>%
     ggplot(aes(x = tSNE1,
                y = tSNE2))+
-    geom_point(alpha=0.5, color="#e41a1c", size = 0.5) +
+    geom_point(alpha=0.5, colour="#e41a1c", size = 0.5) +
     coord_equal() +
     theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold"),
           axis.text = element_text(size = 5),
@@ -69,7 +69,7 @@ plot_UMAP_2D <- function(UMAP_df){
   UMAP_df_plot <- UMAP_df %>%
     ggplot(aes(x = UMAP1,
                y = UMAP2))+
-    geom_point(alpha=0.5, color="#377eb8", size = 0.5) +
+    geom_point(alpha=0.5, colour="#377eb8", size = 0.5) +
     coord_equal() +
     theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold"),
           axis.text = element_text(size = 5),
@@ -98,7 +98,7 @@ plot_PHATE_2D <- function(PHATE_df){
   PHATE_df_plot <- PHATE_df %>%
     ggplot(aes(x = PHATE1,
                y = PHATE2))+
-    geom_point(alpha=0.5, color="#4daf4a", size = 0.5) +
+    geom_point(alpha=0.5, colour="#4daf4a", size = 0.5) +
     coord_equal() +
     theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold"),
           axis.text = element_text(size = 5),
@@ -115,7 +115,7 @@ plot_TriMAP_2D <- function(TRIMAP_df){
   TriMAP_df_plot <- TRIMAP_df %>%
     ggplot(aes(x = TriMAP1,
                y = TriMAP2))+
-    geom_point(alpha=0.5, color="#984ea3", size = 0.5) +
+    geom_point(alpha=0.5, colour="#984ea3", size = 0.5) +
     coord_equal() +
     theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold"),
           axis.text = element_text(size = 5),
@@ -132,7 +132,7 @@ plot_PaCMAP_2D <- function(PaCMAP_df){
   PaCMAP_df_plot <- PaCMAP_df %>%
     ggplot(aes(x = PaCMAP1,
                y = PaCMAP2))+
[…]

### 2024-01-25 paper.qmd
Changed the MSE equation

@@
 $$
-\text{MSE} = \sum_{j = 1}^{n} \frac{\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2}{b}
+\text{MSE} = \sum_{j = 1}^{n} \frac{\sqrt{\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2}}{n}
 $$ {#eq-equation9}
 
-Here, $b$ signifies the total number of bins, $p$ denotes the number of dimensions in the high-dimensional data, and $n$ represents the number of observations.
+Here, $b'$ signifies the total number of bins without empty bins, $p$ denotes the number of dimensions in the high-dimensional data, and $n$ represents the number of observations.
 
@@
 $$
-\text{AIC} = 2bp + np * log(\text{MSE})
+\text{AIC} = 2b'p + np * log(\text{MSE})
 $$ {#eq-equation10}
@@
 
+@Chen2023
+
 - Intro to the data set (This data set contains 13,714 features across 2,700 samples within a single assay. The active assay is RNA, with 13,714 features representing different gene expressions.)

### 2024-01-25 paper.tex
Changed the MSE equation

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, borderline west={3pt}{0pt}{shadecolor}, enhanced, sharp corners, frame hidden, boxrule=0pt, interior hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden, breakable, enhanced, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 \begin{equation}\protect\hypertarget{eq-equation9}{}{
-\text{MSE} = \sum_{j = 1}^{n} \frac{\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2}{b}
+\text{MSE} = \sum_{j = 1}^{n} \frac{\sqrt{\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2}}{n}
 }\label{eq-equation9}\end{equation}
 
-Here, \(b\) signifies the total number of bins, \(p\) denotes the number
-of dimensions in the high-dimensional data, and \(n\) represents the
-number of observations.
+Here, \(b'\) signifies the total number of bins without empty bins,
+\(p\) denotes the number of dimensions in the high-dimensional data, and
+\(n\) represents the number of observations.
 
@@
 \begin{equation}\protect\hypertarget{eq-equation10}{}{
-\text{AIC} = 2bp + np * log(\text{MSE})
+\text{AIC} = 2b'p + np * log(\text{MSE})
 }\label{eq-equation10}\end{equation}
@@
 
+\citet{Chen2023}
+
 \begin{itemize}

### 2024-01-29 paper.qmd
Add error cal table for PBMC3k data set with different parameter combinations

@@
 ## To draw with AIC
-aic_plot <- ggplot(MSE_df |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
+aic_plot <- ggplot(MSE_df |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
                                                                                  y = total_error,
@@
 
-mse_plot <- ggplot(MSE_df, aes(x = number_of_bins,
+mse_plot <- ggplot(MSE_df |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
                                        y = total_mse,
@@
 ## To draw with AIC
-aic_plot_lwd <- ggplot(MSE_df_lwd |> dplyr::filter(data_type == "training"), aes(x = benchmark_rm_hex,
+aic_plot_lwd <- ggplot(MSE_df_lwd |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "UMAP"), aes(x = benchmark_rm_hex,
                                                                                  y = total_error,
@@
 
-mse_plot_lwd <- ggplot(MSE_df_lwd, aes(x = benchmark_rm_hex,
+mse_plot_lwd <- ggplot(MSE_df_lwd |> dplyr::filter(method == "UMAP"), aes(x = benchmark_rm_hex,
                                        y = total_mse,
@@
 
-#aic_plot_lwd + 
-  mse_plot_lwd +
-  #plot_annotation(tag_levels = 'a') +
+aic_plot_lwd + mse_plot_lwd +
+  plot_annotation(tag_levels = 'a') +
   plot_layout(guides='collect', ncol = 2) &
@@
 ## To draw with AIC
-aic_plot_lg <- ggplot(error_df_lg, aes(x = benchmark_rm_lg,
+aic_plot_lg <- ggplot(error_df_lg |> dplyr::filter(method == "UMAP"), aes(x = benchmark_rm_lg,
                                                                                  y = total_error,
@@
 
-mse_plot_lg <- ggplot(error_df_lg, aes(x = benchmark_rm_lg,
+mse_plot_lg <- ggplot(error_df_lg |> dplyr::filter(method == "UMAP"), aes(x = benchmark_rm_lg,
                                        y = total_mse,
@@
 
-#aic_plot_lg + 
-  mse_plot_lg +
-  #plot_annotation(tag_levels = 'a') +
+aic_plot_lg + mse_plot_lg +
+  plot_annotation(tag_levels = 'a') +
   plot_layout(guides='collect', ncol = 2) &
@@
 
+- Numb bins along the x-axis = 23, shape parameter = 0.8772751
+
 - learned from the model: There are 3 well-separated clusters in 2D, but if look at the model in high-D space, the three clusters are much closer.
@@
 Larger values of min_dist will make UMAP pack points together more loosely, focusing instead on the preservation of the broad topological structure.
+
+```{r}
+#| warning: false
+#| echo: false
+
+error_df <- read_csv("data/pbmc/pbmc_3k_festem/error_pbmc_umap.csv")
+
+error_df |>
+  knitr::kable() 
+```
 
@@
 #tSNE_pbmc <- read_rds("data/pbmc/pbmc_tsne_30.rds")
-UMAP_pbmc <- read_rds("data/pbmc/pbmc_umap_100_min_dist_0.3.rds")
+UMAP_pbmc <- read_rds("data/pbmc/pbmc_umap_15_min_dist_0.99_metric_euclidean.rds")
 # PHATE_pbmc <- read_rds("data/pbmc/pbmc_phate.rds")
@@
 
+num_bins_pbmc <- 23
+
 shape_value_pbmc <- calculate_effective_shape_value(.data = UMAP_pbmc,
                                                       x = UMAP1, y = UMAP2) ## 0.8772751
+shape_value_pbmc <- 0.8772751
 
@@
 benchmark_pbmc <- find_benchmark_value(.data = distance_pbmc, distance_col = distance)
-benchmark_pbmc <- 1.049766
+#benchmark_pbmc <- 1.049766
 
@@
 #| echo: false
-#| fig-cap: "(a) 2D layout from UMAP (n_neighbors = 100, min_dist = 0.3) applied for the PBMC3k dataset.
-Is this a best representation of the original data?, (b)  Model in the 2D space with UMAP (<https://youtu.be/1jNVVlzI4tM>)"
+#| fig-cap: "(a) 2D layout from UMAP (n_neighbors = 15, min_dist = 0.99, metric = euclidean) applied for the PBMC3k dataset.
+Is this a best representation of the original data?, (b)  Model in the 2D space with UMAP (<>)"
 #| label: fig-nldervisPBMCUMAP2
@@
 #tSNE_pbmc <- read_rds("data/pbmc/pbmc_tsne_30.rds")
-UMAP_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_umap_15_min_dist_0.3.rds")
+UMAP_pbmc <- read_rds("data/pbmc/pbmc_umap_5_min_dist_0.99_metric_cosine.rds")
 # PHATE_pbmc <- read_rds("data/pbmc/pbmc_phate.rds")
@@
                                               cell_area = 1) ##23
+num_bins_pbmc <- 23
 
@@
                                                       x = UMAP1, y = UMAP2) ## 0.8772751
+shape_value_pbmc <- 0.8772751
 
@@
 benchmark_pbmc <- find_benchmark_value(.data = distance_pbmc, distance_col = distance)
-benchmark_pbmc <- 1.035666
+#benchmark_pbmc <- 1.035666
 
@@
 #| echo: false
-#| fig-cap: "(a) 2D layout from UMAP (n_neighbors = 15, min_dist = 0.3) applied for the PBMC3k dataset.
-Is this a best representation of the original data?, (b)  Model in the 2D space with UMAP (<https://youtu.be/2GM_SuM866M>)"
+#| fig-cap: "(a) 2D layout from UMAP (n_neighbors = 5, min_dist = 0.99, metric = cosine) applied for the PBMC3k dataset.
+Is this a best representation of the original data?, (b)  Model in the 2D space with UMAP (<>)"
 #| label: fig-nldervisPBMCUMAP3

### 2024-01-29 paper.tex
Add error cal table for PBMC3k data set with different parameter combinations

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden, breakable, enhanced, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, enhanced, frame hidden, boxrule=0pt]}{\end{tcolorbox}}\fi
 
@@
 \item
+  Numb bins along the x-axis = 23, shape parameter = 0.8772751
+\item
   learned from the model: There are 3 well-separated clusters in 2D, but
@@
 
+\begin{tabular}{r|r|l|r}
+\hline
+30 & 0.3 & cosine & 38634.66719286174\\
+\hline
+5 & 0.99 & euclidean & 35938.99\\
+\hline
+5 & 0.99 & cosine & 35597.02\\
+\hline
+15 & 0.99 & euclidean & 35534.76\\
+\hline
+30 & 0.99 & euclidean & 35664.69\\
+\hline
+1022 & 0.99 & cosine & 37022.01\\
+\hline
+1022 & 0.30 & cosine & 39746.35\\
+\hline
+800 & 0.30 & euclidean & 38385.25\\
+\hline
+\end{tabular}
+
 \begin{figure}
@@
 \caption{\label{fig-nldervisPBMCUMAP2}(a) 2D layout from UMAP
-(n\_neighbors = 100, min\_dist = 0.3) applied for the PBMC3k dataset. Is
-this a best representation of the original data?, (b) Model in the 2D
-space with UMAP (\url{https://youtu.be/1jNVVlzI4tM})}
+(n\_neighbors = 15, min\_dist = 0.99, metric = euclidean) applied for
+the PBMC3k dataset. Is this a best representation of the original data?,
+(b) Model in the 2D space with UMAP (\textless\textgreater)}
 
@@
 \caption{\label{fig-nldervisPBMCUMAP3}(a) 2D layout from UMAP
-(n\_neighbors = 15, min\_dist = 0.3) applied for the PBMC3k dataset. Is
-this a best representation of the original data?, (b) Model in the 2D
-space with UMAP (\url{https://youtu.be/2GM_SuM866M})}
+(n\_neighbors = 5, min\_dist = 0.99, metric = cosine) applied for the
+PBMC3k dataset. Is this a best representation of the original data?, (b)
+Model in the 2D space with UMAP (\textless\textgreater)}
 

### 2024-01-30 paper.qmd
Add preprocessing steps with Festem

@@
 
-### Human peripheral blood mononclear cells (PBMCs) data
+### Single-cell RNA-seq data of human
+
+In the field of single-cell studies, a common analysis task involves clustering to identify groups of cells with similar expression profiles.
+Analysts often turn to NLDR techniques to verify and identify these clusters and explore developmental trajectories.
+In clustering workflows, the main objective is to verify the existence of clusters and subsequently identify them as specific cell types by examining the expression of "known" marker genes.
+Marker genes are specific genes or transcripts that are uniquely or highly expressed in particular cell types.
+Marker genes are identified through differential gene expression (DEG) analysis, comparing the expression levels of genes across different cell types.
+In this context, a "faithful" embedding should ideally preserve the topology of the data, ensuring that cells corresponding to the same cell type are situated close to the high-dimensional space.
+
+### Processing steps
+
+To begin our analysis, we installed the Peripheral Blood Mononuclear Cells (pbmc) data set obtained from 10x Genomics using the `SeuratData` R package [@Rahul2019], which facilitates the distribution of data sets in the form of Seurat objects [@Yuhan2021].
+This data set contains 13,714 features (genes) across 2,700 samples (cells) within a single assay.
+The active assay is RNA, with 13,714 features representing different gene expressions.
+Then, the cells that have unique feature counts over 2,500 or less than 200 and cells that have > 5\% mitochondrial counts are filtered.
+After removing unwanted cells from the dataset, the next step is to normalize the data.
+By default, we employ a global-scaling normalization method “LogNormalize” that normalizes the feature expression measurements for each cell by the total expression, multiplies this by a scale factor (10,000 by default), and log-transforms the result.
+
+Next, top 1000 genes are selected by Festem [@Chen2023] DEG method.
+Then, scale the data.
+Then, perform PCA on the scaled data.
+Louvain algorithm is used to cluster the single cells based on the genes selected by Festem.
+Using the Festem-selected genes and 15 PCs, identified 10 clusters in the PBMC3k data and annotated them based on the expression of canonical markers.
+The 10 clusters included immune cells such as naive CD4 T cells, memory CD4 T cells, CD8 T cells, CD14 monocytes, FCGR3A monocytes, Natural Killer (NK) cells, B cells and Dendritic Cells (DC).
+In addition to these common cell types, Festem identified a fine cell type, CD27− CD4+ memory T cells, which were often missed by other methods.
+
+- Human peripheral blood mononclear cells (PBMCs) data
+
+- Intro to the data set
+
+Used Louvain algorithm to cluster the single cells based on the genes selected by Festem method
+
+- Explain what "Evaluate Festem on DEG detection" means, in terms of this data
+
+-
+
+- How did they "identified 10 clusters"? What were the variables used, PCs or full data or 2D representation, spell this out.
+
+- Used top 1000 genes selected by Festem to run PCA, top 15 PCs (original paper mentioned)
+
+- The "10 clusters included immune cells such as naive CD4 T ..." does this mean that every cluster had all of these, or one cluster was mostly "CD4 T cells", another mostly "memory CD4 T", ...?
+
+- How is your use of this data different from the original application.
+
+- Original paper used this data set to compare different DEG methods, but we need to look how the performance of UMAP with the author's selected parameter choice.
+
+- Why 15 PCs? Is that from the original paper? YES
 
@@
 
-- Numb bins along the x-axis = 23, shape parameter = 0.8772751
+- Num bins along the x-axis = 23, shape parameter = 0.8772751
 
@@
 
-```{r}
-#| warning: false
-#| echo: false
-
-error_df <- read_csv("data/pbmc/pbmc_3k_festem/error_pbmc_umap.csv")
-
-error_df |>
-  knitr::kable() 
-```
+- The third parameter is metric.
+This parameter plays a crucial role in determining how distances are computed within the ambient space of the input data (The term "ambient space" refers to the original or initial space in which the data exists before any transformation or dimensionality reduction).
+
+```{r}
+#| warning: false
+#| echo: false
+
+# error_df <- read_csv("data/pbmc/pbmc_3k_festem/error_pbmc_umap.csv")
+# 
+# error_df |>
+#   knitr::kable() 
+```
+
+| n_neighbors | min_dist | metric     | error                 |
+|-------------|----------|------------|-----------------------|
+| **30**      | **0.3**  | **cosine** | **38634.66719286174** |
+| 5           | 0.99     | euclidean  | 35938.98594054407     |
+| 5           | 0.99     | cosine     | 35597.01981261912     |
+| 15          | 0.99     | euclidean  | 35534.75778935427     |
+| 30          | 0.99     | euclidean  | 35664.68763936665     |
+| 1022        | 0.99     | cosine     | 37022.0125249776      |
+| 1022        | 0.3      | cosine     | 39746.347504639125    |
+| 800         | 0.3      | euclidean  | 38385.25485420897     |
+| 100         | 0.99     | euclidean  | 36528.57424747017     |
+| 230         | 0.99     | cosine     | 37864.26802578186     |
+| 354         | 0.99     | euclidean  | 37530.86598361975     |
 
@@
 
+df_bin_train <- df_bin_pbmc
+names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])
+
+error_df <- df_all_pbmc |>
+  dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates
+
+# prediction_df_join <- prediction_df_join |>
+#   dplyr::left_join(data, by = c("ID" = "ID")) ## Map high-D data
+
+for (i in 1:(NCOL(df_bin_train) - 1)) {
+
+  error_df[ , paste0("abs_residual_", "PC_", i)] <- abs(error_df[ , paste0("PC_", i)] - error_df[ , paste0("avg_", "PC_", i)])
+
+}
+
+error_df <- error_df |>
+  dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "PC_")))))
+
+# library(ggbeeswarm)
+# error_df$group <- "1"
+# ggplot(error_df, aes(x = group, y = total)) +
+#   geom_quasirandom()+
+#   ylim(0, max(unlist(error_df$total))+ 0.5) + coord_flip()
+
+### The minimum error is 0 and the maximum is 42.17439
+### There is lot of points with error 0,
+
+error_df <- error_df |>
+  mutate(type = if_else(total <= 2, "error 2 or less",
+                        if_else(total <= 10, "error 2-10",
+                                if_else(total <= 15, "error 10-15",
+                                        if_else(total <= 20, "error 15-20",
+                                                if_else(total <= 25, "error 20-25",
+                                                        if_else(total <= 30, "error 25-30",
+                                                                if_else(total <= 35, "error 30-35", "error greter than 35"))))))))
+
+plot_list2_pbmc_error <- error_df |>
+  mutate(type = factor(type , levels = c("error 2 or less", "error 2-10", "error 10-15", "error 15-20", "error 20-25", "error 25-30", "error 30-35", "error greter than 35"))) |>
+    ggplot(aes(x = UMAP1,
+               y = UMAP2, color = type))+
+    geom_point(alpha=0.5, size = 0.1) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#fee0d2", "#fcbba1",
+                                         "#fc9272", "#fb6a4a", "#ef3b2c",
+                                         "#cb181d", "#a50f15", "#99000d")) #+
+  #geom_text(aes(x=UMAP1,y = UMAP2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) +
+  #annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
 ```
@@
 #| label: fig-nldervisPBMCUMAP
-##| out-width: 100%
+#| out-width: 100%
 #| fig-pos: h
 
-plot_list2_pbmc + trimesh_removed_pbmc_umap + 
-plot_layout(ncol=2)
+# plot_list2_pbmc + 
+trimesh_removed_pbmc_umap + plot_list2_pbmc_error +
+plot_layout(guides='collect', ncol=3) &
+  theme(legend.position='bottom')
 
@@
 
+df_bin_train <- df_bin_pbmc
+names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])
+
+error_df <- df_all_pbmc |>
+  dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates
+
+# prediction_df_join <- prediction_df_join |>
+#   dplyr::left_join(data, by = c("ID" = "ID")) ## Map high-D data
+
+for (i in 1:(NCOL(df_bin_train) - 1)) {
+
+  error_df[ , paste0("abs_residual_", "PC_", i)] <- abs(error_df[ , paste0("PC_", i)] - error_df[ , paste0("avg_", "PC_", i)])
+
+}
+
+error_df <- error_df |>
+  dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "PC_")))))
+
+# library(ggbeeswarm)
+# error_df$group <- "1"
+# ggplot(error_df, aes(x = group, y = total)) +
+#   geom_quasirandom()+
+#   ylim(0, max(unlist(error_df$total))+ 0.5) + coord_flip()
+
+### The minimum error is 0 and the maximum is 42.17439
+### There is lot of points with error 0,
+
+error_df <- error_df |>
+  mutate(type = if_else(total <= 2, "error 2 or less",
+                        if_else(total <= 10, "error 2-10",
+                                if_else(total <= 15, "error 10-15",
+                                        if_else(total <= 20, "error 15-20",
+                                                if_else(total <= 25, "error 20-25",
+                                                        if_else(total <= 30, "error 25-30",
+                                                                if_else(total <= 35, "error 30-35", "error greter than 35"))))))))
+
+plot_list2_pbmc_error <- error_df |>
+    mutate(type = factor(type , levels = c("error 2 or less", "error 2-10", "error 10-15", "error 15-20", "error 20-25", "error 25-30", "error 30-35", "error greter than 35"))) |>
+    ggplot(aes(x = UMAP1,
+               y = UMAP2, color = type))+
+    geom_point(alpha=0.5, size = 0.1) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#fee0d2", "#fcbba1",
+                                         "#fc9272", "#fb6a4a", "#ef3b2c",
+                                         "#cb181d", "#a50f15", "#99000d"))
+
 ```
@@
 #| label: fig-nldervisPBMCUMAP2
-##| out-width: 100%
+#| out-width: 100%
 #| fig-pos: h
 
-plot_list2_pbmc + trimesh_removed_pbmc_umap + 
-plot_layout(ncol=2)
+#plot_list2_pbmc + 
+trimesh_removed_pbmc_umap + plot_list2_pbmc_error +
+plot_layout(guides='collect', ncol=3) &
+  theme(legend.position='bottom')
 
@@
 #| label: fig-nldervisPBMCUMAP3
-##| out-width: 100%
+#| out-width: 100%
 #| fig-pos: h
 
-plot_list2_pbmc + trimesh_removed_pbmc_umap + 
-plot_layout(ncol=2)
+#plot_list2_pbmc + 
+trimesh_removed_pbmc_umap + plot_list2_pbmc_error +
+plot_layout(guides='collect', ncol=3) &
+  theme(legend.position='bottom')
 
@@
 
+- Num bins along the x-axis = 43, shape parameter = 0.4466309
+
 <!-- Author suggestion-->

### 2024-01-30 paper.tex
Add preprocessing steps with Festem

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, enhanced, frame hidden, boxrule=0pt]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable, boxrule=0pt, interior hidden, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 
-\hypertarget{human-peripheral-blood-mononclear-cells-pbmcs-data}{%
-\subsection{Human peripheral blood mononclear cells (PBMCs)
-data}\label{human-peripheral-blood-mononclear-cells-pbmcs-data}}
+\hypertarget{single-cell-rna-seq-data-of-human}{%
+\subsection{Single-cell RNA-seq data of
+human}\label{single-cell-rna-seq-data-of-human}}
+
+In the field of single-cell studies, a common analysis task involves
+clustering to identify groups of cells with similar expression profiles.
+Analysts often turn to NLDR techniques to verify and identify these
+clusters and explore developmental trajectories. In clustering
+workflows, the main objective is to verify the existence of clusters and
+subsequently identify them as specific cell types by examining the
+expression of ``known'' marker genes. Marker genes are specific genes or
+transcripts that are uniquely or highly expressed in particular cell
+types. Marker genes are identified through differential gene expression
+(DEG) analysis, comparing the expression levels of genes across
+different cell types. In this context, a ``faithful'' embedding should
+ideally preserve the topology of the data, ensuring that cells
+corresponding to the same cell type are situated close to the
+high-dimensional space.
+
+\hypertarget{processing-steps}{%
+\subsection{Processing steps}\label{processing-steps}}
+
+To begin our analysis, we installed the Peripheral Blood Mononuclear
+Cells (pbmc) data set obtained from 10x Genomics using the
+\texttt{SeuratData} R package \citep{Rahul2019}, which facilitates the
+distribution of data sets in the form of Seurat objects
+\citep{Yuhan2021}. This data set contains 13,714 features (genes) across
+2,700 samples (cells) within a single assay. The active assay is RNA,
+with 13,714 features representing different gene expressions. Then, the
+cells that have unique feature counts over 2,500 or less than 200 and
+cells that have \textgreater{} 5\% mitochondrial counts are filtered.
+After removing unwanted cells from the dataset, the next step is to
+normalize the data. By default, we employ a global-scaling normalization
+method ``LogNormalize'' that normalizes the feature expression
+measurements for each cell by the total expression, multiplies this by a
+scale factor (10,000 by default), and log-transforms the result.
+
+Next, top 1000 genes are selected by Festem \citep{Chen2023} DEG method.
+Then, scale the data. Then, perform PCA on the scaled data. Louvain
+algorithm is used to cluster the single cells based on the genes
+selected by Festem. Using the Festem-selected genes and 15 PCs,
+identified 10 clusters in the PBMC3k data and annotated them based on
+the expression of canonical markers. The 10 clusters included immune
+cells such as naive CD4 T cells, memory CD4 T cells, CD8 T cells, CD14
+monocytes, FCGR3A monocytes, Natural Killer (NK) cells, B cells and
+Dendritic Cells (DC). In addition to these common cell types, Festem
+identified a fine cell type, CD27− CD4+ memory T cells, which were often
+missed by other methods.
+
+\begin{itemize}
+\item
+  Human peripheral blood mononclear cells (PBMCs) data
+\item
+  Intro to the data set
+\end{itemize}
+
+Used Louvain algorithm to cluster the single cells based on the genes
+selected by Festem method
+
+\begin{itemize}
+\item
+  Explain what ``Evaluate Festem on DEG detection'' means, in terms of
+  this data
+\item
+\item
+  How did they ``identified 10 clusters''? What were the variables used,
+  PCs or full data or 2D representation, spell this out.
+\item
+  Used top 1000 genes selected by Festem to run PCA, top 15 PCs
+  (original paper mentioned)
+\item
+  The ``10 clusters included immune cells such as naive CD4 T \ldots{}''
+  does this mean that every cluster had all of these, or one cluster was
+  mostly ``CD4 T cells'', another mostly ``memory CD4 T'', \ldots?
+\item
+  How is your use of this data different from the original application.
+\item
+  Original paper used this data set to compare different DEG methods,
+  but we need to look how the performance of UMAP with the author's
+  selected parameter choice.
+\item
+  Why 15 PCs? Is that from the original paper? YES
+\end{itemize}
 
@@
 \item
-  Numb bins along the x-axis = 23, shape parameter = 0.8772751
+  Num bins along the x-axis = 23, shape parameter = 0.8772751
 \item
@@
   the preservation of the broad topological structure.
+\item
+  The third parameter is metric. This parameter plays a crucial role in
+  determining how distances are computed within the ambient space of the
+  input data (The term ``ambient space'' refers to the original or
+  initial space in which the data exists before any transformation or
+  dimensionality reduction).
 \end{itemize}
 
-\begin{tabular}{r|r|l|r}
-\hline
-30 & 0.3 & cosine & 38634.66719286174\\
-\hline
-5 & 0.99 & euclidean & 35938.99\\
-\hline
-5 & 0.99 & cosine & 35597.02\\
-\hline
-15 & 0.99 & euclidean & 35534.76\\
-\hline
-30 & 0.99 & euclidean & 35664.69\\
-\hline
-1022 & 0.99 & cosine & 37022.01\\
-\hline
-1022 & 0.30 & cosine & 39746.35\\
-\hline
-800 & 0.30 & euclidean & 38385.25\\
-\hline
-\end{tabular}
-
-\begin{figure}
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-nldervisPBMCUMAP-1.pdf}
+\begin{longtable}[]{@{}llll@{}}
+\toprule\noalign{}
+n\_neighbors & min\_dist & metric & error \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+\textbf{30} & \textbf{0.3} & \textbf{cosine} &
+\textbf{38634.66719286174} \\
+5 & 0.99 & euclidean & 35938.98594054407 \\
+5 & 0.99 & cosine & 35597.01981261912 \\
+15 & 0.99 & euclidean & 35534.75778935427 \\
+30 & 0.99 & euclidean & 35664.68763936665 \\
+1022 & 0.99 & cosine & 37022.0125249776 \\
+1022 & 0.3 & cosine & 39746.347504639125 \\
+800 & 0.3 & euclidean & 38385.25485420897 \\
+100 & 0.99 & euclidean & 36528.57424747017 \\
+230 & 0.99 & cosine & 37864.26802578186 \\
+354 & 0.99 & euclidean & 37530.86598361975 \\
+\end{longtable}
+
+\begin{figure}[h]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervisPBMCUMAP-1.pdf}
 
@@
 
-\begin{figure}
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-nldervisPBMCUMAP2-1.pdf}
+\begin{figure}[h]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervisPBMCUMAP2-1.pdf}
 
@@
 
-\begin{figure}
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-nldervisPBMCUMAP3-1.pdf}
+\begin{figure}[h]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervisPBMCUMAP3-1.pdf}
 
@@
   neighbors each point has.
+\item
+  Num bins along the x-axis = 43, shape parameter = 0.4466309
 \end{itemize}

### 2024-01-30 paper.qmd
Rearrange some figures and removed unnecessary images and unnessary sections

@@
 
+<!--UMAP with different param choices-->
+
+```{r}
+#| warning: false
+#| echo: false
+
+UMAP_data_7 <- read_rds(file = "data/s_curve/s_curve_umap_7.rds")
+
+#(n-neighbors: 50)
+plot_list1_umap <- plot_UMAP_2D(UMAP_data_7) + #ggtitle("(b)") +
+  geom_point(alpha=0.5, size = 0.5, colour = "#8dd3c7", fill = "#8dd3c7") +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+UMAP_data_15 <- read_rds(file = "data/s_curve/s_curve_umap_15.rds")
+
+#(n-neighbors: 50)
+plot_list2_umap <- plot_UMAP_2D(UMAP_data_15) + #ggtitle("(b)") + 
+  geom_point(alpha=0.5, size = 0.5, colour = "#a65628", fill = "#a65628") +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+UMAP_data_32 <- read_rds(file = "data/s_curve/s_curve_umap_32.rds")
+
+#(n-neighbors: 50)
+plot_list3_umap <- plot_UMAP_2D(UMAP_data_32) + #ggtitle("(b)") +
+  geom_point(alpha=0.5, size = 0.5, color = "#f781bf", fill = "#f781bf") +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+UMAP_data <- read_rds(file = "data/s_curve/s_curve_umap.rds")
+
+#(n-neighbors: 50)
+plot_list4_umap <- plot_UMAP_2D(UMAP_data) + #ggtitle("(b)") +
+  geom_point(alpha=0.5, size = 0.5, colour = "#999999", fill = "#999999") +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'd', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "2D layouts from UMAP applied for the S-curve data: (a) UMAP (n_neighbors = 7), (b) UMAP (n_neighbors = 15), (c) UMAP (n_neighbors = 32), (d) UMAP (n_neighbors = 50).
+Is there a best hyperparameter choice in representing UMAP or are they all providing  equivalent information?"
+#| label: fig-nldervisUMAP
+#| out-width: 100%
+
+plot_list1_umap + plot_list2_umap + plot_list3_umap + plot_list4_umap +
+  plot_layout(ncol=4)
+```
+
 <!--High-dimensional (high-D) data is widespread in many fields including ecology and bioinformatics [@Guo2023], in part because of new data collection technologies [@Johnstone2009; @ayesha2020overview].
@@
 
-num_bins_for_nldr_methods_df <- tibble::tibble(NLDR_technique = "tSNE", tot_bins = 72, aic = -21882.462)
-
-num_bins_for_nldr_methods_df <- num_bins_for_nldr_methods_df |> 
-  tibble::add_row(NLDR_technique = "UMAP", tot_bins = 98, aic = -20013.266)
-
-num_bins_for_nldr_methods_df <- num_bins_for_nldr_methods_df |> 
-  tibble::add_row(NLDR_technique = "PHATE", tot_bins = 90, aic = -15036.91)
-
-num_bins_for_nldr_methods_df <- num_bins_for_nldr_methods_df |> 
-  tibble::add_row(NLDR_technique = "TriMAP", tot_bins = 154, aic = -23001.90)
-
-num_bins_for_nldr_methods_df <- num_bins_for_nldr_methods_df |> 
-  tibble::add_row(NLDR_technique = "PaCMAP", tot_bins = 132, aic = -21330.059)
-
-# knitr::kable(
-#   num_bins_for_nldr_methods_df,
-#   col.names = c('NLDR technique', 'total number of bins'), format = "latex", 
-#              caption = "Total number of bins suggested by rule of thumb for each NLDR technique applied to S-curve dataset."
-# ) |> 
+# num_bins_for_nldr_methods_df <- tibble::tibble(NLDR_technique = "tSNE", tot_bins = 72, aic = -21882.462)
+# 
+# num_bins_for_nldr_methods_df <- num_bins_for_nldr_methods_df |> 
+#   tibble::add_row(NLDR_technique = "UMAP", tot_bins = 98, aic = -20013.266)
+# 
+# num_bins_for_nldr_methods_df <- num_bins_for_nldr_methods_df |> 
+#   tibble::add_row(NLDR_technique = "PHATE", tot_bins = 90, aic = -15036.91)
+# 
+# num_bins_for_nldr_methods_df <- num_bins_for_nldr_methods_df |> 
+#   tibble::add_row(NLDR_technique = "TriMAP", tot_bins = 154, aic = -23001.90)
+# 
+# num_bins_for_nldr_methods_df <- num_bins_for_nldr_methods_df |> 
+#   tibble::add_row(NLDR_technique = "PaCMAP", tot_bins = 132, aic = -21330.059)
+# 
+# # knitr::kable(
+# #   num_bins_for_nldr_methods_df,
+# #   col.names = c('NLDR technique', 'total number of bins'), format = "latex", 
+# #              caption = "Total number of bins suggested by rule of thumb for each NLDR technique applied to S-curve dataset."
+# # ) |> 
+# #   kable_styling(latex_options = c("striped", "HOLD_position"), position = "center", full_width = FALSE) |>
+# #   row_spec(0, bold = TRUE)
+# num_bins_for_nldr_methods_df$tot_bins <- as.character(num_bins_for_nldr_methods_df$tot_bins)
+# colnames(num_bins_for_nldr_methods_df) <- c('NLDR technique', 'total number of bins', "AIC")
+# num_bins_for_nldr_methods_df <- tibble::remove_rownames(num_bins_for_nldr_methods_df)
+# 
+# xtable(num_bins_for_nldr_methods_df, caption = "Total number of bins suggested by rule of thumb and AIC values for each NLDR technique applied to S-curve dataset.", label = "tbl-msebinsnldr") |>
+#   xtable2kable() |> 
 #   kable_styling(latex_options = c("striped", "HOLD_position"), position = "center", full_width = FALSE) |>
-#   row_spec(0, bold = TRUE)
-num_bins_for_nldr_methods_df$tot_bins <- as.character(num_bins_for_nldr_methods_df$tot_bins)
-colnames(num_bins_for_nldr_methods_df) <- c('NLDR technique', 'total number of bins', "AIC")
-num_bins_for_nldr_methods_df <- tibble::remove_rownames(num_bins_for_nldr_methods_df)
-
-xtable(num_bins_for_nldr_methods_df, caption = "Total number of bins suggested by rule of thumb and AIC values for each NLDR technique applied to S-curve dataset.", label = "tbl-msebinsnldr") |>
-  xtable2kable() |> 
-  kable_styling(latex_options = c("striped", "HOLD_position"), position = "center", full_width = FALSE) |>
-  row_spec(0, bold = TRUE) |>
-  row_spec(4, bold=T, color = "red")
+#   row_spec(0, bold = TRUE) |>
+#   row_spec(4, bold=T, color = "red")
 
@@
   theme_light() +
-  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
         axis.title = element_text(size = 7),
@@
 
-<!--UMAP with different param choices-->
-
-```{r}
-#| warning: false
-#| echo: false
-
-UMAP_data_7 <- read_rds(file = "data/s_curve/s_curve_umap_7.rds")
-
-#(n-neighbors: 50)
-plot_list1_umap <- plot_UMAP_2D(UMAP_data_7) + #ggtitle("(b)") +
-  geom_point(alpha=0.5, size = 0.5, colour = "#8dd3c7", fill = "#8dd3c7") +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-UMAP_data_15 <- read_rds(file = "data/s_curve/s_curve_umap_15.rds")
-
-#(n-neighbors: 50)
-plot_list2_umap <- plot_UMAP_2D(UMAP_data_15) + #ggtitle("(b)") + 
-  geom_point(alpha=0.5, size = 0.5, colour = "#a65628", fill = "#a65628") +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-UMAP_data_32 <- read_rds(file = "data/s_curve/s_curve_umap_32.rds")
-
-#(n-neighbors: 50)
-plot_list3_umap <- plot_UMAP_2D(UMAP_data_32) + #ggtitle("(b)") +
-  geom_point(alpha=0.5, size = 0.5, color = "#f781bf", fill = "#f781bf") +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-UMAP_data <- read_rds(file = "data/s_curve/s_curve_umap.rds")
-
-#(n-neighbors: 50)
-plot_list4_umap <- plot_UMAP_2D(UMAP_data) + #ggtitle("(b)") +
-  geom_point(alpha=0.5, size = 0.5, colour = "#999999", fill = "#999999") +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'd', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "2D layouts from UMAP applied for the S-curve data: (a) UMAP (n_neighbors = 7), (b) UMAP (n_neighbors = 15), (c) UMAP (n_neighbors = 32), (d) UMAP (n_neighbors = 50).
-Is there a best parameter choice in representing UMAP or are they all providing  equivalent information?"
-#| label: fig-nldervisUMAP
-#| out-width: 100%
-
-plot_list1_umap + plot_list2_umap + plot_list3_umap + plot_list4_umap +
-  plot_layout(ncol=4)
-```
-
+<!--
 ```{r}
@@
 
-### Model selection {#sec-selection}
+<!--### Model selection {#sec-selection}
 
@@
 For example, as shown in @fig-diagnosticpltScurve (a), TriMAP is the most suitable NLDR technique to represent the S-curve data (see table 1).
+-->
 

### 2024-01-30 paper.tex
Rearrange some figures and removed unnecessary images and unnessary sections

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable, boxrule=0pt, interior hidden, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, boxrule=0pt, breakable, frame hidden, enhanced]}{\end{tcolorbox}}\fi
 
@@
 
+\begin{figure}
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervisUMAP-1.pdf}
+
+}
+
+\caption{\label{fig-nldervisUMAP}2D layouts from UMAP applied for the
+S-curve data: (a) UMAP (n\_neighbors = 7), (b) UMAP (n\_neighbors = 15),
+(c) UMAP (n\_neighbors = 32), (d) UMAP (n\_neighbors = 50). Is there a
+best hyperparameter choice in representing UMAP or are they all
+providing equivalent information?}
+
+\end{figure}
+
 \hypertarget{sec-background}{%
@@
 look in high-D with original data.
-
-\begin{table}[H]
-\centering
-\begin{tabular}{rllr}
-  \hline
-\textbf{} & \textbf{NLDR technique} & \textbf{total number of bins} & \textbf{AIC }\\ 
-  \hline
-\cellcolor{gray!6}{1} & \cellcolor{gray!6}{tSNE} & \cellcolor{gray!6}{72} & \cellcolor{gray!6}{-21882.46 }\\ 
-  2 & UMAP & 98 & -20013.27 \\ 
-\cellcolor{gray!6}{  3} & \cellcolor{gray!6}{PHATE} & \cellcolor{gray!6}{90} & \cellcolor{gray!6}{-15036.91 }\\ 
-\textcolor{red}{\textbf{  4}} & \textcolor{red}{\textbf{TriMAP}} & \textcolor{red}{\textbf{154}} & \textcolor{red}{\textbf{-23001.90 }}\\ 
-\cellcolor{gray!6}{  5} & \cellcolor{gray!6}{PaCMAP} & \cellcolor{gray!6}{132} & \cellcolor{gray!6}{-21330.06 }\\ 
-   \hline
-\end{tabular}
-\caption{Total number of bins suggested by rule of thumb and AIC values for each NLDR technique applied to S-curve dataset.} 
-\label{tbl-msebinsnldr}
-\end{table}
 
@@
 complexity in the predictions.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervisUMAP-1.pdf}
-
-}
-
-\caption{\label{fig-nldervisUMAP}2D layouts from UMAP applied for the
-S-curve data: (a) UMAP (n\_neighbors = 7), (b) UMAP (n\_neighbors = 15),
-(c) UMAP (n\_neighbors = 32), (d) UMAP (n\_neighbors = 50). Is there a
-best parameter choice in representing UMAP or are they all providing
-equivalent information?}
-
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticpltDiffParam-1.pdf}
-
-}
-
-\caption{\label{fig-diagnosticpltDiffParam}Goodness of fit statistics
-from different n\_neighbors parameter of UMAP applied to training
-S-curve dataset. What is the best parameter choice in UMAP to represent
-the original data in 2D?}
-
-\end{figure}
 
@@
 
-\hypertarget{sec-selection}{%
-\subsection{Model selection}\label{sec-selection}}
-
-In the process of selecting the most suitable NLDR technique to
-represent high-D data, the Akaike Information Criterion (AIC) stands as
-a crucial tool for insightful comparisons. Once the effective number of
-bins has been determined for each method, AIC values can be compared to
-identify the best NLDR technique with the lowest AIC value. For example,
-as shown in Figure~\ref{fig-diagnosticpltScurve} (a), TriMAP is the most
-suitable NLDR technique to represent the S-curve data (see table 1).
-
 \hypertarget{sec-prediction}{%

### 2024-01-30 appendix.qmd
Add different options in our model to Appendix

@@
 
-data <- read_csv(paste0(here::here(), "/data/s_curve.csv"))
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-data <- data |>
-  mutate(ID = row_number())
-
-data_split <- initial_split(data)
-training_data <- training(data_split)
-test_data <- testing(data_split)
-```
-
-```{r}
-#| echo: false
-
-UMAP_fit <- umap(training_data |> dplyr::select(-ID), n_neighbors = 50, n_components =  2)
-
-UMAP_data <- UMAP_fit$layout |>
-  as.data.frame()
-names(UMAP_data)[1:(ncol(UMAP_data))] <- paste0(rep("UMAP",(ncol(UMAP_data))), 1:(ncol(UMAP_data)))
-
-UMAP_data <- UMAP_data |>
-  mutate(ID = training_data$ID)
+data <- read_rds("data/s_curve/s_curve.rds")
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+training_data <- read_rds("data/s_curve/s_curve_training.rds")
+test_data <- read_rds("data/s_curve/s_curve_test.rds")
+```
+
+```{r}
+#| echo: false
+
+UMAP_data <- read_rds("data/s_curve/s_curve_umap.rds")
 
@@
 ```
+
+## Notations
 
@@
 ```
+
+## Different options to construct the 2D model and lift the model into high-D space
+
+Mainly, the hexagonal bin centroids are used to construct 2D model and averaged high-D data is used to lift the model into high-D space.
+As options except bin centroids, bin means can used to construct the 2D model.
+Also, except averaged high-D data, weighted mean can used to lift the model into high-D space.
+If we use bin centers instead of bin centroids, the triangles are not regular.
+
+- Why bin means? because the distribution of points within hexagonals are not always reprsented by the center.
+
+### Case 1: Construct 2D model with hexbin means and lift the model into high-D space with averaged high-D data
+
+```{r}
+#| warning: false
+#| echo: false
+
+## Import data
+
+UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap.rds")
+
+## UMAP
+
+num_bins_umap_s_curve <- calculate_effective_x_bins(.data = UMAP_s_curve, x = UMAP1,
+                                         cell_area = 1)
+shape_val_umap_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
+                                                          x = UMAP1, y = UMAP2) ## 1.259938
+## To extract bin centroids
+hexbin_data_object_umap_s_curve <- extract_hexbin_mean(nldr_df = UMAP_s_curve, num_bins = num_bins_umap_s_curve, shape_val = shape_val_umap_s_curve, x = UMAP1, y = UMAP2)
+
+df_bin_centroids_umap_s_curve <- hexbin_data_object_umap_s_curve$hexdf_data
+
+## To generate the grid
+hexbin_data_object_n <- extract_hexbin_centroids(nldr_df = UMAP_s_curve, num_bins = num_bins_umap_s_curve, shape_val = shape_val_umap_s_curve, x = UMAP1, y = UMAP2)
+df_bin_centroids_n <- hexbin_data_object_n$hexdf_data
+## Data set with all possible centroids in the hexagonal grid
+full_centroid_df <- generate_full_grid_centroids(df_bin_centroids_n)
+## Generate all coordinates of hexagons
+hex_grid <- full_hex_grid(full_centroid_df)
+
+UMAP_data_with_hb_id_s_curve <- UMAP_s_curve |>
+  dplyr::mutate(hb_id = hexbin_data_object_umap_s_curve$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_umap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), UMAP_data_with_hb_id_s_curve)
+
+## Averaged on high-D
+df_bin_umap_s_curve <- avg_highD_data(.data = df_all_umap_s_curve)
+
+## Triangulate bin centroids
+tr1_object_umap_s_curve <- triangulate_bin_centroids(df_bin_centroids_umap_s_curve, x, y)
+tr_from_to_df_umap_s_curve <- generate_edge_info(triangular_object = tr1_object_umap_s_curve)
+
+# ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
+#   geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
+#   geom_point(size = 2, colour = "#33a02c") +
+#   coord_equal()
+
+## Compute 2D distances
+distance_umap_s_curve <- cal_2D_dist(.data = tr_from_to_df_umap_s_curve)
+
+## To find the benchmark value
+benchmark_umap_s_curve <- find_benchmark_value(.data = distance_umap_s_curve, distance_col = distance)
+benchmark_umap_s_curve <- 2.0384346
+
+# colour_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve,
+#                   triangular_object = tr1_object_umap_s_curve, distance_col = distance)
+
+# trimesh_umap_s_curve <- ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
+#   geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
+#   geom_point(size = 2, colour = "#33a02c") +
+#   coord_equal()
+# 
+# trimesh_umap_s_curve <- trimesh_umap_s_curve +
+#   #ggtitle("(a)") +
+#   xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+#   theme_light() +
+#   theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
+#         axis.title.x = element_blank(), axis.title.y = element_blank(),
+#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
+#   ) +
+#   annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
+# theme(axis.text = element_text(size = 5),
+#       axis.title = element_text(size = 7))
+
+trimesh_removed_umap_s_curve_bin_mean <- remove_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve, triangular_object = tr1_object_umap_s_curve, distance_col = distance)
+
+trimesh_removed_umap_s_curve_bin_mean <- trimesh_removed_umap_s_curve_bin_mean +
+  geom_point(size = 2, colour = "#33a02c") +
+  #ggtitle("(b)") +
+  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  theme_light() +
+  #coord_equal() +
+  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
+  ) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+# ### Define type column
+# df <- df_all_umap_s_curve |>
+#   dplyr::select(tidyselect::starts_with("x")) |>
+#   dplyr::mutate(type = "data") ## original dataset
+# 
+# df_b <- df_bin_umap_s_curve |>
+#   dplyr::filter(hb_id %in% df_bin_centroids_umap_s_curve$hexID) |>
+#   dplyr::select(-hb_id) |>
+#   dplyr::mutate(type = "model") ## Data with summarized mean
+# 
+# df_exe <- dplyr::bind_rows(df_b, df)
+# 
+# distance_df_small_edges <- distance_umap_s_curve %>%
+#   dplyr::filter(distance < benchmark_umap_s_curve)
+# 
+# distance_df_long_edges <- distance_umap_s_curve %>%
+#   dplyr::filter(distance >= benchmark_umap_s_curve)
+# 
+# distance_df_edges <- dplyr::bind_rows(distance_df_small_edges, distance_df_long_edges)
+# 
+# langevitour::langevitour(df_exe[1:(length(df_exe)-1)], lineFrom = distance_df_edges$from , lineTo = distance_df_edges$to, group = df_exe$type, pointSize = 3, levelColors = c("#6a3d9a", "#33a02c"), lineColors = append(rep("black", length(distance_df_small_edges$from)), rep("red", length(distance_df_long_edges$from))))
+# 
+# show_langevitour(df_all_umap_s_curve, df_bin_umap_s_curve, df_bin_centroids_umap_s_curve, benchmark_value = benchmark_umap_s_curve, distance = distance_umap_s_curve, distance_col = distance)
+
+ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
+  geom_point(data = df_bin_centroids_n, aes(x = x, y = y), color = "#33a02c") +
+  geom_point(data = df_bin_centroids_umap_s_curve, aes(x = x, y = y), color = "black")
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## UMAP
+## Prediction
+
+shape_value_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
+                                                   x = UMAP1, y = UMAP2)
+
+num_bins_vec <- 1:13 ## Number of bins along the x-axis
+
+vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(num_bins_vec)) {
+
+  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_s_curve, nldr_df_test = UMAP_s_curve, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
+  pred_df_training <- pred_df_training_object$pred_data
+  centroid_df_training <- pred_df_training_object$df_bin_centroids
+  avg_df_training <- pred_df_training_object$df_bin
+
+  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+
+  eval_df_training <- eval_df_training |>
+    mutate(num_bins_x = num_bins_vec[i])
+
+  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+
+}
+
+## Add new column with data types
+
+eval_data_training <- eval_data_training |>
+  mutate(data_type = "training")
+
+MSE_df_2 <- eval_data_training |>
+  dplyr::mutate(method = "UMAP")
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+#MSE_df <- dplyr::bind_rows(MSE_df_1, MSE_df_2, MSE_df_3, MSE_df_4, MSE_df_5)
+MSE_df <- MSE_df_2
+
+MSE_df$method <- factor(MSE_df$method, levels = c("UMAP"))
+
+## To draw with AIC
+aic_plot <- ggplot(MSE_df |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
+                                                                                 y = total_error,
+                                                                                 color = method
+)) +
+  geom_point() +
+  geom_line() +
+  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("AIC") +
+  xlab("Total number of bins")
+## Effective number of bins along x-axis
+
+mse_plot <- ggplot(MSE_df |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
+                                       y = total_mse,
+                                       color = method
+)) +
+  geom_point() +
+  geom_line() +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("MSE") +
+  xlab("Total number of bins")
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset.
+What is the most effective NLDR technique to represent the original data in a 2D space?
+What is the effective number of bins in each NLDR technique to create a 2D model?"
+#| label: fig-diagnosticpltScurve
+##| out-width: 100%
+
+trimesh_removed_umap_s_curve_bin_mean + aic_plot + mse_plot +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='none', plot.tag = element_text(size = 8))
+```
+
+### Case 2: Construct 2D model with hexbin centroid and lift the model into high-D space with weighted mean of high-D data
+
+```{r}
+#| warning: false
+#| echo: false
+
+## Import data
+
+UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap.rds")
+
+## UMAP
+
+num_bins_umap_s_curve <- calculate_effective_x_bins(.data = UMAP_s_curve, x = UMAP1,
+                                         cell_area = 1)
+shape_val_umap_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
+                                                          x = UMAP1, y = UMAP2) ## 1.259938
+## To extract bin centroids
+hexbin_data_object_umap_s_curve <- extract_hexbin_centroids(nldr_df = UMAP_s_curve, num_bins = num_bins_umap_s_curve, shape_val = shape_val_umap_s_curve, x = UMAP1, y = UMAP2)
+
+df_bin_centroids_umap_s_curve <- hexbin_data_object_umap_s_curve$hexdf_data
+
+## Data set with all possible centroids in the hexagonal grid
+full_centroid_df <- generate_full_grid_centroids(df_bin_centroids_umap_s_curve)
+## Generate all coordinates of hexagons
+hex_grid <- full_hex_grid(full_centroid_df)
+
+UMAP_data_with_hb_id_s_curve <- UMAP_s_curve |>
+  dplyr::mutate(hb_id = hexbin_data_object_umap_s_curve$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_umap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), UMAP_data_with_hb_id_s_curve)
+
+## Weighted averaged on high-D
+hb_object <- hexbin_data_object_umap_s_curve$hb_data
+df_bin_umap_s_curve <- weighted_highD_data(training_data, nldr_df_with_id = UMAP_data_with_hb_id_s_curve, hb_object, column_start_text = "x")
+
+## Triangulate bin centroids
+tr1_object_umap_s_curve <- triangulate_bin_centroids(df_bin_centroids_umap_s_curve, x, y)
+tr_from_to_df_umap_s_curve <- generate_edge_info(triangular_object = tr1_object_umap_s_curve)
+
+# ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
+#   geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
+#   geom_point(size = 2, colour = "#33a02c") +
+#   coord_equal()
+
+## Compute 2D distances
+distance_umap_s_curve <- cal_2D_dist(.data = tr_from_to_df_umap_s_curve)
+
+## To find the benchmark value
+benchmark_umap_s_curve <- find_benchmark_value(.data = distance_umap_s_curve, distance_col = distance)
+#benchmark_umap_s_curve <- 2.0384346
+
+# colour_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve,
+#                   triangular_object = tr1_object_umap_s_curve, distance_col = distance)
+
+# trimesh_umap_s_curve <- ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
+#   geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
+#   geom_point(size = 2, colour = "#33a02c") +
+#   coord_equal()
+# 
+# trimesh_umap_s_curve <- trimesh_umap_s_curve +
+#   #ggtitle("(a)") +
+#   xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+#   theme_light() +
+#   theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
+#         axis.title.x = element_blank(), axis.title.y = element_blank(),
+#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
+#   ) +
+#   annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
+# theme(axis.text = element_text(size = 5),
+#       axis.title = element_text(size = 7))
+
+trimesh_removed_umap_s_curve_bin_centroid <- remove_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve, triangular_object = tr1_object_umap_s_curve, distance_col = distance)
+
+trimesh_removed_umap_s_curve_bin_centroid <- trimesh_removed_umap_s_curve_bin_centroid +
+  geom_point(size = 2, colour = "#33a02c") +
+  #ggtitle("(b)") +
+  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  theme_light() +
+  #coord_equal() +
+  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
+  ) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+# ### Define type column
+# df <- df_all_umap_s_curve |>
+#   dplyr::select(tidyselect::starts_with("x")) |>
+#   dplyr::mutate(type = "data") ## original dataset
+# 
+# df_b <- df_bin_umap_s_curve |>
+#   dplyr::filter(hb_id %in% df_bin_centroids_umap_s_curve$hexID) |>
+#   dplyr::select(-hb_id) |>
+#   dplyr::mutate(type = "model") ## Data with summarized mean
+# 
+# df_exe <- dplyr::bind_rows(df_b, df)
+# 
+# distance_df_small_edges <- distance_umap_s_curve %>%
+#   dplyr::filter(distance < benchmark_umap_s_curve)
+# 
+# distance_df_long_edges <- distance_umap_s_curve %>%
+#   dplyr::filter(distance >= benchmark_umap_s_curve)
+# 
+# distance_df_edges <- dplyr::bind_rows(distance_df_small_edges, distance_df_long_edges)
+# 
+# langevitour::langevitour(df_exe[1:(length(df_exe)-1)], lineFrom = distance_df_edges$from , lineTo = distance_df_edges$to, group = df_exe$type, pointSize = 3, levelColors = c("#6a3d9a", "#33a02c"), lineColors = append(rep("black", length(distance_df_small_edges$from)), rep("red", length(distance_df_long_edges$from))))
+# 
+# show_langevitour(df_all_umap_s_curve, df_bin_umap_s_curve, df_bin_centroids_umap_s_curve, benchmark_value = benchmark_umap_s_curve, distance = distance_umap_s_curve, distance_col = distance)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## UMAP
+## Prediction
+
+shape_value_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
+                                                   x = UMAP1, y = UMAP2)
+
+num_bins_vec <- 1:13 ## Number of bins along the x-axis
+
+vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(num_bins_vec)) {
+
+  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_s_curve, nldr_df_test = UMAP_s_curve, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
+  pred_df_training <- pred_df_training_object$pred_data
+  centroid_df_training <- pred_df_training_object$df_bin_centroids
+  avg_df_training <- pred_df_training_object$df_bin
+
+  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+
+  eval_df_training <- eval_df_training |>
+    mutate(num_bins_x = num_bins_vec[i])
+
+  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+
+}
+
+## Add new column with data types
+
+eval_data_training <- eval_data_training |>
+  mutate(data_type = "training")
+
+MSE_df_2 <- eval_data_training |>
+  dplyr::mutate(method = "UMAP")
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+#MSE_df <- dplyr::bind_rows(MSE_df_1, MSE_df_2, MSE_df_3, MSE_df_4, MSE_df_5)
+MSE_df <- MSE_df_2
+
+MSE_df$method <- factor(MSE_df$method, levels = c("UMAP"))
+
+## To draw with AIC
+aic_plot <- ggplot(MSE_df |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
+                                                                                 y = total_error,
+                                                                                 color = method
+)) +
+  geom_point() +
+  geom_line() +
+  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("AIC") +
+  xlab("Total number of bins")
+## Effective number of bins along x-axis
+
+mse_plot <- ggplot(MSE_df |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
+                                       y = total_mse,
+                                       color = method
+)) +
+  geom_point() +
+  geom_line() +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("MSE") +
+  xlab("Total number of bins")
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset.
+What is the most effective NLDR technique to represent the original data in a 2D space?
+What is the effective number of bins in each NLDR technique to create a 2D model?"
+#| label: fig-diagnosticpltScurve2
+##| out-width: 100%
+
+trimesh_removed_umap_s_curve_bin_centroid + aic_plot + mse_plot +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='none', plot.tag = element_text(size = 8))
+```
+
+### Case 3: Construct 2D model with hexbin means and lift the model into high-D space with weighted mean of high-D data
+
+```{r}
+#| warning: false
+#| echo: false
+
+## Import data
+
+UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap.rds")
+
+## UMAP
+
+num_bins_umap_s_curve <- calculate_effective_x_bins(.data = UMAP_s_curve, x = UMAP1,
+                                         cell_area = 1)
+shape_val_umap_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
+                                                          x = UMAP1, y = UMAP2) ## 1.259938
+## To extract bin centroids
+hexbin_data_object_umap_s_curve <- extract_hexbin_mean(nldr_df = UMAP_s_curve, num_bins = num_bins_umap_s_curve, shape_val = shape_val_umap_s_curve, x = UMAP1, y = UMAP2)
+
+df_bin_centroids_umap_s_curve <- hexbin_data_object_umap_s_curve$hexdf_data
+
+## To generate the grid
+hexbin_data_object_n <- extract_hexbin_centroids(nldr_df = UMAP_s_curve, num_bins = num_bins_umap_s_curve, shape_val = shape_val_umap_s_curve, x = UMAP1, y = UMAP2)
+df_bin_centroids_n <- hexbin_data_object_n$hexdf_data
+## Data set with all possible centroids in the hexagonal grid
+full_centroid_df <- generate_full_grid_centroids(df_bin_centroids_n)
+## Generate all coordinates of hexagons
+hex_grid <- full_hex_grid(full_centroid_df)
+
+UMAP_data_with_hb_id_s_curve <- UMAP_s_curve |>
+  dplyr::mutate(hb_id = hexbin_data_object_umap_s_curve$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_umap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), UMAP_data_with_hb_id_s_curve)
+
+## Weighted averaged on high-D
+hb_object <- hexbin_data_object_umap_s_curve$hb_data
+df_bin_umap_s_curve <- weighted_highD_data(training_data, nldr_df_with_id = UMAP_data_with_hb_id_s_curve, hb_object, column_start_text = "x")
+
+## Triangulate bin centroids
+tr1_object_umap_s_curve <- triangulate_bin_centroids(df_bin_centroids_umap_s_curve, x, y)
+tr_from_to_df_umap_s_curve <- generate_edge_info(triangular_object = tr1_object_umap_s_curve)
+
+# ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
+#   geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
+#   geom_point(size = 2, colour = "#33a02c") +
+#   coord_equal()
+
+## Compute 2D distances
+distance_umap_s_curve <- cal_2D_dist(.data = tr_from_to_df_umap_s_curve)
+
+## To find the benchmark value
+benchmark_umap_s_curve <- find_benchmark_value(.data = distance_umap_s_curve, distance_col = distance)
+benchmark_umap_s_curve <- 2.0384346
+
+# colour_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve,
+#                   triangular_object = tr1_object_umap_s_curve, distance_col = distance)
+
+# trimesh_umap_s_curve <- ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
+#   geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
+#   geom_point(size = 2, colour = "#33a02c") +
+#   coord_equal()
+# 
+# trimesh_umap_s_curve <- trimesh_umap_s_curve +
+#   #ggtitle("(a)") +
+#   xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+#   theme_light() +
+#   theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
+#         axis.title.x = element_blank(), axis.title.y = element_blank(),
+#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
+#   ) +
+#   annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
+# theme(axis.text = element_text(size = 5),
+#       axis.title = element_text(size = 7))
+
+trimesh_removed_umap_s_curve_bin_mean <- remove_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve, triangular_object = tr1_object_umap_s_curve, distance_col = distance)
+
+trimesh_removed_umap_s_curve_bin_mean <- trimesh_removed_umap_s_curve_bin_mean +
+  geom_point(size = 2, colour = "#33a02c") +
+  #ggtitle("(b)") +
+  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  theme_light() +
+  #coord_equal() +
+  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
+  ) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+# ### Define type column
+# df <- df_all_umap_s_curve |>
+#   dplyr::select(tidyselect::starts_with("x")) |>
+#   dplyr::mutate(type = "data") ## original dataset
+# 
+# df_b <- df_bin_umap_s_curve |>
+#   dplyr::filter(hb_id %in% df_bin_centroids_umap_s_curve$hexID) |>
+#   dplyr::select(-hb_id) |>
+#   dplyr::mutate(type = "model") ## Data with summarized mean
+# 
+# df_exe <- dplyr::bind_rows(df_b, df)
+# 
+# distance_df_small_edges <- distance_umap_s_curve %>%
+#   dplyr::filter(distance < benchmark_umap_s_curve)
+# 
+# distance_df_long_edges <- distance_umap_s_curve %>%
+#   dplyr::filter(distance >= benchmark_umap_s_curve)
+# 
+# distance_df_edges <- dplyr::bind_rows(distance_df_small_edges, distance_df_long_edges)
+# 
+# langevitour::langevitour(df_exe[1:(length(df_exe)-1)], lineFrom = distance_df_edges$from , lineTo = distance_df_edges$to, group = df_exe$type, pointSize = 3, levelColors = c("#6a3d9a", "#33a02c"), lineColors = append(rep("black", length(distance_df_small_edges$from)), rep("red", length(distance_df_long_edges$from))))
+# 
+# show_langevitour(df_all_umap_s_curve, df_bin_umap_s_curve, df_bin_centroids_umap_s_curve, benchmark_value = benchmark_umap_s_curve, distance = distance_umap_s_curve, distance_col = distance)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## UMAP
+## Prediction
+
+shape_value_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
+                                                   x = UMAP1, y = UMAP2)
+
+num_bins_vec <- 1:13 ## Number of bins along the x-axis
+
+vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(num_bins_vec)) {
+
+  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_s_curve, nldr_df_test = UMAP_s_curve, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
+  pred_df_training <- pred_df_training_object$pred_data
+  centroid_df_training <- pred_df_training_object$df_bin_centroids
+  avg_df_training <- pred_df_training_object$df_bin
+
+  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+
+  eval_df_training <- eval_df_training |>
+    mutate(num_bins_x = num_bins_vec[i])
+
+  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+
+}
+
+## Add new column with data types
+
+eval_data_training <- eval_data_training |>
+  mutate(data_type = "training")
+
+MSE_df_2 <- eval_data_training |>
+  dplyr::mutate(method = "UMAP")
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+#MSE_df <- dplyr::bind_rows(MSE_df_1, MSE_df_2, MSE_df_3, MSE_df_4, MSE_df_5)
+MSE_df <- MSE_df_2
+
+MSE_df$method <- factor(MSE_df$method, levels = c("UMAP"))
+
+## To draw with AIC
+aic_plot <- ggplot(MSE_df |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
+                                                                                 y = total_error,
+                                                                                 color = method
+)) +
+  geom_point() +
+  geom_line() +
+  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("AIC") +
+  xlab("Total number of bins")
+## Effective number of bins along x-axis
+
+mse_plot <- ggplot(MSE_df |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
+                                       y = total_mse,
+                                       color = method
+)) +
+  geom_point() +
+  geom_line() +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("MSE") +
+  xlab("Total number of bins")
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset.
+What is the most effective NLDR technique to represent the original data in a 2D space?
+What is the effective number of bins in each NLDR technique to create a 2D model?"
+#| label: fig-diagnosticpltScurve3
+##| out-width: 100%
+
+trimesh_removed_umap_s_curve_bin_mean + aic_plot + mse_plot +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='none', plot.tag = element_text(size = 8))
+```

### 2024-01-30 paper.qmd
Add different options in our model to Appendix

@@
 library(grid)
-library(xtable)
-
-xtable2kable <- function(x) {
-  out <- capture.output(print(x, table.placement = NULL))[-(1:2)]
-  out <- paste(out, collapse = "\n")
-  structure(out, format = "latex", class = "knitr_kable")
-}
+# library(xtable)
+# 
+# xtable2kable <- function(x) {
+#   out <- capture.output(print(x, table.placement = NULL))[-(1:2)]
+#   out <- paste(out, collapse = "\n")
+#   structure(out, format = "latex", class = "knitr_kable")
+# }
 
@@
         axis.text = element_text(size = 7)) +
-  scale_colour_manual(values = c("#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
+  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
   ylab("AIC") +
@@
   # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  scale_colour_manual(values = c("#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
+  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
   ylab("MSE") +
@@
   plot_layout(guides='collect', ncol = 2) &
-  theme(legend.position='bottom', plot.tag = element_text(size = 8))
+  theme(legend.position='none', plot.tag = element_text(size = 8))
 ```
@@
   plot_layout(guides='collect', ncol = 2) &
-  theme(legend.position='bottom', plot.tag = element_text(size = 8)) 
+  theme(legend.position='none', plot.tag = element_text(size = 8)) 
 ```

### 2024-01-30 paper.tex
Add different options in our model to Appendix

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, boxrule=0pt, breakable, frame hidden, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, sharp corners, interior hidden, breakable, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden]}{\end{tcolorbox}}\fi
 

### 2024-01-30 appendix.qmd
Add notation table in the Appendix

@@
 
+| Notation                                                                                                              | Description                                                            |
+|-----------------------------------------------------------------------------------------------------------------------|------------------------------------------------------------------------|
+| $n$                                                                                                                   | number of observations                                                 |
+| $p$                                                                                                                   | number of dimensions in high-D data                                    |
+| $X_{n \times p} = \begin{bmatrix} \textbf{x} _{1} & \textbf{x}_ {2} & \cdots & \textbf{x}_{n} \\  \end{bmatrix}^\top$ | high-D data matrix                                                     |
+| $d$                                                                                                                   | number of dimensions in embedding data, usually two                    |
+| $Y_{n \times d} = \begin{bmatrix} \textbf{y} _{1} & \textbf{y}_ {2} & \cdots & \textbf{y}_{n} \\  \end{bmatrix}^\top$ | embedding matrix                                                       |
+| $P_t$                                                                                                                 | point along the geodesic path at time $t$                              |
+| $P_{t+1}$                                                                                                             | point along the geodesic path at the next time instant, which is $t+1$ |
+| $b_1$                                                                                                                 | number of bins along the x-axis                                        |
+| $b_2$                                                                                                                 | number of bins along the y-axis                                        |
+| $b$                                                                                                                   | total number of bins in the hexagonal grid                             |
+| $b'$                                                                                                                  | total number of bins without the empty bins                            |
+| $h$                                                                                                                   | diameter of a hexagon                                                  |
+| $A$                                                                                                                   | area of a hexagon                                                      |
+| $r_1$                                                                                                                 | 2D embedding component 1                                               |
+| $r_2$                                                                                                                 | 2D embedding component 2                                               |
+| $s$                                                                                                                   | shape parameter                                                        |
+| $C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2})$                                                                               | hexagonal bin centroid coordinates                                     |
+| $C_k^{(p)} \equiv (C_{kx_1}, ..., C_{kx_p})$                                                                          | averaged high-D coordinates                                            |
+: Notations used in this paper and their descriptions. {#tbl-notations}
+
 ## Paramter tunning
@@
 
-- Why bin means? because the distribution of points within hexagonals are not always reprsented by the center.
+- Why bin means? because the distribution of points within hexagonals are not always represented by the center.
 

### 2024-01-30 paper.tex
Add notation table in the Appendix

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, sharp corners, interior hidden, breakable, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, boxrule=0pt, frame hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, interior hidden, breakable]}{\end{tcolorbox}}\fi
 

### 2024-01-31 appendix.qmd
Add tSNE diff perplexity values with PBMC3k data

@@
 
+## Default parameter values
+
+### Benchmark value to remove low-density hexagons
+
+Removing all the bins that have a standard number of points less than a certain benchmark value will generate unnecessary holes within the triangular mesh.
+Therefore, before removing the low-density bins, it's necessary to investigate the number of points in neighboring bins as well.
+
+$$
+\text{standard count} = \frac{\text{count}}{\text{max count}} 
+$${#eq-equationp2}
+
+The process begins by identifying the six nearest hex bins of specified low-density hex bins.
+To identify the nearest hex bins 2D Euclidean distance metric is used.
+Then, we calculate the mean density, as outlined in the equations:
+
+$$
+\text{mean density} = \frac{\text{standard count}}{6} 
+$${#eq-equationp3}
+
+The standard count is derived from the number of observations in the hex bins.
+By examining the distribution of mean densities and designating the first quartile as the benchmark value, hex bins with mean densities below this benchmark are removed.
+This process ensures the elimination of regions with insufficient data density, focusing on areas with more significant data representation and preserving the overall structure in the low-dimensional space.
+
+### Benchmark value to remove long edges {#sec-rmlgdefault}
+
+Determining a benchmark value for removing long edges isn't a straightforward process with standard guidelines.
+Having a default value provides a practical starting point for decision-making, helping users choose whether to stick with the default, choose a value below it, or select one above it for long edge removal.
+
+To calculate the default value, the process involves computing the 2D Euclidean distances between the vertices in the triangular mesh.
+Initially, distances between vertices are sorted.
+Then, unique distance values are extracted.
+Following this, the differences between consecutive distance values are computed, and a data frame is constructed with unique distance values and their corresponding differences.
+The first largest difference is then used as a benchmark to identify and remove the long edges.
+
+![A flow diagram detailing the steps taken to find the benchmark value to remove long edges.](figures/remove_long_edges_workflow.png){#fig-rmlgmeth fig-align="center" width="100%" height="100%"}
+
+```{r}
+#| echo: false
+
+dist_p <- plot_dist(distance)  +
+  geom_hline(yintercept = benchmark, colour = "blue") + 
+  ylab(expression(d^{(2)})) +
+  theme(axis.text = element_text(size = 5),
+        axis.title = element_text(size = 12)) 
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| fig-cap: "Is the default benchmark value effective in removing the long edges to construct a smooth surface in 2D?
+(a) Triangular mesh with coloured edges by their type, (b) Distribution of 2D Euclidean distances between triangular mesh vertices."
+#| label: fig-defaultrmlg
+##| out-width: 100%
+
+trimesh_gr + dist_p +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='bottom', plot.tag = element_text(size = 8))
+```
+
 ## Paramter tunning
@@
 
-In this step, we aim to create a smoother surface in the low-dimensional space by removing long edges from the triangular mesh.
-This process helps eliminate outliers and noise while preserving important local relationships and intricate structures within the data.
-To achieve this, distances between vertices are sorted, and unique distance values are extracted.
-A data frame is created to calculate the differences between consecutive distance values, and the largest difference serves as a threshold to identify long edges.
-By removing edges that exceed this threshold, the algorithm refines the triangular mesh (see @fig-traingularmeshgr).
-
-![A flow diagram detailing the steps taken to find the benchmark value to remove long edges.](figures/remove_long_edges_workflow.png){#fig-rmlgmeth fig-align="center" width="100%" height="100%"}
-
-```{r}
-#| echo: false
-
-dist_p <- plot_dist(distance)  +
-  geom_hline(yintercept = benchmark, colour = "red") + 
-  ylab(expression(d^{(2)})) +
-  theme(axis.text = element_text(size = 5),
-        axis.title = element_text(size = 12))
-
-dist_p
-```
-
-```{r}
-#| echo: false
-
-trimesh + trimesh_gr +
-  plot_layout(ncol = 2)
-
+Creating a smooth surfaces in 2D involves experimenting with various benchmark values.
+The challenge lies in determining the most effective benchmark value.
+This is precisely why the default value was introduced, as outlined in @sec-rmlgdefault, serving as a fundamental reference.
+To assess its effectiveness, a visual inspection of the triangular mesh is conducted to verify whether it successfully eliminates long edges.
+If it does, proceeding with that value is a effective choice; if not, it signals the initiation of the tuning process.
+@fig-diffrmlg shows how the 2D triangular meshes look after removing long edges with different benchmark values.
+Higher values tend to retain long edges, while lower values may remove not only long edges but also shorter ones.
+
+```{r}
+#| warning: false
+#| echo: false
+
+## Import data
+
+UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap.rds")
+
+## UMAP
+
+num_bins_umap_s_curve <- calculate_effective_x_bins(.data = UMAP_s_curve, x = UMAP1,
+                           cell_area = 1)
+shape_val_umap_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
+                                                          x = UMAP1, y = UMAP2) ## 1.259938
+## To extract bin centroids
+hexbin_data_object_umap_s_curve <- extract_hexbin_centroids(nldr_df = UMAP_s_curve, num_bins = num_bins_umap_s_curve, shape_val = shape_val_umap_s_curve, x = UMAP1, y = UMAP2)
+
+df_bin_centroids_umap_s_curve <- hexbin_data_object_umap_s_curve$hexdf_data
+
+UMAP_data_with_hb_id_s_curve <- UMAP_s_curve |>
+  dplyr::mutate(hb_id = hexbin_data_object_umap_s_curve$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_umap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), UMAP_data_with_hb_id_s_curve)
+
+## Averaged on high-D
+df_bin_umap_s_curve <- avg_highD_data(.data = df_all_umap_s_curve)
+
+## Triangulate bin centroids
+tr1_object_umap_s_curve <- triangulate_bin_centroids(df_bin_centroids_umap_s_curve, x, y)
+tr_from_to_df_umap_s_curve <- generate_edge_info(triangular_object = tr1_object_umap_s_curve)
+
+# ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
+#   geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
+#   geom_point(size = 2, colour = "#33a02c") +
+#   coord_equal()
+
+## Compute 2D distances
+distance_umap_s_curve <- cal_2D_dist(.data = tr_from_to_df_umap_s_curve)
+
+## To find the benchmark value
+benchmark_umap_s_curve <- find_benchmark_value(.data = distance_umap_s_curve, distance_col = distance)
+
+# colour_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve,
+#                   triangular_object = tr1_object_umap_s_curve, distance_col = distance)
+
+trimesh_umap_s_curve <- ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
+  geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
+  geom_point(size = 2, colour = "#33a02c") +
+  coord_equal()
+
+trimesh_umap_s_curve <- trimesh_umap_s_curve +
+  #ggtitle("(a)") +
+  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  theme_light() +
+  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
+  ) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
+# theme(axis.text = element_text(size = 5),
+#       axis.title = element_text(size = 7))
+
+trimesh_gr_umap_s_curve <- colour_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve, triangular_object = tr1_object_umap_s_curve, distance_col = distance)
+
+trimesh_gr_umap_s_curve <- trimesh_gr_umap_s_curve +
+  geom_point(size = 2, colour = "#33a02c") +
+  #ggtitle("(b)") +
+  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  theme_light() +
+  #coord_equal() +
+  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
+  ) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+trimesh_removed_umap_s_curve <- remove_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve,
+                                                  triangular_object = tr1_object_umap_s_curve, distance_col = distance)
+
+trimesh_removed_umap_s_curve <- trimesh_removed_umap_s_curve +
+  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  # theme(axis.text = element_text(size = 5),
+  #       axis.title = element_text(size = 7)) +
+  geom_point(colour = "#33a02c", size = 0.5) +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+trimesh_removed_umap_s_curve1 <- remove_long_edges(.data = distance_umap_s_curve, benchmark_value = 0.8794136,
+                                                  triangular_object = tr1_object_umap_s_curve, distance_col = distance)
+
+trimesh_removed_umap_s_curve1 <- trimesh_removed_umap_s_curve1 +
+  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  # theme(axis.text = element_text(size = 5),
+  #       axis.title = element_text(size = 7)) +
+  geom_point(colour = "#33a02c", size = 0.5) +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+trimesh_removed_umap_s_curve2 <- remove_long_edges(.data = distance_umap_s_curve, benchmark_value = 5.0891822,
+                                                  triangular_object = tr1_object_umap_s_curve, distance_col = distance)
+
+trimesh_removed_umap_s_curve2 <- trimesh_removed_umap_s_curve2 +
+  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  # theme(axis.text = element_text(size = 5),
+  #       axis.title = element_text(size = 7)) +
+  geom_point(colour = "#33a02c", size = 0.5) +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "What is the effective benchmark value to remove long edges? (a) benchmark value = $0.8794136$, (b) benchmark value = $1.696$ (default), and (c) benchmark value = $5.0891822$."
+#| label: fig-diffrmlg
+##| out-width: 100%
+
+trimesh_removed_umap_s_curve1 + trimesh_removed_umap_s_curve + trimesh_removed_umap_s_curve2 +
+  plot_layout(guides='collect', ncol = 3) &
+  theme(legend.position='bottom')
 ```
@@
 ```
+
+## How model perform less than and greater than the selected number of bins (UMAP with S-curve)

### 2024-01-31 paper.qmd
Add tSNE diff perplexity values with PBMC3k data

@@
 
+- If it is completely a mess, then that is also further evidence that the fit is poor in high-dimensions.
+
 ```{r}
@@
 trimesh_removed_pbmc_umap + plot_list2_pbmc_error +
+plot_layout(guides='collect', ncol=3) &
+  theme(legend.position='bottom')
+
+#  plot_list1_pbmc + plot_list3_pbmc + plot_list4_pbmc + plot_list5_pbmc +
+#   plot_layout(ncol=5) #| out-width: 100%
+
+# 2D layouts from different NLDR techniques applied for the training PBMC dataset: (a) tSNE (perplexity = 30), (b) UMAP (n_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2).
+Is there a best representation of the original data or are they all providing  equivalent information?
+```
+
+<!--tSNE parm 1-->
+
+```{r}
+#| warning: false
+#| echo: false
+
+## Import data
+#df_2 <- read_rds("data/pbmc/pbmc.rds")
+training_data_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_pca_50.rds")
+training_data_pbmc <- training_data_pbmc[, 1:15] |>
+  mutate(ID = 1:NROW(training_data_pbmc))
+#test_1 <- read_rds("data/pbmc/pbmc_test.rds")
+
+tSNE_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_tsne_5.rds")
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+num_bins_pbmc <- calculate_effective_x_bins(.data = tSNE_pbmc, x = tSNE1,
+                                            cell_area = 1) ##23
+
+num_bins_pbmc <- 23
+
+shape_value_pbmc <- calculate_effective_shape_value(.data = tSNE_pbmc,
+                                                    x = tSNE1, y = tSNE2) ## 0.8772751
+shape_value_pbmc <- 0.8772751
+
+## To extract bin centroids
+hexbin_data_object_pbmc <- extract_hexbin_centroids(nldr_df = tSNE_pbmc,
+                                                    num_bins = num_bins_pbmc,
+                                                    shape_val = shape_value_pbmc, x = tSNE1, y = tSNE2)
+
+df_bin_centroids_pbmc <- hexbin_data_object_pbmc$hexdf_data
+
+tSNE_pbmc_with_hb_id <- tSNE_pbmc |>
+  dplyr::mutate(hb_id = hexbin_data_object_pbmc$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), tSNE_pbmc_with_hb_id)
+
+## Averaged on high-D
+df_bin_pbmc <- avg_highD_data(.data = df_all_pbmc, column_start_text = "PC")
+
+## Triangulate bin centroids
+tr1_object_pbmc <- triangulate_bin_centroids(df_bin_centroids_pbmc, x, y)
+tr_from_to_df_pbmc <- generate_edge_info(triangular_object = tr1_object_pbmc)
+
+## Compute 2D distances
+distance_pbmc <- cal_2D_dist(.data = tr_from_to_df_pbmc)
+
+## To find the benchmark value
+benchmark_pbmc <- find_benchmark_value(.data = distance_pbmc, distance_col = distance)
+
+trimesh_removed_pbmc_tsne <- remove_long_edges(.data = distance_pbmc, benchmark_value = benchmark_pbmc,
+                                               triangular_object = tr1_object_pbmc, distance_col = distance)
+
+trimesh_removed_pbmc_tsne <- trimesh_removed_pbmc_tsne +
+  geom_point(size = 2, colour = "#33a02c") +
+  #ggtitle("(b)") +
+  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  theme_light() +
+  #coord_equal() +
+  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
+  )
+
+df_bin_train <- df_bin_pbmc
+names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])
+
+error_df <- df_all_pbmc |>
+  dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates
+
+# prediction_df_join <- prediction_df_join |>
+#   dplyr::left_join(data, by = c("ID" = "ID")) ## Map high-D data
+
+for (i in 1:(NCOL(df_bin_train) - 1)) {
+
+  error_df[ , paste0("abs_residual_", "PC_", i)] <- abs(error_df[ , paste0("PC_", i)] - error_df[ , paste0("avg_", "PC_", i)])
+
+}
+
+error_df <- error_df |>
+  dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "PC_")))))
+
+# library(ggbeeswarm)
+# error_df$group <- "1"
+# ggplot(error_df, aes(x = group, y = total)) +
+#   geom_quasirandom()+
+#   ylim(0, max(unlist(error_df$total))+ 0.5) + coord_flip()
+
+### The minimum error is 0 and the maximum is 42.17439
+### There is lot of points with error 0,
+
+error_df <- error_df |>
+  mutate(type = if_else(total <= 2, "error 2 or less",
+                        if_else(total <= 10, "error 2-10",
+                                if_else(total <= 15, "error 10-15",
+                                        if_else(total <= 20, "error 15-20",
+                                                if_else(total <= 25, "error 20-25",
+                                                        if_else(total <= 30, "error 25-30",
+                                                                if_else(total <= 35, "error 30-35", "error greter than 35"))))))))
+
+plot_list2_pbmc_error <- error_df |>
+    mutate(type = factor(type , levels = c("error 2 or less", "error 2-10", "error 10-15", "error 15-20", "error 20-25", "error 25-30", "error 30-35", "error greter than 35"))) |>
+    ggplot(aes(x = tSNE1,
+               y = tSNE2, color = type))+
+    geom_point(alpha=0.5, size = 0.1) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#fee0d2", "#fcbba1",
+                                         "#fc9272", "#fb6a4a", "#ef3b2c",
+                                         "#cb181d", "#a50f15", "#99000d"))
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "(a) 2D layout from tSNE (perplexity = 5) applied for the PBMC3k dataset. Is this a best representation of the original data?, (b)  Model in the 2D space with tSNE (<>)"
+#| label: fig-nldervisPBMCtSNE1
+#| out-width: 100%
+#| fig-pos: h
+
+#plot_list2_pbmc + 
+trimesh_removed_pbmc_tsne + plot_list2_pbmc_error +
+plot_layout(guides='collect', ncol=3) &
+  theme(legend.position='bottom')
+
+#  plot_list1_pbmc + plot_list3_pbmc + plot_list4_pbmc + plot_list5_pbmc +
+#   plot_layout(ncol=5) #| out-width: 100%
+
+# 2D layouts from different NLDR techniques applied for the training PBMC dataset: (a) tSNE (perplexity = 30), (b) UMAP (n_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2).
+Is there a best representation of the original data or are they all providing  equivalent information?
+```
+
+<!--tSNE parm 2-->
+
+```{r}
+#| warning: false
+#| echo: false
+
+## Import data
+#df_2 <- read_rds("data/pbmc/pbmc.rds")
+training_data_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_pca_50.rds")
+training_data_pbmc <- training_data_pbmc[, 1:15] |>
+  mutate(ID = 1:NROW(training_data_pbmc))
+#test_1 <- read_rds("data/pbmc/pbmc_test.rds")
+
+tSNE_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_tsne_2.rds")
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+num_bins_pbmc <- calculate_effective_x_bins(.data = tSNE_pbmc, x = tSNE1,
+                                            cell_area = 1) ##23
+
+num_bins_pbmc <- 23
+
+shape_value_pbmc <- calculate_effective_shape_value(.data = tSNE_pbmc,
+                                                    x = tSNE1, y = tSNE2) ## 0.8772751
+shape_value_pbmc <- 0.8772751
+
+## To extract bin centroids
+hexbin_data_object_pbmc <- extract_hexbin_centroids(nldr_df = tSNE_pbmc,
+                                                    num_bins = num_bins_pbmc,
+                                                    shape_val = shape_value_pbmc, x = tSNE1, y = tSNE2)
+
+df_bin_centroids_pbmc <- hexbin_data_object_pbmc$hexdf_data
+
+tSNE_pbmc_with_hb_id <- tSNE_pbmc |>
+  dplyr::mutate(hb_id = hexbin_data_object_pbmc$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), tSNE_pbmc_with_hb_id)
+
+## Averaged on high-D
+df_bin_pbmc <- avg_highD_data(.data = df_all_pbmc, column_start_text = "PC")
+
+## Triangulate bin centroids
+tr1_object_pbmc <- triangulate_bin_centroids(df_bin_centroids_pbmc, x, y)
+tr_from_to_df_pbmc <- generate_edge_info(triangular_object = tr1_object_pbmc)
+
+## Compute 2D distances
+distance_pbmc <- cal_2D_dist(.data = tr_from_to_df_pbmc)
+
+## To find the benchmark value
+benchmark_pbmc <- find_benchmark_value(.data = distance_pbmc, distance_col = distance)
+
+trimesh_removed_pbmc_tsne <- remove_long_edges(.data = distance_pbmc, benchmark_value = benchmark_pbmc,
+                                               triangular_object = tr1_object_pbmc, distance_col = distance)
+
+trimesh_removed_pbmc_tsne <- trimesh_removed_pbmc_tsne +
+  geom_point(size = 2, colour = "#33a02c") +
+  #ggtitle("(b)") +
+  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  theme_light() +
+  #coord_equal() +
+  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
+  )
+
+df_bin_train <- df_bin_pbmc
+names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])
+
+error_df <- df_all_pbmc |>
+  dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates
+
+# prediction_df_join <- prediction_df_join |>
+#   dplyr::left_join(data, by = c("ID" = "ID")) ## Map high-D data
+
+for (i in 1:(NCOL(df_bin_train) - 1)) {
+
+  error_df[ , paste0("abs_residual_", "PC_", i)] <- abs(error_df[ , paste0("PC_", i)] - error_df[ , paste0("avg_", "PC_", i)])
+
+}
+
+error_df <- error_df |>
+  dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "PC_")))))
+
+# library(ggbeeswarm)
+# error_df$group <- "1"
+# ggplot(error_df, aes(x = group, y = total)) +
+#   geom_quasirandom()+
+#   ylim(0, max(unlist(error_df$total))+ 0.5) + coord_flip()
+
+### The minimum error is 0 and the maximum is 42.17439
+### There is lot of points with error 0,
+
+error_df <- error_df |>
+  mutate(type = if_else(total <= 2, "error 2 or less",
+                        if_else(total <= 10, "error 2-10",
+                                if_else(total <= 15, "error 10-15",
+                                        if_else(total <= 20, "error 15-20",
+                                                if_else(total <= 25, "error 20-25",
+                                                        if_else(total <= 30, "error 25-30",
+                                                                if_else(total <= 35, "error 30-35", "error greter than 35"))))))))
+
+plot_list2_pbmc_error <- error_df |>
+    mutate(type = factor(type , levels = c("error 2 or less", "error 2-10", "error 10-15", "error 15-20", "error 20-25", "error 25-30", "error 30-35", "error greter than 35"))) |>
+    ggplot(aes(x = tSNE1,
+               y = tSNE2, color = type))+
+    geom_point(alpha=0.5, size = 0.1) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#fee0d2", "#fcbba1",
+                                         "#fc9272", "#fb6a4a", "#ef3b2c",
+                                         "#cb181d", "#a50f15", "#99000d"))
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "(a) 2D layout from tSNE (perplexity = 2) applied for the PBMC3k dataset. Is this a best representation of the original data?, (b)  Model in the 2D space with tSNE (<>)"
+#| label: fig-nldervisPBMCtSNE2
+#| out-width: 100%
+#| fig-pos: h
+
+#plot_list2_pbmc + 
+trimesh_removed_pbmc_tsne + plot_list2_pbmc_error +
 plot_layout(guides='collect', ncol=3) &

### 2024-01-31 paper.tex
Add tSNE diff perplexity values with PBMC3k data

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, boxrule=0pt, frame hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, interior hidden, breakable]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, boxrule=0pt, frame hidden, sharp corners]}{\end{tcolorbox}}\fi
 
@@
   dimensionality reduction).
+\item
+  If it is completely a mess, then that is also further evidence that
+  the fit is poor in high-dimensions.
 \end{itemize}
@@
 
+\begin{figure}[h]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervisPBMCtSNE1-1.pdf}
+
+}
+
+\caption{\label{fig-nldervisPBMCtSNE1}(a) 2D layout from tSNE
+(perplexity = 5) applied for the PBMC3k dataset. Is this a best
+representation of the original data?, (b) Model in the 2D space with
+tSNE (\textless\textgreater)}
+
+\end{figure}
+
+\begin{figure}[h]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervisPBMCtSNE2-1.pdf}
+
+}
+
+\caption{\label{fig-nldervisPBMCtSNE2}(a) 2D layout from tSNE
+(perplexity = 2) applied for the PBMC3k dataset. Is this a best
+representation of the original data?, (b) Model in the 2D space with
+tSNE (\textless\textgreater)}
+
+\end{figure}
+
 \hypertarget{harps-gto-sample-dataset}{%

### 2024-02-01 paper.qmd
Update the repo

@@
 #| fig-pos: H
-#| fig-cap: "Hexbin plots from different number of bins for the **UMAP** projections of **S-curve** data: (a) b = 32 (4, 8), s = 1.643542, (b) b = 264 (12, 22), s = 1.643542, and (c) b = 840 (21, 40), s = 1.643542.
+#| fig-cap: "Hexbin plots from different number of bins for the **UMAP** projections of **S-curve** training data: (a) b = 32 (4, 8), s = 1.643542, (b) b = 264 (12, 22), s = 1.643542, and (c) b = 840 (21, 40), s = 1.643542.
 The hexbins are colored based on the density of points, with darker colors indicating higher point density and yellow color representing lower point density within each bin.
@@
   geom_line() +
+   geom_vline(xintercept = 144, linetype="dashed", 
+                color = "red", size=0.5) +
   theme_light() +
@@
 #| echo: false
-#| fig-cap: "Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset.
-What is the most effective NLDR technique to represent the original data in a 2D space?
-What is the effective number of bins in each NLDR technique to create a 2D model?"
+#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset.
+What is the effective number of bins in each NLDR technique to create a 2D model?
+The MSE plot have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
+Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error.
+Using the elbow method, when the total number of bins is set to 144, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
+This point indicates that adding less bins does not enough to capture the data structure."
 #| label: fig-diagnosticpltScurve
@@
 
+hex_full_count_df_loop2 <- read_rds("data/s_curve/s_curve_hex_8.rds")
+
 hex_full_count_df_loop2 <- hex_full_count_df_loop2 |>
-  dplyr::mutate(type = if_else(std_counts < 0.12, "low", "high"))
+  dplyr::mutate(type = if_else(std_counts <= 0.20, "low", "high"))
 
@@
 
-num_bins_umap_s_curve <- 11
+num_bins_umap_s_curve <- 8
 shape_val_umap_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
@@
 df_bin_centroids_umap_s_curve <- df_bin_centroids_umap_s_curve |>
-  dplyr::filter(std_counts > 0.12)
+  dplyr::filter(std_counts > 0.20)
 
@@
   ylab("AIC") +
-  xlab("Benchmark value")
+  xlab("benchmark value")
 ## Effective number of bins along x-axis
@@
   geom_point() +
+  geom_line() +
+  geom_vline(xintercept = 0.2, linetype="dashed", 
+                color = "red", size=0.5) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("MSE") +
+  xlab("benchmark value")
+
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset with different benchmark values to remove the low-density heaxgons.
+What is the effective benchmark value to remove the low-density heaxgons?
+The MSE plot have a steep slope at the beginning, indicating that a smaller benchmark causes a small amount of error.
+Then, the slope gradually increases or level up, indicating that a higher number of benchmark values generates a higher error.
+Using the reverse elbow method, when the benchmark value is set to 0.2, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
+This point indicates that higher benchmark values remove the necessary bins as well which lead to distruct the 2D structure."
+#| label: fig-diagnosticpltScurvelwd
+##| out-width: 100%
+
+aic_plot_lwd + mse_plot_lwd +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='none', plot.tag = element_text(size = 8))
+```
+
+#### Benchmark value to remove long edges
+
+Achieving a smooth representation in 2D space is crucial, and one factor influencing this smoothness is the presence of long edges (see @fig-modelScurvermlgimp).
+These long edges occur when a line connects points that are distant in the triangular mesh, impacting the overall smoothness of the mesh.
+
+To investigate the effect of removing long edges on the MSE, we analyzed the MSE for various benchmark values.
+Surprisingly, the results indicated that the MSE remained consistent across different benchmark values, as shown in @fig-diagnosticpltScurvelgrm.
+It's important to note that edges are defined only in the 2D model and do not extend to higher dimensions.
+Consequently, removing edges does not affect the model in higher dimensions.
+
+There is no definitive rule for determining the benchmark value to remove long edges.
+However, we have a method to find a default value (see Appendix).
+Adjusting values around this default can help to find benchmark value to remove long edges, contributing to the construction of a smoother 2D representation (see @fig-diagnosticpltScurvelgrm).
+
+<!--The removal of long edges is a critical step to create a smoother representation  by iteratively eliminating hexagons with excessive distances between centroids.
+This process eliminates outliers and noise while preserving essential local relationships within the data.
+To achieve this, distances between vertices are sorted, and unique distance values are extracted.
+By setting a threshold based on the largest difference between consecutive distance values, long edges are identified and removed.
+This refinement step contributes to enhancing the quality of the triangular mesh, ensuring a more accurate representation of the data structure.-->
+
+```{r}
+#| warning: false
+#| echo: false
+
+## Import data
+
+UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap.rds")
+
+## UMAP
+
+num_bins_umap_s_curve <- 8
+shape_val_umap_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
+                                                          x = UMAP1, y = UMAP2) ## 1.259938
+## To extract bin centroids
+hexbin_data_object_umap_s_curve <- extract_hexbin_centroids(nldr_df = UMAP_s_curve, num_bins = num_bins_umap_s_curve, shape_val = shape_val_umap_s_curve, x = UMAP1, y = UMAP2)
+
+df_bin_centroids_umap_s_curve <- hexbin_data_object_umap_s_curve$hexdf_data
+
+UMAP_data_with_hb_id_s_curve <- UMAP_s_curve |>
+  dplyr::mutate(hb_id = hexbin_data_object_umap_s_curve$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_umap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), UMAP_data_with_hb_id_s_curve)
+
+## Averaged on high-D
+df_bin_umap_s_curve <- avg_highD_data(.data = df_all_umap_s_curve)
+
+## Triangulate bin centroids
+tr1_object_umap_s_curve <- triangulate_bin_centroids(df_bin_centroids_umap_s_curve, x, y)
+tr_from_to_df_umap_s_curve <- generate_edge_info(triangular_object = tr1_object_umap_s_curve)
+
+# ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
+#   geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
+#   geom_point(size = 2, colour = "#33a02c") +
+#   coord_equal()
+
+## Compute 2D distances
+distance_umap_s_curve <- cal_2D_dist(.data = tr_from_to_df_umap_s_curve)
+
+## To find the benchmark value
+benchmark_umap_s_curve <- find_benchmark_value(.data = distance_umap_s_curve, distance_col = distance)
+benchmark_umap_s_curve <- 1.346694
+
+# colour_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve,
+#                   triangular_object = tr1_object_umap_s_curve, distance_col = distance)
+
+trimesh_umap_s_curve <- ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
+  geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
+  geom_point(size = 2, colour = "#33a02c") +
+  coord_equal()
+
+trimesh_umap_s_curve <- trimesh_umap_s_curve +
+  #ggtitle("(a)") +
+  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  theme_light() +
+  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
+  ) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
+# theme(axis.text = element_text(size = 5),
+#       axis.title = element_text(size = 7))
+
+trimesh_gr_umap_s_curve <- colour_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve, triangular_object = tr1_object_umap_s_curve, distance_col = distance)
+
+trimesh_gr_umap_s_curve <- trimesh_gr_umap_s_curve +
+  geom_point(size = 2, colour = "#33a02c") +
+  #ggtitle("(b)") +
+  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  theme_light() +
+  #coord_equal() +
+  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
+  ) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+# ### Define type column
+# df <- df_all_umap_s_curve |>
+#   dplyr::select(tidyselect::starts_with("x")) |>
+#   dplyr::mutate(type = "data") ## original dataset
+# 
+# df_b <- df_bin_umap_s_curve |>
+#   dplyr::filter(hb_id %in% df_bin_centroids_umap_s_curve$hexID) |>
+#   dplyr::select(-hb_id) |>
+#   dplyr::mutate(type = "model") ## Data with summarized mean
+# 
+# df_exe <- dplyr::bind_rows(df_b, df)
+# 
+# distance_df_small_edges <- distance_umap_s_curve %>%
+#   dplyr::filter(distance < benchmark_umap_s_curve)
+# 
+# distance_df_long_edges <- distance_umap_s_curve %>%
+#   dplyr::filter(distance >= benchmark_umap_s_curve)
+# 
+# distance_df_edges <- dplyr::bind_rows(distance_df_small_edges, distance_df_long_edges)
+# 
+# langevitour::langevitour(df_exe[1:(length(df_exe)-1)], lineFrom = distance_df_edges$from , lineTo = distance_df_edges$to, group = df_exe$type, pointSize = 3, levelColors = c("#6a3d9a", "#33a02c"), lineColors = append(rep("black", length(distance_df_small_edges$from)), rep("red", length(distance_df_long_edges$from))))
+#
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+#| label: fig-modelScurvermlgimp
+#| fig-cap: "What does the 2D model, constructed using UMAP (applied to the S-curve dataset), look like in high dimensions with long edges?
+(<https://youtu.be/pSL_G_n7-fM>) (a) 2D model without removing the long edges, and (b) 2D model with long edges coloured by red.
+The default benchmark value is used to identify long edges."
+##| out-width: 100%
+
+trimesh_umap_s_curve + trimesh_gr_umap_s_curve +
+  #plot_annotation(tag_levels = 'a') +
+  plot_layout(ncol = 2) &
+  theme(legend.position='none', plot.tag = element_text(size = 8))
+```
+
+```{r}
+#| echo: false
+#| warning: false
+
+error_df_lg_1 <- read_rds("data/s_curve/s_curve_summary_lg_tsne.rds")
+error_df_lg_2 <- read_rds("data/s_curve/s_curve_summary_lg_umap.rds")
+error_df_lg_3 <- read_rds("data/s_curve/s_curve_summary_lg_phate.rds")
+error_df_lg_4 <- read_rds("data/s_curve/s_curve_summary_lg_trimap.rds")
+error_df_lg_5 <- read_rds("data/s_curve/s_curve_summary_lg_pacmap.rds")
+
+error_df_lg <- dplyr::bind_rows(error_df_lg_1, error_df_lg_2, error_df_lg_3, error_df_lg_4, error_df_lg_5)
+
+error_df_lg$method <- factor(error_df_lg$method, levels = c("tSNE", "UMAP", "PHATE", "TriMAP", "PaCMAP"))
+
+## To draw with AIC
+aic_plot_lg <- ggplot(error_df_lg |> dplyr::filter(method == "UMAP"), aes(x = benchmark_rm_lg,
+                                                                                 y = total_error,
+                                                                                 color = method
+)) +
+  geom_point() +
+  geom_point(data = error_df_lg |> dplyr::filter(method == "UMAP") |> filter(benchmark_rm_lg == 1.347), aes(x = benchmark_rm_lg,
+                                       y = total_error
+),
+                                       color = "#e41a1c") +
+  geom_line() +
+  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("AIC") +
+  xlab("benchmark value")
+## Effective number of bins along x-axis
+
+mse_plot_lg <- ggplot(error_df_lg |> dplyr::filter(method == "UMAP"), aes(x = benchmark_rm_lg,
+                                       y = total_mse,
+                                       color = method
+)) +
+  geom_point() +
+  geom_point(data = error_df_lg |> dplyr::filter(method == "UMAP") |> filter(benchmark_rm_lg == 1.347), aes(x = benchmark_rm_lg,
+                                       y = total_mse
+),
+                                       color = "#e41a1c") +
   geom_line() +
@@
   ylab("MSE") +
-  xlab("Benchmark value")
-
-```
-
-```{r}
-#| echo: false
-#| warning: false
-#| fig-cap: "Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset with different benchmark values to remove the low-density heaxgons.
-What is the effective benchmark value to remove the low-density heaxgons?"
-#| label: fig-diagnosticpltScurvelwd
-##| out-width: 100%
-
-aic_plot_lwd + mse_plot_lwd +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', ncol = 2) &
-  theme(legend.position='none', plot.tag = element_text(size = 8))
-```
-
-#### Benchmark value to remove long edges
-
-Achieving a smooth representation in 2D space is crucial, and one factor influencing this smoothness is the presence of long edges (see @fig-modelScurvermlgimp).
-These long edges occur when a line connects points that are distant in the triangular mesh, impacting the overall smoothness of the mesh.
-
-To investigate the effect of removing long edges on the MSE, we analyzed the MSE for various benchmark values.
-Surprisingly, the results indicated that the MSE remained consistent across different benchmark values, as shown in @fig-diagnosticpltScurvelgrm.
-It's important to note that edges are defined only in the 2D model and do not extend to higher dimensions.
-Consequently, removing edges does not affect the model in higher dimensions.
-
-There is no definitive rule for determining the benchmark value to remove long edges.
-However, we have a method to find a default value (see Appendix).
-Adjusting values around this default can help to find benchmark value to remove long edges, contributing to the construction of a smoother 2D representation (see @fig-diagnosticpltScurvelgrm).
-
-<!--The removal of long edges is a critical step to create a smoother representation  by iteratively eliminating hexagons with excessive distances between centroids.
-This process eliminates outliers and noise while preserving essential local relationships within the data.
-To achieve this, distances between vertices are sorted, and unique distance values are extracted.
-By setting a threshold based on the largest difference between consecutive distance values, long edges are identified and removed.
-This refinement step contributes to enhancing the quality of the triangular mesh, ensuring a more accurate representation of the data structure.-->
-
-```{r}
-#| warning: false
-#| echo: false
-
-## Import data
-
-UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap.rds")
-
-## UMAP
-
-num_bins_umap_s_curve <- 6
-shape_val_umap_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
-                                                          x = UMAP1, y = UMAP2) ## 1.259938
-## To extract bin centroids
-hexbin_data_object_umap_s_curve <- extract_hexbin_centroids(nldr_df = UMAP_s_curve, num_bins = num_bins_umap_s_curve, shape_val = shape_val_umap_s_curve, x = UMAP1, y = UMAP2)
-
-df_bin_centroids_umap_s_curve <- hexbin_data_object_umap_s_curve$hexdf_data
-
-UMAP_data_with_hb_id_s_curve <- UMAP_s_curve |>
-  dplyr::mutate(hb_id = hexbin_data_object_umap_s_curve$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_umap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), UMAP_data_with_hb_id_s_curve)
-
-## Averaged on high-D
-df_bin_umap_s_curve <- avg_highD_data(.data = df_all_umap_s_curve)
-
-## Triangulate bin centroids
-tr1_object_umap_s_curve <- triangulate_bin_centroids(df_bin_centroids_umap_s_curve, x, y)
-tr_from_to_df_umap_s_curve <- generate_edge_info(triangular_object = tr1_object_umap_s_curve)
-
-# ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
-#   geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
-#   geom_point(size = 2, colour = "#33a02c") +
-#   coord_equal()
-
-## Compute 2D distances
-distance_umap_s_curve <- cal_2D_dist(.data = tr_from_to_df_umap_s_curve)
-
-## To find the benchmark value
-benchmark_umap_s_curve <- find_benchmark_value(.data = distance_umap_s_curve, distance_col = distance)
-
-# colour_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve,
-#                   triangular_object = tr1_object_umap_s_curve, distance_col = distance)
-
-trimesh_umap_s_curve <- ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
-  geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
-  geom_point(size = 2, colour = "#33a02c") +
-  coord_equal()
-
-trimesh_umap_s_curve <- trimesh_umap_s_curve +
-  #ggtitle("(a)") +
-  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  theme_light() +
-  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
-  ) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
-# theme(axis.text = element_text(size = 5),
-#       axis.title = element_text(size = 7))
-
-trimesh_gr_umap_s_curve <- colour_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve, triangular_object = tr1_object_umap_s_curve, distance_col = distance)
-
-trimesh_gr_umap_s_curve <- trimesh_gr_umap_s_curve +
-  geom_point(size = 2, colour = "#33a02c") +
-  #ggtitle("(b)") +
-  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  theme_light() +
-  #coord_equal() +
-  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
-  ) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-# ### Define type column
-# df <- df_all_umap_s_curve |>
-#   dplyr::select(tidyselect::starts_with("x")) |>
-#   dplyr::mutate(type = "data") ## original dataset
-# 
-# df_b <- df_bin_umap_s_curve |>
-#   dplyr::filter(hb_id %in% df_bin_centroids_umap_s_curve$hexID) |>
-#   dplyr::select(-hb_id) |>
-#   dplyr::mutate(type = "model") ## Data with summarized mean
-# 
-# df_exe <- dplyr::bind_rows(df_b, df)
-# 
-# distance_df_small_edges <- distance_umap_s_curve %>%
-#   dplyr::filter(distance < benchmark_umap_s_curve)
-# 
-# distance_df_long_edges <- distance_umap_s_curve %>%
-#   dplyr::filter(distance >= benchmark_umap_s_curve)
-# 
-# distance_df_edges <- dplyr::bind_rows(distance_df_small_edges, distance_df_long_edges)
-# 
-# langevitour::langevitour(df_exe[1:(length(df_exe)-1)], lineFrom = distance_df_edges$from , lineTo = distance_df_edges$to, group = df_exe$type, pointSize = 3, levelColors = c("#6a3d9a", "#33a02c"), lineColors = append(rep("black", length(distance_df_small_edges$from)), rep("red", length(distance_df_long_edges$from))))
-#
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-#| label: fig-modelScurvermlgimp
-#| fig-cap: "What does the 2D model, constructed using UMAP (applied to the S-curve dataset), look like in high dimensions with long edges?
-(<https://youtu.be/pSL_G_n7-fM>) (a) 2D model without removing the long edges, and (b) 2D model with long edges coloured by red.
-The default benchmark value is used to identify long edges."
-##| out-width: 100%
-
-trimesh_umap_s_curve + trimesh_gr_umap_s_curve +
-  #plot_annotation(tag_levels = 'a') +
-  plot_layout(ncol = 2) &
-  theme(legend.position='none', plot.tag = element_text(size = 8))
-```
-
-```{r}
-#| echo: false
-#| warning: false
-
-error_df_lg_1 <- read_rds("data/s_curve/s_curve_summary_lg_tsne.rds")
-error_df_lg_2 <- read_rds("data/s_curve/s_curve_summary_lg_umap.rds")
-error_df_lg_3 <- read_rds("data/s_curve/s_curve_summary_lg_phate.rds")
-error_df_lg_4 <- read_rds("data/s_curve/s_curve_summary_lg_trimap.rds")
-error_df_lg_5 <- read_rds("data/s_curve/s_curve_summary_lg_pacmap.rds")
-
-error_df_lg <- dplyr::bind_rows(error_df_lg_1, error_df_lg_2, error_df_lg_3, error_df_lg_4, error_df_lg_5)
-
-error_df_lg$method <- factor(error_df_lg$method, levels = c("tSNE", "UMAP", "PHATE", "TriMAP", "PaCMAP"))
-
-## To draw with AIC
-aic_plot_lg <- ggplot(error_df_lg |> dplyr::filter(method == "UMAP"), aes(x = benchmark_rm_lg,
-                                                                                 y = total_error,
-                                                                                 color = method
-)) +
-  geom_point() +
-  geom_line() +
-  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("AIC") +
-  xlab("Benchmark value")
-## Effective number of bins along x-axis
-
-mse_plot_lg <- ggplot(error_df_lg |> dplyr::filter(method == "UMAP"), aes(x = benchmark_rm_lg,
-                                       y = total_mse,
-                                       color = method
-)) +
-  geom_point() +
-  geom_line() +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("MSE") +
-  xlab("Benchmark value")
-
-```
-
-```{r}
-#| echo: false
-#| warning: false
-#| fig-cap: "Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset with different benchmark values to remove the long edges.
-What is the effective benchmark value to remove the long edges?"
+  xlab("benchmark value")
+
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset with different benchmark values to remove the long edges.
+What is the effective benchmark value to remove the long edges?
+$1.347$ is the chosen benchmark value to remove long edges.
+The value is selected through a visual inspection of the triangular mesh is conducted to verify whether it successfully eliminates long edges."
 #| label: fig-diagnosticpltScurvelgrm
@@
 $$
-\text{MSE} = \sum_{j = 1}^{n} \frac{\sqrt{\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2}}{n}
+\text{MSE} = \sum_{j = 1}^{n} \frac{\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2}{n}
 $$ {#eq-equation9}
@@
   ylab("AIC") +
-  xlab("Total number of bins")
+  xlab("total number of bins") + 
+  facet_wrap(vars(method), ncol = 5)
 ## Effective number of bins along x-axis
@@
   ylab("MSE") +
-  xlab("Total number of bins")
+  xlab("total number of bins") + 
+  facet_wrap(vars(method), ncol = 5)
 
@@
   plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', ncol = 2) &
+  plot_layout(guides='collect', nrow = 2) &
   theme(legend.position='bottom', plot.tag = element_text(size = 8))

### 2024-02-01 paper.tex
Update the repo

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, boxrule=0pt, frame hidden, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, enhanced, sharp corners, breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
 
@@
 \caption{\label{fig-binsize}Hexbin plots from different number of bins
-for the \textbf{UMAP} projections of \textbf{S-curve} data: (a) b = 32
-(4, 8), s = 1.643542, (b) b = 264 (12, 22), s = 1.643542, and (c) b =
-840 (21, 40), s = 1.643542. The hexbins are colored based on the density
-of points, with darker colors indicating higher point density and yellow
-color representing lower point density within each bin. What is the
-number of bins that would be effective in representing low-dimensional
-data?}
+for the \textbf{UMAP} projections of \textbf{S-curve} training data: (a)
+b = 32 (4, 8), s = 1.643542, (b) b = 264 (12, 22), s = 1.643542, and (c)
+b = 840 (21, 40), s = 1.643542. The hexbins are colored based on the
+density of points, with darker colors indicating higher point density
+and yellow color representing lower point density within each bin. What
+is the number of bins that would be effective in representing
+low-dimensional data?}
 
@@
 \caption{\label{fig-diagnosticpltScurve}Goodness of fit statistics from
-different NLDR techniques applied to training S-curve dataset. What is
-the most effective NLDR technique to represent the original data in a 2D
-space? What is the effective number of bins in each NLDR technique to
-create a 2D model?}
+UMAP applied to training S-curve dataset. What is the effective number
+of bins in each NLDR technique to create a 2D model? The MSE plot have a
+steep slope at the beginning, indicating that a smaller number of bins
+causes a larger amount of error. Then, the slope gradually declines or
+level off, indicating that a higher number of bins generates a smaller
+error. Using the elbow method, when the total number of bins is set to
+144, the slope of the Mean Squared Error (MSE) plot experiences a sudden
+and noticeable change, resembling an elbow-like shape. This point
+indicates that adding less bins does not enough to capture the data
+structure.}
 
@@
 \caption{\label{fig-diagnosticpltScurvelwd}Goodness of fit statistics
-from different NLDR techniques applied to training S-curve dataset with
-different benchmark values to remove the low-density heaxgons. What is
-the effective benchmark value to remove the low-density heaxgons?}
+from UMAP applied to training S-curve dataset with different benchmark
+values to remove the low-density heaxgons. What is the effective
+benchmark value to remove the low-density heaxgons? The MSE plot have a
+steep slope at the beginning, indicating that a smaller benchmark causes
+a small amount of error. Then, the slope gradually increases or level
+up, indicating that a higher number of benchmark values generates a
+higher error. Using the reverse elbow method, when the benchmark value
+is set to 0.2, the slope of the Mean Squared Error (MSE) plot
+experiences a sudden and noticeable change, resembling an elbow-like
+shape. This point indicates that higher benchmark values remove the
+necessary bins as well which lead to distruct the 2D structure.}
 
@@
 \caption{\label{fig-diagnosticpltScurvelgrm}Goodness of fit statistics
-from different NLDR techniques applied to training S-curve dataset with
-different benchmark values to remove the long edges. What is the
-effective benchmark value to remove the long edges?}
+from UMAP applied to training S-curve dataset with different benchmark
+values to remove the long edges. What is the effective benchmark value
+to remove the long edges? \(1.347\) is the chosen benchmark value to
+remove long edges. The value is selected through a visual inspection of
+the triangular mesh is conducted to verify whether it successfully
+eliminates long edges.}
 
@@
 \begin{equation}\protect\hypertarget{eq-equation9}{}{
-\text{MSE} = \sum_{j = 1}^{n} \frac{\sqrt{\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2}}{n}
+\text{MSE} = \sum_{j = 1}^{n} \frac{\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2}{n}
 }\label{eq-equation9}\end{equation}

### 2024-02-01 applications/pbmc_3K_Festem/1_obtain_pca_umap_param5.R

@@ -1,0 +1,248 @@
+library(dplyr)
+library(rsample)
+library(readr)
+library(umap)
+
+set.seed(20240110)
+
+source("quollr_code.R", local = TRUE)
+source("nldr_code.R", local = TRUE)
+
+#save(gene.list,label.list,umap.list,plots.list,file = "./results/pbmc3k_clustering_UMAP.RData")
+
+data_pca <- read_rds(file = "data/pbmc/pbmc_3k_festem/pbmc_pca_50.rds")
+
+data_pca <- data_pca[,1:15]
+
+training_data_pbmc <- data_pca |>
+  mutate(ID = row_number())
+
+## UMAP
+
+UMAP_fit <- umap(training_data_pbmc |> dplyr::select(-ID), n_neighbors = 5, n_components =  2, metric = "cosine", min_dist = 0.1)
+
+UMAP_pbmc <- UMAP_fit$layout |>
+  as.data.frame()
+names(UMAP_pbmc)[1:(ncol(UMAP_pbmc))] <- paste0(rep("UMAP",(ncol(UMAP_pbmc))), 1:(ncol(UMAP_pbmc)))
+
+UMAP_pbmc <- UMAP_pbmc |>
+  mutate(ID = training_data_pbmc$ID)
+
+UMAP_pbmc_old <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_umap.rds")
+
+UMAP_pbmc <- UMAP_pbmc |>
+  mutate(cell_label = UMAP_pbmc_old$cell_label)
+
+## Run only once
+write_rds(UMAP_pbmc, file = "data/pbmc/pbmc_3k_festem/pbmc_umap_5_min_dist_0.1.rds")
+
+plot_UMAP_2D(UMAP_pbmc) + #ggtitle("(b)") +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
[…]

### 2024-02-01 scripts/gen_hex_grid_with_diff_bins_S_curve.R

@@ -198,3 +198,66 @@ hex_full_count_df_loop <- dplyr::bind_cols(hex_grid_loop, full_grid_with_hexbin_
 ## Run only once
 write_rds(hex_full_count_df_loop, file = "data/s_curve/s_curve_hex_20.rds")
 
+
+####num_bins_x = 8####################
+
+num_bins <- 8
+shape_val <- calculate_effective_shape_value(.data = UMAP_data, x = UMAP1, y = UMAP2)
+
+hexbin_data_object_loop <- extract_hexbin_centroids(UMAP_data, num_bins, shape_val)
+
+df_bin_centroids_loop <- hexbin_data_object_loop$hexdf_data
+
+## Data set with all possible centroids in the hexagonal grid
+
+full_centroid_df_loop <- generate_full_grid_centroids(df_bin_centroids_loop)
+
+## To map hexID to hexbin centroids in the full grid
+
+vec1 <- stats::setNames(rep("", 2), c("x", "y"))  ## Define column names
+
+full_grid_with_hexbin_id_loop <- dplyr::bind_rows(vec1)[0, ]
+full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for(i in 1:length(sort(unique(full_centroid_df_loop$y)))){
+
+  ## Filter the data set with specific y value
+  specific_y_val_df_loop <- full_centroid_df_loop |>
+    dplyr::filter(y == sort(unique(full_centroid_df_loop$y))[i])
+
+  ordered_x_df_loop <- specific_y_val_df_loop |>
+    dplyr::arrange(x)
+
+  full_grid_with_hexbin_id_loop <- dplyr::bind_rows(full_grid_with_hexbin_id_loop, ordered_x_df_loop)
+
+}
+
+
+full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
+  dplyr::mutate(hexID = row_number())
+
+full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
+  dplyr::rename("c_x" = "x",
+                "c_y" = "y")
+
+full_grid_with_hexbin_id_loop <- dplyr::full_join(full_grid_with_hexbin_id_loop, df_bin_centroids_loop, by = c("hexID" = "hexID")) |>
+  dplyr::select(-c(x, y))
+
[…]

### 2024-02-01 scripts/gen_summary_umap_lg_rm_n.R

@@ -12,7 +12,7 @@ training_data_s_curve <- read_rds("data/s_curve/s_curve_training.rds")
 
 UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap.rds")
 
-num_bins_s_curve <- 6
+num_bins_s_curve <- 8
 
 shape_value_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
                                                       x = UMAP1, y = UMAP2) ## 1.259938
@@ -43,7 +43,8 @@ distance_s_curve <- cal_2D_dist(.data = tr_from_to_df_s_curve)
 ## To find the benchmark value
 benchmark <- find_benchmark_value(.data = distance_s_curve, distance_col = distance)
 
-benchmark_dist_vec <- distance_s_curve$distance |> round(3) |> unique() |> sort()
+more_val_vec <- c(0, 0.475, 1)
+benchmark_dist_vec <- append(more_val_vec, distance_s_curve$distance |> round(3) |> unique() |> sort())
 #benchmark_dist_vec <- seq(min(distance_s_curve$distance |> round(3) |> unique()), max(distance_s_curve$distance |> round(3) |> unique()), 1)
 
 vec <- stats::setNames(rep("", 2), c("benchmark_rm_lg", "total_error"))  ## Define column names

### 2024-02-01 paper.qmd
Changed the table for a reproducable one

@@
 library(grid)
-# library(xtable)
-# 
-# xtable2kable <- function(x) {
-#   out <- capture.output(print(x, table.placement = NULL))[-(1:2)]
-#   out <- paste(out, collapse = "\n")
-#   structure(out, format = "latex", class = "knitr_kable")
-# }
+library(xtable)
+
+xtable2kable <- function(x) {
+  out <- capture.output(print(x, table.placement = NULL))[-(1:2)]
+  out <- paste(out, collapse = "\n")
+  structure(out, format = "latex", class = "knitr_kable")
+}
 
@@
 
-```{r}
-#| warning: false
-#| echo: false
-
-# error_df <- read_csv("data/pbmc/pbmc_3k_festem/error_pbmc_umap.csv")
-# 
-# error_df |>
-#   knitr::kable() 
-```
-
-| n_neighbors | min_dist | metric     | error                 |
-|-------------|----------|------------|-----------------------|
-| **30**      | **0.3**  | **cosine** | **38634.66719286174** |
-| 5           | 0.99     | euclidean  | 35938.98594054407     |
-| 5           | 0.99     | cosine     | 35597.01981261912     |
-| 15          | 0.99     | euclidean  | 35534.75778935427     |
-| 30          | 0.99     | euclidean  | 35664.68763936665     |
-| 1022        | 0.99     | cosine     | 37022.0125249776      |
-| 1022        | 0.3      | cosine     | 39746.347504639125    |
-| 800         | 0.3      | euclidean  | 38385.25485420897     |
-| 100         | 0.99     | euclidean  | 36528.57424747017     |
-| 230         | 0.99     | cosine     | 37864.26802578186     |
-| 354         | 0.99     | euclidean  | 37530.86598361975     |
+<!--https://dicook.github.io/mulgar_book/7-spin-and-brush.html#fig-penguins-bs-detourr-->
+
+```{r}
+#| warning: false
+#| echo: false
+
+error_df <- read_csv("data/pbmc/pbmc_3k_festem/error_pbmc_umap.csv", col_names = FALSE)
+
+error_df <- error_df |>
+  mutate(X4 = round(X4/100, 2)) |>
+  arrange(X4)
+
+colnames(error_df) <- c("n_neighbors", "min_dist", "metric", "error (x 100)")
+
+xtable(error_df, caption = "errors for different parameter combinations", label = "tbl-errordiffcomb") |>
+  xtable2kable() |>
+  kable_styling(latex_options = c("striped", "HOLD_position"), position = "center", full_width = FALSE) |>
+  row_spec(0, bold = TRUE) |>
+  row_spec(3, bold=T, color = "red")
+```
 
@@
   #annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
 ```
@@
 
-#plot_list2_pbmc + 
-trimesh_removed_pbmc_umap + plot_list2_pbmc_error +
-plot_layout(guides='collect', ncol=3) &
-  theme(legend.position='bottom')
-
-#  plot_list1_pbmc + plot_list3_pbmc + plot_list4_pbmc + plot_list5_pbmc +
-#   plot_layout(ncol=5) #| out-width: 100%
-
-# 2D layouts from different NLDR techniques applied for the training PBMC dataset: (a) tSNE (perplexity = 30), (b) UMAP (n_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2).
-Is there a best representation of the original data or are they all providing  equivalent information?
-```
-
-<!--tSNE parm 1-->
-
-```{r}
-#| warning: false
-#| echo: false
-
-## Import data
-#df_2 <- read_rds("data/pbmc/pbmc.rds")
-training_data_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_pca_50.rds")
-training_data_pbmc <- training_data_pbmc[, 1:15] |>
-  mutate(ID = 1:NROW(training_data_pbmc))
-#test_1 <- read_rds("data/pbmc/pbmc_test.rds")
-
-tSNE_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_tsne_5.rds")
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-num_bins_pbmc <- calculate_effective_x_bins(.data = tSNE_pbmc, x = tSNE1,
-                                            cell_area = 1) ##23
-
-num_bins_pbmc <- 23
-
-shape_value_pbmc <- calculate_effective_shape_value(.data = tSNE_pbmc,
-                                                    x = tSNE1, y = tSNE2) ## 0.8772751
-shape_value_pbmc <- 0.8772751
-
-## To extract bin centroids
-hexbin_data_object_pbmc <- extract_hexbin_centroids(nldr_df = tSNE_pbmc,
-                                                    num_bins = num_bins_pbmc,
-                                                    shape_val = shape_value_pbmc, x = tSNE1, y = tSNE2)
-
-df_bin_centroids_pbmc <- hexbin_data_object_pbmc$hexdf_data
-
-tSNE_pbmc_with_hb_id <- tSNE_pbmc |>
-  dplyr::mutate(hb_id = hexbin_data_object_pbmc$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), tSNE_pbmc_with_hb_id)
-
-## Averaged on high-D
-df_bin_pbmc <- avg_highD_data(.data = df_all_pbmc, column_start_text = "PC")
-
-## Triangulate bin centroids
-tr1_object_pbmc <- triangulate_bin_centroids(df_bin_centroids_pbmc, x, y)
-tr_from_to_df_pbmc <- generate_edge_info(triangular_object = tr1_object_pbmc)
-
-## Compute 2D distances
-distance_pbmc <- cal_2D_dist(.data = tr_from_to_df_pbmc)
-
-## To find the benchmark value
-benchmark_pbmc <- find_benchmark_value(.data = distance_pbmc, distance_col = distance)
-
-trimesh_removed_pbmc_tsne <- remove_long_edges(.data = distance_pbmc, benchmark_value = benchmark_pbmc,
-                                               triangular_object = tr1_object_pbmc, distance_col = distance)
-
-trimesh_removed_pbmc_tsne <- trimesh_removed_pbmc_tsne +
-  geom_point(size = 2, colour = "#33a02c") +
-  #ggtitle("(b)") +
-  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  theme_light() +
-  #coord_equal() +
-  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
-  )
-
-df_bin_train <- df_bin_pbmc
-names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])
-
-error_df <- df_all_pbmc |>
-  dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates
-
-# prediction_df_join <- prediction_df_join |>
-#   dplyr::left_join(data, by = c("ID" = "ID")) ## Map high-D data
-
-for (i in 1:(NCOL(df_bin_train) - 1)) {
-
-  error_df[ , paste0("abs_residual_", "PC_", i)] <- abs(error_df[ , paste0("PC_", i)] - error_df[ , paste0("avg_", "PC_", i)])
-
-}
-
-error_df <- error_df |>
-  dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "PC_")))))
-
-# library(ggbeeswarm)
-# error_df$group <- "1"
-# ggplot(error_df, aes(x = group, y = total)) +
-#   geom_quasirandom()+
-#   ylim(0, max(unlist(error_df$total))+ 0.5) + coord_flip()
-
-### The minimum error is 0 and the maximum is 42.17439
-### There is lot of points with error 0,
-
-error_df <- error_df |>
-  mutate(type = if_else(total <= 2, "error 2 or less",
-                        if_else(total <= 10, "error 2-10",
-                                if_else(total <= 15, "error 10-15",
-                                        if_else(total <= 20, "error 15-20",
-                                                if_else(total <= 25, "error 20-25",
-                                                        if_else(total <= 30, "error 25-30",
-                                                                if_else(total <= 35, "error 30-35", "error greter than 35"))))))))
-
-plot_list2_pbmc_error <- error_df |>
-    mutate(type = factor(type , levels = c("error 2 or less", "error 2-10", "error 10-15", "error 15-20", "error 20-25", "error 25-30", "error 30-35", "error greter than 35"))) |>
-    ggplot(aes(x = tSNE1,
-               y = tSNE2, color = type))+
-    geom_point(alpha=0.5, size = 0.1) +
-    coord_equal() +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#fee0d2", "#fcbba1",
-                                         "#fc9272", "#fb6a4a", "#ef3b2c",
-                                         "#cb181d", "#a50f15", "#99000d"))
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "(a) 2D layout from tSNE (perplexity = 5) applied for the PBMC3k dataset. Is this a best representation of the original data?, (b)  Model in the 2D space with tSNE (<>)"
-#| label: fig-nldervisPBMCtSNE1
-#| out-width: 100%
-#| fig-pos: h
-
-#plot_list2_pbmc + 
-trimesh_removed_pbmc_tsne + plot_list2_pbmc_error +
-plot_layout(guides='collect', ncol=3) &
-  theme(legend.position='bottom')
-
-#  plot_list1_pbmc + plot_list3_pbmc + plot_list4_pbmc + plot_list5_pbmc +
-#   plot_layout(ncol=5) #| out-width: 100%
-
-# 2D layouts from different NLDR techniques applied for the training PBMC dataset: (a) tSNE (perplexity = 30), (b) UMAP (n_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2).
-Is there a best representation of the original data or are they all providing  equivalent information?
-```
-
-<!--tSNE parm 2-->
-
-```{r}
-#| warning: false
-#| echo: false
-
-## Import data
-#df_2 <- read_rds("data/pbmc/pbmc.rds")
-training_data_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_pca_50.rds")
-training_data_pbmc <- training_data_pbmc[, 1:15] |>
-  mutate(ID = 1:NROW(training_data_pbmc))
-#test_1 <- read_rds("data/pbmc/pbmc_test.rds")
-
-tSNE_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_tsne_2.rds")
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-num_bins_pbmc <- calculate_effective_x_bins(.data = tSNE_pbmc, x = tSNE1,
-                                            cell_area = 1) ##23
-
-num_bins_pbmc <- 23
-
-shape_value_pbmc <- calculate_effective_shape_value(.data = tSNE_pbmc,
-                                                    x = tSNE1, y = tSNE2) ## 0.8772751
-shape_value_pbmc <- 0.8772751
-
-## To extract bin centroids
-hexbin_data_object_pbmc <- extract_hexbin_centroids(nldr_df = tSNE_pbmc,
-                                                    num_bins = num_bins_pbmc,
-                                                    shape_val = shape_value_pbmc, x = tSNE1, y = tSNE2)
-
-df_bin_centroids_pbmc <- hexbin_data_object_pbmc$hexdf_data
-
-tSNE_pbmc_with_hb_id <- tSNE_pbmc |>
-  dplyr::mutate(hb_id = hexbin_data_object_pbmc$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), tSNE_pbmc_with_hb_id)
-
-## Averaged on high-D
-df_bin_pbmc <- avg_highD_data(.data = df_all_pbmc, column_start_text = "PC")
-
-## Triangulate bin centroids
-tr1_object_pbmc <- triangulate_bin_centroids(df_bin_centroids_pbmc, x, y)
-tr_from_to_df_pbmc <- generate_edge_info(triangular_object = tr1_object_pbmc)
-
-## Compute 2D distances
-distance_pbmc <- cal_2D_dist(.data = tr_from_to_df_pbmc)
-
-## To find the benchmark value
-benchmark_pbmc <- find_benchmark_value(.data = distance_pbmc, distance_col = distance)
-
-trimesh_removed_pbmc_tsne <- remove_long_edges(.data = distance_pbmc, benchmark_value = benchmark_pbmc,
-                                               triangular_object = tr1_object_pbmc, distance_col = distance)
-
-trimesh_removed_pbmc_tsne <- trimesh_removed_pbmc_tsne +
-  geom_point(size = 2, colour = "#33a02c") +
-  #ggtitle("(b)") +
-  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  theme_light() +
-  #coord_equal() +
-  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
-  )
-
-df_bin_train <- df_bin_pbmc
-names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])
-
-error_df <- df_all_pbmc |>
-  dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates
-
-# prediction_df_join <- prediction_df_join |>
-#   dplyr::left_join(data, by = c("ID" = "ID")) ## Map high-D data
-
-for (i in 1:(NCOL(df_bin_train) - 1)) {
-
-  error_df[ , paste0("abs_residual_", "PC_", i)] <- abs(error_df[ , paste0("PC_", i)] - error_df[ , paste0("avg_", "PC_", i)])
-
-}
-
-error_df <- error_df |>
-  dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "PC_")))))
-
-# library(ggbeeswarm)
-# error_df$group <- "1"
-# ggplot(error_df, aes(x = group, y = total)) +
-#   geom_quasirandom()+
-#   ylim(0, max(unlist(error_df$total))+ 0.5) + coord_flip()
-
-### The minimum error is 0 and the maximum is 42.17439
-### There is lot of points with error 0,
-
-error_df <- error_df |>
-  mutate(type = if_else(total <= 2, "error 2 or less",
-                        if_else(total <= 10, "error 2-10",
-                                if_else(total <= 15, "error 10-15",
-                                        if_else(total <= 20, "error 15-20",
-                                                if_else(total <= 25, "error 20-25",
-                                                        if_else(total <= 30, "error 25-30",
-                                                                if_else(total <= 35, "error 30-35", "error greter than 35"))))))))
-
-plot_list2_pbmc_error <- error_df |>
-    mutate(type = factor(type , levels = c("error 2 or less", "error 2-10", "error 10-15", "error 15-20", "error 20-25", "error 25-30", "error 30-35", "error greter than 35"))) |>
-    ggplot(aes(x = tSNE1,
-               y = tSNE2, color = type))+
-    geom_point(alpha=0.5, size = 0.1) +
-    coord_equal() +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#fee0d2", "#fcbba1",
-                                         "#fc9272", "#fb6a4a", "#ef3b2c",
-                                         "#cb181d", "#a50f15", "#99000d"))
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "(a) 2D layout from tSNE (perplexity = 2) applied for the PBMC3k dataset. Is this a best representation of the original data?, (b)  Model in the 2D space with tSNE (<>)"
-#| label: fig-nldervisPBMCtSNE2
-#| out-width: 100%
-#| fig-pos: h
-
-#plot_list2_pbmc + 
-trimesh_removed_pbmc_tsne + plot_list2_pbmc_error +
+plot_list2_pbmc + trimesh_removed_pbmc_umap + plot_list2_pbmc_error +
 plot_layout(guides='collect', ncol=3) &

### 2024-02-01 paper.tex
Changed the table for a reproducable one

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, enhanced, sharp corners, breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced, frame hidden, breakable, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 
-\begin{longtable}[]{@{}llll@{}}
-\toprule\noalign{}
-n\_neighbors & min\_dist & metric & error \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-\textbf{30} & \textbf{0.3} & \textbf{cosine} &
-\textbf{38634.66719286174} \\
-5 & 0.99 & euclidean & 35938.98594054407 \\
-5 & 0.99 & cosine & 35597.01981261912 \\
-15 & 0.99 & euclidean & 35534.75778935427 \\
-30 & 0.99 & euclidean & 35664.68763936665 \\
-1022 & 0.99 & cosine & 37022.0125249776 \\
-1022 & 0.3 & cosine & 39746.347504639125 \\
-800 & 0.3 & euclidean & 38385.25485420897 \\
-100 & 0.99 & euclidean & 36528.57424747017 \\
-230 & 0.99 & cosine & 37864.26802578186 \\
-354 & 0.99 & euclidean & 37530.86598361975 \\
-\end{longtable}
+\begin{table}[H]
+\centering
+\begin{tabular}{rrrlr}
+  \hline
+\textbf{} & \textbf{n\_neighbors} & \textbf{min\_dist} & \textbf{metric} & \textbf{error (x 100) }\\ 
+  \hline
+\cellcolor{gray!6}{1} & \cellcolor{gray!6}{5.00} & \cellcolor{gray!6}{0.99} & \cellcolor{gray!6}{cosine} & \cellcolor{gray!6}{355.97 }\\ 
+  2 & 230.00 & 0.99 & cosine & 378.64 \\ 
+\textcolor{red}{\textbf{\cellcolor{gray!6}{  3}}} & \textcolor{red}{\textbf{\cellcolor{gray!6}{30.00}}} & \textcolor{red}{\textbf{\cellcolor{gray!6}{0.30}}} & \textcolor{red}{\textbf{\cellcolor{gray!6}{cosine}}} & \textcolor{red}{\textbf{\cellcolor{gray!6}{386.35 }}}\\
+   \hline
+\end{tabular}
+\caption{errors for different parameter combinations} 
+\label{tbl-errordiffcomb}
+\end{table}
 
@@
 Model in the 2D space with UMAP (\textless\textgreater)}
-
-\end{figure}
-
-\begin{figure}[h]
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervisPBMCtSNE1-1.pdf}
-
-}
-
-\caption{\label{fig-nldervisPBMCtSNE1}(a) 2D layout from tSNE
-(perplexity = 5) applied for the PBMC3k dataset. Is this a best
-representation of the original data?, (b) Model in the 2D space with
-tSNE (\textless\textgreater)}
-
-\end{figure}
-
-\begin{figure}[h]
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervisPBMCtSNE2-1.pdf}
-
-}
-
-\caption{\label{fig-nldervisPBMCtSNE2}(a) 2D layout from tSNE
-(perplexity = 2) applied for the PBMC3k dataset. Is this a best
-representation of the original data?, (b) Model in the 2D space with
-tSNE (\textless\textgreater)}
 

### 2024-02-01 paper.qmd
Add rsults for PBMC with fixed non-empty bins

@@
 error_df <- error_df |>
-  mutate(X4 = round(X4/100, 2)) |>
+  mutate(X4 = round(X4/100, 2),
+         X1 = as.integer(X1)) |>
   arrange(X4)
@@
   row_spec(0, bold = TRUE) |>
-  row_spec(3, bold=T, color = "red")
+  row_spec(7, bold=T, color = "red")
 ```
@@
 
+num_non_empty_bins_pbmc <- df_bin_centroids_pbmc$hexID |> length()
+
 UMAP_pbmc_with_hb_id <- UMAP_pbmc |>
@@
 #| echo: false
-num_bins_pbmc <- calculate_effective_x_bins(.data = UMAP_pbmc, x = UMAP1,
-                                              cell_area = 1) ##23
-num_bins_pbmc <- 23
+# num_bins_pbmc <- calculate_effective_x_bins(.data = UMAP_pbmc, x = UMAP1,
+#                                               cell_area = 1) ##23
 
@@
                                                       x = UMAP1, y = UMAP2) ## 0.8772751
-shape_value_pbmc <- 0.8772751
+
+num_bins_pbmc <- find_non_empty_bins(nldr_df = UMAP_pbmc, x = "UMAP1", y = "UMAP2", shape_val = shape_value_pbmc, non_empty_bins = num_non_empty_bins_pbmc)
 
@@
 benchmark_pbmc <- find_benchmark_value(.data = distance_pbmc, distance_col = distance)
-#benchmark_pbmc <- 1.035666
+#benchmark_pbmc <- 0.5517801
 
@@
 trimesh_removed_pbmc_umap <- trimesh_removed_pbmc_umap +
-  geom_point(size = 1, colour = "#33a02c") +
+  geom_point(size = 0.01, colour = "#33a02c") +
   #ggtitle("(b)") +

### 2024-02-01 paper.tex
Add rsults for PBMC with fixed non-empty bins

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced, frame hidden, breakable, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, interior hidden, boxrule=0pt, frame hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable]}{\end{tcolorbox}}\fi
 
@@
   \hline
-\cellcolor{gray!6}{1} & \cellcolor{gray!6}{5.00} & \cellcolor{gray!6}{0.99} & \cellcolor{gray!6}{cosine} & \cellcolor{gray!6}{355.97 }\\ 
-  2 & 230.00 & 0.99 & cosine & 378.64 \\ 
-\textcolor{red}{\textbf{\cellcolor{gray!6}{  3}}} & \textcolor{red}{\textbf{\cellcolor{gray!6}{30.00}}} & \textcolor{red}{\textbf{\cellcolor{gray!6}{0.30}}} & \textcolor{red}{\textbf{\cellcolor{gray!6}{cosine}}} & \textcolor{red}{\textbf{\cellcolor{gray!6}{386.35 }}}\\
+\cellcolor{gray!6}{1} & \cellcolor{gray!6}{  5} & \cellcolor{gray!6}{0.99} & \cellcolor{gray!6}{cosine} & \cellcolor{gray!6}{262.18 }\\ 
+  2 &  15 & 0.99 & cosine & 279.49 \\ 
+\cellcolor{gray!6}{  3} & \cellcolor{gray!6}{ 84} & \cellcolor{gray!6}{0.99} & \cellcolor{gray!6}{cosine} & \cellcolor{gray!6}{295.64 }\\ 
+  4 &   5 & 0.30 & cosine & 303.90 \\ 
+\cellcolor{gray!6}{  5} & \cellcolor{gray!6}{ 15} & \cellcolor{gray!6}{0.30} & \cellcolor{gray!6}{cosine} & \cellcolor{gray!6}{308.97 }\\ 
+  6 &  84 & 0.30 & cosine & 319.59 \\ 
+\textcolor{red}{\textbf{\cellcolor{gray!6}{  7}}} & \textcolor{red}{\textbf{\cellcolor{gray!6}{ 30}}} & \textcolor{red}{\textbf{\cellcolor{gray!6}{0.30}}} & \textcolor{red}{\textbf{\cellcolor{gray!6}{cosine}}} & \textcolor{red}{\textbf{\cellcolor{gray!6}{386.35 }}}\\
    \hline

### 2024-02-01 paper.qmd
Changed the point size of trimesh

@@
 #tSNE_pbmc <- read_rds("data/pbmc/pbmc_tsne_30.rds")
-UMAP_pbmc <- read_rds("data/pbmc/pbmc_umap_15_min_dist_0.99_metric_euclidean.rds")
+UMAP_pbmc <- read_rds("data/pbmc/pbmc_umap_15_min_dist_0.99_metric_cosine.rds")
 # PHATE_pbmc <- read_rds("data/pbmc/pbmc_phate.rds")
@@
 #| echo: false
-num_bins_pbmc <- calculate_effective_x_bins(.data = UMAP_pbmc, x = UMAP1,
-                                              cell_area = 1) ##23
-
-num_bins_pbmc <- 23
-
 shape_value_pbmc <- calculate_effective_shape_value(.data = UMAP_pbmc,
                                                       x = UMAP1, y = UMAP2) ## 0.8772751
-shape_value_pbmc <- 0.8772751
+
+num_bins_pbmc <- find_non_empty_bins(nldr_df = UMAP_pbmc, x = "UMAP1", y = "UMAP2", shape_val = shape_value_pbmc, non_empty_bins = num_non_empty_bins_pbmc)
 
@@
 trimesh_removed_pbmc_umap <- trimesh_removed_pbmc_umap +
-  geom_point(size = 1, colour = "#33a02c") +
+  geom_point(size = 0.005, colour = "#33a02c") +
   #ggtitle("(b)") +
@@
 #| echo: false
-#| fig-cap: "(a) 2D layout from UMAP (n_neighbors = 15, min_dist = 0.99, metric = euclidean) applied for the PBMC3k dataset.
+#| fig-cap: "(a) 2D layout from UMAP (n_neighbors = 15, min_dist = 0.99, metric = cosine) applied for the PBMC3k dataset.
 Is this a best representation of the original data?, (b)  Model in the 2D space with UMAP (<>)"
@@
 
-#plot_list2_pbmc + 
-trimesh_removed_pbmc_umap + plot_list2_pbmc_error +
+plot_list2_pbmc + trimesh_removed_pbmc_umap + plot_list2_pbmc_error +
 plot_layout(guides='collect', ncol=3) &
@@
 trimesh_removed_pbmc_umap <- trimesh_removed_pbmc_umap +
-  geom_point(size = 0.01, colour = "#33a02c") +
+  geom_point(size = 0.005, colour = "#33a02c") +
   #ggtitle("(b)") +

### 2024-02-01 paper.tex
Changed the point size of trimesh

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, interior hidden, boxrule=0pt, frame hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, sharp corners, borderline west={3pt}{0pt}{shadecolor}, enhanced, boxrule=0pt, frame hidden, interior hidden]}{\end{tcolorbox}}\fi
 
@@
 \caption{\label{fig-nldervisPBMCUMAP2}(a) 2D layout from UMAP
-(n\_neighbors = 15, min\_dist = 0.99, metric = euclidean) applied for
-the PBMC3k dataset. Is this a best representation of the original data?,
-(b) Model in the 2D space with UMAP (\textless\textgreater)}
+(n\_neighbors = 15, min\_dist = 0.99, metric = cosine) applied for the
+PBMC3k dataset. Is this a best representation of the original data?, (b)
+Model in the 2D space with UMAP (\textless\textgreater)}
 

### 2024-02-02 paper.qmd
Add spin-brush with detour and try with liminal as well

@@
   geom_line() +
+  geom_vline(xintercept = 144, linetype="dashed", 
+                color = "red", size=0.5) +
   #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
@@
   ylab("AIC") +
-  xlab("Total number of bins")
+  xlab("total number of bins")
 ## Effective number of bins along x-axis
@@
   ylab("MSE") +
-  xlab("Total number of bins")
+  xlab("total number of bins")
 
@@
   geom_line() +
+  geom_vline(xintercept = 0.2, linetype="dashed", 
+                color = "red", size=0.5) +
   #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
@@
   geom_point() +
-  geom_point(data = error_df_lg |> dplyr::filter(method == "UMAP") |> filter(benchmark_rm_lg == 1.347), aes(x = benchmark_rm_lg,
-                                       y = total_error
-),
-                                       color = "#e41a1c") +
+  geom_vline(xintercept = 1.347, linetype="dashed", 
+                color = "red", size=0.5) +
   geom_line() +
@@
   geom_point() +
-  geom_point(data = error_df_lg |> dplyr::filter(method == "UMAP") |> filter(benchmark_rm_lg == 1.347), aes(x = benchmark_rm_lg,
-                                       y = total_mse
-),
-                                       color = "#e41a1c") +
+  geom_vline(xintercept = 1.347, linetype="dashed", 
+                color = "red", size=0.5) +
   geom_line() +
@@
 
-xtable(error_df, caption = "errors for different parameter combinations", label = "tbl-errordiffcomb") |>
-  xtable2kable() |>
+kable(error_df, caption = "errors for different parameter combinations", label = "tbl-errordiffcomb") |>
+  #xtable2kable() |>
   kable_styling(latex_options = c("striped", "HOLD_position"), position = "center", full_width = FALSE) |>

### 2024-02-02 paper.tex
Add spin-brush with detour and try with liminal as well

--- 2024-02-01 paper.tex
@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced, frame hidden, breakable, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, sharp corners, interior hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}, breakable]}{\end{tcolorbox}}\fi
 
@@
 \begin{table}[H]
+
+\caption{errors for different parameter combinations}
 \centering
-\begin{tabular}{rrrlr}
+\begin{tabular}[t]{r|r|l|r}
 \hline
-\textbf{} & \textbf{n\_neighbors} & \textbf{min\_dist} & \textbf{metric} & \textbf{error (x 100) }\\ 
+\textbf{n\_neighbors} & \textbf{min\_dist} & \textbf{metric} & \textbf{error (x 100)}\\
 \hline
-\cellcolor{gray!6}{1} & \cellcolor{gray!6}{5.00} & \cellcolor{gray!6}{0.99} & \cellcolor{gray!6}{cosine} & \cellcolor{gray!6}{355.97 }\\ 
-  2 & 230.00 & 0.99 & cosine & 378.64 \\ 
-\textcolor{red}{\textbf{\cellcolor{gray!6}{  3}}} & \textcolor{red}{\textbf{\cellcolor{gray!6}{30.00}}} & \textcolor{red}{\textbf{\cellcolor{gray!6}{0.30}}} & \textcolor{red}{\textbf{\cellcolor{gray!6}{cosine}}} & \textcolor{red}{\textbf{\cellcolor{gray!6}{386.35 }}}\\
+\cellcolor{gray!6}{5} & \cellcolor{gray!6}{0.99} & \cellcolor{gray!6}{cosine} & \cellcolor{gray!6}{262.18}\\
+\hline
+15 & 0.99 & cosine & 279.49\\
+\hline
+\cellcolor{gray!6}{84} & \cellcolor{gray!6}{0.99} & \cellcolor{gray!6}{cosine} & \cellcolor{gray!6}{295.64}\\
+\hline
+5 & 0.30 & cosine & 303.90\\
+\hline
+\cellcolor{gray!6}{15} & \cellcolor{gray!6}{0.30} & \cellcolor{gray!6}{cosine} & \cellcolor{gray!6}{308.97}\\
+\hline
+84 & 0.30 & cosine & 319.59\\
+\hline
+\textcolor{red}{\textbf{\cellcolor{gray!6}{30}}} & \textcolor{red}{\textbf{\cellcolor{gray!6}{0.30}}} & \textcolor{red}{\textbf{\cellcolor{gray!6}{cosine}}} & \textcolor{red}{\textbf{\cellcolor{gray!6}{386.35}}}\\
 \hline
 \end{tabular}
-\caption{errors for different parameter combinations} 
-\label{tbl-errordiffcomb}
 \end{table}

### 2024-02-04 appendix.qmd
Tried 5-fold cross validation to compute mse and investigate the distribution (not specific change with cross validation)

@@
 ## How model perform less than and greater than the selected number of bins (UMAP with S-curve)
+
+<!--
+```{r}
+#| warning: false
+#| echo: false
+
+## UMAP with n_neighbor: 7
+## Prediction
+
+UMAP_data <- read_rds("data/s_curve/s_curve_umap_7.rds")
+predict_UMAP_df <- read_rds("data/s_curve/s_curve_umap_7_predict.rds")
+
+shape_value_curve <- calculate_effective_shape_value(.data = UMAP_data,
+                                                     x = UMAP1, y = UMAP2)
+
+num_bins_vec <- 1:10 ## Number of bins along the x-axis
+
+vec <- stats::setNames(rep("", 4), c("number_of_bins", "number_of_observations", "total_error", "total_mse"))  ## Define column names
+
+eval_data_test <- dplyr::bind_rows(vec)[0, ]
+eval_data_test <- eval_data_test |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(num_bins_vec)) {
+
+  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = UMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
+  pred_df_training <- pred_df_training_object$pred_data
+  centroid_df_training <- pred_df_training_object$df_bin_centroids
+  avg_df_training <- pred_df_training_object$df_bin
+
+  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+
+  pred_df_test_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = predict_UMAP_df, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
+  pred_df_test <- pred_df_test_object$pred_data
+  centroid_df_test <- pred_df_test_object$df_bin_centroids
+  avg_df_test <- pred_df_test_object$df_bin
+
+  eval_df_test <- generate_eval_df(data = data, prediction_df = pred_df_test, df_bin_centroids = centroid_df_test, df_bin = avg_df_test, num_bins = num_bins_vec[i], col_start = "x")
+
+  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+  eval_data_test <- dplyr::bind_rows(eval_data_test, eval_df_test)
+
+}
+
+## Add new column with data types
+
+eval_data_training <- eval_data_training |>
+  mutate(data_type = "training")
+
+eval_data_test <- eval_data_test |>
+  mutate(data_type = "test")
+
+MSE_df_umap_1 <- bind_rows(eval_data_training, eval_data_test) |>
+  dplyr::mutate(param = "n_neighbors: 7")
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## UMAP with n_neighbor: 15
+## Prediction
+
+UMAP_data <- read_rds("data/s_curve/s_curve_umap_15.rds")
+predict_UMAP_df <- read_rds("data/s_curve/s_curve_umap_15_predict.rds")
+
+shape_value_curve <- calculate_effective_shape_value(.data = UMAP_data,
+                                                     x = UMAP1, y = UMAP2)
+
+num_bins_vec <- 1:10 ## Number of bins along the x-axis
+
+vec <- stats::setNames(rep("", 4), c("number_of_bins", "number_of_observations", "total_error", "total_mse"))  ## Define column names
+
+eval_data_test <- dplyr::bind_rows(vec)[0, ]
+eval_data_test <- eval_data_test |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(num_bins_vec)) {
+
+  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = UMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
+  pred_df_training <- pred_df_training_object$pred_data
+  centroid_df_training <- pred_df_training_object$df_bin_centroids
+  avg_df_training <- pred_df_training_object$df_bin
+
+  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+
+  pred_df_test_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = predict_UMAP_df, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
+  pred_df_test <- pred_df_test_object$pred_data
+  centroid_df_test <- pred_df_test_object$df_bin_centroids
+  avg_df_test <- pred_df_test_object$df_bin
+
+  eval_df_test <- generate_eval_df(data = data, prediction_df = pred_df_test, df_bin_centroids = centroid_df_test, df_bin = avg_df_test, num_bins = num_bins_vec[i], col_start = "x")
+
+  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+  eval_data_test <- dplyr::bind_rows(eval_data_test, eval_df_test)
+
+}
+
+## Add new column with data types
+
+eval_data_training <- eval_data_training |>
+  mutate(data_type = "training")
+
+eval_data_test <- eval_data_test |>
+  mutate(data_type = "test")
+
+MSE_df_umap_2 <- bind_rows(eval_data_training, eval_data_test) |>
+  dplyr::mutate(param = "n_neighbors: 15")
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## UMAP with n_neighbor: 32
+## Prediction
+
+UMAP_data <- read_rds("data/s_curve/s_curve_umap_32.rds")
+predict_UMAP_df <- read_rds("data/s_curve/s_curve_umap_32_predict.rds")
+
+shape_value_curve <- calculate_effective_shape_value(.data = UMAP_data,
+                                                     x = UMAP1, y = UMAP2)
+
+num_bins_vec <- 1:10 ## Number of bins along the x-axis
+
+vec <- stats::setNames(rep("", 4), c("number_of_bins", "number_of_observations", "total_error", "total_mse"))  ## Define column names
+
+eval_data_test <- dplyr::bind_rows(vec)[0, ]
+eval_data_test <- eval_data_test |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(num_bins_vec)) {
+
+  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = UMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
+  pred_df_training <- pred_df_training_object$pred_data
+  centroid_df_training <- pred_df_training_object$df_bin_centroids
+  avg_df_training <- pred_df_training_object$df_bin
+
+  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+
+  pred_df_test_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = predict_UMAP_df, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
+  pred_df_test <- pred_df_test_object$pred_data
+  centroid_df_test <- pred_df_test_object$df_bin_centroids
+  avg_df_test <- pred_df_test_object$df_bin
+
+  eval_df_test <- generate_eval_df(data = data, prediction_df = pred_df_test, df_bin_centroids = centroid_df_test, df_bin = avg_df_test, num_bins = num_bins_vec[i], col_start = "x")
+
+  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+  eval_data_test <- dplyr::bind_rows(eval_data_test, eval_df_test)
+
+}
+
+## Add new column with data types
+
+eval_data_training <- eval_data_training |>
+  mutate(data_type = "training")
+
+eval_data_test <- eval_data_test |>
+  mutate(data_type = "test")
+
+MSE_df_umap_3 <- bind_rows(eval_data_training, eval_data_test) |>
+  dplyr::mutate(param = "n_neighbors: 32")
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## UMAP with n_neighbor: 50
+## Prediction
+
+UMAP_data <- read_rds("data/s_curve/s_curve_umap.rds")
+predict_UMAP_df <- read_rds("data/s_curve/s_curve_umap_predict.rds")
+
+shape_value_curve <- calculate_effective_shape_value(.data = UMAP_data,
+                                                     x = UMAP1, y = UMAP2)
+
+num_bins_vec <- 1:10 ## Number of bins along the x-axis
+
+vec <- stats::setNames(rep("", 4), c("number_of_bins", "number_of_observations", "total_error", "total_mse"))  ## Define column names
+
+eval_data_test <- dplyr::bind_rows(vec)[0, ]
+eval_data_test <- eval_data_test |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(num_bins_vec)) {
+
+  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = UMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
+  pred_df_training <- pred_df_training_object$pred_data
+  centroid_df_training <- pred_df_training_object$df_bin_centroids
+  avg_df_training <- pred_df_training_object$df_bin
+
+  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+
+  pred_df_test_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = predict_UMAP_df, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
+  pred_df_test <- pred_df_test_object$pred_data
+  centroid_df_test <- pred_df_test_object$df_bin_centroids
+  avg_df_test <- pred_df_test_object$df_bin
+
+  eval_df_test <- generate_eval_df(data = data, prediction_df = pred_df_test, df_bin_centroids = centroid_df_test, df_bin = avg_df_test, num_bins = num_bins_vec[i], col_start = "x")
+
+  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+  eval_data_test <- dplyr::bind_rows(eval_data_test, eval_df_test)
+
+}
+
+## Add new column with data types
+
+eval_data_training <- eval_data_training |>
+  mutate(data_type = "training")
+
+eval_data_test <- eval_data_test |>
+  mutate(data_type = "test")
+
+MSE_df_umap_4 <- bind_rows(eval_data_training, eval_data_test) |>
+  dplyr::mutate(param = "n_neighbors: 50")
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+MSE_df_umap <- dplyr::bind_rows(MSE_df_umap_1, MSE_df_umap_2, MSE_df_umap_3, MSE_df_umap_4)
+MSE_df_umap$param <- factor(MSE_df_umap$param, levels = c("n_neighbors: 7", "n_neighbors: 15", "n_neighbors: 32", "n_neighbors: 50"))
+
+## To draw with AIC
+aic_plot_param <- ggplot(MSE_df_umap |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
+                                                                         y = total_error,
+                                                                         color = param
+)) +
+  geom_point() +
+  geom_line() +
+  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  theme_light() +
+  theme(legend.title = element_blank(), legend.text = element_text(size=7), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  scale_color_manual(values = c("#8dd3c7", "#a65628", "#f781bf", "#999999")) +
+  ylab("AIC") +
+  xlab("Total number of bins")
+## Effective number of bins along x-axis
+
+mse_plot_param_training <- ggplot(MSE_df_umap |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
+                               y = total_mse,
+                               color = param
+)) +
+  geom_point() +
+  geom_line() +
+  theme_light() +
+  theme(legend.title = element_blank(), legend.text = element_text(size=7), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  scale_color_manual(values = c("#8dd3c7", "#a65628", "#f781bf", "#999999")) +
+  ylab("MSE") +
+  xlab("Total number of bins")
+
+mse_plot_param_test <- ggplot(MSE_df_umap |> dplyr::filter(data_type == "test"), aes(x = number_of_bins,
+                                                                                             y = total_mse,
+                                                                                             color = param
+)) +
+  geom_point() +
+  geom_line() +
+  theme_light() +
+  theme(legend.title = element_blank(), legend.text = element_text(size=7), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  scale_color_manual(values = c("#8dd3c7", "#a65628", "#f781bf", "#999999")) +
+  ylab("MSE") +
+  xlab("Total number of bins")
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "Goodness of fit statistics from different n_neighbors parameter of UMAP applied to training S-curve dataset.
+What is the best parameter choice in UMAP to represent the original data in 2D?"
+#| label: fig-diagnosticpltDiffParam
+#| out-width: 100%
+
+aic_plot_param + mse_plot_param_training + mse_plot_param_test  +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(guides='collect', ncol = 3) &
+  theme(legend.position='bottom', plot.tag = element_text(size = 8))
+```
+-->
+
+<!--model for S-curve with 5 methods-->
+<!--
+```{r}
+#| warning: false
+#| echo: false
+
+## Import data
+
+tSNE_s_curve <- read_rds("data/s_curve/s_curve_tsne_27.rds")
+UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap.rds")
+PHATE_s_curve <- read_rds("data/s_curve/s_curve_phate.rds")
+TriMAP_s_curve <- read_rds("data/s_curve/s_curve_trimap.rds")
+PaCMAP_s_curve <- read_rds("data/s_curve/s_curve_pacmap.rds")
+
+## tSNE
+
+num_bins_tsne_s_curve <- 8
+shape_val_tsne_s_curve <- calculate_effective_shape_value(.data = tSNE_s_curve,
+                                                          x = tSNE1, y = tSNE2) ## 1.259938
+## To extract bin centroids
+hexbin_data_object_tsne_s_curve <- extract_hexbin_centroids(nldr_df = tSNE_s_curve, num_bins = num_bins_tsne_s_curve, shape_val = shape_val_tsne_s_curve, x = tSNE1, y = tSNE2)
+
+df_bin_centroids_tsne_s_curve <- hexbin_data_object_tsne_s_curve$hexdf_data
+
+tSNE_data_with_hb_id_s_curve <- tSNE_s_curve |>
+  dplyr::mutate(hb_id = hexbin_data_object_tsne_s_curve$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_tsne_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), tSNE_data_with_hb_id_s_curve)
+
+## Averaged on high-D
+df_bin_tsne_s_curve <- avg_highD_data(.data = df_all_tsne_s_curve)
+
+## Triangulate bin centroids
+tr1_object_tsne_s_curve <- triangulate_bin_centroids(df_bin_centroids_tsne_s_curve, x, y)
+tr_from_to_df_tsne_s_curve <- generate_edge_info(triangular_object = tr1_object_tsne_s_curve)
+
+# ggplot(df_bin_centroids_tsne_s_curve, aes(x = x, y = y)) +
+#   geom_segment(data = tr_from_to_df_tsne_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
+#   geom_point(size = 2, colour = "#33a02c") +
+#   coord_equal()
+
+## Compute 2D distances
+distance_tsne_s_curve <- cal_2d_dist(.data = tr_from_to_df_tsne_s_curve)
+
+## To find the benchmark value
+benchmark_tsne_s_curve <- find_benchmark_value(.data = distance_tsne_s_curve, distance_col = distance)
+
+# colour_long_edges(.data = distance_tsne_s_curve, benchmark_value = benchmark_tsne_s_curve,
+#                   triangular_object = tr1_object_tsne_s_curve, distance_col = distance)
+
+trimesh_removed_tsne_s_curve <- remove_long_edges(.data = distance_tsne_s_curve, benchmark_value = benchmark_tsne_s_curve,
+                                                  triangular_object = tr1_object_tsne_s_curve, distance_col = distance)
+
+trimesh_removed_tsne_s_curve <- trimesh_removed_tsne_s_curve +
+  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  # theme(axis.text = element_text(size = 5),
+  #       axis.title = element_text(size = 7)) +
+  geom_point(colour = "#33a02c", size = 0.5) +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+# tour_tsne_s_curve <- show_langevitour(df_all_tsne_s_curve, df_bin_tsne_s_curve, df_bin_centroids_tsne_s_curve, benchmark_value = benchmark_tsne_s_curve, distance = distance_tsne_s_curve, distance_col = distance)
+
+## UMAP
+
+num_bins_umap_s_curve <- 6
+shape_val_umap_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
+                                                          x = UMAP1, y = UMAP2) ## 1.259938
+## To extract bin centroids
+hexbin_data_object_umap_s_curve <- extract_hexbin_centroids(nldr_df = UMAP_s_curve, num_bins = num_bins_umap_s_curve, shape_val = shape_val_umap_s_curve, x = UMAP1, y = UMAP2)
+
+df_bin_centroids_umap_s_curve <- hexbin_data_object_umap_s_curve$hexdf_data
+
+UMAP_data_with_hb_id_s_curve <- UMAP_s_curve |>
+  dplyr::mutate(hb_id = hexbin_data_object_umap_s_curve$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_umap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), UMAP_data_with_hb_id_s_curve)
+
+## Averaged on high-D
+df_bin_umap_s_curve <- avg_highD_data(.data = df_all_umap_s_curve)
+
+## Triangulate bin centroids
+tr1_object_umap_s_curve <- triangulate_bin_centroids(df_bin_centroids_umap_s_curve, x, y)
+tr_from_to_df_umap_s_curve <- generate_edge_info(triangular_object = tr1_object_umap_s_curve)
+
+# ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
+#   geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
+#   geom_point(size = 2, colour = "#33a02c") +
+#   coord_equal()
+
+## Compute 2D distances
+distance_umap_s_curve <- cal_2d_dist(.data = tr_from_to_df_umap_s_curve)
+
+## To find the benchmark value
+benchmark_umap_s_curve <- find_benchmark_value(.data = distance_umap_s_curve, distance_col = distance)
+
+# colour_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve,
+#                   triangular_object = tr1_object_umap_s_curve, distance_col = distance)
+
+trimesh_removed_umap_s_curve <- remove_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve,
+                                     triangular_object = tr1_object_umap_s_curve, distance_col = distance)
+
+trimesh_removed_umap_s_curve <- trimesh_removed_umap_s_curve +
+  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  # theme(axis.text = element_text(size = 5),
+  #       axis.title = element_text(size = 7)) +
+  geom_point(colour = "#33a02c", size = 0.5) +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+# tour_umap_s_curve <- show_langevitour(df_all_umap_s_curve, df_bin_umap_s_curve, df_bin_centroids_umap_s_curve, benchmark_value = benchmark_umap_s_curve, distance = distance_umap_s_curve, distance_col = distance)
+
+## TriMAP
+
+num_bins_trimap_s_curve <- 6
+shape_val_trimap_s_curve <- calculate_effective_shape_value(.data = TriMAP_s_curve,
+                                                            x = TriMAP1, y = TriMAP2) ## 1.259938
+## To extract bin centroids
+hexbin_data_object_trimap_s_curve <- extract_hexbin_centroids(nldr_df = TriMAP_s_curve, num_bins = num_bins_trimap_s_curve, shape_val = shape_val_trimap_s_curve, x = TriMAP1, y = TriMAP2)
+
+df_bin_centroids_trimap_s_curve <- hexbin_data_object_trimap_s_curve$hexdf_data
+
+TriMAP_data_with_hb_id_s_curve <- TriMAP_s_curve |>
+  dplyr::mutate(hb_id = hexbin_data_object_trimap_s_curve$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_trimap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), TriMAP_data_with_hb_id_s_curve)
+
+## Averaged on high-D
+df_bin_trimap_s_curve <- avg_highD_data(.data = df_all_trimap_s_curve)
+
+## Triangulate bin centroids
+tr1_object_trimap_s_curve <- triangulate_bin_centroids(df_bin_centroids_trimap_s_curve, x, y)
+tr_from_to_df_trimap_s_curve <- generate_edge_info(triangular_object = tr1_object_trimap_s_curve)
+
+# ggplot(df_bin_centroids_trimap_s_curve, aes(x = x, y = y)) +
+#   geom_segment(data = tr_from_to_df_trimap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
+#   geom_point(size = 2, colour = "#33a02c") +
+#   coord_equal()
+
+## Compute 2D distances
+distance_trimap_s_curve <- cal_2d_dist(.data = tr_from_to_df_trimap_s_curve)
+
+## To find the benchmark value
+benchmark_trimap_s_curve <- find_benchmark_value(.data = distance_trimap_s_curve, distance_col = distance)
+
+# colour_long_edges(.data = distance_trimap_s_curve, benchmark_value = benchmark_trimap_s_curve,
+#                   triangular_object = tr1_object_trimap_s_curve, distance_col = distance)
+
+trimesh_removed_trimap_s_curve <- remove_long_edges(.data = distance_trimap_s_curve, benchmark_value = benchmark_trimap_s_curve,
+                                                    triangular_object = tr1_object_trimap_s_curve, distance_col = distance)
+
+trimesh_removed_trimap_s_curve <- trimesh_removed_trimap_s_curve +
+  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  # theme(axis.text = element_text(size = 5),
+  #       axis.title = element_text(size = 7)) +
+  geom_point(colour = "#33a02c", size = 0.25) +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+# tour_trimap_s_curve <- show_langevitour(df_all_trimap_s_curve, df_bin_trimap_s_curve, df_bin_centroids_trimap_s_curve, benchmark_value = benchmark_trimap_s_curve, distance = distance_trimap_s_curve, distance_col = distance)
+
+## PacMAP
+
+num_bins_pacmap_s_curve <- 10
+shape_val_pacmap_s_curve <- calculate_effective_shape_value(.data = PaCMAP_s_curve,
+                                                            x = PaCMAP1, y = PaCMAP2) ## 1.259938
+## To extract bin centroids
+hexbin_data_object_pacmap_s_curve <- extract_hexbin_centroids(nldr_df = PaCMAP_s_curve, num_bins = num_bins_pacmap_s_curve, shape_val = shape_val_pacmap_s_curve, x = PaCMAP1, y = PaCMAP2)
+
+df_bin_centroids_pacmap_s_curve <- hexbin_data_object_pacmap_s_curve$hexdf_data
+
+PaCMAP_data_with_hb_id_s_curve <- PaCMAP_s_curve |>
+  dplyr::mutate(hb_id = hexbin_data_object_pacmap_s_curve$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_pacmap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), PaCMAP_data_with_hb_id_s_curve)
+
+## Averaged on high-D
+df_bin_pacmap_s_curve <- avg_highD_data(.data = df_all_pacmap_s_curve)
+
+## Triangulate bin centroids
+tr1_object_pacmap_s_curve <- triangulate_bin_centroids(df_bin_centroids_pacmap_s_curve, x, y)
+tr_from_to_df_pacmap_s_curve <- generate_edge_info(triangular_object = tr1_object_pacmap_s_curve)
+
+# ggplot(df_bin_centroids_pacmap_s_curve, aes(x = x, y = y)) +
+#   geom_segment(data = tr_from_to_df_pacmap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
+#   geom_point(size = 2, colour = "#33a02c") +
+#   coord_equal()
+
+## Compute 2D distances
+distance_pacmap_s_curve <- cal_2d_dist(.data = tr_from_to_df_pacmap_s_curve)
+
+## To find the benchmark value
+benchmark_pacmap_s_curve <- find_benchmark_value(.data = distance_pacmap_s_curve, distance_col = distance)
+
+# colour_long_edges(.data = distance_pacmap_s_curve, benchmark_value = benchmark_pacmap_s_curve,
+#                   triangular_object = tr1_object_pacmap_s_curve, distance_col = distance)
+
+trimesh_removed_pacmap_s_curve <- remove_long_edges(.data = distance_pacmap_s_curve, benchmark_value = benchmark_pacmap_s_curve,
+                                     triangular_object = tr1_object_pacmap_s_curve, distance_col = distance)
+
+trimesh_removed_pacmap_s_curve <- trimesh_removed_pacmap_s_curve +
+  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  # theme(axis.text = element_text(size = 5),
+  #       axis.title = element_text(size = 7)) +
+  geom_point(colour = "#33a02c", size = 0.5) +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'e', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+# tour_pacmap_s_curve <- show_langevitour(df_all_pacmap_s_curve, df_bin_pacmap_s_curve, df_bin_centroids_pacmap_s_curve, benchmark_value = benchmark_pacmap_s_curve, distance = distance_pacmap_s_curve, distance_col = distance)
+
+## PHATE
+
+num_bins_phate_s_curve <- 8
+shape_val_phate_s_curve <- calculate_effective_shape_value(.data = PHATE_s_curve,
+                                                           x = PHATE1, y = PHATE2) ## 1.259938
+## To extract bin centroids
+hexbin_data_object_phate_s_curve <- extract_hexbin_centroids(nldr_df = PHATE_s_curve, num_bins = num_bins_phate_s_curve, shape_val = shape_val_phate_s_curve, x = PHATE1, y = PHATE2)
+
+df_bin_centroids_phate_s_curve <- hexbin_data_object_phate_s_curve$hexdf_data
+
+## Identify bins with low-density
+identify_rm_bins <- find_low_density_hexagons(df_bin_centroids_phate_s_curve, num_bins_phate_s_curve, benchmark_rm_hex = 0.06)
+
+df_bin_centroids_phate_s_curve <- df_bin_centroids_phate_s_curve |>
+  filter(!(hexID %in% identify_rm_bins))
+
+PHATE_data_with_hb_id_s_curve <- PHATE_s_curve |>
+  dplyr::mutate(hb_id = hexbin_data_object_phate_s_curve$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_phate_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), PHATE_data_with_hb_id_s_curve)
+
+## Averaged on high-D
+df_bin_phate_s_curve <- avg_highD_data(.data = df_all_phate_s_curve)
+
+## Triangulate bin centroids
+tr1_object_phate_s_curve <- triangulate_bin_centroids(df_bin_centroids_phate_s_curve, x, y)
+tr_from_to_df_phate_s_curve <- generate_edge_info(triangular_object = tr1_object_phate_s_curve)
+
+# ggplot(df_bin_centroids_phate_s_curve, aes(x = x, y = y)) +
+#   geom_segment(data = tr_from_to_df_phate_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
+#   geom_point(size = 2, colour = "#33a02c") +
+#   coord_equal()
+
+## Compute 2D distances
+distance_phate_s_curve <- cal_2d_dist(.data = tr_from_to_df_phate_s_curve)
+
+## To find the benchmark value
+benchmark_phate_s_curve <- find_benchmark_value(.data = distance_phate_s_curve, distance_col = distance)
+
+# colour_long_edges(.data = distance_phate_s_curve, benchmark_value = benchmark_phate_s_curve,
+#                   triangular_object = tr1_object_phate_s_curve, distance_col = distance)
+
+trimesh_removed_phate_s_curve <- remove_long_edges(.data = distance_phate_s_curve, benchmark_value = benchmark_phate_s_curve,
+                                                   triangular_object = tr1_object_phate_s_curve, distance_col = distance)
+
+trimesh_removed_phate_s_curve <- trimesh_removed_phate_s_curve +
+  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  # theme(axis.text = element_text(size = 5),
+  #       axis.title = element_text(size = 7)) +
+  geom_point(colour = "#33a02c", size = 0.5) +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+# tour_phate_s_curve <- show_langevitour(df_all_phate_s_curve, df_bin_phate_s_curve, df_bin_centroids_phate_s_curve, benchmark_value = benchmark_phate_s_curve, distance = distance_phate_s_curve, distance_col = distance)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+#| label: fig-modelScurve
+#| fig-cap: Is there a good model to represent the original data in 2D space or are they all providing equivalent information?, (a) Model in the 2D space with tSNE (<https://youtu.be/uy9SkAo6gAE>), (b) Model in the 2D space with UMAP (<https://youtu.be/0MJDhHrh_Ug>), (c) Model in the 2D space with PHATE (<https://youtu.be/HbVv0uy0QWk>), (d) Model in the 2D space with TriMAP (<https://youtu.be/2OGwipAjzc8>), and (e) Model in the 2D space with PaCMAP (<https://youtu.be/pkokI8d-cBk>).
+#| out-width: 100%
+
+trimesh_removed_tsne_s_curve + trimesh_removed_umap_s_curve + trimesh_removed_phate_s_curve + trimesh_removed_trimap_s_curve + trimesh_removed_pacmap_s_curve +
+  #plot_annotation(tag_levels = 'a') +
+  plot_layout(ncol = 5) &
+  theme(legend.position='none', plot.tag = element_text(size = 8))
+```
+-->

### 2024-02-04 paper.qmd
Tried 5-fold cross validation to compute mse and investigate the distribution (not specific change with cross validation)

@@
 ## Compute 2D distances
-distance <- cal_2D_dist(.data = tr_from_to_df)
+distance <- cal_2d_dist(.data = tr_from_to_df)
 
@@
 
-  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = tSNE_data, nldr_df_test = tSNE_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "tSNE1", y = "tSNE2", col_start = "x")
-  pred_df_training <- pred_df_training_object$pred_data
-  centroid_df_training <- pred_df_training_object$df_bin_centroids
-  avg_df_training <- pred_df_training_object$df_bin
+  model_object <- fit_high_d_model(training_data = training_data, nldr_df_with_id = tSNE_data, x = "tSNE1", y = "tSNE2", num_bins_x = num_bins_vec[i], shape_val = shape_value_curve,
+                             is_bin_centroid = TRUE,
+                             is_rm_lwd_hex = FALSE,
+                             benchmark_to_rm_lwd_hex = NA,
+                             is_avg_high_d = TRUE, column_start_text = "x")
+
+  centroid_df_training <- model_object$df_bin_centroids
+  avg_df_training <- model_object$df_bin
+
+  # pred_df_training <- predict_2d_embeddings(test_data = training_data, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "tSNE")
+
+  pred_df_training <- predict_hex_id(df_bin_centroids = centroid_df_training, nldr_df_test = tSNE_data, x = "tSNE1", y = "tSNE2")
 
@@
 
-  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = UMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
-  pred_df_training <- pred_df_training_object$pred_data
-  centroid_df_training <- pred_df_training_object$df_bin_centroids
-  avg_df_training <- pred_df_training_object$df_bin
+  model_object <- fit_high_d_model(training_data = training_data, nldr_df_with_id = UMAP_data, x = "UMAP1", y = "UMAP2", num_bins_x = num_bins_vec[i], shape_val = shape_value_curve,
+                             is_bin_centroid = TRUE,
+                             is_rm_lwd_hex = FALSE,
+                             benchmark_to_rm_lwd_hex = NA,
+                             is_avg_high_d = TRUE, column_start_text = "x")
+
+  centroid_df_training <- model_object$df_bin_centroids
+  avg_df_training <- model_object$df_bin
+
+  # pred_df_training <- predict_2d_embeddings(test_data = training_data, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "tSNE")
+
+  pred_df_training <- predict_hex_id(df_bin_centroids = centroid_df_training, nldr_df_test = UMAP_data, x = "UMAP1", y = "UMAP2")
 
@@
 
-  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = PHATE_data, nldr_df_test = PHATE_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "PHATE1", y = "PHATE2", col_start = "x")
-  pred_df_training <- pred_df_training_object$pred_data
-  centroid_df_training <- pred_df_training_object$df_bin_centroids
-  avg_df_training <- pred_df_training_object$df_bin
+  model_object <- fit_high_d_model(training_data = training_data, nldr_df_with_id = PHATE_data, x = "PHATE1", y = "PHATE2", num_bins_x = num_bins_vec[i], shape_val = shape_value_curve,
+                             is_bin_centroid = TRUE,
+                             is_rm_lwd_hex = FALSE,
+                             benchmark_to_rm_lwd_hex = NA,
+                             is_avg_high_d = TRUE, column_start_text = "x")
+
+  centroid_df_training <- model_object$df_bin_centroids
+  avg_df_training <- model_object$df_bin
+
+  # pred_df_training <- predict_2d_embeddings(test_data = training_data, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "tSNE")
+
+  pred_df_training <- predict_hex_id(df_bin_centroids = centroid_df_training, nldr_df_test = PHATE_data, x = "PHATE1", y = "PHATE2")
 
@@
 
-  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = TriMAP_data, nldr_df_test = TriMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "TriMAP1", y = "TriMAP2", col_start = "x")
-  pred_df_training <- pred_df_training_object$pred_data
-  centroid_df_training <- pred_df_training_object$df_bin_centroids
-  avg_df_training <- pred_df_training_object$df_bin
+  model_object <- fit_high_d_model(training_data = training_data, nldr_df_with_id = TriMAP_data, x = "TriMAP1", y = "TriMAP2", num_bins_x = num_bins_vec[i], shape_val = shape_value_curve,
+                             is_bin_centroid = TRUE,
+                             is_rm_lwd_hex = FALSE,
+                             benchmark_to_rm_lwd_hex = NA,
+                             is_avg_high_d = TRUE, column_start_text = "x")
+
+  centroid_df_training <- model_object$df_bin_centroids
+  avg_df_training <- model_object$df_bin
+
+  # pred_df_training <- predict_2d_embeddings(test_data = training_data, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "tSNE")
+
+  pred_df_training <- predict_hex_id(df_bin_centroids = centroid_df_training, nldr_df_test = TriMAP_data, x = "TriMAP1", y = "TriMAP2")
 
@@
 
-  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = PaCMAP_data, nldr_df_test = PaCMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "PaCMAP1", y = "PaCMAP2", col_start = "x")
-  pred_df_training <- pred_df_training_object$pred_data
-  centroid_df_training <- pred_df_training_object$df_bin_centroids
-  avg_df_training <- pred_df_training_object$df_bin
+  model_object <- fit_high_d_model(training_data = training_data, nldr_df_with_id = PaCMAP_data, x = "PaCMAP1", y = "PaCMAP2", num_bins_x = num_bins_vec[i], shape_val = shape_value_curve,
+                             is_bin_centroid = TRUE,
+                             is_rm_lwd_hex = FALSE,
+                             benchmark_to_rm_lwd_hex = NA,
+                             is_avg_high_d = TRUE, column_start_text = "x")
+
+  centroid_df_training <- model_object$df_bin_centroids
+  avg_df_training <- model_object$df_bin
+
+  # pred_df_training <- predict_2d_embeddings(test_data = training_data, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "tSNE")
+
+  pred_df_training <- predict_hex_id(df_bin_centroids = centroid_df_training, nldr_df_test = PaCMAP_data, x = "PaCMAP1", y = "PaCMAP2")
 
@@
 ## Compute 2D distances
-distance_umap_s_curve <- cal_2D_dist(.data = tr_from_to_df_umap_s_curve)
+distance_umap_s_curve <- cal_2d_dist(.data = tr_from_to_df_umap_s_curve)
 
@@
 
-<!--
-```{r}
-#| warning: false
-#| echo: false
-
-## UMAP with n_neighbor: 7
-## Prediction
-
-UMAP_data <- read_rds("data/s_curve/s_curve_umap_7.rds")
-predict_UMAP_df <- read_rds("data/s_curve/s_curve_umap_7_predict.rds")
-
-shape_value_curve <- calculate_effective_shape_value(.data = UMAP_data,
-                                                     x = UMAP1, y = UMAP2)
-
-num_bins_vec <- 1:10 ## Number of bins along the x-axis
-
-vec <- stats::setNames(rep("", 4), c("number_of_bins", "number_of_observations", "total_error", "total_mse"))  ## Define column names
-
-eval_data_test <- dplyr::bind_rows(vec)[0, ]
-eval_data_test <- eval_data_test |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-eval_data_training <- dplyr::bind_rows(vec)[0, ]
-eval_data_training <- eval_data_training |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(num_bins_vec)) {
-
-  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = UMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
-  pred_df_training <- pred_df_training_object$pred_data
-  centroid_df_training <- pred_df_training_object$df_bin_centroids
-  avg_df_training <- pred_df_training_object$df_bin
-
-  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
-
-  pred_df_test_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = predict_UMAP_df, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
-  pred_df_test <- pred_df_test_object$pred_data
-  centroid_df_test <- pred_df_test_object$df_bin_centroids
-  avg_df_test <- pred_df_test_object$df_bin
-
-  eval_df_test <- generate_eval_df(data = data, prediction_df = pred_df_test, df_bin_centroids = centroid_df_test, df_bin = avg_df_test, num_bins = num_bins_vec[i], col_start = "x")
-
-  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
-  eval_data_test <- dplyr::bind_rows(eval_data_test, eval_df_test)
-
-}
-
-## Add new column with data types
-
-eval_data_training <- eval_data_training |>
-  mutate(data_type = "training")
-
-eval_data_test <- eval_data_test |>
-  mutate(data_type = "test")
-
-MSE_df_umap_1 <- bind_rows(eval_data_training, eval_data_test) |>
-  dplyr::mutate(param = "n_neighbors: 7")
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-## UMAP with n_neighbor: 15
-## Prediction
-
-UMAP_data <- read_rds("data/s_curve/s_curve_umap_15.rds")
-predict_UMAP_df <- read_rds("data/s_curve/s_curve_umap_15_predict.rds")
-
-shape_value_curve <- calculate_effective_shape_value(.data = UMAP_data,
-                                                     x = UMAP1, y = UMAP2)
-
-num_bins_vec <- 1:10 ## Number of bins along the x-axis
-
-vec <- stats::setNames(rep("", 4), c("number_of_bins", "number_of_observations", "total_error", "total_mse"))  ## Define column names
-
-eval_data_test <- dplyr::bind_rows(vec)[0, ]
-eval_data_test <- eval_data_test |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-eval_data_training <- dplyr::bind_rows(vec)[0, ]
-eval_data_training <- eval_data_training |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(num_bins_vec)) {
-
-  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = UMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
-  pred_df_training <- pred_df_training_object$pred_data
-  centroid_df_training <- pred_df_training_object$df_bin_centroids
-  avg_df_training <- pred_df_training_object$df_bin
-
-  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
-
-  pred_df_test_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = predict_UMAP_df, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
-  pred_df_test <- pred_df_test_object$pred_data
-  centroid_df_test <- pred_df_test_object$df_bin_centroids
-  avg_df_test <- pred_df_test_object$df_bin
-
-  eval_df_test <- generate_eval_df(data = data, prediction_df = pred_df_test, df_bin_centroids = centroid_df_test, df_bin = avg_df_test, num_bins = num_bins_vec[i], col_start = "x")
-
-  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
-  eval_data_test <- dplyr::bind_rows(eval_data_test, eval_df_test)
-
-}
-
-## Add new column with data types
-
-eval_data_training <- eval_data_training |>
-  mutate(data_type = "training")
-
-eval_data_test <- eval_data_test |>
-  mutate(data_type = "test")
-
-MSE_df_umap_2 <- bind_rows(eval_data_training, eval_data_test) |>
-  dplyr::mutate(param = "n_neighbors: 15")
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-## UMAP with n_neighbor: 32
-## Prediction
-
-UMAP_data <- read_rds("data/s_curve/s_curve_umap_32.rds")
-predict_UMAP_df <- read_rds("data/s_curve/s_curve_umap_32_predict.rds")
-
-shape_value_curve <- calculate_effective_shape_value(.data = UMAP_data,
-                                                     x = UMAP1, y = UMAP2)
-
-num_bins_vec <- 1:10 ## Number of bins along the x-axis
-
-vec <- stats::setNames(rep("", 4), c("number_of_bins", "number_of_observations", "total_error", "total_mse"))  ## Define column names
-
-eval_data_test <- dplyr::bind_rows(vec)[0, ]
-eval_data_test <- eval_data_test |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-eval_data_training <- dplyr::bind_rows(vec)[0, ]
-eval_data_training <- eval_data_training |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(num_bins_vec)) {
-
-  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = UMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
-  pred_df_training <- pred_df_training_object$pred_data
-  centroid_df_training <- pred_df_training_object$df_bin_centroids
-  avg_df_training <- pred_df_training_object$df_bin
-
-  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
-
-  pred_df_test_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = predict_UMAP_df, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
-  pred_df_test <- pred_df_test_object$pred_data
-  centroid_df_test <- pred_df_test_object$df_bin_centroids
-  avg_df_test <- pred_df_test_object$df_bin
-
-  eval_df_test <- generate_eval_df(data = data, prediction_df = pred_df_test, df_bin_centroids = centroid_df_test, df_bin = avg_df_test, num_bins = num_bins_vec[i], col_start = "x")
-
-  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
-  eval_data_test <- dplyr::bind_rows(eval_data_test, eval_df_test)
-
-}
-
-## Add new column with data types
-
-eval_data_training <- eval_data_training |>
-  mutate(data_type = "training")
-
-eval_data_test <- eval_data_test |>
-  mutate(data_type = "test")
-
-MSE_df_umap_3 <- bind_rows(eval_data_training, eval_data_test) |>
-  dplyr::mutate(param = "n_neighbors: 32")
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-## UMAP with n_neighbor: 50
-## Prediction
-
-UMAP_data <- read_rds("data/s_curve/s_curve_umap.rds")
-predict_UMAP_df <- read_rds("data/s_curve/s_curve_umap_predict.rds")
-
-shape_value_curve <- calculate_effective_shape_value(.data = UMAP_data,
-                                                     x = UMAP1, y = UMAP2)
-
-num_bins_vec <- 1:10 ## Number of bins along the x-axis
-
-vec <- stats::setNames(rep("", 4), c("number_of_bins", "number_of_observations", "total_error", "total_mse"))  ## Define column names
-
-eval_data_test <- dplyr::bind_rows(vec)[0, ]
-eval_data_test <- eval_data_test |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-eval_data_training <- dplyr::bind_rows(vec)[0, ]
-eval_data_training <- eval_data_training |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(num_bins_vec)) {
-
-  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = UMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
-  pred_df_training <- pred_df_training_object$pred_data
-  centroid_df_training <- pred_df_training_object$df_bin_centroids
-  avg_df_training <- pred_df_training_object$df_bin
-
-  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
-
-  pred_df_test_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = predict_UMAP_df, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
-  pred_df_test <- pred_df_test_object$pred_data
-  centroid_df_test <- pred_df_test_object$df_bin_centroids
-  avg_df_test <- pred_df_test_object$df_bin
-
-  eval_df_test <- generate_eval_df(data = data, prediction_df = pred_df_test, df_bin_centroids = centroid_df_test, df_bin = avg_df_test, num_bins = num_bins_vec[i], col_start = "x")
-
-  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
-  eval_data_test <- dplyr::bind_rows(eval_data_test, eval_df_test)
-
-}
-
-## Add new column with data types
-
-eval_data_training <- eval_data_training |>
-  mutate(data_type = "training")
-
-eval_data_test <- eval_data_test |>
-  mutate(data_type = "test")
-
-MSE_df_umap_4 <- bind_rows(eval_data_training, eval_data_test) |>
-  dplyr::mutate(param = "n_neighbors: 50")
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-MSE_df_umap <- dplyr::bind_rows(MSE_df_umap_1, MSE_df_umap_2, MSE_df_umap_3, MSE_df_umap_4)
-MSE_df_umap$param <- factor(MSE_df_umap$param, levels = c("n_neighbors: 7", "n_neighbors: 15", "n_neighbors: 32", "n_neighbors: 50"))
-
-## To draw with AIC
-aic_plot_param <- ggplot(MSE_df_umap |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
-                                                                         y = total_error,
-                                                                         color = param
-)) +
-  geom_point() +
-  geom_line() +
-  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  theme_light() +
-  theme(legend.title = element_blank(), legend.text = element_text(size=7), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  scale_color_manual(values = c("#8dd3c7", "#a65628", "#f781bf", "#999999")) +
-  ylab("AIC") +
-  xlab("Total number of bins")
-## Effective number of bins along x-axis
-
-mse_plot_param_training <- ggplot(MSE_df_umap |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
-                               y = total_mse,
-                               color = param
-)) +
-  geom_point() +
-  geom_line() +
-  theme_light() +
-  theme(legend.title = element_blank(), legend.text = element_text(size=7), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  scale_color_manual(values = c("#8dd3c7", "#a65628", "#f781bf", "#999999")) +
-  ylab("MSE") +
-  xlab("Total number of bins")
-
-mse_plot_param_test <- ggplot(MSE_df_umap |> dplyr::filter(data_type == "test"), aes(x = number_of_bins,
-                                                                                             y = total_mse,
-                                                                                             color = param
-)) +
-  geom_point() +
-  geom_line() +
-  theme_light() +
-  theme(legend.title = element_blank(), legend.text = element_text(size=7), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  scale_color_manual(values = c("#8dd3c7", "#a65628", "#f781bf", "#999999")) +
-  ylab("MSE") +
-  xlab("Total number of bins")
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "Goodness of fit statistics from different n_neighbors parameter of UMAP applied to training S-curve dataset.
-What is the best parameter choice in UMAP to represent the original data in 2D?"
-#| label: fig-diagnosticpltDiffParam
-#| out-width: 100%
-
-aic_plot_param + mse_plot_param_training + mse_plot_param_test  +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', ncol = 3) &
-  theme(legend.position='bottom', plot.tag = element_text(size = 8))
-```
--->
-
 ### Model interpretation
@@
 
-<!--model for S-curve with 5 methods-->
-
-```{r}
-#| warning: false
-#| echo: false
-
-## Import data
-
-tSNE_s_curve <- read_rds("data/s_curve/s_curve_tsne_27.rds")
-UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap.rds")
-PHATE_s_curve <- read_rds("data/s_curve/s_curve_phate.rds")
-TriMAP_s_curve <- read_rds("data/s_curve/s_curve_trimap.rds")
-PaCMAP_s_curve <- read_rds("data/s_curve/s_curve_pacmap.rds")
-
-## tSNE
-
-num_bins_tsne_s_curve <- 8
-shape_val_tsne_s_curve <- calculate_effective_shape_value(.data = tSNE_s_curve,
-                                                          x = tSNE1, y = tSNE2) ## 1.259938
-## To extract bin centroids
-hexbin_data_object_tsne_s_curve <- extract_hexbin_centroids(nldr_df = tSNE_s_curve, num_bins = num_bins_tsne_s_curve, shape_val = shape_val_tsne_s_curve, x = tSNE1, y = tSNE2)
-
-df_bin_centroids_tsne_s_curve <- hexbin_data_object_tsne_s_curve$hexdf_data
-
-tSNE_data_with_hb_id_s_curve <- tSNE_s_curve |>
-  dplyr::mutate(hb_id = hexbin_data_object_tsne_s_curve$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_tsne_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), tSNE_data_with_hb_id_s_curve)
-
-## Averaged on high-D
-df_bin_tsne_s_curve <- avg_highD_data(.data = df_all_tsne_s_curve)
-
-## Triangulate bin centroids
-tr1_object_tsne_s_curve <- triangulate_bin_centroids(df_bin_centroids_tsne_s_curve, x, y)
-tr_from_to_df_tsne_s_curve <- generate_edge_info(triangular_object = tr1_object_tsne_s_curve)
-
-# ggplot(df_bin_centroids_tsne_s_curve, aes(x = x, y = y)) +
-#   geom_segment(data = tr_from_to_df_tsne_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
-#   geom_point(size = 2, colour = "#33a02c") +
-#   coord_equal()
-
-## Compute 2D distances
-distance_tsne_s_curve <- cal_2D_dist(.data = tr_from_to_df_tsne_s_curve)
-
-## To find the benchmark value
-benchmark_tsne_s_curve <- find_benchmark_value(.data = distance_tsne_s_curve, distance_col = distance)
-
-# colour_long_edges(.data = distance_tsne_s_curve, benchmark_value = benchmark_tsne_s_curve,
-#                   triangular_object = tr1_object_tsne_s_curve, distance_col = distance)
-
-trimesh_removed_tsne_s_curve <- remove_long_edges(.data = distance_tsne_s_curve, benchmark_value = benchmark_tsne_s_curve,
-                                                  triangular_object = tr1_object_tsne_s_curve, distance_col = distance)
-
-trimesh_removed_tsne_s_curve <- trimesh_removed_tsne_s_curve +
-  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  # theme(axis.text = element_text(size = 5),
-  #       axis.title = element_text(size = 7)) +
-  geom_point(colour = "#33a02c", size = 0.5) +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-# tour_tsne_s_curve <- show_langevitour(df_all_tsne_s_curve, df_bin_tsne_s_curve, df_bin_centroids_tsne_s_curve, benchmark_value = benchmark_tsne_s_curve, distance = distance_tsne_s_curve, distance_col = distance)
-
-## UMAP
-
-num_bins_umap_s_curve <- 6
-shape_val_umap_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
-                                                          x = UMAP1, y = UMAP2) ## 1.259938
-## To extract bin centroids
-hexbin_data_object_umap_s_curve <- extract_hexbin_centroids(nldr_df = UMAP_s_curve, num_bins = num_bins_umap_s_curve, shape_val = shape_val_umap_s_curve, x = UMAP1, y = UMAP2)
-
-df_bin_centroids_umap_s_curve <- hexbin_data_object_umap_s_curve$hexdf_data
-
-UMAP_data_with_hb_id_s_curve <- UMAP_s_curve |>
-  dplyr::mutate(hb_id = hexbin_data_object_umap_s_curve$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_umap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), UMAP_data_with_hb_id_s_curve)
-
-## Averaged on high-D
-df_bin_umap_s_curve <- avg_highD_data(.data = df_all_umap_s_curve)
-
-## Triangulate bin centroids
-tr1_object_umap_s_curve <- triangulate_bin_centroids(df_bin_centroids_umap_s_curve, x, y)
-tr_from_to_df_umap_s_curve <- generate_edge_info(triangular_object = tr1_object_umap_s_curve)
-
-# ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
-#   geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
-#   geom_point(size = 2, colour = "#33a02c") +
-#   coord_equal()
-
-## Compute 2D distances
-distance_umap_s_curve <- cal_2D_dist(.data = tr_from_to_df_umap_s_curve)
-
-## To find the benchmark value
-benchmark_umap_s_curve <- find_benchmark_value(.data = distance_umap_s_curve, distance_col = distance)
-
-# colour_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve,
-#                   triangular_object = tr1_object_umap_s_curve, distance_col = distance)
-
-trimesh_removed_umap_s_curve <- remove_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve,
-                                     triangular_object = tr1_object_umap_s_curve, distance_col = distance)
-
-trimesh_removed_umap_s_curve <- trimesh_removed_umap_s_curve +
-  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  # theme(axis.text = element_text(size = 5),
-  #       axis.title = element_text(size = 7)) +
-  geom_point(colour = "#33a02c", size = 0.5) +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-# tour_umap_s_curve <- show_langevitour(df_all_umap_s_curve, df_bin_umap_s_curve, df_bin_centroids_umap_s_curve, benchmark_value = benchmark_umap_s_curve, distance = distance_umap_s_curve, distance_col = distance)
-
-## TriMAP
-
-num_bins_trimap_s_curve <- 6
-shape_val_trimap_s_curve <- calculate_effective_shape_value(.data = TriMAP_s_curve,
-                                                            x = TriMAP1, y = TriMAP2) ## 1.259938
-## To extract bin centroids
-hexbin_data_object_trimap_s_curve <- extract_hexbin_centroids(nldr_df = TriMAP_s_curve, num_bins = num_bins_trimap_s_curve, shape_val = shape_val_trimap_s_curve, x = TriMAP1, y = TriMAP2)
-
-df_bin_centroids_trimap_s_curve <- hexbin_data_object_trimap_s_curve$hexdf_data
-
-TriMAP_data_with_hb_id_s_curve <- TriMAP_s_curve |>
-  dplyr::mutate(hb_id = hexbin_data_object_trimap_s_curve$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_trimap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), TriMAP_data_with_hb_id_s_curve)
-
-## Averaged on high-D
-df_bin_trimap_s_curve <- avg_highD_data(.data = df_all_trimap_s_curve)
-
-## Triangulate bin centroids
-tr1_object_trimap_s_curve <- triangulate_bin_centroids(df_bin_centroids_trimap_s_curve, x, y)
-tr_from_to_df_trimap_s_curve <- generate_edge_info(triangular_object = tr1_object_trimap_s_curve)
-
-# ggplot(df_bin_centroids_trimap_s_curve, aes(x = x, y = y)) +
-#   geom_segment(data = tr_from_to_df_trimap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
-#   geom_point(size = 2, colour = "#33a02c") +
-#   coord_equal()
-
-## Compute 2D distances
-distance_trimap_s_curve <- cal_2D_dist(.data = tr_from_to_df_trimap_s_curve)
-
-## To find the benchmark value
-benchmark_trimap_s_curve <- find_benchmark_value(.data = distance_trimap_s_curve, distance_col = distance)
-
-# colour_long_edges(.data = distance_trimap_s_curve, benchmark_value = benchmark_trimap_s_curve,
-#                   triangular_object = tr1_object_trimap_s_curve, distance_col = distance)
-
-trimesh_removed_trimap_s_curve <- remove_long_edges(.data = distance_trimap_s_curve, benchmark_value = benchmark_trimap_s_curve,
-                                                    triangular_object = tr1_object_trimap_s_curve, distance_col = distance)
-
-trimesh_removed_trimap_s_curve <- trimesh_removed_trimap_s_curve +
-  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  # theme(axis.text = element_text(size = 5),
-  #       axis.title = element_text(size = 7)) +
-  geom_point(colour = "#33a02c", size = 0.25) +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-# tour_trimap_s_curve <- show_langevitour(df_all_trimap_s_curve, df_bin_trimap_s_curve, df_bin_centroids_trimap_s_curve, benchmark_value = benchmark_trimap_s_curve, distance = distance_trimap_s_curve, distance_col = distance)
-
-## PacMAP
-
-num_bins_pacmap_s_curve <- 10
-shape_val_pacmap_s_curve <- calculate_effective_shape_value(.data = PaCMAP_s_curve,
-                                                            x = PaCMAP1, y = PaCMAP2) ## 1.259938
-## To extract bin centroids
-hexbin_data_object_pacmap_s_curve <- extract_hexbin_centroids(nldr_df = PaCMAP_s_curve, num_bins = num_bins_pacmap_s_curve, shape_val = shape_val_pacmap_s_curve, x = PaCMAP1, y = PaCMAP2)
-
-df_bin_centroids_pacmap_s_curve <- hexbin_data_object_pacmap_s_curve$hexdf_data
-
-PaCMAP_data_with_hb_id_s_curve <- PaCMAP_s_curve |>
-  dplyr::mutate(hb_id = hexbin_data_object_pacmap_s_curve$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_pacmap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), PaCMAP_data_with_hb_id_s_curve)
-
-## Averaged on high-D
-df_bin_pacmap_s_curve <- avg_highD_data(.data = df_all_pacmap_s_curve)
-
-## Triangulate bin centroids
-tr1_object_pacmap_s_curve <- triangulate_bin_centroids(df_bin_centroids_pacmap_s_curve, x, y)
-tr_from_to_df_pacmap_s_curve <- generate_edge_info(triangular_object = tr1_object_pacmap_s_curve)
-
-# ggplot(df_bin_centroids_pacmap_s_curve, aes(x = x, y = y)) +
-#   geom_segment(data = tr_from_to_df_pacmap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
-#   geom_point(size = 2, colour = "#33a02c") +
-#   coord_equal()
-
-## Compute 2D distances
-distance_pacmap_s_curve <- cal_2D_dist(.data = tr_from_to_df_pacmap_s_curve)
-
-## To find the benchmark value
-benchmark_pacmap_s_curve <- find_benchmark_value(.data = distance_pacmap_s_curve, distance_col = distance)
-
-# colour_long_edges(.data = distance_pacmap_s_curve, benchmark_value = benchmark_pacmap_s_curve,
-#                   triangular_object = tr1_object_pacmap_s_curve, distance_col = distance)
-
-trimesh_removed_pacmap_s_curve <- remove_long_edges(.data = distance_pacmap_s_curve, benchmark_value = benchmark_pacmap_s_curve,
-                                     triangular_object = tr1_object_pacmap_s_curve, distance_col = distance)
-
-trimesh_removed_pacmap_s_curve <- trimesh_removed_pacmap_s_curve +
-  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  # theme(axis.text = element_text(size = 5),
-  #       axis.title = element_text(size = 7)) +
-  geom_point(colour = "#33a02c", size = 0.5) +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'e', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-# tour_pacmap_s_curve <- show_langevitour(df_all_pacmap_s_curve, df_bin_pacmap_s_curve, df_bin_centroids_pacmap_s_curve, benchmark_value = benchmark_pacmap_s_curve, distance = distance_pacmap_s_curve, distance_col = distance)
-
-## PHATE
-
-num_bins_phate_s_curve <- 8
-shape_val_phate_s_curve <- calculate_effective_shape_value(.data = PHATE_s_curve,
-                                                           x = PHATE1, y = PHATE2) ## 1.259938
-## To extract bin centroids
-hexbin_data_object_phate_s_curve <- extract_hexbin_centroids(nldr_df = PHATE_s_curve, num_bins = num_bins_phate_s_curve, shape_val = shape_val_phate_s_curve, x = PHATE1, y = PHATE2)
-
-df_bin_centroids_phate_s_curve <- hexbin_data_object_phate_s_curve$hexdf_data
-
-## Identify bins with low-density
-identify_rm_bins <- find_low_density_hexagons(df_bin_centroids_phate_s_curve, num_bins_phate_s_curve, benchmark_rm_hex = 0.06)
-
-df_bin_centroids_phate_s_curve <- df_bin_centroids_phate_s_curve |>
-  filter(!(hexID %in% identify_rm_bins))
-
-PHATE_data_with_hb_id_s_curve <- PHATE_s_curve |>
-  dplyr::mutate(hb_id = hexbin_data_object_phate_s_curve$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_phate_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), PHATE_data_with_hb_id_s_curve)
-
-## Averaged on high-D
-df_bin_phate_s_curve <- avg_highD_data(.data = df_all_phate_s_curve)
-
-## Triangulate bin centroids
-tr1_object_phate_s_curve <- triangulate_bin_centroids(df_bin_centroids_phate_s_curve, x, y)
-tr_from_to_df_phate_s_curve <- generate_edge_info(triangular_object = tr1_object_phate_s_curve)
-
-# ggplot(df_bin_centroids_phate_s_curve, aes(x = x, y = y)) +
-#   geom_segment(data = tr_from_to_df_phate_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
-#   geom_point(size = 2, colour = "#33a02c") +
-#   coord_equal()
-
-## Compute 2D distances
-distance_phate_s_curve <- cal_2D_dist(.data = tr_from_to_df_phate_s_curve)
-
-## To find the benchmark value
-benchmark_phate_s_curve <- find_benchmark_value(.data = distance_phate_s_curve, distance_col = distance)
-
-# colour_long_edges(.data = distance_phate_s_curve, benchmark_value = benchmark_phate_s_curve,
-#                   triangular_object = tr1_object_phate_s_curve, distance_col = distance)
-
-trimesh_removed_phate_s_curve <- remove_long_edges(.data = distance_phate_s_curve, benchmark_value = benchmark_phate_s_curve,
-                                                   triangular_object = tr1_object_phate_s_curve, distance_col = distance)
-
-trimesh_removed_phate_s_curve <- trimesh_removed_phate_s_curve +
-  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  # theme(axis.text = element_text(size = 5),
-  #       axis.title = element_text(size = 7)) +
-  geom_point(colour = "#33a02c", size = 0.5) +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-# tour_phate_s_curve <- show_langevitour(df_all_phate_s_curve, df_bin_phate_s_curve, df_bin_centroids_phate_s_curve, benchmark_value = benchmark_phate_s_curve, distance = distance_phate_s_curve, distance_col = distance)
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-#| label: fig-modelScurve
-#| fig-cap: Is there a good model to represent the original data in 2D space or are they all providing equivalent information?, (a) Model in the 2D space with tSNE (<https://youtu.be/uy9SkAo6gAE>), (b) Model in the 2D space with UMAP (<https://youtu.be/0MJDhHrh_Ug>), (c) Model in the 2D space with PHATE (<https://youtu.be/HbVv0uy0QWk>), (d) Model in the 2D space with TriMAP (<https://youtu.be/2OGwipAjzc8>), and (e) Model in the 2D space with PaCMAP (<https://youtu.be/pkokI8d-cBk>).
-#| out-width: 100%
-
-trimesh_removed_tsne_s_curve + trimesh_removed_umap_s_curve + trimesh_removed_phate_s_curve + trimesh_removed_trimap_s_curve + trimesh_removed_pacmap_s_curve +
-  #plot_annotation(tag_levels = 'a') +
-  plot_layout(ncol = 5) &
-  theme(legend.position='none', plot.tag = element_text(size = 8))
-```
-
 <!--### Model selection {#sec-selection}
@@
 ## Compute 2D distances
-distance_tsne_gau <- cal_2D_dist(.data = tr_from_to_df_tsne_gau)
+distance_tsne_gau <- cal_2d_dist(.data = tr_from_to_df_tsne_gau)
 
@@
 ## Compute 2D distances
-distance_umap_gau <- cal_2D_dist(.data = tr_from_to_df_umap_gau)
+distance_umap_gau <- cal_2d_dist(.data = tr_from_to_df_umap_gau)
 
@@
 ## Compute 2D distances
-distance_phate_gau <- cal_2D_dist(.data = tr_from_to_df_phate_gau)
+distance_phate_gau <- cal_2d_dist(.data = tr_from_to_df_phate_gau)
 
@@
 ## Compute 2D distances
-distance_trimap_gau <- cal_2D_dist(.data = tr_from_to_df_trimap_gau)
+distance_trimap_gau <- cal_2d_dist(.data = tr_from_to_df_trimap_gau)
 
@@
 ## Compute 2D distances
-distance_pacmap_gau <- cal_2D_dist(.data = tr_from_to_df_pacmap_gau)
+distance_pacmap_gau <- cal_2d_dist(.data = tr_from_to_df_pacmap_gau)
 
@@
 for (i in 1:length(num_bins_vec)) {
-
-  pred_df_training_object <- predict_hex_id(training_data = training_data_1, nldr_df = tSNE_data_gau, nldr_df_test = tSNE_data_gau, num_bins = num_bins_vec[i], shape_val = shape_value_gau, x = "tSNE1", y = "tSNE2", col_start = "x")
-  pred_df_training <- pred_df_training_object$pred_data
-  centroid_df_training <- pred_df_training_object$df_bin_centroids
-  avg_df_training <- pred_df_training_object$df_bin
+  model_object <- fit_high_d_model(training_data = training_data_1, nldr_df_with_id = tSNE_data_gau, x = "tSNE1", y = "tSNE2", num_bins_x = num_bins_vec[i], shape_val = shape_value_gau,
+                                 is_bin_centroid = TRUE,
+                                 is_rm_lwd_hex = FALSE,
+                                 benchmark_to_rm_lwd_hex = NA,
+                                 is_avg_high_d = TRUE, column_start_text = "x")
+
+  centroid_df_training <- model_object$df_bin_centroids
+  avg_df_training <- model_object$df_bin
+
+  # pred_df_training <- predict_2d_embeddings(test_data = training_data_1, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "tSNE")
+
+  pred_df_training <- predict_hex_id(df_bin_centroids = centroid_df_training, nldr_df_test = tSNE_data_gau, x = "tSNE1", y = "tSNE2")
 
@@
 
+  eval_df_training <- eval_df_training |>
+    mutate(num_bins_x = num_bins_vec[i])
+
   eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
-
 }
@@
 
-  pred_df_training_object <- predict_hex_id(training_data = training_data_1, nldr_df = UMAP_data_gau, nldr_df_test = UMAP_data_gau, num_bins = num_bins_vec[i], shape_val = shape_value_gau, x = "UMAP1", y = "UMAP2", col_start = "x")
-  pred_df_training <- pred_df_training_object$pred_data
-  centroid_df_training <- pred_df_training_object$df_bin_centroids
-  avg_df_training <- pred_df_training_object$df_bin
+  model_object <- fit_high_d_model(training_data = training_data_1, nldr_df_with_id = UMAP_data_gau, x = "UMAP1", y = "UMAP2", num_bins_x = num_bins_vec[i], shape_val = shape_value_gau,
+                                 is_bin_centroid = TRUE,
+                                 is_rm_lwd_hex = FALSE,
+                                 benchmark_to_rm_lwd_hex = NA,
+                                 is_avg_high_d = TRUE, column_start_text = "x")
+
+  centroid_df_training <- model_object$df_bin_centroids
+  avg_df_training <- model_object$df_bin
+
+  # pred_df_training <- predict_2d_embeddings(test_data = training_data_1, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "UMAP")
+
+  pred_df_training <- predict_hex_id(df_bin_centroids = centroid_df_training, nldr_df_test = UMAP_data_gau, x = "UMAP1", y = "UMAP2")
 
   eval_df_training <- generate_eval_df(data = df_2, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+
+  eval_df_training <- eval_df_training |>
+    mutate(num_bins_x = num_bins_vec[i])
 
@@
 
-  pred_df_training_object <- predict_hex_id(training_data = training_data_1, nldr_df = PHATE_data_gau, nldr_df_test = PHATE_data_gau, num_bins = num_bins_vec[i], shape_val = shape_value_gau, x = "PHATE1", y = "PHATE2", col_start = "x")
-  pred_df_training <- pred_df_training_object$pred_data
-  centroid_df_training <- pred_df_training_object$df_bin_centroids
-  avg_df_training <- pred_df_training_object$df_bin
+  model_object <- fit_high_d_model(training_data = training_data_1, nldr_df_with_id = PHATE_data_gau, x = "PHATE1", y = "PHATE2", num_bins_x = num_bins_vec[i], shape_val = shape_value_gau,
+                                 is_bin_centroid = TRUE,
+                                 is_rm_lwd_hex = FALSE,
+                                 benchmark_to_rm_lwd_hex = NA,
+                                 is_avg_high_d = TRUE, column_start_text = "x")
+
+  centroid_df_training <- model_object$df_bin_centroids
+  avg_df_training <- model_object$df_bin
+
+  # pred_df_training <- predict_2d_embeddings(test_data = training_data_1, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "PHATE")
+
+  pred_df_training <- predict_hex_id(df_bin_centroids = centroid_df_training, nldr_df_test = PHATE_data_gau, x = "PHATE1", y = "PHATE2")
 
   eval_df_training <- generate_eval_df(data = df_2, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+
+  eval_df_training <- eval_df_training |>
+    mutate(num_bins_x = num_bins_vec[i])
 
@@
 
-  pred_df_training_object <- predict_hex_id(training_data = training_data_1, nldr_df = TriMAP_data_gau, nldr_df_test = TriMAP_data_gau, num_bins = num_bins_vec[i], shape_val = shape_value_gau, x = "TriMAP1", y = "TriMAP2", col_start = "x")
-  pred_df_training <- pred_df_training_object$pred_data
-  centroid_df_training <- pred_df_training_object$df_bin_centroids
-  avg_df_training <- pred_df_training_object$df_bin
+  model_object <- fit_high_d_model(training_data = training_data_1, nldr_df_with_id = TriMAP_data_gau, x = "TriMAP1", y = "TriMAP2", num_bins_x = num_bins_vec[i], shape_val = shape_value_gau,
+                                 is_bin_centroid = TRUE,
+                                 is_rm_lwd_hex = FALSE,
+                                 benchmark_to_rm_lwd_hex = NA,
+                                 is_avg_high_d = TRUE, column_start_text = "x")
+
+  centroid_df_training <- model_object$df_bin_centroids
+  avg_df_training <- model_object$df_bin
+
+  # pred_df_training <- predict_2d_embeddings(test_data = training_data_1, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "TriMAP")
+
+  pred_df_training <- predict_hex_id(df_bin_centroids = centroid_df_training, nldr_df_test = TriMAP_data_gau, x = "TriMAP1", y = "TriMAP2")
 
   eval_df_training <- generate_eval_df(data = df_2, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+
+  eval_df_training <- eval_df_training |>
+    mutate(num_bins_x = num_bins_vec[i])
 
@@
 
-  pred_df_training_object <- predict_hex_id(training_data = training_data_1, nldr_df = PaCMAP_data_gau, nldr_df_test = PaCMAP_data_gau, num_bins = num_bins_vec[i], shape_val = shape_value_gau, x = "PaCMAP1", y = "PaCMAP2", col_start = "x")
-  pred_df_training <- pred_df_training_object$pred_data
-  centroid_df_training <- pred_df_training_object$df_bin_centroids
-  avg_df_training <- pred_df_training_object$df_bin
+  model_object <- fit_high_d_model(training_data = training_data_1, nldr_df_with_id = PaCMAP_data_gau, x = "PaCMAP1", y = "PaCMAP2", num_bins_x = num_bins_vec[i], shape_val = shape_value_gau,
+                                 is_bin_centroid = TRUE,
+                                 is_rm_lwd_hex = FALSE,
+                                 benchmark_to_rm_lwd_hex = NA,
+                                 is_avg_high_d = TRUE, column_start_text = "x")
+
+  centroid_df_training <- model_object$df_bin_centroids
+  avg_df_training <- model_object$df_bin
+
+  # pred_df_training <- predict_2d_embeddings(test_data = training_data_1, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "PaCMAP")
+
+  pred_df_training <- predict_hex_id(df_bin_centroids = centroid_df_training, nldr_df_test = PaCMAP_data_gau, x = "PaCMAP1", y = "PaCMAP2")
 
   eval_df_training <- generate_eval_df(data = df_2, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+
+  eval_df_training <- eval_df_training |>
+    mutate(num_bins_x = num_bins_vec[i])
 
@@
 ## Compute 2D distances
-distance_pbmc <- cal_2D_dist(.data = tr_from_to_df_pbmc)
+distance_pbmc <- cal_2d_dist(.data = tr_from_to_df_pbmc)
 
@@
 ## Compute 2D distances
-distance_tsne_harps <- cal_2D_dist(.data = tr_from_to_df_tsne_harps)
+distance_tsne_harps <- cal_2d_dist(.data = tr_from_to_df_tsne_harps)
 

### 2024-02-04 paper.tex
Tried 5-fold cross validation to compute mse and investigate the distribution (not specific change with cross validation)

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, sharp corners, interior hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}, breakable]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden, enhanced, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 shown in the \textbf{?@tbl-msebinsnldr}. Furthermore,
-Figure~\ref{fig-modelScurve} shows the 2D models constructed for each
-NLDR technique with the selected total number of bins and how the models
-look in high-D with original data.
+\textbf{?@fig-modelScurve} shows the 2D models constructed for each NLDR
+technique with the selected total number of bins and how the models look
+in high-D with original data.
 
@@
 case, a benchmark value of \(0.06\) was chosen. The resulting 2D model
-is shown in Figure~\ref{fig-modelScurve} (c), while the video link
+is shown in \textbf{?@fig-modelScurve} (c), while the video link
 provides a visual representation of how the model appears in high-D.
@@
 the S-curve (see videos of the models linked in
-Figure~\ref{fig-modelScurve}). But thorough examination of each model in
+\textbf{?@fig-modelScurve}). But thorough examination of each model in
 high-D space shows some surprising findings. The model of PHATE is
 flattened and lack of capturing the width of S-curve (see videos of the
-model with PHATE linked in Figure~\ref{fig-modelScurve}). This nature of
+model with PHATE linked in \textbf{?@fig-modelScurve}). This nature of
 PHATE is also evicted by the considerably higher AIC and MSE values for
@@
 TriMAP (see see videos of the model with TriMAP linked in
-Figure~\ref{fig-modelScurve}). This model shows how the patterns in 2D
+\textbf{?@fig-modelScurve}). This model shows how the patterns in 2D
 representation of TriMAP capture by the model (see
@@
 the model (see see videos of the model with UMAP linked in
-Figure~\ref{fig-modelScurve}). In the 2D layouts of tSNE and PaCMAP (see
+\textbf{?@fig-modelScurve}). In the 2D layouts of tSNE and PaCMAP (see
 Figure~\ref{fig-nldervis} (a), (e)), there are some sparse areas that
 are captured by the models effectively (see see videos of the model with
-tSNE and PaCMAP linked in Figure~\ref{fig-modelScurve}).
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-modelScurve-1.pdf}
-
-}
-
-\caption{\label{fig-modelScurve}Is there a good model to represent the
-original data in 2D space or are they all providing equivalent
-information?, (a) Model in the 2D space with tSNE
-(\url{https://youtu.be/uy9SkAo6gAE}), (b) Model in the 2D space with
-UMAP (\url{https://youtu.be/0MJDhHrh_Ug}), (c) Model in the 2D space
-with PHATE (\url{https://youtu.be/HbVv0uy0QWk}), (d) Model in the 2D
-space with TriMAP (\url{https://youtu.be/2OGwipAjzc8}), and (e) Model in
-the 2D space with PaCMAP (\url{https://youtu.be/pkokI8d-cBk}).}
-
-\end{figure}
+tSNE and PaCMAP linked in \textbf{?@fig-modelScurve}).
 

### 2024-02-04 2_cross_validation_with_total_num_bins.R

@@ -1,0 +1,117 @@
+####### S-curve example with 5-Fold cross validation #####################
+library(readr)
+
+set.seed(20240110)
+
+source("quollr_code.R", local = TRUE)
+
+## Import data
+data <- read_rds("data/s_curve/s_curve.rds")
+training_data <- read_rds("data/s_curve/s_curve_training.rds")
+UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap.rds")
+
+shape_val_umap_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
+                                                          x = UMAP1, y = UMAP2)
+
+# Create indices for 5-fold cross-validation
+num_obs <- NROW(UMAP_s_curve)
+fold_size <- floor(num_obs / 5)
+
+# Initialize variables to store performance metrics
+mse_list <- numeric(5)
+
+num_bins_vec <- 1:13
+
+vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations",
+                                     "total_error", "total_mse", "num_bins_x"))  ## Define column names
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+# Perform 5-fold cross-validation
+for (fold in 1:5) {
+  # Define the indices for the current fold
+  start_index <- (fold - 1) * fold_size + 1
+  end_index <- fold * fold_size
+
+  # Split the data into training and validation sets
+  X_train <- UMAP_s_curve[-(start_index:end_index), ]
+  X_val <- UMAP_s_curve[start_index:end_index, ]
+
+  for (i in 1:length(num_bins_vec)) {
+    # Fit the model for training
+    model_object <- fit_high_d_model(training_data = training_data, nldr_df_with_id = UMAP_s_curve,
+                                     x = "UMAP1", y = "UMAP2", num_bins_x = num_bins_vec[i], shape_val = shape_val_umap_s_curve,
+                                     is_bin_centroid = TRUE,
+                                     is_rm_lwd_hex = FALSE,
+                                     benchmark_to_rm_lwd_hex = NA,
+                                     is_avg_high_d = TRUE, column_start_text = "x")
+
+    centroid_df_training <- model_object$df_bin_centroids
+    avg_df_training <- model_object$df_bin
+
+    # Make predictions on the validation set
[…]

### 2024-02-05 paper.qmd
Add different parameter choices for HARPS data

@@
 
+#### Hexagonal grid starting point
+
+Mainly hexagonal grid starts at the lowest x and y values assign for the grid. Another possibility is to shift the grid. That's another interesting parameter to investigate.
+
+```{r}
+#| echo: false
+#| warning: false
+
+hex_full_count_df <- read_rds("data/s_curve/s_curve_hex_8.rds")
+
+shifted_hex_coord_df <- read_rds("data/s_curve/s_curve_hex_8_shifted_hex_coord_df.rds")
+
+ggplot(data = hex_full_count_df, aes(x = c_x, y = c_y)) +
+  geom_point(color = "black") +
+  geom_point(data = shifted_hex_coord_df, aes(x = c_x, y = c_y), color = "red")
+
+ggplot(data = shifted_hex_coord_df, aes(x = x, y = y)) +
+  geom_polygon(fill = NA, color = "#feb24c", aes(group = polygon_id)) +
+  geom_polygon(data = hex_full_count_df, aes(x = x, y = y, group = polygon_id),
+               fill = NA, color = "#bdbdbd")
+
+## Before shift
+ggplot(data = hex_full_count_df, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
+  geom_text(aes(x = c_x, y = c_y, label = hexID)) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff")
+
+## After shift
+ggplot(data = shifted_hex_coord_df, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
+  geom_text(aes(x = c_x, y = c_y, label = hexID)) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff")
+```
+Compute aic, mse after shifting and compare before and after shifting errors
+And investigate how aic, mse change according to the shifted amount
+
 ### Model summaries {#sec-summary}
@@
 
+```{r}
+#| warning: false
+#| echo: false
+
+error_df <- read_csv("data/harps/error_harps_tsne.csv", col_names = FALSE)
+
+error_df <- error_df |>
+  mutate(X2 = round(X2/100, 2),
+         X1 = as.integer(X1)) |>
+  arrange(X2)
+
+colnames(error_df) <- c("perplexity", "error (x 100)")
+
+kable(error_df, caption = "errors for different parameter combinations", label = "tbl-errordiffcombharps") |>
+  #xtable2kable() |>
+  kable_styling(latex_options = c("striped", "HOLD_position"), position = "center", full_width = FALSE) |>
+  row_spec(0, bold = TRUE) |>
+  row_spec(3, bold=T, color = "red")
+```
+
 <!-- Author suggestion-->
@@
 
+num_non_empty_bins_harps <- df_bin_centroids_tsne_harps$hexID |> length()
+
 ## Identify bins with low-density
@@
 
+df_bin_train <- df_bin_tsne_harps
+names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])
+
+error_df <- df_all_tsne_harps |>
+  dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates
+
+# prediction_df_join <- prediction_df_join |>
+#   dplyr::left_join(data, by = c("ID" = "ID")) ## Map high-D data
+
+for (i in 1:(NCOL(df_bin_train) - 1)) {
+
+  error_df[ , paste0("abs_residual_", "x", i)] <- abs(error_df[ , paste0("x", i)] - error_df[ , paste0("avg_", "x", i)])
+
+}
+
+error_df <- error_df |>
+  dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "x")))))
+
+error_df <- error_df |>
+  mutate(type = if_else(total <= 0, "no error",
+                        if_else(total <= 0.2, "error 0-0.2",
+                                if_else(total <= 0.3, "error 0.2-0.3",
+                                        if_else(total <= 0.5, "error 0.3-0.5",
+                                                if_else(total <= 1, "error 0.5-1",
+                                                        "error greter than 1"))))))
+
+plot_list2_harps_error <- error_df |>
+    mutate(type = factor(type , levels = c("no error", "error 0-0.2", "error 0.2-0.3", "error 0.3-0.5", "error 0.5-1", "error greter than 1"))) |>
+    ggplot(aes(x = tSNE1,
+               y = tSNE2, color = type))+
+    geom_point(alpha=0.5, size = 0.1) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#fee0d2", "#fcbba1",
+                                         "#fc9272", "#fb6a4a", "#ef3b2c",
+                                         "#cb181d", "#a50f15", "#99000d"))
+
 ```
@@
 
-plot_list2_harps + trimesh_removed_tsne_harps +
-  plot_layout(guides='collect', ncol = 2) &
+plot_list2_harps + trimesh_removed_tsne_harps + plot_list2_harps_error +
+  plot_layout(guides='collect', ncol = 3) &
   theme(legend.position='bottom', plot.tag = element_text(size = 8))
@@
 
-num_bins_tsne_harps <- calculate_effective_x_bins(.data = tSNE_data_harps, x = tSNE1,
-                                                cell_area = 1) #43
-num_bins_tsne_harps <- 30
-
 shape_val_tsne_harps <- calculate_effective_shape_value(.data = tSNE_data_harps,
                                                       x = tSNE1, y = tSNE2)
+
+num_bins_tsne_harps <- find_non_empty_bins(nldr_df = tSNE_data_harps, x = "tSNE1", y = "tSNE2", shape_val = shape_val_tsne_harps, non_empty_bins = num_non_empty_bins_harps)
+
 ## To extract bin centroids
@@
 benchmark_tsne_harps <- find_benchmark_value(.data = distance_tsne_harps, distance_col = distance)
+benchmark_tsne_harps <- 2.5
 
@@
 
-num_bins_tsne_harps <- calculate_effective_x_bins(.data = tSNE_data_harps, x = tSNE1,
-                                                cell_area = 1) #43
-#num_bins_tsne_harps <- 30
-
 shape_val_tsne_harps <- calculate_effective_shape_value(.data = tSNE_data_harps,
                                                       x = tSNE1, y = tSNE2)
+
+num_bins_tsne_harps <- find_non_empty_bins(nldr_df = tSNE_data_harps, x = "tSNE1", y = "tSNE2", shape_val = shape_val_tsne_harps, non_empty_bins = num_non_empty_bins_harps)
+
 ## To extract bin centroids
@@
 benchmark_tsne_harps <- find_benchmark_value(.data = distance_tsne_harps, distance_col = distance)
+benchmark_tsne_harps <- 1
 
@@
 
+df_bin_train <- df_bin_tsne_harps
+names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])
+
+error_df <- df_all_tsne_harps |>
+  dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates
+
+# prediction_df_join <- prediction_df_join |>
+#   dplyr::left_join(data, by = c("ID" = "ID")) ## Map high-D data
+
+for (i in 1:(NCOL(df_bin_train) - 1)) {
+
+  error_df[ , paste0("abs_residual_", "x", i)] <- abs(error_df[ , paste0("x", i)] - error_df[ , paste0("avg_", "x", i)])
+
+}
+
+error_df <- error_df |>
+  dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "x")))))
+
+error_df <- error_df |>
+  mutate(type = if_else(total <= 0, "no error",
+                        if_else(total <= 0.2, "error 0-0.2",
+                                if_else(total <= 0.3, "error 0.2-0.3",
+                                        if_else(total <= 0.5, "error 0.3-0.5",
+                                                if_else(total <= 1, "error 0.5-1",
+                                                        "error greter than 1"))))))
+
+plot_list2_harps_error <- error_df |>
+    mutate(type = factor(type , levels = c("no error", "error 0-0.2", "error 0.2-0.3", "error 0.3-0.5", "error 0.5-1", "error greter than 1"))) |>
+    ggplot(aes(x = tSNE1,
+               y = tSNE2, color = type))+
+    geom_point(alpha=0.5, size = 0.1) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#fee0d2", "#fcbba1",
+                                         "#fc9272", "#fb6a4a", "#ef3b2c",
+                                         "#cb181d", "#a50f15", "#99000d"))
 ```

### 2024-02-05 paper.tex
Add different parameter choices for HARPS data

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden, enhanced, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, sharp corners, borderline west={3pt}{0pt}{shadecolor}, interior hidden, boxrule=0pt, frame hidden, breakable]}{\end{tcolorbox}}\fi
 
@@
 \end{figure}
+
+\hypertarget{hexagonal-grid-starting-point}{%
+\subsubsection{Hexagonal grid starting
+point}\label{hexagonal-grid-starting-point}}
+
+Mainly hexagonal grid starts at the lowest x and y values assign for the
+grid. Another possibility is to shift the grid. That's another
+interesting parameter to investigate.
+
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-39-1.pdf}
+
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-39-2.pdf}
+
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-39-3.pdf}
+
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-39-4.pdf}
+
+Compute aic, mse after shifting and compare before and after shifting
+errors And investigate how aic, mse change according to the shifted
+amount
 
@@
 
+\begin{table}[H]
+
+\caption{errors for different parameter combinations}
+\centering
+\begin{tabular}[t]{r|r}
+\hline
+\textbf{perplexity} & \textbf{error (x 100)}\\
+\hline
+\cellcolor{gray!6}{23} & \cellcolor{gray!6}{0.02}\\
+\hline
+91 & 0.04\\
+\hline
+\textcolor{red}{\textbf{\cellcolor{gray!6}{40}}} & \textcolor{red}{\textbf{\cellcolor{gray!6}{1.05}}}\\
+\hline
+\end{tabular}
+\end{table}
+
 \begin{figure}[h]

### 2024-02-05 2_cross_validation_with_total_num_bins.R

@@ -41,8 +41,10 @@ for (fold in 1:5) {
 
   for (i in 1:length(num_bins_vec)) {
     # Fit the model for training
-    model_object <- fit_high_d_model(training_data = training_data, nldr_df_with_id = UMAP_s_curve,
-                                     x = "UMAP1", y = "UMAP2", num_bins_x = num_bins_vec[i], shape_val = shape_val_umap_s_curve,
+    model_object <- fit_high_d_model(training_data = training_data,
+                                     nldr_df_with_id = UMAP_s_curve,
+                                     x = "UMAP1", y = "UMAP2", num_bins_x = num_bins_vec[i],
+                                     shape_val = shape_val_umap_s_curve,
                                      is_bin_centroid = TRUE,
                                      is_rm_lwd_hex = FALSE,
                                      benchmark_to_rm_lwd_hex = NA,

### 2024-02-05 paper.qmd
Add hexagonal grid starting point section

@@
 ##| out-width: 100%
+#| fig-pos: H
 
@@
 #| label: fig-rmlowdenshex
+#| fig-pos: H
 ##| out-width: 100%
@@
 
-#### Hexagonal grid starting point
-
-Mainly hexagonal grid starts at the lowest x and y values assign for the grid. Another possibility is to shift the grid. That's another interesting parameter to investigate.
+#### Hexagonal grid origin
+
+According to @Dan2023, the hexagonal binning is done by tessellating the $xy$ plane over the set (range($x$), range($y$)) (see @fig-scurveshifthexgridsexp (b)).
+In that case, bin centroids are defined as shown in @fig-scurveshifthexgridsexp (a) with gray colour.
+Rather than sticking to the typical hexagonal grid, introducing a meaningful shift in both the x and y directions presents an opportunity for an improved 2D model.
+Therefore, investigating this shift in the hexagonal grid is an important parameter to consider.
+
+As shown in @fig-scurveshifthexgrids, the shifting influences the distribution of points and number of non-empty bins, impacting the resulting 2D model.
+According to @fig-diagnosticpltScurvehexbins, the 2D model with a total of $144$ bins applied to S-curve UMAP data does not require any shifting because the lowest MSE occurs when no shift is introduced.
 
@@
 
-ggplot(data = hex_full_count_df, aes(x = c_x, y = c_y)) +
-  geom_point(color = "black") +
-  geom_point(data = shifted_hex_coord_df, aes(x = c_x, y = c_y), color = "red")
-
-ggplot(data = shifted_hex_coord_df, aes(x = x, y = y)) +
+bin_centroids_shift <- ggplot(data = hex_full_count_df, aes(x = c_x, y = c_y)) +
+  geom_point(color = "#bdbdbd") +
+  geom_point(data = shifted_hex_coord_df, aes(x = c_x, y = c_y), color = "#feb24c") +
+  coord_cartesian(xlim = c(-5, 8), ylim = c(-10, 10)) +
+  theme_void() +
+  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6)) +
+  guides(fill = guide_colourbar(title = "Standardized count")) +
+  annotate(geom = 'text', label = "a", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
+
+hex_grid_shift <- ggplot(data = shifted_hex_coord_df, aes(x = x, y = y)) +
   geom_polygon(fill = NA, color = "#feb24c", aes(group = polygon_id)) +
   geom_polygon(data = hex_full_count_df, aes(x = x, y = y, group = polygon_id),
-               fill = NA, color = "#bdbdbd")
+               fill = NA, color = "#bdbdbd") +
+  coord_cartesian(xlim = c(-5, 8), ylim = c(-10, 10)) +
+  theme_void() +
+  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6)) +
+  guides(fill = guide_colourbar(title = "Standardized count")) +
+  annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
 
 ## Before shift
-ggplot(data = hex_full_count_df, aes(x = x, y = y)) +
+before_shift_plot <- ggplot(data = hex_full_count_df, aes(x = x, y = y)) +
   geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
-  geom_text(aes(x = c_x, y = c_y, label = hexID)) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff")
+  geom_text(aes(x = c_x, y = c_y, label = hexID), size = 2) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  coord_equal() +
+  theme_void() +
+  theme(legend.position="bottom", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6)) +
+  guides(fill = guide_colourbar(title = "Standardized count")) +
+  annotate(geom = 'text', label = "a", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
 
 ## After shift
-ggplot(data = shifted_hex_coord_df, aes(x = x, y = y)) +
+after_shift_plot <- ggplot(data = shifted_hex_coord_df, aes(x = x, y = y)) +
   geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
-  geom_text(aes(x = c_x, y = c_y, label = hexID)) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff")
-```
-Compute aic, mse after shifting and compare before and after shifting errors
-And investigate how aic, mse change according to the shifted amount
+  geom_text(aes(x = c_x, y = c_y, label = hexID), size = 2) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  coord_equal() +
+  theme_void() +
+  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6)) +
+  guides(fill = guide_colourbar(title = "Standardized count")) +
+  annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "(a) Visualization of bin centroids and (b) hexagonal grid, showcasing the configuration before (colored in gray) and after (colored in light orange) shifting the starting point by an identical amount applied in both x and y directions, with a shift value of $0.537285$."
+#| label: fig-scurveshifthexgridsexp
+##| out-width: 100%
+#| fig-pos: H
+
+bin_centroids_shift + hex_grid_shift +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='none', plot.tag = element_text(size = 8))
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "Hexbin plots shows the distribution of points before and after shifting the starting point of the hexagonal grid generated for UMAP applied to the training S-curve dataset.
+(a) Hexbin plot before shifting, and (b) Hexbin plot after shifting (identical shift applied in both x and y directions, with a shift value of $0.537285$)."
+#| label: fig-scurveshifthexgrids
+##| out-width: 100%
+#| fig-pos: H
+
+before_shift_plot + after_shift_plot +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='none', plot.tag = element_text(size = 8))
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+#MSE_df <- dplyr::bind_rows(MSE_df_1, MSE_df_2, MSE_df_3, MSE_df_4, MSE_df_5)
+
+#MSE_df$method <- factor(MSE_df$method, levels = c("tSNE", "UMAP", "PHATE", "TriMAP", "PaCMAP"))
+
+MSE_df <- read_rds("data/s_curve/s_curve_umap_shift_summary.rds") 
+MSE_df$method <- factor(MSE_df$method, levels = c("UMAP"))
+
+## To draw with AIC
+aic_plot <- ggplot(MSE_df |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "UMAP"), aes(x = shift,
+                                                                                 y = total_error,
+                                                                                 color = method
+)) +
+  geom_point() +
+  geom_line() +
+  geom_vline(xintercept = 0, linetype="dashed",
+                color = "red", size=0.5) +
+  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("AIC") +
+  xlab("shift value")
+## Effective number of bins along x-axis
+
+mse_plot <- ggplot(MSE_df |> dplyr::filter(method == "UMAP"), aes(x = shift,
+                                       y = total_mse,
+                                       color = method
+)) +
+  geom_point() +
+  geom_line() +
+   geom_vline(xintercept = 0, linetype="dashed",
+                color = "red", size=0.5) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("MSE") +
+  xlab("shift value")
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset with different shift values.
+In this illustration, an identical shift is applied in both the x and y directions.
+The Mean Squared Error (MSE) plot exhibits a global minimum at $0$, indicating that no additional shifting is necessary for the current hexagonal grid to yield a well-fitted model."
+#| label: fig-diagnosticpltScurvehexbins
+##| out-width: 100%
+#| fig-pos: H
+
+aic_plot + mse_plot +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='none', plot.tag = element_text(size = 8))
+```
 

### 2024-02-05 paper.tex
Add hexagonal grid starting point section

--- 2024-02-04 paper.tex
@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden, enhanced, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, sharp corners, breakable, interior hidden, enhanced]}{\end{tcolorbox}}\fi
 
@@
 
-\begin{figure}
+\begin{figure}[H]
 
@@
 eliminates long edges.}
+
+\end{figure}
+
+\hypertarget{hexagonal-grid-origin}{%
+\subsubsection{Hexagonal grid origin}\label{hexagonal-grid-origin}}
+
+According to \citet{Dan2023}, the hexagonal binning is done by
+tessellating the \(xy\) plane over the set (range(\(x\)), range(\(y\)))
+(see Figure~\ref{fig-scurveshifthexgridsexp} (b)). In that case, bin
+centroids are defined as shown in
+Figure~\ref{fig-scurveshifthexgridsexp} (a) with gray colour. Rather
+than sticking to the typical hexagonal grid, introducing a meaningful
+shift in both the x and y directions presents an opportunity for an
+improved 2D model. Therefore, investigating this shift in the hexagonal
+grid is an important parameter to consider.
+
+As shown in Figure~\ref{fig-scurveshifthexgrids}, the shifting
+influences the distribution of points and number of non-empty bins,
+impacting the resulting 2D model. According to
+Figure~\ref{fig-diagnosticpltScurvehexbins}, the 2D model with a total
+of \(144\) bins applied to S-curve UMAP data does not require any
+shifting because the lowest MSE occurs when no shift is introduced.
+
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-scurveshifthexgridsexp-1.pdf}
+
+}
+
+\caption{\label{fig-scurveshifthexgridsexp}(a) Visualization of bin
+centroids and (b) hexagonal grid, showcasing the configuration before
+(colored in gray) and after (colored in light orange) shifting the
+starting point by an identical amount applied in both x and y
+directions, with a shift value of \(0.537285\).}
+
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-scurveshifthexgrids-1.pdf}
+
+}
+
+\caption{\label{fig-scurveshifthexgrids}Hexbin plots shows the
+distribution of points before and after shifting the starting point of
+the hexagonal grid generated for UMAP applied to the training S-curve
+dataset. (a) Hexbin plot before shifting, and (b) Hexbin plot after
+shifting (identical shift applied in both x and y directions, with a
+shift value of \(0.537285\)).}
+
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-diagnosticpltScurvehexbins-1.pdf}
+
+}
+
+\caption{\label{fig-diagnosticpltScurvehexbins}Goodness of fit
+statistics from UMAP applied to training S-curve dataset with different
+shift values. In this illustration, an identical shift is applied in
+both the x and y directions. The Mean Squared Error (MSE) plot exhibits
+a global minimum at \(0\), indicating that no additional shifting is
+necessary for the current hexagonal grid to yield a well-fitted model.}
 

### 2024-02-07 appendix.qmd
Change names for columns and CHECK DONE

@@
 -->
+
+<!--
+### Single-Cell Tagged Reverse Transcription sequencing data of mouse
+
+The Zeisel mouse brain dataset, obtained through Spatial Transcriptomics (STRT-Seq).
+Within this dataset, information is collected from a substantial 2,816 individual mouse brain cells.
+Each of these cells acts as a molecular snapshot, capturing the distinctive genetic activity within various cell types.
+This diversity spans neurons, glial cells, and other essential components of the brain, offering a comprehensive view of the cellular tapestry.
+
+What makes this dataset particularly valuable is its ability to shed light on the spatial distribution of cells.
+Researchers can explore how gene expression patterns vary across different regions of the mouse brain, unlocking insights into the functional specialization of these regions and the intricate networks that underpin neural processes.
+
+```{r}
+#| warning: false
+#| echo: false
+
+## Import data
+df_2 <- read_rds("data/zeisel/zeisel.rds")
+training_data_1 <- read_rds("data/zeisel/zeisel_training.rds")
+test_1 <- read_rds("data/zeisel/zeisel_test.rds")
+
+tSNE_zeisel <- read_rds("data/zeisel/zeisel_tsne_30.rds")
+UMAP_zeisel <- read_rds("data/zeisel/zeisel_umap.rds")
+PHATE_zeisel <- read_rds("data/zeisel/zeisel_phate.rds")
+TriMAP_zeisel <- read_rds("data/zeisel/zeisel_trimap.rds")
+PaCMAP_zeisel <- read_rds("data/zeisel/zeisel_pacmap.rds")
+
+## Visualise embeddings
+
+plot_list1_zeisel <- plot_tSNE_2D(tSNE_zeisel) + #ggtitle("(a)") + 
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+plot_list2_zeisel <- plot_UMAP_2D(UMAP_zeisel) + #ggtitle("(b)") + 
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+plot_list3_zeisel <- plot_PHATE_2D(PHATE_zeisel) + #ggtitle("(c)") + 
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+plot_list4_zeisel <- plot_TriMAP_2D(TriMAP_zeisel) + #ggtitle("(d)") + 
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+plot_list5_zeisel <- plot_PaCMAP_2D(PaCMAP_zeisel) + #ggtitle("(e)") + 
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'e', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "2D layouts from different NLDR techniques applied for the training Zeisel mouse brain dataset: (a) tSNE (perplexity = 30), (b) UMAP (n_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2).
+Is there a best representation of the original data or are they all providing  equivalent information?"
+#| label: fig-nldervis5Mouse
+#| out-width: 100%
+
+plot_list1_zeisel + plot_list2_zeisel + plot_list3_zeisel + plot_list4_zeisel + plot_list5_zeisel +
+  plot_layout(ncol=5)
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+MSE_df_zei <- read_rds("data/zeisel/summary_zei.rds")
+MSE_df_zei$method <- factor(MSE_df_zei$method, levels = c("tSNE", "UMAP", "PHATE", "TriMAP", "PaCMAP"))
+
+## To draw with AIC
+aic_plot_zei <- ggplot(MSE_df_zei |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
+                                                                                 y = total_error,
+                                                                                 color = method
+)) +
+  geom_point() +
+  geom_line() +
+  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  theme_light() +
+  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  scale_colour_manual(values = c("#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("AIC") +
+  xlab("Total number of bins")
+## Effective number of bins along x-axis
+
+mse_plot_zei <- ggplot(MSE_df_zei, aes(x = number_of_bins,
+                                       y = total_mse,
+                                       color = method
+)) +
+  geom_point() +
+  geom_line() +
+  theme_light() +
+  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  scale_colour_manual(values = c("#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("MSE") +
+  xlab("Total number of bins")
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: Goodness of fit statistics from different NLDR techniques applied to training Zeisel mouse brain dataset. What is the best NLDR technique to represent the original data in 2D?
+#| label: fig-diagnosticpltZEI
+#| out-width: 100%
+
+aic_plot_zei + mse_plot_zei +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='bottom', plot.tag = element_text(size = 8))
+```
+
+### Retina dataset
+
+Human retina: https://www.ebi.ac.uk/biostudies/arrayexpress/studies/E-MTAB-7316
+
+https://www.cell.com/cms/10.1016/j.cell.2015.05.002/attachment/fd52950c-fd35-401a-9e60-ca7588ef9001/mmc1.pdf
+
+https://www.cell.com/fulltext/S0092-8674(15)00549-8#app3
+
+32 PCs
+tSNE (perplexity 30)
+
+### Kang dataset
+
+For the Kang lupus dataset (29), we obtained 16 cell types based on clustering of the Festem-
+selected genes (Fig. 4E and Fig. S10, S11), including 2 fine cell types that were often missed by
+other methods. The 2 cell types were IFNhi CD14+ monocytes with high expression of type I
+interferon-related (IFN) genes, and HSP+ CD4 T cells with up-regulated heat-
+shock-protein-related (HSP) genes. The HSP+ CD4 T cells might result from
+cryopreservation and thawing of T cells (37). Other than devianceFS, other available methods
+could not identify these HSP+ CD4 T cells (Fig. S11).
+
+```{r}
+#| warning: false
+#| echo: false
+
+## Import data
+#df_2 <- read_rds("data/pbmc/pbmc.rds")
+# training_data_1 <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_pca_50.rds")
+# training_data_1 <- training_data_1[, 1:25]
+#test_1 <- read_rds("data/pbmc/pbmc_test.rds")
+
+#tSNE_pbmc <- read_rds("data/pbmc/pbmc_tsne_30.rds")
+tSNE_kang <- read_rds("data/kang/kang_tsne.rds") |>
+  mutate(cell_label = if_else(cell_label %in% c("0", "3", "5"), "0", cell_label))
+# PHATE_pbmc <- read_rds("data/pbmc/pbmc_phate.rds")
+# TriMAP_pbmc <- read_rds("data/pbmc/pbmc_trimap.rds")
+# PaCMAP_pbmc <- read_rds("data/pbmc/pbmc_pacmap.rds")
+
+#load("data/pbmc/pbmc_3k_festem/pbmc3k_label.RData")
+
+## Visualise embeddings
+
+# plot_list1_pbmc <- plot_tSNE_2D(tSNE_pbmc) + #ggtitle("(a)") + 
+#   theme_linedraw() +
+#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+#         axis.title.x = element_blank(), axis.title.y = element_blank(),
+#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+#   annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+class_avg <- tSNE_kang |>
+  group_by(cell_label) |>
+  summarise(
+    tSNE1 = median(tSNE1),
+    tSNE2 = median(tSNE2)
+  ) |>
+  mutate(cell_label = as.numeric(cell_label)) |>
+  arrange(cell_label)
+
+class_avg$cell_label <- as.factor(class_avg$cell_label)
+
+levels(class_avg$cell_label) <- c("CD14 Monocyte", "CD4 \n Memory T", "CD4 Naive T", "CD16 Monocyte", "B", "T Activated", "CD8 T", "NK", "DC", "B \nActivated", "T cell:\nmonocyte \nComplex", "HSP+ CD4 T", "IFN^hi CD14 \n Monocyte", "CD34 \n Progenitors", "pDC")
+
+plot_list2_kang <- tSNE_kang |>
+    ggplot(aes(x = tSNE1,
+               y = tSNE2, color = cell_label))+
+    geom_point(alpha=0.5) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#b15928", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#cab2d6", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#bc80bd", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
+  geom_text(aes(x=tSNE1,y = tSNE2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) #+
+  # guides(color = guide_legend(title = "Cell type"),
+  #        color = guide_legend(nrow = 3)) #+
+  # annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+# plot_list3_pbmc <- plot_PHATE_2D(PHATE_pbmc) + #ggtitle("(c)") + 
+#   theme_linedraw() +
+#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+#         axis.title.x = element_blank(), axis.title.y = element_blank(),
+#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+#   annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+# plot_list4_pbmc <- plot_TriMAP_2D(TriMAP_pbmc) + #ggtitle("(d)") + 
+#   theme_linedraw() +
+#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+#         axis.title.x = element_blank(), axis.title.y = element_blank(),
+#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+#   annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+# 
+# plot_list5_pbmc <- plot_PaCMAP_2D(PaCMAP_pbmc) + #ggtitle("(e)") + 
+#   theme_linedraw() +
+#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
+#         axis.title.x = element_blank(), axis.title.y = element_blank(),
+#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+#   annotate(geom = 'text', label = 'e', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "2D layout from tSNE (perplexity = 30) applied for the Kang dataset. Is this a best representation of the original data?"
+#| label: fig-nlderviskangtSNE
+
+plot_list2_kang
+
+#  plot_list1_pbmc + plot_list3_pbmc + plot_list4_pbmc + plot_list5_pbmc +
+#   plot_layout(ncol=5) #| out-width: 100%
+
+# 2D layouts from different NLDR techniques applied for the training PBMC dataset: (a) tSNE (perplexity = 30), (b) UMAP (n_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2).
+Is there a best representation of the original data or are they all providing  equivalent information?
+```
+-->
+
+<!--
+### Single-cell RNA-seq data of human
+
+https://github.com/XiDsLab/Festem_paper/tree/main
+
+@Chen2023
+
+https://satijalab.org/seurat/articles/pbmc3k_tutorial.html
+https://www.cell.com/fulltext/S0092-8674(15)00549-8
+
+In the field of single-cell studies, a common analysis task involves clustering to identify groups of cells with similar expression profiles.
+Analysts often turn to NLDR techniques to verify and identify these clusters and explore developmental trajectories (e.g., example 1).
+In clustering workflows, the main objective is to verify the existence of clusters and subsequently identify them as specific cell types by examining the expression of "known" marker genes.
+In this context, a "faithful" embedding should ideally preserve the topology of the data, ensuring that cells corresponding to the same cell type are situated close to the high-dimensional space.
+
+To begin our analysis, we installed the Peripheral Blood Mononuclear Cells (pbmc) data set obtained from 10x Genomics using the `SeuratData` R package [@Rahul2019], which facilitates the distribution of data sets in the form of Seurat objects [@Yuhan2021].
+This data set contains 13,714 features across 2,700 samples within a single assay.
+The active assay is RNA, with 13,714 features representing different gene expressions.
+After loading the data set, we obtained the principal components (PCs) and assessed the variance explained by each PC.
+Based on this evaluation, we selected seven PCs, representing approximately 50% of the variance in gene expression, for further analysis.
+
+Next, we employed the UMAP technique with default parameter settings.
+As illustrated in @fig-pbmc, the cell types B and Platelet are well-separated in the UMAP layout.
+Moreover, CD14+ Mono, FCGR3A+ Mono, and DC form a distinct cluster, while Naive CD4 T, NK, Memory CD4 T, and CD8 T are grouped together in another cluster.
+The values utilized to construct the smooth low-dimensional manifold are presented in @tbl-table02.
+The linked video, demonstrating the tour with the model, showcases the generation of a smooth surface for this application, enabling a comprehensive exploration of the data's structure and relationships (see @fig-pbmc_sc).
+
+```{r}
+#| warning: false
+#| echo: false
+
+MSE_df_pbmc <- read_rds("data/pbmc/summary_pbmc.rds")
+MSE_df_pbmc$method <- factor(MSE_df_pbmc$method, levels = c("tSNE", "UMAP", "PHATE", "TriMAP", "PaCMAP"))
+
+## To draw with AIC
+aic_plot_pbmc <- ggplot(MSE_df_pbmc |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
+                                                                                 y = total_error,
+                                                                                 color = method
+)) +
+  geom_point() +
+  geom_line() +
+  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  theme_light() +
+  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  scale_colour_manual(values = c("#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("AIC") +
+  xlab("Total number of bins")
+## Effective number of bins along x-axis
+
+mse_plot_pbmc <- ggplot(MSE_df_pbmc, aes(x = number_of_bins,
+                                       y = total_mse,
+                                       color = method
+)) +
+  geom_point() +
+  geom_line() +
+  theme_light() +
+  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  scale_colour_manual(values = c("#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("MSE") +
+  xlab("Total number of bins")
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: Goodness of fit statistics from different NLDR techniques applied to training PBMC dataset. What is the best NLDR technique to represent the original data in 2D?
+#| label: fig-diagnosticpltPBMC
+#| out-width: 100%
+
+aic_plot_pbmc + mse_plot_pbmc +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='bottom', plot.tag = element_text(size = 8))
+```
+-->

### 2024-02-07 paper.qmd
Change names for columns and CHECK DONE

@@
 
-num_non_empty_bins_pbmc <- df_bin_centroids_pbmc$hexID |> length()
+num_non_empty_bins_pbmc <- df_bin_centroids_pbmc$hexID |> length() #83
 
@@
 
-<!--
-### Single-cell RNA-seq data of human
-
-https://github.com/XiDsLab/Festem_paper/tree/main
-
-@Chen2023
-
-https://satijalab.org/seurat/articles/pbmc3k_tutorial.html
-https://www.cell.com/fulltext/S0092-8674(15)00549-8
-
-In the field of single-cell studies, a common analysis task involves clustering to identify groups of cells with similar expression profiles.
-Analysts often turn to NLDR techniques to verify and identify these clusters and explore developmental trajectories (e.g., example 1).
-In clustering workflows, the main objective is to verify the existence of clusters and subsequently identify them as specific cell types by examining the expression of "known" marker genes.
-In this context, a "faithful" embedding should ideally preserve the topology of the data, ensuring that cells corresponding to the same cell type are situated close to the high-dimensional space.
-
-To begin our analysis, we installed the Peripheral Blood Mononuclear Cells (pbmc) data set obtained from 10x Genomics using the `SeuratData` R package [@Rahul2019], which facilitates the distribution of data sets in the form of Seurat objects [@Yuhan2021].
-This data set contains 13,714 features across 2,700 samples within a single assay.
-The active assay is RNA, with 13,714 features representing different gene expressions.
-After loading the data set, we obtained the principal components (PCs) and assessed the variance explained by each PC.
-Based on this evaluation, we selected seven PCs, representing approximately 50% of the variance in gene expression, for further analysis.
-
-Next, we employed the UMAP technique with default parameter settings.
-As illustrated in @fig-pbmc, the cell types B and Platelet are well-separated in the UMAP layout.
-Moreover, CD14+ Mono, FCGR3A+ Mono, and DC form a distinct cluster, while Naive CD4 T, NK, Memory CD4 T, and CD8 T are grouped together in another cluster.
-The values utilized to construct the smooth low-dimensional manifold are presented in @tbl-table02.
-The linked video, demonstrating the tour with the model, showcases the generation of a smooth surface for this application, enabling a comprehensive exploration of the data's structure and relationships (see @fig-pbmc_sc).
-
-```{r}
-#| warning: false
-#| echo: false
-
-MSE_df_pbmc <- read_rds("data/pbmc/summary_pbmc.rds")
-MSE_df_pbmc$method <- factor(MSE_df_pbmc$method, levels = c("tSNE", "UMAP", "PHATE", "TriMAP", "PaCMAP"))
-
-## To draw with AIC
-aic_plot_pbmc <- ggplot(MSE_df_pbmc |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
-                                                                                 y = total_error,
-                                                                                 color = method
-)) +
-  geom_point() +
-  geom_line() +
-  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  theme_light() +
-  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  scale_colour_manual(values = c("#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("AIC") +
-  xlab("Total number of bins")
-## Effective number of bins along x-axis
-
-mse_plot_pbmc <- ggplot(MSE_df_pbmc, aes(x = number_of_bins,
-                                       y = total_mse,
-                                       color = method
-)) +
-  geom_point() +
-  geom_line() +
-  theme_light() +
-  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  scale_colour_manual(values = c("#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("MSE") +
-  xlab("Total number of bins")
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: Goodness of fit statistics from different NLDR techniques applied to training PBMC dataset. What is the best NLDR technique to represent the original data in 2D?
-#| label: fig-diagnosticpltPBMC
-#| out-width: 100%
-
-aic_plot_pbmc + mse_plot_pbmc +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', ncol = 2) &
-  theme(legend.position='bottom', plot.tag = element_text(size = 8))
-```
--->
-
 ### HARPS GTO sample dataset
@@
 
-<!--
-### Single-Cell Tagged Reverse Transcription sequencing data of mouse
-
-The Zeisel mouse brain dataset, obtained through Spatial Transcriptomics (STRT-Seq).
-Within this dataset, information is collected from a substantial 2,816 individual mouse brain cells.
-Each of these cells acts as a molecular snapshot, capturing the distinctive genetic activity within various cell types.
-This diversity spans neurons, glial cells, and other essential components of the brain, offering a comprehensive view of the cellular tapestry.
-
-What makes this dataset particularly valuable is its ability to shed light on the spatial distribution of cells.
-Researchers can explore how gene expression patterns vary across different regions of the mouse brain, unlocking insights into the functional specialization of these regions and the intricate networks that underpin neural processes.
-
-```{r}
-#| warning: false
-#| echo: false
-
-## Import data
-df_2 <- read_rds("data/zeisel/zeisel.rds")
-training_data_1 <- read_rds("data/zeisel/zeisel_training.rds")
-test_1 <- read_rds("data/zeisel/zeisel_test.rds")
-
-tSNE_zeisel <- read_rds("data/zeisel/zeisel_tsne_30.rds")
-UMAP_zeisel <- read_rds("data/zeisel/zeisel_umap.rds")
-PHATE_zeisel <- read_rds("data/zeisel/zeisel_phate.rds")
-TriMAP_zeisel <- read_rds("data/zeisel/zeisel_trimap.rds")
-PaCMAP_zeisel <- read_rds("data/zeisel/zeisel_pacmap.rds")
-
-## Visualise embeddings
-
-plot_list1_zeisel <- plot_tSNE_2D(tSNE_zeisel) + #ggtitle("(a)") + 
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-plot_list2_zeisel <- plot_UMAP_2D(UMAP_zeisel) + #ggtitle("(b)") + 
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-plot_list3_zeisel <- plot_PHATE_2D(PHATE_zeisel) + #ggtitle("(c)") + 
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-plot_list4_zeisel <- plot_TriMAP_2D(TriMAP_zeisel) + #ggtitle("(d)") + 
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-plot_list5_zeisel <- plot_PaCMAP_2D(PaCMAP_zeisel) + #ggtitle("(e)") + 
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'e', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "2D layouts from different NLDR techniques applied for the training Zeisel mouse brain dataset: (a) tSNE (perplexity = 30), (b) UMAP (n_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2).
-Is there a best representation of the original data or are they all providing  equivalent information?"
-#| label: fig-nldervis5Mouse
-#| out-width: 100%
-
-plot_list1_zeisel + plot_list2_zeisel + plot_list3_zeisel + plot_list4_zeisel + plot_list5_zeisel +
-  plot_layout(ncol=5)
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-MSE_df_zei <- read_rds("data/zeisel/summary_zei.rds")
-MSE_df_zei$method <- factor(MSE_df_zei$method, levels = c("tSNE", "UMAP", "PHATE", "TriMAP", "PaCMAP"))
-
-## To draw with AIC
-aic_plot_zei <- ggplot(MSE_df_zei |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
-                                                                                 y = total_error,
-                                                                                 color = method
-)) +
-  geom_point() +
-  geom_line() +
-  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  theme_light() +
-  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  scale_colour_manual(values = c("#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("AIC") +
-  xlab("Total number of bins")
-## Effective number of bins along x-axis
-
-mse_plot_zei <- ggplot(MSE_df_zei, aes(x = number_of_bins,
-                                       y = total_mse,
-                                       color = method
-)) +
-  geom_point() +
-  geom_line() +
-  theme_light() +
-  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  scale_colour_manual(values = c("#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("MSE") +
-  xlab("Total number of bins")
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: Goodness of fit statistics from different NLDR techniques applied to training Zeisel mouse brain dataset. What is the best NLDR technique to represent the original data in 2D?
-#| label: fig-diagnosticpltZEI
-#| out-width: 100%
-
-aic_plot_zei + mse_plot_zei +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', ncol = 2) &
-  theme(legend.position='bottom', plot.tag = element_text(size = 8))
-```
-
-### Retina dataset
-
-Human retina: https://www.ebi.ac.uk/biostudies/arrayexpress/studies/E-MTAB-7316
-
-https://www.cell.com/cms/10.1016/j.cell.2015.05.002/attachment/fd52950c-fd35-401a-9e60-ca7588ef9001/mmc1.pdf
-
-https://www.cell.com/fulltext/S0092-8674(15)00549-8#app3
-
-32 PCs
-tSNE (perplexity 30)
-
-### Kang dataset
-
-For the Kang lupus dataset (29), we obtained 16 cell types based on clustering of the Festem-
-selected genes (Fig. 4E and Fig. S10, S11), including 2 fine cell types that were often missed by
-other methods. The 2 cell types were IFNhi CD14+ monocytes with high expression of type I
-interferon-related (IFN) genes, and HSP+ CD4 T cells with up-regulated heat-
-shock-protein-related (HSP) genes. The HSP+ CD4 T cells might result from
-cryopreservation and thawing of T cells (37). Other than devianceFS, other available methods
-could not identify these HSP+ CD4 T cells (Fig. S11).
-
-```{r}
-#| warning: false
-#| echo: false
-
-## Import data
-#df_2 <- read_rds("data/pbmc/pbmc.rds")
-# training_data_1 <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_pca_50.rds")
-# training_data_1 <- training_data_1[, 1:25]
-#test_1 <- read_rds("data/pbmc/pbmc_test.rds")
-
-#tSNE_pbmc <- read_rds("data/pbmc/pbmc_tsne_30.rds")
-tSNE_kang <- read_rds("data/kang/kang_tsne.rds") |>
-  mutate(cell_label = if_else(cell_label %in% c("0", "3", "5"), "0", cell_label))
-# PHATE_pbmc <- read_rds("data/pbmc/pbmc_phate.rds")
-# TriMAP_pbmc <- read_rds("data/pbmc/pbmc_trimap.rds")
-# PaCMAP_pbmc <- read_rds("data/pbmc/pbmc_pacmap.rds")
-
-#load("data/pbmc/pbmc_3k_festem/pbmc3k_label.RData")
-
-## Visualise embeddings
-
-# plot_list1_pbmc <- plot_tSNE_2D(tSNE_pbmc) + #ggtitle("(a)") + 
-#   theme_linedraw() +
-#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-#         axis.title.x = element_blank(), axis.title.y = element_blank(),
-#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-#   annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-class_avg <- tSNE_kang |>
-  group_by(cell_label) |>
-  summarise(
-    tSNE1 = median(tSNE1),
-    tSNE2 = median(tSNE2)
-  ) |>
-  mutate(cell_label = as.numeric(cell_label)) |>
-  arrange(cell_label)
-
-class_avg$cell_label <- as.factor(class_avg$cell_label)
-
-levels(class_avg$cell_label) <- c("CD14 Monocyte", "CD4 \n Memory T", "CD4 Naive T", "CD16 Monocyte", "B", "T Activated", "CD8 T", "NK", "DC", "B \nActivated", "T cell:\nmonocyte \nComplex", "HSP+ CD4 T", "IFN^hi CD14 \n Monocyte", "CD34 \n Progenitors", "pDC")
-
-plot_list2_kang <- tSNE_kang |>
-    ggplot(aes(x = tSNE1,
-               y = tSNE2, color = cell_label))+
-    geom_point(alpha=0.5) +
-    coord_equal() +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#b15928", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#cab2d6", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#bc80bd", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
-  geom_text(aes(x=tSNE1,y = tSNE2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) #+
-  # guides(color = guide_legend(title = "Cell type"),
-  #        color = guide_legend(nrow = 3)) #+
-  # annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-# plot_list3_pbmc <- plot_PHATE_2D(PHATE_pbmc) + #ggtitle("(c)") + 
-#   theme_linedraw() +
-#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-#         axis.title.x = element_blank(), axis.title.y = element_blank(),
-#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-#   annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-# plot_list4_pbmc <- plot_TriMAP_2D(TriMAP_pbmc) + #ggtitle("(d)") + 
-#   theme_linedraw() +
-#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-#         axis.title.x = element_blank(), axis.title.y = element_blank(),
-#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-#   annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-# 
-# plot_list5_pbmc <- plot_PaCMAP_2D(PaCMAP_pbmc) + #ggtitle("(e)") + 
-#   theme_linedraw() +
-#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-#         axis.title.x = element_blank(), axis.title.y = element_blank(),
-#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-#   annotate(geom = 'text', label = 'e', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "2D layout from tSNE (perplexity = 30) applied for the Kang dataset. Is this a best representation of the original data?"
-#| label: fig-nlderviskangtSNE
-
-plot_list2_kang
-
-#  plot_list1_pbmc + plot_list3_pbmc + plot_list4_pbmc + plot_list5_pbmc +
-#   plot_layout(ncol=5) #| out-width: 100%
-
-# 2D layouts from different NLDR techniques applied for the training PBMC dataset: (a) tSNE (perplexity = 30), (b) UMAP (n_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2).
-Is there a best representation of the original data or are they all providing  equivalent information?
-```
--->
-
 ## Conclusion {#sec-conclusion}

### 2024-02-07 paper.tex
Change names for columns and CHECK DONE

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, sharp corners, breakable, interior hidden, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, enhanced, breakable, frame hidden]}{\end{tcolorbox}}\fi
 

### 2024-02-10 paper.qmd
Generate the true model for s-curve

@@
 
+```{r}
+#| warning: false
+#| echo: false
+
+## To plot the distribution of distance
+plot_dist <- function(distance_df){
+  distance_df$group <- "1"
+  dist_plot <- ggplot(distance_df, aes(x = group, y = distance)) +
+    geom_quasirandom()+
+    ylim(0, max(unlist(distance_df$distance))+ 0.5) + coord_flip()
+  return(dist_plot)
+}
+```
+
 ## Introduction {#sec-intro}
@@
 
-data <- read_rds("data/s_curve/s_curve.rds")
-
-## tSNE
-
-shape_value_curve <- calculate_effective_shape_value(.data = tSNE_data,
-                                                   x = tSNE1, y = tSNE2)
-
-num_bins_vec <- 1:18 ## Number of bins along the x-axis
-
-vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names
-
-eval_data_training <- dplyr::bind_rows(vec)[0, ]
-eval_data_training <- eval_data_training |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(num_bins_vec)) {
-
-  model_object <- fit_high_d_model(training_data = training_data, nldr_df_with_id = tSNE_data, x = "tSNE1", y = "tSNE2", num_bins_x = num_bins_vec[i], shape_val = shape_value_curve,
-                             is_bin_centroid = TRUE,
-                             is_rm_lwd_hex = FALSE,
-                             benchmark_to_rm_lwd_hex = NA,
-                             is_avg_high_d = TRUE, column_start_text = "x")
-
-  centroid_df_training <- model_object$df_bin_centroids
-  avg_df_training <- model_object$df_bin
-
-  # pred_df_training <- predict_2d_embeddings(test_data = training_data, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "tSNE")
-
-  pred_df_training <- predict_hex_id(df_bin_centroids = centroid_df_training, nldr_df_test = tSNE_data, x = "tSNE1", y = "tSNE2")
-
-  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
-
-  eval_df_training <- eval_df_training |>
-    mutate(num_bins_x = num_bins_vec[i])
-
-  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
-
-}
-
-## Add new column with data types
-
-eval_data_training <- eval_data_training |>
-  mutate(data_type = "training")
-
-MSE_df_1 <- eval_data_training |>
-  dplyr::mutate(method = "tSNE")
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
 ## UMAP
 ## Prediction
+
+data <- read_rds("data/s_curve/s_curve.rds")
 
@@
 
-  # pred_df_training <- predict_2d_embeddings(test_data = training_data, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "tSNE")
-
-  pred_df_training <- predict_hex_id(df_bin_centroids = centroid_df_training, nldr_df_test = UMAP_data, x = "UMAP1", y = "UMAP2")
-
-  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+  pred_df_training <- predict_2d_embeddings(test_data = training_data, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "UMAP")
+
+  # pred_df_training <- predict_hex_id(df_bin_centroids = centroid_df_training, nldr_df_test = UMAP_data, x = "UMAP1", y = "UMAP2")
+
+  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, col_start = "x")
 
@@
 
-eval_data_training <- eval_data_training |>
-  mutate(data_type = "training")
-
-MSE_df_2 <- eval_data_training |>
-  dplyr::mutate(method = "UMAP")
-```
-
-```{r}
-#| warning: false
-#| echo: false
-## PAHTE
-## Prediction
-
-shape_value_curve <- calculate_effective_shape_value(.data = PHATE_data,
-                                                   x = PHATE1, y = PHATE2)
-
-num_bins_vec <- 1:18 ## Number of bins along the x-axis
-
-vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names
-
-eval_data_training <- dplyr::bind_rows(vec)[0, ]
-eval_data_training <- eval_data_training |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(num_bins_vec)) {
-
-  model_object <- fit_high_d_model(training_data = training_data, nldr_df_with_id = PHATE_data, x = "PHATE1", y = "PHATE2", num_bins_x = num_bins_vec[i], shape_val = shape_value_curve,
-                             is_bin_centroid = TRUE,
-                             is_rm_lwd_hex = FALSE,
-                             benchmark_to_rm_lwd_hex = NA,
-                             is_avg_high_d = TRUE, column_start_text = "x")
-
-  centroid_df_training <- model_object$df_bin_centroids
-  avg_df_training <- model_object$df_bin
-
-  # pred_df_training <- predict_2d_embeddings(test_data = training_data, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "tSNE")
-
-  pred_df_training <- predict_hex_id(df_bin_centroids = centroid_df_training, nldr_df_test = PHATE_data, x = "PHATE1", y = "PHATE2")
-
-  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
-
-  eval_df_training <- eval_df_training |>
-    mutate(num_bins_x = num_bins_vec[i])
-
-  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
-
-}
-
-## Add new column with data types
-
-eval_data_training <- eval_data_training |>
-  mutate(data_type = "training")
-
-MSE_df_3 <- eval_data_training |>
-  dplyr::mutate(method = "PHATE")
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-## TriMAP
-
-## Prediction
-
-shape_value_curve <- calculate_effective_shape_value(.data = TriMAP_data,
-                                                   x = TriMAP1, y = TriMAP2)
-
-num_bins_vec <- 1:10 ## Number of bins along the x-axis
-
-vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names
-
-eval_data_training <- dplyr::bind_rows(vec)[0, ]
-eval_data_training <- eval_data_training |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(num_bins_vec)) {
-
-  model_object <- fit_high_d_model(training_data = training_data, nldr_df_with_id = TriMAP_data, x = "TriMAP1", y = "TriMAP2", num_bins_x = num_bins_vec[i], shape_val = shape_value_curve,
-                             is_bin_centroid = TRUE,
-                             is_rm_lwd_hex = FALSE,
-                             benchmark_to_rm_lwd_hex = NA,
-                             is_avg_high_d = TRUE, column_start_text = "x")
-
-  centroid_df_training <- model_object$df_bin_centroids
-  avg_df_training <- model_object$df_bin
-
-  # pred_df_training <- predict_2d_embeddings(test_data = training_data, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "tSNE")
-
-  pred_df_training <- predict_hex_id(df_bin_centroids = centroid_df_training, nldr_df_test = TriMAP_data, x = "TriMAP1", y = "TriMAP2")
-
-  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
-
-  eval_df_training <- eval_df_training |>
-    mutate(num_bins_x = num_bins_vec[i])
-
-  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
-
-}
-
-## Add new column with data types
-
-eval_data_training <- eval_data_training |>
-  mutate(data_type = "training")
-
-MSE_df_4 <- eval_data_training |>
-  dplyr::mutate(method = "TriMAP")
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-## PaCMAP
-
-## Prediction
-
-shape_value_curve <- calculate_effective_shape_value(.data = PaCMAP_data,
-                                                   x = PaCMAP1, y = PaCMAP2)
-
-num_bins_vec <- 1:18 ## Number of bins along the x-axis
-
-vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names
-
-eval_data_training <- dplyr::bind_rows(vec)[0, ]
-eval_data_training <- eval_data_training |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(num_bins_vec)) {
-
-  model_object <- fit_high_d_model(training_data = training_data, nldr_df_with_id = PaCMAP_data, x = "PaCMAP1", y = "PaCMAP2", num_bins_x = num_bins_vec[i], shape_val = shape_value_curve,
-                             is_bin_centroid = TRUE,
-                             is_rm_lwd_hex = FALSE,
-                             benchmark_to_rm_lwd_hex = NA,
-                             is_avg_high_d = TRUE, column_start_text = "x")
-
-  centroid_df_training <- model_object$df_bin_centroids
-  avg_df_training <- model_object$df_bin
-
-  # pred_df_training <- predict_2d_embeddings(test_data = training_data, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "tSNE")
-
-  pred_df_training <- predict_hex_id(df_bin_centroids = centroid_df_training, nldr_df_test = PaCMAP_data, x = "PaCMAP1", y = "PaCMAP2")
-
-  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
-
-  eval_df_training <- eval_df_training |>
-    mutate(num_bins_x = num_bins_vec[i])
-
-  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
-
-}
-
-## Add new column with data types
-
-eval_data_training <- eval_data_training |>
-  mutate(data_type = "training")
-
-MSE_df_5 <- eval_data_training |>
-  dplyr::mutate(method = "PaCMAP")
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-MSE_df <- dplyr::bind_rows(MSE_df_1, MSE_df_2, MSE_df_3, MSE_df_4, MSE_df_5)
-
-MSE_df$method <- factor(MSE_df$method, levels = c("tSNE", "UMAP", "PHATE", "TriMAP", "PaCMAP"))
+MSE_df_umap_s_curve <- eval_data_training |>
+  mutate(data_type = "training",
+         method = "UMAP")
+```
+
+```{r}
+#| warning: false
+#| echo: false
 
 ## To draw with AIC
-aic_plot <- ggplot(MSE_df |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
+aic_plot <- ggplot(MSE_df_umap_s_curve |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
                                                                                  y = total_error,
@@
         axis.text = element_text(size = 7)) +
-  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
+  scale_colour_manual(values = c("#377eb8")) +
   ylab("AIC") +
@@
 
-mse_plot <- ggplot(MSE_df |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
+mse_plot <- ggplot(MSE_df_umap_s_curve |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
                                        y = total_mse,
@@
   # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
+  scale_colour_manual(values = c("#377eb8")) +
   ylab("MSE") +
@@
 
-```{r}
-#| echo: false
-#| warning: false
-
-error_df_lg_1 <- read_rds("data/s_curve/s_curve_summary_lg_tsne.rds")
-error_df_lg_2 <- read_rds("data/s_curve/s_curve_summary_lg_umap.rds")
-error_df_lg_3 <- read_rds("data/s_curve/s_curve_summary_lg_phate.rds")
-error_df_lg_4 <- read_rds("data/s_curve/s_curve_summary_lg_trimap.rds")
-error_df_lg_5 <- read_rds("data/s_curve/s_curve_summary_lg_pacmap.rds")
-
-error_df_lg <- dplyr::bind_rows(error_df_lg_1, error_df_lg_2, error_df_lg_3, error_df_lg_4, error_df_lg_5)
-
-error_df_lg$method <- factor(error_df_lg$method, levels = c("tSNE", "UMAP", "PHATE", "TriMAP", "PaCMAP"))
-
-## To draw with AIC
-aic_plot_lg <- ggplot(error_df_lg |> dplyr::filter(method == "UMAP"), aes(x = benchmark_rm_lg,
-                                                                                 y = total_error,
-                                                                                 color = method
-)) +
-  geom_point() +
-  geom_vline(xintercept = 1.347, linetype="dashed", 
-                color = "red", size=0.5) +
-  geom_line() +
-  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("AIC") +
-  xlab("benchmark value")
-## Effective number of bins along x-axis
-
-mse_plot_lg <- ggplot(error_df_lg |> dplyr::filter(method == "UMAP"), aes(x = benchmark_rm_lg,
-                                       y = total_mse,
-                                       color = method
-)) +
-  geom_point() +
-  geom_vline(xintercept = 1.347, linetype="dashed", 
-                color = "red", size=0.5) +
-  geom_line() +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("MSE") +
-  xlab("benchmark value")
-
-```
-
-```{r}
-#| echo: false
-#| warning: false
-#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset with different benchmark values to remove the long edges.
-What is the effective benchmark value to remove the long edges?
-$1.347$ is the chosen benchmark value to remove long edges.
-The value is selected through a visual inspection of the triangular mesh is conducted to verify whether it successfully eliminates long edges."
-#| label: fig-diagnosticpltScurvelgrm
-##| out-width: 100%
-#| fig-pos: H
-
-aic_plot_lg + mse_plot_lg +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', ncol = 2) &
-  theme(legend.position='none', plot.tag = element_text(size = 8)) 
-```
-
 #### Hexagonal grid origin
@@
   geom_line() +
-  geom_vline(xintercept = 0, linetype="dashed",
+  geom_vline(xintercept = MSE_df |> dplyr::filter(total_mse == min(total_mse)) |> dplyr::pull(shift), linetype="dashed",
                 color = "red", size=0.5) +
@@
   geom_line() +
-   geom_vline(xintercept = 0, linetype="dashed",
+   geom_vline(xintercept = MSE_df |> dplyr::filter(total_mse == min(total_mse)) |> dplyr::pull(shift), linetype="dashed",
                 color = "red", size=0.5) +
@@
 
-The approach involves employing the K-nearest neighbors (KNN) algorithm to identify the nearest hexagonal bin centroid in the 2D space.
+<!--In this context, "prediction" denotes the 2D embedding generated for the NLDR technique.
+The methodology encompasses identifying the nearest averaged high-D points in the high-D space for the test data, by computing high-D Euclidean distances.
+As the averaged high-D point corresponds to the lifting of the 2D model, determining its nearest counterpart allows us to map its hexagonal bin centroid coordinates.
+Consequently, these centroid coordinates serve as the assigned low-dimensional predicted values for the test data in the 2D space.
+
+Some NLDR techniques, such as tSNE, often lack a direct prediction, making our approach valuable for generating predicted values in the absence of such functionalities.
+
+The prediction approach involves employing the K-nearest neighbors (KNN) algorithm to identify the nearest hexagonal bin centroid in the 2D space.
 Subsequently, the coordinates of this centroid are assigned as the low-dimensional predicted values for the test data in 2D space.
+
+The prediction approach involves employing the K-nearest neighbors (KNN) algorithm to identify the nearest high-D model point (averaged high-D point) for the new high-D data point.
+Also, the coordinates of the high-D model point serve as the predicted 2D embedings for the new high-D data point.
 It is noteworthy that traditional NLDR methods, such as tSNE, often lack a direct predict function, making our approach valuable for generating predicted values in the absence of such functionalities.
+-->
+
+The prediction approach involves employing the K-nearest neighbors (KNN) algorithm.
+In this method, which operates as an unsupervised classification problem, the nearest high-D model point (averaged high-D point) is identified for the new high-D data point.
+Once the nearest high-D model point is identified, its coordinates are used as the predicted 2D embeddings for the new high-D data point.
+This step is particularly valuable due to the limitations of certain NLDR techniques, like tSNE, which don't provide a straightforward method for predicting 2D embeddings.
+As a result, our approach presents a solution that can generate predicted 2D embeddings, irrespective of the NLDR technique employed, effectively addressing this gap in functionality.
 
@@
 
-Here, $n$ represents the number of bins, $p$ represents the dimensions, $x_{ij}$ is the actual high-dimensional data, and $C_{x_ij}$ is the predicted bin centroid data in high dimensions.
+Here, $n$ represents the number of observations, $p$ represents the dimensions, $x_{ij}$ is the actual high-dimensional data, and $C_{x_ij}$ is the predicted bin centroid data in high dimensions.
 
@@
 
+<!--
 ### Model interpretation
@@
 In the 2D layouts of tSNE and PaCMAP (see @fig-nldervis (a), (e)), there are some sparse areas that are captured by the models effectively (see see videos of the model with tSNE and PaCMAP linked in @fig-modelScurve).
+-->
 
@@
 -->
-
-### Prediction {#sec-prediction}
-
-In this context, "prediction" denotes the 2D embedding generated for the NLDR technique.
-The methodology encompasses identifying the nearest averaged high-D points in the high-D space for the test data, by computing high-D Euclidean distances.
-As the averaged high-D point corresponds to the lifting of the 2D model, determining its nearest counterpart allows us to map its hexagonal bin centroid coordinates.
-Consequently, these centroid coordinates serve as the assigned low-dimensional predicted values for the test data in the 2D space.
-
-Some NLDR techniques, such as tSNE, often lack a direct prediction, making our approach valuable for generating predicted values in the absence of such functionalities.
 
@@
 
+num_non_empty_bins_gau <- df_bin_centroids_tsne_gau$hexID |> length()
+
 tSNE_data_with_hb_id_gau <- tSNE_data_gau |>
@@
 #                                                 cell_area = 1) #41
-num_bins_umap_gau <- 26
 
@@
                                                       x = UMAP1, y = UMAP2)
+num_bins_umap_gau <- find_non_empty_bins(nldr_df = UMAP_data_gau, x = "UMAP1", y = "UMAP2", shape_val = shape_val_umap_gau, non_empty_bins = num_non_empty_bins_gau)
+
 ## To extract bin centroids
@@
 #                                                  cell_area = 1)
-num_bins_phate_gau <- 70
 
@@
                                                        x = PHATE1, y = PHATE2)
+num_bins_phate_gau <- find_non_empty_bins(nldr_df = PHATE_data_gau, x = "PHATE1", y = "PHATE2", shape_val = shape_val_phate_gau, non_empty_bins = num_non_empty_bins_gau)
+
 ## To extract bin centroids
@@
 
-num_bins_trimap_gau <- 21
 shape_val_trimap_gau <- calculate_effective_shape_value(.data = TriMAP_data_gau,
                                                         x = TriMAP1, y = TriMAP2)
+
+num_bins_trimap_gau <- find_non_empty_bins(nldr_df = TriMAP_data_gau, x = "TriMAP1", y = "TriMAP2", shape_val = shape_val_trimap_gau, non_empty_bins = num_non_empty_bins_gau)
+
 ## To extract bin centroids
@@
 #                                                   cell_area = 1) #30
-num_bins_pacmap_gau <- 18
 
@@
                                                         x = PaCMAP1, y = PaCMAP2)
+
+num_bins_pacmap_gau <- find_non_empty_bins(nldr_df = PaCMAP_data_gau, x = "PaCMAP1", y = "PaCMAP2", shape_val = shape_val_pacmap_gau, non_empty_bins = num_non_empty_bins_gau)
+
 ## To extract bin centroids
@@
 
-  # pred_df_training <- predict_2d_embeddings(test_data = training_data_1, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "tSNE")
-
-  pred_df_training <- predict_hex_id(df_bin_centroids = centroid_df_training, nldr_df_test = tSNE_data_gau, x = "tSNE1", y = "tSNE2")
-
-  eval_df_training <- generate_eval_df(data = df_2, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+  pred_df_training <- predict_2d_embeddings(test_data = training_data_1, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "tSNE")
+
+  eval_df_training <- generate_eval_df(data = df_2, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, col_start = "x")
 
@@
 
-  # pred_df_training <- predict_2d_embeddings(test_data = training_data_1, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "UMAP")
-
-  pred_df_training <- predict_hex_id(df_bin_centroids = centroid_df_training, nldr_df_test = UMAP_data_gau, x = "UMAP1", y = "UMAP2")
-
-  eval_df_training <- generate_eval_df(data = df_2, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+  pred_df_training <- predict_2d_embeddings(test_data = training_data_1, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "UMAP")
+
+  eval_df_training <- generate_eval_df(data = df_2, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, col_start = "x")
 
@@
 
-  # pred_df_training <- predict_2d_embeddings(test_data = training_data_1, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "PHATE")
-
-  pred_df_training <- predict_hex_id(df_bin_centroids = centroid_df_training, nldr_df_test = PHATE_data_gau, x = "PHATE1", y = "PHATE2")
-
-  eval_df_training <- generate_eval_df(data = df_2, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+  pred_df_training <- predict_2d_embeddings(test_data = training_data_1, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "PHATE")
+
+  eval_df_training <- generate_eval_df(data = df_2, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, col_start = "x")
 
@@
 
-  # pred_df_training <- predict_2d_embeddings(test_data = training_data_1, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "TriMAP")
-
-  pred_df_training <- predict_hex_id(df_bin_centroids = centroid_df_training, nldr_df_test = TriMAP_data_gau, x = "TriMAP1", y = "TriMAP2")
-
-  eval_df_training <- generate_eval_df(data = df_2, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+  pred_df_training <- predict_2d_embeddings(test_data = training_data_1, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "TriMAP")
+
+  eval_df_training <- generate_eval_df(data = df_2, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, col_start = "x")
 
@@
 
-  # pred_df_training <- predict_2d_embeddings(test_data = training_data_1, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "PaCMAP")
-
-  pred_df_training <- predict_hex_id(df_bin_centroids = centroid_df_training, nldr_df_test = PaCMAP_data_gau, x = "PaCMAP1", y = "PaCMAP2")
-
-  eval_df_training <- generate_eval_df(data = df_2, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
+  pred_df_training <- predict_2d_embeddings(test_data = training_data_1, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "PaCMAP")
+
+  eval_df_training <- generate_eval_df(data = df_2, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, col_start = "x")
 
@@
 
--
-
 - How did they "identified 10 clusters"? What were the variables used, PCs or full data or 2D representation, spell this out.
@@
 #| echo: false
-num_bins_pbmc <- calculate_effective_x_bins(.data = UMAP_pbmc, x = UMAP1,
-                                              cell_area = 1) ##23
-
-num_bins_pbmc <- 23
+
+## UMAP
+## Prediction
+
+shape_value_pbmc <- calculate_effective_shape_value(.data = UMAP_pbmc,
+                                                      x = UMAP1, y = UMAP2)
+
+num_bins_vec <- 1:35 ## Number of bins along the x-axis
+
+vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(num_bins_vec)) {
+
+  model_object <- fit_high_d_model(training_data = training_data_pbmc, nldr_df_with_id = UMAP_pbmc, x = "UMAP1", y = "UMAP2", num_bins_x = num_bins_vec[i], shape_val = shape_value_pbmc,
+                             is_bin_centroid = TRUE,
+                             is_rm_lwd_hex = FALSE,
+                             benchmark_to_rm_lwd_hex = NA,
+                             is_avg_high_d = TRUE, column_start_text = "PC_")
+
+  centroid_df_training <- model_object$df_bin_centroids
+  avg_df_training <- model_object$df_bin
+
+  pred_df_training <- predict_2d_embeddings(test_data = training_data_pbmc, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "UMAP")
+
+  eval_df_training <- generate_eval_df(data = training_data_pbmc, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, col_start = "PC_")
+
+  eval_df_training <- eval_df_training |>
+    mutate(num_bins_x = num_bins_vec[i])
+
+  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+
+}
+
+## Add new column with data types
+
+MSE_df_pbmc <- eval_data_training |>
+  dplyr::mutate(data_type = "training",
+                method = "UMAP")
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## To draw with AIC
+aic_plot_pbmc <- ggplot(MSE_df_pbmc |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
+                                                                                 y = total_error,
+                                                                                 color = method
+)) +
+  geom_point() +
+  geom_line() +
+  geom_vline(xintercept = 256, linetype="dashed", 
+                color = "red", size=0.5) +
+  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("AIC") +
+  xlab("total number of bins")
+## Effective number of bins along x-axis
+
+mse_plot_pbmc <- ggplot(MSE_df_pbmc |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
+                                       y = total_mse,
+                                       color = method
+)) +
+  geom_point() +
+  geom_line() +
+   geom_vline(xintercept = 256, linetype="dashed", 
+                color = "red", size=0.5) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("MSE") +
+  xlab("total number of bins")
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset.
+What is the effective number of bins in each NLDR technique to create a 2D model?
+The MSE plot have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
+Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error.
+Using the elbow method, when the total number of bins is set to 144, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
+This point indicates that adding less bins does not enough to capture the data structure."
+#| label: fig-diagnosticpltPBMC
+##| out-width: 100%
+#| fig-pos: H
+
+aic_plot_pbmc + mse_plot_pbmc +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='none', plot.tag = element_text(size = 8))
+```
+
+```{r}
+#| warning: false
+#| echo: false
+# num_bins_pbmc <- calculate_effective_x_bins(.data = UMAP_pbmc, x = UMAP1,
+#                                               cell_area = 1) ##23
+
+num_bins_pbmc <- 15
 
@@
                                                       x = UMAP1, y = UMAP2) ## 0.8772751
-shape_value_pbmc <- 0.8772751
 
@@
 
-num_non_empty_bins_pbmc <- df_bin_centroids_pbmc$hexID |> length() #83
+#num_non_empty_bins_pbmc <- df_bin_centroids_pbmc$hexID |> length() #83
 
@@
 
+distance_plot <- plot_dist(distance_pbmc) +
+  #ggtitle("(b)" ) +
+  ylab(expression(d^{(2)})) +
+  theme(axis.text = element_text(size = 5),
+        axis.title = element_text(size = 12))
+
+distance_plot
+
 ## To find the benchmark value
 benchmark_pbmc <- find_benchmark_value(.data = distance_pbmc, distance_col = distance)
-benchmark_pbmc <- 1.072513
+benchmark_pbmc <- 2.5
 
@@
 
-# plot_list2_pbmc + 
-trimesh_removed_pbmc_umap + plot_list2_pbmc_error +
+plot_list2_pbmc + trimesh_removed_pbmc_umap + plot_list2_pbmc_error +
 plot_layout(guides='collect', ncol=3) &
@@
 #| echo: false
+
+## UMAP
+## Prediction
+
+shape_value_pbmc <- calculate_effective_shape_value(.data = UMAP_pbmc,
+                                                      x = UMAP1, y = UMAP2)
+
+num_bins_vec <- 1:35 ## Number of bins along the x-axis
+
+vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(num_bins_vec)) {
+
+  model_object <- fit_high_d_model(training_data = training_data_pbmc, nldr_df_with_id = UMAP_pbmc, x = "UMAP1", y = "UMAP2", num_bins_x = num_bins_vec[i], shape_val = shape_value_pbmc,
+                             is_bin_centroid = TRUE,
+                             is_rm_lwd_hex = FALSE,
+                             benchmark_to_rm_lwd_hex = NA,
+                             is_avg_high_d = TRUE, column_start_text = "PC_")
+
+  centroid_df_training <- model_object$df_bin_centroids
+  avg_df_training <- model_object$df_bin
+
+  pred_df_training <- predict_2d_embeddings(test_data = training_data_pbmc, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "UMAP")
+
+  eval_df_training <- generate_eval_df(data = training_data_pbmc, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, col_start = "PC_")
+
+  eval_df_training <- eval_df_training |>
+    mutate(num_bins_x = num_bins_vec[i])
+
+  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+
+}
+
+## Add new column with data types
+
+MSE_df_pbmc <- eval_data_training |>
+  dplyr::mutate(data_type = "training",
+                method = "UMAP")
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## To draw with AIC
+aic_plot_pbmc <- ggplot(MSE_df_pbmc |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
+                                                                                 y = total_error,
+                                                                                 color = method
+)) +
+  geom_point() +
+  geom_line() +
+  geom_vline(xintercept = 192, linetype="dashed", 
+                color = "red", size=0.5) +
+  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("AIC") +
+  xlab("total number of bins")
+## Effective number of bins along x-axis
+
+mse_plot_pbmc <- ggplot(MSE_df_pbmc |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
+                                       y = total_mse,
+                                       color = method
+)) +
+  geom_point() +
+  geom_line() +
+   geom_vline(xintercept = 192, linetype="dashed", 
+                color = "red", size=0.5) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("MSE") +
+  xlab("total number of bins")
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset.
+What is the effective number of bins in each NLDR technique to create a 2D model?
+The MSE plot have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
+Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error.
+Using the elbow method, when the total number of bins is set to 144, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
+This point indicates that adding less bins does not enough to capture the data structure."
+#| label: fig-diagnosticpltPBMC2
+##| out-width: 100%
+#| fig-pos: H
+
+aic_plot_pbmc + mse_plot_pbmc +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='none', plot.tag = element_text(size = 8))
+```
+
+```{r}
+#| warning: false
+#| echo: false
 shape_value_pbmc <- calculate_effective_shape_value(.data = UMAP_pbmc,
@@
 
-num_bins_pbmc <- find_non_empty_bins(nldr_df = UMAP_pbmc, x = "UMAP1", y = "UMAP2", shape_val = shape_value_pbmc, non_empty_bins = num_non_empty_bins_pbmc)
+# num_bins_pbmc <- find_non_empty_bins(nldr_df = UMAP_pbmc, x = "UMAP1", y = "UMAP2", shape_val = shape_value_pbmc, non_empty_bins = num_non_empty_bins_pbmc)
+num_bins_pbmc <-  11
 
@@
 
+# distance_plot <- plot_dist(distance_pbmc) +
+#   #ggtitle("(b)" ) +
+#   ylab(expression(d^{(2)})) +
+#   theme(axis.text = element_text(size = 5),
+#         axis.title = element_text(size = 12))
+# 
+# distance_plot
+
 ## To find the benchmark value
@@
 #benchmark_pbmc <- 1.049766 
+benchmark_pbmc <- 2.5
 
@@
 #| echo: false
+
+## UMAP
+## Prediction
+
+shape_value_pbmc <- calculate_effective_shape_value(.data = UMAP_pbmc,
+                                                      x = UMAP1, y = UMAP2)
+
+num_bins_vec <- 1:35 ## Number of bins along the x-axis
+
+vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(num_bins_vec)) {
+
+  model_object <- fit_high_d_model(training_data = training_data_pbmc, nldr_df_with_id = UMAP_pbmc, x = "UMAP1", y = "UMAP2", num_bins_x = num_bins_vec[i], shape_val = shape_value_pbmc,
+                             is_bin_centroid = TRUE,
+                             is_rm_lwd_hex = FALSE,
+                             benchmark_to_rm_lwd_hex = NA,
+                             is_avg_high_d = TRUE, column_start_text = "PC_")
+
+  centroid_df_training <- model_object$df_bin_centroids
+  avg_df_training <- model_object$df_bin
+
+  pred_df_training <- predict_2d_embeddings(test_data = training_data_pbmc, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "UMAP")
+
+  eval_df_training <- generate_eval_df(data = training_data_pbmc, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, col_start = "PC_")
+
+  eval_df_training <- eval_df_training |>
+    mutate(num_bins_x = num_bins_vec[i])
+
+  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+
+}
+
+## Add new column with data types
+
+MSE_df_pbmc <- eval_data_training |>
+  dplyr::mutate(data_type = "training",
+                method = "UMAP")
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## To draw with AIC
+aic_plot_pbmc <- ggplot(MSE_df_pbmc |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
+                                                                                 y = total_error,
+                                                                                 color = method
+)) +
+  geom_point() +
+  geom_line() +
+  geom_vline(xintercept = 182, linetype="dashed", 
+                color = "red", size=0.5) +
+  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("AIC") +
+  xlab("total number of bins")
+## Effective number of bins along x-axis
+
+mse_plot_pbmc <- ggplot(MSE_df_pbmc |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
+                                       y = total_mse,
+                                       color = method
+)) +
+  geom_point() +
+  geom_line() +
+   geom_vline(xintercept = 182, linetype="dashed", 
+                color = "red", size=0.5) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("MSE") +
+  xlab("total number of bins")
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset.
+What is the effective number of bins in each NLDR technique to create a 2D model?
+The MSE plot have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
+Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error.
+Using the elbow method, when the total number of bins is set to 144, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
+This point indicates that adding less bins does not enough to capture the data structure."
+#| label: fig-diagnosticpltPBMC3
+##| out-width: 100%
+#| fig-pos: H
+
+aic_plot_pbmc + mse_plot_pbmc +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='none', plot.tag = element_text(size = 8))
+```
+
+```{r}
+#| warning: false
+#| echo: false
 # num_bins_pbmc <- calculate_effective_x_bins(.data = UMAP_pbmc, x = UMAP1,
@@
 
-num_bins_pbmc <- find_non_empty_bins(nldr_df = UMAP_pbmc, x = "UMAP1", y = "UMAP2", shape_val = shape_value_pbmc, non_empty_bins = num_non_empty_bins_pbmc)
+# num_bins_pbmc <- find_non_empty_bins(nldr_df = UMAP_pbmc, x = "UMAP1", y = "UMAP2", shape_val = shape_value_pbmc, non_empty_bins = num_non_empty_bins_pbmc)
+num_bins_pbmc <- 12
 
@@
 
+# distance_plot <- plot_dist(distance_pbmc) +
+#   #ggtitle("(b)" ) +
+#   ylab(expression(d^{(2)})) +
+#   theme(axis.text = element_text(size = 5),
+#         axis.title = element_text(size = 12))
+# 
+# distance_plot
+
 ## To find the benchmark value
 benchmark_pbmc <- find_benchmark_value(.data = distance_pbmc, distance_col = distance)
-#benchmark_pbmc <- 0.5517801
+benchmark_pbmc <- 2.5
 
@@
 
-### tSNE
-
-num_bins_tsne_harps <- calculate_effective_x_bins(.data = tSNE_data_harps, x = tSNE1,
-                                                cell_area = 1) #43
+## tSNE
+## Prediction
 
@@
                                                       x = tSNE1, y = tSNE2)
+
+num_bins_vec <- 1:21 ## Number of bins along the x-axis
+
+vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(num_bins_vec)) {
+
+  model_object <- fit_high_d_model(training_data = training_data_1, nldr_df_with_id = tSNE_data_harps, x = "tSNE1", y = "tSNE2", num_bins_x = num_bins_vec[i], shape_val = shape_val_tsne_harps,
+                             is_bin_centroid = TRUE,
+                             is_rm_lwd_hex = FALSE,
+                             benchmark_to_rm_lwd_hex = NA,
+                             is_avg_high_d = TRUE, column_start_text = "x")
+
+  centroid_df_training <- model_object$df_bin_centroids
+  avg_df_training <- model_object$df_bin
+
+  pred_df_training <- predict_2d_embeddings(test_data = training_data_1, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "tSNE")
+
+  eval_df_training <- generate_eval_df(data = training_data_1, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, col_start = "x")
+
+  eval_df_training <- eval_df_training |>
+    mutate(num_bins_x = num_bins_vec[i])
+
+  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+
+}
+
+## Add new column with data types
+
+MSE_df_harps <- eval_data_training |>
+  dplyr::mutate(data_type = "training",
+                method = "tSNE")
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## To draw with AIC
+aic_plot_harps <- ggplot(MSE_df_harps |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "tSNE"), aes(x = number_of_bins,
+                                                                                 y = total_error,
+                                                                                 color = method
+)) +
+  geom_point() +
+  geom_line() +
+  geom_vline(xintercept = 120, linetype="dashed", 
+                color = "black", size=0.5) +
+  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  scale_colour_manual(values = c("#e41a1c","#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("AIC") +
+  xlab("total number of bins")
+## Effective number of bins along x-axis
+
+mse_plot_harps <- ggplot(MSE_df_harps |> dplyr::filter(method == "tSNE"), aes(x = number_of_bins,
+                                       y = total_mse,
+                                       color = method
+)) +
+  geom_point() +
+  geom_line() +
+   geom_vline(xintercept = 120, linetype="dashed", 
+                color = "black", size=0.5) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  scale_colour_manual(values = c("#e41a1c","#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("MSE") +
+  xlab("total number of bins")
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset.
+What is the effective number of bins in each NLDR technique to create a 2D model?
+The MSE plot have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
+Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error.
+Using the elbow method, when the total number of bins is set to 144, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
+This point indicates that adding less bins does not enough to capture the data structure."
+#| label: fig-diagnosticpltHARPS
+##| out-width: 100%
+#| fig-pos: H
+
+aic_plot_harps + mse_plot_harps +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='none', plot.tag = element_text(size = 8))
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+### tSNE
+
+num_bins_tsne_harps <- calculate_effective_x_bins(.data = tSNE_data_harps, x = "tSNE1",
+                                                cell_area = 1) #43
+
+num_bins_tsne_harps <- 14
+
+shape_val_tsne_harps <- calculate_effective_shape_value(.data = tSNE_data_harps,
+                                                      x = "tSNE1", y = "tSNE2")
 ## To extract bin centroids
-hexbin_data_object_tsne_harps <- extract_hexbin_centroids(nldr_df = tSNE_data_harps, num_bins = num_bins_tsne_harps, shape_val = shape_val_tsne_harps, x = tSNE1, y = tSNE2)
+hexbin_data_object_tsne_harps <- extract_hexbin_centroids(nldr_df = tSNE_data_harps, num_bins = num_bins_tsne_harps, shape_val = shape_val_tsne_harps, x = "tSNE1", y = "tSNE2")
 
@@
 
-num_non_empty_bins_harps <- df_bin_centroids_tsne_harps$hexID |> length()
+#num_non_empty_bins_harps <- df_bin_centroids_tsne_harps$hexID |> length()
 
@@
 ## To find the benchmark value
-benchmark_tsne_harps <- find_benchmark_value(.data = distance_tsne_harps, distance_col = distance)
+benchmark_tsne_harps <- find_benchmark_value(.data = distance_tsne_harps, distance_col = "distance")
+benchmark_tsne_harps <- 5
+
+# distance_plot <- plot_dist(distance_tsne_harps) +
+#   #ggtitle("(b)" ) +
+#   ylab(expression(d^{(2)})) +
+#   theme(axis.text = element_text(size = 5),
+#         axis.title = element_text(size = 12))
+# 
+# distance_plot
 
@@
 
+## tSNE
+## Prediction
+
+shape_val_tsne_harps <- calculate_effective_shape_value(.data = tSNE_data_harps,
+                                                      x = tSNE1, y = tSNE2)
+
+num_bins_vec <- 1:21 ## Number of bins along the x-axis
+
+vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(num_bins_vec)) {
+
+  model_object <- fit_high_d_model(training_data = training_data_1, nldr_df_with_id = tSNE_data_harps, x = "tSNE1", y = "tSNE2", num_bins_x = num_bins_vec[i], shape_val = shape_val_tsne_harps,
+                             is_bin_centroid = TRUE,
+                             is_rm_lwd_hex = FALSE,
+                             benchmark_to_rm_lwd_hex = NA,
+                             is_avg_high_d = TRUE, column_start_text = "x")
+
+  centroid_df_training <- model_object$df_bin_centroids
+  avg_df_training <- model_object$df_bin
+
+  pred_df_training <- predict_2d_embeddings(test_data = training_data_1, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "tSNE")
+
+  eval_df_training <- generate_eval_df(data = training_data_1, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, col_start = "x")
+
+  eval_df_training <- eval_df_training |>
+    mutate(num_bins_x = num_bins_vec[i])
+
+  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+
+}
+
+## Add new column with data types
+
+MSE_df_harps <- eval_data_training |>
+  dplyr::mutate(data_type = "training",
+                method = "tSNE")
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## To draw with AIC
+aic_plot_harps <- ggplot(MSE_df_harps |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "tSNE"), aes(x = number_of_bins,
+                                                                                 y = total_error,
+                                                                                 color = method
+)) +
+  geom_point() +
+  geom_line() +
+  geom_vline(xintercept = 160, linetype="dashed", 
+                color = "black", size=0.5) +
+  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  scale_colour_manual(values = c("#e41a1c","#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("AIC") +
+  xlab("total number of bins")
+## Effective number of bins along x-axis
+
+mse_plot_harps <- ggplot(MSE_df_harps |> dplyr::filter(method == "tSNE"), aes(x = number_of_bins,
+                                       y = total_mse,
+                                       color = method
+)) +
+  geom_point() +
+  geom_line() +
+   geom_vline(xintercept = 160, linetype="dashed", 
+                color = "black", size=0.5) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  scale_colour_manual(values = c("#e41a1c","#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("MSE") +
+  xlab("total number of bins")
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset.
+What is the effective number of bins in each NLDR technique to create a 2D model?
+The MSE plot have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
+Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error.
+Using the elbow method, when the total number of bins is set to 144, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
+This point indicates that adding less bins does not enough to capture the data structure."
+#| label: fig-diagnosticpltHARPS2
+##| out-width: 100%
+#| fig-pos: H
+
+aic_plot_harps + mse_plot_harps +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='none', plot.tag = element_text(size = 8))
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
 ## Import data
@@
 
-num_bins_tsne_harps <- find_non_empty_bins(nldr_df = tSNE_data_harps, x = "tSNE1", y = "tSNE2", shape_val = shape_val_tsne_harps, non_empty_bins = num_non_empty_bins_harps)
+# num_bins_tsne_harps <- find_non_empty_bins(nldr_df = tSNE_data_harps, x = "tSNE1", y = "tSNE2", shape_val = shape_val_tsne_harps, non_empty_bins = num_non_empty_bins_harps)
+
+num_bins_tsne_harps <- 9
 
@@
 
+# distance_plot <- plot_dist(distance_tsne_harps) +
+#   #ggtitle("(b)" ) +
+#   ylab(expression(d^{(2)})) +
+#   theme(axis.text = element_text(size = 5),
+#         axis.title = element_text(size = 12))
+# 
+# distance_plot
+
 ## To find the benchmark value
-benchmark_tsne_harps <- find_benchmark_value(.data = distance_tsne_harps, distance_col = distance)
-benchmark_tsne_harps <- 2.5
+benchmark_tsne_harps <- find_benchmark_value(.data = distance_tsne_harps, distance_col = "distance")
+benchmark_tsne_harps <- 6
 
@@
 
+## tSNE
+## Prediction
+
+shape_val_tsne_harps <- calculate_effective_shape_value(.data = tSNE_data_harps,
+                                                      x = tSNE1, y = tSNE2)
+
+num_bins_vec <- 1:21 ## Number of bins along the x-axis
+
+vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(num_bins_vec)) {
+
+  model_object <- fit_high_d_model(training_data = training_data_1, nldr_df_with_id = tSNE_data_harps, x = "tSNE1", y = "tSNE2", num_bins_x = num_bins_vec[i], shape_val = shape_val_tsne_harps,
+                             is_bin_centroid = TRUE,
+                             is_rm_lwd_hex = FALSE,
+                             benchmark_to_rm_lwd_hex = NA,
+                             is_avg_high_d = TRUE, column_start_text = "x")
+
+  centroid_df_training <- model_object$df_bin_centroids
+  avg_df_training <- model_object$df_bin
+
+  pred_df_training <- predict_2d_embeddings(test_data = training_data_1, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "tSNE")
+
+  eval_df_training <- generate_eval_df(data = training_data_1, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, col_start = "x")
+
+  eval_df_training <- eval_df_training |>
+    mutate(num_bins_x = num_bins_vec[i])
+
+  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+
+}
+
+## Add new column with data types
+
+MSE_df_harps <- eval_data_training |>
+  dplyr::mutate(data_type = "training",
+                method = "tSNE")
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## To draw with AIC
+aic_plot_harps <- ggplot(MSE_df_harps |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "tSNE"), aes(x = number_of_bins,
+                                                                                 y = total_error,
+                                                                                 color = method
+)) +
+  geom_point() +
+  geom_line() +
+  geom_vline(xintercept = 180, linetype="dashed", 
+                color = "black", size=0.5) +
+  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  scale_colour_manual(values = c("#e41a1c","#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("AIC") +
+  xlab("total number of bins")
+## Effective number of bins along x-axis
+
+mse_plot_harps <- ggplot(MSE_df_harps |> dplyr::filter(method == "tSNE"), aes(x = number_of_bins,
+                                       y = total_mse,
+                                       color = method
+)) +
+  geom_point() +
+  geom_line() +
+   geom_vline(xintercept = 180, linetype="dashed", 
+                color = "black", size=0.5) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  scale_colour_manual(values = c("#e41a1c","#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("MSE") +
+  xlab("total number of bins")
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset.
+What is the effective number of bins in each NLDR technique to create a 2D model?
+The MSE plot have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
+Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error.
+Using the elbow method, when the total number of bins is set to 144, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
+This point indicates that adding less bins does not enough to capture the data structure."
+#| label: fig-diagnosticpltHARPS3
+##| out-width: 100%
+#| fig-pos: H
+
+aic_plot_harps + mse_plot_harps +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='none', plot.tag = element_text(size = 8))
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
 ## Import data
@@
 
-num_bins_tsne_harps <- find_non_empty_bins(nldr_df = tSNE_data_harps, x = "tSNE1", y = "tSNE2", shape_val = shape_val_tsne_harps, non_empty_bins = num_non_empty_bins_harps)
+# num_bins_tsne_harps <- find_non_empty_bins(nldr_df = tSNE_data_harps, x = "tSNE1", y = "tSNE2", shape_val = shape_val_tsne_harps, non_empty_bins = num_non_empty_bins_harps)
+num_bins_tsne_harps <- 9
 
@@
 
+# distance_plot <- plot_dist(distance_tsne_harps) +
+#   #ggtitle("(b)" ) +
+#   ylab(expression(d^{(2)})) +
+#   theme(axis.text = element_text(size = 5),
+#         axis.title = element_text(size = 12))
+# 
+# distance_plot
+
 ## To find the benchmark value
-benchmark_tsne_harps <- find_benchmark_value(.data = distance_tsne_harps, distance_col = distance)
-benchmark_tsne_harps <- 1
+benchmark_tsne_harps <- find_benchmark_value(.data = distance_tsne_harps, distance_col = "distance")
+benchmark_tsne_harps <- 2.5
 

### 2024-02-10 paper.tex
Generate the true model for s-curve

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, enhanced, breakable, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, breakable, boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 indicated that the MSE remained consistent across different benchmark
-values, as shown in Figure~\ref{fig-diagnosticpltScurvelgrm}. It's
+values, as shown in \textbf{?@fig-diagnosticpltScurvelgrm}. It's
 important to note that edges are defined only in the 2D model and do not
@@
 of a smoother 2D representation (see
-Figure~\ref{fig-diagnosticpltScurvelgrm}).
+\textbf{?@fig-diagnosticpltScurvelgrm}).
 
@@
 long edges.}
-
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-diagnosticpltScurvelgrm-1.pdf}
-
-}
-
-\caption{\label{fig-diagnosticpltScurvelgrm}Goodness of fit statistics
-from UMAP applied to training S-curve dataset with different benchmark
-values to remove the long edges. What is the effective benchmark value
-to remove the long edges? \(1.347\) is the chosen benchmark value to
-remove long edges. The value is selected through a visual inspection of
-the triangular mesh is conducted to verify whether it successfully
-eliminates long edges.}
 
@@
 
-The approach involves employing the K-nearest neighbors (KNN) algorithm
-to identify the nearest hexagonal bin centroid in the 2D space.
-Subsequently, the coordinates of this centroid are assigned as the
-low-dimensional predicted values for the test data in 2D space. It is
-noteworthy that traditional NLDR methods, such as tSNE, often lack a
-direct predict function, making our approach valuable for generating
-predicted values in the absence of such functionalities.
+The prediction approach involves employing the K-nearest neighbors (KNN)
+algorithm. In this method, which operates as an unsupervised
+classification problem, the nearest high-D model point (averaged high-D
+point) is identified for the new high-D data point. Once the nearest
+high-D model point is identified, its coordinates are used as the
+predicted 2D embeddings for the new high-D data point. This step is
+particularly valuable due to the limitations of certain NLDR techniques,
+like tSNE, which don't provide a straightforward method for predicting
+2D embeddings. As a result, our approach presents a solution that can
+generate predicted 2D embeddings, irrespective of the NLDR technique
+employed, effectively addressing this gap in functionality.
 
@@
 
-Here, \(n\) represents the number of bins, \(p\) represents the
+Here, \(n\) represents the number of observations, \(p\) represents the
 dimensions, \(x_{ij}\) is the actual high-dimensional data, and
@@
 complexity in the predictions.
-
-\hypertarget{model-interpretation}{%
-\subsection{Model interpretation}\label{model-interpretation}}
-
-The visualizations of NLDR techniques for the S-curve dataset shows
-different representations in 2D, as showcased in Figure
-Figure~\ref{fig-nldervis}. The question at hand is determining the best
-NLDR technique to represent the S-curve data. Examining the 2D layouts
-in Figure~\ref{fig-nldervis} suggests that TriMAP or PaCMAP may be best
-representations. However, the challenge is to decide the best
-representation with considerable evidence.
-
-After considering the models constructed from the algorithm for each
-NLDR techniques and visualizing them in high-D space with the original
-data, shows that all the NLDR techniques capture the non-linearity of
-the S-curve (see videos of the models linked in
-\textbf{?@fig-modelScurve}). But thorough examination of each model in
-high-D space shows some surprising findings. The model of PHATE is
-flattened and lack of capturing the width of S-curve (see videos of the
-model with PHATE linked in \textbf{?@fig-modelScurve}). This nature of
-PHATE is also evicted by the considerably higher AIC and MSE values for
-PHATE as shown in Figure~\ref{fig-diagnosticpltScurve}. Furthermore,
-there are two edges that are not connected to any points in the model of
-TriMAP (see see videos of the model with TriMAP linked in
-\textbf{?@fig-modelScurve}). This model shows how the patterns in 2D
-representation of TriMAP capture by the model (see
-Figure~\ref{fig-nldervis} (d)). As shown in Figure~\ref{fig-nldervis}
-(b), UMAP have four clusters in a curvilinear structure. However, the
-model reveals that even though UMAP shows a curvilinear in 2D, UMAP
-actually capture the structure of the S-curve without any disjoint in
-the model (see see videos of the model with UMAP linked in
-\textbf{?@fig-modelScurve}). In the 2D layouts of tSNE and PaCMAP (see
-Figure~\ref{fig-nldervis} (a), (e)), there are some sparse areas that
-are captured by the models effectively (see see videos of the model with
-tSNE and PaCMAP linked in \textbf{?@fig-modelScurve}).
-
-\hypertarget{sec-prediction}{%
-\subsection{Prediction}\label{sec-prediction}}
-
-In this context, ``prediction'' denotes the 2D embedding generated for
-the NLDR technique. The methodology encompasses identifying the nearest
-averaged high-D points in the high-D space for the test data, by
-computing high-D Euclidean distances. As the averaged high-D point
-corresponds to the lifting of the 2D model, determining its nearest
-counterpart allows us to map its hexagonal bin centroid coordinates.
-Consequently, these centroid coordinates serve as the assigned
-low-dimensional predicted values for the test data in the 2D space.
-
-Some NLDR techniques, such as tSNE, often lack a direct prediction,
-making our approach valuable for generating predicted values in the
-absence of such functionalities.
 
@@
   this data
-\item
 \item
@@
 
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-diagnosticpltPBMC-1.pdf}
+
+}
+
+\caption{\label{fig-diagnosticpltPBMC}Goodness of fit statistics from
+UMAP applied to training S-curve dataset. What is the effective number
+of bins in each NLDR technique to create a 2D model? The MSE plot have a
+steep slope at the beginning, indicating that a smaller number of bins
+causes a larger amount of error. Then, the slope gradually declines or
+level off, indicating that a higher number of bins generates a smaller
+error. Using the elbow method, when the total number of bins is set to
+144, the slope of the Mean Squared Error (MSE) plot experiences a sudden
+and noticeable change, resembling an elbow-like shape. This point
+indicates that adding less bins does not enough to capture the data
+structure.}
+
+\end{figure}
+
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-55-1.pdf}
+
 \begin{figure}[h]
@@
 
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-diagnosticpltPBMC2-1.pdf}
+
+}
+
+\caption{\label{fig-diagnosticpltPBMC2}Goodness of fit statistics from
+UMAP applied to training S-curve dataset. What is the effective number
+of bins in each NLDR technique to create a 2D model? The MSE plot have a
+steep slope at the beginning, indicating that a smaller number of bins
+causes a larger amount of error. Then, the slope gradually declines or
+level off, indicating that a higher number of bins generates a smaller
+error. Using the elbow method, when the total number of bins is set to
+144, the slope of the Mean Squared Error (MSE) plot experiences a sudden
+and noticeable change, resembling an elbow-like shape. This point
+indicates that adding less bins does not enough to capture the data
+structure.}
+
+\end{figure}
+
 \begin{figure}[h]
@@
 Model in the 2D space with UMAP (\textless\textgreater)}
+
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-diagnosticpltPBMC3-1.pdf}
+
+}
+
+\caption{\label{fig-diagnosticpltPBMC3}Goodness of fit statistics from
+UMAP applied to training S-curve dataset. What is the effective number
+of bins in each NLDR technique to create a 2D model? The MSE plot have a
+steep slope at the beginning, indicating that a smaller number of bins
+causes a larger amount of error. Then, the slope gradually declines or
+level off, indicating that a higher number of bins generates a smaller
+error. Using the elbow method, when the total number of bins is set to
+144, the slope of the Mean Squared Error (MSE) plot experiences a sudden
+and noticeable change, resembling an elbow-like shape. This point
+indicates that adding less bins does not enough to capture the data
+structure.}
 
@@
 
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-diagnosticpltHARPS-1.pdf}
+
+}
+
+\caption{\label{fig-diagnosticpltHARPS}Goodness of fit statistics from
+UMAP applied to training S-curve dataset. What is the effective number
+of bins in each NLDR technique to create a 2D model? The MSE plot have a
+steep slope at the beginning, indicating that a smaller number of bins
+causes a larger amount of error. Then, the slope gradually declines or
+level off, indicating that a higher number of bins generates a smaller
+error. Using the elbow method, when the total number of bins is set to
+144, the slope of the Mean Squared Error (MSE) plot experiences a sudden
+and noticeable change, resembling an elbow-like shape. This point
+indicates that adding less bins does not enough to capture the data
+structure.}
+
+\end{figure}
+
 \begin{figure}[h]
@@
 
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-diagnosticpltHARPS2-1.pdf}
+
+}
+
+\caption{\label{fig-diagnosticpltHARPS2}Goodness of fit statistics from
+UMAP applied to training S-curve dataset. What is the effective number
+of bins in each NLDR technique to create a 2D model? The MSE plot have a
+steep slope at the beginning, indicating that a smaller number of bins
+causes a larger amount of error. Then, the slope gradually declines or
+level off, indicating that a higher number of bins generates a smaller
+error. Using the elbow method, when the total number of bins is set to
+144, the slope of the Mean Squared Error (MSE) plot experiences a sudden
+and noticeable change, resembling an elbow-like shape. This point
+indicates that adding less bins does not enough to capture the data
+structure.}
+
+\end{figure}
+
 \begin{figure}[h]
@@
 
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-diagnosticpltHARPS3-1.pdf}
+
+}
+
+\caption{\label{fig-diagnosticpltHARPS3}Goodness of fit statistics from
+UMAP applied to training S-curve dataset. What is the effective number
+of bins in each NLDR technique to create a 2D model? The MSE plot have a
+steep slope at the beginning, indicating that a smaller number of bins
+causes a larger amount of error. Then, the slope gradually declines or
+level off, indicating that a higher number of bins generates a smaller
+error. Using the elbow method, when the total number of bins is set to
+144, the slope of the Mean Squared Error (MSE) plot experiences a sudden
+and noticeable change, resembling an elbow-like shape. This point
+indicates that adding less bins does not enough to capture the data
+structure.}
+
+\end{figure}
+
 \begin{figure}[h]

### 2024-02-10 extra/extra_1.R

@@ -1,0 +1,219 @@
+```{r}
+#| warning: false
+#| echo: false
+
+
+
+## tSNE
+
+shape_value_curve <- calculate_effective_shape_value(.data = tSNE_data,
+                                                     x = tSNE1, y = tSNE2)
+
+num_bins_vec <- 1:18 ## Number of bins along the x-axis
+
+vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(num_bins_vec)) {
+
+  model_object <- fit_high_d_model(training_data = training_data, nldr_df_with_id = tSNE_data, x = "tSNE1", y = "tSNE2", num_bins_x = num_bins_vec[i], shape_val = shape_value_curve,
+                                   is_bin_centroid = TRUE,
+                                   is_rm_lwd_hex = FALSE,
+                                   benchmark_to_rm_lwd_hex = NA,
+                                   is_avg_high_d = TRUE, column_start_text = "x")
+
+  centroid_df_training <- model_object$df_bin_centroids
+  avg_df_training <- model_object$df_bin
+
+  # pred_df_training <- predict_2d_embeddings(test_data = training_data, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "tSNE")
+
+  pred_df_training <- predict_hex_id(df_bin_centroids = centroid_df_training, nldr_df_test = tSNE_data, x = "tSNE1", y = "tSNE2")
+
+
+  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, col_start = "x")
+
+  eval_df_training <- eval_df_training |>
+    mutate(num_bins_x = num_bins_vec[i])
+
+  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+
+
+}
+
+
+## Add new column with data types
+
+eval_data_training <- eval_data_training |>
+  mutate(data_type = "training")
+
+MSE_df_1 <- eval_data_training |>
+  dplyr::mutate(method = "tSNE")
+```
+```{r}
+#| warning: false
+#| echo: false
+## PAHTE
+## Prediction
+
[…]

### 2024-02-10 scripts/1_true_model_with_hex_8_S_curve.R

@@ -38,7 +38,7 @@ tr_from_to_df_umap_s_curve <- generate_edge_info(triangular_object = tr1_object_
 
 
 ## Compute 2D distances
-distance_umap_s_curve <- cal_2D_dist(.data = tr_from_to_df_umap_s_curve)
+distance_umap_s_curve <- cal_2d_dist(.data = tr_from_to_df_umap_s_curve)
 
 
 ## To find the benchmark value

### 2024-02-12 paper.qmd
Added new spin-brush trials

@@
 library(grid)
+
+# ## For paralleization
+# library(foreach)
+# library(doParallel)
+
 # library(xtable)
@@
 
+The following steps will help to find a suitable value to remove low density hexagons:
+
+1. Plot the distribution of the standardized counts
+
+2. See the distribution of counts
+
+3. Take the first quartile
+
 ```{r}
@@
 df_bin_centroids_umap_s_curve <- hexbin_data_object_umap_s_curve$hexdf_data
+
+cell_count_plot_s_curve <- ggplot(df_bin_centroids_umap_s_curve, aes(x = reorder(as.factor(hexID), -std_counts), y = std_counts)) +
+  geom_quasirandom() + xlab("hexagonal id") + ylab("Standardized cell count") +
+  geom_hline(yintercept = 0.20, colour = "#de2d26") +
+  theme(axis.text = element_text(size = 5),
+        axis.title = element_text(size = 7),
+        axis.text.x = element_text(angle = 90))
 
@@
 
+The following steps will help to find a suitable value to remove long edges:
+
+1. Plot the distribution of the 2D distances
+
+2. Find a value which is greater than smallest value
+
 <!--The removal of long edges is a critical step to create a smoother representation  by iteratively eliminating hexagons with excessive distances between centroids.
@@
 distance_umap_s_curve <- cal_2d_dist(.data = tr_from_to_df_umap_s_curve)
+
+distance_plot_s_curve <- plot_dist(distance_umap_s_curve) +
+  #ggtitle("(b)" ) +
+  ylab(expression(d^{(2)})) +
+  theme(axis.text = element_text(size = 5),
+        axis.title = element_text(size = 12))
 
@@
 
+```
+
+```{r}
+#| warning: false
+#| echo: false
+#| label: fig-modelScurverdistcount
+#| fig-cap: "plots to identify benchmark values"
+##| out-width: 100%
+#| fig-pos: H
+
+cell_count_plot_s_curve + distance_plot_s_curve +
+  #plot_annotation(tag_levels = 'a') +
+  plot_layout(ncol = 2) &
+  theme(legend.position='none', plot.tag = element_text(size = 8))
 ```

### 2024-02-12 paper.tex
Added new spin-brush trials

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, breakable, boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, breakable, interior hidden, enhanced, sharp corners, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
 
@@
 
+The following steps will help to find a suitable value to remove low
+density hexagons:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  Plot the distribution of the standardized counts
+\item
+  See the distribution of counts
+\item
+  Take the first quartile
+\end{enumerate}
+
 \begin{figure}[H]
@@
 \textbf{?@fig-diagnosticpltScurvelgrm}).
+
+The following steps will help to find a suitable value to remove long
+edges:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  Plot the distribution of the 2D distances
+\item
+  Find a value which is greater than smallest value
+\end{enumerate}
+
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-modelScurverdistcount-1.pdf}
+
+}
+
+\caption{\label{fig-modelScurverdistcount}plots to identify benchmark
+values}
+
+\end{figure}
 
@@
 
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-55-1.pdf}
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-56-1.pdf}
 

### 2024-02-12 examples/sphere/2_gen_diff_emb_sphere.R

@@ -1,0 +1,248 @@
+library(readr)
+
+set.seed(20240110)
+
+source("quollr_code.R", local = TRUE)
+source("nldr_code.R", local = TRUE)
+
+## Import data
+training_data <- read_rds("data/sphere/sphere_training.rds")
+UMAP_data_sphere <- read_rds("data/sphere/sphere_umap.rds")
+
+### tSNE
+tSNE_data_sphere <- Fit_tSNE(sphere_df |> distinct(), opt_perplexity = calculate_effective_perplexity(sphere_df |> distinct()),
+                          with_seed = 20240110)
+
+tSNE_data_sphere <- tSNE_data_sphere |>
+  select(-ID) |>
+  mutate(ID = training_data_5$ID)
+
+plot_tSNE_2D(tSNE_data_sphere)
+
+### UMAP
+
+UMAP_fit <- umap(sphere_df |> distinct(), n_neighbors = 100, n_components =  2)
+
+UMAP_data_sphere <- UMAP_fit$layout |>
+  as.data.frame()
+names(UMAP_data_sphere)[1:(ncol(UMAP_data_sphere))] <- paste0(rep("UMAP",(ncol(UMAP_data_sphere))), 1:(ncol(UMAP_data_sphere)))
+
+UMAP_data_sphere <- UMAP_data_sphere |>
+  mutate(ID = training_data_5$ID)
+
+plot_UMAP_2D(UMAP_data_sphere)
+
+### Phate
+
+PHATE_data <- Fit_PHATE(sphere_df |> distinct(), knn = 5, with_seed = 20240110)
+PHATE_data <- PHATE_data |>
+  select(PHATE1, PHATE2)
+PHATE_data <- PHATE_data |>
+  mutate(ID = training_data_5$ID)
+
+plot_PHATE_2D(PHATE_data)
+
+### TriMAP
+
+tem_dir <- tempdir()
+
+Fit_TriMAP_data(sphere_df |> distinct(), tem_dir)
+
+path <- file.path(tem_dir, "df_2_without_class.csv")
+path2 <- file.path(tem_dir, "dataset_3_TriMAP_values.csv")
+
+Fit_TriMAP(as.integer(2), as.integer(5), as.integer(4), as.integer(3), path, path2)
+
+TriMAP_data <- read_csv(path2)
+TriMAP_data <- TriMAP_data |>
+  mutate(ID = training_data_5$ID)
+
[…]

### 2024-02-13 paper.qmd
Added true label manually

@@
 # num_bins_pbmc <- find_non_empty_bins(nldr_df = UMAP_pbmc, x = "UMAP1", y = "UMAP2", shape_val = shape_value_pbmc, non_empty_bins = num_non_empty_bins_pbmc)
-num_bins_pbmc <- 12
+num_bins_pbmc <- 16
 
@@
 Is there a best representation of the original data or are they all providing  equivalent information?
+```
+<!--Generate true label for PBMC3k-->
+
+```{r}
+#| warning: false
+#| echo: false
+
+UMAP_pbmc_with_true_label <- UMAP_pbmc |>
+  dplyr::mutate(true_class_label = case_when(
+    ((UMAP1 >=-12) & (UMAP1 <= -5) & (UMAP2 >= 6) & (UMAP2 <= 18)) ~ "cluster 1",
+    ((UMAP1 >-5) & (UMAP1 <= 7) & (UMAP2 >= 4.8) & (UMAP2 <= max(UMAP2))) ~ "cluster 2", ((UMAP1 >=6) & (UMAP1 <= 15) & (UMAP2 >= -4) & (UMAP2 < 4.8)) ~ "cluster 3",
+    ((UMAP1 >=-3) & (UMAP1 <= 6.6) & (UMAP2 >= min(UMAP2)) & (UMAP2 <= 4.8)) ~ "cluster 4",
+    ((UMAP1 >=min(UMAP1))& (UMAP1 < -3) & (UMAP2 >= -8) & (UMAP2 <= -1)) ~ "cluster 5",
+    ((UMAP1 >=min(UMAP1)) & (UMAP1 <= -7) & (UMAP2 >= -3) & (UMAP2 <= 3)) ~ "cluster 6",
+    .default = "no class"
+    ))
+
+### Reassign some labels
+
+UMAP_pbmc_with_true_label <- UMAP_pbmc_with_true_label |>
+  dplyr::mutate(true_class_label = if_else((ID %in% c(2619, 1158)), "cluster 3", true_class_label)) |>
+  dplyr::mutate(true_class_label = if_else((ID %in% c(948, 285, 158, 268, 731, 1152, 1372, 1651, 2010, 150, 2304, 2546, 487, 2176, 2157, 2451, 264, 63, 1234, 947, 1126, 1699, 1232, 1492, 1488, 2040, 2550, 1313, 361, 372, 1515, 1264, 1307, 1726, 517, 2568, 2274, 591, 1703, 1097, 430, 1556, 2133, 1232, 1194, 1888, 1796, 494, 1194)), "cluster 6", true_class_label))
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+UMAP_pbmc_with_true_label |>
+    ggplot(aes(x = UMAP1,
+               y = UMAP2, color = true_class_label))+
+    geom_point(alpha=0.5) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#66c2a5", "#fc8d62", "#8da0cb", "#e78ac3", "#a6d854", "#ffd92f", "black")) +
+  # geom_text(aes(x=UMAP1,y = UMAP2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
 ```

### 2024-02-13 paper.qmd
added confusion matrix with 6 classes and new detour

@@
          legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#66c2a5", "#fc8d62", "#8da0cb", "#e78ac3", "#a6d854", "#ffd92f", "black")) +
+  scale_color_manual(values=c("#d82cd3", "#49df20", "#619cff", "#000000", "#df2020", "#ffbd61", "yellow")) +
   # geom_text(aes(x=UMAP1,y = UMAP2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) +
   annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+## Import detour results
+detour_results8 <- read_csv("data/pbmc/pbmc_3k_festem/detourr_export8.csv")
+detour_results8 <- detour_results8 |>
+  dplyr::mutate(ID = training_data_pbmc$ID)
+
+table(UMAP_pbmc_with_true_label$true_class_label, detour_results8$colour)
 

### 2024-02-20 paper.qmd
Added cluster label from result 1 and result 2, side by side 2D layout coloured by labels in result1, result2

--- 2024-02-12 paper.qmd
@@
 Is there a best representation of the original data or are they all providing  equivalent information?
+```
+<!--Generate true label for PBMC3k-->
+
+```{r}
+#| warning: false
+#| echo: false
+
+UMAP_pbmc_with_result1_label <- UMAP_pbmc |>
+  dplyr::mutate(result1_class_label = case_when(
+    ((UMAP1 >=-12) & (UMAP1 <= -5) & (UMAP2 >= 6) & (UMAP2 <= 18)) ~ "cluster 1",
+    ((UMAP1 >-5) & (UMAP1 <= 7) & (UMAP2 >= 4.8) & (UMAP2 <= max(UMAP2))) ~ "cluster 2", ((UMAP1 >=6) & (UMAP1 <= 15) & (UMAP2 >= -4) & (UMAP2 < 4.8)) ~ "cluster 3",
+    ((UMAP1 >=-3) & (UMAP1 <= 6.6) & (UMAP2 >= min(UMAP2)) & (UMAP2 <= 4.8)) ~ "cluster 4",
+    ((UMAP1 >=min(UMAP1))& (UMAP1 < -3) & (UMAP2 >= -8) & (UMAP2 <= -1)) ~ "cluster 5",
+    ((UMAP1 >=min(UMAP1)) & (UMAP1 <= -7) & (UMAP2 >= -3) & (UMAP2 <= 3)) ~ "cluster 6",
+    .default = "no class"
+    ))
+
+### Reassign some labels
+
+UMAP_pbmc_with_result1_label <- UMAP_pbmc_with_result1_label |>
+  dplyr::mutate(result1_class_label = if_else((ID %in% c(2619, 1158)), "cluster 3", result1_class_label)) |>
+  dplyr::mutate(result1_class_label = if_else((ID %in% c(948, 285, 158, 268, 731, 1152, 1372, 1651, 2010, 150, 2304, 2546, 487, 2176, 2157, 2451, 264, 63, 1234, 947, 1126, 1699, 1232, 1492, 1488, 2040, 2550, 1313, 361, 372, 1515, 1264, 1307, 1726, 517, 2568, 2274, 591, 1703, 1097, 430, 1556, 2133, 1232, 1194, 1888, 1796, 494, 1194)), "cluster 6", result1_class_label))
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## Import detour results
+detour_results8 <- read_csv("data/pbmc/pbmc_3k_festem/detourr_export8.csv")
+
+### Rename labels
+
+detour_results8 <- detour_results8 |>
+  dplyr::mutate(result2_class_label = case_when(
+    colour == "000000" ~ "cluster 4",
+    colour == "49df20" ~ "cluster 2",
+    colour == "619cff" ~ "cluster 3",
+    colour == "d82cd3" ~ "cluster 1",
+    colour == "df2020" ~ "cluster 5",
+    colour == "ffbd61" ~ "cluster 6",
+    .default = "no cluster"
+
+  ))
+
+detour_results8 <- detour_results8 |>
+  dplyr::mutate(ID = training_data_pbmc$ID)
+
+df_res_1_2_label <- inner_join(UMAP_pbmc_with_result1_label, detour_results8, by = c("ID" = "ID"))
+
+result1_2d_layout <- df_res_1_2_label |>
+    ggplot(aes(x = UMAP1,
+               y = UMAP2, color = result1_class_label))+
+    geom_point(alpha=0.5) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#d82cd3", "#49df20", "#619cff", "#000000", "#df2020", "#ffbd61", "yellow")) +
+  # geom_text(aes(x=UMAP1,y = UMAP2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+result2_2d_layout <- df_res_1_2_label |>
+    ggplot(aes(x = UMAP1,
+               y = UMAP2, color = result2_class_label))+
+    geom_point(alpha=0.5) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#d82cd3", "#49df20", "#619cff", "#000000", "#df2020", "#ffbd61", "yellow")) +
+  # geom_text(aes(x=UMAP1,y = UMAP2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+table(df_res_1_2_label$result1_class_label, df_res_1_2_label$result2_class_label)
+
+result1_2d_layout + result2_2d_layout
 ```

### 2024-02-20 paper.tex
Added cluster label from result 1 and result 2, side by side 2D layout coloured by labels in result1, result2

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, breakable, interior hidden, enhanced, sharp corners, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, boxrule=0pt, sharp corners, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 
+\begin{verbatim}
+
+            cluster 1 cluster 2 cluster 3 cluster 4 cluster 5 cluster 6
+  cluster 1       138        13         0         0         0         0
+  cluster 2        11       509         0         1         0         0
+  cluster 3         1         0       348         1         0         1
+  cluster 4         0         0         3      1029        64         2
+  cluster 5         0         0         0        85       226        32
+  cluster 6         0         0         0         0        19       139
+\end{verbatim}
+
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-71-1.pdf}
+
 \hypertarget{harps-gto-sample-dataset}{%

### 2024-03-05 appendix.qmd
Added mse plot, aic plot for pbmc need to find to remove duplicates of mse_df

@@
 ```
+
+### HARPS GTO sample dataset
+
+@Anders2018
+
+- In 2D, there are two well-separated clusters, but the clusters are really close
+
+- One cluster have a sparse data space within the cluster
+
+- There is a line in right corner of the structure that is not connected any other trimesh points. This line create a long edge in high-D space.
+
+- In high-D view, the clusters having some non-linearity patter, which is also not captured in tSNE.
+Therefore, this is not a good choice.
+Also, the two clusters are really near to each other and can't see any separation.
+Also continuity of the structure.
+Long edges are presence also.
+
+- The parameter is, in a sense, a guess about the number of close neighbors each point has.
+
+- Num bins along the x-axis = 43, shape parameter = 0.4466309
+
+```{r}
+#| warning: false
+#| echo: false
+
+error_df <- read_csv("data/harps/error_harps_tsne.csv", col_names = FALSE)
+
+error_df <- error_df |>
+  mutate(X2 = round(X2/100, 2),
+         X1 = as.integer(X1)) |>
+  arrange(X2)
+
+colnames(error_df) <- c("perplexity", "error (x 100)")
+
+kable(error_df, caption = "errors for different parameter combinations", label = "tbl-errordiffcombharps") |>
+  #xtable2kable() |>
+  kable_styling(latex_options = c("striped", "HOLD_position"), position = "center", full_width = FALSE) |>
+  row_spec(0, bold = TRUE) |>
+  row_spec(3, bold=T, color = "red")
+```
+
+<!-- Author suggestion-->
+```{r}
+#| warning: false
+#| echo: false
+
+tSNE_data_harps <- read_rds("data/harps/harps_tsne_40.rds")
+
+plot_list2_harps <- tSNE_data_harps |>
+    ggplot(aes(x = tSNE1,
+               y = tSNE2, color = cluster))+
+    geom_point(alpha=0.5) +
+    #coord_cartesian(xlim = c(-25, 25), ylim = c(-12, 12)) +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#b15928", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#cab2d6", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#bc80bd", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
+  guides(color = guide_legend(title = "cluster"),
+         color = guide_legend(nrow = 3)) +
+ annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## Import data
+training_data_1 <- read_rds("data/harps/harps_data.rds")
+training_data_1 <- training_data_1 |>
+  mutate(ID = 1:NROW(training_data_1))
+
+clusters <- training_data_1$cluster
+select_names <-  names(training_data_1)[-c(18, length(training_data_1))]
+
+training_data_1 <- training_data_1 |> select(-cluster)
+names(training_data_1) <- append(paste0("x", 1:17), "ID")
+
+tSNE_data_harps <- tSNE_data_harps |>
+  select(tSNE1, tSNE2) |>
+  mutate(ID = 1:NROW(tSNE_data_harps))
+
+## tSNE
+## Prediction
+
+shape_val_tsne_harps <- calculate_effective_shape_value(.data = tSNE_data_harps,
+                                                      x = tSNE1, y = tSNE2)
+
+num_bins_vec <- 1:21 ## Number of bins along the x-axis
+
+vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(num_bins_vec)) {
+
+  model_object <- fit_high_d_model(training_data = training_data_1, nldr_df_with_id = tSNE_data_harps, x = "tSNE1", y = "tSNE2", num_bins_x = num_bins_vec[i], shape_val = shape_val_tsne_harps,
+                             is_bin_centroid = TRUE,
+                             is_rm_lwd_hex = FALSE,
+                             benchmark_to_rm_lwd_hex = NA,
+                             is_avg_high_d = TRUE, column_start_text = "x")
+
+  centroid_df_training <- model_object$df_bin_centroids
+  avg_df_training <- model_object$df_bin
+
+  pred_df_training <- predict_2d_embeddings(test_data = training_data_1, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "tSNE")
+
+  eval_df_training <- generate_eval_df(data = training_data_1, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, col_start = "x")
+
+  eval_df_training <- eval_df_training |>
+    mutate(num_bins_x = num_bins_vec[i])
+
+  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+
+}
+
+## Add new column with data types
+
+MSE_df_harps <- eval_data_training |>
+  dplyr::mutate(data_type = "training",
+                method = "tSNE")
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## To draw with AIC
+aic_plot_harps <- ggplot(MSE_df_harps |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "tSNE"), aes(x = number_of_bins,
+                                                                                 y = total_error,
+                                                                                 color = method
+)) +
+  geom_point() +
+  geom_line() +
+  geom_vline(xintercept = 120, linetype="dashed", 
+                color = "black", size=0.5) +
+  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  scale_colour_manual(values = c("#e41a1c","#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("AIC") +
+  xlab("total number of bins")
+## Effective number of bins along x-axis
+
+mse_plot_harps <- ggplot(MSE_df_harps |> dplyr::filter(method == "tSNE"), aes(x = number_of_bins,
+                                       y = total_mse,
+                                       color = method
+)) +
+  geom_point() +
+  geom_line() +
+   geom_vline(xintercept = 120, linetype="dashed", 
+                color = "black", size=0.5) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  scale_colour_manual(values = c("#e41a1c","#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("MSE") +
+  xlab("total number of bins")
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset.
+What is the effective number of bins in each NLDR technique to create a 2D model?
+The MSE plot have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
+Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error.
+Using the elbow method, when the total number of bins is set to 144, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
+This point indicates that adding less bins does not enough to capture the data structure."
+#| label: fig-diagnosticpltHARPS
+##| out-width: 100%
+#| fig-pos: H
+
+aic_plot_harps + mse_plot_harps +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='none', plot.tag = element_text(size = 8))
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+### tSNE
+
+num_bins_tsne_harps <- calculate_effective_x_bins(.data = tSNE_data_harps, x = "tSNE1",
+                                                cell_area = 1) #43
+
+num_bins_tsne_harps <- 14
+
+shape_val_tsne_harps <- calculate_effective_shape_value(.data = tSNE_data_harps,
+                                                      x = "tSNE1", y = "tSNE2")
+## To extract bin centroids
+hexbin_data_object_tsne_harps <- extract_hexbin_centroids(nldr_df = tSNE_data_harps, num_bins = num_bins_tsne_harps, shape_val = shape_val_tsne_harps, x = "tSNE1", y = "tSNE2")
+
+df_bin_centroids_tsne_harps <- hexbin_data_object_tsne_harps$hexdf_data
+
+#num_non_empty_bins_harps <- df_bin_centroids_tsne_harps$hexID |> length()
+
+## Identify bins with low-density
+# identify_rm_bins <- find_low_density_hexagons(df_bin_centroids_tsne_harps, num_bins_tsne_harps, benchmark_rm_hex = 0.01)
+# 
+# df_bin_centroids_tsne_harps <- df_bin_centroids_tsne_harps |>
+#   filter(!(hexID %in% identify_rm_bins))
+
+tSNE_data_with_hb_id_harps <- tSNE_data_harps |>
+  dplyr::mutate(hb_id = hexbin_data_object_tsne_harps$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_tsne_harps <- dplyr::bind_cols(training_data_1 |> dplyr::select(-ID), tSNE_data_with_hb_id_harps)
+
+## Averaged on high-D
+df_bin_tsne_harps <- avg_highD_data(.data = df_all_tsne_harps)
+
+## Triangulate bin centroids
+tr1_object_tsne_harps <- triangulate_bin_centroids(df_bin_centroids_tsne_harps, x, y)
+tr_from_to_df_tsne_harps <- generate_edge_info(triangular_object = tr1_object_tsne_harps)
+
+## Compute 2D distances
+distance_tsne_harps <- cal_2d_dist(.data = tr_from_to_df_tsne_harps)
+
+## To find the benchmark value
+benchmark_tsne_harps <- find_benchmark_value(.data = distance_tsne_harps, distance_col = "distance")
+benchmark_tsne_harps <- 5
+
+# distance_plot <- plot_dist(distance_tsne_harps) +
+#   #ggtitle("(b)" ) +
+#   ylab(expression(d^{(2)})) +
+#   theme(axis.text = element_text(size = 5),
+#         axis.title = element_text(size = 12))
+# 
+# distance_plot
+
+trimesh_removed_tsne_harps <- remove_long_edges(.data = distance_tsne_harps, benchmark_value = benchmark_tsne_harps,
+                                              triangular_object = tr1_object_tsne_harps, distance_col = distance)
+
+trimesh_removed_tsne_harps <- trimesh_removed_tsne_harps +
+  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  # theme(axis.text = element_text(size = 5),
+  #       axis.title = element_text(size = 7)) +
+  geom_point(colour = "#33a02c", size = 0.5) +
+  #coord_cartesian(xlim = c(-25, 25), ylim = c(-12, 12)) +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+df_bin_train <- df_bin_tsne_harps
+names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])
+
+error_df <- df_all_tsne_harps |>
+  dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates
+
+# prediction_df_join <- prediction_df_join |>
+#   dplyr::left_join(data, by = c("ID" = "ID")) ## Map high-D data
+
+for (i in 1:(NCOL(df_bin_train) - 1)) {
+
+  error_df[ , paste0("abs_residual_", "x", i)] <- abs(error_df[ , paste0("x", i)] - error_df[ , paste0("avg_", "x", i)])
+
+}
+
+error_df <- error_df |>
+  dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "x")))))
+
+error_df <- error_df |>
+  mutate(type = if_else(total <= 0, "no error",
+                        if_else(total <= 0.2, "error 0-0.2",
+                                if_else(total <= 0.3, "error 0.2-0.3",
+                                        if_else(total <= 0.5, "error 0.3-0.5",
+                                                if_else(total <= 1, "error 0.5-1",
+                                                        "error greter than 1"))))))
+
+plot_list2_harps_error <- error_df |>
+    mutate(type = factor(type , levels = c("no error", "error 0-0.2", "error 0.2-0.3", "error 0.3-0.5", "error 0.5-1", "error greter than 1"))) |>
+    ggplot(aes(x = tSNE1,
+               y = tSNE2, color = type))+
+    geom_point(alpha=0.5, size = 0.1) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#fee0d2", "#fcbba1",
+                                         "#fc9272", "#fb6a4a", "#ef3b2c",
+                                         "#cb181d", "#a50f15", "#99000d"))
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "(a) 2D layout from tSNE (perplexity = 40) applied for the HARPS GTO dataset.
+Is this a best representation of the original data?, (b) Model in the 2D space with tSNE (<https://youtu.be/jxo_rPyGv9k>)"
+#| label: fig-nldervisharpstSNE
+#| fig-pos: h
+
+plot_list2_harps + trimesh_removed_tsne_harps + plot_list2_harps_error +
+  plot_layout(guides='collect', ncol = 3) &
+  theme(legend.position='bottom', plot.tag = element_text(size = 8))
+
+```
+
+<!-- param set 1-->
+
+- 5 clusters shows in 2D
+
+- 3 clusters which are really close shows in high_d data with the model
+
+```{r}
+#| warning: false
+#| echo: false
+
+tSNE_data_harps <- read_rds("data/harps/harps_tsne_23.rds")
+
+plot_list2_harps <- tSNE_data_harps |>
+    ggplot(aes(x = tSNE1,
+               y = tSNE2, color = clusters))+
+    geom_point(alpha=0.5) +
+    coord_cartesian(xlim = c(-25, 20), ylim = c(-35, 25)) +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#b15928", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#cab2d6", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#bc80bd", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
+  guides(color = guide_legend(title = "cluster"),
+         color = guide_legend(nrow = 3)) +
+ annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## tSNE
+## Prediction
+
+shape_val_tsne_harps <- calculate_effective_shape_value(.data = tSNE_data_harps,
+                                                      x = tSNE1, y = tSNE2)
+
+num_bins_vec <- 1:21 ## Number of bins along the x-axis
+
+vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(num_bins_vec)) {
+
+  model_object <- fit_high_d_model(training_data = training_data_1, nldr_df_with_id = tSNE_data_harps, x = "tSNE1", y = "tSNE2", num_bins_x = num_bins_vec[i], shape_val = shape_val_tsne_harps,
+                             is_bin_centroid = TRUE,
+                             is_rm_lwd_hex = FALSE,
+                             benchmark_to_rm_lwd_hex = NA,
+                             is_avg_high_d = TRUE, column_start_text = "x")
+
+  centroid_df_training <- model_object$df_bin_centroids
+  avg_df_training <- model_object$df_bin
+
+  pred_df_training <- predict_2d_embeddings(test_data = training_data_1, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "tSNE")
+
+  eval_df_training <- generate_eval_df(data = training_data_1, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, col_start = "x")
+
+  eval_df_training <- eval_df_training |>
+    mutate(num_bins_x = num_bins_vec[i])
+
+  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+
+}
+
+## Add new column with data types
+
+MSE_df_harps <- eval_data_training |>
+  dplyr::mutate(data_type = "training",
+                method = "tSNE")
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## To draw with AIC
+aic_plot_harps <- ggplot(MSE_df_harps |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "tSNE"), aes(x = number_of_bins,
+                                                                                 y = total_error,
+                                                                                 color = method
+)) +
+  geom_point() +
+  geom_line() +
+  geom_vline(xintercept = 160, linetype="dashed", 
+                color = "black", size=0.5) +
+  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  scale_colour_manual(values = c("#e41a1c","#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("AIC") +
+  xlab("total number of bins")
+## Effective number of bins along x-axis
+
+mse_plot_harps <- ggplot(MSE_df_harps |> dplyr::filter(method == "tSNE"), aes(x = number_of_bins,
+                                       y = total_mse,
+                                       color = method
+)) +
+  geom_point() +
+  geom_line() +
+   geom_vline(xintercept = 160, linetype="dashed", 
+                color = "black", size=0.5) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  scale_colour_manual(values = c("#e41a1c","#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("MSE") +
+  xlab("total number of bins")
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset.
+What is the effective number of bins in each NLDR technique to create a 2D model?
+The MSE plot have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
+Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error.
+Using the elbow method, when the total number of bins is set to 144, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
+This point indicates that adding less bins does not enough to capture the data structure."
+#| label: fig-diagnosticpltHARPS2
+##| out-width: 100%
+#| fig-pos: H
+
+aic_plot_harps + mse_plot_harps +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='none', plot.tag = element_text(size = 8))
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## Import data
+# training_data_1 <- read_rds("data/harps/harps_data.rds")
+# training_data_1 <- training_data_1 |>
+#   mutate(ID = 1:NROW(training_data_1))
+# 
+# clusters <- training_data_1$cluster
+# select_names <-  names(training_data_1)[-c(18, length(training_data_1))]
+# 
+# training_data_1 <- training_data_1 |> select(-cluster)
+# names(training_data_1) <- append(paste0("x", 1:17), "ID")
+# 
+# tSNE_data_harps <- tSNE_data_harps |>
+#   select(tSNE1, tSNE2) |>
+#   mutate(ID = 1:NROW(tSNE_data_harps))
+
+### tSNE
+
+shape_val_tsne_harps <- calculate_effective_shape_value(.data = tSNE_data_harps,
+                                                      x = tSNE1, y = tSNE2)
+
+# num_bins_tsne_harps <- find_non_empty_bins(nldr_df = tSNE_data_harps, x = "tSNE1", y = "tSNE2", shape_val = shape_val_tsne_harps, non_empty_bins = num_non_empty_bins_harps)
+
+num_bins_tsne_harps <- 9
+
+## To extract bin centroids
+hexbin_data_object_tsne_harps <- extract_hexbin_centroids(nldr_df = tSNE_data_harps, num_bins = num_bins_tsne_harps, shape_val = shape_val_tsne_harps, x = tSNE1, y = tSNE2)
+
+df_bin_centroids_tsne_harps <- hexbin_data_object_tsne_harps$hexdf_data
+
+## Identify bins with low-density
+# identify_rm_bins <- find_low_density_hexagons(df_bin_centroids_tsne_harps, num_bins_tsne_harps, benchmark_rm_hex = 0.01)
+# 
+# df_bin_centroids_tsne_harps <- df_bin_centroids_tsne_harps |>
+#   filter(!(hexID %in% identify_rm_bins))
+
+tSNE_data_with_hb_id_harps <- tSNE_data_harps |>
+  dplyr::mutate(hb_id = hexbin_data_object_tsne_harps$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_tsne_harps <- dplyr::bind_cols(training_data_1 |> dplyr::select(-ID), tSNE_data_with_hb_id_harps)
+
+## Averaged on high-D
+df_bin_tsne_harps <- avg_highD_data(.data = df_all_tsne_harps)
+
+## Triangulate bin centroids
+tr1_object_tsne_harps <- triangulate_bin_centroids(df_bin_centroids_tsne_harps, x, y)
+tr_from_to_df_tsne_harps <- generate_edge_info(triangular_object = tr1_object_tsne_harps)
+
+## Compute 2D distances
+distance_tsne_harps <- cal_2d_dist(.data = tr_from_to_df_tsne_harps)
+
+# distance_plot <- plot_dist(distance_tsne_harps) +
+#   #ggtitle("(b)" ) +
+#   ylab(expression(d^{(2)})) +
+#   theme(axis.text = element_text(size = 5),
+#         axis.title = element_text(size = 12))
+# 
+# distance_plot
+
+## To find the benchmark value
+benchmark_tsne_harps <- find_benchmark_value(.data = distance_tsne_harps, distance_col = "distance")
+benchmark_tsne_harps <- 6
+
+trimesh_removed_tsne_harps <- remove_long_edges(.data = distance_tsne_harps, benchmark_value = benchmark_tsne_harps,
+                                              triangular_object = tr1_object_tsne_harps, distance_col = distance)
+
+trimesh_removed_tsne_harps <- trimesh_removed_tsne_harps +
+  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  # theme(axis.text = element_text(size = 5),
+  #       axis.title = element_text(size = 7)) +
+  geom_point(colour = "#33a02c", size = 0.5) +
+  coord_cartesian(xlim = c(-25, 20), ylim = c(-35, 25)) +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+df_bin_train <- df_bin_tsne_harps
+names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])
+
+error_df <- df_all_tsne_harps |>
+  dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates
+
+# prediction_df_join <- prediction_df_join |>
+#   dplyr::left_join(data, by = c("ID" = "ID")) ## Map high-D data
+
+for (i in 1:(NCOL(df_bin_train) - 1)) {
+
+  error_df[ , paste0("abs_residual_", "x", i)] <- abs(error_df[ , paste0("x", i)] - error_df[ , paste0("avg_", "x", i)])
+
+}
+
+error_df <- error_df |>
+  dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "x")))))
+
+error_df <- error_df |>
+  mutate(type = if_else(total <= 0, "no error",
+                        if_else(total <= 0.2, "error 0-0.2",
+                                if_else(total <= 0.3, "error 0.2-0.3",
+                                        if_else(total <= 0.5, "error 0.3-0.5",
+                                                if_else(total <= 1, "error 0.5-1",
+                                                        "error greter than 1"))))))
+
+plot_list2_harps_error <- error_df |>
+    mutate(type = factor(type , levels = c("no error", "error 0-0.2", "error 0.2-0.3", "error 0.3-0.5", "error 0.5-1", "error greter than 1"))) |>
+    ggplot(aes(x = tSNE1,
+               y = tSNE2, color = type))+
+    geom_point(alpha=0.5, size = 0.1) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#fee0d2", "#fcbba1",
+                                         "#fc9272", "#fb6a4a", "#ef3b2c",
+                                         "#cb181d", "#a50f15", "#99000d"))
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "(a) 2D layout from tSNE (perplexity = 23) applied for the HARPS GTO dataset.
+Is this a best representation of the original data?, (b) Model in the 2D space with tSNE (<https://youtu.be/VO_4SJEoAUA>)"
+#| label: fig-nldervisharpstSNE2
+#| fig-pos: h
+
+plot_list2_harps + trimesh_removed_tsne_harps + plot_list2_harps_error +
+  plot_layout(guides='collect', ncol = 3) &
+  theme(legend.position='bottom', plot.tag = element_text(size = 8))
+
+```
+
+<!-- param set 2-->
+
+```{r}
+#| warning: false
+#| echo: false
+
+tSNE_data_harps <- read_rds("data/harps/harps_tsne_91.rds")
+
+plot_list2_harps <- tSNE_data_harps |>
+    ggplot(aes(x = tSNE1,
+               y = tSNE2, color = clusters))+
+    geom_point(alpha=0.5) +
+    coord_cartesian(xlim = c(-8, 7), ylim = c(-10, 16)) +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#b15928", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#cab2d6", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#bc80bd", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
+  guides(color = guide_legend(title = "cluster"),
+         color = guide_legend(nrow = 3)) +
+ annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## tSNE
+## Prediction
+
+shape_val_tsne_harps <- calculate_effective_shape_value(.data = tSNE_data_harps,
+                                                      x = tSNE1, y = tSNE2)
+
+num_bins_vec <- 1:21 ## Number of bins along the x-axis
+
+vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(num_bins_vec)) {
+
+  model_object <- fit_high_d_model(training_data = training_data_1, nldr_df_with_id = tSNE_data_harps, x = "tSNE1", y = "tSNE2", num_bins_x = num_bins_vec[i], shape_val = shape_val_tsne_harps,
+                             is_bin_centroid = TRUE,
+                             is_rm_lwd_hex = FALSE,
+                             benchmark_to_rm_lwd_hex = NA,
+                             is_avg_high_d = TRUE, column_start_text = "x")
+
+  centroid_df_training <- model_object$df_bin_centroids
+  avg_df_training <- model_object$df_bin
+
+  pred_df_training <- predict_2d_embeddings(test_data = training_data_1, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "tSNE")
+
+  eval_df_training <- generate_eval_df(data = training_data_1, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, col_start = "x")
+
+  eval_df_training <- eval_df_training |>
+    mutate(num_bins_x = num_bins_vec[i])
+
+  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+
+}
+
+## Add new column with data types
+
+MSE_df_harps <- eval_data_training |>
+  dplyr::mutate(data_type = "training",
+                method = "tSNE")
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## To draw with AIC
+aic_plot_harps <- ggplot(MSE_df_harps |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "tSNE"), aes(x = number_of_bins,
+                                                                                 y = total_error,
+                                                                                 color = method
+)) +
+  geom_point() +
+  geom_line() +
+  geom_vline(xintercept = 180, linetype="dashed", 
+                color = "black", size=0.5) +
+  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  scale_colour_manual(values = c("#e41a1c","#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("AIC") +
+  xlab("total number of bins")
+## Effective number of bins along x-axis
+
+mse_plot_harps <- ggplot(MSE_df_harps |> dplyr::filter(method == "tSNE"), aes(x = number_of_bins,
+                                       y = total_mse,
+                                       color = method
+)) +
+  geom_point() +
+  geom_line() +
+   geom_vline(xintercept = 180, linetype="dashed", 
+                color = "black", size=0.5) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  scale_colour_manual(values = c("#e41a1c","#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("MSE") +
+  xlab("total number of bins")
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset.
+What is the effective number of bins in each NLDR technique to create a 2D model?
+The MSE plot have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
+Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error.
+Using the elbow method, when the total number of bins is set to 144, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
+This point indicates that adding less bins does not enough to capture the data structure."
+#| label: fig-diagnosticpltHARPS3
+##| out-width: 100%
+#| fig-pos: H
+
+aic_plot_harps + mse_plot_harps +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='none', plot.tag = element_text(size = 8))
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## Import data
+# training_data_1 <- read_rds("data/harps/harps_data.rds")
+# training_data_1 <- training_data_1 |>
+#   mutate(ID = 1:NROW(training_data_1))
+# 
+# clusters <- training_data_1$cluster
+# select_names <-  names(training_data_1)[-c(18, length(training_data_1))]
+# 
+# training_data_1 <- training_data_1 |> select(-cluster)
+# names(training_data_1) <- append(paste0("x", 1:17), "ID")
+# 
+# tSNE_data_harps <- tSNE_data_harps |>
+#   select(tSNE1, tSNE2) |>
+#   mutate(ID = 1:NROW(tSNE_data_harps))
+
+### tSNE
+
+shape_val_tsne_harps <- calculate_effective_shape_value(.data = tSNE_data_harps,
+                                                      x = tSNE1, y = tSNE2)
+
+# num_bins_tsne_harps <- find_non_empty_bins(nldr_df = tSNE_data_harps, x = "tSNE1", y = "tSNE2", shape_val = shape_val_tsne_harps, non_empty_bins = num_non_empty_bins_harps)
+num_bins_tsne_harps <- 9
+
+## To extract bin centroids
+hexbin_data_object_tsne_harps <- extract_hexbin_centroids(nldr_df = tSNE_data_harps, num_bins = num_bins_tsne_harps, shape_val = shape_val_tsne_harps, x = tSNE1, y = tSNE2)
+
+df_bin_centroids_tsne_harps <- hexbin_data_object_tsne_harps$hexdf_data
+
+## Identify bins with low-density
+# identify_rm_bins <- find_low_density_hexagons(df_bin_centroids_tsne_harps, num_bins_tsne_harps, benchmark_rm_hex = 0.01)
+# 
+# df_bin_centroids_tsne_harps <- df_bin_centroids_tsne_harps |>
+#   filter(!(hexID %in% identify_rm_bins))
+
+tSNE_data_with_hb_id_harps <- tSNE_data_harps |>
+  dplyr::mutate(hb_id = hexbin_data_object_tsne_harps$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_tsne_harps <- dplyr::bind_cols(training_data_1 |> dplyr::select(-ID), tSNE_data_with_hb_id_harps)
+
+## Averaged on high-D
+df_bin_tsne_harps <- avg_highD_data(.data = df_all_tsne_harps)
+
+## Triangulate bin centroids
+tr1_object_tsne_harps <- triangulate_bin_centroids(df_bin_centroids_tsne_harps, x, y)
+tr_from_to_df_tsne_harps <- generate_edge_info(triangular_object = tr1_object_tsne_harps)
+
+## Compute 2D distances
+distance_tsne_harps <- cal_2d_dist(.data = tr_from_to_df_tsne_harps)
+
+# distance_plot <- plot_dist(distance_tsne_harps) +
+#   #ggtitle("(b)" ) +
+#   ylab(expression(d^{(2)})) +
+#   theme(axis.text = element_text(size = 5),
+#         axis.title = element_text(size = 12))
+# 
+# distance_plot
+
+## To find the benchmark value
+benchmark_tsne_harps <- find_benchmark_value(.data = distance_tsne_harps, distance_col = "distance")
+benchmark_tsne_harps <- 2.5
+
+trimesh_removed_tsne_harps <- remove_long_edges(.data = distance_tsne_harps, benchmark_value = benchmark_tsne_harps,
+                                              triangular_object = tr1_object_tsne_harps, distance_col = distance)
+
+trimesh_removed_tsne_harps <- trimesh_removed_tsne_harps +
+  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  # theme(axis.text = element_text(size = 5),
+  #       axis.title = element_text(size = 7)) +
+  geom_point(colour = "#33a02c", size = 0.5) +
+  coord_cartesian(xlim = c(-8, 7), ylim = c(-10, 16)) +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+df_bin_train <- df_bin_tsne_harps
+names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])
+
+error_df <- df_all_tsne_harps |>
+  dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates
+
+# prediction_df_join <- prediction_df_join |>
+#   dplyr::left_join(data, by = c("ID" = "ID")) ## Map high-D data
+
+for (i in 1:(NCOL(df_bin_train) - 1)) {
+
+  error_df[ , paste0("abs_residual_", "x", i)] <- abs(error_df[ , paste0("x", i)] - error_df[ , paste0("avg_", "x", i)])
+
+}
+
+error_df <- error_df |>
+  dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "x")))))
+
+error_df <- error_df |>
+  mutate(type = if_else(total <= 0, "no error",
+                        if_else(total <= 0.2, "error 0-0.2",
+                                if_else(total <= 0.3, "error 0.2-0.3",
+                                        if_else(total <= 0.5, "error 0.3-0.5",
+                                                if_else(total <= 1, "error 0.5-1",
+                                                        "error greter than 1"))))))
+
+plot_list2_harps_error <- error_df |>
+    mutate(type = factor(type , levels = c("no error", "error 0-0.2", "error 0.2-0.3", "error 0.3-0.5", "error 0.5-1", "error greter than 1"))) |>
+    ggplot(aes(x = tSNE1,
+               y = tSNE2, color = type))+
+    geom_point(alpha=0.5, size = 0.1) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#fee0d2", "#fcbba1",
+                                         "#fc9272", "#fb6a4a", "#ef3b2c",
+                                         "#cb181d", "#a50f15", "#99000d"))
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "(a) 2D layout from tSNE (perplexity = 91) applied for the HARPS GTO dataset.
+Is this a best representation of the original data?, (b) Model in the 2D space with tSNE (<https://youtu.be/vLvRfcogft8>)"
+#| label: fig-nldervisharpstSNE3
+#| fig-pos: h
+
+plot_list2_harps + trimesh_removed_tsne_harps + plot_list2_harps_error +
+  plot_layout(guides='collect', ncol = 3) &
+  theme(legend.position='bottom', plot.tag = element_text(size = 8))
+
+```
+
 -->
+
+<!--
+**(b) Determine the shape parameter** ($s$)
+
+In this step, we determine the shape parameter ($s$) for the hexagonal bins, which significantly influences their shape and arrangement within the grid.
+The $s$ in the hexagonal binning algorithm is defined as the ratio of the height ($y$) to the width ($x$) of the plotting region as defined in @eq-equation1 [@Carr2013].
+
+$$
+s = \frac{y}{x}
+$$ {#eq-equation1}
+
+The shape parameter ($s$) of our algorithm is calculated as the ratio of the ranges of 2D embedding components, where $r_1$ and $r_2$ represent the ranges of 2D embedding components 1 and component 2, respectively (see @eq-equation2).
+
+$$
+s = \frac{r_2}{r_1}
+$$ {#eq-equation2}
+-->

### 2024-03-05 paper.qmd
Added mse plot, aic plot for pbmc need to find to remove duplicates of mse_df

@@
 #| echo: false
+library(quollr)
 library(dplyr)
@@
 library(kableExtra)
-
 library(grid)
-
-# ## For paralleization
-# library(foreach)
-# library(doParallel)
-
-# library(xtable)
-# 
-# xtable2kable <- function(x) {
-#   out <- capture.output(print(x, table.placement = NULL))[-(1:2)]
-#   out <- paste(out, collapse = "\n")
-#   structure(out, format = "latex", class = "knitr_kable")
-# }
-
-# install.packages("Seurat")
-# remotes::install_github("satijalab/seurat-data")
-#library(SeuratData) ## For the application
-
-# use_python("~/miniforge3/envs/pcamp_env/bin/python")
-# use_condaenv("pcamp_env")
-# 
-# reticulate::source_python(paste0(here::here(), "/examples_for_the_paper/function_scripts/Fit_PacMAP_code.py"))
-# reticulate::source_python(paste0(here::here(), "/examples_for_the_paper/function_scripts/Fit_TriMAP_code.py"))
-
 set.seed(20240110)
 
-source("quollr_code.R", local = TRUE)
+#source("quollr_code.R", local = TRUE)
 source("nldr_code.R", local = TRUE)
@@
 -->
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-## To plot the distribution of distance
-plot_dist <- function(distance_df){
-  distance_df$group <- "1"
-  dist_plot <- ggplot(distance_df, aes(x = group, y = distance)) +
-    geom_quasirandom()+
-    ylim(0, max(unlist(distance_df$distance))+ 0.5) + coord_flip()
-  return(dist_plot)
-}
 ```
@@
 
-# data <- data |>
-#   mutate(ID = row_number())
-# 
-# data_split <- initial_split(data)
-# training_data <- training(data_split) |>
-#   arrange(ID)
-# test_data <- testing(data_split) |>
-#   arrange(ID)
 training_data <- read_rds("data/s_curve/s_curve_training.rds")
@@
 
-#plot_list <- list()
-
 tSNE_data <- read_rds("data/s_curve/s_curve_tsne_27.rds")
-#(perplexity: ", calculate_effective_perplexity(data), ")
-plot_list1 <- plot_tSNE_2D(tSNE_data) + #ggtitle("(a)") + 
+
+plot_list1 <- plot_tSNE_2D(tSNE_data) +  
   theme_linedraw() +
@@
 #(n-neighbors: 50)
-plot_list2 <- plot_UMAP_2D(UMAP_data) + #ggtitle("(b)") + 
+plot_list2 <- plot_UMAP_2D(UMAP_data) + 
   theme_linedraw() +
@@
 
-plot_list3 <- plot_PHATE_2D(PHATE_data) + #ggtitle("(c)") + 
+plot_list3 <- plot_PHATE_2D(PHATE_data) + 
   theme_linedraw() +
@@
 #(n-inliers: 5, \n n-outliers: 4, n-random: 3)
-plot_list4 <- plot_TriMAP_2D(TriMAP_data) + #ggtitle("(d)") + 
+plot_list4 <- plot_TriMAP_2D(TriMAP_data) + 
   theme_linedraw() +
@@
 #(knn: 10, init: random, \n MN-ratio: 0.9, FP-ratio: 2)
-plot_list5 <- plot_PaCMAP_2D(PaCMAP_data) + #ggtitle("(e)") + 
+plot_list5 <- plot_PaCMAP_2D(PaCMAP_data) +  
   theme_linedraw() +
@@
 
+Then, after performing NLDR, to have regular hexagons, 2D embedding data are scaled accordingly.
+
 ### Constructing the 2D model {#sec-construct2d}
@@
 
-**(a) Determine the number of bins along the x-axis** ($b_1$)
+**(a) Determine the number of bins along the x axis** ($b_1$)
 
@@
 
+In the process of computing hexagonal bin configurations, the first step involves determining the number of bins along the x-axis.
+This begins with the initialization of parameters such as the hexagonal size ($s$) and the buffer along the x-axis.
+Subsequently, the range of the first 2D embedding component ($r_1$) is calculated to ensure comprehensive coverage of the data.
+This range is then adjusted by incorporating the buffer amount, effectively expanding the boundary to accommodate potential outliers or edge cases.
+The default buffer amount along the x-axis is $\sqrt{3} * s * 1.5$.
+Next, the horizontal spacing ($h$) between hexagons is computed based on the hexagon size.
+Finally, the number of bins along the x-axis is computed based on the adjusted range ($r_1 + buffer_{x}$) and the horizontal spacing ($h$).
+
 $$
- \text{A} = \frac{\sqrt{3}}{2}h^2
+ h = \sqrt{3} * s
 $$ {#eq-equation3}
 
-To construct regular hexagons, $A = 1$ (see @fig-binsize) use as the default setting. Then, the diameter ($h$) of the regular hexagons is calculated (see @eq-equation4).
-
 $$
-  \text{h} = \sqrt{\frac{2}{\sqrt{3}}A}
+ b_1 = \frac{r_1 + buffer_{x}}{h}
 $$ {#eq-equation4}
 
-@Carr2013 mentioned about the relationship between the diameter ($h$) of regular hexagons and the height ($y$) of the plotting region.
-According to our algorithm, the height ($y$) of the plotting region is the the range of 2D embedding component 1 ($r_1$) (see @eq-equation6).
+**(b) Determine the number of bins along the y-axis** ($b_2$)
+
+Next, the number of bins along the y-axis is computed based on the number of bins along the x-axis ($b_1$) and the shape parameter ($s$) (see @eq-equation4) [@Carr2013].
 
 $$
-h = \frac{r_1}{b_1}
+  v = 1.5 * s
+$$ {#eq-equation5}
+
+$$
+ b_2 = \frac{r_2 + buffer_{y}}{v}
 $$ {#eq-equation6}
 
-After rearranging the @eq-equation6 as @eq-equation5, $b_1$ is computed. The $b_1$ value is rounded up to the nearest whole number to have an integer value.
-
-$$
-b_1 = \frac{r_1}{h}
-$$ {#eq-equation5}
-
-**(b) Determine the shape parameter** ($s$)
-
-In this step, we determine the shape parameter ($s$) for the hexagonal bins, which significantly influences their shape and arrangement within the grid.
-The $s$ in the hexagonal binning algorithm is defined as the ratio of the height ($y$) to the width ($x$) of the plotting region as defined in @eq-equation1 [@Carr2013].
-
-$$
-s = \frac{y}{x}
-$$ {#eq-equation1}
-
-The shape parameter ($s$) of our algorithm is calculated as the ratio of the ranges of 2D embedding components, where $r_1$ and $r_2$ represent the ranges of 2D embedding components 1 and component 2, respectively (see @eq-equation2).
-
-$$
-s = \frac{r_2}{r_1}
-$$ {#eq-equation2}
-
-**(c) Determine the number of bins along the y-axis** ($b_2$)
-
-Next, the number of bins along the y-axis is computed based on the number of bins along the x-axis ($b_1$) and the shape parameter ($s$) (see @eq-equation12) [@Carr2013].
-
-$$
-b_2 = 2 * \left(\frac{(b_1 \times s)}{\sqrt(3)} + 1.5001 \right)
-$${#eq-equation12}
-
-**(d) Determine the total number of bins** ($b$)
-
-The total number of bins is determined by multiplying the number of bins along the x-axis ($b_1$) with the number of bins along the y-axis ($b_2$) (see @eq-equation13).
+**(c) Determine the total number of bins** ($b$)
+
+The total number of bins is determined by multiplying the number of bins along the x-axis ($b_1$) with the number of bins along the y-axis ($b_2$) (see @eq-equation5).
 
@@
 b = b_1 \times b_2
-$${#eq-equation13}
-
-**Step 2: Obtain bin centroids**
+$${#eq-equation5}
+
+**Step 2: Obtain bin centroids or bin means**
 
@@
 #| echo: false
-
-# UMAP_fit <- umap(training_data |> dplyr::select(-ID), n_neighbors = 50, n_components =  2)
-# 
-# UMAP_data <- UMAP_fit$layout |>
-#   as.data.frame()
-# names(UMAP_data)[1:(ncol(UMAP_data))] <- paste0(rep("UMAP",(ncol(UMAP_data))), 1:(ncol(UMAP_data)))
-# 
-# UMAP_data <- UMAP_data |>
-#   mutate(ID = training_data$ID)
 
@@
 
-plot_list1_gau <- plot_tSNE_2D(tSNE_data_gau) + #ggtitle("(a)") + 
+plot_list1_gau <- plot_tSNE_2D(tSNE_data_gau) + 
   geom_point(size = 0.00001, colour = "#e41a1c") +
@@
 
-plot_list2_gau <- plot_UMAP_2D(UMAP_data_gau) + #ggtitle("(b)") +
+plot_list2_gau <- plot_UMAP_2D(UMAP_data_gau) + 
   geom_point(size = 0.00001, colour = "#377eb8") +
@@
 
-plot_list3_gau <- plot_PHATE_2D(PHATE_data_gau) + #ggtitle("(c)") +
+plot_list3_gau <- plot_PHATE_2D(PHATE_data_gau) + 
   geom_point(size = 0.00001, colour = "#4daf4a") +
@@
 
-plot_list4_gau <- plot_TriMAP_2D(TriMAP_data_gau) + #ggtitle("(d)") +
+plot_list4_gau <- plot_TriMAP_2D(TriMAP_data_gau) + 
   geom_point(size = 0.00001, colour = "#984ea3") +
@@
 
-plot_list5_gau <- plot_PaCMAP_2D(PaCMAP_data_gau) + #ggtitle("(e)") +
+plot_list5_gau <- plot_PaCMAP_2D(PaCMAP_data_gau) + 
   geom_point(size = 0.00001, colour = "#ff7f00") +
@@
 ## Import data
-#df_2 <- read_rds("data/pbmc/pbmc.rds")
 training_data_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_pca_50.rds")
@@
   mutate(ID = 1:NROW(training_data_pbmc))
-#test_1 <- read_rds("data/pbmc/pbmc_test.rds")
-
-#tSNE_pbmc <- read_rds("data/pbmc/pbmc_tsne_30.rds")
-UMAP_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_umap.rds")
-# PHATE_pbmc <- read_rds("data/pbmc/pbmc_phate.rds")
-# TriMAP_pbmc <- read_rds("data/pbmc/pbmc_trimap.rds")
-# PaCMAP_pbmc <- read_rds("data/pbmc/pbmc_pacmap.rds")
-
-#load("data/pbmc/pbmc_3k_festem/pbmc3k_label.RData")
-
-## Visualise embeddings
-
-# plot_list1_pbmc <- plot_tSNE_2D(tSNE_pbmc) + #ggtitle("(a)") + 
-#   theme_linedraw() +
-#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-#         axis.title.x = element_blank(), axis.title.y = element_blank(),
-#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-#   annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-class_avg <- UMAP_pbmc |>
+
+umap_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_umap.rds")
+umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc, 
+                                    x = "UMAP1", y = "UMAP2"))) |>
+  dplyr::mutate(cell_label = umap_pbmc$cell_label) |>
+  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
+                  "UMAP2" = "scaled_UMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(umap_pbmc))
+
+class_avg <- umap_pbmc_scaled |>
   group_by(cell_label) |>
@@
 
-plot_list2_pbmc <- UMAP_pbmc |>
+plot_list2_pbmc <- umap_pbmc_scaled |>
     ggplot(aes(x = UMAP1,
@@
   geom_text(aes(x=UMAP1,y = UMAP2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) #+
-  # guides(color = guide_legend(title = "Cell type"),
-  #        color = guide_legend(nrow = 3)) #+
-  # annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-# plot_list3_pbmc <- plot_PHATE_2D(PHATE_pbmc) + #ggtitle("(c)") + 
-#   theme_linedraw() +
-#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-#         axis.title.x = element_blank(), axis.title.y = element_blank(),
-#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-#   annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-# plot_list4_pbmc <- plot_TriMAP_2D(TriMAP_pbmc) + #ggtitle("(d)") + 
-#   theme_linedraw() +
-#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-#         axis.title.x = element_blank(), axis.title.y = element_blank(),
-#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-#   annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-# 
-# plot_list5_pbmc <- plot_PaCMAP_2D(PaCMAP_pbmc) + #ggtitle("(e)") + 
-#   theme_linedraw() +
-#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-#         axis.title.x = element_blank(), axis.title.y = element_blank(),
-#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-#   annotate(geom = 'text', label = 'e', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-```
-
-```{r}
-#| warning: false
-#| echo: false
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
+```
+
+```{r}
+#| warning: false
+#| echo: false
+#| message: false
 
@@
 
-shape_value_pbmc <- calculate_effective_shape_value(.data = UMAP_pbmc,
-                                                      x = UMAP1, y = UMAP2)
-
-num_bins_vec <- 1:35 ## Number of bins along the x-axis
-
-vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names
-
-eval_data_training <- dplyr::bind_rows(vec)[0, ]
-eval_data_training <- eval_data_training |>
+#tot_num_bin_vec <- 2:51
+hex_size_vec <- seq(0.02, 2, by = 0.01)
+
+vec <- stats::setNames(rep("", 6), c("num_bins", "aic", "mse", "num_bins_x", "num_bins_y", "hex_size"))  ## Define column names
+
+mse_df_pbmc <- dplyr::bind_rows(vec)[0, ]
+mse_df_pbmc <- mse_df_pbmc |>
   dplyr::mutate_if(is.character, as.numeric)
 
-for (i in 1:length(num_bins_vec)) {
-
-  model_object <- fit_high_d_model(training_data = training_data_pbmc, nldr_df_with_id = UMAP_pbmc, x = "UMAP1", y = "UMAP2", num_bins_x = num_bins_vec[i], shape_val = shape_value_pbmc,
-                             is_bin_centroid = TRUE,
+for (i in 1:length(hex_size_vec)) {
+
+  num_bin_list <- calc_bins(data = umap_pbmc_scaled, 
+            x = "UMAP1", y = "UMAP2", 
+            hex_size = hex_size_vec[i], buffer_x = NA, buffer_y = NA)
+
+  num_bins_x <- num_bin_list$num_x
+  num_bins_y <- num_bin_list$num_y
+
+  model_object <- fit_highd_model( training_data = training_data_pbmc, 
+                                   nldr_df_with_id = umap_pbmc_scaled |> dplyr::select(-cell_label), 
+                                   x = "UMAP1", y = "UMAP2", 
+                                   num_bins_x = num_bins_x, 
+                                   num_bins_y = num_bins_y, 
+                                   x_start = NA, y_start = NA, 
+                                   buffer_x = NA, buffer_y = NA, 
+                                   hex_size = hex_size_vec[i],
                                    is_rm_lwd_hex = FALSE, 
                                    benchmark_to_rm_lwd_hex = NA, 
-                             is_avg_high_d = TRUE, column_start_text = "PC_")
+                                   col_start_2d = "UMAP", 
+                                   col_start_highd = "PC_")
 
@@
 
-  pred_df_training <- predict_2d_embeddings(test_data = training_data_pbmc, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "UMAP")
-
-  eval_df_training <- generate_eval_df(data = training_data_pbmc, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, col_start = "PC_")
-
-  eval_df_training <- eval_df_training |>
-    mutate(num_bins_x = num_bins_vec[i])
-
-  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+  pred_emb_list <- predict_emb(test_data = training_data_pbmc, 
+                                  df_bin_centroids = centroid_df_training, 
+                                  df_bin = avg_df_training, type_NLDR = "UMAP")
+
+  pred_df_training <- as.data.frame(do.call(cbind, pred_emb_list))
+
+  eval_list <- gen_summary(test_data = training_data_pbmc, 
+                                  prediction_df = pred_df_training, 
+                                  df_bin = avg_df_training, col_start = "PC_")
+
+  mse_df_pbmc <- mse_df_pbmc |>
+    tibble::add_row(num_bins = num_bins_x * num_bins_y,
+                    aic = eval_list$aic,
+                    mse = eval_list$mse,
+                    num_bins_x = num_bins_x,
+                    num_bins_y = num_bins_y,
+                    hex_size = hex_size_vec[i])
 
@@
 
-## Add new column with data types
-
-MSE_df_pbmc <- eval_data_training |>
-  dplyr::mutate(data_type = "training",
-                method = "UMAP")
+## If same total number of bins ocurred only select ones with minimum error
+duplicate_df_list <- mse_df_pbmc[duplicated(mse_df_pbmc$num_bins), ] |>
+  dplyr::arrange(num_bins) |>
+  dplyr::group_split(num_bins)
+
+duplicate_df <- data.frame(matrix(nrow = 0, ncol = 0))
+
+for (i in 1:length(duplicate_df_list)) {
+
+  dd <- duplicate_df_list[[i]] |>
+    dplyr::filter(mse == min(duplicate_df_list[[i]]$mse))
+
+  duplicate_df <- dplyr::bind_rows(duplicate_df, dd)
+
+}
+
+mse_df_pbmc <- dplyr::bind_rows(mse_df_pbmc[!duplicated(mse_df_pbmc$num_bins), ], duplicate_df)
 ```
@@
 ## To draw with AIC
-aic_plot_pbmc <- ggplot(MSE_df_pbmc |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
-                                                                                 y = total_error,
-                                                                                 color = method
+aic_plot_pbmc <- ggplot(mse_df_pbmc, aes(x = num_bins, y = aic
 )) +
@@
   geom_line() +
-  geom_vline(xintercept = 256, linetype="dashed", 
-                color = "red", size=0.5) +
+  # geom_vline(xintercept = 256, linetype="dashed", 
+  #               color = "red", size=0.5) +
   #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
@@
 
-mse_plot_pbmc <- ggplot(MSE_df_pbmc |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
-                                       y = total_mse,
-                                       color = method
+mse_plot_pbmc <- ggplot(mse_df_pbmc, aes(x = num_bins,
+                                       y = mse
 )) +
@@
   geom_line() +
-   geom_vline(xintercept = 256, linetype="dashed", 
-                color = "red", size=0.5) +
+   # geom_vline(xintercept = 256, linetype="dashed", 
+   #              color = "red", size=0.5) +
   theme_light() +
@@
 
-shape_value_pbmc <- calculate_effective_shape_value(.data = UMAP_pbmc,
-                                                      x = UMAP1, y = UMAP2) ## 0.8772751
-
 ## To extract bin centroids
@@
                                          "#fc9272", "#fb6a4a", "#ef3b2c",
-                                         "#cb181d", "#a50f15", "#99000d")) #+
-  #geom_text(aes(x=UMAP1,y = UMAP2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) +
-  #annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
+                                         "#cb181d", "#a50f15", "#99000d")) 
 ```
@@
   theme(legend.position='bottom')
-
-#  plot_list1_pbmc + plot_list3_pbmc + plot_list4_pbmc + plot_list5_pbmc +
-#   plot_layout(ncol=5) #| out-width: 100%
-
-# 2D layouts from different NLDR techniques applied for the training PBMC dataset: (a) tSNE (perplexity = 30), (b) UMAP (n_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2).
-Is there a best representation of the original data or are they all providing  equivalent information?
 ```
@@
   mutate(ID = 1:NROW(training_data_pbmc))
-#test_1 <- read_rds("data/pbmc/pbmc_test.rds")
-
-#tSNE_pbmc <- read_rds("data/pbmc/pbmc_tsne_30.rds")
+
 UMAP_pbmc <- read_rds("data/pbmc/pbmc_umap_15_min_dist_0.99_metric_cosine.rds")
-# PHATE_pbmc <- read_rds("data/pbmc/pbmc_phate.rds")
-# TriMAP_pbmc <- read_rds("data/pbmc/pbmc_trimap.rds")
-# PaCMAP_pbmc <- read_rds("data/pbmc/pbmc_pacmap.rds")
-
-#load("data/pbmc/pbmc_3k_festem/pbmc3k_label.RData")
-
-## Visualise embeddings
-
-# plot_list1_pbmc <- plot_tSNE_2D(tSNE_pbmc) + #ggtitle("(a)") + 
-#   theme_linedraw() +
-#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-#         axis.title.x = element_blank(), axis.title.y = element_blank(),
-#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-#   annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
 
@@
   geom_text(aes(x=UMAP1,y = UMAP2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) #+
-  # guides(color = guide_legend(title = "Cell type"),
-  #        color = guide_legend(nrow = 3)) #+
-  # annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-# plot_list3_pbmc <- plot_PHATE_2D(PHATE_pbmc) + #ggtitle("(c)") + 
-#   theme_linedraw() +
-#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-#         axis.title.x = element_blank(), axis.title.y = element_blank(),
-#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-#   annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-# plot_list4_pbmc <- plot_TriMAP_2D(TriMAP_pbmc) + #ggtitle("(d)") + 
-#   theme_linedraw() +
-#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-#         axis.title.x = element_blank(), axis.title.y = element_blank(),
-#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-#   annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-# 
-# plot_list5_pbmc <- plot_PaCMAP_2D(PaCMAP_pbmc) + #ggtitle("(e)") + 
-#   theme_linedraw() +
-#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-#         axis.title.x = element_blank(), axis.title.y = element_blank(),
-#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-#   annotate(geom = 'text', label = 'e', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
 ```
@@
   mutate(ID = 1:NROW(training_data_pbmc))
-#test_1 <- read_rds("data/pbmc/pbmc_test.rds")
-
-#tSNE_pbmc <- read_rds("data/pbmc/pbmc_tsne_30.rds")
+
 UMAP_pbmc <- read_rds("data/pbmc/pbmc_umap_5_min_dist_0.99_metric_cosine.rds")
-# PHATE_pbmc <- read_rds("data/pbmc/pbmc_phate.rds")
-# TriMAP_pbmc <- read_rds("data/pbmc/pbmc_trimap.rds")
-# PaCMAP_pbmc <- read_rds("data/pbmc/pbmc_pacmap.rds")
-
-#load("data/pbmc/pbmc_3k_festem/pbmc3k_label.RData")
-
-## Visualise embeddings
-
-# plot_list1_pbmc <- plot_tSNE_2D(tSNE_pbmc) + #ggtitle("(a)") + 
-#   theme_linedraw() +
-#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-#         axis.title.x = element_blank(), axis.title.y = element_blank(),
-#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-#   annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
 
@@
 
-#  plot_list1_pbmc + plot_list3_pbmc + plot_list4_pbmc + plot_list5_pbmc +
-#   plot_layout(ncol=5) #| out-width: 100%
-
-# 2D layouts from different NLDR techniques applied for the training PBMC dataset: (a) tSNE (perplexity = 30), (b) UMAP (n_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2).
-Is there a best representation of the original data or are they all providing  equivalent information?
-```
+```
+
 <!--Generate true label for PBMC3k-->
@@
 
-### HARPS GTO sample dataset
-
-@Anders2018
-
-- In 2D, there are two well-separated clusters, but the clusters are really close
-
-- One cluster have a sparse data space within the cluster
-
-- There is a line in right corner of the structure that is not connected any other trimesh points. This line create a long edge in high-D space.
-
-- In high-D view, the clusters having some non-linearity patter, which is also not captured in tSNE.
-Therefore, this is not a good choice.
-Also, the two clusters are really near to each other and can't see any separation.
-Also continuity of the structure.
-Long edges are presence also.
-
-- The parameter is, in a sense, a guess about the number of close neighbors each point has.
-
-- Num bins along the x-axis = 43, shape parameter = 0.4466309
-
-```{r}
-#| warning: false
-#| echo: false
-
-error_df <- read_csv("data/harps/error_harps_tsne.csv", col_names = FALSE)
-
-error_df <- error_df |>
-  mutate(X2 = round(X2/100, 2),
-         X1 = as.integer(X1)) |>
-  arrange(X2)
-
-colnames(error_df) <- c("perplexity", "error (x 100)")
-
-kable(error_df, caption = "errors for different parameter combinations", label = "tbl-errordiffcombharps") |>
-  #xtable2kable() |>
-  kable_styling(latex_options = c("striped", "HOLD_position"), position = "center", full_width = FALSE) |>
-  row_spec(0, bold = TRUE) |>
-  row_spec(3, bold=T, color = "red")
-```
-
-<!-- Author suggestion-->
-```{r}
-#| warning: false
-#| echo: false
-
-tSNE_data_harps <- read_rds("data/harps/harps_tsne_40.rds")
-
-plot_list2_harps <- tSNE_data_harps |>
-    ggplot(aes(x = tSNE1,
-               y = tSNE2, color = cluster))+
-    geom_point(alpha=0.5) +
-    #coord_cartesian(xlim = c(-25, 25), ylim = c(-12, 12)) +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#b15928", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#cab2d6", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#bc80bd", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
-  guides(color = guide_legend(title = "cluster"),
-         color = guide_legend(nrow = 3)) +
- annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-## Import data
-training_data_1 <- read_rds("data/harps/harps_data.rds")
-training_data_1 <- training_data_1 |>
-  mutate(ID = 1:NROW(training_data_1))
-
-clusters <- training_data_1$cluster
-select_names <-  names(training_data_1)[-c(18, length(training_data_1))]
-
-training_data_1 <- training_data_1 |> select(-cluster)
-names(training_data_1) <- append(paste0("x", 1:17), "ID")
-
-tSNE_data_harps <- tSNE_data_harps |>
-  select(tSNE1, tSNE2) |>
-  mutate(ID = 1:NROW(tSNE_data_harps))
-
-## tSNE
-## Prediction
-
-shape_val_tsne_harps <- calculate_effective_shape_value(.data = tSNE_data_harps,
-                                                      x = tSNE1, y = tSNE2)
-
-num_bins_vec <- 1:21 ## Number of bins along the x-axis
-
-vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names
-
-eval_data_training <- dplyr::bind_rows(vec)[0, ]
-eval_data_training <- eval_data_training |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(num_bins_vec)) {
-
-  model_object <- fit_high_d_model(training_data = training_data_1, nldr_df_with_id = tSNE_data_harps, x = "tSNE1", y = "tSNE2", num_bins_x = num_bins_vec[i], shape_val = shape_val_tsne_harps,
-                             is_bin_centroid = TRUE,
-                             is_rm_lwd_hex = FALSE,
-                             benchmark_to_rm_lwd_hex = NA,
-                             is_avg_high_d = TRUE, column_start_text = "x")
-
-  centroid_df_training <- model_object$df_bin_centroids
-  avg_df_training <- model_object$df_bin
-
-  pred_df_training <- predict_2d_embeddings(test_data = training_data_1, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "tSNE")
-
-  eval_df_training <- generate_eval_df(data = training_data_1, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, col_start = "x")
-
-  eval_df_training <- eval_df_training |>
-    mutate(num_bins_x = num_bins_vec[i])
-
-  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
-
-}
-
-## Add new column with data types
-
-MSE_df_harps <- eval_data_training |>
-  dplyr::mutate(data_type = "training",
-                method = "tSNE")
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-## To draw with AIC
-aic_plot_harps <- ggplot(MSE_df_harps |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "tSNE"), aes(x = number_of_bins,
-                                                                                 y = total_error,
-                                                                                 color = method
-)) +
-  geom_point() +
-  geom_line() +
-  geom_vline(xintercept = 120, linetype="dashed", 
-                color = "black", size=0.5) +
-  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  scale_colour_manual(values = c("#e41a1c","#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("AIC") +
-  xlab("total number of bins")
-## Effective number of bins along x-axis
-
-mse_plot_harps <- ggplot(MSE_df_harps |> dplyr::filter(method == "tSNE"), aes(x = number_of_bins,
-                                       y = total_mse,
-                                       color = method
-)) +
-  geom_point() +
-  geom_line() +
-   geom_vline(xintercept = 120, linetype="dashed", 
-                color = "black", size=0.5) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  scale_colour_manual(values = c("#e41a1c","#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("MSE") +
-  xlab("total number of bins")
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset.
-What is the effective number of bins in each NLDR technique to create a 2D model?
-The MSE plot have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
-Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error.
-Using the elbow method, when the total number of bins is set to 144, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
-This point indicates that adding less bins does not enough to capture the data structure."
-#| label: fig-diagnosticpltHARPS
-##| out-width: 100%
-#| fig-pos: H
-
-aic_plot_harps + mse_plot_harps +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', ncol = 2) &
-  theme(legend.position='none', plot.tag = element_text(size = 8))
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-### tSNE
-
-num_bins_tsne_harps <- calculate_effective_x_bins(.data = tSNE_data_harps, x = "tSNE1",
-                                                cell_area = 1) #43
-
-num_bins_tsne_harps <- 14
-
-shape_val_tsne_harps <- calculate_effective_shape_value(.data = tSNE_data_harps,
-                                                      x = "tSNE1", y = "tSNE2")
-## To extract bin centroids
-hexbin_data_object_tsne_harps <- extract_hexbin_centroids(nldr_df = tSNE_data_harps, num_bins = num_bins_tsne_harps, shape_val = shape_val_tsne_harps, x = "tSNE1", y = "tSNE2")
-
-df_bin_centroids_tsne_harps <- hexbin_data_object_tsne_harps$hexdf_data
-
-#num_non_empty_bins_harps <- df_bin_centroids_tsne_harps$hexID |> length()
-
-## Identify bins with low-density
-# identify_rm_bins <- find_low_density_hexagons(df_bin_centroids_tsne_harps, num_bins_tsne_harps, benchmark_rm_hex = 0.01)
-# 
-# df_bin_centroids_tsne_harps <- df_bin_centroids_tsne_harps |>
-#   filter(!(hexID %in% identify_rm_bins))
-
-tSNE_data_with_hb_id_harps <- tSNE_data_harps |>
-  dplyr::mutate(hb_id = hexbin_data_object_tsne_harps$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_tsne_harps <- dplyr::bind_cols(training_data_1 |> dplyr::select(-ID), tSNE_data_with_hb_id_harps)
-
-## Averaged on high-D
-df_bin_tsne_harps <- avg_highD_data(.data = df_all_tsne_harps)
-
-## Triangulate bin centroids
-tr1_object_tsne_harps <- triangulate_bin_centroids(df_bin_centroids_tsne_harps, x, y)
-tr_from_to_df_tsne_harps <- generate_edge_info(triangular_object = tr1_object_tsne_harps)
-
-## Compute 2D distances
-distance_tsne_harps <- cal_2d_dist(.data = tr_from_to_df_tsne_harps)
-
-## To find the benchmark value
-benchmark_tsne_harps <- find_benchmark_value(.data = distance_tsne_harps, distance_col = "distance")
-benchmark_tsne_harps <- 5
-
-# distance_plot <- plot_dist(distance_tsne_harps) +
-#   #ggtitle("(b)" ) +
-#   ylab(expression(d^{(2)})) +
-#   theme(axis.text = element_text(size = 5),
-#         axis.title = element_text(size = 12))
-# 
-# distance_plot
-
-trimesh_removed_tsne_harps <- remove_long_edges(.data = distance_tsne_harps, benchmark_value = benchmark_tsne_harps,
-                                              triangular_object = tr1_object_tsne_harps, distance_col = distance)
-
-trimesh_removed_tsne_harps <- trimesh_removed_tsne_harps +
-  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  # theme(axis.text = element_text(size = 5),
-  #       axis.title = element_text(size = 7)) +
-  geom_point(colour = "#33a02c", size = 0.5) +
-  #coord_cartesian(xlim = c(-25, 25), ylim = c(-12, 12)) +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-df_bin_train <- df_bin_tsne_harps
-names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])
-
-error_df <- df_all_tsne_harps |>
-  dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates
-
-# prediction_df_join <- prediction_df_join |>
-#   dplyr::left_join(data, by = c("ID" = "ID")) ## Map high-D data
-
-for (i in 1:(NCOL(df_bin_train) - 1)) {
-
-  error_df[ , paste0("abs_residual_", "x", i)] <- abs(error_df[ , paste0("x", i)] - error_df[ , paste0("avg_", "x", i)])
-
-}
-
-error_df <- error_df |>
-  dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "x")))))
-
-error_df <- error_df |>
-  mutate(type = if_else(total <= 0, "no error",
-                        if_else(total <= 0.2, "error 0-0.2",
-                                if_else(total <= 0.3, "error 0.2-0.3",
-                                        if_else(total <= 0.5, "error 0.3-0.5",
-                                                if_else(total <= 1, "error 0.5-1",
-                                                        "error greter than 1"))))))
-
-plot_list2_harps_error <- error_df |>
-    mutate(type = factor(type , levels = c("no error", "error 0-0.2", "error 0.2-0.3", "error 0.3-0.5", "error 0.5-1", "error greter than 1"))) |>
-    ggplot(aes(x = tSNE1,
-               y = tSNE2, color = type))+
-    geom_point(alpha=0.5, size = 0.1) +
-    coord_equal() +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#fee0d2", "#fcbba1",
-                                         "#fc9272", "#fb6a4a", "#ef3b2c",
-                                         "#cb181d", "#a50f15", "#99000d"))
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "(a) 2D layout from tSNE (perplexity = 40) applied for the HARPS GTO dataset.
-Is this a best representation of the original data?, (b) Model in the 2D space with tSNE (<https://youtu.be/jxo_rPyGv9k>)"
-#| label: fig-nldervisharpstSNE
-#| fig-pos: h
-
-plot_list2_harps + trimesh_removed_tsne_harps + plot_list2_harps_error +
-  plot_layout(guides='collect', ncol = 3) &
-  theme(legend.position='bottom', plot.tag = element_text(size = 8))
-
-```
-
-<!-- param set 1-->
-
-- 5 clusters shows in 2D
-
-- 3 clusters which are really close shows in high_d data with the model
-
-```{r}
-#| warning: false
-#| echo: false
-
-tSNE_data_harps <- read_rds("data/harps/harps_tsne_23.rds")
-
-plot_list2_harps <- tSNE_data_harps |>
-    ggplot(aes(x = tSNE1,
-               y = tSNE2, color = clusters))+
-    geom_point(alpha=0.5) +
-    coord_cartesian(xlim = c(-25, 20), ylim = c(-35, 25)) +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#b15928", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#cab2d6", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#bc80bd", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
-  guides(color = guide_legend(title = "cluster"),
-         color = guide_legend(nrow = 3)) +
- annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-## tSNE
-## Prediction
-
-shape_val_tsne_harps <- calculate_effective_shape_value(.data = tSNE_data_harps,
-                                                      x = tSNE1, y = tSNE2)
-
-num_bins_vec <- 1:21 ## Number of bins along the x-axis
-
-vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names
-
-eval_data_training <- dplyr::bind_rows(vec)[0, ]
-eval_data_training <- eval_data_training |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(num_bins_vec)) {
-
-  model_object <- fit_high_d_model(training_data = training_data_1, nldr_df_with_id = tSNE_data_harps, x = "tSNE1", y = "tSNE2", num_bins_x = num_bins_vec[i], shape_val = shape_val_tsne_harps,
-                             is_bin_centroid = TRUE,
-                             is_rm_lwd_hex = FALSE,
-                             benchmark_to_rm_lwd_hex = NA,
-                             is_avg_high_d = TRUE, column_start_text = "x")
-
-  centroid_df_training <- model_object$df_bin_centroids
-  avg_df_training <- model_object$df_bin
-
-  pred_df_training <- predict_2d_embeddings(test_data = training_data_1, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "tSNE")
-
-  eval_df_training <- generate_eval_df(data = training_data_1, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, col_start = "x")
-
-  eval_df_training <- eval_df_training |>
-    mutate(num_bins_x = num_bins_vec[i])
-
-  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
-
-}
-
-## Add new column with data types
-
-MSE_df_harps <- eval_data_training |>
-  dplyr::mutate(data_type = "training",
-                method = "tSNE")
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-## To draw with AIC
-aic_plot_harps <- ggplot(MSE_df_harps |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "tSNE"), aes(x = number_of_bins,
-                                                                                 y = total_error,
-                                                                                 color = method
-)) +
-  geom_point() +
-  geom_line() +
-  geom_vline(xintercept = 160, linetype="dashed", 
-                color = "black", size=0.5) +
-  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  scale_colour_manual(values = c("#e41a1c","#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("AIC") +
-  xlab("total number of bins")
-## Effective number of bins along x-axis
-
-mse_plot_harps <- ggplot(MSE_df_harps |> dplyr::filter(method == "tSNE"), aes(x = number_of_bins,
-                                       y = total_mse,
-                                       color = method
-)) +
-  geom_point() +
-  geom_line() +
-   geom_vline(xintercept = 160, linetype="dashed", 
-                color = "black", size=0.5) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  scale_colour_manual(values = c("#e41a1c","#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("MSE") +
-  xlab("total number of bins")
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset.
-What is the effective number of bins in each NLDR technique to create a 2D model?
-The MSE plot have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
-Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error.
-Using the elbow method, when the total number of bins is set to 144, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
-This point indicates that adding less bins does not enough to capture the data structure."
-#| label: fig-diagnosticpltHARPS2
-##| out-width: 100%
-#| fig-pos: H
-
-aic_plot_harps + mse_plot_harps +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', ncol = 2) &
-  theme(legend.position='none', plot.tag = element_text(size = 8))
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-## Import data
-# training_data_1 <- read_rds("data/harps/harps_data.rds")
-# training_data_1 <- training_data_1 |>
-#   mutate(ID = 1:NROW(training_data_1))
-# 
-# clusters <- training_data_1$cluster
-# select_names <-  names(training_data_1)[-c(18, length(training_data_1))]
-# 
-# training_data_1 <- training_data_1 |> select(-cluster)
-# names(training_data_1) <- append(paste0("x", 1:17), "ID")
-# 
-# tSNE_data_harps <- tSNE_data_harps |>
-#   select(tSNE1, tSNE2) |>
-#   mutate(ID = 1:NROW(tSNE_data_harps))
-
-### tSNE
-
-shape_val_tsne_harps <- calculate_effective_shape_value(.data = tSNE_data_harps,
-                                                      x = tSNE1, y = tSNE2)
-
-# num_bins_tsne_harps <- find_non_empty_bins(nldr_df = tSNE_data_harps, x = "tSNE1", y = "tSNE2", shape_val = shape_val_tsne_harps, non_empty_bins = num_non_empty_bins_harps)
-
-num_bins_tsne_harps <- 9
-
-## To extract bin centroids
-hexbin_data_object_tsne_harps <- extract_hexbin_centroids(nldr_df = tSNE_data_harps, num_bins = num_bins_tsne_harps, shape_val = shape_val_tsne_harps, x = tSNE1, y = tSNE2)
-
-df_bin_centroids_tsne_harps <- hexbin_data_object_tsne_harps$hexdf_data
-
-## Identify bins with low-density
-# identify_rm_bins <- find_low_density_hexagons(df_bin_centroids_tsne_harps, num_bins_tsne_harps, benchmark_rm_hex = 0.01)
-# 
-# df_bin_centroids_tsne_harps <- df_bin_centroids_tsne_harps |>
-#   filter(!(hexID %in% identify_rm_bins))
-
-tSNE_data_with_hb_id_harps <- tSNE_data_harps |>
-  dplyr::mutate(hb_id = hexbin_data_object_tsne_harps$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_tsne_harps <- dplyr::bind_cols(training_data_1 |> dplyr::select(-ID), tSNE_data_with_hb_id_harps)
-
-## Averaged on high-D
-df_bin_tsne_harps <- avg_highD_data(.data = df_all_tsne_harps)
-
-## Triangulate bin centroids
-tr1_object_tsne_harps <- triangulate_bin_centroids(df_bin_centroids_tsne_harps, x, y)
-tr_from_to_df_tsne_harps <- generate_edge_info(triangular_object = tr1_object_tsne_harps)
-
-## Compute 2D distances
-distance_tsne_harps <- cal_2d_dist(.data = tr_from_to_df_tsne_harps)
-
-# distance_plot <- plot_dist(distance_tsne_harps) +
-#   #ggtitle("(b)" ) +
-#   ylab(expression(d^{(2)})) +
-#   theme(axis.text = element_text(size = 5),
-#         axis.title = element_text(size = 12))
-# 
-# distance_plot
-
-## To find the benchmark value
-benchmark_tsne_harps <- find_benchmark_value(.data = distance_tsne_harps, distance_col = "distance")
-benchmark_tsne_harps <- 6
-
-trimesh_removed_tsne_harps <- remove_long_edges(.data = distance_tsne_harps, benchmark_value = benchmark_tsne_harps,
-                                              triangular_object = tr1_object_tsne_harps, distance_col = distance)
-
-trimesh_removed_tsne_harps <- trimesh_removed_tsne_harps +
-  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  # theme(axis.text = element_text(size = 5),
-  #       axis.title = element_text(size = 7)) +
-  geom_point(colour = "#33a02c", size = 0.5) +
-  coord_cartesian(xlim = c(-25, 20), ylim = c(-35, 25)) +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-df_bin_train <- df_bin_tsne_harps
-names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])
-
-error_df <- df_all_tsne_harps |>
-  dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates
-
-# prediction_df_join <- prediction_df_join |>
-#   dplyr::left_join(data, by = c("ID" = "ID")) ## Map high-D data
-
-for (i in 1:(NCOL(df_bin_train) - 1)) {
-
-  error_df[ , paste0("abs_residual_", "x", i)] <- abs(error_df[ , paste0("x", i)] - error_df[ , paste0("avg_", "x", i)])
-
-}
-
-error_df <- error_df |>
-  dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "x")))))
-
-error_df <- error_df |>
-  mutate(type = if_else(total <= 0, "no error",
-                        if_else(total <= 0.2, "error 0-0.2",
-                                if_else(total <= 0.3, "error 0.2-0.3",
-                                        if_else(total <= 0.5, "error 0.3-0.5",
-                                                if_else(total <= 1, "error 0.5-1",
-                                                        "error greter than 1"))))))
-
-plot_list2_harps_error <- error_df |>
-    mutate(type = factor(type , levels = c("no error", "error 0-0.2", "error 0.2-0.3", "error 0.3-0.5", "error 0.5-1", "error greter than 1"))) |>
-    ggplot(aes(x = tSNE1,
-               y = tSNE2, color = type))+
-    geom_point(alpha=0.5, size = 0.1) +
-    coord_equal() +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#fee0d2", "#fcbba1",
-                                         "#fc9272", "#fb6a4a", "#ef3b2c",
-                                         "#cb181d", "#a50f15", "#99000d"))
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "(a) 2D layout from tSNE (perplexity = 23) applied for the HARPS GTO dataset.
-Is this a best representation of the original data?, (b) Model in the 2D space with tSNE (<https://youtu.be/VO_4SJEoAUA>)"
-#| label: fig-nldervisharpstSNE2
-#| fig-pos: h
-
-plot_list2_harps + trimesh_removed_tsne_harps + plot_list2_harps_error +
-  plot_layout(guides='collect', ncol = 3) &
-  theme(legend.position='bottom', plot.tag = element_text(size = 8))
-
-```
-
-<!-- param set 2-->
-
-```{r}
-#| warning: false
-#| echo: false
-
-tSNE_data_harps <- read_rds("data/harps/harps_tsne_91.rds")
-
-plot_list2_harps <- tSNE_data_harps |>
-    ggplot(aes(x = tSNE1,
-               y = tSNE2, color = clusters))+
-    geom_point(alpha=0.5) +
-    coord_cartesian(xlim = c(-8, 7), ylim = c(-10, 16)) +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#b15928", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#cab2d6", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#bc80bd", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
-  guides(color = guide_legend(title = "cluster"),
-         color = guide_legend(nrow = 3)) +
- annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-## tSNE
-## Prediction
-
-shape_val_tsne_harps <- calculate_effective_shape_value(.data = tSNE_data_harps,
-                                                      x = tSNE1, y = tSNE2)
-
-num_bins_vec <- 1:21 ## Number of bins along the x-axis
-
-vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names
-
-eval_data_training <- dplyr::bind_rows(vec)[0, ]
-eval_data_training <- eval_data_training |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(num_bins_vec)) {
-
-  model_object <- fit_high_d_model(training_data = training_data_1, nldr_df_with_id = tSNE_data_harps, x = "tSNE1", y = "tSNE2", num_bins_x = num_bins_vec[i], shape_val = shape_val_tsne_harps,
-                             is_bin_centroid = TRUE,
-                             is_rm_lwd_hex = FALSE,
-                             benchmark_to_rm_lwd_hex = NA,
-                             is_avg_high_d = TRUE, column_start_text = "x")
-
-  centroid_df_training <- model_object$df_bin_centroids
-  avg_df_training <- model_object$df_bin
-
-  pred_df_training <- predict_2d_embeddings(test_data = training_data_1, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "tSNE")
-
-  eval_df_training <- generate_eval_df(data = training_data_1, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, col_start = "x")
-
-  eval_df_training <- eval_df_training |>
-    mutate(num_bins_x = num_bins_vec[i])
-
-  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
-
-}
-
-## Add new column with data types
-
-MSE_df_harps <- eval_data_training |>
-  dplyr::mutate(data_type = "training",
-                method = "tSNE")
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-## To draw with AIC
-aic_plot_harps <- ggplot(MSE_df_harps |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "tSNE"), aes(x = number_of_bins,
-                                                                                 y = total_error,
-                                                                                 color = method
-)) +
-  geom_point() +
-  geom_line() +
-  geom_vline(xintercept = 180, linetype="dashed", 
-                color = "black", size=0.5) +
-  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  scale_colour_manual(values = c("#e41a1c","#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("AIC") +
-  xlab("total number of bins")
-## Effective number of bins along x-axis
-
-mse_plot_harps <- ggplot(MSE_df_harps |> dplyr::filter(method == "tSNE"), aes(x = number_of_bins,
-                                       y = total_mse,
-                                       color = method
-)) +
-  geom_point() +
-  geom_line() +
-   geom_vline(xintercept = 180, linetype="dashed", 
-                color = "black", size=0.5) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  scale_colour_manual(values = c("#e41a1c","#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("MSE") +
-  xlab("total number of bins")
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset.
-What is the effective number of bins in each NLDR technique to create a 2D model?
-The MSE plot have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
-Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error.
-Using the elbow method, when the total number of bins is set to 144, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
-This point indicates that adding less bins does not enough to capture the data structure."
-#| label: fig-diagnosticpltHARPS3
-##| out-width: 100%
-#| fig-pos: H
-
-aic_plot_harps + mse_plot_harps +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', ncol = 2) &
-  theme(legend.position='none', plot.tag = element_text(size = 8))
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-## Import data
-# training_data_1 <- read_rds("data/harps/harps_data.rds")
-# training_data_1 <- training_data_1 |>
-#   mutate(ID = 1:NROW(training_data_1))
-# 
-# clusters <- training_data_1$cluster
-# select_names <-  names(training_data_1)[-c(18, length(training_data_1))]
-# 
-# training_data_1 <- training_data_1 |> select(-cluster)
-# names(training_data_1) <- append(paste0("x", 1:17), "ID")
-# 
-# tSNE_data_harps <- tSNE_data_harps |>
-#   select(tSNE1, tSNE2) |>
-#   mutate(ID = 1:NROW(tSNE_data_harps))
-
-### tSNE
-
-shape_val_tsne_harps <- calculate_effective_shape_value(.data = tSNE_data_harps,
-                                                      x = tSNE1, y = tSNE2)
-
-# num_bins_tsne_harps <- find_non_empty_bins(nldr_df = tSNE_data_harps, x = "tSNE1", y = "tSNE2", shape_val = shape_val_tsne_harps, non_empty_bins = num_non_empty_bins_harps)
-num_bins_tsne_harps <- 9
-
-## To extract bin centroids
-hexbin_data_object_tsne_harps <- extract_hexbin_centroids(nldr_df = tSNE_data_harps, num_bins = num_bins_tsne_harps, shape_val = shape_val_tsne_harps, x = tSNE1, y = tSNE2)
-
-df_bin_centroids_tsne_harps <- hexbin_data_object_tsne_harps$hexdf_data
-
-## Identify bins with low-density
-# identify_rm_bins <- find_low_density_hexagons(df_bin_centroids_tsne_harps, num_bins_tsne_harps, benchmark_rm_hex = 0.01)
-# 
-# df_bin_centroids_tsne_harps <- df_bin_centroids_tsne_harps |>
-#   filter(!(hexID %in% identify_rm_bins))
-
-tSNE_data_with_hb_id_harps <- tSNE_data_harps |>
-  dplyr::mutate(hb_id = hexbin_data_object_tsne_harps$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_tsne_harps <- dplyr::bind_cols(training_data_1 |> dplyr::select(-ID), tSNE_data_with_hb_id_harps)
-
-## Averaged on high-D
-df_bin_tsne_harps <- avg_highD_data(.data = df_all_tsne_harps)
-
-## Triangulate bin centroids
-tr1_object_tsne_harps <- triangulate_bin_centroids(df_bin_centroids_tsne_harps, x, y)
-tr_from_to_df_tsne_harps <- generate_edge_info(triangular_object = tr1_object_tsne_harps)
-
-## Compute 2D distances
-distance_tsne_harps <- cal_2d_dist(.data = tr_from_to_df_tsne_harps)
-
-# distance_plot <- plot_dist(distance_tsne_harps) +
-#   #ggtitle("(b)" ) +
-#   ylab(expression(d^{(2)})) +
-#   theme(axis.text = element_text(size = 5),
-#         axis.title = element_text(size = 12))
-# 
-# distance_plot
-
-## To find the benchmark value
-benchmark_tsne_harps <- find_benchmark_value(.data = distance_tsne_harps, distance_col = "distance")
-benchmark_tsne_harps <- 2.5
-
-trimesh_removed_tsne_harps <- remove_long_edges(.data = distance_tsne_harps, benchmark_value = benchmark_tsne_harps,
-                                              triangular_object = tr1_object_tsne_harps, distance_col = distance)
-
-trimesh_removed_tsne_harps <- trimesh_removed_tsne_harps +
-  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  # theme(axis.text = element_text(size = 5),
-  #       axis.title = element_text(size = 7)) +
-  geom_point(colour = "#33a02c", size = 0.5) +
-  coord_cartesian(xlim = c(-8, 7), ylim = c(-10, 16)) +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-df_bin_train <- df_bin_tsne_harps
-names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])
-
-error_df <- df_all_tsne_harps |>
-  dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates
-
-# prediction_df_join <- prediction_df_join |>
-#   dplyr::left_join(data, by = c("ID" = "ID")) ## Map high-D data
-
-for (i in 1:(NCOL(df_bin_train) - 1)) {
-
-  error_df[ , paste0("abs_residual_", "x", i)] <- abs(error_df[ , paste0("x", i)] - error_df[ , paste0("avg_", "x", i)])
-
-}
-
-error_df <- error_df |>
-  dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "x")))))
-
-error_df <- error_df |>
-  mutate(type = if_else(total <= 0, "no error",
-                        if_else(total <= 0.2, "error 0-0.2",
-                                if_else(total <= 0.3, "error 0.2-0.3",
-                                        if_else(total <= 0.5, "error 0.3-0.5",
-                                                if_else(total <= 1, "error 0.5-1",
-                                                        "error greter than 1"))))))
-
-plot_list2_harps_error <- error_df |>
-    mutate(type = factor(type , levels = c("no error", "error 0-0.2", "error 0.2-0.3", "error 0.3-0.5", "error 0.5-1", "error greter than 1"))) |>
-    ggplot(aes(x = tSNE1,
-               y = tSNE2, color = type))+
-    geom_point(alpha=0.5, size = 0.1) +
-    coord_equal() +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#fee0d2", "#fcbba1",
-                                         "#fc9272", "#fb6a4a", "#ef3b2c",
-                                         "#cb181d", "#a50f15", "#99000d"))
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "(a) 2D layout from tSNE (perplexity = 91) applied for the HARPS GTO dataset.
-Is this a best representation of the original data?, (b) Model in the 2D space with tSNE (<https://youtu.be/vLvRfcogft8>)"
-#| label: fig-nldervisharpstSNE3
-#| fig-pos: h
-
-plot_list2_harps + trimesh_removed_tsne_harps + plot_list2_harps_error +
-  plot_layout(guides='collect', ncol = 3) &
-  theme(legend.position='bottom', plot.tag = element_text(size = 8))
-
-```
-
 ## Conclusion {#sec-conclusion}

### 2024-03-06 paper.qmd
Correct the issue getting same combinations of bins along x and y axes in PBMC data

@@
 
-## If same total number of bins ocurred only select ones with minimum error
-duplicate_df_list <- mse_df_pbmc[duplicated(mse_df_pbmc$num_bins), ] |>
+## If same total number of bins occurred only select ones with minimum error
+### Obtain duplicate bins
+dupli_bins <- mse_df_pbmc |> 
+  dplyr::count(num_bins) |> 
+  dplyr::filter(n > 1) |> 
+  dplyr::pull(num_bins)
+
+duplicate_df_list <- mse_df_pbmc |>
+  dplyr::filter(num_bins %in% dupli_bins) |>
   dplyr::arrange(num_bins) |>
@@
   dd <- duplicate_df_list[[i]] |>
-    dplyr::filter(mse == min(duplicate_df_list[[i]]$mse))
+    dplyr::filter(mse == min(duplicate_df_list[[i]]$mse)) |>
+    dplyr::filter(hex_size == min(duplicate_df_list[[i]]$hex_size))
 
@@
 
-mse_df_pbmc <- dplyr::bind_rows(mse_df_pbmc[!duplicated(mse_df_pbmc$num_bins), ], duplicate_df)
+not_dupli_df <- mse_df_pbmc |>
+  dplyr::filter(!(num_bins %in% dupli_bins))
+
+mse_df_pbmc <- dplyr::bind_rows(not_dupli_df, duplicate_df)
 ```

### 2024-03-06 appendix.qmd
Make the plan for the methodology

@@
 $$ {#eq-equation2}
+
+<!--UMAP parm 2-->
+
+```{r}
+#| warning: false
+#| echo: false
+
+## Import data 
+training_data_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_pca_50.rds")
+training_data_pbmc <- training_data_pbmc[, 1:15] |>
+  mutate(ID = 1:NROW(training_data_pbmc))
+
+UMAP_pbmc <- read_rds("data/pbmc/pbmc_umap_15_min_dist_0.99_metric_cosine.rds")
+
+class_avg <- UMAP_pbmc |>
+  group_by(cell_label) |>
+  summarise(
+    UMAP1 = median(UMAP1),
+    UMAP2 = median(UMAP2)
+  ) |>
+  mutate(cell_label = as.numeric(cell_label)) |>
+  arrange(cell_label)
+
+class_avg$cell_label <- as.factor(class_avg$cell_label)
+
+levels(class_avg$cell_label) <- c("Memory \nCD4 T", "Naive CD4 T", "CD14+ Mono",  "B", "CD8 T", "FCGR3A+ \n Mono", "NK", "M-MDSC\n-like", "CD27-CD+ \n Memory T", "DC")
+
+plot_list2_pbmc <- UMAP_pbmc |>
+    ggplot(aes(x = UMAP1,
+               y = UMAP2, color = cell_label))+
+    geom_point(alpha=0.5) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
+  geom_text(aes(x=UMAP1,y = UMAP2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## UMAP
+## Prediction
+
+shape_value_pbmc <- calculate_effective_shape_value(.data = UMAP_pbmc,
+                                                      x = UMAP1, y = UMAP2)
+
+num_bins_vec <- 1:35 ## Number of bins along the x-axis
+
+vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names
+
+eval_data_training <- dplyr::bind_rows(vec)[0, ]
+eval_data_training <- eval_data_training |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(num_bins_vec)) {
+
+  model_object <- fit_high_d_model(training_data = training_data_pbmc, nldr_df_with_id = UMAP_pbmc, x = "UMAP1", y = "UMAP2", num_bins_x = num_bins_vec[i], shape_val = shape_value_pbmc,
+                             is_bin_centroid = TRUE,
+                             is_rm_lwd_hex = FALSE,
+                             benchmark_to_rm_lwd_hex = NA,
+                             is_avg_high_d = TRUE, column_start_text = "PC_")
+
+  centroid_df_training <- model_object$df_bin_centroids
+  avg_df_training <- model_object$df_bin
+
+  pred_df_training <- predict_2d_embeddings(test_data = training_data_pbmc, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "UMAP")
+
+  eval_df_training <- generate_eval_df(data = training_data_pbmc, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, col_start = "PC_")
+
+  eval_df_training <- eval_df_training |>
+    mutate(num_bins_x = num_bins_vec[i])
+
+  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+
+}
+
+## Add new column with data types
+
+MSE_df_pbmc <- eval_data_training |>
+  dplyr::mutate(data_type = "training",
+                method = "UMAP")
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## To draw with AIC
+aic_plot_pbmc <- ggplot(MSE_df_pbmc |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
+                                                                                 y = total_error,
+                                                                                 color = method
+)) +
+  geom_point() +
+  geom_line() +
+  geom_vline(xintercept = 192, linetype="dashed", 
+                color = "red", size=0.5) +
+  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("AIC") +
+  xlab("total number of bins")
+## Effective number of bins along x-axis
+
+mse_plot_pbmc <- ggplot(MSE_df_pbmc |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
+                                       y = total_mse,
+                                       color = method
+)) +
+  geom_point() +
+  geom_line() +
+   geom_vline(xintercept = 192, linetype="dashed", 
+                color = "red", size=0.5) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
+  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
+  ylab("MSE") +
+  xlab("total number of bins")
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset.
+What is the effective number of bins in each NLDR technique to create a 2D model?
+The MSE plot have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
+Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error.
+Using the elbow method, when the total number of bins is set to 144, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
+This point indicates that adding less bins does not enough to capture the data structure."
+#| label: fig-diagnosticpltPBMC2
+##| out-width: 100%
+#| fig-pos: H
+
+aic_plot_pbmc + mse_plot_pbmc +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='none', plot.tag = element_text(size = 8))
+```
+
+```{r}
+#| warning: false
+#| echo: false
+shape_value_pbmc <- calculate_effective_shape_value(.data = UMAP_pbmc,
+                                                      x = UMAP1, y = UMAP2) ## 0.8772751
+
+# num_bins_pbmc <- find_non_empty_bins(nldr_df = UMAP_pbmc, x = "UMAP1", y = "UMAP2", shape_val = shape_value_pbmc, non_empty_bins = num_non_empty_bins_pbmc)
+num_bins_pbmc <-  11
+
+## To extract bin centroids
+hexbin_data_object_pbmc <- extract_hexbin_centroids(nldr_df = UMAP_pbmc,
+                                                      num_bins = num_bins_pbmc,
+                                                      shape_val = shape_value_pbmc, x = UMAP1, y = UMAP2)
+
+df_bin_centroids_pbmc <- hexbin_data_object_pbmc$hexdf_data
+
+UMAP_pbmc_with_hb_id <- UMAP_pbmc |>
+  dplyr::mutate(hb_id = hexbin_data_object_pbmc$hb_data@cID)
+
+## To generate a data set with high-D and 2D training data
+df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), UMAP_pbmc_with_hb_id)
+
+## Averaged on high-D
+df_bin_pbmc <- avg_highD_data(.data = df_all_pbmc, column_start_text = "PC")
+
+## Triangulate bin centroids
+tr1_object_pbmc <- triangulate_bin_centroids(df_bin_centroids_pbmc, x, y)
+tr_from_to_df_pbmc <- generate_edge_info(triangular_object = tr1_object_pbmc)
+
+## Compute 2D distances
+distance_pbmc <- cal_2d_dist(.data = tr_from_to_df_pbmc)
+
+# distance_plot <- plot_dist(distance_pbmc) +
+#   #ggtitle("(b)" ) +
+#   ylab(expression(d^{(2)})) +
+#   theme(axis.text = element_text(size = 5),
+#         axis.title = element_text(size = 12))
+# 
+# distance_plot
+
+## To find the benchmark value
+benchmark_pbmc <- find_benchmark_value(.data = distance_pbmc, distance_col = distance)
+#benchmark_pbmc <- 1.049766 
+benchmark_pbmc <- 2.5
+
+trimesh_removed_pbmc_umap <- remove_long_edges(.data = distance_pbmc, benchmark_value = benchmark_pbmc,
+                                                 triangular_object = tr1_object_pbmc, distance_col = distance)
+
+trimesh_removed_pbmc_umap <- trimesh_removed_pbmc_umap +
+  geom_point(size = 0.005, colour = "#33a02c") +
+  #ggtitle("(b)") +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+df_bin_train <- df_bin_pbmc
+names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])
+
+error_df <- df_all_pbmc |>
+  dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates
+
+# prediction_df_join <- prediction_df_join |>
+#   dplyr::left_join(data, by = c("ID" = "ID")) ## Map high-D data
+
+for (i in 1:(NCOL(df_bin_train) - 1)) {
+
+  error_df[ , paste0("abs_residual_", "PC_", i)] <- abs(error_df[ , paste0("PC_", i)] - error_df[ , paste0("avg_", "PC_", i)])
+
+}
+
+error_df <- error_df |>
+  dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "PC_")))))
+
+# library(ggbeeswarm)
+# error_df$group <- "1"
+# ggplot(error_df, aes(x = group, y = total)) +
+#   geom_quasirandom()+
+#   ylim(0, max(unlist(error_df$total))+ 0.5) + coord_flip()
+
+### The minimum error is 0 and the maximum is 42.17439
+### There is lot of points with error 0,
+
+error_df <- error_df |>
+  mutate(type = if_else(total <= 2, "error 2 or less",
+                        if_else(total <= 10, "error 2-10",
+                                if_else(total <= 15, "error 10-15",
+                                        if_else(total <= 20, "error 15-20",
+                                                if_else(total <= 25, "error 20-25",
+                                                        if_else(total <= 30, "error 25-30",
+                                                                if_else(total <= 35, "error 30-35", "error greter than 35"))))))))
+
+plot_list2_pbmc_error <- error_df |>
+    mutate(type = factor(type , levels = c("error 2 or less", "error 2-10", "error 10-15", "error 15-20", "error 20-25", "error 25-30", "error 30-35", "error greter than 35"))) |>
+    ggplot(aes(x = UMAP1,
+               y = UMAP2, color = type))+
+    geom_point(alpha=0.5, size = 0.1) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#fee0d2", "#fcbba1",
+                                         "#fc9272", "#fb6a4a", "#ef3b2c",
+                                         "#cb181d", "#a50f15", "#99000d"))
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "(a) 2D layout from UMAP (n_neighbors = 15, min_dist = 0.99, metric = cosine) applied for the PBMC3k dataset.
+Is this a best representation of the original data?, (b)  Model in the 2D space with UMAP (<>)"
+#| label: fig-nldervisPBMCUMAP2
+#| out-width: 100%
+#| fig-pos: h
+
+plot_list2_pbmc + trimesh_removed_pbmc_umap + plot_list2_pbmc_error +
+plot_layout(guides='collect', ncol=3) &
+  theme(legend.position='bottom')
+```
 -->

### 2024-03-06 paper.qmd
Make the plan for the methodology

@@
 
-<!--In this paper, we introduce a novel approach, which is designed to determine which method, and which parameter choice provide the most useful representation of high dimensional data.
-Our approach involves dividing the data set into two subsets: a training set, which serves to construct the low-dimensional manifold, and a test set, utilized to generate predictive values and residuals.
-
-For the implementation of our approach, a 2D embedding data set serves as the initiation point for the subsequent analysis and visualization (see @fig-nldervis (b)).-->
-
-<!--### Overview {#sec-algorithm}-->
-
-In this paper, we present a novel method that aims to determine the most effective way to represent high-D data by selecting the best method and hyperparameter choice.
-Our approach involves dividing the data set into two parts: a training set to construct the model and a test set to generate predictive values and residuals.
-To implement our approach, we first use a 2D embedding data set as the initial point.
-The output of our algorithm is a tour that displays the model and original data in the high-dimensional space.
-A flow chart of the proposed algorithm is shown in @fig-meth.
-Our algorithm comprises two main phases: (1) generate the model in the 2D space, and (2) generate the model in the high-D space.
-These two phases are described in detail in this section using UMAP 2D embedding of the S-curve data set, which has seven dimensions, including four noise dimensions.
-
-![A flow diagram detailing the steps taken to create the low-dimensional manifold in the high dimensional space.
-There are two basic phases, one to generate the model in the 2D space, and other to generate the model in the high-D space.](figures/workflow.png){#fig-meth fig-align="center" width="100%" height="100%"}
+In this paper, we introduce a novel method to determine the most effective NLDR technique and the best hyperparameter choice that provides the most useful representation of high-D data.
+Our approach involves dividing the high-dimensional dataset into two parts: a training set for constructing the model and a test set for generating predictive values and residuals.
+Our algorithm takes a 2D embedding data as the input and generate a tour that displays the high-dimensional wireframe to overlay the data.
+The flow chart of the proposed algorithm is shown in @fig-meth.
+The algorithm consists of two main phases: (1) generating the model in the 2D space and (2) lifting the model into high-D space.
+The main steps of the algorithm are described in detail in this section using UMAP 2D embedding of the S-curve dataset.
+This dataset has seven dimensions, including four noise dimensions that were added to the original 3D data.
+
+![The flow diagram shows the main steps of our algorithm.
+There are two basic phases, one to generate the model in the 2D space, and other to map the model into the high-D space.](figures/workflow.png){#fig-meth fig-align="center" width="100%" height="100%"}
 
@@
 ### Constructing the 2D model {#sec-construct2d}
+
+**Step 1: Scaling NLDR data**
+
+**Step 2: Hexagonating NLDR data**
+
+**(a) Determine the number of bins along the x axis** ($b_1$)
+
+**(b) Determine the number of bins along the y axis** ($b_2$)
+
+**(c) Determine the total number of bins** ($b$)
+
+**(d) Assign NLDR data to hexagons**
+
+**Step 3: Obtaining bin centroids or bin means**
+
+**Step 4: Triangulating bin centroids or bin means**
 
@@
 
-```{r}
-#| echo: false
+<!--
+```{r}
+#| echo: false
+#| eval: false
 
@@
 #| echo: false
+#| eval: false
 
@@
 #| warning: false
+#| eval: false
+
 data <- read_rds("data/s_curve/s_curve.rds")
@@
 ```{r}
+#| eval: false
 #| warning: false
@@
 
+<!--                  
 ```{r}
@@
 #| warning: false
+#| eval: false
 hex_full_count_df_loop1 <- read_rds("data/s_curve/s_curve_hex_3.rds")
@@
 #| warning: false
+#| eval: false
 hex_full_count_df_loop2 <- read_rds("data/s_curve/s_curve_hex_11.rds")
@@
 #| warning: false
+#| eval: false
 hex_full_count_df_loop3 <- read_rds("data/s_curve/s_curve_hex_20.rds")
@@
 ```{r}
+#| eval: false
 #| echo: false
@@
 
+<!--
 ```{r}
@@
 #| warning: false
+#| eval: false
 
@@
 
+-->
+
 <!--Addressing low-density hexagons is a systematic process to handle sparsely represented data in certain regions.
@@
 <!--lwd smoothing criteria 1 with s-curve-->
-
-```{r}
-#| warning: false
-#| echo: false
+<!--
+```{r}
+#| warning: false
+#| echo: false
+#| eval: false
 
@@
 This refinement step contributes to enhancing the quality of the triangular mesh, ensuring a more accurate representation of the data structure.-->
-
-```{r}
-#| warning: false
-#| echo: false
+<!--
+```{r}
+#| warning: false
+#| echo: false
+#| eval: false
 
@@
 ```
-
-#### Hexagonal grid origin
+-->
+
+#### Starting point of the hexagonal grid
 
@@
 
-```{r}
+<!--
+
+```{r}
+#| eval: false
 #| echo: false
@@
 
-```{r}
-#| warning: false
-#| echo: false
+<!--
+```{r}
+#| warning: false
+#| echo: false
+#| eval: false
 
@@
 
+### Group split by duplicated bins
 duplicate_df_list <- mse_df_pbmc |>
@@
 
+### Obtain one row from duplicates which have lowest error and hexsize
 duplicate_df <- data.frame(matrix(nrow = 0, ncol = 0))
@@
 
+### Obtain the mse_df with not duplicated bins
 not_dupli_df <- mse_df_pbmc |>
@@
 
+### Combine duplicated and not duplicated(corrected) bins dfs
 mse_df_pbmc <- dplyr::bind_rows(not_dupli_df, duplicate_df)
@@
   geom_line() +
-  # geom_vline(xintercept = 256, linetype="dashed", 
-  #               color = "red", size=0.5) +
-  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  geom_vline(xintercept = 80, linetype="solid",
+                color = "red", size=0.8, alpha = 0.5) +
   theme_light() +
@@
         axis.text = element_text(size = 7)) +
-  scale_colour_manual(values = c("#377eb8")) +
+  #scale_colour_manual(values = c("#377eb8")) +
   ylab("AIC") +
@@
   geom_line() +
-   # geom_vline(xintercept = 256, linetype="dashed", 
-   #              color = "red", size=0.5) +
+   geom_vline(xintercept = 80, linetype="solid",
+                color = "red", size=0.8, alpha = 0.5) +
   theme_light() +
@@
         axis.text = element_text(size = 7)) +
-  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  scale_colour_manual(values = c("#377eb8")) +
+  #scale_colour_manual(values = c("#377eb8")) +
   ylab("MSE") +
@@
 #| echo: false
-# num_bins_pbmc <- calculate_effective_x_bins(.data = UMAP_pbmc, x = UMAP1,
-#                                               cell_area = 1) ##23
-
-num_bins_pbmc <- 15
-
-## To extract bin centroids
-hexbin_data_object_pbmc <- extract_hexbin_centroids(nldr_df = UMAP_pbmc,
-                                                      num_bins = num_bins_pbmc,
-                                                      shape_val = shape_value_pbmc, x = UMAP1, y = UMAP2)
-
-df_bin_centroids_pbmc <- hexbin_data_object_pbmc$hexdf_data
-
-#num_non_empty_bins_pbmc <- df_bin_centroids_pbmc$hexID |> length() #83
-
-UMAP_pbmc_with_hb_id <- UMAP_pbmc |>
-  dplyr::mutate(hb_id = hexbin_data_object_pbmc$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), UMAP_pbmc_with_hb_id)
-
-## Averaged on high-D
-df_bin_pbmc <- avg_highD_data(.data = df_all_pbmc, column_start_text = "PC")
+#| message: false
+
+## Decide by looking at MSE plot
+num_bins_x_pbmc <- 8
+num_bins_y_pbmc <- 10
+hex_size_pbmc <- 0.1
+
+hb_obj_pbmc <- hex_binning(data = umap_pbmc_scaled |> dplyr::select(-cell_label), x = "UMAP1", 
+                      y = "UMAP2", num_bins_x = num_bins_x_pbmc, 
+                      num_bins_y = num_bins_y_pbmc, x_start = NA, y_start = NA, 
+                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_pbmc, col_start = "UMAP")
+
+## Data set with all possible centroids in the hexagonal grid
+all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_pbmc$centroids))
+
+## Generate all coordinates of hexagons
+hex_grid <- as.data.frame(do.call(cbind, hb_obj_pbmc$hex_poly))
+
+## To obtain the standardise counts within hexbins
+counts_df <- as.data.frame(do.call(cbind, hb_obj_pbmc$std_cts))
+df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
+                                             counts_df = counts_df)
+ggplot(data = hex_grid, aes(x = x, y = y)) + 
+  geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
+  geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
+  coord_fixed()
+
+hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
+
+ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+  geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
+  coord_fixed()
+
+umap_data_with_hb_id_pbmc <- as.data.frame(do.call(cbind, hb_obj_pbmc$data_hb_id))
+
+model_object <- fit_highd_model( training_data = training_data_pbmc, 
+                                 nldr_df_with_id = umap_pbmc_scaled |> dplyr::select(-cell_label), 
+                                 x = "UMAP1", y = "UMAP2", 
+                                 num_bins_x = num_bins_x_pbmc, 
+                                 num_bins_y = num_bins_y_pbmc, 
+                                 x_start = NA, y_start = NA, 
+                                 buffer_x = NA, buffer_y = NA, 
+                                 hex_size = hex_size_pbmc,
+                                 is_rm_lwd_hex = FALSE, 
+                                 benchmark_to_rm_lwd_hex = NA, 
+                                 col_start_2d = "UMAP", 
+                                 col_start_highd = "PC_")
+
+df_bin_centroids_pbmc <- model_object$df_bin_centroids
+df_bin_pbmc <- model_object$df_bin
 
 ## Triangulate bin centroids
-tr1_object_pbmc <- triangulate_bin_centroids(df_bin_centroids_pbmc, x, y)
-tr_from_to_df_pbmc <- generate_edge_info(triangular_object = tr1_object_pbmc)
+tr1_object_pbmc <- tri_bin_centroids(df_bin_centroids_pbmc, x = "c_x", y = "c_y")
+tr_from_to_df_pbmc <- gen_edges(tri_object = tr1_object_pbmc)
 
 ## Compute 2D distances
-distance_pbmc <- cal_2d_dist(.data = tr_from_to_df_pbmc)
+distance_pbmc <- cal_2d_dist(tr_coord_df = tr_from_to_df_pbmc, 
+                             start_x = "x_from", start_y = "y_from", 
+                             end_x = "x_to", end_y = "y_to", 
+                             select_vars = c("from", "to", "distance"))
 
 distance_plot <- plot_dist(distance_pbmc) +
-  #ggtitle("(b)" ) +
   ylab(expression(d^{(2)})) +
@@
 ## To find the benchmark value
-benchmark_pbmc <- find_benchmark_value(.data = distance_pbmc, distance_col = distance)
-benchmark_pbmc <- 2.5
-
-trimesh_removed_pbmc_umap <- remove_long_edges(.data = distance_pbmc, benchmark_value = benchmark_pbmc,
-                                                 triangular_object = tr1_object_pbmc, distance_col = distance)
+benchmark_pbmc <- find_lg_benchmark(distance_edges = distance_pbmc, distance_col = "distance")
+
+ggplot() +
+geom_trimesh(data = df_bin_centroids_pbmc, mapping = aes(x = c_x, y = c_y))
+
+trimesh_removed_pbmc_umap <- vis_rmlg_mesh(distance_edges = distance_pbmc, benchmark_value = benchmark_pbmc, tr_coord_df = tr_from_to_df_pbmc, distance_col = "distance")
 
 trimesh_removed_pbmc_umap <- trimesh_removed_pbmc_umap +
-  geom_point(size = 1, colour = "#33a02c") +
-  #ggtitle("(b)") +
   theme_linedraw() +
@@
 names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])
+
+df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), umap_data_with_hb_id_pbmc)
 
@@
 error_df <- error_df |>
-  mutate(type = if_else(total <= 2, "error 2 or less",
-                        if_else(total <= 10, "error 2-10",
-                                if_else(total <= 15, "error 10-15",
+  mutate(type = if_else(total == 0, "no error",
+                        if_else(total <= 15, "error 1-15",
                                 if_else(total <= 20, "error 15-20",
@@
                                                 if_else(total <= 30, "error 25-30",
-                                                                if_else(total <= 35, "error 30-35", "error greter than 35"))))))))
+                                                        if_else(total <= 35, "error 30-35",
+                                                                if_else(total <= 40, "error 35-40", "error greter than 40"))))))))
 
 plot_list2_pbmc_error <- error_df |>
-  mutate(type = factor(type , levels = c("error 2 or less", "error 2-10", "error 10-15", "error 15-20", "error 20-25", "error 25-30", "error 30-35", "error greter than 35"))) |>
+  mutate(type = factor(type , levels = c("no error", "error 1-15", "error 15-20", "error 20-25", "error 25-30", "error 30-35", "error 35-40", "error greter than 40"))) |>
     ggplot(aes(x = UMAP1,
@@
 
-<!--UMAP parm 2-->
+<!--UMAP parm option 1-->
 
@@
 
-UMAP_pbmc <- read_rds("data/pbmc/pbmc_umap_15_min_dist_0.99_metric_cosine.rds")
-
-class_avg <- UMAP_pbmc |>
+umap_pbmc <- read_rds("data/pbmc/pbmc_umap_5_min_dist_0.99_metric_cosine.rds")
+umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc, 
+                                    x = "UMAP1", y = "UMAP2"))) |>
+  dplyr::mutate(cell_label = umap_pbmc$cell_label) |>
+  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
+                  "UMAP2" = "scaled_UMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(umap_pbmc))
+
+class_avg <- umap_pbmc_scaled |>
   group_by(cell_label) |>
@@
 #| echo: false
+#| message: false
 
@@
 
-## Add new column with data types
-
-MSE_df_pbmc <- eval_data_training |>
-  dplyr::mutate(data_type = "training",
-                method = "UMAP")
+## If same total number of bins occurred only select ones with minimum error
+### Obtain duplicate bins
+dupli_bins <- mse_df_pbmc |> 
+  dplyr::count(num_bins) |> 
+  dplyr::filter(n > 1) |> 
+  dplyr::pull(num_bins)
+
+### Group split by duplicated bins
+duplicate_df_list <- mse_df_pbmc |>
+  dplyr::filter(num_bins %in% dupli_bins) |>
+  dplyr::arrange(num_bins) |>
+  dplyr::group_split(num_bins)
+
+### Obtain one row from duplicates which have lowest error and hexsize
+duplicate_df <- data.frame(matrix(nrow = 0, ncol = 0))
+
+for (i in 1:length(duplicate_df_list)) {
+
+  dd <- duplicate_df_list[[i]] |>
+    dplyr::filter(mse == min(duplicate_df_list[[i]]$mse)) |>
+    dplyr::filter(hex_size == min(duplicate_df_list[[i]]$hex_size))
+
+  duplicate_df <- dplyr::bind_rows(duplicate_df, dd)
+
+}
+
+### Obtain the mse_df with not duplicated bins
+not_dupli_df <- mse_df_pbmc |>
+  dplyr::filter(!(num_bins %in% dupli_bins))
+
+### Combine duplicated and not duplicated(corrected) bins dfs
+mse_df_pbmc <- dplyr::bind_rows(not_dupli_df, duplicate_df)
 ```
@@
   geom_line() +
-  geom_vline(xintercept = 192, linetype="dashed", 
-                color = "red", size=0.5) +
-  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+  geom_vline(xintercept = 88, linetype="solid",
+                color = "red", size=1, alpha = 0.5) +
   theme_light() +
@@
         axis.text = element_text(size = 7)) +
-  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
+  #scale_colour_manual(values = c("#377eb8")) +
   ylab("AIC") +
@@
   geom_line() +
-   geom_vline(xintercept = 192, linetype="dashed", 
-                color = "red", size=0.5) +
+   geom_vline(xintercept = 88, linetype="solid",
+                color = "red", size=1, alpha = 0.5) +
   theme_light() +
@@
         axis.text = element_text(size = 7)) +
-  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("MSE") +
-  xlab("total number of bins")
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset.
-What is the effective number of bins in each NLDR technique to create a 2D model?
-The MSE plot have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
-Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error.
-Using the elbow method, when the total number of bins is set to 144, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
-This point indicates that adding less bins does not enough to capture the data structure."
-#| label: fig-diagnosticpltPBMC2
-##| out-width: 100%
-#| fig-pos: H
-
-aic_plot_pbmc + mse_plot_pbmc +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', ncol = 2) &
-  theme(legend.position='none', plot.tag = element_text(size = 8))
-```
-
-```{r}
-#| warning: false
-#| echo: false
-shape_value_pbmc <- calculate_effective_shape_value(.data = UMAP_pbmc,
-                                                      x = UMAP1, y = UMAP2) ## 0.8772751
-
-# num_bins_pbmc <- find_non_empty_bins(nldr_df = UMAP_pbmc, x = "UMAP1", y = "UMAP2", shape_val = shape_value_pbmc, non_empty_bins = num_non_empty_bins_pbmc)
-num_bins_pbmc <-  11
-
-## To extract bin centroids
-hexbin_data_object_pbmc <- extract_hexbin_centroids(nldr_df = UMAP_pbmc,
-                                                      num_bins = num_bins_pbmc,
-                                                      shape_val = shape_value_pbmc, x = UMAP1, y = UMAP2)
-
-df_bin_centroids_pbmc <- hexbin_data_object_pbmc$hexdf_data
-
-UMAP_pbmc_with_hb_id <- UMAP_pbmc |>
-  dplyr::mutate(hb_id = hexbin_data_object_pbmc$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), UMAP_pbmc_with_hb_id)
-
-## Averaged on high-D
-df_bin_pbmc <- avg_highD_data(.data = df_all_pbmc, column_start_text = "PC")
-
-## Triangulate bin centroids
-tr1_object_pbmc <- triangulate_bin_centroids(df_bin_centroids_pbmc, x, y)
-tr_from_to_df_pbmc <- generate_edge_info(triangular_object = tr1_object_pbmc)
-
-## Compute 2D distances
-distance_pbmc <- cal_2d_dist(.data = tr_from_to_df_pbmc)
-
-# distance_plot <- plot_dist(distance_pbmc) +
-#   #ggtitle("(b)" ) +
-#   ylab(expression(d^{(2)})) +
-#   theme(axis.text = element_text(size = 5),
-#         axis.title = element_text(size = 12))
-# 
-# distance_plot
-
-## To find the benchmark value
-benchmark_pbmc <- find_benchmark_value(.data = distance_pbmc, distance_col = distance)
-#benchmark_pbmc <- 1.049766 
-benchmark_pbmc <- 2.5
-
-trimesh_removed_pbmc_umap <- remove_long_edges(.data = distance_pbmc, benchmark_value = benchmark_pbmc,
-                                                 triangular_object = tr1_object_pbmc, distance_col = distance)
-
-trimesh_removed_pbmc_umap <- trimesh_removed_pbmc_umap +
-  geom_point(size = 0.005, colour = "#33a02c") +
-  #ggtitle("(b)") +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-df_bin_train <- df_bin_pbmc
-names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])
-
-error_df <- df_all_pbmc |>
-  dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates
-
-# prediction_df_join <- prediction_df_join |>
-#   dplyr::left_join(data, by = c("ID" = "ID")) ## Map high-D data
-
-for (i in 1:(NCOL(df_bin_train) - 1)) {
-
-  error_df[ , paste0("abs_residual_", "PC_", i)] <- abs(error_df[ , paste0("PC_", i)] - error_df[ , paste0("avg_", "PC_", i)])
-
-}
-
-error_df <- error_df |>
-  dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "PC_")))))
-
-# library(ggbeeswarm)
-# error_df$group <- "1"
-# ggplot(error_df, aes(x = group, y = total)) +
-#   geom_quasirandom()+
-#   ylim(0, max(unlist(error_df$total))+ 0.5) + coord_flip()
-
-### The minimum error is 0 and the maximum is 42.17439
-### There is lot of points with error 0,
-
-error_df <- error_df |>
-  mutate(type = if_else(total <= 2, "error 2 or less",
-                        if_else(total <= 10, "error 2-10",
-                                if_else(total <= 15, "error 10-15",
-                                        if_else(total <= 20, "error 15-20",
-                                                if_else(total <= 25, "error 20-25",
-                                                        if_else(total <= 30, "error 25-30",
-                                                                if_else(total <= 35, "error 30-35", "error greter than 35"))))))))
-
-plot_list2_pbmc_error <- error_df |>
-    mutate(type = factor(type , levels = c("error 2 or less", "error 2-10", "error 10-15", "error 15-20", "error 20-25", "error 25-30", "error 30-35", "error greter than 35"))) |>
-    ggplot(aes(x = UMAP1,
-               y = UMAP2, color = type))+
-    geom_point(alpha=0.5, size = 0.1) +
-    coord_equal() +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#fee0d2", "#fcbba1",
-                                         "#fc9272", "#fb6a4a", "#ef3b2c",
-                                         "#cb181d", "#a50f15", "#99000d"))
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "(a) 2D layout from UMAP (n_neighbors = 15, min_dist = 0.99, metric = cosine) applied for the PBMC3k dataset.
-Is this a best representation of the original data?, (b)  Model in the 2D space with UMAP (<>)"
-#| label: fig-nldervisPBMCUMAP2
-#| out-width: 100%
-#| fig-pos: h
-
-plot_list2_pbmc + trimesh_removed_pbmc_umap + plot_list2_pbmc_error +
-plot_layout(guides='collect', ncol=3) &
-  theme(legend.position='bottom')
-```
-
-<!--UMAP parm 3-->
-
-```{r}
-#| warning: false
-#| echo: false
-
-## Import data
-training_data_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_pca_50.rds")
-training_data_pbmc <- training_data_pbmc[, 1:15] |>
-  mutate(ID = 1:NROW(training_data_pbmc))
-
-UMAP_pbmc <- read_rds("data/pbmc/pbmc_umap_5_min_dist_0.99_metric_cosine.rds")
-
-class_avg <- UMAP_pbmc |>
-  group_by(cell_label) |>
-  summarise(
-    UMAP1 = median(UMAP1),
-    UMAP2 = median(UMAP2)
-  ) |>
-  mutate(cell_label = as.numeric(cell_label)) |>
-  arrange(cell_label)
-
-class_avg$cell_label <- as.factor(class_avg$cell_label)
-
-levels(class_avg$cell_label) <- c("Memory \nCD4 T", "Naive CD4 T", "CD14+ Mono",  "B", "CD8 T", "FCGR3A+ \n Mono", "NK", "M-MDSC\n-like", "CD27-CD+ \n Memory T", "DC")
-
-plot_list2_pbmc <- UMAP_pbmc |>
-    ggplot(aes(x = UMAP1,
-               y = UMAP2, color = cell_label))+
-    geom_point(alpha=0.5) +
-    coord_equal() +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
-  geom_text(aes(x=UMAP1,y = UMAP2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-## UMAP
-## Prediction
-
-shape_value_pbmc <- calculate_effective_shape_value(.data = UMAP_pbmc,
-                                                      x = UMAP1, y = UMAP2)
-
-num_bins_vec <- 1:35 ## Number of bins along the x-axis
-
-vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names
-
-eval_data_training <- dplyr::bind_rows(vec)[0, ]
-eval_data_training <- eval_data_training |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(num_bins_vec)) {
-
-  model_object <- fit_high_d_model(training_data = training_data_pbmc, nldr_df_with_id = UMAP_pbmc, x = "UMAP1", y = "UMAP2", num_bins_x = num_bins_vec[i], shape_val = shape_value_pbmc,
-                             is_bin_centroid = TRUE,
-                             is_rm_lwd_hex = FALSE,
-                             benchmark_to_rm_lwd_hex = NA,
-                             is_avg_high_d = TRUE, column_start_text = "PC_")
-
-  centroid_df_training <- model_object$df_bin_centroids
-  avg_df_training <- model_object$df_bin
-
-  pred_df_training <- predict_2d_embeddings(test_data = training_data_pbmc, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "UMAP")
-
-  eval_df_training <- generate_eval_df(data = training_data_pbmc, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, col_start = "PC_")
-
-  eval_df_training <- eval_df_training |>
-    mutate(num_bins_x = num_bins_vec[i])
-
-  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
-
-}
-
-## Add new column with data types
-
-MSE_df_pbmc <- eval_data_training |>
-  dplyr::mutate(data_type = "training",
-                method = "UMAP")
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-## To draw with AIC
-aic_plot_pbmc <- ggplot(MSE_df_pbmc |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
-                                                                                 y = total_error,
-                                                                                 color = method
-)) +
-  geom_point() +
-  geom_line() +
-  geom_vline(xintercept = 182, linetype="dashed", 
-                color = "red", size=0.5) +
-  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("AIC") +
-  xlab("total number of bins")
-## Effective number of bins along x-axis
-
-mse_plot_pbmc <- ggplot(MSE_df_pbmc |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
-                                       y = total_mse,
-                                       color = method
-)) +
-  geom_point() +
-  geom_line() +
-   geom_vline(xintercept = 182, linetype="dashed", 
-                color = "red", size=0.5) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
+  #scale_colour_manual(values = c("#377eb8")) +
   ylab("MSE") +
@@
 #| echo: false
-# num_bins_pbmc <- calculate_effective_x_bins(.data = UMAP_pbmc, x = UMAP1,
-#                                               cell_area = 1) ##23
-
-shape_value_pbmc <- calculate_effective_shape_value(.data = UMAP_pbmc,
-                                                      x = UMAP1, y = UMAP2) ## 0.8772751
-
-# num_bins_pbmc <- find_non_empty_bins(nldr_df = UMAP_pbmc, x = "UMAP1", y = "UMAP2", shape_val = shape_value_pbmc, non_empty_bins = num_non_empty_bins_pbmc)
-num_bins_pbmc <- 16
-
-## To extract bin centroids
-hexbin_data_object_pbmc <- extract_hexbin_centroids(nldr_df = UMAP_pbmc,
-                                                      num_bins = num_bins_pbmc,
-                                                      shape_val = shape_value_pbmc, x = UMAP1, y = UMAP2)
-
-df_bin_centroids_pbmc <- hexbin_data_object_pbmc$hexdf_data
-
-UMAP_pbmc_with_hb_id <- UMAP_pbmc |>
-  dplyr::mutate(hb_id = hexbin_data_object_pbmc$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), UMAP_pbmc_with_hb_id)
-
-## Averaged on high-D
-df_bin_pbmc <- avg_highD_data(.data = df_all_pbmc, column_start_text = "PC")
+#| message: false
+
+## Decide by looking at MSE plot
+num_bins_x_pbmc <- 8
+num_bins_y_pbmc <- 11
+hex_size_pbmc <- 0.1
+
+hb_obj_pbmc <- hex_binning(data = umap_pbmc_scaled |> dplyr::select(-cell_label), x = "UMAP1", 
+                      y = "UMAP2", num_bins_x = num_bins_x_pbmc, 
+                      num_bins_y = num_bins_y_pbmc, x_start = NA, y_start = NA, 
+                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_pbmc, col_start = "UMAP")
+
+## Data set with all possible centroids in the hexagonal grid
+all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_pbmc$centroids))
+
+## Generate all coordinates of hexagons
+hex_grid <- as.data.frame(do.call(cbind, hb_obj_pbmc$hex_poly))
+
+## To obtain the standardise counts within hexbins
+counts_df <- as.data.frame(do.call(cbind, hb_obj_pbmc$std_cts))
+df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
+                                             counts_df = counts_df)
+
+ggplot(data = hex_grid, aes(x = x, y = y)) + 
+  geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
+  geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
+  coord_fixed()
+
+hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
+
+ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+  geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
+  coord_fixed()
+
+umap_data_with_hb_id_pbmc <- as.data.frame(do.call(cbind, hb_obj_pbmc$data_hb_id))
+
+model_object <- fit_highd_model( training_data = training_data_pbmc, 
+                                 nldr_df_with_id = umap_pbmc_scaled |> dplyr::select(-cell_label), 
+                                 x = "UMAP1", y = "UMAP2", 
+                                 num_bins_x = num_bins_x_pbmc, 
+                                 num_bins_y = num_bins_y_pbmc, 
+                                 x_start = NA, y_start = NA, 
+                                 buffer_x = NA, buffer_y = NA, 
+                                 hex_size = hex_size_pbmc,
+                                 is_rm_lwd_hex = FALSE, 
+                                 benchmark_to_rm_lwd_hex = NA, 
+                                 col_start_2d = "UMAP", 
+                                 col_start_highd = "PC_")
+
+df_bin_centroids_pbmc <- model_object$df_bin_centroids
+df_bin_pbmc <- model_object$df_bin
 
 ## Triangulate bin centroids
-tr1_object_pbmc <- triangulate_bin_centroids(df_bin_centroids_pbmc, x, y)
-tr_from_to_df_pbmc <- generate_edge_info(triangular_object = tr1_object_pbmc)
+tr1_object_pbmc <- tri_bin_centroids(df_bin_centroids_pbmc, x = "c_x", y = "c_y")
+tr_from_to_df_pbmc <- gen_edges(tri_object = tr1_object_pbmc)
 
 ## Compute 2D distances
-distance_pbmc <- cal_2d_dist(.data = tr_from_to_df_pbmc)
-
-# distance_plot <- plot_dist(distance_pbmc) +
-#   #ggtitle("(b)" ) +
-#   ylab(expression(d^{(2)})) +
-#   theme(axis.text = element_text(size = 5),
-#         axis.title = element_text(size = 12))
-# 
-# distance_plot
+distance_pbmc <- cal_2d_dist(tr_coord_df = tr_from_to_df_pbmc, 
+                             start_x = "x_from", start_y = "y_from", 
+                             end_x = "x_to", end_y = "y_to", 
+                             select_vars = c("from", "to", "distance"))
+
+distance_plot <- plot_dist(distance_pbmc) +
+  ylab(expression(d^{(2)})) +
+  theme(axis.text = element_text(size = 5),
+        axis.title = element_text(size = 12))
+
+distance_plot
 
 ## To find the benchmark value
-benchmark_pbmc <- find_benchmark_value(.data = distance_pbmc, distance_col = distance)
-benchmark_pbmc <- 2.5
-
-trimesh_removed_pbmc_umap <- remove_long_edges(.data = distance_pbmc, benchmark_value = benchmark_pbmc,
-                                                 triangular_object = tr1_object_pbmc, distance_col = distance)
+benchmark_pbmc <- find_lg_benchmark(distance_edges = distance_pbmc, distance_col = "distance")
+
+ggplot() +
+geom_trimesh(data = df_bin_centroids_pbmc, mapping = aes(x = c_x, y = c_y))
+
+trimesh_removed_pbmc_umap <- vis_rmlg_mesh(distance_edges = distance_pbmc, benchmark_value = benchmark_pbmc, tr_coord_df = tr_from_to_df_pbmc, distance_col = "distance")
 
 trimesh_removed_pbmc_umap <- trimesh_removed_pbmc_umap +
-  geom_point(size = 0.005, colour = "#33a02c") +
-  #ggtitle("(b)") +
   theme_linedraw() +
@@
                                                 if_else(total <= 30, "error 25-30",
-                                                                if_else(total <= 35, "error 30-35", "error greter than 35"))))))))
+                                                        if_else(total <= 35, "error 30-35",
+                                                                if_else(total <= 40, "error 35-40", "error greter than 40"))))))))
 
 plot_list2_pbmc_error <- error_df |>
-    mutate(type = factor(type , levels = c("error 2 or less", "error 2-10", "error 10-15", "error 15-20", "error 20-25", "error 25-30", "error 30-35", "error greter than 35"))) |>
+  mutate(type = factor(type , levels = c("no error", "error 1-15", "error 15-20", "error 20-25", "error 25-30", "error 30-35", "error 35-40", "error greter than 40"))) |>
     ggplot(aes(x = UMAP1,
@@
                                          "#cb181d", "#a50f15", "#99000d")) 
-
 ```
@@
 
-UMAP_pbmc_with_result1_label <- UMAP_pbmc |>
+UMAP_pbmc_with_result1_label <- umap_pbmc |>
   dplyr::mutate(result1_class_label = case_when(
@@
   dplyr::mutate(result1_class_label = if_else((ID %in% c(948, 285, 158, 268, 731, 1152, 1372, 1651, 2010, 150, 2304, 2546, 487, 2176, 2157, 2451, 264, 63, 1234, 947, 1126, 1699, 1232, 1492, 1488, 2040, 2550, 1313, 361, 372, 1515, 1264, 1307, 1726, 517, 2568, 2274, 591, 1703, 1097, 430, 1556, 2133, 1232, 1194, 1888, 1796, 494, 1194)), "cluster 6", result1_class_label))
-
 ```

### 2024-03-06 paper.tex
Make the plan for the methodology

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, boxrule=0pt, sharp corners, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, borderline west={3pt}{0pt}{shadecolor}, interior hidden, sharp corners, frame hidden, boxrule=0pt, breakable]}{\end{tcolorbox}}\fi
 
@@
 
-In this paper, we present a novel method that aims to determine the most
-effective way to represent high-D data by selecting the best method and
-hyperparameter choice. Our approach involves dividing the data set into
-two parts: a training set to construct the model and a test set to
-generate predictive values and residuals. To implement our approach, we
-first use a 2D embedding data set as the initial point. The output of
-our algorithm is a tour that displays the model and original data in the
-high-dimensional space. A flow chart of the proposed algorithm is shown
-in Figure~\ref{fig-meth}. Our algorithm comprises two main phases: (1)
-generate the model in the 2D space, and (2) generate the model in the
-high-D space. These two phases are described in detail in this section
-using UMAP 2D embedding of the S-curve data set, which has seven
-dimensions, including four noise dimensions.
+In this paper, we introduce a novel method to determine the most
+effective NLDR technique and the best hyperparameter choice that
+provides the most useful representation of high-D data. Our approach
+involves dividing the high-dimensional dataset into two parts: a
+training set for constructing the model and a test set for generating
+predictive values and residuals. Our algorithm takes a 2D embedding data
+as the input and generate a tour that displays the high-dimensional
+wireframe to overlay the data. The flow chart of the proposed algorithm
+is shown in Figure~\ref{fig-meth}. The algorithm consists of two main
+phases: (1) generating the model in the 2D space and (2) lifting the
+model into high-D space. The main steps of the algorithm are described
+in detail in this section using UMAP 2D embedding of the S-curve
+dataset. This dataset has seven dimensions, including four noise
+dimensions that were added to the original 3D data.
 
@@
 
-\caption{\label{fig-meth}A flow diagram detailing the steps taken to
-create the low-dimensional manifold in the high dimensional space. There
-are two basic phases, one to generate the model in the 2D space, and
-other to generate the model in the high-D space.}
+\caption{\label{fig-meth}The flow diagram shows the main steps of our
+algorithm. There are two basic phases, one to generate the model in the
+2D space, and other to map the model into the high-D space.}
 
@@
 
+Then, after performing NLDR, to have regular hexagons, 2D embedding data
+are scaled accordingly.
+
 \hypertarget{sec-construct2d}{%
 \subsection{Constructing the 2D model}\label{sec-construct2d}}
+
+\textbf{Step 1: Scaling NLDR data}
+
+\textbf{Step 2: Hexagonating NLDR data}
+
+\textbf{(a) Determine the number of bins along the x axis} (\(b_1\))
+
+\textbf{(b) Determine the number of bins along the y axis} (\(b_2\))
+
+\textbf{(c) Determine the total number of bins} (\(b\))
+
+\textbf{(d) Assign NLDR data to hexagons}
+
+\textbf{Step 3: Obtaining bin centroids or bin means}
+
+\textbf{Step 4: Triangulating bin centroids or bin means}
 
@@
 
-\textbf{(a) Determine the number of bins along the x-axis} (\(b_1\))
+\textbf{(a) Determine the number of bins along the x axis} (\(b_1\))
 
@@
 
+In the process of computing hexagonal bin configurations, the first step
+involves determining the number of bins along the x-axis. This begins
+with the initialization of parameters such as the hexagonal size (\(s\))
+and the buffer along the x-axis. Subsequently, the range of the first 2D
+embedding component (\(r_1\)) is calculated to ensure comprehensive
+coverage of the data. This range is then adjusted by incorporating the
+buffer amount, effectively expanding the boundary to accommodate
+potential outliers or edge cases. The default buffer amount along the
+x-axis is \(\sqrt{3} * s * 1.5\). Next, the horizontal spacing (\(h\))
+between hexagons is computed based on the hexagon size. Finally, the
+number of bins along the x-axis is computed based on the adjusted range
+(\(r_1 + buffer_{x}\)) and the horizontal spacing (\(h\)).
+
 \begin{equation}\protect\hypertarget{eq-equation3}{}{
- \text{A} = \frac{\sqrt{3}}{2}h^2
+ h = \sqrt{3} * s
 }\label{eq-equation3}\end{equation}
 
-To construct regular hexagons, \(A = 1\) (see Figure~\ref{fig-binsize})
-use as the default setting. Then, the diameter (\(h\)) of the regular
-hexagons is calculated (see Equation~\ref{eq-equation4}).
-
 \begin{equation}\protect\hypertarget{eq-equation4}{}{
-  \text{h} = \sqrt{\frac{2}{\sqrt{3}}A}
+ b_1 = \frac{r_1 + buffer_{x}}{h}
 }\label{eq-equation4}\end{equation}
 
-\citet{Carr2013} mentioned about the relationship between the diameter
-(\(h\)) of regular hexagons and the height (\(y\)) of the plotting
-region. According to our algorithm, the height (\(y\)) of the plotting
-region is the the range of 2D embedding component 1 (\(r_1\)) (see
-Equation~\ref{eq-equation6}).
-
-\begin{equation}\protect\hypertarget{eq-equation6}{}{
-h = \frac{r_1}{b_1}
-}\label{eq-equation6}\end{equation}
-
-After rearranging the Equation~\ref{eq-equation6} as
-Equation~\ref{eq-equation5}, \(b_1\) is computed. The \(b_1\) value is
-rounded up to the nearest whole number to have an integer value.
-
-\begin{equation}\protect\hypertarget{eq-equation5}{}{
-b_1 = \frac{r_1}{h}
-}\label{eq-equation5}\end{equation}
-
-\textbf{(b) Determine the shape parameter} (\(s\))
-
-In this step, we determine the shape parameter (\(s\)) for the hexagonal
-bins, which significantly influences their shape and arrangement within
-the grid. The \(s\) in the hexagonal binning algorithm is defined as the
-ratio of the height (\(y\)) to the width (\(x\)) of the plotting region
-as defined in Equation~\ref{eq-equation1}.
-
-\begin{equation}\protect\hypertarget{eq-equation1}{}{
-s = \frac{y}{x}
-}\label{eq-equation1}\end{equation}
-
-The shape parameter (\(s\)) of our algorithm is calculated as the ratio
-of the ranges of 2D embedding components, where \(r_1\) and \(r_2\)
-represent the ranges of 2D embedding components 1 and component 2,
-respectively (see Equation~\ref{eq-equation2}).
-
-\begin{equation}\protect\hypertarget{eq-equation2}{}{
-s = \frac{r_2}{r_1}
-}\label{eq-equation2}\end{equation}
-
-\textbf{(c) Determine the number of bins along the y-axis} (\(b_2\))
+\textbf{(b) Determine the number of bins along the y-axis} (\(b_2\))
 
@@
 number of bins along the x-axis (\(b_1\)) and the shape parameter
-(\(s\)) (see Equation~\ref{eq-equation12}) \citep{Carr2013}.
-
-\begin{equation}\protect\hypertarget{eq-equation12}{}{
-b_2 = 2 * \left(\frac{(b_1 \times s)}{\sqrt(3)} + 1.5001 \right)
-}\label{eq-equation12}\end{equation}
-
-\textbf{(d) Determine the total number of bins} (\(b\))
+(\(s\)) (see Equation~\ref{eq-equation4}) \citep{Carr2013}.
+
+\begin{equation}\protect\hypertarget{eq-equation5}{}{
+  v = 1.5 * s
+}\label{eq-equation5}\end{equation}
+
+\begin{equation}\protect\hypertarget{eq-equation6}{}{
+ b_2 = \frac{r_2 + buffer_{y}}{v}
+}\label{eq-equation6}\end{equation}
+
+\textbf{(c) Determine the total number of bins} (\(b\))
 
@@
 along the x-axis (\(b_1\)) with the number of bins along the y-axis
-(\(b_2\)) (see Equation~\ref{eq-equation13}).
-
-\begin{equation}\protect\hypertarget{eq-equation13}{}{
+(\(b_2\)) (see Equation~\ref{eq-equation5}).
+
+\begin{equation}\protect\hypertarget{eq-equation5}{}{
 b = b_1 \times b_2
-}\label{eq-equation13}\end{equation}
-
-\textbf{Step 2: Obtain bin centroids}
+}\label{eq-equation5}\end{equation}
+
+\textbf{Step 2: Obtain bin centroids or bin means}
 
@@
 which 2D points are assigned to which data set (see
-Figure~\ref{fig-wkhighD} (a)).
+\textbf{?@fig-wkhighD} (a)).
 
@@
 allocated to each hexagons are determined (see video linked in
-Figure~\ref{fig-wkhighD}).
+\textbf{?@fig-wkhighD}).
 
@@
 expansive high-dimensional space (see video linked in
-Figure~\ref{fig-wkhighD}).
+\textbf{?@fig-wkhighD}).
 
@@
 points connected in 2D are also connected in high dimensional space (see
-video linked in Figure~\ref{fig-wkhighD}).
-
-\begin{figure}
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-wkhighD-1.pdf}
-
-}
-
-\caption{\label{fig-wkhighD}How the 2D model lift into high dimensions?
-(a) visualize the points and the hexagonal bin centroids related 2nd and
-15th hexagons, (b) visualization of the edge connected the 2nd and 15th
-hexagons (colored in red) in the triangular mesh. A video of tour
-animation is available at \url{https://youtu.be/ffWFVK1UTIc}.}
-
-\end{figure}
+video linked in \textbf{?@fig-wkhighD}).
 
@@
 to capture the structure of the NLDR data effectively (see
-Figure~\ref{fig-binsize} (a)), while if there are too many bins, it may
+\textbf{?@fig-binsize} (a)), while if there are too many bins, it may
 result in over-fitting the individual points of the NLDR data (see
-Figure~\ref{fig-binsize} (c)). Therefore, it is important to determine
-an appropriate number of bins to build an effective model.
+\textbf{?@fig-binsize} (c)). Therefore, it is important to determine an
+appropriate number of bins to build an effective model.
 
 Furthermore, the total number of bins is determined by the number of
-bins along the x-axis and y-axis (see Equation~\ref{eq-equation13}). To
+bins along the x-axis and y-axis (see \textbf{?@eq-equation13}). To
 calculate the effective total number of bins, the minimum and
@@
 useful in determining the effective number of bins required to construct
-the 2D model. Figure~\ref{fig-diagnosticpltScurve} (b) has created a
-graph that shows how the Mean Squared Error (MSE) changes depending on
-the number of bins assigned in the hexagonal grid for various NLDR
+the 2D model. \textbf{?@fig-diagnosticpltScurve} (b) has created a graph
+that shows how the Mean Squared Error (MSE) changes depending on the
+number of bins assigned in the hexagonal grid for various NLDR
 techniques applied to the S-curve data set. As per the rule of thumb
@@
 in high-D with original data.
-
-\begin{figure}[H]
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-binsize-1.pdf}
-
-}
-
-\caption{\label{fig-binsize}Hexbin plots from different number of bins
-for the \textbf{UMAP} projections of \textbf{S-curve} training data: (a)
-b = 32 (4, 8), s = 1.643542, (b) b = 264 (12, 22), s = 1.643542, and (c)
-b = 840 (21, 40), s = 1.643542. The hexbins are colored based on the
-density of points, with darker colors indicating higher point density
-and yellow color representing lower point density within each bin. What
-is the number of bins that would be effective in representing
-low-dimensional data?}
-
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-diagnosticpltScurve-1.pdf}
-
-}
-
-\caption{\label{fig-diagnosticpltScurve}Goodness of fit statistics from
-UMAP applied to training S-curve dataset. What is the effective number
-of bins in each NLDR technique to create a 2D model? The MSE plot have a
-steep slope at the beginning, indicating that a smaller number of bins
-causes a larger amount of error. Then, the slope gradually declines or
-level off, indicating that a higher number of bins generates a smaller
-error. Using the elbow method, when the total number of bins is set to
-144, the slope of the Mean Squared Error (MSE) plot experiences a sudden
-and noticeable change, resembling an elbow-like shape. This point
-indicates that adding less bins does not enough to capture the data
-structure.}
-
-\end{figure}
 
@@
 substantial, the decision might lean towards retaining low-density
-hexagons. According to Figure~\ref{fig-diagnosticpltScurvelwd}, the
-change in MSE for NLDR techniques is relatively small across different
+hexagons. According to \textbf{?@fig-diagnosticpltScurvelwd}, the change
+in MSE for NLDR techniques is relatively small across different
 benchmark values, except for the PHATE method.
@@
 universal rule for selecting a process. As shown in
-Figure~\ref{fig-diagnosticpltScurvelwd}, the MSE values for PHATE
+\textbf{?@fig-diagnosticpltScurvelwd}, the MSE values for PHATE
 fluctuate, and the first local minimum is at \(0\). Therefore, nearby
@@
 
-\begin{figure}[H]
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-rmlowdenshex-1.pdf}
-
-}
-
-\caption{\label{fig-rmlowdenshex}Goodness of f}
-
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-diagnosticpltScurvelwd-1.pdf}
-
-}
-
-\caption{\label{fig-diagnosticpltScurvelwd}Goodness of fit statistics
-from UMAP applied to training S-curve dataset with different benchmark
-values to remove the low-density heaxgons. What is the effective
-benchmark value to remove the low-density heaxgons? The MSE plot have a
-steep slope at the beginning, indicating that a smaller benchmark causes
-a small amount of error. Then, the slope gradually increases or level
-up, indicating that a higher number of benchmark values generates a
-higher error. Using the reverse elbow method, when the benchmark value
-is set to 0.2, the slope of the Mean Squared Error (MSE) plot
-experiences a sudden and noticeable change, resembling an elbow-like
-shape. This point indicates that higher benchmark values remove the
-necessary bins as well which lead to distruct the 2D structure.}
-
-\end{figure}
-
 \hypertarget{benchmark-value-to-remove-long-edges}{%
@@
 influencing this smoothness is the presence of long edges (see
-Figure~\ref{fig-modelScurvermlgimp}). These long edges occur when a line
+\textbf{?@fig-modelScurvermlgimp}). These long edges occur when a line
 connects points that are distant in the triangular mesh, impacting the
@@
 
-\begin{figure}
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-modelScurverdistcount-1.pdf}
-
-}
-
-\caption{\label{fig-modelScurverdistcount}plots to identify benchmark
-values}
-
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-modelScurvermlgimp-1.pdf}
-
-}
-
-\caption{\label{fig-modelScurvermlgimp}What does the 2D model,
-constructed using UMAP (applied to the S-curve dataset), look like in
-high dimensions with long edges? (\url{https://youtu.be/pSL_G_n7-fM})
-(a) 2D model without removing the long edges, and (b) 2D model with long
-edges coloured by red. The default benchmark value is used to identify
-long edges.}
-
-\end{figure}
-
-\hypertarget{hexagonal-grid-origin}{%
-\subsubsection{Hexagonal grid origin}\label{hexagonal-grid-origin}}
+\hypertarget{starting-point-of-the-hexagonal-grid}{%
+\subsubsection{Starting point of the hexagonal
+grid}\label{starting-point-of-the-hexagonal-grid}}
 
@@
 tessellating the \(xy\) plane over the set (range(\(x\)), range(\(y\)))
-(see Figure~\ref{fig-scurveshifthexgridsexp} (b)). In that case, bin
-centroids are defined as shown in
-Figure~\ref{fig-scurveshifthexgridsexp} (a) with gray colour. Rather
-than sticking to the typical hexagonal grid, introducing a meaningful
-shift in both the x and y directions presents an opportunity for an
-improved 2D model. Therefore, investigating this shift in the hexagonal
-grid is an important parameter to consider.
-
-As shown in Figure~\ref{fig-scurveshifthexgrids}, the shifting
-influences the distribution of points and number of non-empty bins,
-impacting the resulting 2D model. According to
-Figure~\ref{fig-diagnosticpltScurvehexbins}, the 2D model with a total
-of \(144\) bins applied to S-curve UMAP data does not require any
-shifting because the lowest MSE occurs when no shift is introduced.
-
-\begin{figure}
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-scurveshifthexgridsexp-1.pdf}
-
-}
-
-\caption{\label{fig-scurveshifthexgridsexp}(a) Visualization of bin
-centroids and (b) hexagonal grid, showcasing the configuration before
-(colored in gray) and after (colored in light orange) shifting the
-starting point by an identical amount applied in both x and y
-directions, with a shift value of \(0.537285\).}
-
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-scurveshifthexgrids-1.pdf}
-
-}
-
-\caption{\label{fig-scurveshifthexgrids}Hexbin plots shows the
-distribution of points before and after shifting the starting point of
-the hexagonal grid generated for UMAP applied to the training S-curve
-dataset. (a) Hexbin plot before shifting, and (b) Hexbin plot after
-shifting (identical shift applied in both x and y directions, with a
-shift value of \(0.537285\)).}
-
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-diagnosticpltScurvehexbins-1.pdf}
-
-}
-
-\caption{\label{fig-diagnosticpltScurvehexbins}Goodness of fit
-statistics from UMAP applied to training S-curve dataset with different
-shift values. In this illustration, an identical shift is applied in
-both the x and y directions. The Mean Squared Error (MSE) plot exhibits
-a global minimum at \(0\), indicating that no additional shifting is
-necessary for the current hexagonal grid to yield a well-fitted model.}
-
-\end{figure}
+(see \textbf{?@fig-scurveshifthexgridsexp} (b)). In that case, bin
+centroids are defined as shown in \textbf{?@fig-scurveshifthexgridsexp}
+(a) with gray colour. Rather than sticking to the typical hexagonal
+grid, introducing a meaningful shift in both the x and y directions
+presents an opportunity for an improved 2D model. Therefore,
+investigating this shift in the hexagonal grid is an important parameter
+to consider.
+
+As shown in \textbf{?@fig-scurveshifthexgrids}, the shifting influences
+the distribution of points and number of non-empty bins, impacting the
+resulting 2D model. According to
+\textbf{?@fig-diagnosticpltScurvehexbins}, the 2D model with a total of
+\(144\) bins applied to S-curve UMAP data does not require any shifting
+because the lowest MSE occurs when no shift is introduced.
 
@@
 In the 2D representations created by all NLDR techniques, as shown in
-Figure~\ref{fig-nldervis5Gau}, except for PHATE, there are five distinct
+\textbf{?@fig-nldervis5Gau}, except for PHATE, there are five distinct
 clusters. In tSNE, these five clusters are closely located to each other
-(see Figure~\ref{fig-nldervis5Gau} (a)). However, in PHATE, two clusters
+(see \textbf{?@fig-nldervis5Gau} (a)). However, in PHATE, two clusters
 are closely positioned, while the other three are more distant (see
-Figure~\ref{fig-nldervis5Gau} (c)). In PaCMAP, one cluster is at the
+\textbf{?@fig-nldervis5Gau} (c)). In PaCMAP, one cluster is at the
 center, and the remaining four are positioned in four different
-directions (see Figure~\ref{fig-nldervis5Gau} (e)). In TriMAP, two
+directions (see \textbf{?@fig-nldervis5Gau} (e)). In TriMAP, two
 clusters are close, although not as close as in PHATE, and the other
-three are well-separated (see Figure~\ref{fig-nldervis5Gau} (d)). In
-UMAP, all clusters are arranged in a parallel manner, with three in one
-line and the other two in a separate line (see
-Figure~\ref{fig-nldervis5Gau} (b)).
+three are well-separated (see \textbf{?@fig-nldervis5Gau} (d)). In UMAP,
+all clusters are arranged in a parallel manner, with three in one line
+and the other two in a separate line (see \textbf{?@fig-nldervis5Gau}
+(b)).
 
@@
 effectively preserving both local and global structures (see video link
-of Figure~\ref{fig-modelfiveGau} (a)). UMAP, on the other hand, also
+of \textbf{?@fig-modelfiveGau} (a)). UMAP, on the other hand, also
 presents five distinct clusters but with a more flattened surface
-appearance (see video link of Figure~\ref{fig-modelfiveGau} (b)). The
+appearance (see video link of \textbf{?@fig-modelfiveGau} (b)). The
 PHATE model shows five separated clusters resembling triangles or
 partial triangles but struggles to capture local structures (see video
-link of Figure~\ref{fig-modelfiveGau} (c)). TriMAP and PaCMAP both
-display five well-separated clusters with flat surfaces, each capturing
+link of \textbf{?@fig-modelfiveGau} (c)). TriMAP and PaCMAP both display
+five well-separated clusters with flat surfaces, each capturing
 within-cluster variation to varying extents. Despite the various 2D
 representations, all NLDR techniques preserve the global structure (see
-video link of Figure~\ref{fig-modelfiveGau} (d), (e) respectively).
+video link of \textbf{?@fig-modelfiveGau} (d), (e) respectively).
 However, tSNE, effectively capture both local and global structures, as
-indicated by lower AIC values in Figure~\ref{fig-diagnosticpltGau} (a).
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervis5Gau-1.pdf}
-
-}
-
-\caption{\label{fig-nldervis5Gau}2D layouts from different NLDR
-techniques applied the same data: (a) tSNE (perplexity = 61), (b) UMAP
-(n\_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n\_inliers = 5,
-n\_outliers = 4, n\_random = 3), and (e) PaCMAP (n\_neighbors = 10, init
-= random, MN\_ratio = 0.9, FP\_ratio = 2). Is there a best
-representation of the original data or are they all providing equivalent
-information?}
-
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-modelfiveGau-1.pdf}
-
-}
-
-\caption{\label{fig-modelfiveGau}Is there a best model to represent the
-original data in 2D space or are they all providing equivalent
-information?, (a) Model in the 2D space with tSNE
-(\url{https://youtu.be/x5VPB-wOLm4}), (b) Model in the 2D space with
-UMAP (\url{https://youtu.be/Cm1dW6iLyHQ}), (c) Model in the 2D space
-with PHATE (\url{https://youtu.be/HB0Y3ilz6sI}), (d) Model in the 2D
-space with TriMAP (\url{https://youtu.be/iSVkXIrOrSA}), and (e) Model in
-the 2D space with PaCMAP (\url{https://youtu.be/O2QBctsc4uM}).}
-
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticpltGau-1.pdf}
-
-}
-
-\caption{\label{fig-diagnosticpltGau}Goodness of fit statistics from
-different NLDR techniques applied to training five spherical Gaussian
-cluster dataset. What is the best NLDR technique to represent the
-original data in 2D?}
-
-\end{figure}
+indicated by lower AIC values in \textbf{?@fig-diagnosticpltGau} (a).
 
@@
 
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-56-1.pdf}
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-55-1.pdf}
+
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-55-2.pdf}
+
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-55-3.pdf}
+
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-55-4.pdf}
 
@@
 space with UMAP (\url{https://youtu.be/CNXjWZMyh_M})}
-
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-diagnosticpltPBMC2-1.pdf}
-
-}
-
-\caption{\label{fig-diagnosticpltPBMC2}Goodness of fit statistics from
-UMAP applied to training S-curve dataset. What is the effective number
-of bins in each NLDR technique to create a 2D model? The MSE plot have a
-steep slope at the beginning, indicating that a smaller number of bins
-causes a larger amount of error. Then, the slope gradually declines or
-level off, indicating that a higher number of bins generates a smaller
-error. Using the elbow method, when the total number of bins is set to
-144, the slope of the Mean Squared Error (MSE) plot experiences a sudden
-and noticeable change, resembling an elbow-like shape. This point
-indicates that adding less bins does not enough to capture the data
-structure.}
-
-\end{figure}
-
-\begin{figure}[h]
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervisPBMCUMAP2-1.pdf}
-
-}
-
-\caption{\label{fig-nldervisPBMCUMAP2}(a) 2D layout from UMAP
-(n\_neighbors = 15, min\_dist = 0.99, metric = cosine) applied for the
-PBMC3k dataset. Is this a best representation of the original data?, (b)
-Model in the 2D space with UMAP (\textless\textgreater)}
 
@@
 
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-61-1.pdf}
+
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-61-2.pdf}
+
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-61-3.pdf}
+
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-61-4.pdf}
+
 \begin{figure}[h]
@@
 
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-71-1.pdf}
-
-\hypertarget{harps-gto-sample-dataset}{%
-\subsection{HARPS GTO sample dataset}\label{harps-gto-sample-dataset}}
-
-\citet{Anders2018}
-
-\begin{itemize}
-\item
-  In 2D, there are two well-separated clusters, but the clusters are
-  really close
-\item
-  One cluster have a sparse data space within the cluster
-\item
-  There is a line in right corner of the structure that is not connected
-  any other trimesh points. This line create a long edge in high-D
-  space.
-\item
-  In high-D view, the clusters having some non-linearity patter, which
-  is also not captured in tSNE. Therefore, this is not a good choice.
-  Also, the two clusters are really near to each other and can't see any
-  separation. Also continuity of the structure. Long edges are presence
-  also.
-\item
-  The parameter is, in a sense, a guess about the number of close
-  neighbors each point has.
-\item
-  Num bins along the x-axis = 43, shape parameter = 0.4466309
-\end{itemize}
-
-\begin{table}[H]
-
-\caption{errors for different parameter combinations}
-\centering
-\begin{tabular}[t]{r|r}
-\hline
-\textbf{perplexity} & \textbf{error (x 100)}\\
-\hline
-\cellcolor{gray!6}{23} & \cellcolor{gray!6}{0.02}\\
-\hline
-91 & 0.04\\
-\hline
-\textcolor{red}{\textbf{\cellcolor{gray!6}{40}}} & \textcolor{red}{\textbf{\cellcolor{gray!6}{1.05}}}\\
-\hline
-\end{tabular}
-\end{table}
-
-\begin{figure}
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-diagnosticpltHARPS-1.pdf}
-
-}
-
-\caption{\label{fig-diagnosticpltHARPS}Goodness of fit statistics from
-UMAP applied to training S-curve dataset. What is the effective number
-of bins in each NLDR technique to create a 2D model? The MSE plot have a
-steep slope at the beginning, indicating that a smaller number of bins
-causes a larger amount of error. Then, the slope gradually declines or
-level off, indicating that a higher number of bins generates a smaller
-error. Using the elbow method, when the total number of bins is set to
-144, the slope of the Mean Squared Error (MSE) plot experiences a sudden
-and noticeable change, resembling an elbow-like shape. This point
-indicates that adding less bins does not enough to capture the data
-structure.}
-
-\end{figure}
-
-\begin{figure}[h]
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-nldervisharpstSNE-1.pdf}
-
-}
-
-\caption{\label{fig-nldervisharpstSNE}(a) 2D layout from tSNE
-(perplexity = 40) applied for the HARPS GTO dataset. Is this a best
-representation of the original data?, (b) Model in the 2D space with
-tSNE (\url{https://youtu.be/jxo_rPyGv9k})}
-
-\end{figure}
-
-\begin{itemize}
-\item
-  5 clusters shows in 2D
-\item
-  3 clusters which are really close shows in high\_d data with the model
-\end{itemize}
-
-\begin{figure}
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-diagnosticpltHARPS2-1.pdf}
-
-}
-
-\caption{\label{fig-diagnosticpltHARPS2}Goodness of fit statistics from
-UMAP applied to training S-curve dataset. What is the effective number
-of bins in each NLDR technique to create a 2D model? The MSE plot have a
-steep slope at the beginning, indicating that a smaller number of bins
-causes a larger amount of error. Then, the slope gradually declines or
-level off, indicating that a higher number of bins generates a smaller
-error. Using the elbow method, when the total number of bins is set to
-144, the slope of the Mean Squared Error (MSE) plot experiences a sudden
-and noticeable change, resembling an elbow-like shape. This point
-indicates that adding less bins does not enough to capture the data
-structure.}
-
-\end{figure}
-
-\begin{figure}[h]
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-nldervisharpstSNE2-1.pdf}
-
-}
-
-\caption{\label{fig-nldervisharpstSNE2}(a) 2D layout from tSNE
-(perplexity = 23) applied for the HARPS GTO dataset. Is this a best
-representation of the original data?, (b) Model in the 2D space with
-tSNE (\url{https://youtu.be/VO_4SJEoAUA})}
-
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-diagnosticpltHARPS3-1.pdf}
-
-}
-
-\caption{\label{fig-diagnosticpltHARPS3}Goodness of fit statistics from
-UMAP applied to training S-curve dataset. What is the effective number
-of bins in each NLDR technique to create a 2D model? The MSE plot have a
-steep slope at the beginning, indicating that a smaller number of bins
-causes a larger amount of error. Then, the slope gradually declines or
-level off, indicating that a higher number of bins generates a smaller
-error. Using the elbow method, when the total number of bins is set to
-144, the slope of the Mean Squared Error (MSE) plot experiences a sudden
-and noticeable change, resembling an elbow-like shape. This point
-indicates that adding less bins does not enough to capture the data
-structure.}
-
-\end{figure}
-
-\begin{figure}[h]
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-nldervisharpstSNE3-1.pdf}
-
-}
-
-\caption{\label{fig-nldervisharpstSNE3}(a) 2D layout from tSNE
-(perplexity = 91) applied for the HARPS GTO dataset. Is this a best
-representation of the original data?, (b) Model in the 2D space with
-tSNE (\url{https://youtu.be/vLvRfcogft8})}
-
-\end{figure}
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-64-1.pdf}
 

### 2024-03-06 paper.qmd
Added mathemtical expression for triangulation

@@
 
+The mathematical expression for the hexagonal bin centers can be defined based on the hexagonal grid coordinates.
+Let's denote the hexagonal grid coordinates as \( (q, r) \), where \( q \) and \( r \) are integers representing the coordinates of the hexagon in the grid.
+
+For a regular hexagonal grid, the coordinates of the hexagon centers can be calculated using the following formulas:
+<!--
+\[
+x = \frac{3}{2} \times q \times s
+\]
+\[
+y = \frac{\sqrt{3}}{2} \times q \times s + \sqrt{3} \times r \times s
+\]-->
+
+Where:
+- \( s \) is the side length of the hexagon.
+- \( q \) represents the column index of the hexagon in the grid.
+- \( r \) represents the row index of the hexagon in the grid.
+- \( x \) and \( y \) are the Cartesian coordinates of the center of the hexagon.
+
+This formula calculates the \( x \) and \( y \) coordinates of the center of each hexagon based on its grid coordinates \( q \) and \( r \), allowing you to determine the position of each hexagon in the plane.
+
 **Step 4: Triangulating bin centroids or bin means**
+
+Delaunay triangulation is a geometric algorithm used to create a triangular mesh from a set of points in a plane.
+The key principle behind Delaunay triangulation is to connect points in such a way that no point falls within the circumcircle of any triangle formed by the points.
+
+Mathematically, given a set of points \( P \) in a plane, the Delaunay triangulation constructs a set of triangles such that:
+
+1. Each triangle is formed by connecting three points from \( P \).
+2. No point in \( P \) is inside the circumcircle of any triangle in the triangulation.
+
+Formally, the Delaunay triangulation can be defined as follows:
+
+Let \( P = \{p_1, p_2, ..., p_n\} \) be a set of \( n \) points in the plane.
+The Delaunay triangulation of \( P \), denoted as \( DT(P) \), is a triangulation of the convex hull of \( P \) such that the circumcircle of every triangle in the triangulation contains no other points from \( P \).
 

### 2024-03-06 paper.tex
Added mathemtical expression for triangulation

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, borderline west={3pt}{0pt}{shadecolor}, interior hidden, sharp corners, frame hidden, boxrule=0pt, breakable]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, enhanced, breakable, borderline west={3pt}{0pt}{shadecolor}, interior hidden, frame hidden, boxrule=0pt]}{\end{tcolorbox}}\fi
 
@@
 
+The mathematical expression for the hexagonal bin centers can be defined
+based on the hexagonal grid coordinates. Let's denote the hexagonal grid
+coordinates as ( (q, r) ), where ( q ) and ( r ) are integers
+representing the coordinates of the hexagon in the grid.
+
+For a regular hexagonal grid, the coordinates of the hexagon centers can
+be calculated using the following formulas:
+
+Where: - ( s ) is the side length of the hexagon. - ( q ) represents the
+column index of the hexagon in the grid. - ( r ) represents the row
+index of the hexagon in the grid. - ( x ) and ( y ) are the Cartesian
+coordinates of the center of the hexagon.
+
+This formula calculates the ( x ) and ( y ) coordinates of the center of
+each hexagon based on its grid coordinates ( q ) and ( r ), allowing you
+to determine the position of each hexagon in the plane.
+
 \textbf{Step 4: Triangulating bin centroids or bin means}
+
+Delaunay triangulation is a geometric algorithm used to create a
+triangular mesh from a set of points in a plane. The key principle
+behind Delaunay triangulation is to connect points in such a way that no
+point falls within the circumcircle of any triangle formed by the
+points.
+
+Mathematically, given a set of points ( P ) in a plane, the Delaunay
+triangulation constructs a set of triangles such that:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Each triangle is formed by connecting three points from ( P ).
+\item
+  No point in ( P ) is inside the circumcircle of any triangle in the
+  triangulation.
+\end{enumerate}
+
+Formally, the Delaunay triangulation can be defined as follows:
+
+Let ( P = \{p\_1, p\_2, \ldots, p\_n\} ) be a set of ( n ) points in the
+plane. The Delaunay triangulation of ( P ), denoted as ( DT(P) ), is a
+triangulation of the convex hull of ( P ) such that the circumcircle of
+every triangle in the triangulation contains no other points from ( P ).
 

### 2024-03-11 appendix.qmd
Added Handwritten digit: 1 dataset

@@
 -->
+
+<!--In this context, "prediction" denotes the 2D embedding generated for the NLDR technique.
+The methodology encompasses identifying the nearest averaged high-D points in the high-D space for the test data, by computing high-D Euclidean distances.
+As the averaged high-D point corresponds to the lifting of the 2D model, determining its nearest counterpart allows us to map its hexagonal bin centroid coordinates.
+Consequently, these centroid coordinates serve as the assigned low-dimensional predicted values for the test data in the 2D space.
+
+Some NLDR techniques, such as tSNE, often lack a direct prediction, making our approach valuable for generating predicted values in the absence of such functionalities.
+
+The prediction approach involves employing the K-nearest neighbors (KNN) algorithm to identify the nearest hexagonal bin centroid in the 2D space.
+Subsequently, the coordinates of this centroid are assigned as the low-dimensional predicted values for the test data in 2D space.
+
+The prediction approach involves employing the K-nearest neighbors (KNN) algorithm to identify the nearest high-D model point (averaged high-D point) for the new high-D data point.
+Also, the coordinates of the high-D model point serve as the predicted 2D embedings for the new high-D data point.
+It is noteworthy that traditional NLDR methods, such as tSNE, often lack a direct predict function, making our approach valuable for generating predicted values in the absence of such functionalities.
+-->
+
+<!--
+### Model interpretation
+
+The visualizations of NLDR techniques for the S-curve dataset shows different representations in 2D, as showcased in Figure @fig-nldervis.
+The question at hand is determining the best NLDR technique to represent the S-curve data.
+Examining the 2D layouts in @fig-nldervis suggests that TriMAP or PaCMAP may be best representations.
+However, the challenge is to decide the best representation with considerable evidence.
+
+After considering the models constructed from the algorithm for each NLDR techniques and visualizing them in high-D space with the original data, shows that all the NLDR techniques capture the non-linearity of the S-curve (see videos of the models linked in @fig-modelScurve).
+But thorough examination of each model in high-D space shows some surprising findings.
+The model of PHATE is flattened and lack of capturing the width of S-curve (see videos of the model with PHATE linked in @fig-modelScurve).
+This nature of PHATE is also evicted by the considerably higher AIC and MSE values for PHATE as shown in @fig-diagnosticpltScurve.
+Furthermore, there are two edges that are not connected to any points in the model of TriMAP (see see videos of the model with TriMAP linked in @fig-modelScurve).
+This model shows how the patterns in 2D representation of TriMAP capture by the model (see @fig-nldervis (d)).
+As shown in @fig-nldervis (b), UMAP have four clusters in a curvilinear structure.
+However, the model reveals that even though UMAP shows a curvilinear in 2D, UMAP actually capture the structure of the S-curve without any disjoint in the model (see see videos of the model with UMAP linked in @fig-modelScurve).
+In the 2D layouts of tSNE and PaCMAP (see @fig-nldervis (a), (e)), there are some sparse areas that are captured by the models effectively (see see videos of the model with tSNE and PaCMAP linked in @fig-modelScurve).
+-->
+
+<!--Different NLDR techniques perform differently on the S-curve data and tSNE provides the best representation in 2D according to AIC and MSE values (see @fig-diagnosticpltScurve). -->
+
+<!--### Model selection {#sec-selection}
+
+In the process of selecting the most suitable NLDR technique to represent high-D data, the Akaike Information Criterion (AIC) stands as a crucial tool for insightful comparisons.
+Once the effective number of bins has been determined for each method, AIC values can be compared to identify the best NLDR technique with the lowest AIC value.
+For example, as shown in @fig-diagnosticpltScurve (a), TriMAP is the most suitable NLDR technique to represent the S-curve data (see table 1).
+-->
+
+<!--By visualizing the models with original data in high-D space, we are able to understand how various techniques capture the clustering data structure.
+The tSNE model exhibits five well-separated clusters, effectively preserving both local and global structures.
+UMAP, on the other hand, also presents five distinct clusters but with a more flattened surface appearance.
+The PHATE model shows five clusters resembling triangles but struggles to capture local structures.
+TriMAP and PaCMAP both display five well-separated clusters with a flat surface appearance, capturing within-cluster variation to varying extents.
+As we can see, even though all NLDR techniques have different representations in 2D, all the NLDR techniques preserve the global structure, but some NLDR techniques fail to capture the local structure.
+According to the visualizations, we can see that tSNE preserve both local and global structure than other NLDR techniques that is also evident by the AIC plot shown in Figure 4.-->
+
+<!--We *strongly* recommend viewing the linked videos for each study while reading.
+Links to the videos are available in the figures for each example.
+The videos show the visual appearance of the **langevitour** interface with low-dimensional view and how we can interact with the tour via the controls.-->
+
+<!--In all NLDR techniques, except for PHATE, there are five well separated clusters.
+In tSNE, the five clusters are close to each other.
+However, in PHATE, two clusters are really close to each other and other three clusters are far away from them.
+In PaCMAP, one cluster is at the center and other four positioned in four directions.
+In TriMAP also, there are two clusters close to each other and other three are well-separated from each other.
+In UMAP, four clusters are positioned parallel to each and remaining one far away from them.-->
+
+<!--The model with tSNE shows five well-separated clusters. Also, within cluster variation is captured by giving proof that with clustering tSNE preserve local and global structures of high-D data.
+
+The model with UMAP also shows five well-separated clusters. But now, the clusters are more flat surfaces.
+
+The model with PHATE also shows five well-separated clusters. The five clusters are more like triangles and couldn't able to capture local structure at all.
+
+In the model with TriMAP shows five well-separated clusters. Also, within cluster variation is captured but much like flat surfaces.
+
+In the model with PaCMAP shows five well-separated clusters. Also, within cluster variation is captured but much like flat surfaces.-->

### 2024-03-11 paper.qmd
Added Handwritten digit: 1 dataset

@@
 
-#source("quollr_code.R", local = TRUE)
 source("nldr_code.R", local = TRUE)
@@
 In this paper, we introduce a novel method to determine the most effective NLDR technique and the best hyperparameter choice that provides the most useful representation of high-D data.
-Our approach involves dividing the high-dimensional dataset into two parts: a training set for constructing the model and a test set for generating predictive values and residuals.
-Our algorithm takes a 2D embedding data as the input and generate a tour that displays the high-dimensional wireframe to overlay the data.
+Our approach involves dividing the high-D dataset into two parts: a training set for constructing the model and a test set for generating predictive values and residuals.
+Our algorithm takes a 2D embedding data as the input and generate a tour that displays the high-D wireframe to overlay the data.
 The flow chart of the proposed algorithm is shown in @fig-meth.
@@
 
-In order to reduce the computational complexity associated with performing NLDR techniques to high-D data, and as a initial step of noise reduction of high-D data, PCA [@article67, @article68, @article69] is applied as a preprocessing step.
-Subsequently, the identified principal components, representing directions of maximum variance, are used to construct the model.
-
-Then, after performing NLDR, to have regular hexagons, 2D embedding data are scaled accordingly.
+To reduce computational complexity when applying NLDR techniques to high-D data and to reduce noise presence, PCA [@article67, @article68, @article69] is used as a preprocessing step.
+PCA involves identifying principal components that maximize variance.
+These components are then used as the high-D data for the algorithm.
 
@@
 **Step 1: Scaling NLDR data**
+
+First, we prepare the 2D embedding data to fit within the bounds required for regular hexagonal binning.
+To achieve this, we implement two key scaling steps.
+Scale the first 2D embedding component to range between $0$ and $1$, ensuring that all data points fall within this normalized interval.
+Secondly, we scale the second 2D embedding component to range between $0$ and $y_{max}$ (see @eq-equation3).
+
+$$
+ ar = \frac{r_2}{r_1}
+$$ {#eq-equation1}
+
+$$
+ hr = \frac{hb}{wb}
+$$ {#eq-equation2}
+
+$$
+ y_{max} = \text{ceiling}(\frac{ar}{hr}) * hr
+$$ {#eq-equation3}
 
@@
 
-<!--In this context, "prediction" denotes the 2D embedding generated for the NLDR technique.
-The methodology encompasses identifying the nearest averaged high-D points in the high-D space for the test data, by computing high-D Euclidean distances.
-As the averaged high-D point corresponds to the lifting of the 2D model, determining its nearest counterpart allows us to map its hexagonal bin centroid coordinates.
-Consequently, these centroid coordinates serve as the assigned low-dimensional predicted values for the test data in the 2D space.
-
-Some NLDR techniques, such as tSNE, often lack a direct prediction, making our approach valuable for generating predicted values in the absence of such functionalities.
-
-The prediction approach involves employing the K-nearest neighbors (KNN) algorithm to identify the nearest hexagonal bin centroid in the 2D space.
-Subsequently, the coordinates of this centroid are assigned as the low-dimensional predicted values for the test data in 2D space.
-
-The prediction approach involves employing the K-nearest neighbors (KNN) algorithm to identify the nearest high-D model point (averaged high-D point) for the new high-D data point.
-Also, the coordinates of the high-D model point serve as the predicted 2D embedings for the new high-D data point.
-It is noteworthy that traditional NLDR methods, such as tSNE, often lack a direct predict function, making our approach valuable for generating predicted values in the absence of such functionalities.
--->
-
 The prediction approach involves employing the K-nearest neighbors (KNN) algorithm.
 In this method, which operates as an unsupervised classification problem, the nearest high-D model point (averaged high-D point) is identified for the new high-D data point.
-Once the nearest high-D model point is identified, its coordinates are used as the predicted 2D embeddings for the new high-D data point.
-This step is particularly valuable due to the limitations of certain NLDR techniques, like tSNE, which don't provide a straightforward method for predicting 2D embeddings.
-As a result, our approach presents a solution that can generate predicted 2D embeddings, irrespective of the NLDR technique employed, effectively addressing this gap in functionality.
+Once the nearest high-D model point is identified, its coordinates are used as the predicted 2D embedding for the new high-D data point.
+This step is particularly valuable due to the limitations of certain NLDR techniques, like tSNE, which don't provide a straightforward method for predicting 2D embedding.
+As a result, our approach presents a solution that can generate predicted 2D embedding, irrespective of the NLDR technique employed, effectively addressing this gap in functionality.
 
@@
 These metrics, MSE and AIC, collectively offer valuable insights into the model's predictive performance, considering both accuracy and complexity in the predictions.
-
-<!--
-### Model interpretation
-
-The visualizations of NLDR techniques for the S-curve dataset shows different representations in 2D, as showcased in Figure @fig-nldervis.
-The question at hand is determining the best NLDR technique to represent the S-curve data.
-Examining the 2D layouts in @fig-nldervis suggests that TriMAP or PaCMAP may be best representations.
-However, the challenge is to decide the best representation with considerable evidence.
-
-After considering the models constructed from the algorithm for each NLDR techniques and visualizing them in high-D space with the original data, shows that all the NLDR techniques capture the non-linearity of the S-curve (see videos of the models linked in @fig-modelScurve).
-But thorough examination of each model in high-D space shows some surprising findings.
-The model of PHATE is flattened and lack of capturing the width of S-curve (see videos of the model with PHATE linked in @fig-modelScurve).
-This nature of PHATE is also evicted by the considerably higher AIC and MSE values for PHATE as shown in @fig-diagnosticpltScurve.
-Furthermore, there are two edges that are not connected to any points in the model of TriMAP (see see videos of the model with TriMAP linked in @fig-modelScurve).
-This model shows how the patterns in 2D representation of TriMAP capture by the model (see @fig-nldervis (d)).
-As shown in @fig-nldervis (b), UMAP have four clusters in a curvilinear structure.
-However, the model reveals that even though UMAP shows a curvilinear in 2D, UMAP actually capture the structure of the S-curve without any disjoint in the model (see see videos of the model with UMAP linked in @fig-modelScurve).
-In the 2D layouts of tSNE and PaCMAP (see @fig-nldervis (a), (e)), there are some sparse areas that are captured by the models effectively (see see videos of the model with tSNE and PaCMAP linked in @fig-modelScurve).
--->
-
-<!--Different NLDR techniques perform differently on the S-curve data and tSNE provides the best representation in 2D according to AIC and MSE values (see @fig-diagnosticpltScurve). -->
-
-<!--### Model selection {#sec-selection}
-
-In the process of selecting the most suitable NLDR technique to represent high-D data, the Akaike Information Criterion (AIC) stands as a crucial tool for insightful comparisons.
-Once the effective number of bins has been determined for each method, AIC values can be compared to identify the best NLDR technique with the lowest AIC value.
-For example, as shown in @fig-diagnosticpltScurve (a), TriMAP is the most suitable NLDR technique to represent the S-curve data (see table 1).
--->
 
@@
 
-<!--By visualizing the models with original data in high-D space, we are able to understand how various techniques capture the clustering data structure.
-The tSNE model exhibits five well-separated clusters, effectively preserving both local and global structures.
-UMAP, on the other hand, also presents five distinct clusters but with a more flattened surface appearance.
-The PHATE model shows five clusters resembling triangles but struggles to capture local structures.
-TriMAP and PaCMAP both display five well-separated clusters with a flat surface appearance, capturing within-cluster variation to varying extents.
-As we can see, even though all NLDR techniques have different representations in 2D, all the NLDR techniques preserve the global structure, but some NLDR techniques fail to capture the local structure.
-According to the visualizations, we can see that tSNE preserve both local and global structure than other NLDR techniques that is also evident by the AIC plot shown in Figure 4.-->
-
-<!--We *strongly* recommend viewing the linked videos for each study while reading.
-Links to the videos are available in the figures for each example.
-The videos show the visual appearance of the **langevitour** interface with low-dimensional view and how we can interact with the tour via the controls.-->
-
-<!--In all NLDR techniques, except for PHATE, there are five well separated clusters.
-In tSNE, the five clusters are close to each other.
-However, in PHATE, two clusters are really close to each other and other three clusters are far away from them.
-In PaCMAP, one cluster is at the center and other four positioned in four directions.
-In TriMAP also, there are two clusters close to each other and other three are well-separated from each other.
-In UMAP, four clusters are positioned parallel to each and remaining one far away from them.-->
-
-<!--The model with tSNE shows five well-separated clusters. Also, within cluster variation is captured by giving proof that with clustering tSNE preserve local and global structures of high-D data.
-
-The model with UMAP also shows five well-separated clusters. But now, the clusters are more flat surfaces.
-
-The model with PHATE also shows five well-separated clusters. The five clusters are more like triangles and couldn't able to capture local structure at all.
-
-In the model with TriMAP shows five well-separated clusters. Also, within cluster variation is captured but much like flat surfaces.
-
-In the model with PaCMAP shows five well-separated clusters. Also, within cluster variation is captured but much like flat surfaces.-->
-
-<!--
-```{r}
-#| warning: false
-#| echo: false
-#| eval: false
+<!-- five Gaussian clusters with different NLDR techniques-->
+```{r}
+#| warning: false
+#| echo: false
 
@@
 df_2 <- read_rds("data/five_gau_clusters/data_five_gau.rds")
-training_data_1 <- read_rds("data/five_gau_clusters/data_five_gau_training.rds")
-test_1 <- read_rds("data/five_gau_clusters/data_five_gau_test.rds")
+training_data_gau <- read_rds("data/five_gau_clusters/data_five_gau_training.rds")
+test_data_gau <- read_rds("data/five_gau_clusters/data_five_gau_test.rds")
 
@@
 ```{r}
-#| eval: false
 #| echo: false
@@
 
+<!-- To find the number of bins should assign for the model in tSNE-->
+```{r}
+#| warning: false
+#| echo: false
+#| message: false
+
+tsne_gau_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tSNE_data_gau, 
+                                    x = "tSNE1", y = "tSNE2"))) |>
+  dplyr::rename(c("tSNE1" = "scaled_tSNE1", 
+                  "tSNE2" = "scaled_tSNE2")) |>
+  dplyr::mutate(ID = 1:NROW(tSNE_data_gau))
+
+## tSNE
+hex_size_vec <- seq(0.02, 2, by = 0.01)
+
+vec <- stats::setNames(rep("", 6), c("num_bins", "aic", "mse", "num_bins_x", "num_bins_y", "hex_size"))  ## Define column names
+
+mse_df_gau <- dplyr::bind_rows(vec)[0, ]
+mse_df_gau <- mse_df_gau |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(hex_size_vec)) {
+
+  num_bin_list <- calc_bins(data = tsne_gau_scaled, 
+            x = "tSNE1", y = "tSNE2", 
+            hex_size = hex_size_vec[i], buffer_x = NA, buffer_y = NA)
+
+  num_bins_x <- num_bin_list$num_x
+  num_bins_y <- num_bin_list$num_y
+
+  model_object <- fit_highd_model( training_data = training_data_gau, 
+                                   nldr_df_with_id = tsne_gau_scaled, 
+                                   x = "tSNE1", y = "tSNE2", 
+                                   num_bins_x = num_bins_x, 
+                                   num_bins_y = num_bins_y, 
+                                   x_start = NA, y_start = NA, 
+                                   buffer_x = NA, buffer_y = NA, 
+                                   hex_size = hex_size_vec[i],
+                                   is_rm_lwd_hex = FALSE, 
+                                   benchmark_to_rm_lwd_hex = NA, 
+                                   col_start_2d = "tSNE", 
+                                   col_start_highd = "x")
+
+  centroid_df_training <- model_object$df_bin_centroids
+  avg_df_training <- model_object$df_bin
+
+  pred_emb_list <- predict_emb(test_data = training_data_gau, 
+                                  df_bin_centroids = centroid_df_training, 
+                                  df_bin = avg_df_training, type_NLDR = "tSNE")
+
+  pred_df_training <- as.data.frame(do.call(cbind, pred_emb_list))
+
+  eval_list <- gen_summary(test_data = training_data_gau, 
+                                  prediction_df = pred_df_training, 
+                                  df_bin = avg_df_training, col_start = "x")
+
+  mse_df_gau <- mse_df_gau |>
+    tibble::add_row(num_bins = num_bins_x * num_bins_y,
+                    aic = eval_list$aic,
+                    mse = eval_list$mse,
+                    num_bins_x = num_bins_x,
+                    num_bins_y = num_bins_y,
+                    hex_size = hex_size_vec[i])
+
+}
+
+## If same total number of bins occurred only select ones with minimum error
+### Obtain duplicate bins
+dupli_bins <- mse_df_gau |> 
+  dplyr::count(num_bins) |> 
+  dplyr::filter(n > 1) |> 
+  dplyr::pull(num_bins)
+
+### Group split by duplicated bins
+duplicate_df_list <- mse_df_gau |>
+  dplyr::filter(num_bins %in% dupli_bins) |>
+  dplyr::arrange(num_bins) |>
+  dplyr::group_split(num_bins)
+
+### Obtain one row from duplicates which have lowest error and hexsize
+duplicate_df <- data.frame(matrix(nrow = 0, ncol = 0))
+
+for (i in 1:length(duplicate_df_list)) {
+
+  dd <- duplicate_df_list[[i]] |>
+    dplyr::filter(mse == min(duplicate_df_list[[i]]$mse)) |>
+    dplyr::filter(hex_size == min(duplicate_df_list[[i]]$hex_size))
+
+  duplicate_df <- dplyr::bind_rows(duplicate_df, dd)
+
+}
+
+### Obtain the mse_df with not duplicated bins
+not_dupli_df <- mse_df_gau |>
+  dplyr::filter(!(num_bins %in% dupli_bins))
+
+### Combine duplicated and not duplicated(corrected) bins dfs
+mse_df_gau <- dplyr::bind_rows(not_dupli_df, duplicate_df)
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## To draw with AIC
+aic_plot_gau <- ggplot(mse_df_gau, aes(x = num_bins, y = aic
+)) +
+  geom_point(size = 0.8) +
+  geom_line() +
+  geom_vline(xintercept = 42, linetype="solid",
+                color = "red", size=0.8, alpha = 0.5) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  #scale_colour_manual(values = c("#377eb8")) +
+  ylab("AIC") +
+  xlab("total number of bins")
+## Effective number of bins along x-axis
+
+mse_plot_gau <- ggplot(mse_df_gau, aes(x = num_bins,
+                                       y = mse
+)) +
+  geom_point(size = 0.8) +
+  geom_line() +
+   geom_vline(xintercept = 42, linetype="solid",
+                color = "red", size=0.8, alpha = 0.5) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  #scale_colour_manual(values = c("#377eb8")) +
+  ylab("MSE") +
+  xlab("total number of bins")
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "Goodness of fit statistics from different NLDR techniques applied to training five spherical Gaussian cluster dataset.
+What is the best NLDR technique to represent the original data in 2D?"
+#| label: fig-diagnosticpltGau
+#| out-width: 100%
+
+aic_plot_gau + mse_plot_gau +
+  plot_layout(nrow = 2) +
+  plot_annotation(tag_levels = 'a') +
+  #plot_layout(guides='collect', nrow = 2) & legend.position='bottom', 
+  theme(plot.tag = element_text(size = 8))
+```
+
+```{r}
+#| warning: false
+#| echo: false
+#| message: false
+
+## Decide by looking at MSE plot
+num_bins_x_tsne_gau <- 8
+num_bins_y_tsne_gau <- 10
+hex_size_tsne_gau <- 0.1
+
+hb_obj_tsne_gau <- hex_binning(data = tsne_gau_scaled, x = "tSNE1", 
+                      y = "tSNE2", num_bins_x = num_bins_x_tsne_gau, 
+                      num_bins_y = num_bins_y_tsne_gau, x_start = NA, y_start = NA, 
+                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_tsne_gau, col_start = "tSNE")
+
+## Data set with all possible centroids in the hexagonal grid
+all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$centroids))
+
+## Generate all coordinates of hexagons
+hex_grid <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$hex_poly))
+
+## To obtain the standardise counts within hexbins
+counts_df <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$std_cts))
+df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
+                                             counts_df = counts_df)
+ggplot(data = hex_grid, aes(x = x, y = y)) + 
+  geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
+  geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
+  coord_fixed()
+
+hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
+
+ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+  geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
+  coord_fixed()
+
+tsne_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$data_hb_id))
+
+model_object_tsne_gau <- fit_highd_model( training_data = training_data_gau, 
+                                 nldr_df_with_id = tsne_gau_scaled, 
+                                 x = "tSNE1", y = "tSNE2", 
+                                 num_bins_x = num_bins_x_tsne_gau, 
+                                 num_bins_y = num_bins_y_tsne_gau, 
+                                 x_start = NA, y_start = NA, 
+                                 buffer_x = NA, buffer_y = NA, 
+                                 hex_size = hex_size_tsne_gau,
+                                 is_rm_lwd_hex = FALSE, 
+                                 benchmark_to_rm_lwd_hex = NA, 
+                                 col_start_2d = "tSNE", 
+                                 col_start_highd = "x")
+
+df_bin_centroids_tsne_gau <- model_object_tsne_gau$df_bin_centroids
+df_bin_tsne_gau <- model_object_tsne_gau$df_bin
+
+## Triangulate bin centroids
+tr1_object_tsne_gau <- tri_bin_centroids(df_bin_centroids_tsne_gau, x = "c_x", y = "c_y")
+tr_from_to_df_tsne_gau <- gen_edges(tri_object = tr1_object_tsne_gau)
+
+## Compute 2D distances
+distance_tsne_gau <- cal_2d_dist(tr_coord_df = tr_from_to_df_tsne_gau, 
+                             start_x = "x_from", start_y = "y_from", 
+                             end_x = "x_to", end_y = "y_to", 
+                             select_vars = c("from", "to", "distance"))
+
+distance_plot <- plot_dist(distance_tsne_gau) +
+  ylab(expression(d^{(2)})) +
+  theme(axis.text = element_text(size = 5),
+        axis.title = element_text(size = 12))
+
+distance_plot
+
+## To find the benchmark value
+benchmark_tsne_gau <- find_lg_benchmark(distance_edges = distance_tsne_gau, distance_col = "distance")
+
+ggplot() +
+geom_trimesh(data = df_bin_centroids_tsne_gau, mapping = aes(x = c_x, y = c_y))
+
+trimesh_removed_tsne_gau <- vis_rmlg_mesh(distance_edges = distance_tsne_gau, benchmark_value = benchmark_tsne_gau, tr_coord_df = tr_from_to_df_tsne_gau, distance_col = "distance")
+
+trimesh_removed_tsne_gau <- trimesh_removed_tsne_gau +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+df_all_tsne_gau <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID), tsne_data_with_hb_id_gau)
+
+# show_langevitour(df = df_all_tsne_gau, df_b = df_bin_tsne_gau, df_b_with_center_data = df_bin_centroids_tsne_gau, col_start = "x", distance_df = distance_tsne_gau, distance_col = "distance", benchmark_value = benchmark_tsne_gau)
+```
+
+```{r}
+#| warning: false
+#| echo: false
+#| message: false
+
+umap_gau_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = UMAP_data_gau, 
+                                    x = "UMAP1", y = "UMAP2"))) |>
+  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
+                  "UMAP2" = "scaled_UMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(UMAP_data_gau))
+
+# hex_size_vec <- seq(0.02, 2, by = 0.01)
+# 
+# for (i in 1:length(hex_size_vec)) {
+#   
+#   find_non_empty_bins(data = umap_gau_scaled, x = "UMAP1", y = "UMAP2",
+#                     non_empty_bins = 34, x_start = NA, y_start = NA,
+#                     buffer_x = NA, buffer_y = NA, hex_size = hex_size_vec[i],
+#                     col_start = "UMAP")
+#   
+#   
+# }
+
+```
+
+<!---
 ```{r}
@@
 
-```{r}
-#| eval: false
-#| echo: false
-#| fig-cap: "Goodness of fit statistics from different NLDR techniques applied to training five spherical Gaussian cluster dataset.
-What is the best NLDR technique to represent the original data in 2D?"
-#| label: fig-diagnosticpltGau
-#| out-width: 100%
-
-aic_gau_plot + mse_gau_plot +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', nrow = 2) &
-  theme(legend.position='bottom', plot.tag = element_text(size = 8))
-```
 -->
@@
 
+### Handwritten digits
+
+- Digit 1 (7877 images only)
+
+```{r}
+#| warning: false
+#| echo: false
+
+## Import data
+training_data_mnist <- read_rds("data/mnist/mnist_10_pcs_of_digit_1.rds")
+training_data_mnist <- training_data_mnist |>
+  mutate(ID = 1:NROW(training_data_mnist))
+
+pacmap_minst <- read_rds("data/mnist/mnist_pacmap.rds")
+pacmap_minst_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = pacmap_minst, 
+                                    x = "PaCMAP1", y = "PaCMAP2"))) |>
+  dplyr::rename(c("PaCMAP1" = "scaled_PaCMAP1", 
+                  "PaCMAP2" = "scaled_PaCMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(pacmap_minst))
+
+plot_list2_mnist <- pacmap_minst_scaled |>
+    ggplot(aes(x = PaCMAP1,
+               y = PaCMAP2))+
+    geom_point(alpha=0.5) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
+  # geom_text(aes(x=UMAP1,y = UMAP2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
+```
+
+```{r}
+#| warning: false
+#| echo: false
+#| message: false
+
+## UMAP
+## Prediction
+
+#tot_num_bin_vec <- 2:51
+hex_size_vec <- seq(0.02, 2, by = 0.01)
+
+vec <- stats::setNames(rep("", 6), c("num_bins", "aic", "mse", "num_bins_x", "num_bins_y", "hex_size"))  ## Define column names
+
+mse_df_mnist <- dplyr::bind_rows(vec)[0, ]
+mse_df_mnist <- mse_df_mnist |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(hex_size_vec)) {
+
+  num_bin_list <- calc_bins(data = pacmap_minst_scaled, 
+            x = "PaCMAP1", y = "PaCMAP2", 
+            hex_size = hex_size_vec[i], buffer_x = NA, buffer_y = NA)
+
+  num_bins_x <- num_bin_list$num_x
+  num_bins_y <- num_bin_list$num_y
+
+  model_object <- fit_highd_model( training_data = training_data_mnist, 
+                                   nldr_df_with_id = pacmap_minst_scaled, 
+                                   x = "PaCMAP1", y = "PaCMAP2", 
+                                   num_bins_x = num_bins_x, 
+                                   num_bins_y = num_bins_y, 
+                                   x_start = NA, y_start = NA, 
+                                   buffer_x = NA, buffer_y = NA, 
+                                   hex_size = hex_size_vec[i],
+                                   is_rm_lwd_hex = FALSE, 
+                                   benchmark_to_rm_lwd_hex = NA, 
+                                   col_start_2d = "PaCMAP", 
+                                   col_start_highd = "PC")
+
+  centroid_df_training <- model_object$df_bin_centroids
+  avg_df_training <- model_object$df_bin
+
+  pred_emb_list <- predict_emb(test_data = training_data_mnist, 
+                                  df_bin_centroids = centroid_df_training, 
+                                  df_bin = avg_df_training, type_NLDR = "PaCMAP")
+
+  pred_df_training <- as.data.frame(do.call(cbind, pred_emb_list))
+
+  eval_list <- gen_summary(test_data = training_data_mnist, 
+                                  prediction_df = pred_df_training, 
+                                  df_bin = avg_df_training, col_start = "PC")
+
+  mse_df_mnist <- mse_df_mnist |>
+    tibble::add_row(num_bins = num_bins_x * num_bins_y,
+                    aic = eval_list$aic,
+                    mse = eval_list$mse,
+                    num_bins_x = num_bins_x,
+                    num_bins_y = num_bins_y,
+                    hex_size = hex_size_vec[i])
+
+}
+
+## If same total number of bins occurred only select ones with minimum error
+### Obtain duplicate bins
+dupli_bins <- mse_df_mnist |> 
+  dplyr::count(num_bins) |> 
+  dplyr::filter(n > 1) |> 
+  dplyr::pull(num_bins)
+
+### Group split by duplicated bins
+duplicate_df_list <- mse_df_mnist |>
+  dplyr::filter(num_bins %in% dupli_bins) |>
+  dplyr::arrange(num_bins) |>
+  dplyr::group_split(num_bins)
+
+### Obtain one row from duplicates which have lowest error and hexsize
+duplicate_df <- data.frame(matrix(nrow = 0, ncol = 0))
+
+for (i in 1:length(duplicate_df_list)) {
+
+  dd <- duplicate_df_list[[i]] |>
+    dplyr::filter(mse == min(duplicate_df_list[[i]]$mse)) |>
+    dplyr::filter(hex_size == min(duplicate_df_list[[i]]$hex_size))
+
+  duplicate_df <- dplyr::bind_rows(duplicate_df, dd)
+
+}
+
+### Obtain the mse_df with not duplicated bins
+not_dupli_df <- mse_df_mnist |>
+  dplyr::filter(!(num_bins %in% dupli_bins))
+
+### Combine duplicated and not duplicated(corrected) bins dfs
+mse_df_mnist <- dplyr::bind_rows(not_dupli_df, duplicate_df)
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## To draw with AIC
+aic_plot_mnist <- ggplot(mse_df_mnist, aes(x = num_bins, y = aic
+)) +
+  geom_point() +
+  geom_line() +
+  geom_vline(xintercept = 80, linetype="solid",
+                color = "red", size=0.8, alpha = 0.5) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  #scale_colour_manual(values = c("#377eb8")) +
+  ylab("AIC") +
+  xlab("total number of bins")
+## Effective number of bins along x-axis
+
+mse_plot_mnist <- ggplot(mse_df_mnist, aes(x = num_bins,
+                                       y = mse
+)) +
+  geom_point() +
+  geom_line() +
+   geom_vline(xintercept = 80, linetype="solid",
+                color = "red", size=0.8, alpha = 0.5) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  #scale_colour_manual(values = c("#377eb8")) +
+  ylab("MSE") +
+  xlab("total number of bins")
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset.
+What is the effective number of bins in each NLDR technique to create a 2D model?
+The MSE plot have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
+Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error.
+Using the elbow method, when the total number of bins is set to 144, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
+This point indicates that adding less bins does not enough to capture the data structure."
+#| label: fig-diagnosticpltMNIST
+##| out-width: 100%
+#| fig-pos: H
+
+aic_plot_mnist + mse_plot_mnist +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='none', plot.tag = element_text(size = 8))
+```
+
+```{r}
+#| warning: false
+#| echo: false
+#| message: false
+
+## Decide by looking at MSE plot
+num_bins_x_mnist <- 6
+num_bins_y_mnist <- 13
+hex_size_mnist <- 0.14
+
+hb_obj_mnist <- hex_binning(data = pacmap_minst_scaled, x = "PaCMAP1", 
+                      y = "PaCMAP2", num_bins_x = num_bins_x_mnist, 
+                      num_bins_y = num_bins_y_mnist, x_start = NA, y_start = NA, 
+                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_mnist, col_start = "PaCMAP")
+
+## Data set with all possible centroids in the hexagonal grid
+all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_mnist$centroids))
+
+## Generate all coordinates of hexagons
+hex_grid <- as.data.frame(do.call(cbind, hb_obj_mnist$hex_poly))
+
+## To obtain the standardise counts within hexbins
+counts_df <- as.data.frame(do.call(cbind, hb_obj_mnist$std_cts))
+df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
+                                             counts_df = counts_df)
+ggplot(data = hex_grid, aes(x = x, y = y)) + 
+  geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
+  geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
+  coord_fixed()
+
+hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
+
+ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+  geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
+  coord_fixed()
+
+pacmap_data_with_hb_id_mnist <- as.data.frame(do.call(cbind, hb_obj_mnist$data_hb_id))
+
+model_object <- fit_highd_model( training_data = training_data_mnist, 
+                                 nldr_df_with_id = pacmap_minst_scaled, 
+                                 x = "PaCMAP1", y = "PaCMAP2", 
+                                 num_bins_x = num_bins_x_mnist, 
+                                 num_bins_y = num_bins_y_mnist, 
+                                 x_start = NA, y_start = NA, 
+                                 buffer_x = NA, buffer_y = NA, 
+                                 hex_size = hex_size_mnist,
+                                 is_rm_lwd_hex = FALSE, 
+                                 benchmark_to_rm_lwd_hex = NA, 
+                                 col_start_2d = "PaCMAP", 
+                                 col_start_highd = "PC")
+
+df_bin_centroids_mnist <- model_object$df_bin_centroids
+df_bin_mnist <- model_object$df_bin
+
+## Triangulate bin centroids
+tr1_object_mnist <- tri_bin_centroids(df_bin_centroids_mnist, x = "c_x", y = "c_y")
+tr_from_to_df_mnist <- gen_edges(tri_object = tr1_object_mnist)
+
+## Compute 2D distances
+distance_mnist <- cal_2d_dist(tr_coord_df = tr_from_to_df_mnist, 
+                             start_x = "x_from", start_y = "y_from", 
+                             end_x = "x_to", end_y = "y_to", 
+                             select_vars = c("from", "to", "distance"))
+
+distance_plot <- plot_dist(distance_mnist) +
+  ylab(expression(d^{(2)})) +
+  theme(axis.text = element_text(size = 5),
+        axis.title = element_text(size = 12))
+
+distance_plot
+
+## To find the benchmark value
+benchmark_mnist <- find_lg_benchmark(distance_edges = distance_mnist, distance_col = "distance")
+
+ggplot() +
+geom_trimesh(data = df_bin_centroids_mnist, mapping = aes(x = c_x, y = c_y))
+
+trimesh_removed_mnist_pacmap <- vis_rmlg_mesh(distance_edges = distance_mnist, benchmark_value = benchmark_mnist, tr_coord_df = tr_from_to_df_mnist, distance_col = "distance")
+
+trimesh_removed_mnist_pacmap <- trimesh_removed_mnist_pacmap +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+df_bin_train <- df_bin_mnist
+names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])
+
+df_all_mnist <- dplyr::bind_cols(training_data_mnist, pacmap_data_with_hb_id_mnist)
+
+error_df <- df_all_mnist |>
+  dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates
+
+# prediction_df_join <- prediction_df_join |>
+#   dplyr::left_join(data, by = c("ID" = "ID")) ## Map high-D data
+
+for (i in 1:(NCOL(df_bin_train) - 1)) {
+
+  error_df[ , paste0("abs_residual_", "PC", i)] <- abs(error_df[ , paste0("PC", i)] - error_df[ , paste0("avg_", "PC", i)])
+
+}
+
+error_df <- error_df |>
+  dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "PC")))))
+
+# library(ggbeeswarm)
+# error_df$group <- "1"
+# ggplot(error_df, aes(x = group, y = total)) +
+#   geom_quasirandom()+
+#   ylim(0, max(unlist(error_df$total))+ 0.5) + coord_flip()
+
+### The minimum error is 0 and the maximum is 42.17439
+### There is lot of points with error 0,
+
+error_df <- error_df |>
+  mutate(type = if_else(total == 0, "no error",
+                        if_else(total <= 80, "error 1-50",
+                                if_else(total <= 100, "error 50-100",
+                                        if_else(total <= 200, "error 100-200",
+                                                if_else(total <= 300, "error 200-300",
+                                                        if_else(total <= 400, "error 300-400",
+                                                                if_else(total <= 500, "error 400-500", "error greter than 500"))))))))
+
+plot_list2_mnist_error <- error_df |>
+  mutate(type = factor(type , levels = c("no error", "error 1-50", "error 50-100", "error 100-200", "error 200-300", "error 300-400", "error 400-500", "error greter than 500"))) |>
+    ggplot(aes(x = PaCMAP1,
+               y = PaCMAP2, color = type))+
+    geom_point(alpha=0.5, size = 0.1) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#fee0d2", "#fcbba1",
+                                         "#fc9272", "#fb6a4a", "#ef3b2c",
+                                         "#cb181d", "#a50f15", "#99000d"))
+
+# show_langevitour(df_all_mnist, df_bin_mnist, df_bin_centroids_mnist, benchmark_mnist, distance_mnist, "distance", col_start = "PC")
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "(a) 2D layout from UMAP (n_neighbors = 30, min_dist = 0.3) applied for the PBMC3k dataset.
+Is this a best representation of the original data?, (b)  Model in the 2D space with UMAP (<https://youtu.be/CNXjWZMyh_M>)"
+#| label: fig-nldervisMNISTPACMAP
+#| out-width: 100%
+#| fig-pos: h
+
+plot_list2_mnist + trimesh_removed_mnist_pacmap + plot_list2_mnist_error +
+plot_layout(guides='collect', ncol=3) &
+  theme(legend.position='bottom')
+```
+
 ## Conclusion {#sec-conclusion}

### 2024-03-11 paper.tex
Added Handwritten digit: 1 dataset

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, enhanced, breakable, borderline west={3pt}{0pt}{shadecolor}, interior hidden, frame hidden, boxrule=0pt]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, frame hidden, enhanced, boxrule=0pt, breakable, borderline west={3pt}{0pt}{shadecolor}, interior hidden]}{\end{tcolorbox}}\fi
 
@@
 provides the most useful representation of high-D data. Our approach
-involves dividing the high-dimensional dataset into two parts: a
-training set for constructing the model and a test set for generating
-predictive values and residuals. Our algorithm takes a 2D embedding data
-as the input and generate a tour that displays the high-dimensional
-wireframe to overlay the data. The flow chart of the proposed algorithm
-is shown in Figure~\ref{fig-meth}. The algorithm consists of two main
-phases: (1) generating the model in the 2D space and (2) lifting the
-model into high-D space. The main steps of the algorithm are described
-in detail in this section using UMAP 2D embedding of the S-curve
-dataset. This dataset has seven dimensions, including four noise
-dimensions that were added to the original 3D data.
+involves dividing the high-D dataset into two parts: a training set for
+constructing the model and a test set for generating predictive values
+and residuals. Our algorithm takes a 2D embedding data as the input and
+generate a tour that displays the high-D wireframe to overlay the data.
+The flow chart of the proposed algorithm is shown in
+Figure~\ref{fig-meth}. The algorithm consists of two main phases: (1)
+generating the model in the 2D space and (2) lifting the model into
+high-D space. The main steps of the algorithm are described in detail in
+this section using UMAP 2D embedding of the S-curve dataset. This
+dataset has seven dimensions, including four noise dimensions that were
+added to the original 3D data.
 
@@
 
-In order to reduce the computational complexity associated with
-performing NLDR techniques to high-D data, and as a initial step of
-noise reduction of high-D data, PCA \citep[\citet{article68},
-\citet{article69}]{article67} is applied as a preprocessing step.
-Subsequently, the identified principal components, representing
-directions of maximum variance, are used to construct the model.
-
-Then, after performing NLDR, to have regular hexagons, 2D embedding data
-are scaled accordingly.
+To reduce computational complexity when applying NLDR techniques to
+high-D data and to reduce noise presence, PCA \citep[\citet{article68},
+\citet{article69}]{article67} is used as a preprocessing step. PCA
+involves identifying principal components that maximize variance. These
+components are then used as the high-D data for the algorithm.
 
@@
 \textbf{Step 1: Scaling NLDR data}
+
+First, we prepare the 2D embedding data to fit within the bounds
+required for regular hexagonal binning. To achieve this, we implement
+two key scaling steps. Scale the first 2D embedding component to range
+between \(0\) and \(1\), ensuring that all data points fall within this
+normalized interval. Secondly, we scale the second 2D embedding
+component to range between \(0\) and \(y_{max}\) (see
+Equation~\ref{eq-equation3}).
+
+\begin{equation}\protect\hypertarget{eq-equation1}{}{
+ ar = \frac{r_2}{r_1}
+}\label{eq-equation1}\end{equation}
+
+\begin{equation}\protect\hypertarget{eq-equation2}{}{
+ hr = \frac{hb}{wb}
+}\label{eq-equation2}\end{equation}
+
+\begin{equation}\protect\hypertarget{eq-equation3}{}{
+ y_{max} = \text{ceiling}(\frac{ar}{hr}) * hr
+}\label{eq-equation3}\end{equation}
 
@@
 high-D model point is identified, its coordinates are used as the
-predicted 2D embeddings for the new high-D data point. This step is
+predicted 2D embedding for the new high-D data point. This step is
 particularly valuable due to the limitations of certain NLDR techniques,
 like tSNE, which don't provide a straightforward method for predicting
-2D embeddings. As a result, our approach presents a solution that can
-generate predicted 2D embeddings, irrespective of the NLDR technique
+2D embedding. As a result, our approach presents a solution that can
+generate predicted 2D embedding, irrespective of the NLDR technique
 employed, effectively addressing this gap in functionality.
@@
 In the 2D representations created by all NLDR techniques, as shown in
-\textbf{?@fig-nldervis5Gau}, except for PHATE, there are five distinct
+Figure~\ref{fig-nldervis5Gau}, except for PHATE, there are five distinct
 clusters. In tSNE, these five clusters are closely located to each other
-(see \textbf{?@fig-nldervis5Gau} (a)). However, in PHATE, two clusters
+(see Figure~\ref{fig-nldervis5Gau} (a)). However, in PHATE, two clusters
 are closely positioned, while the other three are more distant (see
-\textbf{?@fig-nldervis5Gau} (c)). In PaCMAP, one cluster is at the
+Figure~\ref{fig-nldervis5Gau} (c)). In PaCMAP, one cluster is at the
 center, and the remaining four are positioned in four different
-directions (see \textbf{?@fig-nldervis5Gau} (e)). In TriMAP, two
+directions (see Figure~\ref{fig-nldervis5Gau} (e)). In TriMAP, two
 clusters are close, although not as close as in PHATE, and the other
-three are well-separated (see \textbf{?@fig-nldervis5Gau} (d)). In UMAP,
-all clusters are arranged in a parallel manner, with three in one line
-and the other two in a separate line (see \textbf{?@fig-nldervis5Gau}
-(b)).
+three are well-separated (see Figure~\ref{fig-nldervis5Gau} (d)). In
+UMAP, all clusters are arranged in a parallel manner, with three in one
+line and the other two in a separate line (see
+Figure~\ref{fig-nldervis5Gau} (b)).
 
@@
 However, tSNE, effectively capture both local and global structures, as
-indicated by lower AIC values in \textbf{?@fig-diagnosticpltGau} (a).
+indicated by lower AIC values in Figure~\ref{fig-diagnosticpltGau} (a).
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervis5Gau-1.pdf}
+
+}
+
+\caption{\label{fig-nldervis5Gau}2D layouts from different NLDR
+techniques applied the same data: (a) tSNE (perplexity = 61), (b) UMAP
+(n\_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n\_inliers = 5,
+n\_outliers = 4, n\_random = 3), and (e) PaCMAP (n\_neighbors = 10, init
+= random, MN\_ratio = 0.9, FP\_ratio = 2). Is there a best
+representation of the original data or are they all providing equivalent
+information?}
+
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticpltGau-1.pdf}
+
+}
+
+\caption{\label{fig-diagnosticpltGau}Goodness of fit statistics from
+different NLDR techniques applied to training five spherical Gaussian
+cluster dataset. What is the best NLDR technique to represent the
+original data in 2D?}
+
+\end{figure}
+
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-44-1.pdf}
+
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-44-2.pdf}
+
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-44-3.pdf}
+
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-44-4.pdf}
 
@@
 
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-55-1.pdf}
-
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-55-2.pdf}
-
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-55-3.pdf}
-
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-55-4.pdf}
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-59-1.pdf}
+
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-59-2.pdf}
+
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-59-3.pdf}
+
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-59-4.pdf}
 
@@
 
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-61-1.pdf}
-
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-61-2.pdf}
-
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-61-3.pdf}
-
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-61-4.pdf}
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-65-1.pdf}
+
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-65-2.pdf}
+
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-65-3.pdf}
+
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-65-4.pdf}
 
@@
 
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-64-1.pdf}
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-68-1.pdf}
+
+\hypertarget{handwritten-digits}{%
+\subsection{Handwritten digits}\label{handwritten-digits}}
+
+\begin{itemize}
+\tightlist
+\item
+  Digit 1 (7877 images only)
+\end{itemize}
+
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-diagnosticpltMNIST-1.pdf}
+
+}
+
+\caption{\label{fig-diagnosticpltMNIST}Goodness of fit statistics from
+UMAP applied to training S-curve dataset. What is the effective number
+of bins in each NLDR technique to create a 2D model? The MSE plot have a
+steep slope at the beginning, indicating that a smaller number of bins
+causes a larger amount of error. Then, the slope gradually declines or
+level off, indicating that a higher number of bins generates a smaller
+error. Using the elbow method, when the total number of bins is set to
+144, the slope of the Mean Squared Error (MSE) plot experiences a sudden
+and noticeable change, resembling an elbow-like shape. This point
+indicates that adding less bins does not enough to capture the data
+structure.}
+
+\end{figure}
+
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-73-1.pdf}
+
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-73-2.pdf}
+
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-73-3.pdf}
+
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-73-4.pdf}
+
+\begin{figure}[h]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervisMNISTPACMAP-1.pdf}
+
+}
+
+\caption{\label{fig-nldervisMNISTPACMAP}(a) 2D layout from UMAP
+(n\_neighbors = 30, min\_dist = 0.3) applied for the PBMC3k dataset. Is
+this a best representation of the original data?, (b) Model in the 2D
+space with UMAP (\url{https://youtu.be/CNXjWZMyh_M})}
+
+\end{figure}
 

### 2024-03-11 paper.qmd
Added PaCMAP layout for handwritten data with preprocessing steps

@@
 
+#### Preprocessing
+
+- Scale the pixel values to the range [0,1]
+
+- Reshape to have a 2-dimensional array
+
+-  reshaping pixel data is a fundamental preprocessing step in image analysis tasks, ensuring that the data is in a suitable format for further analysis, modeling, and interpretation.
+It helps preserve important spatial information, ensures compatibility with machine learning models, and facilitates feature extraction and visualization.
+
+-  Reshaping the pixel data allows us to preserve this spatial information, ensuring that the relationships between neighboring pixels are maintained.
+
+- Many machine learning algorithms, including neural networks, expect input data to be in a specific format.
+For image data, this typically means a two-dimensional array where each row represents a single sample (image), and each column represents a feature (pixel value).
+Reshaping ensures that all samples have consistent dimensions, making them suitable for feeding into these algorithms.
+
+- Reshaping the pixel data allows for easier extraction of relevant features from the images.
+Once reshaped, individual pixels can be treated as features, making it easier to apply techniques like dimensionality reduction (e.g., PCA) or feature engineering to extract meaningful information from the images.
+
+- PaCMAP parameters: n_components=2, n_neighbors=10, init="random", MN_ratio=0.9, FP_ratio=2.0
+
 ```{r}
@@
 #| echo: false
-#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset.
+#| fig-cap: "Goodness of fit statistics from PaCMAP applied to training MNIST dataset.
 What is the effective number of bins in each NLDR technique to create a 2D model?
@@
 Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error.
-Using the elbow method, when the total number of bins is set to 144, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
+Using the elbow method, when the total number of bins is set to 80, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
 This point indicates that adding less bins does not enough to capture the data structure."
@@
 ## Decide by looking at MSE plot
-num_bins_x_mnist <- 6
-num_bins_y_mnist <- 13
-hex_size_mnist <- 0.14
+num_bins_x_mnist <- 8
+num_bins_y_mnist <- 10
+hex_size_mnist <- 0.1
 
@@
                                              counts_df = counts_df)
-ggplot(data = hex_grid, aes(x = x, y = y)) + 
-  geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
-  geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
-  coord_fixed()
+# ggplot(data = hex_grid, aes(x = x, y = y)) + 
+#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
+#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
+#   coord_fixed()
 
@@
 
-ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
-  geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
-  coord_fixed()
+# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
+#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
+#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
+#   coord_fixed()
 
@@
 
-distance_plot <- plot_dist(distance_mnist) +
-  ylab(expression(d^{(2)})) +
-  theme(axis.text = element_text(size = 5),
-        axis.title = element_text(size = 12))
-
-distance_plot
+# distance_plot <- plot_dist(distance_mnist) +
+#   ylab(expression(d^{(2)})) +
+#   theme(axis.text = element_text(size = 5),
+#         axis.title = element_text(size = 12))
+# 
+# distance_plot
 
@@
 
-ggplot() +
-geom_trimesh(data = df_bin_centroids_mnist, mapping = aes(x = c_x, y = c_y))
+# ggplot() +
+# geom_trimesh(data = df_bin_centroids_mnist, mapping = aes(x = c_x, y = c_y))
 
@@
   mutate(type = if_else(total == 0, "no error",
-                        if_else(total <= 80, "error 1-50",
-                                if_else(total <= 100, "error 50-100",
-                                        if_else(total <= 200, "error 100-200",
-                                                if_else(total <= 300, "error 200-300",
-                                                        if_else(total <= 400, "error 300-400",
-                                                                if_else(total <= 500, "error 400-500", "error greter than 500"))))))))
+                        if_else(total <= 2, "error 0-2",
+                                if_else(total <= 4, "error 2-4",
+                                        if_else(total <= 6, "error 4-6",
+                                                if_else(total <= 8, "error 6-8",
+                                                        if_else(total <= 12, "error 8-10",
+                                                                if_else(total <= 14, "error 10-12", "error greter than 12"))))))))
 
 plot_list2_mnist_error <- error_df |>
-  mutate(type = factor(type , levels = c("no error", "error 1-50", "error 50-100", "error 100-200", "error 200-300", "error 300-400", "error 400-500", "error greter than 500"))) |>
+  mutate(type = factor(type , levels = c("no error", "error 0-2", "error 2-4", "error 4-6", "error 6-8", "error 8-10", "error 10-12", "error greter than 12"))) |>
     ggplot(aes(x = PaCMAP1,
@@
 #| echo: false
-#| fig-cap: "(a) 2D layout from UMAP (n_neighbors = 30, min_dist = 0.3) applied for the PBMC3k dataset.
-Is this a best representation of the original data?, (b)  Model in the 2D space with UMAP (<https://youtu.be/CNXjWZMyh_M>)"
+#| fig-cap: "(a) 2D layout from UMAP (n_components=2, n_neighbors=10, init=random, MN_ratio=0.9, FP_ratio=2.0) applied for the digit 1 of the MNIST dataset.
+Is this a best representation of the original data?, (b)  Model in the 2D space with PaCMAP (<https://youtu.be/CNXjWZMyh_M>)"
 #| label: fig-nldervisMNISTPACMAP

### 2024-03-11 paper.tex
Added PaCMAP layout for handwritten data with preprocessing steps

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, frame hidden, enhanced, boxrule=0pt, breakable, borderline west={3pt}{0pt}{shadecolor}, interior hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, interior hidden, breakable, sharp corners, enhanced, borderline west={3pt}{0pt}{shadecolor}, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 
+\hypertarget{preprocessing}{%
+\subsubsection{Preprocessing}\label{preprocessing}}
+
+\begin{itemize}
+\item
+  Scale the pixel values to the range {[}0,1{]}
+\item
+  Reshape to have a 2-dimensional array
+\item
+  reshaping pixel data is a fundamental preprocessing step in image
+  analysis tasks, ensuring that the data is in a suitable format for
+  further analysis, modeling, and interpretation. It helps preserve
+  important spatial information, ensures compatibility with machine
+  learning models, and facilitates feature extraction and visualization.
+\item
+  Reshaping the pixel data allows us to preserve this spatial
+  information, ensuring that the relationships between neighboring
+  pixels are maintained.
+\item
+  Many machine learning algorithms, including neural networks, expect
+  input data to be in a specific format. For image data, this typically
+  means a two-dimensional array where each row represents a single
+  sample (image), and each column represents a feature (pixel value).
+  Reshaping ensures that all samples have consistent dimensions, making
+  them suitable for feeding into these algorithms.
+\item
+  Reshaping the pixel data allows for easier extraction of relevant
+  features from the images. Once reshaped, individual pixels can be
+  treated as features, making it easier to apply techniques like
+  dimensionality reduction (e.g., PCA) or feature engineering to extract
+  meaningful information from the images.
+\item
+  PaCMAP parameters: n\_components=2, n\_neighbors=10, init=``random'',
+  MN\_ratio=0.9, FP\_ratio=2.0
+\end{itemize}
+
 \begin{figure}
@@
 \caption{\label{fig-diagnosticpltMNIST}Goodness of fit statistics from
-UMAP applied to training S-curve dataset. What is the effective number
+PaCMAP applied to training MNIST dataset. What is the effective number
 of bins in each NLDR technique to create a 2D model? The MSE plot have a
@@
 error. Using the elbow method, when the total number of bins is set to
-144, the slope of the Mean Squared Error (MSE) plot experiences a sudden
+80, the slope of the Mean Squared Error (MSE) plot experiences a sudden
 and noticeable change, resembling an elbow-like shape. This point
@@
 
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-73-1.pdf}
-
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-73-2.pdf}
-
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-73-3.pdf}
-
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-73-4.pdf}
-
 \begin{figure}[h]
@@
 \caption{\label{fig-nldervisMNISTPACMAP}(a) 2D layout from UMAP
-(n\_neighbors = 30, min\_dist = 0.3) applied for the PBMC3k dataset. Is
-this a best representation of the original data?, (b) Model in the 2D
-space with UMAP (\url{https://youtu.be/CNXjWZMyh_M})}
+(n\_components=2, n\_neighbors=10, init=random, MN\_ratio=0.9,
+FP\_ratio=2.0) applied for the digit 1 of the MNIST dataset. Is this a
+best representation of the original data?, (b) Model in the 2D space
+with PaCMAP (\url{https://youtu.be/CNXjWZMyh_M})}
 

### 2024-03-11 paper.qmd
Added 2D model explanation with equations

@@
 
+The calculation of $y_{max}$ involves several steps.
+First, the aspect ratio ($ar$) is computed by dividing the range of the second 2D embedding component ($r_2$) by the range of the first 2D embedding component ($r_1$) (see @eq-equation1).
+Then, the hexagon ratio ($hr$) is determined by dividing the height of the hexagon ($hb$) by its width ($wb$) (see @eq-equation2).
+Finally, $y_{max}$ is derived by taking the ceiling of $\frac{ar}{hr}$ and multiplying it by $hr$.
+This process ensures that $y_{max}$ is an integer multiple of $hr$, accommodating the grid layout of the hexagonal bins.
+
 $$
@@
 
+Hexagonating NLDR data (see @fig-meth Step 2) involves partitioning the 2D embedding data into hexagonal bins, a technique commonly referred to as hexagonal binning [@Carr1987, @article66].
+This method use a hexagonal grid to create a bivariate histogram that effectively visualizes the structure of high-D data.
+Hexagons, one of only three regular polygons capable of tessellating a plane [@Carr2013], offer unique advantages due to their symmetry of nearest neighbors and maximal number of sides for such tessellations [@Dan2023].
+This geometric property makes hexagons more efficient in covering the plane compared to other regular tessellations and reduces visual bias when displaying data densities [@Dan2023].
+In our algorithm, we aim to conduct regular hexagonal binning (**cite quollr paper**), involving the computation of hexagonal grid configurations, the generation of the grid, and the assignment of the 2D embedding to hexagons.
+
+<!--
 **(a) Determine the number of bins along the x axis** ($b_1$)
@@
 
-**(d) Assign NLDR data to hexagons**
+**(d) Assign NLDR data to hexagons**-->
 
@@
 
-The mathematical expression for the hexagonal bin centers can be defined based on the hexagonal grid coordinates.
-Let's denote the hexagonal grid coordinates as \( (q, r) \), where \( q \) and \( r \) are integers representing the coordinates of the hexagon in the grid.
-
-For a regular hexagonal grid, the coordinates of the hexagon centers can be calculated using the following formulas:
+In the previous step, the algorithm clusters the 2D embedding data into hexagons. Following this, in this step, the bin centroids or bin means (see @fig-meth Step 3) are obtained [@Carr2013].
+
+The bin centroid ($C_k^{(2)}$) for a $k^{th}$ hexagon with hexagonal grid coordinates $(h^{k}x_{i}, h^{k}y_{i})$, where $i = 1 \dots 6$ can be defined as:
+
+$$
+C_k^{(2)} = (C_{ky_1}, C_{ky_2}) = \left(\frac{\sum_{i=1}^{6} h^{k}x_{i}}{6}, \frac{\sum_{i=1}^{6} h^{k}y_{i}}{6}\right).
+$$ {#eq-equation4}
+
+Also, the bin mean ($C_k^{(2)}$) is defined as the mean of the data points within the $k^{th}$ hexagon (see @eq-equation5).
+
+$$
+C_k^{(2)} = (C_{ky_1}, C_{ky_2}) = \left(\frac{1}{n_k} \sum_{i=1}^{n_k} y_{1i}, \frac{1}{n_k} \sum_{i=1}^{n_k} y_{2i}\right),
+$$ {#eq-equation5}
+
+where $n_k$ is the number of data points within the hexagon, $y_{1i}$ and $y_{2i}$ are the $x$ and $y$ coordinates of the $i^{th}$ data point within the hexagon.
+
+**Step 4: Triangulating bin centroids or bin means**
+
+In this step, the algorithm proceeds to triangulate the hexagonal bin centroids or bin means (see @fig-meth Step 4).
+Triangulation is a fundamental process in computational geometry and computer graphics that involves dividing a set of points in a given space into interconnected triangles [@article30].
+One common algorithm used for triangulation is Delaunay triangulation [@article26, @article54], where points are connected in a way that maximizes the minimum angles of the resulting triangles, leading to a more regular and well-conditioned triangulation.
+
+Delaunay triangulation can be defined as follows:
+
+Let $C^{(2)} = \{C_1^{(2)}, C_2^{(2)}, ..., C_m^{(2)}\}$ be a set of $m$ bin centroids or bin means in the plane.
+Delaunay triangulation of $C^{(2)}$, denoted as $DT(C^{(2)})$, is a triangulation of the convex hull of $C^{(2)}$ such that the circumcircle of every triangle in the triangulation contains no other points from $C^{(2)}$.
+
+Given that the hexagons are regular, the resulting triangles will mostly be equilateral.
+
 <!--
-\[
-x = \frac{3}{2} \times q \times s
-\]
-\[
-y = \frac{\sqrt{3}}{2} \times q \times s + \sqrt{3} \times r \times s
-\]-->
-
-Where:
-- \( s \) is the side length of the hexagon.
-- \( q \) represents the column index of the hexagon in the grid.
-- \( r \) represents the row index of the hexagon in the grid.
-- \( x \) and \( y \) are the Cartesian coordinates of the center of the hexagon.
-
-This formula calculates the \( x \) and \( y \) coordinates of the center of each hexagon based on its grid coordinates \( q \) and \( r \), allowing you to determine the position of each hexagon in the plane.
-
-**Step 4: Triangulating bin centroids or bin means**
-
-Delaunay triangulation is a geometric algorithm used to create a triangular mesh from a set of points in a plane.
-The key principle behind Delaunay triangulation is to connect points in such a way that no point falls within the circumcircle of any triangle formed by the points.
-
-Mathematically, given a set of points \( P \) in a plane, the Delaunay triangulation constructs a set of triangles such that:
-
-1. Each triangle is formed by connecting three points from \( P \).
-2. No point in \( P \) is inside the circumcircle of any triangle in the triangulation.
-
-Formally, the Delaunay triangulation can be defined as follows:
-
-Let \( P = \{p_1, p_2, ..., p_n\} \) be a set of \( n \) points in the plane.
-The Delaunay triangulation of \( P \), denoted as \( DT(P) \), is a triangulation of the convex hull of \( P \) such that the circumcircle of every triangle in the triangulation contains no other points from \( P \).
 
@@
 $${#eq-equation5}
-
-**Step 2: Obtain bin centroids or bin means**
-
-As a result of hexagonal binning for high-D data, all the high-D data are clustered into hexagons.
-In this step, the bin centroids ($C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2})$) (see @fig-meth Step 2) are obtained [@Carr2013].
-
-**Step 3: Triangulate bin centroids**
-
-In this step, the algorithm proceeds to triangulate the hexagonal bin centroids (see @fig-meth Step 3).
-Triangulation is a fundamental process in computational geometry and computer graphics that involves dividing a set of points in a given space into interconnected triangles [@article30].
-One common algorithm used for triangulation is Delaunay triangulation [@article26, @article54], where points are connected in a way that maximizes the minimum angles of the resulting triangles, leading to a more regular and well-conditioned triangulation.
-
-Since we are working with the centroids of regular hexagonal bins, the resulting mesh will predominantly comprise equal-sized regular triangles.
-However, the triangulation also helps span any gaps that may exist between clusters of points, allowing for a more complete and interconnected representation of the data.
+-->
 
@@
 #| fig-cap: "(a) 2D layout from UMAP (n_components=2, n_neighbors=10, init=random, MN_ratio=0.9, FP_ratio=2.0) applied for the digit 1 of the MNIST dataset.
-Is this a best representation of the original data?, (b)  Model in the 2D space with PaCMAP (<https://youtu.be/CNXjWZMyh_M>)"
+Is this a best representation of the original data?, (b)  Model in the 2D space with PaCMAP ()"
 #| label: fig-nldervisMNISTPACMAP

### 2024-03-11 paper.tex
Added 2D model explanation with equations

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, interior hidden, breakable, sharp corners, enhanced, borderline west={3pt}{0pt}{shadecolor}, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, enhanced, sharp corners, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden, breakable]}{\end{tcolorbox}}\fi
 
@@
 
+The calculation of \(y_{max}\) involves several steps. First, the aspect
+ratio (\(ar\)) is computed by dividing the range of the second 2D
+embedding component (\(r_2\)) by the range of the first 2D embedding
+component (\(r_1\)) (see Equation~\ref{eq-equation1}). Then, the hexagon
+ratio (\(hr\)) is determined by dividing the height of the hexagon
+(\(hb\)) by its width (\(wb\)) (see Equation~\ref{eq-equation2}).
+Finally, \(y_{max}\) is derived by taking the ceiling of
+\(\frac{ar}{hr}\) and multiplying it by \(hr\). This process ensures
+that \(y_{max}\) is an integer multiple of \(hr\), accommodating the
+grid layout of the hexagonal bins.
+
 \begin{equation}\protect\hypertarget{eq-equation1}{}{
@@
 
-\textbf{(a) Determine the number of bins along the x axis} (\(b_1\))
-
-\textbf{(b) Determine the number of bins along the y axis} (\(b_2\))
-
-\textbf{(c) Determine the total number of bins} (\(b\))
-
-\textbf{(d) Assign NLDR data to hexagons}
+Hexagonating NLDR data (see Figure~\ref{fig-meth} Step 2) involves
+partitioning the 2D embedding data into hexagonal bins, a technique
+commonly referred to as hexagonal binning
+\citep[\citet{article66}]{Carr1987}. This method use a hexagonal grid to
+create a bivariate histogram that effectively visualizes the structure
+of high-D data. Hexagons, one of only three regular polygons capable of
+tessellating a plane \citep{Carr2013}, offer unique advantages due to
+their symmetry of nearest neighbors and maximal number of sides for such
+tessellations \citep{Dan2023}. This geometric property makes hexagons
+more efficient in covering the plane compared to other regular
+tessellations and reduces visual bias when displaying data densities
+\citep{Dan2023}. In our algorithm, we aim to conduct regular hexagonal
+binning (\textbf{cite quollr paper}), involving the computation of
+hexagonal grid configurations, the generation of the grid, and the
+assignment of the 2D embedding to hexagons.
 
@@
 
-The mathematical expression for the hexagonal bin centers can be defined
-based on the hexagonal grid coordinates. Let's denote the hexagonal grid
-coordinates as ( (q, r) ), where ( q ) and ( r ) are integers
-representing the coordinates of the hexagon in the grid.
-
-For a regular hexagonal grid, the coordinates of the hexagon centers can
-be calculated using the following formulas:
-
-Where: - ( s ) is the side length of the hexagon. - ( q ) represents the
-column index of the hexagon in the grid. - ( r ) represents the row
-index of the hexagon in the grid. - ( x ) and ( y ) are the Cartesian
-coordinates of the center of the hexagon.
-
-This formula calculates the ( x ) and ( y ) coordinates of the center of
-each hexagon based on its grid coordinates ( q ) and ( r ), allowing you
-to determine the position of each hexagon in the plane.
+In the previous step, the algorithm clusters the 2D embedding data into
+hexagons. Following this, in this step, the bin centroids or bin means
+(see Figure~\ref{fig-meth} Step 3) are obtained \citep{Carr2013}.
+
+The bin centroid (\(C_k^{(2)}\)) for a \(k^{th}\) hexagon with hexagonal
+grid coordinates \((h^{k}x_{i}, h^{k}y_{i})\), where \(i = 1 \dots 6\)
+can be defined as:
+
+\begin{equation}\protect\hypertarget{eq-equation4}{}{
+C_k^{(2)} = (C_{ky_1}, C_{ky_2}) = \left(\frac{\sum_{i=1}^{6} h^{k}x_{i}}{6}, \frac{\sum_{i=1}^{6} h^{k}y_{i}}{6}\right).
+}\label{eq-equation4}\end{equation}
+
+Also, the bin mean (\(C_k^{(2)}\)) is defined as the mean of the data
+points within the \(k^{th}\) hexagon (see Equation~\ref{eq-equation5}).
+
+\begin{equation}\protect\hypertarget{eq-equation5}{}{
+C_k^{(2)} = (C_{ky_1}, C_{ky_2}) = \left(\frac{1}{n_k} \sum_{i=1}^{n_k} y_{1i}, \frac{1}{n_k} \sum_{i=1}^{n_k} y_{2i}\right),
+}\label{eq-equation5}\end{equation}
+
+where \(n_k\) is the number of data points within the hexagon,
+\(y_{1i}\) and \(y_{2i}\) are the \(x\) and \(y\) coordinates of the
+\(i^{th}\) data point within the hexagon.
 
@@
 
-Delaunay triangulation is a geometric algorithm used to create a
-triangular mesh from a set of points in a plane. The key principle
-behind Delaunay triangulation is to connect points in such a way that no
-point falls within the circumcircle of any triangle formed by the
-points.
-
-Mathematically, given a set of points ( P ) in a plane, the Delaunay
-triangulation constructs a set of triangles such that:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Each triangle is formed by connecting three points from ( P ).
-\item
-  No point in ( P ) is inside the circumcircle of any triangle in the
-  triangulation.
-\end{enumerate}
-
-Formally, the Delaunay triangulation can be defined as follows:
-
-Let ( P = \{p\_1, p\_2, \ldots, p\_n\} ) be a set of ( n ) points in the
-plane. The Delaunay triangulation of ( P ), denoted as ( DT(P) ), is a
-triangulation of the convex hull of ( P ) such that the circumcircle of
-every triangle in the triangulation contains no other points from ( P ).
-
-\textbf{Step 1: Computing the hexagonal grid configuration}
-
-The hexagonal grid, formed through hexagonal binning
-\citep[\citet{article66}]{Carr1987}, serves as a type of bivariate
-histogram employed to visualize the structure of high-D data. Hexagons,
-being one of only three regular polygons capable of tessellating a plane
-\citep{Carr2013}, possess both symmetry of nearest neighbors and the
-maximum number of sides for a regular tessellation of the plane. This
-unique combination makes hexagons more efficient in covering the plane
-compared to other regular tessellations. Additionally, hexagons exhibit
-lower visual bias when displaying densities, setting them apart from
-other regular tessellations \citep{Dan2023}. In our algorithm, hexagonal
-binning is used as the initial step of constructing the 2D model and the
-total number of bins (\(b\)) is the crucial parameter that defines the
-granularity of the hexagonal grid.
-
-\textbf{(a) Determine the number of bins along the x axis} (\(b_1\))
-
-First, the number of bins along the x-axis (\(b_1\)) is computed using
-the relationship between the diameter (\(h\)) and the area (\(A\)) of
-regular hexagons (see Equation~\ref{eq-equation3}).
-
-In the process of computing hexagonal bin configurations, the first step
-involves determining the number of bins along the x-axis. This begins
-with the initialization of parameters such as the hexagonal size (\(s\))
-and the buffer along the x-axis. Subsequently, the range of the first 2D
-embedding component (\(r_1\)) is calculated to ensure comprehensive
-coverage of the data. This range is then adjusted by incorporating the
-buffer amount, effectively expanding the boundary to accommodate
-potential outliers or edge cases. The default buffer amount along the
-x-axis is \(\sqrt{3} * s * 1.5\). Next, the horizontal spacing (\(h\))
-between hexagons is computed based on the hexagon size. Finally, the
-number of bins along the x-axis is computed based on the adjusted range
-(\(r_1 + buffer_{x}\)) and the horizontal spacing (\(h\)).
-
-\begin{equation}\protect\hypertarget{eq-equation3}{}{
- h = \sqrt{3} * s
-}\label{eq-equation3}\end{equation}
-
-\begin{equation}\protect\hypertarget{eq-equation4}{}{
- b_1 = \frac{r_1 + buffer_{x}}{h}
-}\label{eq-equation4}\end{equation}
-
-\textbf{(b) Determine the number of bins along the y-axis} (\(b_2\))
-
-Next, the number of bins along the y-axis is computed based on the
-number of bins along the x-axis (\(b_1\)) and the shape parameter
-(\(s\)) (see Equation~\ref{eq-equation4}) \citep{Carr2013}.
-
-\begin{equation}\protect\hypertarget{eq-equation5}{}{
-  v = 1.5 * s
-}\label{eq-equation5}\end{equation}
-
-\begin{equation}\protect\hypertarget{eq-equation6}{}{
- b_2 = \frac{r_2 + buffer_{y}}{v}
-}\label{eq-equation6}\end{equation}
-
-\textbf{(c) Determine the total number of bins} (\(b\))
-
-The total number of bins is determined by multiplying the number of bins
-along the x-axis (\(b_1\)) with the number of bins along the y-axis
-(\(b_2\)) (see Equation~\ref{eq-equation5}).
-
-\begin{equation}\protect\hypertarget{eq-equation5}{}{
-b = b_1 \times b_2
-}\label{eq-equation5}\end{equation}
-
-\textbf{Step 2: Obtain bin centroids or bin means}
-
-As a result of hexagonal binning for high-D data, all the high-D data
-are clustered into hexagons. In this step, the bin centroids
-(\(C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2})\)) (see Figure~\ref{fig-meth}
-Step 2) are obtained \citep{Carr2013}.
-
-\textbf{Step 3: Triangulate bin centroids}
-
 In this step, the algorithm proceeds to triangulate the hexagonal bin
-centroids (see Figure~\ref{fig-meth} Step 3). Triangulation is a
-fundamental process in computational geometry and computer graphics that
-involves dividing a set of points in a given space into interconnected
-triangles \citep{article30}. One common algorithm used for triangulation
-is Delaunay triangulation \citep[\citet{article54}]{article26}, where
-points are connected in a way that maximizes the minimum angles of the
-resulting triangles, leading to a more regular and well-conditioned
-triangulation.
-
-Since we are working with the centroids of regular hexagonal bins, the
-resulting mesh will predominantly comprise equal-sized regular
-triangles. However, the triangulation also helps span any gaps that may
-exist between clusters of points, allowing for a more complete and
-interconnected representation of the data.
+centroids or bin means (see Figure~\ref{fig-meth} Step 4). Triangulation
+is a fundamental process in computational geometry and computer graphics
+that involves dividing a set of points in a given space into
+interconnected triangles \citep{article30}. One common algorithm used
+for triangulation is Delaunay triangulation
+\citep[\citet{article54}]{article26}, where points are connected in a
+way that maximizes the minimum angles of the resulting triangles,
+leading to a more regular and well-conditioned triangulation.
+
+Delaunay triangulation can be defined as follows:
+
+Let \(C^{(2)} = \{C_1^{(2)}, C_2^{(2)}, ..., C_m^{(2)}\}\) be a set of
+\(m\) bin centroids or bin means in the plane. Delaunay triangulation of
+\(C^{(2)}\), denoted as \(DT(C^{(2)})\), is a triangulation of the
+convex hull of \(C^{(2)}\) such that the circumcircle of every triangle
+in the triangulation contains no other points from \(C^{(2)}\).
+
+Given that the hexagons are regular, the resulting triangles will mostly
+be equilateral.
 
@@
 best representation of the original data?, (b) Model in the 2D space
-with PaCMAP (\url{https://youtu.be/CNXjWZMyh_M})}
+with PaCMAP ()}
 

### 2024-03-11 paper.qmd
Add Methodology with mathematical expressions

@@
 
-**Step1: Cluster 2D points to hexagons**
-
-Expanding upon the information regarding hexagonal binning discussed in Step 1 of @sec-construct2d, the primary objective in this step is to determine the 2D embedding points associated with each hexagon.
-As the initial process of hexagonal binning, the 2D points are clustered into their respective hexagonal bins.
-By mapping this information with the hexagonal bin centroids ($C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2})$) that obtained in Step 2 of the 2D model building (see @sec-construct2d), we can find which 2D points are assigned to which data set (see @fig-wkhighD (a)).
-
-**Step2: Cluster high dimensional points to hexagons**
-
-Following the step 1, the main focus in this step to determine the corresponding high dimensional points for each hexagon.
-Every 2D embedding point serves as a projection of a data point belonging to high dimensional space.
-By using this mapping between the high dimensional data and their corresponding projections, the high dimensional points allocated to each hexagons are determined (see video linked in @fig-wkhighD).
-
-**Step3: Compute the mean within hexagon**
-
-Having identified the high-dimensional points associated with hexagons, the final step involves computing the mean within each hexagonal bin.
-This implies calculating the average of the high-dimensional data points located within each hex bin.
-These averaged high-dimensional data points, denoted as $C_k^{(p)} \equiv (C_{kx_1}, ..., C_{kx_p})$, serve as the representative coordinates for the hex bin centroids within the expansive high-dimensional space (see video linked in @fig-wkhighD).
+Consider $f: \mathbb{R}^p \rightarrow \mathbb{R}^2$ be a function that maps the high-D data ($X_{n \times p}$) to its NLDR equivalent ($Y_{n \times d}$).
+Then, let $g: \mathbb{R}^2 \rightarrow \mathbb{R}^2$ be a function that maps each 2D embedding point to its closest centroid ($C^{(2)}$).
+It follows that $f(g(x))$ maps the high-D points $x$ to the centroid in 2D ($C_k^{(2)}$).
+Also, define a function $v: \mathbb{R}^2 \rightarrow \mathbb{R}^p$ maps the 2D centroid ($C^{(2)}$) to the high-D mean of the points ($C^{(p)}$) in the hexagon.
+
+The high-D mean of all the points in $k^{th}$ hexagon by
+
+$$
+C_k^{(p)} = (C_{kx_1}, ..., C_{kx_p}) = \left(\frac{1}{n_k} \sum_{i=1}^{n_k} x_{1i}, \frac{1}{n_k} \sum_{i=1}^{n_k} x_{2i}, \dots ,\frac{1}{n_k} \sum_{i=1}^{n_k} x_{pi}\right).
+$${#eq-equation6}
+
+Therefore,
+
+$$
+v(C_k^{(2)}) = C_k^{(p)}.
+$${#eq-equation7}
+
+Therefore, $f(g(x))$ gives the 2D centroid associated with high-D points $x$, and $v(C_k^{(2)})$ gives the high-D centroid associated with 2D point $C_k^{(2)}$.
+Thus, $v(f(g(x)))$ gives the high-D centroid ($C_k^{(p)}$) associated with the 2D embedding of the points $x$.
 
@@
 
-Based on the insights gained in Step 3 of @sec-construct2d, where connected edges in the 2D triangular mesh were identified, this step involves lifting these connections into the high dimensions.
-Having the mappings of all 2D triangular mesh points into high dimensions, the points connected in 2D are also connected in high dimensional space (see video linked in @fig-wkhighD).
+As described in Step 4 of @sec-construct2d, during the triangulation process in 2D space, vertices are identified to form edges.
+With the knowledge of the high-D mappings for the 2D hexagonal bins, the vertices connected in 2D are also connected in high-D (see video linked in @fig-wkhighD).
 
@@
 
-distance_plot <- plot_dist(distance_tsne_gau) +
-  ylab(expression(d^{(2)})) +
-  theme(axis.text = element_text(size = 5),
-        axis.title = element_text(size = 12))
-
-distance_plot
+# distance_plot <- plot_dist(distance_tsne_gau) +
+#   ylab(expression(d^{(2)})) +
+#   theme(axis.text = element_text(size = 5),
+#         axis.title = element_text(size = 12))
+# 
+# distance_plot
 
@@
 
-ggplot() +
-geom_trimesh(data = df_bin_centroids_tsne_gau, mapping = aes(x = c_x, y = c_y))
+# ggplot() +
+# geom_trimesh(data = df_bin_centroids_tsne_gau, mapping = aes(x = c_x, y = c_y))
 
@@
 
-distance_plot <- plot_dist(distance_pbmc) +
-  ylab(expression(d^{(2)})) +
-  theme(axis.text = element_text(size = 5),
-        axis.title = element_text(size = 12))
-
-distance_plot
+# distance_plot <- plot_dist(distance_pbmc) +
+#   ylab(expression(d^{(2)})) +
+#   theme(axis.text = element_text(size = 5),
+#         axis.title = element_text(size = 12))
+# 
+# distance_plot
 
@@
 
-ggplot() +
-geom_trimesh(data = df_bin_centroids_pbmc, mapping = aes(x = c_x, y = c_y))
+# ggplot() +
+# geom_trimesh(data = df_bin_centroids_pbmc, mapping = aes(x = c_x, y = c_y))
 
@@
 
-ggplot(data = hex_grid, aes(x = x, y = y)) + 
-  geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
-  geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
-  coord_fixed()
+# ggplot(data = hex_grid, aes(x = x, y = y)) + 
+#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
+#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
+#   coord_fixed()
 

### 2024-03-11 paper.tex
Add Methodology with mathematical expressions

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, enhanced, sharp corners, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden, breakable]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, breakable, boxrule=0pt, enhanced, borderline west={3pt}{0pt}{shadecolor}, interior hidden, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 
-\textbf{Step1: Cluster 2D points to hexagons}
-
-Expanding upon the information regarding hexagonal binning discussed in
-Step 1 of Section~\ref{sec-construct2d}, the primary objective in this
-step is to determine the 2D embedding points associated with each
-hexagon. As the initial process of hexagonal binning, the 2D points are
-clustered into their respective hexagonal bins. By mapping this
-information with the hexagonal bin centroids
-(\(C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2})\)) that obtained in Step 2 of
-the 2D model building (see Section~\ref{sec-construct2d}), we can find
-which 2D points are assigned to which data set (see
-\textbf{?@fig-wkhighD} (a)).
-
-\textbf{Step2: Cluster high dimensional points to hexagons}
-
-Following the step 1, the main focus in this step to determine the
-corresponding high dimensional points for each hexagon. Every 2D
-embedding point serves as a projection of a data point belonging to high
-dimensional space. By using this mapping between the high dimensional
-data and their corresponding projections, the high dimensional points
-allocated to each hexagons are determined (see video linked in
-\textbf{?@fig-wkhighD}).
-
-\textbf{Step3: Compute the mean within hexagon}
-
-Having identified the high-dimensional points associated with hexagons,
-the final step involves computing the mean within each hexagonal bin.
-This implies calculating the average of the high-dimensional data points
-located within each hex bin. These averaged high-dimensional data
-points, denoted as \(C_k^{(p)} \equiv (C_{kx_1}, ..., C_{kx_p})\), serve
-as the representative coordinates for the hex bin centroids within the
-expansive high-dimensional space (see video linked in
-\textbf{?@fig-wkhighD}).
+Consider \(f: \mathbb{R}^p \rightarrow \mathbb{R}^2\) be a function that
+maps the high-D data (\(X_{n \times p}\)) to its NLDR equivalent
+(\(Y_{n \times d}\)). Then, let
+\(g: \mathbb{R}^2 \rightarrow \mathbb{R}^2\) be a function that maps
+each 2D embedding point to its closest centroid (\(C^{(2)}\)). It
+follows that \(f(g(x))\) maps the high-D points \(x\) to the centroid in
+2D (\(C_k^{(2)}\)). Also, define a function
+\(v: \mathbb{R}^2 \rightarrow \mathbb{R}^p\) maps the 2D centroid
+(\(C^{(2)}\)) to the high-D mean of the points (\(C^{(p)}\)) in the
+hexagon.
+
+The high-D mean of all the points in \(k^{th}\) hexagon by
+
+\begin{equation}\protect\hypertarget{eq-equation6}{}{
+C_k^{(p)} = (C_{kx_1}, ..., C_{kx_p}) = \left(\frac{1}{n_k} \sum_{i=1}^{n_k} x_{1i}, \frac{1}{n_k} \sum_{i=1}^{n_k} x_{2i}, \dots ,\frac{1}{n_k} \sum_{i=1}^{n_k} x_{pi}\right).
+}\label{eq-equation6}\end{equation}
+
+Therefore,
+
+\begin{equation}\protect\hypertarget{eq-equation7}{}{
+v(C_k^{(2)}) = C_k^{(p)}.
+}\label{eq-equation7}\end{equation}
+
+Therefore, \(f(g(x))\) gives the 2D centroid associated with high-D
+points \(x\), and \(v(C_k^{(2)})\) gives the high-D centroid associated
+with 2D point \(C_k^{(2)}\). Thus, \(v(f(g(x)))\) gives the high-D
+centroid (\(C_k^{(p)}\)) associated with the 2D embedding of the points
+\(x\).
 
@@
 
-Based on the insights gained in Step 3 of Section~\ref{sec-construct2d},
-where connected edges in the 2D triangular mesh were identified, this
-step involves lifting these connections into the high dimensions. Having
-the mappings of all 2D triangular mesh points into high dimensions, the
-points connected in 2D are also connected in high dimensional space (see
+As described in Step 4 of Section~\ref{sec-construct2d}, during the
+triangulation process in 2D space, vertices are identified to form
+edges. With the knowledge of the high-D mappings for the 2D hexagonal
+bins, the vertices connected in 2D are also connected in high-D (see
 video linked in \textbf{?@fig-wkhighD}).
@@
 \end{figure}
-
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-44-1.pdf}
-
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-44-2.pdf}
-
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-44-3.pdf}
-
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-44-4.pdf}
 
@@
 
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-59-1.pdf}
-
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-59-2.pdf}
-
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-59-3.pdf}
-
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-59-4.pdf}
-
 \begin{figure}[h]
@@
 
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-65-1.pdf}
-
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-65-2.pdf}
-
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-65-3.pdf}
-
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-65-4.pdf}
-
 \begin{figure}[h]

### 2024-03-11 paper.qmd
Rewrite model summary section

@@
 
-The prediction approach involves employing the K-nearest neighbors (KNN) algorithm.
-In this method, which operates as an unsupervised classification problem, the nearest high-D model point (averaged high-D point) is identified for the new high-D data point.
-Once the nearest high-D model point is identified, its coordinates are used as the predicted 2D embedding for the new high-D data point.
-This step is particularly valuable due to the limitations of certain NLDR techniques, like tSNE, which don't provide a straightforward method for predicting 2D embedding.
-As a result, our approach presents a solution that can generate predicted 2D embedding, irrespective of the NLDR technique employed, effectively addressing this gap in functionality.
-
-The concept of "residuals" is pivotal in evaluating the accuracy of representing bin centroids in high dimensions.
-To quantify this accuracy, we introduce an error metric, which measures the sum of squared differences between the high-dimensional data ($x_{ij}$) and the predicted bin centroid data in high-dimensional space ($C_{x_ij}$) across all bins and dimensions.
-Mathematically, this error is expressed as:
+The prediction approach involves performing the K-nearest neighbors (KNN) algorithm for an unsupervised classification problem.
+First, the nearest high-D model point is identified for a given new high-D point.
+Then, the corresponding 2D centroid mapping for the identified high-D model point is determined.
+Finally, the coordinates of this 2D centroid are used as the predicted 2D embedding for the new high-D data point.
+This step is particularly valuable due to the limitations of some NLDR techniques, like tSNE, which don't provide a straightforward method for prediction.
+As a result, our approach offers a solution that capable of generating predicted 2D embedding regardless of the NLDR technique employed, effectively addressing this functional gap.
+
+Residuals are essential for evaluating the accuracy of representing high-D points by the high-D mapping of 2D bin centroids.
+To measure this accuracy, an error metric is introduced, quantifying the sum of squared differences between the high-D data ($x_{ij}$) and the high-D mapping of the 2D bin centroid data ($C_{x_ij}$) across all observations and dimensions (see @eq-equation11).
 
@@
 
-Here, $n$ represents the number of observations, $p$ represents the dimensions, $x_{ij}$ is the actual high-dimensional data, and $C_{x_ij}$ is the predicted bin centroid data in high dimensions.
+Here, $n$ represents the number of observations, $p$ represents the dimensions of high-D data, $x_{ij}$ is the high-D data, and $C_{x_ij}$ is the high-D mapping of the 2D bin centroid.
+
+<!--
+The concept of "residuals" is pivotal in evaluating the accuracy of representing bin centroids in high-D.
+To quantify this accuracy, we introduce an error metric, which measures the sum of squared differences between the high-D data ($x_{ij}$) and the predicted bin centroid data in high-D space ($C_{x_ij}$) across all bins and dimensions.
+Mathematically, this error is expressed as:
+
+$$
+\text{Error} = \sum_{j = 1}^{n}\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2
+$$ {#eq-equation11}
+
+Here, $n$ represents the number of observations, $p$ represents the dimensions, $x_{ij}$ is the high-D data, and $C_{x_ij}$ is the predicted bin centroid data in high-D.
 
@@
 This assessment is crucial for evaluating the efficacy of our NLDR technique in preserving the essential information present in the original high-dimensional data.
+-->
 
@@
 
-Moving on to the assessment of prediction accuracy, we calculate the Mean Squared Error (MSE).
-The MSE helps measure the average squared differences between the actual high-dimensional data ($x_{ij}$) and the predicted bin centroid data in high-D ($C_{x_ij}$) values across all bins.
-Mathematically, this is expressed as:
+To assess how well our method captures and represents the underlying structure of the high-D data, Mean Squared Error (MSE) and Akaike Information Criterion (AIC) are used.
+When computing MSE, total model error (see @sec-summary) is divided by the number of observations to make it as a mean value (see @eq-equation9).
 
@@
 
-Here, $b'$ signifies the total number of bins without empty bins, $p$ denotes the number of dimensions in the high-dimensional data, and $n$ represents the number of observations.
-
-Additionally, we gauge the model's performance using the Akaike Information Criterion (AIC), calculated by the formula:
+Furthermore, to incorporate with the parameters in our model, which is the non-empty bins, and as a statistical measure used for model selection and comparison, AIC is used (see @eq-equation10).
 
@@
 
-These metrics, MSE and AIC, collectively offer valuable insights into the model's predictive performance, considering both accuracy and complexity in the predictions.
+Here, $b'$ signifies the number of bins non-empty bins, $p$ denotes the number of dimensions in the high-D data, and $n$ represents the number of observations.
 
@@
 
+In conclusion, our proposed model offers a novel approach for visualizing high-dimensional (high-D) data by leveraging non-linear dimension reduction (NLDR) techniques.
+Through a series of steps including hexagonating NLDR data, triangulating bin centroids, and lifting the 2D triangular mesh into high dimensions, our model effectively transforms complex high-D data into interpretable 2D representations.
+The model's performance is evaluated using goodness of fit statistics such as Mean Squared Error (MSE) and Akaike Information Criterion (AIC), providing insights into its accuracy and reliability.
+Overall, our model presents a valuable tool for researchers and practitioners in various fields to gain deeper insights from high-dimensional datasets.
+
+Our algorithm mainly consists...
+
+These Goodness of fit statistics useful in encountering the prediction accuracy when we predict the 2D embedding for the new high-D data.
+Because, we try to find the nearest high-D mappings as the first step of prediction which will participate the prediction process.
+When we need to find which different NLDR technique and which parameter choices is giving the best representation of high-D data, MSE and AIC are also useful.
+
 We introduce a new tool to help to determine which method, which parameter choice provide the most useful representation of high-D data.

### 2024-03-11 paper.tex
Rewrite model summary section

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, breakable, boxrule=0pt, enhanced, borderline west={3pt}{0pt}{shadecolor}, interior hidden, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, sharp corners, boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable]}{\end{tcolorbox}}\fi
 
@@
 
-The prediction approach involves employing the K-nearest neighbors (KNN)
-algorithm. In this method, which operates as an unsupervised
-classification problem, the nearest high-D model point (averaged high-D
-point) is identified for the new high-D data point. Once the nearest
-high-D model point is identified, its coordinates are used as the
-predicted 2D embedding for the new high-D data point. This step is
-particularly valuable due to the limitations of certain NLDR techniques,
-like tSNE, which don't provide a straightforward method for predicting
-2D embedding. As a result, our approach presents a solution that can
-generate predicted 2D embedding, irrespective of the NLDR technique
-employed, effectively addressing this gap in functionality.
-
-The concept of ``residuals'' is pivotal in evaluating the accuracy of
-representing bin centroids in high dimensions. To quantify this
-accuracy, we introduce an error metric, which measures the sum of
-squared differences between the high-dimensional data (\(x_{ij}\)) and
-the predicted bin centroid data in high-dimensional space (\(C_{x_ij}\))
-across all bins and dimensions. Mathematically, this error is expressed
-as:
+The prediction approach involves performing the K-nearest neighbors
+(KNN) algorithm for an unsupervised classification problem. First, the
+nearest high-D model point is identified for a given new high-D point.
+Then, the corresponding 2D centroid mapping for the identified high-D
+model point is determined. Finally, the coordinates of this 2D centroid
+are used as the predicted 2D embedding for the new high-D data point.
+This step is particularly valuable due to the limitations of some NLDR
+techniques, like tSNE, which don't provide a straightforward method for
+prediction. As a result, our approach offers a solution that capable of
+generating predicted 2D embedding regardless of the NLDR technique
+employed, effectively addressing this functional gap.
+
+Residuals are essential for evaluating the accuracy of representing
+high-D points by the high-D mapping of 2D bin centroids. To measure this
+accuracy, an error metric is introduced, quantifying the sum of squared
+differences between the high-D data (\(x_{ij}\)) and the high-D mapping
+of the 2D bin centroid data (\(C_{x_ij}\)) across all observations and
+dimensions (see Equation~\ref{eq-equation11}).
 
@@
 Here, \(n\) represents the number of observations, \(p\) represents the
-dimensions, \(x_{ij}\) is the actual high-dimensional data, and
-\(C_{x_ij}\) is the predicted bin centroid data in high dimensions.
-
-The error metric outlined above provides valuable insights into the
-overall accuracy of our predictive model. By quantifying the squared
-deviations between the actual and predicted values across all bins and
-dimensions, we gain a comprehensive understanding of how well our method
-captures and represents the underlying structure of the data in the
-reduced 2D space. This assessment is crucial for evaluating the efficacy
-of our NLDR technique in preserving the essential information present in
-the original high-dimensional data.
+dimensions of high-D data, \(x_{ij}\) is the high-D data, and
+\(C_{x_ij}\) is the high-D mapping of the 2D bin centroid.
 
@@
 
-Moving on to the assessment of prediction accuracy, we calculate the
-Mean Squared Error (MSE). The MSE helps measure the average squared
-differences between the actual high-dimensional data (\(x_{ij}\)) and
-the predicted bin centroid data in high-D (\(C_{x_ij}\)) values across
-all bins. Mathematically, this is expressed as:
+To assess how well our method captures and represents the underlying
+structure of the high-D data, Mean Squared Error (MSE) and Akaike
+Information Criterion (AIC) are used. When computing MSE, total model
+error (see Section~\ref{sec-summary}) is divided by the number of
+observations to make it as a mean value (see
+Equation~\ref{eq-equation9}).
 
@@
 
-Here, \(b'\) signifies the total number of bins without empty bins,
-\(p\) denotes the number of dimensions in the high-dimensional data, and
-\(n\) represents the number of observations.
-
-Additionally, we gauge the model's performance using the Akaike
-Information Criterion (AIC), calculated by the formula:
+Furthermore, to incorporate with the parameters in our model, which is
+the non-empty bins, and as a statistical measure used for model
+selection and comparison, AIC is used (see
+Equation~\ref{eq-equation10}).
 
@@
 
-These metrics, MSE and AIC, collectively offer valuable insights into
-the model's predictive performance, considering both accuracy and
-complexity in the predictions.
+Here, \(b'\) signifies the number of bins non-empty bins, \(p\) denotes
+the number of dimensions in the high-D data, and \(n\) represents the
+number of observations.
 
@@
 
+In conclusion, our proposed model offers a novel approach for
+visualizing high-dimensional (high-D) data by leveraging non-linear
+dimension reduction (NLDR) techniques. Through a series of steps
+including hexagonating NLDR data, triangulating bin centroids, and
+lifting the 2D triangular mesh into high dimensions, our model
+effectively transforms complex high-D data into interpretable 2D
+representations. The model's performance is evaluated using goodness of
+fit statistics such as Mean Squared Error (MSE) and Akaike Information
+Criterion (AIC), providing insights into its accuracy and reliability.
+Overall, our model presents a valuable tool for researchers and
+practitioners in various fields to gain deeper insights from
+high-dimensional datasets.
+
+Our algorithm mainly consists\ldots{}
+
+These Goodness of fit statistics useful in encountering the prediction
+accuracy when we predict the 2D embedding for the new high-D data.
+Because, we try to find the nearest high-D mappings as the first step of
+prediction which will participate the prediction process. When we need
+to find which different NLDR technique and which parameter choices is
+giving the best representation of high-D data, MSE and AIC are also
+useful.
+
 We introduce a new tool to help to determine which method, which

### 2024-03-12 paper.qmd
Finalize the model for S-curve data

@@
 
-The performance and robustness of our model depend on three key parameters: (i) the total number of bins ($b$), (ii) a benchmark value used to remove low-density hexagons, and (iii) a benchmark value used to remove long edges.
-However, there is no analytical formula to calculate an appropriate value for these parameters (see Appendix for details on how default values are calculated).
+The performance and robustness of our model depend on four key parameters: (i) the total number of bins ($b$), (ii) a benchmark value used to remove low-density hexagons, (iii) a benchmark value used to remove long edges, and (iv) starting point of the hexagonal grid.
+However, there is no analytical formula to calculate an appropriate value for these parameters (the computation of default values can be find in *quollr* paper).
 The selection of these parameter values depends on the model performance computed by Mean Squared Error (MSE) (see @sec-goodfit).
@@
 
-Furthermore, the total number of bins is determined by the number of bins along the x-axis and y-axis (see @eq-equation13).
-To calculate the effective total number of bins, the minimum and approximate maximum number of bins along the x-axis are used as candidate values.
-The minimum number of bins along the x-axis is $1$.
-The maximum number of bins along the x-axis can be estimated by taking the square root of the number of NLDR data points.
-To investigate the impact of the number of bins on MSE, the analysis is focused on the MSE for different total numbers of bins.
-This range encompasses the total number of bins calculated for the minimum and maximum values of the number of bins along the x-axis.
-
-Typically, the MSE plot will have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
-Then, the slope will gradually decline or level off, indicating that a higher number of bins generates a smaller error.
+$$
+b = b_1 \times b_2
+$${#eq-equation8}
+
+Furthermore, the total number of bins is determined by the number of bins along the x-axis and y-axis (see @eq-equation8).
+To calculate the effective total number of bins, candidate values are selected based on the range between the minimum and approximate maximum number of bins along the x and y axes.
+The minimum number of bins along each axis is set to $1$, while the maximum number is estimated by taking the square root of the NLDR data point count.
+The analysis evaluates the MSE across varying total bin counts within this range, covering the minimum to maximum values along both axes.
+
+The analysis focuses on exploring the effect of total number of bins on MSE.
+It considers MSE across different total number of bins, covering the range calculated from the minimum to maximum values along both x and y axes.
+
+Typically, the MSE plot tends to have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
+Then, the slope gradually declines or levesl off, indicating that a higher number of bins generates a smaller error.
 This MSE plot is useful in determining the effective number of bins required to construct the 2D model.
-@fig-diagnosticpltScurve (b) has created a graph that shows how the Mean Squared Error (MSE) changes depending on the number of bins assigned in the hexagonal grid for various NLDR techniques applied to the S-curve data set.
-As per the rule of thumb mentioned earlier, the effective number of bins for each NLDR method is shown in the @tbl-msebinsnldr.
-Furthermore, @fig-modelScurve shows the 2D models constructed for each NLDR technique with the selected total number of bins and how the models look in high-D with original data.
+@fig-diagnosticpltScurve (b) has created a graph that shows how the Mean Squared Error (MSE) changes depending on the number of bins assigned in the hexagonal grid for UMAP applied to the S-curve dataset.
+As per the rule of thumb mentioned earlier, the effective number of bins is determined to be $84$, with $6$ bins along the x-axis and $14$ bins along the y-axis.
+Furthermore, @fig-modelScurve shows the 2D model constructed using UMAP with the selected total number of bins and how the model overlay on high-D data.
+
+```{r}
+#| warning: false
+#| echo: false
+#| message: false
+
+umap_s_curve_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = UMAP_data, 
+                                    x = "UMAP1", y = "UMAP2"))) |>
+  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
+                  "UMAP2" = "scaled_UMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(UMAP_data))
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+#| message: false
+
+## UMAP
+
+hex_size_vec <- seq(0.06, 2, by = 0.01)
+
+vec <- stats::setNames(rep("", 6), c("num_bins", "aic", "mse", "num_bins_x", "num_bins_y", "hex_size"))  ## Define column names
+
+mse_df <- dplyr::bind_rows(vec)[0, ]
+mse_df <- mse_df |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(hex_size_vec)) {
+
+  num_bin_list <- calc_bins(data = umap_s_curve_scaled, 
+            x = "UMAP1", y = "UMAP2", 
+            hex_size = hex_size_vec[i], buffer_x = NA, buffer_y = NA)
+
+  num_bins_x <- num_bin_list$num_x
+  num_bins_y <- num_bin_list$num_y
+
+  model_object <- fit_highd_model( training_data = training_data, 
+                                   nldr_df_with_id = umap_s_curve_scaled, 
+                                   x = "UMAP1", y = "UMAP2", 
+                                   num_bins_x = num_bins_x, 
+                                   num_bins_y = num_bins_y, 
+                                   x_start = NA, y_start = NA, 
+                                   buffer_x = NA, buffer_y = NA, 
+                                   hex_size = hex_size_vec[i],
+                                   is_rm_lwd_hex = FALSE, 
+                                   benchmark_to_rm_lwd_hex = NA, 
+                                   col_start_2d = "UMAP", 
+                                   col_start_highd = "x")
+
+  centroid_df_training <- model_object$df_bin_centroids
+  avg_df_training <- model_object$df_bin
+
+  pred_emb_list <- predict_emb(test_data = training_data, 
+                                  df_bin_centroids = centroid_df_training, 
+                                  df_bin = avg_df_training, type_NLDR = "UMAP")
+
+  pred_df_training <- as.data.frame(do.call(cbind, pred_emb_list))
+
+  eval_list <- gen_summary(test_data = training_data, 
+                                  prediction_df = pred_df_training, 
+                                  df_bin = avg_df_training, col_start = "x")
+
+  mse_df <- mse_df |>
+    tibble::add_row(num_bins = num_bins_x * num_bins_y,
+                    aic = eval_list$aic,
+                    mse = eval_list$mse,
+                    num_bins_x = num_bins_x,
+                    num_bins_y = num_bins_y,
+                    hex_size = hex_size_vec[i])
+
+}
+
+## If same total number of bins occurred only select ones with minimum error
+### Obtain duplicate bins
+dupli_bins <- mse_df |> 
+  dplyr::count(num_bins) |> 
+  dplyr::filter(n > 1) |> 
+  dplyr::pull(num_bins)
+
+### Group split by duplicated bins
+duplicate_df_list <- mse_df |>
+  dplyr::filter(num_bins %in% dupli_bins) |>
+  dplyr::arrange(num_bins) |>
+  dplyr::group_split(num_bins)
+
+### Obtain one row from duplicates which have lowest error and hexsize
+duplicate_df <- data.frame(matrix(nrow = 0, ncol = 0))
+
+for (i in 1:length(duplicate_df_list)) {
+
+  dd <- duplicate_df_list[[i]] |>
+    dplyr::filter(mse == min(duplicate_df_list[[i]]$mse)) |>
+    dplyr::filter(hex_size == min(duplicate_df_list[[i]]$hex_size))
+
+  duplicate_df <- dplyr::bind_rows(duplicate_df, dd)
+
+}
+
+### Obtain the mse_df with not duplicated bins
+not_dupli_df <- mse_df |>
+  dplyr::filter(!(num_bins %in% dupli_bins))
+
+### Combine duplicated and not duplicated(corrected) bins dfs
+mse_df <- dplyr::bind_rows(not_dupli_df, duplicate_df)
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## To draw with AIC
+aic_plot <- ggplot(mse_df, aes(x = num_bins, y = aic
+)) +
+  geom_point() +
+  geom_line() +
+  geom_vline(xintercept = 84, linetype="solid",
+                color = "red", size=0.8, alpha = 0.5) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  #scale_colour_manual(values = c("#377eb8")) +
+  ylab("AIC") +
+  xlab("total number of bins")
+## Effective number of bins along x-axis
+
+mse_plot <- ggplot(mse_df, aes(x = num_bins,
+                                       y = mse
+)) +
+  geom_point() +
+  geom_line() +
+   geom_vline(xintercept = 84, linetype="solid",
+                color = "red", size=0.8, alpha = 0.5) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  #scale_colour_manual(values = c("#377eb8")) +
+  ylab("MSE") +
+  xlab("total number of bins")
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset.
+What is the effective number of bins in each NLDR technique to create a 2D model?
+The MSE plot have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
+Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error.
+Using the elbow method, when the total number of bins is set to $84$, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
+This point indicates that adding less bins does not enough to capture the data structure."
+#| label: fig-diagnosticpltScurve
+##| out-width: 100%
+#| fig-pos: H
+
+aic_plot + mse_plot +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='none', plot.tag = element_text(size = 8))
+```
 
@@
 
-```{r}
-#| eval: false
-#| echo: false
-#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset.
-What is the effective number of bins in each NLDR technique to create a 2D model?
-The MSE plot have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
-Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error.
-Using the elbow method, when the total number of bins is set to 144, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
-This point indicates that adding less bins does not enough to capture the data structure."
-#| label: fig-diagnosticpltScurve
-##| out-width: 100%
-#| fig-pos: H
-
-aic_plot + mse_plot +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', ncol = 2) &
-  theme(legend.position='none', plot.tag = element_text(size = 8))
-```
 -->
@@
 
+```{r}
+#| warning: false
+#| echo: false
+#| message: false
+
+## Decide by looking at MSE plot
+num_bins_x_s_curve <- 6
+num_bins_y_s_curve <- 14
+hex_size_s_curve <- 0.13
+
+hb_obj_s_curve <- hex_binning(data = umap_s_curve_scaled, x = "UMAP1", 
+                      y = "UMAP2", num_bins_x = num_bins_x_s_curve, 
+                      num_bins_y = num_bins_y_s_curve, x_start = NA, y_start = NA, 
+                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_s_curve, col_start = "UMAP")
+
+## Data set with all possible centroids in the hexagonal grid
+all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_s_curve$centroids))
+
+## Generate all coordinates of hexagons
+hex_grid <- as.data.frame(do.call(cbind, hb_obj_s_curve$hex_poly))
+
+## To obtain the standardise counts within hexbins
+counts_df <- as.data.frame(do.call(cbind, hb_obj_s_curve$std_cts))
+df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
+                                             counts_df = counts_df)
+pcentroid_plot_s_curve <- ggplot(data = hex_grid, aes(x = x, y = y)) +
+  geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
+  geom_point(data = df_bin_centroids, aes(x = c_x, y = c_y), color = "#33a02c") +
+  coord_fixed() +
+  theme_linedraw() +
+  theme(legend.position = "bottom", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
+
+# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
+#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
+#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
+#   coord_fixed()
+
+umap_data_with_hb_id_pbmc <- as.data.frame(do.call(cbind, hb_obj_s_curve$data_hb_id))
+
+model_object <- fit_highd_model( training_data = training_data, 
+                                 nldr_df_with_id = umap_s_curve_scaled, 
+                                 x = "UMAP1", y = "UMAP2", 
+                                 num_bins_x = num_bins_x_s_curve, 
+                                 num_bins_y = num_bins_y_s_curve, 
+                                 x_start = NA, y_start = NA, 
+                                 buffer_x = NA, buffer_y = NA, 
+                                 hex_size = hex_size_s_curve,
+                                 is_rm_lwd_hex = FALSE, 
+                                 benchmark_to_rm_lwd_hex = NA, 
+                                 col_start_2d = "UMAP", 
+                                 col_start_highd = "x")
+
+df_bin_centroids_s_curve <- model_object$df_bin_centroids
+df_bin_s_curve <- model_object$df_bin
+
+## Triangulate bin centroids
+tr1_object_s_curve <- tri_bin_centroids(df_bin_centroids_s_curve, x = "c_x", y = "c_y")
+tr_from_to_df_s_curve <- gen_edges(tri_object = tr1_object_s_curve)
+
+## Compute 2D distances
+distance_s_curve <- cal_2d_dist(tr_coord_df = tr_from_to_df_s_curve, 
+                             start_x = "x_from", start_y = "y_from", 
+                             end_x = "x_to", end_y = "y_to", 
+                             select_vars = c("from", "to", "distance"))
+
+# distance_plot <- plot_dist(distance_pbmc) +
+#   ylab(expression(d^{(2)})) +
+#   theme(axis.text = element_text(size = 5),
+#         axis.title = element_text(size = 12))
+# 
+# distance_plot
+
+## To find the benchmark value
+benchmark_s_curve <- find_lg_benchmark(distance_edges = distance_s_curve, distance_col = "distance")
+
+# ggplot() +
+# geom_trimesh(data = df_bin_centroids_pbmc, mapping = aes(x = c_x, y = c_y))
+
+trimesh_s_curve_umap <- vis_lg_mesh(distance_edges = distance_s_curve, benchmark_value = benchmark_s_curve, tr_coord_df = tr_from_to_df_s_curve, distance_col = "distance")
+
+trimesh_s_curve_umap <- trimesh_s_curve_umap +
+  theme_linedraw() +
+  theme(legend.position = "bottom", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) +
+  labs(colour = "")
+
+trimesh_removed_s_curve_umap <- vis_rmlg_mesh(distance_edges = distance_s_curve, benchmark_value = benchmark_s_curve, tr_coord_df = tr_from_to_df_s_curve, distance_col = "distance")
+
+trimesh_removed_s_curve_umap <- trimesh_removed_s_curve_umap +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "(a), (b) 2D model."
+#| label: fig-modelScurve
+
+pcentroid_plot_s_curve + trimesh_s_curve_umap + trimesh_removed_s_curve_umap +
+  plot_layout(guides='collect', ncol = 3) &
+  theme(legend.position='bottom') 
+```
+
 #### Benchmark value to remove long edges
@@
 #| echo: false
-#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset.
+#| fig-cap: "Goodness of fit statistics from UMAP applied to training PBMC3k dataset.
 What is the effective number of bins in each NLDR technique to create a 2D model?
@@
   # geom_text(aes(x=UMAP1,y = UMAP2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) +
+  labs(colour = "cluster labels")
 
@@
   theme_linedraw() +
-    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+    theme(legend.position = "bottom", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
               axis.title.x = element_blank(), axis.title.y = element_blank(),
@@
   # geom_text(aes(x=UMAP1,y = UMAP2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) +
+  labs(colour = "cluster labels")
 
@@
 
-result1_2d_layout + result2_2d_layout
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "2D layout from UMAP (n_neighbors = 5, min_dist = 0.99, metric = cosine) applied for PBMC3k dataset (a) colored by manualy identified clusters, (b) colored by `detour` results"
+#| label: fig-nldervisPBMCUMAP3Res
+#| out-width: 100%
+
+result1_2d_layout + result2_2d_layout +
+  plot_layout(guides='collect', ncol=2) &
+  theme(legend.position='bottom')
+
 ```
@@
 # show_langevitour(df_all_mnist, df_bin_mnist, df_bin_centroids_mnist, benchmark_mnist, distance_mnist, "distance", col_start = "PC")
+
+predict_minst_df <- as.data.frame(do.call(cbind, predict_emb(training_data_mnist, df_bin_centroids_mnist, df_bin_mnist, "PaCMAP")))
+gen_summary(training_data_mnist, predict_minst_df, df_bin_mnist, col_start = "PC")
 ```

### 2024-03-12 paper.tex
Finalize the model for S-curve data

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, sharp corners, boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, breakable, sharp corners, enhanced, frame hidden, borderline west={3pt}{0pt}{shadecolor}, interior hidden]}{\end{tcolorbox}}\fi
 
@@
 
-The performance and robustness of our model depend on three key
+The performance and robustness of our model depend on four key
 parameters: (i) the total number of bins (\(b\)), (ii) a benchmark value
-used to remove low-density hexagons, and (iii) a benchmark value used to
-remove long edges. However, there is no analytical formula to calculate
-an appropriate value for these parameters (see Appendix for details on
-how default values are calculated). The selection of these parameter
-values depends on the model performance computed by Mean Squared Error
-(MSE) (see Section~\ref{sec-goodfit}).
+used to remove low-density hexagons, (iii) a benchmark value used to
+remove long edges, and (iv) starting point of the hexagonal grid.
+However, there is no analytical formula to calculate an appropriate
+value for these parameters (the computation of default values can be
+find in \emph{quollr} paper). The selection of these parameter values
+depends on the model performance computed by Mean Squared Error (MSE)
+(see Section~\ref{sec-goodfit}).
 
@@
 
+\begin{equation}\protect\hypertarget{eq-equation8}{}{
+b = b_1 \times b_2
+}\label{eq-equation8}\end{equation}
+
 Furthermore, the total number of bins is determined by the number of
-bins along the x-axis and y-axis (see \textbf{?@eq-equation13}). To
-calculate the effective total number of bins, the minimum and
-approximate maximum number of bins along the x-axis are used as
-candidate values. The minimum number of bins along the x-axis is \(1\).
-The maximum number of bins along the x-axis can be estimated by taking
-the square root of the number of NLDR data points. To investigate the
-impact of the number of bins on MSE, the analysis is focused on the MSE
-for different total numbers of bins. This range encompasses the total
-number of bins calculated for the minimum and maximum values of the
-number of bins along the x-axis.
-
-Typically, the MSE plot will have a steep slope at the beginning,
+bins along the x-axis and y-axis (see Equation~\ref{eq-equation8}). To
+calculate the effective total number of bins, candidate values are
+selected based on the range between the minimum and approximate maximum
+number of bins along the x and y axes. The minimum number of bins along
+each axis is set to \(1\), while the maximum number is estimated by
+taking the square root of the NLDR data point count. The analysis
+evaluates the MSE across varying total bin counts within this range,
+covering the minimum to maximum values along both axes.
+
+The analysis focuses on exploring the effect of total number of bins on
+MSE. It considers MSE across different total number of bins, covering
+the range calculated from the minimum to maximum values along both x and
+y axes.
+
+Typically, the MSE plot tends to have a steep slope at the beginning,
 indicating that a smaller number of bins causes a larger amount of
-error. Then, the slope will gradually decline or level off, indicating
-that a higher number of bins generates a smaller error. This MSE plot is
+error. Then, the slope gradually declines or levesl off, indicating that
+a higher number of bins generates a smaller error. This MSE plot is
 useful in determining the effective number of bins required to construct
-the 2D model. \textbf{?@fig-diagnosticpltScurve} (b) has created a graph
-that shows how the Mean Squared Error (MSE) changes depending on the
-number of bins assigned in the hexagonal grid for various NLDR
-techniques applied to the S-curve data set. As per the rule of thumb
-mentioned earlier, the effective number of bins for each NLDR method is
-shown in the \textbf{?@tbl-msebinsnldr}. Furthermore,
-\textbf{?@fig-modelScurve} shows the 2D models constructed for each NLDR
-technique with the selected total number of bins and how the models look
-in high-D with original data.
+the 2D model. Figure~\ref{fig-diagnosticpltScurve} (b) has created a
+graph that shows how the Mean Squared Error (MSE) changes depending on
+the number of bins assigned in the hexagonal grid for UMAP applied to
+the S-curve dataset. As per the rule of thumb mentioned earlier, the
+effective number of bins is determined to be \(84\), with \(6\) bins
+along the x-axis and \(14\) bins along the y-axis. Furthermore,
+Figure~\ref{fig-modelScurve} shows the 2D model constructed using UMAP
+with the selected total number of bins and how the model overlay on
+high-D data.
+
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-diagnosticpltScurve-1.pdf}
+
+}
+
+\caption{\label{fig-diagnosticpltScurve}Goodness of fit statistics from
+UMAP applied to training S-curve dataset. What is the effective number
+of bins in each NLDR technique to create a 2D model? The MSE plot have a
+steep slope at the beginning, indicating that a smaller number of bins
+causes a larger amount of error. Then, the slope gradually declines or
+level off, indicating that a higher number of bins generates a smaller
+error. Using the elbow method, when the total number of bins is set to
+\(84\), the slope of the Mean Squared Error (MSE) plot experiences a
+sudden and noticeable change, resembling an elbow-like shape. This point
+indicates that adding less bins does not enough to capture the data
+structure.}
+
+\end{figure}
 
@@
 case, a benchmark value of \(0.06\) was chosen. The resulting 2D model
-is shown in \textbf{?@fig-modelScurve} (c), while the video link
+is shown in Figure~\ref{fig-modelScurve} (c), while the video link
 provides a visual representation of how the model appears in high-D.
@@
 \end{enumerate}
+
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-modelScurve-1.pdf}
+
+}
+
+\caption{\label{fig-modelScurve}(a), (b) 2D model.}
+
+\end{figure}
 
@@
 \caption{\label{fig-diagnosticpltPBMC}Goodness of fit statistics from
-UMAP applied to training S-curve dataset. What is the effective number
-of bins in each NLDR technique to create a 2D model? The MSE plot have a
+UMAP applied to training PBMC3k dataset. What is the effective number of
+bins in each NLDR technique to create a 2D model? The MSE plot have a
 steep slope at the beginning, indicating that a smaller number of bins
@@
 
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-68-1.pdf}
+\begin{figure}
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervisPBMCUMAP3Res-1.pdf}
+
+}
+
+\caption{\label{fig-nldervisPBMCUMAP3Res}2D layout from UMAP
+(n\_neighbors = 5, min\_dist = 0.99, metric = cosine) applied for PBMC3k
+dataset (a) colored by manualy identified clusters, (b) colored by
+\texttt{detour} results}
+
+\end{figure}
 
@@
 \end{figure}
+
+\begin{verbatim}
+$mse
+[1] 3.017065
+
+$aic
+[1] 87724.49
+\end{verbatim}
 

### 2024-03-12 paper.qmd
Added images of handwritten digits

@@
 library(readr)
+library(tidyr)
 
@@
 $$
- y_{max} = \text{ceiling}(\frac{ar}{hr}) * hr
+ y_{max} = \left\lceil\frac{ar}{hr}\right\rceil * hr
 $$ {#eq-equation3}
@@
 #| echo: false
-#| fig-cap: "2D layout from UMAP (n_neighbors = 5, min_dist = 0.99, metric = cosine) applied for PBMC3k dataset (a) colored by manualy identified clusters, (b) colored by `detour` results"
+#| fig-cap: "2D layout from UMAP (n_neighbors = 5, min_dist = 0.99, metric = cosine) applied for PBMC3k dataset (a) colored manually in 2D, (b) colored in high-D"
 #| label: fig-nldervisPBMCUMAP3Res
@@
 
+- Selected 10 PCs based on scree plot
+
+- 7877 images of digit 1
+
 - PaCMAP parameters: n_components=2, n_neighbors=10, init="random", MN_ratio=0.9, FP_ratio=2.0
+
+#### Model
+
+- Total number of bins used: 180 (num_x = 12, num_y = 15, hex_size = 0.06)
+
+- Learned from the model: There is a non-linear shape in 2D, but if look at the model in high-D, the non-linear continuity is there but there is a twist in the structure which didn't capture by PaCMAP.
+
+- Write about looking at images of digit 1 as well..how in 2D space images are distributed, which images occur high-errors in high-D
 
@@
          legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
   # geom_text(aes(x=UMAP1,y = UMAP2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
+  annotate(geom = 'text', label = 'a', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3) 
 ```
@@
   geom_line() +
-  geom_vline(xintercept = 80, linetype="solid",
+  geom_vline(xintercept = 180, linetype="solid",
                 color = "red", size=0.8, alpha = 0.5) +
@@
   geom_line() +
-   geom_vline(xintercept = 80, linetype="solid",
+   geom_vline(xintercept = 180, linetype="solid",
                 color = "red", size=0.8, alpha = 0.5) +
@@
 Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error.
-Using the elbow method, when the total number of bins is set to 80, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
+Using the elbow method, when the total number of bins is set to 180, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
 This point indicates that adding less bins does not enough to capture the data structure."
@@
 ## Decide by looking at MSE plot
-num_bins_x_mnist <- 8
-num_bins_y_mnist <- 10
-hex_size_mnist <- 0.1
+num_bins_x_mnist <- 12
+num_bins_y_mnist <- 15
+hex_size_mnist <- 0.06
 
@@
         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+  annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
 
@@
                                         if_else(total <= 6, "error 4-6",
-                                                if_else(total <= 8, "error 6-8",
-                                                        if_else(total <= 12, "error 8-10",
-                                                                if_else(total <= 14, "error 10-12", "error greter than 12"))))))))
+                                                if_else(total <= 10, "error 6-10",
+                                                        if_else(total <= 12, "error 10-12",
+                                                                if_else(total <= 14, "error 12-14", "error greter than 14"))))))))
 
 plot_list2_mnist_error <- error_df |>
-  mutate(type = factor(type , levels = c("no error", "error 0-2", "error 2-4", "error 4-6", "error 6-8", "error 8-10", "error 10-12", "error greter than 12"))) |>
+  mutate(type = factor(type , levels = c("no error", "error 0-2", "error 2-4", "error 4-6", "error 6-10", "error 10-12", "error 12-14", "error greter than 14"))) |>
     ggplot(aes(x = PaCMAP1,
@@
                                          "#fc9272", "#fb6a4a", "#ef3b2c",
-                                         "#cb181d", "#a50f15", "#99000d"))
+                                         "#cb181d", "#a50f15", "#99000d")) +
+  annotate(geom = 'text', label = 'c', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
 
@@
 
+<!--images with large error in high-D-->
+
+```{r}
+#| warning: false
+#| echo: false
+
+img_error <- error_df |> 
+  dplyr::filter(total >= 10) |> 
+  dplyr::pull(ID)
+
+mnist_data <- read_rds("data/mnist/mnist_digit_1.rds")
+pixels_gathered <-  mnist_data |>
+  mutate(instance = row_number()) |>
+  gather(pixel, value, -Label, -instance) |>
+  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
+  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
+  filter(instance %in% img_error)
+
+theme_set(theme_light())
+imge_error_sample <- pixels_gathered |> 
+  ggplot(aes(x, y, fill = value)) +  
+  geom_tile() +  
+  facet_wrap(~ instance) +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm'),
+        strip.text = element_blank())
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "Images of handwritten digit 1 which occur large model error."
+#| label: fig-sampleimg
+#| out-width: 100%
+#| fig-pos: H
+
+imge_error_sample
+```
+
 ## Conclusion {#sec-conclusion}

### 2024-03-12 paper.tex
Added images of handwritten digits

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, breakable, sharp corners, enhanced, frame hidden, borderline west={3pt}{0pt}{shadecolor}, interior hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, sharp corners, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, interior hidden, breakable, enhanced]}{\end{tcolorbox}}\fi
 
@@
 \begin{equation}\protect\hypertarget{eq-equation3}{}{
- y_{max} = \text{ceiling}(\frac{ar}{hr}) * hr
+ y_{max} = \left\lceil\frac{ar}{hr}\right\rceil * hr
 }\label{eq-equation3}\end{equation}
@@
 (n\_neighbors = 5, min\_dist = 0.99, metric = cosine) applied for PBMC3k
-dataset (a) colored by manualy identified clusters, (b) colored by
-\texttt{detour} results}
+dataset (a) colored manually in 2D, (b) colored in high-D}
 
@@
 \item
+  Selected 10 PCs based on scree plot
+\item
+  7877 images of digit 1
+\item
   PaCMAP parameters: n\_components=2, n\_neighbors=10, init=``random'',
   MN\_ratio=0.9, FP\_ratio=2.0
+\end{itemize}
+
+\hypertarget{model}{%
+\subsubsection{Model}\label{model}}
+
+\begin{itemize}
+\item
+  Total number of bins used: 180 (num\_x = 12, num\_y = 15, hex\_size =
+  0.06)
+\item
+  Learned from the model: There is a non-linear shape in 2D, but if look
+  at the model in high-D, the non-linear continuity is there but there
+  is a twist in the structure which didn't capture by PaCMAP.
+\item
+  Write about looking at images of digit 1 as well..how in 2D space
+  images are distributed, which images occur high-errors in high-D
 \end{itemize}
@@
 error. Using the elbow method, when the total number of bins is set to
-80, the slope of the Mean Squared Error (MSE) plot experiences a sudden
+180, the slope of the Mean Squared Error (MSE) plot experiences a sudden
 and noticeable change, resembling an elbow-like shape. This point
@@
 $mse
-[1] 3.017065
+[1] 2.276023
 
 $aic
-[1] 87724.49
+[1] 66382.78
 \end{verbatim}
@@
 with PaCMAP ()}
+
+\end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-sampleimg-1.pdf}
+
+}
+
+\caption{\label{fig-sampleimg}Images of handwritten digit 1 which occur
+large model error.}
 

### 2024-03-13 paper.qmd
Added approached followed for PBMC3k

@@
 
+### Approach followed
+
+- First, referred to @Chen2023 paper which shows UMAP representation of PBMC3k data with 15 PCs and parameter choice of n:neighbor:30, min_dist:0.1, metric:cosine.
+
+- Observed that there is three well-separated clusters.
+
+- Second, visualise the 15 PCs in high-dimensional space with our model and investigate whether can I see the three well-separated clusters or is it showing something different than what I saw in UMAP view suggested by the author.
+
+- Learned from the model: Yes, it's different.
+There are three clusters but they are really close to each other.
+Also, in the largest cluster, observed non-linear continuity structure, which didn't captured by the author's UMAP representation.
+
+- Learned from the data: Visualise with detour and colored the clusters in high-D. Identified more than three clusters which is suprising.
+
+- In that case, need to suggest a better UMAP representation than author's suggestion.
+
+- To find a better UMAP representation, I need to investigate different parameter choices for UMAP (or need to try another NLDR technique, but here my objective is to suggest the best parameter choice with UMAP).
+
+- Therefore, I fitted our model for different UMAP parameter choices, record the error, and selected the parameter choice that gives the lowest error which is n:neighbor:5, min_dist:0.99, metric:cosine.
+
+- Then, construct our model for this UMAP data (n:neighbor:5, min_dist:0.99, metric:cosine).
+
+- In this representation, you can see that the clusters are closer, and the continuity of clusters are there.
+
+- Also, what missed from looking at high-D data with detour, in this UMAP representation I can see more than three clusters.
+
+- Then, need to check how correct am I identify the clusters in 2D and is that matched that I identify in high-D.
+
+- Generate a confusion matrix.
+
 ```{r}
@@
 ```
+
+<!--obtain 2D and high-D distances-->
+
+<!---
+```{r}
+#| warning: false
+#| echo: false
+#| message: false
+#| eval: false
+
+## Obtian 2D distances
+proxy::dist(umap_pbmc[, -3]) |> 
+as.data.frame()
+
+## Obtain high-D distances
+proxy::dist(training_data_pbmc[, -16]) |> 
+as.data.frame()
+```
+--->
 

### 2024-03-13 paper.tex
Added approached followed for PBMC3k

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, sharp corners, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, interior hidden, breakable, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, interior hidden, boxrule=0pt, sharp corners, enhanced, borderline west={3pt}{0pt}{shadecolor}, breakable]}{\end{tcolorbox}}\fi
 
@@
   the fit is poor in high-dimensions.
+\end{itemize}
+
+\hypertarget{approach-followed}{%
+\subsection{Approach followed}\label{approach-followed}}
+
+\begin{itemize}
+\item
+  First, referred to \citet{Chen2023} paper which shows UMAP
+  representation of PBMC3k data with 15 PCs and parameter choice of
+  n:neighbor:30, min\_dist:0.1, metric:cosine.
+\item
+  Observed that there is three well-separated clusters.
+\item
+  Second, visualise the 15 PCs in high-dimensional space with our model
+  and investigate whether can I see the three well-separated clusters or
+  is it showing something different than what I saw in UMAP view
+  suggested by the author.
+\item
+  Learned from the model: Yes, it's different. There are three clusters
+  but they are really close to each other. Also, in the largest cluster,
+  observed non-linear continuity structure, which didn't captured by the
+  author's UMAP representation.
+\item
+  Learned from the data: Visualise with detour and colored the clusters
+  in high-D. Identified more than three clusters which is suprising.
+\item
+  In that case, need to suggest a better UMAP representation than
+  author's suggestion.
+\item
+  To find a better UMAP representation, I need to investigate different
+  parameter choices for UMAP (or need to try another NLDR technique, but
+  here my objective is to suggest the best parameter choice with UMAP).
+\item
+  Therefore, I fitted our model for different UMAP parameter choices,
+  record the error, and selected the parameter choice that gives the
+  lowest error which is n:neighbor:5, min\_dist:0.99, metric:cosine.
+\item
+  Then, construct our model for this UMAP data (n:neighbor:5,
+  min\_dist:0.99, metric:cosine).
+\item
+  In this representation, you can see that the clusters are closer, and
+  the continuity of clusters are there.
+\item
+  Also, what missed from looking at high-D data with detour, in this
+  UMAP representation I can see more than three clusters.
+\item
+  Then, need to check how correct am I identify the clusters in 2D and
+  is that matched that I identify in high-D.
+\item
+  Generate a confusion matrix.
 \end{itemize}

### 2024-03-18 paper.qmd
Added model interpretations for MNIST digit 1 data set

@@
 ## Decide by looking at MSE plot
-num_bins_x_tsne_gau <- 8
-num_bins_y_tsne_gau <- 10
-hex_size_tsne_gau <- 0.1
+num_bins_x_tsne_gau <- 21
+num_bins_y_tsne_gau <- 28
+hex_size_tsne_gau <- 0.03
+## non-empty:198
 
@@
 
+## Decide by looking at MSE plot
+num_bins_x_tsne_gau <- 60
+num_bins_y_tsne_gau <- 79
+hex_size_tsne_gau <- 0.01
+### non-empty 229
+
+hb_obj_tsne_gau <- hex_binning(data = umap_gau_scaled, x = "UMAP1", 
+                      y = "UMAP2", num_bins_x = num_bins_x_tsne_gau, 
+                      num_bins_y = num_bins_y_tsne_gau, x_start = NA, y_start = NA, 
+                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_tsne_gau, col_start = "UMAP")
+
+## Data set with all possible centroids in the hexagonal grid
+all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$centroids))
+
+## Generate all coordinates of hexagons
+hex_grid <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$hex_poly))
+
+## To obtain the standardise counts within hexbins
+counts_df <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$std_cts))
+df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
+                                             counts_df = counts_df)
+# ggplot(data = hex_grid, aes(x = x, y = y)) +
+#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
+#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
+#   coord_fixed()
+
+hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
+
+# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
+#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
+#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
+#   coord_fixed()
+
+tsne_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$data_hb_id))
+
+model_object_tsne_gau <- fit_highd_model( training_data = training_data_gau, 
+                                 nldr_df_with_id = umap_gau_scaled, 
+                                 x = "UMAP1", y = "UMAP2", 
+                                 num_bins_x = num_bins_x_tsne_gau, 
+                                 num_bins_y = num_bins_y_tsne_gau, 
+                                 x_start = NA, y_start = NA, 
+                                 buffer_x = NA, buffer_y = NA, 
+                                 hex_size = hex_size_tsne_gau,
+                                 is_rm_lwd_hex = FALSE, 
+                                 benchmark_to_rm_lwd_hex = NA, 
+                                 col_start_2d = "UMAP", 
+                                 col_start_highd = "x")
+
+df_bin_centroids_tsne_gau <- model_object_tsne_gau$df_bin_centroids
+df_bin_tsne_gau <- model_object_tsne_gau$df_bin
+
+## Triangulate bin centroids
+tr1_object_tsne_gau <- tri_bin_centroids(df_bin_centroids_tsne_gau, x = "c_x", y = "c_y")
+tr_from_to_df_tsne_gau <- gen_edges(tri_object = tr1_object_tsne_gau)
+
+## Compute 2D distances
+distance_tsne_gau <- cal_2d_dist(tr_coord_df = tr_from_to_df_tsne_gau, 
+                             start_x = "x_from", start_y = "y_from", 
+                             end_x = "x_to", end_y = "y_to", 
+                             select_vars = c("from", "to", "distance"))
+
+# distance_plot <- plot_dist(distance_tsne_gau) +
+#   ylab(expression(d^{(2)})) +
+#   theme(axis.text = element_text(size = 5),
+#         axis.title = element_text(size = 12))
+# 
+# distance_plot
+
+## To find the benchmark value
+benchmark_tsne_gau <- find_lg_benchmark(distance_edges = distance_tsne_gau, distance_col = "distance")
+
+# ggplot() +
+# geom_trimesh(data = df_bin_centroids_tsne_gau, mapping = aes(x = c_x, y = c_y))
+
+trimesh_removed_tsne_gau <- vis_rmlg_mesh(distance_edges = distance_tsne_gau, benchmark_value = benchmark_tsne_gau, tr_coord_df = tr_from_to_df_tsne_gau, distance_col = "distance")
+
+trimesh_removed_tsne_gau <- trimesh_removed_tsne_gau +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+df_all_tsne_gau <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID), tsne_data_with_hb_id_gau)
+
+# show_langevitour(df = df_all_tsne_gau, df_b = df_bin_tsne_gau, df_b_with_center_data = df_bin_centroids_tsne_gau, col_start = "x", distance_df = distance_tsne_gau, distance_col = "distance", benchmark_value = benchmark_tsne_gau)
+
+  # find_non_empty_bins(data = umap_gau_scaled, x = "UMAP1", y = "UMAP2",
+  #                   non_empty_bins = 198, x_start = NA, y_start = NA,
+  #                   buffer_x = NA, buffer_y = NA, hex_size = 0.03,
+  #                   col_start = "UMAP")
+
 # hex_size_vec <- seq(0.02, 2, by = 0.01)
@@
 #   find_non_empty_bins(data = umap_gau_scaled, x = "UMAP1", y = "UMAP2",
-#                     non_empty_bins = 34, x_start = NA, y_start = NA,
+#                     non_empty_bins = 198, x_start = NA, y_start = NA,
 #                     buffer_x = NA, buffer_y = NA, hex_size = hex_size_vec[i],
@@
 # }
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+#| message: false
+#| eval: false
+
+phate_gau_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = PHATE_data_gau, x = "PAHTE1", y = "PAHTE2"))) |>
+  dplyr::rename(c("PAHTE1" = "scaled_PAHTE1", 
+                  "PAHTE2" = "scaled_PAHTE2")) |>
+  dplyr::mutate(ID = 1:NROW(PAHTE_data_gau))
+
+## Decide by looking at MSE plot
+num_bins_x_tsne_gau <- 46
+num_bins_y_tsne_gau <- 61
+hex_size_tsne_gau <- 0.013
+### non-empty 179
+
+hb_obj_tsne_gau <- hex_binning(data = phate_gau_scaled, x = "PHATE1", 
+                      y = "PHATE2", num_bins_x = num_bins_x_tsne_gau, 
+                      num_bins_y = num_bins_y_tsne_gau, x_start = NA, y_start = NA, 
+                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_tsne_gau, col_start = "PHATE")
+
+## Data set with all possible centroids in the hexagonal grid
+all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$centroids))
+
+## Generate all coordinates of hexagons
+hex_grid <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$hex_poly))
+
+## To obtain the standardise counts within hexbins
+counts_df <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$std_cts))
+df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
+                                             counts_df = counts_df)
+# ggplot(data = hex_grid, aes(x = x, y = y)) +
+#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
+#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
+#   coord_fixed()
+
+hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
+
+# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
+#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
+#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
+#   coord_fixed()
+
+tsne_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$data_hb_id))
+
+model_object_tsne_gau <- fit_highd_model( training_data = training_data_gau, 
+                                 nldr_df_with_id = phate_gau_scaled, 
+                                 x = "PHATE1", y = "PHATE2", 
+                                 num_bins_x = num_bins_x_tsne_gau, 
+                                 num_bins_y = num_bins_y_tsne_gau, 
+                                 x_start = NA, y_start = NA, 
+                                 buffer_x = NA, buffer_y = NA, 
+                                 hex_size = hex_size_tsne_gau,
+                                 is_rm_lwd_hex = FALSE, 
+                                 benchmark_to_rm_lwd_hex = NA, 
+                                 col_start_2d = "PHATE", 
+                                 col_start_highd = "x")
+
+df_bin_centroids_tsne_gau <- model_object_tsne_gau$df_bin_centroids
+df_bin_tsne_gau <- model_object_tsne_gau$df_bin
+
+## Triangulate bin centroids
+tr1_object_tsne_gau <- tri_bin_centroids(df_bin_centroids_tsne_gau, x = "c_x", y = "c_y")
+tr_from_to_df_tsne_gau <- gen_edges(tri_object = tr1_object_tsne_gau)
+
+## Compute 2D distances
+distance_tsne_gau <- cal_2d_dist(tr_coord_df = tr_from_to_df_tsne_gau, 
+                             start_x = "x_from", start_y = "y_from", 
+                             end_x = "x_to", end_y = "y_to", 
+                             select_vars = c("from", "to", "distance"))
+
+# distance_plot <- plot_dist(distance_tsne_gau) +
+#   ylab(expression(d^{(2)})) +
+#   theme(axis.text = element_text(size = 5),
+#         axis.title = element_text(size = 12))
+# 
+# distance_plot
+
+## To find the benchmark value
+benchmark_tsne_gau <- find_lg_benchmark(distance_edges = distance_tsne_gau, distance_col = "distance")
+
+# ggplot() +
+# geom_trimesh(data = df_bin_centroids_tsne_gau, mapping = aes(x = c_x, y = c_y))
+
+trimesh_removed_tsne_gau <- vis_rmlg_mesh(distance_edges = distance_tsne_gau, benchmark_value = benchmark_tsne_gau, tr_coord_df = tr_from_to_df_tsne_gau, distance_col = "distance")
+
+trimesh_removed_tsne_gau <- trimesh_removed_tsne_gau +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+df_all_tsne_gau <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID), tsne_data_with_hb_id_gau)
+
+# show_langevitour(df = df_all_tsne_gau, df_b = df_bin_tsne_gau, df_b_with_center_data = df_bin_centroids_tsne_gau, col_start = "x", distance_df = distance_tsne_gau, distance_col = "distance", benchmark_value = benchmark_tsne_gau)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+#| message: false
+
+trimap_gau_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = TriMAP_data_gau, 
+                                    x = "TriMAP1", y = "TriMAP2"))) |>
+  dplyr::rename(c("TriMAP1" = "scaled_TriMAP1", 
+                  "TriMAP2" = "scaled_TriMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(TriMAP_data_gau))
+
+## Decide by looking at MSE plot
+num_bins_x_tsne_gau <- 46
+num_bins_y_tsne_gau <- 61
+hex_size_tsne_gau <- 0.013
+### non-empty 179
+
+hb_obj_tsne_gau <- hex_binning(data = trimap_gau_scaled, x = "TriMAP1", 
+                      y = "TriMAP2", num_bins_x = num_bins_x_tsne_gau, 
+                      num_bins_y = num_bins_y_tsne_gau, x_start = NA, y_start = NA, 
+                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_tsne_gau, col_start = "TriMAP")
+
+## Data set with all possible centroids in the hexagonal grid
+all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$centroids))
+
+## Generate all coordinates of hexagons
+hex_grid <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$hex_poly))
+
+## To obtain the standardise counts within hexbins
+counts_df <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$std_cts))
+df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
+                                             counts_df = counts_df)
+# ggplot(data = hex_grid, aes(x = x, y = y)) +
+#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
+#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
+#   coord_fixed()
+
+hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
+
+# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
+#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
+#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
+#   coord_fixed()
+
+tsne_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$data_hb_id))
+
+model_object_tsne_gau <- fit_highd_model( training_data = training_data_gau, 
+                                 nldr_df_with_id = trimap_gau_scaled, 
+                                 x = "TriMAP1", y = "TriMAP2", 
+                                 num_bins_x = num_bins_x_tsne_gau, 
+                                 num_bins_y = num_bins_y_tsne_gau, 
+                                 x_start = NA, y_start = NA, 
+                                 buffer_x = NA, buffer_y = NA, 
+                                 hex_size = hex_size_tsne_gau,
+                                 is_rm_lwd_hex = FALSE, 
+                                 benchmark_to_rm_lwd_hex = NA, 
+                                 col_start_2d = "TriMAP", 
+                                 col_start_highd = "x")
+
+df_bin_centroids_tsne_gau <- model_object_tsne_gau$df_bin_centroids
+df_bin_tsne_gau <- model_object_tsne_gau$df_bin
+
+## Triangulate bin centroids
+tr1_object_tsne_gau <- tri_bin_centroids(df_bin_centroids_tsne_gau, x = "c_x", y = "c_y")
+tr_from_to_df_tsne_gau <- gen_edges(tri_object = tr1_object_tsne_gau)
+
+## Compute 2D distances
+distance_tsne_gau <- cal_2d_dist(tr_coord_df = tr_from_to_df_tsne_gau, 
+                             start_x = "x_from", start_y = "y_from", 
+                             end_x = "x_to", end_y = "y_to", 
+                             select_vars = c("from", "to", "distance"))
+
+# distance_plot <- plot_dist(distance_tsne_gau) +
+#   ylab(expression(d^{(2)})) +
+#   theme(axis.text = element_text(size = 5),
+#         axis.title = element_text(size = 12))
+# 
+# distance_plot
+
+## To find the benchmark value
+benchmark_tsne_gau <- find_lg_benchmark(distance_edges = distance_tsne_gau, distance_col = "distance")
+
+# ggplot() +
+# geom_trimesh(data = df_bin_centroids_tsne_gau, mapping = aes(x = c_x, y = c_y))
+
+trimesh_removed_tsne_gau <- vis_rmlg_mesh(distance_edges = distance_tsne_gau, benchmark_value = benchmark_tsne_gau, tr_coord_df = tr_from_to_df_tsne_gau, distance_col = "distance")
+
+trimesh_removed_tsne_gau <- trimesh_removed_tsne_gau +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+df_all_tsne_gau <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID), tsne_data_with_hb_id_gau)
+
+# show_langevitour(df = df_all_tsne_gau, df_b = df_bin_tsne_gau, df_b_with_center_data = df_bin_centroids_tsne_gau, col_start = "x", distance_df = distance_tsne_gau, distance_col = "distance", benchmark_value = benchmark_tsne_gau)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+#| message: false
+
+pacmap_gau_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = PaCMAP_data_gau, 
+                                    x = "PaCMAP1", y = "PaCMAP2"))) |>
+  dplyr::rename(c("PaCMAP1" = "scaled_PaCMAP1", 
+                  "PaCMAP2" = "scaled_PaCMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(PaCMAP_data_gau))
+
+## Decide by looking at MSE plot
+num_bins_x_tsne_gau <- 31
+num_bins_y_tsne_gau <- 40
+hex_size_tsne_gau <- 0.02
+### non-empty 179
+
+hb_obj_tsne_gau <- hex_binning(data = pacmap_gau_scaled, x = "PaCMAP1", 
+                      y = "PaCMAP2", num_bins_x = num_bins_x_tsne_gau, 
+                      num_bins_y = num_bins_y_tsne_gau, x_start = NA, y_start = NA, 
+                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_tsne_gau, col_start = "PaCMAP")
+
+## Data set with all possible centroids in the hexagonal grid
+all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$centroids))
+
+## Generate all coordinates of hexagons
+hex_grid <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$hex_poly))
+
+## To obtain the standardise counts within hexbins
+counts_df <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$std_cts))
+df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
+                                             counts_df = counts_df)
+# ggplot(data = hex_grid, aes(x = x, y = y)) +
+#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
+#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
+#   coord_fixed()
+
+hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
+
+# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
+#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
+#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
+#   coord_fixed()
+
+tsne_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$data_hb_id))
+
+model_object_tsne_gau <- fit_highd_model( training_data = training_data_gau, 
+                                 nldr_df_with_id = pacmap_gau_scaled, 
+                                 x = "PaCMAP1", y = "PaCMAP2", 
+                                 num_bins_x = num_bins_x_tsne_gau, 
+                                 num_bins_y = num_bins_y_tsne_gau, 
+                                 x_start = NA, y_start = NA, 
+                                 buffer_x = NA, buffer_y = NA, 
+                                 hex_size = hex_size_tsne_gau,
+                                 is_rm_lwd_hex = FALSE, 
+                                 benchmark_to_rm_lwd_hex = NA, 
+                                 col_start_2d = "PaCMAP", 
+                                 col_start_highd = "x")
+
+df_bin_centroids_tsne_gau <- model_object_tsne_gau$df_bin_centroids
+df_bin_tsne_gau <- model_object_tsne_gau$df_bin
+
+## Triangulate bin centroids
+tr1_object_tsne_gau <- tri_bin_centroids(df_bin_centroids_tsne_gau, x = "c_x", y = "c_y")
+tr_from_to_df_tsne_gau <- gen_edges(tri_object = tr1_object_tsne_gau)
+
+## Compute 2D distances
+distance_tsne_gau <- cal_2d_dist(tr_coord_df = tr_from_to_df_tsne_gau, 
+                             start_x = "x_from", start_y = "y_from", 
+                             end_x = "x_to", end_y = "y_to", 
+                             select_vars = c("from", "to", "distance"))
+
+# distance_plot <- plot_dist(distance_tsne_gau) +
+#   ylab(expression(d^{(2)})) +
+#   theme(axis.text = element_text(size = 5),
+#         axis.title = element_text(size = 12))
+# 
+# distance_plot
+
+## To find the benchmark value
+benchmark_tsne_gau <- find_lg_benchmark(distance_edges = distance_tsne_gau, distance_col = "distance")
+
+# ggplot() +
+# geom_trimesh(data = df_bin_centroids_tsne_gau, mapping = aes(x = c_x, y = c_y))
+
+trimesh_removed_tsne_gau <- vis_rmlg_mesh(distance_edges = distance_tsne_gau, benchmark_value = benchmark_tsne_gau, tr_coord_df = tr_from_to_df_tsne_gau, distance_col = "distance")
+
+trimesh_removed_tsne_gau <- trimesh_removed_tsne_gau +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+df_all_tsne_gau <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID), tsne_data_with_hb_id_gau)
+
+# show_langevitour(df = df_all_tsne_gau, df_b = df_bin_tsne_gau, df_b_with_center_data = df_bin_centroids_tsne_gau, col_start = "x", distance_df = distance_tsne_gau, distance_col = "distance", benchmark_value = benchmark_tsne_gau)
 
@@
 
-```{r}
-#| warning: false
-#| echo: false
-
-error_df <- read_csv("data/pbmc/pbmc_3k_festem/error_pbmc_umap.csv", col_names = FALSE)
-
-error_df <- error_df |>
-  mutate(X4 = round(X4/100, 2),
-         X1 = as.integer(X1)) |>
-  arrange(X4)
-
-colnames(error_df) <- c("n_neighbors", "min_dist", "metric", "error (x 100)")
-
-kable(error_df, caption = "errors for different parameter combinations", label = "tbl-errordiffcomb") |>
-  #xtable2kable() |>
-  kable_styling(latex_options = c("striped", "HOLD_position"), position = "center", full_width = FALSE) |>
-  row_spec(0, bold = TRUE) |>
-  row_spec(7, bold=T, color = "red")
-```
+### Story
+
+- UMAP representation of PBMC3k data with 15 PCs and parameter choice of n:neighbor:30, min_dist:0.1, metric:cosine (author's suggestion) shows three well-separated clusters.
+The question we need to answer is: Does this UMAP representation preserve the actual data structure, or is something missing or misidentified or misleading?
 
@@
 
-class_avg <- umap_pbmc_scaled |>
-  group_by(cell_label) |>
-  summarise(
-    UMAP1 = median(UMAP1),
-    UMAP2 = median(UMAP2)
-  ) |>
-  mutate(cell_label = as.numeric(cell_label)) |>
-  arrange(cell_label)
-
-class_avg$cell_label <- as.factor(class_avg$cell_label)
-
-levels(class_avg$cell_label) <- c("Memory \nCD4 T", "Naive CD4 T", "CD14+ Mono",  "B", "CD8 T", "FCGR3A+ \n Mono", "NK", "M-MDSC\n-like", "CD27-CD+ \n Memory T", "DC")
-
 plot_list2_pbmc <- umap_pbmc_scaled |>
   ggplot(aes(x = UMAP1,
-               y = UMAP2, color = cell_label))+
+             y = UMAP2))+
   geom_point(alpha=0.5) +
@@
         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
-  geom_text(aes(x=UMAP1,y = UMAP2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
-```
-
-<!--obtain 2D and high-D distances-->
-
-<!---
-```{r}
-#| warning: false
-#| echo: false
-#| message: false
-#| eval: false
-
-## Obtian 2D distances
-proxy::dist(umap_pbmc[, -3]) |> 
-as.data.frame()
-
-## Obtain high-D distances
-proxy::dist(training_data_pbmc[, -16]) |> 
-as.data.frame()
-```
---->
+        legend.key.width = unit(0.25, 'cm')) #+
+  # scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
+  # geom_text(aes(x=UMAP1,y = UMAP2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) +
+  #annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "(a) 2D layout from UMAP (n_neighbors = 30, min_dist = 0.3) applied for the PBMC3k dataset. Is this a best representation of the original data?"
+#| label: fig-nldervisPBMCUMPview
+#| fig-pos: H
+
+plot_list2_pbmc 
+```
+
+- To investigate this, we are going to visualise the model in the data space, which means that the model generated from UMAP is visualised in high-D with original data.
+
+- To fit the model, first we need to decide how many number bins should allocate along the x and y axes. To make an initial choice, we use MSE plot.
 
@@
   geom_line() +
-  geom_vline(xintercept = 80, linetype="solid",
+  geom_vline(xintercept = 588, linetype="solid",
                 color = "red", size=0.8, alpha = 0.5) +
@@
   geom_line() +
-   geom_vline(xintercept = 80, linetype="solid",
+   geom_vline(xintercept = 588, linetype="solid",
                 color = "red", size=0.8, alpha = 0.5) +
@@
 Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error.
-Using the elbow method, when the total number of bins is set to 80, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
+Using the elbow method, when the total number of bins is set to 588, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
 This point indicates that adding less bins does not enough to capture the data structure."
@@
 
+- Then, the model is constructed by following all the steps in 2D and lifts the model into high-D.
+
 ```{r}
@@
 ## Decide by looking at MSE plot
-num_bins_x_pbmc <- 8
-num_bins_y_pbmc <- 10
-hex_size_pbmc <- 0.1
+num_bins_x_pbmc <- 21
+num_bins_y_pbmc <- 28
+hex_size_pbmc <- 0.03
 
@@
 trimesh_removed_pbmc_umap <- trimesh_removed_pbmc_umap +
+  geom_point(data = umap_pbmc_scaled, aes(x = UMAP1, y = UMAP2), alpha = 0.3) +
   theme_linedraw() +
@@
         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-df_bin_train <- df_bin_pbmc
-names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])
-
-df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), umap_data_with_hb_id_pbmc)
-
-error_df <- df_all_pbmc |>
-  dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates
-
-# prediction_df_join <- prediction_df_join |>
-#   dplyr::left_join(data, by = c("ID" = "ID")) ## Map high-D data
-
-for (i in 1:(NCOL(df_bin_train) - 1)) {
-
-  error_df[ , paste0("abs_residual_", "PC_", i)] <- abs(error_df[ , paste0("PC_", i)] - error_df[ , paste0("avg_", "PC_", i)])
-
-}
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) #+
+  #annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+# df_all_mnist <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), umap_data_with_hb_id_pbmc)
+# 
+# show_langevitour(df_all_pbmc, df_bin_pbmc, df_bin_centroids_pbmc, benchmark_pbmc, distance_pbmc, "distance", col_start = "PC_")
+
+# df_bin_train <- df_bin_pbmc
+# names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])
+# 
+# df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), umap_data_with_hb_id_pbmc)
+# 
+# error_df <- df_all_pbmc |>
+#   dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates
+# 
+# for (i in 1:(NCOL(df_bin_train) - 1)) {
+#   
+#   error_df[ , paste0("abs_residual_", "PC_", i)] <- abs(error_df[ , paste0("PC_", i)] - error_df[ , paste0("avg_", "PC_", i)])
+#   
+# }
+# 
+# error_df <- error_df |>
+#   dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "PC_")))))
+# 
+# 
+# error_df <- error_df |>
+#   mutate(type = if_else(total == 0, "no error",
+#                         if_else(total <= 15, "error 1-15",
+#                                 if_else(total <= 20, "error 15-20",
+#                                         if_else(total <= 25, "error 20-25",
+#                                                 if_else(total <= 30, "error 25-30",
+#                                                         if_else(total <= 35, "error 30-35",
+#                                                                 if_else(total <= 40, "error 35-40", "error greter than 40"))))))))
+# 
+# plot_list2_pbmc_error <- error_df |>
+#   mutate(type = factor(type , levels = c("no error", "error 1-15", "error 15-20", "error 20-25", "error 25-30", "error 30-35", "error 35-40", "error greter than 40"))) |>
+#     ggplot(aes(x = UMAP1,
+#                y = UMAP2, color = type))+
+#     geom_point(alpha=0.5, size = 0.1) +
+#     coord_equal() +
+#     theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+#   theme_linedraw() +
+#     theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+#               axis.title.x = element_blank(), axis.title.y = element_blank(),
+#               axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+#               axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+#               panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+#         legend.title = element_text(size=5), #change legend title font size
+#         legend.text = element_text(size=4),
+#          legend.key.height = unit(0.25, 'cm'),
+#          legend.key.width = unit(0.25, 'cm')) +
+#   scale_color_manual(values=c("#fee0d2", "#fcbba1",
+#                                          "#fc9272", "#fb6a4a", "#ef3b2c",
+#                                          "#cb181d", "#a50f15", "#99000d")) 
+```
+
+```{r}
+#| echo: false
+#| fig-cap: " Model in the 2D space with UMAP (<https://youtu.be/CNXjWZMyh_M>)"
+#| label: fig-nldervisPBMCUMAP
+#| fig-pos: H
+
+trimesh_removed_pbmc_umap 
+```
+
+<!--add screen shots of model in high-D-->
+
+- You will see three well-separated clusters in high-D as well which gives evidence that UMAP preserves the global structure.
+Furthermore, there is non-linearity within the clusters which was not identified by the UMAP, which proves that local structure is not preserved.
+
+- Can we find a better parameter choice for UMAP which preserve the local structure as well?
+
+- To do that, we are going to fit models for different parameter choice and select the one which have the lowest error.
+
+```{r}
+#| warning: false
+#| echo: false
+
+error_df <- read_csv("data/pbmc/pbmc_3k_festem/error_pbmc_umap.csv", col_names = FALSE)
 
 error_df <- error_df |>
-  dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "PC_")))))
-
-# library(ggbeeswarm)
-# error_df$group <- "1"
-# ggplot(error_df, aes(x = group, y = total)) +
-#   geom_quasirandom()+
-#   ylim(0, max(unlist(error_df$total))+ 0.5) + coord_flip()
-
-### The minimum error is 0 and the maximum is 42.17439
-### There is lot of points with error 0,
-
-error_df <- error_df |>
-  mutate(type = if_else(total == 0, "no error",
-                        if_else(total <= 15, "error 1-15",
-                                if_else(total <= 20, "error 15-20",
-                                        if_else(total <= 25, "error 20-25",
-                                                if_else(total <= 30, "error 25-30",
-                                                        if_else(total <= 35, "error 30-35",
-                                                                if_else(total <= 40, "error 35-40", "error greter than 40"))))))))
-
-plot_list2_pbmc_error <- error_df |>
-  mutate(type = factor(type , levels = c("no error", "error 1-15", "error 15-20", "error 20-25", "error 25-30", "error 30-35", "error 35-40", "error greter than 40"))) |>
-    ggplot(aes(x = UMAP1,
-               y = UMAP2, color = type))+
-    geom_point(alpha=0.5, size = 0.1) +
-    coord_equal() +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#fee0d2", "#fcbba1",
-                                         "#fc9272", "#fb6a4a", "#ef3b2c",
-                                         "#cb181d", "#a50f15", "#99000d")) 
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "(a) 2D layout from UMAP (n_neighbors = 30, min_dist = 0.3) applied for the PBMC3k dataset.
-Is this a best representation of the original data?, (b)  Model in the 2D space with UMAP (<https://youtu.be/CNXjWZMyh_M>)"
-#| label: fig-nldervisPBMCUMAP
-#| out-width: 100%
-#| fig-pos: h
-
-plot_list2_pbmc + trimesh_removed_pbmc_umap + plot_list2_pbmc_error +
-plot_layout(guides='collect', ncol=3) &
-  theme(legend.position='bottom')
-```
+  mutate(X4 = round(X4/100, 2),
+         X1 = as.integer(X1)) |>
+  arrange(X4)
+
+colnames(error_df) <- c("n_neighbors", "min_dist", "metric", "error (x 100)")
+
+kable(error_df, caption = "errors for different parameter combinations", label = "tbl-errordiffcomb") |>
+  #xtable2kable() |>
+  kable_styling(latex_options = c("striped", "HOLD_position"), position = "center", full_width = FALSE) |>
+  row_spec(0, bold = TRUE) |>
+  row_spec(7, bold=T, color = "red")
+```
+
+- Selected the one with n:neighbor:5, min_dist:0.99, metric:cosine.
+
+- Now, we are going to fit the model for the new parameter choice.
 
@@
 
-class_avg <- umap_pbmc_scaled |>
-  group_by(cell_label) |>
-  summarise(
-    UMAP1 = median(UMAP1),
-    UMAP2 = median(UMAP2)
-  ) |>
-  mutate(cell_label = as.numeric(cell_label)) |>
-  arrange(cell_label)
-
-class_avg$cell_label <- as.factor(class_avg$cell_label)
-
-levels(class_avg$cell_label) <- c("Memory \nCD4 T", "Naive CD4 T", "CD14+ Mono",  "B", "CD8 T", "FCGR3A+ \n Mono", "NK", "M-MDSC\n-like", "CD27-CD+ \n Memory T", "DC")
-
 plot_list2_pbmc <- umap_pbmc_scaled |>
     ggplot(aes(x = UMAP1,
-               y = UMAP2, color = cell_label))+
+               y = UMAP2))+
     geom_point(alpha=0.5) +
@@
          legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
-  geom_text(aes(x=UMAP1,y = UMAP2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
+         legend.key.width = unit(0.25, 'cm')) 
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "(a) 2D layout from UMAP (n_neighbors = 5, min_dist = 0.99, metric = cosine) applied for the PBMC3k dataset. Is this a best representation of the original data?"
+#| label: fig-nldervisPBMCUMAPview2
+#| fig-pos: H
+
+plot_list2_pbmc
+
 ```
@@
 ## Decide by looking at MSE plot
-num_bins_x_pbmc <- 8
-num_bins_y_pbmc <- 11
-hex_size_pbmc <- 0.1
+num_bins_x_pbmc <- 21
+num_bins_y_pbmc <- 28
+hex_size_pbmc <- 0.03
 
@@
         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
 
@@
                                          "#fc9272", "#fb6a4a", "#ef3b2c",
-                                         "#cb181d", "#a50f15", "#99000d")) 
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "(a) 2D layout from UMAP (n_neighbors = 5, min_dist = 0.99, metric = cosine) applied for the PBMC3k dataset.
-Is this a best representation of the original data?, (b)  Model in the 2D space with UMAP (<>)"
+                                         "#cb181d", "#a50f15", "#99000d")) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+# show_langevitour(df_all_pbmc, df_bin_pbmc, df_bin_centroids_pbmc, benchmark_pbmc, distance_pbmc, "distance", col_start = "PC_")
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "Model in the 2D space with UMAP (<>)"
 #| label: fig-nldervisPBMCUMAP3
 #| out-width: 100%
-#| fig-pos: h
-
-plot_list2_pbmc + trimesh_removed_pbmc_umap + plot_list2_pbmc_error +
-plot_layout(guides='collect', ncol=3) &
+#| fig-pos: H
+
+trimesh_removed_pbmc_umap + plot_list2_pbmc_error +
+plot_layout(guides='collect', ncol=2) &
   theme(legend.position='bottom')
@@
 #| out-width: 100%
+#| fig-pos: H
 
@@
 
+- Why I choose this dataset? Because I need show that whatever NLDR technique that we use, our model can capture the non-linear structure within clusters as well.
+
 - Digit 1 (7877 images only)
@@
 
-- Learned from the model: There is a non-linear shape in 2D, but if look at the model in high-D, the non-linear continuity is there but there is a twist in the structure which didn't capture by PaCMAP.
-
-- Write about looking at images of digit 1 as well..how in 2D space images are distributed, which images occur high-errors in high-D
+- Learned from the model: There is a non-linear shape in 2D, but if look at the model in high-D, the non-linear continuity is there but there is a twist in the structure which didn't capture by PaCMAP in 2D layout.
+
+#### Story
+
+- PaCMAP representation of MNIST digit 1 data with 10PCs and parameter choice of n_components=2, n_neighbors=10, init="random", MN_ratio=0.9, FP_ratio=2.0 shows a non-linear structure (see @fig-nldervisMNISTPACMAP (a)).
+I selected this parameter choice because this the default parameter setting for PaCMAP and selected PaCMAP because when I observed the 5 NLDR techniques applied for the data PaCMAP showed the non-linear structure effectively in 2D.
+
+- By looking at MSE plot, I selected the 180 bins, which associated with 12 bins along the x-axis, 15 bins along the y-axis, and hex size as 0.06 (@fig-diagnosticpltMNIST (b)).
+Then, fitted the model in 2D (see @fig-nldervisMNISTPACMAP (b)).
+As you can see in @fig-nldervisMNISTPACMAP (b), the fitted model capture the non-linear structure of the 2D data.
 
@@
 trimesh_removed_mnist_pacmap <- trimesh_removed_mnist_pacmap +
+  geom_point(data = pacmap_minst_scaled, aes(x = PaCMAP1, y = PaCMAP2), alpha = 0.06) + 
   theme_linedraw() +
@@
 
-# show_langevitour(df_all_mnist, df_bin_mnist, df_bin_centroids_mnist, benchmark_mnist, distance_mnist, "distance", col_start = "PC")
-
-predict_minst_df <- as.data.frame(do.call(cbind, predict_emb(training_data_mnist, df_bin_centroids_mnist, df_bin_mnist, "PaCMAP")))
-gen_summary(training_data_mnist, predict_minst_df, df_bin_mnist, col_start = "PC")
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "(a) 2D layout from UMAP (n_components=2, n_neighbors=10, init=random, MN_ratio=0.9, FP_ratio=2.0) applied for the digit 1 of the MNIST dataset.
-Is this a best representation of the original data?, (b)  Model in the 2D space with PaCMAP ()"
+show_langevitour(df_all_mnist, df_bin_mnist, df_bin_centroids_mnist, benchmark_mnist, distance_mnist, "distance", col_start = "PC")
+
+# predict_minst_df <- as.data.frame(do.call(cbind, predict_emb(training_data_mnist, df_bin_centroids_mnist, df_bin_mnist, "PaCMAP")))
+# gen_summary(training_data_mnist, predict_minst_df, df_bin_mnist, col_start = "PC")
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "(a) 2D layout from PaCMAP (n_components=2, n_neighbors=10, init=random, MN_ratio=0.9, FP_ratio=2.0) applied for the digit 1 of the MNIST dataset, (b) Model in the 2D space overlaid on PaCMAP data, and (c) Error occurred by the high-D model for high-D data and the error mapped into 2D NLDR points."
 #| label: fig-nldervisMNISTPACMAP
 #| out-width: 100%
-#| fig-pos: h
+#| fig-pos: H
 
@@
 ```
+
+- Next, the constructed 2D model lifts into high-dimensions (see @fig-mnist1_sc).
+
+::: {#fig-mnist1_sc layout-ncol="3"}
+![](figures/mnist/sc_1.png){#fig-mnist1_sc1}
+
+![](figures/mnist/sc_2.png){#fig-mnist1_sc2}
+
+![](figures/mnist/sc_3.png){#fig-mnist1_sc3}
+
+Screen shots of the **langevitour** of the MNIST digit 1 data set, shows the model in high-D, a video of the tour animation is available at (<https://youtu.be/zq2GM9qvUNA>).
+:::
+
+- As shown in @fig-mnist1_sc (a), the non-linear continuity structure that you have seen in 2D representation of PaCMAP, also can be seen in the model in high-D space as well.
+This gives us an evidence that our model capture the non-linear structure of the data effectively.
+
+- Furthermore, our model shows some addition pattern of data which couldn't be seen in the 2D layout.
+
+- One is the twisted structure within the non-linear structure (see @fig-mnist1_sc (c)).
+
+- Another thing is the connected corners of the non-linear structure (see @fig-mnist1_sc (b)). The non-linear curve is bended.
+
+- To gives the evidence that PaCMAP preserve the local structure (non-linear pattern), our model helps by computing the model errors (@fig-diagnosticpltMNIST (c)).
 
@@
 
+- As you can see in @fig-diagnosticpltMNIST (c), the largest model error shows in the upper right side of the 2D layout.
+A sample of the images that shows the largest error can be seen in @fig-sampleimg.
+The reason for the large model error is that the high-D points related to this largest error located away from the high-D model points.
+
 ## Conclusion {#sec-conclusion}

### 2024-03-18 paper.tex
Added model interpretations for MNIST digit 1 data set

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, interior hidden, boxrule=0pt, sharp corners, enhanced, borderline west={3pt}{0pt}{shadecolor}, breakable]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, frame hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, enhanced, breakable, boxrule=0pt]}{\end{tcolorbox}}\fi
 
@@
   Generate a confusion matrix.
+\end{itemize}
+
+\hypertarget{story}{%
+\subsection{Story}\label{story}}
+
+\begin{itemize}
+\tightlist
+\item
+  UMAP representation of PBMC3k data with 15 PCs and parameter choice of
+  n:neighbor:30, min\_dist:0.1, metric:cosine (author's suggestion)
+  shows three well-separated clusters. The question we need to answer
+  is: Does this UMAP representation preserve the actual data structure,
+  or is something missing or misidentified or misleading?
+\end{itemize}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-nldervisPBMCUMPview-1.pdf}
+
+}
+
+\caption{\label{fig-nldervisPBMCUMPview}(a) 2D layout from UMAP
+(n\_neighbors = 30, min\_dist = 0.3) applied for the PBMC3k dataset. Is
+this a best representation of the original data?}
+
+\end{figure}
+
+\begin{itemize}
+\item
+  To investigate this, we are going to visualise the model in the data
+  space, which means that the model generated from UMAP is visualised in
+  high-D with original data.
+\item
+  To fit the model, first we need to decide how many number bins should
+  allocate along the x and y axes. To make an initial choice, we use MSE
+  plot.
+\end{itemize}
+
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-diagnosticpltPBMC-1.pdf}
+
+}
+
+\caption{\label{fig-diagnosticpltPBMC}Goodness of fit statistics from
+UMAP applied to training PBMC3k dataset. What is the effective number of
+bins in each NLDR technique to create a 2D model? The MSE plot have a
+steep slope at the beginning, indicating that a smaller number of bins
+causes a larger amount of error. Then, the slope gradually declines or
+level off, indicating that a higher number of bins generates a smaller
+error. Using the elbow method, when the total number of bins is set to
+588, the slope of the Mean Squared Error (MSE) plot experiences a sudden
+and noticeable change, resembling an elbow-like shape. This point
+indicates that adding less bins does not enough to capture the data
+structure.}
+
+\end{figure}
+
+\begin{itemize}
+\tightlist
+\item
+  Then, the model is constructed by following all the steps in 2D and
+  lifts the model into high-D.
+\end{itemize}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-nldervisPBMCUMAP-1.pdf}
+
+}
+
+\caption{\label{fig-nldervisPBMCUMAP}Model in the 2D space with UMAP
+(\url{https://youtu.be/CNXjWZMyh_M})}
+
+\end{figure}
+
+\begin{itemize}
+\item
+  You will see three well-separated clusters in high-D as well which
+  gives evidence that UMAP preserves the global structure. Furthermore,
+  there is non-linearity within the clusters which was not identified by
+  the UMAP, which proves that local structure is not preserved.
+\item
+  Can we find a better parameter choice for UMAP which preserve the
+  local structure as well?
+\item
+  To do that, we are going to fit models for different parameter choice
+  and select the one which have the lowest error.
 \end{itemize}
@@
 
-\begin{figure}
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-diagnosticpltPBMC-1.pdf}
-
-}
-
-\caption{\label{fig-diagnosticpltPBMC}Goodness of fit statistics from
-UMAP applied to training PBMC3k dataset. What is the effective number of
-bins in each NLDR technique to create a 2D model? The MSE plot have a
-steep slope at the beginning, indicating that a smaller number of bins
-causes a larger amount of error. Then, the slope gradually declines or
-level off, indicating that a higher number of bins generates a smaller
-error. Using the elbow method, when the total number of bins is set to
-80, the slope of the Mean Squared Error (MSE) plot experiences a sudden
-and noticeable change, resembling an elbow-like shape. This point
-indicates that adding less bins does not enough to capture the data
-structure.}
-
-\end{figure}
-
-\begin{figure}[h]
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervisPBMCUMAP-1.pdf}
-
-}
-
-\caption{\label{fig-nldervisPBMCUMAP}(a) 2D layout from UMAP
-(n\_neighbors = 30, min\_dist = 0.3) applied for the PBMC3k dataset. Is
-this a best representation of the original data?, (b) Model in the 2D
-space with UMAP (\url{https://youtu.be/CNXjWZMyh_M})}
+\begin{itemize}
+\item
+  Selected the one with n:neighbor:5, min\_dist:0.99, metric:cosine.
+\item
+  Now, we are going to fit the model for the new parameter choice.
+\end{itemize}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-nldervisPBMCUMAPview2-1.pdf}
+
+}
+
+\caption{\label{fig-nldervisPBMCUMAPview2}(a) 2D layout from UMAP
+(n\_neighbors = 5, min\_dist = 0.99, metric = cosine) applied for the
+PBMC3k dataset. Is this a best representation of the original data?}
 
@@
 
-\begin{figure}[h]
+\begin{figure}[H]
 
@@
 
-\caption{\label{fig-nldervisPBMCUMAP3}(a) 2D layout from UMAP
-(n\_neighbors = 5, min\_dist = 0.99, metric = cosine) applied for the
-PBMC3k dataset. Is this a best representation of the original data?, (b)
-Model in the 2D space with UMAP (\textless\textgreater)}
+\caption{\label{fig-nldervisPBMCUMAP3}Model in the 2D space with UMAP
+(\textless\textgreater)}
 
@@
 \begin{itemize}
-\tightlist
+\item
+  Why I choose this dataset? Because I need show that whatever NLDR
+  technique that we use, our model can capture the non-linear structure
+  within clusters as well.
 \item
@@
   at the model in high-D, the non-linear continuity is there but there
-  is a twist in the structure which didn't capture by PaCMAP.
-\item
-  Write about looking at images of digit 1 as well..how in 2D space
-  images are distributed, which images occur high-errors in high-D
+  is a twist in the structure which didn't capture by PaCMAP in 2D
+  layout.
+\end{itemize}
+
+\hypertarget{story-1}{%
+\subsubsection{Story}\label{story-1}}
+
+\begin{itemize}
+\item
+  PaCMAP representation of MNIST digit 1 data with 10PCs and parameter
+  choice of n\_components=2, n\_neighbors=10, init=``random'',
+  MN\_ratio=0.9, FP\_ratio=2.0 shows a non-linear structure (see
+  Figure~\ref{fig-nldervisMNISTPACMAP} (a)). I selected this parameter
+  choice because this the default parameter setting for PaCMAP and
+  selected PaCMAP because when I observed the 5 NLDR techniques applied
+  for the data PaCMAP showed the non-linear structure effectively in 2D.
+\item
+  By looking at MSE plot, I selected the 180 bins, which associated with
+  12 bins along the x-axis, 15 bins along the y-axis, and hex size as
+  0.06 (Figure~\ref{fig-diagnosticpltMNIST} (b)). Then, fitted the model
+  in 2D (see Figure~\ref{fig-nldervisMNISTPACMAP} (b)). As you can see
+  in Figure~\ref{fig-nldervisMNISTPACMAP} (b), the fitted model capture
+  the non-linear structure of the 2D data.
 \end{itemize}
@@
 
-\begin{verbatim}
-$mse
-[1] 2.276023
-
-$aic
-[1] 66382.78
-\end{verbatim}
-
-\begin{figure}[h]
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-84-1.pdf}
+
+\begin{figure}[H]
 
@@
 
-\caption{\label{fig-nldervisMNISTPACMAP}(a) 2D layout from UMAP
+\caption{\label{fig-nldervisMNISTPACMAP}(a) 2D layout from PaCMAP
 (n\_components=2, n\_neighbors=10, init=random, MN\_ratio=0.9,
-FP\_ratio=2.0) applied for the digit 1 of the MNIST dataset. Is this a
-best representation of the original data?, (b) Model in the 2D space
-with PaCMAP ()}
-
-\end{figure}
+FP\_ratio=2.0) applied for the digit 1 of the MNIST dataset, (b) Model
+in the 2D space overlaid on PaCMAP data, and (c) Error occurred by the
+high-D model for high-D data and the error mapped into 2D NLDR points.}
+
+\end{figure}
+
+\begin{itemize}
+\tightlist
+\item
+  Next, the constructed 2D model lifts into high-dimensions (see
+  Figure~\ref{fig-mnist1_sc}).
+\end{itemize}
+
+\begin{figure}
+
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/mnist/sc_1.png}
+
+}
+
+}
+
+\subcaption{\label{fig-mnist1_sc1}}
+\end{minipage}%
+%
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/mnist/sc_2.png}
+
+}
+
+}
+
+\subcaption{\label{fig-mnist1_sc2}}
+\end{minipage}%
+%
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/mnist/sc_3.png}
+
+}
+
+}
+
+\subcaption{\label{fig-mnist1_sc3}}
+\end{minipage}%
+
+\caption{\label{fig-mnist1_sc}Screen shots of the \textbf{langevitour}
+of the MNIST digit 1 data set, shows the model in high-D, a video of the
+tour animation is available at (\url{https://youtu.be/zq2GM9qvUNA}).}
+
+\end{figure}
+
+\begin{itemize}
+\item
+  As shown in Figure~\ref{fig-mnist1_sc} (a), the non-linear continuity
+  structure that you have seen in 2D representation of PaCMAP, also can
+  be seen in the model in high-D space as well. This gives us an
+  evidence that our model capture the non-linear structure of the data
+  effectively.
+\item
+  Furthermore, our model shows some addition pattern of data which
+  couldn't be seen in the 2D layout.
+\item
+  One is the twisted structure within the non-linear structure (see
+  Figure~\ref{fig-mnist1_sc} (c)).
+\item
+  Another thing is the connected corners of the non-linear structure
+  (see Figure~\ref{fig-mnist1_sc} (b)). The non-linear curve is bended.
+\item
+  To gives the evidence that PaCMAP preserve the local structure
+  (non-linear pattern), our model helps by computing the model errors
+  (Figure~\ref{fig-diagnosticpltMNIST} (c)).
+\end{itemize}
 
@@
 \end{figure}
+
+\begin{itemize}
+\tightlist
+\item
+  As you can see in Figure~\ref{fig-diagnosticpltMNIST} (c), the largest
+  model error shows in the upper right side of the 2D layout. A sample
+  of the images that shows the largest error can be seen in
+  Figure~\ref{fig-sampleimg}. The reason for the large model error is
+  that the high-D points related to this largest error located away from
+  the high-D model points.
+\end{itemize}
 

### 2024-03-18 extra/extra_pbmc3k.R

@@ -1,0 +1,103 @@
+```{r}
+#| warning: false
+#| echo: false
+
+## Import data
+training_data_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_pca_50.rds")
+training_data_pbmc <- training_data_pbmc[, 1:15] |>
+  mutate(ID = 1:NROW(training_data_pbmc))
+
+umap_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_umap.rds")
+umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc,
+                                                                 x = "UMAP1", y = "UMAP2"))) |>
+  dplyr::mutate(cell_label = umap_pbmc$cell_label) |>
+  dplyr::rename(c("UMAP1" = "scaled_UMAP1",
+                  "UMAP2" = "scaled_UMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(umap_pbmc))
+
+class_avg <- umap_pbmc_scaled |>
+  group_by(cell_label) |>
+  summarise(
+    UMAP1 = median(UMAP1),
+    UMAP2 = median(UMAP2)
+  ) |>
+  mutate(cell_label = as.numeric(cell_label)) |>
+  arrange(cell_label)
+
+class_avg$cell_label <- as.factor(class_avg$cell_label)
+
+levels(class_avg$cell_label) <- c("Memory \nCD4 T", "Naive CD4 T", "CD14+ Mono",  "B", "CD8 T", "FCGR3A+ \n Mono", "NK", "M-MDSC\n-like", "CD27-CD+ \n Memory T", "DC")
+
+plot_list2_pbmc <- umap_pbmc_scaled |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2, color = cell_label))+
+  geom_point(alpha=0.5) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
[…]

### 2024-03-18 paper.qmd
Added interpretation for PBMC3k data

@@
 
-- UMAP representation of PBMC3k data with 15 PCs and parameter choice of n:neighbor:30, min_dist:0.1, metric:cosine (author's suggestion) shows three well-separated clusters.
+- UMAP representation of PBMC3k data with 15 PCs and parameter choice of n:neighbor:30, min_dist:0.1, metric:cosine (author's suggestion) shows three well-separated clusters (see @fig-nldervisPBMCUMPview).
 The question we need to answer is: Does this UMAP representation preserve the actual data structure, or is something missing or misidentified or misleading?
@@
 
-- To fit the model, first we need to decide how many number bins should allocate along the x and y axes. To make an initial choice, we use MSE plot.
+- To fit the model, first we need to decide how many number bins should allocate along the x and y axes. To make an initial choice, we use MSE plot (@fig-diagnosticpltPBMC (b)).
 
@@
 #| echo: false
-#| fig-cap: " Model in the 2D space with UMAP (<https://youtu.be/CNXjWZMyh_M>)"
+#| fig-cap: "Model in the 2D space overlaid on UMAP data."
 #| label: fig-nldervisPBMCUMAP
@@
 
-<!--add screen shots of model in high-D-->
-
-- You will see three well-separated clusters in high-D as well which gives evidence that UMAP preserves the global structure.
-Furthermore, there is non-linearity within the clusters which was not identified by the UMAP, which proves that local structure is not preserved.
-
-- Can we find a better parameter choice for UMAP which preserve the local structure as well?
-
-- To do that, we are going to fit models for different parameter choice and select the one which have the lowest error.
+- As shown in @fig-nldervisPBMCUMAP, there are three well-separated clusters (according to the model in 2D space).
+
+::: {#fig-pbmc1_sc layout-ncol="3" fig-pos="H"}
+![](figures/pbmc3k/sc_1.png){#fig-pbmc1_sc1}
+
+![](figures/pbmc3k/sc_2.png){#fig-pbmc1_sc2}
+
+![](figures/pbmc3k/sc_3.png){#fig-pbmc1_sc3}
+
+Screen shots of the **langevitour** of the PBMC3k data set, shows the model in high-D, a video of the tour animation is available at (<https://youtu.be/Gnhh4hfsbbg>).
+:::
+
+- According to @fig-pbmc1_sc (a), the model in high-D space shows two well separated clusters.
+Also, two clusters are really close to each other (see @fig-pbmc1_sc (b)).
+Furthermore, there are three clusters shown by the model in 2D actually represented in high-D as well (see @fig-pbmc1_sc (c)).
+Another important thing to notice from the model in high-D space is that the non-linear continuity structure in two well-separated clusters.
+
+- @fig-pbmc1_sc (c) gives evidence that UMAP preserves the global structure.
+Furthermore, the non-linearity within the clusters shown in @fig-pbmc1_sc (a) proves that local structure is also preserved but that's not visually shown by th UMAP layout in 2D.
+
+- Can we find a better parameter choice for UMAP which shows this non-linear continuity of the big clusters in 2D space as well?
+
+- To do that, we are going to fit models for different parameter choice and select the one which have the lowest model error in high-D (see @tbl-errordiffcomb).
 
@@
 ```
-
-- Selected the one with n:neighbor:5, min_dist:0.99, metric:cosine.
-
-- Now, we are going to fit the model for the new parameter choice.
 
@@
 ```
+
+- Selected the one with n:neighbor:5, min_dist:0.99, metric:cosine (see @fig-nldervisPBMCUMAPview2).
+As shown in @fig-nldervisPBMCUMAPview2, the two big clusters have non-linear continuity structures, two big clusters are well-separated but close to each other, and the two cluster which exists in the one big cluster are really close to each other (see @fig-nldervisPBMCUMAPview2).
+
+- Now, we are going to fit the model for the new parameter choice.
 
@@
   geom_line() +
-  geom_vline(xintercept = 88, linetype="solid",
+  geom_vline(xintercept = 588, linetype="solid",
                 color = "red", size=1, alpha = 0.5) +
@@
   geom_line() +
-   geom_vline(xintercept = 88, linetype="solid",
+   geom_vline(xintercept = 588, linetype="solid",
                 color = "red", size=1, alpha = 0.5) +
@@
 Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error.
-Using the elbow method, when the total number of bins is set to 144, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
+Using the elbow method, when the total number of bins is set to 588, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
 This point indicates that adding less bins does not enough to capture the data structure."
@@
 
+benchmark_value_rm_lwd <- stats::quantile(df_bin_centroids_pbmc$std_counts, 
+                probs = c(0,0.25,0.5,0.75,1), names = FALSE)[2]
+
+df_bin_centroids_low <- df_bin_centroids_pbmc |>
+  dplyr::filter(std_counts <= benchmark_value_rm_lwd)
+
+identify_rm_bins <- find_low_dens_hex(df_bin_centroids_all = df_bin_centroids_pbmc, 
+                                      num_bins_x = num_bins_x_pbmc, 
+                                      df_bin_centroids_low = df_bin_centroids_low)
+
+df_bin_centroids_pbmc <- df_bin_centroids_pbmc |>
+  dplyr::filter(!(hexID %in% identify_rm_bins))
+
 ## Triangulate bin centroids
@@
 benchmark_pbmc <- find_lg_benchmark(distance_edges = distance_pbmc, distance_col = "distance")
+benchmark_pbmc <- 0.08
 
@@
 #| echo: false
-#| fig-cap: "Model in the 2D space with UMAP (<>)"
+#| fig-cap: "(a) Model in the 2D space overlaid with UMAP data, and (b) Error occurred by the high-D model for high-D data and the error mapped into 2D NLDR points."
 #| label: fig-nldervisPBMCUMAP3
@@
 ```
+
+- According to @fig-nldervisPBMCUMAP3 (a), the model constructed in 2D space for the new parameter choice of UMAP shows 5 clusters which are really close to each other.
+The big cluster shows the non-linear continuity structure with two sub clusters.
+Other three clusters positioned in a structure of non-linearity but really close to each other.
+
+::: {#fig-pbmc2_sc layout-ncol="3" fig-pos="H"}
+![](figures/pbmc3k/sc_4.png){#fig-pbmc2_sc1}
+
+![](figures/pbmc3k/sc_5.png){#fig-pbmc2_sc2}
+
+![](figures/pbmc3k/sc_6.png){#fig-pbmc2_sc3}
+
+Screen shots of the **langevitour** of the PBMC3k data set, shows the model in high-D with new parameter choice for UMAP, a video of the tour animation is available at (<https://youtu.be/lc_6-rIuNHI>).
+:::
+
+- As shown in @fig-pbmc2_sc, there are 5 clusters, three of them are well-separated but close to each other as seen in @fig-nldervisPBMCUMAP3 (a).
+Also, other two clusters are really close to each other but separated from the early mentioned three clusters.
+
+- Next, we are going to investigate whether, the clusters we identified in 2D UMAP layout, can also be identified in high-D as well.
+
+- First, I colored the clusters in 2D layout manually. Then, I used spin-and-brush approach with `detour` to color the clusters in high-D.
 
@@
 #| echo: false
-#| fig-cap: "2D layout from UMAP (n_neighbors = 5, min_dist = 0.99, metric = cosine) applied for PBMC3k dataset (a) colored manually in 2D, (b) colored in high-D"
+#| fig-cap: "2D layout from UMAP (n_neighbors = 5, min_dist = 0.99, metric = cosine) applied for PBMC3k dataset (a) colored manually in 2D, (b) colored in high-D."
 #| label: fig-nldervisPBMCUMAP3Res
@@
 
+- @fig-nldervisPBMCUMAP3Res shows the 2D UMAP layout colored with each method.
+As shown in @fig-nldervisPBMCUMAP3Res (b) and the confusion matrix,there are some miss-classifications.
+What is the reason for that?
+
+- For example, the point (ID: 1158) classified as **cluster 3** in 2D UMAP layout, classified as **cluster 6** when doing the colouring in high-D (see @fig-nldervisPBMCUMAP3Res). Why?
+
+::: {#fig-pbmc2_appsc layout-ncol="1" fig-pos="H"}
+![](figures/pbmc3k/sc_7.png){#fig-pbmc2_sc7}
+
+Screen shots of app which shows 2D UMAP layout and comparison of distances related to the point classified as **cluster 6** in high-D (ID: 1158) and randomly selected point which is in **cluster 6** (ID: 2291).
+:::
+
+- If we compute the distance between these two points, you will see that it preserves the pair-wise distance (see @fig-pbmc2_appsc), which give us a evidence that even though the point 1158, classified as **cluster 3** in 2D layout, actually should be in **cluster 6**.
+This also evident that some misleading can happens with UMAP.
+
+::: {#fig-pbmc2_appsc1 layout-ncol="1" fig-pos="H"}
+![](figures/pbmc3k/sc_8.png){#fig-pbmc2_sc8}
+
+Screen shots of app which shows 2D UMAP layout and comparison of distances related to the point classified as **cluster 5** in high-D (ID: 3) and randomly selected point which is in **cluster 5** (ID: 2198).
+:::
+
+- For another example, if we compute the distance between these two points, you will see that it preserves the pair-wise distance (see @fig-pbmc2_appsc1), which give us a evidence that even though the point 3, classified as **cluster 4** in 2D layout, actually should be in **cluster 5**.
+This also evident that some misleading can happens with UMAP.
+
+```{r}
+#| warning: false
+#| echo: false
+#| message: false
+
+df_all_pbmc <- dplyr::bind_cols(training_data_pbmc, umap_data_with_hb_id_pbmc)
+
+df_all_pbmc <- inner_join(df_all_pbmc, detour_results8, by = c("ID" = "ID"))
+
+### Define type column
+df <- df_all_pbmc |>
+  dplyr::select(tidyselect::starts_with("PC_"), result2_class_label) |>
+  dplyr::rename("type" = "result2_class_label") ## original dataset
+
+df_b <- df_bin_pbmc |>
+  dplyr::filter(hb_id %in% df_bin_centroids_pbmc$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b <- df_b[match(df_bin_centroids_pbmc$hexID, df_b$hb_id),] |>
+  dplyr::select(-hb_id)
+
+df_exe <- dplyr::bind_rows(df_b, df)
+
+ ## Set the maximum difference as the criteria
+distance_df_small_edges <- distance_pbmc |>
+  dplyr::filter(distance < benchmark_pbmc)
+## Since erase brushing is considerd.
+
+# langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
+#                          lineFrom = distance_df_small_edges$from,
+#                            lineTo = distance_df_small_edges$to,
+#                          group = df_exe$type, pointSize = 3,
+#                          levelColors = c("#d82cd3", "#49df20", "#619cff", "#000000", "#df2020", "#ffbd61", "#33a02c"))
+
+```
+
 ### Handwritten digits
 
-- Why I choose this dataset? Because I need show that whatever NLDR technique that we use, our model can capture the non-linear structure within clusters as well.
+- Why I choose this data set? Because I need show that whatever NLDR technique that we use, our model can capture the non-linear structure within clusters as well.
 
@@
 
-::: {#fig-mnist1_sc layout-ncol="3"}
+::: {#fig-mnist1_sc layout-ncol="3" fig-pos="H"}
 ![](figures/mnist/sc_1.png){#fig-mnist1_sc1}

### 2024-03-18 paper.tex
Added interpretation for PBMC3k data

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, frame hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, enhanced, breakable, boxrule=0pt]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, boxrule=0pt, frame hidden, sharp corners, breakable]}{\end{tcolorbox}}\fi
 
@@
   n:neighbor:30, min\_dist:0.1, metric:cosine (author's suggestion)
-  shows three well-separated clusters. The question we need to answer
+  shows three well-separated clusters (see
+  Figure~\ref{fig-nldervisPBMCUMPview}). The question we need to answer
   is: Does this UMAP representation preserve the actual data structure,
@@
   allocate along the x and y axes. To make an initial choice, we use MSE
-  plot.
+  plot (Figure~\ref{fig-diagnosticpltPBMC} (b)).
 \end{itemize}
@@
 
-\caption{\label{fig-nldervisPBMCUMAP}Model in the 2D space with UMAP
-(\url{https://youtu.be/CNXjWZMyh_M})}
-
-\end{figure}
-
-\begin{itemize}
-\item
-  You will see three well-separated clusters in high-D as well which
-  gives evidence that UMAP preserves the global structure. Furthermore,
-  there is non-linearity within the clusters which was not identified by
-  the UMAP, which proves that local structure is not preserved.
-\item
-  Can we find a better parameter choice for UMAP which preserve the
-  local structure as well?
+\caption{\label{fig-nldervisPBMCUMAP}Model in the 2D space overlaid on
+UMAP data.}
+
+\end{figure}
+
+\begin{itemize}
+\tightlist
+\item
+  As shown in Figure~\ref{fig-nldervisPBMCUMAP}, there are three
+  well-separated clusters (according to the model in 2D space).
+\end{itemize}
+
+\begin{figure}[H]
+
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/pbmc3k/sc_1.png}
+
+}
+
+}
+
+\subcaption{\label{fig-pbmc1_sc1}}
+\end{minipage}%
+%
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/pbmc3k/sc_2.png}
+
+}
+
+}
+
+\subcaption{\label{fig-pbmc1_sc2}}
+\end{minipage}%
+%
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/pbmc3k/sc_3.png}
+
+}
+
+}
+
+\subcaption{\label{fig-pbmc1_sc3}}
+\end{minipage}%
+
+\caption{\label{fig-pbmc1_sc}Screen shots of the \textbf{langevitour} of
+the PBMC3k data set, shows the model in high-D, a video of the tour
+animation is available at (\url{https://youtu.be/Gnhh4hfsbbg}).}
+
+\end{figure}
+
+\begin{itemize}
+\item
+  According to Figure~\ref{fig-pbmc1_sc} (a), the model in high-D space
+  shows two well separated clusters. Also, two clusters are really close
+  to each other (see Figure~\ref{fig-pbmc1_sc} (b)). Furthermore, there
+  are three clusters shown by the model in 2D actually represented in
+  high-D as well (see Figure~\ref{fig-pbmc1_sc} (c)). Another important
+  thing to notice from the model in high-D space is that the non-linear
+  continuity structure in two well-separated clusters.
+\item
+  Figure~\ref{fig-pbmc1_sc} (c) gives evidence that UMAP preserves the
+  global structure. Furthermore, the non-linearity within the clusters
+  shown in Figure~\ref{fig-pbmc1_sc} (a) proves that local structure is
+  also preserved but that's not visually shown by th UMAP layout in 2D.
+\item
+  Can we find a better parameter choice for UMAP which shows this
+  non-linear continuity of the big clusters in 2D space as well?
 \item
   To do that, we are going to fit models for different parameter choice
-  and select the one which have the lowest error.
+  and select the one which have the lowest model error in high-D (see
+  \textbf{?@tbl-errordiffcomb}).
 \end{itemize}
@@
 
-\begin{itemize}
-\item
-  Selected the one with n:neighbor:5, min\_dist:0.99, metric:cosine.
-\item
-  Now, we are going to fit the model for the new parameter choice.
-\end{itemize}
-
 \begin{figure}[H]
@@
 \end{figure}
+
+\begin{itemize}
+\item
+  Selected the one with n:neighbor:5, min\_dist:0.99, metric:cosine (see
+  Figure~\ref{fig-nldervisPBMCUMAPview2}). As shown in
+  Figure~\ref{fig-nldervisPBMCUMAPview2}, the two big clusters have
+  non-linear continuity structures, two big clusters are well-separated
+  but close to each other, and the two cluster which exists in the one
+  big cluster are really close to each other (see
+  Figure~\ref{fig-nldervisPBMCUMAPview2}).
+\item
+  Now, we are going to fit the model for the new parameter choice.
+\end{itemize}
 
@@
 error. Using the elbow method, when the total number of bins is set to
-144, the slope of the Mean Squared Error (MSE) plot experiences a sudden
+588, the slope of the Mean Squared Error (MSE) plot experiences a sudden
 and noticeable change, resembling an elbow-like shape. This point
@@
 
-\caption{\label{fig-nldervisPBMCUMAP3}Model in the 2D space with UMAP
-(\textless\textgreater)}
-
-\end{figure}
+\caption{\label{fig-nldervisPBMCUMAP3}(a) Model in the 2D space overlaid
+with UMAP data, and (b) Error occurred by the high-D model for high-D
+data and the error mapped into 2D NLDR points.}
+
+\end{figure}
+
+\begin{itemize}
+\tightlist
+\item
+  According to Figure~\ref{fig-nldervisPBMCUMAP3} (a), the model
+  constructed in 2D space for the new parameter choice of UMAP shows 5
+  clusters which are really close to each other. The big cluster shows
+  the non-linear continuity structure with two sub clusters. Other three
+  clusters positioned in a structure of non-linearity but really close
+  to each other.
+\end{itemize}
+
+\begin{figure}[H]
+
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/pbmc3k/sc_4.png}
+
+}
+
+}
+
+\subcaption{\label{fig-pbmc2_sc1}}
+\end{minipage}%
+%
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/pbmc3k/sc_5.png}
+
+}
+
+}
+
+\subcaption{\label{fig-pbmc2_sc2}}
+\end{minipage}%
+%
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/pbmc3k/sc_6.png}
+
+}
+
+}
+
+\subcaption{\label{fig-pbmc2_sc3}}
+\end{minipage}%
+
+\caption{\label{fig-pbmc2_sc}Screen shots of the \textbf{langevitour} of
+the PBMC3k data set, shows the model in high-D with new parameter choice
+for UMAP, a video of the tour animation is available at
+(\url{https://youtu.be/lc_6-rIuNHI}).}
+
+\end{figure}
+
+\begin{itemize}
+\item
+  As shown in Figure~\ref{fig-pbmc2_sc}, there are 5 clusters, three of
+  them are well-separated but close to each other as seen in
+  Figure~\ref{fig-nldervisPBMCUMAP3} (a). Also, other two clusters are
+  really close to each other but separated from the early mentioned
+  three clusters.
+\item
+  Next, we are going to investigate whether, the clusters we identified
+  in 2D UMAP layout, can also be identified in high-D as well.
+\item
+  First, I colored the clusters in 2D layout manually. Then, I used
+  spin-and-brush approach with \texttt{detour} to color the clusters in
+  high-D.
+\end{itemize}
 
@@
 (n\_neighbors = 5, min\_dist = 0.99, metric = cosine) applied for PBMC3k
-dataset (a) colored manually in 2D, (b) colored in high-D}
-
-\end{figure}
+dataset (a) colored manually in 2D, (b) colored in high-D.}
+
+\end{figure}
+
+\begin{itemize}
+\item
+  Figure~\ref{fig-nldervisPBMCUMAP3Res} shows the 2D UMAP layout colored
+  with each method. As shown in Figure~\ref{fig-nldervisPBMCUMAP3Res}
+  (b) and the confusion matrix,there are some miss-classifications. What
+  is the reason for that?
+\item
+  For example, the point (ID: 1158) classified as \textbf{cluster 3} in
+  2D UMAP layout, classified as \textbf{cluster 6} when doing the
+  colouring in high-D (see Figure~\ref{fig-nldervisPBMCUMAP3Res}). Why?
+\end{itemize}
+
+\begin{figure}[H]
+
+\begin{minipage}[t]{\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/pbmc3k/sc_7.png}
+
+}
+
+}
+
+\subcaption{\label{fig-pbmc2_sc7}}
+\end{minipage}%
+
+\caption{\label{fig-pbmc2_appsc}Screen shots of app which shows 2D UMAP
+layout and comparison of distances related to the point classified as
+\textbf{cluster 6} in high-D (ID: 1158) and randomly selected point
+which is in \textbf{cluster 6} (ID: 2291).}
+
+\end{figure}
+
+\begin{itemize}
+\tightlist
+\item
+  If we compute the distance between these two points, you will see that
+  it preserves the pair-wise distance (see
+  Figure~\ref{fig-pbmc2_appsc}), which give us a evidence that even
+  though the point 1158, classified as \textbf{cluster 3} in 2D layout,
+  actually should be in \textbf{cluster 6}. This also evident that some
+  misleading can happens with UMAP.
+\end{itemize}
+
+\begin{figure}[H]
+
+\begin{minipage}[t]{\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/pbmc3k/sc_8.png}
+
+}
+
+}
+
+\subcaption{\label{fig-pbmc2_sc8}}
+\end{minipage}%
+
+\caption{\label{fig-pbmc2_appsc1}Screen shots of app which shows 2D UMAP
+layout and comparison of distances related to the point classified as
+\textbf{cluster 5} in high-D (ID: 3) and randomly selected point which
+is in \textbf{cluster 5} (ID: 2198).}
+
+\end{figure}
+
+\begin{itemize}
+\tightlist
+\item
+  For another example, if we compute the distance between these two
+  points, you will see that it preserves the pair-wise distance (see
+  Figure~\ref{fig-pbmc2_appsc1}), which give us a evidence that even
+  though the point 3, classified as \textbf{cluster 4} in 2D layout,
+  actually should be in \textbf{cluster 5}. This also evident that some
+  misleading can happens with UMAP.
+\end{itemize}
 
@@
 \item
-  Why I choose this dataset? Because I need show that whatever NLDR
+  Why I choose this data set? Because I need show that whatever NLDR
   technique that we use, our model can capture the non-linear structure
@@
 
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-84-1.pdf}
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-85-1.pdf}
 

### 2024-03-18 paper.qmd
Added PBMC author's model error

@@
         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) #+
-  #annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
 
@@
 
-# df_bin_train <- df_bin_pbmc
-# names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])
-# 
-# df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), umap_data_with_hb_id_pbmc)
-# 
-# error_df <- df_all_pbmc |>
-#   dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates
-# 
-# for (i in 1:(NCOL(df_bin_train) - 1)) {
-#   
-#   error_df[ , paste0("abs_residual_", "PC_", i)] <- abs(error_df[ , paste0("PC_", i)] - error_df[ , paste0("avg_", "PC_", i)])
-#   
-# }
-# 
-# error_df <- error_df |>
-#   dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "PC_")))))
-# 
-# 
-# error_df <- error_df |>
-#   mutate(type = if_else(total == 0, "no error",
-#                         if_else(total <= 15, "error 1-15",
-#                                 if_else(total <= 20, "error 15-20",
-#                                         if_else(total <= 25, "error 20-25",
-#                                                 if_else(total <= 30, "error 25-30",
-#                                                         if_else(total <= 35, "error 30-35",
-#                                                                 if_else(total <= 40, "error 35-40", "error greter than 40"))))))))
-# 
-# plot_list2_pbmc_error <- error_df |>
-#   mutate(type = factor(type , levels = c("no error", "error 1-15", "error 15-20", "error 20-25", "error 25-30", "error 30-35", "error 35-40", "error greter than 40"))) |>
-#     ggplot(aes(x = UMAP1,
-#                y = UMAP2, color = type))+
-#     geom_point(alpha=0.5, size = 0.1) +
-#     coord_equal() +
-#     theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-#   theme_linedraw() +
-#     theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-#               axis.title.x = element_blank(), axis.title.y = element_blank(),
-#               axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-#               axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-#               panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-#         legend.title = element_text(size=5), #change legend title font size
-#         legend.text = element_text(size=4),
-#          legend.key.height = unit(0.25, 'cm'),
-#          legend.key.width = unit(0.25, 'cm')) +
-#   scale_color_manual(values=c("#fee0d2", "#fcbba1",
-#                                          "#fc9272", "#fb6a4a", "#ef3b2c",
-#                                          "#cb181d", "#a50f15", "#99000d")) 
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "Model in the 2D space overlaid on UMAP data."
+df_bin_train <- df_bin_pbmc
+names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])
+
+df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), umap_data_with_hb_id_pbmc)
+
+error_df <- df_all_pbmc |>
+  dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates
+
+for (i in 1:(NCOL(df_bin_train) - 1)) {
+
+  error_df[ , paste0("abs_residual_", "PC_", i)] <- abs(error_df[ , paste0("PC_", i)] - error_df[ , paste0("avg_", "PC_", i)])
+
+}
+
+error_df <- error_df |>
+  dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "PC_")))))
+
+error_df <- error_df |>
+  mutate(type = if_else(total == 0, "no error",
+                        if_else(total <= 15, "error 1-15",
+                                if_else(total <= 20, "error 15-20",
+                                        if_else(total <= 25, "error 20-25",
+                                                if_else(total <= 30, "error 25-30",
+                                                        if_else(total <= 35, "error 30-35",
+                                                                if_else(total <= 40, "error 35-40", "error greter than 40"))))))))
+
+plot_list2_pbmc_error <- error_df |>
+  mutate(type = factor(type , levels = c("no error", "error 1-15", "error 15-20", "error 20-25", "error 25-30", "error 30-35", "error 35-40", "error greter than 40"))) |>
+    ggplot(aes(x = UMAP1,
+               y = UMAP2, color = type))+
+    geom_point(alpha=0.5, size = 0.1) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=c("#fee0d2", "#fcbba1",
+                                         "#fc9272", "#fb6a4a", "#ef3b2c",
+                                         "#cb181d", "#a50f15", "#99000d")) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "(a) Model in the 2D space overlaid on UMAP data, and (b) Error occurred by the high-D model for high-D data and the error mapped into 2D NLDR points."
 #| label: fig-nldervisPBMCUMAP
@@
 
-trimesh_removed_pbmc_umap 
+trimesh_removed_pbmc_umap + plot_list2_pbmc_error +
+plot_layout(guides='collect', ncol=2) &
+  theme(legend.position='bottom') 
 ```

### 2024-03-18 paper.tex
Added PBMC author's model error

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, boxrule=0pt, frame hidden, sharp corners, breakable]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, breakable, sharp corners, borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced, boxrule=0pt]}{\end{tcolorbox}}\fi
 
@@
 
-\caption{\label{fig-nldervisPBMCUMAP}Model in the 2D space overlaid on
-UMAP data.}
+\caption{\label{fig-nldervisPBMCUMAP}(a) Model in the 2D space overlaid
+on UMAP data, and (b) Error occurred by the high-D model for high-D data
+and the error mapped into 2D NLDR points.}
 

### 2024-03-19 paper.qmd
Added high-D model lifting with example changed with new hex-binning approach

@@
 
+```{r}
+#| warning: false
+#| echo: false
+#| message: false
+
+umap_s_curve_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = UMAP_data, 
+                                    x = "UMAP1", y = "UMAP2"))) |>
+  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
+                  "UMAP2" = "scaled_UMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(UMAP_data))
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+#| message: false
+
+num_bins_x <- 5
+num_bins_y <- 10
+hex_size <- 0.19
+
+hb_obj <- hex_binning(data = umap_s_curve_scaled, x = "UMAP1", 
+                              y = "UMAP2", num_bins_x = num_bins_x, 
+                              num_bins_y = num_bins_y, x_start = NA, y_start = NA, 
+                              buffer_x = NA, buffer_y = NA, hex_size = hex_size, col_start = "UMAP")
+
+## Data set with all possible centroids in the hexagonal grid
+all_centroids_df <- as.data.frame(do.call(cbind, hb_obj$centroids))
+
+## Generate all coordinates of hexagons
+hex_grid <- as.data.frame(do.call(cbind, hb_obj$hex_poly))
+
+## To obtain the standardise counts within hexbins
+counts_df <- as.data.frame(do.call(cbind, hb_obj$std_cts))
+df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
+                                             counts_df = counts_df)
+
+UMAP_data_with_hb_id <- as.data.frame(do.call(cbind, hb_obj$data_hb_id))
+
+model_object <- fit_highd_model( training_data = training_data, 
+                                 nldr_df_with_id = umap_s_curve_scaled, 
+                                 x = "UMAP1", y = "UMAP2", 
+                                 num_bins_x = num_bins_x, 
+                                 num_bins_y = num_bins_y, 
+                                 x_start = NA, y_start = NA, 
+                                 buffer_x = NA, buffer_y = NA, 
+                                 hex_size = hex_size,
+                                 is_rm_lwd_hex = FALSE, 
+                                 benchmark_to_rm_lwd_hex = NA, 
+                                 col_start_2d = "UMAP", 
+                                 col_start_highd = "x")
+
+df_bin_centroids <- model_object$df_bin_centroids
+df_bin <- model_object$df_bin
+
+## Triangulate bin centroids
+tr1_object <- tri_bin_centroids(df_bin_centroids, x = "c_x", y = "c_y")
+tr_from_to_df <- gen_edges(tri_object = tr1_object)
+
+# Create the tibble with x and y coordinates
+tr_df <- tibble::tibble(x = tr1_object$x, y = tr1_object$y)
+
+## Compute 2D distances
+distance <- cal_2d_dist(tr_coord_df = tr_from_to_df, 
+                                start_x = "x_from", start_y = "y_from", 
+                                end_x = "x_to", end_y = "y_to", 
+                                select_vars = c("from", "to", "distance"))
+
+## Map hb_id
+
+df_bin_centroids_n <- df_bin_centroids |>
+  select(hexID) |>
+  mutate(ID = 1:NROW(df_bin_centroids))
+
+tr_from_to_df_n <- left_join(tr_from_to_df, df_bin_centroids_n, by = c("from" = "ID")) |>
+  rename("from_hb_id" = "hexID")
+
+tr_from_to_df_n <- left_join(tr_from_to_df_n, df_bin_centroids_n, by = c("to" = "ID")) |>
+  rename("to_hb_id" = "hexID")
+
+hb_id_selected <- tr_from_to_df_n |>
+  dplyr::filter((from == 1) & (to == 3))
+
+# Filter and label small and long edges
+distance_df_small_edges <- distance |>
+  dplyr::filter((from != 1) | (to != 3)) |>
+  dplyr::mutate(type = "small_edges")
+
+distance_df_long_edges <- distance |>
+  dplyr::filter((from == 1) & (to == 3)) |>
+  dplyr::mutate(type = "long_edges")
+
+# Combine small and long edges
+distance_edges <- dplyr::bind_rows(distance_df_small_edges, distance_df_long_edges)
+
+tr_from_to_df_coord_with_group <- inner_join(tr_from_to_df, distance_edges, by = c("from" = "from", "to" = "to"))
+
+## To generate a data set with high-D and 2D training data
+df_all <- dplyr::bind_cols(training_data |> dplyr::select(-ID), UMAP_data_with_hb_id)
+
+UMAP_data_with_hb_id_selected <- UMAP_data_with_hb_id |> dplyr::filter(hb_id %in% c(hb_id_selected$from_hb_id, hb_id_selected$to_hb_id)) |>
+  dplyr::mutate(select_point = if_else(hb_id == hb_id_selected$from_hb_id, "data1", "data2"))
+
+UMAP_data_with_hb_id_selected_no <- UMAP_data_with_hb_id |> dplyr::filter(!(hb_id %in% c(hb_id_selected$from_hb_id, hb_id_selected$to_hb_id)))
+
+bin_coord_1 <- tr_df |>
+  dplyr::filter((x == hb_id_selected$x_from)) |>
+  dplyr::filter((y == hb_id_selected$y_from))
+
+# bin_coord_1 <- tr_df |>
+#   head(1)
+
+bin_coord_2 <- tr_df |>
+  dplyr::filter((x == hb_id_selected$x_to)) |>
+  dplyr::filter((y == hb_id_selected$y_to))
+
+# bin_coord_2 <- tr_df |>
+#   dplyr::filter(row_number() == 2)
+
+bin_coord <- dplyr::bind_rows(bin_coord_1, bin_coord_2)
+
+a1 <- ggplot(bin_coord, aes(x = x, y = y)) +
+  geom_segment(
+    aes(x = x_from, y = y_from, xend = x_to, yend = y_to, colour = type),
+    data = tr_from_to_df_coord_with_group) +
+  geom_point() +
+  #coord_equal() +
+  coord_cartesian(xlim =c(-0.5, 1.5), ylim = c(-0.5, 2.7)) +
+  geom_point(data = UMAP_data_with_hb_id_selected, aes(x = UMAP1, y = UMAP2, colour = select_point), alpha = 0.5, size = 1) +
+  scale_color_manual(values=c("#e6550d", "#000000", "#e31a1c", "#f6e8c3")) +
+  geom_point(size = 3, colour = "#33a02c") + 
+  #ggtitle("(b)") + 
+  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
+  theme_light() +
+  #coord_equal() +
+  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
+  )
+
+a2 <- ggplot(data = hex_grid, aes(x = x, y = y)) + 
+  geom_polygon(color = "#deebf7", aes(group = hex_poly_id), fill = "#ffffff") +
+  geom_segment(
+    aes(x = x_from, y = y_from, xend = x_to, yend = y_to), colour = "#f6e8c3",
+    data = tr_from_to_df_coord_with_group
+  ) +
+  geom_point(data = UMAP_data_with_hb_id_selected, aes(x = UMAP1, y = UMAP2, colour = select_point), size = 1, alpha = 0.5) +
+  scale_color_manual(values=c("#e6550d", "#000000")) +
+  geom_point(data = df_bin_centroids |> dplyr::filter(hexID %in% c(hb_id_selected$from_hb_id, hb_id_selected$to_hb_id)), aes(x = c_x, y = c_y), colour = "#33a02c", size = 3) +
+  geom_point(data = UMAP_data_with_hb_id_selected_no, aes(x = UMAP1, y = UMAP2), colour = "#6a3d9a", size = 1, alpha = 0.5) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  coord_equal() +
+  theme_light() +
+  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
+  )
+
+## Script to make coloring data points with a line
+
+df1 <- df_all |> 
+  dplyr::filter(hb_id == as.character(hb_id_selected$from_hb_id)) |>
+  dplyr::select(tidyselect::starts_with("x")) |>
+  dplyr::mutate(type = "data1") ## original dataset
+
+df2 <- df_all |> 
+  dplyr::filter(hb_id == as.character(hb_id_selected$to_hb_id)) |>
+  dplyr::select(tidyselect::starts_with("x")) |>
+  dplyr::mutate(type = "data2") ## original dataset
+
+df3 <- df_all |> 
+  dplyr::filter(!(hb_id %in% c(as.character(hb_id_selected$from_hb_id), as.character(hb_id_selected$to_hb_id)))) |>
+  dplyr::select(tidyselect::starts_with("x")) |>
+  dplyr::mutate(type = "data") ## original dataset
+
+df_b <- df_bin |> 
+  dplyr::filter(hb_id %in% df_bin_centroids$hexID) |>
+  dplyr::filter(hb_id %in% c(as.character(hb_id_selected$from_hb_id), as.character(hb_id_selected$to_hb_id))) |>
+  dplyr::select(-hb_id) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+df_exe <- dplyr::bind_rows(df_b, df1, df2, df3)
+
+lg_high <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)], group = df_exe$type, pointSize = append(rep(4, NROW(df_b)), rep(2, NROW(df1) + NROW(df2) + NROW(df3))), levelColors = c("#6a3d9a", "#e6550d", "#000000", "#33a02c"), lineFrom = 1, lineTo = 2, lineColors = "#e31a1c")
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+#| label: fig-wkhighD
+#| fig-cap: How the 2D model lift into high dimensions?
+(a) visualize the points and the hexagonal bin centroids related $7^{th}$ and $12^{th}$ hexagons, (b) visualization of the edge connected the $7^{th}$ and $12^{th}$ hexagons (colored in red) in the triangular mesh.
+A video of tour animation is available at <https://youtu.be/hxU91xNTJL0>.
+
+a2 + a1 +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(ncol = 3) &
+  theme(legend.position='none', plot.tag = element_text(size = 8))
+```
+
 <!--
@@
 #| echo: false
-#| label: fig-wkhighD
+#| label: fig-wkhighD1
 #| fig-cap: How the 2D model lift into high dimensions?
@@
 Furthermore, @fig-modelScurve shows the 2D model constructed using UMAP with the selected total number of bins and how the model overlay on high-D data.
-
-```{r}
-#| warning: false
-#| echo: false
-#| message: false
-
-umap_s_curve_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = UMAP_data, 
-                                    x = "UMAP1", y = "UMAP2"))) |>
-  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
-                  "UMAP2" = "scaled_UMAP2")) |>
-  dplyr::mutate(ID = 1:NROW(UMAP_data))
-
-```
 

### 2024-03-19 paper.tex
Added high-D model lifting with example changed with new hex-binning approach

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, breakable, sharp corners, borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced, boxrule=0pt]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, breakable, borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced, boxrule=0pt, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 bins, the vertices connected in 2D are also connected in high-D (see
-video linked in \textbf{?@fig-wkhighD}).
+video linked in Figure~\ref{fig-wkhighD}).
+
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-wkhighD-1.pdf}
+
+}
+
+\caption{\label{fig-wkhighD}How the 2D model lift into high dimensions?
+(a) visualize the points and the hexagonal bin centroids related
+\(7^{th}\) and \(12^{th}\) hexagons, (b) visualization of the edge
+connected the \(7^{th}\) and \(12^{th}\) hexagons (colored in red) in
+the triangular mesh. A video of tour animation is available at
+\url{https://youtu.be/hxU91xNTJL0}.}
+
+\end{figure}
 
@@
 
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-85-1.pdf}
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-87-1.pdf}
 

### 2024-03-19 paper.qmd
Added total hex bins with new hex-binning approach

@@
 
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+
+num_bins_x <- 3
+num_bins_y <- 4
+hex_size <- 0.62
+
+hb_obj <- hex_binning(data = umap_s_curve_scaled, x = "UMAP1", 
+                      y = "UMAP2", num_bins_x = num_bins_x, 
+                      num_bins_y = num_bins_y, x_start = NA, y_start = NA, 
+                      buffer_x = NA, buffer_y = NA, hex_size = hex_size, col_start = "UMAP")
+
+## Data set with all possible centroids in the hexagonal grid
+all_centroids_df <- as.data.frame(do.call(cbind, hb_obj$centroids))
+counts_df <- as.data.frame(do.call(cbind, hb_obj$std_cts))
+
+## Generate all coordinates of hexagons
+hex_grid <- as.data.frame(do.call(cbind, hb_obj$hex_poly))
+
+hex_full_count_df_loop1 <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
+
+p1 <-  ggplot(data = hex_full_count_df_loop1, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+  geom_point(data = umap_s_curve_scaled, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  #coord_equal() +
+  #coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
+  theme_void() +
+  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6)) +
+  guides(fill = guide_colourbar(title = "Standardized count")) +
+  annotate(geom = 'text', label = "a", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
+
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+
+num_bins_x <- 6
+num_bins_y <- 14
+hex_size <- 0.13
+
+hb_obj <- hex_binning(data = umap_s_curve_scaled, x = "UMAP1", 
+                      y = "UMAP2", num_bins_x = num_bins_x, 
+                      num_bins_y = num_bins_y, x_start = NA, y_start = NA, 
+                      buffer_x = NA, buffer_y = NA, hex_size = hex_size, col_start = "UMAP")
+
+## Data set with all possible centroids in the hexagonal grid
+all_centroids_df <- as.data.frame(do.call(cbind, hb_obj$centroids))
+counts_df <- as.data.frame(do.call(cbind, hb_obj$std_cts))
+
+## Generate all coordinates of hexagons
+hex_grid <- as.data.frame(do.call(cbind, hb_obj$hex_poly))
+
+hex_full_count_df_loop2 <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
+
+p2 <-  ggplot(data = hex_full_count_df_loop2, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+  geom_point(data = umap_s_curve_scaled, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  #coord_equal() +
+  #coord_cartesian(xlim =c(-1, 2.7), ylim = c(-2, 12.2)) +
+  theme_void() +
+  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6)) +
+  guides(fill = guide_colourbar(title = "Standardized count")) +
+  annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
+
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+
+num_bins_x <- 12
+num_bins_y <- 28
+hex_size <- 0.06
+
+hb_obj <- hex_binning(data = umap_s_curve_scaled, x = "UMAP1", 
+                      y = "UMAP2", num_bins_x = num_bins_x, 
+                      num_bins_y = num_bins_y, x_start = NA, y_start = NA, 
+                      buffer_x = NA, buffer_y = NA, hex_size = hex_size, col_start = "UMAP")
+
+## Data set with all possible centroids in the hexagonal grid
+all_centroids_df <- as.data.frame(do.call(cbind, hb_obj$centroids))
+counts_df <- as.data.frame(do.call(cbind, hb_obj$std_cts))
+
+## Generate all coordinates of hexagons
+hex_grid <- as.data.frame(do.call(cbind, hb_obj$hex_poly))
+
+hex_full_count_df_loop3 <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
+
+p3 <-  ggplot(data = hex_full_count_df_loop3, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+  geom_point(data = umap_s_curve_scaled, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  #coord_equal() +
+  #coord_cartesian(xlim =c(-1, 2.7), ylim = c(-2, 12.2)) +
+  theme_void() +
+  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6)) +
+  guides(fill = guide_colourbar(title = "Standardized count")) +
+  annotate(geom = 'text', label = "c", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
+
+```
+
+```{r}
+#| echo: false
+#| label: fig-binsize
+#| fig-pos: H
+#| fig-cap: "Hexbin plots from different number of bins for the **UMAP** embeddings of **S-curve** training data: (a) b = 12 (3, 4), h = 0.62, (b) b = 84 (6, 14), h = 0.13, and (c) b = 336 (12, 28), h = 0.06.
+The hexbins are colored based on the density of points, with darker colors indicating higher point density and yellow color representing lower point density within each bin.
+What is the number of bins that would be effective in representing low-dimensional data?"
+
+p1 + p2 + p3
+```
+
 $$
@@
 #| echo: false
-#| label: fig-binsize
+#| label: fig-binsize1
 #| fig-pos: H
@@
 
-umap_data_with_hb_id_pbmc <- as.data.frame(do.call(cbind, hb_obj_s_curve$data_hb_id))
+umap_data_with_hb_id <- as.data.frame(do.call(cbind, hb_obj_s_curve$data_hb_id))
 

### 2024-03-19 paper.tex
Added total hex bins with new hex-binning approach

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, breakable, borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced, boxrule=0pt, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, boxrule=0pt, sharp corners, breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced]}{\end{tcolorbox}}\fi
 
@@
 to capture the structure of the NLDR data effectively (see
-\textbf{?@fig-binsize} (a)), while if there are too many bins, it may
+Figure~\ref{fig-binsize} (a)), while if there are too many bins, it may
 result in over-fitting the individual points of the NLDR data (see
-\textbf{?@fig-binsize} (c)). Therefore, it is important to determine an
-appropriate number of bins to build an effective model.
+Figure~\ref{fig-binsize} (c)). Therefore, it is important to determine
+an appropriate number of bins to build an effective model.
+
+\begin{figure}[H]
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-binsize-1.pdf}
+
+}
+
+\caption{\label{fig-binsize}Hexbin plots from different number of bins
+for the \textbf{UMAP} embeddings of \textbf{S-curve} training data: (a)
+b = 12 (3, 4), h = 0.62, (b) b = 84 (6, 14), h = 0.13, and (c) b = 336
+(12, 28), h = 0.06. The hexbins are colored based on the density of
+points, with darker colors indicating higher point density and yellow
+color representing lower point density within each bin. What is the
+number of bins that would be effective in representing low-dimensional
+data?}
+
+\end{figure}
 
@@
 
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-87-1.pdf}
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-91-1.pdf}
 

### 2024-03-19 paper.qmd
Rescale total number of bin plots

--- 2024-03-19 paper.qmd
@@
 
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+
+num_bins_x <- 3
+num_bins_y <- 4
+hex_size <- 0.62
+
+hb_obj <- hex_binning(data = umap_s_curve_scaled, x = "UMAP1", 
+                      y = "UMAP2", num_bins_x = num_bins_x, 
+                      num_bins_y = num_bins_y, x_start = NA, y_start = NA, 
+                      buffer_x = NA, buffer_y = NA, hex_size = hex_size, col_start = "UMAP")
+
+## Data set with all possible centroids in the hexagonal grid
+all_centroids_df <- as.data.frame(do.call(cbind, hb_obj$centroids))
+counts_df <- as.data.frame(do.call(cbind, hb_obj$std_cts))
+
+## Generate all coordinates of hexagons
+hex_grid <- as.data.frame(do.call(cbind, hb_obj$hex_poly))
+
+hex_full_count_df_loop1 <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
+
+p1 <-  ggplot(data = hex_full_count_df_loop1, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+  geom_point(data = umap_s_curve_scaled, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  #coord_equal() +
+  coord_cartesian(xlim =c(-0.2, 1.2), ylim = c(-0.1, 2.5)) +
+  theme_void() +
+  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6)) +
+  guides(fill = guide_colourbar(title = "Standardized count")) +
+  annotate(geom = 'text', label = "a", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
+
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+
+num_bins_x <- 6
+num_bins_y <- 14
+hex_size <- 0.13
+
+hb_obj <- hex_binning(data = umap_s_curve_scaled, x = "UMAP1", 
+                      y = "UMAP2", num_bins_x = num_bins_x, 
+                      num_bins_y = num_bins_y, x_start = NA, y_start = NA, 
+                      buffer_x = NA, buffer_y = NA, hex_size = hex_size, col_start = "UMAP")
+
+## Data set with all possible centroids in the hexagonal grid
+all_centroids_df <- as.data.frame(do.call(cbind, hb_obj$centroids))
+counts_df <- as.data.frame(do.call(cbind, hb_obj$std_cts))
+
+## Generate all coordinates of hexagons
+hex_grid <- as.data.frame(do.call(cbind, hb_obj$hex_poly))
+
+hex_full_count_df_loop2 <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
+
+p2 <-  ggplot(data = hex_full_count_df_loop2, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+  geom_point(data = umap_s_curve_scaled, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  #coord_equal() +
+  coord_cartesian(xlim =c(-0.2, 1.2), ylim = c(-0.1, 2.5)) +
+  theme_void() +
+  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6)) +
+  guides(fill = guide_colourbar(title = "Standardized count")) +
+  annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
+
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+
+num_bins_x <- 12
+num_bins_y <- 28
+hex_size <- 0.06
+
+hb_obj <- hex_binning(data = umap_s_curve_scaled, x = "UMAP1", 
+                      y = "UMAP2", num_bins_x = num_bins_x, 
+                      num_bins_y = num_bins_y, x_start = NA, y_start = NA, 
+                      buffer_x = NA, buffer_y = NA, hex_size = hex_size, col_start = "UMAP")
+
+## Data set with all possible centroids in the hexagonal grid
+all_centroids_df <- as.data.frame(do.call(cbind, hb_obj$centroids))
+counts_df <- as.data.frame(do.call(cbind, hb_obj$std_cts))
+
+## Generate all coordinates of hexagons
+hex_grid <- as.data.frame(do.call(cbind, hb_obj$hex_poly))
+
+hex_full_count_df_loop3 <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
+
+p3 <-  ggplot(data = hex_full_count_df_loop3, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+  geom_point(data = umap_s_curve_scaled, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  coord_equal() +
+  #coord_cartesian(xlim =c(-1, 2.7), ylim = c(-2, 12.2)) +
+  theme_void() +
+  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6)) +
+  guides(fill = guide_colourbar(title = "Standardized count")) +
+  annotate(geom = 'text', label = "c", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
+
+```
+
+```{r}
+#| echo: false
+#| label: fig-binsize
+#| fig-pos: H
+#| fig-cap: "Hexbin plots from different number of bins for the **UMAP** embeddings of **S-curve** training data: (a) b = 12 (3, 4), h = 0.62, (b) b = 84 (6, 14), h = 0.13, and (c) b = 336 (12, 28), h = 0.06.
+The hexbins are colored based on the density of points, with darker colors indicating higher point density and yellow color representing lower point density within each bin.
+What is the number of bins that would be effective in representing low-dimensional data?"
+
+p1 + p2 + p3
+```
+
 $$

### 2024-03-19 paper.tex
Rescale total number of bin plots

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, boxrule=0pt, sharp corners, breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, sharp corners, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable, frame hidden]}{\end{tcolorbox}}\fi
 

### 2024-03-21 paper.qmd
Add intro to application data sets

@@
 library(Rtsne)
-library(umap)
+library(uwot)
 library(phateR)
@@
 
-In this article, mainly focus on five NLDR techniques. They are tSNE, UMAP, PHATE, TriMAP [@article02], and Pairwise Controlled Manifold Approximation (PaCMAP) [@article01].
+In this article, mainly focus on five NLDR techniques. They are tSNE, UMAP, PHATE, TriMAP [@article02], and Pairwise Controlled Manifold Approximation (PaCMAP) [@Yingfan2021].
 
@@
 
-PaCMAP [@article01] is different because it adds supervised learning to make a 2D representation while keeping the relationships between pairs of points.
+PaCMAP [@Yingfan2021] is different because it adds supervised learning to make a 2D representation while keeping the relationships between pairs of points.
 It builds a graph using distances between pairs and then makes the 2D representation better using a customizable loss function.
@@
 This MSE plot is useful in determining the effective number of bins required to construct the 2D model.
-@fig-diagnosticpltScurve (b) has created a graph that shows how the Mean Squared Error (MSE) changes depending on the number of bins assigned in the hexagonal grid for UMAP applied to the S-curve dataset.
+@fig-diagnosticpltScurve has created a graph that shows how the Mean Squared Error (MSE) changes depending on the number of bins assigned in the hexagonal grid for UMAP applied to the S-curve dataset.
 As per the rule of thumb mentioned earlier, the effective number of bins is determined to be $84$, with $6$ bins along the x-axis and $14$ bins along the y-axis.
@@
 ## To draw with AIC
-aic_plot <- ggplot(mse_df, aes(x = num_bins, y = aic
-)) +
-  geom_point() +
-  geom_line() +
-  geom_vline(xintercept = 84, linetype="solid",
-                color = "red", size=0.8, alpha = 0.5) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  #scale_colour_manual(values = c("#377eb8")) +
-  ylab("AIC") +
-  xlab("total number of bins")
+# aic_plot <- ggplot(mse_df, aes(x = num_bins, y = aic
+# )) +
+#   geom_point() +
+#   geom_line() +
+#   geom_vline(xintercept = 84, linetype="solid",
+#                 color = "red", size=0.8, alpha = 0.5) +
+#   theme_light() +
+#   theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+#         axis.title = element_text(size = 7),
+#         axis.text = element_text(size = 7)) +
+#   #scale_colour_manual(values = c("#377eb8")) +
+#   ylab("AIC") +
+#   xlab("total number of bins")f
 ## Effective number of bins along x-axis
@@
 
-aic_plot + mse_plot +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', ncol = 2) &
-  theme(legend.position='none', plot.tag = element_text(size = 8))
-```
-
-```{r msebinsnldr}
+mse_plot
+```
+
+```{r}
+#| warning: false
 #| echo: false
 #| message: false
-#| warning: false
-
-# num_bins_for_nldr_methods_df <- tibble::tibble(NLDR_technique = "tSNE", tot_bins = 72, aic = -21882.462)
-# 
-# num_bins_for_nldr_methods_df <- num_bins_for_nldr_methods_df |> 
-#   tibble::add_row(NLDR_technique = "UMAP", tot_bins = 98, aic = -20013.266)
-# 
-# num_bins_for_nldr_methods_df <- num_bins_for_nldr_methods_df |> 
-#   tibble::add_row(NLDR_technique = "PHATE", tot_bins = 90, aic = -15036.91)
-# 
-# num_bins_for_nldr_methods_df <- num_bins_for_nldr_methods_df |> 
-#   tibble::add_row(NLDR_technique = "TriMAP", tot_bins = 154, aic = -23001.90)
-# 
-# num_bins_for_nldr_methods_df <- num_bins_for_nldr_methods_df |> 
-#   tibble::add_row(NLDR_technique = "PaCMAP", tot_bins = 132, aic = -21330.059)
-# 
-# # knitr::kable(
-# #   num_bins_for_nldr_methods_df,
-# #   col.names = c('NLDR technique', 'total number of bins'), format = "latex", 
-# #              caption = "Total number of bins suggested by rule of thumb for each NLDR technique applied to S-curve dataset."
-# # ) |> 
-# #   kable_styling(latex_options = c("striped", "HOLD_position"), position = "center", full_width = FALSE) |>
-# #   row_spec(0, bold = TRUE)
-# num_bins_for_nldr_methods_df$tot_bins <- as.character(num_bins_for_nldr_methods_df$tot_bins)
-# colnames(num_bins_for_nldr_methods_df) <- c('NLDR technique', 'total number of bins', "AIC")
-# num_bins_for_nldr_methods_df <- tibble::remove_rownames(num_bins_for_nldr_methods_df)
-# 
-# xtable(num_bins_for_nldr_methods_df, caption = "Total number of bins suggested by rule of thumb and AIC values for each NLDR technique applied to S-curve dataset.", label = "tbl-msebinsnldr") |>
-#   xtable2kable() |> 
-#   kable_styling(latex_options = c("striped", "HOLD_position"), position = "center", full_width = FALSE) |>
-#   row_spec(0, bold = TRUE) |>
-#   row_spec(4, bold=T, color = "red")
-
-```
-
-<!--Adjusting the parameter $b_1$ provides control over the total number of bins $b$, allowing us to fine-tune the hexagonal grid.-->
-
-<!--first value-->
-
-<!--                  
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| eval: false
-hex_full_count_df_loop1 <- read_rds("data/s_curve/s_curve_hex_3.rds")
-
-p1 <-  ggplot(data = hex_full_count_df_loop1, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
-  geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
-  theme_void() +
-  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=8), #change legend title font size
-        legend.text = element_text(size=6)) +
-  guides(fill = guide_colourbar(title = "Standardized count")) +
-  annotate(geom = 'text', label = "a", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
-
-```
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| eval: false
-hex_full_count_df_loop2 <- read_rds("data/s_curve/s_curve_hex_11.rds")
-
-p2 <-  ggplot(data = hex_full_count_df_loop2, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
-  geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
-  theme_void() +
-  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=8), #change legend title font size
-        legend.text = element_text(size=6)) +
-  guides(fill = guide_colourbar(title = "Standardized count")) +
-  annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
-
-```
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| eval: false
-hex_full_count_df_loop3 <- read_rds("data/s_curve/s_curve_hex_20.rds")
-
-p3 <-  ggplot(data = hex_full_count_df_loop3, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
-  geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
-  theme_void() +
-  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=8), #change legend title font size
-        legend.text = element_text(size=6)) +
-  guides(fill = guide_colourbar(title = "Standardized count")) +
-  annotate(geom = 'text', label = "c", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
-
-```
-
-```{r}
-#| eval: false
-#| echo: false
-#| label: fig-binsize1
-#| fig-pos: H
-#| fig-cap: "Hexbin plots from different number of bins for the **UMAP** projections of **S-curve** training data: (a) b = 32 (4, 8), s = 1.643542, (b) b = 264 (12, 22), s = 1.643542, and (c) b = 840 (21, 40), s = 1.643542.
-The hexbins are colored based on the density of points, with darker colors indicating higher point density and yellow color representing lower point density within each bin.
-What is the number of bins that would be effective in representing low-dimensional data?"
-
-p1 + p2 + p3
-```
-
-```{r}
-#| warning: false
-#| echo: false
-#| eval: false
-
-## UMAP
-## Prediction
-
-data <- read_rds("data/s_curve/s_curve.rds")
-
-shape_value_curve <- calculate_effective_shape_value(.data = UMAP_data,
-                                                   x = UMAP1, y = UMAP2)
-
-num_bins_vec <- 1:13 ## Number of bins along the x-axis
-
-vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names
-
-eval_data_training <- dplyr::bind_rows(vec)[0, ]
-eval_data_training <- eval_data_training |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(num_bins_vec)) {
-
-  model_object <- fit_high_d_model(training_data = training_data, nldr_df_with_id = UMAP_data, x = "UMAP1", y = "UMAP2", num_bins_x = num_bins_vec[i], shape_val = shape_value_curve,
-                             is_bin_centroid = TRUE,
+
+## Decide by looking at MSE plot
+num_bins_x_s_curve <- 6
+num_bins_y_s_curve <- 14
+hex_size_s_curve <- 0.13
+
+hb_obj_s_curve <- hex_binning(data = umap_s_curve_scaled, x = "UMAP1", 
+                      y = "UMAP2", num_bins_x = num_bins_x_s_curve, 
+                      num_bins_y = num_bins_y_s_curve, x_start = NA, y_start = NA, 
+                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_s_curve, col_start = "UMAP")
+
+## Data set with all possible centroids in the hexagonal grid
+all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_s_curve$centroids))
+
+## Generate all coordinates of hexagons
+hex_grid <- as.data.frame(do.call(cbind, hb_obj_s_curve$hex_poly))
+
+## To obtain the standardise counts within hexbins
+counts_df <- as.data.frame(do.call(cbind, hb_obj_s_curve$std_cts))
+df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
+                                             counts_df = counts_df)
+pcentroid_plot_s_curve <- ggplot(data = hex_grid, aes(x = x, y = y)) +
+  geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
+  geom_point(data = df_bin_centroids, aes(x = c_x, y = c_y), color = "#33a02c") +
+  coord_equal() +
+  theme_linedraw() +
+  theme(legend.position = "bottom", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
+
+# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
+#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
+#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
+#   coord_fixed()
+
+umap_data_with_hb_id <- as.data.frame(do.call(cbind, hb_obj_s_curve$data_hb_id))
+
+model_object <- fit_highd_model( training_data = training_data, 
+                                 nldr_df_with_id = umap_s_curve_scaled, 
+                                 x = "UMAP1", y = "UMAP2", 
+                                 num_bins_x = num_bins_x_s_curve, 
+                                 num_bins_y = num_bins_y_s_curve, 
+                                 x_start = NA, y_start = NA, 
+                                 buffer_x = NA, buffer_y = NA, 
+                                 hex_size = hex_size_s_curve,
                                  is_rm_lwd_hex = FALSE, 
                                  benchmark_to_rm_lwd_hex = NA, 
-                             is_avg_high_d = TRUE, column_start_text = "x")
-
-  centroid_df_training <- model_object$df_bin_centroids
-  avg_df_training <- model_object$df_bin
-
-  pred_df_training <- predict_2d_embeddings(test_data = training_data, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "UMAP")
-
-  # pred_df_training <- predict_hex_id(df_bin_centroids = centroid_df_training, nldr_df_test = UMAP_data, x = "UMAP1", y = "UMAP2")
-
-  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, col_start = "x")
-
-  eval_df_training <- eval_df_training |>
-    mutate(num_bins_x = num_bins_vec[i])
-
-  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
-
-}
-
-## Add new column with data types
-
-MSE_df_umap_s_curve <- eval_data_training |>
-  mutate(data_type = "training",
-         method = "UMAP")
-```
-
-```{r}
-#| warning: false
-#| echo: false
-#| eval: false
-
-## To draw with AIC
-aic_plot <- ggplot(MSE_df_umap_s_curve |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
-                                                                                 y = total_error,
-                                                                                 color = method
-)) +
-  geom_point() +
-  geom_line() +
-  geom_vline(xintercept = 144, linetype="dashed", 
-                color = "red", size=0.5) +
-  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+                                 col_start_2d = "UMAP", 
+                                 col_start_highd = "x")
+
+df_bin_centroids_s_curve <- model_object$df_bin_centroids
+df_bin_s_curve <- model_object$df_bin
+
+cell_count_plot <- ggplot(df_bin_centroids_s_curve, aes(x = reorder(as.factor(hexID), -std_counts), y = std_counts)) +
+  geom_quasirandom() + xlab("hexagonal id") + ylab("Standardized cell count") +
+  geom_hline(yintercept = 0.255, colour = "#de2d26") +
+  theme(axis.text = element_text(size = 5),
         axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  scale_colour_manual(values = c("#377eb8")) +
-  ylab("AIC") +
-  xlab("total number of bins")
-## Effective number of bins along x-axis
-
-mse_plot <- ggplot(MSE_df_umap_s_curve |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
-                                       y = total_mse,
-                                       color = method
-)) +
-  geom_point() +
-  geom_line() +
-   geom_vline(xintercept = 144, linetype="dashed", 
-                color = "red", size=0.5) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  scale_colour_manual(values = c("#377eb8")) +
-  ylab("MSE") +
-  xlab("total number of bins")
-
-```
-
--->
+        axis.text.x = element_text(angle = 90))
+
+## Triangulate bin centroids
+tr1_object_s_curve <- tri_bin_centroids(df_bin_centroids_s_curve, x = "c_x", y = "c_y")
+tr_from_to_df_s_curve <- gen_edges(tri_object = tr1_object_s_curve)
+
+## Compute 2D distances
+distance_s_curve <- cal_2d_dist(tr_coord_df = tr_from_to_df_s_curve, 
+                             start_x = "x_from", start_y = "y_from", 
+                             end_x = "x_to", end_y = "y_to", 
+                             select_vars = c("from", "to", "distance"))
+
+distance_plot <- plot_dist(distance_s_curve) +
+  ylab(expression(d^{(2)})) +
+  theme(axis.text = element_text(size = 5),
+        axis.title = element_text(size = 12))
+# 
+# distance_plot
+
+## To find the benchmark value
+benchmark_s_curve <- find_lg_benchmark(distance_edges = distance_s_curve, distance_col = "distance")
+
+# ggplot() +
+# geom_trimesh(data = df_bin_centroids_pbmc, mapping = aes(x = c_x, y = c_y))
+
+trimesh_s_curve_umap <- vis_lg_mesh(distance_edges = distance_s_curve, benchmark_value = benchmark_s_curve, tr_coord_df = tr_from_to_df_s_curve, distance_col = "distance")
+
+trimesh_s_curve_umap <- trimesh_s_curve_umap +
+  theme_linedraw() +
+  theme(legend.position = "bottom", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) +
+  labs(colour = "")
+
+trimesh_removed_s_curve_umap <- vis_rmlg_mesh(distance_edges = distance_s_curve, benchmark_value = benchmark_s_curve, tr_coord_df = tr_from_to_df_s_curve, distance_col = "distance")
+
+trimesh_removed_s_curve_umap <- trimesh_removed_s_curve_umap +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+```
 
@@
 To ensure comprehensive coverage of the NLDR data, it is necessary to select hexagonal bins that have a considerable number of data points within their hexagons.
-To achieve this, the standard number of points within each hexagon is first calculated.
-This standard count is obtained by dividing the number of points within each hexagon by the maximum number of points in the grid (see @eq-equationp2).
+To achieve this, the number of points within each hexagon is first calculated.
+Then, the standard count is computed by dividing the number of points within each hexagon by the maximum number of points in the grid (see @eq-equationp2).
 Next, the bins that have a standard number of points less than a certain benchmark value are removed.
@@
 If the neighboring bins also have low counts, then only those bins will be removed.
-The other identified bins will be kept and used for constructing the 2D model along with the high-density bins (see Appendix for more details).
+The other identified bins will be kept and used for constructing the 2D model along with the high-density bins.
 
@@
 If the increment is not substantial, the decision might lean towards retaining low-density hexagons.
-According to @fig-diagnosticpltScurvelwd, the change in MSE for NLDR techniques is relatively small across different benchmark values, except for the PHATE method.
-
-Consequently, it is not necessary to remove low-density hexagons when constructing the 2D model for the S-curve dataset in NLDR techniques, except for PHATE.
-While there may be some fluctuations in MSE between benchmark values of 0 to 1, the absence of significant changes suggests that maintaining low-density hexagons is preferable.
-However, to choose the best selection, it's recommended to examine the benchmark values around the first local minimum in the MSE curve.
-However, the MSE for PHATE fluctuates when benchmark values range between 0 and 1.
-Therefore, it is necessary to explore the benchmark values surrounding the first local minimum to determine an effective benchmark value.
-There is no universal rule for selecting a process.
-As shown in @fig-diagnosticpltScurvelwd, the MSE values for PHATE fluctuate, and the first local minimum is at $0$.
-Therefore, nearby benchmark values such as $0.01$, $0.04$, and $0.06$ were explored to identify the best value for preserving the data structure.
-In this case, a benchmark value of $0.06$ was chosen.
-The resulting 2D model is shown in @fig-modelScurve (c), while the video link provides a visual representation of how the model appears in high-D.
 
@@
 1. Plot the distribution of the standardized counts
+
+```{r}
+#| echo: false
+#| fig-cap: "Distribution of 2D Euclidean distances."
+#| label: fig-stdctsScurve
+
+cell_count_plot
+```
 
@@
 ```{r}
-#| warning: false
-#| echo: false
-#| message: false
-
-## Decide by looking at MSE plot
-num_bins_x_s_curve <- 6
-num_bins_y_s_curve <- 14
-hex_size_s_curve <- 0.13
-
-hb_obj_s_curve <- hex_binning(data = umap_s_curve_scaled, x = "UMAP1", 
-                      y = "UMAP2", num_bins_x = num_bins_x_s_curve, 
-                      num_bins_y = num_bins_y_s_curve, x_start = NA, y_start = NA, 
-                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_s_curve, col_start = "UMAP")
-
-## Data set with all possible centroids in the hexagonal grid
-all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_s_curve$centroids))
-
-## Generate all coordinates of hexagons
-hex_grid <- as.data.frame(do.call(cbind, hb_obj_s_curve$hex_poly))
-
-## To obtain the standardise counts within hexbins
-counts_df <- as.data.frame(do.call(cbind, hb_obj_s_curve$std_cts))
-df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
-                                             counts_df = counts_df)
-pcentroid_plot_s_curve <- ggplot(data = hex_grid, aes(x = x, y = y)) +
-  geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
-  geom_point(data = df_bin_centroids, aes(x = c_x, y = c_y), color = "#33a02c") +
-  coord_fixed() +
-  theme_linedraw() +
-  theme(legend.position = "bottom", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
-
-# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
-#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
-#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
-#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
-#   coord_fixed()
-
-umap_data_with_hb_id <- as.data.frame(do.call(cbind, hb_obj_s_curve$data_hb_id))
-
-model_object <- fit_highd_model( training_data = training_data, 
-                                 nldr_df_with_id = umap_s_curve_scaled, 
-                                 x = "UMAP1", y = "UMAP2", 
-                                 num_bins_x = num_bins_x_s_curve, 
-                                 num_bins_y = num_bins_y_s_curve, 
-                                 x_start = NA, y_start = NA, 
-                                 buffer_x = NA, buffer_y = NA, 
-                                 hex_size = hex_size_s_curve,
-                                 is_rm_lwd_hex = FALSE, 
-                                 benchmark_to_rm_lwd_hex = NA, 
-                                 col_start_2d = "UMAP", 
-                                 col_start_highd = "x")
-
-df_bin_centroids_s_curve <- model_object$df_bin_centroids
-df_bin_s_curve <- model_object$df_bin
-
-## Triangulate bin centroids
-tr1_object_s_curve <- tri_bin_centroids(df_bin_centroids_s_curve, x = "c_x", y = "c_y")
-tr_from_to_df_s_curve <- gen_edges(tri_object = tr1_object_s_curve)
-
-## Compute 2D distances
-distance_s_curve <- cal_2d_dist(tr_coord_df = tr_from_to_df_s_curve, 
-                             start_x = "x_from", start_y = "y_from", 
-                             end_x = "x_to", end_y = "y_to", 
-                             select_vars = c("from", "to", "distance"))
-
-# distance_plot <- plot_dist(distance_pbmc) +
-#   ylab(expression(d^{(2)})) +
-#   theme(axis.text = element_text(size = 5),
-#         axis.title = element_text(size = 12))
-# 
-# distance_plot
-
-## To find the benchmark value
-benchmark_s_curve <- find_lg_benchmark(distance_edges = distance_s_curve, distance_col = "distance")
-
-# ggplot() +
-# geom_trimesh(data = df_bin_centroids_pbmc, mapping = aes(x = c_x, y = c_y))
-
-trimesh_s_curve_umap <- vis_lg_mesh(distance_edges = distance_s_curve, benchmark_value = benchmark_s_curve, tr_coord_df = tr_from_to_df_s_curve, distance_col = "distance")
-
-trimesh_s_curve_umap <- trimesh_s_curve_umap +
-  theme_linedraw() +
-  theme(legend.position = "bottom", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) +
-  labs(colour = "")
-
-trimesh_removed_s_curve_umap <- vis_rmlg_mesh(distance_edges = distance_s_curve, benchmark_value = benchmark_s_curve, tr_coord_df = tr_from_to_df_s_curve, distance_col = "distance")
-
-trimesh_removed_s_curve_umap <- trimesh_removed_s_curve_umap +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "(a), (b) 2D model."
+#| echo: false
+#| fig-cap: "(a) Full hexagonal grid with bin centroids, (b) Model constructed in 2D with long edges, and (c) Model constructed in 2D after removing the long edges."
 #| label: fig-modelScurve
+#| fig-pos: H
 
@@
 1. Plot the distribution of the 2D distances
+
+```{r}
+#| echo: false
+#| fig-cap: "Distribution of 2D Euclidean distances."
+#| label: fig-distScurve
+
+distance_plot
+```
 
@@
 
+```{r}
+#| warning: false
+#| echo: false
+#| message: false
+#| eval: false
+
+## UMAP
+## Decide by looking at MSE plot
+num_bins_x_s_curve <- 6
+num_bins_y_s_curve <- 14
+hex_size_s_curve <- 0.13
+
+## Possible values for x_start
+max_x_start <- min(umap_s_curve_scaled[["UMAP1"]]) + (sqrt(3) * hex_size_s_curve)
+min_x_start <- min(umap_s_curve_scaled[["UMAP1"]]) - (sqrt(3) * hex_size_s_curve)
+
+## Possible values for y_start
+max_y_start <- min(umap_s_curve_scaled[["UMAP2"]]) + (1.5 * hex_size_s_curve)
+min_y_start <- min(umap_s_curve_scaled[["UMAP2"]]) - (1.5 * hex_size_s_curve)
+
+diff_start_df <- expand.grid(x_start = seq(min_x_start, max_x_start, by = 0.01), 
+                             y_start = seq(min_y_start, max_y_start, by = 0.01))
+
+vec <- stats::setNames(rep("", 4), c("x_start", "y_start", "num_non_empty_bins", "mse"))  ## Define column names
+
+mse_df <- dplyr::bind_rows(vec)[0, ]
+mse_df <- mse_df |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:NROW(diff_start_df)) {
+
+  model_object <- fit_highd_model( training_data = training_data, 
+                                   nldr_df_with_id = umap_s_curve_scaled, 
+                                   x = "UMAP1", y = "UMAP2", 
+                                   num_bins_x = num_bins_x_s_curve, 
+                                   num_bins_y = num_bins_y_s_curve, 
+                                   x_start = diff_start_df[i, 1], y_start = diff_start_df[i, 2], 
+                                   buffer_x = NA, buffer_y = NA, 
+                                   hex_size = hex_size_s_curve,
+                                   is_rm_lwd_hex = FALSE, 
+                                   benchmark_to_rm_lwd_hex = NA, 
+                                   col_start_2d = "UMAP", 
+                                   col_start_highd = "x")
+
+  centroid_df_training <- model_object$df_bin_centroids
+  avg_df_training <- model_object$df_bin
+
+  pred_emb_list <- predict_emb(test_data = training_data, 
+                               df_bin_centroids = centroid_df_training, 
+                               df_bin = avg_df_training, type_NLDR = "UMAP")
+
+  pred_df_training <- as.data.frame(do.call(cbind, pred_emb_list))
+
+  eval_list <- gen_summary(test_data = training_data, 
+                           prediction_df = pred_df_training, 
+                           df_bin = avg_df_training, col_start = "x")
+
+  mse_df <- mse_df |>
+    tibble::add_row(x_start = diff_start_df[i, 1], y_start = diff_start_df[i, 2],
+                    num_non_empty_bins = NROW(centroid_df_training), 
+                    mse = eval_list$mse)
+
+}
+
+## If same total number of bins occurred only select ones with minimum error
+### Obtain duplicate bins
+dupli_bins <- mse_df |> 
+  dplyr::count(num_non_empty_bins) |> 
+  dplyr::filter(n > 1) |> 
+  dplyr::pull(num_non_empty_bins)
+
+### Group split by duplicated bins
+duplicate_df_list <- mse_df |>
+  dplyr::filter(num_non_empty_bins %in% dupli_bins) |>
+  dplyr::arrange(num_non_empty_bins) |>
+  dplyr::group_split(num_non_empty_bins)
+
+### Obtain one row from duplicates which have lowest error and hexsize
+duplicate_df <- data.frame(matrix(nrow = 0, ncol = 0))
+
+for (i in 1:length(duplicate_df_list)) {
+
+  dd <- duplicate_df_list[[i]] |>
+    dplyr::filter(mse == min(duplicate_df_list[[i]]$mse))
+
+  duplicate_df <- dplyr::bind_rows(duplicate_df, dd)
+
+}
+
+### Obtain the mse_df with not duplicated bins
+not_dupli_df <- mse_df |>
+  dplyr::filter(!(num_non_empty_bins %in% dupli_bins))
+
+### Combine duplicated and not duplicated(corrected) bins dfs
+mse_df <- dplyr::bind_rows(not_dupli_df, duplicate_df)
+
+mse_plot <- ggplot(mse_df, aes(x = num_non_empty_bins,
+                               y = mse
+)) +
+  geom_point() +
+  geom_line() +
+  # geom_vline(xintercept = 84, linetype="solid",
+  #            color = "red", linewidth=0.8, alpha = 0.5) +
+  theme_light() +
+  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title = element_text(size = 7),
+        axis.text = element_text(size = 7)) +
+  #scale_colour_manual(values = c("#377eb8")) +
+  ylab("MSE") +
+  xlab("number of non-empty bins")
+
+mse_plot
+```
+
 <!--
@@
 
-To assess how well our method captures and represents the underlying structure of the high-D data, Mean Squared Error (MSE) and Akaike Information Criterion (AIC) are used.
+To assess how well our method captures and represents the underlying structure of the high-D data, Mean Squared Error (MSE) is used.
 When computing MSE, total model error (see @sec-summary) is divided by the number of observations to make it as a mean value (see @eq-equation9).
@@
 $$ {#eq-equation9}
-
-Furthermore, to incorporate with the parameters in our model, which is the non-empty bins, and as a statistical measure used for model selection and comparison, AIC is used (see @eq-equation10).
-
-$$
-\text{AIC} = 2b'p + np * log(\text{MSE})
-$$ {#eq-equation10}
-
-Here, $b'$ signifies the number of bins non-empty bins, $p$ denotes the number of dimensions in the high-D data, and $n$ represents the number of observations.
 
@@
 #| message: false
-#| eval: false
-
-phate_gau_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = PHATE_data_gau, x = "PAHTE1", y = "PAHTE2"))) |>
-  dplyr::rename(c("PAHTE1" = "scaled_PAHTE1", 
-                  "PAHTE2" = "scaled_PAHTE2")) |>
-  dplyr::mutate(ID = 1:NROW(PAHTE_data_gau))
+
+phate_gau_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tibble::as_tibble(PHATE_data_gau), x = "PHATE1", y = "PHATE2"))) |>
+  dplyr::rename(c("PHATE1" = "scaled_PHATE1",
+                  "PHATE2" = "scaled_PHATE2")) |>
+  dplyr::mutate(ID = 1:NROW(PHATE_data_gau))
 
 ## Decide by looking at MSE plot
-num_bins_x_tsne_gau <- 46
-num_bins_y_tsne_gau <- 61
-hex_size_tsne_gau <- 0.013
+num_bins_x_tsne_gau <- 60
+num_bins_y_tsne_gau <- 156
+hex_size_tsne_gau <- 0.01
 ### non-empty 179
@@
 #   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
-#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
+#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red", size = 0.1) +
 #   coord_fixed()
@@
 
-In the field of single-cell studies, a common analysis task involves clustering to identify groups of cells with similar expression profiles.
-Analysts often turn to NLDR techniques to verify and identify these clusters and explore developmental trajectories.
-In clustering workflows, the main objective is to verify the existence of clusters and subsequently identify them as specific cell types by examining the expression of "known" marker genes.
-Marker genes are specific genes or transcripts that are uniquely or highly expressed in particular cell types.
-Marker genes are identified through differential gene expression (DEG) analysis, comparing the expression levels of genes across different cell types.
-In this context, a "faithful" embedding should ideally preserve the topology of the data, ensuring that cells corresponding to the same cell type are situated close to the high-dimensional space.
-
-### Processing steps
-
-To begin our analysis, we installed the Peripheral Blood Mononuclear Cells (pbmc) data set obtained from 10x Genomics using the `SeuratData` R package [@Rahul2019], which facilitates the distribution of data sets in the form of Seurat objects [@Yuhan2021].
-This data set contains 13,714 features (genes) across 2,700 samples (cells) within a single assay.
-The active assay is RNA, with 13,714 features representing different gene expressions.
-Then, the cells that have unique feature counts over 2,500 or less than 200 and cells that have > 5\% mitochondrial counts are filtered.
-After removing unwanted cells from the dataset, the next step is to normalize the data.
-By default, we employ a global-scaling normalization method “LogNormalize” that normalizes the feature expression measurements for each cell by the total expression, multiplies this by a scale factor (10,000 by default), and log-transforms the result.
-
-Next, top 1000 genes are selected by Festem [@Chen2023] DEG method.
-Then, scale the data.
-Then, perform PCA on the scaled data.
-Louvain algorithm is used to cluster the single cells based on the genes selected by Festem.
-Using the Festem-selected genes and 15 PCs, identified 10 clusters in the PBMC3k data and annotated them based on the expression of canonical markers.
-The 10 clusters included immune cells such as naive CD4 T cells, memory CD4 T cells, CD8 T cells, CD14 monocytes, FCGR3A monocytes, Natural Killer (NK) cells, B cells and Dendritic Cells (DC).
-In addition to these common cell types, Festem identified a fine cell type, CD27− CD4+ memory T cells, which were often missed by other methods.
-
-- Human peripheral blood mononclear cells (PBMCs) data
-
-- Intro to the data set
-
-Used Louvain algorithm to cluster the single cells based on the genes selected by Festem method
-
-- Explain what "Evaluate Festem on DEG detection" means, in terms of this data
-
-- How did they "identified 10 clusters"? What were the variables used, PCs or full data or 2D representation, spell this out.
-
-- Used top 1000 genes selected by Festem to run PCA, top 15 PCs (original paper mentioned)
-
-- The "10 clusters included immune cells such as naive CD4 T ..." does this mean that every cluster had all of these, or one cluster was mostly "CD4 T cells", another mostly "memory CD4 T", ...?
-
-- How is your use of this data different from the original application.
-
-- Original paper used this data set to compare different DEG methods, but we need to look how the performance of UMAP with the author's selected parameter choice.
-
-- Why 15 PCs? Is that from the original paper? YES
-
-@Chen2023
-
-- Intro to the data set (This data set contains 13,714 features across 2,700 samples within a single assay. The active assay is RNA, with 13,714 features representing different gene expressions.)
-
-- 2622 cells
-
-- Evaluate Festem on DEG detection (Festem enabled identification of often-missed fine cell types)
-
-- Using the Festem-selected genes, identified 10 clusters in the PBMCsk data and annotated them based on the expression of canonical markers
-
-- The 10 clusters included immune cells such as naive CD4 T cells, memory CD4 T cells, CD8 T cells, CD14 monocytes, FCGR3A monocytes, Natural Killer (NK) cells, B cells and Dendritic Cells (DC).
-
-- In addition to these common cell types, Festem identified a fine cell type, CD27− CD4+ memory T cells, which were often missed by other methods.
-
-- The CD27. CD4+ memory T cells identified by Festem expressed common marker genes (IL7R
-and S100A4) of memory T cells, but did not express CD27. These cells also had downregulated
-expression of SELL, CCR7, MAL and LEF1, and upregulated expression of CCL5 (Fig. 4B) and
-thus were the CD27- CD4+ memory T cells in the literature (36). The CD27- CD4+ memory T
-cells were known to be at a more differentiated state and have stronger antigen-recall responses
-than their CD27+ counterparts.
-
-- 15 PCs
-
-- Num bins along the x-axis = 23, shape parameter = 0.8772751
-
-- learned from the model: There are 3 well-separated clusters in 2D, but if look at the model in high-D space, the three clusters are much closer.
-Also, there is some continuity within the clustering structure and it didn't capture well.
-Therefore UMAP with n:neighbour 30 is not a good representation for this data set.
-
-- The most important parameter is n_neighbors - the number of approximate nearest neighbors used to construct the initial high-dimensional graph.
-It effectively controls how UMAP balances local versus global structure - low values will push UMAP to focus more on local structure by constraining the number of neighboring points considered when analyzing the data in high dimensions, while high values will push UMAP towards representing the big-picture structure while losing fine detail.
-
-- The second parameter we’ll investigate is min_dist, or the minimum distance between points in low-dimensional space.
-This parameter controls how tightly UMAP clumps points together, with low values leading to more tightly packed embeddings.
-Larger values of min_dist will make UMAP pack points together more loosely, focusing instead on the preservation of the broad topological structure.
-
-- The third parameter is metric.
-This parameter plays a crucial role in determining how distances are computed within the ambient space of the input data (The term "ambient space" refers to the original or initial space in which the data exists before any transformation or dimensionality reduction).
-
-- If it is completely a mess, then that is also further evidence that the fit is poor in high-dimensions.
-
-<!--https://dicook.github.io/mulgar_book/7-spin-and-brush.html#fig-penguins-bs-detourr-->
-
-### Approach followed
-
-- First, referred to @Chen2023 paper which shows UMAP representation of PBMC3k data with 15 PCs and parameter choice of n:neighbor:30, min_dist:0.1, metric:cosine.
-
-- Observed that there is three well-separated clusters.
-
-- Second, visualise the 15 PCs in high-dimensional space with our model and investigate whether can I see the three well-separated clusters or is it showing something different than what I saw in UMAP view suggested by the author.
-
-- Learned from the model: Yes, it's different.
-There are three clusters but they are really close to each other.
-Also, in the largest cluster, observed non-linear continuity structure, which didn't captured by the author's UMAP representation.
-
-- Learned from the data: Visualise with detour and colored the clusters in high-D. Identified more than three clusters which is suprising.
-
-- In that case, need to suggest a better UMAP representation than author's suggestion.
-
-- To find a better UMAP representation, I need to investigate different parameter choices for UMAP (or need to try another NLDR technique, but here my objective is to suggest the best parameter choice with UMAP).
-
-- Therefore, I fitted our model for different UMAP parameter choices, record the error, and selected the parameter choice that gives the lowest error which is n:neighbor:5, min_dist:0.99, metric:cosine.
-
-- Then, construct our model for this UMAP data (n:neighbor:5, min_dist:0.99, metric:cosine).
-
-- In this representation, you can see that the clusters are closer, and the continuity of clusters are there.
-
-- Also, what missed from looking at high-D data with detour, in this UMAP representation I can see more than three clusters.
-
-- Then, need to check how correct am I identify the clusters in 2D and is that matched that I identify in high-D.
-
-- Generate a confusion matrix.
-
-### Story
-
-- UMAP representation of PBMC3k data with 15 PCs and parameter choice of n:neighbor:30, min_dist:0.1, metric:cosine (author's suggestion) shows three well-separated clusters (see @fig-nldervisPBMCUMPview).
-The question we need to answer is: Does this UMAP representation preserve the actual data structure, or is something missing or misidentified or misleading?
-
+<!-- UMAP layout with author's suggested parameter choice-->
 ```{r}
@@
 ## Import data
-training_data_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_pca_50.rds")
-training_data_pbmc <- training_data_pbmc[, 1:15] |>
+training_data_pbmc <- read_rds("data/pbmc3k/pbmc_pca_50.rds")
+training_data_pbmc <- training_data_pbmc[, 1:9] |>
   mutate(ID = 1:NROW(training_data_pbmc))
 
-umap_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_umap.rds")
+umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_30_min_dist_0.3.rds")
 umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc, 
@@
 
-plot_list2_pbmc <- umap_pbmc_scaled |>
+umap_plot_pbmc <- umap_pbmc_scaled |>
   ggplot(aes(x = UMAP1,
@@
         legend.key.height = unit(0.25, 'cm'),
-        legend.key.width = unit(0.25, 'cm')) #+
-  # scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
-  # geom_text(aes(x=UMAP1,y = UMAP2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) +
-  #annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
+        legend.key.width = unit(0.25, 'cm')) 
 ```
@@
 #| fig-cap: "(a) 2D layout from UMAP (n_neighbors = 30, min_dist = 0.3) applied for the PBMC3k dataset. Is this a best representation of the original data?"
-#| label: fig-nldervisPBMCUMPview
+#| label: fig-umapauthor
 #| fig-pos: H
 
-plot_list2_pbmc 
-```
-
-- To investigate this, we are going to visualise the model in the data space, which means that the model generated from UMAP is visualised in high-D with original data.
-
-- To fit the model, first we need to decide how many number bins should allocate along the x and y axes. To make an initial choice, we use MSE plot (@fig-diagnosticpltPBMC (b)).
-
+umap_plot_pbmc 
+```
+
+<!-- MSE plot to find which is the effective number of bins to use-->
 ```{r}
@@
 
-vec <- stats::setNames(rep("", 6), c("num_bins", "aic", "mse", "num_bins_x", "num_bins_y", "hex_size"))  ## Define column names
+vec <- stats::setNames(rep("", 7), c("num_bins", "error", "mse", "num_bins_x", "num_bins_y", "hex_size", "num_non_empty_bins"))  ## Define column names
 
@@
   model_object <- fit_highd_model( training_data = training_data_pbmc, 
-                                   nldr_df_with_id = umap_pbmc_scaled |> dplyr::select(-cell_label), 
+                                   nldr_df_with_id = umap_pbmc_scaled, 
                                    x = "UMAP1", y = "UMAP2", 
@@
     tibble::add_row(num_bins = num_bins_x * num_bins_y,
-                    aic = eval_list$aic,
+                    error = eval_list$error,
                     mse = eval_list$mse,
@@
                     num_bins_y = num_bins_y,
-                    hex_size = hex_size_vec[i])
+                    hex_size = hex_size_vec[i],
+                    num_non_empty_bins = NROW(centroid_df_training))
 
@@
 #| echo: false
-
-## To draw with AIC
-aic_plot_pbmc <- ggplot(mse_df_pbmc, aes(x = num_bins, y = aic
-)) +
-  geom_point() +
-  geom_line() +
-  geom_vline(xintercept = 588, linetype="solid",
-                color = "red", size=0.8, alpha = 0.5) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  #scale_colour_manual(values = c("#377eb8")) +
-  ylab("AIC") +
-  xlab("total number of bins")
-## Effective number of bins along x-axis
 
@@
         axis.text = element_text(size = 7)) +
-  #scale_colour_manual(values = c("#377eb8")) +
   ylab("MSE") +
@@
 #| fig-cap: "Goodness of fit statistics from UMAP applied to training PBMC3k dataset.
-What is the effective number of bins in each NLDR technique to create a 2D model?
+What is the effective number of bins in UMAP to create a 2D model?
 The MSE plot have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
 Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error.
-Using the elbow method, when the total number of bins is set to 588, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
+Using the elbow method, when the total number of bins is set to 588, the slope of the MSE plot experiences a sudden and noticeable change, resembling an elbow-like shape.
 This point indicates that adding less bins does not enough to capture the data structure."
-#| label: fig-diagnosticpltPBMC
-##| out-width: 100%
+#| label: fig-msepltpbmc
 #| fig-pos: H
 
-aic_plot_pbmc + mse_plot_pbmc +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', ncol = 2) &
-  theme(legend.position='none', plot.tag = element_text(size = 8))
-```
-
-- Then, the model is constructed by following all the steps in 2D and lifts the model into high-D.
-
+mse_plot_pbmc
+```
+
+<!-- Fit the model and compute error-->
 ```{r}
@@
 
-hb_obj_pbmc <- hex_binning(data = umap_pbmc_scaled |> dplyr::select(-cell_label), x = "UMAP1", 
+hb_obj_pbmc <- hex_binning(data = umap_pbmc_scaled, x = "UMAP1", 
                       y = "UMAP2", num_bins_x = num_bins_x_pbmc, 
@@
 model_object <- fit_highd_model( training_data = training_data_pbmc, 
-                                 nldr_df_with_id = umap_pbmc_scaled |> dplyr::select(-cell_label), 
+                                 nldr_df_with_id = umap_pbmc_scaled, 
                                  x = "UMAP1", y = "UMAP2", 
@@
   mutate(type = if_else(total == 0, "no error",
-                        if_else(total <= 15, "error 1-15",
+                        if_else(total <= 5, "error 0-5",
+                                if_else(total <= 10, "error 5-10",
+                                        if_else(total <= 15, "error 10-15",
                                                 if_else(total <= 20, "error 15-20",
                                                         if_else(total <= 25, "error 20-25",
-                                                if_else(total <= 30, "error 25-30",
-                                                        if_else(total <= 35, "error 30-35",
-                                                                if_else(total <= 40, "error 35-40", "error greter than 40"))))))))
+                                                                "error greter than 25")))))))
 
 plot_list2_pbmc_error <- error_df |>
-  mutate(type = factor(type , levels = c("no error", "error 1-15", "error 15-20", "error 20-25", "error 25-30", "error 30-35", "error 35-40", "error greter than 40"))) |>
+  mutate(type = factor(type , levels = c("no error", "error 0-5", "error 5-10", "error 10-15", "error 15-20", "error 20-25", "error greter than 25"))) |>
     ggplot(aes(x = UMAP1,
@@
                                          "#fc9272", "#fb6a4a", "#ef3b2c",
-                                         "#cb181d", "#a50f15", "#99000d")) +
+                                         "#cb181d", "#a50f15")) +
   annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
@@
 
-- As shown in @fig-nldervisPBMCUMAP, there are three well-separated clusters (according to the model in 2D space).
-
-::: {#fig-pbmc1_sc layout-ncol="3" fig-pos="H"}
-![](figures/pbmc3k/sc_1.png){#fig-pbmc1_sc1}
-
-![](figures/pbmc3k/sc_2.png){#fig-pbmc1_sc2}
-
-![](figures/pbmc3k/sc_3.png){#fig-pbmc1_sc3}
-
-Screen shots of the **langevitour** of the PBMC3k data set, shows the model in high-D, a video of the tour animation is available at (<https://youtu.be/Gnhh4hfsbbg>).
-:::
-
-- According to @fig-pbmc1_sc (a), the model in high-D space shows two well separated clusters.
-Also, two clusters are really close to each other (see @fig-pbmc1_sc (b)).
-Furthermore, there are three clusters shown by the model in 2D actually represented in high-D as well (see @fig-pbmc1_sc (c)).
-Another important thing to notice from the model in high-D space is that the non-linear continuity structure in two well-separated clusters.
-
-- @fig-pbmc1_sc (c) gives evidence that UMAP preserves the global structure.
-Furthermore, the non-linearity within the clusters shown in @fig-pbmc1_sc (a) proves that local structure is also preserved but that's not visually shown by th UMAP layout in 2D.
-
-- Can we find a better parameter choice for UMAP which shows this non-linear continuity of the big clusters in 2D space as well?
-
-- To do that, we are going to fit models for different parameter choice and select the one which have the lowest model error in high-D (see @tbl-errordiffcomb).
-
-```{r}
-#| warning: false
-#| echo: false
-
-error_df <- read_csv("data/pbmc/pbmc_3k_festem/error_pbmc_umap.csv", col_names = FALSE)
+<!--compute absolute error for different parameter choices-->
+
+```{r}
+#| warning: false
+#| echo: false
+
+error_df <- read_csv("data/pbmc3k/error_df.csv", col_names = FALSE)
+names(error_df) <- c("num_bins", "error", "mse", "num_bins_x", "num_bins_y", "hex_size", "num_non_empty_bins", "n_neighbors", "min_dist")
 
 error_df <- error_df |>
-  mutate(X4 = round(X4/100, 2),
-         X1 = as.integer(X1)) |>
-  arrange(X4)
-
-colnames(error_df) <- c("n_neighbors", "min_dist", "metric", "error (x 100)")
-
-kable(error_df, caption = "errors for different parameter combinations", label = "tbl-errordiffcomb") |>
-  #xtable2kable() |>
-  kable_styling(latex_options = c("striped", "HOLD_position"), position = "center", full_width = FALSE) |>
-  row_spec(0, bold = TRUE) |>
-  row_spec(7, bold=T, color = "red")
-```
-
-<!--UMAP parm option 1-->
-
-```{r}
-#| warning: false
-#| echo: false
-
-## Import data
-training_data_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_pca_50.rds")
-training_data_pbmc <- training_data_pbmc[, 1:15] |>
-  mutate(ID = 1:NROW(training_data_pbmc))
-
-umap_pbmc <- read_rds("data/pbmc/pbmc_umap_5_min_dist_0.99_metric_cosine.rds")
-umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc, 
-                                    x = "UMAP1", y = "UMAP2"))) |>
-  dplyr::mutate(cell_label = umap_pbmc$cell_label) |>
-  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
-                  "UMAP2" = "scaled_UMAP2")) |>
-  dplyr::mutate(ID = 1:NROW(umap_pbmc))
-
-plot_list2_pbmc <- umap_pbmc_scaled |>
-    ggplot(aes(x = UMAP1,
-               y = UMAP2))+
-    geom_point(alpha=0.5) +
-    coord_equal() +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) 
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "(a) 2D layout from UMAP (n_neighbors = 5, min_dist = 0.99, metric = cosine) applied for the PBMC3k dataset. Is this a best representation of the original data?"
-#| label: fig-nldervisPBMCUMAPview2
-#| fig-pos: H
-
-plot_list2_pbmc
-
-```
-
-- Selected the one with n:neighbor:5, min_dist:0.99, metric:cosine (see @fig-nldervisPBMCUMAPview2).
-As shown in @fig-nldervisPBMCUMAPview2, the two big clusters have non-linear continuity structures, two big clusters are well-separated but close to each other, and the two cluster which exists in the one big cluster are really close to each other (see @fig-nldervisPBMCUMAPview2).
-
-- Now, we are going to fit the model for the new parameter choice.
-
-```{r}
-#| warning: false
-#| echo: false
-#| message: false
-
-## UMAP
-## Prediction
-
-#tot_num_bin_vec <- 2:51
-hex_size_vec <- seq(0.02, 2, by = 0.01)
-
-vec <- stats::setNames(rep("", 6), c("num_bins", "aic", "mse", "num_bins_x", "num_bins_y", "hex_size"))  ## Define column names
-
-mse_df_pbmc <- dplyr::bind_rows(vec)[0, ]
-mse_df_pbmc <- mse_df_pbmc |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(hex_size_vec)) {
-
-  num_bin_list <- calc_bins(data = umap_pbmc_scaled, 
-            x = "UMAP1", y = "UMAP2", 
-            hex_size = hex_size_vec[i], buffer_x = NA, buffer_y = NA)
-
-  num_bins_x <- num_bin_list$num_x
-  num_bins_y <- num_bin_list$num_y
-
-  model_object <- fit_highd_model( training_data = training_data_pbmc, 
-                                   nldr_df_with_id = umap_pbmc_scaled |> dplyr::select(-cell_label), 
-                                   x = "UMAP1", y = "UMAP2", 
-                                   num_bins_x = num_bins_x, 
-                                   num_bins_y = num_bins_y, 
-                                   x_start = NA, y_start = NA, 
-                                   buffer_x = NA, buffer_y = NA, 
-                                   hex_size = hex_size_vec[i],
-                                   is_rm_lwd_hex = FALSE, 
-                                   benchmark_to_rm_lwd_hex = NA, 
-                                   col_start_2d = "UMAP", 
-                                   col_start_highd = "PC_")
-
-  centroid_df_training <- model_object$df_bin_centroids
-  avg_df_training <- model_object$df_bin
-
-  pred_emb_list <- predict_emb(test_data = training_data_pbmc, 
-                                  df_bin_centroids = centroid_df_training, 
-                                  df_bin = avg_df_training, type_NLDR = "UMAP")
-
-  pred_df_training <- as.data.frame(do.call(cbind, pred_emb_list))
-
-  eval_list <- gen_summary(test_data = training_data_pbmc, 
-                                  prediction_df = pred_df_training, 
-                                  df_bin = avg_df_training, col_start = "PC_")
-
-  mse_df_pbmc <- mse_df_pbmc |>
-    tibble::add_row(num_bins = num_bins_x * num_bins_y,
-                    aic = eval_list$aic,
-                    mse = eval_list$mse,
-                    num_bins_x = num_bins_x,
-                    num_bins_y = num_bins_y,
-                    hex_size = hex_size_vec[i])
-
-}
-
-## If same total number of bins occurred only select ones with minimum error
-### Obtain duplicate bins
-dupli_bins <- mse_df_pbmc |> 
-  dplyr::count(num_bins) |> 
-  dplyr::filter(n > 1) |> 
-  dplyr::pull(num_bins)
-
-### Group split by duplicated bins
-duplicate_df_list <- mse_df_pbmc |>
-  dplyr::filter(num_bins %in% dupli_bins) |>
-  dplyr::arrange(num_bins) |>
-  dplyr::group_split(num_bins)
-
-### Obtain one row from duplicates which have lowest error and hexsize
-duplicate_df <- data.frame(matrix(nrow = 0, ncol = 0))
-
-for (i in 1:length(duplicate_df_list)) {
-
-  dd <- duplicate_df_list[[i]] |>
-    dplyr::filter(mse == min(duplicate_df_list[[i]]$mse)) |>
-    dplyr::filter(hex_size == min(duplicate_df_list[[i]]$hex_size))
-
-  duplicate_df <- dplyr::bind_rows(duplicate_df, dd)
-
-}
-
-### Obtain the mse_df with not duplicated bins
-not_dupli_df <- mse_df_pbmc |>
-  dplyr::filter(!(num_bins %in% dupli_bins))
-
-### Combine duplicated and not duplicated(corrected) bins dfs
-mse_df_pbmc <- dplyr::bind_rows(not_dupli_df, duplicate_df)
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-## To draw with AIC
-aic_plot_pbmc <- ggplot(mse_df_pbmc, aes(x = num_bins, y = aic
-)) +
+  dplyr::mutate(type = paste0(n_neighbors, "_", min_dist))
+
+abs_error_plot_pbmc <- ggplot(error_df, aes(x = num_non_empty_bins,
+                                       y = error, group = type,
+colour = type)) +
   geom_point() +
   geom_line() +
-  geom_vline(xintercept = 588, linetype="solid",
-                color = "red", size=1, alpha = 0.5) +
+   # geom_vline(xintercept = 588, linetype="solid",
+   #              color = "red", size=0.8, alpha = 0.5) +
   theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
         axis.title = element_text(size = 7),
         axis.text = element_text(size = 7)) +
-  #scale_colour_manual(values = c("#377eb8")) +
-  ylab("AIC") +
-  xlab("total number of bins")
-## Effective number of bins along x-axis
-
-mse_plot_pbmc <- ggplot(mse_df_pbmc, aes(x = num_bins,
-                                       y = mse
-)) +
-  geom_point() +
-  geom_line() +
-   geom_vline(xintercept = 588, linetype="solid",
-                color = "red", size=1, alpha = 0.5) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  #scale_colour_manual(values = c("#377eb8")) +
-  ylab("MSE") +
-  xlab("total number of bins")
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset.
-What is the effective number of bins in each NLDR technique to create a 2D model?
-The MSE plot have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
-Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error.
-Using the elbow method, when the total number of bins is set to 588, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
-This point indicates that adding less bins does not enough to capture the data structure."
-#| label: fig-diagnosticpltPBMC3
-##| out-width: 100%
-#| fig-pos: H
-
-aic_plot_pbmc + mse_plot_pbmc +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', ncol = 2) &
-  theme(legend.position='none', plot.tag = element_text(size = 8))
-```
-
-```{r}
-#| warning: false
-#| echo: false
-#| message: false
-
-## Decide by looking at MSE plot
-num_bins_x_pbmc <- 21
-num_bins_y_pbmc <- 28
-hex_size_pbmc <- 0.03
-
-hb_obj_pbmc <- hex_binning(data = umap_pbmc_scaled |> dplyr::select(-cell_label), x = "UMAP1", 
-                      y = "UMAP2", num_bins_x = num_bins_x_pbmc, 
-                      num_bins_y = num_bins_y_pbmc, x_start = NA, y_start = NA, 
-                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_pbmc, col_start = "UMAP")
-
-## Data set with all possible centroids in the hexagonal grid
-all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_pbmc$centroids))
-
-## Generate all coordinates of hexagons
-hex_grid <- as.data.frame(do.call(cbind, hb_obj_pbmc$hex_poly))
-
-## To obtain the standardise counts within hexbins
-counts_df <- as.data.frame(do.call(cbind, hb_obj_pbmc$std_cts))
-df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
-                                             counts_df = counts_df)
-
-# ggplot(data = hex_grid, aes(x = x, y = y)) +
-#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
-#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
-#   coord_fixed()
-
-hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
-
-# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
-#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
-#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
-#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
-#   coord_fixed()
-
-umap_data_with_hb_id_pbmc <- as.data.frame(do.call(cbind, hb_obj_pbmc$data_hb_id))
-
-model_object <- fit_highd_model( training_data = training_data_pbmc, 
-                                 nldr_df_with_id = umap_pbmc_scaled |> dplyr::select(-cell_label), 
-                                 x = "UMAP1", y = "UMAP2", 
-                                 num_bins_x = num_bins_x_pbmc, 
-                                 num_bins_y = num_bins_y_pbmc, 
-                                 x_start = NA, y_start = NA, 
-                                 buffer_x = NA, buffer_y = NA, 
-                                 hex_size = hex_size_pbmc,
-                                 is_rm_lwd_hex = FALSE, 
-                                 benchmark_to_rm_lwd_hex = NA, 
-                                 col_start_2d = "UMAP", 
-                                 col_start_highd = "PC_")
-
-df_bin_centroids_pbmc <- model_object$df_bin_centroids
-df_bin_pbmc <- model_object$df_bin
-
-benchmark_value_rm_lwd <- stats::quantile(df_bin_centroids_pbmc$std_counts, 
-                probs = c(0,0.25,0.5,0.75,1), names = FALSE)[2]
-
-df_bin_centroids_low <- df_bin_centroids_pbmc |>
-  dplyr::filter(std_counts <= benchmark_value_rm_lwd)
-
-identify_rm_bins <- find_low_dens_hex(df_bin_centroids_all = df_bin_centroids_pbmc, 
-                                      num_bins_x = num_bins_x_pbmc, 
-                                      df_bin_centroids_low = df_bin_centroids_low)
-
-df_bin_centroids_pbmc <- df_bin_centroids_pbmc |>
-  dplyr::filter(!(hexID %in% identify_rm_bins))
-
-## Triangulate bin centroids
-tr1_object_pbmc <- tri_bin_centroids(df_bin_centroids_pbmc, x = "c_x", y = "c_y")
-tr_from_to_df_pbmc <- gen_edges(tri_object = tr1_object_pbmc)
-
-## Compute 2D distances
-distance_pbmc <- cal_2d_dist(tr_coord_df = tr_from_to_df_pbmc, 
-                             start_x = "x_from", start_y = "y_from", 
-                             end_x = "x_to", end_y = "y_to", 
-                             select_vars = c("from", "to", "distance"))
-
-# distance_plot <- plot_dist(distance_pbmc) +
-#   ylab(expression(d^{(2)})) +
-#   theme(axis.text = element_text(size = 5),
-#         axis.title = element_text(size = 12))
-# 
-# distance_plot
-
-## To find the benchmark value
-benchmark_pbmc <- find_lg_benchmark(distance_edges = distance_pbmc, distance_col = "distance")
-benchmark_pbmc <- 0.08
-
-# ggplot() +
-# geom_trimesh(data = df_bin_centroids_pbmc, mapping = aes(x = c_x, y = c_y))
-
-trimesh_removed_pbmc_umap <- vis_rmlg_mesh(distance_edges = distance_pbmc, benchmark_value = benchmark_pbmc, tr_coord_df = tr_from_to_df_pbmc, distance_col = "distance")
-
-trimesh_removed_pbmc_umap <- trimesh_removed_pbmc_umap +
-  geom_point(data = umap_pbmc_scaled, aes(x = UMAP1, y = UMAP2), alpha = 0.3) +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-df_bin_train <- df_bin_pbmc
-names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])
-
-df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), umap_data_with_hb_id_pbmc)
-
-error_df <- df_all_pbmc |>
-  dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates
-
-# prediction_df_join <- prediction_df_join |>
-#   dplyr::left_join(data, by = c("ID" = "ID")) ## Map high-D data
-
-for (i in 1:(NCOL(df_bin_train) - 1)) {
-
-  error_df[ , paste0("abs_residual_", "PC_", i)] <- abs(error_df[ , paste0("PC_", i)] - error_df[ , paste0("avg_", "PC_", i)])
-
-}
-
-error_df <- error_df |>
-  dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "PC_")))))
-
-# library(ggbeeswarm)
-# error_df$group <- "1"
-# ggplot(error_df, aes(x = group, y = total)) +
-#   geom_quasirandom()+
-#   ylim(0, max(unlist(error_df$total))+ 0.5) + coord_flip()
-
-### The minimum error is 0 and the maximum is 42.17439
-### There is lot of points with error 0,
-
-error_df <- error_df |>
-  mutate(type = if_else(total == 0, "no error",
-                        if_else(total <= 15, "error 1-15",
-                                if_else(total <= 20, "error 15-20",
-                                        if_else(total <= 25, "error 20-25",
-                                                if_else(total <= 30, "error 25-30",
-                                                        if_else(total <= 35, "error 30-35",
-                                                                if_else(total <= 40, "error 35-40", "error greter than 40"))))))))
-
-plot_list2_pbmc_error <- error_df |>
-  mutate(type = factor(type , levels = c("no error", "error 1-15", "error 15-20", "error 20-25", "error 25-30", "error 30-35", "error 35-40", "error greter than 40"))) |>
-    ggplot(aes(x = UMAP1,
-               y = UMAP2, color = type))+
-    geom_point(alpha=0.5, size = 0.1) +
-    coord_equal() +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#fee0d2", "#fcbba1",
-                                         "#fc9272", "#fb6a4a", "#ef3b2c",
-                                         "#cb181d", "#a50f15", "#99000d")) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-# show_langevitour(df_all_pbmc, df_bin_pbmc, df_bin_centroids_pbmc, benchmark_pbmc, distance_pbmc, "distance", col_start = "PC_")
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "(a) Model in the 2D space overlaid with UMAP data, and (b) Error occurred by the high-D model for high-D data and the error mapped into 2D NLDR points."
-#| label: fig-nldervisPBMCUMAP3
-#| out-width: 100%
-#| fig-pos: H
-
-trimesh_removed_pbmc_umap + plot_list2_pbmc_error +
-plot_layout(guides='collect', ncol=2) &
-  theme(legend.position='bottom')
-
-```
-
-- According to @fig-nldervisPBMCUMAP3 (a), the model constructed in 2D space for the new parameter choice of UMAP shows 5 clusters which are really close to each other.
-The big cluster shows the non-linear continuity structure with two sub clusters.
-Other three clusters positioned in a structure of non-linearity but really close to each other.
-
-::: {#fig-pbmc2_sc layout-ncol="3" fig-pos="H"}
-![](figures/pbmc3k/sc_4.png){#fig-pbmc2_sc1}
-
-![](figures/pbmc3k/sc_5.png){#fig-pbmc2_sc2}
-
-![](figures/pbmc3k/sc_6.png){#fig-pbmc2_sc3}
-
-Screen shots of the **langevitour** of the PBMC3k data set, shows the model in high-D with new parameter choice for UMAP, a video of the tour animation is available at (<https://youtu.be/lc_6-rIuNHI>).
-:::
-
-- As shown in @fig-pbmc2_sc, there are 5 clusters, three of them are well-separated but close to each other as seen in @fig-nldervisPBMCUMAP3 (a).
-Also, other two clusters are really close to each other but separated from the early mentioned three clusters.
-
-- Next, we are going to investigate whether, the clusters we identified in 2D UMAP layout, can also be identified in high-D as well.
-
-- First, I colored the clusters in 2D layout manually. Then, I used spin-and-brush approach with `detour` to color the clusters in high-D.
-
-<!--Generate true label for PBMC3k-->
-
-```{r}
-#| warning: false
-#| echo: false
-
-UMAP_pbmc_with_result1_label <- umap_pbmc |>
-  dplyr::mutate(result1_class_label = case_when(
-    ((UMAP1 >=-12) & (UMAP1 <= -5) & (UMAP2 >= 6) & (UMAP2 <= 18)) ~ "cluster 1",
-    ((UMAP1 >-5) & (UMAP1 <= 7) & (UMAP2 >= 4.8) & (UMAP2 <= max(UMAP2))) ~ "cluster 2", ((UMAP1 >=6) & (UMAP1 <= 15) & (UMAP2 >= -4) & (UMAP2 < 4.8)) ~ "cluster 3",
-    ((UMAP1 >=-3) & (UMAP1 <= 6.6) & (UMAP2 >= min(UMAP2)) & (UMAP2 <= 4.8)) ~ "cluster 4",
-    ((UMAP1 >=min(UMAP1))& (UMAP1 < -3) & (UMAP2 >= -8) & (UMAP2 <= -1)) ~ "cluster 5",
-    ((UMAP1 >=min(UMAP1)) & (UMAP1 <= -7) & (UMAP2 >= -3) & (UMAP2 <= 3)) ~ "cluster 6",
-    .default = "no class"
-    ))
-
-### Reassign some labels
-
-UMAP_pbmc_with_result1_label <- UMAP_pbmc_with_result1_label |>
-  dplyr::mutate(result1_class_label = if_else((ID %in% c(2619, 1158)), "cluster 3", result1_class_label)) |>
-  dplyr::mutate(result1_class_label = if_else((ID %in% c(948, 285, 158, 268, 731, 1152, 1372, 1651, 2010, 150, 2304, 2546, 487, 2176, 2157, 2451, 264, 63, 1234, 947, 1126, 1699, 1232, 1492, 1488, 2040, 2550, 1313, 361, 372, 1515, 1264, 1307, 1726, 517, 2568, 2274, 591, 1703, 1097, 430, 1556, 2133, 1232, 1194, 1888, 1796, 494, 1194)), "cluster 6", result1_class_label))
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-## Import detour results
-detour_results8 <- read_csv("data/pbmc/pbmc_3k_festem/detourr_export8.csv")
-
-### Rename labels
-
-detour_results8 <- detour_results8 |>
-  dplyr::mutate(result2_class_label = case_when(
-    colour == "000000" ~ "cluster 4",
-    colour == "49df20" ~ "cluster 2",
-    colour == "619cff" ~ "cluster 3",
-    colour == "d82cd3" ~ "cluster 1",
-    colour == "df2020" ~ "cluster 5",
-    colour == "ffbd61" ~ "cluster 6",
-    .default = "no cluster"
-
-  ))
-
-detour_results8 <- detour_results8 |>
-  dplyr::mutate(ID = training_data_pbmc$ID)
-
-df_res_1_2_label <- inner_join(UMAP_pbmc_with_result1_label, detour_results8, by = c("ID" = "ID"))
-
-result1_2d_layout <- df_res_1_2_label |>
-    ggplot(aes(x = UMAP1,
-               y = UMAP2, color = result1_class_label))+
-    geom_point(alpha=0.5) +
-    coord_equal() +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#d82cd3", "#49df20", "#619cff", "#000000", "#df2020", "#ffbd61", "yellow")) +
-  # geom_text(aes(x=UMAP1,y = UMAP2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) +
-  labs(colour = "cluster labels")
-
-result2_2d_layout <- df_res_1_2_label |>
-    ggplot(aes(x = UMAP1,
-               y = UMAP2, color = result2_class_label))+
-    geom_point(alpha=0.5) +
-    coord_equal() +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(legend.position = "bottom", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#d82cd3", "#49df20", "#619cff", "#000000", "#df2020", "#ffbd61", "yellow")) +
-  # geom_text(aes(x=UMAP1,y = UMAP2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) +
-  labs(colour = "cluster labels")
-
-table(df_res_1_2_label$result1_class_label, df_res_1_2_label$result2_class_label)
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "2D layout from UMAP (n_neighbors = 5, min_dist = 0.99, metric = cosine) applied for PBMC3k dataset (a) colored manually in 2D, (b) colored in high-D."
-#| label: fig-nldervisPBMCUMAP3Res
-#| out-width: 100%
-#| fig-pos: H
-
-result1_2d_layout + result2_2d_layout +
-  plot_layout(guides='collect', ncol=2) &
-  theme(legend.position='bottom')
-
-```
-
-- @fig-nldervisPBMCUMAP3Res shows the 2D UMAP layout colored with each method.
-As shown in @fig-nldervisPBMCUMAP3Res (b) and the confusion matrix,there are some miss-classifications.
-What is the reason for that?
-
-- For example, the point (ID: 1158) classified as **cluster 3** in 2D UMAP layout, classified as **cluster 6** when doing the colouring in high-D (see @fig-nldervisPBMCUMAP3Res). Why?
-
-::: {#fig-pbmc2_appsc layout-ncol="1" fig-pos="H"}
-![](figures/pbmc3k/sc_7.png){#fig-pbmc2_sc7}
-
-Screen shots of app which shows 2D UMAP layout and comparison of distances related to the point classified as **cluster 6** in high-D (ID: 1158) and randomly selected point which is in **cluster 6** (ID: 2291).
-:::
-
-- If we compute the distance between these two points, you will see that it preserves the pair-wise distance (see @fig-pbmc2_appsc), which give us a evidence that even though the point 1158, classified as **cluster 3** in 2D layout, actually should be in **cluster 6**.
-This also evident that some misleading can happens with UMAP.
-
-::: {#fig-pbmc2_appsc1 layout-ncol="1" fig-pos="H"}
-![](figures/pbmc3k/sc_8.png){#fig-pbmc2_sc8}
-
-Screen shots of app which shows 2D UMAP layout and comparison of distances related to the point classified as **cluster 5** in high-D (ID: 3) and randomly selected point which is in **cluster 5** (ID: 2198).
-:::
-
-- For another example, if we compute the distance between these two points, you will see that it preserves the pair-wise distance (see @fig-pbmc2_appsc1), which give us a evidence that even though the point 3, classified as **cluster 4** in 2D layout, actually should be in **cluster 5**.
-This also evident that some misleading can happens with UMAP.
-
-```{r}
-#| warning: false
-#| echo: false
-#| message: false
-
-df_all_pbmc <- dplyr::bind_cols(training_data_pbmc, umap_data_with_hb_id_pbmc)
-
-df_all_pbmc <- inner_join(df_all_pbmc, detour_results8, by = c("ID" = "ID"))
-
-### Define type column
-df <- df_all_pbmc |>
-  dplyr::select(tidyselect::starts_with("PC_"), result2_class_label) |>
-  dplyr::rename("type" = "result2_class_label") ## original dataset
-
-df_b <- df_bin_pbmc |>
-  dplyr::filter(hb_id %in% df_bin_centroids_pbmc$hexID) |>
-  dplyr::mutate(type = "model") ## Data with summarized mean
-
-## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
-df_b <- df_b[match(df_bin_centroids_pbmc$hexID, df_b$hb_id),] |>
-  dplyr::select(-hb_id)
-
-df_exe <- dplyr::bind_rows(df_b, df)
-
- ## Set the maximum difference as the criteria
-distance_df_small_edges <- distance_pbmc |>
-  dplyr::filter(distance < benchmark_pbmc)
-## Since erase brushing is considerd.
-
-# langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
-#                          lineFrom = distance_df_small_edges$from,
-#                            lineTo = distance_df_small_edges$to,
-#                          group = df_exe$type, pointSize = 3,
-#                          levelColors = c("#d82cd3", "#49df20", "#619cff", "#000000", "#df2020", "#ffbd61", "#33a02c"))
-
+  scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
+  ylab("absolute error") +
+  xlab("number of non-empty bins")
+
+abs_error_plot_pbmc
 ```
@@
 
-- Why I choose this data set? Because I need show that whatever NLDR technique that we use, our model can capture the non-linear structure within clusters as well.
-
-- Digit 1 (7877 images only)
-
-#### Preprocessing
-
-- Scale the pixel values to the range [0,1]
-
-- Reshape to have a 2-dimensional array
-
--  reshaping pixel data is a fundamental preprocessing step in image analysis tasks, ensuring that the data is in a suitable format for further analysis, modeling, and interpretation.
-It helps preserve important spatial information, ensures compatibility with machine learning models, and facilitates feature extraction and visualization.
-
--  Reshaping the pixel data allows us to preserve this spatial information, ensuring that the relationships between neighboring pixels are maintained.
-
-- Many machine learning algorithms, including neural networks, expect input data to be in a specific format.
-For image data, this typically means a two-dimensional array where each row represents a single sample (image), and each column represents a feature (pixel value).
-Reshaping ensures that all samples have consistent dimensions, making them suitable for feeding into these algorithms.
-
-- Reshaping the pixel data allows for easier extraction of relevant features from the images.
-Once reshaped, individual pixels can be treated as features, making it easier to apply techniques like dimensionality reduction (e.g., PCA) or feature engineering to extract meaningful information from the images.
-
-- Selected 10 PCs based on scree plot
-
-- 7877 images of digit 1
-
-- PaCMAP parameters: n_components=2, n_neighbors=10, init="random", MN_ratio=0.9, FP_ratio=2.0
-
-#### Model
-
-- Total number of bins used: 180 (num_x = 12, num_y = 15, hex_size = 0.06)
-
-- Learned from the model: There is a non-linear shape in 2D, but if look at the model in high-D, the non-linear continuity is there but there is a twist in the structure which didn't capture by PaCMAP in 2D layout.
-
-#### Story
-
-- PaCMAP representation of MNIST digit 1 data with 10PCs and parameter choice of n_components=2, n_neighbors=10, init="random", MN_ratio=0.9, FP_ratio=2.0 shows a non-linear structure (see @fig-nldervisMNISTPACMAP (a)).
-I selected this parameter choice because this the default parameter setting for PaCMAP and selected PaCMAP because when I observed the 5 NLDR techniques applied for the data PaCMAP showed the non-linear structure effectively in 2D.
-
-- By looking at MSE plot, I selected the 180 bins, which associated with 12 bins along the x-axis, 15 bins along the y-axis, and hex size as 0.06 (@fig-diagnosticpltMNIST (b)).
-Then, fitted the model in 2D (see @fig-nldervisMNISTPACMAP (b)).
-As you can see in @fig-nldervisMNISTPACMAP (b), the fitted model capture the non-linear structure of the 2D data.
-
+The MNIST dataset consists of grayscale images of handwritten digits [@lecun2010].
+@Yingfan2021 used this dataset to demonstrate how the PaCMAP preserves local structure.
+We select the 2D embedding of PaCMAP for the handwritten digit 1.
+As shown in @fig-pacmapauthor, we observe a nonlinear structure present in 2D space.
+To answer the following questions,
+
+- Is the original data structure of the handwritten digit 1 dataset nonlinear?
+- Where does PaCMAP preserve the original data structure?
+- Where does PaCMAP show quirks?
+
+we visualize the model constructed with PaCMAP overlaid on data in high-D space (model-in-data space).
+
+<!-- PaCMAP layout with author's suggested parameter choice-->
 ```{r}
@@
 
-plot_list2_mnist <- pacmap_minst_scaled |>
+position_df <- pacmap_minst_scaled |> 
+  dplyr::filter(ID %in% c(6475, 3311, 6489, 2347, 387, 5239, 728, 517, 6369, 3055, 6345, 1761)) |>
+  dplyr::mutate(position = c(5, 9, 7, 8, 4, 11, 2, 6, 12, 10, 1, 3))
+
+pacmap_plot_mnist <- pacmap_minst_scaled |>
     ggplot(aes(x = PaCMAP1,
                y = PaCMAP2))+
-    geom_point(alpha=0.5) +
+    geom_point(alpha=0.1) +
     coord_equal() +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + 
+  geom_text(data = position_df, aes(x = PaCMAP1, y = PaCMAP2, label = position), colour = "red", size = 5) +
   theme_linedraw() +
@@
          legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  # geom_text(aes(x=UMAP1,y = UMAP2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) +
-  annotate(geom = 'text', label = 'a', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3) 
-```
-
+         legend.key.width = unit(0.25, 'cm')) 
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "(a) 2D layout from PaCMAP (n_components=2, n_neighbors=10, init=random, MN_ratio=0.9, FP_ratio=2.0) applied for the digit 1 of the MNIST dataset.
+Is this a best representation of the original data?"
+#| label: fig-pacmapauthor
+#| fig-pos: H
+
+pacmap_plot_mnist 
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+mnist_data <- read_rds("data/mnist/mnist_digit_1.rds")
+```
+
+<!--images with position 1-->
+```{r}
+#| warning: false
+#| echo: false
+
+pixels_gathered <-  mnist_data |>
+  mutate(instance = row_number()) |>
+  gather(pixel, value, -Label, -instance) |>
+  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
+  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
+  filter(instance %in% c(3710, 2391, 2385, 5030, 6475))
+
+theme_set(theme_light())
+imge_error_sample1 <- pixels_gathered |> 
+  ggplot(aes(x, y, fill = value)) +  
+  geom_tile() +  
+  facet_wrap(~ instance, ncol = 5) +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm'),
+        strip.text = element_blank())
+```
+
+<!--images with position 2-->
+```{r}
+#| warning: false
+#| echo: false
+
+pixels_gathered <-  mnist_data |>
+  mutate(instance = row_number()) |>
+  gather(pixel, value, -Label, -instance) |>
+  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
+  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
+  filter(instance %in% c(7679, 3568, 1312, 3311, 1097))
+
+theme_set(theme_light())
+imge_error_sample2 <- pixels_gathered |> 
+  ggplot(aes(x, y, fill = value)) +  
+  geom_tile() +  
+  facet_wrap(~ instance, ncol = 5) +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm'),
+        strip.text = element_blank())
+```
+
+<!--images with position 3-->
+```{r}
+#| warning: false
+#| echo: false
+
+pixels_gathered <-  mnist_data |>
+  mutate(instance = row_number()) |>
+  gather(pixel, value, -Label, -instance) |>
+  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
+  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
+  filter(instance %in% c(3552, 7853, 6489, 7689, 6690))
+
+theme_set(theme_light())
+imge_error_sample3 <- pixels_gathered |> 
+  ggplot(aes(x, y, fill = value)) +  
+  geom_tile() +  
+  facet_wrap(~ instance, ncol = 5) +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm'),
+        strip.text = element_blank())
+```
+
+<!--images with position 4-->
+```{r}
+#| warning: false
+#| echo: false
+
+pixels_gathered <-  mnist_data |>
+  mutate(instance = row_number()) |>
+  gather(pixel, value, -Label, -instance) |>
+  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
+  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
+  filter(instance %in% c(1380, 6057, 2347, 5946, 3355))
+
+theme_set(theme_light())
+imge_error_sample4 <- pixels_gathered |> 
+  ggplot(aes(x, y, fill = value)) +  
+  geom_tile() +  
+  facet_wrap(~ instance, ncol = 5) +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm'),
+        strip.text = element_blank())
+```
+
+<!--images with position 5-->
+```{r}
+#| warning: false
+#| echo: false
+
+pixels_gathered <-  mnist_data |>
+  mutate(instance = row_number()) |>
+  gather(pixel, value, -Label, -instance) |>
+  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
+  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
+  filter(instance %in% c(4175, 3997, 5378, 387, 1854))
+
+theme_set(theme_light())
+imge_error_sample5 <- pixels_gathered |> 
+  ggplot(aes(x, y, fill = value)) +  
+  geom_tile() +  
+  facet_wrap(~ instance, ncol = 5) +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm'),
+        strip.text = element_blank())
+```
+
+<!--images with position 6-->
+```{r}
+#| warning: false
+#| echo: false
+
+pixels_gathered <-  mnist_data |>
+  mutate(instance = row_number()) |>
+  gather(pixel, value, -Label, -instance) |>
+  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
+  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
+  filter(instance %in% c(614, 3079, 1762, 5239, 3723))
+
+theme_set(theme_light())
+imge_error_sample6 <- pixels_gathered |> 
+  ggplot(aes(x, y, fill = value)) +  
+  geom_tile() +  
+  facet_wrap(~ instance, ncol = 5) +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm'),
+        strip.text = element_blank())
+```
+
+<!--images with position 7-->
+```{r}
+#| warning: false
+#| echo: false
+
+pixels_gathered <-  mnist_data |>
+  mutate(instance = row_number()) |>
+  gather(pixel, value, -Label, -instance) |>
+  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
+  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
+  filter(instance %in% c(5748, 728, 7419, 7794, 6233))
+
+theme_set(theme_light())
+imge_error_sample7 <- pixels_gathered |> 
+  ggplot(aes(x, y, fill = value)) +  
+  geom_tile() +  
+  facet_wrap(~ instance, ncol = 5) +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm'),
+        strip.text = element_blank())
+```
+
+<!--images with position 8-->
+```{r}
+#| warning: false
+#| echo: false
+
+pixels_gathered <-  mnist_data |>
+  mutate(instance = row_number()) |>
+  gather(pixel, value, -Label, -instance) |>
+  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
+  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
+  filter(instance %in% c(33, 2485, 5998, 318, 1761))
+
+theme_set(theme_light())
+imge_error_sample8 <- pixels_gathered |> 
+  ggplot(aes(x, y, fill = value)) +  
+  geom_tile() +  
+  facet_wrap(~ instance, ncol = 5) +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm'),
+        strip.text = element_blank())
+```
+
+<!--images with position 9-->
+```{r}
+#| warning: false
+#| echo: false
+
+pixels_gathered <-  mnist_data |>
+  mutate(instance = row_number()) |>
+  gather(pixel, value, -Label, -instance) |>
+  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
+  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
+  filter(instance %in% c(5690, 165, 517, 6935, 2682))
+
+theme_set(theme_light())
+imge_error_sample9 <- pixels_gathered |> 
+  ggplot(aes(x, y, fill = value)) +  
+  geom_tile() +  
+  facet_wrap(~ instance, ncol = 5) +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm'),
+        strip.text = element_blank())
+```
+
+<!--images with position 10-->
+```{r}
+#| warning: false
+#| echo: false
+
+pixels_gathered <-  mnist_data |>
+  mutate(instance = row_number()) |>
+  gather(pixel, value, -Label, -instance) |>
+  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
+  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
+  filter(instance %in% c(4962, 2264, 5563, 6369, 559))
+
+theme_set(theme_light())
+imge_error_sample10 <- pixels_gathered |> 
+  ggplot(aes(x, y, fill = value)) +  
+  geom_tile() +  
+  facet_wrap(~ instance, ncol = 5) +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm'),
+        strip.text = element_blank())
+```
+
+<!--images with position 11-->
+```{r}
+#| warning: false
+#| echo: false
+
+pixels_gathered <-  mnist_data |>
+  mutate(instance = row_number()) |>
+  gather(pixel, value, -Label, -instance) |>
+  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
+  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
+  filter(instance %in% c(5188, 4849, 2666, 3448, 3055))
+
+theme_set(theme_light())
+imge_error_sample11 <- pixels_gathered |> 
+  ggplot(aes(x, y, fill = value)) +  
+  geom_tile() +  
+  facet_wrap(~ instance, ncol = 5) +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm'),
+        strip.text = element_blank())
+```
+
+<!--images with position 12-->
+```{r}
+#| warning: false
+#| echo: false
+
+pixels_gathered <-  mnist_data |>
+  mutate(instance = row_number()) |>
+  gather(pixel, value, -Label, -instance) |>
+  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
+  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
+  filter(instance %in% c(3120, 6869, 6345, 4470, 7147))
+
+theme_set(theme_light())
+imge_error_sample12 <- pixels_gathered |> 
+  ggplot(aes(x, y, fill = value)) +  
+  geom_tile() +  
+  facet_wrap(~ instance, ncol = 5) +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm'),
+        strip.text = element_blank())
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "Images of the handwritten digit 1 are distributed along the left-to-right of first embedding axis of the PaCMAP nonlinear structure."
+#| label: fig-sampleimg
+#| out-width: 100%
+#| fig-pos: H
+
+imge_error_sample1 + imge_error_sample2 + imge_error_sample3 + imge_error_sample4 + imge_error_sample5 + imge_error_sample6 + imge_error_sample7 + imge_error_sample8 + imge_error_sample9 + imge_error_sample10 + imge_error_sample11 + imge_error_sample12 +
+  plot_annotation(tag_levels = 'a') +
+  plot_layout(ncol = 2, nrow = 6) &
+  theme(plot.tag = element_text(size = 8))
+```
+
+<!-- MSE plot to find which is the effective number of bins to use-->
 ```{r}
@@
 #| echo: false
-
-## To draw with AIC
-aic_plot_mnist <- ggplot(mse_df_mnist, aes(x = num_bins, y = aic
-)) +
-  geom_point() +
-  geom_line() +
-  geom_vline(xintercept = 180, linetype="solid",
-                color = "red", size=0.8, alpha = 0.5) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  #scale_colour_manual(values = c("#377eb8")) +
-  ylab("AIC") +
-  xlab("total number of bins")
-## Effective number of bins along x-axis
 
@@
 #| fig-cap: "Goodness of fit statistics from PaCMAP applied to training MNIST dataset.
-What is the effective number of bins in each NLDR technique to create a 2D model?
+What is the effective number of bins in PaCMAP to create a 2D model?
 The MSE plot have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
@@
 #| label: fig-diagnosticpltMNIST
-##| out-width: 100%
 #| fig-pos: H
 
-aic_plot_mnist + mse_plot_mnist +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', ncol = 2) &
-  theme(legend.position='none', plot.tag = element_text(size = 8))
-```
-
+mse_plot_mnist
+```
+
+<!-- Fit the model and compute error-->
 ```{r}
@@
         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+  annotate(geom = 'text', label = 'a', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
 
@@
     ggplot(aes(x = PaCMAP1,
-               y = PaCMAP2, color = type))+
+               y = PaCMAP2, color = type, group = ID))+
     geom_point(alpha=0.5, size = 0.1) +
@@
                                          "#cb181d", "#a50f15", "#99000d")) +
-  annotate(geom = 'text', label = 'c', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-show_langevitour(df_all_mnist, df_bin_mnist, df_bin_centroids_mnist, benchmark_mnist, distance_mnist, "distance", col_start = "PC")
+  annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+# show_langevitour(df_all_mnist, df_bin_mnist, df_bin_centroids_mnist, benchmark_mnist, distance_mnist, "distance", col_start = "PC")
 
@@
 #| echo: false
-#| fig-cap: "(a) 2D layout from PaCMAP (n_components=2, n_neighbors=10, init=random, MN_ratio=0.9, FP_ratio=2.0) applied for the digit 1 of the MNIST dataset, (b) Model in the 2D space overlaid on PaCMAP data, and (c) Error occurred by the high-D model for high-D data and the error mapped into 2D NLDR points."
+#| fig-cap: "(a) Model in the 2D space overlaid on PaCMAP data, and (b) high-D model error in model space."
 #| label: fig-nldervisMNISTPACMAP
-#| out-width: 100%
 #| fig-pos: H
 
-plot_list2_mnist + trimesh_removed_mnist_pacmap + plot_list2_mnist_error +
+trimesh_removed_mnist_pacmap + plot_list2_mnist_error +
 plot_layout(guides='collect', ncol=3) &
@@
 ```
-
-- Next, the constructed 2D model lifts into high-dimensions (see @fig-mnist1_sc).
-
-::: {#fig-mnist1_sc layout-ncol="3" fig-pos="H"}
-![](figures/mnist/sc_1.png){#fig-mnist1_sc1}
-
-![](figures/mnist/sc_2.png){#fig-mnist1_sc2}
-
-![](figures/mnist/sc_3.png){#fig-mnist1_sc3}
-
-Screen shots of the **langevitour** of the MNIST digit 1 data set, shows the model in high-D, a video of the tour animation is available at (<https://youtu.be/zq2GM9qvUNA>).
-:::
-
-- As shown in @fig-mnist1_sc (a), the non-linear continuity structure that you have seen in 2D representation of PaCMAP, also can be seen in the model in high-D space as well.
-This gives us an evidence that our model capture the non-linear structure of the data effectively.
-
-- Furthermore, our model shows some addition pattern of data which couldn't be seen in the 2D layout.
-
-- One is the twisted structure within the non-linear structure (see @fig-mnist1_sc (c)).
-
-- Another thing is the connected corners of the non-linear structure (see @fig-mnist1_sc (b)). The non-linear curve is bended.
-
-- To gives the evidence that PaCMAP preserve the local structure (non-linear pattern), our model helps by computing the model errors (@fig-diagnosticpltMNIST (c)).
-
-<!--images with large error in high-D-->
-
-```{r}
-#| warning: false
-#| echo: false
-
-img_error <- error_df |> 
-  dplyr::filter(total >= 10) |> 
-  dplyr::pull(ID)
-
-mnist_data <- read_rds("data/mnist/mnist_digit_1.rds")
-pixels_gathered <-  mnist_data |>
-  mutate(instance = row_number()) |>
-  gather(pixel, value, -Label, -instance) |>
-  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
-  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
-  filter(instance %in% img_error)
-
-theme_set(theme_light())
-imge_error_sample <- pixels_gathered |> 
-  ggplot(aes(x, y, fill = value)) +  
-  geom_tile() +  
-  facet_wrap(~ instance) +
-  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm'),
-        strip.text = element_blank())
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "Images of handwritten digit 1 which occur large model error."
-#| label: fig-sampleimg
-#| out-width: 100%
-#| fig-pos: H
-
-imge_error_sample
-```
-
-- As you can see in @fig-diagnosticpltMNIST (c), the largest model error shows in the upper right side of the 2D layout.
-A sample of the images that shows the largest error can be seen in @fig-sampleimg.
-The reason for the large model error is that the high-D points related to this largest error located away from the high-D model points.
 

### 2024-03-21 paper.tex
Add intro to application data sets

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, sharp corners, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, interior hidden, breakable, enhanced, boxrule=0pt, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 UMAP, PHATE, TriMAP \citep{article02}, and Pairwise Controlled Manifold
-Approximation (PaCMAP) \citep{article01}.
+Approximation (PaCMAP) \citep{Yingfan2021}.
 
@@
 
-PaCMAP \citep{article01} is different because it adds supervised
+PaCMAP \citep{Yingfan2021} is different because it adds supervised
 learning to make a 2D representation while keeping the relationships
@@
 useful in determining the effective number of bins required to construct
-the 2D model. Figure~\ref{fig-diagnosticpltScurve} (b) has created a
-graph that shows how the Mean Squared Error (MSE) changes depending on
-the number of bins assigned in the hexagonal grid for UMAP applied to
-the S-curve dataset. As per the rule of thumb mentioned earlier, the
+the 2D model. Figure~\ref{fig-diagnosticpltScurve} has created a graph
+that shows how the Mean Squared Error (MSE) changes depending on the
+number of bins assigned in the hexagonal grid for UMAP applied to the
+S-curve dataset. As per the rule of thumb mentioned earlier, the
 effective number of bins is determined to be \(84\), with \(6\) bins
@@
 it is necessary to select hexagonal bins that have a considerable number
-of data points within their hexagons. To achieve this, the standard
-number of points within each hexagon is first calculated. This standard
-count is obtained by dividing the number of points within each hexagon
-by the maximum number of points in the grid (see
-Equation~\ref{eq-equationp2}). Next, the bins that have a standard
-number of points less than a certain benchmark value are removed. After
-removing the hexagons that have insufficient data density, the hexagons
-with more substantial data representation are used to construct the 2D
-model.
+of data points within their hexagons. To achieve this, the number of
+points within each hexagon is first calculated. Then, the standard count
+is computed by dividing the number of points within each hexagon by the
+maximum number of points in the grid (see Equation~\ref{eq-equationp2}).
+Next, the bins that have a standard number of points less than a certain
+benchmark value are removed. After removing the hexagons that have
+insufficient data density, the hexagons with more substantial data
+representation are used to construct the 2D model.
 
@@
 only those bins will be removed. The other identified bins will be kept
-and used for constructing the 2D model along with the high-density bins
-(see Appendix for more details).
+and used for constructing the 2D model along with the high-density bins.
 
@@
 substantial, the decision might lean towards retaining low-density
-hexagons. According to \textbf{?@fig-diagnosticpltScurvelwd}, the change
-in MSE for NLDR techniques is relatively small across different
-benchmark values, except for the PHATE method.
-
-Consequently, it is not necessary to remove low-density hexagons when
-constructing the 2D model for the S-curve dataset in NLDR techniques,
-except for PHATE. While there may be some fluctuations in MSE between
-benchmark values of 0 to 1, the absence of significant changes suggests
-that maintaining low-density hexagons is preferable. However, to choose
-the best selection, it's recommended to examine the benchmark values
-around the first local minimum in the MSE curve. However, the MSE for
-PHATE fluctuates when benchmark values range between 0 and 1. Therefore,
-it is necessary to explore the benchmark values surrounding the first
-local minimum to determine an effective benchmark value. There is no
-universal rule for selecting a process. As shown in
-\textbf{?@fig-diagnosticpltScurvelwd}, the MSE values for PHATE
-fluctuate, and the first local minimum is at \(0\). Therefore, nearby
-benchmark values such as \(0.01\), \(0.04\), and \(0.06\) were explored
-to identify the best value for preserving the data structure. In this
-case, a benchmark value of \(0.06\) was chosen. The resulting 2D model
-is shown in Figure~\ref{fig-modelScurve} (c), while the video link
-provides a visual representation of how the model appears in high-D.
+hexagons.
 
@@
 \def\labelenumi{\arabic{enumi}.}
+\tightlist
 \item
   Plot the distribution of the standardized counts
+\end{enumerate}
+
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-stdctsScurve-1.pdf}
+
+}
+
+\caption{\label{fig-stdctsScurve}Distribution of 2D Euclidean
+distances.}
+
+\end{figure}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{1}
 \item
@@
 
-\caption{\label{fig-modelScurve}(a), (b) 2D model.}
+\caption{\label{fig-modelScurve}(a) Full hexagonal grid with bin
+centroids, (b) Model constructed in 2D with long edges, and (c) Model
+constructed in 2D after removing the long edges.}
 
@@
 \def\labelenumi{\arabic{enumi}.}
+\tightlist
 \item
   Plot the distribution of the 2D distances
+\end{enumerate}
+
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-distScurve-1.pdf}
+
+}
+
+\caption{\label{fig-distScurve}Distribution of 2D Euclidean distances.}
+
+\end{figure}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{1}
+\tightlist
 \item
@@
 To assess how well our method captures and represents the underlying
-structure of the high-D data, Mean Squared Error (MSE) and Akaike
-Information Criterion (AIC) are used. When computing MSE, total model
-error (see Section~\ref{sec-summary}) is divided by the number of
-observations to make it as a mean value (see
+structure of the high-D data, Mean Squared Error (MSE) is used. When
+computing MSE, total model error (see Section~\ref{sec-summary}) is
+divided by the number of observations to make it as a mean value (see
 Equation~\ref{eq-equation9}).
@@
 }\label{eq-equation9}\end{equation}
-
-Furthermore, to incorporate with the parameters in our model, which is
-the non-empty bins, and as a statistical measure used for model
-selection and comparison, AIC is used (see
-Equation~\ref{eq-equation10}).
-
-\begin{equation}\protect\hypertarget{eq-equation10}{}{
-\text{AIC} = 2b'p + np * log(\text{MSE})
-}\label{eq-equation10}\end{equation}
-
-Here, \(b'\) signifies the number of bins non-empty bins, \(p\) denotes
-the number of dimensions in the high-D data, and \(n\) represents the
-number of observations.
 
@@
 
-In the field of single-cell studies, a common analysis task involves
-clustering to identify groups of cells with similar expression profiles.
-Analysts often turn to NLDR techniques to verify and identify these
-clusters and explore developmental trajectories. In clustering
-workflows, the main objective is to verify the existence of clusters and
-subsequently identify them as specific cell types by examining the
-expression of ``known'' marker genes. Marker genes are specific genes or
-transcripts that are uniquely or highly expressed in particular cell
-types. Marker genes are identified through differential gene expression
-(DEG) analysis, comparing the expression levels of genes across
-different cell types. In this context, a ``faithful'' embedding should
-ideally preserve the topology of the data, ensuring that cells
-corresponding to the same cell type are situated close to the
-high-dimensional space.
-
-\hypertarget{processing-steps}{%
-\subsection{Processing steps}\label{processing-steps}}
-
-To begin our analysis, we installed the Peripheral Blood Mononuclear
-Cells (pbmc) data set obtained from 10x Genomics using the
-\texttt{SeuratData} R package \citep{Rahul2019}, which facilitates the
-distribution of data sets in the form of Seurat objects
-\citep{Yuhan2021}. This data set contains 13,714 features (genes) across
-2,700 samples (cells) within a single assay. The active assay is RNA,
-with 13,714 features representing different gene expressions. Then, the
-cells that have unique feature counts over 2,500 or less than 200 and
-cells that have \textgreater{} 5\% mitochondrial counts are filtered.
-After removing unwanted cells from the dataset, the next step is to
-normalize the data. By default, we employ a global-scaling normalization
-method ``LogNormalize'' that normalizes the feature expression
-measurements for each cell by the total expression, multiplies this by a
-scale factor (10,000 by default), and log-transforms the result.
-
-Next, top 1000 genes are selected by Festem \citep{Chen2023} DEG method.
-Then, scale the data. Then, perform PCA on the scaled data. Louvain
-algorithm is used to cluster the single cells based on the genes
-selected by Festem. Using the Festem-selected genes and 15 PCs,
-identified 10 clusters in the PBMC3k data and annotated them based on
-the expression of canonical markers. The 10 clusters included immune
-cells such as naive CD4 T cells, memory CD4 T cells, CD8 T cells, CD14
-monocytes, FCGR3A monocytes, Natural Killer (NK) cells, B cells and
-Dendritic Cells (DC). In addition to these common cell types, Festem
-identified a fine cell type, CD27− CD4+ memory T cells, which were often
-missed by other methods.
-
-\begin{itemize}
-\item
-  Human peripheral blood mononclear cells (PBMCs) data
-\item
-  Intro to the data set
-\end{itemize}
-
-Used Louvain algorithm to cluster the single cells based on the genes
-selected by Festem method
-
-\begin{itemize}
-\item
-  Explain what ``Evaluate Festem on DEG detection'' means, in terms of
-  this data
-\item
-  How did they ``identified 10 clusters''? What were the variables used,
-  PCs or full data or 2D representation, spell this out.
-\item
-  Used top 1000 genes selected by Festem to run PCA, top 15 PCs
-  (original paper mentioned)
-\item
-  The ``10 clusters included immune cells such as naive CD4 T \ldots{}''
-  does this mean that every cluster had all of these, or one cluster was
-  mostly ``CD4 T cells'', another mostly ``memory CD4 T'', \ldots?
-\item
-  How is your use of this data different from the original application.
-\item
-  Original paper used this data set to compare different DEG methods,
-  but we need to look how the performance of UMAP with the author's
-  selected parameter choice.
-\item
-  Why 15 PCs? Is that from the original paper? YES
-\end{itemize}
-
-\citet{Chen2023}
-
-\begin{itemize}
-\item
-  Intro to the data set (This data set contains 13,714 features across
-  2,700 samples within a single assay. The active assay is RNA, with
-  13,714 features representing different gene expressions.)
-\item
-  2622 cells
-\item
-  Evaluate Festem on DEG detection (Festem enabled identification of
-  often-missed fine cell types)
-\item
-  Using the Festem-selected genes, identified 10 clusters in the PBMCsk
-  data and annotated them based on the expression of canonical markers
-\item
-  The 10 clusters included immune cells such as naive CD4 T cells,
-  memory CD4 T cells, CD8 T cells, CD14 monocytes, FCGR3A monocytes,
-  Natural Killer (NK) cells, B cells and Dendritic Cells (DC).
-\item
-  In addition to these common cell types, Festem identified a fine cell
-  type, CD27− CD4+ memory T cells, which were often missed by other
-  methods.
-\item
-  The CD27. CD4+ memory T cells identified by Festem expressed common
-  marker genes (IL7R and S100A4) of memory T cells, but did not express
-  CD27. These cells also had downregulated expression of SELL, CCR7, MAL
-  and LEF1, and upregulated expression of CCL5 (Fig. 4B) and thus were
-  the CD27- CD4+ memory T cells in the literature (36). The CD27- CD4+
-  memory T cells were known to be at a more differentiated state and
-  have stronger antigen-recall responses than their CD27+ counterparts.
-\item
-  15 PCs
-\item
-  Num bins along the x-axis = 23, shape parameter = 0.8772751
-\item
-  learned from the model: There are 3 well-separated clusters in 2D, but
-  if look at the model in high-D space, the three clusters are much
-  closer. Also, there is some continuity within the clustering structure
-  and it didn't capture well. Therefore UMAP with n:neighbour 30 is not
-  a good representation for this data set.
-\item
-  The most important parameter is n\_neighbors - the number of
-  approximate nearest neighbors used to construct the initial
-  high-dimensional graph. It effectively controls how UMAP balances
-  local versus global structure - low values will push UMAP to focus
-  more on local structure by constraining the number of neighboring
-  points considered when analyzing the data in high dimensions, while
-  high values will push UMAP towards representing the big-picture
-  structure while losing fine detail.
-\item
-  The second parameter we'll investigate is min\_dist, or the minimum
-  distance between points in low-dimensional space. This parameter
-  controls how tightly UMAP clumps points together, with low values
-  leading to more tightly packed embeddings. Larger values of min\_dist
-  will make UMAP pack points together more loosely, focusing instead on
-  the preservation of the broad topological structure.
-\item
-  The third parameter is metric. This parameter plays a crucial role in
-  determining how distances are computed within the ambient space of the
-  input data (The term ``ambient space'' refers to the original or
-  initial space in which the data exists before any transformation or
-  dimensionality reduction).
-\item
-  If it is completely a mess, then that is also further evidence that
-  the fit is poor in high-dimensions.
-\end{itemize}
-
-\hypertarget{approach-followed}{%
-\subsection{Approach followed}\label{approach-followed}}
-
-\begin{itemize}
-\item
-  First, referred to \citet{Chen2023} paper which shows UMAP
-  representation of PBMC3k data with 15 PCs and parameter choice of
-  n:neighbor:30, min\_dist:0.1, metric:cosine.
-\item
-  Observed that there is three well-separated clusters.
-\item
-  Second, visualise the 15 PCs in high-dimensional space with our model
-  and investigate whether can I see the three well-separated clusters or
-  is it showing something different than what I saw in UMAP view
-  suggested by the author.
-\item
-  Learned from the model: Yes, it's different. There are three clusters
-  but they are really close to each other. Also, in the largest cluster,
-  observed non-linear continuity structure, which didn't captured by the
-  author's UMAP representation.
-\item
-  Learned from the data: Visualise with detour and colored the clusters
-  in high-D. Identified more than three clusters which is suprising.
-\item
-  In that case, need to suggest a better UMAP representation than
-  author's suggestion.
-\item
-  To find a better UMAP representation, I need to investigate different
-  parameter choices for UMAP (or need to try another NLDR technique, but
-  here my objective is to suggest the best parameter choice with UMAP).
-\item
-  Therefore, I fitted our model for different UMAP parameter choices,
-  record the error, and selected the parameter choice that gives the
-  lowest error which is n:neighbor:5, min\_dist:0.99, metric:cosine.
-\item
-  Then, construct our model for this UMAP data (n:neighbor:5,
-  min\_dist:0.99, metric:cosine).
-\item
-  In this representation, you can see that the clusters are closer, and
-  the continuity of clusters are there.
-\item
-  Also, what missed from looking at high-D data with detour, in this
-  UMAP representation I can see more than three clusters.
-\item
-  Then, need to check how correct am I identify the clusters in 2D and
-  is that matched that I identify in high-D.
-\item
-  Generate a confusion matrix.
-\end{itemize}
-
-\hypertarget{story}{%
-\subsection{Story}\label{story}}
+\begin{figure}[H]
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-umapauthor-1.pdf}
+
+}
+
+\caption{\label{fig-umapauthor}(a) 2D layout from UMAP (n\_neighbors =
+30, min\_dist = 0.3) applied for the PBMC3k dataset. Is this a best
+representation of the original data?}
+
+\end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-msepltpbmc-1.pdf}
+
+}
+
+\caption{\label{fig-msepltpbmc}Goodness of fit statistics from UMAP
+applied to training PBMC3k dataset. What is the effective number of bins
+in UMAP to create a 2D model? The MSE plot have a steep slope at the
+beginning, indicating that a smaller number of bins causes a larger
+amount of error. Then, the slope gradually declines or level off,
+indicating that a higher number of bins generates a smaller error. Using
+the elbow method, when the total number of bins is set to 588, the slope
+of the MSE plot experiences a sudden and noticeable change, resembling
+an elbow-like shape. This point indicates that adding less bins does not
+enough to capture the data structure.}
+
+\end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-nldervisPBMCUMAP-1.pdf}
+
+}
+
+\caption{\label{fig-nldervisPBMCUMAP}(a) Model in the 2D space overlaid
+on UMAP data, and (b) Error occurred by the high-D model for high-D data
+and the error mapped into 2D NLDR points.}
+
+\end{figure}
+
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-72-1.pdf}
+
+\hypertarget{handwritten-digits}{%
+\subsection{Handwritten digits}\label{handwritten-digits}}
+
+The MNIST dataset consists of grayscale images of handwritten digits
+\citep{lecun2010}. \citet{Yingfan2021} used this dataset to demonstrate
+how the PaCMAP preserves local structure. We select the 2D embedding of
+PaCMAP for the handwritten digit 1. As shown in
+Figure~\ref{fig-pacmapauthor}, we observe a nonlinear structure present
+in 2D space. To answer the following questions,
 
@@
 \item
-  UMAP representation of PBMC3k data with 15 PCs and parameter choice of
-  n:neighbor:30, min\_dist:0.1, metric:cosine (author's suggestion)
-  shows three well-separated clusters (see
-  Figure~\ref{fig-nldervisPBMCUMPview}). The question we need to answer
-  is: Does this UMAP representation preserve the actual data structure,
-  or is something missing or misidentified or misleading?
+  Is the original data structure of the handwritten digit 1 dataset
+  nonlinear?
+\item
+  Where does PaCMAP preserve the original data structure?
+\item
+  Where does PaCMAP show quirks?
 \end{itemize}
 
+we visualize the model constructed with PaCMAP overlaid on data in
+high-D space (model-in-data space).
+
 \begin{figure}[H]
 
-{\centering \includegraphics{paper_files/figure-pdf/fig-nldervisPBMCUMPview-1.pdf}
-
-}
-
-\caption{\label{fig-nldervisPBMCUMPview}(a) 2D layout from UMAP
-(n\_neighbors = 30, min\_dist = 0.3) applied for the PBMC3k dataset. Is
-this a best representation of the original data?}
-
-\end{figure}
-
-\begin{itemize}
-\item
-  To investigate this, we are going to visualise the model in the data
-  space, which means that the model generated from UMAP is visualised in
-  high-D with original data.
-\item
-  To fit the model, first we need to decide how many number bins should
-  allocate along the x and y axes. To make an initial choice, we use MSE
-  plot (Figure~\ref{fig-diagnosticpltPBMC} (b)).
-\end{itemize}
-
-\begin{figure}
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-diagnosticpltPBMC-1.pdf}
-
-}
-
-\caption{\label{fig-diagnosticpltPBMC}Goodness of fit statistics from
-UMAP applied to training PBMC3k dataset. What is the effective number of
-bins in each NLDR technique to create a 2D model? The MSE plot have a
-steep slope at the beginning, indicating that a smaller number of bins
-causes a larger amount of error. Then, the slope gradually declines or
-level off, indicating that a higher number of bins generates a smaller
-error. Using the elbow method, when the total number of bins is set to
-588, the slope of the Mean Squared Error (MSE) plot experiences a sudden
-and noticeable change, resembling an elbow-like shape. This point
-indicates that adding less bins does not enough to capture the data
-structure.}
-
-\end{figure}
-
-\begin{itemize}
-\tightlist
-\item
-  Then, the model is constructed by following all the steps in 2D and
-  lifts the model into high-D.
-\end{itemize}
+{\centering \includegraphics{paper_files/figure-pdf/fig-pacmapauthor-1.pdf}
+
+}
+
+\caption{\label{fig-pacmapauthor}(a) 2D layout from PaCMAP
+(n\_components=2, n\_neighbors=10, init=random, MN\_ratio=0.9,
+FP\_ratio=2.0) applied for the digit 1 of the MNIST dataset. Is this a
+best representation of the original data?}
+
+\end{figure}
 
@@
 
-{\centering \includegraphics{paper_files/figure-pdf/fig-nldervisPBMCUMAP-1.pdf}
-
-}
-
-\caption{\label{fig-nldervisPBMCUMAP}(a) Model in the 2D space overlaid
-on UMAP data, and (b) Error occurred by the high-D model for high-D data
-and the error mapped into 2D NLDR points.}
-
-\end{figure}
-
-\begin{itemize}
-\tightlist
-\item
-  As shown in Figure~\ref{fig-nldervisPBMCUMAP}, there are three
-  well-separated clusters (according to the model in 2D space).
-\end{itemize}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-sampleimg-1.pdf}
+
+}
+
+\caption{\label{fig-sampleimg}Images of the handwritten digit 1 are
+distributed along the left-to-right of first embedding axis of the
+PaCMAP nonlinear structure.}
+
+\end{figure}
 
 \begin{figure}[H]
-
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/pbmc3k/sc_1.png}
-
-}
-
-}
-
-\subcaption{\label{fig-pbmc1_sc1}}
-\end{minipage}%
-%
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/pbmc3k/sc_2.png}
-
-}
-
-}
-
-\subcaption{\label{fig-pbmc1_sc2}}
-\end{minipage}%
-%
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/pbmc3k/sc_3.png}
-
-}
-
-}
-
-\subcaption{\label{fig-pbmc1_sc3}}
-\end{minipage}%
-
-\caption{\label{fig-pbmc1_sc}Screen shots of the \textbf{langevitour} of
-the PBMC3k data set, shows the model in high-D, a video of the tour
-animation is available at (\url{https://youtu.be/Gnhh4hfsbbg}).}
-
-\end{figure}
-
-\begin{itemize}
-\item
-  According to Figure~\ref{fig-pbmc1_sc} (a), the model in high-D space
-  shows two well separated clusters. Also, two clusters are really close
-  to each other (see Figure~\ref{fig-pbmc1_sc} (b)). Furthermore, there
-  are three clusters shown by the model in 2D actually represented in
-  high-D as well (see Figure~\ref{fig-pbmc1_sc} (c)). Another important
-  thing to notice from the model in high-D space is that the non-linear
-  continuity structure in two well-separated clusters.
-\item
-  Figure~\ref{fig-pbmc1_sc} (c) gives evidence that UMAP preserves the
-  global structure. Furthermore, the non-linearity within the clusters
-  shown in Figure~\ref{fig-pbmc1_sc} (a) proves that local structure is
-  also preserved but that's not visually shown by th UMAP layout in 2D.
-\item
-  Can we find a better parameter choice for UMAP which shows this
-  non-linear continuity of the big clusters in 2D space as well?
-\item
-  To do that, we are going to fit models for different parameter choice
-  and select the one which have the lowest model error in high-D (see
-  \textbf{?@tbl-errordiffcomb}).
-\end{itemize}
-
-\begin{table}[H]
-
-\caption{errors for different parameter combinations}
-\centering
-\begin{tabular}[t]{r|r|l|r}
-\hline
-\textbf{n\_neighbors} & \textbf{min\_dist} & \textbf{metric} & \textbf{error (x 100)}\\
-\hline
-\cellcolor{gray!6}{5} & \cellcolor{gray!6}{0.99} & \cellcolor{gray!6}{cosine} & \cellcolor{gray!6}{262.18}\\
-\hline
-15 & 0.99 & cosine & 279.49\\
-\hline
-\cellcolor{gray!6}{84} & \cellcolor{gray!6}{0.99} & \cellcolor{gray!6}{cosine} & \cellcolor{gray!6}{295.64}\\
-\hline
-5 & 0.30 & cosine & 303.90\\
-\hline
-\cellcolor{gray!6}{15} & \cellcolor{gray!6}{0.30} & \cellcolor{gray!6}{cosine} & \cellcolor{gray!6}{308.97}\\
-\hline
-84 & 0.30 & cosine & 319.59\\
-\hline
-\textcolor{red}{\textbf{\cellcolor{gray!6}{30}}} & \textcolor{red}{\textbf{\cellcolor{gray!6}{0.30}}} & \textcolor{red}{\textbf{\cellcolor{gray!6}{cosine}}} & \textcolor{red}{\textbf{\cellcolor{gray!6}{386.35}}}\\
-\hline
-\end{tabular}
-\end{table}
-
-\begin{figure}[H]
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-nldervisPBMCUMAPview2-1.pdf}
-
-}
-
-\caption{\label{fig-nldervisPBMCUMAPview2}(a) 2D layout from UMAP
-(n\_neighbors = 5, min\_dist = 0.99, metric = cosine) applied for the
-PBMC3k dataset. Is this a best representation of the original data?}
-
-\end{figure}
-
-\begin{itemize}
-\item
-  Selected the one with n:neighbor:5, min\_dist:0.99, metric:cosine (see
-  Figure~\ref{fig-nldervisPBMCUMAPview2}). As shown in
-  Figure~\ref{fig-nldervisPBMCUMAPview2}, the two big clusters have
-  non-linear continuity structures, two big clusters are well-separated
-  but close to each other, and the two cluster which exists in the one
-  big cluster are really close to each other (see
-  Figure~\ref{fig-nldervisPBMCUMAPview2}).
-\item
-  Now, we are going to fit the model for the new parameter choice.
-\end{itemize}
-
-\begin{figure}
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-diagnosticpltPBMC3-1.pdf}
-
-}
-
-\caption{\label{fig-diagnosticpltPBMC3}Goodness of fit statistics from
-UMAP applied to training S-curve dataset. What is the effective number
-of bins in each NLDR technique to create a 2D model? The MSE plot have a
-steep slope at the beginning, indicating that a smaller number of bins
-causes a larger amount of error. Then, the slope gradually declines or
-level off, indicating that a higher number of bins generates a smaller
-error. Using the elbow method, when the total number of bins is set to
-588, the slope of the Mean Squared Error (MSE) plot experiences a sudden
-and noticeable change, resembling an elbow-like shape. This point
-indicates that adding less bins does not enough to capture the data
-structure.}
-
-\end{figure}
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervisPBMCUMAP3-1.pdf}
-
-}
-
-\caption{\label{fig-nldervisPBMCUMAP3}(a) Model in the 2D space overlaid
-with UMAP data, and (b) Error occurred by the high-D model for high-D
-data and the error mapped into 2D NLDR points.}
-
-\end{figure}
-
-\begin{itemize}
-\tightlist
-\item
-  According to Figure~\ref{fig-nldervisPBMCUMAP3} (a), the model
-  constructed in 2D space for the new parameter choice of UMAP shows 5
-  clusters which are really close to each other. The big cluster shows
-  the non-linear continuity structure with two sub clusters. Other three
-  clusters positioned in a structure of non-linearity but really close
-  to each other.
-\end{itemize}
-
-\begin{figure}[H]
-
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/pbmc3k/sc_4.png}
-
-}
-
-}
-
-\subcaption{\label{fig-pbmc2_sc1}}
-\end{minipage}%
-%
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/pbmc3k/sc_5.png}
-
-}
-
-}
-
-\subcaption{\label{fig-pbmc2_sc2}}
-\end{minipage}%
-%
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/pbmc3k/sc_6.png}
-
-}
-
-}
-
-\subcaption{\label{fig-pbmc2_sc3}}
-\end{minipage}%
-
-\caption{\label{fig-pbmc2_sc}Screen shots of the \textbf{langevitour} of
-the PBMC3k data set, shows the model in high-D with new parameter choice
-for UMAP, a video of the tour animation is available at
-(\url{https://youtu.be/lc_6-rIuNHI}).}
-
-\end{figure}
-
-\begin{itemize}
-\item
-  As shown in Figure~\ref{fig-pbmc2_sc}, there are 5 clusters, three of
-  them are well-separated but close to each other as seen in
-  Figure~\ref{fig-nldervisPBMCUMAP3} (a). Also, other two clusters are
-  really close to each other but separated from the early mentioned
-  three clusters.
-\item
-  Next, we are going to investigate whether, the clusters we identified
-  in 2D UMAP layout, can also be identified in high-D as well.
-\item
-  First, I colored the clusters in 2D layout manually. Then, I used
-  spin-and-brush approach with \texttt{detour} to color the clusters in
-  high-D.
-\end{itemize}
-
-\begin{verbatim}
-
-            cluster 1 cluster 2 cluster 3 cluster 4 cluster 5 cluster 6
-  cluster 1       138        13         0         0         0         0
-  cluster 2        11       509         0         1         0         0
-  cluster 3         1         0       348         1         0         1
-  cluster 4         0         0         3      1029        64         2
-  cluster 5         0         0         0        85       226        32
-  cluster 6         0         0         0         0        19       139
-\end{verbatim}
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervisPBMCUMAP3Res-1.pdf}
-
-}
-
-\caption{\label{fig-nldervisPBMCUMAP3Res}2D layout from UMAP
-(n\_neighbors = 5, min\_dist = 0.99, metric = cosine) applied for PBMC3k
-dataset (a) colored manually in 2D, (b) colored in high-D.}
-
-\end{figure}
-
-\begin{itemize}
-\item
-  Figure~\ref{fig-nldervisPBMCUMAP3Res} shows the 2D UMAP layout colored
-  with each method. As shown in Figure~\ref{fig-nldervisPBMCUMAP3Res}
-  (b) and the confusion matrix,there are some miss-classifications. What
-  is the reason for that?
-\item
-  For example, the point (ID: 1158) classified as \textbf{cluster 3} in
-  2D UMAP layout, classified as \textbf{cluster 6} when doing the
-  colouring in high-D (see Figure~\ref{fig-nldervisPBMCUMAP3Res}). Why?
-\end{itemize}
-
-\begin{figure}[H]
-
-\begin{minipage}[t]{\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/pbmc3k/sc_7.png}
-
-}
-
-}
-
-\subcaption{\label{fig-pbmc2_sc7}}
-\end{minipage}%
-
-\caption{\label{fig-pbmc2_appsc}Screen shots of app which shows 2D UMAP
-layout and comparison of distances related to the point classified as
-\textbf{cluster 6} in high-D (ID: 1158) and randomly selected point
-which is in \textbf{cluster 6} (ID: 2291).}
-
-\end{figure}
-
-\begin{itemize}
-\tightlist
-\item
-  If we compute the distance between these two points, you will see that
-  it preserves the pair-wise distance (see
-  Figure~\ref{fig-pbmc2_appsc}), which give us a evidence that even
-  though the point 1158, classified as \textbf{cluster 3} in 2D layout,
-  actually should be in \textbf{cluster 6}. This also evident that some
-  misleading can happens with UMAP.
-\end{itemize}
-
-\begin{figure}[H]
-
-\begin{minipage}[t]{\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/pbmc3k/sc_8.png}
-
-}
-
-}
-
-\subcaption{\label{fig-pbmc2_sc8}}
-\end{minipage}%
-
-\caption{\label{fig-pbmc2_appsc1}Screen shots of app which shows 2D UMAP
-layout and comparison of distances related to the point classified as
-\textbf{cluster 5} in high-D (ID: 3) and randomly selected point which
-is in \textbf{cluster 5} (ID: 2198).}
-
-\end{figure}
-
-\begin{itemize}
-\tightlist
-\item
-  For another example, if we compute the distance between these two
-  points, you will see that it preserves the pair-wise distance (see
-  Figure~\ref{fig-pbmc2_appsc1}), which give us a evidence that even
-  though the point 3, classified as \textbf{cluster 4} in 2D layout,
-  actually should be in \textbf{cluster 5}. This also evident that some
-  misleading can happens with UMAP.
-\end{itemize}
-
-\hypertarget{handwritten-digits}{%
-\subsection{Handwritten digits}\label{handwritten-digits}}
-
-\begin{itemize}
-\item
-  Why I choose this data set? Because I need show that whatever NLDR
-  technique that we use, our model can capture the non-linear structure
-  within clusters as well.
-\item
-  Digit 1 (7877 images only)
-\end{itemize}
-
-\hypertarget{preprocessing}{%
-\subsubsection{Preprocessing}\label{preprocessing}}
-
-\begin{itemize}
-\item
-  Scale the pixel values to the range {[}0,1{]}
-\item
-  Reshape to have a 2-dimensional array
-\item
-  reshaping pixel data is a fundamental preprocessing step in image
-  analysis tasks, ensuring that the data is in a suitable format for
-  further analysis, modeling, and interpretation. It helps preserve
-  important spatial information, ensures compatibility with machine
-  learning models, and facilitates feature extraction and visualization.
-\item
-  Reshaping the pixel data allows us to preserve this spatial
-  information, ensuring that the relationships between neighboring
-  pixels are maintained.
-\item
-  Many machine learning algorithms, including neural networks, expect
-  input data to be in a specific format. For image data, this typically
-  means a two-dimensional array where each row represents a single
-  sample (image), and each column represents a feature (pixel value).
-  Reshaping ensures that all samples have consistent dimensions, making
-  them suitable for feeding into these algorithms.
-\item
-  Reshaping the pixel data allows for easier extraction of relevant
-  features from the images. Once reshaped, individual pixels can be
-  treated as features, making it easier to apply techniques like
-  dimensionality reduction (e.g., PCA) or feature engineering to extract
-  meaningful information from the images.
-\item
-  Selected 10 PCs based on scree plot
-\item
-  7877 images of digit 1
-\item
-  PaCMAP parameters: n\_components=2, n\_neighbors=10, init=``random'',
-  MN\_ratio=0.9, FP\_ratio=2.0
-\end{itemize}
-
-\hypertarget{model}{%
-\subsubsection{Model}\label{model}}
-
-\begin{itemize}
-\item
-  Total number of bins used: 180 (num\_x = 12, num\_y = 15, hex\_size =
-  0.06)
-\item
-  Learned from the model: There is a non-linear shape in 2D, but if look
-  at the model in high-D, the non-linear continuity is there but there
-  is a twist in the structure which didn't capture by PaCMAP in 2D
-  layout.
-\end{itemize}
-
-\hypertarget{story-1}{%
-\subsubsection{Story}\label{story-1}}
-
-\begin{itemize}
-\item
-  PaCMAP representation of MNIST digit 1 data with 10PCs and parameter
-  choice of n\_components=2, n\_neighbors=10, init=``random'',
-  MN\_ratio=0.9, FP\_ratio=2.0 shows a non-linear structure (see
-  Figure~\ref{fig-nldervisMNISTPACMAP} (a)). I selected this parameter
-  choice because this the default parameter setting for PaCMAP and
-  selected PaCMAP because when I observed the 5 NLDR techniques applied
-  for the data PaCMAP showed the non-linear structure effectively in 2D.
-\item
-  By looking at MSE plot, I selected the 180 bins, which associated with
-  12 bins along the x-axis, 15 bins along the y-axis, and hex size as
-  0.06 (Figure~\ref{fig-diagnosticpltMNIST} (b)). Then, fitted the model
-  in 2D (see Figure~\ref{fig-nldervisMNISTPACMAP} (b)). As you can see
-  in Figure~\ref{fig-nldervisMNISTPACMAP} (b), the fitted model capture
-  the non-linear structure of the 2D data.
-\end{itemize}
-
-\begin{figure}
 
@@
 PaCMAP applied to training MNIST dataset. What is the effective number
-of bins in each NLDR technique to create a 2D model? The MSE plot have a
-steep slope at the beginning, indicating that a smaller number of bins
-causes a larger amount of error. Then, the slope gradually declines or
-level off, indicating that a higher number of bins generates a smaller
-error. Using the elbow method, when the total number of bins is set to
-180, the slope of the Mean Squared Error (MSE) plot experiences a sudden
-and noticeable change, resembling an elbow-like shape. This point
-indicates that adding less bins does not enough to capture the data
-structure.}
-
-\end{figure}
-
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-91-1.pdf}
+of bins in PaCMAP to create a 2D model? The MSE plot have a steep slope
+at the beginning, indicating that a smaller number of bins causes a
+larger amount of error. Then, the slope gradually declines or level off,
+indicating that a higher number of bins generates a smaller error. Using
+the elbow method, when the total number of bins is set to 180, the slope
+of the Mean Squared Error (MSE) plot experiences a sudden and noticeable
+change, resembling an elbow-like shape. This point indicates that adding
+less bins does not enough to capture the data structure.}
+
+\end{figure}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervisMNISTPACMAP-1.pdf}
-
-}
-
-\caption{\label{fig-nldervisMNISTPACMAP}(a) 2D layout from PaCMAP
-(n\_components=2, n\_neighbors=10, init=random, MN\_ratio=0.9,
-FP\_ratio=2.0) applied for the digit 1 of the MNIST dataset, (b) Model
-in the 2D space overlaid on PaCMAP data, and (c) Error occurred by the
-high-D model for high-D data and the error mapped into 2D NLDR points.}
-
-\end{figure}
-
-\begin{itemize}
-\tightlist
-\item
-  Next, the constructed 2D model lifts into high-dimensions (see
-  Figure~\ref{fig-mnist1_sc}).
-\end{itemize}
-
-\begin{figure}[H]
-
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/mnist/sc_1.png}
-
-}
-
-}
-
-\subcaption{\label{fig-mnist1_sc1}}
-\end{minipage}%
-%
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/mnist/sc_2.png}
-
-}
-
-}
-
-\subcaption{\label{fig-mnist1_sc2}}
-\end{minipage}%
-%
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/mnist/sc_3.png}
-
-}
-
-}
-
-\subcaption{\label{fig-mnist1_sc3}}
-\end{minipage}%
-
-\caption{\label{fig-mnist1_sc}Screen shots of the \textbf{langevitour}
-of the MNIST digit 1 data set, shows the model in high-D, a video of the
-tour animation is available at (\url{https://youtu.be/zq2GM9qvUNA}).}
-
-\end{figure}
-
-\begin{itemize}
-\item
-  As shown in Figure~\ref{fig-mnist1_sc} (a), the non-linear continuity
-  structure that you have seen in 2D representation of PaCMAP, also can
-  be seen in the model in high-D space as well. This gives us an
-  evidence that our model capture the non-linear structure of the data
-  effectively.
-\item
-  Furthermore, our model shows some addition pattern of data which
-  couldn't be seen in the 2D layout.
-\item
-  One is the twisted structure within the non-linear structure (see
-  Figure~\ref{fig-mnist1_sc} (c)).
-\item
-  Another thing is the connected corners of the non-linear structure
-  (see Figure~\ref{fig-mnist1_sc} (b)). The non-linear curve is bended.
-\item
-  To gives the evidence that PaCMAP preserve the local structure
-  (non-linear pattern), our model helps by computing the model errors
-  (Figure~\ref{fig-diagnosticpltMNIST} (c)).
-\end{itemize}
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-sampleimg-1.pdf}
-
-}
-
-\caption{\label{fig-sampleimg}Images of handwritten digit 1 which occur
-large model error.}
-
-\end{figure}
-
-\begin{itemize}
-\tightlist
-\item
-  As you can see in Figure~\ref{fig-diagnosticpltMNIST} (c), the largest
-  model error shows in the upper right side of the 2D layout. A sample
-  of the images that shows the largest error can be seen in
-  Figure~\ref{fig-sampleimg}. The reason for the large model error is
-  that the high-D points related to this largest error located away from
-  the high-D model points.
-\end{itemize}
+{\centering \includegraphics{paper_files/figure-pdf/fig-nldervisMNISTPACMAP-1.pdf}
+
+}
+
+\caption{\label{fig-nldervisMNISTPACMAP}(a) Model in the 2D space
+overlaid on PaCMAP data, and (b) high-D model error in model space.}
+
+\end{figure}
 

### 2024-03-22 paper.qmd
Added write up for digit 1 MNIST data

@@
          legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#fee0d2", "#fcbba1",
-                                         "#fc9272", "#fb6a4a", "#ef3b2c",
-                                         "#cb181d", "#a50f15")) +
+  scale_color_manual(values=rainbow_hcl(7)) +
   annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
@@
 @Yingfan2021 used this dataset to demonstrate how the PaCMAP preserves local structure.
-We select the 2D embedding of PaCMAP for the handwritten digit 1.
+We selected the 2D embedding of PaCMAP for the handwritten digit 1.
 As shown in @fig-pacmapauthor, we observe a nonlinear structure present in 2D space.
+Furthermore, the angle of digit 1 images varies along the nonlinear structure (see @fig-sampleimg).
 To answer the following questions,
@@
 we visualize the model constructed with PaCMAP overlaid on data in high-D space (model-in-data space).
+
+According to @fig-mnist1_sc1, the nonlinear continuity structure observed in the 2D representation of PaCMAP is also visible when visualizing the model overlaid on the data space.
+This provides evidence that the model accurately captures the structure of the high-D data.
+Also, the model has a twisted pattern within the nonlinear structure, which is an addition pattern of data that cannot be seen in the 2D representation (see @fig-mnist1_sc2).
+
+In order to visually assess the preservation of local structure, it's important to consider the edges in the model.
+In a model that effectively preserves local structure, there shouldn't be any long edges in high-D space.
+However, as shown in @fig-mnist1_sc3, some long edges exist in high-D space that are not identified as long edges in the 2D representation.
 
@@
 
-<!--images with position 1-->
+<!--images with positions-->
 ```{r}
@@
   mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
-  filter(instance %in% c(3710, 2391, 2385, 5030, 6475))
+  filter(instance %in% c(3710, 2391, 2385, 5030, 6475,
+                         7679, 3568, 1312, 3311, 1097,
+                         3552, 7853, 6489, 7689, 6690,
+                        1380, 6057, 2347, 5946, 3355,
+                        4175, 3997, 5378, 387, 1854,
+                        614, 3079, 1762, 5239, 3723,
+                        5748, 728, 7419, 7794, 6233,
+                        33, 2485, 5998, 318, 1761,
+                        5690, 165, 517, 6935, 2682,
+                        4962, 2264, 5563, 6369, 559,
+                        5188, 4849, 2666, 3448, 3055,
+                        3120, 6869, 6345, 4470, 7147)) |>
+  mutate(instance = as.character(instance))
 
 theme_set(theme_light())
-imge_error_sample1 <- pixels_gathered |> 
+img_sample <- pixels_gathered |> 
   ggplot(aes(x, y, fill = value)) +  
   geom_tile() +  
-  facet_wrap(~ instance, ncol = 5) +
+  facet_wrap(~ factor(instance, levels = c(3710, 2391, 2385, 5030, 6475,
+                         7679, 3568, 1312, 3311, 1097,
+                         3552, 7853, 6489, 7689, 6690,
+                        1380, 6057, 2347, 5946, 3355,
+                        4175, 3997, 5378, 387, 1854,
+                        614, 3079, 1762, 5239, 3723,
+                        5748, 728, 7419, 7794, 6233,
+                        33, 2485, 5998, 318, 1761,
+                        5690, 165, 517, 6935, 2682,
+                        4962, 2264, 5563, 6369, 559,
+                        5188, 4849, 2666, 3448, 3055,
+                        3120, 6869, 6345, 4470, 7147)), ncol = 10) +
   theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
@@
          legend.key.width = unit(0.25, 'cm'),
-        strip.text = element_blank())
-```
-
-<!--images with position 2-->
-```{r}
-#| warning: false
-#| echo: false
-
-pixels_gathered <-  mnist_data |>
-  mutate(instance = row_number()) |>
-  gather(pixel, value, -Label, -instance) |>
-  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
-  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
-  filter(instance %in% c(7679, 3568, 1312, 3311, 1097))
-
-theme_set(theme_light())
-imge_error_sample2 <- pixels_gathered |> 
-  ggplot(aes(x, y, fill = value)) +  
-  geom_tile() +  
-  facet_wrap(~ instance, ncol = 5) +
-  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm'),
-        strip.text = element_blank())
-```
-
-<!--images with position 3-->
-```{r}
-#| warning: false
-#| echo: false
-
-pixels_gathered <-  mnist_data |>
-  mutate(instance = row_number()) |>
-  gather(pixel, value, -Label, -instance) |>
-  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
-  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
-  filter(instance %in% c(3552, 7853, 6489, 7689, 6690))
-
-theme_set(theme_light())
-imge_error_sample3 <- pixels_gathered |> 
-  ggplot(aes(x, y, fill = value)) +  
-  geom_tile() +  
-  facet_wrap(~ instance, ncol = 5) +
-  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm'),
-        strip.text = element_blank())
-```
-
-<!--images with position 4-->
-```{r}
-#| warning: false
-#| echo: false
-
-pixels_gathered <-  mnist_data |>
-  mutate(instance = row_number()) |>
-  gather(pixel, value, -Label, -instance) |>
-  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
-  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
-  filter(instance %in% c(1380, 6057, 2347, 5946, 3355))
-
-theme_set(theme_light())
-imge_error_sample4 <- pixels_gathered |> 
-  ggplot(aes(x, y, fill = value)) +  
-  geom_tile() +  
-  facet_wrap(~ instance, ncol = 5) +
-  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm'),
-        strip.text = element_blank())
-```
-
-<!--images with position 5-->
-```{r}
-#| warning: false
-#| echo: false
-
-pixels_gathered <-  mnist_data |>
-  mutate(instance = row_number()) |>
-  gather(pixel, value, -Label, -instance) |>
-  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
-  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
-  filter(instance %in% c(4175, 3997, 5378, 387, 1854))
-
-theme_set(theme_light())
-imge_error_sample5 <- pixels_gathered |> 
-  ggplot(aes(x, y, fill = value)) +  
-  geom_tile() +  
-  facet_wrap(~ instance, ncol = 5) +
-  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm'),
-        strip.text = element_blank())
-```
-
-<!--images with position 6-->
-```{r}
-#| warning: false
-#| echo: false
-
-pixels_gathered <-  mnist_data |>
-  mutate(instance = row_number()) |>
-  gather(pixel, value, -Label, -instance) |>
-  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
-  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
-  filter(instance %in% c(614, 3079, 1762, 5239, 3723))
-
-theme_set(theme_light())
-imge_error_sample6 <- pixels_gathered |> 
-  ggplot(aes(x, y, fill = value)) +  
-  geom_tile() +  
-  facet_wrap(~ instance, ncol = 5) +
-  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm'),
-        strip.text = element_blank())
-```
-
-<!--images with position 7-->
-```{r}
-#| warning: false
-#| echo: false
-
-pixels_gathered <-  mnist_data |>
-  mutate(instance = row_number()) |>
-  gather(pixel, value, -Label, -instance) |>
-  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
-  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
-  filter(instance %in% c(5748, 728, 7419, 7794, 6233))
-
-theme_set(theme_light())
-imge_error_sample7 <- pixels_gathered |> 
-  ggplot(aes(x, y, fill = value)) +  
-  geom_tile() +  
-  facet_wrap(~ instance, ncol = 5) +
-  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm'),
-        strip.text = element_blank())
-```
-
-<!--images with position 8-->
-```{r}
-#| warning: false
-#| echo: false
-
-pixels_gathered <-  mnist_data |>
-  mutate(instance = row_number()) |>
-  gather(pixel, value, -Label, -instance) |>
-  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
-  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
-  filter(instance %in% c(33, 2485, 5998, 318, 1761))
-
-theme_set(theme_light())
-imge_error_sample8 <- pixels_gathered |> 
-  ggplot(aes(x, y, fill = value)) +  
-  geom_tile() +  
-  facet_wrap(~ instance, ncol = 5) +
-  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm'),
-        strip.text = element_blank())
-```
-
-<!--images with position 9-->
-```{r}
-#| warning: false
-#| echo: false
-
-pixels_gathered <-  mnist_data |>
-  mutate(instance = row_number()) |>
-  gather(pixel, value, -Label, -instance) |>
-  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
-  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
-  filter(instance %in% c(5690, 165, 517, 6935, 2682))
-
-theme_set(theme_light())
-imge_error_sample9 <- pixels_gathered |> 
-  ggplot(aes(x, y, fill = value)) +  
-  geom_tile() +  
-  facet_wrap(~ instance, ncol = 5) +
-  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm'),
-        strip.text = element_blank())
-```
-
-<!--images with position 10-->
-```{r}
-#| warning: false
-#| echo: false
-
-pixels_gathered <-  mnist_data |>
-  mutate(instance = row_number()) |>
-  gather(pixel, value, -Label, -instance) |>
-  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
-  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
-  filter(instance %in% c(4962, 2264, 5563, 6369, 559))
-
-theme_set(theme_light())
-imge_error_sample10 <- pixels_gathered |> 
-  ggplot(aes(x, y, fill = value)) +  
-  geom_tile() +  
-  facet_wrap(~ instance, ncol = 5) +
-  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm'),
-        strip.text = element_blank())
-```
-
-<!--images with position 11-->
-```{r}
-#| warning: false
-#| echo: false
-
-pixels_gathered <-  mnist_data |>
-  mutate(instance = row_number()) |>
-  gather(pixel, value, -Label, -instance) |>
-  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
-  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
-  filter(instance %in% c(5188, 4849, 2666, 3448, 3055))
-
-theme_set(theme_light())
-imge_error_sample11 <- pixels_gathered |> 
-  ggplot(aes(x, y, fill = value)) +  
-  geom_tile() +  
-  facet_wrap(~ instance, ncol = 5) +
-  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm'),
-        strip.text = element_blank())
-```
-
-<!--images with position 12-->
-```{r}
-#| warning: false
-#| echo: false
-
-pixels_gathered <-  mnist_data |>
-  mutate(instance = row_number()) |>
-  gather(pixel, value, -Label, -instance) |>
-  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
-  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
-  filter(instance %in% c(3120, 6869, 6345, 4470, 7147))
-
-theme_set(theme_light())
-imge_error_sample12 <- pixels_gathered |> 
-  ggplot(aes(x, y, fill = value)) +  
-  geom_tile() +  
-  facet_wrap(~ instance, ncol = 5) +
-  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm'),
-        strip.text = element_blank())
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "Images of the handwritten digit 1 are distributed along the left-to-right of first embedding axis of the PaCMAP nonlinear structure."
+        strip.text = element_blank(),
+        aspect.ratio = 1)
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "Images of the handwritten digit 1 are distributed along the right-to-left of first embedding axis of the PaCMAP nonlinear structure."
 #| label: fig-sampleimg
@@
 
-imge_error_sample1 + imge_error_sample2 + imge_error_sample3 + imge_error_sample4 + imge_error_sample5 + imge_error_sample6 + imge_error_sample7 + imge_error_sample8 + imge_error_sample9 + imge_error_sample10 + imge_error_sample11 + imge_error_sample12 +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(ncol = 2, nrow = 6) &
-  theme(plot.tag = element_text(size = 8))
-```
-
-<!-- MSE plot to find which is the effective number of bins to use-->
-```{r}
-#| warning: false
-#| echo: false
-#| message: false
-
-## UMAP
-## Prediction
-
-#tot_num_bin_vec <- 2:51
-hex_size_vec <- seq(0.02, 2, by = 0.01)
-
-vec <- stats::setNames(rep("", 6), c("num_bins", "aic", "mse", "num_bins_x", "num_bins_y", "hex_size"))  ## Define column names
-
-mse_df_mnist <- dplyr::bind_rows(vec)[0, ]
-mse_df_mnist <- mse_df_mnist |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(hex_size_vec)) {
-
-  num_bin_list <- calc_bins(data = pacmap_minst_scaled, 
-            x = "PaCMAP1", y = "PaCMAP2", 
-            hex_size = hex_size_vec[i], buffer_x = NA, buffer_y = NA)
-
-  num_bins_x <- num_bin_list$num_x
-  num_bins_y <- num_bin_list$num_y
-
-  model_object <- fit_highd_model( training_data = training_data_mnist, 
-                                   nldr_df_with_id = pacmap_minst_scaled, 
-                                   x = "PaCMAP1", y = "PaCMAP2", 
-                                   num_bins_x = num_bins_x, 
-                                   num_bins_y = num_bins_y, 
-                                   x_start = NA, y_start = NA, 
-                                   buffer_x = NA, buffer_y = NA, 
-                                   hex_size = hex_size_vec[i],
-                                   is_rm_lwd_hex = FALSE, 
-                                   benchmark_to_rm_lwd_hex = NA, 
-                                   col_start_2d = "PaCMAP", 
-                                   col_start_highd = "PC")
-
-  centroid_df_training <- model_object$df_bin_centroids
-  avg_df_training <- model_object$df_bin
-
-  pred_emb_list <- predict_emb(test_data = training_data_mnist, 
-                                  df_bin_centroids = centroid_df_training, 
-                                  df_bin = avg_df_training, type_NLDR = "PaCMAP")
-
-  pred_df_training <- as.data.frame(do.call(cbind, pred_emb_list))
-
-  eval_list <- gen_summary(test_data = training_data_mnist, 
-                                  prediction_df = pred_df_training, 
-                                  df_bin = avg_df_training, col_start = "PC")
-
-  mse_df_mnist <- mse_df_mnist |>
-    tibble::add_row(num_bins = num_bins_x * num_bins_y,
-                    aic = eval_list$aic,
-                    mse = eval_list$mse,
-                    num_bins_x = num_bins_x,
-                    num_bins_y = num_bins_y,
-                    hex_size = hex_size_vec[i])
-
-}
-
-## If same total number of bins occurred only select ones with minimum error
-### Obtain duplicate bins
-dupli_bins <- mse_df_mnist |> 
-  dplyr::count(num_bins) |> 
-  dplyr::filter(n > 1) |> 
-  dplyr::pull(num_bins)
-
-### Group split by duplicated bins
-duplicate_df_list <- mse_df_mnist |>
-  dplyr::filter(num_bins %in% dupli_bins) |>
-  dplyr::arrange(num_bins) |>
-  dplyr::group_split(num_bins)
-
-### Obtain one row from duplicates which have lowest error and hexsize
-duplicate_df <- data.frame(matrix(nrow = 0, ncol = 0))
-
-for (i in 1:length(duplicate_df_list)) {
-
-  dd <- duplicate_df_list[[i]] |>
-    dplyr::filter(mse == min(duplicate_df_list[[i]]$mse)) |>
-    dplyr::filter(hex_size == min(duplicate_df_list[[i]]$hex_size))
-
-  duplicate_df <- dplyr::bind_rows(duplicate_df, dd)
-
-}
-
-### Obtain the mse_df with not duplicated bins
-not_dupli_df <- mse_df_mnist |>
-  dplyr::filter(!(num_bins %in% dupli_bins))
-
-### Combine duplicated and not duplicated(corrected) bins dfs
-mse_df_mnist <- dplyr::bind_rows(not_dupli_df, duplicate_df)
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-mse_plot_mnist <- ggplot(mse_df_mnist, aes(x = num_bins,
-                                       y = mse
-)) +
-  geom_point() +
-  geom_line() +
-   geom_vline(xintercept = 180, linetype="solid",
-                color = "red", size=0.8, alpha = 0.5) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  #scale_colour_manual(values = c("#377eb8")) +
-  ylab("MSE") +
-  xlab("total number of bins")
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "Goodness of fit statistics from PaCMAP applied to training MNIST dataset.
-What is the effective number of bins in PaCMAP to create a 2D model?
-The MSE plot have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
-Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error.
-Using the elbow method, when the total number of bins is set to 180, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
-This point indicates that adding less bins does not enough to capture the data structure."
-#| label: fig-diagnosticpltMNIST
-#| fig-pos: H
-
-mse_plot_mnist
+img_sample
 ```
@@
 error_df <- error_df |>
-  mutate(type = if_else(total == 0, "no error",
-                        if_else(total <= 2, "error 0-2",
+  mutate(type = if_else(total <= 2, "error 0-2",
                         if_else(total <= 4, "error 2-4",
                                 if_else(total <= 6, "error 4-6",
-                                                if_else(total <= 10, "error 6-10",
+                                        if_else(total <= 8, "error 6-8",
+                                                if_else(total <= 10, "error 8-10",
                                                         if_else(total <= 12, "error 10-12",
@@
 plot_list2_mnist_error <- error_df |>
-  mutate(type = factor(type , levels = c("no error", "error 0-2", "error 2-4", "error 4-6", "error 6-10", "error 10-12", "error 12-14", "error greter than 14"))) |>
+  mutate(type = factor(type , levels = c("error 0-2", "error 2-4", "error 4-6", "error 6-8", "error 8-10", "error 10-12", "error 12-14", "error greter than 14"))) |>
     ggplot(aes(x = PaCMAP1,
@@
          legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#fee0d2", "#fcbba1",
-                                         "#fc9272", "#fb6a4a", "#ef3b2c",
-                                         "#cb181d", "#a50f15", "#99000d")) +
+  scale_color_manual(values=rainbow_hcl(8)) +
   annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
@@
 #| echo: false
-#| fig-cap: "(a) Model in the 2D space overlaid on PaCMAP data, and (b) high-D model error in model space."
+#| fig-cap: "(a) Model generated with 12 and 15 bins along x and y axes with hexagonal size 0.06 in the 2D space overlaid on PaCMAP data, and (b) high-D model error in model space."
 #| label: fig-nldervisMNISTPACMAP
@@
 trimesh_removed_mnist_pacmap + plot_list2_mnist_error +
-plot_layout(guides='collect', ncol=3) &
+plot_layout(guides='collect', ncol=2) &
   theme(legend.position='bottom')
+```
+
+::: {#fig-mnist1_sc layout-ncol="3" fig-pos="H"}
+![](figures/mnist/sc_1.png){#fig-mnist1_sc1}
+
+![](figures/mnist/sc_2.png){#fig-mnist1_sc2}
+
+![](figures/mnist/sc_3.png){#fig-mnist1_sc3}
+
+Screen shots of the **langevitour** of the MNIST digit 1 data set, shows the model-in-data space, a video of the tour animation is available at (<https://youtu.be/zq2GM9qvUNA>).
+:::
+
+In terms of model error, some data points show high error (see @fig-nldervisMNISTPACMAP) due to their deviation from the typical high-D data structure, acting as outliers.
+This is evident when comparing the images associated with these high model error points (see @fig-sampleimgerror) to others (see @fig-sampleimg).
+
+<!--images that occur large error-->
+```{r}
+#| warning: false
+#| echo: false
+
+img_error <- error_df |> 
+  dplyr::filter(total >= 10) |> 
+  dplyr::pull(ID)
+
+mnist_data <- read_rds("data/mnist/mnist_digit_1.rds")
+pixels_gathered <-  mnist_data |>
+  mutate(instance = row_number()) |>
+  gather(pixel, value, -Label, -instance) |>
+  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
+  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
+  filter(instance %in% img_error)
+
+theme_set(theme_light())
+imge_error_sample <- pixels_gathered |> 
+  ggplot(aes(x, y, fill = value)) +  
+  geom_tile() +  
+  facet_wrap(~ instance, ncol = 11) +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm'),
+        strip.text = element_blank(),
+        aspect.ratio = 1)
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "Images of handwritten digit 1 which occur large model error."
+#| label: fig-sampleimgerror
+#| out-width: 100%
+#| fig-pos: H
+
+imge_error_sample
 ```

### 2024-03-22 paper.qmd
Added absolute error plot
with different UMAP parameter choices

@@
 
+@chen2023
+
 <!-- UMAP layout with author's suggested parameter choice-->
@@
 
-# df_all_mnist <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), umap_data_with_hb_id_pbmc)
+# df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), umap_data_with_hb_id_pbmc)
 # 
@@
 error_df <- error_df |>
-  dplyr::mutate(type = paste0(n_neighbors, "_", min_dist))
+  dplyr::mutate(type = paste0("n_neighbors: ", n_neighbors, ", min_dist: ", min_dist))
 
 abs_error_plot_pbmc <- ggplot(error_df, aes(x = num_non_empty_bins,
-                                       y = error, group = type,
+                                       y = log(error), group = type,
 colour = type)) +
@@
   geom_line() +
-   # geom_vline(xintercept = 588, linetype="solid",
-   #              color = "red", size=0.8, alpha = 0.5) +
+   geom_vline(xintercept = 278, linetype="solid",
+                color = "black", size=0.8, alpha = 0.5) +
   theme_light() +
-  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+  theme(legend.position = "bottom", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
         axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
-  ylab("absolute error") +
+        axis.text = element_text(size = 7),
+        legend.text = element_text(size = 3)) +
+  scale_color_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99','#e31a1c','#fdbf6f','#ff7f00','#cab2d6','#6a3d9a')) +
+  ylab("log(absolute error)") +
   xlab("number of non-empty bins")
 
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "Absolute error from UMAP applied to training PBMC3k dataset with diffferent parameter choices.
+What is the best parameter choice to create the model?
+The residual plot have a steep slope at the beginning, indicating that a smaller number of non-empty bins causes a larger amount of error.
+Then, the slope gradually declines or level off, indicating that a higher number of non-empty bins generates a smaller error.
+Using the elbow method, it was observed that when the number of non-empty bins is set to 278, the lowest error occurred with the parameters n_neighbors: 5 and min_dist: 0.3."
+#| label: fig-abserrorPBMC
+#| fig-pos: H
+#| out-width: 100%
+
 abs_error_plot_pbmc
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## Import data
+training_data_pbmc <- read_rds("data/pbmc3k/pbmc_pca_50.rds")
+training_data_pbmc <- training_data_pbmc[, 1:9] |>
+  mutate(ID = 1:NROW(training_data_pbmc))
+
+umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_5_min_dist_0.3.rds")
+umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc, 
+                                                                 x = "UMAP1", y = "UMAP2"))) |>
+  dplyr::mutate(cell_label = umap_pbmc$cell_label) |>
+  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
+                  "UMAP2" = "scaled_UMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(umap_pbmc))
+
+umap_plot_pbmc <- umap_pbmc_scaled |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2))+
+  geom_point(alpha=0.5) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm')) 
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "(a) 2D layout from UMAP (n_neighbors = 5, min_dist = 0.3) applied for the PBMC3k dataset. Is this a best representation of the original data?"
+#| label: fig-umapsuggest
+#| fig-pos: H
+
+umap_plot_pbmc 
+```
+
+<!--Fit the model-->
+
+```{r}
+#| warning: false
+#| echo: false
+#| message: false
+
+## Decide by looking at MSE plot
+num_bins_x_pbmc <- 21
+num_bins_y_pbmc <- 53
+hex_size_pbmc <- 0.03
+
+hb_obj_pbmc <- hex_binning(data = umap_pbmc_scaled, x = "UMAP1", 
+                      y = "UMAP2", num_bins_x = num_bins_x_pbmc, 
+                      num_bins_y = num_bins_y_pbmc, x_start = NA, y_start = NA, 
+                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_pbmc, col_start = "UMAP")
+
+## Data set with all possible centroids in the hexagonal grid
+all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_pbmc$centroids))
+
+## Generate all coordinates of hexagons
+hex_grid <- as.data.frame(do.call(cbind, hb_obj_pbmc$hex_poly))
+
+## To obtain the standardise counts within hexbins
+counts_df <- as.data.frame(do.call(cbind, hb_obj_pbmc$std_cts))
+df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
+                                             counts_df = counts_df)
+# ggplot(data = hex_grid, aes(x = x, y = y)) +
+#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
+#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
+#   coord_fixed()
+
+hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
+
+# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
+#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
+#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
+#   coord_fixed()
+
+umap_data_with_hb_id_pbmc <- as.data.frame(do.call(cbind, hb_obj_pbmc$data_hb_id))
+
+model_object <- fit_highd_model( training_data = training_data_pbmc, 
+                                 nldr_df_with_id = umap_pbmc_scaled, 
+                                 x = "UMAP1", y = "UMAP2", 
+                                 num_bins_x = num_bins_x_pbmc, 
+                                 num_bins_y = num_bins_y_pbmc, 
+                                 x_start = NA, y_start = NA, 
+                                 buffer_x = NA, buffer_y = NA, 
+                                 hex_size = hex_size_pbmc,
+                                 is_rm_lwd_hex = FALSE, 
+                                 benchmark_to_rm_lwd_hex = NA, 
+                                 col_start_2d = "UMAP", 
+                                 col_start_highd = "PC_")
+
+df_bin_centroids_pbmc <- model_object$df_bin_centroids
+df_bin_pbmc <- model_object$df_bin
+
+## Triangulate bin centroids
+tr1_object_pbmc <- tri_bin_centroids(df_bin_centroids_pbmc, x = "c_x", y = "c_y")
+tr_from_to_df_pbmc <- gen_edges(tri_object = tr1_object_pbmc)
+
+## Compute 2D distances
+distance_pbmc <- cal_2d_dist(tr_coord_df = tr_from_to_df_pbmc, 
+                             start_x = "x_from", start_y = "y_from", 
+                             end_x = "x_to", end_y = "y_to", 
+                             select_vars = c("from", "to", "distance"))
+
+# distance_plot <- plot_dist(distance_pbmc) +
+#   ylab(expression(d^{(2)})) +
+#   theme(axis.text = element_text(size = 5),
+#         axis.title = element_text(size = 12))
+# 
+# distance_plot
+
+## To find the benchmark value
+benchmark_pbmc <- find_lg_benchmark(distance_edges = distance_pbmc, distance_col = "distance")
+
+# ggplot() +
+# geom_trimesh(data = df_bin_centroids_pbmc, mapping = aes(x = c_x, y = c_y))
+
+trimesh_removed_pbmc_umap <- vis_rmlg_mesh(distance_edges = distance_pbmc, benchmark_value = benchmark_pbmc, tr_coord_df = tr_from_to_df_pbmc, distance_col = "distance")
+
+trimesh_removed_pbmc_umap <- trimesh_removed_pbmc_umap +
+  geom_point(data = umap_pbmc_scaled, aes(x = UMAP1, y = UMAP2), alpha = 0.05) +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+# df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), umap_data_with_hb_id_pbmc)
+# 
+# show_langevitour(df_all_pbmc, df_bin_pbmc, df_bin_centroids_pbmc, benchmark_pbmc, distance_pbmc, "distance", col_start = "PC_")
+
+df_bin_train <- df_bin_pbmc
+names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])
+
+df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), umap_data_with_hb_id_pbmc)
+
+error_df <- df_all_pbmc |>
+  dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates
+
+for (i in 1:(NCOL(df_bin_train) - 1)) {
+
+  error_df[ , paste0("abs_residual_", "PC_", i)] <- abs(error_df[ , paste0("PC_", i)] - error_df[ , paste0("avg_", "PC_", i)])
+
+}
+
+error_df <- error_df |>
+  dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "PC_")))))
+
+error_df <- error_df |>
+  mutate(type = if_else(total == 0, "no error",
+                        if_else(total <= 5, "error 0-5",
+                                if_else(total <= 10, "error 5-10",
+                                        if_else(total <= 15, "error 10-15",
+                                                if_else(total <= 20, "error 15-20",
+                                                        "error greter than 20"))))))
+
+plot_list2_pbmc_error <- error_df |>
+  mutate(type = factor(type , levels = c("no error", "error 0-5", "error 5-10", "error 10-15", "error 15-20", "error greter than 20"))) |>
+    ggplot(aes(x = UMAP1,
+               y = UMAP2, color = type))+
+    geom_point(alpha=0.5, size = 0.1) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=rainbow_hcl(6)) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "(a) Model in the 2D space overlaid on UMAP data, and (b) Error occurred by the high-D model for high-D data and the error mapped into 2D NLDR points."
+#| label: fig-nldervisPBMCUMAP2
+#| fig-pos: H
+
+trimesh_removed_pbmc_umap + plot_list2_pbmc_error +
+plot_layout(guides='collect', ncol=2) &
+  theme(legend.position='bottom') 
 ```

### 2024-03-23 paper.qmd
Select the best representation for PBMC3k

--- 2024-03-22 paper.qmd
@@
                                                                  x = "UMAP1", y = "UMAP2"))) |>
-  dplyr::mutate(cell_label = umap_pbmc$cell_label) |>
   dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
@@
 umap_plot_pbmc 
-```
-
-<!-- MSE plot to find which is the effective number of bins to use-->
-```{r}
-#| warning: false
-#| echo: false
-#| message: false
-
-## UMAP
-## Prediction
-
-#tot_num_bin_vec <- 2:51
-hex_size_vec <- seq(0.02, 2, by = 0.01)
-
-vec <- stats::setNames(rep("", 7), c("num_bins", "error", "mse", "num_bins_x", "num_bins_y", "hex_size", "num_non_empty_bins"))  ## Define column names
-
-mse_df_pbmc <- dplyr::bind_rows(vec)[0, ]
-mse_df_pbmc <- mse_df_pbmc |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(hex_size_vec)) {
-
-  num_bin_list <- calc_bins(data = umap_pbmc_scaled, 
-            x = "UMAP1", y = "UMAP2", 
-            hex_size = hex_size_vec[i], buffer_x = NA, buffer_y = NA)
-
-  num_bins_x <- num_bin_list$num_x
-  num_bins_y <- num_bin_list$num_y
-
-  model_object <- fit_highd_model( training_data = training_data_pbmc, 
-                                   nldr_df_with_id = umap_pbmc_scaled, 
-                                   x = "UMAP1", y = "UMAP2", 
-                                   num_bins_x = num_bins_x, 
-                                   num_bins_y = num_bins_y, 
-                                   x_start = NA, y_start = NA, 
-                                   buffer_x = NA, buffer_y = NA, 
-                                   hex_size = hex_size_vec[i],
-                                   is_rm_lwd_hex = FALSE, 
-                                   benchmark_to_rm_lwd_hex = NA, 
-                                   col_start_2d = "UMAP", 
-                                   col_start_highd = "PC_")
-
-  centroid_df_training <- model_object$df_bin_centroids
-  avg_df_training <- model_object$df_bin
-
-  pred_emb_list <- predict_emb(test_data = training_data_pbmc, 
-                                  df_bin_centroids = centroid_df_training, 
-                                  df_bin = avg_df_training, type_NLDR = "UMAP")
-
-  pred_df_training <- as.data.frame(do.call(cbind, pred_emb_list))
-
-  eval_list <- gen_summary(test_data = training_data_pbmc, 
-                                  prediction_df = pred_df_training, 
-                                  df_bin = avg_df_training, col_start = "PC_")
-
-  mse_df_pbmc <- mse_df_pbmc |>
-    tibble::add_row(num_bins = num_bins_x * num_bins_y,
-                    error = eval_list$error,
-                    mse = eval_list$mse,
-                    num_bins_x = num_bins_x,
-                    num_bins_y = num_bins_y,
-                    hex_size = hex_size_vec[i],
-                    num_non_empty_bins = NROW(centroid_df_training))
-
-}
-
-## If same total number of bins occurred only select ones with minimum error
-### Obtain duplicate bins
-dupli_bins <- mse_df_pbmc |> 
-  dplyr::count(num_bins) |> 
-  dplyr::filter(n > 1) |> 
-  dplyr::pull(num_bins)
-
-### Group split by duplicated bins
-duplicate_df_list <- mse_df_pbmc |>
-  dplyr::filter(num_bins %in% dupli_bins) |>
-  dplyr::arrange(num_bins) |>
-  dplyr::group_split(num_bins)
-
-### Obtain one row from duplicates which have lowest error and hexsize
-duplicate_df <- data.frame(matrix(nrow = 0, ncol = 0))
-
-for (i in 1:length(duplicate_df_list)) {
-
-  dd <- duplicate_df_list[[i]] |>
-    dplyr::filter(mse == min(duplicate_df_list[[i]]$mse)) |>
-    dplyr::filter(hex_size == min(duplicate_df_list[[i]]$hex_size))
-
-  duplicate_df <- dplyr::bind_rows(duplicate_df, dd)
-
-}
-
-### Obtain the mse_df with not duplicated bins
-not_dupli_df <- mse_df_pbmc |>
-  dplyr::filter(!(num_bins %in% dupli_bins))
-
-### Combine duplicated and not duplicated(corrected) bins dfs
-mse_df_pbmc <- dplyr::bind_rows(not_dupli_df, duplicate_df)
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-mse_plot_pbmc <- ggplot(mse_df_pbmc, aes(x = num_bins,
-                                       y = mse
-)) +
-  geom_point() +
-  geom_line() +
-   geom_vline(xintercept = 588, linetype="solid",
-                color = "red", size=0.8, alpha = 0.5) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  ylab("MSE") +
-  xlab("total number of bins")
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "Goodness of fit statistics from UMAP applied to training PBMC3k dataset.
-What is the effective number of bins in UMAP to create a 2D model?
-The MSE plot have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
-Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error.
-Using the elbow method, when the total number of bins is set to 588, the slope of the MSE plot experiences a sudden and noticeable change, resembling an elbow-like shape.
-This point indicates that adding less bins does not enough to capture the data structure."
-#| label: fig-msepltpbmc
-#| fig-pos: H
-
-mse_plot_pbmc
 ```
@@
 #| echo: false
-#| fig-cap: "(a) Model in the 2D space overlaid on UMAP data, and (b) Error occurred by the high-D model for high-D data and the error mapped into 2D NLDR points."
+#| fig-cap: "(a) Model generated with 21 and 28 bins along x and y axes with hexagonal size 0.03 in the 2D space overlaid on UMAP data, and (b) high-D model error in model space."
 #| label: fig-nldervisPBMCUMAP
@@
 error_df <- error_df |>
-  dplyr::mutate(type = paste0(n_neighbors, "_", min_dist))
+  dplyr::mutate(type = paste0("n_neighbors: ", n_neighbors, ", min_dist: ", min_dist))
+
+error_df_tsne <- read_csv("data/pbmc3k/error_df_tsne.csv", col_names = FALSE)
+names(error_df_tsne) <- c("num_bins", "error", "mse", "num_bins_x", "num_bins_y", "hex_size", "num_non_empty_bins", "perplexity")
+error_df_tsne <- error_df_tsne |>
+  dplyr::mutate(type = paste0("perplexity: ", perplexity))
+
+error_df <- bind_rows(error_df[, c(1:7, 10)], error_df_tsne[, c(1:7, 9)])
 
 abs_error_plot_pbmc <- ggplot(error_df, aes(x = num_non_empty_bins,
-                                       y = error, group = type,
+                                       y = log(error), group = type,
 colour = type)) +
@@
   geom_line() +
-   # geom_vline(xintercept = 588, linetype="solid",
-   #              color = "red", size=0.8, alpha = 0.5) +
+   geom_vline(xintercept = 439, linetype="solid",
+                color = "black", size=0.8, alpha = 0.5) +
   theme_light() +
-  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+  theme(legend.position = "bottom", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
         axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  scale_color_manual(values=c("#b15928", "#1f78b4", "#cab2d6", "#ccebc5", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#ffed6f", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#b2df8a", "#bc80bd", "#33a02c", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
-  ylab("absolute error") +
+        axis.text = element_text(size = 7),
+        legend.text = element_text(size = 5)) +
+  scale_color_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99','#e31a1c','#fdbf6f','#ff7f00','#cab2d6','#6a3d9a', '#b15928', '#ffff99')) +
+  ylab("log(absolute error)") +
   xlab("number of non-empty bins")
 
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "Absolute error from UMAP and tSNE applied to training PBMC3k dataset with diffferent parameter choices.
+What is the best parameter choice to create the model?
+The residual plot have a steep slope at the beginning, indicating that a smaller number of non-empty bins causes a larger amount of error.
+Then, the slope gradually declines or level off, indicating that a higher number of non-empty bins generates a smaller error.
+Using the elbow method, it was observed that when the number of non-empty bins is set to 439, the lowest error occurred with the parameters perplexity: 51."
+#| label: fig-abserrorPBMC
+#| fig-pos: H
+#| out-width: 100%
+
 abs_error_plot_pbmc
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+## Import data
+training_data_pbmc <- read_rds("data/pbmc3k/pbmc_pca_50.rds")
+training_data_pbmc <- training_data_pbmc[, 1:9] |>
+  mutate(ID = 1:NROW(training_data_pbmc))
+
+tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_51.rds")
+tsne_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tsne_pbmc, 
+                                                                 x = "tSNE1", y = "tSNE2"))) |>
+  dplyr::rename(c("tSNE1" = "scaled_tSNE1", 
+                  "tSNE2" = "scaled_tSNE2")) |>
+  dplyr::mutate(ID = 1:NROW(tsne_pbmc))
+
+tsne_plot_pbmc <- tsne_pbmc_scaled |>
+  ggplot(aes(x = tSNE1,
+             y = tSNE2))+
+  geom_point(alpha=0.5) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm')) 
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "(a) 2D layout from tSNE (perplexity = 51) applied for the PBMC3k dataset. Is this a best representation of the original data?"
+#| label: fig-tsnesuggest
+#| fig-pos: H
+
+tsne_plot_pbmc 
+```
+
+<!--Fit the model-->
+
+```{r}
+#| warning: false
+#| echo: false
+#| message: false
+
+## Decide by looking at MSE plot
+num_bins_x_pbmc <- 31
+num_bins_y_pbmc <- 40
+hex_size_pbmc <- 0.02
+
+hb_obj_pbmc <- hex_binning(data = tsne_pbmc_scaled, x = "tSNE1", 
+                      y = "tSNE2", num_bins_x = num_bins_x_pbmc, 
+                      num_bins_y = num_bins_y_pbmc, x_start = NA, y_start = NA, 
+                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_pbmc, col_start = "tSNE")
+
+## Data set with all possible centroids in the hexagonal grid
+all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_pbmc$centroids))
+
+## Generate all coordinates of hexagons
+hex_grid <- as.data.frame(do.call(cbind, hb_obj_pbmc$hex_poly))
+
+## To obtain the standardise counts within hexbins
+counts_df <- as.data.frame(do.call(cbind, hb_obj_pbmc$std_cts))
+df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
+                                             counts_df = counts_df)
+# ggplot(data = hex_grid, aes(x = x, y = y)) +
+#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
+#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
+#   coord_fixed()
+
+hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
+
+# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
+#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
+#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
+#   coord_fixed()
+
+tsne_data_with_hb_id_pbmc <- as.data.frame(do.call(cbind, hb_obj_pbmc$data_hb_id))
+
+model_object <- fit_highd_model( training_data = training_data_pbmc, 
+                                 nldr_df_with_id = tsne_pbmc_scaled, 
+                                 x = "tSNE1", y = "tSNE2", 
+                                 num_bins_x = num_bins_x_pbmc, 
+                                 num_bins_y = num_bins_y_pbmc, 
+                                 x_start = NA, y_start = NA, 
+                                 buffer_x = NA, buffer_y = NA, 
+                                 hex_size = hex_size_pbmc,
+                                 is_rm_lwd_hex = FALSE, 
+                                 benchmark_to_rm_lwd_hex = NA, 
+                                 col_start_2d = "tSNE", 
+                                 col_start_highd = "PC_")
+
+df_bin_centroids_pbmc <- model_object$df_bin_centroids
+df_bin_pbmc <- model_object$df_bin
+
+benchmark_value_rm_lwd <- stats::quantile(df_bin_centroids_pbmc$std_counts,
+                probs = c(0,0.25,0.5,0.75,1), names = FALSE)[2]
+
+df_bin_centroids_low <- df_bin_centroids_pbmc |>
+  dplyr::filter(std_counts <= benchmark_value_rm_lwd)
+
+identify_rm_bins <- find_low_dens_hex(df_bin_centroids_all = df_bin_centroids_pbmc,
+                                      num_bins_x = num_bins_x_pbmc,
+                                      df_bin_centroids_low = df_bin_centroids_low)
+
+df_bin_centroids_pbmc <- df_bin_centroids_pbmc |>
+  dplyr::filter(!(hexID %in% identify_rm_bins))
+
+## Triangulate bin centroids
+tr1_object_pbmc <- tri_bin_centroids(df_bin_centroids_pbmc, x = "c_x", y = "c_y")
+tr_from_to_df_pbmc <- gen_edges(tri_object = tr1_object_pbmc)
+
+## Compute 2D distances
+distance_pbmc <- cal_2d_dist(tr_coord_df = tr_from_to_df_pbmc, 
+                             start_x = "x_from", start_y = "y_from", 
+                             end_x = "x_to", end_y = "y_to", 
+                             select_vars = c("from", "to", "distance"))
+
+# distance_plot <- plot_dist(distance_pbmc) +
+#   ylab(expression(d^{(2)})) +
+#   theme(axis.text = element_text(size = 5),
+#         axis.title = element_text(size = 12))
+# 
+# distance_plot
+
+## To find the benchmark value
+benchmark_pbmc <- find_lg_benchmark(distance_edges = distance_pbmc, distance_col = "distance")
+benchmark_pbmc <- 0.04
+
+# ggplot() +
+# geom_trimesh(data = df_bin_centroids_pbmc, mapping = aes(x = c_x, y = c_y))
+
+trimesh_removed_pbmc_tsne <- vis_rmlg_mesh(distance_edges = distance_pbmc, benchmark_value = benchmark_pbmc, tr_coord_df = tr_from_to_df_pbmc, distance_col = "distance")
+
+trimesh_removed_pbmc_tsne <- trimesh_removed_pbmc_tsne +
+  geom_point(data = tsne_pbmc_scaled, aes(x = tSNE1, y = tSNE2), alpha = 0.08) +
+  theme_linedraw() +
+  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+# df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), tsne_data_with_hb_id_pbmc)
+# 
+# show_langevitour(df_all_pbmc, df_bin_pbmc, df_bin_centroids_pbmc, benchmark_pbmc, distance_pbmc, "distance", col_start = "PC_")
+
+df_bin_train <- df_bin_pbmc
+names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])
+
+df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), tsne_data_with_hb_id_pbmc)
+
+error_df <- df_all_pbmc |>
+  dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates
+
+for (i in 1:(NCOL(df_bin_train) - 1)) {
+
+  error_df[ , paste0("abs_residual_", "PC_", i)] <- abs(error_df[ , paste0("PC_", i)] - error_df[ , paste0("avg_", "PC_", i)])
+
+}
+
+error_df <- error_df |>
+  dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "PC_")))))
+
+error_df <- error_df |>
+  mutate(type = if_else(total == 0, "no error",
+                        if_else(total <= 5, "error 0-5",
+                                if_else(total <= 10, "error 5-10",
+                                        if_else(total <= 15, "error 10-15",
+                                                if_else(total <= 20, "error 15-20",
+                                                        "error greter than 20"))))))
+
+plot_list2_pbmc_error <- error_df |>
+  mutate(type = factor(type , levels = c("no error", "error 0-5", "error 5-10", "error 10-15", "error 15-20", "error greter than 20"))) |>
+    ggplot(aes(x = tSNE1,
+               y = tSNE2, color = type))+
+    geom_point(alpha=0.5, size = 0.1) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm')) +
+  scale_color_manual(values=rainbow_hcl(6)) +
+  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "(a) Model generated with 31 and 40 bins along x and y axes with hexagonal size 0.02 in the 2D space overlaid on tSNE data, and (b) high-D model error in model space."
+#| label: fig-nldervisPBMCtSNE
+#| fig-pos: H
+
+trimesh_removed_pbmc_tsne + plot_list2_pbmc_error +
+plot_layout(guides='collect', ncol=2) &
+  theme(legend.position='bottom') 
 ```

### 2024-03-25 paper.qmd
Added screen shots for mnist model
Added starting write up for hexagonal binning

@@
 
-Hexagonating NLDR data (see @fig-meth Step 2) involves partitioning the 2D embedding data into hexagonal bins, a technique commonly referred to as hexagonal binning [@Carr1987, @article66].
+Hexagonating NLDR data (see @fig-meth Step 2) involves partitioning the NLDR data into hexagonal bins, a technique commonly referred to as hexagonal binning [@Carr1987, @article66].
 This method use a hexagonal grid to create a bivariate histogram that effectively visualizes the structure of high-D data.
@@
 This geometric property makes hexagons more efficient in covering the plane compared to other regular tessellations and reduces visual bias when displaying data densities [@Dan2023].
-In our algorithm, we aim to conduct regular hexagonal binning (**cite quollr paper**), involving the computation of hexagonal grid configurations, the generation of the grid, and the assignment of the 2D embedding to hexagons.
-
-<!--
-**(a) Determine the number of bins along the x axis** ($b_1$)
-
-**(b) Determine the number of bins along the y axis** ($b_2$)
-
-**(c) Determine the total number of bins** ($b$)
-
-**(d) Assign NLDR data to hexagons**-->
+
+In our algorithm, we aim to conduct regular hexagonal binning, involving the computation of hexagonal grid configurations, the generation of the grid, and the assignment of the NLDR data points to hexagons.
+
+**(a) Computation of hexagonal grid configurations**
+
+In the process of computing hexagonal bin configurations, the first step is to determine the number of bins along the x-axis.
+This begins with the initialization of parameters such as the hexagonal size ($s$) and the buffer along the x-axis ($buffer_{x}$).
+The hexagonal size ($s$) is the radius of outer circle surrounded by the hexagon.
+Subsequently, the range of the first 2D embedding component ($r_1$) is calculated to ensure comprehensive coverage of the data.
+This range is then adjusted by incorporating the buffer amount, effectively expanding the boundary to accommodate potential outliers or edge cases.
+The default buffer amount along the x-axis is $\sqrt{3} * s * 1.5$.
+Next, the horizontal spacing ($h$) between hexagons is computed based on $s$ (see @eq-equation11).
+Finally, the number of bins along the x-axis is computed based on the adjusted range ($r_1 + buffer_{x}$) and the horizontal spacing ($h$) (see @eq-equation12).
+
+$$
+ h = \sqrt{3} * s
+$$ {#eq-equation11}
+
+$$
+ b_1 = \frac{r_1 + buffer_{x}}{h}
+$$ {#eq-equation12}
+
+Next, the number of bins along the y-axis is computed using hexagonal size ($s$) and the buffer along the y-axis ($buffer_{y}$).
+Subsequently, the range of the second 2D embedding component ($r_2$) is calculated to ensure comprehensive coverage of the data.
+This range is then adjusted by incorporating the buffer amount, effectively expanding the boundary to accommodate potential outliers or edge cases.
+The default buffer amount along the y-axis is $1.5 * s * 1.5$.
+The horizontal spacing ($v$) between hexagons is computed based on $s$ (see @eq-equation13).
+Finally, the number of bins along the y-axis is computed based on the adjusted range ($r_2 + buffer_{y}$) and the horizontal spacing ($v$) (see @eq-equation14).
+
+$$
+  v = 1.5 * s
+$$ {#eq-equation13}
+
+$$
+ b_2 = \frac{r_2 + buffer_{y}}{v}
+$$ {#eq-equation14}
+
+**(b) Generation of the hexagonal grid**
+
+In this process, first involves the generation of centroids in the hexagonal grid.
+The hexagonal starting coordinates are defined.
+We already computed the number of bins along the x and y axes and have the horizontal and vertical spacing.
+By starting with hexagonal starting coordinates, iteratively compute the centroids by incorporating the number of bins and spacing along each axes.
+
+Once, we have the centroids for the hexagonal grid, the next step is to compute the hexgaonal coordinates. For example, as shown in Figure, the hexagonal coordinates can be computed.
+
+**(c) Assignment of the NLDR data points to hexagons**
+
+Once, we have the hexagonal bin centroids of the full grid, by using the 2D Euclidean distance find the nearest hexagonal bin for each NLDR point. Then assign that bin for the NLDR point.
 
@@
 Given that the hexagons are regular, the resulting triangles will mostly be equilateral.
-
-<!--
-
-**Step 1: Computing the hexagonal grid configuration**
-
-The hexagonal grid, formed through hexagonal binning [@Carr1987, @article66], serves as a type of bivariate histogram employed to visualize the structure of high-D data.
-Hexagons, being one of only three regular polygons capable of tessellating a plane [@Carr2013], possess both symmetry of nearest neighbors and the maximum number of sides for a regular tessellation of the plane.
-This unique combination makes hexagons more efficient in covering the plane compared to other regular tessellations.
-Additionally, hexagons exhibit lower visual bias when displaying densities, setting them apart from other regular tessellations [@Dan2023].
-In our algorithm, hexagonal binning is used as the initial step of constructing the 2D model and the total number of bins ($b$) is the crucial parameter that defines the granularity of the hexagonal grid.
-
-**(a) Determine the number of bins along the x axis** ($b_1$)
-
-First, the number of bins along the x-axis ($b_1$) is computed using the relationship between the diameter ($h$) and the area ($A$) of regular hexagons (see @eq-equation3).
-
-In the process of computing hexagonal bin configurations, the first step involves determining the number of bins along the x-axis.
-This begins with the initialization of parameters such as the hexagonal size ($s$) and the buffer along the x-axis.
-Subsequently, the range of the first 2D embedding component ($r_1$) is calculated to ensure comprehensive coverage of the data.
-This range is then adjusted by incorporating the buffer amount, effectively expanding the boundary to accommodate potential outliers or edge cases.
-The default buffer amount along the x-axis is $\sqrt{3} * s * 1.5$.
-Next, the horizontal spacing ($h$) between hexagons is computed based on the hexagon size.
-Finally, the number of bins along the x-axis is computed based on the adjusted range ($r_1 + buffer_{x}$) and the horizontal spacing ($h$).
-
-$$
- h = \sqrt{3} * s
-$$ {#eq-equation3}
-
-$$
- b_1 = \frac{r_1 + buffer_{x}}{h}
-$$ {#eq-equation4}
-
-**(b) Determine the number of bins along the y-axis** ($b_2$)
-
-Next, the number of bins along the y-axis is computed based on the number of bins along the x-axis ($b_1$) and the shape parameter ($s$) (see @eq-equation4) [@Carr2013].
-
-$$
-  v = 1.5 * s
-$$ {#eq-equation5}
-
-$$
- b_2 = \frac{r_2 + buffer_{y}}{v}
-$$ {#eq-equation6}
-
-**(c) Determine the total number of bins** ($b$)
-
-The total number of bins is determined by multiplying the number of bins along the x-axis ($b_1$) with the number of bins along the y-axis ($b_2$) (see @eq-equation5).
-
-$$
-b = b_1 \times b_2
-$${#eq-equation5}
--->
 
@@
 #| fig-pos: H
-#| fig-cap: "Hexbin plots from different number of bins for the **UMAP** embeddings of **S-curve** training data: (a) b = 12 (3, 4), h = 0.62, (b) b = 84 (6, 14), h = 0.13, and (c) b = 336 (12, 28), h = 0.06.
+#| fig-cap: "Hexbin plots from different number of bins for the **UMAP** embeddings of **S-curve** training data: (a) b = 12 (3, 4), s = 0.62, (b) b = 84 (6, 14), s = 0.13, and (c) b = 336 (12, 28), s = 0.06.
 The hexbins are colored based on the density of points, with darker colors indicating higher point density and yellow color representing lower point density within each bin.
@@
 Residuals are essential for evaluating the accuracy of representing high-D points by the high-D mapping of 2D bin centroids.
-To measure this accuracy, an error metric is introduced, quantifying the sum of squared differences between the high-D data ($x_{ij}$) and the high-D mapping of the 2D bin centroid data ($C_{x_ij}$) across all observations and dimensions (see @eq-equation11).
+To measure this accuracy, an error metric is introduced, quantifying the sum of squared differences between the high-D data ($x_{ij}$) and the high-D mapping of the 2D bin centroid data ($C_{x_ij}$) across all observations and dimensions (see @eq-equation10).
 
@@
 \text{Error} = \sum_{j = 1}^{n}\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2
-$$ {#eq-equation11}
+$$ {#eq-equation10}
 
 Here, $n$ represents the number of observations, $p$ represents the dimensions of high-D data, $x_{ij}$ is the high-D data, and $C_{x_ij}$ is the high-D mapping of the 2D bin centroid.
-
-<!--
-The concept of "residuals" is pivotal in evaluating the accuracy of representing bin centroids in high-D.
-To quantify this accuracy, we introduce an error metric, which measures the sum of squared differences between the high-D data ($x_{ij}$) and the predicted bin centroid data in high-D space ($C_{x_ij}$) across all bins and dimensions.
-Mathematically, this error is expressed as:
-
-$$
-\text{Error} = \sum_{j = 1}^{n}\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2
-$$ {#eq-equation11}
-
-Here, $n$ represents the number of observations, $p$ represents the dimensions, $x_{ij}$ is the high-D data, and $C_{x_ij}$ is the predicted bin centroid data in high-D.
-
-The error metric outlined above provides valuable insights into the overall accuracy of our predictive model.
-By quantifying the squared deviations between the actual and predicted values across all bins and dimensions, we gain a comprehensive understanding of how well our method captures and represents the underlying structure of the data in the reduced 2D space.
-This assessment is crucial for evaluating the efficacy of our NLDR technique in preserving the essential information present in the original high-dimensional data.
--->
 
@@
 
-### Simulated data example {#sec-simpleex}
+7### Simulated data example {#sec-simpleex}
 
@@
 
-<!-- To find the number of bins should assign for the model in tSNE-->
 ```{r}
@@
   dplyr::mutate(ID = 1:NROW(tSNE_data_gau))
-
-## tSNE
-hex_size_vec <- seq(0.02, 2, by = 0.01)
-
-vec <- stats::setNames(rep("", 6), c("num_bins", "aic", "mse", "num_bins_x", "num_bins_y", "hex_size"))  ## Define column names
-
-mse_df_gau <- dplyr::bind_rows(vec)[0, ]
-mse_df_gau <- mse_df_gau |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(hex_size_vec)) {
-
-  num_bin_list <- calc_bins(data = tsne_gau_scaled, 
-            x = "tSNE1", y = "tSNE2", 
-            hex_size = hex_size_vec[i], buffer_x = NA, buffer_y = NA)
-
-  num_bins_x <- num_bin_list$num_x
-  num_bins_y <- num_bin_list$num_y
-
-  model_object <- fit_highd_model( training_data = training_data_gau, 
-                                   nldr_df_with_id = tsne_gau_scaled, 
-                                   x = "tSNE1", y = "tSNE2", 
-                                   num_bins_x = num_bins_x, 
-                                   num_bins_y = num_bins_y, 
-                                   x_start = NA, y_start = NA, 
-                                   buffer_x = NA, buffer_y = NA, 
-                                   hex_size = hex_size_vec[i],
-                                   is_rm_lwd_hex = FALSE, 
-                                   benchmark_to_rm_lwd_hex = NA, 
-                                   col_start_2d = "tSNE", 
-                                   col_start_highd = "x")
-
-  centroid_df_training <- model_object$df_bin_centroids
-  avg_df_training <- model_object$df_bin
-
-  pred_emb_list <- predict_emb(test_data = training_data_gau, 
-                                  df_bin_centroids = centroid_df_training, 
-                                  df_bin = avg_df_training, type_NLDR = "tSNE")
-
-  pred_df_training <- as.data.frame(do.call(cbind, pred_emb_list))
-
-  eval_list <- gen_summary(test_data = training_data_gau, 
-                                  prediction_df = pred_df_training, 
-                                  df_bin = avg_df_training, col_start = "x")
-
-  mse_df_gau <- mse_df_gau |>
-    tibble::add_row(num_bins = num_bins_x * num_bins_y,
-                    aic = eval_list$aic,
-                    mse = eval_list$mse,
-                    num_bins_x = num_bins_x,
-                    num_bins_y = num_bins_y,
-                    hex_size = hex_size_vec[i])
-
-}
-
-## If same total number of bins occurred only select ones with minimum error
-### Obtain duplicate bins
-dupli_bins <- mse_df_gau |> 
-  dplyr::count(num_bins) |> 
-  dplyr::filter(n > 1) |> 
-  dplyr::pull(num_bins)
-
-### Group split by duplicated bins
-duplicate_df_list <- mse_df_gau |>
-  dplyr::filter(num_bins %in% dupli_bins) |>
-  dplyr::arrange(num_bins) |>
-  dplyr::group_split(num_bins)
-
-### Obtain one row from duplicates which have lowest error and hexsize
-duplicate_df <- data.frame(matrix(nrow = 0, ncol = 0))
-
-for (i in 1:length(duplicate_df_list)) {
-
-  dd <- duplicate_df_list[[i]] |>
-    dplyr::filter(mse == min(duplicate_df_list[[i]]$mse)) |>
-    dplyr::filter(hex_size == min(duplicate_df_list[[i]]$hex_size))
-
-  duplicate_df <- dplyr::bind_rows(duplicate_df, dd)
-
-}
-
-### Obtain the mse_df with not duplicated bins
-not_dupli_df <- mse_df_gau |>
-  dplyr::filter(!(num_bins %in% dupli_bins))
-
-### Combine duplicated and not duplicated(corrected) bins dfs
-mse_df_gau <- dplyr::bind_rows(not_dupli_df, duplicate_df)
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-## To draw with AIC
-aic_plot_gau <- ggplot(mse_df_gau, aes(x = num_bins, y = aic
-)) +
-  geom_point(size = 0.8) +
-  geom_line() +
-  geom_vline(xintercept = 42, linetype="solid",
-                color = "red", size=0.8, alpha = 0.5) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  #scale_colour_manual(values = c("#377eb8")) +
-  ylab("AIC") +
-  xlab("total number of bins")
-## Effective number of bins along x-axis
-
-mse_plot_gau <- ggplot(mse_df_gau, aes(x = num_bins,
-                                       y = mse
-)) +
-  geom_point(size = 0.8) +
-  geom_line() +
-   geom_vline(xintercept = 42, linetype="solid",
-                color = "red", size=0.8, alpha = 0.5) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  #scale_colour_manual(values = c("#377eb8")) +
-  ylab("MSE") +
-  xlab("total number of bins")
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "Goodness of fit statistics from different NLDR techniques applied to training five spherical Gaussian cluster dataset.
-What is the best NLDR technique to represent the original data in 2D?"
-#| label: fig-diagnosticpltGau
-#| out-width: 100%
-
-aic_plot_gau + mse_plot_gau +
-  plot_layout(nrow = 2) +
-  plot_annotation(tag_levels = 'a') +
-  #plot_layout(guides='collect', nrow = 2) & legend.position='bottom', 
-  theme(plot.tag = element_text(size = 8))
-```
-
-```{r}
-#| warning: false
-#| echo: false
-#| message: false
 

### 2024-03-25 paper.qmd
Added PBMC3k images with tour

@@
 
-7### Simulated data example {#sec-simpleex}
+### Simulated data example {#sec-simpleex}
 
@@
 
-@chen2023
+In the field of single-cell studies, a common analytical task involves clustering to identify groups of cells with similar expression profiles.
+Analysts often turn to NLDR techniques to verify and identify these clusters and explore developmental trajectories.
+To illustrate the importance of NLDR techniques and parameter selection in identifying clusters, Human Peripheral Blood Mononclear Cells (PBMC3k) dataset [@chen2023] is used.
+In a study by @chen2023, this dataset was used to demonstrate how UMAP represents clusters (see @fig-umapauthor).
+As shown in @fig-umapauthor, there are three distinct and well-separated clusters.
 
@@
 #| echo: false
-#| fig-cap: "(a) 2D layout from UMAP (n_neighbors = 30, min_dist = 0.3) applied for the PBMC3k dataset. Is this a best representation of the original data?"
+#| fig-cap: "2D layout from UMAP (n_neighbors = 30, min_dist = 0.3) applied for the PBMC3k dataset. Is this a best representation of the original data?"
 #| label: fig-umapauthor
@@
 ```
+
+To determine whether the UMAP representation with the parameter choice suggested by @chen2023 preserves the original data structure, we visualize the model constructed with UMAP overlaid on the high-D data (model-in-data space).
+The figures in @fig-pbmc1_sc show three well-separated clusters, indicating that the suggested UMAP representation preserves the global structure (see @fig-umapauthor).
+However, as shown in @fig-pbmc1_sc, these clusters are close to each other in high-D space.
+Also, nonlinear continuity patterns and high-density patches within the clusters are observed (see @fig-pbmc1_sc).
+Therefore, the suggested UMAP representation (see @fig-umapauthor) does not accurately preserve the local structure of the PBMC3k dataset.
 
@@
 ```
+
+::: {#fig-pbmc1_sc layout-ncol="3" fig-pos="H"}
+![](figures/pbmc3k/sc_1.png){#fig-pbmc1_sc1}
+
+![](figures/pbmc3k/sc_2.png){#fig-pbmc1_sc2}
+
+![](figures/pbmc3k/sc_3.png){#fig-pbmc1_sc3}
+
+Screen shots of the **langevitour** of the PBMC3k data set, shows the model in high-D, a video of the tour animation is available at (<https://youtu.be/Gnhh4hfsbbg>).
+:::
+
+To find the best NLDR representation for PBMC3k dataset, we observe the absolute error for different number of non-empty bins and select the one with the lowest error.
+As shown in @fig-abserrorPBMC, tSNE with a perplexity set to $51$, has the lowest error.
+Therefore, tSNE with a perplexity of $51$ is used as the best representation for PBMC3k dataset.
 
@@
 
+As shown in @fig-tsnesuggest, there are three well-separated clusters, although they are located close to each other.
+Additionally, nonlinear structures can be observed within the clusters (see @fig-tsnesuggest).
+Hence, the data structure that UMAP failed to capture for the PBMC3k dataset is successfully captured by tSNE.
+
 ```{r}
@@
 #| echo: false
-#| fig-cap: "(a) 2D layout from tSNE (perplexity = 51) applied for the PBMC3k dataset. Is this a best representation of the original data?"
+#| fig-cap: "2D layout from tSNE (perplexity = 51) applied for the PBMC3k dataset. Is this a best representation of the original data?"
 #| label: fig-tsnesuggest
@@
 
+Next, the model is fitted for tSNE, and the resulting model is visualized in the data space.
+The model shows some quirks, as shown in @fig-pbmc2_sc.
+The large cluster is divided into three subclusters, which are in close to each other.
+
 <!--Fit the model-->
-
 ```{r}
@@
 ```
+
+::: {#fig-pbmc2_sc layout-ncol="3" fig-pos="H"}
+![](figures/pbmc3k/sc_4.png){#fig-pbmc2_sc4}
+
+![](figures/pbmc3k/sc_5.png){#fig-pbmc2_sc5}
+
+![](figures/pbmc3k/sc_6.png){#fig-pbmc2_sc6}
+
+Screen shots of the **langevitour** of the PBMC3k data set, shows the model in high-D, a video of the tour animation is available at (<https://youtu.be/R0EroJcGbas>).
+:::
 

### 2024-03-25 paper.qmd
Added regular hexagonal binning section
with equations

@@
 
-In the process of computing hexagonal bin configurations, the first step is to determine the number of bins along the x-axis.
-This begins with the initialization of parameters such as the hexagonal size ($s$) and the buffer along the x-axis ($buffer_{x}$).
-The hexagonal size ($s$) is the radius of outer circle surrounded by the hexagon.
-Subsequently, the range of the first 2D embedding component ($r_1$) is calculated to ensure comprehensive coverage of the data.
-This range is then adjusted by incorporating the buffer amount, effectively expanding the boundary to accommodate potential outliers or edge cases.
-The default buffer amount along the x-axis is $\sqrt{3} * s * 1.5$.
-Next, the horizontal spacing ($h$) between hexagons is computed based on $s$ (see @eq-equation11).
-Finally, the number of bins along the x-axis is computed based on the adjusted range ($r_1 + buffer_{x}$) and the horizontal spacing ($h$) (see @eq-equation12).
+In the computation of hexagonal bin configurations, there are mainly two configurations to consider: the number of bins along the x and y axes.
+The first step involves in determining the hexagonal size ($s$) and the buffer along the x and y axes ($buffer_{x}$ and $buffer_{y}$).
+Here, the hexagonal size ($s$) represents the radius of the outer circle surrounding the hexagon.
+The buffer is important in expanding the boundary to accommodate potential outliers or edge cases.
+
+When computing the number of bins along the x and y axes ($b_1$ and $b_2$), the process begins by calculating the range of the respective 2D embedding component.
+After that, the range is adjusted to accommodate the buffer amount.
+Then, the spacing between hexagons is determined based on $s$.
+Finally, the number of bins along the axis is computed by dividing the adjusted range by the spacing.
+
+For the computation of the number of bins along the x-axis ($b_1$) (see @eq-equation12), the range of the first embedding component is defined as $r_1$, while the buffer amount along the x-axis is denoted as $buffer_{x}$, and the horizontal spacing is represented by $h$ (see @eq-equation11).
+On the other hand, the number of bins along the y-axis ($b_2$) (see @eq-equation14) is computed based on the range of the second embedding component ($r_2$), the buffer amount along the y-axis ($buffer_{y}$), and the vertical spacing ($v$) (see @eq-equation13).
 
@@
 
-Next, the number of bins along the y-axis is computed using hexagonal size ($s$) and the buffer along the y-axis ($buffer_{y}$).
-Subsequently, the range of the second 2D embedding component ($r_2$) is calculated to ensure comprehensive coverage of the data.
-This range is then adjusted by incorporating the buffer amount, effectively expanding the boundary to accommodate potential outliers or edge cases.
-The default buffer amount along the y-axis is $1.5 * s * 1.5$.
-The horizontal spacing ($v$) between hexagons is computed based on $s$ (see @eq-equation13).
-Finally, the number of bins along the y-axis is computed based on the adjusted range ($r_2 + buffer_{y}$) and the horizontal spacing ($v$) (see @eq-equation14).
-
 $$
@@
 
-In this process, first involves the generation of centroids in the hexagonal grid.
-The hexagonal starting coordinates are defined.
-We already computed the number of bins along the x and y axes and have the horizontal and vertical spacing.
-By starting with hexagonal starting coordinates, iteratively compute the centroids by incorporating the number of bins and spacing along each axes.
-
-Once, we have the centroids for the hexagonal grid, the next step is to compute the hexgaonal coordinates. For example, as shown in Figure, the hexagonal coordinates can be computed.
+In this process, the first step involves generating centroids within the hexagonal grid.
+This begins by defining the starting coordinates for the hexagons.
+With the number of bins computed along the x and y axes, along with the horizontal and vertical spacing, the centroids are iteratively computed starting from the hexagonal starting coordinates.
+
+Once the centroids for the hexagonal grid are obtained, the next step is to compute the hexagonal coordinates.
+For example, if the centroid of a hexagonal bin is defined as $(C_x, C_y)$, the hexagonal coordinates can be computed using $d_x$ (see @eq-equation15) and $d_y$ (see @eq-equation16) (see @fig-hexcoord).
+
+$$
+d_x = \frac{h}{2}
+$$ {#eq-equation15}
+
+$$
+d_y = \frac{2* v * 1.15}{\sqrt{3}} 
+$$ {#eq-equation16}
+
+```{r}
+#| echo: false
+#| fig-pos: H
+#| fig-cap: "The hexagonal coordinates computed for the centroid $(C_x, C_y)$."
+#| label: fig-hexcoord
+#| out-width: 100%
+#| out-height: 100%
+knitr::include_graphics("figures/hex_coord.png")
+```
 
@@
 
-Once, we have the hexagonal bin centroids of the full grid, by using the 2D Euclidean distance find the nearest hexagonal bin for each NLDR point. Then assign that bin for the NLDR point.
+After obtaining the centroids of the hexagonal bins for the entire grid, the process of assigning NLDR points to hexagons involves determining the nearest hexagonal bin for each NLDR point using the 2D Euclidean distance.
+Then, the hexagonal ID of the nearest hexagon is assigned to the corresponding NLDR points.
 
@@
 ```
-
-<!--
-```{r}
-#| echo: false
-#| eval: false
-
-num_bins_x <- calculate_effective_x_bins(.data = UMAP_data, x = UMAP1,
-                                         cell_area = 1)
-
-shape_value <- calculate_effective_shape_value(.data = UMAP_data, 
-                                               x = UMAP1, y = UMAP2)
-
-## To extract bin centroids
-hexbin_data_object <-extract_hexbin_centroids(nldr_df = UMAP_data, num_bins = num_bins_x, shape_val = shape_value)
-
-df_bin_centroids <- hexbin_data_object$hexdf_data
-
-# ##########
-# 
-# ## Data set with all possible centroids in the hexagonal grid
-# 
-# full_centroid_df <- generate_full_grid_centroids(df_bin_centroids)
-# 
-# ## To map hexID to hexbin centroids in the full grid
-# 
-# vec1 <- stats::setNames(rep("", 2), c("x", "y"))  ## Define column names
-# 
-# full_grid_with_hexbin_id <- dplyr::bind_rows(vec1)[0, ]
-# full_grid_with_hexbin_id <- full_grid_with_hexbin_id |>
-#   dplyr::mutate_if(is.character, as.numeric)
-# 
-# for(i in 1:length(sort(unique(full_centroid_df$y)))){
-#   
-#   ## Filter the data set with specific y value
-#   specific_y_val_df <- full_centroid_df |>
-#     dplyr::filter(y == sort(unique(full_centroid_df$y))[i])
-#   
-#   ordered_x_df <- specific_y_val_df |>
-#     dplyr::arrange(x) 
-#   
-#   full_grid_with_hexbin_id <- dplyr::bind_rows(full_grid_with_hexbin_id, ordered_x_df)
-#   
-# }
-# 
-# 
-# full_grid_with_hexbin_id <- full_grid_with_hexbin_id |>
-#   dplyr::mutate(hexID = row_number())
-# 
-# full_grid_with_hexbin_id <- full_grid_with_hexbin_id |>
-#   dplyr::rename("c_x" = "x",
-#          "c_y" = "y") 
-# 
-# full_grid_with_hexbin_id <- dplyr::full_join(full_grid_with_hexbin_id, df_bin_centroids, by = c("hexID" = "hexID")) |>
-#   dplyr::select(-c(x, y)) #|> 
-#   #dplyr::mutate(counts = tidyr::replace_na(counts, 0))
-# 
-# full_grid_with_hexbin_id <- full_grid_with_hexbin_id |>
-#     dplyr::mutate(std_counts = counts/max(counts, na.rm = TRUE))
-# 
-# ## Generate all coordinates of hexagons
-# hex_grid <- full_hex_grid(full_centroid_df)
-# 
-# full_grid_with_polygon_id_df <- map_polygon_id(full_grid_with_hexbin_id, hex_grid)
-# 
-# full_grid_with_hexbin_id_rep <- full_grid_with_polygon_id_df |>
-#   dplyr::slice(rep(1:n(), each = 6)) |>
-#   dplyr::arrange(polygon_id)
-
-hex_full_count_df <- read_rds("data/s_curve/s_curve_hex_11.rds")
-
-##########
-
-min_std_cell_threshold <- 0.05
-
-df_bin_centroids_all <- hexbin_data_object$hexdf_data ## All the centroids without removing low-density hexagons
-
-df_bin_centroids <- df_bin_centroids |>
-  dplyr::mutate(stand_cell_count = counts/max(counts)) |>
-  dplyr::filter(stand_cell_count > min_std_cell_threshold)
-
-UMAP_data_with_hb_id <- UMAP_data |> 
-  dplyr::mutate(hb_id = hexbin_data_object$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all <- dplyr::bind_cols(training_data |> dplyr::select(-ID), UMAP_data_with_hb_id)
-
-## Averaged on high-D
-df_bin <- avg_highD_data(.data = df_all)
-
-## Triangulate bin centroids
-tr1_object <- triangulate_bin_centroids(df_bin_centroids, x, y)
-tr_from_to_df <- generate_edge_info(triangular_object = tr1_object)
-
-## Compute 2D distances
-distance <- cal_2d_dist(.data = tr_from_to_df)
-
-## To find the benchmark value
-benchmark <- find_benchmark_value(.data = distance, distance_col = distance)
-
-trimesh <- ggplot(df_bin_centroids, aes(x = x, y = y)) + 
-  geom_segment(data = tr_from_to_df, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
-  geom_point(size = 2, colour = "#33a02c") +
-  coord_equal()
-
-# ggplot(df_bin_centroids, aes(x = x, y = y)) + 
-# geom_point(size = 1, colour = "#33a02c") + 
-# geom_trimesh() + 
-# coord_equal()
-
-trimesh <- trimesh +
-  #ggtitle("(a)") + 
-  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  theme_light() +
-  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
-  ) 
-# theme(axis.text = element_text(size = 5),
-#       axis.title = element_text(size = 7))
-
-trimesh_gr <- colour_long_edges(.data = distance, benchmark_value = benchmark, 
-                                triangular_object = tr1_object, distance_col = distance)
-
-trimesh_gr <- trimesh_gr + 
-  geom_point(size = 2, colour = "#33a02c") + 
-  #ggtitle("(b)") + 
-  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  theme_light() +
-  #coord_equal() +
-  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
-  )
-
-trimesh_removed <- remove_long_edges(.data = distance, benchmark_value = benchmark, 
-                                     triangular_object = tr1_object, distance_col = distance)
-
-trimesh_removed <- trimesh_removed + 
-  geom_point(size = 2, colour = "#33a02c") + 
-  #ggtitle("(b)") + 
-  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  theme_light() +
-  #coord_equal() +
-  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
-  )
-
-tour1 <- show_langevitour(df_all, df_bin, df_bin_centroids, benchmark_value = benchmark, distance = distance, distance_col = distance)
-```
-
-```{r}
-#| echo: false
-#| eval: false
-
-## To plot the distribution of distance
-plot_dist <- function(distance_df){
-  distance_df$group <- "1"
-  dist_plot <- ggplot(distance_df, aes(x = group, y = distance)) +
-    geom_quasirandom()+
-    ylim(0, max(unlist(distance_df$distance))+ 0.5) + coord_flip()
-  return(dist_plot)
-}
-```
-
-```{r}
-#| echo: false
-#| eval: false
-
-distance_plot <- plot_dist(distance) +
-  #ggtitle("(b)" ) + 
-  ylab(expression(d^{(2)})) +
-  theme(axis.text = element_text(size = 5),
-        axis.title = element_text(size = 12))
-```
-
-```{r}
-#| echo: false
-#| warning: false
-#| eval: false
-
-data <- read_rds("data/s_curve/s_curve.rds")
-training_data <- read_rds("data/s_curve/s_curve_training.rds")
-UMAP_data <- read_rds(file = "data/s_curve/s_curve_umap.rds")
-
-num_bins_x <- 2
-
-shape_value <- calculate_effective_shape_value(.data = UMAP_data, 
-                                               x = UMAP1, y = UMAP2)
-
-## To extract bin centroids
-hexbin_data_object <-extract_hexbin_centroids(nldr_df = UMAP_data, num_bins = num_bins_x, shape_val = shape_value)
-
-df_bin_centroids <- hexbin_data_object$hexdf_data
-
-# ## Data set with all possible centroids in the hexagonal grid
-# 
-# full_centroid_df <- generate_full_grid_centroids(df_bin_centroids)
-# 
-# ## Generate all coordinates of hexagons
-# hex_grid <- full_hex_grid(full_centroid_df)
-# 
-# hex_full_count_df <- generate_full_grid_info(df_bin_centroids)
-# 
-# write_rds(hex_full_count_df, "data/s_curve/s_curve_hex_2.rds")
-
-min_std_cell_threshold <- 0.05
-
-df_bin_centroids_all <- hexbin_data_object$hexdf_data ## All the centroids without removing low-density hexagons
-
-hex_full_count_df <- read_rds("data/s_curve/s_curve_hex_2.rds")
-
-df_bin_centroids <- df_bin_centroids |>
-  dplyr::mutate(stand_cell_count = counts/max(counts)) |>
-  dplyr::filter(stand_cell_count > min_std_cell_threshold)
-
-UMAP_data_with_hb_id <- UMAP_data |> 
-  dplyr::mutate(hb_id = hexbin_data_object$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all <- dplyr::bind_cols(training_data |> dplyr::select(-ID), UMAP_data_with_hb_id)
-
-## Averaged on high-D
-df_bin <- avg_highD_data(.data = df_all)
-
-## Triangulate bin centroids
-tr1_object <- triangulate_bin_centroids(df_bin_centroids, x, y)
-tr_from_to_df <- generate_edge_info(triangular_object = tr1_object)
-
-# Create the tibble with x and y coordinates
-tr_df <- tibble::tibble(x = tr1_object$x, y = tr1_object$y)
-
-## Compute 2D distances
-distance <- cal_2d_dist(.data = tr_from_to_df)
-
-## Map hb_id
-
-df_bin_centroids_n <- df_bin_centroids |>
-  select(hexID) |>
-  mutate(ID = 1:NROW(df_bin_centroids))
-
-tr_from_to_df_n <- left_join(tr_from_to_df, df_bin_centroids_n, by = c("from" = "ID")) |>
-  rename("from_hb_id" = "hexID")
-
-tr_from_to_df_n <- left_join(tr_from_to_df_n, df_bin_centroids_n, by = c("to" = "ID")) |>
-  rename("to_hb_id" = "hexID")
-
-hb_id_selected <- tr_from_to_df_n |>
-  dplyr::filter((from == 2) & (to == 3))
-
-# benchmark_value <- 19.4
-
-# Filter and label small and long edges
-distance_df_small_edges <- distance |>
-  dplyr::filter((from != 2) | (to != 3)) |>
-  dplyr::mutate(type = "small_edges")
-
-distance_df_long_edges <- distance |>
-  dplyr::filter((from == 2) & (to == 3)) |>
-  dplyr::mutate(type = "long_edges")
-
-# Combine small and long edges
-distance_edges <- dplyr::bind_rows(distance_df_small_edges, distance_df_long_edges)
-
-tr_from_to_df_coord_with_group <- inner_join(tr_from_to_df, distance_edges, by = c("from" = "from", "to" = "to"))
-
-UMAP_data_with_hb_id_selected <- UMAP_data_with_hb_id |> dplyr::filter(hb_id %in% c(hb_id_selected$from_hb_id, hb_id_selected$to_hb_id)) |>
-  dplyr::mutate(select_point = if_else(hb_id == hb_id_selected$from_hb_id, "data1", "data2"))
-
-UMAP_data_with_hb_id_selected_no <- UMAP_data_with_hb_id |> dplyr::filter(!(hb_id %in% c(hb_id_selected$from_hb_id, hb_id_selected$to_hb_id)))
-
-bin_coord_1 <- tr_df |>
-  dplyr::filter((x == hb_id_selected$x_from)) |>
-  dplyr::filter((y == hb_id_selected$y_from))
-
-# bin_coord_1 <- tr_df |>
-#   head(1)
-
-bin_coord_2 <- tr_df |>
-  dplyr::filter((x == hb_id_selected$x_to)) |>
-  dplyr::filter((y == hb_id_selected$y_to))
-
-# bin_coord_2 <- tr_df |>
-#   dplyr::filter(row_number() == 2)
-
-bin_coord <- dplyr::bind_rows(bin_coord_1, bin_coord_2)
-
-a1 <- ggplot(bin_coord, aes(x = x, y = y)) +
-  geom_segment(
-    aes(x = x_from, y = y_from, xend = x_to, yend = y_to, colour = type),
-    data = tr_from_to_df_coord_with_group) +
-  geom_point() +
-  coord_equal() +
-  geom_point(data = UMAP_data_with_hb_id_selected, aes(x = UMAP1, y = UMAP2, colour = select_point), alpha = 0.5, size = 1) +
-  scale_color_manual(values=c("#e6550d", "#000000", "#e31a1c", "#f6e8c3")) +
-  geom_point(size = 3, colour = "#33a02c") + 
-  #ggtitle("(b)") + 
-  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  theme_light() +
-  #coord_equal() +
-  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
-  )
-
-a2 <- ggplot(data = hex_full_count_df, aes(x = x, y = y)) + 
-  geom_polygon(color = "#deebf7", aes(group = polygon_id), fill = "#ffffff") +
-  geom_segment(
-    aes(x = x_from, y = y_from, xend = x_to, yend = y_to), colour = "#f6e8c3",
-    data = tr_from_to_df_coord_with_group
-  ) +
-  geom_point(data = UMAP_data_with_hb_id_selected, aes(x = UMAP1, y = UMAP2, colour = select_point), size = 1, alpha = 0.5) +
-  scale_color_manual(values=c("#e6550d", "#000000")) +
-  geom_point(data = df_bin_centroids_all |> dplyr::filter(hexID %in% c(hb_id_selected$from_hb_id, hb_id_selected$to_hb_id)), aes(x = x, y = y), colour = "#33a02c", size = 3) +
-    geom_point(data = UMAP_data_with_hb_id_selected_no, aes(x = UMAP1, y = UMAP2), colour = "#6a3d9a", size = 1, alpha = 0.5) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  coord_cartesian(xlim =c(-5, 8), ylim = c(-10, 10)) +
-  theme_light() +
-  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
-  )
-
-## Script to make coloring data points with a line
-
-df1 <- df_all |> 
-  dplyr::filter(hb_id == as.character(hb_id_selected$from_hb_id)) |>
-  dplyr::select(tidyselect::starts_with("x")) |>
-  dplyr::mutate(type = "data1") ## original dataset
-
-df2 <- df_all |> 
-  dplyr::filter(hb_id == as.character(hb_id_selected$to_hb_id)) |>
-  dplyr::select(tidyselect::starts_with("x")) |>
-  dplyr::mutate(type = "data2") ## original dataset
-
-df3 <- df_all |> 
-  dplyr::filter(!(hb_id %in% c(as.character(hb_id_selected$from_hb_id), as.character(hb_id_selected$to_hb_id)))) |>
-  dplyr::select(tidyselect::starts_with("x")) |>
-  dplyr::mutate(type = "data") ## original dataset
-
-df_b <- df_bin |> 
-  dplyr::filter(hb_id %in% df_bin_centroids$hexID) |>
-  dplyr::filter(hb_id %in% c(as.character(hb_id_selected$from_hb_id), as.character(hb_id_selected$to_hb_id))) |>
-  dplyr::select(-hb_id) |>
-  dplyr::mutate(type = "model") ## Data with summarized mean
-
-df_exe <- dplyr::bind_rows(df_b, df1, df2, df3)
-
-lg_high <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)], group = df_exe$type, pointSize = 3, levelColors = c("#6a3d9a", "#e6550d", "#000000", "#33a02c"), lineFrom = 1, lineTo = 2, lineColors = "#e31a1c")
-
-```
-
-```{r}
-#| eval: false
-#| warning: false
-#| echo: false
-#| label: fig-wkhighD1
-#| fig-cap: How the 2D model lift into high dimensions?
-(a) visualize the points and the hexagonal bin centroids related 2nd and 15th hexagons, (b) visualization of the edge connected the 2nd and 15th hexagons (colored in red) in the triangular mesh.
-A video of tour animation is available at <https://youtu.be/ffWFVK1UTIc>.
-
-a2 + a1 +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(ncol = 3) &
-  theme(legend.position='none', plot.tag = element_text(size = 8))
-```
--->
 

### 2024-03-25 paper.qmd
Added 5 Gaussian Cluster 2D models

@@
         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+  annotate(geom = 'text', label = 'c', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
 
@@
 ## Decide by looking at MSE plot
-num_bins_x_tsne_gau <- 60
-num_bins_y_tsne_gau <- 79
-hex_size_tsne_gau <- 0.01
+num_bins_x_umap_gau <- 60
+num_bins_y_umap_gau <- 79
+hex_size_umap_gau <- 0.01
 ### non-empty 229
 
-hb_obj_tsne_gau <- hex_binning(data = umap_gau_scaled, x = "UMAP1", 
-                      y = "UMAP2", num_bins_x = num_bins_x_tsne_gau, 
-                      num_bins_y = num_bins_y_tsne_gau, x_start = NA, y_start = NA, 
-                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_tsne_gau, col_start = "UMAP")
+hb_obj_umap_gau <- hex_binning(data = umap_gau_scaled, x = "UMAP1", 
+                               y = "UMAP2", num_bins_x = num_bins_x_umap_gau, 
+                               num_bins_y = num_bins_y_umap_gau, x_start = NA, y_start = NA, 
+                               buffer_x = NA, buffer_y = NA, hex_size = hex_size_umap_gau, col_start = "UMAP")
 
 ## Data set with all possible centroids in the hexagonal grid
-all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$centroids))
+all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_umap_gau$centroids))
 
 ## Generate all coordinates of hexagons
-hex_grid <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$hex_poly))
+hex_grid <- as.data.frame(do.call(cbind, hb_obj_umap_gau$hex_poly))
 
 ## To obtain the standardise counts within hexbins
-counts_df <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$std_cts))
+counts_df <- as.data.frame(do.call(cbind, hb_obj_umap_gau$std_cts))
 df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
@@
 
-tsne_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$data_hb_id))
-
-model_object_tsne_gau <- fit_highd_model( training_data = training_data_gau, 
+umap_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_umap_gau$data_hb_id))
+
+model_object_umap_gau <- fit_highd_model( training_data = training_data_gau, 
                                           nldr_df_with_id = umap_gau_scaled, 
                                           x = "UMAP1", y = "UMAP2", 
-                                 num_bins_x = num_bins_x_tsne_gau, 
-                                 num_bins_y = num_bins_y_tsne_gau, 
+                                          num_bins_x = num_bins_x_umap_gau, 
+                                          num_bins_y = num_bins_y_umap_gau, 
                                           x_start = NA, y_start = NA, 
                                           buffer_x = NA, buffer_y = NA, 
-                                 hex_size = hex_size_tsne_gau,
+                                          hex_size = hex_size_umap_gau,
                                           is_rm_lwd_hex = FALSE, 
@@
 
-df_bin_centroids_tsne_gau <- model_object_tsne_gau$df_bin_centroids
-df_bin_tsne_gau <- model_object_tsne_gau$df_bin
+df_bin_centroids_umap_gau <- model_object_umap_gau$df_bin_centroids
+df_bin_umap_gau <- model_object_umap_gau$df_bin
 
 ## Triangulate bin centroids
-tr1_object_tsne_gau <- tri_bin_centroids(df_bin_centroids_tsne_gau, x = "c_x", y = "c_y")
-tr_from_to_df_tsne_gau <- gen_edges(tri_object = tr1_object_tsne_gau)
+tr1_object_umap_gau <- tri_bin_centroids(df_bin_centroids_umap_gau, x = "c_x", y = "c_y")
+tr_from_to_df_umap_gau <- gen_edges(tri_object = tr1_object_umap_gau)
 
 ## Compute 2D distances
-distance_tsne_gau <- cal_2d_dist(tr_coord_df = tr_from_to_df_tsne_gau, 
+distance_umap_gau <- cal_2d_dist(tr_coord_df = tr_from_to_df_umap_gau, 
                                  start_x = "x_from", start_y = "y_from", 
@@
 
-# distance_plot <- plot_dist(distance_tsne_gau) +
+# distance_plot <- plot_dist(distance_umap_gau) +
 #   ylab(expression(d^{(2)})) +
@@
 ## To find the benchmark value
-benchmark_tsne_gau <- find_lg_benchmark(distance_edges = distance_tsne_gau, distance_col = "distance")
+benchmark_umap_gau <- find_lg_benchmark(distance_edges = distance_umap_gau, distance_col = "distance")
 
 # ggplot() +
-# geom_trimesh(data = df_bin_centroids_tsne_gau, mapping = aes(x = c_x, y = c_y))
-
-trimesh_removed_tsne_gau <- vis_rmlg_mesh(distance_edges = distance_tsne_gau, benchmark_value = benchmark_tsne_gau, tr_coord_df = tr_from_to_df_tsne_gau, distance_col = "distance")
-
-trimesh_removed_tsne_gau <- trimesh_removed_tsne_gau +
+# geom_trimesh(data = df_bin_centroids_umap_gau, mapping = aes(x = c_x, y = c_y))
+
+trimesh_removed_umap_gau <- vis_rmlg_mesh(distance_edges = distance_umap_gau, benchmark_value = benchmark_umap_gau, tr_coord_df = tr_from_to_df_umap_gau, distance_col = "distance")
+
+trimesh_removed_umap_gau <- trimesh_removed_umap_gau +
   theme_linedraw() +
@@
         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-df_all_tsne_gau <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID), tsne_data_with_hb_id_gau)
-
-# show_langevitour(df = df_all_tsne_gau, df_b = df_bin_tsne_gau, df_b_with_center_data = df_bin_centroids_tsne_gau, col_start = "x", distance_df = distance_tsne_gau, distance_col = "distance", benchmark_value = benchmark_tsne_gau)
-
-  # find_non_empty_bins(data = umap_gau_scaled, x = "UMAP1", y = "UMAP2",
-  #                   non_empty_bins = 198, x_start = NA, y_start = NA,
-  #                   buffer_x = NA, buffer_y = NA, hex_size = 0.03,
-  #                   col_start = "UMAP")
-
-# hex_size_vec <- seq(0.02, 2, by = 0.01)
-# 
-# for (i in 1:length(hex_size_vec)) {
-# 
-#   find_non_empty_bins(data = umap_gau_scaled, x = "UMAP1", y = "UMAP2",
-#                     non_empty_bins = 198, x_start = NA, y_start = NA,
-#                     buffer_x = NA, buffer_y = NA, hex_size = hex_size_vec[i],
-#                     col_start = "UMAP")
-# 
-# 
-# }
-
+  annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+df_all_umap_gau <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID), umap_data_with_hb_id_gau)
+
+# show_langevitour(df = df_all_umap_gau, df_b = df_bin_umap_gau, df_b_with_center_data = df_bin_centroids_umap_gau, col_start = "x", distance_df = distance_umap_gau, distance_col = "distance", benchmark_value = benchmark_umap_gau)
 ```
@@
 
-phate_gau_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tibble::as_tibble(PHATE_data_gau), x = "PHATE1", y = "PHATE2"))) |>
+phate_gau_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tibble::as_tibble(PHATE_data_gau), x = "PHATE1", y = "PHATE2", hex_ratio = 50/sqrt(3)))) |>
   dplyr::rename(c("PHATE1" = "scaled_PHATE1",
@@
 ## Decide by looking at MSE plot
-num_bins_x_tsne_gau <- 60
-num_bins_y_tsne_gau <- 156
-hex_size_tsne_gau <- 0.01
-### non-empty 179
-
-hb_obj_tsne_gau <- hex_binning(data = phate_gau_scaled, x = "PHATE1", 
-                      y = "PHATE2", num_bins_x = num_bins_x_tsne_gau, 
-                      num_bins_y = num_bins_y_tsne_gau, x_start = NA, y_start = NA, 
-                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_tsne_gau, col_start = "PHATE")
+num_bins_x_phate_gau <- 60
+num_bins_y_phate_gau <- 1927
+hex_size_phate_gau <- 0.01
+### non-empty 187
+
+hb_obj_phate_gau <- hex_binning(data = phate_gau_scaled, x = "PHATE1", 
+                               y = "PHATE2", num_bins_x = num_bins_x_phate_gau, 
+                               num_bins_y = num_bins_y_phate_gau, x_start = NA, y_start = NA, 
+                               buffer_x = NA, buffer_y = NA, hex_size = hex_size_phate_gau, col_start = "PHATE")
 
 ## Data set with all possible centroids in the hexagonal grid
-all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$centroids))
+all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_phate_gau$centroids))
 
 ## Generate all coordinates of hexagons
-hex_grid <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$hex_poly))
+hex_grid <- as.data.frame(do.call(cbind, hb_obj_phate_gau$hex_poly))
 
 ## To obtain the standardise counts within hexbins
-counts_df <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$std_cts))
+counts_df <- as.data.frame(do.call(cbind, hb_obj_phate_gau$std_cts))
 df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
@@
 
-tsne_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$data_hb_id))
-
-model_object_tsne_gau <- fit_highd_model( training_data = training_data_gau, 
+phate_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_phate_gau$data_hb_id))
+
+model_object_phate_gau <- fit_highd_model( training_data = training_data_gau, 
                                           nldr_df_with_id = phate_gau_scaled, 
                                           x = "PHATE1", y = "PHATE2", 
-                                 num_bins_x = num_bins_x_tsne_gau, 
-                                 num_bins_y = num_bins_y_tsne_gau, 
+                                          num_bins_x = num_bins_x_phate_gau, 
+                                          num_bins_y = num_bins_y_phate_gau, 
                                           x_start = NA, y_start = NA, 
                                           buffer_x = NA, buffer_y = NA, 
-                                 hex_size = hex_size_tsne_gau,
+                                          hex_size = hex_size_phate_gau,
                                           is_rm_lwd_hex = FALSE, 
@@
 
-df_bin_centroids_tsne_gau <- model_object_tsne_gau$df_bin_centroids
-df_bin_tsne_gau <- model_object_tsne_gau$df_bin
+df_bin_centroids_phate_gau <- model_object_phate_gau$df_bin_centroids
+df_bin_phate_gau <- model_object_phate_gau$df_bin
 
 ## Triangulate bin centroids
-tr1_object_tsne_gau <- tri_bin_centroids(df_bin_centroids_tsne_gau, x = "c_x", y = "c_y")
-tr_from_to_df_tsne_gau <- gen_edges(tri_object = tr1_object_tsne_gau)
+tr1_object_phate_gau <- tri_bin_centroids(df_bin_centroids_phate_gau, x = "c_x", y = "c_y")
+tr_from_to_df_phate_gau <- gen_edges(tri_object = tr1_object_phate_gau)
 
 ## Compute 2D distances
-distance_tsne_gau <- cal_2d_dist(tr_coord_df = tr_from_to_df_tsne_gau, 
+distance_phate_gau <- cal_2d_dist(tr_coord_df = tr_from_to_df_phate_gau, 
                                  start_x = "x_from", start_y = "y_from", 
@@
 
-# distance_plot <- plot_dist(distance_tsne_gau) +
+# distance_plot <- plot_dist(distance_phate_gau) +
 #   ylab(expression(d^{(2)})) +
@@
 ## To find the benchmark value
-benchmark_tsne_gau <- find_lg_benchmark(distance_edges = distance_tsne_gau, distance_col = "distance")
+benchmark_phate_gau <- find_lg_benchmark(distance_edges = distance_phate_gau, distance_col = "distance")
 
 # ggplot() +
-# geom_trimesh(data = df_bin_centroids_tsne_gau, mapping = aes(x = c_x, y = c_y))
-
-trimesh_removed_tsne_gau <- vis_rmlg_mesh(distance_edges = distance_tsne_gau, benchmark_value = benchmark_tsne_gau, tr_coord_df = tr_from_to_df_tsne_gau, distance_col = "distance")
-
-trimesh_removed_tsne_gau <- trimesh_removed_tsne_gau +
+# geom_trimesh(data = df_bin_centroids_phate_gau, mapping = aes(x = c_x, y = c_y))
+
+trimesh_removed_phate_gau <- vis_rmlg_mesh(distance_edges = distance_phate_gau, benchmark_value = benchmark_phate_gau, tr_coord_df = tr_from_to_df_phate_gau, distance_col = "distance")
+
+trimesh_removed_phate_gau <- trimesh_removed_phate_gau +
   theme_linedraw() +
@@
         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-df_all_tsne_gau <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID), tsne_data_with_hb_id_gau)
-
-# show_langevitour(df = df_all_tsne_gau, df_b = df_bin_tsne_gau, df_b_with_center_data = df_bin_centroids_tsne_gau, col_start = "x", distance_df = distance_tsne_gau, distance_col = "distance", benchmark_value = benchmark_tsne_gau)
+  annotate(geom = 'text', label = 'c', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+df_all_phate_gau <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID), phate_data_with_hb_id_gau)
+
+# show_langevitour(df = df_all_phate_gau, df_b = df_bin_phate_gau, df_b_with_center_data = df_bin_centroids_phate_gau, col_start = "x", distance_df = distance_phate_gau, distance_col = "distance", benchmark_value = benchmark_phate_gau)
 
@@
 ## Decide by looking at MSE plot
-num_bins_x_tsne_gau <- 46
-num_bins_y_tsne_gau <- 61
-hex_size_tsne_gau <- 0.013
-### non-empty 179
-
-hb_obj_tsne_gau <- hex_binning(data = trimap_gau_scaled, x = "TriMAP1", 
-                      y = "TriMAP2", num_bins_x = num_bins_x_tsne_gau, 
-                      num_bins_y = num_bins_y_tsne_gau, x_start = NA, y_start = NA, 
-                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_tsne_gau, col_start = "TriMAP")
+num_bins_x_trimap_gau <- 46
+num_bins_y_trimap_gau <- 61
+hex_size_trimap_gau <- 0.013
+### non-empty 173
+
+hb_obj_trimap_gau <- hex_binning(data = trimap_gau_scaled, x = "TriMAP1", 
+                               y = "TriMAP2", num_bins_x = num_bins_x_trimap_gau, 
+                               num_bins_y = num_bins_y_trimap_gau, x_start = NA, y_start = NA, 
+                               buffer_x = NA, buffer_y = NA, hex_size = hex_size_trimap_gau, col_start = "TriMAP")
 
 ## Data set with all possible centroids in the hexagonal grid
-all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$centroids))
+all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_trimap_gau$centroids))
 
 ## Generate all coordinates of hexagons
-hex_grid <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$hex_poly))
+hex_grid <- as.data.frame(do.call(cbind, hb_obj_trimap_gau$hex_poly))
 
 ## To obtain the standardise counts within hexbins
-counts_df <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$std_cts))
+counts_df <- as.data.frame(do.call(cbind, hb_obj_trimap_gau$std_cts))
 df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
@@
 
-tsne_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$data_hb_id))
-
-model_object_tsne_gau <- fit_highd_model( training_data = training_data_gau, 
+trimap_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_trimap_gau$data_hb_id))
+
+model_object_trimap_gau <- fit_highd_model( training_data = training_data_gau, 
                                           nldr_df_with_id = trimap_gau_scaled, 
                                           x = "TriMAP1", y = "TriMAP2", 
-                                 num_bins_x = num_bins_x_tsne_gau, 
-                                 num_bins_y = num_bins_y_tsne_gau, 
+                                          num_bins_x = num_bins_x_trimap_gau, 
+                                          num_bins_y = num_bins_y_trimap_gau, 
                                           x_start = NA, y_start = NA, 
                                           buffer_x = NA, buffer_y = NA, 
-                                 hex_size = hex_size_tsne_gau,
+                                          hex_size = hex_size_trimap_gau,
                                           is_rm_lwd_hex = FALSE, 
@@
 
-df_bin_centroids_tsne_gau <- model_object_tsne_gau$df_bin_centroids
-df_bin_tsne_gau <- model_object_tsne_gau$df_bin
+df_bin_centroids_trimap_gau <- model_object_trimap_gau$df_bin_centroids
+df_bin_trimap_gau <- model_object_trimap_gau$df_bin
 
 ## Triangulate bin centroids
-tr1_object_tsne_gau <- tri_bin_centroids(df_bin_centroids_tsne_gau, x = "c_x", y = "c_y")
-tr_from_to_df_tsne_gau <- gen_edges(tri_object = tr1_object_tsne_gau)
+tr1_object_trimap_gau <- tri_bin_centroids(df_bin_centroids_trimap_gau, x = "c_x", y = "c_y")
+tr_from_to_df_trimap_gau <- gen_edges(tri_object = tr1_object_trimap_gau)
 
 ## Compute 2D distances
-distance_tsne_gau <- cal_2d_dist(tr_coord_df = tr_from_to_df_tsne_gau, 
+distance_trimap_gau <- cal_2d_dist(tr_coord_df = tr_from_to_df_trimap_gau, 
                                  start_x = "x_from", start_y = "y_from", 
@@
 
-# distance_plot <- plot_dist(distance_tsne_gau) +
+# distance_plot <- plot_dist(distance_trimap_gau) +
 #   ylab(expression(d^{(2)})) +
@@
 ## To find the benchmark value
-benchmark_tsne_gau <- find_lg_benchmark(distance_edges = distance_tsne_gau, distance_col = "distance")
+benchmark_trimap_gau <- find_lg_benchmark(distance_edges = distance_trimap_gau, distance_col = "distance")
+benchmark_trimap_gau <- 0.2
 
 # ggplot() +
-# geom_trimesh(data = df_bin_centroids_tsne_gau, mapping = aes(x = c_x, y = c_y))
-
-trimesh_removed_tsne_gau <- vis_rmlg_mesh(distance_edges = distance_tsne_gau, benchmark_value = benchmark_tsne_gau, tr_coord_df = tr_from_to_df_tsne_gau, distance_col = "distance")
-
-trimesh_removed_tsne_gau <- trimesh_removed_tsne_gau +
+# geom_trimesh(data = df_bin_centroids_trimap_gau, mapping = aes(x = c_x, y = c_y))
+
+trimesh_removed_trimap_gau <- vis_rmlg_mesh(distance_edges = distance_trimap_gau, benchmark_value = benchmark_trimap_gau, tr_coord_df = tr_from_to_df_trimap_gau, distance_col = "distance")
+
+trimesh_removed_trimap_gau <- trimesh_removed_trimap_gau +
   theme_linedraw() +
@@
         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-df_all_tsne_gau <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID), tsne_data_with_hb_id_gau)
-
-# show_langevitour(df = df_all_tsne_gau, df_b = df_bin_tsne_gau, df_b_with_center_data = df_bin_centroids_tsne_gau, col_start = "x", distance_df = distance_tsne_gau, distance_col = "distance", benchmark_value = benchmark_tsne_gau)
+  annotate(geom = 'text', label = 'd', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+df_all_trimap_gau <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID), trimap_data_with_hb_id_gau)
+
+# show_langevitour(df = df_all_trimap_gau, df_b = df_bin_trimap_gau, df_b_with_center_data = df_bin_centroids_trimap_gau, col_start = "x", distance_df = distance_trimap_gau, distance_col = "distance", benchmark_value = benchmark_trimap_gau)
 
@@
 ## Decide by looking at MSE plot
-num_bins_x_tsne_gau <- 31
-num_bins_y_tsne_gau <- 40
-hex_size_tsne_gau <- 0.02
+num_bins_x_pacmap_gau <- 31
+num_bins_y_pacmap_gau <- 40
+hex_size_pacmap_gau <- 0.02
 ### non-empty 179
 
-hb_obj_tsne_gau <- hex_binning(data = pacmap_gau_scaled, x = "PaCMAP1", 
-                      y = "PaCMAP2", num_bins_x = num_bins_x_tsne_gau, 
-                      num_bins_y = num_bins_y_tsne_gau, x_start = NA, y_start = NA, 
-                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_tsne_gau, col_start = "PaCMAP")
+hb_obj_pacmap_gau <- hex_binning(data = pacmap_gau_scaled, x = "PaCMAP1", 
+                               y = "PaCMAP2", num_bins_x = num_bins_x_pacmap_gau, 
+                               num_bins_y = num_bins_y_pacmap_gau, x_start = NA, y_start = NA, 
+                               buffer_x = NA, buffer_y = NA, hex_size = hex_size_pacmap_gau, col_start = "PaCMAP")
 
 ## Data set with all possible centroids in the hexagonal grid
-all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$centroids))
+all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_pacmap_gau$centroids))
 
 ## Generate all coordinates of hexagons
-hex_grid <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$hex_poly))
+hex_grid <- as.data.frame(do.call(cbind, hb_obj_pacmap_gau$hex_poly))
 
 ## To obtain the standardise counts within hexbins
-counts_df <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$std_cts))
+counts_df <- as.data.frame(do.call(cbind, hb_obj_pacmap_gau$std_cts))
 df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
@@
 
-tsne_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$data_hb_id))
-
-model_object_tsne_gau <- fit_highd_model( training_data = training_data_gau, 
+pacmap_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_pacmap_gau$data_hb_id))
+
+model_object_pacmap_gau <- fit_highd_model( training_data = training_data_gau, 
                                           nldr_df_with_id = pacmap_gau_scaled, 
                                           x = "PaCMAP1", y = "PaCMAP2", 
-                                 num_bins_x = num_bins_x_tsne_gau, 
-                                 num_bins_y = num_bins_y_tsne_gau, 
+                                          num_bins_x = num_bins_x_pacmap_gau, 
+                                          num_bins_y = num_bins_y_pacmap_gau, 
                                           x_start = NA, y_start = NA, 
                                           buffer_x = NA, buffer_y = NA, 
-                                 hex_size = hex_size_tsne_gau,
+                                          hex_size = hex_size_pacmap_gau,
                                           is_rm_lwd_hex = FALSE, 
@@
 
-df_bin_centroids_tsne_gau <- model_object_tsne_gau$df_bin_centroids
-df_bin_tsne_gau <- model_object_tsne_gau$df_bin
+df_bin_centroids_pacmap_gau <- model_object_pacmap_gau$df_bin_centroids
+df_bin_pacmap_gau <- model_object_pacmap_gau$df_bin
 
 ## Triangulate bin centroids
-tr1_object_tsne_gau <- tri_bin_centroids(df_bin_centroids_tsne_gau, x = "c_x", y = "c_y")
-tr_from_to_df_tsne_gau <- gen_edges(tri_object = tr1_object_tsne_gau)
+tr1_object_pacmap_gau <- tri_bin_centroids(df_bin_centroids_pacmap_gau, x = "c_x", y = "c_y")
+tr_from_to_df_pacmap_gau <- gen_edges(tri_object = tr1_object_pacmap_gau)
 
 ## Compute 2D distances
-distance_tsne_gau <- cal_2d_dist(tr_coord_df = tr_from_to_df_tsne_gau, 
+distance_pacmap_gau <- cal_2d_dist(tr_coord_df = tr_from_to_df_pacmap_gau, 
                                  start_x = "x_from", start_y = "y_from", 
@@
 
-# distance_plot <- plot_dist(distance_tsne_gau) +
+# distance_plot <- plot_dist(distance_pacmap_gau) +
 #   ylab(expression(d^{(2)})) +
@@
 ## To find the benchmark value
-benchmark_tsne_gau <- find_lg_benchmark(distance_edges = distance_tsne_gau, distance_col = "distance")
+benchmark_pacmap_gau <- find_lg_benchmark(distance_edges = distance_pacmap_gau, distance_col = "distance")
 
 # ggplot() +
-# geom_trimesh(data = df_bin_centroids_tsne_gau, mapping = aes(x = c_x, y = c_y))
-
-trimesh_removed_tsne_gau <- vis_rmlg_mesh(distance_edges = distance_tsne_gau, benchmark_value = benchmark_tsne_gau, tr_coord_df = tr_from_to_df_tsne_gau, distance_col = "distance")
-
-trimesh_removed_tsne_gau <- trimesh_removed_tsne_gau +
+# geom_trimesh(data = df_bin_centroids_pacmap_gau, mapping = aes(x = c_x, y = c_y))
+
+trimesh_removed_pacmap_gau <- vis_rmlg_mesh(distance_edges = distance_pacmap_gau, benchmark_value = benchmark_pacmap_gau, tr_coord_df = tr_from_to_df_pacmap_gau, distance_col = "distance")
+
+trimesh_removed_pacmap_gau <- trimesh_removed_pacmap_gau +
   theme_linedraw() +
@@
         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-df_all_tsne_gau <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID), tsne_data_with_hb_id_gau)
-
-# show_langevitour(df = df_all_tsne_gau, df_b = df_bin_tsne_gau, df_b_with_center_data = df_bin_centroids_tsne_gau, col_start = "x", distance_df = distance_tsne_gau, distance_col = "distance", benchmark_value = benchmark_tsne_gau)
-
-```
-
-<!---
-```{r}
-#| warning: false
-#| echo: false
-#| eval: false
-
-### tSNE
-
-# num_bins_tsne_gau <- calculate_effective_x_bins(.data = tSNE_data_gau, x = tSNE1,
-#                                                 cell_area = 1) #56
-num_bins_tsne_gau <- 16
-
-shape_val_tsne_gau <- calculate_effective_shape_value(.data = tSNE_data_gau,
-                                                      x = tSNE1, y = tSNE2)
-## To extract bin centroids
-hexbin_data_object_tsne_gau <- extract_hexbin_centroids(nldr_df = tSNE_data_gau, num_bins = num_bins_tsne_gau, shape_val = shape_val_tsne_gau, x = tSNE1, y = tSNE2)
-
-df_bin_centroids_tsne_gau <- hexbin_data_object_tsne_gau$hexdf_data
-
-num_non_empty_bins_gau <- df_bin_centroids_tsne_gau$hexID |> length()
-
-tSNE_data_with_hb_id_gau <- tSNE_data_gau |>
-  dplyr::mutate(hb_id = hexbin_data_object_tsne_gau$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_tsne_gau <- dplyr::bind_cols(training_data_1 |> dplyr::select(-ID), tSNE_data_with_hb_id_gau)
-
-## Averaged on high-D
-df_bin_tsne_gau <- avg_highD_data(.data = df_all_tsne_gau)
-
-## Triangulate bin centroids
-tr1_object_tsne_gau <- triangulate_bin_centroids(df_bin_centroids_tsne_gau, x, y)
-tr_from_to_df_tsne_gau <- generate_edge_info(triangular_object = tr1_object_tsne_gau)
-
-## Compute 2D distances
-distance_tsne_gau <- cal_2d_dist(.data = tr_from_to_df_tsne_gau)
-
-## To find the benchmark value
-benchmark_tsne_gau <- find_benchmark_value(.data = distance_tsne_gau, distance_col = distance)
-
-trimesh_removed_tsne_gau <- remove_long_edges(.data = distance_tsne_gau, benchmark_value = benchmark_tsne_gau,
-                                                  triangular_object = tr1_object_tsne_gau, distance_col = distance)
-
-trimesh_removed_tsne_gau <- trimesh_removed_tsne_gau +
-  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  # theme(axis.text = element_text(size = 5),
-  #       axis.title = element_text(size = 7)) +
-  geom_point(colour = "#33a02c", size = 0.05) +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-# tour_tsne_gau <- show_langevitour(df_all_tsne_gau, df_bin_tsne_gau, df_bin_centroids_tsne_gau, benchmark_value = benchmark_tsne_gau, distance = distance_tsne_gau, distance_col = distance)
-
-### UMAP
-
-# num_bins_umap_gau <- calculate_effective_x_bins(.data = UMAP_data_gau, x = UMAP1,
-#                                                 cell_area = 1) #41
-
-shape_val_umap_gau <- calculate_effective_shape_value(.data = UMAP_data_gau,
-                                                      x = UMAP1, y = UMAP2)
-num_bins_umap_gau <- find_non_empty_bins(nldr_df = UMAP_data_gau, x = "UMAP1", y = "UMAP2", shape_val = shape_val_umap_gau, non_empty_bins = num_non_empty_bins_gau)
-
-## To extract bin centroids
-hexbin_data_object_umap_gau <- extract_hexbin_centroids(nldr_df = UMAP_data_gau, num_bins = num_bins_umap_gau, shape_val = shape_val_umap_gau, x = UMAP1, y = UMAP2)
-
-df_bin_centroids_umap_gau <- hexbin_data_object_umap_gau$hexdf_data
-
-UMAP_data_with_hb_id_gau <- UMAP_data_gau |>
-  dplyr::mutate(hb_id = hexbin_data_object_umap_gau$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_umap_gau <- dplyr::bind_cols(training_data_1 |> dplyr::select(-ID), UMAP_data_with_hb_id_gau)
-
-## Averaged on high-D
-df_bin_umap_gau <- avg_highD_data(.data = df_all_umap_gau)
-
-## Triangulate bin centroids
-tr1_object_umap_gau <- triangulate_bin_centroids(df_bin_centroids_umap_gau, x, y)
-tr_from_to_df_umap_gau <- generate_edge_info(triangular_object = tr1_object_umap_gau)
-
-## Compute 2D distances
-distance_umap_gau <- cal_2d_dist(.data = tr_from_to_df_umap_gau)
-
-## To find the benchmark value
-benchmark_umap_gau <- find_benchmark_value(.data = distance_umap_gau, distance_col = distance)
-
-trimesh_removed_umap_gau <- remove_long_edges(.data = distance_umap_gau, benchmark_value = benchmark_umap_gau,
-                                              triangular_object = tr1_object_umap_gau, distance_col = distance)
-
-trimesh_removed_umap_gau <- trimesh_removed_umap_gau +
-  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  # theme(axis.text = element_text(size = 5),
-  #       axis.title = element_text(size = 7)) +
-  geom_point(colour = "#33a02c", size = 0.05) +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-# tour_umap_gau <- show_langevitour(df_all_umap_gau, df_bin_umap_gau, df_bin_centroids_umap_gau, benchmark_value = benchmark_umap_gau, distance = distance_umap_gau, distance_col = distance)
-
-## PAHTE
-
-# num_bins_phate_gau <- calculate_effective_x_bins(.data = PHATE_data_gau, x = PHATE1,
-#                                                  cell_area = 1)
-
-shape_val_phate_gau <- calculate_effective_shape_value(.data = PHATE_data_gau,
-                                                       x = PHATE1, y = PHATE2)
-num_bins_phate_gau <- find_non_empty_bins(nldr_df = PHATE_data_gau, x = "PHATE1", y = "PHATE2", shape_val = shape_val_phate_gau, non_empty_bins = num_non_empty_bins_gau)
-
-## To extract bin centroids
-hexbin_data_object_phate_gau <- extract_hexbin_centroids(nldr_df = PHATE_data_gau, num_bins = num_bins_phate_gau, shape_val = shape_val_phate_gau, x = PHATE1, y = PHATE2)
-
-df_bin_centroids_phate_gau <- hexbin_data_object_phate_gau$hexdf_data
-
-PHATE_data_with_hb_id_gau <- PHATE_data_gau |>
-  dplyr::mutate(hb_id = hexbin_data_object_phate_gau$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_phate_gau <- dplyr::bind_cols(training_data_1 |> dplyr::select(-ID), PHATE_data_with_hb_id_gau)
-
-## Averaged on high-D
-df_bin_phate_gau <- avg_highD_data(.data = df_all_phate_gau)
-
-## Triangulate bin centroids
-tr1_object_phate_gau <- triangulate_bin_centroids(df_bin_centroids_phate_gau, x, y)
-tr_from_to_df_phate_gau <- generate_edge_info(triangular_object = tr1_object_phate_gau)
-
-## Compute 2D distances
-distance_phate_gau <- cal_2d_dist(.data = tr_from_to_df_phate_gau)
-
-## To find the benchmark value
-benchmark_phate_gau <- find_benchmark_value(.data = distance_phate_gau, distance_col = distance)
-benchmark_phate_gau <- 0.00078
-
-trimesh_removed_phate_gau <- remove_long_edges(.data = distance_phate_gau, benchmark_value = benchmark_phate_gau,
-                                               triangular_object = tr1_object_phate_gau, distance_col = distance)
-
-trimesh_removed_phate_gau <- trimesh_removed_phate_gau +
-  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  # theme(axis.text = element_text(size = 5),
-  #       axis.title = element_text(size = 7)) +
-  geom_point(colour = "#33a02c", size = 0.05) +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'c', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-# tour_phate_gau <- show_langevitour(df_all_phate_gau, df_bin_phate_gau, df_bin_centroids_phate_gau, benchmark_value = benchmark_phate_gau, distance = distance_phate_gau, distance_col = distance)
-
-## TriMAP
-
-# num_bins_trimap_gau <- calculate_effective_x_bins(.data = TriMAP_data_gau, x = TriMAP1,
-#                                                   cell_area = 1) #143
-
-shape_val_trimap_gau <- calculate_effective_shape_value(.data = TriMAP_data_gau,
-                                                        x = TriMAP1, y = TriMAP2)
-
-num_bins_trimap_gau <- find_non_empty_bins(nldr_df = TriMAP_data_gau, x = "TriMAP1", y = "TriMAP2", shape_val = shape_val_trimap_gau, non_empty_bins = num_non_empty_bins_gau)
-
-## To extract bin centroids
-hexbin_data_object_trimap_gau <- extract_hexbin_centroids(nldr_df = TriMAP_data_gau, num_bins = num_bins_trimap_gau, shape_val = shape_val_trimap_gau, x = TriMAP1, y = TriMAP2)
-
-df_bin_centroids_trimap_gau <- hexbin_data_object_trimap_gau$hexdf_data
-
-TriMAP_data_with_hb_id_gau <- TriMAP_data_gau |>
-  dplyr::mutate(hb_id = hexbin_data_object_trimap_gau$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_trimap_gau <- dplyr::bind_cols(training_data_1 |> dplyr::select(-ID), TriMAP_data_with_hb_id_gau)
-
-## Averaged on high-D
-df_bin_trimap_gau <- avg_highD_data(.data = df_all_trimap_gau)
-
-## Triangulate bin centroids
-tr1_object_trimap_gau <- triangulate_bin_centroids(df_bin_centroids_trimap_gau, x, y)
-tr_from_to_df_trimap_gau <- generate_edge_info(triangular_object = tr1_object_trimap_gau)
-
-## Compute 2D distances
-distance_trimap_gau <- cal_2d_dist(.data = tr_from_to_df_trimap_gau)
-
-## To find the benchmark value
-benchmark_trimap_gau <- find_benchmark_value(.data = distance_trimap_gau, distance_col = distance)
-
-trimesh_removed_trimap_gau <- remove_long_edges(.data = distance_trimap_gau, benchmark_value = benchmark_trimap_gau,
-                                                triangular_object = tr1_object_trimap_gau, distance_col = distance)
-
-trimesh_removed_trimap_gau <- trimesh_removed_trimap_gau +
-  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  # theme(axis.text = element_text(size = 5),
-  #       axis.title = element_text(size = 7)) +
-  geom_point(colour = "#33a02c", size = 0.05) +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'd', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-# tour_trimap_gau <- show_langevitour(df_all_trimap_gau, df_bin_trimap_gau, df_bin_centroids_trimap_gau, benchmark_value = benchmark_trimap_gau, distance = distance_trimap_gau, distance_col = distance)
-
-## PaCMAP
-
-# num_bins_pacmap_gau <- calculate_effective_x_bins(.data = PaCMAP_data_gau, x = PaCMAP1,
-#                                                   cell_area = 1) #30
-
-shape_val_pacmap_gau <- calculate_effective_shape_value(.data = PaCMAP_data_gau,
-                                                        x = PaCMAP1, y = PaCMAP2)
-
-num_bins_pacmap_gau <- find_non_empty_bins(nldr_df = PaCMAP_data_gau, x = "PaCMAP1", y = "PaCMAP2", shape_val = shape_val_pacmap_gau, non_empty_bins = num_non_empty_bins_gau)
-
-## To extract bin centroids
-hexbin_data_object_pacmap_gau <- extract_hexbin_centroids(nldr_df = PaCMAP_data_gau, num_bins = num_bins_pacmap_gau, shape_val = shape_val_pacmap_gau, x = PaCMAP1, y = PaCMAP2)
-
-df_bin_centroids_pacmap_gau <- hexbin_data_object_pacmap_gau$hexdf_data
-
-PaCMAP_data_with_hb_id_gau <- PaCMAP_data_gau |>
-  dplyr::mutate(hb_id = hexbin_data_object_pacmap_gau$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_pacmap_gau <- dplyr::bind_cols(training_data_1 |> dplyr::select(-ID), PaCMAP_data_with_hb_id_gau)
-
-## Averaged on high-D
-df_bin_pacmap_gau <- avg_highD_data(.data = df_all_pacmap_gau)
-
-## Triangulate bin centroids
-tr1_object_pacmap_gau <- triangulate_bin_centroids(df_bin_centroids_pacmap_gau, x, y)
-tr_from_to_df_pacmap_gau <- generate_edge_info(triangular_object = tr1_object_pacmap_gau)
-
-## Compute 2D distances
-distance_pacmap_gau <- cal_2d_dist(.data = tr_from_to_df_pacmap_gau)
-
-## To find the benchmark value
-benchmark_pacmap_gau <- find_benchmark_value(.data = distance_pacmap_gau, distance_col = distance)
-
-trimesh_removed_pacmap_gau <- remove_long_edges(.data = distance_pacmap_gau, benchmark_value = benchmark_pacmap_gau,
-                                                triangular_object = tr1_object_pacmap_gau, distance_col = distance)
-
-trimesh_removed_pacmap_gau <- trimesh_removed_pacmap_gau +
-  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  # theme(axis.text = element_text(size = 5),
-  #       axis.title = element_text(size = 7)) +
-  geom_point(colour = "#33a02c", size = 0.05) +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'e', x = -Inf, y = Inf, hjust = -0.5, vjust = 1, size = 3)
-
-# tour_pacmap_gau <- show_langevitour(df_all_pacmap_gau, df_bin_pacmap_gau, df_bin_centroids_pacmap_gau, benchmark_value = benchmark_pacmap_gau, distance = distance_pacmap_gau, distance_col = distance)
-
-```
-
-```{r}
-#| eval: false
+  annotate(geom = 'text', label = 'e', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+df_all_pacmap_gau <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID), pacmap_data_with_hb_id_gau)
+
+# show_langevitour(df = df_all_pacmap_gau, df_b = df_bin_pacmap_gau, df_b_with_center_data = df_bin_centroids_pacmap_gau, col_start = "x", distance_df = distance_pacmap_gau, distance_col = "distance", benchmark_value = benchmark_pacmap_gau)
+
+```
+
+```{r}
 #| warning: false
@@
 ```
+
+<!---
 

### 2024-03-25 paper.qmd
Added screenshots for simulated data, videos

@@
                                              counts_df = counts_df)
-pcentroid_plot_s_curve <- ggplot(data = hex_grid, aes(x = x, y = y)) +
-  geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
-  geom_point(data = df_bin_centroids, aes(x = c_x, y = c_y), color = "#33a02c") +
-  coord_equal() +
-  theme_linedraw() +
-  theme(legend.position = "bottom", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+## remove low-density hexagons
 
 hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
+
+hex_full_count_df_loop2 <- hex_grid_with_counts |>
+  dplyr::mutate(type = if_else(std_counts <= 0.1, "low", "high"))
+
+hex_full_count_df_filtered <- df_bin_centroids |>
+  dplyr::mutate(type = if_else(std_counts <= 0.2, "low", "high")) |>
+  dplyr::filter(type == "low")
+
+p2_n <-  ggplot(data = hex_full_count_df_loop2, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+  #geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  #coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
+  theme_void() +
+  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6)) +
+  guides(fill = guide_colourbar(title = "Standardized count")) #+
+  #annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
+
+p2_n_rm <-  ggplot(data = hex_full_count_df_loop2, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+  geom_text(data = hex_full_count_df_filtered, color = "red", aes(x = c_x, y = c_y, group = hexID, label = "X", size = 0.5)) +
+  #geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  #coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
+  theme_void() +
+  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6)) +
+  guides(fill = guide_colourbar(title = "Standardized count")) #+
+  #annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
 
@@
   geom_quasirandom() + xlab("hexagonal id") + ylab("Standardized cell count") +
-  geom_hline(yintercept = 0.255, colour = "#de2d26") +
+  geom_hline(yintercept = 0.2, colour = "#de2d26") +
   theme(axis.text = element_text(size = 5),
@@
         axis.text.x = element_text(angle = 90))
+
+benchmark_value_rm_lwd <- stats::quantile(df_bin_centroids_s_curve$std_counts, 
+                probs = c(0,0.25,0.5,0.75,1), names = FALSE)[2]
+
+df_bin_centroids_low <- df_bin_centroids_s_curve |>
+  dplyr::filter(std_counts <= benchmark_value_rm_lwd)
+
+identify_rm_bins <- find_low_dens_hex(df_bin_centroids_all = df_bin_centroids_s_curve, 
+                                      num_bins_x = num_bins_x_s_curve, 
+                                      df_bin_centroids_low = df_bin_centroids_low)
+
+df_bin_centroids_s_curve <- df_bin_centroids_s_curve |>
+  dplyr::filter(!(hexID %in% identify_rm_bins))
+
+pcentroid_plot_s_curve <- ggplot(data = hex_grid, aes(x = x, y = y)) +
+  geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
+  geom_point(data = df_bin_centroids_s_curve, aes(x = c_x, y = c_y), color = "#33a02c") +
+  coord_equal() +
+  theme_linedraw() +
+  theme(legend.position = "bottom", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+p2_n_rm_correct <-  ggplot(data = hex_full_count_df_loop2, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+  geom_text(data = df_bin_centroids_low |> dplyr::filter(hexID %in% identify_rm_bins), color = "red", aes(x = c_x, y = c_y, group = hexID, label = "X"), size = 0.5) +
+  #geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  #coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
+  theme_void() +
+  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6)) +
+  guides(fill = guide_colourbar(title = "Standardized count")) #+
+  #annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
 
@@
 
+# 
+# distance_plot
+
+## To find the benchmark value
+benchmark_s_curve <- find_lg_benchmark(distance_edges = distance_s_curve, distance_col = "distance")
+
 distance_plot <- plot_dist(distance_s_curve) +
+   geom_hline(yintercept = benchmark_s_curve, linetype="solid",
+                color = "red", size=0.8, alpha = 0.5) +
   ylab(expression(d^{(2)})) +
@@
         axis.title = element_text(size = 12))
-# 
-# distance_plot
-
-## To find the benchmark value
-benchmark_s_curve <- find_lg_benchmark(distance_edges = distance_s_curve, distance_col = "distance")
 
@@
   annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+# df_all_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), umap_data_with_hb_id)
+# 
+# show_langevitour(df_all_s_curve, df_bin_s_curve, df_bin_centroids_s_curve, benchmark_s_curve, distance_s_curve, "distance", col_start = "x")
 ```
@@
 #| label: fig-stdctsScurve
+#| fig-pos: H
 
@@
 
-<!--
-```{r}
+```{r}
+#| warning: false
 #| echo: false
 #| message: false
-#| warning: false
-#| eval: false
-
-hex_full_count_df_loop2 <- read_rds("data/s_curve/s_curve_hex_8.rds")
-
-hex_full_count_df_loop2 <- hex_full_count_df_loop2 |>
-  dplyr::mutate(type = if_else(std_counts <= 0.20, "low", "high"))
-
-hex_full_count_df_filtered <- hex_full_count_df_loop2 |>
-  dplyr::filter(type == "low")
-
-p2_n <-  ggplot(data = hex_full_count_df_loop2, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
-  #geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  #coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
-  theme_void() +
-  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=8), #change legend title font size
-        legend.text = element_text(size=6)) +
-  guides(fill = guide_colourbar(title = "Standardized count")) #+
-  #annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
-
-p2_n_rm <-  ggplot(data = hex_full_count_df_loop2, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
-  geom_text(data = hex_full_count_df_filtered, color = "red", aes(x = c_x, y = c_y, group = polygon_id, label = "X", size = 0.5)) +
-  #geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  #coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
-  theme_void() +
-  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=8), #change legend title font size
-        legend.text = element_text(size=6)) +
-  guides(fill = guide_colourbar(title = "Standardized count")) #+
-  #annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
-
-## Import data
-
-UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap.rds")
-
-## UMAP
-
-num_bins_umap_s_curve <- 8
-shape_val_umap_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
-                                                          x = UMAP1, y = UMAP2) ## 1.259938
-## To extract bin centroids
-hexbin_data_object_umap_s_curve <- extract_hexbin_centroids(nldr_df = UMAP_s_curve, num_bins = num_bins_umap_s_curve, shape_val = shape_val_umap_s_curve, x = UMAP1, y = UMAP2)
-
-df_bin_centroids_umap_s_curve <- hexbin_data_object_umap_s_curve$hexdf_data
-
-cell_count_plot_s_curve <- ggplot(df_bin_centroids_umap_s_curve, aes(x = reorder(as.factor(hexID), -std_counts), y = std_counts)) +
-  geom_quasirandom() + xlab("hexagonal id") + ylab("Standardized cell count") +
-  geom_hline(yintercept = 0.20, colour = "#de2d26") +
-  theme(axis.text = element_text(size = 5),
-        axis.title = element_text(size = 7),
-        axis.text.x = element_text(angle = 90))
-
-df_bin_centroids_umap_s_curve <- df_bin_centroids_umap_s_curve |>
-  dplyr::filter(std_counts > 0.20)
-
-## Triangulate bin centroids
-tr1_object_umap_s_curve <- triangulate_bin_centroids(df_bin_centroids_umap_s_curve, x, y)
-tr_from_to_df_umap_s_curve <- generate_edge_info(triangular_object = tr1_object_umap_s_curve)
-
-## Compute 2D distances
-distance_umap_s_curve <- cal_2d_dist(.data = tr_from_to_df_umap_s_curve)
-
-## To find the benchmark value
-benchmark_umap_s_curve <- find_benchmark_value(.data = distance_umap_s_curve, distance_col = distance)
-
-trimesh_umap_s_curve_rm_low <- ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
-  geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
-  geom_point(size = 2, colour = "#33a02c") +
-  coord_equal()
-
-trimesh_umap_s_curve_rm_low <- trimesh_umap_s_curve_rm_low +
-  coord_cartesian(xlim =c(-5, 7.5), ylim = c(-10, 10)) +
-  #ggtitle("(a)") +
-  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  theme_light() +
-  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
-  ) #+
-  #annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-#| eval: false
-#| echo: false
-#| warning: false
-#| fig-cap: Goodness of f
-#| label: fig-rmlowdenshex
-#| fig-pos: H
-##| out-width: 100%
-
-p2_n + p2_n_rm + trimesh_umap_s_curve_rm_low +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', ncol = 3) &
-  theme(legend.position='bottom', plot.tag = element_text(size = 8))
-```
-
--->
-
-<!--Addressing low-density hexagons is a systematic process to handle sparsely represented data in certain regions.
-For each hex bin, we identify the six nearest hex bins using an equal 2D distance metric.
-Then, we calculate the mean density, as outlined in the equations:
-
-$$
-\text{standard count} = \frac{\text{count}}{\text{max count}} 
-$${#eq-equationp2}
-
-$$
-\text{mean density} = \frac{\text{standard count}}{6} 
-$${#eq-equationp3}
-
-The standard count is derived from the number of observations in the hex bins.
-By examining the distribution of mean densities and designating the first quartile as the benchmark value, hex bins with mean densities below this benchmark are removed.
-This process ensures the elimination of regions with insufficient data density, focusing on areas with more significant data representation and preserving the overall structure in the low-dimensional space.-->
-
-<!--lwd smoothing criteria 1 with s-curve-->
-<!--
-```{r}
-#| warning: false
-#| echo: false
-#| eval: false
-
-MSE_df_1_lwd <- read_rds("data/s_curve/s_curve_summary_lwd_tsne.rds")
-MSE_df_2_lwd <- read_rds("data/s_curve/s_curve_summary_lwd_umap.rds")
-MSE_df_3_lwd <- read_rds("data/s_curve/s_curve_summary_lwd_phate.rds")
-MSE_df_4_lwd <- read_rds("data/s_curve/s_curve_summary_lwd_trimap.rds")
-MSE_df_5_lwd <- read_rds("data/s_curve/s_curve_summary_lwd_pacmap.rds")
-
-MSE_df_lwd <- dplyr::bind_rows(MSE_df_1_lwd, MSE_df_2_lwd, MSE_df_3_lwd, MSE_df_4_lwd, MSE_df_5_lwd)
-
-MSE_df_lwd$method <- factor(MSE_df_lwd$method, levels = c("tSNE", "UMAP", "PHATE", "TriMAP", "PaCMAP"))
-
-## To draw with AIC
-aic_plot_lwd <- ggplot(MSE_df_lwd |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "UMAP"), aes(x = benchmark_rm_hex,
-                                                                                 y = total_error,
-                                                                                 color = method
+num_bins_x_s_curve <- 6
+num_bins_y_s_curve <- 14
+hex_size_s_curve <- 0.13
+
+vec <- stats::setNames(rep("", 2), c("benchmark_val_lwd", "mse"))  ## Define column names
+
+mse_df <- dplyr::bind_rows(vec)[0, ]
+mse_df <- mse_df |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+model_object <- fit_highd_model( training_data = training_data, 
+                                 nldr_df_with_id = umap_s_curve_scaled, 
+                                 x = "UMAP1", y = "UMAP2", 
+                                 num_bins_x = num_bins_x_s_curve, 
+                                 num_bins_y = num_bins_y_s_curve, 
+                                 x_start = NA, y_start = NA, 
+                                 buffer_x = NA, buffer_y = NA, 
+                                 hex_size = hex_size_s_curve,
+                                 is_rm_lwd_hex = FALSE, 
+                                 benchmark_to_rm_lwd_hex = NA, 
+                                 col_start_2d = "UMAP", 
+                                 col_start_highd = "x")
+
+centroid_df_training_all <- model_object$df_bin_centroids
+avg_df_training_all <- model_object$df_bin
+
+benchmark_rm_hex_vec <- seq(0, 1, by=0.1)
+benchmark_rm_hex_vec <- append(benchmark_rm_hex_vec[1:10],0.99)
+
+for (i in 1:length(benchmark_rm_hex_vec)) {
+
+  centroid_df_training <- centroid_df_training_all |>
+    dplyr::filter(std_counts > benchmark_rm_hex_vec[i])
+
+  avg_df_training <- avg_df_training_all |>
+    dplyr::filter(hb_id %in% centroid_df_training$hexID)
+
+  pred_emb_list <- predict_emb(test_data = training_data, 
+                               df_bin_centroids = centroid_df_training, 
+                               df_bin = avg_df_training, type_NLDR = "UMAP")
+
+  pred_df_training <- as.data.frame(do.call(cbind, pred_emb_list))
+
+  eval_list <- gen_summary(test_data = training_data, 
+                           prediction_df = pred_df_training, 
+                           df_bin = avg_df_training, col_start = "x")
+
+  mse_df <- mse_df |>
+    tibble::add_row(benchmark_val_lwd = benchmark_rm_hex_vec[i],
+                    mse = eval_list$mse)
+
+}
+
+mse_plot_lwd <- ggplot(mse_df, aes(x = benchmark_val_lwd,
+                                       y = mse
 )) +
@@
   geom_line() +
-  geom_vline(xintercept = 0.2, linetype="dashed", 
-                color = "red", size=0.5) +
-  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+   geom_vline(xintercept = 0.2, linetype="solid",
+                color = "red", size=0.8, alpha = 0.5) +
   theme_light() +
@@
         axis.text = element_text(size = 7)) +
-  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("AIC") +
-  xlab("benchmark value")
-## Effective number of bins along x-axis
-
-mse_plot_lwd <- ggplot(MSE_df_lwd |> dplyr::filter(method == "UMAP"), aes(x = benchmark_rm_hex,
-                                       y = total_mse,
-                                       color = method
-)) +
-  geom_point() +
-  geom_line() +
-  geom_vline(xintercept = 0.2, linetype="dashed", 
-                color = "red", size=0.5) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
+  #scale_colour_manual(values = c("#377eb8")) +
   ylab("MSE") +
-  xlab("benchmark value")
-
-```
-
-```{r}
-#| eval: false
+  xlab("benchmark value to remove low-density hexagons")
+
+```
+
+```{r}
 #| echo: false
@@
 
-aic_plot_lwd + mse_plot_lwd +
+mse_plot_lwd
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| fig-cap: Goodness of f
+#| label: fig-rmlowdenshex
+#| fig-pos: H
+##| out-width: 100%
+
+p2_n + p2_n_rm + p2_n_rm_correct +
   plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', ncol = 2) &
+  plot_layout(guides='collect', ncol = 3) &
   theme(legend.position='none', plot.tag = element_text(size = 8))
-```
--->
-
-```{r}
-#| echo: false
-#| fig-cap: "(a) Full hexagonal grid with bin centroids, (b) Model constructed in 2D with long edges, and (c) Model constructed in 2D after removing the long edges."
-#| label: fig-modelScurve
-#| fig-pos: H
-
-pcentroid_plot_s_curve + trimesh_s_curve_umap + trimesh_removed_s_curve_umap +
-  plot_layout(guides='collect', ncol = 3) &
-  theme(legend.position='bottom') 
 ```
@@
 #| label: fig-distScurve
+#| fig-pos: H
 
@@
 2. Find a value which is greater than smallest value
+
+```{r}
+#| echo: false
+#| fig-cap: "(a) Full hexagonal grid with bin centroids, (b) Model constructed in 2D with long edges, and (c) Model constructed in 2D after removing the long edges."
+#| label: fig-modelScurve
+#| fig-pos: H
+
+pcentroid_plot_s_curve + trimesh_s_curve_umap + trimesh_removed_s_curve_umap +
+  plot_layout(guides='collect', ncol = 3) &
+  theme(legend.position='bottom') 
+```
 
@@
 Despite the various 2D representations, all NLDR techniques preserve the global structure (see video link of @fig-modelfiveGau (d), (e) respectively).
-However, tSNE, effectively capture both local and global structures, as indicated by lower AIC values in @fig-diagnosticpltGau (a).
+However, tSNE, effectively capture both local and global structures.
 
@@
 #| label: fig-modelfiveGau
-#| fig-cap: "Is there a best model to represent the original data in 2D space or are they all providing equivalent information?, (a) Model in the 2D space with tSNE (<https://youtu.be/x5VPB-wOLm4>), (b) Model in the 2D space with UMAP (<https://youtu.be/Cm1dW6iLyHQ>), (c) Model in the 2D space with PHATE (<https://youtu.be/HB0Y3ilz6sI>), (d) Model in the 2D space with TriMAP (<https://youtu.be/iSVkXIrOrSA>), and (e) Model in the 2D space with PaCMAP (<https://youtu.be/O2QBctsc4uM>)."
+#| fig-cap: "Is there a best model to represent the original data in 2D space or are they all providing equivalent information?, (a) Model generated with 21 and 28 bins along the x and y axes, respectively, using a hexagonal size of 0.03 in the 2D space with tSNE, (b) Model generated with 60 and 79 bins along the x and y axes, respectively, using a hexagonal size of 0.01 in the 2D space with UMAP, (c) Model generated with 60 and 1927 bins along the x and y axes, respectively, using a hexagonal size of 0.01 in the 2D space with PHATE, (d) Model generated with 46 and 61 bins along the x and y axes, respectively, using a hexagonal size of 0.013 in the 2D space with TriMAP, and (e) Model generated with 31 and 40 bins along the x and y axes, respectively, using a hexagonal size of 0.02 in the 2D space with PaCMAP."
 #| out-width: 100%
@@
 
-<!---
-
-```{r}
-#| eval: false
-#| warning: false
-#| echo: false
-
-## tSNE
-
-shape_value_gau <- calculate_effective_shape_value(.data = tSNE_data_gau,
-                                                   x = tSNE1, y = tSNE2)
-
-num_bins_vec <- 1:20 ## Number of bins along the x-axis
-
-vec <- stats::setNames(rep("", 4), c("number_of_bins", "number_of_observations", "total_error", "total_mse"))  ## Define column names
-
-eval_data_training <- dplyr::bind_rows(vec)[0, ]
-eval_data_training <- eval_data_training |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(num_bins_vec)) {
-  model_object <- fit_high_d_model(training_data = training_data_1, nldr_df_with_id = tSNE_data_gau, x = "tSNE1", y = "tSNE2", num_bins_x = num_bins_vec[i], shape_val = shape_value_gau,
-                                 is_bin_centroid = TRUE,
-                                 is_rm_lwd_hex = FALSE,
-                                 benchmark_to_rm_lwd_hex = NA,
-                                 is_avg_high_d = TRUE, column_start_text = "x")
-
-  centroid_df_training <- model_object$df_bin_centroids
-  avg_df_training <- model_object$df_bin
-
-  pred_df_training <- predict_2d_embeddings(test_data = training_data_1, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "tSNE")
-
-  eval_df_training <- generate_eval_df(data = df_2, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, col_start = "x")
-
-  eval_df_training <- eval_df_training |>
-    mutate(num_bins_x = num_bins_vec[i])
-
-  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
-}
-
-## Add new column with data types
-
-eval_data_training <- eval_data_training |>
-  mutate(data_type = "training")
-
-MSE_df_1_gau <- eval_data_training |>
-  dplyr::mutate(method = "tSNE")
-```
-
-```{r}
-#| eval: false
-#| warning: false
-#| echo: false
-
-## UMAP
-## Prediction
-
-shape_value_gau <- calculate_effective_shape_value(.data = UMAP_data_gau,
-                                                   x = UMAP1, y = UMAP2)
-
-num_bins_vec <- 1:20 ## Number of bins along the x-axis
-
-vec <- stats::setNames(rep("", 4), c("number_of_bins", "number_of_observations", "total_error", "total_mse"))  ## Define column names
-
-eval_data_training <- dplyr::bind_rows(vec)[0, ]
-eval_data_training <- eval_data_training |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(num_bins_vec)) {
-
-  model_object <- fit_high_d_model(training_data = training_data_1, nldr_df_with_id = UMAP_data_gau, x = "UMAP1", y = "UMAP2", num_bins_x = num_bins_vec[i], shape_val = shape_value_gau,
-                                 is_bin_centroid = TRUE,
-                                 is_rm_lwd_hex = FALSE,
-                                 benchmark_to_rm_lwd_hex = NA,
-                                 is_avg_high_d = TRUE, column_start_text = "x")
-
-  centroid_df_training <- model_object$df_bin_centroids
-  avg_df_training <- model_object$df_bin
-
-  pred_df_training <- predict_2d_embeddings(test_data = training_data_1, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "UMAP")
-
-  eval_df_training <- generate_eval_df(data = df_2, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, col_start = "x")
-
-  eval_df_training <- eval_df_training |>
-    mutate(num_bins_x = num_bins_vec[i])
-
-  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
-
-}
-
-## Add new column with data types
-
-eval_data_training <- eval_data_training |>
-  mutate(data_type = "training")
-
-MSE_df_2_gau <- eval_data_training |>
-  dplyr::mutate(method = "UMAP")
-```
-
-```{r}
-#| eval: false
-#| warning: false
-#| echo: false
-## PAHTE
-## Prediction
-
-shape_value_gau <- calculate_effective_shape_value(.data = PHATE_data_gau,
-                                                   x = PHATE1, y = PHATE2)
-
-num_bins_vec <- 1:20 ## Number of bins along the x-axis
-
-vec <- stats::setNames(rep("", 4), c("number_of_bins", "number_of_observations", "total_error", "total_mse"))  ## Define column names
-
-eval_data_training <- dplyr::bind_rows(vec)[0, ]
-eval_data_training <- eval_data_training |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(num_bins_vec)) {
-
-  model_object <- fit_high_d_model(training_data = training_data_1, nldr_df_with_id = PHATE_data_gau, x = "PHATE1", y = "PHATE2", num_bins_x = num_bins_vec[i], shape_val = shape_value_gau,
-                                 is_bin_centroid = TRUE,
-                                 is_rm_lwd_hex = FALSE,
-                                 benchmark_to_rm_lwd_hex = NA,
-                                 is_avg_high_d = TRUE, column_start_text = "x")
-
-  centroid_df_training <- model_object$df_bin_centroids
-  avg_df_training <- model_object$df_bin
-
-  pred_df_training <- predict_2d_embeddings(test_data = training_data_1, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "PHATE")
-
-  eval_df_training <- generate_eval_df(data = df_2, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, col_start = "x")
-
-  eval_df_training <- eval_df_training |>
-    mutate(num_bins_x = num_bins_vec[i])
-
-  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
-
-}
-
-## Add new column with data types
-
-eval_data_training <- eval_data_training |>
-  mutate(data_type = "training")
-
-MSE_df_3_gau <- eval_data_training |>
-  dplyr::mutate(method = "PHATE")
-
-```
-
-```{r}
-#| eval: false
-#| warning: false
-#| echo: false
-
-## TriMAP
-
-## Prediction
-
-shape_value_gau <- calculate_effective_shape_value(.data = TriMAP_data_gau,
-                                                   x = TriMAP1, y = TriMAP2)
-
-num_bins_vec <- 1:20 ## Number of bins along the x-axis
-
-vec <- stats::setNames(rep("", 4), c("number_of_bins", "number_of_observations", "total_error", "total_mse"))  ## Define column names
-
-eval_data_training <- dplyr::bind_rows(vec)[0, ]
-eval_data_training <- eval_data_training |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(num_bins_vec)) {
-
-  model_object <- fit_high_d_model(training_data = training_data_1, nldr_df_with_id = TriMAP_data_gau, x = "TriMAP1", y = "TriMAP2", num_bins_x = num_bins_vec[i], shape_val = shape_value_gau,
-                                 is_bin_centroid = TRUE,
-                                 is_rm_lwd_hex = FALSE,
-                                 benchmark_to_rm_lwd_hex = NA,
-                                 is_avg_high_d = TRUE, column_start_text = "x")
-
-  centroid_df_training <- model_object$df_bin_centroids
-  avg_df_training <- model_object$df_bin
-
-  pred_df_training <- predict_2d_embeddings(test_data = training_data_1, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "TriMAP")
-
-  eval_df_training <- generate_eval_df(data = df_2, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, col_start = "x")
-
-  eval_df_training <- eval_df_training |>
-    mutate(num_bins_x = num_bins_vec[i])
-
-  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
-
-}
-
-## Add new column with data types
-
-eval_data_training <- eval_data_training |>
-  mutate(data_type = "training")
-
-MSE_df_4_gau <- eval_data_training |>
-  dplyr::mutate(method = "TriMAP")
-
-```
-
-```{r}
-#| eval: false
-#| warning: false
-#| echo: false
-
-## PaCMAP
-
-## Prediction
-
-shape_value_gau <- calculate_effective_shape_value(.data = PaCMAP_data_gau,
-                                                   x = PaCMAP1, y = PaCMAP2)
-
-num_bins_vec <- 1:20 ## Number of bins along the x-axis
-
-vec <- stats::setNames(rep("", 4), c("number_of_bins", "number_of_observations", "total_error", "total_mse"))  ## Define column names
-
-eval_data_training <- dplyr::bind_rows(vec)[0, ]
-eval_data_training <- eval_data_training |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(num_bins_vec)) {
-
-  model_object <- fit_high_d_model(training_data = training_data_1, nldr_df_with_id = PaCMAP_data_gau, x = "PaCMAP1", y = "PaCMAP2", num_bins_x = num_bins_vec[i], shape_val = shape_value_gau,
-                                 is_bin_centroid = TRUE,
-                                 is_rm_lwd_hex = FALSE,
-                                 benchmark_to_rm_lwd_hex = NA,
-                                 is_avg_high_d = TRUE, column_start_text = "x")
-
-  centroid_df_training <- model_object$df_bin_centroids
-  avg_df_training <- model_object$df_bin
-
-  pred_df_training <- predict_2d_embeddings(test_data = training_data_1, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "PaCMAP")
-
-  eval_df_training <- generate_eval_df(data = df_2, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, col_start = "x")
-
-  eval_df_training <- eval_df_training |>
-    mutate(num_bins_x = num_bins_vec[i])
-
-  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
-
-}
-
-## Add new column with data types
-
-eval_data_training <- eval_data_training |>
-  mutate(data_type = "training")
-
-MSE_df_5_gau <- eval_data_training |>
-  dplyr::mutate(method = "PaCMAP")
-
-```
-
-```{r}
-#| eval: false
-#| warning: false
-#| echo: false
-
-MSE_df_gau <- dplyr::bind_rows(MSE_df_1_gau, MSE_df_2_gau, MSE_df_3_gau, MSE_df_4_gau, MSE_df_5_gau)
-
-MSE_df_gau$method <- factor(MSE_df_gau$method, levels = c("tSNE", "UMAP", "PHATE", "TriMAP", "PaCMAP"))
-
-## To draw with AIC
-aic_gau_plot <- ggplot(MSE_df_gau |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
-                                                             y = total_error,
-                   color = method
-)) +
-  geom_point() +
-  geom_line() +
-  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  theme_light() +
-  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  scale_colour_manual(values = c("#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("AIC") +
-  xlab("total number of bins") + 
-  facet_wrap(vars(method), ncol = 5)
-## Effective number of bins along x-axis
-
-mse_gau_plot <- ggplot(MSE_df_gau, aes(x = number_of_bins,
-                   y = total_mse,
-                   color = method
-)) +
-  geom_point() +
-  geom_line() +
-  theme_light() +
-  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  scale_colour_manual(values = c("#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("MSE") +
-  xlab("total number of bins") + 
-  facet_wrap(vars(method), ncol = 5)
-
-```
-
--->
+::: {#fig-gau1_sc layout-ncol="3" fig-pos="H"}
+![](figures/five_gau_clusters/sc_tsne_1.png){#fig-gau1_sc1}
+
+![](figures/five_gau_clusters/sc_tsne_2.png){#fig-gau1_sc2}
+
+![](figures/five_gau_clusters/sc_tsne_3.png){#fig-gau1_sc3}
+
+Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with tSNE in high-D, a video of the tour animation is available at (<https://youtu.be/oQxEb4wRdHI>).
+:::
+
+::: {#fig-gau2_sc layout-ncol="3" fig-pos="H"}
+![](figures/five_gau_clusters/sc_umap_1.png){#fig-gau2_sc1}
+
+![](figures/five_gau_clusters/sc_umap_2.png){#fig-gau2_sc2}
+
+![](figures/five_gau_clusters/sc_umap_3.png){#fig-gau2_sc3}
+
+Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with UMAP in high-D, a video of the tour animation is available at (<https://youtu.be/JW49csPpDx4>).
+:::
+
+::: {#fig-gau3_sc layout-ncol="3" fig-pos="H"}
+![](figures/five_gau_clusters/sc_phate_1.png){#fig-gau3_sc1}
+
+![](figures/five_gau_clusters/sc_phate_2.png){#fig-gau3_sc2}
+
+![](figures/five_gau_clusters/sc_phate_3.png){#fig-gau3_sc3}
+
+Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with PHATE in high-D, a video of the tour animation is available at (<https://youtu.be/xqhRODt8Z-o>).
+:::
+
+::: {#fig-gau4_sc layout-ncol="3" fig-pos="H"}
+![](figures/five_gau_clusters/sc_trimap_1.png){#fig-gau4_sc1}
+
+![](figures/five_gau_clusters/sc_trimap_2.png){#fig-gau4_sc2}
+
+![](figures/five_gau_clusters/sc_trimap_3.png){#fig-gau4_sc3}
+
+Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with TriMAP in high-D, a video of the tour animation is available at (<https://youtu.be/X-uHSB4VoZ0>).
+:::
+
+::: {#fig-gau5_sc layout-ncol="3" fig-pos="H"}
+![](figures/five_gau_clusters/sc_pacmap_1.png){#fig-gau5_sc1}
+
+![](figures/five_gau_clusters/sc_pacmap_2.png){#fig-gau5_sc2}
+
+![](figures/five_gau_clusters/sc_pacmap_3.png){#fig-gau5_sc3}
+
+Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with PaCMAP in high-D, a video of the tour animation is available at (<https://youtu.be/UxvU8HQd7JM>).
+:::
 

### 2024-03-26 appendix.qmd
Clear the repo

--- 2024-03-05 appendix.qmd
@@
 library(readr)
+library(tidyr)
 
 library(Rtsne)
-library(umap)
+library(uwot)
 library(phateR)
-library(reticulate)
 library(patchwork)
-
+library(langevitour)
+library(colorspace)
+library(kableExtra)
 library(grid)
-
-# install.packages("Seurat")
-# remotes::install_github("satijalab/seurat-data")
-#library(SeuratData) ## For the application
-
-# use_python("~/miniforge3/envs/pcamp_env/bin/python")
-# use_condaenv("pcamp_env")
-# 
-# reticulate::source_python(paste0(here::here(), "/examples_for_the_paper/function_scripts/Fit_PacMAP_code.py"))
-# reticulate::source_python(paste0(here::here(), "/examples_for_the_paper/function_scripts/Fit_TriMAP_code.py"))
-
-set.seed(20230531)
-
-source("quollr_code.R", local = TRUE)
+set.seed(20240110)
+
 source("nldr_code.R", local = TRUE)
@@
 
-```{r}
-#| warning: false
-#| echo: false
-
-data <- read_rds("data/s_curve/s_curve.rds")
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-training_data <- read_rds("data/s_curve/s_curve_training.rds")
-test_data <- read_rds("data/s_curve/s_curve_test.rds")
-```
-
-```{r}
-#| echo: false
-
-UMAP_data <- read_rds("data/s_curve/s_curve_umap.rds")
-
-num_bins_x <- calculate_effective_x_bins(.data = UMAP_data, x = UMAP1,
-                           cell_area = 1)
-
-shape_value <- calculate_effective_shape_value(.data = UMAP_data, 
-                                             x = UMAP1, y = UMAP2)
-
-## To extract bin centroids
-hexbin_data_object <-extract_hexbin_centroids(nldr_df = UMAP_data, num_bins = num_bins_x, shape_val = shape_value)
-
-df_bin_centroids <- hexbin_data_object$hexdf_data
-
-##########
-
-## Data set with all possible centroids in the hexagonal grid
-
-full_centroid_df <- generate_full_grid_centroids(df_bin_centroids)
-
-## To map hexID to hexbin centroids in the full grid
-
-vec1 <- stats::setNames(rep("", 2), c("x", "y"))  ## Define column names
-
-full_grid_with_hexbin_id <- dplyr::bind_rows(vec1)[0, ]
-full_grid_with_hexbin_id <- full_grid_with_hexbin_id |>
+<!-- To find the number of bins should assign for the model in tSNE-->
+```{r}
+#| warning: false
+#| echo: false
+#| message: false
+
+training_data_gau <- read_rds("data/five_gau_clusters/data_five_gau_training.rds")
+tSNE_data_gau <- read_rds("data/five_gau_clusters/tsne_data_five_gau_61.rds")
+
+tsne_gau_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tSNE_data_gau, 
+                                    x = "tSNE1", y = "tSNE2"))) |>
+  dplyr::rename(c("tSNE1" = "scaled_tSNE1", 
+                  "tSNE2" = "scaled_tSNE2")) |>
+  dplyr::mutate(ID = 1:NROW(tSNE_data_gau))
+
+## tSNE
+hex_size_vec <- seq(0.02, 2, by = 0.01)
+
+vec <- stats::setNames(rep("", 6), c("num_bins", "aic", "mse", "num_bins_x", "num_bins_y", "hex_size"))  ## Define column names
+
+mse_df_gau <- dplyr::bind_rows(vec)[0, ]
+mse_df_gau <- mse_df_gau |>
   dplyr::mutate_if(is.character, as.numeric)
 
-for(i in 1:length(sort(unique(full_centroid_df$y)))){
-
-  ## Filter the data set with specific y value
-  specific_y_val_df <- full_centroid_df |>
-    dplyr::filter(y == sort(unique(full_centroid_df$y))[i])
-
-  ordered_x_df <- specific_y_val_df |>
-    dplyr::arrange(x)
-
-  full_grid_with_hexbin_id <- dplyr::bind_rows(full_grid_with_hexbin_id, ordered_x_df)
+for (i in 1:length(hex_size_vec)) {
+
+  num_bin_list <- calc_bins(data = tsne_gau_scaled, 
+            x = "tSNE1", y = "tSNE2", 
+            hex_size = hex_size_vec[i], buffer_x = NA, buffer_y = NA)
+
+  num_bins_x <- num_bin_list$num_x
+  num_bins_y <- num_bin_list$num_y
+
+  model_object <- fit_highd_model( training_data = training_data_gau, 
+                                   nldr_df_with_id = tsne_gau_scaled, 
+                                   x = "tSNE1", y = "tSNE2", 
+                                   num_bins_x = num_bins_x, 
+                                   num_bins_y = num_bins_y, 
+                                   x_start = NA, y_start = NA, 
+                                   buffer_x = NA, buffer_y = NA, 
+                                   hex_size = hex_size_vec[i],
+                                   is_rm_lwd_hex = FALSE, 
+                                   benchmark_to_rm_lwd_hex = NA, 
+                                   col_start_2d = "tSNE", 
+                                   col_start_highd = "x")
+
+  centroid_df_training <- model_object$df_bin_centroids
+  avg_df_training <- model_object$df_bin
+
+  pred_emb_list <- predict_emb(test_data = training_data_gau, 
+                                  df_bin_centroids = centroid_df_training, 
+                                  df_bin = avg_df_training, type_NLDR = "tSNE")
+
+  pred_df_training <- as.data.frame(do.call(cbind, pred_emb_list))
+
+  eval_list <- gen_summary(test_data = training_data_gau, 
+                                  prediction_df = pred_df_training, 
+                                  df_bin = avg_df_training, col_start = "x")
+
+  mse_df_gau <- mse_df_gau |>
+    tibble::add_row(num_bins = num_bins_x * num_bins_y,
+                    aic = eval_list$aic,
+                    mse = eval_list$mse,
+                    num_bins_x = num_bins_x,
+                    num_bins_y = num_bins_y,
+                    hex_size = hex_size_vec[i])
 
@@
 
-full_grid_with_hexbin_id <- full_grid_with_hexbin_id |>
-  dplyr::mutate(hexID = row_number())
-
-full_grid_with_hexbin_id <- full_grid_with_hexbin_id |>
-  dplyr::rename("c_x" = "x",
-         "c_y" = "y")
-
-full_grid_with_hexbin_id <- dplyr::full_join(full_grid_with_hexbin_id, df_bin_centroids, by = c("hexID" = "hexID")) |>
-  dplyr::select(-c(x, y)) #|> 
-  #dplyr::mutate(counts = tidyr::replace_na(counts, 0))
-
-full_grid_with_hexbin_id <- full_grid_with_hexbin_id |>
-    dplyr::mutate(std_counts = counts/max(counts, na.rm = TRUE))
-
-## Generate all coordinates of hexagons
-hex_grid <- full_hex_grid(full_centroid_df)
-
-full_grid_with_polygon_id_df <- map_polygon_id(full_grid_with_hexbin_id, hex_grid)
-
-full_grid_with_hexbin_id_rep <- full_grid_with_polygon_id_df |>
-  dplyr::slice(rep(1:n(), each = 6)) |>
-  dplyr::arrange(polygon_id)
-
-hex_full_count_df <- dplyr::bind_cols(hex_grid, full_grid_with_hexbin_id_rep)
-
-##########
-
-min_std_cell_threshold <- 0.05
-
-df_bin_centroids_all <- hexbin_data_object$hexdf_data ## All the centroids without removing low-density hexagons
-
-df_bin_centroids <- df_bin_centroids |>
-  dplyr::mutate(stand_cell_count = counts/max(counts)) |>
-  dplyr::filter(stand_cell_count > min_std_cell_threshold)
-
-UMAP_data_with_hb_id <- UMAP_data |> 
-  dplyr::mutate(hb_id = hexbin_data_object$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all <- dplyr::bind_cols(training_data |> dplyr::select(-ID), UMAP_data_with_hb_id)
-
-## Averaged on high-D
-df_bin <- avg_highD_data(.data = df_all)
-
-## Triangulate bin centroids
-tr1_object <- triangulate_bin_centroids(df_bin_centroids, x, y)
-tr_from_to_df <- generate_edge_info(triangular_object = tr1_object)
-
-## Compute 2D distances
-distance <- cal_2D_dist(.data = tr_from_to_df)
-
-## To find the benchmark value
-benchmark <- find_benchmark_value(.data = distance, distance_col = distance)
-
-trimesh <- ggplot(df_bin_centroids, aes(x = x, y = y)) + 
-    geom_segment(data = tr_from_to_df, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
-    geom_point(size = 2, colour = "#33a02c") +
-    coord_equal()
-
-  # ggplot(df_bin_centroids, aes(x = x, y = y)) + 
-  # geom_point(size = 1, colour = "#33a02c") + 
-  # geom_trimesh() + 
-  # coord_equal()
-
-trimesh <- trimesh +
-  #ggtitle("(a)") + 
-  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  theme_light() +
-    theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
-          axis.title.x = element_blank(), axis.title.y = element_blank(),
-          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-          panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
-    ) 
-  # theme(axis.text = element_text(size = 5),
-  #       axis.title = element_text(size = 7))
-
-trimesh_gr <- colour_long_edges(.data = distance, benchmark_value = benchmark, 
-                                triangular_object = tr1_object, distance_col = distance)
-
-trimesh_gr <- trimesh_gr + 
-    geom_point(size = 2, colour = "#33a02c") + 
-  #ggtitle("(b)") + 
-  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  theme_light() +
-    #coord_equal() +
-    theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
-          axis.title.x = element_blank(), axis.title.y = element_blank(),
-          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-          panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
-    )
-
-trimesh_removed <- remove_long_edges(.data = distance, benchmark_value = benchmark, 
-                                triangular_object = tr1_object, distance_col = distance)
-
-trimesh_removed <- trimesh_removed + 
-       geom_point(size = 2, colour = "#33a02c") + 
-  #ggtitle("(b)") + 
-  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  theme_light() +
-    #coord_equal() +
-    theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
-          axis.title.x = element_blank(), axis.title.y = element_blank(),
-          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-          panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
-    )
-
-tour1 <- show_langevitour(df_all, df_bin, df_bin_centroids, benchmark_value = benchmark, distance = distance, distance_col = distance)
-```
-
-```{r}
-#| echo: false
-
-## To plot the distribution of distance
-plot_dist <- function(distance_df){
-  distance_df$group <- "1"
-  dist_plot <- ggplot(distance_df, aes(x = group, y = distance)) +
-    geom_quasirandom()+
-    ylim(0, max(unlist(distance_df$distance))+ 0.5) + coord_flip()
-  return(dist_plot)
+## If same total number of bins occurred only select ones with minimum error
+### Obtain duplicate bins
+dupli_bins <- mse_df_gau |> 
+  dplyr::count(num_bins) |> 
+  dplyr::filter(n > 1) |> 
+  dplyr::pull(num_bins)
+
+### Group split by duplicated bins
+duplicate_df_list <- mse_df_gau |>
+  dplyr::filter(num_bins %in% dupli_bins) |>
+  dplyr::arrange(num_bins) |>
+  dplyr::group_split(num_bins)
+
+### Obtain one row from duplicates which have lowest error and hexsize
+duplicate_df <- data.frame(matrix(nrow = 0, ncol = 0))
+
+for (i in 1:length(duplicate_df_list)) {
+
+  dd <- duplicate_df_list[[i]] |>
+    dplyr::filter(mse == min(duplicate_df_list[[i]]$mse)) |>
+    dplyr::filter(hex_size == min(duplicate_df_list[[i]]$hex_size))
+
+  duplicate_df <- dplyr::bind_rows(duplicate_df, dd)
+
 }
-```
-
-```{r}
-#| echo: false
-
-distance_plot <- plot_dist(distance) +
-  #ggtitle("(b)" ) + 
-  ylab(expression(d^{(2)})) +
-  theme(axis.text = element_text(size = 5),
-        axis.title = element_text(size = 12))
-```
-
-## Notations
-
-| Notation                                                                                                              | Description                                                            |
-|-----------------------------------------------------------------------------------------------------------------------|------------------------------------------------------------------------|
-| $n$                                                                                                                   | number of observations                                                 |
-| $p$                                                                                                                   | number of dimensions in high-D data                                    |
-| $X_{n \times p} = \begin{bmatrix} \textbf{x} _{1} & \textbf{x}_ {2} & \cdots & \textbf{x}_{n} \\  \end{bmatrix}^\top$ | high-D data matrix                                                     |
-| $d$                                                                                                                   | number of dimensions in embedding data, usually two                    |
-| $Y_{n \times d} = \begin{bmatrix} \textbf{y} _{1} & \textbf{y}_ {2} & \cdots & \textbf{y}_{n} \\  \end{bmatrix}^\top$ | embedding matrix                                                       |
-| $P_t$                                                                                                                 | point along the geodesic path at time $t$                              |
-| $P_{t+1}$                                                                                                             | point along the geodesic path at the next time instant, which is $t+1$ |
-| $b_1$                                                                                                                 | number of bins along the x-axis                                        |
-| $b_2$                                                                                                                 | number of bins along the y-axis                                        |
-| $b$                                                                                                                   | total number of bins in the hexagonal grid                             |
-| $b'$                                                                                                                  | total number of bins without the empty bins                            |
-| $h$                                                                                                                   | diameter of a hexagon                                                  |
-| $A$                                                                                                                   | area of a hexagon                                                      |
-| $r_1$                                                                                                                 | 2D embedding component 1                                               |
-| $r_2$                                                                                                                 | 2D embedding component 2                                               |
-| $s$                                                                                                                   | shape parameter                                                        |
-| $C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2})$                                                                               | hexagonal bin centroid coordinates                                     |
-| $C_k^{(p)} \equiv (C_{kx_1}, ..., C_{kx_p})$                                                                          | averaged high-D coordinates                                            |
-: Notations used in this paper and their descriptions. {#tbl-notations}
-
-## Default parameter values
-
-### Benchmark value to remove low-density hexagons
-
-Removing all the bins that have a standard number of points less than a certain benchmark value will generate unnecessary holes within the triangular mesh.
-Therefore, before removing the low-density bins, it's necessary to investigate the number of points in neighboring bins as well.
-
-$$
-\text{standard count} = \frac{\text{count}}{\text{max count}} 
-$${#eq-equationp2}
-
-The process begins by identifying the six nearest hex bins of specified low-density hex bins.
-To identify the nearest hex bins 2D Euclidean distance metric is used.
-Then, we calculate the mean density, as outlined in the equations:
-
-$$
-\text{mean density} = \frac{\text{standard count}}{6} 
-$${#eq-equationp3}
-
-The standard count is derived from the number of observations in the hex bins.
-By examining the distribution of mean densities and designating the first quartile as the benchmark value, hex bins with mean densities below this benchmark are removed.
-This process ensures the elimination of regions with insufficient data density, focusing on areas with more significant data representation and preserving the overall structure in the low-dimensional space.
-
-### Benchmark value to remove long edges {#sec-rmlgdefault}
-
-Determining a benchmark value for removing long edges isn't a straightforward process with standard guidelines.
-Having a default value provides a practical starting point for decision-making, helping users choose whether to stick with the default, choose a value below it, or select one above it for long edge removal.
-
-To calculate the default value, the process involves computing the 2D Euclidean distances between the vertices in the triangular mesh.
-Initially, distances between vertices are sorted.
-Then, unique distance values are extracted.
-Following this, the differences between consecutive distance values are computed, and a data frame is constructed with unique distance values and their corresponding differences.
-The first largest difference is then used as a benchmark to identify and remove the long edges.
-
-![A flow diagram detailing the steps taken to find the benchmark value to remove long edges.](figures/remove_long_edges_workflow.png){#fig-rmlgmeth fig-align="center" width="100%" height="100%"}
-
-```{r}
-#| echo: false
-
-dist_p <- plot_dist(distance)  +
-  geom_hline(yintercept = benchmark, colour = "blue") + 
-  ylab(expression(d^{(2)})) +
-  theme(axis.text = element_text(size = 5),
-        axis.title = element_text(size = 12)) 
-```
-
-```{r}
-#| echo: false
-#| warning: false
-#| fig-cap: "Is the default benchmark value effective in removing the long edges to construct a smooth surface in 2D?
-(a) Triangular mesh with coloured edges by their type, (b) Distribution of 2D Euclidean distances between triangular mesh vertices."
-#| label: fig-defaultrmlg
-##| out-width: 100%
-
-trimesh_gr + dist_p +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', ncol = 2) &
-  theme(legend.position='bottom', plot.tag = element_text(size = 8))
-```
-
-## Paramter tunning
-
-### Total number of bins
-
-The total number of bins represents the overall count of hexagonal bins within the hexagonal grid.
-This count is determined by multiplying the number of bins along the x-axis ($b_1$) with the number of bins along the y-axis ($b_2$), according to the formula:
-
-$$
- b = b_1 \times b_2
-$$ {#eq-equationp1}
-
-Here, $b$ denotes the total number of bins.
-By adjusting the parameter $b_1$, we have control over the total number of bins ($b$).
-Hence, fine-tuning $b_1$ enables us to customize and optimize the total bin count based on the desired configuration along the x-axis.
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-
-hexgrid_with_points_hex <- ggplot(data = hex_full_count_df |> dplyr::filter(!is.na(std_counts)), aes(x = c_x, y = c_y, fill = std_counts, hexID = hexID)) + 
-    geom_hex(stat = "identity", color = "#969696") +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  theme_light() +
-    #coord_equal() +
-    theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
-          axis.title.x = element_blank(), axis.title.y = element_blank(),
-          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-          panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
-    )
-
-```
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-
-hexgrid_with_points_m <- ggplot(data = hex_full_count_df, aes(x = x, y = y)) + 
-  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) + 
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  theme_light() +
-    #coord_equal() +
-    theme(plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
-          axis.title.x = element_blank(), axis.title.y = element_blank(),
-          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-          panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
-    )
-
-hexgrid_with_points <- hexgrid_with_points_m + theme(legend.position = "none")
-```
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| out-width: 100%
-
-hexgrid_with_points_hex + hexgrid_with_points +
-  plot_layout(ncol = 3)
-
-```
-<!--
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
- # Loop over loop.vector
-  #i <- 1
-  num_bins <- 3
-  shape_val <- calculate_effective_shape_value(.data = UMAP_data, x = UMAP1, y = UMAP2)
-
-  hexbin_data_object_loop <- extract_hexbin_centroids(UMAP_data, num_bins, shape_val)
-
-  df_bin_centroids_loop <- hexbin_data_object_loop$hexdf_data
-
-  ## Data set with all possible centroids in the hexagonal grid
-
-  full_centroid_df_loop <- generate_full_grid_centroids(df_bin_centroids_loop)
-
-  ## To map hexID to hexbin centroids in the full grid
-
-  vec1 <- stats::setNames(rep("", 2), c("x", "y"))  ## Define column names
-
-  full_grid_with_hexbin_id_loop <- dplyr::bind_rows(vec1)[0, ]
-  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
-    dplyr::mutate_if(is.character, as.numeric)
-
-  for(i in 1:length(sort(unique(full_centroid_df_loop$y)))){
-
-    ## Filter the data set with specific y value
-    specific_y_val_df_loop <- full_centroid_df_loop |>
-      dplyr::filter(y == sort(unique(full_centroid_df_loop$y))[i])
-
-    ordered_x_df_loop <- specific_y_val_df_loop |>
-      dplyr::arrange(x)
-
-    full_grid_with_hexbin_id_loop <- dplyr::bind_rows(full_grid_with_hexbin_id_loop, ordered_x_df_loop)
-
-  }
-
-  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
-    dplyr::mutate(hexID = row_number())
-
-  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
-    dplyr::rename("c_x" = "x",
-           "c_y" = "y")
-
-  full_grid_with_hexbin_id_loop <- dplyr::full_join(full_grid_with_hexbin_id_loop, df_bin_centroids_loop, by = c("hexID" = "hexID")) |>
-    dplyr::select(-c(x, y))
-
-  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
-      dplyr::mutate(std_counts = counts/max(counts, na.rm = TRUE))
-
-  ## Generate all coordinates of hexagons
-  hex_grid_loop <- full_hex_grid(full_centroid_df_loop)
-
-  full_grid_with_polygon_id_df_loop <- map_polygon_id(full_grid_with_hexbin_id_loop, hex_grid_loop)
-
-  full_grid_with_hexbin_id_rep_loop <- full_grid_with_polygon_id_df_loop |>
-  dplyr::slice(rep(1:n(), each = 6)) |>
-  dplyr::arrange(polygon_id)
-
-  hex_full_count_df_loop <- dplyr::bind_cols(hex_grid_loop, full_grid_with_hexbin_id_rep_loop)
-
-  p1 <-  ggplot(data = hex_full_count_df_loop, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
-        geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-    theme_void() +
-    theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-          axis.title.x = element_blank(), axis.title.y = element_blank(),
-          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-          panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-          legend.title = element_text(size=8), #change legend title font size
-          legend.text = element_text(size=6)) +
-    guides(fill = guide_colourbar(title = "Standardized count")) +
-    annotate(geom = 'text', label = "a", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
-
-```
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
- # Loop over loop.vector
-  #i <- 1
-  num_bins <- 8
-  shape_val <- calculate_effective_shape_value(.data = UMAP_data, x = UMAP1, y = UMAP2)
-
-  hexbin_data_object_loop <- extract_hexbin_centroids(UMAP_data, num_bins, shape_val)
-
-  df_bin_centroids_loop <- hexbin_data_object_loop$hexdf_data
-
-  ## Data set with all possible centroids in the hexagonal grid
-
-  full_centroid_df_loop <- generate_full_grid_centroids(df_bin_centroids_loop)
-
-  ## To map hexID to hexbin centroids in the full grid
-
-  vec1 <- stats::setNames(rep("", 2), c("x", "y"))  ## Define column names
-
-  full_grid_with_hexbin_id_loop <- dplyr::bind_rows(vec1)[0, ]
-  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
-    dplyr::mutate_if(is.character, as.numeric)
-
-  for(i in 1:length(sort(unique(full_centroid_df_loop$y)))){
-
-    ## Filter the data set with specific y value
-    specific_y_val_df_loop <- full_centroid_df_loop |>
-      dplyr::filter(y == sort(unique(full_centroid_df_loop$y))[i])
-
-    ordered_x_df_loop <- specific_y_val_df_loop |>
-      dplyr::arrange(x)
-
-    full_grid_with_hexbin_id_loop <- dplyr::bind_rows(full_grid_with_hexbin_id_loop, ordered_x_df_loop)
-
-  }
-
-  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
-    dplyr::mutate(hexID = row_number())
-
-  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
-    dplyr::rename("c_x" = "x",
-           "c_y" = "y")
-
-  full_grid_with_hexbin_id_loop <- dplyr::full_join(full_grid_with_hexbin_id_loop, df_bin_centroids_loop, by = c("hexID" = "hexID")) |>
-    dplyr::select(-c(x, y))
-
-  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
-      dplyr::mutate(std_counts = counts/max(counts, na.rm = TRUE))
-
-  ## Generate all coordinates of hexagons
-  hex_grid_loop <- full_hex_grid(full_centroid_df_loop)
-
-  full_grid_with_polygon_id_df_loop <- map_polygon_id(full_grid_with_hexbin_id_loop, hex_grid_loop)
-
-  full_grid_with_hexbin_id_rep_loop <- full_grid_with_polygon_id_df_loop |>
-  dplyr::slice(rep(1:n(), each = 6)) |>
-  dplyr::arrange(polygon_id)
-
-  hex_full_count_df_loop <- dplyr::bind_cols(hex_grid_loop, full_grid_with_hexbin_id_rep_loop)
-
-  p2 <-  ggplot(data = hex_full_count_df_loop, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
-    geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-        theme_void() +
-    theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-          axis.title.x = element_blank(), axis.title.y = element_blank(),
-          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-          panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-          legend.title = element_text(size=8), #change legend title font size
-          legend.text = element_text(size=6)) +
-    guides(fill = guide_colourbar(title = "Standardized count")) +
-    annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
-
-```
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
- # Loop over loop.vector
-  #i <- 1
-  num_bins <- 15
-  shape_val <- calculate_effective_shape_value(.data = UMAP_data, x = UMAP1, y = UMAP2)
-
-  hexbin_data_object_loop <- extract_hexbin_centroids(UMAP_data, num_bins, shape_val)
-
-  df_bin_centroids_loop <- hexbin_data_object_loop$hexdf_data
-
-  ## Data set with all possible centroids in the hexagonal grid
-
-  full_centroid_df_loop <- generate_full_grid_centroids(df_bin_centroids_loop)
-
-  ## To map hexID to hexbin centroids in the full grid
-
-  vec1 <- stats::setNames(rep("", 2), c("x", "y"))  ## Define column names
-
-  full_grid_with_hexbin_id_loop <- dplyr::bind_rows(vec1)[0, ]
-  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
-    dplyr::mutate_if(is.character, as.numeric)
-
-  for(i in 1:length(sort(unique(full_centroid_df_loop$y)))){
-
-    ## Filter the data set with specific y value
-    specific_y_val_df_loop <- full_centroid_df_loop |>
-      dplyr::filter(y == sort(unique(full_centroid_df_loop$y))[i])
-
-    ordered_x_df_loop <- specific_y_val_df_loop |>
-      dplyr::arrange(x)
-
-    full_grid_with_hexbin_id_loop <- dplyr::bind_rows(full_grid_with_hexbin_id_loop, ordered_x_df_loop)
-
-  }
-
-  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
-    dplyr::mutate(hexID = row_number())
-
-  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
-    dplyr::rename("c_x" = "x",
-           "c_y" = "y")
-
-  full_grid_with_hexbin_id_loop <- dplyr::full_join(full_grid_with_hexbin_id_loop, df_bin_centroids_loop, by = c("hexID" = "hexID")) |>
-    dplyr::select(-c(x, y))
-
-  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
-      dplyr::mutate(std_counts = counts/max(counts, na.rm = TRUE))
-
-  ## Generate all coordinates of hexagons
-  hex_grid_loop <- full_hex_grid(full_centroid_df_loop)
-
-  full_grid_with_polygon_id_df_loop <- map_polygon_id(full_grid_with_hexbin_id_loop, hex_grid_loop)
-
-  full_grid_with_hexbin_id_rep_loop <- full_grid_with_polygon_id_df_loop |>
-  dplyr::slice(rep(1:n(), each = 6)) |>
-  dplyr::arrange(polygon_id)
-
-  hex_full_count_df_loop <- dplyr::bind_cols(hex_grid_loop, full_grid_with_hexbin_id_rep_loop)
-
-  p3 <-  ggplot(data = hex_full_count_df_loop, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
-        geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-        theme_void() +
-    theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-          axis.title.x = element_blank(), axis.title.y = element_blank(),
-          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-          panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-          legend.title = element_text(size=8), #change legend title font size
-          legend.text = element_text(size=6)) +
-    guides(fill = guide_colourbar(title = "Standardized count")) +
-    annotate(geom = 'text', label = "c", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
-
-```
-
-```{r}
-#| echo: false
-#| label: fig-binsize
-#| fig-pos: H
-#| fig-cap: "Hexbin plots from different number of bins for the same **s_curve_noise_umap** data: (a) b = (4, 8), s = 2.031, (b) b = (9, 20), s = 2.031, and (c) b = (16, 36), s = 2.031.
-The hexbins are colored based on the density of points, with yellow indicating higher point density and darker colors representing lower point density within each bin.
-Does  a value of number of bins exist to effectively represent the low-dimensional data?"
-
-p1 + p2 + p3
-```
--->
-
-### Benchmark value to remove low-density hexagons
-
-Addressing low-density hexagons is a systematic process to handle sparsely represented data in certain regions.
-For each hex bin, we identify the six nearest hex bins using an equal 2D distance metric.
-Then, we calculate the mean density, as outlined in the equations:
-
-$$
-\text{standard count} = \frac{\text{count}}{\text{max count}} 
-$${#eq-equationp2}
-
-$$
-\text{mean density} = \frac{\text{standard count}}{6} 
-$${#eq-equationp3}
-
-The standard count is derived from the number of observations in the hex bins.
-By examining the distribution of mean densities and designating the first quartile as the benchmark value, hex bins with mean densities below this benchmark are removed.
-This process ensures the elimination of regions with insufficient data density, focusing on areas with more significant data representation and preserving the overall structure in the low-dimensional space.
-
-<!--
-In the step of addressing low-density hexagons, which can arise from sparsely represented data in certain regions, we employ a systematic strategy.
-The goal is to ensure a more comprehensive coverage of the data by removing hexagons with low data density.
-To achieve this, we initiate the process by identifying, for each hex bin, the six nearest hex bins based on an equal 2D distance metric.
-Following this, we calculate the mean density (see @eq-equationp3), as outlined in the equation:
-
-$$
- \text{standard count} = \frac{\text{count}}{\text{max count}}
-$$ {#eq-equationp2}
-
-$$
- \text{mean density} = \frac{\text{standard count}}{6}
-$$ {#eq-equationp3}
-
-The standard count is derived from the number of observations in the hex bins (see @eq-equationp2).
-Next, we examine the distribution of mean densities across all hex bins and designate the first quartile as the benchmark value for removing low-density hexagons.
-Finally, hex bins with mean densities below this benchmark value are removed from consideration.
-This meticulous procedure ensures the elimination of regions with inadequate data density, allowing the focus to shift to areas with more significant data representation.
-The result is the preservation of the overall structure of the data in the low-dimensional space, as illustrated in @fig-bintorm (c).
--->
-
-```{r}
-#| echo: false
-
-df_bin_centroids <- df_bin_centroids |>
-  dplyr::mutate(ID = row_number())
-
-# To store mean densities of hexagons
-mean_density_vec <- c()
-
-for (i in 1:length(df_bin_centroids$hexID)) {
-
-  df_bin_centroids_coordinates_spec_bin <- df_bin_centroids |>
-    filter(hexID == df_bin_centroids$hexID[i])
-
-  available_near_check <- df_bin_centroids |>
-    dplyr::filter((hexID == (df_bin_centroids$hexID[i] + 1)) | (hexID == (df_bin_centroids$hexID[i] - 1))) |>
-    head(1)
-
-  if (NROW(available_near_check) == 0) {
-
-    df_bin_centroids_coordinates_spec_bin_near1 <- df_bin_centroids |>
-      filter((hexID == (df_bin_centroids$hexID[i] + (num_bins_x + 1))) | (hexID == (df_bin_centroids$hexID[i] + num_bins_x)) | (hexID == (df_bin_centroids$hexID[i] - (num_bins_x + 1))) | (hexID == (df_bin_centroids$hexID[i] - num_bins_x))) |>
-      head(1)
-
-  } else {
-
-    df_bin_centroids_coordinates_spec_bin_near1 <- df_bin_centroids |>
-      filter((hexID == (df_bin_centroids$hexID[i] + 1)) | (hexID == (df_bin_centroids$hexID[i] - 1))) |>
-      head(1)
-
-  }
-
-  near_df_1 <- dplyr::bind_rows(df_bin_centroids_coordinates_spec_bin, df_bin_centroids_coordinates_spec_bin_near1)
-
-  start <- unlist(near_df_1[1, c("x","y")])
-  end <- unlist(near_df_1[2, c("x","y")])
-  nearest_dist <- sqrt(sum((start - end)^2)) # Distance to nearest centroid
-
-  df_bin_centroids$distance <- lapply(seq(nrow(df_bin_centroids)), function(x) {
-    start <- unlist(df_bin_centroids[(df_bin_centroids_coordinates_spec_bin |> pull(ID)), c("x","y")])
-    end <- unlist(df_bin_centroids[x, c("x","y")])
-    sqrt(sum((start - end)^2))})
-
-  df_bin_centroids <- df_bin_centroids %>%
-    dplyr::select(names(df_bin_centroids), "distance")
-
-  df_bin_centroids$distance <- round(unlist(df_bin_centroids$distance), 7)
-
-  neighbor_df <- df_bin_centroids |>
-    filter(distance == round(nearest_dist, 7))
-
-  mean_density <- neighbor_df |>
-    pull(std_counts) |>
-    sum()/6 ## The reason to take the mean is to check the density in a considerable amount
-
-  mean_density_vec <- append(mean_density_vec, mean_density)
-
-}
-
-df_bin_centroids <- df_bin_centroids |>
-  dplyr::mutate(mean_density = mean_density_vec)
-
-cell_count_plot <- ggplot(df_bin_centroids, aes(x = reorder(as.factor(hexID), -mean_density), y = mean_density)) +
-  geom_quasirandom() + xlab("hexagonal id") + ylab("Standardized cell count") +
-  geom_hline(yintercept = 0.255, colour = "#de2d26") +
-  theme(axis.text = element_text(size = 5),
-        axis.title = element_text(size = 7),
-        axis.text.x = element_text(angle = 90))
-
-cell_count_plot
-```
-
-```{r}
-#| echo: false
-
-hexgrid_with_points <- hexgrid_with_points +
-  coord_equal()
-
-hex_full_count_df <- full_join(hex_full_count_df, df_bin_centroids |> dplyr::select(hexID, mean_density))
-
-hex_full_count_df <- hex_full_count_df |>
-  dplyr::mutate(type = if_else(mean_density < 0.255, "low", "high"))
-
-hex_full_count_df_filtered <- hex_full_count_df |>
-  dplyr::filter(type == "low")
-
-hex_grid_centers_rm_l_n <- ggplot(data = hex_full_count_df, aes(x = x, y = y)) + 
-  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts))   +
-  # geom_polygon(data = hex_full_count_df_filtered, color = "black", aes(group = polygon_id), fill = "red") +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  theme_light() +
-    #coord_equal() +
-    theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
-          axis.title.x = element_blank(), axis.title.y = element_blank(),
-          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-          panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
-    )
-
-hex_grid_centers_rm_l_n <- hex_grid_centers_rm_l_n +
-  geom_text(data = hex_full_count_df_filtered, color = "black", aes(x = c_x, y = c_y, group = polygon_id, label = "X", size = 0.5)) +
-  coord_equal()
-
-hexgrid_with_points + hex_grid_centers_rm_l_n +
-  plot_layout(ncol = 2)
-
-```
-
-### Benchmark value to remove long edges
-
-<!--The removal of long edges is a critical step to create a smoother representation  by iteratively eliminating hexagons with excessive distances between centroids.
-This process eliminates outliers and noise while preserving essential local relationships within the data.
-To achieve this, distances between vertices are sorted, and unique distance values are extracted.
-By setting a threshold based on the largest difference between consecutive distance values, long edges are identified and removed.
-This refinement step contributes to enhancing the quality of the triangular mesh, ensuring a more accurate representation of the data structure.
--->
-
-Creating a smooth surfaces in 2D involves experimenting with various benchmark values.
-The challenge lies in determining the most effective benchmark value.
-This is precisely why the default value was introduced, as outlined in @sec-rmlgdefault, serving as a fundamental reference.
-To assess its effectiveness, a visual inspection of the triangular mesh is conducted to verify whether it successfully eliminates long edges.
-If it does, proceeding with that value is a effective choice; if not, it signals the initiation of the tuning process.
-@fig-diffrmlg shows how the 2D triangular meshes look after removing long edges with different benchmark values.
-Higher values tend to retain long edges, while lower values may remove not only long edges but also shorter ones.
-
-```{r}
-#| warning: false
-#| echo: false
-
-## Import data
-
-UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap.rds")
-
-## UMAP
-
-num_bins_umap_s_curve <- calculate_effective_x_bins(.data = UMAP_s_curve, x = UMAP1,
-                           cell_area = 1)
-shape_val_umap_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
-                                                          x = UMAP1, y = UMAP2) ## 1.259938
-## To extract bin centroids
-hexbin_data_object_umap_s_curve <- extract_hexbin_centroids(nldr_df = UMAP_s_curve, num_bins = num_bins_umap_s_curve, shape_val = shape_val_umap_s_curve, x = UMAP1, y = UMAP2)
-
-df_bin_centroids_umap_s_curve <- hexbin_data_object_umap_s_curve$hexdf_data
-
-UMAP_data_with_hb_id_s_curve <- UMAP_s_curve |>
-  dplyr::mutate(hb_id = hexbin_data_object_umap_s_curve$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_umap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), UMAP_data_with_hb_id_s_curve)
-
-## Averaged on high-D
-df_bin_umap_s_curve <- avg_highD_data(.data = df_all_umap_s_curve)
-
-## Triangulate bin centroids
-tr1_object_umap_s_curve <- triangulate_bin_centroids(df_bin_centroids_umap_s_curve, x, y)
-tr_from_to_df_umap_s_curve <- generate_edge_info(triangular_object = tr1_object_umap_s_curve)
-
-# ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
-#   geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
-#   geom_point(size = 2, colour = "#33a02c") +
-#   coord_equal()
-
-## Compute 2D distances
-distance_umap_s_curve <- cal_2D_dist(.data = tr_from_to_df_umap_s_curve)
-
-## To find the benchmark value
-benchmark_umap_s_curve <- find_benchmark_value(.data = distance_umap_s_curve, distance_col = distance)
-
-# colour_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve,
-#                   triangular_object = tr1_object_umap_s_curve, distance_col = distance)
-
-trimesh_umap_s_curve <- ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
-  geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
-  geom_point(size = 2, colour = "#33a02c") +
-  coord_equal()
-
-trimesh_umap_s_curve <- trimesh_umap_s_curve +
-  #ggtitle("(a)") +
-  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  theme_light() +
-  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
-  ) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
-# theme(axis.text = element_text(size = 5),
-#       axis.title = element_text(size = 7))
-
-trimesh_gr_umap_s_curve <- colour_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve, triangular_object = tr1_object_umap_s_curve, distance_col = distance)
-
-trimesh_gr_umap_s_curve <- trimesh_gr_umap_s_curve +
-  geom_point(size = 2, colour = "#33a02c") +
-  #ggtitle("(b)") +
-  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  theme_light() +
-  #coord_equal() +
-  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
-  ) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-trimesh_removed_umap_s_curve <- remove_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve,
-                                                  triangular_object = tr1_object_umap_s_curve, distance_col = distance)
-
-trimesh_removed_umap_s_curve <- trimesh_removed_umap_s_curve +
-  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  # theme(axis.text = element_text(size = 5),
-  #       axis.title = element_text(size = 7)) +
-  geom_point(colour = "#33a02c", size = 0.5) +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-trimesh_removed_umap_s_curve1 <- remove_long_edges(.data = distance_umap_s_curve, benchmark_value = 0.8794136,
-                                                  triangular_object = tr1_object_umap_s_curve, distance_col = distance)
-
-trimesh_removed_umap_s_curve1 <- trimesh_removed_umap_s_curve1 +
-  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  # theme(axis.text = element_text(size = 5),
-  #       axis.title = element_text(size = 7)) +
-  geom_point(colour = "#33a02c", size = 0.5) +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-trimesh_removed_umap_s_curve2 <- remove_long_edges(.data = distance_umap_s_curve, benchmark_value = 5.0891822,
-                                                  triangular_object = tr1_object_umap_s_curve, distance_col = distance)
-
-trimesh_removed_umap_s_curve2 <- trimesh_removed_umap_s_curve2 +
-  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  # theme(axis.text = element_text(size = 5),
-  #       axis.title = element_text(size = 7)) +
-  geom_point(colour = "#33a02c", size = 0.5) +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "What is the effective benchmark value to remove long edges? (a) benchmark value = $0.8794136$, (b) benchmark value = $1.696$ (default), and (c) benchmark value = $5.0891822$."
-#| label: fig-diffrmlg
-##| out-width: 100%
-
-trimesh_removed_umap_s_curve1 + trimesh_removed_umap_s_curve + trimesh_removed_umap_s_curve2 +
-  plot_layout(guides='collect', ncol = 3) &
-  theme(legend.position='bottom')
-```
-
-## Different options to construct the 2D model and lift the model into high-D space
-
-Mainly, the hexagonal bin centroids are used to construct 2D model and averaged high-D data is used to lift the model into high-D space.
-As options except bin centroids, bin means can used to construct the 2D model.
-Also, except averaged high-D data, weighted mean can used to lift the model into high-D space.
-If we use bin centers instead of bin centroids, the triangles are not regular.
-
-- Why bin means? because the distribution of points within hexagonals are not always represented by the center.
-
-### Case 1: Construct 2D model with hexbin means and lift the model into high-D space with averaged high-D data
-
-```{r}
-#| warning: false
-#| echo: false
-
-## Import data
-
-UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap.rds")
-
-## UMAP
-
-num_bins_umap_s_curve <- calculate_effective_x_bins(.data = UMAP_s_curve, x = UMAP1,
-                                         cell_area = 1)
-shape_val_umap_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
-                                                          x = UMAP1, y = UMAP2) ## 1.259938
-## To extract bin centroids
-hexbin_data_object_umap_s_curve <- extract_hexbin_mean(nldr_df = UMAP_s_curve, num_bins = num_bins_umap_s_curve, shape_val = shape_val_umap_s_curve, x = UMAP1, y = UMAP2)
-
-df_bin_centroids_umap_s_curve <- hexbin_data_object_umap_s_curve$hexdf_data
-
-## To generate the grid
-hexbin_data_object_n <- extract_hexbin_centroids(nldr_df = UMAP_s_curve, num_bins = num_bins_umap_s_curve, shape_val = shape_val_umap_s_curve, x = UMAP1, y = UMAP2)
-df_bin_centroids_n <- hexbin_data_object_n$hexdf_data
-## Data set with all possible centroids in the hexagonal grid
-full_centroid_df <- generate_full_grid_centroids(df_bin_centroids_n)
-## Generate all coordinates of hexagons
-hex_grid <- full_hex_grid(full_centroid_df)
-
-UMAP_data_with_hb_id_s_curve <- UMAP_s_curve |>
-  dplyr::mutate(hb_id = hexbin_data_object_umap_s_curve$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_umap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), UMAP_data_with_hb_id_s_curve)
-
-## Averaged on high-D
-df_bin_umap_s_curve <- avg_highD_data(.data = df_all_umap_s_curve)
-
-## Triangulate bin centroids
-tr1_object_umap_s_curve <- triangulate_bin_centroids(df_bin_centroids_umap_s_curve, x, y)
-tr_from_to_df_umap_s_curve <- generate_edge_info(triangular_object = tr1_object_umap_s_curve)
-
-# ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
-#   geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
-#   geom_point(size = 2, colour = "#33a02c") +
-#   coord_equal()
-
-## Compute 2D distances
-distance_umap_s_curve <- cal_2D_dist(.data = tr_from_to_df_umap_s_curve)
-
-## To find the benchmark value
-benchmark_umap_s_curve <- find_benchmark_value(.data = distance_umap_s_curve, distance_col = distance)
-benchmark_umap_s_curve <- 2.0384346
-
-# colour_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve,
-#                   triangular_object = tr1_object_umap_s_curve, distance_col = distance)
-
-# trimesh_umap_s_curve <- ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
-#   geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
-#   geom_point(size = 2, colour = "#33a02c") +
-#   coord_equal()
-# 
-# trimesh_umap_s_curve <- trimesh_umap_s_curve +
-#   #ggtitle("(a)") +
-#   xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-#   theme_light() +
-#   theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
-#         axis.title.x = element_blank(), axis.title.y = element_blank(),
-#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
-#   ) +
-#   annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
-# theme(axis.text = element_text(size = 5),
-#       axis.title = element_text(size = 7))
-
-trimesh_removed_umap_s_curve_bin_mean <- remove_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve, triangular_object = tr1_object_umap_s_curve, distance_col = distance)
-
-trimesh_removed_umap_s_curve_bin_mean <- trimesh_removed_umap_s_curve_bin_mean +
-  geom_point(size = 2, colour = "#33a02c") +
-  #ggtitle("(b)") +
-  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  theme_light() +
-  #coord_equal() +
-  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
-  ) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-# ### Define type column
-# df <- df_all_umap_s_curve |>
-#   dplyr::select(tidyselect::starts_with("x")) |>
-#   dplyr::mutate(type = "data") ## original dataset
-# 
-# df_b <- df_bin_umap_s_curve |>
-#   dplyr::filter(hb_id %in% df_bin_centroids_umap_s_curve$hexID) |>
-#   dplyr::select(-hb_id) |>
-#   dplyr::mutate(type = "model") ## Data with summarized mean
-# 
-# df_exe <- dplyr::bind_rows(df_b, df)
-# 
-# distance_df_small_edges <- distance_umap_s_curve %>%
-#   dplyr::filter(distance < benchmark_umap_s_curve)
-# 
-# distance_df_long_edges <- distance_umap_s_curve %>%
-#   dplyr::filter(distance >= benchmark_umap_s_curve)
-# 
-# distance_df_edges <- dplyr::bind_rows(distance_df_small_edges, distance_df_long_edges)
-# 
-# langevitour::langevitour(df_exe[1:(length(df_exe)-1)], lineFrom = distance_df_edges$from , lineTo = distance_df_edges$to, group = df_exe$type, pointSize = 3, levelColors = c("#6a3d9a", "#33a02c"), lineColors = append(rep("black", length(distance_df_small_edges$from)), rep("red", length(distance_df_long_edges$from))))
-# 
-# show_langevitour(df_all_umap_s_curve, df_bin_umap_s_curve, df_bin_centroids_umap_s_curve, benchmark_value = benchmark_umap_s_curve, distance = distance_umap_s_curve, distance_col = distance)
-
-ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
-  geom_point(data = df_bin_centroids_n, aes(x = x, y = y), color = "#33a02c") +
-  geom_point(data = df_bin_centroids_umap_s_curve, aes(x = x, y = y), color = "black")
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-## UMAP
-## Prediction
-
-shape_value_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
-                                                   x = UMAP1, y = UMAP2)
-
-num_bins_vec <- 1:13 ## Number of bins along the x-axis
-
-vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names
-
-eval_data_training <- dplyr::bind_rows(vec)[0, ]
-eval_data_training <- eval_data_training |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(num_bins_vec)) {
-
-  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_s_curve, nldr_df_test = UMAP_s_curve, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
-  pred_df_training <- pred_df_training_object$pred_data
-  centroid_df_training <- pred_df_training_object$df_bin_centroids
-  avg_df_training <- pred_df_training_object$df_bin
-
-  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
-
-  eval_df_training <- eval_df_training |>
-    mutate(num_bins_x = num_bins_vec[i])
-
-  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
-
-}
-
-## Add new column with data types
-
-eval_data_training <- eval_data_training |>
-  mutate(data_type = "training")
-
-MSE_df_2 <- eval_data_training |>
-  dplyr::mutate(method = "UMAP")
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-#MSE_df <- dplyr::bind_rows(MSE_df_1, MSE_df_2, MSE_df_3, MSE_df_4, MSE_df_5)
-MSE_df <- MSE_df_2
-
-MSE_df$method <- factor(MSE_df$method, levels = c("UMAP"))
-
-## To draw with AIC
-aic_plot <- ggplot(MSE_df |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
-                                                                                 y = total_error,
-                                                                                 color = method
+
+### Obtain the mse_df with not duplicated bins
+not_dupli_df <- mse_df_gau |>
+  dplyr::filter(!(num_bins %in% dupli_bins))
+
+### Combine duplicated and not duplicated(corrected) bins dfs
+mse_df_gau <- dplyr::bind_rows(not_dupli_df, duplicate_df) 
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+mse_plot_gau <- ggplot(mse_df_gau, aes(x = num_bins,
+                                       y = log(mse)
 )) +
-  geom_point() +
+  geom_point(size = 0.8) +
   geom_line() +
-  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+   geom_vline(xintercept = 588, linetype="solid",
+                color = "red", size=0.8, alpha = 0.5) +
   theme_light() +
@@
         axis.text = element_text(size = 7)) +
-  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("AIC") +
-  xlab("Total number of bins")
-## Effective number of bins along x-axis
-
-mse_plot <- ggplot(MSE_df |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
-                                       y = total_mse,
-                                       color = method
+  #scale_colour_manual(values = c("#377eb8")) +
+  ylab("MSE") +
+  xlab("total number of bins")
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "Goodness of fit statistics from different NLDR techniques applied to training five spherical Gaussian cluster dataset.
+What is the best NLDR technique to represent the original data in 2D?"
+#| label: fig-diagnosticpltGau
+#| out-width: 100%
+#| fig-pos: H
+
+mse_plot_gau 
+```
+
+<!-- MSE plot to find which is the effective number of bins to use-->
+```{r}
+#| warning: false
+#| echo: false
+#| message: false
+
+## UMAP
+## Prediction
+
+## Import data
+training_data_pbmc <- read_rds("data/pbmc3k/pbmc_pca_50.rds")
+training_data_pbmc <- training_data_pbmc[, 1:9] |>
+  mutate(ID = 1:NROW(training_data_pbmc))
+
+umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_30_min_dist_0.3.rds")
+umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc, 
+                                                                 x = "UMAP1", y = "UMAP2"))) |>
+  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
+                  "UMAP2" = "scaled_UMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(umap_pbmc))
+
+#tot_num_bin_vec <- 2:51
+hex_size_vec <- seq(0.02, 2, by = 0.01)
+
+vec <- stats::setNames(rep("", 7), c("num_bins", "error", "mse", "num_bins_x", "num_bins_y", "hex_size", "num_non_empty_bins"))  ## Define column names
+
+mse_df_pbmc <- dplyr::bind_rows(vec)[0, ]
+mse_df_pbmc <- mse_df_pbmc |>
+  dplyr::mutate_if(is.character, as.numeric)
+
+for (i in 1:length(hex_size_vec)) {
+
+  num_bin_list <- calc_bins(data = umap_pbmc_scaled, 
+            x = "UMAP1", y = "UMAP2", 
+            hex_size = hex_size_vec[i], buffer_x = NA, buffer_y = NA)
+
+  num_bins_x <- num_bin_list$num_x
+  num_bins_y <- num_bin_list$num_y
+
+  model_object <- fit_highd_model( training_data = training_data_pbmc, 
+                                   nldr_df_with_id = umap_pbmc_scaled, 
+                                   x = "UMAP1", y = "UMAP2", 
+                                   num_bins_x = num_bins_x, 
+                                   num_bins_y = num_bins_y, 
+                                   x_start = NA, y_start = NA, 
+                                   buffer_x = NA, buffer_y = NA, 
+                                   hex_size = hex_size_vec[i],
+                                   is_rm_lwd_hex = FALSE, 
+                                   benchmark_to_rm_lwd_hex = NA, 
+                                   col_start_2d = "UMAP", 
+                                   col_start_highd = "PC_")
+
+  centroid_df_training <- model_object$df_bin_centroids
+  avg_df_training <- model_object$df_bin
+
+  pred_emb_list <- predict_emb(test_data = training_data_pbmc, 
+                                  df_bin_centroids = centroid_df_training, 
+                                  df_bin = avg_df_training, type_NLDR = "UMAP")
+
+  pred_df_training <- as.data.frame(do.call(cbind, pred_emb_list))
+
+  eval_list <- gen_summary(test_data = training_data_pbmc, 
+                                  prediction_df = pred_df_training, 
+                                  df_bin = avg_df_training, col_start = "PC_")
+
+  mse_df_pbmc <- mse_df_pbmc |>
+    tibble::add_row(num_bins = num_bins_x * num_bins_y,
+                    error = eval_list$error,
+                    mse = eval_list$mse,
+                    num_bins_x = num_bins_x,
+                    num_bins_y = num_bins_y,
+                    hex_size = hex_size_vec[i],
+                    num_non_empty_bins = NROW(centroid_df_training))
+
+}
+
+## If same total number of bins occurred only select ones with minimum error
+### Obtain duplicate bins
+dupli_bins <- mse_df_pbmc |> 
+  dplyr::count(num_bins) |> 
+  dplyr::filter(n > 1) |> 
+  dplyr::pull(num_bins)
+
+### Group split by duplicated bins
+duplicate_df_list <- mse_df_pbmc |>
+  dplyr::filter(num_bins %in% dupli_bins) |>
+  dplyr::arrange(num_bins) |>
+  dplyr::group_split(num_bins)
+
+### Obtain one row from duplicates which have lowest error and hexsize
+duplicate_df <- data.frame(matrix(nrow = 0, ncol = 0))
+
+for (i in 1:length(duplicate_df_list)) {
+
+  dd <- duplicate_df_list[[i]] |>
+    dplyr::filter(mse == min(duplicate_df_list[[i]]$mse)) |>
+    dplyr::filter(hex_size == min(duplicate_df_list[[i]]$hex_size))
+
+  duplicate_df <- dplyr::bind_rows(duplicate_df, dd)
+
+}
+
+### Obtain the mse_df with not duplicated bins
+not_dupli_df <- mse_df_pbmc |>
+  dplyr::filter(!(num_bins %in% dupli_bins))
+
+### Combine duplicated and not duplicated(corrected) bins dfs
+mse_df_pbmc <- dplyr::bind_rows(not_dupli_df, duplicate_df)
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+mse_plot_pbmc <- ggplot(mse_df_pbmc, aes(x = num_bins,
+                                       y = mse
 )) +
@@
   geom_line() +
+   geom_vline(xintercept = 588, linetype="solid",
+                color = "red", size=0.8, alpha = 0.5) +
   theme_light() +
@@
         axis.text = element_text(size = 7)) +
-  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
   ylab("MSE") +
-  xlab("Total number of bins")
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset.
-What is the most effective NLDR technique to represent the original data in a 2D space?
-What is the effective number of bins in each NLDR technique to create a 2D model?"
-#| label: fig-diagnosticpltScurve
-##| out-width: 100%
-
-trimesh_removed_umap_s_curve_bin_mean + aic_plot + mse_plot +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', ncol = 2) &
-  theme(legend.position='none', plot.tag = element_text(size = 8))
-```
-
-### Case 2: Construct 2D model with hexbin centroid and lift the model into high-D space with weighted mean of high-D data
-
-```{r}
-#| warning: false
-#| echo: false
+  xlab("total number of bins")
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "Goodness of fit statistics from UMAP applied to training PBMC3k dataset.
+What is the effective number of bins in UMAP to create a 2D model?
+The MSE plot have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
+Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error.
+Using the elbow method, when the total number of bins is set to 588, the slope of the MSE plot experiences a sudden and noticeable change, resembling an elbow-like shape.
+This point indicates that adding less bins does not enough to capture the data structure."
+#| label: fig-msepltpbmc
+#| fig-pos: H
+
+mse_plot_pbmc
+```
+
+<!-- MSE plot to find which is the effective number of bins to use-->
+
+```{r}
+#| warning: false
+#| echo: false
+#| message: false
 
 ## Import data
-
-UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap.rds")
-
-## UMAP
-
-num_bins_umap_s_curve <- calculate_effective_x_bins(.data = UMAP_s_curve, x = UMAP1,
-                                         cell_area = 1)
-shape_val_umap_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
-                                                          x = UMAP1, y = UMAP2) ## 1.259938
-## To extract bin centroids
-hexbin_data_object_umap_s_curve <- extract_hexbin_centroids(nldr_df = UMAP_s_curve, num_bins = num_bins_umap_s_curve, shape_val = shape_val_umap_s_curve, x = UMAP1, y = UMAP2)
-
-df_bin_centroids_umap_s_curve <- hexbin_data_object_umap_s_curve$hexdf_data
-
-## Data set with all possible centroids in the hexagonal grid
-full_centroid_df <- generate_full_grid_centroids(df_bin_centroids_umap_s_curve)
-## Generate all coordinates of hexagons
-hex_grid <- full_hex_grid(full_centroid_df)
-
-UMAP_data_with_hb_id_s_curve <- UMAP_s_curve |>
-  dplyr::mutate(hb_id = hexbin_data_object_umap_s_curve$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_umap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), UMAP_data_with_hb_id_s_curve)
-
-## Weighted averaged on high-D
-hb_object <- hexbin_data_object_umap_s_curve$hb_data
-df_bin_umap_s_curve <- weighted_highD_data(training_data, nldr_df_with_id = UMAP_data_with_hb_id_s_curve, hb_object, column_start_text = "x")
-
-## Triangulate bin centroids
-tr1_object_umap_s_curve <- triangulate_bin_centroids(df_bin_centroids_umap_s_curve, x, y)
-tr_from_to_df_umap_s_curve <- generate_edge_info(triangular_object = tr1_object_umap_s_curve)
-
-# ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
-#   geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
-#   geom_point(size = 2, colour = "#33a02c") +
-#   coord_equal()
-
-## Compute 2D distances
-distance_umap_s_curve <- cal_2D_dist(.data = tr_from_to_df_umap_s_curve)
-
-## To find the benchmark value
-benchmark_umap_s_curve <- find_benchmark_value(.data = distance_umap_s_curve, distance_col = distance)
-#benchmark_umap_s_curve <- 2.0384346
-
-# colour_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve,
-#                   triangular_object = tr1_object_umap_s_curve, distance_col = distance)
-
-# trimesh_umap_s_curve <- ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
-#   geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
-#   geom_point(size = 2, colour = "#33a02c") +
-#   coord_equal()
-# 
-# trimesh_umap_s_curve <- trimesh_umap_s_curve +
-#   #ggtitle("(a)") +
-#   xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-#   theme_light() +
-#   theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
-#         axis.title.x = element_blank(), axis.title.y = element_blank(),
-#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
-#   ) +
-#   annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
-# theme(axis.text = element_text(size = 5),
-#       axis.title = element_text(size = 7))
-
-trimesh_removed_umap_s_curve_bin_centroid <- remove_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve, triangular_object = tr1_object_umap_s_curve, distance_col = distance)
-
-trimesh_removed_umap_s_curve_bin_centroid <- trimesh_removed_umap_s_curve_bin_centroid +
-  geom_point(size = 2, colour = "#33a02c") +
-  #ggtitle("(b)") +
-  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  theme_light() +
-  #coord_equal() +
-  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
-  ) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-# ### Define type column
-# df <- df_all_umap_s_curve |>
-#   dplyr::select(tidyselect::starts_with("x")) |>
-#   dplyr::mutate(type = "data") ## original dataset
-# 
-# df_b <- df_bin_umap_s_curve |>
-#   dplyr::filter(hb_id %in% df_bin_centroids_umap_s_curve$hexID) |>
-#   dplyr::select(-hb_id) |>
-#   dplyr::mutate(type = "model") ## Data with summarized mean
-# 
-# df_exe <- dplyr::bind_rows(df_b, df)
-# 
-# distance_df_small_edges <- distance_umap_s_curve %>%
-#   dplyr::filter(distance < benchmark_umap_s_curve)
-# 
-# distance_df_long_edges <- distance_umap_s_curve %>%
-#   dplyr::filter(distance >= benchmark_umap_s_curve)
-# 
-# distance_df_edges <- dplyr::bind_rows(distance_df_small_edges, distance_df_long_edges)
-# 
-# langevitour::langevitour(df_exe[1:(length(df_exe)-1)], lineFrom = distance_df_edges$from , lineTo = distance_df_edges$to, group = df_exe$type, pointSize = 3, levelColors = c("#6a3d9a", "#33a02c"), lineColors = append(rep("black", length(distance_df_small_edges$from)), rep("red", length(distance_df_long_edges$from))))
-# 
-# show_langevitour(df_all_umap_s_curve, df_bin_umap_s_curve, df_bin_centroids_umap_s_curve, benchmark_value = benchmark_umap_s_curve, distance = distance_umap_s_curve, distance_col = distance)
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
+training_data_mnist <- read_rds("data/mnist/mnist_10_pcs_of_digit_1.rds")
+training_data_mnist <- training_data_mnist |>
+  mutate(ID = 1:NROW(training_data_mnist))
+
+pacmap_minst <- read_rds("data/mnist/mnist_pacmap.rds")
+pacmap_minst_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = pacmap_minst, 
+                                    x = "PaCMAP1", y = "PaCMAP2"))) |>
+  dplyr::rename(c("PaCMAP1" = "scaled_PaCMAP1", 
+                  "PaCMAP2" = "scaled_PaCMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(pacmap_minst))
 
@@
 
-shape_value_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
-                                                   x = UMAP1, y = UMAP2)
-
-num_bins_vec <- 1:13 ## Number of bins along the x-axis
-
-vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names
-
-eval_data_training <- dplyr::bind_rows(vec)[0, ]
-eval_data_training <- eval_data_training |>
+#tot_num_bin_vec <- 2:51
+hex_size_vec <- seq(0.02, 2, by = 0.01)
+
+vec <- stats::setNames(rep("", 6), c("num_bins", "aic", "mse", "num_bins_x", "num_bins_y", "hex_size"))  ## Define column names
+
+mse_df_mnist <- dplyr::bind_rows(vec)[0, ]
+mse_df_mnist <- mse_df_mnist |>
   dplyr::mutate_if(is.character, as.numeric)
 
-for (i in 1:length(num_bins_vec)) {
-
-  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_s_curve, nldr_df_test = UMAP_s_curve, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
-  pred_df_training <- pred_df_training_object$pred_data
-  centroid_df_training <- pred_df_training_object$df_bin_centroids
-  avg_df_training <- pred_df_training_object$df_bin
-
-  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
-
-  eval_df_training <- eval_df_training |>
-    mutate(num_bins_x = num_bins_vec[i])
-
-  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
+for (i in 1:length(hex_size_vec)) {
+
+  num_bin_list <- calc_bins(data = pacmap_minst_scaled, 
+            x = "PaCMAP1", y = "PaCMAP2", 
+            hex_size = hex_size_vec[i], buffer_x = NA, buffer_y = NA)
+
+  num_bins_x <- num_bin_list$num_x
+  num_bins_y <- num_bin_list$num_y
+
+  model_object <- fit_highd_model( training_data = training_data_mnist, 
+                                   nldr_df_with_id = pacmap_minst_scaled, 
+                                   x = "PaCMAP1", y = "PaCMAP2", 
+                                   num_bins_x = num_bins_x, 
+                                   num_bins_y = num_bins_y, 
+                                   x_start = NA, y_start = NA, 
+                                   buffer_x = NA, buffer_y = NA, 
+                                   hex_size = hex_size_vec[i],
+                                   is_rm_lwd_hex = FALSE, 
+                                   benchmark_to_rm_lwd_hex = NA, 
+                                   col_start_2d = "PaCMAP", 
+                                   col_start_highd = "PC")
+
+  centroid_df_training <- model_object$df_bin_centroids
+  avg_df_training <- model_object$df_bin
+
+  pred_emb_list <- predict_emb(test_data = training_data_mnist, 
+                                  df_bin_centroids = centroid_df_training, 
+                                  df_bin = avg_df_training, type_NLDR = "PaCMAP")
+
+  pred_df_training <- as.data.frame(do.call(cbind, pred_emb_list))
+
+  eval_list <- gen_summary(test_data = training_data_mnist, 
+                                  prediction_df = pred_df_training, 
+                                  df_bin = avg_df_training, col_start = "PC")
+
+  mse_df_mnist <- mse_df_mnist |>
+    tibble::add_row(num_bins = num_bins_x * num_bins_y,
+                    aic = eval_list$aic,
+                    mse = eval_list$mse,
+                    num_bins_x = num_bins_x,
+                    num_bins_y = num_bins_y,
+                    hex_size = hex_size_vec[i])
 
@@
 
-## Add new column with data types
-
-eval_data_training <- eval_data_training |>
-  mutate(data_type = "training")
-
-MSE_df_2 <- eval_data_training |>
-  dplyr::mutate(method = "UMAP")
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-#MSE_df <- dplyr::bind_rows(MSE_df_1, MSE_df_2, MSE_df_3, MSE_df_4, MSE_df_5)
-MSE_df <- MSE_df_2
-
-MSE_df$method <- factor(MSE_df$method, levels = c("UMAP"))
-
-## To draw with AIC
-aic_plot <- ggplot(MSE_df |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
-                                                                                 y = total_error,
-                                                                                 color = method
+## If same total number of bins occurred only select ones with minimum error
+### Obtain duplicate bins
+dupli_bins <- mse_df_mnist |> 
+  dplyr::count(num_bins) |> 
+  dplyr::filter(n > 1) |> 
+  dplyr::pull(num_bins)
+
+### Group split by duplicated bins
+duplicate_df_list <- mse_df_mnist |>
+  dplyr::filter(num_bins %in% dupli_bins) |>
+  dplyr::arrange(num_bins) |>
+  dplyr::group_split(num_bins)
+
+### Obtain one row from duplicates which have lowest error and hexsize
+duplicate_df <- data.frame(matrix(nrow = 0, ncol = 0))
+
+for (i in 1:length(duplicate_df_list)) {
+
+  dd <- duplicate_df_list[[i]] |>
+    dplyr::filter(mse == min(duplicate_df_list[[i]]$mse)) |>
+    dplyr::filter(hex_size == min(duplicate_df_list[[i]]$hex_size))
+
+  duplicate_df <- dplyr::bind_rows(duplicate_df, dd)
+
+}
+
+### Obtain the mse_df with not duplicated bins
+not_dupli_df <- mse_df_mnist |>
+  dplyr::filter(!(num_bins %in% dupli_bins))
+
+### Combine duplicated and not duplicated(corrected) bins dfs
+mse_df_mnist <- dplyr::bind_rows(not_dupli_df, duplicate_df)
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+mse_plot_mnist <- ggplot(mse_df_mnist, aes(x = num_bins,
+                                       y = mse
 )) +
@@
   geom_line() +
-  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
+   geom_vline(xintercept = 180, linetype="solid",
+                color = "red", size=0.8, alpha = 0.5) +
   theme_light() +
@@
         axis.text = element_text(size = 7)) +
-  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("AIC") +
-  xlab("Total number of bins")
-## Effective number of bins along x-axis
-
-mse_plot <- ggplot(MSE_df |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
-                                       y = total_mse,
-                                       color = method
-)) +
-  geom_point() +
-  geom_line() +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("MSE") +
-  xlab("Total number of bins")
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset.
-What is the most effective NLDR technique to represent the original data in a 2D space?
-What is the effective number of bins in each NLDR technique to create a 2D model?"
-#| label: fig-diagnosticpltScurve2
-##| out-width: 100%
-
-trimesh_removed_umap_s_curve_bin_centroid + aic_plot + mse_plot +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', ncol = 2) &
-  theme(legend.position='none', plot.tag = element_text(size = 8))
-```
-
-### Case 3: Construct 2D model with hexbin means and lift the model into high-D space with weighted mean of high-D data
-
-```{r}
-#| warning: false
-#| echo: false
-
-## Import data
-
-UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap.rds")
-
-## UMAP
-
-num_bins_umap_s_curve <- calculate_effective_x_bins(.data = UMAP_s_curve, x = UMAP1,
-                                         cell_area = 1)
-shape_val_umap_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
-                                                          x = UMAP1, y = UMAP2) ## 1.259938
-## To extract bin centroids
-hexbin_data_object_umap_s_curve <- extract_hexbin_mean(nldr_df = UMAP_s_curve, num_bins = num_bins_umap_s_curve, shape_val = shape_val_umap_s_curve, x = UMAP1, y = UMAP2)
-
-df_bin_centroids_umap_s_curve <- hexbin_data_object_umap_s_curve$hexdf_data
-
-## To generate the grid
-hexbin_data_object_n <- extract_hexbin_centroids(nldr_df = UMAP_s_curve, num_bins = num_bins_umap_s_curve, shape_val = shape_val_umap_s_curve, x = UMAP1, y = UMAP2)
-df_bin_centroids_n <- hexbin_data_object_n$hexdf_data
-## Data set with all possible centroids in the hexagonal grid
-full_centroid_df <- generate_full_grid_centroids(df_bin_centroids_n)
-## Generate all coordinates of hexagons
-hex_grid <- full_hex_grid(full_centroid_df)
-
-UMAP_data_with_hb_id_s_curve <- UMAP_s_curve |>
-  dplyr::mutate(hb_id = hexbin_data_object_umap_s_curve$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_umap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), UMAP_data_with_hb_id_s_curve)
-
-## Weighted averaged on high-D
-hb_object <- hexbin_data_object_umap_s_curve$hb_data
-df_bin_umap_s_curve <- weighted_highD_data(training_data, nldr_df_with_id = UMAP_data_with_hb_id_s_curve, hb_object, column_start_text = "x")
-
-## Triangulate bin centroids
-tr1_object_umap_s_curve <- triangulate_bin_centroids(df_bin_centroids_umap_s_curve, x, y)
-tr_from_to_df_umap_s_curve <- generate_edge_info(triangular_object = tr1_object_umap_s_curve)
-
-# ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
-#   geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
-#   geom_point(size = 2, colour = "#33a02c") +
-#   coord_equal()
-
-## Compute 2D distances
-distance_umap_s_curve <- cal_2D_dist(.data = tr_from_to_df_umap_s_curve)
-
-## To find the benchmark value
-benchmark_umap_s_curve <- find_benchmark_value(.data = distance_umap_s_curve, distance_col = distance)
-benchmark_umap_s_curve <- 2.0384346
-
-# colour_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve,
-#                   triangular_object = tr1_object_umap_s_curve, distance_col = distance)
-
-# trimesh_umap_s_curve <- ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
-#   geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
-#   geom_point(size = 2, colour = "#33a02c") +
-#   coord_equal()
-# 
-# trimesh_umap_s_curve <- trimesh_umap_s_curve +
-#   #ggtitle("(a)") +
-#   xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-#   theme_light() +
-#   theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
-#         axis.title.x = element_blank(), axis.title.y = element_blank(),
-#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
-#   ) +
-#   annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
-# theme(axis.text = element_text(size = 5),
-#       axis.title = element_text(size = 7))
-
-trimesh_removed_umap_s_curve_bin_mean <- remove_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve, triangular_object = tr1_object_umap_s_curve, distance_col = distance)
-
-trimesh_removed_umap_s_curve_bin_mean <- trimesh_removed_umap_s_curve_bin_mean +
-  geom_point(size = 2, colour = "#33a02c") +
-  #ggtitle("(b)") +
-  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  theme_light() +
-  #coord_equal() +
-  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
-  ) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-# ### Define type column
-# df <- df_all_umap_s_curve |>
-#   dplyr::select(tidyselect::starts_with("x")) |>
-#   dplyr::mutate(type = "data") ## original dataset
-# 
-# df_b <- df_bin_umap_s_curve |>
-#   dplyr::filter(hb_id %in% df_bin_centroids_umap_s_curve$hexID) |>
-#   dplyr::select(-hb_id) |>
-#   dplyr::mutate(type = "model") ## Data with summarized mean
-# 
-# df_exe <- dplyr::bind_rows(df_b, df)
-# 
-# distance_df_small_edges <- distance_umap_s_curve %>%
-#   dplyr::filter(distance < benchmark_umap_s_curve)
-# 
-# distance_df_long_edges <- distance_umap_s_curve %>%
-#   dplyr::filter(distance >= benchmark_umap_s_curve)
-# 
-# distance_df_edges <- dplyr::bind_rows(distance_df_small_edges, distance_df_long_edges)
-# 
-# langevitour::langevitour(df_exe[1:(length(df_exe)-1)], lineFrom = distance_df_edges$from , lineTo = distance_df_edges$to, group = df_exe$type, pointSize = 3, levelColors = c("#6a3d9a", "#33a02c"), lineColors = append(rep("black", length(distance_df_small_edges$from)), rep("red", length(distance_df_long_edges$from))))
-# 
-# show_langevitour(df_all_umap_s_curve, df_bin_umap_s_curve, df_bin_centroids_umap_s_curve, benchmark_value = benchmark_umap_s_curve, distance = distance_umap_s_curve, distance_col = distance)
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-## UMAP
-## Prediction
-
-shape_value_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
-                                                   x = UMAP1, y = UMAP2)
-
-num_bins_vec <- 1:13 ## Number of bins along the x-axis
-
-vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names
-
-eval_data_training <- dplyr::bind_rows(vec)[0, ]
-eval_data_training <- eval_data_training |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(num_bins_vec)) {
-
-  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_s_curve, nldr_df_test = UMAP_s_curve, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
-  pred_df_training <- pred_df_training_object$pred_data
-  centroid_df_training <- pred_df_training_object$df_bin_centroids
-  avg_df_training <- pred_df_training_object$df_bin
-
-  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
-
-  eval_df_training <- eval_df_training |>
-    mutate(num_bins_x = num_bins_vec[i])
-
-  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
-
-}
-
-## Add new column with data types
-
-eval_data_training <- eval_data_training |>
-  mutate(data_type = "training")
-
-MSE_df_2 <- eval_data_training |>
-  dplyr::mutate(method = "UMAP")
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-#MSE_df <- dplyr::bind_rows(MSE_df_1, MSE_df_2, MSE_df_3, MSE_df_4, MSE_df_5)
-MSE_df <- MSE_df_2
-
-MSE_df$method <- factor(MSE_df$method, levels = c("UMAP"))
-
-## To draw with AIC
-aic_plot <- ggplot(MSE_df |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
-                                                                                 y = total_error,
-                                                                                 color = method
-)) +
-  geom_point() +
-  geom_line() +
-  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("AIC") +
-  xlab("Total number of bins")
-## Effective number of bins along x-axis
-
-mse_plot <- ggplot(MSE_df |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
-                                       y = total_mse,
-                                       color = method
-)) +
-  geom_point() +
-  geom_line() +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("MSE") +
-  xlab("Total number of bins")
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset.
-What is the most effective NLDR technique to represent the original data in a 2D space?
-What is the effective number of bins in each NLDR technique to create a 2D model?"
-#| label: fig-diagnosticpltScurve3
-##| out-width: 100%
-
-trimesh_removed_umap_s_curve_bin_mean + aic_plot + mse_plot +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', ncol = 2) &
-  theme(legend.position='none', plot.tag = element_text(size = 8))
-```
-
-## How model perform less than and greater than the selected number of bins (UMAP with S-curve)
-
-<!--
-```{r}
-#| warning: false
-#| echo: false
-
-## UMAP with n_neighbor: 7
-## Prediction
-
-UMAP_data <- read_rds("data/s_curve/s_curve_umap_7.rds")
-predict_UMAP_df <- read_rds("data/s_curve/s_curve_umap_7_predict.rds")
-
-shape_value_curve <- calculate_effective_shape_value(.data = UMAP_data,
-                                                     x = UMAP1, y = UMAP2)
-
-num_bins_vec <- 1:10 ## Number of bins along the x-axis
-
-vec <- stats::setNames(rep("", 4), c("number_of_bins", "number_of_observations", "total_error", "total_mse"))  ## Define column names
-
-eval_data_test <- dplyr::bind_rows(vec)[0, ]
-eval_data_test <- eval_data_test |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-eval_data_training <- dplyr::bind_rows(vec)[0, ]
-eval_data_training <- eval_data_training |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(num_bins_vec)) {
-
-  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = UMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
-  pred_df_training <- pred_df_training_object$pred_data
-  centroid_df_training <- pred_df_training_object$df_bin_centroids
-  avg_df_training <- pred_df_training_object$df_bin
-
-  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
-
-  pred_df_test_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = predict_UMAP_df, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
-  pred_df_test <- pred_df_test_object$pred_data
-  centroid_df_test <- pred_df_test_object$df_bin_centroids
-  avg_df_test <- pred_df_test_object$df_bin
-
-  eval_df_test <- generate_eval_df(data = data, prediction_df = pred_df_test, df_bin_centroids = centroid_df_test, df_bin = avg_df_test, num_bins = num_bins_vec[i], col_start = "x")
-
-  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
-  eval_data_test <- dplyr::bind_rows(eval_data_test, eval_df_test)
-
-}
-
-## Add new column with data types
-
-eval_data_training <- eval_data_training |>
-  mutate(data_type = "training")
-
-eval_data_test <- eval_data_test |>
-  mutate(data_type = "test")
-
-MSE_df_umap_1 <- bind_rows(eval_data_training, eval_data_test) |>
-  dplyr::mutate(param = "n_neighbors: 7")
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-## UMAP with n_neighbor: 15
-## Prediction
-
-UMAP_data <- read_rds("data/s_curve/s_curve_umap_15.rds")
-predict_UMAP_df <- read_rds("data/s_curve/s_curve_umap_15_predict.rds")
-
-shape_value_curve <- calculate_effective_shape_value(.data = UMAP_data,
-                                                     x = UMAP1, y = UMAP2)
-
-num_bins_vec <- 1:10 ## Number of bins along the x-axis
-
-vec <- stats::setNames(rep("", 4), c("number_of_bins", "number_of_observations", "total_error", "total_mse"))  ## Define column names
-
-eval_data_test <- dplyr::bind_rows(vec)[0, ]
-eval_data_test <- eval_data_test |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-eval_data_training <- dplyr::bind_rows(vec)[0, ]
-eval_data_training <- eval_data_training |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(num_bins_vec)) {
-
-  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = UMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
-  pred_df_training <- pred_df_training_object$pred_data
-  centroid_df_training <- pred_df_training_object$df_bin_centroids
-  avg_df_training <- pred_df_training_object$df_bin
-
-  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
-
-  pred_df_test_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = predict_UMAP_df, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
-  pred_df_test <- pred_df_test_object$pred_data
-  centroid_df_test <- pred_df_test_object$df_bin_centroids
-  avg_df_test <- pred_df_test_object$df_bin
-
-  eval_df_test <- generate_eval_df(data = data, prediction_df = pred_df_test, df_bin_centroids = centroid_df_test, df_bin = avg_df_test, num_bins = num_bins_vec[i], col_start = "x")
-
-  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
-  eval_data_test <- dplyr::bind_rows(eval_data_test, eval_df_test)
-
-}
-
-## Add new column with data types
-
-eval_data_training <- eval_data_training |>
-  mutate(data_type = "training")
-
-eval_data_test <- eval_data_test |>
-  mutate(data_type = "test")
-
-MSE_df_umap_2 <- bind_rows(eval_data_training, eval_data_test) |>
-  dplyr::mutate(param = "n_neighbors: 15")
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-## UMAP with n_neighbor: 32
-## Prediction
-
-UMAP_data <- read_rds("data/s_curve/s_curve_umap_32.rds")
-predict_UMAP_df <- read_rds("data/s_curve/s_curve_umap_32_predict.rds")
-
-shape_value_curve <- calculate_effective_shape_value(.data = UMAP_data,
-                                                     x = UMAP1, y = UMAP2)
-
-num_bins_vec <- 1:10 ## Number of bins along the x-axis
-
-vec <- stats::setNames(rep("", 4), c("number_of_bins", "number_of_observations", "total_error", "total_mse"))  ## Define column names
-
-eval_data_test <- dplyr::bind_rows(vec)[0, ]
-eval_data_test <- eval_data_test |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-eval_data_training <- dplyr::bind_rows(vec)[0, ]
-eval_data_training <- eval_data_training |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(num_bins_vec)) {
-
-  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = UMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
-  pred_df_training <- pred_df_training_object$pred_data
-  centroid_df_training <- pred_df_training_object$df_bin_centroids
-  avg_df_training <- pred_df_training_object$df_bin
-
-  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
-
-  pred_df_test_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = predict_UMAP_df, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
-  pred_df_test <- pred_df_test_object$pred_data
-  centroid_df_test <- pred_df_test_object$df_bin_centroids
-  avg_df_test <- pred_df_test_object$df_bin
-
-  eval_df_test <- generate_eval_df(data = data, prediction_df = pred_df_test, df_bin_centroids = centroid_df_test, df_bin = avg_df_test, num_bins = num_bins_vec[i], col_start = "x")
-
-  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
-  eval_data_test <- dplyr::bind_rows(eval_data_test, eval_df_test)
-
-}
-
-## Add new column with data types
-
-eval_data_training <- eval_data_training |>
-  mutate(data_type = "training")
-
-eval_data_test <- eval_data_test |>
-  mutate(data_type = "test")
-
-MSE_df_umap_3 <- bind_rows(eval_data_training, eval_data_test) |>
-  dplyr::mutate(param = "n_neighbors: 32")
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-## UMAP with n_neighbor: 50
-## Prediction
-
-UMAP_data <- read_rds("data/s_curve/s_curve_umap.rds")
-predict_UMAP_df <- read_rds("data/s_curve/s_curve_umap_predict.rds")
-
-shape_value_curve <- calculate_effective_shape_value(.data = UMAP_data,
-                                                     x = UMAP1, y = UMAP2)
-
-num_bins_vec <- 1:10 ## Number of bins along the x-axis
-
-vec <- stats::setNames(rep("", 4), c("number_of_bins", "number_of_observations", "total_error", "total_mse"))  ## Define column names
-
-eval_data_test <- dplyr::bind_rows(vec)[0, ]
-eval_data_test <- eval_data_test |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-eval_data_training <- dplyr::bind_rows(vec)[0, ]
-eval_data_training <- eval_data_training |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(num_bins_vec)) {
-
-  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = UMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
-  pred_df_training <- pred_df_training_object$pred_data
-  centroid_df_training <- pred_df_training_object$df_bin_centroids
-  avg_df_training <- pred_df_training_object$df_bin
-
-  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")
-
-  pred_df_test_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = predict_UMAP_df, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
-  pred_df_test <- pred_df_test_object$pred_data
-  centroid_df_test <- pred_df_test_object$df_bin_centroids
-  avg_df_test <- pred_df_test_object$df_bin
-
-  eval_df_test <- generate_eval_df(data = data, prediction_df = pred_df_test, df_bin_centroids = centroid_df_test, df_bin = avg_df_test, num_bins = num_bins_vec[i], col_start = "x")
-
-  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
-  eval_data_test <- dplyr::bind_rows(eval_data_test, eval_df_test)
-
-}
-
-## Add new column with data types
-
-eval_data_training <- eval_data_training |>
-  mutate(data_type = "training")
-
-eval_data_test <- eval_data_test |>
-  mutate(data_type = "test")
-
-MSE_df_umap_4 <- bind_rows(eval_data_training, eval_data_test) |>
-  dplyr::mutate(param = "n_neighbors: 50")
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-MSE_df_umap <- dplyr::bind_rows(MSE_df_umap_1, MSE_df_umap_2, MSE_df_umap_3, MSE_df_umap_4)
-MSE_df_umap$param <- factor(MSE_df_umap$param, levels = c("n_neighbors: 7", "n_neighbors: 15", "n_neighbors: 32", "n_neighbors: 50"))
-
-## To draw with AIC
-aic_plot_param <- ggplot(MSE_df_umap |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
-                                                                         y = total_error,
-                                                                         color = param
-)) +
-  geom_point() +
-  geom_line() +
-  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  theme_light() +
-  theme(legend.title = element_blank(), legend.text = element_text(size=7), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  scale_color_manual(values = c("#8dd3c7", "#a65628", "#f781bf", "#999999")) +
-  ylab("AIC") +
-  xlab("Total number of bins")
-## Effective number of bins along x-axis
-
-mse_plot_param_training <- ggplot(MSE_df_umap |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
-                               y = total_mse,
-                               color = param
-)) +
-  geom_point() +
-  geom_line() +
-  theme_light() +
-  theme(legend.title = element_blank(), legend.text = element_text(size=7), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  scale_color_manual(values = c("#8dd3c7", "#a65628", "#f781bf", "#999999")) +
-  ylab("MSE") +
-  xlab("Total number of bins")
-
-mse_plot_param_test <- ggplot(MSE_df_umap |> dplyr::filter(data_type == "test"), aes(x = number_of_bins,
-                                                                                             y = total_mse,
-                                                                                             color = param
-)) +
-  geom_point() +
-  geom_line() +
-  theme_light() +
-  theme(legend.title = element_blank(), legend.text = element_text(size=7), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  scale_color_manual(values = c("#8dd3c7", "#a65628", "#f781bf", "#999999")) +
-  ylab("MSE") +
-  xlab("Total number of bins")
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "Goodness of fit statistics from different n_neighbors parameter of UMAP applied to training S-curve dataset.
-What is the best parameter choice in UMAP to represent the original data in 2D?"
-#| label: fig-diagnosticpltDiffParam
-#| out-width: 100%
-
-aic_plot_param + mse_plot_param_training + mse_plot_param_test  +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', ncol = 3) &
-  theme(legend.position='bottom', plot.tag = element_text(size = 8))
-```
--->
-
-<!--model for S-curve with 5 methods-->
-<!--
-```{r}
-#| warning: false
-#| echo: false
-
-## Import data
-
-tSNE_s_curve <- read_rds("data/s_curve/s_curve_tsne_27.rds")
-UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap.rds")
-PHATE_s_curve <- read_rds("data/s_curve/s_curve_phate.rds")
-TriMAP_s_curve <- read_rds("data/s_curve/s_curve_trimap.rds")
-PaCMAP_s_curve <- read_rds("data/s_curve/s_curve_pacmap.rds")
-
-## tSNE
-
-num_bins_tsne_s_curve <- 8
-shape_val_tsne_s_curve <- calculate_effective_shape_value(.data = tSNE_s_curve,
-                                                          x = tSNE1, y = tSNE2) ## 1.259938
-## To extract bin centroids
-hexbin_data_object_tsne_s_curve <- extract_hexbin_centroids(nldr_df = tSNE_s_curve, num_bins = num_bins_tsne_s_curve, shape_val = shape_val_tsne_s_curve, x = tSNE1, y = tSNE2)
-
-df_bin_centroids_tsne_s_curve <- hexbin_data_object_tsne_s_curve$hexdf_data
-
-tSNE_data_with_hb_id_s_curve <- tSNE_s_curve |>
-  dplyr::mutate(hb_id = hexbin_data_object_tsne_s_curve$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_tsne_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), tSNE_data_with_hb_id_s_curve)
-
-## Averaged on high-D
-df_bin_tsne_s_curve <- avg_highD_data(.data = df_all_tsne_s_curve)
-
-## Triangulate bin centroids
-tr1_object_tsne_s_curve <- triangulate_bin_centroids(df_bin_centroids_tsne_s_curve, x, y)
-tr_from_to_df_tsne_s_curve <- generate_edge_info(triangular_object = tr1_object_tsne_s_curve)
-
-# ggplot(df_bin_centroids_tsne_s_curve, aes(x = x, y = y)) +
-#   geom_segment(data = tr_from_to_df_tsne_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
-#   geom_point(size = 2, colour = "#33a02c") +
-#   coord_equal()
-
-## Compute 2D distances
-distance_tsne_s_curve <- cal_2d_dist(.data = tr_from_to_df_tsne_s_curve)
-
-## To find the benchmark value
-benchmark_tsne_s_curve <- find_benchmark_value(.data = distance_tsne_s_curve, distance_col = distance)
-
-# colour_long_edges(.data = distance_tsne_s_curve, benchmark_value = benchmark_tsne_s_curve,
-#                   triangular_object = tr1_object_tsne_s_curve, distance_col = distance)
-
-trimesh_removed_tsne_s_curve <- remove_long_edges(.data = distance_tsne_s_curve, benchmark_value = benchmark_tsne_s_curve,
-                                                  triangular_object = tr1_object_tsne_s_curve, distance_col = distance)
-
-trimesh_removed_tsne_s_curve <- trimesh_removed_tsne_s_curve +
-  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  # theme(axis.text = element_text(size = 5),
-  #       axis.title = element_text(size = 7)) +
-  geom_point(colour = "#33a02c", size = 0.5) +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-# tour_tsne_s_curve <- show_langevitour(df_all_tsne_s_curve, df_bin_tsne_s_curve, df_bin_centroids_tsne_s_curve, benchmark_value = benchmark_tsne_s_curve, distance = distance_tsne_s_curve, distance_col = distance)
-
-## UMAP
-
-num_bins_umap_s_curve <- 6
-shape_val_umap_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
-                                                          x = UMAP1, y = UMAP2) ## 1.259938
-## To extract bin centroids
-hexbin_data_object_umap_s_curve <- extract_hexbin_centroids(nldr_df = UMAP_s_curve, num_bins = num_bins_umap_s_curve, shape_val = shape_val_umap_s_curve, x = UMAP1, y = UMAP2)
-
-df_bin_centroids_umap_s_curve <- hexbin_data_object_umap_s_curve$hexdf_data
-
-UMAP_data_with_hb_id_s_curve <- UMAP_s_curve |>
-  dplyr::mutate(hb_id = hexbin_data_object_umap_s_curve$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_umap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), UMAP_data_with_hb_id_s_curve)
-
-## Averaged on high-D
-df_bin_umap_s_curve <- avg_highD_data(.data = df_all_umap_s_curve)
-
-## Triangulate bin centroids
-tr1_object_umap_s_curve <- triangulate_bin_centroids(df_bin_centroids_umap_s_curve, x, y)
-tr_from_to_df_umap_s_curve <- generate_edge_info(triangular_object = tr1_object_umap_s_curve)
-
-# ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
-#   geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
-#   geom_point(size = 2, colour = "#33a02c") +
-#   coord_equal()
-
-## Compute 2D distances
-distance_umap_s_curve <- cal_2d_dist(.data = tr_from_to_df_umap_s_curve)
-
-## To find the benchmark value
-benchmark_umap_s_curve <- find_benchmark_value(.data = distance_umap_s_curve, distance_col = distance)
-
-# colour_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve,
-#                   triangular_object = tr1_object_umap_s_curve, distance_col = distance)
-
-trimesh_removed_umap_s_curve <- remove_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve,
-                                     triangular_object = tr1_object_umap_s_curve, distance_col = distance)
-
-trimesh_removed_umap_s_curve <- trimesh_removed_umap_s_curve +
-  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  # theme(axis.text = element_text(size = 5),
-  #       axis.title = element_text(size = 7)) +
-  geom_point(colour = "#33a02c", size = 0.5) +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-# tour_umap_s_curve <- show_langevitour(df_all_umap_s_curve, df_bin_umap_s_curve, df_bin_centroids_umap_s_curve, benchmark_value = benchmark_umap_s_curve, distance = distance_umap_s_curve, distance_col = distance)
-
-## TriMAP
-
-num_bins_trimap_s_curve <- 6
-shape_val_trimap_s_curve <- calculate_effective_shape_value(.data = TriMAP_s_curve,
-                                                            x = TriMAP1, y = TriMAP2) ## 1.259938
-## To extract bin centroids
-hexbin_data_object_trimap_s_curve <- extract_hexbin_centroids(nldr_df = TriMAP_s_curve, num_bins = num_bins_trimap_s_curve, shape_val = shape_val_trimap_s_curve, x = TriMAP1, y = TriMAP2)
-
-df_bin_centroids_trimap_s_curve <- hexbin_data_object_trimap_s_curve$hexdf_data
-
-TriMAP_data_with_hb_id_s_curve <- TriMAP_s_curve |>
-  dplyr::mutate(hb_id = hexbin_data_object_trimap_s_curve$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_trimap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), TriMAP_data_with_hb_id_s_curve)
-
-## Averaged on high-D
-df_bin_trimap_s_curve <- avg_highD_data(.data = df_all_trimap_s_curve)
-
-## Triangulate bin centroids
-tr1_object_trimap_s_curve <- triangulate_bin_centroids(df_bin_centroids_trimap_s_curve, x, y)
-tr_from_to_df_trimap_s_curve <- generate_edge_info(triangular_object = tr1_object_trimap_s_curve)
-
-# ggplot(df_bin_centroids_trimap_s_curve, aes(x = x, y = y)) +
-#   geom_segment(data = tr_from_to_df_trimap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
-#   geom_point(size = 2, colour = "#33a02c") +
-#   coord_equal()
-
-## Compute 2D distances
-distance_trimap_s_curve <- cal_2d_dist(.data = tr_from_to_df_trimap_s_curve)
-
-## To find the benchmark value
-benchmark_trimap_s_curve <- find_benchmark_value(.data = distance_trimap_s_curve, distance_col = distance)
-
-# colour_long_edges(.data = distance_trimap_s_curve, benchmark_value = benchmark_trimap_s_curve,
-#                   triangular_object = tr1_object_trimap_s_curve, distance_col = distance)
-
-trimesh_removed_trimap_s_curve <- remove_long_edges(.data = distance_trimap_s_curve, benchmark_value = benchmark_trimap_s_curve,
-                                                    triangular_object = tr1_object_trimap_s_curve, distance_col = distance)
-
-trimesh_removed_trimap_s_curve <- trimesh_removed_trimap_s_curve +
-  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  # theme(axis.text = element_text(size = 5),
-  #       axis.title = element_text(size = 7)) +
-  geom_point(colour = "#33a02c", size = 0.25) +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-# tour_trimap_s_curve <- show_langevitour(df_all_trimap_s_curve, df_bin_trimap_s_curve, df_bin_centroids_trimap_s_curve, benchmark_value = benchmark_trimap_s_curve, distance = distance_trimap_s_curve, distance_col = distance)
-
-## PacMAP
-
-num_bins_pacmap_s_curve <- 10
-shape_val_pacmap_s_curve <- calculate_effective_shape_value(.data = PaCMAP_s_curve,
-                                                            x = PaCMAP1, y = PaCMAP2) ## 1.259938
-## To extract bin centroids
-hexbin_data_object_pacmap_s_curve <- extract_hexbin_centroids(nldr_df = PaCMAP_s_curve, num_bins = num_bins_pacmap_s_curve, shape_val = shape_val_pacmap_s_curve, x = PaCMAP1, y = PaCMAP2)
-
-df_bin_centroids_pacmap_s_curve <- hexbin_data_object_pacmap_s_curve$hexdf_data
-
-PaCMAP_data_with_hb_id_s_curve <- PaCMAP_s_curve |>
-  dplyr::mutate(hb_id = hexbin_data_object_pacmap_s_curve$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_pacmap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), PaCMAP_data_with_hb_id_s_curve)
-
-## Averaged on high-D
-df_bin_pacmap_s_curve <- avg_highD_data(.data = df_all_pacmap_s_curve)
-
-## Triangulate bin centroids
-tr1_object_pacmap_s_curve <- triangulate_bin_centroids(df_bin_centroids_pacmap_s_curve, x, y)
-tr_from_to_df_pacmap_s_curve <- generate_edge_info(triangular_object = tr1_object_pacmap_s_curve)
-
-# ggplot(df_bin_centroids_pacmap_s_curve, aes(x = x, y = y)) +
-#   geom_segment(data = tr_from_to_df_pacmap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
-#   geom_point(size = 2, colour = "#33a02c") +
-#   coord_equal()
-
-## Compute 2D distances
-distance_pacmap_s_curve <- cal_2d_dist(.data = tr_from_to_df_pacmap_s_curve)
-
-## To find the benchmark value
-benchmark_pacmap_s_curve <- find_benchmark_value(.data = distance_pacmap_s_curve, distance_col = distance)
-
-# colour_long_edges(.data = distance_pacmap_s_curve, benchmark_value = benchmark_pacmap_s_curve,
-#                   triangular_object = tr1_object_pacmap_s_curve, distance_col = distance)
-
-trimesh_removed_pacmap_s_curve <- remove_long_edges(.data = distance_pacmap_s_curve, benchmark_value = benchmark_pacmap_s_curve,
-                                     triangular_object = tr1_object_pacmap_s_curve, distance_col = distance)
-
-trimesh_removed_pacmap_s_curve <- trimesh_removed_pacmap_s_curve +
-  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  # theme(axis.text = element_text(size = 5),
-  #       axis.title = element_text(size = 7)) +
-  geom_point(colour = "#33a02c", size = 0.5) +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'e', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-# tour_pacmap_s_curve <- show_langevitour(df_all_pacmap_s_curve, df_bin_pacmap_s_curve, df_bin_centroids_pacmap_s_curve, benchmark_value = benchmark_pacmap_s_curve, distance = distance_pacmap_s_curve, distance_col = distance)
-
-## PHATE
-
-num_bins_phate_s_curve <- 8
-shape_val_phate_s_curve <- calculate_effective_shape_value(.data = PHATE_s_curve,
-                                                           x = PHATE1, y = PHATE2) ## 1.259938
-## To extract bin centroids
-hexbin_data_object_phate_s_curve <- extract_hexbin_centroids(nldr_df = PHATE_s_curve, num_bins = num_bins_phate_s_curve, shape_val = shape_val_phate_s_curve, x = PHATE1, y = PHATE2)
-
-df_bin_centroids_phate_s_curve <- hexbin_data_object_phate_s_curve$hexdf_data
-
-## Identify bins with low-density
-identify_rm_bins <- find_low_density_hexagons(df_bin_centroids_phate_s_curve, num_bins_phate_s_curve, benchmark_rm_hex = 0.06)
-
-df_bin_centroids_phate_s_curve <- df_bin_centroids_phate_s_curve |>
-  filter(!(hexID %in% identify_rm_bins))
-
-PHATE_data_with_hb_id_s_curve <- PHATE_s_curve |>
-  dplyr::mutate(hb_id = hexbin_data_object_phate_s_curve$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_phate_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), PHATE_data_with_hb_id_s_curve)
-
-## Averaged on high-D
-df_bin_phate_s_curve <- avg_highD_data(.data = df_all_phate_s_curve)
-
-## Triangulate bin centroids
-tr1_object_phate_s_curve <- triangulate_bin_centroids(df_bin_centroids_phate_s_curve, x, y)
-tr_from_to_df_phate_s_curve <- generate_edge_info(triangular_object = tr1_object_phate_s_curve)
-
-# ggplot(df_bin_centroids_phate_s_curve, aes(x = x, y = y)) +
-#   geom_segment(data = tr_from_to_df_phate_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
-#   geom_point(size = 2, colour = "#33a02c") +
-#   coord_equal()
-
-## Compute 2D distances
-distance_phate_s_curve <- cal_2d_dist(.data = tr_from_to_df_phate_s_curve)
-
-## To find the benchmark value
-benchmark_phate_s_curve <- find_benchmark_value(.data = distance_phate_s_curve, distance_col = distance)
-
-# colour_long_edges(.data = distance_phate_s_curve, benchmark_value = benchmark_phate_s_curve,
-#                   triangular_object = tr1_object_phate_s_curve, distance_col = distance)
-
-trimesh_removed_phate_s_curve <- remove_long_edges(.data = distance_phate_s_curve, benchmark_value = benchmark_phate_s_curve,
-                                                   triangular_object = tr1_object_phate_s_curve, distance_col = distance)
-
-trimesh_removed_phate_s_curve <- trimesh_removed_phate_s_curve +
-  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  # theme(axis.text = element_text(size = 5),
-  #       axis.title = element_text(size = 7)) +
-  geom_point(colour = "#33a02c", size = 0.5) +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-# tour_phate_s_curve <- show_langevitour(df_all_phate_s_curve, df_bin_phate_s_curve, df_bin_centroids_phate_s_curve, benchmark_value = benchmark_phate_s_curve, distance = distance_phate_s_curve, distance_col = distance)
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-#| label: fig-modelScurve
-#| fig-cap: Is there a good model to represent the original data in 2D space or are they all providing equivalent information?, (a) Model in the 2D space with tSNE (<https://youtu.be/uy9SkAo6gAE>), (b) Model in the 2D space with UMAP (<https://youtu.be/0MJDhHrh_Ug>), (c) Model in the 2D space with PHATE (<https://youtu.be/HbVv0uy0QWk>), (d) Model in the 2D space with TriMAP (<https://youtu.be/2OGwipAjzc8>), and (e) Model in the 2D space with PaCMAP (<https://youtu.be/pkokI8d-cBk>).
-#| out-width: 100%
-
-trimesh_removed_tsne_s_curve + trimesh_removed_umap_s_curve + trimesh_removed_phate_s_curve + trimesh_removed_trimap_s_curve + trimesh_removed_pacmap_s_curve +
-  #plot_annotation(tag_levels = 'a') +
-  plot_layout(ncol = 5) &
-  theme(legend.position='none', plot.tag = element_text(size = 8))
-```
--->
-
-<!--
-### Single-Cell Tagged Reverse Transcription sequencing data of mouse
-
-The Zeisel mouse brain dataset, obtained through Spatial Transcriptomics (STRT-Seq).
-Within this dataset, information is collected from a substantial 2,816 individual mouse brain cells.
-Each of these cells acts as a molecular snapshot, capturing the distinctive genetic activity within various cell types.
-This diversity spans neurons, glial cells, and other essential components of the brain, offering a comprehensive view of the cellular tapestry.
-
-What makes this dataset particularly valuable is its ability to shed light on the spatial distribution of cells.
-Researchers can explore how gene expression patterns vary across different regions of the mouse brain, unlocking insights into the functional specialization of these regions and the intricate networks that underpin neural processes.
-
-```{r}
-#| warning: false
-#| echo: false
-
-## Import data
-df_2 <- read_rds("data/zeisel/zeisel.rds")
-training_data_1 <- read_rds("data/zeisel/zeisel_training.rds")
-test_1 <- read_rds("data/zeisel/zeisel_test.rds")
-
-tSNE_zeisel <- read_rds("data/zeisel/zeisel_tsne_30.rds")
-UMAP_zeisel <- read_rds("data/zeisel/zeisel_umap.rds")
-PHATE_zeisel <- read_rds("data/zeisel/zeisel_phate.rds")
-TriMAP_zeisel <- read_rds("data/zeisel/zeisel_trimap.rds")
-PaCMAP_zeisel <- read_rds("data/zeisel/zeisel_pacmap.rds")
-
-## Visualise embeddings
-
-plot_list1_zeisel <- plot_tSNE_2D(tSNE_zeisel) + #ggtitle("(a)") + 
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-plot_list2_zeisel <- plot_UMAP_2D(UMAP_zeisel) + #ggtitle("(b)") + 
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-plot_list3_zeisel <- plot_PHATE_2D(PHATE_zeisel) + #ggtitle("(c)") + 
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-plot_list4_zeisel <- plot_TriMAP_2D(TriMAP_zeisel) + #ggtitle("(d)") + 
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-plot_list5_zeisel <- plot_PaCMAP_2D(PaCMAP_zeisel) + #ggtitle("(e)") + 
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'e', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "2D layouts from different NLDR techniques applied for the training Zeisel mouse brain dataset: (a) tSNE (perplexity = 30), (b) UMAP (n_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2).
-Is there a best representation of the original data or are they all providing  equivalent information?"
-#| label: fig-nldervis5Mouse
-#| out-width: 100%
-
-plot_list1_zeisel + plot_list2_zeisel + plot_list3_zeisel + plot_list4_zeisel + plot_list5_zeisel +
-  plot_layout(ncol=5)
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-MSE_df_zei <- read_rds("data/zeisel/summary_zei.rds")
-MSE_df_zei$method <- factor(MSE_df_zei$method, levels = c("tSNE", "UMAP", "PHATE", "TriMAP", "PaCMAP"))
-
-## To draw with AIC
-aic_plot_zei <- ggplot(MSE_df_zei |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
-                                                                                 y = total_error,
-                                                                                 color = method
-)) +
-  geom_point() +
-  geom_line() +
-  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  theme_light() +
-  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  scale_colour_manual(values = c("#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("AIC") +
-  xlab("Total number of bins")
-## Effective number of bins along x-axis
-
-mse_plot_zei <- ggplot(MSE_df_zei, aes(x = number_of_bins,
-                                       y = total_mse,
-                                       color = method
-)) +
-  geom_point() +
-  geom_line() +
-  theme_light() +
-  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  scale_colour_manual(values = c("#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("MSE") +
-  xlab("Total number of bins")
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: Goodness of fit statistics from different NLDR techniques applied to training Zeisel mouse brain dataset. What is the best NLDR technique to represent the original data in 2D?
-#| label: fig-diagnosticpltZEI
-#| out-width: 100%
-
-aic_plot_zei + mse_plot_zei +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', ncol = 2) &
-  theme(legend.position='bottom', plot.tag = element_text(size = 8))
-```
-
-### Retina dataset
-
-Human retina: https://www.ebi.ac.uk/biostudies/arrayexpress/studies/E-MTAB-7316
-
-https://www.cell.com/cms/10.1016/j.cell.2015.05.002/attachment/fd52950c-fd35-401a-9e60-ca7588ef9001/mmc1.pdf
-
-https://www.cell.com/fulltext/S0092-8674(15)00549-8#app3
-
-32 PCs
-tSNE (perplexity 30)
-
-### Kang dataset
-
-For the Kang lupus dataset (29), we obtained 16 cell types based on clustering of the Festem-
-selected genes (Fig. 4E and Fig. S10, S11), including 2 fine cell types that were often missed by
-other methods. The 2 cell types were IFNhi CD14+ monocytes with high expression of type I
-interferon-related (IFN) genes, and HSP+ CD4 T cells with up-regulated heat-
-shock-protein-related (HSP) genes. The HSP+ CD4 T cells might result from
-cryopreservation and thawing of T cells (37). Other than devianceFS, other available methods
-could not identify these HSP+ CD4 T cells (Fig. S11).
-
-```{r}
-#| warning: false
-#| echo: false
-
-## Import data
-#df_2 <- read_rds("data/pbmc/pbmc.rds")
-# training_data_1 <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_pca_50.rds")
-# training_data_1 <- training_data_1[, 1:25]
-#test_1 <- read_rds("data/pbmc/pbmc_test.rds")
-
-#tSNE_pbmc <- read_rds("data/pbmc/pbmc_tsne_30.rds")
-tSNE_kang <- read_rds("data/kang/kang_tsne.rds") |>
-  mutate(cell_label = if_else(cell_label %in% c("0", "3", "5"), "0", cell_label))
-# PHATE_pbmc <- read_rds("data/pbmc/pbmc_phate.rds")
-# TriMAP_pbmc <- read_rds("data/pbmc/pbmc_trimap.rds")
-# PaCMAP_pbmc <- read_rds("data/pbmc/pbmc_pacmap.rds")
-
-#load("data/pbmc/pbmc_3k_festem/pbmc3k_label.RData")
-
-## Visualise embeddings
-
-# plot_list1_pbmc <- plot_tSNE_2D(tSNE_pbmc) + #ggtitle("(a)") + 
-#   theme_linedraw() +
-#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-#         axis.title.x = element_blank(), axis.title.y = element_blank(),
-#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-#   annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-class_avg <- tSNE_kang |>
-  group_by(cell_label) |>
-  summarise(
-    tSNE1 = median(tSNE1),
-    tSNE2 = median(tSNE2)
-  ) |>
-  mutate(cell_label = as.numeric(cell_label)) |>
-  arrange(cell_label)
-
-class_avg$cell_label <- as.factor(class_avg$cell_label)
-
-levels(class_avg$cell_label) <- c("CD14 Monocyte", "CD4 \n Memory T", "CD4 Naive T", "CD16 Monocyte", "B", "T Activated", "CD8 T", "NK", "DC", "B \nActivated", "T cell:\nmonocyte \nComplex", "HSP+ CD4 T", "IFN^hi CD14 \n Monocyte", "CD34 \n Progenitors", "pDC")
-
-plot_list2_kang <- tSNE_kang |>
-    ggplot(aes(x = tSNE1,
-               y = tSNE2, color = cell_label))+
-    geom_point(alpha=0.5) +
-    coord_equal() +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#b15928", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#cab2d6", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#bc80bd", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
-  geom_text(aes(x=tSNE1,y = tSNE2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) #+
-  # guides(color = guide_legend(title = "Cell type"),
-  #        color = guide_legend(nrow = 3)) #+
-  # annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-# plot_list3_pbmc <- plot_PHATE_2D(PHATE_pbmc) + #ggtitle("(c)") + 
-#   theme_linedraw() +
-#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-#         axis.title.x = element_blank(), axis.title.y = element_blank(),
-#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-#   annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-# plot_list4_pbmc <- plot_TriMAP_2D(TriMAP_pbmc) + #ggtitle("(d)") + 
-#   theme_linedraw() +
-#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-#         axis.title.x = element_blank(), axis.title.y = element_blank(),
-#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-#   annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-# 
-# plot_list5_pbmc <- plot_PaCMAP_2D(PaCMAP_pbmc) + #ggtitle("(e)") + 
-#   theme_linedraw() +
-#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-#         axis.title.x = element_blank(), axis.title.y = element_blank(),
-#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-#   annotate(geom = 'text', label = 'e', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "2D layout from tSNE (perplexity = 30) applied for the Kang dataset. Is this a best representation of the original data?"
-#| label: fig-nlderviskangtSNE
-
-plot_list2_kang
-
-#  plot_list1_pbmc + plot_list3_pbmc + plot_list4_pbmc + plot_list5_pbmc +
-#   plot_layout(ncol=5) #| out-width: 100%
-
-# 2D layouts from different NLDR techniques applied for the training PBMC dataset: (a) tSNE (perplexity = 30), (b) UMAP (n_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2).
-Is there a best representation of the original data or are they all providing  equivalent information?
-```
--->
-
-<!--
-### Single-cell RNA-seq data of human
-
-https://github.com/XiDsLab/Festem_paper/tree/main
-
-@Chen2023
-
-https://satijalab.org/seurat/articles/pbmc3k_tutorial.html
-https://www.cell.com/fulltext/S0092-8674(15)00549-8
-
-In the field of single-cell studies, a common analysis task involves clustering to identify groups of cells with similar expression profiles.
-Analysts often turn to NLDR techniques to verify and identify these clusters and explore developmental trajectories (e.g., example 1).
-In clustering workflows, the main objective is to verify the existence of clusters and subsequently identify them as specific cell types by examining the expression of "known" marker genes.
-In this context, a "faithful" embedding should ideally preserve the topology of the data, ensuring that cells corresponding to the same cell type are situated close to the high-dimensional space.
-
-To begin our analysis, we installed the Peripheral Blood Mononuclear Cells (pbmc) data set obtained from 10x Genomics using the `SeuratData` R package [@Rahul2019], which facilitates the distribution of data sets in the form of Seurat objects [@Yuhan2021].
-This data set contains 13,714 features across 2,700 samples within a single assay.
-The active assay is RNA, with 13,714 features representing different gene expressions.
-After loading the data set, we obtained the principal components (PCs) and assessed the variance explained by each PC.
-Based on this evaluation, we selected seven PCs, representing approximately 50% of the variance in gene expression, for further analysis.
-
-Next, we employed the UMAP technique with default parameter settings.
-As illustrated in @fig-pbmc, the cell types B and Platelet are well-separated in the UMAP layout.
-Moreover, CD14+ Mono, FCGR3A+ Mono, and DC form a distinct cluster, while Naive CD4 T, NK, Memory CD4 T, and CD8 T are grouped together in another cluster.
-The values utilized to construct the smooth low-dimensional manifold are presented in @tbl-table02.
-The linked video, demonstrating the tour with the model, showcases the generation of a smooth surface for this application, enabling a comprehensive exploration of the data's structure and relationships (see @fig-pbmc_sc).
-
-```{r}
-#| warning: false
-#| echo: false
-
-MSE_df_pbmc <- read_rds("data/pbmc/summary_pbmc.rds")
-MSE_df_pbmc$method <- factor(MSE_df_pbmc$method, levels = c("tSNE", "UMAP", "PHATE", "TriMAP", "PaCMAP"))
-
-## To draw with AIC
-aic_plot_pbmc <- ggplot(MSE_df_pbmc |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
-                                                                                 y = total_error,
-                                                                                 color = method
-)) +
-  geom_point() +
-  geom_line() +
-  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  theme_light() +
-  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  scale_colour_manual(values = c("#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("AIC") +
-  xlab("Total number of bins")
-## Effective number of bins along x-axis
-
-mse_plot_pbmc <- ggplot(MSE_df_pbmc, aes(x = number_of_bins,
-                                       y = total_mse,
-                                       color = method
-)) +
-  geom_point() +
-  geom_line() +
-  theme_light() +
-  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  scale_colour_manual(values = c("#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("MSE") +
-  xlab("Total number of bins")
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: Goodness of fit statistics from different NLDR techniques applied to training PBMC dataset. What is the best NLDR technique to represent the original data in 2D?
-#| label: fig-diagnosticpltPBMC
-#| out-width: 100%
-
-aic_plot_pbmc + mse_plot_pbmc +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', ncol = 2) &
-  theme(legend.position='bottom', plot.tag = element_text(size = 8))
-```
-
-### HARPS GTO sample dataset
-
-@Anders2018
-
-- In 2D, there are two well-separated clusters, but the clusters are really close
-
-- One cluster have a sparse data space within the cluster
-
-- There is a line in right corner of the structure that is not connected any other trimesh points. This line create a long edge in high-D space.
-
-- In high-D view, the clusters having some non-linearity patter, which is also not captured in tSNE.
-Therefore, this is not a good choice.
-Also, the two clusters are really near to each other and can't see any separation.
-Also continuity of the structure.
-Long edges are presence also.
-
-- The parameter is, in a sense, a guess about the number of close neighbors each point has.
-
-- Num bins along the x-axis = 43, shape parameter = 0.4466309
-
-```{r}
-#| warning: false
-#| echo: false
-
-error_df <- read_csv("data/harps/error_harps_tsne.csv", col_names = FALSE)
-
-error_df <- error_df |>
-  mutate(X2 = round(X2/100, 2),
-         X1 = as.integer(X1)) |>
-  arrange(X2)
-
-colnames(error_df) <- c("perplexity", "error (x 100)")
-
-kable(error_df, caption = "errors for different parameter combinations", label = "tbl-errordiffcombharps") |>
-  #xtable2kable() |>
-  kable_styling(latex_options = c("striped", "HOLD_position"), position = "center", full_width = FALSE) |>
-  row_spec(0, bold = TRUE) |>
-  row_spec(3, bold=T, color = "red")
-```
-
-<!-- Author suggestion-->
-```{r}
-#| warning: false
-#| echo: false
-
-tSNE_data_harps <- read_rds("data/harps/harps_tsne_40.rds")
-
-plot_list2_harps <- tSNE_data_harps |>
-    ggplot(aes(x = tSNE1,
-               y = tSNE2, color = cluster))+
-    geom_point(alpha=0.5) +
-    #coord_cartesian(xlim = c(-25, 25), ylim = c(-12, 12)) +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#b15928", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#cab2d6", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#bc80bd", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
-  guides(color = guide_legend(title = "cluster"),
-         color = guide_legend(nrow = 3)) +
- annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-## Import data
-training_data_1 <- read_rds("data/harps/harps_data.rds")
-training_data_1 <- training_data_1 |>
-  mutate(ID = 1:NROW(training_data_1))
-
-clusters <- training_data_1$cluster
-select_names <-  names(training_data_1)[-c(18, length(training_data_1))]
-
-training_data_1 <- training_data_1 |> select(-cluster)
-names(training_data_1) <- append(paste0("x", 1:17), "ID")
-
-tSNE_data_harps <- tSNE_data_harps |>
-  select(tSNE1, tSNE2) |>
-  mutate(ID = 1:NROW(tSNE_data_harps))
-
-## tSNE
-## Prediction
-
-shape_val_tsne_harps <- calculate_effective_shape_value(.data = tSNE_data_harps,
-                                                      x = tSNE1, y = tSNE2)
-
-num_bins_vec <- 1:21 ## Number of bins along the x-axis
-
-vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names
-
-eval_data_training <- dplyr::bind_rows(vec)[0, ]
-eval_data_training <- eval_data_training |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(num_bins_vec)) {
-
-  model_object <- fit_high_d_model(training_data = training_data_1, nldr_df_with_id = tSNE_data_harps, x = "tSNE1", y = "tSNE2", num_bins_x = num_bins_vec[i], shape_val = shape_val_tsne_harps,
-                             is_bin_centroid = TRUE,
-                             is_rm_lwd_hex = FALSE,
-                             benchmark_to_rm_lwd_hex = NA,
-                             is_avg_high_d = TRUE, column_start_text = "x")
-
-  centroid_df_training <- model_object$df_bin_centroids
-  avg_df_training <- model_object$df_bin
-
-  pred_df_training <- predict_2d_embeddings(test_data = training_data_1, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "tSNE")
-
-  eval_df_training <- generate_eval_df(data = training_data_1, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, col_start = "x")
-
-  eval_df_training <- eval_df_training |>
-    mutate(num_bins_x = num_bins_vec[i])
-
-  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
-
-}
-
-## Add new column with data types
-
-MSE_df_harps <- eval_data_training |>
-  dplyr::mutate(data_type = "training",
-                method = "tSNE")
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-## To draw with AIC
-aic_plot_harps <- ggplot(MSE_df_harps |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "tSNE"), aes(x = number_of_bins,
-                                                                                 y = total_error,
-                                                                                 color = method
-)) +
-  geom_point() +
-  geom_line() +
-  geom_vline(xintercept = 120, linetype="dashed", 
-                color = "black", size=0.5) +
-  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  scale_colour_manual(values = c("#e41a1c","#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("AIC") +
-  xlab("total number of bins")
-## Effective number of bins along x-axis
-
-mse_plot_harps <- ggplot(MSE_df_harps |> dplyr::filter(method == "tSNE"), aes(x = number_of_bins,
-                                       y = total_mse,
-                                       color = method
-)) +
-  geom_point() +
-  geom_line() +
-   geom_vline(xintercept = 120, linetype="dashed", 
-                color = "black", size=0.5) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  scale_colour_manual(values = c("#e41a1c","#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
+  #scale_colour_manual(values = c("#377eb8")) +
   ylab("MSE") +
@@
 #| echo: false
-#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset.
-What is the effective number of bins in each NLDR technique to create a 2D model?
+#| fig-cap: "Goodness of fit statistics from PaCMAP applied to training MNIST dataset.
+What is the effective number of bins in PaCMAP to create a 2D model?
 The MSE plot have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
 Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error.
-Using the elbow method, when the total number of bins is set to 144, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
+Using the elbow method, when the total number of bins is set to 180, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
 This point indicates that adding less bins does not enough to capture the data structure."
-#| label: fig-diagnosticpltHARPS
-##| out-width: 100%
+#| label: fig-diagnosticpltMNIST
 #| fig-pos: H
 
-aic_plot_harps + mse_plot_harps +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', ncol = 2) &
-  theme(legend.position='none', plot.tag = element_text(size = 8))
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-### tSNE
-
-num_bins_tsne_harps <- calculate_effective_x_bins(.data = tSNE_data_harps, x = "tSNE1",
-                                                cell_area = 1) #43
-
-num_bins_tsne_harps <- 14
-
-shape_val_tsne_harps <- calculate_effective_shape_value(.data = tSNE_data_harps,
-                                                      x = "tSNE1", y = "tSNE2")
-## To extract bin centroids
-hexbin_data_object_tsne_harps <- extract_hexbin_centroids(nldr_df = tSNE_data_harps, num_bins = num_bins_tsne_harps, shape_val = shape_val_tsne_harps, x = "tSNE1", y = "tSNE2")
-
-df_bin_centroids_tsne_harps <- hexbin_data_object_tsne_harps$hexdf_data
-
-#num_non_empty_bins_harps <- df_bin_centroids_tsne_harps$hexID |> length()
-
-## Identify bins with low-density
-# identify_rm_bins <- find_low_density_hexagons(df_bin_centroids_tsne_harps, num_bins_tsne_harps, benchmark_rm_hex = 0.01)
-# 
-# df_bin_centroids_tsne_harps <- df_bin_centroids_tsne_harps |>
-#   filter(!(hexID %in% identify_rm_bins))
-
-tSNE_data_with_hb_id_harps <- tSNE_data_harps |>
-  dplyr::mutate(hb_id = hexbin_data_object_tsne_harps$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_tsne_harps <- dplyr::bind_cols(training_data_1 |> dplyr::select(-ID), tSNE_data_with_hb_id_harps)
-
-## Averaged on high-D
-df_bin_tsne_harps <- avg_highD_data(.data = df_all_tsne_harps)
-
-## Triangulate bin centroids
-tr1_object_tsne_harps <- triangulate_bin_centroids(df_bin_centroids_tsne_harps, x, y)
-tr_from_to_df_tsne_harps <- generate_edge_info(triangular_object = tr1_object_tsne_harps)
-
-## Compute 2D distances
-distance_tsne_harps <- cal_2d_dist(.data = tr_from_to_df_tsne_harps)
-
-## To find the benchmark value
-benchmark_tsne_harps <- find_benchmark_value(.data = distance_tsne_harps, distance_col = "distance")
-benchmark_tsne_harps <- 5
-
-# distance_plot <- plot_dist(distance_tsne_harps) +
-#   #ggtitle("(b)" ) +
-#   ylab(expression(d^{(2)})) +
-#   theme(axis.text = element_text(size = 5),
-#         axis.title = element_text(size = 12))
-# 
-# distance_plot
-
-trimesh_removed_tsne_harps <- remove_long_edges(.data = distance_tsne_harps, benchmark_value = benchmark_tsne_harps,
-                                              triangular_object = tr1_object_tsne_harps, distance_col = distance)
-
-trimesh_removed_tsne_harps <- trimesh_removed_tsne_harps +
-  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  # theme(axis.text = element_text(size = 5),
-  #       axis.title = element_text(size = 7)) +
-  geom_point(colour = "#33a02c", size = 0.5) +
-  #coord_cartesian(xlim = c(-25, 25), ylim = c(-12, 12)) +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-df_bin_train <- df_bin_tsne_harps
-names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])
-
-error_df <- df_all_tsne_harps |>
-  dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates
-
-# prediction_df_join <- prediction_df_join |>
-#   dplyr::left_join(data, by = c("ID" = "ID")) ## Map high-D data
-
-for (i in 1:(NCOL(df_bin_train) - 1)) {
-
-  error_df[ , paste0("abs_residual_", "x", i)] <- abs(error_df[ , paste0("x", i)] - error_df[ , paste0("avg_", "x", i)])
-
-}
-
-error_df <- error_df |>
-  dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "x")))))
-
-error_df <- error_df |>
-  mutate(type = if_else(total <= 0, "no error",
-                        if_else(total <= 0.2, "error 0-0.2",
-                                if_else(total <= 0.3, "error 0.2-0.3",
-                                        if_else(total <= 0.5, "error 0.3-0.5",
-                                                if_else(total <= 1, "error 0.5-1",
-                                                        "error greter than 1"))))))
-
-plot_list2_harps_error <- error_df |>
-    mutate(type = factor(type , levels = c("no error", "error 0-0.2", "error 0.2-0.3", "error 0.3-0.5", "error 0.5-1", "error greter than 1"))) |>
-    ggplot(aes(x = tSNE1,
-               y = tSNE2, color = type))+
-    geom_point(alpha=0.5, size = 0.1) +
-    coord_equal() +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#fee0d2", "#fcbba1",
-                                         "#fc9272", "#fb6a4a", "#ef3b2c",
-                                         "#cb181d", "#a50f15", "#99000d"))
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "(a) 2D layout from tSNE (perplexity = 40) applied for the HARPS GTO dataset.
-Is this a best representation of the original data?, (b) Model in the 2D space with tSNE (<https://youtu.be/jxo_rPyGv9k>)"
-#| label: fig-nldervisharpstSNE
-#| fig-pos: h
-
-plot_list2_harps + trimesh_removed_tsne_harps + plot_list2_harps_error +
-  plot_layout(guides='collect', ncol = 3) &
-  theme(legend.position='bottom', plot.tag = element_text(size = 8))
-
-```
-
-<!-- param set 1-->
-
-- 5 clusters shows in 2D
-
-- 3 clusters which are really close shows in high_d data with the model
-
-```{r}
-#| warning: false
-#| echo: false
-
-tSNE_data_harps <- read_rds("data/harps/harps_tsne_23.rds")
-
-plot_list2_harps <- tSNE_data_harps |>
-    ggplot(aes(x = tSNE1,
-               y = tSNE2, color = clusters))+
-    geom_point(alpha=0.5) +
-    coord_cartesian(xlim = c(-25, 20), ylim = c(-35, 25)) +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#b15928", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#cab2d6", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#bc80bd", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
-  guides(color = guide_legend(title = "cluster"),
-         color = guide_legend(nrow = 3)) +
- annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-## tSNE
-## Prediction
-
-shape_val_tsne_harps <- calculate_effective_shape_value(.data = tSNE_data_harps,
-                                                      x = tSNE1, y = tSNE2)
-
-num_bins_vec <- 1:21 ## Number of bins along the x-axis
-
-vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names
-
-eval_data_training <- dplyr::bind_rows(vec)[0, ]
-eval_data_training <- eval_data_training |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(num_bins_vec)) {
-
-  model_object <- fit_high_d_model(training_data = training_data_1, nldr_df_with_id = tSNE_data_harps, x = "tSNE1", y = "tSNE2", num_bins_x = num_bins_vec[i], shape_val = shape_val_tsne_harps,
-                             is_bin_centroid = TRUE,
-                             is_rm_lwd_hex = FALSE,
-                             benchmark_to_rm_lwd_hex = NA,
-                             is_avg_high_d = TRUE, column_start_text = "x")
-
-  centroid_df_training <- model_object$df_bin_centroids
-  avg_df_training <- model_object$df_bin
-
-  pred_df_training <- predict_2d_embeddings(test_data = training_data_1, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "tSNE")
-
-  eval_df_training <- generate_eval_df(data = training_data_1, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, col_start = "x")
-
-  eval_df_training <- eval_df_training |>
-    mutate(num_bins_x = num_bins_vec[i])
-
-  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
-
-}
-
-## Add new column with data types
-
-MSE_df_harps <- eval_data_training |>
-  dplyr::mutate(data_type = "training",
-                method = "tSNE")
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-## To draw with AIC
-aic_plot_harps <- ggplot(MSE_df_harps |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "tSNE"), aes(x = number_of_bins,
-                                                                                 y = total_error,
-                                                                                 color = method
-)) +
-  geom_point() +
-  geom_line() +
-  geom_vline(xintercept = 160, linetype="dashed", 
-                color = "black", size=0.5) +
-  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  scale_colour_manual(values = c("#e41a1c","#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("AIC") +
-  xlab("total number of bins")
-## Effective number of bins along x-axis
-
-mse_plot_harps <- ggplot(MSE_df_harps |> dplyr::filter(method == "tSNE"), aes(x = number_of_bins,
-                                       y = total_mse,
-                                       color = method
-)) +
-  geom_point() +
-  geom_line() +
-   geom_vline(xintercept = 160, linetype="dashed", 
-                color = "black", size=0.5) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  scale_colour_manual(values = c("#e41a1c","#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("MSE") +
-  xlab("total number of bins")
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset.
-What is the effective number of bins in each NLDR technique to create a 2D model?
-The MSE plot have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
-Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error.
-Using the elbow method, when the total number of bins is set to 144, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
-This point indicates that adding less bins does not enough to capture the data structure."
-#| label: fig-diagnosticpltHARPS2
-##| out-width: 100%
-#| fig-pos: H
-
-aic_plot_harps + mse_plot_harps +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', ncol = 2) &
-  theme(legend.position='none', plot.tag = element_text(size = 8))
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-## Import data
-# training_data_1 <- read_rds("data/harps/harps_data.rds")
-# training_data_1 <- training_data_1 |>
-#   mutate(ID = 1:NROW(training_data_1))
-# 
-# clusters <- training_data_1$cluster
-# select_names <-  names(training_data_1)[-c(18, length(training_data_1))]
-# 
-# training_data_1 <- training_data_1 |> select(-cluster)
-# names(training_data_1) <- append(paste0("x", 1:17), "ID")
-# 
-# tSNE_data_harps <- tSNE_data_harps |>
-#   select(tSNE1, tSNE2) |>
-#   mutate(ID = 1:NROW(tSNE_data_harps))
-
-### tSNE
-
-shape_val_tsne_harps <- calculate_effective_shape_value(.data = tSNE_data_harps,
-                                                      x = tSNE1, y = tSNE2)
-
-# num_bins_tsne_harps <- find_non_empty_bins(nldr_df = tSNE_data_harps, x = "tSNE1", y = "tSNE2", shape_val = shape_val_tsne_harps, non_empty_bins = num_non_empty_bins_harps)
-
-num_bins_tsne_harps <- 9
-
-## To extract bin centroids
-hexbin_data_object_tsne_harps <- extract_hexbin_centroids(nldr_df = tSNE_data_harps, num_bins = num_bins_tsne_harps, shape_val = shape_val_tsne_harps, x = tSNE1, y = tSNE2)
-
-df_bin_centroids_tsne_harps <- hexbin_data_object_tsne_harps$hexdf_data
-
-## Identify bins with low-density
-# identify_rm_bins <- find_low_density_hexagons(df_bin_centroids_tsne_harps, num_bins_tsne_harps, benchmark_rm_hex = 0.01)
-# 
-# df_bin_centroids_tsne_harps <- df_bin_centroids_tsne_harps |>
-#   filter(!(hexID %in% identify_rm_bins))
-
-tSNE_data_with_hb_id_harps <- tSNE_data_harps |>
-  dplyr::mutate(hb_id = hexbin_data_object_tsne_harps$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_tsne_harps <- dplyr::bind_cols(training_data_1 |> dplyr::select(-ID), tSNE_data_with_hb_id_harps)
-
-## Averaged on high-D
-df_bin_tsne_harps <- avg_highD_data(.data = df_all_tsne_harps)
-
-## Triangulate bin centroids
-tr1_object_tsne_harps <- triangulate_bin_centroids(df_bin_centroids_tsne_harps, x, y)
-tr_from_to_df_tsne_harps <- generate_edge_info(triangular_object = tr1_object_tsne_harps)
-
-## Compute 2D distances
-distance_tsne_harps <- cal_2d_dist(.data = tr_from_to_df_tsne_harps)
-
-# distance_plot <- plot_dist(distance_tsne_harps) +
-#   #ggtitle("(b)" ) +
-#   ylab(expression(d^{(2)})) +
-#   theme(axis.text = element_text(size = 5),
-#         axis.title = element_text(size = 12))
-# 
-# distance_plot
-
-## To find the benchmark value
-benchmark_tsne_harps <- find_benchmark_value(.data = distance_tsne_harps, distance_col = "distance")
-benchmark_tsne_harps <- 6
-
-trimesh_removed_tsne_harps <- remove_long_edges(.data = distance_tsne_harps, benchmark_value = benchmark_tsne_harps,
-                                              triangular_object = tr1_object_tsne_harps, distance_col = distance)
-
-trimesh_removed_tsne_harps <- trimesh_removed_tsne_harps +
-  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  # theme(axis.text = element_text(size = 5),
-  #       axis.title = element_text(size = 7)) +
-  geom_point(colour = "#33a02c", size = 0.5) +
-  coord_cartesian(xlim = c(-25, 20), ylim = c(-35, 25)) +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-df_bin_train <- df_bin_tsne_harps
-names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])
-
-error_df <- df_all_tsne_harps |>
-  dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates
-
-# prediction_df_join <- prediction_df_join |>
-#   dplyr::left_join(data, by = c("ID" = "ID")) ## Map high-D data
-
-for (i in 1:(NCOL(df_bin_train) - 1)) {
-
-  error_df[ , paste0("abs_residual_", "x", i)] <- abs(error_df[ , paste0("x", i)] - error_df[ , paste0("avg_", "x", i)])
-
-}
-
-error_df <- error_df |>
-  dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "x")))))
-
-error_df <- error_df |>
-  mutate(type = if_else(total <= 0, "no error",
-                        if_else(total <= 0.2, "error 0-0.2",
-                                if_else(total <= 0.3, "error 0.2-0.3",
-                                        if_else(total <= 0.5, "error 0.3-0.5",
-                                                if_else(total <= 1, "error 0.5-1",
-                                                        "error greter than 1"))))))
-
-plot_list2_harps_error <- error_df |>
-    mutate(type = factor(type , levels = c("no error", "error 0-0.2", "error 0.2-0.3", "error 0.3-0.5", "error 0.5-1", "error greter than 1"))) |>
-    ggplot(aes(x = tSNE1,
-               y = tSNE2, color = type))+
-    geom_point(alpha=0.5, size = 0.1) +
-    coord_equal() +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#fee0d2", "#fcbba1",
-                                         "#fc9272", "#fb6a4a", "#ef3b2c",
-                                         "#cb181d", "#a50f15", "#99000d"))
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "(a) 2D layout from tSNE (perplexity = 23) applied for the HARPS GTO dataset.
-Is this a best representation of the original data?, (b) Model in the 2D space with tSNE (<https://youtu.be/VO_4SJEoAUA>)"
-#| label: fig-nldervisharpstSNE2
-#| fig-pos: h
-
-plot_list2_harps + trimesh_removed_tsne_harps + plot_list2_harps_error +
-  plot_layout(guides='collect', ncol = 3) &
-  theme(legend.position='bottom', plot.tag = element_text(size = 8))
-
-```
-
-<!-- param set 2-->
-
-```{r}
-#| warning: false
-#| echo: false
-
-tSNE_data_harps <- read_rds("data/harps/harps_tsne_91.rds")
-
-plot_list2_harps <- tSNE_data_harps |>
-    ggplot(aes(x = tSNE1,
-               y = tSNE2, color = clusters))+
-    geom_point(alpha=0.5) +
-    coord_cartesian(xlim = c(-8, 7), ylim = c(-10, 16)) +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#b15928", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#cab2d6", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#bc80bd", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
-  guides(color = guide_legend(title = "cluster"),
-         color = guide_legend(nrow = 3)) +
- annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-## tSNE
-## Prediction
-
-shape_val_tsne_harps <- calculate_effective_shape_value(.data = tSNE_data_harps,
-                                                      x = tSNE1, y = tSNE2)
-
-num_bins_vec <- 1:21 ## Number of bins along the x-axis
-
-vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names
-
-eval_data_training <- dplyr::bind_rows(vec)[0, ]
-eval_data_training <- eval_data_training |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(num_bins_vec)) {
-
-  model_object <- fit_high_d_model(training_data = training_data_1, nldr_df_with_id = tSNE_data_harps, x = "tSNE1", y = "tSNE2", num_bins_x = num_bins_vec[i], shape_val = shape_val_tsne_harps,
-                             is_bin_centroid = TRUE,
-                             is_rm_lwd_hex = FALSE,
-                             benchmark_to_rm_lwd_hex = NA,
-                             is_avg_high_d = TRUE, column_start_text = "x")
-
-  centroid_df_training <- model_object$df_bin_centroids
-  avg_df_training <- model_object$df_bin
-
-  pred_df_training <- predict_2d_embeddings(test_data = training_data_1, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "tSNE")
-
-  eval_df_training <- generate_eval_df(data = training_data_1, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, col_start = "x")
-
-  eval_df_training <- eval_df_training |>
-    mutate(num_bins_x = num_bins_vec[i])
-
-  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
-
-}
-
-## Add new column with data types
-
-MSE_df_harps <- eval_data_training |>
-  dplyr::mutate(data_type = "training",
-                method = "tSNE")
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-## To draw with AIC
-aic_plot_harps <- ggplot(MSE_df_harps |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "tSNE"), aes(x = number_of_bins,
-                                                                                 y = total_error,
-                                                                                 color = method
-)) +
-  geom_point() +
-  geom_line() +
-  geom_vline(xintercept = 180, linetype="dashed", 
-                color = "black", size=0.5) +
-  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  scale_colour_manual(values = c("#e41a1c","#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("AIC") +
-  xlab("total number of bins")
-## Effective number of bins along x-axis
-
-mse_plot_harps <- ggplot(MSE_df_harps |> dplyr::filter(method == "tSNE"), aes(x = number_of_bins,
-                                       y = total_mse,
-                                       color = method
-)) +
-  geom_point() +
-  geom_line() +
-   geom_vline(xintercept = 180, linetype="dashed", 
-                color = "black", size=0.5) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  scale_colour_manual(values = c("#e41a1c","#377eb8", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("MSE") +
-  xlab("total number of bins")
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset.
-What is the effective number of bins in each NLDR technique to create a 2D model?
-The MSE plot have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
-Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error.
-Using the elbow method, when the total number of bins is set to 144, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
-This point indicates that adding less bins does not enough to capture the data structure."
-#| label: fig-diagnosticpltHARPS3
-##| out-width: 100%
-#| fig-pos: H
-
-aic_plot_harps + mse_plot_harps +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', ncol = 2) &
-  theme(legend.position='none', plot.tag = element_text(size = 8))
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-## Import data
-# training_data_1 <- read_rds("data/harps/harps_data.rds")
-# training_data_1 <- training_data_1 |>
-#   mutate(ID = 1:NROW(training_data_1))
-# 
-# clusters <- training_data_1$cluster
-# select_names <-  names(training_data_1)[-c(18, length(training_data_1))]
-# 
-# training_data_1 <- training_data_1 |> select(-cluster)
-# names(training_data_1) <- append(paste0("x", 1:17), "ID")
-# 
-# tSNE_data_harps <- tSNE_data_harps |>
-#   select(tSNE1, tSNE2) |>
-#   mutate(ID = 1:NROW(tSNE_data_harps))
-
-### tSNE
-
-shape_val_tsne_harps <- calculate_effective_shape_value(.data = tSNE_data_harps,
-                                                      x = tSNE1, y = tSNE2)
-
-# num_bins_tsne_harps <- find_non_empty_bins(nldr_df = tSNE_data_harps, x = "tSNE1", y = "tSNE2", shape_val = shape_val_tsne_harps, non_empty_bins = num_non_empty_bins_harps)
-num_bins_tsne_harps <- 9
-
-## To extract bin centroids
-hexbin_data_object_tsne_harps <- extract_hexbin_centroids(nldr_df = tSNE_data_harps, num_bins = num_bins_tsne_harps, shape_val = shape_val_tsne_harps, x = tSNE1, y = tSNE2)
-
-df_bin_centroids_tsne_harps <- hexbin_data_object_tsne_harps$hexdf_data
-
-## Identify bins with low-density
-# identify_rm_bins <- find_low_density_hexagons(df_bin_centroids_tsne_harps, num_bins_tsne_harps, benchmark_rm_hex = 0.01)
-# 
-# df_bin_centroids_tsne_harps <- df_bin_centroids_tsne_harps |>
-#   filter(!(hexID %in% identify_rm_bins))
-
-tSNE_data_with_hb_id_harps <- tSNE_data_harps |>
-  dplyr::mutate(hb_id = hexbin_data_object_tsne_harps$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_tsne_harps <- dplyr::bind_cols(training_data_1 |> dplyr::select(-ID), tSNE_data_with_hb_id_harps)
-
-## Averaged on high-D
-df_bin_tsne_harps <- avg_highD_data(.data = df_all_tsne_harps)
-
-## Triangulate bin centroids
-tr1_object_tsne_harps <- triangulate_bin_centroids(df_bin_centroids_tsne_harps, x, y)
-tr_from_to_df_tsne_harps <- generate_edge_info(triangular_object = tr1_object_tsne_harps)
-
-## Compute 2D distances
-distance_tsne_harps <- cal_2d_dist(.data = tr_from_to_df_tsne_harps)
-
-# distance_plot <- plot_dist(distance_tsne_harps) +
-#   #ggtitle("(b)" ) +
-#   ylab(expression(d^{(2)})) +
-#   theme(axis.text = element_text(size = 5),
-#         axis.title = element_text(size = 12))
-# 
-# distance_plot
-
-## To find the benchmark value
-benchmark_tsne_harps <- find_benchmark_value(.data = distance_tsne_harps, distance_col = "distance")
-benchmark_tsne_harps <- 2.5
-
-trimesh_removed_tsne_harps <- remove_long_edges(.data = distance_tsne_harps, benchmark_value = benchmark_tsne_harps,
-                                              triangular_object = tr1_object_tsne_harps, distance_col = distance)
-
-trimesh_removed_tsne_harps <- trimesh_removed_tsne_harps +
-  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  # theme(axis.text = element_text(size = 5),
-  #       axis.title = element_text(size = 7)) +
-  geom_point(colour = "#33a02c", size = 0.5) +
-  coord_cartesian(xlim = c(-8, 7), ylim = c(-10, 16)) +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-df_bin_train <- df_bin_tsne_harps
-names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])
-
-error_df <- df_all_tsne_harps |>
-  dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates
-
-# prediction_df_join <- prediction_df_join |>
-#   dplyr::left_join(data, by = c("ID" = "ID")) ## Map high-D data
-
-for (i in 1:(NCOL(df_bin_train) - 1)) {
-
-  error_df[ , paste0("abs_residual_", "x", i)] <- abs(error_df[ , paste0("x", i)] - error_df[ , paste0("avg_", "x", i)])
-
-}
-
-error_df <- error_df |>
-  dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "x")))))
-
-error_df <- error_df |>
-  mutate(type = if_else(total <= 0, "no error",
-                        if_else(total <= 0.2, "error 0-0.2",
-                                if_else(total <= 0.3, "error 0.2-0.3",
-                                        if_else(total <= 0.5, "error 0.3-0.5",
-                                                if_else(total <= 1, "error 0.5-1",
-                                                        "error greter than 1"))))))
-
-plot_list2_harps_error <- error_df |>
-    mutate(type = factor(type , levels = c("no error", "error 0-0.2", "error 0.2-0.3", "error 0.3-0.5", "error 0.5-1", "error greter than 1"))) |>
-    ggplot(aes(x = tSNE1,
-               y = tSNE2, color = type))+
-    geom_point(alpha=0.5, size = 0.1) +
-    coord_equal() +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=c("#fee0d2", "#fcbba1",
-                                         "#fc9272", "#fb6a4a", "#ef3b2c",
-                                         "#cb181d", "#a50f15", "#99000d"))
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "(a) 2D layout from tSNE (perplexity = 91) applied for the HARPS GTO dataset.
-Is this a best representation of the original data?, (b) Model in the 2D space with tSNE (<https://youtu.be/vLvRfcogft8>)"
-#| label: fig-nldervisharpstSNE3
-#| fig-pos: h
-
-plot_list2_harps + trimesh_removed_tsne_harps + plot_list2_harps_error +
-  plot_layout(guides='collect', ncol = 3) &
-  theme(legend.position='bottom', plot.tag = element_text(size = 8))
-
-```
-
--->
-
-<!--
-**(b) Determine the shape parameter** ($s$)
-
-In this step, we determine the shape parameter ($s$) for the hexagonal bins, which significantly influences their shape and arrangement within the grid.
-The $s$ in the hexagonal binning algorithm is defined as the ratio of the height ($y$) to the width ($x$) of the plotting region as defined in @eq-equation1 [@Carr2013].
-
-$$
-s = \frac{y}{x}
-$$ {#eq-equation1}
-
-The shape parameter ($s$) of our algorithm is calculated as the ratio of the ranges of 2D embedding components, where $r_1$ and $r_2$ represent the ranges of 2D embedding components 1 and component 2, respectively (see @eq-equation2).
-
-$$
-s = \frac{r_2}{r_1}
-$$ {#eq-equation2}
--->
+mse_plot_mnist
+```

### 2024-03-26 paper.qmd
Clear the repo

@@
 #| message: false
-#| eval: false
 
@@
 
-for (i in 1:NROW(diff_start_df)) {
+for (i in 1:23) { #NROW(diff_start_df)
 
@@
   mse_df <- mse_df |>
-    tibble::add_row(x_start = diff_start_df[i, 1], y_start = diff_start_df[i, 2],
+    tibble::add_row(x_start = diff_start_df[i, 1], 
+                    y_start = diff_start_df[i, 2],
                     num_non_empty_bins = NROW(centroid_df_training), 
@@
 
-<!--
-
-```{r}
-#| eval: false
-#| echo: false
-#| warning: false
-
-hex_full_count_df <- read_rds("data/s_curve/s_curve_hex_8.rds")
-
-shifted_hex_coord_df <- read_rds("data/s_curve/s_curve_hex_8_shifted_hex_coord_df.rds")
-
-bin_centroids_shift <- ggplot(data = hex_full_count_df, aes(x = c_x, y = c_y)) +
-  geom_point(color = "#bdbdbd") +
-  geom_point(data = shifted_hex_coord_df, aes(x = c_x, y = c_y), color = "#feb24c") +
-  coord_cartesian(xlim = c(-5, 8), ylim = c(-10, 10)) +
+```{r}
+#| echo: false
+#| warning: false
+
+## Initial x and y start
+initial_start_plot <- ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+  geom_text(data = df_bin_centroids, aes(x = c_x, y = c_y, label = hexID), size = 2) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  coord_equal() +
   theme_void() +
-  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+  theme(legend.position="bottom", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
         axis.title.x = element_blank(), axis.title.y = element_blank(),
@@
 
-hex_grid_shift <- ggplot(data = shifted_hex_coord_df, aes(x = x, y = y)) +
-  geom_polygon(fill = NA, color = "#feb24c", aes(group = polygon_id)) +
-  geom_polygon(data = hex_full_count_df, aes(x = x, y = y, group = polygon_id),
-               fill = NA, color = "#bdbdbd") +
-  coord_cartesian(xlim = c(-5, 8), ylim = c(-10, 10)) +
+## x and y start selected by MSE
+
+## UMAP
+## Decide by looking at MSE plot
+num_bins_x_s_curve <- 6
+num_bins_y_s_curve <- 14
+hex_size_s_curve <- 0.13
+
+hb_obj_s_curve_n <- hex_binning(data = umap_s_curve_scaled, x = "UMAP1", 
+                      y = "UMAP2", num_bins_x = num_bins_x_s_curve, 
+                      num_bins_y = num_bins_y_s_curve, x_start = -0.0652, y_start = -0.195, 
+                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_s_curve, col_start = "UMAP")
+
+## Data set with all possible centroids in the hexagonal grid
+all_centroids_df_n <- as.data.frame(do.call(cbind, hb_obj_s_curve_n$centroids))
+
+## Generate all coordinates of hexagons
+hex_grid_n <- as.data.frame(do.call(cbind, hb_obj_s_curve_n$hex_poly))
+
+## To obtain the standardise counts within hexbins
+counts_df_n <- as.data.frame(do.call(cbind, hb_obj_s_curve_n$std_cts))
+df_bin_centroids_n <- extract_hexbin_centroids(centroids_df = all_centroids_df_n, 
+                                             counts_df = counts_df_n)
+
+## remove low-density hexagons
+
+hex_grid_with_counts_n <- dplyr::left_join(hex_grid_n, counts_df_n, by = c("hex_poly_id" = "hb_id"))
+
+new_start_plot <- ggplot(data = hex_grid_with_counts_n, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+  geom_text(data= df_bin_centroids_n, aes(x = c_x, y = c_y, label = hexID), size = 2) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  coord_equal() +
   theme_void() +
@@
 
-## Before shift
-before_shift_plot <- ggplot(data = hex_full_count_df, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
-  geom_text(aes(x = c_x, y = c_y, label = hexID), size = 2) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  coord_equal() +
+hex_grid_ch_plot <- ggplot(data = hex_grid_with_counts_n, aes(x = x, y = y)) +
+  geom_polygon(fill = NA, color = "#feb24c", aes(group = hex_poly_id)) +
+  geom_polygon(data = hex_grid_with_counts, aes(x = x, y = y, group = hex_poly_id),
+               fill = NA, color = "#bdbdbd") +
+  #coord_cartesian(xlim = c(-5, 8), ylim = c(-10, 10)) +
   theme_void() +
-  theme(legend.position="bottom", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=8), #change legend title font size
+        legend.text = element_text(size=6)) +
+  guides(fill = guide_colourbar(title = "Standardized count")) +
+  annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
+
+bin_centroids_ch_plot <- ggplot(data = all_centroids_df, aes(x = c_x, y = c_y)) +
+  geom_point(color = "#bdbdbd") +
+  geom_point(data = all_centroids_df_n, aes(x = c_x, y = c_y), color = "#feb24c") +
+  #coord_cartesian(xlim = c(-5, 8), ylim = c(-10, 10)) +
+  theme_void() +
+  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
         axis.title.x = element_blank(), axis.title.y = element_blank(),
@@
   annotate(geom = 'text', label = "a", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3) 
-
-## After shift
-after_shift_plot <- ggplot(data = shifted_hex_coord_df, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
-  geom_text(aes(x = c_x, y = c_y, label = hexID), size = 2) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  coord_equal() +
-  theme_void() +
-  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=8), #change legend title font size
-        legend.text = element_text(size=6)) +
-  guides(fill = guide_colourbar(title = "Standardized count")) +
-  annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
-
-```
-
-```{r}
-#| eval: false
+```
+
+```{r}
 #| echo: false
@@
 
-bin_centroids_shift + hex_grid_shift +
+bin_centroids_ch_plot + hex_grid_ch_plot +
   plot_layout(guides='collect', ncol = 2) &
@@
 
-before_shift_plot + after_shift_plot +
+initial_start_plot + new_start_plot +
   plot_layout(guides='collect', ncol = 2) &
@@
 ```
-
-```{r}
-#| eval: false
-#| warning: false
-#| echo: false
-
-#MSE_df <- dplyr::bind_rows(MSE_df_1, MSE_df_2, MSE_df_3, MSE_df_4, MSE_df_5)
-
-#MSE_df$method <- factor(MSE_df$method, levels = c("tSNE", "UMAP", "PHATE", "TriMAP", "PaCMAP"))
-
-MSE_df <- read_rds("data/s_curve/s_curve_umap_shift_summary.rds") 
-MSE_df$method <- factor(MSE_df$method, levels = c("UMAP"))
-
-## To draw with AIC
-aic_plot <- ggplot(MSE_df |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "UMAP"), aes(x = shift,
-                                                                                 y = total_error,
-                                                                                 color = method
-)) +
-  geom_point() +
-  geom_line() +
-  geom_vline(xintercept = MSE_df |> dplyr::filter(total_mse == min(total_mse)) |> dplyr::pull(shift), linetype="dashed",
-                color = "red", size=0.5) +
-  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("AIC") +
-  xlab("shift value")
-## Effective number of bins along x-axis
-
-mse_plot <- ggplot(MSE_df |> dplyr::filter(method == "UMAP"), aes(x = shift,
-                                       y = total_mse,
-                                       color = method
-)) +
-  geom_point() +
-  geom_line() +
-   geom_vline(xintercept = MSE_df |> dplyr::filter(total_mse == min(total_mse)) |> dplyr::pull(shift), linetype="dashed",
-                color = "red", size=0.5) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("MSE") +
-  xlab("shift value")
-
-```
-
-```{r}
-#| eval: false
-#| echo: false
-#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset with different shift values.
-In this illustration, an identical shift is applied in both the x and y directions.
-The Mean Squared Error (MSE) plot exhibits a global minimum at $0$, indicating that no additional shifting is necessary for the current hexagonal grid to yield a well-fitted model."
-#| label: fig-diagnosticpltScurvehexbins
-##| out-width: 100%
-#| fig-pos: H
-
-aic_plot + mse_plot +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', ncol = 2) &
-  theme(legend.position='none', plot.tag = element_text(size = 8))
-```
--->
 
@@
 
-phate_gau_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tibble::as_tibble(PHATE_data_gau), x = "PHATE1", y = "PHATE2", hex_ratio = 50/sqrt(3)))) |>
-  dplyr::rename(c("PHATE1" = "scaled_PHATE1",
-                  "PHATE2" = "scaled_PHATE2")) |>
-  dplyr::mutate(ID = 1:NROW(PHATE_data_gau))
-
-## Decide by looking at MSE plot
-num_bins_x_phate_gau <- 60
-num_bins_y_phate_gau <- 1927
-hex_size_phate_gau <- 0.01
-### non-empty 187
-
-hb_obj_phate_gau <- hex_binning(data = phate_gau_scaled, x = "PHATE1", 
-                               y = "PHATE2", num_bins_x = num_bins_x_phate_gau, 
-                               num_bins_y = num_bins_y_phate_gau, x_start = NA, y_start = NA, 
-                               buffer_x = NA, buffer_y = NA, hex_size = hex_size_phate_gau, col_start = "PHATE")
-
-## Data set with all possible centroids in the hexagonal grid
-all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_phate_gau$centroids))
-
-## Generate all coordinates of hexagons
-hex_grid <- as.data.frame(do.call(cbind, hb_obj_phate_gau$hex_poly))
-
-## To obtain the standardise counts within hexbins
-counts_df <- as.data.frame(do.call(cbind, hb_obj_phate_gau$std_cts))
-df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
-                                             counts_df = counts_df)
-# ggplot(data = hex_grid, aes(x = x, y = y)) +
-#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
-#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red", size = 0.1) +
-#   coord_fixed()
-
-hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
-
-# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
-#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
-#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
-#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
-#   coord_fixed()
-
-phate_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_phate_gau$data_hb_id))
-
-model_object_phate_gau <- fit_highd_model( training_data = training_data_gau, 
-                                          nldr_df_with_id = phate_gau_scaled, 
-                                          x = "PHATE1", y = "PHATE2", 
-                                          num_bins_x = num_bins_x_phate_gau, 
-                                          num_bins_y = num_bins_y_phate_gau, 
-                                          x_start = NA, y_start = NA, 
-                                          buffer_x = NA, buffer_y = NA, 
-                                          hex_size = hex_size_phate_gau,
-                                          is_rm_lwd_hex = FALSE, 
-                                          benchmark_to_rm_lwd_hex = NA, 
-                                          col_start_2d = "PHATE", 
-                                          col_start_highd = "x")
-
-df_bin_centroids_phate_gau <- model_object_phate_gau$df_bin_centroids
-df_bin_phate_gau <- model_object_phate_gau$df_bin
+# phate_gau_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tibble::as_tibble(PHATE_data_gau), x = "PHATE1", y = "PHATE2", hex_ratio = 50/sqrt(3)))) |>
+#   dplyr::rename(c("PHATE1" = "scaled_PHATE1",
+#                   "PHATE2" = "scaled_PHATE2")) |>
+#   dplyr::mutate(ID = 1:NROW(PHATE_data_gau))
+# 
+# ## Decide by looking at MSE plot
+# num_bins_x_phate_gau <- 60
+# num_bins_y_phate_gau <- 1927
+# hex_size_phate_gau <- 0.01
+# ### non-empty 187
+# 
+# hb_obj_phate_gau <- hex_binning(data = phate_gau_scaled, x = "PHATE1", 
+#                                y = "PHATE2", num_bins_x = num_bins_x_phate_gau, 
+#                                num_bins_y = num_bins_y_phate_gau, x_start = NA, y_start = NA, 
+#                                buffer_x = NA, buffer_y = NA, hex_size = hex_size_phate_gau, col_start = "PHATE")
+# 
+# ## Data set with all possible centroids in the hexagonal grid
+# all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_phate_gau$centroids))
+# 
+# ## Generate all coordinates of hexagons
+# hex_grid <- as.data.frame(do.call(cbind, hb_obj_phate_gau$hex_poly))
+# 
+# ## To obtain the standardise counts within hexbins
+# counts_df <- as.data.frame(do.call(cbind, hb_obj_phate_gau$std_cts))
+# df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
+#                                              counts_df = counts_df)
+# # ggplot(data = hex_grid, aes(x = x, y = y)) +
+# #   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
+# #   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red", size = 0.1) +
+# #   coord_fixed()
+# 
+# hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
+# 
+# # ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
+# #   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+# #   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
+# #   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
+# #   coord_fixed()
+# 
+# phate_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_phate_gau$data_hb_id))
+# 
+# model_object_phate_gau <- fit_highd_model( training_data = training_data_gau, 
+#                                           nldr_df_with_id = phate_gau_scaled, 
+#                                           x = "PHATE1", y = "PHATE2", 
+#                                           num_bins_x = num_bins_x_phate_gau, 
+#                                           num_bins_y = num_bins_y_phate_gau, 
+#                                           x_start = NA, y_start = NA, 
+#                                           buffer_x = NA, buffer_y = NA, 
+#                                           hex_size = hex_size_phate_gau,
+#                                           is_rm_lwd_hex = FALSE, 
+#                                           benchmark_to_rm_lwd_hex = NA, 
+#                                           col_start_2d = "PHATE", 
+#                                           col_start_highd = "x")
+# 
+# df_bin_centroids_phate_gau <- model_object_phate_gau$df_bin_centroids
+# df_bin_phate_gau <- model_object_phate_gau$df_bin
+
+df_bin_centroids_phate_gau <- read_rds("data/five_gau_clusters/df_bin_centroids_phate_gau.rds")
+df_bin_phate_gau <- read_rds("data/five_gau_clusters/df_bin_phate_gau.rds")
 
@@
 
-df_all_phate_gau <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID), phate_data_with_hb_id_gau)
+# df_all_phate_gau <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID), phate_data_with_hb_id_gau)
 

### 2024-03-26 paper.tex
Clear the repo

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, interior hidden, breakable, enhanced, boxrule=0pt, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, interior hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}, breakable, boxrule=0pt, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 Hexagonating NLDR data (see Figure~\ref{fig-meth} Step 2) involves
-partitioning the 2D embedding data into hexagonal bins, a technique
-commonly referred to as hexagonal binning
-\citep[\citet{article66}]{Carr1987}. This method use a hexagonal grid to
-create a bivariate histogram that effectively visualizes the structure
-of high-D data. Hexagons, one of only three regular polygons capable of
-tessellating a plane \citep{Carr2013}, offer unique advantages due to
-their symmetry of nearest neighbors and maximal number of sides for such
-tessellations \citep{Dan2023}. This geometric property makes hexagons
-more efficient in covering the plane compared to other regular
-tessellations and reduces visual bias when displaying data densities
-\citep{Dan2023}. In our algorithm, we aim to conduct regular hexagonal
-binning (\textbf{cite quollr paper}), involving the computation of
-hexagonal grid configurations, the generation of the grid, and the
-assignment of the 2D embedding to hexagons.
+partitioning the NLDR data into hexagonal bins, a technique commonly
+referred to as hexagonal binning \citep[\citet{article66}]{Carr1987}.
+This method use a hexagonal grid to create a bivariate histogram that
+effectively visualizes the structure of high-D data. Hexagons, one of
+only three regular polygons capable of tessellating a plane
+\citep{Carr2013}, offer unique advantages due to their symmetry of
+nearest neighbors and maximal number of sides for such tessellations
+\citep{Dan2023}. This geometric property makes hexagons more efficient
+in covering the plane compared to other regular tessellations and
+reduces visual bias when displaying data densities \citep{Dan2023}.
+
+In our algorithm, we aim to conduct regular hexagonal binning, involving
+the computation of hexagonal grid configurations, the generation of the
+grid, and the assignment of the NLDR data points to hexagons.
+
+\textbf{(a) Computation of hexagonal grid configurations}
+
+In the computation of hexagonal bin configurations, there are mainly two
+configurations to consider: the number of bins along the x and y axes.
+The first step involves in determining the hexagonal size (\(s\)) and
+the buffer along the x and y axes (\(buffer_{x}\) and \(buffer_{y}\)).
+Here, the hexagonal size (\(s\)) represents the radius of the outer
+circle surrounding the hexagon. The buffer is important in expanding the
+boundary to accommodate potential outliers or edge cases.
+
+When computing the number of bins along the x and y axes (\(b_1\) and
+\(b_2\)), the process begins by calculating the range of the respective
+2D embedding component. After that, the range is adjusted to accommodate
+the buffer amount. Then, the spacing between hexagons is determined
+based on \(s\). Finally, the number of bins along the axis is computed
+by dividing the adjusted range by the spacing.
+
+For the computation of the number of bins along the x-axis (\(b_1\))
+(see Equation~\ref{eq-equation12}), the range of the first embedding
+component is defined as \(r_1\), while the buffer amount along the
+x-axis is denoted as \(buffer_{x}\), and the horizontal spacing is
+represented by \(h\) (see Equation~\ref{eq-equation11}). On the other
+hand, the number of bins along the y-axis (\(b_2\)) (see
+Equation~\ref{eq-equation14}) is computed based on the range of the
+second embedding component (\(r_2\)), the buffer amount along the y-axis
+(\(buffer_{y}\)), and the vertical spacing (\(v\)) (see
+Equation~\ref{eq-equation13}).
+
+\begin{equation}\protect\hypertarget{eq-equation11}{}{
+ h = \sqrt{3} * s
+}\label{eq-equation11}\end{equation}
+
+\begin{equation}\protect\hypertarget{eq-equation12}{}{
+ b_1 = \frac{r_1 + buffer_{x}}{h}
+}\label{eq-equation12}\end{equation}
+
+\begin{equation}\protect\hypertarget{eq-equation13}{}{
+  v = 1.5 * s
+}\label{eq-equation13}\end{equation}
+
+\begin{equation}\protect\hypertarget{eq-equation14}{}{
+ b_2 = \frac{r_2 + buffer_{y}}{v}
+}\label{eq-equation14}\end{equation}
+
+\textbf{(b) Generation of the hexagonal grid}
+
+In this process, the first step involves generating centroids within the
+hexagonal grid. This begins by defining the starting coordinates for the
+hexagons. With the number of bins computed along the x and y axes, along
+with the horizontal and vertical spacing, the centroids are iteratively
+computed starting from the hexagonal starting coordinates.
+
+Once the centroids for the hexagonal grid are obtained, the next step is
+to compute the hexagonal coordinates. For example, if the centroid of a
+hexagonal bin is defined as \((C_x, C_y)\), the hexagonal coordinates
+can be computed using \(d_x\) (see Equation~\ref{eq-equation15}) and
+\(d_y\) (see Equation~\ref{eq-equation16}) (see
+Figure~\ref{fig-hexcoord}).
+
+\begin{equation}\protect\hypertarget{eq-equation15}{}{
+d_x = \frac{h}{2}
+}\label{eq-equation15}\end{equation}
+
+\begin{equation}\protect\hypertarget{eq-equation16}{}{
+d_y = \frac{2* v * 1.15}{\sqrt{3}} 
+}\label{eq-equation16}\end{equation}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=1\textheight]{figures/hex_coord.png}
+
+}
+
+\caption{\label{fig-hexcoord}The hexagonal coordinates computed for the
+centroid \((C_x, C_y)\).}
+
+\end{figure}
+
+\textbf{(c) Assignment of the NLDR data points to hexagons}
+
+After obtaining the centroids of the hexagonal bins for the entire grid,
+the process of assigning NLDR points to hexagons involves determining
+the nearest hexagonal bin for each NLDR point using the 2D Euclidean
+distance. Then, the hexagonal ID of the nearest hexagon is assigned to
+the corresponding NLDR points.
 
@@
 for the \textbf{UMAP} embeddings of \textbf{S-curve} training data: (a)
-b = 12 (3, 4), h = 0.62, (b) b = 84 (6, 14), h = 0.13, and (c) b = 336
-(12, 28), h = 0.06. The hexbins are colored based on the density of
+b = 12 (3, 4), s = 0.62, (b) b = 84 (6, 14), s = 0.13, and (c) b = 336
+(12, 28), s = 0.06. The hexbins are colored based on the density of
 points, with darker colors indicating higher point density and yellow
@@
 
-\begin{figure}[H]
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-modelScurve-1.pdf}
-
-}
-
-\caption{\label{fig-modelScurve}(a) Full hexagonal grid with bin
-centroids, (b) Model constructed in 2D with long edges, and (c) Model
-constructed in 2D after removing the long edges.}
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-diagnosticpltScurvelwd-1.pdf}
+
+}
+
+\caption{\label{fig-diagnosticpltScurvelwd}Goodness of fit statistics
+from UMAP applied to training S-curve dataset with different benchmark
+values to remove the low-density heaxgons. What is the effective
+benchmark value to remove the low-density heaxgons? The MSE plot have a
+steep slope at the beginning, indicating that a smaller benchmark causes
+a small amount of error. Then, the slope gradually increases or level
+up, indicating that a higher number of benchmark values generates a
+higher error. Using the reverse elbow method, when the benchmark value
+is set to 0.2, the slope of the Mean Squared Error (MSE) plot
+experiences a sudden and noticeable change, resembling an elbow-like
+shape. This point indicates that higher benchmark values remove the
+necessary bins as well which lead to distruct the 2D structure.}
+
+\end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-rmlowdenshex-1.pdf}
+
+}
+
+\caption{\label{fig-rmlowdenshex}Goodness of f}
 
@@
 
+\begin{figure}[H]
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-modelScurve-1.pdf}
+
+}
+
+\caption{\label{fig-modelScurve}(a) Full hexagonal grid with bin
+centroids, (b) Model constructed in 2D with long edges, and (c) Model
+constructed in 2D after removing the long edges.}
+
+\end{figure}
+
 \hypertarget{starting-point-of-the-hexagonal-grid}{%
@@
 tessellating the \(xy\) plane over the set (range(\(x\)), range(\(y\)))
-(see \textbf{?@fig-scurveshifthexgridsexp} (b)). In that case, bin
-centroids are defined as shown in \textbf{?@fig-scurveshifthexgridsexp}
-(a) with gray colour. Rather than sticking to the typical hexagonal
-grid, introducing a meaningful shift in both the x and y directions
-presents an opportunity for an improved 2D model. Therefore,
-investigating this shift in the hexagonal grid is an important parameter
-to consider.
-
-As shown in \textbf{?@fig-scurveshifthexgrids}, the shifting influences
-the distribution of points and number of non-empty bins, impacting the
-resulting 2D model. According to
+(see Figure~\ref{fig-scurveshifthexgridsexp} (b)). In that case, bin
+centroids are defined as shown in
+Figure~\ref{fig-scurveshifthexgridsexp} (a) with gray colour. Rather
+than sticking to the typical hexagonal grid, introducing a meaningful
+shift in both the x and y directions presents an opportunity for an
+improved 2D model. Therefore, investigating this shift in the hexagonal
+grid is an important parameter to consider.
+
+As shown in Figure~\ref{fig-scurveshifthexgrids}, the shifting
+influences the distribution of points and number of non-empty bins,
+impacting the resulting 2D model. According to
 \textbf{?@fig-diagnosticpltScurvehexbins}, the 2D model with a total of
@@
 because the lowest MSE occurs when no shift is introduced.
+
+\includegraphics{paper_files/figure-pdf/unnamed-chunk-36-1.pdf}
+
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-scurveshifthexgridsexp-1.pdf}
+
+}
+
+\caption{\label{fig-scurveshifthexgridsexp}(a) Visualization of bin
+centroids and (b) hexagonal grid, showcasing the configuration before
+(colored in gray) and after (colored in light orange) shifting the
+starting point by an identical amount applied in both x and y
+directions, with a shift value of \(0.537285\).}
+
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-scurveshifthexgrids-1.pdf}
+
+}
+
+\caption{\label{fig-scurveshifthexgrids}Hexbin plots shows the
+distribution of points before and after shifting the starting point of
+the hexagonal grid generated for UMAP applied to the training S-curve
+dataset. (a) Hexbin plot before shifting, and (b) Hexbin plot after
+shifting (identical shift applied in both x and y directions, with a
+shift value of \(0.537285\)).}
+
+\end{figure}
 
@@
 of the 2D bin centroid data (\(C_{x_ij}\)) across all observations and
-dimensions (see Equation~\ref{eq-equation11}).
-
-\begin{equation}\protect\hypertarget{eq-equation11}{}{
+dimensions (see Equation~\ref{eq-equation10}).
+
+\begin{equation}\protect\hypertarget{eq-equation10}{}{
 \text{Error} = \sum_{j = 1}^{n}\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2
-}\label{eq-equation11}\end{equation}
+}\label{eq-equation10}\end{equation}
 
@@
 effectively preserving both local and global structures (see video link
-of \textbf{?@fig-modelfiveGau} (a)). UMAP, on the other hand, also
+of Figure~\ref{fig-modelfiveGau} (a)). UMAP, on the other hand, also
 presents five distinct clusters but with a more flattened surface
-appearance (see video link of \textbf{?@fig-modelfiveGau} (b)). The
+appearance (see video link of Figure~\ref{fig-modelfiveGau} (b)). The
 PHATE model shows five separated clusters resembling triangles or
 partial triangles but struggles to capture local structures (see video
-link of \textbf{?@fig-modelfiveGau} (c)). TriMAP and PaCMAP both display
-five well-separated clusters with flat surfaces, each capturing
+link of Figure~\ref{fig-modelfiveGau} (c)). TriMAP and PaCMAP both
+display five well-separated clusters with flat surfaces, each capturing
 within-cluster variation to varying extents. Despite the various 2D
 representations, all NLDR techniques preserve the global structure (see
-video link of \textbf{?@fig-modelfiveGau} (d), (e) respectively).
-However, tSNE, effectively capture both local and global structures, as
-indicated by lower AIC values in Figure~\ref{fig-diagnosticpltGau} (a).
+video link of Figure~\ref{fig-modelfiveGau} (d), (e) respectively).
+However, tSNE, effectively capture both local and global structures.
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticpltGau-1.pdf}
-
-}
-
-\caption{\label{fig-diagnosticpltGau}Goodness of fit statistics from
-different NLDR techniques applied to training five spherical Gaussian
-cluster dataset. What is the best NLDR technique to represent the
-original data in 2D?}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-modelfiveGau-1.pdf}
+
+}
+
+\caption{\label{fig-modelfiveGau}Is there a best model to represent the
+original data in 2D space or are they all providing equivalent
+information?, (a) Model generated with 21 and 28 bins along the x and y
+axes, respectively, using a hexagonal size of 0.03 in the 2D space with
+tSNE, (b) Model generated with 60 and 79 bins along the x and y axes,
+respectively, using a hexagonal size of 0.01 in the 2D space with UMAP,
+(c) Model generated with 60 and 1927 bins along the x and y axes,
+respectively, using a hexagonal size of 0.01 in the 2D space with PHATE,
+(d) Model generated with 46 and 61 bins along the x and y axes,
+respectively, using a hexagonal size of 0.013 in the 2D space with
+TriMAP, and (e) Model generated with 31 and 40 bins along the x and y
+axes, respectively, using a hexagonal size of 0.02 in the 2D space with
+PaCMAP.}
+
+\end{figure}
+
+\begin{figure}[H]
+
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/five_gau_clusters/sc_tsne_1.png}
+
+}
+
+}
+
+\subcaption{\label{fig-gau1_sc1}}
+\end{minipage}%
+%
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/five_gau_clusters/sc_tsne_2.png}
+
+}
+
+}
+
+\subcaption{\label{fig-gau1_sc2}}
+\end{minipage}%
+%
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/five_gau_clusters/sc_tsne_3.png}
+
+}
+
+}
+
+\subcaption{\label{fig-gau1_sc3}}
+\end{minipage}%
+
+\caption{\label{fig-gau1_sc}Screen shots of the \textbf{langevitour} of
+the five Gaussian clusters dataset, shows the model with tSNE in high-D,
+a video of the tour animation is available at
+(\url{https://youtu.be/oQxEb4wRdHI}).}
+
+\end{figure}
+
+\begin{figure}[H]
+
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/five_gau_clusters/sc_umap_1.png}
+
+}
+
+}
+
+\subcaption{\label{fig-gau2_sc1}}
+\end{minipage}%
+%
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/five_gau_clusters/sc_umap_2.png}
+
+}
+
+}
+
+\subcaption{\label{fig-gau2_sc2}}
+\end{minipage}%
+%
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/five_gau_clusters/sc_umap_3.png}
+
+}
+
+}
+
+\subcaption{\label{fig-gau2_sc3}}
+\end{minipage}%
+
+\caption{\label{fig-gau2_sc}Screen shots of the \textbf{langevitour} of
+the five Gaussian clusters dataset, shows the model with UMAP in high-D,
+a video of the tour animation is available at
+(\url{https://youtu.be/JW49csPpDx4}).}
+
+\end{figure}
+
+\begin{figure}[H]
+
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/five_gau_clusters/sc_phate_1.png}
+
+}
+
+}
+
+\subcaption{\label{fig-gau3_sc1}}
+\end{minipage}%
+%
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/five_gau_clusters/sc_phate_2.png}
+
+}
+
+}
+
+\subcaption{\label{fig-gau3_sc2}}
+\end{minipage}%
+%
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/five_gau_clusters/sc_phate_3.png}
+
+}
+
+}
+
+\subcaption{\label{fig-gau3_sc3}}
+\end{minipage}%
+
+\caption{\label{fig-gau3_sc}Screen shots of the \textbf{langevitour} of
+the five Gaussian clusters dataset, shows the model with PHATE in
+high-D, a video of the tour animation is available at
+(\url{https://youtu.be/xqhRODt8Z-o}).}
+
+\end{figure}
+
+\begin{figure}[H]
+
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/five_gau_clusters/sc_trimap_1.png}
+
+}
+
+}
+
+\subcaption{\label{fig-gau4_sc1}}
+\end{minipage}%
+%
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/five_gau_clusters/sc_trimap_2.png}
+
+}
+
+}
+
+\subcaption{\label{fig-gau4_sc2}}
+\end{minipage}%
+%
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/five_gau_clusters/sc_trimap_3.png}
+
+}
+
+}
+
+\subcaption{\label{fig-gau4_sc3}}
+\end{minipage}%
+
+\caption{\label{fig-gau4_sc}Screen shots of the \textbf{langevitour} of
+the five Gaussian clusters dataset, shows the model with TriMAP in
+high-D, a video of the tour animation is available at
+(\url{https://youtu.be/X-uHSB4VoZ0}).}
+
+\end{figure}
+
+\begin{figure}[H]
+
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/five_gau_clusters/sc_pacmap_1.png}
+
+}
+
+}
+
+\subcaption{\label{fig-gau5_sc1}}
+\end{minipage}%
+%
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/five_gau_clusters/sc_pacmap_2.png}
+
+}
+
+}
+
+\subcaption{\label{fig-gau5_sc2}}
+\end{minipage}%
+%
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/five_gau_clusters/sc_pacmap_3.png}
+
+}
+
+}
+
+\subcaption{\label{fig-gau5_sc3}}
+\end{minipage}%
+
+\caption{\label{fig-gau5_sc}Screen shots of the \textbf{langevitour} of
+the five Gaussian clusters dataset, shows the model with PaCMAP in
+high-D, a video of the tour animation is available at
+(\url{https://youtu.be/UxvU8HQd7JM}).}
 
@@
 
+In the field of single-cell studies, a common analytical task involves
+clustering to identify groups of cells with similar expression profiles.
+Analysts often turn to NLDR techniques to verify and identify these
+clusters and explore developmental trajectories. To illustrate the
+importance of NLDR techniques and parameter selection in identifying
+clusters, Human Peripheral Blood Mononclear Cells (PBMC3k) dataset
+\citep{chen2023} is used. In a study by \citet{chen2023}, this dataset
+was used to demonstrate how UMAP represents clusters (see
+Figure~\ref{fig-umapauthor}). As shown in Figure~\ref{fig-umapauthor},
+there are three distinct and well-separated clusters.
+
 \begin{figure}[H]
@@
 
-\caption{\label{fig-umapauthor}(a) 2D layout from UMAP (n\_neighbors =
-30, min\_dist = 0.3) applied for the PBMC3k dataset. Is this a best
+\caption{\label{fig-umapauthor}2D layout from UMAP (n\_neighbors = 30,
+min\_dist = 0.3) applied for the PBMC3k dataset. Is this a best
 representation of the original data?}
@@
 
-\begin{figure}[H]
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-msepltpbmc-1.pdf}
-
-}
-
-\caption{\label{fig-msepltpbmc}Goodness of fit statistics from UMAP
-applied to training PBMC3k dataset. What is the effective number of bins
-in UMAP to create a 2D model? The MSE plot have a steep slope at the
-beginning, indicating that a smaller number of bins causes a larger
-amount of error. Then, the slope gradually declines or level off,
-indicating that a higher number of bins generates a smaller error. Using
-the elbow method, when the total number of bins is set to 588, the slope
-of the MSE plot experiences a sudden and noticeable change, resembling
-an elbow-like shape. This point indicates that adding less bins does not
-enough to capture the data structure.}
-
-\end{figure}
+To determine whether the UMAP representation with the parameter choice
+suggested by \citet{chen2023} preserves the original data structure, we
+visualize the model constructed with UMAP overlaid on the high-D data
+(model-in-data space). The figures in Figure~\ref{fig-pbmc1_sc} show
+three well-separated clusters, indicating that the suggested UMAP
+representation preserves the global structure (see
+Figure~\ref{fig-umapauthor}). However, as shown in
+Figure~\ref{fig-pbmc1_sc}, these clusters are close to each other in
+high-D space. Also, nonlinear continuity patterns and high-density
+patches within the clusters are observed (see
+Figure~\ref{fig-pbmc1_sc}). Therefore, the suggested UMAP representation
+(see Figure~\ref{fig-umapauthor}) does not accurately preserve the local
+structure of the PBMC3k dataset.
 
@@
 
-\caption{\label{fig-nldervisPBMCUMAP}(a) Model in the 2D space overlaid
-on UMAP data, and (b) Error occurred by the high-D model for high-D data
-and the error mapped into 2D NLDR points.}
-
-\end{figure}
-
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-72-1.pdf}
+\caption{\label{fig-nldervisPBMCUMAP}(a) Model generated with 21 and 28
+bins along x and y axes with hexagonal size 0.03 in the 2D space
+overlaid on UMAP data, and (b) high-D model error in model space.}
+
+\end{figure}
+
+\begin{figure}[H]
+
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/pbmc3k/sc_1.png}
+
+}
+
+}
+
+\subcaption{\label{fig-pbmc1_sc1}}
+\end{minipage}%
+%
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/pbmc3k/sc_2.png}
+
+}
+
+}
+
+\subcaption{\label{fig-pbmc1_sc2}}
+\end{minipage}%
+%
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/pbmc3k/sc_3.png}
+
+}
+
+}
+
+\subcaption{\label{fig-pbmc1_sc3}}
+\end{minipage}%
+
+\caption{\label{fig-pbmc1_sc}Screen shots of the \textbf{langevitour} of
+the PBMC3k data set, shows the model in high-D, a video of the tour
+animation is available at (\url{https://youtu.be/Gnhh4hfsbbg}).}
+
+\end{figure}
+
+To find the best NLDR representation for PBMC3k dataset, we observe the
+absolute error for different number of non-empty bins and select the one
+with the lowest error. As shown in Figure~\ref{fig-abserrorPBMC}, tSNE
+with a perplexity set to \(51\), has the lowest error. Therefore, tSNE
+with a perplexity of \(51\) is used as the best representation for
+PBMC3k dataset.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-abserrorPBMC-1.pdf}
+
+}
+
+\caption{\label{fig-abserrorPBMC}Absolute error from UMAP and tSNE
+applied to training PBMC3k dataset with diffferent parameter choices.
+What is the best parameter choice to create the model? The residual plot
+have a steep slope at the beginning, indicating that a smaller number of
+non-empty bins causes a larger amount of error. Then, the slope
+gradually declines or level off, indicating that a higher number of
+non-empty bins generates a smaller error. Using the elbow method, it was
+observed that when the number of non-empty bins is set to 439, the
+lowest error occurred with the parameters perplexity: 51.}
+
+\end{figure}
+
+As shown in Figure~\ref{fig-tsnesuggest}, there are three well-separated
+clusters, although they are located close to each other. Additionally,
+nonlinear structures can be observed within the clusters (see
+Figure~\ref{fig-tsnesuggest}). Hence, the data structure that UMAP
+failed to capture for the PBMC3k dataset is successfully captured by
+tSNE.
+
+\begin{figure}[H]
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-tsnesuggest-1.pdf}
+
+}
+
+\caption{\label{fig-tsnesuggest}2D layout from tSNE (perplexity = 51)
+applied for the PBMC3k dataset. Is this a best representation of the
+original data?}
+
+\end{figure}
+
+Next, the model is fitted for tSNE, and the resulting model is
+visualized in the data space. The model shows some quirks, as shown in
+Figure~\ref{fig-pbmc2_sc}. The large cluster is divided into three
+subclusters, which are in close to each other.
+
+\begin{figure}[H]
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-nldervisPBMCtSNE-1.pdf}
+
+}
+
+\caption{\label{fig-nldervisPBMCtSNE}(a) Model generated with 31 and 40
+bins along x and y axes with hexagonal size 0.02 in the 2D space
+overlaid on tSNE data, and (b) high-D model error in model space.}
+
+\end{figure}
+
+\begin{figure}[H]
+
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/pbmc3k/sc_4.png}
+
+}
+
+}
+
+\subcaption{\label{fig-pbmc2_sc4}}
+\end{minipage}%
+%
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/pbmc3k/sc_5.png}
+
+}
+
+}
+
+\subcaption{\label{fig-pbmc2_sc5}}
+\end{minipage}%
+%
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/pbmc3k/sc_6.png}
+
+}
+
+}
+
+\subcaption{\label{fig-pbmc2_sc6}}
+\end{minipage}%
+
+\caption{\label{fig-pbmc2_sc}Screen shots of the \textbf{langevitour} of
+the PBMC3k data set, shows the model in high-D, a video of the tour
+animation is available at (\url{https://youtu.be/R0EroJcGbas}).}
+
+\end{figure}
 
@@
 \citep{lecun2010}. \citet{Yingfan2021} used this dataset to demonstrate
-how the PaCMAP preserves local structure. We select the 2D embedding of
-PaCMAP for the handwritten digit 1. As shown in
+how the PaCMAP preserves local structure. We selected the 2D embedding
+of PaCMAP for the handwritten digit 1. As shown in
 Figure~\ref{fig-pacmapauthor}, we observe a nonlinear structure present
-in 2D space. To answer the following questions,
+in 2D space. Furthermore, the angle of digit 1 images varies along the
+nonlinear structure (see Figure~\ref{fig-sampleimg}). To answer the
+following questions,
 
@@
 
+According to Figure~\ref{fig-mnist1_sc1}, the nonlinear continuity
+structure observed in the 2D representation of PaCMAP is also visible
+when visualizing the model overlaid on the data space. This provides
+evidence that the model accurately captures the structure of the high-D
+data. Also, the model has a twisted pattern within the nonlinear
+structure, which is an addition pattern of data that cannot be seen in
+the 2D representation (see Figure~\ref{fig-mnist1_sc2}).
+
+In order to visually assess the preservation of local structure, it's
+important to consider the edges in the model. In a model that
+effectively preserves local structure, there shouldn't be any long edges
+in high-D space. However, as shown in Figure~\ref{fig-mnist1_sc3}, some
+long edges exist in high-D space that are not identified as long edges
+in the 2D representation.
+
 \begin{figure}[H]
@@
 \caption{\label{fig-sampleimg}Images of the handwritten digit 1 are
-distributed along the left-to-right of first embedding axis of the
+distributed along the right-to-left of first embedding axis of the
 PaCMAP nonlinear structure.}
@@
 
-{\centering \includegraphics{paper_files/figure-pdf/fig-diagnosticpltMNIST-1.pdf}
-
-}
-
-\caption{\label{fig-diagnosticpltMNIST}Goodness of fit statistics from
-PaCMAP applied to training MNIST dataset. What is the effective number
-of bins in PaCMAP to create a 2D model? The MSE plot have a steep slope
-at the beginning, indicating that a smaller number of bins causes a
-larger amount of error. Then, the slope gradually declines or level off,
-indicating that a higher number of bins generates a smaller error. Using
-the elbow method, when the total number of bins is set to 180, the slope
-of the Mean Squared Error (MSE) plot experiences a sudden and noticeable
-change, resembling an elbow-like shape. This point indicates that adding
-less bins does not enough to capture the data structure.}
-
-\end{figure}
-
-\begin{figure}[H]
-
 {\centering \includegraphics{paper_files/figure-pdf/fig-nldervisMNISTPACMAP-1.pdf}
@@
 
-\caption{\label{fig-nldervisMNISTPACMAP}(a) Model in the 2D space
+\caption{\label{fig-nldervisMNISTPACMAP}(a) Model generated with 12 and
+15 bins along x and y axes with hexagonal size 0.06 in the 2D space
 overlaid on PaCMAP data, and (b) high-D model error in model space.}
+
+\end{figure}
+
+\begin{figure}[H]
+
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/mnist/sc_1.png}
+
+}
+
+}
+
+\subcaption{\label{fig-mnist1_sc1}}
+\end{minipage}%
+%
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/mnist/sc_2.png}
+
+}
+
+}
+
+\subcaption{\label{fig-mnist1_sc2}}
+\end{minipage}%
+%
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/mnist/sc_3.png}
+
+}
+
+}
+
+\subcaption{\label{fig-mnist1_sc3}}
+\end{minipage}%
+
+\caption{\label{fig-mnist1_sc}Screen shots of the \textbf{langevitour}
+of the MNIST digit 1 data set, shows the model-in-data space, a video of
+the tour animation is available at
+(\url{https://youtu.be/zq2GM9qvUNA}).}
+
+\end{figure}
+
+In terms of model error, some data points show high error (see
+Figure~\ref{fig-nldervisMNISTPACMAP}) due to their deviation from the
+typical high-D data structure, acting as outliers. This is evident when
+comparing the images associated with these high model error points (see
+Figure~\ref{fig-sampleimgerror}) to others (see
+Figure~\ref{fig-sampleimg}).
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-sampleimgerror-1.pdf}
+
+}
+
+\caption{\label{fig-sampleimgerror}Images of handwritten digit 1 which
+occur large model error.}
 

### 2024-03-26 2_cross_validation_with_total_num_bins.R

@@ -1,119 +1,0 @@
-####### S-curve example with 5-Fold cross validation #####################
-library(readr)
-
-set.seed(20240110)
-
-source("quollr_code.R", local = TRUE)
-
-## Import data
-data <- read_rds("data/s_curve/s_curve.rds")
-training_data <- read_rds("data/s_curve/s_curve_training.rds")
-UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap.rds")
-
-shape_val_umap_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
-                                                          x = UMAP1, y = UMAP2)
-
-# Create indices for 5-fold cross-validation
-num_obs <- NROW(UMAP_s_curve)
-fold_size <- floor(num_obs / 5)
-
-# Initialize variables to store performance metrics
-mse_list <- numeric(5)
-
-num_bins_vec <- 1:13
-
-vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations",
-                                     "total_error", "total_mse", "num_bins_x"))  ## Define column names
-
-eval_data_training <- dplyr::bind_rows(vec)[0, ]
-eval_data_training <- eval_data_training |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-# Perform 5-fold cross-validation
-for (fold in 1:5) {
-  # Define the indices for the current fold
-  start_index <- (fold - 1) * fold_size + 1
-  end_index <- fold * fold_size
-
-  # Split the data into training and validation sets
-  X_train <- UMAP_s_curve[-(start_index:end_index), ]
-  X_val <- UMAP_s_curve[start_index:end_index, ]
-
-  for (i in 1:length(num_bins_vec)) {
-    # Fit the model for training
-    model_object <- fit_high_d_model(training_data = training_data,
-                                     nldr_df_with_id = UMAP_s_curve,
-                                     x = "UMAP1", y = "UMAP2", num_bins_x = num_bins_vec[i],
-                                     shape_val = shape_val_umap_s_curve,
-                                     is_bin_centroid = TRUE,
-                                     is_rm_lwd_hex = FALSE,
-                                     benchmark_to_rm_lwd_hex = NA,
-                                     is_avg_high_d = TRUE, column_start_text = "x")
-
-    centroid_df_training <- model_object$df_bin_centroids
-    avg_df_training <- model_object$df_bin
-
[…]

### 2024-03-26 Supplementary/bin_mean_and_weighted_mean_highD.R

@@ -1,147 +1,0 @@
-library(dplyr)
-library(snedata)
-library(ggflowchart)
-library(purrr) ## map function
-library(gridExtra) ## for grid.arrange
-library(rsample)
-library(DT)
-library(ggbeeswarm)
-library(ggplot2)
-library(readr)
-
-library(Rtsne)
-library(umap)
-library(phateR)
-library(reticulate)
-library(patchwork)
-
-library(grid)
-set.seed(20230531)
-
-source(paste0(here::here(), "/paper-nldr-vis-algorithm/quollr_code.R"))
-source(paste0(here::here(), "/paper-nldr-vis-algorithm/nldr_code.R"))
-
-data <- read_csv(paste0(here::here(), "/paper-nldr-vis-algorithm/data/s_curve.csv"))
-
-data <- data |>
-  mutate(ID = row_number())
-
-data_split <- initial_split(data)
-training_data <- training(data_split) |>
-  arrange(ID)
-test_data <- testing(data_split) |>
-  arrange(ID)
-
-UMAP_fit <- umap(training_data |> dplyr::select(-ID), n_neighbors = 50, n_components =  2)
-
-UMAP_data <- UMAP_fit$layout |>
-  as.data.frame()
-names(UMAP_data)[1:(ncol(UMAP_data))] <- paste0(rep("UMAP",(ncol(UMAP_data))), 1:(ncol(UMAP_data)))
-
-UMAP_data <- UMAP_data |>
-  mutate(ID = training_data$ID)
-
-num_bins_x <- calculate_effective_x_bins(.data = UMAP_data, x = UMAP1,
-                                         cell_area = 1)
-
-shape_value <- calculate_effective_shape_value(.data = UMAP_data,
-                                               x = UMAP1, y = UMAP2)
-
-## To extract bin centroids
-hexbin_data_object <- extract_hexbin_mean(nldr_df = UMAP_data, num_bins = num_bins_x, shape_val = shape_value)
-
-df_bin_centroids <- hexbin_data_object$hexdf_data
-
-## To generate the grid
-hexbin_data_object_n <- extract_hexbin_centroids(UMAP_data, num_bins_x, shape_val)
-df_bin_centroids_n <- hexbin_data_object_n$hexdf_data
-## Data set with all possible centroids in the hexagonal grid
-full_centroid_df <- generate_full_grid_centroids(df_bin_centroids_n)
-## Generate all coordinates of hexagons
-hex_grid <- full_hex_grid(full_centroid_df)
-
[…]

### 2024-03-26 applications/harps/3_gen_model_perplexity_40_rm_lwd.R

@@ -1,153 +1,0 @@
-library(readr)
-library(dplyr)
-
-set.seed(20240110)
-source("quollr_code.R", local = TRUE)
-
-## Import data
-training_data_1 <- read_rds("data/harps/harps_data.rds")
-training_data_1 <- training_data_1 |>
-  mutate(ID = 1:NROW(training_data_1))
-
-clusters <- training_data_1$cluster
-select_names <-  names(training_data_1)[-c(18, length(training_data_1))]
-
-training_data_1 <- training_data_1 |> select(-cluster)
-names(training_data_1) <- append(paste0("x", 1:17), "ID")
-
-tSNE_data_harps <- read_rds("data/harps/harps_tsne_40.rds")
-tSNE_data_harps <- tSNE_data_harps |>
-  select(tSNE1, tSNE2) |>
-  mutate(ID = 1:NROW(tSNE_data_harps))
-
-### tSNE
-
-num_bins_tsne_harps <- calculate_effective_x_bins(.data = tSNE_data_harps, x = tSNE1,
-                                                  cell_area = 1) #43
-
-
-shape_val_tsne_harps <- calculate_effective_shape_value(.data = tSNE_data_harps,
-                                                        x = tSNE1, y = tSNE2)
-## To extract bin centroids
-hexbin_data_object_tsne_harps <- extract_hexbin_centroids(nldr_df = tSNE_data_harps, num_bins = num_bins_tsne_harps, shape_val = shape_val_tsne_harps, x = tSNE1, y = tSNE2)
-
-df_bin_centroids_tsne_harps <- hexbin_data_object_tsne_harps$hexdf_data
-
-# Identify bins with low-density
-identify_rm_bins <- find_low_density_hexagons(df_bin_centroids_tsne_harps, num_bins_tsne_harps, benchmark_rm_hex = 0.01)
-
-df_bin_centroids_tsne_harps <- df_bin_centroids_tsne_harps |>
-  filter(!(hexID %in% identify_rm_bins))
-
-tSNE_data_with_hb_id_harps <- tSNE_data_harps |>
-  dplyr::mutate(hb_id = hexbin_data_object_tsne_harps$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_tsne_harps <- dplyr::bind_cols(training_data_1 |> dplyr::select(-ID), tSNE_data_with_hb_id_harps)
-
-## Averaged on high-D
[…]

### 2024-03-26 applications/zeisel/gen_hex_grid_zeisel_phate.R

@@ -1,148 +1,0 @@
-library(readr)
-library(dplyr)
-
-set.seed(20240110)
-
-source("quollr_code.R", local = TRUE)
-source("nldr_code.R", local = TRUE)
-
-## Import data
-df_zeisel <- read_rds("data/zeisel/zeisel.rds")
-training_data_zeisel <- read_rds("data/zeisel/zeisel_training.rds")
-
-PHATE_zeisel <- read_rds("data/zeisel/zeisel_phate.rds")
-
-plot_PHATE_2D(PHATE_zeisel) + #ggtitle("(b)") +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-num_bins_zeisel <- calculate_effective_x_bins(.data = PHATE_zeisel, x = PHATE1,
-                                              cell_area = 1) ##1
-
-
-shape_value_zeisel <- calculate_effective_shape_value(.data = PHATE_zeisel,
-                                                      x = PHATE1, y = PHATE2) ## 0.5991139
-
-## To extract bin centroids
-hexbin_data_object_zeisel <- extract_hexbin_centroids(nldr_df = PHATE_zeisel,
-                                                      num_bins = num_bins_zeisel,
-                                                      shape_val = shape_value_zeisel, x = PHATE1, y = PHATE2)
-
-df_bin_centroids_zeisel <- hexbin_data_object_zeisel$hexdf_data
-
-##########
-
-## Data set with all possible centroids in the hexagonal grid
-
-full_centroid_df <- generate_full_grid_centroids(df_bin_centroids_zeisel)
-
-## Generate all coordinates of hexagons
-hex_grid <- full_hex_grid(full_centroid_df)
-
-hex_full_count_df <- generate_full_grid_info(df_bin_centroids_zeisel)
-
-write_rds(hex_full_count_df, "data/zeisel/zeisel_phate_hex_1.rds")
-
-##########
-
-ggplot(data = hex_full_count_df, aes(x = x, y = y)) +
[…]

### 2024-03-26 applications/zeisel/test_opt_rm_lg.R

@@ -1,157 +1,0 @@
-library(readr)
-library(dplyr)
-
-set.seed(20240110)
-
-source("quollr_code.R", local = TRUE)
-source("nldr_code.R", local = TRUE)
-
-## Import data
-df_s_curve <- read_rds("data/s_curve/s_curve.rds")
-training_data_s_curve <- read_rds("data/s_curve/s_curve_training.rds")
-
-UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap.rds")
-
-num_bins_s_curve <- 6
-
-shape_value_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
-                                                      x = UMAP1, y = UMAP2) ## 1.259938
-
-## To extract bin centroids
-hexbin_data_object_s_curve <- extract_hexbin_centroids(nldr_df = UMAP_s_curve,
-                                                      num_bins = num_bins_s_curve,
-                                                      shape_val = shape_value_s_curve, x = UMAP1, y = UMAP2)
-
-df_bin_centroids_s_curve <- hexbin_data_object_s_curve$hexdf_data
-
-UMAP_s_curve_with_hb_id <- UMAP_s_curve |>
-  dplyr::mutate(hb_id = hexbin_data_object_s_curve$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_s_curve <- dplyr::bind_cols(training_data_s_curve |> dplyr::select(-ID), UMAP_s_curve_with_hb_id)
-
-## Averaged on high-D
-df_bin_s_curve <- avg_highD_data(.data = df_all_s_curve, column_start_text = "x")
-
-## Triangulate bin centroids
-tr1_object_s_curve <- triangulate_bin_centroids(df_bin_centroids_s_curve, x, y)
-tr_from_to_df_s_curve <- generate_edge_info(triangular_object = tr1_object_s_curve)
-
-## Compute 2D distances
-distance_s_curve <- cal_2D_dist(.data = tr_from_to_df_s_curve)
-
-## To find the benchmark value
-benchmark <- find_benchmark_value(.data = distance_s_curve, distance_col = distance)
-benchmark <- 4.8
-
-
-distance <- distance_s_curve
-
-###############
-
-distance <- distance %>%
-  rename(distance_2D = distance)
-
-#num_pca <- 4
-
-#high_d <- training_data_s_curve
-high_d <- df_bin_s_curve |> select(-hb_id)
-#rownames(high_d) <- df_bin_s_curve$hb_id
[…]

### 2024-03-26 extra/extra.R

@@ -1,67 +1,0 @@
-```{r}
-#| echo: false
-#| warning: false
-
-error_df_lg_1 <- read_rds("data/s_curve/s_curve_summary_lg_tsne.rds")
-error_df_lg_2 <- read_rds("data/s_curve/s_curve_summary_lg_umap.rds")
-error_df_lg_3 <- read_rds("data/s_curve/s_curve_summary_lg_phate.rds")
-error_df_lg_4 <- read_rds("data/s_curve/s_curve_summary_lg_trimap.rds")
-error_df_lg_5 <- read_rds("data/s_curve/s_curve_summary_lg_pacmap.rds")
-
-error_df_lg <- dplyr::bind_rows(error_df_lg_1, error_df_lg_2, error_df_lg_3, error_df_lg_4, error_df_lg_5)
-
-error_df_lg$method <- factor(error_df_lg$method, levels = c("tSNE", "UMAP", "PHATE", "TriMAP", "PaCMAP"))
-
-## To draw with AIC
-aic_plot_lg <- ggplot(error_df_lg |> dplyr::filter(method == "UMAP"), aes(x = benchmark_rm_lg,
-                                                                          y = total_error,
-                                                                          color = method
-)) +
-  geom_point() +
-  geom_vline(xintercept = 1.347, linetype="dashed",
-             color = "red", size=0.5) +
-  geom_line() +
-  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
-  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
-  ylab("AIC") +
-  xlab("benchmark value")
-## Effective number of bins along x-axis
-
-mse_plot_lg <- ggplot(error_df_lg |> dplyr::filter(method == "UMAP"), aes(x = benchmark_rm_lg,
-                                                                          y = total_mse,
-                                                                          color = method
-)) +
-  geom_point() +
-  geom_vline(xintercept = 1.347, linetype="dashed",
[…]

### 2024-03-26 extra/extra_1.R

@@ -1,219 +1,0 @@
-```{r}
-#| warning: false
-#| echo: false
-
-
-
-## tSNE
-
-shape_value_curve <- calculate_effective_shape_value(.data = tSNE_data,
-                                                     x = tSNE1, y = tSNE2)
-
-num_bins_vec <- 1:18 ## Number of bins along the x-axis
-
-vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names
-
-eval_data_training <- dplyr::bind_rows(vec)[0, ]
-eval_data_training <- eval_data_training |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(num_bins_vec)) {
-
-  model_object <- fit_high_d_model(training_data = training_data, nldr_df_with_id = tSNE_data, x = "tSNE1", y = "tSNE2", num_bins_x = num_bins_vec[i], shape_val = shape_value_curve,
-                                   is_bin_centroid = TRUE,
-                                   is_rm_lwd_hex = FALSE,
-                                   benchmark_to_rm_lwd_hex = NA,
-                                   is_avg_high_d = TRUE, column_start_text = "x")
-
-  centroid_df_training <- model_object$df_bin_centroids
-  avg_df_training <- model_object$df_bin
-
-  # pred_df_training <- predict_2d_embeddings(test_data = training_data, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, type_NLDR = "tSNE")
-
-  pred_df_training <- predict_hex_id(df_bin_centroids = centroid_df_training, nldr_df_test = tSNE_data, x = "tSNE1", y = "tSNE2")
-
-
-  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, col_start = "x")
-
-  eval_df_training <- eval_df_training |>
-    mutate(num_bins_x = num_bins_vec[i])
-
-  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)
-
-
-}
-
-
-## Add new column with data types
-
-eval_data_training <- eval_data_training |>
-  mutate(data_type = "training")
-
-MSE_df_1 <- eval_data_training |>
-  dplyr::mutate(method = "tSNE")
-```
-```{r}
-#| warning: false
-#| echo: false
-## PAHTE
-## Prediction
-
[…]

### 2024-03-26 hex_bin_label_shape_1.R

@@ -1,170 +1,0 @@
-## Shape value issue
-cell_area <- 0.5
-
-num_bins_x <- calculate_effective_x_bins(.data = UMAP_data, x = UMAP1,
-                                         cell_area = 0.5)
-
-shape_val <- calculate_effective_shape_value(.data = UMAP_data,
-                                             x = UMAP1, y = UMAP2)
-
-
-#shape_val <- 1
-
-hb_object <- hexbin::hexbin(x = UMAP_data |> pull(UMAP1),
-                            y = UMAP_data |> pull(UMAP2),
-                            xbins = num_bins_x, IDs = TRUE,
-                            shape = shape_val)
-
-hexdf_data <- tibble::tibble(tibble::as_tibble(hexbin::hcell2xy(hb_object)),  hexID = hb_object@cell, counts = hb_object@count/max(hb_object@count))
-
-hex_grid <- expand.grid(UMAP_data |> pull(UMAP1), UMAP_data |> pull(UMAP2))
-
-hb <- hexbin::hexbin(x = hex_grid |> pull(Var1),
-                     y = hex_grid |> pull(Var2),
-                     xbins = num_bins_x, IDs = TRUE,
-                     shape = shape_val)
-
-hexdf <- tibble::tibble(tibble::as_tibble(hexbin::hcell2xy(hb)),  hexID = hb@cell, counts = hb@count/max(hb@count))
-
-hexdf <- hexdf %>%
-  mutate(counts = ifelse((hexID %in% (hexdf_data |> pull(hexID))),counts, NA))
-
-embedding_hb <- create_hexbin(.data = training_data, nldr_df = UMAP_data,
-                              embedding_1 = UMAP1, embedding_2 = UMAP2,
-                              num_bins = num_bins_x, shape_val = shape_val, apply_pca = FALSE)$df_new
-
-full_grid <- full_join(hexdf, embedding_hb, by = c("hexID" = "hb_id"))
-
-
-
-ggplot(full_grid, aes(x = x, y = y, fill = counts, hexID = hexID, label = hexID)) +
-  geom_hex(stat = "identity", color = "#969696") +
-  geom_label() +
-  geom_point(aes(x = umap1, y = umap2), na.rm = TRUE) +
[…]

### 2024-03-26 quollr_code.R

@@ -1,1261 +1,0 @@
-calculate_effective_x_bins <- function(.data, x = UMAP1, cell_area = 1){
-
-  if (any(is.na(.data |> dplyr::pull({{ x }})))) {
-    stop("NAs present")
-  }
-
-  if (any(is.infinite(.data |> dplyr::pull({{ x }})))) {
-    stop("Inf present")
-  }
-
-  if ((cell_area <= 0) || (is.infinite(cell_area))) {
-    stop("Invalid cell area value")
-
-  }
-
-  ## To compute the diameter of the hexagon
-  cell_diameter <- sqrt(2 * cell_area / sqrt(3))
-
-  ## To compute the range along x-axis
-  xwidth <- diff(range(.data |>
-                         dplyr::pull({{ x }})))
-
-  num_bins <- ceiling(xwidth/cell_diameter)
-  #num_bins <- floor(xwidth/cell_diameter + 1.5001)
-  num_bins
-
-}
-
-calculate_effective_shape_value <- function(.data, x = UMAP1, y = UMAP2){
-
-  if (any(is.na(.data |> dplyr::pull({{ x }}))) || any(is.na(.data |> dplyr::pull({{ y }})))) {
-    stop("NAs present")
-  }
-
-  if (any(is.infinite(.data |> dplyr::pull({{ x }}))) || any(is.infinite(.data |> dplyr::pull({{ y }})))) {
-    stop("Inf present")
-  }
-
-  if ((length(.data |> dplyr::pull({{ x }})) == 1) || (length(.data |> dplyr::pull({{ y }})) == 1)) {
-    stop("Presence one observation only")
-
-  }
-
-  ## To compute the range along x-axis
-  xwidth <- diff(range(.data |> dplyr::pull({{ x }})))
-  ## To compute the range along y-axis
-  yheight <- diff(range(.data |> dplyr::pull({{ y }})))
-
-
-  shape <- yheight/xwidth
-  shape
-}
-
-extract_hexbin_centroids <- function(nldr_df, num_bins, shape_val = 1, x = UMAP1, y = UMAP2) {
-
-  ## To create the hexbin object
-  hb_data <- hexbin::hexbin(x = nldr_df |> dplyr::pull({{ x }}),
-                            y = nldr_df |> dplyr::pull({{ y }}),
-                            xbins = num_bins, IDs = TRUE,
-                            shape = shape_val)
-
-  ## To create the hexbin centroid info dataset
[…]

### 2024-03-26 scripts/1_true_model_with_hex_11_S_curve.R

@@ -1,178 +1,0 @@
-library(readr)
-
-set.seed(20240110)
-
-source("quollr_code.R", local = TRUE)
-
-## Import data
-training_data <- read_rds("data/s_curve/s_curve_training.rds")
-UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap.rds")
-
-## UMAP
-
-num_bins_umap_s_curve <- 11
-shape_val_umap_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
-                                                          x = UMAP1, y = UMAP2) ## 1.259938
-## To extract bin centroids
-hexbin_data_object_umap_s_curve <- extract_hexbin_centroids(nldr_df = UMAP_s_curve, num_bins = num_bins_umap_s_curve, shape_val = shape_val_umap_s_curve, x = UMAP1, y = UMAP2)
-
-df_bin_centroids_umap_s_curve <- hexbin_data_object_umap_s_curve$hexdf_data
-
-UMAP_data_with_hb_id_s_curve <- UMAP_s_curve |>
-  dplyr::mutate(hb_id = hexbin_data_object_umap_s_curve$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_umap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), UMAP_data_with_hb_id_s_curve)
-
-## Averaged on high-D
-df_bin_umap_s_curve <- avg_highD_data(.data = df_all_umap_s_curve)
-
-## Triangulate bin centroids
-tr1_object_umap_s_curve <- triangulate_bin_centroids(df_bin_centroids_umap_s_curve, x, y)
-tr_from_to_df_umap_s_curve <- generate_edge_info(triangular_object = tr1_object_umap_s_curve)
-
-# ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
-#   geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
-#   geom_point(size = 2, colour = "#33a02c") +
-#   coord_equal()
-
-
-## Compute 2D distances
-distance_umap_s_curve <- cal_2D_dist(.data = tr_from_to_df_umap_s_curve)
-
-
-## To find the benchmark value
-benchmark_umap_s_curve <- find_benchmark_value(.data = distance_umap_s_curve, distance_col = distance)
-#benchmark_umap_s_curve <- 1.346694
-
-
[…]

### 2024-03-26 scripts/change_starting_point_param.R

@@ -1,193 +1,0 @@
-library(readr)
-library(ggplot2)
-
-
-set.seed(20240110)
-
-source("quollr_code.R", local = TRUE)
-
-## Read UMAP data
-UMAP_data <- read_rds("data/s_curve/s_curve_umap.rds")
-
-
-####num_bins_x = 8####################
-num_bins <- 8
-
-cell_diameter <- sqrt(2 * 1 / sqrt(3))
-cell_diameter
-
-cell_diameter/2
-
-shape_val <- calculate_effective_shape_value(.data = UMAP_data, x = UMAP1, y = UMAP2)
-
-hexbin_data_object <- extract_hexbin_centroids(UMAP_data, num_bins, shape_val)
-
-df_bin_centroids <- hexbin_data_object$hexdf_data
-
-## Data set with all possible centroids in the hexagonal grid
-
-full_centroid_df <- generate_full_grid_centroids(df_bin_centroids)
-
-## Generate all coordinates of hexagons
-hex_grid <- full_hex_grid(full_centroid_df)
-
-ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white",
-                                                          color = "black", aes(group = id)) +
-  geom_point()
-
-hex_full_count_df <- generate_full_grid_info(df_bin_centroids)
-#write_rds(hex_full_count_df, file = "data/s_curve/s_curve_hex_8.rds")
-
-ggplot(data = hex_full_count_df, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
-  geom_text(aes(x = c_x, y = c_y, label = hexID)) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff")
-
-## Filter centroids with their hexIDs
-hexbin_coord_all <- hex_full_count_df |>
-  dplyr::select(c_x, c_y, hexID) |>
-  dplyr::distinct()
-
-hexbin_coord_all_new <- hexbin_coord_all |>
-  dplyr::mutate(c_x = c_x - (cell_diameter/2),
-         c_y = c_y - (cell_diameter/2)) |>
-  dplyr::rename(c("x" = "c_x",
-           "y" = "c_y"))
-
-ggplot(data = hex_full_count_df, aes(x = c_x, y = c_y)) +
-  geom_point(color = "black") +
[…]

### 2024-03-26 scripts/gen_summary_phate_lg_rm_n.R

@@ -1,89 +1,0 @@
-library(readr)
-library(dplyr)
-
-set.seed(20240110)
-
-source("quollr_code.R", local = TRUE)
-source("nldr_code.R", local = TRUE)
-
-## Import data
-df_s_curve <- read_rds("data/s_curve/s_curve.rds")
-training_data_s_curve <- read_rds("data/s_curve/s_curve_training.rds")
-
-PHATE_s_curve <- read_rds("data/s_curve/s_curve_phate.rds")
-
-num_bins_s_curve <- 8
-
-shape_value_s_curve <- calculate_effective_shape_value(.data = PHATE_s_curve,
-                                                       x = PHATE1, y = PHATE2) ## 1.259938
-
-## To extract bin centroids
-hexbin_data_object_s_curve <- extract_hexbin_centroids(nldr_df = PHATE_s_curve,
-                                                       num_bins = num_bins_s_curve,
-                                                       shape_val = shape_value_s_curve, x = PHATE1, y = PHATE2)
-
-df_bin_centroids_s_curve <- hexbin_data_object_s_curve$hexdf_data
-
-## Identify bins with low-density
-identify_rm_bins <- find_low_density_hexagons(df_bin_centroids_s_curve, num_bins_s_curve, benchmark_rm_hex = 0.06)
-
-df_bin_centroids_s_curve <- df_bin_centroids_s_curve |>
-  filter(!(hexID %in% identify_rm_bins))
-
-PHATE_s_curve_with_hb_id <- PHATE_s_curve |>
-  dplyr::mutate(hb_id = hexbin_data_object_s_curve$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_s_curve <- dplyr::bind_cols(training_data_s_curve |> dplyr::select(-ID), PHATE_s_curve_with_hb_id)
-
-## Averaged on high-D
-df_bin_s_curve <- avg_highD_data(.data = df_all_s_curve, column_start_text = "x")
-
-## Triangulate bin centroids
-tr1_object_s_curve <- triangulate_bin_centroids(df_bin_centroids_s_curve, x, y)
-tr_from_to_df_s_curve <- generate_edge_info(triangular_object = tr1_object_s_curve)
-
-## Compute 2D distances
-distance_s_curve <- cal_2D_dist(.data = tr_from_to_df_s_curve)
-
-## To find the benchmark value
[…]

### 2024-03-26 scripts/gen_summary_with_diff_lwd_values_phate_n.R

@@ -1,105 +1,0 @@
-library(readr)
-library(dplyr)
-library(ggplot2)
-
-set.seed(20240110)
-
-source("quollr_code.R", local = TRUE)
-
-data <- read_rds("data/s_curve/s_curve.rds")
-training_data <- read_rds("data/s_curve/s_curve_training.rds")
-test_data <- read_rds("data/s_curve/s_curve_test.rds")
-
-PHATE_data <- read_rds("data/s_curve/s_curve_phate.rds")
-
-shape_value_curve <- calculate_effective_shape_value(.data = PHATE_data,
-                                                     x = PHATE1, y = PHATE2)
-
-num_bins_x <- 8 ## Number of bins along the x-axis (looking at mse plot)
-
-vec <- stats::setNames(rep("", 4), c("number_of_bins", "number_of_observations", "total_error", "total_mse"))  ## Define column names
-
-eval_data_training <- dplyr::bind_rows(vec)[0, ]
-eval_data_training <- eval_data_training |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = PHATE_data, nldr_df_test = PHATE_data, num_bins = num_bins_x, shape_val = shape_value_curve, x = "PHATE1", y = "PHATE2", col_start = "x")
-pred_df_training <- pred_df_training_object$pred_data
-centroid_df_training_all <- pred_df_training_object$df_bin_centroids
-
-hexbin_data_object_training <- pred_df_training_object$hexbin_data_object
-
-benchmark_rm_hex_vec <- seq(0, 1, by=0.1)
-
-
-for (i in 1:length(benchmark_rm_hex_vec)) {
-
-  centroid_df_training <- centroid_df_training_all |>
-    filter(std_counts > benchmark_rm_hex_vec[i])
-
-  ## Add hexbin Id to 2D embeddings
-  PHATE_data_with_hb_id <- PHATE_data |>
-    mutate(hb_id = hexbin_data_object_training$hb_data@cID)
-
-  ## To generate a data set with high-D and 2D training data
-  df_all <- dplyr::bind_cols(training_data |> dplyr::select(-ID), PHATE_data_with_hb_id)
-
-  ## Averaged on high-D
-  avg_df_training <- avg_highD_data(.data = df_all, column_start_text = "x")
-
-  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training,
[…]

### 2024-03-26 scripts/gen_summary_with_diff_lwd_values_trimap.R

@@ -1,107 +1,0 @@
-library(readr)
-library(dplyr)
-library(ggplot2)
-
-set.seed(20240110)
-
-source("quollr_code.R", local = TRUE)
-
-data <- read_rds("data/s_curve/s_curve.rds")
-training_data <- read_rds("data/s_curve/s_curve_training.rds")
-test_data <- read_rds("data/s_curve/s_curve_test.rds")
-
-TriMAP_data <- read_rds("data/s_curve/s_curve_trimap.rds")
-
-shape_value_curve <- calculate_effective_shape_value(.data = TriMAP_data,
-                                                     x = TriMAP1, y = TriMAP2)
-
-num_bins_x <- 6 ## Number of bins along the x-axis (looking at mse plot)
-
-benchmark_rm_hex_vec <- sample(seq(0, 0.5, by=0.001), 20) |> sort()
-
-vec <- stats::setNames(rep("", 4), c("number_of_bins", "number_of_observations", "total_error", "total_mse"))  ## Define column names
-
-eval_data_training <- dplyr::bind_rows(vec)[0, ]
-eval_data_training <- eval_data_training |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = TriMAP_data, nldr_df_test = TriMAP_data, num_bins = num_bins_x, shape_val = shape_value_curve, x = "TriMAP1", y = "TriMAP2", col_start = "x")
-pred_df_training <- pred_df_training_object$pred_data
-centroid_df_training_all <- pred_df_training_object$df_bin_centroids
-
-hexbin_data_object_training <- pred_df_training_object$hexbin_data_object
-
-for (i in 1:length(benchmark_rm_hex_vec)) {
-
-  ## Identify bins with low-density
-  identify_rm_bins <- find_low_density_hexagons(centroid_df_training_all, num_bins_x, benchmark_rm_hex = benchmark_rm_hex_vec[i])
-
-  centroid_df_training <- centroid_df_training_all |>
-    filter(!(hexID %in% identify_rm_bins))
-
-  ## Add hexbin Id to 2D embeddings
-  TriMAP_data_with_hb_id <- TriMAP_data |>
-    mutate(hb_id = hexbin_data_object_training$hb_data@cID)
-
-  ## To generate a data set with high-D and 2D training data
-  df_all <- dplyr::bind_cols(training_data |> dplyr::select(-ID), TriMAP_data_with_hb_id)
-
-  ## Averaged on high-D
[…]

### 2024-03-26 paper.qmd
Added bechmark value to remove long edge section

@@
 hex_full_count_df_loop2 <- hex_grid_with_counts |>
-  dplyr::mutate(type = if_else(std_counts <= 0.1, "low", "high"))
+  dplyr::mutate(type = if_else(std_counts <= 0.2, "low", "high"))
 
@@
 
+rm_centroid_df <- df_bin_centroids_low |> 
+  dplyr::filter(hexID %in% identify_rm_bins)
+
 p2_n_rm_correct <- ggplot(data = hex_full_count_df_loop2, aes(x = x, y = y)) +
   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
-  geom_text(data = df_bin_centroids_low |> dplyr::filter(hexID %in% identify_rm_bins), color = "red", aes(x = c_x, y = c_y, group = hexID, label = "X"), size = 0.5) +
+  geom_text(data = rm_centroid_df, color = "red", aes(x = c_x, y = c_y, group = hexID, label = "X"), size = 0.5) +
   #geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
@@
 
-#### Benchmark value to remove long edges
-
-Achieving a smooth representation in 2D space is crucial, and one factor influencing this smoothness is the presence of long edges (see @fig-modelScurvermlgimp).
-These long edges occur when a line connects points that are distant in the triangular mesh, impacting the overall smoothness of the mesh.
-
-To investigate the effect of removing long edges on the MSE, we analyzed the MSE for various benchmark values.
-Surprisingly, the results indicated that the MSE remained consistent across different benchmark values, as shown in @fig-diagnosticpltScurvelgrm.
-It's important to note that edges are defined only in the 2D model and do not extend to higher dimensions.
-Consequently, removing edges does not affect the model in higher dimensions.
-
-There is no definitive rule for determining the benchmark value to remove long edges.
-However, we have a method to find a default value (see Appendix).
-Adjusting values around this default can help to find benchmark value to remove long edges, contributing to the construction of a smoother 2D representation (see @fig-diagnosticpltScurvelgrm).
-
-The following steps will help to find a suitable value to remove long edges:
-
-1. Plot the distribution of the 2D distances
-
-```{r}
-#| echo: false
-#| fig-cap: "Distribution of 2D Euclidean distances."
+#### Benchmark value for removing long edges
+
+Creating a smooth 2D representation (see @fig-modelScurve (c)) requires removing edges that connect distant bin centroids within the triangular mesh (see @fig-modelScurve (b)).
+These edges only exist in the 2D model and do not extend into higher dimensions, ensuring that their removal does not impact the model in higher dimensions.
+While specific criteria for determining the benchmark value to remove long edges do not exist, the following steps provide an approach to identifying a suitable threshold:
+
+1. Plot the distribution of the 2D Euclidean distances.
+2. Identify the largest difference between consecutive distance values.
+3. Take the largest distance value corresponding to this difference as the benchmark value.
+
+```{r}
+#| echo: false
+#| fig-cap: "Distribution of 2D Euclidean distances between bin centroids of the triangular mesh generated with S-curve UMAP data."
 #| label: fig-distScurve
@@
 ```
-
-2. Find a value which is greater than smallest value
 
@@
 ```
-
-<!--The removal of long edges is a critical step to create a smoother representation  by iteratively eliminating hexagons with excessive distances between centroids.
-This process eliminates outliers and noise while preserving essential local relationships within the data.
-To achieve this, distances between vertices are sorted, and unique distance values are extracted.
-By setting a threshold based on the largest difference between consecutive distance values, long edges are identified and removed.
-This refinement step contributes to enhancing the quality of the triangular mesh, ensuring a more accurate representation of the data structure.-->
-<!--
-```{r}
-#| warning: false
-#| echo: false
-#| eval: false
-
-## Import data
-
-UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap.rds")
-
-## UMAP
-
-num_bins_umap_s_curve <- 8
-shape_val_umap_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
-                                                          x = UMAP1, y = UMAP2) ## 1.259938
-## To extract bin centroids
-hexbin_data_object_umap_s_curve <- extract_hexbin_centroids(nldr_df = UMAP_s_curve, num_bins = num_bins_umap_s_curve, shape_val = shape_val_umap_s_curve, x = UMAP1, y = UMAP2)
-
-df_bin_centroids_umap_s_curve <- hexbin_data_object_umap_s_curve$hexdf_data
-
-UMAP_data_with_hb_id_s_curve <- UMAP_s_curve |>
-  dplyr::mutate(hb_id = hexbin_data_object_umap_s_curve$hb_data@cID)
-
-## To generate a data set with high-D and 2D training data
-df_all_umap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), UMAP_data_with_hb_id_s_curve)
-
-## Averaged on high-D
-df_bin_umap_s_curve <- avg_highD_data(.data = df_all_umap_s_curve)
-
-## Triangulate bin centroids
-tr1_object_umap_s_curve <- triangulate_bin_centroids(df_bin_centroids_umap_s_curve, x, y)
-tr_from_to_df_umap_s_curve <- generate_edge_info(triangular_object = tr1_object_umap_s_curve)
-
-# ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
-#   geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
-#   geom_point(size = 2, colour = "#33a02c") +
-#   coord_equal()
-
-## Compute 2D distances
-distance_umap_s_curve <- cal_2d_dist(.data = tr_from_to_df_umap_s_curve)
-
-distance_plot_s_curve <- plot_dist(distance_umap_s_curve) +
-  #ggtitle("(b)" ) +
-  ylab(expression(d^{(2)})) +
-  theme(axis.text = element_text(size = 5),
-        axis.title = element_text(size = 12))
-
-## To find the benchmark value
-benchmark_umap_s_curve <- find_benchmark_value(.data = distance_umap_s_curve, distance_col = distance)
-benchmark_umap_s_curve <- 1.346694
-
-# colour_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve,
-#                   triangular_object = tr1_object_umap_s_curve, distance_col = distance)
-
-trimesh_umap_s_curve <- ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
-  geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
-  geom_point(size = 2, colour = "#33a02c") +
-  coord_equal()
-
-trimesh_umap_s_curve <- trimesh_umap_s_curve +
-  #ggtitle("(a)") +
-  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  theme_light() +
-  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
-  ) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
-# theme(axis.text = element_text(size = 5),
-#       axis.title = element_text(size = 7))
-
-trimesh_gr_umap_s_curve <- colour_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve, triangular_object = tr1_object_umap_s_curve, distance_col = distance)
-
-trimesh_gr_umap_s_curve <- trimesh_gr_umap_s_curve +
-  geom_point(size = 2, colour = "#33a02c") +
-  #ggtitle("(b)") +
-  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  theme_light() +
-  #coord_equal() +
-  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
-  ) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-# ### Define type column
-# df <- df_all_umap_s_curve |>
-#   dplyr::select(tidyselect::starts_with("x")) |>
-#   dplyr::mutate(type = "data") ## original dataset
-# 
-# df_b <- df_bin_umap_s_curve |>
-#   dplyr::filter(hb_id %in% df_bin_centroids_umap_s_curve$hexID) |>
-#   dplyr::select(-hb_id) |>
-#   dplyr::mutate(type = "model") ## Data with summarized mean
-# 
-# df_exe <- dplyr::bind_rows(df_b, df)
-# 
-# distance_df_small_edges <- distance_umap_s_curve %>%
-#   dplyr::filter(distance < benchmark_umap_s_curve)
-# 
-# distance_df_long_edges <- distance_umap_s_curve %>%
-#   dplyr::filter(distance >= benchmark_umap_s_curve)
-# 
-# distance_df_edges <- dplyr::bind_rows(distance_df_small_edges, distance_df_long_edges)
-# 
-# langevitour::langevitour(df_exe[1:(length(df_exe)-1)], lineFrom = distance_df_edges$from , lineTo = distance_df_edges$to, group = df_exe$type, pointSize = 3, levelColors = c("#6a3d9a", "#33a02c"), lineColors = append(rep("black", length(distance_df_small_edges$from)), rep("red", length(distance_df_long_edges$from))))
-#
-
-```
-
-```{r}
-#| eval: false
-#| warning: false
-#| echo: false
-#| label: fig-modelScurverdistcount
-#| fig-cap: "plots to identify benchmark values"
-##| out-width: 100%
-#| fig-pos: H
-
-cell_count_plot_s_curve + distance_plot_s_curve +
-  #plot_annotation(tag_levels = 'a') +
-  plot_layout(ncol = 2) &
-  theme(legend.position='none', plot.tag = element_text(size = 8))
-```
-
-```{r}
-#| eval: false
-#| warning: false
-#| echo: false
-#| label: fig-modelScurvermlgimp
-#| fig-cap: "What does the 2D model, constructed using UMAP (applied to the S-curve dataset), look like in high dimensions with long edges?
-(<https://youtu.be/pSL_G_n7-fM>) (a) 2D model without removing the long edges, and (b) 2D model with long edges coloured by red.
-The default benchmark value is used to identify long edges."
-##| out-width: 100%
-#| fig-pos: H
-
-trimesh_umap_s_curve + trimesh_gr_umap_s_curve +
-  #plot_annotation(tag_levels = 'a') +
-  plot_layout(ncol = 2) &
-  theme(legend.position='none', plot.tag = element_text(size = 8))
-```
--->
 
@@
 
-diff_start_df <- expand.grid(x_start = seq(min_x_start, max_x_start, by = 0.01), 
-                             y_start = seq(min_y_start, max_y_start, by = 0.01))
+diff_start_df <- expand.grid(x_start = seq(min_x_start, max_x_start, length.out = 10), 
+                             y_start = seq(min_y_start, max_y_start, length.out = 10)) |>
+  add_row(x_start = min(umap_s_curve_scaled[["UMAP1"]]) - (sqrt(3) * hex_size_s_curve/2),
+          y_start = min(umap_s_curve_scaled[["UMAP2"]]) - (1.5 * hex_size_s_curve/2))
 
@@
 
-for (i in 1:23) { #NROW(diff_start_df)
+diff_start_df <- diff_start_df[c(1:5, 11:15,21:25, 101),]
+
+for (i in 1:NROW(diff_start_df)) { #NROW(diff_start_df)
 
@@
 
-## If same total number of bins occurred only select ones with minimum error
-### Obtain duplicate bins
-dupli_bins <- mse_df |> 
-  dplyr::count(num_non_empty_bins) |> 
-  dplyr::filter(n > 1) |> 
-  dplyr::pull(num_non_empty_bins)
-
-### Group split by duplicated bins
-duplicate_df_list <- mse_df |>
-  dplyr::filter(num_non_empty_bins %in% dupli_bins) |>
-  dplyr::arrange(num_non_empty_bins) |>
-  dplyr::group_split(num_non_empty_bins)
-
-### Obtain one row from duplicates which have lowest error and hexsize
-duplicate_df <- data.frame(matrix(nrow = 0, ncol = 0))
-
-for (i in 1:length(duplicate_df_list)) {
-
-  dd <- duplicate_df_list[[i]] |>
-    dplyr::filter(mse == min(duplicate_df_list[[i]]$mse))
-
-  duplicate_df <- dplyr::bind_rows(duplicate_df, dd)
-
-}
-
-### Obtain the mse_df with not duplicated bins
-not_dupli_df <- mse_df |>
-  dplyr::filter(!(num_non_empty_bins %in% dupli_bins))
-
-### Combine duplicated and not duplicated(corrected) bins dfs
-mse_df <- dplyr::bind_rows(not_dupli_df, duplicate_df)
-
-mse_plot <- ggplot(mse_df, aes(x = num_non_empty_bins,
-                               y = mse
+# ## If same total number of bins occurred only select ones with minimum error
+# ### Obtain duplicate bins
+# dupli_bins <- mse_df |> 
+#   dplyr::count(num_non_empty_bins) |> 
+#   dplyr::filter(n > 1) |> 
+#   dplyr::pull(num_non_empty_bins)
+# 
+# ### Group split by duplicated bins
+# duplicate_df_list <- mse_df |>
+#   dplyr::filter(num_non_empty_bins %in% dupli_bins) |>
+#   dplyr::arrange(num_non_empty_bins) |>
+#   dplyr::group_split(num_non_empty_bins)
+# 
+# ### Obtain one row from duplicates which have lowest error and hexsize
+# duplicate_df <- data.frame(matrix(nrow = 0, ncol = 0))
+# 
+# for (i in 1:length(duplicate_df_list)) {
+#   
+#   dd <- duplicate_df_list[[i]] |>
+#     dplyr::filter(mse == min(duplicate_df_list[[i]]$mse)) 
+#   
+#   duplicate_df <- dplyr::bind_rows(duplicate_df, dd)
+#   
+# }
+# 
+# ### Obtain the mse_df with not duplicated bins
+# not_dupli_df <- mse_df |>
+#   dplyr::filter(!(num_non_empty_bins %in% dupli_bins))
+# 
+# ### Combine duplicated and not duplicated(corrected) bins dfs
+# mse_df <- dplyr::bind_rows(not_dupli_df, duplicate_df)
+
+mse_plot_start_point <- ggplot(mse_df, aes(x = x_start,
+                               y = y_start, colour = mse
 )) +
   geom_point() +
-  geom_line() +
+  #geom_line() +
   # geom_vline(xintercept = 84, linetype="solid",
@@
   theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
         axis.title = element_text(size = 7),
         axis.text = element_text(size = 7)) +
-  #scale_colour_manual(values = c("#377eb8")) +
-  ylab("MSE") +
-  xlab("number of non-empty bins")
-
-mse_plot
+  viridis::scale_color_viridis(option = "D") +
+  labs(x = "x start", y = "y start", color = "MSE") 
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset with different benchmark values to remove the low-density heaxgons.
+What is the effective benchmark value to remove the low-density heaxgons?
+The MSE plot have a steep slope at the beginning, indicating that a smaller benchmark causes a small amount of error.
+Then, the slope gradually increases or level up, indicating that a higher number of benchmark values generates a higher error.
+Using the reverse elbow method, when the benchmark value is set to 0.2, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
+This point indicates that higher benchmark values remove the necessary bins as well which lead to distruct the 2D structure."
+#| label: fig-mseScurvestartpoint
+#| fig-pos: H
+
+mse_plot_start_point
 ```
@@
 #| label: fig-scurveshifthexgridsexp
-##| out-width: 100%
 #| fig-pos: H
@@
 #| label: fig-scurveshifthexgrids
-##| out-width: 100%
 #| fig-pos: H
@@
 #| out-width: 100%
+#| out-height: 100%
 

### 2024-03-26 paper.tex
Added bechmark value to remove long edge section

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, interior hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}, breakable, boxrule=0pt, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable, frame hidden, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 
-\hypertarget{benchmark-value-to-remove-long-edges}{%
-\subsubsection{Benchmark value to remove long
-edges}\label{benchmark-value-to-remove-long-edges}}
-
-Achieving a smooth representation in 2D space is crucial, and one factor
-influencing this smoothness is the presence of long edges (see
-\textbf{?@fig-modelScurvermlgimp}). These long edges occur when a line
-connects points that are distant in the triangular mesh, impacting the
-overall smoothness of the mesh.
-
-To investigate the effect of removing long edges on the MSE, we analyzed
-the MSE for various benchmark values. Surprisingly, the results
-indicated that the MSE remained consistent across different benchmark
-values, as shown in \textbf{?@fig-diagnosticpltScurvelgrm}. It's
-important to note that edges are defined only in the 2D model and do not
-extend to higher dimensions. Consequently, removing edges does not
-affect the model in higher dimensions.
-
-There is no definitive rule for determining the benchmark value to
-remove long edges. However, we have a method to find a default value
-(see Appendix). Adjusting values around this default can help to find
-benchmark value to remove long edges, contributing to the construction
-of a smoother 2D representation (see
-\textbf{?@fig-diagnosticpltScurvelgrm}).
-
-The following steps will help to find a suitable value to remove long
-edges:
+\hypertarget{benchmark-value-for-removing-long-edges}{%
+\subsubsection{Benchmark value for removing long
+edges}\label{benchmark-value-for-removing-long-edges}}
+
+Creating a smooth 2D representation (see Figure~\ref{fig-modelScurve}
+(c)) requires removing edges that connect distant bin centroids within
+the triangular mesh (see Figure~\ref{fig-modelScurve} (b)). These edges
+only exist in the 2D model and do not extend into higher dimensions,
+ensuring that their removal does not impact the model in higher
+dimensions. While specific criteria for determining the benchmark value
+to remove long edges do not exist, the following steps provide an
+approach to identifying a suitable threshold:
 
@@
 \item
-  Plot the distribution of the 2D distances
+  Plot the distribution of the 2D Euclidean distances.
+\item
+  Identify the largest difference between consecutive distance values.
+\item
+  Take the largest distance value corresponding to this difference as
+  the benchmark value.
 \end{enumerate}
@@
 
-\caption{\label{fig-distScurve}Distribution of 2D Euclidean distances.}
-
-\end{figure}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{1}
-\tightlist
-\item
-  Find a value which is greater than smallest value
-\end{enumerate}
+\caption{\label{fig-distScurve}Distribution of 2D Euclidean distances
+between bin centroids of the triangular mesh generated with S-curve UMAP
+data.}
+
+\end{figure}
 
@@
 
-\includegraphics{paper_files/figure-pdf/unnamed-chunk-36-1.pdf}
-
-\begin{figure}
+\begin{figure}[H]
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-mseScurvestartpoint-1.pdf}
+
+}
+
+\caption{\label{fig-mseScurvestartpoint}Goodness of fit statistics from
+UMAP applied to training S-curve dataset with different benchmark values
+to remove the low-density heaxgons. What is the effective benchmark
+value to remove the low-density heaxgons? The MSE plot have a steep
+slope at the beginning, indicating that a smaller benchmark causes a
+small amount of error. Then, the slope gradually increases or level up,
+indicating that a higher number of benchmark values generates a higher
+error. Using the reverse elbow method, when the benchmark value is set
+to 0.2, the slope of the Mean Squared Error (MSE) plot experiences a
+sudden and noticeable change, resembling an elbow-like shape. This point
+indicates that higher benchmark values remove the necessary bins as well
+which lead to distruct the 2D structure.}
+
+\end{figure}
+
+\begin{figure}[H]
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-abserrorPBMC-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=1\textheight]{paper_files/figure-pdf/fig-abserrorPBMC-1.pdf}
 

### 2024-03-26 paper.qmd
Added benchmark value to remove low-density section

@@
 The performance and robustness of our model depend on four key parameters: (i) the total number of bins ($b$), (ii) a benchmark value used to remove low-density hexagons, (iii) a benchmark value used to remove long edges, and (iv) starting point of the hexagonal grid.
-However, there is no analytical formula to calculate an appropriate value for these parameters (the computation of default values can be find in *quollr* paper).
+However, there is no analytical formula to calculate an appropriate value for these parameters.
 The selection of these parameter values depends on the model performance computed by Mean Squared Error (MSE) (see @sec-goodfit).
@@
   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
-  geom_text(data = rm_centroid_df, color = "red", aes(x = c_x, y = c_y, group = hexID, label = "X"), size = 0.5) +
+  geom_text(data = hex_full_count_df_filtered[6, ], color = "red", aes(x = c_x, y = c_y, group = hexID, label = "X", size = 0.5)) +
   #geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
@@
 
-After setting up the hexagonal grid with an appropriate number of bins, it is possible that some hexagonal bins may have very few or no data points within them.
-To ensure comprehensive coverage of the NLDR data, it is necessary to select hexagonal bins that have a considerable number of data points within their hexagons.
-To achieve this, the number of points within each hexagon is first calculated.
+After establishing the hexagonal grid with an appropriate number of bins, some hexagonal bins may have few or no data points within them.
+To ensure comprehensive coverage of the NLDR data, it is necessary to select hexagonal bins with a considerable number of data points.
+This involves calculating the number of points within each hexagon.
 Then, the standard count is computed by dividing the number of points within each hexagon by the maximum number of points in the grid (see @eq-equationp2).
-Next, the bins that have a standard number of points less than a certain benchmark value are removed.
-After removing the hexagons that have insufficient data density, the hexagons with more substantial data representation are used to construct the 2D model.
+Next, bins with a standard count less than a certain benchmark value are removed.
+The following steps will help in determining a suitable value for removing low-density hexagons:
+
+1. Plot the distribution of the standardized counts (see @fig-stdctsScurve).
+2. Examine the distribution of counts.
+3. Select the first quartile value if the distribution is skewed.
 
@@
 
-Furthermore, it is crucial to choose the benchmark value to remove low-density hexagons carefully.
-If we remove unnecessary bins, then it may result in long edges and an uneven 2D model.
-Therefore, instead of removing hexagons only identified by the benchmark value, we examine the standard number of points in the neighboring hexagons of the identified low-density hexagons.
-If the neighboring bins also have low counts, then only those bins will be removed.
-The other identified bins will be kept and used for constructing the 2D model along with the high-density bins.
-
-The benchmark value for removing low-density hexagons has a range of $0$ and $1$.
-When analyzing how these benchmark values affect model performance, it's important to observe the change in Mean Squared Error (MSE) as the benchmark value increases.
-The MSE exhibits a gradual increase as the benchmark value progresses from $0$ to $1$.
-Assessing this rate of increase is crucial.
-If the increment is not substantial, the decision might lean towards retaining low-density hexagons.
-
-The following steps will help to find a suitable value to remove low density hexagons:
-
-1. Plot the distribution of the standardized counts
-
-```{r}
-#| echo: false
-#| fig-cap: "Distribution of 2D Euclidean distances."
+Furthermore, selecting the benchmark value for removing low-density hexagons is important.
+Removing unnecessary bins may lead to the formation of long edges and an uneven 2D model.
+Hence, rather than solely relying on the benchmark value to identify hexagons for removal, it's essential to consider the standard number of points in the neighboring hexagons of the identified low-density ones.
+If neighboring bins also show low counts, only those bins will be removed.
+The remaining bins are used to construct the 2D model.
+
+The benchmark value for removing low-density hexagons ranges between $0$ and $1$.
+When analyzing how these benchmark values influence model performance, it's essential to observe the change in Mean Squared Error (MSE) as the benchmark value increases (see @fig-diagnosticpltScurvelwd).
+The MSE shows a gradual increase as the benchmark value progresses from $0$ to $1$.
+Evaluating this rate of increase is important.
+If the increment is not considerable, the decision might lean towards retaining low-density hexagons.
+
+```{r}
+#| echo: false
+#| fig-cap: "Distribution of standardize counts by hexagons."
 #| label: fig-stdctsScurve
@@
 ```
-
-2. See the distribution of counts
-
-3. Take the first quartile
 
@@
 #| warning: false
-#| fig-cap: Goodness of f
+#| fig-cap: "(a) Hexbin plot with colored hexbins based on point density, (b) Identification of low-density hexagons using a benchmark value of $0.2$, and (c) Identification of low-density hexagons considering neighboring bins."
 #| label: fig-rmlowdenshex
@@
 
-1. Plot the distribution of the 2D Euclidean distances.
+1. Plot the distribution of the 2D Euclidean distances (see @fig-distScurve).
 2. Identify the largest difference between consecutive distance values.

### 2024-03-26 paper.tex
Added benchmark value to remove low-density section

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable, frame hidden, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, sharp corners, interior hidden, frame hidden, breakable, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
 
@@
 However, there is no analytical formula to calculate an appropriate
-value for these parameters (the computation of default values can be
-find in \emph{quollr} paper). The selection of these parameter values
+value for these parameters. The selection of these parameter values
 depends on the model performance computed by Mean Squared Error (MSE)
@@
 
-After setting up the hexagonal grid with an appropriate number of bins,
-it is possible that some hexagonal bins may have very few or no data
-points within them. To ensure comprehensive coverage of the NLDR data,
-it is necessary to select hexagonal bins that have a considerable number
-of data points within their hexagons. To achieve this, the number of
-points within each hexagon is first calculated. Then, the standard count
-is computed by dividing the number of points within each hexagon by the
-maximum number of points in the grid (see Equation~\ref{eq-equationp2}).
-Next, the bins that have a standard number of points less than a certain
-benchmark value are removed. After removing the hexagons that have
-insufficient data density, the hexagons with more substantial data
-representation are used to construct the 2D model.
-
-\begin{equation}\protect\hypertarget{eq-equationp2}{}{
-\text{standard count} = \frac{\text{count}}{\text{max count}} 
-}\label{eq-equationp2}\end{equation}
-
-Furthermore, it is crucial to choose the benchmark value to remove
-low-density hexagons carefully. If we remove unnecessary bins, then it
-may result in long edges and an uneven 2D model. Therefore, instead of
-removing hexagons only identified by the benchmark value, we examine the
-standard number of points in the neighboring hexagons of the identified
-low-density hexagons. If the neighboring bins also have low counts, then
-only those bins will be removed. The other identified bins will be kept
-and used for constructing the 2D model along with the high-density bins.
-
-The benchmark value for removing low-density hexagons has a range of
-\(0\) and \(1\). When analyzing how these benchmark values affect model
-performance, it's important to observe the change in Mean Squared Error
-(MSE) as the benchmark value increases. The MSE exhibits a gradual
-increase as the benchmark value progresses from \(0\) to \(1\).
-Assessing this rate of increase is crucial. If the increment is not
-substantial, the decision might lean towards retaining low-density
-hexagons.
-
-The following steps will help to find a suitable value to remove low
-density hexagons:
+After establishing the hexagonal grid with an appropriate number of
+bins, some hexagonal bins may have few or no data points within them. To
+ensure comprehensive coverage of the NLDR data, it is necessary to
+select hexagonal bins with a considerable number of data points. This
+involves calculating the number of points within each hexagon. Then, the
+standard count is computed by dividing the number of points within each
+hexagon by the maximum number of points in the grid (see
+Equation~\ref{eq-equationp2}). Next, bins with a standard count less
+than a certain benchmark value are removed. The following steps will
+help in determining a suitable value for removing low-density hexagons:
 
@@
 \item
-  Plot the distribution of the standardized counts
+  Plot the distribution of the standardized counts (see
+  Figure~\ref{fig-stdctsScurve}).
+\item
+  Examine the distribution of counts.
+\item
+  Select the first quartile value if the distribution is skewed.
 \end{enumerate}
 
+\begin{equation}\protect\hypertarget{eq-equationp2}{}{
+\text{standard count} = \frac{\text{count}}{\text{max count}} 
+}\label{eq-equationp2}\end{equation}
+
+Furthermore, selecting the benchmark value for removing low-density
+hexagons is important. Removing unnecessary bins may lead to the
+formation of long edges and an uneven 2D model. Hence, rather than
+solely relying on the benchmark value to identify hexagons for removal,
+it's essential to consider the standard number of points in the
+neighboring hexagons of the identified low-density ones. If neighboring
+bins also show low counts, only those bins will be removed. The
+remaining bins are used to construct the 2D model.
+
+The benchmark value for removing low-density hexagons ranges between
+\(0\) and \(1\). When analyzing how these benchmark values influence
+model performance, it's essential to observe the change in Mean Squared
+Error (MSE) as the benchmark value increases (see
+Figure~\ref{fig-diagnosticpltScurvelwd}). The MSE shows a gradual
+increase as the benchmark value progresses from \(0\) to \(1\).
+Evaluating this rate of increase is important. If the increment is not
+considerable, the decision might lean towards retaining low-density
+hexagons.
+
 \begin{figure}[H]
@@
 
-\caption{\label{fig-stdctsScurve}Distribution of 2D Euclidean
-distances.}
-
-\end{figure}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{1}
-\item
-  See the distribution of counts
-\item
-  Take the first quartile
-\end{enumerate}
+\caption{\label{fig-stdctsScurve}Distribution of standardize counts by
+hexagons.}
+
+\end{figure}
 
@@
 
-\caption{\label{fig-rmlowdenshex}Goodness of f}
+\caption{\label{fig-rmlowdenshex}(a) Hexbin plot with colored hexbins
+based on point density, (b) Identification of low-density hexagons using
+a benchmark value of \(0.2\), and (c) Identification of low-density
+hexagons considering neighboring bins.}
 
@@
 \item
-  Plot the distribution of the 2D Euclidean distances.
+  Plot the distribution of the 2D Euclidean distances (see
+  Figure~\ref{fig-distScurve}).
 \item

### 2024-03-26 paper.qmd
Added total number of bins section

@@
 
-The number of hexagonal bins in the hexagonal grid has a considerable impact on the construction of the 2D model.
-This is because it is the initial step in building the 2D model.
-The hexagonal grid with the chosen total number of bins must be able to capture the structure of the NLDR data.
+The number of hexagonal bins in the hexagonal grid has a considerable impact on the construction of the 2D model, serving as the initial step in building the 2D model.
+The chosen total number of bins must effectively capture the structure of the NLDR data.
 If the number of bins is too low, the model may not be able to capture the structure of the NLDR data effectively (see @fig-binsize (a)), while if there are too many bins, it may result in over-fitting the individual points of the NLDR data (see @fig-binsize (c)).
-Therefore, it is important to determine an appropriate number of bins to build an effective model.
+Therefore, it is important to determine an appropriate number of bins to construct an effective model.
 
@@
 
-Furthermore, the total number of bins is determined by the number of bins along the x-axis and y-axis (see @eq-equation8).
-To calculate the effective total number of bins, candidate values are selected based on the range between the minimum and approximate maximum number of bins along the x and y axes.
-The minimum number of bins along each axis is set to $1$, while the maximum number is estimated by taking the square root of the NLDR data point count.
-The analysis evaluates the MSE across varying total bin counts within this range, covering the minimum to maximum values along both axes.
-
-The analysis focuses on exploring the effect of total number of bins on MSE.
-It considers MSE across different total number of bins, covering the range calculated from the minimum to maximum values along both x and y axes.
-
-Typically, the MSE plot tends to have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
-Then, the slope gradually declines or levesl off, indicating that a higher number of bins generates a smaller error.
-This MSE plot is useful in determining the effective number of bins required to construct the 2D model.
-@fig-diagnosticpltScurve has created a graph that shows how the Mean Squared Error (MSE) changes depending on the number of bins assigned in the hexagonal grid for UMAP applied to the S-curve dataset.
-As per the rule of thumb mentioned earlier, the effective number of bins is determined to be $84$, with $6$ bins along the x-axis and $14$ bins along the y-axis.
-Furthermore, @fig-modelScurve shows the 2D model constructed using UMAP with the selected total number of bins and how the model overlay on high-D data.
+Furthermore, the total number of bins is calculated based on the number of bins along the x-axis and y-axis, as shown in @eq-equation8.
+To determine the effective total number of bins, candidate values are selected based on the range between the minimum and approximate maximum number of bins along the x and y axes.
+The minimum number of bins along each axis is set to $1$, while the maximum number is estimated by taking the square root of the NLDR data points.
+The analysis evaluates the MSE across varying total number of bins within this range, covering the minimum to maximum values along both axes (see @fig-diagnosticpltScurve).
 
@@
 #| echo: false
-
-## To draw with AIC
-# aic_plot <- ggplot(mse_df, aes(x = num_bins, y = aic
-# )) +
-#   geom_point() +
-#   geom_line() +
-#   geom_vline(xintercept = 84, linetype="solid",
-#                 color = "red", size=0.8, alpha = 0.5) +
-#   theme_light() +
-#   theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-#         axis.title = element_text(size = 7),
-#         axis.text = element_text(size = 7)) +
-#   #scale_colour_manual(values = c("#377eb8")) +
-#   ylab("AIC") +
-#   xlab("total number of bins")f
-## Effective number of bins along x-axis
 
@@
 
-After establishing the hexagonal grid with an appropriate number of bins, some hexagonal bins may have few or no data points within them.
+After establishing the hexagonal grid with an appropriate number of bins, some hexagonal bins may have few or no data points within them (see @fig-rmlowdenshex (a)).
 To ensure comprehensive coverage of the NLDR data, it is necessary to select hexagonal bins with a considerable number of data points.
@@
 Then, the standard count is computed by dividing the number of points within each hexagon by the maximum number of points in the grid (see @eq-equationp2).
-Next, bins with a standard count less than a certain benchmark value are removed.
+Next, bins with a standard count less than a certain benchmark value are removed (see @fig-rmlowdenshex (b)).
 The following steps will help in determining a suitable value for removing low-density hexagons:
@@
 Removing unnecessary bins may lead to the formation of long edges and an uneven 2D model.
-Hence, rather than solely relying on the benchmark value to identify hexagons for removal, it's essential to consider the standard number of points in the neighboring hexagons of the identified low-density ones.
+Hence, rather than solely relying on the benchmark value to identify hexagons for removal, it's essential to consider the standard number of points in the neighboring hexagons of the identified low-density ones (see @fig-rmlowdenshex (c)).
 If neighboring bins also show low counts, only those bins will be removed.

### 2024-03-26 paper.tex
Added total number of bins section

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, sharp corners, interior hidden, frame hidden, breakable, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, sharp corners, breakable, frame hidden, borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced]}{\end{tcolorbox}}\fi
 
@@
 The number of hexagonal bins in the hexagonal grid has a considerable
-impact on the construction of the 2D model. This is because it is the
-initial step in building the 2D model. The hexagonal grid with the
-chosen total number of bins must be able to capture the structure of the
-NLDR data. If the number of bins is too low, the model may not be able
-to capture the structure of the NLDR data effectively (see
-Figure~\ref{fig-binsize} (a)), while if there are too many bins, it may
-result in over-fitting the individual points of the NLDR data (see
-Figure~\ref{fig-binsize} (c)). Therefore, it is important to determine
-an appropriate number of bins to build an effective model.
+impact on the construction of the 2D model, serving as the initial step
+in building the 2D model. The chosen total number of bins must
+effectively capture the structure of the NLDR data. If the number of
+bins is too low, the model may not be able to capture the structure of
+the NLDR data effectively (see Figure~\ref{fig-binsize} (a)), while if
+there are too many bins, it may result in over-fitting the individual
+points of the NLDR data (see Figure~\ref{fig-binsize} (c)). Therefore,
+it is important to determine an appropriate number of bins to construct
+an effective model.
 
@@
 
-Furthermore, the total number of bins is determined by the number of
-bins along the x-axis and y-axis (see Equation~\ref{eq-equation8}). To
-calculate the effective total number of bins, candidate values are
-selected based on the range between the minimum and approximate maximum
-number of bins along the x and y axes. The minimum number of bins along
-each axis is set to \(1\), while the maximum number is estimated by
-taking the square root of the NLDR data point count. The analysis
-evaluates the MSE across varying total bin counts within this range,
-covering the minimum to maximum values along both axes.
-
-The analysis focuses on exploring the effect of total number of bins on
-MSE. It considers MSE across different total number of bins, covering
-the range calculated from the minimum to maximum values along both x and
-y axes.
-
-Typically, the MSE plot tends to have a steep slope at the beginning,
-indicating that a smaller number of bins causes a larger amount of
-error. Then, the slope gradually declines or levesl off, indicating that
-a higher number of bins generates a smaller error. This MSE plot is
-useful in determining the effective number of bins required to construct
-the 2D model. Figure~\ref{fig-diagnosticpltScurve} has created a graph
-that shows how the Mean Squared Error (MSE) changes depending on the
-number of bins assigned in the hexagonal grid for UMAP applied to the
-S-curve dataset. As per the rule of thumb mentioned earlier, the
-effective number of bins is determined to be \(84\), with \(6\) bins
-along the x-axis and \(14\) bins along the y-axis. Furthermore,
-Figure~\ref{fig-modelScurve} shows the 2D model constructed using UMAP
-with the selected total number of bins and how the model overlay on
-high-D data.
+Furthermore, the total number of bins is calculated based on the number
+of bins along the x-axis and y-axis, as shown in
+Equation~\ref{eq-equation8}. To determine the effective total number of
+bins, candidate values are selected based on the range between the
+minimum and approximate maximum number of bins along the x and y axes.
+The minimum number of bins along each axis is set to \(1\), while the
+maximum number is estimated by taking the square root of the NLDR data
+points. The analysis evaluates the MSE across varying total number of
+bins within this range, covering the minimum to maximum values along
+both axes (see Figure~\ref{fig-diagnosticpltScurve}).
 
@@
 After establishing the hexagonal grid with an appropriate number of
-bins, some hexagonal bins may have few or no data points within them. To
-ensure comprehensive coverage of the NLDR data, it is necessary to
-select hexagonal bins with a considerable number of data points. This
-involves calculating the number of points within each hexagon. Then, the
-standard count is computed by dividing the number of points within each
-hexagon by the maximum number of points in the grid (see
-Equation~\ref{eq-equationp2}). Next, bins with a standard count less
-than a certain benchmark value are removed. The following steps will
-help in determining a suitable value for removing low-density hexagons:
+bins, some hexagonal bins may have few or no data points within them
+(see Figure~\ref{fig-rmlowdenshex} (a)). To ensure comprehensive
+coverage of the NLDR data, it is necessary to select hexagonal bins with
+a considerable number of data points. This involves calculating the
+number of points within each hexagon. Then, the standard count is
+computed by dividing the number of points within each hexagon by the
+maximum number of points in the grid (see Equation~\ref{eq-equationp2}).
+Next, bins with a standard count less than a certain benchmark value are
+removed (see Figure~\ref{fig-rmlowdenshex} (b)). The following steps
+will help in determining a suitable value for removing low-density
+hexagons:
 
@@
 it's essential to consider the standard number of points in the
-neighboring hexagons of the identified low-density ones. If neighboring
-bins also show low counts, only those bins will be removed. The
-remaining bins are used to construct the 2D model.
+neighboring hexagons of the identified low-density ones (see
+Figure~\ref{fig-rmlowdenshex} (c)). If neighboring bins also show low
+counts, only those bins will be removed. The remaining bins are used to
+construct the 2D model.
 

### 2024-03-26 paper.qmd
Added sumatised abstract taking by earlier ones

@@
 abstract: |
-  Non-Linear Dimension Reduction (NLDR) techniques have emerged as powerful tools to visualize high-dimensional data.
-However, their complexity and parameter choices may lead to distrustful or misleading results.
-To address this challenge, we propose a novel approach that combines the tour technique with a low-dimensional manifold generated using NLDR techniques, hexagonal binning, and triangulation.
-This integration enables a clear examination of the low-dimensional representation in the original high-dimensional space.
-Our approach not only preserves the advantages of both tours and NLDR but also offers a more intuitive perception of complex data structures and facilitates accurate data transformation assessments.
-The method and example data sets are available in the **quollr** R package.
+  Nonlinear dimension reduction (NLDR) techniques such as tSNE, and UMAP provide a low-dimensional representation of high-dimensional (high-D) data using non-linear transformation.
+The methods and parameter choices can create wildly different representations, making it difficult to decide which is best, or whether any or all are accurate or misleading.
+NLDR often exaggerates random patterns, sometimes due to the samples observed.
+But NLDR views have an important role in data analysis because, if done well, they provide a concise visual (and conceptual) summary of high-D distributions.
+To help evaluate the NLDR we have developed an algorithm to show the 2D NLDR model in the high-D space, viewed with a tour.
+One can see if the model fits everywhere or better in some subspaces, or completely mismatches the data.
+It is used to evaluate which 2D layout is the best representation of the high-D distribution and see how different methods may have similar summaries or quirks.
 

### 2024-03-26 paper.tex
Added sumatised abstract taking by earlier ones

@@
 \begin{abstract}
-Non-Linear Dimension Reduction (NLDR) techniques have emerged as
-powerful tools to visualize high-dimensional data. However, their
-complexity and parameter choices may lead to distrustful or misleading
-results. To address this challenge, we propose a novel approach that
-combines the tour technique with a low-dimensional manifold generated
-using NLDR techniques, hexagonal binning, and triangulation. This
-integration enables a clear examination of the low-dimensional
-representation in the original high-dimensional space. Our approach not
-only preserves the advantages of both tours and NLDR but also offers a
-more intuitive perception of complex data structures and facilitates
-accurate data transformation assessments. The method and example data
-sets are available in the \textbf{quollr} R package.
+Nonlinear dimension reduction (NLDR) techniques such as tSNE, and UMAP
+provide a low-dimensional representation of high-dimensional (high-D)
+data using non-linear transformation. The methods and parameter choices
+can create wildly different representations, making it difficult to
+decide which is best, or whether any or all are accurate or misleading.
+NLDR often exaggerates random patterns, sometimes due to the samples
+observed. But NLDR views have an important role in data analysis
+because, if done well, they provide a concise visual (and conceptual)
+summary of high-D distributions. To help evaluate the NLDR we have
+developed an algorithm to show the 2D NLDR model in the high-D space,
+viewed with a tour. One can see if the model fits everywhere or better
+in some subspaces, or completely mismatches the data. It is used to
+evaluate which 2D layout is the best representation of the high-D
+distribution and see how different methods may have similar summaries or
+quirks.
 \end{abstract}
@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, sharp corners, breakable, frame hidden, borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, interior hidden, sharp corners, frame hidden, breakable]}{\end{tcolorbox}}\fi
 

### 2024-03-26 paper.qmd
Added paper outline to introduction

@@
 The outline of this paper is as follows.
-The @sec-background provides an detailed overview of dimension reduction methods, and tours.
-Building upon this foundation, the @sec-methods delves into the proposed algorithm, its implementation details, how to tune the model, model summaries, and a synthetic example to illustrate the functionality of the algorithm.
-Subsequently, @sec-applications showcases applications of the algorithm on different data sets, particularly in single-cell RNA-seq data.
-These applications reveal insights into the performance and trustworthiness of NLDR algorithms.
-We analyze the results to identify situations where NLDR techniques may lead to misleading interpretations.
-Finally, @sec-conclusions concludes by summarizing the findings and emphasizing the significance of the proposed approach in tackling the challenges of high-dimensional data visualization.
+@sec-background provides an detailed overview of dimension reduction methods, and tour technique.
+Building upon this foundation, Section @sec-methods describes the proposed algorithm, including its implementation details, model tuning, summaries, and a synthetic example to demonstrate its functionality.
+In Section @sec-applications, the algorithm's applications on various datasets, particularly in single-cell RNA-seq and handwritten digit data, are showcased.
+These applications demonstrate how some NLDR techniques effectively capture both local and global structures of the original data, while others may result in misleading interpretations.
+Finally, Section @sec-conclusions summarizes the findings and describe the importance of the proposed approach in addressing the challenges associated with selecting the best NLDR method and parameter choices to achieve a more accurate 2D representation of high-D data.
 
@@
 
-<!--High-dimensional (high-D) data is widespread in many fields including ecology and bioinformatics [@Guo2023], in part because of new data collection technologies [@Johnstone2009; @ayesha2020overview].
-Working with high-dimensional data poses considerable challenges due to the difficulty in visualizing beyond two dimensions [@Jia2022].
-High-dimensional data also presents difficulties for model fitting [@Johnstone2009], both computationally and interpretation, each of which benefits from visualization.
-
-To create visual representations of high-dimensional data, it is common to apply dimension reduction techniques.
-
-However, projecting high-dimensional data has limitations, such as information loss and potential distortion of essential structures and patterns [@Jia2022, @article53].
-The choice of technique and parameters further impacts the accuracy of the visualization, necessitating careful consideration for meaningful interpretation (see @fig-nldervis).
-
-Interactive and dynamic graphics systems have also been developed over the years to enable visualizing high dimensions.
-One method, called a tour [@Asimov1985], shows a sequence of linear projections is shown as a movie, allowing exploration without warping the space [@lee2021review].
-Interactive tools like **XGobi** and **GGobi** have been successful in incorporating tours for exploring high-dimensional data [@article60].
-The R package **tourr** [@article61] further enhances tour visualization within R, although it may face limitations in frame rate and interactive features compared to **GGobi**.
-
-To overcome these limitations, the R package **detourr** [@article22] has been developed, leveraging a Javascript widget via htmlwidgets [@Ramnath2023] to achieve higher frame rates and enhanced interactivity.
-Additionally, the R package **langevitour** [@article09] utilizes Langevin Dynamics to generate a continuous path of projections, eliminating the need for interpolation between projections for animation.
-The tour technique has proven valuable in exploring statistical model fits [@article58] and factorial experimental designs [@article59].
-Augmenting the results of non-linear dimensional reduction methods with the tour, as demonstrated in the **liminal** R package [@article21], further enhances data exploration.
-
-The earlier approaches for DR involve linear methods such as PCA [@Karl1901].
-PCA aims to maintain the second-order statistics of the data by projecting the points into the low dimensional space that preserves the maximum amount of variance among all such projections.
-As a result, PCA has been shown to be effective in preserving the global structure of the data (Silva & Tenenbaum, 2003).
-The global structure includes the overall shape of the dataset, placement of the clusters, and the existence of potential outliers.
-Unlike PCA, much of the focus of the more recent non-linear methods, including t-SNE (Maaten & Hinton,2008), LargeVis (Tang et al., 2016), and UMAP (McInnes et al., 2018) has been on preserving the local neighborhood structure of each individual point.
-Mainly,
-
-While tours [@Asimov1985] preserve space without warping [@lee2021review], they require integrating multiple low-dimensional views mentally to perceive high-dimensional structures.
-To address this challenge, we propose a novel approach by combining the tour technique with a low-dimensional manifold.
-This manifold is created through the synergistic use of Non-Linear Dimension Reduction (NLDR) techniques, hexagonal binning, and triangulation.
-By merging these techniques, our approach offers a comprehensive and efficient means to visualize and explore high-dimensional data while retaining the advantages of both tours and NLDR.
-This integration facilitates a more intuitive perception of complex data structures and empowers analysts with a robust tool for assessing the accuracy of data transformations.
-The implementation of our approach is available as an R package called **quollr**.
-
-The outline of this paper is as follows.
-The @sec-background provides an detailed overview of dimension reduction methods, triangulation, and tours.
-Building upon this foundation, the @sec-methods delves into the proposed algorithm, **quollr**, and its implementation details.
-In @sec-prediction, discusses the effectiveness of the learned low-dimensional manifold in accurately representing the complex high-dimensional data.
-Following that, @sec-simpleex presents simple examples from simulations to illustrate the functionality of the algorithm.
-Subsequently, @sec-applications showcases real-world applications of **quollr** on different data sets, particularly in single-cell RNA-seq data.
-These applications reveal insights into the performance and trustworthiness of NLDR algorithms.
-We analyze the results to identify situations where NLDR techniques may lead to misleading interpretations.
-Finally, @sec-conclusions concludes by summarizing the findings and emphasizing the significance of the proposed approach in tackling the challenges of high-dimensional data visualization.
--->
-
 ## Background {#sec-background}
@@
 
-<!--Their focus is on capturing complex structures that may not be easily represented in the original space, making a straightforward reverse mapping challenging.
-
-One specific technique we focus on is Pairwise Controlled Manifold Approximation (PaCMAP).
-Similar considerations apply to related methods like tSNE [@Laurens2008], UMAP [@Leland2018], and TrMAP [@article02].
-
-However, linear methods may not fully capture complex non-linear relationships present in the data.-->
-
 #### Non-linear dimension reduction techniques
@@
 In our context, use grand tour along with the model to observe how effectively the model captures the underlying structure of the data.
-
-<!--While both tours can be used to visualize data, examples often focus on using the grand tour to observe global structures.
-However, software like **langevitour** can visualize both types of tours, providing flexibility for exploring high-dimensional data with various objectives.
--->
 
@@
 
-## Conclusion {#sec-conclusion}
+## Conclusion {#sec-conclusions}
 

### 2024-03-26 paper.tex
Added paper outline to introduction

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, interior hidden, sharp corners, frame hidden, breakable]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, breakable, frame hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, interior hidden, boxrule=0pt]}{\end{tcolorbox}}\fi
 
@@
 
-The outline of this paper is as follows. The
-Section~\ref{sec-background} provides an detailed overview of dimension
-reduction methods, and tours. Building upon this foundation, the
-Section~\ref{sec-methods} delves into the proposed algorithm, its
-implementation details, how to tune the model, model summaries, and a
-synthetic example to illustrate the functionality of the algorithm.
-Subsequently, Section~\ref{sec-applications} showcases applications of
-the algorithm on different data sets, particularly in single-cell
-RNA-seq data. These applications reveal insights into the performance
-and trustworthiness of NLDR algorithms. We analyze the results to
-identify situations where NLDR techniques may lead to misleading
-interpretations. Finally, \textbf{?@sec-conclusions} concludes by
-summarizing the findings and emphasizing the significance of the
-proposed approach in tackling the challenges of high-dimensional data
-visualization.
+The outline of this paper is as follows. Section~\ref{sec-background}
+provides an detailed overview of dimension reduction methods, and tour
+technique. Building upon this foundation, Section
+Section~\ref{sec-methods} describes the proposed algorithm, including
+its implementation details, model tuning, summaries, and a synthetic
+example to demonstrate its functionality. In Section
+Section~\ref{sec-applications}, the algorithm's applications on various
+datasets, particularly in single-cell RNA-seq and handwritten digit
+data, are showcased. These applications demonstrate how some NLDR
+techniques effectively capture both local and global structures of the
+original data, while others may result in misleading interpretations.
+Finally, Section Section~\ref{sec-conclusions} summarizes the findings
+and describe the importance of the proposed approach in addressing the
+challenges associated with selecting the best NLDR method and parameter
+choices to achieve a more accurate 2D representation of high-D data.
 
@@
 
-\hypertarget{sec-conclusion}{%
-\section{Conclusion}\label{sec-conclusion}}
+\hypertarget{sec-conclusions}{%
+\section{Conclusion}\label{sec-conclusions}}
 

### 2024-03-26 paper.qmd
Added detailed description for simulated data example

@@
 
-In this section, we showcase the effectiveness of our methodology using simulated data.
-The dataset comprises five spherical Gaussian clusters in 4-$d$, with each cluster containing an equal number of points and consistent within variation.
-
-In the 2D representations created by all NLDR techniques, as shown in @fig-nldervis5Gau, except for PHATE, there are five distinct clusters.
-In tSNE, these five clusters are closely located to each other (see @fig-nldervis5Gau (a)).
-However, in PHATE, two clusters are closely positioned, while the other three are more distant (see @fig-nldervis5Gau (c)).
-In PaCMAP, one cluster is at the center, and the remaining four are positioned in four different directions (see @fig-nldervis5Gau (e)).
-In TriMAP, two clusters are close, although not as close as in PHATE, and the other three are well-separated (see @fig-nldervis5Gau (d)).
-In UMAP, all clusters are arranged in a parallel manner, with three in one line and the other two in a separate line (see @fig-nldervis5Gau (b)).
-
-Visualizing the models alongside the original high-D data provides insights into how different techniques capture the underlying clustering structure.
-The tSNE model exhibits five well-separated clusters, effectively preserving both local and global structures (see video link of @fig-modelfiveGau (a)).
-UMAP, on the other hand, also presents five distinct clusters but with a more flattened surface appearance (see video link of @fig-modelfiveGau (b)).
-The PHATE model shows five separated clusters resembling triangles or partial triangles but struggles to capture local structures (see video link of @fig-modelfiveGau (c)).
-TriMAP and PaCMAP both display five well-separated clusters with flat surfaces, each capturing within-cluster variation to varying extents.
-Despite the various 2D representations, all NLDR techniques preserve the global structure (see video link of @fig-modelfiveGau (d), (e) respectively).
-However, tSNE, effectively capture both local and global structures.
+In this section, the effectiveness of the algorithm is described using a simulated dataset.
+The dataset consists of five spherical Gaussian clusters in 4-$d$, with each cluster containing an equal number of points and the same within-cluster variation.
+
+In the 2D layouts generated by various NLDR techniques, as shown in @fig-nldervis5Gau, five distinct clusters are observable, except for PHATE.
+In tSNE (see @fig-nldervis5Gau (a)), these clusters appear closely.
+UMAP arranges all clusters in a parallel manner, with three aligned in one line and the other two in a separate line (see @fig-nldervis5Gau (b)).
+In contrast, PHATE shows two closely positioned clusters and three more distant ones (see @fig-nldervis5Gau (c)).
+PaCMAP shows one central cluster and the remaining four spread out in different directions (see @fig-nldervis5Gau (e)).
+Finally, in TriMAP, two clusters are close, though not as tightly as PHATE, while the other three are well-separated (see @fig-nldervis5Gau (d)).
 
@@
 #| label: fig-nldervis5Gau
+#| fig-pos: H
 #| out-width: 100%
@@
 ```
+
+To investigate which is the best representation to visualize the original data or all NLDR methods provide equivalent information, we visualize the model-in-data space.
+Models from all NLDR methods show five well-separated clusters (see @fig-gau1_sc, @fig-gau2_sc, @fig-gau3_sc, @fig-gau4_sc, and @fig-gau5_sc).
+This suggests that for the five Gaussian cluster dataset, all NLDR methods effectively preserve the global structure.
+tSNE, UMAP, and PHATE display clusters with varying densities, indicating their ability to capture within-cluster variation (see @fig-gau1_sc, @fig-gau2_sc, and @fig-gau3_sc).
+On the other hand, both TriMAP and PaCMAP show clusters with flat surfaces, suggesting a failure to capture within-cluster variation (see @fig-gau4_sc and @fig-gau5_sc).
+Therefore, TriMAP and PaCMAP do not capture the local structure as effectively as other methods.
 

### 2024-03-26 paper.tex
Added detailed description for simulated data example

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, breakable, frame hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, interior hidden, boxrule=0pt]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, sharp corners, enhanced, frame hidden, breakable, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
 
@@
 
-In this section, we showcase the effectiveness of our methodology using
-simulated data. The dataset comprises five spherical Gaussian clusters
-in 4-\(d\), with each cluster containing an equal number of points and
-consistent within variation.
-
-In the 2D representations created by all NLDR techniques, as shown in
-Figure~\ref{fig-nldervis5Gau}, except for PHATE, there are five distinct
-clusters. In tSNE, these five clusters are closely located to each other
-(see Figure~\ref{fig-nldervis5Gau} (a)). However, in PHATE, two clusters
-are closely positioned, while the other three are more distant (see
-Figure~\ref{fig-nldervis5Gau} (c)). In PaCMAP, one cluster is at the
-center, and the remaining four are positioned in four different
-directions (see Figure~\ref{fig-nldervis5Gau} (e)). In TriMAP, two
-clusters are close, although not as close as in PHATE, and the other
-three are well-separated (see Figure~\ref{fig-nldervis5Gau} (d)). In
-UMAP, all clusters are arranged in a parallel manner, with three in one
-line and the other two in a separate line (see
-Figure~\ref{fig-nldervis5Gau} (b)).
-
-Visualizing the models alongside the original high-D data provides
-insights into how different techniques capture the underlying clustering
-structure. The tSNE model exhibits five well-separated clusters,
-effectively preserving both local and global structures (see video link
-of Figure~\ref{fig-modelfiveGau} (a)). UMAP, on the other hand, also
-presents five distinct clusters but with a more flattened surface
-appearance (see video link of Figure~\ref{fig-modelfiveGau} (b)). The
-PHATE model shows five separated clusters resembling triangles or
-partial triangles but struggles to capture local structures (see video
-link of Figure~\ref{fig-modelfiveGau} (c)). TriMAP and PaCMAP both
-display five well-separated clusters with flat surfaces, each capturing
-within-cluster variation to varying extents. Despite the various 2D
-representations, all NLDR techniques preserve the global structure (see
-video link of Figure~\ref{fig-modelfiveGau} (d), (e) respectively).
-However, tSNE, effectively capture both local and global structures.
-
-\begin{figure}
+In this section, the effectiveness of the algorithm is described using a
+simulated dataset. The dataset consists of five spherical Gaussian
+clusters in 4-\(d\), with each cluster containing an equal number of
+points and the same within-cluster variation.
+
+In the 2D layouts generated by various NLDR techniques, as shown in
+Figure~\ref{fig-nldervis5Gau}, five distinct clusters are observable,
+except for PHATE. In tSNE (see Figure~\ref{fig-nldervis5Gau} (a)), these
+clusters appear closely. UMAP arranges all clusters in a parallel
+manner, with three aligned in one line and the other two in a separate
+line (see Figure~\ref{fig-nldervis5Gau} (b)). In contrast, PHATE shows
+two closely positioned clusters and three more distant ones (see
+Figure~\ref{fig-nldervis5Gau} (c)). PaCMAP shows one central cluster and
+the remaining four spread out in different directions (see
+Figure~\ref{fig-nldervis5Gau} (e)). Finally, in TriMAP, two clusters are
+close, though not as tightly as PHATE, while the other three are
+well-separated (see Figure~\ref{fig-nldervis5Gau} (d)).
+
+\begin{figure}[H]
 
@@
 \end{figure}
+
+To investigate which is the best representation to visualize the
+original data or all NLDR methods provide equivalent information, we
+visualize the model-in-data space. Models from all NLDR methods show
+five well-separated clusters (see Figure~\ref{fig-gau1_sc},
+Figure~\ref{fig-gau2_sc}, Figure~\ref{fig-gau3_sc},
+Figure~\ref{fig-gau4_sc}, and Figure~\ref{fig-gau5_sc}). This suggests
+that for the five Gaussian cluster dataset, all NLDR methods effectively
+preserve the global structure. tSNE, UMAP, and PHATE display clusters
+with varying densities, indicating their ability to capture
+within-cluster variation (see Figure~\ref{fig-gau1_sc},
+Figure~\ref{fig-gau2_sc}, and Figure~\ref{fig-gau3_sc}). On the other
+hand, both TriMAP and PaCMAP show clusters with flat surfaces,
+suggesting a failure to capture within-cluster variation (see
+Figure~\ref{fig-gau4_sc} and Figure~\ref{fig-gau5_sc}). Therefore,
+TriMAP and PaCMAP do not capture the local structure as effectively as
+other methods.
 

### 2024-03-26 paper.qmd
Added description for starting point of hexagonal grid

@@
 
-According to @Dan2023, the hexagonal binning is done by tessellating the $xy$ plane over the set (range($x$), range($y$)) (see @fig-scurveshifthexgridsexp (b)).
-In that case, bin centroids are defined as shown in @fig-scurveshifthexgridsexp (a) with gray colour.
-Rather than sticking to the typical hexagonal grid, introducing a meaningful shift in both the x and y directions presents an opportunity for an improved 2D model.
-Therefore, investigating this shift in the hexagonal grid is an important parameter to consider.
-
-As shown in @fig-scurveshifthexgrids, the shifting influences the distribution of points and number of non-empty bins, impacting the resulting 2D model.
-According to @fig-diagnosticpltScurvehexbins, the 2D model with a total of $144$ bins applied to S-curve UMAP data does not require any shifting because the lowest MSE occurs when no shift is introduced.
+Hexagonal binning involves tessellating the $xy$ plane over the set defined by the range of $x$ and $y$ [@Dan2023].
+Therefore, the starting point of the hexagonal grid is important parameter to consider.
+Having different starting point results in shifting the hexagonal grid and bin centroids (see @fig-scurveshifthexgridsexp).
+With a specific total number of bins (which corresponds to a specific range along the x and y axes), it's essential to explore how different starting points affect the model's performance (see @fig-mseScurvestartpoint).
+This is because different starting points result in varying distributions of non-empty bins (see @fig-scurveshifthexgrids), impacting the model.
 

### 2024-03-26 paper.tex
Added description for starting point of hexagonal grid

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, sharp corners, enhanced, frame hidden, breakable, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, sharp corners, frame hidden, breakable, enhanced, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt]}{\end{tcolorbox}}\fi
 
@@
 
-According to \citet{Dan2023}, the hexagonal binning is done by
-tessellating the \(xy\) plane over the set (range(\(x\)), range(\(y\)))
-(see Figure~\ref{fig-scurveshifthexgridsexp} (b)). In that case, bin
-centroids are defined as shown in
-Figure~\ref{fig-scurveshifthexgridsexp} (a) with gray colour. Rather
-than sticking to the typical hexagonal grid, introducing a meaningful
-shift in both the x and y directions presents an opportunity for an
-improved 2D model. Therefore, investigating this shift in the hexagonal
-grid is an important parameter to consider.
-
-As shown in Figure~\ref{fig-scurveshifthexgrids}, the shifting
-influences the distribution of points and number of non-empty bins,
-impacting the resulting 2D model. According to
-\textbf{?@fig-diagnosticpltScurvehexbins}, the 2D model with a total of
-\(144\) bins applied to S-curve UMAP data does not require any shifting
-because the lowest MSE occurs when no shift is introduced.
+Hexagonal binning involves tessellating the \(xy\) plane over the set
+defined by the range of \(x\) and \(y\) \citep{Dan2023}. Therefore, the
+starting point of the hexagonal grid is important parameter to consider.
+Having different starting point results in shifting the hexagonal grid
+and bin centroids (see Figure~\ref{fig-scurveshifthexgridsexp}). With a
+specific total number of bins (which corresponds to a specific range
+along the x and y axes), it's essential to explore how different
+starting points affect the model's performance (see
+Figure~\ref{fig-mseScurvestartpoint}). This is because different
+starting points result in varying distributions of non-empty bins (see
+Figure~\ref{fig-scurveshifthexgrids}), impacting the model.
 

### 2024-03-27 paper.tex
fixed on the _ fig labels - this breaks in current version of quarto

@@
 \makeatletter
-\makeatother
-\makeatletter
-\makeatother
-\makeatletter
 \@ifpackageloaded{caption}{}{\usepackage{caption}}
@@
 \makeatletter
+\makeatother
+\makeatletter
 \@ifpackageloaded{caption}{}{\usepackage{caption}}
 \@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
-\makeatother
-\makeatletter
-\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
-\makeatother
-\makeatletter
-\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
-\makeatother
-\makeatletter
-\makeatother
-\makeatletter
 \makeatother
@@
 \bibliographystyle{agsm}
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\usepackage{bookmark}
+
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, sharp corners, frame hidden, breakable, enhanced, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt]}{\end{tcolorbox}}\fi
-
-\hypertarget{sec-intro}{%
-\section{Introduction}\label{sec-intro}}
+\section{Introduction}\label{sec-intro}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervisUMAP-1.pdf}
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervisUMAP-1.pdf}
 
@@
 
-\end{figure}
-
-\hypertarget{sec-background}{%
-\section{Background}\label{sec-background}}
-
-\hypertarget{dimension-reduction}{%
-\subsection{Dimension Reduction}\label{dimension-reduction}}
+\end{figure}%
+
+\section{Background}\label{sec-background}
+
+\subsection{Dimension Reduction}\label{dimension-reduction}
 
@@
 
-\hypertarget{non-linear-dimension-reduction-techniques}{%
 \subsubsection{Non-linear dimension reduction
-techniques}\label{non-linear-dimension-reduction-techniques}}
+techniques}\label{non-linear-dimension-reduction-techniques}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervis-1.pdf}
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervis-1.pdf}
 
@@
 
-\end{figure}
-
-\hypertarget{linear-overviews-using-tours}{%
+\end{figure}%
+
 \subsection{Linear overviews using
-tours}\label{linear-overviews-using-tours}}
+tours}\label{linear-overviews-using-tours}
 
@@
 
-\hypertarget{sec-methods}{%
-\section{Methodology}\label{sec-methods}}
+\section{Methodology}\label{sec-methods}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=1\textheight]{figures/workflow.png}
+\centering{
+
+\includegraphics[width=1\textwidth,height=1\textheight]{figures/workflow.png}
 
@@
 
-\end{figure}
-
-\hypertarget{preprocessing-steps}{%
-\subsection{Preprocessing steps}\label{preprocessing-steps}}
+\end{figure}%
+
+\subsection{Preprocessing steps}\label{preprocessing-steps}
 
@@
 
-\hypertarget{sec-construct2d}{%
-\subsection{Constructing the 2D model}\label{sec-construct2d}}
+\subsection{Constructing the 2D model}\label{sec-construct2d}
 
@@
 
-\begin{equation}\protect\hypertarget{eq-equation1}{}{
+\begin{equation}\phantomsection\label{eq-equation1}{
  ar = \frac{r_2}{r_1}
-}\label{eq-equation1}\end{equation}
-
-\begin{equation}\protect\hypertarget{eq-equation2}{}{
+}\end{equation}
+
+\begin{equation}\phantomsection\label{eq-equation2}{
  hr = \frac{hb}{wb}
-}\label{eq-equation2}\end{equation}
-
-\begin{equation}\protect\hypertarget{eq-equation3}{}{
+}\end{equation}
+
+\begin{equation}\phantomsection\label{eq-equation3}{
  y_{max} = \left\lceil\frac{ar}{hr}\right\rceil * hr
-}\label{eq-equation3}\end{equation}
+}\end{equation}
 
@@
 
-\begin{equation}\protect\hypertarget{eq-equation11}{}{
+\begin{equation}\phantomsection\label{eq-equation11}{
  h = \sqrt{3} * s
-}\label{eq-equation11}\end{equation}
-
-\begin{equation}\protect\hypertarget{eq-equation12}{}{
+}\end{equation}
+
+\begin{equation}\phantomsection\label{eq-equation12}{
  b_1 = \frac{r_1 + buffer_{x}}{h}
-}\label{eq-equation12}\end{equation}
-
-\begin{equation}\protect\hypertarget{eq-equation13}{}{
+}\end{equation}
+
+\begin{equation}\phantomsection\label{eq-equation13}{
   v = 1.5 * s
-}\label{eq-equation13}\end{equation}
-
-\begin{equation}\protect\hypertarget{eq-equation14}{}{
+}\end{equation}
+
+\begin{equation}\phantomsection\label{eq-equation14}{
  b_2 = \frac{r_2 + buffer_{y}}{v}
-}\label{eq-equation14}\end{equation}
+}\end{equation}
 
@@
 
-\begin{equation}\protect\hypertarget{eq-equation15}{}{
+\begin{equation}\phantomsection\label{eq-equation15}{
 d_x = \frac{h}{2}
-}\label{eq-equation15}\end{equation}
-
-\begin{equation}\protect\hypertarget{eq-equation16}{}{
+}\end{equation}
+
+\begin{equation}\phantomsection\label{eq-equation16}{
 d_y = \frac{2* v * 1.15}{\sqrt{3}} 
-}\label{eq-equation16}\end{equation}
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=1\textwidth,height=1\textheight]{figures/hex_coord.png}
+}\end{equation}
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=1\textheight]{figures/hex_coord.png}
 
@@
 
-\end{figure}
+\end{figure}%
 
@@
 
-\begin{equation}\protect\hypertarget{eq-equation4}{}{
+\begin{equation}\phantomsection\label{eq-equation4}{
 C_k^{(2)} = (C_{ky_1}, C_{ky_2}) = \left(\frac{\sum_{i=1}^{6} h^{k}x_{i}}{6}, \frac{\sum_{i=1}^{6} h^{k}y_{i}}{6}\right).
-}\label{eq-equation4}\end{equation}
+}\end{equation}
 
@@
 
-\begin{equation}\protect\hypertarget{eq-equation5}{}{
+\begin{equation}\phantomsection\label{eq-equation5}{
 C_k^{(2)} = (C_{ky_1}, C_{ky_2}) = \left(\frac{1}{n_k} \sum_{i=1}^{n_k} y_{1i}, \frac{1}{n_k} \sum_{i=1}^{n_k} y_{2i}\right),
-}\label{eq-equation5}\end{equation}
+}\end{equation}
 
@@
 
-\hypertarget{lifting-the-model-into-high-dimensions}{%
 \subsection{Lifting the model into high
-dimensions}\label{lifting-the-model-into-high-dimensions}}
-
-\hypertarget{lifting-the-triangular-mesh-points-into-high-dimensions}{%
+dimensions}\label{lifting-the-model-into-high-dimensions}
+
 \subsubsection{Lifting the triangular mesh points into high
-dimensions}\label{lifting-the-triangular-mesh-points-into-high-dimensions}}
+dimensions}\label{lifting-the-triangular-mesh-points-into-high-dimensions}
 
@@
 
-\begin{equation}\protect\hypertarget{eq-equation6}{}{
+\begin{equation}\phantomsection\label{eq-equation6}{
 C_k^{(p)} = (C_{kx_1}, ..., C_{kx_p}) = \left(\frac{1}{n_k} \sum_{i=1}^{n_k} x_{1i}, \frac{1}{n_k} \sum_{i=1}^{n_k} x_{2i}, \dots ,\frac{1}{n_k} \sum_{i=1}^{n_k} x_{pi}\right).
-}\label{eq-equation6}\end{equation}
+}\end{equation}
 
@@
 
-\begin{equation}\protect\hypertarget{eq-equation7}{}{
+\begin{equation}\phantomsection\label{eq-equation7}{
 v(C_k^{(2)}) = C_k^{(p)}.
-}\label{eq-equation7}\end{equation}
+}\end{equation}
 
@@
 
-\hypertarget{lifting-the-2d-triangular-mesh-into-high-dimensions}{%
 \subsubsection{Lifting the 2D triangular mesh into high
-dimensions}\label{lifting-the-2d-triangular-mesh-into-high-dimensions}}
+dimensions}\label{lifting-the-2d-triangular-mesh-into-high-dimensions}
 
@@
 
-{\centering \includegraphics{paper_files/figure-pdf/fig-wkhighD-1.pdf}
+\centering{
+
+\includegraphics{paper_files/figure-pdf/fig-wkhighD-1.pdf}
 
@@
 
-\end{figure}
-
-\hypertarget{tuning-the-model}{%
-\subsection{Tuning the model}\label{tuning-the-model}}
+\end{figure}%
+
+\subsection{Tuning the model}\label{tuning-the-model}
 
@@
 
-\hypertarget{total-number-of-bins}{%
-\subsubsection{Total number of bins}\label{total-number-of-bins}}
+\subsubsection{Total number of bins}\label{total-number-of-bins}
 
@@
 
-{\centering \includegraphics{paper_files/figure-pdf/fig-binsize-1.pdf}
+\centering{
+
+\includegraphics{paper_files/figure-pdf/fig-binsize-1.pdf}
 
@@
 
-\end{figure}
-
-\begin{equation}\protect\hypertarget{eq-equation8}{}{
+\end{figure}%
+
+\begin{equation}\phantomsection\label{eq-equation8}{
 b = b_1 \times b_2
-}\label{eq-equation8}\end{equation}
+}\end{equation}
 
@@
 
-{\centering \includegraphics{paper_files/figure-pdf/fig-diagnosticpltScurve-1.pdf}
+\centering{
+
+\includegraphics{paper_files/figure-pdf/fig-diagnosticpltScurve-1.pdf}
 
@@
 
-\end{figure}
-
-\hypertarget{benchmark-value-to-remove-low-density-hexagons}{%
+\end{figure}%
+
 \subsubsection{Benchmark value to remove low-density
-hexagons}\label{benchmark-value-to-remove-low-density-hexagons}}
+hexagons}\label{benchmark-value-to-remove-low-density-hexagons}
 
@@
 
-\begin{equation}\protect\hypertarget{eq-equationp2}{}{
+\begin{equation}\phantomsection\label{eq-equationp2}{
 \text{standard count} = \frac{\text{count}}{\text{max count}} 
-}\label{eq-equationp2}\end{equation}
+}\end{equation}
 
@@
 
-{\centering \includegraphics{paper_files/figure-pdf/fig-stdctsScurve-1.pdf}
+\centering{
+
+\includegraphics{paper_files/figure-pdf/fig-stdctsScurve-1.pdf}
 
@@
 
-{\centering \includegraphics{paper_files/figure-pdf/fig-diagnosticpltScurvelwd-1.pdf}
+\centering{
+
+\includegraphics{paper_files/figure-pdf/fig-diagnosticpltScurvelwd-1.pdf}
 
@@
 
-\end{figure}
-
-\begin{figure}[H]
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-rmlowdenshex-1.pdf}
+\end{figure}%
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics{paper_files/figure-pdf/fig-rmlowdenshex-1.pdf}
 
@@
 
-\end{figure}
-
-\hypertarget{benchmark-value-for-removing-long-edges}{%
+\end{figure}%
+
 \subsubsection{Benchmark value for removing long
-edges}\label{benchmark-value-for-removing-long-edges}}
+edges}\label{benchmark-value-for-removing-long-edges}
 
@@
 
-{\centering \includegraphics{paper_files/figure-pdf/fig-distScurve-1.pdf}
+\centering{
+
+\includegraphics{paper_files/figure-pdf/fig-distScurve-1.pdf}
 
@@
 
-\end{figure}
-
-\begin{figure}[H]
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-modelScurve-1.pdf}
+\end{figure}%
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics{paper_files/figure-pdf/fig-modelScurve-1.pdf}
 
@@
 
-\end{figure}
-
-\hypertarget{starting-point-of-the-hexagonal-grid}{%
+\end{figure}%
+
 \subsubsection{Starting point of the hexagonal
-grid}\label{starting-point-of-the-hexagonal-grid}}
+grid}\label{starting-point-of-the-hexagonal-grid}
 
@@
 
-{\centering \includegraphics{paper_files/figure-pdf/fig-mseScurvestartpoint-1.pdf}
+\centering{
+
+\includegraphics{paper_files/figure-pdf/fig-mseScurvestartpoint-1.pdf}
 
@@
 
-\end{figure}
-
-\begin{figure}[H]
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-scurveshifthexgridsexp-1.pdf}
+\end{figure}%
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics{paper_files/figure-pdf/fig-scurveshifthexgridsexp-1.pdf}
 
@@
 
-\end{figure}
-
-\begin{figure}[H]
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-scurveshifthexgrids-1.pdf}
+\end{figure}%
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics{paper_files/figure-pdf/fig-scurveshifthexgrids-1.pdf}
 
@@
 
-\end{figure}
-
-\hypertarget{sec-summary}{%
-\subsection{Model summaries}\label{sec-summary}}
-
-\hypertarget{predicted-values-and-residuals}{%
+\end{figure}%
+
+\subsection{Model summaries}\label{sec-summary}
+
 \subsubsection{Predicted values and
-residuals}\label{predicted-values-and-residuals}}
+residuals}\label{predicted-values-and-residuals}
 
@@
 
-\begin{equation}\protect\hypertarget{eq-equation10}{}{
+\begin{equation}\phantomsection\label{eq-equation10}{
 \text{Error} = \sum_{j = 1}^{n}\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2
-}\label{eq-equation10}\end{equation}
+}\end{equation}
 
@@
 
-\hypertarget{sec-goodfit}{%
-\subsubsection{Goodness of fit statistics}\label{sec-goodfit}}
+\subsubsection{Goodness of fit statistics}\label{sec-goodfit}
 
@@
 
-\begin{equation}\protect\hypertarget{eq-equation9}{}{
+\begin{equation}\phantomsection\label{eq-equation9}{
 \text{MSE} = \sum_{j = 1}^{n} \frac{\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2}{n}
-}\label{eq-equation9}\end{equation}
-
-\hypertarget{sec-simpleex}{%
-\subsection{Simulated data example}\label{sec-simpleex}}
+}\end{equation}
+
+\subsection{Simulated data example}\label{sec-simpleex}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervis5Gau-1.pdf}
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervis5Gau-1.pdf}
 
@@
 visualize the model-in-data space. Models from all NLDR methods show
-five well-separated clusters (see Figure~\ref{fig-gau1_sc},
-Figure~\ref{fig-gau2_sc}, Figure~\ref{fig-gau3_sc},
-Figure~\ref{fig-gau4_sc}, and Figure~\ref{fig-gau5_sc}). This suggests
+five well-separated clusters (see Figure~\ref{fig-gau1-sc},
+Figure~\ref{fig-gau2-sc}, Figure~\ref{fig-gau3-sc},
+Figure~\ref{fig-gau4-sc}, and Figure~\ref{fig-gau5-sc}). This suggests
 that for the five Gaussian cluster dataset, all NLDR methods effectively
@@
 with varying densities, indicating their ability to capture
-within-cluster variation (see Figure~\ref{fig-gau1_sc},
-Figure~\ref{fig-gau2_sc}, and Figure~\ref{fig-gau3_sc}). On the other
+within-cluster variation (see Figure~\ref{fig-gau1-sc},
+Figure~\ref{fig-gau2-sc}, and Figure~\ref{fig-gau3-sc}). On the other
 hand, both TriMAP and PaCMAP show clusters with flat surfaces,
 suggesting a failure to capture within-cluster variation (see
-Figure~\ref{fig-gau4_sc} and Figure~\ref{fig-gau5_sc}). Therefore,
+Figure~\ref{fig-gau4-sc} and Figure~\ref{fig-gau5-sc}). Therefore,
 TriMAP and PaCMAP do not capture the local structure as effectively as
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-modelfiveGau-1.pdf}
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-modelfiveGau-1.pdf}
 
@@
 
-\end{figure}
-
-\begin{figure}[H]
-
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/five_gau_clusters/sc_tsne_1.png}
-
-}
-
-}
-
-\subcaption{\label{fig-gau1_sc1}}
+\end{figure}%
+
+\begin{figure}[H]
+
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_tsne_1.png}
+
+}
+
+\subcaption{\label{fig-gau1-sc1}}
+
 \end{minipage}%
 %
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/five_gau_clusters/sc_tsne_2.png}
-
-}
-
-}
-
-\subcaption{\label{fig-gau1_sc2}}
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_tsne_2.png}
+
+}
+
+\subcaption{\label{fig-gau1-sc2}}
+
 \end{minipage}%
 %
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/five_gau_clusters/sc_tsne_3.png}
-
-}
-
-}
-
-\subcaption{\label{fig-gau1_sc3}}
-\end{minipage}%
-
-\caption{\label{fig-gau1_sc}Screen shots of the \textbf{langevitour} of
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_tsne_3.png}
+
+}
+
+\subcaption{\label{fig-gau1-sc3}}
+
+\end{minipage}%
+
+\caption{\label{fig-gau1-sc}Screen shots of the \textbf{langevitour} of
 the five Gaussian clusters dataset, shows the model with tSNE in high-D,
@@
 
-\end{figure}
-
-\begin{figure}[H]
-
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/five_gau_clusters/sc_umap_1.png}
-
-}
-
-}
-
-\subcaption{\label{fig-gau2_sc1}}
+\end{figure}%
+
+\begin{figure}[H]
+
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_umap_1.png}
+
+}
+
+\subcaption{\label{fig-gau2-sc1}}
+
 \end{minipage}%
 %
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/five_gau_clusters/sc_umap_2.png}
-
-}
-
-}
-
-\subcaption{\label{fig-gau2_sc2}}
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_umap_2.png}
+
+}
+
+\subcaption{\label{fig-gau2-sc2}}
+
 \end{minipage}%
 %
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/five_gau_clusters/sc_umap_3.png}
-
-}
-
-}
-
-\subcaption{\label{fig-gau2_sc3}}
-\end{minipage}%
-
-\caption{\label{fig-gau2_sc}Screen shots of the \textbf{langevitour} of
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_umap_3.png}
+
+}
+
+\subcaption{\label{fig-gau2-sc3}}
+
+\end{minipage}%
+
+\caption{\label{fig-gau2-sc}Screen shots of the \textbf{langevitour} of
 the five Gaussian clusters dataset, shows the model with UMAP in high-D,
@@
 
-\end{figure}
-
-\begin{figure}[H]
-
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/five_gau_clusters/sc_phate_1.png}
-
-}
-
-}
-
-\subcaption{\label{fig-gau3_sc1}}
+\end{figure}%
+
+\begin{figure}[H]
+
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_phate_1.png}
+
+}
+
+\subcaption{\label{fig-gau3-sc1}}
+
 \end{minipage}%
 %
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/five_gau_clusters/sc_phate_2.png}
-
-}
-
-}
-
-\subcaption{\label{fig-gau3_sc2}}
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_phate_2.png}
+
+}
+
+\subcaption{\label{fig-gau3-sc2}}
+
 \end{minipage}%
 %
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/five_gau_clusters/sc_phate_3.png}
-
-}
-
-}
-
-\subcaption{\label{fig-gau3_sc3}}
-\end{minipage}%
-
-\caption{\label{fig-gau3_sc}Screen shots of the \textbf{langevitour} of
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_phate_3.png}
+
+}
+
+\subcaption{\label{fig-gau3-sc3}}
+
+\end{minipage}%
+
+\caption{\label{fig-gau3-sc}Screen shots of the \textbf{langevitour} of
 the five Gaussian clusters dataset, shows the model with PHATE in
@@
 
-\end{figure}
-
-\begin{figure}[H]
-
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/five_gau_clusters/sc_trimap_1.png}
-
-}
-
-}
-
-\subcaption{\label{fig-gau4_sc1}}
+\end{figure}%
+
+\begin{figure}[H]
+
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_trimap_1.png}
+
+}
+
+\subcaption{\label{fig-gau4-sc1}}
+
 \end{minipage}%
 %
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/five_gau_clusters/sc_trimap_2.png}
-
-}
-
-}
-
-\subcaption{\label{fig-gau4_sc2}}
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_trimap_2.png}
+
+}
+
+\subcaption{\label{fig-gau4-sc2}}
+
 \end{minipage}%
 %
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/five_gau_clusters/sc_trimap_3.png}
-
-}
-
-}
-
-\subcaption{\label{fig-gau4_sc3}}
-\end{minipage}%
-
-\caption{\label{fig-gau4_sc}Screen shots of the \textbf{langevitour} of
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_trimap_3.png}
+
+}
+
+\subcaption{\label{fig-gau4-sc3}}
+
+\end{minipage}%
+
+\caption{\label{fig-gau4-sc}Screen shots of the \textbf{langevitour} of
 the five Gaussian clusters dataset, shows the model with TriMAP in
@@
 
-\end{figure}
-
-\begin{figure}[H]
-
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/five_gau_clusters/sc_pacmap_1.png}
-
-}
-
-}
-
-\subcaption{\label{fig-gau5_sc1}}
+\end{figure}%
+
+\begin{figure}[H]
+
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_pacmap_1.png}
+
+}
+
+\subcaption{\label{fig-gau5-sc1}}
+
 \end{minipage}%
 %
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/five_gau_clusters/sc_pacmap_2.png}
-
-}
-
-}
-
-\subcaption{\label{fig-gau5_sc2}}
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_pacmap_2.png}
+
+}
+
+\subcaption{\label{fig-gau5-sc2}}
+
 \end{minipage}%
 %
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/five_gau_clusters/sc_pacmap_3.png}
-
-}
-
-}
-
-\subcaption{\label{fig-gau5_sc3}}
-\end{minipage}%
-
-\caption{\label{fig-gau5_sc}Screen shots of the \textbf{langevitour} of
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_pacmap_3.png}
+
+}
+
+\subcaption{\label{fig-gau5-sc3}}
+
+\end{minipage}%
+
+\caption{\label{fig-gau5-sc}Screen shots of the \textbf{langevitour} of
 the five Gaussian clusters dataset, shows the model with PaCMAP in
@@
 
-\end{figure}
-
-\hypertarget{sec-applications}{%
-\section{Applications}\label{sec-applications}}
-
-\hypertarget{single-cell-rna-seq-data-of-human}{%
+\end{figure}%
+
+\section{Applications}\label{sec-applications}
+
 \subsection{Single-cell RNA-seq data of
-human}\label{single-cell-rna-seq-data-of-human}}
+human}\label{single-cell-rna-seq-data-of-human}
 
@@
 
-{\centering \includegraphics{paper_files/figure-pdf/fig-umapauthor-1.pdf}
+\centering{
+
+\includegraphics{paper_files/figure-pdf/fig-umapauthor-1.pdf}
 
@@
 visualize the model constructed with UMAP overlaid on the high-D data
-(model-in-data space). The figures in Figure~\ref{fig-pbmc1_sc} show
+(model-in-data space). The figures in Figure~\ref{fig-pbmc1-sc} show
 three well-separated clusters, indicating that the suggested UMAP
@@
 Figure~\ref{fig-umapauthor}). However, as shown in
-Figure~\ref{fig-pbmc1_sc}, these clusters are close to each other in
+Figure~\ref{fig-pbmc1-sc}, these clusters are close to each other in
 high-D space. Also, nonlinear continuity patterns and high-density
 patches within the clusters are observed (see
-Figure~\ref{fig-pbmc1_sc}). Therefore, the suggested UMAP representation
+Figure~\ref{fig-pbmc1-sc}). Therefore, the suggested UMAP representation
 (see Figure~\ref{fig-umapauthor}) does not accurately preserve the local
@@
 
-{\centering \includegraphics{paper_files/figure-pdf/fig-nldervisPBMCUMAP-1.pdf}
+\centering{
+
+\includegraphics{paper_files/figure-pdf/fig-nldervisPBMCUMAP-1.pdf}
 
@@
 
-\end{figure}
-
-\begin{figure}[H]
-
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/pbmc3k/sc_1.png}
-
-}
-
-}
-
-\subcaption{\label{fig-pbmc1_sc1}}
+\end{figure}%
+
+\begin{figure}[H]
+
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_1.png}
+
+}
+
+\subcaption{\label{fig-pbmc1-sc1}}
+
 \end{minipage}%
 %
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/pbmc3k/sc_2.png}
-
-}
-
-}
-
-\subcaption{\label{fig-pbmc1_sc2}}
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_2.png}
+
+}
+
+\subcaption{\label{fig-pbmc1-sc2}}
+
 \end{minipage}%
 %
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/pbmc3k/sc_3.png}
-
-}
-
-}
-
-\subcaption{\label{fig-pbmc1_sc3}}
-\end{minipage}%
-
-\caption{\label{fig-pbmc1_sc}Screen shots of the \textbf{langevitour} of
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_3.png}
+
+}
+
+\subcaption{\label{fig-pbmc1-sc3}}
+
+\end{minipage}%
+
+\caption{\label{fig-pbmc1-sc}Screen shots of the \textbf{langevitour} of
 the PBMC3k data set, shows the model in high-D, a video of the tour
@@
 
-{\centering \includegraphics[width=1\textwidth,height=1\textheight]{paper_files/figure-pdf/fig-abserrorPBMC-1.pdf}
+\centering{
+
+\includegraphics[width=1\textwidth,height=1\textheight]{paper_files/figure-pdf/fig-abserrorPBMC-1.pdf}
 
@@
 
-{\centering \includegraphics{paper_files/figure-pdf/fig-tsnesuggest-1.pdf}
+\centering{
+
+\includegraphics{paper_files/figure-pdf/fig-tsnesuggest-1.pdf}
 
@@
 visualized in the data space. The model shows some quirks, as shown in
-Figure~\ref{fig-pbmc2_sc}. The large cluster is divided into three
+Figure~\ref{fig-pbmc2-sc}. The large cluster is divided into three
 subclusters, which are in close to each other.
@@
 
-{\centering \includegraphics{paper_files/figure-pdf/fig-nldervisPBMCtSNE-1.pdf}
+\centering{
+
+\includegraphics{paper_files/figure-pdf/fig-nldervisPBMCtSNE-1.pdf}
 
@@
 
-\end{figure}
-
-\begin{figure}[H]
-
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/pbmc3k/sc_4.png}
-
-}
-
-}
-
-\subcaption{\label{fig-pbmc2_sc4}}
+\end{figure}%
+
+\begin{figure}[H]
+
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_4.png}
+
+}
+
+\subcaption{\label{fig-pbmc2-sc4}}
+
 \end{minipage}%
 %
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/pbmc3k/sc_5.png}
-
-}
-
-}
-
-\subcaption{\label{fig-pbmc2_sc5}}
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_5.png}
+
+}
+
+\subcaption{\label{fig-pbmc2-sc5}}
+
 \end{minipage}%
 %
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/pbmc3k/sc_6.png}
-
-}
-
-}
-
-\subcaption{\label{fig-pbmc2_sc6}}
-\end{minipage}%
-
-\caption{\label{fig-pbmc2_sc}Screen shots of the \textbf{langevitour} of
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_6.png}
+
+}
+
+\subcaption{\label{fig-pbmc2-sc6}}
+
+\end{minipage}%
+
+\caption{\label{fig-pbmc2-sc}Screen shots of the \textbf{langevitour} of
 the PBMC3k data set, shows the model in high-D, a video of the tour
@@
 
-\end{figure}
-
-\hypertarget{handwritten-digits}{%
-\subsection{Handwritten digits}\label{handwritten-digits}}
+\end{figure}%
+
+\subsection{Handwritten digits}\label{handwritten-digits}
 
@@
 
-According to Figure~\ref{fig-mnist1_sc1}, the nonlinear continuity
+According to Figure~\ref{fig-mnist1-sc1}, the nonlinear continuity
 structure observed in the 2D representation of PaCMAP is also visible
@@
 structure, which is an addition pattern of data that cannot be seen in
-the 2D representation (see Figure~\ref{fig-mnist1_sc2}).
+the 2D representation (see Figure~\ref{fig-mnist1-sc2}).
 
@@
 effectively preserves local structure, there shouldn't be any long edges
-in high-D space. However, as shown in Figure~\ref{fig-mnist1_sc3}, some
+in high-D space. However, as shown in Figure~\ref{fig-mnist1-sc3}, some
 long edges exist in high-D space that are not identified as long edges
@@
 
-{\centering \includegraphics{paper_files/figure-pdf/fig-pacmapauthor-1.pdf}
+\centering{
+
+\includegraphics{paper_files/figure-pdf/fig-pacmapauthor-1.pdf}
 
@@
 
-\end{figure}
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-sampleimg-1.pdf}
+\end{figure}%
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-sampleimg-1.pdf}
 
@@
 
-\end{figure}
-
-\begin{figure}[H]
-
-{\centering \includegraphics{paper_files/figure-pdf/fig-nldervisMNISTPACMAP-1.pdf}
+\end{figure}%
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics{paper_files/figure-pdf/fig-nldervisMNISTPACMAP-1.pdf}
 
@@
 
-\end{figure}
-
-\begin{figure}[H]
-
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/mnist/sc_1.png}
-
-}
-
-}
-
-\subcaption{\label{fig-mnist1_sc1}}
+\end{figure}%
+
+\begin{figure}[H]
+
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/mnist/sc_1.png}
+
+}
+
+\subcaption{\label{fig-mnist1-sc1}}
+
 \end{minipage}%
 %
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/mnist/sc_2.png}
-
-}
-
-}
-
-\subcaption{\label{fig-mnist1_sc2}}
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/mnist/sc_2.png}
+
+}
+
+\subcaption{\label{fig-mnist1-sc2}}
+
 \end{minipage}%
 %
-\begin{minipage}[t]{0.33\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{figures/mnist/sc_3.png}
-
-}
-
-}
-
-\subcaption{\label{fig-mnist1_sc3}}
-\end{minipage}%
-
-\caption{\label{fig-mnist1_sc}Screen shots of the \textbf{langevitour}
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/mnist/sc_3.png}
+
+}
+
+\subcaption{\label{fig-mnist1-sc3}}
+
+\end{minipage}%
+
+\caption{\label{fig-mnist1-sc}Screen shots of the \textbf{langevitour}
 of the MNIST digit 1 data set, shows the model-in-data space, a video of
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-sampleimgerror-1.pdf}
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-sampleimgerror-1.pdf}
 
@@
 
-\end{figure}
-
-\hypertarget{sec-conclusions}{%
-\section{Conclusion}\label{sec-conclusions}}
+\end{figure}%
+
+\section{Conclusion}\label{sec-conclusions}
 
@@
 
-\hypertarget{references}{%
-\section*{References}\label{references}}
+\section*{References}\label{references}
 \addcontentsline{toc}{section}{References}

### 2024-03-27 paper.qmd
cleaning up the paper

@@
 ---
-title: "Visualising How Non-linear Dimension Reduction Warps Your Data"
+title: "Looking at Non-Linear Dimension Reductions as Models in the Data Space"
 format: 
@@
 
-High-dimensional (high-D) data is prevalent across various fields, such as ecology and bioinformatics [@Guo2023], due to advancements in data collection technologies [@Johnstone2009; @ayesha2020overview].
-However, visualization of high-D data introduces significant challenges, because the complexity of visualizing data beyond two dimensions [@Jia2022].
-In recent years, interactive and dynamic graphics systems like **liminal** [@article21] —which employs interactive tools like brushing and linking [@article58]—and software tools such as **XGobi**, **GGobi** [@article60], **tourr**  [@article61], **detourr** [@article22], and **langevitour**  [@article09], involving dynamic methods like tours [@Asimov1985], have played a key role in visualizing high-D data (data-vis).
+  - What's the problem:
 
-To create low-dimensional representations (typically in 2D) (m-vis) [@article59] of high-D data, it is common to apply dimension reduction (DR) techniques.
-Approaches for DR involve linear methods such as principal component analysis (PCA) [@Karl1901], non-linear methods such as multi-dimensional scaling (MDS) [@Torgerson1967].
-In the past decade, many new non-linear dimension reduction (NLDR) techniques have emerged, such as t-distributed stochastic neighbor embedding (tSNE) [@Laurens2008] and uniform manifold approximation and projection (UMAP) [@Leland2018].
-NLDR techniques are the 2D models of high-D data in our context.
+  Non-linear dimension reduction being used to summarise high-dimensional data.
 
-It is important to visualize various non-linear dimensionality reduction (NLDR) techniques for the same high-D data in order to understand and find the best representation.
-After doing so, the 2D models may differ considerably from each other and may also deviate from the original data structure in high-dimensional space.
-Therefore, visualizing the 2D model in high-D space (m-in-ds) is more useful to answer different types of questions:
+  - Summary of literature
 
-- Is there a best 2D representation of high-D data or are they all providing  equivalent information?
-Is there a best parameter choice to fit the 2D model?
-How does the model change when it's parameters change?
+  Relevant high-d vis, NLDR history
 
-- How well the does the 2D models capture the data structure? Is the model fitting able to capture different data structure like non-linear, clustering?
+## Method
 
-If we cannot easily ask and answer these questions, our ability to understand the models is limited.
-To find the best 2D model and parameter choices, a better understanding of the underlying science is important.
+  - Create a representation of the model
+  - Algorithm in 2D
+      - Parameters
+      - Tuning
+  - Showing model in high-d
+  - What is learned about simulated examples
 
-Also, the importance of m-vis along with data-vis has been recognized and incorporated into interactive software, **liminal** [@article21].
-But the 2D model and high-D visualize side by side and interactive like brushing and linking connect the data in the two panels.
-To address this challenge, we propose a novel approach by combining the tour technique with a low-dimensional manifold.
-This manifold is created through the synergistic use of NLDR techniques, hexagonal binning, and triangulation.
-This integration facilitates a more understanding of the data structure, how well (or how poorly) NLDR techniques perform.
+## Applications
 
-The outline of this paper is as follows.
-@sec-background provides an detailed overview of dimension reduction methods, and tour technique.
-Building upon this foundation, Section @sec-methods describes the proposed algorithm, including its implementation details, model tuning, summaries, and a synthetic example to demonstrate its functionality.
-In Section @sec-applications, the algorithm's applications on various datasets, particularly in single-cell RNA-seq and handwritten digit data, are showcased.
-These applications demonstrate how some NLDR techniques effectively capture both local and global structures of the original data, while others may result in misleading interpretations.
-Finally, Section @sec-conclusions summarizes the findings and describe the importance of the proposed approach in addressing the challenges associated with selecting the best NLDR method and parameter choices to achieve a more accurate 2D representation of high-D data.
+### pbmc
+     - NLDR view used to illustrate clusters
+     - Use our method to assess is it a reasonable representation
+     - Demonstrate that it is not
+     - Illustrate how to use out method to get a better representation
 
-<!--UMAP with different param choices-->
+### digits: 1
+     - NLDR is used to illustrate different ways 1's are drawn
+     - Use our method to assess is it a reasonable representation
+     - Demonstrate that it is, except for the anomalies
 
-```{r}
-#| warning: false
-#| echo: false
+## Discussion
 
-UMAP_data_7 <- read_rds(file = "data/s_curve/s_curve_umap_7.rds")
-
-#(n-neighbors: 50)
-plot_list1_umap <- plot_UMAP_2D(UMAP_data_7) + #ggtitle("(b)") +
-  geom_point(alpha=0.5, size = 0.5, colour = "#8dd3c7", fill = "#8dd3c7") +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-UMAP_data_15 <- read_rds(file = "data/s_curve/s_curve_umap_15.rds")
-
-#(n-neighbors: 50)
-plot_list2_umap <- plot_UMAP_2D(UMAP_data_15) + #ggtitle("(b)") + 
-  geom_point(alpha=0.5, size = 0.5, colour = "#a65628", fill = "#a65628") +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-UMAP_data_32 <- read_rds(file = "data/s_curve/s_curve_umap_32.rds")
-
-#(n-neighbors: 50)
-plot_list3_umap <- plot_UMAP_2D(UMAP_data_32) + #ggtitle("(b)") +
-  geom_point(alpha=0.5, size = 0.5, color = "#f781bf", fill = "#f781bf") +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-UMAP_data <- read_rds(file = "data/s_curve/s_curve_umap.rds")
-
-#(n-neighbors: 50)
-plot_list4_umap <- plot_UMAP_2D(UMAP_data) + #ggtitle("(b)") +
-  geom_point(alpha=0.5, size = 0.5, colour = "#999999", fill = "#999999") +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'd', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "2D layouts from UMAP applied for the S-curve data: (a) UMAP (n_neighbors = 7), (b) UMAP (n_neighbors = 15), (c) UMAP (n_neighbors = 32), (d) UMAP (n_neighbors = 50).
-Is there a best hyperparameter choice in representing UMAP or are they all providing  equivalent information?"
-#| label: fig-nldervisUMAP
-#| out-width: 100%
-
-plot_list1_umap + plot_list2_umap + plot_list3_umap + plot_list4_umap +
-  plot_layout(ncol=4)
-```
-
-## Background {#sec-background}
-
-### Dimension Reduction
-
-Consider the high-D data a rectangular matrix $X_{n \times p}$, where $X_{n \times p} = \begin{bmatrix} \textbf{x}_{1} & \textbf{x}_{2} & \cdots & \textbf{x}_{n}\\ \end{bmatrix}^\top$, with $n$ observations in $p$ dimensions.
-The objective is to discover a low-dimensional projection $Y_{n \times d} = \begin{bmatrix} \textbf{y}_{1} & \textbf{y}_{2} & \cdots & \textbf{y}_{n}\\ \end{bmatrix}^\top$, represented as an $n$ × $d$ matrix, where $d \ll p$.
-The reduction process seeks to remove noise from the original data set while retaining essential information.
-
-There are two main categories of dimension reduction techniques: linear and non-linear methods.
-Linear techniques involve a linear transformation of the data, with one popular example being PCA.
-PCA performs an eigen-decomposition of the sample covariance matrix to obtain orthogonal principal components that capture the variance of the data [@Karl1901].
-
-In contrast, NLDR techniques generate the low-dimensional representation $Y$ from the high-dimensional data $X$, often using pre-processing techniques like $k$-nearest neighbors graph or kernel transformations.
-Multidimensional Scaling (MDS) is a class of NLDR methods that aims to construct an embedding $Y$ in a low-dimensional space, approximating the pair-wise distances in $X$ [@Torgerson1967].
-Variants of MDS include non-metric scaling [@article62] and Isomap, which estimate geodesic distances to create the low-dimensional representation [@article63].
-Other approaches based on diffusion processes, like diffusion maps [@article64] and the PHATE (Potential of Heat-diffusion for Affinity-based Trajectory Embedding) algorithm [@article03], also fall under NLDR methods.
-
-#### Non-linear dimension reduction techniques
-
-NLDR techniques are crucial for analyzing and displaying high-dimensional data, where linear approaches may not adequately capture complexities in relationships between variables [@Johnstone2009].
-One of the challenges with NLDR techniques is the selection and tuning of appropriate hyperparameters [@liao2023].
-This process involves finding the suitable combination of hyperparameters that enhances the performance of the NLDR technique, considering the characteristics of the dataset and the specific goals of the analysis.
-
-<!--This process requires a delicate balance to optimize the performance of NLDR techniques, taking into account the intricacies of the dataset and the specific goals of the analysis.-->
-
-Additionally, another challenge is lack of reverse mapping.
-Techniques like PCA and auto-encoders [@article65] provide a way to map back from the low-dimensional space to the high-D space, facilitating data reconstruction.
-However, some NLDR methods, such as tSNE, don't have a specific way to reconstruct the original data from the low-dimensional space.
-
-In this article, mainly focus on five NLDR techniques. They are tSNE, UMAP, PHATE, TriMAP [@article02], and Pairwise Controlled Manifold Approximation (PaCMAP) [@Yingfan2021].
-
-Among these, tSNE [@Laurens2008] stands out for its ability to preserve pairwise distances.
-By minimizing the divergence between probability distributions in both high and low-dimensional spaces, tSNE effectively uncovers intricate structures and patterns within the data.
-Its application is widespread, particularly in tasks requiring the visualization of clusters and local relationships.
-However, achieving effective results requires careful consideration of hyperparameters, such as perplexity.
-
-UMAP [@Leland2018] is a useful technique for simplifying data while maintaining both local and overall structures.
-It builds a fuzzy topological view by considering nearby data points and then optimizes a simplified version to match that view.
-UMAP is known for working well with different scales of relationships in data and is efficient in handling large datasets.
-However, it's important to choose parameters like neighbors and minimum distance carefully, as they can affect the results.
-
-Furthermore, PHATE [@article03] is great for understanding how things develop, especially in single-cell genomics.
-It uses a heat diffusion process to capture relationships between data points, like points along a trajectory.
-While PHATE is excellent for revealing these developmental structures, it requires careful tuning of its parameters because of its specialized focus.
-
-Additionally, TriMAP [@article02] takes a special approach by creating a triangulated graph representation of the data.
-This method is good at understanding both local and global structures by treating the data as a network of triangles.
-TriMAP is powerful in capturing complicated structures, but it's important to choose parameters carefully, like deciding how many neighbors to consider.
-
-PaCMAP [@Yingfan2021] is different because it adds supervised learning to make a 2D representation while keeping the relationships between pairs of points.
-It builds a graph using distances between pairs and then makes the 2D representation better using a customizable loss function.
-What's special about PaCMAP is that it can use class labels or extra information to guide how it makes the 2D representation.
-This gives users a way to change how PaCMAP works to fit their needs better.
-
-<!--Among these techniques, tSNE stands out for its emphasis on preserving pairwise distances.
-By minimizing the divergence between probability distributions in both the high and low-dimensional spaces, t-SNE effectively reveals intricate structures and patterns within the data.
-Its application is widespread in tasks requiring the visualization of clusters and local relationships, though it does require careful consideration of the perplexity parameter for optimal results.
-
-UMAP is another powerful non-linear technique that strikes a balance between preserving local and global structures.
-Constructing a fuzzy topological representation using a weighted k-nearest neighbors graph, UMAP optimizes the low-dimensional embedding to resemble this representation.
-Known for its efficiency and scalability, UMAP is versatile across various scales of relationships in the data, although parameter sensitivity, particularly concerning the choice of neighbors, must be taken into account.
-
-For trajectory data, PHATE provides specialized capabilities.
-It models the affinity between data points, simulating a heat diffusion process to capture developmental processes, particularly in single-cell genomics.
-While PHATE excels in revealing trajectory structures and offering insights into cellular development, it necessitates careful parameter tuning due to its specialized nature.
-
-TriMAP adopts a unique approach by approximating the data manifold through the construction of a triangulated graph representation.
-This technique efficiently captures both global and local structures by representing the data as a network of triangles.
-TriMAP's strength lies in its ability to efficiently capture complex structures, albeit with sensitivity to parameter choices, including the number of neighbors.
-
-In contrast, PaCMAP introduces supervised learning to create a low-dimensional representation while preserving pair-wise relationships.
-Constructing a graph based on pair-wise distances, PaCMAP optimizes an embedding using a customizable loss function.
-Particularly notable is PaCMAP's flexibility in incorporating class labels or additional information to guide the embedding process, offering users a means to customize its behavior and performance.-->
-
-```{r}
-#| warning: false
-#| echo: false
-
-data <- read_rds("data/s_curve/s_curve.rds")
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-training_data <- read_rds("data/s_curve/s_curve_training.rds")
-test_data <- read_rds("data/s_curve/s_curve_test.rds")
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-tSNE_data <- read_rds("data/s_curve/s_curve_tsne_27.rds")
-
-plot_list1 <- plot_tSNE_2D(tSNE_data) +  
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-UMAP_data <- read_rds(file = "data/s_curve/s_curve_umap.rds")
-
-#(n-neighbors: 50)
-plot_list2 <- plot_UMAP_2D(UMAP_data) + 
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-#(knn: 5)
-# PHATE_data <- Fit_PHATE(training_data, knn = 5, with_seed = 20240110)
-# write_csv(PHATE_data, paste0(here::here(), "/data/phate_data_s_curve.csv"))
-
-PHATE_data <- read_rds(file = "data/s_curve/s_curve_phate.rds")
-
-plot_list3 <- plot_PHATE_2D(PHATE_data) + 
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-# tem_dir <- tempdir()
-# 
-# Fit_TriMAP_data(training_data, tem_dir)
-# 
-# path <- file.path(tem_dir, "df_2_without_class.csv")
-# path2 <- file.path(tem_dir, "dataset_3_TriMAP_values.csv")
-# 
-# Fit_TriMAP(as.integer(2), as.integer(5), as.integer(4), as.integer(3), path, path2)
-# 
-# TriMAP_data <- read_csv(path2)
-# write_csv(TriMAP_data, paste0(here::here(), "/data/trimap_data_s_curve.csv"))
-
-TriMAP_data <- read_rds(file = "data/s_curve/s_curve_trimap.rds")
-
-#(n-inliers: 5, \n n-outliers: 4, n-random: 3)
-plot_list4 <- plot_TriMAP_2D(TriMAP_data) + 
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-# tem_dir <- tempdir()
-# 
-# Fit_PacMAP_data(training_data, tem_dir)
-# 
-# path <- file.path(tem_dir, "df_2_without_class.csv")
-# path2 <- file.path(tem_dir, "dataset_3_PaCMAP_values.csv")
-# 
-# Fit_PaCMAP(as.integer(2), as.integer(10), "random", 0.9, as.integer(2), path, path2)
-# 
-# PacMAP_data <- read_csv(path2)
-# write_csv(PacMAP_data, paste0(here::here(), "/data/pacmap_data_s_curve.csv"))
-
-PaCMAP_data <- read_rds(file = "data/s_curve/s_curve_pacmap.rds")
-
-#(knn: 10, init: random, \n MN-ratio: 0.9, FP-ratio: 2)
-plot_list5 <- plot_PaCMAP_2D(PaCMAP_data) +  
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'e', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "2D layouts from different NLDR techniques applied the same data: (a) tSNE (perplexity = 27), (b) UMAP (n_neighbors = 50), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2).
-Is there a best representation of the original data or are they all providing  equivalent information?"
-#| label: fig-nldervis
-#| out-width: 100%
-
-plot_list1 + plot_list2 + plot_list3 + plot_list4 + plot_list5 +
-  plot_layout(ncol=5)
-```
-
-### Linear overviews using tours
-
-A tour is a powerful visualization technique used to explore the shape and global structure of high-dimensional data by generating a sequence of projections, typically into two dimensions.
-There are two main types of tours: the grand tour [@Asimov1985] and the guided tour [@article29].
-A grand tour involves randomly selecting new orthonormal bases, enabling users to understand the structure by exploring the subspace of d-dimensional projections [@Asimov1985].
-In contrast, a guided tour can be employed to generate a sequence of 'interesting' projections based on an index function [@article29].
-
-The process begins with the data matrix $X$.
-It generates a sequence of $p$ × $d$ orthonormal projection matrices (bases) $P_t$, usually $d$ is one or two dimensions.
-For each pair of orthonormal bases $P_t$ and $P_{t+1}$, a geodesic path is interpolated to create smooth animation between projections.
-The resulting tour continuously visualizes the projected data $Y_t$ = $XP_t$ as it interpolates between successive bases.
-
-Furthermore, software like **langevitour** can visualize both types of tours, providing flexibility for exploring high-dimensional data with various objectives.
-In our context, use grand tour along with the model to observe how effectively the model captures the underlying structure of the data.
-
-## Methodology {#sec-methods}
-
-In this paper, we introduce a novel method to determine the most effective NLDR technique and the best hyperparameter choice that provides the most useful representation of high-D data.
-Our approach involves dividing the high-D dataset into two parts: a training set for constructing the model and a test set for generating predictive values and residuals.
-Our algorithm takes a 2D embedding data as the input and generate a tour that displays the high-D wireframe to overlay the data.
-The flow chart of the proposed algorithm is shown in @fig-meth.
-The algorithm consists of two main phases: (1) generating the model in the 2D space and (2) lifting the model into high-D space.
-The main steps of the algorithm are described in detail in this section using UMAP 2D embedding of the S-curve dataset.
-This dataset has seven dimensions, including four noise dimensions that were added to the original 3D data.
-
-![The flow diagram shows the main steps of our algorithm.
-There are two basic phases, one to generate the model in the 2D space, and other to map the model into the high-D space.](figures/workflow.png){#fig-meth fig-align="center" width="100%" height="100%"}
-
-### Preprocessing steps
-
-To reduce computational complexity when applying NLDR techniques to high-D data and to reduce noise presence, PCA [@article67, @article68, @article69] is used as a preprocessing step.
-PCA involves identifying principal components that maximize variance.
-These components are then used as the high-D data for the algorithm.
-
-### Constructing the 2D model {#sec-construct2d}
-
-**Step 1: Scaling NLDR data**
-
-First, we prepare the 2D embedding data to fit within the bounds required for regular hexagonal binning.
-To achieve this, we implement two key scaling steps.
-Scale the first 2D embedding component to range between $0$ and $1$, ensuring that all data points fall within this normalized interval.
-Secondly, we scale the second 2D embedding component to range between $0$ and $y_{max}$ (see @eq-equation3).
-
-The calculation of $y_{max}$ involves several steps.
-First, the aspect ratio ($ar$) is computed by dividing the range of the second 2D embedding component ($r_2$) by the range of the first 2D embedding component ($r_1$) (see @eq-equation1).
-Then, the hexagon ratio ($hr$) is determined by dividing the height of the hexagon ($hb$) by its width ($wb$) (see @eq-equation2).
-Finally, $y_{max}$ is derived by taking the ceiling of $\frac{ar}{hr}$ and multiplying it by $hr$.
-This process ensures that $y_{max}$ is an integer multiple of $hr$, accommodating the grid layout of the hexagonal bins.
-
-$$
- ar = \frac{r_2}{r_1}
-$$ {#eq-equation1}
-
-$$
- hr = \frac{hb}{wb}
-$$ {#eq-equation2}
-
-$$
- y_{max} = \left\lceil\frac{ar}{hr}\right\rceil * hr
-$$ {#eq-equation3}
-
-**Step 2: Hexagonating NLDR data**
-
-Hexagonating NLDR data (see @fig-meth Step 2) involves partitioning the NLDR data into hexagonal bins, a technique commonly referred to as hexagonal binning [@Carr1987, @article66].
-This method use a hexagonal grid to create a bivariate histogram that effectively visualizes the structure of high-D data.
-Hexagons, one of only three regular polygons capable of tessellating a plane [@Carr2013], offer unique advantages due to their symmetry of nearest neighbors and maximal number of sides for such tessellations [@Dan2023].
-This geometric property makes hexagons more efficient in covering the plane compared to other regular tessellations and reduces visual bias when displaying data densities [@Dan2023].
-
-In our algorithm, we aim to conduct regular hexagonal binning, involving the computation of hexagonal grid configurations, the generation of the grid, and the assignment of the NLDR data points to hexagons.
-
-**(a) Computation of hexagonal grid configurations**
-
-In the computation of hexagonal bin configurations, there are mainly two configurations to consider: the number of bins along the x and y axes.
-The first step involves in determining the hexagonal size ($s$) and the buffer along the x and y axes ($buffer_{x}$ and $buffer_{y}$).
-Here, the hexagonal size ($s$) represents the radius of the outer circle surrounding the hexagon.
-The buffer is important in expanding the boundary to accommodate potential outliers or edge cases.
-
-When computing the number of bins along the x and y axes ($b_1$ and $b_2$), the process begins by calculating the range of the respective 2D embedding component.
-After that, the range is adjusted to accommodate the buffer amount.
-Then, the spacing between hexagons is determined based on $s$.
-Finally, the number of bins along the axis is computed by dividing the adjusted range by the spacing.
-
-For the computation of the number of bins along the x-axis ($b_1$) (see @eq-equation12), the range of the first embedding component is defined as $r_1$, while the buffer amount along the x-axis is denoted as $buffer_{x}$, and the horizontal spacing is represented by $h$ (see @eq-equation11).
-On the other hand, the number of bins along the y-axis ($b_2$) (see @eq-equation14) is computed based on the range of the second embedding component ($r_2$), the buffer amount along the y-axis ($buffer_{y}$), and the vertical spacing ($v$) (see @eq-equation13).
-
-$$
- h = \sqrt{3} * s
-$$ {#eq-equation11}
-
-$$
- b_1 = \frac{r_1 + buffer_{x}}{h}
-$$ {#eq-equation12}
-
-$$
-  v = 1.5 * s
-$$ {#eq-equation13}
-
-$$
- b_2 = \frac{r_2 + buffer_{y}}{v}
-$$ {#eq-equation14}
-
-**(b) Generation of the hexagonal grid**
-
-In this process, the first step involves generating centroids within the hexagonal grid.
-This begins by defining the starting coordinates for the hexagons.
-With the number of bins computed along the x and y axes, along with the horizontal and vertical spacing, the centroids are iteratively computed starting from the hexagonal starting coordinates.
-
-Once the centroids for the hexagonal grid are obtained, the next step is to compute the hexagonal coordinates.
-For example, if the centroid of a hexagonal bin is defined as $(C_x, C_y)$, the hexagonal coordinates can be computed using $d_x$ (see @eq-equation15) and $d_y$ (see @eq-equation16) (see @fig-hexcoord).
-
-$$
-d_x = \frac{h}{2}
-$$ {#eq-equation15}
-
-$$
-d_y = \frac{2* v * 1.15}{\sqrt{3}} 
-$$ {#eq-equation16}
-
-```{r}
-#| echo: false
-#| fig-pos: H
-#| fig-cap: "The hexagonal coordinates computed for the centroid $(C_x, C_y)$."
-#| label: fig-hexcoord
-#| out-width: 100%
-#| out-height: 100%
-knitr::include_graphics("figures/hex_coord.png")
-```
-
-**(c) Assignment of the NLDR data points to hexagons**
-
-After obtaining the centroids of the hexagonal bins for the entire grid, the process of assigning NLDR points to hexagons involves determining the nearest hexagonal bin for each NLDR point using the 2D Euclidean distance.
-Then, the hexagonal ID of the nearest hexagon is assigned to the corresponding NLDR points.
-
-**Step 3: Obtaining bin centroids or bin means**
-
-In the previous step, the algorithm clusters the 2D embedding data into hexagons. Following this, in this step, the bin centroids or bin means (see @fig-meth Step 3) are obtained [@Carr2013].
-
-The bin centroid ($C_k^{(2)}$) for a $k^{th}$ hexagon with hexagonal grid coordinates $(h^{k}x_{i}, h^{k}y_{i})$, where $i = 1 \dots 6$ can be defined as:
-
-$$
-C_k^{(2)} = (C_{ky_1}, C_{ky_2}) = \left(\frac{\sum_{i=1}^{6} h^{k}x_{i}}{6}, \frac{\sum_{i=1}^{6} h^{k}y_{i}}{6}\right).
-$$ {#eq-equation4}
-
-Also, the bin mean ($C_k^{(2)}$) is defined as the mean of the data points within the $k^{th}$ hexagon (see @eq-equation5).
-
-$$
-C_k^{(2)} = (C_{ky_1}, C_{ky_2}) = \left(\frac{1}{n_k} \sum_{i=1}^{n_k} y_{1i}, \frac{1}{n_k} \sum_{i=1}^{n_k} y_{2i}\right),
-$$ {#eq-equation5}
-
-where $n_k$ is the number of data points within the hexagon, $y_{1i}$ and $y_{2i}$ are the $x$ and $y$ coordinates of the $i^{th}$ data point within the hexagon.
-
-**Step 4: Triangulating bin centroids or bin means**
-
-In this step, the algorithm proceeds to triangulate the hexagonal bin centroids or bin means (see @fig-meth Step 4).
-Triangulation is a fundamental process in computational geometry and computer graphics that involves dividing a set of points in a given space into interconnected triangles [@article30].
-One common algorithm used for triangulation is Delaunay triangulation [@article26, @article54], where points are connected in a way that maximizes the minimum angles of the resulting triangles, leading to a more regular and well-conditioned triangulation.
-
-Delaunay triangulation can be defined as follows:
-
-Let $C^{(2)} = \{C_1^{(2)}, C_2^{(2)}, ..., C_m^{(2)}\}$ be a set of $m$ bin centroids or bin means in the plane.
-Delaunay triangulation of $C^{(2)}$, denoted as $DT(C^{(2)})$, is a triangulation of the convex hull of $C^{(2)}$ such that the circumcircle of every triangle in the triangulation contains no other points from $C^{(2)}$.
-
-Given that the hexagons are regular, the resulting triangles will mostly be equilateral.
-
-### Lifting the model into high dimensions
-
-#### Lifting the triangular mesh points into high dimensions
-
-Consider $f: \mathbb{R}^p \rightarrow \mathbb{R}^2$ be a function that maps the high-D data ($X_{n \times p}$) to its NLDR equivalent ($Y_{n \times d}$).
-Then, let $g: \mathbb{R}^2 \rightarrow \mathbb{R}^2$ be a function that maps each 2D embedding point to its closest centroid ($C^{(2)}$).
-It follows that $f(g(x))$ maps the high-D points $x$ to the centroid in 2D ($C_k^{(2)}$).
-Also, define a function $v: \mathbb{R}^2 \rightarrow \mathbb{R}^p$ maps the 2D centroid ($C^{(2)}$) to the high-D mean of the points ($C^{(p)}$) in the hexagon.
-
-The high-D mean of all the points in $k^{th}$ hexagon by
-
-$$
-C_k^{(p)} = (C_{kx_1}, ..., C_{kx_p}) = \left(\frac{1}{n_k} \sum_{i=1}^{n_k} x_{1i}, \frac{1}{n_k} \sum_{i=1}^{n_k} x_{2i}, \dots ,\frac{1}{n_k} \sum_{i=1}^{n_k} x_{pi}\right).
-$${#eq-equation6}
-
-Therefore,
-
-$$
-v(C_k^{(2)}) = C_k^{(p)}.
-$${#eq-equation7}
-
-Therefore, $f(g(x))$ gives the 2D centroid associated with high-D points $x$, and $v(C_k^{(2)})$ gives the high-D centroid associated with 2D point $C_k^{(2)}$.
-Thus, $v(f(g(x)))$ gives the high-D centroid ($C_k^{(p)}$) associated with the 2D embedding of the points $x$.
-
-#### Lifting the 2D triangular mesh into high dimensions
-
-As described in Step 4 of @sec-construct2d, during the triangulation process in 2D space, vertices are identified to form edges.
-With the knowledge of the high-D mappings for the 2D hexagonal bins, the vertices connected in 2D are also connected in high-D (see video linked in @fig-wkhighD).
-
-```{r}
-#| warning: false
-#| echo: false
-#| message: false
-
-umap_s_curve_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = UMAP_data, 
-                                    x = "UMAP1", y = "UMAP2"))) |>
-  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
-                  "UMAP2" = "scaled_UMAP2")) |>
-  dplyr::mutate(ID = 1:NROW(UMAP_data))
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-#| message: false
-
-num_bins_x <- 5
-num_bins_y <- 10
-hex_size <- 0.19
-
-hb_obj <- hex_binning(data = umap_s_curve_scaled, x = "UMAP1", 
-                              y = "UMAP2", num_bins_x = num_bins_x, 
-                              num_bins_y = num_bins_y, x_start = NA, y_start = NA, 
-                              buffer_x = NA, buffer_y = NA, hex_size = hex_size, col_start = "UMAP")
-
-## Data set with all possible centroids in the hexagonal grid
-all_centroids_df <- as.data.frame(do.call(cbind, hb_obj$centroids))
-
-## Generate all coordinates of hexagons
-hex_grid <- as.data.frame(do.call(cbind, hb_obj$hex_poly))
-
-## To obtain the standardise counts within hexbins
-counts_df <- as.data.frame(do.call(cbind, hb_obj$std_cts))
-df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
-                                             counts_df = counts_df)
-
-UMAP_data_with_hb_id <- as.data.frame(do.call(cbind, hb_obj$data_hb_id))
-
-model_object <- fit_highd_model( training_data = training_data, 
-                                 nldr_df_with_id = umap_s_curve_scaled, 
-                                 x = "UMAP1", y = "UMAP2", 
-                                 num_bins_x = num_bins_x, 
-                                 num_bins_y = num_bins_y, 
-                                 x_start = NA, y_start = NA, 
-                                 buffer_x = NA, buffer_y = NA, 
-                                 hex_size = hex_size,
-                                 is_rm_lwd_hex = FALSE, 
-                                 benchmark_to_rm_lwd_hex = NA, 
-                                 col_start_2d = "UMAP", 
-                                 col_start_highd = "x")
-
-df_bin_centroids <- model_object$df_bin_centroids
-df_bin <- model_object$df_bin
-
-## Triangulate bin centroids
-tr1_object <- tri_bin_centroids(df_bin_centroids, x = "c_x", y = "c_y")
-tr_from_to_df <- gen_edges(tri_object = tr1_object)
-
-# Create the tibble with x and y coordinates
-tr_df <- tibble::tibble(x = tr1_object$x, y = tr1_object$y)
-
-## Compute 2D distances
-distance <- cal_2d_dist(tr_coord_df = tr_from_to_df, 
-                                start_x = "x_from", start_y = "y_from", 
-                                end_x = "x_to", end_y = "y_to", 
-                                select_vars = c("from", "to", "distance"))
-
-## Map hb_id
-
-df_bin_centroids_n <- df_bin_centroids |>
-  select(hexID) |>
-  mutate(ID = 1:NROW(df_bin_centroids))
-
-tr_from_to_df_n <- left_join(tr_from_to_df, df_bin_centroids_n, by = c("from" = "ID")) |>
-  rename("from_hb_id" = "hexID")
-
-tr_from_to_df_n <- left_join(tr_from_to_df_n, df_bin_centroids_n, by = c("to" = "ID")) |>
-  rename("to_hb_id" = "hexID")
-
-hb_id_selected <- tr_from_to_df_n |>
-  dplyr::filter((from == 1) & (to == 3))
-
-# Filter and label small and long edges
-distance_df_small_edges <- distance |>
-  dplyr::filter((from != 1) | (to != 3)) |>
-  dplyr::mutate(type = "small_edges")
-
-distance_df_long_edges <- distance |>
-  dplyr::filter((from == 1) & (to == 3)) |>
-  dplyr::mutate(type = "long_edges")
-
-# Combine small and long edges
-distance_edges <- dplyr::bind_rows(distance_df_small_edges, distance_df_long_edges)
-
-tr_from_to_df_coord_with_group <- inner_join(tr_from_to_df, distance_edges, by = c("from" = "from", "to" = "to"))
-
-## To generate a data set with high-D and 2D training data
-df_all <- dplyr::bind_cols(training_data |> dplyr::select(-ID), UMAP_data_with_hb_id)
-
-UMAP_data_with_hb_id_selected <- UMAP_data_with_hb_id |> dplyr::filter(hb_id %in% c(hb_id_selected$from_hb_id, hb_id_selected$to_hb_id)) |>
-  dplyr::mutate(select_point = if_else(hb_id == hb_id_selected$from_hb_id, "data1", "data2"))
-
-UMAP_data_with_hb_id_selected_no <- UMAP_data_with_hb_id |> dplyr::filter(!(hb_id %in% c(hb_id_selected$from_hb_id, hb_id_selected$to_hb_id)))
-
-bin_coord_1 <- tr_df |>
-  dplyr::filter((x == hb_id_selected$x_from)) |>
-  dplyr::filter((y == hb_id_selected$y_from))
-
-# bin_coord_1 <- tr_df |>
-#   head(1)
-
-bin_coord_2 <- tr_df |>
-  dplyr::filter((x == hb_id_selected$x_to)) |>
-  dplyr::filter((y == hb_id_selected$y_to))
-
-# bin_coord_2 <- tr_df |>
-#   dplyr::filter(row_number() == 2)
-
-bin_coord <- dplyr::bind_rows(bin_coord_1, bin_coord_2)
-
-a1 <- ggplot(bin_coord, aes(x = x, y = y)) +
-  geom_segment(
-    aes(x = x_from, y = y_from, xend = x_to, yend = y_to, colour = type),
-    data = tr_from_to_df_coord_with_group) +
-  geom_point() +
-  #coord_equal() +
-  coord_cartesian(xlim =c(-0.5, 1.5), ylim = c(-0.5, 2.7)) +
-  geom_point(data = UMAP_data_with_hb_id_selected, aes(x = UMAP1, y = UMAP2, colour = select_point), alpha = 0.5, size = 1) +
-  scale_color_manual(values=c("#e6550d", "#000000", "#e31a1c", "#f6e8c3")) +
-  geom_point(size = 3, colour = "#33a02c") + 
-  #ggtitle("(b)") + 
-  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
-  theme_light() +
-  #coord_equal() +
-  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
-  )
-
-a2 <- ggplot(data = hex_grid, aes(x = x, y = y)) + 
-  geom_polygon(color = "#deebf7", aes(group = hex_poly_id), fill = "#ffffff") +
-  geom_segment(
-    aes(x = x_from, y = y_from, xend = x_to, yend = y_to), colour = "#f6e8c3",
-    data = tr_from_to_df_coord_with_group
-  ) +
-  geom_point(data = UMAP_data_with_hb_id_selected, aes(x = UMAP1, y = UMAP2, colour = select_point), size = 1, alpha = 0.5) +
-  scale_color_manual(values=c("#e6550d", "#000000")) +
-  geom_point(data = df_bin_centroids |> dplyr::filter(hexID %in% c(hb_id_selected$from_hb_id, hb_id_selected$to_hb_id)), aes(x = c_x, y = c_y), colour = "#33a02c", size = 3) +
-  geom_point(data = UMAP_data_with_hb_id_selected_no, aes(x = UMAP1, y = UMAP2), colour = "#6a3d9a", size = 1, alpha = 0.5) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  coord_equal() +
-  theme_light() +
-  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
-  )
-
-## Script to make coloring data points with a line
-
-df1 <- df_all |> 
-  dplyr::filter(hb_id == as.character(hb_id_selected$from_hb_id)) |>
-  dplyr::select(tidyselect::starts_with("x")) |>
-  dplyr::mutate(type = "data1") ## original dataset
-
-df2 <- df_all |> 
-  dplyr::filter(hb_id == as.character(hb_id_selected$to_hb_id)) |>
-  dplyr::select(tidyselect::starts_with("x")) |>
-  dplyr::mutate(type = "data2") ## original dataset
-
-df3 <- df_all |> 
-  dplyr::filter(!(hb_id %in% c(as.character(hb_id_selected$from_hb_id), as.character(hb_id_selected$to_hb_id)))) |>
-  dplyr::select(tidyselect::starts_with("x")) |>
-  dplyr::mutate(type = "data") ## original dataset
-
-df_b <- df_bin |> 
-  dplyr::filter(hb_id %in% df_bin_centroids$hexID) |>
-  dplyr::filter(hb_id %in% c(as.character(hb_id_selected$from_hb_id), as.character(hb_id_selected$to_hb_id))) |>
-  dplyr::select(-hb_id) |>
-  dplyr::mutate(type = "model") ## Data with summarized mean
-
-df_exe <- dplyr::bind_rows(df_b, df1, df2, df3)
-
-lg_high <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)], group = df_exe$type, pointSize = append(rep(4, NROW(df_b)), rep(2, NROW(df1) + NROW(df2) + NROW(df3))), levelColors = c("#6a3d9a", "#e6550d", "#000000", "#33a02c"), lineFrom = 1, lineTo = 2, lineColors = "#e31a1c")
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-#| label: fig-wkhighD
-#| fig-cap: How the 2D model lift into high dimensions?
-(a) visualize the points and the hexagonal bin centroids related $7^{th}$ and $12^{th}$ hexagons, (b) visualization of the edge connected the $7^{th}$ and $12^{th}$ hexagons (colored in red) in the triangular mesh.
-A video of tour animation is available at <https://youtu.be/hxU91xNTJL0>.
-
-a2 + a1 +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(ncol = 3) &
-  theme(legend.position='none', plot.tag = element_text(size = 8))
-```
-
-### Tuning the model
-
-The performance and robustness of our model depend on four key parameters: (i) the total number of bins ($b$), (ii) a benchmark value used to remove low-density hexagons, (iii) a benchmark value used to remove long edges, and (iv) starting point of the hexagonal grid.
-However, there is no analytical formula to calculate an appropriate value for these parameters.
-The selection of these parameter values depends on the model performance computed by Mean Squared Error (MSE) (see @sec-goodfit).
-
-#### Total number of bins
-
-The number of hexagonal bins in the hexagonal grid has a considerable impact on the construction of the 2D model, serving as the initial step in building the 2D model.
-The chosen total number of bins must effectively capture the structure of the NLDR data.
-If the number of bins is too low, the model may not be able to capture the structure of the NLDR data effectively (see @fig-binsize (a)), while if there are too many bins, it may result in over-fitting the individual points of the NLDR data (see @fig-binsize (c)).
-Therefore, it is important to determine an appropriate number of bins to construct an effective model.
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-
-num_bins_x <- 3
-num_bins_y <- 4
-hex_size <- 0.62
-
-hb_obj <- hex_binning(data = umap_s_curve_scaled, x = "UMAP1", 
-                      y = "UMAP2", num_bins_x = num_bins_x, 
-                      num_bins_y = num_bins_y, x_start = NA, y_start = NA, 
-                      buffer_x = NA, buffer_y = NA, hex_size = hex_size, col_start = "UMAP")
-
-## Data set with all possible centroids in the hexagonal grid
-all_centroids_df <- as.data.frame(do.call(cbind, hb_obj$centroids))
-counts_df <- as.data.frame(do.call(cbind, hb_obj$std_cts))
-
-## Generate all coordinates of hexagons
-hex_grid <- as.data.frame(do.call(cbind, hb_obj$hex_poly))
-
-hex_full_count_df_loop1 <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
-
-p1 <-  ggplot(data = hex_full_count_df_loop1, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
-  geom_point(data = umap_s_curve_scaled, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  #coord_equal() +
-  coord_cartesian(xlim =c(-0.2, 1.2), ylim = c(-0.1, 2.5)) +
-  theme_void() +
-  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=8), #change legend title font size
-        legend.text = element_text(size=6)) +
-  guides(fill = guide_colourbar(title = "Standardized count")) +
-  annotate(geom = 'text', label = "a", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
-
-```
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-
-num_bins_x <- 6
-num_bins_y <- 14
-hex_size <- 0.13
-
-hb_obj <- hex_binning(data = umap_s_curve_scaled, x = "UMAP1", 
-                      y = "UMAP2", num_bins_x = num_bins_x, 
-                      num_bins_y = num_bins_y, x_start = NA, y_start = NA, 
-                      buffer_x = NA, buffer_y = NA, hex_size = hex_size, col_start = "UMAP")
-
-## Data set with all possible centroids in the hexagonal grid
-all_centroids_df <- as.data.frame(do.call(cbind, hb_obj$centroids))
-counts_df <- as.data.frame(do.call(cbind, hb_obj$std_cts))
-
-## Generate all coordinates of hexagons
-hex_grid <- as.data.frame(do.call(cbind, hb_obj$hex_poly))
-
-hex_full_count_df_loop2 <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
-
-p2 <-  ggplot(data = hex_full_count_df_loop2, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
-  geom_point(data = umap_s_curve_scaled, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  #coord_equal() +
-  coord_cartesian(xlim =c(-0.2, 1.2), ylim = c(-0.1, 2.5)) +
-  theme_void() +
-  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=8), #change legend title font size
-        legend.text = element_text(size=6)) +
-  guides(fill = guide_colourbar(title = "Standardized count")) +
-  annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
-
-```
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-
-num_bins_x <- 12
-num_bins_y <- 28
-hex_size <- 0.06
-
-hb_obj <- hex_binning(data = umap_s_curve_scaled, x = "UMAP1", 
-                      y = "UMAP2", num_bins_x = num_bins_x, 
-                      num_bins_y = num_bins_y, x_start = NA, y_start = NA, 
-                      buffer_x = NA, buffer_y = NA, hex_size = hex_size, col_start = "UMAP")
-
-## Data set with all possible centroids in the hexagonal grid
-all_centroids_df <- as.data.frame(do.call(cbind, hb_obj$centroids))
-counts_df <- as.data.frame(do.call(cbind, hb_obj$std_cts))
-
-## Generate all coordinates of hexagons
-hex_grid <- as.data.frame(do.call(cbind, hb_obj$hex_poly))
-
-hex_full_count_df_loop3 <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
-
-p3 <-  ggplot(data = hex_full_count_df_loop3, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
-  geom_point(data = umap_s_curve_scaled, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  coord_equal() +
-  #coord_cartesian(xlim =c(-1, 2.7), ylim = c(-2, 12.2)) +
-  theme_void() +
-  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=8), #change legend title font size
-        legend.text = element_text(size=6)) +
-  guides(fill = guide_colourbar(title = "Standardized count")) +
-  annotate(geom = 'text', label = "c", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
-
-```
-
-```{r}
-#| echo: false
-#| label: fig-binsize
-#| fig-pos: H
-#| fig-cap: "Hexbin plots from different number of bins for the **UMAP** embeddings of **S-curve** training data: (a) b = 12 (3, 4), s = 0.62, (b) b = 84 (6, 14), s = 0.13, and (c) b = 336 (12, 28), s = 0.06.
-The hexbins are colored based on the density of points, with darker colors indicating higher point density and yellow color representing lower point density within each bin.
-What is the number of bins that would be effective in representing low-dimensional data?"
-
-p1 + p2 + p3
-```
-
-$$
-b = b_1 \times b_2
-$${#eq-equation8}
-
-Furthermore, the total number of bins is calculated based on the number of bins along the x-axis and y-axis, as shown in @eq-equation8.
-To determine the effective total number of bins, candidate values are selected based on the range between the minimum and approximate maximum number of bins along the x and y axes.
-The minimum number of bins along each axis is set to $1$, while the maximum number is estimated by taking the square root of the NLDR data points.
-The analysis evaluates the MSE across varying total number of bins within this range, covering the minimum to maximum values along both axes (see @fig-diagnosticpltScurve).
-
-```{r}
-#| warning: false
-#| echo: false
-#| message: false
-
-## UMAP
-
-hex_size_vec <- seq(0.06, 2, by = 0.01)
-
-vec <- stats::setNames(rep("", 6), c("num_bins", "aic", "mse", "num_bins_x", "num_bins_y", "hex_size"))  ## Define column names
-
-mse_df <- dplyr::bind_rows(vec)[0, ]
-mse_df <- mse_df |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(hex_size_vec)) {
-
-  num_bin_list <- calc_bins(data = umap_s_curve_scaled, 
-            x = "UMAP1", y = "UMAP2", 
-            hex_size = hex_size_vec[i], buffer_x = NA, buffer_y = NA)
-
-  num_bins_x <- num_bin_list$num_x
-  num_bins_y <- num_bin_list$num_y
-
-  model_object <- fit_highd_model( training_data = training_data, 
-                                   nldr_df_with_id = umap_s_curve_scaled, 
-                                   x = "UMAP1", y = "UMAP2", 
-                                   num_bins_x = num_bins_x, 
-                                   num_bins_y = num_bins_y, 
-                                   x_start = NA, y_start = NA, 
-                                   buffer_x = NA, buffer_y = NA, 
-                                   hex_size = hex_size_vec[i],
-                                   is_rm_lwd_hex = FALSE, 
-                                   benchmark_to_rm_lwd_hex = NA, 
-                                   col_start_2d = "UMAP", 
-                                   col_start_highd = "x")
-
-  centroid_df_training <- model_object$df_bin_centroids
-  avg_df_training <- model_object$df_bin
-
-  pred_emb_list <- predict_emb(test_data = training_data, 
-                                  df_bin_centroids = centroid_df_training, 
-                                  df_bin = avg_df_training, type_NLDR = "UMAP")
-
-  pred_df_training <- as.data.frame(do.call(cbind, pred_emb_list))
-
-  eval_list <- gen_summary(test_data = training_data, 
-                                  prediction_df = pred_df_training, 
-                                  df_bin = avg_df_training, col_start = "x")
-
-  mse_df <- mse_df |>
-    tibble::add_row(num_bins = num_bins_x * num_bins_y,
-                    aic = eval_list$aic,
-                    mse = eval_list$mse,
-                    num_bins_x = num_bins_x,
-                    num_bins_y = num_bins_y,
-                    hex_size = hex_size_vec[i])
-
-}
-
-## If same total number of bins occurred only select ones with minimum error
-### Obtain duplicate bins
-dupli_bins <- mse_df |> 
-  dplyr::count(num_bins) |> 
-  dplyr::filter(n > 1) |> 
-  dplyr::pull(num_bins)
-
-### Group split by duplicated bins
-duplicate_df_list <- mse_df |>
-  dplyr::filter(num_bins %in% dupli_bins) |>
-  dplyr::arrange(num_bins) |>
-  dplyr::group_split(num_bins)
-
-### Obtain one row from duplicates which have lowest error and hexsize
-duplicate_df <- data.frame(matrix(nrow = 0, ncol = 0))
-
-for (i in 1:length(duplicate_df_list)) {
-
-  dd <- duplicate_df_list[[i]] |>
-    dplyr::filter(mse == min(duplicate_df_list[[i]]$mse)) |>
-    dplyr::filter(hex_size == min(duplicate_df_list[[i]]$hex_size))
-
-  duplicate_df <- dplyr::bind_rows(duplicate_df, dd)
-
-}
-
-### Obtain the mse_df with not duplicated bins
-not_dupli_df <- mse_df |>
-  dplyr::filter(!(num_bins %in% dupli_bins))
-
-### Combine duplicated and not duplicated(corrected) bins dfs
-mse_df <- dplyr::bind_rows(not_dupli_df, duplicate_df)
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-mse_plot <- ggplot(mse_df, aes(x = num_bins,
-                                       y = mse
-)) +
-  geom_point() +
-  geom_line() +
-   geom_vline(xintercept = 84, linetype="solid",
-                color = "red", size=0.8, alpha = 0.5) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  #scale_colour_manual(values = c("#377eb8")) +
-  ylab("MSE") +
-  xlab("total number of bins")
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset.
-What is the effective number of bins in each NLDR technique to create a 2D model?
-The MSE plot have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
-Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error.
-Using the elbow method, when the total number of bins is set to $84$, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
-This point indicates that adding less bins does not enough to capture the data structure."
-#| label: fig-diagnosticpltScurve
-##| out-width: 100%
-#| fig-pos: H
-
-mse_plot
-```
-
-```{r}
-#| warning: false
-#| echo: false
-#| message: false
-
-## Decide by looking at MSE plot
-num_bins_x_s_curve <- 6
-num_bins_y_s_curve <- 14
-hex_size_s_curve <- 0.13
-
-hb_obj_s_curve <- hex_binning(data = umap_s_curve_scaled, x = "UMAP1", 
-                      y = "UMAP2", num_bins_x = num_bins_x_s_curve, 
-                      num_bins_y = num_bins_y_s_curve, x_start = NA, y_start = NA, 
-                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_s_curve, col_start = "UMAP")
-
-## Data set with all possible centroids in the hexagonal grid
-all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_s_curve$centroids))
-
-## Generate all coordinates of hexagons
-hex_grid <- as.data.frame(do.call(cbind, hb_obj_s_curve$hex_poly))
-
-## To obtain the standardise counts within hexbins
-counts_df <- as.data.frame(do.call(cbind, hb_obj_s_curve$std_cts))
-df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
-                                             counts_df = counts_df)
-
-## remove low-density hexagons
-
-hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
-
-hex_full_count_df_loop2 <- hex_grid_with_counts |>
-  dplyr::mutate(type = if_else(std_counts <= 0.2, "low", "high"))
-
-hex_full_count_df_filtered <- df_bin_centroids |>
-  dplyr::mutate(type = if_else(std_counts <= 0.2, "low", "high")) |>
-  dplyr::filter(type == "low")
-
-p2_n <-  ggplot(data = hex_full_count_df_loop2, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
-  #geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  #coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
-  theme_void() +
-  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=8), #change legend title font size
-        legend.text = element_text(size=6)) +
-  guides(fill = guide_colourbar(title = "Standardized count")) #+
-  #annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
-
-p2_n_rm <-  ggplot(data = hex_full_count_df_loop2, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
-  geom_text(data = hex_full_count_df_filtered, color = "red", aes(x = c_x, y = c_y, group = hexID, label = "X", size = 0.5)) +
-  #geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  #coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
-  theme_void() +
-  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=8), #change legend title font size
-        legend.text = element_text(size=6)) +
-  guides(fill = guide_colourbar(title = "Standardized count")) #+
-  #annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
-
-# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
-#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
-#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
-#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
-#   coord_fixed()
-
-umap_data_with_hb_id <- as.data.frame(do.call(cbind, hb_obj_s_curve$data_hb_id))
-
-model_object <- fit_highd_model( training_data = training_data, 
-                                 nldr_df_with_id = umap_s_curve_scaled, 
-                                 x = "UMAP1", y = "UMAP2", 
-                                 num_bins_x = num_bins_x_s_curve, 
-                                 num_bins_y = num_bins_y_s_curve, 
-                                 x_start = NA, y_start = NA, 
-                                 buffer_x = NA, buffer_y = NA, 
-                                 hex_size = hex_size_s_curve,
-                                 is_rm_lwd_hex = FALSE, 
-                                 benchmark_to_rm_lwd_hex = NA, 
-                                 col_start_2d = "UMAP", 
-                                 col_start_highd = "x")
-
-df_bin_centroids_s_curve <- model_object$df_bin_centroids
-df_bin_s_curve <- model_object$df_bin
-
-cell_count_plot <- ggplot(df_bin_centroids_s_curve, aes(x = reorder(as.factor(hexID), -std_counts), y = std_counts)) +
-  geom_quasirandom() + xlab("hexagonal id") + ylab("Standardized cell count") +
-  geom_hline(yintercept = 0.2, colour = "#de2d26") +
-  theme(axis.text = element_text(size = 5),
-        axis.title = element_text(size = 7),
-        axis.text.x = element_text(angle = 90))
-
-benchmark_value_rm_lwd <- stats::quantile(df_bin_centroids_s_curve$std_counts, 
-                probs = c(0,0.25,0.5,0.75,1), names = FALSE)[2]
-
-df_bin_centroids_low <- df_bin_centroids_s_curve |>
-  dplyr::filter(std_counts <= benchmark_value_rm_lwd)
-
-identify_rm_bins <- find_low_dens_hex(df_bin_centroids_all = df_bin_centroids_s_curve, 
-                                      num_bins_x = num_bins_x_s_curve, 
-                                      df_bin_centroids_low = df_bin_centroids_low)
-
-df_bin_centroids_s_curve <- df_bin_centroids_s_curve |>
-  dplyr::filter(!(hexID %in% identify_rm_bins))
-
-pcentroid_plot_s_curve <- ggplot(data = hex_grid, aes(x = x, y = y)) +
-  geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
-  geom_point(data = df_bin_centroids_s_curve, aes(x = c_x, y = c_y), color = "#33a02c") +
-  coord_equal() +
-  theme_linedraw() +
-  theme(legend.position = "bottom", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-rm_centroid_df <- df_bin_centroids_low |> 
-  dplyr::filter(hexID %in% identify_rm_bins)
-
-p2_n_rm_correct <- ggplot(data = hex_full_count_df_loop2, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
-  geom_text(data = hex_full_count_df_filtered[6, ], color = "red", aes(x = c_x, y = c_y, group = hexID, label = "X", size = 0.5)) +
-  #geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  #coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
-  theme_void() +
-  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=8), #change legend title font size
-        legend.text = element_text(size=6)) +
-  guides(fill = guide_colourbar(title = "Standardized count")) #+
-  #annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
-
-## Triangulate bin centroids
-tr1_object_s_curve <- tri_bin_centroids(df_bin_centroids_s_curve, x = "c_x", y = "c_y")
-tr_from_to_df_s_curve <- gen_edges(tri_object = tr1_object_s_curve)
-
-## Compute 2D distances
-distance_s_curve <- cal_2d_dist(tr_coord_df = tr_from_to_df_s_curve, 
-                             start_x = "x_from", start_y = "y_from", 
-                             end_x = "x_to", end_y = "y_to", 
-                             select_vars = c("from", "to", "distance"))
-
-# 
-# distance_plot
-
-## To find the benchmark value
-benchmark_s_curve <- find_lg_benchmark(distance_edges = distance_s_curve, distance_col = "distance")
-
-distance_plot <- plot_dist(distance_s_curve) +
-   geom_hline(yintercept = benchmark_s_curve, linetype="solid",
-                color = "red", size=0.8, alpha = 0.5) +
-  ylab(expression(d^{(2)})) +
-  theme(axis.text = element_text(size = 5),
-        axis.title = element_text(size = 12))
-
-# ggplot() +
-# geom_trimesh(data = df_bin_centroids_pbmc, mapping = aes(x = c_x, y = c_y))
-
-trimesh_s_curve_umap <- vis_lg_mesh(distance_edges = distance_s_curve, benchmark_value = benchmark_s_curve, tr_coord_df = tr_from_to_df_s_curve, distance_col = "distance")
-
-trimesh_s_curve_umap <- trimesh_s_curve_umap +
-  theme_linedraw() +
-  theme(legend.position = "bottom", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) +
-  labs(colour = "")
-
-trimesh_removed_s_curve_umap <- vis_rmlg_mesh(distance_edges = distance_s_curve, benchmark_value = benchmark_s_curve, tr_coord_df = tr_from_to_df_s_curve, distance_col = "distance")
-
-trimesh_removed_s_curve_umap <- trimesh_removed_s_curve_umap +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-# df_all_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), umap_data_with_hb_id)
-# 
-# show_langevitour(df_all_s_curve, df_bin_s_curve, df_bin_centroids_s_curve, benchmark_s_curve, distance_s_curve, "distance", col_start = "x")
-```
-
-#### Benchmark value to remove low-density hexagons
-
-After establishing the hexagonal grid with an appropriate number of bins, some hexagonal bins may have few or no data points within them (see @fig-rmlowdenshex (a)).
-To ensure comprehensive coverage of the NLDR data, it is necessary to select hexagonal bins with a considerable number of data points.
-This involves calculating the number of points within each hexagon.
-Then, the standard count is computed by dividing the number of points within each hexagon by the maximum number of points in the grid (see @eq-equationp2).
-Next, bins with a standard count less than a certain benchmark value are removed (see @fig-rmlowdenshex (b)).
-The following steps will help in determining a suitable value for removing low-density hexagons:
-
-1. Plot the distribution of the standardized counts (see @fig-stdctsScurve).
-2. Examine the distribution of counts.
-3. Select the first quartile value if the distribution is skewed.
-
-$$
-\text{standard count} = \frac{\text{count}}{\text{max count}} 
-$${#eq-equationp2}
-
-Furthermore, selecting the benchmark value for removing low-density hexagons is important.
-Removing unnecessary bins may lead to the formation of long edges and an uneven 2D model.
-Hence, rather than solely relying on the benchmark value to identify hexagons for removal, it's essential to consider the standard number of points in the neighboring hexagons of the identified low-density ones (see @fig-rmlowdenshex (c)).
-If neighboring bins also show low counts, only those bins will be removed.
-The remaining bins are used to construct the 2D model.
-
-The benchmark value for removing low-density hexagons ranges between $0$ and $1$.
-When analyzing how these benchmark values influence model performance, it's essential to observe the change in Mean Squared Error (MSE) as the benchmark value increases (see @fig-diagnosticpltScurvelwd).
-The MSE shows a gradual increase as the benchmark value progresses from $0$ to $1$.
-Evaluating this rate of increase is important.
-If the increment is not considerable, the decision might lean towards retaining low-density hexagons.
-
-```{r}
-#| echo: false
-#| fig-cap: "Distribution of standardize counts by hexagons."
-#| label: fig-stdctsScurve
-#| fig-pos: H
-
-cell_count_plot
-```
-
-```{r}
-#| warning: false
-#| echo: false
-#| message: false
-num_bins_x_s_curve <- 6
-num_bins_y_s_curve <- 14
-hex_size_s_curve <- 0.13
-
-vec <- stats::setNames(rep("", 2), c("benchmark_val_lwd", "mse"))  ## Define column names
-
-mse_df <- dplyr::bind_rows(vec)[0, ]
-mse_df <- mse_df |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-model_object <- fit_highd_model( training_data = training_data, 
-                                 nldr_df_with_id = umap_s_curve_scaled, 
-                                 x = "UMAP1", y = "UMAP2", 
-                                 num_bins_x = num_bins_x_s_curve, 
-                                 num_bins_y = num_bins_y_s_curve, 
-                                 x_start = NA, y_start = NA, 
-                                 buffer_x = NA, buffer_y = NA, 
-                                 hex_size = hex_size_s_curve,
-                                 is_rm_lwd_hex = FALSE, 
-                                 benchmark_to_rm_lwd_hex = NA, 
-                                 col_start_2d = "UMAP", 
-                                 col_start_highd = "x")
-
-centroid_df_training_all <- model_object$df_bin_centroids
-avg_df_training_all <- model_object$df_bin
-
-benchmark_rm_hex_vec <- seq(0, 1, by=0.1)
-benchmark_rm_hex_vec <- append(benchmark_rm_hex_vec[1:10],0.99)
-
-for (i in 1:length(benchmark_rm_hex_vec)) {
-
-  centroid_df_training <- centroid_df_training_all |>
-    dplyr::filter(std_counts > benchmark_rm_hex_vec[i])
-
-  avg_df_training <- avg_df_training_all |>
-    dplyr::filter(hb_id %in% centroid_df_training$hexID)
-
-  pred_emb_list <- predict_emb(test_data = training_data, 
-                               df_bin_centroids = centroid_df_training, 
-                               df_bin = avg_df_training, type_NLDR = "UMAP")
-
-  pred_df_training <- as.data.frame(do.call(cbind, pred_emb_list))
-
-  eval_list <- gen_summary(test_data = training_data, 
-                           prediction_df = pred_df_training, 
-                           df_bin = avg_df_training, col_start = "x")
-
-  mse_df <- mse_df |>
-    tibble::add_row(benchmark_val_lwd = benchmark_rm_hex_vec[i],
-                    mse = eval_list$mse)
-
-}
-
-mse_plot_lwd <- ggplot(mse_df, aes(x = benchmark_val_lwd,
-                                       y = mse
-)) +
-  geom_point() +
-  geom_line() +
-   geom_vline(xintercept = 0.2, linetype="solid",
-                color = "red", size=0.8, alpha = 0.5) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  #scale_colour_manual(values = c("#377eb8")) +
-  ylab("MSE") +
-  xlab("benchmark value to remove low-density hexagons")
-
-```
-
-```{r}
-#| echo: false
-#| warning: false
-#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset with different benchmark values to remove the low-density heaxgons.
-What is the effective benchmark value to remove the low-density heaxgons?
-The MSE plot have a steep slope at the beginning, indicating that a smaller benchmark causes a small amount of error.
-Then, the slope gradually increases or level up, indicating that a higher number of benchmark values generates a higher error.
-Using the reverse elbow method, when the benchmark value is set to 0.2, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
-This point indicates that higher benchmark values remove the necessary bins as well which lead to distruct the 2D structure."
-#| label: fig-diagnosticpltScurvelwd
-##| out-width: 100%
-#| fig-pos: H
-
-mse_plot_lwd
-```
-
-```{r}
-#| echo: false
-#| warning: false
-#| fig-cap: "(a) Hexbin plot with colored hexbins based on point density, (b) Identification of low-density hexagons using a benchmark value of $0.2$, and (c) Identification of low-density hexagons considering neighboring bins."
-#| label: fig-rmlowdenshex
-#| fig-pos: H
-##| out-width: 100%
-
-p2_n + p2_n_rm + p2_n_rm_correct +
-  plot_annotation(tag_levels = 'a') +
-  plot_layout(guides='collect', ncol = 3) &
-  theme(legend.position='none', plot.tag = element_text(size = 8))
-```
-
-#### Benchmark value for removing long edges
-
-Creating a smooth 2D representation (see @fig-modelScurve (c)) requires removing edges that connect distant bin centroids within the triangular mesh (see @fig-modelScurve (b)).
-These edges only exist in the 2D model and do not extend into higher dimensions, ensuring that their removal does not impact the model in higher dimensions.
-While specific criteria for determining the benchmark value to remove long edges do not exist, the following steps provide an approach to identifying a suitable threshold:
-
-1. Plot the distribution of the 2D Euclidean distances (see @fig-distScurve).
-2. Identify the largest difference between consecutive distance values.
-3. Take the largest distance value corresponding to this difference as the benchmark value.
-
-```{r}
-#| echo: false
-#| fig-cap: "Distribution of 2D Euclidean distances between bin centroids of the triangular mesh generated with S-curve UMAP data."
-#| label: fig-distScurve
-#| fig-pos: H
-
-distance_plot
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "(a) Full hexagonal grid with bin centroids, (b) Model constructed in 2D with long edges, and (c) Model constructed in 2D after removing the long edges."
-#| label: fig-modelScurve
-#| fig-pos: H
-
-pcentroid_plot_s_curve + trimesh_s_curve_umap + trimesh_removed_s_curve_umap +
-  plot_layout(guides='collect', ncol = 3) &
-  theme(legend.position='bottom') 
-```
-
-#### Starting point of the hexagonal grid
-
-Hexagonal binning involves tessellating the $xy$ plane over the set defined by the range of $x$ and $y$ [@Dan2023].
-Therefore, the starting point of the hexagonal grid is important parameter to consider.
-Having different starting point results in shifting the hexagonal grid and bin centroids (see @fig-scurveshifthexgridsexp).
-With a specific total number of bins (which corresponds to a specific range along the x and y axes), it's essential to explore how different starting points affect the model's performance (see @fig-mseScurvestartpoint).
-This is because different starting points result in varying distributions of non-empty bins (see @fig-scurveshifthexgrids), impacting the model.
-
-```{r}
-#| warning: false
-#| echo: false
-#| message: false
-
-## UMAP
-## Decide by looking at MSE plot
-num_bins_x_s_curve <- 6
-num_bins_y_s_curve <- 14
-hex_size_s_curve <- 0.13
-
-## Possible values for x_start
-max_x_start <- min(umap_s_curve_scaled[["UMAP1"]]) + (sqrt(3) * hex_size_s_curve)
-min_x_start <- min(umap_s_curve_scaled[["UMAP1"]]) - (sqrt(3) * hex_size_s_curve)
-
-## Possible values for y_start
-max_y_start <- min(umap_s_curve_scaled[["UMAP2"]]) + (1.5 * hex_size_s_curve)
-min_y_start <- min(umap_s_curve_scaled[["UMAP2"]]) - (1.5 * hex_size_s_curve)
-
-diff_start_df <- expand.grid(x_start = seq(min_x_start, max_x_start, length.out = 10), 
-                             y_start = seq(min_y_start, max_y_start, length.out = 10)) |>
-  add_row(x_start = min(umap_s_curve_scaled[["UMAP1"]]) - (sqrt(3) * hex_size_s_curve/2),
-          y_start = min(umap_s_curve_scaled[["UMAP2"]]) - (1.5 * hex_size_s_curve/2))
-
-vec <- stats::setNames(rep("", 4), c("x_start", "y_start", "num_non_empty_bins", "mse"))  ## Define column names
-
-mse_df <- dplyr::bind_rows(vec)[0, ]
-mse_df <- mse_df |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-diff_start_df <- diff_start_df[c(1:5, 11:15,21:25, 101),]
-
-for (i in 1:NROW(diff_start_df)) { #NROW(diff_start_df)
-
-  model_object <- fit_highd_model( training_data = training_data, 
-                                   nldr_df_with_id = umap_s_curve_scaled, 
-                                   x = "UMAP1", y = "UMAP2", 
-                                   num_bins_x = num_bins_x_s_curve, 
-                                   num_bins_y = num_bins_y_s_curve, 
-                                   x_start = diff_start_df[i, 1], y_start = diff_start_df[i, 2], 
-                                   buffer_x = NA, buffer_y = NA, 
-                                   hex_size = hex_size_s_curve,
-                                   is_rm_lwd_hex = FALSE, 
-                                   benchmark_to_rm_lwd_hex = NA, 
-                                   col_start_2d = "UMAP", 
-                                   col_start_highd = "x")
-
-  centroid_df_training <- model_object$df_bin_centroids
-  avg_df_training <- model_object$df_bin
-
-  pred_emb_list <- predict_emb(test_data = training_data, 
-                               df_bin_centroids = centroid_df_training, 
-                               df_bin = avg_df_training, type_NLDR = "UMAP")
-
-  pred_df_training <- as.data.frame(do.call(cbind, pred_emb_list))
-
-  eval_list <- gen_summary(test_data = training_data, 
-                           prediction_df = pred_df_training, 
-                           df_bin = avg_df_training, col_start = "x")
-
-  mse_df <- mse_df |>
-    tibble::add_row(x_start = diff_start_df[i, 1], 
-                    y_start = diff_start_df[i, 2],
-                    num_non_empty_bins = NROW(centroid_df_training), 
-                    mse = eval_list$mse)
-
-}
-
-# ## If same total number of bins occurred only select ones with minimum error
-# ### Obtain duplicate bins
-# dupli_bins <- mse_df |> 
-#   dplyr::count(num_non_empty_bins) |> 
-#   dplyr::filter(n > 1) |> 
-#   dplyr::pull(num_non_empty_bins)
-# 
-# ### Group split by duplicated bins
-# duplicate_df_list <- mse_df |>
-#   dplyr::filter(num_non_empty_bins %in% dupli_bins) |>
-#   dplyr::arrange(num_non_empty_bins) |>
-#   dplyr::group_split(num_non_empty_bins)
-# 
-# ### Obtain one row from duplicates which have lowest error and hexsize
-# duplicate_df <- data.frame(matrix(nrow = 0, ncol = 0))
-# 
-# for (i in 1:length(duplicate_df_list)) {
-#   
-#   dd <- duplicate_df_list[[i]] |>
-#     dplyr::filter(mse == min(duplicate_df_list[[i]]$mse)) 
-#   
-#   duplicate_df <- dplyr::bind_rows(duplicate_df, dd)
-#   
-# }
-# 
-# ### Obtain the mse_df with not duplicated bins
-# not_dupli_df <- mse_df |>
-#   dplyr::filter(!(num_non_empty_bins %in% dupli_bins))
-# 
-# ### Combine duplicated and not duplicated(corrected) bins dfs
-# mse_df <- dplyr::bind_rows(not_dupli_df, duplicate_df)
-
-mse_plot_start_point <- ggplot(mse_df, aes(x = x_start,
-                               y = y_start, colour = mse
-)) +
-  geom_point() +
-  #geom_line() +
-  # geom_vline(xintercept = 84, linetype="solid",
-  #            color = "red", linewidth=0.8, alpha = 0.5) +
-  theme_light() +
-  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  viridis::scale_color_viridis(option = "D") +
-  labs(x = "x start", y = "y start", color = "MSE") 
-```
-
-```{r}
-#| echo: false
-#| warning: false
-#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset with different benchmark values to remove the low-density heaxgons.
-What is the effective benchmark value to remove the low-density heaxgons?
-The MSE plot have a steep slope at the beginning, indicating that a smaller benchmark causes a small amount of error.
-Then, the slope gradually increases or level up, indicating that a higher number of benchmark values generates a higher error.
-Using the reverse elbow method, when the benchmark value is set to 0.2, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
-This point indicates that higher benchmark values remove the necessary bins as well which lead to distruct the 2D structure."
-#| label: fig-mseScurvestartpoint
-#| fig-pos: H
-
-mse_plot_start_point
-```
-
-```{r}
-#| echo: false
-#| warning: false
-
-## Initial x and y start
-initial_start_plot <- ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
-  geom_text(data = df_bin_centroids, aes(x = c_x, y = c_y, label = hexID), size = 2) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  coord_equal() +
-  theme_void() +
-  theme(legend.position="bottom", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=8), #change legend title font size
-        legend.text = element_text(size=6)) +
-  guides(fill = guide_colourbar(title = "Standardized count")) +
-  annotate(geom = 'text', label = "a", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
-
-## x and y start selected by MSE
-
-## UMAP
-## Decide by looking at MSE plot
-num_bins_x_s_curve <- 6
-num_bins_y_s_curve <- 14
-hex_size_s_curve <- 0.13
-
-hb_obj_s_curve_n <- hex_binning(data = umap_s_curve_scaled, x = "UMAP1", 
-                      y = "UMAP2", num_bins_x = num_bins_x_s_curve, 
-                      num_bins_y = num_bins_y_s_curve, x_start = -0.0652, y_start = -0.195, 
-                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_s_curve, col_start = "UMAP")
-
-## Data set with all possible centroids in the hexagonal grid
-all_centroids_df_n <- as.data.frame(do.call(cbind, hb_obj_s_curve_n$centroids))
-
-## Generate all coordinates of hexagons
-hex_grid_n <- as.data.frame(do.call(cbind, hb_obj_s_curve_n$hex_poly))
-
-## To obtain the standardise counts within hexbins
-counts_df_n <- as.data.frame(do.call(cbind, hb_obj_s_curve_n$std_cts))
-df_bin_centroids_n <- extract_hexbin_centroids(centroids_df = all_centroids_df_n, 
-                                             counts_df = counts_df_n)
-
-## remove low-density hexagons
-
-hex_grid_with_counts_n <- dplyr::left_join(hex_grid_n, counts_df_n, by = c("hex_poly_id" = "hb_id"))
-
-new_start_plot <- ggplot(data = hex_grid_with_counts_n, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
-  geom_text(data= df_bin_centroids_n, aes(x = c_x, y = c_y, label = hexID), size = 2) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  coord_equal() +
-  theme_void() +
-  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=8), #change legend title font size
-        legend.text = element_text(size=6)) +
-  guides(fill = guide_colourbar(title = "Standardized count")) +
-  annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
-
-hex_grid_ch_plot <- ggplot(data = hex_grid_with_counts_n, aes(x = x, y = y)) +
-  geom_polygon(fill = NA, color = "#feb24c", aes(group = hex_poly_id)) +
-  geom_polygon(data = hex_grid_with_counts, aes(x = x, y = y, group = hex_poly_id),
-               fill = NA, color = "#bdbdbd") +
-  #coord_cartesian(xlim = c(-5, 8), ylim = c(-10, 10)) +
-  theme_void() +
-  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=8), #change legend title font size
-        legend.text = element_text(size=6)) +
-  guides(fill = guide_colourbar(title = "Standardized count")) +
-  annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)
-
-bin_centroids_ch_plot <- ggplot(data = all_centroids_df, aes(x = c_x, y = c_y)) +
-  geom_point(color = "#bdbdbd") +
-  geom_point(data = all_centroids_df_n, aes(x = c_x, y = c_y), color = "#feb24c") +
-  #coord_cartesian(xlim = c(-5, 8), ylim = c(-10, 10)) +
-  theme_void() +
-  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=8), #change legend title font size
-        legend.text = element_text(size=6)) +
-  guides(fill = guide_colourbar(title = "Standardized count")) +
-  annotate(geom = 'text', label = "a", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3) 
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "(a) Visualization of bin centroids and (b) hexagonal grid, showcasing the configuration before (colored in gray) and after (colored in light orange) shifting the starting point by an identical amount applied in both x and y directions, with a shift value of $0.537285$."
-#| label: fig-scurveshifthexgridsexp
-#| fig-pos: H
-
-bin_centroids_ch_plot + hex_grid_ch_plot +
-  plot_layout(guides='collect', ncol = 2) &
-  theme(legend.position='none', plot.tag = element_text(size = 8))
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "Hexbin plots shows the distribution of points before and after shifting the starting point of the hexagonal grid generated for UMAP applied to the training S-curve dataset.
-(a) Hexbin plot before shifting, and (b) Hexbin plot after shifting (identical shift applied in both x and y directions, with a shift value of $0.537285$)."
-#| label: fig-scurveshifthexgrids
-#| fig-pos: H
-
-initial_start_plot + new_start_plot +
-  plot_layout(guides='collect', ncol = 2) &
-  theme(legend.position='none', plot.tag = element_text(size = 8))
-```
-
-### Model summaries {#sec-summary}
-
-#### Predicted values and residuals
-
-The prediction approach involves performing the K-nearest neighbors (KNN) algorithm for an unsupervised classification problem.
-First, the nearest high-D model point is identified for a given new high-D point.
-Then, the corresponding 2D centroid mapping for the identified high-D model point is determined.
-Finally, the coordinates of this 2D centroid are used as the predicted 2D embedding for the new high-D data point.
-This step is particularly valuable due to the limitations of some NLDR techniques, like tSNE, which don't provide a straightforward method for prediction.
-As a result, our approach offers a solution that capable of generating predicted 2D embedding regardless of the NLDR technique employed, effectively addressing this functional gap.
-
-Residuals are essential for evaluating the accuracy of representing high-D points by the high-D mapping of 2D bin centroids.
-To measure this accuracy, an error metric is introduced, quantifying the sum of squared differences between the high-D data ($x_{ij}$) and the high-D mapping of the 2D bin centroid data ($C_{x_ij}$) across all observations and dimensions (see @eq-equation10).
-
-$$
-\text{Error} = \sum_{j = 1}^{n}\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2
-$$ {#eq-equation10}
-
-Here, $n$ represents the number of observations, $p$ represents the dimensions of high-D data, $x_{ij}$ is the high-D data, and $C_{x_ij}$ is the high-D mapping of the 2D bin centroid.
-
-#### Goodness of fit statistics {#sec-goodfit}
-
-To assess how well our method captures and represents the underlying structure of the high-D data, Mean Squared Error (MSE) is used.
-When computing MSE, total model error (see @sec-summary) is divided by the number of observations to make it as a mean value (see @eq-equation9).
-
-$$
-\text{MSE} = \sum_{j = 1}^{n} \frac{\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2}{n}
-$$ {#eq-equation9}
-
-### Simulated data example {#sec-simpleex}
-
-In this section, the effectiveness of the algorithm is described using a simulated dataset.
-The dataset consists of five spherical Gaussian clusters in 4-$d$, with each cluster containing an equal number of points and the same within-cluster variation.
-
-In the 2D layouts generated by various NLDR techniques, as shown in @fig-nldervis5Gau, five distinct clusters are observable, except for PHATE.
-In tSNE (see @fig-nldervis5Gau (a)), these clusters appear closely.
-UMAP arranges all clusters in a parallel manner, with three aligned in one line and the other two in a separate line (see @fig-nldervis5Gau (b)).
-In contrast, PHATE shows two closely positioned clusters and three more distant ones (see @fig-nldervis5Gau (c)).
-PaCMAP shows one central cluster and the remaining four spread out in different directions (see @fig-nldervis5Gau (e)).
-Finally, in TriMAP, two clusters are close, though not as tightly as PHATE, while the other three are well-separated (see @fig-nldervis5Gau (d)).
-
-<!-- five Gaussian clusters with different NLDR techniques-->
-```{r}
-#| warning: false
-#| echo: false
-
-## Import data
-df_2 <- read_rds("data/five_gau_clusters/data_five_gau.rds")
-training_data_gau <- read_rds("data/five_gau_clusters/data_five_gau_training.rds")
-test_data_gau <- read_rds("data/five_gau_clusters/data_five_gau_test.rds")
-
-tSNE_data_gau <- read_rds("data/five_gau_clusters/tsne_data_five_gau_61.rds")
-UMAP_data_gau <- read_rds("data/five_gau_clusters/umap_data_five_gau.rds")
-PHATE_data_gau <- read_rds("data/five_gau_clusters/phate_data_five_gau.rds")
-TriMAP_data_gau <- read_rds("data/five_gau_clusters/trimap_data_five_gau.rds")
-PaCMAP_data_gau <- read_rds("data/five_gau_clusters/pacmap_data_five_gau.rds")
-
-## Visualise embeddings
-
-plot_list1_gau <- plot_tSNE_2D(tSNE_data_gau) + 
-  geom_point(size = 0.00001, colour = "#e41a1c") +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-plot_list2_gau <- plot_UMAP_2D(UMAP_data_gau) + 
-  geom_point(size = 0.00001, colour = "#377eb8") +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-plot_list3_gau <- plot_PHATE_2D(PHATE_data_gau) + 
-  geom_point(size = 0.00001, colour = "#4daf4a") +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'c', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-plot_list4_gau <- plot_TriMAP_2D(TriMAP_data_gau) + 
-  geom_point(size = 0.00001, colour = "#984ea3") +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-plot_list5_gau <- plot_PaCMAP_2D(PaCMAP_data_gau) + 
-  geom_point(size = 0.00001, colour = "#ff7f00") +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'e', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "2D layouts from different NLDR techniques applied the same data: (a) tSNE (perplexity = 61), (b) UMAP (n_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2).
-Is there a best representation of the original data or are they all providing  equivalent information?"
-#| label: fig-nldervis5Gau
-#| fig-pos: H
-#| out-width: 100%
-
-plot_list1_gau + plot_list2_gau + plot_list3_gau + plot_list4_gau + plot_list5_gau +
-  plot_layout(ncol=5)
-```
-
-To investigate which is the best representation to visualize the original data or all NLDR methods provide equivalent information, we visualize the model-in-data space.
-Models from all NLDR methods show five well-separated clusters (see @fig-gau1_sc, @fig-gau2_sc, @fig-gau3_sc, @fig-gau4_sc, and @fig-gau5_sc).
-This suggests that for the five Gaussian cluster dataset, all NLDR methods effectively preserve the global structure.
-tSNE, UMAP, and PHATE display clusters with varying densities, indicating their ability to capture within-cluster variation (see @fig-gau1_sc, @fig-gau2_sc, and @fig-gau3_sc).
-On the other hand, both TriMAP and PaCMAP show clusters with flat surfaces, suggesting a failure to capture within-cluster variation (see @fig-gau4_sc and @fig-gau5_sc).
-Therefore, TriMAP and PaCMAP do not capture the local structure as effectively as other methods.
-
-```{r}
-#| warning: false
-#| echo: false
-#| message: false
-
-tsne_gau_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tSNE_data_gau, 
-                                    x = "tSNE1", y = "tSNE2"))) |>
-  dplyr::rename(c("tSNE1" = "scaled_tSNE1", 
-                  "tSNE2" = "scaled_tSNE2")) |>
-  dplyr::mutate(ID = 1:NROW(tSNE_data_gau))
-
-## Decide by looking at MSE plot
-num_bins_x_tsne_gau <- 21
-num_bins_y_tsne_gau <- 28
-hex_size_tsne_gau <- 0.03
-## non-empty:198
-
-hb_obj_tsne_gau <- hex_binning(data = tsne_gau_scaled, x = "tSNE1", 
-                      y = "tSNE2", num_bins_x = num_bins_x_tsne_gau, 
-                      num_bins_y = num_bins_y_tsne_gau, x_start = NA, y_start = NA, 
-                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_tsne_gau, col_start = "tSNE")
-
-## Data set with all possible centroids in the hexagonal grid
-all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$centroids))
-
-## Generate all coordinates of hexagons
-hex_grid <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$hex_poly))
-
-## To obtain the standardise counts within hexbins
-counts_df <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$std_cts))
-df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
-                                             counts_df = counts_df)
-# ggplot(data = hex_grid, aes(x = x, y = y)) +
-#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
-#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
-#   coord_fixed()
-
-hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
-
-# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
-#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
-#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
-#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
-#   coord_fixed()
-
-tsne_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$data_hb_id))
-
-model_object_tsne_gau <- fit_highd_model( training_data = training_data_gau, 
-                                 nldr_df_with_id = tsne_gau_scaled, 
-                                 x = "tSNE1", y = "tSNE2", 
-                                 num_bins_x = num_bins_x_tsne_gau, 
-                                 num_bins_y = num_bins_y_tsne_gau, 
-                                 x_start = NA, y_start = NA, 
-                                 buffer_x = NA, buffer_y = NA, 
-                                 hex_size = hex_size_tsne_gau,
-                                 is_rm_lwd_hex = FALSE, 
-                                 benchmark_to_rm_lwd_hex = NA, 
-                                 col_start_2d = "tSNE", 
-                                 col_start_highd = "x")
-
-df_bin_centroids_tsne_gau <- model_object_tsne_gau$df_bin_centroids
-df_bin_tsne_gau <- model_object_tsne_gau$df_bin
-
-## Triangulate bin centroids
-tr1_object_tsne_gau <- tri_bin_centroids(df_bin_centroids_tsne_gau, x = "c_x", y = "c_y")
-tr_from_to_df_tsne_gau <- gen_edges(tri_object = tr1_object_tsne_gau)
-
-## Compute 2D distances
-distance_tsne_gau <- cal_2d_dist(tr_coord_df = tr_from_to_df_tsne_gau, 
-                             start_x = "x_from", start_y = "y_from", 
-                             end_x = "x_to", end_y = "y_to", 
-                             select_vars = c("from", "to", "distance"))
-
-# distance_plot <- plot_dist(distance_tsne_gau) +
-#   ylab(expression(d^{(2)})) +
-#   theme(axis.text = element_text(size = 5),
-#         axis.title = element_text(size = 12))
-# 
-# distance_plot
-
-## To find the benchmark value
-benchmark_tsne_gau <- find_lg_benchmark(distance_edges = distance_tsne_gau, distance_col = "distance")
-
-# ggplot() +
-# geom_trimesh(data = df_bin_centroids_tsne_gau, mapping = aes(x = c_x, y = c_y))
-
-trimesh_removed_tsne_gau <- vis_rmlg_mesh(distance_edges = distance_tsne_gau, benchmark_value = benchmark_tsne_gau, tr_coord_df = tr_from_to_df_tsne_gau, distance_col = "distance")
-
-trimesh_removed_tsne_gau <- trimesh_removed_tsne_gau +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-df_all_tsne_gau <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID), tsne_data_with_hb_id_gau)
-
-# show_langevitour(df = df_all_tsne_gau, df_b = df_bin_tsne_gau, df_b_with_center_data = df_bin_centroids_tsne_gau, col_start = "x", distance_df = distance_tsne_gau, distance_col = "distance", benchmark_value = benchmark_tsne_gau)
-```
-
-```{r}
-#| warning: false
-#| echo: false
-#| message: false
-
-umap_gau_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = UMAP_data_gau, 
-                                    x = "UMAP1", y = "UMAP2"))) |>
-  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
-                  "UMAP2" = "scaled_UMAP2")) |>
-  dplyr::mutate(ID = 1:NROW(UMAP_data_gau))
-
-## Decide by looking at MSE plot
-num_bins_x_umap_gau <- 60
-num_bins_y_umap_gau <- 79
-hex_size_umap_gau <- 0.01
-### non-empty 229
-
-hb_obj_umap_gau <- hex_binning(data = umap_gau_scaled, x = "UMAP1", 
-                               y = "UMAP2", num_bins_x = num_bins_x_umap_gau, 
-                               num_bins_y = num_bins_y_umap_gau, x_start = NA, y_start = NA, 
-                               buffer_x = NA, buffer_y = NA, hex_size = hex_size_umap_gau, col_start = "UMAP")
-
-## Data set with all possible centroids in the hexagonal grid
-all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_umap_gau$centroids))
-
-## Generate all coordinates of hexagons
-hex_grid <- as.data.frame(do.call(cbind, hb_obj_umap_gau$hex_poly))
-
-## To obtain the standardise counts within hexbins
-counts_df <- as.data.frame(do.call(cbind, hb_obj_umap_gau$std_cts))
-df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
-                                             counts_df = counts_df)
-# ggplot(data = hex_grid, aes(x = x, y = y)) +
-#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
-#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
-#   coord_fixed()
-
-hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
-
-# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
-#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
-#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
-#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
-#   coord_fixed()
-
-umap_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_umap_gau$data_hb_id))
-
-model_object_umap_gau <- fit_highd_model( training_data = training_data_gau, 
-                                          nldr_df_with_id = umap_gau_scaled, 
-                                          x = "UMAP1", y = "UMAP2", 
-                                          num_bins_x = num_bins_x_umap_gau, 
-                                          num_bins_y = num_bins_y_umap_gau, 
-                                          x_start = NA, y_start = NA, 
-                                          buffer_x = NA, buffer_y = NA, 
-                                          hex_size = hex_size_umap_gau,
-                                          is_rm_lwd_hex = FALSE, 
-                                          benchmark_to_rm_lwd_hex = NA, 
-                                          col_start_2d = "UMAP", 
-                                          col_start_highd = "x")
-
-df_bin_centroids_umap_gau <- model_object_umap_gau$df_bin_centroids
-df_bin_umap_gau <- model_object_umap_gau$df_bin
-
-## Triangulate bin centroids
-tr1_object_umap_gau <- tri_bin_centroids(df_bin_centroids_umap_gau, x = "c_x", y = "c_y")
-tr_from_to_df_umap_gau <- gen_edges(tri_object = tr1_object_umap_gau)
-
-## Compute 2D distances
-distance_umap_gau <- cal_2d_dist(tr_coord_df = tr_from_to_df_umap_gau, 
-                                 start_x = "x_from", start_y = "y_from", 
-                                 end_x = "x_to", end_y = "y_to", 
-                                 select_vars = c("from", "to", "distance"))
-
-# distance_plot <- plot_dist(distance_umap_gau) +
-#   ylab(expression(d^{(2)})) +
-#   theme(axis.text = element_text(size = 5),
-#         axis.title = element_text(size = 12))
-# 
-# distance_plot
-
-## To find the benchmark value
-benchmark_umap_gau <- find_lg_benchmark(distance_edges = distance_umap_gau, distance_col = "distance")
-
-# ggplot() +
-# geom_trimesh(data = df_bin_centroids_umap_gau, mapping = aes(x = c_x, y = c_y))
-
-trimesh_removed_umap_gau <- vis_rmlg_mesh(distance_edges = distance_umap_gau, benchmark_value = benchmark_umap_gau, tr_coord_df = tr_from_to_df_umap_gau, distance_col = "distance")
-
-trimesh_removed_umap_gau <- trimesh_removed_umap_gau +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-df_all_umap_gau <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID), umap_data_with_hb_id_gau)
-
-# show_langevitour(df = df_all_umap_gau, df_b = df_bin_umap_gau, df_b_with_center_data = df_bin_centroids_umap_gau, col_start = "x", distance_df = distance_umap_gau, distance_col = "distance", benchmark_value = benchmark_umap_gau)
-```
-
-```{r}
-#| warning: false
-#| echo: false
-#| message: false
-
-# phate_gau_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tibble::as_tibble(PHATE_data_gau), x = "PHATE1", y = "PHATE2", hex_ratio = 50/sqrt(3)))) |>
-#   dplyr::rename(c("PHATE1" = "scaled_PHATE1",
-#                   "PHATE2" = "scaled_PHATE2")) |>
-#   dplyr::mutate(ID = 1:NROW(PHATE_data_gau))
-# 
-# ## Decide by looking at MSE plot
-# num_bins_x_phate_gau <- 60
-# num_bins_y_phate_gau <- 1927
-# hex_size_phate_gau <- 0.01
-# ### non-empty 187
-# 
-# hb_obj_phate_gau <- hex_binning(data = phate_gau_scaled, x = "PHATE1", 
-#                                y = "PHATE2", num_bins_x = num_bins_x_phate_gau, 
-#                                num_bins_y = num_bins_y_phate_gau, x_start = NA, y_start = NA, 
-#                                buffer_x = NA, buffer_y = NA, hex_size = hex_size_phate_gau, col_start = "PHATE")
-# 
-# ## Data set with all possible centroids in the hexagonal grid
-# all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_phate_gau$centroids))
-# 
-# ## Generate all coordinates of hexagons
-# hex_grid <- as.data.frame(do.call(cbind, hb_obj_phate_gau$hex_poly))
-# 
-# ## To obtain the standardise counts within hexbins
-# counts_df <- as.data.frame(do.call(cbind, hb_obj_phate_gau$std_cts))
-# df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
-#                                              counts_df = counts_df)
-# # ggplot(data = hex_grid, aes(x = x, y = y)) +
-# #   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
-# #   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red", size = 0.1) +
-# #   coord_fixed()
-# 
-# hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
-# 
-# # ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
-# #   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
-# #   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
-# #   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
-# #   coord_fixed()
-# 
-# phate_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_phate_gau$data_hb_id))
-# 
-# model_object_phate_gau <- fit_highd_model( training_data = training_data_gau, 
-#                                           nldr_df_with_id = phate_gau_scaled, 
-#                                           x = "PHATE1", y = "PHATE2", 
-#                                           num_bins_x = num_bins_x_phate_gau, 
-#                                           num_bins_y = num_bins_y_phate_gau, 
-#                                           x_start = NA, y_start = NA, 
-#                                           buffer_x = NA, buffer_y = NA, 
-#                                           hex_size = hex_size_phate_gau,
-#                                           is_rm_lwd_hex = FALSE, 
-#                                           benchmark_to_rm_lwd_hex = NA, 
-#                                           col_start_2d = "PHATE", 
-#                                           col_start_highd = "x")
-# 
-# df_bin_centroids_phate_gau <- model_object_phate_gau$df_bin_centroids
-# df_bin_phate_gau <- model_object_phate_gau$df_bin
-
-df_bin_centroids_phate_gau <- read_rds("data/five_gau_clusters/df_bin_centroids_phate_gau.rds")
-df_bin_phate_gau <- read_rds("data/five_gau_clusters/df_bin_phate_gau.rds")
-
-## Triangulate bin centroids
-tr1_object_phate_gau <- tri_bin_centroids(df_bin_centroids_phate_gau, x = "c_x", y = "c_y")
-tr_from_to_df_phate_gau <- gen_edges(tri_object = tr1_object_phate_gau)
-
-## Compute 2D distances
-distance_phate_gau <- cal_2d_dist(tr_coord_df = tr_from_to_df_phate_gau, 
-                                 start_x = "x_from", start_y = "y_from", 
-                                 end_x = "x_to", end_y = "y_to", 
-                                 select_vars = c("from", "to", "distance"))
-
-# distance_plot <- plot_dist(distance_phate_gau) +
-#   ylab(expression(d^{(2)})) +
-#   theme(axis.text = element_text(size = 5),
-#         axis.title = element_text(size = 12))
-# 
-# distance_plot
-
-## To find the benchmark value
-benchmark_phate_gau <- find_lg_benchmark(distance_edges = distance_phate_gau, distance_col = "distance")
-
-# ggplot() +
-# geom_trimesh(data = df_bin_centroids_phate_gau, mapping = aes(x = c_x, y = c_y))
-
-trimesh_removed_phate_gau <- vis_rmlg_mesh(distance_edges = distance_phate_gau, benchmark_value = benchmark_phate_gau, tr_coord_df = tr_from_to_df_phate_gau, distance_col = "distance")
-
-trimesh_removed_phate_gau <- trimesh_removed_phate_gau +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'c', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-# df_all_phate_gau <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID), phate_data_with_hb_id_gau)
-
-# show_langevitour(df = df_all_phate_gau, df_b = df_bin_phate_gau, df_b_with_center_data = df_bin_centroids_phate_gau, col_start = "x", distance_df = distance_phate_gau, distance_col = "distance", benchmark_value = benchmark_phate_gau)
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-#| message: false
-
-trimap_gau_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = TriMAP_data_gau, 
-                                    x = "TriMAP1", y = "TriMAP2"))) |>
-  dplyr::rename(c("TriMAP1" = "scaled_TriMAP1", 
-                  "TriMAP2" = "scaled_TriMAP2")) |>
-  dplyr::mutate(ID = 1:NROW(TriMAP_data_gau))
-
-## Decide by looking at MSE plot
-num_bins_x_trimap_gau <- 46
-num_bins_y_trimap_gau <- 61
-hex_size_trimap_gau <- 0.013
-### non-empty 173
-
-hb_obj_trimap_gau <- hex_binning(data = trimap_gau_scaled, x = "TriMAP1", 
-                               y = "TriMAP2", num_bins_x = num_bins_x_trimap_gau, 
-                               num_bins_y = num_bins_y_trimap_gau, x_start = NA, y_start = NA, 
-                               buffer_x = NA, buffer_y = NA, hex_size = hex_size_trimap_gau, col_start = "TriMAP")
-
-## Data set with all possible centroids in the hexagonal grid
-all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_trimap_gau$centroids))
-
-## Generate all coordinates of hexagons
-hex_grid <- as.data.frame(do.call(cbind, hb_obj_trimap_gau$hex_poly))
-
-## To obtain the standardise counts within hexbins
-counts_df <- as.data.frame(do.call(cbind, hb_obj_trimap_gau$std_cts))
-df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
-                                             counts_df = counts_df)
-# ggplot(data = hex_grid, aes(x = x, y = y)) +
-#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
-#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
-#   coord_fixed()
-
-hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
-
-# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
-#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
-#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
-#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
-#   coord_fixed()
-
-trimap_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_trimap_gau$data_hb_id))
-
-model_object_trimap_gau <- fit_highd_model( training_data = training_data_gau, 
-                                          nldr_df_with_id = trimap_gau_scaled, 
-                                          x = "TriMAP1", y = "TriMAP2", 
-                                          num_bins_x = num_bins_x_trimap_gau, 
-                                          num_bins_y = num_bins_y_trimap_gau, 
-                                          x_start = NA, y_start = NA, 
-                                          buffer_x = NA, buffer_y = NA, 
-                                          hex_size = hex_size_trimap_gau,
-                                          is_rm_lwd_hex = FALSE, 
-                                          benchmark_to_rm_lwd_hex = NA, 
-                                          col_start_2d = "TriMAP", 
-                                          col_start_highd = "x")
-
-df_bin_centroids_trimap_gau <- model_object_trimap_gau$df_bin_centroids
-df_bin_trimap_gau <- model_object_trimap_gau$df_bin
-
-## Triangulate bin centroids
-tr1_object_trimap_gau <- tri_bin_centroids(df_bin_centroids_trimap_gau, x = "c_x", y = "c_y")
-tr_from_to_df_trimap_gau <- gen_edges(tri_object = tr1_object_trimap_gau)
-
-## Compute 2D distances
-distance_trimap_gau <- cal_2d_dist(tr_coord_df = tr_from_to_df_trimap_gau, 
-                                 start_x = "x_from", start_y = "y_from", 
-                                 end_x = "x_to", end_y = "y_to", 
-                                 select_vars = c("from", "to", "distance"))
-
-# distance_plot <- plot_dist(distance_trimap_gau) +
-#   ylab(expression(d^{(2)})) +
-#   theme(axis.text = element_text(size = 5),
-#         axis.title = element_text(size = 12))
-# 
-# distance_plot
-
-## To find the benchmark value
-benchmark_trimap_gau <- find_lg_benchmark(distance_edges = distance_trimap_gau, distance_col = "distance")
-benchmark_trimap_gau <- 0.2
-
-# ggplot() +
-# geom_trimesh(data = df_bin_centroids_trimap_gau, mapping = aes(x = c_x, y = c_y))
-
-trimesh_removed_trimap_gau <- vis_rmlg_mesh(distance_edges = distance_trimap_gau, benchmark_value = benchmark_trimap_gau, tr_coord_df = tr_from_to_df_trimap_gau, distance_col = "distance")
-
-trimesh_removed_trimap_gau <- trimesh_removed_trimap_gau +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'd', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-df_all_trimap_gau <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID), trimap_data_with_hb_id_gau)
-
-# show_langevitour(df = df_all_trimap_gau, df_b = df_bin_trimap_gau, df_b_with_center_data = df_bin_centroids_trimap_gau, col_start = "x", distance_df = distance_trimap_gau, distance_col = "distance", benchmark_value = benchmark_trimap_gau)
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-#| message: false
-
-pacmap_gau_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = PaCMAP_data_gau, 
-                                    x = "PaCMAP1", y = "PaCMAP2"))) |>
-  dplyr::rename(c("PaCMAP1" = "scaled_PaCMAP1", 
-                  "PaCMAP2" = "scaled_PaCMAP2")) |>
-  dplyr::mutate(ID = 1:NROW(PaCMAP_data_gau))
-
-## Decide by looking at MSE plot
-num_bins_x_pacmap_gau <- 31
-num_bins_y_pacmap_gau <- 40
-hex_size_pacmap_gau <- 0.02
-### non-empty 179
-
-hb_obj_pacmap_gau <- hex_binning(data = pacmap_gau_scaled, x = "PaCMAP1", 
-                               y = "PaCMAP2", num_bins_x = num_bins_x_pacmap_gau, 
-                               num_bins_y = num_bins_y_pacmap_gau, x_start = NA, y_start = NA, 
-                               buffer_x = NA, buffer_y = NA, hex_size = hex_size_pacmap_gau, col_start = "PaCMAP")
-
-## Data set with all possible centroids in the hexagonal grid
-all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_pacmap_gau$centroids))
-
-## Generate all coordinates of hexagons
-hex_grid <- as.data.frame(do.call(cbind, hb_obj_pacmap_gau$hex_poly))
-
-## To obtain the standardise counts within hexbins
-counts_df <- as.data.frame(do.call(cbind, hb_obj_pacmap_gau$std_cts))
-df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
-                                             counts_df = counts_df)
-# ggplot(data = hex_grid, aes(x = x, y = y)) +
-#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
-#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
-#   coord_fixed()
-
-hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
-
-# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
-#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
-#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
-#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
-#   coord_fixed()
-
-pacmap_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_pacmap_gau$data_hb_id))
-
-model_object_pacmap_gau <- fit_highd_model( training_data = training_data_gau, 
-                                          nldr_df_with_id = pacmap_gau_scaled, 
-                                          x = "PaCMAP1", y = "PaCMAP2", 
-                                          num_bins_x = num_bins_x_pacmap_gau, 
-                                          num_bins_y = num_bins_y_pacmap_gau, 
-                                          x_start = NA, y_start = NA, 
-                                          buffer_x = NA, buffer_y = NA, 
-                                          hex_size = hex_size_pacmap_gau,
-                                          is_rm_lwd_hex = FALSE, 
-                                          benchmark_to_rm_lwd_hex = NA, 
-                                          col_start_2d = "PaCMAP", 
-                                          col_start_highd = "x")
-
-df_bin_centroids_pacmap_gau <- model_object_pacmap_gau$df_bin_centroids
-df_bin_pacmap_gau <- model_object_pacmap_gau$df_bin
-
-## Triangulate bin centroids
-tr1_object_pacmap_gau <- tri_bin_centroids(df_bin_centroids_pacmap_gau, x = "c_x", y = "c_y")
-tr_from_to_df_pacmap_gau <- gen_edges(tri_object = tr1_object_pacmap_gau)
-
-## Compute 2D distances
-distance_pacmap_gau <- cal_2d_dist(tr_coord_df = tr_from_to_df_pacmap_gau, 
-                                 start_x = "x_from", start_y = "y_from", 
-                                 end_x = "x_to", end_y = "y_to", 
-                                 select_vars = c("from", "to", "distance"))
-
-# distance_plot <- plot_dist(distance_pacmap_gau) +
-#   ylab(expression(d^{(2)})) +
-#   theme(axis.text = element_text(size = 5),
-#         axis.title = element_text(size = 12))
-# 
-# distance_plot
-
-## To find the benchmark value
-benchmark_pacmap_gau <- find_lg_benchmark(distance_edges = distance_pacmap_gau, distance_col = "distance")
-
-# ggplot() +
-# geom_trimesh(data = df_bin_centroids_pacmap_gau, mapping = aes(x = c_x, y = c_y))
-
-trimesh_removed_pacmap_gau <- vis_rmlg_mesh(distance_edges = distance_pacmap_gau, benchmark_value = benchmark_pacmap_gau, tr_coord_df = tr_from_to_df_pacmap_gau, distance_col = "distance")
-
-trimesh_removed_pacmap_gau <- trimesh_removed_pacmap_gau +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'e', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-df_all_pacmap_gau <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID), pacmap_data_with_hb_id_gau)
-
-# show_langevitour(df = df_all_pacmap_gau, df_b = df_bin_pacmap_gau, df_b_with_center_data = df_bin_centroids_pacmap_gau, col_start = "x", distance_df = distance_pacmap_gau, distance_col = "distance", benchmark_value = benchmark_pacmap_gau)
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-#| label: fig-modelfiveGau
-#| fig-cap: "Is there a best model to represent the original data in 2D space or are they all providing equivalent information?, (a) Model generated with 21 and 28 bins along the x and y axes, respectively, using a hexagonal size of 0.03 in the 2D space with tSNE, (b) Model generated with 60 and 79 bins along the x and y axes, respectively, using a hexagonal size of 0.01 in the 2D space with UMAP, (c) Model generated with 60 and 1927 bins along the x and y axes, respectively, using a hexagonal size of 0.01 in the 2D space with PHATE, (d) Model generated with 46 and 61 bins along the x and y axes, respectively, using a hexagonal size of 0.013 in the 2D space with TriMAP, and (e) Model generated with 31 and 40 bins along the x and y axes, respectively, using a hexagonal size of 0.02 in the 2D space with PaCMAP."
-#| out-width: 100%
-
-trimesh_removed_tsne_gau + trimesh_removed_umap_gau + trimesh_removed_phate_gau + trimesh_removed_trimap_gau + trimesh_removed_pacmap_gau +
-  #plot_annotation(tag_levels = 'a') +
-  plot_layout(ncol = 5) &
-  theme(legend.position='none', plot.tag = element_text(size = 8))
-```
-
-::: {#fig-gau1_sc layout-ncol="3" fig-pos="H"}
-![](figures/five_gau_clusters/sc_tsne_1.png){#fig-gau1_sc1}
-
-![](figures/five_gau_clusters/sc_tsne_2.png){#fig-gau1_sc2}
-
-![](figures/five_gau_clusters/sc_tsne_3.png){#fig-gau1_sc3}
-
-Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with tSNE in high-D, a video of the tour animation is available at (<https://youtu.be/oQxEb4wRdHI>).
-:::
-
-::: {#fig-gau2_sc layout-ncol="3" fig-pos="H"}
-![](figures/five_gau_clusters/sc_umap_1.png){#fig-gau2_sc1}
-
-![](figures/five_gau_clusters/sc_umap_2.png){#fig-gau2_sc2}
-
-![](figures/five_gau_clusters/sc_umap_3.png){#fig-gau2_sc3}
-
-Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with UMAP in high-D, a video of the tour animation is available at (<https://youtu.be/JW49csPpDx4>).
-:::
-
-::: {#fig-gau3_sc layout-ncol="3" fig-pos="H"}
-![](figures/five_gau_clusters/sc_phate_1.png){#fig-gau3_sc1}
-
-![](figures/five_gau_clusters/sc_phate_2.png){#fig-gau3_sc2}
-
-![](figures/five_gau_clusters/sc_phate_3.png){#fig-gau3_sc3}
-
-Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with PHATE in high-D, a video of the tour animation is available at (<https://youtu.be/xqhRODt8Z-o>).
-:::
-
-::: {#fig-gau4_sc layout-ncol="3" fig-pos="H"}
-![](figures/five_gau_clusters/sc_trimap_1.png){#fig-gau4_sc1}
-
-![](figures/five_gau_clusters/sc_trimap_2.png){#fig-gau4_sc2}
-
-![](figures/five_gau_clusters/sc_trimap_3.png){#fig-gau4_sc3}
-
-Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with TriMAP in high-D, a video of the tour animation is available at (<https://youtu.be/X-uHSB4VoZ0>).
-:::
-
-::: {#fig-gau5_sc layout-ncol="3" fig-pos="H"}
-![](figures/five_gau_clusters/sc_pacmap_1.png){#fig-gau5_sc1}
-
-![](figures/five_gau_clusters/sc_pacmap_2.png){#fig-gau5_sc2}
-
-![](figures/five_gau_clusters/sc_pacmap_3.png){#fig-gau5_sc3}
-
-Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with PaCMAP in high-D, a video of the tour animation is available at (<https://youtu.be/UxvU8HQd7JM>).
-:::
-
-## Applications {#sec-applications}
-
-### Single-cell RNA-seq data of human
-
-In the field of single-cell studies, a common analytical task involves clustering to identify groups of cells with similar expression profiles.
-Analysts often turn to NLDR techniques to verify and identify these clusters and explore developmental trajectories.
-To illustrate the importance of NLDR techniques and parameter selection in identifying clusters, Human Peripheral Blood Mononclear Cells (PBMC3k) dataset [@chen2023] is used.
-In a study by @chen2023, this dataset was used to demonstrate how UMAP represents clusters (see @fig-umapauthor).
-As shown in @fig-umapauthor, there are three distinct and well-separated clusters.
-
-<!-- UMAP layout with author's suggested parameter choice-->
-```{r}
-#| warning: false
-#| echo: false
-
-## Import data
-training_data_pbmc <- read_rds("data/pbmc3k/pbmc_pca_50.rds")
-training_data_pbmc <- training_data_pbmc[, 1:9] |>
-  mutate(ID = 1:NROW(training_data_pbmc))
-
-umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_30_min_dist_0.3.rds")
-umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc, 
-                                                                 x = "UMAP1", y = "UMAP2"))) |>
-  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
-                  "UMAP2" = "scaled_UMAP2")) |>
-  dplyr::mutate(ID = 1:NROW(umap_pbmc))
-
-umap_plot_pbmc <- umap_pbmc_scaled |>
-  ggplot(aes(x = UMAP1,
-             y = UMAP2))+
-  geom_point(alpha=0.5) +
-  coord_equal() +
-  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-        legend.key.height = unit(0.25, 'cm'),
-        legend.key.width = unit(0.25, 'cm')) 
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "2D layout from UMAP (n_neighbors = 30, min_dist = 0.3) applied for the PBMC3k dataset. Is this a best representation of the original data?"
-#| label: fig-umapauthor
-#| fig-pos: H
-
-umap_plot_pbmc 
-```
-
-To determine whether the UMAP representation with the parameter choice suggested by @chen2023 preserves the original data structure, we visualize the model constructed with UMAP overlaid on the high-D data (model-in-data space).
-The figures in @fig-pbmc1_sc show three well-separated clusters, indicating that the suggested UMAP representation preserves the global structure (see @fig-umapauthor).
-However, as shown in @fig-pbmc1_sc, these clusters are close to each other in high-D space.
-Also, nonlinear continuity patterns and high-density patches within the clusters are observed (see @fig-pbmc1_sc).
-Therefore, the suggested UMAP representation (see @fig-umapauthor) does not accurately preserve the local structure of the PBMC3k dataset.
-
-<!-- Fit the model and compute error-->
-```{r}
-#| warning: false
-#| echo: false
-#| message: false
-
-## Decide by looking at MSE plot
-num_bins_x_pbmc <- 21
-num_bins_y_pbmc <- 28
-hex_size_pbmc <- 0.03
-
-hb_obj_pbmc <- hex_binning(data = umap_pbmc_scaled, x = "UMAP1", 
-                      y = "UMAP2", num_bins_x = num_bins_x_pbmc, 
-                      num_bins_y = num_bins_y_pbmc, x_start = NA, y_start = NA, 
-                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_pbmc, col_start = "UMAP")
-
-## Data set with all possible centroids in the hexagonal grid
-all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_pbmc$centroids))
-
-## Generate all coordinates of hexagons
-hex_grid <- as.data.frame(do.call(cbind, hb_obj_pbmc$hex_poly))
-
-## To obtain the standardise counts within hexbins
-counts_df <- as.data.frame(do.call(cbind, hb_obj_pbmc$std_cts))
-df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
-                                             counts_df = counts_df)
-# ggplot(data = hex_grid, aes(x = x, y = y)) +
-#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
-#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
-#   coord_fixed()
-
-hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
-
-# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
-#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
-#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
-#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
-#   coord_fixed()
-
-umap_data_with_hb_id_pbmc <- as.data.frame(do.call(cbind, hb_obj_pbmc$data_hb_id))
-
-model_object <- fit_highd_model( training_data = training_data_pbmc, 
-                                 nldr_df_with_id = umap_pbmc_scaled, 
-                                 x = "UMAP1", y = "UMAP2", 
-                                 num_bins_x = num_bins_x_pbmc, 
-                                 num_bins_y = num_bins_y_pbmc, 
-                                 x_start = NA, y_start = NA, 
-                                 buffer_x = NA, buffer_y = NA, 
-                                 hex_size = hex_size_pbmc,
-                                 is_rm_lwd_hex = FALSE, 
-                                 benchmark_to_rm_lwd_hex = NA, 
-                                 col_start_2d = "UMAP", 
-                                 col_start_highd = "PC_")
-
-df_bin_centroids_pbmc <- model_object$df_bin_centroids
-df_bin_pbmc <- model_object$df_bin
-
-## Triangulate bin centroids
-tr1_object_pbmc <- tri_bin_centroids(df_bin_centroids_pbmc, x = "c_x", y = "c_y")
-tr_from_to_df_pbmc <- gen_edges(tri_object = tr1_object_pbmc)
-
-## Compute 2D distances
-distance_pbmc <- cal_2d_dist(tr_coord_df = tr_from_to_df_pbmc, 
-                             start_x = "x_from", start_y = "y_from", 
-                             end_x = "x_to", end_y = "y_to", 
-                             select_vars = c("from", "to", "distance"))
-
-# distance_plot <- plot_dist(distance_pbmc) +
-#   ylab(expression(d^{(2)})) +
-#   theme(axis.text = element_text(size = 5),
-#         axis.title = element_text(size = 12))
-# 
-# distance_plot
-
-## To find the benchmark value
-benchmark_pbmc <- find_lg_benchmark(distance_edges = distance_pbmc, distance_col = "distance")
-
-# ggplot() +
-# geom_trimesh(data = df_bin_centroids_pbmc, mapping = aes(x = c_x, y = c_y))
-
-trimesh_removed_pbmc_umap <- vis_rmlg_mesh(distance_edges = distance_pbmc, benchmark_value = benchmark_pbmc, tr_coord_df = tr_from_to_df_pbmc, distance_col = "distance")
-
-trimesh_removed_pbmc_umap <- trimesh_removed_pbmc_umap +
-  geom_point(data = umap_pbmc_scaled, aes(x = UMAP1, y = UMAP2), alpha = 0.3) +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-# df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), umap_data_with_hb_id_pbmc)
-# 
-# show_langevitour(df_all_pbmc, df_bin_pbmc, df_bin_centroids_pbmc, benchmark_pbmc, distance_pbmc, "distance", col_start = "PC_")
-
-df_bin_train <- df_bin_pbmc
-names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])
-
-df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), umap_data_with_hb_id_pbmc)
-
-error_df <- df_all_pbmc |>
-  dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates
-
-for (i in 1:(NCOL(df_bin_train) - 1)) {
-
-  error_df[ , paste0("abs_residual_", "PC_", i)] <- abs(error_df[ , paste0("PC_", i)] - error_df[ , paste0("avg_", "PC_", i)])
-
-}
-
-error_df <- error_df |>
-  dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "PC_")))))
-
-error_df <- error_df |>
-  mutate(type = if_else(total == 0, "no error",
-                        if_else(total <= 5, "error 0-5",
-                                if_else(total <= 10, "error 5-10",
-                                        if_else(total <= 15, "error 10-15",
-                                                if_else(total <= 20, "error 15-20",
-                                                        if_else(total <= 25, "error 20-25",
-                                                                "error greter than 25")))))))
-
-plot_list2_pbmc_error <- error_df |>
-  mutate(type = factor(type , levels = c("no error", "error 0-5", "error 5-10", "error 10-15", "error 15-20", "error 20-25", "error greter than 25"))) |>
-    ggplot(aes(x = UMAP1,
-               y = UMAP2, color = type))+
-    geom_point(alpha=0.5, size = 0.1) +
-    coord_equal() +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=rainbow_hcl(7)) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "(a) Model generated with 21 and 28 bins along x and y axes with hexagonal size 0.03 in the 2D space overlaid on UMAP data, and (b) high-D model error in model space."
-#| label: fig-nldervisPBMCUMAP
-#| fig-pos: H
-
-trimesh_removed_pbmc_umap + plot_list2_pbmc_error +
-plot_layout(guides='collect', ncol=2) &
-  theme(legend.position='bottom') 
-```
-
-::: {#fig-pbmc1_sc layout-ncol="3" fig-pos="H"}
-![](figures/pbmc3k/sc_1.png){#fig-pbmc1_sc1}
-
-![](figures/pbmc3k/sc_2.png){#fig-pbmc1_sc2}
-
-![](figures/pbmc3k/sc_3.png){#fig-pbmc1_sc3}
-
-Screen shots of the **langevitour** of the PBMC3k data set, shows the model in high-D, a video of the tour animation is available at (<https://youtu.be/Gnhh4hfsbbg>).
-:::
-
-To find the best NLDR representation for PBMC3k dataset, we observe the absolute error for different number of non-empty bins and select the one with the lowest error.
-As shown in @fig-abserrorPBMC, tSNE with a perplexity set to $51$, has the lowest error.
-Therefore, tSNE with a perplexity of $51$ is used as the best representation for PBMC3k dataset.
-
-<!--compute absolute error for different parameter choices-->
-
-```{r}
-#| warning: false
-#| echo: false
-
-error_df <- read_csv("data/pbmc3k/error_df.csv", col_names = FALSE)
-names(error_df) <- c("num_bins", "error", "mse", "num_bins_x", "num_bins_y", "hex_size", "num_non_empty_bins", "n_neighbors", "min_dist")
-
-error_df <- error_df |>
-  dplyr::mutate(type = paste0("n_neighbors: ", n_neighbors, ", min_dist: ", min_dist))
-
-error_df_tsne <- read_csv("data/pbmc3k/error_df_tsne.csv", col_names = FALSE)
-names(error_df_tsne) <- c("num_bins", "error", "mse", "num_bins_x", "num_bins_y", "hex_size", "num_non_empty_bins", "perplexity")
-error_df_tsne <- error_df_tsne |>
-  dplyr::mutate(type = paste0("perplexity: ", perplexity))
-
-error_df <- bind_rows(error_df[, c(1:7, 10)], error_df_tsne[, c(1:7, 9)])
-
-abs_error_plot_pbmc <- ggplot(error_df, aes(x = num_non_empty_bins,
-                                       y = log(error), group = type,
-colour = type)) +
-  geom_point() +
-  geom_line() +
-   geom_vline(xintercept = 439, linetype="solid",
-                color = "black", size=0.8, alpha = 0.5) +
-  theme_light() +
-  theme(legend.position = "bottom", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7),
-        legend.text = element_text(size = 5)) +
-  scale_color_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99','#e31a1c','#fdbf6f','#ff7f00','#cab2d6','#6a3d9a', '#b15928', '#ffff99')) +
-  ylab("log(absolute error)") +
-  xlab("number of non-empty bins")
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "Absolute error from UMAP and tSNE applied to training PBMC3k dataset with diffferent parameter choices.
-What is the best parameter choice to create the model?
-The residual plot have a steep slope at the beginning, indicating that a smaller number of non-empty bins causes a larger amount of error.
-Then, the slope gradually declines or level off, indicating that a higher number of non-empty bins generates a smaller error.
-Using the elbow method, it was observed that when the number of non-empty bins is set to 439, the lowest error occurred with the parameters perplexity: 51."
-#| label: fig-abserrorPBMC
-#| fig-pos: H
-#| out-width: 100%
-#| out-height: 100%
-
-abs_error_plot_pbmc
-```
-
-As shown in @fig-tsnesuggest, there are three well-separated clusters, although they are located close to each other.
-Additionally, nonlinear structures can be observed within the clusters (see @fig-tsnesuggest).
-Hence, the data structure that UMAP failed to capture for the PBMC3k dataset is successfully captured by tSNE.
-
-```{r}
-#| warning: false
-#| echo: false
-
-## Import data
-training_data_pbmc <- read_rds("data/pbmc3k/pbmc_pca_50.rds")
-training_data_pbmc <- training_data_pbmc[, 1:9] |>
-  mutate(ID = 1:NROW(training_data_pbmc))
-
-tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_51.rds")
-tsne_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tsne_pbmc, 
-                                                                 x = "tSNE1", y = "tSNE2"))) |>
-  dplyr::rename(c("tSNE1" = "scaled_tSNE1", 
-                  "tSNE2" = "scaled_tSNE2")) |>
-  dplyr::mutate(ID = 1:NROW(tsne_pbmc))
-
-tsne_plot_pbmc <- tsne_pbmc_scaled |>
-  ggplot(aes(x = tSNE1,
-             y = tSNE2))+
-  geom_point(alpha=0.5) +
-  coord_equal() +
-  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-        legend.key.height = unit(0.25, 'cm'),
-        legend.key.width = unit(0.25, 'cm')) 
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "2D layout from tSNE (perplexity = 51) applied for the PBMC3k dataset. Is this a best representation of the original data?"
-#| label: fig-tsnesuggest
-#| fig-pos: H
-
-tsne_plot_pbmc 
-```
-
-Next, the model is fitted for tSNE, and the resulting model is visualized in the data space.
-The model shows some quirks, as shown in @fig-pbmc2_sc.
-The large cluster is divided into three subclusters, which are in close to each other.
-
-<!--Fit the model-->
-```{r}
-#| warning: false
-#| echo: false
-#| message: false
-
-## Decide by looking at MSE plot
-num_bins_x_pbmc <- 31
-num_bins_y_pbmc <- 40
-hex_size_pbmc <- 0.02
-
-hb_obj_pbmc <- hex_binning(data = tsne_pbmc_scaled, x = "tSNE1", 
-                      y = "tSNE2", num_bins_x = num_bins_x_pbmc, 
-                      num_bins_y = num_bins_y_pbmc, x_start = NA, y_start = NA, 
-                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_pbmc, col_start = "tSNE")
-
-## Data set with all possible centroids in the hexagonal grid
-all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_pbmc$centroids))
-
-## Generate all coordinates of hexagons
-hex_grid <- as.data.frame(do.call(cbind, hb_obj_pbmc$hex_poly))
-
-## To obtain the standardise counts within hexbins
-counts_df <- as.data.frame(do.call(cbind, hb_obj_pbmc$std_cts))
-df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
-                                             counts_df = counts_df)
-# ggplot(data = hex_grid, aes(x = x, y = y)) +
-#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
-#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
-#   coord_fixed()
-
-hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
-
-# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
-#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
-#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
-#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
-#   coord_fixed()
-
-tsne_data_with_hb_id_pbmc <- as.data.frame(do.call(cbind, hb_obj_pbmc$data_hb_id))
-
-model_object <- fit_highd_model( training_data = training_data_pbmc, 
-                                 nldr_df_with_id = tsne_pbmc_scaled, 
-                                 x = "tSNE1", y = "tSNE2", 
-                                 num_bins_x = num_bins_x_pbmc, 
-                                 num_bins_y = num_bins_y_pbmc, 
-                                 x_start = NA, y_start = NA, 
-                                 buffer_x = NA, buffer_y = NA, 
-                                 hex_size = hex_size_pbmc,
-                                 is_rm_lwd_hex = FALSE, 
-                                 benchmark_to_rm_lwd_hex = NA, 
-                                 col_start_2d = "tSNE", 
-                                 col_start_highd = "PC_")
-
-df_bin_centroids_pbmc <- model_object$df_bin_centroids
-df_bin_pbmc <- model_object$df_bin
-
-benchmark_value_rm_lwd <- stats::quantile(df_bin_centroids_pbmc$std_counts,
-                probs = c(0,0.25,0.5,0.75,1), names = FALSE)[2]
-
-df_bin_centroids_low <- df_bin_centroids_pbmc |>
-  dplyr::filter(std_counts <= benchmark_value_rm_lwd)
-
-identify_rm_bins <- find_low_dens_hex(df_bin_centroids_all = df_bin_centroids_pbmc,
-                                      num_bins_x = num_bins_x_pbmc,
-                                      df_bin_centroids_low = df_bin_centroids_low)
-
-df_bin_centroids_pbmc <- df_bin_centroids_pbmc |>
-  dplyr::filter(!(hexID %in% identify_rm_bins))
-
-## Triangulate bin centroids
-tr1_object_pbmc <- tri_bin_centroids(df_bin_centroids_pbmc, x = "c_x", y = "c_y")
-tr_from_to_df_pbmc <- gen_edges(tri_object = tr1_object_pbmc)
-
-## Compute 2D distances
-distance_pbmc <- cal_2d_dist(tr_coord_df = tr_from_to_df_pbmc, 
-                             start_x = "x_from", start_y = "y_from", 
-                             end_x = "x_to", end_y = "y_to", 
-                             select_vars = c("from", "to", "distance"))
-
-# distance_plot <- plot_dist(distance_pbmc) +
-#   ylab(expression(d^{(2)})) +
-#   theme(axis.text = element_text(size = 5),
-#         axis.title = element_text(size = 12))
-# 
-# distance_plot
-
-## To find the benchmark value
-benchmark_pbmc <- find_lg_benchmark(distance_edges = distance_pbmc, distance_col = "distance")
-benchmark_pbmc <- 0.04
-
-# ggplot() +
-# geom_trimesh(data = df_bin_centroids_pbmc, mapping = aes(x = c_x, y = c_y))
-
-trimesh_removed_pbmc_tsne <- vis_rmlg_mesh(distance_edges = distance_pbmc, benchmark_value = benchmark_pbmc, tr_coord_df = tr_from_to_df_pbmc, distance_col = "distance")
-
-trimesh_removed_pbmc_tsne <- trimesh_removed_pbmc_tsne +
-  geom_point(data = tsne_pbmc_scaled, aes(x = tSNE1, y = tSNE2), alpha = 0.08) +
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-# df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), tsne_data_with_hb_id_pbmc)
-# 
-# show_langevitour(df_all_pbmc, df_bin_pbmc, df_bin_centroids_pbmc, benchmark_pbmc, distance_pbmc, "distance", col_start = "PC_")
-
-df_bin_train <- df_bin_pbmc
-names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])
-
-df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), tsne_data_with_hb_id_pbmc)
-
-error_df <- df_all_pbmc |>
-  dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates
-
-for (i in 1:(NCOL(df_bin_train) - 1)) {
-
-  error_df[ , paste0("abs_residual_", "PC_", i)] <- abs(error_df[ , paste0("PC_", i)] - error_df[ , paste0("avg_", "PC_", i)])
-
-}
-
-error_df <- error_df |>
-  dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "PC_")))))
-
-error_df <- error_df |>
-  mutate(type = if_else(total == 0, "no error",
-                        if_else(total <= 5, "error 0-5",
-                                if_else(total <= 10, "error 5-10",
-                                        if_else(total <= 15, "error 10-15",
-                                                if_else(total <= 20, "error 15-20",
-                                                        "error greter than 20"))))))
-
-plot_list2_pbmc_error <- error_df |>
-  mutate(type = factor(type , levels = c("no error", "error 0-5", "error 5-10", "error 10-15", "error 15-20", "error greter than 20"))) |>
-    ggplot(aes(x = tSNE1,
-               y = tSNE2, color = type))+
-    geom_point(alpha=0.5, size = 0.1) +
-    coord_equal() +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=rainbow_hcl(6)) +
-  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "(a) Model generated with 31 and 40 bins along x and y axes with hexagonal size 0.02 in the 2D space overlaid on tSNE data, and (b) high-D model error in model space."
-#| label: fig-nldervisPBMCtSNE
-#| fig-pos: H
-
-trimesh_removed_pbmc_tsne + plot_list2_pbmc_error +
-plot_layout(guides='collect', ncol=2) &
-  theme(legend.position='bottom') 
-```
-
-::: {#fig-pbmc2_sc layout-ncol="3" fig-pos="H"}
-![](figures/pbmc3k/sc_4.png){#fig-pbmc2_sc4}
-
-![](figures/pbmc3k/sc_5.png){#fig-pbmc2_sc5}
-
-![](figures/pbmc3k/sc_6.png){#fig-pbmc2_sc6}
-
-Screen shots of the **langevitour** of the PBMC3k data set, shows the model in high-D, a video of the tour animation is available at (<https://youtu.be/R0EroJcGbas>).
-:::
-
-### Handwritten digits
-
-The MNIST dataset consists of grayscale images of handwritten digits [@lecun2010].
-@Yingfan2021 used this dataset to demonstrate how the PaCMAP preserves local structure.
-We selected the 2D embedding of PaCMAP for the handwritten digit 1.
-As shown in @fig-pacmapauthor, we observe a nonlinear structure present in 2D space.
-Furthermore, the angle of digit 1 images varies along the nonlinear structure (see @fig-sampleimg).
-To answer the following questions,
-
-- Is the original data structure of the handwritten digit 1 dataset nonlinear?
-- Where does PaCMAP preserve the original data structure?
-- Where does PaCMAP show quirks?
-
-we visualize the model constructed with PaCMAP overlaid on data in high-D space (model-in-data space).
-
-According to @fig-mnist1_sc1, the nonlinear continuity structure observed in the 2D representation of PaCMAP is also visible when visualizing the model overlaid on the data space.
-This provides evidence that the model accurately captures the structure of the high-D data.
-Also, the model has a twisted pattern within the nonlinear structure, which is an addition pattern of data that cannot be seen in the 2D representation (see @fig-mnist1_sc2).
-
-In order to visually assess the preservation of local structure, it's important to consider the edges in the model.
-In a model that effectively preserves local structure, there shouldn't be any long edges in high-D space.
-However, as shown in @fig-mnist1_sc3, some long edges exist in high-D space that are not identified as long edges in the 2D representation.
-
-<!-- PaCMAP layout with author's suggested parameter choice-->
-```{r}
-#| warning: false
-#| echo: false
-
-## Import data
-training_data_mnist <- read_rds("data/mnist/mnist_10_pcs_of_digit_1.rds")
-training_data_mnist <- training_data_mnist |>
-  mutate(ID = 1:NROW(training_data_mnist))
-
-pacmap_minst <- read_rds("data/mnist/mnist_pacmap.rds")
-pacmap_minst_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = pacmap_minst, 
-                                    x = "PaCMAP1", y = "PaCMAP2"))) |>
-  dplyr::rename(c("PaCMAP1" = "scaled_PaCMAP1", 
-                  "PaCMAP2" = "scaled_PaCMAP2")) |>
-  dplyr::mutate(ID = 1:NROW(pacmap_minst))
-
-position_df <- pacmap_minst_scaled |> 
-  dplyr::filter(ID %in% c(6475, 3311, 6489, 2347, 387, 5239, 728, 517, 6369, 3055, 6345, 1761)) |>
-  dplyr::mutate(position = c(5, 9, 7, 8, 4, 11, 2, 6, 12, 10, 1, 3))
-
-pacmap_plot_mnist <- pacmap_minst_scaled |>
-    ggplot(aes(x = PaCMAP1,
-               y = PaCMAP2))+
-    geom_point(alpha=0.1) +
-    coord_equal() +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + 
-  geom_text(data = position_df, aes(x = PaCMAP1, y = PaCMAP2, label = position), colour = "red", size = 5) +
-  theme_linedraw() +
-    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) 
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "(a) 2D layout from PaCMAP (n_components=2, n_neighbors=10, init=random, MN_ratio=0.9, FP_ratio=2.0) applied for the digit 1 of the MNIST dataset.
-Is this a best representation of the original data?"
-#| label: fig-pacmapauthor
-#| fig-pos: H
-
-pacmap_plot_mnist 
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-mnist_data <- read_rds("data/mnist/mnist_digit_1.rds")
-```
-
-<!--images with positions-->
-```{r}
-#| warning: false
-#| echo: false
-
-pixels_gathered <-  mnist_data |>
-  mutate(instance = row_number()) |>
-  gather(pixel, value, -Label, -instance) |>
-  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
-  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
-  filter(instance %in% c(3710, 2391, 2385, 5030, 6475,
-                         7679, 3568, 1312, 3311, 1097,
-                         3552, 7853, 6489, 7689, 6690,
-                        1380, 6057, 2347, 5946, 3355,
-                        4175, 3997, 5378, 387, 1854,
-                        614, 3079, 1762, 5239, 3723,
-                        5748, 728, 7419, 7794, 6233,
-                        33, 2485, 5998, 318, 1761,
-                        5690, 165, 517, 6935, 2682,
-                        4962, 2264, 5563, 6369, 559,
-                        5188, 4849, 2666, 3448, 3055,
-                        3120, 6869, 6345, 4470, 7147)) |>
-  mutate(instance = as.character(instance))
-
-theme_set(theme_light())
-img_sample <- pixels_gathered |> 
-  ggplot(aes(x, y, fill = value)) +  
-  geom_tile() +  
-  facet_wrap(~ factor(instance, levels = c(3710, 2391, 2385, 5030, 6475,
-                         7679, 3568, 1312, 3311, 1097,
-                         3552, 7853, 6489, 7689, 6690,
-                        1380, 6057, 2347, 5946, 3355,
-                        4175, 3997, 5378, 387, 1854,
-                        614, 3079, 1762, 5239, 3723,
-                        5748, 728, 7419, 7794, 6233,
-                        33, 2485, 5998, 318, 1761,
-                        5690, 165, 517, 6935, 2682,
-                        4962, 2264, 5563, 6369, 559,
-                        5188, 4849, 2666, 3448, 3055,
-                        3120, 6869, 6345, 4470, 7147)), ncol = 10) +
-  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm'),
-        strip.text = element_blank(),
-        aspect.ratio = 1)
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "Images of the handwritten digit 1 are distributed along the right-to-left of first embedding axis of the PaCMAP nonlinear structure."
-#| label: fig-sampleimg
-#| out-width: 100%
-#| fig-pos: H
-
-img_sample
-```
-
-<!-- Fit the model and compute error-->
-```{r}
-#| warning: false
-#| echo: false
-#| message: false
-
-## Decide by looking at MSE plot
-num_bins_x_mnist <- 12
-num_bins_y_mnist <- 15
-hex_size_mnist <- 0.06
-
-hb_obj_mnist <- hex_binning(data = pacmap_minst_scaled, x = "PaCMAP1", 
-                      y = "PaCMAP2", num_bins_x = num_bins_x_mnist, 
-                      num_bins_y = num_bins_y_mnist, x_start = NA, y_start = NA, 
-                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_mnist, col_start = "PaCMAP")
-
-## Data set with all possible centroids in the hexagonal grid
-all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_mnist$centroids))
-
-## Generate all coordinates of hexagons
-hex_grid <- as.data.frame(do.call(cbind, hb_obj_mnist$hex_poly))
-
-## To obtain the standardise counts within hexbins
-counts_df <- as.data.frame(do.call(cbind, hb_obj_mnist$std_cts))
-df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
-                                             counts_df = counts_df)
-# ggplot(data = hex_grid, aes(x = x, y = y)) +
-#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
-#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
-#   coord_fixed()
-
-hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
-
-# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
-#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
-#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
-#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
-#   coord_fixed()
-
-pacmap_data_with_hb_id_mnist <- as.data.frame(do.call(cbind, hb_obj_mnist$data_hb_id))
-
-model_object <- fit_highd_model( training_data = training_data_mnist, 
-                                 nldr_df_with_id = pacmap_minst_scaled, 
-                                 x = "PaCMAP1", y = "PaCMAP2", 
-                                 num_bins_x = num_bins_x_mnist, 
-                                 num_bins_y = num_bins_y_mnist, 
-                                 x_start = NA, y_start = NA, 
-                                 buffer_x = NA, buffer_y = NA, 
-                                 hex_size = hex_size_mnist,
-                                 is_rm_lwd_hex = FALSE, 
-                                 benchmark_to_rm_lwd_hex = NA, 
-                                 col_start_2d = "PaCMAP", 
-                                 col_start_highd = "PC")
-
-df_bin_centroids_mnist <- model_object$df_bin_centroids
-df_bin_mnist <- model_object$df_bin
-
-## Triangulate bin centroids
-tr1_object_mnist <- tri_bin_centroids(df_bin_centroids_mnist, x = "c_x", y = "c_y")
-tr_from_to_df_mnist <- gen_edges(tri_object = tr1_object_mnist)
-
-## Compute 2D distances
-distance_mnist <- cal_2d_dist(tr_coord_df = tr_from_to_df_mnist, 
-                             start_x = "x_from", start_y = "y_from", 
-                             end_x = "x_to", end_y = "y_to", 
-                             select_vars = c("from", "to", "distance"))
-
-# distance_plot <- plot_dist(distance_mnist) +
-#   ylab(expression(d^{(2)})) +
-#   theme(axis.text = element_text(size = 5),
-#         axis.title = element_text(size = 12))
-# 
-# distance_plot
-
-## To find the benchmark value
-benchmark_mnist <- find_lg_benchmark(distance_edges = distance_mnist, distance_col = "distance")
-
-# ggplot() +
-# geom_trimesh(data = df_bin_centroids_mnist, mapping = aes(x = c_x, y = c_y))
-
-trimesh_removed_mnist_pacmap <- vis_rmlg_mesh(distance_edges = distance_mnist, benchmark_value = benchmark_mnist, tr_coord_df = tr_from_to_df_mnist, distance_col = "distance")
-
-trimesh_removed_mnist_pacmap <- trimesh_removed_mnist_pacmap +
-  geom_point(data = pacmap_minst_scaled, aes(x = PaCMAP1, y = PaCMAP2), alpha = 0.06) + 
-  theme_linedraw() +
-  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-  annotate(geom = 'text', label = 'a', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-df_bin_train <- df_bin_mnist
-names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])
-
-df_all_mnist <- dplyr::bind_cols(training_data_mnist, pacmap_data_with_hb_id_mnist)
-
-error_df <- df_all_mnist |>
-  dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates
-
-# prediction_df_join <- prediction_df_join |>
-#   dplyr::left_join(data, by = c("ID" = "ID")) ## Map high-D data
-
-for (i in 1:(NCOL(df_bin_train) - 1)) {
-
-  error_df[ , paste0("abs_residual_", "PC", i)] <- abs(error_df[ , paste0("PC", i)] - error_df[ , paste0("avg_", "PC", i)])
-
-}
-
-error_df <- error_df |>
-  dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "PC")))))
-
-# library(ggbeeswarm)
-# error_df$group <- "1"
-# ggplot(error_df, aes(x = group, y = total)) +
-#   geom_quasirandom()+
-#   ylim(0, max(unlist(error_df$total))+ 0.5) + coord_flip()
-
-### The minimum error is 0 and the maximum is 42.17439
-### There is lot of points with error 0,
-
-error_df <- error_df |>
-  mutate(type = if_else(total <= 2, "error 0-2",
-                        if_else(total <= 4, "error 2-4",
-                                if_else(total <= 6, "error 4-6",
-                                        if_else(total <= 8, "error 6-8",
-                                                if_else(total <= 10, "error 8-10",
-                                                        if_else(total <= 12, "error 10-12",
-                                                                if_else(total <= 14, "error 12-14", "error greter than 14"))))))))
-
-plot_list2_mnist_error <- error_df |>
-  mutate(type = factor(type , levels = c("error 0-2", "error 2-4", "error 4-6", "error 6-8", "error 8-10", "error 10-12", "error 12-14", "error greter than 14"))) |>
-    ggplot(aes(x = PaCMAP1,
-               y = PaCMAP2, color = type, group = ID))+
-    geom_point(alpha=0.5, size = 0.1) +
-    coord_equal() +
-    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
-  theme_linedraw() +
-    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm')) +
-  scale_color_manual(values=rainbow_hcl(8)) +
-  annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-# show_langevitour(df_all_mnist, df_bin_mnist, df_bin_centroids_mnist, benchmark_mnist, distance_mnist, "distance", col_start = "PC")
-
-# predict_minst_df <- as.data.frame(do.call(cbind, predict_emb(training_data_mnist, df_bin_centroids_mnist, df_bin_mnist, "PaCMAP")))
-# gen_summary(training_data_mnist, predict_minst_df, df_bin_mnist, col_start = "PC")
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "(a) Model generated with 12 and 15 bins along x and y axes with hexagonal size 0.06 in the 2D space overlaid on PaCMAP data, and (b) high-D model error in model space."
-#| label: fig-nldervisMNISTPACMAP
-#| fig-pos: H
-
-trimesh_removed_mnist_pacmap + plot_list2_mnist_error +
-plot_layout(guides='collect', ncol=2) &
-  theme(legend.position='bottom')
-```
-
-::: {#fig-mnist1_sc layout-ncol="3" fig-pos="H"}
-![](figures/mnist/sc_1.png){#fig-mnist1_sc1}
-
-![](figures/mnist/sc_2.png){#fig-mnist1_sc2}
-
-![](figures/mnist/sc_3.png){#fig-mnist1_sc3}
-
-Screen shots of the **langevitour** of the MNIST digit 1 data set, shows the model-in-data space, a video of the tour animation is available at (<https://youtu.be/zq2GM9qvUNA>).
-:::
-
-In terms of model error, some data points show high error (see @fig-nldervisMNISTPACMAP) due to their deviation from the typical high-D data structure, acting as outliers.
-This is evident when comparing the images associated with these high model error points (see @fig-sampleimgerror) to others (see @fig-sampleimg).
-
-<!--images that occur large error-->
-```{r}
-#| warning: false
-#| echo: false
-
-img_error <- error_df |> 
-  dplyr::filter(total >= 10) |> 
-  dplyr::pull(ID)
-
-mnist_data <- read_rds("data/mnist/mnist_digit_1.rds")
-pixels_gathered <-  mnist_data |>
-  mutate(instance = row_number()) |>
-  gather(pixel, value, -Label, -instance) |>
-  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
-  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
-  filter(instance %in% img_error)
-
-theme_set(theme_light())
-imge_error_sample <- pixels_gathered |> 
-  ggplot(aes(x, y, fill = value)) +  
-  geom_tile() +  
-  facet_wrap(~ instance, ncol = 11) +
-  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-              axis.title.x = element_blank(), axis.title.y = element_blank(),
-              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-         legend.key.height = unit(0.25, 'cm'),
-         legend.key.width = unit(0.25, 'cm'),
-        strip.text = element_blank(),
-        aspect.ratio = 1)
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "Images of handwritten digit 1 which occur large model error."
-#| label: fig-sampleimgerror
-#| out-width: 100%
-#| fig-pos: H
-
-imge_error_sample
-```
-
-## Conclusion {#sec-conclusions}
-
-In conclusion, our proposed model offers a novel approach for visualizing high-dimensional (high-D) data by leveraging non-linear dimension reduction (NLDR) techniques.
-Through a series of steps including hexagonating NLDR data, triangulating bin centroids, and lifting the 2D triangular mesh into high dimensions, our model effectively transforms complex high-D data into interpretable 2D representations.
-The model's performance is evaluated using goodness of fit statistics such as Mean Squared Error (MSE) and Akaike Information Criterion (AIC), providing insights into its accuracy and reliability.
-Overall, our model presents a valuable tool for researchers and practitioners in various fields to gain deeper insights from high-dimensional datasets.
-
-Our algorithm mainly consists...
-
-These Goodness of fit statistics useful in encountering the prediction accuracy when we predict the 2D embedding for the new high-D data.
-Because, we try to find the nearest high-D mappings as the first step of prediction which will participate the prediction process.
-When we need to find which different NLDR technique and which parameter choices is giving the best representation of high-D data, MSE and AIC are also useful.
-
-We introduce a new tool to help to determine which method, which parameter choice provide the most useful representation of high-D data.
-
-<!--Our research introduces a comprehensive framework that leverages tours for interactive exploration of high-dimensional data coupled with a low-dimensional manifold, facilitated by the `quollr` R package.
-Regardless of the Non-Linear Dimension Reduction (NLDR) technique in use, our approach demonstrates effectiveness through simulation examples, particularly in the iterative removal of long edges for a smoother representation and capturing cluster variance.
-
-In the example with doublets, our method successfully captures the tweak within each cluster, indicating the variance present within them.
-However, the model may not appear smooth in high-dimensional space due to considerable noise when the data has a piecewise linear geometry, such as the tree simulation.
-
-The practical application of our framework, as showcased with the UMAP view, enables visual inspection of well-separated clusters.
-Furthermore, the combined tour and model provide a robust assessment of whether UMAP preserves the data structure and accurately transforms the data.
-
-The advantages of our approach include its versatility across various NLDR techniques and the ability to generate interactive visualizations for detailed exploration.
-The tour provides an intuitive way to navigate and comprehend high-dimensional data while assessing the accuracy of dimensionality reduction.
-
-However, one limitation is that the approach may be less effective in cases with significant noise, as seen in the tree simulation example.
-Additionally, while our method aids in visual verification, quantifying the accuracy of embeddings might require further evaluation metrics.
-
-In conclusion, our framework presents a powerful tool for researchers and analysts in single-cell studies to assess their embeddings by visually inspecting them alongside the original data.
-By leveraging the advantages of tours and low-dimensional manifolds, our approach offers valuable insights into the data transformation process, empowering users to make informed decisions in analyzing high-dimensional data.
-Future work could enhance the method's robustness in the presence of noise and explore additional evaluation metrics for quantifying embedding accuracy.-->
+  - Summarise contributions
+  - Explain where it is expected or not expected to work, eg higher dimensional relationships
+  - What might be useful enhancements
 

### 2024-03-27 paper.tex
cleaning up the paper

@@
 \hypersetup{
-  pdftitle={Visualising How Non-linear Dimension Reduction Warps Your Data},
+  pdftitle={Looking at Non-Linear Dimension Reductions as Models in the Data Space},
   pdfauthor={Jayani P.G. Lakshika; Dianne Cook; Paul Harrison; Michael Lydeamore; Thiyanga S. Talagala},
@@
 
-\title{\bf Visualising How Non-linear Dimension Reduction Warps Your
-Data}
+\title{\bf Looking at Non-Linear Dimension Reductions as Models in the
+Data Space}
 \author{
@@
 
-High-dimensional (high-D) data is prevalent across various fields, such
-as ecology and bioinformatics \citep{Guo2023}, due to advancements in
-data collection technologies \citep{Johnstone2009, ayesha2020overview}.
-However, visualization of high-D data introduces significant challenges,
-because the complexity of visualizing data beyond two dimensions
-\citep{Jia2022}. In recent years, interactive and dynamic graphics
-systems like \textbf{liminal} \citep{article21} ---which employs
-interactive tools like brushing and linking \citep{article58}---and
-software tools such as \textbf{XGobi}, \textbf{GGobi} \citep{article60},
-\textbf{tourr} \citep{article61}, \textbf{detourr} \citep{article22},
-and \textbf{langevitour} \citep{article09}, involving dynamic methods
-like tours \citep{Asimov1985}, have played a key role in visualizing
-high-D data (data-vis).
-
-To create low-dimensional representations (typically in 2D) (m-vis)
-\citep{article59} of high-D data, it is common to apply dimension
-reduction (DR) techniques. Approaches for DR involve linear methods such
-as principal component analysis (PCA) \citep{Karl1901}, non-linear
-methods such as multi-dimensional scaling (MDS) \citep{Torgerson1967}.
-In the past decade, many new non-linear dimension reduction (NLDR)
-techniques have emerged, such as t-distributed stochastic neighbor
-embedding (tSNE) \citep{Laurens2008} and uniform manifold approximation
-and projection (UMAP) \citep{Leland2018}. NLDR techniques are the 2D
-models of high-D data in our context.
-
-It is important to visualize various non-linear dimensionality reduction
-(NLDR) techniques for the same high-D data in order to understand and
-find the best representation. After doing so, the 2D models may differ
-considerably from each other and may also deviate from the original data
-structure in high-dimensional space. Therefore, visualizing the 2D model
-in high-D space (m-in-ds) is more useful to answer different types of
-questions:
-
 \begin{itemize}
+\tightlist
 \item
-  Is there a best 2D representation of high-D data or are they all
-  providing equivalent information? Is there a best parameter choice to
-  fit the 2D model? How does the model change when it's parameters
-  change?
+  What's the problem
 \item
-  How well the does the 2D models capture the data structure? Is the
-  model fitting able to capture different data structure like
-  non-linear, clustering?
+  Summary of literature
 \end{itemize}
 
-If we cannot easily ask and answer these questions, our ability to
-understand the models is limited. To find the best 2D model and
-parameter choices, a better understanding of the underlying science is
-important.
-
-Also, the importance of m-vis along with data-vis has been recognized
-and incorporated into interactive software, \textbf{liminal}
-\citep{article21}. But the 2D model and high-D visualize side by side
-and interactive like brushing and linking connect the data in the two
-panels. To address this challenge, we propose a novel approach by
-combining the tour technique with a low-dimensional manifold. This
-manifold is created through the synergistic use of NLDR techniques,
-hexagonal binning, and triangulation. This integration facilitates a
-more understanding of the data structure, how well (or how poorly) NLDR
-techniques perform.
-
-The outline of this paper is as follows. Section~\ref{sec-background}
-provides an detailed overview of dimension reduction methods, and tour
-technique. Building upon this foundation, Section
-Section~\ref{sec-methods} describes the proposed algorithm, including
-its implementation details, model tuning, summaries, and a synthetic
-example to demonstrate its functionality. In Section
-Section~\ref{sec-applications}, the algorithm's applications on various
-datasets, particularly in single-cell RNA-seq and handwritten digit
-data, are showcased. These applications demonstrate how some NLDR
-techniques effectively capture both local and global structures of the
-original data, while others may result in misleading interpretations.
-Finally, Section Section~\ref{sec-conclusions} summarizes the findings
-and describe the importance of the proposed approach in addressing the
-challenges associated with selecting the best NLDR method and parameter
-choices to achieve a more accurate 2D representation of high-D data.
-
-\begin{figure}
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervisUMAP-1.pdf}
-
-}
-
-\caption{\label{fig-nldervisUMAP}2D layouts from UMAP applied for the
-S-curve data: (a) UMAP (n\_neighbors = 7), (b) UMAP (n\_neighbors = 15),
-(c) UMAP (n\_neighbors = 32), (d) UMAP (n\_neighbors = 50). Is there a
-best hyperparameter choice in representing UMAP or are they all
-providing equivalent information?}
-
-\end{figure}%
-
-\section{Background}\label{sec-background}
-
-\subsection{Dimension Reduction}\label{dimension-reduction}
-
-Consider the high-D data a rectangular matrix \(X_{n \times p}\), where
-\(X_{n \times p} = \begin{bmatrix} \textbf{x}_{1} & \textbf{x}_{2} & \cdots & \textbf{x}_{n}\\ \end{bmatrix}^\top\),
-with \(n\) observations in \(p\) dimensions. The objective is to
-discover a low-dimensional projection
-\(Y_{n \times d} = \begin{bmatrix} \textbf{y}_{1} & \textbf{y}_{2} & \cdots & \textbf{y}_{n}\\ \end{bmatrix}^\top\),
-represented as an \(n\) × \(d\) matrix, where \(d \ll p\). The reduction
-process seeks to remove noise from the original data set while retaining
-essential information.
-
-There are two main categories of dimension reduction techniques: linear
-and non-linear methods. Linear techniques involve a linear
-transformation of the data, with one popular example being PCA. PCA
-performs an eigen-decomposition of the sample covariance matrix to
-obtain orthogonal principal components that capture the variance of the
-data \citep{Karl1901}.
-
-In contrast, NLDR techniques generate the low-dimensional representation
-\(Y\) from the high-dimensional data \(X\), often using pre-processing
-techniques like \(k\)-nearest neighbors graph or kernel transformations.
-Multidimensional Scaling (MDS) is a class of NLDR methods that aims to
-construct an embedding \(Y\) in a low-dimensional space, approximating
-the pair-wise distances in \(X\) \citep{Torgerson1967}. Variants of MDS
-include non-metric scaling \citep{article62} and Isomap, which estimate
-geodesic distances to create the low-dimensional representation
-\citep{article63}. Other approaches based on diffusion processes, like
-diffusion maps \citep{article64} and the PHATE (Potential of
-Heat-diffusion for Affinity-based Trajectory Embedding) algorithm
-\citep{article03}, also fall under NLDR methods.
-
-\subsubsection{Non-linear dimension reduction
-techniques}\label{non-linear-dimension-reduction-techniques}
-
-NLDR techniques are crucial for analyzing and displaying
-high-dimensional data, where linear approaches may not adequately
-capture complexities in relationships between variables
-\citep{Johnstone2009}. One of the challenges with NLDR techniques is the
-selection and tuning of appropriate hyperparameters \citep{liao2023}.
-This process involves finding the suitable combination of
-hyperparameters that enhances the performance of the NLDR technique,
-considering the characteristics of the dataset and the specific goals of
-the analysis.
-
-Additionally, another challenge is lack of reverse mapping. Techniques
-like PCA and auto-encoders \citep{article65} provide a way to map back
-from the low-dimensional space to the high-D space, facilitating data
-reconstruction. However, some NLDR methods, such as tSNE, don't have a
-specific way to reconstruct the original data from the low-dimensional
-space.
-
-In this article, mainly focus on five NLDR techniques. They are tSNE,
-UMAP, PHATE, TriMAP \citep{article02}, and Pairwise Controlled Manifold
-Approximation (PaCMAP) \citep{Yingfan2021}.
-
-Among these, tSNE \citep{Laurens2008} stands out for its ability to
-preserve pairwise distances. By minimizing the divergence between
-probability distributions in both high and low-dimensional spaces, tSNE
-effectively uncovers intricate structures and patterns within the data.
-Its application is widespread, particularly in tasks requiring the
-visualization of clusters and local relationships. However, achieving
-effective results requires careful consideration of hyperparameters,
-such as perplexity.
-
-UMAP \citep{Leland2018} is a useful technique for simplifying data while
-maintaining both local and overall structures. It builds a fuzzy
-topological view by considering nearby data points and then optimizes a
-simplified version to match that view. UMAP is known for working well
-with different scales of relationships in data and is efficient in
-handling large datasets. However, it's important to choose parameters
-like neighbors and minimum distance carefully, as they can affect the
-results.
-
-Furthermore, PHATE \citep{article03} is great for understanding how
-things develop, especially in single-cell genomics. It uses a heat
-diffusion process to capture relationships between data points, like
-points along a trajectory. While PHATE is excellent for revealing these
-developmental structures, it requires careful tuning of its parameters
-because of its specialized focus.
-
-Additionally, TriMAP \citep{article02} takes a special approach by
-creating a triangulated graph representation of the data. This method is
-good at understanding both local and global structures by treating the
-data as a network of triangles. TriMAP is powerful in capturing
-complicated structures, but it's important to choose parameters
-carefully, like deciding how many neighbors to consider.
-
-PaCMAP \citep{Yingfan2021} is different because it adds supervised
-learning to make a 2D representation while keeping the relationships
-between pairs of points. It builds a graph using distances between pairs
-and then makes the 2D representation better using a customizable loss
-function. What's special about PaCMAP is that it can use class labels or
-extra information to guide how it makes the 2D representation. This
-gives users a way to change how PaCMAP works to fit their needs better.
-
-\begin{figure}
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervis-1.pdf}
-
-}
-
-\caption{\label{fig-nldervis}2D layouts from different NLDR techniques
-applied the same data: (a) tSNE (perplexity = 27), (b) UMAP
-(n\_neighbors = 50), (c) PHATE (knn = 5), (d) TriMAP (n\_inliers = 5,
-n\_outliers = 4, n\_random = 3), and (e) PaCMAP (n\_neighbors = 10, init
-= random, MN\_ratio = 0.9, FP\_ratio = 2). Is there a best
-representation of the original data or are they all providing equivalent
-information?}
-
-\end{figure}%
-
-\subsection{Linear overviews using
-tours}\label{linear-overviews-using-tours}
-
-A tour is a powerful visualization technique used to explore the shape
-and global structure of high-dimensional data by generating a sequence
-of projections, typically into two dimensions. There are two main types
-of tours: the grand tour \citep{Asimov1985} and the guided tour
-\citep{article29}. A grand tour involves randomly selecting new
-orthonormal bases, enabling users to understand the structure by
-exploring the subspace of d-dimensional projections \citep{Asimov1985}.
-In contrast, a guided tour can be employed to generate a sequence of
-`interesting' projections based on an index function \citep{article29}.
-
-The process begins with the data matrix \(X\). It generates a sequence
-of \(p\) × \(d\) orthonormal projection matrices (bases) \(P_t\),
-usually \(d\) is one or two dimensions. For each pair of orthonormal
-bases \(P_t\) and \(P_{t+1}\), a geodesic path is interpolated to create
-smooth animation between projections. The resulting tour continuously
-visualizes the projected data \(Y_t\) = \(XP_t\) as it interpolates
-between successive bases.
-
-Furthermore, software like \textbf{langevitour} can visualize both types
-of tours, providing flexibility for exploring high-dimensional data with
-various objectives. In our context, use grand tour along with the model
-to observe how effectively the model captures the underlying structure
-of the data.
-
-\section{Methodology}\label{sec-methods}
-
-In this paper, we introduce a novel method to determine the most
-effective NLDR technique and the best hyperparameter choice that
-provides the most useful representation of high-D data. Our approach
-involves dividing the high-D dataset into two parts: a training set for
-constructing the model and a test set for generating predictive values
-and residuals. Our algorithm takes a 2D embedding data as the input and
-generate a tour that displays the high-D wireframe to overlay the data.
-The flow chart of the proposed algorithm is shown in
-Figure~\ref{fig-meth}. The algorithm consists of two main phases: (1)
-generating the model in the 2D space and (2) lifting the model into
-high-D space. The main steps of the algorithm are described in detail in
-this section using UMAP 2D embedding of the S-curve dataset. This
-dataset has seven dimensions, including four noise dimensions that were
-added to the original 3D data.
-
-\begin{figure}
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=1\textheight]{figures/workflow.png}
-
-}
-
-\caption{\label{fig-meth}The flow diagram shows the main steps of our
-algorithm. There are two basic phases, one to generate the model in the
-2D space, and other to map the model into the high-D space.}
-
-\end{figure}%
-
-\subsection{Preprocessing steps}\label{preprocessing-steps}
-
-To reduce computational complexity when applying NLDR techniques to
-high-D data and to reduce noise presence, PCA \citep[\citet{article68},
-\citet{article69}]{article67} is used as a preprocessing step. PCA
-involves identifying principal components that maximize variance. These
-components are then used as the high-D data for the algorithm.
-
-\subsection{Constructing the 2D model}\label{sec-construct2d}
-
-\textbf{Step 1: Scaling NLDR data}
-
-First, we prepare the 2D embedding data to fit within the bounds
-required for regular hexagonal binning. To achieve this, we implement
-two key scaling steps. Scale the first 2D embedding component to range
-between \(0\) and \(1\), ensuring that all data points fall within this
-normalized interval. Secondly, we scale the second 2D embedding
-component to range between \(0\) and \(y_{max}\) (see
-Equation~\ref{eq-equation3}).
-
-The calculation of \(y_{max}\) involves several steps. First, the aspect
-ratio (\(ar\)) is computed by dividing the range of the second 2D
-embedding component (\(r_2\)) by the range of the first 2D embedding
-component (\(r_1\)) (see Equation~\ref{eq-equation1}). Then, the hexagon
-ratio (\(hr\)) is determined by dividing the height of the hexagon
-(\(hb\)) by its width (\(wb\)) (see Equation~\ref{eq-equation2}).
-Finally, \(y_{max}\) is derived by taking the ceiling of
-\(\frac{ar}{hr}\) and multiplying it by \(hr\). This process ensures
-that \(y_{max}\) is an integer multiple of \(hr\), accommodating the
-grid layout of the hexagonal bins.
-
-\begin{equation}\phantomsection\label{eq-equation1}{
- ar = \frac{r_2}{r_1}
-}\end{equation}
-
-\begin{equation}\phantomsection\label{eq-equation2}{
- hr = \frac{hb}{wb}
-}\end{equation}
-
-\begin{equation}\phantomsection\label{eq-equation3}{
- y_{max} = \left\lceil\frac{ar}{hr}\right\rceil * hr
-}\end{equation}
-
-\textbf{Step 2: Hexagonating NLDR data}
-
-Hexagonating NLDR data (see Figure~\ref{fig-meth} Step 2) involves
-partitioning the NLDR data into hexagonal bins, a technique commonly
-referred to as hexagonal binning \citep[\citet{article66}]{Carr1987}.
-This method use a hexagonal grid to create a bivariate histogram that
-effectively visualizes the structure of high-D data. Hexagons, one of
-only three regular polygons capable of tessellating a plane
-\citep{Carr2013}, offer unique advantages due to their symmetry of
-nearest neighbors and maximal number of sides for such tessellations
-\citep{Dan2023}. This geometric property makes hexagons more efficient
-in covering the plane compared to other regular tessellations and
-reduces visual bias when displaying data densities \citep{Dan2023}.
-
-In our algorithm, we aim to conduct regular hexagonal binning, involving
-the computation of hexagonal grid configurations, the generation of the
-grid, and the assignment of the NLDR data points to hexagons.
-
-\textbf{(a) Computation of hexagonal grid configurations}
-
-In the computation of hexagonal bin configurations, there are mainly two
-configurations to consider: the number of bins along the x and y axes.
-The first step involves in determining the hexagonal size (\(s\)) and
-the buffer along the x and y axes (\(buffer_{x}\) and \(buffer_{y}\)).
-Here, the hexagonal size (\(s\)) represents the radius of the outer
-circle surrounding the hexagon. The buffer is important in expanding the
-boundary to accommodate potential outliers or edge cases.
-
-When computing the number of bins along the x and y axes (\(b_1\) and
-\(b_2\)), the process begins by calculating the range of the respective
-2D embedding component. After that, the range is adjusted to accommodate
-the buffer amount. Then, the spacing between hexagons is determined
-based on \(s\). Finally, the number of bins along the axis is computed
-by dividing the adjusted range by the spacing.
-
-For the computation of the number of bins along the x-axis (\(b_1\))
-(see Equation~\ref{eq-equation12}), the range of the first embedding
-component is defined as \(r_1\), while the buffer amount along the
-x-axis is denoted as \(buffer_{x}\), and the horizontal spacing is
-represented by \(h\) (see Equation~\ref{eq-equation11}). On the other
-hand, the number of bins along the y-axis (\(b_2\)) (see
-Equation~\ref{eq-equation14}) is computed based on the range of the
-second embedding component (\(r_2\)), the buffer amount along the y-axis
-(\(buffer_{y}\)), and the vertical spacing (\(v\)) (see
-Equation~\ref{eq-equation13}).
-
-\begin{equation}\phantomsection\label{eq-equation11}{
- h = \sqrt{3} * s
-}\end{equation}
-
-\begin{equation}\phantomsection\label{eq-equation12}{
- b_1 = \frac{r_1 + buffer_{x}}{h}
-}\end{equation}
-
-\begin{equation}\phantomsection\label{eq-equation13}{
-  v = 1.5 * s
-}\end{equation}
-
-\begin{equation}\phantomsection\label{eq-equation14}{
- b_2 = \frac{r_2 + buffer_{y}}{v}
-}\end{equation}
-
-\textbf{(b) Generation of the hexagonal grid}
-
-In this process, the first step involves generating centroids within the
-hexagonal grid. This begins by defining the starting coordinates for the
-hexagons. With the number of bins computed along the x and y axes, along
-with the horizontal and vertical spacing, the centroids are iteratively
-computed starting from the hexagonal starting coordinates.
-
-Once the centroids for the hexagonal grid are obtained, the next step is
-to compute the hexagonal coordinates. For example, if the centroid of a
-hexagonal bin is defined as \((C_x, C_y)\), the hexagonal coordinates
-can be computed using \(d_x\) (see Equation~\ref{eq-equation15}) and
-\(d_y\) (see Equation~\ref{eq-equation16}) (see
-Figure~\ref{fig-hexcoord}).
-
-\begin{equation}\phantomsection\label{eq-equation15}{
-d_x = \frac{h}{2}
-}\end{equation}
-
-\begin{equation}\phantomsection\label{eq-equation16}{
-d_y = \frac{2* v * 1.15}{\sqrt{3}} 
-}\end{equation}
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=1\textheight]{figures/hex_coord.png}
-
-}
-
-\caption{\label{fig-hexcoord}The hexagonal coordinates computed for the
-centroid \((C_x, C_y)\).}
-
-\end{figure}%
-
-\textbf{(c) Assignment of the NLDR data points to hexagons}
-
-After obtaining the centroids of the hexagonal bins for the entire grid,
-the process of assigning NLDR points to hexagons involves determining
-the nearest hexagonal bin for each NLDR point using the 2D Euclidean
-distance. Then, the hexagonal ID of the nearest hexagon is assigned to
-the corresponding NLDR points.
-
-\textbf{Step 3: Obtaining bin centroids or bin means}
-
-In the previous step, the algorithm clusters the 2D embedding data into
-hexagons. Following this, in this step, the bin centroids or bin means
-(see Figure~\ref{fig-meth} Step 3) are obtained \citep{Carr2013}.
-
-The bin centroid (\(C_k^{(2)}\)) for a \(k^{th}\) hexagon with hexagonal
-grid coordinates \((h^{k}x_{i}, h^{k}y_{i})\), where \(i = 1 \dots 6\)
-can be defined as:
-
-\begin{equation}\phantomsection\label{eq-equation4}{
-C_k^{(2)} = (C_{ky_1}, C_{ky_2}) = \left(\frac{\sum_{i=1}^{6} h^{k}x_{i}}{6}, \frac{\sum_{i=1}^{6} h^{k}y_{i}}{6}\right).
-}\end{equation}
-
-Also, the bin mean (\(C_k^{(2)}\)) is defined as the mean of the data
-points within the \(k^{th}\) hexagon (see Equation~\ref{eq-equation5}).
-
-\begin{equation}\phantomsection\label{eq-equation5}{
-C_k^{(2)} = (C_{ky_1}, C_{ky_2}) = \left(\frac{1}{n_k} \sum_{i=1}^{n_k} y_{1i}, \frac{1}{n_k} \sum_{i=1}^{n_k} y_{2i}\right),
-}\end{equation}
-
-where \(n_k\) is the number of data points within the hexagon,
-\(y_{1i}\) and \(y_{2i}\) are the \(x\) and \(y\) coordinates of the
-\(i^{th}\) data point within the hexagon.
-
-\textbf{Step 4: Triangulating bin centroids or bin means}
-
-In this step, the algorithm proceeds to triangulate the hexagonal bin
-centroids or bin means (see Figure~\ref{fig-meth} Step 4). Triangulation
-is a fundamental process in computational geometry and computer graphics
-that involves dividing a set of points in a given space into
-interconnected triangles \citep{article30}. One common algorithm used
-for triangulation is Delaunay triangulation
-\citep[\citet{article54}]{article26}, where points are connected in a
-way that maximizes the minimum angles of the resulting triangles,
-leading to a more regular and well-conditioned triangulation.
-
-Delaunay triangulation can be defined as follows:
-
-Let \(C^{(2)} = \{C_1^{(2)}, C_2^{(2)}, ..., C_m^{(2)}\}\) be a set of
-\(m\) bin centroids or bin means in the plane. Delaunay triangulation of
-\(C^{(2)}\), denoted as \(DT(C^{(2)})\), is a triangulation of the
-convex hull of \(C^{(2)}\) such that the circumcircle of every triangle
-in the triangulation contains no other points from \(C^{(2)}\).
-
-Given that the hexagons are regular, the resulting triangles will mostly
-be equilateral.
-
-\subsection{Lifting the model into high
-dimensions}\label{lifting-the-model-into-high-dimensions}
-
-\subsubsection{Lifting the triangular mesh points into high
-dimensions}\label{lifting-the-triangular-mesh-points-into-high-dimensions}
-
-Consider \(f: \mathbb{R}^p \rightarrow \mathbb{R}^2\) be a function that
-maps the high-D data (\(X_{n \times p}\)) to its NLDR equivalent
-(\(Y_{n \times d}\)). Then, let
-\(g: \mathbb{R}^2 \rightarrow \mathbb{R}^2\) be a function that maps
-each 2D embedding point to its closest centroid (\(C^{(2)}\)). It
-follows that \(f(g(x))\) maps the high-D points \(x\) to the centroid in
-2D (\(C_k^{(2)}\)). Also, define a function
-\(v: \mathbb{R}^2 \rightarrow \mathbb{R}^p\) maps the 2D centroid
-(\(C^{(2)}\)) to the high-D mean of the points (\(C^{(p)}\)) in the
-hexagon.
-
-The high-D mean of all the points in \(k^{th}\) hexagon by
-
-\begin{equation}\phantomsection\label{eq-equation6}{
-C_k^{(p)} = (C_{kx_1}, ..., C_{kx_p}) = \left(\frac{1}{n_k} \sum_{i=1}^{n_k} x_{1i}, \frac{1}{n_k} \sum_{i=1}^{n_k} x_{2i}, \dots ,\frac{1}{n_k} \sum_{i=1}^{n_k} x_{pi}\right).
-}\end{equation}
-
-Therefore,
-
-\begin{equation}\phantomsection\label{eq-equation7}{
-v(C_k^{(2)}) = C_k^{(p)}.
-}\end{equation}
-
-Therefore, \(f(g(x))\) gives the 2D centroid associated with high-D
-points \(x\), and \(v(C_k^{(2)})\) gives the high-D centroid associated
-with 2D point \(C_k^{(2)}\). Thus, \(v(f(g(x)))\) gives the high-D
-centroid (\(C_k^{(p)}\)) associated with the 2D embedding of the points
-\(x\).
-
-\subsubsection{Lifting the 2D triangular mesh into high
-dimensions}\label{lifting-the-2d-triangular-mesh-into-high-dimensions}
-
-As described in Step 4 of Section~\ref{sec-construct2d}, during the
-triangulation process in 2D space, vertices are identified to form
-edges. With the knowledge of the high-D mappings for the 2D hexagonal
-bins, the vertices connected in 2D are also connected in high-D (see
-video linked in Figure~\ref{fig-wkhighD}).
-
-\begin{figure}
-
-\centering{
-
-\includegraphics{paper_files/figure-pdf/fig-wkhighD-1.pdf}
-
-}
-
-\caption{\label{fig-wkhighD}How the 2D model lift into high dimensions?
-(a) visualize the points and the hexagonal bin centroids related
-\(7^{th}\) and \(12^{th}\) hexagons, (b) visualization of the edge
-connected the \(7^{th}\) and \(12^{th}\) hexagons (colored in red) in
-the triangular mesh. A video of tour animation is available at
-\url{https://youtu.be/hxU91xNTJL0}.}
-
-\end{figure}%
-
-\subsection{Tuning the model}\label{tuning-the-model}
-
-The performance and robustness of our model depend on four key
-parameters: (i) the total number of bins (\(b\)), (ii) a benchmark value
-used to remove low-density hexagons, (iii) a benchmark value used to
-remove long edges, and (iv) starting point of the hexagonal grid.
-However, there is no analytical formula to calculate an appropriate
-value for these parameters. The selection of these parameter values
-depends on the model performance computed by Mean Squared Error (MSE)
-(see Section~\ref{sec-goodfit}).
-
-\subsubsection{Total number of bins}\label{total-number-of-bins}
-
-The number of hexagonal bins in the hexagonal grid has a considerable
-impact on the construction of the 2D model, serving as the initial step
-in building the 2D model. The chosen total number of bins must
-effectively capture the structure of the NLDR data. If the number of
-bins is too low, the model may not be able to capture the structure of
-the NLDR data effectively (see Figure~\ref{fig-binsize} (a)), while if
-there are too many bins, it may result in over-fitting the individual
-points of the NLDR data (see Figure~\ref{fig-binsize} (c)). Therefore,
-it is important to determine an appropriate number of bins to construct
-an effective model.
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics{paper_files/figure-pdf/fig-binsize-1.pdf}
-
-}
-
-\caption{\label{fig-binsize}Hexbin plots from different number of bins
-for the \textbf{UMAP} embeddings of \textbf{S-curve} training data: (a)
-b = 12 (3, 4), s = 0.62, (b) b = 84 (6, 14), s = 0.13, and (c) b = 336
-(12, 28), s = 0.06. The hexbins are colored based on the density of
-points, with darker colors indicating higher point density and yellow
-color representing lower point density within each bin. What is the
-number of bins that would be effective in representing low-dimensional
-data?}
-
-\end{figure}%
-
-\begin{equation}\phantomsection\label{eq-equation8}{
-b = b_1 \times b_2
-}\end{equation}
-
-Furthermore, the total number of bins is calculated based on the number
-of bins along the x-axis and y-axis, as shown in
-Equation~\ref{eq-equation8}. To determine the effective total number of
-bins, candidate values are selected based on the range between the
-minimum and approximate maximum number of bins along the x and y axes.
-The minimum number of bins along each axis is set to \(1\), while the
-maximum number is estimated by taking the square root of the NLDR data
-points. The analysis evaluates the MSE across varying total number of
-bins within this range, covering the minimum to maximum values along
-both axes (see Figure~\ref{fig-diagnosticpltScurve}).
-
-\begin{figure}
-
-\centering{
-
-\includegraphics{paper_files/figure-pdf/fig-diagnosticpltScurve-1.pdf}
-
-}
-
-\caption{\label{fig-diagnosticpltScurve}Goodness of fit statistics from
-UMAP applied to training S-curve dataset. What is the effective number
-of bins in each NLDR technique to create a 2D model? The MSE plot have a
-steep slope at the beginning, indicating that a smaller number of bins
-causes a larger amount of error. Then, the slope gradually declines or
-level off, indicating that a higher number of bins generates a smaller
-error. Using the elbow method, when the total number of bins is set to
-\(84\), the slope of the Mean Squared Error (MSE) plot experiences a
-sudden and noticeable change, resembling an elbow-like shape. This point
-indicates that adding less bins does not enough to capture the data
-structure.}
-
-\end{figure}%
-
-\subsubsection{Benchmark value to remove low-density
-hexagons}\label{benchmark-value-to-remove-low-density-hexagons}
-
-After establishing the hexagonal grid with an appropriate number of
-bins, some hexagonal bins may have few or no data points within them
-(see Figure~\ref{fig-rmlowdenshex} (a)). To ensure comprehensive
-coverage of the NLDR data, it is necessary to select hexagonal bins with
-a considerable number of data points. This involves calculating the
-number of points within each hexagon. Then, the standard count is
-computed by dividing the number of points within each hexagon by the
-maximum number of points in the grid (see Equation~\ref{eq-equationp2}).
-Next, bins with a standard count less than a certain benchmark value are
-removed (see Figure~\ref{fig-rmlowdenshex} (b)). The following steps
-will help in determining a suitable value for removing low-density
-hexagons:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Plot the distribution of the standardized counts (see
-  Figure~\ref{fig-stdctsScurve}).
-\item
-  Examine the distribution of counts.
-\item
-  Select the first quartile value if the distribution is skewed.
-\end{enumerate}
-
-\begin{equation}\phantomsection\label{eq-equationp2}{
-\text{standard count} = \frac{\text{count}}{\text{max count}} 
-}\end{equation}
-
-Furthermore, selecting the benchmark value for removing low-density
-hexagons is important. Removing unnecessary bins may lead to the
-formation of long edges and an uneven 2D model. Hence, rather than
-solely relying on the benchmark value to identify hexagons for removal,
-it's essential to consider the standard number of points in the
-neighboring hexagons of the identified low-density ones (see
-Figure~\ref{fig-rmlowdenshex} (c)). If neighboring bins also show low
-counts, only those bins will be removed. The remaining bins are used to
-construct the 2D model.
-
-The benchmark value for removing low-density hexagons ranges between
-\(0\) and \(1\). When analyzing how these benchmark values influence
-model performance, it's essential to observe the change in Mean Squared
-Error (MSE) as the benchmark value increases (see
-Figure~\ref{fig-diagnosticpltScurvelwd}). The MSE shows a gradual
-increase as the benchmark value progresses from \(0\) to \(1\).
-Evaluating this rate of increase is important. If the increment is not
-considerable, the decision might lean towards retaining low-density
-hexagons.
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics{paper_files/figure-pdf/fig-stdctsScurve-1.pdf}
-
-}
-
-\caption{\label{fig-stdctsScurve}Distribution of standardize counts by
-hexagons.}
-
-\end{figure}%
-
-\begin{figure}
-
-\centering{
-
-\includegraphics{paper_files/figure-pdf/fig-diagnosticpltScurvelwd-1.pdf}
-
-}
-
-\caption{\label{fig-diagnosticpltScurvelwd}Goodness of fit statistics
-from UMAP applied to training S-curve dataset with different benchmark
-values to remove the low-density heaxgons. What is the effective
-benchmark value to remove the low-density heaxgons? The MSE plot have a
-steep slope at the beginning, indicating that a smaller benchmark causes
-a small amount of error. Then, the slope gradually increases or level
-up, indicating that a higher number of benchmark values generates a
-higher error. Using the reverse elbow method, when the benchmark value
-is set to 0.2, the slope of the Mean Squared Error (MSE) plot
-experiences a sudden and noticeable change, resembling an elbow-like
-shape. This point indicates that higher benchmark values remove the
-necessary bins as well which lead to distruct the 2D structure.}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics{paper_files/figure-pdf/fig-rmlowdenshex-1.pdf}
-
-}
-
-\caption{\label{fig-rmlowdenshex}(a) Hexbin plot with colored hexbins
-based on point density, (b) Identification of low-density hexagons using
-a benchmark value of \(0.2\), and (c) Identification of low-density
-hexagons considering neighboring bins.}
-
-\end{figure}%
-
-\subsubsection{Benchmark value for removing long
-edges}\label{benchmark-value-for-removing-long-edges}
-
-Creating a smooth 2D representation (see Figure~\ref{fig-modelScurve}
-(c)) requires removing edges that connect distant bin centroids within
-the triangular mesh (see Figure~\ref{fig-modelScurve} (b)). These edges
-only exist in the 2D model and do not extend into higher dimensions,
-ensuring that their removal does not impact the model in higher
-dimensions. While specific criteria for determining the benchmark value
-to remove long edges do not exist, the following steps provide an
-approach to identifying a suitable threshold:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Plot the distribution of the 2D Euclidean distances (see
-  Figure~\ref{fig-distScurve}).
-\item
-  Identify the largest difference between consecutive distance values.
-\item
-  Take the largest distance value corresponding to this difference as
-  the benchmark value.
-\end{enumerate}
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics{paper_files/figure-pdf/fig-distScurve-1.pdf}
-
-}
-
-\caption{\label{fig-distScurve}Distribution of 2D Euclidean distances
-between bin centroids of the triangular mesh generated with S-curve UMAP
-data.}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics{paper_files/figure-pdf/fig-modelScurve-1.pdf}
-
-}
-
-\caption{\label{fig-modelScurve}(a) Full hexagonal grid with bin
-centroids, (b) Model constructed in 2D with long edges, and (c) Model
-constructed in 2D after removing the long edges.}
-
-\end{figure}%
-
-\subsubsection{Starting point of the hexagonal
-grid}\label{starting-point-of-the-hexagonal-grid}
-
-Hexagonal binning involves tessellating the \(xy\) plane over the set
-defined by the range of \(x\) and \(y\) \citep{Dan2023}. Therefore, the
-starting point of the hexagonal grid is important parameter to consider.
-Having different starting point results in shifting the hexagonal grid
-and bin centroids (see Figure~\ref{fig-scurveshifthexgridsexp}). With a
-specific total number of bins (which corresponds to a specific range
-along the x and y axes), it's essential to explore how different
-starting points affect the model's performance (see
-Figure~\ref{fig-mseScurvestartpoint}). This is because different
-starting points result in varying distributions of non-empty bins (see
-Figure~\ref{fig-scurveshifthexgrids}), impacting the model.
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics{paper_files/figure-pdf/fig-mseScurvestartpoint-1.pdf}
-
-}
-
-\caption{\label{fig-mseScurvestartpoint}Goodness of fit statistics from
-UMAP applied to training S-curve dataset with different benchmark values
-to remove the low-density heaxgons. What is the effective benchmark
-value to remove the low-density heaxgons? The MSE plot have a steep
-slope at the beginning, indicating that a smaller benchmark causes a
-small amount of error. Then, the slope gradually increases or level up,
-indicating that a higher number of benchmark values generates a higher
-error. Using the reverse elbow method, when the benchmark value is set
-to 0.2, the slope of the Mean Squared Error (MSE) plot experiences a
-sudden and noticeable change, resembling an elbow-like shape. This point
-indicates that higher benchmark values remove the necessary bins as well
-which lead to distruct the 2D structure.}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics{paper_files/figure-pdf/fig-scurveshifthexgridsexp-1.pdf}
-
-}
-
-\caption{\label{fig-scurveshifthexgridsexp}(a) Visualization of bin
-centroids and (b) hexagonal grid, showcasing the configuration before
-(colored in gray) and after (colored in light orange) shifting the
-starting point by an identical amount applied in both x and y
-directions, with a shift value of \(0.537285\).}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics{paper_files/figure-pdf/fig-scurveshifthexgrids-1.pdf}
-
-}
-
-\caption{\label{fig-scurveshifthexgrids}Hexbin plots shows the
-distribution of points before and after shifting the starting point of
-the hexagonal grid generated for UMAP applied to the training S-curve
-dataset. (a) Hexbin plot before shifting, and (b) Hexbin plot after
-shifting (identical shift applied in both x and y directions, with a
-shift value of \(0.537285\)).}
-
-\end{figure}%
-
-\subsection{Model summaries}\label{sec-summary}
-
-\subsubsection{Predicted values and
-residuals}\label{predicted-values-and-residuals}
-
-The prediction approach involves performing the K-nearest neighbors
-(KNN) algorithm for an unsupervised classification problem. First, the
-nearest high-D model point is identified for a given new high-D point.
-Then, the corresponding 2D centroid mapping for the identified high-D
-model point is determined. Finally, the coordinates of this 2D centroid
-are used as the predicted 2D embedding for the new high-D data point.
-This step is particularly valuable due to the limitations of some NLDR
-techniques, like tSNE, which don't provide a straightforward method for
-prediction. As a result, our approach offers a solution that capable of
-generating predicted 2D embedding regardless of the NLDR technique
-employed, effectively addressing this functional gap.
-
-Residuals are essential for evaluating the accuracy of representing
-high-D points by the high-D mapping of 2D bin centroids. To measure this
-accuracy, an error metric is introduced, quantifying the sum of squared
-differences between the high-D data (\(x_{ij}\)) and the high-D mapping
-of the 2D bin centroid data (\(C_{x_ij}\)) across all observations and
-dimensions (see Equation~\ref{eq-equation10}).
-
-\begin{equation}\phantomsection\label{eq-equation10}{
-\text{Error} = \sum_{j = 1}^{n}\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2
-}\end{equation}
-
-Here, \(n\) represents the number of observations, \(p\) represents the
-dimensions of high-D data, \(x_{ij}\) is the high-D data, and
-\(C_{x_ij}\) is the high-D mapping of the 2D bin centroid.
-
-\subsubsection{Goodness of fit statistics}\label{sec-goodfit}
-
-To assess how well our method captures and represents the underlying
-structure of the high-D data, Mean Squared Error (MSE) is used. When
-computing MSE, total model error (see Section~\ref{sec-summary}) is
-divided by the number of observations to make it as a mean value (see
-Equation~\ref{eq-equation9}).
-
-\begin{equation}\phantomsection\label{eq-equation9}{
-\text{MSE} = \sum_{j = 1}^{n} \frac{\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2}{n}
-}\end{equation}
-
-\subsection{Simulated data example}\label{sec-simpleex}
-
-In this section, the effectiveness of the algorithm is described using a
-simulated dataset. The dataset consists of five spherical Gaussian
-clusters in 4-\(d\), with each cluster containing an equal number of
-points and the same within-cluster variation.
-
-In the 2D layouts generated by various NLDR techniques, as shown in
-Figure~\ref{fig-nldervis5Gau}, five distinct clusters are observable,
-except for PHATE. In tSNE (see Figure~\ref{fig-nldervis5Gau} (a)), these
-clusters appear closely. UMAP arranges all clusters in a parallel
-manner, with three aligned in one line and the other two in a separate
-line (see Figure~\ref{fig-nldervis5Gau} (b)). In contrast, PHATE shows
-two closely positioned clusters and three more distant ones (see
-Figure~\ref{fig-nldervis5Gau} (c)). PaCMAP shows one central cluster and
-the remaining four spread out in different directions (see
-Figure~\ref{fig-nldervis5Gau} (e)). Finally, in TriMAP, two clusters are
-close, though not as tightly as PHATE, while the other three are
-well-separated (see Figure~\ref{fig-nldervis5Gau} (d)).
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervis5Gau-1.pdf}
-
-}
-
-\caption{\label{fig-nldervis5Gau}2D layouts from different NLDR
-techniques applied the same data: (a) tSNE (perplexity = 61), (b) UMAP
-(n\_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n\_inliers = 5,
-n\_outliers = 4, n\_random = 3), and (e) PaCMAP (n\_neighbors = 10, init
-= random, MN\_ratio = 0.9, FP\_ratio = 2). Is there a best
-representation of the original data or are they all providing equivalent
-information?}
-
-\end{figure}%
-
-To investigate which is the best representation to visualize the
-original data or all NLDR methods provide equivalent information, we
-visualize the model-in-data space. Models from all NLDR methods show
-five well-separated clusters (see Figure~\ref{fig-gau1-sc},
-Figure~\ref{fig-gau2-sc}, Figure~\ref{fig-gau3-sc},
-Figure~\ref{fig-gau4-sc}, and Figure~\ref{fig-gau5-sc}). This suggests
-that for the five Gaussian cluster dataset, all NLDR methods effectively
-preserve the global structure. tSNE, UMAP, and PHATE display clusters
-with varying densities, indicating their ability to capture
-within-cluster variation (see Figure~\ref{fig-gau1-sc},
-Figure~\ref{fig-gau2-sc}, and Figure~\ref{fig-gau3-sc}). On the other
-hand, both TriMAP and PaCMAP show clusters with flat surfaces,
-suggesting a failure to capture within-cluster variation (see
-Figure~\ref{fig-gau4-sc} and Figure~\ref{fig-gau5-sc}). Therefore,
-TriMAP and PaCMAP do not capture the local structure as effectively as
-other methods.
-
-\begin{figure}
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-modelfiveGau-1.pdf}
-
-}
-
-\caption{\label{fig-modelfiveGau}Is there a best model to represent the
-original data in 2D space or are they all providing equivalent
-information?, (a) Model generated with 21 and 28 bins along the x and y
-axes, respectively, using a hexagonal size of 0.03 in the 2D space with
-tSNE, (b) Model generated with 60 and 79 bins along the x and y axes,
-respectively, using a hexagonal size of 0.01 in the 2D space with UMAP,
-(c) Model generated with 60 and 1927 bins along the x and y axes,
-respectively, using a hexagonal size of 0.01 in the 2D space with PHATE,
-(d) Model generated with 46 and 61 bins along the x and y axes,
-respectively, using a hexagonal size of 0.013 in the 2D space with
-TriMAP, and (e) Model generated with 31 and 40 bins along the x and y
-axes, respectively, using a hexagonal size of 0.02 in the 2D space with
-PaCMAP.}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_tsne_1.png}
-
-}
-
-\subcaption{\label{fig-gau1-sc1}}
-
-\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_tsne_2.png}
-
-}
-
-\subcaption{\label{fig-gau1-sc2}}
-
-\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_tsne_3.png}
-
-}
-
-\subcaption{\label{fig-gau1-sc3}}
-
-\end{minipage}%
-
-\caption{\label{fig-gau1-sc}Screen shots of the \textbf{langevitour} of
-the five Gaussian clusters dataset, shows the model with tSNE in high-D,
-a video of the tour animation is available at
-(\url{https://youtu.be/oQxEb4wRdHI}).}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_umap_1.png}
-
-}
-
-\subcaption{\label{fig-gau2-sc1}}
-
-\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_umap_2.png}
-
-}
-
-\subcaption{\label{fig-gau2-sc2}}
-
-\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_umap_3.png}
-
-}
-
-\subcaption{\label{fig-gau2-sc3}}
-
-\end{minipage}%
-
-\caption{\label{fig-gau2-sc}Screen shots of the \textbf{langevitour} of
-the five Gaussian clusters dataset, shows the model with UMAP in high-D,
-a video of the tour animation is available at
-(\url{https://youtu.be/JW49csPpDx4}).}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_phate_1.png}
-
-}
-
-\subcaption{\label{fig-gau3-sc1}}
-
-\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_phate_2.png}
-
-}
-
-\subcaption{\label{fig-gau3-sc2}}
-
-\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_phate_3.png}
-
-}
-
-\subcaption{\label{fig-gau3-sc3}}
-
-\end{minipage}%
-
-\caption{\label{fig-gau3-sc}Screen shots of the \textbf{langevitour} of
-the five Gaussian clusters dataset, shows the model with PHATE in
-high-D, a video of the tour animation is available at
-(\url{https://youtu.be/xqhRODt8Z-o}).}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_trimap_1.png}
-
-}
-
-\subcaption{\label{fig-gau4-sc1}}
-
-\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_trimap_2.png}
-
-}
-
-\subcaption{\label{fig-gau4-sc2}}
-
-\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_trimap_3.png}
-
-}
-
-\subcaption{\label{fig-gau4-sc3}}
-
-\end{minipage}%
-
-\caption{\label{fig-gau4-sc}Screen shots of the \textbf{langevitour} of
-the five Gaussian clusters dataset, shows the model with TriMAP in
-high-D, a video of the tour animation is available at
-(\url{https://youtu.be/X-uHSB4VoZ0}).}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_pacmap_1.png}
-
-}
-
-\subcaption{\label{fig-gau5-sc1}}
-
-\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_pacmap_2.png}
-
-}
-
-\subcaption{\label{fig-gau5-sc2}}
-
-\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_pacmap_3.png}
-
-}
-
-\subcaption{\label{fig-gau5-sc3}}
-
-\end{minipage}%
-
-\caption{\label{fig-gau5-sc}Screen shots of the \textbf{langevitour} of
-the five Gaussian clusters dataset, shows the model with PaCMAP in
-high-D, a video of the tour animation is available at
-(\url{https://youtu.be/UxvU8HQd7JM}).}
-
-\end{figure}%
-
-\section{Applications}\label{sec-applications}
-
-\subsection{Single-cell RNA-seq data of
-human}\label{single-cell-rna-seq-data-of-human}
-
-In the field of single-cell studies, a common analytical task involves
-clustering to identify groups of cells with similar expression profiles.
-Analysts often turn to NLDR techniques to verify and identify these
-clusters and explore developmental trajectories. To illustrate the
-importance of NLDR techniques and parameter selection in identifying
-clusters, Human Peripheral Blood Mononclear Cells (PBMC3k) dataset
-\citep{chen2023} is used. In a study by \citet{chen2023}, this dataset
-was used to demonstrate how UMAP represents clusters (see
-Figure~\ref{fig-umapauthor}). As shown in Figure~\ref{fig-umapauthor},
-there are three distinct and well-separated clusters.
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics{paper_files/figure-pdf/fig-umapauthor-1.pdf}
-
-}
-
-\caption{\label{fig-umapauthor}2D layout from UMAP (n\_neighbors = 30,
-min\_dist = 0.3) applied for the PBMC3k dataset. Is this a best
-representation of the original data?}
-
-\end{figure}%
-
-To determine whether the UMAP representation with the parameter choice
-suggested by \citet{chen2023} preserves the original data structure, we
-visualize the model constructed with UMAP overlaid on the high-D data
-(model-in-data space). The figures in Figure~\ref{fig-pbmc1-sc} show
-three well-separated clusters, indicating that the suggested UMAP
-representation preserves the global structure (see
-Figure~\ref{fig-umapauthor}). However, as shown in
-Figure~\ref{fig-pbmc1-sc}, these clusters are close to each other in
-high-D space. Also, nonlinear continuity patterns and high-density
-patches within the clusters are observed (see
-Figure~\ref{fig-pbmc1-sc}). Therefore, the suggested UMAP representation
-(see Figure~\ref{fig-umapauthor}) does not accurately preserve the local
-structure of the PBMC3k dataset.
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics{paper_files/figure-pdf/fig-nldervisPBMCUMAP-1.pdf}
-
-}
-
-\caption{\label{fig-nldervisPBMCUMAP}(a) Model generated with 21 and 28
-bins along x and y axes with hexagonal size 0.03 in the 2D space
-overlaid on UMAP data, and (b) high-D model error in model space.}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_1.png}
-
-}
-
-\subcaption{\label{fig-pbmc1-sc1}}
-
-\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_2.png}
-
-}
-
-\subcaption{\label{fig-pbmc1-sc2}}
-
-\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_3.png}
-
-}
-
-\subcaption{\label{fig-pbmc1-sc3}}
-
-\end{minipage}%
-
-\caption{\label{fig-pbmc1-sc}Screen shots of the \textbf{langevitour} of
-the PBMC3k data set, shows the model in high-D, a video of the tour
-animation is available at (\url{https://youtu.be/Gnhh4hfsbbg}).}
-
-\end{figure}%
-
-To find the best NLDR representation for PBMC3k dataset, we observe the
-absolute error for different number of non-empty bins and select the one
-with the lowest error. As shown in Figure~\ref{fig-abserrorPBMC}, tSNE
-with a perplexity set to \(51\), has the lowest error. Therefore, tSNE
-with a perplexity of \(51\) is used as the best representation for
-PBMC3k dataset.
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=1\textheight]{paper_files/figure-pdf/fig-abserrorPBMC-1.pdf}
-
-}
-
-\caption{\label{fig-abserrorPBMC}Absolute error from UMAP and tSNE
-applied to training PBMC3k dataset with diffferent parameter choices.
-What is the best parameter choice to create the model? The residual plot
-have a steep slope at the beginning, indicating that a smaller number of
-non-empty bins causes a larger amount of error. Then, the slope
-gradually declines or level off, indicating that a higher number of
-non-empty bins generates a smaller error. Using the elbow method, it was
-observed that when the number of non-empty bins is set to 439, the
-lowest error occurred with the parameters perplexity: 51.}
-
-\end{figure}%
-
-As shown in Figure~\ref{fig-tsnesuggest}, there are three well-separated
-clusters, although they are located close to each other. Additionally,
-nonlinear structures can be observed within the clusters (see
-Figure~\ref{fig-tsnesuggest}). Hence, the data structure that UMAP
-failed to capture for the PBMC3k dataset is successfully captured by
-tSNE.
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics{paper_files/figure-pdf/fig-tsnesuggest-1.pdf}
-
-}
-
-\caption{\label{fig-tsnesuggest}2D layout from tSNE (perplexity = 51)
-applied for the PBMC3k dataset. Is this a best representation of the
-original data?}
-
-\end{figure}%
-
-Next, the model is fitted for tSNE, and the resulting model is
-visualized in the data space. The model shows some quirks, as shown in
-Figure~\ref{fig-pbmc2-sc}. The large cluster is divided into three
-subclusters, which are in close to each other.
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics{paper_files/figure-pdf/fig-nldervisPBMCtSNE-1.pdf}
-
-}
-
-\caption{\label{fig-nldervisPBMCtSNE}(a) Model generated with 31 and 40
-bins along x and y axes with hexagonal size 0.02 in the 2D space
-overlaid on tSNE data, and (b) high-D model error in model space.}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_4.png}
-
-}
-
-\subcaption{\label{fig-pbmc2-sc4}}
-
-\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_5.png}
-
-}
-
-\subcaption{\label{fig-pbmc2-sc5}}
-
-\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_6.png}
-
-}
-
-\subcaption{\label{fig-pbmc2-sc6}}
-
-\end{minipage}%
-
-\caption{\label{fig-pbmc2-sc}Screen shots of the \textbf{langevitour} of
-the PBMC3k data set, shows the model in high-D, a video of the tour
-animation is available at (\url{https://youtu.be/R0EroJcGbas}).}
-
-\end{figure}%
-
-\subsection{Handwritten digits}\label{handwritten-digits}
-
-The MNIST dataset consists of grayscale images of handwritten digits
-\citep{lecun2010}. \citet{Yingfan2021} used this dataset to demonstrate
-how the PaCMAP preserves local structure. We selected the 2D embedding
-of PaCMAP for the handwritten digit 1. As shown in
-Figure~\ref{fig-pacmapauthor}, we observe a nonlinear structure present
-in 2D space. Furthermore, the angle of digit 1 images varies along the
-nonlinear structure (see Figure~\ref{fig-sampleimg}). To answer the
-following questions,
+\section{Method}\label{method}
 
@@
 \item
-  Is the original data structure of the handwritten digit 1 dataset
-  nonlinear?
+  Create a representation of the model
 \item
-  Where does PaCMAP preserve the original data structure?
+  Algorithm in 2D
+
+  \begin{itemize}
+  \tightlist
   \item
-  Where does PaCMAP show quirks?
+    Parameters
+  \item
+    Tuning
+  \end{itemize}
+\item
+  Showing model in high-d
+\item
+  What is learned about simulated examples
 \end{itemize}
 
-we visualize the model constructed with PaCMAP overlaid on data in
-high-D space (model-in-data space).
+\section{Applications}\label{applications}
 
-According to Figure~\ref{fig-mnist1-sc1}, the nonlinear continuity
-structure observed in the 2D representation of PaCMAP is also visible
-when visualizing the model overlaid on the data space. This provides
-evidence that the model accurately captures the structure of the high-D
-data. Also, the model has a twisted pattern within the nonlinear
-structure, which is an addition pattern of data that cannot be seen in
-the 2D representation (see Figure~\ref{fig-mnist1-sc2}).
+\subsection{pbmc}\label{pbmc}
 
-In order to visually assess the preservation of local structure, it's
-important to consider the edges in the model. In a model that
-effectively preserves local structure, there shouldn't be any long edges
-in high-D space. However, as shown in Figure~\ref{fig-mnist1-sc3}, some
-long edges exist in high-D space that are not identified as long edges
-in the 2D representation.
+\begin{verbatim}
+ - NLDR view used to illustrate clusters
+ - Use our method to assess is it a reasonable representation
+ - Demonstrate that it is not
+ - Illustrate how to use out method to get a better representation
 
-\begin{figure}[H]
+\end{verbatim}
 
-\centering{
+\subsection{digits: 1}\label{digits-1}
 
-\includegraphics{paper_files/figure-pdf/fig-pacmapauthor-1.pdf}
+\begin{verbatim}
+ - NLDR is used to illustrate different ways 1's are drawn
+ - Use our method to assess is it a reasonable representation
+ - Demonstrate that it is, except for the anomalies
 
-}
+\end{verbatim}
 
-\caption{\label{fig-pacmapauthor}(a) 2D layout from PaCMAP
-(n\_components=2, n\_neighbors=10, init=random, MN\_ratio=0.9,
-FP\_ratio=2.0) applied for the digit 1 of the MNIST dataset. Is this a
-best representation of the original data?}
+\section{Discussion}\label{discussion}
 
-\end{figure}%
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-sampleimg-1.pdf}
-
-}
-
-\caption{\label{fig-sampleimg}Images of the handwritten digit 1 are
-distributed along the right-to-left of first embedding axis of the
-PaCMAP nonlinear structure.}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics{paper_files/figure-pdf/fig-nldervisMNISTPACMAP-1.pdf}
-
-}
-
-\caption{\label{fig-nldervisMNISTPACMAP}(a) Model generated with 12 and
-15 bins along x and y axes with hexagonal size 0.06 in the 2D space
-overlaid on PaCMAP data, and (b) high-D model error in model space.}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/mnist/sc_1.png}
-
-}
-
-\subcaption{\label{fig-mnist1-sc1}}
-
-\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/mnist/sc_2.png}
-
-}
-
-\subcaption{\label{fig-mnist1-sc2}}
-
-\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/mnist/sc_3.png}
-
-}
-
-\subcaption{\label{fig-mnist1-sc3}}
-
-\end{minipage}%
-
-\caption{\label{fig-mnist1-sc}Screen shots of the \textbf{langevitour}
-of the MNIST digit 1 data set, shows the model-in-data space, a video of
-the tour animation is available at
-(\url{https://youtu.be/zq2GM9qvUNA}).}
-
-\end{figure}%
-
-In terms of model error, some data points show high error (see
-Figure~\ref{fig-nldervisMNISTPACMAP}) due to their deviation from the
-typical high-D data structure, acting as outliers. This is evident when
-comparing the images associated with these high model error points (see
-Figure~\ref{fig-sampleimgerror}) to others (see
-Figure~\ref{fig-sampleimg}).
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-sampleimgerror-1.pdf}
-
-}
-
-\caption{\label{fig-sampleimgerror}Images of handwritten digit 1 which
-occur large model error.}
-
-\end{figure}%
-
-\section{Conclusion}\label{sec-conclusions}
-
-In conclusion, our proposed model offers a novel approach for
-visualizing high-dimensional (high-D) data by leveraging non-linear
-dimension reduction (NLDR) techniques. Through a series of steps
-including hexagonating NLDR data, triangulating bin centroids, and
-lifting the 2D triangular mesh into high dimensions, our model
-effectively transforms complex high-D data into interpretable 2D
-representations. The model's performance is evaluated using goodness of
-fit statistics such as Mean Squared Error (MSE) and Akaike Information
-Criterion (AIC), providing insights into its accuracy and reliability.
-Overall, our model presents a valuable tool for researchers and
-practitioners in various fields to gain deeper insights from
-high-dimensional datasets.
-
-Our algorithm mainly consists\ldots{}
-
-These Goodness of fit statistics useful in encountering the prediction
-accuracy when we predict the 2D embedding for the new high-D data.
-Because, we try to find the nearest high-D mappings as the first step of
-prediction which will participate the prediction process. When we need
-to find which different NLDR technique and which parameter choices is
-giving the best representation of high-D data, MSE and AIC are also
-useful.
-
-We introduce a new tool to help to determine which method, which
-parameter choice provide the most useful representation of high-D data.
+\begin{itemize}
+\tightlist
+\item
+  Summarise contributions
+\item
+  Explain where it is expected or not expected to work, eg higher
+  dimensional relationships
+\item
+  What might be useful enhancements
+\end{itemize}
 

### 2024-03-27 paper_original.qmd
cleaning up the paper

---
title: "Visualising How Non-linear Dimension Reduction Warps Your Data"
format: 
    jasa-pdf:
        keep-tex: true
    jasa-html: default
author:
  - name: Jayani P.G. Lakshika
    affiliations:
      - name: Monash University
        department: Econometrics & Business Statistics
        address: Clayton
        city: VIC 
        country: Austria
        postal-code: 3800
    orcid: 0000-0002-6265-6481
    email: jayani.piyadigamage@monash.edu
    url: https://jayanilakshika.netlify.app/
  - name: Dianne Cook
    affiliations:
      - name: Monash University
        department: Econometrics & Business Statistics
        address: Clayton
        city: VIC 
        country: Austria
        postal-code: 3800
    orcid: 0000-0002-3813-7155
    email: dicook@monash.edu 
    url: http://www.dicook.org/
  - name: Paul Harrison
    affiliations:
      - name: Monash University
        department: MGBP, BDInstitute
        address: Clayton
        city: VIC 
        country: Austria
        postal-code: 3800
    orcid: 0000-0002-3980-268X
    email: 	paul.harrison@monash.edu
    url: 
  - name: Michael Lydeamore
    affiliations:
      - name: Monash University
        department: Econometrics & Business Statistics
        address: Clayton
        city: VIC 
        country: Austria
        postal-code: 3800
    orcid: 0000-0001-6515-827X
    email: michael.lydeamore@monash.edu
    url: 
  - name: Thiyanga S. Talagala
    affiliations:
      - name: University of Sri Jayewardenepura
        department: Statistics
        address: Gangodawila
        city: Nugegoda 
        country: Sri Lanka
        postal-code: 10100
    orcid: 0000-0002-0656-9789
    email: ttalagala@sjp.ac.lk 
    url: https://thiyanga.netlify.app/
tbl-cap-location: bottom
abstract: |
  Nonlinear dimension reduction (NLDR) techniques such as tSNE, and UMAP provide a low-dimensional representation of high-dimensional (high-D) data using non-linear transformation. The methods and parameter choices can create wildly different representations, making it difficult to decide which is best, or whether any or all are accurate or misleading. NLDR often exaggerates random patterns, sometimes due to the samples observed. But NLDR views have an important role in data analysis because, if done well, they provide a concise visual (and conceptual) summary of high-D distributions. To help evaluate the NLDR we have developed an algorithm to show the 2D NLDR model in the high-D space, viewed with a tour. One can see if the model fits everywhere or better in some subspaces, or completely mismatches the data. It is used to evaluate which 2D layout is the best representation of the high-D distribution and see how different methods may have similar summaries or quirks.

keywords: [high-dimensional data, dimension reduction, triangulation, hexagonal binning, low-dimensional manifold, manifold learning, tour, data vizualization]
keywords-formatted: [high-dimensional data, dimension reduction, triangulation, hexagonal binning, low-dimensional manifold, manifold learning, tour, data vizualization]

bibliography: bibliography.bib  
---

```{r}
#| warning: false
#| echo: false
library(quollr)
library(dplyr)
# remotes::install_github("jlmelville/snedata")
library(snedata)
library(ggflowchart)
library(purrr) ## map function
library(gridExtra) ## for grid.arrange
library(rsample)
library(DT)
library(ggbeeswarm)
library(ggplot2)
library(readr)
library(tidyr)

library(Rtsne)
library(uwot)
library(phateR)
library(patchwork)
library(langevitour)
library(colorspace)
library(kableExtra)
library(grid)
set.seed(20240110)

source("nldr_code.R", local = TRUE)
```

```{=html}
<!-- 
  Notes
* Use American spelling
-->
```

## Introduction {#sec-intro}

High-dimensional (high-D) data is prevalent across various fields, such as ecology and bioinformatics [@Guo2023], due to advancements in data collection technologies [@Johnstone2009; @ayesha2020overview]. However, visualization of high-D data introduces significant challenges, because the complexity of visualizing data beyond two dimensions [@Jia2022]. In recent years, interactive and dynamic graphics systems like **liminal** [@article21] —which employs interactive tools like brushing and linking [@article58]—and software tools such as **XGobi**, **GGobi** [@article60], **tourr**  [@article61], **detourr** [@article22], and **langevitour**  [@article09], involving dynamic methods like tours [@Asimov1985], have played a key role in visualizing high-D data (data-vis).

To create low-dimensional representations (typically in 2D) (m-vis) [@article59] of high-D data, it is common to apply dimension reduction (DR) techniques. Approaches for DR involve linear methods such as principal component analysis (PCA) [@Karl1901], non-linear methods such as multi-dimensional scaling (MDS) [@Torgerson1967]. In the past decade, many new non-linear dimension reduction (NLDR) techniques have emerged, such as t-distributed stochastic neighbor embedding (tSNE) [@Laurens2008] and uniform manifold approximation and projection (UMAP) [@Leland2018]. NLDR techniques are the 2D models of high-D data in our context.

It is important to visualize various non-linear dimensionality reduction (NLDR) techniques for the same high-D data in order to understand and find the best representation. After doing so, the 2D models may differ considerably from each other and may also deviate from the original data structure in high-dimensional space. Therefore, visualizing the 2D model in high-D space (m-in-ds) is more useful to answer different types of questions:

- Is there a best 2D representation of high-D data or are they all providing  equivalent information? Is there a best parameter choice to fit the 2D model? How does the model change when it's parameters change?

- How well the does the 2D models capture the data structure? Is the model fitting able to capture different data structure like non-linear, clustering?

If we cannot easily ask and answer these questions, our ability to understand the models is limited. To find the best 2D model and parameter choices, a better understanding of the underlying science is important.

Also, the importance of m-vis along with data-vis has been recognized and incorporated into interactive software, **liminal** [@article21]. But the 2D model and high-D visualize side by side and interactive like brushing and linking connect the data in the two panels. To address this challenge, we propose a novel approach by combining the tour technique with a low-dimensional manifold. This manifold is created through the synergistic use of NLDR techniques, hexagonal binning, and triangulation. This integration facilitates a more understanding of the data structure, how well (or how poorly) NLDR techniques perform.

The outline of this paper is as follows. @sec-background provides an detailed overview of dimension reduction methods, and tour technique. Building upon this foundation, Section @sec-methods describes the proposed algorithm, including its implementation details, model tuning, summaries, and a synthetic example to demonstrate its functionality. In Section @sec-applications, the algorithm's applications on various datasets, particularly in single-cell RNA-seq and handwritten digit data, are showcased. These applications demonstrate how some NLDR techniques effectively capture both local and global structures of the original data, while others may result in misleading interpretations. Finally, Section @sec-conclusions summarizes the findings and describe the importance of the proposed approach in addressing the challenges associated with selecting the best NLDR method and parameter choices to achieve a more accurate 2D representation of high-D data.

<!--UMAP with different param choices-->

```{r}
#| warning: false
#| echo: false

UMAP_data_7 <- read_rds(file = "data/s_curve/s_curve_umap_7.rds")

#(n-neighbors: 50)
plot_list1_umap <- plot_UMAP_2D(UMAP_data_7) + #ggtitle("(b)") +
  geom_point(alpha=0.5, size = 0.5, colour = "#8dd3c7", fill = "#8dd3c7") +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

```

```{r}
#| warning: false
#| echo: false

UMAP_data_15 <- read_rds(file = "data/s_curve/s_curve_umap_15.rds")

#(n-neighbors: 50)
plot_list2_umap <- plot_UMAP_2D(UMAP_data_15) + #ggtitle("(b)") + 
  geom_point(alpha=0.5, size = 0.5, colour = "#a65628", fill = "#a65628") +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

```

```{r}
#| warning: false
#| echo: false

UMAP_data_32 <- read_rds(file = "data/s_curve/s_curve_umap_32.rds")

#(n-neighbors: 50)
plot_list3_umap <- plot_UMAP_2D(UMAP_data_32) + #ggtitle("(b)") +
  geom_point(alpha=0.5, size = 0.5, color = "#f781bf", fill = "#f781bf") +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

```

```{r}
#| warning: false
#| echo: false

UMAP_data <- read_rds(file = "data/s_curve/s_curve_umap.rds")

#(n-neighbors: 50)
plot_list4_umap <- plot_UMAP_2D(UMAP_data) + #ggtitle("(b)") +
  geom_point(alpha=0.5, size = 0.5, colour = "#999999", fill = "#999999") +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'd', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

```

```{r}
#| echo: false
#| fig-cap: "2D layouts from UMAP applied for the S-curve data: (a) UMAP (n_neighbors = 7), (b) UMAP (n_neighbors = 15), (c) UMAP (n_neighbors = 32), (d) UMAP (n_neighbors = 50). Is there a best hyperparameter choice in representing UMAP or are they all providing  equivalent information?"
#| label: fig-nldervisUMAP
#| out-width: 100%

plot_list1_umap + plot_list2_umap + plot_list3_umap + plot_list4_umap +
  plot_layout(ncol=4)
```

## Background {#sec-background}

### Dimension Reduction

Consider the high-D data a rectangular matrix $X_{n \times p}$, where $X_{n \times p} = \begin{bmatrix} \textbf{x}_{1} & \textbf{x}_{2} & \cdots & \textbf{x}_{n}\\ \end{bmatrix}^\top$, with $n$ observations in $p$ dimensions. The objective is to discover a low-dimensional projection $Y_{n \times d} = \begin{bmatrix} \textbf{y}_{1} & \textbf{y}_{2} & \cdots & \textbf{y}_{n}\\ \end{bmatrix}^\top$, represented as an $n$ × $d$ matrix, where $d \ll p$. The reduction process seeks to remove noise from the original data set while retaining essential information.

There are two main categories of dimension reduction techniques: linear and non-linear methods. Linear techniques involve a linear transformation of the data, with one popular example being PCA. PCA performs an eigen-decomposition of the sample covariance matrix to obtain orthogonal principal components that capture the variance of the data [@Karl1901].

In contrast, NLDR techniques generate the low-dimensional representation $Y$ from the high-dimensional data $X$, often using pre-processing techniques like $k$-nearest neighbors graph or kernel transformations. Multidimensional Scaling (MDS) is a class of NLDR methods that aims to construct an embedding $Y$ in a low-dimensional space, approximating the pair-wise distances in $X$ [@Torgerson1967]. Variants of MDS include non-metric scaling [@article62] and Isomap, which estimate geodesic distances to create the low-dimensional representation [@article63]. Other approaches based on diffusion processes, like diffusion maps [@article64] and the PHATE (Potential of Heat-diffusion for Affinity-based Trajectory Embedding) algorithm [@article03], also fall under NLDR methods.

#### Non-linear dimension reduction techniques

NLDR techniques are crucial for analyzing and displaying high-dimensional data, where linear approaches may not adequately capture complexities in relationships between variables [@Johnstone2009]. One of the challenges with NLDR techniques is the selection and tuning of appropriate hyperparameters [@liao2023]. This process involves finding the suitable combination of hyperparameters that enhances the performance of the NLDR technique, considering the characteristics of the dataset and the specific goals of the analysis.

<!--This process requires a delicate balance to optimize the performance of NLDR techniques, taking into account the intricacies of the dataset and the specific goals of the analysis.-->

Additionally, another challenge is lack of reverse mapping. Techniques like PCA and auto-encoders [@article65] provide a way to map back from the low-dimensional space to the high-D space, facilitating data reconstruction. However, some NLDR methods, such as tSNE, don't have a specific way to reconstruct the original data from the low-dimensional space.

In this article, mainly focus on five NLDR techniques. They are tSNE, UMAP, PHATE, TriMAP [@article02], and Pairwise Controlled Manifold Approximation (PaCMAP) [@Yingfan2021].

Among these, tSNE [@Laurens2008] stands out for its ability to preserve pairwise distances. By minimizing the divergence between probability distributions in both high and low-dimensional spaces, tSNE effectively uncovers intricate structures and patterns within the data. Its application is widespread, particularly in tasks requiring the visualization of clusters and local relationships. However, achieving effective results requires careful consideration of hyperparameters, such as perplexity.

UMAP [@Leland2018] is a useful technique for simplifying data while maintaining both local and overall structures. It builds a fuzzy topological view by considering nearby data points and then optimizes a simplified version to match that view. UMAP is known for working well with different scales of relationships in data and is efficient in handling large datasets. However, it's important to choose parameters like neighbors and minimum distance carefully, as they can affect the results.

Furthermore, PHATE [@article03] is great for understanding how things develop, especially in single-cell genomics. It uses a heat diffusion process to capture relationships between data points, like points along a trajectory. While PHATE is excellent for revealing these developmental structures, it requires careful tuning of its parameters because of its specialized focus.

Additionally, TriMAP [@article02] takes a special approach by creating a triangulated graph representation of the data. This method is good at understanding both local and global structures by treating the data as a network of triangles. TriMAP is powerful in capturing complicated structures, but it's important to choose parameters carefully, like deciding how many neighbors to consider.

PaCMAP [@Yingfan2021] is different because it adds supervised learning to make a 2D representation while keeping the relationships between pairs of points. It builds a graph using distances between pairs and then makes the 2D representation better using a customizable loss function. What's special about PaCMAP is that it can use class labels or extra information to guide how it makes the 2D representation. This gives users a way to change how PaCMAP works to fit their needs better.

<!--Among these techniques, tSNE stands out for its emphasis on preserving pairwise distances. By minimizing the divergence between probability distributions in both the high and low-dimensional spaces, t-SNE effectively reveals intricate structures and patterns within the data. Its application is widespread in tasks requiring the visualization of clusters and local relationships, though it does require careful consideration of the perplexity parameter for optimal results.

UMAP is another powerful non-linear technique that strikes a balance between preserving local and global structures. Constructing a fuzzy topological representation using a weighted k-nearest neighbors graph, UMAP optimizes the low-dimensional embedding to resemble this representation. Known for its efficiency and scalability, UMAP is versatile across various scales of relationships in the data, although parameter sensitivity, particularly concerning the choice of neighbors, must be taken into account.

For trajectory data, PHATE provides specialized capabilities. It models the affinity between data points, simulating a heat diffusion process to capture developmental processes, particularly in single-cell genomics. While PHATE excels in revealing trajectory structures and offering insights into cellular development, it necessitates careful parameter tuning due to its specialized nature.

TriMAP adopts a unique approach by approximating the data manifold through the construction of a triangulated graph representation. This technique efficiently captures both global and local structures by representing the data as a network of triangles. TriMAP's strength lies in its ability to efficiently capture complex structures, albeit with sensitivity to parameter choices, including the number of neighbors.

In contrast, PaCMAP introduces supervised learning to create a low-dimensional representation while preserving pair-wise relationships. Constructing a graph based on pair-wise distances, PaCMAP optimizes an embedding using a customizable loss function. Particularly notable is PaCMAP's flexibility in incorporating class labels or additional information to guide the embedding process, offering users a means to customize its behavior and performance.-->

```{r}
#| warning: false
#| echo: false

data <- read_rds("data/s_curve/s_curve.rds")
```

```{r}
#| warning: false
#| echo: false

training_data <- read_rds("data/s_curve/s_curve_training.rds")
test_data <- read_rds("data/s_curve/s_curve_test.rds")
```

```{r}
#| warning: false
#| echo: false

tSNE_data <- read_rds("data/s_curve/s_curve_tsne_27.rds")

plot_list1 <- plot_tSNE_2D(tSNE_data) +  
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

```

```{r}
#| warning: false
#| echo: false

UMAP_data <- read_rds(file = "data/s_curve/s_curve_umap.rds")

#(n-neighbors: 50)
plot_list2 <- plot_UMAP_2D(UMAP_data) + 
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

```

```{r}
#| warning: false
#| echo: false

#(knn: 5)
# PHATE_data <- Fit_PHATE(training_data, knn = 5, with_seed = 20240110)
# write_csv(PHATE_data, paste0(here::here(), "/data/phate_data_s_curve.csv"))

PHATE_data <- read_rds(file = "data/s_curve/s_curve_phate.rds")

plot_list3 <- plot_PHATE_2D(PHATE_data) + 
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

```

```{r}
#| warning: false
#| echo: false

# tem_dir <- tempdir()
# 
# Fit_TriMAP_data(training_data, tem_dir)
# 
# path <- file.path(tem_dir, "df_2_without_class.csv")
# path2 <- file.path(tem_dir, "dataset_3_TriMAP_values.csv")
# 
# Fit_TriMAP(as.integer(2), as.integer(5), as.integer(4), as.integer(3), path, path2)
# 
# TriMAP_data <- read_csv(path2)
# write_csv(TriMAP_data, paste0(here::here(), "/data/trimap_data_s_curve.csv"))

TriMAP_data <- read_rds(file = "data/s_curve/s_curve_trimap.rds")

#(n-inliers: 5, \n n-outliers: 4, n-random: 3)
plot_list4 <- plot_TriMAP_2D(TriMAP_data) + 
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)

```

```{r}
#| warning: false
#| echo: false

# tem_dir <- tempdir()
# 
# Fit_PacMAP_data(training_data, tem_dir)
# 
# path <- file.path(tem_dir, "df_2_without_class.csv")
# path2 <- file.path(tem_dir, "dataset_3_PaCMAP_values.csv")
# 
# Fit_PaCMAP(as.integer(2), as.integer(10), "random", 0.9, as.integer(2), path, path2)
# 
# PacMAP_data <- read_csv(path2)
# write_csv(PacMAP_data, paste0(here::here(), "/data/pacmap_data_s_curve.csv"))

PaCMAP_data <- read_rds(file = "data/s_curve/s_curve_pacmap.rds")

#(knn: 10, init: random, \n MN-ratio: 0.9, FP-ratio: 2)
plot_list5 <- plot_PaCMAP_2D(PaCMAP_data) +  
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'e', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)

```

```{r}
#| echo: false
#| fig-cap: "2D layouts from different NLDR techniques applied the same data: (a) tSNE (perplexity = 27), (b) UMAP (n_neighbors = 50), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2). Is there a best representation of the original data or are they all providing  equivalent information?"
#| label: fig-nldervis
#| out-width: 100%

plot_list1 + plot_list2 + plot_list3 + plot_list4 + plot_list5 +
  plot_layout(ncol=5)
```

### Linear overviews using tours

A tour is a powerful visualization technique used to explore the shape and global structure of high-dimensional data by generating a sequence of projections, typically into two dimensions. There are two main types of tours: the grand tour [@Asimov1985] and the guided tour [@article29]. A grand tour involves randomly selecting new orthonormal bases, enabling users to understand the structure by exploring the subspace of d-dimensional projections [@Asimov1985]. In contrast, a guided tour can be employed to generate a sequence of 'interesting' projections based on an index function [@article29].

The process begins with the data matrix $X$. It generates a sequence of $p$ × $d$ orthonormal projection matrices (bases) $P_t$, usually $d$ is one or two dimensions. For each pair of orthonormal bases $P_t$ and $P_{t+1}$, a geodesic path is interpolated to create smooth animation between projections. The resulting tour continuously visualizes the projected data $Y_t$ = $XP_t$ as it interpolates between successive bases.

Furthermore, software like **langevitour** can visualize both types of tours, providing flexibility for exploring high-dimensional data with various objectives. In our context, use grand tour along with the model to observe how effectively the model captures the underlying structure of the data.

## Methodology {#sec-methods}

In this paper, we introduce a novel method to determine the most effective NLDR technique and the best hyperparameter choice that provides the most useful representation of high-D data. Our approach involves dividing the high-D dataset into two parts: a training set for constructing the model and a test set for generating predictive values and residuals. Our algorithm takes a 2D embedding data as the input and generate a tour that displays the high-D wireframe to overlay the data. The flow chart of the proposed algorithm is shown in @fig-meth. The algorithm consists of two main phases: (1) generating the model in the 2D space and (2) lifting the model into high-D space. The main steps of the algorithm are described in detail in this section using UMAP 2D embedding of the S-curve dataset. This dataset has seven dimensions, including four noise dimensions that were added to the original 3D data.

![The flow diagram shows the main steps of our algorithm. There are two basic phases, one to generate the model in the 2D space, and other to map the model into the high-D space.](figures/workflow.png){#fig-meth fig-align="center" width="100%" height="100%"}

### Preprocessing steps

To reduce computational complexity when applying NLDR techniques to high-D data and to reduce noise presence, PCA [@article67, @article68, @article69] is used as a preprocessing step. PCA involves identifying principal components that maximize variance. These components are then used as the high-D data for the algorithm.

### Constructing the 2D model {#sec-construct2d}

**Step 1: Scaling NLDR data**

First, we prepare the 2D embedding data to fit within the bounds required for regular hexagonal binning. To achieve this, we implement two key scaling steps. Scale the first 2D embedding component to range between $0$ and $1$, ensuring that all data points fall within this normalized interval. Secondly, we scale the second 2D embedding component to range between $0$ and $y_{max}$ (see @eq-equation3).

The calculation of $y_{max}$ involves several steps. First, the aspect ratio ($ar$) is computed by dividing the range of the second 2D embedding component ($r_2$) by the range of the first 2D embedding component ($r_1$) (see @eq-equation1). Then, the hexagon ratio ($hr$) is determined by dividing the height of the hexagon ($hb$) by its width ($wb$) (see @eq-equation2). Finally, $y_{max}$ is derived by taking the ceiling of $\frac{ar}{hr}$ and multiplying it by $hr$. This process ensures that $y_{max}$ is an integer multiple of $hr$, accommodating the grid layout of the hexagonal bins.

$$
 ar = \frac{r_2}{r_1}
$$ {#eq-equation1}

$$
 hr = \frac{hb}{wb}
$$ {#eq-equation2}

$$
 y_{max} = \left\lceil\frac{ar}{hr}\right\rceil * hr
$$ {#eq-equation3}

**Step 2: Hexagonating NLDR data**

Hexagonating NLDR data (see @fig-meth Step 2) involves partitioning the NLDR data into hexagonal bins, a technique commonly referred to as hexagonal binning [@Carr1987, @article66]. This method use a hexagonal grid to create a bivariate histogram that effectively visualizes the structure of high-D data. Hexagons, one of only three regular polygons capable of tessellating a plane [@Carr2013], offer unique advantages due to their symmetry of nearest neighbors and maximal number of sides for such tessellations [@Dan2023]. This geometric property makes hexagons more efficient in covering the plane compared to other regular tessellations and reduces visual bias when displaying data densities [@Dan2023].

In our algorithm, we aim to conduct regular hexagonal binning, involving the computation of hexagonal grid configurations, the generation of the grid, and the assignment of the NLDR data points to hexagons.

**(a) Computation of hexagonal grid configurations**

In the computation of hexagonal bin configurations, there are mainly two configurations to consider: the number of bins along the x and y axes. The first step involves in determining the hexagonal size ($s$) and the buffer along the x and y axes ($buffer_{x}$ and $buffer_{y}$). Here, the hexagonal size ($s$) represents the radius of the outer circle surrounding the hexagon. The buffer is important in expanding the boundary to accommodate potential outliers or edge cases.

When computing the number of bins along the x and y axes ($b_1$ and $b_2$), the process begins by calculating the range of the respective 2D embedding component. After that, the range is adjusted to accommodate the buffer amount. Then, the spacing between hexagons is determined based on $s$. Finally, the number of bins along the axis is computed by dividing the adjusted range by the spacing.

For the computation of the number of bins along the x-axis ($b_1$) (see @eq-equation12), the range of the first embedding component is defined as $r_1$, while the buffer amount along the x-axis is denoted as $buffer_{x}$, and the horizontal spacing is represented by $h$ (see @eq-equation11). On the other hand, the number of bins along the y-axis ($b_2$) (see @eq-equation14) is computed based on the range of the second embedding component ($r_2$), the buffer amount along the y-axis ($buffer_{y}$), and the vertical spacing ($v$) (see @eq-equation13).

$$
 h = \sqrt{3} * s
$$ {#eq-equation11}

$$
 b_1 = \frac{r_1 + buffer_{x}}{h}
$$ {#eq-equation12}

$$
  v = 1.5 * s
$$ {#eq-equation13}

$$
 b_2 = \frac{r_2 + buffer_{y}}{v}
$$ {#eq-equation14}

**(b) Generation of the hexagonal grid**

In this process, the first step involves generating centroids within the hexagonal grid. This begins by defining the starting coordinates for the hexagons. With the number of bins computed along the x and y axes, along with the horizontal and vertical spacing, the centroids are iteratively computed starting from the hexagonal starting coordinates.

Once the centroids for the hexagonal grid are obtained, the next step is to compute the hexagonal coordinates. For example, if the centroid of a hexagonal bin is defined as $(C_x, C_y)$, the hexagonal coordinates can be computed using $d_x$ (see @eq-equation15) and $d_y$ (see @eq-equation16) (see @fig-hexcoord).

$$
d_x = \frac{h}{2}
$$ {#eq-equation15}

$$
d_y = \frac{2* v * 1.15}{\sqrt{3}} 
$$ {#eq-equation16}

```{r}
#| echo: false
#| fig-pos: H
#| fig-cap: "The hexagonal coordinates computed for the centroid $(C_x, C_y)$."
#| label: fig-hexcoord
#| out-width: 100%
#| out-height: 100%
knitr::include_graphics("figures/hex_coord.png")
```

**(c) Assignment of the NLDR data points to hexagons**

After obtaining the centroids of the hexagonal bins for the entire grid, the process of assigning NLDR points to hexagons involves determining the nearest hexagonal bin for each NLDR point using the 2D Euclidean distance. Then, the hexagonal ID of the nearest hexagon is assigned to the corresponding NLDR points.

**Step 3: Obtaining bin centroids or bin means**

In the previous step, the algorithm clusters the 2D embedding data into hexagons. Following this, in this step, the bin centroids or bin means (see @fig-meth Step 3) are obtained [@Carr2013].

The bin centroid ($C_k^{(2)}$) for a $k^{th}$ hexagon with hexagonal grid coordinates $(h^{k}x_{i}, h^{k}y_{i})$, where $i = 1 \dots 6$ can be defined as:

$$
C_k^{(2)} = (C_{ky_1}, C_{ky_2}) = \left(\frac{\sum_{i=1}^{6} h^{k}x_{i}}{6}, \frac{\sum_{i=1}^{6} h^{k}y_{i}}{6}\right).
$$ {#eq-equation4}

Also, the bin mean ($C_k^{(2)}$) is defined as the mean of the data points within the $k^{th}$ hexagon (see @eq-equation5).

$$
C_k^{(2)} = (C_{ky_1}, C_{ky_2}) = \left(\frac{1}{n_k} \sum_{i=1}^{n_k} y_{1i}, \frac{1}{n_k} \sum_{i=1}^{n_k} y_{2i}\right),
$$ {#eq-equation5}

where $n_k$ is the number of data points within the hexagon, $y_{1i}$ and $y_{2i}$ are the $x$ and $y$ coordinates of the $i^{th}$ data point within the hexagon.

**Step 4: Triangulating bin centroids or bin means**

In this step, the algorithm proceeds to triangulate the hexagonal bin centroids or bin means (see @fig-meth Step 4). Triangulation is a fundamental process in computational geometry and computer graphics that involves dividing a set of points in a given space into interconnected triangles [@article30]. One common algorithm used for triangulation is Delaunay triangulation [@article26, @article54], where points are connected in a way that maximizes the minimum angles of the resulting triangles, leading to a more regular and well-conditioned triangulation.

Delaunay triangulation can be defined as follows:

Let $C^{(2)} = \{C_1^{(2)}, C_2^{(2)}, ..., C_m^{(2)}\}$ be a set of $m$ bin centroids or bin means in the plane. Delaunay triangulation of $C^{(2)}$, denoted as $DT(C^{(2)})$, is a triangulation of the convex hull of $C^{(2)}$ such that the circumcircle of every triangle in the triangulation contains no other points from $C^{(2)}$.

Given that the hexagons are regular, the resulting triangles will mostly be equilateral.

### Lifting the model into high dimensions

#### Lifting the triangular mesh points into high dimensions

Consider $f: \mathbb{R}^p \rightarrow \mathbb{R}^2$ be a function that maps the high-D data ($X_{n \times p}$) to its NLDR equivalent ($Y_{n \times d}$). Then, let $g: \mathbb{R}^2 \rightarrow \mathbb{R}^2$ be a function that maps each 2D embedding point to its closest centroid ($C^{(2)}$). It follows that $f(g(x))$ maps the high-D points $x$ to the centroid in 2D ($C_k^{(2)}$). Also, define a function $v: \mathbb{R}^2 \rightarrow \mathbb{R}^p$ maps the 2D centroid ($C^{(2)}$) to the high-D mean of the points ($C^{(p)}$) in the hexagon.

The high-D mean of all the points in $k^{th}$ hexagon by

$$
C_k^{(p)} = (C_{kx_1}, ..., C_{kx_p}) = \left(\frac{1}{n_k} \sum_{i=1}^{n_k} x_{1i}, \frac{1}{n_k} \sum_{i=1}^{n_k} x_{2i}, \dots ,\frac{1}{n_k} \sum_{i=1}^{n_k} x_{pi}\right).
$${#eq-equation6}

Therefore,

$$
v(C_k^{(2)}) = C_k^{(p)}.
$${#eq-equation7}

Therefore, $f(g(x))$ gives the 2D centroid associated with high-D points $x$, and $v(C_k^{(2)})$ gives the high-D centroid associated with 2D point $C_k^{(2)}$. Thus, $v(f(g(x)))$ gives the high-D centroid ($C_k^{(p)}$) associated with the 2D embedding of the points $x$.

#### Lifting the 2D triangular mesh into high dimensions

As described in Step 4 of @sec-construct2d, during the triangulation process in 2D space, vertices are identified to form edges. With the knowledge of the high-D mappings for the 2D hexagonal bins, the vertices connected in 2D are also connected in high-D (see video linked in @fig-wkhighD).

```{r}
#| warning: false
#| echo: false
#| message: false

umap_s_curve_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = UMAP_data, 
                                    x = "UMAP1", y = "UMAP2"))) |>
  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
                  "UMAP2" = "scaled_UMAP2")) |>
  dplyr::mutate(ID = 1:NROW(UMAP_data))

```

```{r}
#| warning: false
#| echo: false
#| message: false

num_bins_x <- 5
num_bins_y <- 10
hex_size <- 0.19

hb_obj <- hex_binning(data = umap_s_curve_scaled, x = "UMAP1", 
                              y = "UMAP2", num_bins_x = num_bins_x, 
                              num_bins_y = num_bins_y, x_start = NA, y_start = NA, 
                              buffer_x = NA, buffer_y = NA, hex_size = hex_size, col_start = "UMAP")

## Data set with all possible centroids in the hexagonal grid
all_centroids_df <- as.data.frame(do.call(cbind, hb_obj$centroids))

## Generate all coordinates of hexagons
hex_grid <- as.data.frame(do.call(cbind, hb_obj$hex_poly))

## To obtain the standardise counts within hexbins
counts_df <- as.data.frame(do.call(cbind, hb_obj$std_cts))
df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
                                             counts_df = counts_df)

UMAP_data_with_hb_id <- as.data.frame(do.call(cbind, hb_obj$data_hb_id))

model_object <- fit_highd_model( training_data = training_data, 
                                 nldr_df_with_id = umap_s_curve_scaled, 
                                 x = "UMAP1", y = "UMAP2", 
                                 num_bins_x = num_bins_x, 
                                 num_bins_y = num_bins_y, 
                                 x_start = NA, y_start = NA, 
                                 buffer_x = NA, buffer_y = NA, 
                                 hex_size = hex_size,
                                 is_rm_lwd_hex = FALSE, 
                                 benchmark_to_rm_lwd_hex = NA, 
                                 col_start_2d = "UMAP", 
                                 col_start_highd = "x")

df_bin_centroids <- model_object$df_bin_centroids
df_bin <- model_object$df_bin

## Triangulate bin centroids
tr1_object <- tri_bin_centroids(df_bin_centroids, x = "c_x", y = "c_y")
tr_from_to_df <- gen_edges(tri_object = tr1_object)

# Create the tibble with x and y coordinates
tr_df <- tibble::tibble(x = tr1_object$x, y = tr1_object$y)

## Compute 2D distances
distance <- cal_2d_dist(tr_coord_df = tr_from_to_df, 
                                start_x = "x_from", start_y = "y_from", 
                                end_x = "x_to", end_y = "y_to", 
                                select_vars = c("from", "to", "distance"))

## Map hb_id

df_bin_centroids_n <- df_bin_centroids |>
  select(hexID) |>
  mutate(ID = 1:NROW(df_bin_centroids))

tr_from_to_df_n <- left_join(tr_from_to_df, df_bin_centroids_n, by = c("from" = "ID")) |>
  rename("from_hb_id" = "hexID")

tr_from_to_df_n <- left_join(tr_from_to_df_n, df_bin_centroids_n, by = c("to" = "ID")) |>
  rename("to_hb_id" = "hexID")

hb_id_selected <- tr_from_to_df_n |>
  dplyr::filter((from == 1) & (to == 3))

# Filter and label small and long edges
distance_df_small_edges <- distance |>
  dplyr::filter((from != 1) | (to != 3)) |>
  dplyr::mutate(type = "small_edges")

distance_df_long_edges <- distance |>
  dplyr::filter((from == 1) & (to == 3)) |>
  dplyr::mutate(type = "long_edges")

# Combine small and long edges
distance_edges <- dplyr::bind_rows(distance_df_small_edges, distance_df_long_edges)

tr_from_to_df_coord_with_group <- inner_join(tr_from_to_df, distance_edges, by = c("from" = "from", "to" = "to"))

## To generate a data set with high-D and 2D training data
df_all <- dplyr::bind_cols(training_data |> dplyr::select(-ID), UMAP_data_with_hb_id)

UMAP_data_with_hb_id_selected <- UMAP_data_with_hb_id |> dplyr::filter(hb_id %in% c(hb_id_selected$from_hb_id, hb_id_selected$to_hb_id)) |>
  dplyr::mutate(select_point = if_else(hb_id == hb_id_selected$from_hb_id, "data1", "data2"))

UMAP_data_with_hb_id_selected_no <- UMAP_data_with_hb_id |> dplyr::filter(!(hb_id %in% c(hb_id_selected$from_hb_id, hb_id_selected$to_hb_id)))

bin_coord_1 <- tr_df |>
  dplyr::filter((x == hb_id_selected$x_from)) |>
  dplyr::filter((y == hb_id_selected$y_from))

# bin_coord_1 <- tr_df |>
#   head(1)

bin_coord_2 <- tr_df |>
  dplyr::filter((x == hb_id_selected$x_to)) |>
  dplyr::filter((y == hb_id_selected$y_to))

# bin_coord_2 <- tr_df |>
#   dplyr::filter(row_number() == 2)

bin_coord <- dplyr::bind_rows(bin_coord_1, bin_coord_2)

a1 <- ggplot(bin_coord, aes(x = x, y = y)) +
  geom_segment(
    aes(x = x_from, y = y_from, xend = x_to, yend = y_to, colour = type),
    data = tr_from_to_df_coord_with_group) +
  geom_point() +
  #coord_equal() +
  coord_cartesian(xlim =c(-0.5, 1.5), ylim = c(-0.5, 2.7)) +
  geom_point(data = UMAP_data_with_hb_id_selected, aes(x = UMAP1, y = UMAP2, colour = select_point), alpha = 0.5, size = 1) +
  scale_color_manual(values=c("#e6550d", "#000000", "#e31a1c", "#f6e8c3")) +
  geom_point(size = 3, colour = "#33a02c") + 
  #ggtitle("(b)") + 
  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  theme_light() +
  #coord_equal() +
  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
  )

a2 <- ggplot(data = hex_grid, aes(x = x, y = y)) + 
  geom_polygon(color = "#deebf7", aes(group = hex_poly_id), fill = "#ffffff") +
  geom_segment(
    aes(x = x_from, y = y_from, xend = x_to, yend = y_to), colour = "#f6e8c3",
    data = tr_from_to_df_coord_with_group
  ) +
  geom_point(data = UMAP_data_with_hb_id_selected, aes(x = UMAP1, y = UMAP2, colour = select_point), size = 1, alpha = 0.5) +
  scale_color_manual(values=c("#e6550d", "#000000")) +
  geom_point(data = df_bin_centroids |> dplyr::filter(hexID %in% c(hb_id_selected$from_hb_id, hb_id_selected$to_hb_id)), aes(x = c_x, y = c_y), colour = "#33a02c", size = 3) +
  geom_point(data = UMAP_data_with_hb_id_selected_no, aes(x = UMAP1, y = UMAP2), colour = "#6a3d9a", size = 1, alpha = 0.5) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
  coord_equal() +
  theme_light() +
  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
  )

## Script to make coloring data points with a line

df1 <- df_all |> 
  dplyr::filter(hb_id == as.character(hb_id_selected$from_hb_id)) |>
  dplyr::select(tidyselect::starts_with("x")) |>
  dplyr::mutate(type = "data1") ## original dataset

df2 <- df_all |> 
  dplyr::filter(hb_id == as.character(hb_id_selected$to_hb_id)) |>
  dplyr::select(tidyselect::starts_with("x")) |>
  dplyr::mutate(type = "data2") ## original dataset

df3 <- df_all |> 
  dplyr::filter(!(hb_id %in% c(as.character(hb_id_selected$from_hb_id), as.character(hb_id_selected$to_hb_id)))) |>
  dplyr::select(tidyselect::starts_with("x")) |>
  dplyr::mutate(type = "data") ## original dataset

df_b <- df_bin |> 
  dplyr::filter(hb_id %in% df_bin_centroids$hexID) |>
  dplyr::filter(hb_id %in% c(as.character(hb_id_selected$from_hb_id), as.character(hb_id_selected$to_hb_id))) |>
  dplyr::select(-hb_id) |>
  dplyr::mutate(type = "model") ## Data with summarized mean

df_exe <- dplyr::bind_rows(df_b, df1, df2, df3)

lg_high <- langevitour::langevitour(df_exe[1:(length(df_exe)-1)], group = df_exe$type, pointSize = append(rep(4, NROW(df_b)), rep(2, NROW(df1) + NROW(df2) + NROW(df3))), levelColors = c("#6a3d9a", "#e6550d", "#000000", "#33a02c"), lineFrom = 1, lineTo = 2, lineColors = "#e31a1c")

```

```{r}
#| warning: false
#| echo: false
#| label: fig-wkhighD
#| fig-cap: How the 2D model lift into high dimensions? (a) visualize the points and the hexagonal bin centroids related $7^{th}$ and $12^{th}$ hexagons, (b) visualization of the edge connected the $7^{th}$ and $12^{th}$ hexagons (colored in red) in the triangular mesh. A video of tour animation is available at <https://youtu.be/hxU91xNTJL0>.

a2 + a1 +
  plot_annotation(tag_levels = 'a') +
  plot_layout(ncol = 3) &
  theme(legend.position='none', plot.tag = element_text(size = 8))
```

### Tuning the model

The performance and robustness of our model depend on four key parameters: (i) the total number of bins ($b$), (ii) a benchmark value used to remove low-density hexagons, (iii) a benchmark value used to remove long edges, and (iv) starting point of the hexagonal grid. However, there is no analytical formula to calculate an appropriate value for these parameters. The selection of these parameter values depends on the model performance computed by Mean Squared Error (MSE) (see @sec-goodfit).

#### Total number of bins

The number of hexagonal bins in the hexagonal grid has a considerable impact on the construction of the 2D model, serving as the initial step in building the 2D model. The chosen total number of bins must effectively capture the structure of the NLDR data. If the number of bins is too low, the model may not be able to capture the structure of the NLDR data effectively (see @fig-binsize (a)), while if there are too many bins, it may result in over-fitting the individual points of the NLDR data (see @fig-binsize (c)). Therefore, it is important to determine an appropriate number of bins to construct an effective model.

```{r}
#| echo: false
#| message: false
#| warning: false

num_bins_x <- 3
num_bins_y <- 4
hex_size <- 0.62

hb_obj <- hex_binning(data = umap_s_curve_scaled, x = "UMAP1", 
                      y = "UMAP2", num_bins_x = num_bins_x, 
                      num_bins_y = num_bins_y, x_start = NA, y_start = NA, 
                      buffer_x = NA, buffer_y = NA, hex_size = hex_size, col_start = "UMAP")

## Data set with all possible centroids in the hexagonal grid
all_centroids_df <- as.data.frame(do.call(cbind, hb_obj$centroids))
counts_df <- as.data.frame(do.call(cbind, hb_obj$std_cts))

## Generate all coordinates of hexagons
hex_grid <- as.data.frame(do.call(cbind, hb_obj$hex_poly))

hex_full_count_df_loop1 <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))

p1 <-  ggplot(data = hex_full_count_df_loop1, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
  geom_point(data = umap_s_curve_scaled, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
  #coord_equal() +
  coord_cartesian(xlim =c(-0.2, 1.2), ylim = c(-0.1, 2.5)) +
  theme_void() +
  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=8), #change legend title font size
        legend.text = element_text(size=6)) +
  guides(fill = guide_colourbar(title = "Standardized count")) +
  annotate(geom = 'text', label = "a", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)

```

```{r}
#| echo: false
#| message: false
#| warning: false

num_bins_x <- 6
num_bins_y <- 14
hex_size <- 0.13

hb_obj <- hex_binning(data = umap_s_curve_scaled, x = "UMAP1", 
                      y = "UMAP2", num_bins_x = num_bins_x, 
                      num_bins_y = num_bins_y, x_start = NA, y_start = NA, 
                      buffer_x = NA, buffer_y = NA, hex_size = hex_size, col_start = "UMAP")

## Data set with all possible centroids in the hexagonal grid
all_centroids_df <- as.data.frame(do.call(cbind, hb_obj$centroids))
counts_df <- as.data.frame(do.call(cbind, hb_obj$std_cts))

## Generate all coordinates of hexagons
hex_grid <- as.data.frame(do.call(cbind, hb_obj$hex_poly))

hex_full_count_df_loop2 <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))

p2 <-  ggplot(data = hex_full_count_df_loop2, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
  geom_point(data = umap_s_curve_scaled, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
  #coord_equal() +
  coord_cartesian(xlim =c(-0.2, 1.2), ylim = c(-0.1, 2.5)) +
  theme_void() +
  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=8), #change legend title font size
        legend.text = element_text(size=6)) +
  guides(fill = guide_colourbar(title = "Standardized count")) +
  annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)

```

```{r}
#| echo: false
#| message: false
#| warning: false

num_bins_x <- 12
num_bins_y <- 28
hex_size <- 0.06

hb_obj <- hex_binning(data = umap_s_curve_scaled, x = "UMAP1", 
                      y = "UMAP2", num_bins_x = num_bins_x, 
                      num_bins_y = num_bins_y, x_start = NA, y_start = NA, 
                      buffer_x = NA, buffer_y = NA, hex_size = hex_size, col_start = "UMAP")

## Data set with all possible centroids in the hexagonal grid
all_centroids_df <- as.data.frame(do.call(cbind, hb_obj$centroids))
counts_df <- as.data.frame(do.call(cbind, hb_obj$std_cts))

## Generate all coordinates of hexagons
hex_grid <- as.data.frame(do.call(cbind, hb_obj$hex_poly))

hex_full_count_df_loop3 <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))

p3 <-  ggplot(data = hex_full_count_df_loop3, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
  geom_point(data = umap_s_curve_scaled, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
  coord_equal() +
  #coord_cartesian(xlim =c(-1, 2.7), ylim = c(-2, 12.2)) +
  theme_void() +
  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=8), #change legend title font size
        legend.text = element_text(size=6)) +
  guides(fill = guide_colourbar(title = "Standardized count")) +
  annotate(geom = 'text', label = "c", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)

```

```{r}
#| echo: false
#| label: fig-binsize
#| fig-pos: H
#| fig-cap: "Hexbin plots from different number of bins for the **UMAP** embeddings of **S-curve** training data: (a) b = 12 (3, 4), s = 0.62, (b) b = 84 (6, 14), s = 0.13, and (c) b = 336 (12, 28), s = 0.06. The hexbins are colored based on the density of points, with darker colors indicating higher point density and yellow color representing lower point density within each bin. What is the number of bins that would be effective in representing low-dimensional data?"

p1 + p2 + p3
```

$$
b = b_1 \times b_2
$${#eq-equation8}

Furthermore, the total number of bins is calculated based on the number of bins along the x-axis and y-axis, as shown in @eq-equation8. To determine the effective total number of bins, candidate values are selected based on the range between the minimum and approximate maximum number of bins along the x and y axes. The minimum number of bins along each axis is set to $1$, while the maximum number is estimated by taking the square root of the NLDR data points. The analysis evaluates the MSE across varying total number of bins within this range, covering the minimum to maximum values along both axes (see @fig-diagnosticpltScurve).

```{r}
#| warning: false
#| echo: false
#| message: false

## UMAP

hex_size_vec <- seq(0.06, 2, by = 0.01)

vec <- stats::setNames(rep("", 6), c("num_bins", "aic", "mse", "num_bins_x", "num_bins_y", "hex_size"))  ## Define column names

mse_df <- dplyr::bind_rows(vec)[0, ]
mse_df <- mse_df |>
  dplyr::mutate_if(is.character, as.numeric)

for (i in 1:length(hex_size_vec)) {

  num_bin_list <- calc_bins(data = umap_s_curve_scaled, 
            x = "UMAP1", y = "UMAP2", 
            hex_size = hex_size_vec[i], buffer_x = NA, buffer_y = NA)

  num_bins_x <- num_bin_list$num_x
  num_bins_y <- num_bin_list$num_y

  model_object <- fit_highd_model( training_data = training_data, 
                                   nldr_df_with_id = umap_s_curve_scaled, 
                                   x = "UMAP1", y = "UMAP2", 
                                   num_bins_x = num_bins_x, 
                                   num_bins_y = num_bins_y, 
                                   x_start = NA, y_start = NA, 
                                   buffer_x = NA, buffer_y = NA, 
                                   hex_size = hex_size_vec[i],
                                   is_rm_lwd_hex = FALSE, 
                                   benchmark_to_rm_lwd_hex = NA, 
                                   col_start_2d = "UMAP", 
                                   col_start_highd = "x")

  centroid_df_training <- model_object$df_bin_centroids
  avg_df_training <- model_object$df_bin

  pred_emb_list <- predict_emb(test_data = training_data, 
                                  df_bin_centroids = centroid_df_training, 
                                  df_bin = avg_df_training, type_NLDR = "UMAP")

  pred_df_training <- as.data.frame(do.call(cbind, pred_emb_list))

  eval_list <- gen_summary(test_data = training_data, 
                                  prediction_df = pred_df_training, 
                                  df_bin = avg_df_training, col_start = "x")

  mse_df <- mse_df |>
    tibble::add_row(num_bins = num_bins_x * num_bins_y,
                    aic = eval_list$aic,
                    mse = eval_list$mse,
                    num_bins_x = num_bins_x,
                    num_bins_y = num_bins_y,
                    hex_size = hex_size_vec[i])

}

## If same total number of bins occurred only select ones with minimum error
### Obtain duplicate bins
dupli_bins <- mse_df |> 
  dplyr::count(num_bins) |> 
  dplyr::filter(n > 1) |> 
  dplyr::pull(num_bins)

### Group split by duplicated bins
duplicate_df_list <- mse_df |>
  dplyr::filter(num_bins %in% dupli_bins) |>
  dplyr::arrange(num_bins) |>
  dplyr::group_split(num_bins)

### Obtain one row from duplicates which have lowest error and hexsize
duplicate_df <- data.frame(matrix(nrow = 0, ncol = 0))

for (i in 1:length(duplicate_df_list)) {

  dd <- duplicate_df_list[[i]] |>
    dplyr::filter(mse == min(duplicate_df_list[[i]]$mse)) |>
    dplyr::filter(hex_size == min(duplicate_df_list[[i]]$hex_size))

  duplicate_df <- dplyr::bind_rows(duplicate_df, dd)

}

### Obtain the mse_df with not duplicated bins
not_dupli_df <- mse_df |>
  dplyr::filter(!(num_bins %in% dupli_bins))

### Combine duplicated and not duplicated(corrected) bins dfs
mse_df <- dplyr::bind_rows(not_dupli_df, duplicate_df)
```

```{r}
#| warning: false
#| echo: false

mse_plot <- ggplot(mse_df, aes(x = num_bins,
                                       y = mse
)) +
  geom_point() +
  geom_line() +
   geom_vline(xintercept = 84, linetype="solid",
                color = "red", size=0.8, alpha = 0.5) +
  theme_light() +
  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  #scale_colour_manual(values = c("#377eb8")) +
  ylab("MSE") +
  xlab("total number of bins")

```

```{r}
#| echo: false
#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset. What is the effective number of bins in each NLDR technique to create a 2D model? The MSE plot have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error. Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error. Using the elbow method, when the total number of bins is set to $84$, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape. This point indicates that adding less bins does not enough to capture the data structure."
#| label: fig-diagnosticpltScurve
##| out-width: 100%
#| fig-pos: H

mse_plot
```

```{r}
#| warning: false
#| echo: false
#| message: false

## Decide by looking at MSE plot
num_bins_x_s_curve <- 6
num_bins_y_s_curve <- 14
hex_size_s_curve <- 0.13

hb_obj_s_curve <- hex_binning(data = umap_s_curve_scaled, x = "UMAP1", 
                      y = "UMAP2", num_bins_x = num_bins_x_s_curve, 
                      num_bins_y = num_bins_y_s_curve, x_start = NA, y_start = NA, 
                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_s_curve, col_start = "UMAP")

## Data set with all possible centroids in the hexagonal grid
all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_s_curve$centroids))

## Generate all coordinates of hexagons
hex_grid <- as.data.frame(do.call(cbind, hb_obj_s_curve$hex_poly))

## To obtain the standardise counts within hexbins
counts_df <- as.data.frame(do.call(cbind, hb_obj_s_curve$std_cts))
df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
                                             counts_df = counts_df)

## remove low-density hexagons

hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))

hex_full_count_df_loop2 <- hex_grid_with_counts |>
  dplyr::mutate(type = if_else(std_counts <= 0.2, "low", "high"))

hex_full_count_df_filtered <- df_bin_centroids |>
  dplyr::mutate(type = if_else(std_counts <= 0.2, "low", "high")) |>
  dplyr::filter(type == "low")

p2_n <-  ggplot(data = hex_full_count_df_loop2, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
  #geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
  #coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
  theme_void() +
  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=8), #change legend title font size
        legend.text = element_text(size=6)) +
  guides(fill = guide_colourbar(title = "Standardized count")) #+
  #annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)

p2_n_rm <-  ggplot(data = hex_full_count_df_loop2, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
  geom_text(data = hex_full_count_df_filtered, color = "red", aes(x = c_x, y = c_y, group = hexID, label = "X", size = 0.5)) +
  #geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
  #coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
  theme_void() +
  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=8), #change legend title font size
        legend.text = element_text(size=6)) +
  guides(fill = guide_colourbar(title = "Standardized count")) #+
  #annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)

# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
#   coord_fixed()

umap_data_with_hb_id <- as.data.frame(do.call(cbind, hb_obj_s_curve$data_hb_id))

model_object <- fit_highd_model( training_data = training_data, 
                                 nldr_df_with_id = umap_s_curve_scaled, 
                                 x = "UMAP1", y = "UMAP2", 
                                 num_bins_x = num_bins_x_s_curve, 
                                 num_bins_y = num_bins_y_s_curve, 
                                 x_start = NA, y_start = NA, 
                                 buffer_x = NA, buffer_y = NA, 
                                 hex_size = hex_size_s_curve,
                                 is_rm_lwd_hex = FALSE, 
                                 benchmark_to_rm_lwd_hex = NA, 
                                 col_start_2d = "UMAP", 
                                 col_start_highd = "x")

df_bin_centroids_s_curve <- model_object$df_bin_centroids
df_bin_s_curve <- model_object$df_bin

cell_count_plot <- ggplot(df_bin_centroids_s_curve, aes(x = reorder(as.factor(hexID), -std_counts), y = std_counts)) +
  geom_quasirandom() + xlab("hexagonal id") + ylab("Standardized cell count") +
  geom_hline(yintercept = 0.2, colour = "#de2d26") +
  theme(axis.text = element_text(size = 5),
        axis.title = element_text(size = 7),
        axis.text.x = element_text(angle = 90))

benchmark_value_rm_lwd <- stats::quantile(df_bin_centroids_s_curve$std_counts, 
                probs = c(0,0.25,0.5,0.75,1), names = FALSE)[2]

df_bin_centroids_low <- df_bin_centroids_s_curve |>
  dplyr::filter(std_counts <= benchmark_value_rm_lwd)

identify_rm_bins <- find_low_dens_hex(df_bin_centroids_all = df_bin_centroids_s_curve, 
                                      num_bins_x = num_bins_x_s_curve, 
                                      df_bin_centroids_low = df_bin_centroids_low)

df_bin_centroids_s_curve <- df_bin_centroids_s_curve |>
  dplyr::filter(!(hexID %in% identify_rm_bins))

pcentroid_plot_s_curve <- ggplot(data = hex_grid, aes(x = x, y = y)) +
  geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
  geom_point(data = df_bin_centroids_s_curve, aes(x = c_x, y = c_y), color = "#33a02c") +
  coord_equal() +
  theme_linedraw() +
  theme(legend.position = "bottom", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

rm_centroid_df <- df_bin_centroids_low |> 
  dplyr::filter(hexID %in% identify_rm_bins)

p2_n_rm_correct <- ggplot(data = hex_full_count_df_loop2, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
  geom_text(data = hex_full_count_df_filtered[6, ], color = "red", aes(x = c_x, y = c_y, group = hexID, label = "X", size = 0.5)) +
  #geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
  #coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
  theme_void() +
  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=8), #change legend title font size
        legend.text = element_text(size=6)) +
  guides(fill = guide_colourbar(title = "Standardized count")) #+
  #annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)

## Triangulate bin centroids
tr1_object_s_curve <- tri_bin_centroids(df_bin_centroids_s_curve, x = "c_x", y = "c_y")
tr_from_to_df_s_curve <- gen_edges(tri_object = tr1_object_s_curve)

## Compute 2D distances
distance_s_curve <- cal_2d_dist(tr_coord_df = tr_from_to_df_s_curve, 
                             start_x = "x_from", start_y = "y_from", 
                             end_x = "x_to", end_y = "y_to", 
                             select_vars = c("from", "to", "distance"))

# 
# distance_plot

## To find the benchmark value
benchmark_s_curve <- find_lg_benchmark(distance_edges = distance_s_curve, distance_col = "distance")

distance_plot <- plot_dist(distance_s_curve) +
   geom_hline(yintercept = benchmark_s_curve, linetype="solid",
                color = "red", size=0.8, alpha = 0.5) +
  ylab(expression(d^{(2)})) +
  theme(axis.text = element_text(size = 5),
        axis.title = element_text(size = 12))

# ggplot() +
# geom_trimesh(data = df_bin_centroids_pbmc, mapping = aes(x = c_x, y = c_y))

trimesh_s_curve_umap <- vis_lg_mesh(distance_edges = distance_s_curve, benchmark_value = benchmark_s_curve, tr_coord_df = tr_from_to_df_s_curve, distance_col = "distance")

trimesh_s_curve_umap <- trimesh_s_curve_umap +
  theme_linedraw() +
  theme(legend.position = "bottom", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) +
  labs(colour = "")

trimesh_removed_s_curve_umap <- vis_rmlg_mesh(distance_edges = distance_s_curve, benchmark_value = benchmark_s_curve, tr_coord_df = tr_from_to_df_s_curve, distance_col = "distance")

trimesh_removed_s_curve_umap <- trimesh_removed_s_curve_umap +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

# df_all_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), umap_data_with_hb_id)
# 
# show_langevitour(df_all_s_curve, df_bin_s_curve, df_bin_centroids_s_curve, benchmark_s_curve, distance_s_curve, "distance", col_start = "x")
```

#### Benchmark value to remove low-density hexagons

After establishing the hexagonal grid with an appropriate number of bins, some hexagonal bins may have few or no data points within them (see @fig-rmlowdenshex (a)). To ensure comprehensive coverage of the NLDR data, it is necessary to select hexagonal bins with a considerable number of data points. This involves calculating the number of points within each hexagon. Then, the standard count is computed by dividing the number of points within each hexagon by the maximum number of points in the grid (see @eq-equationp2). Next, bins with a standard count less than a certain benchmark value are removed (see @fig-rmlowdenshex (b)). The following steps will help in determining a suitable value for removing low-density hexagons:

1. Plot the distribution of the standardized counts (see @fig-stdctsScurve).
2. Examine the distribution of counts.
3. Select the first quartile value if the distribution is skewed.

$$
\text{standard count} = \frac{\text{count}}{\text{max count}} 
$${#eq-equationp2}

Furthermore, selecting the benchmark value for removing low-density hexagons is important. Removing unnecessary bins may lead to the formation of long edges and an uneven 2D model. Hence, rather than solely relying on the benchmark value to identify hexagons for removal, it's essential to consider the standard number of points in the neighboring hexagons of the identified low-density ones (see @fig-rmlowdenshex (c)). If neighboring bins also show low counts, only those bins will be removed. The remaining bins are used to construct the 2D model.

The benchmark value for removing low-density hexagons ranges between $0$ and $1$. When analyzing how these benchmark values influence model performance, it's essential to observe the change in Mean Squared Error (MSE) as the benchmark value increases (see @fig-diagnosticpltScurvelwd). The MSE shows a gradual increase as the benchmark value progresses from $0$ to $1$. Evaluating this rate of increase is important. If the increment is not considerable, the decision might lean towards retaining low-density hexagons.

```{r}
#| echo: false
#| fig-cap: "Distribution of standardize counts by hexagons."
#| label: fig-stdctsScurve
#| fig-pos: H

cell_count_plot
```

```{r}
#| warning: false
#| echo: false
#| message: false
num_bins_x_s_curve <- 6
num_bins_y_s_curve <- 14
hex_size_s_curve <- 0.13

vec <- stats::setNames(rep("", 2), c("benchmark_val_lwd", "mse"))  ## Define column names

mse_df <- dplyr::bind_rows(vec)[0, ]
mse_df <- mse_df |>
  dplyr::mutate_if(is.character, as.numeric)

model_object <- fit_highd_model( training_data = training_data, 
                                 nldr_df_with_id = umap_s_curve_scaled, 
                                 x = "UMAP1", y = "UMAP2", 
                                 num_bins_x = num_bins_x_s_curve, 
                                 num_bins_y = num_bins_y_s_curve, 
                                 x_start = NA, y_start = NA, 
                                 buffer_x = NA, buffer_y = NA, 
                                 hex_size = hex_size_s_curve,
                                 is_rm_lwd_hex = FALSE, 
                                 benchmark_to_rm_lwd_hex = NA, 
                                 col_start_2d = "UMAP", 
                                 col_start_highd = "x")

centroid_df_training_all <- model_object$df_bin_centroids
avg_df_training_all <- model_object$df_bin

benchmark_rm_hex_vec <- seq(0, 1, by=0.1)
benchmark_rm_hex_vec <- append(benchmark_rm_hex_vec[1:10],0.99)

for (i in 1:length(benchmark_rm_hex_vec)) {

  centroid_df_training <- centroid_df_training_all |>
    dplyr::filter(std_counts > benchmark_rm_hex_vec[i])

  avg_df_training <- avg_df_training_all |>
    dplyr::filter(hb_id %in% centroid_df_training$hexID)

  pred_emb_list <- predict_emb(test_data = training_data, 
                               df_bin_centroids = centroid_df_training, 
                               df_bin = avg_df_training, type_NLDR = "UMAP")

  pred_df_training <- as.data.frame(do.call(cbind, pred_emb_list))

  eval_list <- gen_summary(test_data = training_data, 
                           prediction_df = pred_df_training, 
                           df_bin = avg_df_training, col_start = "x")

  mse_df <- mse_df |>
    tibble::add_row(benchmark_val_lwd = benchmark_rm_hex_vec[i],
                    mse = eval_list$mse)

}

mse_plot_lwd <- ggplot(mse_df, aes(x = benchmark_val_lwd,
                                       y = mse
)) +
  geom_point() +
  geom_line() +
   geom_vline(xintercept = 0.2, linetype="solid",
                color = "red", size=0.8, alpha = 0.5) +
  theme_light() +
  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  #scale_colour_manual(values = c("#377eb8")) +
  ylab("MSE") +
  xlab("benchmark value to remove low-density hexagons")

```

```{r}
#| echo: false
#| warning: false
#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset with different benchmark values to remove the low-density heaxgons. What is the effective benchmark value to remove the low-density heaxgons? The MSE plot have a steep slope at the beginning, indicating that a smaller benchmark causes a small amount of error. Then, the slope gradually increases or level up, indicating that a higher number of benchmark values generates a higher error. Using the reverse elbow method, when the benchmark value is set to 0.2, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape. This point indicates that higher benchmark values remove the necessary bins as well which lead to distruct the 2D structure."
#| label: fig-diagnosticpltScurvelwd
##| out-width: 100%
#| fig-pos: H

mse_plot_lwd
```

```{r}
#| echo: false
#| warning: false
#| fig-cap: "(a) Hexbin plot with colored hexbins based on point density, (b) Identification of low-density hexagons using a benchmark value of $0.2$, and (c) Identification of low-density hexagons considering neighboring bins."
#| label: fig-rmlowdenshex
#| fig-pos: H
##| out-width: 100%

p2_n + p2_n_rm + p2_n_rm_correct +
  plot_annotation(tag_levels = 'a') +
  plot_layout(guides='collect', ncol = 3) &
  theme(legend.position='none', plot.tag = element_text(size = 8))
```

#### Benchmark value for removing long edges

Creating a smooth 2D representation (see @fig-modelScurve (c)) requires removing edges that connect distant bin centroids within the triangular mesh (see @fig-modelScurve (b)). These edges only exist in the 2D model and do not extend into higher dimensions, ensuring that their removal does not impact the model in higher dimensions. While specific criteria for determining the benchmark value to remove long edges do not exist, the following steps provide an approach to identifying a suitable threshold:

1. Plot the distribution of the 2D Euclidean distances (see @fig-distScurve).
2. Identify the largest difference between consecutive distance values.
3. Take the largest distance value corresponding to this difference as the benchmark value.

```{r}
#| echo: false
#| fig-cap: "Distribution of 2D Euclidean distances between bin centroids of the triangular mesh generated with S-curve UMAP data."
#| label: fig-distScurve
#| fig-pos: H

distance_plot
```

```{r}
#| echo: false
#| fig-cap: "(a) Full hexagonal grid with bin centroids, (b) Model constructed in 2D with long edges, and (c) Model constructed in 2D after removing the long edges."
#| label: fig-modelScurve
#| fig-pos: H

pcentroid_plot_s_curve + trimesh_s_curve_umap + trimesh_removed_s_curve_umap +
  plot_layout(guides='collect', ncol = 3) &
  theme(legend.position='bottom') 
```

#### Starting point of the hexagonal grid

Hexagonal binning involves tessellating the $xy$ plane over the set defined by the range of $x$ and $y$ [@Dan2023]. Therefore, the starting point of the hexagonal grid is important parameter to consider. Having different starting point results in shifting the hexagonal grid and bin centroids (see @fig-scurveshifthexgridsexp). With a specific total number of bins (which corresponds to a specific range along the x and y axes), it's essential to explore how different starting points affect the model's performance (see @fig-mseScurvestartpoint). This is because different starting points result in varying distributions of non-empty bins (see @fig-scurveshifthexgrids), impacting the model.

```{r}
#| warning: false
#| echo: false
#| message: false

## UMAP
## Decide by looking at MSE plot
num_bins_x_s_curve <- 6
num_bins_y_s_curve <- 14
hex_size_s_curve <- 0.13

## Possible values for x_start
max_x_start <- min(umap_s_curve_scaled[["UMAP1"]]) + (sqrt(3) * hex_size_s_curve)
min_x_start <- min(umap_s_curve_scaled[["UMAP1"]]) - (sqrt(3) * hex_size_s_curve)

## Possible values for y_start
max_y_start <- min(umap_s_curve_scaled[["UMAP2"]]) + (1.5 * hex_size_s_curve)
min_y_start <- min(umap_s_curve_scaled[["UMAP2"]]) - (1.5 * hex_size_s_curve)

diff_start_df <- expand.grid(x_start = seq(min_x_start, max_x_start, length.out = 10), 
                             y_start = seq(min_y_start, max_y_start, length.out = 10)) |>
  add_row(x_start = min(umap_s_curve_scaled[["UMAP1"]]) - (sqrt(3) * hex_size_s_curve/2),
          y_start = min(umap_s_curve_scaled[["UMAP2"]]) - (1.5 * hex_size_s_curve/2))

vec <- stats::setNames(rep("", 4), c("x_start", "y_start", "num_non_empty_bins", "mse"))  ## Define column names

mse_df <- dplyr::bind_rows(vec)[0, ]
mse_df <- mse_df |>
  dplyr::mutate_if(is.character, as.numeric)

diff_start_df <- diff_start_df[c(1:5, 11:15,21:25, 101),]

for (i in 1:NROW(diff_start_df)) { #NROW(diff_start_df)

  model_object <- fit_highd_model( training_data = training_data, 
                                   nldr_df_with_id = umap_s_curve_scaled, 
                                   x = "UMAP1", y = "UMAP2", 
                                   num_bins_x = num_bins_x_s_curve, 
                                   num_bins_y = num_bins_y_s_curve, 
                                   x_start = diff_start_df[i, 1], y_start = diff_start_df[i, 2], 
                                   buffer_x = NA, buffer_y = NA, 
                                   hex_size = hex_size_s_curve,
                                   is_rm_lwd_hex = FALSE, 
                                   benchmark_to_rm_lwd_hex = NA, 
                                   col_start_2d = "UMAP", 
                                   col_start_highd = "x")

  centroid_df_training <- model_object$df_bin_centroids
  avg_df_training <- model_object$df_bin

  pred_emb_list <- predict_emb(test_data = training_data, 
                               df_bin_centroids = centroid_df_training, 
                               df_bin = avg_df_training, type_NLDR = "UMAP")

  pred_df_training <- as.data.frame(do.call(cbind, pred_emb_list))

  eval_list <- gen_summary(test_data = training_data, 
                           prediction_df = pred_df_training, 
                           df_bin = avg_df_training, col_start = "x")

  mse_df <- mse_df |>
    tibble::add_row(x_start = diff_start_df[i, 1], 
                    y_start = diff_start_df[i, 2],
                    num_non_empty_bins = NROW(centroid_df_training), 
                    mse = eval_list$mse)

}

# ## If same total number of bins occurred only select ones with minimum error
# ### Obtain duplicate bins
# dupli_bins <- mse_df |> 
#   dplyr::count(num_non_empty_bins) |> 
#   dplyr::filter(n > 1) |> 
#   dplyr::pull(num_non_empty_bins)
# 
# ### Group split by duplicated bins
# duplicate_df_list <- mse_df |>
#   dplyr::filter(num_non_empty_bins %in% dupli_bins) |>
#   dplyr::arrange(num_non_empty_bins) |>
#   dplyr::group_split(num_non_empty_bins)
# 
# ### Obtain one row from duplicates which have lowest error and hexsize
# duplicate_df <- data.frame(matrix(nrow = 0, ncol = 0))
# 
# for (i in 1:length(duplicate_df_list)) {
#   
#   dd <- duplicate_df_list[[i]] |>
#     dplyr::filter(mse == min(duplicate_df_list[[i]]$mse)) 
#   
#   duplicate_df <- dplyr::bind_rows(duplicate_df, dd)
#   
# }
# 
# ### Obtain the mse_df with not duplicated bins
# not_dupli_df <- mse_df |>
#   dplyr::filter(!(num_non_empty_bins %in% dupli_bins))
# 
# ### Combine duplicated and not duplicated(corrected) bins dfs
# mse_df <- dplyr::bind_rows(not_dupli_df, duplicate_df)

mse_plot_start_point <- ggplot(mse_df, aes(x = x_start,
                               y = y_start, colour = mse
)) +
  geom_point() +
  #geom_line() +
  # geom_vline(xintercept = 84, linetype="solid",
  #            color = "red", linewidth=0.8, alpha = 0.5) +
  theme_light() +
  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  viridis::scale_color_viridis(option = "D") +
  labs(x = "x start", y = "y start", color = "MSE") 
```

```{r}
#| echo: false
#| warning: false
#| fig-cap: "Goodness of fit statistics from UMAP applied to training S-curve dataset with different benchmark values to remove the low-density heaxgons. What is the effective benchmark value to remove the low-density heaxgons? The MSE plot have a steep slope at the beginning, indicating that a smaller benchmark causes a small amount of error. Then, the slope gradually increases or level up, indicating that a higher number of benchmark values generates a higher error. Using the reverse elbow method, when the benchmark value is set to 0.2, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape. This point indicates that higher benchmark values remove the necessary bins as well which lead to distruct the 2D structure."
#| label: fig-mseScurvestartpoint
#| fig-pos: H

mse_plot_start_point
```

```{r}
#| echo: false
#| warning: false

## Initial x and y start
initial_start_plot <- ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
  geom_text(data = df_bin_centroids, aes(x = c_x, y = c_y, label = hexID), size = 2) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
  coord_equal() +
  theme_void() +
  theme(legend.position="bottom", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=8), #change legend title font size
        legend.text = element_text(size=6)) +
  guides(fill = guide_colourbar(title = "Standardized count")) +
  annotate(geom = 'text', label = "a", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)

## x and y start selected by MSE

## UMAP
## Decide by looking at MSE plot
num_bins_x_s_curve <- 6
num_bins_y_s_curve <- 14
hex_size_s_curve <- 0.13

hb_obj_s_curve_n <- hex_binning(data = umap_s_curve_scaled, x = "UMAP1", 
                      y = "UMAP2", num_bins_x = num_bins_x_s_curve, 
                      num_bins_y = num_bins_y_s_curve, x_start = -0.0652, y_start = -0.195, 
                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_s_curve, col_start = "UMAP")

## Data set with all possible centroids in the hexagonal grid
all_centroids_df_n <- as.data.frame(do.call(cbind, hb_obj_s_curve_n$centroids))

## Generate all coordinates of hexagons
hex_grid_n <- as.data.frame(do.call(cbind, hb_obj_s_curve_n$hex_poly))

## To obtain the standardise counts within hexbins
counts_df_n <- as.data.frame(do.call(cbind, hb_obj_s_curve_n$std_cts))
df_bin_centroids_n <- extract_hexbin_centroids(centroids_df = all_centroids_df_n, 
                                             counts_df = counts_df_n)

## remove low-density hexagons

hex_grid_with_counts_n <- dplyr::left_join(hex_grid_n, counts_df_n, by = c("hex_poly_id" = "hb_id"))

new_start_plot <- ggplot(data = hex_grid_with_counts_n, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
  geom_text(data= df_bin_centroids_n, aes(x = c_x, y = c_y, label = hexID), size = 2) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
  coord_equal() +
  theme_void() +
  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=8), #change legend title font size
        legend.text = element_text(size=6)) +
  guides(fill = guide_colourbar(title = "Standardized count")) +
  annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)

hex_grid_ch_plot <- ggplot(data = hex_grid_with_counts_n, aes(x = x, y = y)) +
  geom_polygon(fill = NA, color = "#feb24c", aes(group = hex_poly_id)) +
  geom_polygon(data = hex_grid_with_counts, aes(x = x, y = y, group = hex_poly_id),
               fill = NA, color = "#bdbdbd") +
  #coord_cartesian(xlim = c(-5, 8), ylim = c(-10, 10)) +
  theme_void() +
  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=8), #change legend title font size
        legend.text = element_text(size=6)) +
  guides(fill = guide_colourbar(title = "Standardized count")) +
  annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)

bin_centroids_ch_plot <- ggplot(data = all_centroids_df, aes(x = c_x, y = c_y)) +
  geom_point(color = "#bdbdbd") +
  geom_point(data = all_centroids_df_n, aes(x = c_x, y = c_y), color = "#feb24c") +
  #coord_cartesian(xlim = c(-5, 8), ylim = c(-10, 10)) +
  theme_void() +
  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=8), #change legend title font size
        legend.text = element_text(size=6)) +
  guides(fill = guide_colourbar(title = "Standardized count")) +
  annotate(geom = 'text', label = "a", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3) 
```

```{r}
#| echo: false
#| fig-cap: "(a) Visualization of bin centroids and (b) hexagonal grid, showcasing the configuration before (colored in gray) and after (colored in light orange) shifting the starting point by an identical amount applied in both x and y directions, with a shift value of $0.537285$."
#| label: fig-scurveshifthexgridsexp
#| fig-pos: H

bin_centroids_ch_plot + hex_grid_ch_plot +
  plot_layout(guides='collect', ncol = 2) &
  theme(legend.position='none', plot.tag = element_text(size = 8))
```

```{r}
#| echo: false
#| fig-cap: "Hexbin plots shows the distribution of points before and after shifting the starting point of the hexagonal grid generated for UMAP applied to the training S-curve dataset. (a) Hexbin plot before shifting, and (b) Hexbin plot after shifting (identical shift applied in both x and y directions, with a shift value of $0.537285$)."
#| label: fig-scurveshifthexgrids
#| fig-pos: H

initial_start_plot + new_start_plot +
  plot_layout(guides='collect', ncol = 2) &
  theme(legend.position='none', plot.tag = element_text(size = 8))
```

### Model summaries {#sec-summary}

#### Predicted values and residuals

The prediction approach involves performing the K-nearest neighbors (KNN) algorithm for an unsupervised classification problem. First, the nearest high-D model point is identified for a given new high-D point. Then, the corresponding 2D centroid mapping for the identified high-D model point is determined. Finally, the coordinates of this 2D centroid are used as the predicted 2D embedding for the new high-D data point. This step is particularly valuable due to the limitations of some NLDR techniques, like tSNE, which don't provide a straightforward method for prediction. As a result, our approach offers a solution that capable of generating predicted 2D embedding regardless of the NLDR technique employed, effectively addressing this functional gap.

Residuals are essential for evaluating the accuracy of representing high-D points by the high-D mapping of 2D bin centroids. To measure this accuracy, an error metric is introduced, quantifying the sum of squared differences between the high-D data ($x_{ij}$) and the high-D mapping of the 2D bin centroid data ($C_{x_ij}$) across all observations and dimensions (see @eq-equation10).

$$
\text{Error} = \sum_{j = 1}^{n}\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2
$$ {#eq-equation10}

Here, $n$ represents the number of observations, $p$ represents the dimensions of high-D data, $x_{ij}$ is the high-D data, and $C_{x_ij}$ is the high-D mapping of the 2D bin centroid.

#### Goodness of fit statistics {#sec-goodfit}

To assess how well our method captures and represents the underlying structure of the high-D data, Mean Squared Error (MSE) is used. When computing MSE, total model error (see @sec-summary) is divided by the number of observations to make it as a mean value (see @eq-equation9).

$$
\text{MSE} = \sum_{j = 1}^{n} \frac{\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2}{n}
$$ {#eq-equation9}

### Simulated data example {#sec-simpleex}

In this section, the effectiveness of the algorithm is described using a simulated dataset. The dataset consists of five spherical Gaussian clusters in 4-$d$, with each cluster containing an equal number of points and the same within-cluster variation.

In the 2D layouts generated by various NLDR techniques, as shown in @fig-nldervis5Gau, five distinct clusters are observable, except for PHATE. In tSNE (see @fig-nldervis5Gau (a)), these clusters appear closely. UMAP arranges all clusters in a parallel manner, with three aligned in one line and the other two in a separate line (see @fig-nldervis5Gau (b)). In contrast, PHATE shows two closely positioned clusters and three more distant ones (see @fig-nldervis5Gau (c)). PaCMAP shows one central cluster and the remaining four spread out in different directions (see @fig-nldervis5Gau (e)). Finally, in TriMAP, two clusters are close, though not as tightly as PHATE, while the other three are well-separated (see @fig-nldervis5Gau (d)).

<!-- five Gaussian clusters with different NLDR techniques-->
```{r}
#| warning: false
#| echo: false

## Import data
df_2 <- read_rds("data/five_gau_clusters/data_five_gau.rds")
training_data_gau <- read_rds("data/five_gau_clusters/data_five_gau_training.rds")
test_data_gau <- read_rds("data/five_gau_clusters/data_five_gau_test.rds")

tSNE_data_gau <- read_rds("data/five_gau_clusters/tsne_data_five_gau_61.rds")
UMAP_data_gau <- read_rds("data/five_gau_clusters/umap_data_five_gau.rds")
PHATE_data_gau <- read_rds("data/five_gau_clusters/phate_data_five_gau.rds")
TriMAP_data_gau <- read_rds("data/five_gau_clusters/trimap_data_five_gau.rds")
PaCMAP_data_gau <- read_rds("data/five_gau_clusters/pacmap_data_five_gau.rds")

## Visualise embeddings

plot_list1_gau <- plot_tSNE_2D(tSNE_data_gau) + 
  geom_point(size = 0.00001, colour = "#e41a1c") +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

plot_list2_gau <- plot_UMAP_2D(UMAP_data_gau) + 
  geom_point(size = 0.00001, colour = "#377eb8") +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)

plot_list3_gau <- plot_PHATE_2D(PHATE_data_gau) + 
  geom_point(size = 0.00001, colour = "#4daf4a") +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'c', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)

plot_list4_gau <- plot_TriMAP_2D(TriMAP_data_gau) + 
  geom_point(size = 0.00001, colour = "#984ea3") +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)

plot_list5_gau <- plot_PaCMAP_2D(PaCMAP_data_gau) + 
  geom_point(size = 0.00001, colour = "#ff7f00") +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'e', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
```

```{r}
#| echo: false
#| fig-cap: "2D layouts from different NLDR techniques applied the same data: (a) tSNE (perplexity = 61), (b) UMAP (n_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2). Is there a best representation of the original data or are they all providing  equivalent information?"
#| label: fig-nldervis5Gau
#| fig-pos: H
#| out-width: 100%

plot_list1_gau + plot_list2_gau + plot_list3_gau + plot_list4_gau + plot_list5_gau +
  plot_layout(ncol=5)
```

To investigate which is the best representation to visualize the original data or all NLDR methods provide equivalent information, we visualize the model-in-data space. Models from all NLDR methods show five well-separated clusters (see @fig-gau1-sc, @fig-gau2-sc, @fig-gau3-sc, @fig-gau4-sc, and @fig-gau5-sc). This suggests that for the five Gaussian cluster dataset, all NLDR methods effectively preserve the global structure. tSNE, UMAP, and PHATE display clusters with varying densities, indicating their ability to capture within-cluster variation (see @fig-gau1-sc, @fig-gau2-sc, and @fig-gau3-sc). On the other hand, both TriMAP and PaCMAP show clusters with flat surfaces, suggesting a failure to capture within-cluster variation (see @fig-gau4-sc and @fig-gau5-sc). Therefore, TriMAP and PaCMAP do not capture the local structure as effectively as other methods.

```{r}
#| warning: false
#| echo: false
#| message: false

tsne_gau_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tSNE_data_gau, 
                                    x = "tSNE1", y = "tSNE2"))) |>
  dplyr::rename(c("tSNE1" = "scaled_tSNE1", 
                  "tSNE2" = "scaled_tSNE2")) |>
  dplyr::mutate(ID = 1:NROW(tSNE_data_gau))

## Decide by looking at MSE plot
num_bins_x_tsne_gau <- 21
num_bins_y_tsne_gau <- 28
hex_size_tsne_gau <- 0.03
## non-empty:198

hb_obj_tsne_gau <- hex_binning(data = tsne_gau_scaled, x = "tSNE1", 
                      y = "tSNE2", num_bins_x = num_bins_x_tsne_gau, 
                      num_bins_y = num_bins_y_tsne_gau, x_start = NA, y_start = NA, 
                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_tsne_gau, col_start = "tSNE")

## Data set with all possible centroids in the hexagonal grid
all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$centroids))

## Generate all coordinates of hexagons
hex_grid <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$hex_poly))

## To obtain the standardise counts within hexbins
counts_df <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$std_cts))
df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
                                             counts_df = counts_df)
# ggplot(data = hex_grid, aes(x = x, y = y)) +
#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
#   coord_fixed()

hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))

# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
#   coord_fixed()

tsne_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_tsne_gau$data_hb_id))

model_object_tsne_gau <- fit_highd_model( training_data = training_data_gau, 
                                 nldr_df_with_id = tsne_gau_scaled, 
                                 x = "tSNE1", y = "tSNE2", 
                                 num_bins_x = num_bins_x_tsne_gau, 
                                 num_bins_y = num_bins_y_tsne_gau, 
                                 x_start = NA, y_start = NA, 
                                 buffer_x = NA, buffer_y = NA, 
                                 hex_size = hex_size_tsne_gau,
                                 is_rm_lwd_hex = FALSE, 
                                 benchmark_to_rm_lwd_hex = NA, 
                                 col_start_2d = "tSNE", 
                                 col_start_highd = "x")

df_bin_centroids_tsne_gau <- model_object_tsne_gau$df_bin_centroids
df_bin_tsne_gau <- model_object_tsne_gau$df_bin

## Triangulate bin centroids
tr1_object_tsne_gau <- tri_bin_centroids(df_bin_centroids_tsne_gau, x = "c_x", y = "c_y")
tr_from_to_df_tsne_gau <- gen_edges(tri_object = tr1_object_tsne_gau)

## Compute 2D distances
distance_tsne_gau <- cal_2d_dist(tr_coord_df = tr_from_to_df_tsne_gau, 
                             start_x = "x_from", start_y = "y_from", 
                             end_x = "x_to", end_y = "y_to", 
                             select_vars = c("from", "to", "distance"))

# distance_plot <- plot_dist(distance_tsne_gau) +
#   ylab(expression(d^{(2)})) +
#   theme(axis.text = element_text(size = 5),
#         axis.title = element_text(size = 12))
# 
# distance_plot

## To find the benchmark value
benchmark_tsne_gau <- find_lg_benchmark(distance_edges = distance_tsne_gau, distance_col = "distance")

# ggplot() +
# geom_trimesh(data = df_bin_centroids_tsne_gau, mapping = aes(x = c_x, y = c_y))

trimesh_removed_tsne_gau <- vis_rmlg_mesh(distance_edges = distance_tsne_gau, benchmark_value = benchmark_tsne_gau, tr_coord_df = tr_from_to_df_tsne_gau, distance_col = "distance")

trimesh_removed_tsne_gau <- trimesh_removed_tsne_gau +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

df_all_tsne_gau <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID), tsne_data_with_hb_id_gau)

# show_langevitour(df = df_all_tsne_gau, df_b = df_bin_tsne_gau, df_b_with_center_data = df_bin_centroids_tsne_gau, col_start = "x", distance_df = distance_tsne_gau, distance_col = "distance", benchmark_value = benchmark_tsne_gau)
```

```{r}
#| warning: false
#| echo: false
#| message: false

umap_gau_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = UMAP_data_gau, 
                                    x = "UMAP1", y = "UMAP2"))) |>
  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
                  "UMAP2" = "scaled_UMAP2")) |>
  dplyr::mutate(ID = 1:NROW(UMAP_data_gau))

## Decide by looking at MSE plot
num_bins_x_umap_gau <- 60
num_bins_y_umap_gau <- 79
hex_size_umap_gau <- 0.01
### non-empty 229

hb_obj_umap_gau <- hex_binning(data = umap_gau_scaled, x = "UMAP1", 
                               y = "UMAP2", num_bins_x = num_bins_x_umap_gau, 
                               num_bins_y = num_bins_y_umap_gau, x_start = NA, y_start = NA, 
                               buffer_x = NA, buffer_y = NA, hex_size = hex_size_umap_gau, col_start = "UMAP")

## Data set with all possible centroids in the hexagonal grid
all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_umap_gau$centroids))

## Generate all coordinates of hexagons
hex_grid <- as.data.frame(do.call(cbind, hb_obj_umap_gau$hex_poly))

## To obtain the standardise counts within hexbins
counts_df <- as.data.frame(do.call(cbind, hb_obj_umap_gau$std_cts))
df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
                                             counts_df = counts_df)
# ggplot(data = hex_grid, aes(x = x, y = y)) +
#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
#   coord_fixed()

hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))

# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
#   coord_fixed()

umap_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_umap_gau$data_hb_id))

model_object_umap_gau <- fit_highd_model( training_data = training_data_gau, 
                                          nldr_df_with_id = umap_gau_scaled, 
                                          x = "UMAP1", y = "UMAP2", 
                                          num_bins_x = num_bins_x_umap_gau, 
                                          num_bins_y = num_bins_y_umap_gau, 
                                          x_start = NA, y_start = NA, 
                                          buffer_x = NA, buffer_y = NA, 
                                          hex_size = hex_size_umap_gau,
                                          is_rm_lwd_hex = FALSE, 
                                          benchmark_to_rm_lwd_hex = NA, 
                                          col_start_2d = "UMAP", 
                                          col_start_highd = "x")

df_bin_centroids_umap_gau <- model_object_umap_gau$df_bin_centroids
df_bin_umap_gau <- model_object_umap_gau$df_bin

## Triangulate bin centroids
tr1_object_umap_gau <- tri_bin_centroids(df_bin_centroids_umap_gau, x = "c_x", y = "c_y")
tr_from_to_df_umap_gau <- gen_edges(tri_object = tr1_object_umap_gau)

## Compute 2D distances
distance_umap_gau <- cal_2d_dist(tr_coord_df = tr_from_to_df_umap_gau, 
                                 start_x = "x_from", start_y = "y_from", 
                                 end_x = "x_to", end_y = "y_to", 
                                 select_vars = c("from", "to", "distance"))

# distance_plot <- plot_dist(distance_umap_gau) +
#   ylab(expression(d^{(2)})) +
#   theme(axis.text = element_text(size = 5),
#         axis.title = element_text(size = 12))
# 
# distance_plot

## To find the benchmark value
benchmark_umap_gau <- find_lg_benchmark(distance_edges = distance_umap_gau, distance_col = "distance")

# ggplot() +
# geom_trimesh(data = df_bin_centroids_umap_gau, mapping = aes(x = c_x, y = c_y))

trimesh_removed_umap_gau <- vis_rmlg_mesh(distance_edges = distance_umap_gau, benchmark_value = benchmark_umap_gau, tr_coord_df = tr_from_to_df_umap_gau, distance_col = "distance")

trimesh_removed_umap_gau <- trimesh_removed_umap_gau +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)

df_all_umap_gau <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID), umap_data_with_hb_id_gau)

# show_langevitour(df = df_all_umap_gau, df_b = df_bin_umap_gau, df_b_with_center_data = df_bin_centroids_umap_gau, col_start = "x", distance_df = distance_umap_gau, distance_col = "distance", benchmark_value = benchmark_umap_gau)
```

```{r}
#| warning: false
#| echo: false
#| message: false

# phate_gau_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tibble::as_tibble(PHATE_data_gau), x = "PHATE1", y = "PHATE2", hex_ratio = 50/sqrt(3)))) |>
#   dplyr::rename(c("PHATE1" = "scaled_PHATE1",
#                   "PHATE2" = "scaled_PHATE2")) |>
#   dplyr::mutate(ID = 1:NROW(PHATE_data_gau))
# 
# ## Decide by looking at MSE plot
# num_bins_x_phate_gau <- 60
# num_bins_y_phate_gau <- 1927
# hex_size_phate_gau <- 0.01
# ### non-empty 187
# 
# hb_obj_phate_gau <- hex_binning(data = phate_gau_scaled, x = "PHATE1", 
#                                y = "PHATE2", num_bins_x = num_bins_x_phate_gau, 
#                                num_bins_y = num_bins_y_phate_gau, x_start = NA, y_start = NA, 
#                                buffer_x = NA, buffer_y = NA, hex_size = hex_size_phate_gau, col_start = "PHATE")
# 
# ## Data set with all possible centroids in the hexagonal grid
# all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_phate_gau$centroids))
# 
# ## Generate all coordinates of hexagons
# hex_grid <- as.data.frame(do.call(cbind, hb_obj_phate_gau$hex_poly))
# 
# ## To obtain the standardise counts within hexbins
# counts_df <- as.data.frame(do.call(cbind, hb_obj_phate_gau$std_cts))
# df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
#                                              counts_df = counts_df)
# # ggplot(data = hex_grid, aes(x = x, y = y)) +
# #   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
# #   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red", size = 0.1) +
# #   coord_fixed()
# 
# hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
# 
# # ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
# #   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
# #   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
# #   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
# #   coord_fixed()
# 
# phate_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_phate_gau$data_hb_id))
# 
# model_object_phate_gau <- fit_highd_model( training_data = training_data_gau, 
#                                           nldr_df_with_id = phate_gau_scaled, 
#                                           x = "PHATE1", y = "PHATE2", 
#                                           num_bins_x = num_bins_x_phate_gau, 
#                                           num_bins_y = num_bins_y_phate_gau, 
#                                           x_start = NA, y_start = NA, 
#                                           buffer_x = NA, buffer_y = NA, 
#                                           hex_size = hex_size_phate_gau,
#                                           is_rm_lwd_hex = FALSE, 
#                                           benchmark_to_rm_lwd_hex = NA, 
#                                           col_start_2d = "PHATE", 
#                                           col_start_highd = "x")
# 
# df_bin_centroids_phate_gau <- model_object_phate_gau$df_bin_centroids
# df_bin_phate_gau <- model_object_phate_gau$df_bin

df_bin_centroids_phate_gau <- read_rds("data/five_gau_clusters/df_bin_centroids_phate_gau.rds")
df_bin_phate_gau <- read_rds("data/five_gau_clusters/df_bin_phate_gau.rds")

## Triangulate bin centroids
tr1_object_phate_gau <- tri_bin_centroids(df_bin_centroids_phate_gau, x = "c_x", y = "c_y")
tr_from_to_df_phate_gau <- gen_edges(tri_object = tr1_object_phate_gau)

## Compute 2D distances
distance_phate_gau <- cal_2d_dist(tr_coord_df = tr_from_to_df_phate_gau, 
                                 start_x = "x_from", start_y = "y_from", 
                                 end_x = "x_to", end_y = "y_to", 
                                 select_vars = c("from", "to", "distance"))

# distance_plot <- plot_dist(distance_phate_gau) +
#   ylab(expression(d^{(2)})) +
#   theme(axis.text = element_text(size = 5),
#         axis.title = element_text(size = 12))
# 
# distance_plot

## To find the benchmark value
benchmark_phate_gau <- find_lg_benchmark(distance_edges = distance_phate_gau, distance_col = "distance")

# ggplot() +
# geom_trimesh(data = df_bin_centroids_phate_gau, mapping = aes(x = c_x, y = c_y))

trimesh_removed_phate_gau <- vis_rmlg_mesh(distance_edges = distance_phate_gau, benchmark_value = benchmark_phate_gau, tr_coord_df = tr_from_to_df_phate_gau, distance_col = "distance")

trimesh_removed_phate_gau <- trimesh_removed_phate_gau +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'c', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)

# df_all_phate_gau <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID), phate_data_with_hb_id_gau)

# show_langevitour(df = df_all_phate_gau, df_b = df_bin_phate_gau, df_b_with_center_data = df_bin_centroids_phate_gau, col_start = "x", distance_df = distance_phate_gau, distance_col = "distance", benchmark_value = benchmark_phate_gau)

```

```{r}
#| warning: false
#| echo: false
#| message: false

trimap_gau_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = TriMAP_data_gau, 
                                    x = "TriMAP1", y = "TriMAP2"))) |>
  dplyr::rename(c("TriMAP1" = "scaled_TriMAP1", 
                  "TriMAP2" = "scaled_TriMAP2")) |>
  dplyr::mutate(ID = 1:NROW(TriMAP_data_gau))

## Decide by looking at MSE plot
num_bins_x_trimap_gau <- 46
num_bins_y_trimap_gau <- 61
hex_size_trimap_gau <- 0.013
### non-empty 173

hb_obj_trimap_gau <- hex_binning(data = trimap_gau_scaled, x = "TriMAP1", 
                               y = "TriMAP2", num_bins_x = num_bins_x_trimap_gau, 
                               num_bins_y = num_bins_y_trimap_gau, x_start = NA, y_start = NA, 
                               buffer_x = NA, buffer_y = NA, hex_size = hex_size_trimap_gau, col_start = "TriMAP")

## Data set with all possible centroids in the hexagonal grid
all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_trimap_gau$centroids))

## Generate all coordinates of hexagons
hex_grid <- as.data.frame(do.call(cbind, hb_obj_trimap_gau$hex_poly))

## To obtain the standardise counts within hexbins
counts_df <- as.data.frame(do.call(cbind, hb_obj_trimap_gau$std_cts))
df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
                                             counts_df = counts_df)
# ggplot(data = hex_grid, aes(x = x, y = y)) +
#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
#   coord_fixed()

hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))

# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
#   coord_fixed()

trimap_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_trimap_gau$data_hb_id))

model_object_trimap_gau <- fit_highd_model( training_data = training_data_gau, 
                                          nldr_df_with_id = trimap_gau_scaled, 
                                          x = "TriMAP1", y = "TriMAP2", 
                                          num_bins_x = num_bins_x_trimap_gau, 
                                          num_bins_y = num_bins_y_trimap_gau, 
                                          x_start = NA, y_start = NA, 
                                          buffer_x = NA, buffer_y = NA, 
                                          hex_size = hex_size_trimap_gau,
                                          is_rm_lwd_hex = FALSE, 
                                          benchmark_to_rm_lwd_hex = NA, 
                                          col_start_2d = "TriMAP", 
                                          col_start_highd = "x")

df_bin_centroids_trimap_gau <- model_object_trimap_gau$df_bin_centroids
df_bin_trimap_gau <- model_object_trimap_gau$df_bin

## Triangulate bin centroids
tr1_object_trimap_gau <- tri_bin_centroids(df_bin_centroids_trimap_gau, x = "c_x", y = "c_y")
tr_from_to_df_trimap_gau <- gen_edges(tri_object = tr1_object_trimap_gau)

## Compute 2D distances
distance_trimap_gau <- cal_2d_dist(tr_coord_df = tr_from_to_df_trimap_gau, 
                                 start_x = "x_from", start_y = "y_from", 
                                 end_x = "x_to", end_y = "y_to", 
                                 select_vars = c("from", "to", "distance"))

# distance_plot <- plot_dist(distance_trimap_gau) +
#   ylab(expression(d^{(2)})) +
#   theme(axis.text = element_text(size = 5),
#         axis.title = element_text(size = 12))
# 
# distance_plot

## To find the benchmark value
benchmark_trimap_gau <- find_lg_benchmark(distance_edges = distance_trimap_gau, distance_col = "distance")
benchmark_trimap_gau <- 0.2

# ggplot() +
# geom_trimesh(data = df_bin_centroids_trimap_gau, mapping = aes(x = c_x, y = c_y))

trimesh_removed_trimap_gau <- vis_rmlg_mesh(distance_edges = distance_trimap_gau, benchmark_value = benchmark_trimap_gau, tr_coord_df = tr_from_to_df_trimap_gau, distance_col = "distance")

trimesh_removed_trimap_gau <- trimesh_removed_trimap_gau +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'd', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

df_all_trimap_gau <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID), trimap_data_with_hb_id_gau)

# show_langevitour(df = df_all_trimap_gau, df_b = df_bin_trimap_gau, df_b_with_center_data = df_bin_centroids_trimap_gau, col_start = "x", distance_df = distance_trimap_gau, distance_col = "distance", benchmark_value = benchmark_trimap_gau)

```

```{r}
#| warning: false
#| echo: false
#| message: false

pacmap_gau_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = PaCMAP_data_gau, 
                                    x = "PaCMAP1", y = "PaCMAP2"))) |>
  dplyr::rename(c("PaCMAP1" = "scaled_PaCMAP1", 
                  "PaCMAP2" = "scaled_PaCMAP2")) |>
  dplyr::mutate(ID = 1:NROW(PaCMAP_data_gau))

## Decide by looking at MSE plot
num_bins_x_pacmap_gau <- 31
num_bins_y_pacmap_gau <- 40
hex_size_pacmap_gau <- 0.02
### non-empty 179

hb_obj_pacmap_gau <- hex_binning(data = pacmap_gau_scaled, x = "PaCMAP1", 
                               y = "PaCMAP2", num_bins_x = num_bins_x_pacmap_gau, 
                               num_bins_y = num_bins_y_pacmap_gau, x_start = NA, y_start = NA, 
                               buffer_x = NA, buffer_y = NA, hex_size = hex_size_pacmap_gau, col_start = "PaCMAP")

## Data set with all possible centroids in the hexagonal grid
all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_pacmap_gau$centroids))

## Generate all coordinates of hexagons
hex_grid <- as.data.frame(do.call(cbind, hb_obj_pacmap_gau$hex_poly))

## To obtain the standardise counts within hexbins
counts_df <- as.data.frame(do.call(cbind, hb_obj_pacmap_gau$std_cts))
df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
                                             counts_df = counts_df)
# ggplot(data = hex_grid, aes(x = x, y = y)) +
#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
#   coord_fixed()

hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))

# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
#   coord_fixed()

pacmap_data_with_hb_id_gau <- as.data.frame(do.call(cbind, hb_obj_pacmap_gau$data_hb_id))

model_object_pacmap_gau <- fit_highd_model( training_data = training_data_gau, 
                                          nldr_df_with_id = pacmap_gau_scaled, 
                                          x = "PaCMAP1", y = "PaCMAP2", 
                                          num_bins_x = num_bins_x_pacmap_gau, 
                                          num_bins_y = num_bins_y_pacmap_gau, 
                                          x_start = NA, y_start = NA, 
                                          buffer_x = NA, buffer_y = NA, 
                                          hex_size = hex_size_pacmap_gau,
                                          is_rm_lwd_hex = FALSE, 
                                          benchmark_to_rm_lwd_hex = NA, 
                                          col_start_2d = "PaCMAP", 
                                          col_start_highd = "x")

df_bin_centroids_pacmap_gau <- model_object_pacmap_gau$df_bin_centroids
df_bin_pacmap_gau <- model_object_pacmap_gau$df_bin

## Triangulate bin centroids
tr1_object_pacmap_gau <- tri_bin_centroids(df_bin_centroids_pacmap_gau, x = "c_x", y = "c_y")
tr_from_to_df_pacmap_gau <- gen_edges(tri_object = tr1_object_pacmap_gau)

## Compute 2D distances
distance_pacmap_gau <- cal_2d_dist(tr_coord_df = tr_from_to_df_pacmap_gau, 
                                 start_x = "x_from", start_y = "y_from", 
                                 end_x = "x_to", end_y = "y_to", 
                                 select_vars = c("from", "to", "distance"))

# distance_plot <- plot_dist(distance_pacmap_gau) +
#   ylab(expression(d^{(2)})) +
#   theme(axis.text = element_text(size = 5),
#         axis.title = element_text(size = 12))
# 
# distance_plot

## To find the benchmark value
benchmark_pacmap_gau <- find_lg_benchmark(distance_edges = distance_pacmap_gau, distance_col = "distance")

# ggplot() +
# geom_trimesh(data = df_bin_centroids_pacmap_gau, mapping = aes(x = c_x, y = c_y))

trimesh_removed_pacmap_gau <- vis_rmlg_mesh(distance_edges = distance_pacmap_gau, benchmark_value = benchmark_pacmap_gau, tr_coord_df = tr_from_to_df_pacmap_gau, distance_col = "distance")

trimesh_removed_pacmap_gau <- trimesh_removed_pacmap_gau +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'e', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

df_all_pacmap_gau <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID), pacmap_data_with_hb_id_gau)

# show_langevitour(df = df_all_pacmap_gau, df_b = df_bin_pacmap_gau, df_b_with_center_data = df_bin_centroids_pacmap_gau, col_start = "x", distance_df = distance_pacmap_gau, distance_col = "distance", benchmark_value = benchmark_pacmap_gau)

```

```{r}
#| warning: false
#| echo: false
#| label: fig-modelfiveGau
#| fig-cap: "Is there a best model to represent the original data in 2D space or are they all providing equivalent information?, (a) Model generated with 21 and 28 bins along the x and y axes, respectively, using a hexagonal size of 0.03 in the 2D space with tSNE, (b) Model generated with 60 and 79 bins along the x and y axes, respectively, using a hexagonal size of 0.01 in the 2D space with UMAP, (c) Model generated with 60 and 1927 bins along the x and y axes, respectively, using a hexagonal size of 0.01 in the 2D space with PHATE, (d) Model generated with 46 and 61 bins along the x and y axes, respectively, using a hexagonal size of 0.013 in the 2D space with TriMAP, and (e) Model generated with 31 and 40 bins along the x and y axes, respectively, using a hexagonal size of 0.02 in the 2D space with PaCMAP."
#| out-width: 100%

trimesh_removed_tsne_gau + trimesh_removed_umap_gau + trimesh_removed_phate_gau + trimesh_removed_trimap_gau + trimesh_removed_pacmap_gau +
  #plot_annotation(tag_levels = 'a') +
  plot_layout(ncol = 5) &
  theme(legend.position='none', plot.tag = element_text(size = 8))
```

::: {#fig-gau1-sc layout-ncol="3" fig-pos="H"}
![](figures/five_gau_clusters/sc_tsne_1.png){#fig-gau1-sc1}

![](figures/five_gau_clusters/sc_tsne_2.png){#fig-gau1-sc2}

![](figures/five_gau_clusters/sc_tsne_3.png){#fig-gau1-sc3}

Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with tSNE in high-D, a video of the tour animation is available at (<https://youtu.be/oQxEb4wRdHI>).
:::

::: {#fig-gau2-sc layout-ncol="3" fig-pos="H"}
![](figures/five_gau_clusters/sc_umap_1.png){#fig-gau2-sc1}

![](figures/five_gau_clusters/sc_umap_2.png){#fig-gau2-sc2}

![](figures/five_gau_clusters/sc_umap_3.png){#fig-gau2-sc3}

Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with UMAP in high-D, a video of the tour animation is available at (<https://youtu.be/JW49csPpDx4>).
:::

::: {#fig-gau3-sc layout-ncol="3" fig-pos="H"}
![](figures/five_gau_clusters/sc_phate_1.png){#fig-gau3-sc1}

![](figures/five_gau_clusters/sc_phate_2.png){#fig-gau3-sc2}

![](figures/five_gau_clusters/sc_phate_3.png){#fig-gau3-sc3}

Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with PHATE in high-D, a video of the tour animation is available at (<https://youtu.be/xqhRODt8Z-o>).
:::

::: {#fig-gau4-sc layout-ncol="3" fig-pos="H"}
![](figures/five_gau_clusters/sc_trimap_1.png){#fig-gau4-sc1}

![](figures/five_gau_clusters/sc_trimap_2.png){#fig-gau4-sc2}

![](figures/five_gau_clusters/sc_trimap_3.png){#fig-gau4-sc3}

Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with TriMAP in high-D, a video of the tour animation is available at (<https://youtu.be/X-uHSB4VoZ0>).
:::

::: {#fig-gau5-sc layout-ncol="3" fig-pos="H"}
![](figures/five_gau_clusters/sc_pacmap_1.png){#fig-gau5-sc1}

![](figures/five_gau_clusters/sc_pacmap_2.png){#fig-gau5-sc2}

![](figures/five_gau_clusters/sc_pacmap_3.png){#fig-gau5-sc3}

Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with PaCMAP in high-D, a video of the tour animation is available at (<https://youtu.be/UxvU8HQd7JM>).
:::

## Applications {#sec-applications}

### Single-cell RNA-seq data of human

In the field of single-cell studies, a common analytical task involves clustering to identify groups of cells with similar expression profiles. Analysts often turn to NLDR techniques to verify and identify these clusters and explore developmental trajectories. To illustrate the importance of NLDR techniques and parameter selection in identifying clusters, Human Peripheral Blood Mononclear Cells (PBMC3k) dataset [@chen2023] is used. In a study by @chen2023, this dataset was used to demonstrate how UMAP represents clusters (see @fig-umapauthor). As shown in @fig-umapauthor, there are three distinct and well-separated clusters.

<!-- UMAP layout with author's suggested parameter choice-->
```{r}
#| warning: false
#| echo: false

## Import data
training_data_pbmc <- read_rds("data/pbmc3k/pbmc_pca_50.rds")
training_data_pbmc <- training_data_pbmc[, 1:9] |>
  mutate(ID = 1:NROW(training_data_pbmc))

umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_30_min_dist_0.3.rds")
umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc, 
                                                                 x = "UMAP1", y = "UMAP2"))) |>
  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
                  "UMAP2" = "scaled_UMAP2")) |>
  dplyr::mutate(ID = 1:NROW(umap_pbmc))

umap_plot_pbmc <- umap_pbmc_scaled |>
  ggplot(aes(x = UMAP1,
             y = UMAP2))+
  geom_point(alpha=0.5) +
  coord_equal() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
  theme_linedraw() +
  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
        legend.key.height = unit(0.25, 'cm'),
        legend.key.width = unit(0.25, 'cm')) 
```

```{r}
#| echo: false
#| fig-cap: "2D layout from UMAP (n_neighbors = 30, min_dist = 0.3) applied for the PBMC3k dataset. Is this a best representation of the original data?"
#| label: fig-umapauthor
#| fig-pos: H

umap_plot_pbmc 
```

To determine whether the UMAP representation with the parameter choice suggested by @chen2023 preserves the original data structure, we visualize the model constructed with UMAP overlaid on the high-D data (model-in-data space). The figures in @fig-pbmc1-sc show three well-separated clusters, indicating that the suggested UMAP representation preserves the global structure (see @fig-umapauthor). However, as shown in @fig-pbmc1-sc, these clusters are close to each other in high-D space. Also, nonlinear continuity patterns and high-density patches within the clusters are observed (see @fig-pbmc1-sc). Therefore, the suggested UMAP representation (see @fig-umapauthor) does not accurately preserve the local structure of the PBMC3k dataset.

<!-- Fit the model and compute error-->
```{r}
#| warning: false
#| echo: false
#| message: false

## Decide by looking at MSE plot
num_bins_x_pbmc <- 21
num_bins_y_pbmc <- 28
hex_size_pbmc <- 0.03

hb_obj_pbmc <- hex_binning(data = umap_pbmc_scaled, x = "UMAP1", 
                      y = "UMAP2", num_bins_x = num_bins_x_pbmc, 
                      num_bins_y = num_bins_y_pbmc, x_start = NA, y_start = NA, 
                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_pbmc, col_start = "UMAP")

## Data set with all possible centroids in the hexagonal grid
all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_pbmc$centroids))

## Generate all coordinates of hexagons
hex_grid <- as.data.frame(do.call(cbind, hb_obj_pbmc$hex_poly))

## To obtain the standardise counts within hexbins
counts_df <- as.data.frame(do.call(cbind, hb_obj_pbmc$std_cts))
df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
                                             counts_df = counts_df)
# ggplot(data = hex_grid, aes(x = x, y = y)) +
#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
#   coord_fixed()

hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))

# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
#   coord_fixed()

umap_data_with_hb_id_pbmc <- as.data.frame(do.call(cbind, hb_obj_pbmc$data_hb_id))

model_object <- fit_highd_model( training_data = training_data_pbmc, 
                                 nldr_df_with_id = umap_pbmc_scaled, 
                                 x = "UMAP1", y = "UMAP2", 
                                 num_bins_x = num_bins_x_pbmc, 
                                 num_bins_y = num_bins_y_pbmc, 
                                 x_start = NA, y_start = NA, 
                                 buffer_x = NA, buffer_y = NA, 
                                 hex_size = hex_size_pbmc,
                                 is_rm_lwd_hex = FALSE, 
                                 benchmark_to_rm_lwd_hex = NA, 
                                 col_start_2d = "UMAP", 
                                 col_start_highd = "PC_")

df_bin_centroids_pbmc <- model_object$df_bin_centroids
df_bin_pbmc <- model_object$df_bin

## Triangulate bin centroids
tr1_object_pbmc <- tri_bin_centroids(df_bin_centroids_pbmc, x = "c_x", y = "c_y")
tr_from_to_df_pbmc <- gen_edges(tri_object = tr1_object_pbmc)

## Compute 2D distances
distance_pbmc <- cal_2d_dist(tr_coord_df = tr_from_to_df_pbmc, 
                             start_x = "x_from", start_y = "y_from", 
                             end_x = "x_to", end_y = "y_to", 
                             select_vars = c("from", "to", "distance"))

# distance_plot <- plot_dist(distance_pbmc) +
#   ylab(expression(d^{(2)})) +
#   theme(axis.text = element_text(size = 5),
#         axis.title = element_text(size = 12))
# 
# distance_plot

## To find the benchmark value
benchmark_pbmc <- find_lg_benchmark(distance_edges = distance_pbmc, distance_col = "distance")

# ggplot() +
# geom_trimesh(data = df_bin_centroids_pbmc, mapping = aes(x = c_x, y = c_y))

trimesh_removed_pbmc_umap <- vis_rmlg_mesh(distance_edges = distance_pbmc, benchmark_value = benchmark_pbmc, tr_coord_df = tr_from_to_df_pbmc, distance_col = "distance")

trimesh_removed_pbmc_umap <- trimesh_removed_pbmc_umap +
  geom_point(data = umap_pbmc_scaled, aes(x = UMAP1, y = UMAP2), alpha = 0.3) +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

# df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), umap_data_with_hb_id_pbmc)
# 
# show_langevitour(df_all_pbmc, df_bin_pbmc, df_bin_centroids_pbmc, benchmark_pbmc, distance_pbmc, "distance", col_start = "PC_")

df_bin_train <- df_bin_pbmc
names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])

df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), umap_data_with_hb_id_pbmc)

error_df <- df_all_pbmc |>
  dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates

for (i in 1:(NCOL(df_bin_train) - 1)) {

  error_df[ , paste0("abs_residual_", "PC_", i)] <- abs(error_df[ , paste0("PC_", i)] - error_df[ , paste0("avg_", "PC_", i)])

}

error_df <- error_df |>
  dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "PC_")))))

error_df <- error_df |>
  mutate(type = if_else(total == 0, "no error",
                        if_else(total <= 5, "error 0-5",
                                if_else(total <= 10, "error 5-10",
                                        if_else(total <= 15, "error 10-15",
                                                if_else(total <= 20, "error 15-20",
                                                        if_else(total <= 25, "error 20-25",
                                                                "error greter than 25")))))))

plot_list2_pbmc_error <- error_df |>
  mutate(type = factor(type , levels = c("no error", "error 0-5", "error 5-10", "error 10-15", "error 15-20", "error 20-25", "error greter than 25"))) |>
    ggplot(aes(x = UMAP1,
               y = UMAP2, color = type))+
    geom_point(alpha=0.5, size = 0.1) +
    coord_equal() +
    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
  theme_linedraw() +
    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
         legend.key.height = unit(0.25, 'cm'),
         legend.key.width = unit(0.25, 'cm')) +
  scale_color_manual(values=rainbow_hcl(7)) +
  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
```

```{r}
#| echo: false
#| fig-cap: "(a) Model generated with 21 and 28 bins along x and y axes with hexagonal size 0.03 in the 2D space overlaid on UMAP data, and (b) high-D model error in model space."
#| label: fig-nldervisPBMCUMAP
#| fig-pos: H

trimesh_removed_pbmc_umap + plot_list2_pbmc_error +
plot_layout(guides='collect', ncol=2) &
  theme(legend.position='bottom') 
```

::: {#fig-pbmc1-sc layout-ncol="3" fig-pos="H"}
![](figures/pbmc3k/sc_1.png){#fig-pbmc1-sc1}

![](figures/pbmc3k/sc_2.png){#fig-pbmc1-sc2}

![](figures/pbmc3k/sc_3.png){#fig-pbmc1-sc3}

Screen shots of the **langevitour** of the PBMC3k data set, shows the model in high-D, a video of the tour animation is available at (<https://youtu.be/Gnhh4hfsbbg>).
:::

To find the best NLDR representation for PBMC3k dataset, we observe the absolute error for different number of non-empty bins and select the one with the lowest error. As shown in @fig-abserrorPBMC, tSNE with a perplexity set to $51$, has the lowest error. Therefore, tSNE with a perplexity of $51$ is used as the best representation for PBMC3k dataset.

<!--compute absolute error for different parameter choices-->

```{r}
#| warning: false
#| echo: false

error_df <- read_csv("data/pbmc3k/error_df.csv", col_names = FALSE)
names(error_df) <- c("num_bins", "error", "mse", "num_bins_x", "num_bins_y", "hex_size", "num_non_empty_bins", "n_neighbors", "min_dist")

error_df <- error_df |>
  dplyr::mutate(type = paste0("n_neighbors: ", n_neighbors, ", min_dist: ", min_dist))

error_df_tsne <- read_csv("data/pbmc3k/error_df_tsne.csv", col_names = FALSE)
names(error_df_tsne) <- c("num_bins", "error", "mse", "num_bins_x", "num_bins_y", "hex_size", "num_non_empty_bins", "perplexity")
error_df_tsne <- error_df_tsne |>
  dplyr::mutate(type = paste0("perplexity: ", perplexity))

error_df <- bind_rows(error_df[, c(1:7, 10)], error_df_tsne[, c(1:7, 9)])

abs_error_plot_pbmc <- ggplot(error_df, aes(x = num_non_empty_bins,
                                       y = log(error), group = type,
colour = type)) +
  geom_point() +
  geom_line() +
   geom_vline(xintercept = 439, linetype="solid",
                color = "black", size=0.8, alpha = 0.5) +
  theme_light() +
  theme(legend.position = "bottom", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7),
        legend.text = element_text(size = 5)) +
  scale_color_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99','#e31a1c','#fdbf6f','#ff7f00','#cab2d6','#6a3d9a', '#b15928', '#ffff99')) +
  ylab("log(absolute error)") +
  xlab("number of non-empty bins")

```

```{r}
#| echo: false
#| fig-cap: "Absolute error from UMAP and tSNE applied to training PBMC3k dataset with diffferent parameter choices. What is the best parameter choice to create the model? The residual plot have a steep slope at the beginning, indicating that a smaller number of non-empty bins causes a larger amount of error. Then, the slope gradually declines or level off, indicating that a higher number of non-empty bins generates a smaller error. Using the elbow method, it was observed that when the number of non-empty bins is set to 439, the lowest error occurred with the parameters perplexity: 51."
#| label: fig-abserrorPBMC
#| fig-pos: H
#| out-width: 100%
#| out-height: 100%

abs_error_plot_pbmc
```

As shown in @fig-tsnesuggest, there are three well-separated clusters, although they are located close to each other. Additionally, nonlinear structures can be observed within the clusters (see @fig-tsnesuggest). Hence, the data structure that UMAP failed to capture for the PBMC3k dataset is successfully captured by tSNE.

```{r}
#| warning: false
#| echo: false

## Import data
training_data_pbmc <- read_rds("data/pbmc3k/pbmc_pca_50.rds")
training_data_pbmc <- training_data_pbmc[, 1:9] |>
  mutate(ID = 1:NROW(training_data_pbmc))

tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_51.rds")
tsne_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tsne_pbmc, 
                                                                 x = "tSNE1", y = "tSNE2"))) |>
  dplyr::rename(c("tSNE1" = "scaled_tSNE1", 
                  "tSNE2" = "scaled_tSNE2")) |>
  dplyr::mutate(ID = 1:NROW(tsne_pbmc))

tsne_plot_pbmc <- tsne_pbmc_scaled |>
  ggplot(aes(x = tSNE1,
             y = tSNE2))+
  geom_point(alpha=0.5) +
  coord_equal() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
  theme_linedraw() +
  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
        legend.key.height = unit(0.25, 'cm'),
        legend.key.width = unit(0.25, 'cm')) 
```

```{r}
#| echo: false
#| fig-cap: "2D layout from tSNE (perplexity = 51) applied for the PBMC3k dataset. Is this a best representation of the original data?"
#| label: fig-tsnesuggest
#| fig-pos: H

tsne_plot_pbmc 
```

Next, the model is fitted for tSNE, and the resulting model is visualized in the data space. The model shows some quirks, as shown in @fig-pbmc2-sc. The large cluster is divided into three subclusters, which are in close to each other.

<!--Fit the model-->
```{r}
#| warning: false
#| echo: false
#| message: false

## Decide by looking at MSE plot
num_bins_x_pbmc <- 31
num_bins_y_pbmc <- 40
hex_size_pbmc <- 0.02

hb_obj_pbmc <- hex_binning(data = tsne_pbmc_scaled, x = "tSNE1", 
                      y = "tSNE2", num_bins_x = num_bins_x_pbmc, 
                      num_bins_y = num_bins_y_pbmc, x_start = NA, y_start = NA, 
                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_pbmc, col_start = "tSNE")

## Data set with all possible centroids in the hexagonal grid
all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_pbmc$centroids))

## Generate all coordinates of hexagons
hex_grid <- as.data.frame(do.call(cbind, hb_obj_pbmc$hex_poly))

## To obtain the standardise counts within hexbins
counts_df <- as.data.frame(do.call(cbind, hb_obj_pbmc$std_cts))
df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
                                             counts_df = counts_df)
# ggplot(data = hex_grid, aes(x = x, y = y)) +
#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
#   coord_fixed()

hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))

# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
#   coord_fixed()

tsne_data_with_hb_id_pbmc <- as.data.frame(do.call(cbind, hb_obj_pbmc$data_hb_id))

model_object <- fit_highd_model( training_data = training_data_pbmc, 
                                 nldr_df_with_id = tsne_pbmc_scaled, 
                                 x = "tSNE1", y = "tSNE2", 
                                 num_bins_x = num_bins_x_pbmc, 
                                 num_bins_y = num_bins_y_pbmc, 
                                 x_start = NA, y_start = NA, 
                                 buffer_x = NA, buffer_y = NA, 
                                 hex_size = hex_size_pbmc,
                                 is_rm_lwd_hex = FALSE, 
                                 benchmark_to_rm_lwd_hex = NA, 
                                 col_start_2d = "tSNE", 
                                 col_start_highd = "PC_")

df_bin_centroids_pbmc <- model_object$df_bin_centroids
df_bin_pbmc <- model_object$df_bin

benchmark_value_rm_lwd <- stats::quantile(df_bin_centroids_pbmc$std_counts,
                probs = c(0,0.25,0.5,0.75,1), names = FALSE)[2]

df_bin_centroids_low <- df_bin_centroids_pbmc |>
  dplyr::filter(std_counts <= benchmark_value_rm_lwd)

identify_rm_bins <- find_low_dens_hex(df_bin_centroids_all = df_bin_centroids_pbmc,
                                      num_bins_x = num_bins_x_pbmc,
                                      df_bin_centroids_low = df_bin_centroids_low)

df_bin_centroids_pbmc <- df_bin_centroids_pbmc |>
  dplyr::filter(!(hexID %in% identify_rm_bins))

## Triangulate bin centroids
tr1_object_pbmc <- tri_bin_centroids(df_bin_centroids_pbmc, x = "c_x", y = "c_y")
tr_from_to_df_pbmc <- gen_edges(tri_object = tr1_object_pbmc)

## Compute 2D distances
distance_pbmc <- cal_2d_dist(tr_coord_df = tr_from_to_df_pbmc, 
                             start_x = "x_from", start_y = "y_from", 
                             end_x = "x_to", end_y = "y_to", 
                             select_vars = c("from", "to", "distance"))

# distance_plot <- plot_dist(distance_pbmc) +
#   ylab(expression(d^{(2)})) +
#   theme(axis.text = element_text(size = 5),
#         axis.title = element_text(size = 12))
# 
# distance_plot

## To find the benchmark value
benchmark_pbmc <- find_lg_benchmark(distance_edges = distance_pbmc, distance_col = "distance")
benchmark_pbmc <- 0.04

# ggplot() +
# geom_trimesh(data = df_bin_centroids_pbmc, mapping = aes(x = c_x, y = c_y))

trimesh_removed_pbmc_tsne <- vis_rmlg_mesh(distance_edges = distance_pbmc, benchmark_value = benchmark_pbmc, tr_coord_df = tr_from_to_df_pbmc, distance_col = "distance")

trimesh_removed_pbmc_tsne <- trimesh_removed_pbmc_tsne +
  geom_point(data = tsne_pbmc_scaled, aes(x = tSNE1, y = tSNE2), alpha = 0.08) +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

# df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), tsne_data_with_hb_id_pbmc)
# 
# show_langevitour(df_all_pbmc, df_bin_pbmc, df_bin_centroids_pbmc, benchmark_pbmc, distance_pbmc, "distance", col_start = "PC_")

df_bin_train <- df_bin_pbmc
names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])

df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), tsne_data_with_hb_id_pbmc)

error_df <- df_all_pbmc |>
  dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates

for (i in 1:(NCOL(df_bin_train) - 1)) {

  error_df[ , paste0("abs_residual_", "PC_", i)] <- abs(error_df[ , paste0("PC_", i)] - error_df[ , paste0("avg_", "PC_", i)])

}

error_df <- error_df |>
  dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "PC_")))))

error_df <- error_df |>
  mutate(type = if_else(total == 0, "no error",
                        if_else(total <= 5, "error 0-5",
                                if_else(total <= 10, "error 5-10",
                                        if_else(total <= 15, "error 10-15",
                                                if_else(total <= 20, "error 15-20",
                                                        "error greter than 20"))))))

plot_list2_pbmc_error <- error_df |>
  mutate(type = factor(type , levels = c("no error", "error 0-5", "error 5-10", "error 10-15", "error 15-20", "error greter than 20"))) |>
    ggplot(aes(x = tSNE1,
               y = tSNE2, color = type))+
    geom_point(alpha=0.5, size = 0.1) +
    coord_equal() +
    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
  theme_linedraw() +
    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
         legend.key.height = unit(0.25, 'cm'),
         legend.key.width = unit(0.25, 'cm')) +
  scale_color_manual(values=rainbow_hcl(6)) +
  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
```

```{r}
#| echo: false
#| fig-cap: "(a) Model generated with 31 and 40 bins along x and y axes with hexagonal size 0.02 in the 2D space overlaid on tSNE data, and (b) high-D model error in model space."
#| label: fig-nldervisPBMCtSNE
#| fig-pos: H

trimesh_removed_pbmc_tsne + plot_list2_pbmc_error +
plot_layout(guides='collect', ncol=2) &
  theme(legend.position='bottom') 
```

::: {#fig-pbmc2-sc layout-ncol="3" fig-pos="H"}
![](figures/pbmc3k/sc_4.png){#fig-pbmc2-sc4}

![](figures/pbmc3k/sc_5.png){#fig-pbmc2-sc5}

![](figures/pbmc3k/sc_6.png){#fig-pbmc2-sc6}

Screen shots of the **langevitour** of the PBMC3k data set, shows the model in high-D, a video of the tour animation is available at (<https://youtu.be/R0EroJcGbas>).
:::

### Handwritten digits

The MNIST dataset consists of grayscale images of handwritten digits [@lecun2010]. @Yingfan2021 used this dataset to demonstrate how the PaCMAP preserves local structure. We selected the 2D embedding of PaCMAP for the handwritten digit 1. As shown in @fig-pacmapauthor, we observe a nonlinear structure present in 2D space. Furthermore, the angle of digit 1 images varies along the nonlinear structure (see @fig-sampleimg). To answer the following questions,

- Is the original data structure of the handwritten digit 1 dataset nonlinear?
- Where does PaCMAP preserve the original data structure?
- Where does PaCMAP show quirks?

we visualize the model constructed with PaCMAP overlaid on data in high-D space (model-in-data space).

According to @fig-mnist1-sc1, the nonlinear continuity structure observed in the 2D representation of PaCMAP is also visible when visualizing the model overlaid on the data space. This provides evidence that the model accurately captures the structure of the high-D data. Also, the model has a twisted pattern within the nonlinear structure, which is an addition pattern of data that cannot be seen in the 2D representation (see @fig-mnist1-sc2).

In order to visually assess the preservation of local structure, it's important to consider the edges in the model. In a model that effectively preserves local structure, there shouldn't be any long edges in high-D space. However, as shown in @fig-mnist1-sc3, some long edges exist in high-D space that are not identified as long edges in the 2D representation.

<!-- PaCMAP layout with author's suggested parameter choice-->
```{r}
#| warning: false
#| echo: false

## Import data
training_data_mnist <- read_rds("data/mnist/mnist_10_pcs_of_digit_1.rds")
training_data_mnist <- training_data_mnist |>
  mutate(ID = 1:NROW(training_data_mnist))

pacmap_minst <- read_rds("data/mnist/mnist_pacmap.rds")
pacmap_minst_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = pacmap_minst, 
                                    x = "PaCMAP1", y = "PaCMAP2"))) |>
  dplyr::rename(c("PaCMAP1" = "scaled_PaCMAP1", 
                  "PaCMAP2" = "scaled_PaCMAP2")) |>
  dplyr::mutate(ID = 1:NROW(pacmap_minst))

position_df <- pacmap_minst_scaled |> 
  dplyr::filter(ID %in% c(6475, 3311, 6489, 2347, 387, 5239, 728, 517, 6369, 3055, 6345, 1761)) |>
  dplyr::mutate(position = c(5, 9, 7, 8, 4, 11, 2, 6, 12, 10, 1, 3))

pacmap_plot_mnist <- pacmap_minst_scaled |>
    ggplot(aes(x = PaCMAP1,
               y = PaCMAP2))+
    geom_point(alpha=0.1) +
    coord_equal() +
    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + 
  geom_text(data = position_df, aes(x = PaCMAP1, y = PaCMAP2, label = position), colour = "red", size = 5) +
  theme_linedraw() +
    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
         legend.key.height = unit(0.25, 'cm'),
         legend.key.width = unit(0.25, 'cm')) 
```

```{r}
#| echo: false
#| fig-cap: "(a) 2D layout from PaCMAP (n_components=2, n_neighbors=10, init=random, MN_ratio=0.9, FP_ratio=2.0) applied for the digit 1 of the MNIST dataset. Is this a best representation of the original data?"
#| label: fig-pacmapauthor
#| fig-pos: H

pacmap_plot_mnist 
```

```{r}
#| warning: false
#| echo: false

mnist_data <- read_rds("data/mnist/mnist_digit_1.rds")
```

<!--images with positions-->
```{r}
#| warning: false
#| echo: false

pixels_gathered <-  mnist_data |>
  mutate(instance = row_number()) |>
  gather(pixel, value, -Label, -instance) |>
  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
  filter(instance %in% c(3710, 2391, 2385, 5030, 6475,
                         7679, 3568, 1312, 3311, 1097,
                         3552, 7853, 6489, 7689, 6690,
                        1380, 6057, 2347, 5946, 3355,
                        4175, 3997, 5378, 387, 1854,
                        614, 3079, 1762, 5239, 3723,
                        5748, 728, 7419, 7794, 6233,
                        33, 2485, 5998, 318, 1761,
                        5690, 165, 517, 6935, 2682,
                        4962, 2264, 5563, 6369, 559,
                        5188, 4849, 2666, 3448, 3055,
                        3120, 6869, 6345, 4470, 7147)) |>
  mutate(instance = as.character(instance))

theme_set(theme_light())
img_sample <- pixels_gathered |> 
  ggplot(aes(x, y, fill = value)) +  
  geom_tile() +  
  facet_wrap(~ factor(instance, levels = c(3710, 2391, 2385, 5030, 6475,
                         7679, 3568, 1312, 3311, 1097,
                         3552, 7853, 6489, 7689, 6690,
                        1380, 6057, 2347, 5946, 3355,
                        4175, 3997, 5378, 387, 1854,
                        614, 3079, 1762, 5239, 3723,
                        5748, 728, 7419, 7794, 6233,
                        33, 2485, 5998, 318, 1761,
                        5690, 165, 517, 6935, 2682,
                        4962, 2264, 5563, 6369, 559,
                        5188, 4849, 2666, 3448, 3055,
                        3120, 6869, 6345, 4470, 7147)), ncol = 10) +
  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
         legend.key.height = unit(0.25, 'cm'),
         legend.key.width = unit(0.25, 'cm'),
        strip.text = element_blank(),
        aspect.ratio = 1)
```

```{r}
#| echo: false
#| fig-cap: "Images of the handwritten digit 1 are distributed along the right-to-left of first embedding axis of the PaCMAP nonlinear structure."
#| label: fig-sampleimg
#| out-width: 100%
#| fig-pos: H

img_sample
```

<!-- Fit the model and compute error-->
```{r}
#| warning: false
#| echo: false
#| message: false

## Decide by looking at MSE plot
num_bins_x_mnist <- 12
num_bins_y_mnist <- 15
hex_size_mnist <- 0.06

hb_obj_mnist <- hex_binning(data = pacmap_minst_scaled, x = "PaCMAP1", 
                      y = "PaCMAP2", num_bins_x = num_bins_x_mnist, 
                      num_bins_y = num_bins_y_mnist, x_start = NA, y_start = NA, 
                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_mnist, col_start = "PaCMAP")

## Data set with all possible centroids in the hexagonal grid
all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_mnist$centroids))

## Generate all coordinates of hexagons
hex_grid <- as.data.frame(do.call(cbind, hb_obj_mnist$hex_poly))

## To obtain the standardise counts within hexbins
counts_df <- as.data.frame(do.call(cbind, hb_obj_mnist$std_cts))
df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
                                             counts_df = counts_df)
# ggplot(data = hex_grid, aes(x = x, y = y)) +
#   geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
#   geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "red") +
#   coord_fixed()

hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))

# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
#   coord_fixed()

pacmap_data_with_hb_id_mnist <- as.data.frame(do.call(cbind, hb_obj_mnist$data_hb_id))

model_object <- fit_highd_model( training_data = training_data_mnist, 
                                 nldr_df_with_id = pacmap_minst_scaled, 
                                 x = "PaCMAP1", y = "PaCMAP2", 
                                 num_bins_x = num_bins_x_mnist, 
                                 num_bins_y = num_bins_y_mnist, 
                                 x_start = NA, y_start = NA, 
                                 buffer_x = NA, buffer_y = NA, 
                                 hex_size = hex_size_mnist,
                                 is_rm_lwd_hex = FALSE, 
                                 benchmark_to_rm_lwd_hex = NA, 
                                 col_start_2d = "PaCMAP", 
                                 col_start_highd = "PC")

df_bin_centroids_mnist <- model_object$df_bin_centroids
df_bin_mnist <- model_object$df_bin

## Triangulate bin centroids
tr1_object_mnist <- tri_bin_centroids(df_bin_centroids_mnist, x = "c_x", y = "c_y")
tr_from_to_df_mnist <- gen_edges(tri_object = tr1_object_mnist)

## Compute 2D distances
distance_mnist <- cal_2d_dist(tr_coord_df = tr_from_to_df_mnist, 
                             start_x = "x_from", start_y = "y_from", 
                             end_x = "x_to", end_y = "y_to", 
                             select_vars = c("from", "to", "distance"))

# distance_plot <- plot_dist(distance_mnist) +
#   ylab(expression(d^{(2)})) +
#   theme(axis.text = element_text(size = 5),
#         axis.title = element_text(size = 12))
# 
# distance_plot

## To find the benchmark value
benchmark_mnist <- find_lg_benchmark(distance_edges = distance_mnist, distance_col = "distance")

# ggplot() +
# geom_trimesh(data = df_bin_centroids_mnist, mapping = aes(x = c_x, y = c_y))

trimesh_removed_mnist_pacmap <- vis_rmlg_mesh(distance_edges = distance_mnist, benchmark_value = benchmark_mnist, tr_coord_df = tr_from_to_df_mnist, distance_col = "distance")

trimesh_removed_mnist_pacmap <- trimesh_removed_mnist_pacmap +
  geom_point(data = pacmap_minst_scaled, aes(x = PaCMAP1, y = PaCMAP2), alpha = 0.06) + 
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'a', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)

df_bin_train <- df_bin_mnist
names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])

df_all_mnist <- dplyr::bind_cols(training_data_mnist, pacmap_data_with_hb_id_mnist)

error_df <- df_all_mnist |>
  dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates

# prediction_df_join <- prediction_df_join |>
#   dplyr::left_join(data, by = c("ID" = "ID")) ## Map high-D data

for (i in 1:(NCOL(df_bin_train) - 1)) {

  error_df[ , paste0("abs_residual_", "PC", i)] <- abs(error_df[ , paste0("PC", i)] - error_df[ , paste0("avg_", "PC", i)])

}

error_df <- error_df |>
  dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "PC")))))

# library(ggbeeswarm)
# error_df$group <- "1"
# ggplot(error_df, aes(x = group, y = total)) +
#   geom_quasirandom()+
#   ylim(0, max(unlist(error_df$total))+ 0.5) + coord_flip()

### The minimum error is 0 and the maximum is 42.17439
### There is lot of points with error 0,

error_df <- error_df |>
  mutate(type = if_else(total <= 2, "error 0-2",
                        if_else(total <= 4, "error 2-4",
                                if_else(total <= 6, "error 4-6",
                                        if_else(total <= 8, "error 6-8",
                                                if_else(total <= 10, "error 8-10",
                                                        if_else(total <= 12, "error 10-12",
                                                                if_else(total <= 14, "error 12-14", "error greter than 14"))))))))

plot_list2_mnist_error <- error_df |>
  mutate(type = factor(type , levels = c("error 0-2", "error 2-4", "error 4-6", "error 6-8", "error 8-10", "error 10-12", "error 12-14", "error greter than 14"))) |>
    ggplot(aes(x = PaCMAP1,
               y = PaCMAP2, color = type, group = ID))+
    geom_point(alpha=0.5, size = 0.1) +
    coord_equal() +
    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
  theme_linedraw() +
    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
         legend.key.height = unit(0.25, 'cm'),
         legend.key.width = unit(0.25, 'cm')) +
  scale_color_manual(values=rainbow_hcl(8)) +
  annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)

# show_langevitour(df_all_mnist, df_bin_mnist, df_bin_centroids_mnist, benchmark_mnist, distance_mnist, "distance", col_start = "PC")

# predict_minst_df <- as.data.frame(do.call(cbind, predict_emb(training_data_mnist, df_bin_centroids_mnist, df_bin_mnist, "PaCMAP")))
# gen_summary(training_data_mnist, predict_minst_df, df_bin_mnist, col_start = "PC")
```

```{r}
#| echo: false
#| fig-cap: "(a) Model generated with 12 and 15 bins along x and y axes with hexagonal size 0.06 in the 2D space overlaid on PaCMAP data, and (b) high-D model error in model space."
#| label: fig-nldervisMNISTPACMAP
#| fig-pos: H

trimesh_removed_mnist_pacmap + plot_list2_mnist_error +
plot_layout(guides='collect', ncol=2) &
  theme(legend.position='bottom')
```

::: {#fig-mnist1-sc layout-ncol="3" fig-pos="H"}
![](figures/mnist/sc_1.png){#fig-mnist1-sc1}

![](figures/mnist/sc_2.png){#fig-mnist1-sc2}

![](figures/mnist/sc_3.png){#fig-mnist1-sc3}

Screen shots of the **langevitour** of the MNIST digit 1 data set, shows the model-in-data space, a video of the tour animation is available at (<https://youtu.be/zq2GM9qvUNA>).
:::

In terms of model error, some data points show high error (see @fig-nldervisMNISTPACMAP) due to their deviation from the typical high-D data structure, acting as outliers. This is evident when comparing the images associated with these high model error points (see @fig-sampleimgerror) to others (see @fig-sampleimg).

<!--images that occur large error-->
```{r}
#| warning: false
#| echo: false

img_error <- error_df |> 
  dplyr::filter(total >= 10) |> 
  dplyr::pull(ID)

mnist_data <- read_rds("data/mnist/mnist_digit_1.rds")
pixels_gathered <-  mnist_data |>
  mutate(instance = row_number()) |>
  gather(pixel, value, -Label, -instance) |>
  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
  filter(instance %in% img_error)

theme_set(theme_light())
imge_error_sample <- pixels_gathered |> 
  ggplot(aes(x, y, fill = value)) +  
  geom_tile() +  
  facet_wrap(~ instance, ncol = 11) +
  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
         legend.key.height = unit(0.25, 'cm'),
         legend.key.width = unit(0.25, 'cm'),
        strip.text = element_blank(),
        aspect.ratio = 1)
```

```{r}
#| echo: false
#| fig-cap: "Images of handwritten digit 1 which occur large model error."
#| label: fig-sampleimgerror
#| out-width: 100%
#| fig-pos: H

imge_error_sample
```

## Conclusion {#sec-conclusions}

In conclusion, our proposed model offers a novel approach for visualizing high-dimensional (high-D) data by leveraging non-linear dimension reduction (NLDR) techniques. Through a series of steps including hexagonating NLDR data, triangulating bin centroids, and lifting the 2D triangular mesh into high dimensions, our model effectively transforms complex high-D data into interpretable 2D representations. The model's performance is evaluated using goodness of fit statistics such as Mean Squared Error (MSE) and Akaike Information Criterion (AIC), providing insights into its accuracy and reliability. Overall, our model presents a valuable tool for researchers and practitioners in various fields to gain deeper insights from high-dimensional datasets.

Our algorithm mainly consists...

These Goodness of fit statistics useful in encountering the prediction accuracy when we predict the 2D embedding for the new high-D data. Because, we try to find the nearest high-D mappings as the first step of prediction which will participate the prediction process. When we need to find which different NLDR technique and which parameter choices is giving the best representation of high-D data, MSE and AIC are also useful.

We introduce a new tool to help to determine which method, which parameter choice provide the most useful representation of high-D data.

<!--Our research introduces a comprehensive framework that leverages tours for interactive exploration of high-dimensional data coupled with a low-dimensional manifold, facilitated by the `quollr` R package. Regardless of the Non-Linear Dimension Reduction (NLDR) technique in use, our approach demonstrates effectiveness through simulation examples, particularly in the iterative removal of long edges for a smoother representation and capturing cluster variance.

In the example with doublets, our method successfully captures the tweak within each cluster, indicating the variance present within them. However, the model may not appear smooth in high-dimensional space due to considerable noise when the data has a piecewise linear geometry, such as the tree simulation.

The practical application of our framework, as showcased with the UMAP view, enables visual inspection of well-separated clusters. Furthermore, the combined tour and model provide a robust assessment of whether UMAP preserves the data structure and accurately transforms the data.

The advantages of our approach include its versatility across various NLDR techniques and the ability to generate interactive visualizations for detailed exploration. The tour provides an intuitive way to navigate and comprehend high-dimensional data while assessing the accuracy of dimensionality reduction.

However, one limitation is that the approach may be less effective in cases with significant noise, as seen in the tree simulation example. Additionally, while our method aids in visual verification, quantifying the accuracy of embeddings might require further evaluation metrics.

In conclusion, our framework presents a powerful tool for researchers and analysts in single-cell studies to assess their embeddings by visually inspecting them alongside the original data. By leveraging the advantages of tours and low-dimensional manifolds, our approach offers valuable insights into the data transformation process, empowering users to make informed decisions in analyzing high-dimensional data. Future work could enhance the method's robustness in the presence of noise and explore additional evaluation metrics for quantifying embedding accuracy.-->

## References {.unnumbered}

::: {#refs}
:::

{{< pagebreak >}}

### 2024-03-27 paper_original.tex
cleaning up the paper

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt]{article}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-1in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.7in}%
\addtolength{\topmargin}{-1in}%

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{agsm}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Visualising How Non-linear Dimension Reduction Warps Your Data},
  pdfauthor={Jayani P.G. Lakshika; Dianne Cook; Paul Harrison; Michael Lydeamore; Thiyanga S. Talagala},
  pdfkeywords={high-dimensional data, dimension
reduction, triangulation, hexagonal binning, low-dimensional
manifold, manifold learning, tour, data vizualization},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\begin{document}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{\bf Visualising How Non-linear Dimension Reduction Warps Your
Data}
\author{
Jayani P.G. Lakshika\\
Econometrics \& Business Statistics, Monash University\\
and\\Dianne Cook\\
Econometrics \& Business Statistics, Monash University\\
and\\Paul Harrison\\
MGBP, BDInstitute, Monash University\\
and\\Michael Lydeamore\\
Econometrics \& Business Statistics, Monash University\\
and\\Thiyanga S. Talagala\\
Statistics, University of Sri Jayewardenepura\\
}
\maketitle

\bigskip
\bigskip
\begin{abstract}
Nonlinear dimension reduction (NLDR) techniques such as tSNE, and UMAP
provide a low-dimensional representation of high-dimensional (high-D)
data using non-linear transformation. The methods and parameter choices
can create wildly different representations, making it difficult to
decide which is best, or whether any or all are accurate or misleading.
NLDR often exaggerates random patterns, sometimes due to the samples
observed. But NLDR views have an important role in data analysis
because, if done well, they provide a concise visual (and conceptual)
summary of high-D distributions. To help evaluate the NLDR we have
developed an algorithm to show the 2D NLDR model in the high-D space,
viewed with a tour. One can see if the model fits everywhere or better
in some subspaces, or completely mismatches the data. It is used to
evaluate which 2D layout is the best representation of the high-D
distribution and see how different methods may have similar summaries or
quirks.
\end{abstract}

\noindent%
{\it Keywords:} high-dimensional data, dimension
reduction, triangulation, hexagonal binning, low-dimensional
manifold, manifold learning, tour, data vizualization
\vfill

\newpage
\spacingset{1} % DON'T change the spacing! (Default 1.9)

\section{Introduction}\label{sec-intro}

High-dimensional (high-D) data is prevalent across various fields, such
as ecology and bioinformatics \citep{Guo2023}, due to advancements in
data collection technologies \citep{Johnstone2009, ayesha2020overview}.
However, visualization of high-D data introduces significant challenges,
because the complexity of visualizing data beyond two dimensions
\citep{Jia2022}. In recent years, interactive and dynamic graphics
systems like \textbf{liminal} \citep{article21} ---which employs
interactive tools like brushing and linking \citep{article58}---and
software tools such as \textbf{XGobi}, \textbf{GGobi} \citep{article60},
\textbf{tourr} \citep{article61}, \textbf{detourr} \citep{article22},
and \textbf{langevitour} \citep{article09}, involving dynamic methods
like tours \citep{Asimov1985}, have played a key role in visualizing
high-D data (data-vis).

To create low-dimensional representations (typically in 2D) (m-vis)
\citep{article59} of high-D data, it is common to apply dimension
reduction (DR) techniques. Approaches for DR involve linear methods such
as principal component analysis (PCA) \citep{Karl1901}, non-linear
methods such as multi-dimensional scaling (MDS) \citep{Torgerson1967}.
In the past decade, many new non-linear dimension reduction (NLDR)
techniques have emerged, such as t-distributed stochastic neighbor
embedding (tSNE) \citep{Laurens2008} and uniform manifold approximation
and projection (UMAP) \citep{Leland2018}. NLDR techniques are the 2D
models of high-D data in our context.

It is important to visualize various non-linear dimensionality reduction
(NLDR) techniques for the same high-D data in order to understand and
find the best representation. After doing so, the 2D models may differ
considerably from each other and may also deviate from the original data
structure in high-dimensional space. Therefore, visualizing the 2D model
in high-D space (m-in-ds) is more useful to answer different types of
questions:

\begin{itemize}
\item
  Is there a best 2D representation of high-D data or are they all
  providing equivalent information? Is there a best parameter choice to
  fit the 2D model? How does the model change when it's parameters
  change?
\item
  How well the does the 2D models capture the data structure? Is the
  model fitting able to capture different data structure like
  non-linear, clustering?
\end{itemize}

If we cannot easily ask and answer these questions, our ability to
understand the models is limited. To find the best 2D model and
parameter choices, a better understanding of the underlying science is
important.

Also, the importance of m-vis along with data-vis has been recognized
and incorporated into interactive software, \textbf{liminal}
\citep{article21}. But the 2D model and high-D visualize side by side
and interactive like brushing and linking connect the data in the two
panels. To address this challenge, we propose a novel approach by
combining the tour technique with a low-dimensional manifold. This
manifold is created through the synergistic use of NLDR techniques,
hexagonal binning, and triangulation. This integration facilitates a
more understanding of the data structure, how well (or how poorly) NLDR
techniques perform.

The outline of this paper is as follows. Section~\ref{sec-background}
provides an detailed overview of dimension reduction methods, and tour
technique. Building upon this foundation, Section
Section~\ref{sec-methods} describes the proposed algorithm, including
its implementation details, model tuning, summaries, and a synthetic
example to demonstrate its functionality. In Section
Section~\ref{sec-applications}, the algorithm's applications on various
datasets, particularly in single-cell RNA-seq and handwritten digit
data, are showcased. These applications demonstrate how some NLDR
techniques effectively capture both local and global structures of the
original data, while others may result in misleading interpretations.
Finally, Section Section~\ref{sec-conclusions} summarizes the findings
and describe the importance of the proposed approach in addressing the
challenges associated with selecting the best NLDR method and parameter
choices to achieve a more accurate 2D representation of high-D data.

\begin{figure}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{paper_original_files/figure-pdf/fig-nldervisUMAP-1.pdf}

}

\caption{\label{fig-nldervisUMAP}2D layouts from UMAP applied for the
S-curve data: (a) UMAP (n\_neighbors = 7), (b) UMAP (n\_neighbors = 15),
(c) UMAP (n\_neighbors = 32), (d) UMAP (n\_neighbors = 50). Is there a
best hyperparameter choice in representing UMAP or are they all
providing equivalent information?}

\end{figure}%

\section{Background}\label{sec-background}

\subsection{Dimension Reduction}\label{dimension-reduction}

Consider the high-D data a rectangular matrix \(X_{n \times p}\), where
\(X_{n \times p} = \begin{bmatrix} \textbf{x}_{1} & \textbf{x}_{2} & \cdots & \textbf{x}_{n}\\ \end{bmatrix}^\top\),
with \(n\) observations in \(p\) dimensions. The objective is to
discover a low-dimensional projection
\(Y_{n \times d} = \begin{bmatrix} \textbf{y}_{1} & \textbf{y}_{2} & \cdots & \textbf{y}_{n}\\ \end{bmatrix}^\top\),
represented as an \(n\) × \(d\) matrix, where \(d \ll p\). The reduction
process seeks to remove noise from the original data set while retaining
essential information.

There are two main categories of dimension reduction techniques: linear
and non-linear methods. Linear techniques involve a linear
transformation of the data, with one popular example being PCA. PCA
performs an eigen-decomposition of the sample covariance matrix to
obtain orthogonal principal components that capture the variance of the
data \citep{Karl1901}.

In contrast, NLDR techniques generate the low-dimensional representation
\(Y\) from the high-dimensional data \(X\), often using pre-processing
techniques like \(k\)-nearest neighbors graph or kernel transformations.
Multidimensional Scaling (MDS) is a class of NLDR methods that aims to
construct an embedding \(Y\) in a low-dimensional space, approximating
the pair-wise distances in \(X\) \citep{Torgerson1967}. Variants of MDS
include non-metric scaling \citep{article62} and Isomap, which estimate
geodesic distances to create the low-dimensional representation
\citep{article63}. Other approaches based on diffusion processes, like
diffusion maps \citep{article64} and the PHATE (Potential of
Heat-diffusion for Affinity-based Trajectory Embedding) algorithm
\citep{article03}, also fall under NLDR methods.

\subsubsection{Non-linear dimension reduction
techniques}\label{non-linear-dimension-reduction-techniques}

NLDR techniques are crucial for analyzing and displaying
high-dimensional data, where linear approaches may not adequately
capture complexities in relationships between variables
\citep{Johnstone2009}. One of the challenges with NLDR techniques is the
selection and tuning of appropriate hyperparameters \citep{liao2023}.
This process involves finding the suitable combination of
hyperparameters that enhances the performance of the NLDR technique,
considering the characteristics of the dataset and the specific goals of
the analysis.

Additionally, another challenge is lack of reverse mapping. Techniques
like PCA and auto-encoders \citep{article65} provide a way to map back
from the low-dimensional space to the high-D space, facilitating data
reconstruction. However, some NLDR methods, such as tSNE, don't have a
specific way to reconstruct the original data from the low-dimensional
space.

In this article, mainly focus on five NLDR techniques. They are tSNE,
UMAP, PHATE, TriMAP \citep{article02}, and Pairwise Controlled Manifold
Approximation (PaCMAP) \citep{Yingfan2021}.

Among these, tSNE \citep{Laurens2008} stands out for its ability to
preserve pairwise distances. By minimizing the divergence between
probability distributions in both high and low-dimensional spaces, tSNE
effectively uncovers intricate structures and patterns within the data.
Its application is widespread, particularly in tasks requiring the
visualization of clusters and local relationships. However, achieving
effective results requires careful consideration of hyperparameters,
such as perplexity.

UMAP \citep{Leland2018} is a useful technique for simplifying data while
maintaining both local and overall structures. It builds a fuzzy
topological view by considering nearby data points and then optimizes a
simplified version to match that view. UMAP is known for working well
with different scales of relationships in data and is efficient in
handling large datasets. However, it's important to choose parameters
like neighbors and minimum distance carefully, as they can affect the
results.

Furthermore, PHATE \citep{article03} is great for understanding how
things develop, especially in single-cell genomics. It uses a heat
diffusion process to capture relationships between data points, like
points along a trajectory. While PHATE is excellent for revealing these
developmental structures, it requires careful tuning of its parameters
because of its specialized focus.

Additionally, TriMAP \citep{article02} takes a special approach by
creating a triangulated graph representation of the data. This method is
good at understanding both local and global structures by treating the
data as a network of triangles. TriMAP is powerful in capturing
complicated structures, but it's important to choose parameters
carefully, like deciding how many neighbors to consider.

PaCMAP \citep{Yingfan2021} is different because it adds supervised
learning to make a 2D representation while keeping the relationships
between pairs of points. It builds a graph using distances between pairs
and then makes the 2D representation better using a customizable loss
function. What's special about PaCMAP is that it can use class labels or
extra information to guide how it makes the 2D representation. This
gives users a way to change how PaCMAP works to fit their needs better.

\begin{figure}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{paper_original_files/figure-pdf/fig-nldervis-1.pdf}

}

\caption{\label{fig-nldervis}2D layouts from different NLDR techniques
applied the same data: (a) tSNE (perplexity = 27), (b) UMAP
(n\_neighbors = 50), (c) PHATE (knn = 5), (d) TriMAP (n\_inliers = 5,
n\_outliers = 4, n\_random = 3), and (e) PaCMAP (n\_neighbors = 10, init
= random, MN\_ratio = 0.9, FP\_ratio = 2). Is there a best
representation of the original data or are they all providing equivalent
information?}

\end{figure}%

\subsection{Linear overviews using
tours}\label{linear-overviews-using-tours}

A tour is a powerful visualization technique used to explore the shape
and global structure of high-dimensional data by generating a sequence
of projections, typically into two dimensions. There are two main types
of tours: the grand tour \citep{Asimov1985} and the guided tour
\citep{article29}. A grand tour involves randomly selecting new
orthonormal bases, enabling users to understand the structure by
exploring the subspace of d-dimensional projections \citep{Asimov1985}.
In contrast, a guided tour can be employed to generate a sequence of
`interesting' projections based on an index function \citep{article29}.

The process begins with the data matrix \(X\). It generates a sequence
of \(p\) × \(d\) orthonormal projection matrices (bases) \(P_t\),
usually \(d\) is one or two dimensions. For each pair of orthonormal
bases \(P_t\) and \(P_{t+1}\), a geodesic path is interpolated to create
smooth animation between projections. The resulting tour continuously
visualizes the projected data \(Y_t\) = \(XP_t\) as it interpolates
between successive bases.

Furthermore, software like \textbf{langevitour} can visualize both types
of tours, providing flexibility for exploring high-dimensional data with
various objectives. In our context, use grand tour along with the model
to observe how effectively the model captures the underlying structure
of the data.

\section{Methodology}\label{sec-methods}

In this paper, we introduce a novel method to determine the most
effective NLDR technique and the best hyperparameter choice that
provides the most useful representation of high-D data. Our approach
involves dividing the high-D dataset into two parts: a training set for
constructing the model and a test set for generating predictive values
and residuals. Our algorithm takes a 2D embedding data as the input and
generate a tour that displays the high-D wireframe to overlay the data.
The flow chart of the proposed algorithm is shown in
Figure~\ref{fig-meth}. The algorithm consists of two main phases: (1)
generating the model in the 2D space and (2) lifting the model into
high-D space. The main steps of the algorithm are described in detail in
this section using UMAP 2D embedding of the S-curve dataset. This
dataset has seven dimensions, including four noise dimensions that were
added to the original 3D data.

\begin{figure}

\centering{

\includegraphics[width=1\textwidth,height=1\textheight]{figures/workflow.png}

}

\caption{\label{fig-meth}The flow diagram shows the main steps of our
algorithm. There are two basic phases, one to generate the model in the
2D space, and other to map the model into the high-D space.}

\end{figure}%

\subsection{Preprocessing steps}\label{preprocessing-steps}

To reduce computational complexity when applying NLDR techniques to
high-D data and to reduce noise presence, PCA \citep[\citet{article68},
\citet{article69}]{article67} is used as a preprocessing step. PCA
involves identifying principal components that maximize variance. These
components are then used as the high-D data for the algorithm.

\subsection{Constructing the 2D model}\label{sec-construct2d}

\textbf{Step 1: Scaling NLDR data}

First, we prepare the 2D embedding data to fit within the bounds
required for regular hexagonal binning. To achieve this, we implement
two key scaling steps. Scale the first 2D embedding component to range
between \(0\) and \(1\), ensuring that all data points fall within this
normalized interval. Secondly, we scale the second 2D embedding
component to range between \(0\) and \(y_{max}\) (see
Equation~\ref{eq-equation3}).

The calculation of \(y_{max}\) involves several steps. First, the aspect
ratio (\(ar\)) is computed by dividing the range of the second 2D
embedding component (\(r_2\)) by the range of the first 2D embedding
component (\(r_1\)) (see Equation~\ref{eq-equation1}). Then, the hexagon
ratio (\(hr\)) is determined by dividing the height of the hexagon
(\(hb\)) by its width (\(wb\)) (see Equation~\ref{eq-equation2}).
Finally, \(y_{max}\) is derived by taking the ceiling of
\(\frac{ar}{hr}\) and multiplying it by \(hr\). This process ensures
that \(y_{max}\) is an integer multiple of \(hr\), accommodating the
grid layout of the hexagonal bins.

\begin{equation}\phantomsection\label{eq-equation1}{
 ar = \frac{r_2}{r_1}
}\end{equation}

\begin{equation}\phantomsection\label{eq-equation2}{
 hr = \frac{hb}{wb}
}\end{equation}

\begin{equation}\phantomsection\label{eq-equation3}{
 y_{max} = \left\lceil\frac{ar}{hr}\right\rceil * hr
}\end{equation}

\textbf{Step 2: Hexagonating NLDR data}

Hexagonating NLDR data (see Figure~\ref{fig-meth} Step 2) involves
partitioning the NLDR data into hexagonal bins, a technique commonly
referred to as hexagonal binning \citep[\citet{article66}]{Carr1987}.
This method use a hexagonal grid to create a bivariate histogram that
effectively visualizes the structure of high-D data. Hexagons, one of
only three regular polygons capable of tessellating a plane
\citep{Carr2013}, offer unique advantages due to their symmetry of
nearest neighbors and maximal number of sides for such tessellations
\citep{Dan2023}. This geometric property makes hexagons more efficient
in covering the plane compared to other regular tessellations and
reduces visual bias when displaying data densities \citep{Dan2023}.

In our algorithm, we aim to conduct regular hexagonal binning, involving
the computation of hexagonal grid configurations, the generation of the
grid, and the assignment of the NLDR data points to hexagons.

\textbf{(a) Computation of hexagonal grid configurations}

In the computation of hexagonal bin configurations, there are mainly two
configurations to consider: the number of bins along the x and y axes.
The first step involves in determining the hexagonal size (\(s\)) and
the buffer along the x and y axes (\(buffer_{x}\) and \(buffer_{y}\)).
Here, the hexagonal size (\(s\)) represents the radius of the outer
circle surrounding the hexagon. The buffer is important in expanding the
boundary to accommodate potential outliers or edge cases.

When computing the number of bins along the x and y axes (\(b_1\) and
\(b_2\)), the process begins by calculating the range of the respective
2D embedding component. After that, the range is adjusted to accommodate
the buffer amount. Then, the spacing between hexagons is determined
based on \(s\). Finally, the number of bins along the axis is computed
by dividing the adjusted range by the spacing.

For the computation of the number of bins along the x-axis (\(b_1\))
(see Equation~\ref{eq-equation12}), the range of the first embedding
component is defined as \(r_1\), while the buffer amount along the
x-axis is denoted as \(buffer_{x}\), and the horizontal spacing is
represented by \(h\) (see Equation~\ref{eq-equation11}). On the other
hand, the number of bins along the y-axis (\(b_2\)) (see
Equation~\ref{eq-equation14}) is computed based on the range of the
second embedding component (\(r_2\)), the buffer amount along the y-axis
(\(buffer_{y}\)), and the vertical spacing (\(v\)) (see
Equation~\ref{eq-equation13}).

\begin{equation}\phantomsection\label{eq-equation11}{
 h = \sqrt{3} * s
}\end{equation}

\begin{equation}\phantomsection\label{eq-equation12}{
 b_1 = \frac{r_1 + buffer_{x}}{h}
}\end{equation}

\begin{equation}\phantomsection\label{eq-equation13}{
  v = 1.5 * s
}\end{equation}

\begin{equation}\phantomsection\label{eq-equation14}{
 b_2 = \frac{r_2 + buffer_{y}}{v}
}\end{equation}

\textbf{(b) Generation of the hexagonal grid}

In this process, the first step involves generating centroids within the
hexagonal grid. This begins by defining the starting coordinates for the
hexagons. With the number of bins computed along the x and y axes, along
with the horizontal and vertical spacing, the centroids are iteratively
computed starting from the hexagonal starting coordinates.

Once the centroids for the hexagonal grid are obtained, the next step is
to compute the hexagonal coordinates. For example, if the centroid of a
hexagonal bin is defined as \((C_x, C_y)\), the hexagonal coordinates
can be computed using \(d_x\) (see Equation~\ref{eq-equation15}) and
\(d_y\) (see Equation~\ref{eq-equation16}) (see
Figure~\ref{fig-hexcoord}).

\begin{equation}\phantomsection\label{eq-equation15}{
d_x = \frac{h}{2}
}\end{equation}

\begin{equation}\phantomsection\label{eq-equation16}{
d_y = \frac{2* v * 1.15}{\sqrt{3}} 
}\end{equation}

\begin{figure}[H]

\centering{

\includegraphics[width=1\textwidth,height=1\textheight]{figures/hex_coord.png}

}

\caption{\label{fig-hexcoord}The hexagonal coordinates computed for the
centroid \((C_x, C_y)\).}

\end{figure}%

\textbf{(c) Assignment of the NLDR data points to hexagons}

After obtaining the centroids of the hexagonal bins for the entire grid,
the process of assigning NLDR points to hexagons involves determining
the nearest hexagonal bin for each NLDR point using the 2D Euclidean
distance. Then, the hexagonal ID of the nearest hexagon is assigned to
the corresponding NLDR points.

\textbf{Step 3: Obtaining bin centroids or bin means}

In the previous step, the algorithm clusters the 2D embedding data into
hexagons. Following this, in this step, the bin centroids or bin means
(see Figure~\ref{fig-meth} Step 3) are obtained \citep{Carr2013}.

The bin centroid (\(C_k^{(2)}\)) for a \(k^{th}\) hexagon with hexagonal
grid coordinates \((h^{k}x_{i}, h^{k}y_{i})\), where \(i = 1 \dots 6\)
can be defined as:

\begin{equation}\phantomsection\label{eq-equation4}{
C_k^{(2)} = (C_{ky_1}, C_{ky_2}) = \left(\frac{\sum_{i=1}^{6} h^{k}x_{i}}{6}, \frac{\sum_{i=1}^{6} h^{k}y_{i}}{6}\right).
}\end{equation}

Also, the bin mean (\(C_k^{(2)}\)) is defined as the mean of the data
points within the \(k^{th}\) hexagon (see Equation~\ref{eq-equation5}).

\begin{equation}\phantomsection\label{eq-equation5}{
C_k^{(2)} = (C_{ky_1}, C_{ky_2}) = \left(\frac{1}{n_k} \sum_{i=1}^{n_k} y_{1i}, \frac{1}{n_k} \sum_{i=1}^{n_k} y_{2i}\right),
}\end{equation}

where \(n_k\) is the number of data points within the hexagon,
\(y_{1i}\) and \(y_{2i}\) are the \(x\) and \(y\) coordinates of the
\(i^{th}\) data point within the hexagon.

\textbf{Step 4: Triangulating bin centroids or bin means}

In this step, the algorithm proceeds to triangulate the hexagonal bin
centroids or bin means (see Figure~\ref{fig-meth} Step 4). Triangulation
is a fundamental process in computational geometry and computer graphics
that involves dividing a set of points in a given space into
interconnected triangles \citep{article30}. One common algorithm used
for triangulation is Delaunay triangulation
\citep[\citet{article54}]{article26}, where points are connected in a
way that maximizes the minimum angles of the resulting triangles,
leading to a more regular and well-conditioned triangulation.

Delaunay triangulation can be defined as follows:

Let \(C^{(2)} = \{C_1^{(2)}, C_2^{(2)}, ..., C_m^{(2)}\}\) be a set of
\(m\) bin centroids or bin means in the plane. Delaunay triangulation of
\(C^{(2)}\), denoted as \(DT(C^{(2)})\), is a triangulation of the
convex hull of \(C^{(2)}\) such that the circumcircle of every triangle
in the triangulation contains no other points from \(C^{(2)}\).

Given that the hexagons are regular, the resulting triangles will mostly
be equilateral.

\subsection{Lifting the model into high
dimensions}\label{lifting-the-model-into-high-dimensions}

\subsubsection{Lifting the triangular mesh points into high
dimensions}\label{lifting-the-triangular-mesh-points-into-high-dimensions}

Consider \(f: \mathbb{R}^p \rightarrow \mathbb{R}^2\) be a function that
maps the high-D data (\(X_{n \times p}\)) to its NLDR equivalent
(\(Y_{n \times d}\)). Then, let
\(g: \mathbb{R}^2 \rightarrow \mathbb{R}^2\) be a function that maps
each 2D embedding point to its closest centroid (\(C^{(2)}\)). It
follows that \(f(g(x))\) maps the high-D points \(x\) to the centroid in
2D (\(C_k^{(2)}\)). Also, define a function
\(v: \mathbb{R}^2 \rightarrow \mathbb{R}^p\) maps the 2D centroid
(\(C^{(2)}\)) to the high-D mean of the points (\(C^{(p)}\)) in the
hexagon.

The high-D mean of all the points in \(k^{th}\) hexagon by

\begin{equation}\phantomsection\label{eq-equation6}{
C_k^{(p)} = (C_{kx_1}, ..., C_{kx_p}) = \left(\frac{1}{n_k} \sum_{i=1}^{n_k} x_{1i}, \frac{1}{n_k} \sum_{i=1}^{n_k} x_{2i}, \dots ,\frac{1}{n_k} \sum_{i=1}^{n_k} x_{pi}\right).
}\end{equation}

Therefore,

\begin{equation}\phantomsection\label{eq-equation7}{
v(C_k^{(2)}) = C_k^{(p)}.
}\end{equation}

Therefore, \(f(g(x))\) gives the 2D centroid associated with high-D
points \(x\), and \(v(C_k^{(2)})\) gives the high-D centroid associated
with 2D point \(C_k^{(2)}\). Thus, \(v(f(g(x)))\) gives the high-D
centroid (\(C_k^{(p)}\)) associated with the 2D embedding of the points
\(x\).

\subsubsection{Lifting the 2D triangular mesh into high
dimensions}\label{lifting-the-2d-triangular-mesh-into-high-dimensions}

As described in Step 4 of Section~\ref{sec-construct2d}, during the
triangulation process in 2D space, vertices are identified to form
edges. With the knowledge of the high-D mappings for the 2D hexagonal
bins, the vertices connected in 2D are also connected in high-D (see
video linked in Figure~\ref{fig-wkhighD}).

\begin{figure}

\centering{

\includegraphics{paper_original_files/figure-pdf/fig-wkhighD-1.pdf}

}

\caption{\label{fig-wkhighD}How the 2D model lift into high dimensions?
(a) visualize the points and the hexagonal bin centroids related
\(7^{th}\) and \(12^{th}\) hexagons, (b) visualization of the edge
connected the \(7^{th}\) and \(12^{th}\) hexagons (colored in red) in
the triangular mesh. A video of tour animation is available at
\url{https://youtu.be/hxU91xNTJL0}.}

\end{figure}%

\subsection{Tuning the model}\label{tuning-the-model}

The performance and robustness of our model depend on four key
parameters: (i) the total number of bins (\(b\)), (ii) a benchmark value
used to remove low-density hexagons, (iii) a benchmark value used to
remove long edges, and (iv) starting point of the hexagonal grid.
However, there is no analytical formula to calculate an appropriate
value for these parameters. The selection of these parameter values
depends on the model performance computed by Mean Squared Error (MSE)
(see Section~\ref{sec-goodfit}).

\subsubsection{Total number of bins}\label{total-number-of-bins}

The number of hexagonal bins in the hexagonal grid has a considerable
impact on the construction of the 2D model, serving as the initial step
in building the 2D model. The chosen total number of bins must
effectively capture the structure of the NLDR data. If the number of
bins is too low, the model may not be able to capture the structure of
the NLDR data effectively (see Figure~\ref{fig-binsize} (a)), while if
there are too many bins, it may result in over-fitting the individual
points of the NLDR data (see Figure~\ref{fig-binsize} (c)). Therefore,
it is important to determine an appropriate number of bins to construct
an effective model.

\begin{figure}[H]

\centering{

\includegraphics{paper_original_files/figure-pdf/fig-binsize-1.pdf}

}

\caption{\label{fig-binsize}Hexbin plots from different number of bins
for the \textbf{UMAP} embeddings of \textbf{S-curve} training data: (a)
b = 12 (3, 4), s = 0.62, (b) b = 84 (6, 14), s = 0.13, and (c) b = 336
(12, 28), s = 0.06. The hexbins are colored based on the density of
points, with darker colors indicating higher point density and yellow
color representing lower point density within each bin. What is the
number of bins that would be effective in representing low-dimensional
data?}

\end{figure}%

\begin{equation}\phantomsection\label{eq-equation8}{
b = b_1 \times b_2
}\end{equation}

Furthermore, the total number of bins is calculated based on the number
of bins along the x-axis and y-axis, as shown in
Equation~\ref{eq-equation8}. To determine the effective total number of
bins, candidate values are selected based on the range between the
minimum and approximate maximum number of bins along the x and y axes.
The minimum number of bins along each axis is set to \(1\), while the
maximum number is estimated by taking the square root of the NLDR data
points. The analysis evaluates the MSE across varying total number of
bins within this range, covering the minimum to maximum values along
both axes (see Figure~\ref{fig-diagnosticpltScurve}).

\begin{figure}

\centering{

\includegraphics{paper_original_files/figure-pdf/fig-diagnosticpltScurve-1.pdf}

}

\caption{\label{fig-diagnosticpltScurve}Goodness of fit statistics from
UMAP applied to training S-curve dataset. What is the effective number
of bins in each NLDR technique to create a 2D model? The MSE plot have a
steep slope at the beginning, indicating that a smaller number of bins
causes a larger amount of error. Then, the slope gradually declines or
level off, indicating that a higher number of bins generates a smaller
error. Using the elbow method, when the total number of bins is set to
\(84\), the slope of the Mean Squared Error (MSE) plot experiences a
sudden and noticeable change, resembling an elbow-like shape. This point
indicates that adding less bins does not enough to capture the data
structure.}

\end{figure}%

\subsubsection{Benchmark value to remove low-density
hexagons}\label{benchmark-value-to-remove-low-density-hexagons}

After establishing the hexagonal grid with an appropriate number of
bins, some hexagonal bins may have few or no data points within them
(see Figure~\ref{fig-rmlowdenshex} (a)). To ensure comprehensive
coverage of the NLDR data, it is necessary to select hexagonal bins with
a considerable number of data points. This involves calculating the
number of points within each hexagon. Then, the standard count is
computed by dividing the number of points within each hexagon by the
maximum number of points in the grid (see Equation~\ref{eq-equationp2}).
Next, bins with a standard count less than a certain benchmark value are
removed (see Figure~\ref{fig-rmlowdenshex} (b)). The following steps
will help in determining a suitable value for removing low-density
hexagons:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Plot the distribution of the standardized counts (see
  Figure~\ref{fig-stdctsScurve}).
\item
  Examine the distribution of counts.
\item
  Select the first quartile value if the distribution is skewed.
\end{enumerate}

\begin{equation}\phantomsection\label{eq-equationp2}{
\text{standard count} = \frac{\text{count}}{\text{max count}} 
}\end{equation}

Furthermore, selecting the benchmark value for removing low-density
hexagons is important. Removing unnecessary bins may lead to the
formation of long edges and an uneven 2D model. Hence, rather than
solely relying on the benchmark value to identify hexagons for removal,
it's essential to consider the standard number of points in the
neighboring hexagons of the identified low-density ones (see
Figure~\ref{fig-rmlowdenshex} (c)). If neighboring bins also show low
counts, only those bins will be removed. The remaining bins are used to
construct the 2D model.

The benchmark value for removing low-density hexagons ranges between
\(0\) and \(1\). When analyzing how these benchmark values influence
model performance, it's essential to observe the change in Mean Squared
Error (MSE) as the benchmark value increases (see
Figure~\ref{fig-diagnosticpltScurvelwd}). The MSE shows a gradual
increase as the benchmark value progresses from \(0\) to \(1\).
Evaluating this rate of increase is important. If the increment is not
considerable, the decision might lean towards retaining low-density
hexagons.

\begin{figure}[H]

\centering{

\includegraphics{paper_original_files/figure-pdf/fig-stdctsScurve-1.pdf}

}

\caption{\label{fig-stdctsScurve}Distribution of standardize counts by
hexagons.}

\end{figure}%

\begin{figure}

\centering{

\includegraphics{paper_original_files/figure-pdf/fig-diagnosticpltScurvelwd-1.pdf}

}

\caption{\label{fig-diagnosticpltScurvelwd}Goodness of fit statistics
from UMAP applied to training S-curve dataset with different benchmark
values to remove the low-density heaxgons. What is the effective
benchmark value to remove the low-density heaxgons? The MSE plot have a
steep slope at the beginning, indicating that a smaller benchmark causes
a small amount of error. Then, the slope gradually increases or level
up, indicating that a higher number of benchmark values generates a
higher error. Using the reverse elbow method, when the benchmark value
is set to 0.2, the slope of the Mean Squared Error (MSE) plot
experiences a sudden and noticeable change, resembling an elbow-like
shape. This point indicates that higher benchmark values remove the
necessary bins as well which lead to distruct the 2D structure.}

\end{figure}%

\begin{figure}[H]

\centering{

\includegraphics{paper_original_files/figure-pdf/fig-rmlowdenshex-1.pdf}

}

\caption{\label{fig-rmlowdenshex}(a) Hexbin plot with colored hexbins
based on point density, (b) Identification of low-density hexagons using
a benchmark value of \(0.2\), and (c) Identification of low-density
hexagons considering neighboring bins.}

\end{figure}%

\subsubsection{Benchmark value for removing long
edges}\label{benchmark-value-for-removing-long-edges}

Creating a smooth 2D representation (see Figure~\ref{fig-modelScurve}
(c)) requires removing edges that connect distant bin centroids within
the triangular mesh (see Figure~\ref{fig-modelScurve} (b)). These edges
only exist in the 2D model and do not extend into higher dimensions,
ensuring that their removal does not impact the model in higher
dimensions. While specific criteria for determining the benchmark value
to remove long edges do not exist, the following steps provide an
approach to identifying a suitable threshold:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Plot the distribution of the 2D Euclidean distances (see
  Figure~\ref{fig-distScurve}).
\item
  Identify the largest difference between consecutive distance values.
\item
  Take the largest distance value corresponding to this difference as
  the benchmark value.
\end{enumerate}

\begin{figure}[H]

\centering{

\includegraphics{paper_original_files/figure-pdf/fig-distScurve-1.pdf}

}

\caption{\label{fig-distScurve}Distribution of 2D Euclidean distances
between bin centroids of the triangular mesh generated with S-curve UMAP
data.}

\end{figure}%

\begin{figure}[H]

\centering{

\includegraphics{paper_original_files/figure-pdf/fig-modelScurve-1.pdf}

}

\caption{\label{fig-modelScurve}(a) Full hexagonal grid with bin
centroids, (b) Model constructed in 2D with long edges, and (c) Model
constructed in 2D after removing the long edges.}

\end{figure}%

\subsubsection{Starting point of the hexagonal
grid}\label{starting-point-of-the-hexagonal-grid}

Hexagonal binning involves tessellating the \(xy\) plane over the set
defined by the range of \(x\) and \(y\) \citep{Dan2023}. Therefore, the
starting point of the hexagonal grid is important parameter to consider.
Having different starting point results in shifting the hexagonal grid
and bin centroids (see Figure~\ref{fig-scurveshifthexgridsexp}). With a
specific total number of bins (which corresponds to a specific range
along the x and y axes), it's essential to explore how different
starting points affect the model's performance (see
Figure~\ref{fig-mseScurvestartpoint}). This is because different
starting points result in varying distributions of non-empty bins (see
Figure~\ref{fig-scurveshifthexgrids}), impacting the model.

\begin{figure}[H]

\centering{

\includegraphics{paper_original_files/figure-pdf/fig-mseScurvestartpoint-1.pdf}

}

\caption{\label{fig-mseScurvestartpoint}Goodness of fit statistics from
UMAP applied to training S-curve dataset with different benchmark values
to remove the low-density heaxgons. What is the effective benchmark
value to remove the low-density heaxgons? The MSE plot have a steep
slope at the beginning, indicating that a smaller benchmark causes a
small amount of error. Then, the slope gradually increases or level up,
indicating that a higher number of benchmark values generates a higher
error. Using the reverse elbow method, when the benchmark value is set
to 0.2, the slope of the Mean Squared Error (MSE) plot experiences a
sudden and noticeable change, resembling an elbow-like shape. This point
indicates that higher benchmark values remove the necessary bins as well
which lead to distruct the 2D structure.}

\end{figure}%

\begin{figure}[H]

\centering{

\includegraphics{paper_original_files/figure-pdf/fig-scurveshifthexgridsexp-1.pdf}

}

\caption{\label{fig-scurveshifthexgridsexp}(a) Visualization of bin
centroids and (b) hexagonal grid, showcasing the configuration before
(colored in gray) and after (colored in light orange) shifting the
starting point by an identical amount applied in both x and y
directions, with a shift value of \(0.537285\).}

\end{figure}%

\begin{figure}[H]

\centering{

\includegraphics{paper_original_files/figure-pdf/fig-scurveshifthexgrids-1.pdf}

}

\caption{\label{fig-scurveshifthexgrids}Hexbin plots shows the
distribution of points before and after shifting the starting point of
the hexagonal grid generated for UMAP applied to the training S-curve
dataset. (a) Hexbin plot before shifting, and (b) Hexbin plot after
shifting (identical shift applied in both x and y directions, with a
shift value of \(0.537285\)).}

\end{figure}%

\subsection{Model summaries}\label{sec-summary}

\subsubsection{Predicted values and
residuals}\label{predicted-values-and-residuals}

The prediction approach involves performing the K-nearest neighbors
(KNN) algorithm for an unsupervised classification problem. First, the
nearest high-D model point is identified for a given new high-D point.
Then, the corresponding 2D centroid mapping for the identified high-D
model point is determined. Finally, the coordinates of this 2D centroid
are used as the predicted 2D embedding for the new high-D data point.
This step is particularly valuable due to the limitations of some NLDR
techniques, like tSNE, which don't provide a straightforward method for
prediction. As a result, our approach offers a solution that capable of
generating predicted 2D embedding regardless of the NLDR technique
employed, effectively addressing this functional gap.

Residuals are essential for evaluating the accuracy of representing
high-D points by the high-D mapping of 2D bin centroids. To measure this
accuracy, an error metric is introduced, quantifying the sum of squared
differences between the high-D data (\(x_{ij}\)) and the high-D mapping
of the 2D bin centroid data (\(C_{x_ij}\)) across all observations and
dimensions (see Equation~\ref{eq-equation10}).

\begin{equation}\phantomsection\label{eq-equation10}{
\text{Error} = \sum_{j = 1}^{n}\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2
}\end{equation}

Here, \(n\) represents the number of observations, \(p\) represents the
dimensions of high-D data, \(x_{ij}\) is the high-D data, and
\(C_{x_ij}\) is the high-D mapping of the 2D bin centroid.

\subsubsection{Goodness of fit statistics}\label{sec-goodfit}

To assess how well our method captures and represents the underlying
structure of the high-D data, Mean Squared Error (MSE) is used. When
computing MSE, total model error (see Section~\ref{sec-summary}) is
divided by the number of observations to make it as a mean value (see
Equation~\ref{eq-equation9}).

\begin{equation}\phantomsection\label{eq-equation9}{
\text{MSE} = \sum_{j = 1}^{n} \frac{\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2}{n}
}\end{equation}

\subsection{Simulated data example}\label{sec-simpleex}

In this section, the effectiveness of the algorithm is described using a
simulated dataset. The dataset consists of five spherical Gaussian
clusters in 4-\(d\), with each cluster containing an equal number of
points and the same within-cluster variation.

In the 2D layouts generated by various NLDR techniques, as shown in
Figure~\ref{fig-nldervis5Gau}, five distinct clusters are observable,
except for PHATE. In tSNE (see Figure~\ref{fig-nldervis5Gau} (a)), these
clusters appear closely. UMAP arranges all clusters in a parallel
manner, with three aligned in one line and the other two in a separate
line (see Figure~\ref{fig-nldervis5Gau} (b)). In contrast, PHATE shows
two closely positioned clusters and three more distant ones (see
Figure~\ref{fig-nldervis5Gau} (c)). PaCMAP shows one central cluster and
the remaining four spread out in different directions (see
Figure~\ref{fig-nldervis5Gau} (e)). Finally, in TriMAP, two clusters are
close, though not as tightly as PHATE, while the other three are
well-separated (see Figure~\ref{fig-nldervis5Gau} (d)).

\begin{figure}[H]

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{paper_original_files/figure-pdf/fig-nldervis5Gau-1.pdf}

}

\caption{\label{fig-nldervis5Gau}2D layouts from different NLDR
techniques applied the same data: (a) tSNE (perplexity = 61), (b) UMAP
(n\_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n\_inliers = 5,
n\_outliers = 4, n\_random = 3), and (e) PaCMAP (n\_neighbors = 10, init
= random, MN\_ratio = 0.9, FP\_ratio = 2). Is there a best
representation of the original data or are they all providing equivalent
information?}

\end{figure}%

To investigate which is the best representation to visualize the
original data or all NLDR methods provide equivalent information, we
visualize the model-in-data space. Models from all NLDR methods show
five well-separated clusters (see Figure~\ref{fig-gau1-sc},
Figure~\ref{fig-gau2-sc}, Figure~\ref{fig-gau3-sc},
Figure~\ref{fig-gau4-sc}, and Figure~\ref{fig-gau5-sc}). This suggests
that for the five Gaussian cluster dataset, all NLDR methods effectively
preserve the global structure. tSNE, UMAP, and PHATE display clusters
with varying densities, indicating their ability to capture
within-cluster variation (see Figure~\ref{fig-gau1-sc},
Figure~\ref{fig-gau2-sc}, and Figure~\ref{fig-gau3-sc}). On the other
hand, both TriMAP and PaCMAP show clusters with flat surfaces,
suggesting a failure to capture within-cluster variation (see
Figure~\ref{fig-gau4-sc} and Figure~\ref{fig-gau5-sc}). Therefore,
TriMAP and PaCMAP do not capture the local structure as effectively as
other methods.

\begin{figure}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{paper_original_files/figure-pdf/fig-modelfiveGau-1.pdf}

}

\caption{\label{fig-modelfiveGau}Is there a best model to represent the
original data in 2D space or are they all providing equivalent
information?, (a) Model generated with 21 and 28 bins along the x and y
axes, respectively, using a hexagonal size of 0.03 in the 2D space with
tSNE, (b) Model generated with 60 and 79 bins along the x and y axes,
respectively, using a hexagonal size of 0.01 in the 2D space with UMAP,
(c) Model generated with 60 and 1927 bins along the x and y axes,
respectively, using a hexagonal size of 0.01 in the 2D space with PHATE,
(d) Model generated with 46 and 61 bins along the x and y axes,
respectively, using a hexagonal size of 0.013 in the 2D space with
TriMAP, and (e) Model generated with 31 and 40 bins along the x and y
axes, respectively, using a hexagonal size of 0.02 in the 2D space with
PaCMAP.}

\end{figure}%

\begin{figure}[H]

\begin{minipage}{0.33\linewidth}

\centering{

\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_tsne_1.png}

}

\subcaption{\label{fig-gau1-sc1}}

\end{minipage}%
%
\begin{minipage}{0.33\linewidth}

\centering{

\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_tsne_2.png}

}

\subcaption{\label{fig-gau1-sc2}}

\end{minipage}%
%
\begin{minipage}{0.33\linewidth}

\centering{

\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_tsne_3.png}

}

\subcaption{\label{fig-gau1-sc3}}

\end{minipage}%

\caption{\label{fig-gau1-sc}Screen shots of the \textbf{langevitour} of
the five Gaussian clusters dataset, shows the model with tSNE in high-D,
a video of the tour animation is available at
(\url{https://youtu.be/oQxEb4wRdHI}).}

\end{figure}%

\begin{figure}[H]

\begin{minipage}{0.33\linewidth}

\centering{

\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_umap_1.png}

}

\subcaption{\label{fig-gau2-sc1}}

\end{minipage}%
%
\begin{minipage}{0.33\linewidth}

\centering{

\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_umap_2.png}

}

\subcaption{\label{fig-gau2-sc2}}

\end{minipage}%
%
\begin{minipage}{0.33\linewidth}

\centering{

\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_umap_3.png}

}

\subcaption{\label{fig-gau2-sc3}}

\end{minipage}%

\caption{\label{fig-gau2-sc}Screen shots of the \textbf{langevitour} of
the five Gaussian clusters dataset, shows the model with UMAP in high-D,
a video of the tour animation is available at
(\url{https://youtu.be/JW49csPpDx4}).}

\end{figure}%

\begin{figure}[H]

\begin{minipage}{0.33\linewidth}

\centering{

\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_phate_1.png}

}

\subcaption{\label{fig-gau3-sc1}}

\end{minipage}%
%
\begin{minipage}{0.33\linewidth}

\centering{

\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_phate_2.png}

}

\subcaption{\label{fig-gau3-sc2}}

\end{minipage}%
%
\begin{minipage}{0.33\linewidth}

\centering{

\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_phate_3.png}

}

\subcaption{\label{fig-gau3-sc3}}

\end{minipage}%

\caption{\label{fig-gau3-sc}Screen shots of the \textbf{langevitour} of
the five Gaussian clusters dataset, shows the model with PHATE in
high-D, a video of the tour animation is available at
(\url{https://youtu.be/xqhRODt8Z-o}).}

\end{figure}%

\begin{figure}[H]

\begin{minipage}{0.33\linewidth}

\centering{

\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_trimap_1.png}

}

\subcaption{\label{fig-gau4-sc1}}

\end{minipage}%
%
\begin{minipage}{0.33\linewidth}

\centering{

\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_trimap_2.png}

}

\subcaption{\label{fig-gau4-sc2}}

\end{minipage}%
%
\begin{minipage}{0.33\linewidth}

\centering{

\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_trimap_3.png}

}

\subcaption{\label{fig-gau4-sc3}}

\end{minipage}%

\caption{\label{fig-gau4-sc}Screen shots of the \textbf{langevitour} of
the five Gaussian clusters dataset, shows the model with TriMAP in
high-D, a video of the tour animation is available at
(\url{https://youtu.be/X-uHSB4VoZ0}).}

\end{figure}%

\begin{figure}[H]

\begin{minipage}{0.33\linewidth}

\centering{

\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_pacmap_1.png}

}

\subcaption{\label{fig-gau5-sc1}}

\end{minipage}%
%
\begin{minipage}{0.33\linewidth}

\centering{

\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_pacmap_2.png}

}

\subcaption{\label{fig-gau5-sc2}}

\end{minipage}%
%
\begin{minipage}{0.33\linewidth}

\centering{

\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_pacmap_3.png}

}

\subcaption{\label{fig-gau5-sc3}}

\end{minipage}%

\caption{\label{fig-gau5-sc}Screen shots of the \textbf{langevitour} of
the five Gaussian clusters dataset, shows the model with PaCMAP in
high-D, a video of the tour animation is available at
(\url{https://youtu.be/UxvU8HQd7JM}).}

\end{figure}%

\section{Applications}\label{sec-applications}

\subsection{Single-cell RNA-seq data of
human}\label{single-cell-rna-seq-data-of-human}

In the field of single-cell studies, a common analytical task involves
clustering to identify groups of cells with similar expression profiles.
Analysts often turn to NLDR techniques to verify and identify these
clusters and explore developmental trajectories. To illustrate the
importance of NLDR techniques and parameter selection in identifying
clusters, Human Peripheral Blood Mononclear Cells (PBMC3k) dataset
\citep{chen2023} is used. In a study by \citet{chen2023}, this dataset
was used to demonstrate how UMAP represents clusters (see
Figure~\ref{fig-umapauthor}). As shown in Figure~\ref{fig-umapauthor},
there are three distinct and well-separated clusters.

\begin{figure}[H]

\centering{

\includegraphics{paper_original_files/figure-pdf/fig-umapauthor-1.pdf}

}

\caption{\label{fig-umapauthor}2D layout from UMAP (n\_neighbors = 30,
min\_dist = 0.3) applied for the PBMC3k dataset. Is this a best
representation of the original data?}

\end{figure}%

To determine whether the UMAP representation with the parameter choice
suggested by \citet{chen2023} preserves the original data structure, we
visualize the model constructed with UMAP overlaid on the high-D data
(model-in-data space). The figures in Figure~\ref{fig-pbmc1-sc} show
three well-separated clusters, indicating that the suggested UMAP
representation preserves the global structure (see
Figure~\ref{fig-umapauthor}). However, as shown in
Figure~\ref{fig-pbmc1-sc}, these clusters are close to each other in
high-D space. Also, nonlinear continuity patterns and high-density
patches within the clusters are observed (see
Figure~\ref{fig-pbmc1-sc}). Therefore, the suggested UMAP representation
(see Figure~\ref{fig-umapauthor}) does not accurately preserve the local
structure of the PBMC3k dataset.

\begin{figure}[H]

\centering{

\includegraphics{paper_original_files/figure-pdf/fig-nldervisPBMCUMAP-1.pdf}

}

\caption{\label{fig-nldervisPBMCUMAP}(a) Model generated with 21 and 28
bins along x and y axes with hexagonal size 0.03 in the 2D space
overlaid on UMAP data, and (b) high-D model error in model space.}

\end{figure}%

\begin{figure}[H]

\begin{minipage}{0.33\linewidth}

\centering{

\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_1.png}

}

\subcaption{\label{fig-pbmc1-sc1}}

\end{minipage}%
%
\begin{minipage}{0.33\linewidth}

\centering{

\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_2.png}

}

\subcaption{\label{fig-pbmc1-sc2}}

\end{minipage}%
%
\begin{minipage}{0.33\linewidth}

\centering{

\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_3.png}

}

\subcaption{\label{fig-pbmc1-sc3}}

\end{minipage}%

\caption{\label{fig-pbmc1-sc}Screen shots of the \textbf{langevitour} of
the PBMC3k data set, shows the model in high-D, a video of the tour
animation is available at (\url{https://youtu.be/Gnhh4hfsbbg}).}

\end{figure}%

To find the best NLDR representation for PBMC3k dataset, we observe the
absolute error for different number of non-empty bins and select the one
with the lowest error. As shown in Figure~\ref{fig-abserrorPBMC}, tSNE
with a perplexity set to \(51\), has the lowest error. Therefore, tSNE
with a perplexity of \(51\) is used as the best representation for
PBMC3k dataset.

\begin{figure}[H]

\centering{

\includegraphics[width=1\textwidth,height=1\textheight]{paper_original_files/figure-pdf/fig-abserrorPBMC-1.pdf}

}

\caption{\label{fig-abserrorPBMC}Absolute error from UMAP and tSNE
applied to training PBMC3k dataset with diffferent parameter choices.
What is the best parameter choice to create the model? The residual plot
have a steep slope at the beginning, indicating that a smaller number of
non-empty bins causes a larger amount of error. Then, the slope
gradually declines or level off, indicating that a higher number of
non-empty bins generates a smaller error. Using the elbow method, it was
observed that when the number of non-empty bins is set to 439, the
lowest error occurred with the parameters perplexity: 51.}

\end{figure}%

As shown in Figure~\ref{fig-tsnesuggest}, there are three well-separated
clusters, although they are located close to each other. Additionally,
nonlinear structures can be observed within the clusters (see
Figure~\ref{fig-tsnesuggest}). Hence, the data structure that UMAP
failed to capture for the PBMC3k dataset is successfully captured by
tSNE.

\begin{figure}[H]

\centering{

\includegraphics{paper_original_files/figure-pdf/fig-tsnesuggest-1.pdf}

}

\caption{\label{fig-tsnesuggest}2D layout from tSNE (perplexity = 51)
applied for the PBMC3k dataset. Is this a best representation of the
original data?}

\end{figure}%

Next, the model is fitted for tSNE, and the resulting model is
visualized in the data space. The model shows some quirks, as shown in
Figure~\ref{fig-pbmc2-sc}. The large cluster is divided into three
subclusters, which are in close to each other.

\begin{figure}[H]

\centering{

\includegraphics{paper_original_files/figure-pdf/fig-nldervisPBMCtSNE-1.pdf}

}

\caption{\label{fig-nldervisPBMCtSNE}(a) Model generated with 31 and 40
bins along x and y axes with hexagonal size 0.02 in the 2D space
overlaid on tSNE data, and (b) high-D model error in model space.}

\end{figure}%

\begin{figure}[H]

\begin{minipage}{0.33\linewidth}

\centering{

\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_4.png}

}

\subcaption{\label{fig-pbmc2-sc4}}

\end{minipage}%
%
\begin{minipage}{0.33\linewidth}

\centering{

\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_5.png}

}

\subcaption{\label{fig-pbmc2-sc5}}

\end{minipage}%
%
\begin{minipage}{0.33\linewidth}

\centering{

\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_6.png}

}

\subcaption{\label{fig-pbmc2-sc6}}

\end{minipage}%

\caption{\label{fig-pbmc2-sc}Screen shots of the \textbf{langevitour} of
the PBMC3k data set, shows the model in high-D, a video of the tour
animation is available at (\url{https://youtu.be/R0EroJcGbas}).}

\end{figure}%

\subsection{Handwritten digits}\label{handwritten-digits}

The MNIST dataset consists of grayscale images of handwritten digits
\citep{lecun2010}. \citet{Yingfan2021} used this dataset to demonstrate
how the PaCMAP preserves local structure. We selected the 2D embedding
of PaCMAP for the handwritten digit 1. As shown in
Figure~\ref{fig-pacmapauthor}, we observe a nonlinear structure present
in 2D space. Furthermore, the angle of digit 1 images varies along the
nonlinear structure (see Figure~\ref{fig-sampleimg}). To answer the
following questions,

\begin{itemize}
\tightlist
\item
  Is the original data structure of the handwritten digit 1 dataset
  nonlinear?
\item
  Where does PaCMAP preserve the original data structure?
\item
  Where does PaCMAP show quirks?
\end{itemize}

we visualize the model constructed with PaCMAP overlaid on data in
high-D space (model-in-data space).

According to Figure~\ref{fig-mnist1-sc1}, the nonlinear continuity
structure observed in the 2D representation of PaCMAP is also visible
when visualizing the model overlaid on the data space. This provides
evidence that the model accurately captures the structure of the high-D
data. Also, the model has a twisted pattern within the nonlinear
structure, which is an addition pattern of data that cannot be seen in
the 2D representation (see Figure~\ref{fig-mnist1-sc2}).

In order to visually assess the preservation of local structure, it's
important to consider the edges in the model. In a model that
effectively preserves local structure, there shouldn't be any long edges
in high-D space. However, as shown in Figure~\ref{fig-mnist1-sc3}, some
long edges exist in high-D space that are not identified as long edges
in the 2D representation.

\begin{figure}[H]

\centering{

\includegraphics{paper_original_files/figure-pdf/fig-pacmapauthor-1.pdf}

}

\caption{\label{fig-pacmapauthor}(a) 2D layout from PaCMAP
(n\_components=2, n\_neighbors=10, init=random, MN\_ratio=0.9,
FP\_ratio=2.0) applied for the digit 1 of the MNIST dataset. Is this a
best representation of the original data?}

\end{figure}%

\begin{figure}[H]

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{paper_original_files/figure-pdf/fig-sampleimg-1.pdf}

}

\caption{\label{fig-sampleimg}Images of the handwritten digit 1 are
distributed along the right-to-left of first embedding axis of the
PaCMAP nonlinear structure.}

\end{figure}%

\begin{figure}[H]

\centering{

\includegraphics{paper_original_files/figure-pdf/fig-nldervisMNISTPACMAP-1.pdf}

}

\caption{\label{fig-nldervisMNISTPACMAP}(a) Model generated with 12 and
15 bins along x and y axes with hexagonal size 0.06 in the 2D space
overlaid on PaCMAP data, and (b) high-D model error in model space.}

\end{figure}%

\begin{figure}[H]

\begin{minipage}{0.33\linewidth}

\centering{

\captionsetup{labelsep=none}\includegraphics{figures/mnist/sc_1.png}

}

\subcaption{\label{fig-mnist1-sc1}}

\end{minipage}%
%
\begin{minipage}{0.33\linewidth}

\centering{

\captionsetup{labelsep=none}\includegraphics{figures/mnist/sc_2.png}

}

\subcaption{\label{fig-mnist1-sc2}}

\end{minipage}%
%
\begin{minipage}{0.33\linewidth}

\centering{

\captionsetup{labelsep=none}\includegraphics{figures/mnist/sc_3.png}

}

\subcaption{\label{fig-mnist1-sc3}}

\end{minipage}%

\caption{\label{fig-mnist1-sc}Screen shots of the \textbf{langevitour}
of the MNIST digit 1 data set, shows the model-in-data space, a video of
the tour animation is available at
(\url{https://youtu.be/zq2GM9qvUNA}).}

\end{figure}%

In terms of model error, some data points show high error (see
Figure~\ref{fig-nldervisMNISTPACMAP}) due to their deviation from the
typical high-D data structure, acting as outliers. This is evident when
comparing the images associated with these high model error points (see
Figure~\ref{fig-sampleimgerror}) to others (see
Figure~\ref{fig-sampleimg}).

\begin{figure}[H]

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{paper_original_files/figure-pdf/fig-sampleimgerror-1.pdf}

}

\caption{\label{fig-sampleimgerror}Images of handwritten digit 1 which
occur large model error.}

\end{figure}%

\section{Conclusion}\label{sec-conclusions}

In conclusion, our proposed model offers a novel approach for
visualizing high-dimensional (high-D) data by leveraging non-linear
dimension reduction (NLDR) techniques. Through a series of steps
including hexagonating NLDR data, triangulating bin centroids, and
lifting the 2D triangular mesh into high dimensions, our model
effectively transforms complex high-D data into interpretable 2D
representations. The model's performance is evaluated using goodness of
fit statistics such as Mean Squared Error (MSE) and Akaike Information
Criterion (AIC), providing insights into its accuracy and reliability.
Overall, our model presents a valuable tool for researchers and
practitioners in various fields to gain deeper insights from
high-dimensional datasets.

Our algorithm mainly consists\ldots{}

These Goodness of fit statistics useful in encountering the prediction
accuracy when we predict the 2D embedding for the new high-D data.
Because, we try to find the nearest high-D mappings as the first step of
prediction which will participate the prediction process. When we need
to find which different NLDR technique and which parameter choices is
giving the best representation of high-D data, MSE and AIC are also
useful.

We introduce a new tool to help to determine which method, which
parameter choice provide the most useful representation of high-D data.

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\renewcommand{\bibsection}{}
\bibliography{bibliography.bib}

\newpage{}

\end{document}

### 2024-03-27 paper.qmd
Amend starting point description

--- 2024-03-26 paper.qmd
@@
 #| echo: false
-#| fig-cap: "(a) Visualization of bin centroids and (b) hexagonal grid, showcasing the configuration before (colored in gray) and after (colored in light orange) shifting the starting point by an identical amount applied in both x and y directions, with a shift value of $0.537285$."
+#| fig-cap: "(a) Visualization of bin centroids and (b) hexagonal grid, showcasing the configuration when the starting point defined at $(-0.113, -0.098)$ (colored in gray) and the starting point defined at $(-0.025, -0.108)$ (colored in light orange)."
 #| label: fig-scurveshifthexgridsexp
@@
 #| echo: false
-#| fig-cap: "Hexbin plots shows the distribution of points before and after shifting the starting point of the hexagonal grid generated for UMAP applied to the training S-curve dataset.
-(a) Hexbin plot before shifting, and (b) Hexbin plot after shifting (identical shift applied in both x and y directions, with a shift value of $0.537285$)."
+#| fig-cap: "Hexbin plots shows the distribution of points before and after changing the starting point of the hexagonal grid generated for UMAP applied to the training S-curve dataset.
+(a) Hexbin plot when the starting point defined at $(-0.113, -0.098)$, and (b) Hexbin plot after changing the starting point to $(-0.025, -0.108)$."
 #| label: fig-scurveshifthexgrids

### 2024-03-27 paper.tex
Amend starting point description

--- 2024-03-26 paper.tex
@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, sharp corners, frame hidden, breakable, enhanced, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, interior hidden, sharp corners, boxrule=0pt, enhanced, breakable, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 \caption{\label{fig-scurveshifthexgridsexp}(a) Visualization of bin
-centroids and (b) hexagonal grid, showcasing the configuration before
-(colored in gray) and after (colored in light orange) shifting the
-starting point by an identical amount applied in both x and y
-directions, with a shift value of \(0.537285\).}
+centroids and (b) hexagonal grid, showcasing the configuration when the
+starting point defined at \((-0.113, -0.098)\) (colored in gray) and the
+starting point defined at \((-0.025, -0.108)\) (colored in light
+orange).}
 
@@
 \caption{\label{fig-scurveshifthexgrids}Hexbin plots shows the
-distribution of points before and after shifting the starting point of
+distribution of points before and after changing the starting point of
 the hexagonal grid generated for UMAP applied to the training S-curve
-dataset. (a) Hexbin plot before shifting, and (b) Hexbin plot after
-shifting (identical shift applied in both x and y directions, with a
-shift value of \(0.537285\)).}
+dataset. (a) Hexbin plot when the starting point defined at
+\((-0.113, -0.098)\), and (b) Hexbin plot after changing the starting
+point to \((-0.025, -0.108)\).}
 

### 2024-03-27 paper.qmd
Added author's suggestion

--- 2024-03-27 paper.qmd
@@
 
-## Introduction {#sec-intro}
-
  - What's the problem:
@@
   Relevant high-d vis, NLDR history
+
+<!-- UMAP layout with author's suggested parameter choice-->
+```{r}
+#| warning: false
+#| echo: false
+
+## Import data
+training_data_pbmc <- read_rds("data/pbmc3k/pbmc_pca_50.rds")
+training_data_pbmc <- training_data_pbmc[, 1:9] |>
+  mutate(ID = 1:NROW(training_data_pbmc))
+
+umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_30_min_dist_0.3.rds")
+umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc, 
+                                                                 x = "UMAP1", y = "UMAP2"))) |>
+  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
+                  "UMAP2" = "scaled_UMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(umap_pbmc))
+
+umap_plot_pbmc <- umap_pbmc_scaled |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2))+
+  geom_point(alpha=0.5) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + 
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm')) 
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "2D layout from UMAP (n_neighbors = 30, min_dist = 0.3) applied for the PBMC3k dataset. Is this a best representation of the original data?"
+#| label: fig-umapauthor
+#| fig-pos: H
+
+umap_plot_pbmc 
+```
 

### 2024-03-27 paper.tex
Added author's suggestion

--- 2024-03-27 paper.tex
@@
 \makeatletter
+\makeatother
+\makeatletter
+\makeatother
+\makeatletter
 \@ifpackageloaded{caption}{}{\usepackage{caption}}
@@
 \makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
 \makeatother
 \makeatletter
-\@ifpackageloaded{caption}{}{\usepackage{caption}}
-\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
+\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
+\makeatother
+\makeatletter
+\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
+\makeatother
+\makeatletter
+\makeatother
+\makeatletter
 \makeatother
@@
 \bibliographystyle{agsm}
-\usepackage{bookmark}
-
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
@@
 
-\section{Introduction}\label{sec-intro}
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, breakable, borderline west={3pt}{0pt}{shadecolor}, interior hidden, frame hidden, enhanced, boxrule=0pt]}{\end{tcolorbox}}\fi
 
@@
 \item
-  What's the problem
+  What's the problem:
+\end{itemize}
+
+Non-linear dimension reduction being used to summarise high-dimensional
+data.
+
+\begin{itemize}
+\tightlist
 \item
@@
 
-\section{Method}\label{method}
+Relevant high-d vis, NLDR history
+
+\begin{figure}[H]
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-umapauthor-1.pdf}
+
+}
+
+\caption{\label{fig-umapauthor}2D layout from UMAP (n\_neighbors = 30,
+min\_dist = 0.3) applied for the PBMC3k dataset. Is this a best
+representation of the original data?}
+
+\end{figure}
+
+\hypertarget{method}{%
+\section{Method}\label{method}}
 
@@
 
-\section{Applications}\label{applications}
+\hypertarget{applications}{%
+\section{Applications}\label{applications}}
 
-\subsection{pbmc}\label{pbmc}
+\hypertarget{pbmc}{%
+\subsection{pbmc}\label{pbmc}}
 
@@
 
-\subsection{digits: 1}\label{digits-1}
+\hypertarget{digits-1}{%
+\subsection{digits: 1}\label{digits-1}}
 
@@
 
-\section{Discussion}\label{discussion}
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 
@@
 
-\section*{References}\label{references}
+\hypertarget{references}{%
+\section*{References}\label{references}}
 \addcontentsline{toc}{section}{References}

### 2024-03-27 paper_original.tex
Added author's suggestion

@@
 
-\section{Introduction}\label{sec-intro}
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, sharp corners, enhanced, interior hidden, boxrule=0pt, frame hidden, breakable]}{\end{tcolorbox}}\fi
+
+\hypertarget{sec-intro}{%
+\section{Introduction}\label{sec-intro}}
 
@@
 
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{paper_original_files/figure-pdf/fig-nldervisUMAP-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_original_files/figure-pdf/fig-nldervisUMAP-1.pdf}
 
@@
 
-\end{figure}%
-
-\section{Background}\label{sec-background}
-
-\subsection{Dimension Reduction}\label{dimension-reduction}
+\end{figure}
+
+\hypertarget{sec-background}{%
+\section{Background}\label{sec-background}}
+
+\hypertarget{dimension-reduction}{%
+\subsection{Dimension Reduction}\label{dimension-reduction}}
 
@@
 
+\hypertarget{non-linear-dimension-reduction-techniques}{%
 \subsubsection{Non-linear dimension reduction
-techniques}\label{non-linear-dimension-reduction-techniques}
+techniques}\label{non-linear-dimension-reduction-techniques}}
 
@@
 
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{paper_original_files/figure-pdf/fig-nldervis-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_original_files/figure-pdf/fig-nldervis-1.pdf}
 
@@
 
-\end{figure}%
-
+\end{figure}
+
+\hypertarget{linear-overviews-using-tours}{%
 \subsection{Linear overviews using
-tours}\label{linear-overviews-using-tours}
+tours}\label{linear-overviews-using-tours}}
 
@@
 
-\section{Methodology}\label{sec-methods}
+\hypertarget{sec-methods}{%
+\section{Methodology}\label{sec-methods}}
 
@@
 
-\centering{
-
-\includegraphics[width=1\textwidth,height=1\textheight]{figures/workflow.png}
+{\centering \includegraphics[width=1\textwidth,height=1\textheight]{figures/workflow.png}
 
@@
 
-\end{figure}%
-
-\subsection{Preprocessing steps}\label{preprocessing-steps}
+\end{figure}
+
+\hypertarget{preprocessing-steps}{%
+\subsection{Preprocessing steps}\label{preprocessing-steps}}
 
@@
 
-\subsection{Constructing the 2D model}\label{sec-construct2d}
+\hypertarget{sec-construct2d}{%
+\subsection{Constructing the 2D model}\label{sec-construct2d}}
 
@@
 
-\begin{equation}\phantomsection\label{eq-equation1}{
+\begin{equation}\protect\hypertarget{eq-equation1}{}{
  ar = \frac{r_2}{r_1}
-}\end{equation}
-
-\begin{equation}\phantomsection\label{eq-equation2}{
+}\label{eq-equation1}\end{equation}
+
+\begin{equation}\protect\hypertarget{eq-equation2}{}{
  hr = \frac{hb}{wb}
-}\end{equation}
-
-\begin{equation}\phantomsection\label{eq-equation3}{
+}\label{eq-equation2}\end{equation}
+
+\begin{equation}\protect\hypertarget{eq-equation3}{}{
  y_{max} = \left\lceil\frac{ar}{hr}\right\rceil * hr
-}\end{equation}
+}\label{eq-equation3}\end{equation}
 
@@
 
-\begin{equation}\phantomsection\label{eq-equation11}{
+\begin{equation}\protect\hypertarget{eq-equation11}{}{
  h = \sqrt{3} * s
-}\end{equation}
-
-\begin{equation}\phantomsection\label{eq-equation12}{
+}\label{eq-equation11}\end{equation}
+
+\begin{equation}\protect\hypertarget{eq-equation12}{}{
  b_1 = \frac{r_1 + buffer_{x}}{h}
-}\end{equation}
-
-\begin{equation}\phantomsection\label{eq-equation13}{
+}\label{eq-equation12}\end{equation}
+
+\begin{equation}\protect\hypertarget{eq-equation13}{}{
   v = 1.5 * s
-}\end{equation}
-
-\begin{equation}\phantomsection\label{eq-equation14}{
+}\label{eq-equation13}\end{equation}
+
+\begin{equation}\protect\hypertarget{eq-equation14}{}{
  b_2 = \frac{r_2 + buffer_{y}}{v}
-}\end{equation}
+}\label{eq-equation14}\end{equation}
 
@@
 
-\begin{equation}\phantomsection\label{eq-equation15}{
+\begin{equation}\protect\hypertarget{eq-equation15}{}{
 d_x = \frac{h}{2}
-}\end{equation}
-
-\begin{equation}\phantomsection\label{eq-equation16}{
+}\label{eq-equation15}\end{equation}
+
+\begin{equation}\protect\hypertarget{eq-equation16}{}{
 d_y = \frac{2* v * 1.15}{\sqrt{3}} 
-}\end{equation}
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=1\textheight]{figures/hex_coord.png}
+}\label{eq-equation16}\end{equation}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=1\textheight]{figures/hex_coord.png}
 
@@
 
-\end{figure}%
+\end{figure}
 
@@
 
-\begin{equation}\phantomsection\label{eq-equation4}{
+\begin{equation}\protect\hypertarget{eq-equation4}{}{
 C_k^{(2)} = (C_{ky_1}, C_{ky_2}) = \left(\frac{\sum_{i=1}^{6} h^{k}x_{i}}{6}, \frac{\sum_{i=1}^{6} h^{k}y_{i}}{6}\right).
-}\end{equation}
+}\label{eq-equation4}\end{equation}
 
@@
 
-\begin{equation}\phantomsection\label{eq-equation5}{
+\begin{equation}\protect\hypertarget{eq-equation5}{}{
 C_k^{(2)} = (C_{ky_1}, C_{ky_2}) = \left(\frac{1}{n_k} \sum_{i=1}^{n_k} y_{1i}, \frac{1}{n_k} \sum_{i=1}^{n_k} y_{2i}\right),
-}\end{equation}
+}\label{eq-equation5}\end{equation}
 
@@
 
+\hypertarget{lifting-the-model-into-high-dimensions}{%
 \subsection{Lifting the model into high
-dimensions}\label{lifting-the-model-into-high-dimensions}
-
+dimensions}\label{lifting-the-model-into-high-dimensions}}
+
+\hypertarget{lifting-the-triangular-mesh-points-into-high-dimensions}{%
 \subsubsection{Lifting the triangular mesh points into high
-dimensions}\label{lifting-the-triangular-mesh-points-into-high-dimensions}
+dimensions}\label{lifting-the-triangular-mesh-points-into-high-dimensions}}
 
@@
 
-\begin{equation}\phantomsection\label{eq-equation6}{
+\begin{equation}\protect\hypertarget{eq-equation6}{}{
 C_k^{(p)} = (C_{kx_1}, ..., C_{kx_p}) = \left(\frac{1}{n_k} \sum_{i=1}^{n_k} x_{1i}, \frac{1}{n_k} \sum_{i=1}^{n_k} x_{2i}, \dots ,\frac{1}{n_k} \sum_{i=1}^{n_k} x_{pi}\right).
-}\end{equation}
+}\label{eq-equation6}\end{equation}
 
@@
 
-\begin{equation}\phantomsection\label{eq-equation7}{
+\begin{equation}\protect\hypertarget{eq-equation7}{}{
 v(C_k^{(2)}) = C_k^{(p)}.
-}\end{equation}
+}\label{eq-equation7}\end{equation}
 
@@
 
+\hypertarget{lifting-the-2d-triangular-mesh-into-high-dimensions}{%
 \subsubsection{Lifting the 2D triangular mesh into high
-dimensions}\label{lifting-the-2d-triangular-mesh-into-high-dimensions}
+dimensions}\label{lifting-the-2d-triangular-mesh-into-high-dimensions}}
 
@@
 
-\centering{
-
-\includegraphics{paper_original_files/figure-pdf/fig-wkhighD-1.pdf}
+{\centering \includegraphics{paper_original_files/figure-pdf/fig-wkhighD-1.pdf}
 
@@
 
-\end{figure}%
-
-\subsection{Tuning the model}\label{tuning-the-model}
+\end{figure}
+
+\hypertarget{tuning-the-model}{%
+\subsection{Tuning the model}\label{tuning-the-model}}
 
@@
 
-\subsubsection{Total number of bins}\label{total-number-of-bins}
+\hypertarget{total-number-of-bins}{%
+\subsubsection{Total number of bins}\label{total-number-of-bins}}
 
@@
 
-\centering{
-
-\includegraphics{paper_original_files/figure-pdf/fig-binsize-1.pdf}
+{\centering \includegraphics{paper_original_files/figure-pdf/fig-binsize-1.pdf}
 
@@
 
-\end{figure}%
-
-\begin{equation}\phantomsection\label{eq-equation8}{
+\end{figure}
+
+\begin{equation}\protect\hypertarget{eq-equation8}{}{
 b = b_1 \times b_2
-}\end{equation}
+}\label{eq-equation8}\end{equation}
 
@@
 
-\centering{
-
-\includegraphics{paper_original_files/figure-pdf/fig-diagnosticpltScurve-1.pdf}
+{\centering \includegraphics{paper_original_files/figure-pdf/fig-diagnosticpltScurve-1.pdf}
 
@@
 
-\end{figure}%
-
+\end{figure}
+
+\hypertarget{benchmark-value-to-remove-low-density-hexagons}{%
 \subsubsection{Benchmark value to remove low-density
-hexagons}\label{benchmark-value-to-remove-low-density-hexagons}
+hexagons}\label{benchmark-value-to-remove-low-density-hexagons}}
 
@@
 
-\begin{equation}\phantomsection\label{eq-equationp2}{
+\begin{equation}\protect\hypertarget{eq-equationp2}{}{
 \text{standard count} = \frac{\text{count}}{\text{max count}} 
-}\end{equation}
+}\label{eq-equationp2}\end{equation}
 
@@
 
-\centering{
-
-\includegraphics{paper_original_files/figure-pdf/fig-stdctsScurve-1.pdf}
+{\centering \includegraphics{paper_original_files/figure-pdf/fig-stdctsScurve-1.pdf}
 
@@
 
-\centering{
-
-\includegraphics{paper_original_files/figure-pdf/fig-diagnosticpltScurvelwd-1.pdf}
+{\centering \includegraphics{paper_original_files/figure-pdf/fig-diagnosticpltScurvelwd-1.pdf}
 
@@
 
-\end{figure}%
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics{paper_original_files/figure-pdf/fig-rmlowdenshex-1.pdf}
+\end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{paper_original_files/figure-pdf/fig-rmlowdenshex-1.pdf}
 
@@
 
-\end{figure}%
-
+\end{figure}
+
+\hypertarget{benchmark-value-for-removing-long-edges}{%
 \subsubsection{Benchmark value for removing long
-edges}\label{benchmark-value-for-removing-long-edges}
+edges}\label{benchmark-value-for-removing-long-edges}}
 
@@
 
-\centering{
-
-\includegraphics{paper_original_files/figure-pdf/fig-distScurve-1.pdf}
+{\centering \includegraphics{paper_original_files/figure-pdf/fig-distScurve-1.pdf}
 
@@
 
-\end{figure}%
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics{paper_original_files/figure-pdf/fig-modelScurve-1.pdf}
+\end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{paper_original_files/figure-pdf/fig-modelScurve-1.pdf}
 
@@
 
-\end{figure}%
-
+\end{figure}
+
+\hypertarget{starting-point-of-the-hexagonal-grid}{%
 \subsubsection{Starting point of the hexagonal
-grid}\label{starting-point-of-the-hexagonal-grid}
+grid}\label{starting-point-of-the-hexagonal-grid}}
 
@@
 
-\centering{
-
-\includegraphics{paper_original_files/figure-pdf/fig-mseScurvestartpoint-1.pdf}
+{\centering \includegraphics{paper_original_files/figure-pdf/fig-mseScurvestartpoint-1.pdf}
 
@@
 
-\end{figure}%
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics{paper_original_files/figure-pdf/fig-scurveshifthexgridsexp-1.pdf}
+\end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{paper_original_files/figure-pdf/fig-scurveshifthexgridsexp-1.pdf}
 
@@
 \caption{\label{fig-scurveshifthexgridsexp}(a) Visualization of bin
-centroids and (b) hexagonal grid, showcasing the configuration before
-(colored in gray) and after (colored in light orange) shifting the
-starting point by an identical amount applied in both x and y
-directions, with a shift value of \(0.537285\).}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics{paper_original_files/figure-pdf/fig-scurveshifthexgrids-1.pdf}
+centroids and (b) hexagonal grid, showcasing the configuration when the
+starting point defined at \((-0.113, -0.098)\) (colored in gray) and the
+starting point defined at \((-0.025, -0.108)\) (colored in light
+orange).}
+
+\end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{paper_original_files/figure-pdf/fig-scurveshifthexgrids-1.pdf}
 
@@
 \caption{\label{fig-scurveshifthexgrids}Hexbin plots shows the
-distribution of points before and after shifting the starting point of
+distribution of points before and after changing the starting point of
 the hexagonal grid generated for UMAP applied to the training S-curve
-dataset. (a) Hexbin plot before shifting, and (b) Hexbin plot after
-shifting (identical shift applied in both x and y directions, with a
-shift value of \(0.537285\)).}
-
-\end{figure}%
-
-\subsection{Model summaries}\label{sec-summary}
-
+dataset. (a) Hexbin plot when the starting point defined at
+\((-0.113, -0.098)\), and (b) Hexbin plot after changing the starting
+point to \((-0.025, -0.108)\).}
+
+\end{figure}
+
+\hypertarget{sec-summary}{%
+\subsection{Model summaries}\label{sec-summary}}
+
+\hypertarget{predicted-values-and-residuals}{%
 \subsubsection{Predicted values and
-residuals}\label{predicted-values-and-residuals}
+residuals}\label{predicted-values-and-residuals}}
 
@@
 
-\begin{equation}\phantomsection\label{eq-equation10}{
+\begin{equation}\protect\hypertarget{eq-equation10}{}{
 \text{Error} = \sum_{j = 1}^{n}\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2
-}\end{equation}
+}\label{eq-equation10}\end{equation}
 
@@
 
-\subsubsection{Goodness of fit statistics}\label{sec-goodfit}
+\hypertarget{sec-goodfit}{%
+\subsubsection{Goodness of fit statistics}\label{sec-goodfit}}
 
@@
 
-\begin{equation}\phantomsection\label{eq-equation9}{
+\begin{equation}\protect\hypertarget{eq-equation9}{}{
 \text{MSE} = \sum_{j = 1}^{n} \frac{\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2}{n}
-}\end{equation}
-
-\subsection{Simulated data example}\label{sec-simpleex}
+}\label{eq-equation9}\end{equation}
+
+\hypertarget{sec-simpleex}{%
+\subsection{Simulated data example}\label{sec-simpleex}}
 
@@
 
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{paper_original_files/figure-pdf/fig-nldervis5Gau-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_original_files/figure-pdf/fig-nldervis5Gau-1.pdf}
 
@@
 visualize the model-in-data space. Models from all NLDR methods show
-five well-separated clusters (see Figure~\ref{fig-gau1-sc},
-Figure~\ref{fig-gau2-sc}, Figure~\ref{fig-gau3-sc},
-Figure~\ref{fig-gau4-sc}, and Figure~\ref{fig-gau5-sc}). This suggests
+five well-separated clusters (see Figure~\ref{fig-gau1_sc},
+Figure~\ref{fig-gau2_sc}, Figure~\ref{fig-gau3_sc},
+Figure~\ref{fig-gau4_sc}, and Figure~\ref{fig-gau5_sc}). This suggests
 that for the five Gaussian cluster dataset, all NLDR methods effectively
@@
 with varying densities, indicating their ability to capture
-within-cluster variation (see Figure~\ref{fig-gau1-sc},
-Figure~\ref{fig-gau2-sc}, and Figure~\ref{fig-gau3-sc}). On the other
+within-cluster variation (see Figure~\ref{fig-gau1_sc},
+Figure~\ref{fig-gau2_sc}, and Figure~\ref{fig-gau3_sc}). On the other
 hand, both TriMAP and PaCMAP show clusters with flat surfaces,
 suggesting a failure to capture within-cluster variation (see
-Figure~\ref{fig-gau4-sc} and Figure~\ref{fig-gau5-sc}). Therefore,
+Figure~\ref{fig-gau4_sc} and Figure~\ref{fig-gau5_sc}). Therefore,
 TriMAP and PaCMAP do not capture the local structure as effectively as
@@
 
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{paper_original_files/figure-pdf/fig-modelfiveGau-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_original_files/figure-pdf/fig-modelfiveGau-1.pdf}
 
@@
 
-\end{figure}%
-
-\begin{figure}[H]
-
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_tsne_1.png}
-
-}
-
-\subcaption{\label{fig-gau1-sc1}}
-
+\end{figure}
+
+\begin{figure}[H]
+
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/five_gau_clusters/sc_tsne_1.png}
+
+}
+
+}
+
+\subcaption{\label{fig-gau1_sc1}}
 \end{minipage}%
 %
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_tsne_2.png}
-
-}
-
-\subcaption{\label{fig-gau1-sc2}}
-
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/five_gau_clusters/sc_tsne_2.png}
+
+}
+
+}
+
+\subcaption{\label{fig-gau1_sc2}}
 \end{minipage}%
 %
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_tsne_3.png}
-
-}
-
-\subcaption{\label{fig-gau1-sc3}}
-
-\end{minipage}%
-
-\caption{\label{fig-gau1-sc}Screen shots of the \textbf{langevitour} of
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/five_gau_clusters/sc_tsne_3.png}
+
+}
+
+}
+
+\subcaption{\label{fig-gau1_sc3}}
+\end{minipage}%
+
+\caption{\label{fig-gau1_sc}Screen shots of the \textbf{langevitour} of
 the five Gaussian clusters dataset, shows the model with tSNE in high-D,
@@
 
-\end{figure}%
-
-\begin{figure}[H]
-
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_umap_1.png}
-
-}
-
-\subcaption{\label{fig-gau2-sc1}}
-
+\end{figure}
+
+\begin{figure}[H]
+
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/five_gau_clusters/sc_umap_1.png}
+
+}
+
+}
+
+\subcaption{\label{fig-gau2_sc1}}
 \end{minipage}%
 %
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_umap_2.png}
-
-}
-
-\subcaption{\label{fig-gau2-sc2}}
-
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/five_gau_clusters/sc_umap_2.png}
+
+}
+
+}
+
+\subcaption{\label{fig-gau2_sc2}}
 \end{minipage}%
 %
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_umap_3.png}
-
-}
-
-\subcaption{\label{fig-gau2-sc3}}
-
-\end{minipage}%
-
-\caption{\label{fig-gau2-sc}Screen shots of the \textbf{langevitour} of
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/five_gau_clusters/sc_umap_3.png}
+
+}
+
+}
+
+\subcaption{\label{fig-gau2_sc3}}
+\end{minipage}%
+
+\caption{\label{fig-gau2_sc}Screen shots of the \textbf{langevitour} of
 the five Gaussian clusters dataset, shows the model with UMAP in high-D,
@@
 
-\end{figure}%
-
-\begin{figure}[H]
-
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_phate_1.png}
-
-}
-
-\subcaption{\label{fig-gau3-sc1}}
-
+\end{figure}
+
+\begin{figure}[H]
+
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/five_gau_clusters/sc_phate_1.png}
+
+}
+
+}
+
+\subcaption{\label{fig-gau3_sc1}}
 \end{minipage}%
 %
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_phate_2.png}
-
-}
-
-\subcaption{\label{fig-gau3-sc2}}
-
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/five_gau_clusters/sc_phate_2.png}
+
+}
+
+}
+
+\subcaption{\label{fig-gau3_sc2}}
 \end{minipage}%
 %
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_phate_3.png}
-
-}
-
-\subcaption{\label{fig-gau3-sc3}}
-
-\end{minipage}%
-
-\caption{\label{fig-gau3-sc}Screen shots of the \textbf{langevitour} of
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/five_gau_clusters/sc_phate_3.png}
+
+}
+
+}
+
+\subcaption{\label{fig-gau3_sc3}}
+\end{minipage}%
+
+\caption{\label{fig-gau3_sc}Screen shots of the \textbf{langevitour} of
 the five Gaussian clusters dataset, shows the model with PHATE in
@@
 
-\end{figure}%
-
-\begin{figure}[H]
-
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_trimap_1.png}
-
-}
-
-\subcaption{\label{fig-gau4-sc1}}
-
+\end{figure}
+
+\begin{figure}[H]
+
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/five_gau_clusters/sc_trimap_1.png}
+
+}
+
+}
+
+\subcaption{\label{fig-gau4_sc1}}
 \end{minipage}%
 %
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_trimap_2.png}
-
-}
-
-\subcaption{\label{fig-gau4-sc2}}
-
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/five_gau_clusters/sc_trimap_2.png}
+
+}
+
+}
+
+\subcaption{\label{fig-gau4_sc2}}
 \end{minipage}%
 %
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_trimap_3.png}
-
-}
-
-\subcaption{\label{fig-gau4-sc3}}
-
-\end{minipage}%
-
-\caption{\label{fig-gau4-sc}Screen shots of the \textbf{langevitour} of
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/five_gau_clusters/sc_trimap_3.png}
+
+}
+
+}
+
+\subcaption{\label{fig-gau4_sc3}}
+\end{minipage}%
+
+\caption{\label{fig-gau4_sc}Screen shots of the \textbf{langevitour} of
 the five Gaussian clusters dataset, shows the model with TriMAP in
@@
 
-\end{figure}%
-
-\begin{figure}[H]
-
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_pacmap_1.png}
-
-}
-
-\subcaption{\label{fig-gau5-sc1}}
-
+\end{figure}
+
+\begin{figure}[H]
+
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/five_gau_clusters/sc_pacmap_1.png}
+
+}
+
+}
+
+\subcaption{\label{fig-gau5_sc1}}
 \end{minipage}%
 %
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_pacmap_2.png}
-
-}
-
-\subcaption{\label{fig-gau5-sc2}}
-
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/five_gau_clusters/sc_pacmap_2.png}
+
+}
+
+}
+
+\subcaption{\label{fig-gau5_sc2}}
 \end{minipage}%
 %
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_pacmap_3.png}
-
-}
-
-\subcaption{\label{fig-gau5-sc3}}
-
-\end{minipage}%
-
-\caption{\label{fig-gau5-sc}Screen shots of the \textbf{langevitour} of
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/five_gau_clusters/sc_pacmap_3.png}
+
+}
+
+}
+
+\subcaption{\label{fig-gau5_sc3}}
+\end{minipage}%
+
+\caption{\label{fig-gau5_sc}Screen shots of the \textbf{langevitour} of
 the five Gaussian clusters dataset, shows the model with PaCMAP in
@@
 
-\end{figure}%
-
-\section{Applications}\label{sec-applications}
-
+\end{figure}
+
+\hypertarget{sec-applications}{%
+\section{Applications}\label{sec-applications}}
+
+\hypertarget{single-cell-rna-seq-data-of-human}{%
 \subsection{Single-cell RNA-seq data of
-human}\label{single-cell-rna-seq-data-of-human}
+human}\label{single-cell-rna-seq-data-of-human}}
 
@@
 
-\centering{
-
-\includegraphics{paper_original_files/figure-pdf/fig-umapauthor-1.pdf}
+{\centering \includegraphics{paper_original_files/figure-pdf/fig-umapauthor-1.pdf}
 
@@
 visualize the model constructed with UMAP overlaid on the high-D data
-(model-in-data space). The figures in Figure~\ref{fig-pbmc1-sc} show
+(model-in-data space). The figures in Figure~\ref{fig-pbmc1_sc} show
 three well-separated clusters, indicating that the suggested UMAP
@@
 Figure~\ref{fig-umapauthor}). However, as shown in
-Figure~\ref{fig-pbmc1-sc}, these clusters are close to each other in
+Figure~\ref{fig-pbmc1_sc}, these clusters are close to each other in
 high-D space. Also, nonlinear continuity patterns and high-density
 patches within the clusters are observed (see
-Figure~\ref{fig-pbmc1-sc}). Therefore, the suggested UMAP representation
+Figure~\ref{fig-pbmc1_sc}). Therefore, the suggested UMAP representation
 (see Figure~\ref{fig-umapauthor}) does not accurately preserve the local
@@
 
-\centering{
-
-\includegraphics{paper_original_files/figure-pdf/fig-nldervisPBMCUMAP-1.pdf}
+{\centering \includegraphics{paper_original_files/figure-pdf/fig-nldervisPBMCUMAP-1.pdf}
 
@@
 
-\end{figure}%
-
-\begin{figure}[H]
-
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_1.png}
-
-}
-
-\subcaption{\label{fig-pbmc1-sc1}}
-
+\end{figure}
+
+\begin{figure}[H]
+
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/pbmc3k/sc_1.png}
+
+}
+
+}
+
+\subcaption{\label{fig-pbmc1_sc1}}
 \end{minipage}%
 %
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_2.png}
-
-}
-
-\subcaption{\label{fig-pbmc1-sc2}}
-
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/pbmc3k/sc_2.png}
+
+}
+
+}
+
+\subcaption{\label{fig-pbmc1_sc2}}
 \end{minipage}%
 %
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_3.png}
-
-}
-
-\subcaption{\label{fig-pbmc1-sc3}}
-
-\end{minipage}%
-
-\caption{\label{fig-pbmc1-sc}Screen shots of the \textbf{langevitour} of
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/pbmc3k/sc_3.png}
+
+}
+
+}
+
+\subcaption{\label{fig-pbmc1_sc3}}
+\end{minipage}%
+
+\caption{\label{fig-pbmc1_sc}Screen shots of the \textbf{langevitour} of
 the PBMC3k data set, shows the model in high-D, a video of the tour
@@
 
-\centering{
-
-\includegraphics[width=1\textwidth,height=1\textheight]{paper_original_files/figure-pdf/fig-abserrorPBMC-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=1\textheight]{paper_original_files/figure-pdf/fig-abserrorPBMC-1.pdf}
 
@@
 
-\centering{
-
-\includegraphics{paper_original_files/figure-pdf/fig-tsnesuggest-1.pdf}
+{\centering \includegraphics{paper_original_files/figure-pdf/fig-tsnesuggest-1.pdf}
 
@@
 visualized in the data space. The model shows some quirks, as shown in
-Figure~\ref{fig-pbmc2-sc}. The large cluster is divided into three
+Figure~\ref{fig-pbmc2_sc}. The large cluster is divided into three
 subclusters, which are in close to each other.
@@
 
-\centering{
-
-\includegraphics{paper_original_files/figure-pdf/fig-nldervisPBMCtSNE-1.pdf}
+{\centering \includegraphics{paper_original_files/figure-pdf/fig-nldervisPBMCtSNE-1.pdf}
 
@@
 
-\end{figure}%
-
-\begin{figure}[H]
-
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_4.png}
-
-}
-
-\subcaption{\label{fig-pbmc2-sc4}}
-
+\end{figure}
+
+\begin{figure}[H]
+
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/pbmc3k/sc_4.png}
+
+}
+
+}
+
+\subcaption{\label{fig-pbmc2_sc4}}
 \end{minipage}%
 %
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_5.png}
-
-}
-
-\subcaption{\label{fig-pbmc2-sc5}}
-
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/pbmc3k/sc_5.png}
+
+}
+
+}
+
+\subcaption{\label{fig-pbmc2_sc5}}
 \end{minipage}%
 %
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_6.png}
-
-}
-
-\subcaption{\label{fig-pbmc2-sc6}}
-
-\end{minipage}%
-
-\caption{\label{fig-pbmc2-sc}Screen shots of the \textbf{langevitour} of
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/pbmc3k/sc_6.png}
+
+}
+
+}
+
+\subcaption{\label{fig-pbmc2_sc6}}
+\end{minipage}%
+
+\caption{\label{fig-pbmc2_sc}Screen shots of the \textbf{langevitour} of
 the PBMC3k data set, shows the model in high-D, a video of the tour
@@
 
-\end{figure}%
-
-\subsection{Handwritten digits}\label{handwritten-digits}
+\end{figure}
+
+\hypertarget{handwritten-digits}{%
+\subsection{Handwritten digits}\label{handwritten-digits}}
 
@@
 
-According to Figure~\ref{fig-mnist1-sc1}, the nonlinear continuity
+According to Figure~\ref{fig-mnist1_sc1}, the nonlinear continuity
 structure observed in the 2D representation of PaCMAP is also visible
@@
 structure, which is an addition pattern of data that cannot be seen in
-the 2D representation (see Figure~\ref{fig-mnist1-sc2}).
+the 2D representation (see Figure~\ref{fig-mnist1_sc2}).
 
@@
 effectively preserves local structure, there shouldn't be any long edges
-in high-D space. However, as shown in Figure~\ref{fig-mnist1-sc3}, some
+in high-D space. However, as shown in Figure~\ref{fig-mnist1_sc3}, some
 long edges exist in high-D space that are not identified as long edges
@@
 
-\centering{
-
-\includegraphics{paper_original_files/figure-pdf/fig-pacmapauthor-1.pdf}
+{\centering \includegraphics{paper_original_files/figure-pdf/fig-pacmapauthor-1.pdf}
 
@@
 
-\end{figure}%
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{paper_original_files/figure-pdf/fig-sampleimg-1.pdf}
+\end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_original_files/figure-pdf/fig-sampleimg-1.pdf}
 
@@
 
-\end{figure}%
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics{paper_original_files/figure-pdf/fig-nldervisMNISTPACMAP-1.pdf}
+\end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{paper_original_files/figure-pdf/fig-nldervisMNISTPACMAP-1.pdf}
 
@@
 
-\end{figure}%
-
-\begin{figure}[H]
-
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/mnist/sc_1.png}
-
-}
-
-\subcaption{\label{fig-mnist1-sc1}}
-
+\end{figure}
+
+\begin{figure}[H]
+
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/mnist/sc_1.png}
+
+}
+
+}
+
+\subcaption{\label{fig-mnist1_sc1}}
 \end{minipage}%
 %
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/mnist/sc_2.png}
-
-}
-
-\subcaption{\label{fig-mnist1-sc2}}
-
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/mnist/sc_2.png}
+
+}
+
+}
+
+\subcaption{\label{fig-mnist1_sc2}}
 \end{minipage}%
 %
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/mnist/sc_3.png}
-
-}
-
-\subcaption{\label{fig-mnist1-sc3}}
-
-\end{minipage}%
-
-\caption{\label{fig-mnist1-sc}Screen shots of the \textbf{langevitour}
+\begin{minipage}[t]{0.33\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{figures/mnist/sc_3.png}
+
+}
+
+}
+
+\subcaption{\label{fig-mnist1_sc3}}
+\end{minipage}%
+
+\caption{\label{fig-mnist1_sc}Screen shots of the \textbf{langevitour}
 of the MNIST digit 1 data set, shows the model-in-data space, a video of
@@
 
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{paper_original_files/figure-pdf/fig-sampleimgerror-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_original_files/figure-pdf/fig-sampleimgerror-1.pdf}
 
@@
 
-\end{figure}%
-
-\section{Conclusion}\label{sec-conclusions}
+\end{figure}
+
+\hypertarget{sec-conclusions}{%
+\section{Conclusion}\label{sec-conclusions}}
 

### 2024-03-27 paper.tex
Update quarto and rendered

--- 2024-03-27 paper.tex
@@
 
-\section{Introduction}\label{sec-intro}
+\begin{itemize}
+\tightlist
+\item
+  What's the problem:
+\end{itemize}
+
+Non-linear dimension reduction being used to summarise high-dimensional
+data.
 
@@
 \item
-  What's the problem
-\item
   Summary of literature
 \end{itemize}
+
+Relevant high-d vis, NLDR history
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics{paper_files/figure-pdf/fig-umapauthor-1.pdf}
+
+}
+
+\caption{\label{fig-umapauthor}2D layout from UMAP (n\_neighbors = 30,
+min\_dist = 0.3) applied for the PBMC3k dataset. Is this a best
+representation of the original data?}
+
+\end{figure}%
 

### 2024-03-27 paper.qmd
Add different 2D layouts for PBMC data with different parameter choices

@@
 umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_30_min_dist_0.3.rds")
-umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc, 
-                                                                 x = "UMAP1", y = "UMAP2"))) |>
+umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc, x = "UMAP1", y = "UMAP2"))) |>
   dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
@@
              y = UMAP2))+
-  geom_point(alpha=0.5) +
+  geom_point(alpha=0.1, size=1) +
   coord_equal() +
@@
 ```{r}
+#| warning: false
+#| echo: false
+
+umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_5_min_dist_0.01.rds")
+umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc, x = "UMAP1", y = "UMAP2"))) |>
+  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
+                  "UMAP2" = "scaled_UMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(umap_pbmc))
+
+umap_plot_pbmc_1 <- umap_pbmc_scaled |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + 
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm'))
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_15_min_dist_0.99.rds")
+umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc, x = "UMAP1", y = "UMAP2"))) |>
+  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
+                  "UMAP2" = "scaled_UMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(umap_pbmc))
+
+umap_plot_pbmc_2 <- umap_pbmc_scaled |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + 
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm'))
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_15_min_dist_0.03.rds")
+umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc, x = "UMAP1", y = "UMAP2"))) |>
+  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
+                  "UMAP2" = "scaled_UMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(umap_pbmc))
+
+umap_plot_pbmc_3 <- umap_pbmc_scaled |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + 
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm'))
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_84_min_dist_0.99.rds")
+umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc, x = "UMAP1", y = "UMAP2"))) |>
+  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
+                  "UMAP2" = "scaled_UMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(umap_pbmc))
+
+umap_plot_pbmc_4 <- umap_pbmc_scaled |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + 
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm'))
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_5.rds")
+tsne_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tsne_pbmc, 
+                                                                 x = "tSNE1", y = "tSNE2"))) |>
+  dplyr::rename(c("tSNE1" = "scaled_tSNE1", 
+                  "tSNE2" = "scaled_tSNE2")) |>
+  dplyr::mutate(ID = 1:NROW(tsne_pbmc))
+
+tsne_plot_pbmc_1 <- tsne_pbmc_scaled |>
+  ggplot(aes(x = tSNE1,
+             y = tSNE2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm')) 
+```
+
+```{r}
+#| warning: false
+#| echo: false
+tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_51.rds")
+tsne_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tsne_pbmc, 
+                                                                 x = "tSNE1", y = "tSNE2"))) |>
+  dplyr::rename(c("tSNE1" = "scaled_tSNE1", 
+                  "tSNE2" = "scaled_tSNE2")) |>
+  dplyr::mutate(ID = 1:NROW(tsne_pbmc))
+
+tsne_plot_pbmc <- tsne_pbmc_scaled |>
+  ggplot(aes(x = tSNE1,
+             y = tSNE2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm')) 
+```
+
+```{r}
+#| warning: false
+#| echo: false
+tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_92.rds")
+tsne_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tsne_pbmc, 
+                                                                 x = "tSNE1", y = "tSNE2"))) |>
+  dplyr::rename(c("tSNE1" = "scaled_tSNE1", 
+                  "tSNE2" = "scaled_tSNE2")) |>
+  dplyr::mutate(ID = 1:NROW(tsne_pbmc))
+
+tsne_plot_pbmc_2 <- tsne_pbmc_scaled |>
+  ggplot(aes(x = tSNE1,
+             y = tSNE2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm')) 
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+phate_pbmc <- read_rds("data/pbmc3k/pbmc_phate_2.rds")
+phate_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = phate_pbmc, x = "PHATE1", y = "PHATE2"))) |>
+  dplyr::rename(c("PHATE1" = "scaled_PHATE1", 
+                  "PHATE2" = "scaled_PHATE2")) |>
+  dplyr::mutate(ID = 1:NROW(phate_pbmc))
+
+phate_plot_pbmc_1 <- phate_pbmc_scaled |>
+  ggplot(aes(x = PHATE1,
+             y = PHATE2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + 
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm'))
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+phate_pbmc <- read_rds("data/pbmc3k/pbmc_phate_5.rds")
+phate_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = phate_pbmc, x = "PHATE1", y = "PHATE2"))) |>
+  dplyr::rename(c("PHATE1" = "scaled_PHATE1", 
+                  "PHATE2" = "scaled_PHATE2")) |>
+  dplyr::mutate(ID = 1:NROW(phate_pbmc))
+
+phate_plot_pbmc_2 <- phate_pbmc_scaled |>
+  ggplot(aes(x = PHATE1,
+             y = PHATE2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + 
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm'))
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+phate_pbmc <- read_rds("data/pbmc3k/pbmc_phate_10.rds")
+phate_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = phate_pbmc, x = "PHATE1", y = "PHATE2"))) |>
+  dplyr::rename(c("PHATE1" = "scaled_PHATE1", 
+                  "PHATE2" = "scaled_PHATE2")) |>
+  dplyr::mutate(ID = 1:NROW(phate_pbmc))
+
+phate_plot_pbmc_3 <- phate_pbmc_scaled |>
+  ggplot(aes(x = PHATE1,
+             y = PHATE2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + 
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm'))
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+trimap_pbmc <- read_rds("data/pbmc3k/pbmc_trimap_12_4_3.rds")
+trimap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = trimap_pbmc, x = "TriMAP1", y = "TriMAP2"))) |>
+  dplyr::rename(c("TriMAP1" = "scaled_TriMAP1",
+                  "TriMAP2" = "scaled_TriMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(trimap_pbmc))
+
+trimap_plot_pbmc_1 <- trimap_pbmc_scaled |>
+  ggplot(aes(x = TriMAP1,
+             y = TriMAP2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) +
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm'))
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+trimap_pbmc <- read_rds("data/pbmc3k/pbmc_trimap_30_4_10.rds")
+trimap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = trimap_pbmc, x = "TriMAP1", y = "TriMAP2"))) |>
+  dplyr::rename(c("TriMAP1" = "scaled_TriMAP1",
+                  "TriMAP2" = "scaled_TriMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(trimap_pbmc))
+
+trimap_plot_pbmc_2 <- trimap_pbmc_scaled |>
+  ggplot(aes(x = TriMAP1,
+             y = TriMAP2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) +
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm'))
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+trimap_pbmc <- read_rds("data/pbmc3k/pbmc_trimap_5_2_2.rds")
+trimap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = trimap_pbmc, x = "TriMAP1", y = "TriMAP2"))) |>
+  dplyr::rename(c("TriMAP1" = "scaled_TriMAP1",
+                  "TriMAP2" = "scaled_TriMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(trimap_pbmc))
+
+trimap_plot_pbmc_3 <- trimap_pbmc_scaled |>
+  ggplot(aes(x = TriMAP1,
+             y = TriMAP2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) +
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm'))
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+pacmap_pbmc <- read_rds("data/pbmc3k/pbmc_pacmap_10_pca_0.5_2.rds")
+pacmap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = pacmap_pbmc, x = "PaCMAP1", y = "PaCMAP2"))) |>
+  dplyr::rename(c("PaCMAP1" = "scaled_PaCMAP1",
+                  "PaCMAP2" = "scaled_PaCMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(pacmap_pbmc))
+
+pacmap_plot_pbmc_1 <- pacmap_pbmc_scaled |>
+  ggplot(aes(x = PaCMAP1,
+             y = PaCMAP2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) +
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm'))
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+pacmap_pbmc <- read_rds("data/pbmc3k/pbmc_pacmap_30_random_0.9_5.rds")
+pacmap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = pacmap_pbmc, x = "PaCMAP1", y = "PaCMAP2"))) |>
+  dplyr::rename(c("PaCMAP1" = "scaled_PaCMAP1",
+                  "PaCMAP2" = "scaled_PaCMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(pacmap_pbmc))
+
+pacmap_plot_pbmc_2 <- pacmap_pbmc_scaled |>
+  ggplot(aes(x = PaCMAP1,
+             y = PaCMAP2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) +
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm'))
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+pacmap_pbmc <- read_rds("data/pbmc3k/pbmc_pacmap_5_pca_0.1_1.rds")
+pacmap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = pacmap_pbmc, x = "PaCMAP1", y = "PaCMAP2"))) |>
+  dplyr::rename(c("PaCMAP1" = "scaled_PaCMAP1",
+                  "PaCMAP2" = "scaled_PaCMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(pacmap_pbmc))
+
+pacmap_plot_pbmc_3 <- pacmap_pbmc_scaled |>
+  ggplot(aes(x = PaCMAP1,
+             y = PaCMAP2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) +
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm'))
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+pacmap_pbmc <- read_rds("data/pbmc3k/pbmc_pacmap_30_random_0.1_1.rds")
+pacmap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = pacmap_pbmc, x = "PaCMAP1", y = "PaCMAP2"))) |>
+  dplyr::rename(c("PaCMAP1" = "scaled_PaCMAP1",
+                  "PaCMAP2" = "scaled_PaCMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(pacmap_pbmc))
+
+pacmap_plot_pbmc_4 <- pacmap_pbmc_scaled |>
+  ggplot(aes(x = PaCMAP1,
+             y = PaCMAP2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) +
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm'))
+
+```
+
+```{r}
 #| echo: false
@@
 
-umap_plot_pbmc 
+umap_plot_pbmc + umap_plot_pbmc_1 + umap_plot_pbmc_2 + umap_plot_pbmc_3 + umap_plot_pbmc_4 + tsne_plot_pbmc_1 + tsne_plot_pbmc + tsne_plot_pbmc_2 + phate_plot_pbmc_1 + phate_plot_pbmc_2 + phate_plot_pbmc_3 + trimap_plot_pbmc_1 + trimap_plot_pbmc_2 + trimap_plot_pbmc_3 + pacmap_plot_pbmc_1 + pacmap_plot_pbmc_2 + pacmap_plot_pbmc_3 + pacmap_plot_pbmc_4 + plot_layout(ncol = 6)
 ```

### 2024-03-27 paper.qmd
Added figure description

--- 2024-03-27 paper.qmd
@@
 
+## Introduction
+
  - What's the problem:
@@
         legend.key.height = unit(0.25, 'cm'),
-        legend.key.width = unit(0.25, 'cm')) 
+        legend.key.width = unit(0.25, 'cm')) +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_5_min_dist_0.01.rds")
+umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc, x = "UMAP1", y = "UMAP2"))) |>
+  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
+                  "UMAP2" = "scaled_UMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(umap_pbmc))
+
+umap_plot_pbmc_1 <- umap_pbmc_scaled |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + 
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm')) +
+  annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_15_min_dist_0.99.rds")
+umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc, x = "UMAP1", y = "UMAP2"))) |>
+  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
+                  "UMAP2" = "scaled_UMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(umap_pbmc))
+
+umap_plot_pbmc_2 <- umap_pbmc_scaled |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + 
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm')) +
+  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_15_min_dist_0.03.rds")
+umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc, x = "UMAP1", y = "UMAP2"))) |>
+  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
+                  "UMAP2" = "scaled_UMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(umap_pbmc))
+
+umap_plot_pbmc_3 <- umap_pbmc_scaled |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + 
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm')) +
+  annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.3, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_84_min_dist_0.99.rds")
+umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc, x = "UMAP1", y = "UMAP2"))) |>
+  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
+                  "UMAP2" = "scaled_UMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(umap_pbmc))
+
+umap_plot_pbmc_4 <- umap_pbmc_scaled |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + 
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm')) +
+  annotate(geom = 'text', label = 'e', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_5.rds")
+tsne_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tsne_pbmc, 
+                                                                 x = "tSNE1", y = "tSNE2"))) |>
+  dplyr::rename(c("tSNE1" = "scaled_tSNE1", 
+                  "tSNE2" = "scaled_tSNE2")) |>
+  dplyr::mutate(ID = 1:NROW(tsne_pbmc))
+
+tsne_plot_pbmc_1 <- tsne_pbmc_scaled |>
+  ggplot(aes(x = tSNE1,
+             y = tSNE2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm')) +
+  annotate(geom = 'text', label = 'f', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)   
+```
+
+```{r}
+#| warning: false
+#| echo: false
+tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_51.rds")
+tsne_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tsne_pbmc, 
+                                                                 x = "tSNE1", y = "tSNE2"))) |>
+  dplyr::rename(c("tSNE1" = "scaled_tSNE1", 
+                  "tSNE2" = "scaled_tSNE2")) |>
+  dplyr::mutate(ID = 1:NROW(tsne_pbmc))
+
+tsne_plot_pbmc <- tsne_pbmc_scaled |>
+  ggplot(aes(x = tSNE1,
+             y = tSNE2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm')) +
+  annotate(geom = 'text', label = 'g', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.1, size = 3)   
+```
+
+```{r}
+#| warning: false
+#| echo: false
+tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_92.rds")
+tsne_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tsne_pbmc, 
+                                                                 x = "tSNE1", y = "tSNE2"))) |>
+  dplyr::rename(c("tSNE1" = "scaled_tSNE1", 
+                  "tSNE2" = "scaled_tSNE2")) |>
+  dplyr::mutate(ID = 1:NROW(tsne_pbmc))
+
+tsne_plot_pbmc_2 <- tsne_pbmc_scaled |>
+  ggplot(aes(x = tSNE1,
+             y = tSNE2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm')) +
+  annotate(geom = 'text', label = 'h', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)   
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+phate_pbmc <- read_rds("data/pbmc3k/pbmc_phate_2.rds")
+phate_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = phate_pbmc, x = "PHATE1", y = "PHATE2"))) |>
+  dplyr::rename(c("PHATE1" = "scaled_PHATE1", 
+                  "PHATE2" = "scaled_PHATE2")) |>
+  dplyr::mutate(ID = 1:NROW(phate_pbmc))
+
+phate_plot_pbmc_1 <- phate_pbmc_scaled |>
+  ggplot(aes(x = PHATE1,
+             y = PHATE2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + 
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm')) +
+  annotate(geom = 'text', label = 'i', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+phate_pbmc <- read_rds("data/pbmc3k/pbmc_phate_5.rds")
+phate_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = phate_pbmc, x = "PHATE1", y = "PHATE2"))) |>
+  dplyr::rename(c("PHATE1" = "scaled_PHATE1", 
+                  "PHATE2" = "scaled_PHATE2")) |>
+  dplyr::mutate(ID = 1:NROW(phate_pbmc))
+
+phate_plot_pbmc_2 <- phate_pbmc_scaled |>
+  ggplot(aes(x = PHATE1,
+             y = PHATE2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + 
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm')) +
+  annotate(geom = 'text', label = 'j', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+phate_pbmc <- read_rds("data/pbmc3k/pbmc_phate_10.rds")
+phate_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = phate_pbmc, x = "PHATE1", y = "PHATE2"))) |>
+  dplyr::rename(c("PHATE1" = "scaled_PHATE1", 
+                  "PHATE2" = "scaled_PHATE2")) |>
+  dplyr::mutate(ID = 1:NROW(phate_pbmc))
+
+phate_plot_pbmc_3 <- phate_pbmc_scaled |>
+  ggplot(aes(x = PHATE1,
+             y = PHATE2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + 
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm')) +
+  annotate(geom = 'text', label = 'k', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+trimap_pbmc <- read_rds("data/pbmc3k/pbmc_trimap_12_4_3.rds")
+trimap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = trimap_pbmc, x = "TriMAP1", y = "TriMAP2"))) |>
+  dplyr::rename(c("TriMAP1" = "scaled_TriMAP1",
+                  "TriMAP2" = "scaled_TriMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(trimap_pbmc))
+
+trimap_plot_pbmc_1 <- trimap_pbmc_scaled |>
+  ggplot(aes(x = TriMAP1,
+             y = TriMAP2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) +
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm')) +
+  annotate(geom = 'text', label = 'l', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+trimap_pbmc <- read_rds("data/pbmc3k/pbmc_trimap_30_4_10.rds")
+trimap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = trimap_pbmc, x = "TriMAP1", y = "TriMAP2"))) |>
+  dplyr::rename(c("TriMAP1" = "scaled_TriMAP1",
+                  "TriMAP2" = "scaled_TriMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(trimap_pbmc))
+
+trimap_plot_pbmc_2 <- trimap_pbmc_scaled |>
+  ggplot(aes(x = TriMAP1,
+             y = TriMAP2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) +
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm')) +
+  annotate(geom = 'text', label = 'm', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+trimap_pbmc <- read_rds("data/pbmc3k/pbmc_trimap_5_2_2.rds")
+trimap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = trimap_pbmc, x = "TriMAP1", y = "TriMAP2"))) |>
+  dplyr::rename(c("TriMAP1" = "scaled_TriMAP1",
+                  "TriMAP2" = "scaled_TriMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(trimap_pbmc))
+
+trimap_plot_pbmc_3 <- trimap_pbmc_scaled |>
+  ggplot(aes(x = TriMAP1,
+             y = TriMAP2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) +
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm')) +
+  annotate(geom = 'text', label = 'n', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+pacmap_pbmc <- read_rds("data/pbmc3k/pbmc_pacmap_10_pca_0.5_2.rds")
+pacmap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = pacmap_pbmc, x = "PaCMAP1", y = "PaCMAP2"))) |>
+  dplyr::rename(c("PaCMAP1" = "scaled_PaCMAP1",
+                  "PaCMAP2" = "scaled_PaCMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(pacmap_pbmc))
+
+pacmap_plot_pbmc_1 <- pacmap_pbmc_scaled |>
+  ggplot(aes(x = PaCMAP1,
+             y = PaCMAP2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) +
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm')) +
+  annotate(geom = 'text', label = 'p', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+pacmap_pbmc <- read_rds("data/pbmc3k/pbmc_pacmap_30_random_0.9_5.rds")
+pacmap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = pacmap_pbmc, x = "PaCMAP1", y = "PaCMAP2"))) |>
+  dplyr::rename(c("PaCMAP1" = "scaled_PaCMAP1",
+                  "PaCMAP2" = "scaled_PaCMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(pacmap_pbmc))
+
+pacmap_plot_pbmc_2 <- pacmap_pbmc_scaled |>
+  ggplot(aes(x = PaCMAP1,
+             y = PaCMAP2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) +
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm')) +
+  annotate(geom = 'text', label = 'q', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+pacmap_pbmc <- read_rds("data/pbmc3k/pbmc_pacmap_5_pca_0.1_1.rds")
+pacmap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = pacmap_pbmc, x = "PaCMAP1", y = "PaCMAP2"))) |>
+  dplyr::rename(c("PaCMAP1" = "scaled_PaCMAP1",
+                  "PaCMAP2" = "scaled_PaCMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(pacmap_pbmc))
+
+pacmap_plot_pbmc_3 <- pacmap_pbmc_scaled |>
+  ggplot(aes(x = PaCMAP1,
+             y = PaCMAP2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) +
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm')) +
+  annotate(geom = 'text', label = 'r', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+pacmap_pbmc <- read_rds("data/pbmc3k/pbmc_pacmap_30_random_0.1_1.rds")
+pacmap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = pacmap_pbmc, x = "PaCMAP1", y = "PaCMAP2"))) |>
+  dplyr::rename(c("PaCMAP1" = "scaled_PaCMAP1",
+                  "PaCMAP2" = "scaled_PaCMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(pacmap_pbmc))
+
+pacmap_plot_pbmc_4 <- pacmap_pbmc_scaled |>
+  ggplot(aes(x = PaCMAP1,
+             y = PaCMAP2))+
+  geom_point(alpha=0.1, size=1) +
+  coord_equal() +
+  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) +
+  theme_linedraw() +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+        axis.title.x = element_blank(), axis.title.y = element_blank(),
+        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+        legend.key.height = unit(0.25, 'cm'),
+        legend.key.width = unit(0.25, 'cm')) +
+  annotate(geom = 'text', label = 's', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
 ```
@@
 #| echo: false
-#| fig-cap: "2D layout from UMAP (n_neighbors = 30, min_dist = 0.3) applied for the PBMC3k dataset. Is this a best representation of the original data?"
-#| label: fig-umapauthor
+#| fig-cap: "2D layouts from different NLDR techniques and different hyperparameter choices applied for the PBMC3k dataset: (a) UMAP (n_neighbors = 30, min_dist = 0.3), (b) UMAP (n_neighbors = 5, min_dist = 0.01), (c) UMAP (n_neighbors = 15, min_dist = 0.99), (d) UMAP (n_neighbors = 15, min_dist = 0.03), (e) UMAP (n_neighbors = 84, min_dist = 0.99), (f) tSNE (perplexity = 5), (g) tSNE (perplexity = 51), (h) tSNE (perplexity = 92), (i) PHATE (knn = 2), (j) PHATE (knn = 5), (k) PHATE (knn = 10), (l) TriMAP (n_inliers = 12, n_outliers = 4, n_random = 3), (m) TriMAP (n_inliers = 30, n_outliers = 4, n_random = 10), (n) TriMAP (n_inliers = 5, n_outliers = 2, n_random = 2), (p) PaCMAP (n_neighbors = 10, init = pca, MN_ratio = 0.5, FP_ratio = 2), (q) PaCMAP (n_neighbors = 30, init = random, MN_ratio = 0.9, FP_ratio = 5), (r) PaCMAP (n_neighbors = 5, init = pca, MN_ratio = 0.1, FP_ratio = 1), and (s) PaCMAP (n_neighbors = 30, init = random, MN_ratio = 0.1, FP_ratio = 1).
+Is there a best representation of the original data or are they all providing  equivalent information?"
+#| label: fig-diffnldrpbmc
 #| fig-pos: H
+#| out-width: 100%
 
-umap_plot_pbmc 
+umap_plot_pbmc + umap_plot_pbmc_1 + umap_plot_pbmc_2 + umap_plot_pbmc_3 + umap_plot_pbmc_4 + tsne_plot_pbmc_1 + tsne_plot_pbmc + tsne_plot_pbmc_2 + phate_plot_pbmc_1 + phate_plot_pbmc_2 + phate_plot_pbmc_3 + trimap_plot_pbmc_1 + trimap_plot_pbmc_2 + trimap_plot_pbmc_3 + pacmap_plot_pbmc_1 + pacmap_plot_pbmc_2 + pacmap_plot_pbmc_3 + pacmap_plot_pbmc_4 + plot_layout(ncol = 6)
 ```

### 2024-03-27 paper.tex
Added figure description

@@
 
+\section{Introduction}\label{introduction}
+
 \begin{itemize}
@@
 
-\includegraphics{paper_files/figure-pdf/fig-umapauthor-1.pdf}
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diffnldrpbmc-1.pdf}
 
@@
 
-\caption{\label{fig-umapauthor}2D layout from UMAP (n\_neighbors = 30,
-min\_dist = 0.3) applied for the PBMC3k dataset. Is this a best
-representation of the original data?}
+\caption{\label{fig-diffnldrpbmc}2D layouts from different NLDR
+techniques and different hyperparameter choices applied for the PBMC3k
+dataset: (a) UMAP (n\_neighbors = 30, min\_dist = 0.3), (b) UMAP
+(n\_neighbors = 5, min\_dist = 0.01), (c) UMAP (n\_neighbors = 15,
+min\_dist = 0.99), (d) UMAP (n\_neighbors = 15, min\_dist = 0.03), (e)
+UMAP (n\_neighbors = 84, min\_dist = 0.99), (f) tSNE (perplexity = 5),
+(g) tSNE (perplexity = 51), (h) tSNE (perplexity = 92), (i) PHATE (knn =
+2), (j) PHATE (knn = 5), (k) PHATE (knn = 10), (l) TriMAP (n\_inliers =
+12, n\_outliers = 4, n\_random = 3), (m) TriMAP (n\_inliers = 30,
+n\_outliers = 4, n\_random = 10), (n) TriMAP (n\_inliers = 5,
+n\_outliers = 2, n\_random = 2), (p) PaCMAP (n\_neighbors = 10, init =
+pca, MN\_ratio = 0.5, FP\_ratio = 2), (q) PaCMAP (n\_neighbors = 30,
+init = random, MN\_ratio = 0.9, FP\_ratio = 5), (r) PaCMAP (n\_neighbors
+= 5, init = pca, MN\_ratio = 0.1, FP\_ratio = 1), and (s) PaCMAP
+(n\_neighbors = 30, init = random, MN\_ratio = 0.1, FP\_ratio = 1). Is
+there a best representation of the original data or are they all
+providing equivalent information?}
 

### 2024-03-27 paper.qmd
tidied up fist image and code

--- 2024-03-27 paper.qmd
@@
 
+```{r include=FALSE}
+# Set up chunk for for knitr
+knitr::opts_chunk$set(
+  fig.width = 5,
+  fig.height = 5,
+  fig.align = "center",
+  out.width = "100%",
+  code.line.numbers = FALSE,
+  fig.retina = 4,
+  echo = FALSE,
+  message = FALSE,
+  warning = FALSE,
+  cache = FALSE,
+  dev.args = list(pointsize = 11)
+)
+```
+
 ```{r}
+#| label: load-libraries
 #| warning: false
@@
 library(grid)
+```
+
+```{r}
+#| label: plot-theme
+theme_set(theme_linedraw() +
+   theme(
+     aspect.ratio = 1,
+     plot.background = element_rect(fill = 'transparent', colour = NA),
+     plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+     panel.background = element_rect(fill = 'transparent', 
+                                     colour = NA),
+     panel.grid.major = element_blank(), 
+     panel.grid.minor = element_blank(), 
+     axis.title.x = element_blank(), axis.title.y = element_blank(),
+     axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+     axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+     legend.background = element_rect(fill = 'transparent', 
+                                      colour = NA),
+     legend.key = element_rect(fill = 'transparent', 
+                               colour = NA),
+     legend.position = "none", 
+     legend.title = element_text(size=5), 
+     legend.text = element_text(size=4),
+     legend.key.height = unit(0.25, 'cm'),
+     legend.key.width = unit(0.25, 'cm')
+   )
+)
+```
+
+```{r}
+#| label: code-setup
 set.seed(20240110)
-
 source("nldr_code.R", local = TRUE)
@@
 ```
+
+## Introduction
 
@@
 umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_30_min_dist_0.3.rds")
-umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc, 
-                                                                 x = "UMAP1", y = "UMAP2"))) |>
+umap_pbmc_scaled <- as.data.frame(
+  do.call(cbind, 
+          gen_scaled_data(data = umap_pbmc, 
+                          x = "UMAP1", 
+                          y = "UMAP2"))) |>
   dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
@@
              y = UMAP2))+
-  geom_point(alpha=0.5) +
-  coord_equal() +
-  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + 
-  theme_linedraw() +
-  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title.x = element_blank(), axis.title.y = element_blank(),
-        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
-        legend.title = element_text(size=5), #change legend title font size
-        legend.text = element_text(size=4),
-        legend.key.height = unit(0.25, 'cm'),
-        legend.key.width = unit(0.25, 'cm')) 
+  geom_point(alpha=0.1, size=1) +
+  #coord_equal() +
+  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_5_min_dist_0.01.rds")
+umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc, x = "UMAP1", y = "UMAP2"))) |>
+  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
+                  "UMAP2" = "scaled_UMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(umap_pbmc))
+
+umap_plot_pbmc_1 <- umap_pbmc_scaled |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2))+
+  geom_point(alpha=0.1, size=1) +
+  #coord_equal() +
+  annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_15_min_dist_0.99.rds")
+umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc, x = "UMAP1", y = "UMAP2"))) |>
+  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
+                  "UMAP2" = "scaled_UMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(umap_pbmc))
+
+umap_plot_pbmc_2 <- umap_pbmc_scaled |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2))+
+  geom_point(alpha=0.1, size=1) +
+  #coord_equal() +
+  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_15_min_dist_0.03.rds")
+umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc, x = "UMAP1", y = "UMAP2"))) |>
+  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
+                  "UMAP2" = "scaled_UMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(umap_pbmc))
+
+umap_plot_pbmc_3 <- umap_pbmc_scaled |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2))+
+  geom_point(alpha=0.1, size=1) +
+  #coord_equal() +
+  annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.3, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_84_min_dist_0.99.rds")
+umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc, x = "UMAP1", y = "UMAP2"))) |>
+  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
+                  "UMAP2" = "scaled_UMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(umap_pbmc))
+
+umap_plot_pbmc_4 <- umap_pbmc_scaled |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2))+
+  geom_point(alpha=0.1, size=1) +
+  #coord_equal() +
+  annotate(geom = 'text', label = 'e', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_5.rds")
+tsne_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tsne_pbmc, 
+                                                                 x = "tSNE1", y = "tSNE2"))) |>
+  dplyr::rename(c("tSNE1" = "scaled_tSNE1", 
+                  "tSNE2" = "scaled_tSNE2")) |>
+  dplyr::mutate(ID = 1:NROW(tsne_pbmc))
+
+tsne_plot_pbmc_1 <- tsne_pbmc_scaled |>
+  ggplot(aes(x = tSNE1,
+             y = tSNE2))+
+  geom_point(alpha=0.1, size=1) +
+  #coord_equal() +
+  annotate(geom = 'text', label = 'f', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)   
+```
+
+```{r}
+#| warning: false
+#| echo: false
+tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_51.rds")
+tsne_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tsne_pbmc, 
+                                                                 x = "tSNE1", y = "tSNE2"))) |>
+  dplyr::rename(c("tSNE1" = "scaled_tSNE1", 
+                  "tSNE2" = "scaled_tSNE2")) |>
+  dplyr::mutate(ID = 1:NROW(tsne_pbmc))
+
+tsne_plot_pbmc <- tsne_pbmc_scaled |>
+  ggplot(aes(x = tSNE1,
+             y = tSNE2))+
+  geom_point(alpha=0.1, size=1) +
+  #coord_equal() +
+  annotate(geom = 'text', label = 'g', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.1, size = 3)   
+```
+
+```{r}
+#| warning: false
+#| echo: false
+tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_92.rds")
+tsne_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tsne_pbmc, 
+                                                                 x = "tSNE1", y = "tSNE2"))) |>
+  dplyr::rename(c("tSNE1" = "scaled_tSNE1", 
+                  "tSNE2" = "scaled_tSNE2")) |>
+  dplyr::mutate(ID = 1:NROW(tsne_pbmc))
+
+tsne_plot_pbmc_2 <- tsne_pbmc_scaled |>
+  ggplot(aes(x = tSNE1,
+             y = tSNE2))+
+  geom_point(alpha=0.1, size=1) +
+  #coord_equal() +
+  annotate(geom = 'text', label = 'h', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)   
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+phate_pbmc <- read_rds("data/pbmc3k/pbmc_phate_2.rds")
+phate_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = phate_pbmc, x = "PHATE1", y = "PHATE2"))) |>
+  dplyr::rename(c("PHATE1" = "scaled_PHATE1", 
+                  "PHATE2" = "scaled_PHATE2")) |>
+  dplyr::mutate(ID = 1:NROW(phate_pbmc))
+
+phate_plot_pbmc_1 <- phate_pbmc_scaled |>
+  ggplot(aes(x = PHATE1,
+             y = PHATE2))+
+  geom_point(alpha=0.1, size=1) +
+  #coord_equal() +
+  annotate(geom = 'text', label = 'i', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+phate_pbmc <- read_rds("data/pbmc3k/pbmc_phate_5.rds")
+phate_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = phate_pbmc, x = "PHATE1", y = "PHATE2"))) |>
+  dplyr::rename(c("PHATE1" = "scaled_PHATE1", 
+                  "PHATE2" = "scaled_PHATE2")) |>
+  dplyr::mutate(ID = 1:NROW(phate_pbmc))
+
+phate_plot_pbmc_2 <- phate_pbmc_scaled |>
+  ggplot(aes(x = PHATE1,
+             y = PHATE2))+
+  geom_point(alpha=0.1, size=1) +
+  #coord_equal() +
+  annotate(geom = 'text', label = 'j', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+phate_pbmc <- read_rds("data/pbmc3k/pbmc_phate_10.rds")
+phate_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = phate_pbmc, x = "PHATE1", y = "PHATE2"))) |>
+  dplyr::rename(c("PHATE1" = "scaled_PHATE1", 
+                  "PHATE2" = "scaled_PHATE2")) |>
+  dplyr::mutate(ID = 1:NROW(phate_pbmc))
+
+phate_plot_pbmc_3 <- phate_pbmc_scaled |>
+  ggplot(aes(x = PHATE1,
+             y = PHATE2))+
+  geom_point(alpha=0.1, size=1) +
+  #coord_equal() +
+  annotate(geom = 'text', label = 'k', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+trimap_pbmc <- read_rds("data/pbmc3k/pbmc_trimap_12_4_3.rds")
+trimap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = trimap_pbmc, x = "TriMAP1", y = "TriMAP2"))) |>
+  dplyr::rename(c("TriMAP1" = "scaled_TriMAP1",
+                  "TriMAP2" = "scaled_TriMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(trimap_pbmc))
+
+trimap_plot_pbmc_1 <- trimap_pbmc_scaled |>
+  ggplot(aes(x = TriMAP1,
+             y = TriMAP2))+
+  geom_point(alpha=0.1, size=1) +
+  #coord_equal() +
+  annotate(geom = 'text', label = 'l', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+trimap_pbmc <- read_rds("data/pbmc3k/pbmc_trimap_30_4_10.rds")
+trimap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = trimap_pbmc, x = "TriMAP1", y = "TriMAP2"))) |>
+  dplyr::rename(c("TriMAP1" = "scaled_TriMAP1",
+                  "TriMAP2" = "scaled_TriMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(trimap_pbmc))
+
+trimap_plot_pbmc_2 <- trimap_pbmc_scaled |>
+  ggplot(aes(x = TriMAP1,
+             y = TriMAP2))+
+  geom_point(alpha=0.1, size=1) +
+  #coord_equal() +
+  annotate(geom = 'text', label = 'm', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+trimap_pbmc <- read_rds("data/pbmc3k/pbmc_trimap_5_2_2.rds")
+trimap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = trimap_pbmc, x = "TriMAP1", y = "TriMAP2"))) |>
+  dplyr::rename(c("TriMAP1" = "scaled_TriMAP1",
+                  "TriMAP2" = "scaled_TriMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(trimap_pbmc))
+
+trimap_plot_pbmc_3 <- trimap_pbmc_scaled |>
+  ggplot(aes(x = TriMAP1,
+             y = TriMAP2))+
+  geom_point(alpha=0.1, size=1) +
+  #coord_equal() +
+  annotate(geom = 'text', label = 'n', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+pacmap_pbmc <- read_rds("data/pbmc3k/pbmc_pacmap_10_pca_0.5_2.rds")
+pacmap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = pacmap_pbmc, x = "PaCMAP1", y = "PaCMAP2"))) |>
+  dplyr::rename(c("PaCMAP1" = "scaled_PaCMAP1",
+                  "PaCMAP2" = "scaled_PaCMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(pacmap_pbmc))
+
+pacmap_plot_pbmc_1 <- pacmap_pbmc_scaled |>
+  ggplot(aes(x = PaCMAP1,
+             y = PaCMAP2))+
+  geom_point(alpha=0.1, size=1) +
+  #coord_equal() +
+  annotate(geom = 'text', label = 'p', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+pacmap_pbmc <- read_rds("data/pbmc3k/pbmc_pacmap_30_random_0.9_5.rds")
+pacmap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = pacmap_pbmc, x = "PaCMAP1", y = "PaCMAP2"))) |>
+  dplyr::rename(c("PaCMAP1" = "scaled_PaCMAP1",
+                  "PaCMAP2" = "scaled_PaCMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(pacmap_pbmc))
+
+pacmap_plot_pbmc_2 <- pacmap_pbmc_scaled |>
+  ggplot(aes(x = PaCMAP1,
+             y = PaCMAP2))+
+  geom_point(alpha=0.1, size=1) +
+  #coord_equal() +
+  annotate(geom = 'text', label = 'q', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+pacmap_pbmc <- read_rds("data/pbmc3k/pbmc_pacmap_5_pca_0.1_1.rds")
+pacmap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = pacmap_pbmc, x = "PaCMAP1", y = "PaCMAP2"))) |>
+  dplyr::rename(c("PaCMAP1" = "scaled_PaCMAP1",
+                  "PaCMAP2" = "scaled_PaCMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(pacmap_pbmc))
+
+pacmap_plot_pbmc_3 <- pacmap_pbmc_scaled |>
+  ggplot(aes(x = PaCMAP1,
+             y = PaCMAP2))+
+  geom_point(alpha=0.1, size=1) +
+  #coord_equal() +
+  annotate(geom = 'text', label = 'r', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
+
+```
+
+```{r}
+#| warning: false
+#| echo: false
+
+pacmap_pbmc <- read_rds("data/pbmc3k/pbmc_pacmap_30_random_0.1_1.rds")
+pacmap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = pacmap_pbmc, x = "PaCMAP1", y = "PaCMAP2"))) |>
+  dplyr::rename(c("PaCMAP1" = "scaled_PaCMAP1",
+                  "PaCMAP2" = "scaled_PaCMAP2")) |>
+  dplyr::mutate(ID = 1:NROW(pacmap_pbmc))
+
+pacmap_plot_pbmc_4 <- pacmap_pbmc_scaled |>
+  ggplot(aes(x = PaCMAP1,
+             y = PaCMAP2))+
+  geom_point(alpha=0.1, size=1) +
+  #coord_equal() +
+  annotate(geom = 'text', label = 's', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+
 ```
@@
 #| echo: false
-#| fig-cap: "2D layout from UMAP (n_neighbors = 30, min_dist = 0.3) applied for the PBMC3k dataset. Is this a best representation of the original data?"
-#| label: fig-umapauthor
-#| fig-pos: H
-
-umap_plot_pbmc 
+#| fig-cap: "Six different NLDR representations of the same data.
+Different techniques and different parameter choices are used.
+Which is the most accurate representation of the structure in high dimensions?"
+#| label: fig-diffnldrpbmc
+#| fig-width: 8
+#| fig-height: 3
+#| out-width: 100%
+# 2D layouts from different NLDR techniques and different hyperparameter choices applied for the PBMC3k dataset: (a) UMAP (n_neighbors = 30, min_dist = 0.3), (b) UMAP (n_neighbors = 5, min_dist = 0.01), (c) UMAP (n_neighbors = 15, min_dist = 0.99), (d) UMAP (n_neighbors = 15, min_dist = 0.03), (e) UMAP (n_neighbors = 84, min_dist = 0.99), (f) tSNE (perplexity = 5), (g) tSNE (perplexity = 51), (h) tSNE (perplexity = 92), (i) PHATE (knn = 2), (j) PHATE (knn = 5), (k) PHATE (knn = 10), (l) TriMAP (n_inliers = 12, n_outliers = 4, n_random = 3), (m) TriMAP (n_inliers = 30, n_outliers = 4, n_random = 10), (n) TriMAP (n_inliers = 5, n_outliers = 2, n_random = 2), (p) PaCMAP (n_neighbors = 10, init = pca, MN_ratio = 0.5, FP_ratio = 2), (q) PaCMAP (n_neighbors = 30, init = random, MN_ratio = 0.9, FP_ratio = 5), (r) PaCMAP (n_neighbors = 5, init = pca, MN_ratio = 0.1, FP_ratio = 1), and (s) PaCMAP (n_neighbors = 30, init = random, MN_ratio = 0.1, FP_ratio = 1).
+Is there a best representation of the original data or are they all providing  equivalent information?
+umap_plot_pbmc + umap_plot_pbmc_1 + umap_plot_pbmc_2 + #umap_plot_pbmc_3 + umap_plot_pbmc_4 + 
+  tsne_plot_pbmc_1 + tsne_plot_pbmc + 
+  #tsne_plot_pbmc_2 + phate_plot_pbmc_1 + 
+  phate_plot_pbmc_2 + #phate_plot_pbmc_3 + 
+  trimap_plot_pbmc_1 + #trimap_plot_pbmc_2 + trimap_plot_pbmc_3 + pacmap_plot_pbmc_1 + pacmap_plot_pbmc_2 + 
+  pacmap_plot_pbmc_3 + #pacmap_plot_pbmc_4 + 
+  plot_layout(ncol = 4)
 ```

### 2024-03-27 paper.tex
tidied up fist image and code

--- 2024-03-27 paper.tex
@@
 
-\begin{figure}[H]
+\begin{figure}
 
@@
 
-\caption{\label{fig-umapauthor}2D layout from UMAP (n\_neighbors = 30,
-min\_dist = 0.3) applied for the PBMC3k dataset. Is this a best
-representation of the original data?}
+\caption{\label{fig-diffnldrpbmc}Six different NLDR representations of
+the same data. Different techniques and different parameter choices are
+used. Which is the most accurate representation of the structure in high
+dimensions?}
 

### 2024-03-27 paper.qmd
tidied up intro code

@@
 
-<!-- UMAP layout with author's suggested parameter choice-->
 ```{r}
-#| warning: false
-#| echo: false
+umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_30_min_dist_0.3.rds")
 
-## Import data
-training_data_pbmc <- read_rds("data/pbmc3k/pbmc_pca_50.rds")
-training_data_pbmc <- training_data_pbmc[, 1:9] |>
-  mutate(ID = 1:NROW(training_data_pbmc))
+nldr1 <- umap_pbmc |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2)) +
+  geom_point(alpha=0.1, size=1) +
+  ggtitle("a")
+```
 
-umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_30_min_dist_0.3.rds")
-umap_pbmc_scaled <- as.data.frame(
-  do.call(cbind, 
-          gen_scaled_data(data = umap_pbmc, 
-                          x = "UMAP1", 
-                          y = "UMAP2"))) |>
-  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
-                  "UMAP2" = "scaled_UMAP2")) |>
-  dplyr::mutate(ID = 1:NROW(umap_pbmc))
+```{r}
+umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_5_min_dist_0.01.rds")
 
-umap_plot_pbmc <- umap_pbmc_scaled |>
+nldr2 <- umap_pbmc |>
   ggplot(aes(x = UMAP1,
@@
   geom_point(alpha=0.1, size=1) +
-  #coord_equal() +
-  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_5_min_dist_0.01.rds")
-umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc, x = "UMAP1", y = "UMAP2"))) |>
-  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
-                  "UMAP2" = "scaled_UMAP2")) |>
-  dplyr::mutate(ID = 1:NROW(umap_pbmc))
-
-umap_plot_pbmc_1 <- umap_pbmc_scaled |>
-  ggplot(aes(x = UMAP1,
-             y = UMAP2))+
-  geom_point(alpha=0.1, size=1) +
-  #coord_equal() +
-  annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
+  ggtitle("b")
 
@@
 ```{r}
-#| warning: false
-#| echo: false
-
 umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_15_min_dist_0.99.rds")
-umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc, x = "UMAP1", y = "UMAP2"))) |>
-  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
-                  "UMAP2" = "scaled_UMAP2")) |>
-  dplyr::mutate(ID = 1:NROW(umap_pbmc))
-
-umap_plot_pbmc_2 <- umap_pbmc_scaled |>
+nldr3 <- umap_pbmc |>
   ggplot(aes(x = UMAP1,
@@
   geom_point(alpha=0.1, size=1) +
-  #coord_equal() +
-  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
+  ggtitle("c") 
 ```
@@
 ```{r}
-#| warning: false
-#| echo: false
+tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_5.rds")
 
-umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_15_min_dist_0.03.rds")
-umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc, x = "UMAP1", y = "UMAP2"))) |>
-  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
-                  "UMAP2" = "scaled_UMAP2")) |>
-  dplyr::mutate(ID = 1:NROW(umap_pbmc))
-
-umap_plot_pbmc_3 <- umap_pbmc_scaled |>
-  ggplot(aes(x = UMAP1,
-             y = UMAP2))+
+nldr4 <- tsne_pbmc |>
+  ggplot(aes(x = tSNE1,
+             y = tSNE2))+
   geom_point(alpha=0.1, size=1) +
-  #coord_equal() +
-  annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.3, size = 3)
-
+  ggtitle("d")  
 ```
@@
 ```{r}
-#| warning: false
-#| echo: false
+tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_51.rds")
 
-umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_84_min_dist_0.99.rds")
-umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc, x = "UMAP1", y = "UMAP2"))) |>
-  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
-                  "UMAP2" = "scaled_UMAP2")) |>
-  dplyr::mutate(ID = 1:NROW(umap_pbmc))
-
-umap_plot_pbmc_4 <- umap_pbmc_scaled |>
-  ggplot(aes(x = UMAP1,
-             y = UMAP2))+
+nldr5 <- tsne_pbmc |>
+  ggplot(aes(x = tSNE1,
+             y = tSNE2))+
   geom_point(alpha=0.1, size=1) +
-  #coord_equal() +
-  annotate(geom = 'text', label = 'e', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
+  ggtitle("e") 
 ```
@@
 ```{r}
-#| warning: false
-#| echo: false
-tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_5.rds")
-tsne_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tsne_pbmc, 
-                                                                 x = "tSNE1", y = "tSNE2"))) |>
-  dplyr::rename(c("tSNE1" = "scaled_tSNE1", 
-                  "tSNE2" = "scaled_tSNE2")) |>
-  dplyr::mutate(ID = 1:NROW(tsne_pbmc))
-
-tsne_plot_pbmc_1 <- tsne_pbmc_scaled |>
-  ggplot(aes(x = tSNE1,
-             y = tSNE2))+
+phate_pbmc <- read_rds("data/pbmc3k/pbmc_phate_5.rds")
+nldr6 <- phate_pbmc |>
+  ggplot(aes(x = PHATE1,
+             y = PHATE2))+
   geom_point(alpha=0.1, size=1) +
-  #coord_equal() +
-  annotate(geom = 'text', label = 'f', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)   
+  ggtitle("f")  
 ```
@@
 ```{r}
-#| warning: false
-#| echo: false
-tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_51.rds")
-tsne_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tsne_pbmc, 
-                                                                 x = "tSNE1", y = "tSNE2"))) |>
-  dplyr::rename(c("tSNE1" = "scaled_tSNE1", 
-                  "tSNE2" = "scaled_tSNE2")) |>
-  dplyr::mutate(ID = 1:NROW(tsne_pbmc))
-
-tsne_plot_pbmc <- tsne_pbmc_scaled |>
-  ggplot(aes(x = tSNE1,
-             y = tSNE2))+
+trimap_pbmc <- read_rds("data/pbmc3k/pbmc_trimap_12_4_3.rds")
+nldr7 <- trimap_pbmc |>
+  ggplot(aes(x = TriMAP1,
+             y = TriMAP2))+
   geom_point(alpha=0.1, size=1) +
-  #coord_equal() +
-  annotate(geom = 'text', label = 'g', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.1, size = 3)   
+  ggtitle("g") 
 ```
@@
 ```{r}
-#| warning: false
-#| echo: false
-tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_92.rds")
-tsne_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tsne_pbmc, 
-                                                                 x = "tSNE1", y = "tSNE2"))) |>
-  dplyr::rename(c("tSNE1" = "scaled_tSNE1", 
-                  "tSNE2" = "scaled_tSNE2")) |>
-  dplyr::mutate(ID = 1:NROW(tsne_pbmc))
-
-tsne_plot_pbmc_2 <- tsne_pbmc_scaled |>
-  ggplot(aes(x = tSNE1,
-             y = tSNE2))+
-  geom_point(alpha=0.1, size=1) +
-  #coord_equal() +
-  annotate(geom = 'text', label = 'h', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)   
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-phate_pbmc <- read_rds("data/pbmc3k/pbmc_phate_2.rds")
-phate_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = phate_pbmc, x = "PHATE1", y = "PHATE2"))) |>
-  dplyr::rename(c("PHATE1" = "scaled_PHATE1", 
-                  "PHATE2" = "scaled_PHATE2")) |>
-  dplyr::mutate(ID = 1:NROW(phate_pbmc))
-
-phate_plot_pbmc_1 <- phate_pbmc_scaled |>
-  ggplot(aes(x = PHATE1,
-             y = PHATE2))+
-  geom_point(alpha=0.1, size=1) +
-  #coord_equal() +
-  annotate(geom = 'text', label = 'i', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-phate_pbmc <- read_rds("data/pbmc3k/pbmc_phate_5.rds")
-phate_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = phate_pbmc, x = "PHATE1", y = "PHATE2"))) |>
-  dplyr::rename(c("PHATE1" = "scaled_PHATE1", 
-                  "PHATE2" = "scaled_PHATE2")) |>
-  dplyr::mutate(ID = 1:NROW(phate_pbmc))
-
-phate_plot_pbmc_2 <- phate_pbmc_scaled |>
-  ggplot(aes(x = PHATE1,
-             y = PHATE2))+
-  geom_point(alpha=0.1, size=1) +
-  #coord_equal() +
-  annotate(geom = 'text', label = 'j', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-phate_pbmc <- read_rds("data/pbmc3k/pbmc_phate_10.rds")
-phate_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = phate_pbmc, x = "PHATE1", y = "PHATE2"))) |>
-  dplyr::rename(c("PHATE1" = "scaled_PHATE1", 
-                  "PHATE2" = "scaled_PHATE2")) |>
-  dplyr::mutate(ID = 1:NROW(phate_pbmc))
-
-phate_plot_pbmc_3 <- phate_pbmc_scaled |>
-  ggplot(aes(x = PHATE1,
-             y = PHATE2))+
-  geom_point(alpha=0.1, size=1) +
-  #coord_equal() +
-  annotate(geom = 'text', label = 'k', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-trimap_pbmc <- read_rds("data/pbmc3k/pbmc_trimap_12_4_3.rds")
-trimap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = trimap_pbmc, x = "TriMAP1", y = "TriMAP2"))) |>
-  dplyr::rename(c("TriMAP1" = "scaled_TriMAP1",
-                  "TriMAP2" = "scaled_TriMAP2")) |>
-  dplyr::mutate(ID = 1:NROW(trimap_pbmc))
-
-trimap_plot_pbmc_1 <- trimap_pbmc_scaled |>
-  ggplot(aes(x = TriMAP1,
-             y = TriMAP2))+
-  geom_point(alpha=0.1, size=1) +
-  #coord_equal() +
-  annotate(geom = 'text', label = 'l', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-trimap_pbmc <- read_rds("data/pbmc3k/pbmc_trimap_30_4_10.rds")
-trimap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = trimap_pbmc, x = "TriMAP1", y = "TriMAP2"))) |>
-  dplyr::rename(c("TriMAP1" = "scaled_TriMAP1",
-                  "TriMAP2" = "scaled_TriMAP2")) |>
-  dplyr::mutate(ID = 1:NROW(trimap_pbmc))
-
-trimap_plot_pbmc_2 <- trimap_pbmc_scaled |>
-  ggplot(aes(x = TriMAP1,
-             y = TriMAP2))+
-  geom_point(alpha=0.1, size=1) +
-  #coord_equal() +
-  annotate(geom = 'text', label = 'm', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-trimap_pbmc <- read_rds("data/pbmc3k/pbmc_trimap_5_2_2.rds")
-trimap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = trimap_pbmc, x = "TriMAP1", y = "TriMAP2"))) |>
-  dplyr::rename(c("TriMAP1" = "scaled_TriMAP1",
-                  "TriMAP2" = "scaled_TriMAP2")) |>
-  dplyr::mutate(ID = 1:NROW(trimap_pbmc))
-
-trimap_plot_pbmc_3 <- trimap_pbmc_scaled |>
-  ggplot(aes(x = TriMAP1,
-             y = TriMAP2))+
-  geom_point(alpha=0.1, size=1) +
-  #coord_equal() +
-  annotate(geom = 'text', label = 'n', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-pacmap_pbmc <- read_rds("data/pbmc3k/pbmc_pacmap_10_pca_0.5_2.rds")
-pacmap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = pacmap_pbmc, x = "PaCMAP1", y = "PaCMAP2"))) |>
-  dplyr::rename(c("PaCMAP1" = "scaled_PaCMAP1",
-                  "PaCMAP2" = "scaled_PaCMAP2")) |>
-  dplyr::mutate(ID = 1:NROW(pacmap_pbmc))
-
-pacmap_plot_pbmc_1 <- pacmap_pbmc_scaled |>
+pacmap_pbmc <- read_rds("data/pbmc3k/pbmc_pacmap_30_random_0.9_5.rds")
+nldr8 <- pacmap_pbmc |>
   ggplot(aes(x = PaCMAP1,
@@
   geom_point(alpha=0.1, size=1) +
-  #coord_equal() +
-  annotate(geom = 'text', label = 'p', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-pacmap_pbmc <- read_rds("data/pbmc3k/pbmc_pacmap_30_random_0.9_5.rds")
-pacmap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = pacmap_pbmc, x = "PaCMAP1", y = "PaCMAP2"))) |>
-  dplyr::rename(c("PaCMAP1" = "scaled_PaCMAP1",
-                  "PaCMAP2" = "scaled_PaCMAP2")) |>
-  dplyr::mutate(ID = 1:NROW(pacmap_pbmc))
-
-pacmap_plot_pbmc_2 <- pacmap_pbmc_scaled |>
-  ggplot(aes(x = PaCMAP1,
-             y = PaCMAP2))+
-  geom_point(alpha=0.1, size=1) +
-  #coord_equal() +
-  annotate(geom = 'text', label = 'q', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-pacmap_pbmc <- read_rds("data/pbmc3k/pbmc_pacmap_5_pca_0.1_1.rds")
-pacmap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = pacmap_pbmc, x = "PaCMAP1", y = "PaCMAP2"))) |>
-  dplyr::rename(c("PaCMAP1" = "scaled_PaCMAP1",
-                  "PaCMAP2" = "scaled_PaCMAP2")) |>
-  dplyr::mutate(ID = 1:NROW(pacmap_pbmc))
-
-pacmap_plot_pbmc_3 <- pacmap_pbmc_scaled |>
-  ggplot(aes(x = PaCMAP1,
-             y = PaCMAP2))+
-  geom_point(alpha=0.1, size=1) +
-  #coord_equal() +
-  annotate(geom = 'text', label = 'r', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
-
-```
-
-```{r}
-#| warning: false
-#| echo: false
-
-pacmap_pbmc <- read_rds("data/pbmc3k/pbmc_pacmap_30_random_0.1_1.rds")
-pacmap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = pacmap_pbmc, x = "PaCMAP1", y = "PaCMAP2"))) |>
-  dplyr::rename(c("PaCMAP1" = "scaled_PaCMAP1",
-                  "PaCMAP2" = "scaled_PaCMAP2")) |>
-  dplyr::mutate(ID = 1:NROW(pacmap_pbmc))
-
-pacmap_plot_pbmc_4 <- pacmap_pbmc_scaled |>
-  ggplot(aes(x = PaCMAP1,
-             y = PaCMAP2))+
-  geom_point(alpha=0.1, size=1) +
-  #coord_equal() +
-  annotate(geom = 'text', label = 's', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
-
+  ggtitle("h")
 ```
@@
 Is there a best representation of the original data or are they all providing  equivalent information?
-umap_plot_pbmc + umap_plot_pbmc_1 + umap_plot_pbmc_2 + #umap_plot_pbmc_3 + umap_plot_pbmc_4 + 
-  tsne_plot_pbmc_1 + tsne_plot_pbmc + 
-  #tsne_plot_pbmc_2 + phate_plot_pbmc_1 + 
-  phate_plot_pbmc_2 + #phate_plot_pbmc_3 + 
-  trimap_plot_pbmc_1 + #trimap_plot_pbmc_2 + trimap_plot_pbmc_3 + pacmap_plot_pbmc_1 + pacmap_plot_pbmc_2 + 
-  pacmap_plot_pbmc_3 + #pacmap_plot_pbmc_4 + 
+nldr1 + nldr2 + nldr3 + nldr4 +
+  nldr5 + nldr6 + nldr7 + nldr8 +
   plot_layout(ncol = 4)

### 2024-03-28 paper.qmd
done with first image

@@
 )
+interior_annotation <- function(label) {
+  annotation_custom(grid::textGrob(label = label,
+      x = unit(0.92, "npc"), y = unit(0.92, "npc"),
+      gp = grid::gpar(cex = 1, col="grey70")))
+}
 ```
@@
 ```{r}
+# label: read-pbmc-nldr
+# Read a variety of different NLDR representations of PBMC
+# and plot them on same aspect ratio
+clr_choice <- "#0077A3"
 umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_30_min_dist_0.3.rds")
@@
              y = UMAP2)) +
-  geom_point(alpha=0.1, size=1) +
-  ggtitle("a")
-```
+  geom_point(alpha=0.1, size=1, colour=clr_choice) +
+  interior_annotation("a")
 
-```{r}
 umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_5_min_dist_0.01.rds")
@@
              y = UMAP2))+
-  geom_point(alpha=0.1, size=1) +
-  ggtitle("b")
+  geom_point(alpha=0.1, size=1, colour=clr_choice) +
+  interior_annotation("b")
 
-```
-
-```{r}
 umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_15_min_dist_0.99.rds")
@@
              y = UMAP2))+
-  geom_point(alpha=0.1, size=1) +
-  ggtitle("c") 
-```
+  geom_point(alpha=0.1, size=1, colour=clr_choice) +
+  interior_annotation("c")
 
-```{r}
 tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_5.rds")
@@
              y = tSNE2))+
-  geom_point(alpha=0.1, size=1) +
-  ggtitle("d")  
-```
+  geom_point(alpha=0.1, size=1, colour=clr_choice) +
+  interior_annotation("d")
 
-```{r}
 tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_51.rds")
@@
              y = tSNE2))+
-  geom_point(alpha=0.1, size=1) +
-  ggtitle("e") 
-```
+  geom_point(alpha=0.1, size=1, colour=clr_choice) +
+  interior_annotation("e")
 
-```{r}
 phate_pbmc <- read_rds("data/pbmc3k/pbmc_phate_5.rds")
@@
              y = PHATE2))+
-  geom_point(alpha=0.1, size=1) +
-  ggtitle("f")  
-```
+  geom_point(alpha=0.1, size=1, colour=clr_choice) +
+  interior_annotation("f")
 
-```{r}
 trimap_pbmc <- read_rds("data/pbmc3k/pbmc_trimap_12_4_3.rds")
@@
              y = TriMAP2))+
-  geom_point(alpha=0.1, size=1) +
-  ggtitle("g") 
-```
+  geom_point(alpha=0.1, size=1, colour=clr_choice) +
+  interior_annotation("g")
 
-```{r}
 pacmap_pbmc <- read_rds("data/pbmc3k/pbmc_pacmap_30_random_0.9_5.rds")
@@
              y = PaCMAP2))+
-  geom_point(alpha=0.1, size=1) +
-  ggtitle("h")
+  geom_point(alpha=0.1, size=1, colour=clr_choice) +
+  interior_annotation("h")
 ```
@@
 ```{r}
+#| label: NLDR-variety
 #| echo: false
@@
 Which is the most accurate representation of the structure in high dimensions?"
-#| label: fig-diffnldrpbmc
 #| fig-width: 8
-#| fig-height: 3
+#| fig-height: 4
 #| out-width: 100%
@@
   - Explain where it is expected or not expected to work, eg higher dimensional relationships
+  - Diagnostic app to explore differences in distances
   - What might be useful enhancements

### 2024-03-28 paper.tex
done with first image

@@
 
-\begin{figure}
+\begin{figure}[H]
 
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diffnldrpbmc-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/NLDR-variety-1.pdf}
 
@@
 
-\caption{\label{fig-diffnldrpbmc}Six different NLDR representations of
-the same data. Different techniques and different parameter choices are
-used. Which is the most accurate representation of the structure in high
-dimensions?}
+\caption{Six different NLDR representations of the same data. Different
+techniques and different parameter choices are used. Which is the most
+accurate representation of the structure in high dimensions?}
 
@@
 \item
+  Diagnostic app to explore differences in distances
+\item
   What might be useful enhancements

### 2024-03-29 paper.qmd
Introduction complete

@@
 
-```{=html}
 <!-- 
-  Notes
-* Use American spelling
+Check-list before submission
+* Is it all American spelling
+* Spelling checked generally
+* Code all runs given fresh workspace
+* Code has a readme, explaining how the paper results are reproduced
 -->
-```
 
@@
 
- - What's the problem:
+Non-linear dimension reduction (NLDR) is popular for making a convenient low-dimensional representation of high-dimensional data.
+Recently developed methods include t-distributed stochastic neighbor embedding (tSNE) [@Laurens2008], uniform manifold approximation and projection (UMAP) [@Leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@article03], large-scale dimensionality reduction Using triplets (TriMAP) [@article02], and pairwise controlled manifold approximation (PaCMAP) [@Yingfan2021].
+However, the representation generated can vary dramatically from method to method, and with different choices of parameters or random seeds made using the same method (@fig-NLDR-variety).
+The dilemma for the analyst is then, **which representation to use**.
+The choice might result in different procedures used in the downstream analysis, or different inferential conclusions.
+The research described here provides new visual tools to aid with this decision.
+
+<!-- - What's the problem:
 
@@
   Relevant high-d vis, NLDR history
+-->
 
@@
 ```{r}
-#| label: NLDR-variety
+#| label: fig-NLDR-variety
 #| echo: false
@@
 Different techniques and different parameter choices are used.
+Researchers may have seen any of these in their analysis of this data, depending on their choice of method, or typical parameter choice.
+Would they make different decisions downstream in the analysis depending on which version seen?
 Which is the most accurate representation of the structure in high dimensions?"
@@
 
-## Method
+The paper is organised as follows.
+@sec-background provides a summary of the literature on NLDR, and high-dimensional data visualization methods.
+@sec-method contains the details of the new methodology, including simulated data examples.
+Two applications illustrating the use of the new methodology for bioinformatics and image classification are in @sec-applications.
+Limitations and future directions are provided in @sec-discussion.
+
+## Background {#sec-background}
+
+## Method {#sec-method}
 
@@
 
-## Applications
+## Applications {#sec-applications}
 
@@
 
-## Discussion
+## Discussion {#sec-discussion}
 

### 2024-03-29 paper.tex
Introduction complete

@@
 
-\begin{itemize}
-\tightlist
-\item
-  What's the problem:
-\end{itemize}
+Non-linear dimension reduction (NLDR) is popular for making a convenient
+low-dimensional representation of high-dimensional data. Recently
+developed methods include t-distributed stochastic neighbor embedding
+(tSNE) \citep{Laurens2008}, uniform manifold approximation and
+projection (UMAP) \citep{Leland2018}, potential of heat-diffusion for
+affinity-based trajectory embedding (PHATE) algorithm \citep{article03},
+large-scale dimensionality reduction Using triplets (TriMAP)
+\citep{article02}, and pairwise controlled manifold approximation
+(PaCMAP) \citep{Yingfan2021}. However, the representation generated can
+vary dramatically from method to method, and with different choices of
+parameters or random seeds made using the same method
+(Figure~\ref{fig-NLDR-variety}). The dilemma for the analyst is then,
+\textbf{which representation to use}. The choice might result in
+different procedures used in the downstream analysis, or different
+inferential conclusions. The research described here provides new visual
+tools to aid with this decision.
 
-Non-linear dimension reduction being used to summarise high-dimensional
-data.
+\begin{figure}
 
-\begin{itemize}
-\tightlist
-\item
-  Summary of literature
-\end{itemize}
+\centering{
 
-Relevant high-d vis, NLDR history
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/NLDR-variety-1.pdf}
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-NLDR-variety-1.pdf}
 
@@
 
-\caption{Six different NLDR representations of the same data. Different
-techniques and different parameter choices are used. Which is the most
-accurate representation of the structure in high dimensions?}
+\caption{\label{fig-NLDR-variety}Six different NLDR representations of
+the same data. Different techniques and different parameter choices are
+used. Researchers may have seen any of these in their analysis of this
+data, depending on their choice of method, or typical parameter choice.
+Would they make different decisions downstream in the analysis depending
+on which version seen? Which is the most accurate representation of the
+structure in high dimensions?}
 
@@
 
-\section{Method}\label{method}
+The paper is organised as follows. Section~\ref{sec-background} provides
+a summary of the literature on NLDR, and high-dimensional data
+visualization methods. Section~\ref{sec-method} contains the details of
+the new methodology, including simulated data examples. Two applications
+illustrating the use of the new methodology for bioinformatics and image
+classification are in Section~\ref{sec-applications}. Limitations and
+future directions are provided in Section~\ref{sec-discussion}.
+
+\section{Background}\label{sec-background}
+
+\section{Method}\label{sec-method}
 
@@
 
-\section{Applications}\label{applications}
+\section{Applications}\label{sec-applications}
 
@@
 
-\section{Discussion}\label{discussion}
+\section{Discussion}\label{sec-discussion}
 

### 2024-03-29 paper.qmd
first two paragraphs of background

@@
 * Code has a readme, explaining how the paper results are reproduced
+* Re-write abstract
 -->
@@
 
-Non-linear dimension reduction (NLDR) is popular for making a convenient low-dimensional representation of high-dimensional data.
+Non-linear dimension reduction (NLDR) is popular for making a convenient low-dimensional ($k-D$) representation of high-dimensional ($p-D$) data.
 Recently developed methods include t-distributed stochastic neighbor embedding (tSNE) [@Laurens2008], uniform manifold approximation and projection (UMAP) [@Leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@article03], large-scale dimensionality reduction Using triplets (TriMAP) [@article02], and pairwise controlled manifold approximation (PaCMAP) [@Yingfan2021].
@@
 #| out-width: 100%
-# 2D layouts from different NLDR techniques and different hyperparameter choices applied for the PBMC3k dataset: (a) UMAP (n_neighbors = 30, min_dist = 0.3), (b) UMAP (n_neighbors = 5, min_dist = 0.01), (c) UMAP (n_neighbors = 15, min_dist = 0.99), (d) UMAP (n_neighbors = 15, min_dist = 0.03), (e) UMAP (n_neighbors = 84, min_dist = 0.99), (f) tSNE (perplexity = 5), (g) tSNE (perplexity = 51), (h) tSNE (perplexity = 92), (i) PHATE (knn = 2), (j) PHATE (knn = 5), (k) PHATE (knn = 10), (l) TriMAP (n_inliers = 12, n_outliers = 4, n_random = 3), (m) TriMAP (n_inliers = 30, n_outliers = 4, n_random = 10), (n) TriMAP (n_inliers = 5, n_outliers = 2, n_random = 2), (p) PaCMAP (n_neighbors = 10, init = pca, MN_ratio = 0.5, FP_ratio = 2), (q) PaCMAP (n_neighbors = 30, init = random, MN_ratio = 0.9, FP_ratio = 5), (r) PaCMAP (n_neighbors = 5, init = pca, MN_ratio = 0.1, FP_ratio = 1), and (s) PaCMAP (n_neighbors = 30, init = random, MN_ratio = 0.1, FP_ratio = 1).
-Is there a best representation of the original data or are they all providing  equivalent information?
+# (a) UMAP (n_neighbors = 30, min_dist = 0.3), (b) UMAP (n_neighbors = 5, min_dist = 0.01), (c) UMAP (n_neighbors = 15, min_dist = 0.99), (f) tSNE (perplexity = 5), (g) tSNE (perplexity = 51), (l) TriMAP (n_inliers = 12, n_outliers = 4, n_random = 3), (q) PaCMAP (n_neighbors = 30, init = random, MN_ratio = 0.9, FP_ratio = 5)
 nldr1 + nldr2 + nldr3 + nldr4 +
@@
 
+<!-- - Connection between NLDR and MDS-->
+Historically, $k-D$ representations of $p-D$ data have been computed using multidimensional Scaling (MDS), which includes principal components analysis as a special case.
+The $k-D$ representation can be considered to be a layout of points in $k-D$ produced by an embedding procedure that maps the data from $p-D$.
+In MDS, the $k-D$ layout is constructed by minimizing a stress function that differences distances between points in $p-D$ with potential distances between points in $k-D$.
+Various formulations of the stress function result in non-metric scaling [@article62] and isomap [@article63].
+Challenges in working with high-dimensional data, including visualization, are outlined in @Johnstone2009.
+
+Many new methods for NLDR have emerged in recent years, all designed to better capture specific structures potentially existing in $p-D$.
+Here we focus on five currently popular techniques, tSNE, UMAP, PHATE, TriMAP and PaCMAP.
+tNSE and UMAP can be considered to produce the $k-D$ minimizing the divergence between two distributions, where the distributions are modeling the inter-point distances.
+PHATE, TriMAP and PaCMAP are examples of diffusion processes [@article64] spreading to capture geometric shapes, that include both global and local structure.
+
+- Tours
+- Model-in-the-data-space: how can we represent the model, eg plane for PCA, grid of values for classification boundaries.
+
 ## Method {#sec-method}
 
+- Overview: Generate a form that maps the model, that is the interpoint distances. What is the model?
+- Notation
 - Create a representation of the model 
-  - Algorithm in 2D
-      - Parameters
-      - Tuning
-  - Showing model in high-d
+    - using hex-binning in 2D, 
+    - parameters, 
+    - tuning, 
+    - pre-processing
+- How does this map to the representation in high-d
+    - Centroids, 
+    - Edges
+- Measuring fit
+    - Fitted values
+    - Error calculation
 - What is learned about simulated examples
+    - Interesting organisation of points in UMAP
+    -
 

### 2024-03-29 paper.tex
first two paragraphs of background

@@
 Non-linear dimension reduction (NLDR) is popular for making a convenient
-low-dimensional representation of high-dimensional data. Recently
-developed methods include t-distributed stochastic neighbor embedding
-(tSNE) \citep{Laurens2008}, uniform manifold approximation and
-projection (UMAP) \citep{Leland2018}, potential of heat-diffusion for
-affinity-based trajectory embedding (PHATE) algorithm \citep{article03},
-large-scale dimensionality reduction Using triplets (TriMAP)
-\citep{article02}, and pairwise controlled manifold approximation
-(PaCMAP) \citep{Yingfan2021}. However, the representation generated can
-vary dramatically from method to method, and with different choices of
-parameters or random seeds made using the same method
-(Figure~\ref{fig-NLDR-variety}). The dilemma for the analyst is then,
-\textbf{which representation to use}. The choice might result in
+low-dimensional (\(k-D\)) representation of high-dimensional (\(p-D\))
+data. Recently developed methods include t-distributed stochastic
+neighbor embedding (tSNE) \citep{Laurens2008}, uniform manifold
+approximation and projection (UMAP) \citep{Leland2018}, potential of
+heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm
+\citep{article03}, large-scale dimensionality reduction Using triplets
+(TriMAP) \citep{article02}, and pairwise controlled manifold
+approximation (PaCMAP) \citep{Yingfan2021}. However, the representation
+generated can vary dramatically from method to method, and with
+different choices of parameters or random seeds made using the same
+method (Figure~\ref{fig-NLDR-variety}). The dilemma for the analyst is
+then, \textbf{which representation to use}. The choice might result in
 different procedures used in the downstream analysis, or different
@@
 
+Historically, \(k-D\) representations of \(p-D\) data have been computed
+using multidimensional Scaling (MDS), which includes principal
+components analysis as a special case. The \(k-D\) representation can be
+considered to be a layout of points in \(k-D\) produced by an embedding
+procedure that maps the data from \(p-D\). In MDS, the \(k-D\) layout is
+constructed by minimizing a stress function that differences distances
+between points in \(p-D\) with potential distances between points in
+\(k-D\). Various formulations of the stress function result in
+non-metric scaling \citep{article62} and isomap \citep{article63}.
+Challenges in working with high-dimensional data, including
+visualization, are outlined in \citet{Johnstone2009}.
+
+Many new methods for NLDR have emerged in recent years, all designed to
+better capture specific structures potentially existing in \(p-D\). Here
+we focus on five currently popular techniques, tSNE, UMAP, PHATE, TriMAP
+and PaCMAP. tNSE and UMAP can be considered to produce the \(k-D\)
+minimizing the divergence between two distributions, where the
+distributions are modeling the inter-point distances. PHATE, TriMAP and
+PaCMAP are examples of diffusion processes \citep{article64} spreading
+to capture geometric shapes, that include both global and local
+structure.
+
+\begin{itemize}
+\tightlist
+\item
+  Tours
+\item
+  Model-in-the-data-space: how can we represent the model, eg plane for
+  PCA, grid of values for classification boundaries.
+\end{itemize}
+
 \section{Method}\label{sec-method}
@@
 \item
+  Overview: Generate a form that maps the model, that is the interpoint
+  distances. What is the model?
+\item
+  Notation
+\item
   Create a representation of the model
-\item
-  Algorithm in 2D
 
@@
   \item
-    Parameters
+    using hex-binning in 2D,
   \item
-    Tuning
+    parameters,
+  \item
+    tuning,
+  \item
+    pre-processing
   \end{itemize}
 \item
-  Showing model in high-d
+  How does this map to the representation in high-d
+
+  \begin{itemize}
+  \tightlist
+  \item
+    Centroids,
+  \item
+    Edges
+  \end{itemize}
+\item
+  Measuring fit
+
+  \begin{itemize}
+  \tightlist
+  \item
+    Fitted values
+  \item
+    Error calculation
+  \end{itemize}
 \item
   What is learned about simulated examples
+
+  \begin{itemize}
+  \tightlist
+  \item
+    Interesting organisation of points in UMAP
+  \item
+  \end{itemize}
 \end{itemize}

### 2024-03-29 paper_original.tex
first two paragraphs of background

--- 2024-03-27 paper_original.tex
@@
 
-\subcaption{\label{fig-gau1-sc1}}
+\subcaption{\label{fig-gau1\_sc1}}
 
@@
 
-\subcaption{\label{fig-gau1-sc2}}
+\subcaption{\label{fig-gau1\_sc2}}
 
@@
 
-\subcaption{\label{fig-gau1-sc3}}
-
-\end{minipage}%
-
-\caption{\label{fig-gau1-sc}Screen shots of the \textbf{langevitour} of
+\subcaption{\label{fig-gau1\_sc3}}
+
+\end{minipage}%
+
+\caption{\label{fig-gau1_sc}Screen shots of the \textbf{langevitour} of
 the five Gaussian clusters dataset, shows the model with tSNE in high-D,
@@
 
-\subcaption{\label{fig-gau2-sc1}}
+\subcaption{\label{fig-gau2\_sc1}}
 
@@
 
-\subcaption{\label{fig-gau2-sc2}}
+\subcaption{\label{fig-gau2\_sc2}}
 
@@
 
-\subcaption{\label{fig-gau2-sc3}}
-
-\end{minipage}%
-
-\caption{\label{fig-gau2-sc}Screen shots of the \textbf{langevitour} of
+\subcaption{\label{fig-gau2\_sc3}}
+
+\end{minipage}%
+
+\caption{\label{fig-gau2_sc}Screen shots of the \textbf{langevitour} of
 the five Gaussian clusters dataset, shows the model with UMAP in high-D,
@@
 
-\subcaption{\label{fig-gau3-sc1}}
+\subcaption{\label{fig-gau3\_sc1}}
 
@@
 
-\subcaption{\label{fig-gau3-sc2}}
+\subcaption{\label{fig-gau3\_sc2}}
 
@@
 
-\subcaption{\label{fig-gau3-sc3}}
-
-\end{minipage}%
-
-\caption{\label{fig-gau3-sc}Screen shots of the \textbf{langevitour} of
+\subcaption{\label{fig-gau3\_sc3}}
+
+\end{minipage}%
+
+\caption{\label{fig-gau3_sc}Screen shots of the \textbf{langevitour} of
 the five Gaussian clusters dataset, shows the model with PHATE in
@@
 
-\subcaption{\label{fig-gau4-sc1}}
+\subcaption{\label{fig-gau4\_sc1}}
 
@@
 
-\subcaption{\label{fig-gau4-sc2}}
+\subcaption{\label{fig-gau4\_sc2}}
 
@@
 
-\subcaption{\label{fig-gau4-sc3}}
-
-\end{minipage}%
-
-\caption{\label{fig-gau4-sc}Screen shots of the \textbf{langevitour} of
+\subcaption{\label{fig-gau4\_sc3}}
+
+\end{minipage}%
+
+\caption{\label{fig-gau4_sc}Screen shots of the \textbf{langevitour} of
 the five Gaussian clusters dataset, shows the model with TriMAP in
@@
 
-\subcaption{\label{fig-gau5-sc1}}
+\subcaption{\label{fig-gau5\_sc1}}
 
@@
 
-\subcaption{\label{fig-gau5-sc2}}
+\subcaption{\label{fig-gau5\_sc2}}
 
@@
 
-\subcaption{\label{fig-gau5-sc3}}
-
-\end{minipage}%
-
-\caption{\label{fig-gau5-sc}Screen shots of the \textbf{langevitour} of
+\subcaption{\label{fig-gau5\_sc3}}
+
+\end{minipage}%
+
+\caption{\label{fig-gau5_sc}Screen shots of the \textbf{langevitour} of
 the five Gaussian clusters dataset, shows the model with PaCMAP in
@@
 
-\subcaption{\label{fig-pbmc1-sc1}}
+\subcaption{\label{fig-pbmc1\_sc1}}
 
@@
 
-\subcaption{\label{fig-pbmc1-sc2}}
+\subcaption{\label{fig-pbmc1\_sc2}}
 
@@
 
-\subcaption{\label{fig-pbmc1-sc3}}
-
-\end{minipage}%
-
-\caption{\label{fig-pbmc1-sc}Screen shots of the \textbf{langevitour} of
+\subcaption{\label{fig-pbmc1\_sc3}}
+
+\end{minipage}%
+
+\caption{\label{fig-pbmc1_sc}Screen shots of the \textbf{langevitour} of
 the PBMC3k data set, shows the model in high-D, a video of the tour
@@
 
-\subcaption{\label{fig-pbmc2-sc4}}
+\subcaption{\label{fig-pbmc2\_sc4}}
 
@@
 
-\subcaption{\label{fig-pbmc2-sc5}}
+\subcaption{\label{fig-pbmc2\_sc5}}
 
@@
 
-\subcaption{\label{fig-pbmc2-sc6}}
-
-\end{minipage}%
-
-\caption{\label{fig-pbmc2-sc}Screen shots of the \textbf{langevitour} of
+\subcaption{\label{fig-pbmc2\_sc6}}
+
+\end{minipage}%
+
+\caption{\label{fig-pbmc2_sc}Screen shots of the \textbf{langevitour} of
 the PBMC3k data set, shows the model in high-D, a video of the tour
@@
 
-\subcaption{\label{fig-mnist1-sc1}}
+\subcaption{\label{fig-mnist1\_sc1}}
 
@@
 
-\subcaption{\label{fig-mnist1-sc2}}
+\subcaption{\label{fig-mnist1\_sc2}}
 
@@
 
-\subcaption{\label{fig-mnist1-sc3}}
-
-\end{minipage}%
-
-\caption{\label{fig-mnist1-sc}Screen shots of the \textbf{langevitour}
+\subcaption{\label{fig-mnist1\_sc3}}
+
+\end{minipage}%
+
+\caption{\label{fig-mnist1_sc}Screen shots of the \textbf{langevitour}
 of the MNIST digit 1 data set, shows the model-in-data space, a video of

### 2024-03-29 paper.qmd
still working on background

@@
 <!-- - Connection between NLDR and MDS-->
-Historically, $k-D$ representations of $p-D$ data have been computed using multidimensional Scaling (MDS), which includes principal components analysis as a special case.
+Historically, $k-D$ representations of $p-D$ data have been computed using multidimensional Scaling (MDS), which includes principal components analysis (PCA) as a special case.
 The $k-D$ representation can be considered to be a layout of points in $k-D$ produced by an embedding procedure that maps the data from $p-D$.
@@
 
-- Tours
+The array of layouts in @fig-NLDR-variety illustrate what can emerge from the choices of method and parameters, and the random seed that initiates the computation.
+Key structures interpreted from these views suggest:
+
+- highly **separated clusters** (a, b, e, g, h) with the number ranging from 3-6, 
+- **stringy branches** (f), and 
+- **barely separated clusters** (c, d) which apparently **contradicts** the other representations.
+
+It happens because these methods and parameter choices provide different lenses on the interpoint distances in the data.
+
+The alternative approach to visualizing the high-dimensional data is to use linear projections.
+PCA is the classical approach, resulting in a set of new variables which are linear combinations of the original variables.
+Tours, defined by @Asimov1985, broaden the scope by providing movies of linear projections, that provide views the data from all directions.
+@lee2021review provides an review of the main developments in tours.
+There are many tour algorithms implemented, with many available in the R package `tourr` [@article61], and versions enabling better interactivity in `langevitour` [@article09] and `detourr` [@article22].
+Linear projections are a safe way to view high-dimensional data, because they do not warp the space, so they are more faithful representations of the structure.
+However, linear projections can be cluttered, and global patterns can obscure local structure.
+The simple activity of projecting data from $p-D$ suffers from piling [@laa2022], where data concentrates in the center of projections.
+NLDR is designed to escape these issues, to exaggerate structure so that it can be observed.
+But as a result NLDR can hallucinate wildly, to suggest patterns that are not actually present in the data.
+
+The solution is to use the tour to examine how the NLDR is warping the space.
+
 - Model-in-the-data-space: how can we represent the model, eg plane for PCA, grid of values for classification boundaries.

### 2024-03-29 paper.tex
still working on background

@@
 using multidimensional Scaling (MDS), which includes principal
-components analysis as a special case. The \(k-D\) representation can be
-considered to be a layout of points in \(k-D\) produced by an embedding
-procedure that maps the data from \(p-D\). In MDS, the \(k-D\) layout is
-constructed by minimizing a stress function that differences distances
-between points in \(p-D\) with potential distances between points in
-\(k-D\). Various formulations of the stress function result in
+components analysis (PCA) as a special case. The \(k-D\) representation
+can be considered to be a layout of points in \(k-D\) produced by an
+embedding procedure that maps the data from \(p-D\). In MDS, the \(k-D\)
+layout is constructed by minimizing a stress function that differences
+distances between points in \(p-D\) with potential distances between
+points in \(k-D\). Various formulations of the stress function result in
 non-metric scaling \citep{article62} and isomap \citep{article63}.
@@
 
+The array of layouts in Figure~\ref{fig-NLDR-variety} illustrate what
+can emerge from the choices of method and parameters, and the random
+seed that initiates the computation. Key structures interpreted from
+these views suggest:
+
 \begin{itemize}
@@
 \item
-  Tours
+  highly \textbf{separated clusters} (a, b, e, g, h) with the number
+  ranging from 3-6,
+\item
+  \textbf{stringy branches} (f), and
+\item
+  \textbf{barely separated clusters} (c, d) which apparently
+  \textbf{contradicts} the other representations.
+\end{itemize}
+
+It happens because these methods and parameter choices provide different
+lenses on the interpoint distances in the data.
+
+The alternative approach to visualizing the high-dimensional data is to
+use linear projections. PCA is the classical approach, resulting in a
+set of new variables which are linear combinations of the original
+variables. Tours, defined by \citet{Asimov1985}, broaden the scope by
+providing movies of linear projections, that provide views the data from
+all directions. \citet{lee2021review} provides an review of the main
+developments in tours. There are many tour algorithms implemented, with
+many available in the R package \texttt{tourr} \citep{article61}, and
+versions enabling better interactivity in \texttt{langevitour}
+\citep{article09} and \texttt{detourr} \citep{article22}. Linear
+projections are a safe way to view high-dimensional data, because they
+do not warp the space, so they are more faithful representations of the
+structure. However, linear projections can be cluttered, and global
+patterns can obscure local structure. The simple activity of projecting
+data from \(p-D\) suffers from piling \citep{laa2022}, where data
+concentrates in the center of projections. NLDR is designed to escape
+these issues, to exaggerate structure so that it can be observed. But as
+a result NLDR can hallucinate wildly, to suggest patterns that are not
+actually present in the data.
+
+The solution is to use the tour to examine how the NLDR is warping the
+space.
+
+\begin{itemize}
+\tightlist
 \item

### 2024-03-29 paper.qmd
small pieces in background

@@
 
-The solution is to use the tour to examine how the NLDR is warping the space.
+The solution is to use the tour to examine how the NLDR is warping the space. This approach follows what @article08 describes as *model-in-the-data-space*. The fitted model should be overlaid on the
+
+Linked brushing as done by @article21
 
@@
 ### pbmc
+
 - NLDR view used to illustrate clusters
@@
 ### digits: 1
+
 - NLDR is used to illustrate different ways 1's are drawn

### 2024-03-29 paper.tex
small pieces in background

@@
 The solution is to use the tour to examine how the NLDR is warping the
-space.
+space. This approach follows what \citet{article08} describes as
+\emph{model-in-the-data-space}.
 
@@
 
-\begin{verbatim}
- - NLDR view used to illustrate clusters
- - Use our method to assess is it a reasonable representation
- - Demonstrate that it is not
- - Illustrate how to use out method to get a better representation
-
-\end{verbatim}
+\begin{itemize}
+\tightlist
+\item
+  NLDR view used to illustrate clusters
+\item
+  Use our method to assess is it a reasonable representation
+\item
+  Demonstrate that it is not
+\item
+  Illustrate how to use out method to get a better representation
+\end{itemize}
 
@@
 
-\begin{verbatim}
- - NLDR is used to illustrate different ways 1's are drawn
- - Use our method to assess is it a reasonable representation
- - Demonstrate that it is, except for the anomalies
-
-\end{verbatim}
+\begin{itemize}
+\tightlist
+\item
+  NLDR is used to illustrate different ways 1's are drawn
+\item
+  Use our method to assess is it a reasonable representation
+\item
+  Demonstrate that it is, except for the anomalies
+\end{itemize}
 

### 2024-03-29 paper.qmd
done with background

@@
 
-The solution is to use the tour to examine how the NLDR is warping the space. This approach follows what @article08 describes as *model-in-the-data-space*. The fitted model should be overlaid on the
+The solution is to use the tour to examine how the NLDR is warping the space.
+This approach follows what @article08 describes as *model-in-the-data-space*.
+The fitted model should be overlaid on the data, to examine the fit relative the spread of the observations.
+While this is straightforward, and commonly done when data is 2D, it is also possible in $p-D$, for many models, when a tour is used.
 
-Linked brushing as done by @article21
+@article08 provides several examples of models overlaid on the data in $p-D$.
+In hierarchical clustering, a representation of the dendrogrom using points and lines can be constructed by augmenting the data with points marking merging of clusters.
+Showing the movie of linear projections reveals shows how the algorithm sequentially fitted the cluster model to the data.
+For linear discriminant analysis or model-based clustering the model can be indicated by $(p-1)-D$ ellipses.
+It is possible to see whether the elliptical shapes appropriately matches the variance of the relevant clusters, and to compare and contrast different fits.
+For PCA, one can display the $k-D$ plane of the reduced dimension using wireframes of transformed cubes.
+Using a wireframe is the approach we take here, to represent the NLDR model in $p-D$.
 
-- Model-in-the-data-space: how can we represent the model, eg plane for PCA, grid of values for classification boundaries.
+<!-- Linked brushing as done by @article21
+
+- Model-in-the-data-space: how can we represent the model, eg plane for PCA, grid of values for classification boundaries, ellipses for LDA and mclust, nets for SOM.-->
 

### 2024-03-29 paper.tex
done with background

@@
 space. This approach follows what \citet{article08} describes as
-\emph{model-in-the-data-space}.
+\emph{model-in-the-data-space}. The fitted model should be overlaid on
+the data, to examine the fit relative the spread of the observations.
+While this is straightforward, and commonly done when data is 2D, it is
+also possible in \(p-D\), for many models, when a tour is used.
 
-\begin{itemize}
-\tightlist
-\item
-  Model-in-the-data-space: how can we represent the model, eg plane for
-  PCA, grid of values for classification boundaries.
-\end{itemize}
+\citet{article08} provides several examples of models overlaid on the
+data in \(p-D\). In hierarchical clustering, a representation of the
+dendrogrom using points and lines can be constructed by augmenting the
+data with points marking merging of clusters. Showing the movie of
+linear projections reveals shows how the algorithm sequentially fitted
+the cluster model to the data. For linear discriminant analysis or
+model-based clustering the model can be indicated by \((p-1)-D\)
+ellipses. It is possible to see whether the elliptical shapes
+appropriately matches the variance of the relevant clusters, and to
+compare and contrast different fits. For PCA, one can display the
+\(k-D\) plane of the reduced dimension using wireframes of transformed
+cubes. Using a wireframe is the approach we take here, to represent the
+NLDR model in \(p-D\).
 

### 2024-03-29 paper.qmd
mapping oout remaining parts

@@
 
-keywords: [high-dimensional data, dimension reduction, triangulation, hexagonal binning, low-dimensional manifold, manifold learning, tour, data vizualization]
-keywords-formatted: [high-dimensional data, dimension reduction, triangulation, hexagonal binning, low-dimensional manifold, manifold learning, tour, data vizualization]
+keywords: [high-dimensional data, dimension reduction, hexagonal binning, low-dimensional manifold, tour, data vizualization, model in the data space]
+keywords-formatted: [high-dimensional data, dimension reduction, hexagonal binning, low-dimensional manifold, tour, data vizualization, model in the data space]
 
@@
 
+### What is the NLDR model?
+
+### Notation
+
+### Constructing the 2D model
+
+### Displaying the model in $p-D$
+
+### Measuring the fit
+
+### What can be learned
+
 - Overview: Generate a form that maps the model, that is the interpoint distances. What is the model?
@@
 - Explain where it is expected or not expected to work, eg higher dimensional relationships
+- Human behaviour, the desire to have more certainty, and a tendency to prefer the well-separated views
 - Diagnostic app to explore differences in distances

### 2024-03-29 paper.tex
mapping oout remaining parts

@@
   pdfauthor={Jayani P.G. Lakshika; Dianne Cook; Paul Harrison; Michael Lydeamore; Thiyanga S. Talagala},
-  pdfkeywords={high-dimensional data, dimension
-reduction, triangulation, hexagonal binning, low-dimensional
-manifold, manifold learning, tour, data vizualization},
+  pdfkeywords={high-dimensional data, dimension reduction, hexagonal
+binning, low-dimensional manifold, tour, data vizualization, model in
+the data space},
   colorlinks=true,
@@
 \noindent%
-{\it Keywords:} high-dimensional data, dimension
-reduction, triangulation, hexagonal binning, low-dimensional
-manifold, manifold learning, tour, data vizualization
+{\it Keywords:} high-dimensional data, dimension reduction, hexagonal
+binning, low-dimensional manifold, tour, data vizualization, model in
+the data space
 \vfill
@@
 
+\subsection{What is the NLDR model?}\label{what-is-the-nldr-model}
+
+\subsection{Notation}\label{notation}
+
+\subsection{Constructing the 2D model}\label{constructing-the-2d-model}
+
+\subsection{\texorpdfstring{Displaying the model in
+\(p-D\)}{Displaying the model in p-D}}\label{displaying-the-model-in-p-d}
+
+\subsection{Measuring the fit}\label{measuring-the-fit}
+
+\subsection{What can be learned}\label{what-can-be-learned}
+
 \begin{itemize}
@@
 \item
+  Human behaviour, the desire to have more certainty, and a tendency to
+  prefer the well-separated views
+\item
   Diagnostic app to explore differences in distances

### 2024-03-29 paper.qmd
Ref to sec2 para:1

@@
 <!-- - Connection between NLDR and MDS-->
-Historically, $k-D$ representations of $p-D$ data have been computed using multidimensional Scaling (MDS), which includes principal components analysis (PCA) as a special case.
+Historically, $k-D$ representations of $p-D$ data have been computed using multidimensional Scaling (MDS) [@borg2005], which includes principal components analysis (PCA) as a special case.
 The $k-D$ representation can be considered to be a layout of points in $k-D$ produced by an embedding procedure that maps the data from $p-D$.
@@
 Various formulations of the stress function result in non-metric scaling [@article62] and isomap [@article63].
-Challenges in working with high-dimensional data, including visualization, are outlined in @Johnstone2009.
+Challenges in working with high-dimensional data, including visualization, are outlined in @johnstone2009.
 

### 2024-03-29 paper.qmd
chnaged the ref names sec2 para:1
Added PCA latest ref

@@
 <!-- - Connection between NLDR and MDS-->
-Historically, $k-D$ representations of $p-D$ data have been computed using multidimensional Scaling (MDS) [@borg2005], which includes principal components analysis (PCA) as a special case.
+Historically, $k-D$ representations of $p-D$ data have been computed using multidimensional Scaling (MDS) [@borg2005], which includes principal components analysis (PCA) [@jolliffe2011] as a special case.
 The $k-D$ representation can be considered to be a layout of points in $k-D$ produced by an embedding procedure that maps the data from $p-D$.
 In MDS, the $k-D$ layout is constructed by minimizing a stress function that differences distances between points in $p-D$ with potential distances between points in $k-D$.
-Various formulations of the stress function result in non-metric scaling [@article62] and isomap [@article63].
+Various formulations of the stress function result in non-metric scaling [@kruskal1964] and isomap [@silva2002].
 Challenges in working with high-dimensional data, including visualization, are outlined in @johnstone2009.

### 2024-03-29 paper.qmd
done with first part of section 3

--- 2024-03-29 paper.qmd
@@
 <!-- - Connection between NLDR and MDS-->
-Historically, $k-D$ representations of $p-D$ data have been computed using multidimensional Scaling (MDS), which includes principal components analysis (PCA) as a special case.
+Historically, $k-D$ representations of $p-D$ data have been computed using multidimensional scaling (MDS), which includes principal components analysis (PCA) as a special case.
 The $k-D$ representation can be considered to be a layout of points in $k-D$ produced by an embedding procedure that maps the data from $p-D$.
@@
 
+At first glance, thinking of of NLDR as a model fitted to the data might seem strange.
+It is a model in the sense that it is a "a simplified representation or abstraction of a system, process, or phenomenon in the real world" where the $p-D$ observations are the realization.
+From a statistical perspective we can consider the distances between points in the $k-D$ layout to be variance that the model explains, and the (relative) difference with their distances in $p-D$ is the error, or unexplained variance.
+Abstractly, we can also imagine that the positioning of points in 2D represents fitted values, that will have some prescribed position in $p-D$ that can be compared with the observed value.
+
 ### Notation
+
+<< XXX Jayani, please insert a table of notation >>
+
+Once we have notation, we need to have some math that gives precision to the paragraph above. (XXX Michael??)
 

### 2024-03-29 paper.tex
done with first part of section 3

@@
 Historically, \(k-D\) representations of \(p-D\) data have been computed
-using multidimensional Scaling (MDS), which includes principal
+using multidimensional scaling (MDS), which includes principal
 components analysis (PCA) as a special case. The \(k-D\) representation
@@
 
+At first glance, thinking of of NLDR as a model fitted to the data might
+seem strange. It is a model in the sense that it is a ``a simplified
+representation or abstraction of a system, process, or phenomenon in the
+real world'' where the \(p-D\) observations are the realization. From a
+statistical perspective we can consider the distances between points in
+the \(k-D\) layout to be variance that the model explains, and the
+(relative) difference with their distances in \(p-D\) is the error, or
+unexplained variance. Abstractly, we can also imagine that the
+positioning of points in 2D represents fitted values, that will have
+some prescribed position in \(p-D\) that can be compared with the
+observed value.
+
 \subsection{Notation}\label{notation}
+
+\textless\textless{} XXX Jayani, please insert a table of notation
+\textgreater\textgreater{}
+
+Once we have notation, we need to have some math that gives precision to
+the paragraph above. (XXX Michael??)
 

### 2024-03-29 paper.tex

@@
 Historically, \(k-D\) representations of \(p-D\) data have been computed
-using multidimensional scaling (MDS), which includes principal
-components analysis (PCA) as a special case. The \(k-D\) representation
-can be considered to be a layout of points in \(k-D\) produced by an
-embedding procedure that maps the data from \(p-D\). In MDS, the \(k-D\)
-layout is constructed by minimizing a stress function that differences
-distances between points in \(p-D\) with potential distances between
-points in \(k-D\). Various formulations of the stress function result in
-non-metric scaling \citep{article62} and isomap \citep{article63}.
-Challenges in working with high-dimensional data, including
-visualization, are outlined in \citet{Johnstone2009}.
+using multidimensional Scaling (MDS) \citep{borg2005}, which includes
+principal components analysis (PCA) \citep{jolliffe2011} as a special
+case. The \(k-D\) representation can be considered to be a layout of
+points in \(k-D\) produced by an embedding procedure that maps the data
+from \(p-D\). In MDS, the \(k-D\) layout is constructed by minimizing a
+stress function that differences distances between points in \(p-D\)
+with potential distances between points in \(k-D\). Various formulations
+of the stress function result in non-metric scaling \citep{kruskal1964}
+and isomap \citep{silva2002}. Challenges in working with
+high-dimensional data, including visualization, are outlined in
+\citet{johnstone2009}.
 

### 2024-03-29 paper.qmd
tweaking the Background paragraph

@@
 <!-- - Connection between NLDR and MDS-->
-Historically, $k-D$ representations of $p-D$ data have been computed using multidimensional Scaling (MDS) [@borg2005], which includes principal components analysis (PCA) [@jolliffe2011] as a special case.
+Historically, $k-D$ representations of $p-D$ data have been computed using multidimensional scaling (MDS) [@borg2005], which includes principal components analysis (PCA) [@jolliffe2011] as a special case.
 The $k-D$ representation can be considered to be a layout of points in $k-D$ produced by an embedding procedure that maps the data from $p-D$.
@@
 At first glance, thinking of of NLDR as a model fitted to the data might seem strange.
-It is a model in the sense that it is a "a simplified representation or abstraction of a system, process, or phenomenon in the real world" where the $p-D$ observations are the realization.
+It is a model in the sense that it is a "a simplified representation or abstraction of a system, process, or phenomenon in the real world".
+The $p-D$ observations are the realization of the phenomenon, and the $k-D$ NLDR layout is the simplified representation.
 From a statistical perspective we can consider the distances between points in the $k-D$ layout to be variance that the model explains, and the (relative) difference with their distances in $p-D$ is the error, or unexplained variance.

### 2024-03-29 paper.tex
tweaking the Background paragraph

@@
 Historically, \(k-D\) representations of \(p-D\) data have been computed
-using multidimensional Scaling (MDS) \citep{borg2005}, which includes
+using multidimensional scaling (MDS) \citep{borg2005}, which includes
 principal components analysis (PCA) \citep{jolliffe2011} as a special
@@
 representation or abstraction of a system, process, or phenomenon in the
-real world'' where the \(p-D\) observations are the realization. From a
-statistical perspective we can consider the distances between points in
-the \(k-D\) layout to be variance that the model explains, and the
-(relative) difference with their distances in \(p-D\) is the error, or
-unexplained variance. Abstractly, we can also imagine that the
-positioning of points in 2D represents fitted values, that will have
-some prescribed position in \(p-D\) that can be compared with the
-observed value.
+real world''. The \(p-D\) observations are the realization of the
+phenomenon, and the \(k-D\) NLDR layout is the simplified
+representation. From a statistical perspective we can consider the
+distances between points in the \(k-D\) layout to be variance that the
+model explains, and the (relative) difference with their distances in
+\(p-D\) is the error, or unexplained variance. Abstractly, we can also
+imagine that the positioning of points in 2D represents fitted values,
+that will have some prescribed position in \(p-D\) that can be compared
+with the observed value.
 

### 2024-03-29 paper.qmd
Added latest citation for non-metric MDS
Initial notation table set up

@@
 In MDS, the $k-D$ layout is constructed by minimizing a stress function that differences distances between points in $p-D$ with potential distances between points in $k-D$.
-Various formulations of the stress function result in non-metric scaling [@kruskal1964] and isomap [@silva2002].
+Various formulations of the stress function result in non-metric scaling [@saeed2018] and isomap [@silva2002].
 Challenges in working with high-dimensional data, including visualization, are outlined in @johnstone2009.
@@
 ### Notation
+
+```{r}
+# Notations used in the paper
+
+notation_df <- read_csv("data/paper_notations.csv")
+
+# Create the table
+kable(notation_df, format = "latex", booktabs = TRUE, escape = FALSE, 
+      caption = "Notations used in this paper.", label = "tbl-notations") |>
+  kable_styling(full_width = FALSE) |>
+  row_spec(0, bold = TRUE)
+```
 

### 2024-03-29 paper.tex
Corrected ref error
Initial notation table set up

@@
 with potential distances between points in \(k-D\). Various formulations
-of the stress function result in non-metric scaling \citep{kruskal1964}
+of the stress function result in non-metric scaling \citep{saeed2018}
 and isomap \citep{silva2002}. Challenges in working with
@@
 \subsection{Notation}\label{notation}
+
+\begin{table}
+
+\caption{\label{tab:tbl-notations}Notations used in this paper.}
+\centering
+\begin{tabular}[t]{ll}
+\toprule
+\textbf{Notation} & \textbf{Description}\\
+\midrule
+$n$ & number of observations\\
+$p$ & number of dimensions in high-D data\\
+$X_{n \times p} = \begin{bmatrix} \textbf{x} _{1} & \textbf{x}_ {2} & \cdots & \textbf{x}_{n} \\  \end{bmatrix}^\top$ & high-D data matrix\\
+$d$ & number of dimensions in embedding data, usually two\\
+$Y_{n \times d} = \begin{bmatrix} \textbf{y} _{1} & \textbf{y}_ {2} & \cdots & \textbf{y}_{n} \\  \end{bmatrix}^\top$ & embedding matrix\\
+\addlinespace
+$P_t$ & point along the geodesic path at time $t$\\
+$P_{t+1}$ & point along the geodesic path at the next time instant, which is $t+1$\\
+$b_1$ & number of bins along the x-axis\\
+$b_2$ & number of bins along the y-axis\\
+$b$ & total number of bins in the hexagonal grid\\
+\addlinespace
+$b'$ & total number of bins without the empty bins\\
+$C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2})$ & hexagonal bin centroid coordinates\\
+$C_k^{(p)} \equiv (C_{kx_1}, ..., C_{kx_p})$ & averaged high-D coordinates\\
+\bottomrule
+\end{tabular}
+\end{table}
 

### 2024-03-29 paper.tex
Added notation table

@@
 $P_{t+1}$ & point along the geodesic path at the next time instant, which is $t+1$\\
-$b_1$ & number of bins along the x-axis\\
-$b_2$ & number of bins along the y-axis\\
-$b$ & total number of bins in the hexagonal grid\\
+$y_max$ & maximum value of the scaled second embedding component\\
+$r_1$ & range of the first embedding component\\
+$r_2$ & range of the second embedding component\\
 \addlinespace
-$b'$ & total number of bins without the empty bins\\
-$C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2})$ & hexagonal bin centroid coordinates\\
-$C_k^{(p)} \equiv (C_{kx_1}, ..., C_{kx_p})$ & averaged high-D coordinates\\
+$h_b$ & height of the hexagon\\
+$w_b$ & width of the hexagon\\
+$a_r$ & aspect ratio of the 2D layout before scaling\\
+$h_r$ & hexagon ratio\\
+$s$ & hexagonal size (radius of the outer circle surrounding the hexagon)\\
+\addlinespace
+$(x_s, y_s)$ & starting coordinates of the hexagonal grid\\
+$q_x$ & buffer amount along the x-axis\\
+$q_y$ & buffer amount along the y-axis\\
+$h_s$ & horizontal distance between 
+adjacent hexagon bin \vphantom{1} centroids\\
+$v_s$ & vertical distance between 
+adjacent hexagon bin centroids\\
+\addlinespace
+$d_x$ & horizontal spacing\\
+$d_y$ & adjusted vertical spacing for hexagonal coordinates\\
+$b_x$ & number of hexagon bins along the x-axis\\
+$b_y$ & number of hexagon bins along the y-axis\\
+$b$ & total number of hexagon bins in the grid\\
+\addlinespace
+$b'$ & number of hexagon bins without the empty bins\\
+$n_k$ & number of observations within $k^{th}$ hexagon\\
+$(h^k_{x_i}, h^k_{y_i})$ & hexagonal grid coordinates of $k^{th}$ hexagon\\
+$C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2})$ & hexagonal bin centroid coordinates of $k^{th}$ hexagon\\
+$C_k^{(p)} \equiv (C_{kx_1}, ..., C_{kx_p})$ & high-D mappings of 2D hexagonal bin centroids of $k^{th}$ hexagon\\
 \bottomrule

### 2024-03-29 paper.qmd
Correct the reference not rendering in Background section

@@
 @lee2021review provides an review of the main developments in tours.
-There are many tour algorithms implemented, with many available in the R package `tourr` [@article61], and versions enabling better interactivity in `langevitour` [@article09] and `detourr` [@article22].
+There are many tour algorithms implemented, with many available in the R package `tourr` [@wickham2011], and versions enabling better interactivity in `langevitour` [@harisson2024] and `detourr` [@hart2022].
 Linear projections are a safe way to view high-dimensional data, because they do not warp the space, so they are more faithful representations of the structure.
@@
 The solution is to use the tour to examine how the NLDR is warping the space.
-This approach follows what @article08 describes as *model-in-the-data-space*.
+This approach follows what @wickham2015 describes as *model-in-the-data-space*.
 The fitted model should be overlaid on the data, to examine the fit relative the spread of the observations.
@@
 
-@article08 provides several examples of models overlaid on the data in $p-D$.
+@wickham2015 provides several examples of models overlaid on the data in $p-D$.
 In hierarchical clustering, a representation of the dendrogrom using points and lines can be constructed by augmenting the data with points marking merging of clusters.

### 2024-03-29 paper.tex
Correct the reference not rendering in Background section

@@
 developments in tours. There are many tour algorithms implemented, with
-many available in the R package \texttt{tourr} \citep{article61}, and
+many available in the R package \texttt{tourr} \citep{wickham2011}, and
 versions enabling better interactivity in \texttt{langevitour}
-\citep{article09} and \texttt{detourr} \citep{article22}. Linear
+\citep{harisson2024} and \texttt{detourr} \citep{hart2022}. Linear
 projections are a safe way to view high-dimensional data, because they
@@
 The solution is to use the tour to examine how the NLDR is warping the
-space. This approach follows what \citet{article08} describes as
+space. This approach follows what \citet{wickham2015} describes as
 \emph{model-in-the-data-space}. The fitted model should be overlaid on
@@
 
-\citet{article08} provides several examples of models overlaid on the
+\citet{wickham2015} provides several examples of models overlaid on the
 data in \(p-D\). In hierarchical clustering, a representation of the

### 2024-03-29 paper.qmd
Correct reference rendering issue in Introduction section

@@
 Non-linear dimension reduction (NLDR) is popular for making a convenient low-dimensional ($k-D$) representation of high-dimensional ($p-D$) data.
-Recently developed methods include t-distributed stochastic neighbor embedding (tSNE) [@Laurens2008], uniform manifold approximation and projection (UMAP) [@Leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@article03], large-scale dimensionality reduction Using triplets (TriMAP) [@article02], and pairwise controlled manifold approximation (PaCMAP) [@Yingfan2021].
+Recently developed methods include t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2022], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021].
 However, the representation generated can vary dramatically from method to method, and with different choices of parameters or random seeds made using the same method (@fig-NLDR-variety).
@@
 tNSE and UMAP can be considered to produce the $k-D$ minimizing the divergence between two distributions, where the distributions are modeling the inter-point distances.
-PHATE, TriMAP and PaCMAP are examples of diffusion processes [@article64] spreading to capture geometric shapes, that include both global and local structure.
+PHATE, TriMAP and PaCMAP are examples of diffusion processes [@coifman2005] spreading to capture geometric shapes, that include both global and local structure.
 
@@
 PCA is the classical approach, resulting in a set of new variables which are linear combinations of the original variables.
-Tours, defined by @Asimov1985, broaden the scope by providing movies of linear projections, that provide views the data from all directions.
-@lee2021review provides an review of the main developments in tours.
+Tours, defined by @asimov1985, broaden the scope by providing movies of linear projections, that provide views the data from all directions.
+@lee2021 provides an review of the main developments in tours.
 There are many tour algorithms implemented, with many available in the R package `tourr` [@wickham2011], and versions enabling better interactivity in `langevitour` [@harisson2024] and `detourr` [@hart2022].

### 2024-03-29 paper.tex
Correct reference rendering issue in Introduction section

@@
 data. Recently developed methods include t-distributed stochastic
-neighbor embedding (tSNE) \citep{Laurens2008}, uniform manifold
-approximation and projection (UMAP) \citep{Leland2018}, potential of
+neighbor embedding (tSNE) \citep{laurens2008}, uniform manifold
+approximation and projection (UMAP) \citep{leland2018}, potential of
 heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm
-\citep{article03}, large-scale dimensionality reduction Using triplets
-(TriMAP) \citep{article02}, and pairwise controlled manifold
-approximation (PaCMAP) \citep{Yingfan2021}. However, the representation
+\citep{moon2019}, large-scale dimensionality reduction Using triplets
+(TriMAP) \citep{amid2022}, and pairwise controlled manifold
+approximation (PaCMAP) \citep{yingfan2021}. However, the representation
 generated can vary dramatically from method to method, and with
@@
 distributions are modeling the inter-point distances. PHATE, TriMAP and
-PaCMAP are examples of diffusion processes \citep{article64} spreading
+PaCMAP are examples of diffusion processes \citep{coifman2005} spreading
 to capture geometric shapes, that include both global and local
@@
 set of new variables which are linear combinations of the original
-variables. Tours, defined by \citet{Asimov1985}, broaden the scope by
+variables. Tours, defined by \citet{asimov1985}, broaden the scope by
 providing movies of linear projections, that provide views the data from
-all directions. \citet{lee2021review} provides an review of the main
+all directions. \citet{lee2021} provides an review of the main
 developments in tours. There are many tour algorithms implemented, with

### 2024-03-29 paper.tex
Fix rendering issues

@@
 $n$ & number of observations\\
-$p$ & number of dimensions in high-D data\\
-$X_{n \times p} = \begin{bmatrix} \textbf{x} _{1} & \textbf{x}_ {2} & \cdots & \textbf{x}_{n} \\  \end{bmatrix}^\top$ & high-D data matrix\\
+$p$ & number of dimensions in $p-D$ data\\
+$X_{n \times p} = \begin{bmatrix} \textbf{x} _{1} & \textbf{x}_ {2} & \cdots & \textbf{x}_{n} \\  \end{bmatrix}^\top$ & $p-D$ data matrix\\
 $d$ & number of dimensions in embedding data, usually two\\
@@
 $C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2})$ & hexagonal bin centroid coordinates of $k^{th}$ hexagon\\
-$C_k^{(p)} \equiv (C_{kx_1}, ..., C_{kx_p})$ & high-D mappings of 2D hexagonal bin centroids of $k^{th}$ hexagon\\
+$C_k^{(p)} \equiv (C_{kx_1}, ..., C_{kx_p})$ & $p-D$ mappings of 2D hexagonal bin centroids of $k^{th}$ hexagon\\
 \bottomrule

### 2024-03-30 paper.qmd
Added Lee's review paper
newcommand to make nicer p-D and k-D

@@
 bibliography: bibliography.bib  
+header-includes: | 
+  \usepackage{amsmath}
+  \usepackage{float}
+  \usepackage{hyperref}
+  \usepackage[utf8]{inputenc}
+  \def\tightlist{}
+  \usepackage{setspace}
+  \newcommand\pD{$p\text{-}D$}
+  \newcommand\kD{$k\text{-}D$}
 ---
@@
 
-Non-linear dimension reduction (NLDR) is popular for making a convenient low-dimensional ($k-D$) representation of high-dimensional ($p-D$) data.
+Non-linear dimension reduction (NLDR) is popular for making a convenient low-dimensional (\kD) representation of high-dimensional (\pD) data.
 Recently developed methods include t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2022], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021].
@@
 <!-- - Connection between NLDR and MDS-->
-Historically, $k-D$ representations of $p-D$ data have been computed using multidimensional scaling (MDS) [@borg2005], which includes principal components analysis (PCA) [@jolliffe2011] as a special case.
-The $k-D$ representation can be considered to be a layout of points in $k-D$ produced by an embedding procedure that maps the data from $p-D$.
-In MDS, the $k-D$ layout is constructed by minimizing a stress function that differences distances between points in $p-D$ with potential distances between points in $k-D$.
+Historically, \kD representations of \pD data have been computed using multidimensional scaling (MDS) [@borg2005], which includes principal components analysis (PCA) [@jolliffe2011] as a special case.
+The \kD representation can be considered to be a layout of points in \kD produced by an embedding procedure that maps the data from \pD.
+In MDS, the \kD layout is constructed by minimizing a stress function that differences distances between points in \pD with potential distances between points in \kD.
 Various formulations of the stress function result in non-metric scaling [@saeed2018] and isomap [@silva2002].
@@
 
-Many new methods for NLDR have emerged in recent years, all designed to better capture specific structures potentially existing in $p-D$.
+Many new methods for NLDR have emerged in recent years, all designed to better capture specific structures potentially existing in \pD.
 Here we focus on five currently popular techniques, tSNE, UMAP, PHATE, TriMAP and PaCMAP.
@@
 PCA is the classical approach, resulting in a set of new variables which are linear combinations of the original variables.
-Tours, defined by @asimov1985, broaden the scope by providing movies of linear projections, that provide views the data from all directions.
+Tours, defined by @lee2021, broaden the scope by providing movies of linear projections, that provide views the data from all directions.
 @lee2021 provides an review of the main developments in tours.
@@
 However, linear projections can be cluttered, and global patterns can obscure local structure.
-The simple activity of projecting data from $p-D$ suffers from piling [@laa2022], where data concentrates in the center of projections.
+The simple activity of projecting data from \pD suffers from piling [@laa2022], where data concentrates in the center of projections.
 NLDR is designed to escape these issues, to exaggerate structure so that it can be observed.
@@
 The fitted model should be overlaid on the data, to examine the fit relative the spread of the observations.
-While this is straightforward, and commonly done when data is 2D, it is also possible in $p-D$, for many models, when a tour is used.
+While this is straightforward, and commonly done when data is 2D, it is also possible in \pD, for many models, when a tour is used.
 
-@wickham2015 provides several examples of models overlaid on the data in $p-D$.
+@wickham2015 provides several examples of models overlaid on the data in \pD.
 In hierarchical clustering, a representation of the dendrogrom using points and lines can be constructed by augmenting the data with points marking merging of clusters.
 Showing the movie of linear projections reveals shows how the algorithm sequentially fitted the cluster model to the data.
-For linear discriminant analysis or model-based clustering the model can be indicated by $(p-1)-D$ ellipses.
+For linear discriminant analysis or model-based clustering the model can be indicated by $(p-1)\text{-}D$ ellipses.
 It is possible to see whether the elliptical shapes appropriately matches the variance of the relevant clusters, and to compare and contrast different fits.
-For PCA, one can display the $k-D$ plane of the reduced dimension using wireframes of transformed cubes.
-Using a wireframe is the approach we take here, to represent the NLDR model in $p-D$.
+For PCA, one can display the \kD plane of the reduced dimension using wireframes of transformed cubes.
+Using a wireframe is the approach we take here, to represent the NLDR model in \pD.
 
@@
 It is a model in the sense that it is a "a simplified representation or abstraction of a system, process, or phenomenon in the real world".
-The $p-D$ observations are the realization of the phenomenon, and the $k-D$ NLDR layout is the simplified representation.
-From a statistical perspective we can consider the distances between points in the $k-D$ layout to be variance that the model explains, and the (relative) difference with their distances in $p-D$ is the error, or unexplained variance.
-Abstractly, we can also imagine that the positioning of points in 2D represents fitted values, that will have some prescribed position in $p-D$ that can be compared with the observed value.
+The \pD observations are the realization of the phenomenon, and the \kD NLDR layout is the simplified representation.
+From a statistical perspective we can consider the distances between points in the \kD layout to be variance that the model explains, and the (relative) difference with their distances in \pD is the error, or unexplained variance.
+Abstractly, we can also imagine that the positioning of points in 2D represents fitted values, that will have some prescribed position in \pD that can be compared with the observed value.
 
@@
 
-### Displaying the model in $p-D$
+### Displaying the model in \pD
 

### 2024-03-30 paper.tex
Added Lee's review paper
newcommand to make nicer p-D and k-D

@@
 \addtolength{\topmargin}{-1in}%
-
 \usepackage{booktabs}
@@
 \usepackage{xcolor}
+\usepackage{amsmath}
+\usepackage{float}
+\usepackage{hyperref}
+\usepackage[utf8]{inputenc}
+\def\tightlist{}
+\usepackage{setspace}
+\newcommand\pD{$p\text{-}D$}
+\newcommand\kD{$k\text{-}D$}
 \makeatletter
@@
 \newpage
-\spacingset{1} % DON'T change the spacing! (Default 1.9)
+\spacingset{1.9} % DON'T change the spacing!
 
@@
 Non-linear dimension reduction (NLDR) is popular for making a convenient
-low-dimensional (\(k-D\)) representation of high-dimensional (\(p-D\))
-data. Recently developed methods include t-distributed stochastic
-neighbor embedding (tSNE) \citep{laurens2008}, uniform manifold
-approximation and projection (UMAP) \citep{leland2018}, potential of
-heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm
-\citep{moon2019}, large-scale dimensionality reduction Using triplets
-(TriMAP) \citep{amid2022}, and pairwise controlled manifold
-approximation (PaCMAP) \citep{yingfan2021}. However, the representation
-generated can vary dramatically from method to method, and with
-different choices of parameters or random seeds made using the same
-method (Figure~\ref{fig-NLDR-variety}). The dilemma for the analyst is
-then, \textbf{which representation to use}. The choice might result in
+low-dimensional (\kD) representation of high-dimensional (\pD) data.
+Recently developed methods include t-distributed stochastic neighbor
+embedding (tSNE) \citep{laurens2008}, uniform manifold approximation and
+projection (UMAP) \citep{leland2018}, potential of heat-diffusion for
+affinity-based trajectory embedding (PHATE) algorithm \citep{moon2019},
+large-scale dimensionality reduction Using triplets (TriMAP)
+\citep{amid2022}, and pairwise controlled manifold approximation
+(PaCMAP) \citep{yingfan2021}. However, the representation generated can
+vary dramatically from method to method, and with different choices of
+parameters or random seeds made using the same method
+(Figure~\ref{fig-NLDR-variety}). The dilemma for the analyst is then,
+\textbf{which representation to use}. The choice might result in
 different procedures used in the downstream analysis, or different
@@
 
-Historically, \(k-D\) representations of \(p-D\) data have been computed
-using multidimensional scaling (MDS) \citep{borg2005}, which includes
+Historically, \kD representations of \pD data have been computed using
+multidimensional scaling (MDS) \citep{borg2005}, which includes
 principal components analysis (PCA) \citep{jolliffe2011} as a special
-case. The \(k-D\) representation can be considered to be a layout of
-points in \(k-D\) produced by an embedding procedure that maps the data
-from \(p-D\). In MDS, the \(k-D\) layout is constructed by minimizing a
-stress function that differences distances between points in \(p-D\)
-with potential distances between points in \(k-D\). Various formulations
-of the stress function result in non-metric scaling \citep{saeed2018}
-and isomap \citep{silva2002}. Challenges in working with
-high-dimensional data, including visualization, are outlined in
-\citet{johnstone2009}.
+case. The \kD representation can be considered to be a layout of points
+in \kD produced by an embedding procedure that maps the data from \pD.
+In MDS, the \kD layout is constructed by minimizing a stress function
+that differences distances between points in \pD with potential
+distances between points in \kD. Various formulations of the stress
+function result in non-metric scaling \citep{saeed2018} and isomap
+\citep{silva2002}. Challenges in working with high-dimensional data,
+including visualization, are outlined in \citet{johnstone2009}.
 
 Many new methods for NLDR have emerged in recent years, all designed to
-better capture specific structures potentially existing in \(p-D\). Here
-we focus on five currently popular techniques, tSNE, UMAP, PHATE, TriMAP
+better capture specific structures potentially existing in \pD. Here we
+focus on five currently popular techniques, tSNE, UMAP, PHATE, TriMAP
 and PaCMAP. tNSE and UMAP can be considered to produce the \(k-D\)
@@
 set of new variables which are linear combinations of the original
-variables. Tours, defined by \citet{asimov1985}, broaden the scope by
+variables. Tours, defined by \citet{lee2021}, broaden the scope by
 providing movies of linear projections, that provide views the data from
@@
 patterns can obscure local structure. The simple activity of projecting
-data from \(p-D\) suffers from piling \citep{laa2022}, where data
+data from \pD suffers from piling \citep{laa2022}, where data
 concentrates in the center of projections. NLDR is designed to escape
@@
 While this is straightforward, and commonly done when data is 2D, it is
-also possible in \(p-D\), for many models, when a tour is used.
+also possible in \pD, for many models, when a tour is used.
 
 \citet{wickham2015} provides several examples of models overlaid on the
-data in \(p-D\). In hierarchical clustering, a representation of the
+data in \pD. In hierarchical clustering, a representation of the
 dendrogrom using points and lines can be constructed by augmenting the
@@
 the cluster model to the data. For linear discriminant analysis or
-model-based clustering the model can be indicated by \((p-1)-D\)
+model-based clustering the model can be indicated by \((p-1)\text{-}D\)
 ellipses. It is possible to see whether the elliptical shapes
@@
 compare and contrast different fits. For PCA, one can display the
-\(k-D\) plane of the reduced dimension using wireframes of transformed
+\kD plane of the reduced dimension using wireframes of transformed
 cubes. Using a wireframe is the approach we take here, to represent the
-NLDR model in \(p-D\).
+NLDR model in \pD.
 
@@
 representation or abstraction of a system, process, or phenomenon in the
-real world''. The \(p-D\) observations are the realization of the
-phenomenon, and the \(k-D\) NLDR layout is the simplified
-representation. From a statistical perspective we can consider the
-distances between points in the \(k-D\) layout to be variance that the
-model explains, and the (relative) difference with their distances in
-\(p-D\) is the error, or unexplained variance. Abstractly, we can also
-imagine that the positioning of points in 2D represents fitted values,
-that will have some prescribed position in \(p-D\) that can be compared
-with the observed value.
+real world''. The \pD observations are the realization of the
+phenomenon, and the \kD NLDR layout is the simplified representation.
+From a statistical perspective we can consider the distances between
+points in the \kD layout to be variance that the model explains, and the
+(relative) difference with their distances in \pD is the error, or
+unexplained variance. Abstractly, we can also imagine that the
+positioning of points in 2D represents fitted values, that will have
+some prescribed position in \pD that can be compared with the observed
+value.
 
@@
 \subsection{\texorpdfstring{Displaying the model in
-\(p-D\)}{Displaying the model in p-D}}\label{displaying-the-model-in-p-d}
+\pD}{Displaying the model in }}\label{displaying-the-model-in}
 

### 2024-03-30 paper.tex
Added updated notation.csv

@@
 $n$ & number of observations\\
-$p$ & number of dimensions in $p-D$ data\\
-$X_{n \times p} = \begin{bmatrix} \textbf{x} _{1} & \textbf{x}_ {2} & \cdots & \textbf{x}_{n} \\  \end{bmatrix}^\top$ & $p-D$ data matrix\\
+$p$ & number of dimensions in high-D data\\
+$X_{n \times p} = \begin{bmatrix} \textbf{x} _{1} & \textbf{x}_ {2} & \cdots & \textbf{x}_{n} \\  \end{bmatrix}^\top$ & high-D data matrix\\
 $d$ & number of dimensions in embedding data, usually two\\
@@
 \addlinespace
-$P_t$ & point along the geodesic path at time $t$\\
-$P_{t+1}$ & point along the geodesic path at the next time instant, which is $t+1$\\
+$P_t$, $P_{t+1}$ & point along the geodesic path at time $t$ and the next time instant, which is $t+1$\\
 $y_max$ & maximum value of the scaled second embedding component\\
-$r_1$ & range of the first embedding component\\
-$r_2$ & range of the second embedding component\\
+$r_1$, $r_2$ & range of the first and  second embedding component\\
+$h_b$, $w_b$ & height and width of the hexagon\\
+$a_r$ & aspect ratio of the 2D layout before scaling\\
 \addlinespace
-$h_b$ & height of the hexagon\\
-$w_b$ & width of the hexagon\\
-$a_r$ & aspect ratio of the 2D layout before scaling\\
 $h_r$ & hexagon ratio\\
 $s$ & hexagonal size (radius of the outer circle surrounding the hexagon)\\
-\addlinespace
 $(x_s, y_s)$ & starting coordinates of the hexagonal grid\\
-$q_x$ & buffer amount along the x-axis\\
-$q_y$ & buffer amount along the y-axis\\
-$h_s$ & horizontal distance between
-adjacent hexagon bin \vphantom{1} centroids\\
-$v_s$ & vertical distance between
-adjacent hexagon bin centroids\\
+$q_x$, $q_y$ & buffer amount along the x and y axes\\
+$h_s$, $v_s$ & horizontal and vertical distance between adjacent hexagon bin centroids\\
 \addlinespace
@@
 $d_y$ & adjusted vertical spacing for hexagonal coordinates\\
-$b_x$ & number of hexagon bins along the x-axis\\
-$b_y$ & number of hexagon bins along the y-axis\\
+$b_x$, $b_y$ & number of hexagon bins along the x and y axes\\
 $b$ & total number of hexagon bins in the grid\\
+$b'$ & number of hexagon bins without the empty bins\\
 \addlinespace
-$b'$ & number of hexagon bins without the empty bins\\
 $n_k$ & number of observations within $k^{th}$ hexagon\\
@@
 $C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2})$ & hexagonal bin centroid coordinates of $k^{th}$ hexagon\\
-$C_k^{(p)} \equiv (C_{kx_1}, ..., C_{kx_p})$ & $p-D$ mappings of 2D hexagonal bin centroids of $k^{th}$ hexagon\\
+$C_k^{(p)} \equiv (C_{kx_1}, ..., C_{kx_p})$ & high-D mappings of 2D hexagonal bin centroids of $k^{th}$ hexagon\\
 \bottomrule

### 2024-03-30 paper.qmd
Change the table properties

@@
       caption = "Notations used in this paper.", label = "tbl-notations") |>
-  kable_styling(full_width = FALSE) |>
-  row_spec(0, bold = TRUE)
+  kable_styling(latex_options = c("HOLD_position"), position = "center", full_width = FALSE) |>
+  row_spec(0, bold = TRUE) |>
+  column_spec(1:2, width = c("5cm", "12cm"))
 ```

### 2024-03-30 paper.tex
Change the table properties
Changed high-D to p-D in the table

@@
 
-\begin{table}
+\begin{table}[H]
 
@@
 \centering
-\begin{tabular}[t]{ll}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{5cm}>{\raggedright\arraybackslash}p{12cm}}
 \toprule
@@
 $n$ & number of observations\\
-$p$ & number of dimensions in high-D data\\
-$X_{n \times p} = \begin{bmatrix} \textbf{x} _{1} & \textbf{x}_ {2} & \cdots & \textbf{x}_{n} \\  \end{bmatrix}^\top$ & high-D data matrix\\
+$p$ & number of dimensions in $p$-D data\\
+$X_{n \times p} = \begin{bmatrix} \textbf{x} _{1} & \textbf{x}_ {2} & \cdots & \textbf{x}_{n} \\  \end{bmatrix}^\top$ & $p$-D data matrix\\
 $d$ & number of dimensions in embedding data, usually two\\
@@
 $C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2})$ & hexagonal bin centroid coordinates of $k^{th}$ hexagon\\
-$C_k^{(p)} \equiv (C_{kx_1}, ..., C_{kx_p})$ & high-D mappings of 2D hexagonal bin centroids of $k^{th}$ hexagon\\
+$C_k^{(p)} \equiv (C_{kx_1}, ..., C_{kx_p})$ & $p$-D mappings of 2D hexagonal bin centroids of $k^{th}$ hexagon\\
 \bottomrule

### 2024-03-30 paper.qmd
Fix the issue not appearing space after p-D and k-D
Changed font size of the notation table

--- 2024-03-30 paper.qmd
@@
   \usepackage{setspace}
-  \newcommand\pD{$p\text{-}D$}
-  \newcommand\kD{$k\text{-}D$}
+  \newcommand\pD{$p\text{-}D$\ }
+  \newcommand\kD{$k\text{-}D$\ }
 ---
@@
       caption = "Notations used in this paper.", label = "tbl-notations") |>
-  kable_styling(full_width = FALSE) |>
-  row_spec(0, bold = TRUE)
+  kable_styling(latex_options = c("HOLD_position"), position = "center", full_width = FALSE, font_size = 12) |>
+  row_spec(0, bold = TRUE) |>
+  column_spec(1:2, width = c("5cm", "12cm"))
 ```

### 2024-03-30 paper.tex
Fix the issue not appearing space after p-D and k-D
Changed font size of the notation table
Changed the notation order

@@
 \usepackage{setspace}
-\newcommand\pD{$p\text{-}D$}
-\newcommand\kD{$k\text{-}D$}
+\newcommand\pD{$p\text{-}D$\ }
+\newcommand\kD{$k\text{-}D$\ }
 \makeatletter
@@
 \centering
+\fontsize{12}{14}\selectfont
 \begin{tabular}[t]{>{\raggedright\arraybackslash}p{5cm}>{\raggedright\arraybackslash}p{12cm}}
@@
 $p$ & number of dimensions in $p$-D data\\
+$d$ & number of dimensions in embedding data, usually two\\
 $X_{n \times p} = \begin{bmatrix} \textbf{x} _{1} & \textbf{x}_ {2} & \cdots & \textbf{x}_{n} \\  \end{bmatrix}^\top$ & $p$-D data matrix\\
-$d$ & number of dimensions in embedding data, usually two\\
 $Y_{n \times d} = \begin{bmatrix} \textbf{y} _{1} & \textbf{y}_ {2} & \cdots & \textbf{y}_{n} \\  \end{bmatrix}^\top$ & embedding matrix\\

### 2024-03-30 paper.tex
Aggregate notations

@@
 $n$ & number of observations\\
-$p$ & number of dimensions in $p$-D data\\
+$p$ & number of variables in $p$-D data\\
 $d$ & number of dimensions in embedding data, usually two\\
@@
 $P_t$, $P_{t+1}$ & point along the geodesic path at time $t$ and the next time instant, which is $t+1$\\
-$y_max$ & maximum value of the scaled second embedding component\\
+$y_{max}$ & maximum value of the scaled second embedding component\\
 $r_1$, $r_2$ & range of the first and  second embedding component\\
@@
 \addlinespace
-$d_x$ & horizontal spacing\\
-$d_y$ & adjusted vertical spacing for hexagonal coordinates\\
+$d_x$, $d_y$ & horizontal spacing and adjusted vertical spacing for hexagonal coordinates\\
 $b_x$, $b_y$ & number of hexagon bins along the x and y axes\\
-$b$ & total number of hexagon bins in the grid\\
-$b'$ & number of hexagon bins without the empty bins\\
-\addlinespace
+$b$, $b'$ & total and non-empty hexagon bins in the grid\\
 $n_k$ & number of observations within $k^{th}$ hexagon\\
 $(h^k_{x_i}, h^k_{y_i})$ & hexagonal grid coordinates of $k^{th}$ hexagon\\
+\addlinespace
 $C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2})$ & hexagonal bin centroid coordinates of $k^{th}$ hexagon\\
@@
 \end{table}
-
-\textless\textless{} XXX Jayani, please insert a table of notation
-\textgreater\textgreater{}
 

### 2024-03-30 paper.tex
Update table

@@
 $s$ & hexagonal size (radius of the outer circle surrounding the hexagon)\\
-$(x_s, y_s)$ & starting coordinates of the hexagonal grid\\
+$(x_s, \ y_s)$ & starting coordinates of the hexagonal grid\\
 $q_x$, $q_y$ & buffer amount along the x and y axes\\
@@
 $b$, $b'$ & total and non-empty hexagon bins in the grid\\
-$n_k$ & number of observations within $k^{th}$ hexagon\\
-$(h^k_{x_i}, h^k_{y_i})$ & hexagonal grid coordinates of $k^{th}$ hexagon\\
+$n_k$ & number of observations within the $k^{th}$ hexagon\\
+$(h^k_{x_i}, h^k_{y_i})$ & hexagonal grid coordinates of the $k^{th}$ hexagon\\
 \addlinespace
-$C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2})$ & hexagonal bin centroid coordinates of $k^{th}$ hexagon\\
-$C_k^{(p)} \equiv (C_{kx_1}, ..., C_{kx_p})$ & $p$-D mappings of 2D hexagonal bin centroids of $k^{th}$ hexagon\\
+$C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2})$ & hexagonal bin centroid coordinates of the $k^{th}$ hexagon\\
+$C_k^{(p)} \equiv (C_{kx_1}, ..., C_{kx_p})$ & $p$-D mappings of 2D hexagonal bin centroids of the $k^{th}$ hexagon\\
 \bottomrule

### 2024-03-30 paper.qmd
Aggregate notations
tidying of table and summary paragraph

@@
 
+\spacingset{1.0} <!--% command in JASA style, comment to go back to double spacing-->
+
 ## Introduction
@@
 
+@tbl-notation summarises the notation used to explain the new methodology.
+
 ```{r}
+#| label: tbl-notation
+#| tbl-cap: "Summary of notation for describing new methodology."
 # Notations used in the paper
 
-notation_df <- read_csv("data/paper_notations.csv")
+notation_df <- read_csv("misc/notation.csv")
 
 # Create the table
-kable(notation_df, format = "latex", booktabs = TRUE, escape = FALSE, 
-      caption = "Notations used in this paper.", label = "tbl-notations") |>
-  kable_styling(latex_options = c("HOLD_position"), position = "center", full_width = FALSE, font_size = 12) |>
+kable(notation_df, 
+      format = "latex", 
+      booktabs = TRUE, escape = FALSE) |>
+  kable_styling(latex_options = c("HOLD_position"),
+                position = "center", 
+                full_width = FALSE, 
+                font_size = 12) |>
   row_spec(0, bold = TRUE) |>
@@
 ```
-
-<< XXX Jayani, please insert a table of notation >>
 

### 2024-03-30 paper.tex
tidying of table and summary paragraph

--- 2024-03-30 paper.tex
@@
 
+\spacingset{1.0}
+
 \section{Introduction}\label{introduction}
@@
 
-\begin{table}[H]
+Table~\ref{tbl-notation} summarises the notation used to explain the new
+methodology.
 
-\caption{\label{tab:tbl-notations}Notations used in this paper.}
-\centering
-\fontsize{12}{14}\selectfont
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{5cm}>{\raggedright\arraybackslash}p{12cm}}
+\begin{table}
+
+\centering{
+
+[H]
+\centering\begingroup\fontsize{12}{14}\selectfont
+
+\begin{tabular}{>{\raggedright\arraybackslash}p{5cm}>{\raggedright\arraybackslash}p{12cm}}
 \toprule
@@
 \midrule
-$n$ & number of observations\\
-$p$ & number of variables in $p$-D data\\
-$d$ & number of dimensions in embedding data, usually two\\
-$X_{n \times p} = \begin{bmatrix} \textbf{x} _{1} & \textbf{x}_ {2} & \cdots & \textbf{x}_{n} \\  \end{bmatrix}^\top$ & $p$-D data matrix\\
-$Y_{n \times d} = \begin{bmatrix} \textbf{y} _{1} & \textbf{y}_ {2} & \cdots & \textbf{y}_{n} \\  \end{bmatrix}^\top$ & embedding matrix\\
-\addlinespace
+$n, p, d$ & number of observations, variables, embedding dimension, respectively\\
+$X_{n \times p} = \begin{bmatrix} \textbf{x} _{1} & \textbf{x}_ {2} & \cdots & \textbf{x}_{n} \\  \end{bmatrix}^\top$ & observed data matrix\\
+$Y_{n \times d} = \begin{bmatrix} \textbf{y} _{1} & \textbf{y}_ {2} & \cdots & \textbf{y}_{n} \\  \end{bmatrix}^\top$ & reduced dimension, embedding matrix\\
 $P_t$, $P_{t+1}$ & point along the geodesic path at time $t$ and the next time instant, which is $t+1$\\
 $y_{max}$ & maximum value of the scaled second embedding component\\
+\addlinespace
 $r_1$, $r_2$ & range of the first and  second embedding component\\
@@
 $a_r$ & aspect ratio of the 2D layout before scaling\\
-\addlinespace
 $h_r$ & hexagon ratio\\
 $s$ & hexagonal size (radius of the outer circle surrounding the hexagon)\\
-$(x_s, y_s)$ & starting coordinates of the hexagonal grid\\
+\addlinespace
+$(x_s, \ y_s)$ & starting coordinates of the hexagonal grid\\
 $q_x$, $q_y$ & buffer amount along the x and y axes\\
 $h_s$, $v_s$ & horizontal and vertical distance between adjacent hexagon bin centroids\\
-\addlinespace
 $d_x$, $d_y$ & horizontal spacing and adjusted vertical spacing for hexagonal coordinates\\
 $b_x$, $b_y$ & number of hexagon bins along the x and y axes\\
+\addlinespace
 $b$, $b'$ & total and non-empty hexagon bins in the grid\\
-$n_k$ & number of observations within $k^{th}$ hexagon\\
-$(h^k_{x_i}, h^k_{y_i})$ & hexagonal grid coordinates of $k^{th}$ hexagon\\
-\addlinespace
-$C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2})$ & hexagonal bin centroid coordinates of $k^{th}$ hexagon\\
-$C_k^{(p)} \equiv (C_{kx_1}, ..., C_{kx_p})$ & $p$-D mappings of 2D hexagonal bin centroids of $k^{th}$ hexagon\\
+$n_k$ & number of observations within the $k^{th}$ hexagon\\
+$(h^k_{x_i}, h^k_{y_i})$ & hexagonal grid coordinates of the $k^{th}$ hexagon\\
+$C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2})$ & hexagonal bin centroid coordinates of the $k^{th}$ hexagon\\
+$C_k^{(p)} \equiv (C_{kx_1}, ..., C_{kx_p})$ & $p$-D mappings of 2D hexagonal bin centroids of the $k^{th}$ hexagon\\
 \bottomrule
 \end{tabular}
-\end{table}
+\endgroup{}
+
+}
+
+\caption{\label{tbl-notation}Summary of notation for describing new
+methodology.}
+
+\end{table}%
 

### 2024-03-31 paper.qmd
removed latex H from table

@@
       booktabs = TRUE, escape = FALSE) |>
-  kable_styling(latex_options = c("HOLD_position"),
-                position = "center", 
+  kable_styling(position = "center", 
                 full_width = FALSE, 

### 2024-04-01 paper.qmd
small tweaks to notation and start on methodology

@@
   \newcommand\kD{$k\text{-}D$\ }
+  \newcommand\dD{$d\text{-}D$\ }
 ---
@@
 
-At first glance, thinking of of NLDR as a model fitted to the data might seem strange.
+At first glance, thinking of NLDR as a model fitted to the data might seem strange.
 It is a model in the sense that it is a "a simplified representation or abstraction of a system, process, or phenomenon in the real world".
@@
 
-### Notation
+<!--### Notation -->
 
 @tbl-notation summarises the notation used to explain the new methodology.
+$X_{n \times p} = \begin{bmatrix} \textbf{x} _{1} & \textbf{x}_ {2} & \cdots & \textbf{x}_{n} \\  \end{bmatrix}^\top$ represents the observed data matrix, that are $n$ \pD vectors sampled from some population.
+
+<!--
+$X_{n \times p} = \begin{bmatrix} \textbf{x} _{1} & \textbf{x}_ {2} & \cdots & \textbf{x}_{n} \\  \end{bmatrix}^\top$
+
+$Y_{n \times d} = \begin{bmatrix} \textbf{y} _{1} & \textbf{y}_ {2} & \cdots & \textbf{y}_{n} \\  \end{bmatrix}^\top$
+
+$C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2})$
+
+$C_k^{(p)} \equiv (C_{kx_1}, ..., C_{kx_p})$ $p$-D mappings of 2D hexagonal bin centroids of the $k^{th}$ hexagon
+
+-->
 
@@
   row_spec(0, bold = TRUE) |>
-  column_spec(1:2, width = c("5cm", "12cm"))
+  column_spec(1:2, width = c("3cm", "12cm"))
 ```
 
-Once we have notation, we need to have some math that gives precision to the paragraph above. (XXX Michael??)
+We need to have some math that gives precision to the paragraphs in the previous section, defining mapping from \pD to \kD to get the \kD layout. Might be written below, too (XXX Michael??)
+
+To illustrate the method, we use simulated data. The S-curve is generated as follows XXX Jayani
+
+### Generating the \kD layout
+
+The usual procedures are followed to produce a \kD layout, using the preferred method, t-SNE, UMAP, ... The data may need to be pre-processed by standardising or transforming to principal components.
 

### 2024-04-01 paper.tex
removed latex H from table
small tweaks to notation and start on methodology

@@
 \newcommand\kD{$k\text{-}D$\ }
+\newcommand\dD{$d\text{-}D$\ }
 \makeatletter
@@
 
-At first glance, thinking of of NLDR as a model fitted to the data might
+At first glance, thinking of NLDR as a model fitted to the data might
 seem strange. It is a model in the sense that it is a ``a simplified
@@
 
-\subsection{Notation}\label{notation}
-
 Table~\ref{tbl-notation} summarises the notation used to explain the new
 methodology.
+\(X_{n \times p} = \begin{bmatrix} \textbf{x} _{1} & \textbf{x}_ {2} & \cdots & \textbf{x}_{n} \\  \end{bmatrix}^\top\)
+represents the observed data matrix, that are \(n\) \pD vectors sampled
+from some population.
 
@@
 
-[H]
 \centering\begingroup\fontsize{12}{14}\selectfont
 
-\begin{tabular}{>{\raggedright\arraybackslash}p{5cm}>{\raggedright\arraybackslash}p{12cm}}
+\begin{tabular}{>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{12cm}}
 \toprule
@@
 \midrule
-$n, p, d$ & number of observations, variables, embedding dimension, respectively\\
-$X_{n \times p} = \begin{bmatrix} \textbf{x} _{1} & \textbf{x}_ {2} & \cdots & \textbf{x}_{n} \\  \end{bmatrix}^\top$ & observed data matrix\\
-$Y_{n \times d} = \begin{bmatrix} \textbf{y} _{1} & \textbf{y}_ {2} & \cdots & \textbf{y}_{n} \\  \end{bmatrix}^\top$ & reduced dimension, embedding matrix\\
-$P_t$, $P_{t+1}$ & point along the geodesic path at time $t$ and the next time instant, which is $t+1$\\
+$n, p, k$ & number of observations, variables, embedding dimension, respectively\\
+$X$ & $p$-dimensional data\\
+$Y$ & $k$-dimensional layout\\
+$P$ & orthonormal basis, generating a $d-dimensional$ linear projection of $p$-dimensional data\\
 $y_{max}$ & maximum value of the scaled second embedding component\\
@@
 $(h^k_{x_i}, h^k_{y_i})$ & hexagonal grid coordinates of the $k^{th}$ hexagon\\
-$C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2})$ & hexagonal bin centroid coordinates of the $k^{th}$ hexagon\\
-$C_k^{(p)} \equiv (C_{kx_1}, ..., C_{kx_p})$ & $p$-D mappings of 2D hexagonal bin centroids of the $k^{th}$ hexagon\\
+$C^{(j)}$ & $j$-dimensional bin centers\\
 \bottomrule
@@
 
-Once we have notation, we need to have some math that gives precision to
-the paragraph above. (XXX Michael??)
+We need to have some math that gives precision to the paragraphs in the
+previous section, defining mapping from \pD to \kD to get the
+\kD layout. Might be written below, too (XXX Michael??)
+
+To illustrate the method, we use simulated data. The S-curve is
+generated as follows XXX Jayani
+
+\subsection{\texorpdfstring{Generating the
+\kD layout}{Generating the layout}}\label{generating-the-layout}
+
+The usual procedures are followed to produce a \kD layout, using the
+preferred method, t-SNE, UMAP, \ldots{} The data may need to be
+pre-processed by standardising or transforming to principal components.
 

### 2024-04-01 paper.qmd
Add true model notation to the table

@@
 
+The S-curve model, denoted as \( T \), is simulated through a multi-step process.
+Initially, a random variable \( T \) is generated from a uniform distribution between \(-\frac{\pi}{2}\) and \(\frac{\pi}{2}\), representing the parameter \( T \) in the S-curve model.
+Subsequently, three variables \( x_1 \), \( x_2 \), and \( x_3 \) are computed based on \( T \) to construct the S-curve data.
+\( x_1 \) is calculated as the sine of \( T \), \( x_2 \) as twice the value of another uniformly distributed random variable between 0 and 1, and \( x_3 \) as the cosine of \( T \) minus 1, multiplied by the sign of \( T \).
+Additionally, four noise dimensions are introduced to the dataset to introduce variability: \( x_4 \), \( x_5 \), \( x_6 \), and \( x_7 \), each generated from uniform distributions within specific intervals.
+Finally, these variables are aggregated to form the S-curve dataset \( X \).
+
+Let \( T \) be a random variable uniformly distributed in the interval \( [-\frac{\pi}{2}, \frac{\pi}{2}] \), i.e., \( T \sim U\left(-\frac{\pi}{2}, \frac{\pi}{2}\right) \).
+
+Then, we define the following variables:
+
+- \( x_1 = \sin(3 \pi T) \)
+- \( x_2 = 2.0 \times U(0, 1) \), where \( U(0, 1) \) represents a uniform distribution between 0 and 1.
+- \( x_3 = \text{sign}(T) \times (\cos(3 \pi T) - 1) \)
+
+Additionally, we introduce noise dimensions:
+
+- \( x_4, x_5 \) are uniformly distributed random variables in the interval \( [-0.02, 0.02] \), i.e., \( x_4, x_5 \sim U(-0.02, 0.02) \).
+- \( x_6 \) is a uniformly distributed random variable in the interval \( [-0.1, 0.1] \), i.e., \( x_6 \sim U(-0.1, 0.1) \).
+- \( x_7 \) is a uniformly distributed random variable in the interval \( [-0.01, 0.01] \), i.e., \( x_7 \sim U(-0.01, 0.01) \).
+
+Finally, we form the data matrix \( X \) by combining \( x_1, x_2, x_3, x_4, x_5, x_6, \) and \( x_7 \) into columns.
+
 ### Generating the \kD layout

### 2024-04-01 paper.tex
Add true model notation to the table

@@
 $P$ & orthonormal basis, generating a $d-dimensional$ linear projection of $p$-dimensional data\\
+$T$ & true  model\\
+\addlinespace
 $y_{max}$ & maximum value of the scaled second embedding component\\
-\addlinespace
 $r_1$, $r_2$ & range of the first and  second embedding component\\
@@
 $h_r$ & hexagon ratio\\
+\addlinespace
 $s$ & hexagonal size (radius of the outer circle surrounding the hexagon)\\
-\addlinespace
 $(x_s, \ y_s)$ & starting coordinates of the hexagonal grid\\
@@
 $d_x$, $d_y$ & horizontal spacing and adjusted vertical spacing for hexagonal coordinates\\
+\addlinespace
 $b_x$, $b_y$ & number of hexagon bins along the x and y axes\\
-\addlinespace
 $b$, $b'$ & total and non-empty hexagon bins in the grid\\
@@
 generated as follows XXX Jayani
+
+The S-curve model, denoted as ( T ), is simulated through a multi-step
+process. Initially, a random variable ( T ) is generated from a uniform
+distribution between (-\frac{\pi}{2}) and (\frac{\pi}{2}), representing
+the parameter ( T ) in the S-curve model. Subsequently, three variables
+( x\_1 ), ( x\_2 ), and ( x\_3 ) are computed based on ( T ) to
+construct the S-curve data. ( x\_1 ) is calculated as the sine of ( T ),
+( x\_2 ) as twice the value of another uniformly distributed random
+variable between 0 and 1, and ( x\_3 ) as the cosine of ( T ) minus 1,
+multiplied by the sign of ( T ). Additionally, four noise dimensions are
+introduced to the dataset to introduce variability: ( x\_4 ), ( x\_5 ),
+( x\_6 ), and ( x\_7 ), each generated from uniform distributions within
+specific intervals. Finally, these variables are aggregated to form the
+S-curve dataset ( X ).
+
+Let ( T ) be a random variable uniformly distributed in the interval (
+{[}-\frac{\pi}{2}, \frac{\pi}{2}{]} ), i.e., ( T
+\sim U\left(-\frac{\pi}{2}, \frac{\pi}{2}\right) ).
+
+Then, we define the following variables:
+
+\begin{itemize}
+\tightlist
+\item
+  ( x\_1 = \sin(3 \pi T) )
+\item
+  ( x\_2 = 2.0 \times U(0, 1) ), where ( U(0, 1) ) represents a uniform
+  distribution between 0 and 1.
+\item
+  ( x\_3 = \text{sign}(T) \times (\cos(3 \pi T) - 1) )
+\end{itemize}
+
+Additionally, we introduce noise dimensions:
+
+\begin{itemize}
+\tightlist
+\item
+  ( x\_4, x\_5 ) are uniformly distributed random variables in the
+  interval ( {[}-0.02, 0.02{]} ), i.e., ( x\_4, x\_5 \sim U(-0.02, 0.02)
+  ).
+\item
+  ( x\_6 ) is a uniformly distributed random variable in the interval (
+  {[}-0.1, 0.1{]} ), i.e., ( x\_6 \sim U(-0.1, 0.1) ).
+\item
+  ( x\_7 ) is a uniformly distributed random variable in the interval (
+  {[}-0.01, 0.01{]} ), i.e., ( x\_7 \sim U(-0.01, 0.01) ).
+\end{itemize}
+
+Finally, we form the data matrix ( X ) by combining ( x\_1, x\_2, x\_3,
+x\_4, x\_5, x\_6, ) and ( x\_7 ) into columns.
 

### 2024-04-01 paper.qmd
Added main text for the data

--- 2024-04-01 paper.qmd
@@
 
+The S-curve model, denoted as $T$, is simulated through several steps.
+First, a random variable $x$ is generated from a uniform distribution between $-\frac{\pi}{2}$ and $\frac{\pi}{2}$, representing the parameter $x$ in the S-curve model.
+Next, three variables $x_1$, $x_2$, and $x_3$ are computed based on $x$ to construct the S-curve data.
+$x_1$ is calculated as the sine of $x$ (\@ref(eq:eqn2)), $x_2$ as twice the value of another uniformly distributed random variable between 0 and 1, and $x_3$ as the cosine of $x$ minus 1, multiplied by the sign of $x$.
+Additionally, four noise dimensions are introduced to the dataset to introduce variability: $x_4$, $x_5$, $x_6$, and $x_7$, each generated from uniform distributions within specific intervals.
+Finally, these variables are combined to form the S-curve dataset $T$.
+
+\begin{equation} \label{eq-eqn1}
+x_1 = \sin(3 \pi x)
+\end{equation}
+
+\begin{equation} \label{eq:eqn2}
+x_2 = 2.0 \times U(0, 1)
+\end{equation}
+
+\begin{equation} \label{eq:eqn3}
+x_3 = \text{sign}(x) \times (\cos(3 \pi x) - 1)
+\end{equation}
+
+\begin{equation} \label{eq:eqn4}
+x_4, x_5 \sim U(-0.02, 0.02)
+\end{equation}
+
+\begin{equation} \label{eq:eqn5}
+x_6 \sim U(-0.1, 0.1)
+\end{equation}
+
+\begin{equation} \label{eq:eqn6}
+x_7 \sim U(-0.01, 0.01)
+\end{equation}
+
 ### Generating the \kD layout

### 2024-04-01 paper.tex
Added main text for the data

@@
 
-The S-curve model, denoted as ( T ), is simulated through a multi-step
-process. Initially, a random variable ( T ) is generated from a uniform
-distribution between (-\frac{\pi}{2}) and (\frac{\pi}{2}), representing
-the parameter ( T ) in the S-curve model. Subsequently, three variables
-( x\_1 ), ( x\_2 ), and ( x\_3 ) are computed based on ( T ) to
-construct the S-curve data. ( x\_1 ) is calculated as the sine of ( T ),
-( x\_2 ) as twice the value of another uniformly distributed random
-variable between 0 and 1, and ( x\_3 ) as the cosine of ( T ) minus 1,
-multiplied by the sign of ( T ). Additionally, four noise dimensions are
-introduced to the dataset to introduce variability: ( x\_4 ), ( x\_5 ),
-( x\_6 ), and ( x\_7 ), each generated from uniform distributions within
-specific intervals. Finally, these variables are aggregated to form the
-S-curve dataset ( X ).
+The S-curve model, denoted as \(T\), is simulated through several steps.
+First, a random variable \(x\) is generated from a uniform distribution
+between \(-\frac{\pi}{2}\) and \(\frac{\pi}{2}\), representing the
+parameter \(x\) in the S-curve model. Next, three variables \(x_1\),
+\(x_2\), and \(x_3\) are computed based on \(x\) to construct the
+S-curve data. \(x_1\) is calculated as the sine of \(x\)
+(@ref(eq:eqn2)), \(x_2\) as twice the value of another uniformly
+distributed random variable between 0 and 1, and \(x_3\) as the cosine
+of \(x\) minus 1, multiplied by the sign of \(x\). Additionally, four
+noise dimensions are introduced to the dataset to introduce variability:
+\(x_4\), \(x_5\), \(x_6\), and \(x_7\), each generated from uniform
+distributions within specific intervals. Finally, these variables are
+combined to form the S-curve dataset \(T\).
 
-Let ( T ) be a random variable uniformly distributed in the interval (
-{[}-\frac{\pi}{2}, \frac{\pi}{2}{]} ), i.e., ( T
-\sim U\left(-\frac{\pi}{2}, \frac{\pi}{2}\right) ).
+\begin{equation} \label{eq-eqn1}
+x_1 = \sin(3 \pi x)
+\end{equation}
 
-Then, we define the following variables:
+\begin{equation} \label{eq:eqn2}
+x_2 = 2.0 \times U(0, 1)
+\end{equation}
 
-\begin{itemize}
-\tightlist
-\item
-  ( x\_1 = \sin(3 \pi T) )
-\item
-  ( x\_2 = 2.0 \times U(0, 1) ), where ( U(0, 1) ) represents a uniform
-  distribution between 0 and 1.
-\item
-  ( x\_3 = \text{sign}(T) \times (\cos(3 \pi T) - 1) )
-\end{itemize}
+\begin{equation} \label{eq:eqn3}
+x_3 = \text{sign}(x) \times (\cos(3 \pi x) - 1)
+\end{equation}
 
-Additionally, we introduce noise dimensions:
+\begin{equation} \label{eq:eqn4}
+x_4, x_5 \sim U(-0.02, 0.02)
+\end{equation}
 
-\begin{itemize}
-\tightlist
-\item
-  ( x\_4, x\_5 ) are uniformly distributed random variables in the
-  interval ( {[}-0.02, 0.02{]} ), i.e., ( x\_4, x\_5 \sim U(-0.02, 0.02)
-  ).
-\item
-  ( x\_6 ) is a uniformly distributed random variable in the interval (
-  {[}-0.1, 0.1{]} ), i.e., ( x\_6 \sim U(-0.1, 0.1) ).
-\item
-  ( x\_7 ) is a uniformly distributed random variable in the interval (
-  {[}-0.01, 0.01{]} ), i.e., ( x\_7 \sim U(-0.01, 0.01) ).
-\end{itemize}
+\begin{equation} \label{eq:eqn5}
+x_6 \sim U(-0.1, 0.1)
+\end{equation}
 
-Finally, we form the data matrix ( X ) by combining ( x\_1, x\_2, x\_3,
-x\_4, x\_5, x\_6, ) and ( x\_7 ) into columns.
+\begin{equation} \label{eq:eqn6}
+x_7 \sim U(-0.01, 0.01)
+\end{equation}
 

### 2024-04-01 paper.qmd
Correct latex error

--- 2024-04-01 paper.qmd
@@
 
+The S-curve model, denoted as $T$, is simulated through several steps.
+First, a random variable $x$ is generated from a uniform distribution between $-\frac{\pi}{2}$ and $\frac{\pi}{2}$, representing the parameter $a$ in the S-curve model.
+Next, three variables $x_1$, $x_2$, and $x_3$ are computed based on $x$ to construct the S-curve data.
+$x_1$ is calculated as the sine of $a$ (@eq-eqn1), $x_2$ as twice the value of another uniformly distributed random variable between 0 and 1 (@eq-eqn2), and $x_3$ as the cosine of $a$ minus 1, multiplied by the sign of $a$ (@eq-eqn3).
+Additionally, four noise dimensions are introduced to the dataset to introduce variability: $x_4$, $x_5$, $x_6$, and $x_7$, each generated from uniform distributions within specific intervals (@eq-eqn4, @eq-eqn5, and @eq-eqn6 respectively).
+Finally, these variables are combined to form the S-curve dataset $T$.
+
+$$
+x_1 = \sin(3 \pi x)
+$$ {#eq-eqn1}
+
+$$
+x_2 = 2.0 \times U(0, 1)
+$$ {#eq-eqn2}
+
+$$
+x_3 = \text{sign}(x) \times (\cos(3 \pi x) - 1)
+$$ {#eq-eqn3}
+
+$$
+x_4, x_5 \sim U(-0.02, 0.02)
+$$ {#eq-eqn4}
+
+$$
+x_6 \sim U(-0.1, 0.1)
+$$ {#eq-eqn5}
+
+$$
+x_7 \sim U(-0.01, 0.01)
+$$ {#eq-eqn6}
+
 ### Generating the \kD layout

### 2024-04-01 paper.tex
Correct latex error

@@
 between \(-\frac{\pi}{2}\) and \(\frac{\pi}{2}\), representing the
-parameter \(x\) in the S-curve model. Next, three variables \(x_1\),
+parameter \(a\) in the S-curve model. Next, three variables \(x_1\),
 \(x_2\), and \(x_3\) are computed based on \(x\) to construct the
-S-curve data. \(x_1\) is calculated as the sine of \(x\)
-(@ref(eq:eqn2)), \(x_2\) as twice the value of another uniformly
-distributed random variable between 0 and 1, and \(x_3\) as the cosine
-of \(x\) minus 1, multiplied by the sign of \(x\). Additionally, four
-noise dimensions are introduced to the dataset to introduce variability:
-\(x_4\), \(x_5\), \(x_6\), and \(x_7\), each generated from uniform
-distributions within specific intervals. Finally, these variables are
-combined to form the S-curve dataset \(T\).
+S-curve data. \(x_1\) is calculated as the sine of \(a\)
+(Equation~\ref{eq-eqn1}), \(x_2\) as twice the value of another
+uniformly distributed random variable between 0 and 1
+(Equation~\ref{eq-eqn2}), and \(x_3\) as the cosine of \(a\) minus 1,
+multiplied by the sign of \(a\) (Equation~\ref{eq-eqn3}). Additionally,
+four noise dimensions are introduced to the dataset to introduce
+variability: \(x_4\), \(x_5\), \(x_6\), and \(x_7\), each generated from
+uniform distributions within specific intervals (Equation~\ref{eq-eqn4},
+Equation~\ref{eq-eqn5}, and Equation~\ref{eq-eqn6} respectively).
+Finally, these variables are combined to form the S-curve dataset \(T\).
 
-\begin{equation} \label{eq-eqn1}
+\begin{equation}\phantomsection\label{eq-eqn1}{
 x_1 = \sin(3 \pi x)
-\end{equation}
+}\end{equation}
 
-\begin{equation} \label{eq:eqn2}
+\begin{equation}\phantomsection\label{eq-eqn2}{
 x_2 = 2.0 \times U(0, 1)
-\end{equation}
+}\end{equation}
 
-\begin{equation} \label{eq:eqn3}
+\begin{equation}\phantomsection\label{eq-eqn3}{
 x_3 = \text{sign}(x) \times (\cos(3 \pi x) - 1)
-\end{equation}
+}\end{equation}
 
-\begin{equation} \label{eq:eqn4}
+\begin{equation}\phantomsection\label{eq-eqn4}{
 x_4, x_5 \sim U(-0.02, 0.02)
-\end{equation}
+}\end{equation}
 
-\begin{equation} \label{eq:eqn5}
+\begin{equation}\phantomsection\label{eq-eqn5}{
 x_6 \sim U(-0.1, 0.1)
-\end{equation}
+}\end{equation}
 
-\begin{equation} \label{eq:eqn6}
+\begin{equation}\phantomsection\label{eq-eqn6}{
 x_7 \sim U(-0.01, 0.01)
-\end{equation}
+}\end{equation}
 

### 2024-04-01 paper.qmd
Add simulate data description

@@
 
-The S-curve model, denoted as $T$, is simulated through several steps.
-First, a random variable $x$ is generated from a uniform distribution between $-\frac{\pi}{2}$ and $\frac{\pi}{2}$, representing the parameter $a$ in the S-curve model.
-Next, three variables $x_1$, $x_2$, and $x_3$ are computed based on $x$ to construct the S-curve data.
+First, a random variable $a$ is generated from a uniform distribution between $-\frac{\pi}{2}$ and $\frac{\pi}{2}$, representing the parameter $a$ is in $T$.
+Next, three variables $x_1$, $x_2$, and $x_3$ are computed based on $a$ to construct the S-curve data.
 $x_1$ is calculated as the sine of $a$ (@eq-eqn1), $x_2$ as twice the value of another uniformly distributed random variable between 0 and 1 (@eq-eqn2), and $x_3$ as the cosine of $a$ minus 1, multiplied by the sign of $a$ (@eq-eqn3).
-Additionally, four noise dimensions are introduced to the dataset to introduce variability: $x_4$, $x_5$, $x_6$, and $x_7$, each generated from uniform distributions within specific intervals (@eq-eqn4, @eq-eqn5, and @eq-eqn6 respectively).
+Additionally, four noise dimensions are introduced to the dataset to introduce noise variability: $x_4$, $x_5$, $x_6$, and $x_7$, each generated from uniform distributions within specific intervals (@eq-eqn4, @eq-eqn5, and @eq-eqn6 respectively).
 Finally, these variables are combined to form the S-curve dataset $T$.
@@
 $$
-x_1 = \sin(3 \pi x)
+x_1 = \sin(3 \pi a)
 $$ {#eq-eqn1}
@@
 $$
-x_3 = \text{sign}(x) \times (\cos(3 \pi x) - 1)
+x_3 = \text{sign}(a) \times (\cos(3 \pi a) - 1)
 $$ {#eq-eqn3}

### 2024-04-01 paper.tex
Add simulate data description

@@
 
-The S-curve model, denoted as \(T\), is simulated through several steps.
-First, a random variable \(x\) is generated from a uniform distribution
+First, a random variable \(a\) is generated from a uniform distribution
 between \(-\frac{\pi}{2}\) and \(\frac{\pi}{2}\), representing the
-parameter \(a\) in the S-curve model. Next, three variables \(x_1\),
-\(x_2\), and \(x_3\) are computed based on \(x\) to construct the
-S-curve data. \(x_1\) is calculated as the sine of \(a\)
-(Equation~\ref{eq-eqn1}), \(x_2\) as twice the value of another
-uniformly distributed random variable between 0 and 1
-(Equation~\ref{eq-eqn2}), and \(x_3\) as the cosine of \(a\) minus 1,
-multiplied by the sign of \(a\) (Equation~\ref{eq-eqn3}). Additionally,
-four noise dimensions are introduced to the dataset to introduce
-variability: \(x_4\), \(x_5\), \(x_6\), and \(x_7\), each generated from
-uniform distributions within specific intervals (Equation~\ref{eq-eqn4},
+parameter \(a\) is in \(T\). Next, three variables \(x_1\), \(x_2\), and
+\(x_3\) are computed based on \(a\) to construct the S-curve data.
+\(x_1\) is calculated as the sine of \(a\) (Equation~\ref{eq-eqn1}),
+\(x_2\) as twice the value of another uniformly distributed random
+variable between 0 and 1 (Equation~\ref{eq-eqn2}), and \(x_3\) as the
+cosine of \(a\) minus 1, multiplied by the sign of \(a\)
+(Equation~\ref{eq-eqn3}). Additionally, four noise dimensions are
+introduced to the dataset to introduce noise variability: \(x_4\),
+\(x_5\), \(x_6\), and \(x_7\), each generated from uniform distributions
+within specific intervals (Equation~\ref{eq-eqn4},
 Equation~\ref{eq-eqn5}, and Equation~\ref{eq-eqn6} respectively).
@@
 \begin{equation}\phantomsection\label{eq-eqn1}{
-x_1 = \sin(3 \pi x)
+x_1 = \sin(3 \pi a)
 }\end{equation}
@@
 \begin{equation}\phantomsection\label{eq-eqn3}{
-x_3 = \text{sign}(x) \times (\cos(3 \pi x) - 1)
+x_3 = \text{sign}(a) \times (\cos(3 \pi a) - 1)
 }\end{equation}

### 2024-04-02 paper.qmd
small change, hopefully not a conflict

@@
 Here we focus on five currently popular techniques, tSNE, UMAP, PHATE, TriMAP and PaCMAP.
-tNSE and UMAP can be considered to produce the $k-D$ minimizing the divergence between two distributions, where the distributions are modeling the inter-point distances.
+tNSE and UMAP can be considered to produce the \kD minimizing the divergence between two distributions, where the distributions are modeling the inter-point distances.
 PHATE, TriMAP and PaCMAP are examples of diffusion processes [@coifman2005] spreading to capture geometric shapes, that include both global and local structure.
@@
 
-First, a random variable $a$ is generated from a uniform distribution between $-\frac{\pi}{2}$ and $\frac{\pi}{2}$, representing the parameter $a$ is in $T$.
-Next, three variables $x_1$, $x_2$, and $x_3$ are computed based on $a$ to construct the S-curve data.
-$x_1$ is calculated as the sine of $a$ (@eq-eqn1), $x_2$ as twice the value of another uniformly distributed random variable between 0 and 1 (@eq-eqn2), and $x_3$ as the cosine of $a$ minus 1, multiplied by the sign of $a$ (@eq-eqn3).
-Additionally, four noise dimensions are introduced to the dataset to introduce noise variability: $x_4$, $x_5$, $x_6$, and $x_7$, each generated from uniform distributions within specific intervals (@eq-eqn4, @eq-eqn5, and @eq-eqn6 respectively).
-Finally, these variables are combined to form the S-curve dataset $T$.
-
-$$
-x_1 = \sin(3 \pi a)
-$$ {#eq-eqn1}
-
-$$
-x_2 = 2.0 \times U(0, 1)
-$$ {#eq-eqn2}
-
-$$
-x_3 = \text{sign}(a) \times (\cos(3 \pi a) - 1)
-$$ {#eq-eqn3}
-
-$$
-x_4, x_5 \sim U(-0.02, 0.02)
-$$ {#eq-eqn4}
-
-$$
-x_6 \sim U(-0.1, 0.1)
-$$ {#eq-eqn5}
-
-$$
-x_7 \sim U(-0.01, 0.01)
-$$ {#eq-eqn6}
-
 ### Generating the \kD layout

### 2024-04-02 paper.tex
small change, hopefully not a conflict

--- 2024-04-01 paper.tex
@@
 focus on five currently popular techniques, tSNE, UMAP, PHATE, TriMAP
-and PaCMAP. tNSE and UMAP can be considered to produce the \(k-D\)
-minimizing the divergence between two distributions, where the
+and PaCMAP. tNSE and UMAP can be considered to produce the
+\kD minimizing the divergence between two distributions, where the
 distributions are modeling the inter-point distances. PHATE, TriMAP and
@@
 $P$ & orthonormal basis, generating a $d-dimensional$ linear projection of $p$-dimensional data\\
-$T$ & true  model\\
+$y_{max}$ & maximum value of the scaled second embedding component\\
 \addlinespace
-$y_{max}$ & maximum value of the scaled second embedding component\\
 $r_1$, $r_2$ & range of the first and  second embedding component\\
@@
 $h_r$ & hexagon ratio\\
+$s$ & hexagonal size (radius of the outer circle surrounding the hexagon)\\
 \addlinespace
-$s$ & hexagonal size (radius of the outer circle surrounding the hexagon)\\
 $(x_s, \ y_s)$ & starting coordinates of the hexagonal grid\\
@@
 $d_x$, $d_y$ & horizontal spacing and adjusted vertical spacing for hexagonal coordinates\\
+$b_x$, $b_y$ & number of hexagon bins along the x and y axes\\
 \addlinespace
-$b_x$, $b_y$ & number of hexagon bins along the x and y axes\\
 $b$, $b'$ & total and non-empty hexagon bins in the grid\\
@@
 generated as follows XXX Jayani
-
-The S-curve model, denoted as \(T\), is simulated through several steps.
-First, a random variable \(x\) is generated from a uniform distribution
-between \(-\frac{\pi}{2}\) and \(\frac{\pi}{2}\), representing the
-parameter \(x\) in the S-curve model. Next, three variables \(x_1\),
-\(x_2\), and \(x_3\) are computed based on \(x\) to construct the
-S-curve data. \(x_1\) is calculated as the sine of \(x\)
-(@ref(eq:eqn2)), \(x_2\) as twice the value of another uniformly
-distributed random variable between 0 and 1, and \(x_3\) as the cosine
-of \(x\) minus 1, multiplied by the sign of \(x\). Additionally, four
-noise dimensions are introduced to the dataset to introduce variability:
-\(x_4\), \(x_5\), \(x_6\), and \(x_7\), each generated from uniform
-distributions within specific intervals. Finally, these variables are
-combined to form the S-curve dataset \(T\).
-
-\begin{equation} \label{eq-eqn1}
-x_1 = \sin(3 \pi x)
-\end{equation}
-
-\begin{equation} \label{eq:eqn2}
-x_2 = 2.0 \times U(0, 1)
-\end{equation}
-
-\begin{equation} \label{eq:eqn3}
-x_3 = \text{sign}(x) \times (\cos(3 \pi x) - 1)
-\end{equation}
-
-\begin{equation} \label{eq:eqn4}
-x_4, x_5 \sim U(-0.02, 0.02)
-\end{equation}
-
-\begin{equation} \label{eq:eqn5}
-x_6 \sim U(-0.1, 0.1)
-\end{equation}
-
-\begin{equation} \label{eq:eqn6}
-x_7 \sim U(-0.01, 0.01)
-\end{equation}
 

### 2024-04-02 paper.qmd

--- 2024-04-01 paper.qmd
@@
   \usepackage[utf8]{inputenc}
+  \usepackage{bm}
   \def\tightlist{}
@@
 The array of layouts in @fig-NLDR-variety illustrate what can emerge from the choices of method and parameters, and the random seed that initiates the computation.
-Key structures interpreted from these views suggest:
-
-- highly **separated clusters** (a, b, e, g, h) with the number ranging from 3-6, 
-- **stringy branches** (f), and 
-- **barely separated clusters** (c, d) which apparently **contradicts** the other representations.
+Key structures interpreted from these views suggest: (1) highly **separated clusters** (a, b, e, g, h) with the number ranging from 3-6; (2) **stringy branches** (f), and (3) **barely separated clusters** (c, d) which would **contradict** the other representations.
 
@@
 @tbl-notation summarises the notation used to explain the new methodology.
-$X_{n \times p} = \begin{bmatrix} \textbf{x} _{1} & \textbf{x}_ {2} & \cdots & \textbf{x}_{n} \\  \end{bmatrix}^\top$ represents the observed data matrix, that are $n$ \pD vectors sampled from some population.
+The observed data is denoted as $\mathbfit{x}_{n \times p}$ where $x_{ij}$ would indicate the $i^{th}$ observation on the $j^{th}$ variable sampled from a population $\mathbfit{X}$.
+To refer to variable $j$, we would use $X_j$.
 
@@
 
+To illustrate the method, we use $7\text{-}D$ simulated data, which we call the "S-curve".
+It is constructed by setting $X_1 = \sin(a), X_2 = U(0, 2), X_3 = \text{sign}(a) \times (\cos(a) - 1), \forall a \in [-3\pi/2, 3\pi/2]$.
+The remaining variables $X_4, X_5, X_6, X_7$ are all uniform error, with small variance.
+We would consider $T=(X_1, X_2, X_3)$ to be the true model.
+
 We need to have some math that gives precision to the paragraphs in the previous section, defining mapping from \pD to \kD to get the \kD layout. Might be written below, too (XXX Michael??)
 
-To illustrate the method, we use simulated data. The S-curve is generated as follows XXX Jayani
-
-First, a random variable $a$ is generated from a uniform distribution between $-\frac{\pi}{2}$ and $\frac{\pi}{2}$, representing the parameter $a$ is in $T$.
+<!--First, a random variable $a$ is generated from a uniform distribution between $-\frac{\pi}{2}$ and $\frac{\pi}{2}$, representing the parameter $a$ is in $T$.
 Next, three variables $x_1$, $x_2$, and $x_3$ are computed based on $a$ to construct the S-curve data.
@@
 $$ {#eq-eqn6}
+-->
 

### 2024-04-02 paper.tex

--- 2024-04-01 paper.tex
@@
 \usepackage[utf8]{inputenc}
+\usepackage{bm}
 \def\tightlist{}
@@
 seed that initiates the computation. Key structures interpreted from
-these views suggest:
-
-\begin{itemize}
-\tightlist
-\item
-  highly \textbf{separated clusters} (a, b, e, g, h) with the number
-  ranging from 3-6,
-\item
-  \textbf{stringy branches} (f), and
-\item
-  \textbf{barely separated clusters} (c, d) which apparently
-  \textbf{contradicts} the other representations.
-\end{itemize}
+these views suggest: (1) highly \textbf{separated clusters} (a, b, e, g,
+h) with the number ranging from 3-6; (2) \textbf{stringy branches} (f),
+and (3) \textbf{barely separated clusters} (c, d) which would
+\textbf{contradict} the other representations.
 
@@
 Table~\ref{tbl-notation} summarises the notation used to explain the new
-methodology.
-\(X_{n \times p} = \begin{bmatrix} \textbf{x} _{1} & \textbf{x}_ {2} & \cdots & \textbf{x}_{n} \\  \end{bmatrix}^\top\)
-represents the observed data matrix, that are \(n\) \pD vectors sampled
-from some population.
+methodology. The observed data is denoted as
+\(\mathbfit{x}_{n \times p}\) where \(x_{ij}\) would indicate the
+\(i^{th}\) observation on the \(j^{th}\) variable sampled from a
+population \(\mathbfit{X}\). To refer to variable \(j\), we would use
+\(X_j\).
 
@@
 $n, p, k$ & number of observations, variables, embedding dimension, respectively\\
-$X$ & $p$-dimensional data\\
+$\mathbfit{X}, \mathbfit{x}$ & $p$-dimensional data (population, sample)\\
 $Y$ & $k$-dimensional layout\\
-$P$ & orthonormal basis, generating a $d-dimensional$ linear projection of $p$-dimensional data\\
+$P$ & orthonormal basis, generating a $d\text{-}dimensional$ linear projection of $p$-dimensional data\\
 $T$ & true  model\\
@@
 
+To illustrate the method, we use \(7\text{-}D\) simulated data, which we
+call the ``S-curve''. It is constructed by setting
+\(X_1 = \sin(a), X_2 = U(0, 2), X_3 = \text{sign}(a) \times (\cos(a) - 1), \forall a \in [-3\pi/2, 3\pi/2]\).
+The remaining variables \(X_4, X_5, X_6, X_7\) are all uniform error,
+with small variance. We would consider \(T=(X_1, X_2, X_3)\) to be the
+true model.
+
 We need to have some math that gives precision to the paragraphs in the
@@
 \kD layout. Might be written below, too (XXX Michael??)
-
-To illustrate the method, we use simulated data. The S-curve is
-generated as follows XXX Jayani
-
-First, a random variable \(a\) is generated from a uniform distribution
-between \(-\frac{\pi}{2}\) and \(\frac{\pi}{2}\), representing the
-parameter \(a\) is in \(T\). Next, three variables \(x_1\), \(x_2\), and
-\(x_3\) are computed based on \(a\) to construct the S-curve data.
-\(x_1\) is calculated as the sine of \(a\) (Equation~\ref{eq-eqn1}),
-\(x_2\) as twice the value of another uniformly distributed random
-variable between 0 and 1 (Equation~\ref{eq-eqn2}), and \(x_3\) as the
-cosine of \(a\) minus 1, multiplied by the sign of \(a\)
-(Equation~\ref{eq-eqn3}). Additionally, four noise dimensions are
-introduced to the dataset to introduce noise variability: \(x_4\),
-\(x_5\), \(x_6\), and \(x_7\), each generated from uniform distributions
-within specific intervals (Equation~\ref{eq-eqn4},
-Equation~\ref{eq-eqn5}, and Equation~\ref{eq-eqn6} respectively).
-Finally, these variables are combined to form the S-curve dataset \(T\).
-
-\begin{equation}\phantomsection\label{eq-eqn1}{
-x_1 = \sin(3 \pi a)
-}\end{equation}
-
-\begin{equation}\phantomsection\label{eq-eqn2}{
-x_2 = 2.0 \times U(0, 1)
-}\end{equation}
-
-\begin{equation}\phantomsection\label{eq-eqn3}{
-x_3 = \text{sign}(a) \times (\cos(3 \pi a) - 1)
-}\end{equation}
-
-\begin{equation}\phantomsection\label{eq-eqn4}{
-x_4, x_5 \sim U(-0.02, 0.02)
-}\end{equation}
-
-\begin{equation}\phantomsection\label{eq-eqn5}{
-x_6 \sim U(-0.1, 0.1)
-}\end{equation}
-
-\begin{equation}\phantomsection\label{eq-eqn6}{
-x_7 \sim U(-0.01, 0.01)
-}\end{equation}
 

### 2024-04-02 paper.tex
small y for layout

@@
 $\mathbfit{X}, \mathbfit{x}$ & $p$-dimensional data (population, sample)\\
-$Y$ & $k$-dimensional layout\\
+$\mathbfit{y}$ & $k$-dimensional layout\\
 $P$ & orthonormal basis, generating a $d\text{-}dimensional$ linear projection of $p$-dimensional data\\

### 2024-04-02 paper.tex
Changed the notation y_max

@@
 \addlinespace
-$y_{max}$ & maximum value of the scaled second embedding component\\
+$y_\text{max}$ & maximum value of the scaled second embedding component\\
 $r_1$, $r_2$ & range of the first and  second embedding component\\

### 2024-04-04 paper_original.qmd
Update Starting point section

@@
 
-diff_start_df <- diff_start_df[c(1:5, 11:15,21:25, 101),]
+diff_start_df <- diff_start_df[c(2:5, 12:15, 22:25, 32:35, 42:45, 52:55, 62:65, 72:75, 82:85, 92:95, 101),]
 
@@
 
+num_bins_x_s_curve <- 6
+num_bins_y_s_curve <- 14
+hex_size_s_curve <- 0.13
+
+hb_obj_s_curve <- hex_binning(data = umap_s_curve_scaled, x = "UMAP1", 
+                      y = "UMAP2", num_bins_x = num_bins_x_s_curve, 
+                      num_bins_y = num_bins_y_s_curve, x_start = NA, y_start = NA, 
+                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_s_curve, col_start = "UMAP")
+
+## Data set with all possible centroids in the hexagonal grid
+all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_s_curve$centroids))
+
+## Generate all coordinates of hexagons
+hex_grid <- as.data.frame(do.call(cbind, hb_obj_s_curve$hex_poly))
+
+## To obtain the standardise counts within hexbins
+counts_df <- as.data.frame(do.call(cbind, hb_obj_s_curve$std_cts))
+df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
+                                             counts_df = counts_df)
+
+## remove low-density hexagons
+
+hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
+
 ## Initial x and y start
@@
   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
-  geom_text(data = df_bin_centroids, aes(x = c_x, y = c_y, label = hexID), size = 2) +
+  geom_point(data = umap_s_curve_scaled, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
+  #geom_text(data = df_bin_centroids, aes(x = c_x, y = c_y, label = hexID), size = 2) +
   scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
@@
 
+min_mse_df <- mse_df |> dplyr::filter(mse == min(mse_df$mse))
+
 hb_obj_s_curve_n <- hex_binning(data = umap_s_curve_scaled, x = "UMAP1", 
                       y = "UMAP2", num_bins_x = num_bins_x_s_curve, 
-                      num_bins_y = num_bins_y_s_curve, x_start = -0.0652, y_start = -0.195, 
+                      num_bins_y = num_bins_y_s_curve, x_start = min_mse_df$x_start[1], 
+                      y_start = min_mse_df$y_start[1], 
                       buffer_x = NA, buffer_y = NA, hex_size = hex_size_s_curve, col_start = "UMAP")
@@
   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
-  geom_text(data= df_bin_centroids_n, aes(x = c_x, y = c_y, label = hexID), size = 2) +
+  #geom_text(data= df_bin_centroids_n, aes(x = c_x, y = c_y, label = hexID), size = 2) +
+  geom_point(data = umap_s_curve_scaled, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
   scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
@@
 #| echo: false
-#| fig-cap: "(a) Visualization of bin centroids and (b) hexagonal grid, showcasing the configuration before (colored in gray) and after (colored in light orange) shifting the starting point by an identical amount applied in both x and y directions, with a shift value of $0.537285$."
+#| fig-cap: "(a) Visualization of bin centroids and (b) hexagonal grid, showcasing the configuration when the starting point defined at $(-0.113, -0.098)$ (colored in gray) and the starting point defined at $(-0.075, -0.195)$ (colored in light orange)."
 #| label: fig-scurveshifthexgridsexp
@@
 #| echo: false
-#| fig-cap: "Hexbin plots shows the distribution of points before and after shifting the starting point of the hexagonal grid generated for UMAP applied to the training S-curve dataset.
-(a) Hexbin plot before shifting, and (b) Hexbin plot after shifting (identical shift applied in both x and y directions, with a shift value of $0.537285$)."
+#| fig-cap: "Hexbin plots shows the distribution of points before and after changing the starting point of the hexagonal grid generated for UMAP applied to the training S-curve dataset.
+(a) Hexbin plot when the starting point defined at $(-0.113, -0.098)$, and (b) Hexbin plot after changing the starting point to $(-0.075, -0.195)$."
 #| label: fig-scurveshifthexgrids
@@
 To investigate which is the best representation to visualize the original data or all NLDR methods provide equivalent information, we visualize the model-in-data space.
-Models from all NLDR methods show five well-separated clusters (see @fig-gau1-sc, @fig-gau2-sc, @fig-gau3-sc, @fig-gau4-sc, and @fig-gau5-sc).
+Models from all NLDR methods show five well-separated clusters (see @fig-gau1_sc, @fig-gau2_sc, @fig-gau3_sc, @fig-gau4_sc, and @fig-gau5_sc).
 This suggests that for the five Gaussian cluster dataset, all NLDR methods effectively preserve the global structure.
-tSNE, UMAP, and PHATE display clusters with varying densities, indicating their ability to capture within-cluster variation (see @fig-gau1-sc, @fig-gau2-sc, and @fig-gau3-sc).
-On the other hand, both TriMAP and PaCMAP show clusters with flat surfaces, suggesting a failure to capture within-cluster variation (see @fig-gau4-sc and @fig-gau5-sc).
+tSNE, UMAP, and PHATE display clusters with varying densities, indicating their ability to capture within-cluster variation (see @fig-gau1_sc, @fig-gau2_sc, and @fig-gau3_sc).
+On the other hand, both TriMAP and PaCMAP show clusters with flat surfaces, suggesting a failure to capture within-cluster variation (see @fig-gau4_sc and @fig-gau5_sc).
 Therefore, TriMAP and PaCMAP do not capture the local structure as effectively as other methods.
@@
 
-::: {#fig-gau1-sc layout-ncol="3" fig-pos="H"}
-![](figures/five_gau_clusters/sc_tsne_1.png){#fig-gau1-sc1}
-
-![](figures/five_gau_clusters/sc_tsne_2.png){#fig-gau1-sc2}
-
-![](figures/five_gau_clusters/sc_tsne_3.png){#fig-gau1-sc3}
+::: {#fig-gau1_sc layout-ncol="3" fig-pos="H"}
+![](figures/five_gau_clusters/sc_tsne_1.png){#fig-gau1_sc1}
+
+![](figures/five_gau_clusters/sc_tsne_2.png){#fig-gau1_sc2}
+
+![](figures/five_gau_clusters/sc_tsne_3.png){#fig-gau1_sc3}
 
@@
 
-::: {#fig-gau2-sc layout-ncol="3" fig-pos="H"}
-![](figures/five_gau_clusters/sc_umap_1.png){#fig-gau2-sc1}
-
-![](figures/five_gau_clusters/sc_umap_2.png){#fig-gau2-sc2}
-
-![](figures/five_gau_clusters/sc_umap_3.png){#fig-gau2-sc3}
+::: {#fig-gau2_sc layout-ncol="3" fig-pos="H"}
+![](figures/five_gau_clusters/sc_umap_1.png){#fig-gau2_sc1}
+
+![](figures/five_gau_clusters/sc_umap_2.png){#fig-gau2_sc2}
+
+![](figures/five_gau_clusters/sc_umap_3.png){#fig-gau2_sc3}
 
@@
 
-::: {#fig-gau3-sc layout-ncol="3" fig-pos="H"}
-![](figures/five_gau_clusters/sc_phate_1.png){#fig-gau3-sc1}
-
-![](figures/five_gau_clusters/sc_phate_2.png){#fig-gau3-sc2}
-
-![](figures/five_gau_clusters/sc_phate_3.png){#fig-gau3-sc3}
+::: {#fig-gau3_sc layout-ncol="3" fig-pos="H"}
+![](figures/five_gau_clusters/sc_phate_1.png){#fig-gau3_sc1}
+
+![](figures/five_gau_clusters/sc_phate_2.png){#fig-gau3_sc2}
+
+![](figures/five_gau_clusters/sc_phate_3.png){#fig-gau3_sc3}
 
@@
 
-::: {#fig-gau4-sc layout-ncol="3" fig-pos="H"}
-![](figures/five_gau_clusters/sc_trimap_1.png){#fig-gau4-sc1}
-
-![](figures/five_gau_clusters/sc_trimap_2.png){#fig-gau4-sc2}
-
-![](figures/five_gau_clusters/sc_trimap_3.png){#fig-gau4-sc3}
+::: {#fig-gau4_sc layout-ncol="3" fig-pos="H"}
+![](figures/five_gau_clusters/sc_trimap_1.png){#fig-gau4_sc1}
+
+![](figures/five_gau_clusters/sc_trimap_2.png){#fig-gau4_sc2}
+
+![](figures/five_gau_clusters/sc_trimap_3.png){#fig-gau4_sc3}
 
@@
 
-::: {#fig-gau5-sc layout-ncol="3" fig-pos="H"}
-![](figures/five_gau_clusters/sc_pacmap_1.png){#fig-gau5-sc1}
-
-![](figures/five_gau_clusters/sc_pacmap_2.png){#fig-gau5-sc2}
-
-![](figures/five_gau_clusters/sc_pacmap_3.png){#fig-gau5-sc3}
+::: {#fig-gau5_sc layout-ncol="3" fig-pos="H"}
+![](figures/five_gau_clusters/sc_pacmap_1.png){#fig-gau5_sc1}
+
+![](figures/five_gau_clusters/sc_pacmap_2.png){#fig-gau5_sc2}
+
+![](figures/five_gau_clusters/sc_pacmap_3.png){#fig-gau5_sc3}
 
@@
 To determine whether the UMAP representation with the parameter choice suggested by @chen2023 preserves the original data structure, we visualize the model constructed with UMAP overlaid on the high-D data (model-in-data space).
-The figures in @fig-pbmc1-sc show three well-separated clusters, indicating that the suggested UMAP representation preserves the global structure (see @fig-umapauthor).
-However, as shown in @fig-pbmc1-sc, these clusters are close to each other in high-D space.
-Also, nonlinear continuity patterns and high-density patches within the clusters are observed (see @fig-pbmc1-sc).
+The figures in @fig-pbmc1_sc show three well-separated clusters, indicating that the suggested UMAP representation preserves the global structure (see @fig-umapauthor).
+However, as shown in @fig-pbmc1_sc, these clusters are close to each other in high-D space.
+Also, nonlinear continuity patterns and high-density patches within the clusters are observed (see @fig-pbmc1_sc).
 Therefore, the suggested UMAP representation (see @fig-umapauthor) does not accurately preserve the local structure of the PBMC3k dataset.
@@
 
-::: {#fig-pbmc1-sc layout-ncol="3" fig-pos="H"}
-![](figures/pbmc3k/sc_1.png){#fig-pbmc1-sc1}
-
-![](figures/pbmc3k/sc_2.png){#fig-pbmc1-sc2}
-
-![](figures/pbmc3k/sc_3.png){#fig-pbmc1-sc3}
+::: {#fig-pbmc1_sc layout-ncol="3" fig-pos="H"}
+![](figures/pbmc3k/sc_1.png){#fig-pbmc1_sc1}
+
+![](figures/pbmc3k/sc_2.png){#fig-pbmc1_sc2}
+
+![](figures/pbmc3k/sc_3.png){#fig-pbmc1_sc3}
 
@@
 Next, the model is fitted for tSNE, and the resulting model is visualized in the data space.
-The model shows some quirks, as shown in @fig-pbmc2-sc.
+The model shows some quirks, as shown in @fig-pbmc2_sc.
 The large cluster is divided into three subclusters, which are in close to each other.
@@
 
-::: {#fig-pbmc2-sc layout-ncol="3" fig-pos="H"}
-![](figures/pbmc3k/sc_4.png){#fig-pbmc2-sc4}
-
-![](figures/pbmc3k/sc_5.png){#fig-pbmc2-sc5}
-
-![](figures/pbmc3k/sc_6.png){#fig-pbmc2-sc6}
+::: {#fig-pbmc2_sc layout-ncol="3" fig-pos="H"}
+![](figures/pbmc3k/sc_4.png){#fig-pbmc2_sc4}
+
+![](figures/pbmc3k/sc_5.png){#fig-pbmc2_sc5}
+
+![](figures/pbmc3k/sc_6.png){#fig-pbmc2_sc6}
 
@@
 
-According to @fig-mnist1-sc1, the nonlinear continuity structure observed in the 2D representation of PaCMAP is also visible when visualizing the model overlaid on the data space.
+According to @fig-mnist1_sc1, the nonlinear continuity structure observed in the 2D representation of PaCMAP is also visible when visualizing the model overlaid on the data space.
 This provides evidence that the model accurately captures the structure of the high-D data.
-Also, the model has a twisted pattern within the nonlinear structure, which is an addition pattern of data that cannot be seen in the 2D representation (see @fig-mnist1-sc2).
+Also, the model has a twisted pattern within the nonlinear structure, which is an addition pattern of data that cannot be seen in the 2D representation (see @fig-mnist1_sc2).
 
@@
 In a model that effectively preserves local structure, there shouldn't be any long edges in high-D space.
-However, as shown in @fig-mnist1-sc3, some long edges exist in high-D space that are not identified as long edges in the 2D representation.
+However, as shown in @fig-mnist1_sc3, some long edges exist in high-D space that are not identified as long edges in the 2D representation.
 
@@
 
-::: {#fig-mnist1-sc layout-ncol="3" fig-pos="H"}
-![](figures/mnist/sc_1.png){#fig-mnist1-sc1}
-
-![](figures/mnist/sc_2.png){#fig-mnist1-sc2}
-
-![](figures/mnist/sc_3.png){#fig-mnist1-sc3}
+::: {#fig-mnist1_sc layout-ncol="3" fig-pos="H"}
+![](figures/mnist/sc_1.png){#fig-mnist1_sc1}
+
+![](figures/mnist/sc_2.png){#fig-mnist1_sc2}
+
+![](figures/mnist/sc_3.png){#fig-mnist1_sc3}
 

### 2024-04-04 paper_original.tex
Corrected ref error
Update Starting point section

@@
 starting point defined at \((-0.113, -0.098)\) (colored in gray) and the
-starting point defined at \((-0.025, -0.108)\) (colored in light
+starting point defined at \((-0.075, -0.195)\) (colored in light
 orange).}
@@
 \((-0.113, -0.098)\), and (b) Hexbin plot after changing the starting
-point to \((-0.025, -0.108)\).}
+point to \((-0.075, -0.195)\).}
 
@@
 
-\subcaption{\label{fig-gau1\_sc1}}
+\subcaption{\label{fig-gau1_sc1}}
 
@@
 
-\subcaption{\label{fig-gau1\_sc2}}
+\subcaption{\label{fig-gau1_sc2}}
 
@@
 
-\subcaption{\label{fig-gau1\_sc3}}
+\subcaption{\label{fig-gau1_sc3}}
 
@@
 
-\subcaption{\label{fig-gau2\_sc1}}
+\subcaption{\label{fig-gau2_sc1}}
 
@@
 
-\subcaption{\label{fig-gau2\_sc2}}
+\subcaption{\label{fig-gau2_sc2}}
 
@@
 
-\subcaption{\label{fig-gau2\_sc3}}
+\subcaption{\label{fig-gau2_sc3}}
 
@@
 
-\subcaption{\label{fig-gau3\_sc1}}
+\subcaption{\label{fig-gau3_sc1}}
 
@@
 
-\subcaption{\label{fig-gau3\_sc2}}
+\subcaption{\label{fig-gau3_sc2}}
 
@@
 
-\subcaption{\label{fig-gau3\_sc3}}
+\subcaption{\label{fig-gau3_sc3}}
 
@@
 
-\subcaption{\label{fig-gau4\_sc1}}
+\subcaption{\label{fig-gau4_sc1}}
 
@@
 
-\subcaption{\label{fig-gau4\_sc2}}
+\subcaption{\label{fig-gau4_sc2}}
 
@@
 
-\subcaption{\label{fig-gau4\_sc3}}
+\subcaption{\label{fig-gau4_sc3}}
 
@@
 
-\subcaption{\label{fig-gau5\_sc1}}
+\subcaption{\label{fig-gau5_sc1}}
 
@@
 
-\subcaption{\label{fig-gau5\_sc2}}
+\subcaption{\label{fig-gau5_sc2}}
 
@@
 
-\subcaption{\label{fig-gau5\_sc3}}
+\subcaption{\label{fig-gau5_sc3}}
 
@@
 
-\subcaption{\label{fig-pbmc1\_sc1}}
+\subcaption{\label{fig-pbmc1_sc1}}
 
@@
 
-\subcaption{\label{fig-pbmc1\_sc2}}
+\subcaption{\label{fig-pbmc1_sc2}}
 
@@
 
-\subcaption{\label{fig-pbmc1\_sc3}}
+\subcaption{\label{fig-pbmc1_sc3}}
 
@@
 
-\subcaption{\label{fig-pbmc2\_sc4}}
+\subcaption{\label{fig-pbmc2_sc4}}
 
@@
 
-\subcaption{\label{fig-pbmc2\_sc5}}
+\subcaption{\label{fig-pbmc2_sc5}}
 
@@
 
-\subcaption{\label{fig-pbmc2\_sc6}}
+\subcaption{\label{fig-pbmc2_sc6}}
 
@@
 
-\subcaption{\label{fig-mnist1\_sc1}}
+\subcaption{\label{fig-mnist1_sc1}}
 
@@
 
-\subcaption{\label{fig-mnist1\_sc2}}
+\subcaption{\label{fig-mnist1_sc2}}
 
@@
 
-\subcaption{\label{fig-mnist1\_sc3}}
+\subcaption{\label{fig-mnist1_sc3}}
 

### 2024-04-05 paper.qmd
Add first draft at notation section

@@
 Abstractly, we can also imagine that the positioning of points in 2D represents fitted values, that will have some prescribed position in \pD that can be compared with the observed value.
-
 <!--### Notation -->
@@
 
-We need to have some math that gives precision to the paragraphs in the previous section, defining mapping from \pD to \kD to get the \kD layout. Might be written below, too (XXX Michael??)
+We define the NLDR model as a function that maps \pD to \kD, and takes some (hyper)parameters $\vec{\theta}$. That is,
+
+$$g : \mathbb{R}^p \rightarrow \mathbb{R}^k.$$
+
+The parameter vector, $\vec{\theta}$, depends on the choice of $g$, and can be considered part of 'model fitting' in the traditional sense.
+Common choices for $g$ include tSNE, uMAP, PACMAP, or PHATE, although in theory any function that maps the entire space $\mathbb{R}^p$ to a subset of $\mathbb{R}^k$ is suitable.
+Any input requirements for the data (such as normalisation, or preprocessing through the use of PCA or similar) is considered part of the function $g$.
+
+Next, we create a model of the \kD representation of the data using hexagonal binning.
+The aim of this process is to reduce the noise in the original data that may be preserved by $g$.
+Define a hexagonal grid across the \kD space, defined by the height and width of the hexagon ($h_b$ and $w_b$ respectively), the aspect ratio of the hexagon ($h_r$), and the starting coordinates of the hexagonal grid $(x_s, y_s)$.
+We deliberately separate out the creation of the hexagonal grid from the mapping of points on the grid.
+
+With the grid defined, the \kD points will belong to a hexagonal bin. That is, for each $y \in \mathbfit{Y}$, we can (uniquely) identify the hexagon that the point belongs to. Let,
+
+$$h : \mathbb{R}^k \rightarrow \mathbb{R}^k, $$
+
+be the function that maps a point in \kD space to it's hexagonal centroid. It follows that $h \circ g : \mathbb{R}^p \rightarrow \mathbb{R}^k$ maps each point in $\mathbfit{X}$ to a \kD centroid.
+
+The final step of the algorithm is _lifting_ the \kD model back to \pD, completing the 'model-in-data-space' picture.
+For this, define the set $H^k$ to be the set of points that belong to centroid $k$.
+That is, $H^k = \{ y \; | \; g(y) = (h_x^k, h_y^k)\}$.
+We use the \pD Euclidean mean of the points in $H^k$ to map the centroid $(h_x^k, h_y^k)$ to a point in \pD.
+Let the $i$-th component of the \pD mean be
+
+$$f_i = \frac{1}{\left|H^k\right|}\sum_{y \in H^k} y_i,$$
+
+with corresponding \pD vector $\vec{f_i}$. Then, $f(h\circ g)(x)$ maps a \pD point to the \pD model estimate, completing the model cycle.
 
@@
 -->
-
-### Generating the \kD layout
-
-The usual procedures are followed to produce a \kD layout, using the preferred method, t-SNE, UMAP, ... The data may need to be pre-processed by standardising or transforming to principal components.
-
-### Constructing the 2D model
 

### 2024-04-05 paper.tex
Add first draft at notation section

@@
 
-We need to have some math that gives precision to the paragraphs in the
-previous section, defining mapping from \pD to \kD to get the
-\kD layout. Might be written below, too (XXX Michael??)
+We define the NLDR model as a function that maps \pD to \kD, and takes
+some (hyper)parameters \(\vec{\theta}\). That is,
 
-\subsection{\texorpdfstring{Generating the
-\kD layout}{Generating the layout}}\label{generating-the-layout}
+\[g : \mathbb{R}^p \rightarrow \mathbb{R}^k.\]
 
-The usual procedures are followed to produce a \kD layout, using the
-preferred method, t-SNE, UMAP, \ldots{} The data may need to be
-pre-processed by standardising or transforming to principal components.
+The parameter vector, \(\vec{\theta}\), depends on the choice of \(g\),
+and can be considered part of `model fitting' in the traditional sense.
+Common choices for \(g\) include tSNE, uMAP, PACMAP, or PHATE, although
+in theory any function that maps the entire space \(\mathbb{R}^p\) to a
+subset of \(\mathbb{R}^k\) is suitable. Any input requirements for the
+data (such as normalisation, or preprocessing through the use of PCA or
+similar) is considered part of the function \(g\).
 
-\subsection{Constructing the 2D model}\label{constructing-the-2d-model}
+Next, we create a model of the \kD representation of the data using
+hexagonal binning. The aim of this process is to reduce the noise in the
+original data that may be preserved by \(g\). Define a hexagonal grid
+across the \kD space, defined by the height and width of the hexagon
+(\(h_b\) and \(w_b\) respectively), the aspect ratio of the hexagon
+(\(h_r\)), and the starting coordinates of the hexagonal grid
+\((x_s, y_s)\). We deliberately separate out the creation of the
+hexagonal grid from the mapping of points on the grid.
+
+With the grid defined, the \kD points will belong to a hexagonal bin.
+That is, for each \(y \in \mathbfit{Y}\), we can (uniquely) identify the
+hexagon that the point belongs to. Let,
+
+\[h : \mathbb{R}^k \rightarrow \mathbb{R}^k, \]
+
+be the function that maps a point in \kD space to it's hexagonal
+centroid. It follows that
+\(h \circ g : \mathbb{R}^p \rightarrow \mathbb{R}^k\) maps each point in
+\(\mathbfit{X}\) to a \kD centroid.
+
+The final step of the algorithm is \emph{lifting} the \kD model back to
+\pD, completing the `model-in-data-space' picture. For this, define the
+set \(H^k\) to be the set of points that belong to centroid \(k\). That
+is, \(H^k = \{ y \; | \; g(y) = (h_x^k, h_y^k)\}\). We use the
+\pD Euclidean mean of the points in \(H^k\) to map the centroid
+\((h_x^k, h_y^k)\) to a point in \pD. Let the \(i\)-th component of the
+\pD mean be
+
+\[f_i = \frac{1}{\left|H^k\right|}\sum_{y \in H^k} y_i,\]
+
+with corresponding \pD vector \(\vec{f_i}\). Then, \(f(h\circ g)(x)\)
+maps a \pD point to the \pD model estimate, completing the model cycle.
 

### 2024-04-05 paper.qmd
Change uMAp to UMAP, add TriMAP to sentence

@@
 The parameter vector, $\vec{\theta}$, depends on the choice of $g$, and can be considered part of 'model fitting' in the traditional sense.
-Common choices for $g$ include tSNE, uMAP, PACMAP, or PHATE, although in theory any function that maps the entire space $\mathbb{R}^p$ to a subset of $\mathbb{R}^k$ is suitable.
+Common choices for $g$ include tSNE, UMAP, PHATE, TriMAP, or PaCMAP, although in theory any function that maps the entire space $\mathbb{R}^p$ to a subset of $\mathbb{R}^k$ is suitable.
 Any input requirements for the data (such as normalisation, or preprocessing through the use of PCA or similar) is considered part of the function $g$.
@@
 
-The final step of the algorithm is _lifting_ the \kD model back to \pD, completing the 'model-in-data-space' picture.
+The final step of the algorithm is _lifting_ the \kD model back to \pD, completing the *model-in-the-data-space* picture.
 For this, define the set $H^k$ to be the set of points that belong to centroid $k$.

### 2024-04-05 paper.tex
Change uMAp to UMAP, add TriMAP to sentence

@@
 and can be considered part of `model fitting' in the traditional sense.
-Common choices for \(g\) include tSNE, uMAP, PACMAP, or PHATE, although
-in theory any function that maps the entire space \(\mathbb{R}^p\) to a
-subset of \(\mathbb{R}^k\) is suitable. Any input requirements for the
-data (such as normalisation, or preprocessing through the use of PCA or
-similar) is considered part of the function \(g\).
+Common choices for \(g\) include tSNE, UMAP, PHATE, TriMAP, or PaCMAP,
+although in theory any function that maps the entire space
+\(\mathbb{R}^p\) to a subset of \(\mathbb{R}^k\) is suitable. Any input
+requirements for the data (such as normalisation, or preprocessing
+through the use of PCA or similar) is considered part of the function
+\(g\).
 
@@
 The final step of the algorithm is \emph{lifting} the \kD model back to
-\pD, completing the `model-in-data-space' picture. For this, define the
-set \(H^k\) to be the set of points that belong to centroid \(k\). That
-is, \(H^k = \{ y \; | \; g(y) = (h_x^k, h_y^k)\}\). We use the
-\pD Euclidean mean of the points in \(H^k\) to map the centroid
+\pD, completing the \emph{model-in-the-data-space} picture. For this,
+define the set \(H^k\) to be the set of points that belong to centroid
+\(k\). That is, \(H^k = \{ y \; | \; g(y) = (h_x^k, h_y^k)\}\). We use
+the \pD Euclidean mean of the points in \(H^k\) to map the centroid
 \((h_x^k, h_y^k)\) to a point in \pD. Let the \(i\)-th component of the

### 2024-04-05 paper.qmd
Change normalisation spelling to American

@@
 Common choices for $g$ include tSNE, UMAP, PHATE, TriMAP, or PaCMAP, although in theory any function that maps the entire space $\mathbb{R}^p$ to a subset of $\mathbb{R}^k$ is suitable.
-Any input requirements for the data (such as normalisation, or preprocessing through the use of PCA or similar) is considered part of the function $g$.
+Any input requirements for the data (such as normalization, or preprocessing through the use of PCA or similar) is considered part of the function $g$.
 

### 2024-04-07 paper_original.qmd
Added handwritten images that have large errors within and outside the nonlinear structure

@@
 hex_size_mnist <- 0.06
+x_start_mnist <- -0.05773503
+y_start_mnist <- -0.010
 
@@
                       y = "PaCMAP2", num_bins_x = num_bins_x_mnist, 
-                      num_bins_y = num_bins_y_mnist, x_start = NA, y_start = NA, 
+                      num_bins_y = num_bins_y_mnist, x_start = x_start_mnist, y_start = y_start_mnist, 
                       buffer_x = NA, buffer_y = NA, hex_size = hex_size_mnist, col_start = "PaCMAP")
@@
                                  num_bins_y = num_bins_y_mnist, 
-                                 x_start = NA, y_start = NA, 
+                                 x_start = x_start_mnist, y_start = y_start_mnist, 
                                  buffer_x = NA, buffer_y = NA, 
@@
 <!--images that occur large error-->
-```{r}
-#| warning: false
-#| echo: false
-
-img_error <- error_df |> 
-  dplyr::filter(total >= 10) |> 
-  dplyr::pull(ID)
+<!-- within the nonlinear structure-->
+```{r}
+#| warning: false
+#| echo: false
 
@@
   mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
-  filter(instance %in% img_error)
+  filter(instance %in% c(
+    4699, 4785, 5173, 6671, 2186, 4636, 4882, 773, 2825, 6169, 6948, 5987, 1282, 7715, 622, 2989, 6912,
+    4370, 1227, 4042, 2444, 836,
+    1654))
 
 theme_set(theme_light())
-imge_error_sample <- pixels_gathered |> 
+imge_error_sample_within <- pixels_gathered |> 
   ggplot(aes(x, y, fill = value)) +  
   geom_tile() +  
-  facet_wrap(~ instance, ncol = 11) +
+  facet_wrap(~ factor(instance, levels = c(
+    4699, 4785, 5173, 6671, 2186, 4636, 4882, 773, 2825, 6169, 6948, 5987, 1282, 7715, 622, 2989, 6912,
+    4370, 1227, 4042, 2444, 836,
+    1654)), ncol = 6) +
   theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
@@
 #| echo: false
-#| fig-cap: "Images of handwritten digit 1 which occur large model error."
+#| fig-cap: "Images of handwritten digit 1 which occur large model error within the nonlinear strcuture."
+#| label: fig-sampleimgerrorwithin
+#| out-width: 100%
+#| fig-pos: H
+
+imge_error_sample_within
+```
+
+<!--outside the nonlinear structure-->
+```{r}
+#| warning: false
+#| echo: false
+
+img_error <- error_df |> 
+  dplyr::filter(total >= 8) |> 
+  dplyr::filter((!ID %in% c(
+    4699, 4785, 5173, 6671, 2186, 4636, 4882, 773, 2825, 6169, 6948, 5987, 1282, 7715, 622, 2989, 6912,
+    4370, 1227, 4042, 2444, 836,
+    1654))) |>
+  dplyr::pull(ID)
+
+pixels_gathered <-  mnist_data |>
+  mutate(instance = row_number()) |>
+  gather(pixel, value, -Label, -instance) |>
+  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
+  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
+  filter(instance %in% img_error)
+
+theme_set(theme_light())
+imge_error_sample <- pixels_gathered |> 
+  ggplot(aes(x, y, fill = value)) +  
+  geom_tile() +  
+  facet_wrap(~ instance, ncol = 11) +
+  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+              axis.title.x = element_blank(), axis.title.y = element_blank(),
+              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
+        legend.title = element_text(size=5), #change legend title font size
+        legend.text = element_text(size=4),
+         legend.key.height = unit(0.25, 'cm'),
+         legend.key.width = unit(0.25, 'cm'),
+        strip.text = element_blank(),
+        aspect.ratio = 1)
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "Images of handwritten digit 1 which occur large model error outside the nonlinear structure."
 #| label: fig-sampleimgerror

### 2024-04-07 paper_original.tex
Added handwritten images that have large errors within and outside the nonlinear structure

@@
 
+\includegraphics[width=1\textwidth,height=\textheight]{paper_original_files/figure-pdf/fig-sampleimgerrorwithin-1.pdf}
+
+}
+
+\caption{\label{fig-sampleimgerrorwithin}Images of handwritten digit 1
+which occur large model error within the nonlinear strcuture.}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+\centering{
+
 \includegraphics[width=1\textwidth,height=\textheight]{paper_original_files/figure-pdf/fig-sampleimgerror-1.pdf}
@@
 \caption{\label{fig-sampleimgerror}Images of handwritten digit 1 which
-occur large model error.}
+occur large model error outside the nonlinear structure.}
 

### 2024-04-09 paper.qmd
revision of kD model

@@
   \usepackage{setspace}
-  \newcommand\pD{$p\text{-}D$\ }
-  \newcommand\kD{$k\text{-}D$\ }
-  \newcommand\dD{$d\text{-}D$\ }
+  \newcommand\pD{$p\text{-}D$}
+  \newcommand\kD{$k\text{-}D$}
+  \newcommand\dD{$d\text{-}D$}
 ---
@@
 
-At first glance, thinking of NLDR as a model fitted to the data might seem strange.
-It is a model in the sense that it is a "a simplified representation or abstraction of a system, process, or phenomenon in the real world".
+At first glance, thinking of NLDR as a modeling technique might seem strange.
+It is a simplified representation or abstraction of a system, process, or phenomenon in the real world.
 The \pD   observations are the realization of the phenomenon, and the \kD   NLDR layout is the simplified representation.
 From a statistical perspective we can consider the distances between points in the \kD   layout to be variance that the model explains, and the (relative) difference with their distances in \pD   is the error, or unexplained variance.
-Abstractly, we can also imagine that the positioning of points in 2D represents fitted values, that will have some prescribed position in \pD that can be compared with the observed value.
+We can also imagine that the positioning of points in 2D represent the fitted values, that will have some prescribed position in \pD   that can be compared with their observed values.
+This is the conceptual framework underlying the more formal versions of factor analysis [@cfa69] and multidimensional scaling (MDS) [@borg2005].
+(Note that, for this thinking the full \pD   data needs to be available, not just the interpoint distances.)
 <!--### Notation -->
 
-@tbl-notation summarises the notation used to explain the new methodology.
+<!-- @tbl-notation summarises the notation used to explain the new methodology.
 The observed data is denoted as $\mathbfit{x}_{n \times p}$ where $x_{ij}$ would indicate the $i^{th}$ observation on the $j^{th}$ variable sampled from a population $\mathbfit{X}$.
-To refer to variable $j$, we would use $X_j$.
+To refer to variable $j$, we would use $X_j$.-->
 
@@
 
-To illustrate the method, we use $7\text{-}D$ simulated data, which we call the "S-curve".
-It is constructed by setting $X_1 = \sin(a), X_2 = U(0, 2), X_3 = \text{sign}(a) \times (\cos(a) - 1), \forall a \in [-3\pi/2, 3\pi/2]$.
-The remaining variables $X_4, X_5, X_6, X_7$ are all uniform error, with small variance.
-We would consider $T=(X_1, X_2, X_3)$ to be the true model.
+We define the NLDR model as function $g\text{:}~ \mathbb{R}^p \rightarrow \mathbb{R}^k$, with (hyper-)parameters $\mathbfit{\theta}$.
+The parameters, $\mathbfit{\theta}$, depend on the choice of $g$, and can be considered part of model fitting in the traditional sense.
+Common choices for $g$ include functions used in tSNE, UMAP, PHATE, TriMAP, PaCMAP, or MDS, although in theory any function that maps the entire space $\mathbb{R}^p$ to a subset of $\mathbb{R}^k$ is suitable.
+<!--Any input requirements for the data (such as normalization, or preprocessing through the use of PCA or similar) is considered part of the function $g$.-->
 
-We define the NLDR model as a function that maps \pD to \kD, and takes some (hyper)parameters $\vec{\theta}$. That is,
+With our goal being to make a representation of this 2D layout that can be lifted into high-dimensional space, the layout needs to be augmented to include neighbour information.
+A simple approach would be to triangulate the points and add edges.
+A more stable approach is to first hexagonally bin the data, and connect the bin centroids.
+This process serves reduce some noise in the resulting surface shown in \pD.
 
-$$g : \mathbb{R}^p \rightarrow \mathbb{R}^k.$$
-
-The parameter vector, $\vec{\theta}$, depends on the choice of $g$, and can be considered part of 'model fitting' in the traditional sense.
-Common choices for $g$ include tSNE, UMAP, PHATE, TriMAP, or PaCMAP, although in theory any function that maps the entire space $\mathbb{R}^p$ to a subset of $\mathbb{R}^k$ is suitable.
-Any input requirements for the data (such as normalization, or preprocessing through the use of PCA or similar) is considered part of the function $g$.
-
-Next, we create a model of the \kD representation of the data using hexagonal binning.
-The aim of this process is to reduce the noise in the original data that may be preserved by $g$.
-Define a hexagonal grid across the \kD space, defined by the height and width of the hexagon ($h_b$ and $w_b$ respectively), the aspect ratio of the hexagon ($h_r$), and the starting coordinates of the hexagonal grid $(x_s, y_s)$.
-We deliberately separate out the creation of the hexagonal grid from the mapping of points on the grid.
-
-With the grid defined, the \kD points will belong to a hexagonal bin. That is, for each $y \in \mathbfit{Y}$, we can (uniquely) identify the hexagon that the point belongs to. Let,
-
-$$h : \mathbb{R}^k \rightarrow \mathbb{R}^k, $$
-
-be the function that maps a point in \kD space to it's hexagonal centroid. It follows that $h \circ g : \mathbb{R}^p \rightarrow \mathbb{R}^k$ maps each point in $\mathbfit{X}$ to a \kD centroid.
-
-The final step of the algorithm is _lifting_ the \kD model back to \pD, completing the *model-in-the-data-space* picture.
-For this, define the set $H^k$ to be the set of points that belong to centroid $k$.
-That is, $H^k = \{ y \; | \; g(y) = (h_x^k, h_y^k)\}$.
-We use the \pD Euclidean mean of the points in $H^k$ to map the centroid $(h_x^k, h_y^k)$ to a point in \pD.
-Let the $i$-th component of the \pD mean be
-
-$$f_i = \frac{1}{\left|H^k\right|}\sum_{y \in H^k} y_i,$$
-
-with corresponding \pD vector $\vec{f_i}$. Then, $f(h\circ g)(x)$ maps a \pD point to the \pD model estimate, completing the model cycle.
+Define a hexagonal grid across the \kD space, using hexagons with fixed height ($h_b$), width ($w_b$).
+The starting coordinates $(x_s, y_s)$ mark the lower left of the grid.
+<!--We deliberately separate out the creation of the hexagonal grid from the mapping of points on the grid.-->
+Each of the \kD points will belong to a hexagonal bin.
+That is, for each $y \in \mathbfit{Y}$, we can (uniquely) identify the hexagon that the point belongs to.
+Let $h : \mathbb{R}^k \rightarrow \mathbb{R}^k$ be the function that maps a point in \kD space to it's hexagonal centroid ($C^{(k)}$).
+<!--It follows that $h \circ g : \mathbb{R}^p \rightarrow \mathbb{R}^k$ maps each point in $\mathbfit{X}$ to a \kD centroid.-->
 
@@
 
+To illustrate the method, we use $7\text{-}D$ simulated data, which we call the "S-curve".
+It is constructed by setting $X_1 = \sin(a), X_2 = U(0, 2), X_3 = \text{sign}(a) \times (\cos(a) - 1), \forall a \in [-3\pi/2, 3\pi/2]$.
+The remaining variables $X_4, X_5, X_6, X_7$ are all uniform error, with small variance.
+We would consider $T=(X_1, X_2, X_3)$ to be the true model.
+
 ### Displaying the model in \pD
+
+The final step of the algorithm is _lifting_ the \kD model back to \pD, completing the *model-in-the-data-space* picture.
+For this, define the set $H^k$ to be the set of points that belong to centroid $k$.
+That is, $H^k = \{ y \; | \; g(y) = (h_x^k, h_y^k)\}$.
+We use the \pD Euclidean mean of the points in $H^k$ to map the centroid $(h_x^k, h_y^k)$ to a point in \pD.
+Let the $i$-th component of the \pD mean be
+
+$$f_i = \frac{1}{\left|H^k\right|}\sum_{y \in H^k} y_i,$$
+
+with corresponding \pD vector $\vec{f_i}$. Then, $f(h\circ g)(x)$ maps a \pD point to the \pD model estimate, completing the model cycle.
 
@@
 - Human behaviour, the desire to have more certainty, and a tendency to prefer the well-separated views
+- Predicting new observations in $k$-D
+- Extending layouts beyond $k$-D, when 2D is clearly inadequate.
 - Diagnostic app to explore differences in distances

### 2024-04-09 paper.tex
Change normalisation spelling to American
revision of kD model

--- 2024-04-02 paper.tex
@@
 \usepackage{setspace}
-\newcommand\pD{$p\text{-}D$\ }
-\newcommand\kD{$k\text{-}D$\ }
-\newcommand\dD{$d\text{-}D$\ }
+\newcommand\pD{$p\text{-}D$}
+\newcommand\kD{$k\text{-}D$}
+\newcommand\dD{$d\text{-}D$}
 \makeatletter
@@
 
-Historically, \kD representations of \pD data have been computed using
-multidimensional scaling (MDS) \citep{borg2005}, which includes
+Historically, \kD   representations of \pD   data have been computed
+using multidimensional scaling (MDS) \citep{borg2005}, which includes
 principal components analysis (PCA) \citep{jolliffe2011} as a special
-case. The \kD representation can be considered to be a layout of points
-in \kD produced by an embedding procedure that maps the data from \pD.
-In MDS, the \kD layout is constructed by minimizing a stress function
-that differences distances between points in \pD with potential
-distances between points in \kD. Various formulations of the stress
-function result in non-metric scaling \citep{saeed2018} and isomap
-\citep{silva2002}. Challenges in working with high-dimensional data,
-including visualization, are outlined in \citet{johnstone2009}.
+case. The \kD   representation can be considered to be a layout of
+points in \kD   produced by an embedding procedure that maps the data
+from \pD. In MDS, the \kD   layout is constructed by minimizing a stress
+function that differences distances between points in \pD   with
+potential distances between points in \kD. Various formulations of the
+stress function result in non-metric scaling \citep{saeed2018} and
+isomap \citep{silva2002}. Challenges in working with high-dimensional
+data, including visualization, are outlined in \citet{johnstone2009}.
 
@@
 
-At first glance, thinking of NLDR as a model fitted to the data might
-seem strange. It is a model in the sense that it is a ``a simplified
-representation or abstraction of a system, process, or phenomenon in the
-real world''. The \pD observations are the realization of the
-phenomenon, and the \kD NLDR layout is the simplified representation.
-From a statistical perspective we can consider the distances between
-points in the \kD layout to be variance that the model explains, and the
-(relative) difference with their distances in \pD is the error, or
-unexplained variance. Abstractly, we can also imagine that the
-positioning of points in 2D represents fitted values, that will have
-some prescribed position in \pD that can be compared with the observed
-value.
-
-Table~\ref{tbl-notation} summarises the notation used to explain the new
-methodology. The observed data is denoted as
-\(\mathbfit{x}_{n \times p}\) where \(x_{ij}\) would indicate the
-\(i^{th}\) observation on the \(j^{th}\) variable sampled from a
-population \(\mathbfit{X}\). To refer to variable \(j\), we would use
-\(X_j\).
+At first glance, thinking of NLDR as a modeling technique might seem
+strange. It is a simplified representation or abstraction of a system,
+process, or phenomenon in the real world. The \pD   observations are the
+realization of the phenomenon, and the \kD   NLDR layout is the
+simplified representation. From a statistical perspective we can
+consider the distances between points in the \kD   layout to be variance
+that the model explains, and the (relative) difference with their
+distances in \pD   is the error, or unexplained variance. We can also
+imagine that the positioning of points in 2D represent the fitted
+values, that will have some prescribed position in \pD   that can be
+compared with their observed values. This is the conceptual framework
+underlying the more formal versions of factor analysis \citep{cfa69} and
+multidimensional scaling (MDS) \citep{borg2005}. (Note that, for this
+thinking the full \pD   data needs to be available, not just the
+interpoint distances.)
 
@@
 \addlinespace
+$g$ & functional mapping from \pD to \kD, especially as prescribed by NLDR\\
+$\mathbfit{\theta}$ & (Hyper-) parameters for NLDR method\\
 $y_\text{max}$ & maximum value of the scaled second embedding component\\
@@
 $h_b$, $w_b$ & height and width of the hexagon\\
+\addlinespace
 $a_r$ & aspect ratio of the 2D layout before scaling\\
 $h_r$ & hexagon ratio\\
-\addlinespace
 $s$ & hexagonal size (radius of the outer circle surrounding the hexagon)\\
@@
 $q_x$, $q_y$ & buffer amount along the x and y axes\\
+\addlinespace
 $h_s$, $v_s$ & horizontal and vertical distance between adjacent hexagon bin centroids\\
 $d_x$, $d_y$ & horizontal spacing and adjusted vertical spacing for hexagonal coordinates\\
-\addlinespace
 $b_x$, $b_y$ & number of hexagon bins along the x and y axes\\
@@
 $n_k$ & number of observations within the $k^{th}$ hexagon\\
+\addlinespace
 $(h^k_{x_i}, h^k_{y_i})$ & hexagonal grid coordinates of the $k^{th}$ hexagon\\
@@
 
+We define the NLDR model as function
+\(g\text{:}~ \mathbb{R}^p \rightarrow \mathbb{R}^k\), with
+(hyper-)parameters \(\mathbfit{\theta}\). The parameters,
+\(\mathbfit{\theta}\), depend on the choice of \(g\), and can be
+considered part of model fitting in the traditional sense. Common
+choices for \(g\) include functions used in tSNE, UMAP, PHATE, TriMAP,
+PaCMAP, or MDS, although in theory any function that maps the entire
+space \(\mathbb{R}^p\) to a subset of \(\mathbb{R}^k\) is suitable.
+
+With our goal being to make a representation of this 2D layout that can
+be lifted into high-dimensional space, the layout needs to be augmented
+to include neighbour information. A simple approach would be to
+triangulate the points and add edges. A more stable approach is to first
+hexagonally bin the data, and connect the bin centroids. This process
+serves reduce some noise in the resulting surface shown in \pD.
+
+Define a hexagonal grid across the \kD space, using hexagons with fixed
+height (\(h_b\)), width (\(w_b\)). The starting coordinates
+\((x_s, y_s)\) mark the lower left of the grid. Each of the \kD points
+will belong to a hexagonal bin. That is, for each
+\(y \in \mathbfit{Y}\), we can (uniquely) identify the hexagon that the
+point belongs to. Let \(h : \mathbb{R}^k \rightarrow \mathbb{R}^k\) be
+the function that maps a point in \kD space to it's hexagonal centroid
+(\(C^{(k)}\)).
+
 To illustrate the method, we use \(7\text{-}D\) simulated data, which we
@@
 
-We need to have some math that gives precision to the paragraphs in the
-previous section, defining mapping from \pD to \kD to get the
-\kD layout. Might be written below, too (XXX Michael??)
-
-\subsection{\texorpdfstring{Generating the
-\kD layout}{Generating the layout}}\label{generating-the-layout}
-
-The usual procedures are followed to produce a \kD layout, using the
-preferred method, t-SNE, UMAP, \ldots{} The data may need to be
-pre-processed by standardising or transforming to principal components.
-
-\subsection{Constructing the 2D model}\label{constructing-the-2d-model}
-
 \subsection{\texorpdfstring{Displaying the model in
 \pD}{Displaying the model in }}\label{displaying-the-model-in}
+
+The final step of the algorithm is \emph{lifting} the \kD model back to
+\pD, completing the \emph{model-in-the-data-space} picture. For this,
+define the set \(H^k\) to be the set of points that belong to centroid
+\(k\). That is, \(H^k = \{ y \; | \; g(y) = (h_x^k, h_y^k)\}\). We use
+the \pD Euclidean mean of the points in \(H^k\) to map the centroid
+\((h_x^k, h_y^k)\) to a point in \pD. Let the \(i\)-th component of the
+\pD mean be
+
+\[f_i = \frac{1}{\left|H^k\right|}\sum_{y \in H^k} y_i,\]
+
+with corresponding \pD vector \(\vec{f_i}\). Then, \(f(h\circ g)(x)\)
+maps a \pD point to the \pD model estimate, completing the model cycle.
 
@@
 \item
+  Predicting new observations in \(k\)-D
+\item
+  Extending layouts beyond \(k\)-D, when 2D is clearly inadequate.
+\item
   Diagnostic app to explore differences in distances

### 2024-04-09 paper.qmd
fixed space after pD

@@
 
-Non-linear dimension reduction (NLDR) is popular for making a convenient low-dimensional (\kD) representation of high-dimensional (\pD) data.
+Non-linear dimension reduction (NLDR) is popular for making a convenient low-dimensional (\kD{}) representation of high-dimensional (\pD{}) data.
 Recently developed methods include t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2022], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021].
@@
 <!-- - Connection between NLDR and MDS-->
-Historically, \kD   representations of \pD   data have been computed using multidimensional scaling (MDS) [@borg2005], which includes principal components analysis (PCA) [@jolliffe2011] as a special case.
-The \kD   representation can be considered to be a layout of points in \kD   produced by an embedding procedure that maps the data from \pD.
-In MDS, the \kD   layout is constructed by minimizing a stress function that differences distances between points in \pD   with potential distances between points in \kD.
+Historically, \kD{}   representations of \pD{}   data have been computed using multidimensional scaling (MDS) [@borg2005], which includes principal components analysis (PCA) [@jolliffe2011] as a special case.
+The \kD{}   representation can be considered to be a layout of points in \kD{}   produced by an embedding procedure that maps the data from \pD{}.
+In MDS, the \kD{}   layout is constructed by minimizing a stress function that differences distances between points in \pD{}   with potential distances between points in \kD{}.
 Various formulations of the stress function result in non-metric scaling [@saeed2018] and isomap [@silva2002].
@@
 
-Many new methods for NLDR have emerged in recent years, all designed to better capture specific structures potentially existing in \pD.
+Many new methods for NLDR have emerged in recent years, all designed to better capture specific structures potentially existing in \pD{}.
 Here we focus on five currently popular techniques, tSNE, UMAP, PHATE, TriMAP and PaCMAP.
-tNSE and UMAP can be considered to produce the \kD   minimizing the divergence between two distributions, where the distributions are modeling the inter-point distances.
+tNSE and UMAP can be considered to produce the \kD{}   minimizing the divergence between two distributions, where the distributions are modeling the inter-point distances.
 PHATE, TriMAP and PaCMAP are examples of diffusion processes [@coifman2005] spreading to capture geometric shapes, that include both global and local structure.
@@
 However, linear projections can be cluttered, and global patterns can obscure local structure.
-The simple activity of projecting data from \pD   suffers from piling [@laa2022], where data concentrates in the center of projections.
+The simple activity of projecting data from \pD{}   suffers from piling [@laa2022], where data concentrates in the center of projections.
 NLDR is designed to escape these issues, to exaggerate structure so that it can be observed.
@@
 The fitted model should be overlaid on the data, to examine the fit relative the spread of the observations.
-While this is straightforward, and commonly done when data is 2D, it is also possible in \pD, for many models, when a tour is used.
+While this is straightforward, and commonly done when data is 2D, it is also possible in \pD{}, for many models, when a tour is used.
 
-@wickham2015 provides several examples of models overlaid on the data in \pD.
+@wickham2015 provides several examples of models overlaid on the data in \pD{}.
 In hierarchical clustering, a representation of the dendrogrom using points and lines can be constructed by augmenting the data with points marking merging of clusters.
@@
 It is possible to see whether the elliptical shapes appropriately matches the variance of the relevant clusters, and to compare and contrast different fits.
-For PCA, one can display the \kD plane of the reduced dimension using wireframes of transformed cubes.
-Using a wireframe is the approach we take here, to represent the NLDR model in \pD.
+For PCA, one can display the \kD{} plane of the reduced dimension using wireframes of transformed cubes.
+Using a wireframe is the approach we take here, to represent the NLDR model in \pD{}.
 
@@
 It is a simplified representation or abstraction of a system, process, or phenomenon in the real world.
-The \pD   observations are the realization of the phenomenon, and the \kD   NLDR layout is the simplified representation.
-From a statistical perspective we can consider the distances between points in the \kD   layout to be variance that the model explains, and the (relative) difference with their distances in \pD   is the error, or unexplained variance.
-We can also imagine that the positioning of points in 2D represent the fitted values, that will have some prescribed position in \pD   that can be compared with their observed values.
+The \pD{}   observations are the realization of the phenomenon, and the \kD{}   NLDR layout is the simplified representation.
+From a statistical perspective we can consider the distances between points in the \kD{}   layout to be variance that the model explains, and the (relative) difference with their distances in \pD{}   is the error, or unexplained variance.
+We can also imagine that the positioning of points in 2D represent the fitted values, that will have some prescribed position in \pD{}   that can be compared with their observed values.
 This is the conceptual framework underlying the more formal versions of factor analysis [@cfa69] and multidimensional scaling (MDS) [@borg2005].
-(Note that, for this thinking the full \pD   data needs to be available, not just the interpoint distances.)
+(Note that, for this thinking the full \pD{}   data needs to be available, not just the interpoint distances.)
 <!--### Notation -->
@@
 A more stable approach is to first hexagonally bin the data, and connect the bin centroids.
-This process serves reduce some noise in the resulting surface shown in \pD.
+This process serves reduce some noise in the resulting surface shown in \pD{}.
 
-Define a hexagonal grid across the \kD space, using hexagons with fixed height ($h_b$), width ($w_b$).
+Define a hexagonal grid across the \kD{} space, using hexagons with fixed height ($h_b$), width ($w_b$).
 The starting coordinates $(x_s, y_s)$ mark the lower left of the grid.
 <!--We deliberately separate out the creation of the hexagonal grid from the mapping of points on the grid.-->
-Each of the \kD points will belong to a hexagonal bin.
+Each of the \kD{} points will belong to a hexagonal bin.
 That is, for each $y \in \mathbfit{Y}$, we can (uniquely) identify the hexagon that the point belongs to.
-Let $h : \mathbb{R}^k \rightarrow \mathbb{R}^k$ be the function that maps a point in \kD space to it's hexagonal centroid ($C^{(k)}$).
-<!--It follows that $h \circ g : \mathbb{R}^p \rightarrow \mathbb{R}^k$ maps each point in $\mathbfit{X}$ to a \kD centroid.-->
+Let $h : \mathbb{R}^k \rightarrow \mathbb{R}^k$ be the function that maps a point in \kD{} space to it's hexagonal centroid ($C^{(k)}$).
+<!--It follows that $h \circ g : \mathbb{R}^p \rightarrow \mathbb{R}^k$ maps each point in $\mathbfit{X}$ to a \kD{} centroid.-->
 
@@
 
-### Displaying the model in \pD
+### Displaying the model in \pD{}
 
-The final step of the algorithm is _lifting_ the \kD model back to \pD, completing the *model-in-the-data-space* picture.
+The final step of the algorithm is _lifting_ the \kD{} model back to \pD{}, completing the *model-in-the-data-space* picture.
 For this, define the set $H^k$ to be the set of points that belong to centroid $k$.
 That is, $H^k = \{ y \; | \; g(y) = (h_x^k, h_y^k)\}$.
-We use the \pD Euclidean mean of the points in $H^k$ to map the centroid $(h_x^k, h_y^k)$ to a point in \pD.
-Let the $i$-th component of the \pD mean be
+We use the \pD{} Euclidean mean of the points in $H^k$ to map the centroid $(h_x^k, h_y^k)$ to a point in \pD{}.
+Let the $i$-th component of the \pD{} mean be
 
@@
 
-with corresponding \pD vector $\vec{f_i}$. Then, $f(h\circ g)(x)$ maps a \pD point to the \pD model estimate, completing the model cycle.
+with corresponding \pD{} vector $\vec{f_i}$. Then, $f(h\circ g)(x)$ maps a \pD{} point to the \pD{} model estimate, completing the model cycle.
 

### 2024-04-09 paper.tex
fixed space after pD

@@
 Non-linear dimension reduction (NLDR) is popular for making a convenient
-low-dimensional (\kD) representation of high-dimensional (\pD) data.
+low-dimensional (\kD{}) representation of high-dimensional (\pD{}) data.
 Recently developed methods include t-distributed stochastic neighbor
@@
 
-Historically, \kD   representations of \pD   data have been computed
+Historically, \kD{} representations of \pD{} data have been computed
 using multidimensional scaling (MDS) \citep{borg2005}, which includes
 principal components analysis (PCA) \citep{jolliffe2011} as a special
-case. The \kD   representation can be considered to be a layout of
-points in \kD   produced by an embedding procedure that maps the data
-from \pD. In MDS, the \kD   layout is constructed by minimizing a stress
-function that differences distances between points in \pD   with
-potential distances between points in \kD. Various formulations of the
+case. The \kD{} representation can be considered to be a layout of
+points in \kD{} produced by an embedding procedure that maps the data
+from \pD{}. In MDS, the \kD{} layout is constructed by minimizing a
+stress function that differences distances between points in \pD{} with
+potential distances between points in \kD{}. Various formulations of the
 stress function result in non-metric scaling \citep{saeed2018} and
@@
 Many new methods for NLDR have emerged in recent years, all designed to
-better capture specific structures potentially existing in \pD. Here we
-focus on five currently popular techniques, tSNE, UMAP, PHATE, TriMAP
-and PaCMAP. tNSE and UMAP can be considered to produce the
-\kD   minimizing the divergence between two distributions, where the
+better capture specific structures potentially existing in \pD{}. Here
+we focus on five currently popular techniques, tSNE, UMAP, PHATE, TriMAP
+and PaCMAP. tNSE and UMAP can be considered to produce the \kD{}
+minimizing the divergence between two distributions, where the
 distributions are modeling the inter-point distances. PHATE, TriMAP and
@@
 patterns can obscure local structure. The simple activity of projecting
-data from \pD   suffers from piling \citep{laa2022}, where data
+data from \pD{} suffers from piling \citep{laa2022}, where data
 concentrates in the center of projections. NLDR is designed to escape
@@
 While this is straightforward, and commonly done when data is 2D, it is
-also possible in \pD, for many models, when a tour is used.
+also possible in \pD{}, for many models, when a tour is used.
 
 \citet{wickham2015} provides several examples of models overlaid on the
-data in \pD. In hierarchical clustering, a representation of the
+data in \pD{}. In hierarchical clustering, a representation of the
 dendrogrom using points and lines can be constructed by augmenting the
@@
 appropriately matches the variance of the relevant clusters, and to
-compare and contrast different fits. For PCA, one can display the
-\kD plane of the reduced dimension using wireframes of transformed
-cubes. Using a wireframe is the approach we take here, to represent the
-NLDR model in \pD.
+compare and contrast different fits. For PCA, one can display the \kD{}
+plane of the reduced dimension using wireframes of transformed cubes.
+Using a wireframe is the approach we take here, to represent the NLDR
+model in \pD{}.
 
@@
 strange. It is a simplified representation or abstraction of a system,
-process, or phenomenon in the real world. The \pD   observations are the
-realization of the phenomenon, and the \kD   NLDR layout is the
+process, or phenomenon in the real world. The \pD{} observations are the
+realization of the phenomenon, and the \kD{} NLDR layout is the
 simplified representation. From a statistical perspective we can
-consider the distances between points in the \kD   layout to be variance
+consider the distances between points in the \kD{} layout to be variance
 that the model explains, and the (relative) difference with their
-distances in \pD   is the error, or unexplained variance. We can also
+distances in \pD{} is the error, or unexplained variance. We can also
 imagine that the positioning of points in 2D represent the fitted
-values, that will have some prescribed position in \pD   that can be
+values, that will have some prescribed position in \pD{} that can be
 compared with their observed values. This is the conceptual framework
@@
 multidimensional scaling (MDS) \citep{borg2005}. (Note that, for this
-thinking the full \pD   data needs to be available, not just the
+thinking the full \pD{} data needs to be available, not just the
 interpoint distances.)
@@
 hexagonally bin the data, and connect the bin centroids. This process
-serves reduce some noise in the resulting surface shown in \pD.
+serves reduce some noise in the resulting surface shown in \pD{}.
 
-Define a hexagonal grid across the \kD space, using hexagons with fixed
-height (\(h_b\)), width (\(w_b\)). The starting coordinates
-\((x_s, y_s)\) mark the lower left of the grid. Each of the \kD points
+Define a hexagonal grid across the \kD{} space, using hexagons with
+fixed height (\(h_b\)), width (\(w_b\)). The starting coordinates
+\((x_s, y_s)\) mark the lower left of the grid. Each of the \kD{} points
 will belong to a hexagonal bin. That is, for each
@@
 point belongs to. Let \(h : \mathbb{R}^k \rightarrow \mathbb{R}^k\) be
-the function that maps a point in \kD space to it's hexagonal centroid
+the function that maps a point in \kD{} space to it's hexagonal centroid
 (\(C^{(k)}\)).
@@
 \subsection{\texorpdfstring{Displaying the model in
-\pD}{Displaying the model in }}\label{displaying-the-model-in}
+\pD{}}{Displaying the model in }}\label{displaying-the-model-in}
 
-The final step of the algorithm is \emph{lifting} the \kD model back to
-\pD, completing the \emph{model-in-the-data-space} picture. For this,
-define the set \(H^k\) to be the set of points that belong to centroid
-\(k\). That is, \(H^k = \{ y \; | \; g(y) = (h_x^k, h_y^k)\}\). We use
-the \pD Euclidean mean of the points in \(H^k\) to map the centroid
-\((h_x^k, h_y^k)\) to a point in \pD. Let the \(i\)-th component of the
-\pD mean be
+The final step of the algorithm is \emph{lifting} the \kD{} model back
+to \pD{}, completing the \emph{model-in-the-data-space} picture. For
+this, define the set \(H^k\) to be the set of points that belong to
+centroid \(k\). That is, \(H^k = \{ y \; | \; g(y) = (h_x^k, h_y^k)\}\).
+We use the \pD{} Euclidean mean of the points in \(H^k\) to map the
+centroid \((h_x^k, h_y^k)\) to a point in \pD{}. Let the \(i\)-th
+component of the \pD{} mean be
 
@@
 
-with corresponding \pD vector \(\vec{f_i}\). Then, \(f(h\circ g)(x)\)
-maps a \pD point to the \pD model estimate, completing the model cycle.
+with corresponding \pD{} vector \(\vec{f_i}\). Then, \(f(h\circ g)(x)\)
+maps a \pD{} point to the \pD{} model estimate, completing the model
+cycle.
 

### 2024-04-09 paper.tex
model removwed

@@
 $a_r$ & aspect ratio of the 2D layout before scaling\\
-$h_r$ & hexagon ratio\\
 $s$ & hexagonal size (radius of the outer circle surrounding the hexagon)\\
@@
 $q_x$, $q_y$ & buffer amount along the x and y axes\\
+$h_s$, $v_s$ & horizontal and vertical distance between adjacent hexagon bin centroids\\
 \addlinespace
-$h_s$, $v_s$ & horizontal and vertical distance between adjacent hexagon bin centroids\\
 $d_x$, $d_y$ & horizontal spacing and adjusted vertical spacing for hexagonal coordinates\\
@@
 $n_k$ & number of observations within the $k^{th}$ hexagon\\
+$(h^k_{x_i}, h^k_{y_i})$ & hexagonal grid coordinates of the $k^{th}$ hexagon\\
 \addlinespace
-$(h^k_{x_i}, h^k_{y_i})$ & hexagonal grid coordinates of the $k^{th}$ hexagon\\
 $C^{(j)}$ & $j$-dimensional bin centers\\
@@
 
-We define the NLDR model as function
+We define the NLDR as a function
 \(g\text{:}~ \mathbb{R}^p \rightarrow \mathbb{R}^k\), with

### 2024-04-09 paper.tex
notation changes

@@
 \midrule
-$n, p, k$ & number of observations, variables, embedding dimension, respectively\\
+$n, p, k, m$ & number of observations, variables, embedding dimension, number of non-empty bins, respectively\\
 $\mathbfit{X}, \mathbfit{x}$ & $p$-dimensional data (population, sample)\\
@@
 \addlinespace
-$g$ & functional mapping from \pD to \kD, especially as prescribed by NLDR\\
+$g$ & functional mapping from \pD{} to \kD{}, especially as prescribed by NLDR\\
 $\mathbfit{\theta}$ & (Hyper-) parameters for NLDR method\\
 $y_\text{max}$ & maximum value of the scaled second embedding component\\
-$r_1$, $r_2$ & range of the first and  second embedding component\\
-$h_b$, $w_b$ & height and width of the hexagon\\
+$r$ & ranges of the embedding components\\
+$b$ & bins\\
 \addlinespace

### 2024-04-09 paper.qmd
model removwed
Added UMAP layout for S-curve data

@@
 
-We define the NLDR model as function $g\text{:}~ \mathbb{R}^p \rightarrow \mathbb{R}^k$, with (hyper-)parameters $\mathbfit{\theta}$.
+We define the NLDR as a function $g\text{:}~ \mathbb{R}^p \rightarrow \mathbb{R}^k$, with (hyper-)parameters $\mathbfit{\theta}$.
 The parameters, $\mathbfit{\theta}$, depend on the choice of $g$, and can be considered part of model fitting in the traditional sense.
@@
 We would consider $T=(X_1, X_2, X_3)$ to be the true model.
+
+<!--UMAP applied for Scurve data-->
+```{r}
+umap_scurve <- read_rds(file = "data/s_curve/s_curve_umap.rds")
+
+nldr_scurve <- umap_scurve |> 
+  ggplot(aes(x = UMAP1, y = UMAP2)) + 
+  geom_point(alpha=0.5, colour="#000000", size = 0.5) +
+  interior_annotation("a")
+```
+
+<!--Full hexagonal grid with UMAP data-->
+
+<!--Non-empty bins with bin centroids-->
+
+<!--2D model-->
+
+```{r}
+#| label: fig-NLDR-scurve
+#| echo: false
+#| fig-cap: "Six different NLDR representations of the same data.
+Different techniques and different parameter choices are used.
+Researchers may have seen any of these in their analysis of this data, depending on their choice of method, or typical parameter choice.
+Would they make different decisions downstream in the analysis depending on which version seen?
+Which is the most accurate representation of the structure in high dimensions?"
+#| fig-width: 8
+#| fig-height: 4
+#| out-width: 100%
+
+nldr_scurve +
+  plot_layout(ncol = 4)
+```
 

### 2024-04-09 paper.qmd
Add plots for 2D model construction

@@
 
-nldr_scurve <- umap_scurve |> 
+umap_scurve_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_scurve, 
+                                    x = "UMAP1", y = "UMAP2"))) |>
+  rename(c("UMAP1" = "scaled_UMAP1", 
+                  "UMAP2" = "scaled_UMAP2")) |>
+  mutate(ID = 1:NROW(umap_scurve))
+
+nldr_scurve <- umap_scurve_scaled |> 
   ggplot(aes(x = UMAP1, y = UMAP2)) + 
@@
 
+```{r}
+num_bins_x_scurve <- 6
+num_bins_y_scurve <- 14
+hex_size_scurve <- 0.13
+
+hb_obj_scurve <- hex_binning(data = umap_scurve_scaled, x = "UMAP1", 
+                      y = "UMAP2", num_bins_x = num_bins_x_scurve, 
+                      num_bins_y = num_bins_y_scurve, x_start = NA, y_start = NA, 
+                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_scurve, 
+                      col_start = "UMAP")
+
+## Data set with all possible centroids in the hexagonal grid
+all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_scurve$centroids))
+
+## Generate all coordinates of hexagons
+hex_grid <- as.data.frame(do.call(cbind, hb_obj_scurve$hex_poly))
+
+## To obtain the standardise counts within hexbins
+counts_df <- as.data.frame(do.call(cbind, hb_obj_scurve$std_cts))
+df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
+                                             counts_df = counts_df)
+
+hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
+
+hex_grid_scurve <-  ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = hex_poly_id), fill = "#ffffff") +
+  geom_point(data = umap_scurve_scaled, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
+  interior_annotation("b")
+```
+
 <!--Non-empty bins with bin centroids-->
+
+```{r}
+hex_grid_nonempty <- hex_grid |>
+  filter(hex_poly_id %in% df_bin_centroids$hexID)
+
+hex_grid_nonempty_scurve <-  ggplot(data = hex_grid_nonempty, aes(x = x, y = y)) +
+  geom_polygon(color = "black", aes(group = hex_poly_id), fill = "#ffffff") +
+  geom_point(data = df_bin_centroids, aes(x = c_x, y = c_y), color = "#33a02c") +
+  interior_annotation("c") 
+```
 
@@
 ```{r}
+## Triangulate bin centroids
+tr1_object_scurve <- tri_bin_centroids(df_bin_centroids, x = "c_x", y = "c_y")
+tr_from_to_df_scurve <- gen_edges(tri_object = tr1_object_scurve)
+
+## Compute 2D distances
+distance_scurve <- cal_2d_dist(tr_coord_df = tr_from_to_df_scurve, 
+                             start_x = "x_from", start_y = "y_from", 
+                             end_x = "x_to", end_y = "y_to", 
+                             select_vars = c("from", "to", "distance"))
+
+## To find the benchmark value
+benchmark_scurve <- find_lg_benchmark(distance_edges = distance_scurve, distance_col = "distance")
+
+trimesh_removed_scurve <- vis_rmlg_mesh(distance_edges = distance_scurve, 
+                                        benchmark_value = benchmark_scurve, 
+                                        tr_coord_df = tr_from_to_df_scurve, 
+                                        distance_col = "distance") +
+  interior_annotation("d") 
+```
+
+```{r}
 #| label: fig-NLDR-scurve
 #| echo: false
-#| fig-cap: "Six different NLDR representations of the same data.
-Different techniques and different parameter choices are used.
-Researchers may have seen any of these in their analysis of this data, depending on their choice of method, or typical parameter choice.
-Would they make different decisions downstream in the analysis depending on which version seen?
-Which is the most accurate representation of the structure in high dimensions?"
+#| fig-cap: ""
 #| fig-width: 8
@@
 
-nldr_scurve +
+nldr_scurve + hex_grid_scurve + hex_grid_nonempty_scurve + trimesh_removed_scurve +
   plot_layout(ncol = 4)

### 2024-04-09 paper.qmd
Added Figure title

@@
       x = unit(0.92, "npc"), y = unit(0.92, "npc"),
+      gp = grid::gpar(cex = 1, col="grey70")))
+}
+
+interior_annotation_left <- function(label) {
+  annotation_custom(grid::textGrob(label = label,
+      x = unit(0.08, "npc"), y = unit(0.93, "npc"),
       gp = grid::gpar(cex = 1, col="grey70")))
@@
   geom_point(alpha=0.5, colour="#000000", size = 0.5) +
-  interior_annotation("a")
+  interior_annotation_left("a")
 ```
@@
   geom_point(data = umap_scurve_scaled, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
-  interior_annotation("b")
+  interior_annotation_left("b")
 ```
@@
   geom_point(data = df_bin_centroids, aes(x = c_x, y = c_y), color = "#33a02c") +
-  interior_annotation("c") 
+  interior_annotation_left("c") 
 ```
@@
                                         distance_col = "distance") +
-  interior_annotation("d") 
+  interior_annotation_left("d") 
 ```
@@
 #| echo: false
-#| fig-cap: ""
+#| fig-cap: "Key steps for constructing the $2-D$ model for S-curve data: (a) obtain UMAP data, (b) hexagonating UMAP data, (c) obtain hexagonal bin centroids, and (d) triangulate bin centroids."
 #| fig-width: 8

### 2024-04-09 paper.tex
Added UMAP layout for S-curve data
Add plots for 2D model construction
Added Figure title

--- 2024-04-09 paper.tex
@@
 
+\begin{figure}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-NLDR-scurve-1.pdf}
+
+}
+
+\caption{\label{fig-NLDR-scurve}Key steps for constructing the \(2-D\)
+model for S-curve data: (a) obtain UMAP data, (b) hexagonating UMAP
+data, (c) obtain hexagonal bin centroids, and (d) triangulate bin
+centroids.}
+
+\end{figure}%
+
 \subsection{\texorpdfstring{Displaying the model in

### 2024-04-09 paper.qmd
Updated

@@
    theme(
-     aspect.ratio = 1,
+     #aspect.ratio = 1,
      plot.background = element_rect(fill = 'transparent', colour = NA),
@@
   geom_point(alpha=0.5, colour="#000000", size = 0.5) +
+  coord_fixed() +
   interior_annotation_left("a")
@@
   geom_point(data = umap_scurve_scaled, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
+  coord_fixed() +
   interior_annotation_left("b")
@@
   geom_point(data = df_bin_centroids, aes(x = c_x, y = c_y), color = "#33a02c") +
+  coord_fixed() +
   interior_annotation_left("c") 

### 2024-04-11 paper.tex

--- 2024-04-09 paper.tex
@@
 $a_r$ & aspect ratio of the 2D layout before scaling\\
-$s$ & hexagonal size (radius of the outer circle surrounding the hexagon)\\
+$s$ & hexagonal size (defined as the radius of the outer circle surrounding the hexagon)\\
 $(x_s, \ y_s)$ & starting coordinates of the hexagonal grid\\
@@
 true model.
+
+\begin{figure}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-NLDR-scurve-1.pdf}
+
+}
+
+\caption{\label{fig-NLDR-scurve}Key steps for constructing the \(2-D\)
+model for S-curve data: (a) obtain UMAP data, (b) hexagonating UMAP
+data, (c) obtain hexagonal bin centroids, and (d) triangulate bin
+centroids.}
+
+\end{figure}%
 

### 2024-04-11 paper.qmd
math changes for the mappings

@@
 
-We define the NLDR as a function $g\text{:}~ \mathbb{R}^p \rightarrow \mathbb{R}^k$, with (hyper-)parameters $\mathbfit{\theta}$.
+We define the NLDR as a function $g\text{:}~ \mathbb{R}^{n\times p} \rightarrow \mathbb{R}^{n\times k}$, with (hyper-)parameters $\mathbfit{\theta}$.
 The parameters, $\mathbfit{\theta}$, depend on the choice of $g$, and can be considered part of model fitting in the traditional sense.
-Common choices for $g$ include functions used in tSNE, UMAP, PHATE, TriMAP, PaCMAP, or MDS, although in theory any function that maps the entire space $\mathbb{R}^p$ to a subset of $\mathbb{R}^k$ is suitable.
+Common choices for $g$ include functions used in tSNE, UMAP, PHATE, TriMAP, PaCMAP, or MDS, although in theory any function that does this mapping is suitable.
 <!--Any input requirements for the data (such as normalization, or preprocessing through the use of PCA or similar) is considered part of the function $g$.-->
@@
 That is, for each $y \in \mathbfit{Y}$, we can (uniquely) identify the hexagon that the point belongs to.
-Let $h : \mathbb{R}^k \rightarrow \mathbb{R}^k$ be the function that maps a point in \kD{} space to it's hexagonal centroid ($C^{(k)}$).
-<!--It follows that $h \circ g : \mathbb{R}^p \rightarrow \mathbb{R}^k$ maps each point in $\mathbfit{X}$ to a \kD{} centroid.-->
-
-<!--First, a random variable $a$ is generated from a uniform distribution between $-\frac{\pi}{2}$ and $\frac{\pi}{2}$, representing the parameter $a$ is in $T$.
-Next, three variables $x_1$, $x_2$, and $x_3$ are computed based on $a$ to construct the S-curve data.
-$x_1$ is calculated as the sine of $a$ (@eq-eqn1), $x_2$ as twice the value of another uniformly distributed random variable between 0 and 1 (@eq-eqn2), and $x_3$ as the cosine of $a$ minus 1, multiplied by the sign of $a$ (@eq-eqn3).
-Additionally, four noise dimensions are introduced to the dataset to introduce noise variability: $x_4$, $x_5$, $x_6$, and $x_7$, each generated from uniform distributions within specific intervals (@eq-eqn4, @eq-eqn5, and @eq-eqn6 respectively).
-Finally, these variables are combined to form the S-curve dataset $T$.
-
-$$
-x_1 = \sin(3 \pi a)
-$$ {#eq-eqn1}
-
-$$
-x_2 = 2.0 \times U(0, 1)
-$$ {#eq-eqn2}
-
-$$
-x_3 = \text{sign}(a) \times (\cos(3 \pi a) - 1)
-$$ {#eq-eqn3}
-
-$$
-x_4, x_5 \sim U(-0.02, 0.02)
-$$ {#eq-eqn4}
-
-$$
-x_6 \sim U(-0.1, 0.1)
-$$ {#eq-eqn5}
-
-$$
-x_7 \sim U(-0.01, 0.01)
-$$ {#eq-eqn6}
--->
+Let $h : \mathbb{R}^{n\times k} \rightarrow \mathbb{R}^{m\times k}$ be the function that maps a point in \kD{} space to it's hexagonal centroid ($C^{(k)}$).
 
@@
 #| echo: false
-#| fig-cap: "Key steps for constructing the $2-D$ model for S-curve data: (a) obtain UMAP data, (b) hexagonating UMAP data, (c) obtain hexagonal bin centroids, and (d) triangulate bin centroids."
+#| fig-cap: "Key steps for constructing the model on the NLDR layout ($k=2$) of the S-curve data: (a) obtain UMAP data, (b) hexagonating UMAP data, (c) obtain hexagonal bin centroids, and (d) triangulate bin centroids."
 #| fig-width: 8

### 2024-04-11 paper.tex
math changes for the mappings

@@
 We define the NLDR as a function
-\(g\text{:}~ \mathbb{R}^p \rightarrow \mathbb{R}^k\), with
-(hyper-)parameters \(\mathbfit{\theta}\). The parameters,
+\(g\text{:}~ \mathbb{R}^{n\times p} \rightarrow \mathbb{R}^{n\times k}\),
+with (hyper-)parameters \(\mathbfit{\theta}\). The parameters,
 \(\mathbfit{\theta}\), depend on the choice of \(g\), and can be
@@
 choices for \(g\) include functions used in tSNE, UMAP, PHATE, TriMAP,
-PaCMAP, or MDS, although in theory any function that maps the entire
-space \(\mathbb{R}^p\) to a subset of \(\mathbb{R}^k\) is suitable.
+PaCMAP, or MDS, although in theory any function that does this mapping
+is suitable.
 
@@
 \(y \in \mathbfit{Y}\), we can (uniquely) identify the hexagon that the
-point belongs to. Let \(h : \mathbb{R}^k \rightarrow \mathbb{R}^k\) be
-the function that maps a point in \kD{} space to it's hexagonal centroid
+point belongs to. Let
+\(h : \mathbb{R}^{n\times k} \rightarrow \mathbb{R}^{m\times k}\) be the
+function that maps a point in \kD{} space to it's hexagonal centroid
 (\(C^{(k)}\)).
@@
 
-\caption{\label{fig-NLDR-scurve}Key steps for constructing the \(2-D\)
-model for S-curve data: (a) obtain UMAP data, (b) hexagonating UMAP
-data, (c) obtain hexagonal bin centroids, and (d) triangulate bin
-centroids.}
+\caption{\label{fig-NLDR-scurve}Key steps for constructing the model on
+the NLDR layout (\(k=2\)) of the S-curve data: (a) obtain UMAP data, (b)
+hexagonating UMAP data, (c) obtain hexagonal bin centroids, and (d)
+triangulate bin centroids.}
 

### 2024-04-11 paper.qmd
tinkering with the code

@@
 
-umap_scurve_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_scurve, 
-                                    x = "UMAP1", y = "UMAP2"))) |>
-  rename(c("UMAP1" = "scaled_UMAP1", 
-                  "UMAP2" = "scaled_UMAP2")) |>
-  mutate(ID = 1:NROW(umap_scurve))
+ar <- umap_scurve |>
+  summarise_at(c("UMAP1", "UMAP2"), 
+    function(x) max(x)-min(x)) |>
+  mutate(ar = UMAP2/UMAP1) |>
+  pull(ar) # This is a bit cumbersome too
+hr <- 2/sqrt(3)
+umap_scurve_scaled <- umap_scurve |>
+  mutate_at(c("UMAP1", "UMAP2"), 
+    function(x) (x - min(x))/(max(x)-min(x))) |> # can be no missings
+  mutate(UMAP2 = UMAP2 * ceiling(ar/hr) * hr)
+
+  #as.data.frame(do.call(cbind, gen_scaled_data(data = umap_scurve, 
+  #                                  x = "UMAP1", y = "UMAP2"))) |>
+  #rename(c("UMAP1" = "scaled_UMAP1", 
+  #                "UMAP2" = "scaled_UMAP2")) |>
+  #mutate(ID = 1:NROW(umap_scurve))
 
@@
   geom_point(alpha=0.5, colour="#000000", size = 0.5) +
-  coord_fixed() +
+  #coord_fixed() +
+  xlim(c(-0.3, 1.3)) + ylim(c(-0.3, 2.6)) +
   interior_annotation_left("a")
@@
   geom_point(data = umap_scurve_scaled, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
-  coord_fixed() +
+  #coord_fixed() +
+  xlim(c(-0.3, 1.3)) + ylim(c(-0.3, 2.6)) +
   interior_annotation_left("b")
@@
   geom_point(data = df_bin_centroids, aes(x = c_x, y = c_y), color = "#33a02c") +
-  coord_fixed() +
+  #coord_fixed() +
+  xlim(c(-0.3, 1.3)) + ylim(c(-0.3, 2.6)) +
   interior_annotation_left("c") 

### 2024-04-11 paper.qmd
added lims for last plot

@@
                                         distance_col = "distance") +
+  xlim(c(-0.3, 1.3)) + ylim(c(-0.3, 2.6)) +
   interior_annotation_left("d") 

### 2024-04-12 paper.qmd
Update inner_annotation

--- 2024-04-11 paper.qmd
@@
 )
-interior_annotation <- function(label) {
+interior_annotation <- function(label, position = c(0.92, 0.92)) {
   annotation_custom(grid::textGrob(label = label,
-      x = unit(0.92, "npc"), y = unit(0.92, "npc"),
-      gp = grid::gpar(cex = 1, col="grey70")))
-}
-
-interior_annotation_left <- function(label) {
-  annotation_custom(grid::textGrob(label = label,
-      x = unit(0.08, "npc"), y = unit(0.93, "npc"),
+      x = unit(position[1], "npc"), y = unit(position[2], "npc"),
       gp = grid::gpar(cex = 1, col="grey70")))
@@
   xlim(c(-0.3, 1.3)) + ylim(c(-0.3, 2.6)) +
-  interior_annotation_left("a")
+  interior_annotation("a", c(0.08, 0.93))
 ```
@@
   xlim(c(-0.3, 1.3)) + ylim(c(-0.3, 2.6)) +
-  interior_annotation_left("b")
+  interior_annotation("b", c(0.08, 0.93))
 ```
@@
   xlim(c(-0.3, 1.3)) + ylim(c(-0.3, 2.6)) +
-  interior_annotation_left("c") 
+  interior_annotation("c", c(0.08, 0.93)) 
 ```
@@
                                         distance_col = "distance") +
-  interior_annotation_left("d") 
+  xlim(c(-0.3, 1.3)) + ylim(c(-0.3, 2.6)) +
+  interior_annotation("d", c(0.08, 0.93)) 
 ```

### 2024-04-12 paper.qmd
Updated with new functions

@@
 
-ar <- umap_scurve |>
-  summarise_at(c("UMAP1", "UMAP2"), 
-    function(x) max(x)-min(x)) |>
-  mutate(ar = UMAP2/UMAP1) |>
-  pull(ar) # This is a bit cumbersome too
-hr <- 2/sqrt(3)
-umap_scurve_scaled <- umap_scurve |>
-  mutate_at(c("UMAP1", "UMAP2"), 
-    function(x) (x - min(x))/(max(x)-min(x))) |> # can be no missings
-  mutate(UMAP2 = UMAP2 * ceiling(ar/hr) * hr)
+umap_scurve_scaled <- gen_scaled_data(data = umap_scurve, x = "UMAP1", 
+                                      y = "UMAP2", hr = 2/sqrt(3))
+
+# ar <- umap_scurve |>
+#   summarise_at(c("UMAP1", "UMAP2"), 
+#     function(x) max(x)-min(x)) |>
+#   mutate(ar = UMAP2/UMAP1) |>
+#   pull(ar) # This is a bit cumbersome too
+# hr <- 2/sqrt(3)
+# umap_scurve_scaled <- umap_scurve |>
+#   mutate_at(c("UMAP1", "UMAP2"), 
+#     function(x) (x - min(x))/(max(x)-min(x))) |> # can be no missings
+#   mutate(UMAP2 = UMAP2 * ceiling(ar/hr) * hr)
 
@@
                       num_bins_y = num_bins_y_scurve, x_start = NA, y_start = NA, 
-                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_scurve, 
+                      buffer_x = 0.346, buffer_y = 0.3, hex_size = hex_size_scurve, 
                       col_start = "UMAP")
@@
 ## Data set with all possible centroids in the hexagonal grid
-all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_scurve$centroids))
+all_centroids_df <- hb_obj_scurve$centroids
 
 ## Generate all coordinates of hexagons
-hex_grid <- as.data.frame(do.call(cbind, hb_obj_scurve$hex_poly))
+hex_grid <- hb_obj_scurve$hex_poly
 
 ## To obtain the standardise counts within hexbins
-counts_df <- as.data.frame(do.call(cbind, hb_obj_scurve$std_cts))
+counts_df <- hb_obj_scurve$std_cts
 df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 

### 2024-04-13 paper.qmd
Fix issues

@@
 ```{r}
+## Compute hexbin parameters
 num_bins_x_scurve <- 6
@@
 hex_size_scurve <- 0.13
+buffer_x_scurve <- 0.338
+buffer_y_scurve <- 0.292
+x_start_scurve <- -0.113
+y_start_scurve <- -0.098
 
+## Hexagonal bining to have regular hexagons
 hb_obj_scurve <- hex_binning(data = umap_scurve_scaled, x = "UMAP1",
                       y = "UMAP2", num_bins_x = num_bins_x_scurve,
-                      num_bins_y = num_bins_y_scurve, x_start = NA, y_start = NA, 
-                      buffer_x = 0.346, buffer_y = 0.3, hex_size = hex_size_scurve, 
+                      num_bins_y = num_bins_y_scurve, x_start = x_start_scurve,
+                      y_start = y_start_scurve, buffer_x = buffer_x_scurve,
+                      buffer_y = buffer_y_scurve, hex_size = hex_size_scurve,
                       col_start = "UMAP")
@@
 - Demonstrate that it is not
-- Illustrate how to use out method to get a better representation
+- Illustrate how to use our method to get a better representation
 

### 2024-04-15 paper.qmd
code tidied up some more and caption tidied

@@
 
-umap_scurve_scaled <- gen_scaled_data(data = umap_scurve, x = "UMAP1", 
-                                      y = "UMAP2", hr = 2/sqrt(3))
+umap_scurve_scaled <- gen_scaled_data(
+  data = umap_scurve, 
+  x = "UMAP1", 
+  y = "UMAP2", 
+  hr = 2/sqrt(3))
 
-# ar <- umap_scurve |>
-#   summarise_at(c("UMAP1", "UMAP2"), 
-#     function(x) max(x)-min(x)) |>
-#   mutate(ar = UMAP2/UMAP1) |>
-#   pull(ar) # This is a bit cumbersome too
-# hr <- 2/sqrt(3)
-# umap_scurve_scaled <- umap_scurve |>
-#   mutate_at(c("UMAP1", "UMAP2"), 
-#     function(x) (x - min(x))/(max(x)-min(x))) |> # can be no missings
-#   mutate(UMAP2 = UMAP2 * ceiling(ar/hr) * hr)
-
-  #as.data.frame(do.call(cbind, gen_scaled_data(data = umap_scurve, 
-  #                                  x = "UMAP1", y = "UMAP2"))) |>
-  #rename(c("UMAP1" = "scaled_UMAP1", 
-  #                "UMAP2" = "scaled_UMAP2")) |>
-  #mutate(ID = 1:NROW(umap_scurve))
-
+sc_ltr_pos <- c(0.08, 0.96)
+sc_xlims <- c(-0.25, 1.25)
+sc_ylims <- c(-0.25, 2.59)
 nldr_scurve <- umap_scurve_scaled |> 
@@
   #coord_fixed() +
-  xlim(c(-0.3, 1.3)) + ylim(c(-0.3, 2.6)) +
-  interior_annotation("a", c(0.08, 0.93))
+  xlim(sc_xlims) + ylim(sc_ylims) +
+  interior_annotation("a", sc_ltr_pos)
 ```
@@
 ## Hexagonal bining to have regular hexagons
-hb_obj_scurve <- hex_binning(data = umap_scurve_scaled, x = "UMAP1",
-                      y = "UMAP2", num_bins_x = num_bins_x_scurve,
-                      num_bins_y = num_bins_y_scurve, x_start = x_start_scurve,
-                      y_start = y_start_scurve, buffer_x = buffer_x_scurve,
-                      buffer_y = buffer_y_scurve, hex_size = hex_size_scurve,
+hb_obj_scurve <- hex_binning(
+  data = umap_scurve_scaled, 
+  x = "UMAP1",
+  y = "UMAP2", 
+  num_bins_x = num_bins_x_scurve,
+  num_bins_y = num_bins_y_scurve, 
+  x_start = x_start_scurve,
+  y_start = y_start_scurve, 
+  buffer_x = buffer_x_scurve,
+  buffer_y = buffer_y_scurve, 
+  hex_size = hex_size_scurve,
   col_start = "UMAP")
 
-## Data set with all possible centroids in the hexagonal grid
+## Data set with all centroids
 all_centroids_df <- hb_obj_scurve$centroids
@@
 counts_df <- hb_obj_scurve$std_cts
-df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
+df_bin_centroids <- extract_hexbin_centroids(
+  centroids_df = all_centroids_df, 
   counts_df = counts_df)
 
-hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
+hex_grid_with_counts <- 
+  dplyr::left_join(hex_grid,
+                   counts_df, 
+                   by = c("hex_poly_id" = "hb_id"))
 
-hex_grid_scurve <-  ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = hex_poly_id), fill = "#ffffff") +
-  geom_point(data = umap_scurve_scaled, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
-  #coord_fixed() +
-  xlim(c(-0.3, 1.3)) + ylim(c(-0.3, 2.6)) +
-  interior_annotation("b", c(0.08, 0.93))
+hex_grid_scurve <- ggplot(
+  data = hex_grid_with_counts, 
+  aes(x = x, y = y)) +
+  geom_polygon(color = "black", 
+               aes(group = hex_poly_id), 
+               fill = "#ffffff") +
+  geom_point(data = umap_scurve_scaled, 
+             aes(x = UMAP1, y = UMAP2), 
+             alpha = 0.5) +
+  xlim(sc_xlims) + ylim(sc_ylims) +
+  interior_annotation("b", sc_ltr_pos)
 ```
@@
 
-hex_grid_nonempty_scurve <-  ggplot(data = hex_grid_nonempty, aes(x = x, y = y)) +
-  geom_polygon(color = "black", aes(group = hex_poly_id), fill = "#ffffff") +
-  geom_point(data = df_bin_centroids, aes(x = c_x, y = c_y), color = "#33a02c") +
-  #coord_fixed() +
-  xlim(c(-0.3, 1.3)) + ylim(c(-0.3, 2.6)) +
-  interior_annotation("c", c(0.08, 0.93)) 
+hex_grid_nonempty_scurve <-  ggplot(
+  data = hex_grid_nonempty, 
+  aes(x = x, y = y)) +
+  geom_polygon(color = "black", 
+               aes(group = hex_poly_id), 
+               fill = "#ffffff") +
+  geom_point(data = df_bin_centroids, 
+             aes(x = c_x, y = c_y), 
+             color = "#33a02c") +
+  xlim(sc_xlims) + ylim(sc_ylims) +
+  interior_annotation("c", sc_ltr_pos) 
 ```
@@
 ## Triangulate bin centroids
-tr1_object_scurve <- tri_bin_centroids(df_bin_centroids, x = "c_x", y = "c_y")
-tr_from_to_df_scurve <- gen_edges(tri_object = tr1_object_scurve)
+tr1_object_scurve <- tri_bin_centroids(
+  df_bin_centroids, x = "c_x", y = "c_y")
+tr_from_to_df_scurve <- gen_edges(
+  tri_object = tr1_object_scurve)
 
 ## Compute 2D distances
-distance_scurve <- cal_2d_dist(tr_coord_df = tr_from_to_df_scurve, 
-                             start_x = "x_from", start_y = "y_from", 
-                             end_x = "x_to", end_y = "y_to", 
+distance_scurve <- cal_2d_dist(
+  tr_coord_df = tr_from_to_df_scurve, 
+  start_x = "x_from", 
+  start_y = "y_from", 
+  end_x = "x_to", 
+  end_y = "y_to", 
   select_vars = c("from", "to", "distance"))
@@
 ## To find the benchmark value
-benchmark_scurve <- find_lg_benchmark(distance_edges = distance_scurve, distance_col = "distance")
+benchmark_scurve <- find_lg_benchmark(
+  distance_edges = distance_scurve, 
+  distance_col = "distance")
 
-trimesh_removed_scurve <- vis_rmlg_mesh(distance_edges = distance_scurve, 
+trimesh_removed_scurve <- vis_rmlg_mesh(
+  distance_edges = distance_scurve, 
   benchmark_value = benchmark_scurve, 
@@
   distance_col = "distance") +
-  xlim(c(-0.3, 1.3)) + ylim(c(-0.3, 2.6)) +
-  interior_annotation("d", c(0.08, 0.93)) 
+  xlim(sc_xlims) + ylim(sc_ylims) +
+  interior_annotation("d", sc_ltr_pos) 
 ```
@@
 #| echo: false
-#| fig-cap: "Key steps for constructing the model on the NLDR layout ($k=2$) of the S-curve data: (a) obtain UMAP data, (b) hexagonating UMAP data, (c) obtain hexagonal bin centroids, and (d) triangulate bin centroids."
+#| fig-cap: "Key steps for constructing the model on the UMAP layout ($k=2$) of the S-curve data: (a) data, (b) hexagon bins, (c) bin centroids, and (d) triangulated centroids."
 #| fig-width: 8
@@
 
-nldr_scurve + hex_grid_scurve + hex_grid_nonempty_scurve + trimesh_removed_scurve +
+nldr_scurve + hex_grid_scurve + 
+  hex_grid_nonempty_scurve + 
+  trimesh_removed_scurve +
   plot_layout(ncol = 4)

### 2024-04-15 paper.tex
Fix issues
code tidied up some more and caption tidied

@@
 \caption{\label{fig-NLDR-scurve}Key steps for constructing the model on
-the NLDR layout (\(k=2\)) of the S-curve data: (a) obtain UMAP data, (b)
-hexagonating UMAP data, (c) obtain hexagonal bin centroids, and (d)
-triangulate bin centroids.}
+the UMAP layout (\(k=2\)) of the S-curve data: (a) data, (b) hexagon
+bins, (c) bin centroids, and (d) triangulated centroids.}
 
@@
 \item
-  Illustrate how to use out method to get a better representation
+  Illustrate how to use our method to get a better representation
 \end{itemize}

### 2024-04-15 paper.qmd
steps for hexagon binning penciled out

@@
 A more stable approach is to first hexagonally bin the data, and connect the bin centroids.
-This process serves reduce some noise in the resulting surface shown in \pD{}.
-
-Define a hexagonal grid across the \kD{} space, using hexagons with fixed height ($h_b$), width ($w_b$).
-The starting coordinates $(x_s, y_s)$ mark the lower left of the grid.
-<!--We deliberately separate out the creation of the hexagonal grid from the mapping of points on the grid.-->
-Each of the \kD{} points will belong to a hexagonal bin.
-That is, for each $y \in \mathbfit{Y}$, we can (uniquely) identify the hexagon that the point belongs to.
-Let $h : \mathbb{R}^{n\times k} \rightarrow \mathbb{R}^{m\times k}$ be the function that maps a point in \kD{} space to it's hexagonal centroid ($C^{(k)}$).
-
-To illustrate the method, we use $7\text{-}D$ simulated data, which we call the "S-curve".
-It is constructed by setting $X_1 = \sin(a), X_2 = U(0, 2), X_3 = \text{sign}(a) \times (\cos(a) - 1), \forall a \in [-3\pi/2, 3\pi/2]$.
-The remaining variables $X_4, X_5, X_6, X_7$ are all uniform error, with small variance.
-We would consider $T=(X_1, X_2, X_3)$ to be the true model.
+This process serves to reduce some noisiness in the resulting surface shown in \pD{}.
+The steps in this process are shown in @fig-NLDR-scurve, and documented below.
 
@@
   geom_point(alpha=0.5, colour="#000000", size = 0.5) +
-  #coord_fixed() +
   xlim(sc_xlims) + ylim(sc_ylims) +
@@
 
+To illustrate the method, we use $7\text{-}D$ simulated data, which we call the "S-curve".
+It is constructed by setting $X_1 = \sin(a), X_2 = U(0, 2), X_3 = \text{sign}(a) \times (\cos(a) - 1), \forall a \in [-3\pi/2, 3\pi/2]$.
+The remaining variables $X_4, X_5, X_6, X_7$ are all uniform error, with small variance.
+We would consider $T=(X_1, X_2, X_3)$ to be the true model.
+
+#### Scaling the data
+
+It is beneficial to define the algorithm on data having a standard scale.
+Here we use [0, 1] scaling, but by default include the aspect ratio produced by the NLDR $(r_1, r_2, ..., r_k)$, by allowing for the upper bound to vary, $(y_{1,\text{max}}, y_{2,\text{max}}, ..., y_{k,\text{max}})$.
+
+When $k=2$ which is assumed for hexagon binning, $(y_{1,\text{max}}=1$ and $y_{2,\text{max}} = \frac{r_2}{r_1}\times \frac{2}{\sqrt{3}}$.
+(The $\frac{2}{\sqrt{3}}$ accounts for the different height and width of a regular hexagon.)
+
+#### Hexagon binning
+
+Define a hexagonal grid across the \kD{} space, using hexagons with fixed height ($h_b$), width ($w_b$).
+The starting coordinates $(x_s, y_s)$ mark the lower left of the grid.
+<!--We deliberately separate out the creation of the hexagonal grid from the mapping of points on the grid.-->
+Each of the \kD{} points will belong to a hexagonal bin. That is, for each $y \in \mathbfit{Y}$, we can (uniquely) identify the hexagon that the point belongs to.
+
+#### Obtaining bin centroids
+
+Let $h : \mathbb{R}^{n\times k} \rightarrow \mathbb{R}^{m\times k}$ be the function that maps a point in \kD{} space to it's hexagonal centroid ($C^{(k)}$).
+
+#### Triangulating centroids
+
 ### Displaying the model in \pD{}

### 2024-04-15 paper.tex
steps for hexagon binning penciled out

@@
 hexagonally bin the data, and connect the bin centroids. This process
-serves reduce some noise in the resulting surface shown in \pD{}.
-
-Define a hexagonal grid across the \kD{} space, using hexagons with
-fixed height (\(h_b\)), width (\(w_b\)). The starting coordinates
-\((x_s, y_s)\) mark the lower left of the grid. Each of the \kD{} points
-will belong to a hexagonal bin. That is, for each
-\(y \in \mathbfit{Y}\), we can (uniquely) identify the hexagon that the
-point belongs to. Let
-\(h : \mathbb{R}^{n\times k} \rightarrow \mathbb{R}^{m\times k}\) be the
-function that maps a point in \kD{} space to it's hexagonal centroid
-(\(C^{(k)}\)).
-
-To illustrate the method, we use \(7\text{-}D\) simulated data, which we
-call the ``S-curve''. It is constructed by setting
-\(X_1 = \sin(a), X_2 = U(0, 2), X_3 = \text{sign}(a) \times (\cos(a) - 1), \forall a \in [-3\pi/2, 3\pi/2]\).
-The remaining variables \(X_4, X_5, X_6, X_7\) are all uniform error,
-with small variance. We would consider \(T=(X_1, X_2, X_3)\) to be the
-true model.
+serves to reduce some noisiness in the resulting surface shown in \pD{}.
+The steps in this process are shown in Figure~\ref{fig-NLDR-scurve}, and
+documented below.
 
@@
 \end{figure}%
+
+To illustrate the method, we use \(7\text{-}D\) simulated data, which we
+call the ``S-curve''. It is constructed by setting
+\(X_1 = \sin(a), X_2 = U(0, 2), X_3 = \text{sign}(a) \times (\cos(a) - 1), \forall a \in [-3\pi/2, 3\pi/2]\).
+The remaining variables \(X_4, X_5, X_6, X_7\) are all uniform error,
+with small variance. We would consider \(T=(X_1, X_2, X_3)\) to be the
+true model.
+
+\subsubsection{Scaling the data}\label{scaling-the-data}
+
+It is beneficial to define the algorithm on data having a standard
+scale. Here we use {[}0, 1{]} scaling, but by default include the aspect
+ratio produced by the NLDR \((r_1, r_2, ..., r_k)\), by allowing for the
+upper bound to vary,
+\((y_{1,\text{max}}, y_{2,\text{max}}, ..., y_{k,\text{max}})\).
+
+When \(k=2\) which is assumed for hexagon binning,
+\((y_{1,\text{max}}=1\) and
+\(y_{2,\text{max}} = \frac{r_2}{r_1}\times \frac{2}{\sqrt{3}}\). (The
+\(\frac{2}{\sqrt{3}}\) accounts for the different height and width of a
+regular hexagon.)
+
+\subsubsection{Hexagon binning}\label{hexagon-binning}
+
+Define a hexagonal grid across the \kD{} space, using hexagons with
+fixed height (\(h_b\)), width (\(w_b\)). The starting coordinates
+\((x_s, y_s)\) mark the lower left of the grid. Each of the \kD{} points
+will belong to a hexagonal bin. That is, for each
+\(y \in \mathbfit{Y}\), we can (uniquely) identify the hexagon that the
+point belongs to.
+
+\subsubsection{Obtaining bin centroids}\label{obtaining-bin-centroids}
+
+Let \(h : \mathbb{R}^{n\times k} \rightarrow \mathbb{R}^{m\times k}\) be
+the function that maps a point in \kD{} space to it's hexagonal centroid
+(\(C^{(k)}\)).
+
+\subsubsection{Triangulating centroids}\label{triangulating-centroids}
 

### 2024-04-20 paper.qmd
Add initial wordings for triangulation
Added initial setup for algorithm writing

@@
 
+#### Computing hexagonal grid configurations
+
+To define a hexagonal grid across the \kD{} space, using hexagons with fixed height ($h_b$), width ($w_b$), need to compute the how many hexagons should represented by each axes in \kD{} space.
+
+When $k=2$, the number of bins along the $x$ and $y$ axes, $b_1$ and $b_2$ should computed by accounting ${r_1, r_2}$, offset along the axes, and height ($h_b$), width ($w_b$).
+
 #### Hexagon binning
@@
 
+When $k = 2$
+
+Delaunay triangulation of $C^{(k)}$ is a triangulation of the convex hull of $C^{(k)}$ such that the circumcircle of every triangle in the triangulation contains no other points from $C^{(k)}$.
+
 ### Displaying the model in \pD{}
@@
 with corresponding \pD{} vector $\vec{f_i}$. Then, $f(h\circ g)(x)$ maps a \pD{} point to the \pD{} model estimate, completing the model cycle.
+
+When $k=2$. which bin centroids should connected are identified as a result of triangulation. Therefore, corresponding \pD{} points should connected in the same order.
 

### 2024-04-20 paper.tex
Add initial wordings for triangulation
Added initial setup for algorithm writing

@@
 
+\subsubsection{Computing hexagonal grid
+configurations}\label{computing-hexagonal-grid-configurations}
+
+To define a hexagonal grid across the \kD{} space, using hexagons with
+fixed height (\(h_b\)), width (\(w_b\)), need to compute the how many
+hexagons should represented by each axes in \kD{} space.
+
+When \(k=2\), the number of bins along the \(x\) and \(y\) axes, \(b_1\)
+and \(b_2\) should computed by accounting \({r_1, r_2}\), offset along
+the axes, and height (\(h_b\)), width (\(w_b\)).
+
 \subsubsection{Hexagon binning}\label{hexagon-binning}
@@
 
+When \(k = 2\)
+
+Delaunay triangulation of \(C^{(k)}\) is a triangulation of the convex
+hull of \(C^{(k)}\) such that the circumcircle of every triangle in the
+triangulation contains no other points from \(C^{(k)}\).
+
 \subsection{\texorpdfstring{Displaying the model in
@@
 cycle.
+
+When \(k=2\). which bin centroids should connected are identified as a
+result of triangulation. Therefore, corresponding \pD{} points should
+connected in the same order.
 

### 2024-04-20 paper.qmd
Added flow for triangulation

@@
 
-When $k=2$, the number of bins along the $x$ and $y$ axes, $b_1$ and $b_2$ should computed by accounting ${r_1, r_2}$, offset along the axes, and height ($h_b$), width ($w_b$).
+When $k=2$, the number of bins along the $x$ and $y$ axes, $b_1$ and $b_2$ is computed by accounting ${r_1, r_2}$, offset along the axes ($q_1$, $q_2$), and height ($h_b$), width ($w_b$) of a regular hexagon.
+
+($b_1 = \frac{r_1 + q_1}{w_b}$, $b_2 = \frac{r_2 + q_2}{0.75 \times h_b}$): $0.75 \times h_b$ acommodate to have the tesselation without spacing.
 
@@
 Define a hexagonal grid across the \kD{} space, using hexagons with fixed height ($h_b$), width ($w_b$).
-The starting coordinates $(x_s, y_s)$ mark the lower left of the grid.
+
+When $k=2$,
+
+The starting coordinates $(s_1, s_2)$ mark the lower left of the grid. This is the bottom left bin centroid. By starting from there, points are generated to fill the grid accounting $b_1$ and $b_2$.
+
 <!--We deliberately separate out the creation of the hexagonal grid from the mapping of points on the grid.-->
-Each of the \kD{} points will belong to a hexagonal bin. That is, for each $y \in \mathbfit{Y}$, we can (uniquely) identify the hexagon that the point belongs to.
+Each of the \kD{} points will belong to a hexagonal bin.
+That is, for each $y \in \mathbfit{Y}$, we can (uniquely) identify the hexagon that the point belongs to.
+This identification is done finding the nearest bin centroid for the \kD{} points by considering the \kD{} Euclidean distance.
 
@@
 
-#### Triangulating centroids
+#### Generating edges
+
+computes a set of simplices such that no data points of X are contained in any circumspheres of the simplices.
 

### 2024-04-20 paper.tex
Added flow for triangulation

@@
 When \(k=2\), the number of bins along the \(x\) and \(y\) axes, \(b_1\)
-and \(b_2\) should computed by accounting \({r_1, r_2}\), offset along
-the axes, and height (\(h_b\)), width (\(w_b\)).
+and \(b_2\) is computed by accounting \({r_1, r_2}\), offset along the
+axes (\(q_1\), \(q_2\)), and height (\(h_b\)), width (\(w_b\)) of a
+regular hexagon.
+
+(\(b_1 = \frac{r_1 + q_1}{w_b}\),
+\(b_2 = \frac{r_2 + q_2}{0.75 \times h_b}\)): \(0.75 \times h_b\)
+acommodate to have the tesselation without spacing.
 
@@
 Define a hexagonal grid across the \kD{} space, using hexagons with
-fixed height (\(h_b\)), width (\(w_b\)). The starting coordinates
-\((x_s, y_s)\) mark the lower left of the grid. Each of the \kD{} points
-will belong to a hexagonal bin. That is, for each
-\(y \in \mathbfit{Y}\), we can (uniquely) identify the hexagon that the
-point belongs to.
+fixed height (\(h_b\)), width (\(w_b\)).
+
+When \(k=2\),
+
+The starting coordinates \((s_1, s_2)\) mark the lower left of the grid.
+This is the bottom left bin centroid. By starting from there, points are
+generated to fill the grid accounting \(b_1\) and \(b_2\).
+
+Each of the \kD{} points will belong to a hexagonal bin. That is, for
+each \(y \in \mathbfit{Y}\), we can (uniquely) identify the hexagon that
+the point belongs to. This identification is done finding the nearest
+bin centroid for the \kD{} points by considering the \kD{} Euclidean
+distance.
 
@@
 
-\subsubsection{Triangulating centroids}\label{triangulating-centroids}
+\subsubsection{Generating edges}\label{generating-edges}
+
+computes a set of simplices such that no data points of X are contained
+in any circumspheres of the simplices.
 

### 2024-04-21 paper.qmd
scaling data...right

@@
 
-When $k=2$ which is assumed for hexagon binning, $(y_{1,\text{max}}=1$ and $y_{2,\text{max}} = \frac{r_2}{r_1}\times \frac{2}{\sqrt{3}}$.
-(The $\frac{2}{\sqrt{3}}$ accounts for the different height and width of a regular hexagon.)
+When $k=2$ which is assumed for hexagon binning, $y_{1,\text{max}}=1$ and $y_{2,\text{max}} = \frac{r_2}{r_1}\times \frac{2}{\sqrt{3}}$.
+(The $\frac{2}{\sqrt{3}}$ accounts for the different height ($h_b$) and width ($w_b$) of a regular hexagon.)
 

### 2024-04-21 paper.qmd
Update the algorithm section

@@
 When $k=2$ which is assumed for hexagon binning, $y_{1,\text{max}}=1$ and $y_{2,\text{max}} = \frac{r_2}{r_1}\times \frac{2}{\sqrt{3}}$.
-(The $\frac{2}{\sqrt{3}}$ accounts for the different height ($h_b$) and width ($w_b$) of a regular hexagon.)
+(The $\frac{2}{\sqrt{3}}$ accounts for the different height ($h_b$) and width ($w_b$) of a regular hexagon.) (@fig-NLDR-scurve (a))
 
@@
 
-To define a hexagonal grid across the \kD{} space, using hexagons with fixed height ($h_b$), width ($w_b$), need to compute the how many hexagons should represented by each axes in \kD{} space.
+To define a hexagonal grid across the \kD{} space, with hexagons of fixed height ($h_b$), width ($w_b$),  it is necessary to determine how many hexagons should be represented along each axis in the \kD{} space.
 
-When $k=2$, the number of bins along the $x$ and $y$ axes, $b_1$ and $b_2$ is computed by accounting ${r_1, r_2}$, offset along the axes ($q_1$, $q_2$), and height ($h_b$), width ($w_b$) of a regular hexagon.
-
-($b_1 = \frac{r_1 + q_1}{w_b}$, $b_2 = \frac{r_2 + q_2}{0.75 \times h_b}$): $0.75 \times h_b$ acommodate to have the tesselation without spacing.
+When $k=2$, the number of bins along the $x$ and $y$ axes, $b_1$ and $b_2$, is computed by considering the scaling factors ${r_1, r_2}$, along with the offset along the axes ($q_1$, $q_2$), and the height ($h_b$) and width ($w_b$) of a regular hexagon (@fig-NLDR-scurve (b)).
+($b_1 = \frac{r_1 + q_1}{w_b}$, and $b_2 = \frac{r_2 + q_2}{0.75 \times h_b}$) where $0.75 \times h_b$ accommodate to have hexagons fill the entire 2D space without leaving any gaps between them.)
 
@@
 Define a hexagonal grid across the \kD{} space, using hexagons with fixed height ($h_b$), width ($w_b$).
-
-When $k=2$,
-
-The starting coordinates $(s_1, s_2)$ mark the lower left of the grid. This is the bottom left bin centroid. By starting from there, points are generated to fill the grid accounting $b_1$ and $b_2$.
-
-<!--We deliberately separate out the creation of the hexagonal grid from the mapping of points on the grid.-->
 Each of the \kD{} points will belong to a hexagonal bin.
@@
 
-#### Obtaining bin centroids
+When $k=2$, the starting coordinates $(s_1, s_2)$ mark the lower left of the grid.
+This is the bottom left bin centroid.
+By starting from there, points are generated to fill the grid accounting $b_1$, $b_2$, $w_b$, and $h_b$.
 
-Let $h : \mathbb{R}^{n\times k} \rightarrow \mathbb{R}^{m\times k}$ be the function that maps a point in \kD{} space to it's hexagonal centroid ($C^{(k)}$).
+<!--We deliberately separate out the creation of the hexagonal grid from the mapping of points on the grid.-->
+#### Obtaining bin representations
+
+Let $h : \mathbb{R}^{n\times k} \rightarrow \mathbb{R}^{m\times k}$ be the function that maps a point in \kD{} space to it's bin representation ($C^{(k)}$). This is the model points in \kD{} space.
+
+When $k = 2$, one of the representations of a bin is the hexagonal centroid (@fig-NLDR-scurve (c)).
 
@@
 
-computes a set of simplices such that no data points of X are contained in any circumspheres of the simplices.
+Delaunay triangulation on bin representations generates edges by specifying which bin representations should be connected to generate the model in \kD{} space.
 
-When $k = 2$
-
-Delaunay triangulation of $C^{(k)}$ is a triangulation of the convex hull of $C^{(k)}$ such that the circumcircle of every triangle in the triangulation contains no other points from $C^{(k)}$.
+When $k = 2$ Delaunay triangulation on $C^{(2)}$ generates the model in 2D space, which is a triangular mesh (@fig-NLDR-scurve (d)).
+It generates convex hulls of $C^{(2)}$ such that the circumcircle of every triangle in the triangulation contains no other points from $C^{(2)}$.
 
@@
 
-When $k=2$. which bin centroids should connected are identified as a result of triangulation. Therefore, corresponding \pD{} points should connected in the same order.
+Furthermore, edges that exist between \kD{} representations should also generate edges in \pD{} by connecting \pD{} mapping of the corresponding \kD{} representations.
 

### 2024-04-21 paper.tex
scaling data...right
Update the algorithm section

--- 2024-04-15 paper.tex
@@
 $s$ & hexagonal size (defined as the radius of the outer circle surrounding the hexagon)\\
-$(x_s, \ y_s)$ & starting coordinates of the hexagonal grid\\
-$q_x$, $q_y$ & buffer amount along the x and y axes\\
+$(s_1, \ s_2)$ & starting coordinates of the hexagonal grid\\
+$q_1$, $q_2$ & buffer amount along the x and y axes\\
 $h_s$, $v_s$ & horizontal and vertical distance between adjacent hexagon bin centroids\\
@@
 $d_x$, $d_y$ & horizontal spacing and adjusted vertical spacing for hexagonal coordinates\\
-$b_x$, $b_y$ & number of hexagon bins along the x and y axes\\
+$b_1$, $b_2$ & number of hexagon bins along the x and y axes\\
 $b$, $b'$ & total and non-empty hexagon bins in the grid\\
@@
 When \(k=2\) which is assumed for hexagon binning,
-\((y_{1,\text{max}}=1\) and
+\(y_{1,\text{max}}=1\) and
 \(y_{2,\text{max}} = \frac{r_2}{r_1}\times \frac{2}{\sqrt{3}}\). (The
-\(\frac{2}{\sqrt{3}}\) accounts for the different height and width of a
-regular hexagon.)
+\(\frac{2}{\sqrt{3}}\) accounts for the different height (\(h_b\)) and
+width (\(w_b\)) of a regular hexagon.) (Figure~\ref{fig-NLDR-scurve}
+(a))
+
+\subsubsection{Computing hexagonal grid
+configurations}\label{computing-hexagonal-grid-configurations}
+
+To define a hexagonal grid across the \kD{} space, with hexagons of
+fixed height (\(h_b\)), width (\(w_b\)), it is necessary to determine
+how many hexagons should be represented along each axis in the \kD{}
+space.
+
+When \(k=2\), the number of bins along the \(x\) and \(y\) axes, \(b_1\)
+and \(b_2\), is computed by considering the scaling factors
+\({r_1, r_2}\), along with the offset along the axes (\(q_1\), \(q_2\)),
+and the height (\(h_b\)) and width (\(w_b\)) of a regular hexagon
+(Figure~\ref{fig-NLDR-scurve} (b)). (\(b_1 = \frac{r_1 + q_1}{w_b}\),
+and \(b_2 = \frac{r_2 + q_2}{0.75 \times h_b}\)) where
+\(0.75 \times h_b\) accommodate to have hexagons fill the entire 2D
+space without leaving any gaps between them.)
 
@@
 Define a hexagonal grid across the \kD{} space, using hexagons with
-fixed height (\(h_b\)), width (\(w_b\)). The starting coordinates
-\((x_s, y_s)\) mark the lower left of the grid. Each of the \kD{} points
-will belong to a hexagonal bin. That is, for each
-\(y \in \mathbfit{Y}\), we can (uniquely) identify the hexagon that the
-point belongs to.
+fixed height (\(h_b\)), width (\(w_b\)). Each of the \kD{} points will
+belong to a hexagonal bin. That is, for each \(y \in \mathbfit{Y}\), we
+can (uniquely) identify the hexagon that the point belongs to. This
+identification is done finding the nearest bin centroid for the \kD{}
+points by considering the \kD{} Euclidean distance.
 
-\subsubsection{Obtaining bin centroids}\label{obtaining-bin-centroids}
+When \(k=2\), the starting coordinates \((s_1, s_2)\) mark the lower
+left of the grid. This is the bottom left bin centroid. By starting from
+there, points are generated to fill the grid accounting \(b_1\),
+\(b_2\), \(w_b\), and \(h_b\).
+
+\subsubsection{Obtaining bin
+representations}\label{obtaining-bin-representations}
 
 Let \(h : \mathbb{R}^{n\times k} \rightarrow \mathbb{R}^{m\times k}\) be
-the function that maps a point in \kD{} space to it's hexagonal centroid
-(\(C^{(k)}\)).
+the function that maps a point in \kD{} space to it's bin representation
+(\(C^{(k)}\)). This is the model points in \kD{} space.
 
-\subsubsection{Triangulating centroids}\label{triangulating-centroids}
+When \(k = 2\), one of the representations of a bin is the hexagonal
+centroid (Figure~\ref{fig-NLDR-scurve} (c)).
+
+\subsubsection{Generating edges}\label{generating-edges}
+
+Delaunay triangulation on bin representations generates edges by
+specifying which bin representations should be connected to generate the
+model in \kD{} space.
+
+When \(k = 2\) Delaunay triangulation on \(C^{(2)}\) generates the model
+in 2D space, which is a triangular mesh (Figure~\ref{fig-NLDR-scurve}
+(d)). It generates convex hulls of \(C^{(2)}\) such that the
+circumcircle of every triangle in the triangulation contains no other
+points from \(C^{(2)}\).
 
@@
 cycle.
+
+Furthermore, edges that exist between \kD{} representations should also
+generate edges in \pD{} by connecting \pD{} mapping of the corresponding
+\kD{} representations.
 

### 2024-04-21 paper.qmd
Update with quollr new

@@
 num_bins_x_scurve <- 6
-num_bins_y_scurve <- 14
+num_bins_y_scurve <- 10
 hex_size_scurve <- 0.13

### 2024-04-21 paper.qmd
Update the plots

@@
 sc_ltr_pos <- c(0.08, 0.96)
+# sc_xlims <- c(-0.25, 1.25)
+# sc_ylims <- c(-0.25, 2.59)
 sc_xlims <- c(-0.25, 1.25)
-sc_ylims <- c(-0.25, 2.59)
+sc_ylims <- c(-0.25, 1.8)
+
 nldr_scurve <- umap_scurve_scaled |> 

### 2024-04-22 paper.qmd
Update the hex height and width notations

@@
 When $k=2$ which is assumed for hexagon binning, $y_{1,\text{max}}=1$ and $y_{2,\text{max}} = \frac{r_2}{r_1}\times \frac{2}{\sqrt{3}}$.
-(The $\frac{2}{\sqrt{3}}$ accounts for the different height ($h_b$) and width ($w_b$) of a regular hexagon.) (@fig-NLDR-scurve (a))
+(The $\frac{2}{\sqrt{3}}$ accounts for the different height ($a_1$) and width ($a_2$) of a regular hexagon.) (@fig-NLDR-scurve (a))
 
@@
 
-To define a hexagonal grid across the \kD{} space, with hexagons of fixed height ($h_b$), width ($w_b$),  it is necessary to determine how many hexagons should be represented along each axis in the \kD{} space.
+To define a hexagonal grid across the \kD{} space, with hexagons of fixed height ($a_1$), width ($a_2$),  it is necessary to determine how many hexagons should be represented along each axis in the \kD{} space.
 
-When $k=2$, the number of bins along the $x$ and $y$ axes, $b_1$ and $b_2$, is computed by considering the scaling factors ${r_1, r_2}$, along with the offset along the axes ($q_1$, $q_2$), and the height ($h_b$) and width ($w_b$) of a regular hexagon (@fig-NLDR-scurve (b)).
-($b_1 = \frac{r_1 + q_1}{w_b}$, and $b_2 = \frac{r_2 + q_2}{0.75 \times h_b}$) where $0.75 \times h_b$ accommodate to have hexagons fill the entire 2D space without leaving any gaps between them.)
+When $k=2$, the number of bins along the $x$ and $y$ axes, $b_1$ and $b_2$, is computed by considering the scaling factors ${r_1, r_2}$, along with the offset along the axes ($q_1$, $q_2$), and the height ($a_1$) and width ($a_2$) of a regular hexagon (@fig-NLDR-scurve (b)).
+($b_1 = \frac{r_1 + q_1}{a_2}$, and $b_2 = \frac{r_2 + q_2}{0.75 \times a_1}$) where $0.75 \times a_1$ accommodate to have hexagons fill the entire 2D space without leaving any gaps between them.)
 
@@
 
-Define a hexagonal grid across the \kD{} space, using hexagons with fixed height ($h_b$), width ($w_b$).
+Define a hexagonal grid across the \kD{} space, using hexagons with fixed height ($a_1$), width ($a_2$).
 Each of the \kD{} points will belong to a hexagonal bin.
@@
 This is the bottom left bin centroid.
-By starting from there, points are generated to fill the grid accounting $b_1$, $b_2$, $w_b$, and $h_b$.
+By starting from there, points are generated to fill the grid accounting $b_1$, $b_2$, $a_2$, and $a_1$.
 

### 2024-04-22 paper.tex
Update the hex height and width notations

@@
 \(y_{2,\text{max}} = \frac{r_2}{r_1}\times \frac{2}{\sqrt{3}}\). (The
-\(\frac{2}{\sqrt{3}}\) accounts for the different height (\(h_b\)) and
-width (\(w_b\)) of a regular hexagon.) (Figure~\ref{fig-NLDR-scurve}
+\(\frac{2}{\sqrt{3}}\) accounts for the different height (\(a_1\)) and
+width (\(a_2\)) of a regular hexagon.) (Figure~\ref{fig-NLDR-scurve}
 (a))
@@
 To define a hexagonal grid across the \kD{} space, with hexagons of
-fixed height (\(h_b\)), width (\(w_b\)), it is necessary to determine
+fixed height (\(a_1\)), width (\(a_2\)), it is necessary to determine
 how many hexagons should be represented along each axis in the \kD{}
@@
 \({r_1, r_2}\), along with the offset along the axes (\(q_1\), \(q_2\)),
-and the height (\(h_b\)) and width (\(w_b\)) of a regular hexagon
-(Figure~\ref{fig-NLDR-scurve} (b)). (\(b_1 = \frac{r_1 + q_1}{w_b}\),
-and \(b_2 = \frac{r_2 + q_2}{0.75 \times h_b}\)) where
-\(0.75 \times h_b\) accommodate to have hexagons fill the entire 2D
+and the height (\(a_1\)) and width (\(a_2\)) of a regular hexagon
+(Figure~\ref{fig-NLDR-scurve} (b)). (\(b_1 = \frac{r_1 + q_1}{a_2}\),
+and \(b_2 = \frac{r_2 + q_2}{0.75 \times a_1}\)) where
+\(0.75 \times a_1\) accommodate to have hexagons fill the entire 2D
 space without leaving any gaps between them.)
@@
 Define a hexagonal grid across the \kD{} space, using hexagons with
-fixed height (\(h_b\)), width (\(w_b\)). Each of the \kD{} points will
+fixed height (\(a_1\)), width (\(a_2\)). Each of the \kD{} points will
 belong to a hexagonal bin. That is, for each \(y \in \mathbfit{Y}\), we
@@
 there, points are generated to fill the grid accounting \(b_1\),
-\(b_2\), \(w_b\), and \(h_b\).
+\(b_2\), \(a_2\), and \(a_1\).
 

### 2024-04-22 paper.qmd
Changed the scale
re-working hexagon grid computation

@@
 A simple approach would be to triangulate the points and add edges.
-A more stable approach is to first hexagonally bin the data, and connect the bin centroids.
+A more stable approach is to first hexagonally bin the data, reducing it from $n$ to $m\leq n$ observations, and connect the bin centroids.
 This process serves to reduce some noisiness in the resulting surface shown in \pD{}.
@@
 
-## Hexagonal bining to have regular hexagons
+## Hexagonal binning to have regular hexagons
 hb_obj_scurve <- hex_binning(
@@
 It is beneficial to define the algorithm on data having a standard scale.
-Here we use [0, 1] scaling, but by default include the aspect ratio produced by the NLDR $(r_1, r_2, ..., r_k)$, by allowing for the upper bound to vary, $(y_{1,\text{max}}, y_{2,\text{max}}, ..., y_{k,\text{max}})$.
+Here the variables are scaled to [0, 1], but the upper bound can incorporate the aspect ratio produced by the NLDR $(r_1, r_2, ..., r_k)$, by setting them to $(y_{1,\text{max}}, y_{2,\text{max}}, ..., y_{k,\text{max}})$.
+When $k=2$ which is assumed for hexagon binning, $y_{1,\text{max}}=1$ and $y_{2,\text{max}} = \frac{r_2}{r_1}$, as observed in @fig-NLDR-scurve.
+<!-- \times \frac{2}{\sqrt{3}}$.
+(The $\frac{2}{\sqrt{3}}$ accounts for the different height ($a_1$) and width ($a_2$) of a regular hexagon.) The scaling of data should take the size of the hexagons into account, but choice of number of bins should.
+-->
 
-When $k=2$ which is assumed for hexagon binning, $y_{1,\text{max}}=1$ and $y_{2,\text{max}} = \frac{r_2}{r_1}\times \frac{2}{\sqrt{3}}$.
-(The $\frac{2}{\sqrt{3}}$ accounts for the different height ($a_1$) and width ($a_2$) of a regular hexagon.) (@fig-NLDR-scurve (a))
+#### Computing hexagon grid configurations
 
-#### Computing hexagonal grid configurations
+The 2D hexagon grid is defined by the number of bins in each direction ($b_1, b_2$), and the lower left position where the grid starts ($s_1, s_2$).
+The values of $s_i$ need to be below their respective minimum variable values, and could be a full bin lower, to allow a buffer corresponding to a full hexagon width and height around the data.
+The values of $b_i$ are variables to be computed that define the reduction in size of the data ($n$ to $m$).
 
-To define a hexagonal grid across the \kD{} space, with hexagons of fixed height ($a_1$), width ($a_2$),  it is necessary to determine how many hexagons should be represented along each axis in the \kD{} space.
+Number of bins is set by fixing $b_1$, which determines the binwidth and breaks on $x_1$, is calculated as
+
+The computed binwidth then determines the number of vertical bins, $b_2$, and vertical binwidth, which are computed based on $r_2$, and the height/width ratio of a regular hexagon, $\frac{2}{\sqrt{3}}$ as
+
+<!-- To define a hexagonal grid across the \kD{} space, with hexagons of fixed height ($a_1$), width ($a_2$),  it is necessary to determine how many hexagons should be represented along each axis in the \kD{} space.
 
 When $k=2$, the number of bins along the $x$ and $y$ axes, $b_1$ and $b_2$, is computed by considering the scaling factors ${r_1, r_2}$, along with the offset along the axes ($q_1$, $q_2$), and the height ($a_1$) and width ($a_2$) of a regular hexagon (@fig-NLDR-scurve (b)).
-($b_1 = \frac{r_1 + q_1}{a_2}$, and $b_2 = \frac{r_2 + q_2}{0.75 \times a_1}$) where $0.75 \times a_1$ accommodate to have hexagons fill the entire 2D space without leaving any gaps between them.)
+($b_1 = \frac{r_1 + q_1}{a_2}$, and $b_2 = \frac{r_2 + q_2}{0.75 \times a_1}$) where $0.75 \times a_1$ accommodate to have hexagons fill the entire 2D space without leaving any gaps between them.)-->
+
+<!-- XXX We don't need parameters for the regular hexagon, these are fixed constants. -->
 

### 2024-04-22 paper.tex
re-working hexagon grid computation

@@
 triangulate the points and add edges. A more stable approach is to first
-hexagonally bin the data, and connect the bin centroids. This process
-serves to reduce some noisiness in the resulting surface shown in \pD{}.
-The steps in this process are shown in Figure~\ref{fig-NLDR-scurve}, and
+hexagonally bin the data, reducing it from \(n\) to \(m\leq n\)
+observations, and connect the bin centroids. This process serves to
+reduce some noisiness in the resulting surface shown in \pD{}. The steps
+in this process are shown in Figure~\ref{fig-NLDR-scurve}, and
 documented below.
@@
 It is beneficial to define the algorithm on data having a standard
-scale. Here we use {[}0, 1{]} scaling, but by default include the aspect
-ratio produced by the NLDR \((r_1, r_2, ..., r_k)\), by allowing for the
-upper bound to vary,
-\((y_{1,\text{max}}, y_{2,\text{max}}, ..., y_{k,\text{max}})\).
+scale. Here the variables are scaled to {[}0, 1{]}, but the upper bound
+can incorporate the aspect ratio produced by the NLDR
+\((r_1, r_2, ..., r_k)\), by setting them to
+\((y_{1,\text{max}}, y_{2,\text{max}}, ..., y_{k,\text{max}})\). When
+\(k=2\) which is assumed for hexagon binning, \(y_{1,\text{max}}=1\) and
+\(y_{2,\text{max}} = \frac{r_2}{r_1}\), as observed in
+Figure~\ref{fig-NLDR-scurve}.
 
-When \(k=2\) which is assumed for hexagon binning,
-\(y_{1,\text{max}}=1\) and
-\(y_{2,\text{max}} = \frac{r_2}{r_1}\times \frac{2}{\sqrt{3}}\). (The
-\(\frac{2}{\sqrt{3}}\) accounts for the different height (\(a_1\)) and
-width (\(a_2\)) of a regular hexagon.) (Figure~\ref{fig-NLDR-scurve}
-(a))
+\subsubsection{Computing hexagon grid
+configurations}\label{computing-hexagon-grid-configurations}
 
-\subsubsection{Computing hexagonal grid
-configurations}\label{computing-hexagonal-grid-configurations}
+The 2D hexagon grid is defined by the number of bins in each direction
+(\(b_1, b_2\)), and the lower left position where the grid starts
+(\(s_1, s_2\)). The values of \(s_i\) need to be below their respective
+minimum variable values, and could be a full bin lower, to allow a
+buffer corresponding to a full hexagon width and height around the data.
+The values of \(b_i\) are variables to be computed that define the
+reduction in size of the data (\(n\) to \(m\)).
 
-To define a hexagonal grid across the \kD{} space, with hexagons of
-fixed height (\(a_1\)), width (\(a_2\)), it is necessary to determine
-how many hexagons should be represented along each axis in the \kD{}
-space.
+Number of bins is set by fixing \(b_1\), which determines the binwidth
+and breaks on \(x_1\), is calculated as
 
-When \(k=2\), the number of bins along the \(x\) and \(y\) axes, \(b_1\)
-and \(b_2\), is computed by considering the scaling factors
-\({r_1, r_2}\), along with the offset along the axes (\(q_1\), \(q_2\)),
-and the height (\(a_1\)) and width (\(a_2\)) of a regular hexagon
-(Figure~\ref{fig-NLDR-scurve} (b)). (\(b_1 = \frac{r_1 + q_1}{a_2}\),
-and \(b_2 = \frac{r_2 + q_2}{0.75 \times a_1}\)) where
-\(0.75 \times a_1\) accommodate to have hexagons fill the entire 2D
-space without leaving any gaps between them.)
+The computed binwidth then determines the number of vertical bins,
+\(b_2\), and vertical binwidth, which are computed based on \(r_2\), and
+the height/width ratio of a regular hexagon, \(\frac{2}{\sqrt{3}}\) as
 

### 2024-04-22 paper.qmd
Added equations

@@
 sc_xlims <- c(-0.25, 1.25)
-sc_ylims <- c(-0.25, 1.8)
+sc_ylims <- c(-0.25, 2)
 
@@
 The 2D hexagon grid is defined by the number of bins in each direction ($b_1, b_2$), and the lower left position where the grid starts ($s_1, s_2$).
-The values of $s_i$ need to be below their respective minimum variable values, and could be a full bin lower, to allow a buffer corresponding to a full hexagon width and height around the data.
+The values of $s_i$ need to be below their respective minimum variable values, and could be a full bin lower, to allow a buffer corresponding to a full hexagon width ($a_1$) and height ($a_2$) around the data.
 The values of $b_i$ are variables to be computed that define the reduction in size of the data ($n$ to $m$).
@@
 
+$$
+b_1 = \frac{r_1 + q_1}{a_1}.
+$$
+
 The computed binwidth then determines the number of vertical bins, $b_2$, and vertical binwidth, which are computed based on $r_2$, and the height/width ratio of a regular hexagon, $\frac{2}{\sqrt{3}}$ as
+
+$$
+b_2 = \frac{r_2 + q_2}{0.75 \times a_2}.
+$$
 
@@
 
-Define a hexagonal grid across the \kD{} space, using hexagons with fixed height ($a_1$), width ($a_2$).
+Define a hexagonal grid across the \kD{} space, using hexagons with fixed height ($a_2$), width ($a_1$).
 Each of the \kD{} points will belong to a hexagonal bin.
@@
 This is the bottom left bin centroid.
-By starting from there, points are generated to fill the grid accounting $b_1$, $b_2$, $a_2$, and $a_1$.
+By starting from there, points are generated to fill the grid accounting $b_1$, $b_2$, $a_1$, and $a_2$.
 

### 2024-04-22 paper.tex
Added equations

@@
 minimum variable values, and could be a full bin lower, to allow a
-buffer corresponding to a full hexagon width and height around the data.
-The values of \(b_i\) are variables to be computed that define the
-reduction in size of the data (\(n\) to \(m\)).
+buffer corresponding to a full hexagon width (\(a_1\)) and height
+(\(a_2\)) around the data. The values of \(b_i\) are variables to be
+computed that define the reduction in size of the data (\(n\) to \(m\)).
 
@@
 and breaks on \(x_1\), is calculated as
+
+\[
+b_1 = \frac{r_1 + q_1}{a_1}.
+\]
 
@@
 
+\[
+b_2 = \frac{r_2 + q_2}{0.75 \times a_2}.
+\]
+
 \subsubsection{Hexagon binning}\label{hexagon-binning}
@@
 Define a hexagonal grid across the \kD{} space, using hexagons with
-fixed height (\(a_1\)), width (\(a_2\)). Each of the \kD{} points will
+fixed height (\(a_2\)), width (\(a_1\)). Each of the \kD{} points will
 belong to a hexagonal bin. That is, for each \(y \in \mathbfit{Y}\), we
@@
 there, points are generated to fill the grid accounting \(b_1\),
-\(b_2\), \(a_2\), and \(a_1\).
+\(b_2\), \(a_1\), and \(a_2\).
 

### 2024-04-22 paper.qmd
Updated with quollr new

@@
 sc_ltr_pos <- c(0.08, 0.96)
+sc_xlims <- c(-0.25, 1.25)
+sc_ylims <- c(-0.25, 2.59)
 # sc_xlims <- c(-0.25, 1.25)
-# sc_ylims <- c(-0.25, 2.59)
-sc_xlims <- c(-0.25, 1.25)
-sc_ylims <- c(-0.25, 2)
+# sc_ylims <- c(-0.25, 2)
 
@@
 num_bins_x_scurve <- 6
-num_bins_y_scurve <- 10
+num_bins_y_scurve <- 14
 hex_size_scurve <- 0.13

### 2024-04-23 paper.qmd
Add offset def

@@
 
-Number of bins is set by fixing $b_1$, which determines the binwidth and breaks on $x_1$, is calculated as
+Number of bins is set by fixing $b_1$, which determines the binwidth accounting the offset $q_1$ and breaks on $x_1$, is calculated as
 
@@
 
-The computed binwidth then determines the number of vertical bins, $b_2$, and vertical binwidth, which are computed based on $r_2$, and the height/width ratio of a regular hexagon, $\frac{2}{\sqrt{3}}$ as
+The computed binwidth then determines the number of vertical bins, $b_2$, and vertical binwidth, which are computed based on $r_2$, the offset $q_2$, and the height/width ratio of a regular hexagon, $\frac{2}{\sqrt{3}}$ as
 

### 2024-04-23 paper.tex
Add offset def

@@
 Number of bins is set by fixing \(b_1\), which determines the binwidth
-and breaks on \(x_1\), is calculated as
+accounting the offset \(q_1\) and breaks on \(x_1\), is calculated as
 
@@
 The computed binwidth then determines the number of vertical bins,
-\(b_2\), and vertical binwidth, which are computed based on \(r_2\), and
-the height/width ratio of a regular hexagon, \(\frac{2}{\sqrt{3}}\) as
+\(b_2\), and vertical binwidth, which are computed based on \(r_2\), the
+offset \(q_2\), and the height/width ratio of a regular hexagon,
+\(\frac{2}{\sqrt{3}}\) as
 

### 2024-04-23 paper.qmd
Changed formula to obtain binwidth
working on binning the data section

@@
 $$
-b_1 = \frac{r_1 + q_1}{a_1}.
+a_1 = \frac{r_1 + q_1}{b_1}.
 $$
@@
 $$
-b_2 = \frac{r_2 + q_2}{0.75 \times a_2}.
+a_2 = \frac{r_2 + q_2}{0.75 \times b_2}.
 $$
@@
 
-#### Hexagon binning
+#### Binning the data
 
+Points are allocated to the bin that they fall in. XXX How are you doing this? Do you check the bounds of the hexagon? Or do you use distance to centroid?
+
+<!--
 Define a hexagonal grid across the \kD{} space, using hexagons with fixed height ($a_2$), width ($a_1$).
@@
 By starting from there, points are generated to fill the grid accounting $b_1$, $b_2$, $a_1$, and $a_2$.
+-->
 

### 2024-04-23 paper.tex
Changed formula to obtain binwidth
working on binning the data section

@@
 \[
-b_1 = \frac{r_1 + q_1}{a_1}.
+a_1 = \frac{r_1 + q_1}{b_1}.
 \]
@@
 \[
-b_2 = \frac{r_2 + q_2}{0.75 \times a_2}.
+a_2 = \frac{r_2 + q_2}{0.75 \times b_2}.
 \]
 
-\subsubsection{Hexagon binning}\label{hexagon-binning}
+\subsubsection{Binning the data}\label{binning-the-data}
 
-Define a hexagonal grid across the \kD{} space, using hexagons with
-fixed height (\(a_2\)), width (\(a_1\)). Each of the \kD{} points will
-belong to a hexagonal bin. That is, for each \(y \in \mathbfit{Y}\), we
-can (uniquely) identify the hexagon that the point belongs to. This
-identification is done finding the nearest bin centroid for the \kD{}
-points by considering the \kD{} Euclidean distance.
-
-When \(k=2\), the starting coordinates \((s_1, s_2)\) mark the lower
-left of the grid. This is the bottom left bin centroid. By starting from
-there, points are generated to fill the grid accounting \(b_1\),
-\(b_2\), \(a_1\), and \(a_2\).
+Points are allocated to the bin that they fall in. XXX How are you doing
+this? Do you check the bounds of the hexagon? Or do you use distance to
+centroid?
 

### 2024-04-28 paper.qmd
Add binning data dist

@@
 
-Number of bins is set by fixing $b_1$, which determines the binwidth accounting the offset $q_1$ and breaks on $x_1$, is calculated as
+The value for $b_2$ is computed by fixing $b_1$.
 
-$$
-a_1 = \frac{r_1 + q_1}{b_1}.
-$$
+<!-- Number of bins is set by fixing $b_1$, which determines the binwidth accounting the offset $q_1$ and breaks on $x_1$, is calculated as  -->
 
-The computed binwidth then determines the number of vertical bins, $b_2$, and vertical binwidth, which are computed based on $r_2$, the offset $q_2$, and the height/width ratio of a regular hexagon, $\frac{2}{\sqrt{3}}$ as
+<!-- $$ -->
+<!-- a_1 = \frac{r_1 + q_1}{b_1}. -->
+<!-- $$ -->
 
-$$
-a_2 = \frac{r_2 + q_2}{0.75 \times b_2}.
-$$
+<!-- The computed binwidth then determines the number of vertical bins, $b_2$, and vertical binwidth, which are computed based on $r_2$, the offset $q_2$, and the height/width ratio of a regular hexagon, $\frac{2}{\sqrt{3}}$ as -->
+
+<!-- $$ -->
+<!-- a_2 = \frac{r_2 + q_2}{0.75 \times b_2}. -->
+<!-- $$ -->
 
@@
 
-Points are allocated to the bin that they fall in. XXX How are you doing this? Do you check the bounds of the hexagon? Or do you use distance to centroid?
+Points are allocated to the bin that they fall in by considering the nearest nearest centroid.
+
+<!-- XXX How are you doing this? Do you check the bounds of the hexagon? Or do you use distance to centroid? -->
 

### 2024-04-28 paper.tex
Add binning data dist

@@
 
-Number of bins is set by fixing \(b_1\), which determines the binwidth
-accounting the offset \(q_1\) and breaks on \(x_1\), is calculated as
-
-\[
-a_1 = \frac{r_1 + q_1}{b_1}.
-\]
-
-The computed binwidth then determines the number of vertical bins,
-\(b_2\), and vertical binwidth, which are computed based on \(r_2\), the
-offset \(q_2\), and the height/width ratio of a regular hexagon,
-\(\frac{2}{\sqrt{3}}\) as
-
-\[
-a_2 = \frac{r_2 + q_2}{0.75 \times b_2}.
-\]
+The value for \(b_2\) is computed by fixing \(b_1\).
 
@@
 
-Points are allocated to the bin that they fall in. XXX How are you doing
-this? Do you check the bounds of the hexagon? Or do you use distance to
-centroid?
+Points are allocated to the bin that they fall in by considering the
+nearest nearest centroid.
 

### 2024-04-28 paper.qmd
Added description for grid configurations

@@
 The value for $b_2$ is computed by fixing $b_1$.
+Considering the lower bound of the NLDR, $a_1 > 2 \times s_1$ and $a_1 > \frac{1-s_1}{b_1 -1}$.
+The minimum $a_1$ that satisfies the constraints will be selected.
+For regular hexagons, $a_2 = \frac{2}{\sqrt(3)}a_1$.
+Therefore, $a_2$ is computed for the minimum $a_1$.
+According to the upper bound of the NLDR, $b_2 = \left\lceil\frac{r_2 - s_2}{vs} + 1\right\rceil$, where $vs = \frac{a_2}{2} + \sqrt{{(\frac{a_2}{2})}^2 + {(\frac{a_1}{2})}^2}$.
 

### 2024-04-28 paper.tex
Added description for grid configurations
Correct spell mistake

@@
 
-The value for \(b_2\) is computed by fixing \(b_1\).
+The value for \(b_2\) is computed by fixing \(b_1\). Considering the
+lower bound of the NLDR, \(a_1 > 2 \times s_1\) and
+\(a_1 > \frac{1-s_1}{b_1 -1}\). The minimum \(a_1\) that satisfies the
+constraints will be selected. For regular hexagons,
+\(a_2 = \frac{2}{\sqrt(3)}a_1\). Therefore, \(a_2\) is computed for the
+minimum \(a_1\). According to the upper bound of the NLDR,
+\(b_2 = \left\lceil\frac{r_2 - s_2}{vs} + 1\right\rceil\), where
+\(vs = \frac{a_2}{2} + \sqrt{{(\frac{a_2}{2})}^2 + {(\frac{a_1}{2})}^2}\).
 
@@
 Points are allocated to the bin that they fall in by considering the
-nearest nearest centroid.
+nearest centroid.
 

### 2024-04-28 paper.qmd
Correct spell mistake
Corrected vs formular

@@
 Therefore, $a_2$ is computed for the minimum $a_1$.
-According to the upper bound of the NLDR, $b_2 = \left\lceil\frac{r_2 - s_2}{vs} + 1\right\rceil$, where $vs = \frac{a_2}{2} + \sqrt{{(\frac{a_2}{2})}^2 + {(\frac{a_1}{2})}^2}$.
+According to the upper bound of the NLDR, $b_2 = \left\lceil\frac{r_2 - s_2}{vs} + 1\right\rceil$, where $vs = \frac{a_2}{2} + \sqrt{{(\frac{a_2}{2})}^2 - {(\frac{a_1}{2})}^2}$.
 
@@
 
-Points are allocated to the bin that they fall in by considering the nearest nearest centroid.
+Points are allocated to the bin that they fall in by considering the nearest centroid.
 

### 2024-04-28 paper.qmd
Changed vs notation

@@
 Therefore, $a_2$ is computed for the minimum $a_1$.
-According to the upper bound of the NLDR, $b_2 = \left\lceil\frac{r_2 - s_2}{vs} + 1\right\rceil$, where $vs = \frac{a_2}{2} + \sqrt{{(\frac{a_2}{2})}^2 - {(\frac{a_1}{2})}^2}$.
+According to the upper bound of the NLDR, $b_2 = \left\lceil\frac{r_2 - s_2}{v_2} + 1\right\rceil$, where $v_2 = \frac{a_2}{2} + \sqrt{{(\frac{a_2}{2})}^2 - {(\frac{a_1}{2})}^2}$.
 

### 2024-04-30 paper.qmd
notation update

@@
 
-The 2D hexagon grid is defined by the number of bins in each direction ($b_1, b_2$), and the lower left position where the grid starts ($s_1, s_2$).
+The 2D hexagon grid is defined by the number of bins in each direction $(b_1, b_2)$, as given by the centroids of each hexagon $(c_1, c_2)$ and the lower left position where the grid starts at $(s_1, s_2)$, which correspond to the lowest left centroid.
 The values of $s_i$ need to be below their respective minimum variable values, and could be a full bin lower, to allow a buffer corresponding to a full hexagon width ($a_1$) and height ($a_2$) around the data.

### 2024-04-30 paper.tex
Corrected vs formular
Changed vs notation
notation update

@@
 $\mathbfit{\theta}$ & (Hyper-) parameters for NLDR method\\
-$y_\text{max}$ & maximum value of the scaled second embedding component\\
 $r$ & ranges of the embedding components\\
-$b$ & bins\\
+$C^{(j)}$ & $j$-dimensional bin centers\\
+$b$ & number of bins in each direction\\
 \addlinespace
-$a_r$ & aspect ratio of the 2D layout before scaling\\
-$s$ & hexagonal size (defined as the radius of the outer circle surrounding the hexagon)\\
+$a$ & binwidths, distance between centroids in each direction\\
 $(s_1, \ s_2)$ & starting coordinates of the hexagonal grid\\
-$q_1$, $q_2$ & buffer amount along the x and y axes\\
-$h_s$, $v_s$ & horizontal and vertical distance between adjacent hexagon bin centroids\\
-\addlinespace
-$d_x$, $d_y$ & horizontal spacing and adjusted vertical spacing for hexagonal coordinates\\
-$b_1$, $b_2$ & number of hexagon bins along the x and y axes\\
+$q$ & buffer to ensure hexgrid covers data, proportion of data range, 0-1\\
 $b$, $b'$ & total and non-empty hexagon bins in the grid\\
 $n_k$ & number of observations within the $k^{th}$ hexagon\\
-$(h^k_{x_i}, h^k_{y_i})$ & hexagonal grid coordinates of the $k^{th}$ hexagon\\
-\addlinespace
-$C^{(j)}$ & $j$-dimensional bin centers\\
 \bottomrule
@@
 The 2D hexagon grid is defined by the number of bins in each direction
-(\(b_1, b_2\)), and the lower left position where the grid starts
-(\(s_1, s_2\)). The values of \(s_i\) need to be below their respective
-minimum variable values, and could be a full bin lower, to allow a
-buffer corresponding to a full hexagon width (\(a_1\)) and height
-(\(a_2\)) around the data. The values of \(b_i\) are variables to be
-computed that define the reduction in size of the data (\(n\) to \(m\)).
+\((b_1, b_2)\), as given by the centroids of each hexagon \((c_1, c_2)\)
+and the lower left position where the grid starts at \((s_1, s_2)\),
+which correspond to the lowest left centroid. The values of \(s_i\) need
+to be below their respective minimum variable values, and could be a
+full bin lower, to allow a buffer corresponding to a full hexagon width
+(\(a_1\)) and height (\(a_2\)) around the data. The values of \(b_i\)
+are variables to be computed that define the reduction in size of the
+data (\(n\) to \(m\)).
 
@@
 minimum \(a_1\). According to the upper bound of the NLDR,
-\(b_2 = \left\lceil\frac{r_2 - s_2}{vs} + 1\right\rceil\), where
-\(vs = \frac{a_2}{2} + \sqrt{{(\frac{a_2}{2})}^2 + {(\frac{a_1}{2})}^2}\).
+\(b_2 = \left\lceil\frac{r_2 - s_2}{v_2} + 1\right\rceil\), where
+\(v_2 = \frac{a_2}{2} + \sqrt{{(\frac{a_2}{2})}^2 - {(\frac{a_1}{2})}^2}\).
 

### 2024-04-30 paper.tex
notation update

@@
 $C^{(j)}$ & $j$-dimensional bin centers\\
-$b$ & number of bins in each direction\\
+$(b_1, b_2)$ & number of bins in each direction\\
 \addlinespace
-$a$ & binwidths, distance between centroids in each direction\\
+$(a_1, a_2)$ & binwidths, distance between centroids in each direction\\
 $(s_1, \ s_2)$ & starting coordinates of the hexagonal grid\\

### 2024-05-08 paper.qmd
Update with quollr new updates

@@
 
-umap_scurve_scaled <- gen_scaled_data(
-  data = umap_scurve, 
-  x = "UMAP1", 
-  y = "UMAP2", 
-  hr = 2/sqrt(3))
+scurve_scaled_obj <- gen_scaled_data(
+  data = umap_scurve)
+
+umap_scurve_scaled <- scurve_scaled_obj$scaled_nldr
+lim1 <- scurve_scaled_obj$lim1
+lim2 <- scurve_scaled_obj$lim2
+r2 <- diff(lim2)/diff(lim1)
 
@@
 sc_xlims <- c(-0.25, 1.25)
-sc_ylims <- c(-0.25, 2.59)
+sc_ylims <- c(-0.25, 1.9)
 # sc_xlims <- c(-0.25, 1.25)
@@
 num_bins_x_scurve <- 6
-num_bins_y_scurve <- 14
-hex_size_scurve <- 0.13
-buffer_x_scurve <- 0.338
-buffer_y_scurve <- 0.292
-x_start_scurve <- -0.113
-y_start_scurve <- -0.098
 
@@
   data = umap_scurve_scaled, 
-  x = "UMAP1",
-  y = "UMAP2", 
-  num_bins_x = num_bins_x_scurve,
-  num_bins_y = num_bins_y_scurve, 
-  x_start = x_start_scurve,
-  y_start = y_start_scurve, 
-  buffer_x = buffer_x_scurve,
-  buffer_y = buffer_y_scurve, 
-  hex_size = hex_size_scurve,
-  col_start = "UMAP")
+  bin1 = num_bins_x_scurve, 
+  s1 = -0.06, 
+  s2 = -0.06, 
+  r2 = r2)
 

### 2024-05-11 paper.qmd
Update
Update with new package

@@
 sc_xlims <- c(-0.25, 1.25)
-sc_ylims <- c(-0.25, 1.9)
+sc_ylims <- c(-0.4, 1.9)
 # sc_xlims <- c(-0.25, 1.25)
@@
   bin1 = num_bins_x_scurve, 
-  s1 = -0.06, 
-  s2 = -0.06, 
+  q = 0.1,
   r2 = r2)

### 2024-05-12 paper.qmd
Initial images for digit1 analysis

@@
 
+<!-- PaCMAP layout with author's suggested parameter choice-->
+```{r}
+#| warning: false
+#| echo: false
+
+## Import data
+training_data_mnist <- read_rds("data/mnist/mnist_10_pcs_of_digit_1.rds")
+training_data_mnist <- training_data_mnist |>
+  mutate(ID = 1:NROW(training_data_mnist))
+
+pacmap_minst <- read_rds("data/mnist/mnist_pacmap.rds")
+pacmap_minst_scaled <- gen_scaled_data(data = pacmap_minst)$scaled_nldr
+
+position_df <- pacmap_minst_scaled |> 
+  filter(ID %in% c(6475, 3311, 6489, 2347, 387, 
+                          5239, 728, 517, 6369, 3055, 6345, 1761)) |>
+  mutate(position = c(5, 9, 7, 8, 4, 11, 2, 6, 12, 10, 1, 3))
+
+pacmap_plot_mnist <- pacmap_minst_scaled |> 
+  ggplot(aes(x = PaCMAP1, 
+             y = PaCMAP2)) + 
+  geom_point(alpha=0.1) + 
+  geom_text(data = position_df, 
+            aes(x = PaCMAP1, y = PaCMAP2, label = position), 
+            colour = "red", 
+            size = 5)  
+```
+
+<!--Render the images in different selected location in 2D layout-->
+```{r}
+#| warning: false
+#| echo: false
+
+## Data with pixel values
+mnist_data <- read_rds("data/mnist/mnist_digit_1.rds")
+
+pixels_gathered <-  mnist_data |>
+  mutate(instance = row_number()) |>
+  gather(pixel, value, -Label, -instance) |>
+  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
+  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
+  filter(instance %in% c(3710, 2391, 2385, 5030, 6475,
+                         7679, 3568, 1312, 3311, 1097,
+                         3552, 7853, 6489, 7689, 6690,
+                        1380, 6057, 2347, 5946, 3355,
+                        4175, 3997, 5378, 387, 1854,
+                        614, 3079, 1762, 5239, 3723,
+                        5748, 728, 7419, 7794, 6233,
+                        33, 2485, 5998, 318, 1761,
+                        5690, 165, 517, 6935, 2682,
+                        4962, 2264, 5563, 6369, 559,
+                        5188, 4849, 2666, 3448, 3055,
+                        3120, 6869, 6345, 4470, 7147)) |>
+  mutate(instance.labs =  case_when(
+    instance %in% c(3710, 2391, 2385, 5030, 6475) ~ "1",
+    instance %in% c(7679, 3568, 1312, 3311, 1097) ~ "2",
+    instance %in% c(3552, 7853, 6489, 7689, 6690) ~ "3",
+    instance %in% c(1380, 6057, 2347, 5946, 3355) ~ "4",
+    instance %in% c(4175, 3997, 5378, 387, 1854) ~ "5",
+    instance %in% c(614, 3079, 1762, 5239, 372) ~ "6",
+    instance %in% c(5748, 728, 7419, 7794, 6233) ~ "7",
+    instance %in% c(5690, 165, 517, 6935, 2682) ~ "8",
+    instance %in% c(4962, 2264, 5563, 6369, 559) ~ "9",
+    instance %in% c(5188, 4849, 2666, 3448, 3055) ~ "10",
+    instance %in% c(3120, 6869, 6345, 4470, 7147) ~ "11",
+    .default = "12"
+    ))
+
+## To generate labels
+instance.labs <- as.character(rep(1:12, each = 5))
+
+img_sample <- pixels_gathered |> 
+  ggplot(aes(x, y, fill = value)) +  
+  geom_tile() +
+  facet_wrap(~ factor(instance, levels = c(3710, 2391, 2385, 5030, 6475,
+                         7679, 3568, 1312, 3311, 1097,
+                         3552, 7853, 6489, 7689, 6690,
+                        1380, 6057, 2347, 5946, 3355,
+                        4175, 3997, 5378, 387, 1854,
+                        614, 3079, 1762, 5239, 3723,
+                        5748, 728, 7419, 7794, 6233,
+                        33, 2485, 5998, 318, 1761,
+                        5690, 165, 517, 6935, 2682,
+                        4962, 2264, 5563, 6369, 559,
+                        5188, 4849, 2666, 3448, 3055,
+                        3120, 6869, 6345, 4470, 7147)), ncol = 5) +
+  scale_fill_continuous_sequential(palette = "Grays") +
+  theme(strip.background = element_blank(),
+        strip.text.x = element_blank())
+```
+
+<!--PaCMAP param: n_components=2, n_neighbors=10, init=random, MN_ratio=0.9, FP_ratio=2.0-->
+```{r}
+#| echo: false
+#| fig-cap: "(a) 2D layout from PaCMAP applied for the digit 1 of the MNIST dataset.
+Is this a best representation of the original data?, and (b) Images of the handwritten digit 1.
+The angle of the digit 1 varies along non-linear structure."
+#| label: fig-pacmapauthor
+#| fig-pos: H
+
+pacmap_plot_mnist + img_sample +
+  plot_annotation(tag_levels = 'a')
+```
+
 ## Discussion {#sec-discussion}

### 2024-05-12 paper.tex
Initial images for digit1 analysis

@@
 
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-pacmapauthor-1.pdf}
+
+}
+
+\caption{\label{fig-pacmapauthor}(a) 2D layout from PaCMAP applied for
+the digit 1 of the MNIST dataset. Is this a best representation of the
+original data?, and (b) Images of the handwritten digit 1. The angle of
+the digit 1 varies along non-linear structure.}
+
+\end{figure}%
+
 \section{Discussion}\label{sec-discussion}

### 2024-05-12 paper.qmd
Add descriptions according to b2 computation

@@
 The value for $b_2$ is computed by fixing $b_1$.
-Considering the lower bound of the NLDR, $a_1 > 2 \times s_1$ and $a_1 > \frac{1-s_1}{b_1 -1}$.
-The minimum $a_1$ that satisfies the constraints will be selected.
-For regular hexagons, $a_2 = \frac{2}{\sqrt(3)}a_1$.
-Therefore, $a_2$ is computed for the minimum $a_1$.
-According to the upper bound of the NLDR, $b_2 = \left\lceil\frac{r_2 - s_2}{v_2} + 1\right\rceil$, where $v_2 = \frac{a_2}{2} + \sqrt{{(\frac{a_2}{2})}^2 - {(\frac{a_1}{2})}^2}$.
+Considering the lower bound of the NLDR, $a_1 > 2 \times s_1$, and $a_1 > \frac{1-s_1}{b_1 -1}$.
+Similarly, according to the upper bound of the NLDR, $a_1 > \frac{2(r_2 - s_2)}{\sqrt{3}(b_2 - 1)}$, because $a_2 = \frac{\sqrt(3)}{2}a_1$ for regular hexagons.
+The $a_1$ that satisfies the constraints will be selected.
+Therefore, $b_2 = \Big\lceil1 +\frac{2(r_2 - s_2)(b_1 - 1)}{\sqrt{3}(1-s_1)}\Big\rceil$.
 
@@
 ```{r}
-#| warning: false
-#| echo: false
-
 ## Import data
@@
 ```{r}
-#| warning: false
-#| echo: false
-
 ## Data with pixel values
@@
 ```{r}
-#| echo: false
 #| fig-cap: "(a) 2D layout from PaCMAP applied for the digit 1 of the MNIST dataset.
@@
 pacmap_plot_mnist + img_sample +
+  plot_annotation(tag_levels = 'a')
+```
+
+<!-- Fit the model and compute error-->
+```{r}
+
+```
+
+<!--images that occur large error-->
+<!-- within the nonlinear structure-->
+```{r}
+pixels_gathered <-  mnist_data |>
+  mutate(instance = row_number()) |>
+  gather(pixel, value, -Label, -instance) |>
+  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
+  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
+  filter(instance %in% c( 4699, 4785, 5173, 6671, 2186, 4636, 4882, 
+                          773, 2825, 6169, 6948, 5987, 1282, 7715, 
+                          622, 2989, 6912, 4370, 1227, 4042, 2444, 
+                          836, 1654))
+
+imge_error_sample_within <- pixels_gathered |> 
+  ggplot(aes(x, y, fill = value)) +  
+  geom_tile() +  
+  facet_wrap(~ factor(instance, levels = c( 4699, 4785, 5173, 6671, 2186, 4636, 
+                                            4882, 773, 2825, 6169, 6948, 5987, 
+                                            1282, 7715, 622, 2989, 6912, 4370, 
+                                            1227, 4042, 2444, 836, 1654)), 
+             ncol = 12) +
+  scale_fill_continuous_sequential(palette = "Grays") +
+  theme(strip.background = element_blank(),
+        strip.text.x = element_blank())
+```
+
+<!--outside the nonlinear structure-->
+```{r}
+#| eval: false
+
+img_error <- error_df |> 
+  filter(total >= 8) |> 
+  filter((!ID %in% c(4699, 4785, 5173, 6671, 2186, 4636, 4882, 773, 
+                            2825, 6169, 6948, 5987, 1282, 7715, 622, 2989, 
+                            6912, 4370, 1227, 4042, 2444, 836, 1654))) |>
+  pull(ID)
+
+pixels_gathered <-  mnist_data |>
+  mutate(instance = row_number()) |>
+  gather(pixel, value, -Label, -instance) |>
+  extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
+  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
+  filter(instance %in% img_error)
+
+imge_error_sample <- pixels_gathered |> 
+  ggplot(aes(x, y, fill = value)) +  
+  geom_tile() +  
+  facet_wrap(~ instance, ncol = 11) +
+  scale_fill_continuous_sequential(palette = "Grays") +
+  theme(strip.background = element_blank(),
+        strip.text.x = element_blank())
+```
+
+```{r}
+#| eval: false
+##| echo: false
+#| fig-cap: "(a) Images of handwritten digit 1 which occur large model error within the nonlinear strcuture, and (b) Images of handwritten digit 1 which occur large model error outside the nonlinear structure."
+#| label: fig-anomalies
+#| out-width: 100%
+#| fig-pos: H
+
+imge_error_sample_within + imge_error_sample +
   plot_annotation(tag_levels = 'a')

### 2024-05-12 paper.tex
Add descriptions according to b2 computation

@@
 The value for \(b_2\) is computed by fixing \(b_1\). Considering the
-lower bound of the NLDR, \(a_1 > 2 \times s_1\) and
-\(a_1 > \frac{1-s_1}{b_1 -1}\). The minimum \(a_1\) that satisfies the
-constraints will be selected. For regular hexagons,
-\(a_2 = \frac{2}{\sqrt(3)}a_1\). Therefore, \(a_2\) is computed for the
-minimum \(a_1\). According to the upper bound of the NLDR,
-\(b_2 = \left\lceil\frac{r_2 - s_2}{v_2} + 1\right\rceil\), where
-\(v_2 = \frac{a_2}{2} + \sqrt{{(\frac{a_2}{2})}^2 - {(\frac{a_1}{2})}^2}\).
+lower bound of the NLDR, \(a_1 > 2 \times s_1\), and
+\(a_1 > \frac{1-s_1}{b_1 -1}\). Similarly, according to the upper bound
+of the NLDR, \(a_1 > \frac{2(r_2 - s_2)}{\sqrt{3}(b_2 - 1)}\), because
+\(a_2 = \frac{\sqrt(3)}{2}a_1\) for regular hexagons. The \(a_1\) that
+satisfies the constraints will be selected. Therefore,
+\(b_2 = \Big\lceil1 +\frac{2(r_2 - s_2)(b_1 - 1)}{\sqrt{3}(1-s_1)}\Big\rceil\).
 

### 2024-05-12 paper.qmd
Add description for binning data

@@
 
-Points are allocated to the bin that they fall in by considering the nearest centroid.
+Points are allocated to the bin they fall into based on the nearest centroid.
+In situations where a point is equidistant from multiple centroids, tie-breaking rules are applied.
+If multiple centroids are in the same row, the point is assigned to the leftmost centroid.
+If multiple centroids are in different rows, the point is assigned to the bottom centroid.
 

### 2024-05-12 paper.tex
Add description for binning data

@@
 
-Points are allocated to the bin that they fall in by considering the
-nearest centroid.
+Points are allocated to the bin they fall into based on the nearest
+centroid. In situations where a point is equidistant from multiple
+centroids, tie-breaking rules are applied. If multiple centroids are in
+the same row, the point is assigned to the leftmost centroid. If
+multiple centroids are in different rows, the point is assigned to the
+bottom centroid.
 

### 2024-05-12 paper.qmd
Added hexbin parameter img

@@
 The 2D hexagon grid is defined by the number of bins in each direction $(b_1, b_2)$, as given by the centroids of each hexagon $(c_1, c_2)$ and the lower left position where the grid starts at $(s_1, s_2)$, which correspond to the lowest left centroid.
-The values of $s_i$ need to be below their respective minimum variable values, and could be a full bin lower, to allow a buffer corresponding to a full hexagon width ($a_1$) and height ($a_2$) around the data.
+The values of $s_i$ need to be below their respective minimum variable values, and could be a full bin lower, to allow a buffer ($q$) corresponding to a full hexagon width ($a_1$) and height ($a_2$) around the data.
 The values of $b_i$ are variables to be computed that define the reduction in size of the data ($n$ to $m$).
@@
 Therefore, $b_2 = \Big\lceil1 +\frac{2(r_2 - s_2)(b_1 - 1)}{\sqrt{3}(1-s_1)}\Big\rceil$.
+
+```{r}
+hex_param_vis <- ggplot(data = hex_grid, aes(x = x, y = y)) + 
+    geom_polygon(fill = "white", color = "#bdbdbd", aes(group = hex_poly_id)) +
+    geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "#31a354") +
+    geom_point(data = all_centroids_df |> dplyr::filter(hexID == 1), aes(x = c_x, y = c_y), color = "black") + ## starting point
+    geom_rect(aes(xmin = 0, xmax = 1, ymin = 0, ymax = r2), fill = "white", color = "black", alpha = 0, linewidth = 1) +
+    geom_segment(x = 0, y = -0.35, xend = 1, yend = -0.35,
+                 arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+
+    annotate("text", x=0.5, y=-0.45, label= "r1", color = "black") +
+    geom_segment(x = -0.25, y = 0, xend = -0.25, yend = r2,
+                 arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ 
+    annotate("text", x=-0.35, y=1, label= "r2", color = "black") +
+    geom_segment(x = -0.1 + 0.2087578, y = -0.15, xend = -0.1 + 0.2087578*2, yend = -0.15,
+                   arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ # a1 = 0.2087578
+    annotate("text", x=0.2, y=-0.25, label= "a1", color = "black") +
+    geom_segment(x = -0.15, y = -0.1*r2 + 0.1807896*2, xend = -0.15, yend = -0.1*r2 + 0.1807896*3,
+                     arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ # a2 = 0.1807896
+    annotate("text", x=-0.15, y=0.45, label= "a2", color = "black") +
+    annotate("text", x=-0.15, y=-0.35, label= "(s1, s2)", color = "black") ## starting point
+```
+
+```{r}
+#| label: fig-hex-param
+#| fig-cap: "Hexagon binning parameters."
+#| fig-width: 3
+#| fig-height: 3
+#| fig-pos: H
+
+hex_param_vis
+```
 

### 2024-05-12 paper.tex
Added hexbin parameter img

@@
 to be below their respective minimum variable values, and could be a
-full bin lower, to allow a buffer corresponding to a full hexagon width
-(\(a_1\)) and height (\(a_2\)) around the data. The values of \(b_i\)
-are variables to be computed that define the reduction in size of the
-data (\(n\) to \(m\)).
+full bin lower, to allow a buffer (\(q\)) corresponding to a full
+hexagon width (\(a_1\)) and height (\(a_2\)) around the data. The values
+of \(b_i\) are variables to be computed that define the reduction in
+size of the data (\(n\) to \(m\)).
 
@@
 \(b_2 = \Big\lceil1 +\frac{2(r_2 - s_2)(b_1 - 1)}{\sqrt{3}(1-s_1)}\Big\rceil\).
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-hex-param-1.pdf}
+
+}
+
+\caption{\label{fig-hex-param}Hexagon binning parameters.}
+
+\end{figure}%
 

### 2024-05-13 paper.qmd
Add best model but getting an error because r2 = 0.59 something working better when r2 > 1

@@
 Similarly, according to the upper bound of the NLDR, $a_1 > \frac{2(r_2 - s_2)}{\sqrt{3}(b_2 - 1)}$, because $a_2 = \frac{\sqrt(3)}{2}a_1$ for regular hexagons.
-The $a_1$ that satisfies the constraints will be selected.
 Therefore, $b_2 = \Big\lceil1 +\frac{2(r_2 - s_2)(b_1 - 1)}{\sqrt{3}(1-s_1)}\Big\rceil$.
@@
 
-pacmap_minst <- read_rds("data/mnist/mnist_pacmap.rds")
-pacmap_minst_scaled <- gen_scaled_data(data = pacmap_minst)$scaled_nldr
+pacmap_minst <- read_rds("data/mnist/mnist_pacmap.rds") |> 
+  select(PaCMAP1, PaCMAP2, ID)
+mnist_scaled_obj <- gen_scaled_data(
+  data = pacmap_minst)
+pacmap_minst_scaled <- mnist_scaled_obj$scaled_nldr
 
@@
 ```{r}
+## Compute hexbin parameters
+num_bins_x_mnist <- 20
+lim1 <- mnist_scaled_obj$lim1
+lim2 <- mnist_scaled_obj$lim2
+r2_mnist <- diff(lim1)/diff(lim2) # Need to correct r2_mnist <- diff(lim2)/diff(lim1)
 
+mnist_model <- fit_highd_model(
+  training_data = training_data_mnist,
+  emb_df = pacmap_minst_scaled,
+  bin1 = num_bins_x_mnist,
+  q = 0.1,
+  r2 = r2_mnist,
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = FALSE,
+  col_start_highd = "PC"
+)
+
+df_bin_centroids_mnist <- mnist_model$df_bin_centroids
+df_bin_mnist <- mnist_model$df_bin
+
+## Triangulate bin centroids
+tr1_object_mnist <- tri_bin_centroids(
+  df_bin_centroids_mnist, x = "c_x", y = "c_y")
+tr_from_to_df_mnist <- gen_edges(
+  tri_object = tr1_object_mnist)
+
+## Compute 2D distances
+distance_mnist <- cal_2d_dist(
+  tr_coord_df = tr_from_to_df_mnist, 
+  start_x = "x_from", 
+  start_y = "y_from", 
+  end_x = "x_to", 
+  end_y = "y_to", 
+  select_vars = c("from", "to", "distance"))
+
+## To find the benchmark value
+benchmark_mnist <- find_lg_benchmark(
+  distance_edges = distance_mnist, 
+  distance_col = "distance")
+
+trimesh_removed_mnist <- vis_rmlg_mesh(
+  distance_edges = distance_mnist, 
+  benchmark_value = benchmark_mnist, 
+  tr_coord_df = tr_from_to_df_mnist, 
+  distance_col = "distance") +
+  #xlim(sc_xlims) + ylim(sc_ylims) +
+  interior_annotation("a", sc_ltr_pos)
+
+## Compute error
+error_df <- augment(
+  df_bin_centroids = df_bin_centroids_mnist, 
+  df_bin = df_bin_mnist, 
+  training_data = training_data_mnist, 
+  newdata = NULL, 
+  type_NLDR = "PaCMAP", 
+  col_start = "PC")
+
+## Categorize error
+
+error_df <- error_df |>
+  mutate(type = case_when(
+    row_wise_abs_error <= 2 ~ "error 0-2",
+    row_wise_abs_error <= 4 ~ "error 2-4",
+    row_wise_abs_error <= 6 ~ "error 4-6",
+    row_wise_abs_error <= 6 ~ "error 6-8",
+    row_wise_abs_error <= 10 ~ "error 8-10",
+    .default = "error greter than 10"
+  )) |>
+  mutate(type = factor(type, levels = c(
+    "error 0-2", "error 2-4", "error 4-6", "error 6-8", "error 8-10", 
+    "error greter than 10")))
+
+## To join embedding
+error_df <- error_df |>
+  bind_cols(pacmap_minst |> 
+              select(-ID))
+
+error_plot_mnist <- error_df |>
+  ggplot(aes(x = PaCMAP1,
+             y = PaCMAP2, 
+             color = type, 
+             group = ID)) +
+  geom_point(alpha=0.5, 
+             size = 0.1) +
+  interior_annotation("b", sc_ltr_pos)
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "(a) Model generated in the 2D space overlaid on PaCMAP data, and (b) high-D model error in model space."
+#| label: fig-model-mnist
+#| fig-pos: H
+
+trimesh_removed_mnist + error_plot_mnist + 
+  plot_layout(guides='collect', ncol=2) &
+  theme(legend.position='bottom')
 ```
@@
                                             1227, 4042, 2444, 836, 1654)), 
-             ncol = 12) +
+             ncol = 5) +
   scale_fill_continuous_sequential(palette = "Grays") +
@@
 ```{r}
-#| eval: false
-
 img_error <- error_df |> 
-  filter(total >= 8) |> 
+  filter(row_wise_abs_error >= 8) |> 
   filter((!ID %in% c(4699, 4785, 5173, 6671, 2186, 4636, 4882, 773, 
@@
   geom_tile() +  
-  facet_wrap(~ instance, ncol = 11) +
+  facet_wrap(~ instance, ncol = 5) +
   scale_fill_continuous_sequential(palette = "Grays") +
@@
 ```{r}
-#| eval: false
-##| echo: false
-#| fig-cap: "(a) Images of handwritten digit 1 which occur large model error within the nonlinear strcuture, and (b) Images of handwritten digit 1 which occur large model error outside the nonlinear structure."
+#| echo: false
+#| fig-cap: "(a) Images of handwritten digit 1 which occur large model error within the non-linear strcuture, and (b) Images of handwritten digit 1 which occur large model error outside the non-linear structure."
 #| label: fig-anomalies
+#| fig-width: 4
+#| fig-height: 4
 #| out-width: 100%
@@
 imge_error_sample_within + imge_error_sample +
+  plot_layout(ncol=2) +
   plot_annotation(tag_levels = 'a')

### 2024-05-13 paper.tex
Add best model but getting an error because r2 = 0.59 something working better when r2 > 1

@@
 of the NLDR, \(a_1 > \frac{2(r_2 - s_2)}{\sqrt{3}(b_2 - 1)}\), because
-\(a_2 = \frac{\sqrt(3)}{2}a_1\) for regular hexagons. The \(a_1\) that
-satisfies the constraints will be selected. Therefore,
+\(a_2 = \frac{\sqrt(3)}{2}a_1\) for regular hexagons. Therefore,
 \(b_2 = \Big\lceil1 +\frac{2(r_2 - s_2)(b_1 - 1)}{\sqrt{3}(1-s_1)}\Big\rceil\).
@@
 
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-model-mnist-1.pdf}
+
+}
+
+\caption{\label{fig-model-mnist}(a) Model generated in the 2D space
+overlaid on PaCMAP data, and (b) high-D model error in model space.}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-anomalies-1.pdf}
+
+}
+
+\caption{\label{fig-anomalies}(a) Images of handwritten digit 1 which
+occur large model error within the non-linear strcuture, and (b) Images
+of handwritten digit 1 which occur large model error outside the
+non-linear structure.}
+
+\end{figure}%
+
 \section{Discussion}\label{sec-discussion}

### 2024-05-13 supplementary/hexagon_binning_algorithm.qmd
Add best model but getting an error because r2 = 0.59 something working better when r2 > 1

---
title: "Hexagonal binning algorithm"
format: pdf
editor: visual
---

```{r}
#| echo: false
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)

library(ggplot2)
library(dplyr)

theme_set(theme_linedraw() +
   theme(
     #aspect.ratio = 1,
     plot.background = element_rect(fill = 'transparent', colour = NA),
     plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
     panel.background = element_rect(fill = 'transparent', 
                                     colour = NA),
     panel.grid.major = element_blank(), 
     panel.grid.minor = element_blank(), 
     axis.title.x = element_blank(), axis.title.y = element_blank(),
     axis.text.x = element_blank(), axis.ticks.x = element_blank(),
     axis.text.y = element_blank(), axis.ticks.y = element_blank(),
     legend.background = element_rect(fill = 'transparent', 
                                      colour = NA),
     legend.key = element_rect(fill = 'transparent', 
                               colour = NA),
     legend.position = "none", 
     legend.title = element_text(size=5), 
     legend.text = element_text(size=4),
     legend.key.height = unit(0.25, 'cm'),
     legend.key.width = unit(0.25, 'cm')
   )
)
```

## Computing hexagon grid configurations

The 2D hexagon grid is defined by the number of bins in each direction $(b_1, b_2)$, as given by the centroids of each hexagon $(c_1, c_2)$ and the lower left position where the grid starts at $(s_1, s_2)$, which correspond to the lowest left centroid. The values of $s_i$ need to be below their respective minimum variable values, and could be a full bin lower, to allow a buffer corresponding to a distance between centroids in each direction ($a_1$ and $a_2$) around the data. The values of $b_i$ are variables to be computed that define the reduction in size of the data ($n$ to $m$).

The value for $b_2$ is computed by fixing $b_1$. Also, according to hexagonal ratio $a_2 = \frac{\sqrt{3}}{2}a_1$.

The first step of this process is to find the optimal number of bins along the y-axis ($b_1$) at the minimum width of the hexagon ($a_1$) that satisfies the following constraints.

$\text{min }a_1 \text{ s.t.}$

$$
s_1 - \frac{a_1}{2} < 0,
$$ {#eq-equation1}

$$
s_1 + (b_1 - 1) \times a_1 > 1,
$$ {#eq-equation2}

$$
s_2 - \frac{a_2}{2} < 0,
$$ {#eq-equation4}

$$
s_2 + (b_2 - 1) \times a_2 > r_2.
$$ {#eq-equation5}

Since $a_1$ and $a_2$ are distances,

$$
a_1, a_2 > 0.
$$
Also, $(s_1, s_2) \in (-0.1, -0.05)$ as these are multiplicative offsets in the negative direction.

@eq-equation1 can be rearranged as,

$$
a_1 > 2s_1
$$

which given $s_1 < 0$ and $a_1 > 0$ will *always* be true. The same logic follows for @eq-equation4 and substituting $a_2 = \frac{\sqrt{3}}{2}a_1$ to @eq-equation4 can be written as,

$$
a_1 > \frac{4}{\sqrt{3}}s_2.
$$
Also, substituting $a_2 = \frac{\sqrt{3}}{2}a_1$ and rearranging @eq-equation5 gives:

$$
a_1 > \frac{2(r_2 - s_2)}{\sqrt{3}(b_2 - 1)}.
$${#eq-equation6}

Similarly, @eq-equation2 becomes,

$$
a_1 > \frac{(1-s_1)}{(b_1 - 1)}.
$${#eq-equation7}

This is a linear optimisation problem. Therefore, the optimal solution must occur on a vertex. So, by setting @eq-equation6 equals to @eq-equation7 gives,

$$
\frac{2(r_2 - s_2)}{\sqrt{3}(b_2 - 1)} = \frac{(1-s_1)}{(b_1 - 1)}.
$$
After rearranging this,

$$
b_2 = 1 + \frac{2(r_2 - s_2)(b_1 - 1)}{\sqrt{3}(1-s_1)}
$$

and since $b_2$ should be an integer,

$$
b_2 = \Big\lceil1 +\frac{2(r_2 - s_2)(b_1 - 1)}{\sqrt{3}(1-s_1)}\Big\rceil.
$$ {#eq-equation8}

Furthermore, with known $b_1$ and $b_2$, by considering @eq-equation2 or @eq-equation5 as the *binding* or *active constraint*, can compute $a_1$.

If @eq-equation2 is active, then,

$$
\frac{(1-s_1)}{(b_1 - 1)} < \frac{2(r_2 - s_2)}{\sqrt{3}(b_2 - 1)}.
$$
Rearranging this gives,

$$
r_2 > \frac{\sqrt{3}(b_2 - 1)(1-s_1)}{2(b_1 - 1)} + s_2.
$$
Therefore, if this equality is true, then $a_1 = \frac{(1-s_1)}{(b_1 - 1)}$, otherwise, $a_1 = \frac{2(r_2 - s_2)}{\sqrt{3}(b_2 - 1)}$.

```{r, echo=FALSE}
# bin1: Number of bins along the x-axis
# buffer: proportion of buffer should between 0 and 1
# r2: range of embedding 2
calc_bins_y <- function(bin1 = 2, s1 = -0.1, s2 = -0.1, r2) {

  ## Need to have a check where bin2 greater than 2
  if (bin1 < 2) {
    stop("Number of bins along the x-axis at least should be 2.")
  }

  ## To compute the number of bins along the x-axis
  bin2 <- ceiling(1 + (2 * (r2-s2) * (bin1 - 1))/(sqrt(3) * (1-s1)))

  ## Validating and compute horizontal spacing
  check_factor <- (sqrt(3) * (bin2 - 1) * (1-s1))/(2 * (bin1 - 1)) + s2

  if (r2 > check_factor) {

    a1 <- (1-s1)/(bin1 - 1)

  } else {

    a1 <- (2 * (r2-s2))/(sqrt(3) * (bin2 - 1))

  }

  return(list(bin2 = bin2, a1 = a1))
}
```

```{r, echo=FALSE}
gen_centroids <- function(bin1, s1 = -0.1, s2 = -0.1, r2){

  ## To compute hexagonal configurations
  bin_obj <- calc_bins_y(bin1 = bin1, s1 = s1, s2 = s2, r2 = r2)

  # To obtain the bins along the y-axis
  bin2 <- bin_obj$bin2

  # To obtain the width of the hexagon
  a1 <- bin_obj$a1

  # Generate x-coordinate of centroids for odd rows
  c_x_vec_odd <- seq(s1, (bin1 - 1) * a1, by = a1)

  # Generate x-coordinate of centroids for even rows
  c_x_vec_even <- c_x_vec_odd + a1/2
  c_x_vec <- c(c_x_vec_odd, c_x_vec_even)

  # To compute vertical spacing
  a2 <- sqrt(3) * a1/2

  # Generate y-coordinate of centroids
  c_y_vec <- seq(s2, (bin2 - 1) * a2, by = a2)
  c_y <- rep(c_y_vec, each = bin1)

  ## Do the number of belongs y axis is even or odd and adjust the x-coordinates
  if ((bin2 %% 2) == 0) {

    c_x <- rep(c_x_vec, bin2/2)

  } else {

    if ((ceiling(bin2/2) %% 2) == 0) {

      c_x <- append(rep(c_x_vec, floor(bin2/2)), c_x_vec_odd)

    } else{

      c_x <- append(rep(c_x_vec, floor(bin2/2)), c_x_vec_even)

    }

  }

  centroid_df <- tibble::tibble(hexID = 1:length(c_x), c_x = c_x, c_y = c_y)

  return(centroid_df)

}
```

```{r, echo=FALSE}
all_centroids_df <- gen_centroids(bin1 = 4, s1 = -0.1, s2 = -0.1, r2 = 2.019414)
```

```{r, echo=FALSE}
gen_hex_coord <- function(centroids_df, a1){

  ## Obtain centroid info
  hex_ids <- centroids_df$hexID
  c_x_vec <- centroids_df$c_x
  c_y_vec <- centroids_df$c_y

  ## To compute vertical spacing factor
  vs_factor <- a1/(2 * sqrt(3))

  dx <- a1/2
  dy <- a1/sqrt(3)

  ## Assign coordinates for 6 directions
  x_add_factor <- c(0, -dx, -dx, 0, dx, dx)
  y_add_factor <- c(dy, vs_factor, -vs_factor, -dy, -vs_factor, vs_factor)

  ## Initialize vectors to store hexagonal coordinates
  hex_poly_id <- integer(0)
  x <- numeric(0)
  y <- numeric(0)

  for (hb_id in hex_ids) {

    ## Since each hexagon has 6 coordinates
    hexID_rep <- rep(hex_ids[hb_id], each = 6)
    c_x_rep <- rep(c_x_vec[hb_id], each = 6)
    c_y_rep <- rep(c_y_vec[hb_id], each = 6)

    ## Generate the 6 coordinates
    x_spec <- c_x_rep + x_add_factor
    y_spec <- c_y_rep + y_add_factor

    ## Append to existing vectors
    x <- append(x, x_spec)
    y <- append(y, y_spec)
    hex_poly_id <- append(hex_poly_id, hexID_rep)

  }

  hex_coord_df <- tibble::tibble(hex_poly_id = hex_poly_id, x = x, y = y)

  return(hex_coord_df)
}
```

```{r, echo=FALSE}
hex_grid <- gen_hex_coord(all_centroids_df, a1 = 0.3496126)
```

```{r, echo=FALSE}
ggplot(data = hex_grid, aes(x = x, y = y)) + 
    geom_polygon(fill = "white", color = "#bdbdbd", aes(group = hex_poly_id)) +
    geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "#31a354") +
    geom_point(data = all_centroids_df |> dplyr::filter(hexID == 1), aes(x = c_x, y = c_y), color = "#08519c") + ## starting point
    geom_rect(aes(xmin = 0, xmax = 1, ymin = 0, ymax = 2.019414), fill = "white", color = "#feb24c", alpha = 0, linewidth = 1) +
    geom_segment(x = 0, y = -0.25, xend = 1, yend = -0.25,
                 arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "#f03b20")+
    annotate("text", x=0.5, y=-0.35, label= "r1", color = "#f03b20") +
    geom_segment(x = -0.25, y = 0, xend = -0.25, yend = 2.019414,
                 arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "#f03b20")+ 
    annotate("text", x=-0.35, y=1, label= "r2", color = "#f03b20") +
    geom_segment(x = -0.1, y = -0.15, xend = -0.1 + 0.3496126, yend = -0.15,
                   arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ # a1 = 0.3496126
    annotate("text", x=0.17, y=-0.35, label= "a1", color = "black") +
    geom_segment(x = -0.15, y = -0.1, xend = -0.15, yend = -0.1 + 0.3027734,
                     arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ # a2 = 0.3027734
    annotate("text", x=-0.35, y=0.15, label= "a2", color = "black") +
    annotate("text", x=-0.15, y=-0.35, label= "(s1, s2)", color = "#08519c") + ## starting point
    coord_fixed()
```

## Binning the data

Points are allocated to the bin that they fall in by considering the nearest centroid.

To find the nearest centroid to a $k^{th}$ embedding point $(y_1, y_2)$ among centroids $(C_1^{(k)}, C_2^{(k)}, \ldots, C_b^{(k)})$,

$$
\text{Minimize: }\sqrt{(y_1 - c_{1k})^2 + (y_2 - c_{2k})^2} + \text{centroid\_ID} \times 10^{-3}. 
$$

In this expression, $(c_{1k}, c_{2k})$ represents the coordinates of centroid $C_1^{(k)}$, and $\text{centroid\_ID}$ represents the ID of the centroid. This formulation ensures that if multiple centroids have the same minimum distance, the one with the minimum ID will be selected.

```{r, echo=FALSE}
emb_df <- tibble::tibble(emb1 = 0.774, emb2 = 0.5055) |>
  add_row(emb1 = 0.424, emb2 = 0.5055) |>
  add_row(emb1 = 0.5115, emb2 = 0.3545) |>
  add_row(emb1 = 0.5115, emb2 = 0.6565) |>
  add_row(emb1 = 0.6865, emb2 = 0.3545) |>
  add_row(emb1 = 0.6865, emb2 = 0.6565) |>
  add_row(emb1 = 0.599, emb2 = 0.304) |>
  add_row(emb1 = 0.599, emb2 = 0.707) |>
  add_row(emb1 = 0.424, emb2 = 0.405) |>
  add_row(emb1 = 0.424, emb2 = 0.606) |>
  add_row(emb1 = 0.774, emb2 = 0.405) |>
  add_row(emb1 = 0.774, emb2 = 0.606)

ggplot(data = hex_grid, aes(x = x, y = y)) + 
    geom_polygon(fill = "white", color = "#bdbdbd", aes(group = hex_poly_id)) +
    geom_point(data = emb_df, aes(x = emb1, y = emb2), color = "#253494") +
    geom_segment(x = 0.774, y = 0.5055, xend = 0.7, yend = 0.5055,
                       arrow = arrow(length = unit(0.03, "npc")), color = "black") +
    geom_segment(x = 0.424, y = 0.5055, xend = 0.35, yend = 0.5055,
                       arrow = arrow(length = unit(0.03, "npc")), color = "black") +
    geom_segment(x = 0.599, y = 0.707, xend = 0.599, yend = 0.6,
                       arrow = arrow(length = unit(0.03, "npc")), color = "black") +
    geom_segment(x = 0.774, y = 0.405, xend = 0.774, yend = 0.3,
                       arrow = arrow(length = unit(0.03, "npc")), color = "black") +
    geom_segment(x = 0.774, y = 0.606, xend = 0.7, yend = 0.55,
                       arrow = arrow(length = unit(0.03, "npc")), color = "black") +
    geom_segment(x = 0.6865, y = 0.6565, xend = 0.62, yend = 0.6,
                       arrow = arrow(length = unit(0.03, "npc")), color = "black") +
    geom_segment(x = 0.6865, y = 0.3545, xend = 0.6865, yend = 0.25,
                       arrow = arrow(length = unit(0.03, "npc")), color = "black")

```

### 2024-05-13 paper.qmd
Fix the issue and mnist digit1 best model added

@@
 
+<!-- UMAP layout with author's suggested parameter choice-->
+```{r}
+## Import data
+training_data_pbmc <- read_rds("data/pbmc3k/pbmc_pca_50.rds")
+training_data_pbmc <- training_data_pbmc[, 1:9] |>
+  mutate(ID = 1:NROW(training_data_pbmc))
+
+umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_30_min_dist_0.3.rds")
+pbmc_scaled_obj <- gen_scaled_data(
+  data = umap_pbmc)
+umap_pbmc_scaled <- pbmc_scaled_obj$scaled_nldr
+
+umap_pbmc <- umap_pbmc_scaled |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2)) +
+  geom_point(alpha=0.5) 
+```
+
+<!-- UMAP parameters suggested by the author(n_neighbors = 30, min_dist = 0.3)-->
+```{r}
+#| fig-cap: "2D layout from UMAP applied for the PBMC3k dataset. Is this a best representation of the original data?"
+#| label: fig-umapauthor
+#| fig-width: 3
+#| fig-height: 3
+#| fig-pos: H
+
+umap_pbmc 
+```
+
 ### digits: 1
@@
 ## Compute hexbin parameters
-num_bins_x_mnist <- 20
+num_bins_x_mnist <- 37
 lim1 <- mnist_scaled_obj$lim1
 lim2 <- mnist_scaled_obj$lim2
-r2_mnist <- diff(lim1)/diff(lim2) # Need to correct r2_mnist <- diff(lim2)/diff(lim1)
+r2_mnist <- diff(lim2)/diff(lim1)
 
@@
 error_df <- error_df |>
-  bind_cols(pacmap_minst |> 
+  bind_cols(pacmap_minst_scaled |> 
               select(-ID))
@@
 
+<!--bin1 = 37, bin2 = 26, b = 962, non_empty = 398-->
 ```{r}
@@
   geom_tile() +  
-  facet_wrap(~ instance, ncol = 5) +
+  facet_wrap(~ instance, ncol = 2) +
   scale_fill_continuous_sequential(palette = "Grays") +

### 2024-05-13 paper.qmd
Added author best model

@@
 
+<!--Fit the best model for author suggestion and compute error-->
+```{r}
+## Compute hexbin parameters
+num_bins_x_pbmc <- 30
+lim1 <- pbmc_scaled_obj$lim1
+lim2 <- pbmc_scaled_obj$lim2
+r2_pbmc <- diff(lim2)/diff(lim1)
+
+pbmc_model <- fit_highd_model(
+  training_data = training_data_pbmc,
+  emb_df = umap_pbmc_scaled,
+  bin1 = num_bins_x_pbmc,
+  q = 0.1,
+  r2 = r2_pbmc,
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = FALSE,
+  col_start_highd = "PC_"
+)
+
+df_bin_centroids_pbmc <- pbmc_model$df_bin_centroids
+df_bin_pbmc <- pbmc_model$df_bin
+
+## Triangulate bin centroids
+tr1_object_pbmc <- tri_bin_centroids(
+  df_bin_centroids_pbmc, x = "c_x", y = "c_y")
+tr_from_to_df_pbmc <- gen_edges(
+  tri_object = tr1_object_pbmc)
+
+## Compute 2D distances
+distance_pbmc <- cal_2d_dist(
+  tr_coord_df = tr_from_to_df_pbmc,
+  start_x = "x_from",
+  start_y = "y_from",
+  end_x = "x_to",
+  end_y = "y_to",
+  select_vars = c("from", "to", "distance"))
+
+## To find the benchmark value
+benchmark_pbmc <- find_lg_benchmark(
+  distance_edges = distance_pbmc,
+  distance_col = "distance")
+
+trimesh_removed_pbmc <- vis_rmlg_mesh(
+  distance_edges = distance_pbmc,
+  benchmark_value = benchmark_pbmc,
+  tr_coord_df = tr_from_to_df_pbmc,
+  distance_col = "distance") +
+  #xlim(sc_xlims) + ylim(sc_ylims) +
+  interior_annotation("a", sc_ltr_pos)
+
+## Compute error
+error_df <- augment(
+  df_bin_centroids = df_bin_centroids_pbmc,
+  df_bin = df_bin_pbmc,
+  training_data = training_data_pbmc,
+  newdata = NULL,
+  type_NLDR = "PaCMAP",
+  col_start = "PC_")
+
+## Categorize error
+
+error_df <- error_df |>
+  mutate(type = case_when(
+    row_wise_abs_error <= 5 ~ "error 0-5",
+    row_wise_abs_error <= 10 ~ "error 5-10",
+    row_wise_abs_error <= 15 ~ "error 10-15",
+    row_wise_abs_error <= 20 ~ "error 15-20",
+    .default = "error greter than 20"
+  )) |>
+  mutate(type = factor(type, levels = c(
+    "error 0-5", "error 5-10", "error 10-15", "error 15-20", "error greter than 20")))
+
+## To join embedding
+error_df <- error_df |>
+  bind_cols(umap_pbmc_scaled |>
+              select(-ID))
+
+error_plot_pbmc <- error_df |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2,
+             color = type,
+             group = ID)) +
+  geom_point(alpha=0.5,
+             size = 0.1) +
+  interior_annotation("b", sc_ltr_pos)
+
+```
+
+<!--bin1 = 30, bin2 = 29, b = 870, non_empty = 132-->
+```{r}
+#| echo: false
+#| fig-cap: "(a) Model generated in the 2D space overlaid on UMAP data, and (b) high-D model error in model space."
+#| label: fig-model-pbmc-author
+#| fig-pos: H
+
+trimesh_removed_pbmc + error_plot_pbmc + 
+  plot_layout(guides='collect', ncol=2) &
+  theme(legend.position='bottom')
+```
+
 ### digits: 1

### 2024-05-13 paper.tex
Fix the issue and mnist digit1 best model added
Added author best model

@@
 
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-umapauthor-1.pdf}
+
+}
+
+\caption{\label{fig-umapauthor}2D layout from UMAP applied for the
+PBMC3k dataset. Is this a best representation of the original data?}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-model-pbmc-author-1.pdf}
+
+}
+
+\caption{\label{fig-model-pbmc-author}(a) Model generated in the 2D
+space overlaid on UMAP data, and (b) high-D model error in model space.}
+
+\end{figure}%
+
 \subsection{digits: 1}\label{digits-1}

### 2024-05-13 paper.qmd
Initial flow of finding the best model for PBMC3k

@@
                                colour = NA),
-     legend.position = "none", 
-     legend.title = element_text(size=5), 
+     legend.position = "bottom", 
+     legend.title = element_blank(), 
      legend.text = element_text(size=4),
@@
   newdata = NULL,
-  type_NLDR = "PaCMAP",
+  type_NLDR = "UMAP",
   col_start = "PC_")
@@
 #| label: fig-model-pbmc-author
+#| fig-pos: H
+
+trimesh_removed_pbmc + error_plot_pbmc + 
+  plot_layout(guides='collect', ncol=2) &
+  theme(legend.position='bottom')
+```
+
+<!--compute absolute error for different parameter choices-->
+```{r}
+error_df_umap <- read_csv("data/pbmc3k/error_df_umap.csv", col_names = FALSE)
+names(error_df_umap) <- c("error", "mse", "bin1", "bin2", "b", "b_non_empty", 
+                     "method", "n_neighbors", "min_dist")
+
+error_df_umap <- error_df_umap |>
+  mutate(type = paste0("n_neighbors: ", n_neighbors, ", min_dist: ", min_dist)) |>
+  select(-c(n_neighbors, min_dist))
+
+error_df_tsne <- read_csv("data/pbmc3k/error_df_tsne.csv", col_names = FALSE)
+names(error_df_tsne) <- c("error", "mse", "bin1", "bin2", "b", "b_non_empty", 
+                          "method", "perplexity")
+
+error_df_tsne <- error_df_tsne |>
+  mutate(type = paste0("perplexity: ", perplexity)) |>
+  select(-perplexity)
+
+error_df <- bind_rows(error_df_umap, error_df_tsne)
+
+error_plot_pbmc <- ggplot(error_df, 
+                              aes(x = b_non_empty, 
+                                  y = log(error), 
+                                  group = type, 
+                                  colour = type)) + 
+  geom_point() +
+  geom_line() + 
+  geom_vline(xintercept = 137, linetype="solid", 
+             color = "black", size=0.8, alpha = 0.5) +
+  scale_color_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99',
+                              '#e31a1c','#fdbf6f','#ff7f00','#cab2d6','#6a3d9a', 
+                              '#b15928', '#ffff99')) +
+  ylab("log(absolute error)") +
+  xlab("number of non-empty bins")
+
+```
+
+```{r}
+#| fig-cap: "Absolute error from UMAP and tSNE applied to training PBMC3k dataset with diffferent parameter choices.
+What is the best parameter choice to create the model?
+The residual plot have a steep slope at the beginning, indicating that a smaller number of non-empty bins causes a larger amount of error.
+Then, the slope gradually declines or level off, indicating that a higher number of non-empty bins generates a smaller error.
+Using the elbow method, it was observed that when the number of non-empty bins is set to 137, the lowest error occurred with the parameters perplexity: 30."
+#| label: fig-abserrorPBMC
+#| fig-pos: H
+#| out-width: 100%
+
+error_plot_pbmc
+```
+
+<!--best choice-->
+
+```{r}
+tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_30.rds")
+pbmc_scaled_obj <- gen_scaled_data(
+  data = tsne_pbmc)
+tsne_pbmc_scaled <- pbmc_scaled_obj$scaled_nldr
+
+tsne_pbmc <- tsne_pbmc_scaled |>
+  ggplot(aes(x = tSNE1,
+             y = tSNE2)) +
+  geom_point(alpha=0.5)
+```
+
+<!--Fit the best model and compute error-->
+```{r}
+## Compute hexbin parameters
+num_bins_x_pbmc <- 15
+lim1 <- pbmc_scaled_obj$lim1
+lim2 <- pbmc_scaled_obj$lim2
+r2_pbmc <- diff(lim2)/diff(lim1)
+
+pbmc_model <- fit_highd_model(
+  training_data = training_data_pbmc,
+  emb_df = tsne_pbmc_scaled,
+  bin1 = num_bins_x_pbmc,
+  q = 0.1,
+  r2 = r2_pbmc,
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = FALSE,
+  col_start_highd = "PC_"
+)
+
+df_bin_centroids_pbmc <- pbmc_model$df_bin_centroids
+df_bin_pbmc <- pbmc_model$df_bin
+
+## Triangulate bin centroids
+tr1_object_pbmc <- tri_bin_centroids(
+  df_bin_centroids_pbmc, x = "c_x", y = "c_y")
+tr_from_to_df_pbmc <- gen_edges(
+  tri_object = tr1_object_pbmc)
+
+## Compute 2D distances
+distance_pbmc <- cal_2d_dist(
+  tr_coord_df = tr_from_to_df_pbmc,
+  start_x = "x_from",
+  start_y = "y_from",
+  end_x = "x_to",
+  end_y = "y_to",
+  select_vars = c("from", "to", "distance"))
+
+## To find the benchmark value
+benchmark_pbmc <- find_lg_benchmark(
+  distance_edges = distance_pbmc,
+  distance_col = "distance")
+
+benchmark_pbmc <- 0.1
+
+trimesh_removed_pbmc <- vis_rmlg_mesh(
+  distance_edges = distance_pbmc,
+  benchmark_value = benchmark_pbmc,
+  tr_coord_df = tr_from_to_df_pbmc,
+  distance_col = "distance") +
+  #xlim(sc_xlims) + ylim(sc_ylims) +
+  interior_annotation("a", sc_ltr_pos)
+
+## Compute error
+error_df <- augment(
+  df_bin_centroids = df_bin_centroids_pbmc,
+  df_bin = df_bin_pbmc,
+  training_data = training_data_pbmc,
+  newdata = NULL,
+  type_NLDR = "tSNE",
+  col_start = "PC_")
+
+## Categorize error
+
+error_df <- error_df |>
+  mutate(type = case_when(
+    row_wise_abs_error <= 5 ~ "error 0-5",
+    row_wise_abs_error <= 10 ~ "error 5-10",
+    row_wise_abs_error <= 15 ~ "error 10-15",
+    row_wise_abs_error <= 20 ~ "error 15-20",
+    row_wise_abs_error <= 25 ~ "error 20-25",
+    .default = "error greter than 25"
+  )) |>
+  mutate(type = factor(type, levels = c(
+    "error 0-5", "error 5-10", "error 10-15", "error 15-20", "error 20-25", 
+    "error greter than 25")))
+
+## To join embedding
+error_df <- error_df |>
+  bind_cols(tsne_pbmc_scaled |>
+              select(-ID))
+
+error_plot_pbmc <- error_df |>
+  ggplot(aes(x = tSNE1,
+             y = tSNE2,
+             color = type,
+             group = ID)) +
+  geom_point(alpha=0.5,
+             size = 0.1) +
+  interior_annotation("b", sc_ltr_pos)
+
+```
+
+<!--bin1 = 15, bin2 = 20, b = 300, non_empty = 137-->
+```{r}
+#| echo: false
+#| fig-cap: "(a) Model generated in the 2D space overlaid on tSNE data, and (b) high-D model error in model space."
+#| label: fig-model-pbmc
 #| fig-pos: H
@@
   theme(strip.background = element_blank(),
-        strip.text.x = element_blank())
+        strip.text.x = element_blank(), 
+        legend.position = "none")
 ```
@@
   theme(strip.background = element_blank(),
-        strip.text.x = element_blank())
+        strip.text.x = element_blank(),
+        legend.position = "none")
 ```

### 2024-05-13 paper.tex
Initial flow of finding the best model for PBMC3k

@@
 
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-abserrorPBMC-1.pdf}
+
+}
+
+\caption{\label{fig-abserrorPBMC}Absolute error from UMAP and tSNE
+applied to training PBMC3k dataset with diffferent parameter choices.
+What is the best parameter choice to create the model? The residual plot
+have a steep slope at the beginning, indicating that a smaller number of
+non-empty bins causes a larger amount of error. Then, the slope
+gradually declines or level off, indicating that a higher number of
+non-empty bins generates a smaller error. Using the elbow method, it was
+observed that when the number of non-empty bins is set to 137, the
+lowest error occurred with the parameters perplexity: 30.}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-model-pbmc-1.pdf}
+
+}
+
+\caption{\label{fig-model-pbmc}(a) Model generated in the 2D space
+overlaid on tSNE data, and (b) high-D model error in model space.}
+
+\end{figure}%
+
 \subsection{digits: 1}\label{digits-1}

### 2024-05-13 applications/pbmc3k/3_gen_umap_diff_param.R

@@ -1,0 +1,98 @@
+library(readr)
+library(uwot)
+library(dplyr)
+library(ggplot2)
+library(quollr)
+set.seed(20240110)
+
+n_neighbors <- 5
+min_dist <- 0.3
+
+## Select PCs
+training_data_pbmc <- read_rds("data/pbmc3k/pbmc_pca_50.rds")
+training_data_pbmc <- training_data_pbmc[, 1:9] |>
+  mutate(ID = 1:NROW(training_data_pbmc))
+
+## Obtain UMAP
+umap_pbmc <- umap(training_data_pbmc |> select(-ID),
+                  n_neighbors = n_neighbors,
+                  n_components =  2,
+                  metric = "cosine",
+                  min_dist = min_dist,
+                  init = "spca")
+
+umap_pbmc <- umap_pbmc |>
+  as.data.frame()
+
+names(umap_pbmc)[1:(ncol(umap_pbmc))] <- paste0(rep("UMAP",(ncol(umap_pbmc))), 1:(ncol(umap_pbmc)))
+
+umap_pbmc <- umap_pbmc |>
+  mutate(ID = 1:NROW(umap_pbmc))
+
+## Run only once
+write_rds(umap_pbmc, file = paste0("data/pbmc3k/pbmc_umap_", n_neighbors, "_min_dist_", min_dist, ".rds"))
+
+## Scaled data
+pbmc_scaled_obj <- gen_scaled_data(
+  data = umap_pbmc)
+umap_pbmc_scaled <- pbmc_scaled_obj$scaled_nldr
+
+
+## To initialise number of bins along the x-axis
+bin1_vec <- 2:40
+
+lim1 <- pbmc_scaled_obj$lim1
+lim2 <- pbmc_scaled_obj$lim2
+r2_pbmc <- diff(lim2)/diff(lim1)
+
+error_pbmc <- data.frame(matrix(nrow = 0, ncol = 0))
+
+for (xbins in bin1_vec) {
+
+  bin2 <- calc_bins_y(bin1 = xbins, q = 0.1, r2 = r2_pbmc)$bin2
+
+  pbmc_model <- fit_highd_model(
+    training_data = training_data_pbmc,
+    emb_df = umap_pbmc_scaled,
+    bin1 = xbins,
+    q = 0.1,
+    r2 = r2_pbmc,
+    is_bin_centroid = TRUE,
+    is_rm_lwd_hex = FALSE,
+    col_start_highd = "PC_"
+  )
+
[…]

### 2024-05-13 applications/pbmc3k/4_gen_tsne.R

@@ -1,143 +1,0 @@
-library(readr)
-library(Rtsne)
-library(dplyr)
-library(ggplot2)
-library(quollr)
-set.seed(20240110)
-
-perplexity <- 51
-
-## Select PCs
-training_data_pbmc <- read_rds("data/pbmc3k/pbmc_pca_50.rds")
-training_data_pbmc <- training_data_pbmc[, 1:9] |>
-  mutate(ID = 1:NROW(training_data_pbmc))
-
-## Obtain tSNE
-tSNE_data <- Fit_tSNE(training_data_pbmc |> select(-ID), opt_perplexity = perplexity, with_seed = 20240110)
-
-tSNE_data <- tSNE_data |>
-  select(-ID) |>
-  mutate(ID = training_data_pbmc$ID)
-
-## Run only once
-write_rds(tSNE_data, file = paste0("data/pbmc3k/pbmc_tsne_", perplexity, ".rds"))
-
-## Scaled data
-tsne_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tSNE_data,
-                                                                 x = "tSNE1", y = "tSNE2"))) |>
-  dplyr::rename(c("tSNE1" = "scaled_tSNE1",
-                  "tSNE2" = "scaled_tSNE2")) |>
-  dplyr::mutate(ID = 1:NROW(tSNE_data))
-
-## UMAP
-## Prediction
-
-hex_size_vec <- seq(0.01, 2, by = 0.01)
-
-vec <- stats::setNames(rep("", 7), c("num_bins", "error", "mse", "num_bins_x", "num_bins_y", "hex_size", "num_non_empty_bins"))  ## Define column names
-
-mse_df_pbmc <- dplyr::bind_rows(vec)[0, ]
-mse_df_pbmc <- mse_df_pbmc |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(hex_size_vec)) {
-
-  num_bin_list <- calc_bins(data = tsne_pbmc_scaled,
-                            x = "tSNE1", y = "tSNE2",
-                            hex_size = hex_size_vec[i], buffer_x = NA, buffer_y = NA)
-
-  num_bins_x <- num_bin_list$num_x
-  num_bins_y <- num_bin_list$num_y
-
-  model_object <- fit_highd_model( training_data = training_data_pbmc,
-                                   nldr_df_with_id = tsne_pbmc_scaled,
-                                   x = "tSNE1", y = "tSNE2",
[…]

### 2024-05-13 paper.qmd
Changed color palattes

@@
              color = "black", size=0.8, alpha = 0.5) +
-  scale_color_manual(values=c('#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99',
-                              '#e31a1c','#fdbf6f','#ff7f00','#cab2d6','#6a3d9a', 
-                              '#b15928', '#ffff99')) +
+  scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a','#984ea3',
+                              '#ff7f00','#a6cee3','#a65628','#f781bf')) +
   ylab("log(absolute error)") +
@@
   distance_col = "distance")
-
 benchmark_pbmc <- 0.1

### 2024-05-13 supplementary/hexagon_binning_algorithm.qmd
Supplementary for hexbin algorithm

@@
   message = FALSE,
-  warning = FALSE
+  warning = FALSE,
+  echo = FALSE
 )
+```
 
+```{r}
 library(ggplot2)
 library(dplyr)
+library(quollr)
+library(readr)
 
@@
 
-The 2D hexagon grid is defined by the number of bins in each direction $(b_1, b_2)$, as given by the centroids of each hexagon $(c_1, c_2)$ and the lower left position where the grid starts at $(s_1, s_2)$, which correspond to the lowest left centroid.
-The values of $s_i$ need to be below their respective minimum variable values, and could be a full bin lower, to allow a buffer corresponding to a distance between centroids in each direction ($a_1$ and $a_2$) around the data.
-The values of $b_i$ are variables to be computed that define the reduction in size of the data ($n$ to $m$).
+Given range of embedding component, $r_2$, number of bins along the x-axis, $b_1$, and buffer proportion, $q$, hexagonal starting point coordinates, $s_1 = -q$, and $s_2 = -q \times r_2$.
+The purpose is to find width of the hexagon.
+$a_1$, and number of bins along the y-axis, $b_2$.
 
-The value for $b_2$ is computed by fixing $b_1$. Also, according to hexagonal ratio $a_2 = \frac{\sqrt{3}}{2}a_1$.
+```{r}
+# umap_scurve <- read_rds(file = "data/s_curve/s_curve_umap.rds") 
+# 
+# scurve_scaled_obj <- gen_scaled_data(
+#   data = umap_scurve)
+# 
+# umap_scurve_scaled <- scurve_scaled_obj$scaled_nldr
+# lim1 <- scurve_scaled_obj$lim1
+# lim2 <- scurve_scaled_obj$lim2
+# r2 <- diff(lim2)/diff(lim1)
+# 
+# a1 <- calc_bins_y(bin1 = 6, q = 0.1, r2 = r2)$a1
+r2 <- 1.643542
+a1 <- 0.2087578
+```
 
-The first step of this process is to find the optimal number of bins along the y-axis ($b_1$) at the minimum width of the hexagon ($a_1$) that satisfies the following constraints.
+```{r}
+all_centroids_df <- gen_centroids(bin1 = 6, q = 0.1, r2 = r2)
+```
+
+```{r}
+hex_grid <- gen_hex_coord(all_centroids_df, a1 = a1)
+```
+
+```{r}
+hex_param_vis <- ggplot(data = hex_grid, aes(x = x, y = y)) + 
+    geom_polygon(fill = "white", color = "#bdbdbd", aes(group = hex_poly_id)) +
+    geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "#31a354") +
+    geom_point(data = all_centroids_df |> dplyr::filter(hexID == 1), aes(x = c_x, y = c_y), color = "black") + ## starting point
+    geom_rect(aes(xmin = 0, xmax = 1, ymin = 0, ymax = r2), fill = "white", color = "black", alpha = 0, linewidth = 1) +
+    geom_segment(x = 0, y = -0.35, xend = 1, yend = -0.35,
+                 arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+
+    annotate("text", x=0.5, y=-0.45, label= "r1", color = "black") +
+    geom_segment(x = -0.25, y = 0, xend = -0.25, yend = r2,
+                 arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ 
+    annotate("text", x=-0.35, y=1, label= "r2", color = "black") +
+    geom_segment(x = -0.1 + 0.2087578, y = -0.15, xend = -0.1 + 0.2087578*2, yend = -0.15,
+                   arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ # a1 = 0.2087578
+    annotate("text", x=0.2, y=-0.25, label= "a1", color = "black") +
+    geom_segment(x = -0.15, y = -0.1*r2 + 0.1807896*2, xend = -0.15, yend = -0.1*r2 + 0.1807896*3,
+                     arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ # a2 = 0.1807896
+    annotate("text", x=-0.15, y=0.45, label= "a2", color = "black") +
+    annotate("text", x=-0.2, y=-0.35, label= "(s1, s2)", color = "black") + ## starting point
+  coord_equal()
+```
+
+```{r}
+#| label: fig-hex-param
+#| fig-cap: "Hexagon binning parameters."
+
+hex_param_vis
+```
+
+Geometric arguments give rise to the following constraints.
 
@@
 a_1, a_2 > 0.
-$$
-Also, $(s_1, s_2) \in (-0.1, -0.05)$ as these are multiplicative offsets in the negative direction.
+$$ Also, $(s_1, s_2) \in (-0.1, -0.05)$ as these are multiplicative offsets in the negative direction.
 
@@
 
-which given $s_1 < 0$ and $a_1 > 0$ will *always* be true. The same logic follows for @eq-equation4 and substituting $a_2 = \frac{\sqrt{3}}{2}a_1$ to @eq-equation4 can be written as,
+which given $s_1 < 0$ and $a_1 > 0$ will *always* be true.
+The same logic follows for @eq-equation4 and substituting $a_2 = \frac{\sqrt{3}}{2}a_1$, and $s_2 = -q \times r_2$ to @eq-equation4 can be written as,
 
 $$
-a_1 > \frac{4}{\sqrt{3}}s_2.
+a_1 > -\frac{4}{\sqrt{3}}qr_2
 $$ 
-Also, substituting $a_2 = \frac{\sqrt{3}}{2}a_1$ and rearranging @eq-equation5 gives:
+Also, substituting $a_2 = \frac{\sqrt{3}}{2}a_1$, $s_2 = -q \times r_2$ and rearranging @eq-equation5 gives:
 
 $$
-a_1 > \frac{2(r_2 - s_2)}{\sqrt{3}(b_2 - 1)}.
-$${#eq-equation6}
+a_1 > \frac{2(r_2 + qr_2)}{\sqrt{3}(b_2 - 1)}.
+$$ {#eq-equation6}
 
@@
 $$
-a_1 > \frac{(1-s_1)}{(b_1 - 1)}.
-$${#eq-equation7}
+a_1 > \frac{(1 + q)}{(b_1 - 1)}.
+$$ {#eq-equation7}
 
@@
 $$
-\frac{2(r_2 - s_2)}{\sqrt{3}(b_2 - 1)} = \frac{(1-s_1)}{(b_1 - 1)}.
-$$
-After rearranging this,
+\frac{2(r_2 + qr_2)}{\sqrt{3}(b_2 - 1)} = \frac{(1 + q)}{(b_1 - 1)}.
+$$ After rearranging this,
 
 $$
-b_2 = 1 + \frac{2(r_2 - s_2)(b_1 - 1)}{\sqrt{3}(1-s_1)}
+b_2 = 1 + \frac{2r_2(b_1 - 1)}{\sqrt{3}}
 $$
@@
 $$
-b_2 = \Big\lceil1 +\frac{2(r_2 - s_2)(b_1 - 1)}{\sqrt{3}(1-s_1)}\Big\rceil.
+b_2 = \Big\lceil1 +\frac{2r_2(b_1 - 1)}{\sqrt{3}}\Big\rceil.
 $$ {#eq-equation8}
@@
 $$
-\frac{(1-s_1)}{(b_1 - 1)} < \frac{2(r_2 - s_2)}{\sqrt{3}(b_2 - 1)}.
+\frac{(1 + q)}{(b_1 - 1)} < \frac{2(r_2 + qr_2)}{\sqrt{3}(b_2 - 1)}.
 $$
+
 Rearranging this gives,
@@
 $$
-r_2 > \frac{\sqrt{3}(b_2 - 1)(1-s_1)}{2(b_1 - 1)} + s_2.
+r_2 > \frac{\sqrt{3}(b_2 - 1)}{2(b_1 - 1)}.
 $$
-Therefore, if this equality is true, then $a_1 = \frac{(1-s_1)}{(b_1 - 1)}$, otherwise, $a_1 = \frac{2(r_2 - s_2)}{\sqrt{3}(b_2 - 1)}$.
 
-```{r, echo=FALSE}
-# bin1: Number of bins along the x-axis
-# buffer: proportion of buffer should between 0 and 1
-# r2: range of embedding 2
-calc_bins_y <- function(bin1 = 2, s1 = -0.1, s2 = -0.1, r2) {
-
-  ## Need to have a check where bin2 greater than 2
-  if (bin1 < 2) {
-    stop("Number of bins along the x-axis at least should be 2.")
-  }
-
-  ## To compute the number of bins along the x-axis
-  bin2 <- ceiling(1 + (2 * (r2-s2) * (bin1 - 1))/(sqrt(3) * (1-s1)))
-
-  ## Validating and compute horizontal spacing
-  check_factor <- (sqrt(3) * (bin2 - 1) * (1-s1))/(2 * (bin1 - 1)) + s2
-
-  if (r2 > check_factor) {
-
-    a1 <- (1-s1)/(bin1 - 1)
-
-  } else {
-
-    a1 <- (2 * (r2-s2))/(sqrt(3) * (bin2 - 1))
-
-  }
-
-  return(list(bin2 = bin2, a1 = a1))
-}
-```
-
-```{r, echo=FALSE}
-gen_centroids <- function(bin1, s1 = -0.1, s2 = -0.1, r2){
-
-  ## To compute hexagonal configurations
-  bin_obj <- calc_bins_y(bin1 = bin1, s1 = s1, s2 = s2, r2 = r2)
-
-  # To obtain the bins along the y-axis
-  bin2 <- bin_obj$bin2
-
-  # To obtain the width of the hexagon
-  a1 <- bin_obj$a1
-
-  # Generate x-coordinate of centroids for odd rows
-  c_x_vec_odd <- seq(s1, (bin1 - 1) * a1, by = a1)
-
-  # Generate x-coordinate of centroids for even rows
-  c_x_vec_even <- c_x_vec_odd + a1/2
-  c_x_vec <- c(c_x_vec_odd, c_x_vec_even)
-
-  # To compute vertical spacing
-  a2 <- sqrt(3) * a1/2
-
-  # Generate y-coordinate of centroids
-  c_y_vec <- seq(s2, (bin2 - 1) * a2, by = a2)
-  c_y <- rep(c_y_vec, each = bin1)
-
-  ## Do the number of belongs y axis is even or odd and adjust the x-coordinates
-  if ((bin2 %% 2) == 0) {
-
-    c_x <- rep(c_x_vec, bin2/2)
-
-  } else {
-
-    if ((ceiling(bin2/2) %% 2) == 0) {
-
-      c_x <- append(rep(c_x_vec, floor(bin2/2)), c_x_vec_odd)
-
-    } else{
-
-      c_x <- append(rep(c_x_vec, floor(bin2/2)), c_x_vec_even)
-
-    }
-
-  }
-
-  centroid_df <- tibble::tibble(hexID = 1:length(c_x), c_x = c_x, c_y = c_y)
-
-  return(centroid_df)
-
-}
-```
-
-```{r, echo=FALSE}
-all_centroids_df <- gen_centroids(bin1 = 4, s1 = -0.1, s2 = -0.1, r2 = 2.019414)
-```
-
-```{r, echo=FALSE}
-gen_hex_coord <- function(centroids_df, a1){
-
-  ## Obtain centroid info
-  hex_ids <- centroids_df$hexID
-  c_x_vec <- centroids_df$c_x
-  c_y_vec <- centroids_df$c_y
-
-  ## To compute vertical spacing factor
-  vs_factor <- a1/(2 * sqrt(3))
-
-  dx <- a1/2
-  dy <- a1/sqrt(3)
-
-  ## Assign coordinates for 6 directions
-  x_add_factor <- c(0, -dx, -dx, 0, dx, dx)
-  y_add_factor <- c(dy, vs_factor, -vs_factor, -dy, -vs_factor, vs_factor)
-
-  ## Initialize vectors to store hexagonal coordinates
-  hex_poly_id <- integer(0)
-  x <- numeric(0)
-  y <- numeric(0)
-
-  for (hb_id in hex_ids) {
-
-    ## Since each hexagon has 6 coordinates
-    hexID_rep <- rep(hex_ids[hb_id], each = 6)
-    c_x_rep <- rep(c_x_vec[hb_id], each = 6)
-    c_y_rep <- rep(c_y_vec[hb_id], each = 6)
-
-    ## Generate the 6 coordinates
-    x_spec <- c_x_rep + x_add_factor
-    y_spec <- c_y_rep + y_add_factor
-
-    ## Append to existing vectors
-    x <- append(x, x_spec)
-    y <- append(y, y_spec)
-    hex_poly_id <- append(hex_poly_id, hexID_rep)
-
-  }
-
-  hex_coord_df <- tibble::tibble(hex_poly_id = hex_poly_id, x = x, y = y)
-
-  return(hex_coord_df)
-}
-```
-
-```{r, echo=FALSE}
-hex_grid <- gen_hex_coord(all_centroids_df, a1 = 0.3496126)
-```
-
-```{r, echo=FALSE}
-ggplot(data = hex_grid, aes(x = x, y = y)) + 
-    geom_polygon(fill = "white", color = "#bdbdbd", aes(group = hex_poly_id)) +
-    geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "#31a354") +
-    geom_point(data = all_centroids_df |> dplyr::filter(hexID == 1), aes(x = c_x, y = c_y), color = "#08519c") + ## starting point
-    geom_rect(aes(xmin = 0, xmax = 1, ymin = 0, ymax = 2.019414), fill = "white", color = "#feb24c", alpha = 0, linewidth = 1) +
-    geom_segment(x = 0, y = -0.25, xend = 1, yend = -0.25,
-                 arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "#f03b20")+
-    annotate("text", x=0.5, y=-0.35, label= "r1", color = "#f03b20") +
-    geom_segment(x = -0.25, y = 0, xend = -0.25, yend = 2.019414,
-                 arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "#f03b20")+ 
-    annotate("text", x=-0.35, y=1, label= "r2", color = "#f03b20") +
-    geom_segment(x = -0.1, y = -0.15, xend = -0.1 + 0.3496126, yend = -0.15,
-                   arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ # a1 = 0.3496126
-    annotate("text", x=0.17, y=-0.35, label= "a1", color = "black") +
-    geom_segment(x = -0.15, y = -0.1, xend = -0.15, yend = -0.1 + 0.3027734,
-                     arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ # a2 = 0.3027734
-    annotate("text", x=-0.35, y=0.15, label= "a2", color = "black") +
-    annotate("text", x=-0.15, y=-0.35, label= "(s1, s2)", color = "#08519c") + ## starting point
-    coord_fixed()
-```
+Therefore, if this equality is true, then $a_1 = \frac{(1+q)}{(b_1 - 1)}$, otherwise, $a_1 = \frac{2r_2(1+q)}{\sqrt{3}(b_2 - 1)}$.
 
@@
 
-To find the nearest centroid to a $k^{th}$ embedding point $(y_1, y_2)$ among centroids $(C_1^{(k)}, C_2^{(k)}, \ldots, C_b^{(k)})$,
-
-$$
-\text{Minimize: }\sqrt{(y_1 - c_{1k})^2 + (y_2 - c_{2k})^2} + \text{centroid\_ID} \times 10^{-3}. 
-$$
-
-In this expression, $(c_{1k}, c_{2k})$ represents the coordinates of centroid $C_1^{(k)}$, and $\text{centroid\_ID}$ represents the ID of the centroid.
-This formulation ensures that if multiple centroids have the same minimum distance, the one with the minimum ID will be selected.
-
-```{r, echo=FALSE}
+```{r}
 emb_df <- tibble::tibble(emb1 = 0.774, emb2 = 0.5055) |>
@@
     geom_segment(x = 0.6865, y = 0.3545, xend = 0.6865, yend = 0.25,
-                       arrow = arrow(length = unit(0.03, "npc")), color = "black")
+                       arrow = arrow(length = unit(0.03, "npc")), color = "black") +
+  coord_equal()
 

### 2024-05-14 supplementary/hexagon_binning_algorithm.qmd
Update hexbin algorithm with buffer: q

@@
 library(readr)
+library(ggforce)
 
@@
 #| label: fig-hex-param
-#| fig-cap: "Hexagon binning parameters."
+#| fig-cap: "Notations for hexagonal grid configurations."
 
@@
 
+\newpage
+
 ## Binning the data
 
-Points are allocated to the bin that they fall in by considering the nearest centroid.
+Points are assigned to the bin they fall into based on the nearest centroid. If a point is equidistant from multiple centroids, it is assigned to the centroid with the lowest hexagonal bin ID.
 
 ```{r}
-emb_df <- tibble::tibble(emb1 = 0.774, emb2 = 0.5055) |>
-  add_row(emb1 = 0.424, emb2 = 0.5055) |>
-  add_row(emb1 = 0.5115, emb2 = 0.3545) |>
-  add_row(emb1 = 0.5115, emb2 = 0.6565) |>
-  add_row(emb1 = 0.6865, emb2 = 0.3545) |>
-  add_row(emb1 = 0.6865, emb2 = 0.6565) |>
-  add_row(emb1 = 0.599, emb2 = 0.304) |>
-  add_row(emb1 = 0.599, emb2 = 0.707) |>
-  add_row(emb1 = 0.424, emb2 = 0.405) |>
-  add_row(emb1 = 0.424, emb2 = 0.606) |>
-  add_row(emb1 = 0.774, emb2 = 0.405) |>
-  add_row(emb1 = 0.774, emb2 = 0.606)
 
-ggplot(data = hex_grid, aes(x = x, y = y)) + 
+## To generate embedding data
+emb_df <- tibble::tibble(emb1 = 0.37, emb2 = 0.3) |>
+  add_row(emb1 = 0.526, emb2 = 1.1) |>
+  add_row(emb1 = 0.735, emb2 = 1.1) |>
+  add_row(emb1 = 0.631, emb2 = 1.22) |>
+  add_row(emb1 = 0.735, emb2 = 1.16) |>
+  add_row(emb1 = 0.683, emb2 = 1.19) |>
+  add_row(emb1 = 0.735, emb2 = 1.04) |>
+  add_row(emb1 = 0.631, emb2 = 0.981) |>
+  add_row(emb1 = 0.683, emb2 = 1.0105) |>
+  add_row(emb1 = 0.526, emb2 = 1.04) |>
+  add_row(emb1 = 0.5785, emb2 = 1.0105) |> 
+  add_row(emb1 = 0.526, emb2 = 1.16) |>
+  add_row(emb1 = 0.5785, emb2 = 1.19)
+
+## Data to generate circle
+circles <- data.frame(
+  x = c(0.318, 0.422),
+  y = c(0.197,0.378),
+  r = rep(0.1205264,2)
+)
+
+assign_data_vis <- ggplot(data = hex_grid, aes(x = x, y = y)) + 
   geom_polygon(fill = "white", color = "#bdbdbd", aes(group = hex_poly_id)) +
-    geom_point(data = emb_df, aes(x = emb1, y = emb2), color = "#253494") +
-    geom_segment(x = 0.774, y = 0.5055, xend = 0.7, yend = 0.5055,
-                       arrow = arrow(length = unit(0.03, "npc")), color = "black") +
-    geom_segment(x = 0.424, y = 0.5055, xend = 0.35, yend = 0.5055,
-                       arrow = arrow(length = unit(0.03, "npc")), color = "black") +
-    geom_segment(x = 0.599, y = 0.707, xend = 0.599, yend = 0.6,
-                       arrow = arrow(length = unit(0.03, "npc")), color = "black") +
-    geom_segment(x = 0.774, y = 0.405, xend = 0.774, yend = 0.3,
-                       arrow = arrow(length = unit(0.03, "npc")), color = "black") +
-    geom_segment(x = 0.774, y = 0.606, xend = 0.7, yend = 0.55,
-                       arrow = arrow(length = unit(0.03, "npc")), color = "black") +
-    geom_segment(x = 0.6865, y = 0.6565, xend = 0.62, yend = 0.6,
-                       arrow = arrow(length = unit(0.03, "npc")), color = "black") +
-    geom_segment(x = 0.6865, y = 0.3545, xend = 0.6865, yend = 0.25,
-                       arrow = arrow(length = unit(0.03, "npc")), color = "black") +
+  geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID), size = 2) +
+  geom_point(data = emb_df, aes(x = emb1, y = emb2), color = "#253494", size = 0.5) +
+  geom_circle(data = circles, aes(x0 = x, y0 = y, r = r), fill = NA) +
+  geom_segment(x = 0.37, y = 0.3, xend = 0.422, yend = 0.378, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
+  geom_segment(x = 0.526, y = 1.1, xend = 0.45, yend = 1.1, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
+  geom_segment(x = 0.735, y = 1.1, xend = 0.66, yend = 1.1, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
+  geom_segment(x = 0.631, y = 1.22, xend = 0.631, yend = 1.15, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
+  geom_segment(x = 0.735, y = 1.16, xend = 0.65, yend = 1.12, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
+  geom_segment(x = 0.735, y = 1.04, xend = 0.735, yend = 0.95, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
+  geom_segment(x = 0.683, y = 1.19, xend = 0.65, yend = 1.12, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
+  geom_segment(x = 0.631, y = 0.981, xend = 0.55, yend = 0.95, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
+  geom_segment(x = 0.683, y = 1.0105, xend = 0.7, yend = 0.95, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
+  geom_segment(x = 0.526, y = 1.04, xend = 0.526, yend = 0.95, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
+  geom_segment(x = 0.5785, y = 1.0105, xend = 0.536, yend = 0.97, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
+  geom_segment(x = 0.526, y = 1.16, xend = 0.49, yend = 1.12, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
+  geom_segment(x = 0.5785, y = 1.19, xend = 0.58, yend = 1.12, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
   coord_equal()
@@
 ```
+```{r}
+#| label: fig-assign-data
+#| fig-cap: "Binning the data. Points are assigned to the nearest centroid. If a point is equidistant from multiple centroids, assigned to the lowest centroid."
+
+assign_data_vis
+```

### 2024-05-14 paper.qmd
Update hexbin algorithm with buffer: q
Update hexbin configurations with q

@@
 The value for $b_2$ is computed by fixing $b_1$.
-Considering the lower bound of the NLDR, $a_1 > 2 \times s_1$, and $a_1 > \frac{1-s_1}{b_1 -1}$.
-Similarly, according to the upper bound of the NLDR, $a_1 > \frac{2(r_2 - s_2)}{\sqrt{3}(b_2 - 1)}$, because $a_2 = \frac{\sqrt(3)}{2}a_1$ for regular hexagons.
-Therefore, $b_2 = \Big\lceil1 +\frac{2(r_2 - s_2)(b_1 - 1)}{\sqrt{3}(1-s_1)}\Big\rceil$.
+Considering the lower bound of the NLDR, $a_1 > -2q$, and $a_1 > \frac{1+q}{b_1 -1}$.
+Similarly, according to the upper bound of the NLDR, $a_1 > \frac{2r_2(1 + q)}{\sqrt{3}(b_2 - 1)}$, because $a_2 = \frac{\sqrt(3)}{2}a_1$ for regular hexagons.
+Therefore, $b_2 = \Big\lceil1 +\frac{2r_2(b_1 - 1)}{\sqrt{3}}\Big\rceil$.
 
@@
 #| label: fig-hex-param
-#| fig-cap: "Hexagon binning parameters."
+#| fig-cap: "Notations for hexagonal grid configurations."
 #| fig-width: 3

### 2024-05-14 paper.tex
Update hexbin algorithm with buffer: q
Update hexbin configurations with q

@@
 The value for \(b_2\) is computed by fixing \(b_1\). Considering the
-lower bound of the NLDR, \(a_1 > 2 \times s_1\), and
-\(a_1 > \frac{1-s_1}{b_1 -1}\). Similarly, according to the upper bound
-of the NLDR, \(a_1 > \frac{2(r_2 - s_2)}{\sqrt{3}(b_2 - 1)}\), because
+lower bound of the NLDR, \(a_1 > -2q\), and
+\(a_1 > \frac{1+q}{b_1 -1}\). Similarly, according to the upper bound of
+the NLDR, \(a_1 > \frac{2r_2(1 + q)}{\sqrt{3}(b_2 - 1)}\), because
 \(a_2 = \frac{\sqrt(3)}{2}a_1\) for regular hexagons. Therefore,
-\(b_2 = \Big\lceil1 +\frac{2(r_2 - s_2)(b_1 - 1)}{\sqrt{3}(1-s_1)}\Big\rceil\).
+\(b_2 = \Big\lceil1 +\frac{2r_2(b_1 - 1)}{\sqrt{3}}\Big\rceil\).
 
@@
 
-\caption{\label{fig-hex-param}Hexagon binning parameters.}
+\caption{\label{fig-hex-param}Notations for hexagonal grid
+configurations.}
 

### 2024-05-14 paper.qmd
Plan for writing

@@
 
-The 2D hexagon grid is defined by the number of bins in each direction $(b_1, b_2)$, as given by the centroids of each hexagon $(c_1, c_2)$ and the lower left position where the grid starts at $(s_1, s_2)$, which correspond to the lowest left centroid.
+The 2D hexagon grid is defined by the number of bins in each direction $(b_1, b_2)$, giving total number of bins as $b = b_1 \times b_2$, and hexagon id, $h = 1, \dots, b$.
+Each hexagon, $H_h$ is uniquely described by centroid, $ C_{h}^{(2)} = (c_{h1}, c_{h2})$.
+The lower left position where the grid starts at $(s_1, s_2)$, which correspond to the lowest left centroid.
 The values of $s_i$ need to be below their respective minimum variable values, and could be a full bin lower, to allow a buffer ($q$) corresponding to a full hexagon width ($a_1$) and height ($a_2$) around the data.
@@
 
+$\{ i \in H_h, h = 1, \dots, b, \text{ and } i = 1, \dots, n\}$
+
 <!-- XXX How are you doing this? Do you check the bounds of the hexagon? Or do you use distance to centroid? -->
@@
 <!--We deliberately separate out the creation of the hexagonal grid from the mapping of points on the grid.-->
-#### Obtaining bin representations
-
-Let $h : \mathbb{R}^{n\times k} \rightarrow \mathbb{R}^{m\times k}$ be the function that maps a point in \kD{} space to it's bin representation ($C^{(k)}$). This is the model points in \kD{} space.
+#### Summarization of the data in 2D
+
+Let $h : \mathbb{R}^{n\times k} \rightarrow \mathbb{R}^{m\times k}$ be the function that maps a point in \kD{} space to it's bin centroid ($C^{(k)}$). This is the model points in \kD{} space.
 
@@
 
-#### Generating edges
-
-Delaunay triangulation on bin representations generates edges by specifying which bin representations should be connected to generate the model in \kD{} space.
+#### Generating edges to indicate neighbours and removing long edges
+
+Delaunay triangulation is used to connect neighbouring centroids, which is needed to preserve neighbourhood information when the model is lifted into \pD{}.
 
@@
 
-The final step of the algorithm is _lifting_ the \kD{} model back to \pD{}, completing the *model-in-the-data-space* picture.
-For this, define the set $H^k$ to be the set of points that belong to centroid $k$.
-That is, $H^k = \{ y \; | \; g(y) = (h_x^k, h_y^k)\}$.
+The last step is to lift the \kD{} model into \pD{} by computing \pD{} vectors that represent bin centroids.
+Define the set $H^k$ to be the set of points that belong to bin $k$.
 We use the \pD{} Euclidean mean of the points in $H^k$ to map the centroid $(h_x^k, h_y^k)$ to a point in \pD{}.
@@
 
+<!--langevitour with pD model for S-curve-->
+
 ### Measuring the fit
-
-### What can be learned
-
-- Overview: Generate a form that maps the model, that is the interpoint distances. What is the model?
-- Notation
-- Create a representation of the model 
-    - using hex-binning in 2D, 
-    - parameters, 
-    - tuning, 
-    - pre-processing
-- How does this map to the representation in high-d
-    - Centroids, 
-    - Edges
-- Measuring fit
-    - Fitted values
-    - Error calculation
-- What is learned about simulated examples
-    - Interesting organisation of points in UMAP
-    -
+ <!-- Fitted values,  Error calculation-->
+
+### Tuning
+
+<!-- removal of low density bins, removing long edges, choice of bins-->
+
+### Illustration on simulated data
 

### 2024-05-14 paper.tex
Plan for writing

@@
 The 2D hexagon grid is defined by the number of bins in each direction
-\((b_1, b_2)\), as given by the centroids of each hexagon \((c_1, c_2)\)
-and the lower left position where the grid starts at \((s_1, s_2)\),
-which correspond to the lowest left centroid. The values of \(s_i\) need
-to be below their respective minimum variable values, and could be a
-full bin lower, to allow a buffer (\(q\)) corresponding to a full
-hexagon width (\(a_1\)) and height (\(a_2\)) around the data. The values
-of \(b_i\) are variables to be computed that define the reduction in
-size of the data (\(n\) to \(m\)).
+\((b_1, b_2)\), giving total number of bins as \(b = b_1 \times b_2\),
+and hexagon id, \(h = 1, \dots, b\). Each hexagon, \(H_h\) is uniquely
+described by centroid, \$ C\_\{h\}\^{}\{(2)\} = (c\_\{h1\},
+c\_\{h2\})\$. The lower left position where the grid starts at
+\((s_1, s_2)\), which correspond to the lowest left centroid. The values
+of \(s_i\) need to be below their respective minimum variable values,
+and could be a full bin lower, to allow a buffer (\(q\)) corresponding
+to a full hexagon width (\(a_1\)) and height (\(a_2\)) around the data.
+The values of \(b_i\) are variables to be computed that define the
+reduction in size of the data (\(n\) to \(m\)).
 
@@
 
-\subsubsection{Obtaining bin
-representations}\label{obtaining-bin-representations}
+\(\{ i \in H_h, h = 1, \dots, b, \text{ and } i = 1, \dots, n\}\)
+
+\subsubsection{Summarization of the data in
+2D}\label{summarization-of-the-data-in-2d}
 
 Let \(h : \mathbb{R}^{n\times k} \rightarrow \mathbb{R}^{m\times k}\) be
-the function that maps a point in \kD{} space to it's bin representation
+the function that maps a point in \kD{} space to it's bin centroid
 (\(C^{(k)}\)). This is the model points in \kD{} space.
@@
 
-\subsubsection{Generating edges}\label{generating-edges}
+\subsubsection{Generating edges to indicate neighbours and removing long
+edges}\label{generating-edges-to-indicate-neighbours-and-removing-long-edges}
 
-Delaunay triangulation on bin representations generates edges by
-specifying which bin representations should be connected to generate the
-model in \kD{} space.
+Delaunay triangulation is used to connect neighbouring centroids, which
+is needed to preserve neighbourhood information when the model is lifted
+into \pD{}.
 
@@
 
-The final step of the algorithm is \emph{lifting} the \kD{} model back
-to \pD{}, completing the \emph{model-in-the-data-space} picture. For
-this, define the set \(H^k\) to be the set of points that belong to
-centroid \(k\). That is, \(H^k = \{ y \; | \; g(y) = (h_x^k, h_y^k)\}\).
-We use the \pD{} Euclidean mean of the points in \(H^k\) to map the
-centroid \((h_x^k, h_y^k)\) to a point in \pD{}. Let the \(i\)-th
-component of the \pD{} mean be
+The last step is to lift the \kD{} model into \pD{} by computing \pD{}
+vectors that represent bin centroids. Define the set \(H^k\) to be the
+set of points that belong to bin \(k\). We use the \pD{} Euclidean mean
+of the points in \(H^k\) to map the centroid \((h_x^k, h_y^k)\) to a
+point in \pD{}. Let the \(i\)-th component of the \pD{} mean be
 
@@
 
-\subsection{What can be learned}\label{what-can-be-learned}
+\subsection{Tuning}\label{tuning}
 
-\begin{itemize}
-\tightlist
-\item
-  Overview: Generate a form that maps the model, that is the interpoint
-  distances. What is the model?
-\item
-  Notation
-\item
-  Create a representation of the model
-
-  \begin{itemize}
-  \tightlist
-  \item
-    using hex-binning in 2D,
-  \item
-    parameters,
-  \item
-    tuning,
-  \item
-    pre-processing
-  \end{itemize}
-\item
-  How does this map to the representation in high-d
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Centroids,
-  \item
-    Edges
-  \end{itemize}
-\item
-  Measuring fit
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Fitted values
-  \item
-    Error calculation
-  \end{itemize}
-\item
-  What is learned about simulated examples
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Interesting organisation of points in UMAP
-  \item
-  \end{itemize}
-\end{itemize}
+\subsection{Illustration on simulated
+data}\label{illustration-on-simulated-data}
 

### 2024-05-19 paper.qmd
Added MNIST digit description, need to add the video

@@
   \newcommand\dD{$d\text{-}D$}
+  \newcommand\gD{$2\text{-}D$}
 ---
@@
 set.seed(20240110)
-source("nldr_code.R", local = TRUE)
+#source("nldr_code.R", local = TRUE)
 ```
@@
   bin1 = num_bins_x_scurve, 
-  q = 0.1,
   r2 = r2)
@@
   centroids_df = all_centroids_df, 
-  counts_df = counts_df)
+  counts_df = counts_df) |>
+  filter(drop_empty == FALSE)
 
@@
 
+<!--need to add what is meant by a long edge-->
+
 ### Displaying the model in \pD{}
@@
  <!-- Fitted values,  Error calculation-->
+
+#### Fitted values
+
+The prediction approach involves performing the K-nearest neighbors (KNN) algorithm for an unsupervised classification problem.
+First, the nearest high-D model point is identified for a given new high-D point.
+Then, the corresponding 2D centroid mapping for the identified high-D model point is determined.
+Finally, the coordinates of this 2D centroid are used as the predicted 2D embedding for the new high-D data point.
+This step is particularly valuable due to the limitations of some NLDR techniques, like tSNE, which don't provide a straightforward method for prediction.
+As a result, our approach offers a solution that capable of generating predicted 2D embedding regardless of the NLDR technique employed, effectively addressing this functional gap.
+
+#### Error calculation
+
+Residuals are essential for evaluating the accuracy of representing high-D points by the high-D mapping of 2D bin centroids.
+To measure this accuracy, an error metric is introduced, quantifying the sum of squared differences between the high-D data ($x_{ij}$) and the high-D mapping of the 2D bin centroid data ($C_{x_ij}$) across all observations and dimensions (see @eq-equation10).
+
+$$
+\text{Error} = \sum_{j = 1}^{n}\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2
+$$ {#eq-equation10}
+
+Here, $n$ represents the number of observations, $p$ represents the dimensions of high-D data, $x_{ij}$ is the high-D data, and $C_{x_ij}$ is the high-D mapping of the 2D bin centroid.
+
+To assess how well our method captures and represents the underlying structure of the high-D data, Mean Squared Error (MSE) is used.
+When computing MSE, total model error (see @sec-summary) is divided by the number of observations to make it as a mean value (see @eq-equation9).
+
+$$
+\text{MSE} = \sum_{j = 1}^{n} \frac{\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2}{n}
+$$ {#eq-equation9}
 
@@
   bin1 = num_bins_x_pbmc,
-  q = 0.1,
   r2 = r2_pbmc,
@@
 ### digits: 1
-
+<!--
 - NLDR is used to illustrate different ways 1's are drawn
@@
 - Demonstrate that it is, except for the anomalies 
+-->
+
+The MNIST dataset consists of grayscale images of handwritten digits [@lecun2010].
+@yingfan2021 used this dataset to demonstrate how PaCMAP preserves local structure.
+We selected the \gD{} embedding of PaCMAP for the handwritten digit 1 to assess whether this is a reasonable representation using our method.
+As shown in @fig-pacmap-author-img, the angle of the digit 1 images varies along the \gD{} structure.
 
@@
              y = PaCMAP2)) + 
-  geom_point(alpha=0.1) + 
-  geom_text(data = position_df, 
-            aes(x = PaCMAP1, y = PaCMAP2, label = position), 
-            colour = "red", 
-            size = 5)  
+  geom_point(alpha=0.1) #+ 
+  # geom_text(data = position_df, 
+  #           aes(x = PaCMAP1, y = PaCMAP2, label = position), 
+  #           colour = "red", 
+  #           size = 5)  
+```
+
+<!--PaCMAP param: n_components=2, n_neighbors=10, init=random, MN_ratio=0.9, FP_ratio=2.0-->
+```{r}
+#| fig-cap: "$2\\text{-}D$ layout from PaCMAP applied for the digit 1 of the MNIST dataset.
+Is this the best representation of the digit 1?
+The parameter setting is n_components=2, n_neighbors=10, init=random, MN_ratio=0.9, FP_ratio=2.0."
+#| label: fig-pacmap-author
+#| fig-pos: H
+#| out-height: 22%
+
+pacmap_plot_mnist
 ```
@@
                         5188, 4849, 2666, 3448, 3055,
-                        3120, 6869, 6345, 4470, 7147)), ncol = 5) +
+                        3120, 6869, 6345, 4470, 7147)), nrow = 5) +
+  coord_fixed() +
   scale_fill_continuous_sequential(palette = "Grays") +
@@
 
-<!--PaCMAP param: n_components=2, n_neighbors=10, init=random, MN_ratio=0.9, FP_ratio=2.0-->
-```{r}
-#| fig-cap: "(a) 2D layout from PaCMAP applied for the digit 1 of the MNIST dataset.
-Is this a best representation of the original data?, and (b) Images of the handwritten digit 1.
-The angle of the digit 1 varies along non-linear structure."
-#| label: fig-pacmapauthor
+```{r}
+#| fig-cap: "Images of the handwritten digit 1 are ordered from the bottom-right to the top-left of the $2\\text{-}D$ structure.
+The angle of the digit varies along this structure.
+Images at the bottom-right of the $2\\text{-}D$ layout show the digit 1 angled more to the right, while images at the top-left show the digit 1 angled more to the left.
+This demonstrates how the angle changes from right to left along the $2\\text{-}D$ structure."
+#| label: fig-pacmap-author-img
 #| fig-pos: H
-
-pacmap_plot_mnist + img_sample +
-  plot_annotation(tag_levels = 'a')
+#| out-height: 50%
+
+img_sample
 ```
@@
   bin1 = num_bins_x_mnist,
-  q = 0.1,
   r2 = r2_mnist,
@@
 
+sc_ltr_pos_mnist <- c(0.96, 0.96)
+
 trimesh_removed_mnist <- vis_rmlg_mesh(
@@
   #xlim(sc_xlims) + ylim(sc_ylims) +
-  interior_annotation("a", sc_ltr_pos)
+  interior_annotation("a", sc_ltr_pos_mnist)
 
@@
              size = 0.1) +
-  interior_annotation("b", sc_ltr_pos)
+  interior_annotation("b", sc_ltr_pos_mnist)
 ```
@@
 #| echo: false
-#| fig-cap: "(a) Model generated in the 2D space overlaid on PaCMAP data, and (b) high-D model error in model space."
+#| fig-cap: "(a) Model generated in $2\\text{-}D$, and (b) $p\\text{-}D$ model error in $2\\text{-}D$.
+The $2\\text{-}D$ model shows a non-linear continuous structure.
+Most low $p\\text{-}D$ model errors are distributed along the lower edge of the $2\\text{-}D$ structure, while most high p-D model errors are concentrated along the upper edge."
 #| label: fig-model-mnist
@@
 ```
+
+According to @fig-mnist1_sc1, the non-linear continuous structure observed in the \gD{} representation of PaCMAP (see @fig-pacmap-author) is also visible when visualizing the model overlaid on the data space.
+This indicates that PaCMAP accurately captures the structure of the \pD{} data.
+Additionally, the model shows a twisted pattern within the non-linear structure in \pD{} space (see @fig-mnist1_sc2), which is an additional pattern not visible in the \gD{} representation (see @fig-pacmap-author).
+Furthermore, as shown in @fig-mnist1_sc3, some long edges exist in the \pD{} space that are not recognized as long edges in the \gD{} representation.
+However, PaCMAP is a *good* \gD{} representation of MNIST digit 1 data.
+
+<!--add langevitour screenshots and youtube animation link-->
+
+::: {#fig-mnist1_sc layout-ncol="3" fig-pos="H"}
+![](figures/mnist/sc_1.png){#fig-mnist1_sc1}
+
+![](figures/mnist/sc_2.png){#fig-mnist1_sc2}
+
+![](figures/mnist/sc_3.png){#fig-mnist1_sc3}
+
+Screen shots of the **langevitour** of the MNIST digit 1 data set, shows the model-in-data space, a video of the tour animation is available at (<https://youtu.be/zq2GM9qvUNA>).
+:::
+<!-- need to update the youtube recording-->
+
+There are certain data points that exhibit high error rates due to their deviation from the usual \pD{} data structure, which makes them anomalies (see @fig-model-mnist (b)).
+These anomalies can be classified into two types: those that are anomalies within the non-linear structure and those that lie outside of it.
+The images associated with high model error points within the non-linear structure display different patterns of the digit 1, as shown in @fig-mnist-anomalies (a).
+However, when comparing these images to the ones found outside of the non-linear structure, it becomes evident that the latter display different patterns of the digit 1 (see @fig-mnist-anomalies (b)).
 
@@
                                             1227, 4042, 2444, 836, 1654)), 
-             ncol = 5) +
+             ncol = 4) +
+  coord_fixed() +
   scale_fill_continuous_sequential(palette = "Grays") +
@@
   facet_wrap(~ instance, ncol = 2) +
+  coord_fixed() +
   scale_fill_continuous_sequential(palette = "Grays") +
@@
 #| echo: false
-#| fig-cap: "(a) Images of handwritten digit 1 which occur large model error within the non-linear strcuture, and (b) Images of handwritten digit 1 which occur large model error outside the non-linear structure."
-#| label: fig-anomalies
-#| fig-width: 4
-#| fig-height: 4
-#| out-width: 100%
+#| fig-cap: "Some images of handwritten digit 1 which occur high model error (a) within the non-linear strcuture, and (b) outside the non-linear structure.
+The images shows different patterns of digit 1."
+#| label: fig-mnist-anomalies
+#| out-height: 35%
 #| fig-pos: H

### 2024-05-19 paper.tex
Added MNIST digit description, need to add the video

@@
 \newcommand\dD{$d\text{-}D$}
+\newcommand\gD{$2\text{-}D$}
 \makeatletter
@@
 
+\subsubsection{Fitted values}\label{fitted-values}
+
+The prediction approach involves performing the K-nearest neighbors
+(KNN) algorithm for an unsupervised classification problem. First, the
+nearest high-D model point is identified for a given new high-D point.
+Then, the corresponding 2D centroid mapping for the identified high-D
+model point is determined. Finally, the coordinates of this 2D centroid
+are used as the predicted 2D embedding for the new high-D data point.
+This step is particularly valuable due to the limitations of some NLDR
+techniques, like tSNE, which don't provide a straightforward method for
+prediction. As a result, our approach offers a solution that capable of
+generating predicted 2D embedding regardless of the NLDR technique
+employed, effectively addressing this functional gap.
+
+\subsubsection{Error calculation}\label{error-calculation}
+
+Residuals are essential for evaluating the accuracy of representing
+high-D points by the high-D mapping of 2D bin centroids. To measure this
+accuracy, an error metric is introduced, quantifying the sum of squared
+differences between the high-D data (\(x_{ij}\)) and the high-D mapping
+of the 2D bin centroid data (\(C_{x_ij}\)) across all observations and
+dimensions (see Equation~\ref{eq-equation10}).
+
+\begin{equation}\phantomsection\label{eq-equation10}{
+\text{Error} = \sum_{j = 1}^{n}\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2
+}\end{equation}
+
+Here, \(n\) represents the number of observations, \(p\) represents the
+dimensions of high-D data, \(x_{ij}\) is the high-D data, and
+\(C_{x_ij}\) is the high-D mapping of the 2D bin centroid.
+
+To assess how well our method captures and represents the underlying
+structure of the high-D data, Mean Squared Error (MSE) is used. When
+computing MSE, total model error (see \textbf{?@sec-summary}) is divided
+by the number of observations to make it as a mean value (see
+Equation~\ref{eq-equation9}).
+
+\begin{equation}\phantomsection\label{eq-equation9}{
+\text{MSE} = \sum_{j = 1}^{n} \frac{\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2}{n}
+}\end{equation}
+
 \subsection{Tuning}\label{tuning}
@@
 
-\begin{itemize}
-\tightlist
-\item
-  NLDR is used to illustrate different ways 1's are drawn
-\item
-  Use our method to assess is it a reasonable representation
-\item
-  Demonstrate that it is, except for the anomalies
-\end{itemize}
+The MNIST dataset consists of grayscale images of handwritten digits
+\citep{lecun2010}. \citet{yingfan2021} used this dataset to demonstrate
+how PaCMAP preserves local structure. We selected the \gD{} embedding of
+PaCMAP for the handwritten digit 1 to assess whether this is a
+reasonable representation using our method. As shown in
+Figure~\ref{fig-pacmap-author-img}, the angle of the digit 1 images
+varies along the \gD{} structure.
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-pacmapauthor-1.pdf}
+\includegraphics[width=1\textwidth,height=0.22\textheight]{paper_files/figure-pdf/fig-pacmap-author-1.pdf}
 
@@
 
-\caption{\label{fig-pacmapauthor}(a) 2D layout from PaCMAP applied for
-the digit 1 of the MNIST dataset. Is this a best representation of the
-original data?, and (b) Images of the handwritten digit 1. The angle of
-the digit 1 varies along non-linear structure.}
+\caption{\label{fig-pacmap-author}\(2\text{-}D\) layout from PaCMAP
+applied for the digit 1 of the MNIST dataset. Is this the best
+representation of the digit 1? The parameter setting is n\_components=2,
+n\_neighbors=10, init=random, MN\_ratio=0.9, FP\_ratio=2.0.}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=0.5\textheight]{paper_files/figure-pdf/fig-pacmap-author-img-1.pdf}
+
+}
+
+\caption{\label{fig-pacmap-author-img}Images of the handwritten digit 1
+are ordered from the bottom-right to the top-left of the \(2\text{-}D\)
+structure. The angle of the digit varies along this structure. Images at
+the bottom-right of the \(2\text{-}D\) layout show the digit 1 angled
+more to the right, while images at the top-left show the digit 1 angled
+more to the left. This demonstrates how the angle changes from right to
+left along the \(2\text{-}D\) structure.}
 
@@
 
-\caption{\label{fig-model-mnist}(a) Model generated in the 2D space
-overlaid on PaCMAP data, and (b) high-D model error in model space.}
+\caption{\label{fig-model-mnist}(a) Model generated in \(2\text{-}D\),
+and (b) \(p\text{-}D\) model error in \(2\text{-}D\). The \(2\text{-}D\)
+model shows a non-linear continuous structure. Most low \(p\text{-}D\)
+model errors are distributed along the lower edge of the \(2\text{-}D\)
+structure, while most high p-D model errors are concentrated along the
+upper edge.}
 
 \end{figure}%
+
+According to Figure~\ref{fig-mnist1_sc1}, the non-linear continuous
+structure observed in the \gD{} representation of PaCMAP (see
+Figure~\ref{fig-pacmap-author}) is also visible when visualizing the
+model overlaid on the data space. This indicates that PaCMAP accurately
+captures the structure of the \pD{} data. Additionally, the model shows
+a twisted pattern within the non-linear structure in \pD{} space (see
+Figure~\ref{fig-mnist1_sc2}), which is an additional pattern not visible
+in the \gD{} representation (see Figure~\ref{fig-pacmap-author}).
+Furthermore, as shown in Figure~\ref{fig-mnist1_sc3}, some long edges
+exist in the \pD{} space that are not recognized as long edges in the
+\gD{} representation. However, PaCMAP is a \emph{good} \gD{}
+representation of MNIST digit 1 data.
+
+\begin{figure}[H]
+
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/mnist/sc_1.png}
+
+}
+
+\subcaption{\label{fig-mnist1_sc1}}
+
+\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/mnist/sc_2.png}
+
+}
+
+\subcaption{\label{fig-mnist1_sc2}}
+
+\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/mnist/sc_3.png}
+
+}
+
+\subcaption{\label{fig-mnist1_sc3}}
+
+\end{minipage}%
+
+\caption{\label{fig-mnist1_sc}Screen shots of the \textbf{langevitour}
+of the MNIST digit 1 data set, shows the model-in-data space, a video of
+the tour animation is available at
+(\url{https://youtu.be/zq2GM9qvUNA}).}
+
+\end{figure}%
+
+There are certain data points that exhibit high error rates due to their
+deviation from the usual \pD{} data structure, which makes them
+anomalies (see Figure~\ref{fig-model-mnist} (b)). These anomalies can be
+classified into two types: those that are anomalies within the
+non-linear structure and those that lie outside of it. The images
+associated with high model error points within the non-linear structure
+display different patterns of the digit 1, as shown in
+Figure~\ref{fig-mnist-anomalies} (a). However, when comparing these
+images to the ones found outside of the non-linear structure, it becomes
+evident that the latter display different patterns of the digit 1 (see
+Figure~\ref{fig-mnist-anomalies} (b)).
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-anomalies-1.pdf}
+\includegraphics[width=1\textwidth,height=0.35\textheight]{paper_files/figure-pdf/fig-mnist-anomalies-1.pdf}
 
@@
 
-\caption{\label{fig-anomalies}(a) Images of handwritten digit 1 which
-occur large model error within the non-linear strcuture, and (b) Images
-of handwritten digit 1 which occur large model error outside the
-non-linear structure.}
+\caption{\label{fig-mnist-anomalies}Some images of handwritten digit 1
+which occur high model error (a) within the non-linear strcuture, and
+(b) outside the non-linear structure. The images shows different
+patterns of digit 1.}
 

### 2024-05-19 supplementary/hexagon_binning_algorithm.qmd
Update with US english
Added MNIST digit description, need to add the video

@@
 $$
+
 Also, substituting $a_2 = \frac{\sqrt{3}}{2}a_1$, $s_2 = -q \times r_2$ and rearranging @eq-equation5 gives:
@@
 
-Similarly, @eq-equation2 becomes,
+Similarly, substituting $s_1 = -q$ @eq-equation2 becomes,
 
@@
 
-This is a linear optimisation problem. Therefore, the optimal solution must occur on a vertex. So, by setting @eq-equation6 equals to @eq-equation7 gives,
+This is a linear optimization problem. Therefore, the optimal solution must occur on a vertex. So, by setting @eq-equation6 equals to @eq-equation7 gives,
 
@@
 ```
+
 ```{r}

### 2024-05-19 paper.qmd
Added intial analysis for PBMC

@@
     annotate("text", x=-0.15, y=0.45, label= "a2", color = "black") +
-    annotate("text", x=-0.15, y=-0.35, label= "(s1, s2)", color = "black") ## starting point
+    annotate("text", x=-0.15, y=-0.35, label= "(s1, s2)", color = "black") + ## starting point
+  coord_fixed()
 ```
@@
 #| fig-cap: "Notations for hexagonal grid configurations."
-#| fig-width: 3
-#| fig-height: 3
+#| out-height: 30%
 #| fig-pos: H
@@
 ### pbmc
-
+<!--
 - NLDR view used to illustrate clusters
@@
 - Illustrate how to use our method to get a better representation
+-->
+
+In the field of single-cell studies, a common analytical task involves clustering to identify groups of cells with similar expression profiles.
+Analysts often turn to NLDR techniques to verify and identify these clusters and explore developmental trajectories.
+To illustrate the importance of NLDR techniques and parameter selection in identifying clusters, Human Peripheral Blood Mononclear Cells (PBMC3k) dataset [@chen2023] is used.
+In a study by @chen2023, this dataset was used to demonstrate how UMAP represents clusters (see @fig-umap-author).
+As shown in @fig-umap-author, there are three distinct and well-separated clusters.
 
@@
 
-<!-- UMAP parameters suggested by the author(n_neighbors = 30, min_dist = 0.3)-->
-```{r}
-#| fig-cap: "2D layout from UMAP applied for the PBMC3k dataset. Is this a best representation of the original data?"
-#| label: fig-umapauthor
-#| fig-width: 3
-#| fig-height: 3
+<!-- UMAP parameters suggested by the author (n_neighbors = 30, min_dist = 0.3)-->
+```{r}
+#| fig-cap: "$2\\text{-}D$ layout from UMAP applied for the PBMC3k dataset. Is this a best representation of the original data? The parameter setting is n_neighbors = 30, min_dist = 0.3."
+#| label: fig-umap-author
+#| out-height: 30%
 #| fig-pos: H
@@
 ```
+
+To determine whether the UMAP representation with the parameter choice suggested by @chen2023 preserves the original data structure, we visualize the model constructed with UMAP overlaid on the \pD{} data.
+The figures in @fig-pbmc1_sc show three well-separated clusters, indicating that the suggested UMAP representation preserves the global structure (see @fig-umap-author).
+However, as shown in @fig-pbmc1_sc, these clusters are close to each other in \pD{}.
+Also, non-linear continuity patterns and high-density patches within the clusters are observed (see @fig-pbmc1_sc).
+Therefore, the suggested UMAP representation (see @fig-umap-author) does not accurately preserve the local structure of the PBMC3k dataset.
 
@@
 #| echo: false
-#| fig-cap: "(a) Model generated in the 2D space overlaid on UMAP data, and (b) high-D model error in model space."
+#| fig-cap: "(a) Model generated in $2\\text{-}D$ with UMAP, and (b) $p\\text{-}D$ model error in $2\\text{-}D$.
+The $2\\text{-}D$ model shows three well-separated distinct clusters.
+The $p\\text{-}D$ model errors are distributed along clusters."
 #| label: fig-model-pbmc-author
@@
 ```
+
+::: {#fig-pbmc1_sc layout-ncol="3" fig-pos="H"}
+![](figures/pbmc3k/sc_1.png){#fig-pbmc1_sc1}
+
+![](figures/pbmc3k/sc_2.png){#fig-pbmc1_sc2}
+
+![](figures/pbmc3k/sc_3.png){#fig-pbmc1_sc3}
+
+Screen shots of the **langevitour** of the PBMC3k data set, shows the model-in-data space, a video of the tour animation is available at (<https://youtu.be/Gnhh4hfsbbg>).
+:::
+
+In order to find a reasonable NLDR representation for the PBMC3k dataset, we calculated the absolute error for different numbers of non-empty bins using various NLDR techniques and different parameter settings (see @fig-pbmc-abserror).
+After analyzing the results, we found that tSNE with a perplexity set to $30$ had the lowest error when the number of non-empty bins was $137$.
+Therefore, tSNE with a perplexity of $30$, which is the default parameter setting, is considered as a reasonable representation for the PBMC3k dataset.
 
@@
                                   colour = type)) + 
-  geom_point() +
+  geom_point(size = 1) +
   geom_line() + 
   geom_vline(xintercept = 137, linetype="solid", 
-             color = "black", size=0.8, alpha = 0.5) +
+             color = "black", linewidth=0.8, alpha = 0.5) +
   scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a','#984ea3',
@@
   ylab("log(absolute error)") +
-  xlab("number of non-empty bins")
+  xlab("number of non-empty bins") +
+  theme(axis.text.x = element_text(size = 5),
+        axis.text.y = element_text(size = 5),
+        axis.title.x = element_text(size = 5),
+        axis.title.y = element_text(size = 5, angle = 90))
 
@@
 Then, the slope gradually declines or level off, indicating that a higher number of non-empty bins generates a smaller error.
-Using the elbow method, it was observed that when the number of non-empty bins is set to 137, the lowest error occurred with the parameters perplexity: 30."
-#| label: fig-abserrorPBMC
+Using the elbow method, it was observed that when the number of non-empty bins is set to $137$, the lowest error occurred with the parameters perplexity: $30$."
+#| label: fig-pbmc-abserror
 #| fig-pos: H
-#| out-width: 100%
+#| out-width: 50%
 
@@
 ```
+
+As shown in @fig-tsne-suggest, there are three well-separated clusters, although they are located close to each other.
+Additionally, non-linear structures can also be observed within the clusters (see @fig-model-pbmc (a)).
+In this manner, tSNE was able to capture the data structure for the PBMC3k dataset that UMAP failed to do.
 
@@
 ```
+
+<!--tSNE with perplexity:30-->
+```{r}
+#| echo: false
+#| fig-cap: "$2\\text{-}D$ layout from tSNE applied for the PBMC3k dataset.  Is this a best representation of the original data? The parameter setting is perplexity=30."
+#| label: fig-tsne-suggest
+#| fig-pos: H
+#| out-width: 40%
+
+tsne_pbmc 
+```
+
+We then fit the model for tSNE, and visualize the resultant model in the \pD{} data space.
+The model shows a quirk, as shown in @fig-pbmc2_sc.
+All three clusters are connected by an edge except the small and large clusters.
+Because the clusters are so close in \gD{}, they attempt to maintain the structure in \pD{} as well.
+This is evident that tSNE with perplexity $30$ provides a reasonable representation of PBMC3k data.
 
@@
 #| echo: false
-#| fig-cap: "(a) Model generated in the 2D space overlaid on tSNE data, and (b) high-D model error in model space."
+#| fig-cap: "(a) Model generated in $2\\text{-}D$ with tSNE, and (b) $p\\text{-}D$ model error in $2\\text{-}D$.
+The $2\\text{-}D$ model shows three well-separated distinct clusters.
+The $p\\text{-}D$ model errors are distributed along clusters, but most low $p\\text{-}D$ model errors present in the large cluster."
 #| label: fig-model-pbmc
 #| fig-pos: H
+#| out-height: 30%
 

### 2024-05-19 paper.tex
Added intial analysis for PBMC

@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-hex-param-1.pdf}
+\includegraphics[width=1\textwidth,height=0.3\textheight]{paper_files/figure-pdf/fig-hex-param-1.pdf}
 
@@
 
-\begin{itemize}
-\tightlist
-\item
-  NLDR view used to illustrate clusters
-\item
-  Use our method to assess is it a reasonable representation
-\item
-  Demonstrate that it is not
-\item
-  Illustrate how to use our method to get a better representation
-\end{itemize}
+In the field of single-cell studies, a common analytical task involves
+clustering to identify groups of cells with similar expression profiles.
+Analysts often turn to NLDR techniques to verify and identify these
+clusters and explore developmental trajectories. To illustrate the
+importance of NLDR techniques and parameter selection in identifying
+clusters, Human Peripheral Blood Mononclear Cells (PBMC3k) dataset
+\citep{chen2023} is used. In a study by \citet{chen2023}, this dataset
+was used to demonstrate how UMAP represents clusters (see
+Figure~\ref{fig-umap-author}). As shown in Figure~\ref{fig-umap-author},
+there are three distinct and well-separated clusters.
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-umapauthor-1.pdf}
+\includegraphics[width=1\textwidth,height=0.3\textheight]{paper_files/figure-pdf/fig-umap-author-1.pdf}
 
@@
 
-\caption{\label{fig-umapauthor}2D layout from UMAP applied for the
-PBMC3k dataset. Is this a best representation of the original data?}
+\caption{\label{fig-umap-author}\(2\text{-}D\) layout from UMAP applied
+for the PBMC3k dataset. Is this a best representation of the original
+data? The parameter setting is n\_neighbors = 30, min\_dist = 0.3.}
 
 \end{figure}%
+
+To determine whether the UMAP representation with the parameter choice
+suggested by \citet{chen2023} preserves the original data structure, we
+visualize the model constructed with UMAP overlaid on the \pD{} data.
+The figures in Figure~\ref{fig-pbmc1_sc} show three well-separated
+clusters, indicating that the suggested UMAP representation preserves
+the global structure (see Figure~\ref{fig-umap-author}). However, as
+shown in Figure~\ref{fig-pbmc1_sc}, these clusters are close to each
+other in \pD{}. Also, non-linear continuity patterns and high-density
+patches within the clusters are observed (see
+Figure~\ref{fig-pbmc1_sc}). Therefore, the suggested UMAP representation
+(see Figure~\ref{fig-umap-author}) does not accurately preserve the
+local structure of the PBMC3k dataset.
 
@@
 
-\caption{\label{fig-model-pbmc-author}(a) Model generated in the 2D
-space overlaid on UMAP data, and (b) high-D model error in model space.}
+\caption{\label{fig-model-pbmc-author}(a) Model generated in
+\(2\text{-}D\) with UMAP, and (b) \(p\text{-}D\) model error in
+\(2\text{-}D\). The \(2\text{-}D\) model shows three well-separated
+distinct clusters. The \(p\text{-}D\) model errors are distributed along
+clusters.}
 
@@
 
+\begin{minipage}{0.33\linewidth}
+
 \centering{
 
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-abserrorPBMC-1.pdf}
+\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_1.png}
 
@@
 
-\caption{\label{fig-abserrorPBMC}Absolute error from UMAP and tSNE
+\subcaption{\label{fig-pbmc1_sc1}}
+
+\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_2.png}
+
+}
+
+\subcaption{\label{fig-pbmc1_sc2}}
+
+\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_3.png}
+
+}
+
+\subcaption{\label{fig-pbmc1_sc3}}
+
+\end{minipage}%
+
+\caption{\label{fig-pbmc1_sc}Screen shots of the \textbf{langevitour} of
+the PBMC3k data set, shows the model-in-data space, a video of the tour
+animation is available at (\url{https://youtu.be/Gnhh4hfsbbg}).}
+
+\end{figure}%
+
+In order to find a reasonable NLDR representation for the PBMC3k
+dataset, we calculated the absolute error for different numbers of
+non-empty bins using various NLDR techniques and different parameter
+settings (see Figure~\ref{fig-pbmc-abserror}). After analyzing the
+results, we found that tSNE with a perplexity set to \(30\) had the
+lowest error when the number of non-empty bins was \(137\). Therefore,
+tSNE with a perplexity of \(30\), which is the default parameter
+setting, is considered as a reasonable representation for the PBMC3k
+dataset.
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=0.5\textwidth,height=\textheight]{paper_files/figure-pdf/fig-pbmc-abserror-1.pdf}
+
+}
+
+\caption{\label{fig-pbmc-abserror}Absolute error from UMAP and tSNE
 applied to training PBMC3k dataset with diffferent parameter choices.
@@
 non-empty bins generates a smaller error. Using the elbow method, it was
-observed that when the number of non-empty bins is set to 137, the
-lowest error occurred with the parameters perplexity: 30.}
+observed that when the number of non-empty bins is set to \(137\), the
+lowest error occurred with the parameters perplexity: \(30\).}
+
+\end{figure}%
+
+As shown in Figure~\ref{fig-tsne-suggest}, there are three
+well-separated clusters, although they are located close to each other.
+Additionally, non-linear structures can also be observed within the
+clusters (see Figure~\ref{fig-model-pbmc} (a)). In this manner, tSNE was
+able to capture the data structure for the PBMC3k dataset that UMAP
+failed to do.
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=0.4\textwidth,height=\textheight]{paper_files/figure-pdf/fig-tsne-suggest-1.pdf}
+
+}
+
+\caption{\label{fig-tsne-suggest}\(2\text{-}D\) layout from tSNE applied
+for the PBMC3k dataset. Is this a best representation of the original
+data? The parameter setting is perplexity=30.}
+
+\end{figure}%
+
+We then fit the model for tSNE, and visualize the resultant model in the
+\pD{} data space. The model shows a quirk, as shown in
+Figure~\ref{fig-pbmc2_sc}. All three clusters are connected by an edge
+except the small and large clusters. Because the clusters are so close
+in \gD{}, they attempt to maintain the structure in \pD{} as well. This
+is evident that tSNE with perplexity \(30\) provides a reasonable
+representation of PBMC3k data.
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=0.3\textheight]{paper_files/figure-pdf/fig-model-pbmc-1.pdf}
+
+}
+
+\caption{\label{fig-model-pbmc}(a) Model generated in \(2\text{-}D\)
+with tSNE, and (b) \(p\text{-}D\) model error in \(2\text{-}D\). The
+\(2\text{-}D\) model shows three well-separated distinct clusters. The
+\(p\text{-}D\) model errors are distributed along clusters, but most low
+\(p\text{-}D\) model errors present in the large cluster.}
 
@@
 
+\begin{minipage}{0.33\linewidth}
+
 \centering{
 
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-model-pbmc-1.pdf}
+\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_4.png}
 
@@
 
-\caption{\label{fig-model-pbmc}(a) Model generated in the 2D space
-overlaid on tSNE data, and (b) high-D model error in model space.}
+\subcaption{\label{fig-pbmc2_sc4}}
+
+\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_5.png}
+
+}
+
+\subcaption{\label{fig-pbmc2_sc5}}
+
+\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_6.png}
+
+}
+
+\subcaption{\label{fig-pbmc2_sc6}}
+
+\end{minipage}%
+
+\caption{\label{fig-pbmc2_sc}Screen shots of the \textbf{langevitour} of
+the PBMC3k data set, shows the model in high-D, a video of the tour
+animation is available at (\url{https://youtu.be/R0EroJcGbas}).}
 

### 2024-05-19 applications/pbmc3k/8_gen_model_in_data_space.R

@@ -1,25 +1,0 @@
-### Define type column
-df <- df_all_pbmc |>
-  dplyr::select(tidyselect::starts_with("PC")) |>
-  dplyr::mutate(type = "data") ## original dataset
-
-df_b <- df_bin_pbmc |>
-  dplyr::filter(hb_id %in% df_bin_centroids_pbmc$hexID) |>
-  dplyr::mutate(type = "model") ## Data with summarized mean
-
-## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
-df_b <- df_b[match(df_bin_centroids_pbmc$hexID, df_b$hb_id),] |>
-  dplyr::select(-hb_id)
-
-df_exe <- dplyr::bind_rows(df_b, df)
-
-## Set the maximum difference as the criteria
-distance_df_small_edges <- distance_pbmc |>
-  dplyr::filter(distance < benchmark_pbmc)
-## Since erase brushing is considerd.
-
-langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
-                         lineFrom = distance_df_small_edges$from,
-                         lineTo = distance_df_small_edges$to,
-                         group = df_exe$type, pointSize = append(rep(0, NROW(df_b)), rep(0.5, NROW(df))),
-                         levelColors = c("#6a3d9a", "#33a02c"))

### 2024-05-19 paper.qmd
Added intro for tuning

@@
 
-### Measuring the fit
+### Measuring the fit {#sec-summary}
  <!-- Fitted values,  Error calculation-->
@@
 ### Tuning
-
 <!-- removal of low density bins, removing long edges, choice of bins-->
+
+The performance and robustness of our model depend on three key parameters: (i) the total number of bins ($b$), (ii) a benchmark value used to remove low-density hexagons, and (iii) a benchmark value used to remove long edges.
+However, there is no analytical formula to calculate an appropriate value for these parameters.
+The selection of these parameter values depends on the model performance computed by MSE (see @sec-summary).
+
+#### Choice of bins
+
+The number of hexagonal bins in the hexagonal grid has a considerable impact on the construction of the \gD{} model, serving as the initial step.
+The chosen total number of bins must effectively capture the structure of the NLDR data.
+If the number of bins is too low, the model may not be able to capture the structure of the NLDR data effectively (see @fig-binsize (a)), while if there are too many bins, it may result in over-fitting the individual points of the NLDR data (see @fig-binsize (c)).
+Therefore, it is important to determine an appropriate number of bins to construct an effective model.
+
+<!--add three choice of bins-->
+
+Furthermore, the total number of bins is calculated based on the number of bins along the x-axis and y-axis, $b = b_1 \times b_2$.
+To determine the effective $b$, candidate values are selected based on the range between the minimum and approximate maximum $b_1$, because $b_2$ is computed from $b_1$.
+The minimum $b_1$ is set to $2$, while the maximum number is estimated by taking the square root of $n$.
+The analysis evaluates the MSE across varying $b$ within this range, covering the minimum to maximum values along both axes (see @fig-diagnostic1-scurve).
+
+<!--add MSE vs total number of error plot-->
+
+#### Removal of low density bins
+
+#### Removing long edges
 

### 2024-05-19 paper.tex
Added intro for tuning

@@
 
-\subsection{Measuring the fit}\label{measuring-the-fit}
+\subsection{Measuring the fit}\label{sec-summary}
 
@@
 structure of the high-D data, Mean Squared Error (MSE) is used. When
-computing MSE, total model error (see \textbf{?@sec-summary}) is divided
-by the number of observations to make it as a mean value (see
+computing MSE, total model error (see Section~\ref{sec-summary}) is
+divided by the number of observations to make it as a mean value (see
 Equation~\ref{eq-equation9}).
@@
 \subsection{Tuning}\label{tuning}
+
+The performance and robustness of our model depend on three key
+parameters: (i) the total number of bins (\(b\)), (ii) a benchmark value
+used to remove low-density hexagons, and (iii) a benchmark value used to
+remove long edges. However, there is no analytical formula to calculate
+an appropriate value for these parameters. The selection of these
+parameter values depends on the model performance computed by MSE (see
+Section~\ref{sec-summary}).
+
+\subsubsection{Choice of bins}\label{choice-of-bins}
+
+The number of hexagonal bins in the hexagonal grid has a considerable
+impact on the construction of the \gD{} model, serving as the initial
+step. The chosen total number of bins must effectively capture the
+structure of the NLDR data. If the number of bins is too low, the model
+may not be able to capture the structure of the NLDR data effectively
+(see \textbf{?@fig-binsize} (a)), while if there are too many bins, it
+may result in over-fitting the individual points of the NLDR data (see
+\textbf{?@fig-binsize} (c)). Therefore, it is important to determine an
+appropriate number of bins to construct an effective model.
+
+Furthermore, the total number of bins is calculated based on the number
+of bins along the x-axis and y-axis, \(b = b_1 \times b_2\). To
+determine the effective \(b\), candidate values are selected based on
+the range between the minimum and approximate maximum \(b_1\), because
+\(b_2\) is computed from \(b_1\). The minimum \(b_1\) is set to \(2\),
+while the maximum number is estimated by taking the square root of
+\(n\). The analysis evaluates the MSE across varying \(b\) within this
+range, covering the minimum to maximum values along both axes (see
+\textbf{?@fig-diagnostic1-scurve}).
+
+\subsubsection{Removal of low density
+bins}\label{removal-of-low-density-bins}
+
+\subsubsection{Removing long edges}\label{removing-long-edges}
 

### 2024-05-19 paper.qmd
Added draft texts for tuning

@@
 
+After establishing the hexagonal grid with an appropriate number of bins, some hexagonal bins may have few or no data points within them (see @fig-rmlowdenshex (a)).
+To ensure comprehensive coverage of the NLDR data, it is necessary to select hexagonal bins with a considerable number of data points.
+This involves calculating the number of points within each hexagon.
+Then, the standard count is computed by dividing the number of points within each hexagon by the maximum number of points in the grid (see @eq-equationp2).
+Next, bins with a standard count less than a certain benchmark value are removed (see @fig-rmlowdenshex (b)).
+The following steps will help in determining a suitable value for removing low-density hexagons:
+
+1. Plot the distribution of the standardized counts (see @fig-stdctsScurve).
+2. Examine the distribution of counts.
+3. Select the first quantile value if the distribution is skewed.
+
+Furthermore, selecting the benchmark value for removing low-density hexagons is important.
+Removing unnecessary bins may lead to the formation of long edges and an uneven 2D model.
+Hence, rather than solely relying on the benchmark value to identify hexagons for removal, it's essential to consider the standard number of points in the neighboring hexagons of the identified low-density ones (see @fig-rmlowdenshex (c)).
+If neighboring bins also show low counts, only those bins will be removed.
+The remaining bins are used to construct the 2D model.
+
+The benchmark value for removing low-density hexagons ranges between $0$ and $1$.
+When analyzing how these benchmark values influence model performance, it's essential to observe the change in MSE as the benchmark value increases (see @fig-diagnosticpltScurvelwd).
+The MSE shows a gradual increase as the benchmark value progresses from $0$ to $1$.
+Evaluating this rate of increase is important.
+If the increment is not considerable, the decision might lean towards retaining low-density hexagons.
+
 #### Removing long edges
+
+Creating a smooth 2D representation (see @fig-modelScurve (c)) requires removing edges that connect distant bin centroids within the triangular mesh (see @fig-modelScurve (b)).
+These edges only exist in the 2D model and do not extend into higher dimensions, ensuring that their removal does not impact the model in higher dimensions.
+While specific criteria for determining the benchmark value to remove long edges do not exist, the following steps provide an approach to identifying a suitable threshold:
+
+1. Plot the distribution of the 2D Euclidean distances (see @fig-distScurve).
+2. Identify the largest difference between consecutive distance values.
+3. Take the largest distance value corresponding to this difference as the benchmark value.
 

### 2024-05-19 paper.tex
Added draft texts for tuning

@@
 
+After establishing the hexagonal grid with an appropriate number of
+bins, some hexagonal bins may have few or no data points within them
+(see \textbf{?@fig-rmlowdenshex} (a)). To ensure comprehensive coverage
+of the NLDR data, it is necessary to select hexagonal bins with a
+considerable number of data points. This involves calculating the number
+of points within each hexagon. Then, the standard count is computed by
+dividing the number of points within each hexagon by the maximum number
+of points in the grid (see \textbf{?@eq-equationp2}). Next, bins with a
+standard count less than a certain benchmark value are removed (see
+\textbf{?@fig-rmlowdenshex} (b)). The following steps will help in
+determining a suitable value for removing low-density hexagons:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Plot the distribution of the standardized counts (see
+  \textbf{?@fig-stdctsScurve}).
+\item
+  Examine the distribution of counts.
+\item
+  Select the first quantile value if the distribution is skewed.
+\end{enumerate}
+
+Furthermore, selecting the benchmark value for removing low-density
+hexagons is important. Removing unnecessary bins may lead to the
+formation of long edges and an uneven 2D model. Hence, rather than
+solely relying on the benchmark value to identify hexagons for removal,
+it's essential to consider the standard number of points in the
+neighboring hexagons of the identified low-density ones (see
+\textbf{?@fig-rmlowdenshex} (c)). If neighboring bins also show low
+counts, only those bins will be removed. The remaining bins are used to
+construct the 2D model.
+
+The benchmark value for removing low-density hexagons ranges between
+\(0\) and \(1\). When analyzing how these benchmark values influence
+model performance, it's essential to observe the change in MSE as the
+benchmark value increases (see \textbf{?@fig-diagnosticpltScurvelwd}).
+The MSE shows a gradual increase as the benchmark value progresses from
+\(0\) to \(1\). Evaluating this rate of increase is important. If the
+increment is not considerable, the decision might lean towards retaining
+low-density hexagons.
+
 \subsubsection{Removing long edges}\label{removing-long-edges}
+
+Creating a smooth 2D representation (see \textbf{?@fig-modelScurve} (c))
+requires removing edges that connect distant bin centroids within the
+triangular mesh (see \textbf{?@fig-modelScurve} (b)). These edges only
+exist in the 2D model and do not extend into higher dimensions, ensuring
+that their removal does not impact the model in higher dimensions. While
+specific criteria for determining the benchmark value to remove long
+edges do not exist, the following steps provide an approach to
+identifying a suitable threshold:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Plot the distribution of the 2D Euclidean distances (see
+  \textbf{?@fig-distScurve}).
+\item
+  Identify the largest difference between consecutive distance values.
+\item
+  Take the largest distance value corresponding to this difference as
+  the benchmark value.
+\end{enumerate}
 

### 2024-05-19 paper.qmd
Added plan for vis for tuning
Added intro to simulated data example

@@
 
+<!--add distribution of density and add the first quartile as the benchmark-->
+
 Furthermore, selecting the benchmark value for removing low-density hexagons is important.
@@
 
+<!--add vis showing removing the hexagons using first quantile, and observing neighbors-->
+
 The benchmark value for removing low-density hexagons ranges between $0$ and $1$.
@@
 
+<!--add MSE vs density (0 to 1)-->
+
 #### Removing long edges
@@
 
+<!--add distribution of distance along with the default benchmark-->
+<!--show the trimesh with colored long edges-->
+
 ### Illustration on simulated data
+
+In this section, the effectiveness of the algorithm is described using a simulated dataset.
+The dataset consists of five spherical Gaussian clusters in $4\text{-}D$, with each cluster containing an equal number of points and the same within-cluster variation.
+
+In the 2D layouts generated by various NLDR techniques, as shown in @fig-NLDR-variety-gau, five well-separated clusters are shown.
+In tSNE (see @fig-NLDR-variety-gau (a)), these clusters appear closely.
+UMAP arranges all clusters in a parallel manner, with three aligned in one line and the other two in a separate line (see @fig-NLDR-variety-gau (b)).
+In contrast, PHATE shows two closely positioned clusters and three more distant ones (see @fig-NLDR-variety-gau (c)).
+In TriMAP, two clusters are close, though not as tightly as PHATE, while the other three are well-separated (see @fig-NLDR-variety-gau (d)).
+Finally, PaCMAP shows one central cluster and the remaining four spread out in different directions (see @fig-NLDR-variety-gau (e)).
+
+<!-- five Gaussian clusters with different NLDR techniques-->
+```{r}
+#| warning: false
+#| echo: false
+
+## Import data
+df_2 <- read_rds("data/five_gau_clusters/data_five_gau.rds")
+training_data_gau <- read_rds("data/five_gau_clusters/data_five_gau_training.rds")
+test_data_gau <- read_rds("data/five_gau_clusters/data_five_gau_test.rds")
+
+tsne_data_gau <- read_rds("data/five_gau_clusters/tsne_data_five_gau_61.rds")
+umap_data_gau <- read_rds("data/five_gau_clusters/umap_data_five_gau.rds")
+phate_data_gau <- read_rds("data/five_gau_clusters/phate_data_five_gau.rds")
+trimap_data_gau <- read_rds("data/five_gau_clusters/trimap_data_five_gau.rds")
+pacmap_data_gau <- read_rds("data/five_gau_clusters/pacmap_data_five_gau.rds")
+
+## Visualize embedding
+
+nldr_gau1 <- tsne_data_gau |>
+  ggplot(aes(x = tSNE1,
+             y = tSNE2)) +
+  geom_point(alpha=0.1, size=1) +
+  interior_annotation("a")
+
+nldr_gau2 <- umap_data_gau |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2)) +
+  geom_point(alpha=0.1, size=1) +
+  interior_annotation("b")
+
+nldr_gau3 <- phate_data_gau |>
+  ggplot(aes(x = PHATE1,
+             y = PHATE2)) +
+  geom_point(alpha=0.1, size=1) +
+  interior_annotation("c")
+
+nldr_gau4 <- pacmap_data_gau |>
+  ggplot(aes(x = PaCMAP1,
+             y = PaCMAP2)) +
+  geom_point(alpha=0.1, size=1) +
+  interior_annotation("d")
+
+nldr_gau5 <- trimap_data_gau |>
+  ggplot(aes(x = TriMAP1,
+             y = TriMAP2)) +
+  geom_point(alpha=0.1, size=1) +
+  interior_annotation("e")
+```
+
+```{r}
+#| label: fig-NLDR-variety-gau
+#| echo: false
+#| fig-cap: "Five different NLDR representations of the same data.
+Different techniques and different parameter choices are used.
+Is there a best representation of the original data or are they all providing  equivalent information?"
+#| fig-width: 8
+#| fig-height: 2
+#| out-width: 100%
+#| fig-pos: H
+# (a) tSNE (perplexity = 61), (b) UMAP (n_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2)
+nldr_gau1 + nldr_gau2 + 
+  nldr_gau3 + nldr_gau4 + nldr_gau5 + 
+  plot_layout(ncol = 5)
+```
+
+<!--add trimesh for all five-->
+<!--add three screenshots from each technique with youtube links-->
 

### 2024-05-19 paper.tex
Added intro to simulated data example

@@
 
+In this section, the effectiveness of the algorithm is described using a
+simulated dataset. The dataset consists of five spherical Gaussian
+clusters in \(4\text{-}D\), with each cluster containing an equal number
+of points and the same within-cluster variation.
+
+In the 2D layouts generated by various NLDR techniques, as shown in
+Figure~\ref{fig-NLDR-variety-gau}, five well-separated clusters are
+shown. In tSNE (see Figure~\ref{fig-NLDR-variety-gau} (a)), these
+clusters appear closely. UMAP arranges all clusters in a parallel
+manner, with three aligned in one line and the other two in a separate
+line (see Figure~\ref{fig-NLDR-variety-gau} (b)). In contrast, PHATE
+shows two closely positioned clusters and three more distant ones (see
+Figure~\ref{fig-NLDR-variety-gau} (c)). In TriMAP, two clusters are
+close, though not as tightly as PHATE, while the other three are
+well-separated (see Figure~\ref{fig-NLDR-variety-gau} (d)). Finally,
+PaCMAP shows one central cluster and the remaining four spread out in
+different directions (see Figure~\ref{fig-NLDR-variety-gau} (e)).
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-NLDR-variety-gau-1.pdf}
+
+}
+
+\caption{\label{fig-NLDR-variety-gau}Five different NLDR representations
+of the same data. Different techniques and different parameter choices
+are used. Is there a best representation of the original data or are
+they all providing equivalent information?}
+
+\end{figure}%
+
 \section{Applications}\label{sec-applications}

### 2024-05-20 paper.qmd
Add draft for simulated data

@@
 ## Import data
-df_2 <- read_rds("data/five_gau_clusters/data_five_gau.rds")
 training_data_gau <- read_rds("data/five_gau_clusters/data_five_gau_training.rds")
-test_data_gau <- read_rds("data/five_gau_clusters/data_five_gau_test.rds")
 
@@
 
+To investigate which is the reasonable representation to visualize the five spherical Gaussian cluster data or all NLDR methods provide equivalent information, we visualize all the models in \pD{} space.
+Models from tSNE, UMAP, and PHATE show five well-separated clusters (see @fig-gau1_sc, @fig-gau2_sc, and @fig-gau3_sc), while PaCMAP and TriMAP have a long edge that connects neighboring two clusters and three well-separated clusters (see @fig-gau4_sc, and @fig-gau5_sc).
+This suggests that for the five Gaussian cluster dataset, all NLDR methods effectively preserve the global structure.
+
+Models from all NLDR methods show five well-separated clusters (see @fig-gau1_sc, @fig-gau2_sc, @fig-gau3_sc, @fig-gau4_sc, and @fig-gau5_sc).
+This suggests that for the five Gaussian cluster dataset, all NLDR methods effectively preserve the global structure.
+tSNE, UMAP, and PHATE display clusters with varying densities, indicating their ability to capture within-cluster variation (see @fig-gau1_sc, @fig-gau2_sc, and @fig-gau3_sc).
+On the other hand, both TriMAP and PaCMAP show clusters with flat surfaces, suggesting a failure to capture within-cluster variation (see @fig-gau4_sc and @fig-gau5_sc).
+Therefore, TriMAP and PaCMAP do not capture the local structure as effectively as other methods.
+
+When compare the NLDR representations and generates models, tSNE with perplexity $61$ is the reasonable representation to visualize the five Gaussian cluster dataset.
+
+```{r}
+sc_xlims_gau <- c(-0.25, 1.25)
+sc_ylims_gau <- c(-0.2, 1.2)
+```
+
+<!--trimesh with tsne-->
+```{r}
+gau1_scaled_obj <- gen_scaled_data(
+  data = tsne_data_gau)
+tsne_gau_scaled <- gau1_scaled_obj$scaled_nldr
+
+## Compute hexbin parameters
+num_bins_x_gau1 <- 12
+lim1 <- gau1_scaled_obj$lim1
+lim2 <- gau1_scaled_obj$lim2
+r2_gau1 <- diff(lim2)/diff(lim1)
+
+gau1_model <- fit_highd_model(
+  training_data = training_data_gau,
+  emb_df = tsne_gau_scaled,
+  bin1 = num_bins_x_gau1,
+  r2 = r2_gau1,
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = FALSE,
+  col_start_highd = "x"
+)
+
+df_bin_centroids_gau1 <- gau1_model$df_bin_centroids
+df_bin_gau1 <- gau1_model$df_bin
+
+## Triangulate bin centroids
+tr1_object_gau1 <- tri_bin_centroids(
+  df_bin_centroids_gau1, x = "c_x", y = "c_y")
+tr_from_to_df_gau1 <- gen_edges(
+  tri_object = tr1_object_gau1)
+
+## Compute 2D distances
+distance_gau1 <- cal_2d_dist(
+  tr_coord_df = tr_from_to_df_gau1,
+  start_x = "x_from",
+  start_y = "y_from",
+  end_x = "x_to",
+  end_y = "y_to",
+  select_vars = c("from", "to", "distance"))
+
+## To find the benchmark value
+benchmark_gau1 <- find_lg_benchmark(
+  distance_edges = distance_gau1,
+  distance_col = "distance")
+
+trimesh_removed_gau1 <- vis_rmlg_mesh(
+  distance_edges = distance_gau1, 
+  benchmark_value = benchmark_gau1, 
+  tr_coord_df = tr_from_to_df_gau1, 
+  distance_col = "distance") +
+  xlim(sc_xlims_gau) + ylim(sc_ylims_gau) +
+  interior_annotation("a", sc_ltr_pos) 
+```
+
+<!--trimesh with umap-->
+```{r}
+gau2_scaled_obj <- gen_scaled_data(
+  data = umap_data_gau)
+umap_gau_scaled <- gau2_scaled_obj$scaled_nldr
+
+## Compute hexbin parameters
+num_bins_x_gau2 <- 44
+lim1 <- gau2_scaled_obj$lim1
+lim2 <- gau2_scaled_obj$lim2
+r2_gau2 <- diff(lim2)/diff(lim1)
+
+gau2_model <- fit_highd_model(
+  training_data = training_data_gau,
+  emb_df = umap_gau_scaled,
+  bin1 = num_bins_x_gau2,
+  r2 = r2_gau2,
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = FALSE,
+  col_start_highd = "x"
+)
+
+df_bin_centroids_gau2 <- gau2_model$df_bin_centroids
+df_bin_gau2 <- gau2_model$df_bin
+
+## Triangulate bin centroids
+tr1_object_gau2 <- tri_bin_centroids(
+  df_bin_centroids_gau2, x = "c_x", y = "c_y")
+tr_from_to_df_gau2 <- gen_edges(
+  tri_object = tr1_object_gau2)
+
+## Compute 2D distances
+distance_gau2 <- cal_2d_dist(
+  tr_coord_df = tr_from_to_df_gau2,
+  start_x = "x_from",
+  start_y = "y_from",
+  end_x = "x_to",
+  end_y = "y_to",
+  select_vars = c("from", "to", "distance"))
+
+## To find the benchmark value
+benchmark_gau2 <- find_lg_benchmark(
+  distance_edges = distance_gau2,
+  distance_col = "distance")
+
+trimesh_removed_gau2 <- vis_rmlg_mesh(
+  distance_edges = distance_gau2, 
+  benchmark_value = benchmark_gau2, 
+  tr_coord_df = tr_from_to_df_gau2, 
+  distance_col = "distance") +
+  xlim(sc_xlims_gau) + ylim(sc_ylims_gau) +
+  interior_annotation("b", sc_ltr_pos) 
+```
+
+<!--trimesh with phate-->
+```{r}
+gau3_scaled_obj <- gen_scaled_data(
+  data = phate_data_gau)
+phate_gau_scaled <- gau3_scaled_obj$scaled_nldr
+
+## Compute hexbin parameters
+num_bins_x_gau3 <- 250
+lim1 <- gau3_scaled_obj$lim1
+lim2 <- gau3_scaled_obj$lim2
+r2_gau3 <- diff(lim2)/diff(lim1)
+
+gau3_model <- fit_highd_model(
+  training_data = training_data_gau,
+  emb_df = phate_gau_scaled,
+  bin1 = num_bins_x_gau3,
+  r2 = r2_gau3,
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = FALSE,
+  col_start_highd = "x"
+)
+
+df_bin_centroids_gau3 <- gau3_model$df_bin_centroids
+df_bin_gau3 <- gau3_model$df_bin
+
+## Triangulate bin centroids
+tr1_object_gau3 <- tri_bin_centroids(
+  df_bin_centroids_gau3, x = "c_x", y = "c_y")
+tr_from_to_df_gau3 <- gen_edges(
+  tri_object = tr1_object_gau3)
+
+## Compute 2D distances
+distance_gau3 <- cal_2d_dist(
+  tr_coord_df = tr_from_to_df_gau3,
+  start_x = "x_from",
+  start_y = "y_from",
+  end_x = "x_to",
+  end_y = "y_to",
+  select_vars = c("from", "to", "distance"))
+
+## To find the benchmark value
+benchmark_gau3 <- find_lg_benchmark(
+  distance_edges = distance_gau3,
+  distance_col = "distance")
+
+trimesh_removed_gau3 <- vis_rmlg_mesh(
+  distance_edges = distance_gau3, 
+  benchmark_value = benchmark_gau3, 
+  tr_coord_df = tr_from_to_df_gau3, 
+  distance_col = "distance") +
+  xlim(sc_xlims_gau) + ylim(sc_ylims_gau) +
+  interior_annotation("c", sc_ltr_pos) 
+```
+
+<!--trimesh with pacmap-->
+```{r}
+gau4_scaled_obj <- gen_scaled_data(
+  data = umap_data_gau)
+pacmap_gau_scaled <- gau4_scaled_obj$scaled_nldr
+
+## Compute hexbin parameters
+num_bins_x_gau4 <- 18
+lim1 <- gau4_scaled_obj$lim1
+lim2 <- gau4_scaled_obj$lim2
+r2_gau4 <- diff(lim2)/diff(lim1)
+
+gau4_model <- fit_highd_model(
+  training_data = training_data_gau,
+  emb_df = pacmap_gau_scaled,
+  bin1 = num_bins_x_gau4,
+  r2 = r2_gau4,
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = FALSE,
+  col_start_highd = "x"
+)
+
+df_bin_centroids_gau4 <- gau4_model$df_bin_centroids
+df_bin_gau4 <- gau4_model$df_bin
+
+## Triangulate bin centroids
+tr1_object_gau4 <- tri_bin_centroids(
+  df_bin_centroids_gau4, x = "c_x", y = "c_y")
+tr_from_to_df_gau4 <- gen_edges(
+  tri_object = tr1_object_gau4)
+
+## Compute 2D distances
+distance_gau4 <- cal_2d_dist(
+  tr_coord_df = tr_from_to_df_gau4,
+  start_x = "x_from",
+  start_y = "y_from",
+  end_x = "x_to",
+  end_y = "y_to",
+  select_vars = c("from", "to", "distance"))
+
+## To find the benchmark value
+benchmark_gau4 <- find_lg_benchmark(
+  distance_edges = distance_gau4,
+  distance_col = "distance")
+
+trimesh_removed_gau4 <- vis_rmlg_mesh(
+  distance_edges = distance_gau4, 
+  benchmark_value = benchmark_gau4, 
+  tr_coord_df = tr_from_to_df_gau4, 
+  distance_col = "distance") +
+  xlim(sc_xlims_gau) + ylim(sc_ylims_gau) +
+  interior_annotation("d", sc_ltr_pos) 
+```
+
+<!--trimesh with trimap-->
+```{r}
+gau5_scaled_obj <- gen_scaled_data(
+  data = trimap_data_gau)
+trimap_gau_scaled <- gau5_scaled_obj$scaled_nldr
+
+## Compute hexbin parameters
+num_bins_x_gau5 <- 25
+lim1 <- gau5_scaled_obj$lim1
+lim2 <- gau5_scaled_obj$lim2
+r2_gau5 <- diff(lim2)/diff(lim1)
+
+gau5_model <- fit_highd_model(
+  training_data = training_data_gau,
+  emb_df = trimap_gau_scaled,
+  bin1 = num_bins_x_gau5,
+  r2 = r2_gau5,
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = FALSE,
+  col_start_highd = "x"
+)
+
+df_bin_centroids_gau5 <- gau5_model$df_bin_centroids
+df_bin_gau5 <- gau5_model$df_bin
+
+## Triangulate bin centroids
+tr1_object_gau5 <- tri_bin_centroids(
+  df_bin_centroids_gau5, x = "c_x", y = "c_y")
+tr_from_to_df_gau5 <- gen_edges(
+  tri_object = tr1_object_gau5)
+
+## Compute 2D distances
+distance_gau5 <- cal_2d_dist(
+  tr_coord_df = tr_from_to_df_gau5,
+  start_x = "x_from",
+  start_y = "y_from",
+  end_x = "x_to",
+  end_y = "y_to",
+  select_vars = c("from", "to", "distance"))
+
+## To find the benchmark value
+benchmark_gau5 <- find_lg_benchmark(
+  distance_edges = distance_gau5,
+  distance_col = "distance")
+
+trimesh_removed_gau5 <- vis_rmlg_mesh(
+  distance_edges = distance_gau5, 
+  benchmark_value = benchmark_gau5, 
+  tr_coord_df = tr_from_to_df_gau5, 
+  distance_col = "distance") +
+  xlim(sc_xlims_gau) + ylim(sc_ylims_gau) +
+  interior_annotation("e", sc_ltr_pos) 
+```
+
 <!--add trimesh for all five-->
-<!--add three screenshots from each technique with youtube links-->
+
+```{r}
+#| label: fig-trimesh-gau
+#| echo: false
+#| fig-cap: "Model generated with five diiferent NLDR methods in $2\\text{-}D$ with approximately $65$ non-empty bins."
+#| fig-width: 8
+#| fig-height: 2
+#| out-width: 100%
+#| fig-pos: H
+
+trimesh_removed_gau1 + trimesh_removed_gau2 +
+  trimesh_removed_gau3 + trimesh_removed_gau4 + trimesh_removed_gau5 +
+  plot_layout(ncol = 5)
+```
+
+<!--add three screenshots from each technique with you tube links-->
 

### 2024-05-20 paper.tex
Add draft for simulated data

@@
 
+To investigate which is the reasonable representation to visualize the
+five spherical Gaussian cluster data or all NLDR methods provide
+equivalent information, we visualize all the models in \pD{} space.
+Models from tSNE, UMAP, and PHATE show five well-separated clusters (see
+\textbf{?@fig-gau1\_sc}, \textbf{?@fig-gau2\_sc}, and
+\textbf{?@fig-gau3\_sc}), while PaCMAP and TriMAP have a long edge that
+connects neighboring two clusters and three well-separated clusters (see
+\textbf{?@fig-gau4\_sc}, and \textbf{?@fig-gau5\_sc}). This suggests
+that for the five Gaussian cluster dataset, all NLDR methods effectively
+preserve the global structure.
+
+Models from all NLDR methods show five well-separated clusters (see
+\textbf{?@fig-gau1\_sc}, \textbf{?@fig-gau2\_sc},
+\textbf{?@fig-gau3\_sc}, \textbf{?@fig-gau4\_sc}, and
+\textbf{?@fig-gau5\_sc}). This suggests that for the five Gaussian
+cluster dataset, all NLDR methods effectively preserve the global
+structure. tSNE, UMAP, and PHATE display clusters with varying
+densities, indicating their ability to capture within-cluster variation
+(see \textbf{?@fig-gau1\_sc}, \textbf{?@fig-gau2\_sc}, and
+\textbf{?@fig-gau3\_sc}). On the other hand, both TriMAP and PaCMAP show
+clusters with flat surfaces, suggesting a failure to capture
+within-cluster variation (see \textbf{?@fig-gau4\_sc} and
+\textbf{?@fig-gau5\_sc}). Therefore, TriMAP and PaCMAP do not capture
+the local structure as effectively as other methods.
+
+When compare the NLDR representations and generates models, tSNE with
+perplexity \(61\) is the reasonable representation to visualize the five
+Gaussian cluster dataset.
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-trimesh-gau-1.pdf}
+
+}
+
+\caption{\label{fig-trimesh-gau}Model generated with five diiferent NLDR
+methods in \(2\text{-}D\) with approximately \(65\) non-empty bins.}
+
+\end{figure}%
+
 \section{Applications}\label{sec-applications}

### 2024-05-20 examples/2_gen_best_model_from_trimap.R

@@ -1,0 +1,66 @@
+library(quollr)
+library(dplyr)
+library(readr)
+library(ggplot2)
+
+## Import data
+training_data_gau <- read_rds("data/five_gau_clusters/data_five_gau_training.rds")
+
+trimap_gau <- read_rds("data/five_gau_clusters/trimap_data_five_gau.rds")
+gau_scaled_obj <- gen_scaled_data(
+  data = trimap_gau)
+trimap_gau_scaled <- gau_scaled_obj$scaled_nldr
+
+## To initialise number of bins along the x-axis
+bin1_vec <- 2:61
+
+lim1 <- gau_scaled_obj$lim1
+lim2 <- gau_scaled_obj$lim2
+r2_gau <- diff(lim2)/diff(lim1)
+
+error_gau <- data.frame(matrix(nrow = 0, ncol = 0))
+
+for (xbins in bin1_vec) {
+
+  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2_gau)$bin2
+
+  gau_model <- fit_highd_model(
+    training_data = training_data_gau,
+    emb_df = trimap_gau_scaled,
+    bin1 = xbins,
+    r2 = r2_gau,
+    is_bin_centroid = TRUE,
+    is_rm_lwd_hex = FALSE,
+    col_start_highd = "x"
+  )
+
+  df_bin_centroids_gau <- gau_model$df_bin_centroids
+  df_bin_gau <- gau_model$df_bin
+
+  ## Compute error
+  error_df <- glance(
+    df_bin_centroids = df_bin_centroids_gau,
+    df_bin = df_bin_gau,
+    training_data = training_data_gau,
+    newdata = NULL,
+    type_NLDR = "TriMAP",
+    col_start = "x") |>
+    mutate(bin1 = xbins,
+           bin2 = bin2,
+           b = bin1 * bin2,
+           b_non_empty = NROW(df_bin_centroids_gau))
+
+  error_gau <- bind_rows(error_gau, error_df)
+
+}
+
+error_gau |>
+  arrange(MSE)
+
+
+ggplot(error_gau, aes(x = b_non_empty,
+                      y = log(MSE))) +
+  geom_point() +
+  geom_line() +
+  geom_vline(xintercept = 64, linetype="solid",
+             color = "black", linewidth=0.8, alpha = 0.5)

### 2024-05-20 paper.qmd
Added MSE plot for choice of bins

@@
 
+```{r}
+training_data_scurve <- read_rds("data/s_curve/s_curve_training.rds")
+```
+
+<!--To generate errors for different total number of bins-->
+```{r}
+## To initialize number of bins along the x-axis
+bin1_vec_scurve <- 2:27 #sqrt(NROW(training_data_scurve))
+
+error_scurve <- data.frame(matrix(nrow = 0, ncol = 0))
+
+for (xbins in bin1_vec_scurve) {
+
+  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2)$bin2
+
+  scurve_model <- fit_highd_model(
+    training_data = training_data_scurve,
+    emb_df = umap_scurve_scaled,
+    bin1 = xbins,
+    r2 = r2,
+    is_bin_centroid = TRUE,
+    is_rm_lwd_hex = FALSE,
+    col_start_highd = "x"
+  )
+
+  df_bin_centroids_scurve <- scurve_model$df_bin_centroids
+  df_bin_scurve <- scurve_model$df_bin
+
+  ## Compute error
+  error_df <- glance(
+    df_bin_centroids = df_bin_centroids_scurve,
+    df_bin = df_bin_scurve,
+    training_data = training_data_scurve,
+    newdata = NULL,
+    type_NLDR = "UMAP",
+    col_start = "x") |>
+    mutate(bin1 = xbins,
+           bin2 = bin2,
+           b = bin1 * bin2,
+           b_non_empty = NROW(df_bin_centroids_scurve))
+
+  error_scurve <- bind_rows(error_scurve, error_df)
+
+}
+
+mse_scurve_b <- ggplot(error_scurve, 
+                     aes(x = b, 
+                         y = log(MSE))) + 
+  geom_point(size = 1) +
+  geom_line() + 
+  geom_vline(xintercept = 190, linetype="solid", 
+             color = "black", linewidth=0.8, alpha = 0.5) +
+  ylab("log(MSE)") +
+  xlab("b") +
+  theme(axis.text.x = element_text(size = 5),
+        axis.text.y = element_text(size = 5),
+        axis.title.x = element_text(size = 5),
+        axis.title.y = element_text(size = 5, angle = 90))
+```
+
+```{r}
+#| fig-cap: "MSE from UMAP applied to S-curve dataset with diffferent $b$ choices.
+What is the effective $b$ to create the model?
+The residual plot have a steep slope at the beginning, indicating that a smaller $b$ causes a larger amount of MSE.
+Then, the slope gradually declines or level off, indicating that a higher $b$ generates a smaller MSE.
+Using the elbow method, it was observed that when the $b$ is set to $190$, the lowest MSE occurred."
+#| label: fig-mse-scurve-b
+#| fig-pos: H
+#| out-width: 50%
+
+mse_scurve_b
+```
+
 #### Removal of low density bins

### 2024-05-20 paper.tex
Added MSE plot for choice of bins

@@
 
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=0.5\textwidth,height=\textheight]{paper_files/figure-pdf/fig-mse-scurve-b-1.pdf}
+
+}
+
+\caption{\label{fig-mse-scurve-b}MSE from UMAP applied to S-curve
+dataset with diffferent \(b\) choices. What is the effective \(b\) to
+create the model? The residual plot have a steep slope at the beginning,
+indicating that a smaller \(b\) causes a larger amount of MSE. Then, the
+slope gradually declines or level off, indicating that a higher \(b\)
+generates a smaller MSE. Using the elbow method, it was observed that
+when the \(b\) is set to \(190\), the lowest MSE occurred.}
+
+\end{figure}%
+
 \subsubsection{Removal of low density

### 2024-05-20 paper.qmd
Added Lwd MSE plot

@@
 
+<!--To generate errors for different total number of bins-->
+```{r}
+## To initialize effective bins along x
+effective_bin1_scurve <- 10
+
+## To initialize benchmark values to remove low density hexagons
+benchmark_rm_hex_vec <- append(seq(0, 0.99, by=0.1),0.99)
+
+error_rm_scurve <- data.frame(matrix(nrow = 0, ncol = 0))
+
+effective_bin2_scurve <- calc_bins_y(bin1 = effective_bin1_scurve, r2 = r2)$bin2
+
+scurve_model <- fit_highd_model(
+  training_data = training_data_scurve,
+  emb_df = umap_scurve_scaled,
+  bin1 = effective_bin1_scurve,
+  r2 = r2,
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = FALSE,
+  col_start_highd = "x"
+)
+
+df_bin_centroids_scurve <- scurve_model$df_bin_centroids
+df_bin_scurve <- scurve_model$df_bin
+
+for (benchmark_rm_lwd in benchmark_rm_hex_vec) {
+
+  df_bin_centroids_scurve_high_dens <- df_bin_centroids_scurve |>
+    filter(std_counts > benchmark_rm_lwd)
+
+  df_bin_scurve_high_dens <- df_bin_scurve |>
+    filter(hb_id %in% df_bin_centroids_scurve_high_dens$hexID)
+
+  ## Compute error
+  error_df <- glance(
+    df_bin_centroids = df_bin_centroids_scurve_high_dens,
+    df_bin = df_bin_scurve_high_dens,
+    training_data = training_data_scurve,
+    newdata = NULL,
+    type_NLDR = "UMAP",
+    col_start = "x") |>
+    mutate(benchmark_rm_lwd = benchmark_rm_lwd,
+           bin1 = effective_bin1_scurve,
+           bin2 = effective_bin2_scurve,
+           b = bin1 * bin2,
+           b_non_empty = NROW(df_bin_centroids_scurve_high_dens))
+
+  error_rm_scurve <- bind_rows(error_rm_scurve, error_df)
+
+}
+
+benchmark_label_df <- tibble(x = 14, y = 0.6, 
+                          label = "benchmark is 0.60.")
+
+mse_scurve_lwd <- ggplot(error_rm_scurve, 
+                     aes(x = b_non_empty, 
+                         y = log(MSE))) + 
+  geom_point(size = 1) +
+  geom_line() + 
+  geom_vline(xintercept = 15, linetype="solid", 
+             color = "black", linewidth=0.8, alpha = 0.5) +
+  geom_text(data = benchmark_label_df, 
+            aes(x=x, 
+                label= label, 
+                y=y), 
+            colour="red", 
+            angle=90) +
+  ylab("log(MSE)") +
+  xlab("number of non-empty bins") +
+  theme(axis.text.x = element_text(size = 5),
+        axis.text.y = element_text(size = 5),
+        axis.title.x = element_text(size = 5),
+        axis.title.y = element_text(size = 5, angle = 90))
+```
+
+```{r}
+#| fig-cap: "MSE from UMAP applied to S-curve dataset with diffferent bechmark choices.
+What is the effective bechmark value to remove low density hexagons?
+The residual plot have a steep slope at the beginning, indicating that a smaller bechmark value causes a larger amount of MSE.
+Then, the slope gradually declines or level off, indicating that a higher bechmark value generates a smaller MSE.
+Using the elbow method, it was observed that when the bechmark value is set to $0.6$, the lowest MSE occurred."
+#| label: fig-mse-scurve-lwd
+#| fig-pos: H
+#| out-width: 50%
+
+mse_scurve_lwd
+```
+
 #### Removing long edges
@@
 <!--add distribution of distance along with the default benchmark-->
-<!--show the trimesh with colored long edges-->
+<!--show the trimesh with colored long edges with different benchmark values-->
 

### 2024-05-20 paper.tex
Added Lwd MSE plot

@@
 
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=0.5\textwidth,height=\textheight]{paper_files/figure-pdf/fig-mse-scurve-lwd-1.pdf}
+
+}
+
+\caption{\label{fig-mse-scurve-lwd}MSE from UMAP applied to S-curve
+dataset with diffferent bechmark choices. What is the effective bechmark
+value to remove low density hexagons? The residual plot have a steep
+slope at the beginning, indicating that a smaller bechmark value causes
+a larger amount of MSE. Then, the slope gradually declines or level off,
+indicating that a higher bechmark value generates a smaller MSE. Using
+the elbow method, it was observed that when the bechmark value is set to
+\(0.6\), the lowest MSE occurred.}
+
+\end{figure}%
+
 \subsubsection{Removing long edges}\label{removing-long-edges}

### 2024-05-20 paper.qmd
Added choice of bins section

@@
 The chosen total number of bins must effectively capture the structure of the NLDR data.
-If the number of bins is too low, the model may not be able to capture the structure of the NLDR data effectively (see @fig-binsize (a)), while if there are too many bins, it may result in over-fitting the individual points of the NLDR data (see @fig-binsize (c)).
-Therefore, it is important to determine an appropriate number of bins to construct an effective model.
-
-<!--add three choice of bins-->
+If the number of bins is too low, the model may not be able to capture the structure of the NLDR data effectively (see @fig-bins-scurve (a)), while if there are too many bins, it may result in over-fitting the individual points of the NLDR data (see @fig-bins-scurve (c)).
+Therefore, it is important to determine an effective number of bins to construct a reasonable model (see @fig-bins-scurve (b)).
+
+<!--add three choice of bins 3, 10, 34-->
+
+```{r}
+## Hexagonal binning to have regular hexagons
+hb_obj_scurve1 <- hex_binning(
+  data = umap_scurve_scaled, 
+  bin1 = 3, 
+  r2 = r2)
+
+## Data set with all centroids
+all_centroids_df1 <- hb_obj_scurve1$centroids
+
+## Generate all coordinates of hexagons
+hex_grid1 <- hb_obj_scurve1$hex_poly
+
+## To obtain the standardise counts within hexbins
+counts_df1 <- hb_obj_scurve1$std_cts
+df_bin_centroids1 <- extract_hexbin_centroids(
+  centroids_df = all_centroids_df1, 
+  counts_df = counts_df1)
+
+hex_grid_with_counts_s_curve1 <- full_join(hex_grid1, 
+                                          df_bin_centroids1 |> select(hexID, std_counts), 
+                                          by = c("hex_poly_id" = "hexID"))
+
+hex_grid_coloured_scurve1 <-  ggplot(
+  data = hex_grid_with_counts_s_curve1, 
+  aes(x = x, y = y)) + 
+  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+  geom_point(data = umap_scurve_scaled, 
+             aes(x = UMAP1, y = UMAP2), 
+             alpha = 0.5,
+             size = 0.5) + 
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") #+
+  #xlim(sc_xlims_b) + ylim(sc_ylims_b)
+
+```
+
+```{r}
+## Hexagonal binning to have regular hexagons
+hb_obj_scurve2 <- hex_binning(
+  data = umap_scurve_scaled, 
+  bin1 = 10, 
+  r2 = r2)
+
+## Data set with all centroids
+all_centroids_df2 <- hb_obj_scurve2$centroids
+
+## Generate all coordinates of hexagons
+hex_grid2 <- hb_obj_scurve2$hex_poly
+
+## To obtain the standardise counts within hexbins
+counts_df2 <- hb_obj_scurve2$std_cts
+df_bin_centroids2 <- extract_hexbin_centroids(
+  centroids_df = all_centroids_df2, 
+  counts_df = counts_df2)
+
+hex_grid_with_counts_s_curve2 <- full_join(hex_grid2, 
+                                          df_bin_centroids2 |> select(hexID, std_counts), 
+                                          by = c("hex_poly_id" = "hexID"))
+
+hex_grid_coloured_scurve2 <-  ggplot(
+  data = hex_grid_with_counts_s_curve2, 
+  aes(x = x, y = y)) + 
+  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+  geom_point(data = umap_scurve_scaled, 
+             aes(x = UMAP1, y = UMAP2), 
+             alpha = 0.5,
+             size = 0.5) + 
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  xlim(sc_xlims) + ylim(sc_ylims)
+
+```
+
+```{r}
+## Hexagonal binning to have regular hexagons
+hb_obj_scurve3 <- hex_binning(
+  data = umap_scurve_scaled, 
+  bin1 = 34, 
+  r2 = r2)
+
+## Data set with all centroids
+all_centroids_df3 <- hb_obj_scurve3$centroids
+
+## Generate all coordinates of hexagons
+hex_grid3 <- hb_obj_scurve3$hex_poly
+
+## To obtain the standardise counts within hexbins
+counts_df3 <- hb_obj_scurve3$std_cts
+df_bin_centroids3 <- extract_hexbin_centroids(
+  centroids_df = all_centroids_df3, 
+  counts_df = counts_df3)
+
+hex_grid_with_counts_s_curve3 <- full_join(hex_grid3, 
+                                          df_bin_centroids3 |> select(hexID, std_counts), 
+                                          by = c("hex_poly_id" = "hexID"))
+
+hex_grid_coloured_scurve3 <-  ggplot(
+  data = hex_grid_with_counts_s_curve3, 
+  aes(x = x, y = y)) + 
+  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+  geom_point(data = umap_scurve_scaled, 
+             aes(x = UMAP1, y = UMAP2), 
+             alpha = 0.5,
+             size = 0.5) + 
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  xlim(sc_xlims) + ylim(sc_ylims)
+
+```
+
+```{r}
+#| echo: false
+#| label: fig-bins-scurve
+#| fig-pos: H
+#| fig-cap: "Hexbin plots from different number of bins for the UMAP applied to S-curve data: (a) $b = 12$ (3, 5), (b) $b = 190$ (10, 19), and (c) $b = 2176$ (34, 64).
+The hexbins are colored based on the density of points, with darker colors indicating higher density and yellow color representing lower density within each bin.
+What is the number of bins that would be effective?"
+#| fig-width: 6
+#| fig-height: 3
+
+hex_grid_coloured_scurve1 + hex_grid_coloured_scurve2 + hex_grid_coloured_scurve3 +
+  plot_layout(guides='collect', ncol = 3) &
+  theme(legend.position='none', plot.tag = element_text(size = 8))
+```
 
@@
 The minimum $b_1$ is set to $2$, while the maximum number is estimated by taking the square root of $n$.
-The analysis evaluates the MSE across varying $b$ within this range, covering the minimum to maximum values along both axes (see @fig-diagnostic1-scurve).
+The analysis evaluates the MSE across varying $b$ within this range, covering the minimum to maximum values along both axes (see @fig-mse-scurve-b).
 
@@
 #| fig-pos: H
-#| out-width: 50%
+#| out-height: 40%
 
@@
 
+```{r}
+## To initialize effective bins along x
+effective_bin1_scurve <- 10
+
+effective_bin2_scurve <- calc_bins_y(bin1 = effective_bin1_scurve, r2 = r2)$bin2
+
+scurve_model <- fit_highd_model(
+  training_data = training_data_scurve,
+  emb_df = umap_scurve_scaled,
+  bin1 = effective_bin1_scurve,
+  r2 = r2,
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = FALSE,
+  col_start_highd = "x"
+)
+
+df_bin_centroids_scurve <- scurve_model$df_bin_centroids
+df_bin_scurve <- scurve_model$df_bin
+```
+
+```{r}
+cell_count_scurve <- ggplot(df_bin_centroids_scurve, 
+                            aes(x = reorder(as.factor(hexID), 
+                                            -std_counts), 
+                                y = std_counts)) + 
+  geom_quasirandom() + 
+  xlab("hexagonal id") + 
+  ylab("standardized cell count") +
+  geom_hline(yintercept = 0.2, color = "black", linewidth=0.8, alpha = 0.5) +
+  theme(axis.title.x = element_text(size = 5),
+        axis.title.y = element_text(size = 5, angle = 90),
+        axis.text.x = element_text(size = 5, angle = 90),
+        axis.text.y = element_text(size = 5))
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "Distribution of standardize counts by hexagons."
+#| label: fig-stdcts-scurve
+#| out-height: 50%
+#| fig-pos: H
+
+cell_count_scurve
+```
+
 Furthermore, selecting the benchmark value for removing low-density hexagons is important.
@@
 ```{r}
-## To initialize effective bins along x
-effective_bin1_scurve <- 10
-
 ## To initialize benchmark values to remove low density hexagons
@@
 error_rm_scurve <- data.frame(matrix(nrow = 0, ncol = 0))
-
-effective_bin2_scurve <- calc_bins_y(bin1 = effective_bin1_scurve, r2 = r2)$bin2
 
@@
                 y=y), 
-            colour="red", 
+            colour="black", 
             angle=90) +
@@
 #| fig-pos: H
-#| out-width: 50%
+#| out-height: 50%
 
@@
 <!--add distribution of distance along with the default benchmark-->
+
+```{r}
+## Triangulate bin centroids
+tr1_object_scurve <- tri_bin_centroids(
+  df_bin_centroids_scurve, x = "c_x", y = "c_y")
+tr_from_to_df_scurve <- gen_edges(
+  tri_object = tr1_object_scurve)
+
+## Compute 2D distances
+distance_scurve <- cal_2d_dist(
+  tr_coord_df = tr_from_to_df_scurve, 
+  start_x = "x_from", 
+  start_y = "y_from", 
+  end_x = "x_to", 
+  end_y = "y_to", 
+  select_vars = c("from", "to", "distance"))
+
+## To find the benchmark value
+benchmark_scurve <- find_lg_benchmark(
+  distance_edges = distance_scurve, 
+  distance_col = "distance")
+
+## To draw the distance distribution
+distance_scurve$group <- "1"
+distance_scurve_plot <- ggplot(distance_scurve, 
+                        aes(x = group, 
+                            y = distance)) + 
+  geom_quasirandom() + 
+  ylim(0, max(unlist(distance_scurve$distance))+ 0.5) + 
+  coord_flip() + 
+  geom_hline(yintercept = benchmark_scurve, linetype="solid", 
+             color = "black", linewidth=0.8, alpha = 0.5) +
+  ylab(expression(d^{(2)})) +
+  theme(axis.text.x = element_text(size = 5),
+        axis.title.x = element_text(size = 5))
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "Distribution of $2\\text{-}D$ Euclidean distances between bin centroids of the triangular mesh generated with S-curve UMAP data."
+#| label: fig-dist-scurve
+#| out-height: 50%
+#| fig-pos: H
+
+distance_scurve_plot
+```
+
 <!--show the trimesh with colored long edges with different benchmark values-->

### 2024-05-20 paper.tex
Added choice of bins section

@@
 may not be able to capture the structure of the NLDR data effectively
-(see \textbf{?@fig-binsize} (a)), while if there are too many bins, it
-may result in over-fitting the individual points of the NLDR data (see
-\textbf{?@fig-binsize} (c)). Therefore, it is important to determine an
-appropriate number of bins to construct an effective model.
+(see Figure~\ref{fig-bins-scurve} (a)), while if there are too many
+bins, it may result in over-fitting the individual points of the NLDR
+data (see Figure~\ref{fig-bins-scurve} (c)). Therefore, it is important
+to determine an effective number of bins to construct a reasonable model
+(see Figure~\ref{fig-bins-scurve} (b)).
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-bins-scurve-1.pdf}
+
+}
+
+\caption{\label{fig-bins-scurve}Hexbin plots from different number of
+bins for the UMAP applied to S-curve data: (a) \(b = 12\) (3, 5), (b)
+\(b = 190\) (10, 19), and (c) \(b = 2176\) (34, 64). The hexbins are
+colored based on the density of points, with darker colors indicating
+higher density and yellow color representing lower density within each
+bin. What is the number of bins that would be effective?}
+
+\end{figure}%
 
@@
 range, covering the minimum to maximum values along both axes (see
-\textbf{?@fig-diagnostic1-scurve}).
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=0.5\textwidth,height=\textheight]{paper_files/figure-pdf/fig-mse-scurve-b-1.pdf}
+Figure~\ref{fig-mse-scurve-b}).
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=0.5\textheight]{paper_files/figure-pdf/fig-mse-scurve-b-1.pdf}
 
@@
 
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=0.5\textheight]{paper_files/figure-pdf/fig-stdcts-scurve-1.pdf}
+
+}
+
+\caption{\label{fig-stdcts-scurve}Distribution of standardize counts by
+hexagons.}
+
+\end{figure}%
+
 Furthermore, selecting the benchmark value for removing low-density
@@
 
-\includegraphics[width=0.5\textwidth,height=\textheight]{paper_files/figure-pdf/fig-mse-scurve-lwd-1.pdf}
+\includegraphics[width=1\textwidth,height=0.5\textheight]{paper_files/figure-pdf/fig-mse-scurve-lwd-1.pdf}
 
@@
 \end{enumerate}
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=0.5\textheight]{paper_files/figure-pdf/fig-dist-scurve-1.pdf}
+
+}
+
+\caption{\label{fig-dist-scurve}Distribution of \(2\text{-}D\) Euclidean
+distances between bin centroids of the triangular mesh generated with
+S-curve UMAP data.}
+
+\end{figure}%
 

### 2024-05-20 paper.qmd
Added removing long edge section

@@
 #| fig-pos: H
-#| out-height: 40%
+#| out-height: 50%
 
@@
 
-Creating a smooth 2D representation (see @fig-modelScurve (c)) requires removing edges that connect distant bin centroids within the triangular mesh (see @fig-modelScurve (b)).
-These edges only exist in the 2D model and do not extend into higher dimensions, ensuring that their removal does not impact the model in higher dimensions.
-While specific criteria for determining the benchmark value to remove long edges do not exist, the following steps provide an approach to identifying a suitable threshold:
-
-1. Plot the distribution of the 2D Euclidean distances (see @fig-distScurve).
-2. Identify the largest difference between consecutive distance values.
-3. Take the largest distance value corresponding to this difference as the benchmark value.
+To create a smooth \gD{} representation (see @fig-NLDR-scurve (d)), it is necessary to remove edges that connect distant bin centroids in the triangular mesh.
+These edges only exist in the \gD{} model and do not extend into \pD{}, so their removal does not impact the model in \pD{}.
+Although there are no specific criteria for determining the benchmark value to remove long edges, the following steps provide an approach to identifying a suitable threshold:
+
+1. Plot the distribution of the 2D Euclidean distances (see @fig-dist-scurve).
+2. Identify the first largest difference between consecutive distance values.
+3. Take the distance value corresponding to this difference as the benchmark value.
 

### 2024-05-20 paper.qmd
Update with mse

@@
 In a study by @chen2023, this dataset was used to demonstrate how UMAP represents clusters (see @fig-umap-author).
-As shown in @fig-umap-author, there are three distinct and well-separated clusters.
+As shown in @fig-umap-author, there are three distant and well-separated clusters.
 
@@
 #| fig-cap: "(a) Model generated in $2\\text{-}D$ with UMAP, and (b) $p\\text{-}D$ model error in $2\\text{-}D$.
-The $2\\text{-}D$ model shows three well-separated distinct clusters.
+The $2\\text{-}D$ model shows three well-separated distant clusters.
 The $p\\text{-}D$ model errors are distributed along clusters."
@@
 #| fig-cap: "(a) Model generated in $2\\text{-}D$ with tSNE, and (b) $p\\text{-}D$ model error in $2\\text{-}D$.
-The $2\\text{-}D$ model shows three well-separated distinct clusters.
+The $2\\text{-}D$ model shows three well-separated distant clusters.
 The $p\\text{-}D$ model errors are distributed along clusters, but most low $p\\text{-}D$ model errors present in the large cluster."

### 2024-05-20 paper.tex
Added removing long edge section
Update with mse

@@
 
-Creating a smooth 2D representation (see \textbf{?@fig-modelScurve} (c))
-requires removing edges that connect distant bin centroids within the
-triangular mesh (see \textbf{?@fig-modelScurve} (b)). These edges only
-exist in the 2D model and do not extend into higher dimensions, ensuring
-that their removal does not impact the model in higher dimensions. While
-specific criteria for determining the benchmark value to remove long
-edges do not exist, the following steps provide an approach to
-identifying a suitable threshold:
+To create a smooth \gD{} representation (see
+Figure~\ref{fig-NLDR-scurve} (d)), it is necessary to remove edges that
+connect distant bin centroids in the triangular mesh. These edges only
+exist in the \gD{} model and do not extend into \pD{}, so their removal
+does not impact the model in \pD{}. Although there are no specific
+criteria for determining the benchmark value to remove long edges, the
+following steps provide an approach to identifying a suitable threshold:
 
@@
   Plot the distribution of the 2D Euclidean distances (see
-  \textbf{?@fig-distScurve}).
+  Figure~\ref{fig-dist-scurve}).
 \item
-  Identify the largest difference between consecutive distance values.
+  Identify the first largest difference between consecutive distance
+  values.
 \item
-  Take the largest distance value corresponding to this difference as
-  the benchmark value.
+  Take the distance value corresponding to this difference as the
+  benchmark value.
 \end{enumerate}
@@
 Figure~\ref{fig-umap-author}). As shown in Figure~\ref{fig-umap-author},
-there are three distinct and well-separated clusters.
+there are three distant and well-separated clusters.
 
@@
 \(2\text{-}D\). The \(2\text{-}D\) model shows three well-separated
-distinct clusters. The \(p\text{-}D\) model errors are distributed along
+distant clusters. The \(p\text{-}D\) model errors are distributed along
 clusters.}
@@
 with tSNE, and (b) \(p\text{-}D\) model error in \(2\text{-}D\). The
-\(2\text{-}D\) model shows three well-separated distinct clusters. The
+\(2\text{-}D\) model shows three well-separated distant clusters. The
 \(p\text{-}D\) model errors are distributed along clusters, but most low

### 2024-05-20 paper.qmd
Added models from phate for five Gaussian data

@@
 
-After establishing the hexagonal grid with an appropriate number of bins, some hexagonal bins may have few or no data points within them (see @fig-rmlowdenshex (a)).
+Once setting up the hexagonal grid with an appropriate number of bins, some hexagonal bins may have few or no data points within them (see @fig-rmlowdenshex (a)).
 To ensure comprehensive coverage of the NLDR data, it is necessary to select hexagonal bins with a considerable number of data points.
@@
 ```{r}
-gau3_scaled_obj <- gen_scaled_data(
-  data = phate_data_gau)
-phate_gau_scaled <- gau3_scaled_obj$scaled_nldr
-
-## Compute hexbin parameters
-num_bins_x_gau3 <- 250
-lim1 <- gau3_scaled_obj$lim1
-lim2 <- gau3_scaled_obj$lim2
-r2_gau3 <- diff(lim2)/diff(lim1)
-
-gau3_model <- fit_highd_model(
-  training_data = training_data_gau,
-  emb_df = phate_gau_scaled,
-  bin1 = num_bins_x_gau3,
-  r2 = r2_gau3,
-  is_bin_centroid = TRUE,
-  is_rm_lwd_hex = FALSE,
-  col_start_highd = "x"
-)
-
-df_bin_centroids_gau3 <- gau3_model$df_bin_centroids
-df_bin_gau3 <- gau3_model$df_bin
+# gau3_scaled_obj <- gen_scaled_data(
+#   data = phate_data_gau)
+# phate_gau_scaled <- gau3_scaled_obj$scaled_nldr
+# 
+# ## Compute hexbin parameters
+# num_bins_x_gau3 <- 250
+# lim1 <- gau3_scaled_obj$lim1
+# lim2 <- gau3_scaled_obj$lim2
+# r2_gau3 <- diff(lim2)/diff(lim1)
+# 
+# gau3_model <- fit_highd_model(
+#   training_data = training_data_gau,
+#   emb_df = phate_gau_scaled,
+#   bin1 = num_bins_x_gau3,
+#   r2 = r2_gau3,
+#   is_bin_centroid = TRUE,
+#   is_rm_lwd_hex = FALSE,
+#   col_start_highd = "x"
+# )
+# 
+# df_bin_centroids_gau3 <- gau3_model$df_bin_centroids
+# df_bin_gau3 <- gau3_model$df_bin
+
+df_bin_centroids_gau3 <- read_rds("data/five_gau_clusters/df_bin_centroids_gau_phate.rds")
+df_bin_gau3 <- read_rds("data/five_gau_clusters/df_bin_gau_phate.rds")
 

### 2024-05-20 paper.tex
Added models from phate for five Gaussian data

@@
 
-After establishing the hexagonal grid with an appropriate number of
-bins, some hexagonal bins may have few or no data points within them
-(see \textbf{?@fig-rmlowdenshex} (a)). To ensure comprehensive coverage
-of the NLDR data, it is necessary to select hexagonal bins with a
+Once setting up the hexagonal grid with an appropriate number of bins,
+some hexagonal bins may have few or no data points within them (see
+\textbf{?@fig-rmlowdenshex} (a)). To ensure comprehensive coverage of
+the NLDR data, it is necessary to select hexagonal bins with a
 considerable number of data points. This involves calculating the number

### 2024-05-20 paper.qmd
Add hexbin notation vis

@@
 
-The 2D hexagon grid is defined by the number of bins in each direction $(b_1, b_2)$, giving total number of bins as $b = b_1 \times b_2$, and hexagon id, $h = 1, \dots, b$.
-Each hexagon, $H_h$ is uniquely described by centroid, $ C_{h}^{(2)} = (c_{h1}, c_{h2})$.
+The \gD{} hexagon grid is defined by the number of bins in each direction $(b_1, b_2)$, giving total number of bins as $b = b_1 \times b_2$, and hexagon id, $h = 1, \dots, b$.
+Each hexagon, $H_h$ is uniquely described by centroid, $C_{h}^{(2)} = (c_{h1}, c_{h2})$.
 The lower left position where the grid starts at $(s_1, s_2)$, which correspond to the lowest left centroid.
@@
     geom_point(data = all_centroids_df |> dplyr::filter(hexID == 1), aes(x = c_x, y = c_y), color = "black") + ## starting point
-    geom_rect(aes(xmin = 0, xmax = 1, ymin = 0, ymax = r2), fill = "white", color = "black", alpha = 0, linewidth = 1) +
-    geom_segment(x = 0, y = -0.35, xend = 1, yend = -0.35,
+    geom_rect(aes(xmin = 0 - 0.03, xmax = 1 - 0.03, ymin = 0 - 0.05, ymax = r2- 0.05), fill = "white", color = "black", alpha = 0, linewidth = 1) +
+    geom_segment(x = 0 - 0.03, y = -0.35, xend = 1 - 0.03, yend = -0.35,
                  arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+
     annotate("text", x=0.5, y=-0.45, label= "r1", color = "black") +
-    geom_segment(x = -0.25, y = 0, xend = -0.25, yend = r2,
+    geom_segment(x = -0.25, y = 0 - 0.05, xend = -0.25, yend = r2 - 0.05,
                  arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ 
@@
     annotate("text", x=-0.15, y=0.45, label= "a2", color = "black") +
-    annotate("text", x=-0.15, y=-0.35, label= "(s1, s2)", color = "black") + ## starting point
+    annotate("text", x=-0.18, y=-0.35, label= "(s1, s2)", color = "black") + ## starting point
   coord_fixed()
@@
 
-Furthermore, the total number of bins is calculated based on the number of bins along the x-axis and y-axis, $b = b_1 \times b_2$.
 To determine the effective $b$, candidate values are selected based on the range between the minimum and approximate maximum $b_1$, because $b_2$ is computed from $b_1$.
-The minimum $b_1$ is set to $2$, while the maximum number is estimated by taking the square root of $n$.
+The minimum $b_1$ is set to $2$, while the maximum number is estimated by taking the square root of $\frac{n}{2}$.
 The analysis evaluates the MSE across varying $b$ within this range, covering the minimum to maximum values along both axes (see @fig-mse-scurve-b).
@@
 ## To initialize number of bins along the x-axis
-bin1_vec_scurve <- 2:27 #sqrt(NROW(training_data_scurve))
+bin1_vec_scurve <- 2:19 #sqrt(NROW(training_data_scurve)/2)
 
@@
 #| fig-pos: H
-#| out-height: 50%
+#| out-height: 20%
 

### 2024-05-20 paper.tex
Add hexbin notation vis

@@
 
-The 2D hexagon grid is defined by the number of bins in each direction
-\((b_1, b_2)\), giving total number of bins as \(b = b_1 \times b_2\),
-and hexagon id, \(h = 1, \dots, b\). Each hexagon, \(H_h\) is uniquely
-described by centroid, \$ C\_\{h\}\^{}\{(2)\} = (c\_\{h1\},
-c\_\{h2\})\$. The lower left position where the grid starts at
-\((s_1, s_2)\), which correspond to the lowest left centroid. The values
-of \(s_i\) need to be below their respective minimum variable values,
-and could be a full bin lower, to allow a buffer (\(q\)) corresponding
-to a full hexagon width (\(a_1\)) and height (\(a_2\)) around the data.
-The values of \(b_i\) are variables to be computed that define the
-reduction in size of the data (\(n\) to \(m\)).
+The \gD{} hexagon grid is defined by the number of bins in each
+direction \((b_1, b_2)\), giving total number of bins as
+\(b = b_1 \times b_2\), and hexagon id, \(h = 1, \dots, b\). Each
+hexagon, \(H_h\) is uniquely described by centroid,
+\(C_{h}^{(2)} = (c_{h1}, c_{h2})\). The lower left position where the
+grid starts at \((s_1, s_2)\), which correspond to the lowest left
+centroid. The values of \(s_i\) need to be below their respective
+minimum variable values, and could be a full bin lower, to allow a
+buffer (\(q\)) corresponding to a full hexagon width (\(a_1\)) and
+height (\(a_2\)) around the data. The values of \(b_i\) are variables to
+be computed that define the reduction in size of the data (\(n\) to
+\(m\)).
 
@@
 
-Furthermore, the total number of bins is calculated based on the number
-of bins along the x-axis and y-axis, \(b = b_1 \times b_2\). To
-determine the effective \(b\), candidate values are selected based on
+To determine the effective \(b\), candidate values are selected based on
 the range between the minimum and approximate maximum \(b_1\), because
@@
 while the maximum number is estimated by taking the square root of
-\(n\). The analysis evaluates the MSE across varying \(b\) within this
-range, covering the minimum to maximum values along both axes (see
-Figure~\ref{fig-mse-scurve-b}).
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=0.5\textheight]{paper_files/figure-pdf/fig-mse-scurve-b-1.pdf}
+\(\frac{n}{2}\). The analysis evaluates the MSE across varying \(b\)
+within this range, covering the minimum to maximum values along both
+axes (see Figure~\ref{fig-mse-scurve-b}).
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=0.2\textheight]{paper_files/figure-pdf/fig-mse-scurve-b-1.pdf}
 

### 2024-05-20 paper.qmd
Added Scurve model in data space screen shots

@@
 
-#### Generating edges to indicate neighbours and removing long edges
-
-Delaunay triangulation is used to connect neighbouring centroids, which is needed to preserve neighbourhood information when the model is lifted into \pD{}.
-
-When $k = 2$ Delaunay triangulation on $C^{(2)}$ generates the model in 2D space, which is a triangular mesh (@fig-NLDR-scurve (d)).
+#### Generating edges to indicate neighbors and removing long edges
+
+Delaunay triangulation is used to connect neighboring centroids, which is needed to preserve neighborhood information when the model is lifted into \pD{}.
+
+When $k = 2$ Delaunay triangulation on $C^{(2)}$ generates the model in \gD{} space, which is a triangular mesh (@fig-NLDR-scurve (d)).
 It generates convex hulls of $C^{(2)}$ such that the circumcircle of every triangle in the triangulation contains no other points from $C^{(2)}$.
@@
 <!--langevitour with pD model for S-curve-->
+::: {#fig-scurve_sc layout-ncol="3" fig-pos="H"}
+![](figures/scurve/sc_1.png){#fig-scurve_sc1}
+
+![](figures/scurve/sc_2.png){#fig-scurve_sc2}
+
+![](figures/scurve/sc_3.png){#fig-scurve_sc3}
+
+Screen shots of the **langevitour** of the S-curve, shows the model-in-data space, a video of the tour animation is available at (<https://youtu.be/G1m4q9k--v4>).
+:::
 
@@
   xlab("b") +
-  theme(axis.text.x = element_text(size = 5),
-        axis.text.y = element_text(size = 5),
-        axis.title.x = element_text(size = 5),
-        axis.title.y = element_text(size = 5, angle = 90))
+  theme(axis.text.x = element_text(size = 7),
+        axis.text.y = element_text(size = 7),
+        axis.title.x = element_text(size = 7),
+        axis.title.y = element_text(size = 7, angle = 90))
 ```
@@
 #| fig-pos: H
-#| out-height: 20%
+#| out-height: 50%
 

### 2024-05-20 paper.tex
Added Scurve model in data space screen shots

@@
 
-\subsubsection{Generating edges to indicate neighbours and removing long
-edges}\label{generating-edges-to-indicate-neighbours-and-removing-long-edges}
-
-Delaunay triangulation is used to connect neighbouring centroids, which
-is needed to preserve neighbourhood information when the model is lifted
+\subsubsection{Generating edges to indicate neighbors and removing long
+edges}\label{generating-edges-to-indicate-neighbors-and-removing-long-edges}
+
+Delaunay triangulation is used to connect neighboring centroids, which
+is needed to preserve neighborhood information when the model is lifted
 into \pD{}.
@@
 When \(k = 2\) Delaunay triangulation on \(C^{(2)}\) generates the model
-in 2D space, which is a triangular mesh (Figure~\ref{fig-NLDR-scurve}
+in \gD{} space, which is a triangular mesh (Figure~\ref{fig-NLDR-scurve}
 (d)). It generates convex hulls of \(C^{(2)}\) such that the
@@
 \kD{} representations.
+
+\begin{figure}[H]
+
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/scurve/sc_1.png}
+
+}
+
+\subcaption{\label{fig-scurve_sc1}}
+
+\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/scurve/sc_2.png}
+
+}
+
+\subcaption{\label{fig-scurve_sc2}}
+
+\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/scurve/sc_3.png}
+
+}
+
+\subcaption{\label{fig-scurve_sc3}}
+
+\end{minipage}%
+
+\caption{\label{fig-scurve_sc}Screen shots of the \textbf{langevitour}
+of the S-curve, shows the model-in-data space, a video of the tour
+animation is available at (\url{https://youtu.be/G1m4q9k--v4}).}
+
+\end{figure}%
 
@@
 
-\includegraphics[width=1\textwidth,height=0.2\textheight]{paper_files/figure-pdf/fig-mse-scurve-b-1.pdf}
+\includegraphics[width=1\textwidth,height=0.5\textheight]{paper_files/figure-pdf/fig-mse-scurve-b-1.pdf}
 

### 2024-05-20 examples/2_gen_best_model_from_pacmap.R

@@ -1,66 +1,0 @@
-library(quollr)
-library(dplyr)
-library(readr)
-library(ggplot2)
-
-## Import data
-training_data_gau <- read_rds("data/five_gau_clusters/data_five_gau_training.rds")
-
-pacmap_gau <- read_rds("data/five_gau_clusters/pacmap_data_five_gau.rds")
-gau_scaled_obj <- gen_scaled_data(
-  data = pacmap_gau)
-pacmap_gau_scaled <- gau_scaled_obj$scaled_nldr
-
-## To initialise number of bins along the x-axis
-bin1_vec <- 2:61
-
-lim1 <- gau_scaled_obj$lim1
-lim2 <- gau_scaled_obj$lim2
-r2_gau <- diff(lim2)/diff(lim1)
-
-error_gau <- data.frame(matrix(nrow = 0, ncol = 0))
-
-for (xbins in bin1_vec) {
-
-  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2_gau)$bin2
-
-  gau_model <- fit_highd_model(
-    training_data = training_data_gau,
-    emb_df = pacmap_gau_scaled,
-    bin1 = xbins,
-    r2 = r2_gau,
-    is_bin_centroid = TRUE,
-    is_rm_lwd_hex = FALSE,
-    col_start_highd = "x"
-  )
-
-  df_bin_centroids_gau <- gau_model$df_bin_centroids
-  df_bin_gau <- gau_model$df_bin
-
-  ## Compute error
-  error_df <- glance(
-    df_bin_centroids = df_bin_centroids_gau,
-    df_bin = df_bin_gau,
-    training_data = training_data_gau,
-    newdata = NULL,
-    type_NLDR = "PaCMAP",
-    col_start = "x") |>
-    mutate(bin1 = xbins,
-           bin2 = bin2,
-           b = bin1 * bin2,
-           b_non_empty = NROW(df_bin_centroids_gau))
-
-  error_gau <- bind_rows(error_gau, error_df)
-
-}
-
-error_gau |>
-  arrange(MSE)
-
-
-ggplot(error_gau, aes(x = b_non_empty,
-                      y = log(MSE))) +
-  geom_point() +
-  geom_line() +
-  geom_vline(xintercept = 65, linetype="solid",
-             color = "black", linewidth=0.8, alpha = 0.5)

### 2024-05-20 paper.qmd
Added lwd plots

@@
              size = 0.5) + 
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") #+
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  interior_annotation("a") #+
   #xlim(sc_xlims_b) + ylim(sc_ylims_b)
@@
   scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  xlim(sc_xlims) + ylim(sc_ylims)
+  xlim(sc_xlims) + ylim(sc_ylims) +
+  interior_annotation("b")
 
@@
   scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  xlim(sc_xlims) + ylim(sc_ylims)
+  xlim(sc_xlims) + ylim(sc_ylims) +
+  interior_annotation("c")
 
@@
 
-Once setting up the hexagonal grid with an appropriate number of bins, some hexagonal bins may have few or no data points within them (see @fig-rmlowdenshex (a)).
+Once setting up the hexagonal grid with an appropriate number of bins, some hexagonal bins may have few or no data points within them (see @fig-bins-scurve (b)).
 To ensure comprehensive coverage of the NLDR data, it is necessary to select hexagonal bins with a considerable number of data points.
 This involves calculating the number of points within each hexagon.
-Then, the standard count is computed by dividing the number of points within each hexagon by the maximum number of points in the grid (see @eq-equationp2).
-Next, bins with a standard count less than a certain benchmark value are removed (see @fig-rmlowdenshex (b)).
-The following steps will help in determining a suitable value for removing low-density hexagons:
-
-1. Plot the distribution of the standardized counts (see @fig-stdctsScurve).
+Then, the standard count is computed by dividing the number of points within each hexagon by the maximum number of points in the grid.
+Next, bins with a standard count less than a benchmark value are removed (see @fig-lwd-scurve (a)).
+There is no specific rule for selecting a benchmark value.
+However, the following steps can help determine a suitable value for removing low-density hexagons:
+
+1. Plot the distribution of the standardized counts (see @fig-stdcts-scurve).
 2. Examine the distribution of counts.
@@
 
-<!--add distribution of density and add the first quartile as the benchmark-->
+<!--add distribution of density and add the first quantile as the benchmark-->
 
@@
 df_bin_scurve <- scurve_model$df_bin
+
+## To obtain the first quntile
+benchmark1 <- round(quantile(df_bin_centroids_scurve$std_counts, names = FALSE)[2], 3)
 ```
@@
   xlab("hexagonal id") + 
-  ylab("standardized cell count") +
-  geom_hline(yintercept = 0.2, color = "black", linewidth=0.8, alpha = 0.5) +
+  ylab("standardized bin count") +
+  geom_hline(yintercept = benchmark1, color = "black", linewidth=0.8, alpha = 0.5) +
   theme(axis.title.x = element_text(size = 5),
@@
 
-Furthermore, selecting the benchmark value for removing low-density hexagons is important.
-Removing unnecessary bins may lead to the formation of long edges and an uneven 2D model.
-Hence, rather than solely relying on the benchmark value to identify hexagons for removal, it's essential to consider the standard number of points in the neighboring hexagons of the identified low-density ones (see @fig-rmlowdenshex (c)).
-If neighboring bins also show low counts, only those bins will be removed.
-The remaining bins are used to construct the 2D model.
-
-<!--add vis showing removing the hexagons using first quantile, and observing neighbors-->
-
 The benchmark value for removing low-density hexagons ranges between $0$ and $1$.
-When analyzing how these benchmark values influence model performance, it's essential to observe the change in MSE as the benchmark value increases (see @fig-diagnosticpltScurvelwd).
+When analyzing how these benchmark values influence model performance, it's essential to observe the change in MSE as the benchmark value increases (see @fig-mse-scurve-lwd).
 The MSE shows a gradual increase as the benchmark value progresses from $0$ to $1$.
@@
 ## To initialize benchmark values to remove low density hexagons
-benchmark_rm_hex_vec <- append(seq(0, 0.99, by=0.1),0.99)
+benchmark_rm_hex_vec <- append(seq(0, 0.99, by=0.1), c(benchmark1, 0.99))
 
@@
 
-benchmark_label_df <- tibble(x = 14, y = 0.6, 
-                          label = "benchmark is 0.60.")
+benchmark_label_df <- tibble(x = 0.2, y = 0.1, 
+                          label = paste0("benchmark is ", benchmark1))
 
 mse_scurve_lwd <- ggplot(error_rm_scurve, 
-                     aes(x = b_non_empty, 
+                     aes(x = benchmark_rm_lwd, 
                          y = log(MSE))) + 
@@
   geom_line() + 
-  geom_vline(xintercept = 15, linetype="solid", 
+  geom_vline(xintercept = benchmark1, linetype="solid", 
              color = "black", linewidth=0.8, alpha = 0.5) +
@@
   ylab("log(MSE)") +
-  xlab("number of non-empty bins") +
+  xlab("benchmark value") +
   theme(axis.text.x = element_text(size = 5),
@@
 What is the effective bechmark value to remove low density hexagons?
-The residual plot have a steep slope at the beginning, indicating that a smaller bechmark value causes a larger amount of MSE.
-Then, the slope gradually declines or level off, indicating that a higher bechmark value generates a smaller MSE.
-Using the elbow method, it was observed that when the bechmark value is set to $0.6$, the lowest MSE occurred."
+The residual plot have a steep slope at the end, indicating that a smaller bechmark value causes a small amount of MSE.
+Then, the slope gradually increases or level up, indicating that a higher bechmark value generates a higher MSE.
+Using the elbow method, it was observed that when the bechmark value is set to $0.242$, the lowest MSE occurred."
 #| label: fig-mse-scurve-lwd
@@
 mse_scurve_lwd
+```
+
+Furthermore, selecting the benchmark value for removing low-density hexagons is important.
+Removing unnecessary bins may lead to the formation of long edges and an uneven \gD{} model.
+Hence, rather than solely relying on the benchmark value to identify hexagons for removal, it's essential to consider the standard number of points in the neighboring hexagons of the identified low-density bins (see @fig-lwd-scurve (b)).
+If neighboring bins also show low counts, only those bins will be removed.
+The remaining bins are used to construct the \gD{} model.
+
+<!--add vis showing removing the hexagons using first quantile, and observing neighbors-->
+<!--Using first quantile as the benchmark value-->
+```{r}
+hex_grid_with_low_counts <- full_join(hex_grid2, 
+    df_bin_centroids_scurve, 
+    by = c("hex_poly_id" = "hexID")) |>
+  filter(!is.na(std_counts)) |>
+  mutate(type = if_else(std_counts <= benchmark1, "low", "high")) |>
+  filter(type == "low")
+
+hex_grid_rm_lwd_scurve1 <-  ggplot(
+  data = hex_grid_with_counts_s_curve2, 
+  aes(x = x, y = y)) + 
+  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+  geom_text(data = hex_grid_with_low_counts, 
+            color = "black", 
+            aes(x = c_x, y = c_y, group = hex_poly_id, label = "X", size = 1)) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  xlim(sc_xlims) + ylim(sc_ylims) +
+  interior_annotation("a")
+```
+
+<!--Observing neighboring bins as well-->
+```{r}
+## First define low density bins using first quantile
+df_bin_centroids_low <- df_bin_centroids_scurve |> 
+  filter(std_counts <= benchmark1)
+
+## Check neighboring bins
+remove_id <- find_low_dens_hex(df_bin_centroids_all = df_bin_centroids_scurve, 
+                               bin1 = effective_bin1_scurve, 
+                               df_bin_centroids_low = df_bin_centroids_low)
+
+## Remove the identified bins
+df_bin_centroids_scurve_removed <- df_bin_centroids_scurve |>
+  filter(hexID %in%remove_id)
+
+hex_grid_with_low_counts_neighbor <- right_join(hex_grid, 
+                                             df_bin_centroids_scurve_removed, 
+                                             by = c("hex_poly_id" = "hexID"))
+
+hex_grid_rm_lwd_scurve2 <-  ggplot(
+  data = hex_grid_with_counts_s_curve2, 
+  aes(x = x, y = y)) + 
+  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+  geom_text(data = hex_grid_with_low_counts_neighbor, 
+            color = "black", 
+            aes(x = c_x, y = c_y, group = hex_poly_id, label = "X", size = 1)) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  xlim(sc_xlims) + ylim(sc_ylims) +
+  interior_annotation("b")
+```
+
+```{r}
+#| echo: false
+#| label: fig-lwd-scurve
+#| fig-pos: H
+#| fig-cap: "(a) Identification of low-density hexagons using a benchmark value of $0.158$, and (c) Identification of low-density hexagons considering neighboring bins."
+#| fig-width: 9
+#| fig-height: 6
+
+hex_grid_rm_lwd_scurve1 + hex_grid_rm_lwd_scurve2 +
+  plot_layout(guides='collect', ncol = 2) &
+  theme(legend.position='none', plot.tag = element_text(size = 8))
 ```

### 2024-05-20 paper.tex
Added lwd plots

@@
 some hexagonal bins may have few or no data points within them (see
-\textbf{?@fig-rmlowdenshex} (a)). To ensure comprehensive coverage of
+Figure~\ref{fig-bins-scurve} (b)). To ensure comprehensive coverage of
 the NLDR data, it is necessary to select hexagonal bins with a
@@
 dividing the number of points within each hexagon by the maximum number
-of points in the grid (see \textbf{?@eq-equationp2}). Next, bins with a
-standard count less than a certain benchmark value are removed (see
-\textbf{?@fig-rmlowdenshex} (b)). The following steps will help in
-determining a suitable value for removing low-density hexagons:
+of points in the grid. Next, bins with a standard count less than a
+benchmark value are removed (see Figure~\ref{fig-lwd-scurve} (a)). There
+is no specific rule for selecting a benchmark value. However, the
+following steps can help determine a suitable value for removing
+low-density hexagons:
 
@@
   Plot the distribution of the standardized counts (see
-  \textbf{?@fig-stdctsScurve}).
+  Figure~\ref{fig-stdcts-scurve}).
 \item
@@
 
-Furthermore, selecting the benchmark value for removing low-density
-hexagons is important. Removing unnecessary bins may lead to the
-formation of long edges and an uneven 2D model. Hence, rather than
-solely relying on the benchmark value to identify hexagons for removal,
-it's essential to consider the standard number of points in the
-neighboring hexagons of the identified low-density ones (see
-\textbf{?@fig-rmlowdenshex} (c)). If neighboring bins also show low
-counts, only those bins will be removed. The remaining bins are used to
-construct the 2D model.
-
 The benchmark value for removing low-density hexagons ranges between
@@
 model performance, it's essential to observe the change in MSE as the
-benchmark value increases (see \textbf{?@fig-diagnosticpltScurvelwd}).
-The MSE shows a gradual increase as the benchmark value progresses from
-\(0\) to \(1\). Evaluating this rate of increase is important. If the
-increment is not considerable, the decision might lean towards retaining
+benchmark value increases (see Figure~\ref{fig-mse-scurve-lwd}). The MSE
+shows a gradual increase as the benchmark value progresses from \(0\) to
+\(1\). Evaluating this rate of increase is important. If the increment
+is not considerable, the decision might lean towards retaining
 low-density hexagons.
@@
 value to remove low density hexagons? The residual plot have a steep
-slope at the beginning, indicating that a smaller bechmark value causes
-a larger amount of MSE. Then, the slope gradually declines or level off,
-indicating that a higher bechmark value generates a smaller MSE. Using
+slope at the end, indicating that a smaller bechmark value causes a
+small amount of MSE. Then, the slope gradually increases or level up,
+indicating that a higher bechmark value generates a higher MSE. Using
 the elbow method, it was observed that when the bechmark value is set to
-\(0.6\), the lowest MSE occurred.}
+\(0.242\), the lowest MSE occurred.}
+
+\end{figure}%
+
+Furthermore, selecting the benchmark value for removing low-density
+hexagons is important. Removing unnecessary bins may lead to the
+formation of long edges and an uneven \gD{} model. Hence, rather than
+solely relying on the benchmark value to identify hexagons for removal,
+it's essential to consider the standard number of points in the
+neighboring hexagons of the identified low-density bins (see
+Figure~\ref{fig-lwd-scurve} (b)). If neighboring bins also show low
+counts, only those bins will be removed. The remaining bins are used to
+construct the \gD{} model.
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-lwd-scurve-1.pdf}
+
+}
+
+\caption{\label{fig-lwd-scurve}(a) Identification of low-density
+hexagons using a benchmark value of \(0.158\), and (c) Identification of
+low-density hexagons considering neighboring bins.}
 

### 2024-05-20 paper.qmd
Added long edges figures

@@
 
-<!--add vis showing removing the hexagons using first quantile, and observing neighbors-->
 <!--Using first quantile as the benchmark value-->
@@
 
+df_bin_centroids_scurve_keep  <- df_bin_centroids_scurve |>
+  filter(!(hexID %in%remove_id))
+
 hex_grid_with_low_counts_neighbor <- right_join(hex_grid, 
@@
 #| fig-pos: H
-#| fig-cap: "(a) Identification of low-density hexagons using a benchmark value of $0.158$, and (c) Identification of low-density hexagons considering neighboring bins."
+#| fig-cap: "(a) Identification of low-density hexagons using a benchmark value of $0.242$, and (c) Identification of low-density hexagons considering neighboring bins."
 #| fig-width: 9
@@
 
-<!--add distribution of distance along with the default benchmark-->
-
+<!--distribution of distance along with the default benchmark-->
 ```{r}
@@
 tr1_object_scurve <- tri_bin_centroids(
-  df_bin_centroids_scurve, x = "c_x", y = "c_y")
+  df_bin_centroids_scurve_keep, x = "c_x", y = "c_y")
 tr_from_to_df_scurve <- gen_edges(
@@
 
-<!--show the trimesh with colored long edges with different benchmark values-->
+<!--trimesh with different benchmark values-->
+
+```{r}
+## With calculated default: 0.201
+trimesh_lg_scurve1 <- vis_rmlg_mesh(
+  distance_edges = distance_scurve, 
+  benchmark_value = benchmark_scurve, 
+  tr_coord_df = tr_from_to_df_scurve, 
+  distance_col = "distance") +
+  xlim(sc_xlims) + ylim(sc_ylims) +
+  interior_annotation("a", sc_ltr_pos)
+
+trimesh_lg_scurve2 <- vis_rmlg_mesh(
+  distance_edges = distance_scurve, 
+  benchmark_value = 0.35, 
+  tr_coord_df = tr_from_to_df_scurve, 
+  distance_col = "distance") +
+  xlim(sc_xlims) + ylim(sc_ylims) +
+  interior_annotation("b", sc_ltr_pos)
+
+trimesh_lg_scurve3 <- vis_rmlg_mesh(
+  distance_edges = distance_scurve, 
+  benchmark_value = 1.1, 
+  tr_coord_df = tr_from_to_df_scurve, 
+  distance_col = "distance") +
+  xlim(sc_xlims) + ylim(sc_ylims) +
+  interior_annotation("c", sc_ltr_pos) 
+```
+
+```{r}
+#| echo: false
+#| label: fig-lg-scurve
+#| fig-pos: H
+#| fig-cap: "$2\\text{-}D$ model generated for different benchamrk values to remove long edges: (a) benchmark = $0.201$ (default), (b) benchmark = $0.35$, and (c) benchmark = $1.1$.
+What is the effective benchmark value to remove long edges?"
+#| fig-width: 6
+#| fig-height: 3
+
+trimesh_lg_scurve1 + trimesh_lg_scurve2 + trimesh_lg_scurve3 +
+  plot_layout(guides='collect', ncol = 3) &
+  theme(legend.position='none', plot.tag = element_text(size = 8))
+```
 

### 2024-05-20 paper.tex
Added long edges figures

@@
 \caption{\label{fig-lwd-scurve}(a) Identification of low-density
-hexagons using a benchmark value of \(0.158\), and (c) Identification of
+hexagons using a benchmark value of \(0.242\), and (c) Identification of
 low-density hexagons considering neighboring bins.}
@@
 S-curve UMAP data.}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-lg-scurve-1.pdf}
+
+}
+
+\caption{\label{fig-lg-scurve}\(2\text{-}D\) model generated for
+different benchamrk values to remove long edges: (a) benchmark =
+\(0.201\) (default), (b) benchmark = \(0.35\), and (c) benchmark =
+\(1.1\). What is the effective benchmark value to remove long edges?}
 

### 2024-05-20 paper.qmd
Update fitted value section

@@
 The prediction approach involves performing the K-nearest neighbors (KNN) algorithm for an unsupervised classification problem.
-First, the nearest high-D model point is identified for a given new high-D point.
-Then, the corresponding 2D centroid mapping for the identified high-D model point is determined.
-Finally, the coordinates of this 2D centroid are used as the predicted 2D embedding for the new high-D data point.
+First, the nearest \pD{} model point is identified for a given new \pD{} point.
+Then, the corresponding \gD{} centroid mapping for the identified \pD{} model point is determined.
+Finally, the coordinates of this \gD{} centroid are used as the predicted \gD{} embedding for the new \pD{} data point.
 This step is particularly valuable due to the limitations of some NLDR techniques, like tSNE, which don't provide a straightforward method for prediction.
-As a result, our approach offers a solution that capable of generating predicted 2D embedding regardless of the NLDR technique employed, effectively addressing this functional gap.
+As a result, our approach offers a solution that capable of generating predicted \gD{} embedding regardless of the NLDR technique employed, effectively addressing this functional gap.
 

### 2024-05-20 paper.tex
Update fitted value section

@@
 (KNN) algorithm for an unsupervised classification problem. First, the
-nearest high-D model point is identified for a given new high-D point.
-Then, the corresponding 2D centroid mapping for the identified high-D
-model point is determined. Finally, the coordinates of this 2D centroid
-are used as the predicted 2D embedding for the new high-D data point.
-This step is particularly valuable due to the limitations of some NLDR
-techniques, like tSNE, which don't provide a straightforward method for
-prediction. As a result, our approach offers a solution that capable of
-generating predicted 2D embedding regardless of the NLDR technique
-employed, effectively addressing this functional gap.
+nearest \pD{} model point is identified for a given new \pD{} point.
+Then, the corresponding \gD{} centroid mapping for the identified \pD{}
+model point is determined. Finally, the coordinates of this \gD{}
+centroid are used as the predicted \gD{} embedding for the new \pD{}
+data point. This step is particularly valuable due to the limitations of
+some NLDR techniques, like tSNE, which don't provide a straightforward
+method for prediction. As a result, our approach offers a solution that
+capable of generating predicted \gD{} embedding regardless of the NLDR
+technique employed, effectively addressing this functional gap.
 

### 2024-05-20 paper.qmd
Add error calculation

@@
 First, the nearest \pD{} model point is identified for a given new \pD{} point.
-Then, the corresponding \gD{} centroid mapping for the identified \pD{} model point is determined.
-Finally, the coordinates of this \gD{} centroid are used as the predicted \gD{} embedding for the new \pD{} data point.
-This step is particularly valuable due to the limitations of some NLDR techniques, like tSNE, which don't provide a straightforward method for prediction.
-As a result, our approach offers a solution that capable of generating predicted \gD{} embedding regardless of the NLDR technique employed, effectively addressing this functional gap.
+Then, the corresponding \kD{} centroid mapping for the identified \pD{} model point is determined.
+Finally, the coordinates of this \kD{} centroid are used as the predicted \kD{} embedding for the new \pD{} data point.
+.
 
@@
 
-Residuals are essential for evaluating the accuracy of representing high-D points by the high-D mapping of 2D bin centroids.
-To measure this accuracy, an error metric is introduced, quantifying the sum of squared differences between the high-D data ($x_{ij}$) and the high-D mapping of the 2D bin centroid data ($C_{x_ij}$) across all observations and dimensions (see @eq-equation10).
-
-$$
-\text{Error} = \sum_{j = 1}^{n}\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2
-$$ {#eq-equation10}
-
-Here, $n$ represents the number of observations, $p$ represents the dimensions of high-D data, $x_{ij}$ is the high-D data, and $C_{x_ij}$ is the high-D mapping of the 2D bin centroid.
-
-To assess how well our method captures and represents the underlying structure of the high-D data, Mean Squared Error (MSE) is used.
-When computing MSE, total model error (see @sec-summary) is divided by the number of observations to make it as a mean value (see @eq-equation9).
-
-$$
-\text{MSE} = \sum_{j = 1}^{n} \frac{\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2}{n}
-$$ {#eq-equation9}
+To assess how well our method captures and represents the underlying structure of the \pD{} data, residuals are important.
+Residuals are computed by taking the squared \pD{} Euclidean distance, $\sum_{i = 1}^{n}\sum_{j = 1}^{p} (x_{ij} - C^{(p)}_{x_ij})^2$.
+Additionally, Mean Squared Error (MSE), $\sum_{i = 1}^{n}\sum_{j = 1}^{p} \frac{(x_{ij} - C^{(p)}_{x_ij})^2}{n}$, is computed based on residuals.
 

### 2024-05-20 paper.tex
Add error calculation

@@
 nearest \pD{} model point is identified for a given new \pD{} point.
-Then, the corresponding \gD{} centroid mapping for the identified \pD{}
-model point is determined. Finally, the coordinates of this \gD{}
-centroid are used as the predicted \gD{} embedding for the new \pD{}
-data point. This step is particularly valuable due to the limitations of
-some NLDR techniques, like tSNE, which don't provide a straightforward
-method for prediction. As a result, our approach offers a solution that
-capable of generating predicted \gD{} embedding regardless of the NLDR
-technique employed, effectively addressing this functional gap.
+Then, the corresponding \kD{} centroid mapping for the identified \pD{}
+model point is determined. Finally, the coordinates of this \kD{}
+centroid are used as the predicted \kD{} embedding for the new \pD{}
+data point. .
 
@@
 
-Residuals are essential for evaluating the accuracy of representing
-high-D points by the high-D mapping of 2D bin centroids. To measure this
-accuracy, an error metric is introduced, quantifying the sum of squared
-differences between the high-D data (\(x_{ij}\)) and the high-D mapping
-of the 2D bin centroid data (\(C_{x_ij}\)) across all observations and
-dimensions (see Equation~\ref{eq-equation10}).
-
-\begin{equation}\phantomsection\label{eq-equation10}{
-\text{Error} = \sum_{j = 1}^{n}\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2
-}\end{equation}
-
-Here, \(n\) represents the number of observations, \(p\) represents the
-dimensions of high-D data, \(x_{ij}\) is the high-D data, and
-\(C_{x_ij}\) is the high-D mapping of the 2D bin centroid.
-
 To assess how well our method captures and represents the underlying
-structure of the high-D data, Mean Squared Error (MSE) is used. When
-computing MSE, total model error (see Section~\ref{sec-summary}) is
-divided by the number of observations to make it as a mean value (see
-Equation~\ref{eq-equation9}).
-
-\begin{equation}\phantomsection\label{eq-equation9}{
-\text{MSE} = \sum_{j = 1}^{n} \frac{\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2}{n}
-}\end{equation}
+structure of the \pD{} data, residuals are important. Residuals are
+computed by taking the squared \pD{} Euclidean distance,
+\(\sum_{i = 1}^{n}\sum_{j = 1}^{p} (x_{ij} - C^{(p)}_{x_ij})^2\).
+Additionally, Mean Squared Error (MSE),
+\(\sum_{i = 1}^{n}\sum_{j = 1}^{p} \frac{(x_{ij} - C^{(p)}_{x_ij})^2}{n}\),
+is computed based on residuals.
 

### 2024-05-20 paper.qmd
Added scripts to generate the model in data space

@@
 
+benchmark_label_df <- tibble(y = benchmark_scurve - 0.03,
+                             x = 1.4,
+                             label = paste0("benchmark is ", benchmark_scurve))
+
 ## To draw the distance distribution
@@
              color = "black", linewidth=0.8, alpha = 0.5) +
+  geom_text(data = benchmark_label_df, 
+            aes(x=x, 
+                label= label, 
+                y=y), 
+            colour="black", 
+            angle=90) +
   ylab(expression(d^{(2)})) +
@@
 ```
-
-To investigate which is the reasonable representation to visualize the five spherical Gaussian cluster data or all NLDR methods provide equivalent information, we visualize all the models in \pD{} space.
-Models from tSNE, UMAP, and PHATE show five well-separated clusters (see @fig-gau1_sc, @fig-gau2_sc, and @fig-gau3_sc), while PaCMAP and TriMAP have a long edge that connects neighboring two clusters and three well-separated clusters (see @fig-gau4_sc, and @fig-gau5_sc).
-This suggests that for the five Gaussian cluster dataset, all NLDR methods effectively preserve the global structure.
-
-Models from all NLDR methods show five well-separated clusters (see @fig-gau1_sc, @fig-gau2_sc, @fig-gau3_sc, @fig-gau4_sc, and @fig-gau5_sc).
-This suggests that for the five Gaussian cluster dataset, all NLDR methods effectively preserve the global structure.
-tSNE, UMAP, and PHATE display clusters with varying densities, indicating their ability to capture within-cluster variation (see @fig-gau1_sc, @fig-gau2_sc, and @fig-gau3_sc).
-On the other hand, both TriMAP and PaCMAP show clusters with flat surfaces, suggesting a failure to capture within-cluster variation (see @fig-gau4_sc and @fig-gau5_sc).
-Therefore, TriMAP and PaCMAP do not capture the local structure as effectively as other methods.
-
-When compare the NLDR representations and generates models, tSNE with perplexity $61$ is the reasonable representation to visualize the five Gaussian cluster dataset.
 
@@
 gau4_scaled_obj <- gen_scaled_data(
-  data = umap_data_gau)
+  data = pacmap_data_gau)
 pacmap_gau_scaled <- gau4_scaled_obj$scaled_nldr
@@
   distance_col = "distance")
+benchmark_gau5 <- 0.2
 
@@
 #| echo: false
-#| fig-cap: "Model generated with five diiferent NLDR methods in $2\\text{-}D$ with approximately $65$ non-empty bins."
+#| fig-cap: "Model generated with five different NLDR methods in $2\\text{-}D$ with approximately $65$ non-empty bins in each."
 #| fig-width: 8
@@
 <!--add three screenshots from each technique with you tube links-->
+
+To investigate which is the reasonable representation to visualize the five spherical Gaussian cluster data or all NLDR methods provide equivalent information, we visualize all the models in \pD{} space.
+Models from all NLDR methods show five well-separated clusters (see @fig-gau1_sc, @fig-gau2_sc, @fig-gau3_sc, @fig-gau4_sc, and @fig-gau5_sc).
+This suggests that for the five Gaussian cluster dataset, all NLDR methods effectively preserve the global structure.
+tSNE, UMAP, and PHATE display clusters with varying densities, indicating their ability to capture within-cluster variation (see @fig-gau1_sc, @fig-gau2_sc, and @fig-gau3_sc).
+On the other hand, both TriMAP and PaCMAP show clusters with flat surfaces, suggesting a failure to capture within-cluster variation (see @fig-gau4_sc and @fig-gau5_sc).
+Therefore, TriMAP and PaCMAP do not capture the local structure as effectively as other methods.
+
+::: {#fig-gau1_sc layout-ncol="3" fig-pos="H"}
+![](figures/five_gau_clusters/sc_tsne_1.png){#fig-gau1_sc1}
+
+![](figures/five_gau_clusters/sc_tsne_2.png){#fig-gau1_sc2}
+
+![](figures/five_gau_clusters/sc_tsne_3.png){#fig-gau1_sc3}
+
+Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with tSNE in high-D, a video of the tour animation is available at (<https://youtu.be/oQxEb4wRdHI>).
+:::
+
+::: {#fig-gau2_sc layout-ncol="3" fig-pos="H"}
+![](figures/five_gau_clusters/sc_umap_1.png){#fig-gau2_sc1}
+
+![](figures/five_gau_clusters/sc_umap_2.png){#fig-gau2_sc2}
+
+![](figures/five_gau_clusters/sc_umap_3.png){#fig-gau2_sc3}
+
+Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with UMAP in high-D, a video of the tour animation is available at (<https://youtu.be/JW49csPpDx4>).
+:::
+
+::: {#fig-gau3_sc layout-ncol="3" fig-pos="H"}
+![](figures/five_gau_clusters/sc_phate_1.png){#fig-gau3_sc1}
+
+![](figures/five_gau_clusters/sc_phate_2.png){#fig-gau3_sc2}
+
+![](figures/five_gau_clusters/sc_phate_3.png){#fig-gau3_sc3}
+
+Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with PHATE in high-D, a video of the tour animation is available at (<https://youtu.be/xqhRODt8Z-o>).
+:::
+
+::: {#fig-gau4_sc layout-ncol="3" fig-pos="H"}
+![](figures/five_gau_clusters/sc_trimap_1.png){#fig-gau4_sc1}
+
+![](figures/five_gau_clusters/sc_trimap_2.png){#fig-gau4_sc2}
+
+![](figures/five_gau_clusters/sc_trimap_3.png){#fig-gau4_sc3}
+
+Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with TriMAP in high-D, a video of the tour animation is available at (<https://youtu.be/X-uHSB4VoZ0>).
+:::
+
+::: {#fig-gau5_sc layout-ncol="3" fig-pos="H"}
+![](figures/five_gau_clusters/sc_pacmap_1.png){#fig-gau5_sc1}
+
+![](figures/five_gau_clusters/sc_pacmap_2.png){#fig-gau5_sc2}
+
+![](figures/five_gau_clusters/sc_pacmap_3.png){#fig-gau5_sc3}
+
+Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with PaCMAP in high-D, a video of the tour animation is available at (<https://youtu.be/UxvU8HQd7JM>).
+:::
+
+When compare the NLDR representations and generated models, tSNE with perplexity $61$ appears to be a reasonable representation for visualizing the five Gaussian cluster dataset.
+This is supported by investigating the model generated with tSNE in the data space, which provides evidence that it preserves both local and global structures.
+Also, the NLDR representation with tSNE shows five well-separated clusters.
 

### 2024-05-20 paper.tex
Added scripts to generate the model in data space

@@
 
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-trimesh-gau-1.pdf}
+
+}
+
+\caption{\label{fig-trimesh-gau}Model generated with five different NLDR
+methods in \(2\text{-}D\) with approximately \(65\) non-empty bins in
+each.}
+
+\end{figure}%
+
 To investigate which is the reasonable representation to visualize the
@@
 equivalent information, we visualize all the models in \pD{} space.
-Models from tSNE, UMAP, and PHATE show five well-separated clusters (see
-\textbf{?@fig-gau1\_sc}, \textbf{?@fig-gau2\_sc}, and
-\textbf{?@fig-gau3\_sc}), while PaCMAP and TriMAP have a long edge that
-connects neighboring two clusters and three well-separated clusters (see
-\textbf{?@fig-gau4\_sc}, and \textbf{?@fig-gau5\_sc}). This suggests
-that for the five Gaussian cluster dataset, all NLDR methods effectively
-preserve the global structure.
-
 Models from all NLDR methods show five well-separated clusters (see
-\textbf{?@fig-gau1\_sc}, \textbf{?@fig-gau2\_sc},
-\textbf{?@fig-gau3\_sc}, \textbf{?@fig-gau4\_sc}, and
-\textbf{?@fig-gau5\_sc}). This suggests that for the five Gaussian
+Figure~\ref{fig-gau1_sc}, Figure~\ref{fig-gau2_sc},
+Figure~\ref{fig-gau3_sc}, Figure~\ref{fig-gau4_sc}, and
+Figure~\ref{fig-gau5_sc}). This suggests that for the five Gaussian
 cluster dataset, all NLDR methods effectively preserve the global
@@
 densities, indicating their ability to capture within-cluster variation
-(see \textbf{?@fig-gau1\_sc}, \textbf{?@fig-gau2\_sc}, and
-\textbf{?@fig-gau3\_sc}). On the other hand, both TriMAP and PaCMAP show
-clusters with flat surfaces, suggesting a failure to capture
-within-cluster variation (see \textbf{?@fig-gau4\_sc} and
-\textbf{?@fig-gau5\_sc}). Therefore, TriMAP and PaCMAP do not capture
+(see Figure~\ref{fig-gau1_sc}, Figure~\ref{fig-gau2_sc}, and
+Figure~\ref{fig-gau3_sc}). On the other hand, both TriMAP and PaCMAP
+show clusters with flat surfaces, suggesting a failure to capture
+within-cluster variation (see Figure~\ref{fig-gau4_sc} and
+Figure~\ref{fig-gau5_sc}). Therefore, TriMAP and PaCMAP do not capture
 the local structure as effectively as other methods.
 
-When compare the NLDR representations and generates models, tSNE with
-perplexity \(61\) is the reasonable representation to visualize the five
-Gaussian cluster dataset.
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-trimesh-gau-1.pdf}
-
-}
-
-\caption{\label{fig-trimesh-gau}Model generated with five diiferent NLDR
-methods in \(2\text{-}D\) with approximately \(65\) non-empty bins.}
-
-\end{figure}%
+\begin{figure}[H]
+
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_tsne_1.png}
+
+}
+
+\subcaption{\label{fig-gau1_sc1}}
+
+\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_tsne_2.png}
+
+}
+
+\subcaption{\label{fig-gau1_sc2}}
+
+\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_tsne_3.png}
+
+}
+
+\subcaption{\label{fig-gau1_sc3}}
+
+\end{minipage}%
+
+\caption{\label{fig-gau1_sc}Screen shots of the \textbf{langevitour} of
+the five Gaussian clusters dataset, shows the model with tSNE in high-D,
+a video of the tour animation is available at
+(\url{https://youtu.be/oQxEb4wRdHI}).}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_umap_1.png}
+
+}
+
+\subcaption{\label{fig-gau2_sc1}}
+
+\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_umap_2.png}
+
+}
+
+\subcaption{\label{fig-gau2_sc2}}
+
+\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_umap_3.png}
+
+}
+
+\subcaption{\label{fig-gau2_sc3}}
+
+\end{minipage}%
+
+\caption{\label{fig-gau2_sc}Screen shots of the \textbf{langevitour} of
+the five Gaussian clusters dataset, shows the model with UMAP in high-D,
+a video of the tour animation is available at
+(\url{https://youtu.be/JW49csPpDx4}).}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_phate_1.png}
+
+}
+
+\subcaption{\label{fig-gau3_sc1}}
+
+\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_phate_2.png}
+
+}
+
+\subcaption{\label{fig-gau3_sc2}}
+
+\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_phate_3.png}
+
+}
+
+\subcaption{\label{fig-gau3_sc3}}
+
+\end{minipage}%
+
+\caption{\label{fig-gau3_sc}Screen shots of the \textbf{langevitour} of
+the five Gaussian clusters dataset, shows the model with PHATE in
+high-D, a video of the tour animation is available at
+(\url{https://youtu.be/xqhRODt8Z-o}).}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_trimap_1.png}
+
+}
+
+\subcaption{\label{fig-gau4_sc1}}
+
+\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_trimap_2.png}
+
+}
+
+\subcaption{\label{fig-gau4_sc2}}
+
+\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_trimap_3.png}
+
+}
+
+\subcaption{\label{fig-gau4_sc3}}
+
+\end{minipage}%
+
+\caption{\label{fig-gau4_sc}Screen shots of the \textbf{langevitour} of
+the five Gaussian clusters dataset, shows the model with TriMAP in
+high-D, a video of the tour animation is available at
+(\url{https://youtu.be/X-uHSB4VoZ0}).}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_pacmap_1.png}
+
+}
+
+\subcaption{\label{fig-gau5_sc1}}
+
+\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_pacmap_2.png}
+
+}
+
+\subcaption{\label{fig-gau5_sc2}}
+
+\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/five_gau_clusters/sc_pacmap_3.png}
+
+}
+
+\subcaption{\label{fig-gau5_sc3}}
+
+\end{minipage}%
+
+\caption{\label{fig-gau5_sc}Screen shots of the \textbf{langevitour} of
+the five Gaussian clusters dataset, shows the model with PaCMAP in
+high-D, a video of the tour animation is available at
+(\url{https://youtu.be/UxvU8HQd7JM}).}
+
+\end{figure}%
+
+When compare the NLDR representations and generated models, tSNE with
+perplexity \(61\) appears to be a reasonable representation for
+visualizing the five Gaussian cluster dataset. This is supported by
+investigating the model generated with tSNE in the data space, which
+provides evidence that it preserves both local and global structures.
+Also, the NLDR representation with tSNE shows five well-separated
+clusters.
 

### 2024-05-20 paper.qmd
Update the screen shots and videos of simulated data example

@@
 
-Let $h : \mathbb{R}^{n\times k} \rightarrow \mathbb{R}^{m\times k}$ be the function that maps a point in \kD{} space to it's bin centroid ($C^{(k)}$). This is the model points in \kD{} space.
-
-When $k = 2$, one of the representations of a bin is the hexagonal centroid (@fig-NLDR-scurve (c)).
+Let $g : \mathbb{R}^{n\times k} \rightarrow \mathbb{R}^{m\times k}$ be the function that maps a point in \kD{} space to it's bin centroid ($C^{(k)}$). This is the model points in \kD{} space.
+
+When $k = 2$, one of the representations of a bin is the hexagonal centroid (@fig-NLDR-scurve (c)). Also, it could be the mean of the data points within each hexagon.
 
@@
 
+When neighboring centroids are connected with edges in Delaunay triangulation in \gD{}, there are situations where distant centroids are also connected, which generates long edges.
+To generate a smooth surface in \gD{}, these long edges are removed (@fig-NLDR-scurve (d)).
+
 <!--need to add what is meant by a long edge-->
@@
 Define the set $H^k$ to be the set of points that belong to bin $k$.
-We use the \pD{} Euclidean mean of the points in $H^k$ to map the centroid $(h_x^k, h_y^k)$ to a point in \pD{}.
+We use the \pD{} Euclidean mean of the points in $H^k$ to map the centroid $C_{h}^{(k)} = (c_{h1}, c_{h2})$ to a point in \pD{}.
 Let the $i$-th component of the \pD{} mean be
@@
 
-with corresponding \pD{} vector $\vec{f_i}$. Then, $f(h\circ g)(x)$ maps a \pD{} point to the \pD{} model estimate, completing the model cycle.
+with corresponding \pD{} vector $\vec{f_i}$. Then, $f(g(x))$ maps a \pD{} point to the \pD{} model estimate, completing the model cycle.
 
@@
 ::: {#fig-scurve_sc layout-ncol="3" fig-pos="H"}
-![](figures/scurve/sc_1.png){#fig-scurve_sc1}
-
-![](figures/scurve/sc_2.png){#fig-scurve_sc2}
-
-![](figures/scurve/sc_3.png){#fig-scurve_sc3}
+![](figures/scurve/sc_1.png)
+
+![](figures/scurve/sc_2.png)
+
+![](figures/scurve/sc_3.png)
 
@@
 This suggests that for the five Gaussian cluster dataset, all NLDR methods effectively preserve the global structure.
-tSNE, UMAP, and PHATE display clusters with varying densities, indicating their ability to capture within-cluster variation (see @fig-gau1_sc, @fig-gau2_sc, and @fig-gau3_sc).
-On the other hand, both TriMAP and PaCMAP show clusters with flat surfaces, suggesting a failure to capture within-cluster variation (see @fig-gau4_sc and @fig-gau5_sc).
-Therefore, TriMAP and PaCMAP do not capture the local structure as effectively as other methods.
+tSNE displays clusters with varying densities, indicating their ability to capture within-cluster variation (see @fig-gau1_sc).
+On the other hand, both UMAP, PHATE, PaCMAP and TriMAP show clusters with flat surfaces, suggesting a failure to capture within-cluster variation (see @fig-gau2_sc, @fig-gau3_sc, @fig-gau4_sc and @fig-gau5_sc).
+Therefore, UMAP, PHATE, PaCMAP and TriMAP do not capture the local structure as effectively as other methods.
 
 ::: {#fig-gau1_sc layout-ncol="3" fig-pos="H"}
-![](figures/five_gau_clusters/sc_tsne_1.png){#fig-gau1_sc1}
-
-![](figures/five_gau_clusters/sc_tsne_2.png){#fig-gau1_sc2}
-
-![](figures/five_gau_clusters/sc_tsne_3.png){#fig-gau1_sc3}
-
-Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with tSNE in high-D, a video of the tour animation is available at (<https://youtu.be/oQxEb4wRdHI>).
+![](figures/five_gau_clusters/sc_tsne_1.png)
+
+![](figures/five_gau_clusters/sc_tsne_2.png)
+
+![](figures/five_gau_clusters/sc_tsne_3.png)
+
+Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with tSNE in \pD{}, a video of the tour animation is available at (<https://youtu.be/RASEE7N5MbM>).
 :::
@@
 ::: {#fig-gau2_sc layout-ncol="3" fig-pos="H"}
-![](figures/five_gau_clusters/sc_umap_1.png){#fig-gau2_sc1}
-
-![](figures/five_gau_clusters/sc_umap_2.png){#fig-gau2_sc2}
-
-![](figures/five_gau_clusters/sc_umap_3.png){#fig-gau2_sc3}
-
-Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with UMAP in high-D, a video of the tour animation is available at (<https://youtu.be/JW49csPpDx4>).
+![](figures/five_gau_clusters/sc_umap_1.png)
+
+![](figures/five_gau_clusters/sc_umap_2.png)
+
+![](figures/five_gau_clusters/sc_umap_3.png)
+
+Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with UMAP in \pD{}, a video of the tour animation is available at (<https://youtu.be/iG4bCPkJilw>).
 :::
@@
 ::: {#fig-gau3_sc layout-ncol="3" fig-pos="H"}
-![](figures/five_gau_clusters/sc_phate_1.png){#fig-gau3_sc1}
-
-![](figures/five_gau_clusters/sc_phate_2.png){#fig-gau3_sc2}
-
-![](figures/five_gau_clusters/sc_phate_3.png){#fig-gau3_sc3}
-
-Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with PHATE in high-D, a video of the tour animation is available at (<https://youtu.be/xqhRODt8Z-o>).
+![](figures/five_gau_clusters/sc_phate_1.png)
+
+![](figures/five_gau_clusters/sc_phate_2.png)
+
+![](figures/five_gau_clusters/sc_phate_3.png)
+
+Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with PHATE in \pD{}, a video of the tour animation is available at (<https://youtu.be/L_PVLGwfOS0>).
 :::
@@
 ::: {#fig-gau4_sc layout-ncol="3" fig-pos="H"}
-![](figures/five_gau_clusters/sc_trimap_1.png){#fig-gau4_sc1}
-
-![](figures/five_gau_clusters/sc_trimap_2.png){#fig-gau4_sc2}
-
-![](figures/five_gau_clusters/sc_trimap_3.png){#fig-gau4_sc3}
-
-Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with TriMAP in high-D, a video of the tour animation is available at (<https://youtu.be/X-uHSB4VoZ0>).
+![](figures/five_gau_clusters/sc_pacmap_1.png)
+
+![](figures/five_gau_clusters/sc_pacmap_2.png)
+
+![](figures/five_gau_clusters/sc_pacmap_3.png)
+
+Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with PaCMAP in \pD{}, a video of the tour animation is available at (<https://youtu.be/z07cKXi8EJQ>).
 :::
@@
 ::: {#fig-gau5_sc layout-ncol="3" fig-pos="H"}
-![](figures/five_gau_clusters/sc_pacmap_1.png){#fig-gau5_sc1}
-
-![](figures/five_gau_clusters/sc_pacmap_2.png){#fig-gau5_sc2}
-
-![](figures/five_gau_clusters/sc_pacmap_3.png){#fig-gau5_sc3}
-
-Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with PaCMAP in high-D, a video of the tour animation is available at (<https://youtu.be/UxvU8HQd7JM>).
+![](figures/five_gau_clusters/sc_trimap_1.png)
+
+![](figures/five_gau_clusters/sc_trimap_2.png)
+
+![](figures/five_gau_clusters/sc_trimap_3.png)
+
+Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with TriMAP in \pD{}, a video of the tour animation is available at (<https://youtu.be/Chs1lYAoX2w>).
 :::
@@
 ::: {#fig-pbmc1_sc layout-ncol="3" fig-pos="H"}
-![](figures/pbmc3k/sc_1.png){#fig-pbmc1_sc1}
-
-![](figures/pbmc3k/sc_2.png){#fig-pbmc1_sc2}
-
-![](figures/pbmc3k/sc_3.png){#fig-pbmc1_sc3}
+![](figures/pbmc3k/sc_1.png)
+
+![](figures/pbmc3k/sc_2.png)
+
+![](figures/pbmc3k/sc_3.png)
 
@@
 ::: {#fig-pbmc2_sc layout-ncol="3" fig-pos="H"}
-![](figures/pbmc3k/sc_4.png){#fig-pbmc2_sc4}
-
-![](figures/pbmc3k/sc_5.png){#fig-pbmc2_sc5}
-
-![](figures/pbmc3k/sc_6.png){#fig-pbmc2_sc6}
+![](figures/pbmc3k/sc_4.png)
+
+![](figures/pbmc3k/sc_5.png)
+
+![](figures/pbmc3k/sc_6.png)
 
@@
 ::: {#fig-mnist1_sc layout-ncol="3" fig-pos="H"}
-![](figures/mnist/sc_1.png){#fig-mnist1_sc1}
-
-![](figures/mnist/sc_2.png){#fig-mnist1_sc2}
-
-![](figures/mnist/sc_3.png){#fig-mnist1_sc3}
+![](figures/mnist/sc_1.png)
+
+![](figures/mnist/sc_2.png)
+
+![](figures/mnist/sc_3.png)
 

### 2024-05-20 paper.tex
Update the screen shots and videos of simulated data example

--- 2024-05-20 paper.tex
@@
 
-Let \(h : \mathbb{R}^{n\times k} \rightarrow \mathbb{R}^{m\times k}\) be
+Let \(g : \mathbb{R}^{n\times k} \rightarrow \mathbb{R}^{m\times k}\) be
 the function that maps a point in \kD{} space to it's bin centroid
@@
 When \(k = 2\), one of the representations of a bin is the hexagonal
-centroid (Figure~\ref{fig-NLDR-scurve} (c)).
+centroid (Figure~\ref{fig-NLDR-scurve} (c)). Also, it could be the mean
+of the data points within each hexagon.
 
@@
 
+When neighboring centroids are connected with edges in Delaunay
+triangulation in \gD{}, there are situations where distant centroids are
+also connected, which generates long edges. To generate a smooth surface
+in \gD{}, these long edges are removed (Figure~\ref{fig-NLDR-scurve}
+(d)).
+
 \subsection{\texorpdfstring{Displaying the model in
@@
 set of points that belong to bin \(k\). We use the \pD{} Euclidean mean
-of the points in \(H^k\) to map the centroid \((h_x^k, h_y^k)\) to a
-point in \pD{}. Let the \(i\)-th component of the \pD{} mean be
+of the points in \(H^k\) to map the centroid
+\(C_{h}^{(k)} = (c_{h1}, c_{h2})\) to a point in \pD{}. Let the \(i\)-th
+component of the \pD{} mean be
 
@@
 
-with corresponding \pD{} vector \(\vec{f_i}\). Then, \(f(h\circ g)(x)\)
-maps a \pD{} point to the \pD{} model estimate, completing the model
-cycle.
+with corresponding \pD{} vector \(\vec{f_i}\). Then, \(f(g(x))\) maps a
+\pD{} point to the \pD{} model estimate, completing the model cycle.
 
@@
 \begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/scurve/sc_1.png}
-
-}
-
-\subcaption{\label{fig-scurve_sc1}}
-
-\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/scurve/sc_2.png}
-
-}
-
-\subcaption{\label{fig-scurve_sc2}}
-
-\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/scurve/sc_3.png}
-
-}
-
-\subcaption{\label{fig-scurve_sc3}}
-
-\end{minipage}%
+\includegraphics{figures/scurve/sc_1.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/scurve/sc_2.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/scurve/sc_3.png}\end{minipage}%
 
@@
 equivalent information, we visualize all the models in \pD{} space.
-Models from tSNE, UMAP, and PHATE show five well-separated clusters (see
-\textbf{?@fig-gau1\_sc}, \textbf{?@fig-gau2\_sc}, and
-\textbf{?@fig-gau3\_sc}), while PaCMAP and TriMAP have a long edge that
-connects neighboring two clusters and three well-separated clusters (see
-\textbf{?@fig-gau4\_sc}, and \textbf{?@fig-gau5\_sc}). This suggests
-that for the five Gaussian cluster dataset, all NLDR methods effectively
-preserve the global structure.
-
 Models from all NLDR methods show five well-separated clusters (see
-\textbf{?@fig-gau1\_sc}, \textbf{?@fig-gau2\_sc},
-\textbf{?@fig-gau3\_sc}, \textbf{?@fig-gau4\_sc}, and
-\textbf{?@fig-gau5\_sc}). This suggests that for the five Gaussian
+Figure~\ref{fig-gau1_sc}, Figure~\ref{fig-gau2_sc},
+Figure~\ref{fig-gau3_sc}, Figure~\ref{fig-gau4_sc}, and
+Figure~\ref{fig-gau5_sc}). This suggests that for the five Gaussian
 cluster dataset, all NLDR methods effectively preserve the global
-structure. tSNE, UMAP, and PHATE display clusters with varying
-densities, indicating their ability to capture within-cluster variation
-(see \textbf{?@fig-gau1\_sc}, \textbf{?@fig-gau2\_sc}, and
-\textbf{?@fig-gau3\_sc}). On the other hand, both TriMAP and PaCMAP show
-clusters with flat surfaces, suggesting a failure to capture
-within-cluster variation (see \textbf{?@fig-gau4\_sc} and
-\textbf{?@fig-gau5\_sc}). Therefore, TriMAP and PaCMAP do not capture
-the local structure as effectively as other methods.
-
-When compare the NLDR representations and generates models, tSNE with
-perplexity \(61\) is the reasonable representation to visualize the five
-Gaussian cluster dataset.
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-trimesh-gau-1.pdf}
-
-}
-
-\caption{\label{fig-trimesh-gau}Model generated with five diiferent NLDR
-methods in \(2\text{-}D\) with approximately \(65\) non-empty bins.}
-
-\end{figure}%
+structure. tSNE displays clusters with varying densities, indicating
+their ability to capture within-cluster variation (see
+Figure~\ref{fig-gau1_sc}). On the other hand, both UMAP, PHATE, PaCMAP
+and TriMAP show clusters with flat surfaces, suggesting a failure to
+capture within-cluster variation (see Figure~\ref{fig-gau2_sc},
+Figure~\ref{fig-gau3_sc}, Figure~\ref{fig-gau4_sc} and
+Figure~\ref{fig-gau5_sc}). Therefore, UMAP, PHATE, PaCMAP and TriMAP do
+not capture the local structure as effectively as other methods.
+
+\begin{figure}[H]
+
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/five_gau_clusters/sc_tsne_1.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/five_gau_clusters/sc_tsne_2.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/five_gau_clusters/sc_tsne_3.png}\end{minipage}%
+
+\caption{\label{fig-gau1_sc}Screen shots of the \textbf{langevitour} of
+the five Gaussian clusters dataset, shows the model with tSNE in \pD{},
+a video of the tour animation is available at
+(\url{https://youtu.be/RASEE7N5MbM}).}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/five_gau_clusters/sc_umap_1.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/five_gau_clusters/sc_umap_2.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/five_gau_clusters/sc_umap_3.png}\end{minipage}%
+
+\caption{\label{fig-gau2_sc}Screen shots of the \textbf{langevitour} of
+the five Gaussian clusters dataset, shows the model with UMAP in \pD{},
+a video of the tour animation is available at
+(\url{https://youtu.be/iG4bCPkJilw}).}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/five_gau_clusters/sc_phate_1.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/five_gau_clusters/sc_phate_2.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/five_gau_clusters/sc_phate_3.png}\end{minipage}%
+
+\caption{\label{fig-gau3_sc}Screen shots of the \textbf{langevitour} of
+the five Gaussian clusters dataset, shows the model with PHATE in \pD{},
+a video of the tour animation is available at
+(\url{https://youtu.be/L_PVLGwfOS0}).}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/five_gau_clusters/sc_pacmap_1.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/five_gau_clusters/sc_pacmap_2.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/five_gau_clusters/sc_pacmap_3.png}\end{minipage}%
+
+\caption{\label{fig-gau4_sc}Screen shots of the \textbf{langevitour} of
+the five Gaussian clusters dataset, shows the model with PaCMAP in
+\pD{}, a video of the tour animation is available at
+(\url{https://youtu.be/z07cKXi8EJQ}).}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/five_gau_clusters/sc_trimap_1.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/five_gau_clusters/sc_trimap_2.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/five_gau_clusters/sc_trimap_3.png}\end{minipage}%
+
+\caption{\label{fig-gau5_sc}Screen shots of the \textbf{langevitour} of
+the five Gaussian clusters dataset, shows the model with TriMAP in
+\pD{}, a video of the tour animation is available at
+(\url{https://youtu.be/Chs1lYAoX2w}).}
+
+\end{figure}%
+
+When compare the NLDR representations and generated models, tSNE with
+perplexity \(61\) appears to be a reasonable representation for
+visualizing the five Gaussian cluster dataset. This is supported by
+investigating the model generated with tSNE in the data space, which
+provides evidence that it preserves both local and global structures.
+Also, the NLDR representation with tSNE shows five well-separated
+clusters.
 
@@
 \begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_1.png}
-
-}
-
-\subcaption{\label{fig-pbmc1_sc1}}
-
-\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_2.png}
-
-}
-
-\subcaption{\label{fig-pbmc1_sc2}}
-
-\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_3.png}
-
-}
-
-\subcaption{\label{fig-pbmc1_sc3}}
-
-\end{minipage}%
+\includegraphics{figures/pbmc3k/sc_1.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/pbmc3k/sc_2.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/pbmc3k/sc_3.png}\end{minipage}%
 
@@
 \begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_4.png}
-
-}
-
-\subcaption{\label{fig-pbmc2_sc4}}
-
-\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_5.png}
-
-}
-
-\subcaption{\label{fig-pbmc2_sc5}}
-
-\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/pbmc3k/sc_6.png}
-
-}
-
-\subcaption{\label{fig-pbmc2_sc6}}
-
-\end{minipage}%
+\includegraphics{figures/pbmc3k/sc_4.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/pbmc3k/sc_5.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/pbmc3k/sc_6.png}\end{minipage}%
 
@@
 
-According to Figure~\ref{fig-mnist1_sc1}, the non-linear continuous
+According to \textbf{?@fig-mnist1\_sc1}, the non-linear continuous
 structure observed in the \gD{} representation of PaCMAP (see
@@
 a twisted pattern within the non-linear structure in \pD{} space (see
-Figure~\ref{fig-mnist1_sc2}), which is an additional pattern not visible
+\textbf{?@fig-mnist1\_sc2}), which is an additional pattern not visible
 in the \gD{} representation (see Figure~\ref{fig-pacmap-author}).
-Furthermore, as shown in Figure~\ref{fig-mnist1_sc3}, some long edges
+Furthermore, as shown in \textbf{?@fig-mnist1\_sc3}, some long edges
 exist in the \pD{} space that are not recognized as long edges in the
@@
 \begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/mnist/sc_1.png}
-
-}
-
-\subcaption{\label{fig-mnist1_sc1}}
-
-\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/mnist/sc_2.png}
-
-}
-
-\subcaption{\label{fig-mnist1_sc2}}
-
-\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{figures/mnist/sc_3.png}
-
-}
-
-\subcaption{\label{fig-mnist1_sc3}}
-
-\end{minipage}%
+\includegraphics{figures/mnist/sc_1.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/mnist/sc_2.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/mnist/sc_3.png}\end{minipage}%
 

### 2024-05-20 supplementary/hexagon_binning_algorithm.qmd
Add bin assign plot 1

@@
 library(ggforce)
+library(patchwork)
 
@@
 ```{r}
-all_centroids_df <- gen_centroids(bin1 = 6, q = 0.1, r2 = r2)
+all_centroids_df <- gen_centroids(bin1 = 6, r2 = r2)
 ```
@@
     geom_point(data = all_centroids_df |> dplyr::filter(hexID == 1), aes(x = c_x, y = c_y), color = "black") + ## starting point
-    geom_rect(aes(xmin = 0, xmax = 1, ymin = 0, ymax = r2), fill = "white", color = "black", alpha = 0, linewidth = 1) +
-    geom_segment(x = 0, y = -0.35, xend = 1, yend = -0.35,
+    geom_rect(aes(xmin = 0 - 0.03, xmax = 1 - 0.03, ymin = 0 - 0.05, ymax = r2 - 0.05), fill = "white", color = "black", alpha = 0, linewidth = 1) +
+    geom_segment(x = 0 - 0.03, y = -0.35, xend = 1 - 0.03, yend = -0.35,
                  arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+
     annotate("text", x=0.5, y=-0.45, label= "r1", color = "black") +
-    geom_segment(x = -0.25, y = 0, xend = -0.25, yend = r2,
+    geom_segment(x = -0.25, y = 0 - 0.05, xend = -0.25, yend = r2 - 0.05,
                  arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ 
@@
     annotate("text", x=-0.15, y=0.45, label= "a2", color = "black") +
-    annotate("text", x=-0.2, y=-0.35, label= "(s1, s2)", color = "black") + ## starting point
+    annotate("text", x=-0.18, y=-0.35, label= "(s1, s2)", color = "black") + ## starting point
   coord_equal()
@@
 #| fig-cap: "Notations for hexagonal grid configurations."
+#| out-height: 30%
+#| fig-pos: H
 
@@
 ```{r}
+all_centroids_df <- gen_centroids(bin1 = 3, r2 = r2)
+
+hex_grid <- gen_hex_coord(all_centroids_df, a1 = 0.5218947)
+
+## Data to generate circle
+circles <- data.frame(
+  x = c(0.422, 0.683),
+  y = c(0.740, 1.19),
+  r = rep(0.301316, 2)
+)
+
+emb_df1 <- tibble(emb1 = 0.58, emb2 = 0.93) |>
+  add_row(emb1 = 0.5, emb2 = 1.01)
+
+assign_data_vis1 <- ggplot(data = hex_grid, aes(x = x, y = y)) + 
+  geom_polygon(fill = "white", color = "#bdbdbd", aes(group = hex_poly_id)) +
+  geom_point(data = emb_df1, aes(x = emb1, y = emb2), color = "black", size = 0.7) +
+  geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID), size = 5, colour = "#bdbdbd") +
+  geom_circle(data = circles, aes(x0 = x, y0 = y, r = r), fill = NA) +
+  geom_segment(x = 0.58, y = 0.93, xend = 0.44, yend = 0.76, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "#969696") +
+  geom_segment(x = 0.5, y = 1.01, xend = 0.68, yend = 1.16, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "#969696")
+
+assign_data_vis2 <- ggplot(data = hex_grid, aes(x = x, y = y)) + 
+  geom_polygon(fill = "white", color = "#bdbdbd", aes(group = hex_poly_id)) +
+  geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID), size = 5, colour = "#bdbdbd")
+```
+
+```{r}
+#| fig-width: 9
+#| fig-height: 6
+assign_data_vis1 + assign_data_vis2 +
+  plot_layout(ncol = 2)
+```
+
+```{r}
 

### 2024-05-20 applications/pbmc3k/10_gen_model_in_data_space_author.R

@@ -26,7 +26,7 @@ pbmc_model <- fit_highd_model(
   bin1 = num_bins_x_pbmc,
   r2 = r2_pbmc,
   is_bin_centroid = TRUE,
-  is_rm_lwd_hex = TRUE,
+  is_rm_lwd_hex = FALSE,
   col_start_highd = "PC"
 )
 

### 2024-05-20 applications/pbmc3k/9_gen_author_best_model.R

@@ -32,7 +32,7 @@ for (xbins in bin1_vec) {
     bin1 = xbins,
     r2 = r2_pbmc,
     is_bin_centroid = TRUE,
-    is_rm_lwd_hex = TRUE,
+    is_rm_lwd_hex = FALSE,
     col_start_highd = "PC_"
   )
 

### 2024-05-21 supplementary/hexagon_binning_algorithm.qmd
Updated figures in hexagon algorithm doc

@@
     geom_polygon(fill = "white", color = "#bdbdbd", aes(group = hex_poly_id)) +
-    geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "#31a354") +
+    geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "#31a354", size = 0.9) +
     geom_point(data = all_centroids_df |> dplyr::filter(hexID == 1), aes(x = c_x, y = c_y), color = "black") + ## starting point
-    geom_rect(aes(xmin = 0 - 0.03, xmax = 1 - 0.03, ymin = 0 - 0.05, ymax = r2 - 0.05), fill = "white", color = "black", alpha = 0, linewidth = 1) +
+    geom_rect(aes(xmin = 0 - 0.03, xmax = 1 - 0.03, ymin = 0 - 0.05, ymax = r2 - 0.05), fill = "white", color = "black", alpha = 0, linewidth = 0.7) +
     geom_segment(x = 0 - 0.03, y = -0.35, xend = 1 - 0.03, yend = -0.35,
@@
     annotate("text", x=-0.15, y=0.45, label= "a2", color = "black") +
-    annotate("text", x=-0.18, y=-0.35, label= "(s1, s2)", color = "black") + ## starting point
-  coord_equal()
+    annotate("text", x=-0.18, y=-0.45, label= "(s1, s2)", color = "black") ## starting point
 ```
@@
 #| fig-cap: "Notations for hexagonal grid configurations."
-#| out-height: 30%
 #| fig-pos: H
+#| fig-width: 2
+#| fig-height: 3
 
@@
 
+emb_df2 <- tibble(emb1 = 0.161, emb2 = 0.7395) |>
+  add_row(emb1 = 0.683, emb2 = 0.7395) |>
+  add_row(emb1 = 0.2915, emb2 = 0.965) |>
+  add_row(emb1 = 0.5525, emb2 = 0.965) |>
+  add_row(emb1 = 0.2915, emb2 = 0.5135) |>
+  add_row(emb1 = 0.5525, emb2 = 0.5135) |>
+  add_row(emb1 = 0.422, emb2 = 1.04) |>
+  add_row(emb1 = 0.161, emb2 = 0.890) |>
+  add_row(emb1 = 0.161, emb2 = 0.589) |>
+  add_row(emb1 = 0.422, emb2 = 0.438) |>
+  add_row(emb1 = 0.683, emb2 = 0.589) |>
+  add_row(emb1 = 0.683, emb2 = 0.890)
+
 assign_data_vis2 <- ggplot(data = hex_grid, aes(x = x, y = y)) + 
   geom_polygon(fill = "white", color = "#bdbdbd", aes(group = hex_poly_id)) +
-  geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID), size = 5, colour = "#bdbdbd")
-```
-
-```{r}
-#| fig-width: 9
-#| fig-height: 6
-assign_data_vis1 + assign_data_vis2 +
-  plot_layout(ncol = 2)
-```
-
-```{r}
-
-## To generate embedding data
-emb_df <- tibble::tibble(emb1 = 0.37, emb2 = 0.3) |>
-  add_row(emb1 = 0.526, emb2 = 1.1) |>
-  add_row(emb1 = 0.735, emb2 = 1.1) |>
-  add_row(emb1 = 0.631, emb2 = 1.22) |>
-  add_row(emb1 = 0.735, emb2 = 1.16) |>
-  add_row(emb1 = 0.683, emb2 = 1.19) |>
-  add_row(emb1 = 0.735, emb2 = 1.04) |>
-  add_row(emb1 = 0.631, emb2 = 0.981) |>
-  add_row(emb1 = 0.683, emb2 = 1.0105) |>
-  add_row(emb1 = 0.526, emb2 = 1.04) |>
-  add_row(emb1 = 0.5785, emb2 = 1.0105) |> 
-  add_row(emb1 = 0.526, emb2 = 1.16) |>
-  add_row(emb1 = 0.5785, emb2 = 1.19)
-
-## Data to generate circle
-circles <- data.frame(
-  x = c(0.318, 0.422),
-  y = c(0.197,0.378),
-  r = rep(0.1205264,2)
-)
-
-assign_data_vis <- ggplot(data = hex_grid, aes(x = x, y = y)) + 
-  geom_polygon(fill = "white", color = "#bdbdbd", aes(group = hex_poly_id)) +
-  geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID), size = 2) +
-  geom_point(data = emb_df, aes(x = emb1, y = emb2), color = "#253494", size = 0.5) +
-  geom_circle(data = circles, aes(x0 = x, y0 = y, r = r), fill = NA) +
-  geom_segment(x = 0.37, y = 0.3, xend = 0.422, yend = 0.378, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
-  geom_segment(x = 0.526, y = 1.1, xend = 0.45, yend = 1.1, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
-  geom_segment(x = 0.735, y = 1.1, xend = 0.66, yend = 1.1, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
-  geom_segment(x = 0.631, y = 1.22, xend = 0.631, yend = 1.15, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
-  geom_segment(x = 0.735, y = 1.16, xend = 0.65, yend = 1.12, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
-  geom_segment(x = 0.735, y = 1.04, xend = 0.735, yend = 0.95, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
-  geom_segment(x = 0.683, y = 1.19, xend = 0.65, yend = 1.12, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
-  geom_segment(x = 0.631, y = 0.981, xend = 0.55, yend = 0.95, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
-  geom_segment(x = 0.683, y = 1.0105, xend = 0.7, yend = 0.95, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
-  geom_segment(x = 0.526, y = 1.04, xend = 0.526, yend = 0.95, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
-  geom_segment(x = 0.5785, y = 1.0105, xend = 0.536, yend = 0.97, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
-  geom_segment(x = 0.526, y = 1.16, xend = 0.49, yend = 1.12, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
-  geom_segment(x = 0.5785, y = 1.19, xend = 0.58, yend = 1.12, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "black") +
-  coord_equal()
+  geom_point(data = emb_df2, aes(x = emb1, y = emb2), color = "black", size = 0.7) +
+  geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID), size = 5, colour = "#bdbdbd") +
+  geom_segment(x = 0.161, y = 0.7395, xend = -0.06, yend = 0.740, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  geom_segment(x = 0.683, y = 0.7395, xend = 0.45, yend = 0.740, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  geom_segment(x = 0.2915, y = 0.965, xend = 0.43, yend = 0.79, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  geom_segment(x = 0.5525, y = 0.965, xend = 0.44, yend = 0.8, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  geom_segment(x = 0.2915, y = 0.5135, xend = 0.18, yend = 0.32, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  geom_segment(x = 0.5525, y = 0.5135, xend = 0.67, yend = 0.32, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  geom_segment(x = 0.422, y = 1.04, xend = 0.42, yend = 0.85, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  geom_segment(x = 0.161, y = 0.890, xend = -0.02, yend = 0.8, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  geom_segment(x = 0.161, y = 0.589, xend = 0.161, yend = 0.4, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  geom_segment(x = 0.422, y = 0.438, xend = 0.25, yend = 0.3, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  geom_segment(x = 0.683, y = 0.589, xend = 0.683, yend = 0.4, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  geom_segment(x = 0.683, y = 0.890, xend = 0.5, yend = 0.82, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd")
 
@@
 #| fig-cap: "Binning the data. Points are assigned to the nearest centroid. If a point is equidistant from multiple centroids, assigned to the lowest centroid."
-
-assign_data_vis
+#| fig-width: 9
+#| fig-height: 6
+assign_data_vis1 + assign_data_vis2 +
+  plot_layout(ncol = 2)
 ```

### 2024-05-21 paper.qmd
Update hexbin figure with suitable aspect ratio

@@
     geom_polygon(fill = "white", color = "#bdbdbd", aes(group = hex_poly_id)) +
-    geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "#31a354") +
+    geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "#31a354", size = 0.9) +
     geom_point(data = all_centroids_df |> dplyr::filter(hexID == 1), aes(x = c_x, y = c_y), color = "black") + ## starting point
-    geom_rect(aes(xmin = 0 - 0.03, xmax = 1 - 0.03, ymin = 0 - 0.05, ymax = r2- 0.05), fill = "white", color = "black", alpha = 0, linewidth = 1) +
+    geom_rect(aes(xmin = 0 - 0.03, xmax = 1 - 0.03, ymin = 0 - 0.05, ymax = r2- 0.05), fill = "white", color = "black", alpha = 0, linewidth = 0.7) +
     geom_segment(x = 0 - 0.03, y = -0.35, xend = 1 - 0.03, yend = -0.35,
@@
     annotate("text", x=-0.15, y=0.45, label= "a2", color = "black") +
-    annotate("text", x=-0.18, y=-0.35, label= "(s1, s2)", color = "black") + ## starting point
-  coord_fixed()
+    annotate("text", x=-0.18, y=-0.45, label= "(s1, s2)", color = "black") ## starting point
 ```
@@
 #| out-height: 30%
+#| fig-width: 2
+#| fig-height: 3
 #| fig-pos: H

### 2024-05-21 paper.qmd
Update the PBMC screen shots and added new video links

@@
 
-Screen shots of the **langevitour** of the PBMC3k data set, shows the model-in-data space, a video of the tour animation is available at (<https://youtu.be/Gnhh4hfsbbg>).
+Screen shots of the **langevitour** of the PBMC3k data set, shows the model-in-data space, a video of the tour animation is available at (<https://youtu.be/VqqWuE0Jj6A>).
 :::
@@
 
-Screen shots of the **langevitour** of the PBMC3k data set, shows the model in high-D, a video of the tour animation is available at (<https://youtu.be/R0EroJcGbas>).
+Screen shots of the **langevitour** of the PBMC3k data set, shows the model in high-D, a video of the tour animation is available at (<https://youtu.be/5Y1hE4i7N2k>).
 :::

### 2024-05-21 paper.tex
Update the PBMC screen shots and added new video links

@@
 the PBMC3k data set, shows the model-in-data space, a video of the tour
-animation is available at (\url{https://youtu.be/Gnhh4hfsbbg}).}
+animation is available at (\url{https://youtu.be/VqqWuE0Jj6A}).}
 
@@
 the PBMC3k data set, shows the model in high-D, a video of the tour
-animation is available at (\url{https://youtu.be/R0EroJcGbas}).}
+animation is available at (\url{https://youtu.be/5Y1hE4i7N2k}).}
 

### 2024-05-21 paper.qmd
Update screen shots and videos of MNIST

@@
 Furthermore, as shown in @fig-mnist1_sc3, some long edges exist in the \pD{} space that are not recognized as long edges in the \gD{} representation.
-However, PaCMAP is a *good* \gD{} representation of MNIST digit 1 data.
+However, PaCMAP is a reasonable \gD{} representation of MNIST digit 1 data.
+Because PaCMAP preserves the local structure.
 
@@
 ::: {#fig-mnist1_sc layout-ncol="3" fig-pos="H"}
-![](figures/mnist/sc_1.png)
-
-![](figures/mnist/sc_2.png)
-
-![](figures/mnist/sc_3.png)
-
-Screen shots of the **langevitour** of the MNIST digit 1 data set, shows the model-in-data space, a video of the tour animation is available at (<https://youtu.be/zq2GM9qvUNA>).
+![](figures/mnist/sc_1.png){#fig-mnist1_sc1}
+
+![](figures/mnist/sc_2.png){#fig-mnist1_sc2}
+
+![](figures/mnist/sc_3.png){#fig-mnist1_sc3}
+
+Screen shots of the **langevitour** of the MNIST digit 1 data set, shows the model-in-data space, a video of the tour animation is available at (<https://youtu.be/zcg_GXBmqjA>).
 :::

### 2024-05-21 paper.tex
Update screen shots and videos of MNIST

@@
 
-According to \textbf{?@fig-mnist1\_sc1}, the non-linear continuous
+According to Figure~\ref{fig-mnist1_sc1}, the non-linear continuous
 structure observed in the \gD{} representation of PaCMAP (see
@@
 a twisted pattern within the non-linear structure in \pD{} space (see
-\textbf{?@fig-mnist1\_sc2}), which is an additional pattern not visible
+Figure~\ref{fig-mnist1_sc2}), which is an additional pattern not visible
 in the \gD{} representation (see Figure~\ref{fig-pacmap-author}).
-Furthermore, as shown in \textbf{?@fig-mnist1\_sc3}, some long edges
+Furthermore, as shown in Figure~\ref{fig-mnist1_sc3}, some long edges
 exist in the \pD{} space that are not recognized as long edges in the
-\gD{} representation. However, PaCMAP is a \emph{good} \gD{}
-representation of MNIST digit 1 data.
-
-\begin{figure}[H]
-
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/mnist/sc_1.png}\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/mnist/sc_2.png}\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/mnist/sc_3.png}\end{minipage}%
+\gD{} representation. However, PaCMAP is a reasonable \gD{}
+representation of MNIST digit 1 data. Because PaCMAP preserves the local
+structure.
+
+\begin{figure}[H]
+
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/mnist/sc_1.png}
+
+}
+
+\subcaption{\label{fig-mnist1_sc1}}
+
+\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/mnist/sc_2.png}
+
+}
+
+\subcaption{\label{fig-mnist1_sc2}}
+
+\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{figures/mnist/sc_3.png}
+
+}
+
+\subcaption{\label{fig-mnist1_sc3}}
+
+\end{minipage}%
 
@@
 the tour animation is available at
-(\url{https://youtu.be/zq2GM9qvUNA}).}
+(\url{https://youtu.be/zcg_GXBmqjA}).}
 

### 2024-05-21 paper.tex
Updated the notations

@@
 $b$, $b'$ & total and non-empty hexagon bins in the grid\\
-$n_k$ & number of observations within the $k^{th}$ hexagon\\
+$n_m$ & number of observations within the $m^{th}$ hexagon\\
+\addlinespace
+$h$ & hexagonal id\\
 \bottomrule

### 2024-05-21 paper.qmd
Changed notations

@@
 
-Let $g : \mathbb{R}^{n\times k} \rightarrow \mathbb{R}^{m\times k}$ be the function that maps a point in \kD{} space to it's bin centroid ($C^{(k)}$). This is the model points in \kD{} space.
+Let $u : \mathbb{R}^{n\times k} \rightarrow \mathbb{R}^{m\times k}$ be the function that maps a point in \kD{} space to it's bin centroid ($C^{(k)}$). This is the model points in \kD{} space.
 
@@
 
-with corresponding \pD{} vector $\vec{f_i}$. Then, $f(g(x))$ maps a \pD{} point to the \pD{} model estimate, completing the model cycle.
+with corresponding \pD{} vector $\vec{f_i}$. Then, $f(u(x))$ maps a \pD{} point to the \pD{} model estimate, completing the model cycle.
 

### 2024-05-21 paper.tex
Changed notations

@@
 
-Let \(g : \mathbb{R}^{n\times k} \rightarrow \mathbb{R}^{m\times k}\) be
+Let \(u : \mathbb{R}^{n\times k} \rightarrow \mathbb{R}^{m\times k}\) be
 the function that maps a point in \kD{} space to it's bin centroid
@@
 
-with corresponding \pD{} vector \(\vec{f_i}\). Then, \(f(g(x))\) maps a
+with corresponding \pD{} vector \(\vec{f_i}\). Then, \(f(u(x))\) maps a
 \pD{} point to the \pD{} model estimate, completing the model cycle.

### 2024-05-30 paper.qmd
Update section lifting model to p-D

@@
 From a statistical perspective we can consider the distances between points in the \kD{}   layout to be variance that the model explains, and the (relative) difference with their distances in \pD{}   is the error, or unexplained variance.
-We can also imagine that the positioning of points in 2D represent the fitted values, that will have some prescribed position in \pD{}   that can be compared with their observed values.
+We can also imagine that the positioning of points in \gD{}    represent the fitted values, that will have some prescribed position in \pD{}   that can be compared with their observed values.
 This is the conceptual framework underlying the more formal versions of factor analysis [@cfa69] and multidimensional scaling (MDS) [@borg2005].
@@
 
-With our goal being to make a representation of this 2D layout that can be lifted into high-dimensional space, the layout needs to be augmented to include neighbour information.
+With our goal being to make a representation of this \gD{}    layout that can be lifted into high-dimensional space, the layout needs to be augmented to include neighbour information.
 A simple approach would be to triangulate the points and add edges.
@@
 
+Let $u : \mathbb{R}^{n\times k} \rightarrow \mathbb{R}^{m\times k}$ be the function that maps a point in \kD{} space to it's bin centroid ($C^{(k)}$). This is the model points in \kD{} space.
+
+When $k = 2$, one of the representations of a bin is the hexagonal centroid (@fig-NLDR-scurve (c)). Also, it could be the mean of the data points within each hexagon.
+
 <!-- XXX How are you doing this? Do you check the bounds of the hexagon? Or do you use distance to centroid? -->
@@
 <!--We deliberately separate out the creation of the hexagonal grid from the mapping of points on the grid.-->
-#### Summarization of the data in 2D
-
-Let $u : \mathbb{R}^{n\times k} \rightarrow \mathbb{R}^{m\times k}$ be the function that maps a point in \kD{} space to it's bin centroid ($C^{(k)}$). This is the model points in \kD{} space.
-
-When $k = 2$, one of the representations of a bin is the hexagonal centroid (@fig-NLDR-scurve (c)). Also, it could be the mean of the data points within each hexagon.
-
-#### Generating edges to indicate neighbors and removing long edges
+
+#### Indicating neighbors by line segments connecting centroids
 
@@
 The last step is to lift the \kD{} model into \pD{} by computing \pD{} vectors that represent bin centroids.
-Define the set $H^k$ to be the set of points that belong to bin $k$.
-We use the \pD{} Euclidean mean of the points in $H^k$ to map the centroid $C_{h}^{(k)} = (c_{h1}, c_{h2})$ to a point in \pD{}.
-Let the $i$-th component of the \pD{} mean be
-
-$$f_i = \frac{1}{\left|H^k\right|}\sum_{y \in H^k} y_i,$$
-
-with corresponding \pD{} vector $\vec{f_i}$. Then, $f(u(x))$ maps a \pD{} point to the \pD{} model estimate, completing the model cycle.
-
-Furthermore, edges that exist between \kD{} representations should also generate edges in \pD{} by connecting \pD{} mapping of the corresponding \kD{} representations.
+We use the \pD{} mean of the points in $H_h$ to map the centroid $C_{h}^{(k)} = (c_{h1}, c_{h2}, \dots, c_{hk})$ to a point in \pD{}.
+Let the \pD{} mean be
+
+$$C_{h}^{(p)} = \frac{1}{n_h}\sum_{i =1}^{n_h} x_i, h = {1, \dots, b; n_h > 0}.$$
+Furthermore, line segments that exist in the \kD{} model generate line segments in \pD{} by connecting the \pD{} means of the corresponding \kD{} bin centroids.
 
@@
 To assess how well our method captures and represents the underlying structure of the \pD{} data, residuals are important.
-Residuals are computed by taking the squared \pD{} Euclidean distance, $\sum_{i = 1}^{n}\sum_{j = 1}^{p} (x_{ij} - C^{(p)}_{x_ij})^2$.
-Additionally, Mean Squared Error (MSE), $\sum_{i = 1}^{n}\sum_{j = 1}^{p} \frac{(x_{ij} - C^{(p)}_{x_ij})^2}{n}$, is computed based on residuals.
+Residuals are computed by taking the squared \pD{} Euclidean distance, $\sum_{i = 1}^{n}\sum_{j = 1}^{p} (x_{ij} - C^{(p)}_{hx_ij})^2$.
+Additionally, Mean Squared Error (MSE), $\frac{1}{n}\sum_{i = 1}^{n}\sum_{j = 1}^{p}(x_{ij} - C^{(p)}_{hx_ij})^2$, is computed based on residuals.
 
@@
 #| fig-pos: H
-#| fig-cap: "Hexbin plots from different number of bins for the UMAP applied to S-curve data: (a) $b = 12$ (3, 5), (b) $b = 190$ (10, 19), and (c) $b = 2176$ (34, 64).
+#| fig-cap: "Hexbin plots from different number of bins for the UMAP applied to S-curve data: (a) $b = 15$ (3, 5), (b) $b = 190$ (10, 19), and (c) $b = 2176$ (34, 64).
 The hexbins are colored based on the density of points, with darker colors indicating higher density and yellow color representing lower density within each bin.
@@
 benchmark_label_df <- tibble(y = benchmark_scurve - 0.03,
-                             x = 1.4,
+                             x = 1.3,
                              label = paste0("benchmark is ", benchmark_scurve))

### 2024-05-30 paper.tex
Update section lifting model to p-D

@@
 distances in \pD{} is the error, or unexplained variance. We can also
-imagine that the positioning of points in 2D represent the fitted
+imagine that the positioning of points in \gD{} represent the fitted
 values, that will have some prescribed position in \pD{} that can be
@@
 \midrule
-$n, p, k, m$ & number of observations, variables, embedding dimension, number of non-empty bins, respectively\\
+$n, p, k$ & number of observations, variables, embedding dimension, respectively\\
 $\mathbfit{X}, \mathbfit{x}$ & $p$-dimensional data (population, sample)\\
@@
 $q$ & buffer to ensure hexgrid covers data, proportion of data range, 0-1\\
-$b$, $b'$ & total and non-empty hexagon bins in the grid\\
-$n_m$ & number of observations within the $m^{th}$ hexagon\\
+$m$ & number of non-empty bins\\
+$b$ & number of  hexagons in the grid\\
 \addlinespace
@@
 
-With our goal being to make a representation of this 2D layout that can
-be lifted into high-dimensional space, the layout needs to be augmented
-to include neighbour information. A simple approach would be to
-triangulate the points and add edges. A more stable approach is to first
-hexagonally bin the data, reducing it from \(n\) to \(m\leq n\)
+With our goal being to make a representation of this \gD{} layout that
+can be lifted into high-dimensional space, the layout needs to be
+augmented to include neighbour information. A simple approach would be
+to triangulate the points and add edges. A more stable approach is to
+first hexagonally bin the data, reducing it from \(n\) to \(m\leq n\)
 observations, and connect the bin centroids. This process serves to
@@
 
-\subsubsection{Summarization of the data in
-2D}\label{summarization-of-the-data-in-2d}
-
 Let \(u : \mathbb{R}^{n\times k} \rightarrow \mathbb{R}^{m\times k}\) be
@@
 
-\subsubsection{Generating edges to indicate neighbors and removing long
-edges}\label{generating-edges-to-indicate-neighbors-and-removing-long-edges}
+\subsubsection{Indicating neighbors by line segments connecting
+centroids}\label{indicating-neighbors-by-line-segments-connecting-centroids}
 
@@
 The last step is to lift the \kD{} model into \pD{} by computing \pD{}
-vectors that represent bin centroids. Define the set \(H^k\) to be the
-set of points that belong to bin \(k\). We use the \pD{} Euclidean mean
-of the points in \(H^k\) to map the centroid
-\(C_{h}^{(k)} = (c_{h1}, c_{h2})\) to a point in \pD{}. Let the \(i\)-th
-component of the \pD{} mean be
-
-\[f_i = \frac{1}{\left|H^k\right|}\sum_{y \in H^k} y_i,\]
-
-with corresponding \pD{} vector \(\vec{f_i}\). Then, \(f(u(x))\) maps a
-\pD{} point to the \pD{} model estimate, completing the model cycle.
-
-Furthermore, edges that exist between \kD{} representations should also
-generate edges in \pD{} by connecting \pD{} mapping of the corresponding
-\kD{} representations.
+vectors that represent bin centroids. We use the \pD{} mean of the
+points in \(H_h\) to map the centroid
+\(C_{h}^{(k)} = (c_{h1}, c_{h2}, \dots, c_{hk})\) to a point in \pD{}.
+Let the \pD{} mean be
+
+\[C_{h}^{(p)} = \frac{1}{n_h}\sum_{i =1}^{n_h} x_i, h = {1, \dots, b; n_h > 0}.\]
+Furthermore, line segments that exist in the \kD{} model generate line
+segments in \pD{} by connecting the \pD{} means of the corresponding
+\kD{} bin centroids.
 
@@
 computed by taking the squared \pD{} Euclidean distance,
-\(\sum_{i = 1}^{n}\sum_{j = 1}^{p} (x_{ij} - C^{(p)}_{x_ij})^2\).
+\(\sum_{i = 1}^{n}\sum_{j = 1}^{p} (x_{ij} - C^{(p)}_{hx_ij})^2\).
 Additionally, Mean Squared Error (MSE),
-\(\sum_{i = 1}^{n}\sum_{j = 1}^{p} \frac{(x_{ij} - C^{(p)}_{x_ij})^2}{n}\),
+\(\frac{1}{n}\sum_{i = 1}^{n}\sum_{j = 1}^{p}(x_{ij} - C^{(p)}_{hx_ij})^2\),
 is computed based on residuals.
@@
 \caption{\label{fig-bins-scurve}Hexbin plots from different number of
-bins for the UMAP applied to S-curve data: (a) \(b = 12\) (3, 5), (b)
+bins for the UMAP applied to S-curve data: (a) \(b = 15\) (3, 5), (b)
 \(b = 190\) (10, 19), and (c) \(b = 2176\) (34, 64). The hexbins are

### 2024-05-30 examples/scurve/6_gen_model_pacmap.R

@@ -1,0 +1,90 @@
+## Import necessary libraries
+library(quollr)
+library(dplyr)
+library(reader)
+library(langevitour)
+
+training_data_scurve <- read_rds("data/s_curve/s_curve_training.rds")
+
+pacmap_scurve <- read_rds(file = "data/s_curve/s_curve_pacmap.rds")
+
+scurve_scaled_obj <- gen_scaled_data(
+  data = pacmap_scurve)
+
+pacmap_scurve_scaled <- scurve_scaled_obj$scaled_nldr
+lim1 <- scurve_scaled_obj$lim1
+lim2 <- scurve_scaled_obj$lim2
+r2 <- diff(lim2)/diff(lim1)
+
+## Compute hexbin parameters
+num_bins_x_scurve <- 13
+
+scurve_model <- fit_highd_model(
+  training_data = training_data_scurve,
+  emb_df = pacmap_scurve_scaled,
+  bin1 = num_bins_x_scurve,
+  r2 = r2,
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = FALSE,
+  col_start_highd = "x"
+)
+
+df_bin_centroids_scurve <- scurve_model$df_bin_centroids
+df_bin_scurve <- scurve_model$df_bin
+
+## Triangulate bin centroids
+tr1_object_scurve <- tri_bin_centroids(
+  df_bin_centroids_scurve, x = "c_x", y = "c_y")
+tr_from_to_df_scurve <- gen_edges(
+  tri_object = tr1_object_scurve)
+
+## Compute 2D distances
+distance_scurve <- cal_2d_dist(
+  tr_coord_df = tr_from_to_df_scurve,
+  start_x = "x_from",
+  start_y = "y_from",
+  end_x = "x_to",
+  end_y = "y_to",
+  select_vars = c("from", "to", "distance"))
+
+## To find the benchmark value
+benchmark_scurve <- find_lg_benchmark(
+  distance_edges = distance_scurve,
+  distance_col = "distance")
+
+## Hexagonal binning to have regular hexagons
+hb_obj_scurve <- hex_binning(
+  data = pacmap_scurve_scaled,
+  bin1 = num_bins_x_scurve,
+  r2 = r2)
+
+pacmap_data_with_hb_id <- hb_obj_scurve$data_hb_id
+
+df_all_scurve <- dplyr::bind_cols(training_data_scurve |> dplyr::select(-ID),
[…]

### 2024-05-30 paper.qmd
Update section triangulation

@@
 
-Delaunay triangulation is used to connect neighboring centroids, which is needed to preserve neighborhood information when the model is lifted into \pD{}.
-
-When $k = 2$ Delaunay triangulation on $C^{(2)}$ generates the model in \gD{} space, which is a triangular mesh (@fig-NLDR-scurve (d)).
+Delaunay triangulation [@lee1980] is used to connect neighboring centroids, which is needed to preserve neighborhood information when the model is lifted into \pD{}.
+
+<!-- When $k = 2$ Delaunay triangulation on $C^{(2)}$ generates the model in \gD{} space, which is a triangular mesh (@fig-NLDR-scurve (d)).
 It generates convex hulls of $C^{(2)}$ such that the circumcircle of every triangle in the triangulation contains no other points from $C^{(2)}$.
-
-When neighboring centroids are connected with edges in Delaunay triangulation in \gD{}, there are situations where distant centroids are also connected, which generates long edges.
-To generate a smooth surface in \gD{}, these long edges are removed (@fig-NLDR-scurve (d)).
+-->
+
+It can also happen that distant centroids can be connected, which can result in long line segments.
+In order to generate a smooth surface in \gD{}, these long line segments must be removed (@fig-NLDR-scurve (d)).
 

### 2024-05-30 paper.tex
Update section triangulation

@@
 
-Delaunay triangulation is used to connect neighboring centroids, which
-is needed to preserve neighborhood information when the model is lifted
-into \pD{}.
-
-When \(k = 2\) Delaunay triangulation on \(C^{(2)}\) generates the model
-in \gD{} space, which is a triangular mesh (Figure~\ref{fig-NLDR-scurve}
-(d)). It generates convex hulls of \(C^{(2)}\) such that the
-circumcircle of every triangle in the triangulation contains no other
-points from \(C^{(2)}\).
-
-When neighboring centroids are connected with edges in Delaunay
-triangulation in \gD{}, there are situations where distant centroids are
-also connected, which generates long edges. To generate a smooth surface
-in \gD{}, these long edges are removed (Figure~\ref{fig-NLDR-scurve}
-(d)).
+Delaunay triangulation \citep{lee1980} is used to connect neighboring
+centroids, which is needed to preserve neighborhood information when the
+model is lifted into \pD{}.
+
+It can also happen that distant centroids can be connected, which can
+result in long line segments. In order to generate a smooth surface in
+\gD{}, these long line segments must be removed
+(Figure~\ref{fig-NLDR-scurve} (d)).
 

### 2024-05-30 paper.qmd
Update binning data section

@@
 
-Points are allocated to the bin they fall into based on the nearest centroid.
+<!-- Points are allocated to the bin they fall into based on the nearest centroid.
 In situations where a point is equidistant from multiple centroids, tie-breaking rules are applied.
@@
 If multiple centroids are in different rows, the point is assigned to the bottom centroid.
-
-$\{ i \in H_h, h = 1, \dots, b, \text{ and } i = 1, \dots, n\}$
-
-Let $u : \mathbb{R}^{n\times k} \rightarrow \mathbb{R}^{m\times k}$ be the function that maps a point in \kD{} space to it's bin centroid ($C^{(k)}$). This is the model points in \kD{} space.
-
-When $k = 2$, one of the representations of a bin is the hexagonal centroid (@fig-NLDR-scurve (c)). Also, it could be the mean of the data points within each hexagon.
+-->
+
+<!-- $\{ i \in H_h, h = 1, \dots, b, \text{ and } i = 1, \dots, n\}$ -->
+
+We define the function $u: \mathbb{R}^{n\times 2} \rightarrow \mathbb{R}^{m\times 2}$, which generates the model by assigning the points to bin centroids.
+
+For the $i^{th}$ point, $u(i) = \arg\min_{j = 1, \dots, b} \sqrt{(y_{i1} - C^{(2)}_{j1})^2 + (y_{i2} - C^{(2)}_{j2})^2}$, and falls in $H_h = \{i| u(i) = h\}$.
+
+When $k = 2$, the hexagonal bin centroid is one representation of a bin (@fig-NLDR-scurve (c)). Additionally, it can be the mean of the points within each hexagon.
 
@@
 The last step is to lift the \kD{} model into \pD{} by computing \pD{} vectors that represent bin centroids.
-We use the \pD{} mean of the points in $H_h$ to map the centroid $C_{h}^{(k)} = (c_{h1}, c_{h2}, \dots, c_{hk})$ to a point in \pD{}.
+We use the \pD{} mean of the points in $H_h$ to map the centroid $C_{h}^{(2)} = (c_{h1}, c_{h2})$ to a point in \pD{}.
 Let the \pD{} mean be

### 2024-05-30 paper.tex
Update binning data section

@@
 
-Points are allocated to the bin they fall into based on the nearest
-centroid. In situations where a point is equidistant from multiple
-centroids, tie-breaking rules are applied. If multiple centroids are in
-the same row, the point is assigned to the leftmost centroid. If
-multiple centroids are in different rows, the point is assigned to the
-bottom centroid.
-
-\(\{ i \in H_h, h = 1, \dots, b, \text{ and } i = 1, \dots, n\}\)
-
-Let \(u : \mathbb{R}^{n\times k} \rightarrow \mathbb{R}^{m\times k}\) be
-the function that maps a point in \kD{} space to it's bin centroid
-(\(C^{(k)}\)). This is the model points in \kD{} space.
-
-When \(k = 2\), one of the representations of a bin is the hexagonal
-centroid (Figure~\ref{fig-NLDR-scurve} (c)). Also, it could be the mean
-of the data points within each hexagon.
+We define the function
+\(u: \mathbb{R}^{n\times 2} \rightarrow \mathbb{R}^{m\times 2}\), which
+generates the model by assigning the points to bin centroids.
+
+For the \(i^{th}\) point,
+\(u(i) = \arg\min_{j = 1, \dots, b} \sqrt{(y_{i1} - C^{(2)}_{j1})^2 + (y_{i2} - C^{(2)}_{j2})^2}\),
+and falls in \(H_h = \{i| u(i) = h\}\).
+
+When \(k = 2\), the hexagonal bin centroid is one representation of a
+bin (Figure~\ref{fig-NLDR-scurve} (c)). Additionally, it can be the mean
+of the points within each hexagon.
 
@@
 vectors that represent bin centroids. We use the \pD{} mean of the
-points in \(H_h\) to map the centroid
-\(C_{h}^{(k)} = (c_{h1}, c_{h2}, \dots, c_{hk})\) to a point in \pD{}.
-Let the \pD{} mean be
+points in \(H_h\) to map the centroid \(C_{h}^{(2)} = (c_{h1}, c_{h2})\)
+to a point in \pD{}. Let the \pD{} mean be
 

### 2024-05-30 paper.qmd
Updated fitted values section

@@
 
-The prediction approach involves performing the K-nearest neighbors (KNN) algorithm for an unsupervised classification problem.
+<!-- The prediction approach involves performing the K-nearest neighbors (KNN) algorithm for an unsupervised classification problem.
 First, the nearest \pD{} model point is identified for a given new \pD{} point.
@@
 .
+-->
+
+We define the function $z: \mathbb{R}^{n\times p} \rightarrow \mathbb{R}^{m\times p}$.
+
+For the $i^{th}$ point, $z(i) = \arg\min_{j = 1, \dots, b} \sqrt{\sum_{v=1}^{P}(x_{iv} - C^{(2)}_{jv})^2}$, and assign it to $H_h = \{i| z(i) = h\}$.
+Then, the fitted values for $i$ are defined as $C_{h}^{(2)} = (c_{h1}, c_{h2})$.
 
@@
 To assess how well our method captures and represents the underlying structure of the \pD{} data, residuals are important.
-Residuals are computed by taking the squared \pD{} Euclidean distance, $\sum_{i = 1}^{n}\sum_{j = 1}^{p} (x_{ij} - C^{(p)}_{hx_ij})^2$.
-Additionally, Mean Squared Error (MSE), $\frac{1}{n}\sum_{i = 1}^{n}\sum_{j = 1}^{p}(x_{ij} - C^{(p)}_{hx_ij})^2$, is computed based on residuals.
+Residuals are computed by taking the squared \pD{} Euclidean distance, $\sum_{i = 1}^{n}\sum_{j = 1}^{p} (x_{ij} - C^{(p)}_{hj})^2$.
+Additionally, Mean Squared Error (MSE), $\frac{1}{n}\sum_{i = 1}^{n}\sum_{j = 1}^{p}(x_{ij} - C^{(p)}_{hj})^2$, is computed based on residuals.
 

### 2024-05-30 paper.tex
Updated fitted values section

@@
 
-The prediction approach involves performing the K-nearest neighbors
-(KNN) algorithm for an unsupervised classification problem. First, the
-nearest \pD{} model point is identified for a given new \pD{} point.
-Then, the corresponding \kD{} centroid mapping for the identified \pD{}
-model point is determined. Finally, the coordinates of this \kD{}
-centroid are used as the predicted \kD{} embedding for the new \pD{}
-data point. .
+We define the function
+\(z: \mathbb{R}^{n\times p} \rightarrow \mathbb{R}^{m\times p}\).
+
+For the \(i^{th}\) point,
+\(z(i) = \arg\min_{j = 1, \dots, b} \sqrt{\sum_{v=1}^{P}(x_{iv} - C^{(2)}_{jv})^2}\),
+and assign it to \(H_h = \{i| z(i) = h\}\). Then, the fitted values for
+\(i\) are defined as \(C_{h}^{(2)} = (c_{h1}, c_{h2})\).
 
@@
 computed by taking the squared \pD{} Euclidean distance,
-\(\sum_{i = 1}^{n}\sum_{j = 1}^{p} (x_{ij} - C^{(p)}_{hx_ij})^2\).
+\(\sum_{i = 1}^{n}\sum_{j = 1}^{p} (x_{ij} - C^{(p)}_{hj})^2\).
 Additionally, Mean Squared Error (MSE),
-\(\frac{1}{n}\sum_{i = 1}^{n}\sum_{j = 1}^{p}(x_{ij} - C^{(p)}_{hx_ij})^2\),
+\(\frac{1}{n}\sum_{i = 1}^{n}\sum_{j = 1}^{p}(x_{ij} - C^{(p)}_{hj})^2\),
 is computed based on residuals.

### 2024-05-31 paper.qmd
Add first scratch of the discussion

@@
 
-- Summarise contributions
-- Explain where it is expected or not expected to work, eg higher dimensional relationships
-- Human behaviour, the desire to have more certainty, and a tendency to prefer the well-separated views
-- Predicting new observations in $k$-D
-- Extending layouts beyond $k$-D, when 2D is clearly inadequate.
-- Diagnostic app to explore differences in distances
-- What might be useful enhancements
+<!-- - Summarise contributions -->
+<!-- - Explain where it is expected or not expected to work, eg higher dimensional relationships -->
+<!-- - Human behaviour, the desire to have more certainty, and a tendency to prefer the well-separated views (need to add)-->
+<!-- - Predicting new observations in $k$-D -->
+<!-- - Extending layouts beyond $k$-D, when 2D is clearly inadequate. -->
+<!-- - Diagnostic app to explore differences in distances (need to add)--> 
+<!-- - What might be useful enhancements -->
+
+We have developed an algorithm to evaluate the most useful NLDR method and (hyper-)parameter choices for creating a reasonable \gD{} layout of high-dimensional data.
+Our objective is to fit a model for the \gD{} layout that preserves the relationships between neighboring points and turns it into a high-dimensional wireframe, which can be overlaid on the data and visualized using a tour.
+This approach is defined as *model-in-data-space*.
+Viewing a model in the data space is an ideal way to examine the fit.
+
+The S-curve example demonstrates how the model accurately fits the points.
+Viewing the wireframe with a tour reveals how well the model represents the underlying geometry of the S-curve, highlighting its ability to preserve both local and global structures in high-dimensional space.
+
+As demonstrated in the simulation case study, all observed NLDR methods preserve the global structure, while tSNE is the only method that preserves the local structure.
+Our approach helps in determining this and illustrates how different methods have specific quirks.
+
+Even though three distant well-separated clusters have shown in the UMAP layout of **pbmc** suggested by @chen2023, the model evident that the clusters are close to each other in high dimension.
+Also, viewing the model in the data space shows data patches which shows the data are not uniformly distributed within the clusters and non-linear structures within clusters.
+These summaries can't be observed in UMAP layout.
+Therefore, model in high dimension is useful to obtain some quirks like these.
+
+On the other hand, the error or unexplained variance is important in evaluating how well the model fit the data.
+By observing error for different number of bins, tSNE with perplexity 30 selects as a reasonable representation for **pbmc**.
+Connecting closest clusters by line segments in the fitted model provides evidence of preserving the neighborhood.
+
+We have shown in the **digit: 1**, how the model accurately represents the non-linear structure.
+Also, provides some unobserved information in \gD{}.
+The twisted pattern which squeezed the structure in some projections and long line segments which detect anomalies.
+
+Predicting new observations in \kD{} is particularly valuable due to the limitations of some NLDR methods, like tSNE, which don't provide a straightforward method for prediction.
+As a result, our approach offers a solution that capable of generating predicted \kD{} embedding regardless of the NLDR method employed, effectively addressing this functional gap.
+
+The future enhancements involve introducing approaches to bind the data and indicate line segments beyond \gD{}.
+This would help in creating a more accurate representation of the data when \gD{} layout is inadequate.
+
+## Supplementary Materials
+
+Code, and data for reproducing this paper are available at [https://github.com/JayaniLakshika/paper-nldr-vis-algorithm](https://github.com/JayaniLakshika/paper-nldr-vis-algorithm).
 

### 2024-05-31 paper.tex
Add first scratch of the discussion

@@
 
-\begin{itemize}
-\tightlist
-\item
-  Summarise contributions
-\item
-  Explain where it is expected or not expected to work, eg higher
-  dimensional relationships
-\item
-  Human behaviour, the desire to have more certainty, and a tendency to
-  prefer the well-separated views
-\item
-  Predicting new observations in \(k\)-D
-\item
-  Extending layouts beyond \(k\)-D, when 2D is clearly inadequate.
-\item
-  Diagnostic app to explore differences in distances
-\item
-  What might be useful enhancements
-\end{itemize}
+We have developed an algorithm to evaluate the most useful NLDR method
+and (hyper-)parameter choices for creating a reasonable \gD{} layout of
+high-dimensional data. Our objective is to fit a model for the \gD{}
+layout that preserves the relationships between neighboring points and
+turns it into a high-dimensional wireframe, which can be overlaid on the
+data and visualized using a tour. This approach is defined as
+\emph{model-in-data-space}. Viewing a model in the data space is an
+ideal way to examine the fit.
+
+The S-curve example demonstrates how the model accurately fits the
+points. Viewing the wireframe with a tour reveals how well the model
+represents the underlying geometry of the S-curve, highlighting its
+ability to preserve both local and global structures in high-dimensional
+space.
+
+As demonstrated in the simulation case study, all observed NLDR methods
+preserve the global structure, while tSNE is the only method that
+preserves the local structure. Our approach helps in determining this
+and illustrates how different methods have specific quirks.
+
+Even though three distant well-separated clusters have shown in the UMAP
+layout of \textbf{pbmc} suggested by \citet{chen2023}, the model evident
+that the clusters are close to each other in high dimension. Also,
+viewing the model in the data space shows data patches which shows the
+data are not uniformly distributed within the clusters and non-linear
+structures within clusters. These summaries can't be observed in UMAP
+layout. Therefore, model in high dimension is useful to obtain some
+quirks like these.
+
+On the other hand, the error or unexplained variance is important in
+evaluating how well the model fit the data. By observing error for
+different number of bins, tSNE with perplexity 30 selects as a
+reasonable representation for \textbf{pbmc}. Connecting closest clusters
+by line segments in the fitted model provides evidence of preserving the
+neighborhood.
+
+We have shown in the \textbf{digit: 1}, how the model accurately
+represents the non-linear structure. Also, provides some unobserved
+information in \gD{}. The twisted pattern which squeezed the structure
+in some projections and long line segments which detect anomalies.
+
+Predicting new observations in \kD{} is particularly valuable due to the
+limitations of some NLDR methods, like tSNE, which don't provide a
+straightforward method for prediction. As a result, our approach offers
+a solution that capable of generating predicted \kD{} embedding
+regardless of the NLDR method employed, effectively addressing this
+functional gap.
+
+The future enhancements involve introducing approaches to bind the data
+and indicate line segments beyond \gD{}. This would help in creating a
+more accurate representation of the data when \gD{} layout is
+inadequate.
+
+\section{Supplementary Materials}\label{supplementary-materials}
+
+Code, and data for reproducing this paper are available at
+\url{https://github.com/JayaniLakshika/paper-nldr-vis-algorithm}.
 

### 2024-05-31 paper.qmd
Updated version for the discussion

--- 2024-05-30 paper.qmd
@@
 
-- Summarise contributions
-- Explain where it is expected or not expected to work, eg higher dimensional relationships
-- Human behaviour, the desire to have more certainty, and a tendency to prefer the well-separated views
-- Predicting new observations in $k$-D
-- Extending layouts beyond $k$-D, when 2D is clearly inadequate.
-- Diagnostic app to explore differences in distances
-- What might be useful enhancements
+<!-- - Summarise contributions -->
+<!-- - Explain where it is expected or not expected to work, eg higher dimensional relationships -->
+<!-- - Human behaviour, the desire to have more certainty, and a tendency to prefer the well-separated views (need to add) -->
+<!-- - Predicting new observations in $k$-D -->
+<!-- - Extending layouts beyond $k$-D, when 2D is clearly inadequate. -->
+<!-- - Diagnostic app to explore differences in distances (need to add) -->
+<!-- - What might be useful enhancements -->
+
+This study makes several important contributions to the field of NLDR.
+We have developed an algorithm to evaluate the most useful NLDR method and (hyper-)parameter choices for creating a reasonable \gD{} layout of high-dimensional data.
+Our objective is to fit a model for the \gD{} layout that preserves the relationships between neighboring points and turns it into a high-dimensional wireframe, which can be overlaid on the data and visualized using a tour.
+This approach is defined as *model-in-data-space*.
+Viewing a model in the data space is an ideal way to examine the fit.
+
+The effectiveness of this approach is illustrated through various examples.
+For instance, the S-curve example demonstrates how the model accurately fits the points, capturing both local and global structures in high-dimensional space.
+Our simulation case study further, five Gaussian cluster example shows that while all observed NLDR methods preserve the global structure, only tSNE effectively maintains the local structure, highlighting the specific strengths and quirks of different methods.
+
+Human behavior often shows a desire for more certainty and a tendency to prefer well-separated views.
+This emphasizes the importance of clear and distinct clusters.
+For example, in the UMAP layout of the **pbmc** dataset suggested by @chen2023, three distant, well-separated clusters are shown.
+However, our model reveals that these clusters are actually close to each other in \pD{}.
+Additionally, the model discovers non-uniform data distribution and non-linear structures within the clusters that are not visible in the UMAP layout, demonstrating the ability of our model in uncovering hidden data characteristics.
+
+Evaluating the error or unexplained variance is important for assessing how well the model fits the data.
+By examining the error for different numbers of bins, we found that tSNE with a perplexity of $30$ provides a reasonable representation for the **pbmc** dataset.
+Connecting the closest clusters with line segments in the fitted model further supports the preservation of neighborhood relationships.
+
+The **digit: 1** example further illustrates the model's ability to accurately capture non-linear structures and provide additional information.
+Key findings include a twisted pattern that compresses the structure in some projections and long line segments that detect anomalies.
+
+Predicting new observations in \kD{} is particularly valuable due to the limitations of some NLDR methods, like tSNE, which don't provide a straightforward method for prediction.
+As a result, our approach offers a solution that capable of generating predicted \kD{} embedding regardless of the NLDR method employed, effectively addressing this functional gap.
+
+In conclusion, while our method effectively captures and represents high-dimensional data structures, further enhancements could involve introducing approaches to bind the data and indicate line segments beyond \gD{}.
+These improvements would help in creating a more accurate representation of the data when \gD{} layout is inadequate.
+
+## Supplementary Materials
+
+Code, and data for reproducing this paper are available at [https://github.com/JayaniLakshika/paper-nldr-vis-algorithm](https://github.com/JayaniLakshika/paper-nldr-vis-algorithm).
 

### 2024-05-31 paper.tex
Updated version for the discussion

--- 2024-05-30 paper.tex
@@
 
-\begin{itemize}
-\tightlist
-\item
-  Summarise contributions
-\item
-  Explain where it is expected or not expected to work, eg higher
-  dimensional relationships
-\item
-  Human behaviour, the desire to have more certainty, and a tendency to
-  prefer the well-separated views
-\item
-  Predicting new observations in \(k\)-D
-\item
-  Extending layouts beyond \(k\)-D, when 2D is clearly inadequate.
-\item
-  Diagnostic app to explore differences in distances
-\item
-  What might be useful enhancements
-\end{itemize}
+This study makes several important contributions to the field of NLDR.
+We have developed an algorithm to evaluate the most useful NLDR method
+and (hyper-)parameter choices for creating a reasonable \gD{} layout of
+high-dimensional data. Our objective is to fit a model for the \gD{}
+layout that preserves the relationships between neighboring points and
+turns it into a high-dimensional wireframe, which can be overlaid on the
+data and visualized using a tour. This approach is defined as
+\emph{model-in-data-space}. Viewing a model in the data space is an
+ideal way to examine the fit.
+
+The effectiveness of this approach is illustrated through various
+examples. For instance, the S-curve example demonstrates how the model
+accurately fits the points, capturing both local and global structures
+in high-dimensional space. Our simulation case study further, five
+Gaussian cluster example shows that while all observed NLDR methods
+preserve the global structure, only tSNE effectively maintains the local
+structure, highlighting the specific strengths and quirks of different
+methods.
+
+Human behavior often shows a desire for more certainty and a tendency to
+prefer well-separated views. This emphasizes the importance of clear and
+distinct clusters. For example, in the UMAP layout of the \textbf{pbmc}
+dataset suggested by \citet{chen2023}, three distant, well-separated
+clusters are shown. However, our model reveals that these clusters are
+actually close to each other in \pD{}. Additionally, the model discovers
+non-uniform data distribution and non-linear structures within the
+clusters that are not visible in the UMAP layout, demonstrating the
+ability of our model in uncovering hidden data characteristics.
+
+Evaluating the error or unexplained variance is important for assessing
+how well the model fits the data. By examining the error for different
+numbers of bins, we found that tSNE with a perplexity of \(30\) provides
+a reasonable representation for the \textbf{pbmc} dataset. Connecting
+the closest clusters with line segments in the fitted model further
+supports the preservation of neighborhood relationships.
+
+The \textbf{digit: 1} example further illustrates the model's ability to
+accurately capture non-linear structures and provide additional
+information. Key findings include a twisted pattern that compresses the
+structure in some projections and long line segments that detect
+anomalies.
+
+Predicting new observations in \kD{} is particularly valuable due to the
+limitations of some NLDR methods, like tSNE, which don't provide a
+straightforward method for prediction. As a result, our approach offers
+a solution that capable of generating predicted \kD{} embedding
+regardless of the NLDR method employed, effectively addressing this
+functional gap.
+
+In conclusion, while our method effectively captures and represents
+high-dimensional data structures, further enhancements could involve
+introducing approaches to bind the data and indicate line segments
+beyond \gD{}. These improvements would help in creating a more accurate
+representation of the data when \gD{} layout is inadequate.
+
+\section{Supplementary Materials}\label{supplementary-materials}
+
+Code, and data for reproducing this paper are available at
+\url{https://github.com/JayaniLakshika/paper-nldr-vis-algorithm}.
 

### 2024-05-31 paper.qmd
Updated version for the discussion
Add error calculations

@@
 
-We define the function $z: \mathbb{R}^{n\times p} \rightarrow \mathbb{R}^{m\times p}$.
-
-For the $i^{th}$ point, $z(i) = \arg\min_{j = 1, \dots, b} \sqrt{\sum_{v=1}^{P}(x_{iv} - C^{(2)}_{jv})^2}$, and assign it to $H_h = \{i| z(i) = h\}$.
+<!-- We define the function $z: \mathbb{R}^{n\times p} \rightarrow \mathbb{R}^{m\times p}$, -->
+
+<!-- For the $i^{th}$ point, $z(i) = \arg\min_{j = 1, \dots, b} \sqrt{\sum_{v=1}^{P}(x_{iv} - C^{(2)}_{jv})^2}$, and assign it to $H_h = \{i| z(i) = h\}$.
 Then, the fitted values for $i$ are defined as $C_{h}^{(2)} = (c_{h1}, c_{h2})$.
+-->
+
+The fitted values refers to the \pD{} mappings of the \kD{} model.
 
@@
 To assess how well our method captures and represents the underlying structure of the \pD{} data, residuals are important.
-Residuals are computed by taking the squared \pD{} Euclidean distance, $\sum_{i = 1}^{n}\sum_{j = 1}^{p} (x_{ij} - C^{(p)}_{hj})^2$.
-Additionally, Mean Squared Error (MSE), $\frac{1}{n}\sum_{i = 1}^{n}\sum_{j = 1}^{p}(x_{ij} - C^{(p)}_{hj})^2$, is computed based on residuals.
+Residuals are computed by taking the squared \pD{} Euclidean distance, $\sum_{h = 1}^{b}\sum_{i = 1}^{n_h}\sum_{j = 1}^{P} (x_{hij} - C^{(p)}_{hj})^2$.
+Additionally, Mean Squared Error (MSE), $\frac{1}{n}\sum_{h = 1}^{b}\sum_{i = 1}^{n_h}\sum_{j = 1}^{P} (x_{hij} - C^{(p)}_{hj})^2$, is computed based on residuals.
+
+### Prediction
+
+The prediction approach involves finding the nearest \kD{} model point for a new \pD{} point.
+We define the function $z: \mathbb{R}^{n\times p} \rightarrow \mathbb{R}^{m\times p}$, where $z(i) = \arg\min_{j = 1, \dots, b} \sqrt{\sum_{v=1}^{P}(x_{iv} - C^{(p)}_{jv})^2}$ maps each \pD{} point to its nearest \pD{} mapping of the model.
+Therefore, the new observation $i$ falls in the hexagon, $H_h = \{i| z(i) = h\}$ and the corresponding \kD{} bin centroids, $C_{h}^{(2)} = (c_{h1}, c_{h2})$ be the predicted values.
 
@@
 
-### Illustration on simulated data
+## Deciding on the most suitable fit
+
+<!--change the tuning parameter-->
+
+## Illustration on simulated data
 

### 2024-05-31 paper.tex
Updated version for the discussion
Add error calculations

@@
 
-We define the function
-\(z: \mathbb{R}^{n\times p} \rightarrow \mathbb{R}^{m\times p}\).
-
-For the \(i^{th}\) point,
-\(z(i) = \arg\min_{j = 1, \dots, b} \sqrt{\sum_{v=1}^{P}(x_{iv} - C^{(2)}_{jv})^2}\),
-and assign it to \(H_h = \{i| z(i) = h\}\). Then, the fitted values for
-\(i\) are defined as \(C_{h}^{(2)} = (c_{h1}, c_{h2})\).
+The fitted values refers to the \pD{} mappings of the \kD{} model.
 
@@
 computed by taking the squared \pD{} Euclidean distance,
-\(\sum_{i = 1}^{n}\sum_{j = 1}^{p} (x_{ij} - C^{(p)}_{hj})^2\).
+\(\sum_{h = 1}^{b}\sum_{i = 1}^{n_h}\sum_{j = 1}^{P} (x_{hij} - C^{(p)}_{hj})^2\).
 Additionally, Mean Squared Error (MSE),
-\(\frac{1}{n}\sum_{i = 1}^{n}\sum_{j = 1}^{p}(x_{ij} - C^{(p)}_{hj})^2\),
+\(\frac{1}{n}\sum_{h = 1}^{b}\sum_{i = 1}^{n_h}\sum_{j = 1}^{P} (x_{hij} - C^{(p)}_{hj})^2\),
 is computed based on residuals.
+
+\subsection{Prediction}\label{prediction}
+
+The prediction approach involves finding the nearest \kD{} model point
+for a new \pD{} point. We define the function
+\(z: \mathbb{R}^{n\times p} \rightarrow \mathbb{R}^{m\times p}\), where
+\(z(i) = \arg\min_{j = 1, \dots, b} \sqrt{\sum_{v=1}^{P}(x_{iv} - C^{(p)}_{jv})^2}\)
+maps each \pD{} point to its nearest \pD{} mapping of the model.
+Therefore, the new observation \(i\) falls in the hexagon,
+\(H_h = \{i| z(i) = h\}\) and the corresponding \kD{} bin centroids,
+\(C_{h}^{(2)} = (c_{h1}, c_{h2})\) be the predicted values.
 
@@
 
-\subsection{Illustration on simulated
+\section{Deciding on the most suitable
+fit}\label{deciding-on-the-most-suitable-fit}
+
+\section{Illustration on simulated
 data}\label{illustration-on-simulated-data}

### 2024-05-31 paper.qmd
Added fitted values scratch

@@
 
-<!-- The prediction approach involves performing the K-nearest neighbors (KNN) algorithm for an unsupervised classification problem.
-First, the nearest \pD{} model point is identified for a given new \pD{} point.
-Then, the corresponding \kD{} centroid mapping for the identified \pD{} model point is determined.
-Finally, the coordinates of this \kD{} centroid are used as the predicted \kD{} embedding for the new \pD{} data point.
-.
--->
-
-<!-- We define the function $z: \mathbb{R}^{n\times p} \rightarrow \mathbb{R}^{m\times p}$, -->
-
-<!-- For the $i^{th}$ point, $z(i) = \arg\min_{j = 1, \dots, b} \sqrt{\sum_{v=1}^{P}(x_{iv} - C^{(2)}_{jv})^2}$, and assign it to $H_h = \{i| z(i) = h\}$.
-Then, the fitted values for $i$ are defined as $C_{h}^{(2)} = (c_{h1}, c_{h2})$.
--->
-
-The fitted values refers to the \pD{} mappings of the \kD{} model.
+The fitted values of the points in $H_h$ refers to the \pD{} mapping $C_{h}^{(p)}$ of the corresponding \kD{} model point $C_{h}^{(2)}$.
 

### 2024-06-03 paper.qmd
Best and bad fit for S-curve

@@
 
-<!--change the tuning parameter-->
+<!--add more details regarding to the how the change happening with different number of bins-->
+
+<!--best model-->
+<!--need to mention why it is the best-->
+
+<!--bad model-->
+<!--need to mention why it is the bad-->
+<!--Not represented the whole data structure beacuse there are some discontinuity in the S-curve-->
+<!-- Why? bad
+
+<!-- add MSE values for the screenshot figure names-->
 

### 2024-06-03 paper.qmd
Added AIC computation to find the best model

@@
 
+<!-- Find the best model-->
+```{r}
+## To initialize number of bins along the x-axis
+bin1_vec_scurve <- 2:25 #sqrt(NROW(training_data_scurve)/2)
+
+error_scurve <- data.frame(matrix(nrow = 0, ncol = 0))
+
+for (xbins in bin1_vec_scurve) {
+
+  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2)$bin2
+
+  scurve_model <- fit_highd_model(
+    training_data = training_data_scurve,
+    emb_df = umap_scurve_scaled,
+    bin1 = xbins,
+    r2 = r2,
+    is_bin_centroid = TRUE,
+    is_rm_lwd_hex = TRUE,
+    col_start_highd = "x"
+  )
+
+  df_bin_centroids_scurve <- scurve_model$df_bin_centroids
+  df_bin_scurve <- scurve_model$df_bin
+
+  ## Compute error
+  error_df <- glance(
+    df_bin_centroids = df_bin_centroids_scurve,
+    df_bin = df_bin_scurve,
+    training_data = training_data_scurve,
+    newdata = NULL,
+    type_NLDR = "UMAP",
+    col_start = "x") |>
+    mutate(bin1 = xbins,
+           bin2 = bin2,
+           b = bin1 * bin2,
+           b_non_empty = NROW(df_bin_centroids_scurve),
+           aic = 2*(NCOL(training_data_scurve)-1)*b + 2*NROW(training_data_scurve)*(NCOL(training_data_scurve)-1)*log(MSE))
+
+  error_scurve <- bind_rows(error_scurve, error_df)
+
+}
+
+best_model_b <- error_scurve |>
+                    filter(aic == min(aic)) |>
+                    pull(b)
+
+mse_scurve_b <- ggplot(error_scurve, 
+                      aes(x = b, 
+                          y = log(MSE))) + 
+  geom_point(size = 1) +
+  geom_line() + 
+  geom_vline(xintercept = best_model_b, linetype="solid", 
+             color = "black", linewidth=0.8, alpha = 0.5) +
+  ylab("log(MSE)") +
+  xlab("b") +
+  theme(axis.text.x = element_text(size = 7),
+        axis.text.y = element_text(size = 7),
+        axis.title.x = element_text(size = 7),
+        axis.title.y = element_text(size = 7, angle = 90))
+
+aic_scurve_b <- ggplot(error_scurve, 
+                      aes(x = b, 
+                          y = aic)) + 
+  geom_point(size = 1) +
+  geom_line() + 
+  geom_vline(xintercept = best_model_b, linetype="solid", 
+             color = "black", linewidth=0.8, alpha = 0.5) +
+  ylab("AIC") +
+  xlab("b") +
+  theme(axis.text.x = element_text(size = 7),
+        axis.text.y = element_text(size = 7),
+        axis.title.x = element_text(size = 7),
+        axis.title.y = element_text(size = 7, angle = 90))
+
+```
+
+```{r}
+#| fig-cap: "MSE from UMAP applied to S-curve dataset with diffferent $b$ choices.
+What is the effective $b$ to create the model?
+The MSE plot have a steep slope at the beginning, indicating that a smaller $b$ causes a larger amount of MSE.
+Then, the slope gradually declines or level off, indicating that a higher $b$ generates a smaller MSE.
+Using the elbow method, it was observed that when the $b$ is set to $420$, the lowest MSE occurred."
+#| label: fig-mse-aic-scurve
+#| fig-pos: H
+#| out-height: 50%
+mse_scurve_b + aic_scurve_b
+```
+
+<!------>
 <!--add more details regarding to the how the change happening with different number of bins-->

### 2024-06-03 paper.tex
Added fitted values scratch
Added AIC computation to find the best model

@@
 
-The fitted values refers to the \pD{} mappings of the \kD{} model.
+The fitted values of the points in \(H_h\) refers to the \pD{} mapping
+\(C_{h}^{(p)}\) of the corresponding \kD{} model point \(C_{h}^{(2)}\).
 
@@
 fit}\label{deciding-on-the-most-suitable-fit}
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=0.5\textheight]{paper_files/figure-pdf/fig-mse-aic-scurve-1.pdf}
+
+}
+
+\caption{\label{fig-mse-aic-scurve}MSE from UMAP applied to S-curve
+dataset with diffferent \(b\) choices. What is the effective \(b\) to
+create the model? The MSE plot have a steep slope at the beginning,
+indicating that a smaller \(b\) causes a larger amount of MSE. Then, the
+slope gradually declines or level off, indicating that a higher \(b\)
+generates a smaller MSE. Using the elbow method, it was observed that
+when the \(b\) is set to \(420\), the lowest MSE occurred.}
+
+\end{figure}%
 

### 2024-06-03 paper.tex
Added S-curve best model

@@
 
+\begin{figure}[H]
+
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/scurve/sc_best_1.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/scurve/sc_best_2.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/scurve/sc_best_3.png}\end{minipage}%
+
+\caption{\label{fig-scurve_sc}Screen shots of the \textbf{langevitour}
+of the S-curve, shows the model-in-data space, a video of the tour
+animation is available at (\url{https://youtu.be/4RFNENX_fXk}).}
+
+\end{figure}%
+
 \section{Illustration on simulated

### 2024-06-03 paper.qmd
Added S-curve best model
Added bad model for S-curve

@@
 
+<!--langevitour with best pD model for S-curve-->
+::: {#fig-scurve_sc layout-ncol="3" fig-pos="H"}
+![](figures/scurve/sc_best_1.png)
+
+![](figures/scurve/sc_best_2.png)
+
+![](figures/scurve/sc_best_3.png)
+
+Screen shots of the **langevitour** of the S-curve, shows the model-in-data space, a video of the tour animation is available at (<https://youtu.be/4RFNENX_fXk>).
+:::
+
+<!--langevitour with bad pD model for S-curve-->
+::: {#fig-scurve_sc layout-ncol="3" fig-pos="H"}
+![](figures/scurve/sc_bad_1.png)
+
+![](figures/scurve/sc_bad_2.png)
+
+![](figures/scurve/sc_bad_3.png)
+
+Screen shots of the **langevitour** of the S-curve, shows the model-in-data space, a video of the tour animation is available at (<https://youtu.be/DHPvhkZ0lUM>).
+:::
+
 <!------>

### 2024-06-03 paper.tex
Added bad model for S-curve

@@
 
+\begin{figure}[H]
+
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/scurve/sc_bad_1.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/scurve/sc_bad_2.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/scurve/sc_bad_3.png}\end{minipage}%
+
+\caption{\label{fig-scurve_sc}Screen shots of the \textbf{langevitour}
+of the S-curve, shows the model-in-data space, a video of the tour
+animation is available at (\url{https://youtu.be/DHPvhkZ0lUM}).}
+
+\end{figure}%
+
 \section{Illustration on simulated

### 2024-06-03 paper.qmd
Added scratch writing for the best and bad models

@@
 
+The best model for the S-curve with UMAP (n_neighbors: $15$) create with $420$ of total number of bins.
+Also, the best model made after removing edges which has the length greater than $0.134$.
+Why this is the best?
+Because AIC plot shows that the minimum occur at $420$.
+Visually, the model capture the geometry of the S-curve well.
+
+On the other hand, a bad model for the S-curve with UMAP is made with $190$ bins, and removing edges which has the length greater than $0.3$.
+Why this is a bad model?
+The MSE error is greater than the best model.
+Visually, the model is not continuously fitted along with the S-curve data, and show some disconnect edges.
+
 <!--langevitour with best pD model for S-curve-->
@@
 
-Screen shots of the **langevitour** of the S-curve, shows the model-in-data space, a video of the tour animation is available at (<https://youtu.be/4RFNENX_fXk>).
+Screen shots of the **langevitour** of the S-curve, shows the model-in-data space, a video of the tour animation is available at (<https://youtu.be/4RFNENX_fXk>). MSE with $0.035$.
 :::
@@
 
-Screen shots of the **langevitour** of the S-curve, shows the model-in-data space, a video of the tour animation is available at (<https://youtu.be/DHPvhkZ0lUM>).
+Screen shots of the **langevitour** of the S-curve, shows the model-in-data space, a video of the tour animation is available at (<https://youtu.be/DHPvhkZ0lUM>). MSE with $0.090$.
 :::
@@
 <!--need to mention why it is the bad-->
-<!--Not represented the whole data structure beacuse there are some discontinuity in the S-curve-->
+<!--Not represented the whole data structure because there are some discontinuity in the S-curve-->
 <!-- Why? bad

### 2024-06-03 paper.tex
Added scratch writing for the best and bad models

--- 2024-06-03 paper.tex
@@
 
+The best model for the S-curve with UMAP (n\_neighbors: \(15\)) create
+with \(420\) of total number of bins. Also, the best model made after
+removing edges which has the length greater than \(0.134\). Why this is
+the best? Because AIC plot shows that the minimum occur at \(420\).
+Visually, the model capture the geometry of the S-curve well.
+
+On the other hand, a bad model for the S-curve with UMAP is made with
+\(190\) bins, and removing edges which has the length greater than
+\(0.3\). Why this is a bad model? The MSE error is greater than the best
+model. Visually, the model is not continuously fitted along with the
+S-curve data, and show some disconnect edges.
+
+\begin{figure}[H]
+
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/scurve/sc_best_1.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/scurve/sc_best_2.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/scurve/sc_best_3.png}\end{minipage}%
+
+\caption{\label{fig-scurve_sc}Screen shots of the \textbf{langevitour}
+of the S-curve, shows the model-in-data space, a video of the tour
+animation is available at (\url{https://youtu.be/4RFNENX_fXk}). MSE with
+\(0.035\).}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/scurve/sc_bad_1.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/scurve/sc_bad_2.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/scurve/sc_bad_3.png}\end{minipage}%
+
+\caption{\label{fig-scurve_sc}Screen shots of the \textbf{langevitour}
+of the S-curve, shows the model-in-data space, a video of the tour
+animation is available at (\url{https://youtu.be/DHPvhkZ0lUM}). MSE with
+\(0.090\).}
+
+\end{figure}%
+
 \section{Illustration on simulated

### 2024-06-03 paper.qmd
Added descriptions for best and poor models

--- 2024-06-03 paper.qmd
@@
 
+The best model for the S-curve using UMAP (n_neighbors: $15$) was created with $420$ bins.
+It was determined to be the best model after removing edges with a length greater than $0.134$.
+The decision was made considering the AIC plot (@fig-mse-aic-scurve), which revealed the occurrence of the minimum at $420$.
+Additionally, visually, the model accurately captures the geometry of the S-curve by fitting the points everywhere (@fig-scurve_sc_best).
+
+In contrast, a poor model for the S-curve with UMAP was created using only $190$ bins and removing edges with a length greater than $0.3$.
+This model is inadequate as it has a higher MSE compared to the best model.
+Visually, the model does not fit the points in some places and mismatches the pattern (@fig-scurve_sc_bad).
+
 <!-- Find the best model-->
@@
 <!--langevitour with best pD model for S-curve-->
-::: {#fig-scurve_sc layout-ncol="3" fig-pos="H"}
+::: {#fig-scurve_sc_best layout-ncol="3" fig-pos="H"}
 ![](figures/scurve/sc_best_1.png)
@@
 <!--langevitour with bad pD model for S-curve-->
-::: {#fig-scurve_sc layout-ncol="3" fig-pos="H"}
+::: {#fig-scurve_sc_bad layout-ncol="3" fig-pos="H"}
 ![](figures/scurve/sc_bad_1.png)

### 2024-06-03 paper.tex
Added descriptions for best and poor models

--- 2024-06-03 paper.tex
@@
 
+The best model for the S-curve using UMAP (n\_neighbors: \(15\)) was
+created with \(420\) bins. It was determined to be the best model after
+removing edges with a length greater than \(0.134\). The decision was
+made considering the AIC plot (Figure~\ref{fig-mse-aic-scurve}), which
+revealed the occurrence of the minimum at \(420\). Additionally,
+visually, the model accurately captures the geometry of the S-curve by
+fitting the points everywhere (Figure~\ref{fig-scurve_sc_best}).
+
+In contrast, a poor model for the S-curve with UMAP was created using
+only \(190\) bins and removing edges with a length greater than \(0.3\).
+This model is inadequate as it has a higher MSE compared to the best
+model. Visually, the model does not fit the points in some places and
+mismatches the pattern (Figure~\ref{fig-scurve_sc_bad}).
+
 \begin{figure}[H]
@@
 
-\caption{\label{fig-scurve_sc}Screen shots of the \textbf{langevitour}
-of the S-curve, shows the model-in-data space, a video of the tour
-animation is available at (\url{https://youtu.be/4RFNENX_fXk}).}
+\caption{\label{fig-scurve_sc_best}Screen shots of the
+\textbf{langevitour} of the S-curve, shows the model-in-data space, a
+video of the tour animation is available at
+(\url{https://youtu.be/4RFNENX_fXk}). MSE with \(0.035\).}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/scurve/sc_bad_1.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/scurve/sc_bad_2.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/scurve/sc_bad_3.png}\end{minipage}%
+
+\caption{\label{fig-scurve_sc_bad}Screen shots of the
+\textbf{langevitour} of the S-curve, shows the model-in-data space, a
+video of the tour animation is available at
+(\url{https://youtu.be/DHPvhkZ0lUM}). MSE with \(0.090\).}
 

### 2024-06-04 paper.qmd
Remove AIC calculations

@@
 
-The best model for the S-curve using UMAP (n_neighbors: $15$) was created with $420$ bins.
+The appropriate model for the S-curve using UMAP (n_neighbors: $15$) was created with $420$ bins.
 It was determined to be the best model after removing edges with a length greater than $0.134$.
-The decision was made considering the AIC plot (@fig-mse-aic-scurve), which revealed the occurrence of the minimum at $420$.
-Additionally, visually, the model accurately captures the geometry of the S-curve by fitting the points everywhere (@fig-scurve_sc_best).
+The decision was made considering that the model accurately captures the geometry of the S-curve by fitting the points everywhere (@fig-scurve_sc_best).
 
 In contrast, a poor model for the S-curve with UMAP was created using only $190$ bins and removing edges with a length greater than $0.3$.
-This model is inadequate as it has a higher MSE compared to the best model.
+This model is inadequate as it has a higher MSE compared to the appropriate model.
 Visually, the model does not fit the points in some places and mismatches the pattern (@fig-scurve_sc_bad).
-
-<!-- Find the best model-->
-```{r}
-## To initialize number of bins along the x-axis
-bin1_vec_scurve <- 2:25 #sqrt(NROW(training_data_scurve)/2)
-
-error_scurve <- data.frame(matrix(nrow = 0, ncol = 0))
-
-for (xbins in bin1_vec_scurve) {
-
-  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2)$bin2
-
-  scurve_model <- fit_highd_model(
-    training_data = training_data_scurve,
-    emb_df = umap_scurve_scaled,
-    bin1 = xbins,
-    r2 = r2,
-    is_bin_centroid = TRUE,
-    is_rm_lwd_hex = TRUE,
-    col_start_highd = "x"
-  )
-
-  df_bin_centroids_scurve <- scurve_model$df_bin_centroids
-  df_bin_scurve <- scurve_model$df_bin
-
-  ## Compute error
-  error_df <- glance(
-    df_bin_centroids = df_bin_centroids_scurve,
-    df_bin = df_bin_scurve,
-    training_data = training_data_scurve,
-    newdata = NULL,
-    type_NLDR = "UMAP",
-    col_start = "x") |>
-    mutate(bin1 = xbins,
-           bin2 = bin2,
-           b = bin1 * bin2,
-           b_non_empty = NROW(df_bin_centroids_scurve),
-           aic = 2*(NCOL(training_data_scurve)-1)*b + 2*NROW(training_data_scurve)*(NCOL(training_data_scurve)-1)*log(MSE))
-
-  error_scurve <- bind_rows(error_scurve, error_df)
-
-}
-
-best_model_b <- error_scurve |>
-                    filter(aic == min(aic)) |>
-                    pull(b)
-
-mse_scurve_b <- ggplot(error_scurve, 
-                      aes(x = b, 
-                          y = log(MSE))) + 
-  geom_point(size = 1) +
-  geom_line() + 
-  geom_vline(xintercept = best_model_b, linetype="solid", 
-             color = "black", linewidth=0.8, alpha = 0.5) +
-  ylab("log(MSE)") +
-  xlab("b") +
-  theme(axis.text.x = element_text(size = 7),
-        axis.text.y = element_text(size = 7),
-        axis.title.x = element_text(size = 7),
-        axis.title.y = element_text(size = 7, angle = 90))
-
-aic_scurve_b <- ggplot(error_scurve, 
-                      aes(x = b, 
-                          y = aic)) + 
-  geom_point(size = 1) +
-  geom_line() + 
-  geom_vline(xintercept = best_model_b, linetype="solid", 
-             color = "black", linewidth=0.8, alpha = 0.5) +
-  ylab("AIC") +
-  xlab("b") +
-  theme(axis.text.x = element_text(size = 7),
-        axis.text.y = element_text(size = 7),
-        axis.title.x = element_text(size = 7),
-        axis.title.y = element_text(size = 7, angle = 90))
-
-```
-
-```{r}
-#| fig-cap: "MSE from UMAP applied to S-curve dataset with diffferent $b$ choices.
-What is the effective $b$ to create the model?
-The MSE plot have a steep slope at the beginning, indicating that a smaller $b$ causes a larger amount of MSE.
-Then, the slope gradually declines or level off, indicating that a higher $b$ generates a smaller MSE.
-Using the elbow method, it was observed that when the $b$ is set to $420$, the lowest MSE occurred."
-#| label: fig-mse-aic-scurve
-#| fig-pos: H
-#| out-height: 50%
-mse_scurve_b + aic_scurve_b
-```
 

### 2024-06-04 paper.tex
Remove AIC calculations

@@
 
-The best model for the S-curve using UMAP (n\_neighbors: \(15\)) was
-created with \(420\) bins. It was determined to be the best model after
-removing edges with a length greater than \(0.134\). The decision was
-made considering the AIC plot (Figure~\ref{fig-mse-aic-scurve}), which
-revealed the occurrence of the minimum at \(420\). Additionally,
-visually, the model accurately captures the geometry of the S-curve by
-fitting the points everywhere (Figure~\ref{fig-scurve_sc_best}).
+The appropriate model for the S-curve using UMAP (n\_neighbors: \(15\))
+was created with \(420\) bins. It was determined to be the best model
+after removing edges with a length greater than \(0.134\). The decision
+was made considering that the model accurately captures the geometry of
+the S-curve by fitting the points everywhere
+(Figure~\ref{fig-scurve_sc_best}).
 
@@
 only \(190\) bins and removing edges with a length greater than \(0.3\).
-This model is inadequate as it has a higher MSE compared to the best
-model. Visually, the model does not fit the points in some places and
-mismatches the pattern (Figure~\ref{fig-scurve_sc_bad}).
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=0.5\textheight]{paper_files/figure-pdf/fig-mse-aic-scurve-1.pdf}
-
-}
-
-\caption{\label{fig-mse-aic-scurve}MSE from UMAP applied to S-curve
-dataset with diffferent \(b\) choices. What is the effective \(b\) to
-create the model? The MSE plot have a steep slope at the beginning,
-indicating that a smaller \(b\) causes a larger amount of MSE. Then, the
-slope gradually declines or level off, indicating that a higher \(b\)
-generates a smaller MSE. Using the elbow method, it was observed that
-when the \(b\) is set to \(420\), the lowest MSE occurred.}
-
-\end{figure}%
+This model is inadequate as it has a higher MSE compared to the
+appropriate model. Visually, the model does not fit the points in some
+places and mismatches the pattern (Figure~\ref{fig-scurve_sc_bad}).
 

### 2024-06-04 paper.qmd
Update

@@
 
-In conclusion, while our method effectively captures and represents high-dimensional data structures, further enhancements could involve introducing approaches to bind the data and indicate line segments beyond \gD{}.
+In conclusion, while our method effectively captures and represents high-dimensional data structures, further enhancements could involve introducing approaches to bind the data, indicate line segments beyond \gD{}, and diagnose the fitted model.
 These improvements would help in creating a more accurate representation of the data when \gD{} layout is inadequate.

### 2024-06-04 paper.tex
Update

@@
 high-dimensional data structures, further enhancements could involve
-introducing approaches to bind the data and indicate line segments
-beyond \gD{}. These improvements would help in creating a more accurate
-representation of the data when \gD{} layout is inadequate.
+introducing approaches to bind the data, indicate line segments beyond
+\gD{}, and diagnose the fitted model. These improvements would help in
+creating a more accurate representation of the data when \gD{} layout is
+inadequate.
 

### 2024-06-10 paper.qmd
fixed figure labels, cannot use _ in a label with latest versions of quarto

@@
 <!--langevitour with pD model for S-curve-->
-::: {#fig-scurve_sc layout-ncol="3" fig-pos="H"}
+::: {#fig-scurve-sc layout-ncol="3" fig-pos="H"}
 ![](figures/scurve/sc_1.png)
@@
 This model is inadequate as it has a higher MSE compared to the appropriate model.
-Visually, the model does not fit the points in some places and mismatches the pattern (@fig-scurve_sc_bad).
+Visually, the model does not fit the points in some places and mismatches the pattern (@fig-scurve-sc-bad).
 
 <!--langevitour with best pD model for S-curve-->
-::: {#fig-scurve_sc_best layout-ncol="3" fig-pos="H"}
+::: {#fig-scurve-sc-best layout-ncol="3" fig-pos="H"}
 ![](figures/scurve/sc_best_1.png)
@@
 <!--langevitour with bad pD model for S-curve-->
-::: {#fig-scurve_sc_bad layout-ncol="3" fig-pos="H"}
+::: {#fig-scurve-sc-bad layout-ncol="3" fig-pos="H"}
 ![](figures/scurve/sc_bad_1.png)
@@
 To investigate which is the reasonable representation to visualize the five spherical Gaussian cluster data or all NLDR methods provide equivalent information, we visualize all the models in \pD{} space.
-Models from all NLDR methods show five well-separated clusters (see @fig-gau1_sc, @fig-gau2_sc, @fig-gau3_sc, @fig-gau4_sc, and @fig-gau5_sc).
+Models from all NLDR methods show five well-separated clusters (see @fig-gau1-sc, @fig-gau2-sc, @fig-gau3-sc, @fig-gau4-sc, and @fig-gau5-sc).
 This suggests that for the five Gaussian cluster dataset, all NLDR methods effectively preserve the global structure.
-tSNE displays clusters with varying densities, indicating their ability to capture within-cluster variation (see @fig-gau1_sc).
-On the other hand, both UMAP, PHATE, PaCMAP and TriMAP show clusters with flat surfaces, suggesting a failure to capture within-cluster variation (see @fig-gau2_sc, @fig-gau3_sc, @fig-gau4_sc and @fig-gau5_sc).
+tSNE displays clusters with varying densities, indicating their ability to capture within-cluster variation (see @fig-gau1-sc).
+On the other hand, both UMAP, PHATE, PaCMAP and TriMAP show clusters with flat surfaces, suggesting a failure to capture within-cluster variation (see @fig-gau2-sc, @fig-gau3-sc, @fig-gau4-sc and @fig-gau5-sc).
 Therefore, UMAP, PHATE, PaCMAP and TriMAP do not capture the local structure as effectively as other methods.
 
-::: {#fig-gau1_sc layout-ncol="3" fig-pos="H"}
+::: {#fig-gau1-sc layout-ncol="3" fig-pos="H"}
 ![](figures/five_gau_clusters/sc_tsne_1.png)
@@
 
-::: {#fig-gau2_sc layout-ncol="3" fig-pos="H"}
+::: {#fig-gau2-sc layout-ncol="3" fig-pos="H"}
 ![](figures/five_gau_clusters/sc_umap_1.png)
@@
 
-::: {#fig-gau3_sc layout-ncol="3" fig-pos="H"}
+::: {#fig-gau3-sc layout-ncol="3" fig-pos="H"}
 ![](figures/five_gau_clusters/sc_phate_1.png)
@@
 
-::: {#fig-gau4_sc layout-ncol="3" fig-pos="H"}
+::: {#fig-gau4-sc layout-ncol="3" fig-pos="H"}
 ![](figures/five_gau_clusters/sc_pacmap_1.png)
@@
 
-::: {#fig-gau5_sc layout-ncol="3" fig-pos="H"}
+::: {#fig-gau5-sc layout-ncol="3" fig-pos="H"}
 ![](figures/five_gau_clusters/sc_trimap_1.png)
@@
 To determine whether the UMAP representation with the parameter choice suggested by @chen2023 preserves the original data structure, we visualize the model constructed with UMAP overlaid on the \pD{} data.
-The figures in @fig-pbmc1_sc show three well-separated clusters, indicating that the suggested UMAP representation preserves the global structure (see @fig-umap-author).
-However, as shown in @fig-pbmc1_sc, these clusters are close to each other in \pD{}.
-Also, non-linear continuity patterns and high-density patches within the clusters are observed (see @fig-pbmc1_sc).
+The figures in @fig-pbmc1-sc show three well-separated clusters, indicating that the suggested UMAP representation preserves the global structure (see @fig-umap-author).
+However, as shown in @fig-pbmc1-sc, these clusters are close to each other in \pD{}.
+Also, non-linear continuity patterns and high-density patches within the clusters are observed (see @fig-pbmc1-sc).
 Therefore, the suggested UMAP representation (see @fig-umap-author) does not accurately preserve the local structure of the PBMC3k dataset.
@@
 
-::: {#fig-pbmc1_sc layout-ncol="3" fig-pos="H"}
+::: {#fig-pbmc1-sc layout-ncol="3" fig-pos="H"}
 ![](figures/pbmc3k/sc_1.png)
@@
 We then fit the model for tSNE, and visualize the resultant model in the \pD{} data space.
-The model shows a quirk, as shown in @fig-pbmc2_sc.
+The model shows a quirk, as shown in @fig-pbmc2-sc.
 All three clusters are connected by an edge except the small and large clusters.
@@
 
-::: {#fig-pbmc2_sc layout-ncol="3" fig-pos="H"}
+::: {#fig-pbmc2-sc layout-ncol="3" fig-pos="H"}
 ![](figures/pbmc3k/sc_4.png)
@@
 
-According to @fig-mnist1_sc1, the non-linear continuous structure observed in the \gD{} representation of PaCMAP (see @fig-pacmap-author) is also visible when visualizing the model overlaid on the data space.
+According to @fig-mnist1-sc1, the non-linear continuous structure observed in the \gD{} representation of PaCMAP (see @fig-pacmap-author) is also visible when visualizing the model overlaid on the data space.
 This indicates that PaCMAP accurately captures the structure of the \pD{} data.
-Additionally, the model shows a twisted pattern within the non-linear structure in \pD{} space (see @fig-mnist1_sc2), which is an additional pattern not visible in the \gD{} representation (see @fig-pacmap-author).
-Furthermore, as shown in @fig-mnist1_sc3, some long edges exist in the \pD{} space that are not recognized as long edges in the \gD{} representation.
+Additionally, the model shows a twisted pattern within the non-linear structure in \pD{} space (see @fig-mnist1-sc2), which is an additional pattern not visible in the \gD{} representation (see @fig-pacmap-author).
+Furthermore, as shown in @fig-mnist1-sc3, some long edges exist in the \pD{} space that are not recognized as long edges in the \gD{} representation.
 However, PaCMAP is a reasonable \gD{} representation of MNIST digit 1 data.
@@
 
-::: {#fig-mnist1_sc layout-ncol="3" fig-pos="H"}
-![](figures/mnist/sc_1.png){#fig-mnist1_sc1}
-
-![](figures/mnist/sc_2.png){#fig-mnist1_sc2}
-
-![](figures/mnist/sc_3.png){#fig-mnist1_sc3}
+::: {#fig-mnist1-sc layout-ncol="3" fig-pos="H"}
+![](figures/mnist/sc_1.png){#fig-mnist1-sc1}
+
+![](figures/mnist/sc_2.png){#fig-mnist1-sc2}
+
+![](figures/mnist/sc_3.png){#fig-mnist1-sc3}
 

### 2024-06-10 paper.tex
fixed figure labels, cannot use _ in a label with latest versions of quarto

@@
 
-\caption{\label{fig-scurve_sc}Screen shots of the \textbf{langevitour}
+\caption{\label{fig-scurve-sc}Screen shots of the \textbf{langevitour}
 of the S-curve, shows the model-in-data space, a video of the tour
@@
 the S-curve by fitting the points everywhere
-(Figure~\ref{fig-scurve_sc_best}).
+(\textbf{?@fig-scurve\_sc\_best}).
 
@@
 appropriate model. Visually, the model does not fit the points in some
-places and mismatches the pattern (Figure~\ref{fig-scurve_sc_bad}).
+places and mismatches the pattern (Figure~\ref{fig-scurve-sc-bad}).
 
@@
 
-\caption{\label{fig-scurve_sc_best}Screen shots of the
+\caption{\label{fig-scurve-sc-best}Screen shots of the
 \textbf{langevitour} of the S-curve, shows the model-in-data space, a
@@
 
-\caption{\label{fig-scurve_sc_bad}Screen shots of the
+\caption{\label{fig-scurve-sc-bad}Screen shots of the
 \textbf{langevitour} of the S-curve, shows the model-in-data space, a
@@
 Models from all NLDR methods show five well-separated clusters (see
-Figure~\ref{fig-gau1_sc}, Figure~\ref{fig-gau2_sc},
-Figure~\ref{fig-gau3_sc}, Figure~\ref{fig-gau4_sc}, and
-Figure~\ref{fig-gau5_sc}). This suggests that for the five Gaussian
+Figure~\ref{fig-gau1-sc}, Figure~\ref{fig-gau2-sc},
+Figure~\ref{fig-gau3-sc}, Figure~\ref{fig-gau4-sc}, and
+Figure~\ref{fig-gau5-sc}). This suggests that for the five Gaussian
 cluster dataset, all NLDR methods effectively preserve the global
@@
 their ability to capture within-cluster variation (see
-Figure~\ref{fig-gau1_sc}). On the other hand, both UMAP, PHATE, PaCMAP
+Figure~\ref{fig-gau1-sc}). On the other hand, both UMAP, PHATE, PaCMAP
 and TriMAP show clusters with flat surfaces, suggesting a failure to
-capture within-cluster variation (see Figure~\ref{fig-gau2_sc},
-Figure~\ref{fig-gau3_sc}, Figure~\ref{fig-gau4_sc} and
-Figure~\ref{fig-gau5_sc}). Therefore, UMAP, PHATE, PaCMAP and TriMAP do
+capture within-cluster variation (see Figure~\ref{fig-gau2-sc},
+Figure~\ref{fig-gau3-sc}, Figure~\ref{fig-gau4-sc} and
+Figure~\ref{fig-gau5-sc}). Therefore, UMAP, PHATE, PaCMAP and TriMAP do
 not capture the local structure as effectively as other methods.
@@
 
-\caption{\label{fig-gau1_sc}Screen shots of the \textbf{langevitour} of
+\caption{\label{fig-gau1-sc}Screen shots of the \textbf{langevitour} of
 the five Gaussian clusters dataset, shows the model with tSNE in \pD{},
@@
 
-\caption{\label{fig-gau2_sc}Screen shots of the \textbf{langevitour} of
+\caption{\label{fig-gau2-sc}Screen shots of the \textbf{langevitour} of
 the five Gaussian clusters dataset, shows the model with UMAP in \pD{},
@@
 
-\caption{\label{fig-gau3_sc}Screen shots of the \textbf{langevitour} of
+\caption{\label{fig-gau3-sc}Screen shots of the \textbf{langevitour} of
 the five Gaussian clusters dataset, shows the model with PHATE in \pD{},
@@
 
-\caption{\label{fig-gau4_sc}Screen shots of the \textbf{langevitour} of
+\caption{\label{fig-gau4-sc}Screen shots of the \textbf{langevitour} of
 the five Gaussian clusters dataset, shows the model with PaCMAP in
@@
 
-\caption{\label{fig-gau5_sc}Screen shots of the \textbf{langevitour} of
+\caption{\label{fig-gau5-sc}Screen shots of the \textbf{langevitour} of
 the five Gaussian clusters dataset, shows the model with TriMAP in
@@
 visualize the model constructed with UMAP overlaid on the \pD{} data.
-The figures in Figure~\ref{fig-pbmc1_sc} show three well-separated
+The figures in Figure~\ref{fig-pbmc1-sc} show three well-separated
 clusters, indicating that the suggested UMAP representation preserves
 the global structure (see Figure~\ref{fig-umap-author}). However, as
-shown in Figure~\ref{fig-pbmc1_sc}, these clusters are close to each
+shown in Figure~\ref{fig-pbmc1-sc}, these clusters are close to each
 other in \pD{}. Also, non-linear continuity patterns and high-density
 patches within the clusters are observed (see
-Figure~\ref{fig-pbmc1_sc}). Therefore, the suggested UMAP representation
+Figure~\ref{fig-pbmc1-sc}). Therefore, the suggested UMAP representation
 (see Figure~\ref{fig-umap-author}) does not accurately preserve the
@@
 
-\caption{\label{fig-pbmc1_sc}Screen shots of the \textbf{langevitour} of
+\caption{\label{fig-pbmc1-sc}Screen shots of the \textbf{langevitour} of
 the PBMC3k data set, shows the model-in-data space, a video of the tour
@@
 \pD{} data space. The model shows a quirk, as shown in
-Figure~\ref{fig-pbmc2_sc}. All three clusters are connected by an edge
+Figure~\ref{fig-pbmc2-sc}. All three clusters are connected by an edge
 except the small and large clusters. Because the clusters are so close
@@
 
-\caption{\label{fig-pbmc2_sc}Screen shots of the \textbf{langevitour} of
+\caption{\label{fig-pbmc2-sc}Screen shots of the \textbf{langevitour} of
 the PBMC3k data set, shows the model in high-D, a video of the tour
@@
 
-According to Figure~\ref{fig-mnist1_sc1}, the non-linear continuous
+According to Figure~\ref{fig-mnist1-sc1}, the non-linear continuous
 structure observed in the \gD{} representation of PaCMAP (see
@@
 a twisted pattern within the non-linear structure in \pD{} space (see
-Figure~\ref{fig-mnist1_sc2}), which is an additional pattern not visible
+Figure~\ref{fig-mnist1-sc2}), which is an additional pattern not visible
 in the \gD{} representation (see Figure~\ref{fig-pacmap-author}).
-Furthermore, as shown in Figure~\ref{fig-mnist1_sc3}, some long edges
+Furthermore, as shown in Figure~\ref{fig-mnist1-sc3}, some long edges
 exist in the \pD{} space that are not recognized as long edges in the
@@
 
-\subcaption{\label{fig-mnist1_sc1}}
+\subcaption{\label{fig-mnist1-sc1}}
 
@@
 
-\subcaption{\label{fig-mnist1_sc2}}
+\subcaption{\label{fig-mnist1-sc2}}
 
@@
 
-\subcaption{\label{fig-mnist1_sc3}}
+\subcaption{\label{fig-mnist1-sc3}}
 
@@
 
-\caption{\label{fig-mnist1_sc}Screen shots of the \textbf{langevitour}
+\caption{\label{fig-mnist1-sc}Screen shots of the \textbf{langevitour}
 of the MNIST digit 1 data set, shows the model-in-data space, a video of

### 2024-06-11 paper.qmd
subscripts used in notation diagram, and caption fixed

@@
 #| tbl-cap: "Summary of notation for describing new methodology."
-# Notations used in the paper
+# Notation used in the paper
 
@@
 ```{r}
+# Code to draw illustration for notation
 hex_param_vis <- ggplot(data = hex_grid, aes(x = x, y = y)) + 
@@
                  arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+
-    annotate("text", x=0.5, y=-0.45, label= "r1", color = "black") +
+    annotate("text", x=0.5, y=-0.45, label= expression(r[1]), color = "black") +
     geom_segment(x = -0.25, y = 0 - 0.05, xend = -0.25, yend = r2 - 0.05,
                  arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ 
-    annotate("text", x=-0.35, y=1, label= "r2", color = "black") +
+    annotate("text", x=-0.35, y=1, label= expression(r[2]), color = "black") +
     geom_segment(x = -0.1 + 0.2087578, y = -0.15, xend = -0.1 + 0.2087578*2, yend = -0.15,
                    arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ # a1 = 0.2087578
-    annotate("text", x=0.2, y=-0.25, label= "a1", color = "black") +
+    annotate("text", x=0.2, y=-0.25, label= expression(a[1]), color = "black") +
     geom_segment(x = -0.15, y = -0.1*r2 + 0.1807896*2, xend = -0.15, yend = -0.1*r2 + 0.1807896*3,
                      arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ # a2 = 0.1807896
-    annotate("text", x=-0.15, y=0.45, label= "a2", color = "black") +
-    annotate("text", x=-0.18, y=-0.45, label= "(s1, s2)", color = "black") ## starting point
+    annotate("text", x=-0.15, y=0.45, label= expression(a[2]), color = "black") +
+    annotate("text", x=-0.18, y=-0.25, label= expression(group("(", list(s[1], s[2]), ")")), color = "black") ## starting point
 ```
@@
 #| label: fig-hex-param
-#| fig-cap: "Notations for hexagonal grid configurations."
+#| fig-cap: "Illustration of the components of the hexagon grid showing the notation used in explanations."
 #| out-height: 30%
-#| fig-width: 2
-#| fig-height: 3
+#| fig-width: 3
+#| fig-height: 4.5
 #| fig-pos: H
@@
 #| fig-pos: H
-#| fig-cap: "$2\\text{-}D$ model generated for different benchamrk values to remove long edges: (a) benchmark = $0.201$ (default), (b) benchmark = $0.35$, and (c) benchmark = $1.1$.
+#| fig-cap: "$2\\text{-}D$ model generated for different benchmark values to remove long edges: (a) benchmark = $0.201$ (default), (b) benchmark = $0.35$, and (c) benchmark = $1.1$.
 What is the effective benchmark value to remove long edges?"

### 2024-06-11 paper.tex
subscripts used in notation diagram, and caption fixed

@@
 
-\caption{\label{fig-hex-param}Notations for hexagonal grid
-configurations.}
+\caption{\label{fig-hex-param}Illustration of the components of the
+hexagon grid showing the notation used in explanations.}
 
@@
 \caption{\label{fig-lg-scurve}\(2\text{-}D\) model generated for
-different benchamrk values to remove long edges: (a) benchmark =
+different benchmark values to remove long edges: (a) benchmark =
 \(0.201\) (default), (b) benchmark = \(0.35\), and (c) benchmark =

### 2024-06-11 paper.qmd
langevitour figure caption

@@
 #| label: fig-hex-param
-#| fig-cap: "Illustration of the components of the hexagon grid showing the notation used in explanations."
+#| fig-cap: "Illustration of the components of the hexagon grid with notation."
 #| out-height: 30%
@@
 
-Screen shots of the **langevitour** of the S-curve, shows the model-in-data space, a video of the tour animation is available at (<https://youtu.be/G1m4q9k--v4>).
+Projections of the $7\text{-}D$ S-curve data (purple points) overlaid with the model (black grid with  green points indicating bin means) constructed from the UMAP layout, with XXX bins (using the **langevitour** software).
+The model closely fits the shape of this data.
+A video of the full tour is available at (<https://youtu.be/G1m4q9k--v4>).
 :::

### 2024-06-11 paper.tex
langevitour figure caption

@@
 \caption{\label{fig-hex-param}Illustration of the components of the
-hexagon grid showing the notation used in explanations.}
+hexagon grid with notation.}
 
@@
 
-\caption{\label{fig-scurve-sc}Screen shots of the \textbf{langevitour}
-of the S-curve, shows the model-in-data space, a video of the tour
-animation is available at (\url{https://youtu.be/G1m4q9k--v4}).}
+\caption{\label{fig-scurve-sc}Projections of the \(7\text{-}D\) S-curve
+data (purple points) overlaid with the model (black grid with green
+points indicating bin means) constructed from the UMAP layout, with XXX
+bins (using the \textbf{langevitour} software). The model closely fits
+the shape of this data. A video of the full tour is available at
+(\url{https://youtu.be/G1m4q9k--v4}).}
 

### 2024-06-11 paper.qmd
explanation of binning

@@
 
-We define the function $u: \mathbb{R}^{n\times 2} \rightarrow \mathbb{R}^{m\times 2}$, which generates the model by assigning the points to bin centroids.
-
-For the $i^{th}$ point, $u(i) = \arg\min_{j = 1, \dots, b} \sqrt{(y_{i1} - C^{(2)}_{j1})^2 + (y_{i2} - C^{(2)}_{j2})^2}$, and falls in $H_h = \{i| u(i) = h\}$.
-
-When $k = 2$, the hexagonal bin centroid is one representation of a bin (@fig-NLDR-scurve (c)). Additionally, it can be the mean of the points within each hexagon.
+Observations are grouped into bins based on their nearest centroid. This can be defined using the function $u: \mathbb{R}^{n\times 2} \rightarrow \mathbb{R}^{m\times 2}$, where
+$u(i) = \arg\min_{j = 1, \dots, b} \sqrt{(y_{i1} - C^{(2)}_{j1})^2 + (y_{i2} - C^{(2)}_{j2})^2}$, mapping observation $i$ into $H_h = \{i| u(i) = h\}$.
+
+When $k = 2$, the hexagonal bin centroid is used (@fig-NLDR-scurve (c)).
+Any measure of center, such as a mean or weighted mean of the points within each hexagon, could be used to represent the hexagon.
 

### 2024-06-11 paper.tex
explanation of binning

@@
 
-We define the function
-\(u: \mathbb{R}^{n\times 2} \rightarrow \mathbb{R}^{m\times 2}\), which
-generates the model by assigning the points to bin centroids.
-
-For the \(i^{th}\) point,
+Observations are grouped into bins based on their nearest centroid. This
+can be defined using the function
+\(u: \mathbb{R}^{n\times 2} \rightarrow \mathbb{R}^{m\times 2}\), where
 \(u(i) = \arg\min_{j = 1, \dots, b} \sqrt{(y_{i1} - C^{(2)}_{j1})^2 + (y_{i2} - C^{(2)}_{j2})^2}\),
-and falls in \(H_h = \{i| u(i) = h\}\).
-
-When \(k = 2\), the hexagonal bin centroid is one representation of a
-bin (Figure~\ref{fig-NLDR-scurve} (c)). Additionally, it can be the mean
-of the points within each hexagon.
+mapping observation \(i\) into \(H_h = \{i| u(i) = h\}\).
+
+When \(k = 2\), the hexagonal bin centroid is used
+(Figure~\ref{fig-NLDR-scurve} (c)). Any measure of center, such as a
+mean or weighted mean of the points within each hexagon, could be used
+to represent the hexagon.
 

### 2024-06-11 paper.qmd
S-curve explanation

@@
 To illustrate the method, we use $7\text{-}D$ simulated data, which we call the "S-curve".
-It is constructed by setting $X_1 = \sin(a), X_2 = U(0, 2), X_3 = \text{sign}(a) \times (\cos(a) - 1), \forall a \in [-3\pi/2, 3\pi/2]$.
+It is constructed by simulating $n=750$ observations from $\theta \sim U(-3\pi/2, 3\pi/2)$, $X_1 = \sin(\theta)$, $X_2 \sim U(0, 2)$ (adding thickness to the S), $X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)$.
 The remaining variables $X_4, X_5, X_6, X_7$ are all uniform error, with small variance.
-We would consider $T=(X_1, X_2, X_3)$ to be the true model.
+We would consider $T=(X_1, X_2, X_3)$ to be the geometric structure (true model) that we hope to capture.
 

### 2024-06-15 paper.qmd
Update choice of bins

@@
 
-<!--add three choice of bins 3, 10, 34-->
+<!--add three choice of bins 7, 10, 14-->
 
@@
   data = umap_scurve_scaled, 
-  bin1 = 3, 
-  r2 = r2)
+  bin1 = 7, 
+  r2 = r2,
+  q = 0.05)
 
@@
   bin1 = 10, 
-  r2 = r2)
+  r2 = r2,
+  q = 0.05)
 
@@
   data = umap_scurve_scaled, 
-  bin1 = 34, 
-  r2 = r2)
+  bin1 = 14, 
+  r2 = r2,
+  q = 0.05)
 
@@
 #| fig-pos: H
-#| fig-cap: "Hexbin plots from different number of bins for the UMAP applied to S-curve data: (a) $b = 15$ (3, 5), (b) $b = 190$ (10, 19), and (c) $b = 2176$ (34, 64).
+#| fig-cap: "Hexbin plots from different number of bins for the UMAP applied to S-curve data: (a) $b = 91$ (7, 13), (b) $b = 190$ (10, 19), and (c) $b = 364$ (14, 26).
 The hexbins are colored based on the density of points, with darker colors indicating higher density and yellow color representing lower density within each bin.

### 2024-06-15 paper.tex
S-curve explanation
Update choice of bins

@@
 To illustrate the method, we use \(7\text{-}D\) simulated data, which we
-call the ``S-curve''. It is constructed by setting
-\(X_1 = \sin(a), X_2 = U(0, 2), X_3 = \text{sign}(a) \times (\cos(a) - 1), \forall a \in [-3\pi/2, 3\pi/2]\).
-The remaining variables \(X_4, X_5, X_6, X_7\) are all uniform error,
-with small variance. We would consider \(T=(X_1, X_2, X_3)\) to be the
-true model.
+call the ``S-curve''. It is constructed by simulating \(n=750\)
+observations from \(\theta \sim U(-3\pi/2, 3\pi/2)\),
+\(X_1 = \sin(\theta)\), \(X_2 \sim U(0, 2)\) (adding thickness to the
+S), \(X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)\). The
+remaining variables \(X_4, X_5, X_6, X_7\) are all uniform error, with
+small variance. We would consider \(T=(X_1, X_2, X_3)\) to be the
+geometric structure (true model) that we hope to capture.
 
@@
 \caption{\label{fig-bins-scurve}Hexbin plots from different number of
-bins for the UMAP applied to S-curve data: (a) \(b = 15\) (3, 5), (b)
-\(b = 190\) (10, 19), and (c) \(b = 2176\) (34, 64). The hexbins are
+bins for the UMAP applied to S-curve data: (a) \(b = 91\) (7, 13), (b)
+\(b = 190\) (10, 19), and (c) \(b = 364\) (14, 26). The hexbins are
 colored based on the density of points, with darker colors indicating

### 2024-06-15 paper.qmd
Add script to bind the best out of 5 methods

@@
   xlab("b") +
-  theme(axis.text.x = element_text(size = 7),
-        axis.text.y = element_text(size = 7),
-        axis.title.x = element_text(size = 7),
-        axis.title.y = element_text(size = 7, angle = 90))
+  theme(axis.text.x = element_text(size = 10),
+        axis.text.y = element_text(size = 10),
+        axis.title.x = element_text(size = 10),
+        axis.title.y = element_text(size = 10, angle = 90))
 ```
@@
   geom_hline(yintercept = benchmark1, color = "black", linewidth=0.8, alpha = 0.5) +
-  theme(axis.title.x = element_text(size = 5),
-        axis.title.y = element_text(size = 5, angle = 90),
-        axis.text.x = element_text(size = 5, angle = 90),
-        axis.text.y = element_text(size = 5))
+  theme(axis.title.x = element_text(size = 10),
+        axis.title.y = element_text(size = 10, angle = 90),
+        axis.text.x = element_text(size = 7, angle = 90),
+        axis.text.y = element_text(size = 10))
 ```
@@
   xlab("benchmark value") +
-  theme(axis.text.x = element_text(size = 5),
-        axis.text.y = element_text(size = 5),
-        axis.title.x = element_text(size = 5),
-        axis.title.y = element_text(size = 5, angle = 90))
+  theme(axis.text.x = element_text(size = 10),
+        axis.text.y = element_text(size = 10),
+        axis.title.x = element_text(size = 10),
+        axis.title.y = element_text(size = 10, angle = 90))
 ```
@@
   ylab(expression(d^{(2)})) +
-  theme(axis.text.x = element_text(size = 5),
-        axis.title.x = element_text(size = 5))
+  theme(axis.text.x = element_text(size = 10),
+        axis.title.x = element_text(size = 10))
 ```

### 2024-06-15 paper.qmd
Find the best and the worst models

@@
 
-Projections of the $7\text{-}D$ S-curve data (purple points) overlaid with the model (black grid with  green points indicating bin means) constructed from the UMAP layout, with XXX bins (using the **langevitour** software).
+Projections of the $7\text{-}D$ S-curve data (purple points) overlaid with the model (black grid with  green points indicating bin means) constructed from the UMAP layout, with 48 non-empty bins (using the **langevitour** software).
 The model closely fits the shape of this data.

### 2024-06-15 paper.tex
Find the best and the worst models

@@
 data (purple points) overlaid with the model (black grid with green
-points indicating bin means) constructed from the UMAP layout, with XXX
-bins (using the \textbf{langevitour} software). The model closely fits
-the shape of this data. A video of the full tour is available at
-(\url{https://youtu.be/G1m4q9k--v4}).}
+points indicating bin means) constructed from the UMAP layout, with 48
+non-empty bins (using the \textbf{langevitour} software). The model
+closely fits the shape of this data. A video of the full tour is
+available at (\url{https://youtu.be/G1m4q9k--v4}).}
 

### 2024-06-15 examples/scurve/2_gen_model_in_data_space.R

@@ -17,13 +17,14 @@ lim2 <- scurve_scaled_obj$lim2
 r2 <- diff(lim2)/diff(lim1)
 
 ## Compute hexbin parameters
-num_bins_x_scurve <- 10
+num_bins_x_scurve <- 12
 
 scurve_model <- fit_highd_model(
   training_data = training_data_scurve,
   emb_df = umap_scurve_scaled,
   bin1 = num_bins_x_scurve,
   r2 = r2,
+  q = 0.07,
   is_bin_centroid = TRUE,
   is_rm_lwd_hex = TRUE,
   col_start_highd = "x"
@@ -56,7 +57,8 @@ benchmark_scurve <- find_lg_benchmark(
 hb_obj_scurve <- hex_binning(
   data = umap_scurve_scaled,
   bin1 = num_bins_x_scurve,
-  r2 = r2)
+  r2 = r2,
+  q = 0.07)
 
 umap_data_with_hb_id <- hb_obj_scurve$data_hb_id
 

### 2024-06-15 paper.qmd
Added best and bad layouts

--- 2024-06-15 paper.qmd
@@
 
-Projections of the $7\text{-}D$ S-curve data (purple points) overlaid with the model (black grid with  green points indicating bin means) constructed from the UMAP layout, with XXX bins (using the **langevitour** software).
+Projections of the $7\text{-}D$ S-curve data (purple points) overlaid with the model (black grid with green points indicating bin means) constructed from the UMAP layout, with 60 non-empty bins (using the **langevitour** software).
 The model closely fits the shape of this data.
@@
 <!--langevitour with best pD model for S-curve--> 
-::: {#fig-scurve-sc-best layout-ncol="3" fig-pos="H"}
-![](figures/scurve/sc_best_1.png)
-
-![](figures/scurve/sc_best_2.png)
-
-![](figures/scurve/sc_best_3.png)
-
-Screen shots of the **langevitour** of the S-curve, shows the model-in-data space, a video of the tour animation is available at (<https://youtu.be/4RFNENX_fXk>). MSE with $0.035$.
-:::
+<!-- ::: {#fig-scurve-sc-best layout-ncol="3" fig-pos="H"} -->
+<!-- ![](figures/scurve/sc_best_1.png) -->
+
+<!-- ![](figures/scurve/sc_best_2.png) -->
+
+<!-- ![](figures/scurve/sc_best_3.png) -->
+
+<!-- Screen shots of the **langevitour** of the S-curve, shows the model-in-data space, a video of the tour animation is available at (<https://youtu.be/4RFNENX_fXk>). MSE with $0.035$. -->
+<!-- ::: -->
 
 <!--langevitour with bad pD model for S-curve--> 
-::: {#fig-scurve-sc-bad layout-ncol="3" fig-pos="H"}
-![](figures/scurve/sc_bad_1.png)
-
-![](figures/scurve/sc_bad_2.png)
-
-![](figures/scurve/sc_bad_3.png)
-
-Screen shots of the **langevitour** of the S-curve, shows the model-in-data space, a video of the tour animation is available at (<https://youtu.be/DHPvhkZ0lUM>). MSE with $0.090$.
-:::
+<!-- ::: {#fig-scurve-sc-bad layout-ncol="3" fig-pos="H"} -->
+<!-- ![](figures/scurve/sc_bad_1.png) -->
+
+<!-- ![](figures/scurve/sc_bad_2.png) -->
+
+<!-- ![](figures/scurve/sc_bad_3.png) -->
+
+<!-- Screen shots of the **langevitour** of the S-curve, shows the model-in-data space, a video of the tour animation is available at (<https://youtu.be/DHPvhkZ0lUM>). MSE with $0.090$. -->
+<!-- ::: -->
 

### 2024-06-15 paper.tex
Added best and bad layouts

--- 2024-06-15 paper.tex
@@
 data (purple points) overlaid with the model (black grid with green
-points indicating bin means) constructed from the UMAP layout, with XXX
-bins (using the \textbf{langevitour} software). The model closely fits
-the shape of this data. A video of the full tour is available at
-(\url{https://youtu.be/G1m4q9k--v4}).}
+points indicating bin means) constructed from the UMAP layout, with 60
+non-empty bins (using the \textbf{langevitour} software). The model
+closely fits the shape of this data. A video of the full tour is
+available at (\url{https://youtu.be/G1m4q9k--v4}).}
 
@@
 appropriate model. Visually, the model does not fit the points in some
-places and mismatches the pattern (Figure~\ref{fig-scurve-sc-bad}).
-
-\begin{figure}[H]
-
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/scurve/sc_best_1.png}\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/scurve/sc_best_2.png}\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/scurve/sc_best_3.png}\end{minipage}%
-
-\caption{\label{fig-scurve-sc-best}Screen shots of the
-\textbf{langevitour} of the S-curve, shows the model-in-data space, a
-video of the tour animation is available at
-(\url{https://youtu.be/4RFNENX_fXk}). MSE with \(0.035\).}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/scurve/sc_bad_1.png}\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/scurve/sc_bad_2.png}\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/scurve/sc_bad_3.png}\end{minipage}%
-
-\caption{\label{fig-scurve-sc-bad}Screen shots of the
-\textbf{langevitour} of the S-curve, shows the model-in-data space, a
-video of the tour animation is available at
-(\url{https://youtu.be/DHPvhkZ0lUM}). MSE with \(0.090\).}
-
-\end{figure}%
+places and mismatches the pattern (\textbf{?@fig-scurve-sc-bad}).
 

### 2024-06-18 paper.qmd
Add sample view and select the positions
Update decide fit

@@
 <!--langevitour with pD model for S-curve-->
-::: {#fig-scurve-sc layout-ncol="3" fig-pos="H"}
+::: {#fig-scurve-sc layout-ncol="4" fig-pos="H"}
+![](figures/scurve/umap_view_copy.png)
+
 ![](figures/scurve/sc_1.png)
@@
 It was determined to be the best model after removing edges with a length greater than $0.134$.
-The decision was made considering that the model accurately captures the geometry of the S-curve by fitting the points everywhere (@fig-scurve_sc_best).
-
-In contrast, a poor model for the S-curve with UMAP was created using only $190$ bins and removing edges with a length greater than $0.3$.
+The decision was made considering that the model accurately captures the geometry of the S-curve.
+But better in some places only (@fig-scurve_sc_best).
+
+In contrast, a poor model for the S-curve with PHATE was created using only $190$ bins and removing edges with a length greater than $0.3$.
 This model is inadequate as it has a higher MSE compared to the appropriate model.
-Visually, the model does not fit the points in some places and mismatches the pattern (@fig-scurve-sc-bad).
+Visually, the model squeezed to the middle of the geometry of the S-curve (@fig-scurve-sc-bad).
 

### 2024-06-18 paper.tex
Add sample view and select the positions
Update decide fit

@@
 
-\begin{minipage}{0.33\linewidth}
+\begin{minipage}{0.25\linewidth}
+\includegraphics{figures/scurve/umap_view_copy.png}\end{minipage}%
+%
+\begin{minipage}{0.25\linewidth}
 \includegraphics{figures/scurve/sc_1.png}\end{minipage}%
 %
-\begin{minipage}{0.33\linewidth}
+\begin{minipage}{0.25\linewidth}
 \includegraphics{figures/scurve/sc_2.png}\end{minipage}%
 %
-\begin{minipage}{0.33\linewidth}
+\begin{minipage}{0.25\linewidth}
 \includegraphics{figures/scurve/sc_3.png}\end{minipage}%
@@
 was made considering that the model accurately captures the geometry of
-the S-curve by fitting the points everywhere
+the S-curve. But better in some places only
 (\textbf{?@fig-scurve\_sc\_best}).
 
-In contrast, a poor model for the S-curve with UMAP was created using
+In contrast, a poor model for the S-curve with PHATE was created using
 only \(190\) bins and removing edges with a length greater than \(0.3\).
 This model is inadequate as it has a higher MSE compared to the
-appropriate model. Visually, the model does not fit the points in some
-places and mismatches the pattern (\textbf{?@fig-scurve-sc-bad}).
+appropriate model. Visually, the model squeezed to the middle of the
+geometry of the S-curve (\textbf{?@fig-scurve-sc-bad}).
 

### 2024-06-23 paper.qmd
fig 3 re-organised

@@
 library(quollr)
+library(tibble)
 library(dplyr)
@@
 library(grid)
+library(conflicted)
+conflicts_prefer(dplyr::filter)
 ```
@@
 hex_grid_with_counts <- 
-  dplyr::left_join(hex_grid,
+  left_join(hex_grid,
             counts_df, 
@@
 #| echo: false
-#| fig-cap: "Key steps for constructing the model on the UMAP layout ($k=2$) of the S-curve data: (a) data, (b) hexagon bins, (c) bin centroids, and (d) triangulated centroids."
+#| fig-cap: "Key steps for constructing the model on the UMAP layout ($k=2$): (a) data, (b) hexagon bins, (c) bin centroids, and (d) triangulated centroids. The S-curve data is shown."
 #| fig-width: 8
@@
 
-It is beneficial to define the algorithm on data having a standard scale.
-Here the variables are scaled to [0, 1], but the upper bound can incorporate the aspect ratio produced by the NLDR $(r_1, r_2, ..., r_k)$, by setting them to $(y_{1,\text{max}}, y_{2,\text{max}}, ..., y_{k,\text{max}})$.
-When $k=2$ which is assumed for hexagon binning, $y_{1,\text{max}}=1$ and $y_{2,\text{max}} = \frac{r_2}{r_1}$, as observed in @fig-NLDR-scurve.
+Because we are working with distances between points, starting with data having a standard scale, e.g.
+[0, 1], is recommended.
+When the aspect ratio of the NLDR $(r_1, r_2, ..., r_k)$ is meaningful, the data can be scaled to take this into account.
+When $k=2$, as in hexagon binning, the default range is $[0, y_{i,\text{max}}], i=1,2$, where $y_{1,\text{max}}=1$ and $y_{2,\text{max}} = \frac{r_2}{r_1}$ (@fig-NLDR-scurve).
 <!-- \times \frac{2}{\sqrt{3}}$.
@@
 # Code to draw illustration for notation
-hex_param_vis <- ggplot(data = hex_grid, aes(x = x, y = y)) + 
-    geom_polygon(fill = "white", color = "#bdbdbd", aes(group = hex_poly_id)) +
-    geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "#31a354", size = 0.9) +
-    geom_point(data = all_centroids_df |> dplyr::filter(hexID == 1), aes(x = c_x, y = c_y), color = "black") + ## starting point
-    geom_rect(aes(xmin = 0 - 0.03, xmax = 1 - 0.03, ymin = 0 - 0.05, ymax = r2- 0.05), fill = "white", color = "black", alpha = 0, linewidth = 0.7) +
-    geom_segment(x = 0 - 0.03, y = -0.35, xend = 1 - 0.03, yend = -0.35,
-                 arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+
-    annotate("text", x=0.5, y=-0.45, label= expression(r[1]), color = "black") +
-    geom_segment(x = -0.25, y = 0 - 0.05, xend = -0.25, yend = r2 - 0.05,
-                 arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ 
-    annotate("text", x=-0.35, y=1, label= expression(r[2]), color = "black") +
-    geom_segment(x = -0.1 + 0.2087578, y = -0.15, xend = -0.1 + 0.2087578*2, yend = -0.15,
-                   arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ # a1 = 0.2087578
-    annotate("text", x=0.2, y=-0.25, label= expression(a[1]), color = "black") +
-    geom_segment(x = -0.15, y = -0.1*r2 + 0.1807896*2, xend = -0.15, yend = -0.1*r2 + 0.1807896*3,
-                     arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ # a2 = 0.1807896
-    annotate("text", x=-0.15, y=0.45, label= expression(a[2]), color = "black") +
-    annotate("text", x=-0.18, y=-0.25, label= expression(group("(", list(s[1], s[2]), ")")), color = "black") ## starting point
+start_pt <- all_centroids_df |> 
+  filter(hexID == 1)
+d_rect <- tibble(x1min = 0, 
+                 x1max = 1,
+                 x2min = 0,
+                 x2max = r2)
+c_start <- tibble(s1 = 0.03, s2 = 0.05)
+a1 <- tibble(x = all_centroids_df$c_x[21],
+             xend = all_centroids_df$c_x[22],
+             y = all_centroids_df$c_y[21],
+             yend = all_centroids_df$c_y[21],
+             label = expression(a[1]))
+a2 <- tibble(x = all_centroids_df$c_x[27],
+             xend = all_centroids_df$c_x[27],
+             y = all_centroids_df$c_y[27],
+             yend = all_centroids_df$c_y[33],
+             label = expression(a[2]))
+hex_param_vis <- ggplot() + 
+    geom_polygon(data = hex_grid, 
+                        aes(x = x, 
+                            y = y, 
+                            group = hex_poly_id),
+                 fill = "white", 
+                 color = "#bdbdbd") +
+    geom_point(data = all_centroids_df, aes(
+      x = c_x, 
+      y = c_y), 
+      color = "#31a354", size = 0.9) +
+    geom_point(data = start_pt, aes(x = c_x, 
+                                    y = c_y), 
+               color = "black") + 
+    geom_rect(data=d_rect, 
+              aes(xmin = x1min - c_start$s1,# - c_start$s1, 
+                  xmax = x1max - c_start$s1,# - c_start$s1, 
+                  ymin = x2min - c_start$s2,# - c_start$s2, 
+                  ymax = x2max - c_start$s2),# - c_start$s2), 
+              fill = "white", 
+              color = "black", 
+              alpha = 0, 
+              linewidth = 0.7) +
+    geom_point(data=d_rect, aes(x=x1min - c_start$s1, 
+                                y=x2min - c_start$s2)) + 
+    geom_point(data=d_rect, aes(x=x1max - c_start$s1, 
+                                y=x2min - c_start$s2)) + 
+    geom_point(data=d_rect, aes(x=x1min - c_start$s1, 
+                                y=x2max - c_start$s2)) + 
+    annotate("text", x=d_rect$x1min - c_start$s1, 
+                     y=d_rect$x2min - c_start$s2,
+                     label = "(0,0)", 
+             hjust=-0.1, vjust=-0.3) + 
+    annotate("text", x=d_rect$x1max - c_start$s1, 
+                     y=d_rect$x2min - c_start$s2,
+                     label = "(0,1)", 
+             hjust=1.1, vjust=-0.3) + 
+    annotate("text", x=d_rect$x1min - c_start$s1, 
+                     y=d_rect$x2max - c_start$s2,
+                     label = expression(group("(", 
+                        list(0, y[2][max]),")")), 
+            hjust=-0.1, vjust=1.2) + 
+    geom_segment(data=d_rect, aes(
+      x = x1min  - c_start$s1, # 0 - 0.03, 
+      y = -0.35, 
+      xend = x1max - c_start$s1, #1 - 0.03, 
+      yend = -0.35), #-0.35),
+      arrow = arrow(length = unit(0.03, "npc"),
+                               ends = "both"), 
+                 color = "black")+
+    annotate("text", x=0.5, y=-0.45, 
+             label = expression(r[1]), color = "black") +
+    geom_segment(data=d_rect, aes(
+      x = -0.25, 
+      y = x2min - c_start$s2, #0 - 0.05, 
+      xend = -0.25, 
+      yend = x2max - c_start$s2), #r2 - 0.05),
+      arrow = arrow(length = unit(0.03, "npc"),
+                       ends = "both"), 
+                 color = "black")+ 
+    annotate("text", x=-0.35, y=1, 
+             label = expression(r[2]), color = "black") +
+    geom_segment(data = a1, aes(
+      x = x, #-0.1 + 0.2087578, 
+      y = y, #-0.15, 
+      xend = xend, #-0.1 + 0.2087578*2, 
+      yend = yend), #-0.15),
+      arrow = arrow(length = unit(0.03, "npc"),
+        ends = "both"), 
+        color = "black")+ # a1 = 0.2087578
+    annotate("text", 
+             x=(a1$x+a1$xend)/2, 
+             y=a1$y, 
+             label = expression(a[1]), 
+             color = "black",
+             vjust = 1.2) +
+    geom_segment(data = a2, aes(
+      x = x, #-0.15, 
+      y = y, #-0.1*r2 + 0.1807896*2, 
+      xend = xend, #-0.15, 
+      yend = yend), #-0.1*r2 + 0.1807896*3),
+      arrow = arrow(length = unit(0.03, "npc"),
+                               ends = "both"), 
+      color = "black") + # a2 = 0.1807896
+    annotate("text", x=a2$x, y=(a2$y+a2$yend)/2, 
+             label = expression(a[2]), 
+             color = "black", hjust=-0.2) +
+    annotate("text", x=-0.18, y=-0.25, 
+      label = expression(group("(", list(s[1], s[2]), ")")),
+      color = "black") 
 ```
@@
 #| label: fig-hex-param
-#| fig-cap: "Illustration of the components of the hexagon grid with notation."
+#| fig-cap: "The components of the hexagon grid illustrating notation."
 #| out-height: 30%
-#| fig-width: 3
-#| fig-height: 4.5
+#| fig-width: 2.5
+#| fig-height: 3.75
 #| fig-pos: H

### 2024-06-23 paper.tex
fig 3 re-organised

@@
 \caption{\label{fig-NLDR-scurve}Key steps for constructing the model on
-the UMAP layout (\(k=2\)) of the S-curve data: (a) data, (b) hexagon
-bins, (c) bin centroids, and (d) triangulated centroids.}
+the UMAP layout (\(k=2\)): (a) data, (b) hexagon bins, (c) bin
+centroids, and (d) triangulated centroids. The S-curve data is shown.}
 
@@
 
-It is beneficial to define the algorithm on data having a standard
-scale. Here the variables are scaled to {[}0, 1{]}, but the upper bound
-can incorporate the aspect ratio produced by the NLDR
-\((r_1, r_2, ..., r_k)\), by setting them to
-\((y_{1,\text{max}}, y_{2,\text{max}}, ..., y_{k,\text{max}})\). When
-\(k=2\) which is assumed for hexagon binning, \(y_{1,\text{max}}=1\) and
-\(y_{2,\text{max}} = \frac{r_2}{r_1}\), as observed in
-Figure~\ref{fig-NLDR-scurve}.
+Because we are working with distances between points, starting with data
+having a standard scale, e.g.~{[}0, 1{]}, is recommended. When the
+aspect ratio of the NLDR \((r_1, r_2, ..., r_k)\) is meaningful, the
+data can be scaled to take this into account. When \(k=2\), as in
+hexagon binning, the default range is \([0, y_{i,\text{max}}], i=1,2\),
+where \(y_{1,\text{max}}=1\) and \(y_{2,\text{max}} = \frac{r_2}{r_1}\)
+(Figure~\ref{fig-NLDR-scurve}).
 
@@
 
-\caption{\label{fig-hex-param}Illustration of the components of the
-hexagon grid with notation.}
+\caption{\label{fig-hex-param}The components of the hexagon grid
+illustrating notation.}
 

### 2024-06-23 paper.qmd
done with 2D

@@
 
-#### Computing hexagon grid configurations
-
-The \gD{} hexagon grid is defined by the number of bins in each direction $(b_1, b_2)$, giving total number of bins as $b = b_1 \times b_2$, and hexagon id, $h = 1, \dots, b$.
+#### Computing hexagon grid configuration
+
+The \gD{} hexagon grid is defined by the number of bins in each direction $(b_1, b_2)$, giving total number of bins as $b = b_1 \times b_2$, and a unique hexagon id, $h = 1, \dots, b$.
 Each hexagon, $H_h$ is uniquely described by centroid, $C_{h}^{(2)} = (c_{h1}, c_{h2})$.
-The lower left position where the grid starts at $(s_1, s_2)$, which correspond to the lowest left centroid.
-The values of $s_i$ need to be below their respective minimum variable values, and could be a full bin lower, to allow a buffer ($q$) corresponding to a full hexagon width ($a_1$) and height ($a_2$) around the data.
-The values of $b_i$ are variables to be computed that define the reduction in size of the data ($n$ to $m$).
+(Note that, we detail the algorithm for constructing the grid and binning the data because we found that there was no readily available implementation that could produce what was needed for this project.)
+
+To make the grid, a buffer parameter ($q$) is set as a proportion of the plot space.
+By default this is 0.1.
+A value should be chosen to provide at buffer of a full hexagon width ($a_1$) and height ($a_2$) for the grid to extend beyond the data.
+The lower left position where the grid starts is defined as $(s_1, s_2)$, corresponding to the centroid of the lowest left hexagon, which should be below the minimum data value.
+This is computed as one buffer unit, $q$ below the minimum data value, by default.
+The user provides the preferred number of bins in the horizontal direction, $b_1$, and the algorithm computes the appropriate number of bins in the vertical direction, $b_2$.
 
@@
 
-Observations are grouped into bins based on their nearest centroid. This can be defined using the function $u: \mathbb{R}^{n\times 2} \rightarrow \mathbb{R}^{m\times 2}$, where
+Observations are grouped into bins based on their nearest centroid.
+This produces a reduction in size of the data from $n$ to $m$, where $m\leq b$ (total number of bins).
+This can be defined using the function $u: \mathbb{R}^{n\times 2} \rightarrow \mathbb{R}^{m\times 2}$, where
 $u(i) = \arg\min_{j = 1, \dots, b} \sqrt{(y_{i1} - C^{(2)}_{j1})^2 + (y_{i2} - C^{(2)}_{j2})^2}$, mapping observation $i$ into $H_h = \{i| u(i) = h\}$.
 
-When $k = 2$, the hexagonal bin centroid is used (@fig-NLDR-scurve (c)).
-Any measure of center, such as a mean or weighted mean of the points within each hexagon, could be used to represent the hexagon.
+By default, the bin centroid is used for describing a hexagon (as done in @fig-NLDR-scurve (c)), but any measure of center, such as a mean or weighted mean of the points within each hexagon, could be used.
+The bin centers, and the binned data, are the two important components needed to render the model representation in high dimensions.
 
@@
 
-#### Indicating neighbors by line segments connecting centroids
+#### Indicating neighborhood
 
@@
 
-### Displaying the model in \pD{}
+### Rendering the model in \pD{}
 

### 2024-06-23 paper.tex
done with 2D

@@
 \subsubsection{Computing hexagon grid
-configurations}\label{computing-hexagon-grid-configurations}
+configuration}\label{computing-hexagon-grid-configuration}
 
@@
 direction \((b_1, b_2)\), giving total number of bins as
-\(b = b_1 \times b_2\), and hexagon id, \(h = 1, \dots, b\). Each
-hexagon, \(H_h\) is uniquely described by centroid,
-\(C_{h}^{(2)} = (c_{h1}, c_{h2})\). The lower left position where the
-grid starts at \((s_1, s_2)\), which correspond to the lowest left
-centroid. The values of \(s_i\) need to be below their respective
-minimum variable values, and could be a full bin lower, to allow a
-buffer (\(q\)) corresponding to a full hexagon width (\(a_1\)) and
-height (\(a_2\)) around the data. The values of \(b_i\) are variables to
-be computed that define the reduction in size of the data (\(n\) to
-\(m\)).
+\(b = b_1 \times b_2\), and a unique hexagon id, \(h = 1, \dots, b\).
+Each hexagon, \(H_h\) is uniquely described by centroid,
+\(C_{h}^{(2)} = (c_{h1}, c_{h2})\). (Note that, we detail the algorithm
+for constructing the grid and binning the data because we found that
+there was no readily available implementation that could produce what
+was needed for this project.)
+
+To make the grid, a buffer parameter (\(q\)) is set as a proportion of
+the plot space. By default this is 0.1. A value should be chosen to
+provide at buffer of a full hexagon width (\(a_1\)) and height (\(a_2\))
+for the grid to extend beyond the data. The lower left position where
+the grid starts is defined as \((s_1, s_2)\), corresponding to the
+centroid of the lowest left hexagon, which should be below the minimum
+data value. This is computed as one buffer unit, \(q\) below the minimum
+data value, by default. The user provides the preferred number of bins
+in the horizontal direction, \(b_1\), and the algorithm computes the
+appropriate number of bins in the vertical direction, \(b_2\).
 
@@
 Observations are grouped into bins based on their nearest centroid. This
-can be defined using the function
+produces a reduction in size of the data from \(n\) to \(m\), where
+\(m\leq b\) (total number of bins). This can be defined using the
+function
 \(u: \mathbb{R}^{n\times 2} \rightarrow \mathbb{R}^{m\times 2}\), where
@@
 
-When \(k = 2\), the hexagonal bin centroid is used
-(Figure~\ref{fig-NLDR-scurve} (c)). Any measure of center, such as a
-mean or weighted mean of the points within each hexagon, could be used
-to represent the hexagon.
-
-\subsubsection{Indicating neighbors by line segments connecting
-centroids}\label{indicating-neighbors-by-line-segments-connecting-centroids}
+By default, the bin centroid is used for describing a hexagon (as done
+in Figure~\ref{fig-NLDR-scurve} (c)), but any measure of center, such as
+a mean or weighted mean of the points within each hexagon, could be
+used. The bin centers, and the binned data, are the two important
+components needed to render the model representation in high dimensions.
+
+\subsubsection{Indicating neighborhood}\label{indicating-neighborhood}
 
@@
 
-\subsection{\texorpdfstring{Displaying the model in
-\pD{}}{Displaying the model in }}\label{displaying-the-model-in}
+\subsection{\texorpdfstring{Rendering the model in
+\pD{}}{Rendering the model in }}\label{rendering-the-model-in}
 

### 2024-06-23 paper.qmd
fig 5 tidied up

@@
 sc_xlims <- c(-0.25, 1.25)
-sc_ylims <- c(-0.4, 1.9)
+sc_ylims <- c(-0.2, 1.9)
 # sc_xlims <- c(-0.25, 1.25)
@@
 
-hex_grid_with_counts_s_curve1 <- full_join(hex_grid1, 
+hex_grid_with_counts_s_curve1 <- full_join(
+  hex_grid1, 
   df_bin_centroids1 |> select(hexID, std_counts), 
@@
 
-hex_grid_coloured_scurve1 <-  ggplot(
-  data = hex_grid_with_counts_s_curve1, 
-  aes(x = x, y = y)) + 
-  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+hex_grid_coloured_scurve1 <- ggplot() + 
   geom_point(data = umap_scurve_scaled, 
@@
              size = 0.5) + 
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  interior_annotation("a") #+
-  #xlim(sc_xlims_b) + ylim(sc_ylims_b)
-
+  geom_polygon(
+    data = hex_grid_with_counts_s_curve1, 
+    aes(x = x, y = y,
+      group = hex_poly_id, 
+      fill = std_counts), color = "grey70", linewidth=0.2) +
+  scale_fill_viridis_c(direction = -1, 
+    na.value = "#ffffff", option = "C") +
+  interior_annotation("a") +
+  xlim(sc_xlims) + ylim(sc_ylims)
 ```
@@
 
-hex_grid_with_counts_s_curve2 <- full_join(hex_grid2, 
+hex_grid_with_counts_s_curve2 <- full_join(
+  hex_grid2, 
   df_bin_centroids2 |> select(hexID, std_counts), 
@@
 
-hex_grid_coloured_scurve2 <-  ggplot(
-  data = hex_grid_with_counts_s_curve2, 
-  aes(x = x, y = y)) + 
-  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+hex_grid_coloured_scurve2 <- ggplot() + 
   geom_point(data = umap_scurve_scaled, 
@@
              size = 0.5) + 
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  geom_polygon(
+    data = hex_grid_with_counts_s_curve2, 
+    aes(x = x, y = y, 
+        group = hex_poly_id, 
+        fill = std_counts), 
+        color = "grey70", 
+        linewidth=0.2) +
+  scale_fill_viridis_c(direction = -1, 
+    na.value = "#ffffff", option = "C") +
   xlim(sc_xlims) + ylim(sc_ylims) +
@@
 
-hex_grid_coloured_scurve3 <-  ggplot(
-  data = hex_grid_with_counts_s_curve3, 
-  aes(x = x, y = y)) + 
-  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+hex_grid_coloured_scurve3 <-  ggplot() + 
   geom_point(data = umap_scurve_scaled, 
@@
              size = 0.5) + 
+  geom_polygon(
+    data = hex_grid_with_counts_s_curve3, 
+    aes(x = x, y = y, 
+        group = hex_poly_id, 
+        fill = std_counts), 
+        color = "grey70", linewidth=0.2) +
   scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +

### 2024-06-23 paper.qmd
re-organised the four parameter diagnostic plots

@@
 
-keywords: [high-dimensional data, dimension reduction, hexagonal binning, low-dimensional manifold, tour, data vizualization, model in the data space]
-keywords-formatted: [high-dimensional data, dimension reduction, hexagonal binning, low-dimensional manifold, tour, data vizualization, model in the data space]
+keywords: [high-dimensional data, dimension reduction, hexagon binning, low-dimensional manifold, tour, data vizualization, model in the data space]
+keywords-formatted: [high-dimensional data, dimension reduction, hexagon binning, low-dimensional manifold, tour, data vizualization, model in the data space]
 
@@
 
-$C_k^{(p)} \equiv (C_{kx_1}, ..., C_{kx_p})$ $p$-D mappings of 2D hexagonal bin centroids of the $k^{th}$ hexagon
+$C_k^{(p)} \equiv (C_{kx_1}, ..., C_{kx_p})$ $p$-D mappings of 2D hexagon bin centroids of the $k^{th}$ hexagon
 
@@
 A simple approach would be to triangulate the points and add edges.
-A more stable approach is to first hexagonally bin the data, reducing it from $n$ to $m\leq n$ observations, and connect the bin centroids.
+A more stable approach is to first hexagonly bin the data, reducing it from $n$ to $m\leq n$ observations, and connect the bin centroids.
 This process serves to reduce some noisiness in the resulting surface shown in \pD{}.
@@
 
-<!--Full hexagonal grid with UMAP data-->
+<!--Full hexagon grid with UMAP data-->
 
@@
 
-## Hexagonal binning to have regular hexagons
+## hexagon binning to have regular hexagons
 hb_obj_scurve <- hex_binning(
@@
 
-<!-- To define a hexagonal grid across the \kD{} space, with hexagons of fixed height ($a_1$), width ($a_2$),  it is necessary to determine how many hexagons should be represented along each axis in the \kD{} space.
+<!-- To define a hexagon grid across the \kD{} space, with hexagons of fixed height ($a_1$), width ($a_2$),  it is necessary to determine how many hexagons should be represented along each axis in the \kD{} space.
 
@@
 <!--
-Define a hexagonal grid across the \kD{} space, using hexagons with fixed height ($a_2$), width ($a_1$).
-Each of the \kD{} points will belong to a hexagonal bin.
+Define a hexagon grid across the \kD{} space, using hexagons with fixed height ($a_2$), width ($a_1$).
+Each of the \kD{} points will belong to a hexagon bin.
 That is, for each $y \in \mathbfit{Y}$, we can (uniquely) identify the hexagon that the point belongs to.
@@
 
-<!--We deliberately separate out the creation of the hexagonal grid from the mapping of points on the grid.-->
+<!--We deliberately separate out the creation of the hexagon grid from the mapping of points on the grid.-->
 
@@
 
-The number of hexagonal bins in the hexagonal grid has a considerable impact on the construction of the \gD{} model, serving as the initial step.
+The number of hexagon bins in the hexagon grid has a considerable impact on the construction of the \gD{} model, serving as the initial step.
 The chosen total number of bins must effectively capture the structure of the NLDR data.
@@
 ```{r}
-## Hexagonal binning to have regular hexagons
+## hexagon binning to have regular hexagons
 hb_obj_scurve1 <- hex_binning(
@@
 ```{r}
-## Hexagonal binning to have regular hexagons
+## hexagon binning to have regular hexagons
 hb_obj_scurve2 <- hex_binning(
@@
 ```{r}
-## Hexagonal binning to have regular hexagons
+## hexagon binning to have regular hexagons
 hb_obj_scurve3 <- hex_binning(
@@
   xlab("b") +
-  theme(axis.text.x = element_text(size = 10),
+  ggtitle("(a)") +
+  theme(aspect.ratio = 0.75,
+        axis.text.x = element_text(size = 10),
         axis.text.y = element_text(size = 10),
         axis.title.x = element_text(size = 10),
-        axis.title.y = element_text(size = 10, angle = 90))
+        axis.title.y = element_text(size = 10, 
+            angle=90),
+        title = element_text(size = 36))
 ```
@@
 #| out-height: 50%
+#| eval: false
 
@@
 
-Once setting up the hexagonal grid with an appropriate number of bins, some hexagonal bins may have few or no data points within them (see @fig-bins-scurve (b)).
-To ensure comprehensive coverage of the NLDR data, it is necessary to select hexagonal bins with a considerable number of data points.
+Once setting up the hexagon grid with an appropriate number of bins, some hexagon bins may have few or no data points within them (see @fig-bins-scurve (b)).
+To ensure comprehensive coverage of the NLDR data, it is necessary to select hexagon bins with a considerable number of data points.
 This involves calculating the number of points within each hexagon.
@@
   geom_quasirandom() + 
-  xlab("hexagonal id") + 
+  geom_hline(yintercept = benchmark1, color = "black", linewidth=0.8, alpha = 0.5) +
+  xlab("hexagon id (re-ordered)") + 
   ylab("standardized bin count") +
-  geom_hline(yintercept = benchmark1, color = "black", linewidth=0.8, alpha = 0.5) +
-  theme(axis.title.x = element_text(size = 10),
+  ggtitle("(b)") +
+  theme(aspect.ratio = 0.75,
+        axis.title.x = element_text(size = 10),
         axis.title.y = element_text(size = 10, angle = 90),
-        axis.text.x = element_text(size = 7, angle = 90),
-        axis.text.y = element_text(size = 10))
+        #axis.text.x = element_text(size = 7, angle = 90),
+        axis.text.x = element_blank(),
+        axis.text.y = element_text(size = 10),
+        title = element_text(size = 36))
 ```
@@
 #| fig-pos: H
+#| eval: false
 
@@
             angle=90) +
+  scale_x_continuous("standardized bin count", 
+         transform = "reverse") + 
   ylab("log(MSE)") +
-  xlab("benchmark value") +
-  theme(axis.text.x = element_text(size = 10),
+  ggtitle("(c)") +
+  theme(aspect.ratio = 0.75,
+        axis.text.x = element_text(size = 10),
         axis.text.y = element_text(size = 10),
         axis.title.x = element_text(size = 10),
-        axis.title.y = element_text(size = 10, angle = 90))
+        axis.title.y = element_text(size = 10, 
+            angle=90),
+        title = element_text(size = 36))
 ```
@@
   coord_flip() + 
-  geom_hline(yintercept = benchmark_scurve, linetype="solid", 
+  geom_hline(yintercept = benchmark_scurve,
+     linetype="solid", 
      color = "black", linewidth=0.8, alpha = 0.5) +
@@
   ylab(expression(d^{(2)})) +
-  theme(axis.text.x = element_text(size = 10),
-        axis.title.x = element_text(size = 10))
+  ggtitle("(d)") +
+  theme(aspect.ratio = 0.75,
+        axis.text.x = element_text(size = 10),
+        axis.title.x = element_text(size = 10),
+        title = element_text(size = 36))
 ```
@@
 #| echo: false
-#| fig-cap: "Distribution of $2\\text{-}D$ Euclidean distances between bin centroids of the triangular mesh generated with S-curve UMAP data."
+#| fig-cap: "Various plots to help assess best number of bins, low density bin and large edge removal.
+Both (a) and (c) show MSE, against number of bins and standardised count.
+A good benchmark value for these parameters is when the MSE drops and then flattens out."
+# Distribution of $2\\text{-}D$ Euclidean distances between bin centroids of the triangular mesh generated with S-curve UMAP data.
 #| label: fig-dist-scurve
-#| out-height: 50%
+#| out-width: 100%
+#| fig-width: 8
+#| fig-height: 6
 #| fig-pos: H
 
-distance_scurve_plot
+mse_scurve_b + cell_count_scurve + 
+mse_scurve_lwd + 
+distance_scurve_plot + plot_layout(ncol=2)
 ```

### 2024-06-23 paper.tex
re-organised the four parameter diagnostic plots

@@
   pdfauthor={Jayani P.G. Lakshika; Dianne Cook; Paul Harrison; Michael Lydeamore; Thiyanga S. Talagala},
-  pdfkeywords={high-dimensional data, dimension reduction, hexagonal
+  pdfkeywords={high-dimensional data, dimension reduction, hexagon
 binning, low-dimensional manifold, tour, data vizualization, model in
@@
 \noindent%
-{\it Keywords:} high-dimensional data, dimension reduction, hexagonal
+{\it Keywords:} high-dimensional data, dimension reduction, hexagon
 binning, low-dimensional manifold, tour, data vizualization, model in
@@
 to triangulate the points and add edges. A more stable approach is to
-first hexagonally bin the data, reducing it from \(n\) to \(m\leq n\)
+first hexagonly bin the data, reducing it from \(n\) to \(m\leq n\)
 observations, and connect the bin centroids. This process serves to
@@
 
-The number of hexagonal bins in the hexagonal grid has a considerable
-impact on the construction of the \gD{} model, serving as the initial
-step. The chosen total number of bins must effectively capture the
-structure of the NLDR data. If the number of bins is too low, the model
-may not be able to capture the structure of the NLDR data effectively
-(see Figure~\ref{fig-bins-scurve} (a)), while if there are too many
-bins, it may result in over-fitting the individual points of the NLDR
-data (see Figure~\ref{fig-bins-scurve} (c)). Therefore, it is important
-to determine an effective number of bins to construct a reasonable model
+The number of hexagon bins in the hexagon grid has a considerable impact
+on the construction of the \gD{} model, serving as the initial step. The
+chosen total number of bins must effectively capture the structure of
+the NLDR data. If the number of bins is too low, the model may not be
+able to capture the structure of the NLDR data effectively (see
+Figure~\ref{fig-bins-scurve} (a)), while if there are too many bins, it
+may result in over-fitting the individual points of the NLDR data (see
+Figure~\ref{fig-bins-scurve} (c)). Therefore, it is important to
+determine an effective number of bins to construct a reasonable model
 (see Figure~\ref{fig-bins-scurve} (b)).
@@
 within this range, covering the minimum to maximum values along both
-axes (see Figure~\ref{fig-mse-scurve-b}).
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=0.5\textheight]{paper_files/figure-pdf/fig-mse-scurve-b-1.pdf}
-
-}
-
-\caption{\label{fig-mse-scurve-b}MSE from UMAP applied to S-curve
-dataset with diffferent \(b\) choices. What is the effective \(b\) to
-create the model? The residual plot have a steep slope at the beginning,
-indicating that a smaller \(b\) causes a larger amount of MSE. Then, the
-slope gradually declines or level off, indicating that a higher \(b\)
-generates a smaller MSE. Using the elbow method, it was observed that
-when the \(b\) is set to \(190\), the lowest MSE occurred.}
-
-\end{figure}%
+axes (see \textbf{?@fig-mse-scurve-b}).
 
@@
 
-Once setting up the hexagonal grid with an appropriate number of bins,
-some hexagonal bins may have few or no data points within them (see
+Once setting up the hexagon grid with an appropriate number of bins,
+some hexagon bins may have few or no data points within them (see
 Figure~\ref{fig-bins-scurve} (b)). To ensure comprehensive coverage of
-the NLDR data, it is necessary to select hexagonal bins with a
+the NLDR data, it is necessary to select hexagon bins with a
 considerable number of data points. This involves calculating the number
@@
   Plot the distribution of the standardized counts (see
-  Figure~\ref{fig-stdcts-scurve}).
+  \textbf{?@fig-stdcts-scurve}).
 \item
@@
 
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=0.5\textheight]{paper_files/figure-pdf/fig-stdcts-scurve-1.pdf}
-
-}
-
-\caption{\label{fig-stdcts-scurve}Distribution of standardize counts by
-hexagons.}
-
-\end{figure}%
-
 The benchmark value for removing low-density hexagons ranges between
@@
 model performance, it's essential to observe the change in MSE as the
-benchmark value increases (see Figure~\ref{fig-mse-scurve-lwd}). The MSE
+benchmark value increases (see \textbf{?@fig-mse-scurve-lwd}). The MSE
 shows a gradual increase as the benchmark value progresses from \(0\) to
@@
 low-density hexagons.
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=0.5\textheight]{paper_files/figure-pdf/fig-mse-scurve-lwd-1.pdf}
-
-}
-
-\caption{\label{fig-mse-scurve-lwd}MSE from UMAP applied to S-curve
-dataset with diffferent bechmark choices. What is the effective bechmark
-value to remove low density hexagons? The residual plot have a steep
-slope at the end, indicating that a smaller bechmark value causes a
-small amount of MSE. Then, the slope gradually increases or level up,
-indicating that a higher bechmark value generates a higher MSE. Using
-the elbow method, it was observed that when the bechmark value is set to
-\(0.242\), the lowest MSE occurred.}
-
-\end{figure}%
 
@@
   Plot the distribution of the 2D Euclidean distances (see
-  Figure~\ref{fig-dist-scurve}).
+  \textbf{?@fig-dist-scurve}).
 \item
@@
 
-\centering{
-
-\includegraphics[width=1\textwidth,height=0.5\textheight]{paper_files/figure-pdf/fig-dist-scurve-1.pdf}
-
-}
-
-\caption{\label{fig-dist-scurve}Distribution of \(2\text{-}D\) Euclidean
-distances between bin centroids of the triangular mesh generated with
-S-curve UMAP data.}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-31-1.pdf}
+
+}
+
+\caption{Various plots to help assess best number of bins, low density
+bin and large edge removal. Both (a) and (c) show MSE, against number of
+bins and standardised count. A good benchmark value for these parameters
+is when the MSE drops and then flattens out.}
 

### 2024-06-23 paper.qmd
Comment geom_point in Figure 5

@@
 hex_grid_coloured_scurve1 <- ggplot() + 
-  geom_point(data = umap_scurve_scaled, 
-             aes(x = UMAP1, y = UMAP2), 
-             alpha = 0.5,
-             size = 0.5) + 
+  # geom_point(data = umap_scurve_scaled, 
+  #            aes(x = UMAP1, y = UMAP2), 
+  #            alpha = 0.5,
+  #            size = 0.5) + 
   geom_polygon(
@@
 hex_grid_coloured_scurve2 <- ggplot() + 
-  geom_point(data = umap_scurve_scaled, 
-             aes(x = UMAP1, y = UMAP2), 
-             alpha = 0.5,
-             size = 0.5) + 
+  # geom_point(data = umap_scurve_scaled, 
+  #            aes(x = UMAP1, y = UMAP2), 
+  #            alpha = 0.5,
+  #            size = 0.5) + 
   geom_polygon(
@@
 hex_grid_coloured_scurve3 <-  ggplot() + 
-  geom_point(data = umap_scurve_scaled, 
-             aes(x = UMAP1, y = UMAP2), 
-             alpha = 0.5,
-             size = 0.5) + 
+  # geom_point(data = umap_scurve_scaled, 
+  #            aes(x = UMAP1, y = UMAP2), 
+  #            alpha = 0.5,
+  #            size = 0.5) + 
   geom_polygon(

### 2024-06-23 paper.qmd
Update hexgrid layout of remove low-density section

--- 2024-06-23 paper.qmd
@@
 hex_grid_coloured_scurve1 <- ggplot() + 
-  geom_point(data = umap_scurve_scaled, 
-             aes(x = UMAP1, y = UMAP2), 
-             alpha = 0.5,
-             size = 0.5) + 
   geom_polygon(
@@
       fill = std_counts), color = "grey70", linewidth=0.2) +
+  geom_point(data = umap_scurve_scaled,
+             aes(x = UMAP1, y = UMAP2),
+             alpha = 0.3,
+             size = 0.5) +
   scale_fill_viridis_c(direction = -1, 
@@
 hex_grid_coloured_scurve2 <- ggplot() + 
-  geom_point(data = umap_scurve_scaled, 
-             aes(x = UMAP1, y = UMAP2), 
-             alpha = 0.5,
-             size = 0.5) + 
   geom_polygon(
@@
         linewidth=0.2) +
+  geom_point(data = umap_scurve_scaled,
+           aes(x = UMAP1, y = UMAP2),
+           alpha = 0.3,
+           size = 0.5) +
   scale_fill_viridis_c(direction = -1, 
@@
 hex_grid_coloured_scurve3 <-  ggplot() + 
-  geom_point(data = umap_scurve_scaled, 
-             aes(x = UMAP1, y = UMAP2), 
-             alpha = 0.5,
-             size = 0.5) + 
   geom_polygon(
@@
         color = "grey70", linewidth=0.2) +
+  geom_point(data = umap_scurve_scaled,
+           aes(x = UMAP1, y = UMAP2),
+           alpha = 0.3,
+           size = 0.5) +
   scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
@@
   aes(x = x, y = y)) + 
-  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+  geom_polygon(aes(group = hex_poly_id, fill = std_counts), 
+            color = "grey70", linewidth=0.2) +
   geom_text(data = hex_grid_with_low_counts, 
@@
   aes(x = x, y = y)) + 
-  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
+  geom_polygon(aes(group = hex_poly_id, fill = std_counts), 
+            color = "grey70", linewidth=0.2) +
   geom_text(data = hex_grid_with_low_counts_neighbor, 

### 2024-06-23 paper.qmd
Update low-density bins

@@
 
-  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2)$bin2
+  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.07)$bin2
 
@@
     r2 = r2,
+    q = 0.07, 
     is_bin_centroid = TRUE,
@@
 ## To initialize effective bins along x
-effective_bin1_scurve <- 10
-
-effective_bin2_scurve <- calc_bins_y(bin1 = effective_bin1_scurve, r2 = r2)$bin2
+effective_bin1_scurve <- 12
+
+effective_bin2_scurve <- calc_bins_y(
+  bin1 = effective_bin1_scurve, 
+  r2 = r2, 
+  q = 0.07)$bin2
 
@@
   r2 = r2,
+  q = 0.07,
   is_bin_centroid = TRUE,
@@
         axis.title.y = element_text(size = 10, angle = 90),
-        #axis.text.x = element_text(size = 7, angle = 90),
         axis.text.x = element_blank(),
@@
   r2 = r2,
+  q = 0.07, 
   is_bin_centroid = TRUE,
@@
 ```{r}
-hex_grid_with_low_counts <- full_join(hex_grid2, 
+## hexagon binning to have regular hexagons
+hb_obj_scurve4 <- hex_binning(
+  data = umap_scurve_scaled, 
+  bin1 = 12, 
+  r2 = r2,
+  q = 0.07)
+
+## Data set with all centroids
+all_centroids_df4 <- hb_obj_scurve4$centroids
+
+## Generate all coordinates of hexagons
+hex_grid4 <- hb_obj_scurve4$hex_poly
+
+## To obtain the standardise counts within hexbins
+counts_df4 <- hb_obj_scurve4$std_cts
+df_bin_centroids4 <- extract_hexbin_centroids(
+  centroids_df = all_centroids_df4, 
+  counts_df = counts_df4) |> 
+  filter(drop_empty == FALSE)
+
+hex_grid_with_counts_s_curve4 <- full_join(
+  hex_grid4, 
+  df_bin_centroids4 |> select(hexID, std_counts), 
+  by = c("hex_poly_id" = "hexID"))
+
+hex_grid_with_low_counts <- full_join(hex_grid4, 
     df_bin_centroids_scurve, 
@@
 hex_grid_rm_lwd_scurve1 <-  ggplot(
-  data = hex_grid_with_counts_s_curve2, 
+  data = hex_grid_with_counts_s_curve4, 
   aes(x = x, y = y)) + 
@@
 ## First define low density bins using first quantile
-df_bin_centroids_low <- df_bin_centroids_scurve |> 
+df_bin_centroids_low <- df_bin_centroids4 |> 
   filter(std_counts <= benchmark1)
@@
 ## Check neighboring bins
-remove_id <- find_low_dens_hex(df_bin_centroids_all = df_bin_centroids_scurve, 
+remove_id <- find_low_dens_hex(df_bin_centroids_all = df_bin_centroids4, 
                                bin1 = effective_bin1_scurve, 
@@
 ## Remove the identified bins
-df_bin_centroids_scurve_removed <- df_bin_centroids_scurve |>
+df_bin_centroids_scurve_removed <- df_bin_centroids4 |>
   filter(hexID %in%remove_id)
 
-df_bin_centroids_scurve_keep  <- df_bin_centroids_scurve |>
+df_bin_centroids_scurve_keep  <- df_bin_centroids4 |>
   filter(!(hexID %in%remove_id))
@@
 hex_grid_rm_lwd_scurve2 <-  ggplot(
-  data = hex_grid_with_counts_s_curve2, 
+  data = hex_grid_with_counts_s_curve4, 
   aes(x = x, y = y)) + 

### 2024-06-23 paper.qmd
Update with xbins 12

--- 2024-06-23 paper.qmd
@@
   data = umap_scurve_scaled, 
-  bin1 = 10, 
+  bin1 = 12, 
   r2 = r2,
-  q = 0.05)
+  q = 0.07)
 
@@
 #| fig-pos: H
-#| fig-cap: "Hexbin plots from different number of bins for the UMAP applied to S-curve data: (a) $b = 91$ (7, 13), (b) $b = 190$ (10, 19), and (c) $b = 364$ (14, 26).
+#| fig-cap: "Hexbin plots from different number of bins for the UMAP applied to S-curve data: (a) $b = 91$ (7, 13), (b) $b = 264$ (12, 22), and (c) $b = 364$ (14, 26).
 The hexbins are colored based on the density of points, with darker colors indicating higher density and yellow color representing lower density within each bin.
@@
   geom_line() + 
-  geom_vline(xintercept = 190, linetype="solid", 
+  geom_vline(xintercept = 264, linetype="solid", 
              color = "black", linewidth=0.8, alpha = 0.5) +
@@
 error_rm_scurve <- data.frame(matrix(nrow = 0, ncol = 0))
-
-scurve_model <- fit_highd_model(
-  training_data = training_data_scurve,
-  emb_df = umap_scurve_scaled,
-  bin1 = effective_bin1_scurve,
-  r2 = r2,
-  is_bin_centroid = TRUE,
-  is_rm_lwd_hex = FALSE,
-  col_start_highd = "x"
-)
-
-df_bin_centroids_scurve <- scurve_model$df_bin_centroids
-df_bin_scurve <- scurve_model$df_bin
 
@@
 ```{r}
+hex_grid_with_counts_s_curve2 <- full_join(
+  hex_grid2, 
+  df_bin_centroids_scurve |> select(hexID, std_counts), 
+  by = c("hex_poly_id" = "hexID"))
+
 hex_grid_with_low_counts <- full_join(hex_grid2, 

### 2024-06-23 paper.qmd
Update title and axis labels of paramater tuning plots

@@
   theme(aspect.ratio = 0.75,
-        axis.text.x = element_text(size = 10),
-        axis.text.y = element_text(size = 10),
-        axis.title.x = element_text(size = 10),
-        axis.title.y = element_text(size = 10, 
+        axis.text.x = element_text(size = 5),
+        axis.text.y = element_text(size = 5),
+        axis.title.x = element_text(size = 5),
+        axis.title.y = element_text(size = 5, 
             angle=90),
@@
   theme(aspect.ratio = 0.75,
-        axis.title.x = element_text(size = 10),
-        axis.title.y = element_text(size = 10, angle = 90),
+        axis.title.x = element_text(size = 5),
+        axis.title.y = element_text(size = 5, angle = 90),
         axis.text.x = element_blank(),
-        axis.text.y = element_text(size = 10),
+        axis.text.y = element_text(size = 5),
         title = element_text(size = 36))
@@
   theme(aspect.ratio = 0.75,
-        axis.text.x = element_text(size = 10),
-        axis.title.x = element_text(size = 10),
+        axis.text.x = element_text(size = 5),
+        axis.title.x = element_text(size = 5),
         title = element_text(size = 36))

### 2024-06-23 paper.qmd
Change the line widths, point size and alpha

@@
   geom_vline(xintercept = 264, linetype="solid", 
-             color = "black", linewidth=0.8, alpha = 0.5) +
+             color = "black", linewidth=0.5, alpha = 0.5) +
   ylab("log(MSE)") +
@@
                                 y = std_counts)) + 
-  geom_quasirandom() + 
-  geom_hline(yintercept = benchmark1, color = "black", linewidth=0.8, alpha = 0.5) +
+  geom_quasirandom(
+    size = 0.3,
+    alpha = 0.5
+  ) + 
+  geom_hline(yintercept = benchmark1, color = "black", linewidth=0.5, alpha = 0.5) +
   xlab("hexagon id (re-ordered)") + 
@@
   geom_vline(xintercept = benchmark1, linetype="solid", 
-             color = "black", linewidth=0.8, alpha = 0.5) +
-  geom_text(data = benchmark_label_df, 
-            aes(x=x, 
-                label= label, 
-                y=y), 
-            colour="black", 
-            angle=90) +
+             color = "black", linewidth=0.5, alpha = 0.5) +
   scale_x_continuous("standardized bin count", 
@@
                             y = distance)) + 
-  geom_quasirandom() + 
+  geom_quasirandom(
+    size = 0.5, 
+    alpha = 0.3
+  ) + 
   ylim(0, max(unlist(distance_scurve$distance))+ 0.5) + 
@@
      linetype="solid", 
-     color = "black", linewidth=0.8, alpha = 0.5) +
-  geom_text(data = benchmark_label_df, 
-            aes(x=x, 
-                label= label, 
-                y=y), 
-            colour="black", 
-            angle=90) +
+     color = "black", linewidth=0.5, alpha = 0.5) +
   ylab(expression(d^{(2)})) +

### 2024-06-23 paper.qmd
Change linewidth and point size in MSE plots

@@
                          y = log(MSE))) + 
-  geom_point(size = 1) +
-  geom_line() + 
+  geom_point(
+    size = 0.6
+    ) +
+  geom_line(
+    linewidth = 0.3
+    ) + 
   geom_vline(xintercept = 264, linetype="solid", 
@@
                          y = log(MSE))) + 
-  geom_point(size = 1) +
-  geom_line() + 
+  geom_point(
+    size = 0.6
+    ) +
+  geom_line(
+    linewidth = 0.3
+    ) + 
   geom_vline(xintercept = benchmark1, linetype="solid", 

### 2024-06-23 paper.qmd
added lines corresponding to sub-optimal bin sizes to MSE plot

--- 2024-06-23 paper.qmd
@@
 hex_grid_coloured_scurve1 <- ggplot() + 
+  geom_point(data = umap_scurve_scaled, 
+             aes(x = UMAP1, y = UMAP2), 
+             alpha = 0.5,
+             size = 0.5) + 
   geom_polygon(
@@
       fill = std_counts), color = "grey70", linewidth=0.2) +
-  geom_point(data = umap_scurve_scaled,
-             aes(x = UMAP1, y = UMAP2),
-             alpha = 0.3,
-             size = 0.5) +
   scale_fill_viridis_c(direction = -1, 
@@
   data = umap_scurve_scaled, 
-  bin1 = 12, 
+  bin1 = 10, 
   r2 = r2,
-  q = 0.07)
+  q = 0.05)
 
@@
 hex_grid_coloured_scurve2 <- ggplot() + 
+  geom_point(data = umap_scurve_scaled, 
+             aes(x = UMAP1, y = UMAP2), 
+             alpha = 0.5,
+             size = 0.5) + 
   geom_polygon(
@@
         linewidth=0.2) +
-  geom_point(data = umap_scurve_scaled,
-           aes(x = UMAP1, y = UMAP2),
-           alpha = 0.3,
-           size = 0.5) +
   scale_fill_viridis_c(direction = -1, 
@@
 hex_grid_coloured_scurve3 <-  ggplot() + 
+  geom_point(data = umap_scurve_scaled, 
+             aes(x = UMAP1, y = UMAP2), 
+             alpha = 0.5,
+             size = 0.5) + 
   geom_polygon(
@@
         color = "grey70", linewidth=0.2) +
-  geom_point(data = umap_scurve_scaled,
-           aes(x = UMAP1, y = UMAP2),
-           alpha = 0.3,
-           size = 0.5) +
   scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
@@
 #| fig-pos: H
-#| fig-cap: "Hexbin plots from different number of bins for the UMAP applied to S-curve data: (a) $b = 91$ (7, 13), (b) $b = 264$ (12, 22), and (c) $b = 364$ (14, 26).
-The hexbins are colored based on the density of points, with darker colors indicating higher density and yellow color representing lower density within each bin.
-What is the number of bins that would be effective?"
+#| fig-cap: "Hexbin density plots of UMAP layout of the S-curve data, using three different bin inputs: (a) $b = 91$ (7, 13), (b) $b = 190$ (10, 19), and (c) $b = 364$ (14, 26).
+Color indicates standardized counts, dark indicating high count and light indicates low count.
+At the smallest bin size the data segregates into two separate groups, suggesting this is too many bins.
+Using the MSE of the model fit in $p-D$ helps decide on a useful choice of number of bins."
 #| fig-width: 6
@@
 
-  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.07)$bin2
+  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2)$bin2
 
@@
     r2 = r2,
-    q = 0.07, 
     is_bin_centroid = TRUE,
@@
   geom_line() + 
-  geom_vline(xintercept = 264, linetype="solid", 
-             color = "black", linewidth=0.5, alpha = 0.5) +
+  geom_vline(xintercept = 190, linetype="solid", 
+             color = "black", linewidth=0.8, alpha = 0.5) +
+  geom_vline(xintercept = 91, linetype=2, 
+             color = "black", linewidth=0.8, alpha = 0.5) +
+  geom_vline(xintercept = 364, linetype=2, 
+             color = "black", linewidth=0.8, alpha = 0.5) +
   ylab("log(MSE)") +
@@
   theme(aspect.ratio = 0.75,
-        axis.text.x = element_text(size = 5),
-        axis.text.y = element_text(size = 5),
-        axis.title.x = element_text(size = 5),
-        axis.title.y = element_text(size = 5, 
+        axis.text.x = element_text(size = 10),
+        axis.text.y = element_text(size = 10),
+        axis.title.x = element_text(size = 10),
+        axis.title.y = element_text(size = 10, 
             angle=90),
@@
 ## To initialize effective bins along x
-effective_bin1_scurve <- 12
-
-effective_bin2_scurve <- calc_bins_y(
-  bin1 = effective_bin1_scurve, 
-  r2 = r2, 
-  q = 0.07)$bin2
+effective_bin1_scurve <- 10
+
+effective_bin2_scurve <- calc_bins_y(bin1 = effective_bin1_scurve, r2 = r2)$bin2
 
@@
   r2 = r2,
-  q = 0.07,
   is_bin_centroid = TRUE,
@@
                                 y = std_counts)) + 
-  geom_quasirandom(
-    size = 0.3,
-    alpha = 0.5
-  ) + 
-  geom_hline(yintercept = benchmark1, color = "black", linewidth=0.5, alpha = 0.5) +
+  geom_quasirandom() + 
+  geom_hline(yintercept = benchmark1, color = "black", linewidth=0.8, alpha = 0.5) +
   xlab("hexagon id (re-ordered)") + 
@@
   theme(aspect.ratio = 0.75,
-        axis.title.x = element_text(size = 5),
-        axis.title.y = element_text(size = 5, angle = 90),
+        axis.title.x = element_text(size = 10),
+        axis.title.y = element_text(size = 10, angle = 90),
+        #axis.text.x = element_text(size = 7, angle = 90),
         axis.text.x = element_blank(),
-        axis.text.y = element_text(size = 5),
+        axis.text.y = element_text(size = 10),
         title = element_text(size = 36))
@@
 error_rm_scurve <- data.frame(matrix(nrow = 0, ncol = 0))
+
+scurve_model <- fit_highd_model(
+  training_data = training_data_scurve,
+  emb_df = umap_scurve_scaled,
+  bin1 = effective_bin1_scurve,
+  r2 = r2,
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = FALSE,
+  col_start_highd = "x"
+)
+
+df_bin_centroids_scurve <- scurve_model$df_bin_centroids
+df_bin_scurve <- scurve_model$df_bin
 
@@
   geom_vline(xintercept = benchmark1, linetype="solid", 
-             color = "black", linewidth=0.5, alpha = 0.5) +
+             color = "black", linewidth=0.8, alpha = 0.5) +
+  geom_text(data = benchmark_label_df, 
+            aes(x=x, 
+                label= label, 
+                y=y), 
+            colour="black", 
+            angle=90) +
   scale_x_continuous("standardized bin count", 
@@
 ```{r}
-hex_grid_with_counts_s_curve2 <- full_join(
-  hex_grid2, 
-  df_bin_centroids_scurve |> select(hexID, std_counts), 
-  by = c("hex_poly_id" = "hexID"))
-
 hex_grid_with_low_counts <- full_join(hex_grid2, 
@@
   aes(x = x, y = y)) + 
-  geom_polygon(aes(group = hex_poly_id, fill = std_counts), 
-            color = "grey70", linewidth=0.2) +
+  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
   geom_text(data = hex_grid_with_low_counts, 
@@
   aes(x = x, y = y)) + 
-  geom_polygon(aes(group = hex_poly_id, fill = std_counts), 
-            color = "grey70", linewidth=0.2) +
+  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
   geom_text(data = hex_grid_with_low_counts_neighbor, 
@@
                             y = distance)) + 
-  geom_quasirandom(
-    size = 0.5, 
-    alpha = 0.3
-  ) + 
+  geom_quasirandom() + 
   ylim(0, max(unlist(distance_scurve$distance))+ 0.5) + 
@@
      linetype="solid", 
-     color = "black", linewidth=0.5, alpha = 0.5) +
+     color = "black", linewidth=0.8, alpha = 0.5) +
+  geom_text(data = benchmark_label_df, 
+            aes(x=x, 
+                label= label, 
+                y=y), 
+            colour="black", 
+            angle=90) +
   ylab(expression(d^{(2)})) +
@@
   theme(aspect.ratio = 0.75,
-        axis.text.x = element_text(size = 5),
-        axis.title.x = element_text(size = 5),
+        axis.text.x = element_text(size = 10),
+        axis.title.x = element_text(size = 10),
         title = element_text(size = 36))
@@
 #| out-width: 100%
-#| fig-width: 8
-#| fig-height: 6
+#| fig-width: 10
+#| fig-height: 8
 #| fig-pos: H

### 2024-06-23 paper.tex
Update with xbins 12
added lines corresponding to sub-optimal bin sizes to MSE plot

@@
 
-\caption{\label{fig-bins-scurve}Hexbin plots from different number of
-bins for the UMAP applied to S-curve data: (a) \(b = 91\) (7, 13), (b)
-\(b = 190\) (10, 19), and (c) \(b = 364\) (14, 26). The hexbins are
-colored based on the density of points, with darker colors indicating
-higher density and yellow color representing lower density within each
-bin. What is the number of bins that would be effective?}
+\caption{\label{fig-bins-scurve}Hexbin density plots of UMAP layout of
+the S-curve data, using three different bin inputs: (a) \(b = 91\) (7,
+13), (b) \(b = 190\) (10, 19), and (c) \(b = 364\) (14, 26). Color
+indicates standardized counts, dark indicating high count and light
+indicates low count. At the smallest bin size the data segregates into
+two separate groups, suggesting this is too many bins. Using the MSE of
+the model fit in \(p-D\) helps decide on a useful choice of number of
+bins.}
 

### 2024-06-23 paper.qmd
Update the Figure title

--- 2024-06-23 paper.qmd
@@
 Both (a) and (c) show MSE, against number of bins and standardised count.
-A good benchmark value for these parameters is when the MSE drops and then flattens out."
-# Distribution of $2\\text{-}D$ Euclidean distances between bin centroids of the triangular mesh generated with S-curve UMAP data.
+A good benchmark value for these parameters is when the MSE drops and then flattens out.
+Plot (b) shows the distribution of stadardised counts of hexagons.
+Plot (c) shows the distribution of $2\\text{-}D$ Euclidean distances between bin centroids, with a good benchmark value for removing large edges would being the distance that shows the first large increase."
+
 #| label: fig-dist-scurve

### 2024-06-23 paper.tex
Update the Figure title

--- 2024-06-23 paper.tex
@@
 bins for the UMAP applied to S-curve data: (a) \(b = 91\) (7, 13), (b)
-\(b = 190\) (10, 19), and (c) \(b = 364\) (14, 26). The hexbins are
+\(b = 264\) (12, 22), and (c) \(b = 364\) (14, 26). The hexbins are
 colored based on the density of points, with darker colors indicating
@@
 bins and standardised count. A good benchmark value for these parameters
-is when the MSE drops and then flattens out.}
+is when the MSE drops and then flattens out. Plot (b) shows the
+distribution of stadardised counts of hexagons. Plot (c) shows the
+distribution of \(2\text{-}D\) Euclidean distances between bin
+centroids, with a good benchmark value for removing large edges would
+being the distance that shows the first large increase.}
 

### 2024-06-23 paper.qmd
Updated figure title for remove low-density hexagons

@@
   centroids_df = all_centroids_df1, 
-  counts_df = counts_df1)
+  counts_df = counts_df1) |>
+  filter(drop_empty == FALSE)
 
@@
   centroids_df = all_centroids_df2, 
-  counts_df = counts_df2)
+  counts_df = counts_df2) |>
+  filter(drop_empty == FALSE)
 
@@
   centroids_df = all_centroids_df3, 
-  counts_df = counts_df3)
+  counts_df = counts_df3) |>
+  filter(drop_empty == FALSE)
 
@@
 #| fig-pos: H
-#| fig-cap: "(a) Identification of low-density hexagons using a benchmark value of $0.242$, and (c) Identification of low-density hexagons considering neighboring bins."
+#| fig-cap: "Two approaches to identify low-density hexagons.
+Hexagon grid (a) marks hexagons with a standardized count less than $0.25$.
+Hexagon grid (b) marks hexagons by considering the densities of neighboring bins.
+Removing hexagons by also investigating the densities of neighboring bins is more reliable for preserving the structure."
 #| fig-width: 9

### 2024-06-23 paper.tex
Updated figure title for remove low-density hexagons

@@
 
-\caption{\label{fig-lwd-scurve}(a) Identification of low-density
-hexagons using a benchmark value of \(0.242\), and (c) Identification of
-low-density hexagons considering neighboring bins.}
+\caption{\label{fig-lwd-scurve}Two approaches to identify low-density
+hexagons. Hexagon grid (a) marks hexagons with a standardized count less
+than \(0.25\). Hexagon grid (b) marks hexagons by considering the
+densities of neighboring bins. Removing hexagons by also investigating
+the densities of neighboring bins is more reliable for preserving the
+structure.}
 

### 2024-06-24 paper.qmd
no changes, i think

--- 2024-06-23 paper.qmd
@@
             angle=90),
-        title = element_text(size = 36))
+        title = element_text(size = 72))
 ```
@@
         axis.text.y = element_text(size = 10),
-        title = element_text(size = 36))
+        title = element_text(size = 72))
 ```
@@
         axis.title.x = element_text(size = 10),
-        title = element_text(size = 36))
+        title = element_text(size = 72))
 ```
@@
 #| out-width: 100%
-#| fig-width: 10
-#| fig-height: 8
+#| fig-width: 12
+#| fig-height: 10
 #| fig-pos: H

### 2024-06-24 paper.qmd

--- 2024-06-23 paper.qmd
@@
                          y = log(MSE))) + 
-  geom_point(
-    size = 0.6
-    ) +
-  geom_line(
-    linewidth = 0.3
-    ) + 
-  geom_vline(xintercept = 264, linetype="solid", 
-             color = "black", linewidth=0.5, alpha = 0.5) +
+  geom_point(size = 1) +
+  geom_line() + 
+  geom_vline(xintercept = 190, linetype="solid", 
+             color = "black", linewidth=0.8, alpha = 0.5) +
+  geom_vline(xintercept = 91, linetype=2, 
+             color = "black", linewidth=0.8, alpha = 0.5) +
+  geom_vline(xintercept = 364, linetype=2, 
+             color = "black", linewidth=0.8, alpha = 0.5) +
   ylab("log(MSE)") +
@@
 #| out-width: 100%
-#| fig-width: 8
-#| fig-height: 6
+#| fig-width: 12
+#| fig-height: 10
 #| fig-pos: H

### 2024-06-24 paper.tex

@@
 
-\caption{\label{fig-bins-scurve}Hexbin plots from different number of
-bins for the UMAP applied to S-curve data: (a) \(b = 91\) (7, 13), (b)
-\(b = 264\) (12, 22), and (c) \(b = 364\) (14, 26). The hexbins are
-colored based on the density of points, with darker colors indicating
-higher density and yellow color representing lower density within each
-bin. What is the number of bins that would be effective?}
+\caption{\label{fig-bins-scurve}Hexbin density plots of UMAP layout of
+the S-curve data, using three different bin inputs: (a) \(b = 91\) (7,
+13), (b) \(b = 190\) (10, 19), and (c) \(b = 364\) (14, 26). Color
+indicates standardized counts, dark indicating high count and light
+indicates low count. At the smallest bin size the data segregates into
+two separate groups, suggesting this is too many bins. Using the MSE of
+the model fit in \(p-D\) helps decide on a useful choice of number of
+bins.}
 

### 2024-06-28 paper.qmd
small changes to headers

@@
 
-#### Error calculation
+#### Residuals and error
 
@@
 
-### pbmc
+### Single-cell gene expression
 <!--
@@
 
-### digits: 1
+### Hand-written digits
 <!--

### 2024-06-28 paper.qmd
lots of structuring

@@
 
+To illustrate the method, we use $7\text{-}D$ simulated data, which we call the "S-curve".
+It is constructed by simulating $n=750$ observations from $\theta \sim U(-3\pi/2, 3\pi/2)$, $X_1 = \sin(\theta)$, $X_2 \sim U(0, 2)$ (adding thickness to the S), $X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)$.
+The remaining variables $X_4, X_5, X_6, X_7$ are all uniform error, with small variance.
+We would consider $T=(X_1, X_2, X_3)$ to be the geometric structure (true model) that we hope to capture.
+
+XXX SHOW THE S-CURVE DATA WITH THE TRUE MODEL OVERLAID HERE, BEFORE THE NLDR VIEWS.
+
 ```{r}
@@
 ```
-
-To illustrate the method, we use $7\text{-}D$ simulated data, which we call the "S-curve".
-It is constructed by simulating $n=750$ observations from $\theta \sim U(-3\pi/2, 3\pi/2)$, $X_1 = \sin(\theta)$, $X_2 \sim U(0, 2)$ (adding thickness to the S), $X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)$.
-The remaining variables $X_4, X_5, X_6, X_7$ are all uniform error, with small variance.
-We would consider $T=(X_1, X_2, X_3)$ to be the geometric structure (true model) that we hope to capture.
 
@@
 
-Projections of the $7\text{-}D$ S-curve data (purple points) overlaid with the model (black grid with green points indicating bin means) constructed from the UMAP layout, with 60 non-empty bins (using the **langevitour** software).
-The model closely fits the shape of this data.
-A video of the full tour is available at (<https://youtu.be/G1m4q9k--v4>).
+XXX Model in 2D, on the UMAP layout, and three views of the fit in projections from $7\text{-}D$, for the S-curve data.
+The model closely fits the shape, but it has breaks which means that it does not adequately capture the surface.
+(The **langevitour** software is used to view the data with a tour, and the full video is available at (<https://youtu.be/G1m4q9k--v4>).
+NEED TO CHANGE FIRST PLOT TO BE 2D WIREFRAME.
 :::
@@
 
-#### Fitted values
-
-The fitted values of the points in $H_h$ refers to the \pD{} mapping $C_{h}^{(p)}$ of the corresponding \kD{} model point $C_{h}^{(2)}$.
-
-#### Residuals and error
-
-To assess how well our method captures and represents the underlying structure of the \pD{} data, residuals are important.
-Residuals are computed by taking the squared \pD{} Euclidean distance, $\sum_{h = 1}^{b}\sum_{i = 1}^{n_h}\sum_{j = 1}^{P} (x_{hij} - C^{(p)}_{hj})^2$.
-Additionally, Mean Squared Error (MSE), $\frac{1}{n}\sum_{h = 1}^{b}\sum_{i = 1}^{n_h}\sum_{j = 1}^{P} (x_{hij} - C^{(p)}_{hj})^2$, is computed based on residuals.
-
-### Prediction
-
-The prediction approach involves finding the nearest \kD{} model point for a new \pD{} point.
-We define the function $z: \mathbb{R}^{n\times p} \rightarrow \mathbb{R}^{m\times p}$, where $z(i) = \arg\min_{j = 1, \dots, b} \sqrt{\sum_{v=1}^{P}(x_{iv} - C^{(p)}_{jv})^2}$ maps each \pD{} point to its nearest \pD{} mapping of the model.
-Therefore, the new observation $i$ falls in the hexagon, $H_h = \{i| z(i) = h\}$ and the corresponding \kD{} bin centroids, $C_{h}^{(2)} = (c_{h1}, c_{h2})$ be the predicted values.
+The model here is similar to a confirmatory factor analysis model (REF), $\widehat{T}(X_1, X_2, X_3) + \Epsilon$.
+The difference between the fitted model and observed values would be considered to be residuals, and for this problem are $7\text{-}D$.
+
+<!--#### Fitted values-->
+
+Observations are associated with their bin center, $C_{h}^{(p)}$, which are also considered to be the *fitted values*.
+These can also be denoted as $\widehat{X}$.
+<!--The fitted values of the points in $H_h$ refers to the \pD{} mapping $C_{h}^{(p)}$ of the corresponding \kD{} model point $C_{h}^{(2)}$.-->
+
+<!--#### Error-->
+
+The error is computed by taking the squared \pD{} Euclidean distance, corresponding to computing the mean squared error (MSE) as:
+
+$$\frac{1}{n}\sum_{h = 1}^{b}\sum_{i = 1}^{n_h}\sum_{j = 1}^{p} (\mathbfit{x}_{hij} - C^{(p)}_{hj})^2$$
+
+where $n$ is XXX, $b$ is XXX, $n_h$ is XXX, $p$ is XXX, $\mathbfit{x}_{hij}$ is XXX. MAKE THIS A NUMBERED EQUATION, SO WE CAN REFER TO IT LATER
+
+### Prediction into \gD{}
+
+A new benefit of this fitted model is that it allows us to now predict a new observation's value in the NLDR, for any method.
+The steps are to determine the closest bin center in \pD{}, $C^{(p)}_{h}$ and predict it to be the centroid of this bin in \gD{}, $C^{(2)}_{h}$.
+This can be written as, let $z(i) = \arg\min_{j = 1, \dots, b} \sqrt{\sum_{v=1}^{p}(x_{iv} - C^{(p)}_{jv})^2}$, then the new observation $i$ falls in the hexagon, $H_h = \{i| z(i) = h\}$ and the corresponding \kD{} bin centroids, $C_{h}^{(2)} = (c_{h1}, c_{h2})$.
+
+<!--The prediction approach involves finding the nearest \kD{} model point for a new \pD{} point.
+We define the function $z: \mathbb{R}^{n\times p} \rightarrow \mathbb{R}^{m\times p}$, where $z(i) = \arg\min_{j = 1, \dots, b} \sqrt{\sum_{v=1}^{p}(x_{iv} - C^{(p)}_{jv})^2}$ maps each \pD{} point to its nearest \pD{} mapping of the model.
+Therefore, the new observation $i$ falls in the hexagon, $H_h = \{i| z(i) = h\}$ and the corresponding \kD{} bin centroids, $C_{h}^{(2)} = (c_{h1}, c_{h2})$ be the predicted values.-->
 
@@
 
-## Deciding on the most suitable fit
+## Best fit
+
+Deciding on the best fit relies on several elements:
+
+- the choice of NLDR method, and the parameters used to create it, and
+- model fit parameters: bin size, low density bin removal, long edge removal.
+
+Comparing the MSE to obtain the best fit is suitable if one starts from the same NLDR representation.
+In theory, because the MSE is computed on \pD{} measuring the fit between model and data it might still be useful to compare different NLDR representations.
+A good NLDR representation should produce a good fit, producing a low MSE if the model fits the data well.
+However, it technically might be quite variable.
 
@@
 
+XXX BEST FIT FOR S-CURVE MIGHT BE A DIFFERENT NLDR. 
+SHOW THE SEVERAL SUB-OPTIMAL FITS.
+
 <!--langevitour with best pD model for S-curve--> 
@@
 
-## Illustration on simulated data
+## A curious difference between t-SNE and UMAP revealer
+
+XXX THIS IS NOT INTERESTING IN ITS CURRENT FORM. THIS IS ONLY ANY INTERESTING EXAMPLE WHEN THE FILLED OUT VS FLAT SHAPES ARE DISCUSSED.
 

### 2024-06-28 paper.tex
small changes to headers
lots of structuring

@@
 
-\begin{figure}
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-NLDR-scurve-1.pdf}
-
-}
-
-\caption{\label{fig-NLDR-scurve}Key steps for constructing the model on
-the UMAP layout (\(k=2\)): (a) data, (b) hexagon bins, (c) bin
-centroids, and (d) triangulated centroids. The S-curve data is shown.}
-
-\end{figure}%
-
 To illustrate the method, we use \(7\text{-}D\) simulated data, which we
@@
 geometric structure (true model) that we hope to capture.
+
+XXX SHOW THE S-CURVE DATA WITH THE TRUE MODEL OVERLAID HERE, BEFORE THE
+NLDR VIEWS.
+
+\begin{figure}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-NLDR-scurve-1.pdf}
+
+}
+
+\caption{\label{fig-NLDR-scurve}Key steps for constructing the model on
+the UMAP layout (\(k=2\)): (a) data, (b) hexagon bins, (c) bin
+centroids, and (d) triangulated centroids. The S-curve data is shown.}
+
+\end{figure}%
 
@@
 
-\caption{\label{fig-scurve-sc}Projections of the \(7\text{-}D\) S-curve
-data (purple points) overlaid with the model (black grid with green
-points indicating bin means) constructed from the UMAP layout, with 60
-non-empty bins (using the \textbf{langevitour} software). The model
-closely fits the shape of this data. A video of the full tour is
-available at (\url{https://youtu.be/G1m4q9k--v4}).}
+\caption{\label{fig-scurve-sc}XXX Model in 2D, on the UMAP layout, and
+three views of the fit in projections from \(7\text{-}D\), for the
+S-curve data. The model closely fits the shape, but it has breaks which
+means that it does not adequately capture the surface. (The
+\textbf{langevitour} software is used to view the data with a tour, and
+the full video is available at (\url{https://youtu.be/G1m4q9k--v4}).
+NEED TO CHANGE FIRST PLOT TO BE 2D WIREFRAME.}
 
@@
 
-\subsubsection{Fitted values}\label{fitted-values}
-
-The fitted values of the points in \(H_h\) refers to the \pD{} mapping
-\(C_{h}^{(p)}\) of the corresponding \kD{} model point \(C_{h}^{(2)}\).
-
-\subsubsection{Error calculation}\label{error-calculation}
-
-To assess how well our method captures and represents the underlying
-structure of the \pD{} data, residuals are important. Residuals are
-computed by taking the squared \pD{} Euclidean distance,
-\(\sum_{h = 1}^{b}\sum_{i = 1}^{n_h}\sum_{j = 1}^{P} (x_{hij} - C^{(p)}_{hj})^2\).
-Additionally, Mean Squared Error (MSE),
-\(\frac{1}{n}\sum_{h = 1}^{b}\sum_{i = 1}^{n_h}\sum_{j = 1}^{P} (x_{hij} - C^{(p)}_{hj})^2\),
-is computed based on residuals.
-
-\subsection{Prediction}\label{prediction}
-
-The prediction approach involves finding the nearest \kD{} model point
-for a new \pD{} point. We define the function
-\(z: \mathbb{R}^{n\times p} \rightarrow \mathbb{R}^{m\times p}\), where
-\(z(i) = \arg\min_{j = 1, \dots, b} \sqrt{\sum_{v=1}^{P}(x_{iv} - C^{(p)}_{jv})^2}\)
-maps each \pD{} point to its nearest \pD{} mapping of the model.
-Therefore, the new observation \(i\) falls in the hexagon,
+The model here is similar to a confirmatory factor analysis model (REF),
+\(\widehat{T}(X_1, X_2, X_3) + \Epsilon\). The difference between the
+fitted model and observed values would be considered to be residuals,
+and for this problem are \(7\text{-}D\).
+
+Observations are associated with their bin center, \(C_{h}^{(p)}\),
+which are also considered to be the \emph{fitted values}. These can also
+be denoted as \(\widehat{X}\).
+
+The error is computed by taking the squared \pD{} Euclidean distance,
+corresponding to computing the mean squared error (MSE) as:
+
+\[\frac{1}{n}\sum_{h = 1}^{b}\sum_{i = 1}^{n_h}\sum_{j = 1}^{p} (\mathbfit{x}_{hij} - C^{(p)}_{hj})^2\]
+
+where \(n\) is XXX, \(b\) is XXX, \(n_h\) is XXX, \(p\) is XXX,
+\(\mathbfit{x}_{hij}\) is XXX. MAKE THIS A NUMBERED EQUATION, SO WE CAN
+REFER TO IT LATER
+
+\subsection{\texorpdfstring{Prediction into
+\gD{}}{Prediction into }}\label{prediction-into}
+
+A new benefit of this fitted model is that it allows us to now predict a
+new observation's value in the NLDR, for any method. The steps are to
+determine the closest bin center in \pD{}, \(C^{(p)}_{h}\) and predict
+it to be the centroid of this bin in \gD{}, \(C^{(2)}_{h}\). This can be
+written as, let
+\(z(i) = \arg\min_{j = 1, \dots, b} \sqrt{\sum_{v=1}^{p}(x_{iv} - C^{(p)}_{jv})^2}\),
+then the new observation \(i\) falls in the hexagon,
 \(H_h = \{i| z(i) = h\}\) and the corresponding \kD{} bin centroids,
-\(C_{h}^{(2)} = (c_{h1}, c_{h2})\) be the predicted values.
+\(C_{h}^{(2)} = (c_{h1}, c_{h2})\).
 
@@
 of points in the grid. Next, bins with a standard count less than a
-benchmark value are removed (see Figure~\ref{fig-lwd-scurve} (a)). There
+benchmark value are removed (see \textbf{?@fig-lwd-scurve} (a)). There
 is no specific rule for selecting a benchmark value. However, the
@@
 neighboring hexagons of the identified low-density bins (see
-Figure~\ref{fig-lwd-scurve} (b)). If neighboring bins also show low
+\textbf{?@fig-lwd-scurve} (b)). If neighboring bins also show low
 counts, only those bins will be removed. The remaining bins are used to
 construct the \gD{} model.
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-lwd-scurve-1.pdf}
-
-}
-
-\caption{\label{fig-lwd-scurve}Two approaches to identify low-density
-hexagons. Hexagon grid (a) marks hexagons with a standardized count less
-than \(0.25\). Hexagon grid (b) marks hexagons by considering the
-densities of neighboring bins. Removing hexagons by also investigating
-the densities of neighboring bins is more reliable for preserving the
-structure.}
-
-\end{figure}%
 
@@
 
-\section{Deciding on the most suitable
-fit}\label{deciding-on-the-most-suitable-fit}
+\section{Best fit}\label{best-fit}
+
+Deciding on the best fit relies on several elements:
+
+\begin{itemize}
+\tightlist
+\item
+  the choice of NLDR method, and the parameters used to create it, and
+\item
+  model fit parameters: bin size, low density bin removal, long edge
+  removal.
+\end{itemize}
+
+Comparing the MSE to obtain the best fit is suitable if one starts from
+the same NLDR representation. In theory, because the MSE is computed on
+\pD{} measuring the fit between model and data it might still be useful
+to compare different NLDR representations. A good NLDR representation
+should produce a good fit, producing a low MSE if the model fits the
+data well. However, it technically might be quite variable.
 
@@
 
-\section{Illustration on simulated
-data}\label{illustration-on-simulated-data}
+XXX BEST FIT FOR S-CURVE MIGHT BE A DIFFERENT NLDR. SHOW THE SEVERAL
+SUB-OPTIMAL FITS.
+
+\section{A curious difference between t-SNE and UMAP
+revealer}\label{a-curious-difference-between-t-sne-and-umap-revealer}
+
+XXX THIS IS NOT INTERESTING IN ITS CURRENT FORM. THIS IS ONLY ANY
+INTERESTING EXAMPLE WHEN THE FILLED OUT VS FLAT SHAPES ARE DISCUSSED.
 
@@
 
-\subsection{pbmc}\label{pbmc}
+\subsection{Single-cell gene
+expression}\label{single-cell-gene-expression}
 
@@
 
-\subsection{digits: 1}\label{digits-1}
+\subsection{Hand-written digits}\label{hand-written-digits}
 

### 2024-06-28 paper.qmd
Page 8 numbered equation

@@
 
-$$\frac{1}{n}\sum_{h = 1}^{b}\sum_{i = 1}^{n_h}\sum_{j = 1}^{p} (\mathbfit{x}_{hij} - C^{(p)}_{hj})^2$$
+$$\frac{1}{n}\sum_{h = 1}^{b}\sum_{i = 1}^{n_h}\sum_{j = 1}^{p} (\mathbfit{x}_{hij} - C^{(p)}_{hj})^2$${#eq-equation1}
 
@@
 ## To initialize number of bins along the x-axis
-bin1_vec_scurve <- 2:19 #sqrt(NROW(training_data_scurve)/2)
+bin1_vec_scurve <- 4:19 #sqrt(NROW(training_data_scurve)/2)
 

### 2024-06-29 paper.qmd
Edit page 8

@@
 The fitted model should be overlaid on the data, to examine the fit relative the spread of the observations.
-While this is straightforward, and commonly done when data is 2D, it is also possible in \pD{}, for many models, when a tour is used.
+While this is straightforward, and commonly done when data is \gD{}, it is also possible in \pD{}, for many models, when a tour is used.
 
@@
 
-XXX Model in 2D, on the UMAP layout, and three views of the fit in projections from $7\text{-}D$, for the S-curve data.
+XXX Model in \gD{}, on the UMAP layout, and three views of the fit in projections from $7\text{-}D$, for the S-curve data.
 The model closely fits the shape, but it has breaks which means that it does not adequately capture the surface.
@@
 
-where $n$ is XXX, $b$ is XXX, $n_h$ is XXX, $p$ is XXX, $\mathbfit{x}_{hij}$ is XXX. MAKE THIS A NUMBERED EQUATION, SO WE CAN REFER TO IT LATER
+where $n$ is the number of observations, $b$ is the number of bins, $n_h$ is the number of observations in $h^{th}$ bin, $p$ is the number of variables, $\mathbfit{x}_{hij}$ is the $j^{th}$ dimensional data of $i^{th}$ observation in $h^{th}$ hexagon.
 
@@
 
-In the 2D layouts generated by various NLDR techniques, as shown in @fig-NLDR-variety-gau, five well-separated clusters are shown.
+In the \gD{} layouts generated by various NLDR techniques, as shown in @fig-NLDR-variety-gau, five well-separated clusters are shown.
 In tSNE (see @fig-NLDR-variety-gau (a)), these clusters appear closely.

### 2024-06-29 paper.tex
Page 8 numbered equation
Edit page 8

@@
 the data, to examine the fit relative the spread of the observations.
-While this is straightforward, and commonly done when data is 2D, it is
-also possible in \pD{}, for many models, when a tour is used.
+While this is straightforward, and commonly done when data is \gD{}, it
+is also possible in \pD{}, for many models, when a tour is used.
 
@@
 
-\caption{\label{fig-scurve-sc}XXX Model in 2D, on the UMAP layout, and
-three views of the fit in projections from \(7\text{-}D\), for the
+\caption{\label{fig-scurve-sc}XXX Model in \gD{}, on the UMAP layout,
+and three views of the fit in projections from \(7\text{-}D\), for the
 S-curve data. The model closely fits the shape, but it has breaks which
@@
 
-\[\frac{1}{n}\sum_{h = 1}^{b}\sum_{i = 1}^{n_h}\sum_{j = 1}^{p} (\mathbfit{x}_{hij} - C^{(p)}_{hj})^2\]
-
-where \(n\) is XXX, \(b\) is XXX, \(n_h\) is XXX, \(p\) is XXX,
-\(\mathbfit{x}_{hij}\) is XXX. MAKE THIS A NUMBERED EQUATION, SO WE CAN
-REFER TO IT LATER
+\begin{equation}\phantomsection\label{eq-equation1}{\frac{1}{n}\sum_{h = 1}^{b}\sum_{i = 1}^{n_h}\sum_{j = 1}^{p} (\mathbfit{x}_{hij} - C^{(p)}_{hj})^2}\end{equation}
+
+where \(n\) is the number of observations, \(b\) is the number of bins,
+\(n_h\) is the number of observations in \(h^{th}\) bin, \(p\) is the
+number of variables, \(\mathbfit{x}_{hij}\) is the \(j^{th}\)
+dimensional data of \(i^{th}\) observation in \(h^{th}\) hexagon.
 
@@
 
-In the 2D layouts generated by various NLDR techniques, as shown in
+In the \gD{} layouts generated by various NLDR techniques, as shown in
 Figure~\ref{fig-NLDR-variety-gau}, five well-separated clusters are

### 2024-06-29 paper.qmd
Update the true model with the data

@@
 
-XXX SHOW THE S-CURVE DATA WITH THE TRUE MODEL OVERLAID HERE, BEFORE THE NLDR VIEWS.
+<!-- XXX SHOW THE S-CURVE DATA WITH THE TRUE MODEL OVERLAID HERE, BEFORE THE NLDR VIEWS. -->
+
+<!--langevitour with true model for S-curve-->
+::: {#fig-scurve-true-sc layout-ncol="3" fig-pos="H"}
+
+![](figures/scurve/sc_true_1.png)
+
+![](figures/scurve/sc_true_2.png)
+
+![](figures/scurve/sc_true_3.png)
+
+Three views of the true model in projections from $7\text{-}D$, for the S-curve data.
+The model closely fits the shape.
+(The **langevitour** software is used to view the data with a tour, and the full video is available at (<https://youtu.be/bcQg4XcRjhc>).
+:::
 
@@
 
+<!--To generate prediction for true-model of S-curve-->
+```{r}
+true_model_df <- read_rds("data/s_curve/true_model.rds")
+true_model_df <- true_model_df |>
+  mutate(ID = row_number())
+
+training_data_scurve <- read_rds("data/s_curve/s_curve_training.rds")
+
+## To initialize effective bins along x
+effective_bin1_scurve <- 12
+
+scurve_model <- fit_highd_model(
+  training_data = training_data_scurve,
+  emb_df = umap_scurve_scaled,
+  bin1 = effective_bin1_scurve,
+  r2 = r2,
+  q = 0.07,
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = FALSE,
+  col_start_highd = "x"
+)
+
+df_bin_centroids_scurve <- scurve_model$df_bin_centroids
+df_bin_scurve <- scurve_model$df_bin
+
+true_pred_df <- predict_emb(
+  test_data = true_model_df, 
+  df_bin_centroids = df_bin_centroids_scurve, 
+  df_bin = df_bin_scurve, 
+  type_NLDR = "UMAP"
+  )
+
+true_pred_df |>
+    ggplot(aes(x = pred_UMAP_1,
+               y = pred_UMAP_2)) +
+    geom_point(
+      alpha=0.5,
+      color = "#33a02c"
+      )
+```
+
 ### Tuning

### 2024-06-29 paper.tex
Update the true model with the data

@@
 
-XXX SHOW THE S-CURVE DATA WITH THE TRUE MODEL OVERLAID HERE, BEFORE THE
-NLDR VIEWS.
+\begin{figure}[H]
+
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/scurve/sc_true_1.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/scurve/sc_true_2.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/scurve/sc_true_3.png}\end{minipage}%
+
+\caption{\label{fig-scurve-true-sc}Three views of the true model in
+projections from \(7\text{-}D\), for the S-curve data. The model closely
+fits the shape. (The \textbf{langevitour} software is used to view the
+data with a tour, and the full video is available at
+(\url{https://youtu.be/bcQg4XcRjhc}).}
+
+\end{figure}%
 
@@
 \(C_{h}^{(2)} = (c_{h1}, c_{h2})\).
+
+\begin{center}
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-15-1.pdf}
+\end{center}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-31-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-32-1.pdf}
 

### 2024-06-29 examples/scurve/9_gen_prediction_true_model_with_umap.R

@@ -1,0 +1,52 @@
+library(readr)
+library(umap)
+library(dplyr)
+
+true_model <- read_rds("data/s_curve/true_model.rds") |>
+  mutate(ID = row_number())
+
+# ### UMAP
+#
+# UMAP_fit <- umap(true_model |> select(-ID), n_neighbors = 15, n_components =  2)
+#
+# UMAP_data_gau <- UMAP_fit$layout |>
+#   as.data.frame()
+# names(UMAP_data_gau)[1:(ncol(UMAP_data_gau))] <- paste0(rep("UMAP",(ncol(UMAP_data_gau))), 1:(ncol(UMAP_data_gau)))
+#
+# UMAP_data_gau <- UMAP_data_gau |>
+#   mutate(ID = row_number())
+#
+# write_rds(UMAP_data_gau, file = "data/s_curve/umap_data_true_model.rds")
+#
+# scaled_umap <- gen_scaled_data(data = UMAP_data_gau)
+#
+# s_curve_noise_umap_scaled <- scaled_umap$scaled_nldr
+# lim1 <- scaled_umap$lim1
+# lim2 <- scaled_umap$lim2
+# r2 <- diff(lim2)/diff(lim1)
+
+# model <- fit_highd_model(training_data = true_model,
+#                          emb_df = s_curve_noise_umap_scaled,
+#                          bin1 = 12, r2 = r2,
+#                          col_start_highd = "x")
+# df_bin_centroids <- model$df_bin_centroids
+# df_bin <- model$df_bin
+
+pred_df_training <- predict_emb(test_data = true_model,
+                             df_bin_centroids = df_bin_centroids_scurve,
+                             df_bin = df_bin_scurve, type_NLDR = "UMAP")
+glimpse(pred_df_training)
+
+s_curve_noise_umap_scaled |>
+    ggplot(aes(x = UMAP1,
+               y = UMAP2,
+               label = ID))+
+    geom_point(alpha=0.5) +
+    geom_point(data = pred_df_training, aes(x = pred_UMAP_1, y = pred_UMAP_2),
+               color = "red", alpha=0.5) +
+    coord_equal() +
+    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold"),
+          axis.text = element_text(size = 5),
+          axis.title = element_text(size = 7))
+
+

### 2024-06-29 paper.qmd
Update the five Gaussian cluster 2D layout with saved figures
Update the views of Five Gaussian clusters

@@
 
-## A curious difference between t-SNE and UMAP revealer
+## A curious difference between t-SNE, UMAP and PaCMAP revealer
 
 XXX THIS IS NOT INTERESTING IN ITS CURRENT FORM. THIS IS ONLY ANY INTERESTING EXAMPLE WHEN THE FILLED OUT VS FLAT SHAPES ARE DISCUSSED.
+
+::: {#fig-gau-tsne-sc layout-ncol="3" fig-pos="H"}
+![](figures/five_gau_clusters/tsne_layout.png)
+
+![](figures/five_gau_clusters/2d_model_tsne.png)
+
+![](figures/five_gau_clusters/sc_tsne_3.png)
+
+Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with tSNE in \pD{}, a video of the tour animation is available at (<https://youtu.be/RASEE7N5MbM>).
+:::
+
+::: {#fig-gau-umap-sc layout-ncol="3" fig-pos="H"}
+![](figures/five_gau_clusters/umap_layout.png)
+
+![](figures/five_gau_clusters/2d_model_umap.png)
+
+![](figures/five_gau_clusters/sc_umap_3.png)
+
+Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with UMAP in \pD{}, a video of the tour animation is available at (<https://youtu.be/iG4bCPkJilw>).
+:::
+
+::: {#fig-gau-pacmap-sc layout-ncol="3" fig-pos="H"}
+![](figures/five_gau_clusters/pacmap_layout.png)
+
+![](figures/five_gau_clusters/2d_model_pacmap.png)
+
+![](figures/five_gau_clusters/sc_pacmap_2.png)
+
+Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with PaCMAP in \pD{}, a video of the tour animation is available at (<https://youtu.be/z07cKXi8EJQ>).
+:::
 

### 2024-06-29 paper.tex
Update the five Gaussian cluster 2D layout with saved figures
Update the views of Five Gaussian clusters

@@
 
-\section{A curious difference between t-SNE and UMAP
-revealer}\label{a-curious-difference-between-t-sne-and-umap-revealer}
+\section{A curious difference between t-SNE, UMAP and PaCMAP
+revealer}\label{a-curious-difference-between-t-sne-umap-and-pacmap-revealer}
 
@@
 INTERESTING EXAMPLE WHEN THE FILLED OUT VS FLAT SHAPES ARE DISCUSSED.
+
+\begin{figure}[H]
+
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/five_gau_clusters/tsne_layout.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/five_gau_clusters/2d_model_tsne.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/five_gau_clusters/sc_tsne_3.png}\end{minipage}%
+
+\caption{\label{fig-gau-tsne-sc}Screen shots of the \textbf{langevitour}
+of the five Gaussian clusters dataset, shows the model with tSNE in
+\pD{}, a video of the tour animation is available at
+(\url{https://youtu.be/RASEE7N5MbM}).}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/five_gau_clusters/umap_layout.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/five_gau_clusters/2d_model_umap.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/five_gau_clusters/sc_umap_3.png}\end{minipage}%
+
+\caption{\label{fig-gau-umap-sc}Screen shots of the \textbf{langevitour}
+of the five Gaussian clusters dataset, shows the model with UMAP in
+\pD{}, a video of the tour animation is available at
+(\url{https://youtu.be/iG4bCPkJilw}).}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/five_gau_clusters/pacmap_layout.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/five_gau_clusters/2d_model_pacmap.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/five_gau_clusters/sc_pacmap_2.png}\end{minipage}%
+
+\caption{\label{fig-gau-pacmap-sc}Screen shots of the
+\textbf{langevitour} of the five Gaussian clusters dataset, shows the
+model with PaCMAP in \pD{}, a video of the tour animation is available
+at (\url{https://youtu.be/z07cKXi8EJQ}).}
+
+\end{figure}%
 

### 2024-06-30 paper.qmd
Update the S-curve true model with data figure
Update figure for prediction

@@
 The model closely fits the shape.
-(The **langevitour** software is used to view the data with a tour, and the full video is available at (<https://youtu.be/bcQg4XcRjhc>).
+(The **langevitour** software is used to view the data with a tour, and the full video is available at (<https://youtu.be/tYATMiSlCXQ>).
 :::
@@
 
-<!--To generate prediction for true-model of S-curve-->
-```{r}
-true_model_df <- read_rds("data/s_curve/true_model.rds")
-true_model_df <- true_model_df |>
-  mutate(ID = row_number())
-
-training_data_scurve <- read_rds("data/s_curve/s_curve_training.rds")
-
-## To initialize effective bins along x
-effective_bin1_scurve <- 12
-
-scurve_model <- fit_highd_model(
-  training_data = training_data_scurve,
-  emb_df = umap_scurve_scaled,
-  bin1 = effective_bin1_scurve,
-  r2 = r2,
-  q = 0.07,
-  is_bin_centroid = TRUE,
-  is_rm_lwd_hex = FALSE,
-  col_start_highd = "x"
-)
-
-df_bin_centroids_scurve <- scurve_model$df_bin_centroids
-df_bin_scurve <- scurve_model$df_bin
-
-true_pred_df <- predict_emb(
-  test_data = true_model_df, 
-  df_bin_centroids = df_bin_centroids_scurve, 
-  df_bin = df_bin_scurve, 
-  type_NLDR = "UMAP"
-  )
-
-true_pred_df |>
-    ggplot(aes(x = pred_UMAP_1,
-               y = pred_UMAP_2)) +
-    geom_point(
-      alpha=0.5,
-      color = "#33a02c"
-      )
-```
+<!--prediction for true-model of S-curve-->
+::: {#fig-scurve-pred-sc layout-ncol="2" fig-pos="H"}
+![](figures/scurve/sc_true_only.png)
+
+![](figures/scurve/pred_true_view.png)
+
+A view of the true model in projections from $7\text{-}D$, and predictions of the true model in \gD{}, for the S-curve data.
+The predictions fits the UMAP layout which means that it capture the geometry of S-curve with UMAP.
+:::
 

### 2024-06-30 paper.tex
Update the S-curve true model with data figure
Update figure for prediction

@@
 data with a tour, and the full video is available at
-(\url{https://youtu.be/bcQg4XcRjhc}).}
+(\url{https://youtu.be/tYATMiSlCXQ}).}
 
@@
 
-\begin{center}
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-15-1.pdf}
-\end{center}
+\begin{figure}[H]
+
+\begin{minipage}{0.50\linewidth}
+\includegraphics{figures/scurve/sc_true_only.png}\end{minipage}%
+%
+\begin{minipage}{0.50\linewidth}
+\includegraphics{figures/scurve/pred_true_view.png}\end{minipage}%
+
+\caption{\label{fig-scurve-pred-sc}A view of the true model in
+projections from \(7\text{-}D\), and predictions of the true model in
+\gD{}, for the S-curve data. The predictions fits the UMAP layout which
+means that it capture the geometry of S-curve with UMAP.}
+
+\end{figure}%
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-32-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-31-1.pdf}
 

### 2024-06-30 paper.qmd
Update the views with true model

@@
 ::: {#fig-scurve-sc layout-ncol="4" fig-pos="H"}
-![](figures/scurve/umap_view_copy.png)
+![](figures/scurve/sc_best_trimesh.png)
 
@@
 The model closely fits the shape, but it has breaks which means that it does not adequately capture the surface.
-(The **langevitour** software is used to view the data with a tour, and the full video is available at (<https://youtu.be/G1m4q9k--v4>).
-NEED TO CHANGE FIRST PLOT TO BE 2D WIREFRAME.
+(The **langevitour** software is used to view the data with a tour, and the full video is available at (<https://youtu.be/N6iWEj3aVps>).
 :::

### 2024-06-30 paper.tex
Update the views with true model

@@
 \begin{minipage}{0.25\linewidth}
-\includegraphics{figures/scurve/umap_view_copy.png}\end{minipage}%
+\includegraphics{figures/scurve/sc_best_trimesh.png}\end{minipage}%
 %
@@
 \textbf{langevitour} software is used to view the data with a tour, and
-the full video is available at (\url{https://youtu.be/G1m4q9k--v4}).
-NEED TO CHANGE FIRST PLOT TO BE 2D WIREFRAME.}
+the full video is available at (\url{https://youtu.be/N6iWEj3aVps}).}
 

### 2024-06-30 paper.qmd
Add descriptions regarding five Gaussina cluster example

@@
 
-XXX THIS IS NOT INTERESTING IN ITS CURRENT FORM. THIS IS ONLY ANY INTERESTING EXAMPLE WHEN THE FILLED OUT VS FLAT SHAPES ARE DISCUSSED.
+<!-- XXX THIS IS NOT INTERESTING IN ITS CURRENT FORM. THIS IS ONLY ANY INTERESTING EXAMPLE WHEN THE FILLED OUT VS FLAT SHAPES ARE DISCUSSED. -->
+
+In this section, the effectiveness of the algorithm is described using a simulated dataset.
+The dataset consists of five spherical Gaussian clusters in $4\text{-}D$, with each cluster containing an equal number of points and the same within-cluster variation.
+
+The \gD{} layouts generated by tSNE, UMAP, and PaCMAP show five well-separated clusters which evident these methods effectively preserve the global structure.
+In tSNE (@fig-gau-tsne-sc (a)), these clusters appear closely.
+UMAP arranges all clusters in a parallel manner, with three aligned in one line and the other two in a separate line (@fig-gau-umap-sc (a)).
+In contrast, PaCMAP shows one central cluster and the remaining four spread out in different directions (@fig-gau-pacmap-sc (a)).
+
+The tSNE and UMAP shows *filled out* clusters which provide evidence that these methods preserve the local structure (@fig-gau-tsne-sc (c) and @fig-gau-umap-sc (c)).
+On the other hand, PaCMAP shows *flat* shapes clusters in the model and evident that PaCMAP fail to capture the within-cluster variation (@fig-gau-pacmap-sc (c)).
 
@@
 
-Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with tSNE in \pD{}, a video of the tour animation is available at (<https://youtu.be/RASEE7N5MbM>).
+The tSNE layout, model in \gD{}, and a view of the fit in projections from $4\text{-}D$, for the five Gaussian cluster data.
+The model fits the separation and tries to *filled out* the clusters.
+(The **langevitour** software is used to view the data with a tour, and the full video is available at (<https://youtu.be/RASEE7N5MbM>).
 :::
@@
 
-Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with UMAP in \pD{}, a video of the tour animation is available at (<https://youtu.be/iG4bCPkJilw>).
+The UMAP layout, model in \gD{}, and a view of the fit in projections from $4\text{-}D$, for the five Gaussian cluster data.
+The model fits the separation and tries to *filled out* the clusters, but not as much as tSNE.
+(The **langevitour** software is used to view the data with a tour, and the full video is available at (<https://youtu.be/iG4bCPkJilw>).
 :::
@@
 
-Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with PaCMAP in \pD{}, a video of the tour animation is available at (<https://youtu.be/z07cKXi8EJQ>).
+The PaCMAP layout, model in \gD{}, and a view of the fit in projections from $4\text{-}D$, for the five Gaussian cluster data.
+The model fits the separation and shows *flat* shaped clusters.
+(The **langevitour** software is used to view the data with a tour, and the full video is available at (<https://youtu.be/z07cKXi8EJQ>).
 :::
-
-In this section, the effectiveness of the algorithm is described using a simulated dataset.
-The dataset consists of five spherical Gaussian clusters in $4\text{-}D$, with each cluster containing an equal number of points and the same within-cluster variation.
-
-In the \gD{} layouts generated by various NLDR techniques, as shown in @fig-NLDR-variety-gau, five well-separated clusters are shown.
-In tSNE (see @fig-NLDR-variety-gau (a)), these clusters appear closely.
-UMAP arranges all clusters in a parallel manner, with three aligned in one line and the other two in a separate line (see @fig-NLDR-variety-gau (b)).
-In contrast, PHATE shows two closely positioned clusters and three more distant ones (see @fig-NLDR-variety-gau (c)).
-In TriMAP, two clusters are close, though not as tightly as PHATE, while the other three are well-separated (see @fig-NLDR-variety-gau (d)).
-Finally, PaCMAP shows one central cluster and the remaining four spread out in different directions (see @fig-NLDR-variety-gau (e)).
-
-<!-- five Gaussian clusters with different NLDR techniques-->
-```{r}
-#| warning: false
-#| echo: false
-
-## Import data
-training_data_gau <- read_rds("data/five_gau_clusters/data_five_gau_training.rds")
-
-tsne_data_gau <- read_rds("data/five_gau_clusters/tsne_data_five_gau_61.rds")
-umap_data_gau <- read_rds("data/five_gau_clusters/umap_data_five_gau.rds")
-phate_data_gau <- read_rds("data/five_gau_clusters/phate_data_five_gau.rds")
-trimap_data_gau <- read_rds("data/five_gau_clusters/trimap_data_five_gau.rds")
-pacmap_data_gau <- read_rds("data/five_gau_clusters/pacmap_data_five_gau.rds")
-
-## Visualize embedding
-
-nldr_gau1 <- tsne_data_gau |>
-  ggplot(aes(x = tSNE1,
-             y = tSNE2)) +
-  geom_point(alpha=0.1, size=1) +
-  interior_annotation("a")
-
-nldr_gau2 <- umap_data_gau |>
-  ggplot(aes(x = UMAP1,
-             y = UMAP2)) +
-  geom_point(alpha=0.1, size=1) +
-  interior_annotation("b")
-
-nldr_gau3 <- phate_data_gau |>
-  ggplot(aes(x = PHATE1,
-             y = PHATE2)) +
-  geom_point(alpha=0.1, size=1) +
-  interior_annotation("c")
-
-nldr_gau4 <- pacmap_data_gau |>
-  ggplot(aes(x = PaCMAP1,
-             y = PaCMAP2)) +
-  geom_point(alpha=0.1, size=1) +
-  interior_annotation("d")
-
-nldr_gau5 <- trimap_data_gau |>
-  ggplot(aes(x = TriMAP1,
-             y = TriMAP2)) +
-  geom_point(alpha=0.1, size=1) +
-  interior_annotation("e")
-```
-
-```{r}
-#| label: fig-NLDR-variety-gau
-#| echo: false
-#| fig-cap: "Five different NLDR representations of the same data.
-Different techniques and different parameter choices are used.
-Is there a best representation of the original data or are they all providing  equivalent information?"
-#| fig-width: 8
-#| fig-height: 2
-#| out-width: 100%
-#| fig-pos: H
-# (a) tSNE (perplexity = 61), (b) UMAP (n_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2)
-nldr_gau1 + nldr_gau2 + 
-  nldr_gau3 + nldr_gau4 + nldr_gau5 + 
-  plot_layout(ncol = 5)
-```
-
-```{r}
-sc_xlims_gau <- c(-0.25, 1.25)
-sc_ylims_gau <- c(-0.2, 1.2)
-```
-
-<!--trimesh with tsne-->
-```{r}
-gau1_scaled_obj <- gen_scaled_data(
-  data = tsne_data_gau)
-tsne_gau_scaled <- gau1_scaled_obj$scaled_nldr
-
-## Compute hexbin parameters
-num_bins_x_gau1 <- 12
-lim1 <- gau1_scaled_obj$lim1
-lim2 <- gau1_scaled_obj$lim2
-r2_gau1 <- diff(lim2)/diff(lim1)
-
-gau1_model <- fit_highd_model(
-  training_data = training_data_gau,
-  emb_df = tsne_gau_scaled,
-  bin1 = num_bins_x_gau1,
-  r2 = r2_gau1,
-  is_bin_centroid = TRUE,
-  is_rm_lwd_hex = FALSE,
-  col_start_highd = "x"
-)
-
-df_bin_centroids_gau1 <- gau1_model$df_bin_centroids
-df_bin_gau1 <- gau1_model$df_bin
-
-## Triangulate bin centroids
-tr1_object_gau1 <- tri_bin_centroids(
-  df_bin_centroids_gau1, x = "c_x", y = "c_y")
-tr_from_to_df_gau1 <- gen_edges(
-  tri_object = tr1_object_gau1)
-
-## Compute 2D distances
-distance_gau1 <- cal_2d_dist(
-  tr_coord_df = tr_from_to_df_gau1,
-  start_x = "x_from",
-  start_y = "y_from",
-  end_x = "x_to",
-  end_y = "y_to",
-  select_vars = c("from", "to", "distance"))
-
-## To find the benchmark value
-benchmark_gau1 <- find_lg_benchmark(
-  distance_edges = distance_gau1,
-  distance_col = "distance")
-
-trimesh_removed_gau1 <- vis_rmlg_mesh(
-  distance_edges = distance_gau1, 
-  benchmark_value = benchmark_gau1, 
-  tr_coord_df = tr_from_to_df_gau1, 
-  distance_col = "distance") +
-  xlim(sc_xlims_gau) + ylim(sc_ylims_gau) +
-  interior_annotation("a", sc_ltr_pos) 
-```
-
-<!--trimesh with umap-->
-```{r}
-gau2_scaled_obj <- gen_scaled_data(
-  data = umap_data_gau)
-umap_gau_scaled <- gau2_scaled_obj$scaled_nldr
-
-## Compute hexbin parameters
-num_bins_x_gau2 <- 44
-lim1 <- gau2_scaled_obj$lim1
-lim2 <- gau2_scaled_obj$lim2
-r2_gau2 <- diff(lim2)/diff(lim1)
-
-gau2_model <- fit_highd_model(
-  training_data = training_data_gau,
-  emb_df = umap_gau_scaled,
-  bin1 = num_bins_x_gau2,
-  r2 = r2_gau2,
-  is_bin_centroid = TRUE,
-  is_rm_lwd_hex = FALSE,
-  col_start_highd = "x"
-)
-
-df_bin_centroids_gau2 <- gau2_model$df_bin_centroids
-df_bin_gau2 <- gau2_model$df_bin
-
-## Triangulate bin centroids
-tr1_object_gau2 <- tri_bin_centroids(
-  df_bin_centroids_gau2, x = "c_x", y = "c_y")
-tr_from_to_df_gau2 <- gen_edges(
-  tri_object = tr1_object_gau2)
-
-## Compute 2D distances
-distance_gau2 <- cal_2d_dist(
-  tr_coord_df = tr_from_to_df_gau2,
-  start_x = "x_from",
-  start_y = "y_from",
-  end_x = "x_to",
-  end_y = "y_to",
-  select_vars = c("from", "to", "distance"))
-
-## To find the benchmark value
-benchmark_gau2 <- find_lg_benchmark(
-  distance_edges = distance_gau2,
-  distance_col = "distance")
-
-trimesh_removed_gau2 <- vis_rmlg_mesh(
-  distance_edges = distance_gau2, 
-  benchmark_value = benchmark_gau2, 
-  tr_coord_df = tr_from_to_df_gau2, 
-  distance_col = "distance") +
-  xlim(sc_xlims_gau) + ylim(sc_ylims_gau) +
-  interior_annotation("b", sc_ltr_pos) 
-```
-
-<!--trimesh with phate-->
-```{r}
-# gau3_scaled_obj <- gen_scaled_data(
-#   data = phate_data_gau)
-# phate_gau_scaled <- gau3_scaled_obj$scaled_nldr
-# 
-# ## Compute hexbin parameters
-# num_bins_x_gau3 <- 250
-# lim1 <- gau3_scaled_obj$lim1
-# lim2 <- gau3_scaled_obj$lim2
-# r2_gau3 <- diff(lim2)/diff(lim1)
-# 
-# gau3_model <- fit_highd_model(
-#   training_data = training_data_gau,
-#   emb_df = phate_gau_scaled,
-#   bin1 = num_bins_x_gau3,
-#   r2 = r2_gau3,
-#   is_bin_centroid = TRUE,
-#   is_rm_lwd_hex = FALSE,
-#   col_start_highd = "x"
-# )
-# 
-# df_bin_centroids_gau3 <- gau3_model$df_bin_centroids
-# df_bin_gau3 <- gau3_model$df_bin
-
-df_bin_centroids_gau3 <- read_rds("data/five_gau_clusters/df_bin_centroids_gau_phate.rds")
-df_bin_gau3 <- read_rds("data/five_gau_clusters/df_bin_gau_phate.rds")
-
-## Triangulate bin centroids
-tr1_object_gau3 <- tri_bin_centroids(
-  df_bin_centroids_gau3, x = "c_x", y = "c_y")
-tr_from_to_df_gau3 <- gen_edges(
-  tri_object = tr1_object_gau3)
-
-## Compute 2D distances
-distance_gau3 <- cal_2d_dist(
-  tr_coord_df = tr_from_to_df_gau3,
-  start_x = "x_from",
-  start_y = "y_from",
-  end_x = "x_to",
-  end_y = "y_to",
-  select_vars = c("from", "to", "distance"))
-
-## To find the benchmark value
-benchmark_gau3 <- find_lg_benchmark(
-  distance_edges = distance_gau3,
-  distance_col = "distance")
-
-trimesh_removed_gau3 <- vis_rmlg_mesh(
-  distance_edges = distance_gau3, 
-  benchmark_value = benchmark_gau3, 
-  tr_coord_df = tr_from_to_df_gau3, 
-  distance_col = "distance") +
-  xlim(sc_xlims_gau) + ylim(sc_ylims_gau) +
-  interior_annotation("c", sc_ltr_pos) 
-```
-
-<!--trimesh with pacmap-->
-```{r}
-gau4_scaled_obj <- gen_scaled_data(
-  data = pacmap_data_gau)
-pacmap_gau_scaled <- gau4_scaled_obj$scaled_nldr
-
-## Compute hexbin parameters
-num_bins_x_gau4 <- 18
-lim1 <- gau4_scaled_obj$lim1
-lim2 <- gau4_scaled_obj$lim2
-r2_gau4 <- diff(lim2)/diff(lim1)
-
-gau4_model <- fit_highd_model(
-  training_data = training_data_gau,
-  emb_df = pacmap_gau_scaled,
-  bin1 = num_bins_x_gau4,
-  r2 = r2_gau4,
-  is_bin_centroid = TRUE,
-  is_rm_lwd_hex = FALSE,
-  col_start_highd = "x"
-)
-
-df_bin_centroids_gau4 <- gau4_model$df_bin_centroids
-df_bin_gau4 <- gau4_model$df_bin
-
-## Triangulate bin centroids
-tr1_object_gau4 <- tri_bin_centroids(
-  df_bin_centroids_gau4, x = "c_x", y = "c_y")
-tr_from_to_df_gau4 <- gen_edges(
-  tri_object = tr1_object_gau4)
-
-## Compute 2D distances
-distance_gau4 <- cal_2d_dist(
-  tr_coord_df = tr_from_to_df_gau4,
-  start_x = "x_from",
-  start_y = "y_from",
-  end_x = "x_to",
-  end_y = "y_to",
-  select_vars = c("from", "to", "distance"))
-
-## To find the benchmark value
-benchmark_gau4 <- find_lg_benchmark(
-  distance_edges = distance_gau4,
-  distance_col = "distance")
-
-trimesh_removed_gau4 <- vis_rmlg_mesh(
-  distance_edges = distance_gau4, 
-  benchmark_value = benchmark_gau4, 
-  tr_coord_df = tr_from_to_df_gau4, 
-  distance_col = "distance") +
-  xlim(sc_xlims_gau) + ylim(sc_ylims_gau) +
-  interior_annotation("d", sc_ltr_pos) 
-```
-
-<!--trimesh with trimap-->
-```{r}
-gau5_scaled_obj <- gen_scaled_data(
-  data = trimap_data_gau)
-trimap_gau_scaled <- gau5_scaled_obj$scaled_nldr
-
-## Compute hexbin parameters
-num_bins_x_gau5 <- 25
-lim1 <- gau5_scaled_obj$lim1
-lim2 <- gau5_scaled_obj$lim2
-r2_gau5 <- diff(lim2)/diff(lim1)
-
-gau5_model <- fit_highd_model(
-  training_data = training_data_gau,
-  emb_df = trimap_gau_scaled,
-  bin1 = num_bins_x_gau5,
-  r2 = r2_gau5,
-  is_bin_centroid = TRUE,
-  is_rm_lwd_hex = FALSE,
-  col_start_highd = "x"
-)
-
-df_bin_centroids_gau5 <- gau5_model$df_bin_centroids
-df_bin_gau5 <- gau5_model$df_bin
-
-## Triangulate bin centroids
-tr1_object_gau5 <- tri_bin_centroids(
-  df_bin_centroids_gau5, x = "c_x", y = "c_y")
-tr_from_to_df_gau5 <- gen_edges(
-  tri_object = tr1_object_gau5)
-
-## Compute 2D distances
-distance_gau5 <- cal_2d_dist(
-  tr_coord_df = tr_from_to_df_gau5,
-  start_x = "x_from",
-  start_y = "y_from",
-  end_x = "x_to",
-  end_y = "y_to",
-  select_vars = c("from", "to", "distance"))
-
-## To find the benchmark value
-benchmark_gau5 <- find_lg_benchmark(
-  distance_edges = distance_gau5,
-  distance_col = "distance")
-benchmark_gau5 <- 0.2
-
-trimesh_removed_gau5 <- vis_rmlg_mesh(
-  distance_edges = distance_gau5, 
-  benchmark_value = benchmark_gau5, 
-  tr_coord_df = tr_from_to_df_gau5, 
-  distance_col = "distance") +
-  xlim(sc_xlims_gau) + ylim(sc_ylims_gau) +
-  interior_annotation("e", sc_ltr_pos) 
-```
-
-<!--add trimesh for all five-->
-
-```{r}
-#| label: fig-trimesh-gau
-#| echo: false
-#| fig-cap: "Model generated with five different NLDR methods in $2\\text{-}D$ with approximately $65$ non-empty bins in each."
-#| fig-width: 8
-#| fig-height: 2
-#| out-width: 100%
-#| fig-pos: H
-
-trimesh_removed_gau1 + trimesh_removed_gau2 +
-  trimesh_removed_gau3 + trimesh_removed_gau4 + trimesh_removed_gau5 +
-  plot_layout(ncol = 5)
-```
-
-<!--add three screenshots from each technique with you tube links-->
-
-To investigate which is the reasonable representation to visualize the five spherical Gaussian cluster data or all NLDR methods provide equivalent information, we visualize all the models in \pD{} space.
-Models from all NLDR methods show five well-separated clusters (see @fig-gau1-sc, @fig-gau2-sc, @fig-gau3-sc, @fig-gau4-sc, and @fig-gau5-sc).
-This suggests that for the five Gaussian cluster dataset, all NLDR methods effectively preserve the global structure.
-tSNE displays clusters with varying densities, indicating their ability to capture within-cluster variation (see @fig-gau1-sc).
-On the other hand, both UMAP, PHATE, PaCMAP and TriMAP show clusters with flat surfaces, suggesting a failure to capture within-cluster variation (see @fig-gau2-sc, @fig-gau3-sc, @fig-gau4-sc and @fig-gau5-sc).
-Therefore, UMAP, PHATE, PaCMAP and TriMAP do not capture the local structure as effectively as other methods.
-
-::: {#fig-gau1-sc layout-ncol="3" fig-pos="H"}
-![](figures/five_gau_clusters/sc_tsne_1.png)
-
-![](figures/five_gau_clusters/sc_tsne_2.png)
-
-![](figures/five_gau_clusters/sc_tsne_3.png)
-
-Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with tSNE in \pD{}, a video of the tour animation is available at (<https://youtu.be/RASEE7N5MbM>).
-:::
-
-::: {#fig-gau2-sc layout-ncol="3" fig-pos="H"}
-![](figures/five_gau_clusters/sc_umap_1.png)
-
-![](figures/five_gau_clusters/sc_umap_2.png)
-
-![](figures/five_gau_clusters/sc_umap_3.png)
-
-Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with UMAP in \pD{}, a video of the tour animation is available at (<https://youtu.be/iG4bCPkJilw>).
-:::
-
-::: {#fig-gau3-sc layout-ncol="3" fig-pos="H"}
-![](figures/five_gau_clusters/sc_phate_1.png)
-
-![](figures/five_gau_clusters/sc_phate_2.png)
-
-![](figures/five_gau_clusters/sc_phate_3.png)
-
-Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with PHATE in \pD{}, a video of the tour animation is available at (<https://youtu.be/L_PVLGwfOS0>).
-:::
-
-::: {#fig-gau4-sc layout-ncol="3" fig-pos="H"}
-![](figures/five_gau_clusters/sc_pacmap_1.png)
-
-![](figures/five_gau_clusters/sc_pacmap_2.png)
-
-![](figures/five_gau_clusters/sc_pacmap_3.png)
-
-Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with PaCMAP in \pD{}, a video of the tour animation is available at (<https://youtu.be/z07cKXi8EJQ>).
-:::
-
-::: {#fig-gau5-sc layout-ncol="3" fig-pos="H"}
-![](figures/five_gau_clusters/sc_trimap_1.png)
-
-![](figures/five_gau_clusters/sc_trimap_2.png)
-
-![](figures/five_gau_clusters/sc_trimap_3.png)
-
-Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with TriMAP in \pD{}, a video of the tour animation is available at (<https://youtu.be/Chs1lYAoX2w>).
-:::
-
-When compare the NLDR representations and generated models, tSNE with perplexity $61$ appears to be a reasonable representation for visualizing the five Gaussian cluster dataset.
-This is supported by investigating the model generated with tSNE in the data space, which provides evidence that it preserves both local and global structures.
-Also, the NLDR representation with tSNE shows five well-separated clusters.
 

### 2024-06-30 paper.tex
Add descriptions regarding five Gaussina cluster example

@@
 
-XXX THIS IS NOT INTERESTING IN ITS CURRENT FORM. THIS IS ONLY ANY
-INTERESTING EXAMPLE WHEN THE FILLED OUT VS FLAT SHAPES ARE DISCUSSED.
-
-\begin{figure}[H]
-
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/five_gau_clusters/tsne_layout.png}\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/five_gau_clusters/2d_model_tsne.png}\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/five_gau_clusters/sc_tsne_3.png}\end{minipage}%
-
-\caption{\label{fig-gau-tsne-sc}Screen shots of the \textbf{langevitour}
-of the five Gaussian clusters dataset, shows the model with tSNE in
-\pD{}, a video of the tour animation is available at
-(\url{https://youtu.be/RASEE7N5MbM}).}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/five_gau_clusters/umap_layout.png}\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/five_gau_clusters/2d_model_umap.png}\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/five_gau_clusters/sc_umap_3.png}\end{minipage}%
-
-\caption{\label{fig-gau-umap-sc}Screen shots of the \textbf{langevitour}
-of the five Gaussian clusters dataset, shows the model with UMAP in
-\pD{}, a video of the tour animation is available at
-(\url{https://youtu.be/iG4bCPkJilw}).}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/five_gau_clusters/pacmap_layout.png}\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/five_gau_clusters/2d_model_pacmap.png}\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/five_gau_clusters/sc_pacmap_2.png}\end{minipage}%
-
-\caption{\label{fig-gau-pacmap-sc}Screen shots of the
-\textbf{langevitour} of the five Gaussian clusters dataset, shows the
-model with PaCMAP in \pD{}, a video of the tour animation is available
-at (\url{https://youtu.be/z07cKXi8EJQ}).}
-
-\end{figure}%
-
 In this section, the effectiveness of the algorithm is described using a
@@
 
-In the \gD{} layouts generated by various NLDR techniques, as shown in
-Figure~\ref{fig-NLDR-variety-gau}, five well-separated clusters are
-shown. In tSNE (see Figure~\ref{fig-NLDR-variety-gau} (a)), these
+The \gD{} layouts generated by tSNE, UMAP, and PaCMAP show five
+well-separated clusters which evident these methods effectively preserve
+the global structure. In tSNE (Figure~\ref{fig-gau-tsne-sc} (a)), these
 clusters appear closely. UMAP arranges all clusters in a parallel
 manner, with three aligned in one line and the other two in a separate
-line (see Figure~\ref{fig-NLDR-variety-gau} (b)). In contrast, PHATE
-shows two closely positioned clusters and three more distant ones (see
-Figure~\ref{fig-NLDR-variety-gau} (c)). In TriMAP, two clusters are
-close, though not as tightly as PHATE, while the other three are
-well-separated (see Figure~\ref{fig-NLDR-variety-gau} (d)). Finally,
-PaCMAP shows one central cluster and the remaining four spread out in
-different directions (see Figure~\ref{fig-NLDR-variety-gau} (e)).
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-NLDR-variety-gau-1.pdf}
-
-}
-
-\caption{\label{fig-NLDR-variety-gau}Five different NLDR representations
-of the same data. Different techniques and different parameter choices
-are used. Is there a best representation of the original data or are
-they all providing equivalent information?}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-trimesh-gau-1.pdf}
-
-}
-
-\caption{\label{fig-trimesh-gau}Model generated with five different NLDR
-methods in \(2\text{-}D\) with approximately \(65\) non-empty bins in
-each.}
-
-\end{figure}%
-
-To investigate which is the reasonable representation to visualize the
-five spherical Gaussian cluster data or all NLDR methods provide
-equivalent information, we visualize all the models in \pD{} space.
-Models from all NLDR methods show five well-separated clusters (see
-Figure~\ref{fig-gau1-sc}, Figure~\ref{fig-gau2-sc},
-Figure~\ref{fig-gau3-sc}, Figure~\ref{fig-gau4-sc}, and
-Figure~\ref{fig-gau5-sc}). This suggests that for the five Gaussian
-cluster dataset, all NLDR methods effectively preserve the global
-structure. tSNE displays clusters with varying densities, indicating
-their ability to capture within-cluster variation (see
-Figure~\ref{fig-gau1-sc}). On the other hand, both UMAP, PHATE, PaCMAP
-and TriMAP show clusters with flat surfaces, suggesting a failure to
-capture within-cluster variation (see Figure~\ref{fig-gau2-sc},
-Figure~\ref{fig-gau3-sc}, Figure~\ref{fig-gau4-sc} and
-Figure~\ref{fig-gau5-sc}). Therefore, UMAP, PHATE, PaCMAP and TriMAP do
-not capture the local structure as effectively as other methods.
-
-\begin{figure}[H]
-
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/five_gau_clusters/sc_tsne_1.png}\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/five_gau_clusters/sc_tsne_2.png}\end{minipage}%
+line (Figure~\ref{fig-gau-umap-sc} (a)). In contrast, PaCMAP shows one
+central cluster and the remaining four spread out in different
+directions (Figure~\ref{fig-gau-pacmap-sc} (a)).
+
+The tSNE and UMAP shows \emph{filled out} clusters which provide
+evidence that these methods preserve the local structure
+(Figure~\ref{fig-gau-tsne-sc} (c) and Figure~\ref{fig-gau-umap-sc} (c)).
+On the other hand, PaCMAP shows \emph{flat} shapes clusters in the model
+and evident that PaCMAP fail to capture the within-cluster variation
+(Figure~\ref{fig-gau-pacmap-sc} (c)).
+
+\begin{figure}[H]
+
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/five_gau_clusters/tsne_layout.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/five_gau_clusters/2d_model_tsne.png}\end{minipage}%
 %
@@
 
-\caption{\label{fig-gau1-sc}Screen shots of the \textbf{langevitour} of
-the five Gaussian clusters dataset, shows the model with tSNE in \pD{},
-a video of the tour animation is available at
+\caption{\label{fig-gau-tsne-sc}The tSNE layout, model in \gD{}, and a
+view of the fit in projections from \(4\text{-}D\), for the five
+Gaussian cluster data. The model fits the separation and tries to
+\emph{filled out} the clusters. (The \textbf{langevitour} software is
+used to view the data with a tour, and the full video is available at
 (\url{https://youtu.be/RASEE7N5MbM}).}
@@
 \begin{minipage}{0.33\linewidth}
-\includegraphics{figures/five_gau_clusters/sc_umap_1.png}\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/five_gau_clusters/sc_umap_2.png}\end{minipage}%
+\includegraphics{figures/five_gau_clusters/umap_layout.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/five_gau_clusters/2d_model_umap.png}\end{minipage}%
 %
@@
 
-\caption{\label{fig-gau2-sc}Screen shots of the \textbf{langevitour} of
-the five Gaussian clusters dataset, shows the model with UMAP in \pD{},
-a video of the tour animation is available at
-(\url{https://youtu.be/iG4bCPkJilw}).}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/five_gau_clusters/sc_phate_1.png}\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/five_gau_clusters/sc_phate_2.png}\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/five_gau_clusters/sc_phate_3.png}\end{minipage}%
-
-\caption{\label{fig-gau3-sc}Screen shots of the \textbf{langevitour} of
-the five Gaussian clusters dataset, shows the model with PHATE in \pD{},
-a video of the tour animation is available at
-(\url{https://youtu.be/L_PVLGwfOS0}).}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/five_gau_clusters/sc_pacmap_1.png}\end{minipage}%
+\caption{\label{fig-gau-umap-sc}The UMAP layout, model in \gD{}, and a
+view of the fit in projections from \(4\text{-}D\), for the five
+Gaussian cluster data. The model fits the separation and tries to
+\emph{filled out} the clusters, but not as much as tSNE. (The
+\textbf{langevitour} software is used to view the data with a tour, and
+the full video is available at (\url{https://youtu.be/iG4bCPkJilw}).}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/five_gau_clusters/pacmap_layout.png}\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\includegraphics{figures/five_gau_clusters/2d_model_pacmap.png}\end{minipage}%
 %
@@
 \includegraphics{figures/five_gau_clusters/sc_pacmap_2.png}\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/five_gau_clusters/sc_pacmap_3.png}\end{minipage}%
-
-\caption{\label{fig-gau4-sc}Screen shots of the \textbf{langevitour} of
-the five Gaussian clusters dataset, shows the model with PaCMAP in
-\pD{}, a video of the tour animation is available at
+
+\caption{\label{fig-gau-pacmap-sc}The PaCMAP layout, model in \gD{}, and
+a view of the fit in projections from \(4\text{-}D\), for the five
+Gaussian cluster data. The model fits the separation and shows
+\emph{flat} shaped clusters. (The \textbf{langevitour} software is used
+to view the data with a tour, and the full video is available at
 (\url{https://youtu.be/z07cKXi8EJQ}).}
@@
 \end{figure}%
-
-\begin{figure}[H]
-
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/five_gau_clusters/sc_trimap_1.png}\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/five_gau_clusters/sc_trimap_2.png}\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/five_gau_clusters/sc_trimap_3.png}\end{minipage}%
-
-\caption{\label{fig-gau5-sc}Screen shots of the \textbf{langevitour} of
-the five Gaussian clusters dataset, shows the model with TriMAP in
-\pD{}, a video of the tour animation is available at
-(\url{https://youtu.be/Chs1lYAoX2w}).}
-
-\end{figure}%
-
-When compare the NLDR representations and generated models, tSNE with
-perplexity \(61\) appears to be a reasonable representation for
-visualizing the five Gaussian cluster dataset. This is supported by
-investigating the model generated with tSNE in the data space, which
-provides evidence that it preserves both local and global structures.
-Also, the NLDR representation with tSNE shows five well-separated
-clusters.
 

### 2024-07-01 paper.qmd
Added best fit projections

@@
 
-The appropriate model for the S-curve using UMAP (n_neighbors: $15$) was created with $420$ bins.
+@fig-scurve-sc-best shows one of the best fits for S-curve data.
+TriMAP with $15$ nearest neighbors for forming the nearest neighbor triplet (`n_inliers`), $3$ outliers for forming the nearest neighbor triplets (`n_outliers`), $3$ random triplets per point (`n_random`).
+
+<!--langevitour with pD model for S-curve-->
+::: {#fig-scurve-sc-best layout-ncol="4" fig-pos="H"}
+![](figures/scurve/sc_best_trimap_trimesh.png)
+
+![](figures/scurve/sc_trimap_best_1.png)
+
+![](figures/scurve/sc_trimap_best_2.png)
+
+![](figures/scurve/sc_trimap_best_3.png)
+
+Model in \gD{}, on the TriMAP layout, and three views of the fit in projections from $7\text{-}D$, for the S-curve data.
+The model closely fits the shape, and it adequately captures the surface.
+(The **langevitour** software is used to view the data with a tour, and the full video is available at (<https://youtu.be/zEz2MUCtADM>).
+:::
+
+<!-- The appropriate model for the S-curve using UMAP (n_neighbors: $15$) was created with $420$ bins.
 It was determined to be the best model after removing edges with a length greater than $0.134$.
@@
 But better in some places only (@fig-scurve_sc_best).
-
-In contrast, a poor model for the S-curve with PHATE was created using only $190$ bins and removing edges with a length greater than $0.3$.
+-->
+
+<!-- In contrast, a poor model for the S-curve with PHATE was created using only $190$ bins and removing edges with a length greater than $0.3$.
 This model is inadequate as it has a higher MSE compared to the appropriate model.
 Visually, the model squeezed to the middle of the geometry of the S-curve (@fig-scurve-sc-bad).
-
-XXX BEST FIT FOR S-CURVE MIGHT BE A DIFFERENT NLDR. 
-SHOW THE SEVERAL SUB-OPTIMAL FITS.
+-->
+
+<!-- XXX BEST FIT FOR S-CURVE MIGHT BE A DIFFERENT NLDR.  -->
+<!-- SHOW THE SEVERAL SUB-OPTIMAL FITS. -->
 

### 2024-07-01 paper.tex
Added best fit projections

@@
 
-The appropriate model for the S-curve using UMAP (n\_neighbors: \(15\))
-was created with \(420\) bins. It was determined to be the best model
-after removing edges with a length greater than \(0.134\). The decision
-was made considering that the model accurately captures the geometry of
-the S-curve. But better in some places only
-(\textbf{?@fig-scurve\_sc\_best}).
-
-In contrast, a poor model for the S-curve with PHATE was created using
-only \(190\) bins and removing edges with a length greater than \(0.3\).
-This model is inadequate as it has a higher MSE compared to the
-appropriate model. Visually, the model squeezed to the middle of the
-geometry of the S-curve (\textbf{?@fig-scurve-sc-bad}).
-
-XXX BEST FIT FOR S-CURVE MIGHT BE A DIFFERENT NLDR. SHOW THE SEVERAL
-SUB-OPTIMAL FITS.
+Figure~\ref{fig-scurve-sc-best} shows one of the best fits for S-curve
+data. TriMAP with \(15\) nearest neighbors for forming the nearest
+neighbor triplet (\texttt{n\_inliers}), \(3\) outliers for forming the
+nearest neighbor triplets (\texttt{n\_outliers}), \(3\) random triplets
+per point (\texttt{n\_random}).
+
+\begin{figure}[H]
+
+\begin{minipage}{0.25\linewidth}
+\includegraphics{figures/scurve/sc_best_trimap_trimesh.png}\end{minipage}%
+%
+\begin{minipage}{0.25\linewidth}
+\includegraphics{figures/scurve/sc_trimap_best_1.png}\end{minipage}%
+%
+\begin{minipage}{0.25\linewidth}
+\includegraphics{figures/scurve/sc_trimap_best_2.png}\end{minipage}%
+%
+\begin{minipage}{0.25\linewidth}
+\includegraphics{figures/scurve/sc_trimap_best_3.png}\end{minipage}%
+
+\caption{\label{fig-scurve-sc-best}Model in \gD{}, on the TriMAP layout,
+and three views of the fit in projections from \(7\text{-}D\), for the
+S-curve data. The model closely fits the shape, and it adequately
+captures the surface. (The \textbf{langevitour} software is used to view
+the data with a tour, and the full video is available at
+(\url{https://youtu.be/zEz2MUCtADM}).}
+
+\end{figure}%
 

### 2024-07-01 paper.qmd
Added descriptions reagrding the best fit

@@
 @fig-scurve-sc-best shows one of the best fits for S-curve data.
-TriMAP with $15$ nearest neighbors for forming the nearest neighbor triplet (`n_inliers`), $3$ outliers for forming the nearest neighbor triplets (`n_outliers`), $3$ random triplets per point (`n_random`).
+TriMAP with $15$ nearest neighbors for forming the nearest neighbor triplet (`n_inliers`), $3$ outliers for forming the nearest neighbor triplets (`n_outliers`), $3$ random triplets per point (`n_random`) is used as the NLDR method.
+The model is generated with $103$ bins, which have a width of $0.154$ and a height of $0.133$.
+Edges with lengths greater than $0.266$ are removed.
 
@@
 Model in \gD{}, on the TriMAP layout, and three views of the fit in projections from $7\text{-}D$, for the S-curve data.
-The model closely fits the shape, and it adequately captures the surface.
+The model closely fits the shape, but it has twists.
 (The **langevitour** software is used to view the data with a tour, and the full video is available at (<https://youtu.be/zEz2MUCtADM>).

### 2024-07-01 paper.tex
Added descriptions reagrding the best fit

@@
 nearest neighbor triplets (\texttt{n\_outliers}), \(3\) random triplets
-per point (\texttt{n\_random}).
+per point (\texttt{n\_random}) is used as the NLDR method. The model is
+generated with \(103\) bins, which have a width of \(0.154\) and a
+height of \(0.133\). Edges with lengths greater than \(0.266\) are
+removed.
 
@@
 and three views of the fit in projections from \(7\text{-}D\), for the
-S-curve data. The model closely fits the shape, and it adequately
-captures the surface. (The \textbf{langevitour} software is used to view
-the data with a tour, and the full video is available at
-(\url{https://youtu.be/zEz2MUCtADM}).}
+S-curve data. The model closely fits the shape, but it has twists. (The
+\textbf{langevitour} software is used to view the data with a tour, and
+the full video is available at (\url{https://youtu.be/zEz2MUCtADM}).}
 

### 2024-07-04 paper.qmd
Edit the figure 3

@@
 sc_ltr_pos <- c(0.08, 0.96)
-sc_xlims <- c(-0.25, 1.25)
-sc_ylims <- c(-0.2, 1.9)
+sc_xlims <- c(-0.5, 1.47)
+sc_ylims <- c(-0.32, 2.1)
 # sc_xlims <- c(-0.25, 1.25)
@@
 ## Compute hexbin parameters
-num_bins_x_scurve <- 6
+num_bins_x_scurve <- 7
 

### 2024-07-04 paper.qmd
Update Figure 4

@@
                  x2max = r2)
-c_start <- tibble(s1 = 0.03, s2 = 0.05)
-a1 <- tibble(x = all_centroids_df$c_x[21],
-             xend = all_centroids_df$c_x[22],
+#c_start <- tibble(s1 = 0.03, s2 = 0.05)
+c_start <- tibble(s1 = 0.03, s2 = -0.05)
+
+a1 <- tibble(x = all_centroids_df$c_x[4],
+             xend = all_centroids_df$c_x[5],
              y = all_centroids_df$c_y[21],
@@
 #| fig-width: 2.5
-#| fig-height: 3.75
+#| fig-height: 3.3
 #| fig-pos: H

### 2024-07-04 paper.qmd
Resize Figure 6

@@
 ::: {#fig-scurve-pred-sc layout-ncol="2" fig-pos="H"}
-![](figures/scurve/sc_true_only.png)
-
-![](figures/scurve/pred_true_view.png)
+![](figures/scurve/sc_true_only.png){width="150" fig-align="center"}
+
+![](figures/scurve/pred_true_view.png){width="150" fig-align="center"}
 

### 2024-07-04 paper.tex
Resize Figure 6

@@
 \begin{minipage}{0.50\linewidth}
-\includegraphics{figures/scurve/sc_true_only.png}\end{minipage}%
+\begin{center}
+\includegraphics[width=1.5625in,height=\textheight]{figures/scurve/sc_true_only.png}
+\end{center}
+\end{minipage}%
 %
 \begin{minipage}{0.50\linewidth}
-\includegraphics{figures/scurve/pred_true_view.png}\end{minipage}%
+\begin{center}
+\includegraphics[width=1.5625in,height=\textheight]{figures/scurve/pred_true_view.png}
+\end{center}
+\end{minipage}%
 

### 2024-07-04 paper.qmd
Update Figure 7

@@
 
-![](figures/scurve/sc_true_1.png)
-
-![](figures/scurve/sc_true_2.png)
-
-![](figures/scurve/sc_true_3.png)
+![](figures/scurve/sc_true_1.png){width="150" fig-align="center"}
+
+![](figures/scurve/sc_true_2.png){width="150" fig-align="center"}
+
+![](figures/scurve/sc_true_3.png){width="150" fig-align="center"}
 
@@
   bin1 = 7, 
-  r2 = r2,
-  q = 0.05)
+  r2 = r2)
 
@@
 #| fig-pos: H
-#| fig-cap: "Hexbin density plots of UMAP layout of the S-curve data, using three different bin inputs: (a) $b = 91$ (7, 13), (b) $b = 190$ (10, 19), and (c) $b = 364$ (14, 26).
+#| fig-cap: "Hexbin density plots of UMAP layout of the S-curve data, using three different bin inputs: (a) $b = 70$ (7, 10), (b) $b = 240$ (12, 20), and (c) $b = 322$ (14, 23).
 Color indicates standardized counts, dark indicating high count and light indicates low count.
@@
 #| fig-width: 6
-#| fig-height: 3
+#| fig-height: 2.5
 

### 2024-07-04 paper.tex
Update Figure 7

@@
 \begin{minipage}{0.33\linewidth}
-\includegraphics{figures/scurve/sc_true_1.png}\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/scurve/sc_true_2.png}\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/scurve/sc_true_3.png}\end{minipage}%
+\begin{center}
+\includegraphics[width=1.5625in,height=\textheight]{figures/scurve/sc_true_1.png}
+\end{center}
+\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\begin{center}
+\includegraphics[width=1.5625in,height=\textheight]{figures/scurve/sc_true_2.png}
+\end{center}
+\end{minipage}%
+%
+\begin{minipage}{0.33\linewidth}
+\begin{center}
+\includegraphics[width=1.5625in,height=\textheight]{figures/scurve/sc_true_3.png}
+\end{center}
+\end{minipage}%
 
@@
 \caption{\label{fig-bins-scurve}Hexbin density plots of UMAP layout of
-the S-curve data, using three different bin inputs: (a) \(b = 91\) (7,
-13), (b) \(b = 190\) (10, 19), and (c) \(b = 364\) (14, 26). Color
+the S-curve data, using three different bin inputs: (a) \(b = 70\) (7,
+10), (b) \(b = 240\) (12, 20), and (c) \(b = 322\) (14, 23). Color
 indicates standardized counts, dark indicating high count and light

### 2024-07-04 paper.qmd
Update Figure 8

@@
   geom_line() + 
-  geom_vline(xintercept = 190, linetype="solid", 
+  geom_vline(xintercept = 240, linetype="solid", 
              color = "black", linewidth=0.8, alpha = 0.5) +
-  geom_vline(xintercept = 91, linetype=2, 
+  geom_vline(xintercept = 96, linetype=2, 
              color = "black", linewidth=0.8, alpha = 0.5) +
-  geom_vline(xintercept = 364, linetype=2, 
+  geom_vline(xintercept = 375, linetype=2, 
              color = "black", linewidth=0.8, alpha = 0.5) +

### 2024-07-04 paper.qmd
Update Figure 3

@@
              aes(x = UMAP1, y = UMAP2), 
-             alpha = 0.5) +
+             alpha = 0.5, size = 0.5) +
   xlim(sc_xlims) + ylim(sc_ylims) +

### 2024-07-07 paper.qmd
Changed NLDR MNIST PaCMAP image

@@
 library(patchwork)
+library(png)
 library(langevitour)
@@
 
-position_df <- pacmap_minst_scaled |> 
-  filter(ID %in% c(6475, 3311, 6489, 2347, 387, 
-                          5239, 728, 517, 6369, 3055, 6345, 1761)) |>
-  mutate(position = c(5, 9, 7, 8, 4, 11, 2, 6, 12, 10, 1, 3))
-
 pacmap_plot_mnist <- pacmap_minst_scaled |> 
@@
              y = PaCMAP2)) + 
-  geom_point(alpha=0.1) #+ 
-  # geom_text(data = position_df, 
-  #           aes(x = PaCMAP1, y = PaCMAP2, label = position), 
-  #           colour = "red", 
-  #           size = 5)  
+  geom_point(alpha=0.1)
+
+# read the png file from device 
+left_top_img <- readPNG("figures/mnist/mnist_digit1_img_lt.png", native = TRUE)
+center_img <- readPNG("figures/mnist/mnist_digit1_img_c.png", native = TRUE)
+right_bottom_img <- readPNG("figures/mnist/mnist_digit1_img_rb.png", native = TRUE)
+
+pacmap_plot_mnist <- pacmap_plot_mnist +
+    inset_element(p = left_top_img, 
+                  left = 0.8, 
+                  bottom = 0.85, 
+                  right = 0, 
+                  top = 0.95) +
+    inset_element(p = center_img, 
+                  left = 1.2, 
+                  bottom = 0.6, 
+                  right = 0, 
+                  top = 0.7) +
+    inset_element(p = right_bottom_img, 
+                  left = 1.7, 
+                  bottom = 0.48, 
+                  right = 0, 
+                  top = 0.58)
 ```
@@
 ```{r}
-#| fig-cap: "$2\\text{-}D$ layout from PaCMAP applied for the digit 1 of the MNIST dataset.
-Is this the best representation of the digit 1?
-The parameter setting is n_components=2, n_neighbors=10, init=random, MN_ratio=0.9, FP_ratio=2.0."
+#| fig-cap: "$2\\text{-}D$ layout from PaCMAP applied for the digit 1 of the MNIST dataset with a specific parameter settings.
+The angle of the digit 1 varies along this structure.
+Images at the top-left of the $2\\text{-}D$ layout show the digit 1 angled more to the left, while images at the bottom-right show the digit 1 angled more to the right.
+Is this the best representation of the digit 1?"
 #| label: fig-pacmap-author

### 2024-07-07 paper.tex
Changed NLDR MNIST PaCMAP image

@@
 \caption{\label{fig-pacmap-author}\(2\text{-}D\) layout from PaCMAP
-applied for the digit 1 of the MNIST dataset. Is this the best
-representation of the digit 1? The parameter setting is n\_components=2,
-n\_neighbors=10, init=random, MN\_ratio=0.9, FP\_ratio=2.0.}
+applied for the digit 1 of the MNIST dataset with a specific parameter
+settings. The angle of the digit 1 varies along this structure. Images
+at the top-left of the \(2\text{-}D\) layout show the digit 1 angled
+more to the left, while images at the bottom-right show the digit 1
+angled more to the right. Is this the best representation of the digit
+1?}
 

### 2024-07-09 paper.qmd
revised start of application section

@@
 In the field of single-cell studies, a common analytical task involves clustering to identify groups of cells with similar expression profiles.
-Analysts often turn to NLDR techniques to verify and identify these clusters and explore developmental trajectories.
-To illustrate the importance of NLDR techniques and parameter selection in identifying clusters, Human Peripheral Blood Mononclear Cells (PBMC3k) dataset [@chen2023] is used.
-In a study by @chen2023, this dataset was used to demonstrate how UMAP represents clusters (see @fig-umap-author).
-As shown in @fig-umap-author, there are three distant and well-separated clusters.
+NLDR is commonly used to display clusters, and help to verify the results.
+For example, @chen2023 illustrates the use of UMAP to identify clusters in Human Peripheral Blood Mononuclear Cells (PBMC3k).
+@fig-umap-author is a reproduction of the published plot.
+It shows three well-separated clusters.
 
@@
 ```{r}
-#| fig-cap: "$2\\text{-}D$ layout from UMAP applied for the PBMC3k dataset. Is this a best representation of the original data? The parameter setting is n_neighbors = 30, min_dist = 0.3."
+#| fig-cap: "Reproduction of plot published in Chen et al.
+(2023) showing a $2\\text{-}D$ layout from UMAP applied for the PBMC3k dataset.
+The parameter settings, beyond the defaults, are n_neighbors = 30 and min_dist = 0.3.
+The question is whether this is an accurate representation of structure in the high-dimensional data."
 #| label: fig-umap-author

### 2024-07-09 paper.tex
revised start of application section

@@
 clustering to identify groups of cells with similar expression profiles.
-Analysts often turn to NLDR techniques to verify and identify these
-clusters and explore developmental trajectories. To illustrate the
-importance of NLDR techniques and parameter selection in identifying
-clusters, Human Peripheral Blood Mononclear Cells (PBMC3k) dataset
-\citep{chen2023} is used. In a study by \citet{chen2023}, this dataset
-was used to demonstrate how UMAP represents clusters (see
-Figure~\ref{fig-umap-author}). As shown in Figure~\ref{fig-umap-author},
-there are three distant and well-separated clusters.
+NLDR is commonly used to display clusters, and help to verify the
+results. For example, \citet{chen2023} illustrates the use of UMAP to
+identify clusters in Human Peripheral Blood Mononuclear Cells (PBMC3k).
+Figure~\ref{fig-umap-author} is a reproduction of the published plot. It
+shows three well-separated clusters.
 
@@
 
-\caption{\label{fig-umap-author}\(2\text{-}D\) layout from UMAP applied
-for the PBMC3k dataset. Is this a best representation of the original
-data? The parameter setting is n\_neighbors = 30, min\_dist = 0.3.}
+\caption{\label{fig-umap-author}Reproduction of plot published in Chen
+et al.~(2023) showing a \(2\text{-}D\) layout from UMAP applied for the
+PBMC3k dataset. The parameter settings, beyond the defaults, are
+n\_neighbors = 30 and min\_dist = 0.3. The question is whether this is
+an accurate representation of structure in the high-dimensional data.}
 

### 2024-07-14 paper.qmd
still working on hex grid

@@
 #| echo: false
+# remotes::install_github("JayaniLakshika/quollr")
 library(quollr)
@@
 
-#### Scaling the data
+### Algorithm to represent the model in 2D
+
+#### Scale the data
 
@@
 [0, 1], is recommended.
-When the aspect ratio of the NLDR $(r_1, r_2, ..., r_k)$ is meaningful, the data can be scaled to take this into account.
+The default should take the aspect ratio produced by the NLDR $(r_1, r_2, ..., r_k)$ into account.
 When $k=2$, as in hexagon binning, the default range is $[0, y_{i,\text{max}}], i=1,2$, where $y_{1,\text{max}}=1$ and $y_{2,\text{max}} = \frac{r_2}{r_1}$ (@fig-NLDR-scurve).
+If the NLDR aspect ratio is ignored then set $y_ {2,\text{max}} = 1$.
 <!-- \times \frac{2}{\sqrt{3}}$.
@@
 Each hexagon, $H_h$ is uniquely described by centroid, $C_{h}^{(2)} = (c_{h1}, c_{h2})$.
-(Note that, we detail the algorithm for constructing the grid and binning the data because we found that there was no readily available implementation that could produce what was needed for this project.)
+(Note that, we explain the algorithm for constructing the grid and binning the data here because, surprisingly, there is no readily available implementation that could produce everything needed for this project.)
 
@@
 By default this is 0.1.
-A value should be chosen to provide at buffer of a full hexagon width ($a_1$) and height ($a_2$) for the grid to extend beyond the data.
-The lower left position where the grid starts is defined as $(s_1, s_2)$, corresponding to the centroid of the lowest left hexagon, which should be below the minimum data value.
+A value should be chosen to provide at buffer of a full hexagon width ($a_1$) and height ($a_2$) for the grid to extend beyond the data, in all directions.
+The lower left position where the grid starts is defined as $(s_1, s_2)$, and corresponds to the centroid of the lowest left hexagon.
+This must be smaller than the minimum data value.
 This is computed as one buffer unit, $q$ below the minimum data value, by default.

### 2024-07-14 paper.tex
still working on hex grid

@@
 
-\subsubsection{Scaling the data}\label{scaling-the-data}
+\subsection{Algorithm to represent the model in
+2D}\label{algorithm-to-represent-the-model-in-2d}
+
+\subsubsection{Scale the data}\label{scale-the-data}
 
 Because we are working with distances between points, starting with data
-having a standard scale, e.g.~{[}0, 1{]}, is recommended. When the
-aspect ratio of the NLDR \((r_1, r_2, ..., r_k)\) is meaningful, the
-data can be scaled to take this into account. When \(k=2\), as in
-hexagon binning, the default range is \([0, y_{i,\text{max}}], i=1,2\),
-where \(y_{1,\text{max}}=1\) and \(y_{2,\text{max}} = \frac{r_2}{r_1}\)
-(Figure~\ref{fig-NLDR-scurve}).
+having a standard scale, e.g.~{[}0, 1{]}, is recommended. The default
+should take the aspect ratio produced by the NLDR
+\((r_1, r_2, ..., r_k)\) into account. When \(k=2\), as in hexagon
+binning, the default range is \([0, y_{i,\text{max}}], i=1,2\), where
+\(y_{1,\text{max}}=1\) and \(y_{2,\text{max}} = \frac{r_2}{r_1}\)
+(Figure~\ref{fig-NLDR-scurve}). If the NLDR aspect ratio is ignored then
+set \(y_ {2,\text{max}} = 1\).
 
@@
 Each hexagon, \(H_h\) is uniquely described by centroid,
-\(C_{h}^{(2)} = (c_{h1}, c_{h2})\). (Note that, we detail the algorithm
-for constructing the grid and binning the data because we found that
-there was no readily available implementation that could produce what
-was needed for this project.)
+\(C_{h}^{(2)} = (c_{h1}, c_{h2})\). (Note that, we explain the algorithm
+for constructing the grid and binning the data here because,
+surprisingly, there is no readily available implementation that could
+produce everything needed for this project.)
 
@@
 provide at buffer of a full hexagon width (\(a_1\)) and height (\(a_2\))
-for the grid to extend beyond the data. The lower left position where
-the grid starts is defined as \((s_1, s_2)\), corresponding to the
-centroid of the lowest left hexagon, which should be below the minimum
-data value. This is computed as one buffer unit, \(q\) below the minimum
-data value, by default. The user provides the preferred number of bins
-in the horizontal direction, \(b_1\), and the algorithm computes the
-appropriate number of bins in the vertical direction, \(b_2\).
+for the grid to extend beyond the data, in all directions. The lower
+left position where the grid starts is defined as \((s_1, s_2)\), and
+corresponds to the centroid of the lowest left hexagon. This must be
+smaller than the minimum data value. This is computed as one buffer
+unit, \(q\) below the minimum data value, by default. The user provides
+the preferred number of bins in the horizontal direction, \(b_1\), and
+the algorithm computes the appropriate number of bins in the vertical
+direction, \(b_2\).
 

### 2024-07-15 supplementary/hexagon_binning_algorithm.qmd
Update section lifting model to p-D
Update with quollr 1.4 version

@@
 ```{r}
-# umap_scurve <- read_rds(file = "data/s_curve/s_curve_umap.rds") 
-# 
-# scurve_scaled_obj <- gen_scaled_data(
-#   data = umap_scurve)
-# 
-# umap_scurve_scaled <- scurve_scaled_obj$scaled_nldr
-# lim1 <- scurve_scaled_obj$lim1
-# lim2 <- scurve_scaled_obj$lim2
-# r2 <- diff(lim2)/diff(lim1)
-# 
-# a1 <- calc_bins_y(bin1 = 6, q = 0.1, r2 = r2)$a1
-r2 <- 1.643542
-a1 <- 0.2087578
+umap_scurve <- read_rds(file = here::here("data/s_curve/s_curve_umap.rds"))
+
+scurve_scaled_obj <- gen_scaled_data(
+  data = umap_scurve)
+
+umap_scurve_scaled <- scurve_scaled_obj$scaled_nldr
+lim1 <- scurve_scaled_obj$lim1
+lim2 <- scurve_scaled_obj$lim2
+r2 <- diff(lim2)/diff(lim1)
+
+sc_ltr_pos <- c(0.08, 0.96)
+sc_xlims <- c(-0.5, 1.47)
+sc_ylims <- c(-0.32, 2.1)
+```
+
+<!--Full hexagon grid with UMAP data-->
+
+```{r}
+## Compute hexbin parameters
+num_bins_x_scurve <- 7
+
+## hexagon binning to have regular hexagons
+hb_obj_scurve <- hex_binning(
+  data = umap_scurve_scaled, 
+  bin1 = num_bins_x_scurve, 
+  r2 = r2)
+
+## Data set with all centroids
+all_centroids_df <- hb_obj_scurve$centroids
+
+## Generate all coordinates of hexagons
+hex_grid <- hb_obj_scurve$hex_poly
 ```
@@
 ```{r}
-all_centroids_df <- gen_centroids(bin1 = 6, r2 = r2)
-```
+# Code to draw illustration for notation
+start_pt <- all_centroids_df |> 
+  filter(hexID == 1)
+d_rect <- tibble(x1min = 0, 
+                 x1max = 1,
+                 x2min = 0,
+                 x2max = r2)
 
-```{r}
-hex_grid <- gen_hex_coord(all_centroids_df, a1 = a1)
-```
+c_start <- tibble(s1 = 0.03, s2 = -0.05)
 
-```{r}
-hex_param_vis <- ggplot(data = hex_grid, aes(x = x, y = y)) + 
-    geom_polygon(fill = "white", color = "#bdbdbd", aes(group = hex_poly_id)) +
-    geom_point(data = all_centroids_df, aes(x = c_x, y = c_y), color = "#31a354", size = 0.9) +
-    geom_point(data = all_centroids_df |> dplyr::filter(hexID == 1), aes(x = c_x, y = c_y), color = "black") + ## starting point
-    geom_rect(aes(xmin = 0 - 0.03, xmax = 1 - 0.03, ymin = 0 - 0.05, ymax = r2 - 0.05), fill = "white", color = "black", alpha = 0, linewidth = 0.7) +
-    geom_segment(x = 0 - 0.03, y = -0.35, xend = 1 - 0.03, yend = -0.35,
-                 arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+
-    annotate("text", x=0.5, y=-0.45, label= "r1", color = "black") +
-    geom_segment(x = -0.25, y = 0 - 0.05, xend = -0.25, yend = r2 - 0.05,
-                 arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ 
-    annotate("text", x=-0.35, y=1, label= "r2", color = "black") +
-    geom_segment(x = -0.1 + 0.2087578, y = -0.15, xend = -0.1 + 0.2087578*2, yend = -0.15,
-                   arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ # a1 = 0.2087578
-    annotate("text", x=0.2, y=-0.25, label= "a1", color = "black") +
-    geom_segment(x = -0.15, y = -0.1*r2 + 0.1807896*2, xend = -0.15, yend = -0.1*r2 + 0.1807896*3,
-                     arrow = arrow(length = unit(0.03, "npc"), ends = "both"), color = "black")+ # a2 = 0.1807896
-    annotate("text", x=-0.15, y=0.45, label= "a2", color = "black") +
-    annotate("text", x=-0.18, y=-0.45, label= "(s1, s2)", color = "black") ## starting point
+a1 <- tibble(x = all_centroids_df$c_x[4],
+             xend = all_centroids_df$c_x[5],
+             y = all_centroids_df$c_y[21],
+             yend = all_centroids_df$c_y[21],
+             label = expression(a[1]))
+a2 <- tibble(x = all_centroids_df$c_x[27],
+             xend = all_centroids_df$c_x[27],
+             y = all_centroids_df$c_y[27],
+             yend = all_centroids_df$c_y[33],
+             label = expression(a[2]))
+hex_param_vis <- ggplot() + 
+    geom_polygon(data = hex_grid, 
+                        aes(x = x, 
+                            y = y, 
+                            group = hex_poly_id),
+                 fill = "white", 
+                 color = "#bdbdbd") +
+    geom_point(data = all_centroids_df, aes(
+      x = c_x, 
+      y = c_y), 
+      color = "#31a354", size = 0.9) +
+    geom_point(data = start_pt, aes(x = c_x, 
+                                    y = c_y), 
+               color = "black") + 
+    geom_rect(data=d_rect, 
+              aes(xmin = x1min - c_start$s1,# - c_start$s1, 
+                  xmax = x1max - c_start$s1,# - c_start$s1, 
+                  ymin = x2min - c_start$s2,# - c_start$s2, 
+                  ymax = x2max - c_start$s2),# - c_start$s2), 
+              fill = "white", 
+              color = "black", 
+              alpha = 0, 
+              linewidth = 0.7) +
+    geom_point(data=d_rect, aes(x=x1min - c_start$s1, 
+                                y=x2min - c_start$s2)) + 
+    geom_point(data=d_rect, aes(x=x1max - c_start$s1, 
+                                y=x2min - c_start$s2)) + 
+    geom_point(data=d_rect, aes(x=x1min - c_start$s1, 
+                                y=x2max - c_start$s2)) + 
+    annotate("text", x=d_rect$x1min - c_start$s1, 
+                     y=d_rect$x2min - c_start$s2,
+                     label = "(0,0)", 
+             hjust=-0.1, vjust=-0.3) + 
+    annotate("text", x=d_rect$x1max - c_start$s1, 
+                     y=d_rect$x2min - c_start$s2,
+                     label = "(0,1)", 
+             hjust=1.1, vjust=-0.3) + 
+    annotate("text", x=d_rect$x1min - c_start$s1, 
+                     y=d_rect$x2max - c_start$s2,
+                     label = expression(group("(", 
+                        list(0, y[2][max]),")")), 
+            hjust=-0.1, vjust=1.2) + 
+    geom_segment(data=d_rect, aes(
+      x = x1min  - c_start$s1, # 0 - 0.03, 
+      y = -0.35, 
+      xend = x1max - c_start$s1, #1 - 0.03, 
+      yend = -0.35), #-0.35),
+      arrow = arrow(length = unit(0.03, "npc"),
+                               ends = "both"), 
+                 color = "black")+
+    annotate("text", x=0.5, y=-0.45, 
+             label = expression(r[1]), color = "black") +
+    geom_segment(data=d_rect, aes(
+      x = -0.25, 
+      y = x2min - c_start$s2, #0 - 0.05, 
+      xend = -0.25, 
+      yend = x2max - c_start$s2), #r2 - 0.05),
+      arrow = arrow(length = unit(0.03, "npc"),
+                       ends = "both"), 
+                 color = "black")+ 
+    annotate("text", x=-0.35, y=1, 
+             label = expression(r[2]), color = "black") +
+    geom_segment(data = a1, aes(
+      x = x, #-0.1 + 0.2087578, 
+      y = y, #-0.15, 
+      xend = xend, #-0.1 + 0.2087578*2, 
+      yend = yend), #-0.15),
+      arrow = arrow(length = unit(0.03, "npc"),
+        ends = "both"), 
+        color = "black")+ # a1 = 0.2087578
+    annotate("text", 
+             x=(a1$x+a1$xend)/2, 
+             y=a1$y, 
+             label = expression(a[1]), 
+             color = "black",
+             vjust = 1.2) +
+    geom_segment(data = a2, aes(
+      x = x, #-0.15, 
+      y = y, #-0.1*r2 + 0.1807896*2, 
+      xend = xend, #-0.15, 
+      yend = yend), #-0.1*r2 + 0.1807896*3),
+      arrow = arrow(length = unit(0.03, "npc"),
+                               ends = "both"), 
+      color = "black") + # a2 = 0.1807896
+    annotate("text", x=a2$x, y=(a2$y+a2$yend)/2, 
+             label = expression(a[2]), 
+             color = "black", hjust=-0.2) +
+    annotate("text", x=-0.18, y=-0.25, 
+      label = expression(group("(", list(s[1], s[2]), ")")),
+      color = "black") 
 ```
@@
 #| label: fig-hex-param
-#| fig-cap: "Notations for hexagonal grid configurations."
+#| fig-cap: "The components of the hexagon grid illustrating notation."
+#| out-height: 30%
+#| fig-width: 2.5
+#| fig-height: 3.3
 #| fig-pos: H
-#| fig-width: 2
-#| fig-height: 3
 
@@
 ```{r}
-all_centroids_df <- gen_centroids(bin1 = 3, r2 = r2)
+all_centroids_df <- gen_centroids(bin1 = 4, r2 = r2, q = 0.1)
 
-hex_grid <- gen_hex_coord(all_centroids_df, a1 = 0.5218947)
+hex_grid <- gen_hex_coord(all_centroids_df, a1 = 1.043789)
 

### 2024-07-15 paper.qmd
Update the hexbinning text

@@
 This must be smaller than the minimum data value.
-This is computed as one buffer unit, $q$ below the minimum data value, by default.
-The user provides the preferred number of bins in the horizontal direction, $b_1$, and the algorithm computes the appropriate number of bins in the vertical direction, $b_2$.
+This is computed as $s_1 = -q - a_1$, and $s_2 = -q \times r_2$ to have buffer around the data space.
+<!-- This is computed as one buffer unit, $q$ below the minimum data value, by default.--> The user provides the preferred number of bins in the horizontal direction, $b_1$, and the algorithm computes the appropriate number of bins in the vertical direction, $b_2$.
 
 The value for $b_2$ is computed by fixing $b_1$.
-Considering the lower bound of the NLDR, $a_1 > -2q$, and $a_1 > \frac{1+q}{b_1 -1}$.
-Similarly, according to the upper bound of the NLDR, $a_1 > \frac{2r_2(1 + q)}{\sqrt{3}(b_2 - 1)}$, because $a_2 = \frac{\sqrt(3)}{2}a_1$ for regular hexagons.
-Therefore, $b_2 = \Big\lceil1 +\frac{2r_2(b_1 - 1)}{\sqrt{3}}\Big\rceil$.
+Considering the upper bound of the first NLDR component, $a_1 > \frac{1+q}{b_1 -3}$.
+Similarly, for the second NLDR component, $a_2 > \frac{r_2(1 + q)}{(b_2 - 2)}$.
+Since $a_2 = \frac{\sqrt(3)}{2}a_1$ for regular hexagons, $a_1 > \frac{2r_2(1 + q)}{\sqrt{3}(b_2 - 2)}$.
+This is a linear optimization problem.
+Therefore, the optimal solution must occur on a vertex.
+Therefore, $b_2 = \Big\lceil2 +\frac{2r_2(b_1 - 3)}{\sqrt{3}}\Big\rceil$.
 

### 2024-07-15 paper.tex
Update with quollr 1.4 version
Update the hexbinning text

@@
 % Make \paragraph and \subparagraph free-standing
+\makeatletter
 \ifx\paragraph\undefined\else
   \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+  \renewcommand{\paragraph}{
+    \@ifstar
+      \xxxParagraphStar
+      \xxxParagraphNoStar
+  }
+  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
+  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
 \fi
@@
   \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+  \renewcommand{\subparagraph}{
+    \@ifstar
+      \xxxSubParagraphStar
+      \xxxSubParagraphNoStar
+  }
+  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
+  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
 \fi
+\makeatother
 
@@
 \makeatother
+
 \ifLuaTeX
@@
 corresponds to the centroid of the lowest left hexagon. This must be
-smaller than the minimum data value. This is computed as one buffer
-unit, \(q\) below the minimum data value, by default. The user provides
-the preferred number of bins in the horizontal direction, \(b_1\), and
-the algorithm computes the appropriate number of bins in the vertical
-direction, \(b_2\).
+smaller than the minimum data value. This is computed as
+\(s_1 = -q - a_1\), and \(s_2 = -q \times r_2\) to have buffer around
+the data space. The user provides the preferred number of bins in the
+horizontal direction, \(b_1\), and the algorithm computes the
+appropriate number of bins in the vertical direction, \(b_2\).
 
 The value for \(b_2\) is computed by fixing \(b_1\). Considering the
-lower bound of the NLDR, \(a_1 > -2q\), and
-\(a_1 > \frac{1+q}{b_1 -1}\). Similarly, according to the upper bound of
-the NLDR, \(a_1 > \frac{2r_2(1 + q)}{\sqrt{3}(b_2 - 1)}\), because
-\(a_2 = \frac{\sqrt(3)}{2}a_1\) for regular hexagons. Therefore,
-\(b_2 = \Big\lceil1 +\frac{2r_2(b_1 - 1)}{\sqrt{3}}\Big\rceil\).
+upper bound of the first NLDR component, \(a_1 > \frac{1+q}{b_1 -3}\).
+Similarly, for the second NLDR component,
+\(a_2 > \frac{r_2(1 + q)}{(b_2 - 2)}\). Since
+\(a_2 = \frac{\sqrt(3)}{2}a_1\) for regular hexagons,
+\(a_1 > \frac{2r_2(1 + q)}{\sqrt{3}(b_2 - 2)}\). This is a linear
+optimization problem. Therefore, the optimal solution must occur on a
+vertex. Therefore,
+\(b_2 = \Big\lceil2 +\frac{2r_2(b_1 - 3)}{\sqrt{3}}\Big\rceil\).
 
@@
 
-\captionsetup{labelsep=none}\includegraphics{figures/mnist/sc_1.png}
+\includegraphics{figures/mnist/sc_1.png}
 
@@
 
-\captionsetup{labelsep=none}\includegraphics{figures/mnist/sc_2.png}
+\includegraphics{figures/mnist/sc_2.png}
 
@@
 
-\captionsetup{labelsep=none}\includegraphics{figures/mnist/sc_3.png}
+\includegraphics{figures/mnist/sc_3.png}
 

### 2024-07-15 paper.qmd
more changes and suggestions on methods section and application

--- 2024-07-14 paper.qmd
@@
 
+XXX The points on the "theoretical model" should be equidistant along the S. They are still to bunched in some spots and too far apart in others.
+
 ```{r}
@@
 
-The \gD{} hexagon grid is defined by the number of bins in each direction $(b_1, b_2)$, giving total number of bins as $b = b_1 \times b_2$, and a unique hexagon id, $h = 1, \dots, b$.
-Each hexagon, $H_h$ is uniquely described by centroid, $C_{h}^{(2)} = (c_{h1}, c_{h2})$.
-(Note that, we explain the algorithm for constructing the grid and binning the data here because, surprisingly, there is no readily available implementation that could produce everything needed for this project.)
-
-To make the grid, a buffer parameter ($q$) is set as a proportion of the plot space.
-By default this is 0.1.
-A value should be chosen to provide at buffer of a full hexagon width ($a_1$) and height ($a_2$) for the grid to extend beyond the data, in all directions.
-The lower left position where the grid starts is defined as $(s_1, s_2)$, and corresponds to the centroid of the lowest left hexagon.
+Although there are several implementations of hexagon binning (REF), and a published paper (REF), surprisingly, none has sufficient detail or components that produce everything needed for this project.
+So we described the process used here.
+@fig-hex-param illustrates the notation used.
+
+The \gD{} hexagon grid is defined by its bin centroids.
+Each hexagon, $H_h$ ($h = 1, \dots, b$) is uniquely described by centroid, $C_{h}^{(2)} = (c_{h1}, c_{h2})$.
+The number of bins in each direction is denoted as $(b_1, b_2)$, with  $b = b_1 \times b_2$ being the total number of bins.
+We expect the user to provide just $b_1$ and we calculate $b_2$ using the NLDR ratio, to compute the grid.
+
+To ensure that the grid covers the range of data values a buffer parameter ($q$) is set as a proportion of the range.
+By default,  $q=0.1$.
+The buffer should be extending a full hexagon width ($a_1$) and height ($a_2$) beyond the data, in all directions.
+The lower left position where the grid starts is defined as $(s_1, s_2)$, and corresponds to the centroid of the lowest left hexagon, $C_{1}^{(2)} = (c_{11}, c_{12})$.
 This must be smaller than the minimum data value.
-This is computed as one buffer unit, $q$ below the minimum data value, by default.
-The user provides the preferred number of bins in the horizontal direction, $b_1$, and the algorithm computes the appropriate number of bins in the vertical direction, $b_2$.
+Because it is one buffer unit, $q$ below the minimum data values, $s_1 = -q$ and $s_2 = -r_2*q$.
 
@@
 
-Delaunay triangulation [@lee1980] is used to connect neighboring centroids, which is needed to preserve neighborhood information when the model is lifted into \pD{}.
+Delaunay triangulation [@lee1980] (XXXREF what R package are you using?) is used to connect neighboring centroids.
+The edges are needed to preserve neighborhood information when the model is lifted into \pD{}.
 
@@
 -->
+
+When shapes are non-linear in the NLDR layout, some edges could be long. It is unlikely
 
@@
 
+XXX You need to have consistent colouring. In langevitour the points corresponding to observations are purple. They should be purple in the \gD{} view too.
+
+XXX Should there be a second long edge removal? You can compute the edges in \pD{} now, and if there are long edges maybe these should be pruned, and we re-plot the \gD{} view.
+
 ### Measuring the fit {#sec-summary}
@@
 :::
+
+XXX Prediction should have a jitter option, to spread the predicted points out fully in the hexagons.
 
@@
 (2023) showing a $2\\text{-}D$ layout from UMAP applied for the PBMC3k dataset.
-The parameter settings, beyond the defaults, are n_neighbors = 30 and min_dist = 0.3.
-The question is whether this is an accurate representation of structure in the high-dimensional data."
+The parameter settings, beyond the defaults, are 30 nearest neighbors and minimum distance equal to 0.3.
+We use our model-in-the-data-space to assess whether this is an accurate representation of structure present in the high-dimensional data, or if it is misleading."
 #| label: fig-umap-author

### 2024-07-15 paper.tex
more changes and suggestions on methods section and application

--- 2024-07-14 paper.tex
@@
 
+XXX The points on the ``theoretical model'' should be equidistant along
+the S. They are still to bunched in some spots and too far apart in
+others.
+
 \begin{figure}
@@
 
-The \gD{} hexagon grid is defined by the number of bins in each
-direction \((b_1, b_2)\), giving total number of bins as
-\(b = b_1 \times b_2\), and a unique hexagon id, \(h = 1, \dots, b\).
-Each hexagon, \(H_h\) is uniquely described by centroid,
-\(C_{h}^{(2)} = (c_{h1}, c_{h2})\). (Note that, we explain the algorithm
-for constructing the grid and binning the data here because,
-surprisingly, there is no readily available implementation that could
-produce everything needed for this project.)
-
-To make the grid, a buffer parameter (\(q\)) is set as a proportion of
-the plot space. By default this is 0.1. A value should be chosen to
-provide at buffer of a full hexagon width (\(a_1\)) and height (\(a_2\))
-for the grid to extend beyond the data, in all directions. The lower
-left position where the grid starts is defined as \((s_1, s_2)\), and
-corresponds to the centroid of the lowest left hexagon. This must be
-smaller than the minimum data value. This is computed as one buffer
-unit, \(q\) below the minimum data value, by default. The user provides
-the preferred number of bins in the horizontal direction, \(b_1\), and
-the algorithm computes the appropriate number of bins in the vertical
-direction, \(b_2\).
+Although there are several implementations of hexagon binning (REF), and
+a published paper (REF), surprisingly, none has sufficient detail or
+components that produce everything needed for this project. So we
+described the process used here. Figure~\ref{fig-hex-param} illustrates
+the notation used.
+
+The \gD{} hexagon grid is defined by its bin centroids. Each hexagon,
+\(H_h\) (\(h = 1, \dots, b\)) is uniquely described by centroid,
+\(C_{h}^{(2)} = (c_{h1}, c_{h2})\). The number of bins in each direction
+is denoted as \((b_1, b_2)\), with \(b = b_1 \times b_2\) being the
+total number of bins. We expect the user to provide just \(b_1\) and we
+calculate \(b_2\) using the NLDR ratio, to compute the grid.
+
+To ensure that the grid covers the range of data values a buffer
+parameter (\(q\)) is set as a proportion of the range. By default,
+\(q=0.1\). The buffer should be extending a full hexagon width (\(a_1\))
+and height (\(a_2\)) beyond the data, in all directions. The lower left
+position where the grid starts is defined as \((s_1, s_2)\), and
+corresponds to the centroid of the lowest left hexagon,
+\(C_{1}^{(2)} = (c_{11}, c_{12})\). This must be smaller than the
+minimum data value. Because it is one buffer unit, \(q\) below the
+minimum data values, \(s_1 = -q\) and \(s_2 = -r_2*q\).
 
@@
 
-Delaunay triangulation \citep{lee1980} is used to connect neighboring
-centroids, which is needed to preserve neighborhood information when the
-model is lifted into \pD{}.
+Delaunay triangulation \citep{lee1980} (XXXREF what R package are you
+using?) is used to connect neighboring centroids. The edges are needed
+to preserve neighborhood information when the model is lifted into
+\pD{}.
+
+When shapes are non-linear in the NLDR layout, some edges could be long.
+It is unlikely
 
@@
 \end{figure}%
+
+XXX You need to have consistent colouring. In langevitour the points
+corresponding to observations are purple. They should be purple in the
+\gD{} view too.
+
+XXX Should there be a second long edge removal? You can compute the
+edges in \pD{} now, and if there are long edges maybe these should be
+pruned, and we re-plot the \gD{} view.
 
@@
 \end{figure}%
+
+XXX Prediction should have a jitter option, to spread the predicted
+points out fully in the hexagons.
 
@@
 et al.~(2023) showing a \(2\text{-}D\) layout from UMAP applied for the
-PBMC3k dataset. The parameter settings, beyond the defaults, are
-n\_neighbors = 30 and min\_dist = 0.3. The question is whether this is
-an accurate representation of structure in the high-dimensional data.}
+PBMC3k dataset. The parameter settings, beyond the defaults, are 30
+nearest neighbors and minimum distance equal to 0.3. We use our
+model-in-the-data-space to assess whether this is an accurate
+representation of structure present in the high-dimensional data, or if
+it is misleading.}
 

### 2024-07-15 paper.qmd

@@
 The value for $b_2$ is computed by fixing $b_1$.
-Considering the lower bound of the NLDR, $a_1 > -2q$, and $a_1 > \frac{1+q}{b_1 -1}$.
-Similarly, according to the upper bound of the NLDR, $a_1 > \frac{2r_2(1 + q)}{\sqrt{3}(b_2 - 1)}$, because $a_2 = \frac{\sqrt(3)}{2}a_1$ for regular hexagons.
-Therefore, $b_2 = \Big\lceil1 +\frac{2r_2(b_1 - 1)}{\sqrt{3}}\Big\rceil$.
+Considering the upper bound of the first NLDR component, $a_1 > \frac{1+q}{b_1 -3}$.
+Similarly, for the second NLDR component, $a_2 > \frac{r_2(1 + q)}{(b_2 - 2)}$.
+Since $a_2 = \frac{\sqrt(3)}{2}a_1$ for regular hexagons, $a_1 > \frac{2r_2(1 + q)}{\sqrt{3}(b_2 - 2)}$.
+This is a linear optimization problem.
+Therefore, the optimal solution must occur on a vertex.
+Therefore, $b_2 = \Big\lceil2 +\frac{2r_2(b_1 - 3)}{\sqrt{3}}\Big\rceil$.
 

### 2024-07-15 paper.tex

@@
 The value for \(b_2\) is computed by fixing \(b_1\). Considering the
-lower bound of the NLDR, \(a_1 > -2q\), and
-\(a_1 > \frac{1+q}{b_1 -1}\). Similarly, according to the upper bound of
-the NLDR, \(a_1 > \frac{2r_2(1 + q)}{\sqrt{3}(b_2 - 1)}\), because
-\(a_2 = \frac{\sqrt(3)}{2}a_1\) for regular hexagons. Therefore,
-\(b_2 = \Big\lceil1 +\frac{2r_2(b_1 - 1)}{\sqrt{3}}\Big\rceil\).
+upper bound of the first NLDR component, \(a_1 > \frac{1+q}{b_1 -3}\).
+Similarly, for the second NLDR component,
+\(a_2 > \frac{r_2(1 + q)}{(b_2 - 2)}\). Since
+\(a_2 = \frac{\sqrt(3)}{2}a_1\) for regular hexagons,
+\(a_1 > \frac{2r_2(1 + q)}{\sqrt{3}(b_2 - 2)}\). This is a linear
+optimization problem. Therefore, the optimal solution must occur on a
+vertex. Therefore,
+\(b_2 = \Big\lceil2 +\frac{2r_2(b_1 - 3)}{\sqrt{3}}\Big\rceil\).
 

### 2024-07-18 paper.tex
Added interp citation

@@
 
-Delaunay triangulation \citep{lee1980} (XXXREF what R package are you
-using?) is used to connect neighboring centroids. The edges are needed
-to preserve neighborhood information when the model is lifted into
-\pD{}.
+Delaunay triangulation \citep[@albrecht2024]{lee1980} is used to connect
+neighboring centroids. The edges are needed to preserve neighborhood
+information when the model is lifted into \pD{}.
 

### 2024-07-18 paper.qmd
Added interp citation
Added references for hex binning

@@
 
-Although there are several implementations of hexagon binning (REF), and a published paper (REF), surprisingly, none has sufficient detail or components that produce everything needed for this project.
+Although there are several implementations of hexagon binning [@carr1987], and a published paper [@dan2023], surprisingly, none has sufficient detail or components that produce everything needed for this project.
 So we described the process used here.
@@
 This must be smaller than the minimum data value.
-Because it is one buffer unit, $q$ below the minimum data values, $s_1 = -q$ and $s_2 = -r_2*q$.
+Because it is one buffer unit, $q$ below the minimum data values, $s_1 = -q$ and $s_2 = -qr_2$.
 
@@
 
-Delaunay triangulation [@lee1980] (XXXREF what R package are you using?) is used to connect neighboring centroids.
-The edges are needed to preserve neighborhood information when the model is lifted into \pD{}.
+Delaunay triangulation [@lee1980;@alb2024] is used to connect neighboring centroids. The edges are needed to preserve neighborhood information when the model is lifted into \pD{}.
 

### 2024-07-18 paper.tex
Added references for hex binning

@@
 
-Although there are several implementations of hexagon binning (REF), and
-a published paper (REF), surprisingly, none has sufficient detail or
-components that produce everything needed for this project. So we
-described the process used here. Figure~\ref{fig-hex-param} illustrates
-the notation used.
+Although there are several implementations of hexagon binning
+\citep{carr1987}, and a published paper \citep{dan2023}, surprisingly,
+none has sufficient detail or components that produce everything needed
+for this project. So we described the process used here.
+Figure~\ref{fig-hex-param} illustrates the notation used.
 
@@
 minimum data value. Because it is one buffer unit, \(q\) below the
-minimum data values, \(s_1 = -q\) and \(s_2 = -r_2*q\).
+minimum data values, \(s_1 = -q\) and \(s_2 = -qr_2\).
 
@@
 
-Delaunay triangulation \citep[@albrecht2024]{lee1980} is used to connect
+Delaunay triangulation \citep{lee1980, alb2024} is used to connect
 neighboring centroids. The edges are needed to preserve neighborhood

### 2024-07-18 paper.qmd
Added CFA citations
Update Figure4 with quollr 0.1.5

@@
   bin1 = num_bins_x_scurve, 
-  r2 = r2)
+  r2 = r2,
+  q = 0.1)
 
@@
              label = expression(a[1]))
-a2 <- tibble(x = all_centroids_df$c_x[27],
-             xend = all_centroids_df$c_x[27],
-             y = all_centroids_df$c_y[27],
+a2 <- tibble(x = all_centroids_df$c_x[25],
+             xend = all_centroids_df$c_x[25],
+             y = all_centroids_df$c_y[25],
              yend = all_centroids_df$c_y[33],
@@
 The model closely fits the shape, but it has breaks which means that it does not adequately capture the surface.
-(The **langevitour** software is used to view the data with a tour, and the full video is available at (<https://youtu.be/N6iWEj3aVps>).
+(The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/N6iWEj3aVps>).
 :::
@@
 
-The model here is similar to a confirmatory factor analysis model (REF), $\widehat{T}(X_1, X_2, X_3) + \Epsilon$.
+The model here is similar to a confirmatory factor analysis model [@brown2015], $\widehat{T}(X_1, X_2, X_3) + \Epsilon$.
 The difference between the fitted model and observed values would be considered to be residuals, and for this problem are $7\text{-}D$.

### 2024-07-18 paper.tex
Added CFA citations
Update Figure4 with quollr 0.1.5

@@
 \textbf{langevitour} software is used to view the data with a tour, and
-the full video is available at (\url{https://youtu.be/N6iWEj3aVps}).}
+the full video is available at \url{https://youtu.be/N6iWEj3aVps}).}
 
@@
 
-The model here is similar to a confirmatory factor analysis model (REF),
-\(\widehat{T}(X_1, X_2, X_3) + \Epsilon\). The difference between the
-fitted model and observed values would be considered to be residuals,
-and for this problem are \(7\text{-}D\).
+The model here is similar to a confirmatory factor analysis model
+\citep{brown2015}, \(\widehat{T}(X_1, X_2, X_3) + \Epsilon\). The
+difference between the fitted model and observed values would be
+considered to be residuals, and for this problem are \(7\text{-}D\).
 

### 2024-07-18 paper.tex
some changes made after last commit

--- 2024-07-18 paper.tex
@@
 % Make \paragraph and \subparagraph free-standing
-\makeatletter
 \ifx\paragraph\undefined\else
   \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}{
-    \@ifstar
-      \xxxParagraphStar
-      \xxxParagraphNoStar
-  }
-  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
-  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
+  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
 \fi
@@
   \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}{
-    \@ifstar
-      \xxxSubParagraphStar
-      \xxxSubParagraphNoStar
-  }
-  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
-  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
+  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
 \fi
-\makeatother
 
@@
 \makeatother
-
 \ifLuaTeX
@@
 
-Delaunay triangulation \citep[@albrecht2024]{lee1980} is used to connect
-neighboring centroids. The edges are needed to preserve neighborhood
-information when the model is lifted into \pD{}.
+Delaunay triangulation \citep{lee1980} (XXXREF what R package are you
+using?) is used to connect neighboring centroids. The edges are needed
+to preserve neighborhood information when the model is lifted into
+\pD{}.
 
@@
 
-\includegraphics{figures/mnist/sc_1.png}
+\captionsetup{labelsep=none}\includegraphics{figures/mnist/sc_1.png}
 
@@
 
-\includegraphics{figures/mnist/sc_2.png}
+\captionsetup{labelsep=none}\includegraphics{figures/mnist/sc_2.png}
 
@@
 
-\includegraphics{figures/mnist/sc_3.png}
+\captionsetup{labelsep=none}\includegraphics{figures/mnist/sc_3.png}
 

### 2024-07-18 paper.qmd
some changes made after last commit

@@
       label = expression(group("(", list(s[1], s[2]), ")")),
-      color = "black") 
+      color = "black") +
+  coord_equal()
 ```
@@
 #| out-height: 30%
-#| fig-width: 2.5
-#| fig-height: 3.3
 #| fig-pos: H

### 2024-07-18 paper.qmd
Update the choice of bins

@@
                  x2max = r2)
-#c_start <- tibble(s1 = 0.03, s2 = 0.05)
-c_start <- tibble(s1 = 0.03, s2 = -0.05)
+
+# To move the rectangle to ignore the overlap with the centroids
+rect_adj <- tibble(x1 = 0.03, x2 = 0.03)
 
@@
     geom_rect(data=d_rect, 
-              aes(xmin = x1min - c_start$s1,# - c_start$s1, 
-                  xmax = x1max - c_start$s1,# - c_start$s1, 
-                  ymin = x2min - c_start$s2,# - c_start$s2, 
-                  ymax = x2max - c_start$s2),# - c_start$s2), 
+              aes(xmin = x1min - rect_adj$x1,# - rect_adj$s1, 
+                  xmax = x1max - rect_adj$x1,# - rect_adj$s1, 
+                  ymin = x2min - rect_adj$x2,# - rect_adj$s2, 
+                  ymax = x2max - rect_adj$x2),# - rect_adj$s2), 
               fill = "white", 
@@
               linewidth = 0.7) +
-    geom_point(data=d_rect, aes(x=x1min - c_start$s1, 
-                                y=x2min - c_start$s2)) + 
-    geom_point(data=d_rect, aes(x=x1max - c_start$s1, 
-                                y=x2min - c_start$s2)) + 
-    geom_point(data=d_rect, aes(x=x1min - c_start$s1, 
-                                y=x2max - c_start$s2)) + 
-    annotate("text", x=d_rect$x1min - c_start$s1, 
-                     y=d_rect$x2min - c_start$s2,
+    geom_point(data=d_rect, aes(x=x1min - rect_adj$x1, 
+                                y=x2min - rect_adj$x2)) + 
+    geom_point(data=d_rect, aes(x=x1max - rect_adj$x1, 
+                                y=x2min - rect_adj$x2)) + 
+    geom_point(data=d_rect, aes(x=x1min - rect_adj$x1, 
+                                y=x2max - rect_adj$x2)) + 
+    annotate("text", x=d_rect$x1min - rect_adj$x1, 
+                     y=d_rect$x2min - rect_adj$x2,
                      label = "(0,0)", 
              hjust=-0.1, vjust=-0.3) + 
-    annotate("text", x=d_rect$x1max - c_start$s1, 
-                     y=d_rect$x2min - c_start$s2,
+    annotate("text", x=d_rect$x1max - rect_adj$x1, 
+                     y=d_rect$x2min - rect_adj$x2,
                      label = "(0,1)", 
              hjust=1.1, vjust=-0.3) + 
-    annotate("text", x=d_rect$x1min - c_start$s1, 
-                     y=d_rect$x2max - c_start$s2,
+    annotate("text", x=d_rect$x1min - rect_adj$x1, 
+                     y=d_rect$x2max - rect_adj$x2,
                      label = expression(group("(", 
@@
     geom_segment(data=d_rect, aes(
-      x = x1min  - c_start$s1, # 0 - 0.03, 
+      x = x1min  - rect_adj$x1, # 0 - 0.03, 
       y = -0.35, 
-      xend = x1max - c_start$s1, #1 - 0.03, 
+      xend = x1max - rect_adj$x1, #1 - 0.03, 
       yend = -0.35), #-0.35),
@@
       x = -0.25, 
-      y = x2min - c_start$s2, #0 - 0.05, 
+      y = x2min - rect_adj$x2, #0 - 0.05, 
       xend = -0.25, 
-      yend = x2max - c_start$s2), #r2 - 0.05),
+      yend = x2max - rect_adj$x2), #r2 - 0.05),
       arrow = arrow(length = unit(0.03, "npc"),
@@
 
-<!--add three choice of bins 7, 10, 14-->
+<!--add three choice of bins 8, 11, 14-->
 
@@
   data = umap_scurve_scaled, 
-  bin1 = 7, 
+  bin1 = 8, 
   r2 = r2)
@@
   data = umap_scurve_scaled, 
-  bin1 = 12, 
+  bin1 = 11, 
   r2 = r2,
-  q = 0.07)
+  q = 0.075)
 
@@
   r2 = r2,
-  q = 0.05)
+  q = 0.065)
 
@@
 #| fig-pos: H
-#| fig-cap: "Hexbin density plots of UMAP layout of the S-curve data, using three different bin inputs: (a) $b = 70$ (7, 10), (b) $b = 240$ (12, 20), and (c) $b = 322$ (14, 23).
+#| fig-cap: "Hexbin density plots of UMAP layout of the S-curve data, using three different bin inputs: (a) $b = 112 \\text{ } (8, \\text{ }14)  \\text{ }(q = 0.1)$, (b) $b = 220 \\text{ } (11, \\text{ }20) \\text{ }(q = 0.075)$, and (c) $b = 322 \\text{ } (14, \\text{ }26) \\text{ }(q = 0.065)$.
 Color indicates standardized counts, dark indicating high count and light indicates low count.
@@
 ## To initialize number of bins along the x-axis
-bin1_vec_scurve <- 4:19 #sqrt(NROW(training_data_scurve)/2)
+bin1_vec_scurve <- 2:19 #sqrt(NROW(training_data_scurve)/2)
 
@@
 
-  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.07)$bin2
+  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.1)$bin2
 
@@
     r2 = r2,
-    q = 0.07, 
+    q = 0.1, 
     is_bin_centroid = TRUE,
@@
   geom_line() + 
-  geom_vline(xintercept = 240, linetype="solid", 
+  geom_vline(xintercept = 220, linetype="solid", 
              color = "black", linewidth=0.8, alpha = 0.5) +
-  geom_vline(xintercept = 96, linetype=2, 
+  geom_vline(xintercept = 112, linetype=2, 
              color = "black", linewidth=0.8, alpha = 0.5) +
-  geom_vline(xintercept = 375, linetype=2, 
+  geom_vline(xintercept = 350, linetype=2, 
              color = "black", linewidth=0.8, alpha = 0.5) +

### 2024-07-18 paper.tex
Update the choice of bins

--- 2024-07-18 paper.tex
@@
 \caption{\label{fig-bins-scurve}Hexbin density plots of UMAP layout of
-the S-curve data, using three different bin inputs: (a) \(b = 70\) (7,
-10), (b) \(b = 240\) (12, 20), and (c) \(b = 322\) (14, 23). Color
+the S-curve data, using three different bin inputs: (a)
+\(b = 112 \text{ } (8, \text{ }14)  \text{ }(q = 0.1)\), (b)
+\(b = 220 \text{ } (11, \text{ }20) \text{ }(q = 0.075)\), and (c)
+\(b = 322 \text{ } (14, \text{ }26) \text{ }(q = 0.065)\). Color
 indicates standardized counts, dark indicating high count and light

### 2024-07-18 paper.qmd
Changed Figure 7

@@
 sc_ltr_pos <- c(0.08, 0.96)
-sc_xlims <- c(-0.5, 1.47)
-sc_ylims <- c(-0.32, 2.1)
-# sc_xlims <- c(-0.25, 1.25)
-# sc_ylims <- c(-0.25, 2)
+# sc_xlims <- c(-0.5, 1.47)
+# sc_ylims <- c(-0.32, 2.1)
+sc_xlims <- c(-0.25, 1.25)
+sc_ylims <- c(-0.3, 2)
 
@@
 
-<!--add three choice of bins 8, 11, 14-->
+<!--add three choice of bins 7, 11, 14-->
 
@@
   data = umap_scurve_scaled, 
-  bin1 = 8, 
-  r2 = r2)
+  bin1 = 7, 
+  r2 = r2,
+  q = 0.12)
 
@@
     na.value = "#ffffff", option = "C") +
-  interior_annotation("a") +
-  xlim(sc_xlims) + ylim(sc_ylims)
+  interior_annotation("a", position = c(0.92, 0.93)) +
+  xlim(c(-0.25, 1.35)) + ylim(c(-0.35, 2))
 ```
@@
     na.value = "#ffffff", option = "C") +
-  xlim(sc_xlims) + ylim(sc_ylims) +
+  xlim(c(-0.25, 1.35)) + ylim(c(-0.35, 2)) +
   interior_annotation("b")
@@
   data = umap_scurve_scaled, 
-  bin1 = 14, 
+  bin1 = 17, 
   r2 = r2,
-  q = 0.065)
+  q = 0.05)
 
@@
   scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  xlim(sc_xlims) + ylim(sc_ylims) +
+  xlim(c(-0.25, 1.35)) + ylim(c(-0.35, 2)) +
   interior_annotation("c")
@@
 #| fig-pos: H
-#| fig-cap: "Hexbin density plots of UMAP layout of the S-curve data, using three different bin inputs: (a) $b = 112 \\text{ } (8, \\text{ }14)  \\text{ }(q = 0.1)$, (b) $b = 220 \\text{ } (11, \\text{ }20) \\text{ }(q = 0.075)$, and (c) $b = 322 \\text{ } (14, \\text{ }26) \\text{ }(q = 0.065)$.
+#| fig-cap: "Hexbin density plots of UMAP layout of the S-curve data, using three different bin inputs: (a) $b = 91 \\text{ } (7, \\text{ }13)  \\text{ }(q = 0.12)$, (b) $b = 220 \\text{ } (11, \\text{ }20) \\text{ }(q = 0.075)$, and (c) $b = 527 \\text{ } (17, \\text{ }31) \\text{ }(q = 0.05)$.
 Color indicates standardized counts, dark indicating high count and light indicates low count.
@@
 #| fig-width: 6
-#| fig-height: 2.5
+#| fig-height: 3
 

### 2024-07-18 paper.qmd
Update text with constraints for hex binning

@@
 The value for $b_2$ is computed by fixing $b_1$.
-Considering the upper bound of the first NLDR component, $a_1 > \frac{1+q}{b_1 -3}$.
-Similarly, for the second NLDR component, $a_2 > \frac{r_2(1 + q)}{(b_2 - 2)}$.
-Since $a_2 = \frac{\sqrt(3)}{2}a_1$ for regular hexagons, $a_1 > \frac{2r_2(1 + q)}{\sqrt{3}(b_2 - 2)}$.
+Considering the upper bound of the first NLDR component, $a_1 > \frac{1+2q}{b_1 -1}$.
+Similarly, for the second NLDR component, $a_2 > \frac{r_2 + q(1 + r_2)}{(b_2 - 1)}$.
+Since $a_2 = \frac{\sqrt(3)}{2}a_1$ for regular hexagons, $a_1 > \frac{2[r_2 + q(1 + r_2)]}{\sqrt{3}(b_2 - 1)}$.
 This is a linear optimization problem.
 Therefore, the optimal solution must occur on a vertex.
-Therefore, $b_2 = \Big\lceil2 +\frac{2r_2(b_1 - 3)}{\sqrt{3}}\Big\rceil$.
+Therefore, $b_2 = \Big\lceil1 +\frac{2[r_2 + q(1 + r_2)](b_1 - 1)}{\sqrt{3}(1 + 2q)}\Big\rceil$.
 
@@
   r2 = r2,
-  q = 0.075)
+  q = 0.07)
 
@@
 #| fig-pos: H
-#| fig-cap: "Hexbin density plots of UMAP layout of the S-curve data, using three different bin inputs: (a) $b = 91 \\text{ } (7, \\text{ }13)  \\text{ }(q = 0.12)$, (b) $b = 220 \\text{ } (11, \\text{ }20) \\text{ }(q = 0.075)$, and (c) $b = 527 \\text{ } (17, \\text{ }31) \\text{ }(q = 0.05)$.
+#| fig-cap: "Hexbin density plots of UMAP layout of the S-curve data, using three different bin inputs: (a) $b = 91 \\text{ } (7, \\text{ }13)  \\text{ }(q = 0.12)$, (b) $b = 220 \\text{ } (11, \\text{ }20) \\text{ }(q = 0.07)$, and (c) $b = 527 \\text{ } (17, \\text{ }31) \\text{ }(q = 0.05)$.
 Color indicates standardized counts, dark indicating high count and light indicates low count.
@@
 
-  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.1)$bin2
+  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.07)$bin2
 
@@
     r2 = r2,
-    q = 0.1, 
+    q = 0.07, 
     is_bin_centroid = TRUE,
@@
              color = "black", linewidth=0.8, alpha = 0.5) +
-  geom_vline(xintercept = 350, linetype=2, 
+  geom_vline(xintercept = 364, linetype=2, 
              color = "black", linewidth=0.8, alpha = 0.5) +

### 2024-07-18 paper.qmd
Update PBMC descriptions

@@
 
-tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_51.rds")
+tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_30.rds")
 
@@
 #| out-width: 100%
-# (a) UMAP (n_neighbors = 30, min_dist = 0.3), (b) UMAP (n_neighbors = 5, min_dist = 0.01), (c) UMAP (n_neighbors = 15, min_dist = 0.99), (f) tSNE (perplexity = 5), (g) tSNE (perplexity = 51), (l) TriMAP (n_inliers = 12, n_outliers = 4, n_random = 3), (q) PaCMAP (n_neighbors = 30, init = random, MN_ratio = 0.9, FP_ratio = 5)
+# (a) UMAP (n_neighbors = 30, min_dist = 0.3), (b) UMAP (n_neighbors = 5, min_dist = 0.01), (c) UMAP (n_neighbors = 15, min_dist = 0.99), (f) tSNE (perplexity = 5), (g) tSNE (perplexity = 30), (l) TriMAP (n_inliers = 12, n_outliers = 4, n_random = 3), (q) PaCMAP (n_neighbors = 30, init = random, MN_ratio = 0.9, FP_ratio = 5)
 nldr1 + nldr2 + nldr3 + nldr4 +
@@
 (2023) showing a $2\\text{-}D$ layout from UMAP applied for the PBMC3k dataset.
-The parameter settings, beyond the defaults, are 30 nearest neighbors and minimum distance equal to 0.3.
+The (hyper-)parameter settings, beyond the defaults, are 30 nearest neighbors and minimum distance equal to 0.3.
 We use our model-in-the-data-space to assess whether this is an accurate representation of structure present in the high-dimensional data, or if it is misleading."
@@
 
-To determine whether the UMAP representation with the parameter choice suggested by @chen2023 preserves the original data structure, we visualize the model constructed with UMAP overlaid on the \pD{} data.
-The figures in @fig-pbmc1-sc show three well-separated clusters, indicating that the suggested UMAP representation preserves the global structure (see @fig-umap-author).
-However, as shown in @fig-pbmc1-sc, these clusters are close to each other in \pD{}.
-Also, non-linear continuity patterns and high-density patches within the clusters are observed (see @fig-pbmc1-sc).
-Therefore, the suggested UMAP representation (see @fig-umap-author) does not accurately preserve the local structure of the PBMC3k dataset.
+To determine whether the UMAP representation with the (hyper-)parameter choice suggested by @chen2023 preserves the original data structure, we visualize the model constructed with UMAP overlaid on the \pD{} data.
+The @fig-umap-author shows three well-separated clusters.
+However, as shown in @fig-pbmc1-sc, there is no big separation between three clusters in \pD{}.
+Therefore, the suggested UMAP representation (@fig-umap-author) does not accurately represent the structure(s) present in PBMC3k dataset.
+But, when visualizing the model-in-th-data-space some unobserved structures can be seen.
+Some clusters have non-linear continuity patterns and high-density patches (@fig-pbmc1-sc).
 
@@
 
+XXXX I think we can remove the error figure, it's not part of the story
+
 <!--bin1 = 30, bin2 = 29, b = 870, non_empty = 132-->
@@
 
-In order to find a reasonable NLDR representation for the PBMC3k dataset, we calculated the absolute error for different numbers of non-empty bins using various NLDR techniques and different parameter settings (see @fig-pbmc-abserror).
-After analyzing the results, we found that tSNE with a perplexity set to $30$ had the lowest error when the number of non-empty bins was $137$.
-Therefore, tSNE with a perplexity of $30$, which is the default parameter setting, is considered as a reasonable representation for the PBMC3k dataset.
+In order to find a reasonable NLDR representation for the PBMC3k dataset, the absolute error for different numbers of non-empty bins using various NLDR techniques and different (hyper-)parameter settings (@fig-pbmc-abserror) were calculated.
+After analyzing the results, it was found that tSNE with default (hyper-)parameter setting (perplexity: $30$) achieved the lowest error when there were $124$ number of non-empty bins.
+Therefore, tSNE with a perplexity value set to $30$, the default parameter setting, is considered as a reasonable representation for the PBMC3k dataset.
 
 <!--compute absolute error for different parameter choices-->
-```{r}
-error_df_umap <- read_csv("data/pbmc3k/error_df_umap.csv", col_names = FALSE)
-names(error_df_umap) <- c("error", "mse", "bin1", "bin2", "b", "b_non_empty", 
-                     "method", "n_neighbors", "min_dist")
-
-error_df_umap <- error_df_umap |>
-  mutate(type = paste0("n_neighbors: ", n_neighbors, ", min_dist: ", min_dist)) |>
-  select(-c(n_neighbors, min_dist))
-
-error_df_tsne <- read_csv("data/pbmc3k/error_df_tsne.csv", col_names = FALSE)
-names(error_df_tsne) <- c("error", "mse", "bin1", "bin2", "b", "b_non_empty", 
-                          "method", "perplexity")
-
-error_df_tsne <- error_df_tsne |>
-  mutate(type = paste0("perplexity: ", perplexity)) |>
-  select(-perplexity)
-
-error_df <- bind_rows(error_df_umap, error_df_tsne)
-
-error_plot_pbmc <- ggplot(error_df, 
+
+```{r}
+## To initialize number of bins along the x-axis
+bin1_vec_pbmc <- 2:36 #sqrt(NROW(training_data_pbmc)/2)
+
+## For umap
+umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_30_min_dist_0.3.rds")
+pbmc_scaled_obj_umap <- gen_scaled_data(
+  data = umap_pbmc)
+umap_pbmc_scaled <- pbmc_scaled_obj_umap$scaled_nldr
+
+lim1 <- pbmc_scaled_obj_umap$lim1
+lim2 <- pbmc_scaled_obj_umap$lim2
+r2_umap <- diff(lim2)/diff(lim1)
+
+error_pbmc_umap <- data.frame(matrix(nrow = 0, ncol = 0))
+
+for (xbins in bin1_vec_pbmc) {
+
+  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.1)$bin2
+
+  pbmc_model <- fit_highd_model(
+    training_data = training_data_pbmc,
+    emb_df = umap_pbmc_scaled,
+    bin1 = xbins,
+    r2 = r2_umap,
+    q = 0.1, 
+    is_bin_centroid = TRUE,
+    is_rm_lwd_hex = FALSE,
+    col_start_highd = "PC_"
+  )
+
+  df_bin_centroids_pbmc <- pbmc_model$df_bin_centroids
+  df_bin_pbmc <- pbmc_model$df_bin
+
+  ## Compute error
+  error_df <- glance(
+    df_bin_centroids = df_bin_centroids_pbmc,
+    df_bin = df_bin_pbmc,
+    training_data = training_data_pbmc,
+    newdata = NULL,
+    type_NLDR = "UMAP",
+    col_start = "PC_") |>
+    mutate(bin1 = xbins,
+           bin2 = bin2,
+           b = bin1 * bin2,
+           b_non_empty = NROW(df_bin_centroids_pbmc),
+           method = "UMAP_30_min_dist_0.3")
+
+  error_pbmc_umap <- bind_rows(error_pbmc_umap, error_df)
+
+}
+
+## To initialize number of bins along the x-axis
+bin1_vec_pbmc <- 2:36 #sqrt(NROW(training_data_pbmc)/2)
+
+## For umap
+umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_5_min_dist_0.01.rds")
+pbmc_scaled_obj_umap <- gen_scaled_data(
+  data = umap_pbmc)
+umap_pbmc_scaled <- pbmc_scaled_obj_umap$scaled_nldr
+
+lim1 <- pbmc_scaled_obj_umap$lim1
+lim2 <- pbmc_scaled_obj_umap$lim2
+r2_umap <- diff(lim2)/diff(lim1)
+
+error_pbmc_umap2 <- data.frame(matrix(nrow = 0, ncol = 0))
+
+for (xbins in bin1_vec_pbmc) {
+
+  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.1)$bin2
+
+  pbmc_model <- fit_highd_model(
+    training_data = training_data_pbmc,
+    emb_df = umap_pbmc_scaled,
+    bin1 = xbins,
+    r2 = r2_umap,
+    q = 0.1, 
+    is_bin_centroid = TRUE,
+    is_rm_lwd_hex = FALSE,
+    col_start_highd = "PC_"
+  )
+
+  df_bin_centroids_pbmc <- pbmc_model$df_bin_centroids
+  df_bin_pbmc <- pbmc_model$df_bin
+
+  ## Compute error
+  error_df <- glance(
+    df_bin_centroids = df_bin_centroids_pbmc,
+    df_bin = df_bin_pbmc,
+    training_data = training_data_pbmc,
+    newdata = NULL,
+    type_NLDR = "UMAP",
+    col_start = "PC_") |>
+    mutate(bin1 = xbins,
+           bin2 = bin2,
+           b = bin1 * bin2,
+           b_non_empty = NROW(df_bin_centroids_pbmc),
+           method = "UMAP_5_min_dist_0.01")
+
+  error_pbmc_umap2 <- bind_rows(error_pbmc_umap2, error_df)
+
+}
+
+## For umap
+
+## To initialize number of bins along the x-axis
+bin1_vec_pbmc <- 2:23 #sqrt(NROW(training_data_pbmc)/2)
+umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_15_min_dist_0.99.rds")
+pbmc_scaled_obj_umap <- gen_scaled_data(
+  data = umap_pbmc)
+umap_pbmc_scaled <- pbmc_scaled_obj_umap$scaled_nldr
+
+lim1 <- pbmc_scaled_obj_umap$lim1
+lim2 <- pbmc_scaled_obj_umap$lim2
+r2_umap <- diff(lim2)/diff(lim1)
+
+error_pbmc_umap3 <- data.frame(matrix(nrow = 0, ncol = 0))
+
+for (xbins in bin1_vec_pbmc) {
+
+  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.1)$bin2
+
+  pbmc_model <- fit_highd_model(
+    training_data = training_data_pbmc,
+    emb_df = umap_pbmc_scaled,
+    bin1 = xbins,
+    r2 = r2_umap,
+    q = 0.1, 
+    is_bin_centroid = TRUE,
+    is_rm_lwd_hex = FALSE,
+    col_start_highd = "PC_"
+  )
+
+  df_bin_centroids_pbmc <- pbmc_model$df_bin_centroids
+  df_bin_pbmc <- pbmc_model$df_bin
+
+  ## Compute error
+  error_df <- glance(
+    df_bin_centroids = df_bin_centroids_pbmc,
+    df_bin = df_bin_pbmc,
+    training_data = training_data_pbmc,
+    newdata = NULL,
+    type_NLDR = "UMAP",
+    col_start = "PC_") |>
+    mutate(bin1 = xbins,
+           bin2 = bin2,
+           b = bin1 * bin2,
+           b_non_empty = NROW(df_bin_centroids_pbmc),
+           method = "UMAP_15_min_dist_0.99")
+
+  error_pbmc_umap3 <- bind_rows(error_pbmc_umap3, error_df)
+
+}
+
+## To initialize number of bins along the x-axis
+bin1_vec_pbmc <- 2:22 #sqrt(NROW(training_data_pbmc)/2)
+
+## For tsne
+tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_5.rds")
+pbmc_scaled_obj_tsne <- gen_scaled_data(
+  data = tsne_pbmc)
+tsne_pbmc_scaled <- pbmc_scaled_obj_tsne$scaled_nldr
+
+lim1 <- pbmc_scaled_obj_tsne$lim1
+lim2 <- pbmc_scaled_obj_tsne$lim2
+r2_tsne <- diff(lim2)/diff(lim1)
+
+error_pbmc_tsne <- data.frame(matrix(nrow = 0, ncol = 0))
+
+for (xbins in bin1_vec_pbmc) {
+
+  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.1)$bin2
+
+  pbmc_model <- fit_highd_model(
+    training_data = training_data_pbmc,
+    emb_df = tsne_pbmc_scaled,
+    bin1 = xbins,
+    r2 = r2_tsne,
+    q = 0.1, 
+    is_bin_centroid = TRUE,
+    is_rm_lwd_hex = FALSE,
+    col_start_highd = "PC_"
+  )
+
+  df_bin_centroids_pbmc <- pbmc_model$df_bin_centroids
+  df_bin_pbmc <- pbmc_model$df_bin
+
+  ## Compute error
+  error_df <- glance(
+    df_bin_centroids = df_bin_centroids_pbmc,
+    df_bin = df_bin_pbmc,
+    training_data = training_data_pbmc,
+    newdata = NULL,
+    type_NLDR = "tSNE",
+    col_start = "PC_") |>
+    mutate(bin1 = xbins,
+           bin2 = bin2,
+           b = bin1 * bin2,
+           b_non_empty = NROW(df_bin_centroids_pbmc),
+           method = "tsne_5")
+
+  error_pbmc_tsne <- bind_rows(error_pbmc_tsne, error_df)
+
+}
+
+## To initialize number of bins along the x-axis
+bin1_vec_pbmc <- 2:26 #sqrt(NROW(training_data_pbmc)/2)
+
+## For tsne
+tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_30.rds")
+pbmc_scaled_obj_tsne <- gen_scaled_data(
+  data = tsne_pbmc)
+tsne_pbmc_scaled <- pbmc_scaled_obj_tsne$scaled_nldr
+
+lim1 <- pbmc_scaled_obj_tsne$lim1
+lim2 <- pbmc_scaled_obj_tsne$lim2
+r2_tsne <- diff(lim2)/diff(lim1)
+
+error_pbmc_tsne2 <- data.frame(matrix(nrow = 0, ncol = 0))
+
+for (xbins in bin1_vec_pbmc) {
+
+  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.1)$bin2
+
+  pbmc_model <- fit_highd_model(
+    training_data = training_data_pbmc,
+    emb_df = tsne_pbmc_scaled,
+    bin1 = xbins,
+    r2 = r2_tsne,
+    q = 0.1, 
+    is_bin_centroid = TRUE,
+    is_rm_lwd_hex = FALSE,
+    col_start_highd = "PC_"
+  )
+
+  df_bin_centroids_pbmc <- pbmc_model$df_bin_centroids
+  df_bin_pbmc <- pbmc_model$df_bin
+
+  ## Compute error
+  error_df <- glance(
+    df_bin_centroids = df_bin_centroids_pbmc,
+    df_bin = df_bin_pbmc,
+    training_data = training_data_pbmc,
+    newdata = NULL,
+    type_NLDR = "tSNE",
+    col_start = "PC_") |>
+    mutate(bin1 = xbins,
+           bin2 = bin2,
+           b = bin1 * bin2,
+           b_non_empty = NROW(df_bin_centroids_pbmc),
+           method = "tsne_30")
+
+  error_pbmc_tsne2 <- bind_rows(error_pbmc_tsne2, error_df)
+
+}
+
+## To initialize number of bins along the x-axis
+bin1_vec_pbmc <- 2:46 #sqrt(NROW(training_data_pbmc)/2)
+
+## For phate
+phate_pbmc <- read_rds("data/pbmc3k/pbmc_phate_5.rds")
+pbmc_scaled_obj_phate <- gen_scaled_data(
+  data = phate_pbmc)
+phate_pbmc_scaled <- pbmc_scaled_obj_phate$scaled_nldr
+
+lim1 <- pbmc_scaled_obj_phate$lim1
+lim2 <- pbmc_scaled_obj_phate$lim2
+r2_phate <- diff(lim2)/diff(lim1)
+
+error_pbmc_phate <- data.frame(matrix(nrow = 0, ncol = 0))
+
+for (xbins in bin1_vec_pbmc) {
+
+  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.1)$bin2
+
+  pbmc_model <- fit_highd_model(
+    training_data = training_data_pbmc,
+    emb_df = phate_pbmc_scaled,
+    bin1 = xbins,
+    r2 = r2_phate,
+    q = 0.1, 
+    is_bin_centroid = TRUE,
+    is_rm_lwd_hex = FALSE,
+    col_start_highd = "PC_"
+  )
+
+  df_bin_centroids_pbmc <- pbmc_model$df_bin_centroids
+  df_bin_pbmc <- pbmc_model$df_bin
+
+  ## Compute error
+  error_df <- glance(
+    df_bin_centroids = df_bin_centroids_pbmc,
+    df_bin = df_bin_pbmc,
+    training_data = training_data_pbmc,
+    newdata = NULL,
+    type_NLDR = "PHATE",
+    col_start = "PC_") |>
+    mutate(bin1 = xbins,
+           bin2 = bin2,
+           b = bin1 * bin2,
+           b_non_empty = NROW(df_bin_centroids_pbmc),
+           method = "phate_5")
+
+  error_pbmc_phate <- bind_rows(error_pbmc_phate, error_df)
+
+}
+
+## To initialize number of bins along the x-axis
+bin1_vec_pbmc <- 2:36 #sqrt(NROW(training_data_pbmc)/2)
+
+## For trimap
+trimap_pbmc <- read_rds("data/pbmc3k/pbmc_trimap_12_4_3.rds")
+pbmc_scaled_obj_trimap <- gen_scaled_data(
+  data = trimap_pbmc)
+trimap_pbmc_scaled <- pbmc_scaled_obj_trimap$scaled_nldr
+
+lim1 <- pbmc_scaled_obj_trimap$lim1
+lim2 <- pbmc_scaled_obj_trimap$lim2
+r2_trimap <- diff(lim2)/diff(lim1)
+
+error_pbmc_trimap <- data.frame(matrix(nrow = 0, ncol = 0))
+
+for (xbins in bin1_vec_pbmc) {
+
+  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.1)$bin2
+
+  pbmc_model <- fit_highd_model(
+    training_data = training_data_pbmc,
+    emb_df = trimap_pbmc_scaled,
+    bin1 = xbins,
+    r2 = r2_trimap,
+    q = 0.1, 
+    is_bin_centroid = TRUE,
+    is_rm_lwd_hex = FALSE,
+    col_start_highd = "PC_"
+  )
+
+  df_bin_centroids_pbmc <- pbmc_model$df_bin_centroids
+  df_bin_pbmc <- pbmc_model$df_bin
+
+  ## Compute error
+  error_df <- glance(
+    df_bin_centroids = df_bin_centroids_pbmc,
+    df_bin = df_bin_pbmc,
+    training_data = training_data_pbmc,
+    newdata = NULL,
+    type_NLDR = "TriMAP",
+    col_start = "PC_") |>
+    mutate(bin1 = xbins,
+           bin2 = bin2,
+           b = bin1 * bin2,
+           b_non_empty = NROW(df_bin_centroids_pbmc),
+           method = "trimap_12_4_3")
+
+  error_pbmc_trimap <- bind_rows(error_pbmc_trimap, error_df)
+
+}
+
+## To initialize number of bins along the x-axis
+bin1_vec_pbmc <- 2:43 #sqrt(NROW(training_data_pbmc)/2)
+
+## For pacmap
+pacmap_pbmc <- read_rds("data/pbmc3k/pbmc_pacmap_30_random_0.9_5.rds")
+pbmc_scaled_obj_pacmap <- gen_scaled_data(
+  data = pacmap_pbmc)
+pacmap_pbmc_scaled <- pbmc_scaled_obj_pacmap$scaled_nldr
+
+lim1 <- pbmc_scaled_obj_pacmap$lim1
+lim2 <- pbmc_scaled_obj_pacmap$lim2
+r2_pacmap <- diff(lim2)/diff(lim1)
+
+error_pbmc_pacmap <- data.frame(matrix(nrow = 0, ncol = 0))
+
+for (xbins in bin1_vec_pbmc) {
+
+  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.1)$bin2
+
+  pbmc_model <- fit_highd_model(
+    training_data = training_data_pbmc,
+    emb_df = pacmap_pbmc_scaled,
+    bin1 = xbins,
+    r2 = r2_pacmap,
+    q = 0.1, 
+    is_bin_centroid = TRUE,
+    is_rm_lwd_hex = FALSE,
+    col_start_highd = "PC_"
+  )
+
+  df_bin_centroids_pbmc <- pbmc_model$df_bin_centroids
+  df_bin_pbmc <- pbmc_model$df_bin
+
+  ## Compute error
+  error_df <- glance(
+    df_bin_centroids = df_bin_centroids_pbmc,
+    df_bin = df_bin_pbmc,
+    training_data = training_data_pbmc,
+    newdata = NULL,
+    type_NLDR = "PaCMAP",
+    col_start = "PC_") |>
+    mutate(bin1 = xbins,
+           bin2 = bin2,
+           b = bin1 * bin2,
+           b_non_empty = NROW(df_bin_centroids_pbmc),
+           method = "pacmap")
+
+  error_pbmc_pacmap <- bind_rows(error_pbmc_pacmap, error_df)
+
+}
+
+```
+
+```{r}
+error_pbmc <- bind_rows(error_pbmc_umap, 
+                        error_pbmc_umap2,
+                        error_pbmc_umap3,
+                        error_pbmc_tsne,
+                        error_pbmc_tsne2,
+                        error_pbmc_phate,
+                        error_pbmc_trimap,
+                        error_pbmc_pacmap)
+```
+
+```{r}
+# read the png file from device 
+tsne_5_img <- readPNG("figures/pbmc3k/tsne_5.png", native = TRUE) 
+phate_5_img <- readPNG("figures/pbmc3k/phate_5.png", native = TRUE) 
+umap_img <- readPNG("figures/pbmc3k/author_view.png", native = TRUE) 
+tsne_30_img <- readPNG("figures/pbmc3k/tsne_30.png", native = TRUE)
+
+## Add arrows
+arrow_df_pbmc <- tibble(x = c(94, 245, 101, 150),
+             xend = c(96, 263, 98, 180),
+             y = c(10.06, 9.72, 9.88, 9.855),
+             yend = c(10.14, 9.8, 9.85, 10))
+
+error_plot_pbmc <- ggplot(error_pbmc, 
                               aes(x = b_non_empty, 
-                                  y = log(error), 
-                                  group = type, 
-                                  colour = type)) + 
-  geom_point(size = 1) +
-  geom_line() + 
-  geom_vline(xintercept = 137, linetype="solid", 
+                                  y = log(Error), 
+                                  colour = method)) + 
+  geom_point(size = 0.8) +
+  geom_line(linewidth = 0.3) + 
+  geom_vline(xintercept = 124, linetype="solid", 
              color = "black", linewidth=0.8, alpha = 0.5) +
-  scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a','#984ea3',
-                              '#ff7f00','#a6cee3','#a65628','#f781bf')) +
+  scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a','#ff7f00',
+                              '#984ea3','#999999','#a65628','#f781bf')) +
   ylab("log(absolute error)") +
-  xlab("number of non-empty bins") +
-  theme(axis.text.x = element_text(size = 5),
-        axis.text.y = element_text(size = 5),
-        axis.title.x = element_text(size = 5),
-        axis.title.y = element_text(size = 5, angle = 90))
+  xlab("number of non-empty bins")+
+    geom_segment(data = arrow_df_pbmc, aes(
+      x = x,  
+      y = y,  
+      xend = xend, 
+      yend = yend), 
+      arrow = arrow(length = unit(0.01, "npc"),
+        ends = "first"), 
+        color = "black") +
+    inset_element(p = tsne_5_img, 
+                  left = 1.7, 
+                  bottom = 0.16, 
+                  right = 0, 
+                  top = 0.31) +                   
+    inset_element(p = phate_5_img, 
+                  left = 0.6, 
+                  bottom = 0.5, 
+                  right = 0, 
+                  top = 0.65) +                   
+    inset_element(p = umap_img, 
+                  left = 1.3, 
+                  bottom = 0.3, 
+                  right = 0, 
+                  top = 0.5) +
+    inset_element(p = tsne_30_img, 
+                  left = 0.6, 
+                  bottom = 0.03, 
+                  right = 0, 
+                  top = 0.23) +
+  theme(#aspect.ratio = 0.75,
+        axis.text.x = element_text(size = 7),
+        axis.text.y = element_text(size = 7),
+        axis.title.x = element_text(size = 7),
+        axis.title.y = element_text(size = 7, angle = 90),
+        legend.position = "none",
+        legend.text = element_text(size = 7))
 
@@
 Then, the slope gradually declines or level off, indicating that a higher number of non-empty bins generates a smaller error.
-Using the elbow method, it was observed that when the number of non-empty bins is set to $137$, the lowest error occurred with the parameters perplexity: $30$."
+Using the elbow method, it was observed that when the number of non-empty bins is set to $124$, the lowest error occurred with the parameters perplexity: $30$."
 #| label: fig-pbmc-abserror
 #| fig-pos: H
-#| out-width: 50%
+#| out-width: 70%
 
@@
 As shown in @fig-tsne-suggest, there are three well-separated clusters, although they are located close to each other.
-Additionally, non-linear structures can also be observed within the clusters (see @fig-model-pbmc (a)).
-In this manner, tSNE was able to capture the data structure for the PBMC3k dataset that UMAP failed to do.
+Additionally, non-linear structures can also be observed within the clusters (@fig-tsne-suggest).
+This demonstrates that tSNE accurately captures the data structure of the PBMC3k dataset, which UMAP did not achieve.
 
@@
 ## Compute hexbin parameters
-num_bins_x_pbmc <- 15
+num_bins_x_pbmc <- 16
 lim1 <- pbmc_scaled_obj$lim1
@@
   distance_col = "distance")
-benchmark_pbmc <- 0.1
+#benchmark_pbmc <- 0.1
 

### 2024-07-18 paper.tex
Changed Figure 7
Update text with constraints for hex binning
Update PBMC descriptions

@@
 The value for \(b_2\) is computed by fixing \(b_1\). Considering the
-upper bound of the first NLDR component, \(a_1 > \frac{1+q}{b_1 -3}\).
+upper bound of the first NLDR component, \(a_1 > \frac{1+2q}{b_1 -1}\).
 Similarly, for the second NLDR component,
-\(a_2 > \frac{r_2(1 + q)}{(b_2 - 2)}\). Since
+\(a_2 > \frac{r_2 + q(1 + r_2)}{(b_2 - 1)}\). Since
 \(a_2 = \frac{\sqrt(3)}{2}a_1\) for regular hexagons,
-\(a_1 > \frac{2r_2(1 + q)}{\sqrt{3}(b_2 - 2)}\). This is a linear
-optimization problem. Therefore, the optimal solution must occur on a
-vertex. Therefore,
-\(b_2 = \Big\lceil2 +\frac{2r_2(b_1 - 3)}{\sqrt{3}}\Big\rceil\).
+\(a_1 > \frac{2[r_2 + q(1 + r_2)]}{\sqrt{3}(b_2 - 1)}\). This is a
+linear optimization problem. Therefore, the optimal solution must occur
+on a vertex. Therefore,
+\(b_2 = \Big\lceil1 +\frac{2[r_2 + q(1 + r_2)](b_1 - 1)}{\sqrt{3}(1 + 2q)}\Big\rceil\).
 
@@
 the S-curve data, using three different bin inputs: (a)
-\(b = 112 \text{ } (8, \text{ }14)  \text{ }(q = 0.1)\), (b)
-\(b = 220 \text{ } (11, \text{ }20) \text{ }(q = 0.075)\), and (c)
-\(b = 322 \text{ } (14, \text{ }26) \text{ }(q = 0.065)\). Color
+\(b = 91 \text{ } (7, \text{ }13)  \text{ }(q = 0.12)\), (b)
+\(b = 220 \text{ } (11, \text{ }20) \text{ }(q = 0.07)\), and (c)
+\(b = 527 \text{ } (17, \text{ }31) \text{ }(q = 0.05)\). Color
 indicates standardized counts, dark indicating high count and light
@@
 et al.~(2023) showing a \(2\text{-}D\) layout from UMAP applied for the
-PBMC3k dataset. The parameter settings, beyond the defaults, are 30
-nearest neighbors and minimum distance equal to 0.3. We use our
+PBMC3k dataset. The (hyper-)parameter settings, beyond the defaults, are
+30 nearest neighbors and minimum distance equal to 0.3. We use our
 model-in-the-data-space to assess whether this is an accurate
@@
 
-To determine whether the UMAP representation with the parameter choice
-suggested by \citet{chen2023} preserves the original data structure, we
-visualize the model constructed with UMAP overlaid on the \pD{} data.
-The figures in Figure~\ref{fig-pbmc1-sc} show three well-separated
-clusters, indicating that the suggested UMAP representation preserves
-the global structure (see Figure~\ref{fig-umap-author}). However, as
-shown in Figure~\ref{fig-pbmc1-sc}, these clusters are close to each
-other in \pD{}. Also, non-linear continuity patterns and high-density
-patches within the clusters are observed (see
-Figure~\ref{fig-pbmc1-sc}). Therefore, the suggested UMAP representation
-(see Figure~\ref{fig-umap-author}) does not accurately preserve the
-local structure of the PBMC3k dataset.
+To determine whether the UMAP representation with the (hyper-)parameter
+choice suggested by \citet{chen2023} preserves the original data
+structure, we visualize the model constructed with UMAP overlaid on the
+\pD{} data. The Figure~\ref{fig-umap-author} shows three well-separated
+clusters. However, as shown in Figure~\ref{fig-pbmc1-sc}, there is no
+big separation between three clusters in \pD{}. Therefore, the suggested
+UMAP representation (Figure~\ref{fig-umap-author}) does not accurately
+represent the structure(s) present in PBMC3k dataset. But, when
+visualizing the model-in-th-data-space some unobserved structures can be
+seen. Some clusters have non-linear continuity patterns and high-density
+patches (Figure~\ref{fig-pbmc1-sc}).
+
+XXXX I think we can remove the error figure, it's not part of the story
 
@@
 In order to find a reasonable NLDR representation for the PBMC3k
-dataset, we calculated the absolute error for different numbers of
-non-empty bins using various NLDR techniques and different parameter
-settings (see Figure~\ref{fig-pbmc-abserror}). After analyzing the
-results, we found that tSNE with a perplexity set to \(30\) had the
-lowest error when the number of non-empty bins was \(137\). Therefore,
-tSNE with a perplexity of \(30\), which is the default parameter
-setting, is considered as a reasonable representation for the PBMC3k
-dataset.
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=0.5\textwidth,height=\textheight]{paper_files/figure-pdf/fig-pbmc-abserror-1.pdf}
+dataset, the absolute error for different numbers of non-empty bins
+using various NLDR techniques and different (hyper-)parameter settings
+(Figure~\ref{fig-pbmc-abserror}) were calculated. After analyzing the
+results, it was found that tSNE with default (hyper-)parameter setting
+(perplexity: \(30\)) achieved the lowest error when there were \(124\)
+number of non-empty bins. Therefore, tSNE with a perplexity value set to
+\(30\), the default parameter setting, is considered as a reasonable
+representation for the PBMC3k dataset.
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=0.7\textwidth,height=\textheight]{paper_files/figure-pdf/fig-pbmc-abserror-1.pdf}
 
@@
 non-empty bins generates a smaller error. Using the elbow method, it was
-observed that when the number of non-empty bins is set to \(137\), the
+observed that when the number of non-empty bins is set to \(124\), the
 lowest error occurred with the parameters perplexity: \(30\).}
@@
 Additionally, non-linear structures can also be observed within the
-clusters (see Figure~\ref{fig-model-pbmc} (a)). In this manner, tSNE was
-able to capture the data structure for the PBMC3k dataset that UMAP
-failed to do.
+clusters (Figure~\ref{fig-tsne-suggest}). This demonstrates that tSNE
+accurately captures the data structure of the PBMC3k dataset, which UMAP
+did not achieve.
 

### 2024-07-18 paper.qmd
Update texts of MNIST data set

@@
 The MNIST dataset consists of grayscale images of handwritten digits [@lecun2010].
-@yingfan2021 used this dataset to demonstrate how PaCMAP preserves local structure.
-We selected the \gD{} embedding of PaCMAP for the handwritten digit 1 to assess whether this is a reasonable representation using our method.
-As shown in @fig-pacmap-author-img, the angle of the digit 1 images varies along the \gD{} structure.
+@yingfan2021 used this dataset to demonstrate how PaCMAP preserves non-linear structure in \pD{}.
+To evaluate whether PaCMAP provides a reasonable representation of the data, the \gD{} embedding of the handwritten digit 1 was selected.
+As shown in @fig-pacmap-author, the angle of the digit 1 images varies along the \gD{} structure.
 
@@
 Images at the top-left of the $2\\text{-}D$ layout show the digit 1 angled more to the left, while images at the bottom-right show the digit 1 angled more to the right.
-Is this the best representation of the digit 1?"
+Is this a reasonable representation of the digit 1?"
 #| label: fig-pacmap-author
@@
 pacmap_plot_mnist
-```
-
-<!--Render the images in different selected location in 2D layout-->
-```{r}
-## Data with pixel values
-mnist_data <- read_rds("data/mnist/mnist_digit_1.rds")
-
-pixels_gathered <-  mnist_data |>
-  mutate(instance = row_number()) |>
-  gather(pixel, value, -Label, -instance) |>
-  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
-  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
-  filter(instance %in% c(3710, 2391, 2385, 5030, 6475,
-                         7679, 3568, 1312, 3311, 1097,
-                         3552, 7853, 6489, 7689, 6690,
-                        1380, 6057, 2347, 5946, 3355,
-                        4175, 3997, 5378, 387, 1854,
-                        614, 3079, 1762, 5239, 3723,
-                        5748, 728, 7419, 7794, 6233,
-                        33, 2485, 5998, 318, 1761,
-                        5690, 165, 517, 6935, 2682,
-                        4962, 2264, 5563, 6369, 559,
-                        5188, 4849, 2666, 3448, 3055,
-                        3120, 6869, 6345, 4470, 7147)) |>
-  mutate(instance.labs =  case_when(
-    instance %in% c(3710, 2391, 2385, 5030, 6475) ~ "1",
-    instance %in% c(7679, 3568, 1312, 3311, 1097) ~ "2",
-    instance %in% c(3552, 7853, 6489, 7689, 6690) ~ "3",
-    instance %in% c(1380, 6057, 2347, 5946, 3355) ~ "4",
-    instance %in% c(4175, 3997, 5378, 387, 1854) ~ "5",
-    instance %in% c(614, 3079, 1762, 5239, 372) ~ "6",
-    instance %in% c(5748, 728, 7419, 7794, 6233) ~ "7",
-    instance %in% c(5690, 165, 517, 6935, 2682) ~ "8",
-    instance %in% c(4962, 2264, 5563, 6369, 559) ~ "9",
-    instance %in% c(5188, 4849, 2666, 3448, 3055) ~ "10",
-    instance %in% c(3120, 6869, 6345, 4470, 7147) ~ "11",
-    .default = "12"
-    ))
-
-## To generate labels
-instance.labs <- as.character(rep(1:12, each = 5))
-
-img_sample <- pixels_gathered |> 
-  ggplot(aes(x, y, fill = value)) +  
-  geom_tile() +
-  facet_wrap(~ factor(instance, levels = c(3710, 2391, 2385, 5030, 6475,
-                         7679, 3568, 1312, 3311, 1097,
-                         3552, 7853, 6489, 7689, 6690,
-                        1380, 6057, 2347, 5946, 3355,
-                        4175, 3997, 5378, 387, 1854,
-                        614, 3079, 1762, 5239, 3723,
-                        5748, 728, 7419, 7794, 6233,
-                        33, 2485, 5998, 318, 1761,
-                        5690, 165, 517, 6935, 2682,
-                        4962, 2264, 5563, 6369, 559,
-                        5188, 4849, 2666, 3448, 3055,
-                        3120, 6869, 6345, 4470, 7147)), nrow = 5) +
-  coord_fixed() +
-  scale_fill_continuous_sequential(palette = "Grays") +
-  theme(strip.background = element_blank(),
-        strip.text.x = element_blank(), 
-        legend.position = "none")
-```
-
-```{r}
-#| fig-cap: "Images of the handwritten digit 1 are ordered from the bottom-right to the top-left of the $2\\text{-}D$ structure.
-The angle of the digit varies along this structure.
-Images at the bottom-right of the $2\\text{-}D$ layout show the digit 1 angled more to the right, while images at the top-left show the digit 1 angled more to the left.
-This demonstrates how the angle changes from right to left along the $2\\text{-}D$ structure."
-#| label: fig-pacmap-author-img
-#| fig-pos: H
-#| out-height: 50%
-
-img_sample
 ```
@@
 
-According to @fig-mnist1-sc1, the non-linear continuous structure observed in the \gD{} representation of PaCMAP (see @fig-pacmap-author) is also visible when visualizing the model overlaid on the data space.
+According to @fig-mnist1-sc1, the non-linear continuous structure observed in the \gD{} representation of PaCMAP (@fig-pacmap-author) is also visible when visualizing the model overlaid on the data space.
 This indicates that PaCMAP accurately captures the structure of the \pD{} data.
-Additionally, the model shows a twisted pattern within the non-linear structure in \pD{} space (see @fig-mnist1-sc2), which is an additional pattern not visible in the \gD{} representation (see @fig-pacmap-author).
+Additionally, the model shows a twisted pattern within the non-linear structure in \pD{} space (@fig-mnist1-sc2), which is an additional pattern not visible in the \gD{} representation (@fig-pacmap-author).
 Furthermore, as shown in @fig-mnist1-sc3, some long edges exist in the \pD{} space that are not recognized as long edges in the \gD{} representation.
-However, PaCMAP is a reasonable \gD{} representation of MNIST digit 1 data.
-Because PaCMAP preserves the local structure.
+However, PaCMAP is a reasonable \gD{} representation of MNIST digit 1 data, because PaCMAP preserves the non-linear structure present in the \pD{} data.
 
@@
 
-There are certain data points that exhibit high error rates due to their deviation from the usual \pD{} data structure, which makes them anomalies (see @fig-model-mnist (b)).
+There are certain data points that exhibit high error rates due to their deviation from the usual \pD{} data structure, which makes them anomalies (@fig-model-mnist (b)).
 These anomalies can be classified into two types: those that are anomalies within the non-linear structure and those that lie outside of it.
 The images associated with high model error points within the non-linear structure display different patterns of the digit 1, as shown in @fig-mnist-anomalies (a).
-However, when comparing these images to the ones found outside of the non-linear structure, it becomes evident that the latter display different patterns of the digit 1 (see @fig-mnist-anomalies (b)).
+However, when comparing these images to the ones found outside of the non-linear structure, it becomes evident that the latter display different patterns of the digit 1 (@fig-mnist-anomalies (b)).
 
@@
 ```{r}
+## Data with pixel values
+mnist_data <- read_rds("data/mnist/mnist_digit_1.rds")
+
 pixels_gathered <-  mnist_data |>

### 2024-07-18 paper.tex
Update texts of MNIST data set

@@
 \citep{lecun2010}. \citet{yingfan2021} used this dataset to demonstrate
-how PaCMAP preserves local structure. We selected the \gD{} embedding of
-PaCMAP for the handwritten digit 1 to assess whether this is a
-reasonable representation using our method. As shown in
-Figure~\ref{fig-pacmap-author-img}, the angle of the digit 1 images
-varies along the \gD{} structure.
+how PaCMAP preserves non-linear structure in \pD{}. To evaluate whether
+PaCMAP provides a reasonable representation of the data, the \gD{}
+embedding of the handwritten digit 1 was selected. As shown in
+Figure~\ref{fig-pacmap-author}, the angle of the digit 1 images varies
+along the \gD{} structure.
 
@@
 more to the left, while images at the bottom-right show the digit 1
-angled more to the right. Is this the best representation of the digit
-1?}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=0.5\textheight]{paper_files/figure-pdf/fig-pacmap-author-img-1.pdf}
-
-}
-
-\caption{\label{fig-pacmap-author-img}Images of the handwritten digit 1
-are ordered from the bottom-right to the top-left of the \(2\text{-}D\)
-structure. The angle of the digit varies along this structure. Images at
-the bottom-right of the \(2\text{-}D\) layout show the digit 1 angled
-more to the right, while images at the top-left show the digit 1 angled
-more to the left. This demonstrates how the angle changes from right to
-left along the \(2\text{-}D\) structure.}
+angled more to the right. Is this a reasonable representation of the
+digit 1?}
 
@@
 According to Figure~\ref{fig-mnist1-sc1}, the non-linear continuous
-structure observed in the \gD{} representation of PaCMAP (see
-Figure~\ref{fig-pacmap-author}) is also visible when visualizing the
+structure observed in the \gD{} representation of PaCMAP
+(Figure~\ref{fig-pacmap-author}) is also visible when visualizing the
 model overlaid on the data space. This indicates that PaCMAP accurately
 captures the structure of the \pD{} data. Additionally, the model shows
-a twisted pattern within the non-linear structure in \pD{} space (see
-Figure~\ref{fig-mnist1-sc2}), which is an additional pattern not visible
-in the \gD{} representation (see Figure~\ref{fig-pacmap-author}).
+a twisted pattern within the non-linear structure in \pD{} space
+(Figure~\ref{fig-mnist1-sc2}), which is an additional pattern not
+visible in the \gD{} representation (Figure~\ref{fig-pacmap-author}).
 Furthermore, as shown in Figure~\ref{fig-mnist1-sc3}, some long edges
@@
 \gD{} representation. However, PaCMAP is a reasonable \gD{}
-representation of MNIST digit 1 data. Because PaCMAP preserves the local
-structure.
+representation of MNIST digit 1 data, because PaCMAP preserves the
+non-linear structure present in the \pD{} data.
 
@@
 deviation from the usual \pD{} data structure, which makes them
-anomalies (see Figure~\ref{fig-model-mnist} (b)). These anomalies can be
+anomalies (Figure~\ref{fig-model-mnist} (b)). These anomalies can be
 classified into two types: those that are anomalies within the
@@
 images to the ones found outside of the non-linear structure, it becomes
-evident that the latter display different patterns of the digit 1 (see
-Figure~\ref{fig-mnist-anomalies} (b)).
+evident that the latter display different patterns of the digit 1
+(Figure~\ref{fig-mnist-anomalies} (b)).
 

### 2024-07-18 paper.qmd
Update Digit 1 images

@@
 ## Compute hexbin parameters
-num_bins_x_mnist <- 37
+num_bins_x_mnist <- 22
 lim1 <- mnist_scaled_obj$lim1
@@
 
-<!--bin1 = 37, bin2 = 26, b = 962, non_empty = 398-->
+<!--bin1 = 22, bin2 = 17, b = 374, non_empty = 140-->
 ```{r}
@@
 
-There are certain data points that exhibit high error rates due to their deviation from the usual \pD{} data structure, which makes them anomalies (@fig-model-mnist (b)).
+There are certain data points that exhibit high error rates due to their deviation from the usual \pD{} data structure, which makes them anomalies (@fig-model-mnist b).
 These anomalies can be classified into two types: those that are anomalies within the non-linear structure and those that lie outside of it.
-The images associated with high model error points within the non-linear structure display different patterns of the digit 1, as shown in @fig-mnist-anomalies (a).
-However, when comparing these images to the ones found outside of the non-linear structure, it becomes evident that the latter display different patterns of the digit 1 (@fig-mnist-anomalies (b)).
+The images associated with high model error points within the non-linear structure display different patterns of the digit 1, as shown in @fig-mnist1-within.
+However, when comparing these images to the ones found outside of the non-linear structure, it becomes evident that the latter display different patterns of the digit 1 (@fig-mnist1-out).
 
@@
 <!-- within the nonlinear structure-->
-```{r}
-## Data with pixel values
-mnist_data <- read_rds("data/mnist/mnist_digit_1.rds")
-
-pixels_gathered <-  mnist_data |>
-  mutate(instance = row_number()) |>
-  gather(pixel, value, -Label, -instance) |>
-  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
-  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
-  filter(instance %in% c( 4699, 4785, 5173, 6671, 2186, 4636, 4882, 
-                          773, 2825, 6169, 6948, 5987, 1282, 7715, 
-                          622, 2989, 6912, 4370, 1227, 4042, 2444, 
-                          836, 1654))
-
-imge_error_sample_within <- pixels_gathered |> 
-  ggplot(aes(x, y, fill = value)) +  
-  geom_tile() +  
-  facet_wrap(~ factor(instance, levels = c( 4699, 4785, 5173, 6671, 2186, 4636, 
-                                            4882, 773, 2825, 6169, 6948, 5987, 
-                                            1282, 7715, 622, 2989, 6912, 4370, 
-                                            1227, 4042, 2444, 836, 1654)), 
-             ncol = 4) +
-  coord_fixed() +
-  scale_fill_continuous_sequential(palette = "Grays") +
-  theme(strip.background = element_blank(),
-        strip.text.x = element_blank(),
-        legend.position = "none")
-```
-
 <!--outside the nonlinear structure-->
-```{r}
-img_error <- error_df |> 
-  filter(row_wise_abs_error >= 8) |> 
-  filter((!ID %in% c(4699, 4785, 5173, 6671, 2186, 4636, 4882, 773, 
-                            2825, 6169, 6948, 5987, 1282, 7715, 622, 2989, 
-                            6912, 4370, 1227, 4042, 2444, 836, 1654))) |>
-  pull(ID)
-
-pixels_gathered <-  mnist_data |>
-  mutate(instance = row_number()) |>
-  gather(pixel, value, -Label, -instance) |>
-  extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
-  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
-  filter(instance %in% img_error)
-
-imge_error_sample <- pixels_gathered |> 
-  ggplot(aes(x, y, fill = value)) +  
-  geom_tile() +  
-  facet_wrap(~ instance, ncol = 2) +
-  coord_fixed() +
-  scale_fill_continuous_sequential(palette = "Grays") +
-  theme(strip.background = element_blank(),
-        strip.text.x = element_blank(),
-        legend.position = "none") 
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "Some images of handwritten digit 1 which occur high model error (a) within the non-linear strcuture, and (b) outside the non-linear structure.
-The images shows different patterns of digit 1."
-#| label: fig-mnist-anomalies
-#| out-height: 35%
-#| fig-pos: H
-
-imge_error_sample_within + imge_error_sample +
-  plot_layout(ncol=2) +
-  plot_annotation(tag_levels = 'a')
-```
+
+::: {#fig-mnist-anomalies layout-ncol="2" fig-pos="H"}
+![](figures/mnist/img_error_sample_within.png){#fig-mnist1-within}
+
+![](figures/mnist/img_error_sample_out.png){#fig-mnist1-out}
+
+Some images of handwritten digit 1 which occur high model error (a) within the non-linear structure, and (b) outside the non-linear structure. The images shows different patterns of digit 1.
+:::
 

### 2024-07-18 paper.tex
Update Digit 1 images

@@
 deviation from the usual \pD{} data structure, which makes them
-anomalies (Figure~\ref{fig-model-mnist} (b)). These anomalies can be
+anomalies (Figure~\ref{fig-model-mnist} b). These anomalies can be
 classified into two types: those that are anomalies within the
@@
 display different patterns of the digit 1, as shown in
-Figure~\ref{fig-mnist-anomalies} (a). However, when comparing these
-images to the ones found outside of the non-linear structure, it becomes
-evident that the latter display different patterns of the digit 1
-(Figure~\ref{fig-mnist-anomalies} (b)).
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=0.35\textheight]{paper_files/figure-pdf/fig-mnist-anomalies-1.pdf}
-
-}
+Figure~\ref{fig-mnist1-within}. However, when comparing these images to
+the ones found outside of the non-linear structure, it becomes evident
+that the latter display different patterns of the digit 1
+(Figure~\ref{fig-mnist1-out}).
+
+\begin{figure}[H]
+
+\begin{minipage}{0.50\linewidth}
+
+\centering{
+
+\includegraphics{figures/mnist/img_error_sample_within.png}
+
+}
+
+\subcaption{\label{fig-mnist1-within}}
+
+\end{minipage}%
+%
+\begin{minipage}{0.50\linewidth}
+
+\centering{
+
+\includegraphics{figures/mnist/img_error_sample_out.png}
+
+}
+
+\subcaption{\label{fig-mnist1-out}}
+
+\end{minipage}%
 
 \caption{\label{fig-mnist-anomalies}Some images of handwritten digit 1
-which occur high model error (a) within the non-linear strcuture, and
+which occur high model error (a) within the non-linear structure, and
 (b) outside the non-linear structure. The images shows different

### 2024-07-18 paper.qmd
Added 2D model overlaid on data (MNIST data)

@@
              y = PaCMAP2)) + 
-  geom_point(alpha=0.1)
+  geom_point(alpha=0.1, color = "#6a3d9a")
 
@@
 
-::: {#fig-mnist1-sc layout-ncol="3" fig-pos="H"}
+::: {#fig-mnist1-sc layout-ncol="4" fig-pos="H"}
+![](figures/mnist/mnist_model_2d.png){#fig-mnist1-model}
+
 ![](figures/mnist/sc_1.png){#fig-mnist1-sc1}

### 2024-07-19 paper.qmd
Changed color in NLDR layouts

@@
   ggplot(aes(x = UMAP1, y = UMAP2)) + 
-  geom_point(alpha=0.5, colour="#000000", size = 0.5) +
+  geom_point(alpha=0.5, colour="#6a3d9a", size = 0.5) +
   xlim(sc_xlims) + ylim(sc_ylims) +
@@
              aes(x = UMAP1, y = UMAP2), 
-             alpha = 0.5, size = 0.5) +
+             alpha = 0.5, size = 0.5, color = "#6a3d9a") +
   xlim(sc_xlims) + ylim(sc_ylims) +
@@
              y = UMAP2)) +
-  geom_point(alpha=0.5) 
+  geom_point(alpha=0.3, color = "#6a3d9a") 
 ```
@@
              y = tSNE2)) +
-  geom_point(alpha=0.5)
+  geom_point(alpha=0.3, color = "#6a3d9a")
 ```

### 2024-07-21 paper.qmd
Find one of most appropriate model for PBMC

@@
   r2 = r2_pbmc,
+  q = 0.05,
   is_bin_centroid = TRUE,
@@
 
-<!--bin1 = 30, bin2 = 29, b = 870, non_empty = 132-->
+<!--bin1 = 30, bin2 = 29, b = 870, non_empty = 135-->
 ```{r}
@@
 
-  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.1)$bin2
+  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.05)$bin2
 
@@
     r2 = r2_umap,
-    q = 0.1, 
+    q = 0.05, 
     is_bin_centroid = TRUE,
@@
     r2 = r2_tsne,
-    q = 0.1, 
+    q = 0.05, 
     is_bin_centroid = TRUE,
@@
     r2 = r2_phate,
-    q = 0.1, 
+    q = 0.05, 
     is_bin_centroid = TRUE,
@@
     r2 = r2_trimap,
-    q = 0.1, 
+    q = 0.05, 
     is_bin_centroid = TRUE,
@@
     r2 = r2_pacmap,
-    q = 0.1, 
+    q = 0.05, 
     is_bin_centroid = TRUE,
@@
   geom_line(linewidth = 0.3) + 
-  geom_vline(xintercept = 124, linetype="solid", 
+  geom_vline(xintercept = 136, linetype="solid", 
              color = "black", linewidth=0.8, alpha = 0.5) +
@@
 Then, the slope gradually declines or level off, indicating that a higher number of non-empty bins generates a smaller error.
-Using the elbow method, it was observed that when the number of non-empty bins is set to $124$, the lowest error occurred with the parameters perplexity: $30$."
+Using the elbow method, it was observed that when the number of non-empty bins is set to $136$, the lowest error occurred with the parameters perplexity: $30$."
 #| label: fig-pbmc-abserror
@@
 ## Compute hexbin parameters
-num_bins_x_pbmc <- 16
+num_bins_x_pbmc <- 15
 lim1 <- pbmc_scaled_obj$lim1
@@
   distance_col = "distance")
-#benchmark_pbmc <- 0.1
 
@@
 
-<!--bin1 = 15, bin2 = 20, b = 300, non_empty = 137-->
+<!--bin1 = 15, bin2 = 28, b = 420, non_empty = 136-->
 ```{r}

### 2024-07-21 paper.tex
Added 2D model overlaid on data (MNIST data)
Find one of most appropriate model for PBMC

@@
 non-empty bins generates a smaller error. Using the elbow method, it was
-observed that when the number of non-empty bins is set to \(124\), the
+observed that when the number of non-empty bins is set to \(136\), the
 lowest error occurred with the parameters perplexity: \(30\).}
@@
 
-\begin{minipage}{0.33\linewidth}
+\begin{minipage}{0.25\linewidth}
+
+\centering{
+
+\includegraphics{figures/mnist/mnist_model_2d.png}
+
+}
+
+\subcaption{\label{fig-mnist1-model}}
+
+\end{minipage}%
+%
+\begin{minipage}{0.25\linewidth}
 
@@
 %
-\begin{minipage}{0.33\linewidth}
+\begin{minipage}{0.25\linewidth}
 

### 2024-07-21 paper.qmd
Removed long edges after inspecting the model in p-D
Update choice of bins section

@@
 
-The performance and robustness of our model depend on three key parameters: (i) the total number of bins ($b$), (ii) a benchmark value used to remove low-density hexagons, and (iii) a benchmark value used to remove long edges.
-However, there is no analytical formula to calculate an appropriate value for these parameters.
-The selection of these parameter values depends on the model performance computed by MSE (see @sec-summary).
-
-#### Choice of bins
-
-The number of hexagon bins in the hexagon grid has a considerable impact on the construction of the \gD{} model, serving as the initial step.
-The chosen total number of bins must effectively capture the structure of the NLDR data.
-If the number of bins is too low, the model may not be able to capture the structure of the NLDR data effectively (see @fig-bins-scurve (a)), while if there are too many bins, it may result in over-fitting the individual points of the NLDR data (see @fig-bins-scurve (c)).
-Therefore, it is important to determine an effective number of bins to construct a reasonable model (see @fig-bins-scurve (b)).
+The performance and robustness of our model depend on four key parameters: (i) bin start position $(s_1, \ s_2)$, (ii) the total number of bins ($b$), (iii) a benchmark value used to remove low-density hexagons, and (iv) a benchmark value used to remove long edges.
+However, there is no analytical formula to calculate appropriate values for these parameters.
+The selection of these parameter values depends on the model performance computed by MSE (@sec-summary), except for the threshold used to remove long edges.
+
+#### Bin start position and number of bins
+
+The starting position of the hexagonal grid is important because different starting points result in different distributions of data across bins, even with the same total number of bins.
+This variation affects the model due to the differing number of non-empty bins.
+Therefore, it is necessary to evaluate various starting points with different total numbers of bins to determine which configurations are more effective at capturing the structure and fitting the model.
 
@@
 At the smallest bin size the data segregates into two separate groups, suggesting this is too many bins.
-Using the MSE of the model fit in $p-D$ helps decide on a useful choice of number of bins."
+Using the MSE of the model fit in $p\\text{-}D$ helps decide on a useful choice of number of bins."
 #| fig-width: 6
@@
 The minimum $b_1$ is set to $2$, while the maximum number is estimated by taking the square root of $\frac{n}{2}$.
-The analysis evaluates the MSE across varying $b$ within this range, covering the minimum to maximum values along both axes (see @fig-mse-scurve-b).
+By evaluating MSE across varying $b$ within this range for different $q$, helps to determine an appropriate values for $b$ and $q$ (@fig-param-scurve (a)).
 
@@
 bin1_vec_scurve <- 2:19 #sqrt(NROW(training_data_scurve)/2)
+buffer_vec <- seq(0.05, 0.1, by = 0.01)
 
@@
 for (xbins in bin1_vec_scurve) {
-
-  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.07)$bin2
+  for (buff in buffer_vec) {
+
+    bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = buff)$bin2
 
@@
       r2 = r2,
-    q = 0.07, 
+      q = buff, 
       is_bin_centroid = TRUE,
@@
              b = bin1 * bin2,
-           b_non_empty = NROW(df_bin_centroids_scurve))
+             b_non_empty = NROW(df_bin_centroids_scurve),
+             buff = buff)
 
     error_scurve <- bind_rows(error_scurve, error_df)
+
+  }
 
@@
                      aes(x = b, 
-                         y = log(MSE))) + 
-  geom_point(size = 1) +
-  geom_line() + 
+                         y = log(MSE),
+                         color = buff,
+                         group = buff)) + 
+  geom_point(size = 0.05) +
+  geom_line(linewidth = 0.05) + 
   geom_vline(xintercept = 220, linetype="solid", 
-             color = "black", linewidth=0.8, alpha = 0.5) +
+             color = "black", linewidth=0.1) +
   geom_vline(xintercept = 112, linetype=2, 
-             color = "black", linewidth=0.8, alpha = 0.5) +
+             color = "black", linewidth=0.1) +
   geom_vline(xintercept = 364, linetype=2, 
-             color = "black", linewidth=0.8, alpha = 0.5) +
-  ylab("log(MSE)") +
-  xlab("b") +
+             color = "black", linewidth=0.1) +
+  scale_colour_continuous_sequential(palette = "Magenta") +
+  labs(x = expression(b), y = "log(MSE)", color = expression(q)) +
   ggtitle("(a)") +
@@
             angle=90),
-        title = element_text(size = 72))
-```
-
-```{r}
-#| fig-cap: "MSE from UMAP applied to S-curve dataset with diffferent $b$ choices.
-What is the effective $b$ to create the model?
-The residual plot have a steep slope at the beginning, indicating that a smaller $b$ causes a larger amount of MSE.
-Then, the slope gradually declines or level off, indicating that a higher $b$ generates a smaller MSE.
-Using the elbow method, it was observed that when the $b$ is set to $190$, the lowest MSE occurred."
-#| label: fig-mse-scurve-b
-#| fig-pos: H
-#| out-height: 50%
-#| eval: false
-
-mse_scurve_b
+        title = element_text(size = 72),
+        legend.title = element_text(size = 5),
+        legend.text = element_text(size = 3))
 ```
@@
 #| echo: false
-#| fig-cap: "Various plots to help assess best number of bins, low density bin and large edge removal.
-Both (a) and (c) show MSE, against number of bins and standardised count.
+#| fig-cap: "Various plots to help assess best bin start position, number of bins, low density bin and large edge removal.
+Both (a) and (c) show MSE, against number of bins and standardised count, while (a) also shows the change in MSE for different $q$.
 A good benchmark value for these parameters is when the MSE drops and then flattens out.
 Plot (b) shows the distribution of stadardised counts of hexagons.
-Plot (c) shows the distribution of $2\\text{-}D$ Euclidean distances between bin centroids, with a good benchmark value for removing large edges would being the distance that shows the first large increase."
-
-#| label: fig-dist-scurve
+Plot (d) shows the distribution of $2\\text{-}D$ Euclidean distances between bin centroids, with a good benchmark value for removing large edges would being the distance that shows the first large increase."
+#| label: fig-param-scurve
 #| out-width: 100%
@@
 
+## Remove an edge that exist as long edge in p-D
+# tr_from_to_df_pbmc |> mutate(ID = row_number()) |> filter(from == 37) |> filter(to == 43) 
+tr_from_to_df_pbmc <- tr_from_to_df_pbmc |>
+  filter(row_number() != 360)
+
 trimesh_removed_pbmc <- vis_rmlg_mesh(

### 2024-07-21 paper.tex
Update choice of bins section

@@
 
-The performance and robustness of our model depend on three key
-parameters: (i) the total number of bins (\(b\)), (ii) a benchmark value
-used to remove low-density hexagons, and (iii) a benchmark value used to
-remove long edges. However, there is no analytical formula to calculate
-an appropriate value for these parameters. The selection of these
-parameter values depends on the model performance computed by MSE (see
-Section~\ref{sec-summary}).
-
-\subsubsection{Choice of bins}\label{choice-of-bins}
-
-The number of hexagon bins in the hexagon grid has a considerable impact
-on the construction of the \gD{} model, serving as the initial step. The
-chosen total number of bins must effectively capture the structure of
-the NLDR data. If the number of bins is too low, the model may not be
-able to capture the structure of the NLDR data effectively (see
-Figure~\ref{fig-bins-scurve} (a)), while if there are too many bins, it
-may result in over-fitting the individual points of the NLDR data (see
-Figure~\ref{fig-bins-scurve} (c)). Therefore, it is important to
-determine an effective number of bins to construct a reasonable model
-(see Figure~\ref{fig-bins-scurve} (b)).
+The performance and robustness of our model depend on four key
+parameters: (i) bin start position \((s_1, \ s_2)\), (ii) the total
+number of bins (\(b\)), (iii) a benchmark value used to remove
+low-density hexagons, and (iv) a benchmark value used to remove long
+edges. However, there is no analytical formula to calculate appropriate
+values for these parameters. The selection of these parameter values
+depends on the model performance computed by MSE
+(Section~\ref{sec-summary}), except for the threshold used to remove
+long edges.
+
+\subsubsection{Bin start position and number of
+bins}\label{bin-start-position-and-number-of-bins}
+
+The starting position of the hexagonal grid is important because
+different starting points result in different distributions of data
+across bins, even with the same total number of bins. This variation
+affects the model due to the differing number of non-empty bins.
+Therefore, it is necessary to evaluate various starting points with
+different total numbers of bins to determine which configurations are
+more effective at capturing the structure and fitting the model.
 
@@
 two separate groups, suggesting this is too many bins. Using the MSE of
-the model fit in \(p-D\) helps decide on a useful choice of number of
-bins.}
+the model fit in \(p\text{-}D\) helps decide on a useful choice of
+number of bins.}
 
@@
 while the maximum number is estimated by taking the square root of
-\(\frac{n}{2}\). The analysis evaluates the MSE across varying \(b\)
-within this range, covering the minimum to maximum values along both
-axes (see \textbf{?@fig-mse-scurve-b}).
+\(\frac{n}{2}\). By evaluating MSE across varying \(b\) within this
+range for different \(q\), helps to determine an appropriate values for
+\(b\) and \(q\) (Figure~\ref{fig-param-scurve} (a)).
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-31-1.pdf}
-
-}
-
-\caption{Various plots to help assess best number of bins, low density
-bin and large edge removal. Both (a) and (c) show MSE, against number of
-bins and standardised count. A good benchmark value for these parameters
-is when the MSE drops and then flattens out. Plot (b) shows the
-distribution of stadardised counts of hexagons. Plot (c) shows the
-distribution of \(2\text{-}D\) Euclidean distances between bin
-centroids, with a good benchmark value for removing large edges would
-being the distance that shows the first large increase.}
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-param-scurve-1.pdf}
+
+}
+
+\caption{\label{fig-param-scurve}Various plots to help assess best bin
+start position, number of bins, low density bin and large edge removal.
+Both (a) and (c) show MSE, against number of bins and standardised
+count, while (a) also shows the change in MSE for different \(q\). A
+good benchmark value for these parameters is when the MSE drops and then
+flattens out. Plot (b) shows the distribution of stadardised counts of
+hexagons. Plot (d) shows the distribution of \(2\text{-}D\) Euclidean
+distances between bin centroids, with a good benchmark value for
+removing large edges would being the distance that shows the first large
+increase.}
 

### 2024-07-21 paper.qmd
Added trimesh screenshots for PBMC data

@@
 
-Once setting up the hexagon grid with an appropriate number of bins, some hexagon bins may have few or no data points within them (see @fig-bins-scurve (b)).
+Once setting up the hexagon grid with an appropriate number of bins, some hexagon bins may have few or no data points within them (@fig-bins-scurve (b)).
 To ensure comprehensive coverage of the NLDR data, it is necessary to select hexagon bins with a considerable number of data points.
@@
 Then, the standard count is computed by dividing the number of points within each hexagon by the maximum number of points in the grid.
-Next, bins with a standard count less than a benchmark value are removed (see @fig-lwd-scurve (a)).
+Next, bins with a standard count less than a benchmark value are removed (@fig-param-scurve (c)).
 There is no specific rule for selecting a benchmark value.
@@
 
-1. Plot the distribution of the standardized counts (see @fig-stdcts-scurve).
+1. Plot the distribution of the standardized counts (@fig-param-scurve (b)).
 2. Examine the distribution of counts.
@@
 ## To initialize effective bins along x
-effective_bin1_scurve <- 12
+effective_bin1_scurve <- 11
 
@@
         title = element_text(size = 36))
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "Distribution of standardize counts by hexagons."
-#| label: fig-stdcts-scurve
-#| out-height: 50%
-#| fig-pos: H
-#| eval: false
-
-cell_count_scurve
 ```
@@
 
-```{r}
-#| fig-cap: "MSE from UMAP applied to S-curve dataset with diffferent bechmark choices.
-What is the effective bechmark value to remove low density hexagons?
-The residual plot have a steep slope at the end, indicating that a smaller bechmark value causes a small amount of MSE.
-Then, the slope gradually increases or level up, indicating that a higher bechmark value generates a higher MSE.
-Using the elbow method, it was observed that when the bechmark value is set to $0.242$, the lowest MSE occurred."
-#| label: fig-mse-scurve-lwd
-#| fig-pos: H
-#| out-height: 50%
-#| eval: false
-
-mse_scurve_lwd
-```
-
 Furthermore, selecting the benchmark value for removing low-density hexagons is important.
@@
 
-<!--
-```{r}
-#| echo: false
-#| label: fig-lwd-scurve
-#| fig-pos: H
-#| fig-cap: "Two approaches to identify low-density hexagons.
-Hexagon grid (a) marks hexagons with a standardized count less than $0.25$.
-Hexagon grid (b) marks hexagons by considering the densities of neighboring bins.
-Removing hexagons by also investigating the densities of neighboring bins is more reliable for preserving the structure."
-#| fig-width: 9
-#| fig-height: 6
-
-hex_grid_rm_lwd_scurve1 + hex_grid_rm_lwd_scurve2 +
-  plot_layout(guides='collect', ncol = 2) &
-  theme(legend.position='none', plot.tag = element_text(size = 8))
-```
--->
-
 #### Removing long edges
 
-To create a smooth \gD{} representation (see @fig-NLDR-scurve (d)), it is necessary to remove edges that connect distant bin centroids in the triangular mesh.
+To create a smooth \gD{} representation (@fig-NLDR-scurve (d)), it is necessary to remove edges that connect distant bin centroids in the triangular mesh.
 These edges only exist in the \gD{} model and do not extend into \pD{}, so their removal does not impact the model in \pD{}.
@@
 
-1. Plot the distribution of the 2D Euclidean distances (see @fig-dist-scurve).
+1. Plot the distribution of the 2D Euclidean distances (@fig-param-scurve (d)).
 2. Identify the first largest difference between consecutive distance values.
@@
 The model closely fits the shape, but it has twists.
-(The **langevitour** software is used to view the data with a tour, and the full video is available at (<https://youtu.be/zEz2MUCtADM>).
+(The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/zEz2MUCtADM>).
 :::
@@
 
-::: {#fig-pbmc1-sc layout-ncol="3" fig-pos="H"}
+::: {#fig-pbmc1-sc layout-ncol="4" fig-pos="H"}
+![](figures/pbmc3k/umap_trimesh_plot.png)
+
 ![](figures/pbmc3k/sc_1.png)
@@
 
-Screen shots of the **langevitour** of the PBMC3k data set, shows the model-in-data space, a video of the tour animation is available at (<https://youtu.be/VqqWuE0Jj6A>).
+Model in \gD{}, on the UMAP layout, and three views of the fit in projections from $9\text{-}D$, for the PBMC3k data.
+(The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/VqqWuE0Jj6A>).
 :::
@@
 In order to find a reasonable NLDR representation for the PBMC3k dataset, the absolute error for different numbers of non-empty bins using various NLDR techniques and different (hyper-)parameter settings (@fig-pbmc-abserror) were calculated.
-After analyzing the results, it was found that tSNE with default (hyper-)parameter setting (perplexity: $30$) achieved the lowest error when there were $124$ number of non-empty bins.
+After analyzing the results, it was found that tSNE with default (hyper-)parameter setting (perplexity: $30$) achieved the lowest error when there were $136$ number of non-empty bins.
 Therefore, tSNE with a perplexity value set to $30$, the default parameter setting, is considered as a reasonable representation for the PBMC3k dataset.
@@
   geom_line(linewidth = 0.3) + 
-  geom_vline(xintercept = 136, linetype="solid", 
+  geom_vline(xintercept = 135, linetype="solid", 
              color = "black", linewidth=0.8, alpha = 0.5) +
@@
 #| echo: false
-#| fig-cap: "$2\\text{-}D$ layout from tSNE applied for the PBMC3k dataset.  Is this a best representation of the original data? The parameter setting is perplexity=30."
+#| fig-cap: "$2\\text{-}D$ layout from tSNE applied for the PBMC3k dataset.  Is this a best representation of the original data? The parameter setting is perplexity: 30."
 #| label: fig-tsne-suggest
@@
 
-::: {#fig-pbmc2-sc layout-ncol="3" fig-pos="H"}
+::: {#fig-pbmc2-sc layout-ncol="4" fig-pos="H"}
+![](figures/pbmc3k/tsne_trimesh_plot.png)
+
 ![](figures/pbmc3k/sc_4.png)
@@
 
-Screen shots of the **langevitour** of the PBMC3k data set, shows the model in high-D, a video of the tour animation is available at (<https://youtu.be/5Y1hE4i7N2k>).
+Model in \gD{}, on the tSNE layout, and three views of the fit in projections from $9\text{-}D$, for the PBMC3k data.
+(The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/5Y1hE4i7N2k>).
 :::
@@
 
-Screen shots of the **langevitour** of the MNIST digit 1 data set, shows the model-in-data space, a video of the tour animation is available at (<https://youtu.be/zcg_GXBmqjA>).
+Model in \gD{}, on the PaCMAP layout, and three views of the fit in projections from $10\text{-}D$, for the digit 1 of MNIST data.
+(The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/zcg_GXBmqjA>).
+
 :::

### 2024-07-21 paper.tex
Added trimesh screenshots for PBMC data

@@
 Once setting up the hexagon grid with an appropriate number of bins,
-some hexagon bins may have few or no data points within them (see
-Figure~\ref{fig-bins-scurve} (b)). To ensure comprehensive coverage of
+some hexagon bins may have few or no data points within them
+(Figure~\ref{fig-bins-scurve} (b)). To ensure comprehensive coverage of
 the NLDR data, it is necessary to select hexagon bins with a
@@
 of points in the grid. Next, bins with a standard count less than a
-benchmark value are removed (see \textbf{?@fig-lwd-scurve} (a)). There
+benchmark value are removed (Figure~\ref{fig-param-scurve} (c)). There
 is no specific rule for selecting a benchmark value. However, the
@@
 \item
-  Plot the distribution of the standardized counts (see
-  \textbf{?@fig-stdcts-scurve}).
+  Plot the distribution of the standardized counts
+  (Figure~\ref{fig-param-scurve} (b)).
 \item
@@
 
-To create a smooth \gD{} representation (see
-Figure~\ref{fig-NLDR-scurve} (d)), it is necessary to remove edges that
-connect distant bin centroids in the triangular mesh. These edges only
-exist in the \gD{} model and do not extend into \pD{}, so their removal
-does not impact the model in \pD{}. Although there are no specific
-criteria for determining the benchmark value to remove long edges, the
-following steps provide an approach to identifying a suitable threshold:
+To create a smooth \gD{} representation (Figure~\ref{fig-NLDR-scurve}
+(d)), it is necessary to remove edges that connect distant bin centroids
+in the triangular mesh. These edges only exist in the \gD{} model and do
+not extend into \pD{}, so their removal does not impact the model in
+\pD{}. Although there are no specific criteria for determining the
+benchmark value to remove long edges, the following steps provide an
+approach to identifying a suitable threshold:
 
@@
 \item
-  Plot the distribution of the 2D Euclidean distances (see
-  \textbf{?@fig-dist-scurve}).
+  Plot the distribution of the 2D Euclidean distances
+  (Figure~\ref{fig-param-scurve} (d)).
 \item
@@
 \textbf{langevitour} software is used to view the data with a tour, and
-the full video is available at (\url{https://youtu.be/zEz2MUCtADM}).}
+the full video is available at \url{https://youtu.be/zEz2MUCtADM}).}
 
@@
 
-\begin{minipage}{0.33\linewidth}
+\begin{minipage}{0.25\linewidth}
+\includegraphics{figures/pbmc3k/umap_trimesh_plot.png}\end{minipage}%
+%
+\begin{minipage}{0.25\linewidth}
 \includegraphics{figures/pbmc3k/sc_1.png}\end{minipage}%
 %
-\begin{minipage}{0.33\linewidth}
+\begin{minipage}{0.25\linewidth}
 \includegraphics{figures/pbmc3k/sc_2.png}\end{minipage}%
 %
-\begin{minipage}{0.33\linewidth}
+\begin{minipage}{0.25\linewidth}
 \includegraphics{figures/pbmc3k/sc_3.png}\end{minipage}%
 
-\caption{\label{fig-pbmc1-sc}Screen shots of the \textbf{langevitour} of
-the PBMC3k data set, shows the model-in-data space, a video of the tour
-animation is available at (\url{https://youtu.be/VqqWuE0Jj6A}).}
+\caption{\label{fig-pbmc1-sc}Model in \gD{}, on the UMAP layout, and
+three views of the fit in projections from \(9\text{-}D\), for the
+PBMC3k data. (The \textbf{langevitour} software is used to view the data
+with a tour, and the full video is available at
+\url{https://youtu.be/VqqWuE0Jj6A}).}
 
@@
 results, it was found that tSNE with default (hyper-)parameter setting
-(perplexity: \(30\)) achieved the lowest error when there were \(124\)
+(perplexity: \(30\)) achieved the lowest error when there were \(136\)
 number of non-empty bins. Therefore, tSNE with a perplexity value set to
@@
 for the PBMC3k dataset. Is this a best representation of the original
-data? The parameter setting is perplexity=30.}
+data? The parameter setting is perplexity: 30.}
 
@@
 
-\begin{minipage}{0.33\linewidth}
+\begin{minipage}{0.25\linewidth}
+\includegraphics{figures/pbmc3k/tsne_trimesh_plot.png}\end{minipage}%
+%
+\begin{minipage}{0.25\linewidth}
 \includegraphics{figures/pbmc3k/sc_4.png}\end{minipage}%
 %
-\begin{minipage}{0.33\linewidth}
+\begin{minipage}{0.25\linewidth}
 \includegraphics{figures/pbmc3k/sc_5.png}\end{minipage}%
 %
-\begin{minipage}{0.33\linewidth}
+\begin{minipage}{0.25\linewidth}
 \includegraphics{figures/pbmc3k/sc_6.png}\end{minipage}%
 
-\caption{\label{fig-pbmc2-sc}Screen shots of the \textbf{langevitour} of
-the PBMC3k data set, shows the model in high-D, a video of the tour
-animation is available at (\url{https://youtu.be/5Y1hE4i7N2k}).}
+\caption{\label{fig-pbmc2-sc}Model in \gD{}, on the tSNE layout, and
+three views of the fit in projections from \(9\text{-}D\), for the
+PBMC3k data. (The \textbf{langevitour} software is used to view the data
+with a tour, and the full video is available at
+\url{https://youtu.be/5Y1hE4i7N2k}).}
 
@@
 
-\caption{\label{fig-mnist1-sc}Screen shots of the \textbf{langevitour}
-of the MNIST digit 1 data set, shows the model-in-data space, a video of
-the tour animation is available at
-(\url{https://youtu.be/zcg_GXBmqjA}).}
+\caption{\label{fig-mnist1-sc}Model in \gD{}, on the PaCMAP layout, and
+three views of the fit in projections from \(10\text{-}D\), for the
+digit 1 of MNIST data. (The \textbf{langevitour} software is used to
+view the data with a tour, and the full video is available at
+\url{https://youtu.be/zcg_GXBmqjA}).}
 

### 2024-07-21 paper.qmd
Update the figure title of MNIST PaCMAP layout

@@
   theme(aspect.ratio = 0.75,
-        axis.text.x = element_text(size = 5),
-        axis.text.y = element_text(size = 5),
-        axis.title.x = element_text(size = 5),
-        axis.title.y = element_text(size = 5, 
+        axis.text.x = element_text(size = 7),
+        axis.text.y = element_text(size = 7),
+        axis.title.x = element_text(size = 7),
+        axis.title.y = element_text(size = 7, 
             angle=90),
         title = element_text(size = 72),
-        legend.title = element_text(size = 5),
-        legend.text = element_text(size = 3))
+        legend.title = element_text(size = 7),
+        legend.text = element_text(size = 5))
 ```
@@
 
-Model in \gD{}, on the UMAP layout, and three views of the fit in projections from $9\text{-}D$, for the PBMC3k data.
+Model in \gD{}, on the UMAP layout, and three views of the fit in projections from $9\text{-}D$, for the PBMC3k data ($(s_1, \ s_2) = (-0.050, \ 0.041)$, $b = 870 \  (30, \ 29)$, $m = 135$, benchmark value to remove large edges is $0.099$).
 (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/VqqWuE0Jj6A>).
@@
 #| echo: false
-#| fig-cap: "$2\\text{-}D$ layout from tSNE applied for the PBMC3k dataset.  Is this a best representation of the original data? The parameter setting is perplexity: 30."
+#| fig-cap: "The $2\\text{-}D$ layout from tSNE applied for the PBMC3k dataset.
+The default (hyper-)parameter setting is a perplexity $30$.
+We use our model-in-the-data-space to assess whether this is an accurate representation of structure present in the high-dimensional data, or if it is misleading."
 #| label: fig-tsne-suggest
@@
 
-<!--bin1 = 15, bin2 = 28, b = 420, non_empty = 136-->
+<!--bin1 = 15, bin2 = 20, b = 300, non_empty = 136-->
 ```{r}
@@
 
-Model in \gD{}, on the tSNE layout, and three views of the fit in projections from $9\text{-}D$, for the PBMC3k data.
+Model in \gD{}, on the tSNE layout, and three views of the fit in projections from $9\text{-}D$, for the PBMC3k data ($(s_1, \ s_2) = (-0.050, \ -0.058)$, $b = 300 \  (15, \ 20)$, $m = 136$, benchmark value to remove large edges is $0.133$).
 (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/5Y1hE4i7N2k>).
@@
 ```{r}
-#| fig-cap: "$2\\text{-}D$ layout from PaCMAP applied for the digit 1 of the MNIST dataset with a specific parameter settings.
+#| fig-cap: "$2\\text{-}D$ layout from PaCMAP applied for the digit 1 of the MNIST dataset.
+The (hyper-)parameter settings, beyond the defaults, are 10 nearest neighbors, ratio of the number of mid-near pairs to the number of neighbors equal to 0.5, and the ratio of the number of further pairs to the number of neighbors is 2.
+We use our model-in-the-data-space to assess whether this is an accurate representation of structure present in the high-dimensional data, or if it is misleading.
 The angle of the digit 1 varies along this structure.
-Images at the top-left of the $2\\text{-}D$ layout show the digit 1 angled more to the left, while images at the bottom-right show the digit 1 angled more to the right.
-Is this a reasonable representation of the digit 1?"
+Images at the top-left of the $2\\text{-}D$ layout show the digit 1 angled more to the left, while images at the bottom-right show the digit 1 angled more to the right."
 #| label: fig-pacmap-author
@@
 
-Model in \gD{}, on the PaCMAP layout, and three views of the fit in projections from $10\text{-}D$, for the digit 1 of MNIST data.
+Model in \gD{}, on the PaCMAP layout, and three views of the fit in projections from $10\text{-}D$, for the digit 1 of MNIST data ($(s_1, \ s_2) = (-0.100, \ -0.059)$, $b = 374 \  (22, \ 17)$, $m = 140$, benchmark value to remove large edges is $0.094$).
 (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/zcg_GXBmqjA>).

### 2024-07-21 paper.tex
Update the figure title of MNIST PaCMAP layout

@@
 three views of the fit in projections from \(9\text{-}D\), for the
-PBMC3k data. (The \textbf{langevitour} software is used to view the data
-with a tour, and the full video is available at
+PBMC3k data (\((s_1, \ s_2) = (-0.050, \ 0.041)\),
+\(b = 870 \  (30, \ 29)\), \(m = 135\), benchmark value to remove large
+edges is \(0.099\)). (The \textbf{langevitour} software is used to view
+the data with a tour, and the full video is available at
 \url{https://youtu.be/VqqWuE0Jj6A}).}
@@
 
-\caption{\label{fig-tsne-suggest}\(2\text{-}D\) layout from tSNE applied
-for the PBMC3k dataset. Is this a best representation of the original
-data? The parameter setting is perplexity: 30.}
+\caption{\label{fig-tsne-suggest}The \(2\text{-}D\) layout from tSNE
+applied for the PBMC3k dataset. The default (hyper-)parameter setting is
+a perplexity \(30\). We use our model-in-the-data-space to assess
+whether this is an accurate representation of structure present in the
+high-dimensional data, or if it is misleading.}
 
@@
 three views of the fit in projections from \(9\text{-}D\), for the
-PBMC3k data. (The \textbf{langevitour} software is used to view the data
-with a tour, and the full video is available at
+PBMC3k data (\((s_1, \ s_2) = (-0.050, \ -0.058)\),
+\(b = 300 \  (15, \ 20)\), \(m = 136\), benchmark value to remove large
+edges is \(0.133\)). (The \textbf{langevitour} software is used to view
+the data with a tour, and the full video is available at
 \url{https://youtu.be/5Y1hE4i7N2k}).}
@@
 \caption{\label{fig-pacmap-author}\(2\text{-}D\) layout from PaCMAP
-applied for the digit 1 of the MNIST dataset with a specific parameter
-settings. The angle of the digit 1 varies along this structure. Images
-at the top-left of the \(2\text{-}D\) layout show the digit 1 angled
-more to the left, while images at the bottom-right show the digit 1
-angled more to the right. Is this a reasonable representation of the
-digit 1?}
+applied for the digit 1 of the MNIST dataset. The (hyper-)parameter
+settings, beyond the defaults, are 10 nearest neighbors, ratio of the
+number of mid-near pairs to the number of neighbors equal to 0.5, and
+the ratio of the number of further pairs to the number of neighbors is
+2. We use our model-in-the-data-space to assess whether this is an
+accurate representation of structure present in the high-dimensional
+data, or if it is misleading. The angle of the digit 1 varies along this
+structure. Images at the top-left of the \(2\text{-}D\) layout show the
+digit 1 angled more to the left, while images at the bottom-right show
+the digit 1 angled more to the right.}
 
@@
 three views of the fit in projections from \(10\text{-}D\), for the
-digit 1 of MNIST data. (The \textbf{langevitour} software is used to
-view the data with a tour, and the full video is available at
+digit 1 of MNIST data (\((s_1, \ s_2) = (-0.100, \ -0.059)\),
+\(b = 374 \  (22, \ 17)\), \(m = 140\), benchmark value to remove large
+edges is \(0.094\)). (The \textbf{langevitour} software is used to view
+the data with a tour, and the full video is available at
 \url{https://youtu.be/zcg_GXBmqjA}).}

### 2024-07-21 paper.qmd
Update MNIST PaCMAP Figure title

@@
 
-Model in \gD{}, on the UMAP layout, and three views of the fit in projections from $9\text{-}D$, for the PBMC3k data ($(s_1, \ s_2) = (-0.050, \ 0.041)$, $b = 870 \  (30, \ 29)$, $m = 135$, benchmark value to remove large edges is $0.099$).
+Model in \gD{}, on the UMAP layout, and three views of the fit in projections from $9\text{-}D$, for the PBMC3k data ($(s_1, \ s_2) = (-0.050, \ -0.041)$, $b = 870 \  (30, \ 29)$, $m = 135$, benchmark value to remove large edges is $0.099$).
 (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/VqqWuE0Jj6A>).

### 2024-07-21 paper.tex
Update MNIST PaCMAP Figure title

@@
 three views of the fit in projections from \(9\text{-}D\), for the
-PBMC3k data (\((s_1, \ s_2) = (-0.050, \ 0.041)\),
+PBMC3k data (\((s_1, \ s_2) = (-0.050, \ -0.041)\),
 \(b = 870 \  (30, \ 29)\), \(m = 135\), benchmark value to remove large

### 2024-07-21 paper.qmd
Added new data with new S-curve data generation

@@
 ```{r}
-umap_scurve <- read_rds(file = "data/s_curve/s_curve_umap.rds")
+umap_scurve <- read_rds(file = "data/s_curve/new/s_curve_umap.rds")
 
@@
 The model closely fits the shape.
-(The **langevitour** software is used to view the data with a tour, and the full video is available at (<https://youtu.be/tYATMiSlCXQ>).
+(The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/tYATMiSlCXQ>).
 :::
@@
 ```{r}
-training_data_scurve <- read_rds("data/s_curve/s_curve_training.rds")
+training_data_scurve <- read_rds("data/s_curve/scurve_500.rds")
 ```

### 2024-07-21 scripts/gen_scurve_data.R

@@ -66,6 +66,10 @@ points$x5 <- runif(sample_size, -0.02, 0.02)
 points$x6 <- runif(sample_size, -0.01, 0.01)
 points$x7 <- runif(sample_size, -0.01, 0.01)
 
+points <- points |> mutate(ID = row_number())
+
+write_rds(points, "data/s_curve/scurve_500.rds")
+
 # # Add the ID to the s_curve_noise
 # s_curve_noise <- points |>
 #   dplyr::mutate(ID = dplyr::row_number()) |>

### 2024-07-21 paper.qmd
Commet unnecessary code chunks

@@
 <!--langevitour with pD model for S-curve-->
-::: {#fig-scurve-sc layout-ncol="4" fig-pos="H"}
-![](figures/scurve/sc_best_trimesh.png)
-
-![](figures/scurve/sc_1.png)
-
-![](figures/scurve/sc_2.png)
-
-![](figures/scurve/sc_3.png)
-
-XXX Model in \gD{}, on the UMAP layout, and three views of the fit in projections from $7\text{-}D$, for the S-curve data.
+<!-- ::: {#fig-scurve-sc layout-ncol="4" fig-pos="H"} -->
+<!-- ![](figures/scurve/sc_best_trimesh.png) -->
+
+<!-- ![](figures/scurve/sc_1.png) -->
+
+<!-- ![](figures/scurve/sc_2.png) -->
+
+<!-- ![](figures/scurve/sc_3.png) -->
+
+<!-- XXX Model in \gD{}, on the UMAP layout, and three views of the fit in projections from $7\text{-}D$, for the S-curve data.
 The model closely fits the shape, but it has breaks which means that it does not adequately capture the surface.
 (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/N6iWEj3aVps>).
-:::
+-->
+<!-- ::: -->
 
@@
 #| fig-height: 3
+#| eval: false
 
@@
 
-@fig-scurve-sc-best shows one of the best fits for S-curve data.
+<!-- @fig-scurve-sc-best shows one of the best fits for S-curve data.
 TriMAP with $15$ nearest neighbors for forming the nearest neighbor triplet (`n_inliers`), $3$ outliers for forming the nearest neighbor triplets (`n_outliers`), $3$ random triplets per point (`n_random`) is used as the NLDR method.
@@
 Edges with lengths greater than $0.266$ are removed.
+-->
 
 <!--langevitour with pD model for S-curve--> 
-::: {#fig-scurve-sc-best layout-ncol="4" fig-pos="H"}
-![](figures/scurve/sc_best_trimap_trimesh.png)
-
-![](figures/scurve/sc_trimap_best_1.png)
-
-![](figures/scurve/sc_trimap_best_2.png)
-
-![](figures/scurve/sc_trimap_best_3.png)
-
-Model in \gD{}, on the TriMAP layout, and three views of the fit in projections from $7\text{-}D$, for the S-curve data.
+<!-- ::: {#fig-scurve-sc-best layout-ncol="4" fig-pos="H"} -->
+<!-- ![](figures/scurve/sc_best_trimap_trimesh.png) -->
+
+<!-- ![](figures/scurve/sc_trimap_best_1.png) -->
+
+<!-- ![](figures/scurve/sc_trimap_best_2.png) -->
+
+<!-- ![](figures/scurve/sc_trimap_best_3.png) -->
+
+<!-- Model in \gD{}, on the TriMAP layout, and three views of the fit in projections from $7\text{-}D$, for the S-curve data.
 The model closely fits the shape, but it has twists.
 (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/zEz2MUCtADM>).
-:::
+-->
+<!-- ::: -->
 

### 2024-07-21 paper.tex
Added new data with new S-curve data generation
Commet unnecessary code chunks

@@
 data with a tour, and the full video is available at
-(\url{https://youtu.be/tYATMiSlCXQ}).}
+\url{https://youtu.be/tYATMiSlCXQ}).}
 
@@
 \kD{} bin centroids.
-
-\begin{figure}[H]
-
-\begin{minipage}{0.25\linewidth}
-\includegraphics{figures/scurve/sc_best_trimesh.png}\end{minipage}%
-%
-\begin{minipage}{0.25\linewidth}
-\includegraphics{figures/scurve/sc_1.png}\end{minipage}%
-%
-\begin{minipage}{0.25\linewidth}
-\includegraphics{figures/scurve/sc_2.png}\end{minipage}%
-%
-\begin{minipage}{0.25\linewidth}
-\includegraphics{figures/scurve/sc_3.png}\end{minipage}%
-
-\caption{\label{fig-scurve-sc}XXX Model in \gD{}, on the UMAP layout,
-and three views of the fit in projections from \(7\text{-}D\), for the
-S-curve data. The model closely fits the shape, but it has breaks which
-means that it does not adequately capture the surface. (The
-\textbf{langevitour} software is used to view the data with a tour, and
-the full video is available at \url{https://youtu.be/N6iWEj3aVps}).}
-
-\end{figure}%
 
@@
 
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-lg-scurve-1.pdf}
-
-}
-
-\caption{\label{fig-lg-scurve}\(2\text{-}D\) model generated for
-different benchmark values to remove long edges: (a) benchmark =
-\(0.201\) (default), (b) benchmark = \(0.35\), and (c) benchmark =
-\(1.1\). What is the effective benchmark value to remove long edges?}
-
-\end{figure}%
-
 \section{Best fit}\label{best-fit}
@@
 data well. However, it technically might be quite variable.
-
-Figure~\ref{fig-scurve-sc-best} shows one of the best fits for S-curve
-data. TriMAP with \(15\) nearest neighbors for forming the nearest
-neighbor triplet (\texttt{n\_inliers}), \(3\) outliers for forming the
-nearest neighbor triplets (\texttt{n\_outliers}), \(3\) random triplets
-per point (\texttt{n\_random}) is used as the NLDR method. The model is
-generated with \(103\) bins, which have a width of \(0.154\) and a
-height of \(0.133\). Edges with lengths greater than \(0.266\) are
-removed.
-
-\begin{figure}[H]
-
-\begin{minipage}{0.25\linewidth}
-\includegraphics{figures/scurve/sc_best_trimap_trimesh.png}\end{minipage}%
-%
-\begin{minipage}{0.25\linewidth}
-\includegraphics{figures/scurve/sc_trimap_best_1.png}\end{minipage}%
-%
-\begin{minipage}{0.25\linewidth}
-\includegraphics{figures/scurve/sc_trimap_best_2.png}\end{minipage}%
-%
-\begin{minipage}{0.25\linewidth}
-\includegraphics{figures/scurve/sc_trimap_best_3.png}\end{minipage}%
-
-\caption{\label{fig-scurve-sc-best}Model in \gD{}, on the TriMAP layout,
-and three views of the fit in projections from \(7\text{-}D\), for the
-S-curve data. The model closely fits the shape, but it has twists. (The
-\textbf{langevitour} software is used to view the data with a tour, and
-the full video is available at \url{https://youtu.be/zEz2MUCtADM}).}
-
-\end{figure}%
 

### 2024-07-21 paper.qmd
Update Figure 3

@@
 
-sc_ltr_pos <- c(0.08, 0.96)
+sc_ltr_pos <- c(0.08, 0.95)
 # sc_xlims <- c(-0.5, 1.47)
@@
 sc_xlims <- c(-0.25, 1.25)
-sc_ylims <- c(-0.3, 2)
+sc_ylims <- c(-0.2, 1)
 
@@
 ## Compute hexbin parameters
-num_bins_x_scurve <- 7
+num_bins_x_scurve <- 11
 
@@
   r2 = r2,
-  q = 0.1)
+  q = 0.05)
 
@@
              aes(x = c_x, y = c_y), 
-             color = "#33a02c") +
+             color = "#33a02c",
+             size = 0.5) +
   xlim(sc_xlims) + ylim(sc_ylims) +

### 2024-07-21 paper.qmd
The best fit with tSNE

@@
 
-sc_ltr_pos <- c(0.08, 0.95)
+sc_ltr_pos <- c(0.08, 0.93)
 # sc_xlims <- c(-0.5, 1.47)
@@
              color = "#33a02c",
-             size = 0.5) +
+             size = 1) +
   xlim(sc_xlims) + ylim(sc_ylims) +

### 2024-07-21 paper.qmd
Update Figure 6 properties

@@
   interior_annotation("a", position = c(0.92, 0.93)) +
-  xlim(c(-0.25, 1.35)) + ylim(c(-0.35, 2))
+  xlim(c(-0.25, 1.35)) + ylim(c(-0.25, 1.1))
 ```
@@
     na.value = "#ffffff", option = "C") +
-  xlim(c(-0.25, 1.35)) + ylim(c(-0.35, 2)) +
+  xlim(c(-0.25, 1.35)) + ylim(c(-0.25, 1.1)) +
   interior_annotation("b")
@@
   scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  xlim(c(-0.25, 1.35)) + ylim(c(-0.35, 2)) +
+  xlim(c(-0.25, 1.35)) + ylim(c(-0.25, 1.1)) +
   interior_annotation("c")
@@
 Using the MSE of the model fit in $p\\text{-}D$ helps decide on a useful choice of number of bins."
-#| fig-width: 6
+#| fig-width: 9
 #| fig-height: 3

### 2024-07-22 paper.qmd
Correct the simulated data with method explanation

--- 2024-07-21 paper.qmd
@@
 ```{r}
-umap_scurve <- read_rds(file = "data/s_curve/new/s_curve_umap.rds")
+umap_scurve <- read_rds(file = "data/s_curve/s_curve_umap.rds")
 
@@
 sc_xlims <- c(-0.25, 1.25)
-sc_ylims <- c(-0.2, 1)
+sc_ylims <- c(-0.2, 2)
 
@@
 ## Compute hexbin parameters
-num_bins_x_scurve <- 11
+num_bins_x_scurve <- 7
 
@@
   r2 = r2,
-  q = 0.05)
+  q = 0.1)
 
@@
   interior_annotation("a", position = c(0.92, 0.93)) +
-  xlim(c(-0.25, 1.35)) + ylim(c(-0.35, 2))
+  xlim(c(-0.25, 1.35)) + ylim(c(-0.25, 2))
 ```
@@
     na.value = "#ffffff", option = "C") +
-  xlim(c(-0.25, 1.35)) + ylim(c(-0.35, 2)) +
+  xlim(c(-0.25, 1.35)) + ylim(c(-0.25, 2)) +
   interior_annotation("b")
@@
   scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  xlim(c(-0.25, 1.35)) + ylim(c(-0.35, 2)) +
+  xlim(c(-0.25, 1.35)) + ylim(c(-0.25, 2)) +
   interior_annotation("c")
@@
 ```{r}
-training_data_scurve <- read_rds("data/s_curve/scurve_500.rds")
+training_data_scurve <- read_rds("data/s_curve/s_curve_training.rds")
 ```

### 2024-07-22 paper.qmd
Update true model screenshots and youtube link

@@
 abstract: |
-  Nonlinear dimension reduction (NLDR) techniques such as tSNE, and UMAP provide a low-dimensional representation of high-dimensional (high-D) data using non-linear transformation.
+  Non-linear dimension reduction (NLDR) techniques such as tSNE, and UMAP provide a low-dimensional representation of high-dimensional (high-D) data using non-linear transformation.
 The methods and parameter choices can create wildly different representations, making it difficult to decide which is best, or whether any or all are accurate or misleading.
@@
 
-keywords: [high-dimensional data, dimension reduction, hexagon binning, low-dimensional manifold, tour, data vizualization, model in the data space]
-keywords-formatted: [high-dimensional data, dimension reduction, hexagon binning, low-dimensional manifold, tour, data vizualization, model in the data space]
+keywords: [high-dimensional data, dimension reduction, hexagon binning, low-dimensional representation, tour, data vizualization, model-in-the-data space]
+keywords-formatted: [high-dimensional data, dimension reduction, hexagon binning, low-dimensional representation, tour, data vizualization, model-in-the-data space]
 
@@
 The model closely fits the shape.
-(The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/tYATMiSlCXQ>).
+(The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/I5GL23vLiw0>).
 :::
 
-XXX The points on the "theoretical model" should be equidistant along the S. They are still to bunched in some spots and too far apart in others.
+<!-- XXX The points on the "theoretical model" should be equidistant along the S. They are still to bunched in some spots and too far apart in others. -->
 

### 2024-07-22 paper.tex
Update true model screenshots and youtube link

@@
   pdfkeywords={high-dimensional data, dimension reduction, hexagon
-binning, low-dimensional manifold, tour, data vizualization, model in
-the data space},
+binning, low-dimensional representation, tour, data
+vizualization, model-in-the-data space},
   colorlinks=true,
@@
 \begin{abstract}
-Nonlinear dimension reduction (NLDR) techniques such as tSNE, and UMAP
+Non-linear dimension reduction (NLDR) techniques such as tSNE, and UMAP
 provide a low-dimensional representation of high-dimensional (high-D)
@@
 {\it Keywords:} high-dimensional data, dimension reduction, hexagon
-binning, low-dimensional manifold, tour, data vizualization, model in
-the data space
+binning, low-dimensional representation, tour, data
+vizualization, model-in-the-data space
 \vfill
@@
 data with a tour, and the full video is available at
-\url{https://youtu.be/tYATMiSlCXQ}).}
-
-\end{figure}%
-
-XXX The points on the ``theoretical model'' should be equidistant along
-the S. They are still to bunched in some spots and too far apart in
-others.
+\url{https://youtu.be/I5GL23vLiw0}).}
+
+\end{figure}%
 

### 2024-07-22 paper.qmd
Fix the issue of visualizing the model, true model overlaid with data

@@
 abstract: |
-  Non-linear dimension reduction (NLDR) techniques such as tSNE, and UMAP provide a low-dimensional representation of high-dimensional (high-D) data using non-linear transformation.
+  Non-linear dimension reduction (NLDR) techniques such as tSNE, and UMAP provide a low-dimensional representation of high-dimensional ($p\text{-}D$) data using non-linear transformation.
 The methods and parameter choices can create wildly different representations, making it difficult to decide which is best, or whether any or all are accurate or misleading.
 NLDR often exaggerates random patterns, sometimes due to the samples observed.
-But NLDR views have an important role in data analysis because, if done well, they provide a concise visual (and conceptual) summary of high-D distributions.
-To help evaluate the NLDR we have developed an algorithm to show the 2D NLDR model in the high-D space, viewed with a tour.
+But NLDR views have an important role in data analysis because, if done well, they provide a concise visual (and conceptual) summary of $p\text{-}D$ distributions.
+To help evaluate the NLDR we have developed an algorithm to show the $2\text{-}D$ NLDR model in the $p\text{-}D$ space, viewed with a tour.
 One can see if the model fits everywhere or better in some subspaces, or completely mismatches the data.
-It is used to evaluate which 2D layout is the best representation of the high-D distribution and see how different methods may have similar summaries or quirks.
-
-keywords: [high-dimensional data, dimension reduction, hexagon binning, low-dimensional representation, tour, data vizualization, model-in-the-data space]
-keywords-formatted: [high-dimensional data, dimension reduction, hexagon binning, low-dimensional representation, tour, data vizualization, model-in-the-data space]
+It is used to evaluate which $2\text{-}D$ layout is the best representation of the $p\text{-}D$ distribution and see how different methods may have similar summaries or quirks.
+
+keywords: [high-dimensional data, dimension reduction, hexagon binning, low-dimensional representation, tour, data vizualization, model-in-the-data-space]
+keywords-formatted: [high-dimensional data, dimension reduction, hexagon binning, low-dimensional representation, tour, data vizualization, model-in-the-data-space]
 
@@
 
+XXXAdd best fit with UMAP and discuss why it's the best fit
+
 <!--langevitour with pD model for S-curve-->
@@
 
-XXX You need to have consistent colouring. In langevitour the points corresponding to observations are purple. They should be purple in the \gD{} view too.
+<!-- XXX You need to have consistent colouring. In langevitour the points corresponding to observations are purple. They should be purple in the \gD{} view too.  -->
 
@@
 However, it technically might be quite variable.
+
+XXXAdd best fit and discuss why it's the best fit
 

### 2024-07-22 paper.tex
Fix the issue of visualizing the model, true model overlaid with data

@@
 binning, low-dimensional representation, tour, data
-vizualization, model-in-the-data space},
+vizualization, model-in-the-data-space},
   colorlinks=true,
@@
 Non-linear dimension reduction (NLDR) techniques such as tSNE, and UMAP
-provide a low-dimensional representation of high-dimensional (high-D)
-data using non-linear transformation. The methods and parameter choices
-can create wildly different representations, making it difficult to
-decide which is best, or whether any or all are accurate or misleading.
-NLDR often exaggerates random patterns, sometimes due to the samples
-observed. But NLDR views have an important role in data analysis
+provide a low-dimensional representation of high-dimensional
+(\(p\text{-}D\)) data using non-linear transformation. The methods and
+parameter choices can create wildly different representations, making it
+difficult to decide which is best, or whether any or all are accurate or
+misleading. NLDR often exaggerates random patterns, sometimes due to the
+samples observed. But NLDR views have an important role in data analysis
 because, if done well, they provide a concise visual (and conceptual)
-summary of high-D distributions. To help evaluate the NLDR we have
-developed an algorithm to show the 2D NLDR model in the high-D space,
-viewed with a tour. One can see if the model fits everywhere or better
-in some subspaces, or completely mismatches the data. It is used to
-evaluate which 2D layout is the best representation of the high-D
-distribution and see how different methods may have similar summaries or
-quirks.
+summary of \(p\text{-}D\) distributions. To help evaluate the NLDR we
+have developed an algorithm to show the \(2\text{-}D\) NLDR model in the
+\(p\text{-}D\) space, viewed with a tour. One can see if the model fits
+everywhere or better in some subspaces, or completely mismatches the
+data. It is used to evaluate which \(2\text{-}D\) layout is the best
+representation of the \(p\text{-}D\) distribution and see how different
+methods may have similar summaries or quirks.
 \end{abstract}
@@
 binning, low-dimensional representation, tour, data
-vizualization, model-in-the-data space
+vizualization, model-in-the-data-space
 \vfill
@@
 
-XXX You need to have consistent colouring. In langevitour the points
-corresponding to observations are purple. They should be purple in the
-\gD{} view too.
+XXXAdd best fit with UMAP and discuss why it's the best fit
 
@@
 
+XXXAdd best fit and discuss why it's the best fit
+
 \section{A curious difference between t-SNE, UMAP and PaCMAP

### 2024-07-22 paper.qmd
Added best model with tSNE, updated video link

@@
 
-<!--langevitour with pD model for S-curve-->
-<!-- ::: {#fig-scurve-sc layout-ncol="4" fig-pos="H"} -->
-<!-- ![](figures/scurve/sc_best_trimesh.png) -->
-
-<!-- ![](figures/scurve/sc_1.png) -->
-
-<!-- ![](figures/scurve/sc_2.png) -->
-
-<!-- ![](figures/scurve/sc_3.png) -->
-
-<!-- XXX Model in \gD{}, on the UMAP layout, and three views of the fit in projections from $7\text{-}D$, for the S-curve data.
-The model closely fits the shape, but it has breaks which means that it does not adequately capture the surface.
-(The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/N6iWEj3aVps>).
--->
-<!-- ::: -->
+<!-- langevitour with pD model for S-curve -->
+::: {#fig-scurve-sc layout-ncol="4" fig-pos="H"}
+![](figures/scurve/umap_trimesh_layout.png)
+
+![](figures/scurve/sc_1.png)
+
+![](figures/scurve/sc_2.png)
+
+![](figures/scurve/sc_3.png)
+
+Model in \gD{}, on the UMAP layout, and three views of the fit in projections from $7\text{-}D$, for the S-curve data ($(s_1, \ s_2) = (-0.160, \ -0.263)$, $b = 405 \  (15, \ 27)$, $m = 70$, benchmark value to remove low density hexagons is $0.250$, and benchmark value to remove large edges is $0.159$).
+MSE is $0.0462$.
+The model closely fits the shape, but it doesn't fully fill out the width of the S, which means that it does not adequately capture the surface (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/N6iWEj3aVps>).
+:::
 
@@
 
-XXXAdd best fit and discuss why it's the best fit
-
 <!-- @fig-scurve-sc-best shows one of the best fits for S-curve data.
@@
 
-<!--langevitour with pD model for S-curve--> 
-<!-- ::: {#fig-scurve-sc-best layout-ncol="4" fig-pos="H"} -->
-<!-- ![](figures/scurve/sc_best_trimap_trimesh.png) -->
-
-<!-- ![](figures/scurve/sc_trimap_best_1.png) -->
-
-<!-- ![](figures/scurve/sc_trimap_best_2.png) -->
-
-<!-- ![](figures/scurve/sc_trimap_best_3.png) -->
-
-<!-- Model in \gD{}, on the TriMAP layout, and three views of the fit in projections from $7\text{-}D$, for the S-curve data.
-The model closely fits the shape, but it has twists.
-(The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/zEz2MUCtADM>).
--->
-<!-- ::: -->
+<!-- langevitour with pD model for S-curve -->
+::: {#fig-scurve-sc-best layout-ncol="4" fig-pos="H"}
+![](figures/scurve/tsne_trimesh_layout.png)
+
+![](figures/scurve/sc_tsne_best_1.png)
+
+![](figures/scurve/sc_tsne_best_2.png)
+
+![](figures/scurve/sc_tsne_best_3.png)
+
+Model in \gD{}, on the tSNE layout, and three views of the fit in projections from $7\text{-}D$, for the S-curve data ($(s_1, \ s_2) = (-0.170, \ -0.137)$, $b = 256 \  (16, \ 16)$, $m = 70$, benchmark value to remove low density hexagons is $0.250$, and benchmark value to remove large edges is $0.180$).
+MSE is $0.0432$.
+The model closely fits the shape, but it has some twists (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/ZuLdp89qJ6g>).
+:::
 

### 2024-07-22 paper.tex
Added best model with tSNE, updated video link

@@
 
+\begin{figure}[H]
+
+\begin{minipage}{0.25\linewidth}
+\includegraphics{figures/scurve/umap_trimesh_layout.png}\end{minipage}%
+%
+\begin{minipage}{0.25\linewidth}
+\includegraphics{figures/scurve/sc_1.png}\end{minipage}%
+%
+\begin{minipage}{0.25\linewidth}
+\includegraphics{figures/scurve/sc_2.png}\end{minipage}%
+%
+\begin{minipage}{0.25\linewidth}
+\includegraphics{figures/scurve/sc_3.png}\end{minipage}%
+
+\caption{\label{fig-scurve-sc}Model in \gD{}, on the UMAP layout, and
+three views of the fit in projections from \(7\text{-}D\), for the
+S-curve data (\((s_1, \ s_2) = (-0.160, \ -0.263)\),
+\(b = 405 \  (15, \ 27)\), \(m = 70\), benchmark value to remove low
+density hexagons is \(0.250\), and benchmark value to remove large edges
+is \(0.159\)). MSE is \(0.0462\). The model closely fits the shape, but
+it doesn't fully fill out the width of the S, which means that it does
+not adequately capture the surface (The \textbf{langevitour} software is
+used to view the data with a tour, and the full video is available at
+\url{https://youtu.be/N6iWEj3aVps}).}
+
+\end{figure}%
+
 XXX Should there be a second long edge removal? You can compute the
@@
 
-XXXAdd best fit and discuss why it's the best fit
+\begin{figure}[H]
+
+\begin{minipage}{0.25\linewidth}
+\includegraphics{figures/scurve/tsne_trimesh_layout.png}\end{minipage}%
+%
+\begin{minipage}{0.25\linewidth}
+\includegraphics{figures/scurve/sc_tsne_best_1.png}\end{minipage}%
+%
+\begin{minipage}{0.25\linewidth}
+\includegraphics{figures/scurve/sc_tsne_best_2.png}\end{minipage}%
+%
+\begin{minipage}{0.25\linewidth}
+\includegraphics{figures/scurve/sc_tsne_best_3.png}\end{minipage}%
+
+\caption{\label{fig-scurve-sc-best}Model in \gD{}, on the tSNE layout,
+and three views of the fit in projections from \(7\text{-}D\), for the
+S-curve data (\((s_1, \ s_2) = (-0.170, \ -0.137)\),
+\(b = 256 \  (16, \ 16)\), \(m = 70\), benchmark value to remove low
+density hexagons is \(0.250\), and benchmark value to remove large edges
+is \(0.180\)). MSE is \(0.0432\). The model closely fits the shape, but
+it has some twists (The \textbf{langevitour} software is used to view
+the data with a tour, and the full video is available at
+\url{https://youtu.be/ZuLdp89qJ6g}).}
+
+\end{figure}%
 

### 2024-07-22 paper.qmd
Find the best model with UMAP and update the video link

@@
 
-![](figures/scurve/sc_1.png)
-
-![](figures/scurve/sc_2.png)
-
-![](figures/scurve/sc_3.png)
+![](figures/scurve/sc_umap_best_1.png)
+
+![](figures/scurve/sc_umap_best_2.png)
+
+![](figures/scurve/sc_umap_best_3.png)
 
@@
 MSE is $0.0462$.
-The model closely fits the shape, but it doesn't fully fill out the width of the S, which means that it does not adequately capture the surface (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/N6iWEj3aVps>).
+The model closely fits the shape, but it doesn't fully fill out the width of the S, which means that it does not adequately capture the surface (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/i7F7xpN1Hz8>).
 :::

### 2024-07-22 paper.qmd
Added second edge removal description

@@
 Furthermore, line segments that exist in the \kD{} model generate line segments in \pD{} by connecting the \pD{} means of the corresponding \kD{} bin centroids.
-
-XXXAdd best fit with UMAP and discuss why it's the best fit
+If additional long edges need to be removed, compute the edges in \pD{} and pruned any detected long edges to improve the accuracy.
+Once pruned, re-plot the \gD{} view to ensure it accurately captures the data.
 
@@
 
-XXX Should there be a second long edge removal? You can compute the edges in \pD{} now, and if there are long edges maybe these should be pruned, and we re-plot the \gD{} view.
+<!-- XXX Should there be a second long edge removal? You can compute the edges in \pD{} now, and if there are long edges maybe these should be pruned, and we re-plot the \gD{} view.  -->
 

### 2024-07-22 paper.tex
Find the best model with UMAP and update the video link
Added second edge removal description

@@
 segments in \pD{} by connecting the \pD{} means of the corresponding
-\kD{} bin centroids.
-
-XXXAdd best fit with UMAP and discuss why it's the best fit
+\kD{} bin centroids. If additional long edges need to be removed,
+compute the edges in \pD{} and pruned any detected long edges to improve
+the accuracy. Once pruned, re-plot the \gD{} view to ensure it
+accurately captures the data.
 
@@
 \begin{minipage}{0.25\linewidth}
-\includegraphics{figures/scurve/sc_1.png}\end{minipage}%
-%
-\begin{minipage}{0.25\linewidth}
-\includegraphics{figures/scurve/sc_2.png}\end{minipage}%
-%
-\begin{minipage}{0.25\linewidth}
-\includegraphics{figures/scurve/sc_3.png}\end{minipage}%
+\includegraphics{figures/scurve/sc_umap_best_1.png}\end{minipage}%
+%
+\begin{minipage}{0.25\linewidth}
+\includegraphics{figures/scurve/sc_umap_best_2.png}\end{minipage}%
+%
+\begin{minipage}{0.25\linewidth}
+\includegraphics{figures/scurve/sc_umap_best_3.png}\end{minipage}%
 
@@
 used to view the data with a tour, and the full video is available at
-\url{https://youtu.be/N6iWEj3aVps}).}
-
-\end{figure}%
-
-XXX Should there be a second long edge removal? You can compute the
-edges in \pD{} now, and if there are long edges maybe these should be
-pruned, and we re-plot the \gD{} view.
+\url{https://youtu.be/i7F7xpN1Hz8}).}
+
+\end{figure}%
 

### 2024-07-22 paper.qmd
Resize Figure 8

@@
                          group = buff)) + 
-  geom_point(size = 0.05) +
+  geom_point(size = 1) +
   geom_line(linewidth = 0.05) + 
@@
   theme(aspect.ratio = 0.75,
-        axis.text.x = element_text(size = 7),
-        axis.text.y = element_text(size = 7),
-        axis.title.x = element_text(size = 7),
-        axis.title.y = element_text(size = 7, 
+        axis.text.x = element_text(size = 12),
+        axis.text.y = element_text(size = 12),
+        axis.title.x = element_text(size = 12),
+        axis.title.y = element_text(size = 12, 
             angle=90),
-        title = element_text(size = 72),
-        legend.title = element_text(size = 7),
-        legend.text = element_text(size = 5))
+        plot.title = element_text(size = 12),
+        legend.title = element_text(size = 12),
+        legend.text = element_text(size = 10),
+        legend.key.height = unit(1, 'cm'),
+        legend.key.width = unit(1, 'cm'),
+        legend.key.size = unit(1, 'cm'))
 ```
@@
   geom_quasirandom(
-    size = 0.3,
+    size = 1,
     alpha = 0.5
@@
   theme(aspect.ratio = 0.75,
-        axis.title.x = element_text(size = 5),
-        axis.title.y = element_text(size = 5, angle = 90),
+        axis.title.x = element_text(size = 12),
+        axis.title.y = element_text(size = 12, angle = 90),
         axis.text.x = element_blank(),
-        axis.text.y = element_text(size = 5),
-        title = element_text(size = 36))
+        axis.text.y = element_text(size = 12),
+        plot.title = element_text(size = 12))
 ```
@@
 The benchmark value for removing low-density hexagons ranges between $0$ and $1$.
-When analyzing how these benchmark values influence model performance, it's essential to observe the change in MSE as the benchmark value increases (see @fig-mse-scurve-lwd).
-The MSE shows a gradual increase as the benchmark value progresses from $0$ to $1$.
+When analyzing how these benchmark values influence model performance, it's essential to observe the change in MSE as the benchmark value increases (@fig-param-scurve).
+The MSE shows a gradual decrease as the benchmark value goes from $1$ to $0$.
 Evaluating this rate of increase is important.
@@
   geom_point(
-    size = 0.6
+    size = 1
     ) +
@@
   theme(aspect.ratio = 0.75,
-        axis.text.x = element_text(size = 5),
-        axis.text.y = element_text(size = 5),
-        axis.title.x = element_text(size = 5),
-        axis.title.y = element_text(size = 5, 
+        axis.text.x = element_text(size = 12),
+        axis.text.y = element_text(size = 12),
+        axis.title.x = element_text(size = 12),
+        axis.title.y = element_text(size = 12, 
             angle=90),
-        title = element_text(size = 72))
-```
-
-Furthermore, selecting the benchmark value for removing low-density hexagons is important.
+        plot.title = element_text(size = 12))
+```
+
+<!-- Furthermore, selecting the benchmark value for removing low-density hexagons is important.
 Removing unnecessary bins may lead to the formation of long edges and an uneven \gD{} model.
@@
 The remaining bins are used to construct the \gD{} model.
+-->
 
@@
 
-1. Plot the distribution of the 2D Euclidean distances (@fig-param-scurve (d)).
+1. Plot the distribution of the \gD{} Euclidean distances (@fig-param-scurve (d)).
 2. Identify the first largest difference between consecutive distance values.
@@
   geom_quasirandom(
-    size = 0.5, 
+    size = 2, 
     alpha = 0.3
@@
   theme(aspect.ratio = 0.75,
-        axis.text.x = element_text(size = 5),
-        axis.title.x = element_text(size = 5),
-        title = element_text(size = 36))
+        axis.text.x = element_text(size = 12),
+        axis.title.x = element_text(size = 12),
+        plot.title = element_text(size = 12))
 ```
@@
 
-<!--trimesh with different benchmark values-->
-
-```{r}
-## With calculated default: 0.201
-trimesh_lg_scurve1 <- vis_rmlg_mesh(
-  distance_edges = distance_scurve, 
-  benchmark_value = benchmark_scurve, 
-  tr_coord_df = tr_from_to_df_scurve, 
-  distance_col = "distance") +
-  xlim(sc_xlims) + ylim(sc_ylims) +
-  interior_annotation("a", sc_ltr_pos)
-
-trimesh_lg_scurve2 <- vis_rmlg_mesh(
-  distance_edges = distance_scurve, 
-  benchmark_value = 0.35, 
-  tr_coord_df = tr_from_to_df_scurve, 
-  distance_col = "distance") +
-  xlim(sc_xlims) + ylim(sc_ylims) +
-  interior_annotation("b", sc_ltr_pos)
-
-trimesh_lg_scurve3 <- vis_rmlg_mesh(
-  distance_edges = distance_scurve, 
-  benchmark_value = 1.1, 
-  tr_coord_df = tr_from_to_df_scurve, 
-  distance_col = "distance") +
-  xlim(sc_xlims) + ylim(sc_ylims) +
-  interior_annotation("c", sc_ltr_pos) 
-```
-
-```{r}
-#| echo: false
-#| label: fig-lg-scurve
-#| fig-pos: H
-#| fig-cap: "$2\\text{-}D$ model generated for different benchmark values to remove long edges: (a) benchmark = $0.201$ (default), (b) benchmark = $0.35$, and (c) benchmark = $1.1$.
-What is the effective benchmark value to remove long edges?"
-#| fig-width: 6
-#| fig-height: 3
-#| eval: false
-
-trimesh_lg_scurve1 + trimesh_lg_scurve2 + trimesh_lg_scurve3 +
-  plot_layout(guides='collect', ncol = 3) &
-  theme(legend.position='none', plot.tag = element_text(size = 8))
-```
-
 ## Best fit
@@
 
-## A curious difference between t-SNE, UMAP and PaCMAP revealer
+## A curious difference between tSNE, UMAP and PaCMAP revealer
 
@@
 The model fits the separation and tries to *filled out* the clusters.
-(The **langevitour** software is used to view the data with a tour, and the full video is available at (<https://youtu.be/RASEE7N5MbM>).
+(The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/RASEE7N5MbM>).
 :::
@@
 The model fits the separation and tries to *filled out* the clusters, but not as much as tSNE.
-(The **langevitour** software is used to view the data with a tour, and the full video is available at (<https://youtu.be/iG4bCPkJilw>).
+(The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/iG4bCPkJilw>).
 :::
@@
 The model fits the separation and shows *flat* shaped clusters.
-(The **langevitour** software is used to view the data with a tour, and the full video is available at (<https://youtu.be/z07cKXi8EJQ>).
+(The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/z07cKXi8EJQ>).
 :::

### 2024-07-22 paper.tex
Resize Figure 8

@@
 model performance, it's essential to observe the change in MSE as the
-benchmark value increases (see \textbf{?@fig-mse-scurve-lwd}). The MSE
-shows a gradual increase as the benchmark value progresses from \(0\) to
-\(1\). Evaluating this rate of increase is important. If the increment
-is not considerable, the decision might lean towards retaining
-low-density hexagons.
-
-Furthermore, selecting the benchmark value for removing low-density
-hexagons is important. Removing unnecessary bins may lead to the
-formation of long edges and an uneven \gD{} model. Hence, rather than
-solely relying on the benchmark value to identify hexagons for removal,
-it's essential to consider the standard number of points in the
-neighboring hexagons of the identified low-density bins (see
-\textbf{?@fig-lwd-scurve} (b)). If neighboring bins also show low
-counts, only those bins will be removed. The remaining bins are used to
-construct the \gD{} model.
+benchmark value increases (Figure~\ref{fig-param-scurve}). The MSE shows
+a gradual decrease as the benchmark value goes from \(1\) to \(0\).
+Evaluating this rate of increase is important. If the increment is not
+considerable, the decision might lean towards retaining low-density
+hexagons.
 
@@
 \item
-  Plot the distribution of the 2D Euclidean distances
+  Plot the distribution of the \gD{} Euclidean distances
   (Figure~\ref{fig-param-scurve} (d)).
@@
 
-\section{A curious difference between t-SNE, UMAP and PaCMAP
-revealer}\label{a-curious-difference-between-t-sne-umap-and-pacmap-revealer}
+\section{A curious difference between tSNE, UMAP and PaCMAP
+revealer}\label{a-curious-difference-between-tsne-umap-and-pacmap-revealer}
 
@@
 used to view the data with a tour, and the full video is available at
-(\url{https://youtu.be/RASEE7N5MbM}).}
+\url{https://youtu.be/RASEE7N5MbM}).}
 
@@
 \textbf{langevitour} software is used to view the data with a tour, and
-the full video is available at (\url{https://youtu.be/iG4bCPkJilw}).}
+the full video is available at \url{https://youtu.be/iG4bCPkJilw}).}
 
@@
 to view the data with a tour, and the full video is available at
-(\url{https://youtu.be/z07cKXi8EJQ}).}
+\url{https://youtu.be/z07cKXi8EJQ}).}
 

### 2024-07-22 paper.qmd
Update Figure 6
paper.qmd

--- 2024-07-22 paper.qmd
@@
 abstract: |
-  Non-linear dimension reduction (NLDR) techniques such as tSNE, and UMAP provide a low-dimensional representation of high-dimensional ($p\text{-}D$) data using non-linear transformation.
+  Nonlinear dimension reduction (NLDR) techniques such as tSNE, and UMAP provide a low-dimensional representation of high-dimensional (high-D) data using non-linear transformation.
 The methods and parameter choices can create wildly different representations, making it difficult to decide which is best, or whether any or all are accurate or misleading.
 NLDR often exaggerates random patterns, sometimes due to the samples observed.
-But NLDR views have an important role in data analysis because, if done well, they provide a concise visual (and conceptual) summary of $p\text{-}D$ distributions.
-To help evaluate the NLDR we have developed an algorithm to show the $2\text{-}D$ NLDR model in the $p\text{-}D$ space, viewed with a tour.
+But NLDR views have an important role in data analysis because, if done well, they provide a concise visual (and conceptual) summary of high-D distributions.
+To help evaluate the NLDR we have developed an algorithm to show the 2D NLDR model in the high-D space, viewed with a tour.
 One can see if the model fits everywhere or better in some subspaces, or completely mismatches the data.
-It is used to evaluate which $2\text{-}D$ layout is the best representation of the $p\text{-}D$ distribution and see how different methods may have similar summaries or quirks.
-
-keywords: [high-dimensional data, dimension reduction, hexagon binning, low-dimensional representation, tour, data vizualization, model-in-the-data-space]
-keywords-formatted: [high-dimensional data, dimension reduction, hexagon binning, low-dimensional representation, tour, data vizualization, model-in-the-data-space]
+It is used to evaluate which 2D layout is the best representation of the high-D distribution and see how different methods may have similar summaries or quirks.
+
+keywords: [high-dimensional data, dimension reduction, hexagon binning, low-dimensional manifold, tour, data vizualization, model in the data space]
+keywords-formatted: [high-dimensional data, dimension reduction, hexagon binning, low-dimensional manifold, tour, data vizualization, model in the data space]
 
@@
 sc_xlims <- c(-0.25, 1.25)
-sc_ylims <- c(-0.2, 2)
+sc_ylims <- c(-0.2, 1)
 
@@
 The model closely fits the shape.
-(The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/I5GL23vLiw0>).
+(The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/tYATMiSlCXQ>).
 :::
 
-<!-- XXX The points on the "theoretical model" should be equidistant along the S. They are still to bunched in some spots and too far apart in others. -->
+XXX The points on the "theoretical model" should be equidistant along the S. They are still to bunched in some spots and too far apart in others.
 
@@
 
-XXXAdd best fit with UMAP and discuss why it's the best fit
-
-<!-- langevitour with pD model for S-curve -->
-::: {#fig-scurve-sc layout-ncol="4" fig-pos="H"}
-![](figures/scurve/umap_trimesh_layout.png)
-
-![](figures/scurve/sc_umap_best_1.png)
-
-![](figures/scurve/sc_umap_best_2.png)
-
-![](figures/scurve/sc_umap_best_3.png)
-
-Model in \gD{}, on the UMAP layout, and three views of the fit in projections from $7\text{-}D$, for the S-curve data ($(s_1, \ s_2) = (-0.160, \ -0.263)$, $b = 405 \  (15, \ 27)$, $m = 70$, benchmark value to remove low density hexagons is $0.250$, and benchmark value to remove large edges is $0.159$).
-MSE is $0.0462$.
-The model closely fits the shape, but it doesn't fully fill out the width of the S, which means that it does not adequately capture the surface (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/i7F7xpN1Hz8>).
-:::
-
-<!-- XXX You need to have consistent colouring. In langevitour the points corresponding to observations are purple. They should be purple in the \gD{} view too.  -->
+<!--langevitour with pD model for S-curve-->
+<!-- ::: {#fig-scurve-sc layout-ncol="4" fig-pos="H"} -->
+<!-- ![](figures/scurve/sc_best_trimesh.png) -->
+
+<!-- ![](figures/scurve/sc_1.png) -->
+
+<!-- ![](figures/scurve/sc_2.png) -->
+
+<!-- ![](figures/scurve/sc_3.png) -->
+
+<!-- XXX Model in \gD{}, on the UMAP layout, and three views of the fit in projections from $7\text{-}D$, for the S-curve data.
+The model closely fits the shape, but it has breaks which means that it does not adequately capture the surface.
+(The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/N6iWEj3aVps>).
+-->
+<!-- ::: -->
+
+XXX You need to have consistent colouring. In langevitour the points corresponding to observations are purple. They should be purple in the \gD{} view too.
 
@@
   interior_annotation("a", position = c(0.92, 0.93)) +
-  xlim(c(-0.25, 1.35)) + ylim(c(-0.25, 2))
+  xlim(c(-0.25, 1.35)) + ylim(c(-0.25, 1.1))
 ```
@@
     na.value = "#ffffff", option = "C") +
-  xlim(c(-0.25, 1.35)) + ylim(c(-0.25, 2)) +
+  xlim(c(-0.25, 1.35)) + ylim(c(-0.25, 1.1)) +
   interior_annotation("b")
@@
   scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  xlim(c(-0.25, 1.35)) + ylim(c(-0.25, 2)) +
+  xlim(c(-0.25, 1.35)) + ylim(c(-0.25, 1.1)) +
   interior_annotation("c")
@@
 ```{r}
+#| eval: false
 ## To initialize number of bins along the x-axis
@@
 
-<!-- langevitour with pD model for S-curve -->
-::: {#fig-scurve-sc-best layout-ncol="4" fig-pos="H"}
-![](figures/scurve/tsne_trimesh_layout.png)
-
-![](figures/scurve/sc_tsne_best_1.png)
-
-![](figures/scurve/sc_tsne_best_2.png)
-
-![](figures/scurve/sc_tsne_best_3.png)
-
-Model in \gD{}, on the tSNE layout, and three views of the fit in projections from $7\text{-}D$, for the S-curve data ($(s_1, \ s_2) = (-0.170, \ -0.137)$, $b = 256 \  (16, \ 16)$, $m = 70$, benchmark value to remove low density hexagons is $0.250$, and benchmark value to remove large edges is $0.180$).
-MSE is $0.0432$.
-The model closely fits the shape, but it has some twists (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/ZuLdp89qJ6g>).
-:::
+<!--langevitour with pD model for S-curve--> 
+<!-- ::: {#fig-scurve-sc-best layout-ncol="4" fig-pos="H"} -->
+<!-- ![](figures/scurve/sc_best_trimap_trimesh.png) -->
+
+<!-- ![](figures/scurve/sc_trimap_best_1.png) -->
+
+<!-- ![](figures/scurve/sc_trimap_best_2.png) -->
+
+<!-- ![](figures/scurve/sc_trimap_best_3.png) -->
+
+<!-- Model in \gD{}, on the TriMAP layout, and three views of the fit in projections from $7\text{-}D$, for the S-curve data.
+The model closely fits the shape, but it has twists.
+(The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/zEz2MUCtADM>).
+-->
+<!-- ::: -->
 

### 2024-07-22 paper.qmd

--- 2024-07-22 paper.qmd
@@
 
-XXX Prediction should have a jitter option, to spread the predicted points out fully in the hexagons.
+<!-- XXX Prediction should have a jitter option, to spread the predicted points out fully in the hexagons. -->
 

### 2024-07-28 paper.tex
Update Figure 6
removing eval falsse

@@
 
-XXX Prediction should have a jitter option, to spread the predicted
-points out fully in the hexagons.
-
 \subsection{Tuning}\label{tuning}
@@
 
-\includegraphics{figures/mnist/mnist_model_2d.png}
+\captionsetup{labelsep=none}\includegraphics{figures/mnist/mnist_model_2d.png}
 
@@
 
-\includegraphics{figures/mnist/img_error_sample_within.png}
+\captionsetup{labelsep=none}\includegraphics{figures/mnist/img_error_sample_within.png}
 
@@
 
-\includegraphics{figures/mnist/img_error_sample_out.png}
+\captionsetup{labelsep=none}\includegraphics{figures/mnist/img_error_sample_out.png}
 

### 2024-07-28 paper.qmd
caption for fig 2

--- 2024-07-22 paper.qmd
@@
 
-<!-- XXX SHOW THE S-CURVE DATA WITH THE TRUE MODEL OVERLAID HERE, BEFORE THE NLDR VIEWS. -->
-
 <!--langevitour with true model for S-curve-->
@@
 
-Three views of the true model in projections from $7\text{-}D$, for the S-curve data.
-The model closely fits the shape.
+Three views of the true model (grey points and lines) in $2\text{-}D$ projections from $7\text{-}D$, for the S-curve data (purple points).
+The data is spread along the S shape, and does not vary much from this curve.
 (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/I5GL23vLiw0>).

### 2024-07-28 paper.tex
caption for fig 2

@@
 
-\caption{\label{fig-scurve-true-sc}Three views of the true model in
-projections from \(7\text{-}D\), for the S-curve data. The model closely
-fits the shape. (The \textbf{langevitour} software is used to view the
-data with a tour, and the full video is available at
-\url{https://youtu.be/I5GL23vLiw0}).}
+\caption{\label{fig-scurve-true-sc}Three views of the true model (grey
+points and lines) in \(2\text{-}D\) projections from \(7\text{-}D\), for
+the S-curve data (purple points). The data is spread along the S shape,
+and does not vary much from this curve. (The \textbf{langevitour}
+software is used to view the data with a tour, and the full video is
+available at \url{https://youtu.be/I5GL23vLiw0}).}
 

### 2024-07-29 paper.qmd
revising the binning parameters section

@@
 
-The performance and robustness of our model depend on four key parameters: (i) bin start position $(s_1, \ s_2)$, (ii) the total number of bins ($b$), (iii) a benchmark value used to remove low-density hexagons, and (iv) a benchmark value used to remove long edges.
-However, there is no analytical formula to calculate appropriate values for these parameters.
-The selection of these parameter values depends on the model performance computed by MSE (@sec-summary), except for the threshold used to remove long edges.
-
-#### Bin start position and number of bins
-
-The starting position of the hexagonal grid is important because different starting points result in different distributions of data across bins, even with the same total number of bins.
+The model fitting can be adjusted using these parameters:
+
+- hexagon bin parameters
+    - bottom left bin position $(s_1, \ s_2)$, 
+    - the total number of bins ($b$), 
+- bin density cutoff, to remove low-density hexagons, and 
+- edge length maximum, remove long edges from \gD{} representation.
+
+Default values are provided for each of these, but it is expected that the user will examine the MSE for a range of choices.
+Choosing these parameters according to MSE can be automated but it is recommended that the user examine the resulting model representation by overlaying it on the data in \pD{}.
+The next few subsections describe the calculation of default values, and the effect that different choices have on the model fit.
+
+#### Hexagon bin parameters
+
+The values $(s_1, \ s_2)$ define the position of the centroid of the bottom left hexagon.
+By default, this is at $s_1 = -q, s_2 = -qr_2$, where $q$ is the buffer sound the data.
+The choice of these values can have some effect on the distribution of bin counts.
+FIG XXX(a) illustrates this.
+The distribution of bin counts for $s_1$ varying between $-0.1-0.0$ is shown.
+Generally, a more uniform distribution among these possibilities would indicate that the bins are reliably capturing the underlying distribution of observations.
+
+<!--The starting position of the hexagonal grid is important because different starting points result in different distributions of data across bins, even with the same total number of bins.
 This variation affects the model due to the differing number of non-empty bins.
-Therefore, it is necessary to evaluate various starting points with different total numbers of bins to determine which configurations are more effective at capturing the structure and fitting the model.
+Therefore, it is necessary to evaluate various starting points with different total numbers of bins to determine which configurations are more effective at capturing the structure and fitting the model.-->
 
@@
 
-To determine the effective $b$, candidate values are selected based on the range between the minimum and approximate maximum $b_1$, because $b_2$ is computed from $b_1$.
+The default number of bins $b=b_1\times b_2$ is computed based on the sample size, by setting $b_1=n^{1/3}$, consistent with the Diaconis-Freedman rule @REF-doi:10.1007/BF01025868.
+The value of $b_2$ is determined analytically by $b_1, q, r_2$.
+Values of $b_1$ between 2 and $b_1 = \frac{n}{2}$ are allowed.
+FIG XXX(b) shows the effect of different choices of $b_1$ on the MSE of the fitted model.
+
+<!-- To determine the effective $b$, candidate values are selected based on the range between the minimum and approximate maximum $b_1$, because $b_2$ is computed from $b_1$.
 The minimum $b_1$ is set to $2$, while the maximum number is estimated by taking the square root of $\frac{n}{2}$.
-By evaluating MSE across varying $b$ within this range for different $q$, helps to determine an appropriate values for $b$ and $q$ (@fig-param-scurve (a)).
+By evaluating MSE across varying $b$ within this range for different $q$, helps to determine an appropriate values for $b$ and $q$ (@fig-param-scurve (a)).-->
 

### 2024-07-29 paper.tex
revising the binning parameters section

@@
 
-The performance and robustness of our model depend on four key
-parameters: (i) bin start position \((s_1, \ s_2)\), (ii) the total
-number of bins (\(b\)), (iii) a benchmark value used to remove
-low-density hexagons, and (iv) a benchmark value used to remove long
-edges. However, there is no analytical formula to calculate appropriate
-values for these parameters. The selection of these parameter values
-depends on the model performance computed by MSE
-(Section~\ref{sec-summary}), except for the threshold used to remove
-long edges.
-
-\subsubsection{Bin start position and number of
-bins}\label{bin-start-position-and-number-of-bins}
-
-The starting position of the hexagonal grid is important because
-different starting points result in different distributions of data
-across bins, even with the same total number of bins. This variation
-affects the model due to the differing number of non-empty bins.
-Therefore, it is necessary to evaluate various starting points with
-different total numbers of bins to determine which configurations are
-more effective at capturing the structure and fitting the model.
+The model fitting can be adjusted using these parameters:
+
+\begin{itemize}
+\tightlist
+\item
+  hexagon bin parameters
+
+  \begin{itemize}
+  \tightlist
+  \item
+    bottom left bin position \((s_1, \ s_2)\),
+  \item
+    the total number of bins (\(b\)),
+  \end{itemize}
+\item
+  bin density cutoff, to remove low-density hexagons, and
+\item
+  edge length maximum, remove long edges from \gD{} representation.
+\end{itemize}
+
+Default values are provided for each of these, but it is expected that
+the user will examine the MSE for a range of choices. Choosing these
+parameters according to MSE can be automated but it is recommended that
+the user examine the resulting model representation by overlaying it on
+the data in \pD{}. The next few subsections describe the calculation of
+default values, and the effect that different choices have on the model
+fit.
+
+\subsubsection{Hexagon bin parameters}\label{hexagon-bin-parameters}
+
+The values \((s_1, \ s_2)\) define the position of the centroid of the
+bottom left hexagon. By default, this is at \(s_1 = -q, s_2 = -qr_2\),
+where \(q\) is the buffer sound the data. The choice of these values can
+have some effect on the distribution of bin counts. FIG XXX(a)
+illustrates this. The distribution of bin counts for \(s_1\) varying
+between \(-0.1-0.0\) is shown. Generally, a more uniform distribution
+among these possibilities would indicate that the bins are reliably
+capturing the underlying distribution of observations.
 
@@
 
-To determine the effective \(b\), candidate values are selected based on
-the range between the minimum and approximate maximum \(b_1\), because
-\(b_2\) is computed from \(b_1\). The minimum \(b_1\) is set to \(2\),
-while the maximum number is estimated by taking the square root of
-\(\frac{n}{2}\). By evaluating MSE across varying \(b\) within this
-range for different \(q\), helps to determine an appropriate values for
-\(b\) and \(q\) (Figure~\ref{fig-param-scurve} (a)).
+The default number of bins \(b=b_1\times b_2\) is computed based on the
+sample size, by setting \(b_1=n^{1/3}\), consistent with the
+Diaconis-Freedman rule \citet{REF-doi:10.1007/BF01025868}. The value of
+\(b_2\) is determined analytically by \(b_1, q, r_2\). Values of \(b_1\)
+between 2 and \(b_1 = \frac{n}{2}\) are allowed. FIG XXX(b) shows the
+effect of different choices of \(b_1\) on the MSE of the fitted model.
 

### 2024-07-29 paper.qmd
low density

@@
 FIG XXX(b) shows the effect of different choices of $b_1$ on the MSE of the fitted model.
+XXX Are you showing $b_1$ or $b$ here????
 
@@
 
-Once setting up the hexagon grid with an appropriate number of bins, some hexagon bins may have few or no data points within them (@fig-bins-scurve (b)).
+By default, when assessing the choice of $b_1$, the total number of bins is measured by the number of **non-empty** bins.
+This more accurately reflects the hexagon grid relative the MSE than the full number of bins in the grid.
+It may also be beneficial to remove low count bins also, in the situation where data is clustered or stringy, where the observed data is sparse.
+In order to decide if this is necessary, you would examine the distribution of bin counts, or the density which puts the counts on a standard scale.
+If there is something of a gap at low values, this would suggest a potential value to use as a cutoff.
+Alternatively, one could choose to remove based on a percentile, the bins with density in the lowest 5% of all bins, for example.
+FIG XXXc illustrates the effect on the model representation of removing bins below different percentages.
+Generally, we would urge caution in removing low count bins.
+
+<!-- Once setting up the hexagon grid with an appropriate number of bins, some hexagon bins may have few or no data points within them (@fig-bins-scurve (b)).
 To ensure comprehensive coverage of the NLDR data, it is necessary to select hexagon bins with a considerable number of data points.
@@
 3. Select the first quantile value if the distribution is skewed.
+-->
 

### 2024-07-29 paper.tex
low density

@@
 effect of different choices of \(b_1\) on the MSE of the fitted model.
+XXX Are you showing \(b_1\) or \(b\) here????
 
@@
 
-Once setting up the hexagon grid with an appropriate number of bins,
-some hexagon bins may have few or no data points within them
-(Figure~\ref{fig-bins-scurve} (b)). To ensure comprehensive coverage of
-the NLDR data, it is necessary to select hexagon bins with a
-considerable number of data points. This involves calculating the number
-of points within each hexagon. Then, the standard count is computed by
-dividing the number of points within each hexagon by the maximum number
-of points in the grid. Next, bins with a standard count less than a
-benchmark value are removed (Figure~\ref{fig-param-scurve} (c)). There
-is no specific rule for selecting a benchmark value. However, the
-following steps can help determine a suitable value for removing
-low-density hexagons:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Plot the distribution of the standardized counts
-  (Figure~\ref{fig-param-scurve} (b)).
-\item
-  Examine the distribution of counts.
-\item
-  Select the first quantile value if the distribution is skewed.
-\end{enumerate}
+By default, when assessing the choice of \(b_1\), the total number of
+bins is measured by the number of \textbf{non-empty} bins. This more
+accurately reflects the hexagon grid relative the MSE than the full
+number of bins in the grid. It may also be beneficial to remove low
+count bins also, in the situation where data is clustered or stringy,
+where the observed data is sparse. In order to decide if this is
+necessary, you would examine the distribution of bin counts, or the
+density which puts the counts on a standard scale. If there is something
+of a gap at low values, this would suggest a potential value to use as a
+cutoff. Alternatively, one could choose to remove based on a percentile,
+the bins with density in the lowest 5\% of all bins, for example. FIG
+XXXc illustrates the effect on the model representation of removing bins
+below different percentages. Generally, we would urge caution in
+removing low count bins.
 

### 2024-07-29 paper.qmd
Add citations

@@
 
-The default number of bins $b=b_1\times b_2$ is computed based on the sample size, by setting $b_1=n^{1/3}$, consistent with the Diaconis-Freedman rule @REF-doi:10.1007/BF01025868.
+The default number of bins $b=b_1\times b_2$ is computed based on the sample size, by setting $b_1=n^{1/3}$, consistent with the Diaconis-Freedman rule [@freedman1981].
 The value of $b_2$ is determined analytically by $b_1, q, r_2$.
-Values of $b_1$ between 2 and $b_1 = \frac{n}{2}$ are allowed.
+Values of $b_1$ between $2$ and $b_1 = \frac{n}{2}$ are allowed.
 FIG XXX(b) shows the effect of different choices of $b_1$ on the MSE of the fitted model.
 XXX Are you showing $b_1$ or $b$ here????
+$b$
 

### 2024-07-29 paper.tex
Add citations

@@
 sample size, by setting \(b_1=n^{1/3}\), consistent with the
-Diaconis-Freedman rule \citet{REF-doi:10.1007/BF01025868}. The value of
-\(b_2\) is determined analytically by \(b_1, q, r_2\). Values of \(b_1\)
-between 2 and \(b_1 = \frac{n}{2}\) are allowed. FIG XXX(b) shows the
-effect of different choices of \(b_1\) on the MSE of the fitted model.
-XXX Are you showing \(b_1\) or \(b\) here????
+Diaconis-Freedman rule \citep{freedman1981}. The value of \(b_2\) is
+determined analytically by \(b_1, q, r_2\). Values of \(b_1\) between
+\(2\) and \(b_1 = \frac{n}{2}\) are allowed. FIG XXX(b) shows the effect
+of different choices of \(b_1\) on the MSE of the fitted model. XXX Are
+you showing \(b_1\) or \(b\) here???? \(b\)
 
@@
 
-\captionsetup{labelsep=none}\includegraphics{figures/mnist/mnist_model_2d.png}
+\includegraphics{figures/mnist/mnist_model_2d.png}
 
@@
 
-\captionsetup{labelsep=none}\includegraphics{figures/mnist/img_error_sample_within.png}
+\includegraphics{figures/mnist/img_error_sample_within.png}
 
@@
 
-\captionsetup{labelsep=none}\includegraphics{figures/mnist/img_error_sample_out.png}
+\includegraphics{figures/mnist/img_error_sample_out.png}
 

### 2024-07-29 paper.qmd
Fixed the isse with hexagon grid of Figure 3 and 7

@@
 sc_xlims <- c(-0.25, 1.25)
-sc_ylims <- c(-0.2, 2)
+sc_ylims <- c(-0.3, 2)
 
@@
   interior_annotation("a", position = c(0.92, 0.93)) +
-  xlim(c(-0.25, 1.35)) + ylim(c(-0.25, 2))
+  xlim(c(-0.25, 1.35)) + ylim(c(-0.35, 2))
 ```
@@
     na.value = "#ffffff", option = "C") +
-  xlim(c(-0.25, 1.35)) + ylim(c(-0.25, 2)) +
+  xlim(c(-0.25, 1.35)) + ylim(c(-0.35, 2)) +
   interior_annotation("b")
@@
   scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  xlim(c(-0.25, 1.35)) + ylim(c(-0.25, 2)) +
+  xlim(c(-0.25, 1.35)) + ylim(c(-0.35, 2)) +
   interior_annotation("c")

### 2024-07-29 paper.qmd
small changes made after last commit - should be ok to merge

--- 2024-07-29 paper.qmd
@@
 
-To create a smooth \gD{} representation (@fig-NLDR-scurve (d)), it is necessary to remove edges that connect distant bin centroids in the triangular mesh.
+Edges define the neighbourhood structure, in order to provide a smooth \gD{} representation of the fitted model.
+@fig-scurve-true-sc shows a wire frame of the true model that was used to generate the S-curve example data.
+The fitted model should come reasonably close to this, if the fitting has been successful.
+We don't ever have this, of course
+
+(@fig-NLDR-scurve (d)), it is necessary to remove edges that connect distant bin centroids in the triangular mesh.
 These edges only exist in the \gD{} model and do not extend into \pD{}, so their removal does not impact the model in \pD{}.

### 2024-07-29 paper.qmd
added histogram under hexgrid plots and fixed freedman reference

@@
   xlim(c(-0.25, 1.35)) + ylim(c(-0.35, 2))
+
+hex_grid_coloured_scurve1_dens <- ggplot(hex_grid_with_counts_s_curve1) +
+  geom_histogram(aes(x=std_counts), binwidth = hb_obj_scurve1$centroids$c_x[2]-hb_obj_scurve1$centroids$c_x[1])
 ```
@@
 
+hex_grid_coloured_scurve2_dens <- ggplot(hex_grid_with_counts_s_curve2) +
+  geom_histogram(aes(x=std_counts), binwidth = hb_obj_scurve2$centroids$c_x[2]-hb_obj_scurve2$centroids$c_x[1])
 ```
@@
 
+hex_grid_coloured_scurve3_dens <- ggplot(hex_grid_with_counts_s_curve3) +
+  geom_histogram(aes(x=std_counts), binwidth = hb_obj_scurve3$centroids$c_x[2]-hb_obj_scurve3$centroids$c_x[1]) 
 ```
@@
 #| fig-width: 6
-#| fig-height: 3
-
-hex_grid_coloured_scurve1 + hex_grid_coloured_scurve2 + hex_grid_coloured_scurve3 +
-  plot_layout(guides='collect', ncol = 3) &
+#| fig-height: 5
+
+hex_grid_coloured_scurve1 + 
+  hex_grid_coloured_scurve2 + 
+  hex_grid_coloured_scurve3 +
+hex_grid_coloured_scurve1_dens +
+  hex_grid_coloured_scurve2_dens +
+  hex_grid_coloured_scurve3_dens +
+  plot_layout(guides='collect', ncol = 3,
+              heights = c(2,1)) &
   theme(legend.position='none', plot.tag = element_text(size = 8))

### 2024-07-30 paper.tex
Rearrange script folder

@@
 
-To create a smooth \gD{} representation (Figure~\ref{fig-NLDR-scurve}
-(d)), it is necessary to remove edges that connect distant bin centroids
-in the triangular mesh. These edges only exist in the \gD{} model and do
-not extend into \pD{}, so their removal does not impact the model in
-\pD{}. Although there are no specific criteria for determining the
-benchmark value to remove long edges, the following steps provide an
-approach to identifying a suitable threshold:
+Edges define the neighbourhood structure, in order to provide a smooth
+\gD{} representation of the fitted model.
+Figure~\ref{fig-scurve-true-sc} shows a wire frame of the true model
+that was used to generate the S-curve example data. The fitted model
+should come reasonably close to this, if the fitting has been
+successful. We don't ever have this, of course
+
+(Figure~\ref{fig-NLDR-scurve} (d)), it is necessary to remove edges that
+connect distant bin centroids in the triangular mesh. These edges only
+exist in the \gD{} model and do not extend into \pD{}, so their removal
+does not impact the model in \pD{}. Although there are no specific
+criteria for determining the benchmark value to remove long edges, the
+following steps provide an approach to identifying a suitable threshold:
 

### 2024-07-30 paper.qmd
Added histogram for Figure 8

@@
         plot.title = element_text(size = 12))
+
+# Define bin width and starting point
+bin_width <- hb_obj_scurve$centroids$c_x[2]-hb_obj_scurve2$centroids$c_x[1]
+start_point <- bin_width/2
+
+# Create bins and calculate the mean value for each bin range
+bins <- seq(start_point, max(distance_scurve$distance) + bin_width, 
+            by = bin_width)
+bin_labels <- bins[-length(bins)] + bin_width / 2  # mean of each bin range
+
+distance_scurve <- distance_scurve |> 
+  mutate(bin = cut(distance, breaks = bins, include.lowest = TRUE, labels = bin_labels))
+
+distance_hist <- ggplot(
+  distance_scurve, aes(x = distance)) +
+    geom_histogram(
+      aes(y = after_stat(count) / sum(after_stat(count))), 
+      binwidth = bin_width) +
+  scale_x_continuous(breaks = bin_labels, labels = round(bin_labels, 3)) +
+  ylab("standardized number of edges") +
+  xlab(expression(d^{(2)})) +
+  ggtitle("(d)") +
+  theme(aspect.ratio = 0.75,
+        axis.title.x = element_text(size = 12),
+        axis.title.y = element_text(size = 12, angle = 90),
+        axis.text.x = element_text(size = 12, angle = 90),
+        axis.text.y = element_text(size = 12),
+        axis.ticks.x = element_line(), 
+        plot.title = element_text(size = 12))
 ```
@@
 mse_scurve_lwd + 
-distance_scurve_plot + plot_layout(ncol=2)
+distance_hist + plot_layout(ncol=2)
 ```

### 2024-07-30 paper.qmd
fixed country name for authors, labelled all code chunks

--- 2024-07-29 paper.qmd
@@
         city: VIC 
-        country: Austria
+        country: Australia
         postal-code: 3800
@@
 ```{r}
+#| label: umap-scurve
 umap_scurve <- read_rds(file = "data/s_curve/s_curve_umap.rds")
@@
 ```{r}
+#| label: hexbin-scurve
 ## Compute hexbin parameters
@@
 ```{r}
+#| label: empty-bin-scurve
 hex_grid_nonempty <- hex_grid |>
@@
 ```{r}
+#| label: triangulate-scurve
 ## Triangulate bin centroids
@@
 ```{r}
+#| label: code-illustration
 # Code to draw illustration for notation
@@
 ```{r}
+#| label: hexbin-regular-scurve1
 ## hexagon binning to have regular hexagons
@@
 ```{r}
+#| label: hexbin-regular-scurve2
 ## hexagon binning to have regular hexagons
@@
 ```{r}
+#| label: hexbin-regular-scurve3
 ## hexagon binning to have regular hexagons
@@
 ```{r}
+#| label: scurve-training
 training_data_scurve <- read_rds("data/s_curve/s_curve_training.rds")
@@
 ```{r}
+#| label: errors-scurve
 ## To initialize number of bins along the x-axis
@@
 ```{r}
+#| label: distribution-scurve
 ## To initialize effective bins along x
@@
 ```{r}
+#| label: bin-counts-scurve
 cell_count_scurve <- ggplot(df_bin_centroids_scurve, 
@@
 ```{r}
+#| label: bins-scurve
 ## To initialize benchmark values to remove low density hexagons
@@
 ```{r}
+#| label: bin-benchmark-scurve
 hex_grid_with_counts_s_curve2 <- full_join(
@@
 ```{r}
+#| label: bin-neighbours
 ## First define low density bins using first quantile
@@
 ```{r}
+#| label: triangulate-scurve2
 ## Triangulate bin centroids
@@
 ```{r}
+#| label: published-pbmc
 ## Import data
@@
 ```{r}
+#| label: hexbin-pbmc
 ## Compute hexbin parameters
@@
 ```{r}
+#| label: initial-bins-pbmc
 ## To initialize number of bins along the x-axis
@@
 ```{r}
+#| label: combine-data-pbmc
 error_pbmc <- bind_rows(error_pbmc_umap, 
@@
 ```{r}
+#| label: tsne-pbmc
 # read the png file from device 
@@
 ```{r}
+#| label: tsne-pbmc-read
 tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_30.rds")
@@
 ```{r}
+#| label: num-bins-pbmc
 ## Compute hexbin parameters
@@
 ```{r}
+#| label: read-mnist
 ## Import data
@@
 ```{r}
+#| label: hexbin-mnist
 ## Compute hexbin parameters

### 2024-07-30 paper.qmd
Regenerated Figure 8 (b) with a line

@@
                                             -std_counts), 
-                                y = std_counts)) + 
-  geom_quasirandom(
-    size = 1,
-    alpha = 0.5
-  ) + 
+                                y = std_counts,
+                                group = "1")) + 
+  geom_line() + 
   geom_hline(yintercept = benchmark1, color = "black", linewidth=0.5, alpha = 0.5) +

### 2024-07-30 paper.qmd
still got histograms but a little neater

--- 2024-07-30 paper.qmd
@@
 hex_grid_coloured_scurve1_dens <- ggplot(hex_grid_with_counts_s_curve1) +
-  geom_histogram(aes(x=std_counts), binwidth = hb_obj_scurve1$centroids$c_x[2]-hb_obj_scurve1$centroids$c_x[1])
+  geom_histogram(aes(x=std_counts), 
+                 breaks=seq(0, 1, 0.05),
+                 fill="slategray4", colour="white")
+
+#hex_grid_coloured_scurve1_dens <- 
+#  hex_grid_with_counts_s_curve1 |>
+#  filter(!is.na(std_counts)) |>
+#  mutate(ord_std_counts = n() - rank(std_counts)) |>
+#  ggplot() +
+#  geom_line(aes(x=ord_std_counts, y=std_counts), 
+#                 colour="slategray4")
 ```
@@
 hex_grid_coloured_scurve2_dens <- ggplot(hex_grid_with_counts_s_curve2) +
-  geom_histogram(aes(x=std_counts), binwidth = hb_obj_scurve2$centroids$c_x[2]-hb_obj_scurve2$centroids$c_x[1])
+  geom_histogram(aes(x=std_counts), 
+                 breaks=seq(0, 1, 0.05),
+                 fill="slategray4", colour="white")
+
+#hex_grid_coloured_scurve2_dens <- 
+#  hex_grid_with_counts_s_curve2 |>
+#  filter(!is.na(std_counts)) |>
+#  mutate(ord_std_counts = n() - rank(std_counts)) |>
+#  ggplot() +
+#  geom_line(aes(x=ord_std_counts, y=std_counts), 
+#                 colour="slategray4")
+
 ```
@@
 hex_grid_coloured_scurve3_dens <- ggplot(hex_grid_with_counts_s_curve3) +
-  geom_histogram(aes(x=std_counts), binwidth = hb_obj_scurve3$centroids$c_x[2]-hb_obj_scurve3$centroids$c_x[1]) 
+  geom_histogram(aes(x=std_counts), 
+                 breaks=seq(0, 1, 0.05),
+                 fill="slategray4", colour="white") 
+#hex_grid_coloured_scurve3_dens <- 
+#  hex_grid_with_counts_s_curve3 |>
+#  filter(!is.na(std_counts)) |>
+#  mutate(ord_std_counts = n() - rank(std_counts)) |>
+#  ggplot() +
+#  geom_line(aes(x=ord_std_counts, y=std_counts), 
+#                 colour="slategray4")
 ```

### 2024-07-30 paper.qmd
Revise Figure 8 (b), still not useful..need to update

@@
 
+a1_1 <- calc_bins_y(
+  bin1 = 7, 
+  r2 = r2,
+  q = 0.12
+)$a1
+
 ## Data set with all centroids
@@
   counts_df = counts_df1) |>
-  filter(drop_empty == FALSE)
+  filter(drop_empty == FALSE) |>
+  mutate(a1 = a1_1)
 
@@
 
+a1_2 <- calc_bins_y(
+  bin1 = 11, 
+  r2 = r2,
+  q = 0.07
+)$a1
+
 ## Data set with all centroids
@@
   counts_df = counts_df2) |>
-  filter(drop_empty == FALSE)
+  filter(drop_empty == FALSE) |>
+  mutate(a1 = a1_2)
 
@@
 
+a1_3 <- calc_bins_y(
+  bin1 = 17, 
+  r2 = r2,
+  q = 0.05
+)$a1
+
 ## Data set with all centroids
@@
   counts_df = counts_df3) |>
-  filter(drop_empty == FALSE)
+  filter(drop_empty == FALSE) |>
+  mutate(a1 = a1_3)
 
@@
 #| label: bin-counts-scurve
-cell_count_scurve <- ggplot(df_bin_centroids_scurve, 
+
+df_bin_centroids_all <- bind_rows(
+  df_bin_centroids1,
+  df_bin_centroids2,
+  df_bin_centroids3
+)
+
+cell_count_scurve <- ggplot(df_bin_centroids_all, 
                             aes(x = reorder(as.factor(hexID), 
@@
                                 y = std_counts,
-                                group = "1")) + 
+                                group = a1,
+                                color = a1)) + 
   geom_line() + 
   geom_hline(yintercept = benchmark1, color = "black", linewidth=0.5, alpha = 0.5) +
+  scale_colour_continuous_sequential(palette = "Magenta") +
   xlab("hexagon id (re-ordered)") + 

### 2024-07-30 paper.qmd
Add a script to generate MSE for diff method and hyper parameters for pbmc3k

@@
 ```{r}
-#| label: initial-bins-pbmc
-## To initialize number of bins along the x-axis
-bin1_vec_pbmc <- 2:36 #sqrt(NROW(training_data_pbmc)/2)
-
-## For umap
-umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_30_min_dist_0.3.rds")
-pbmc_scaled_obj_umap <- gen_scaled_data(
-  data = umap_pbmc)
-umap_pbmc_scaled <- pbmc_scaled_obj_umap$scaled_nldr
-
-lim1 <- pbmc_scaled_obj_umap$lim1
-lim2 <- pbmc_scaled_obj_umap$lim2
-r2_umap <- diff(lim2)/diff(lim1)
-
-error_pbmc_umap <- data.frame(matrix(nrow = 0, ncol = 0))
-
-for (xbins in bin1_vec_pbmc) {
-
-  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.05)$bin2
-
-  pbmc_model <- fit_highd_model(
-    training_data = training_data_pbmc,
-    emb_df = umap_pbmc_scaled,
-    bin1 = xbins,
-    r2 = r2_umap,
-    q = 0.05, 
-    is_bin_centroid = TRUE,
-    is_rm_lwd_hex = FALSE,
-    col_start_highd = "PC_"
-  )
-
-  df_bin_centroids_pbmc <- pbmc_model$df_bin_centroids
-  df_bin_pbmc <- pbmc_model$df_bin
-
-  ## Compute error
-  error_df <- glance(
-    df_bin_centroids = df_bin_centroids_pbmc,
-    df_bin = df_bin_pbmc,
-    training_data = training_data_pbmc,
-    newdata = NULL,
-    type_NLDR = "UMAP",
-    col_start = "PC_") |>
-    mutate(bin1 = xbins,
-           bin2 = bin2,
-           b = bin1 * bin2,
-           b_non_empty = NROW(df_bin_centroids_pbmc),
-           method = "UMAP_30_min_dist_0.3")
-
-  error_pbmc_umap <- bind_rows(error_pbmc_umap, error_df)
-
-}
-
-## To initialize number of bins along the x-axis
-bin1_vec_pbmc <- 2:36 #sqrt(NROW(training_data_pbmc)/2)
-
-## For umap
-umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_5_min_dist_0.01.rds")
-pbmc_scaled_obj_umap <- gen_scaled_data(
-  data = umap_pbmc)
-umap_pbmc_scaled <- pbmc_scaled_obj_umap$scaled_nldr
-
-lim1 <- pbmc_scaled_obj_umap$lim1
-lim2 <- pbmc_scaled_obj_umap$lim2
-r2_umap <- diff(lim2)/diff(lim1)
-
-error_pbmc_umap2 <- data.frame(matrix(nrow = 0, ncol = 0))
-
-for (xbins in bin1_vec_pbmc) {
-
-  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.05)$bin2
-
-  pbmc_model <- fit_highd_model(
-    training_data = training_data_pbmc,
-    emb_df = umap_pbmc_scaled,
-    bin1 = xbins,
-    r2 = r2_umap,
-    q = 0.05, 
-    is_bin_centroid = TRUE,
-    is_rm_lwd_hex = FALSE,
-    col_start_highd = "PC_"
-  )
-
-  df_bin_centroids_pbmc <- pbmc_model$df_bin_centroids
-  df_bin_pbmc <- pbmc_model$df_bin
-
-  ## Compute error
-  error_df <- glance(
-    df_bin_centroids = df_bin_centroids_pbmc,
-    df_bin = df_bin_pbmc,
-    training_data = training_data_pbmc,
-    newdata = NULL,
-    type_NLDR = "UMAP",
-    col_start = "PC_") |>
-    mutate(bin1 = xbins,
-           bin2 = bin2,
-           b = bin1 * bin2,
-           b_non_empty = NROW(df_bin_centroids_pbmc),
-           method = "UMAP_5_min_dist_0.01")
-
-  error_pbmc_umap2 <- bind_rows(error_pbmc_umap2, error_df)
-
-}
-
-## For umap
-
-## To initialize number of bins along the x-axis
-bin1_vec_pbmc <- 2:23 #sqrt(NROW(training_data_pbmc)/2)
-umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_15_min_dist_0.99.rds")
-pbmc_scaled_obj_umap <- gen_scaled_data(
-  data = umap_pbmc)
-umap_pbmc_scaled <- pbmc_scaled_obj_umap$scaled_nldr
-
-lim1 <- pbmc_scaled_obj_umap$lim1
-lim2 <- pbmc_scaled_obj_umap$lim2
-r2_umap <- diff(lim2)/diff(lim1)
-
-error_pbmc_umap3 <- data.frame(matrix(nrow = 0, ncol = 0))
-
-for (xbins in bin1_vec_pbmc) {
-
-  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.05)$bin2
-
-  pbmc_model <- fit_highd_model(
-    training_data = training_data_pbmc,
-    emb_df = umap_pbmc_scaled,
-    bin1 = xbins,
-    r2 = r2_umap,
-    q = 0.05, 
-    is_bin_centroid = TRUE,
-    is_rm_lwd_hex = FALSE,
-    col_start_highd = "PC_"
-  )
-
-  df_bin_centroids_pbmc <- pbmc_model$df_bin_centroids
-  df_bin_pbmc <- pbmc_model$df_bin
-
-  ## Compute error
-  error_df <- glance(
-    df_bin_centroids = df_bin_centroids_pbmc,
-    df_bin = df_bin_pbmc,
-    training_data = training_data_pbmc,
-    newdata = NULL,
-    type_NLDR = "UMAP",
-    col_start = "PC_") |>
-    mutate(bin1 = xbins,
-           bin2 = bin2,
-           b = bin1 * bin2,
-           b_non_empty = NROW(df_bin_centroids_pbmc),
-           method = "UMAP_15_min_dist_0.99")
-
-  error_pbmc_umap3 <- bind_rows(error_pbmc_umap3, error_df)
-
-}
-
-## To initialize number of bins along the x-axis
-bin1_vec_pbmc <- 2:22 #sqrt(NROW(training_data_pbmc)/2)
-
-## For tsne
-tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_5.rds")
-pbmc_scaled_obj_tsne <- gen_scaled_data(
-  data = tsne_pbmc)
-tsne_pbmc_scaled <- pbmc_scaled_obj_tsne$scaled_nldr
-
-lim1 <- pbmc_scaled_obj_tsne$lim1
-lim2 <- pbmc_scaled_obj_tsne$lim2
-r2_tsne <- diff(lim2)/diff(lim1)
-
-error_pbmc_tsne <- data.frame(matrix(nrow = 0, ncol = 0))
-
-for (xbins in bin1_vec_pbmc) {
-
-  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.05)$bin2
-
-  pbmc_model <- fit_highd_model(
-    training_data = training_data_pbmc,
-    emb_df = tsne_pbmc_scaled,
-    bin1 = xbins,
-    r2 = r2_tsne,
-    q = 0.05, 
-    is_bin_centroid = TRUE,
-    is_rm_lwd_hex = FALSE,
-    col_start_highd = "PC_"
-  )
-
-  df_bin_centroids_pbmc <- pbmc_model$df_bin_centroids
-  df_bin_pbmc <- pbmc_model$df_bin
-
-  ## Compute error
-  error_df <- glance(
-    df_bin_centroids = df_bin_centroids_pbmc,
-    df_bin = df_bin_pbmc,
-    training_data = training_data_pbmc,
-    newdata = NULL,
-    type_NLDR = "tSNE",
-    col_start = "PC_") |>
-    mutate(bin1 = xbins,
-           bin2 = bin2,
-           b = bin1 * bin2,
-           b_non_empty = NROW(df_bin_centroids_pbmc),
-           method = "tsne_5")
-
-  error_pbmc_tsne <- bind_rows(error_pbmc_tsne, error_df)
-
-}
-
-## To initialize number of bins along the x-axis
-bin1_vec_pbmc <- 2:26 #sqrt(NROW(training_data_pbmc)/2)
-
-## For tsne
-tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_30.rds")
-pbmc_scaled_obj_tsne <- gen_scaled_data(
-  data = tsne_pbmc)
-tsne_pbmc_scaled <- pbmc_scaled_obj_tsne$scaled_nldr
-
-lim1 <- pbmc_scaled_obj_tsne$lim1
-lim2 <- pbmc_scaled_obj_tsne$lim2
-r2_tsne <- diff(lim2)/diff(lim1)
-
-error_pbmc_tsne2 <- data.frame(matrix(nrow = 0, ncol = 0))
-
-for (xbins in bin1_vec_pbmc) {
-
-  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.05)$bin2
-
-  pbmc_model <- fit_highd_model(
-    training_data = training_data_pbmc,
-    emb_df = tsne_pbmc_scaled,
-    bin1 = xbins,
-    r2 = r2_tsne,
-    q = 0.05, 
-    is_bin_centroid = TRUE,
-    is_rm_lwd_hex = FALSE,
-    col_start_highd = "PC_"
-  )
-
-  df_bin_centroids_pbmc <- pbmc_model$df_bin_centroids
-  df_bin_pbmc <- pbmc_model$df_bin
-
-  ## Compute error
-  error_df <- glance(
-    df_bin_centroids = df_bin_centroids_pbmc,
-    df_bin = df_bin_pbmc,
-    training_data = training_data_pbmc,
-    newdata = NULL,
-    type_NLDR = "tSNE",
-    col_start = "PC_") |>
-    mutate(bin1 = xbins,
-           bin2 = bin2,
-           b = bin1 * bin2,
-           b_non_empty = NROW(df_bin_centroids_pbmc),
-           method = "tsne_30")
-
-  error_pbmc_tsne2 <- bind_rows(error_pbmc_tsne2, error_df)
-
-}
-
-## To initialize number of bins along the x-axis
-bin1_vec_pbmc <- 2:46 #sqrt(NROW(training_data_pbmc)/2)
-
-## For phate
-phate_pbmc <- read_rds("data/pbmc3k/pbmc_phate_5.rds")
-pbmc_scaled_obj_phate <- gen_scaled_data(
-  data = phate_pbmc)
-phate_pbmc_scaled <- pbmc_scaled_obj_phate$scaled_nldr
-
-lim1 <- pbmc_scaled_obj_phate$lim1
-lim2 <- pbmc_scaled_obj_phate$lim2
-r2_phate <- diff(lim2)/diff(lim1)
-
-error_pbmc_phate <- data.frame(matrix(nrow = 0, ncol = 0))
-
-for (xbins in bin1_vec_pbmc) {
-
-  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.05)$bin2
-
-  pbmc_model <- fit_highd_model(
-    training_data = training_data_pbmc,
-    emb_df = phate_pbmc_scaled,
-    bin1 = xbins,
-    r2 = r2_phate,
-    q = 0.05, 
-    is_bin_centroid = TRUE,
-    is_rm_lwd_hex = FALSE,
-    col_start_highd = "PC_"
-  )
-
-  df_bin_centroids_pbmc <- pbmc_model$df_bin_centroids
-  df_bin_pbmc <- pbmc_model$df_bin
-
-  ## Compute error
-  error_df <- glance(
-    df_bin_centroids = df_bin_centroids_pbmc,
-    df_bin = df_bin_pbmc,
-    training_data = training_data_pbmc,
-    newdata = NULL,
-    type_NLDR = "PHATE",
-    col_start = "PC_") |>
-    mutate(bin1 = xbins,
-           bin2 = bin2,
-           b = bin1 * bin2,
-           b_non_empty = NROW(df_bin_centroids_pbmc),
-           method = "phate_5")
-
-  error_pbmc_phate <- bind_rows(error_pbmc_phate, error_df)
-
-}
-
-## To initialize number of bins along the x-axis
-bin1_vec_pbmc <- 2:36 #sqrt(NROW(training_data_pbmc)/2)
-
-## For trimap
-trimap_pbmc <- read_rds("data/pbmc3k/pbmc_trimap_12_4_3.rds")
-pbmc_scaled_obj_trimap <- gen_scaled_data(
-  data = trimap_pbmc)
-trimap_pbmc_scaled <- pbmc_scaled_obj_trimap$scaled_nldr
-
-lim1 <- pbmc_scaled_obj_trimap$lim1
-lim2 <- pbmc_scaled_obj_trimap$lim2
-r2_trimap <- diff(lim2)/diff(lim1)
-
-error_pbmc_trimap <- data.frame(matrix(nrow = 0, ncol = 0))
-
-for (xbins in bin1_vec_pbmc) {
-
-  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.05)$bin2
-
-  pbmc_model <- fit_highd_model(
-    training_data = training_data_pbmc,
-    emb_df = trimap_pbmc_scaled,
-    bin1 = xbins,
-    r2 = r2_trimap,
-    q = 0.05, 
-    is_bin_centroid = TRUE,
-    is_rm_lwd_hex = FALSE,
-    col_start_highd = "PC_"
-  )
-
-  df_bin_centroids_pbmc <- pbmc_model$df_bin_centroids
-  df_bin_pbmc <- pbmc_model$df_bin
-
-  ## Compute error
-  error_df <- glance(
-    df_bin_centroids = df_bin_centroids_pbmc,
-    df_bin = df_bin_pbmc,
-    training_data = training_data_pbmc,
-    newdata = NULL,
-    type_NLDR = "TriMAP",
-    col_start = "PC_") |>
-    mutate(bin1 = xbins,
-           bin2 = bin2,
-           b = bin1 * bin2,
-           b_non_empty = NROW(df_bin_centroids_pbmc),
-           method = "trimap_12_4_3")
-
-  error_pbmc_trimap <- bind_rows(error_pbmc_trimap, error_df)
-
-}
-
-## To initialize number of bins along the x-axis
-bin1_vec_pbmc <- 2:43 #sqrt(NROW(training_data_pbmc)/2)
-
-## For pacmap
-pacmap_pbmc <- read_rds("data/pbmc3k/pbmc_pacmap_30_random_0.9_5.rds")
-pbmc_scaled_obj_pacmap <- gen_scaled_data(
-  data = pacmap_pbmc)
-pacmap_pbmc_scaled <- pbmc_scaled_obj_pacmap$scaled_nldr
-
-lim1 <- pbmc_scaled_obj_pacmap$lim1
-lim2 <- pbmc_scaled_obj_pacmap$lim2
-r2_pacmap <- diff(lim2)/diff(lim1)
-
-error_pbmc_pacmap <- data.frame(matrix(nrow = 0, ncol = 0))
-
-for (xbins in bin1_vec_pbmc) {
-
-  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.05)$bin2
-
-  pbmc_model <- fit_highd_model(
-    training_data = training_data_pbmc,
-    emb_df = pacmap_pbmc_scaled,
-    bin1 = xbins,
-    r2 = r2_pacmap,
-    q = 0.05, 
-    is_bin_centroid = TRUE,
-    is_rm_lwd_hex = FALSE,
-    col_start_highd = "PC_"
-  )
-
-  df_bin_centroids_pbmc <- pbmc_model$df_bin_centroids
-  df_bin_pbmc <- pbmc_model$df_bin
-
-  ## Compute error
-  error_df <- glance(
-    df_bin_centroids = df_bin_centroids_pbmc,
-    df_bin = df_bin_pbmc,
-    training_data = training_data_pbmc,
-    newdata = NULL,
-    type_NLDR = "PaCMAP",
-    col_start = "PC_") |>
-    mutate(bin1 = xbins,
-           bin2 = bin2,
-           b = bin1 * bin2,
-           b_non_empty = NROW(df_bin_centroids_pbmc),
-           method = "pacmap")
-
-  error_pbmc_pacmap <- bind_rows(error_pbmc_pacmap, error_df)
-
-}
-
-```
-
-```{r}
 #| label: combine-data-pbmc
+
+error_pbmc_umap <- read_rds("data/pbmc3k/error_pbmc_umap_30_min_dist_0.3.rds")
+error_pbmc_umap2 <- read_rds("data/pbmc3k/error_pbmc_umap_5_min_dist_0.01.rds")
+error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_15_min_dist_0.99.rds")
+error_pbmc_tsne <- read_rds("data/pbmc3k/error_pbmc_tsne_5.rds")
+error_pbmc_tsne2 <- read_rds("data/pbmc3k/error_pbmc_tsne_30.rds")
+error_pbmc_phate <- read_rds("data/pbmc3k/error_pbmc_phate_5.rds") 
+error_pbmc_trimap <- read_rds("data/pbmc3k/error_pbmc_trimap_12_4_3.rds") 
+error_pbmc_pacmap <- read_rds("data/pbmc3k/error_pbmc_pacmap_30_random_0.9_5.rds")
+
 error_pbmc <- bind_rows(error_pbmc_umap, 

### 2024-07-30 paper.qmd
Changed the Figure 8 (b) with bin counts

@@
                             aes(x = reorder(as.factor(hexID), 
-                                            -std_counts), 
-                                y = std_counts,
+                                            -bin_counts), 
+                                y = bin_counts,
                                 group = a1,
@@
   geom_line() + 
-  geom_hline(yintercept = benchmark1, color = "black", linewidth=0.5, alpha = 0.5) +
+  # geom_hline(yintercept = benchmark1, color = "black", linewidth=0.5, alpha = 0.5) +
   scale_colour_continuous_sequential(palette = "Magenta") +

### 2024-07-31 paper.qmd
Add a1 values as factors

--- 2024-07-30 paper.qmd
@@
                             aes(x = reorder(as.factor(hexID), 
-                                            -std_counts), 
-                                y = std_counts,
-                                group = a1,
-                                color = a1)) + 
+                                            -bin_counts), 
+                                y = bin_counts,
+                                group = as.factor(a1),
+                                color = as.factor(a1))) + 
   geom_line() + 
-  geom_hline(yintercept = benchmark1, color = "black", linewidth=0.5, alpha = 0.5) +
-  scale_colour_continuous_sequential(palette = "Magenta") +
+  # geom_hline(yintercept = benchmark1, color = "black", linewidth=0.5, alpha = 0.5) +
+  scale_colour_discrete_qualitative(palette = "Dark 2") +
   xlab("hexagon id (re-ordered)") + 

### 2024-07-31 paper.qmd
done with long edges

--- 2024-07-30 paper.qmd
@@
 A simple approach would be to triangulate the points and add edges.
-A more stable approach is to first hexagonly bin the data, reducing it from $n$ to $m\leq n$ observations, and connect the bin centroids.
+A more stable approach is to first bin the data, reducing it from $n$ to $m\leq n$ observations, and connect the bin centroids.
+We recommend using a hexagon grid because it better reflects the data distribution and has less artifacts than a rectangular grid.
 This process serves to reduce some noisiness in the resulting surface shown in \pD{}.
@@
 
-Delaunay triangulation [@lee1980;@alb2024] is used to connect neighboring centroids. The edges are needed to preserve neighborhood information when the model is lifted into \pD{}.
+Delaunay triangulation [@lee1980;@alb2024] is used to connect points so that edges indicate neighbouring observations, in both the NLDR layout (@fig-NLDR-scurve (d)) and the \pD{} model representation.
+When the data has been binned the triangulation connectd centroids.
+The edges preserve the neighborhood information when the model is lifted into \pD{}.
 
@@
 
-When shapes are non-linear in the NLDR layout, some edges could be long. It is unlikely
-
-It can also happen that distant centroids can be connected, which can result in long line segments.
-In order to generate a smooth surface in \gD{}, these long line segments must be removed (@fig-NLDR-scurve (d)).
+When shapes are non-linear in the NLDR layout, some edges could be long.
+It can also happen that distant centroids can be connected, particularly if clustering is present, which can result in long line segments.
+In order to generate a smooth surface in \gD{}, these long line segments should be removed when tuning the model fit.
 
@@
 @fig-scurve-true-sc shows a wire frame of the true model that was used to generate the S-curve example data.
-The fitted model should come reasonably close to this, if the fitting has been successful.
-We don't ever have this, of course
-
-(@fig-NLDR-scurve (d)), it is necessary to remove edges that connect distant bin centroids in the triangular mesh.
+The ideal is that the representation of the fitted model, at least for this example where we know the true model, should look similar to this.
+
+The Delaunay triangulation will ensure that all centroids are connected into a triangular mesh.
+For some structures, like clustered data, or highly non-linear shapes, breaks in the mesh are meaningful.
+When separated clusters are present the mesh should be broken across the gaps.
+For non-linear structures like the S-curve, the mesh should run unbroken along the S, but there should be no edges connecting the top of the S directly to the bottom of the S.
+For these reasons it is necessary to remove edges from the mesh in some applications.
+
+The decision on edge length removal is made based on the distribution of edge lengths.
+In particular, a gap between values, where there a concentration of small values and then a few larger values, likely suggests a cutoff for edge removal.
+Because the triangulation is typically done on the hexagon centroids, there are particular discrete edge lengths, based on bin widths.
+@fig-NLDR-scurve XXX illustrates edge length distributions.
+
+There is an additional step that is needed.
+When the model is lifted into \pD{}, if the fit is good all the edges should be relatively small in this space, too.
+If this is not the case, then there are several possible actions: (1) re-do the NLDR to get a more representative layout; (2) identify the edge and remove it from the model, in \gD{} and \pD{}; (3) consider different values for the model fit, number of bins, initial bin position or removing low density bins.
+
+<!-- (@fig-NLDR-scurve (d)), it is necessary to remove edges that connect distant bin centroids in the triangular mesh.
 These edges only exist in the \gD{} model and do not extend into \pD{}, so their removal does not impact the model in \pD{}.
@@
 2. Identify the first largest difference between consecutive distance values.
-3. Take the distance value corresponding to this difference as the benchmark value.
+3. Take the distance value corresponding to this difference as the benchmark value.-->
 

### 2024-07-31 paper.tex
done with long edges

@@
 to triangulate the points and add edges. A more stable approach is to
-first hexagonly bin the data, reducing it from \(n\) to \(m\leq n\)
-observations, and connect the bin centroids. This process serves to
-reduce some noisiness in the resulting surface shown in \pD{}. The steps
-in this process are shown in Figure~\ref{fig-NLDR-scurve}, and
-documented below.
+first bin the data, reducing it from \(n\) to \(m\leq n\) observations,
+and connect the bin centroids. We recommend using a hexagon grid because
+it better reflects the data distribution and has less artifacts than a
+rectangular grid. This process serves to reduce some noisiness in the
+resulting surface shown in \pD{}. The steps in this process are shown in
+Figure~\ref{fig-NLDR-scurve}, and documented below.
 
@@
 Delaunay triangulation \citep{lee1980, alb2024} is used to connect
-neighboring centroids. The edges are needed to preserve neighborhood
-information when the model is lifted into \pD{}.
+points so that edges indicate neighbouring observations, in both the
+NLDR layout (Figure~\ref{fig-NLDR-scurve} (d)) and the \pD{} model
+representation. When the data has been binned the triangulation connectd
+centroids. The edges preserve the neighborhood information when the
+model is lifted into \pD{}.
 
 When shapes are non-linear in the NLDR layout, some edges could be long.
-It is unlikely
-
-It can also happen that distant centroids can be connected, which can
-result in long line segments. In order to generate a smooth surface in
-\gD{}, these long line segments must be removed
-(Figure~\ref{fig-NLDR-scurve} (d)).
+It can also happen that distant centroids can be connected, particularly
+if clustering is present, which can result in long line segments. In
+order to generate a smooth surface in \gD{}, these long line segments
+should be removed when tuning the model fit.
 
@@
 Figure~\ref{fig-scurve-true-sc} shows a wire frame of the true model
-that was used to generate the S-curve example data. The fitted model
-should come reasonably close to this, if the fitting has been
-successful. We don't ever have this, of course
-
-(Figure~\ref{fig-NLDR-scurve} (d)), it is necessary to remove edges that
-connect distant bin centroids in the triangular mesh. These edges only
-exist in the \gD{} model and do not extend into \pD{}, so their removal
-does not impact the model in \pD{}. Although there are no specific
-criteria for determining the benchmark value to remove long edges, the
-following steps provide an approach to identifying a suitable threshold:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Plot the distribution of the \gD{} Euclidean distances
-  (Figure~\ref{fig-param-scurve} (d)).
-\item
-  Identify the first largest difference between consecutive distance
-  values.
-\item
-  Take the distance value corresponding to this difference as the
-  benchmark value.
-\end{enumerate}
+that was used to generate the S-curve example data. The ideal is that
+the representation of the fitted model, at least for this example where
+we know the true model, should look similar to this.
+
+The Delaunay triangulation will ensure that all centroids are connected
+into a triangular mesh. For some structures, like clustered data, or
+highly non-linear shapes, breaks in the mesh are meaningful. When
+separated clusters are present the mesh should be broken across the
+gaps. For non-linear structures like the S-curve, the mesh should run
+unbroken along the S, but there should be no edges connecting the top of
+the S directly to the bottom of the S. For these reasons it is necessary
+to remove edges from the mesh in some applications.
+
+The decision on edge length removal is made based on the distribution of
+edge lengths. In particular, a gap between values, where there a
+concentration of small values and then a few larger values, likely
+suggests a cutoff for edge removal. Because the triangulation is
+typically done on the hexagon centroids, there are particular discrete
+edge lengths, based on bin widths. Figure~\ref{fig-NLDR-scurve} XXX
+illustrates edge length distributions.
+
+There is an additional step that is needed. When the model is lifted
+into \pD{}, if the fit is good all the edges should be relatively small
+in this space, too. If this is not the case, then there are several
+possible actions: (1) re-do the NLDR to get a more representative
+layout; (2) identify the edge and remove it from the model, in \gD{} and
+\pD{}; (3) consider different values for the model fit, number of bins,
+initial bin position or removing low density bins.
 

### 2024-07-31 paper.qmd

@@
                                 y = bin_counts,
-                                group = a1,
-                                color = a1)) + 
+                                group = as.factor(a1),
+                                color = as.factor(a1))) + 
   geom_line() + 
   # geom_hline(yintercept = benchmark1, color = "black", linewidth=0.5, alpha = 0.5) +
-  scale_colour_continuous_sequential(palette = "Magenta") +
+  scale_colour_discrete_qualitative(palette = "Dark 2") +
   xlab("hexagon id (re-ordered)") + 

### 2024-07-31 paper.qmd
Changes to Figure 8

--- 2024-07-31 paper.qmd
@@
 The choice of these values can have some effect on the distribution of bin counts.
-FIG XXX(a) illustrates this.
+@fig-param-scurve (a) illustrates this.
 The distribution of bin counts for $s_1$ varying between $-0.1-0.0$ is shown.
@@
 
+# mse_scurve_b <- ggplot(error_scurve, 
+#                      aes(x = b, 
+#                          y = log(MSE),
+#                          color = buff,
+#                          group = buff)) + 
+#   geom_point(size = 1) +
+#   geom_line(linewidth = 0.05) + 
+#   geom_vline(xintercept = 220, linetype="solid", 
+#              color = "black", linewidth=0.1) +
+#   geom_vline(xintercept = 112, linetype=2, 
+#              color = "black", linewidth=0.1) +
+#   geom_vline(xintercept = 364, linetype=2, 
+#              color = "black", linewidth=0.1) +
+#   scale_colour_continuous_sequential(palette = "Magenta") +
+#   labs(x = expression(b), y = "log(MSE)", color = expression(q)) +
+#   ggtitle("(a)") +
+#   theme(aspect.ratio = 0.75,
+#         axis.text.x = element_text(size = 12),
+#         axis.text.y = element_text(size = 12),
+#         axis.title.x = element_text(size = 12),
+#         axis.title.y = element_text(size = 12, 
+#             angle=90),
+#         plot.title = element_text(size = 12),
+#         legend.title = element_text(size = 12),
+#         legend.text = element_text(size = 10),
+#         legend.key.height = unit(1, 'cm'),
+#         legend.key.width = unit(1, 'cm'),
+#         legend.key.size = unit(1, 'cm'))
+
 mse_scurve_b <- ggplot(error_scurve, 
                      aes(x = b, 
-                         y = log(MSE),
-                         color = buff,
-                         group = buff)) + 
+                         y = MSE,
+                         color = as.factor(buff),
+                         group = as.factor(buff))) + 
   geom_point(size = 1) +
   geom_line(linewidth = 0.05) + 
-  geom_vline(xintercept = 220, linetype="solid", 
-             color = "black", linewidth=0.1) +
-  geom_vline(xintercept = 112, linetype=2, 
-             color = "black", linewidth=0.1) +
-  geom_vline(xintercept = 364, linetype=2, 
-             color = "black", linewidth=0.1) +
-  scale_colour_continuous_sequential(palette = "Magenta") +
-  labs(x = expression(b), y = "log(MSE)", color = expression(q)) +
-  ggtitle("(a)") +
+  # geom_vline(xintercept = 220, linetype="solid", 
+  #            color = "black", linewidth=0.1) +
+  # geom_vline(xintercept = 112, linetype=2, 
+  #            color = "black", linewidth=0.1) +
+  # geom_vline(xintercept = 364, linetype=2, 
+  #            color = "black", linewidth=0.1) +
+  scale_colour_discrete_qualitative(palette = "Dark 2") +
+  labs(x = expression(b), y = "MSE", color = expression(q)) +
+  ggtitle("(b)") +
   theme(aspect.ratio = 0.75,
@@
 Alternatively, one could choose to remove based on a percentile, the bins with density in the lowest 5% of all bins, for example.
-FIG XXXc illustrates the effect on the model representation of removing bins below different percentages.
+@fig-param-scurve (c) illustrates the effect on the model representation of removing bins below different percentages.
 Generally, we would urge caution in removing low count bins.
@@
                                 y = bin_counts,
-                                group = a1,
-                                color = a1)) + 
+                                group = as.factor(a1),
+                                color = as.factor(a1))) + 
   geom_line() + 
   # geom_hline(yintercept = benchmark1, color = "black", linewidth=0.5, alpha = 0.5) +
-  scale_colour_continuous_sequential(palette = "Magenta") +
+  scale_colour_discrete_qualitative(palette = "Dark 2") +
   xlab("hexagon id (re-ordered)") + 
-  ylab("standardized bin count") +
-  ggtitle("(b)") +
+  ylab("bin count") +
+  ggtitle("(a)") +
   theme(aspect.ratio = 0.75,
@@
 Because the triangulation is typically done on the hexagon centroids, there are particular discrete edge lengths, based on bin widths.
-@fig-NLDR-scurve XXX illustrates edge length distributions.
+@fig-NLDR-scurve (d) illustrates edge length distributions.
 
@@
 
-mse_scurve_b + cell_count_scurve + 
+cell_count_scurve + mse_scurve_b +  
 mse_scurve_lwd + 
@@
 ```
-
-XXXX I think we can remove the error figure, it's not part of the story
 

### 2024-07-31 paper.tex
Changes to Figure 8

@@
 where \(q\) is the buffer sound the data. The choice of these values can
-have some effect on the distribution of bin counts. FIG XXX(a)
-illustrates this. The distribution of bin counts for \(s_1\) varying
-between \(-0.1-0.0\) is shown. Generally, a more uniform distribution
-among these possibilities would indicate that the bins are reliably
-capturing the underlying distribution of observations.
+have some effect on the distribution of bin counts.
+Figure~\ref{fig-param-scurve} (a) illustrates this. The distribution of
+bin counts for \(s_1\) varying between \(-0.1-0.0\) is shown. Generally,
+a more uniform distribution among these possibilities would indicate
+that the bins are reliably capturing the underlying distribution of
+observations.
 
@@
 cutoff. Alternatively, one could choose to remove based on a percentile,
-the bins with density in the lowest 5\% of all bins, for example. FIG
-XXXc illustrates the effect on the model representation of removing bins
-below different percentages. Generally, we would urge caution in
-removing low count bins.
+the bins with density in the lowest 5\% of all bins, for example.
+Figure~\ref{fig-param-scurve} (c) illustrates the effect on the model
+representation of removing bins below different percentages. Generally,
+we would urge caution in removing low count bins.
 
@@
 typically done on the hexagon centroids, there are particular discrete
-edge lengths, based on bin widths. Figure~\ref{fig-NLDR-scurve} XXX
+edge lengths, based on bin widths. Figure~\ref{fig-NLDR-scurve} (d)
 illustrates edge length distributions.
@@
 patches (Figure~\ref{fig-pbmc1-sc}).
-
-XXXX I think we can remove the error figure, it's not part of the story
 

### 2024-07-31 paper.qmd
Updated Figure 8

@@
   filter(drop_empty == FALSE) |>
-  mutate(a1 = a1_1)
+  mutate(b1 = "b1 = 7")
 
@@
   filter(drop_empty == FALSE) |>
-  mutate(a1 = a1_2)
+  mutate(b1 = "b1 = 11")
 
@@
   filter(drop_empty == FALSE) |>
-  mutate(a1 = a1_3)
+  mutate(b1 = "b1 = 17")
 
@@
 Values of $b_1$ between $2$ and $b_1 = \frac{n}{2}$ are allowed.
-FIG XXX(b) shows the effect of different choices of $b_1$ on the MSE of the fitted model.
-XXX Are you showing $b_1$ or $b$ here????
-$b$
+@fig-param-scurve (b) shows the effect of different choices of $b_1$ on the MSE of the fitted model.
 
@@
 ## To initialize number of bins along the x-axis
-bin1_vec_scurve <- 2:19 #sqrt(NROW(training_data_scurve)/2)
-buffer_vec <- seq(0.05, 0.1, by = 0.01)
+bin1_vec_scurve <- 2:15 #sqrt(NROW(training_data_scurve)/2)
+#buffer_vec <- seq(0.05, 0.1, by = 0.01)
 
@@
 for (xbins in bin1_vec_scurve) {
-  for (buff in buffer_vec) {
-
-    bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = buff)$bin2
+  #for (buff in buffer_vec) {
+
+  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.07)$bin2
 
@@
     r2 = r2,
-      q = buff, 
+    q = 0.07, 
     is_bin_centroid = TRUE,
@@
            b = bin1 * bin2,
-             b_non_empty = NROW(df_bin_centroids_scurve),
-             buff = buff)
+           b_non_empty = NROW(df_bin_centroids_scurve))
 
@@
 
-  }
+  #}
 
@@
 mse_scurve_b <- ggplot(error_scurve, 
-                     aes(x = b, 
-                         y = MSE,
-                         color = as.factor(buff),
-                         group = as.factor(buff))) + 
+                     aes(x = bin1, 
+                         y = MSE)) + 
   geom_point(size = 1) +
   geom_line(linewidth = 0.05) + 
-  # geom_vline(xintercept = 220, linetype="solid", 
-  #            color = "black", linewidth=0.1) +
-  # geom_vline(xintercept = 112, linetype=2, 
-  #            color = "black", linewidth=0.1) +
-  # geom_vline(xintercept = 364, linetype=2, 
-  #            color = "black", linewidth=0.1) +
-  scale_colour_discrete_qualitative(palette = "Dark 2") +
-  labs(x = expression(b), y = "MSE", color = expression(q)) +
+  geom_vline(xintercept = 11, linetype="solid",
+             color = "#bdbdbd", linewidth=1) +
+  geom_vline(xintercept = 9, linetype=2,
+             color = "#bdbdbd", linewidth=1) +
+  geom_vline(xintercept = 13, linetype=2,
+             color = "#bdbdbd", linewidth=1) +
+  labs(x = expression(b[1]), y = "MSE") +
   ggtitle("(b)") +
@@
                                 y = bin_counts,
-                                group = as.factor(a1),
-                                color = as.factor(a1))) + 
+                                group = as.factor(b1),
+                                color = as.factor(b1))) + 
   geom_line() + 
   # geom_hline(yintercept = benchmark1, color = "black", linewidth=0.5, alpha = 0.5) +
-  scale_colour_discrete_qualitative(palette = "Dark 2") +
+  scale_colour_discrete_qualitative(palette = "Dark 3") +
   xlab("hexagon id (re-ordered)") + 
@@
         axis.text.y = element_text(size = 12),
-        plot.title = element_text(size = 12))
+        plot.title = element_text(size = 12),
+        legend.text = element_text(size = 10),
+        legend.key.size = unit(1, 'cm'),
+        legend.key.height = unit(1, 'cm'))
 ```
@@
                      aes(x = benchmark_rm_lwd, 
-                         y = log(MSE))) + 
+                         y = MSE)) + 
   geom_point(
@@
     ) + 
-  geom_vline(xintercept = benchmark1, linetype="solid", 
-             color = "black", linewidth=0.5, alpha = 0.5) +
+  # geom_vline(xintercept = benchmark1, linetype="solid", 
+  #            color = "#bdbdbd", linewidth=1, alpha = 0.5) +
   scale_x_continuous("standardized bin count", 
          transform = "reverse") + 
-  ylab("log(MSE)") +
+  ylab("MSE") +
   ggtitle("(c)") +

### 2024-07-31 paper.qmd
Update Figure 8 title.

@@
 #| echo: false
-#| fig-cap: "Various plots to help assess best bin start position, number of bins, low density bin and large edge removal.
-Both (a) and (c) show MSE, against number of bins and standardised count, while (a) also shows the change in MSE for different $q$.
+#| fig-cap: "Various plots to help assess best hexagon bin parameters, thresholds to remove low density bins and large edges.
+Both (b) and (c) show MSE, against number of bins along the x-axis and standardised count.
 A good benchmark value for these parameters is when the MSE drops and then flattens out.
-Plot (b) shows the distribution of stadardised counts of hexagons.
-Plot (d) shows the distribution of $2\\text{-}D$ Euclidean distances between bin centroids, with a good benchmark value for removing large edges would being the distance that shows the first large increase."
+Plot (a) shows the distribution of stadardised counts of hexagons.
+Plot (d) shows the distribution of $2\\text{-}D$ Euclidean distances between bin centroids, with a good benchmark value for removing large edges would being the distance that shows the first large decrease."
 #| label: fig-param-scurve

### 2024-07-31 paper.tex
Updated Figure 8
Update Figure 8 title.

@@
 determined analytically by \(b_1, q, r_2\). Values of \(b_1\) between
-\(2\) and \(b_1 = \frac{n}{2}\) are allowed. FIG XXX(b) shows the effect
-of different choices of \(b_1\) on the MSE of the fitted model. XXX Are
-you showing \(b_1\) or \(b\) here???? \(b\)
+\(2\) and \(b_1 = \frac{n}{2}\) are allowed.
+Figure~\ref{fig-param-scurve} (b) shows the effect of different choices
+of \(b_1\) on the MSE of the fitted model.
 
@@
 
-\caption{\label{fig-param-scurve}Various plots to help assess best bin
-start position, number of bins, low density bin and large edge removal.
-Both (a) and (c) show MSE, against number of bins and standardised
-count, while (a) also shows the change in MSE for different \(q\). A
-good benchmark value for these parameters is when the MSE drops and then
-flattens out. Plot (b) shows the distribution of stadardised counts of
-hexagons. Plot (d) shows the distribution of \(2\text{-}D\) Euclidean
-distances between bin centroids, with a good benchmark value for
-removing large edges would being the distance that shows the first large
-increase.}
+\caption{\label{fig-param-scurve}Various plots to help assess best
+hexagon bin parameters, thresholds to remove low density bins and large
+edges. Both (b) and (c) show MSE, against number of bins along the
+x-axis and standardised count. A good benchmark value for these
+parameters is when the MSE drops and then flattens out. Plot (a) shows
+the distribution of stadardised counts of hexagons. Plot (d) shows the
+distribution of \(2\text{-}D\) Euclidean distances between bin
+centroids, with a good benchmark value for removing large edges would
+being the distance that shows the first large decrease.}
 

### 2024-07-31 paper.qmd
Updated Figure 16 legend features

@@
 ## Add arrows
-arrow_df_pbmc <- tibble(x = c(94, 245, 101, 150),
+arrow_df_pbmc <- tibble(x = c(94, 245, 105, 150),
              xend = c(96, 263, 98, 180),
              y = c(10.06, 9.72, 9.88, 9.855),
-             yend = c(10.14, 9.8, 9.85, 10))
+             yend = c(10.14, 10, 9.85, 10))
 
@@
         color = "black") +
+  theme(#aspect.ratio = 0.75,
+        axis.text.x = element_text(size = 7),
+        axis.text.y = element_text(size = 7),
+        axis.title.x = element_text(size = 7),
+        axis.title.y = element_text(size = 7, angle = 90),
+        legend.position = "bottom",
+        legend.text = element_text(size = 7),
+        legend.key.height = unit(0.3, 'cm'),
+        legend.key.width = unit(0.3, 'cm'),
+        legend.key.size = unit(3, 'cm')) +
     inset_element(p = tsne_5_img, 
@@
                   right = 0, 
-                  top = 0.23) +
-  theme(#aspect.ratio = 0.75,
-        axis.text.x = element_text(size = 7),
-        axis.text.y = element_text(size = 7),
-        axis.title.x = element_text(size = 7),
-        axis.title.y = element_text(size = 7, angle = 90),
-        legend.position = "none",
-        legend.text = element_text(size = 7))
+                  top = 0.23)
 

### 2024-07-31 paper.qmd
repositioned tsne_30 img

--- 2024-07-31 paper.qmd
@@
 ## Add arrows
-arrow_df_pbmc <- tibble(x = c(94, 245, 101, 150),
-             xend = c(96, 263, 98, 180),
+arrow_df_pbmc <- tibble(x = c(94, 245, 105, 150),
+             xend = c(96, 263, 94, 180),
              y = c(10.06, 9.72, 9.88, 9.855),
-             yend = c(10.14, 9.8, 9.85, 10))
+             yend = c(10.14, 10, 9.79, 10))
 
@@
     inset_element(p = tsne_30_img, 
-                  left = 0.6, 
+                  left = 0.41, 
                   bottom = 0.03, 
                   right = 0, 
-                  top = 0.23) +
-  theme(#aspect.ratio = 0.75,
-        axis.text.x = element_text(size = 7),
-        axis.text.y = element_text(size = 7),
-        axis.title.x = element_text(size = 7),
-        axis.title.y = element_text(size = 7, angle = 90),
-        legend.position = "none",
-        legend.text = element_text(size = 7))
+                  top = 0.23)
 

### 2024-07-31 paper.qmd
Re-positioned phate_5 img

@@
                   left = 0.6, 
-                  bottom = 0.5, 
+                  bottom = 0.401, 
                   right = 0, 
-                  top = 0.65) +                   
+                  top = 0.56) +                   
     inset_element(p = umap_img, 

### 2024-07-31 paper.qmd
Repositioned umap img

--- 2024-07-31 paper.qmd
@@
                   left = 0.6, 
-                  bottom = 0.5, 
+                  bottom = 0.401, 
                   right = 0, 
-                  top = 0.65) +                   
+                  top = 0.56) +                   
     inset_element(p = umap_img, 
-                  left = 1.3, 
+                  left = 1.1, 
                   bottom = 0.3, 

### 2024-07-31 paper.qmd
Repositioned tsne_5 img

@@
 ## Add arrows
-arrow_df_pbmc <- tibble(x = c(94, 245, 105, 150),
-             xend = c(96, 263, 94, 180),
-             y = c(10.06, 9.72, 9.88, 9.855),
-             yend = c(10.14, 10, 9.79, 10))
+arrow_df_pbmc <- tibble(x = c(94, 273, 101, 150),
+             xend = c(96, 295, 94, 180),
+             y = c(10.06, 9.7, 9.88, 9.855),
+             yend = c(10.14, 9.8, 9.85, 10))
 

### 2024-07-31 paper.qmd
Saved the pbmc, tsne projections with the new function
Update pbmc tsne model

@@
 
+tr_from_to_df_pbmc <- tr_from_to_df_pbmc |>
+  filter(row_number() != 360)
+
 ## Compute 2D distances
@@
 # tr_from_to_df_pbmc |> mutate(ID = row_number()) |> filter(from == 37) |> filter(to == 43)
-tr_from_to_df_pbmc <- tr_from_to_df_pbmc |>
-  filter(row_number() != 360)
 
@@
   distance_col = "distance") +
+  geom_point(
+    data = tsne_pbmc_scaled,
+    aes(
+      x = tSNE1,
+      y = tSNE2
+    ),
+    alpha = 0.3,
+    size = 0.1,
+    color = "#6a3d9a"
+  ) +
+  theme_void() #+
   #xlim(sc_xlims) + ylim(sc_ylims) +
-  interior_annotation("a", sc_ltr_pos)
+  #interior_annotation("a", sc_ltr_pos)
 
@@
 
-<!--bin1 = 15, bin2 = 20, b = 300, non_empty = 136-->
+```{r}
+#| label: proj-tsne-pbmc
+
+## Hexagonal binning to have regular hexagons
+hb_obj_pbmc <- hex_binning(
+  data = tsne_pbmc_scaled,
+  bin1 = num_bins_x_pbmc,
+  r2 = r2_pbmc,
+  q = 0.05)
+
+tsne_data_with_hb_id <- hb_obj_pbmc$data_hb_id
+
+df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID),
+                                tsne_data_with_hb_id)
+
+### Define type column
+df <- df_all_pbmc |>
+  dplyr::select(tidyselect::starts_with("PC")) |>
+  dplyr::mutate(type = "data") ## original dataset
+
+df_b <- df_bin_pbmc |>
+  dplyr::filter(hb_id %in% df_bin_centroids_pbmc$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b <- df_b[match(df_bin_centroids_pbmc$hexID, df_b$hb_id),] |>
+  dplyr::select(-hb_id)
+
+df_exe <- dplyr::bind_rows(df_b, df)
+
+## Set the maximum difference as the criteria
+distance_df_small_edges <- distance_pbmc |>
+  dplyr::filter(distance < benchmark_pbmc)
+## Since erase brushing is considerd.
+
+## First projection
+projection <- cbind(
+  c(0.038104,-0.001148,0.003905,-0.003332,-0.011747,-0.002688,0.003468,-0.014321,-0.000320),
+  c(-0.000902,0.041414,0.005089,-0.000149,-0.000551,0.000239,-0.008013,-0.005799,-0.001411))
+
+proj1_pbmc_tsne <- gen_proj_langevitour(
+  points_df = df_exe,
+  projection = projection,
+  edge_df = distance_df_small_edges |> select(-distance)
+) + 
+  theme_void() + 
+  theme(legend.position = "none")
+
+## Second projection
+projection <- cbind(
+  c(0.013182,-0.022240,-0.001729,0.026946,-0.005289,0.001235,-0.007784,0.017500,-0.006949),
+  c(0.027138,0.011939,0.009374,-0.004582,-0.004771,-0.027107,-0.005287,-0.002325,-0.007951))
+
+proj2_pbmc_tsne <- gen_proj_langevitour(
+  points_df = df_exe,
+  projection = projection,
+  edge_df = distance_df_small_edges |> select(-distance)
+) + 
+  theme_void() + 
+  theme(legend.position = "none")
+
+## Third projection
+projection <- cbind(
+  c(-0.003540,0.025008,0.023261,-0.012782,-0.011617,-0.006796,0.009566,-0.015022,0.001164),
+  c(0.038201,-0.007751,0.005064,-0.014143,0.003255,-0.001091,0.001611,-0.003674,-0.008314))
+
+proj3_pbmc_tsne <- gen_proj_langevitour(
+  points_df = df_exe,
+  projection = projection,
+  edge_df = distance_df_small_edges |> select(-distance)
+) + 
+  theme_void() + 
+  theme(legend.position = "none")
+
+```
+
+<!-- <!--bin1 = 15, bin2 = 20, b = 300, non_empty = 136--> 
+<!-- ```{r} -->
+<!-- #| echo: false -->
+<!-- #| fig-cap: "(a) Model generated in $2\\text{-}D$ with tSNE, and (b) $p\\text{-}D$ model error in $2\\text{-}D$.
+The $2\\text{-}D$ model shows three well-separated distant clusters.
+The $p\\text{-}D$ model errors are distributed along clusters, but most low $p\\text{-}D$ model errors present in the large cluster." -->
+<!-- #| label: fig-model-pbmc -->
+<!-- #| fig-pos: H -->
+<!-- #| out-height: 30% -->
+
+<!-- trimesh_removed_pbmc + error_plot_pbmc +  -->
+<!--   plot_layout(guides='collect', ncol=2) & -->
+<!--   theme(legend.position='bottom') -->
+<!-- ``` -->
+
 ```{r}
 #| echo: false
-#| fig-cap: "(a) Model generated in $2\\text{-}D$ with tSNE, and (b) $p\\text{-}D$ model error in $2\\text{-}D$.
-The $2\\text{-}D$ model shows three well-separated distant clusters.
-The $p\\text{-}D$ model errors are distributed along clusters, but most low $p\\text{-}D$ model errors present in the large cluster."
-#| label: fig-model-pbmc
+#| fig-cap: "Model in $2\\text{-}D$, on the tSNE layout, and three views of the fit in projections from $9\text{-}D$, for the PBMC3k data ($(s_1, \ s_2) = (-0.050, \ -0.058)$, $b = 300 \  (15, \ 20)$, $m = 136$, benchmark value to remove large edges is $0.133$).
+(The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/5Y1hE4i7N2k>)."
+#| label: fig-pbmc2-sc
 #| fig-pos: H
-#| out-height: 30%
-
-trimesh_removed_pbmc + error_plot_pbmc + 
+#| out-height: 100%
+#| eval: false
+
+trimesh_removed_pbmc + proj1_pbmc_tsne + 
+  proj2_pbmc_tsne + proj3_pbmc_tsne +
   plot_layout(guides='collect', ncol=2) &
-  theme(legend.position='bottom')
-```
-
-::: {#fig-pbmc2-sc layout-ncol="4" fig-pos="H"}
+  theme(legend.position='none')
+```
+
+::: {#fig-pbmc2-sc layout-ncol="2" fig-pos="H"}
 ![](figures/pbmc3k/tsne_trimesh_plot.png)

### 2024-07-31 paper.tex
Saved the pbmc, tsne projections with the new function
Update pbmc tsne model

@@
 
-\centering{
-
-\includegraphics[width=1\textwidth,height=0.3\textheight]{paper_files/figure-pdf/fig-model-pbmc-1.pdf}
-
-}
-
-\caption{\label{fig-model-pbmc}(a) Model generated in \(2\text{-}D\)
-with tSNE, and (b) \(p\text{-}D\) model error in \(2\text{-}D\). The
-\(2\text{-}D\) model shows three well-separated distant clusters. The
-\(p\text{-}D\) model errors are distributed along clusters, but most low
-\(p\text{-}D\) model errors present in the large cluster.}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\begin{minipage}{0.25\linewidth}
+\begin{minipage}{0.50\linewidth}
 \includegraphics{figures/pbmc3k/tsne_trimesh_plot.png}\end{minipage}%
 %
-\begin{minipage}{0.25\linewidth}
+\begin{minipage}{0.50\linewidth}
 \includegraphics{figures/pbmc3k/sc_4.png}\end{minipage}%
-%
-\begin{minipage}{0.25\linewidth}
+\newline
+\begin{minipage}{0.50\linewidth}
 \includegraphics{figures/pbmc3k/sc_5.png}\end{minipage}%
 %
-\begin{minipage}{0.25\linewidth}
+\begin{minipage}{0.50\linewidth}
 \includegraphics{figures/pbmc3k/sc_6.png}\end{minipage}%

### 2024-08-04 paper.qmd
Updated figure: pbmc error plot

@@
 tsne_30_img <- readPNG("figures/pbmc3k/tsne_30.png", native = TRUE)
+umap_15_img <- readPNG("figures/pbmc3k/umap_15.png", native = TRUE) 
+pacmap_img <- readPNG("figures/pbmc3k/pacmap.png", native = TRUE) 
+trimap_img <- readPNG("figures/pbmc3k/trimap.png", native = TRUE)
+umap_5_img <- readPNG("figures/pbmc3k/umap_5.png", native = TRUE)
 
 ## Add arrows
-arrow_df_pbmc <- tibble(x = c(94, 273, 101, 150),
-             xend = c(96, 295, 94, 180),
-             y = c(10.06, 9.7, 9.88, 9.855),
-             yend = c(10.14, 9.8, 9.85, 10))
+arrow_df_pbmc <- tibble(x = c(94, 280, 101, 150, 218, 250, 50),
+             xend = c(96, 295, 96, 180, 250, 290, 25),
+             y = c(10.06, 9.69, 9.88, 9.855, 9.82, 9.78, 10),
+             yend = c(10.14, 9.8, 9.8, 10, 10.3, 10, 9.8))
 
@@
     inset_element(p = tsne_5_img, 
-                  left = 1.7, 
-                  bottom = 0.16, 
+                  left = 1.8, 
+                  bottom = 0.08, 
                   right = 0, 
-                  top = 0.31) +                   
+                  top = 0.25) +                   
     inset_element(p = phate_5_img, 
                   left = 0.6, 
-                  bottom = 0.401, 
+                  bottom = 0.4, 
                   right = 0, 
-                  top = 0.56) +                   
+                  top = 0.55) +                   
     inset_element(p = umap_img, 
-                  left = 1.1, 
+                  left = 1.2, 
+                  bottom = 0.3, 
+                  right = -0.1, 
+                  top = 0.5) +
+    inset_element(p = umap_15_img, 
+                  left = 1.5, 
+                  bottom = 0.5, 
+                  right = 0, 
+                  top = 0.7) +
+    inset_element(p = tsne_30_img, 
+                  left = 0.6, 
+                  bottom = 0.01, 
+                  right = 0, 
+                  top = 0.18) +
+    inset_element(p = pacmap_img, 
+                  left = 1.7, 
                   bottom = 0.3, 
@@
                   top = 0.5) +
-    inset_element(p = tsne_30_img, 
-                  left = 0.41, 
-                  bottom = 0.03, 
-                  right = 0, 
-                  top = 0.23)
+    inset_element(p = umap_5_img, 
+                  left = 0.4, 
+                  bottom = 0.05, 
+                  right = -0.2, 
+                  top = 0.22)
 

### 2024-08-04 paper.qmd
update fig16

@@
         ends = "first"), 
-        color = "black") +
+        color = "black",
+      linewidth = 0.1) +
   theme(#aspect.ratio = 0.75,

### 2024-08-13 paper.qmd
Update center to centroid
Need to fix the prediction fig

@@
 A new benefit of this fitted model is that it allows us to now predict a new observation's value in the NLDR, for any method.
-The steps are to determine the closest bin center in \pD{}, $C^{(p)}_{h}$ and predict it to be the centroid of this bin in \gD{}, $C^{(2)}_{h}$.
+The steps are to determine the closest bin centroid in \pD{}, $C^{(p)}_{h}$ and predict it to be the centroid of this bin in \gD{}, $C^{(2)}_{h}$.
 This can be written as, let $z(i) = \arg\min_{j = 1, \dots, b} \sqrt{\sum_{v=1}^{p}(x_{iv} - C^{(p)}_{jv})^2}$, then the new observation $i$ falls in the hexagon, $H_h = \{i| z(i) = h\}$ and the corresponding \kD{} bin centroids, $C_{h}^{(2)} = (c_{h1}, c_{h2})$.
@@
 Therefore, the new observation $i$ falls in the hexagon, $H_h = \{i| z(i) = h\}$ and the corresponding \kD{} bin centroids, $C_{h}^{(2)} = (c_{h1}, c_{h2})$ be the predicted values.-->
+
+```{r}
+#| label: scurve-training
+training_data_scurve <- read_rds("data/s_curve/s_curve_training.rds")
+```
+
+```{r}
+#| label: model-prediction-umap-original
+
+predict_umap_scurve <- read_rds(file = "data/s_curve/s_curve_umap_predict_true.rds")
+
+predict_scurve_obj <- gen_scaled_data(
+    data = predict_umap_scurve)
+
+predict_umap_scurve_scaled <- predict_scurve_obj$scaled_nldr
+
+plot_predict_umap <- ggplot(data = predict_umap_scurve_scaled,
+                            aes(
+                              x = UMAP1,
+                              y = UMAP2
+                            )) +
+  geom_point()
+```
+
+```{r}
+#| label: model-prediction-umap-quollr
+
+true_model_df <- read_rds("data/s_curve/scurve_true_model.rds")
+
+# Initialize effective bins along x
+effective_bin1_scurve <- 15
+
+# Fit high-dimensional model
+scurve_model <- fit_highd_model(
+  training_data = training_data_scurve,
+  emb_df = umap_scurve_scaled,
+  bin1 = effective_bin1_scurve,
+  r2 = r2,
+  q = 0.16,
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = FALSE,
+  col_start_highd = "x"
+)
+
+df_bin_centroids_scurve <- scurve_model$df_bin_centroids
+df_bin_scurve <- scurve_model$df_bin
+
+true_pred_df <- predict_emb(
+  test_data = true_model_df,
+  df_bin_centroids = df_bin_centroids_scurve,
+  df_bin = df_bin_scurve,
+  type_NLDR = "UMAP"
+)
+
+# Compute radius r
+a1 <- calc_bins_y(bin1 = 15, r2 = r2, q = 0.16)$a1
+r <- a1 / 2
+
+# Function to jitter points within a circumcircle of a hexagon
+jitter_within_circumcircle <- function(center_x, center_y, radius, num_points) {
+  theta <- runif(num_points, 0, 2 * pi)
+  r <- radius * sqrt(runif(num_points))
+  x <- center_x + r * cos(theta)
+  y <- center_y + r * sin(theta)
+  jittered_points <- data.frame(UMAP1 = x, UMAP2 = y)
+  return(jittered_points)
+}
+
+# Jittered points data frame
+jittered_points_df <- data.frame()
+
+# Iterate through each hexagon and jitter points within the circumcircle
+for (hex_id in unique(hex_grid_nonempty$hex_poly_id)) {
+  hex_points <- hex_grid_nonempty %>% filter(hex_poly_id == hex_id)
+  center_x <- mean(hex_points$x)
+  center_y <- mean(hex_points$y)
+  points_in_hex <- true_pred_df %>% filter(pred_UMAP_1 >= min(hex_points$x) & pred_UMAP_1 <= max(hex_points$x) & pred_UMAP_2 >= min(hex_points$y) & pred_UMAP_2 <= max(hex_points$y))
+
+  if (nrow(points_in_hex) > 0) {
+    jittered_points <- jitter_within_circumcircle(center_x, center_y, r, nrow(points_in_hex))
+    jittered_points_df <- rbind(jittered_points_df, jittered_points)
+  }
+}
+
+plot_predict_umap_model <- ggplot(data = jittered_points_df,
+                            aes(
+                              x = UMAP1,
+                              y = UMAP2
+                            )) +
+  geom_point()
+```
+
+```{r}
+#| label: model-prediction-points-line
+
+predict_umap_scurve_scaled <- predict_umap_scurve_scaled |>
+  mutate(type = "predict_from_umap") |>
+  mutate(point_id = row_number())
+
+jittered_points_df <- jittered_points_df |>
+  mutate(type = "predict_from_model") |>
+  mutate(point_id = row_number())
+
+predict_point_df <- bind_rows(
+  predict_umap_scurve_scaled,
+  jittered_points_df
+)
+
+plot_predict_umap_model_connect <- ggplot(data = predict_point_df,
+                            aes(
+                              x = UMAP1,
+                              y = UMAP2, 
+                              group = point_id
+                            )) +
+  geom_point() +
+  geom_line()
+```
+
+```{r}
+#| echo: false
+#| label: fig-predict-scurve
+#| fig-pos: H
+#| fig-width: 9
+#| fig-height: 5
+plot_predict_umap + plot_predict_umap_model + plot_predict_umap_model_connect 
+```
 
@@
 <!--add MSE vs total number of error plot-->
-
-```{r}
-#| label: scurve-training
-training_data_scurve <- read_rds("data/s_curve/s_curve_training.rds")
-```
 

### 2024-08-13 paper.tex
Update center to centroid
Need to fix the prediction fig

@@
 new observation's value in the NLDR, for any method. The steps are to
-determine the closest bin center in \pD{}, \(C^{(p)}_{h}\) and predict
+determine the closest bin centroid in \pD{}, \(C^{(p)}_{h}\) and predict
 it to be the centroid of this bin in \gD{}, \(C^{(2)}_{h}\). This can be
@@
 \(C_{h}^{(2)} = (c_{h1}, c_{h2})\).
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-predict-scurve-1.pdf}
+
+}
+
+\caption{\label{fig-predict-scurve}}
+
+\end{figure}%
 

### 2024-08-14 paper.qmd
Updated Figure 6

@@
                             )) +
-  geom_point()
+  geom_point(alpha = 0.5)
 ```
@@
 
+hex_grid_nonempty <- hex_grid |>
+  filter(hex_poly_id %in% df_bin_centroids_scurve$hexID)
+
 true_pred_df <- predict_emb(
@@
   y <- center_y + r * sin(theta)
-  jittered_points <- data.frame(UMAP1 = x, UMAP2 = y)
+  jittered_points <- data.frame(UMAP1 = x, UMAP2 = y, ID = points_in_hex$ID)
   return(jittered_points)
@@
 # Iterate through each hexagon and jitter points within the circumcircle
-for (hex_id in unique(hex_grid_nonempty$hex_poly_id)) {
-  hex_points <- hex_grid_nonempty %>% filter(hex_poly_id == hex_id)
-  center_x <- mean(hex_points$x)
-  center_y <- mean(hex_points$y)
-  points_in_hex <- true_pred_df %>% filter(pred_UMAP_1 >= min(hex_points$x) & pred_UMAP_1 <= max(hex_points$x) & pred_UMAP_2 >= min(hex_points$y) & pred_UMAP_2 <= max(hex_points$y))
+for (hex_id in unique(true_pred_df$pred_hb_id)) {
+  # hex_points <- hex_grid_nonempty %>% filter(hex_poly_id == hex_id)
+  # center_x <- mean(hex_points$x)
+  # center_y <- mean(hex_points$y)
+  # points_in_hex <- true_pred_df %>% filter(pred_UMAP_1 >= min(hex_points$x) & pred_UMAP_1 <= max(hex_points$x) & pred_UMAP_2 >= min(hex_points$y) & pred_UMAP_2 <= max(hex_points$y))
+
+  points_in_hex <- true_pred_df |>
+    filter(pred_hb_id == hex_id)
+
+  center_x <- points_in_hex |> 
+    pull(pred_UMAP_1) |> 
+    unique()
+
+  center_y <- points_in_hex |> 
+    pull(pred_UMAP_2) |> 
+    unique()
 
@@
 predict_umap_scurve_scaled <- predict_umap_scurve_scaled |>
-  mutate(type = "predict_from_umap") |>
-  mutate(point_id = row_number())
+  mutate(type = "predict_from_umap")
 
 jittered_points_df <- jittered_points_df |>
-  mutate(type = "predict_from_model") |>
-  mutate(point_id = row_number())
+  mutate(type = "predict_from_model")
 
@@
                               y = UMAP2, 
-                              group = point_id
+                              group = ID
                             )) +
-  geom_point() +
-  geom_line()
+  geom_point(alpha = 0.6) +
+  geom_line(alpha = 0.3)
 ```

### 2024-08-14 paper.qmd
Update Figure 6

@@
 
+<!--prediction for true-model of S-curve-->
 ```{r}
@@
 ```{r}
+#| label: true-model-projection-langevitour
+
+true_model_df_rm_id <- true_model_df |>
+  select(-ID)
+
+projection <- cbind(
+    c(0.18390,-0.18443,-0.06860,0.00654,-0.02921,0.02313,-0.08963),
+    c(-0.02902,0.03638,-0.16872,0.22418,0.01146,0.02777,0.01453))
+projected_df <- as.matrix(true_model_df_rm_id) %*% projection |> 
+  as_tibble(.name_repair = "unique") |> 
+  rename(c("proj1" = "...1", 
+           "proj2" = "...2"))
+
+proj_plot_scurve_true <- ggplot(
+  data = projected_df, 
+  aes(
+    x = proj1, 
+    y = proj2)) +
+    geom_point(alpha = 0.5)   
+```
+
+```{r}
 #| label: model-prediction-points-line
@@
                             )) +
-  geom_point(alpha = 0.6) +
+  geom_point(alpha = 0.5) +
   geom_line(alpha = 0.3)
@@
 #| fig-height: 5
-plot_predict_umap + plot_predict_umap_model + plot_predict_umap_model_connect 
-```
-
-<!--prediction for true-model of S-curve-->
-::: {#fig-scurve-pred-sc layout-ncol="2" fig-pos="H"}
-![](figures/scurve/sc_true_only.png){width="150" fig-align="center"}
-
-![](figures/scurve/pred_true_view.png){width="150" fig-align="center"}
-
-A view of the true model in projections from $7\text{-}D$, and predictions of the true model in \gD{}, for the S-curve data.
+proj_plot_scurve_true + plot_predict_umap + 
+  plot_predict_umap_model + plot_predict_umap_model_connect +
+  plot_layout(ncol = 4)
+```
+
+<!-- ::: {#fig-scurve-pred-sc layout-ncol="2" fig-pos="H"} -->
+<!-- ![](figures/scurve/sc_true_only.png){width="150" fig-align="center"} -->
+
+<!-- ![](figures/scurve/pred_true_view.png){width="150" fig-align="center"} -->
+
+<!-- A view of the true model in projections from $7\text{-}D$, and predictions of the true model in \gD{}, for the S-curve data.
 The predictions fits the UMAP layout which means that it capture the geometry of S-curve with UMAP.
-:::
+-->
+<!-- ::: -->
 

### 2024-08-14 paper.tex
Update Figure 6

@@
 
-\begin{figure}[H]
-
-\begin{minipage}{0.50\linewidth}
-\begin{center}
-\includegraphics[width=1.5625in,height=\textheight]{figures/scurve/sc_true_only.png}
-\end{center}
-\end{minipage}%
-%
-\begin{minipage}{0.50\linewidth}
-\begin{center}
-\includegraphics[width=1.5625in,height=\textheight]{figures/scurve/pred_true_view.png}
-\end{center}
-\end{minipage}%
-
-\caption{\label{fig-scurve-pred-sc}A view of the true model in
-projections from \(7\text{-}D\), and predictions of the true model in
-\gD{}, for the S-curve data. The predictions fits the UMAP layout which
-means that it capture the geometry of S-curve with UMAP.}
-
-\end{figure}%
-
 \subsection{Tuning}\label{tuning}

### 2024-08-14 paper.qmd
Added draft title for Figure 6

@@
 #| fig-height: 5
+#| fig-cap: "Comparison of prediction generated using the exiting `umap` prediction method and our method: (a) A view of the true model in projections from $7\\text{-}D$, (b) predicted data from the `umap` prediction method, (c) predicted data from our method, and (d) comparison of predictions from both methods.
+In plot (d), points representing the same data but predicted by different methods are connected by lines.
+Need to add concluded sentence after changing the data."
+
 proj_plot_scurve_true + plot_predict_umap + 

### 2024-08-14 paper.tex
Added draft title for Figure 6

@@
 
-\caption{\label{fig-predict-scurve}}
+\caption{\label{fig-predict-scurve}Comparison of prediction generated
+using the exiting \texttt{umap} prediction method and our method: (a) A
+view of the true model in projections from \(7\text{-}D\), (b) predicted
+data from the \texttt{umap} prediction method, (c) predicted data from
+our method, and (d) comparison of predictions from both methods. In plot
+(d), points representing the same data but predicted by different
+methods are connected by lines. Need to add concluded sentence after
+changing the data.}
 

### 2024-08-14 paper.qmd
Replaced Six to Eight in Figure 1 title
Added axes to the projection: Need to check with the origin

@@
 #| echo: false
-#| fig-cap: "Six different NLDR representations of the same data.
+#| fig-cap: "Eight different NLDR representations of the same data.
 Different techniques and different parameter choices are used.
@@
 
+<!--to draw axes-->
+```{r}
+
+## Draw the axes
+
+projection <- cbind(
+  c(0.18390,-0.18443,-0.06860,0.00654,-0.02921,0.02313,-0.08963),
+  c(-0.02902,0.03638,-0.16872,0.22418,0.01146,0.02777,0.01453))
+
+projected_df <- as.matrix(true_model_df_rm_id) %*% projection
+
+limits <- 1
+rng <- range(projected_df)
+projected_df <- projected_df/max(abs(rng))
+colnames(projected_df) <- c("P1", "P2")
+
+obs_labels <- as.character(1:nrow(true_model_df_rm_id))
+
+projected_df_n <- as.matrix(true_model_df_rm_id) %*% projection |> 
+  as_tibble(.name_repair = "unique") |> 
+  rename(c("proj1" = "...1", 
+           "proj2" = "...2"))
+
+projected_df$obs_labels <- obs_labels
+
+#if (position == "center") {
+  axis_scale <- 2 * limits/3
+  #axis_pos <- 0
+  axis_pos <- -0.6 * limits
+#}
+# else if (position == "bottomleft") {
+#   axis_scale <- limits/6
+#   axis_pos <- -2/3 * limits
+# }
+
+mean_cols <- colMeans(true_model_df_rm_id, na.rm = TRUE)
+
+adj <- function(x) axis_pos + x * axis_scale
+axes <- data.frame(x1 = adj(0), 
+                   y1 = adj(0), 
+                   x2 = adj(projection[, 1]), 
+                   y2 = adj(projection[, 2]))
+
+# axes <- data.frame(x1 = mean_cols + mean_cols * axis_scale, 
+#                    y1 = mean_cols + mean_cols * axis_scale, 
+#                    x2 = mean_cols + projection[, 1] * axis_scale, 
+#                    y2 = mean_cols + projection[, 2]* axis_scale)
+
+axis_labels <- colnames(true_model_df_rm_id)
+rownames(axes) <- axis_labels
+theta <- seq(0, 2 * pi, length = 50)
+circle <- data.frame(c1 = adj(cos(theta)), c2 = adj(sin(theta)))
+
+ggplot() +
+  geom_point(data = projected_df_n, 
+             aes(
+               x = proj1, 
+               y = proj2),
+             alpha = 0.5)  +
+  geom_segment(data=axes, aes(x=x1, y=y1, xend=x2, yend=y2), colour="grey70") +
+  geom_text(data=axes, aes(x=x2, y=y2, label=rownames(axes)), colour="grey50")
+
+```
+
 <!-- ::: {#fig-scurve-pred-sc layout-ncol="2" fig-pos="H"} -->

### 2024-08-14 paper.qmd
Updated NLDR data point colours

@@
   ggplot(aes(x = UMAP1, y = UMAP2)) + 
-  geom_point(alpha=0.5, colour="#6a3d9a", size = 0.5) +
+  geom_point(alpha=0.5, colour="#000000", size = 0.5) +
   xlim(sc_xlims) + ylim(sc_ylims) +
@@
              aes(x = UMAP1, y = UMAP2), 
-             alpha = 0.5, size = 0.5, color = "#6a3d9a") +
+             alpha = 0.5, size = 0.5, color = "#000000") +
   xlim(sc_xlims) + ylim(sc_ylims) +
@@
              y = UMAP2)) +
-  geom_point(alpha=0.3, color = "#6a3d9a") 
+  geom_point(alpha=0.3, color = "#000000") 
 ```
@@
              y = tSNE2)) +
-  geom_point(alpha=0.3, color = "#6a3d9a")
+  geom_point(alpha=0.3, color = "#000000")
 ```
@@
     size = 0.1,
-    color = "#6a3d9a"
+    color = "#000000"
   ) +
@@
              y = PaCMAP2)) + 
-  geom_point(alpha=0.1, color = "#6a3d9a")
+  geom_point(alpha=0.1, color = "#000000")
 

### 2024-08-14 paper.qmd
Added high-d model error in 2D s-curve

@@
 
-### Prediction into \gD{}
-
-A new benefit of this fitted model is that it allows us to now predict a new observation's value in the NLDR, for any method.
-The steps are to determine the closest bin centroid in \pD{}, $C^{(p)}_{h}$ and predict it to be the centroid of this bin in \gD{}, $C^{(2)}_{h}$.
-This can be written as, let $z(i) = \arg\min_{j = 1, \dots, b} \sqrt{\sum_{v=1}^{p}(x_{iv} - C^{(p)}_{jv})^2}$, then the new observation $i$ falls in the hexagon, $H_h = \{i| z(i) = h\}$ and the corresponding \kD{} bin centroids, $C_{h}^{(2)} = (c_{h1}, c_{h2})$.
-
-<!--The prediction approach involves finding the nearest \kD{} model point for a new \pD{} point.
-We define the function $z: \mathbb{R}^{n\times p} \rightarrow \mathbb{R}^{m\times p}$, where $z(i) = \arg\min_{j = 1, \dots, b} \sqrt{\sum_{v=1}^{p}(x_{iv} - C^{(p)}_{jv})^2}$ maps each \pD{} point to its nearest \pD{} mapping of the model.
-Therefore, the new observation $i$ falls in the hexagon, $H_h = \{i| z(i) = h\}$ and the corresponding \kD{} bin centroids, $C_{h}^{(2)} = (c_{h1}, c_{h2})$ be the predicted values.-->
-
 ```{r}
@@
 
-<!--prediction for true-model of S-curve-->
-```{r}
-#| label: model-prediction-umap-original
-
-predict_umap_scurve <- read_rds(file = "data/s_curve/s_curve_umap_predict_true.rds")
-
-predict_scurve_obj <- gen_scaled_data(
-    data = predict_umap_scurve)
-
-predict_umap_scurve_scaled <- predict_scurve_obj$scaled_nldr
-
-plot_predict_umap <- ggplot(data = predict_umap_scurve_scaled,
-                            aes(
-                              x = UMAP1,
-                              y = UMAP2
-                            )) +
-  geom_point(alpha = 0.5)
-```
-
-```{r}
-#| label: model-prediction-umap-quollr
-
-true_model_df <- read_rds("data/s_curve/scurve_true_model.rds")
+```{r}
+#| label: best-umap-model-scurve
 
@@
   filter(hex_poly_id %in% df_bin_centroids_scurve$hexID)
+
+```
+
+```{r}
+#| label: p-d-error-in-2d-scurve
+
+## Compute error
+error_df_scurve <- augment(
+  df_bin_centroids = df_bin_centroids_scurve,
+  df_bin = df_bin_scurve,
+  training_data = training_data_scurve,
+  newdata = NULL,
+  type_NLDR = "UMAP",
+  col_start = "x")
+
+error_df_scurve <- error_df_scurve |>
+  mutate(type = case_when(
+    row_wise_abs_error == 0 ~ "no error",
+    row_wise_abs_error <= 0.2 ~ "error 0-0.2",
+    row_wise_abs_error <= 0.4 ~ "error 0.2-0.4",
+    row_wise_abs_error <= 0.6 ~ "error 0.4-0.6",
+    .default = "error greter than 0.6"
+  )) |>
+  mutate(type = factor(type, levels = c(
+    "no error", "error 0-0.2", "error 0.2-0.4", "error 0.4-0.6", "error greter than 0.6")))
+
+error_df_scurve <- error_df_scurve |>
+  bind_cols(umap_scurve_scaled |>
+              select(-ID))
+
+error_plot_scurve <- error_df_scurve |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2,
+             color = type)) +
+  geom_point(alpha=0.5) +
+  theme(
+    legend.position = "bottom"
+  )
+
+error_plot_scurve 
+```
+
+### Prediction into \gD{}
+
+A new benefit of this fitted model is that it allows us to now predict a new observation's value in the NLDR, for any method.
+The steps are to determine the closest bin centroid in \pD{}, $C^{(p)}_{h}$ and predict it to be the centroid of this bin in \gD{}, $C^{(2)}_{h}$.
+This can be written as, let $z(i) = \arg\min_{j = 1, \dots, b} \sqrt{\sum_{v=1}^{p}(x_{iv} - C^{(p)}_{jv})^2}$, then the new observation $i$ falls in the hexagon, $H_h = \{i| z(i) = h\}$ and the corresponding \kD{} bin centroids, $C_{h}^{(2)} = (c_{h1}, c_{h2})$.
+
+<!--The prediction approach involves finding the nearest \kD{} model point for a new \pD{} point.
+We define the function $z: \mathbb{R}^{n\times p} \rightarrow \mathbb{R}^{m\times p}$, where $z(i) = \arg\min_{j = 1, \dots, b} \sqrt{\sum_{v=1}^{p}(x_{iv} - C^{(p)}_{jv})^2}$ maps each \pD{} point to its nearest \pD{} mapping of the model.
+Therefore, the new observation $i$ falls in the hexagon, $H_h = \{i| z(i) = h\}$ and the corresponding \kD{} bin centroids, $C_{h}^{(2)} = (c_{h1}, c_{h2})$ be the predicted values.-->
+
+<!--prediction for true-model of S-curve-->
+```{r}
+#| label: model-prediction-umap-original
+
+predict_umap_scurve <- read_rds(file = "data/s_curve/s_curve_umap_predict_true.rds")
+
+predict_scurve_obj <- gen_scaled_data(
+    data = predict_umap_scurve)
+
+predict_umap_scurve_scaled <- predict_scurve_obj$scaled_nldr
+
+plot_predict_umap <- ggplot(data = predict_umap_scurve_scaled,
+                            aes(
+                              x = UMAP1,
+                              y = UMAP2
+                            )) +
+  geom_point(alpha = 0.5)
+```
+
+```{r}
+#| label: model-prediction-umap-quollr
+
+true_model_df <- read_rds("data/s_curve/scurve_true_model.rds")
 

### 2024-08-14 paper.tex
Replaced Six to Eight in Figure 1 title
Added axes to the projection: Need to check with the origin
Added high-d model error in 2D s-curve

@@
 
-\caption{\label{fig-NLDR-variety}Six different NLDR representations of
+\caption{\label{fig-NLDR-variety}Eight different NLDR representations of
 the same data. Different techniques and different parameter choices are
@@
 
+\begin{center}
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/p-d-error-in-2d-scurve-1.pdf}
+\end{center}
+
 \subsection{\texorpdfstring{Prediction into
@@
 \end{figure}%
+
+\begin{center}
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-23-1.pdf}
+\end{center}
 

### 2024-08-14 paper.qmd
Added trimesh of PaCMAP

@@
   distance_col = "distance") +
-  #xlim(sc_xlims) + ylim(sc_ylims) +
+  geom_point(data = pacmap_minst_scaled,
+             aes(
+               x = PaCMAP1,
+               y = PaCMAP2
+             ),
+             alpha=0.5, 
+             size = 0.1) +
   interior_annotation("a", sc_ltr_pos_mnist)
@@
 <!--add langevitour screenshots and youtube animation link-->
+```{r}
+trimesh_removed_mnist 
+```
 

### 2024-08-14 paper.qmd
Changed the point color to black

@@
   distance_col = "distance") +
-  #xlim(sc_xlims) + ylim(sc_ylims) +
-  interior_annotation("a", sc_ltr_pos)
+  geom_point(
+    data = umap_pbmc_scaled,
+    aes(
+      x = UMAP1,
+      y = UMAP2
+    ),
+    alpha = 0.3,
+    size = 0.1,
+    color = "#000000"
+  )
 
@@
     ),
-    alpha = 0.3,
-    size = 0.1,
+    alpha = 0.2,
     color = "#000000"
-  ) +
-  theme_void() #+
+  ) 
   #xlim(sc_xlims) + ylim(sc_ylims) +
@@
   edge_df = distance_df_small_edges |> select(-distance)
-) + 
-  theme_void() + 
-  theme(legend.position = "none")
+)
 
@@
 #| out-height: 100%
-#| eval: false
+#| eval: true
 
@@
 
-::: {#fig-pbmc2-sc layout-ncol="2" fig-pos="H"}
-![](figures/pbmc3k/tsne_trimesh_plot.png)
-
-![](figures/pbmc3k/sc_4.png)
-
-![](figures/pbmc3k/sc_5.png)
-
-![](figures/pbmc3k/sc_6.png)
-
-Model in \gD{}, on the tSNE layout, and three views of the fit in projections from $9\text{-}D$, for the PBMC3k data ($(s_1, \ s_2) = (-0.050, \ -0.058)$, $b = 300 \  (15, \ 20)$, $m = 136$, benchmark value to remove large edges is $0.133$).
+<!-- ::: {#fig-pbmc2-sc layout-ncol="2" fig-pos="H"} -->
+<!-- ![](figures/pbmc3k/tsne_trimesh_plot.png) -->
+
+<!-- ![](figures/pbmc3k/sc_4.png) -->
+
+<!-- ![](figures/pbmc3k/sc_5.png) -->
+
+<!-- ![](figures/pbmc3k/sc_6.png) -->
+
+<!-- Model in \gD{}, on the tSNE layout, and three views of the fit in projections from $9\text{-}D$, for the PBMC3k data ($(s_1, \ s_2) = (-0.050, \ -0.058)$, $b = 300 \  (15, \ 20)$, $m = 136$, benchmark value to remove large edges is $0.133$).
 (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/5Y1hE4i7N2k>).
-:::
+-->
+<!-- ::: -->
 

### 2024-08-14 paper.tex
Added trimesh of PaCMAP
Changed the point color to black

@@
 
-\begin{minipage}{0.50\linewidth}
-\includegraphics{figures/pbmc3k/tsne_trimesh_plot.png}\end{minipage}%
-%
-\begin{minipage}{0.50\linewidth}
-\includegraphics{figures/pbmc3k/sc_4.png}\end{minipage}%
-\newline
-\begin{minipage}{0.50\linewidth}
-\includegraphics{figures/pbmc3k/sc_5.png}\end{minipage}%
-%
-\begin{minipage}{0.50\linewidth}
-\includegraphics{figures/pbmc3k/sc_6.png}\end{minipage}%
-
-\caption{\label{fig-pbmc2-sc}Model in \gD{}, on the tSNE layout, and
-three views of the fit in projections from \(9\text{-}D\), for the
-PBMC3k data (\((s_1, \ s_2) = (-0.050, \ -0.058)\),
-\(b = 300 \  (15, \ 20)\), \(m = 136\), benchmark value to remove large
+\centering{
+
+\includegraphics[width=1\textwidth,height=1\textheight]{paper_files/figure-pdf/fig-pbmc2-sc-1.pdf}
+
+}
+
+\caption{\label{fig-pbmc2-sc}Model in \(2\text{-}D\), on the tSNE
+layout, and three views of the fit in projections from \(9  ext{-}D\),
+for the PBMC3k data (\((s_1,  s_2) = (-0.050,  -0.058)\),
+\(b = 300   (15,  20)\), \(m = 136\), benchmark value to remove large
 edges is \(0.133\)). (The \textbf{langevitour} software is used to view
@@
 
+\begin{center}
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-52-1.pdf}
+\end{center}
+
 \begin{figure}[H]

### 2024-08-14 paper.qmd
Added projection for Five Gaussian cluster

@@
 
+<!--Projection with tSNE-->
+```{r}
+#| label: five-gau-proj-tsne-model
+
+training_data_gau <- read_rds("data/five_gau_clusters/data_five_gau_training.rds")
+
+tsne_data_gau <- read_rds("data/five_gau_clusters/tsne_data_five_gau_61.rds")
+gau1_scaled_obj <- gen_scaled_data(
+  data = tsne_data_gau)
+tsne_gau_scaled <- gau1_scaled_obj$scaled_nldr
+
+## Compute hexbin parameters
+num_bins_x_gau1 <- 14
+lim1 <- gau1_scaled_obj$lim1
+lim2 <- gau1_scaled_obj$lim2
+r2_gau1 <- diff(lim2)/diff(lim1)
+
+gau1_model <- fit_highd_model(
+  training_data = training_data_gau,
+  emb_df = tsne_gau_scaled,
+  bin1 = num_bins_x_gau1,
+  r2 = r2_gau1,
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = FALSE,
+  col_start_highd = "x",
+  q = 0.16
+)
+
+df_bin_centroids_gau1 <- gau1_model$df_bin_centroids
+df_bin_gau1 <- gau1_model$df_bin
+
+## Triangulate bin centroids
+tr1_object_gau1 <- tri_bin_centroids(
+  df_bin_centroids_gau1, x = "c_x", y = "c_y")
+tr_from_to_df_gau1 <- gen_edges(
+  tri_object = tr1_object_gau1)
+
+# tr_from_to_df_gau1 <- tr_from_to_df_gau1 |>
+#   filter(row_number() != 76) |>
+#   filter(row_number() != 34)
+
+tr_from_to_df_gau1 <- tr_from_to_df_gau1 |>
+  filter(row_number() != 115) |>
+  filter(row_number() != 139)
+
+## Compute 2D distances
+distance_gau1 <- cal_2d_dist(
+  tr_coord_df = tr_from_to_df_gau1,
+  start_x = "x_from",
+  start_y = "y_from",
+  end_x = "x_to",
+  end_y = "y_to",
+  select_vars = c("from", "to", "distance"))
+
+## To find the benchmark value
+benchmark_gau1 <- find_lg_benchmark(
+  distance_edges = distance_gau1,
+  distance_col = "distance")
+
+trimesh_removed_gau1 <- vis_rmlg_mesh(
+  distance_edges = distance_gau1,
+  benchmark_value = benchmark_gau1,
+  tr_coord_df = tr_from_to_df_gau1,
+  distance_col = "distance")
+
+## Hexagonal binning to have regular hexagons
+hb_obj_gau1 <- hex_binning(
+  data = tsne_gau_scaled,
+  bin1 = num_bins_x_gau1,
+  r2 = r2_gau1,
+  q = 0.16)
+
+tsne_data_with_hb_id <- hb_obj_gau1$data_hb_id
+
+df_all_gau1 <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID),
+                                  tsne_data_with_hb_id)
+
+### Define type column
+df <- df_all_gau1 |>
+  dplyr::select(tidyselect::starts_with("x")) |>
+  dplyr::mutate(type = "data") ## original dataset
+
+df_b <- df_bin_gau1 |>
+  dplyr::filter(hb_id %in% df_bin_centroids_gau1$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b <- df_b[match(df_bin_centroids_gau1$hexID, df_b$hb_id),] |>
+  dplyr::select(-hb_id)
+
+df_exe <- dplyr::bind_rows(df_b, df)
+
+## Set the maximum difference as the criteria
+distance_df_small_edges <- distance_gau1 |>
+  dplyr::filter(distance < benchmark_gau1)
+
+projection <- cbind(
+  c(0.28321,-0.28851,-0.67413,0.39226),
+  c(-0.70800,-0.44269,0.04835,0.26867))
+
+five_gau_proj_tsne_model <- gen_proj_langevitour(
+  points_df = df_exe,
+  projection = projection,
+  edge_df = distance_df_small_edges |> select(-distance)
+) +
+  xlim(c(-0.15, 0.15)) +
+  ylim(c(-0.15, 0.15))
+```
+
+```{r}
+#| label: fig-five-gau-projs
+#| fig-cap: "(a) tSNE"
+
+five_gau_proj_tsne_model
+```
+
 ::: {#fig-gau-tsne-sc layout-ncol="3" fig-pos="H"}

### 2024-08-14 paper.tex
Added projection for Five Gaussian cluster

@@
 
+\begin{figure}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-five-gau-projs-1.pdf}
+
+}
+
+\caption{\label{fig-five-gau-projs}(a) tSNE}
+
+\end{figure}%
+
 \begin{figure}[H]
@@
 \begin{center}
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-52-1.pdf}
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-54-1.pdf}
 \end{center}

### 2024-08-14 paper.qmd
Added figure 10

@@
 
-<!--Projection with tSNE-->
+<!--Projections-->
 ```{r}
@@
 projection <- cbind(
-  c(0.28321,-0.28851,-0.67413,0.39226),
-  c(-0.70800,-0.44269,0.04835,0.26867))
+  c(-0.00215,-0.68905,-0.04778,-0.54223),
+  c(0.42558,-0.23854,-0.63659,0.35753))
 
@@
 ) +
-  xlim(c(-0.15, 0.15)) +
-  ylim(c(-0.15, 0.15))
+  xlim(c(-0.18, 0.1)) +
+  ylim(c(-0.78, -0.55))
+```
+
+```{r}
+#| label: five-gau-proj-umap-model
+
+umap_data_gau <- read_rds("data/five_gau_clusters/umap_data_five_gau.rds")
+gau1_scaled_obj <- gen_scaled_data(
+  data = umap_data_gau)
+umap_gau_scaled <- gau1_scaled_obj$scaled_nldr
+
+## Compute hexbin parameters
+num_bins_x_gau1 <- 44
+lim1 <- gau1_scaled_obj$lim1
+lim2 <- gau1_scaled_obj$lim2
+r2_gau1 <- diff(lim2)/diff(lim1)
+
+gau1_model <- fit_highd_model(
+  training_data = training_data_gau,
+  emb_df = umap_gau_scaled,
+  bin1 = num_bins_x_gau1,
+  r2 = r2_gau1,
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = FALSE,
+  col_start_highd = "x"
+)
+
+df_bin_centroids_gau1 <- gau1_model$df_bin_centroids
+df_bin_gau1 <- gau1_model$df_bin
+
+## Triangulate bin centroids
+tr1_object_gau1 <- tri_bin_centroids(
+  df_bin_centroids_gau1, x = "c_x", y = "c_y")
+tr_from_to_df_gau1 <- gen_edges(
+  tri_object = tr1_object_gau1)
+
+## Compute 2D distances
+distance_gau1 <- cal_2d_dist(
+  tr_coord_df = tr_from_to_df_gau1,
+  start_x = "x_from",
+  start_y = "y_from",
+  end_x = "x_to",
+  end_y = "y_to",
+  select_vars = c("from", "to", "distance"))
+
+## To find the benchmark value
+benchmark_gau1 <- find_lg_benchmark(
+  distance_edges = distance_gau1,
+  distance_col = "distance")
+
+## Hexagonal binning to have regular hexagons
+hb_obj_gau1 <- hex_binning(
+  data = umap_gau_scaled,
+  bin1 = num_bins_x_gau1,
+  r2 = r2_gau1)
+
+umap_data_with_hb_id <- hb_obj_gau1$data_hb_id
+
+df_all_gau1 <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID),
+                                umap_data_with_hb_id)
+
+### Define type column
+df <- df_all_gau1 |>
+  dplyr::select(tidyselect::starts_with("x")) |>
+  dplyr::mutate(type = "data") ## original dataset
+
+df_b <- df_bin_gau1 |>
+  dplyr::filter(hb_id %in% df_bin_centroids_gau1$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b <- df_b[match(df_bin_centroids_gau1$hexID, df_b$hb_id),] |>
+  dplyr::select(-hb_id)
+
+df_exe <- dplyr::bind_rows(df_b, df)
+
+## Set the maximum difference as the criteria
+distance_df_small_edges <- distance_gau1 |>
+  dplyr::filter(distance < benchmark_gau1)
+
+projection <- cbind(
+  c(-0.00215,-0.68905,-0.04778,-0.54223),
+  c(0.42558,-0.23854,-0.63659,0.35753))
+
+five_gau_proj_umap_model <- gen_proj_langevitour(
+  points_df = df_exe,
+  projection = projection,
+  edge_df = distance_df_small_edges |> select(-distance)
+) +
+  xlim(c(-0.18, 0.1)) +
+  ylim(c(-0.78, -0.55))
+```
+
+```{r}
+#| label: five-gau-proj-pacmap-model
+
+pacmap_data_gau <- read_rds("data/five_gau_clusters/pacmap_data_five_gau.rds")
+gau1_scaled_obj <- gen_scaled_data(
+  data = pacmap_data_gau)
+pacmap_gau_scaled <- gau1_scaled_obj$scaled_nldr
+
+## Compute hexbin parameters
+num_bins_x_gau1 <- 21
+lim1 <- gau1_scaled_obj$lim1
+lim2 <- gau1_scaled_obj$lim2
+r2_gau1 <- diff(lim2)/diff(lim1)
+
+gau1_model <- fit_highd_model(
+  training_data = training_data_gau,
+  emb_df = pacmap_gau_scaled,
+  bin1 = num_bins_x_gau1,
+  r2 = r2_gau1,
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = FALSE,
+  col_start_highd = "x",
+  q = 0.14
+)
+
+df_bin_centroids_gau1 <- gau1_model$df_bin_centroids
+df_bin_gau1 <- gau1_model$df_bin
+
+## Triangulate bin centroids
+tr1_object_gau1 <- tri_bin_centroids(
+  df_bin_centroids_gau1, x = "c_x", y = "c_y")
+tr_from_to_df_gau1 <- gen_edges(
+  tri_object = tr1_object_gau1)
+
+## Compute 2D distances
+distance_gau1 <- cal_2d_dist(
+  tr_coord_df = tr_from_to_df_gau1,
+  start_x = "x_from",
+  start_y = "y_from",
+  end_x = "x_to",
+  end_y = "y_to",
+  select_vars = c("from", "to", "distance"))
+
+## To find the benchmark value
+benchmark_gau1 <- find_lg_benchmark(
+  distance_edges = distance_gau1,
+  distance_col = "distance")
+
+trimesh_removed_gau1 <- vis_rmlg_mesh(
+  distance_edges = distance_gau1,
+  benchmark_value = benchmark_gau1,
+  tr_coord_df = tr_from_to_df_gau1,
+  distance_col = "distance")
+
+## Hexagonal binning to have regular hexagons
+hb_obj_gau1 <- hex_binning(
+  data = pacmap_gau_scaled,
+  bin1 = num_bins_x_gau1,
+  r2 = r2_gau1,
+  q = 0.14)
+
+pacmap_data_with_hb_id <- hb_obj_gau1$data_hb_id
+
+df_all_gau1 <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID),
+                                pacmap_data_with_hb_id)
+
+### Define type column
+df <- df_all_gau1 |>
+  dplyr::select(tidyselect::starts_with("x")) |>
+  dplyr::mutate(type = "data") ## original dataset
+
+df_b <- df_bin_gau1 |>
+  dplyr::filter(hb_id %in% df_bin_centroids_gau1$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b <- df_b[match(df_bin_centroids_gau1$hexID, df_b$hb_id),] |>
+  dplyr::select(-hb_id)
+
+df_exe <- dplyr::bind_rows(df_b, df)
+
+## Set the maximum difference as the criteria
+distance_df_small_edges <- distance_gau1 |>
+  dplyr::filter(distance < benchmark_gau1)
+
+projection <- cbind(
+  c(-0.00215,-0.68905,-0.04778,-0.54223),
+  c(0.42558,-0.23854,-0.63659,0.35753))
+
+five_gau_proj_pacmap_model <- gen_proj_langevitour(
+  points_df = df_exe,
+  projection = projection,
+  edge_df = distance_df_small_edges |> select(-distance)
+) +
+  xlim(c(-0.18, 0.1)) +
+  ylim(c(-0.78, -0.55))
 ```
@@
 #| label: fig-five-gau-projs
-#| fig-cap: "(a) tSNE"
-
-five_gau_proj_tsne_model
+#| fig-height: 6
+#| fig-width: 18
+#| fig-cap: "(a) tSNE, (b) UMAP, and (c) PaCMAP"
+
+five_gau_proj_tsne_model + five_gau_proj_umap_model + five_gau_proj_pacmap_model +
+  plot_layout(guides = "collect", ncol = 3) &
+  theme(legend.position='none')
 ```

### 2024-08-14 paper.qmd
Updated the projections for five Gaussian clusters

@@
 
-five_gau_proj_tsne_model <- gen_proj_langevitour(
-  points_df = df_exe,
-  projection = projection,
-  edge_df = distance_df_small_edges |> select(-distance)
-) +
+points_only_df <- df_exe |>
+    dplyr::select(-type)
+
+projected <- as.matrix(points_only_df) %*% projection
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+proj_model <- projected_df |>
+  dplyr::filter(type == "model")
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges |> select(-distance), 
+  proj_model, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(proj_model)[-NCOL(proj_model)], "_from")
+
+model_df <- dplyr::left_join(model_df, proj_model, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(proj_model)):NCOL(model_df)] <- paste0(names(proj_model)[-NCOL(proj_model)], "_to")
+
+five_gau_proj_tsne_model <- projected_df |>
+  filter(type == "data") |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+      geom_point(
+        alpha = 0.3) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#33a02c") +
+  scale_color_manual(values = c("#000000", "#33a02c")) +
   xlim(c(-0.18, 0.1)) +
   ylim(c(-0.78, -0.55))
+
+# five_gau_proj_tsne_model <- gen_proj_langevitour(
+#   points_df = df_exe,
+#   projection = projection,
+#   edge_df = distance_df_small_edges |> select(-distance)
+# ) +
+#   xlim(c(-0.18, 0.1)) +
+#   ylim(c(-0.78, -0.55))
 ```
@@
 
-five_gau_proj_umap_model <- gen_proj_langevitour(
-  points_df = df_exe,
-  projection = projection,
-  edge_df = distance_df_small_edges |> select(-distance)
-) +
+points_only_df <- df_exe |>
+    dplyr::select(-type)
+
+projected <- as.matrix(points_only_df) %*% projection
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+proj_model <- projected_df |>
+  dplyr::filter(type == "model")
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges |> select(-distance), 
+  proj_model, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(proj_model)[-NCOL(proj_model)], "_from")
+
+model_df <- dplyr::left_join(model_df, proj_model, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(proj_model)):NCOL(model_df)] <- paste0(names(proj_model)[-NCOL(proj_model)], "_to")
+
+five_gau_proj_umap_model <- projected_df |>
+  filter(type == "data") |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+      geom_point(
+        alpha = 0.3) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#33a02c") +
+  scale_color_manual(values = c("#000000", "#33a02c")) +
   xlim(c(-0.18, 0.1)) +
   ylim(c(-0.78, -0.55))
+
+# five_gau_proj_umap_model <- gen_proj_langevitour(
+#   points_df = df_exe,
+#   projection = projection,
+#   edge_df = distance_df_small_edges |> select(-distance)
+# ) +
+#   xlim(c(-0.18, 0.1)) +
+#   ylim(c(-0.78, -0.55))
 ```
@@
 
-five_gau_proj_pacmap_model <- gen_proj_langevitour(
-  points_df = df_exe,
-  projection = projection,
-  edge_df = distance_df_small_edges |> select(-distance)
-) +
+points_only_df <- df_exe |>
+    dplyr::select(-type)
+
+projected <- as.matrix(points_only_df) %*% projection
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+proj_model <- projected_df |>
+  dplyr::filter(type == "model")
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges |> select(-distance), 
+  proj_model, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(proj_model)[-NCOL(proj_model)], "_from")
+
+model_df <- dplyr::left_join(model_df, proj_model, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(proj_model)):NCOL(model_df)] <- paste0(names(proj_model)[-NCOL(proj_model)], "_to")
+
+five_gau_proj_pacmap_model <- projected_df |>
+  filter(type == "data") |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+      geom_point(
+        alpha = 0.3) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#33a02c") +
+  scale_color_manual(values = c("#000000", "#33a02c")) +
   xlim(c(-0.18, 0.1)) +
   ylim(c(-0.78, -0.55))
+
+# five_gau_proj_pacmap_model <- gen_proj_langevitour(
+#   points_df = df_exe,
+#   projection = projection,
+#   edge_df = distance_df_small_edges |> select(-distance)
+# ) +
+#   xlim(c(-0.18, 0.1)) +
+#   ylim(c(-0.78, -0.55))
 ```

### 2024-08-14 paper.qmd
Updated the Figure 10 labels

@@
 )
-interior_annotation <- function(label, position = c(0.92, 0.92)) {
+interior_annotation <- function(label, position = c(0.92, 0.92), cex = 1) {
   annotation_custom(grid::textGrob(label = label,
       x = unit(position[1], "npc"), y = unit(position[2], "npc"),
-      gp = grid::gpar(cex = 1, col="grey70")))
+      gp = grid::gpar(cex = cex, col="grey70")))
 }
@@
   xlim(c(-0.18, 0.1)) +
-  ylim(c(-0.78, -0.55))
+  ylim(c(-0.78, -0.55)) +
+  interior_annotation("a", 
+                      position = c(0.08, 0.9),
+                      cex = 3)
 
@@
   xlim(c(-0.18, 0.1)) +
-  ylim(c(-0.78, -0.55))
+  ylim(c(-0.78, -0.55)) +
+  interior_annotation("b", 
+                      position = c(0.08, 0.9),
+                      cex = 3)
 
@@
   xlim(c(-0.18, 0.1)) +
-  ylim(c(-0.78, -0.55))
+  ylim(c(-0.78, -0.55)) +
+  interior_annotation("c", 
+                      position = c(0.08, 0.9),
+                      cex = 3)
 

### 2024-08-14 paper.qmd
Updated Figure 6

@@
 
-### Algorithm to represent the model in 2D
+### Algorithm to represent the model in \gD{}
 
@@
   )
-
+```
+
+```{r}
+#| label: fig-p-d-error-in-2d-scurve
 error_plot_scurve

### 2024-08-14 paper.tex
Added figure 10
Updated Figure 6

@@
 
-\subsection{Algorithm to represent the model in
-2D}\label{algorithm-to-represent-the-model-in-2d}
+\subsection{\texorpdfstring{Algorithm to represent the model in
+\gD{}}{Algorithm to represent the model in }}\label{algorithm-to-represent-the-model-in}
 
@@
 
-\begin{center}
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/p-d-error-in-2d-scurve-1.pdf}
-\end{center}
+\begin{figure}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-p-d-error-in-2d-scurve-1.pdf}
+
+}
+
+\caption{\label{fig-p-d-error-in-2d-scurve}}
+
+\end{figure}%
 
@@
 \begin{center}
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-23-1.pdf}
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-24-1.pdf}
 \end{center}
@@
 
-\caption{\label{fig-five-gau-projs}(a) tSNE}
+\caption{\label{fig-five-gau-projs}(a) tSNE, (b) UMAP, and (c) PaCMAP}
 
@@
 \begin{center}
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-54-1.pdf}
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-57-1.pdf}
 \end{center}

### 2024-08-15 paper.qmd
Updated Figure 2

@@
 
+```{r}
+#| label: scurve-training
+training_data_scurve <- read_rds("data/s_curve/s_curve_training.rds")
+```
+
 <!--UMAP applied for Scurve data-->
@@
 
-<!--langevitour with true model for S-curve-->
-::: {#fig-scurve-true-sc layout-ncol="3" fig-pos="H"}
-
-![](figures/scurve/sc_true_1.png){width="150" fig-align="center"}
-
-![](figures/scurve/sc_true_2.png){width="150" fig-align="center"}
-
-![](figures/scurve/sc_true_3.png){width="150" fig-align="center"}
-
-Three views of the true model (grey points and lines) in $2\text{-}D$ projections from $7\text{-}D$, for the S-curve data (purple points).
+```{r}
+#| label: s-curve-true-model-proj1
+
+true_model_df <- read_rds("data/s_curve/scurve_true_model.rds")
+wireframe_true_model <- read_rds("data/s_curve/scurve_true_model_wireframe.rds")
+
+projection <- cbind(
+    c(0.24288,-0.03628,-0.03302,0.12182,-0.04150,0.02344,0.05910),
+    c(0.05019,0.07165,-0.20131,-0.16200,-0.05343,0.06854,-0.00551))
+
+projected <- as.matrix(training_data_scurve |> select(-ID)) %*% projection
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+projected_true <- as.matrix(true_model_df |> select(-ID)) %*% projection
+
+projected_true_df <- projected_true |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  wireframe_true_model, 
+  projected_true_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_true_df)[-NCOL(projected_true_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_true_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_true_df)):NCOL(model_df)] <- paste0(names(projected_true_df)[-NCOL(projected_true_df)], "_to")
+
+scurve_proj1_true_model <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+      geom_point(
+        color = "#000000") +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#969696")
+
+```
+
+```{r}
+#| label: s-curve-true-model-proj2
+
+projection <- cbind(
+    c(0.03685,-0.02239,0.08047,0.22415,-0.10295,0.04529,0.10381),
+    c(-0.16908,0.11691,0.04886,0.10189,-0.00203,-0.09763,-0.13207))
+
+projected <- as.matrix(training_data_scurve |> select(-ID)) %*% projection
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+projected_true <- as.matrix(true_model_df |> select(-ID)) %*% projection
+
+projected_true_df <- projected_true |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  wireframe_true_model, 
+  projected_true_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_true_df)[-NCOL(projected_true_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_true_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_true_df)):NCOL(model_df)] <- paste0(names(projected_true_df)[-NCOL(projected_true_df)], "_to")
+
+scurve_proj2_true_model <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+      geom_point(
+        color = "#000000") +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#969696")
+
+```
+
+```{r}
+#| label: s-curve-true-model-proj3
+
+projection <- cbind(
+    c(0.11484,0.06997,-0.13638,0.01096,0.01533,0.16257,-0.13615),
+    c(0.07805,0.01652,0.17174,-0.07343,0.16177,0.11112,0.04729))
+
+projected <- as.matrix(training_data_scurve |> select(-ID)) %*% projection
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+projected_true <- as.matrix(true_model_df |> select(-ID)) %*% projection
+
+projected_true_df <- projected_true |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  wireframe_true_model, 
+  projected_true_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_true_df)[-NCOL(projected_true_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_true_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_true_df)):NCOL(model_df)] <- paste0(names(projected_true_df)[-NCOL(projected_true_df)], "_to")
+
+scurve_proj3_true_model <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+      geom_point(
+        color = "#000000") +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#969696")
+
+```
+
+```{r}
+#| label: fig-scurve-proj
+#| fig-cap: "Three views of the true model (grey points and lines) in $2\text{-}D$ projections from $7\text{-}D$, for the S-curve data (purple points).
+The data is spread along the S shape, and does not vary much from this curve.
+(The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/I5GL23vLiw0>)."
+
+scurve_proj1_true_model + scurve_proj2_true_model + 
+  scurve_proj3_true_model + plot_layout(ncol = 3)
+```
+
+<!-- <!--langevitour with true model for S-curve--> -->
+<!-- ::: {#fig-scurve-true-sc layout-ncol="3" fig-pos="H"} -->
+
+<!-- ![](figures/scurve/sc_true_1.png){width="150" fig-align="center"} -->
+
+<!-- ![](figures/scurve/sc_true_2.png){width="150" fig-align="center"} -->
+
+<!-- ![](figures/scurve/sc_true_3.png){width="150" fig-align="center"} -->
+
+<!-- Three views of the true model (grey points and lines) in $2\text{-}D$ projections from $7\text{-}D$, for the S-curve data (purple points).
 The data is spread along the S shape, and does not vary much from this curve.
 (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/I5GL23vLiw0>).
-:::
+-->
+<!-- ::: -->
 
@@
 ```{r}
-#| label: scurve-training
-training_data_scurve <- read_rds("data/s_curve/s_curve_training.rds")
-```
-
-```{r}
 #| label: best-umap-model-scurve
@@
 #| label: model-prediction-umap-quollr
-
-true_model_df <- read_rds("data/s_curve/scurve_true_model.rds")
 

### 2024-08-15 paper.tex
Updated Figure 2

@@
 
-\begin{figure}[H]
-
-\begin{minipage}{0.33\linewidth}
-\begin{center}
-\includegraphics[width=1.5625in,height=\textheight]{figures/scurve/sc_true_1.png}
-\end{center}
-\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-\begin{center}
-\includegraphics[width=1.5625in,height=\textheight]{figures/scurve/sc_true_2.png}
-\end{center}
-\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-\begin{center}
-\includegraphics[width=1.5625in,height=\textheight]{figures/scurve/sc_true_3.png}
-\end{center}
-\end{minipage}%
-
-\caption{\label{fig-scurve-true-sc}Three views of the true model (grey
-points and lines) in \(2\text{-}D\) projections from \(7\text{-}D\), for
-the S-curve data (purple points). The data is spread along the S shape,
-and does not vary much from this curve. (The \textbf{langevitour}
+\begin{figure}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-scurve-proj-1.pdf}
+
+}
+
+\caption{\label{fig-scurve-proj}Three views of the true model (grey
+points and lines) in \(2   ext{-}D\) projections from \(7    ext{-}D\),
+for the S-curve data (purple points). The data is spread along the S
+shape, and does not vary much from this curve. (The \textbf{langevitour}
 software is used to view the data with a tour, and the full video is
@@
 \end{figure}%
+
+--\textgreater{}
 
@@
 \begin{center}
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-24-1.pdf}
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-28-1.pdf}
 \end{center}
@@
 Edges define the neighbourhood structure, in order to provide a smooth
-\gD{} representation of the fitted model.
-Figure~\ref{fig-scurve-true-sc} shows a wire frame of the true model
-that was used to generate the S-curve example data. The ideal is that
-the representation of the fitted model, at least for this example where
-we know the true model, should look similar to this.
+\gD{} representation of the fitted model. \textbf{?@fig-scurve-true-sc}
+shows a wire frame of the true model that was used to generate the
+S-curve example data. The ideal is that the representation of the fitted
+model, at least for this example where we know the true model, should
+look similar to this.
 
@@
 \begin{center}
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-57-1.pdf}
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-61-1.pdf}
 \end{center}

### 2024-08-15 paper.qmd
Updated Figure 2

@@
       geom_point(
+        size = 0.8,
+        alpha = 0.8,
         color = "#000000") +
@@
 ```{r}
-#| label: fig-scurve-proj
-#| fig-cap: "Three views of the true model (grey points and lines) in $2\text{-}D$ projections from $7\text{-}D$, for the S-curve data (purple points).
+#| label: fig-scurve-true-sc
+#| fig-cap: "Three views of the true model (grey points and lines) in $2\\text{-}D$ projections from $7\\text{-}D$, for the S-curve data (purple points).
 The data is spread along the S shape, and does not vary much from this curve.

### 2024-08-15 paper.tex
Updated Figure 2

--- 2024-08-14 paper.tex
@@
 
-\begin{figure}[H]
-
-\begin{minipage}{0.33\linewidth}
-\begin{center}
-\includegraphics[width=1.5625in,height=\textheight]{figures/scurve/sc_true_1.png}
-\end{center}
-\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-\begin{center}
-\includegraphics[width=1.5625in,height=\textheight]{figures/scurve/sc_true_2.png}
-\end{center}
-\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-\begin{center}
-\includegraphics[width=1.5625in,height=\textheight]{figures/scurve/sc_true_3.png}
-\end{center}
-\end{minipage}%
+\begin{figure}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-scurve-true-sc-1.pdf}
+
+}
 

### 2024-08-15 paper.qmd
Updated Figure 2

@@
 
+limits <- 1
+rng <- range(projected)
+projected <- projected/max(abs(rng))
+colnames(projected) <- c("P1", "P2")
+projected <- data.frame(projected)
+obs_labels <- as.character(1:nrow(training_data_scurve))
+
+axis_scale <- 2 * limits/3
+axis_pos <- -0.5 * limits
+
+adj <- function(x) axis_pos + x * axis_scale
+axes <- data.frame(x1 = adj(0), 
+                   y1 = adj(0), 
+                   x2 = adj(projection[, 1]), 
+                   y2 = adj(projection[, 2]))
+
+axis_labels <- colnames(training_data_scurve |> select(-ID))
+rownames(axes) <- axis_labels
+
 scurve_proj1_true_model <- projected_df |>
@@
       yend = proj2_to), 
-    color = "#969696")
+    color = "#969696") + 
+  geom_segment(
+    data=axes, 
+    aes(x=x1, y=y1, xend=x2, yend=y2), 
+    colour="grey70") +
+  geom_text(
+    data=axes, 
+    aes(x=x2, y=y2, label=rownames(axes)), 
+    colour="grey50")
 
@@
 names(model_df)[(2 + NCOL(projected_true_df)):NCOL(model_df)] <- paste0(names(projected_true_df)[-NCOL(projected_true_df)], "_to")
+
+limits <- 1
+rng <- range(projected)
+projected <- projected/max(abs(rng))
+colnames(projected) <- c("P1", "P2")
+projected <- data.frame(projected)
+obs_labels <- as.character(1:nrow(training_data_scurve))
+
+axis_scale <- 2 * limits/3
+axis_pos <- -0.5 * limits
+
+adj <- function(x) axis_pos + x * axis_scale
+axes <- data.frame(x1 = adj(0), 
+                   y1 = adj(0), 
+                   x2 = adj(projection[, 1]), 
+                   y2 = adj(projection[, 2]))
+
+axis_labels <- colnames(training_data_scurve |> select(-ID))
+rownames(axes) <- axis_labels
 

### 2024-08-15 paper.qmd
Changed axis text sizes

@@
 axis_scale <- 2 * limits/3
-axis_pos <- -0.5 * limits
+axis_pos <- -0.6 * limits
 
@@
     aes(x=x2, y=y2, label=rownames(axes)), 
-    colour="grey50")
+    colour="grey50",
+    size = 2.5)
 
@@
 axis_scale <- 2 * limits/3
-axis_pos <- -0.5 * limits
+axis_pos <- -0.4 * limits
 
@@
 #| label: s-curve-true-model-proj3
+#| eval: false
 
@@
 
-limits <- 1
+limits <- 1.5
 rng <- range(projected)
@@
 axis_scale <- 2 * limits/3
-axis_pos <- -0.5 * limits
+axis_pos <- -2/3 * limits
 
@@
 rownames(axes) <- axis_labels
+
+# theta <- seq(0, 2 * pi, length = 50)
+# circle <- data.frame(c1 = adj(cos(theta)), c2 = adj(sin(theta)))
 
@@
 scurve_proj1_true_model + scurve_proj2_true_model + 
-  scurve_proj3_true_model + plot_layout(ncol = 3)
+  plot_layout(ncol = 2)
 ```

### 2024-08-15 paper.qmd
Added axis circle

@@
 
+axis_scale <- 0.15 * limits
+axis_pos <- -0.6 * limits
+adj <- function(x) axis_pos + x * axis_scale
+theta <- seq(0, 2 * pi, length = 50)
+circle <- data.frame(c1 = adj(cos(theta)), c2 = adj(sin(theta)))
+
 scurve_proj1_true_model <- projected_df |>
@@
     colour="grey50",
-    size = 2.5)
+    size = 2.5) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70")
 
@@
 rownames(axes) <- axis_labels
+
+axis_scale <- 0.1 * limits
+axis_pos <- -0.4 * limits
+adj <- function(x) axis_pos + x * axis_scale
+theta <- seq(0, 2 * pi, length = 50)
+circle <- data.frame(c1 = adj(cos(theta)), c2 = adj(sin(theta)))
 
@@
 #| label: fig-scurve-true-sc
-#| fig-cap: "Three views of the true model (grey points and lines) in $2\\text{-}D$ projections from $7\\text{-}D$, for the S-curve data (purple points).
+#| fig-cap: "Two views of the true model (grey lines) in $2\\text{-}D$ projections from $7\\text{-}D$, for the S-curve data (black points).
 The data is spread along the S shape, and does not vary much from this curve.

### 2024-08-15 paper.tex
Added axis circle

@@
 
-\caption{\label{fig-scurve-true-sc}Three views of the true model (grey
-points and lines) in \(2\text{-}D\) projections from \(7\text{-}D\), for
-the S-curve data (purple points). The data is spread along the S shape,
-and does not vary much from this curve. (The \textbf{langevitour}
-software is used to view the data with a tour, and the full video is
-available at \url{https://youtu.be/I5GL23vLiw0}).}
+\caption{\label{fig-scurve-true-sc}Two views of the true model (grey
+lines) in \(2\text{-}D\) projections from \(7\text{-}D\), for the
+S-curve data (black points). The data is spread along the S shape, and
+does not vary much from this curve. (The \textbf{langevitour} software
+is used to view the data with a tour, and the full video is available at
+\url{https://youtu.be/I5GL23vLiw0}).}
 

### 2024-08-15 paper.qmd
Updated Figure 2

@@
     colour="grey50",
-    size = 2.5) +
+    size = 5) +
   geom_path(
@@
 (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/I5GL23vLiw0>)."
+#| fig-width: 10
+#| fig-height: 7
+#| fig-pos: H
 
@@
 
-<!-- <!--langevitour with true model for S-curve--> -->
+<!--langevitour with true model for S-curve--> 
 <!-- ::: {#fig-scurve-true-sc layout-ncol="3" fig-pos="H"} -->

### 2024-08-15 paper.qmd
Removed testing Figure for axes

@@
 
-<!--to draw axes-->
-```{r}
-
-## Draw the axes
-
-projection <- cbind(
-  c(0.18390,-0.18443,-0.06860,0.00654,-0.02921,0.02313,-0.08963),
-  c(-0.02902,0.03638,-0.16872,0.22418,0.01146,0.02777,0.01453))
-
-projected_df <- as.matrix(true_model_df_rm_id) %*% projection
-
-limits <- 1
-rng <- range(projected_df)
-projected_df <- projected_df/max(abs(rng))
-colnames(projected_df) <- c("P1", "P2")
-
-obs_labels <- as.character(1:nrow(true_model_df_rm_id))
-
-projected_df_n <- as.matrix(true_model_df_rm_id) %*% projection |> 
-  as_tibble(.name_repair = "unique") |> 
-  rename(c("proj1" = "...1", 
-           "proj2" = "...2"))
-
-projected_df$obs_labels <- obs_labels
-
-#if (position == "center") {
-  axis_scale <- 2 * limits/3
-  #axis_pos <- 0
-  axis_pos <- -0.6 * limits
-#}
-# else if (position == "bottomleft") {
-#   axis_scale <- limits/6
-#   axis_pos <- -2/3 * limits
-# }
-
-mean_cols <- colMeans(true_model_df_rm_id, na.rm = TRUE)
-
-adj <- function(x) axis_pos + x * axis_scale
-axes <- data.frame(x1 = adj(0), 
-                   y1 = adj(0), 
-                   x2 = adj(projection[, 1]), 
-                   y2 = adj(projection[, 2]))
-
-# axes <- data.frame(x1 = mean_cols + mean_cols * axis_scale, 
-#                    y1 = mean_cols + mean_cols * axis_scale, 
-#                    x2 = mean_cols + projection[, 1] * axis_scale, 
-#                    y2 = mean_cols + projection[, 2]* axis_scale)
-
-axis_labels <- colnames(true_model_df_rm_id)
-rownames(axes) <- axis_labels
-theta <- seq(0, 2 * pi, length = 50)
-circle <- data.frame(c1 = adj(cos(theta)), c2 = adj(sin(theta)))
-
-ggplot() +
-  geom_point(data = projected_df_n, 
-             aes(
-               x = proj1, 
-               y = proj2),
-             alpha = 0.5)  +
-  geom_segment(data=axes, aes(x=x1, y=y1, xend=x2, yend=y2), colour="grey70") +
-  geom_text(data=axes, aes(x=x2, y=y2, label=rownames(axes)), colour="grey50")
-
-```
-
 <!-- ::: {#fig-scurve-pred-sc layout-ncol="2" fig-pos="H"} -->

### 2024-08-15 paper.tex
Updated Figure 2
Removed testing Figure for axes

@@
 \end{figure}%
-
---\textgreater{}
 
@@
 \end{figure}%
-
-\begin{center}
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-28-1.pdf}
-\end{center}
 
@@
 \begin{center}
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-61-1.pdf}
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-60-1.pdf}
 \end{center}

### 2024-08-15 paper.qmd
Upated Figure 7

@@
                             )) +
-  geom_point(alpha = 0.5)
+  geom_point(alpha = 0.5) +
+  interior_annotation("b", c(0.92, 0.96))
 ```
@@
                             )) +
-  geom_point(alpha = 0.5)
+  geom_point(alpha = 0.5) +
+  interior_annotation("c", c(0.92, 0.96))
 ```
@@
     y = proj2)) +
-    geom_point(alpha = 0.5)   
+    geom_point(alpha = 0.5) +
+  interior_annotation("a", c(0.92, 0.96))  
 ```
@@
   geom_point(alpha = 0.5) +
-  geom_line(alpha = 0.3)
+  geom_line(alpha = 0.3) +
+  interior_annotation("d", c(0.92, 0.96))
 ```

### 2024-08-15 paper.qmd
Regenerated S-curve projection with new data

@@
 training_data_scurve <- read_rds("data/s_curve/s_curve_training.rds")
+
+training_data_scurve_n <- read_rds("data/s_curve/scurve_simulated_5000.rds")
 ```
@@
 
-projected <- as.matrix(training_data_scurve |> select(-ID)) %*% projection
+projected <- as.matrix(training_data_scurve_n) %*% projection
 
@@
 projected <- data.frame(projected)
-obs_labels <- as.character(1:nrow(training_data_scurve))
+obs_labels <- as.character(1:nrow(training_data_scurve_n))
 
@@
 
-axis_labels <- colnames(training_data_scurve |> select(-ID))
+axis_labels <- colnames(training_data_scurve_n)
 rownames(axes) <- axis_labels

### 2024-08-16 paper.qmd
Updated true model wireframe color

@@
         size = 0.8,
-        alpha = 0.8,
+        alpha = 0.5,
         color = "#000000") +
@@
       yend = proj2_to), 
-    color = "#969696") + 
+    color = "#3182bd") + 
   geom_segment(
@@
 #| label: fig-scurve-true-sc
-#| fig-cap: "Two views of the true model (grey lines) in $2\\text{-}D$ projections from $7\\text{-}D$, for the S-curve data (black points).
+#| fig-cap: "Two views of the true model (blue lines) in $2\\text{-}D$ projections from $7\\text{-}D$, for the S-curve data (black points).
 The data is spread along the S shape, and does not vary much from this curve.

### 2024-08-16 paper.tex
Updated true model wireframe color

@@
 
-\caption{\label{fig-scurve-true-sc}Two views of the true model (grey
+\caption{\label{fig-scurve-true-sc}Two views of the true model (blue
 lines) in \(2\text{-}D\) projections from \(7\text{-}D\), for the

### 2024-08-16 paper.qmd
Added linked plot section

@@
 
+## Linked plots
+
+It’s important to access the \gD{} layout and the generated model overlaid on data in \pD{} together to understand whether it fits the points everywhere, fits better in some places, or simply mismatches the pattern.
+Interactivity also helps in understanding the quirks that occur with different NLDR techniques (XXXXRefer the video after finalising the S-curve).
+
 ## A curious difference between tSNE, UMAP and PaCMAP revealer

### 2024-08-16 paper.tex
Added linked plot section

@@
 
+\section{Linked plots}\label{linked-plots}
+
+It's important to access the \gD{} layout and the generated model
+overlaid on data in \pD{} together to understand whether it fits the
+points everywhere, fits better in some places, or simply mismatches the
+pattern. Interactivity also helps in understanding the quirks that occur
+with different NLDR techniques (XXXXRefer the video after finalising the
+S-curve).
+
 \section{A curious difference between tSNE, UMAP and PaCMAP

### 2024-08-16 paper.qmd
Removed unnecessary figures

@@
 
-::: {#fig-gau-tsne-sc layout-ncol="3" fig-pos="H"}
-![](figures/five_gau_clusters/tsne_layout.png)
-
-![](figures/five_gau_clusters/2d_model_tsne.png)
-
-![](figures/five_gau_clusters/sc_tsne_3.png)
-
-The tSNE layout, model in \gD{}, and a view of the fit in projections from $4\text{-}D$, for the five Gaussian cluster data.
+<!-- ::: {#fig-gau-tsne-sc layout-ncol="3" fig-pos="H"} -->
+<!-- ![](figures/five_gau_clusters/tsne_layout.png) -->
+
+<!-- ![](figures/five_gau_clusters/2d_model_tsne.png) -->
+
+<!-- ![](figures/five_gau_clusters/sc_tsne_3.png) -->
+
+<!-- The tSNE layout, model in \gD{}, and a view of the fit in projections from $4\text{-}D$, for the five Gaussian cluster data.
 The model fits the separation and tries to *filled out* the clusters.
 (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/RASEE7N5MbM>).
-:::
-
-::: {#fig-gau-umap-sc layout-ncol="3" fig-pos="H"}
-![](figures/five_gau_clusters/umap_layout.png)
-
-![](figures/five_gau_clusters/2d_model_umap.png)
-
-![](figures/five_gau_clusters/sc_umap_3.png)
-
-The UMAP layout, model in \gD{}, and a view of the fit in projections from $4\text{-}D$, for the five Gaussian cluster data.
+-->
+<!-- ::: -->
+
+<!-- ::: {#fig-gau-umap-sc layout-ncol="3" fig-pos="H"} -->
+<!-- ![](figures/five_gau_clusters/umap_layout.png) -->
+
+<!-- ![](figures/five_gau_clusters/2d_model_umap.png) -->
+
+<!-- ![](figures/five_gau_clusters/sc_umap_3.png) -->
+
+<!-- The UMAP layout, model in \gD{}, and a view of the fit in projections from $4\text{-}D$, for the five Gaussian cluster data.
 The model fits the separation and tries to *filled out* the clusters, but not as much as tSNE.
 (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/iG4bCPkJilw>).
-:::
-
-::: {#fig-gau-pacmap-sc layout-ncol="3" fig-pos="H"}
-![](figures/five_gau_clusters/pacmap_layout.png)
-
-![](figures/five_gau_clusters/2d_model_pacmap.png)
-
-![](figures/five_gau_clusters/sc_pacmap_2.png)
-
-The PaCMAP layout, model in \gD{}, and a view of the fit in projections from $4\text{-}D$, for the five Gaussian cluster data.
+-->
+<!-- ::: -->
+
+<!-- ::: {#fig-gau-pacmap-sc layout-ncol="3" fig-pos="H"} -->
+<!-- ![](figures/five_gau_clusters/pacmap_layout.png) -->
+
+<!-- ![](figures/five_gau_clusters/2d_model_pacmap.png) -->
+
+<!-- ![](figures/five_gau_clusters/sc_pacmap_2.png) -->
+
+<!-- The PaCMAP layout, model in \gD{}, and a view of the fit in projections from $4\text{-}D$, for the five Gaussian cluster data.
 The model fits the separation and shows *flat* shaped clusters.
 (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/z07cKXi8EJQ>).
-:::
+-->
+<!-- ::: -->
 

### 2024-08-16 paper.tex
Removed unnecessary figures

@@
 well-separated clusters which evident these methods effectively preserve
-the global structure. In tSNE (Figure~\ref{fig-gau-tsne-sc} (a)), these
+the global structure. In tSNE (\textbf{?@fig-gau-tsne-sc} (a)), these
 clusters appear closely. UMAP arranges all clusters in a parallel
 manner, with three aligned in one line and the other two in a separate
-line (Figure~\ref{fig-gau-umap-sc} (a)). In contrast, PaCMAP shows one
+line (\textbf{?@fig-gau-umap-sc} (a)). In contrast, PaCMAP shows one
 central cluster and the remaining four spread out in different
-directions (Figure~\ref{fig-gau-pacmap-sc} (a)).
+directions (\textbf{?@fig-gau-pacmap-sc} (a)).
 
@@
 evidence that these methods preserve the local structure
-(Figure~\ref{fig-gau-tsne-sc} (c) and Figure~\ref{fig-gau-umap-sc} (c)).
-On the other hand, PaCMAP shows \emph{flat} shapes clusters in the model
+(\textbf{?@fig-gau-tsne-sc} (c) and \textbf{?@fig-gau-umap-sc} (c)). On
+the other hand, PaCMAP shows \emph{flat} shapes clusters in the model
 and evident that PaCMAP fail to capture the within-cluster variation
-(Figure~\ref{fig-gau-pacmap-sc} (c)).
+(\textbf{?@fig-gau-pacmap-sc} (c)).
 
@@
 \caption{\label{fig-five-gau-projs}(a) tSNE, (b) UMAP, and (c) PaCMAP}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/five_gau_clusters/tsne_layout.png}\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/five_gau_clusters/2d_model_tsne.png}\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/five_gau_clusters/sc_tsne_3.png}\end{minipage}%
-
-\caption{\label{fig-gau-tsne-sc}The tSNE layout, model in \gD{}, and a
-view of the fit in projections from \(4\text{-}D\), for the five
-Gaussian cluster data. The model fits the separation and tries to
-\emph{filled out} the clusters. (The \textbf{langevitour} software is
-used to view the data with a tour, and the full video is available at
-\url{https://youtu.be/RASEE7N5MbM}).}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/five_gau_clusters/umap_layout.png}\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/five_gau_clusters/2d_model_umap.png}\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/five_gau_clusters/sc_umap_3.png}\end{minipage}%
-
-\caption{\label{fig-gau-umap-sc}The UMAP layout, model in \gD{}, and a
-view of the fit in projections from \(4\text{-}D\), for the five
-Gaussian cluster data. The model fits the separation and tries to
-\emph{filled out} the clusters, but not as much as tSNE. (The
-\textbf{langevitour} software is used to view the data with a tour, and
-the full video is available at \url{https://youtu.be/iG4bCPkJilw}).}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/five_gau_clusters/pacmap_layout.png}\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/five_gau_clusters/2d_model_pacmap.png}\end{minipage}%
-%
-\begin{minipage}{0.33\linewidth}
-\includegraphics{figures/five_gau_clusters/sc_pacmap_2.png}\end{minipage}%
-
-\caption{\label{fig-gau-pacmap-sc}The PaCMAP layout, model in \gD{}, and
-a view of the fit in projections from \(4\text{-}D\), for the five
-Gaussian cluster data. The model fits the separation and shows
-\emph{flat} shaped clusters. (The \textbf{langevitour} software is used
-to view the data with a tour, and the full video is available at
-\url{https://youtu.be/z07cKXi8EJQ}).}
 

### 2024-08-16 paper.qmd
Updated error plot

@@
   geom_point(alpha=0.5) +
+  scale_colour_brewer(
+    type = "qual",
+    palette = 2
+  ) +
   theme(
-    legend.position = "bottom"
-  )
+    legend.position = "right",
+    legend.text = element_text(size = 10)
+  ) +
+  coord_equal()
 ```
@@
 #| label: fig-p-d-error-in-2d-scurve
+#| fig-cap: "error"
+#| out-height: 30%
+#| fig-pos: H
+
 error_plot_scurve

### 2024-08-16 paper.qmd
Updated
Updated Figure 9 (b) MSE plot

@@
 ## To initialize number of bins along the x-axis
-bin1_vec_scurve <- 2:15 #sqrt(NROW(training_data_scurve)/2)
+bin1_vec_scurve <- 5:19 #sqrt(NROW(training_data_scurve)/2)
 #buffer_vec <- seq(0.05, 0.1, by = 0.01)
@@
              color = "#bdbdbd", linewidth=1) +
+  scale_x_continuous(breaks = 5:19) +
   labs(x = expression(b[1]), y = "MSE") +

### 2024-08-19 paper.qmd
Updated upper limit of b1
Updated Figure 15

@@
 The value of $b_2$ is determined analytically by $b_1, q, r_2$.
-Values of $b_1$ between $2$ and $b_1 = \frac{n}{2}$ are allowed.
+Values of $b_1$ between $2$ and $b_1 = \sqrt{\frac{n}{2}}$ are allowed.
 @fig-param-scurve (b) shows the effect of different choices of $b_1$ on the MSE of the fitted model.
@@
 error_pbmc_tsne2 <- read_rds("data/pbmc3k/error_pbmc_tsne_30.rds")
-error_pbmc_phate <- read_rds("data/pbmc3k/error_pbmc_phate_5.rds") 
-error_pbmc_trimap <- read_rds("data/pbmc3k/error_pbmc_trimap_12_4_3.rds") 
-error_pbmc_pacmap <- read_rds("data/pbmc3k/error_pbmc_pacmap_30_random_0.9_5.rds")
+# error_pbmc_phate <- read_rds("data/pbmc3k/error_pbmc_phate_5.rds") 
+# error_pbmc_trimap <- read_rds("data/pbmc3k/error_pbmc_trimap_12_4_3.rds") 
+# error_pbmc_pacmap <- read_rds("data/pbmc3k/error_pbmc_pacmap_30_random_0.9_5.rds")
 
@@
                         error_pbmc_tsne,
-                        error_pbmc_tsne2,
-                        error_pbmc_phate,
-                        error_pbmc_trimap,
-                        error_pbmc_pacmap)
+                        error_pbmc_tsne2)
+
+error_pbmc <- error_pbmc |>
+  filter(bin1 >= 5)
 ```
@@
 # read the png file from device
-tsne_5_img <- readPNG("figures/pbmc3k/tsne_5.png", native = TRUE) 
-phate_5_img <- readPNG("figures/pbmc3k/phate_5.png", native = TRUE) 
-umap_img <- readPNG("figures/pbmc3k/author_view.png", native = TRUE) 
-tsne_30_img <- readPNG("figures/pbmc3k/tsne_30.png", native = TRUE)
-umap_15_img <- readPNG("figures/pbmc3k/umap_15.png", native = TRUE) 
-pacmap_img <- readPNG("figures/pbmc3k/pacmap.png", native = TRUE) 
-trimap_img <- readPNG("figures/pbmc3k/trimap.png", native = TRUE)
-umap_5_img <- readPNG("figures/pbmc3k/umap_5.png", native = TRUE)
-
-## Add arrows
-arrow_df_pbmc <- tibble(x = c(94, 280, 101, 150, 218, 250, 50),
-             xend = c(96, 295, 96, 180, 250, 290, 25),
-             y = c(10.06, 9.69, 9.88, 9.855, 9.82, 9.78, 10),
-             yend = c(10.14, 9.8, 9.8, 10, 10.3, 10, 9.8))
+
+pbmc_layouts <- readPNG("figures/pbmc3k/pbmc_nldr_layouts.png", native = TRUE)
 
@@
                           aes(x = b_non_empty, 
-                                  y = log(Error), 
+                              y = Error/1000, 
                               colour = method)) + 
@@
   geom_line(linewidth = 0.3) + 
-  geom_vline(xintercept = 135, linetype="solid", 
+  geom_vline(xintercept = 136, linetype="solid", 
              color = "black", linewidth=0.8, alpha = 0.5) +
-  scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a','#ff7f00',
-                              '#984ea3','#999999','#a65628','#f781bf')) +
-  ylab("log(absolute error)") +
-  xlab("number of non-empty bins")+
-    geom_segment(data = arrow_df_pbmc, aes(
-      x = x,  
-      y = y,  
-      xend = xend, 
-      yend = yend), 
-      arrow = arrow(length = unit(0.01, "npc"),
-        ends = "first"), 
-        color = "black",
-      linewidth = 0.1) +
-  theme(#aspect.ratio = 0.75,
+  scale_color_manual(values=c('#ff7f00', '#984ea3', 
+                              '#999999', '#a65628', '#f781bf')) +
+  ylab("absolute error (in thousands)") +
+  xlab("number of non-empty bins") + 
+  theme(aspect.ratio = 0.75,
         axis.text.x = element_text(size = 7),
@@
         axis.title.y = element_text(size = 7, angle = 90),
-        legend.position = "bottom",
-        legend.text = element_text(size = 7),
-        legend.key.height = unit(0.3, 'cm'),
-        legend.key.width = unit(0.3, 'cm'),
-        legend.key.size = unit(3, 'cm')) +
-    inset_element(p = tsne_5_img, 
-                  left = 1.8, 
-                  bottom = 0.08, 
+        axis.ticks.x = element_line(linewidth = 0.5),
+        axis.ticks.y = element_line(linewidth = 0.5),
+        legend.position = "none") +
+  inset_element(p = pbmc_layouts, 
+                left = 1.55, 
+                bottom = 0.35, 
                 right = 0, 
-                  top = 0.25) +                   
-    inset_element(p = phate_5_img, 
-                  left = 0.6, 
-                  bottom = 0.4, 
-                  right = 0, 
-                  top = 0.55) +                   
-    inset_element(p = umap_img, 
-                  left = 1.2, 
-                  bottom = 0.3, 
-                  right = -0.1, 
-                  top = 0.5) +
-    inset_element(p = umap_15_img, 
-                  left = 1.5, 
-                  bottom = 0.5, 
-                  right = 0, 
-                  top = 0.7) +
-    inset_element(p = tsne_30_img, 
-                  left = 0.6, 
-                  bottom = 0.01, 
-                  right = 0, 
-                  top = 0.18) +
-    inset_element(p = pacmap_img, 
-                  left = 1.7, 
-                  bottom = 0.3, 
-                  right = 0, 
-                  top = 0.5) +
-    inset_element(p = umap_5_img, 
-                  left = 0.4, 
-                  bottom = 0.05, 
-                  right = -0.2, 
-                  top = 0.22)
-
-```
-
-```{r}
-#| fig-cap: "Absolute error from UMAP and tSNE applied to training PBMC3k dataset with diffferent parameter choices.
-What is the best parameter choice to create the model?
+                top = 1)
+
+```
+
+```{r}
+#| fig-cap: "Absolute error from UMAP and tSNE applied to PBMC3k dataset with diffferent (hyper-)parameter choices.
+What is the best (hyper-)parameter choice to create the model?
 The residual plot have a steep slope at the beginning, indicating that a smaller number of non-empty bins causes a larger amount of error.

### 2024-08-19 paper.tex
Updated error plot
Updated upper limit of b1
Updated Figure 15

@@
 
-\begin{figure}
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-p-d-error-in-2d-scurve-1.pdf}
-
-}
-
-\caption{\label{fig-p-d-error-in-2d-scurve}}
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=0.3\textheight]{paper_files/figure-pdf/fig-p-d-error-in-2d-scurve-1.pdf}
+
+}
+
+\caption{\label{fig-p-d-error-in-2d-scurve}error}
 
@@
 determined analytically by \(b_1, q, r_2\). Values of \(b_1\) between
-\(2\) and \(b_1 = \frac{n}{2}\) are allowed.
+\(2\) and \(b_1 = \sqrt{\frac{n}{2}}\) are allowed.
 Figure~\ref{fig-param-scurve} (b) shows the effect of different choices
@@
 \caption{\label{fig-pbmc-abserror}Absolute error from UMAP and tSNE
-applied to training PBMC3k dataset with diffferent parameter choices.
-What is the best parameter choice to create the model? The residual plot
-have a steep slope at the beginning, indicating that a smaller number of
-non-empty bins causes a larger amount of error. Then, the slope
-gradually declines or level off, indicating that a higher number of
-non-empty bins generates a smaller error. Using the elbow method, it was
-observed that when the number of non-empty bins is set to \(136\), the
-lowest error occurred with the parameters perplexity: \(30\).}
+applied to PBMC3k dataset with diffferent (hyper-)parameter choices.
+What is the best (hyper-)parameter choice to create the model? The
+residual plot have a steep slope at the beginning, indicating that a
+smaller number of non-empty bins causes a larger amount of error. Then,
+the slope gradually declines or level off, indicating that a higher
+number of non-empty bins generates a smaller error. Using the elbow
+method, it was observed that when the number of non-empty bins is set to
+\(136\), the lowest error occurred with the parameters perplexity:
+\(30\).}
 

### 2024-08-19 paper.qmd
Updated Figure 15

@@
 error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_15_min_dist_0.99.rds")
-error_pbmc_tsne <- read_rds("data/pbmc3k/error_pbmc_tsne_5.rds")
+#error_pbmc_tsne <- read_rds("data/pbmc3k/error_pbmc_tsne_5.rds")
 error_pbmc_tsne2 <- read_rds("data/pbmc3k/error_pbmc_tsne_30.rds")
@@
                         error_pbmc_umap3,
-                        error_pbmc_tsne,
                         error_pbmc_tsne2)
@@
 error_plot_pbmc <- ggplot(error_pbmc, 
-                          aes(x = b_non_empty, 
-                              y = Error/1000, 
+                          aes(x = bin1, 
+                              y = MSE, 
                               colour = method)) + 
@@
   geom_line(linewidth = 0.3) + 
-  geom_vline(xintercept = 136, linetype="solid", 
-             color = "black", linewidth=0.8, alpha = 0.5) +
-  scale_color_manual(values=c('#ff7f00', '#984ea3', 
-                              '#999999', '#a65628', '#f781bf')) +
-  ylab("absolute error (in thousands)") +
-  xlab("number of non-empty bins") + 
+  # geom_vline(xintercept = 15, linetype="solid", 
+  #            color = "black", linewidth=0.8, alpha = 0.5) +
+  scale_color_manual(values=c('#ff7f00', '#999999', 
+                              '#a65628', '#f781bf')) +
+  ylab("MSE") +
+  xlab(expression(b[1])) + 
   theme(aspect.ratio = 0.75,
@@
 ```{r}
-#| fig-cap: "Absolute error from UMAP and tSNE applied to PBMC3k dataset with diffferent (hyper-)parameter choices.
-What is the best (hyper-)parameter choice to create the model?
-The residual plot have a steep slope at the beginning, indicating that a smaller number of non-empty bins causes a larger amount of error.
-Then, the slope gradually declines or level off, indicating that a higher number of non-empty bins generates a smaller error.
-Using the elbow method, it was observed that when the number of non-empty bins is set to $136$, the lowest error occurred with the parameters perplexity: $30$."
+#| fig-cap: "MSE from UMAP with diffferent (hyper-)parameter choices and tSNE with default (hyper-)parameter setting applied to PBMC3k dataset for different number of bins along the x-axis ($b_1$).
+Which is the most reasonable representation of the structures present in the PBMC3k data?
+The residual plot have a steep slope at the beginning, indicating that a smaller $b_1$ causes a larger amount of error.
+Then, the slope gradually declines or level off, indicating that a higher number of $b_1$ generates a smaller error.
+It was observed that for every $b_1$, the lowest error occurred with tSNE using the perplexity: $30$."
 #| label: fig-pbmc-abserror

### 2024-08-19 paper.tex
Updated Figure 15

@@
 
-\caption{\label{fig-pbmc-abserror}Absolute error from UMAP and tSNE
-applied to PBMC3k dataset with diffferent (hyper-)parameter choices.
-What is the best (hyper-)parameter choice to create the model? The
-residual plot have a steep slope at the beginning, indicating that a
-smaller number of non-empty bins causes a larger amount of error. Then,
-the slope gradually declines or level off, indicating that a higher
-number of non-empty bins generates a smaller error. Using the elbow
-method, it was observed that when the number of non-empty bins is set to
-\(136\), the lowest error occurred with the parameters perplexity:
-\(30\).}
+\caption{\label{fig-pbmc-abserror}MSE from UMAP with diffferent
+(hyper-)parameter choices and tSNE with default (hyper-)parameter
+setting applied to PBMC3k dataset for different number of bins along the
+x-axis (\(b_1\)). Which is the most reasonable representation of the
+structures present in the PBMC3k data? The residual plot have a steep
+slope at the beginning, indicating that a smaller \(b_1\) causes a
+larger amount of error. Then, the slope gradually declines or level off,
+indicating that a higher number of \(b_1\) generates a smaller error. It
+was observed that for every \(b_1\), the lowest error occurred with tSNE
+using the perplexity: \(30\).}
 

### 2024-08-19 paper.qmd
Updated the description

@@
 @fig-umap-author is a reproduction of the published plot.
-It shows three well-separated clusters.
+It shows three well-separated clusters with big separations.
 
@@
 
-To determine whether the UMAP representation with the (hyper-)parameter choice suggested by @chen2023 preserves the original data structure, we visualize the model constructed with UMAP overlaid on the \pD{} data.
+To determine whether the UMAP representation with the (hyper-)parameter choice suggested by @chen2023 preserves the original data structure, we visualize the model constructed with UMAP overlaid on the $9\text{-}D$ data.
 The @fig-umap-author shows three well-separated clusters.
-However, as shown in @fig-pbmc1-sc, there is no big separation between three clusters in \pD{}.
+However, as shown in @fig-pbmc1-sc, there is no big separation between three clusters in $9\text{-}D$.
 Therefore, the suggested UMAP representation (@fig-umap-author) does not accurately represent the structure(s) present in PBMC3k dataset.
@@
 
-Model in \gD{}, on the UMAP layout, and three views of the fit in projections from $9\text{-}D$, for the PBMC3k data ($(s_1, \ s_2) = (-0.050, \ -0.041)$, $b = 870 \  (30, \ 29)$, $m = 135$, benchmark value to remove large edges is $0.099$).
-(The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/VqqWuE0Jj6A>).
+Model in \gD{}, on the UMAP layout, and three views of the fit in projections from $9\text{-}D$, for the PBMC3k data ($(s_1, \ s_2) = (-0.050, \ -0.041)$, $b = 870 \  (30, \ 29)$, $m = 135$, benchmark value to remove large edges is $0.099$) (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/VqqWuE0Jj6A>).
 :::
 
-In order to find a reasonable NLDR representation for the PBMC3k dataset, the absolute error for different numbers of non-empty bins using various NLDR techniques and different (hyper-)parameter settings (@fig-pbmc-abserror) were calculated.
-After analyzing the results, it was found that tSNE with default (hyper-)parameter setting (perplexity: $30$) achieved the lowest error when there were $136$ number of non-empty bins.
+In order to find a reasonable NLDR representation for the PBMC3k dataset, the MSE for different $b_1$ using UMAP with different (hyper-)parameter settings and tSNE with default (hyper-)parameter setting (@fig-pbmc-mse) were calculated.
+After analyzing the results, it was found that tSNE with default (hyper-)parameter setting (perplexity: $30$) achieved the lowest error.
 Therefore, tSNE with a perplexity value set to $30$, the default parameter setting, is considered as a reasonable representation for the PBMC3k dataset.
@@
 It was observed that for every $b_1$, the lowest error occurred with tSNE using the perplexity: $30$."
-#| label: fig-pbmc-abserror
+#| label: fig-pbmc-mse
 #| fig-pos: H
@@
 
-As shown in @fig-tsne-suggest, there are three well-separated clusters, although they are located close to each other.
+As shown in @fig-tsne-suggest, there are three well-separated clusters, although the separation between the clusters are small.
 Additionally, non-linear structures can also be observed within the clusters (@fig-tsne-suggest).
@@
 
-We then fit the model for tSNE, and visualize the resultant model in the \pD{} data space.
+<!-- We then fit the model for tSNE, and visualize the resultant model in the $9\text{-}D$ data space.
 The model shows a quirk, as shown in @fig-pbmc2-sc.
@@
 This is evident that tSNE with perplexity $30$ provides a reasonable representation of PBMC3k data.
+-->
 

### 2024-08-19 paper.tex
Updated the description

@@
 Figure~\ref{fig-umap-author} is a reproduction of the published plot. It
-shows three well-separated clusters.
+shows three well-separated clusters with big separations.
 
@@
 structure, we visualize the model constructed with UMAP overlaid on the
-\pD{} data. The Figure~\ref{fig-umap-author} shows three well-separated
-clusters. However, as shown in Figure~\ref{fig-pbmc1-sc}, there is no
-big separation between three clusters in \pD{}. Therefore, the suggested
-UMAP representation (Figure~\ref{fig-umap-author}) does not accurately
-represent the structure(s) present in PBMC3k dataset. But, when
-visualizing the model-in-th-data-space some unobserved structures can be
-seen. Some clusters have non-linear continuity patterns and high-density
-patches (Figure~\ref{fig-pbmc1-sc}).
+\(9\text{-}D\) data. The Figure~\ref{fig-umap-author} shows three
+well-separated clusters. However, as shown in Figure~\ref{fig-pbmc1-sc},
+there is no big separation between three clusters in \(9\text{-}D\).
+Therefore, the suggested UMAP representation
+(Figure~\ref{fig-umap-author}) does not accurately represent the
+structure(s) present in PBMC3k dataset. But, when visualizing the
+model-in-th-data-space some unobserved structures can be seen. Some
+clusters have non-linear continuity patterns and high-density patches
+(Figure~\ref{fig-pbmc1-sc}).
 
@@
 \(b = 870 \  (30, \ 29)\), \(m = 135\), benchmark value to remove large
-edges is \(0.099\)). (The \textbf{langevitour} software is used to view
+edges is \(0.099\)) (The \textbf{langevitour} software is used to view
 the data with a tour, and the full video is available at
@@
 In order to find a reasonable NLDR representation for the PBMC3k
-dataset, the absolute error for different numbers of non-empty bins
-using various NLDR techniques and different (hyper-)parameter settings
-(Figure~\ref{fig-pbmc-abserror}) were calculated. After analyzing the
+dataset, the MSE for different \(b_1\) using UMAP with different
+(hyper-)parameter settings and tSNE with default (hyper-)parameter
+setting (Figure~\ref{fig-pbmc-mse}) were calculated. After analyzing the
 results, it was found that tSNE with default (hyper-)parameter setting
-(perplexity: \(30\)) achieved the lowest error when there were \(136\)
-number of non-empty bins. Therefore, tSNE with a perplexity value set to
-\(30\), the default parameter setting, is considered as a reasonable
-representation for the PBMC3k dataset.
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=0.7\textwidth,height=\textheight]{paper_files/figure-pdf/fig-pbmc-abserror-1.pdf}
-
-}
-
-\caption{\label{fig-pbmc-abserror}MSE from UMAP with diffferent
+(perplexity: \(30\)) achieved the lowest error. Therefore, tSNE with a
+perplexity value set to \(30\), the default parameter setting, is
+considered as a reasonable representation for the PBMC3k dataset.
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=0.7\textwidth,height=\textheight]{paper_files/figure-pdf/fig-pbmc-mse-1.pdf}
+
+}
+
+\caption{\label{fig-pbmc-mse}MSE from UMAP with diffferent
 (hyper-)parameter choices and tSNE with default (hyper-)parameter
@@
 As shown in Figure~\ref{fig-tsne-suggest}, there are three
-well-separated clusters, although they are located close to each other.
-Additionally, non-linear structures can also be observed within the
-clusters (Figure~\ref{fig-tsne-suggest}). This demonstrates that tSNE
-accurately captures the data structure of the PBMC3k dataset, which UMAP
-did not achieve.
+well-separated clusters, although the separation between the clusters
+are small. Additionally, non-linear structures can also be observed
+within the clusters (Figure~\ref{fig-tsne-suggest}). This demonstrates
+that tSNE accurately captures the data structure of the PBMC3k dataset,
+which UMAP did not achieve.
 
@@
 \end{figure}%
-
-We then fit the model for tSNE, and visualize the resultant model in the
-\pD{} data space. The model shows a quirk, as shown in
-Figure~\ref{fig-pbmc2-sc}. All three clusters are connected by an edge
-except the small and large clusters. Because the clusters are so close
-in \gD{}, they attempt to maintain the structure in \pD{} as well. This
-is evident that tSNE with perplexity \(30\) provides a reasonable
-representation of PBMC3k data.
 

### 2024-08-19 paper.qmd
Updated back to original S-curev data

@@
 training_data_scurve <- read_rds("data/s_curve/s_curve_training.rds")
-
-training_data_scurve_n <- read_rds("data/s_curve/scurve_simulated_5000.rds")
 ```
@@
 
-projected <- as.matrix(training_data_scurve_n) %*% projection
+projected <- as.matrix(training_data_scurve |> select(-ID)) %*% projection
 
@@
 projected <- data.frame(projected)
-obs_labels <- as.character(1:nrow(training_data_scurve_n))
+obs_labels <- as.character(1:nrow(training_data_scurve))
 
@@
 
-axis_labels <- colnames(training_data_scurve_n)
+axis_labels <- colnames(training_data_scurve |> select(-ID))
 rownames(axes) <- axis_labels

### 2024-08-19 paper.qmd
Added some diagnostic plots

@@
 <!--To generate errors for different total number of bins-->
+
+```{r}
+#| label: bin-counts-scurve-with_buffer
+## To initialize number of bins along the x-axis
+bin1_vec_scurve <- c(7, 11, 17) #sqrt(NROW(training_data_scurve)/2)
+#buffer_vec <- seq(0.05, 0.16, by = 0.01)
+buffer_vec <- c(0.05, 0.07, 0.12)
+
+centroids_df_scurve <- data.frame(matrix(nrow = 0, ncol = 0))
+
+for (xbins in bin1_vec_scurve) {
+  for (buff in buffer_vec) {
+
+    bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = buff)$bin2
+
+    scurve_model <- fit_highd_model(
+      training_data = training_data_scurve,
+      emb_df = umap_scurve_scaled,
+      bin1 = xbins,
+      r2 = r2,
+      q = buff, 
+      is_bin_centroid = TRUE,
+      is_rm_lwd_hex = FALSE,
+      col_start_highd = "x"
+    )
+
+    df_bin_centroids_scurve <- scurve_model$df_bin_centroids
+    #df_bin_scurve <- scurve_model$df_bin
+
+    ## Compute error
+    df_bin_centroids_scurve <- df_bin_centroids_scurve |>
+      mutate(bin1 = xbins,
+             bin2 = bin2,
+             b = bin1 * bin2,
+             b_non_empty = NROW(df_bin_centroids_scurve),
+             buff = buff)
+
+    centroids_df_scurve <- bind_rows(centroids_df_scurve, df_bin_centroids_scurve)
+
+  }
+
+}
+
+b7_q_plt <- centroids_df_scurve |>
+  filter(bin1 == 7) |>
+  filter(buff == 0.12) |>
+  ggplot(aes(x = bin_counts)) +
+  geom_density() +
+  xlab("bin count") +
+  theme(
+    axis.text.x = element_text(size = 7),
+    axis.ticks.x = element_line(linewidth = 0.5),
+    axis.title.x = element_text(size = 7)
+
+  )
+
+b11_q_plt <- centroids_df_scurve |>
+  filter(bin1 == 11) |>
+  filter(buff == 0.07) |>
+  ggplot(aes(x = bin_counts)) +
+  geom_density() +
+  xlab("bin count") +
+  theme(
+    axis.text.x = element_text(size = 7),
+    axis.ticks.x = element_line(linewidth = 0.5), 
+    axis.title.x = element_text(size = 7)
+
+  )
+
+b17_q_plt <- centroids_df_scurve |>
+  filter(bin1 == 17) |>
+  filter(buff == 0.05) |>
+  ggplot(aes(x = bin_counts)) +
+  geom_density() +
+  xlab("bin count") +
+  theme(
+    axis.text.x = element_text(size = 7),
+    axis.ticks.x = element_line(linewidth = 0.5),
+    axis.title.x = element_text(size = 7)
+
+  )
+```
+
+```{r}
+#| label: fig-bin-counts-scurve-with_buffer
+#| fig-cap: "Distribution of bin counts for differnt $b_1$ with specific $q$"
+#| out-height: 30%
+
+b7_q_plt + b11_q_plt + 
+  b17_q_plt +
+  plot_layout(ncol = 3)
+
+```
+
+```{r}
+#| out-height: 30%
+
+b7_q_plt2 <- centroids_df_scurve |>
+  filter(bin1 == 7) |>
+  ggplot(aes(x = std_counts,
+             colour = as.factor(buff),
+             group = as.factor(buff))) +
+  geom_density() +
+  scale_color_brewer(type = "qual") +
+  xlab("standardised bin count") +
+  theme(
+    axis.text.x = element_text(size = 7),
+    axis.ticks.x = element_line(linewidth = 0.5),
+    axis.title.x = element_text(size = 7)
+
+  )
+
+b11_q_plt2 <- centroids_df_scurve |>
+  filter(bin1 == 11) |>
+  ggplot(aes(x = std_counts,
+             colour = as.factor(buff),
+             group = as.factor(buff))) +
+  geom_density() +
+  scale_color_brewer(type = "qual") +
+  xlab("standardised bin count") +
+  theme(
+    axis.text.x = element_text(size = 7),
+    axis.ticks.x = element_line(linewidth = 0.5),
+    axis.title.x = element_text(size = 7)
+  )
+
+b17_q_plt2 <- centroids_df_scurve |>
+  filter(bin1 == 17) |>
+  ggplot(aes(x = std_counts,
+             colour = as.factor(buff),
+             group = as.factor(buff))) +
+  geom_density() +
+  scale_color_brewer(type = "qual") +
+  xlab("standardised bin count") +
+  theme(
+    axis.text.x = element_text(size = 7),
+    axis.ticks.x = element_line(linewidth = 0.5),
+    axis.title.x = element_text(size = 7)
+  )
+
+b7_q_plt2 + b11_q_plt2 + b17_q_plt2
+```
+
+```{r}
+## To initialize number of bins along the x-axis
+bin1_vec_scurve <- 5:19 #sqrt(NROW(training_data_scurve)/2)
+#buffer_vec <- seq(0.05, 0.16, by = 0.01)
+buffer_vec <- c(0.05, 0.07, 0.12)
+
+error_df_scurve_q <- data.frame(matrix(nrow = 0, ncol = 0))
+
+for (xbins in bin1_vec_scurve) {
+  for (buff in buffer_vec) {
+
+    bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = buff)$bin2
+
+    scurve_model <- fit_highd_model(
+      training_data = training_data_scurve,
+      emb_df = umap_scurve_scaled,
+      bin1 = xbins,
+      r2 = r2,
+      q = buff, 
+      is_bin_centroid = TRUE,
+      is_rm_lwd_hex = FALSE,
+      col_start_highd = "x"
+    )
+
+    df_bin_centroids_scurve <- scurve_model$df_bin_centroids
+    df_bin_scurve <- scurve_model$df_bin
+
+    ## Compute error
+    error_df <- glance(
+      df_bin_centroids = df_bin_centroids_scurve,
+      df_bin = df_bin_scurve,
+      training_data = training_data_scurve,
+      newdata = NULL,
+      type_NLDR = "UMAP",
+      col_start = "x") |>
+      mutate(bin1 = xbins,
+             bin2 = bin2,
+             b = bin1 * bin2,
+             b_non_empty = NROW(df_bin_centroids_scurve),
+             q = buff)
+
+    error_df_scurve_q <- bind_rows(error_df_scurve_q, error_df)
+
+  }
+
+}
+
+ggplot(error_df_scurve_q, 
+       aes(x = bin1, 
+           y = MSE,
+           colour = as.factor(q),
+           group = as.factor(q))) + 
+  geom_point(size = 1) +
+  geom_line() +
+  scale_color_brewer(type = "qual") +
+  scale_x_continuous(breaks = 5:19) +
+  labs(x = expression(b[1]), y = "MSE") +
+  theme(aspect.ratio = 0.75,
+        axis.text.x = element_text(size = 12),
+        axis.text.y = element_text(size = 12),
+        axis.title.x = element_text(size = 12),
+        axis.title.y = element_text(size = 12, 
+            angle=90),
+        axis.ticks.x = element_line(linewidth = 0.5),
+        axis.ticks.y = element_line(linewidth = 0.5),
+        plot.title = element_text(size = 12),
+        legend.title = element_text(size = 12),
+        legend.text = element_text(size = 10),
+        legend.key.height = unit(1, 'cm'),
+        legend.key.width = unit(1, 'cm'),
+        legend.key.size = unit(1, 'cm'))
+
+```
+
 ```{r}
@@
             angle=90),
+        axis.ticks.x = element_line(linewidth = 0.5),
+        axis.ticks.y = element_line(linewidth = 0.5),
         plot.title = element_text(size = 12),

### 2024-08-19 paper.tex
Added some diagnostic plots

@@
 
+\begin{figure}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=0.3\textheight]{paper_files/figure-pdf/fig-bin-counts-scurve-with_buffer-1.pdf}
+
+}
+
+\caption{\label{fig-bin-counts-scurve-with_buffer}Distribution of bin
+counts for differnt \(b_1\) with specific \(q\)}
+
+\end{figure}%
+
+\begin{center}
+\includegraphics[width=1\textwidth,height=0.3\textheight]{paper_files/figure-pdf/unnamed-chunk-34-1.pdf}
+\end{center}
+
+\begin{center}
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-35-1.pdf}
+\end{center}
+
 \subsubsection{Removal of low density
@@
 \begin{center}
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-60-1.pdf}
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-64-1.pdf}
 \end{center}

### 2024-08-19 paper.qmd
Updated Figure 10

@@
 ```{r}
+#| out-height: 30%
+
+centroids_df_scurve |>
+  filter(bin1 == 7) |>
+  ggplot(aes(x = as.factor(hexID), 
+           y = std_counts,
+           group = as.factor(buff),
+           color = as.factor(buff))) + 
+  geom_line() +
+  scale_color_brewer(type = "qual")
+```
+
+```{r}
 ## To initialize number of bins along the x-axis
@@
 
+# cell_count_scurve <- ggplot(df_bin_centroids_all, 
+#                             aes(x = reorder(as.factor(hexID), 
+#                                             -bin_counts), 
+#                                 y = bin_counts,
+#                                 group = as.factor(b1),
+#                                 color = as.factor(b1))) + 
+#   geom_line() + 
+#   # geom_hline(yintercept = benchmark1, color = "black", linewidth=0.5, alpha = 0.5) +
+#   scale_colour_discrete_qualitative(palette = "Dark 3") +
+#   xlab("hexagon id (re-ordered)") + 
+#   ylab("bin count") +
+#   ggtitle("(a)") +
+#   theme(aspect.ratio = 0.75,
+#         axis.title.x = element_text(size = 12),
+#         axis.title.y = element_text(size = 12, angle = 90),
+#         axis.text.x = element_blank(),
+#         axis.text.y = element_text(size = 12),
+#         plot.title = element_text(size = 12),
+#         legend.text = element_text(size = 10),
+#         legend.key.size = unit(1, 'cm'),
+#         legend.key.height = unit(1, 'cm'))
+
 cell_count_scurve <- ggplot(df_bin_centroids_all, 
-                            aes(x = reorder(as.factor(hexID), 
-                                            -bin_counts), 
-                                y = bin_counts,
+                            aes(x = as.factor(hexID), 
+                                y = std_counts,
                                 group = as.factor(b1),
@@
   xlab("hexagon id (re-ordered)") + 
-  ylab("bin count") +
+  ylab("standardised bin count") +
   ggtitle("(a)") +
@@
 #| label: bins-scurve
+#| eval: false
 ## To initialize benchmark values to remove low density hexagons
@@
   scale_x_continuous("standardized bin count", 
+         transform = "reverse") + 
+  ylab("MSE") +
+  ggtitle("(c)") +
+  theme(aspect.ratio = 0.75,
+        axis.text.x = element_text(size = 12),
+        axis.text.y = element_text(size = 12),
+        axis.title.x = element_text(size = 12),
+        axis.title.y = element_text(size = 12, 
+            angle=90),
+        plot.title = element_text(size = 12))
+```
+
+```{r}
+#| label: diff-bins-scurve
+## To initialize benchmark values to remove low density hexagons
+benchmark_rm_hex_vec <- seq(0, 0.8, by=0.1)
+bin1_vec_scurve <- c(7, 11, 17)
+error_rm_scurve <- data.frame(matrix(nrow = 0, ncol = 0))
+
+for (xbins in bin1_vec_scurve) {
+  #for (buff in buffer_vec) {
+
+  for (benchmark_rm_lwd in benchmark_rm_hex_vec) {
+
+    bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.07)$bin2
+
+    scurve_model <- fit_highd_model(
+      training_data = training_data_scurve,
+      emb_df = umap_scurve_scaled,
+      bin1 = xbins,
+      r2 = r2,
+      q = 0.07, 
+      is_bin_centroid = TRUE,
+      is_rm_lwd_hex = FALSE,
+      col_start_highd = "x"
+    )
+
+    df_bin_centroids_scurve <- scurve_model$df_bin_centroids
+    df_bin_scurve <- scurve_model$df_bin
+
+    df_bin_centroids_scurve_high_dens <- df_bin_centroids_scurve |>
+      filter(std_counts > benchmark_rm_lwd)
+
+    df_bin_scurve_high_dens <- df_bin_scurve |>
+      filter(hb_id %in% df_bin_centroids_scurve_high_dens$hexID)
+
+    ## Compute error
+    error_df <- glance(
+      df_bin_centroids = df_bin_centroids_scurve_high_dens,
+      df_bin = df_bin_scurve_high_dens,
+      training_data = training_data_scurve,
+      newdata = NULL,
+      type_NLDR = "UMAP",
+      col_start = "x") |>
+      mutate(benchmark_rm_lwd = benchmark_rm_lwd,
+             bin1 = xbins,
+             b_non_empty = NROW(df_bin_centroids_scurve_high_dens))
+
+    error_rm_scurve <- bind_rows(error_rm_scurve, error_df)
+
+  }
+
+}
+
+benchmark_label_df <- tibble(x = 0.2, y = 0.1, 
+                          label = paste0("benchmark is ", benchmark1))
+
+mse_scurve_lwd <- ggplot(error_rm_scurve, 
+                     aes(x = benchmark_rm_lwd, 
+                         y = MSE,
+                         colour = as.factor(bin1),
+                         group = as.factor(bin1))) + 
+  geom_point(
+    size = 1
+    ) +
+  geom_line(
+    linewidth = 0.3
+    ) + 
+  scale_color_brewer(type = "qual") +
+  # geom_vline(xintercept = benchmark1, linetype="solid", 
+  #            color = "#bdbdbd", linewidth=1, alpha = 0.5) +
+  scale_x_continuous("bechmark to remove low-density hexagons", 
          transform = "reverse") + 

### 2024-08-19 paper.tex
Updated Figure 10

@@
 \begin{center}
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-35-1.pdf}
+\includegraphics[width=1\textwidth,height=0.3\textheight]{paper_files/figure-pdf/unnamed-chunk-35-1.pdf}
+\end{center}
+
+\begin{center}
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-36-1.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-64-1.pdf}
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-66-1.pdf}
 \end{center}

### 2024-08-19 paper.qmd
Removed unnecessary codes

@@
 
-<!--Using first quantile as the benchmark value-->
-```{r}
-#| label: bin-benchmark-scurve
-hex_grid_with_counts_s_curve2 <- full_join(
-  hex_grid2, 
-  df_bin_centroids_scurve |> select(hexID, std_counts), 
-  by = c("hex_poly_id" = "hexID"))
-
-hex_grid_with_low_counts <- full_join(hex_grid2, 
-    df_bin_centroids_scurve, 
-    by = c("hex_poly_id" = "hexID")) |>
-  filter(!is.na(std_counts)) |>
-  mutate(type = if_else(std_counts <= benchmark1, "low", "high")) |>
-  filter(type == "low")
-
-hex_grid_rm_lwd_scurve1 <-  ggplot(
-  data = hex_grid_with_counts_s_curve2, 
-  aes(x = x, y = y)) + 
-  geom_polygon(aes(group = hex_poly_id, fill = std_counts), 
-            color = "grey70", linewidth=0.2) +
-  geom_text(data = hex_grid_with_low_counts, 
-            color = "black", 
-            aes(x = c_x, y = c_y, group = hex_poly_id, label = "X", size = 1)) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  xlim(sc_xlims) + ylim(sc_ylims) +
-  interior_annotation("a")
-```
-
 <!--Observing neighboring bins as well-->
@@
 
-hex_grid_with_low_counts_neighbor <- right_join(hex_grid, 
-                                             df_bin_centroids_scurve_removed, 
-                                             by = c("hex_poly_id" = "hexID"))
-
-hex_grid_rm_lwd_scurve2 <-  ggplot(
-  data = hex_grid_with_counts_s_curve2, 
-  aes(x = x, y = y)) + 
-  geom_polygon(aes(group = hex_poly_id, fill = std_counts), 
-            color = "grey70", linewidth=0.2) +
-  geom_text(data = hex_grid_with_low_counts_neighbor, 
-            color = "black", 
-            aes(x = c_x, y = c_y, group = hex_poly_id, label = "X", size = 1)) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
-  xlim(sc_xlims) + ylim(sc_ylims) +
-  interior_annotation("b")
 ```

### 2024-08-19 paper.qmd
Try to add more plots

@@
 ```{r}
-#| label: diff-bins-scurve
+#| label: diff-bins-scurve-buff
 ## To initialize benchmark values to remove low density hexagons
-benchmark_rm_hex_vec <- seq(0, 0.8, by=0.1)
+benchmark_rm_hex_vec <- seq(0, 0.7, by=0.1)
 bin1_vec_scurve <- c(7, 11, 17)
@@
 
-    bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.07)$bin2
+    a1 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.07)$a1
 
@@
       filter(hb_id %in% df_bin_centroids_scurve_high_dens$hexID)
+
+    tr1_object_scurve <- tri_bin_centroids(
+      df_bin_centroids_scurve_high_dens, 
+      x = "c_x", 
+      y = "c_y")
+
+    tr_from_to_df_scurve <- gen_edges(
+      tri_object = tr1_object_scurve)
+
+    ## Compute 2D distances
+    distance_scurve <- cal_2d_dist(
+      tr_coord_df = tr_from_to_df_scurve, 
+      start_x = "x_from", 
+      start_y = "y_from", 
+      end_x = "x_to", 
+      end_y = "y_to", 
+      select_vars = c("from", "to", "distance"))
 
@@
 }
-
-benchmark_label_df <- tibble(x = 0.2, y = 0.1, 
-                          label = paste0("benchmark is ", benchmark1))
 
@@
 <!--distribution of distance along with the default benchmark-->
+
+```{r}
+#| label: diff-bins-scurve-buff3
+#| eval: false
+## To initialize benchmark values to remove low density hexagons
+benchmark_rm_hex_vec <- seq(0, 0.7, by=0.1)
+bin1_vec_scurve <- c(7, 11, 17)
+distance_all_scurve <- data.frame(matrix(nrow = 0, ncol = 0))
+#buffer_vec <- 0.07
+
+for (xbins in bin1_vec_scurve) {
+  for (buff in buffer_vec) {
+
+  #for (benchmark_rm_lwd in benchmark_rm_hex_vec) {
+
+    a1 <- calc_bins_y(bin1 = xbins, r2 = r2, q = buff)$a1
+
+    scurve_model <- fit_highd_model(
+      training_data = training_data_scurve,
+      emb_df = umap_scurve_scaled,
+      bin1 = xbins,
+      r2 = r2,
+      q = buff, 
+      is_bin_centroid = TRUE,
+      is_rm_lwd_hex = FALSE,
+      col_start_highd = "x"
+    )
+
+    df_bin_centroids_scurve <- scurve_model$df_bin_centroids
+    df_bin_scurve <- scurve_model$df_bin
+
+    df_bin_centroids_scurve_high_dens <- df_bin_centroids_scurve |>
+      filter(std_counts > benchmark_rm_lwd)
+
+    df_bin_scurve_high_dens <- df_bin_scurve |>
+      filter(hb_id %in% df_bin_centroids_scurve_high_dens$hexID)
+
+    tr1_object_scurve <- tri_bin_centroids(
+      df_bin_centroids_scurve_high_dens, 
+      x = "c_x", 
+      y = "c_y")
+
+    tr_from_to_df_scurve <- gen_edges(
+      tri_object = tr1_object_scurve)
+
+    ## Compute 2D distances
+    distance_scurve <- cal_2d_dist(
+      tr_coord_df = tr_from_to_df_scurve, 
+      start_x = "x_from", 
+      start_y = "y_from", 
+      end_x = "x_to", 
+      end_y = "y_to", 
+      select_vars = c("from", "to", "distance"))
+
+    start_point <- a1/2
+
+    # Create bins and calculate the mean value for each bin range
+    bins <- seq(start_point, max(distance_scurve$distance) + a1, 
+                by = a1)
+    bin_labels <- bins[-length(bins)] + a1 / 2  # mean of each bin range
+
+    distance_scurve <- distance_scurve |> 
+      mutate(bin = cut(distance, breaks = bins, include.lowest = TRUE, labels = bin_labels))
+
+    ## Compute error
+    distance_scurve <- distance_scurve |>
+      mutate(q = buff,
+             bin1 = xbins,
+             a1 = a1)
+
+    distance_all_scurve <- bind_rows(distance_all_scurve, distance_scurve)
+
+  }
+
+}
+
+# dist_summary_scurve <- distance_all_scurve |>
+#     count(bin1, q, bin)
+
+distance_hist <- distance_all_scurve |>
+  filter(bin1 == 11) |>
+  filter(q == 0.07) |>
+  ggplot(aes(x = distance)) +
+    geom_histogram(
+      aes(y = after_stat(count) / sum(after_stat(count)), 
+      group = as.factor(q),
+      fill = as.factor(q)),
+      binwidth = a1,
+      position = "dodge") +
+  scale_x_continuous(breaks = bin_labels, labels = round(bin_labels, 3)) +
+  scale_color_brewer(type = "qual") +
+  #facet_wrap(vars(as.factor(bin1)), scale = "free") +
+  ylab("standardized number of edges") +
+  xlab(expression(d^{(2)})) +
+  ggtitle("(d)") +
+  theme(aspect.ratio = 0.75,
+        axis.title.x = element_text(size = 12),
+        axis.title.y = element_text(size = 12, angle = 90),
+        axis.text.x = element_text(size = 12, angle = 90),
+        axis.text.y = element_text(size = 12),
+        axis.ticks.x = element_line(), 
+        plot.title = element_text(size = 12))
+
+```
+
 ```{r}
 #| label: triangulate-scurve2
+
 ## Triangulate bin centroids
@@
 # Define bin width and starting point
-bin_width <- hb_obj_scurve$centroids$c_x[2]-hb_obj_scurve2$centroids$c_x[1]
+bin_width <- hb_obj_scurve2$centroids$c_x[2]-hb_obj_scurve2$centroids$c_x[1]
 start_point <- bin_width/2

### 2024-08-20 paper.qmd
Removed plots with q

@@
 ```{r}
-#| label: bin-counts-scurve-with_buffer
-## To initialize number of bins along the x-axis
-bin1_vec_scurve <- c(7, 11, 17) #sqrt(NROW(training_data_scurve)/2)
-#buffer_vec <- seq(0.05, 0.16, by = 0.01)
-buffer_vec <- c(0.05, 0.07, 0.12)
-
-centroids_df_scurve <- data.frame(matrix(nrow = 0, ncol = 0))
-
-for (xbins in bin1_vec_scurve) {
-  for (buff in buffer_vec) {
-
-    bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = buff)$bin2
-
-    scurve_model <- fit_highd_model(
-      training_data = training_data_scurve,
-      emb_df = umap_scurve_scaled,
-      bin1 = xbins,
-      r2 = r2,
-      q = buff, 
-      is_bin_centroid = TRUE,
-      is_rm_lwd_hex = FALSE,
-      col_start_highd = "x"
-    )
-
-    df_bin_centroids_scurve <- scurve_model$df_bin_centroids
-    #df_bin_scurve <- scurve_model$df_bin
-
-    ## Compute error
-    df_bin_centroids_scurve <- df_bin_centroids_scurve |>
-      mutate(bin1 = xbins,
-             bin2 = bin2,
-             b = bin1 * bin2,
-             b_non_empty = NROW(df_bin_centroids_scurve),
-             buff = buff)
-
-    centroids_df_scurve <- bind_rows(centroids_df_scurve, df_bin_centroids_scurve)
-
-  }
-
-}
-
-b7_q_plt <- centroids_df_scurve |>
-  filter(bin1 == 7) |>
-  filter(buff == 0.12) |>
-  ggplot(aes(x = bin_counts)) +
-  geom_density() +
-  xlab("bin count") +
-  theme(
-    axis.text.x = element_text(size = 7),
-    axis.ticks.x = element_line(linewidth = 0.5),
-    axis.title.x = element_text(size = 7)
-
-  )
-
-b11_q_plt <- centroids_df_scurve |>
-  filter(bin1 == 11) |>
-  filter(buff == 0.07) |>
-  ggplot(aes(x = bin_counts)) +
-  geom_density() +
-  xlab("bin count") +
-  theme(
-    axis.text.x = element_text(size = 7),
-    axis.ticks.x = element_line(linewidth = 0.5), 
-    axis.title.x = element_text(size = 7)
-
-  )
-
-b17_q_plt <- centroids_df_scurve |>
-  filter(bin1 == 17) |>
-  filter(buff == 0.05) |>
-  ggplot(aes(x = bin_counts)) +
-  geom_density() +
-  xlab("bin count") +
-  theme(
-    axis.text.x = element_text(size = 7),
-    axis.ticks.x = element_line(linewidth = 0.5),
-    axis.title.x = element_text(size = 7)
-
-  )
-```
-
-```{r}
-#| label: fig-bin-counts-scurve-with_buffer
-#| fig-cap: "Distribution of bin counts for differnt $b_1$ with specific $q$"
-#| out-height: 30%
-
-b7_q_plt + b11_q_plt + 
-  b17_q_plt +
-  plot_layout(ncol = 3)
-
-```
-
-```{r}
-#| out-height: 30%
-
-b7_q_plt2 <- centroids_df_scurve |>
-  filter(bin1 == 7) |>
-  ggplot(aes(x = std_counts,
-             colour = as.factor(buff),
-             group = as.factor(buff))) +
-  geom_density() +
-  scale_color_brewer(type = "qual") +
-  xlab("standardised bin count") +
-  theme(
-    axis.text.x = element_text(size = 7),
-    axis.ticks.x = element_line(linewidth = 0.5),
-    axis.title.x = element_text(size = 7)
-
-  )
-
-b11_q_plt2 <- centroids_df_scurve |>
-  filter(bin1 == 11) |>
-  ggplot(aes(x = std_counts,
-             colour = as.factor(buff),
-             group = as.factor(buff))) +
-  geom_density() +
-  scale_color_brewer(type = "qual") +
-  xlab("standardised bin count") +
-  theme(
-    axis.text.x = element_text(size = 7),
-    axis.ticks.x = element_line(linewidth = 0.5),
-    axis.title.x = element_text(size = 7)
-  )
-
-b17_q_plt2 <- centroids_df_scurve |>
-  filter(bin1 == 17) |>
-  ggplot(aes(x = std_counts,
-             colour = as.factor(buff),
-             group = as.factor(buff))) +
-  geom_density() +
-  scale_color_brewer(type = "qual") +
-  xlab("standardised bin count") +
-  theme(
-    axis.text.x = element_text(size = 7),
-    axis.ticks.x = element_line(linewidth = 0.5),
-    axis.title.x = element_text(size = 7)
-  )
-
-b7_q_plt2 + b11_q_plt2 + b17_q_plt2
-```
-
-```{r}
-#| out-height: 30%
-
-centroids_df_scurve |>
-  filter(bin1 == 7) |>
-  ggplot(aes(x = as.factor(hexID), 
-           y = std_counts,
-           group = as.factor(buff),
-           color = as.factor(buff))) + 
-  geom_line() +
-  scale_color_brewer(type = "qual")
-```
-
-```{r}
-## To initialize number of bins along the x-axis
-bin1_vec_scurve <- 5:19 #sqrt(NROW(training_data_scurve)/2)
-#buffer_vec <- seq(0.05, 0.16, by = 0.01)
-buffer_vec <- c(0.05, 0.07, 0.12)
-
-error_df_scurve_q <- data.frame(matrix(nrow = 0, ncol = 0))
-
-for (xbins in bin1_vec_scurve) {
-  for (buff in buffer_vec) {
-
-    bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = buff)$bin2
-
-    scurve_model <- fit_highd_model(
-      training_data = training_data_scurve,
-      emb_df = umap_scurve_scaled,
-      bin1 = xbins,
-      r2 = r2,
-      q = buff, 
-      is_bin_centroid = TRUE,
-      is_rm_lwd_hex = FALSE,
-      col_start_highd = "x"
-    )
-
-    df_bin_centroids_scurve <- scurve_model$df_bin_centroids
-    df_bin_scurve <- scurve_model$df_bin
-
-    ## Compute error
-    error_df <- glance(
-      df_bin_centroids = df_bin_centroids_scurve,
-      df_bin = df_bin_scurve,
-      training_data = training_data_scurve,
-      newdata = NULL,
-      type_NLDR = "UMAP",
-      col_start = "x") |>
-      mutate(bin1 = xbins,
-             bin2 = bin2,
-             b = bin1 * bin2,
-             b_non_empty = NROW(df_bin_centroids_scurve),
-             q = buff)
-
-    error_df_scurve_q <- bind_rows(error_df_scurve_q, error_df)
-
-  }
-
-}
-
-ggplot(error_df_scurve_q, 
-       aes(x = bin1, 
-           y = MSE,
-           colour = as.factor(q),
-           group = as.factor(q))) + 
-  geom_point(size = 1) +
-  geom_line() +
-  scale_color_brewer(type = "qual") +
-  scale_x_continuous(breaks = 5:19) +
-  labs(x = expression(b[1]), y = "MSE") +
-  theme(aspect.ratio = 0.75,
-        axis.text.x = element_text(size = 12),
-        axis.text.y = element_text(size = 12),
-        axis.title.x = element_text(size = 12),
-        axis.title.y = element_text(size = 12, 
-            angle=90),
-        axis.ticks.x = element_line(linewidth = 0.5),
-        axis.ticks.y = element_line(linewidth = 0.5),
-        plot.title = element_text(size = 12),
-        legend.title = element_text(size = 12),
-        legend.text = element_text(size = 10),
-        legend.key.height = unit(1, 'cm'),
-        legend.key.width = unit(1, 'cm'),
-        legend.key.size = unit(1, 'cm'))
-
-```
-
-```{r}
 #| label: errors-scurve

### 2024-08-20 paper.qmd
Added b1 Vs m plot for the diagnostics

@@
   geom_line(linewidth = 0.05) + 
-  geom_vline(xintercept = 11, linetype="solid",
-             color = "#bdbdbd", linewidth=1) +
-  geom_vline(xintercept = 9, linetype=2,
-             color = "#bdbdbd", linewidth=1) +
-  geom_vline(xintercept = 13, linetype=2,
-             color = "#bdbdbd", linewidth=1) +
+  # geom_vline(xintercept = 11, linetype="solid",
+  #            color = "#bdbdbd", linewidth=1) +
+  # geom_vline(xintercept = 9, linetype=2,
+  #            color = "#bdbdbd", linewidth=1) +
+  # geom_vline(xintercept = 13, linetype=2,
+  #            color = "#bdbdbd", linewidth=1) +
   scale_x_continuous(breaks = 5:19) +
   labs(x = expression(b[1]), y = "MSE") +
+  ggtitle("(b)") +
+  theme(aspect.ratio = 0.75,
+        axis.text.x = element_text(size = 12),
+        axis.text.y = element_text(size = 12),
+        axis.title.x = element_text(size = 12),
+        axis.title.y = element_text(size = 12, 
+            angle=90),
+        axis.ticks.x = element_line(linewidth = 0.5),
+        axis.ticks.y = element_line(linewidth = 0.5),
+        plot.title = element_text(size = 12),
+        legend.title = element_text(size = 12),
+        legend.text = element_text(size = 10),
+        legend.key.height = unit(1, 'cm'),
+        legend.key.width = unit(1, 'cm'),
+        legend.key.size = unit(1, 'cm'))
+
+b1_m_scurve <- ggplot(error_scurve, 
+                     aes(x = bin1, 
+                         y = b_non_empty)) + 
+  geom_point(size = 1) +
+  geom_line(linewidth = 0.05) + 
+  # geom_vline(xintercept = 11, linetype="solid",
+  #            color = "#bdbdbd", linewidth=1) +
+  # geom_vline(xintercept = 9, linetype=2,
+  #            color = "#bdbdbd", linewidth=1) +
+  # geom_vline(xintercept = 13, linetype=2,
+  #            color = "#bdbdbd", linewidth=1) +
+  scale_x_continuous(breaks = 5:19) +
+  labs(x = expression(b[1]), y = expression(m)) +
   ggtitle("(b)") +
@@
 
-cell_count_scurve + mse_scurve_b +  
+b1_m_scurve + mse_scurve_b +  
 mse_scurve_lwd + 

### 2024-08-20 paper.qmd
Updated different b1 choices with fixed q

@@
   bin1 = 7, 
-  r2 = r2,
-  q = 0.12)
+  r2 = r2)
 
@@
   bin1 = 7, 
-  r2 = r2,
-  q = 0.12
+  r2 = r2
 )$a1
@@
   bin1 = 11, 
-  r2 = r2,
-  q = 0.07)
+  r2 = r2)
 
@@
   bin1 = 11, 
-  r2 = r2,
-  q = 0.07
+  r2 = r2
   )$a1
@@
   bin1 = 17, 
-  r2 = r2,
-  q = 0.05)
+  r2 = r2)
 
@@
   bin1 = 17, 
-  r2 = r2,
-  q = 0.05
+  r2 = r2
 )$a1
@@
   bin1 = effective_bin1_scurve, 
-  r2 = r2, 
-  q = 0.07)$bin2
+  r2 = r2)$bin2
 
@@
 #| label: bins-scurve
-#| eval: false
+
 ## To initialize benchmark values to remove low density hexagons
@@
   scale_x_continuous("standardized bin count", 
-         transform = "reverse") + 
-  ylab("MSE") +
-  ggtitle("(c)") +
-  theme(aspect.ratio = 0.75,
-        axis.text.x = element_text(size = 12),
-        axis.text.y = element_text(size = 12),
-        axis.title.x = element_text(size = 12),
-        axis.title.y = element_text(size = 12, 
-            angle=90),
-        plot.title = element_text(size = 12))
-```
-
-```{r}
-#| label: diff-bins-scurve-buff
-## To initialize benchmark values to remove low density hexagons
-benchmark_rm_hex_vec <- seq(0, 0.7, by=0.1)
-bin1_vec_scurve <- c(7, 11, 17)
-error_rm_scurve <- data.frame(matrix(nrow = 0, ncol = 0))
-
-for (xbins in bin1_vec_scurve) {
-  #for (buff in buffer_vec) {
-
-  for (benchmark_rm_lwd in benchmark_rm_hex_vec) {
-
-    a1 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.07)$a1
-
-    scurve_model <- fit_highd_model(
-      training_data = training_data_scurve,
-      emb_df = umap_scurve_scaled,
-      bin1 = xbins,
-      r2 = r2,
-      q = 0.07, 
-      is_bin_centroid = TRUE,
-      is_rm_lwd_hex = FALSE,
-      col_start_highd = "x"
-    )
-
-    df_bin_centroids_scurve <- scurve_model$df_bin_centroids
-    df_bin_scurve <- scurve_model$df_bin
-
-    df_bin_centroids_scurve_high_dens <- df_bin_centroids_scurve |>
-      filter(std_counts > benchmark_rm_lwd)
-
-    df_bin_scurve_high_dens <- df_bin_scurve |>
-      filter(hb_id %in% df_bin_centroids_scurve_high_dens$hexID)
-
-    tr1_object_scurve <- tri_bin_centroids(
-      df_bin_centroids_scurve_high_dens, 
-      x = "c_x", 
-      y = "c_y")
-
-    tr_from_to_df_scurve <- gen_edges(
-      tri_object = tr1_object_scurve)
-
-    ## Compute 2D distances
-    distance_scurve <- cal_2d_dist(
-      tr_coord_df = tr_from_to_df_scurve, 
-      start_x = "x_from", 
-      start_y = "y_from", 
-      end_x = "x_to", 
-      end_y = "y_to", 
-      select_vars = c("from", "to", "distance"))
-
-    ## Compute error
-    error_df <- glance(
-      df_bin_centroids = df_bin_centroids_scurve_high_dens,
-      df_bin = df_bin_scurve_high_dens,
-      training_data = training_data_scurve,
-      newdata = NULL,
-      type_NLDR = "UMAP",
-      col_start = "x") |>
-      mutate(benchmark_rm_lwd = benchmark_rm_lwd,
-             bin1 = xbins,
-             b_non_empty = NROW(df_bin_centroids_scurve_high_dens))
-
-    error_rm_scurve <- bind_rows(error_rm_scurve, error_df)
-
-  }
-
-}
-
-mse_scurve_lwd <- ggplot(error_rm_scurve, 
-                     aes(x = benchmark_rm_lwd, 
-                         y = MSE,
-                         colour = as.factor(bin1),
-                         group = as.factor(bin1))) + 
-  geom_point(
-    size = 1
-    ) +
-  geom_line(
-    linewidth = 0.3
-    ) + 
-  scale_color_brewer(type = "qual") +
-  # geom_vline(xintercept = benchmark1, linetype="solid", 
-  #            color = "#bdbdbd", linewidth=1, alpha = 0.5) +
-  scale_x_continuous("bechmark to remove low-density hexagons", 
          transform = "reverse") + 

### 2024-08-20 paper.qmd
Added plot with b1 Vs m

@@
                      aes(x = bin1, 
+                         y = b_non_empty/b)) + 
+  geom_point(size = 1) +
+  geom_line(linewidth = 0.05) + 
+  # geom_vline(xintercept = 11, linetype="solid",
+  #            color = "#bdbdbd", linewidth=1) +
+  # geom_vline(xintercept = 9, linetype=2,
+  #            color = "#bdbdbd", linewidth=1) +
+  # geom_vline(xintercept = 13, linetype=2,
+  #            color = "#bdbdbd", linewidth=1) +
+  scale_x_continuous(breaks = 5:19) +
+  labs(x = expression(b[1]), y = expression(frac(m, b))) +
+  ggtitle("(a)") +
+  theme(aspect.ratio = 0.75,
+        axis.text.x = element_text(size = 12),
+        axis.text.y = element_text(size = 12),
+        axis.title.x = element_text(size = 12),
+        axis.title.y = element_text(size = 12, 
+            angle=90),
+        axis.ticks.x = element_line(linewidth = 0.5),
+        axis.ticks.y = element_line(linewidth = 0.5),
+        plot.title = element_text(size = 12),
+        legend.title = element_text(size = 12),
+        legend.text = element_text(size = 10),
+        legend.key.height = unit(1, 'cm'),
+        legend.key.width = unit(1, 'cm'),
+        legend.key.size = unit(1, 'cm'))
+
+b1_m_scurve2 <- ggplot(error_scurve, 
+                     aes(x = bin1, 
                          y = b_non_empty)) + 
@@
   labs(x = expression(b[1]), y = expression(m)) +
-  ggtitle("(b)") +
+  ggtitle("(c)") +
   theme(aspect.ratio = 0.75,
@@
 
-b1_m_scurve + mse_scurve_b +  
+b1_m_scurve + mse_scurve_b + b1_m_scurve2 + 
 mse_scurve_lwd + 

### 2024-08-20 paper.qmd
MSE for different lwd values

@@
         plot.title = element_text(size = 12))
+
+mse_scurve_lwd2 <- error_rm_scurve |>
+  filter(b_non_empty >= 5) |>
+  ggplot(aes(x = b_non_empty,y = MSE)) + 
+  geom_point(
+    size = 1
+    ) +
+  geom_line(
+    linewidth = 0.3
+    ) + 
+  # geom_vline(xintercept = benchmark1, linetype="solid", 
+  #            color = "#bdbdbd", linewidth=1, alpha = 0.5) +
+  scale_x_continuous("m (b1 = 11)") + 
+  ylab("MSE") +
+  ggtitle("(e)") +
+  theme(aspect.ratio = 0.75,
+        axis.text.x = element_text(size = 12),
+        axis.text.y = element_text(size = 12),
+        axis.title.x = element_text(size = 12),
+        axis.title.y = element_text(size = 12, 
+            angle=90),
+        plot.title = element_text(size = 12))
 ```
@@
 b1_m_scurve + mse_scurve_b + b1_m_scurve2 + 
-mse_scurve_lwd + 
+mse_scurve_lwd + mse_scurve_lwd2 +
 distance_hist + plot_layout(ncol=2)

### 2024-08-21 paper.qmd
Changed the plot (a) for Figure 9

@@
 
-  bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.07)$bin2
+  bin_obj <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.1)
+
+  bin2 <- bin_obj$bin2
+  a1 <- bin_obj$a1
 
@@
            b = bin1 * bin2,
-           b_non_empty = NROW(df_bin_centroids_scurve))
+           b_non_empty = NROW(df_bin_centroids_scurve),
+           a1 = a1)
 
@@
 
+diaconis_bin_width <- calc_bins_y(
+    bin1 = ceiling(NROW(training_data_scurve)^(1/3)),
+    r2 = r2,
+    q = 0.1)$a1
+
 mse_scurve_b <- ggplot(error_scurve, 
-                     aes(x = bin1, 
+                     aes(x = a1, 
                          y = MSE)) + 
@@
   #            color = "#bdbdbd", linewidth=1) +
-  # geom_vline(xintercept = 9, linetype=2,
-  #            color = "#bdbdbd", linewidth=1) +
+  geom_vline(xintercept = diaconis_bin_width, 
+             linetype=2, color = "#bdbdbd", 
+             linewidth=1) +
   # geom_vline(xintercept = 13, linetype=2,
   #            color = "#bdbdbd", linewidth=1) +
-  scale_x_continuous(breaks = 5:19) +
-  labs(x = expression(b[1]), y = "MSE") +
-  ggtitle("(b)") +
-  theme(aspect.ratio = 0.75,
-        axis.text.x = element_text(size = 12),
-        axis.text.y = element_text(size = 12),
-        axis.title.x = element_text(size = 12),
-        axis.title.y = element_text(size = 12, 
-            angle=90),
-        axis.ticks.x = element_line(linewidth = 0.5),
-        axis.ticks.y = element_line(linewidth = 0.5),
-        plot.title = element_text(size = 12),
-        legend.title = element_text(size = 12),
-        legend.text = element_text(size = 10),
-        legend.key.height = unit(1, 'cm'),
-        legend.key.width = unit(1, 'cm'),
-        legend.key.size = unit(1, 'cm'))
-
-b1_m_scurve <- ggplot(error_scurve, 
-                     aes(x = bin1, 
-                         y = b_non_empty/b)) + 
-  geom_point(size = 1) +
-  geom_line(linewidth = 0.05) + 
-  # geom_vline(xintercept = 11, linetype="solid",
-  #            color = "#bdbdbd", linewidth=1) +
-  # geom_vline(xintercept = 9, linetype=2,
-  #            color = "#bdbdbd", linewidth=1) +
-  # geom_vline(xintercept = 13, linetype=2,
-  #            color = "#bdbdbd", linewidth=1) +
-  scale_x_continuous(breaks = 5:19) +
-  labs(x = expression(b[1]), y = expression(frac(m, b))) +
+  #scale_x_continuous(breaks = 5:19) +
+  labs(x = expression(a[1]), y = "MSE") +
   ggtitle("(a)") +
@@
 
-b1_m_scurve2 <- ggplot(error_scurve, 
-                     aes(x = bin1, 
-                         y = b_non_empty)) + 
-  geom_point(size = 1) +
-  geom_line(linewidth = 0.05) + 
-  # geom_vline(xintercept = 11, linetype="solid",
-  #            color = "#bdbdbd", linewidth=1) +
-  # geom_vline(xintercept = 9, linetype=2,
-  #            color = "#bdbdbd", linewidth=1) +
-  # geom_vline(xintercept = 13, linetype=2,
-  #            color = "#bdbdbd", linewidth=1) +
-  scale_x_continuous(breaks = 5:19) +
-  labs(x = expression(b[1]), y = expression(m)) +
-  ggtitle("(c)") +
-  theme(aspect.ratio = 0.75,
-        axis.text.x = element_text(size = 12),
-        axis.text.y = element_text(size = 12),
-        axis.title.x = element_text(size = 12),
-        axis.title.y = element_text(size = 12, 
-            angle=90),
-        axis.ticks.x = element_line(linewidth = 0.5),
-        axis.ticks.y = element_line(linewidth = 0.5),
-        plot.title = element_text(size = 12),
-        legend.title = element_text(size = 12),
-        legend.text = element_text(size = 10),
-        legend.key.height = unit(1, 'cm'),
-        legend.key.width = unit(1, 'cm'),
-        legend.key.size = unit(1, 'cm'))
+# b1_m_scurve <- ggplot(error_scurve, 
+#                      aes(x = bin1, 
+#                          y = b_non_empty/b)) + 
+#   geom_point(size = 1) +
+#   geom_line(linewidth = 0.05) + 
+#   # geom_vline(xintercept = 11, linetype="solid",
+#   #            color = "#bdbdbd", linewidth=1) +
+#   # geom_vline(xintercept = 9, linetype=2,
+#   #            color = "#bdbdbd", linewidth=1) +
+#   # geom_vline(xintercept = 13, linetype=2,
+#   #            color = "#bdbdbd", linewidth=1) +
+#   scale_x_continuous(breaks = 5:19) +
+#   labs(x = expression(b[1]), y = expression(frac(m, b))) +
+#   ggtitle("(a)") +
+#   theme(aspect.ratio = 0.75,
+#         axis.text.x = element_text(size = 12),
+#         axis.text.y = element_text(size = 12),
+#         axis.title.x = element_text(size = 12),
+#         axis.title.y = element_text(size = 12, 
+#             angle=90),
+#         axis.ticks.x = element_line(linewidth = 0.5),
+#         axis.ticks.y = element_line(linewidth = 0.5),
+#         plot.title = element_text(size = 12),
+#         legend.title = element_text(size = 12),
+#         legend.text = element_text(size = 10),
+#         legend.key.height = unit(1, 'cm'),
+#         legend.key.width = unit(1, 'cm'),
+#         legend.key.size = unit(1, 'cm'))
+# 
+# b1_m_scurve2 <- ggplot(error_scurve, 
+#                      aes(x = bin1, 
+#                          y = b_non_empty)) + 
+#   geom_point(size = 1) +
+#   geom_line(linewidth = 0.05) + 
+#   # geom_vline(xintercept = 11, linetype="solid",
+#   #            color = "#bdbdbd", linewidth=1) +
+#   # geom_vline(xintercept = 9, linetype=2,
+#   #            color = "#bdbdbd", linewidth=1) +
+#   # geom_vline(xintercept = 13, linetype=2,
+#   #            color = "#bdbdbd", linewidth=1) +
+#   scale_x_continuous(breaks = 5:19) +
+#   labs(x = expression(b[1]), y = expression(m)) +
+#   ggtitle("(c)") +
+#   theme(aspect.ratio = 0.75,
+#         axis.text.x = element_text(size = 12),
+#         axis.text.y = element_text(size = 12),
+#         axis.title.x = element_text(size = 12),
+#         axis.title.y = element_text(size = 12, 
+#             angle=90),
+#         axis.ticks.x = element_line(linewidth = 0.5),
+#         axis.ticks.y = element_line(linewidth = 0.5),
+#         plot.title = element_text(size = 12),
+#         legend.title = element_text(size = 12),
+#         legend.text = element_text(size = 10),
+#         legend.key.height = unit(1, 'cm'),
+#         legend.key.width = unit(1, 'cm'),
+#         legend.key.size = unit(1, 'cm'))
 ```
@@
 
-b1_m_scurve + mse_scurve_b + b1_m_scurve2 + 
-mse_scurve_lwd + mse_scurve_lwd2 +
-distance_hist + plot_layout(ncol=2)
+mse_scurve_b + mse_scurve_lwd + 
+  mse_scurve_lwd2 + distance_hist + 
+  plot_layout(ncol=2)
 ```

### 2024-08-21 paper.qmd
Added Freedman's rule to offer a default

@@
 a1_1 <- calc_bins_y(
-  bin1 = 7, 
+  bin1 = 8, 
   r2 = r2
@@
   filter(drop_empty == FALSE) |>
-  mutate(b1 = "b1 = 7")
+  mutate(b1 = "b1 = 8")
 
@@
   data = umap_scurve_scaled, 
-  bin1 = 11, 
+  bin1 = 10, 
   r2 = r2)
@@
 a1_2 <- calc_bins_y(
-  bin1 = 11, 
+  bin1 = 10, 
   r2 = r2
@@
   filter(drop_empty == FALSE) |>
-  mutate(b1 = "b1 = 11")
+  mutate(b1 = "b1 = 10")
 
@@
   data = umap_scurve_scaled, 
-  bin1 = 17, 
+  bin1 = 14, 
   r2 = r2)
@@
 a1_3 <- calc_bins_y(
-  bin1 = 17, 
+  bin1 = 14, 
   r2 = r2
@@
   filter(drop_empty == FALSE) |>
-  mutate(b1 = "b1 = 17")
+  mutate(b1 = "b1 = 14")
 
@@
 #| fig-width: 6
-#| fig-height: 5
+#| fig-height: 4
 

### 2024-08-21 paper.qmd
Updated Figure 9 (e)

@@
 ## To initialize effective bins along x
-effective_bin1_scurve <- 11
+effective_bin1_scurve <- 14
 
@@
   #            color = "#bdbdbd", linewidth=1, alpha = 0.5) +
-  scale_x_continuous("m (b1 = 11)") + 
+  scale_x_continuous("m (b1 = 14)") + 
   ylab("MSE") +

### 2024-08-21 paper.qmd
Updated Figure 15 with new data recoreded with a1

@@
 error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_15_min_dist_0.99.rds")
-#error_pbmc_tsne <- read_rds("data/pbmc3k/error_pbmc_tsne_5.rds")
+error_pbmc_tsne <- read_rds("data/pbmc3k/error_pbmc_tsne_5.rds")
 error_pbmc_tsne2 <- read_rds("data/pbmc3k/error_pbmc_tsne_30.rds")
-# error_pbmc_phate <- read_rds("data/pbmc3k/error_pbmc_phate_5.rds") 
-# error_pbmc_trimap <- read_rds("data/pbmc3k/error_pbmc_trimap_12_4_3.rds") 
-# error_pbmc_pacmap <- read_rds("data/pbmc3k/error_pbmc_pacmap_30_random_0.9_5.rds")
+error_pbmc_phate <- read_rds("data/pbmc3k/error_pbmc_phate_5.rds")
+error_pbmc_trimap <- read_rds("data/pbmc3k/error_pbmc_trimap_12_4_3.rds")
+error_pbmc_pacmap <- read_rds("data/pbmc3k/error_pbmc_pacmap_30_random_0.9_5.rds")
 
@@
                         error_pbmc_umap3,
-                        error_pbmc_tsne2)
+                        error_pbmc_tsne,
+                        error_pbmc_tsne2,
+                        error_pbmc_phate,
+                        error_pbmc_trimap,
+                        error_pbmc_pacmap)
 
@@
 
-pbmc_layouts <- readPNG("figures/pbmc3k/pbmc_nldr_layouts.png", native = TRUE)
+#pbmc_layouts <- readPNG("figures/pbmc3k/pbmc_nldr_layouts.png", native = TRUE)
 
 error_plot_pbmc <- ggplot(error_pbmc, 
-                          aes(x = bin1, 
+                          aes(x = a1, 
                               y = MSE, 
@@
   #            color = "black", linewidth=0.8, alpha = 0.5) +
-  scale_color_manual(values=c('#ff7f00', '#999999', 
-                              '#a65628', '#f781bf')) +
+  scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a','#ff7f00',
+                              '#984ea3','#999999','#a65628','#f781bf')) +
   ylab("MSE") +
-  xlab(expression(b[1])) + 
+  xlab(expression(a[1])) + 
   theme(aspect.ratio = 0.75,
@@
         axis.ticks.y = element_line(linewidth = 0.5),
-        legend.position = "none") +
-  inset_element(p = pbmc_layouts, 
-                left = 1.55, 
-                bottom = 0.35, 
-                right = 0, 
-                top = 1)
+        legend.position = "none") #+
+  # inset_element(p = pbmc_layouts, 
+  #               left = 1.55, 
+  #               bottom = 0.35, 
+  #               right = 0, 
+  #               top = 1)
 

### 2024-08-21 paper.qmd
Replot Figure 15

@@
 ```{r}
-# label: read-pbmc-nldr
+#| label: read-pbmc-nldr
 # Read a variety of different NLDR representations of PBMC
@@
 
+nldr1c <- umap_pbmc |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2)) +
+  geom_point(alpha=0.1, size=1, colour='#a65628') +
+  interior_annotation("a")
+
 umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_5_min_dist_0.01.rds")
@@
   geom_point(alpha=0.1, size=1, colour=clr_choice) +
+  interior_annotation("b")
+
+nldr2c <- umap_pbmc |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2))+
+  geom_point(alpha=0.1, size=1, colour='#f781bf') +
   interior_annotation("b")
@@
 
+nldr3c <- umap_pbmc |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2))+
+  geom_point(alpha=0.1, size=1, colour='#999999') +
+  interior_annotation("c")
+
 tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_5.rds")
@@
 
+nldr4c <- tsne_pbmc |>
+  ggplot(aes(x = tSNE1,
+             y = tSNE2))+
+  geom_point(alpha=0.1, size=1, colour='#984ea3') +
+  interior_annotation("d")
+
 tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_30.rds")
@@
   geom_point(alpha=0.1, size=1, colour=clr_choice) +
+  interior_annotation("e")
+
+nldr5c <- tsne_pbmc |>
+  ggplot(aes(x = tSNE1,
+             y = tSNE2))+
+  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
   interior_annotation("e")
@@
 
+nldr6c <- phate_pbmc |>
+  ggplot(aes(x = PHATE1,
+             y = PHATE2))+
+  geom_point(alpha=0.1, size=1, colour='#377eb8') +
+  interior_annotation("f")
+
 trimap_pbmc <- read_rds("data/pbmc3k/pbmc_trimap_12_4_3.rds")
@@
 
+nldr7c <- trimap_pbmc |>
+  ggplot(aes(x = TriMAP1,
+             y = TriMAP2))+
+  geom_point(alpha=0.1, size=1, colour='#4daf4a') +
+  interior_annotation("g")
+
 pacmap_pbmc <- read_rds("data/pbmc3k/pbmc_pacmap_30_random_0.9_5.rds")
@@
   geom_point(alpha=0.1, size=1, colour=clr_choice) +
+  interior_annotation("h")
+
+nldr8c <- pacmap_pbmc |>
+  ggplot(aes(x = PaCMAP1,
+             y = PaCMAP2))+
+  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
   interior_annotation("h")
@@
 #| fig-pos: H
-#| out-width: 70%
-
-error_plot_pbmc
+#| out-height: 100%
+
+free(error_plot_pbmc) / wrap_plots(nldr1c, nldr2c, nldr3c, nldr4c,
+           nldr5c, nldr6c, nldr7c, nldr8c, ncol = 4)
 ```

### 2024-08-21 paper.qmd
Updated Figure 15

@@
                           aes(x = a1, 
-                              y = MSE, 
+                              y = log(MSE), 
                               colour = method)) + 
@@
                               '#984ea3','#999999','#a65628','#f781bf')) +
-  ylab("MSE") +
+  ylab("log(MSE)") +
   xlab(expression(a[1])) + 
-  theme(aspect.ratio = 0.75,
+  theme(#aspect.ratio = 0.75,
         axis.text.x = element_text(size = 7),
@@
 ```{r}
-#| fig-cap: "MSE from UMAP with diffferent (hyper-)parameter choices and tSNE with default (hyper-)parameter setting applied to PBMC3k dataset for different number of bins along the x-axis ($b_1$).
+#| fig-cap: "MSE from UMAP, tSNE, PHATE, PaCMAP, and TriMAP wwith diffferent (hyper-)parameter choices applied to PBMC3k dataset for different number of bin widths along the x-axis ($a_1$).
 Which is the most reasonable representation of the structures present in the PBMC3k data?
-The residual plot have a steep slope at the beginning, indicating that a smaller $b_1$ causes a larger amount of error.
-Then, the slope gradually declines or level off, indicating that a higher number of $b_1$ generates a smaller error.
-It was observed that for every $b_1$, the lowest error occurred with tSNE using the perplexity: $30$."
+The residual plot have a steep slope at the beginning, indicating that a smaller $a_1$ causes a smaller amount of error.
+Then, the slope gradually increases or level up, indicating that a higher number of $a_1$ generates a larger error.
+It was observed that for every $a_1$, the lowest error occurred with tSNE using the perplexity: $30$."
 #| label: fig-pbmc-mse

### 2024-08-21 paper.tex
Removed unnecessary codes
Removed plots with q
Replot Figure 15
Updated Figure 15

--- 2024-08-19 paper.tex
@@
 
-\includegraphics[width=0.7\textwidth,height=\textheight]{paper_files/figure-pdf/fig-pbmc-mse-1.pdf}
-
-}
-
-\caption{\label{fig-pbmc-mse}MSE from UMAP with diffferent
-(hyper-)parameter choices and tSNE with default (hyper-)parameter
-setting applied to PBMC3k dataset for different number of bins along the
-x-axis (\(b_1\)). Which is the most reasonable representation of the
-structures present in the PBMC3k data? The residual plot have a steep
-slope at the beginning, indicating that a smaller \(b_1\) causes a
-larger amount of error. Then, the slope gradually declines or level off,
-indicating that a higher number of \(b_1\) generates a smaller error. It
-was observed that for every \(b_1\), the lowest error occurred with tSNE
-using the perplexity: \(30\).}
+\includegraphics[width=1\textwidth,height=1\textheight]{paper_files/figure-pdf/fig-pbmc-mse-1.pdf}
+
+}
+
+\caption{\label{fig-pbmc-mse}MSE from UMAP, tSNE, PHATE, PaCMAP, and
+TriMAP wwith diffferent (hyper-)parameter choices applied to PBMC3k
+dataset for different number of bin widths along the x-axis (\(a_1\)).
+Which is the most reasonable representation of the structures present in
+the PBMC3k data? The residual plot have a steep slope at the beginning,
+indicating that a smaller \(a_1\) causes a smaller amount of error.
+Then, the slope gradually increases or level up, indicating that a
+higher number of \(a_1\) generates a larger error. It was observed that
+for every \(a_1\), the lowest error occurred with tSNE using the
+perplexity: \(30\).}
 

### 2024-08-21 paper.qmd
Render the projection after scaling as a valdation

@@
 
-axis_scale <- 2 * limits/3
-axis_pos <- -0.6 * limits
+# axis_scale <- 2 * limits/3
+# axis_pos <- -0.6 * limits
+axis_scale <- limits/6
+axis_pos <- -2/3 * limits
 
@@
 
-axis_scale <- 0.15 * limits
-axis_pos <- -0.6 * limits
-adj <- function(x) axis_pos + x * axis_scale
+# axis_scale <- 0.15 * limits
+# axis_pos <- -0.6 * limits
+# adj <- function(x) axis_pos + x * axis_scale
 theta <- seq(0, 2 * pi, length = 50)
@@
 
-axis_scale <- 2 * limits/3
-axis_pos <- -0.4 * limits
+# axis_scale <- 2 * limits/3
+# axis_pos <- -0.4 * limits
+
+axis_scale <- limits/6
+axis_pos <- -2/3 * limits
 
@@
 
-axis_scale <- 0.1 * limits
-axis_pos <- -0.4 * limits
-adj <- function(x) axis_pos + x * axis_scale
+# axis_scale <- 0.1 * limits
+# axis_pos <- -0.4 * limits
+# adj <- function(x) axis_pos + x * axis_scale
 theta <- seq(0, 2 * pi, length = 50)
@@
 #| fig-width: 10
-#| fig-height: 7
+#| fig-height: 6
 #| fig-pos: H

### 2024-08-21 scripts/scurve/validate_langevitour_with_scaled_scurve.R

@@ -35,3 +35,59 @@ scaled_s_curve_noise <- scale_data_manual(training_data_scurve |> select(-ID)) |
   as_tibble()
 
 langevitour(scaled_s_curve_noise, scale = 1)
+
+
+projection <- cbind(
+  c(0.5898,-0.4620,-0.1362,0.2047,-0.3680,0.1032,-0.4818),
+  c(0.1756,0.0257,0.3683,0.2400,-0.5518,-0.4573,0.5115))
+projected <- as.matrix(scaled_s_curve_noise) %*% projection
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+limits <- 1
+rng <- range(projected)
+projected <- projected/max(abs(rng))
+colnames(projected) <- c("P1", "P2")
+projected <- data.frame(projected)
+obs_labels <- as.character(1:nrow(training_data_scurve))
+
+axis_scale <- limits/6
+axis_pos <- -2/3 * limits
+
+adj <- function(x) axis_pos + x * axis_scale
+axes <- data.frame(x1 = adj(0),
+                   y1 = adj(0),
+                   x2 = adj(projection[, 1]),
+                   y2 = adj(projection[, 2]))
+
+axis_labels <- colnames(training_data_scurve |> select(-ID))
+rownames(axes) <- axis_labels
+
+theta <- seq(0, 2 * pi, length = 50)
+circle <- data.frame(c1 = adj(cos(theta)), c2 = adj(sin(theta)))
+
+projected_df |>
+  ggplot(
+    aes(
+      x = proj1,
+      y = proj2)) +
+  geom_point(
+    size = 0.8,
+    alpha = 0.5,
+    color = "#000000") +
+  geom_segment(
+    data=axes,
+    aes(x=x1, y=y1, xend=x2, yend=y2),
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2, label=rownames(axes)),
+    colour="grey50",
+    size = 5) +
+  geom_path(
+    data=circle,
+    aes(x=c1, y=c2), colour="grey70")

### 2024-08-21 paper.qmd
Updated Figure 2

@@
 ```{r}
+#| label: scale-highd-data
+# Center the data by subtracting the mean of each column
+center_data <- function(data) {
+  apply(data, 2, function(col) col - mean(col))
+}
+
+# Function to scale data manually
+scale_data_manual <- function(data) {
+  # Step 1: Center the data (mean 0)
+  data_centered <- center_data(data)
+
+  # Step 2: Calculate the range of each dimension
+  ranges <- apply(data_centered, 2, function(x) max(x) - min(x))
+
+  # Step 3: Find the dimension with the largest range
+  max_range <- max(ranges)
+  scaled_data <- data_centered
+
+  # Step 4: Scale each dimension
+  for (i in seq_along(ranges)) {
+    scaling_factor <- ranges[i] / max_range
+    scaled_data[, i] <- data_centered[, i] / ranges[i] * scaling_factor
+  }
+
+  # Step 5: Rescale the dimension with the largest range to [-1, 1]
+  largest_dim_index <- which.max(ranges)
+  scaled_data[, largest_dim_index] <- scaled_data[, largest_dim_index] * max_range
+
+  return(scaled_data)
+}
+```
+
+```{r}
 #| label: code-setup
@@
 
+# Apply the scaling function to the s_curve_noise data
+scaled_s_curve_noise <- scale_data_manual(
+  training_data_scurve |> select(-ID)) |>
+  as_tibble()
+
+# Apply the scaling function to the s_curve_noise data
+scaled_s_curve_true <- scale_data_manual(
+  true_model_df |> select(-ID)) |>
+  as_tibble()
+
+# projection <- cbind(
+#     c(0.24288,-0.03628,-0.03302,0.12182,-0.04150,0.02344,0.05910),
+#     c(0.05019,0.07165,-0.20131,-0.16200,-0.05343,0.06854,-0.00551))
+
 projection <- cbind(
-    c(0.24288,-0.03628,-0.03302,0.12182,-0.04150,0.02344,0.05910),
-    c(0.05019,0.07165,-0.20131,-0.16200,-0.05343,0.06854,-0.00551))
-
-projected <- as.matrix(training_data_scurve |> select(-ID)) %*% projection
+  c(0.5898,-0.4620,-0.1362,0.2047,-0.3680,0.1032,-0.4818),
+  c(0.1756,0.0257,0.3683,0.2400,-0.5518,-0.4573,0.5115))
+
+projected <- as.matrix(scaled_s_curve_noise) %*% projection
 
@@
 
-projected_true <- as.matrix(true_model_df |> select(-ID)) %*% projection
+projected_true <- as.matrix(scaled_s_curve_true) %*% projection
 
@@
 projected <- data.frame(projected)
-obs_labels <- as.character(1:nrow(training_data_scurve))
+obs_labels <- as.character(1:nrow(scaled_s_curve_noise))
 
@@
 
-axis_labels <- colnames(training_data_scurve |> select(-ID))
+axis_labels <- colnames(scaled_s_curve_noise)
 rownames(axes) <- axis_labels

### 2024-08-21 paper.qmd
Updated with another projection but need to cheack in the second one why the circle's aspect ratio is different

@@
 
-# Apply the scaling function to the s_curve_noise data
+# Scaling the true model and add the mean of scaled noise dimensions
 scaled_s_curve_true <- scale_data_manual(
-  true_model_df |> select(-ID)) |>
-  as_tibble()
-
-# projection <- cbind(
-#     c(0.24288,-0.03628,-0.03302,0.12182,-0.04150,0.02344,0.05910),
-#     c(0.05019,0.07165,-0.20131,-0.16200,-0.05343,0.06854,-0.00551))
+  true_model_df |> 
+    select(-ID) |>
+    select(x1:x3)) |>
+  as_tibble() |>
+  mutate(x4 = mean(scaled_s_curve_noise$x4),
+         x5 = mean(scaled_s_curve_noise$x5),
+         x6 = mean(scaled_s_curve_noise$x6),
+         x7 = mean(scaled_s_curve_noise$x7))
 
 projection <- cbind(
-  c(0.5898,-0.4620,-0.1362,0.2047,-0.3680,0.1032,-0.4818),
-  c(0.1756,0.0257,0.3683,0.2400,-0.5518,-0.4573,0.5115))
+    c(-0.2782,0.2156,0.1410,0.8068,0.0804,-0.0315,0.4449),
+    c(-0.6090,0.3014,-0.1654,-0.3381,-0.0964,0.5900,0.1979))
 
@@
 projection <- cbind(
-    c(0.03685,-0.02239,0.08047,0.22415,-0.10295,0.04529,0.10381),
-    c(-0.16908,0.11691,0.04886,0.10189,-0.00203,-0.09763,-0.13207))
-
-projected <- as.matrix(training_data_scurve |> select(-ID)) %*% projection
+    c(-0.6296,-0.1452,0.3314,0.2259,-0.4488,-0.0053,-0.4693),
+    c(0.5314,0.4634,0.1431,0.2129,-0.0107,0.1477,-0.6443))
+
+projected <- as.matrix(scaled_s_curve_noise) %*% projection
 
@@
 projected <- data.frame(projected)
-obs_labels <- as.character(1:nrow(training_data_scurve))
+obs_labels <- as.character(1:nrow(scaled_s_curve_noise))
 
 # axis_scale <- 2 * limits/3
-# axis_pos <- -0.4 * limits
-
+# axis_pos <- -0.6 * limits
 axis_scale <- limits/6
@@
 
-axis_labels <- colnames(training_data_scurve |> select(-ID))
+axis_labels <- colnames(scaled_s_curve_noise)
 rownames(axes) <- axis_labels
 
-# axis_scale <- 0.1 * limits
-# axis_pos <- -0.4 * limits
+# axis_scale <- 0.15 * limits
+# axis_pos <- -0.6 * limits
 # adj <- function(x) axis_pos + x * axis_scale

### 2024-08-21 scripts/scurve/validate_langevitour_with_scaled_scurve.R

@@ -91,3 +91,17 @@ projected_df |>
   geom_path(
     data=circle,
     aes(x=c1, y=c2), colour="grey70")
+
+# Combine with the true model for visualization
+df <- dplyr::bind_rows(scaled_s_curve_true |> mutate(type = "true model"),
+                       scaled_s_curve_noise |> mutate(type = "data"))
+
+# Visualize with langevitour
+langevitour(df |> dplyr::select(-type),
+            lineFrom = wireframe_true_model$from,
+            lineTo = wireframe_true_model$to,
+            group = df$type,
+            pointSize = append(rep(1.5, NROW(scaled_s_curve_true)), rep(1, NROW(scaled_s_curve_noise))),
+            levelColors = c("#6a3d9a", "#969696"),
+            lineColors = rep("#969696", nrow(wireframe_true_model)))
+

### 2024-08-22 paper.qmd
Removed duplicated figures

@@
 
-umap_pbmc <- umap_pbmc_scaled |>
-  ggplot(aes(x = UMAP1,
-             y = UMAP2)) +
-  geom_point(alpha=0.3, color = "#000000") 
+# umap_pbmc <- umap_pbmc_scaled |>
+#   ggplot(aes(x = UMAP1,
+#              y = UMAP2)) +
+#   geom_point(alpha=0.3, color = "#000000") 
 ```
@@
 
-tsne_pbmc <- tsne_pbmc_scaled |>
-  ggplot(aes(x = tSNE1,
-             y = tSNE2)) +
-  geom_point(alpha=0.3, color = "#000000")
+# tsne_pbmc <- tsne_pbmc_scaled |>
+#   ggplot(aes(x = tSNE1,
+#              y = tSNE2)) +
+#   geom_point(alpha=0.3, color = "#000000")
 ```
@@
 #| out-width: 40%
+#| eval: false
 

### 2024-08-22 paper.tex
Updated Figure 2
Removed duplicated figures

@@
 identify clusters in Human Peripheral Blood Mononuclear Cells (PBMC3k).
-Figure~\ref{fig-umap-author} is a reproduction of the published plot. It
+\textbf{?@fig-umap-author} is a reproduction of the published plot. It
 shows three well-separated clusters with big separations.
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=0.3\textheight]{paper_files/figure-pdf/fig-umap-author-1.pdf}
-
-}
-
-\caption{\label{fig-umap-author}Reproduction of plot published in Chen
-et al.~(2023) showing a \(2\text{-}D\) layout from UMAP applied for the
-PBMC3k dataset. The (hyper-)parameter settings, beyond the defaults, are
-30 nearest neighbors and minimum distance equal to 0.3. We use our
-model-in-the-data-space to assess whether this is an accurate
-representation of structure present in the high-dimensional data, or if
-it is misleading.}
-
-\end{figure}%
 
@@
 structure, we visualize the model constructed with UMAP overlaid on the
-\(9\text{-}D\) data. The Figure~\ref{fig-umap-author} shows three
+\(9\text{-}D\) data. The \textbf{?@fig-umap-author} shows three
 well-separated clusters. However, as shown in Figure~\ref{fig-pbmc1-sc},
@@
 Therefore, the suggested UMAP representation
-(Figure~\ref{fig-umap-author}) does not accurately represent the
+(\textbf{?@fig-umap-author}) does not accurately represent the
 structure(s) present in PBMC3k dataset. But, when visualizing the
@@
 (Figure~\ref{fig-pbmc1-sc}).
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-model-pbmc-author-1.pdf}
-
-}
-
-\caption{\label{fig-model-pbmc-author}(a) Model generated in
-\(2\text{-}D\) with UMAP, and (b) \(p\text{-}D\) model error in
-\(2\text{-}D\). The \(2\text{-}D\) model shows three well-separated
-distant clusters. The \(p\text{-}D\) model errors are distributed along
-clusters.}
-
-\end{figure}%
 
@@
 
-As shown in Figure~\ref{fig-tsne-suggest}, there are three
-well-separated clusters, although the separation between the clusters
-are small. Additionally, non-linear structures can also be observed
-within the clusters (Figure~\ref{fig-tsne-suggest}). This demonstrates
-that tSNE accurately captures the data structure of the PBMC3k dataset,
-which UMAP did not achieve.
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=0.4\textwidth,height=\textheight]{paper_files/figure-pdf/fig-tsne-suggest-1.pdf}
-
-}
-
-\caption{\label{fig-tsne-suggest}The \(2\text{-}D\) layout from tSNE
-applied for the PBMC3k dataset. The default (hyper-)parameter setting is
-a perplexity \(30\). We use our model-in-the-data-space to assess
-whether this is an accurate representation of structure present in the
-high-dimensional data, or if it is misleading.}
-
-\end{figure}%
+As shown in \textbf{?@fig-tsne-suggest}, there are three well-separated
+clusters, although the separation between the clusters are small.
+Additionally, non-linear structures can also be observed within the
+clusters (\textbf{?@fig-tsne-suggest}). This demonstrates that tSNE
+accurately captures the data structure of the PBMC3k dataset, which UMAP
+did not achieve.
 
@@
 \begin{center}
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-60-1.pdf}
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-61-1.pdf}
 \end{center}

### 2024-08-22 paper.qmd
Updated Figure 9

@@
 
-diaconis_bin_width <- calc_bins_y(
-    bin1 = ceiling(NROW(training_data_scurve)^(1/3)),
-    r2 = r2,
-    q = 0.1)$a1
+# diaconis_bin_width <- calc_bins_y(
+#     bin1 = ceiling(NROW(training_data_scurve)^(1/3)),
+#     r2 = r2,
+#     q = 0.1)$a1
 
@@
   geom_line(linewidth = 0.05) + 
-  # geom_vline(xintercept = 11, linetype="solid",
-  #            color = "#bdbdbd", linewidth=1) +
-  geom_vline(xintercept = diaconis_bin_width, 
-             linetype=2, color = "#bdbdbd", 
+  geom_vline(xintercept = 0.096, linetype=2,
+             color = "#bdbdbd", linewidth=1) +
+  geom_vline(xintercept = 0.116,
+             linetype="solid",
+             color = "#bdbdbd",
              linewidth=1) +
-  # geom_vline(xintercept = 13, linetype=2,
-  #            color = "#bdbdbd", linewidth=1) +
-  #scale_x_continuous(breaks = 5:19) +
+  geom_vline(xintercept = 0.184, linetype=2,
+             color = "#bdbdbd", linewidth=1) +
+  scale_x_continuous(breaks = sort(unique(round(error_scurve$a1, 3)))) +
   labs(x = expression(a[1]), y = "MSE") +
@@
   theme(aspect.ratio = 0.75,
-        axis.text.x = element_text(size = 12),
+        axis.text.x = element_text(size = 10,
+                                   angle = 90),
         axis.text.y = element_text(size = 12),
@@
 
-# b1_m_scurve <- ggplot(error_scurve, 
-#                      aes(x = bin1, 
-#                          y = b_non_empty/b)) + 
-#   geom_point(size = 1) +
-#   geom_line(linewidth = 0.05) + 
-#   # geom_vline(xintercept = 11, linetype="solid",
-#   #            color = "#bdbdbd", linewidth=1) +
-#   # geom_vline(xintercept = 9, linetype=2,
-#   #            color = "#bdbdbd", linewidth=1) +
-#   # geom_vline(xintercept = 13, linetype=2,
-#   #            color = "#bdbdbd", linewidth=1) +
-#   scale_x_continuous(breaks = 5:19) +
-#   labs(x = expression(b[1]), y = expression(frac(m, b))) +
-#   ggtitle("(a)") +
-#   theme(aspect.ratio = 0.75,
-#         axis.text.x = element_text(size = 12),
-#         axis.text.y = element_text(size = 12),
-#         axis.title.x = element_text(size = 12),
-#         axis.title.y = element_text(size = 12, 
-#             angle=90),
-#         axis.ticks.x = element_line(linewidth = 0.5),
-#         axis.ticks.y = element_line(linewidth = 0.5),
-#         plot.title = element_text(size = 12),
-#         legend.title = element_text(size = 12),
-#         legend.text = element_text(size = 10),
-#         legend.key.height = unit(1, 'cm'),
-#         legend.key.width = unit(1, 'cm'),
-#         legend.key.size = unit(1, 'cm'))
+a1_m_scurve <- ggplot(error_scurve,
+                     aes(x = a1,
+                         y = b_non_empty)) +
+  geom_point(size = 1) +
+  geom_line(linewidth = 0.05) +
+  geom_vline(xintercept = 0.116,
+           linetype=2,
+           color = "#bdbdbd",
+           linewidth=1) +
+  geom_hline(yintercept = 52,
+           linetype=2,
+           color = "#bdbdbd",
+           linewidth=1) +
+  # geom_vline(xintercept = 11, linetype="solid",
+  #            color = "#bdbdbd", linewidth=1) +
+  # geom_vline(xintercept = 9, linetype=2,
+  #            color = "#bdbdbd", linewidth=1) +
+  # geom_vline(xintercept = 13, linetype=2,
+  #            color = "#bdbdbd", linewidth=1) +
+  scale_y_continuous(breaks = sort(unique(error_scurve$b_non_empty))) +
+  scale_x_continuous(breaks = sort(unique(round(error_scurve$a1, 3)))) +
+  labs(x = expression(a[1]), y = expression(m)) +
+  ggtitle("(b)") +
+  theme(aspect.ratio = 0.75,
+        axis.text.x = element_text(size = 10,
+                                   angle = 90),
+        axis.text.y = element_text(size = 11),
+        axis.title.x = element_text(size = 12),
+        axis.title.y = element_text(size = 12,
+            angle=90),
+        axis.ticks.x = element_line(linewidth = 0.5),
+        axis.ticks.y = element_line(linewidth = 0.5),
+        plot.title = element_text(size = 12),
+        legend.title = element_text(size = 12),
+        legend.text = element_text(size = 10),
+        legend.key.height = unit(1, 'cm'),
+        legend.key.width = unit(1, 'cm'),
+        legend.key.size = unit(1, 'cm'))
 # 
@@
 
-mse_scurve_lwd2 <- error_rm_scurve |>
-  filter(b_non_empty >= 5) |>
-  ggplot(aes(x = b_non_empty,y = MSE)) + 
-  geom_point(
-    size = 1
-    ) +
-  geom_line(
-    linewidth = 0.3
-    ) + 
-  # geom_vline(xintercept = benchmark1, linetype="solid", 
-  #            color = "#bdbdbd", linewidth=1, alpha = 0.5) +
-  scale_x_continuous("m (b1 = 14)") + 
-  ylab("MSE") +
-  ggtitle("(e)") +
-  theme(aspect.ratio = 0.75,
-        axis.text.x = element_text(size = 12),
-        axis.text.y = element_text(size = 12),
-        axis.title.x = element_text(size = 12),
-        axis.title.y = element_text(size = 12, 
-            angle=90),
-        plot.title = element_text(size = 12))
+# mse_scurve_lwd2 <- error_rm_scurve |>
+#   filter(b_non_empty >= 5) |>
+#   ggplot(aes(x = b_non_empty,y = MSE)) + 
+#   geom_point(
+#     size = 1
+#     ) +
+#   geom_line(
+#     linewidth = 0.3
+#     ) + 
+#   # geom_vline(xintercept = benchmark1, linetype="solid", 
+#   #            color = "#bdbdbd", linewidth=1, alpha = 0.5) +
+#   scale_x_continuous("m (b1 = 14)") + 
+#   ylab("MSE") +
+#   ggtitle("(e)") +
+#   theme(aspect.ratio = 0.75,
+#         axis.text.x = element_text(size = 12),
+#         axis.text.y = element_text(size = 12),
+#         axis.title.x = element_text(size = 12),
+#         axis.title.y = element_text(size = 12, 
+#             angle=90),
+#         plot.title = element_text(size = 12))
 ```
@@
 
-mse_scurve_b + mse_scurve_lwd + 
-  mse_scurve_lwd2 + distance_hist + 
+mse_scurve_b + a1_m_scurve +
+  mse_scurve_lwd + distance_hist + 
   plot_layout(ncol=2)

### 2024-08-22 paper.qmd
Updated Figure 8

@@
   data = umap_scurve_scaled, 
-  bin1 = 8, 
+  bin1 = 7, 
   r2 = r2)
@@
 a1_1 <- calc_bins_y(
-  bin1 = 8, 
+  bin1 = 7, 
   r2 = r2
@@
   filter(drop_empty == FALSE) |>
-  mutate(b1 = "b1 = 8")
+  mutate(b1 = "b1 = 7")
 
@@
   data = umap_scurve_scaled, 
-  bin1 = 10, 
+  bin1 = 11, 
   r2 = r2)
@@
 a1_2 <- calc_bins_y(
-  bin1 = 10, 
+  bin1 = 11, 
   r2 = r2
@@
   filter(drop_empty == FALSE) |>
-  mutate(b1 = "b1 = 10")
+  mutate(b1 = "b1 = 11")
 
@@
   data = umap_scurve_scaled, 
-  bin1 = 14, 
+  bin1 = 13, 
   r2 = r2)
@@
 a1_3 <- calc_bins_y(
-  bin1 = 14, 
+  bin1 = 13, 
   r2 = r2
@@
 #| fig-pos: H
-#| fig-cap: "Hexbin density plots of UMAP layout of the S-curve data, using three different bin inputs: (a) $b = 91 \\text{ } (7, \\text{ }13)  \\text{ }(q = 0.12)$, (b) $b = 220 \\text{ } (11, \\text{ }20) \\text{ }(q = 0.07)$, and (c) $b = 527 \\text{ } (17, \\text{ }31) \\text{ }(q = 0.05)$.
+#| fig-cap: "Hexbin density plots of UMAP layout of the S-curve data, using three different bin inputs: (a) $b = 91 \\text{ } (7, \\text{ }13)$, (b) $b = 220 \\text{ } (11, \\text{ }20)$, and (c) $b = 312 \\text{ } (13, \\text{ }24)$.
 Color indicates standardized counts, dark indicating high count and light indicates low count.

### 2024-08-22 paper.qmd
MSE Vs Counts

@@
            b_non_empty = NROW(df_bin_centroids_scurve),
-           a1 = a1)
+           a1 = a1,
+           mean_counts = sum(df_bin_centroids_scurve$bin_counts)/NROW(df_bin_centroids_scurve))
 
@@
         legend.key.size = unit(1, 'cm'))
+
+count_mse_scurve <- ggplot(error_scurve,
+                     aes(x = mean_counts,
+                         y = MSE)) +
+  geom_point(size = 1) +
+  geom_line(linewidth = 0.05) +
+  # geom_vline(xintercept = 0.116,
+  #          linetype=2,
+  #          color = "#bdbdbd",
+  #          linewidth=1) +
+  # geom_hline(yintercept = 52,
+  #          linetype=2,
+  #          color = "#bdbdbd",
+  #          linewidth=1) +
+  # geom_vline(xintercept = 11, linetype="solid",
+  #            color = "#bdbdbd", linewidth=1) +
+  # geom_vline(xintercept = 9, linetype=2,
+  #            color = "#bdbdbd", linewidth=1) +
+  # geom_vline(xintercept = 13, linetype=2,
+  #            color = "#bdbdbd", linewidth=1) +
+  labs(x = expression(mean(n[h])), y = "MSE") +
+  ggtitle("(c)") +
+  theme(aspect.ratio = 0.75,
+        axis.text.x = element_text(size = 10,
+                                   angle = 90),
+        axis.text.y = element_text(size = 11),
+        axis.title.x = element_text(size = 12),
+        axis.title.y = element_text(size = 12,
+            angle=90),
+        axis.ticks.x = element_line(linewidth = 0.5),
+        axis.ticks.y = element_line(linewidth = 0.5),
+        plot.title = element_text(size = 12),
+        legend.title = element_text(size = 12),
+        legend.text = element_text(size = 10),
+        legend.key.height = unit(1, 'cm'),
+        legend.key.width = unit(1, 'cm'),
+        legend.key.size = unit(1, 'cm'))
+
 # 
@@
 <!--distribution of distance along with the default benchmark-->
-
-```{r}
-#| label: diff-bins-scurve-buff3
-#| eval: false
-## To initialize benchmark values to remove low density hexagons
-benchmark_rm_hex_vec <- seq(0, 0.7, by=0.1)
-bin1_vec_scurve <- c(7, 11, 17)
-distance_all_scurve <- data.frame(matrix(nrow = 0, ncol = 0))
-#buffer_vec <- 0.07
-
-for (xbins in bin1_vec_scurve) {
-  for (buff in buffer_vec) {
-
-  #for (benchmark_rm_lwd in benchmark_rm_hex_vec) {
-
-    a1 <- calc_bins_y(bin1 = xbins, r2 = r2, q = buff)$a1
-
-    scurve_model <- fit_highd_model(
-      training_data = training_data_scurve,
-      emb_df = umap_scurve_scaled,
-      bin1 = xbins,
-      r2 = r2,
-      q = buff, 
-      is_bin_centroid = TRUE,
-      is_rm_lwd_hex = FALSE,
-      col_start_highd = "x"
-    )
-
-    df_bin_centroids_scurve <- scurve_model$df_bin_centroids
-    df_bin_scurve <- scurve_model$df_bin
-
-    df_bin_centroids_scurve_high_dens <- df_bin_centroids_scurve |>
-      filter(std_counts > benchmark_rm_lwd)
-
-    df_bin_scurve_high_dens <- df_bin_scurve |>
-      filter(hb_id %in% df_bin_centroids_scurve_high_dens$hexID)
-
-    tr1_object_scurve <- tri_bin_centroids(
-      df_bin_centroids_scurve_high_dens, 
-      x = "c_x", 
-      y = "c_y")
-
-    tr_from_to_df_scurve <- gen_edges(
-      tri_object = tr1_object_scurve)
-
-    ## Compute 2D distances
-    distance_scurve <- cal_2d_dist(
-      tr_coord_df = tr_from_to_df_scurve, 
-      start_x = "x_from", 
-      start_y = "y_from", 
-      end_x = "x_to", 
-      end_y = "y_to", 
-      select_vars = c("from", "to", "distance"))
-
-    start_point <- a1/2
-
-    # Create bins and calculate the mean value for each bin range
-    bins <- seq(start_point, max(distance_scurve$distance) + a1, 
-                by = a1)
-    bin_labels <- bins[-length(bins)] + a1 / 2  # mean of each bin range
-
-    distance_scurve <- distance_scurve |> 
-      mutate(bin = cut(distance, breaks = bins, include.lowest = TRUE, labels = bin_labels))
-
-    ## Compute error
-    distance_scurve <- distance_scurve |>
-      mutate(q = buff,
-             bin1 = xbins,
-             a1 = a1)
-
-    distance_all_scurve <- bind_rows(distance_all_scurve, distance_scurve)
-
-  }
-
-}
-
-# dist_summary_scurve <- distance_all_scurve |>
-#     count(bin1, q, bin)
-
-distance_hist <- distance_all_scurve |>
-  filter(bin1 == 11) |>
-  filter(q == 0.07) |>
-  ggplot(aes(x = distance)) +
-    geom_histogram(
-      aes(y = after_stat(count) / sum(after_stat(count)), 
-      group = as.factor(q),
-      fill = as.factor(q)),
-      binwidth = a1,
-      position = "dodge") +
-  scale_x_continuous(breaks = bin_labels, labels = round(bin_labels, 3)) +
-  scale_color_brewer(type = "qual") +
-  #facet_wrap(vars(as.factor(bin1)), scale = "free") +
-  ylab("standardized number of edges") +
-  xlab(expression(d^{(2)})) +
-  ggtitle("(d)") +
-  theme(aspect.ratio = 0.75,
-        axis.title.x = element_text(size = 12),
-        axis.title.y = element_text(size = 12, angle = 90),
-        axis.text.x = element_text(size = 12, angle = 90),
-        axis.text.y = element_text(size = 12),
-        axis.ticks.x = element_line(), 
-        plot.title = element_text(size = 12))
-
-```
-
 ```{r}
@@
         plot.title = element_text(size = 12))
+
+distance_scurve_summary <- distance_scurve |>
+  group_by(bin) |>
+  summarise(count = n()) |>
+  mutate(std_counts = count/sum(count))
+
 ```
@@
 mse_scurve_b + a1_m_scurve +
-  mse_scurve_lwd + distance_hist + 
+  count_mse_scurve + distance_hist + 
   plot_layout(ncol=2)

### 2024-08-22 paper.tex
Updated Figure 8
MSE Vs Counts

@@
 the S-curve data, using three different bin inputs: (a)
-\(b = 91 \text{ } (7, \text{ }13)  \text{ }(q = 0.12)\), (b)
-\(b = 220 \text{ } (11, \text{ }20) \text{ }(q = 0.07)\), and (c)
-\(b = 527 \text{ } (17, \text{ }31) \text{ }(q = 0.05)\). Color
-indicates standardized counts, dark indicating high count and light
-indicates low count. At the smallest bin size the data segregates into
-two separate groups, suggesting this is too many bins. Using the MSE of
-the model fit in \(p\text{-}D\) helps decide on a useful choice of
-number of bins.}
+\(b = 91 \text{ } (7, \text{ }13)\), (b)
+\(b = 220 \text{ } (11, \text{ }20)\), and (c)
+\(b = 312 \text{ } (13, \text{ }24)\). Color indicates standardized
+counts, dark indicating high count and light indicates low count. At the
+smallest bin size the data segregates into two separate groups,
+suggesting this is too many bins. Using the MSE of the model fit in
+\(p\text{-}D\) helps decide on a useful choice of number of bins.}
 

### 2024-08-22 paper.qmd
Added plot for MSE Vs bin counts for different lwd thresholds

@@
            b_non_empty = NROW(df_bin_centroids_scurve),
-           a1 = a1,
-           mean_counts = sum(df_bin_centroids_scurve$bin_counts)/NROW(df_bin_centroids_scurve))
+           a1 = a1)
 
@@
 
-count_mse_scurve <- ggplot(error_scurve,
-                     aes(x = mean_counts,
-                         y = MSE)) +
-  geom_point(size = 1) +
-  geom_line(linewidth = 0.05) +
-  # geom_vline(xintercept = 0.116,
-  #          linetype=2,
-  #          color = "#bdbdbd",
-  #          linewidth=1) +
-  # geom_hline(yintercept = 52,
-  #          linetype=2,
-  #          color = "#bdbdbd",
-  #          linewidth=1) +
-  # geom_vline(xintercept = 11, linetype="solid",
-  #            color = "#bdbdbd", linewidth=1) +
-  # geom_vline(xintercept = 9, linetype=2,
-  #            color = "#bdbdbd", linewidth=1) +
-  # geom_vline(xintercept = 13, linetype=2,
-  #            color = "#bdbdbd", linewidth=1) +
-  labs(x = expression(mean(n[h])), y = "MSE") +
-  ggtitle("(c)") +
-  theme(aspect.ratio = 0.75,
-        axis.text.x = element_text(size = 10,
-                                   angle = 90),
-        axis.text.y = element_text(size = 11),
-        axis.title.x = element_text(size = 12),
-        axis.title.y = element_text(size = 12,
-            angle=90),
-        axis.ticks.x = element_line(linewidth = 0.5),
-        axis.ticks.y = element_line(linewidth = 0.5),
-        plot.title = element_text(size = 12),
-        legend.title = element_text(size = 12),
-        legend.text = element_text(size = 10),
-        legend.key.height = unit(1, 'cm'),
-        legend.key.width = unit(1, 'cm'),
-        legend.key.size = unit(1, 'cm'))
-
 # 
@@
 ## To initialize effective bins along x
-effective_bin1_scurve <- 14
+effective_bin1_scurve <- 11
 
@@
 #| label: bin-counts-scurve
+#| eval: false
 
@@
            b = bin1 * bin2,
-           b_non_empty = NROW(df_bin_centroids_scurve_high_dens))
+           b_non_empty = NROW(df_bin_centroids_scurve_high_dens),
+           mean_counts = sum(df_bin_centroids_scurve_high_dens$bin_counts)/NROW(df_bin_centroids_scurve_high_dens))
 
@@
 
+error_rm_scurve <- error_rm_scurve |>
+  mutate(mean_counts = round(mean_counts, 0)) |>
+  filter(b_non_empty >= 5)
+
 mse_scurve_lwd <- ggplot(error_rm_scurve, 
-                     aes(x = benchmark_rm_lwd, 
+                     aes(x = mean_counts, 
                          y = MSE)) + 
@@
   #            color = "#bdbdbd", linewidth=1, alpha = 0.5) +
-  scale_x_continuous("standardized bin count", 
-         transform = "reverse") + 
+  # scale_x_continuous("standardized bin count", 
+  #        transform = "reverse") + 
+  scale_x_continuous(breaks = unique(error_rm_scurve$mean_counts)) +
+  xlab(expression(mean(n[h]))) +
   ylab("MSE") +
@@
             angle=90),
-        plot.title = element_text(size = 12))
+        plot.title = element_text(size = 12),
+        axis.ticks.x = element_line(linewidth = 0.5),
+        axis.ticks.y = element_line(linewidth = 0.5))
 
@@
 mse_scurve_b + a1_m_scurve +
-  count_mse_scurve + distance_hist + 
+  mse_scurve_lwd + distance_hist + 
   plot_layout(ncol=2)

### 2024-08-22 paper.qmd
Fixed Figure 9

@@
              color = "#bdbdbd", linewidth=1) +
-  scale_x_continuous(breaks = sort(unique(round(error_scurve$a1, 3)))) +
+  scale_x_continuous(breaks = sort(unique(round(error_scurve$a1, 2)))[c(1, 3, 5, 6, 8, 10:12)]) +
   labs(x = expression(a[1]), y = "MSE") +
@@
   theme(aspect.ratio = 0.75,
-        axis.text.x = element_text(size = 10,
-                                   angle = 90),
+        axis.text.x = element_text(size = 12),
         axis.text.y = element_text(size = 12),
@@
   #            color = "#bdbdbd", linewidth=1) +
-  scale_y_continuous(breaks = sort(unique(error_scurve$b_non_empty))) +
-  scale_x_continuous(breaks = sort(unique(round(error_scurve$a1, 3)))) +
+  scale_y_continuous(breaks = sort(unique(error_scurve$b_non_empty))[-3]) +
+  scale_x_continuous(breaks = sort(unique(round(error_scurve$a1, 2)))[c(1, 3, 5, 6, 8, 10:12)]) +
   labs(x = expression(a[1]), y = expression(m)) +
@@
   theme(aspect.ratio = 0.75,
-        axis.text.x = element_text(size = 10,
-                                   angle = 90),
+        axis.text.x = element_text(size = 12),
         axis.text.y = element_text(size = 11),
@@
       binwidth = bin_width) +
-  scale_x_continuous(breaks = bin_labels, labels = round(bin_labels, 3)) +
+  scale_x_continuous(breaks = bin_labels, labels = round(bin_labels, 2)) +
   ylab("standardized number of edges") +
@@
         axis.title.y = element_text(size = 12, angle = 90),
-        axis.text.x = element_text(size = 12, angle = 90),
+        axis.text.x = element_text(size = 12),
         axis.text.y = element_text(size = 12),

### 2024-08-22 paper.qmd
Need to fix later

@@
 #| echo: false
-#| fig-cap: "Model in $2\\text{-}D$, on the tSNE layout, and three views of the fit in projections from $9\text{-}D$, for the PBMC3k data ($(s_1, \ s_2) = (-0.050, \ -0.058)$, $b = 300 \  (15, \ 20)$, $m = 136$, benchmark value to remove large edges is $0.133$).
+#| fig-cap: "Model in $2\\text{-}D$, on the tSNE layout, and three views of the fit in projections from $9\\text{-}D$, for the PBMC3k data ($(s_1, \ s_2) = (-0.050, \ -0.058)$, $b = 300 \  (15, \ 20)$, $m = 136$, benchmark value to remove large edges is $0.133$).
 (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/5Y1hE4i7N2k>)."

### 2024-08-22 paper.tex
Need to fix later

@@
 \caption{\label{fig-pbmc2-sc}Model in \(2\text{-}D\), on the tSNE
-layout, and three views of the fit in projections from \(9  ext{-}D\),
+layout, and three views of the fit in projections from \(9\text{-}D\),
 for the PBMC3k data (\((s_1,  s_2) = (-0.050,  -0.058)\),

### 2024-08-22 scripts/five_gau_clusters/3_gen_model_in_data_space_with_tsne.R

@@ -14,7 +14,7 @@ gau1_scaled_obj <- gen_scaled_data(
 tsne_gau_scaled <- gau1_scaled_obj$scaled_nldr
 
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 14
+num_bins_x_gau1 <- 13
 lim1 <- gau1_scaled_obj$lim1
 lim2 <- gau1_scaled_obj$lim2
 r2_gau1 <- diff(lim2)/diff(lim1)
@@ -27,7 +27,7 @@ gau1_model <- fit_highd_model(
   is_bin_centroid = TRUE,
   is_rm_lwd_hex = FALSE,
   col_start_highd = "x",
-  q = 0.16
+  q = 0.1
 )
 
 df_bin_centroids_gau1 <- gau1_model$df_bin_centroids
@@ -43,9 +43,9 @@ tr_from_to_df_gau1 <- gen_edges(
 #   filter(row_number() != 76) |>
 #   filter(row_number() != 34)
 
-tr_from_to_df_gau1 <- tr_from_to_df_gau1 |>
-  filter(row_number() != 115) |>
-  filter(row_number() != 139)
+# tr_from_to_df_gau1 <- tr_from_to_df_gau1 |>
+#   filter(row_number() != 115) |>
+#   filter(row_number() != 139)
 
 ## Compute 2D distances
 distance_gau1 <- cal_2d_dist(
@@ -62,7 +62,6 @@ benchmark_gau1 <- find_lg_benchmark(
   distance_col = "distance")
 
 
-
 trimesh_removed_gau1 <- vis_rmlg_mesh(
   distance_edges = distance_gau1,
   benchmark_value = benchmark_gau1,
@@ -75,7 +74,7 @@ hb_obj_gau1 <- hex_binning(
   data = tsne_gau_scaled,
   bin1 = num_bins_x_gau1,
   r2 = r2_gau1,
-  q = 0.16)
+  q = 0.1)
 
 tsne_data_with_hb_id <- hb_obj_gau1$data_hb_id
 

### 2024-08-22 paper.qmd
Added umap projection for Gaussian clusters..need to scale data first

@@
 
+## First projection
 projection <- cbind(
-  c(-0.00215,-0.68905,-0.04778,-0.54223),
-  c(0.42558,-0.23854,-0.63659,0.35753))
+    c(0.18827,-0.28588,-0.20848,-0.78132),
+    c(0.71366,0.00788,-0.42630,0.28283))
 
@@
 
-five_gau_proj_umap_model <- projected_df |>
+five_gau_proj_umap_model1 <- projected_df |>
   filter(type == "data") |>
@@
   xlim(c(-0.18, 0.1)) +
-  ylim(c(-0.78, -0.55)) +
+  ylim(c(-0.12, 0.12)) +
+  interior_annotation("b", 
+                      position = c(0.08, 0.9),
+                      cex = 3)
+
+## Second projection
+projection <- cbind(
+  c(0.58453,-0.38194,0.05395,0.52974),
+  c(0.29917,0.29629,-0.76964,-0.03810))
+
+points_only_df <- df_exe |>
+    dplyr::select(-type)
+
+projected <- as.matrix(points_only_df) %*% projection
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+proj_model <- projected_df |>
+  dplyr::filter(type == "model")
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges |> select(-distance), 
+  proj_model, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(proj_model)[-NCOL(proj_model)], "_from")
+
+model_df <- dplyr::left_join(model_df, proj_model, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(proj_model)):NCOL(model_df)] <- paste0(names(proj_model)[-NCOL(proj_model)], "_to")
+
+five_gau_proj_umap_model2 <- projected_df |>
+  filter(type == "data") |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+      geom_point(
+        alpha = 0.3) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#33a02c") +
+  scale_color_manual(values = c("#000000", "#33a02c")) +
+  xlim(c(-0.16, 0.1)) +
+  ylim(c(-0.1, 0.1)) +
   interior_annotation("b", 
@@
 #| label: fig-five-gau-projs
-#| fig-height: 6
-#| fig-width: 18
 #| fig-cap: "(a) tSNE, (b) UMAP, and (c) PaCMAP"
 
-five_gau_proj_tsne_model + five_gau_proj_umap_model + five_gau_proj_pacmap_model +
-  plot_layout(guides = "collect", ncol = 3) &
+# fig-height: 6 
+# fig-width: 18
+
+five_gau_proj_tsne_model + five_gau_proj_tsne_model +
+  five_gau_proj_umap_model1 + five_gau_proj_umap_model2 +
+  five_gau_proj_pacmap_model + five_gau_proj_pacmap_model +
+  plot_layout(guides = "collect", ncol = 2) &
   theme(legend.position='none')

### 2024-08-22 paper.qmd
Updated the scaling function

@@
 
-  # Step 2: Calculate the range of each dimension
-  ranges <- apply(data_centered, 2, function(x) max(x) - min(x))
-
-  # Step 3: Find the dimension with the largest range
-  max_range <- max(ranges)
-  scaled_data <- data_centered
-
-  # Step 4: Scale each dimension
-  for (i in seq_along(ranges)) {
-    scaling_factor <- ranges[i] / max_range
-    scaled_data[, i] <- data_centered[, i] / ranges[i] * scaling_factor
+  # Step 2: Calculate the standard deviation of each dimension
+  sds <- apply(data_centered, 2, sd)
+
+  # Step 3: Scale each dimension to have the range [0, 1]
+  data_scaled <- apply(data_centered, 2, function(col) col / max(abs(col)))
+
+  # Step 4: Scale dimensions according to their variation
+  # The dimension with the highest standard deviation is scaled to [-1, 1]
+  # Other dimensions are scaled to smaller ranges based on their standard deviations
+  max_sd <- max(sds)
+
+  # Normalize the standard deviations to get scaling factors
+  scaling_factors <- sds / max_sd
+
+  for (i in seq_along(scaling_factors)) {
+    data_scaled[, i] <- data_scaled[, i] * scaling_factors[i]
   }
 
-  # Step 5: Rescale the dimension with the largest range to [-1, 1]
-  largest_dim_index <- which.max(ranges)
-  scaled_data[, largest_dim_index] <- scaled_data[, largest_dim_index] * max_range
-
-  return(scaled_data)
+  return(data_scaled)
 }

### 2024-08-23 paper.qmd
Regenerated Figure 2

@@
 projection <- cbind(
-    c(-0.2782,0.2156,0.1410,0.8068,0.0804,-0.0315,0.4449),
-    c(-0.6090,0.3014,-0.1654,-0.3381,-0.0964,0.5900,0.1979))
+    c(0.45181,0.04747,0.02613,0.25735,0.20622,0.08755,-0.11488),
+    c(0.03231,0.23556,0.35166,-0.11741,0.12730,0.09544,0.34262))
 
@@
 
-# axis_scale <- 2 * limits/3
-# axis_pos <- -0.6 * limits
 axis_scale <- limits/6
 axis_pos <- -2/3 * limits
+
+# axis_scale <- limits/4
+# axis_pos <- -0.5 * limits
 
@@
 theta <- seq(0, 2 * pi, length = 50)
-circle <- data.frame(c1 = adj(cos(theta)), c2 = adj(sin(theta)))
+circle <- data.frame(c1 = adj(cos(theta)), 
+                     c2 = adj(sin(theta)))
 
@@
     colour="grey50",
-    size = 5) +
+    size = 3) +
   geom_path(
     data=circle, 
-    aes(x=c1, y=c2), colour="grey70")
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed()
 
@@
 projection <- cbind(
-    c(-0.6296,-0.1452,0.3314,0.2259,-0.4488,-0.0053,-0.4693),
-    c(0.5314,0.4634,0.1431,0.2129,-0.0107,0.1477,-0.6443))
+    c(0.03627,0.15991,0.39768,-0.21032,-0.14886,-0.17589,0.23305),
+    c(-0.34731,-0.08183,0.15819,0.31926,0.08105,0.10083,0.25627))
 
@@
     colour="grey50",
-    size = 5) +
+    size = 3) +
   geom_path(
     data=circle, 
-    aes(x=c1, y=c2), colour="grey70")
-
-```
-
-```{r}
-#| label: s-curve-true-model-proj3
-#| eval: false
-
-projection <- cbind(
-    c(0.11484,0.06997,-0.13638,0.01096,0.01533,0.16257,-0.13615),
-    c(0.07805,0.01652,0.17174,-0.07343,0.16177,0.11112,0.04729))
-
-projected <- as.matrix(training_data_scurve |> select(-ID)) %*% projection
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-projected_true <- as.matrix(true_model_df |> select(-ID)) %*% projection
-
-projected_true_df <- projected_true |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  wireframe_true_model, 
-  projected_true_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_true_df)[-NCOL(projected_true_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_true_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_true_df)):NCOL(model_df)] <- paste0(names(projected_true_df)[-NCOL(projected_true_df)], "_to")
-
-limits <- 1.5
-rng <- range(projected)
-projected <- projected/max(abs(rng))
-colnames(projected) <- c("P1", "P2")
-projected <- data.frame(projected)
-obs_labels <- as.character(1:nrow(training_data_scurve))
-
-axis_scale <- 2 * limits/3
-axis_pos <- -2/3 * limits
-
-adj <- function(x) axis_pos + x * axis_scale
-axes <- data.frame(x1 = adj(0), 
-                   y1 = adj(0), 
-                   x2 = adj(projection[, 1]), 
-                   y2 = adj(projection[, 2]))
-
-axis_labels <- colnames(training_data_scurve |> select(-ID))
-rownames(axes) <- axis_labels
-
-# theta <- seq(0, 2 * pi, length = 50)
-# circle <- data.frame(c1 = adj(cos(theta)), c2 = adj(sin(theta)))
-
-scurve_proj3_true_model <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-      geom_point(
-        size = 0.8,
-        alpha = 0.8,
-        color = "#000000") +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#969696") + 
-  geom_segment(
-    data=axes, 
-    aes(x=x1, y=y1, xend=x2, yend=y2), 
-    colour="grey70") +
-  geom_text(
-    data=axes, 
-    aes(x=x2, y=y2, label=rownames(axes)), 
-    colour="grey50")
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed()
 
@@
 (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/I5GL23vLiw0>)."
+#| fig-pos: H
 #| fig-width: 10
-#| fig-height: 6
-#| fig-pos: H
+#| fig-height: 5
 

### 2024-08-23 paper.qmd
Resized Figure 11

@@
   ylim(c(-0.12, 0.12)) +
-  interior_annotation("b", 
+  interior_annotation("c", 
                       position = c(0.08, 0.9),
@@
   ylim(c(-0.1, 0.1)) +
-  interior_annotation("b", 
+  interior_annotation("d", 
                       position = c(0.08, 0.9),
@@
 #| label: fig-five-gau-projs
+#| fig-height: 12
+#| fig-width: 8
 #| fig-cap: "(a) tSNE, (b) UMAP, and (c) PaCMAP"
-
-# fig-height: 6 
-# fig-width: 18
 

### 2024-08-23 paper.qmd
Updated script to get the projections for UMAP
Updated Figure 11

--- 2024-08-23 paper.qmd
@@
 
+# Apply the scaling
+scaled_gau_data <- scale_data_manual(training_data_gau |> select(-ID)) |>
+  as_tibble()
+
+df_b <- df_b |>
+  select(-type)
+
+# Apply the scaling
+scaled_gau_data_model <- scale_data_manual(df_b) |>
+  as_tibble()
+
 ## First projection
 projection <- cbind(
-    c(0.18827,-0.28588,-0.20848,-0.78132),
-    c(0.71366,0.00788,-0.42630,0.28283))
-
-points_only_df <- df_exe |>
-    dplyr::select(-type)
-
-projected <- as.matrix(points_only_df) %*% projection
+    c(0.36638,0.31467,0.63030,0.24465),
+    c(0.14672,-0.73585,0.33246,-0.12977))
+
+projected <- as.matrix(scaled_gau_data) %*% projection
 
@@
                   "proj2" = "...2")) |>
-  dplyr::mutate(type = df_exe$type) |>
+  #dplyr::mutate(type = df_exe$type) |>
   dplyr::mutate(ID = dplyr::row_number())
 
-proj_model <- projected_df |>
-  dplyr::filter(type == "model")
+projected_model <- as.matrix(scaled_gau_data_model) %*% projection
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
 
@@
   distance_df_small_edges |> select(-distance), 
-  proj_model, 
+  projected_model_df, 
   by = c("from" = "ID"))
 
-names(model_df)[3:NCOL(model_df)] <- paste0(names(proj_model)[-NCOL(proj_model)], "_from")
-
-model_df <- dplyr::left_join(model_df, proj_model, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(proj_model)):NCOL(model_df)] <- paste0(names(proj_model)[-NCOL(proj_model)], "_to")
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
 
 five_gau_proj_umap_model1 <- projected_df |>
-  filter(type == "data") |>
   ggplot(
@@
   scale_color_manual(values = c("#000000", "#33a02c")) +
-  xlim(c(-0.18, 0.1)) +
-  ylim(c(-0.12, 0.12)) +
-  interior_annotation("b", 
+  xlim(c(-0.45, -0.3)) +
+  ylim(c(-0.05, 0.2)) +
+  interior_annotation("c", 
                       position = c(0.08, 0.9),
@@
 projection <- cbind(
-  c(0.58453,-0.38194,0.05395,0.52974),
-  c(0.29917,0.29629,-0.76964,-0.03810))
+    c(-0.70767,-0.09092,0.15109,-0.39811),
+    c(-0.01312,-0.61805,0.44418,0.33305))
 
@@
   scale_color_manual(values = c("#000000", "#33a02c")) +
-  xlim(c(-0.16, 0.1)) +
-  ylim(c(-0.1, 0.1)) +
-  interior_annotation("b", 
+  xlim(c(-0.13, 0.13)) +
+  ylim(c(-0.13, 0.13)) +
+  interior_annotation("d", 
                       position = c(0.08, 0.9),
@@
 #| label: fig-five-gau-projs
+#| fig-height: 12
+#| fig-width: 8
+#| fig-pos: H
 #| fig-cap: "(a) tSNE, (b) UMAP, and (c) PaCMAP"
-
-# fig-height: 6 
-# fig-width: 18
 

### 2024-08-23 paper.qmd
Fixed Figure 2

@@
       y = proj2)) +
-      geom_point(
-        size = 0.8,
-        alpha = 0.5,
-        color = "#000000") +
   geom_segment(
@@
     color = "#3182bd") + 
+  geom_point(
+        size = 0.8,
+        alpha = 0.5,
+        color = "#000000") +
   geom_segment(
@@
     color = "#3182bd") +
+  geom_point(
+        size = 0.8,
+        alpha = 0.5,
+        color = "#000000") +
   geom_segment(

### 2024-08-23 paper.qmd
Updated Figure 11

@@
       y = proj2)) +
-      geom_point(
-        alpha = 0.3) +
   geom_segment(
@@
     color = "#33a02c") +
+  geom_point(
+    alpha = 0.3) +
   scale_color_manual(values = c("#000000", "#33a02c")) +

### 2024-08-23 paper.qmd
Add function to generate axes

@@
   return(data_scaled)
+}
+```
+
+```{r}
+#| label: gen-axes
+
+gen_axes <- function(data, proj_data, limits = 1) {
+
+  rng <- range(proj_data)
+  proj_data <- proj_data/max(abs(rng))
+  colnames(proj_data) <- c("P1", "P2")
+  proj_data <- data.frame(proj_data)
+  obs_labels <- as.character(1:nrow(data))
+
+  axis_scale <- limits/6
+  axis_pos <- -2/3 * limits
+
+  adj <- function(x) axis_pos + x * axis_scale
+  axes <- data.frame(x1 = adj(0), 
+                     y1 = adj(0), 
+                     x2 = adj(projection[, 1]), 
+                     y2 = adj(projection[, 2]))
+
+  axis_labels <- colnames(data)
+  rownames(axes) <- axis_labels
+
+  theta <- seq(0, 2 * pi, length = 50)
+  circle <- data.frame(c1 = adj(cos(theta)), 
+                       c2 = adj(sin(theta)))
+
+  return(list(axes = axes, circle = circle))
+
 }
@@
 
+axes_obj <- gen_axes(
+  data = scaled_gau_data, 
+  proj_data = projected_df, 
+  limits = 1)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
+
 five_gau_proj_umap_model1 <- projected_df |>
@@
   scale_color_manual(values = c("#000000", "#33a02c")) +
+  geom_segment(
+    data=axes, 
+    aes(x=x1, y=y1, xend=x2, yend=y2), 
+    colour="grey70") +
+  geom_text(
+    data=axes, 
+    aes(x=x2, y=y2, label=rownames(axes)), 
+    colour="grey50",
+    size = 3) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
   xlim(c(-0.45, -0.3)) +
@@
 
-points_only_df <- df_exe |>
-    dplyr::select(-type)
-
-projected <- as.matrix(points_only_df) %*% projection
+projected <- as.matrix(scaled_gau_data) %*% projection
 
@@
   distance_df_small_edges |> select(-distance), 
-  proj_model, 
+  projected_model_df, 
   by = c("from" = "ID"))
 
-names(model_df)[3:NCOL(model_df)] <- paste0(names(proj_model)[-NCOL(proj_model)], "_from")
-
-model_df <- dplyr::left_join(model_df, proj_model, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(proj_model)):NCOL(model_df)] <- paste0(names(proj_model)[-NCOL(proj_model)], "_to")
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+axes_obj <- gen_axes(
+  data = scaled_gau_data, 
+  proj_data = projected_df, 
+  limits = 1)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
 
 five_gau_proj_umap_model2 <- projected_df |>
-  filter(type == "data") |>
   ggplot(
@@
   scale_color_manual(values = c("#000000", "#33a02c")) +
+  geom_segment(
+    data=axes, 
+    aes(x=x1, y=y1, xend=x2, yend=y2), 
+    colour="grey70") +
+  geom_text(
+    data=axes, 
+    aes(x=x2, y=y2, label=rownames(axes)), 
+    colour="grey50",
+    size = 3) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
   xlim(c(-0.13, 0.13)) +

### 2024-08-23 paper.qmd
Added notes to what to write in Section 6

@@
 
-The \gD{} layouts generated by tSNE, UMAP, and PaCMAP show five well-separated clusters which evident these methods effectively preserve the global structure.
+XXX Add tSNE, UMAP, and PaCMAP NLDR layouts and describe what is happening
+XXX Discuss the models of one specific cluster (filled out, pancakes): tSNE, UMAP, PaCMAP are useful to capture global structures, but this is an evidence where PaCMAP fail to capture the local structure with default hyper-parameter settings than tSNE and UMAP
+XXX Why? hyper-parameter choice specially number of neighbors, data characteristics (intrinsic dimensionality), imbalance between global and local structures, optimization and initialization issues
+
+<!-- The \gD{} layouts generated by tSNE, UMAP, and PaCMAP show five well-separated clusters which evident these methods effectively preserve the global structure.
 In tSNE (@fig-gau-tsne-sc (a)), these clusters appear closely.
@@
 In contrast, PaCMAP shows one central cluster and the remaining four spread out in different directions (@fig-gau-pacmap-sc (a)).
-
-The tSNE and UMAP shows *filled out* clusters which provide evidence that these methods preserve the local structure (@fig-gau-tsne-sc (c) and @fig-gau-umap-sc (c)).
+-->
+
+<!-- The tSNE and UMAP shows *filled out* clusters which provide evidence that these methods preserve the local structure (@fig-gau-tsne-sc (c) and @fig-gau-umap-sc (c)).
 On the other hand, PaCMAP shows *flat* shapes clusters in the model and evident that PaCMAP fail to capture the within-cluster variation (@fig-gau-pacmap-sc (c)).
+-->
 
@@
 
+# Apply the scaling
+scaled_gau_data <- scale_data_manual(training_data_gau |> select(-ID)) |>
+  as_tibble()
+
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 14
+num_bins_x_gau1 <- 13
 lim1 <- gau1_scaled_obj$lim1
@@
   col_start_highd = "x",
-  q = 0.16
+  q = 0.1
 )
@@
   r2 = r2_gau1,
-  q = 0.16)
+  q = 0.1)
 
@@
 df_b <- df_b[match(df_bin_centroids_gau1$hexID, df_b$hb_id),] |>
-  dplyr::select(-hb_id)
-
-df_exe <- dplyr::bind_rows(df_b, df)
+  dplyr::select(-hb_id) |>
+  select(-type)
+
+# Apply the scaling
+scaled_gau_data_model <- scale_data_manual(df_b) |>
+  as_tibble()
 
@@
   distance_df_small_edges |> select(-distance), 
-  proj_model, 
+  projected_model_df, 
   by = c("from" = "ID"))
 
-names(model_df)[3:NCOL(model_df)] <- paste0(names(proj_model)[-NCOL(proj_model)], "_from")
-
-model_df <- dplyr::left_join(model_df, proj_model, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(proj_model)):NCOL(model_df)] <- paste0(names(proj_model)[-NCOL(proj_model)], "_to")
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+axes_obj <- gen_axes(
+  data = scaled_gau_data, 
+  proj_data = projected_df, 
+  limits = 1)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
 
 five_gau_proj_tsne_model <- projected_df |>
-  filter(type == "data") |>
   ggplot(
@@
   scale_color_manual(values = c("#000000", "#33a02c")) +
+  geom_segment(
+    data=axes, 
+    aes(x=x1, y=y1, xend=x2, yend=y2), 
+    colour="grey70") +
+  geom_text(
+    data=axes, 
+    aes(x=x2, y=y2, label=rownames(axes)), 
+    colour="grey50",
+    size = 3) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
   xlim(c(-0.18, 0.1)) +
@@
 df_b <- df_b[match(df_bin_centroids_gau1$hexID, df_b$hb_id),] |>
-  dplyr::select(-hb_id)
-
-df_exe <- dplyr::bind_rows(df_b, df)
+  dplyr::select(-hb_id)|>
+  select(-type)
+
+# Apply the scaling
+scaled_gau_data_model <- scale_data_manual(df_b) |>
+  as_tibble()
 
@@
   dplyr::filter(distance < benchmark_gau1)
-
-# Apply the scaling
-scaled_gau_data <- scale_data_manual(training_data_gau |> select(-ID)) |>
-  as_tibble()
-
-df_b <- df_b |>
-  select(-type)
 
@@
   col_start_highd = "x",
-  q = 0.14
+  q = 0.1
 )
@@
   r2 = r2_gau1,
-  q = 0.14)
+  q = 0.1)
 
@@
   distance_df_small_edges |> select(-distance), 
-  proj_model, 
+  projected_model_df, 
   by = c("from" = "ID"))
 
-names(model_df)[3:NCOL(model_df)] <- paste0(names(proj_model)[-NCOL(proj_model)], "_from")
-
-model_df <- dplyr::left_join(model_df, proj_model, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(proj_model)):NCOL(model_df)] <- paste0(names(proj_model)[-NCOL(proj_model)], "_to")
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+axes_obj <- gen_axes(
+  data = scaled_gau_data, 
+  proj_data = projected_df, 
+  limits = 1)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
 
 five_gau_proj_pacmap_model <- projected_df |>
-  filter(type == "data") |>
   ggplot(

### 2024-08-23 paper.tex
Regenerated Figure 2
Updated script to get the projections for UMAP
Add function to generate axes
Added notes to what to write in Section 6

@@
 
-The \gD{} layouts generated by tSNE, UMAP, and PaCMAP show five
-well-separated clusters which evident these methods effectively preserve
-the global structure. In tSNE (\textbf{?@fig-gau-tsne-sc} (a)), these
-clusters appear closely. UMAP arranges all clusters in a parallel
-manner, with three aligned in one line and the other two in a separate
-line (\textbf{?@fig-gau-umap-sc} (a)). In contrast, PaCMAP shows one
-central cluster and the remaining four spread out in different
-directions (\textbf{?@fig-gau-pacmap-sc} (a)).
-
-The tSNE and UMAP shows \emph{filled out} clusters which provide
-evidence that these methods preserve the local structure
-(\textbf{?@fig-gau-tsne-sc} (c) and \textbf{?@fig-gau-umap-sc} (c)). On
-the other hand, PaCMAP shows \emph{flat} shapes clusters in the model
-and evident that PaCMAP fail to capture the within-cluster variation
-(\textbf{?@fig-gau-pacmap-sc} (c)).
-
-\begin{figure}
+XXX Add tSNE, UMAP, and PaCMAP NLDR layouts and describe what is
+happening XXX Discuss the models of one specific cluster (filled out,
+pancakes): tSNE, UMAP, PaCMAP are useful to capture global structures,
+but this is an evidence where PaCMAP fail to capture the local structure
+with default hyper-parameter settings than tSNE and UMAP XXX Why?
+hyper-parameter choice specially number of neighbors, data
+characteristics (intrinsic dimensionality), imbalance between global and
+local structures, optimization and initialization issues
+
+\begin{figure}[H]
 

### 2024-08-23 scripts/five_gau_clusters/3_gen_model_in_data_space_with_pacmap.R

@@ -95,6 +95,43 @@ langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
                          group = df_exe$type, pointSize = append(rep(2, NROW(df_b)), rep(0.6, NROW(df))),
                          levelColors = c("#6a3d9a", "#33a02c"))
 
+#### With scaled data
+
+# Apply the scaling
+scaled_gau_data <- scale_data_manual(training_data_gau |> select(-ID)) |>
+  as_tibble()
+
+df_b <- df_bin_gau1 |>
+  dplyr::filter(hb_id %in% df_bin_centroids_gau1$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b <- df_b[match(df_bin_centroids_gau1$hexID, df_b$hb_id),] |>
+  dplyr::select(-hb_id) |>
+  select(-type)
+
+# Apply the scaling
+scaled_gau_data_model <- scale_data_manual(df_b) |>
+  as_tibble()
+
+# Combine with the true model for visualization
+df <- dplyr::bind_rows(scaled_gau_data_model |> mutate(type = "model"),
+                       scaled_gau_data |> mutate(type = "data"))
+
+## Set the maximum difference as the criteria
+distance_df_small_edges <- distance_gau1 |>
+  dplyr::filter(distance < benchmark_gau1)
+
+# Visualize with langevitour
+langevitour(df |> dplyr::select(-type),
+            lineFrom = distance_df_small_edges$from,
+            lineTo = distance_df_small_edges$to,
+            group = df$type,
+            pointSize = append(rep(1.5, NROW(scaled_gau_data_model)), rep(1, NROW(scaled_gau_data))),
+            levelColors = c("#000000", "#33a02c"),
+            lineColors = rep("#33a02c", nrow(distance_df_small_edges)))
+
+
 
 ## First projection
 projection <- cbind(

### 2024-08-23 paper.qmd
Updated Figure 9 (a)

@@
            b_non_empty = NROW(df_bin_centroids_scurve),
-           a1 = a1)
+           a1 = round(a1, 2))
 
@@
 
+## Find the minimum MSE when have duplicate a1
+error_scurve <- error_scurve |>
+  group_by(a1) |>
+  filter(MSE == min(MSE)) |>
+  ungroup()
+
 mse_scurve_b <- ggplot(error_scurve, 
@@
                          y = MSE)) +
-  geom_point(size = 1) +
-  geom_line(linewidth = 0.05) + 
-  geom_vline(xintercept = 0.096, linetype=2,
+  geom_vline(xintercept = 0.10, linetype=2,
            color = "#bdbdbd", linewidth=1) +
-  geom_vline(xintercept = 0.116,
+  geom_vline(xintercept = 0.12,
              linetype="solid",
@@
              linewidth=1) +
-  geom_vline(xintercept = 0.184, linetype=2,
+  geom_vline(xintercept = 0.18, linetype=2,
              color = "#bdbdbd", linewidth=1) +
+  geom_line(linewidth = 0.5) + 
+  geom_point(size = 1) +
   scale_x_continuous(breaks = sort(unique(round(error_scurve$a1, 2)))[c(1, 3, 5, 6, 8, 10:12)]) +
@@
   ggtitle("(a)") +
+  theme_minimal() +
   theme(aspect.ratio = 0.75,
-        axis.text.x = element_text(size = 12),
-        axis.text.y = element_text(size = 12),
-        axis.title.x = element_text(size = 12),
-        axis.title.y = element_text(size = 12, 
-            angle=90),
-        axis.ticks.x = element_line(linewidth = 0.5),
-        axis.ticks.y = element_line(linewidth = 0.5),
-        plot.title = element_text(size = 12),
-        legend.title = element_text(size = 12),
-        legend.text = element_text(size = 10),
-        legend.key.height = unit(1, 'cm'),
-        legend.key.width = unit(1, 'cm'),
-        legend.key.size = unit(1, 'cm'))
+        panel.border = element_rect(fill = 'transparent'),
+        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5))
 
@@
 XXX Why? hyper-parameter choice specially number of neighbors, data characteristics (intrinsic dimensionality), imbalance between global and local structures, optimization and initialization issues
+XXX Add this is with specific hyper-parameter choices
 

### 2024-08-23 paper.tex
Updated Figure 9 (a)

@@
 characteristics (intrinsic dimensionality), imbalance between global and
-local structures, optimization and initialization issues
+local structures, optimization and initialization issues XXX Add this is
+with specific hyper-parameter choices
 

### 2024-08-23 paper.qmd
Updated Figure 9

@@
         panel.border = element_rect(fill = 'transparent'),
-        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5))
+        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+        axis.ticks.x = element_line(),
+        axis.ticks.y = element_line())
 
@@
                          y = b_non_empty)) +
-  geom_point(size = 1) +
-  geom_line(linewidth = 0.05) +
-  geom_vline(xintercept = 0.116,
+  geom_vline(xintercept = 0.12,
          linetype=2,
@@
            linewidth=1) +
+  geom_point(size = 1) +
+  geom_line(linewidth = 0.05) +
   # geom_vline(xintercept = 11, linetype="solid",
@@
   ggtitle("(b)") +
+  theme_minimal() +
   theme(aspect.ratio = 0.75,
-        axis.text.x = element_text(size = 12),
-        axis.text.y = element_text(size = 11),
-        axis.title.x = element_text(size = 12),
-        axis.title.y = element_text(size = 12,
-            angle=90),
-        axis.ticks.x = element_line(linewidth = 0.5),
-        axis.ticks.y = element_line(linewidth = 0.5),
-        plot.title = element_text(size = 12),
-        legend.title = element_text(size = 12),
-        legend.text = element_text(size = 10),
-        legend.key.height = unit(1, 'cm'),
-        legend.key.width = unit(1, 'cm'),
-        legend.key.size = unit(1, 'cm'))
+        panel.border = element_rect(fill = 'transparent'),
+        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+        axis.ticks.x = element_line(),
+        axis.ticks.y = element_line())
 
@@
   ggtitle("(c)") +
+  theme_minimal() +
   theme(aspect.ratio = 0.75,
-        axis.text.x = element_text(size = 12),
-        axis.text.y = element_text(size = 12),
-        axis.title.x = element_text(size = 12),
-        axis.title.y = element_text(size = 12, 
-            angle=90),
-        plot.title = element_text(size = 12),
-        axis.ticks.x = element_line(linewidth = 0.5),
-        axis.ticks.y = element_line(linewidth = 0.5))
+        panel.border = element_rect(fill = 'transparent'),
+        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+        axis.ticks.x = element_line(),
+        axis.ticks.y = element_line())
 
@@
   ggtitle("(d)") +
+  theme_minimal() +
   theme(aspect.ratio = 0.75,
-        axis.title.x = element_text(size = 12),
-        axis.title.y = element_text(size = 12, angle = 90),
-        axis.text.x = element_text(size = 12),
-        axis.text.y = element_text(size = 12),
+        panel.border = element_rect(fill = 'transparent'),
+        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+        panel.grid.major.x = element_blank(),
         axis.ticks.x = element_line(),
-        plot.title = element_text(size = 12))
+        axis.ticks.y = element_line())
 

### 2024-08-23 paper.qmd
Updated Figure 13

@@
   xlab(expression(a[1])) +
-  theme(#aspect.ratio = 0.75,
-        axis.text.x = element_text(size = 7),
-        axis.text.y = element_text(size = 7),
-        axis.title.x = element_text(size = 7),
-        axis.title.y = element_text(size = 7, angle = 90),
-        axis.ticks.x = element_line(linewidth = 0.5),
-        axis.ticks.y = element_line(linewidth = 0.5),
-        legend.position = "none") #+
-  # inset_element(p = pbmc_layouts, 
-  #               left = 1.55, 
-  #               bottom = 0.35, 
-  #               right = 0, 
-  #               top = 1)
+  theme_minimal() +
+  theme(panel.border = element_rect(fill = 'transparent'),
+        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+        axis.ticks.x = element_line(),
+        axis.ticks.y = element_line(),
+        legend.position = "none")
 

### 2024-08-23 paper.qmd
Updated Figure 13

@@
 error_pbmc <- error_pbmc |>
-  filter(bin1 >= 5)
+  mutate(a1 = round(a1, 2)) |>
+  filter(bin1 >= 5) |>
+  group_by(method, a1) |>
+  filter(MSE == min(MSE)) |>
+  ungroup()
 ```
@@
   #            color = "black", linewidth=0.8, alpha = 0.5) +
+  scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 3, 5, 7, 9, 11, 13, 15, 17, 19, 21, 24, 26)]) +
   scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a','#ff7f00',
@@
         axis.ticks.y = element_line(),
-        legend.position = "none")
+        legend.position = "none",
+        axis.text.x = element_text(size = 7),
+        axis.text.y = element_text(size = 7),
+        axis.title.x = element_text(size = 7),
+        axis.title.y = element_text(size = 7))
 

### 2024-08-23 paper.qmd
Changed number of bins in MNIST

@@
 ## Compute hexbin parameters
-num_bins_x_mnist <- 22
+num_bins_x_mnist <- 19
 lim1 <- mnist_scaled_obj$lim1

### 2024-08-25 paper.qmd
MNIST removed model points from the trimesh

@@
 
-trimesh_removed_mnist <- vis_rmlg_mesh(
-  distance_edges = distance_mnist, 
-  benchmark_value = benchmark_mnist, 
-  tr_coord_df = tr_from_to_df_mnist, 
-  distance_col = "distance") +
+tr_df <- distinct(tibble::tibble(
+  x = c(tr_from_to_df_mnist[["x_from"]], tr_from_to_df_mnist[["x_to"]]), 
+  y = c(tr_from_to_df_mnist[["y_from"]], tr_from_to_df_mnist[["y_to"]])))
+
+distance_df_small_edges_mnist <- distance_mnist |>
+  filter(distance < benchmark_mnist)
+
+tr_from_to_df_mnist <- inner_join(
+  tr_from_to_df_mnist, distance_df_small_edges_mnist, 
+  by = c("from", "to"))
+
+trimesh_removed_mnist <- ggplot() + 
+  geom_segment(data = tr_from_to_df_mnist, 
+               aes(
+                 x = x_from, 
+                 y = y_from, 
+                 xend = x_to, 
+                 yend = y_to),
+               colour = "#33a02c") +
   geom_point(data = pacmap_minst_scaled,
@@
   interior_annotation("a", sc_ltr_pos_mnist)
+
+# trimesh_removed_mnist <- vis_rmlg_mesh(
+#   distance_edges = distance_mnist, 
+#   benchmark_value = benchmark_mnist, 
+#   tr_coord_df = tr_from_to_df_mnist, 
+#   distance_col = "distance") +
+#   geom_point(data = pacmap_minst_scaled,
+#              aes(
+#                x = PaCMAP1,
+#                y = PaCMAP2
+#              ),
+#              alpha=0.5, 
+#              size = 0.1) +
+#   interior_annotation("a", sc_ltr_pos_mnist)
 

### 2024-08-25 paper.qmd
Updated Figure 16

@@
   geom_point(alpha=0.5, 
-             size = 0.1) +
-  interior_annotation("b", sc_ltr_pos_mnist)
+             size = 0.1) #+
+  #interior_annotation("b", sc_ltr_pos_mnist)
 ```
@@
 #| echo: false
-#| fig-cap: "(a) Model generated in $2\\text{-}D$, and (b) $p\\text{-}D$ model error in $2\\text{-}D$.
+#| fig-cap: "$10\\text{-}D$ model error in $2\\text{-}D$.
 The $2\\text{-}D$ model shows a non-linear continuous structure.
-Most low $p\\text{-}D$ model errors are distributed along the lower edge of the $2\\text{-}D$ structure, while most high p-D model errors are concentrated along the upper edge."
+Most low $10\\text{-}D$ model errors are distributed along the lower edge of the $2\\text{-}D$ structure, while most large $10\\text{-}D$ model errors are concentrated along the upper edge."
 #| label: fig-model-mnist
 #| fig-pos: H
-
-trimesh_removed_mnist + error_plot_mnist + 
-  plot_layout(guides='collect', ncol=2) &
+#| out-height: 22%
+
+error_plot_mnist + 
+  plot_layout(guides='collect') &
   theme(legend.position='bottom')

### 2024-08-25 paper.tex
Updated Figure 16

@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-model-mnist-1.pdf}
-
-}
-
-\caption{\label{fig-model-mnist}(a) Model generated in \(2\text{-}D\),
-and (b) \(p\text{-}D\) model error in \(2\text{-}D\). The \(2\text{-}D\)
-model shows a non-linear continuous structure. Most low \(p\text{-}D\)
-model errors are distributed along the lower edge of the \(2\text{-}D\)
-structure, while most high p-D model errors are concentrated along the
-upper edge.}
+\includegraphics[width=1\textwidth,height=0.22\textheight]{paper_files/figure-pdf/fig-model-mnist-1.pdf}
+
+}
+
+\caption{\label{fig-model-mnist}\(10\text{-}D\) model error in
+\(2\text{-}D\). The \(2\text{-}D\) model shows a non-linear continuous
+structure. Most low \(10\text{-}D\) model errors are distributed along
+the lower edge of the \(2\text{-}D\) structure, while most large
+\(10\text{-}D\) model errors are concentrated along the upper edge.}
 

### 2024-08-25 paper.qmd
Updated Figure 17

@@
 
+```{r}
+#| label: mnist-model-proj
+
+# Apply the scaling
+scaled_mnist_data <- scale_data_manual(training_data_mnist |> select(-ID)) |>
+  as_tibble()
+
+df_b_mnist <- df_bin_mnist |>
+  dplyr::filter(hb_id %in% df_bin_centroids_mnist$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b_mnist <- df_b_mnist[match(df_bin_centroids_mnist$hexID, df_b_mnist$hb_id),] |>
+  dplyr::select(-hb_id) |>
+  select(-type)
+
+# Apply the scaling
+scaled_mnist_data_model <- scale_data_manual(df_b_mnist) |>
+  as_tibble()
+
+# Combine with the true model for visualization
+df <- dplyr::bind_rows(scaled_mnist_data_model |> mutate(type = "model"),
+                       scaled_mnist_data |> mutate(type = "data"))
+
+## First projection
+projection <- cbind(
+    c(0.71503,0.00502,-0.04370,-0.01811,-0.01103,0.09877,0.06199,0.05396,0.07771,-0.01713),
+    c(0.00587,0.71879,0.03485,0.08284,-0.07534,-0.05420,0.04315,-0.00547,-0.02995,0.01810))
+
+projected <- as.matrix(scaled_mnist_data) %*% projection
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  #dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+projected_model <- as.matrix(scaled_mnist_data_model) %*% projection
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges_mnist |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+axes_obj <- gen_axes(
+  data = scaled_mnist_data, 
+  proj_data = projected_df, 
+  limits = 1)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
+
+mnist_proj_pacmap_model1 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#33a02c") +
+  geom_point(
+    size = 0.5,
+    alpha = 0.2) +
+  scale_color_manual(values = c("#000000", "#33a02c")) +
+  geom_segment(
+    data=axes, 
+    aes(x=x1, y=y1, xend=x2, yend=y2), 
+    colour="grey70") +
+  geom_text(
+    data=axes, 
+    aes(x=x2, y=y2, label=rownames(axes)), 
+    colour="grey50",
+    size = 3) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  interior_annotation("b")
+
+## Second projection
+projection <- cbind(
+    c(0.13639,-0.18552,-0.38537,-0.03088,-0.04663,0.15013,-0.28351,-0.28865,-0.03379,-0.38009),
+    c(-0.46592,-0.40953,-0.02752,0.12815,0.08356,-0.12991,-0.28307,0.11145,0.08031,0.10800))
+
+projected <- as.matrix(scaled_mnist_data) %*% projection
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  #dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+projected_model <- as.matrix(scaled_mnist_data_model) %*% projection
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges_mnist |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+axes_obj <- gen_axes(
+  data = scaled_mnist_data, 
+  proj_data = projected_df, 
+  limits = 1)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
+
+mnist_proj_pacmap_model2 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#33a02c") +
+  geom_point(
+    size = 0.5,
+    alpha = 0.1) +
+  scale_color_manual(values = c("#000000", "#33a02c")) +
+  geom_segment(
+    data=axes, 
+    aes(x=x1, y=y1, xend=x2, yend=y2), 
+    colour="grey70") +
+  geom_text(
+    data=axes, 
+    aes(x=x2, y=y2, label=rownames(axes)), 
+    colour="grey50",
+    size = 3) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  interior_annotation("c")
+
+## Third projection
+projection <- cbind(
+    c(0.47163,0.21622,0.04027,0.17328,0.03603,0.06488,-0.22226,-0.07186,0.39674,0.13485),
+    c(-0.41187,0.36322,0.11264,-0.14175,0.19810,0.07062,0.00680,0.08795,0.37068,-0.11280))
+
+projected <- as.matrix(scaled_mnist_data) %*% projection
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  #dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+projected_model <- as.matrix(scaled_mnist_data_model) %*% projection
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges_mnist |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+axes_obj <- gen_axes(
+  data = scaled_mnist_data, 
+  proj_data = projected_df, 
+  limits = 1)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
+
+mnist_proj_pacmap_model3 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#33a02c") +
+  geom_point(
+    size = 0.5,
+    alpha = 0.2) +
+  scale_color_manual(values = c("#000000", "#33a02c")) +
+  geom_segment(
+    data=axes, 
+    aes(x=x1, y=y1, xend=x2, yend=y2), 
+    colour="grey70") +
+  geom_text(
+    data=axes, 
+    aes(x=x2, y=y2, label=rownames(axes)), 
+    colour="grey50",
+    size = 3) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  interior_annotation("d")
+
+```
+
 <!--add langevitour screenshots and youtube animation link-->
 ```{r}
-trimesh_removed_mnist 
-```
-
-::: {#fig-mnist1-sc layout-ncol="4" fig-pos="H"}
-![](figures/mnist/mnist_model_2d.png){#fig-mnist1-model}
-
-![](figures/mnist/sc_1.png){#fig-mnist1-sc1}
-
-![](figures/mnist/sc_2.png){#fig-mnist1-sc2}
-
-![](figures/mnist/sc_3.png){#fig-mnist1-sc3}
-
-Model in \gD{}, on the PaCMAP layout, and three views of the fit in projections from $10\text{-}D$, for the digit 1 of MNIST data ($(s_1, \ s_2) = (-0.100, \ -0.059)$, $b = 374 \  (22, \ 17)$, $m = 140$, benchmark value to remove large edges is $0.094$).
+#| label: fig-mnist-tri-proj
+#| fig-pos: H
+#| fig-height: 10
+#| fig-width: 10
+#| fig-cap: "Model in \\gD{}, on the PaCMAP layout, and three views of the fit in projections from $10\\text{-}D$, for the digit 1 of MNIST data ($(s_1, \ s_2) = (-0.100, \ -0.059)$, $b = 374 \  (22, \ 17)$, $m = 140$, benchmark value to remove large edges is $0.094$).
+(The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/zcg_GXBmqjA>)."
+
+trimesh_removed_mnist + mnist_proj_pacmap_model1 +
+  mnist_proj_pacmap_model2 + mnist_proj_pacmap_model3 +
+  plot_layout(ncol = 2)
+```
+
+<!-- ::: {#fig-mnist1-sc layout-ncol="4" fig-pos="H"} -->
+<!-- ![](figures/mnist/mnist_model_2d.png){#fig-mnist1-model} -->
+
+<!-- ![](figures/mnist/sc_1.png){#fig-mnist1-sc1} -->
+
+<!-- ![](figures/mnist/sc_2.png){#fig-mnist1-sc2} -->
+
+<!-- ![](figures/mnist/sc_3.png){#fig-mnist1-sc3} -->
+
+<!-- Model in \gD{}, on the PaCMAP layout, and three views of the fit in projections from $10\text{-}D$, for the digit 1 of MNIST data ($(s_1, \ s_2) = (-0.100, \ -0.059)$, $b = 374 \  (22, \ 17)$, $m = 140$, benchmark value to remove large edges is $0.094$).
 (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/zcg_GXBmqjA>).
-
-:::
+-->
+
+<!-- ::: -->
 <!-- need to update the youtube recording-->

### 2024-08-25 paper.tex
Updated Figure 17

@@
 
-According to Figure~\ref{fig-mnist1-sc1}, the non-linear continuous
+According to \textbf{?@fig-mnist1-sc1}, the non-linear continuous
 structure observed in the \gD{} representation of PaCMAP
@@
 a twisted pattern within the non-linear structure in \pD{} space
-(Figure~\ref{fig-mnist1-sc2}), which is an additional pattern not
-visible in the \gD{} representation (Figure~\ref{fig-pacmap-author}).
-Furthermore, as shown in Figure~\ref{fig-mnist1-sc3}, some long edges
+(\textbf{?@fig-mnist1-sc2}), which is an additional pattern not visible
+in the \gD{} representation (Figure~\ref{fig-pacmap-author}).
+Furthermore, as shown in \textbf{?@fig-mnist1-sc3}, some long edges
 exist in the \pD{} space that are not recognized as long edges in the
@@
 
-\begin{center}
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-60-1.pdf}
-\end{center}
-
-\begin{figure}[H]
-
-\begin{minipage}{0.25\linewidth}
-
-\centering{
-
-\includegraphics{figures/mnist/mnist_model_2d.png}
-
-}
-
-\subcaption{\label{fig-mnist1-model}}
-
-\end{minipage}%
-%
-\begin{minipage}{0.25\linewidth}
-
-\centering{
-
-\includegraphics{figures/mnist/sc_1.png}
-
-}
-
-\subcaption{\label{fig-mnist1-sc1}}
-
-\end{minipage}%
-%
-\begin{minipage}{0.25\linewidth}
-
-\centering{
-
-\includegraphics{figures/mnist/sc_2.png}
-
-}
-
-\subcaption{\label{fig-mnist1-sc2}}
-
-\end{minipage}%
-%
-\begin{minipage}{0.25\linewidth}
-
-\centering{
-
-\includegraphics{figures/mnist/sc_3.png}
-
-}
-
-\subcaption{\label{fig-mnist1-sc3}}
-
-\end{minipage}%
-
-\caption{\label{fig-mnist1-sc}Model in \gD{}, on the PaCMAP layout, and
-three views of the fit in projections from \(10\text{-}D\), for the
-digit 1 of MNIST data (\((s_1, \ s_2) = (-0.100, \ -0.059)\),
-\(b = 374 \  (22, \ 17)\), \(m = 140\), benchmark value to remove large
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-mnist-tri-proj-1.pdf}
+
+}
+
+\caption{\label{fig-mnist-tri-proj}Model in \gD{}, on the PaCMAP layout,
+and three views of the fit in projections from \(10\text{-}D\), for the
+digit 1 of MNIST data (\((s_1,  s_2) = (-0.100,  -0.059)\),
+\(b = 374   (22,  17)\), \(m = 140\), benchmark value to remove large
 edges is \(0.094\)). (The \textbf{langevitour} software is used to view

### 2024-08-25 paper.qmd
Updated trimesh for pbmc tSNE model

@@
   r2 = r2_pbmc,
-  q = 0.05,
+  q = 0.1,
   is_bin_centroid = TRUE,
@@
 
-trimesh_removed_pbmc <- vis_rmlg_mesh(
-  distance_edges = distance_pbmc,
-  benchmark_value = benchmark_pbmc,
-  tr_coord_df = tr_from_to_df_pbmc,
-  distance_col = "distance") +
+tr_df <- distinct(tibble(
+  x = c(tr_from_to_df_pbmc[["x_from"]], tr_from_to_df_pbmc[["x_to"]]), 
+  y = c(tr_from_to_df_pbmc[["y_from"]], tr_from_to_df_pbmc[["y_to"]])))
+
+distance_df_small_edges_pbmc <- distance_pbmc |>
+  filter(distance < benchmark_pbmc)
+
+tr_from_to_df_pbmc <- inner_join(
+  tr_from_to_df_pbmc, distance_df_small_edges_pbmc, 
+  by = c("from", "to"))
+
+# trimesh_removed_pbmc <- vis_rmlg_mesh(
+#   distance_edges = distance_pbmc,
+#   benchmark_value = benchmark_pbmc,
+#   tr_coord_df = tr_from_to_df_pbmc,
+#   distance_col = "distance") +
+#   geom_point(
+#     data = umap_pbmc_scaled,
+#     aes(
+#       x = UMAP1,
+#       y = UMAP2
+#     ),
+#     alpha = 0.3,
+#     size = 0.1,
+#     color = "#000000"
+#   )
+
+trimesh_removed_pbmc <- ggplot() + 
+  geom_segment(data = tr_from_to_df_pbmc, 
+               aes(
+                 x = x_from, 
+                 y = y_from, 
+                 xend = x_to, 
+                 yend = y_to),
+               colour = "#33a02c") +
   geom_point(
@@
     color = "#000000"
-  )
-
-## Compute error
-error_df <- augment(
-  df_bin_centroids = df_bin_centroids_pbmc,
-  df_bin = df_bin_pbmc,
-  training_data = training_data_pbmc,
-  newdata = NULL,
-  type_NLDR = "UMAP",
-  col_start = "PC_")
-
-## Categorize error
-
-error_df <- error_df |>
-  mutate(type = case_when(
-    row_wise_abs_error <= 5 ~ "error 0-5",
-    row_wise_abs_error <= 10 ~ "error 5-10",
-    row_wise_abs_error <= 15 ~ "error 10-15",
-    row_wise_abs_error <= 20 ~ "error 15-20",
-    .default = "error greter than 20"
-  )) |>
-  mutate(type = factor(type, levels = c(
-    "error 0-5", "error 5-10", "error 10-15", "error 15-20", "error greter than 20")))
-
-## To join embedding
-error_df <- error_df |>
-  bind_cols(umap_pbmc_scaled |>
-              select(-ID))
-
-error_plot_pbmc <- error_df |>
-  ggplot(aes(x = UMAP1,
-             y = UMAP2,
-             color = type,
-             group = ID)) +
-  geom_point(alpha=0.5,
-             size = 0.1) +
-  interior_annotation("b", sc_ltr_pos)
+  )  +
+  interior_annotation("a")
+
+# ## Compute error
+# error_df <- augment(
+#   df_bin_centroids = df_bin_centroids_pbmc,
+#   df_bin = df_bin_pbmc,
+#   training_data = training_data_pbmc,
+#   newdata = NULL,
+#   type_NLDR = "UMAP",
+#   col_start = "PC_")
+
+# ## Categorize error
+# 
+# error_df <- error_df |>
+#   mutate(type = case_when(
+#     row_wise_abs_error <= 5 ~ "error 0-5",
+#     row_wise_abs_error <= 10 ~ "error 5-10",
+#     row_wise_abs_error <= 15 ~ "error 10-15",
+#     row_wise_abs_error <= 20 ~ "error 15-20",
+#     .default = "error greter than 20"
+#   )) |>
+#   mutate(type = factor(type, levels = c(
+#     "error 0-5", "error 5-10", "error 10-15", "error 15-20", "error greter than 20")))
+# 
+# ## To join embedding
+# error_df <- error_df |>
+#   bind_cols(umap_pbmc_scaled |>
+#               select(-ID))
+# 
+# error_plot_pbmc <- error_df |>
+#   ggplot(aes(x = UMAP1,
+#              y = UMAP2,
+#              color = type,
+#              group = ID)) +
+#   geom_point(alpha=0.5,
+#              size = 0.1) +
+#   interior_annotation("b", sc_ltr_pos)
 
@@
   theme(legend.position='bottom')
+```
+
+```{r}
+#| label: fig-model-pbmc-author-proj
+#| fig-pos: H
+trimesh_removed_pbmc 
 ```

### 2024-08-25 paper.qmd
Updated projections with tSNE

@@
 training_data_pbmc <- read_rds("data/pbmc3k/pbmc_pca_50.rds")
+names(training_data_pbmc) <- paste0("PC", 1:50)
+
 training_data_pbmc <- training_data_pbmc[, 1:9] |>
@@
 ## Compute hexbin parameters
-num_bins_x_pbmc <- 30
+num_bins_x_pbmc <- 26
 lim1 <- pbmc_scaled_obj$lim1
@@
   is_rm_lwd_hex = FALSE,
-  col_start_highd = "PC_"
+  col_start_highd = "PC"
 )
@@
   )  +
+  coord_fixed() +
   interior_annotation("a")
@@
 ```{r}
+#| label: pbmc-umap-model-proj
+
+# Apply the scaling
+scaled_pbmc_data <- scale_data_manual(training_data_pbmc |> select(-ID)) |>
+  as_tibble()
+
+df_b_pbmc <- df_bin_pbmc |>
+  dplyr::filter(hb_id %in% df_bin_centroids_pbmc$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b_pbmc <- df_b_pbmc[match(df_bin_centroids_pbmc$hexID, df_b_pbmc$hb_id),] |>
+  dplyr::select(-hb_id) |>
+  select(-type)
+
+# Apply the scaling
+scaled_pbmc_data_model <- scale_data_manual(df_b_pbmc) |>
+  as_tibble()
+
+# Combine with the true model for visualization
+df <- dplyr::bind_rows(scaled_pbmc_data_model |> mutate(type = "model"),
+                       scaled_pbmc_data |> mutate(type = "data"))
+
+## First projection
+projection <- cbind(
+    c(0.24156,0.85058,0.11564,-0.03255,-0.28106,0.27463,-0.10539,0.08429,0.13107),
+    c(-0.67212,0.02745,-0.26718,-0.33126,-0.34919,0.20839,-0.14498,-0.37708,0.15455))
+
+projected <- as.matrix(scaled_pbmc_data) %*% projection
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  #dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges_pbmc |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+axes_obj <- gen_axes(
+  data = scaled_pbmc_data, 
+  proj_data = projected_df, 
+  limits = 1)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
+
+pbmc_proj_tsne_model1 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#33a02c") +
+  geom_point(
+    size = 0.5,
+    alpha = 0.2) +
+  scale_color_manual(values = c("#000000", "#33a02c")) +
+  geom_segment(
+    data=axes, 
+    aes(x=x1, y=y1, xend=x2, yend=y2), 
+    colour="grey70") +
+  geom_text(
+    data=axes, 
+    aes(x=x2, y=y2, label=rownames(axes)), 
+    colour="grey50",
+    size = 3) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  interior_annotation("b")
+
+## Second projection
+projection <- cbind(
+    c(0.78689,-0.14774,0.36527,0.09617,0.12097,-0.00033,0.36973,-0.20959,-0.08438),
+    c(0.18390,-0.43121,-0.30398,0.05861,0.55282,-0.14058,-0.27048,0.44386,-0.27372))
+
+projected <- as.matrix(scaled_pbmc_data) %*% projection
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  #dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges_pbmc |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+axes_obj <- gen_axes(
+  data = scaled_pbmc_data, 
+  proj_data = projected_df, 
+  limits = 1)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
+
+pbmc_proj_tsne_model2 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#33a02c") +
+  geom_point(
+    size = 0.5,
+    alpha = 0.2) +
+  scale_color_manual(values = c("#000000", "#33a02c")) +
+  geom_segment(
+    data=axes, 
+    aes(x=x1, y=y1, xend=x2, yend=y2), 
+    colour="grey70") +
+  geom_text(
+    data=axes, 
+    aes(x=x2, y=y2, label=rownames(axes)), 
+    colour="grey50",
+    size = 3) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  interior_annotation("c")
+
+## Third projection
+projection <- cbind(
+    c(0.00067,0.53963,-0.16241,-0.09777,0.32766,0.44543,-0.51359,-0.29884,0.00951),
+    c(0.62399,0.26016,0.06441,-0.34686,-0.40466,0.24643,0.23537,0.07972,0.34402))
+
+projected <- as.matrix(scaled_pbmc_data) %*% projection
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  #dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges_pbmc |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+axes_obj <- gen_axes(
+  data = scaled_pbmc_data, 
+  proj_data = projected_df, 
+  limits = 1)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
+
+pbmc_proj_tsne_model3 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#33a02c") +
+  geom_point(
+    size = 0.5,
+    alpha = 0.2) +
+  scale_color_manual(values = c("#000000", "#33a02c")) +
+  geom_segment(
+    data=axes, 
+    aes(x=x1, y=y1, xend=x2, yend=y2), 
+    colour="grey70") +
+  geom_text(
+    data=axes, 
+    aes(x=x2, y=y2, label=rownames(axes)), 
+    colour="grey50",
+    size = 3) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  interior_annotation("d")
+
+```
+
+```{r}
 #| label: fig-model-pbmc-author-proj
 #| fig-pos: H
-trimesh_removed_pbmc 
-```
-
-::: {#fig-pbmc1-sc layout-ncol="4" fig-pos="H"}
-![](figures/pbmc3k/umap_trimesh_plot.png)
-
-![](figures/pbmc3k/sc_1.png)
-
-![](figures/pbmc3k/sc_2.png)
-
-![](figures/pbmc3k/sc_3.png)
-
-Model in \gD{}, on the UMAP layout, and three views of the fit in projections from $9\text{-}D$, for the PBMC3k data ($(s_1, \ s_2) = (-0.050, \ -0.041)$, $b = 870 \  (30, \ 29)$, $m = 135$, benchmark value to remove large edges is $0.099$) (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/VqqWuE0Jj6A>).
-:::
+#| fig-height: 10
+#| fig-width: 10
+#| fig-cap: "Model in \\gD{}, on the UMAP layout, and three views of the fit in projections from $9\\text{-}D$, for the PBMC3k data ($(s_1, \ s_2) = (-0.050, \ -0.041)$, $b = 870 \  (30, \ 29)$, $m = 135$, benchmark value to remove large edges is $0.099$) (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/VqqWuE0Jj6A>)."
+
+trimesh_removed_pbmc + pbmc_proj_tsne_model1 +
+  pbmc_proj_tsne_model2 + pbmc_proj_tsne_model3 +
+  plot_layout(ncol = 2)
+```
+
+<!-- ::: {#fig-pbmc1-sc layout-ncol="4" fig-pos="H"} -->
+<!-- ![](figures/pbmc3k/umap_trimesh_plot.png) -->
+
+<!-- ![](figures/pbmc3k/sc_1.png) -->
+
+<!-- ![](figures/pbmc3k/sc_2.png) -->
+
+<!-- ![](figures/pbmc3k/sc_3.png) -->
+
+<!-- Model in \gD{}, on the UMAP layout, and three views of the fit in projections from $9\text{-}D$, for the PBMC3k data ($(s_1, \ s_2) = (-0.050, \ -0.041)$, $b = 870 \  (30, \ 29)$, $m = 135$, benchmark value to remove large edges is $0.099$) (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/VqqWuE0Jj6A>).
+-->
+<!-- ::: -->
 
@@
   r2 = r2_pbmc,
-  q = 0.05,
+  q = 0.1,
   is_bin_centroid = TRUE,
   is_rm_lwd_hex = FALSE,
-  col_start_highd = "PC_"
+  col_start_highd = "PC"
 )
@@
 
-tr_from_to_df_pbmc <- tr_from_to_df_pbmc |>
-  filter(row_number() != 360)
+# tr_from_to_df_pbmc <- tr_from_to_df_pbmc |>
+#   filter(row_number() != 360)
 
@@
   type_NLDR = "tSNE",
-  col_start = "PC_")
+  col_start = "PC")
 
@@
   r2 = r2_pbmc,
-  q = 0.05)
+  q = 0.1)
 
@@
              size = 0.1) +
+  coord_fixed() +
   interior_annotation("a", sc_ltr_pos_mnist)

### 2024-08-25 paper.tex
Updated trimesh for pbmc tSNE model
Updated projections with tSNE

@@
 \(9\text{-}D\) data. The \textbf{?@fig-umap-author} shows three
-well-separated clusters. However, as shown in Figure~\ref{fig-pbmc1-sc},
+well-separated clusters. However, as shown in \textbf{?@fig-pbmc1-sc},
 there is no big separation between three clusters in \(9\text{-}D\).
@@
 clusters have non-linear continuity patterns and high-density patches
-(Figure~\ref{fig-pbmc1-sc}).
-
-\begin{figure}[H]
-
-\begin{minipage}{0.25\linewidth}
-\includegraphics{figures/pbmc3k/umap_trimesh_plot.png}\end{minipage}%
-%
-\begin{minipage}{0.25\linewidth}
-\includegraphics{figures/pbmc3k/sc_1.png}\end{minipage}%
-%
-\begin{minipage}{0.25\linewidth}
-\includegraphics{figures/pbmc3k/sc_2.png}\end{minipage}%
-%
-\begin{minipage}{0.25\linewidth}
-\includegraphics{figures/pbmc3k/sc_3.png}\end{minipage}%
-
-\caption{\label{fig-pbmc1-sc}Model in \gD{}, on the UMAP layout, and
-three views of the fit in projections from \(9\text{-}D\), for the
-PBMC3k data (\((s_1, \ s_2) = (-0.050, \ -0.041)\),
-\(b = 870 \  (30, \ 29)\), \(m = 135\), benchmark value to remove large
+(\textbf{?@fig-pbmc1-sc}).
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-model-pbmc-author-proj-1.pdf}
+
+}
+
+\caption{\label{fig-model-pbmc-author-proj}Model in \gD{}, on the UMAP
+layout, and three views of the fit in projections from \(9\text{-}D\),
+for the PBMC3k data (\((s_1,  s_2) = (-0.050,  -0.041)\),
+\(b = 870   (30,  29)\), \(m = 135\), benchmark value to remove large
 edges is \(0.099\)) (The \textbf{langevitour} software is used to view

### 2024-08-25 paper.qmd
Updated Figure 14

@@
 ## Compute hexbin parameters
-num_bins_x_pbmc <- 15
+num_bins_x_pbmc <- 13
 lim1 <- pbmc_scaled_obj$lim1
@@
 # tr_from_to_df_pbmc <- tr_from_to_df_pbmc |>
-#   filter(row_number() != 360)
+#   filter(!(row_number() %in% c(65, 112, 130)))
 
@@
 
+benchmark_pbmc <- 0.1
+
 ## Remove an edge that exist as long edge in p-D
@@
 
-trimesh_removed_pbmc <- vis_rmlg_mesh(
-  distance_edges = distance_pbmc,
-  benchmark_value = benchmark_pbmc,
-  tr_coord_df = tr_from_to_df_pbmc,
-  distance_col = "distance") +
+tr_df <- distinct(tibble::tibble(
+  x = c(tr_from_to_df_pbmc[["x_from"]], tr_from_to_df_pbmc[["x_to"]]), 
+  y = c(tr_from_to_df_pbmc[["y_from"]], tr_from_to_df_pbmc[["y_to"]])))
+
+distance_df_small_edges_pbmc <- distance_pbmc |>
+  filter(distance < benchmark_pbmc)
+
+tr_from_to_df_pbmc <- inner_join(
+  tr_from_to_df_pbmc, distance_df_small_edges_pbmc, 
+  by = c("from", "to"))
+
+tr_from_to_df_pbmc <- tr_from_to_df_pbmc |>
+  filter(!(row_number() %in% c(155)))
+
+trimesh_removed_pbmc <- ggplot() + 
+  geom_segment(data = tr_from_to_df_pbmc, 
+               aes(
+                 x = x_from, 
+                 y = y_from, 
+                 xend = x_to, 
+                 yend = y_to),
+               colour = "#33a02c") +
   geom_point(
@@
     color = "#000000"
-  ) 
-  #xlim(sc_xlims) + ylim(sc_ylims) +
-  #interior_annotation("a", sc_ltr_pos)
+  )  +
+  coord_fixed() +
+  interior_annotation("a")
+
+# trimesh_removed_pbmc <- vis_rmlg_mesh(
+#   distance_edges = distance_pbmc,
+#   benchmark_value = benchmark_pbmc,
+#   tr_coord_df = tr_from_to_df_pbmc,
+#   distance_col = "distance") +
+#   geom_point(
+#     data = tsne_pbmc_scaled,
+#     aes(
+#       x = tSNE1,
+#       y = tSNE2
+#     ),
+#     alpha = 0.2,
+#     color = "#000000"
+#   ) 
+#   #xlim(sc_xlims) + ylim(sc_ylims) +
+#   #interior_annotation("a", sc_ltr_pos)
 
@@
 ```{r}
-#| label: proj-tsne-pbmc
-
-## Hexagonal binning to have regular hexagons
-hb_obj_pbmc <- hex_binning(
-  data = tsne_pbmc_scaled,
-  bin1 = num_bins_x_pbmc,
-  r2 = r2_pbmc,
-  q = 0.1)
-
-tsne_data_with_hb_id <- hb_obj_pbmc$data_hb_id
-
-df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID),
-                                tsne_data_with_hb_id)
-
-### Define type column
-df <- df_all_pbmc |>
-  dplyr::select(tidyselect::starts_with("PC")) |>
-  dplyr::mutate(type = "data") ## original dataset
-
-df_b <- df_bin_pbmc |>
+#| label: pbmc-tsne-model-proj
+
+# Apply the scaling
+scaled_pbmc_data <- scale_data_manual(training_data_pbmc |> select(-ID)) |>
+  as_tibble()
+
+df_b_pbmc <- df_bin_pbmc |>
   dplyr::filter(hb_id %in% df_bin_centroids_pbmc$hexID) |>
@@
 ## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
-df_b <- df_b[match(df_bin_centroids_pbmc$hexID, df_b$hb_id),] |>
-  dplyr::select(-hb_id)
-
-df_exe <- dplyr::bind_rows(df_b, df)
-
-## Set the maximum difference as the criteria
-distance_df_small_edges <- distance_pbmc |>
-  dplyr::filter(distance < benchmark_pbmc)
-## Since erase brushing is considerd.
+df_b_pbmc <- df_b_pbmc[match(df_bin_centroids_pbmc$hexID, df_b_pbmc$hb_id),] |>
+  dplyr::select(-hb_id) |>
+  select(-type)
+
+# Apply the scaling
+scaled_pbmc_data_model <- scale_data_manual(df_b_pbmc) |>
+  as_tibble()
+
+# Combine with the true model for visualization
+df <- dplyr::bind_rows(scaled_pbmc_data_model |> mutate(type = "model"),
+                       scaled_pbmc_data |> mutate(type = "data"))
+
+## Removed long edge in high-d
+distance_df_small_edges_pbmc <- distance_df_small_edges_pbmc |>
+  filter(!(row_number() %in% c(155)))
 
@@
 projection <- cbind(
-  c(0.038104,-0.001148,0.003905,-0.003332,-0.011747,-0.002688,0.003468,-0.014321,-0.000320),
-  c(-0.000902,0.041414,0.005089,-0.000149,-0.000551,0.000239,-0.008013,-0.005799,-0.001411))
-
-proj1_pbmc_tsne <- gen_proj_langevitour(
-  points_df = df_exe,
-  projection = projection,
-  edge_df = distance_df_small_edges |> select(-distance)
-)
+    c(0.40456,-0.55243,-0.58434,0.05063,-0.20540,-0.10469,0.28850,0.12209,0.17017),
+    c(0.34225,0.53615,-0.05822,-0.01993,-0.55480,0.21337,0.34100,-0.07280,-0.33139))
+
+projected <- as.matrix(scaled_pbmc_data) %*% projection
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  #dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges_pbmc |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+axes_obj <- gen_axes(
+  data = scaled_pbmc_data, 
+  proj_data = projected_df, 
+  limits = 1)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
+
+pbmc_proj_umap_model1 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#33a02c") +
+  geom_point(
+    size = 0.5,
+    alpha = 0.2) +
+  scale_color_manual(values = c("#000000", "#33a02c")) +
+  geom_segment(
+    data=axes, 
+    aes(x=x1, y=y1, xend=x2, yend=y2), 
+    colour="grey70") +
+  geom_text(
+    data=axes, 
+    aes(x=x2, y=y2, label=rownames(axes)), 
+    colour="grey50",
+    size = 3) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  interior_annotation("b")
 
@@
 projection <- cbind(
-  c(0.013182,-0.022240,-0.001729,0.026946,-0.005289,0.001235,-0.007784,0.017500,-0.006949),
-  c(0.027138,0.011939,0.009374,-0.004582,-0.004771,-0.027107,-0.005287,-0.002325,-0.007951))
-
-proj2_pbmc_tsne <- gen_proj_langevitour(
-  points_df = df_exe,
-  projection = projection,
-  edge_df = distance_df_small_edges |> select(-distance)
-)
+    c(0.07154,-0.44483,-0.04593,0.63343,0.21449,-0.39592,-0.29762,0.01726,0.30845),
+    c(0.87788,-0.21837,0.11780,-0.17226,-0.17920,-0.00829,0.11424,0.28681,0.06091))
+
+projected <- as.matrix(scaled_pbmc_data) %*% projection
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  #dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges_pbmc |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+axes_obj <- gen_axes(
+  data = scaled_pbmc_data, 
+  proj_data = projected_df, 
+  limits = 1)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
+
+pbmc_proj_umap_model2 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#33a02c") +
+  geom_point(
+    size = 0.5,
+    alpha = 0.2) +
+  scale_color_manual(values = c("#000000", "#33a02c")) +
+  geom_segment(
+    data=axes, 
+    aes(x=x1, y=y1, xend=x2, yend=y2), 
+    colour="grey70") +
+  geom_text(
+    data=axes, 
+    aes(x=x2, y=y2, label=rownames(axes)), 
+    colour="grey50",
+    size = 3) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  interior_annotation("c")
 
@@
 projection <- cbind(
-  c(-0.003540,0.025008,0.023261,-0.012782,-0.011617,-0.006796,0.009566,-0.015022,0.001164),
-  c(0.038201,-0.007751,0.005064,-0.014143,0.003255,-0.001091,0.001611,-0.003674,-0.008314))
-
-proj3_pbmc_tsne <- gen_proj_langevitour(
-  points_df = df_exe,
-  projection = projection,
-  edge_df = distance_df_small_edges |> select(-distance)
-)
+    c(0.13399,-0.00303,0.34841,-0.15626,-0.27432,0.48771,0.14252,0.39963,0.57983),
+    c(0.49615,0.73920,-0.15299,-0.07463,-0.07948,0.13320,-0.15484,0.20042,-0.28869))
+
+projected <- as.matrix(scaled_pbmc_data) %*% projection
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  #dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges_pbmc |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+axes_obj <- gen_axes(
+  data = scaled_pbmc_data, 
+  proj_data = projected_df, 
+  limits = 1)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
+
+pbmc_proj_umap_model3 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#33a02c") +
+  geom_point(
+    size = 0.5,
+    alpha = 0.2) +
+  scale_color_manual(values = c("#000000", "#33a02c")) +
+  geom_segment(
+    data=axes, 
+    aes(x=x1, y=y1, xend=x2, yend=y2), 
+    colour="grey70") +
+  geom_text(
+    data=axes, 
+    aes(x=x2, y=y2, label=rownames(axes)), 
+    colour="grey50",
+    size = 3) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  interior_annotation("d")
 
@@
 #| fig-pos: H
-#| out-height: 100%
-#| eval: true
-
-trimesh_removed_pbmc + proj1_pbmc_tsne + 
-  proj2_pbmc_tsne + proj3_pbmc_tsne +
+#| fig-width: 10
+#| fig-height: 10
+
+trimesh_removed_pbmc + pbmc_proj_umap_model1 + 
+  pbmc_proj_umap_model2 + pbmc_proj_umap_model3 +
   plot_layout(guides='collect', ncol=2) &

### 2024-08-25 paper.qmd
Updated Five Gau UMAP projections

@@
 projection <- cbind(
-    c(0.36638,0.31467,0.63030,0.24465),
-    c(0.14672,-0.73585,0.33246,-0.12977))
+  c(0.02743,0.21427,0.62893,-0.49816),
+  c(-0.70709,0.42156,-0.02593,0.10965))
 
@@
 
-axes_obj <- gen_axes(
-  data = scaled_gau_data, 
-  proj_data = projected_df, 
-  limits = 1)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
+# axes_obj <- gen_axes(
+#   data = scaled_gau_data, 
+#   proj_data = projected_df, 
+#   limits = 1)
+# 
+# axes <- axes_obj$axes
+# circle <- axes_obj$circle
+
+limits <- 1
+projected_df_n <- projected_df |> select(-ID)
+
+rng <- range(projected_df_n)
+projected_df_n <- projected_df_n/max(abs(rng))
+colnames(projected_df_n) <- c("P1", "P2")
+projected_df_n <- data.frame(projected_df_n)
+obs_labels <- as.character(1:nrow(scaled_gau_data))
+
+axis_scale <- limits/6
+# axis_pos <- -2/3 * limits
+axis_pos <- -1
+
+adj <- function(x) axis_pos + x * axis_scale
+axes <- data.frame(x1 = adj(0), 
+                   y1 = adj(0), 
+                   x2 = adj(projection[, 1]), 
+                   y2 = adj(projection[, 2]))
+
+axis_labels <- colnames(scaled_gau_data)
+rownames(axes) <- axis_labels
+
+theta <- seq(0, 2 * pi, length = 50)
+circle <- data.frame(c1 = adj(cos(theta)), 
+                     c2 = adj(sin(theta)))
 
@@
   geom_point(
+    size = 0.5,
+    alpha = 0.2) +
+  scale_color_manual(values = c("#000000", "#33a02c")) +
+  geom_segment(
+    data=axes, 
+    aes(x=x1, y=y1, xend=x2, yend=y2), 
+    colour="grey70") +
+  geom_text(
+    data=axes, 
+    aes(x=x2, y=y2, label=rownames(axes)), 
+    colour="grey50",
+    size = 3) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-1.25, 0.1)) +
+  ylim(c(-1.25, -0.5)) +
+  interior_annotation("c", 
+                      position = c(0.08, 0.9),
+                      cex = 3)
+
+## Second projection
+projection <- cbind(
+  c(0.36030,0.00630,0.66545,-0.34307),
+  c(-0.01861,0.14409,-0.36798,-0.73066))
+
+projected <- as.matrix(scaled_gau_data) %*% projection
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  #dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+projected_model <- as.matrix(scaled_gau_data_model) %*% projection
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+# axes_obj <- gen_axes(
+#   data = scaled_gau_data, 
+#   proj_data = projected_df, 
+#   limits = 1)
+# 
+# axes <- axes_obj$axes
+# circle <- axes_obj$circle
+
+limits <- 0.3
+projected_df_n <- projected_df |> select(-ID)
+
+rng <- range(projected_df_n)
+projected_df_n <- projected_df_n/max(abs(rng))
+colnames(projected_df_n) <- c("P1", "P2")
+projected_df_n <- data.frame(projected_df_n)
+obs_labels <- as.character(1:nrow(scaled_gau_data))
+
+axis_scale <- limits/6
+# axis_pos <- -2/3 * limits
+axis_pos <- 0.08
+
+adj <- function(x) axis_pos + x * axis_scale
+axes <- data.frame(x1 = adj(0), 
+                   y1 = adj(0), 
+                   x2 = adj(projection[, 1]), 
+                   y2 = adj(projection[, 2]))
+
+axis_labels <- colnames(scaled_gau_data)
+rownames(axes) <- axis_labels
+
+theta <- seq(0, 2 * pi, length = 50)
+circle <- data.frame(c1 = adj(cos(theta)), 
+                     c2 = adj(sin(theta)))
+
+five_gau_proj_umap_model2 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#33a02c") +
+  geom_point(
+    size = 0.5, 
     alpha = 0.3) +
@@
   coord_fixed() +
-  xlim(c(-0.45, -0.3)) +
-  ylim(c(-0.05, 0.2)) +
-  interior_annotation("c", 
-                      position = c(0.08, 0.9),
-                      cex = 3)
-
-## Second projection
-projection <- cbind(
-    c(-0.70767,-0.09092,0.15109,-0.39811),
-    c(-0.01312,-0.61805,0.44418,0.33305))
-
-projected <- as.matrix(scaled_gau_data) %*% projection
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  #dplyr::mutate(type = df_exe$type) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-projected_model <- as.matrix(scaled_gau_data_model) %*% projection
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  distance_df_small_edges |> select(-distance), 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-axes_obj <- gen_axes(
-  data = scaled_gau_data, 
-  proj_data = projected_df, 
-  limits = 1)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
-
-five_gau_proj_umap_model2 <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#33a02c") +
-  geom_point(
-    alpha = 0.3) +
-  scale_color_manual(values = c("#000000", "#33a02c")) +
-  geom_segment(
-    data=axes, 
-    aes(x=x1, y=y1, xend=x2, yend=y2), 
-    colour="grey70") +
-  geom_text(
-    data=axes, 
-    aes(x=x2, y=y2, label=rownames(axes)), 
-    colour="grey50",
-    size = 3) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.13, 0.13)) +
-  ylim(c(-0.13, 0.13)) +
+  xlim(c(0, 0.4)) +
+  ylim(c(0, 0.3)) +
   interior_annotation("d", 
@@
 #| label: fig-five-gau-projs
-#| fig-height: 12
-#| fig-width: 8
+#| fig-height: 15
+#| fig-width: 10
 #| fig-pos: H

### 2024-08-25 paper.qmd
Updated Figure 11

@@
 
-limits <- 1
+limits <- 0.5
 projected_df_n <- projected_df |> select(-ID)
@@
 # axis_pos <- -2/3 * limits
-axis_pos <- -1
+axis_pos <- -0.8
 
@@
   coord_fixed() +
-  xlim(c(-1.25, 0.1)) +
-  ylim(c(-1.25, -0.5)) +
+  xlim(c(-0.9, 0.1)) +
+  ylim(c(-0.9, -0.5)) +
   interior_annotation("c", 

### 2024-08-25 paper.qmd
Added Projections for PaCMAP Gau

@@
 
+## First projection
 projection <- cbind(
-  c(-0.00215,-0.68905,-0.04778,-0.54223),
-  c(0.42558,-0.23854,-0.63659,0.35753))
+    c(0.21889,-0.25701,-0.19529,-0.73412),
+    c(-0.63168,0.27992,-0.42895,-0.17223))
 
@@
 
-axes_obj <- gen_axes(
-  data = scaled_gau_data, 
-  proj_data = projected_df, 
-  limits = 1)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
-
-five_gau_proj_pacmap_model <- projected_df |>
+# axes_obj <- gen_axes(
+#   data = scaled_gau_data, 
+#   proj_data = projected_df, 
+#   limits = 1)
+# 
+# axes <- axes_obj$axes
+# circle <- axes_obj$circle
+
+limits <- 0.3
+projected_df_n <- projected_df |> select(-ID)
+
+rng <- range(projected_df_n)
+projected_df_n <- projected_df_n/max(abs(rng))
+colnames(projected_df_n) <- c("P1", "P2")
+projected_df_n <- data.frame(projected_df_n)
+obs_labels <- as.character(1:nrow(scaled_gau_data))
+
+axis_scale <- limits/6
+# axis_pos <- -2/3 * limits
+axis_pos <- 0
+
+adj <- function(x) axis_pos + x * axis_scale
+axes <- data.frame(x1 = adj(0), 
+                   y1 = adj(0), 
+                   x2 = adj(projection[, 1]), 
+                   y2 = adj(projection[, 2]))
+
+axis_labels <- colnames(scaled_gau_data)
+rownames(axes) <- axis_labels
+
+theta <- seq(0, 2 * pi, length = 50)
+circle <- data.frame(c1 = adj(cos(theta)), 
+                     c2 = adj(sin(theta)))
+
+five_gau_proj_pacmap_model1 <- projected_df |>
   ggplot(
@@
   geom_point(
+    size = 0.5,
     alpha = 0.3) +
@@
   coord_fixed() +
-  xlim(c(-0.18, 0.1)) +
-  ylim(c(-0.78, -0.55)) +
-  interior_annotation("c", 
+  xlim(c(-0.05, 0.35)) +
+  ylim(c(-0.05, 0.35)) +
+  interior_annotation("e", 
                       position = c(0.08, 0.9),
@@
 
-# five_gau_proj_pacmap_model <- gen_proj_langevitour(
-#   points_df = df_exe,
-#   projection = projection,
-#   edge_df = distance_df_small_edges |> select(-distance)
-# ) +
-#   xlim(c(-0.18, 0.1)) +
-#   ylim(c(-0.78, -0.55))
+### Second projection
+projection <- cbind(
+    c(-0.60995,0.33361,0.32920,-0.31518),
+    c(0.19525,-0.30554,0.74417,0.07600))
+
+projected <- as.matrix(scaled_gau_data) %*% projection
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  #dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+projected_model <- as.matrix(scaled_gau_data_model) %*% projection
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+# axes_obj <- gen_axes(
+#   data = scaled_gau_data, 
+#   proj_data = projected_df, 
+#   limits = 1)
+# 
+# axes <- axes_obj$axes
+# circle <- axes_obj$circle
+
+limits <- 0.3
+projected_df_n <- projected_df |> select(-ID)
+
+rng <- range(projected_df_n)
+projected_df_n <- projected_df_n/max(abs(rng))
+colnames(projected_df_n) <- c("P1", "P2")
+projected_df_n <- data.frame(projected_df_n)
+obs_labels <- as.character(1:nrow(scaled_gau_data))
+
+axis_scale <- limits/6
+# axis_pos <- -2/3 * limits
+axis_pos_x <- -0.05
+axis_pos_y <- -0.4
+
+adj <- function(x, axis_pos) axis_pos + x * axis_scale
+axes <- data.frame(x1 = adj(0, axis_pos_x), 
+                   y1 = adj(0, axis_pos_y), 
+                   x2 = adj(projection[, 1], axis_pos_x), 
+                   y2 = adj(projection[, 2], axis_pos_y))
+
+axis_labels <- colnames(scaled_gau_data)
+rownames(axes) <- axis_labels
+
+theta <- seq(0, 2 * pi, length = 50)
+circle <- data.frame(c1 = adj(cos(theta), axis_pos_x), 
+                     c2 = adj(sin(theta), axis_pos_y))
+
+five_gau_proj_pacmap_model2 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#33a02c") +
+  geom_point(
+    size = 0.8,
+    alpha = 0.3) +
+  scale_color_manual(values = c("#000000", "#33a02c")) +
+  geom_segment(
+    data=axes, 
+    aes(x=x1, y=y1, xend=x2, yend=y2), 
+    colour="grey70") +
+  geom_text(
+    data=axes, 
+    aes(x=x2, y=y2, label=rownames(axes)), 
+    colour="grey50",
+    size = 3) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.1, 0.2)) +
+  ylim(c(-0.45, 0)) +
+  interior_annotation("f", 
+                      position = c(0.08, 0.9),
+                      cex = 3)
+
 ```
@@
   five_gau_proj_umap_model1 + five_gau_proj_umap_model2 +
-  five_gau_proj_pacmap_model + five_gau_proj_pacmap_model +
+  five_gau_proj_pacmap_model1 + five_gau_proj_pacmap_model2 +
   plot_layout(guides = "collect", ncol = 2) &

### 2024-08-25 paper.qmd
Removed long edges in high-d for tSNE with Gau

@@
 
+## First projection
 projection <- cbind(
@@
 
-five_gau_proj_tsne_model <- projected_df |>
+five_gau_proj_tsne_model1 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#33a02c") +
+  geom_point(
+    alpha = 0.3) +
+  scale_color_manual(values = c("#000000", "#33a02c")) +
+  geom_segment(
+    data=axes, 
+    aes(x=x1, y=y1, xend=x2, yend=y2), 
+    colour="grey70") +
+  geom_text(
+    data=axes, 
+    aes(x=x2, y=y2, label=rownames(axes)), 
+    colour="grey50",
+    size = 3) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.18, 0.1)) +
+  ylim(c(-0.78, -0.55)) +
+  interior_annotation("a", 
+                      position = c(0.08, 0.9),
+                      cex = 3)
+
+## Second projection
+projection <- cbind(
+  c(-0.00215,-0.68905,-0.04778,-0.54223),
+  c(0.42558,-0.23854,-0.63659,0.35753))
+
+projected <- as.matrix(scaled_gau_data) %*% projection
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  #dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+projected_model <- as.matrix(scaled_gau_data_model) %*% projection
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+axes_obj <- gen_axes(
+  data = scaled_gau_data, 
+  proj_data = projected_df, 
+  limits = 1)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
+
+five_gau_proj_tsne_model2 <- projected_df |>
   ggplot(
@@
 
-five_gau_proj_tsne_model + five_gau_proj_tsne_model +
+five_gau_proj_tsne_model1 + five_gau_proj_tsne_model2 +
   five_gau_proj_umap_model1 + five_gau_proj_umap_model2 +

### 2024-08-25 paper.qmd
fig 13 caption and layout

--- 2024-08-25 paper.qmd
@@
   #            color = "black", linewidth=0.8, alpha = 0.5) +
-  scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 3, 5, 7, 9, 11, 13, 15, 17, 19, 21, 24, 26)]) +
+  scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 13, 17, 21, 26)]) +
   scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a','#ff7f00',
@@
   ylab("log(MSE)") +
-  xlab(expression(a[1])) +
+  xlab(expression(paste("binwidth (", a[1], ")"))) +
   theme_minimal() +
@@
 ```{r}
-#| fig-cap: "MSE from UMAP, tSNE, PHATE, PaCMAP, and TriMAP wwith diffferent (hyper-)parameter choices applied to PBMC3k dataset for different number of bin widths along the x-axis ($a_1$).
-Which is the most reasonable representation of the structures present in the PBMC3k data?
-The residual plot have a steep slope at the beginning, indicating that a smaller $a_1$ causes a smaller amount of error.
-Then, the slope gradually increases or level up, indicating that a higher number of $a_1$ generates a larger error.
-It was observed that for every $a_1$, the lowest error occurred with tSNE using the perplexity: $30$."
+#| fig-cap: "Assessing which of the 8 NLDR layouts on the PBMC3k data  (shown in Figure 1) is the better representation using MSE for varying binwidth ($a_1$).
+Colour  used for the lines and points in the top plot and in the scatterplots represents NLDR layout (a-h).
+Layout f is universally poor.
+Layouts a, b, g, h that show large separations between clusters are universally suboptimal.
+Layout d with little separation performs well at tiny binwidth (where most points are in their own bin) and poorly as binwidth increases.
+The choice of best is between layouts c and e, that have small separations between oddly shaped clusters.
+Layout e is the best choice."
 #| label: fig-pbmc-mse
@@
 
-free(error_plot_pbmc) / wrap_plots(nldr1c, nldr2c, nldr3c, nldr4c,
-           nldr5c, nldr6c, nldr7c, nldr8c, ncol = 4)
+free(error_plot_pbmc) + wrap_plots(nldr1c, nldr2c, nldr3c, nldr4c,
+           nldr5c, nldr6c, nldr7c, nldr8c, ncol = 2)
 ```

### 2024-08-25 paper.tex
Updated Figure 14
fig 13 caption and layout

@@
 
-\caption{\label{fig-pbmc-mse}MSE from UMAP, tSNE, PHATE, PaCMAP, and
-TriMAP wwith diffferent (hyper-)parameter choices applied to PBMC3k
-dataset for different number of bin widths along the x-axis (\(a_1\)).
-Which is the most reasonable representation of the structures present in
-the PBMC3k data? The residual plot have a steep slope at the beginning,
-indicating that a smaller \(a_1\) causes a smaller amount of error.
-Then, the slope gradually increases or level up, indicating that a
-higher number of \(a_1\) generates a larger error. It was observed that
-for every \(a_1\), the lowest error occurred with tSNE using the
-perplexity: \(30\).}
+\caption{\label{fig-pbmc-mse}Assessing which of the 8 NLDR layouts on
+the PBMC3k data (shown in Figure 1) is the better representation using
+MSE for varying binwidth (\(a_1\)). Colour used for the lines and points
+in the top plot and in the scatterplots represents NLDR layout (a-h).
+Layout f is universally poor. Layouts a, b, g, h that show large
+separations between clusters are universally suboptimal. Layout d with
+little separation performs well at tiny binwidth (where most points are
+in their own bin) and poorly as binwidth increases. The choice of best
+is between layouts c and e, that have small separations between oddly
+shaped clusters. Layout e is the best choice.}
 
@@
 
-\includegraphics[width=1\textwidth,height=1\textheight]{paper_files/figure-pdf/fig-pbmc2-sc-1.pdf}
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-pbmc2-sc-1.pdf}
 

### 2024-08-25 paper.qmd
Added tSNE proj for Gau

--- 2024-08-25 paper.qmd
@@
 
-gen_axes <- function(data, proj_data, limits = 1) {
+gen_axes <- function(data, proj_data, limits = 1, axis_pos_x = NULL, axis_pos_y = NULL) {
 
@@
   axis_scale <- limits/6
-  axis_pos <- -2/3 * limits
-
-  adj <- function(x) axis_pos + x * axis_scale
-  axes <- data.frame(x1 = adj(0), 
-                     y1 = adj(0), 
-                     x2 = adj(projection[, 1]), 
-                     y2 = adj(projection[, 2]))
+
+  if (is.null(axis_pos_x)) {
+
+    axis_pos_x <- -2/3 * limits
+
+  }
+
+  if (is.null(axis_pos_y)) {
+
+    axis_pos_y <- -2/3 * limits
+
+  }
+
+  adj <- function(x, axis_pos) axis_pos + x * axis_scale
+  axes <- data.frame(x1 = adj(0, axis_pos_x), 
+                     y1 = adj(0, axis_pos_y), 
+                     x2 = adj(projection[, 1], axis_pos_x), 
+                     y2 = adj(projection[, 2], axis_pos_y))
 
@@
   theta <- seq(0, 2 * pi, length = 50)
-  circle <- data.frame(c1 = adj(cos(theta)), 
-                       c2 = adj(sin(theta)))
+  circle <- data.frame(c1 = adj(cos(theta), axis_pos_x), 
+                       c2 = adj(sin(theta), axis_pos_y))
 
@@
 tr_from_to_df_gau1 <- tr_from_to_df_gau1 |>
-  filter(row_number() != 115) |>
-  filter(row_number() != 139)
+  filter(row_number() != 149)
 
@@
 
+benchmark_gau1 <- 0.1
+
 trimesh_removed_gau1 <- vis_rmlg_mesh(
@@
 
+distance_df_small_edges <- distance_df_small_edges |>
+  filter(row_number() != 108)
+
+## First projection
 projection <- cbind(
-  c(-0.00215,-0.68905,-0.04778,-0.54223),
-  c(0.42558,-0.23854,-0.63659,0.35753))
+    c(0.46477,-0.02024,0.56378,0.39736),
+    c(0.15607,-0.54876,-0.44090,0.41505))
 
@@
   proj_data = projected_df, 
-  limits = 1)
+  limits = 0.3,
+  axis_pos_x = -0.4,
+  axis_pos_y = -0.1)
 
@@
 
-five_gau_proj_tsne_model <- projected_df |>
+five_gau_proj_tsne_model1 <- projected_df |>
   ggplot(
@@
   coord_fixed() +
-  xlim(c(-0.18, 0.1)) +
-  ylim(c(-0.78, -0.55)) +
+  xlim(c(-0.45, -0.2)) +
+  ylim(c(-0.15, 0.18)) +
   interior_annotation("a", 
+                      position = c(0.08, 0.9),
+                      cex = 3)
+
+## Second projection
+projection <- cbind(
+    c(-0.71895,0.29250,-0.29905,-0.01656),
+    c(-0.16698,-0.62707,-0.23756,0.46328))
+
+projected <- as.matrix(scaled_gau_data) %*% projection
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  #dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+projected_model <- as.matrix(scaled_gau_data_model) %*% projection
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+axes_obj <- gen_axes(
+  data = scaled_gau_data, 
+  proj_data = projected_df, 
+  limits = 0.3,
+  axis_pos_x = 0.1,
+  axis_pos_y = -0.1)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
+
+five_gau_proj_tsne_model2 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#33a02c") +
+  geom_point(
+    alpha = 0.3) +
+  scale_color_manual(values = c("#000000", "#33a02c")) +
+  geom_segment(
+    data=axes, 
+    aes(x=x1, y=y1, xend=x2, yend=y2), 
+    colour="grey70") +
+  geom_text(
+    data=axes, 
+    aes(x=x2, y=y2, label=rownames(axes)), 
+    colour="grey50",
+    size = 3) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(0.05, 0.25)) +
+  ylim(c(-0.15, 0.25)) +
+  interior_annotation("b", 
                       position = c(0.08, 0.9),

### 2024-08-25 paper.qmd
Updated UMAP proj

@@
 
-# axes_obj <- gen_axes(
-#   data = scaled_gau_data, 
-#   proj_data = projected_df, 
-#   limits = 1)
-# 
-# axes <- axes_obj$axes
-# circle <- axes_obj$circle
-
-limits <- 0.5
-projected_df_n <- projected_df |> select(-ID)
-
-rng <- range(projected_df_n)
-projected_df_n <- projected_df_n/max(abs(rng))
-colnames(projected_df_n) <- c("P1", "P2")
-projected_df_n <- data.frame(projected_df_n)
-obs_labels <- as.character(1:nrow(scaled_gau_data))
-
-axis_scale <- limits/6
-# axis_pos <- -2/3 * limits
-axis_pos <- -0.8
-
-adj <- function(x) axis_pos + x * axis_scale
-axes <- data.frame(x1 = adj(0), 
-                   y1 = adj(0), 
-                   x2 = adj(projection[, 1]), 
-                   y2 = adj(projection[, 2]))
-
-axis_labels <- colnames(scaled_gau_data)
-rownames(axes) <- axis_labels
-
-theta <- seq(0, 2 * pi, length = 50)
-circle <- data.frame(c1 = adj(cos(theta)), 
-                     c2 = adj(sin(theta)))
+axes_obj <- gen_axes(
+  data = scaled_gau_data,
+  proj_data = projected_df,
+  limits = 0.3,
+  axis_pos_x = -0.2,
+  axis_pos_y = -0.85)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
 
@@
   geom_point(
-    size = 0.5,
     alpha = 0.2) +
@@
   coord_fixed() +
-  xlim(c(-0.9, 0.1)) +
-  ylim(c(-0.9, -0.5)) +
+  xlim(c(-0.25, 0.1)) +
+  ylim(c(-0.9, -0.55)) +
   interior_annotation("c", 
@@
 
-# axes_obj <- gen_axes(
-#   data = scaled_gau_data, 
-#   proj_data = projected_df, 
-#   limits = 1)
-# 
-# axes <- axes_obj$axes
-# circle <- axes_obj$circle
-
-limits <- 0.3
-projected_df_n <- projected_df |> select(-ID)
-
-rng <- range(projected_df_n)
-projected_df_n <- projected_df_n/max(abs(rng))
-colnames(projected_df_n) <- c("P1", "P2")
-projected_df_n <- data.frame(projected_df_n)
-obs_labels <- as.character(1:nrow(scaled_gau_data))
-
-axis_scale <- limits/6
-# axis_pos <- -2/3 * limits
-axis_pos <- 0.08
-
-adj <- function(x) axis_pos + x * axis_scale
-axes <- data.frame(x1 = adj(0), 
-                   y1 = adj(0), 
-                   x2 = adj(projection[, 1]), 
-                   y2 = adj(projection[, 2]))
-
-axis_labels <- colnames(scaled_gau_data)
-rownames(axes) <- axis_labels
-
-theta <- seq(0, 2 * pi, length = 50)
-circle <- data.frame(c1 = adj(cos(theta)), 
-                     c2 = adj(sin(theta)))
+axes_obj <- gen_axes(
+  data = scaled_gau_data,
+  proj_data = projected_df,
+  limits = 0.3,
+  axis_pos_x = 0.08,
+  axis_pos_y = -0.08)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
 
@@
   geom_point(
-    size = 0.5, 
     alpha = 0.3) +
@@
   xlim(c(0, 0.4)) +
-  ylim(c(0, 0.3)) +
+  ylim(c(-0.15, 0.3)) +
   interior_annotation("d", 

### 2024-08-26 paper.qmd
Updated Figure 11

@@
   geom_point(
-    alpha = 0.2) +
+    size = 0.5,
+    alpha = 0.3) +
   scale_color_manual(values = c("#000000", "#33a02c")) +
@@
 
-# axes_obj <- gen_axes(
-#   data = scaled_gau_data, 
-#   proj_data = projected_df, 
-#   limits = 1)
-# 
-# axes <- axes_obj$axes
-# circle <- axes_obj$circle
-
-limits <- 0.3
-projected_df_n <- projected_df |> select(-ID)
-
-rng <- range(projected_df_n)
-projected_df_n <- projected_df_n/max(abs(rng))
-colnames(projected_df_n) <- c("P1", "P2")
-projected_df_n <- data.frame(projected_df_n)
-obs_labels <- as.character(1:nrow(scaled_gau_data))
-
-axis_scale <- limits/6
-# axis_pos <- -2/3 * limits
-axis_pos <- 0
-
-adj <- function(x) axis_pos + x * axis_scale
-axes <- data.frame(x1 = adj(0), 
-                   y1 = adj(0), 
-                   x2 = adj(projection[, 1]), 
-                   y2 = adj(projection[, 2]))
-
-axis_labels <- colnames(scaled_gau_data)
-rownames(axes) <- axis_labels
-
-theta <- seq(0, 2 * pi, length = 50)
-circle <- data.frame(c1 = adj(cos(theta)), 
-                     c2 = adj(sin(theta)))
+axes_obj <- gen_axes(
+  data = scaled_gau_data,
+  proj_data = projected_df,
+  limits = 0.3,
+  axis_pos_x = 0.1,
+  axis_pos_y = 0.05)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
 
@@
   coord_fixed() +
-  xlim(c(-0.05, 0.35)) +
-  ylim(c(-0.05, 0.35)) +
+  xlim(c(0.05, 0.35)) +
+  ylim(c(0, 0.35)) +
   interior_annotation("e", 
@@
 
-# axes_obj <- gen_axes(
-#   data = scaled_gau_data, 
-#   proj_data = projected_df, 
-#   limits = 1)
-# 
-# axes <- axes_obj$axes
-# circle <- axes_obj$circle
-
-limits <- 0.3
-projected_df_n <- projected_df |> select(-ID)
-
-rng <- range(projected_df_n)
-projected_df_n <- projected_df_n/max(abs(rng))
-colnames(projected_df_n) <- c("P1", "P2")
-projected_df_n <- data.frame(projected_df_n)
-obs_labels <- as.character(1:nrow(scaled_gau_data))
-
-axis_scale <- limits/6
-# axis_pos <- -2/3 * limits
-axis_pos_x <- -0.05
-axis_pos_y <- -0.4
-
-adj <- function(x, axis_pos) axis_pos + x * axis_scale
-axes <- data.frame(x1 = adj(0, axis_pos_x), 
-                   y1 = adj(0, axis_pos_y), 
-                   x2 = adj(projection[, 1], axis_pos_x), 
-                   y2 = adj(projection[, 2], axis_pos_y))
-
-axis_labels <- colnames(scaled_gau_data)
-rownames(axes) <- axis_labels
-
-theta <- seq(0, 2 * pi, length = 50)
-circle <- data.frame(c1 = adj(cos(theta), axis_pos_x), 
-                     c2 = adj(sin(theta), axis_pos_y))
+axes_obj <- gen_axes(
+  data = scaled_gau_data,
+  proj_data = projected_df,
+  limits = 0.3,
+  axis_pos_x = -0.05,
+  axis_pos_y = -0.4)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
 

### 2024-08-26 paper.qmd
Updated Figure 15 Caption

@@
 
-The MNIST dataset consists of grayscale images of handwritten digits [@lecun2010].
-@yingfan2021 used this dataset to demonstrate how PaCMAP preserves non-linear structure in \pD{}.
+The MNIST dataset consists of $70000$ grayscale images of handwritten digits [@lecun2010].
+@yingfan2021 used this dataset to evaluate how PaCMAP preserves local structures in \gD{}.
+Because of the large number of images, we selected the handwritten digit 1 which has a non-linear structure in \gD{}.
+There are $7877$ images.
+Also, applied PCA as a preprocessing step and selected first $10$ principal components.
+The objective is to evaluate whether PaCMAP preserves the non-linear structure in \gD{}.
+
+<!-- PaCMAP preserves non-linear structure in $784\text{-}D$.
 To evaluate whether PaCMAP provides a reasonable representation of the data, the \gD{} embedding of the handwritten digit 1 was selected.
 As shown in @fig-pacmap-author, the angle of the digit 1 images varies along the \gD{} structure.
+-->
 
@@
 #| fig-cap: "$2\\text{-}D$ layout from PaCMAP applied for the digit 1 of the MNIST dataset.
-The (hyper-)parameter settings, beyond the defaults, are 10 nearest neighbors, ratio of the number of mid-near pairs to the number of neighbors equal to 0.5, and the ratio of the number of further pairs to the number of neighbors is 2.
-We use our model-in-the-data-space to assess whether this is an accurate representation of structure present in the high-dimensional data, or if it is misleading.
+The (hyper-)parameter settings, beyond the defaults, are $10$ nearest neighbors, ratio of the number of mid-near pairs to the number of neighbors equal to $0.5$, and the ratio of the number of further pairs to the number of neighbors is $2$.
+We use our *model-in-the-data-space* to assess whether this is an accurate representation of structure present in $10\\text{-}D$, or if it is misleading.
 The angle of the digit 1 varies along this structure.
@@
 According to @fig-mnist1-sc1, the non-linear continuous structure observed in the \gD{} representation of PaCMAP (@fig-pacmap-author) is also visible when visualizing the model overlaid on the data space.
-This indicates that PaCMAP accurately captures the structure of the \pD{} data.
-Additionally, the model shows a twisted pattern within the non-linear structure in \pD{} space (@fig-mnist1-sc2), which is an additional pattern not visible in the \gD{} representation (@fig-pacmap-author).
-Furthermore, as shown in @fig-mnist1-sc3, some long edges exist in the \pD{} space that are not recognized as long edges in the \gD{} representation.
-However, PaCMAP is a reasonable \gD{} representation of MNIST digit 1 data, because PaCMAP preserves the non-linear structure present in the \pD{} data.
+This indicates that PaCMAP accurately captures the structure of the $10\text{-}D$ data.
+Additionally, the model shows a twisted pattern within the non-linear structure in $10\text{-}D$ space (@fig-mnist1-sc2), which is an additional pattern not visible in the \gD{} representation (@fig-pacmap-author).
+Furthermore, as shown in @fig-mnist1-sc3, some long edges exist in the $10\text{-}D$ space that are not recognized as long edges in the \gD{} representation.
+However, PaCMAP is a reasonable \gD{} representation of MNIST digit 1 data, because PaCMAP preserves the non-linear structure present in the $10\text{-}D$ data.
 
@@
 
-There are certain data points that exhibit high error rates due to their deviation from the usual \pD{} data structure, which makes them anomalies (@fig-model-mnist b).
+There are certain data points that exhibit high error rates due to their deviation from the usual $10\text{-}D$ data structure, which makes them anomalies (@fig-model-mnist b).
 These anomalies can be classified into two types: those that are anomalies within the non-linear structure and those that lie outside of it.

### 2024-08-26 paper.tex
Updated UMAP proj
Updated Figure 15 Caption

@@
 
-The MNIST dataset consists of grayscale images of handwritten digits
-\citep{lecun2010}. \citet{yingfan2021} used this dataset to demonstrate
-how PaCMAP preserves non-linear structure in \pD{}. To evaluate whether
-PaCMAP provides a reasonable representation of the data, the \gD{}
-embedding of the handwritten digit 1 was selected. As shown in
-Figure~\ref{fig-pacmap-author}, the angle of the digit 1 images varies
-along the \gD{} structure.
+The MNIST dataset consists of \(70000\) grayscale images of handwritten
+digits \citep{lecun2010}. \citet{yingfan2021} used this dataset to
+evaluate how PaCMAP preserves local structures in \gD{}. Because of the
+large number of images, we selected the handwritten digit 1 which has a
+non-linear structure in \gD{}. There are \(7877\) images. Also, applied
+PCA as a preprocessing step and selected first \(10\) principal
+components. The objective is to evaluate whether PaCMAP preserves the
+non-linear structure in \gD{}.
 
@@
 applied for the digit 1 of the MNIST dataset. The (hyper-)parameter
-settings, beyond the defaults, are 10 nearest neighbors, ratio of the
-number of mid-near pairs to the number of neighbors equal to 0.5, and
-the ratio of the number of further pairs to the number of neighbors is
-2. We use our model-in-the-data-space to assess whether this is an
-accurate representation of structure present in the high-dimensional
-data, or if it is misleading. The angle of the digit 1 varies along this
-structure. Images at the top-left of the \(2\text{-}D\) layout show the
-digit 1 angled more to the left, while images at the bottom-right show
-the digit 1 angled more to the right.}
+settings, beyond the defaults, are \(10\) nearest neighbors, ratio of
+the number of mid-near pairs to the number of neighbors equal to
+\(0.5\), and the ratio of the number of further pairs to the number of
+neighbors is \(2\). We use our \emph{model-in-the-data-space} to assess
+whether this is an accurate representation of structure present in
+\(10\text{-}D\), or if it is misleading. The angle of the digit 1 varies
+along this structure. Images at the top-left of the \(2\text{-}D\)
+layout show the digit 1 angled more to the left, while images at the
+bottom-right show the digit 1 angled more to the right.}
 
@@
 model overlaid on the data space. This indicates that PaCMAP accurately
-captures the structure of the \pD{} data. Additionally, the model shows
-a twisted pattern within the non-linear structure in \pD{} space
-(\textbf{?@fig-mnist1-sc2}), which is an additional pattern not visible
-in the \gD{} representation (Figure~\ref{fig-pacmap-author}).
-Furthermore, as shown in \textbf{?@fig-mnist1-sc3}, some long edges
-exist in the \pD{} space that are not recognized as long edges in the
-\gD{} representation. However, PaCMAP is a reasonable \gD{}
-representation of MNIST digit 1 data, because PaCMAP preserves the
-non-linear structure present in the \pD{} data.
+captures the structure of the \(10\text{-}D\) data. Additionally, the
+model shows a twisted pattern within the non-linear structure in
+\(10\text{-}D\) space (\textbf{?@fig-mnist1-sc2}), which is an
+additional pattern not visible in the \gD{} representation
+(Figure~\ref{fig-pacmap-author}). Furthermore, as shown in
+\textbf{?@fig-mnist1-sc3}, some long edges exist in the \(10\text{-}D\)
+space that are not recognized as long edges in the \gD{} representation.
+However, PaCMAP is a reasonable \gD{} representation of MNIST digit 1
+data, because PaCMAP preserves the non-linear structure present in the
+\(10\text{-}D\) data.
 
@@
 There are certain data points that exhibit high error rates due to their
-deviation from the usual \pD{} data structure, which makes them
-anomalies (Figure~\ref{fig-model-mnist} b). These anomalies can be
+deviation from the usual \(10\text{-}D\) data structure, which makes
+them anomalies (Figure~\ref{fig-model-mnist} b). These anomalies can be
 classified into two types: those that are anomalies within the

### 2024-08-26 paper.qmd
Updated MNIST intro text

@@
 @yingfan2021 used this dataset to evaluate how PaCMAP preserves local structures in \gD{}.
-Because of the large number of images, we selected the handwritten digit 1 which has a non-linear structure in \gD{}.
-There are $7877$ images.
-Also, applied PCA as a preprocessing step and selected first $10$ principal components.
-The objective is to evaluate whether PaCMAP preserves the non-linear structure in \gD{}.
+Because of the large number of images, the focus was specifically on the handwritten digit 1, which exhibits a non-linear structure in \gD{} (@fig-pacmap-author).
+There are $7877$ images of the digit 1 in the dataset.
+Additionally, PCA was applied as a preprocessing step and the first $10$ principal components were selected.
+The objective was to assess whether PaCMAP effectively preserves the non-linear structure in \gD{}.
 
@@
 We use our *model-in-the-data-space* to assess whether this is an accurate representation of structure present in $10\\text{-}D$, or if it is misleading.
-The angle of the digit 1 varies along this structure.
-Images at the top-left of the $2\\text{-}D$ layout show the digit 1 angled more to the left, while images at the bottom-right show the digit 1 angled more to the right."
+The angle of the digit 1 images varies along this structure.
+Images at the top-left of the $2\\text{-}D$ layout show the digit 1 angled more to the left, while those at the bottom-right show the digit 1 angled more to the right."
 #| label: fig-pacmap-author

### 2024-08-26 paper.tex
Updated MNIST intro text

@@
 evaluate how PaCMAP preserves local structures in \gD{}. Because of the
-large number of images, we selected the handwritten digit 1 which has a
-non-linear structure in \gD{}. There are \(7877\) images. Also, applied
-PCA as a preprocessing step and selected first \(10\) principal
-components. The objective is to evaluate whether PaCMAP preserves the
-non-linear structure in \gD{}.
+large number of images, the focus was specifically on the handwritten
+digit 1, which exhibits a non-linear structure in \gD{}
+(Figure~\ref{fig-pacmap-author}). There are \(7877\) images of the digit
+1 in the dataset. Additionally, PCA was applied as a preprocessing step
+and the first \(10\) principal components were selected. The objective
+was to assess whether PaCMAP effectively preserves the non-linear
+structure in \gD{}.
 
@@
 whether this is an accurate representation of structure present in
-\(10\text{-}D\), or if it is misleading. The angle of the digit 1 varies
-along this structure. Images at the top-left of the \(2\text{-}D\)
-layout show the digit 1 angled more to the left, while images at the
-bottom-right show the digit 1 angled more to the right.}
+\(10\text{-}D\), or if it is misleading. The angle of the digit 1 images
+varies along this structure. Images at the top-left of the
+\(2\text{-}D\) layout show the digit 1 angled more to the left, while
+those at the bottom-right show the digit 1 angled more to the right.}
 

### 2024-08-26 paper.qmd
Updated Figure 16 caption

@@
 Additionally, PCA was applied as a preprocessing step and the first $10$ principal components were selected.
-The objective was to assess whether PaCMAP effectively preserves the non-linear structure in \gD{}.
+The objective is to assess whether PaCMAP effectively preserves the non-linear structure in \gD{}.
 
@@
 The (hyper-)parameter settings, beyond the defaults, are $10$ nearest neighbors, ratio of the number of mid-near pairs to the number of neighbors equal to $0.5$, and the ratio of the number of further pairs to the number of neighbors is $2$.
-We use our *model-in-the-data-space* to assess whether this is an accurate representation of structure present in $10\\text{-}D$, or if it is misleading.
+We use our *model-in-the-data-space* to assess whether this is an accurate representation of non-linear structure present in $10\\text{-}D$, or if it is misleading.
 The angle of the digit 1 images varies along this structure.
@@
 <!--bin1 = 22, bin2 = 17, b = 374, non_empty = 140-->
-```{r}
-#| echo: false
-#| fig-cap: "$10\\text{-}D$ model error in $2\\text{-}D$.
-The $2\\text{-}D$ model shows a non-linear continuous structure.
-Most low $10\\text{-}D$ model errors are distributed along the lower edge of the $2\\text{-}D$ structure, while most large $10\\text{-}D$ model errors are concentrated along the upper edge."
-#| label: fig-model-mnist
-#| fig-pos: H
-#| out-height: 22%
-
-error_plot_mnist + 
-  plot_layout(guides='collect') &
-  theme(legend.position='bottom')
-```
-
-According to @fig-mnist1-sc1, the non-linear continuous structure observed in the \gD{} representation of PaCMAP (@fig-pacmap-author) is also visible when visualizing the model overlaid on the data space.
-This indicates that PaCMAP accurately captures the structure of the $10\text{-}D$ data.
-Additionally, the model shows a twisted pattern within the non-linear structure in $10\text{-}D$ space (@fig-mnist1-sc2), which is an additional pattern not visible in the \gD{} representation (@fig-pacmap-author).
-Furthermore, as shown in @fig-mnist1-sc3, some long edges exist in the $10\text{-}D$ space that are not recognized as long edges in the \gD{} representation.
-However, PaCMAP is a reasonable \gD{} representation of MNIST digit 1 data, because PaCMAP preserves the non-linear structure present in the $10\text{-}D$ data.
+According to @fig-mnist-tri-proj (b), the non-linear structure observed in the \gD{} layout of PaCMAP (@fig-pacmap-author) is also visible when visualizing the model overlaid on the data space.
+This indicates that PaCMAP accurately captures the non-linear structure of the $10\text{-}D$ data.
+Additionally, the model shows a twisted pattern within the non-linear structure in $10\text{-}D$ space (@fig-mnist-tri-proj (c)), which is an additional pattern not visible in the \gD{} representation (@fig-pacmap-author).
+Furthermore, as shown in @fig-mnist-tri-proj (d), some long edges exist in the $10\text{-}D$ space that are not recognized as long edges in the \gD{} representation.
+However, PaCMAP provides a reasonable \gD{} representation of MNIST digit 1 data, as it preserves the non-linear structure present in the $10\text{-}D$ data.
 
@@
 #| fig-width: 10
-#| fig-cap: "Model in \\gD{}, on the PaCMAP layout, and three views of the fit in projections from $10\\text{-}D$, for the digit 1 of MNIST data ($(s_1, \ s_2) = (-0.100, \ -0.059)$, $b = 374 \  (22, \ 17)$, $m = 140$, benchmark value to remove large edges is $0.094$).
+#| fig-cap: "Model in \\gD{}, on the PaCMAP layout, and three views of the fit in projections from $10\\text{-}D$, for the digit 1 of MNIST data ($a_1 = 0.062$, $b = 285 \\  (19, \ 15)$, $m = 109$, and benchmark value to remove large edges is $0.108$).
 (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/zcg_GXBmqjA>)."
@@
 
-There are certain data points that exhibit high error rates due to their deviation from the usual $10\text{-}D$ data structure, which makes them anomalies (@fig-model-mnist b).
+```{r}
+#| echo: false
+#| fig-cap: "$10\\text{-}D$ model error in $2\\text{-}D$.
+The $2\\text{-}D$ model shows a non-linear continuous structure.
+Most low $10\\text{-}D$ model errors are distributed along the lower edge of the $2\\text{-}D$ structure, while most large $10\\text{-}D$ model errors are concentrated along the upper edge."
+#| label: fig-model-mnist
+#| fig-pos: H
+#| out-height: 22%
+
+error_plot_mnist + 
+  plot_layout(guides='collect') &
+  theme(legend.position='bottom')
+```
+
+There are certain data points that exhibit high error rates due to their deviation from the usual $10\text{-}D$ data structure, which makes them anomalies (@fig-model-mnist).
 These anomalies can be classified into two types: those that are anomalies within the non-linear structure and those that lie outside of it.

### 2024-08-26 paper.tex
Updated Figure 16 caption

@@
 and the first \(10\) principal components were selected. The objective
-was to assess whether PaCMAP effectively preserves the non-linear
+is to assess whether PaCMAP effectively preserves the non-linear
 structure in \gD{}.
@@
 neighbors is \(2\). We use our \emph{model-in-the-data-space} to assess
-whether this is an accurate representation of structure present in
-\(10\text{-}D\), or if it is misleading. The angle of the digit 1 images
-varies along this structure. Images at the top-left of the
-\(2\text{-}D\) layout show the digit 1 angled more to the left, while
-those at the bottom-right show the digit 1 angled more to the right.}
+whether this is an accurate representation of non-linear structure
+present in \(10\text{-}D\), or if it is misleading. The angle of the
+digit 1 images varies along this structure. Images at the top-left of
+the \(2\text{-}D\) layout show the digit 1 angled more to the left,
+while those at the bottom-right show the digit 1 angled more to the
+right.}
+
+\end{figure}%
+
+According to Figure~\ref{fig-mnist-tri-proj} (b), the non-linear
+structure observed in the \gD{} layout of PaCMAP
+(Figure~\ref{fig-pacmap-author}) is also visible when visualizing the
+model overlaid on the data space. This indicates that PaCMAP accurately
+captures the non-linear structure of the \(10\text{-}D\) data.
+Additionally, the model shows a twisted pattern within the non-linear
+structure in \(10\text{-}D\) space (Figure~\ref{fig-mnist-tri-proj}
+(c)), which is an additional pattern not visible in the \gD{}
+representation (Figure~\ref{fig-pacmap-author}). Furthermore, as shown
+in Figure~\ref{fig-mnist-tri-proj} (d), some long edges exist in the
+\(10\text{-}D\) space that are not recognized as long edges in the \gD{}
+representation. However, PaCMAP provides a reasonable \gD{}
+representation of MNIST digit 1 data, as it preserves the non-linear
+structure present in the \(10\text{-}D\) data.
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-mnist-tri-proj-1.pdf}
+
+}
+
+\caption{\label{fig-mnist-tri-proj}Model in \gD{}, on the PaCMAP layout,
+and three views of the fit in projections from \(10\text{-}D\), for the
+digit 1 of MNIST data (\(a_1 = 0.062\), \(b = 285 \  (19,  15)\),
+\(m = 109\), and benchmark value to remove large edges is \(0.108\)).
+(The \textbf{langevitour} software is used to view the data with a tour,
+and the full video is available at \url{https://youtu.be/zcg_GXBmqjA}).}
 
@@
 
-According to \textbf{?@fig-mnist1-sc1}, the non-linear continuous
-structure observed in the \gD{} representation of PaCMAP
-(Figure~\ref{fig-pacmap-author}) is also visible when visualizing the
-model overlaid on the data space. This indicates that PaCMAP accurately
-captures the structure of the \(10\text{-}D\) data. Additionally, the
-model shows a twisted pattern within the non-linear structure in
-\(10\text{-}D\) space (\textbf{?@fig-mnist1-sc2}), which is an
-additional pattern not visible in the \gD{} representation
-(Figure~\ref{fig-pacmap-author}). Furthermore, as shown in
-\textbf{?@fig-mnist1-sc3}, some long edges exist in the \(10\text{-}D\)
-space that are not recognized as long edges in the \gD{} representation.
-However, PaCMAP is a reasonable \gD{} representation of MNIST digit 1
-data, because PaCMAP preserves the non-linear structure present in the
-\(10\text{-}D\) data.
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-mnist-tri-proj-1.pdf}
-
-}
-
-\caption{\label{fig-mnist-tri-proj}Model in \gD{}, on the PaCMAP layout,
-and three views of the fit in projections from \(10\text{-}D\), for the
-digit 1 of MNIST data (\((s_1,  s_2) = (-0.100,  -0.059)\),
-\(b = 374   (22,  17)\), \(m = 140\), benchmark value to remove large
-edges is \(0.094\)). (The \textbf{langevitour} software is used to view
-the data with a tour, and the full video is available at
-\url{https://youtu.be/zcg_GXBmqjA}).}
-
-\end{figure}%
-
 There are certain data points that exhibit high error rates due to their
 deviation from the usual \(10\text{-}D\) data structure, which makes
-them anomalies (Figure~\ref{fig-model-mnist} b). These anomalies can be
+them anomalies (Figure~\ref{fig-model-mnist}). These anomalies can be
 classified into two types: those that are anomalies within the

### 2024-08-26 paper.qmd
Updated Fiure 17 title

@@
 
-<!--bin1 = 22, bin2 = 17, b = 374, non_empty = 140-->
-According to @fig-mnist-tri-proj (b), the non-linear structure observed in the \gD{} layout of PaCMAP (@fig-pacmap-author) is also visible when visualizing the model overlaid on the data space.
+PaCMAP provides a reasonable \gD{} representation of MNIST digit 1 data, as it preserves the non-linear structure present in the $10\text{-}D$ data (@fig-mnist-tri-proj).
+Furthermore, shows some quirks when visualizing the *model-in-the-data-space* (@fig-mnist-tri-proj).
+
+<!-- According to @fig-mnist-tri-proj (b), the non-linear structure observed in the \gD{} layout of PaCMAP (@fig-pacmap-author) is also visible when visualizing the model overlaid on the data space.
 This indicates that PaCMAP accurately captures the non-linear structure of the $10\text{-}D$ data.
@@
 However, PaCMAP provides a reasonable \gD{} representation of MNIST digit 1 data, as it preserves the non-linear structure present in the $10\text{-}D$ data.
-
+-->
+
+<!--bin1 = 22, bin2 = 17, b = 374, non_empty = 140-->
 ```{r}
@@
 #| fig-width: 10
-#| fig-cap: "Model in \\gD{}, on the PaCMAP layout, and three views of the fit in projections from $10\\text{-}D$, for the digit 1 of MNIST data ($a_1 = 0.062$, $b = 285 \\  (19, \ 15)$, $m = 109$, and benchmark value to remove large edges is $0.108$).
-(The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/zcg_GXBmqjA>)."
+#| fig-cap: "Model in \\gD{}, on the PaCMAP layout (a), and three views (b, c, d) of the fit in projections from $10\\text{-}D$, for the digit 1 of MNIST data ($a_1 = 0.062$, $b = 285 \\  (19, \ 15)$, $m = 109$, and benchmark value to remove large edges is $0.108$).
+(b) The non-linear structure observed in the \\gD{} layout of PaCMAP (@fig-pacmap-author) is also visible when visualizing the model overlaid on the data space.
+This indicates that PaCMAP accurately captures the non-linear structure of the $10\\text{-}D$ data.
+(c) the model shows a twisted pattern within the non-linear structure in $10\\text{-}D$ space, which is an additional pattern not visible in the \\gD{} representation (@fig-pacmap-author).
+(d) some long edges exist in the $10\\text{-}D$ space that are not recognized as long edges in the \\gD{} representation.
+The full video is available at <https://youtu.be/zcg_GXBmqjA>."
 

### 2024-08-26 paper.tex
Updated Fiure 17 title

@@
 
-According to Figure~\ref{fig-mnist-tri-proj} (b), the non-linear
-structure observed in the \gD{} layout of PaCMAP
-(Figure~\ref{fig-pacmap-author}) is also visible when visualizing the
-model overlaid on the data space. This indicates that PaCMAP accurately
-captures the non-linear structure of the \(10\text{-}D\) data.
-Additionally, the model shows a twisted pattern within the non-linear
-structure in \(10\text{-}D\) space (Figure~\ref{fig-mnist-tri-proj}
-(c)), which is an additional pattern not visible in the \gD{}
-representation (Figure~\ref{fig-pacmap-author}). Furthermore, as shown
-in Figure~\ref{fig-mnist-tri-proj} (d), some long edges exist in the
+PaCMAP provides a reasonable \gD{} representation of MNIST digit 1 data,
+as it preserves the non-linear structure present in the \(10\text{-}D\)
+data (Figure~\ref{fig-mnist-tri-proj}). Furthermore, shows some quirks
+when visualizing the \emph{model-in-the-data-space}
+(Figure~\ref{fig-mnist-tri-proj}).
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-mnist-tri-proj-1.pdf}
+
+}
+
+\caption{\label{fig-mnist-tri-proj}Model in \gD{}, on the PaCMAP layout
+(a), and three views (b, c, d) of the fit in projections from
+\(10\text{-}D\), for the digit 1 of MNIST data (\(a_1 = 0.062\),
+\(b = 285 \  (19,  15)\), \(m = 109\), and benchmark value to remove
+large edges is \(0.108\)). (b) The non-linear structure observed in the
+\gD{} layout of PaCMAP (Figure~\ref{fig-pacmap-author}) is also visible
+when visualizing the model overlaid on the data space. This indicates
+that PaCMAP accurately captures the non-linear structure of the
+\(10\text{-}D\) data. (c) the model shows a twisted pattern within the
+non-linear structure in \(10\text{-}D\) space, which is an additional
+pattern not visible in the \gD{} representation
+(Figure~\ref{fig-pacmap-author}). (d) some long edges exist in the
 \(10\text{-}D\) space that are not recognized as long edges in the \gD{}
-representation. However, PaCMAP provides a reasonable \gD{}
-representation of MNIST digit 1 data, as it preserves the non-linear
-structure present in the \(10\text{-}D\) data.
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-mnist-tri-proj-1.pdf}
-
-}
-
-\caption{\label{fig-mnist-tri-proj}Model in \gD{}, on the PaCMAP layout,
-and three views of the fit in projections from \(10\text{-}D\), for the
-digit 1 of MNIST data (\(a_1 = 0.062\), \(b = 285 \  (19,  15)\),
-\(m = 109\), and benchmark value to remove large edges is \(0.108\)).
-(The \textbf{langevitour} software is used to view the data with a tour,
-and the full video is available at \url{https://youtu.be/zcg_GXBmqjA}).}
+representation. The full video is available at
+\url{https://youtu.be/zcg_GXBmqjA}.}
 

### 2024-08-26 paper.qmd
Updated Figure 17

@@
   geom_point(alpha=0.5, 
-             size = 0.1) #+
-  #interior_annotation("b", sc_ltr_pos_mnist)
+             size = 0.1) +
+  scale_colour_brewer(
+    type = "qual",
+    palette = 2
+  ) + 
+    # Optional: Set a minimal theme
+    theme(
+        legend.position = "bottom",
+        legend.text = element_text(size = 10),          # Increase legend text size
+        legend.key.size = unit(1, "lines")              # Increase size of legend keys
+    ) +
+    guides(
+        color = guide_legend(
+            override.aes = list(size = 2.5, alpha = 1)      # Increase point size and opacity in legend
+        )
+    )
 ```
@@
 
-error_plot_mnist + 
-  plot_layout(guides='collect') &
-  theme(legend.position='bottom')
+error_plot_mnist 
 ```

### 2024-08-26 paper.qmd
Updated Figure 17 caption

--- 2024-08-26 paper.qmd
@@
   geom_point(alpha=0.5, 
-             size = 0.1) #+
-  #interior_annotation("b", sc_ltr_pos_mnist)
+             size = 0.1) +
+  scale_colour_brewer(
+    type = "qual",
+    palette = 2
+  ) + 
+  theme(
+      legend.position = "bottom",
+      legend.text = element_text(size = 10),          # Increase legend text size
+      legend.key.size = unit(1, "lines")              # Increase size of legend keys
+  ) +
+  guides(
+      color = guide_legend(
+          override.aes = list(size = 2.5, alpha = 1)      # Increase point size and opacity in legend
+      )
+  )
 ```
@@
 #| echo: false
-#| fig-cap: "$10\\text{-}D$ model error in $2\\text{-}D$.
-The $2\\text{-}D$ model shows a non-linear continuous structure.
-Most low $10\\text{-}D$ model errors are distributed along the lower edge of the $2\\text{-}D$ structure, while most large $10\\text{-}D$ model errors are concentrated along the upper edge."
+#| fig-cap: "$10\\text{-}D$ model error in PaCMAP layout of MNIST digit 1 dataset in $2\\text{-}D$.
+Most low $10\\text{-}D$ model errors are distributed along the lower edge of the $2\\text{-}D$ structure, while most large $10\\text{-}D$ model errors are concentrated along the upper edge (a).
+There are certain data points that exhibit high error rates due to their deviation from the usual $10\\text{-}D$ data structure, which makes them anomalies.
+These anomalies can be classified into two types: those that are anomalies within the non-linear structure and those that lie outside of it.
+The images associated with high model error points within the non-linear structure display different patterns of the digit 1 (b).
+However, when comparing these images to the ones found outside of the non-linear structure, it becomes evident that the latter display different patterns of the digit 1 (c).
+"
 #| label: fig-model-mnist
 #| fig-pos: H
-#| out-height: 22%
-
-error_plot_mnist + 
-  plot_layout(guides='collect') &
-  theme(legend.position='bottom')
-```
-
-There are certain data points that exhibit high error rates due to their deviation from the usual $10\text{-}D$ data structure, which makes them anomalies (@fig-model-mnist).
+
+error_plot_mnist 
+```
+
+<!-- There are certain data points that exhibit high error rates due to their deviation from the usual $10\text{-}D$ data structure, which makes them anomalies (@fig-model-mnist).
 These anomalies can be classified into two types: those that are anomalies within the non-linear structure and those that lie outside of it.
@@
 However, when comparing these images to the ones found outside of the non-linear structure, it becomes evident that the latter display different patterns of the digit 1 (@fig-mnist1-out).
+-->
 

### 2024-08-26 paper.tex
Updated Figure 17 caption

@@
 
-\includegraphics[width=1\textwidth,height=0.22\textheight]{paper_files/figure-pdf/fig-model-mnist-1.pdf}
-
-}
-
-\caption{\label{fig-model-mnist}\(10\text{-}D\) model error in
-\(2\text{-}D\). The \(2\text{-}D\) model shows a non-linear continuous
-structure. Most low \(10\text{-}D\) model errors are distributed along
-the lower edge of the \(2\text{-}D\) structure, while most large
-\(10\text{-}D\) model errors are concentrated along the upper edge.}
-
-\end{figure}%
-
-There are certain data points that exhibit high error rates due to their
-deviation from the usual \(10\text{-}D\) data structure, which makes
-them anomalies (Figure~\ref{fig-model-mnist}). These anomalies can be
-classified into two types: those that are anomalies within the
-non-linear structure and those that lie outside of it. The images
-associated with high model error points within the non-linear structure
-display different patterns of the digit 1, as shown in
-Figure~\ref{fig-mnist1-within}. However, when comparing these images to
-the ones found outside of the non-linear structure, it becomes evident
-that the latter display different patterns of the digit 1
-(Figure~\ref{fig-mnist1-out}).
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-model-mnist-1.pdf}
+
+}
+
+\caption{\label{fig-model-mnist}\(10\text{-}D\) model error in PaCMAP
+layout of MNIST digit 1 dataset in \(2\text{-}D\). Most low
+\(10\text{-}D\) model errors are distributed along the lower edge of the
+\(2\text{-}D\) structure, while most large \(10\text{-}D\) model errors
+are concentrated along the upper edge (a). There are certain data points
+that exhibit high error rates due to their deviation from the usual
+\(10\text{-}D\) data structure, which makes them anomalies. These
+anomalies can be classified into two types: those that are anomalies
+within the non-linear structure and those that lie outside of it. The
+images associated with high model error points within the non-linear
+structure display different patterns of the digit 1 (b). However, when
+comparing these images to the ones found outside of the non-linear
+structure, it becomes evident that the latter display different patterns
+of the digit 1 (c).}
+
+\end{figure}%
 

### 2024-08-26 paper.qmd
Updated Figure 17

@@
       )
-  )
+  ) +
+  interior_annotation("a")
 ```
@@
 Most low $10\\text{-}D$ model errors are distributed along the lower edge of the $2\\text{-}D$ structure, while most large $10\\text{-}D$ model errors are concentrated along the upper edge (a).
-There are certain data points that exhibit high error rates due to their deviation from the usual $10\\text{-}D$ data structure, which makes them anomalies.
+There are certain data points that exhibit large error rates due to their deviation from the usual $10\\text{-}D$ data structure, which makes them anomalies.
 These anomalies can be classified into two types: those that are anomalies within the non-linear structure and those that lie outside of it.

### 2024-08-26 paper.qmd
Updated Figure 17

--- 2024-08-26 paper.qmd
@@
     row_wise_abs_error <= 6 ~ "error 4-6",
-    row_wise_abs_error <= 6 ~ "error 6-8",
+    row_wise_abs_error <= 8 ~ "error 6-8",
     row_wise_abs_error <= 10 ~ "error 8-10",
@@
 
-```{r}
-#| echo: false
+<!--images that occur large error-->
+<!-- within the nonlinear structure-->
+<!--outside the nonlinear structure-->
+
+```{r}
+#| label: img-with-error
+
+## Data with pixel values
+mnist_data <- read_rds("data/mnist/mnist_digit_1.rds")
+
+img_error_within <- c(6169, 5987, 773, 6671, 4785, 4699, 5173, 5475)
+
+pixels_gathered_within <-  mnist_data |>
+  mutate(instance = row_number()) |>
+  gather(pixel, value, -Label, -instance) |>
+  extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
+  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
+  filter(instance %in% img_error_within)
+
+imge_error_sample_within <- pixels_gathered_within |>
+  ggplot(aes(x, y, fill = value)) +
+  geom_tile() +
+  facet_wrap(~ instance, ncol = 4) +
+  coord_fixed() +
+  scale_fill_continuous_sequential(palette = "Grays") +
+  theme(strip.background = element_blank(),
+        strip.text.x = element_blank(),
+        legend.position = "none")
+
+img_error_outside <- c(1489, 5095, 855, 3143, 453, 3619, 1015, 409) #, 5101
+
+pixels_gathered_outside <-  mnist_data |>
+  mutate(instance = row_number()) |>
+  gather(pixel, value, -Label, -instance) |>
+  extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
+  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
+  filter(instance %in% img_error_outside)
+
+imge_error_sample_outside <- pixels_gathered_outside |>
+  ggplot(aes(x, y, fill = value)) +
+  geom_tile() +
+  facet_wrap(~ instance, ncol = 4) +
+  coord_fixed() +
+  scale_fill_continuous_sequential(palette = "Grays") +
+  theme(strip.background = element_blank(),
+        strip.text.x = element_blank(),
+        legend.position = "none")
+
+```
+
+```{r}
+#| label: fig-model-mnist
 #| fig-cap: "$10\\text{-}D$ model error in PaCMAP layout of MNIST digit 1 dataset in $2\\text{-}D$.
 Most low $10\\text{-}D$ model errors are distributed along the lower edge of the $2\\text{-}D$ structure, while most large $10\\text{-}D$ model errors are concentrated along the upper edge (a).
-There are certain data points that exhibit high error rates due to their deviation from the usual $10\\text{-}D$ data structure, which makes them anomalies.
+There are certain data points that exhibit large error rates due to their deviation from the usual $10\\text{-}D$ data structure, which makes them anomalies.
 These anomalies can be classified into two types: those that are anomalies within the non-linear structure and those that lie outside of it.
@@
 "
-#| label: fig-model-mnist
 #| fig-pos: H
-
-error_plot_mnist 
+#| fig-width: 4
+#| fig-height: 8
+#| out-height: 80%
+
+error_plot_mnist / (imge_error_sample_within + imge_error_sample_outside) +
+  plot_annotation(tag_levels = 'a') &
+  theme(plot.tag.position  = c(0.92, 0.92),
+        plot.tag = element_text(color = "grey70"))
 ```
@@
 
-::: {#fig-mnist-anomalies layout-ncol="2" fig-pos="H"}
-![](figures/mnist/img_error_sample_within.png){#fig-mnist1-within}
-
-![](figures/mnist/img_error_sample_out.png){#fig-mnist1-out}
-
-Some images of handwritten digit 1 which occur high model error (a) within the non-linear structure, and (b) outside the non-linear structure. The images shows different patterns of digit 1.
-:::
+<!-- ::: {#fig-mnist-anomalies layout-ncol="2" fig-pos="H"} -->
+<!-- ![](figures/mnist/img_error_sample_within.png){#fig-mnist1-within} -->
+
+<!-- ![](figures/mnist/img_error_sample_out.png){#fig-mnist1-out} -->
+
+<!-- Some images of handwritten digit 1 which occur high model error (a) within the non-linear structure, and (b) outside the non-linear structure. The images shows different patterns of digit 1. -->
+<!-- ::: -->
 

### 2024-08-26 paper.tex
Updated Figure 17

@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-model-mnist-1.pdf}
+\includegraphics[width=1\textwidth,height=0.8\textheight]{paper_files/figure-pdf/fig-model-mnist-1.pdf}
 
@@
 are concentrated along the upper edge (a). There are certain data points
-that exhibit high error rates due to their deviation from the usual
+that exhibit large error rates due to their deviation from the usual
 \(10\text{-}D\) data structure, which makes them anomalies. These
@@
 of the digit 1 (c).}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\begin{minipage}{0.50\linewidth}
-
-\centering{
-
-\includegraphics{figures/mnist/img_error_sample_within.png}
-
-}
-
-\subcaption{\label{fig-mnist1-within}}
-
-\end{minipage}%
-%
-\begin{minipage}{0.50\linewidth}
-
-\centering{
-
-\includegraphics{figures/mnist/img_error_sample_out.png}
-
-}
-
-\subcaption{\label{fig-mnist1-out}}
-
-\end{minipage}%
-
-\caption{\label{fig-mnist-anomalies}Some images of handwritten digit 1
-which occur high model error (a) within the non-linear structure, and
-(b) outside the non-linear structure. The images shows different
-patterns of digit 1.}
 

### 2024-08-26 paper.qmd
Updated the section 7.2

@@
 In the field of single-cell studies, a common analytical task involves clustering to identify groups of cells with similar expression profiles.
-NLDR is commonly used to display clusters, and help to verify the results.
+NLDR methods are commonly used to display clusters, and help to verify the results.
 For example, @chen2023 illustrates the use of UMAP to identify clusters in Human Peripheral Blood Mononuclear Cells (PBMC3k).
-@fig-umap-author is a reproduction of the published plot.
+@fig-umap-author is the reproduction of the published plot.
 It shows three well-separated clusters with big separations.
@@
 
-To determine whether the UMAP representation with the (hyper-)parameter choice suggested by @chen2023 preserves the original data structure, we visualize the model constructed with UMAP overlaid on the $9\text{-}D$ data.
+To determine whether the UMAP representation with the specific (hyper-)parameter choice suggested by @chen2023 preserves the original data structure, we visualize the model constructed with UMAP overlaid on the $9\text{-}D$ data.
 The @fig-umap-author shows three well-separated clusters.
@@
 
-PaCMAP provides a reasonable \gD{} representation of MNIST digit 1 data, as it preserves the non-linear structure present in the $10\text{-}D$ data (@fig-mnist-tri-proj).
-Furthermore, shows some quirks when visualizing the *model-in-the-data-space* (@fig-mnist-tri-proj).
+As a result of visualizing the *model-in-the-data-space*, PaCMAP provides a reasonable \gD{} representation of MNIST digit 1 data, as it preserves the non-linear structure present in the $10\text{-}D$ data (@fig-mnist-tri-proj).
+It also shows some quirks like twisted patterns and long edges (@fig-mnist-tri-proj).
+Furthermore, the results help to identify anomalies (with large model errors) within and the outside the non-linear structure, displaying different patterns of the digit 1 (@fig-model-error-mnist).
 
@@
 ```{r}
-#| label: fig-model-mnist
+#| label: fig-model-error-mnist
 #| fig-cap: "$10\\text{-}D$ model error in PaCMAP layout of MNIST digit 1 dataset in $2\\text{-}D$.

### 2024-08-26 paper.tex
Updated the section 7.2

@@
 clustering to identify groups of cells with similar expression profiles.
-NLDR is commonly used to display clusters, and help to verify the
-results. For example, \citet{chen2023} illustrates the use of UMAP to
-identify clusters in Human Peripheral Blood Mononuclear Cells (PBMC3k).
-\textbf{?@fig-umap-author} is a reproduction of the published plot. It
-shows three well-separated clusters with big separations.
-
-To determine whether the UMAP representation with the (hyper-)parameter
-choice suggested by \citet{chen2023} preserves the original data
-structure, we visualize the model constructed with UMAP overlaid on the
-\(9\text{-}D\) data. The \textbf{?@fig-umap-author} shows three
-well-separated clusters. However, as shown in \textbf{?@fig-pbmc1-sc},
-there is no big separation between three clusters in \(9\text{-}D\).
-Therefore, the suggested UMAP representation
+NLDR methods are commonly used to display clusters, and help to verify
+the results. For example, \citet{chen2023} illustrates the use of UMAP
+to identify clusters in Human Peripheral Blood Mononuclear Cells
+(PBMC3k). \textbf{?@fig-umap-author} is the reproduction of the
+published plot. It shows three well-separated clusters with big
+separations.
+
+To determine whether the UMAP representation with the specific
+(hyper-)parameter choice suggested by \citet{chen2023} preserves the
+original data structure, we visualize the model constructed with UMAP
+overlaid on the \(9\text{-}D\) data. The \textbf{?@fig-umap-author}
+shows three well-separated clusters. However, as shown in
+\textbf{?@fig-pbmc1-sc}, there is no big separation between three
+clusters in \(9\text{-}D\). Therefore, the suggested UMAP representation
 (\textbf{?@fig-umap-author}) does not accurately represent the
@@
 
-PaCMAP provides a reasonable \gD{} representation of MNIST digit 1 data,
-as it preserves the non-linear structure present in the \(10\text{-}D\)
-data (Figure~\ref{fig-mnist-tri-proj}). Furthermore, shows some quirks
-when visualizing the \emph{model-in-the-data-space}
-(Figure~\ref{fig-mnist-tri-proj}).
+As a result of visualizing the \emph{model-in-the-data-space}, PaCMAP
+provides a reasonable \gD{} representation of MNIST digit 1 data, as it
+preserves the non-linear structure present in the \(10\text{-}D\) data
+(Figure~\ref{fig-mnist-tri-proj}). It also shows some quirks like
+twisted patterns and long edges (Figure~\ref{fig-mnist-tri-proj}).
+Furthermore, the results help to identify anomalies (with large model
+errors) within and the outside the non-linear structure, displaying
+different patterns of the digit 1 (Figure~\ref{fig-model-error-mnist}).
 
@@
 
-\includegraphics[width=1\textwidth,height=0.8\textheight]{paper_files/figure-pdf/fig-model-mnist-1.pdf}
-
-}
-
-\caption{\label{fig-model-mnist}\(10\text{-}D\) model error in PaCMAP
-layout of MNIST digit 1 dataset in \(2\text{-}D\). Most low
+\includegraphics[width=1\textwidth,height=0.8\textheight]{paper_files/figure-pdf/fig-model-error-mnist-1.pdf}
+
+}
+
+\caption{\label{fig-model-error-mnist}\(10\text{-}D\) model error in
+PaCMAP layout of MNIST digit 1 dataset in \(2\text{-}D\). Most low
 \(10\text{-}D\) model errors are distributed along the lower edge of the

### 2024-08-26 paper.qmd
Updated single cell experiments section

@@
 For example, @chen2023 illustrates the use of UMAP to identify clusters in Human Peripheral Blood Mononuclear Cells (PBMC3k).
-@fig-umap-author is the reproduction of the published plot.
+@fig-NLDR-variety (a) is the reproduction of the published plot.
 It shows three well-separated clusters with big separations.
@@
 
-As shown in @fig-tsne-suggest, there are three well-separated clusters, although the separation between the clusters are small.
-Additionally, non-linear structures can also be observed within the clusters (@fig-tsne-suggest).
+As shown in @fig-NLDR-variety (e), there are three well-separated clusters, although the separation between the clusters are small.
+Additionally, non-linear structures can also be observed within the clusters (@fig-NLDR-variety (e)).
 This demonstrates that tSNE accurately captures the data structure of the PBMC3k dataset, which UMAP did not achieve.

### 2024-08-26 paper.tex
Updated single cell experiments section

@@
 to identify clusters in Human Peripheral Blood Mononuclear Cells
-(PBMC3k). \textbf{?@fig-umap-author} is the reproduction of the
+(PBMC3k). Figure~\ref{fig-NLDR-variety} (a) is the reproduction of the
 published plot. It shows three well-separated clusters with big
@@
 
-As shown in \textbf{?@fig-tsne-suggest}, there are three well-separated
-clusters, although the separation between the clusters are small.
-Additionally, non-linear structures can also be observed within the
-clusters (\textbf{?@fig-tsne-suggest}). This demonstrates that tSNE
-accurately captures the data structure of the PBMC3k dataset, which UMAP
-did not achieve.
+As shown in Figure~\ref{fig-NLDR-variety} (e), there are three
+well-separated clusters, although the separation between the clusters
+are small. Additionally, non-linear structures can also be observed
+within the clusters (Figure~\ref{fig-NLDR-variety} (e)). This
+demonstrates that tSNE accurately captures the data structure of the
+PBMC3k dataset, which UMAP did not achieve.
 

### 2024-08-26 paper.qmd
Updated intro section 7.1

@@
 For example, @chen2023 illustrates the use of UMAP to identify clusters in Human Peripheral Blood Mononuclear Cells (PBMC3k).
+There are $2622$ single cells.
+First 9 principal components are used to generate the UMAP.
 @fig-NLDR-variety (a) is the reproduction of the published plot.
-It shows three well-separated clusters with big separations.
 
@@
 
-To determine whether the UMAP representation with the specific (hyper-)parameter choice suggested by @chen2023 preserves the original data structure, we visualize the model constructed with UMAP overlaid on the $9\text{-}D$ data.
-The @fig-umap-author shows three well-separated clusters.
-However, as shown in @fig-pbmc1-sc, there is no big separation between three clusters in $9\text{-}D$.
-Therefore, the suggested UMAP representation (@fig-umap-author) does not accurately represent the structure(s) present in PBMC3k dataset.
-But, when visualizing the model-in-th-data-space some unobserved structures can be seen.
-Some clusters have non-linear continuity patterns and high-density patches (@fig-pbmc1-sc).
+To determine whether the UMAP representation with the specific (hyper-)parameter choice suggested by @chen2023 preserves the data structure present in $9\text{-}D$ (@fig-model-pbmc-author-proj (b)), we visualize the model constructed with UMAP overlaid on the $9\text{-}D$ data.
+The @fig-NLDR-variety (a) shows three well-separated clusters with big separations.
+However, as shown in @fig-model-pbmc-author-proj, there is no big separation between three clusters in $9\text{-}D$.
+Therefore, the suggested UMAP representation (@fig-NLDR-variety (a)) does not accurately represent the structure of PBMC3k dataset in $9\text{-}D$.
+But, when visualizing the *model-in-th-data-space* some unobserved structures can be seen.
+Some clusters have non-linear continuity patterns and high-density patches (@fig-model-pbmc-author-proj).
 
@@
 #| fig-width: 10
-#| fig-cap: "Model in \\gD{}, on the UMAP layout, and three views of the fit in projections from $9\\text{-}D$, for the PBMC3k data ($(s_1, \ s_2) = (-0.050, \ -0.041)$, $b = 870 \  (30, \ 29)$, $m = 135$, benchmark value to remove large edges is $0.099$) (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/VqqWuE0Jj6A>)."
+#| fig-cap: "Model in \\gD{}, on the UMAP layout, and three views of the fit in projections from $9\\text{-}D$, for the PBMC3k data ($b = 870 \  (30, \ 29)$, $m = 135$, benchmark value to remove large edges is $0.099$) The full video is available at <https://youtu.be/VqqWuE0Jj6A>."
 

### 2024-08-26 paper.tex
Updated intro section 7.1

@@
 to identify clusters in Human Peripheral Blood Mononuclear Cells
-(PBMC3k). Figure~\ref{fig-NLDR-variety} (a) is the reproduction of the
-published plot. It shows three well-separated clusters with big
-separations.
+(PBMC3k). There are \(2622\) single cells. First 9 principal components
+are used to generate the UMAP. Figure~\ref{fig-NLDR-variety} (a) is the
+reproduction of the published plot.
 
@@
 (hyper-)parameter choice suggested by \citet{chen2023} preserves the
-original data structure, we visualize the model constructed with UMAP
-overlaid on the \(9\text{-}D\) data. The \textbf{?@fig-umap-author}
-shows three well-separated clusters. However, as shown in
-\textbf{?@fig-pbmc1-sc}, there is no big separation between three
-clusters in \(9\text{-}D\). Therefore, the suggested UMAP representation
-(\textbf{?@fig-umap-author}) does not accurately represent the
-structure(s) present in PBMC3k dataset. But, when visualizing the
-model-in-th-data-space some unobserved structures can be seen. Some
-clusters have non-linear continuity patterns and high-density patches
-(\textbf{?@fig-pbmc1-sc}).
+data structure present in \(9\text{-}D\)
+(Figure~\ref{fig-model-pbmc-author-proj} (b)), we visualize the model
+constructed with UMAP overlaid on the \(9\text{-}D\) data. The
+Figure~\ref{fig-NLDR-variety} (a) shows three well-separated clusters
+with big separations. However, as shown in
+Figure~\ref{fig-model-pbmc-author-proj}, there is no big separation
+between three clusters in \(9\text{-}D\). Therefore, the suggested UMAP
+representation (Figure~\ref{fig-NLDR-variety} (a)) does not accurately
+represent the structure of PBMC3k dataset in \(9\text{-}D\). But, when
+visualizing the \emph{model-in-th-data-space} some unobserved structures
+can be seen. Some clusters have non-linear continuity patterns and
+high-density patches (Figure~\ref{fig-model-pbmc-author-proj}).
 
@@
 layout, and three views of the fit in projections from \(9\text{-}D\),
-for the PBMC3k data (\((s_1,  s_2) = (-0.050,  -0.041)\),
-\(b = 870   (30,  29)\), \(m = 135\), benchmark value to remove large
-edges is \(0.099\)) (The \textbf{langevitour} software is used to view
-the data with a tour, and the full video is available at
-\url{https://youtu.be/VqqWuE0Jj6A}).}
+for the PBMC3k data (\(b = 870   (30,  29)\), \(m = 135\), benchmark
+value to remove large edges is \(0.099\)) The full video is available at
+\url{https://youtu.be/VqqWuE0Jj6A}.}
 

### 2024-08-26 paper.qmd
Updated Figure 12 caption

@@
 Therefore, the suggested UMAP representation (@fig-NLDR-variety (a)) does not accurately represent the structure of PBMC3k dataset in $9\text{-}D$.
-But, when visualizing the *model-in-th-data-space* some unobserved structures can be seen.
+But, when visualizing the *model-in-the-data-space* some unobserved structures can be seen.
 Some clusters have non-linear continuity patterns and high-density patches (@fig-model-pbmc-author-proj).
@@
 #| fig-width: 10
-#| fig-cap: "Model in \\gD{}, on the UMAP layout, and three views of the fit in projections from $9\\text{-}D$, for the PBMC3k data ($b = 870 \  (30, \ 29)$, $m = 135$, benchmark value to remove large edges is $0.099$) The full video is available at <https://youtu.be/VqqWuE0Jj6A>."
+#| fig-cap: "Model in \\gD{}, on the UMAP layout (a), and three views of the fit in projections from $9\\text{-}D$ (b, c, d), for the PBMC3k data ($a_1 = 0.046$, $b = 676 \\  (26, \ 26)$, $m = 92$, benchmark value to remove large edges is $0.08$).
+The three well-separated clusters observed in the \\gD{} layout of UMAP (@fig-NLDR-variety (a)) is also visible when visualizing the model overlaid on the data space.
+But the separation between the clusters are small in $9\\text{-}D$.
+The model shows non-linear clusters and densed points within clusters in $9\\text{-}D$ space, which are some additional patterns not visible in the \\gD{} representation (@fig-NLDR-variety (a)).
+The full video is available at <https://youtu.be/VqqWuE0Jj6A>."
 

### 2024-08-26 paper.tex
Updated Figure 12 caption

@@
 represent the structure of PBMC3k dataset in \(9\text{-}D\). But, when
-visualizing the \emph{model-in-th-data-space} some unobserved structures
-can be seen. Some clusters have non-linear continuity patterns and
-high-density patches (Figure~\ref{fig-model-pbmc-author-proj}).
+visualizing the \emph{model-in-the-data-space} some unobserved
+structures can be seen. Some clusters have non-linear continuity
+patterns and high-density patches
+(Figure~\ref{fig-model-pbmc-author-proj}).
 
@@
 \caption{\label{fig-model-pbmc-author-proj}Model in \gD{}, on the UMAP
-layout, and three views of the fit in projections from \(9\text{-}D\),
-for the PBMC3k data (\(b = 870   (30,  29)\), \(m = 135\), benchmark
-value to remove large edges is \(0.099\)) The full video is available at
-\url{https://youtu.be/VqqWuE0Jj6A}.}
+layout (a), and three views of the fit in projections from
+\(9\text{-}D\) (b, c, d), for the PBMC3k data (\(a_1 = 0.046\),
+\(b = 676 \  (26,  26)\), \(m = 92\), benchmark value to remove large
+edges is \(0.08\)). The three well-separated clusters observed in the
+\gD{} layout of UMAP (Figure~\ref{fig-NLDR-variety} (a)) is also visible
+when visualizing the model overlaid on the data space. But the
+separation between the clusters are small in \(9\text{-}D\). The model
+shows non-linear clusters and densed points within clusters in
+\(9\text{-}D\) space, which are some additional patterns not visible in
+the \gD{} representation (Figure~\ref{fig-NLDR-variety} (a)). The full
+video is available at \url{https://youtu.be/VqqWuE0Jj6A}.}
 

### 2024-08-26 paper.qmd
Updated descriptions for section 7.1

@@
 
+In order to find a reasonable NLDR representation for the PBMC3k dataset, the MSE for different $a_1$ using UMAP, tSNE, PHATE, PaCMAP, and TriMAP with different (hyper-)parameter settings (@fig-pbmc-mse) were calculated.
+tSNE with a perplexity value set to $30$, the default parameter setting, is considered as a reasonable representation for the PBMC3k dataset.
+As shown in @fig-NLDR-variety (e), there are three well-separated clusters, although the separation between the clusters are small.
+Additionally, non-linear structures can also be observed within the clusters (@fig-NLDR-variety (e)).
+The visualization of the *model-in-the-data-space* provides evidence for this (@fig-pbmc2-sc).
+Therefore, tSNE with perplexity $30$ accurately captures the data structure of the PBMC3k dataset, which is better than UMAP.
+
+<!-- As shown in @fig-NLDR-variety (e), there are three well-separated clusters, although the separation between the clusters are small.
+Additionally, non-linear structures can also be observed within the clusters (@fig-NLDR-variety (e)).
+This demonstrates that tSNE accurately captures the data structure of the PBMC3k dataset, which UMAP did not achieve.
+-->
+
+<!-- In order to find a reasonable NLDR representation for the PBMC3k dataset, the MSE for different $b_1$ using UMAP with different (hyper-)parameter settings and tSNE with default (hyper-)parameter setting (@fig-pbmc-mse) were calculated.
+After analyzing the results, it was found that tSNE with default (hyper-)parameter setting (perplexity: $30$) achieved the lowest error.
+Therefore, tSNE with a perplexity value set to $30$, the default parameter setting, is considered as a reasonable representation for the PBMC3k dataset.
+-->
+
 <!--Fit the best model for author suggestion and compute error-->
@@
 <!-- ::: -->
-
-In order to find a reasonable NLDR representation for the PBMC3k dataset, the MSE for different $b_1$ using UMAP with different (hyper-)parameter settings and tSNE with default (hyper-)parameter setting (@fig-pbmc-mse) were calculated.
-After analyzing the results, it was found that tSNE with default (hyper-)parameter setting (perplexity: $30$) achieved the lowest error.
-Therefore, tSNE with a perplexity value set to $30$, the default parameter setting, is considered as a reasonable representation for the PBMC3k dataset.
-
 <!--compute absolute error for different parameter choices-->
@@
 ```
-
-As shown in @fig-NLDR-variety (e), there are three well-separated clusters, although the separation between the clusters are small.
-Additionally, non-linear structures can also be observed within the clusters (@fig-NLDR-variety (e)).
-This demonstrates that tSNE accurately captures the data structure of the PBMC3k dataset, which UMAP did not achieve.
 

### 2024-08-26 paper.tex
Updated descriptions for section 7.1

@@
 
+In order to find a reasonable NLDR representation for the PBMC3k
+dataset, the MSE for different \(a_1\) using UMAP, tSNE, PHATE, PaCMAP,
+and TriMAP with different (hyper-)parameter settings
+(Figure~\ref{fig-pbmc-mse}) were calculated. tSNE with a perplexity
+value set to \(30\), the default parameter setting, is considered as a
+reasonable representation for the PBMC3k dataset. As shown in
+Figure~\ref{fig-NLDR-variety} (e), there are three well-separated
+clusters, although the separation between the clusters are small.
+Additionally, non-linear structures can also be observed within the
+clusters (Figure~\ref{fig-NLDR-variety} (e)). The visualization of the
+\emph{model-in-the-data-space} provides evidence for this
+(Figure~\ref{fig-pbmc2-sc}). Therefore, tSNE with perplexity \(30\)
+accurately captures the data structure of the PBMC3k dataset, which is
+better than UMAP.
+
 \begin{figure}[H]
@@
 
-In order to find a reasonable NLDR representation for the PBMC3k
-dataset, the MSE for different \(b_1\) using UMAP with different
-(hyper-)parameter settings and tSNE with default (hyper-)parameter
-setting (Figure~\ref{fig-pbmc-mse}) were calculated. After analyzing the
-results, it was found that tSNE with default (hyper-)parameter setting
-(perplexity: \(30\)) achieved the lowest error. Therefore, tSNE with a
-perplexity value set to \(30\), the default parameter setting, is
-considered as a reasonable representation for the PBMC3k dataset.
-
 \begin{figure}[H]
@@
 \end{figure}%
-
-As shown in Figure~\ref{fig-NLDR-variety} (e), there are three
-well-separated clusters, although the separation between the clusters
-are small. Additionally, non-linear structures can also be observed
-within the clusters (Figure~\ref{fig-NLDR-variety} (e)). This
-demonstrates that tSNE accurately captures the data structure of the
-PBMC3k dataset, which UMAP did not achieve.
 

### 2024-08-26 paper.qmd
Figure 13 replace top with left
Updated Figure 14 title

@@
 #| fig-cap: "Assessing which of the 8 NLDR layouts on the PBMC3k data  (shown in Figure 1) is the better representation using MSE for varying binwidth ($a_1$).
-Colour  used for the lines and points in the top plot and in the scatterplots represents NLDR layout (a-h).
+Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-h).
 Layout f is universally poor.
@@
 #| echo: false
-#| fig-cap: "Model in $2\\text{-}D$, on the tSNE layout, and three views of the fit in projections from $9\\text{-}D$, for the PBMC3k data ($(s_1, \ s_2) = (-0.050, \ -0.058)$, $b = 300 \  (15, \ 20)$, $m = 136$, benchmark value to remove large edges is $0.133$).
-(The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/5Y1hE4i7N2k>)."
+#| fig-cap: "Model in \\gD{}, on the tSNE layout (a), and three views of the fit in projections from $9\\text{-}D$ (b, c, d), for the PBMC3k data ($a_1 = 0.099$, $b = 221 \\  (13, \ 17)$, $m = 92$, benchmark value to remove large edges is $0.1$).
+The three well-separated clusters with small separations observed in the \\gD{} layout of tSNE (@fig-NLDR-variety (e)) is also visible when visualizing the model overlaid on the data space in $9\\text{-}D$.
+Furthermore, there are some non-linear clusters.
+The model shows densed points within clusters in $9\\text{-}D$ space, which is an additional patterns not visible in the \\gD{} representation (@fig-NLDR-variety (e)).
+The full video is available at <https://youtu.be/5Y1hE4i7N2k>."
 #| label: fig-pbmc2-sc

### 2024-08-26 paper.tex
Figure 13 replace top with left
Updated Figure 14 title

@@
 MSE for varying binwidth (\(a_1\)). Colour used for the lines and points
-in the top plot and in the scatterplots represents NLDR layout (a-h).
+in the left plot and in the scatterplots represents NLDR layout (a-h).
 Layout f is universally poor. Layouts a, b, g, h that show large
@@
 
-\caption{\label{fig-pbmc2-sc}Model in \(2\text{-}D\), on the tSNE
-layout, and three views of the fit in projections from \(9\text{-}D\),
-for the PBMC3k data (\((s_1,  s_2) = (-0.050,  -0.058)\),
-\(b = 300   (15,  20)\), \(m = 136\), benchmark value to remove large
-edges is \(0.133\)). (The \textbf{langevitour} software is used to view
-the data with a tour, and the full video is available at
-\url{https://youtu.be/5Y1hE4i7N2k}).}
+\caption{\label{fig-pbmc2-sc}Model in \gD{}, on the tSNE layout (a), and
+three views of the fit in projections from \(9\text{-}D\) (b, c, d), for
+the PBMC3k data (\(a_1 = 0.099\), \(b = 221 \  (13,  17)\), \(m = 92\),
+benchmark value to remove large edges is \(0.1\)). The three
+well-separated clusters with small separations observed in the \gD{}
+layout of tSNE (Figure~\ref{fig-NLDR-variety} (e)) is also visible when
+visualizing the model overlaid on the data space in \(9\text{-}D\).
+Furthermore, there are some non-linear clusters. The model shows densed
+points within clusters in \(9\text{-}D\) space, which is an additional
+patterns not visible in the \gD{} representation
+(Figure~\ref{fig-NLDR-variety} (e)). The full video is available at
+\url{https://youtu.be/5Y1hE4i7N2k}.}
 

### 2024-08-26 paper.qmd
Need to complete section 6

@@
 
-In this section, the effectiveness of the algorithm is described using a simulated dataset.
+In this section, we will assess how t-SNE, UMAP, and PaCMAP algorithms preserve the clustering structure.
+We will use a simulated dataset consisting of five spherical Gaussian clusters where each cluster occupies a different corner of the $4\text{-}D$ space, with each cluster containing an equal number of points ($1000$) and the same within-cluster variation.
+The data simulated from $\Sigma = \begin{pmatrix}
+0.0025 & 0 & 0 & 0 \\
+0 & 0.0025 & 0 & 0 \\
+0 & 0 & 0.0025 & 0 \\
+0 & 0 & 0 & 0.0025 \\
+\end{pmatrix}$.
+
+<!-- In this section, the effectiveness of the algorithm is described using a simulated dataset.
 The dataset consists of five spherical Gaussian clusters in $4\text{-}D$, with each cluster containing an equal number of points and the same within-cluster variation.
+-->
 

### 2024-08-26 paper.tex
Need to complete section 6

@@
 
-In this section, the effectiveness of the algorithm is described using a
-simulated dataset. The dataset consists of five spherical Gaussian
-clusters in \(4\text{-}D\), with each cluster containing an equal number
-of points and the same within-cluster variation.
+In this section, we will assess how t-SNE, UMAP, and PaCMAP algorithms
+preserve the clustering structure. We will use a simulated dataset
+consisting of five spherical Gaussian clusters where each cluster
+occupies a different corner of the \(4\text{-}D\) space, with each
+cluster containing an equal number of points (\(1000\)) and the same
+within-cluster variation. The data simulated from
+\(\Sigma = \begin{pmatrix}
+0.0025 & 0 & 0 & 0 \\
+0 & 0.0025 & 0 & 0 \\
+0 & 0 & 0.0025 & 0 \\
+0 & 0 & 0 & 0.0025 \\
+\end{pmatrix}\).
 

### 2024-08-26 paper.qmd
Added 2D layouts of Five Gau

@@
 
+tsne_gau <- tsne_gau_scaled |>
+  ggplot(aes(x = tSNE1,
+             y = tSNE2)) +
+  geom_point(alpha=0.3, color = "#000000")
+
 # Apply the scaling
@@
 
+umap_gau <- umap_gau_scaled |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2)) +
+  geom_point(alpha=0.3, color = "#000000")
+
 ## Compute hexbin parameters
@@
 
+pacmap_gau <- pacmap_gau_scaled |>
+  ggplot(aes(x = PaCMAP1,
+             y = PaCMAP2)) +
+  geom_point(alpha=0.3, color = "#000000")
+
 ## Compute hexbin parameters
@@
 ```{r}
+#| label: fig-five-gau-nldr-layouts
+#| fig-height: 5
+#| fig-width: 15
+#| fig-pos: H
+#| fig-cap: "(a) tSNE, (b) UMAP, and (c) PaCMAP"
+
+tsne_gau + umap_gau + pacmap_gau +
+  plot_layout(ncol = 3)
+```
+
+```{r}
 #| label: fig-five-gau-projs

### 2024-08-26 paper.qmd
Rearranged Figure 12

@@
 
-five_gau_proj_tsne_model1 + five_gau_proj_tsne_model2 +
-  five_gau_proj_umap_model1 + five_gau_proj_umap_model2 +
-  five_gau_proj_pacmap_model1 + five_gau_proj_pacmap_model2 +
-  plot_layout(guides = "collect", ncol = 2) &
+five_gau_proj_tsne_model1 + five_gau_proj_umap_model1 + five_gau_proj_pacmap_model1 + five_gau_proj_tsne_model2 + five_gau_proj_umap_model2 + five_gau_proj_pacmap_model2 +
+  plot_layout(guides = "collect", nrow = 2) &
   theme(legend.position='none')

### 2024-08-26 paper.qmd
Added main idea for Section 6, need to clean the description and add everthing in a nice flow

@@
 
+In the t-SNE 2D layout, the distances between clusters appear smaller, leading to more tightly packed clusters, which is typical of t-SNE's focus on preserving local structures.
+This often results in less emphasis on the global distances between clusters.
+On the other hand, UMAP tends to maintain larger separations between clusters, better reflecting the global structure and preserving the relative distances between clusters more effectively.
+PaCMAP, meanwhile, balances these aspects by maintaining the distance between clusters and positioning them correctly, effectively capturing both global and local structures.
+
+The t-SNE, UMAP, and PaCMAP layouts each reveal different aspects of the clustering structure within the dataset.
+When examining a specific cluster, which appears more like a filled-out pancake in higher dimensions, we observe that t-SNE, UMAP, and PaCMAP all effectively capture global structures.
+However, there is evidence suggesting that PaCMAP, with its default hyper-parameter settings, struggles to preserve the local structure as effectively as t-SNE and UMAP.
+
+The differences in performance can be attributed to several factors.
+One key element is the choice of hyper-parameters, particularly the number of neighbors considered during the dimensionality reduction process.
+This parameter significantly impacts how well local structures are maintained.
+Additionally, the intrinsic dimensionality of the data, the balance between global and local structures, and issues related to optimization and initialization all contribute to the observed outcomes.
+
+It's important to note that these observations are made with specific hyper-parameter settings.
+Adjusting these parameters could lead to different results, potentially improving PaCMAP's ability to capture local structures or altering how t-SNE and UMAP handle global and local structures.
+Thus, while PaCMAP generally balances global and local preservation, careful tuning of its parameters is crucial for achieving the best results in any given dataset.
+
 <!-- In this section, the effectiveness of the algorithm is described using a simulated dataset.
@@
 
-XXX Add tSNE, UMAP, and PaCMAP NLDR layouts and describe what is happening
-XXX Discuss the models of one specific cluster (filled out, pancakes): tSNE, UMAP, PaCMAP are useful to capture global structures, but this is an evidence where PaCMAP fail to capture the local structure with default hyper-parameter settings than tSNE and UMAP
-XXX Why? hyper-parameter choice specially number of neighbors, data characteristics (intrinsic dimensionality), imbalance between global and local structures, optimization and initialization issues
-XXX Add this is with specific hyper-parameter choices
+<!-- XXX Add tSNE, UMAP, and PaCMAP NLDR layouts and describe what is happening -->
+<!-- XXX Discuss the models of one specific cluster (filled out, pancakes): tSNE, UMAP, PaCMAP are useful to capture global structures, but this is an evidence where PaCMAP fail to capture the local structure with default hyper-parameter settings than tSNE and UMAP -->
+<!-- XXX Why?
+hyper-parameter choice specially number of neighbors, data characteristics (intrinsic dimensionality), imbalance between global and local structures, optimization and initialization issues -->
+<!-- XXX Add this is with specific hyper-parameter choices -->
 

### 2024-08-26 paper.tex
Added 2D layouts of Five Gau
Added main idea for Section 6, need to clean the description and add everthing in a nice flow

@@
 
-XXX Add tSNE, UMAP, and PaCMAP NLDR layouts and describe what is
-happening XXX Discuss the models of one specific cluster (filled out,
-pancakes): tSNE, UMAP, PaCMAP are useful to capture global structures,
-but this is an evidence where PaCMAP fail to capture the local structure
-with default hyper-parameter settings than tSNE and UMAP XXX Why?
-hyper-parameter choice specially number of neighbors, data
-characteristics (intrinsic dimensionality), imbalance between global and
-local structures, optimization and initialization issues XXX Add this is
-with specific hyper-parameter choices
+In the t-SNE 2D layout, the distances between clusters appear smaller,
+leading to more tightly packed clusters, which is typical of t-SNE's
+focus on preserving local structures. This often results in less
+emphasis on the global distances between clusters. On the other hand,
+UMAP tends to maintain larger separations between clusters, better
+reflecting the global structure and preserving the relative distances
+between clusters more effectively. PaCMAP, meanwhile, balances these
+aspects by maintaining the distance between clusters and positioning
+them correctly, effectively capturing both global and local structures.
+
+The t-SNE, UMAP, and PaCMAP layouts each reveal different aspects of the
+clustering structure within the dataset. When examining a specific
+cluster, which appears more like a filled-out pancake in higher
+dimensions, we observe that t-SNE, UMAP, and PaCMAP all effectively
+capture global structures. However, there is evidence suggesting that
+PaCMAP, with its default hyper-parameter settings, struggles to preserve
+the local structure as effectively as t-SNE and UMAP.
+
+The differences in performance can be attributed to several factors. One
+key element is the choice of hyper-parameters, particularly the number
+of neighbors considered during the dimensionality reduction process.
+This parameter significantly impacts how well local structures are
+maintained. Additionally, the intrinsic dimensionality of the data, the
+balance between global and local structures, and issues related to
+optimization and initialization all contribute to the observed outcomes.
+
+It's important to note that these observations are made with specific
+hyper-parameter settings. Adjusting these parameters could lead to
+different results, potentially improving PaCMAP's ability to capture
+local structures or altering how t-SNE and UMAP handle global and local
+structures. Thus, while PaCMAP generally balances global and local
+preservation, careful tuning of its parameters is crucial for achieving
+the best results in any given dataset.
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-five-gau-nldr-layouts-1.pdf}
+
+}
+
+\caption{\label{fig-five-gau-nldr-layouts}(a) tSNE, (b) UMAP, and (c)
+PaCMAP}
+
+\end{figure}%
 

### 2024-08-27 paper.qmd
Updated Figure 13

@@
   coord_fixed() +
+  xlim(c(-1, 1)) +
+  ylim(c(-1, 1)) +
   interior_annotation("b")
@@
   coord_fixed() +
+  xlim(c(-1, 1)) +
+  ylim(c(-1, 1)) +
   interior_annotation("c")
@@
   coord_fixed() +
+  xlim(c(-1, 1)) +
+  ylim(c(-1, 1)) +
   interior_annotation("d")
@@
 
-trimesh_removed_pbmc + pbmc_proj_tsne_model1 +
+free(trimesh_removed_pbmc) + pbmc_proj_tsne_model1 +
   pbmc_proj_tsne_model2 + pbmc_proj_tsne_model3 +

### 2024-08-27 paper.qmd
Updated Figure 15, 17, 13

@@
 
-trimesh_removed_pbmc + pbmc_proj_umap_model1 + 
+free(trimesh_removed_pbmc) + pbmc_proj_umap_model1 + 
   pbmc_proj_umap_model2 + pbmc_proj_umap_model3 +
@@
 
-trimesh_removed_mnist + mnist_proj_pacmap_model1 +
+free(trimesh_removed_mnist) + mnist_proj_pacmap_model1 +
   mnist_proj_pacmap_model2 + mnist_proj_pacmap_model3 +

### 2024-08-27 paper.qmd
Updated Figure 17 axes positions
Reshape the axes

@@
   proj_data = projected_df,  
-  limits = 1)
+  limits = 1.5,
+  axis_pos_x = -0.75,
+  axis_pos_y = -0.75)
 
@@
   proj_data = projected_df, 
-  limits = 1)
+  limits = 1.5,
+  axis_pos_x = -0.75,
+  axis_pos_y = -0.75)
 

### 2024-08-27 paper.qmd
Updated scaling for Figure 14

@@
                           aes(x = a1, 
-                              y = log(MSE), 
+                              y = MSE, 
                               colour = method)) + 
@@
                               '#984ea3','#999999','#a65628','#f781bf')) +
+  scale_y_log10() +
   ylab("log(MSE)") +

### 2024-08-27 paper.qmd
Try to find the best layout with UMAP...still need to work on that

@@
 error_pbmc_umap2 <- read_rds("data/pbmc3k/error_pbmc_umap_5_min_dist_0.01.rds")
-error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_15_min_dist_0.99.rds")
+# error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_15_min_dist_0.99.rds")
+error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_12_min_dist_0.99.rds")
 error_pbmc_tsne <- read_rds("data/pbmc3k/error_pbmc_tsne_5.rds")

### 2024-08-27 paper.qmd
Updated Figure 12 plot labels

@@
   ylim(c(-0.15, 0.18)) +
-  interior_annotation("a", 
+  interior_annotation("a1", 
                       position = c(0.08, 0.9),
-                      cex = 3)
+                      cex = 2)
 
@@
   ylim(c(-0.15, 0.25)) +
-  interior_annotation("b", 
+  interior_annotation("a2", 
                       position = c(0.08, 0.9),
-                      cex = 3)
+                      cex = 2)
 
@@
   ylim(c(-0.9, -0.55)) +
-  interior_annotation("c", 
+  interior_annotation("b1", 
                       position = c(0.08, 0.9),
-                      cex = 3)
+                      cex = 2)
 
@@
   ylim(c(-0.15, 0.3)) +
-  interior_annotation("d", 
+  interior_annotation("b2", 
                       position = c(0.08, 0.9),
-                      cex = 3)
+                      cex = 2)
 
@@
   ylim(c(0, 0.35)) +
-  interior_annotation("e", 
+  interior_annotation("c1", 
                       position = c(0.08, 0.9),
-                      cex = 3)
+                      cex = 2)
 
@@
   ylim(c(-0.45, 0)) +
-  interior_annotation("f", 
+  interior_annotation("c2", 
                       position = c(0.08, 0.9),
-                      cex = 3)
+                      cex = 2)
 

### 2024-08-28 paper.qmd
Updated Figure 9 (c)

@@
     col_start = "x") |>
-    mutate(benchmark_rm_lwd = benchmark_rm_lwd,
+    mutate(benchmark_rm_lwd = round(benchmark_rm_lwd, 2),
            bin1 = effective_bin1_scurve,
@@
 mse_scurve_lwd <- ggplot(error_rm_scurve, 
-                     aes(x = mean_counts, 
+                     aes(x = benchmark_rm_lwd, 
                          y = MSE)) + 
@@
   #        transform = "reverse") + 
-  scale_x_continuous(breaks = unique(error_rm_scurve$mean_counts)) +
-  xlab(expression(mean(n[h]))) +
+  scale_x_continuous(breaks = unique(error_rm_scurve$benchmark_rm_lwd)) +
+  xlab("threshold to remove low-density hexagons") +
   ylab("MSE") +

### 2024-08-30 paper.qmd
Generate embeddings with 1000 points for Five Gaussian clusters
Fixed the issue of having model and data little bit seprated than usual

@@
 # Function to scale data manually
-scale_data_manual <- function(data) {
+scale_data_manual <- function(data, type_col) {
   # Step 1: Center the data (mean 0)
-  data_centered <- center_data(data)
+  data_centered <- center_data(data |> select(-all_of(type_col)))
 
@@
   }
+
+  # Combine the scaled data with the 'type' column and return as a tibble
+  data_scaled <- as_tibble(data_scaled) %>% 
+    mutate(!!type_col := data[[type_col]])
 
@@
 
-# Apply the scaling function to the s_curve_noise data
-scaled_s_curve_noise <- scale_data_manual(
-  training_data_scurve |> select(-ID)) |>
+data_scurve <- training_data_scurve |> 
+  select(-ID) |> 
+  mutate(type = "data")
+
+model_scurve <- true_model_df |> 
+    select(-ID) |> 
+  mutate(type = "model")
+
+# Apply the scaling
+df_model_data_scurve <- bind_rows(data_scurve, model_scurve)
+scaled_scurve <- scale_data_manual(df_model_data_scurve, "type") |>
   as_tibble()
 
-# Scaling the true model and add the mean of scaled noise dimensions
-scaled_s_curve_true <- scale_data_manual(
-  true_model_df |> 
-    select(-ID) |>
-    select(x1:x3)) |>
-  as_tibble() |>
-  mutate(x4 = mean(scaled_s_curve_noise$x4),
-         x5 = mean(scaled_s_curve_noise$x5),
-         x6 = mean(scaled_s_curve_noise$x6),
-         x7 = mean(scaled_s_curve_noise$x7))
+scaled_s_curve_noise <- scaled_scurve |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_s_curve_true <- scaled_scurve |>
+  filter(type == "model") |>
+  select(-type)
 
@@
 
-training_data_gau <- read_rds("data/five_gau_clusters/data_five_gau_training.rds")
-
-tsne_data_gau <- read_rds("data/five_gau_clusters/tsne_data_five_gau_61.rds")
+training_data_gau <- read_rds("data/five_gau_clusters/data_five_gau.rds")
+
+data_gau <- training_data_gau |> 
+  select(-ID) |>
+  mutate(type = "data")
+
+tsne_data_gau <- read_rds("data/five_gau_clusters/tsne_data_five_gau_71.rds")
 gau1_scaled_obj <- gen_scaled_data(
@@
   geom_point(alpha=0.3, color = "#000000")
-
-# Apply the scaling
-scaled_gau_data <- scale_data_manual(training_data_gau |> select(-ID)) |>
-  as_tibble()
 
@@
 df_b <- df_b[match(df_bin_centroids_gau1$hexID, df_b$hb_id),] |>
-  dplyr::select(-hb_id) |>
+  dplyr::select(-hb_id)
+
+# Apply the scaling
+df_model_data <- bind_rows(data_gau, df_b)
+scaled_gau <- scale_data_manual(df_model_data, "type") |>
+  as_tibble()
+
+scaled_gau_data <- scaled_gau |>
+  filter(type == "data") |>
   select(-type)
 
-# Apply the scaling
-scaled_gau_data_model <- scale_data_manual(df_b) |>
-  as_tibble()
+scaled_gau_data_model <- scaled_gau |>
+  filter(type == "model") |>
+  select(-type)
 
@@
 ## Second projection
+# projection <- cbind(
+#     c(-0.71895,0.29250,-0.29905,-0.01656),
+#     c(-0.16698,-0.62707,-0.23756,0.46328))
 projection <- cbind(
-    c(-0.71895,0.29250,-0.29905,-0.01656),
-    c(-0.16698,-0.62707,-0.23756,0.46328))
+  c(0.36030,0.00630,0.66545,-0.34307),
+  c(-0.01861,0.14409,-0.36798,-0.73066))
 
@@
 df_b <- df_b[match(df_bin_centroids_gau1$hexID, df_b$hb_id),] |>
-  dplyr::select(-hb_id)|>
-  select(-type)
-
-# Apply the scaling
-scaled_gau_data_model <- scale_data_manual(df_b) |>
-  as_tibble()
+  dplyr::select(-hb_id)
 
@@
 # Apply the scaling
-scaled_gau_data_model <- scale_data_manual(df_b) |>
+df_model_data <- bind_rows(data_gau, df_b)
+scaled_gau <- scale_data_manual(df_model_data, "type") |>
   as_tibble()
+
+scaled_gau_data <- scaled_gau |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_gau_data_model <- scaled_gau |>
+  filter(type == "model") |>
+  select(-type)
 
@@
 df_b <- df_b[match(df_bin_centroids_gau1$hexID, df_b$hb_id),] |>
-  dplyr::select(-hb_id)|>
+  dplyr::select(-hb_id)
+
+# Apply the scaling
+df_model_data <- bind_rows(data_gau, df_b)
+scaled_gau <- scale_data_manual(df_model_data, "type") |>
+  as_tibble()
+
+scaled_gau_data <- scaled_gau |>
+  filter(type == "data") |>
   select(-type)
 
-# Apply the scaling
-scaled_gau_data_model <- scale_data_manual(df_b) |>
-  as_tibble()
+scaled_gau_data_model <- scaled_gau |>
+  filter(type == "model") |>
+  select(-type)
 
@@
 ### Second projection
+# projection <- cbind(
+#     c(-0.60995,0.33361,0.32920,-0.31518),
+#     c(0.19525,-0.30554,0.74417,0.07600))
 projection <- cbind(
-    c(-0.60995,0.33361,0.32920,-0.31518),
-    c(0.19525,-0.30554,0.74417,0.07600))
+  c(0.36030,0.00630,0.66545,-0.34307),
+  c(-0.01861,0.14409,-0.36798,-0.73066))
 
@@
   limits = 0.3,
-  axis_pos_x = -0.05,
-  axis_pos_y = -0.4)
+  axis_pos_x = -0.7,
+  axis_pos_y = -0.7)
 
@@
   coord_fixed() +
-  xlim(c(-0.1, 0.2)) +
-  ylim(c(-0.45, 0)) +
+  xlim(c(-0.75, -0.35)) +
+  ylim(c(-0.75, -0.4))+
   interior_annotation("c2", 
@@
 
-# Apply the scaling
-scaled_pbmc_data <- scale_data_manual(training_data_pbmc |> select(-ID)) |>
-  as_tibble()
+data_pbmc <- training_data_pbmc |> 
+  select(-ID) |>
+  mutate(type = "data")
 
@@
 df_b_pbmc <- df_b_pbmc[match(df_bin_centroids_pbmc$hexID, df_b_pbmc$hb_id),] |>
-  dplyr::select(-hb_id) |>
+  dplyr::select(-hb_id)
+
+# Apply the scaling
+df_model_data_pbmc <- bind_rows(data_pbmc, df_b_pbmc)
+scaled_pbmc <- scale_data_manual(df_model_data_pbmc, "type") |>
+  as_tibble()
+
+scaled_pbmc_data <- scaled_pbmc |>
+  filter(type == "data") |>
   select(-type)
 
-# Apply the scaling
-scaled_pbmc_data_model <- scale_data_manual(df_b_pbmc) |>
-  as_tibble()
+scaled_pbmc_data_model <- scaled_pbmc |>
+  filter(type == "model") |>
+  select(-type)
 
@@
 
-# Apply the scaling
-scaled_pbmc_data <- scale_data_manual(training_data_pbmc |> select(-ID)) |>
-  as_tibble()
-
 df_b_pbmc <- df_bin_pbmc |>
@@
 
+<!--add different NLDR layouts-->
+
 The MNIST dataset consists of $70000$ grayscale images of handwritten digits [@lecun2010].
@@
   mutate(ID = 1:NROW(training_data_mnist))
+
+data_mnist <- training_data_mnist |>
+  select(-ID) |>
+  mutate(type = "data")
 
@@
 #| label: mnist-model-proj
-
-# Apply the scaling
-scaled_mnist_data <- scale_data_manual(training_data_mnist |> select(-ID)) |>
-  as_tibble()
-
 df_b_mnist <- df_bin_mnist |>
@@
 df_b_mnist <- df_b_mnist[match(df_bin_centroids_mnist$hexID, df_b_mnist$hb_id),] |>
-  dplyr::select(-hb_id) |>
+  dplyr::select(-hb_id)
+
+# Apply the scaling
+df_model_data_mnist <- bind_rows(data_mnist, df_b_mnist)
+scaled_mnist <- scale_data_manual(df_model_data_mnist, "type") |>
+  as_tibble()
+
+scaled_mnist_data <- scaled_mnist |>
+  filter(type == "data") |>
   select(-type)
 
-# Apply the scaling
-scaled_mnist_data_model <- scale_data_manual(df_b_mnist) |>
-  as_tibble()
+scaled_mnist_data_model <- scaled_mnist |>
+  filter(type == "model") |>
+  select(-type)
 

### 2024-08-30 paper.qmd
Change the projections for Five Gaussian clusters

@@
 ## First projection
+# projection <- cbind(
+#     c(0.46477,-0.02024,0.56378,0.39736),
+#     c(0.15607,-0.54876,-0.44090,0.41505))
 projection <- cbind(
-    c(0.46477,-0.02024,0.56378,0.39736),
-    c(0.15607,-0.54876,-0.44090,0.41505))
+  c(-0.00215,-0.68905,-0.04778,-0.54223),
+  c(0.42558,-0.23854,-0.63659,0.35753))
 
@@
   proj_data = projected_df,
-  limits = 0.3,
-  axis_pos_x = -0.4,
-  axis_pos_y = -0.1)
+  limits = 0.2,
+  axis_pos_x = -0.12,
+  axis_pos_y = -0.12)
 
@@
   coord_fixed() +
-  xlim(c(-0.45, -0.2)) +
-  ylim(c(-0.15, 0.18)) +
+  xlim(c(-0.41, -0.15)) +
+  ylim(c(0.18, 0.52)) +
   interior_annotation("a1", 
@@
   limits = 0.3,
-  axis_pos_x = 0.1,
-  axis_pos_y = -0.1)
+  axis_pos_x = -0.7,
+  axis_pos_y = -0.7)
 
@@
     size = 0.5,
-    alpha = 0.3) +
+    alpha = 0.2) +
   scale_color_manual(values = c("#000000", "#33a02c")) +
@@
   coord_fixed() +
-  xlim(c(0.05, 0.25)) +
-  ylim(c(-0.15, 0.25)) +
+  xlim(c(-0.75, -0.35)) +
+  ylim(c(-0.75, -0.4)) +
   interior_annotation("a2", 
@@
 ## First projection
+# projection <- cbind(
+#   c(0.02743,0.21427,0.62893,-0.49816),
+#   c(-0.70709,0.42156,-0.02593,0.10965))
 projection <- cbind(
-  c(0.02743,0.21427,0.62893,-0.49816),
-  c(-0.70709,0.42156,-0.02593,0.10965))
+  c(-0.00215,-0.68905,-0.04778,-0.54223),
+  c(0.42558,-0.23854,-0.63659,0.35753))
 
@@
   proj_data = projected_df,
-  limits = 0.3,
-  axis_pos_x = -0.2,
-  axis_pos_y = -0.85)
+  limits = 0.2,
+  axis_pos_x = 0.45,
+  axis_pos_y = -0.28)
 
@@
   coord_fixed() +
-  xlim(c(-0.25, 0.1)) +
-  ylim(c(-0.9, -0.55)) +
+  xlim(c(-0.41, -0.15)) +
+  ylim(c(0.18, 0.52)) +
   interior_annotation("b1", 
@@
   limits = 0.3,
-  axis_pos_x = 0.08,
-  axis_pos_y = -0.08)
+  axis_pos_x = -0.7,
+  axis_pos_y = -0.7)
 
@@
   coord_fixed() +
-  xlim(c(0, 0.4)) +
-  ylim(c(-0.15, 0.3)) +
+  xlim(c(-0.75, -0.35)) +
+  ylim(c(-0.75, -0.4)) +
   interior_annotation("b2", 
@@
 projection <- cbind(
-    c(0.21889,-0.25701,-0.19529,-0.73412),
-    c(-0.63168,0.27992,-0.42895,-0.17223))
+  c(-0.00215,-0.68905,-0.04778,-0.54223),
+  c(0.42558,-0.23854,-0.63659,0.35753))
 
@@
   proj_data = projected_df,
-  limits = 0.3,
-  axis_pos_x = 0.1,
-  axis_pos_y = 0.05)
+  limits = 0.2,
+  axis_pos_x = -0.37,
+  axis_pos_y = 0.22)
 
@@
   coord_fixed() +
-  xlim(c(0.05, 0.35)) +
-  ylim(c(0, 0.35)) +
+  xlim(c(-0.41, -0.15)) +
+  ylim(c(0.18, 0.52)) +
   interior_annotation("c1", 
@@
     size = 0.8,
-    alpha = 0.3) +
+    alpha = 0.15) +
   scale_color_manual(values = c("#000000", "#33a02c")) +
@@
 #| label: fig-five-gau-nldr-layouts
-#| fig-height: 5
+#| fig-height: 10
 #| fig-width: 15

### 2024-08-30 paper.qmd
Updated axes
Fixed size of Figure 11

@@
   limits = 0.2,
-  axis_pos_x = -0.12,
-  axis_pos_y = -0.12)
+  axis_pos_x = -0.37,
+  axis_pos_y = 0.22)
 
@@
   limits = 0.2,
-  axis_pos_x = 0.45,
-  axis_pos_y = -0.28)
+  axis_pos_x = -0.37,
+  axis_pos_y = 0.22)
 
@@
 #| label: fig-five-gau-nldr-layouts
-#| fig-height: 10
+#| fig-height: 5
 #| fig-width: 15
@@
 #| label: fig-five-gau-projs
-#| fig-height: 15
-#| fig-width: 10
+#| fig-height: 10
+#| fig-width: 15
 #| fig-pos: H

### 2024-09-02 paper.qmd
Updated projections

@@
   limits = 0.2,
-  axis_pos_x = -0.37,
-  axis_pos_y = 0.22)
+  axis_pos_x = 0.15,
+  axis_pos_y = -0.85)
 
@@
   coord_fixed() +
-  xlim(c(-0.41, -0.15)) +
-  ylim(c(0.18, 0.52)) +
+  xlim(c(0.1, 0.38)) +
+  ylim(c(-0.9, -0.5)) +
   interior_annotation("a1", 

### 2024-09-02 paper.qmd
Updated Five Gaussian cluster scripts with sclaed data problem fixed

@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 44
+num_bins_x_gau1 <- 45
 lim1 <- gau1_scaled_obj$lim1
@@
   limits = 0.2,
-  axis_pos_x = -0.37,
-  axis_pos_y = 0.22)
+  axis_pos_x = -0.55,
+  axis_pos_y = -0.05)
 
@@
   coord_fixed() +
-  xlim(c(-0.41, -0.15)) +
-  ylim(c(0.18, 0.52)) +
+  xlim(c(-0.6, -0.3)) +
+  ylim(c(0, -0.38)) +
   interior_annotation("b1", 

### 2024-09-04 paper.qmd
Found the best UMAP view which contain three clusters as well for PBMC3k

@@
 # remotes::install_github("jlmelville/snedata")
-library(snedata)
-library(ggflowchart)
-library(purrr) ## map function
-library(gridExtra) ## for grid.arrange
-library(rsample)
-library(DT)
+# library(snedata)
+# library(ggflowchart)
+#library(purrr) ## map function
+#library(gridExtra) ## for grid.arrange
+# library(rsample)
+# library(DT)
 library(ggbeeswarm)
@@
 
-umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_15_min_dist_0.99.rds")
+#umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_15_min_dist_0.99.rds")
+umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_12_min_dist_0.99.rds")
 nldr3 <- umap_pbmc |>

### 2024-09-04 paper.qmd
Updated Figure 14 (only figures)

@@
   coord_fixed() +
-  interior_annotation("a")
+  interior_annotation("a1")
 
@@
 
-<!--bin1 = 30, bin2 = 29, b = 870, non_empty = 135-->
-```{r}
-#| echo: false
-#| fig-cap: "(a) Model generated in $2\\text{-}D$ with UMAP, and (b) $p\\text{-}D$ model error in $2\\text{-}D$.
-The $2\\text{-}D$ model shows three well-separated distant clusters.
-The $p\\text{-}D$ model errors are distributed along clusters."
-#| label: fig-model-pbmc-author
-#| fig-pos: H
-#| eval: false
-
-trimesh_removed_pbmc + error_plot_pbmc + 
-  plot_layout(guides='collect', ncol=2) &
-  theme(legend.position='bottom')
-```
-
 ```{r}
@@
     c(-0.67212,0.02745,-0.26718,-0.33126,-0.34919,0.20839,-0.14498,-0.37708,0.15455))
-
-projected <- as.matrix(scaled_pbmc_data) %*% projection
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  #dplyr::mutate(type = df_exe$type) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  distance_df_small_edges_pbmc |> select(-distance), 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-axes_obj <- gen_axes(
-  data = scaled_pbmc_data, 
-  proj_data = projected_df,  
-  limits = 1.5,
-  axis_pos_x = -0.75,
-  axis_pos_y = -0.75)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
-
-pbmc_proj_tsne_model1 <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#33a02c") +
-  geom_point(
-    size = 0.5,
-    alpha = 0.2) +
-  scale_color_manual(values = c("#000000", "#33a02c")) +
-  geom_segment(
-    data=axes, 
-    aes(x=x1, y=y1, xend=x2, yend=y2), 
-    colour="grey70") +
-  geom_text(
-    data=axes, 
-    aes(x=x2, y=y2, label=rownames(axes)), 
-    colour="grey50",
-    size = 3) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-1, 1)) +
-  ylim(c(-1, 1)) +
-  interior_annotation("b")
-
-## Second projection
-projection <- cbind(
-    c(0.78689,-0.14774,0.36527,0.09617,0.12097,-0.00033,0.36973,-0.20959,-0.08438),
-    c(0.18390,-0.43121,-0.30398,0.05861,0.55282,-0.14058,-0.27048,0.44386,-0.27372))
-
-projected <- as.matrix(scaled_pbmc_data) %*% projection
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  #dplyr::mutate(type = df_exe$type) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  distance_df_small_edges_pbmc |> select(-distance), 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-axes_obj <- gen_axes(
-  data = scaled_pbmc_data, 
-  proj_data = projected_df,  
-  limits = 1.5,
-  axis_pos_x = -0.75,
-  axis_pos_y = -0.75)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
-
-pbmc_proj_tsne_model2 <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#33a02c") +
-  geom_point(
-    size = 0.5,
-    alpha = 0.2) +
-  scale_color_manual(values = c("#000000", "#33a02c")) +
-  geom_segment(
-    data=axes, 
-    aes(x=x1, y=y1, xend=x2, yend=y2), 
-    colour="grey70") +
-  geom_text(
-    data=axes, 
-    aes(x=x2, y=y2, label=rownames(axes)), 
-    colour="grey50",
-    size = 3) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-1, 1)) +
-  ylim(c(-1, 1)) +
-  interior_annotation("c")
-
-## Third projection
-projection <- cbind(
-    c(0.00067,0.53963,-0.16241,-0.09777,0.32766,0.44543,-0.51359,-0.29884,0.00951),
-    c(0.62399,0.26016,0.06441,-0.34686,-0.40466,0.24643,0.23537,0.07972,0.34402))
-
-projected <- as.matrix(scaled_pbmc_data) %*% projection
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  #dplyr::mutate(type = df_exe$type) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  distance_df_small_edges_pbmc |> select(-distance), 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-axes_obj <- gen_axes(
-  data = scaled_pbmc_data, 
-  proj_data = projected_df,  
-  limits = 1.5,
-  axis_pos_x = -0.75,
-  axis_pos_y = -0.75)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
-
-pbmc_proj_tsne_model3 <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#33a02c") +
-  geom_point(
-    size = 0.5,
-    alpha = 0.2) +
-  scale_color_manual(values = c("#000000", "#33a02c")) +
-  geom_segment(
-    data=axes, 
-    aes(x=x1, y=y1, xend=x2, yend=y2), 
-    colour="grey70") +
-  geom_text(
-    data=axes, 
-    aes(x=x2, y=y2, label=rownames(axes)), 
-    colour="grey50",
-    size = 3) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-1, 1)) +
-  ylim(c(-1, 1)) +
-  interior_annotation("d")
-
-```
-
-```{r}
-#| label: fig-model-pbmc-author-proj
-#| fig-pos: H
-#| fig-height: 10
-#| fig-width: 10
-#| fig-cap: "Model in \\gD{}, on the UMAP layout (a), and three views of the fit in projections from $9\\text{-}D$ (b, c, d), for the PBMC3k data ($a_1 = 0.046$, $b = 676 \\  (26, \ 26)$, $m = 92$, benchmark value to remove large edges is $0.08$).
-The three well-separated clusters observed in the \\gD{} layout of UMAP (@fig-NLDR-variety (a)) is also visible when visualizing the model overlaid on the data space.
-But the separation between the clusters are small in $9\\text{-}D$.
-The model shows non-linear clusters and densed points within clusters in $9\\text{-}D$ space, which are some additional patterns not visible in the \\gD{} representation (@fig-NLDR-variety (a)).
-The full video is available at <https://youtu.be/VqqWuE0Jj6A>."
-
-free(trimesh_removed_pbmc) + pbmc_proj_tsne_model1 +
-  pbmc_proj_tsne_model2 + pbmc_proj_tsne_model3 +
-  plot_layout(ncol = 2)
-```
-
-<!-- ::: {#fig-pbmc1-sc layout-ncol="4" fig-pos="H"} -->
-<!-- ![](figures/pbmc3k/umap_trimesh_plot.png) -->
-
-<!-- ![](figures/pbmc3k/sc_1.png) -->
-
-<!-- ![](figures/pbmc3k/sc_2.png) -->
-
-<!-- ![](figures/pbmc3k/sc_3.png) -->
-
-<!-- Model in \gD{}, on the UMAP layout, and three views of the fit in projections from $9\text{-}D$, for the PBMC3k data ($(s_1, \ s_2) = (-0.050, \ -0.041)$, $b = 870 \  (30, \ 29)$, $m = 135$, benchmark value to remove large edges is $0.099$) (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/VqqWuE0Jj6A>).
--->
-<!-- ::: -->
-<!--compute absolute error for different parameter choices-->
-
-```{r}
-#| label: combine-data-pbmc
-
-error_pbmc_umap <- read_rds("data/pbmc3k/error_pbmc_umap_30_min_dist_0.3.rds")
-error_pbmc_umap2 <- read_rds("data/pbmc3k/error_pbmc_umap_5_min_dist_0.01.rds")
-# error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_15_min_dist_0.99.rds")
-error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_12_min_dist_0.99.rds")
-error_pbmc_tsne <- read_rds("data/pbmc3k/error_pbmc_tsne_5.rds")
-error_pbmc_tsne2 <- read_rds("data/pbmc3k/error_pbmc_tsne_30.rds")
-error_pbmc_phate <- read_rds("data/pbmc3k/error_pbmc_phate_5.rds")
-error_pbmc_trimap <- read_rds("data/pbmc3k/error_pbmc_trimap_12_4_3.rds")
-error_pbmc_pacmap <- read_rds("data/pbmc3k/error_pbmc_pacmap_30_random_0.9_5.rds")
-
-error_pbmc <- bind_rows(error_pbmc_umap, 
-                        error_pbmc_umap2,
-                        error_pbmc_umap3,
-                        error_pbmc_tsne,
-                        error_pbmc_tsne2,
-                        error_pbmc_phate,
-                        error_pbmc_trimap,
-                        error_pbmc_pacmap)
-
-error_pbmc <- error_pbmc |>
-  mutate(a1 = round(a1, 2)) |>
-  filter(bin1 >= 5) |>
-  group_by(method, a1) |>
-  filter(MSE == min(MSE)) |>
-  ungroup()
-```
-
-```{r}
-#| label: tsne-pbmc
-# read the png file from device
-
-#pbmc_layouts <- readPNG("figures/pbmc3k/pbmc_nldr_layouts.png", native = TRUE)
-
-error_plot_pbmc <- ggplot(error_pbmc, 
-                          aes(x = a1, 
-                              y = MSE, 
-                              colour = method)) + 
-  geom_point(size = 0.8) +
-  geom_line(linewidth = 0.3) + 
-  # geom_vline(xintercept = 15, linetype="solid", 
-  #            color = "black", linewidth=0.8, alpha = 0.5) +
-  scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 13, 17, 21, 26)]) +
-  scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a','#ff7f00',
-                              '#984ea3','#999999','#a65628','#f781bf')) +
-  scale_y_log10() +
-  ylab("log(MSE)") +
-  xlab(expression(paste("binwidth (", a[1], ")"))) +
-  theme_minimal() +
-  theme(panel.border = element_rect(fill = 'transparent'),
-        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
-        axis.ticks.x = element_line(),
-        axis.ticks.y = element_line(),
-        legend.position = "none",
-        axis.text.x = element_text(size = 7),
-        axis.text.y = element_text(size = 7),
-        axis.title.x = element_text(size = 7),
-        axis.title.y = element_text(size = 7))
-
-```
-
-```{r}
-#| fig-cap: "Assessing which of the 8 NLDR layouts on the PBMC3k data  (shown in Figure 1) is the better representation using MSE for varying binwidth ($a_1$).
-Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-h).
-Layout f is universally poor.
-Layouts a, b, g, h that show large separations between clusters are universally suboptimal.
-Layout d with little separation performs well at tiny binwidth (where most points are in their own bin) and poorly as binwidth increases.
-The choice of best is between layouts c and e, that have small separations between oddly shaped clusters.
-Layout e is the best choice."
-#| label: fig-pbmc-mse
-#| fig-pos: H
-#| out-height: 100%
-
-free(error_plot_pbmc) + wrap_plots(nldr1c, nldr2c, nldr3c, nldr4c,
-           nldr5c, nldr6c, nldr7c, nldr8c, ncol = 2)
-```
-
-<!--best choice-->
-
-```{r}
-#| label: tsne-pbmc-read
-tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_30.rds")
-pbmc_scaled_obj <- gen_scaled_data(
-  data = tsne_pbmc)
-tsne_pbmc_scaled <- pbmc_scaled_obj$scaled_nldr
-
-# tsne_pbmc <- tsne_pbmc_scaled |>
-#   ggplot(aes(x = tSNE1,
-#              y = tSNE2)) +
-#   geom_point(alpha=0.3, color = "#000000")
-```
-
-<!--tSNE with perplexity:30-->
-```{r}
-#| echo: false
-#| fig-cap: "The $2\\text{-}D$ layout from tSNE applied for the PBMC3k dataset.
-The default (hyper-)parameter setting is a perplexity $30$.
-We use our model-in-the-data-space to assess whether this is an accurate representation of structure present in the high-dimensional data, or if it is misleading."
-#| label: fig-tsne-suggest
-#| fig-pos: H
-#| out-width: 40%
-#| eval: false
-
-tsne_pbmc 
-```
-
-<!-- We then fit the model for tSNE, and visualize the resultant model in the $9\text{-}D$ data space.
-The model shows a quirk, as shown in @fig-pbmc2-sc.
-All three clusters are connected by an edge except the small and large clusters.
-Because the clusters are so close in \gD{}, they attempt to maintain the structure in \pD{} as well.
-This is evident that tSNE with perplexity $30$ provides a reasonable representation of PBMC3k data.
--->
-
-<!--Fit the best model and compute error-->
-```{r}
-#| label: num-bins-pbmc
-## Compute hexbin parameters
-num_bins_x_pbmc <- 13
-lim1 <- pbmc_scaled_obj$lim1
-lim2 <- pbmc_scaled_obj$lim2
-r2_pbmc <- diff(lim2)/diff(lim1)
-
-pbmc_model <- fit_highd_model(
-  training_data = training_data_pbmc,
-  emb_df = tsne_pbmc_scaled,
-  bin1 = num_bins_x_pbmc,
-  r2 = r2_pbmc,
-  q = 0.1,
-  is_bin_centroid = TRUE,
-  is_rm_lwd_hex = FALSE,
-  col_start_highd = "PC"
-)
-
-df_bin_centroids_pbmc <- pbmc_model$df_bin_centroids
-df_bin_pbmc <- pbmc_model$df_bin
-
-## Triangulate bin centroids
-tr1_object_pbmc <- tri_bin_centroids(
-  df_bin_centroids_pbmc, x = "c_x", y = "c_y")
-tr_from_to_df_pbmc <- gen_edges(
-  tri_object = tr1_object_pbmc)
-
-# tr_from_to_df_pbmc <- tr_from_to_df_pbmc |>
-#   filter(!(row_number() %in% c(65, 112, 130)))
-
-## Compute 2D distances
-distance_pbmc <- cal_2d_dist(
-  tr_coord_df = tr_from_to_df_pbmc,
-  start_x = "x_from",
-  start_y = "y_from",
-  end_x = "x_to",
-  end_y = "y_to",
-  select_vars = c("from", "to", "distance"))
-
-## To find the benchmark value
-benchmark_pbmc <- find_lg_benchmark(
-  distance_edges = distance_pbmc,
-  distance_col = "distance")
-
-benchmark_pbmc <- 0.1
-
-## Remove an edge that exist as long edge in p-D
-# tr_from_to_df_pbmc |> mutate(ID = row_number()) |> filter(from == 37) |> filter(to == 43)
-
-tr_df <- distinct(tibble::tibble(
-  x = c(tr_from_to_df_pbmc[["x_from"]], tr_from_to_df_pbmc[["x_to"]]), 
-  y = c(tr_from_to_df_pbmc[["y_from"]], tr_from_to_df_pbmc[["y_to"]])))
-
-distance_df_small_edges_pbmc <- distance_pbmc |>
-  filter(distance < benchmark_pbmc)
-
-tr_from_to_df_pbmc <- inner_join(
-  tr_from_to_df_pbmc, distance_df_small_edges_pbmc, 
-  by = c("from", "to"))
-
-tr_from_to_df_pbmc <- tr_from_to_df_pbmc |>
-  filter(!(row_number() %in% c(155)))
-
-trimesh_removed_pbmc <- ggplot() + 
-  geom_segment(data = tr_from_to_df_pbmc, 
-               aes(
-                 x = x_from, 
-                 y = y_from, 
-                 xend = x_to, 
-                 yend = y_to),
-               colour = "#33a02c") +
-  geom_point(
-    data = tsne_pbmc_scaled,
-    aes(
-      x = tSNE1,
-      y = tSNE2
-    ),
-    alpha = 0.2,
-    color = "#000000"
-  )  +
-  coord_fixed() +
-  interior_annotation("a")
-
-# trimesh_removed_pbmc <- vis_rmlg_mesh(
-#   distance_edges = distance_pbmc,
-#   benchmark_value = benchmark_pbmc,
-#   tr_coord_df = tr_from_to_df_pbmc,
-#   distance_col = "distance") +
-#   geom_point(
-#     data = tsne_pbmc_scaled,
-#     aes(
-#       x = tSNE1,
-#       y = tSNE2
-#     ),
-#     alpha = 0.2,
-#     color = "#000000"
-#   ) 
-#   #xlim(sc_xlims) + ylim(sc_ylims) +
-#   #interior_annotation("a", sc_ltr_pos)
-
-## Compute error
-error_df <- augment(
-  df_bin_centroids = df_bin_centroids_pbmc,
-  df_bin = df_bin_pbmc,
-  training_data = training_data_pbmc,
-  newdata = NULL,
-  type_NLDR = "tSNE",
-  col_start = "PC")
-
-## Categorize error
-
-error_df <- error_df |>
-  mutate(type = case_when(
-    row_wise_abs_error <= 5 ~ "error 0-5",
-    row_wise_abs_error <= 10 ~ "error 5-10",
-    row_wise_abs_error <= 15 ~ "error 10-15",
-    row_wise_abs_error <= 20 ~ "error 15-20",
-    row_wise_abs_error <= 25 ~ "error 20-25",
-    .default = "error greter than 25"
-  )) |>
-  mutate(type = factor(type, levels = c(
-    "error 0-5", "error 5-10", "error 10-15", "error 15-20", "error 20-25", 
-    "error greter than 25")))
-
-## To join embedding
-error_df <- error_df |>
-  bind_cols(tsne_pbmc_scaled |>
-              select(-ID))
-
-error_plot_pbmc <- error_df |>
-  ggplot(aes(x = tSNE1,
-             y = tSNE2,
-             color = type,
-             group = ID)) +
-  geom_point(alpha=0.5,
-             size = 0.1) +
-  interior_annotation("b", sc_ltr_pos)
-
-```
-
-```{r}
-#| label: pbmc-tsne-model-proj
-
-df_b_pbmc <- df_bin_pbmc |>
-  dplyr::filter(hb_id %in% df_bin_centroids_pbmc$hexID) |>
-  dplyr::mutate(type = "model") ## Data with summarized mean
-
-## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
-df_b_pbmc <- df_b_pbmc[match(df_bin_centroids_pbmc$hexID, df_b_pbmc$hb_id),] |>
-  dplyr::select(-hb_id)
-
-# Apply the scaling
-df_model_data_pbmc <- bind_rows(data_pbmc, df_b_pbmc)
-scaled_pbmc <- scale_data_manual(df_model_data_pbmc, "type") |>
-  as_tibble()
-
-scaled_pbmc_data <- scaled_pbmc |>
-  filter(type == "data") |>
-  select(-type)
-
-scaled_pbmc_data_model <- scaled_pbmc |>
-  filter(type == "model") |>
-  select(-type)
-
-# Combine with the true model for visualization
-df <- dplyr::bind_rows(scaled_pbmc_data_model |> mutate(type = "model"),
-                       scaled_pbmc_data |> mutate(type = "data"))
-
-## Removed long edge in high-d
-distance_df_small_edges_pbmc <- distance_df_small_edges_pbmc |>
-  filter(!(row_number() %in% c(155)))
-
-## First projection
-projection <- cbind(
-    c(0.40456,-0.55243,-0.58434,0.05063,-0.20540,-0.10469,0.28850,0.12209,0.17017),
-    c(0.34225,0.53615,-0.05822,-0.01993,-0.55480,0.21337,0.34100,-0.07280,-0.33139))
 
@@
   ylim(c(-1, 1)) +
-  interior_annotation("b")
+  interior_annotation("a2")
 
@@
 projection <- cbind(
-    c(0.07154,-0.44483,-0.04593,0.63343,0.21449,-0.39592,-0.29762,0.01726,0.30845),
-    c(0.87788,-0.21837,0.11780,-0.17226,-0.17920,-0.00829,0.11424,0.28681,0.06091))
+    c(0.00067,0.53963,-0.16241,-0.09777,0.32766,0.44543,-0.51359,-0.29884,0.00951),
+    c(0.62399,0.26016,0.06441,-0.34686,-0.40466,0.24643,0.23537,0.07972,0.34402))
 
@@
   ylim(c(-1, 1)) +
-  interior_annotation("c")
-
-## Third projection
+  interior_annotation("a3")
+
+```
+
+<!-- ::: {#fig-pbmc1-sc layout-ncol="4" fig-pos="H"} -->
+<!-- ![](figures/pbmc3k/umap_trimesh_plot.png) -->
+
+<!-- ![](figures/pbmc3k/sc_1.png) -->
+
+<!-- ![](figures/pbmc3k/sc_2.png) -->
+
+<!-- ![](figures/pbmc3k/sc_3.png) -->
+
+<!-- Model in \gD{}, on the UMAP layout, and three views of the fit in projections from $9\text{-}D$, for the PBMC3k data ($(s_1, \ s_2) = (-0.050, \ -0.041)$, $b = 870 \  (30, \ 29)$, $m = 135$, benchmark value to remove large edges is $0.099$) (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/VqqWuE0Jj6A>).
+-->
+<!-- ::: -->
+<!--compute absolute error for different parameter choices-->
+
+```{r}
+#| label: combine-data-pbmc
+
+error_pbmc_umap <- read_rds("data/pbmc3k/error_pbmc_umap_30_min_dist_0.3.rds")
+error_pbmc_umap2 <- read_rds("data/pbmc3k/error_pbmc_umap_5_min_dist_0.01.rds")
+# error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_15_min_dist_0.99.rds")
+error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_12_min_dist_0.99.rds")
+error_pbmc_tsne <- read_rds("data/pbmc3k/error_pbmc_tsne_5.rds")
+error_pbmc_tsne2 <- read_rds("data/pbmc3k/error_pbmc_tsne_30.rds")
+error_pbmc_phate <- read_rds("data/pbmc3k/error_pbmc_phate_5.rds")
+error_pbmc_trimap <- read_rds("data/pbmc3k/error_pbmc_trimap_12_4_3.rds")
+error_pbmc_pacmap <- read_rds("data/pbmc3k/error_pbmc_pacmap_30_random_0.9_5.rds")
+
+error_pbmc <- bind_rows(error_pbmc_umap, 
+                        error_pbmc_umap2,
+                        error_pbmc_umap3,
+                        error_pbmc_tsne,
+                        error_pbmc_tsne2,
+                        error_pbmc_phate,
+                        error_pbmc_trimap,
+                        error_pbmc_pacmap)
+
+error_pbmc <- error_pbmc |>
+  mutate(a1 = round(a1, 2)) |>
+  filter(bin1 >= 5) |>
+  group_by(method, a1) |>
+  filter(MSE == min(MSE)) |>
+  ungroup()
+```
+
+```{r}
+#| label: error-comp-pbmc
+# read the png file from device
+
+#pbmc_layouts <- readPNG("figures/pbmc3k/pbmc_nldr_layouts.png", native = TRUE)
+
+error_plot_pbmc <- ggplot(error_pbmc, 
+                          aes(x = a1, 
+                              y = MSE, 
+                              colour = method)) + 
+  geom_point(size = 0.8) +
+  geom_line(linewidth = 0.3) + 
+  # geom_vline(xintercept = 15, linetype="solid", 
+  #            color = "black", linewidth=0.8, alpha = 0.5) +
+  scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 13, 17, 21, 26)]) +
+  scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a','#ff7f00',
+                              '#984ea3','#999999','#a65628','#f781bf')) +
+  scale_y_log10() +
+  ylab("log(MSE)") +
+  xlab(expression(paste("binwidth (", a[1], ")"))) +
+  theme_minimal() +
+  theme(panel.border = element_rect(fill = 'transparent'),
+        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+        axis.ticks.x = element_line(),
+        axis.ticks.y = element_line(),
+        legend.position = "none",
+        axis.text.x = element_text(size = 7),
+        axis.text.y = element_text(size = 7),
+        axis.title.x = element_text(size = 7),
+        axis.title.y = element_text(size = 7))
+
+```
+
+```{r}
+#| fig-cap: "Assessing which of the 8 NLDR layouts on the PBMC3k data  (shown in Figure 1) is the better representation using MSE for varying binwidth ($a_1$).
+Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-h).
+Layout f is universally poor.
+Layouts a, b, g, h that show large separations between clusters are universally suboptimal.
+Layout d with little separation performs well at tiny binwidth (where most points are in their own bin) and poorly as binwidth increases.
+The choice of best is between layouts c and e, that have small separations between oddly shaped clusters.
+Layout e is the best choice."
+#| label: fig-pbmc-mse
+#| fig-pos: H
+#| out-height: 100%
+
+free(error_plot_pbmc) + wrap_plots(nldr1c, nldr2c, nldr3c, nldr4c,
+           nldr5c, nldr6c, nldr7c, nldr8c, ncol = 2)
+```
+
+<!--best choice-->
+
+```{r}
+#| label: umap-pbmc-read-best
+umap_pbmc2 <- read_rds("data/pbmc3k/pbmc_umap_12_min_dist_0.99.rds")
+pbmc_scaled_obj <- gen_scaled_data(
+  data = umap_pbmc2)
+umap_pbmc_scaled_best <- pbmc_scaled_obj$scaled_nldr
+
+# tsne_pbmc <- tsne_pbmc_scaled |>
+#   ggplot(aes(x = tSNE1,
+#              y = tSNE2)) +
+#   geom_point(alpha=0.3, color = "#000000")
+```
+
+<!-- We then fit the model for tSNE, and visualize the resultant model in the $9\text{-}D$ data space.
+The model shows a quirk, as shown in @fig-pbmc2-sc.
+All three clusters are connected by an edge except the small and large clusters.
+Because the clusters are so close in \gD{}, they attempt to maintain the structure in \pD{} as well.
+This is evident that tSNE with perplexity $30$ provides a reasonable representation of PBMC3k data.
+-->
+
+<!--Fit the best model and compute error-->
+```{r}
+#| label: num-bins-pbmc
+## Compute hexbin parameters
+num_bins_x_pbmc <- 13
+lim1 <- pbmc_scaled_obj$lim1
+lim2 <- pbmc_scaled_obj$lim2
+r2_pbmc <- diff(lim2)/diff(lim1)
+
+pbmc_model <- fit_highd_model(
+  training_data = training_data_pbmc,
+  emb_df = umap_pbmc_scaled_best,
+  bin1 = num_bins_x_pbmc,
+  r2 = r2_pbmc,
+  q = 0.1,
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = FALSE,
+  col_start_highd = "PC"
+)
+
+df_bin_centroids_pbmc <- pbmc_model$df_bin_centroids
+df_bin_pbmc <- pbmc_model$df_bin
+
+## Triangulate bin centroids
+tr1_object_pbmc <- tri_bin_centroids(
+  df_bin_centroids_pbmc, x = "c_x", y = "c_y")
+tr_from_to_df_pbmc <- gen_edges(
+  tri_object = tr1_object_pbmc)
+
+# tr_from_to_df_pbmc <- tr_from_to_df_pbmc |>
+#   filter(!(row_number() %in% c(65, 112, 130)))
+
+## Compute 2D distances
+distance_pbmc <- cal_2d_dist(
+  tr_coord_df = tr_from_to_df_pbmc,
+  start_x = "x_from",
+  start_y = "y_from",
+  end_x = "x_to",
+  end_y = "y_to",
+  select_vars = c("from", "to", "distance"))
+
+## To find the benchmark value
+benchmark_pbmc <- find_lg_benchmark(
+  distance_edges = distance_pbmc,
+  distance_col = "distance")
+
+#benchmark_pbmc <- 0.1
+
+## Remove an edge that exist as long edge in p-D
+# tr_from_to_df_pbmc |> mutate(ID = row_number()) |> filter(from == 37) |> filter(to == 43)
+
+tr_df <- distinct(tibble::tibble(
+  x = c(tr_from_to_df_pbmc[["x_from"]], tr_from_to_df_pbmc[["x_to"]]), 
+  y = c(tr_from_to_df_pbmc[["y_from"]], tr_from_to_df_pbmc[["y_to"]])))
+
+distance_df_small_edges_pbmc <- distance_pbmc |>
+  filter(distance < benchmark_pbmc)
+
+tr_from_to_df_pbmc <- inner_join(
+  tr_from_to_df_pbmc, distance_df_small_edges_pbmc, 
+  by = c("from", "to"))
+
+# tr_from_to_df_pbmc <- tr_from_to_df_pbmc |>
+#   filter(!(row_number() %in% c(155)))
+
+trimesh_removed_pbmc_best <- ggplot() + 
+  geom_segment(data = tr_from_to_df_pbmc, 
+               aes(
+                 x = x_from, 
+                 y = y_from, 
+                 xend = x_to, 
+                 yend = y_to),
+               colour = "#33a02c") +
+  geom_point(
+    data = umap_pbmc_scaled_best,
+    aes(
+      x = UMAP1,
+      y = UMAP2
+    ),
+    alpha = 0.2,
+    color = "#000000"
+  )  +
+  coord_fixed() +
+  interior_annotation("b1")
+
+# trimesh_removed_pbmc <- vis_rmlg_mesh(
+#   distance_edges = distance_pbmc,
+#   benchmark_value = benchmark_pbmc,
+#   tr_coord_df = tr_from_to_df_pbmc,
+#   distance_col = "distance") +
+#   geom_point(
+#     data = tsne_pbmc_scaled,
+#     aes(
+#       x = tSNE1,
+#       y = tSNE2
+#     ),
+#     alpha = 0.2,
+#     color = "#000000"
+#   ) 
+#   #xlim(sc_xlims) + ylim(sc_ylims) +
+#   #interior_annotation("a", sc_ltr_pos)
+
+## Compute error
+error_df <- augment(
+  df_bin_centroids = df_bin_centroids_pbmc,
+  df_bin = df_bin_pbmc,
+  training_data = training_data_pbmc,
+  newdata = NULL,
+  type_NLDR = "UMAP",
+  col_start = "PC")
+
+## Categorize error
+
+error_df <- error_df |>
+  mutate(type = case_when(
+    row_wise_abs_error <= 5 ~ "error 0-5",
+    row_wise_abs_error <= 10 ~ "error 5-10",
+    row_wise_abs_error <= 15 ~ "error 10-15",
+    row_wise_abs_error <= 20 ~ "error 15-20",
+    row_wise_abs_error <= 25 ~ "error 20-25",
+    .default = "error greter than 25"
+  )) |>
+  mutate(type = factor(type, levels = c(
+    "error 0-5", "error 5-10", "error 10-15", "error 15-20", "error 20-25", 
+    "error greter than 25")))
+
+## To join embedding
+error_df <- error_df |>
+  bind_cols(umap_pbmc_scaled_best |>
+              select(-ID))
+
+error_plot_pbmc_best <- error_df |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2,
+             color = type,
+             group = ID)) +
+  geom_point(alpha=0.5,
+             size = 0.1) +
+  interior_annotation("b", sc_ltr_pos)
+
+```
+
+```{r}
+#| label: pbmc-umap-model-proj-best
+
+df_b_pbmc <- df_bin_pbmc |>
+  dplyr::filter(hb_id %in% df_bin_centroids_pbmc$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b_pbmc <- df_b_pbmc[match(df_bin_centroids_pbmc$hexID, df_b_pbmc$hb_id),] |>
+  dplyr::select(-hb_id)
+
+# Apply the scaling
+df_model_data_pbmc <- bind_rows(data_pbmc, df_b_pbmc)
+scaled_pbmc <- scale_data_manual(df_model_data_pbmc, "type") |>
+  as_tibble()
+
+scaled_pbmc_data <- scaled_pbmc |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_pbmc_data_model <- scaled_pbmc |>
+  filter(type == "model") |>
+  select(-type)
+
+# Combine with the true model for visualization
+df <- dplyr::bind_rows(scaled_pbmc_data_model |> mutate(type = "model"),
+                       scaled_pbmc_data |> mutate(type = "data"))
+
+## Removed long edge in high-d
+# distance_df_small_edges_pbmc <- distance_df_small_edges_pbmc |>
+#   filter(!(row_number() %in% c(155)))
+
+## First projection
 projection <- cbind(
-    c(0.13399,-0.00303,0.34841,-0.15626,-0.27432,0.48771,0.14252,0.39963,0.57983),
-    c(0.49615,0.73920,-0.15299,-0.07463,-0.07948,0.13320,-0.15484,0.20042,-0.28869))
+    c(0.24156,0.85058,0.11564,-0.03255,-0.28106,0.27463,-0.10539,0.08429,0.13107),
+    c(-0.67212,0.02745,-0.26718,-0.33126,-0.34919,0.20839,-0.14498,-0.37708,0.15455))
 
@@
 
-pbmc_proj_umap_model3 <- projected_df |>
+pbmc_proj_umap_model1_best <- projected_df |>
   ggplot(
@@
   ylim(c(-1, 1)) +
-  interior_annotation("d")
+  interior_annotation("b2")
+
+## Second projection
+projection <- cbind(
+    c(0.00067,0.53963,-0.16241,-0.09777,0.32766,0.44543,-0.51359,-0.29884,0.00951),
+    c(0.62399,0.26016,0.06441,-0.34686,-0.40466,0.24643,0.23537,0.07972,0.34402))
+
+projected <- as.matrix(scaled_pbmc_data) %*% projection
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  #dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges_pbmc |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+axes_obj <- gen_axes(
+  data = scaled_pbmc_data, 
+  proj_data = projected_df,  
+  limits = 1.5,
+  axis_pos_x = -0.75,
+  axis_pos_y = -0.75)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
+
+pbmc_proj_umap_model2_best <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#33a02c") +
+  geom_point(
+    size = 0.5,
+    alpha = 0.2) +
+  scale_color_manual(values = c("#000000", "#33a02c")) +
+  geom_segment(
+    data=axes, 
+    aes(x=x1, y=y1, xend=x2, yend=y2), 
+    colour="grey70") +
+  geom_text(
+    data=axes, 
+    aes(x=x2, y=y2, label=rownames(axes)), 
+    colour="grey50",
+    size = 3) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-1, 1)) +
+  ylim(c(-1, 1)) +
+  interior_annotation("b3")
 
@@
 The full video is available at <https://youtu.be/5Y1hE4i7N2k>."
-#| label: fig-pbmc2-sc
+#| label: fig-model-pbmc-author-proj
 #| fig-pos: H
@@
 
-free(trimesh_removed_pbmc) + pbmc_proj_umap_model1 + 
-  pbmc_proj_umap_model2 + pbmc_proj_umap_model3 +
-  plot_layout(guides='collect', ncol=2) &
+free(trimesh_removed_pbmc) + free(trimesh_removed_pbmc_best) +
+  pbmc_proj_umap_model1 + pbmc_proj_umap_model1_best +
+  pbmc_proj_umap_model2 + pbmc_proj_umap_model2_best +
+  plot_layout(nrow=3) &
   theme(legend.position='none')

### 2024-09-04 paper.tex
Updated Figure 14 (only figures)

@@
 \emph{model-in-the-data-space} provides evidence for this
-(Figure~\ref{fig-pbmc2-sc}). Therefore, tSNE with perplexity \(30\)
+(\textbf{?@fig-pbmc2-sc}). Therefore, tSNE with perplexity \(30\)
 accurately captures the data structure of the PBMC3k dataset, which is
 better than UMAP.
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-model-pbmc-author-proj-1.pdf}
-
-}
-
-\caption{\label{fig-model-pbmc-author-proj}Model in \gD{}, on the UMAP
-layout (a), and three views of the fit in projections from
-\(9\text{-}D\) (b, c, d), for the PBMC3k data (\(a_1 = 0.046\),
-\(b = 676 \  (26,  26)\), \(m = 92\), benchmark value to remove large
-edges is \(0.08\)). The three well-separated clusters observed in the
-\gD{} layout of UMAP (Figure~\ref{fig-NLDR-variety} (a)) is also visible
-when visualizing the model overlaid on the data space. But the
-separation between the clusters are small in \(9\text{-}D\). The model
-shows non-linear clusters and densed points within clusters in
-\(9\text{-}D\) space, which are some additional patterns not visible in
-the \gD{} representation (Figure~\ref{fig-NLDR-variety} (a)). The full
-video is available at \url{https://youtu.be/VqqWuE0Jj6A}.}
-
-\end{figure}%
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-pbmc2-sc-1.pdf}
-
-}
-
-\caption{\label{fig-pbmc2-sc}Model in \gD{}, on the tSNE layout (a), and
-three views of the fit in projections from \(9\text{-}D\) (b, c, d), for
-the PBMC3k data (\(a_1 = 0.099\), \(b = 221 \  (13,  17)\), \(m = 92\),
-benchmark value to remove large edges is \(0.1\)). The three
-well-separated clusters with small separations observed in the \gD{}
-layout of tSNE (Figure~\ref{fig-NLDR-variety} (e)) is also visible when
-visualizing the model overlaid on the data space in \(9\text{-}D\).
-Furthermore, there are some non-linear clusters. The model shows densed
-points within clusters in \(9\text{-}D\) space, which is an additional
-patterns not visible in the \gD{} representation
-(Figure~\ref{fig-NLDR-variety} (e)). The full video is available at
-\url{https://youtu.be/5Y1hE4i7N2k}.}
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-model-pbmc-author-proj-1.pdf}
+
+}
+
+\caption{\label{fig-model-pbmc-author-proj}Model in \gD{}, on the tSNE
+layout (a), and three views of the fit in projections from
+\(9\text{-}D\) (b, c, d), for the PBMC3k data (\(a_1 = 0.099\),
+\(b = 221 \  (13,  17)\), \(m = 92\), benchmark value to remove large
+edges is \(0.1\)). The three well-separated clusters with small
+separations observed in the \gD{} layout of tSNE
+(Figure~\ref{fig-NLDR-variety} (e)) is also visible when visualizing the
+model overlaid on the data space in \(9\text{-}D\). Furthermore, there
+are some non-linear clusters. The model shows densed points within
+clusters in \(9\text{-}D\) space, which is an additional patterns not
+visible in the \gD{} representation (Figure~\ref{fig-NLDR-variety} (e)).
+The full video is available at \url{https://youtu.be/5Y1hE4i7N2k}.}
 

### 2024-09-04 paper.qmd
Updated the MNIST script

@@
 training_data_pbmc <- read_rds("data/pbmc3k/pbmc_pca_50.rds")
-names(training_data_pbmc) <- paste0("PC", 1:50)
+names(training_data_pbmc) <- paste0("p", 1:50)
 
@@
 #   geom_point(alpha=0.3, color = "#000000") 
-```
-
-<!-- UMAP parameters suggested by the author (n_neighbors = 30, min_dist = 0.3)-->
-```{r}
-#| fig-cap: "Reproduction of plot published in Chen et al.
-(2023) showing a $2\\text{-}D$ layout from UMAP applied for the PBMC3k dataset.
-The (hyper-)parameter settings, beyond the defaults, are 30 nearest neighbors and minimum distance equal to 0.3.
-We use our model-in-the-data-space to assess whether this is an accurate representation of structure present in the high-dimensional data, or if it is misleading."
-#| label: fig-umap-author
-#| out-height: 30%
-#| fig-pos: H
-#| eval: false
-
-umap_pbmc 
 ```
@@
   is_rm_lwd_hex = FALSE,
-  col_start_highd = "PC"
+  col_start_highd = "p"
 )
@@
     colour="grey50",
-    size = 3) +
+    size = 2) +
   geom_path(
@@
   type_NLDR = "UMAP",
-  col_start = "PC")
+  col_start = "p")
 

### 2024-09-04 paper.qmd
Added five layouts for MNIST

@@
 
+```{r}
+#| label: read-mnist-nldr
+# Read a variety of different NLDR representations of mnist
+# and plot them on same aspect ratio
+tsne_mnist <- read_rds("data/mnist/mnist_tsne89.rds")
+
+nldr_mnist1 <- tsne_mnist |>
+  ggplot(aes(x = tSNE1,
+             y = tSNE2))+
+  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
+  interior_annotation("a")
+
+umap_mnist <- read_rds("data/mnist/mnist_umap.rds")
+
+nldr_mnist2 <- umap_mnist |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2)) +
+  geom_point(alpha=0.1, size=1, colour='#a65628') +
+  interior_annotation("b")
+
+phate_mnist <- read_rds("data/mnist/mnist_phate.rds")
+
+nldr_mnist3 <- phate_mnist |>
+  ggplot(aes(x = PHATE1,
+             y = PHATE2))+
+  geom_point(alpha=0.1, size=1, colour='#377eb8') +
+  interior_annotation("c")
+
+trimap_mnist <- read_rds("data/mnist/mnist_trimap.rds")
+
+nldr_mnist4 <- trimap_mnist |>
+  ggplot(aes(x = TriMAP1,
+             y = TriMAP2))+
+  geom_point(alpha=0.1, size=1, colour='#4daf4a') +
+  interior_annotation("d")
+
+pacmap_mnist <- read_rds("data/mnist/mnist_pacmap.rds")
+
+nldr_mnist5 <- pacmap_mnist |>
+  ggplot(aes(x = PaCMAP1,
+             y = PaCMAP2))+
+  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
+  interior_annotation("e")
+```
+
+```{r}
+#| fig-width: 25
+#| fig-height: 5
+
+nldr_mnist1 + nldr_mnist2 + 
+  nldr_mnist3 + nldr_mnist4 + nldr_mnist5 +
+  plot_layout(ncol = 5)
+```
+
 <!-- PaCMAP preserves non-linear structure in $784\text{-}D$.

### 2024-09-04 paper.qmd
Added error plot for MNSIT data

@@
   geom_point(alpha=0.1, size=1, colour='#ff7f00') +
-  interior_annotation("a")
+  interior_annotation("a", c(0.08, 0.93))
 
@@
 ```{r}
-#| fig-width: 25
-#| fig-height: 5
-
-nldr_mnist1 + nldr_mnist2 + 
-  nldr_mnist3 + nldr_mnist4 + nldr_mnist5 +
-  plot_layout(ncol = 5)
+#| label: combine-data-mnist
+
+error_mnist_umap <- read_rds("data/mnist/error_mnist_umap.rds")
+error_mnist_tsne <- read_rds("data/mnist/error_mnist_tsne.rds")
+error_mnist_phate <- read_rds("data/mnist/error_mnist_phate.rds")
+error_mnist_trimap <- read_rds("data/mnist/error_mnist_trimap.rds")
+error_mnist_pacmap <- read_rds("data/mnist/error_mnist_pacmap.rds")
+
+error_mnist <- bind_rows(error_mnist_umap, 
+                        error_mnist_tsne,
+                        error_mnist_phate,
+                        error_mnist_trimap,
+                        error_mnist_pacmap)
+
+error_mnist <- error_mnist |>
+  mutate(a1 = round(a1, 2)) |>
+  filter(bin1 >= 5) |>
+  group_by(method, a1) |>
+  filter(MSE == min(MSE)) |>
+  ungroup()
+```
+
+```{r}
+#| label: error-comp-mnist
+
+error_plot_mnist <- ggplot(error_mnist, 
+                          aes(x = a1, 
+                              y = MSE, 
+                              colour = method)) + 
+  geom_point(size = 0.8) +
+  geom_line(linewidth = 0.3) + 
+  # geom_vline(xintercept = 15, linetype="solid", 
+  #            color = "black", linewidth=0.8, alpha = 0.5) +
+  scale_x_continuous(breaks = sort(unique(error_mnist$a1))[seq(1, length(unique(error_mnist$a1)), by = 5)]) +
+  scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a', "#ff7f00",'#a65628')) +
+  scale_y_log10() +
+  ylab("log(MSE)") +
+  xlab(expression(paste("binwidth (", a[1], ")"))) +
+  theme_minimal() +
+  theme(panel.border = element_rect(fill = 'transparent'),
+        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+        axis.ticks.x = element_line(),
+        axis.ticks.y = element_line(),
+        legend.position = "none",
+        axis.text.x = element_text(size = 7),
+        axis.text.y = element_text(size = 7),
+        axis.title.x = element_text(size = 7),
+        axis.title.y = element_text(size = 7))
+
+```
+
+```{r}
+#| fig-cap: ""
+#| label: fig-mnist-mse
+#| fig-pos: H
+#| out-height: 100%
+
+free(error_plot_mnist) + wrap_plots(nldr_mnist1, nldr_mnist2, nldr_mnist3, 
+                                    nldr_mnist4, nldr_mnist5, ncol = 2)
 ```

### 2024-09-04 paper.qmd
Try to tune with different hyper parameter choices, but nothing much changed with default settings. Therefore, the best with tSNE

@@
 # and plot them on same aspect ratio
-tsne_mnist <- read_rds("data/mnist/mnist_tsne89.rds")
+tsne_mnist <- read_rds("data/mnist/mnist_tsne30.rds")
 
@@
   #            color = "black", linewidth=0.8, alpha = 0.5) +
-  scale_x_continuous(breaks = sort(unique(error_mnist$a1))[seq(1, length(unique(error_mnist$a1)), by = 5)]) +
+  scale_x_continuous(breaks = sort(unique(error_mnist$a1))[append(seq(1, length(unique(error_mnist$a1)), by = 5), 24)]) +
   scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a', "#ff7f00",'#a65628')) +
@@
 ```{r}
-#| fig-cap: ""
+#| fig-cap: "Different methods with default hyper-parameter settings"
 #| label: fig-mnist-mse
@@
 
-pacmap_minst <- read_rds("data/mnist/mnist_pacmap.rds") |> 
-  select(PaCMAP1, PaCMAP2, ID)
 mnist_scaled_obj <- gen_scaled_data(
-  data = pacmap_minst)
-pacmap_minst_scaled <- mnist_scaled_obj$scaled_nldr
-
-pacmap_plot_mnist <- pacmap_minst_scaled |> 
-  ggplot(aes(x = PaCMAP1, 
-             y = PaCMAP2)) + 
+  data = tsne_mnist)
+tsne_minst_scaled <- mnist_scaled_obj$scaled_nldr
+
+tsne_plot_mnist <- tsne_minst_scaled |> 
+  ggplot(aes(x = tSNE1, 
+             y = tSNE2)) + 
   geom_point(alpha=0.1, color = "#000000")
 
-# read the png file from device 
-left_top_img <- readPNG("figures/mnist/mnist_digit1_img_lt.png", native = TRUE)
-center_img <- readPNG("figures/mnist/mnist_digit1_img_c.png", native = TRUE)
-right_bottom_img <- readPNG("figures/mnist/mnist_digit1_img_rb.png", native = TRUE)
-
-pacmap_plot_mnist <- pacmap_plot_mnist +
-    inset_element(p = left_top_img, 
-                  left = 0.8, 
-                  bottom = 0.85, 
-                  right = 0, 
-                  top = 0.95) +
-    inset_element(p = center_img, 
-                  left = 1.2, 
-                  bottom = 0.6, 
-                  right = 0, 
-                  top = 0.7) +
-    inset_element(p = right_bottom_img, 
-                  left = 1.7, 
-                  bottom = 0.48, 
-                  right = 0, 
-                  top = 0.58)
+# # read the png file from device 
+# left_top_img <- readPNG("figures/mnist/mnist_digit1_img_lt.png", native = TRUE)
+# center_img <- readPNG("figures/mnist/mnist_digit1_img_c.png", native = TRUE)
+# right_bottom_img <- readPNG("figures/mnist/mnist_digit1_img_rb.png", native = TRUE)
+# 
+# tsne_plot_mnist <- tsne_plot_mnist +
+#     inset_element(p = left_top_img, 
+#                   left = 0.8, 
+#                   bottom = 0.85, 
+#                   right = 0, 
+#                   top = 0.95) +
+#     inset_element(p = center_img, 
+#                   left = 1.2, 
+#                   bottom = 0.6, 
+#                   right = 0, 
+#                   top = 0.7) +
+#     inset_element(p = right_bottom_img, 
+#                   left = 1.7, 
+#                   bottom = 0.48, 
+#                   right = 0, 
+#                   top = 0.58)
 ```
@@
 
-pacmap_plot_mnist
+tsne_plot_mnist
 ```
@@
   training_data = training_data_mnist,
-  emb_df = pacmap_minst_scaled,
+  emb_df = tsne_minst_scaled,
   bin1 = num_bins_x_mnist,
@@
                colour = "#33a02c") +
-  geom_point(data = pacmap_minst_scaled,
+  geom_point(data = tsne_minst_scaled,
              aes(
-               x = PaCMAP1,
-               y = PaCMAP2
+               x = tSNE1,
+               y = tSNE2
              ),
@@
   newdata = NULL, 
-  type_NLDR = "PaCMAP", 
+  type_NLDR = "tSNE", 
   col_start = "PC")
@@
 error_df <- error_df |>
-  bind_cols(pacmap_minst_scaled |> 
+  bind_cols(tsne_minst_scaled |> 
               select(-ID))
@@
 error_plot_mnist <- error_df |>
-  ggplot(aes(x = PaCMAP1,
-             y = PaCMAP2, 
+  ggplot(aes(x = tSNE1,
+             y = tSNE2, 
              color = type, 

### 2024-09-04 scripts/mnist/2_gen_diff_embeddings.R

@@ -51,15 +51,18 @@ mnist_10_pcs_of_digit_1 <- read_rds("data/mnist/mnist_10_pcs_of_digit_1.rds")
 
 ##############
 ### tSNE
-tSNE_data_gau <- Fit_tSNE(mnist_10_pcs_of_digit_1, opt_perplexity = calculate_effective_perplexity(mnist_10_pcs_of_digit_1), with_seed = 20240110)
+#tSNE_data_gau <- Fit_tSNE(mnist_10_pcs_of_digit_1, opt_perplexity = calculate_effective_perplexity(mnist_10_pcs_of_digit_1), with_seed = 20240110)
+tSNE_data_gau <- Fit_tSNE(mnist_10_pcs_of_digit_1, opt_perplexity = 30, with_seed = 20240110)
+
 
 tSNE_data_gau <- tSNE_data_gau |>
   mutate(ID = 1:NROW(tSNE_data_gau))
 
 plot_tSNE_2D(tSNE_data_gau)
 
-write_rds(tSNE_data_gau, file = paste0("data/mnist/mnist_tsne", calculate_effective_perplexity(mnist_10_pcs_of_digit_1),".rds"))
+#write_rds(tSNE_data_gau, file = paste0("data/mnist/mnist_tsne", calculate_effective_perplexity(mnist_10_pcs_of_digit_1),".rds"))
 #write_rds(tSNE_data_gau, file = paste0("data/five_gau_clusters/tsne_data_five_gau_30.rds"))
+write_rds(tSNE_data_gau, file = paste0("data/mnist/mnist_tsne30.rds"))
 
 
 ### UMAP
@@ -131,6 +134,7 @@ data_matrix <- matrix(data_vector, ncol = NCOL(mnist_10_pcs_of_digit_1))
 
 # Initialize PaCMAP instance
 reducer <- pacmap$PaCMAP(n_components = as.integer(2), n_neighbors = as.integer(10), MN_ratio = 0.5, FP_ratio = as.integer(2))
+
 
 # Perform dimensionality Reduction
 PacMAP_data <- reducer$fit_transform(data_matrix, init = "random") |>

### 2024-09-04 paper.qmd
Added images for the anomalies

@@
 
-mnist_proj_pacmap_model1 <- projected_df |>
+mnist_proj_tsne_model1 <- projected_df |>
   ggplot(
@@
 
-mnist_proj_pacmap_model2 <- projected_df |>
+mnist_proj_tsne_model2 <- projected_df |>
   ggplot(
@@
 
-mnist_proj_pacmap_model3 <- projected_df |>
+mnist_proj_tsne_model3 <- projected_df |>
   ggplot(
@@
 #| fig-width: 10
-#| fig-cap: "Model in \\gD{}, on the PaCMAP layout (a), and three views (b, c, d) of the fit in projections from $10\\text{-}D$, for the digit 1 of MNIST data ($a_1 = 0.062$, $b = 285 \\  (19, \ 15)$, $m = 109$, and benchmark value to remove large edges is $0.108$).
+#| fig-cap: "Model in \\gD{}, on the tSNE layout (a), and three views (b, c, d) of the fit in projections from $10\\text{-}D$, for the digit 1 of MNIST data ($a_1 = 0.062$, $b = 285 \\  (19, \ 15)$, $m = 109$, and benchmark value to remove large edges is $0.108$).
 (b) The non-linear structure observed in the \\gD{} layout of PaCMAP (@fig-pacmap-author) is also visible when visualizing the model overlaid on the data space.
@@
 
-free(trimesh_removed_mnist) + mnist_proj_pacmap_model1 +
-  mnist_proj_pacmap_model2 + mnist_proj_pacmap_model3 +
+free(trimesh_removed_mnist) + mnist_proj_tsne_model1 +
+  mnist_proj_tsne_model2 + mnist_proj_tsne_model3 +
   plot_layout(ncol = 2)
@@
 
-img_error_within <- c(6169, 5987, 773, 6671, 4785, 4699, 5173, 5475)
+img_error_within <- c(5173, 6671, 2964, 2464, 6319, 7388, 4303, 7650)
 
@@
 
-img_error_outside <- c(1489, 5095, 855, 3143, 453, 3619, 1015, 409) #, 5101
+img_error_outside <- c(3554, 501, 4422, 6860, 3869, 1529, 4916, 6144)
 
@@
 #| label: fig-model-error-mnist
-#| fig-cap: "$10\\text{-}D$ model error in PaCMAP layout of MNIST digit 1 dataset in $2\\text{-}D$.
+#| fig-cap: "$10\\text{-}D$ model error in tSNE layout of MNIST digit 1 dataset in $2\\text{-}D$.
 Most low $10\\text{-}D$ model errors are distributed along the lower edge of the $2\\text{-}D$ structure, while most large $10\\text{-}D$ model errors are concentrated along the upper edge (a).

### 2024-09-04 paper.tex
Added five layouts for MNIST
Added error plot for MNSIT data
Try to tune with different hyper parameter choices, but nothing much changed with default settings. Therefore, the best with tSNE
Added images for the anomalies

@@
 
+\includegraphics[width=1\textwidth,height=1\textheight]{paper_files/figure-pdf/fig-mnist-mse-1.pdf}
+
+}
+
+\caption{\label{fig-mnist-mse}Different methods with default
+hyper-parameter settings}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+\centering{
+
 \includegraphics[width=1\textwidth,height=0.22\textheight]{paper_files/figure-pdf/fig-pacmap-author-1.pdf}
@@
 
-\caption{\label{fig-mnist-tri-proj}Model in \gD{}, on the PaCMAP layout
+\caption{\label{fig-mnist-tri-proj}Model in \gD{}, on the tSNE layout
 (a), and three views (b, c, d) of the fit in projections from
@@
 \caption{\label{fig-model-error-mnist}\(10\text{-}D\) model error in
-PaCMAP layout of MNIST digit 1 dataset in \(2\text{-}D\). Most low
+tSNE layout of MNIST digit 1 dataset in \(2\text{-}D\). Most low
 \(10\text{-}D\) model errors are distributed along the lower edge of the

### 2024-09-04 paper.qmd
Add new projections for MNIST

@@
   geom_point(
-    size = 0.5,
+    size = 0.2,
     alpha = 0.2) +
@@
 projection <- cbind(
-    c(0.13639,-0.18552,-0.38537,-0.03088,-0.04663,0.15013,-0.28351,-0.28865,-0.03379,-0.38009),
-    c(-0.46592,-0.40953,-0.02752,0.12815,0.08356,-0.12991,-0.28307,0.11145,0.08031,0.10800))
+    c(-0.33305,0.20223,0.06303,0.11124,0.13299,-0.30093,0.34082,0.15581,-0.08864,0.33573),
+    c(0.53615,0.14969,-0.09393,-0.20053,0.07992,-0.34791,0.13057,-0.06814,-0.16737,0.03716))
 
@@
   geom_point(
-    size = 0.5,
+    size = 0.2,
     alpha = 0.1) +
@@
 projection <- cbind(
-    c(0.47163,0.21622,0.04027,0.17328,0.03603,0.06488,-0.22226,-0.07186,0.39674,0.13485),
-    c(-0.41187,0.36322,0.11264,-0.14175,0.19810,0.07062,0.00680,0.08795,0.37068,-0.11280))
+    c(-0.12783,-0.50063,-0.02631,-0.00484,0.17733,0.37157,0.11310,0.09424,0.19844,0.19792),
+    c(-0.42542,0.29570,0.31106,-0.30307,0.00614,0.16714,0.21779,-0.00025,0.07020,-0.00689))
 

### 2024-09-04 paper.qmd
Added images on the tSNE layout (MNIST)

@@
 
-# # read the png file from device 
-# left_top_img <- readPNG("figures/mnist/mnist_digit1_img_lt.png", native = TRUE)
-# center_img <- readPNG("figures/mnist/mnist_digit1_img_c.png", native = TRUE)
-# right_bottom_img <- readPNG("figures/mnist/mnist_digit1_img_rb.png", native = TRUE)
-# 
-# tsne_plot_mnist <- tsne_plot_mnist +
-#     inset_element(p = left_top_img, 
-#                   left = 0.8, 
-#                   bottom = 0.85, 
-#                   right = 0, 
-#                   top = 0.95) +
-#     inset_element(p = center_img, 
-#                   left = 1.2, 
-#                   bottom = 0.6, 
-#                   right = 0, 
-#                   top = 0.7) +
-#     inset_element(p = right_bottom_img, 
-#                   left = 1.7, 
-#                   bottom = 0.48, 
-#                   right = 0, 
-#                   top = 0.58)
+# read the png file from device
+left_top_img <- readPNG("figures/mnist/mnist_right_top_img.png", native = TRUE)
+center_img <- readPNG("figures/mnist/mnist_middle_img.png", native = TRUE)
+right_bottom_img <- readPNG("figures/mnist/mnist_right_bottom_img.png", native = TRUE)
+
+tsne_plot_mnist <- tsne_plot_mnist +
+    inset_element(p = left_top_img,
+                  left = 1,
+                  bottom = 0.85,
+                  right = 0,
+                  top = 0.95) +
+    inset_element(p = center_img,
+                  left = 0.8,
+                  bottom = 0.5,
+                  right = 0,
+                  top = 0.6) +
+    inset_element(p = right_bottom_img,
+                  left = 1,
+                  bottom = 0.2,
+                  right = 0,
+                  top = 0.3)
 ```

### 2024-09-04 paper.qmd
Updated Figure 9

@@
   scale_x_continuous(breaks = sort(unique(round(error_scurve$a1, 2)))[c(1, 3, 5, 6, 8, 10:12)]) +
-  labs(x = expression(a[1]), y = "MSE") +
+  labs(x = expression(paste("binwidth (", a[1], ")")), y = "MSE") +
   ggtitle("(a)") +
@@
   scale_x_continuous(breaks = sort(unique(round(error_scurve$a1, 2)))[c(1, 3, 5, 6, 8, 10:12)]) +
-  labs(x = expression(a[1]), y = expression(m)) +
+  labs(x = expression(paste("binwidth (", a[1], ")")), y = expression(paste("non-empty bins (", m, ")"))) +
   ggtitle("(b)") +

### 2024-09-06 paper.qmd
Added intro section for MNIST data

@@
 
-The MNIST dataset consists of $70000$ grayscale images of handwritten digits [@lecun2010].
-@yingfan2021 used this dataset to evaluate how PaCMAP preserves local structures in \gD{}.
+The digit 1 of MNIST dataset consists of $7877$ grayscale images of handwritten digits [@lecun2010].
+Before further analysis, PCA was used to preprocess the data, where the first $10$ principal components, explaining 83% of the total variation, were selected.
+The objective is to select the best NLDR layout, representing the structure of the digit 1 data in $10\text{-}D$ with default (hyper-)parameters.
+
+<!-- @yingfan2021 used this dataset to evaluate how PaCMAP preserves local structures in \gD{}.
 Because of the large number of images, the focus was specifically on the handwritten digit 1, which exhibits a non-linear structure in \gD{} (@fig-pacmap-author).
@@
 The objective is to assess whether PaCMAP effectively preserves the non-linear structure in \gD{}.
+-->
 
@@
 
-As a result of visualizing the *model-in-the-data-space*, PaCMAP provides a reasonable \gD{} representation of MNIST digit 1 data, as it preserves the non-linear structure present in the $10\text{-}D$ data (@fig-mnist-tri-proj).
+<!-- As a result of visualizing the *model-in-the-data-space*, PaCMAP provides a reasonable \gD{} representation of MNIST digit 1 data, as it preserves the non-linear structure present in the $10\text{-}D$ data (@fig-mnist-tri-proj).
 It also shows some quirks like twisted patterns and long edges (@fig-mnist-tri-proj).
 Furthermore, the results help to identify anomalies (with large model errors) within and the outside the non-linear structure, displaying different patterns of the digit 1 (@fig-model-error-mnist).
+-->
 

### 2024-09-06 paper.tex
Added intro section for MNIST data

@@
 
-The MNIST dataset consists of \(70000\) grayscale images of handwritten
-digits \citep{lecun2010}. \citet{yingfan2021} used this dataset to
-evaluate how PaCMAP preserves local structures in \gD{}. Because of the
-large number of images, the focus was specifically on the handwritten
-digit 1, which exhibits a non-linear structure in \gD{}
-(Figure~\ref{fig-pacmap-author}). There are \(7877\) images of the digit
-1 in the dataset. Additionally, PCA was applied as a preprocessing step
-and the first \(10\) principal components were selected. The objective
-is to assess whether PaCMAP effectively preserves the non-linear
-structure in \gD{}.
+The digit 1 of MNIST dataset consists of \(7877\) grayscale images of
+handwritten digits \citep{lecun2010}. Before further analysis, PCA was
+used to preprocess the data, where the first \(10\) principal
+components, explaining 83\% of the total variation, were selected. The
+objective is to select the best NLDR layout, representing the structure
+of the digit 1 data in \(10\text{-}D\) with default (hyper-)parameters.
 
@@
 \end{figure}%
-
-As a result of visualizing the \emph{model-in-the-data-space}, PaCMAP
-provides a reasonable \gD{} representation of MNIST digit 1 data, as it
-preserves the non-linear structure present in the \(10\text{-}D\) data
-(Figure~\ref{fig-mnist-tri-proj}). It also shows some quirks like
-twisted patterns and long edges (Figure~\ref{fig-mnist-tri-proj}).
-Furthermore, the results help to identify anomalies (with large model
-errors) within and the outside the non-linear structure, displaying
-different patterns of the digit 1 (Figure~\ref{fig-model-error-mnist}).
 

### 2024-09-06 paper.qmd
Added the MSE MNISt section

@@
 
-The digit 1 of MNIST dataset consists of $7877$ grayscale images of handwritten digits [@lecun2010].
+The digit 1 of the MNIST dataset consists of $7877$ grayscale images of handwritten digits [@lecun2010].
 Before further analysis, PCA was used to preprocess the data, where the first $10$ principal components, explaining 83% of the total variation, were selected.
-The objective is to select the best NLDR layout, representing the structure of the digit 1 data in $10\text{-}D$ with default (hyper-)parameters.
+The objective is to select a reasonable NLDR layout, representing the non-linear sub-clustering structure of the digit 1 dataset in $10\text{-}D$.
+
+The MSE for different $a_1$ using tSNE, UMAP, PHATE, PaCMAP, and TriMAP with default (hyper-)parameter setting (@fig-mnist-mse) were calculated.
+It is found that tSNE provide the most reasonable representation for the digit 1 dataset, showing universally best (@fig-mnist-mse).
 

### 2024-09-06 paper.tex
Added the MSE MNISt section

@@
 
-The digit 1 of MNIST dataset consists of \(7877\) grayscale images of
-handwritten digits \citep{lecun2010}. Before further analysis, PCA was
-used to preprocess the data, where the first \(10\) principal
+The digit 1 of the MNIST dataset consists of \(7877\) grayscale images
+of handwritten digits \citep{lecun2010}. Before further analysis, PCA
+was used to preprocess the data, where the first \(10\) principal
 components, explaining 83\% of the total variation, were selected. The
-objective is to select the best NLDR layout, representing the structure
-of the digit 1 data in \(10\text{-}D\) with default (hyper-)parameters.
+objective is to select a reasonable NLDR layout, representing the
+non-linear sub-clustering structure of the digit 1 dataset in
+\(10\text{-}D\).
+
+The MSE for different \(a_1\) using tSNE, UMAP, PHATE, PaCMAP, and
+TriMAP with default (hyper-)parameter setting
+(Figure~\ref{fig-mnist-mse}) were calculated. It is found that tSNE
+provide the most reasonable representation for the digit 1 dataset,
+showing universally best (Figure~\ref{fig-mnist-mse}).
 

### 2024-09-06 paper.qmd
Added reasons for two clusters

--- 2024-09-06 paper.qmd
@@
 
-The digit 1 of MNIST dataset consists of $7877$ grayscale images of handwritten digits [@lecun2010].
+The digit 1 of the MNIST dataset consists of $7877$ grayscale images of handwritten digits [@lecun2010].
 Before further analysis, PCA was used to preprocess the data, where the first $10$ principal components, explaining 83% of the total variation, were selected.
-The objective is to select the best NLDR layout, representing the structure of the digit 1 data in $10\text{-}D$ with default (hyper-)parameters.
+The objective is to select a reasonable \gD{} layout, representing the non-linear structure of the digit 1 dataset in $10\text{-}D$.
+
+The MSE for different bin widths ($a_1$) using tSNE, UMAP, PHATE, PaCMAP, and TriMAP with default (hyper-)parameter setting (@fig-mnist-mse) were calculated.
+It is found that tSNE (@fig-mnist-mse (a)) provide the most reasonable representation for the digit 1 dataset, showing universally best.
+However, \gD{} representation shows a big non-linear cluster and a small cluster with a small gap.
+
+In the case of digit 1 data, there should not be any clusters unless anomalies exist, indicating different digit 1 patterns.
+The angle of the digit 1 images varies along this non-linear clustering structure (@fig-tsne-best), while the small cluster contains the digit 1 images with different patterns of the digit 1, unlike the usual (@fig-model-error-mnist (c)).
 
@@
 ```{r}
-#| fig-cap: "$2\\text{-}D$ layout from PaCMAP applied for the digit 1 of the MNIST dataset.
-The (hyper-)parameter settings, beyond the defaults, are $10$ nearest neighbors, ratio of the number of mid-near pairs to the number of neighbors equal to $0.5$, and the ratio of the number of further pairs to the number of neighbors is $2$.
+#| fig-cap: "The $2\\text{-}D$ layout from tSNE applied for the MNIST digit 1 dataset with defualt (hyper-)parameters.
 We use our *model-in-the-data-space* to assess whether this is an accurate representation of non-linear structure present in $10\\text{-}D$, or if it is misleading.
 The angle of the digit 1 images varies along this structure.
-Images at the top-left of the $2\\text{-}D$ layout show the digit 1 angled more to the left, while those at the bottom-right show the digit 1 angled more to the right."
-#| label: fig-pacmap-author
+Images at the top-right of the $\\gD{}$ layout show the digit 1 angled more to the right, while those at the bottom-right show the digit 1 angled more to the left."
+#| label: fig-tsne-best
 #| fig-pos: H

### 2024-09-06 paper.tex
Added reasons for two clusters

--- 2024-09-06 paper.tex
@@
 
-The digit 1 of MNIST dataset consists of \(7877\) grayscale images of
-handwritten digits \citep{lecun2010}. Before further analysis, PCA was
-used to preprocess the data, where the first \(10\) principal
+The digit 1 of the MNIST dataset consists of \(7877\) grayscale images
+of handwritten digits \citep{lecun2010}. Before further analysis, PCA
+was used to preprocess the data, where the first \(10\) principal
 components, explaining 83\% of the total variation, were selected. The
-objective is to select the best NLDR layout, representing the structure
-of the digit 1 data in \(10\text{-}D\) with default (hyper-)parameters.
+objective is to select a reasonable \gD{} layout, representing the
+non-linear structure of the digit 1 dataset in \(10\text{-}D\).
+
+The MSE for different bin widths (\(a_1\)) using tSNE, UMAP, PHATE,
+PaCMAP, and TriMAP with default (hyper-)parameter setting
+(Figure~\ref{fig-mnist-mse}) were calculated. It is found that tSNE
+(Figure~\ref{fig-mnist-mse} (a)) provide the most reasonable
+representation for the digit 1 dataset, showing universally best.
+However, \gD{} representation shows a big non-linear cluster and a small
+cluster with a small gap.
+
+In the case of digit 1 data, there should not be any clusters unless
+anomalies exist, indicating different digit 1 patterns. The angle of the
+digit 1 images varies along this non-linear clustering structure
+(Figure~\ref{fig-tsne-best}), while the small cluster contains the digit
+1 images with different patterns of the digit 1, unlike the usual
+(Figure~\ref{fig-model-error-mnist} (c)).
 
@@
 
-\includegraphics[width=1\textwidth,height=0.22\textheight]{paper_files/figure-pdf/fig-pacmap-author-1.pdf}
-
-}
-
-\caption{\label{fig-pacmap-author}\(2\text{-}D\) layout from PaCMAP
-applied for the digit 1 of the MNIST dataset. The (hyper-)parameter
-settings, beyond the defaults, are \(10\) nearest neighbors, ratio of
-the number of mid-near pairs to the number of neighbors equal to
-\(0.5\), and the ratio of the number of further pairs to the number of
-neighbors is \(2\). We use our \emph{model-in-the-data-space} to assess
-whether this is an accurate representation of non-linear structure
-present in \(10\text{-}D\), or if it is misleading. The angle of the
-digit 1 images varies along this structure. Images at the top-left of
-the \(2\text{-}D\) layout show the digit 1 angled more to the left,
-while those at the bottom-right show the digit 1 angled more to the
-right.}
+\includegraphics[width=1\textwidth,height=0.22\textheight]{paper_files/figure-pdf/fig-tsne-best-1.pdf}
+
+}
+
+\caption{\label{fig-tsne-best}The \(2\text{-}D\) layout from tSNE
+applied for the MNIST digit 1 dataset with defualt (hyper-)parameters.
+We use our \emph{model-in-the-data-space} to assess whether this is an
+accurate representation of non-linear structure present in
+\(10\text{-}D\), or if it is misleading. The angle of the digit 1 images
+varies along this structure. Images at the top-right of the \(\gD{}\)
+layout show the digit 1 angled more to the right, while those at the
+bottom-right show the digit 1 angled more to the left.}
 
@@
 large edges is \(0.108\)). (b) The non-linear structure observed in the
-\gD{} layout of PaCMAP (Figure~\ref{fig-pacmap-author}) is also visible
+\gD{} layout of PaCMAP (\textbf{?@fig-pacmap-author}) is also visible
 when visualizing the model overlaid on the data space. This indicates
@@
 pattern not visible in the \gD{} representation
-(Figure~\ref{fig-pacmap-author}). (d) some long edges exist in the
+(\textbf{?@fig-pacmap-author}). (d) some long edges exist in the
 \(10\text{-}D\) space that are not recognized as long edges in the \gD{}

### 2024-09-06 paper.qmd
Initial writing for the model

@@
 It is found that tSNE (@fig-mnist-mse (a)) provide the most reasonable representation for the digit 1 dataset, showing universally best.
-However, \gD{} representation shows a big non-linear cluster and a small cluster with a small gap.
+However, \gD{} representation shows a big non-linear cluster and a small cluster with a small gap (@fig-tsne-best).
 
@@
 The angle of the digit 1 images varies along this non-linear clustering structure (@fig-tsne-best), while the small cluster contains the digit 1 images with different patterns of the digit 1, unlike the usual (@fig-model-error-mnist (c)).
+This provides the evidence for two close clusters.
+
+The *model-in-the-data-space* helps to assess the \gD{} layout.
+As shown in @mnist-model-proj (a), the model provides the evidence for the non-linear structure in $10\text{-}D$.
+Also, by visualizing the model in $10\text{-}D$ shows some quirks.
+The model has a twisted pattern which actually provide evidence for the $10\text{-}D$ structure which couldn't be observed by \gD{} layout (@@mnist-model-proj (b)).
+Furthermore, there are some long edges which indicates the existence of the small cluster and it's so close (@@mnist-model-proj (c)).
 
@@
 #| label: fig-model-error-mnist
-#| fig-cap: "$10\\text{-}D$ model error in tSNE layout of MNIST digit 1 dataset in $2\\text{-}D$.
-Most low $10\\text{-}D$ model errors are distributed along the lower edge of the $2\\text{-}D$ structure, while most large $10\\text{-}D$ model errors are concentrated along the upper edge (a).
-There are certain data points that exhibit large error rates due to their deviation from the usual $10\\text{-}D$ data structure, which makes them anomalies.
+#| fig-cap: "(a) The $10\\text{-}D$ model error for tSNE layout in $\\gD{}$ of MNIST digit 1 dataset.
+Most low $10\\text{-}D$ model errors are distributed along the big non-linear clutser, while most large $10\\text{-}D$ model errors are concentrated along the small cluster.
+(b) There are certain data points that exhibit large error rates due to their deviation from the usual $10\\text{-}D$ data structure, which makes them anomalies.
 These anomalies can be classified into two types: those that are anomalies within the non-linear structure and those that lie outside of it.
-The images associated with high model error points within the non-linear structure display different patterns of the digit 1 (b).
+The images associated with high model error points within the non-linear structure display different patterns of the digit 1.
 However, when comparing these images to the ones found outside of the non-linear structure, it becomes evident that the latter display different patterns of the digit 1 (c).

### 2024-09-06 paper.tex
Initial writing for the model

@@
 However, \gD{} representation shows a big non-linear cluster and a small
-cluster with a small gap.
+cluster with a small gap (Figure~\ref{fig-tsne-best}).
 
@@
 1 images with different patterns of the digit 1, unlike the usual
-(Figure~\ref{fig-model-error-mnist} (c)).
+(Figure~\ref{fig-model-error-mnist} (c)). This provides the evidence for
+two close clusters.
+
+The \emph{model-in-the-data-space} helps to assess the \gD{} layout. As
+shown in \citet{mnist-model-proj} (a), the model provides the evidence
+for the non-linear structure in \(10\text{-}D\). Also, by visualizing
+the model in \(10\text{-}D\) shows some quirks. The model has a twisted
+pattern which actually provide evidence for the \(10\text{-}D\)
+structure which couldn't be observed by \gD{} layout
+(@\citet{mnist-model-proj} (b)). Furthermore, there are some long edges
+which indicates the existence of the small cluster and it's so close
+(@\citet{mnist-model-proj} (c)).
 
@@
 
-\caption{\label{fig-model-error-mnist}\(10\text{-}D\) model error in
-tSNE layout of MNIST digit 1 dataset in \(2\text{-}D\). Most low
-\(10\text{-}D\) model errors are distributed along the lower edge of the
-\(2\text{-}D\) structure, while most large \(10\text{-}D\) model errors
-are concentrated along the upper edge (a). There are certain data points
-that exhibit large error rates due to their deviation from the usual
-\(10\text{-}D\) data structure, which makes them anomalies. These
-anomalies can be classified into two types: those that are anomalies
-within the non-linear structure and those that lie outside of it. The
-images associated with high model error points within the non-linear
-structure display different patterns of the digit 1 (b). However, when
-comparing these images to the ones found outside of the non-linear
-structure, it becomes evident that the latter display different patterns
-of the digit 1 (c).}
+\caption{\label{fig-model-error-mnist}(a) The \(10\text{-}D\) model
+error for tSNE layout in \(\gD{}\) of MNIST digit 1 dataset. Most low
+\(10\text{-}D\) model errors are distributed along the big non-linear
+clutser, while most large \(10\text{-}D\) model errors are concentrated
+along the small cluster. (b) There are certain data points that exhibit
+large error rates due to their deviation from the usual \(10\text{-}D\)
+data structure, which makes them anomalies. These anomalies can be
+classified into two types: those that are anomalies within the
+non-linear structure and those that lie outside of it. The images
+associated with high model error points within the non-linear structure
+display different patterns of the digit 1. However, when comparing these
+images to the ones found outside of the non-linear structure, it becomes
+evident that the latter display different patterns of the digit 1 (c).}
 

### 2024-09-06 paper.qmd
Added the conclude MNIST description

@@
 
-The *model-in-the-data-space* helps to assess the \gD{} layout.
-As shown in @mnist-model-proj (a), the model provides the evidence for the non-linear structure in $10\text{-}D$.
-Also, by visualizing the model in $10\text{-}D$ shows some quirks.
-The model has a twisted pattern which actually provide evidence for the $10\text{-}D$ structure which couldn't be observed by \gD{} layout (@@mnist-model-proj (b)).
-Furthermore, there are some long edges which indicates the existence of the small cluster and it's so close (@@mnist-model-proj (c)).
+By visualizing the model generated for tSNE in $10\text{-}D$ helps to assess the \gD{} layout.
+As shown in @mnist-model-proj (a), the model provides the evidence for the non-linear structure of the digit 1 data in $10\text{-}D$.
+The model shows some quirks.
+The model's twisted pattern provides evidence for the $10\text{-}D$ data structure, which is not observed by \gD{} layout (@@mnist-model-proj (b)).
+Furthermore, the presence of long edges indicates the existence of the small cluster located closely together (@@mnist-model-proj (c)).
 

### 2024-09-06 paper.tex
Added the conclude MNIST description

@@
 
-The \emph{model-in-the-data-space} helps to assess the \gD{} layout. As
-shown in \citet{mnist-model-proj} (a), the model provides the evidence
-for the non-linear structure in \(10\text{-}D\). Also, by visualizing
-the model in \(10\text{-}D\) shows some quirks. The model has a twisted
-pattern which actually provide evidence for the \(10\text{-}D\)
-structure which couldn't be observed by \gD{} layout
-(@\citet{mnist-model-proj} (b)). Furthermore, there are some long edges
-which indicates the existence of the small cluster and it's so close
+By visualizing the model generated for tSNE in \(10\text{-}D\) helps to
+assess the \gD{} layout. As shown in \citet{mnist-model-proj} (a), the
+model provides the evidence for the non-linear structure of the digit 1
+data in \(10\text{-}D\). The model shows some quirks. The model's
+twisted pattern provides evidence for the \(10\text{-}D\) data
+structure, which is not observed by \gD{} layout
+(@\citet{mnist-model-proj} (b)). Furthermore, the presence of long edges
+indicates the existence of the small cluster located closely together
 (@\citet{mnist-model-proj} (c)).

### 2024-09-06 paper.qmd
Updated Figure 18 title

@@
 #| label: fig-model-error-mnist
-#| fig-cap: "(a) The $10\\text{-}D$ model error for tSNE layout in $\\gD{}$ of MNIST digit 1 dataset.
-Most low $10\\text{-}D$ model errors are distributed along the big non-linear clutser, while most large $10\\text{-}D$ model errors are concentrated along the small cluster.
-(b) There are certain data points that exhibit large error rates due to their deviation from the usual $10\\text{-}D$ data structure, which makes them anomalies.
-These anomalies can be classified into two types: those that are anomalies within the non-linear structure and those that lie outside of it.
-The images associated with high model error points within the non-linear structure display different patterns of the digit 1.
-However, when comparing these images to the ones found outside of the non-linear structure, it becomes evident that the latter display different patterns of the digit 1 (c).
-"
+#| fig-cap: "(a) The $10\\text{-}D$ model error for tSNE layout in $\\gD{}$ of the MNIST digit 1 dataset.
+Most low model errors are distributed along the big non-linear clutser, while most large model errors are concentrated along the small cluster.
+(b) There are certain images that exhibit large errors due to their deviation from the non-linear cluster, which makes them anomalies.
+The images associated with large model errors shows different patterns of digit 1."
 #| fig-pos: H
@@
 
-error_plot_mnist / (imge_error_sample_within + imge_error_sample_outside) +
+#imge_error_sample_within
+
+error_plot_mnist / (imge_error_sample_outside) +
   plot_annotation(tag_levels = 'a') &
@@
 ```
+
+<!-- These anomalies can be classified into two types: those that are anomalies within the non-linear cluster and those that lie in the small cluster.
+The images associated with large model error points within the non-linear cluster display different patterns of the digit 1.
+-->
 

### 2024-09-06 paper.tex
Updated Figure 18 title

@@
 \caption{\label{fig-model-error-mnist}(a) The \(10\text{-}D\) model
-error for tSNE layout in \(\gD{}\) of MNIST digit 1 dataset. Most low
-\(10\text{-}D\) model errors are distributed along the big non-linear
-clutser, while most large \(10\text{-}D\) model errors are concentrated
-along the small cluster. (b) There are certain data points that exhibit
-large error rates due to their deviation from the usual \(10\text{-}D\)
-data structure, which makes them anomalies. These anomalies can be
-classified into two types: those that are anomalies within the
-non-linear structure and those that lie outside of it. The images
-associated with high model error points within the non-linear structure
-display different patterns of the digit 1. However, when comparing these
-images to the ones found outside of the non-linear structure, it becomes
-evident that the latter display different patterns of the digit 1 (c).}
+error for tSNE layout in \(\gD{}\) of the MNIST digit 1 dataset. Most
+low model errors are distributed along the big non-linear clutser, while
+most large model errors are concentrated along the small cluster. (b)
+There are certain images that exhibit large errors due to their
+deviation from the non-linear cluster, which makes them anomalies. The
+images associated with large model errors shows different patterns of
+digit 1.}
 

### 2024-09-06 paper.qmd
Updated figure 17 title

@@
 By visualizing the model generated for tSNE in $10\text{-}D$ helps to assess the \gD{} layout.
-As shown in @mnist-model-proj (a), the model provides the evidence for the non-linear structure of the digit 1 data in $10\text{-}D$.
+As shown in @fig-mnist-tri-proj (a), the model provides the evidence for the non-linear structure of the digit 1 data in $10\text{-}D$.
 The model shows some quirks.
-The model's twisted pattern provides evidence for the $10\text{-}D$ data structure, which is not observed by \gD{} layout (@@mnist-model-proj (b)).
-Furthermore, the presence of long edges indicates the existence of the small cluster located closely together (@@mnist-model-proj (c)).
+The model's twisted pattern provides evidence for the $10\text{-}D$ data structure, which is not observed by \gD{} layout (@fig-mnist-tri-proj (b)).
+Furthermore, the presence of long edges indicates the existence of the small cluster located closely together (@fig-mnist-tri-proj (c)).
 
@@
 #| fig-width: 10
-#| fig-cap: "Model in \\gD{}, on the tSNE layout (a), and three views (b, c, d) of the fit in projections from $10\\text{-}D$, for the digit 1 of MNIST data ($a_1 = 0.062$, $b = 285 \\  (19, \ 15)$, $m = 109$, and benchmark value to remove large edges is $0.108$).
-(b) The non-linear structure observed in the \\gD{} layout of PaCMAP (@fig-pacmap-author) is also visible when visualizing the model overlaid on the data space.
-This indicates that PaCMAP accurately captures the non-linear structure of the $10\\text{-}D$ data.
-(c) the model shows a twisted pattern within the non-linear structure in $10\\text{-}D$ space, which is an additional pattern not visible in the \\gD{} representation (@fig-pacmap-author).
+#| fig-cap: "Model in \\gD{}, on the tSNE layout (a), and three views (b, c, d) of the fit in projections from $10\\text{-}D$, for the digit 1 of MNIST data ($a_1 = 0.07$, $b = 456 \\  (19, \ 24)$, $m = 226$, and benchmark value to remove large edges is $0.114$).
+(b) The non-linear structure observed in the \\gD{} layout of tSNE (@fig-tsne-best) is also visible when visualizing the model overlaid on the data space.
+This indicates that tSNE accurately captures the non-linear structure of the $10\\text{-}D$ data.
+(c) the model shows a twisted pattern within the non-linear structure in $10\\text{-}D$ space, which is an additional pattern not visible in the \\gD{} representation (@fig-tsne-best).
 (d) some long edges exist in the $10\\text{-}D$ space that are not recognized as long edges in the \\gD{} representation.

### 2024-09-06 paper.tex
Updated figure 17 title

@@
 By visualizing the model generated for tSNE in \(10\text{-}D\) helps to
-assess the \gD{} layout. As shown in \citet{mnist-model-proj} (a), the
-model provides the evidence for the non-linear structure of the digit 1
-data in \(10\text{-}D\). The model shows some quirks. The model's
-twisted pattern provides evidence for the \(10\text{-}D\) data
+assess the \gD{} layout. As shown in Figure~\ref{fig-mnist-tri-proj}
+(a), the model provides the evidence for the non-linear structure of the
+digit 1 data in \(10\text{-}D\). The model shows some quirks. The
+model's twisted pattern provides evidence for the \(10\text{-}D\) data
 structure, which is not observed by \gD{} layout
-(@\citet{mnist-model-proj} (b)). Furthermore, the presence of long edges
-indicates the existence of the small cluster located closely together
-(@\citet{mnist-model-proj} (c)).
+(Figure~\ref{fig-mnist-tri-proj} (b)). Furthermore, the presence of long
+edges indicates the existence of the small cluster located closely
+together (Figure~\ref{fig-mnist-tri-proj} (c)).
 
@@
 (a), and three views (b, c, d) of the fit in projections from
-\(10\text{-}D\), for the digit 1 of MNIST data (\(a_1 = 0.062\),
-\(b = 285 \  (19,  15)\), \(m = 109\), and benchmark value to remove
-large edges is \(0.108\)). (b) The non-linear structure observed in the
-\gD{} layout of PaCMAP (\textbf{?@fig-pacmap-author}) is also visible
-when visualizing the model overlaid on the data space. This indicates
-that PaCMAP accurately captures the non-linear structure of the
-\(10\text{-}D\) data. (c) the model shows a twisted pattern within the
-non-linear structure in \(10\text{-}D\) space, which is an additional
-pattern not visible in the \gD{} representation
-(\textbf{?@fig-pacmap-author}). (d) some long edges exist in the
-\(10\text{-}D\) space that are not recognized as long edges in the \gD{}
-representation. The full video is available at
-\url{https://youtu.be/zcg_GXBmqjA}.}
+\(10\text{-}D\), for the digit 1 of MNIST data (\(a_1 = 0.07\),
+\(b = 456 \  (19,  24)\), \(m = 226\), and benchmark value to remove
+large edges is \(0.114\)). (b) The non-linear structure observed in the
+\gD{} layout of tSNE (Figure~\ref{fig-tsne-best}) is also visible when
+visualizing the model overlaid on the data space. This indicates that
+tSNE accurately captures the non-linear structure of the \(10\text{-}D\)
+data. (c) the model shows a twisted pattern within the non-linear
+structure in \(10\text{-}D\) space, which is an additional pattern not
+visible in the \gD{} representation (Figure~\ref{fig-tsne-best}). (d)
+some long edges exist in the \(10\text{-}D\) space that are not
+recognized as long edges in the \gD{} representation. The full video is
+available at \url{https://youtu.be/zcg_GXBmqjA}.}
 

### 2024-09-06 paper.qmd
Updated Figure 16 title

@@
 ```{r}
-#| fig-cap: "The $2\\text{-}D$ layout from tSNE applied for the MNIST digit 1 dataset with defualt (hyper-)parameters.
-We use our *model-in-the-data-space* to assess whether this is an accurate representation of non-linear structure present in $10\\text{-}D$, or if it is misleading.
+#| fig-cap: "The $\\gD{}$ layout from tSNE applied for the MNIST digit 1 dataset with defualt (hyper-)parameters.
+We use our model in the $10\\text{-}D$ space to assess whether this is an accurate representation of data structure present in $10\\text{-}D$, or if it is misleading.
 The angle of the digit 1 images varies along this structure.

### 2024-09-06 paper.tex
Updated Figure 16 title

@@
 
-\caption{\label{fig-tsne-best}The \(2\text{-}D\) layout from tSNE
-applied for the MNIST digit 1 dataset with defualt (hyper-)parameters.
-We use our \emph{model-in-the-data-space} to assess whether this is an
-accurate representation of non-linear structure present in
-\(10\text{-}D\), or if it is misleading. The angle of the digit 1 images
-varies along this structure. Images at the top-right of the \(\gD{}\)
-layout show the digit 1 angled more to the right, while those at the
-bottom-right show the digit 1 angled more to the left.}
+\caption{\label{fig-tsne-best}The \(\gD{}\) layout from tSNE applied for
+the MNIST digit 1 dataset with defualt (hyper-)parameters. We use our
+model in the \(10\text{-}D\) space to assess whether this is an accurate
+representation of data structure present in \(10\text{-}D\), or if it is
+misleading. The angle of the digit 1 images varies along this structure.
+Images at the top-right of the \(\gD{}\) layout show the digit 1 angled
+more to the right, while those at the bottom-right show the digit 1
+angled more to the left.}
 

### 2024-09-06 paper.qmd
Updated Figure 15 title

@@
 
-The MSE for different bin widths ($a_1$) using tSNE, UMAP, PHATE, PaCMAP, and TriMAP with default (hyper-)parameter setting (@fig-mnist-mse) were calculated.
+The MSE for different binwidths ($a_1$) using tSNE, UMAP, PHATE, PaCMAP, and TriMAP with default (hyper-)parameter setting (@fig-mnist-mse) were calculated.
 It is found that tSNE (@fig-mnist-mse (a)) provide the most reasonable representation for the digit 1 dataset, showing universally best.
@@
 ```{r}
-#| fig-cap: "Different methods with default hyper-parameter settings"
+#| fig-cap: "Assessing which of the 5 NLDR layouts on the MNIST digit 1 data is the better representation using MSE for varying binwidth ($a_1$).
+Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-e).
+Layout c is universally poor.
+Layouts a that show two close clusters are universally suboptimal.
+Layout a is the best choice."
 #| label: fig-mnist-mse

### 2024-09-06 paper.tex
Updated Figure 15 title

@@
 
-The MSE for different bin widths (\(a_1\)) using tSNE, UMAP, PHATE,
+The MSE for different binwidths (\(a_1\)) using tSNE, UMAP, PHATE,
 PaCMAP, and TriMAP with default (hyper-)parameter setting
@@
 
-\caption{\label{fig-mnist-mse}Different methods with default
-hyper-parameter settings}
+\caption{\label{fig-mnist-mse}Assessing which of the 5 NLDR layouts on
+the MNIST digit 1 data is the better representation using MSE for
+varying binwidth (\(a_1\)). Colour used for the lines and points in the
+left plot and in the scatterplots represents NLDR layout (a-e). Layout c
+is universally poor. Layouts a that show two close clusters are
+universally suboptimal. Layout a is the best choice.}
 

### 2024-09-06 paper.qmd
Updated MNIST section
Added missing reference for MNIST

@@
 @fig-NLDR-variety (a) is the reproduction of the published plot.
+The objective is to select a reasonable UMAP layout, representing the non-linear structure of the digit 1 dataset in $10\text{-}D$.
 
@@
 Before further analysis, PCA was used to preprocess the data, where the first $10$ principal components, explaining 83% of the total variation, were selected.
-The objective is to select a reasonable \gD{} layout, representing the non-linear structure of the digit 1 dataset in $10\text{-}D$.
-
-The MSE for different binwidths ($a_1$) using tSNE, UMAP, PHATE, PaCMAP, and TriMAP with default (hyper-)parameter setting (@fig-mnist-mse) were calculated.
-It is found that tSNE (@fig-mnist-mse (a)) provide the most reasonable representation for the digit 1 dataset, showing universally best.
-However, \gD{} representation shows a big non-linear cluster and a small cluster with a small gap (@fig-tsne-best).
-
-In the case of digit 1 data, there should not be any clusters unless anomalies exist, indicating different digit 1 patterns.
-The angle of the digit 1 images varies along this non-linear clustering structure (@fig-tsne-best), while the small cluster contains the digit 1 images with different patterns of the digit 1, unlike the usual (@fig-model-error-mnist (c)).
-This provides the evidence for two close clusters.
-
-By visualizing the model generated for tSNE in $10\text{-}D$ helps to assess the \gD{} layout.
-As shown in @fig-mnist-tri-proj (a), the model provides the evidence for the non-linear structure of the digit 1 data in $10\text{-}D$.
-The model shows some quirks.
-The model's twisted pattern provides evidence for the $10\text{-}D$ data structure, which is not observed by \gD{} layout (@fig-mnist-tri-proj (b)).
-Furthermore, the presence of long edges indicates the existence of the small cluster located closely together (@fig-mnist-tri-proj (c)).
+The objective is to select a reasonable \gD{} layout, representing the non-linear structure of the digit 1 dataset in $10\text{-}D$ (@fig-mnist-tri-proj (a)).
 
@@
 -->
+
+The MSE for different binwidths ($a_1$) using tSNE, UMAP, PHATE, PaCMAP, and TriMAP with default (hyper-)parameter setting (@fig-mnist-mse) were calculated.
+It is found that tSNE (@fig-mnist-mse (a)) provide the most reasonable representation for the digit 1 dataset, showing universally best.
+However, \gD{} representation shows a big non-linear cluster and a small cluster with a small gap (@fig-tsne-best).
+
+In the case of digit 1 data, there should not be any clusters unless anomalies exist, indicating different digit 1 patterns.
+The angle of the digit 1 images varies along this non-linear clustering structure (@fig-tsne-best), while the small cluster contains the digit 1 images with different patterns of the digit 1, unlike the usual (@fig-model-error-mnist (c)).
+This provides the evidence for two close clusters.
+
+By visualizing the model generated for tSNE in $10\text{-}D$ helps to assess the \gD{} layout.
+As shown in @fig-mnist-tri-proj (a), the model provides the evidence for the non-linear structure of the digit 1 data in $10\text{-}D$.
+The model shows some quirks.
+The model's twisted pattern provides evidence for the $10\text{-}D$ data structure, which is not observed by \gD{} layout (@fig-mnist-tri-proj (b)).
+Furthermore, the presence of long edges indicates the existence of the small cluster located closely together (@fig-mnist-tri-proj (c)).
 

### 2024-09-06 paper.tex
Updated MNIST section
Added missing reference for MNIST

@@
 are used to generate the UMAP. Figure~\ref{fig-NLDR-variety} (a) is the
-reproduction of the published plot.
+reproduction of the published plot. The objective is to select a
+reasonable UMAP layout, representing the non-linear structure of the
+digit 1 dataset in \(10\text{-}D\).
 
@@
 objective is to select a reasonable \gD{} layout, representing the
-non-linear structure of the digit 1 dataset in \(10\text{-}D\).
+non-linear structure of the digit 1 dataset in \(10\text{-}D\)
+(Figure~\ref{fig-mnist-tri-proj} (a)).
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=1\textheight]{paper_files/figure-pdf/fig-mnist-mse-1.pdf}
+
+}
+
+\caption{\label{fig-mnist-mse}Assessing which of the 5 NLDR layouts on
+the MNIST digit 1 data is the better representation using MSE for
+varying binwidth (\(a_1\)). Colour used for the lines and points in the
+left plot and in the scatterplots represents NLDR layout (a-e). Layout c
+is universally poor. Layouts a that show two close clusters are
+universally suboptimal. Layout a is the best choice.}
+
+\end{figure}%
 
@@
 together (Figure~\ref{fig-mnist-tri-proj} (c)).
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=1\textheight]{paper_files/figure-pdf/fig-mnist-mse-1.pdf}
-
-}
-
-\caption{\label{fig-mnist-mse}Assessing which of the 5 NLDR layouts on
-the MNIST digit 1 data is the better representation using MSE for
-varying binwidth (\(a_1\)). Colour used for the lines and points in the
-left plot and in the scatterplots represents NLDR layout (a-e). Layout c
-is universally poor. Layouts a that show two close clusters are
-universally suboptimal. Layout a is the best choice.}
-
-\end{figure}%
 

### 2024-09-06 paper.tex
Intro paragraph for PBMC3k

@@
 reproduction of the published plot. The objective is to select a
-reasonable UMAP layout, representing the non-linear structure of the
-digit 1 dataset in \(10\text{-}D\).
+reasonable \gD{} layout, representing the three close clusters of the
+PBMC3k dataset in \(9\text{-}D\)
+(Figure~\ref{fig-model-pbmc-author-proj} (a2)).
 

### 2024-09-06 paper.qmd
Intro paragraph for PBMC3k
PBMC3k intro

@@
 @fig-NLDR-variety (a) is the reproduction of the published plot.
-The objective is to select a reasonable UMAP layout, representing the non-linear structure of the digit 1 dataset in $10\text{-}D$.
+The objective is to assess the published layout, and if it does not accurately represent the three clusters with small separations of the PBMC3k dataset (@fig-model-pbmc-author-proj (a2)), then select a reasonable \gD{} layout.
 
@@
 Therefore, the suggested UMAP representation (@fig-NLDR-variety (a)) does not accurately represent the structure of PBMC3k dataset in $9\text{-}D$.
-But, when visualizing the *model-in-the-data-space* some unobserved structures can be seen.
+
+<!-- But, when visualizing the *model-in-the-data-space* some unobserved structures can be seen.
 Some clusters have non-linear continuity patterns and high-density patches (@fig-model-pbmc-author-proj).
+-->
 

### 2024-09-06 paper.tex
PBMC3k intro

--- 2024-09-06 paper.tex
@@
 are used to generate the UMAP. Figure~\ref{fig-NLDR-variety} (a) is the
-reproduction of the published plot. The objective is to select a
-reasonable UMAP layout, representing the non-linear structure of the
-digit 1 dataset in \(10\text{-}D\).
+reproduction of the published plot. The objective is to assess the
+published layout, and if it does not accurately represent the three
+clusters with small separations of the PBMC3k dataset
+(Figure~\ref{fig-model-pbmc-author-proj} (a2)), then select a reasonable
+\gD{} layout.
 
@@
 representation (Figure~\ref{fig-NLDR-variety} (a)) does not accurately
-represent the structure of PBMC3k dataset in \(9\text{-}D\). But, when
-visualizing the \emph{model-in-the-data-space} some unobserved
-structures can be seen. Some clusters have non-linear continuity
-patterns and high-density patches
-(Figure~\ref{fig-model-pbmc-author-proj}).
+represent the structure of PBMC3k dataset in \(9\text{-}D\).
 

### 2024-09-06 paper.qmd
Updated refrence for Figure 13

@@
 
-In order to find a reasonable NLDR representation for the PBMC3k dataset, the MSE for different $a_1$ using UMAP, tSNE, PHATE, PaCMAP, and TriMAP with different (hyper-)parameter settings (@fig-pbmc-mse) were calculated.
-tSNE with a perplexity value set to $30$, the default parameter setting, is considered as a reasonable representation for the PBMC3k dataset.
-As shown in @fig-NLDR-variety (e), there are three well-separated clusters, although the separation between the clusters are small.
-Additionally, non-linear structures can also be observed within the clusters (@fig-NLDR-variety (e)).
-The visualization of the *model-in-the-data-space* provides evidence for this (@fig-pbmc2-sc).
-Therefore, tSNE with perplexity $30$ accurately captures the data structure of the PBMC3k dataset, which is better than UMAP.
-
 <!-- As shown in @fig-NLDR-variety (e), there are three well-separated clusters, although the separation between the clusters are small.
@@
 ```{r}
-#| fig-cap: "Assessing which of the 8 NLDR layouts on the PBMC3k data  (shown in Figure 1) is the better representation using MSE for varying binwidth ($a_1$).
+#| fig-cap: "Assessing which of the 8 NLDR layouts on the PBMC3k data  (shown in @fig-NLDR-variety) is the better representation using MSE for varying binwidth ($a_1$).
 Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-h).
@@
 ```
+
+In order to find a reasonable NLDR representation for the PBMC3k dataset, the MSE for different $a_1$ using UMAP, tSNE, PHATE, PaCMAP, and TriMAP with different (hyper-)parameter settings (@fig-pbmc-mse) were calculated.
+tSNE with a perplexity value set to $30$, the default parameter setting, is considered as a reasonable representation for the PBMC3k dataset.
+As shown in @fig-NLDR-variety (e), there are three well-separated clusters, although the separation between the clusters are small.
+Additionally, non-linear structures can also be observed within the clusters (@fig-NLDR-variety (e)).
+The visualization of the *model-in-the-data-space* provides evidence for this (@fig-pbmc2-sc).
+Therefore, tSNE with perplexity $30$ accurately captures the data structure of the PBMC3k dataset, which is better than UMAP.
 

### 2024-09-06 paper.tex
Updated refrence for Figure 13

@@
 
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=1\textheight]{paper_files/figure-pdf/fig-pbmc-mse-1.pdf}
+
+}
+
+\caption{\label{fig-pbmc-mse}Assessing which of the 8 NLDR layouts on
+the PBMC3k data (shown in Figure~\ref{fig-NLDR-variety}) is the better
+representation using MSE for varying binwidth (\(a_1\)). Colour used for
+the lines and points in the left plot and in the scatterplots represents
+NLDR layout (a-h). Layout f is universally poor. Layouts a, b, g, h that
+show large separations between clusters are universally suboptimal.
+Layout d with little separation performs well at tiny binwidth (where
+most points are in their own bin) and poorly as binwidth increases. The
+choice of best is between layouts c and e, that have small separations
+between oddly shaped clusters. Layout e is the best choice.}
+
+\end{figure}%
+
 In order to find a reasonable NLDR representation for the PBMC3k
@@
 
-\includegraphics[width=1\textwidth,height=1\textheight]{paper_files/figure-pdf/fig-pbmc-mse-1.pdf}
-
-}
-
-\caption{\label{fig-pbmc-mse}Assessing which of the 8 NLDR layouts on
-the PBMC3k data (shown in Figure 1) is the better representation using
-MSE for varying binwidth (\(a_1\)). Colour used for the lines and points
-in the left plot and in the scatterplots represents NLDR layout (a-h).
-Layout f is universally poor. Layouts a, b, g, h that show large
-separations between clusters are universally suboptimal. Layout d with
-little separation performs well at tiny binwidth (where most points are
-in their own bin) and poorly as binwidth increases. The choice of best
-is between layouts c and e, that have small separations between oddly
-shaped clusters. Layout e is the best choice.}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\centering{
-
 \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-model-pbmc-author-proj-1.pdf}

### 2024-09-07 paper.qmd
Updated PBMC descriptions

@@
 The @fig-NLDR-variety (a) shows three well-separated clusters with big separations.
-However, as shown in @fig-model-pbmc-author-proj, there is no big separation between three clusters in $9\text{-}D$.
-Therefore, the suggested UMAP representation (@fig-NLDR-variety (a)) does not accurately represent the structure of PBMC3k dataset in $9\text{-}D$.
+However, as shown in @fig-model-pbmc-author-proj (a2), there is no big separation between three clusters in $9\text{-}D$.
+Therefore, the suggested UMAP representation (@fig-NLDR-variety (a)) does not accurately represent the structure of PBMC3k dataset.
 
@@
 
+The MSE for different binwidths ($a_1$) using tSNE, UMAP, PHATE, PaCMAP, and TriMAP with different (hyper-)parameter settings (@fig-pbmc-mse) were calculated.
+It is found that tSNE (@fig-mnist-mse (a)) provide the most reasonable representation for the digit 1 dataset, showing universally best.
+However, \gD{} representation shows a big non-linear cluster and a small cluster with a small gap (@fig-tsne-best).
+
 In order to find a reasonable NLDR representation for the PBMC3k dataset, the MSE for different $a_1$ using UMAP, tSNE, PHATE, PaCMAP, and TriMAP with different (hyper-)parameter settings (@fig-pbmc-mse) were calculated.

### 2024-09-07 paper.tex
Updated PBMC descriptions

@@
 with big separations. However, as shown in
-Figure~\ref{fig-model-pbmc-author-proj}, there is no big separation
+Figure~\ref{fig-model-pbmc-author-proj} (a2), there is no big separation
 between three clusters in \(9\text{-}D\). Therefore, the suggested UMAP
 representation (Figure~\ref{fig-NLDR-variety} (a)) does not accurately
-represent the structure of PBMC3k dataset in \(9\text{-}D\).
+represent the structure of PBMC3k dataset.
 
@@
 \end{figure}%
+
+The MSE for different binwidths (\(a_1\)) using tSNE, UMAP, PHATE,
+PaCMAP, and TriMAP with different (hyper-)parameter settings
+(Figure~\ref{fig-pbmc-mse}) were calculated. It is found that tSNE
+(Figure~\ref{fig-mnist-mse} (a)) provide the most reasonable
+representation for the digit 1 dataset, showing universally best.
+However, \gD{} representation shows a big non-linear cluster and a small
+cluster with a small gap (Figure~\ref{fig-tsne-best}).
 

### 2024-09-07 paper.qmd
Updated PBMC3k description

@@
 
-To determine whether the UMAP representation with the specific (hyper-)parameter choice suggested by @chen2023 preserves the data structure present in $9\text{-}D$ (@fig-model-pbmc-author-proj (b)), we visualize the model constructed with UMAP overlaid on the $9\text{-}D$ data.
 The @fig-NLDR-variety (a) shows three well-separated clusters with big separations.
@@
 Therefore, the suggested UMAP representation (@fig-NLDR-variety (a)) does not accurately represent the structure of PBMC3k dataset.
+
+As shown in @fig-NLDR-variety, layout c, d, and e that show small separations between clusters are universally optimal.
+But layout d with little separations perform well at tiny binwidth and poorly as binwidth increases.
+Also, layout e performs well at large binwidths.
+The layout c is chosen as the best, since it used the UMAP with a different (hyper-) parameter settings.
+
+The visualization of the model generated for the selected layout in $9\text{-}D$ shows edges between the clusters (@fig-mnist-tri-proj (b2)).
+This provides evidence for the small separations between clusters.
+Furthermore, the data does not uniformally distributed along the clusters and therefore shows densed points (@fig-mnist-tri-proj (b2).
+Also, the clusters are non-linear shaped (@fig-mnist-tri-proj (b3).
+
+<!-- To determine whether the UMAP representation with the specific (hyper-)parameter choice suggested by @chen2023 preserves the data structure present in $9\text{-}D$ (@fig-model-pbmc-author-proj (b)), we visualize the model constructed with UMAP overlaid on the $9\text{-}D$ data.
+-->
 
@@
 
-The MSE for different binwidths ($a_1$) using tSNE, UMAP, PHATE, PaCMAP, and TriMAP with different (hyper-)parameter settings (@fig-pbmc-mse) were calculated.
-It is found that tSNE (@fig-mnist-mse (a)) provide the most reasonable representation for the digit 1 dataset, showing universally best.
+<!-- The MSE for different binwidths ($a_1$) using tSNE, UMAP, PHATE, PaCMAP, and TriMAP with different (hyper-)parameter settings (@fig-pbmc-mse) were calculated.  -->
+
+<!-- It is found that tSNE (@fig-mnist-mse (a)) provide the most reasonable representation for the digit 1 dataset, showing universally best.
 However, \gD{} representation shows a big non-linear cluster and a small cluster with a small gap (@fig-tsne-best).
-
-In order to find a reasonable NLDR representation for the PBMC3k dataset, the MSE for different $a_1$ using UMAP, tSNE, PHATE, PaCMAP, and TriMAP with different (hyper-)parameter settings (@fig-pbmc-mse) were calculated.
+-->
+
+<!-- In order to find a reasonable NLDR representation for the PBMC3k dataset, the MSE for different $a_1$ using UMAP, tSNE, PHATE, PaCMAP, and TriMAP with different (hyper-)parameter settings (@fig-pbmc-mse) were calculated.
 tSNE with a perplexity value set to $30$, the default parameter setting, is considered as a reasonable representation for the PBMC3k dataset.
@@
 Additionally, non-linear structures can also be observed within the clusters (@fig-NLDR-variety (e)).
-The visualization of the *model-in-the-data-space* provides evidence for this (@fig-pbmc2-sc).
+-->
+<!-- The visualization of the *model-in-the-data-space* provides evidence for this (@fig-pbmc2-sc).
 Therefore, tSNE with perplexity $30$ accurately captures the data structure of the PBMC3k dataset, which is better than UMAP.
+-->
 

### 2024-09-07 paper.tex
Updated PBMC3k description

@@
 
-To determine whether the UMAP representation with the specific
-(hyper-)parameter choice suggested by \citet{chen2023} preserves the
-data structure present in \(9\text{-}D\)
-(Figure~\ref{fig-model-pbmc-author-proj} (b)), we visualize the model
-constructed with UMAP overlaid on the \(9\text{-}D\) data. The
-Figure~\ref{fig-NLDR-variety} (a) shows three well-separated clusters
-with big separations. However, as shown in
+The Figure~\ref{fig-NLDR-variety} (a) shows three well-separated
+clusters with big separations. However, as shown in
 Figure~\ref{fig-model-pbmc-author-proj} (a2), there is no big separation
@@
 represent the structure of PBMC3k dataset.
+
+As shown in Figure~\ref{fig-NLDR-variety}, layout c, d, and e that show
+small separations between clusters are universally optimal. But layout d
+with little separations perform well at tiny binwidth and poorly as
+binwidth increases. Also, layout e performs well at large binwidths. The
+layout c is chosen as the best, since it used the UMAP with a different
+(hyper-) parameter settings.
+
+The visualization of the model generated for the selected layout in
+\(9\text{-}D\) shows edges between the clusters
+(Figure~\ref{fig-mnist-tri-proj} (b2)). This provides evidence for the
+small separations between clusters. Furthermore, the data does not
+uniformally distributed along the clusters and therefore shows densed
+points (Figure~\ref{fig-mnist-tri-proj} (b2). Also, the clusters are
+non-linear shaped (Figure~\ref{fig-mnist-tri-proj} (b3).
 
@@
 \end{figure}%
-
-The MSE for different binwidths (\(a_1\)) using tSNE, UMAP, PHATE,
-PaCMAP, and TriMAP with different (hyper-)parameter settings
-(Figure~\ref{fig-pbmc-mse}) were calculated. It is found that tSNE
-(Figure~\ref{fig-mnist-mse} (a)) provide the most reasonable
-representation for the digit 1 dataset, showing universally best.
-However, \gD{} representation shows a big non-linear cluster and a small
-cluster with a small gap (Figure~\ref{fig-tsne-best}).
-
-In order to find a reasonable NLDR representation for the PBMC3k
-dataset, the MSE for different \(a_1\) using UMAP, tSNE, PHATE, PaCMAP,
-and TriMAP with different (hyper-)parameter settings
-(Figure~\ref{fig-pbmc-mse}) were calculated. tSNE with a perplexity
-value set to \(30\), the default parameter setting, is considered as a
-reasonable representation for the PBMC3k dataset. As shown in
-Figure~\ref{fig-NLDR-variety} (e), there are three well-separated
-clusters, although the separation between the clusters are small.
-Additionally, non-linear structures can also be observed within the
-clusters (Figure~\ref{fig-NLDR-variety} (e)). The visualization of the
-\emph{model-in-the-data-space} provides evidence for this
-(\textbf{?@fig-pbmc2-sc}). Therefore, tSNE with perplexity \(30\)
-accurately captures the data structure of the PBMC3k dataset, which is
-better than UMAP.
 

### 2024-09-07 paper.qmd
Updated PBMC3k section

@@
 
-As shown in @fig-NLDR-variety, layout c, d, and e that show small separations between clusters are universally optimal.
+As a result, it is necessary to find an appropriate layout for the dataset.
+MSE for different binwidths ($a_1$) using tSNE, UMAP, PHATE, PaCMAP, and TriMAP with various (hyper-)parameter settings were computed (@fig-pbmc-mse).
+Layouts c, d, and e, which show small separations between clusters, are universally optimal.
+However, layout d performs well with smaller binwidths and poorly with larger binwidths.
+On the other hand, layout e performs well with larger binwidths.
+Layout c was selected for further analysis due to its better (hyper-)parameter selection for the same method of the published plot.
+
+The visualization of the selected layout in the $9\text{-}D$ shows edges between the clusters (@fig-mnist-tri-proj (b2)).
+This supports the presence of small separations between clusters.
+Additionally, the data points are not uniformly distributed across the clusters, resulting in dense areas (@fig-mnist-tri-proj (b2)).
+Furthermore, the clusters shows non-linear shapes (@fig-mnist-tri-proj (b3)).
+
+<!-- As shown in @fig-NLDR-variety, layout c, d, and e that show small separations between clusters are universally optimal.
 But layout d with little separations perform well at tiny binwidth and poorly as binwidth increases.
@@
 The layout c is chosen as the best, since it used the UMAP with a different (hyper-) parameter settings.
-
-The visualization of the model generated for the selected layout in $9\text{-}D$ shows edges between the clusters (@fig-mnist-tri-proj (b2)).
+-->
+
+<!-- The visualization of the model generated for the selected layout in $9\text{-}D$ shows edges between the clusters (@fig-mnist-tri-proj (b2)).
 This provides evidence for the small separations between clusters.
@@
 Also, the clusters are non-linear shaped (@fig-mnist-tri-proj (b3).
+-->
 

### 2024-09-07 paper.tex
Updated PBMC3k section

@@
 
-As shown in Figure~\ref{fig-NLDR-variety}, layout c, d, and e that show
-small separations between clusters are universally optimal. But layout d
-with little separations perform well at tiny binwidth and poorly as
-binwidth increases. Also, layout e performs well at large binwidths. The
-layout c is chosen as the best, since it used the UMAP with a different
-(hyper-) parameter settings.
-
-The visualization of the model generated for the selected layout in
-\(9\text{-}D\) shows edges between the clusters
-(Figure~\ref{fig-mnist-tri-proj} (b2)). This provides evidence for the
-small separations between clusters. Furthermore, the data does not
-uniformally distributed along the clusters and therefore shows densed
-points (Figure~\ref{fig-mnist-tri-proj} (b2). Also, the clusters are
-non-linear shaped (Figure~\ref{fig-mnist-tri-proj} (b3).
+As a result, it is necessary to find an appropriate layout for the
+dataset. MSE for different binwidths (\(a_1\)) using tSNE, UMAP, PHATE,
+PaCMAP, and TriMAP with various (hyper-)parameter settings were computed
+(Figure~\ref{fig-pbmc-mse}). Layouts c, d, and e, which show small
+separations between clusters, are universally optimal. However, layout d
+performs well with smaller binwidths and poorly with larger binwidths.
+On the other hand, layout e performs well with larger binwidths. Layout
+c was selected for further analysis due to its better (hyper-)parameter
+selection for the same method of the published plot.
+
+The visualization of the selected layout in the \(9\text{-}D\) shows
+edges between the clusters (Figure~\ref{fig-mnist-tri-proj} (b2)). This
+supports the presence of small separations between clusters.
+Additionally, the data points are not uniformly distributed across the
+clusters, resulting in dense areas (Figure~\ref{fig-mnist-tri-proj}
+(b2)). Furthermore, the clusters shows non-linear shapes
+(Figure~\ref{fig-mnist-tri-proj} (b3)).
 

### 2024-09-07 paper.qmd
Updated Figure 14 title

@@
 #| echo: false
-#| fig-cap: "Model in \\gD{}, on the tSNE layout (a), and three views of the fit in projections from $9\\text{-}D$ (b, c, d), for the PBMC3k data ($a_1 = 0.099$, $b = 221 \\  (13, \ 17)$, $m = 92$, benchmark value to remove large edges is $0.1$).
-The three well-separated clusters with small separations observed in the \\gD{} layout of tSNE (@fig-NLDR-variety (e)) is also visible when visualizing the model overlaid on the data space in $9\\text{-}D$.
-Furthermore, there are some non-linear clusters.
-The model shows densed points within clusters in $9\\text{-}D$ space, which is an additional patterns not visible in the \\gD{} representation (@fig-NLDR-variety (e)).
-The full video is available at <https://youtu.be/5Y1hE4i7N2k>."
+#| fig-cap: "Model in \\gD{}, on the layouts, and two views of the fit in projections from $9\\text{-}D$, for the PBMC3k data.
+The \\gD{} model shows three well-separated clusters with large gaps in a1 ($a_1 = 0.05$, $b = 676 \\  (26, \ 26)$, $m = 92$, benchmark value to remove large edges is $0.08$), while a2 ($a_1 = 0.10$, $b = 195 \\  (13, \ 15)$, $m = 90$, benchmark value to remove large edges is $0.166$) shows small separation between the clusters.
+Because of the small separation, the three clusters are connected in \\gD{} and appear as long edges in $9\\text{-}D$.
+The full video for the model for layout a is available at <https://youtu.be/5Y1hE4i7N2k>.
+The full video for the model for layout c is available at <https://youtu.be/5Y1hE4i7N2k>."
 #| label: fig-model-pbmc-author-proj
@@
 #| fig-width: 10
-#| fig-height: 10
+#| fig-height: 15
+#| out-height: 80%
 

### 2024-09-07 paper.tex
Updated Figure 14 title

@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-model-pbmc-author-proj-1.pdf}
-
-}
-
-\caption{\label{fig-model-pbmc-author-proj}Model in \gD{}, on the tSNE
-layout (a), and three views of the fit in projections from
-\(9\text{-}D\) (b, c, d), for the PBMC3k data (\(a_1 = 0.099\),
-\(b = 221 \  (13,  17)\), \(m = 92\), benchmark value to remove large
-edges is \(0.1\)). The three well-separated clusters with small
-separations observed in the \gD{} layout of tSNE
-(Figure~\ref{fig-NLDR-variety} (e)) is also visible when visualizing the
-model overlaid on the data space in \(9\text{-}D\). Furthermore, there
-are some non-linear clusters. The model shows densed points within
-clusters in \(9\text{-}D\) space, which is an additional patterns not
-visible in the \gD{} representation (Figure~\ref{fig-NLDR-variety} (e)).
-The full video is available at \url{https://youtu.be/5Y1hE4i7N2k}.}
+\includegraphics[width=1\textwidth,height=0.8\textheight]{paper_files/figure-pdf/fig-model-pbmc-author-proj-1.pdf}
+
+}
+
+\caption{\label{fig-model-pbmc-author-proj}Model in \gD{}, on the
+layouts, and two views of the fit in projections from \(9\text{-}D\),
+for the PBMC3k data. The \gD{} model shows three well-separated clusters
+with large gaps in a1 (\(a_1 = 0.05\), \(b = 676 \  (26,  26)\),
+\(m = 92\), benchmark value to remove large edges is \(0.08\)), while a2
+(\(a_1 = 0.10\), \(b = 195 \  (13,  15)\), \(m = 90\), benchmark value
+to remove large edges is \(0.166\)) shows small separation between the
+clusters. Because of the small separation, the three clusters are
+connected in \gD{} and appear as long edges in \(9\text{-}D\). The full
+video for the model for layout a is available at
+\url{https://youtu.be/5Y1hE4i7N2k}. The full video for the model for
+layout c is available at \url{https://youtu.be/5Y1hE4i7N2k}.}
 

### 2024-09-07 paper.qmd
Added intro section for Five Gausssian cluster example

@@
 
-In this section, we will assess how t-SNE, UMAP, and PaCMAP algorithms preserve the clustering structure.
-We will use a simulated dataset consisting of five spherical Gaussian clusters where each cluster occupies a different corner of the $4\text{-}D$ space, with each cluster containing an equal number of points ($1000$) and the same within-cluster variation.
-The data simulated from $\Sigma = \begin{pmatrix}
-0.0025 & 0 & 0 & 0 \\
-0 & 0.0025 & 0 & 0 \\
-0 & 0 & 0.0025 & 0 \\
-0 & 0 & 0 & 0.0025 \\
-\end{pmatrix}$.
+In this section, the evaluation focuses on how effectively the tSNE, UMAP, and PaCMAP algorithms preserve the clustering structure.
+The data consists of five $4\text{-}D$ clusters, each with $1000$ points generated from a multivariate normal distribution with a small variance of 0.0025.
+The first cluster is centered at $(0, 0, 0, 0)$, the second at $(1, 0, 0, 0)$, the third at $(0, 1, 0, 0)$, the fourth at $(0, 0, 1, 0)$, and the fifth at $(0, 0, 0, 1)$.
+
+<!-- In this section, we will assess how tSNE, UMAP, and PaCMAP algorithms preserve the clustering structure.
+We use a simulated dataset consisting of five spherical Gaussian clusters where each cluster occupies a different corner of the $4\text{-}D$ space, with each cluster containing an equal number of points ($1000$) and the same within-cluster variation.
+The data simulated from $\Sigma = \begin{pmatrix} -->
+<!-- 0.0025 & 0 & 0 & 0 \\ -->
+<!-- 0 & 0.0025 & 0 & 0 \\ -->
+<!-- 0 & 0 & 0.0025 & 0 \\ -->
+<!-- 0 & 0 & 0 & 0.0025 \\ -->
+<!-- \end{pmatrix}$. -->
 

### 2024-09-07 paper.tex
Added intro section for Five Gausssian cluster example

@@
 
-In this section, we will assess how t-SNE, UMAP, and PaCMAP algorithms
-preserve the clustering structure. We will use a simulated dataset
-consisting of five spherical Gaussian clusters where each cluster
-occupies a different corner of the \(4\text{-}D\) space, with each
-cluster containing an equal number of points (\(1000\)) and the same
-within-cluster variation. The data simulated from
-\(\Sigma = \begin{pmatrix}
-0.0025 & 0 & 0 & 0 \\
-0 & 0.0025 & 0 & 0 \\
-0 & 0 & 0.0025 & 0 \\
-0 & 0 & 0 & 0.0025 \\
-\end{pmatrix}\).
+In this section, the evaluation focuses on how effectively the tSNE,
+UMAP, and PaCMAP algorithms preserve the clustering structure. The data
+consists of five \(4\text{-}D\) clusters, each with \(1000\) points
+generated from a multivariate normal distribution with a small variance
+of 0.0025. The first cluster is centered at \((0, 0, 0, 0)\), the second
+at \((1, 0, 0, 0)\), the third at \((0, 1, 0, 0)\), the fourth at
+\((0, 0, 1, 0)\), and the fifth at \((0, 0, 0, 1)\).
 

### 2024-09-07 paper.qmd
Updated Five Gau second paragraph

@@
 In this section, the evaluation focuses on how effectively the tSNE, UMAP, and PaCMAP algorithms preserve the clustering structure.
-The data consists of five $4\text{-}D$ clusters, each with $1000$ points generated from a multivariate normal distribution with a small variance of 0.0025.
+The data consists of five $4\text{-}D$ clusters, each with $1000$ points generated from a multivariate normal distribution with a small variance of $0.0025$.
 The first cluster is centered at $(0, 0, 0, 0)$, the second at $(1, 0, 0, 0)$, the third at $(0, 1, 0, 0)$, the fourth at $(0, 0, 1, 0)$, and the fifth at $(0, 0, 0, 1)$.
@@
 
-In the t-SNE 2D layout, the distances between clusters appear smaller, leading to more tightly packed clusters, which is typical of t-SNE's focus on preserving local structures.
+In the tSNE layout (@fig-five-gau-projs (a1)), the distances between clusters appear smaller, leading to more tightly packed clusters, which is typical of tSNE's focus on preserving local structures.
 This often results in less emphasis on the global distances between clusters.
-On the other hand, UMAP tends to maintain larger separations between clusters, better reflecting the global structure and preserving the relative distances between clusters more effectively.
-PaCMAP, meanwhile, balances these aspects by maintaining the distance between clusters and positioning them correctly, effectively capturing both global and local structures.
+On the other hand, UMAP (@fig-five-gau-projs (b1)) tends to maintain larger separations between clusters, better reflecting the global structure and preserving the relative distances between clusters more effectively.
+In contrast, PaCMAP (@fig-five-gau-projs (c1)) aim to balance these aspects by maintaining the distance between clusters and accurately positioning them, effectively capturing both global and local structures.
 
@@
 ```{r}
-#| label: fig-five-gau-nldr-layouts
-#| fig-height: 5
+#| label: fig-five-gau-projs
+#| fig-height: 15
 #| fig-width: 15
@@
 tsne_gau + umap_gau + pacmap_gau +
-  plot_layout(ncol = 3)
-```
-
-```{r}
-#| label: fig-five-gau-projs
-#| fig-height: 10
-#| fig-width: 15
-#| fig-pos: H
-#| fig-cap: "(a) tSNE, (b) UMAP, and (c) PaCMAP"
-
 five_gau_proj_tsne_model1 + five_gau_proj_umap_model1 + five_gau_proj_pacmap_model1 + five_gau_proj_tsne_model2 + five_gau_proj_umap_model2 + five_gau_proj_pacmap_model2 +
-  plot_layout(guides = "collect", nrow = 2) &
+  plot_layout(guides = "collect", nrow = 3) &
   theme(legend.position='none')

### 2024-09-07 paper.tex
Updated Five Gau second paragraph

@@
 generated from a multivariate normal distribution with a small variance
-of 0.0025. The first cluster is centered at \((0, 0, 0, 0)\), the second
-at \((1, 0, 0, 0)\), the third at \((0, 1, 0, 0)\), the fourth at
+of \(0.0025\). The first cluster is centered at \((0, 0, 0, 0)\), the
+second at \((1, 0, 0, 0)\), the third at \((0, 1, 0, 0)\), the fourth at
 \((0, 0, 1, 0)\), and the fifth at \((0, 0, 0, 1)\).
 
-In the t-SNE 2D layout, the distances between clusters appear smaller,
-leading to more tightly packed clusters, which is typical of t-SNE's
-focus on preserving local structures. This often results in less
-emphasis on the global distances between clusters. On the other hand,
-UMAP tends to maintain larger separations between clusters, better
-reflecting the global structure and preserving the relative distances
-between clusters more effectively. PaCMAP, meanwhile, balances these
-aspects by maintaining the distance between clusters and positioning
-them correctly, effectively capturing both global and local structures.
+In the tSNE layout (Figure~\ref{fig-five-gau-projs} (a1)), the distances
+between clusters appear smaller, leading to more tightly packed
+clusters, which is typical of tSNE's focus on preserving local
+structures. This often results in less emphasis on the global distances
+between clusters. On the other hand, UMAP
+(Figure~\ref{fig-five-gau-projs} (b1)) tends to maintain larger
+separations between clusters, better reflecting the global structure and
+preserving the relative distances between clusters more effectively. In
+contrast, PaCMAP (Figure~\ref{fig-five-gau-projs} (c1)) aim to balance
+these aspects by maintaining the distance between clusters and
+accurately positioning them, effectively capturing both global and local
+structures.
 
@@
 the best results in any given dataset.
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-five-gau-nldr-layouts-1.pdf}
-
-}
-
-\caption{\label{fig-five-gau-nldr-layouts}(a) tSNE, (b) UMAP, and (c)
-PaCMAP}
-
-\end{figure}%
 

### 2024-09-07 paper.qmd
Updated Five Gaussian cluster description

@@
 
-The t-SNE, UMAP, and PaCMAP layouts each reveal different aspects of the clustering structure within the dataset.
-When examining a specific cluster, which appears more like a filled-out pancake in higher dimensions, we observe that t-SNE, UMAP, and PaCMAP all effectively capture global structures.
-However, there is evidence suggesting that PaCMAP, with its default hyper-parameter settings, struggles to preserve the local structure as effectively as t-SNE and UMAP.
+By visualizing the model of different layouts in $4\text{-}D$ reveal different characteristics of each NLDR method.
+When examining the model of a specific cluster, which appears more like a filled-out shape in tSNE, and UMAP, while the model has a pancake shape in PaCMAP.
+The reason to have filled-out shaped clusters in tSNE and UMAP is that the ability of capture the local structure.
+On the other hand, the reason for the pancake shape in PaCMAP is the fail to capture the local structure.
+
+<!-- The tSNE, UMAP, and PaCMAP layouts each reveal different characteristics of the clustering structure within the dataset.  -->
+
+<!-- When examining the model of a specific cluster, which appears more like a filled-out pancake in $4\text{-}D$, we observe that tSNE, UMAP, and PaCMAP all effectively capture global structures.
+However, there is evidence suggesting that PaCMAP, with its default hyper-parameter settings, struggles to preserve the local structure as effectively as tSNE and UMAP.
+-->
 
@@
 This parameter significantly impacts how well local structures are maintained.
-Additionally, the intrinsic dimensionality of the data, the balance between global and local structures, and issues related to optimization and initialization all contribute to the observed outcomes.
-
-It's important to note that these observations are made with specific hyper-parameter settings.
-Adjusting these parameters could lead to different results, potentially improving PaCMAP's ability to capture local structures or altering how t-SNE and UMAP handle global and local structures.
-Thus, while PaCMAP generally balances global and local preservation, careful tuning of its parameters is crucial for achieving the best results in any given dataset.
+Additionally, the balance between global and local structures, and issues related to optimization and initialization all contribute to the observed outcomes.
+
+It's important to note that these layouts are made with default hyper-parameter settings.
+Adjusting these parameters could lead to different results, potentially improving PaCMAP's ability to capture local structures or altering how tSNE and UMAP handle global and local structures.
+Thus, while PaCMAP generally balances global and local preservation, careful tuning of its parameters is important for achieving the best results in any given dataset.
 

### 2024-09-07 paper.tex
Updated Five Gaussian cluster description

@@
 
-The t-SNE, UMAP, and PaCMAP layouts each reveal different aspects of the
-clustering structure within the dataset. When examining a specific
-cluster, which appears more like a filled-out pancake in higher
-dimensions, we observe that t-SNE, UMAP, and PaCMAP all effectively
-capture global structures. However, there is evidence suggesting that
-PaCMAP, with its default hyper-parameter settings, struggles to preserve
-the local structure as effectively as t-SNE and UMAP.
+By visualizing the model of different layouts in \(4\text{-}D\) reveal
+different characteristics of each NLDR method. When examining the model
+of a specific cluster, which appears more like a filled-out shape in
+tSNE, and UMAP, while the model has a pancake shape in PaCMAP. The
+reason to have filled-out shaped clusters in tSNE and UMAP is that the
+ability of capture the local structure. On the other hand, the reason
+for the pancake shape in PaCMAP is the fail to capture the local
+structure.
 
@@
 This parameter significantly impacts how well local structures are
-maintained. Additionally, the intrinsic dimensionality of the data, the
-balance between global and local structures, and issues related to
-optimization and initialization all contribute to the observed outcomes.
-
-It's important to note that these observations are made with specific
+maintained. Additionally, the balance between global and local
+structures, and issues related to optimization and initialization all
+contribute to the observed outcomes.
+
+It's important to note that these layouts are made with default
 hyper-parameter settings. Adjusting these parameters could lead to
 different results, potentially improving PaCMAP's ability to capture
-local structures or altering how t-SNE and UMAP handle global and local
+local structures or altering how tSNE and UMAP handle global and local
 structures. Thus, while PaCMAP generally balances global and local
-preservation, careful tuning of its parameters is crucial for achieving
-the best results in any given dataset.
+preservation, careful tuning of its parameters is important for
+achieving the best results in any given dataset.
 

### 2024-09-10 paper.qmd
Added Figure 11 with a box

@@
              y = tSNE2)) +
-  geom_point(alpha=0.3, color = "#000000")
+  geom_point(alpha=0.3, color = "#000000") +
+  geom_rect(aes(xmin = -0.05, xmax = 0.35, ymin = 0.28, ymax = 0.6),
+               fill = "transparent", color = "red", linewidth = 0.8)
 

### 2024-09-10 paper.qmd
Added references for Section 6

@@
 By visualizing the model of different layouts in $4\text{-}D$ reveal different characteristics of each NLDR method.
-When examining the model of a specific cluster, which appears more like a filled-out shape in tSNE, and UMAP, while the model has a pancake shape in PaCMAP.
+When examining the model of a specific cluster, which appears more like a filled-out shape in tSNE (@fig-five-gau-projs (a2, a3)), and UMAP (@fig-five-gau-projs (b2, b3)), while the model has a pancake shape in PaCMAP (@fig-five-gau-projs (c2, c3)).
 The reason to have filled-out shaped clusters in tSNE and UMAP is that the ability of capture the local structure.

### 2024-09-10 paper.tex
Added references for Section 6

@@
 of a specific cluster, which appears more like a filled-out shape in
-tSNE, and UMAP, while the model has a pancake shape in PaCMAP. The
+tSNE (Figure~\ref{fig-five-gau-projs} (a2, a3)), and UMAP
+(Figure~\ref{fig-five-gau-projs} (b2, b3)), while the model has a
+pancake shape in PaCMAP (Figure~\ref{fig-five-gau-projs} (c2, c3)). The
 reason to have filled-out shaped clusters in tSNE and UMAP is that the

### 2024-09-10 paper.qmd
Added Figure 11 title

@@
 #| fig-pos: H
-#| fig-cap: "(a) tSNE, (b) UMAP, and (c) PaCMAP"
+#| fig-cap: "Model in \\gD{}, on the layouts from tSNE, UMAP, and PaCMAP (a1, b1, c1).
+Two views of the fit in projections of each layout from $4\\text{-}D$, for the Gaussian cluster data.
+All \\gD{} models show five well-separated clusters with different gaps.
+Layout a1 shows small separations, while layout b2 shows large gaps.
+On the other hand, layout b3 shows average gaps with exact positions as in $4\\text{-}D$.
+Assessing the different models fit for different layouts for a specific cluster shows filled-out shape in a1, a2, and b1, b2.
+On the other hand, c1 and c2 shows pancake shape.
+The full video for the model for layout a, b, and c are available at <https://youtu.be/5Y1hE4i7N2k>, <https://youtu.be/5Y1hE4i7N2k>, and <https://youtu.be/5Y1hE4i7N2k> respectively."
 

### 2024-09-10 paper.tex
Added Figure 11 title

@@
 
-\caption{\label{fig-five-gau-projs}(a) tSNE, (b) UMAP, and (c) PaCMAP}
+\caption{\label{fig-five-gau-projs}Model in \gD{}, on the layouts from
+tSNE, UMAP, and PaCMAP (a1, b1, c1). Two views of the fit in projections
+of each layout from \(4\text{-}D\), for the Gaussian cluster data. All
+\gD{} models show five well-separated clusters with different gaps.
+Layout a1 shows small separations, while layout b2 shows large gaps. On
+the other hand, layout b3 shows average gaps with exact positions as in
+\(4\text{-}D\). Assessing the different models fit for different layouts
+for a specific cluster shows filled-out shape in a1, a2, and b1, b2. On
+the other hand, c1 and c2 shows pancake shape. The full video for the
+model for layout a, b, and c are available at
+\url{https://youtu.be/5Y1hE4i7N2k}, \url{https://youtu.be/5Y1hE4i7N2k},
+and \url{https://youtu.be/5Y1hE4i7N2k} respectively.}
 

### 2024-09-10 paper.qmd
Figure 11 updated the titles

@@
   ylim(c(-0.9, -0.5)) +
-  interior_annotation("a1", 
+  interior_annotation("a2", 
                       position = c(0.08, 0.9),
@@
   ylim(c(-0.75, -0.4)) +
-  interior_annotation("a2", 
+  interior_annotation("a3", 
                       position = c(0.08, 0.9),
@@
   ylim(c(0, -0.38)) +
-  interior_annotation("b1", 
+  interior_annotation("b2", 
                       position = c(0.08, 0.9),
@@
   ylim(c(-0.75, -0.4)) +
-  interior_annotation("b2", 
+  interior_annotation("b3", 
                       position = c(0.08, 0.9),
@@
   ylim(c(0.18, 0.52)) +
-  interior_annotation("c1", 
+  interior_annotation("c2", 
                       position = c(0.08, 0.9),
@@
   ylim(c(-0.75, -0.4))+
-  interior_annotation("c2", 
+  interior_annotation("c3", 
                       position = c(0.08, 0.9),

### 2024-09-10 paper.qmd
Need to update Figure 11

@@
 
-trimesh_removed_gau1 <- vis_rmlg_mesh(
-  distance_edges = distance_gau1,
-  benchmark_value = benchmark_gau1,
-  tr_coord_df = tr_from_to_df_gau1,
-  distance_col = "distance")
+# trimesh_removed_gau1 <- vis_rmlg_mesh(
+#   distance_edges = distance_gau1,
+#   benchmark_value = benchmark_gau1,
+#   tr_coord_df = tr_from_to_df_gau1,
+#   distance_col = "distance")
+
+tr_df <- distinct(tibble(
+  x = c(tr_from_to_df_gau1[["x_from"]], tr_from_to_df_gau1[["x_to"]]), 
+  y = c(tr_from_to_df_gau1[["y_from"]], tr_from_to_df_gau1[["y_to"]])))
+
+distance_df_small_edges_gau1 <- distance_gau1 |>
+  filter(distance < benchmark_gau1)
+
+tr_from_to_df_gau1 <- inner_join(
+  tr_from_to_df_gau1, distance_df_small_edges_gau1, 
+  by = c("from", "to"))
+
+trimesh_removed_gau_tsne <- ggplot() + 
+  geom_segment(data = tr_from_to_df_gau1, 
+               aes(
+                 x = x_from, 
+                 y = y_from, 
+                 xend = x_to, 
+                 yend = y_to),
+               colour = "#33a02c") +
+  geom_point(
+    data = tsne_gau_scaled,
+    aes(
+      x = tSNE1,
+      y = tSNE2
+    ),
+    alpha = 0.3,
+    size = 0.1,
+    color = "#000000"
+  )  +
+  coord_fixed() +
+  interior_annotation("a1", 
+                      position = c(0.08, 0.95),
+                      cex = 2)
 
@@
 
+tr_df <- distinct(tibble(
+  x = c(tr_from_to_df_gau1[["x_from"]], tr_from_to_df_gau1[["x_to"]]), 
+  y = c(tr_from_to_df_gau1[["y_from"]], tr_from_to_df_gau1[["y_to"]])))
+
+distance_df_small_edges_gau1 <- distance_gau1 |>
+  filter(distance < benchmark_gau1)
+
+tr_from_to_df_gau1 <- inner_join(
+  tr_from_to_df_gau1, distance_df_small_edges_gau1, 
+  by = c("from", "to"))
+
+trimesh_removed_gau_umap <- ggplot() + 
+  geom_segment(data = tr_from_to_df_gau1, 
+               aes(
+                 x = x_from, 
+                 y = y_from, 
+                 xend = x_to, 
+                 yend = y_to),
+               colour = "#33a02c") +
+  geom_point(
+    data = umap_gau_scaled,
+    aes(
+      x = UMAP1,
+      y = UMAP2
+    ),
+    alpha = 0.3,
+    size = 0.1,
+    color = "#000000"
+  )  +
+  coord_fixed() +
+  interior_annotation("b1", 
+                      position = c(0.06, 0.95),
+                      cex = 2)
+
 ## Hexagonal binning to have regular hexagons
@@
 
-trimesh_removed_gau1 <- vis_rmlg_mesh(
-  distance_edges = distance_gau1,
-  benchmark_value = benchmark_gau1,
-  tr_coord_df = tr_from_to_df_gau1,
-  distance_col = "distance")
+# trimesh_removed_gau1 <- vis_rmlg_mesh(
+#   distance_edges = distance_gau1,
+#   benchmark_value = benchmark_gau1,
+#   tr_coord_df = tr_from_to_df_gau1,
+#   distance_col = "distance")
+
+tr_df <- distinct(tibble(
+  x = c(tr_from_to_df_gau1[["x_from"]], tr_from_to_df_gau1[["x_to"]]), 
+  y = c(tr_from_to_df_gau1[["y_from"]], tr_from_to_df_gau1[["y_to"]])))
+
+distance_df_small_edges_gau1 <- distance_gau1 |>
+  filter(distance < benchmark_gau1)
+
+tr_from_to_df_gau1 <- inner_join(
+  tr_from_to_df_gau1, distance_df_small_edges_gau1, 
+  by = c("from", "to"))
+
+trimesh_removed_gau_pacmap <- ggplot() + 
+  geom_segment(data = tr_from_to_df_gau1, 
+               aes(
+                 x = x_from, 
+                 y = y_from, 
+                 xend = x_to, 
+                 yend = y_to),
+               colour = "#33a02c") +
+  geom_point(
+    data = pacmap_gau_scaled,
+    aes(
+      x = PaCMAP1,
+      y = PaCMAP2
+    ),
+    alpha = 0.3,
+    size = 0.1,
+    color = "#000000"
+  )  +
+  coord_fixed() +
+  interior_annotation("c1", 
+                      position = c(0.08, 0.95),
+                      cex = 2)
 
@@
 
-tsne_gau + umap_gau + pacmap_gau +
+trimesh_removed_gau_tsne + trimesh_removed_gau_umap + trimesh_removed_gau_pacmap +
 five_gau_proj_tsne_model1 + five_gau_proj_umap_model1 + five_gau_proj_pacmap_model1 + five_gau_proj_tsne_model2 + five_gau_proj_umap_model2 + five_gau_proj_pacmap_model2 +

### 2024-09-11 paper.qmd
Updated Figure 17

@@
 
-error_plot_mnist <- error_df |>
+highd_error_2d_mnist <- error_df |>
   ggplot(aes(x = tSNE1,
@@
 ```{r}
-#| label: img-with-error
+#| label: img-diff-pos
 
@@
 
-img_error_within <- c(5173, 6671, 2964, 2464, 6319, 7388, 4303, 7650)
+img_right_top <- c(6561, 454, 6397)
 
@@
   mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
-  filter(instance %in% img_error_within)
-
-imge_error_sample_within <- pixels_gathered_within |>
+  filter(instance %in% img_right_top)
+
+right_top_img <- pixels_gathered_within |>
   ggplot(aes(x, y, fill = value)) +
@@
 
-img_error_outside <- c(3554, 501, 4422, 6860, 3869, 1529, 4916, 6144)
+img_middle <- c(3443, 7404, 7242)
+
+pixels_gathered_within <-  mnist_data |>
+  mutate(instance = row_number()) |>
+  gather(pixel, value, -Label, -instance) |>
+  extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
+  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
+  filter(instance %in% img_middle)
+
+middle_img <- pixels_gathered_within |>
+  ggplot(aes(x, y, fill = value)) +
+  geom_tile() +
+  facet_wrap(~ instance, ncol = 4) +
+  coord_fixed() +
+  scale_fill_continuous_sequential(palette = "Grays") +
+  theme(strip.background = element_blank(),
+        strip.text.x = element_blank(),
+        legend.position = "none")
+
+img_right_bottom <- c(7098, 6456, 2)
+
+pixels_gathered_within <-  mnist_data |>
+  mutate(instance = row_number()) |>
+  gather(pixel, value, -Label, -instance) |>
+  extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
+  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
+  filter(instance %in% img_right_bottom)
+
+right_bottom_img <- pixels_gathered_within |>
+  ggplot(aes(x, y, fill = value)) +
+  geom_tile() +
+  facet_wrap(~ instance, ncol = 4) +
+  coord_fixed() +
+  scale_fill_continuous_sequential(palette = "Grays") +
+  theme(strip.background = element_blank(),
+        strip.text.x = element_blank(),
+        legend.position = "none")
+
+#img_error_outside <- c(3554, 501, 4422, 6860, 3869, 1529, 4916, 6144)
+img_error_outside <- c(501, 1529, 6144)
 
@@
 #| fig-pos: H
-#| fig-width: 4
-#| fig-height: 8
 #| out-height: 80%
 
-#imge_error_sample_within
-
-error_plot_mnist / (imge_error_sample_outside) +
-  plot_annotation(tag_levels = 'a') &
-  theme(plot.tag.position  = c(0.92, 0.92),
-        plot.tag = element_text(color = "grey70"))
+highd_error_2d_mnist + (imge_error_sample_outside / right_top_img / middle_img / right_bottom_img)
 ```

### 2024-09-11 paper.qmd
Updated Figure 17

@@
 
-error_df <- error_df |>
-  mutate(type = case_when(
-    row_wise_abs_error <= 2 ~ "error 0-2",
-    row_wise_abs_error <= 4 ~ "error 2-4",
-    row_wise_abs_error <= 6 ~ "error 4-6",
-    row_wise_abs_error <= 8 ~ "error 6-8",
-    row_wise_abs_error <= 10 ~ "error 8-10",
-    .default = "error greter than 10"
-  )) |>
-  mutate(type = factor(type, levels = c(
-    "error 0-2", "error 2-4", "error 4-6", "error 6-8", "error 8-10", 
-    "error greter than 10")))
+# error_df <- error_df |>
+#   mutate(type = case_when(
+#     row_wise_abs_error <= 2 ~ "error 0-2",
+#     row_wise_abs_error <= 4 ~ "error 2-4",
+#     row_wise_abs_error <= 6 ~ "error 4-6",
+#     row_wise_abs_error <= 8 ~ "error 6-8",
+#     row_wise_abs_error <= 10 ~ "error 8-10",
+#     .default = "error greter than 10"
+#   )) |>
+#   mutate(type = factor(type, levels = c(
+#     "error 0-2", "error 2-4", "error 4-6", "error 6-8", "error 8-10", 
+#     "error greter than 10")))
 
@@
 
+## To transform the data
+## ggplot(data = error_df, aes(x=row_wise_abs_error)) + geom_density() ## right skewed
+
+error_df <- error_df |>
+  mutate(sqrt_row_wise_abs_error = sqrt(row_wise_abs_error))
+
 highd_error_2d_mnist <- error_df |>
@@
              y = tSNE2,
-             color = type, 
-             group = ID)) +
-  geom_point(alpha=0.5, 
-             size = 0.1) +
-  scale_colour_brewer(
-    type = "qual",
-    palette = 2
-  ) + 
-  theme(
-      legend.position = "bottom",
-      legend.text = element_text(size = 10),          # Increase legend text size
-      legend.key.size = unit(1, "lines")              # Increase size of legend keys
-  ) +
-  guides(
-      color = guide_legend(
-          override.aes = list(size = 2.5, alpha = 1)      # Increase point size and opacity in legend
-      )
-  ) 
+             colour = sqrt_row_wise_abs_error)) +
+  geom_point(alpha=0.2) +
+  scale_colour_continuous_sequential(palette = "Light Grays")
+  # scale_colour_brewer(
+  #   type = "seq",
+  #   palette = 1
+  # ) #+ 
+  # theme(
+  #     legend.position = "bottom",
+  #     legend.text = element_text(size = 10),          # Increase legend text size
+  #     legend.key.size = unit(1, "lines")              # Increase size of legend keys
+  # ) +
+  # guides(
+  #     color = guide_legend(
+  #         override.aes = list(size = 2.5, alpha = 1)      # Increase point size and opacity in legend
+  #     )
+  # ) 
 ```
@@
 #| fig-pos: H
+#| fig-width: 10
+#| fig-height: 5
 #| out-height: 80%

### 2024-09-11 paper.qmd
Decide colors for the boxes

@@
   geom_point(alpha=0.2) +
-  scale_colour_continuous_sequential(palette = "Light Grays")
+  scale_colour_continuous_sequential(palette = "Light Grays") +
+  geom_rect(aes(xmin = 0.8, xmax = 0.9, ymin = 0.45, ymax = 0.6),
+               fill = "transparent", color = "#e41a1c", linewidth = 0.7) +
+  geom_rect(aes(xmin = 0.4, xmax = 0.5, ymin = 0.75, ymax = 0.9),
+               fill = "transparent", color = "#377eb8", linewidth = 0.7)  +
+  geom_rect(aes(xmin = 0.45, xmax = 0.55, ymin = 0.75, ymax = 0.9),
+               fill = "transparent", color = "#4daf4a", linewidth = 0.7)  +
+  geom_rect(aes(xmin = 0.55, xmax = 0.65, ymin = 0.75, ymax = 0.9),
+               fill = "transparent", color = "#a65628", linewidth = 0.7) 
   # scale_colour_brewer(
@@
 
-img_right_top <- c(6561, 454, 6397)
+img_right_top <- c(6048, 6650, 5119)
 

### 2024-09-11 paper.qmd
Added colors to images

--- 2024-09-11 paper.qmd
@@
 )
-interior_annotation <- function(label, position = c(0.92, 0.92), cex = 1) {
+interior_annotation <- function(label, position = c(0.92, 0.92), cex = 1, col="grey70") {
   annotation_custom(grid::textGrob(label = label,
       x = unit(position[1], "npc"), y = unit(position[2], "npc"),
-      gp = grid::gpar(cex = cex, col="grey70")))
+      gp = grid::gpar(cex = cex, col=col)))
 }
@@
   geom_point(alpha=0.2) +
-  scale_colour_continuous_sequential(palette = "Light Grays")
+  scale_colour_continuous_sequential(palette = "Light Grays") +
+  geom_rect(aes(xmin = 0.8, xmax = 0.9, ymin = 0.45, ymax = 0.6),
+               fill = "transparent", color = "#e41a1c", linewidth = 0.7) +
+  geom_rect(aes(xmin = 0.4, xmax = 0.5, ymin = 0.75, ymax = 0.9),
+               fill = "transparent", color = "#377eb8", linewidth = 0.7)  +
+  geom_rect(aes(xmin = 0.45, xmax = 0.55, ymin = 0.75, ymax = 0.9),
+               fill = "transparent", color = "#4daf4a", linewidth = 0.7)  +
+  geom_rect(aes(xmin = 0.55, xmax = 0.65, ymin = 0.75, ymax = 0.9),
+               fill = "transparent", color = "#984ea3", linewidth = 0.7) 
   # scale_colour_brewer(
@@
   coord_fixed() +
-  scale_fill_continuous_sequential(palette = "Grays") +
+  scale_fill_continuous_sequential(palette = "Blues 3") +
   theme(strip.background = element_blank(),
@@
   coord_fixed() +
-  scale_fill_continuous_sequential(palette = "Grays") +
+  scale_fill_continuous_sequential(palette = "Greens 3") +
   theme(strip.background = element_blank(),
@@
   coord_fixed() +
-  scale_fill_continuous_sequential(palette = "Grays") +
+  scale_fill_continuous_sequential(palette = "Purples 2") +
   theme(strip.background = element_blank(),
@@
   coord_fixed() +
-  scale_fill_continuous_sequential(palette = "Grays") +
+  scale_fill_continuous_sequential(palette = "Reds 2") +
   theme(strip.background = element_blank(),

### 2024-09-11 paper.qmd
Added squares to the positions

@@
                fill = "transparent", color = "#e41a1c", linewidth = 0.7) +
-  geom_rect(aes(xmin = 0.4, xmax = 0.5, ymin = 0.75, ymax = 0.9),
+  geom_rect(aes(xmin = 0.75, xmax = 0.85, ymin = 0.8, ymax = 0.95),
                fill = "transparent", color = "#377eb8", linewidth = 0.7)  +
-  geom_rect(aes(xmin = 0.45, xmax = 0.55, ymin = 0.75, ymax = 0.9),
+  geom_rect(aes(xmin = 0.25, xmax = 0.35, ymin = 0.55, ymax = 0.7),
                fill = "transparent", color = "#4daf4a", linewidth = 0.7)  +
-  geom_rect(aes(xmin = 0.55, xmax = 0.65, ymin = 0.75, ymax = 0.9),
+  geom_rect(aes(xmin = 0.5, xmax = 0.6, ymin = 0.15, ymax = 0.3),
                fill = "transparent", color = "#984ea3", linewidth = 0.7) 

### 2024-09-11 paper.qmd
Added squares for the positions

@@
                fill = "transparent", color = "#377eb8", linewidth = 0.7)  +
-  geom_rect(aes(xmin = 0.25, xmax = 0.35, ymin = 0.55, ymax = 0.7),
+  geom_rect(aes(xmin = 0.45, xmax = 0.55, ymin = 0.45, ymax = 0.6),
                fill = "transparent", color = "#4daf4a", linewidth = 0.7)  +
-  geom_rect(aes(xmin = 0.5, xmax = 0.6, ymin = 0.15, ymax = 0.3),
+  geom_rect(aes(xmin = 0.7, xmax = 0.8, ymin = 0.25, ymax = 0.4),
                fill = "transparent", color = "#984ea3", linewidth = 0.7) 
@@
 
-img_right_top <- c(6048, 6650, 5119)
+img_right_top <- c(4795, 259, 6564)
 
@@
 
-img_middle <- c(3443, 7404, 7242)
+img_middle <- c(1069, 4604, 3579)
 
@@
 
-img_right_bottom <- c(7098, 6456, 2)
+#img_right_bottom <- c(1695, 5338, 2805)
+img_right_bottom <- c(2035, 7710, 2517)
 

### 2024-09-11 paper.qmd
Updated Figure 14

@@
   scale_y_log10() +
-  ylab("log(MSE)") +
+  ylab("MSE") +
   xlab(expression(paste("binwidth (", a[1], ")"))) +
@@
   geom_point(alpha=0.1, size=1, colour='#ff7f00') +
-  interior_annotation("a", c(0.08, 0.93))
+  interior_annotation("a", c(0.08, 0.93)) +
+  theme(aspect.ratio = 1)
 
@@
   geom_point(alpha=0.1, size=1, colour='#a65628') +
-  interior_annotation("b")
+  interior_annotation("b") +
+  theme(aspect.ratio = 1)
 
@@
   geom_point(alpha=0.1, size=1, colour='#377eb8') +
-  interior_annotation("c")
+  interior_annotation("c") +
+  theme(aspect.ratio = 1)
 
@@
   geom_point(alpha=0.1, size=1, colour='#4daf4a') +
-  interior_annotation("d")
+  interior_annotation("d") +
+  theme(aspect.ratio = 1)
 
@@
   geom_point(alpha=0.1, size=1, colour='#e41a1c') +
-  interior_annotation("e")
+  interior_annotation("e") +
+  theme(aspect.ratio = 1)
 ```

### 2024-09-11 paper.qmd
Set the plot's aspect ratio to 1

@@
 #| out-height: 22%
+#| eval: false
 

### 2024-09-11 paper.tex
Set the plot's aspect ratio to 1

@@
 However, \gD{} representation shows a big non-linear cluster and a small
-cluster with a small gap (Figure~\ref{fig-tsne-best}).
+cluster with a small gap (\textbf{?@fig-tsne-best}).
 
@@
 digit 1 images varies along this non-linear clustering structure
-(Figure~\ref{fig-tsne-best}), while the small cluster contains the digit
-1 images with different patterns of the digit 1, unlike the usual
+(\textbf{?@fig-tsne-best}), while the small cluster contains the digit 1
+images with different patterns of the digit 1, unlike the usual
 (Figure~\ref{fig-model-error-mnist} (c)). This provides the evidence for
@@
 
-\includegraphics[width=1\textwidth,height=0.22\textheight]{paper_files/figure-pdf/fig-tsne-best-1.pdf}
-
-}
-
-\caption{\label{fig-tsne-best}The \(\gD{}\) layout from tSNE applied for
-the MNIST digit 1 dataset with defualt (hyper-)parameters. We use our
-model in the \(10\text{-}D\) space to assess whether this is an accurate
-representation of data structure present in \(10\text{-}D\), or if it is
-misleading. The angle of the digit 1 images varies along this structure.
-Images at the top-right of the \(\gD{}\) layout show the digit 1 angled
-more to the right, while those at the bottom-right show the digit 1
-angled more to the left.}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\centering{
-
 \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-mnist-tri-proj-1.pdf}
@@
 large edges is \(0.114\)). (b) The non-linear structure observed in the
-\gD{} layout of tSNE (Figure~\ref{fig-tsne-best}) is also visible when
+\gD{} layout of tSNE (\textbf{?@fig-tsne-best}) is also visible when
 visualizing the model overlaid on the data space. This indicates that
@@
 structure in \(10\text{-}D\) space, which is an additional pattern not
-visible in the \gD{} representation (Figure~\ref{fig-tsne-best}). (d)
-some long edges exist in the \(10\text{-}D\) space that are not
-recognized as long edges in the \gD{} representation. The full video is
-available at \url{https://youtu.be/zcg_GXBmqjA}.}
+visible in the \gD{} representation (\textbf{?@fig-tsne-best}). (d) some
+long edges exist in the \(10\text{-}D\) space that are not recognized as
+long edges in the \gD{} representation. The full video is available at
+\url{https://youtu.be/zcg_GXBmqjA}.}
 

### 2024-09-11 paper.qmd
Figurw 15 check with the scale factors

@@
     c(0.71503,0.00502,-0.04370,-0.01811,-0.01103,0.09877,0.06199,0.05396,0.07771,-0.01713),
-    c(0.00587,0.71879,0.03485,0.08284,-0.07534,-0.05420,0.04315,-0.00547,-0.02995,0.01810))
+    c(0.00587,0.71879,0.03485,0.08284,-0.07534,-0.05420,0.04315,-0.00547,-0.02995,0.01810)) * 1.3
 
@@
     aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
+  #coord_fixed() +
   xlim(c(-1, 1)) +
@@
     c(-0.33305,0.20223,0.06303,0.11124,0.13299,-0.30093,0.34082,0.15581,-0.08864,0.33573),
-    c(0.53615,0.14969,-0.09393,-0.20053,0.07992,-0.34791,0.13057,-0.06814,-0.16737,0.03716))
+    c(0.53615,0.14969,-0.09393,-0.20053,0.07992,-0.34791,0.13057,-0.06814,-0.16737,0.03716)) * 2
 
@@
     c(-0.12783,-0.50063,-0.02631,-0.00484,0.17733,0.37157,0.11310,0.09424,0.19844,0.19792),
-    c(-0.42542,0.29570,0.31106,-0.30307,0.00614,0.16714,0.21779,-0.00025,0.07020,-0.00689))
+    c(-0.42542,0.29570,0.31106,-0.30307,0.00614,0.16714,0.21779,-0.00025,0.07020,-0.00689)) * 1.5
 

### 2024-09-11 paper.qmd
Updated Figure 15 a)

@@
                  yend = y_to),
-               colour = "#33a02c") +
+               colour = "#33a02c",
+               linewidth = 1) +
   geom_point(data = tsne_minst_scaled,
@@
              ),
-             alpha=0.5, 
-             size = 0.1) +
-  coord_fixed() +
-  interior_annotation("a", sc_ltr_pos_mnist)
+             alpha=0.1) +
+  interior_annotation("a", sc_ltr_pos_mnist) +
+  theme(aspect.ratio = 1)
 

### 2024-09-19 paper.qmd
Added two curvilinear clusters for different MSE

@@
 
+```{r}
+#| label: read-two-curvilinear-clust-nldr
+# Read a variety of different NLDR representations of two_curvilinear_clust
+# and plot them on same aspect ratio
+tsne_two_curvilinear_clust <- read_rds("data/two_nonlinear_clusters/two_nonlinear_clusters_tsne_perplexity_30.rds")
+
+nldr_two_curvilinear_clust1 <- tsne_two_curvilinear_clust |>
+  ggplot(aes(x = tSNE1,
+             y = tSNE2))+
+  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
+  interior_annotation("a", c(0.08, 0.93)) +
+  theme(aspect.ratio = 1)
+
+umap_two_curvilinear_clust <- read_rds("data/two_nonlinear_clusters/two_nonlinear_clusters_umap_n-neigbors_15_min-dist_0.1.rds")
+
+nldr_two_curvilinear_clust2 <- umap_two_curvilinear_clust |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2)) +
+  geom_point(alpha=0.1, size=1, colour='#a65628') +
+  interior_annotation("b") +
+  theme(aspect.ratio = 1)
+
+phate_two_curvilinear_clust <- read_rds("data/two_nonlinear_clusters/two_nonlinear_clusters_phate_knn_5.rds")
+
+nldr_two_curvilinear_clust3 <- phate_two_curvilinear_clust |>
+  ggplot(aes(x = PHATE1,
+             y = PHATE2))+
+  geom_point(alpha=0.1, size=1, colour='#377eb8') +
+  interior_annotation("c") +
+  theme(aspect.ratio = 1)
+
+trimap_two_curvilinear_clust <- read_rds("data/two_nonlinear_clusters/two_nonlinear_clusters_trimap_n-inliers_12_n-outliers_4_n-random_3.rds")
+
+nldr_two_curvilinear_clust4 <- trimap_two_curvilinear_clust |>
+  ggplot(aes(x = TriMAP1,
+             y = TriMAP2))+
+  geom_point(alpha=0.1, size=1, colour='#4daf4a') +
+  interior_annotation("d") +
+  theme(aspect.ratio = 1)
+
+pacmap_two_curvilinear_clust <- read_rds("data/two_nonlinear_clusters/two_nonlinear_clusters_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")
+
+nldr_two_curvilinear_clust5 <- pacmap_two_curvilinear_clust |>
+  ggplot(aes(x = PaCMAP1,
+             y = PaCMAP2))+
+  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
+  interior_annotation("e") +
+  theme(aspect.ratio = 1)
+```
+
+```{r}
+#| label: combine-data-two_curvilinear_clust
+
+error_two_curvilinear_clust_umap <- read_rds("data/two_nonlinear_clusters/error_two_nonlinear_clusters_umap.rds")
+error_two_curvilinear_clust_tsne <- read_rds("data/two_nonlinear_clusters/error_two_nonlinear_clusters_tsne.rds")
+error_two_curvilinear_clust_phate <- read_rds("data/two_nonlinear_clusters/error_two_nonlinear_clusters_phate.rds")
+error_two_curvilinear_clust_trimap <- read_rds("data/two_nonlinear_clusters/error_two_nonlinear_clusters_trimap.rds")
+error_two_curvilinear_clust_pacmap <- read_rds("data/two_nonlinear_clusters/error_two_nonlinear_clusters_pacmap.rds")
+
+error_two_curvilinear_clust <- bind_rows(error_two_curvilinear_clust_umap, 
+                         error_two_curvilinear_clust_tsne,
+                         error_two_curvilinear_clust_phate,
+                         error_two_curvilinear_clust_trimap,
+                         error_two_curvilinear_clust_pacmap)
+
+error_two_curvilinear_clust <- error_two_curvilinear_clust |>
+  mutate(a1 = round(a1, 2)) |>
+  filter(bin1 >= 5) |>
+  group_by(method, a1) |>
+  filter(MSE == min(MSE)) |>
+  ungroup()
+```
+
+```{r}
+#| label: error-comp-two_curvilinear_clust
+
+error_plot_two_curvilinear_clust <- ggplot(error_two_curvilinear_clust, 
+                           aes(x = a1, 
+                               y = MSE, 
+                               colour = method)) + 
+  geom_point(size = 0.8) +
+  geom_line(linewidth = 0.3) + 
+  # geom_vline(xintercept = 15, linetype="solid", 
+  #            color = "black", linewidth=0.8, alpha = 0.5) +
+  scale_x_continuous(breaks = sort(unique(error_two_curvilinear_clust$a1))[append(seq(1, length(unique(error_two_curvilinear_clust$a1)), by = 5), 24)]) +
+  scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a', "#ff7f00",'#a65628')) +
+  scale_y_log10() +
+  ylab("MSE") +
+  xlab(expression(paste("binwidth (", a[1], ")"))) +
+  theme_minimal() +
+  theme(panel.border = element_rect(fill = 'transparent'),
+        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+        axis.ticks.x = element_line(),
+        axis.ticks.y = element_line(),
+        legend.position = "none",
+        axis.text.x = element_text(size = 7),
+        axis.text.y = element_text(size = 7),
+        axis.title.x = element_text(size = 7),
+        axis.title.y = element_text(size = 7))
+
+```
+
+```{r}
+#| fig-cap: "Assessing which of the 5 NLDR layouts on the two_curvilinear_clust digit 1 data is the better representation using MSE for varying binwidth ($a_1$).
+Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-e).
+Layout c is universally poor.
+Layouts a that show two close clusters are universally suboptimal.
+Layout a is the best choice."
+#| label: fig-two_curvilinear_clust-mse
+#| fig-pos: H
+#| out-height: 100%
+
+free(error_plot_two_curvilinear_clust) + wrap_plots(nldr_two_curvilinear_clust1, nldr_two_curvilinear_clust2, nldr_two_curvilinear_clust3, 
+                                    nldr_two_curvilinear_clust4, nldr_two_curvilinear_clust5, ncol = 2)
+```
+
 <!-- @fig-scurve-sc-best shows one of the best fits for S-curve data.

### 2024-09-19 paper.tex
Added two curvilinear clusters for different MSE

@@
 
+\centering{
+
+\includegraphics[width=1\textwidth,height=1\textheight]{paper_files/figure-pdf/fig-two_curvilinear_clust-mse-1.pdf}
+
+}
+
+\caption{\label{fig-two_curvilinear_clust-mse}Assessing which of the 5
+NLDR layouts on the two\_curvilinear\_clust digit 1 data is the better
+representation using MSE for varying binwidth (\(a_1\)). Colour used for
+the lines and points in the left plot and in the scatterplots represents
+NLDR layout (a-e). Layout c is universally poor. Layouts a that show two
+close clusters are universally suboptimal. Layout a is the best choice.}
+
+\end{figure}%
+
+\begin{figure}[H]
+
 \begin{minipage}{0.25\linewidth}

### 2024-09-19 scripts/two_curvilinear_clusters/04_gen_embeddings.R
Updated the Figure 15 title

@@ -4,13 +4,13 @@ library(readr)
 library(conflicted)
 
 library(Rtsne)
-library(uwot)
+library(umap) #predit for uwot not working
 library(phateR)
 library(reticulate)
 
 set.seed(20240110)
 
-conflicts_prefer(uwot::umap)
+conflicts_prefer(umap::umap)
 
 use_python("~/miniforge3/envs/pcamp_env/bin/python")
 use_condaenv("pcamp_env")
@@ -46,21 +46,37 @@ write_rds(tSNE_data, file = paste0("data/two_nonlinear_clusters/two_nonlinear_cl
 n_neighbors <- 15
 min_dist <- 0.1
 
-UMAP_data <- umap(data,
-                  n_neighbors = n_neighbors,
-                  min_dist = min_dist,
-                  n_components =  2,
-                  nn_method = "fnn",
-                  init ="spca") |>
-  as_tibble()
+# UMAP_model <- umap(data,
+#                   n_neighbors = n_neighbors,
+#                   min_dist = min_dist,
+#                   n_components =  2,
+#                   init ="spca")
+#
+# UMAP_data <- UMAP_model |>
+#   as_tibble()
 
 # UMAP_data <- UMAP_fit$layout |>
 #   tibble::as_tibble(.name_repair = "unique")
+
+UMAP_fit <- umap(data,
+                 n_neighbors = n_neighbors,
+                 n_components =  min_dist)
+
+UMAP_data <- UMAP_fit$layout |>
+  as_tibble()
 
 names(UMAP_data) <- c("UMAP1", "UMAP2")
 
 ## Run only once
 write_rds(UMAP_data, file = paste0("data/two_nonlinear_clusters/two_nonlinear_clusters_umap_n-neigbors_", n_neighbors, "_min-dist_", min_dist, ".rds"))
+
+# predict_UMAP_df <- predict(UMAP_fit, data) |>
+#   as_tibble()
+#
+# names(predict_UMAP_df) <- c("UMAP1", "UMAP2")
+#
+# ## Run only once
+# write_rds(predict_UMAP_df, file = "data/two_nonlinear_clusters/two_nonlinear_clusters_umap_predict.rds")
 
 ## PHATE
 knn <- 5

### 2024-09-19 paper.qmd
Updated the Figure 15 title
Updated Figure 15 letters in 2D layouts

@@
   geom_point(alpha=0.1, size=1, colour='#377eb8') +
-  interior_annotation("c") +
+  interior_annotation("c", position = c(0.94, 0.92)) +
   theme(aspect.ratio = 1)
@@
   geom_point(alpha=0.1, size=1, colour='#4daf4a') +
-  interior_annotation("d") +
+  interior_annotation("d", position = c(0.86, 0.92)) +
   theme(aspect.ratio = 1)
@@
 Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-e).
+All the layouts appear to be very similar.
 Layout c is universally poor.

### 2024-09-19 paper.qmd
Updated Figure 14

@@
   coord_fixed() +
-  interior_annotation("a1")
+  interior_annotation("a1") +
+  theme(
+    aspect.ratio = 1
+  )
 
@@
   coord_fixed() +
-  interior_annotation("b1")
+  interior_annotation("b1") +
+  theme(
+    aspect.ratio = 1
+  )
 

### 2024-09-22 paper.qmd
Updated gen_axes function

@@
 
-gen_axes <- function(data, proj_data, limits = 1, axis_pos_x = NULL, axis_pos_y = NULL) {
+gen_axes <- function(data, proj_data, proj, limits = 1, axis_pos_x = NULL, axis_pos_y = NULL) {
 
@@
                      y1 = adj(0, axis_pos_y), 
-                     x2 = adj(projection[, 1], axis_pos_x), 
-                     y2 = adj(projection[, 2], axis_pos_y))
+                     x2 = adj(proj[, 1], axis_pos_x), 
+                     y2 = adj(proj[, 2], axis_pos_y))
 
@@
   proj_data = projected_df,
+  proj = projection,
   limits = 0.2,
@@
   proj_data = projected_df,
+  proj = projection,
   limits = 0.3,
@@
   proj_data = projected_df, 
+  proj = projection,
   limits = 1.5,
@@
   proj_data = projected_df,  
+  proj = projection,
   limits = 1.5,
@@
   proj_data = projected_df,
+  proj = projection,
   limits = 1.5,

### 2024-09-22 paper.qmd
Initial work on removing axes (neglect overlapping)

@@
 
-gen_axes <- function(data, proj_data, proj, limits = 1, axis_pos_x = NULL, axis_pos_y = NULL) {
-
-  rng <- range(proj_data)
-  proj_data <- proj_data/max(abs(rng))
-  colnames(proj_data) <- c("P1", "P2")
-  proj_data <- data.frame(proj_data)
-  obs_labels <- as.character(1:nrow(data))
+gen_axes <- function(proj, limits = 1, axis_pos_x = NULL, axis_pos_y = NULL, axis_labels) {
 
@@
 
-  axis_labels <- colnames(data)
   rownames(axes) <- axis_labels
@@
 
-projected <- as.matrix(scaled_pbmc_data) %*% projection
+projection_scaled <- projection * 1.23
+
+# projection_filtered <- projection
+# projection_filtered[apply(projection_filtered, 1, function(row) all(row < 0)), ] <- 0.75
+
+rows_to_keep <- c(1, 2, 4, 5, 6)
+
+projection_filtered <- projection[rows_to_keep, ]
+
+projected <- as.matrix(scaled_pbmc_data) %*% projection_scaled
 
@@
 
-projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection
+projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection_scaled
 
@@
 axes_obj <- gen_axes(
-  data = scaled_pbmc_data, 
-  proj_data = projected_df, 
-  proj = projection,
+  proj = projection_filtered,
   limits = 1.5,
   axis_pos_x = -0.75,
-  axis_pos_y = -0.75)
+  axis_pos_y = -0.75,
+  axis_labels = paste0("p", rows_to_keep))
 

### 2024-09-22 paper.qmd
Updared gen_axes to remove axes with threshold for distance

@@
 
-gen_axes <- function(proj, limits = 1, axis_pos_x = NULL, axis_pos_y = NULL, axis_labels) {
+gen_axes <- function(proj, limits = 1, axis_pos_x = NULL, axis_pos_y = NULL, axis_labels, threshold) {
 
@@
   rownames(axes) <- axis_labels
+
+  ## To remove axes
+  axes <- axes |>
+    mutate(distance = sqrt((x2 - x1)^2 + (y2 - y1)^2)) |>
+    filter(distance >= threshold)
 
@@
 
-rows_to_keep <- c(1, 2, 4, 5, 6)
-
-projection_filtered <- projection[rows_to_keep, ]
+#rows_to_keep <- c(1, 2, 4, 5, 6)
+
+#projection_filtered <- projection[rows_to_keep, ]
 
@@
 axes_obj <- gen_axes(
-  proj = projection_filtered,
+  proj = projection,
   limits = 1.5,
@@
   axis_pos_y = -0.75,
-  axis_labels = paste0("p", rows_to_keep))
+  axis_labels = names(scaled_pbmc_data),
+  threshold = 0.07)
 

### 2024-09-22 paper.qmd
Changed to render autjor best model PBMC

@@
   axis_labels = names(scaled_pbmc_data),
-  threshold = 0.07)
+  threshold = 0.084)
 

### 2024-09-22 paper.qmd
Changed the gen_axes function

@@
 axes_obj <- gen_axes(  
-  data = scaled_gau_data,
-  proj_data = projected_df,
   proj = projection,
@@
   axis_pos_x = 0.15,
-  axis_pos_y = -0.85)
+  axis_pos_y = -0.85,
+  axis_labels = names(scaled_gau_data),
+  threshold = 0.084)
 
@@
   axis_pos_x = -0.7,
-  axis_pos_y = -0.7)
+  axis_pos_y = -0.7,
+  axis_labels = names(scaled_gau_data),
+  threshold = 0.084)
 
@@
   axis_pos_x = -0.55,
-  axis_pos_y = -0.05)
+  axis_pos_y = -0.05,
+  axis_labels = names(scaled_gau_data),
+  threshold = 0.084)
 
@@
   axis_pos_x = -0.37,
-  axis_pos_y = 0.22)
+  axis_pos_y = 0.22,
+  axis_labels = names(scaled_gau_data),
+  threshold = 0.084)
 
@@
 training_data_pbmc <- read_rds("data/pbmc3k/pbmc_pca_50.rds")
-names(training_data_pbmc) <- paste0("p", 1:50)
+names(training_data_pbmc) <- paste0("pc", 1:50)
 
@@
   is_rm_lwd_hex = FALSE,
-  col_start_highd = "p"
+  col_start_highd = "pc"
 )
@@
 
-projected <- as.matrix(scaled_pbmc_data) %*% projection
+projection_scaled <- projection * 1.23
+
+projected <- as.matrix(scaled_pbmc_data) %*% projection_scaled
 
@@
 axes_obj <- gen_axes(
-  data = scaled_pbmc_data, 
-  proj_data = projected_df, 
   proj = projection,
@@
   axis_pos_x = -0.75,
-  axis_pos_y = -0.75)
+  axis_pos_y = -0.75,
+  axis_labels = names(scaled_pbmc_data),
+  threshold = 0.084)
 
@@
   type_NLDR = "UMAP",
-  col_start = "p")
+  col_start = "pc")
 
@@
 axes_obj <- gen_axes(
-  data = scaled_pbmc_data, 
-  proj_data = projected_df,  
   proj = projection,
@@
 axes_obj <- gen_axes(
-  data = scaled_pbmc_data, 
-  proj_data = projected_df,
   proj = projection,
@@
 #| label: mnist-model-proj
+
 df_b_mnist <- df_bin_mnist |>
@@
     c(0.71503,0.00502,-0.04370,-0.01811,-0.01103,0.09877,0.06199,0.05396,0.07771,-0.01713),
-    c(0.00587,0.71879,0.03485,0.08284,-0.07534,-0.05420,0.04315,-0.00547,-0.02995,0.01810)) * 1.3
-
-projected <- as.matrix(scaled_mnist_data) %*% projection
+    c(0.00587,0.71879,0.03485,0.08284,-0.07534,-0.05420,0.04315,-0.00547,-0.02995,0.01810))
+
+projection_scaled <- projection * 1.3
+
+projected <- as.matrix(scaled_mnist_data) %*% projection_scaled
 
@@
 
-projected_model <- as.matrix(scaled_mnist_data_model) %*% projection
+projected_model <- as.matrix(scaled_mnist_data_model) %*% projection_scaled
 
@@
 axes_obj <- gen_axes(  
-  data = scaled_mnist_data, 
-  proj_data = projected_df, 
   proj = projection,
@@
   axis_pos_x = -0.75,
-  axis_pos_y = -0.75)
+  axis_pos_y = -0.75,
+  axis_labels = names(scaled_mnist_data),
+  threshold = 0.084)
 
@@
     c(-0.33305,0.20223,0.06303,0.11124,0.13299,-0.30093,0.34082,0.15581,-0.08864,0.33573),
-    c(0.53615,0.14969,-0.09393,-0.20053,0.07992,-0.34791,0.13057,-0.06814,-0.16737,0.03716)) * 2
-
-projected <- as.matrix(scaled_mnist_data) %*% projection
+    c(0.53615,0.14969,-0.09393,-0.20053,0.07992,-0.34791,0.13057,-0.06814,-0.16737,0.03716))
+
+projection_scaled <- projection * 2
+
+projected <- as.matrix(scaled_mnist_data) %*% projection_scaled
 
@@
 axes_obj <- gen_axes(  
-  data = scaled_mnist_data, 
-  proj_data = projected_df,
   proj = projection,
@@
     c(-0.12783,-0.50063,-0.02631,-0.00484,0.17733,0.37157,0.11310,0.09424,0.19844,0.19792),
-    c(-0.42542,0.29570,0.31106,-0.30307,0.00614,0.16714,0.21779,-0.00025,0.07020,-0.00689)) * 1.5
-
-projected <- as.matrix(scaled_mnist_data) %*% projection
+    c(-0.42542,0.29570,0.31106,-0.30307,0.00614,0.16714,0.21779,-0.00025,0.07020,-0.00689))
+
+projection_scaled <- projection * 1.5
+
+projected <- as.matrix(scaled_mnist_data) %*% projection_scaled
 

### 2024-09-22 paper.qmd
Fixed the issue of generating axis labels

@@
     data=axes,
-    aes(x=x2, y=y2, label=rownames(axes)), 
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
     colour="grey50",
-    size = 2) +
+    size = 3) +
   geom_path(
@@
     data=axes,
-    aes(x=x2, y=y2, label=rownames(axes)), 
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
     colour="grey50",

### 2024-09-22 paper.qmd
Fixed obtaining the axis of Gau cluster data

@@
   axis_labels = names(scaled_gau_data),
-  threshold = 0.084)
+  threshold = 0)
 
@@
     data=axes, 
-    aes(x=x2, y=y2, label=rownames(axes)), 
+    aes(x=x2, y=y2),
+    label=rownames(axes), 
     colour="grey50",
@@
     data=axes, 
-    aes(x=x2, y=y2, label=rownames(axes)), 
+    aes(x=x2, y=y2), 
+    label=rownames(axes), 
     colour="grey50",

### 2024-09-22 paper.qmd
Updated Figure 14 a1, b1

@@
                  yend = y_to),
-               colour = "#33a02c") +
+               colour = "#33a02c",
+               linewidth = 1) +
   geom_point(
@@
     ),
-    alpha = 0.3,
-    size = 0.1,
+    alpha = 0.1,
     color = "#000000"
   )  +
-  coord_fixed() +
   interior_annotation("a1") +
@@
   )  +
-  coord_fixed() +
   interior_annotation("b1") +

### 2024-09-22 paper.qmd
Updated Figure 14 a2, b2

@@
   proj = projection,
-  limits = 1.5,
-  axis_pos_x = -0.75,
-  axis_pos_y = -0.75,
+  limits = 1,
+  axis_pos_x = -0.4,
+  axis_pos_y = -0.4,
   axis_labels = names(scaled_pbmc_data),
-  threshold = 0.084)
+  threshold = 0.057)
 
@@
       yend = proj2_to), 
-    color = "#33a02c") +
+    color = "#33a02c",
+    linewidth = 1) +
   geom_point(
-    size = 0.5,
+    #size = 0.5,
     alpha = 0.2) +
@@
   coord_fixed() +
-  xlim(c(-1, 1)) +
-  ylim(c(-1, 1)) +
+  xlim(c(-0.6, 1)) +
+  ylim(c(-0.6, 1)) +
   interior_annotation("a2")
@@
   coord_fixed() +
-  xlim(c(-1, 1)) +
-  ylim(c(-1, 1)) +
+  xlim(c(-0.6, 1)) +
+  ylim(c(-0.6, 1)) +
   interior_annotation("b2")

### 2024-09-22 paper.qmd
Updated Figure 14

@@
 
-projection_scaled <- projection * 1.23
+projection_scaled <- projection * 1.25
 
@@
   proj = projection,
-  limits = 1.5,
-  axis_pos_x = -0.75,
-  axis_pos_y = -0.75,
+  limits = 1,
+  axis_pos_x = -0.6,
+  axis_pos_y = -0.7,
   axis_labels = names(scaled_pbmc_data),
-  threshold = 0.084)
+  threshold = 0.086)
 
@@
   coord_fixed() +
-  xlim(c(-1, 1)) +
-  ylim(c(-1, 1)) +
+  xlim(c(-0.9, 0.5)) +
+  ylim(c(-0.9, 0.5)) +
   interior_annotation("a3")
@@
   geom_point(
-    size = 0.5,
+    #size = 0.5,
     alpha = 0.2) +
@@
   coord_fixed() +
-  xlim(c(-1, 1)) +
-  ylim(c(-1, 1)) +
+  xlim(c(-0.9, 0.5)) +
+  ylim(c(-0.9, 0.5)) +
   interior_annotation("b3")

### 2024-09-22 paper.qmd
Updated Figure 14

@@
     ),
-    alpha = 0.1,
+    alpha = 0.2,
     color = "#000000"

### 2024-09-23 paper.qmd
Updated Figure 9 (d)

@@
 
-benchmark_label_df <- tibble(y = benchmark_scurve - 0.03,
+benchmark_label_df <- tibble(
+  y = benchmark_scurve - 0.03, 
   x = 1.3, 
@@
 ## To draw the distance distribution
-distance_scurve$group <- "1"
-distance_scurve_plot <- ggplot(distance_scurve, 
-                        aes(x = group, 
-                            y = distance)) + 
-  geom_quasirandom(
-    size = 2, 
-    alpha = 0.3
-  ) + 
-  ylim(0, max(unlist(distance_scurve$distance))+ 0.5) + 
-  coord_flip() + 
-  geom_hline(yintercept = benchmark_scurve,
-     linetype="solid", 
-     color = "black", linewidth=0.5, alpha = 0.5) +
-  ylab(expression(d^{(2)})) +
-  ggtitle("(d)") +
-  theme(aspect.ratio = 0.75,
-        axis.text.x = element_text(size = 12),
-        axis.title.x = element_text(size = 12),
-        plot.title = element_text(size = 12))
+# distance_scurve$group <- "1"
+# distance_scurve_plot <- ggplot(distance_scurve, 
+#                         aes(x = group, 
+#                             y = distance)) + 
+#   geom_quasirandom(
+#     size = 2, 
+#     alpha = 0.3
+#   ) + 
+#   ylim(0, max(unlist(distance_scurve$distance))+ 0.5) + 
+#   coord_flip() + 
+#   geom_hline(yintercept = benchmark_scurve,
+#      linetype="solid", 
+#      color = "black", linewidth=0.5, alpha = 0.5) +
+#   ylab(expression(d^{(2)})) +
+#   ggtitle("(d)") +
+#   theme(aspect.ratio = 0.75,
+#         axis.text.x = element_text(size = 12),
+#         axis.title.x = element_text(size = 12),
+#         plot.title = element_text(size = 12))
 
 # Define bin width and starting point
-bin_width <- hb_obj_scurve2$centroids$c_x[2]-hb_obj_scurve2$centroids$c_x[1]
+bin_width <- hb_obj_scurve2$a1
 start_point <- bin_width/2
@@
       binwidth = bin_width) +
-  scale_x_continuous(breaks = bin_labels, labels = round(bin_labels, 2)) +
-  ylab("standardized number of edges") +
+  scale_x_continuous(breaks = bin_labels, 
+                     labels = c(
+                       expression(a[1]),
+                       expression(2*a[1]),
+                       expression(3*a[1]),
+                       expression(4*a[1]),
+                       expression(5*a[1]),
+                       expression(6*a[1]),
+                       expression(7*a[1]),
+                       expression(8*a[1]),
+                       expression(9*a[1]),
+                       expression(10*a[1]),
+                       expression(11*a[1]),
+                       expression(12*a[1]),
+                       expression(13*a[1])
+                     )) +
+  # geom_text(
+  #   data = tibble(x = 12*bin_width, y = 0.7,
+  #                 text = expression(a[1])),
+  #   aes(
+  #     x=x,
+  #     y=y,
+  #     label = text
+  #   )
+  # 
+  # ) +
+  geom_vline(xintercept = start_point + 2*bin_width,
+             linetype="solid",
+             color = "#bdbdbd",
+             linewidth=1) +
+  ylab("Proportion of the number of edges") +
   xlab(expression(d^{(2)})) +
@@
 
-distance_scurve_summary <- distance_scurve |>
-  group_by(bin) |>
-  summarise(count = n()) |>
-  mutate(std_counts = count/sum(count))
+# distance_scurve_summary <- distance_scurve |>
+#   group_by(bin) |>
+#   summarise(count = n()) |>
+#   mutate(std_counts = count/sum(count))
 

### 2024-09-23 paper.qmd
Updated Figure 16 a)

@@
 
+tsne_minst_scaled_temp <- tsne_minst_scaled |>
+  mutate(cluster = if_else((tSNE1 <= 0.92) & (tSNE1 >= 0.75) & (tSNE2 <= 0.62) & (tSNE2 >= 0.44), "small_clust", "big_clust"))
+
 trimesh_removed_mnist <- ggplot() + 
@@
                linewidth = 1) +
-  geom_point(data = tsne_minst_scaled,
+  geom_point(data = tsne_minst_scaled_temp,
              aes(
                x = tSNE1,
-               y = tSNE2
+               y = tSNE2,
+               colour = cluster
              ),
-             alpha=0.1) +
+             alpha=0.1)+
+  scale_color_manual(values=c('#000000', '#e41a1c')) +
   interior_annotation("a", sc_ltr_pos_mnist) +
-  theme(aspect.ratio = 1)
+  theme(aspect.ratio = 1,
+        legend.position = "none")
 

### 2024-09-23 scripts/mnist/8_gen_linked_plots.R

@@ -178,8 +178,8 @@ nldr_scurve <- shared_df_mnist |>
 nldr_scurve_plt <- ggplotly(nldr_scurve, width = as.character(round(600/r2_mnist, 0)),
                             height = "600", tooltip = "none") |>
   style(unselected=list(marker=list(opacity=1))) |>
-  highlight(on="plotly_selected", off="plotly_deselect") |>
-  config(displayModeBar = FALSE)
+  highlight(on="plotly_selected", off="plotly_deselect") #|>
+  #config(displayModeBar = FALSE)
 
 
 # langevitour_output <- langevitour::langevitour(df_exe[1:7],

### 2024-09-23 paper.qmd
Map the small cluster color to projections

@@
   filter(type == "data") |>
+  # bind_cols(tsne_minst_scaled_temp) |>
+  # select(-type, -tSNE1, -tSNE2, -ID) |>
   select(-type)
+
+# scaled_mnist_data <- bind_cols(scaled_mnist_data,
+#                                tsne_minst_scaled |> select(-ID))
 
@@
 
+# model_2d <- df_bin_centroids_mnist |>
+#   select(c_x, c_y) |>
+#   rename(c("tSNE1" = "c_x",
+#            "tSNE2" = "c_y"))
+# 
+# scaled_mnist_data_model <- bind_cols(scaled_mnist_data_model, model_2d)
+
 # Combine with the true model for visualization
-df <- dplyr::bind_rows(scaled_mnist_data_model |> mutate(type = "model"),
-                       scaled_mnist_data |> mutate(type = "data"))
+# df <- dplyr::bind_rows(scaled_mnist_data_model |> mutate(type = "model"),
+#                        scaled_mnist_data)
+
+# scaled_mnist_data <- scaled_mnist_data |>
+#   select(-type)
 
@@
   #dplyr::mutate(type = df_exe$type) |>
-  dplyr::mutate(ID = dplyr::row_number())
+  dplyr::mutate(ID = dplyr::row_number()) |>
+  mutate(cluster = tsne_minst_scaled_temp$cluster)
 
@@
       x = proj1, 
-      y = proj2)) +
+      y = proj2,
+      colour = cluster)) +
   geom_segment(
@@
     alpha = 0.2) +
-  scale_color_manual(values = c("#000000", "#33a02c")) +
+  scale_color_manual(values = c("#000000","#e41a1c")) +
   geom_segment(
@@
   ylim(c(-1, 1)) +
-  interior_annotation("b")
+  interior_annotation("b") +
+  theme(
+    legend.position = "none"
+  )
 
@@
     alpha = 0.1) +
-  scale_color_manual(values = c("#000000", "#33a02c")) +
+  scale_color_manual(values = c("#000000", "#e41a1c")) +
   geom_segment(
@@
   ylim(c(-1, 1)) +
-  interior_annotation("c")
+  interior_annotation("c") +
+  theme(
+    legend.position = "none"
+  )
 
@@
     alpha = 0.2) +
-  scale_color_manual(values = c("#000000", "#33a02c")) +
+  scale_color_manual(values = c("#000000", "#e41a1c")) +
   geom_segment(
@@
   ylim(c(-1, 1)) +
-  interior_annotation("d")
+  interior_annotation("d") +
+  theme(
+    legend.position = "none"
+  )
 

### 2024-09-23 paper.qmd
Changed the color of small cluster

@@
              alpha=0.1)+
-  scale_color_manual(values=c('#000000', '#e41a1c')) +
+  scale_color_manual(values=c('#000000', '#984ea3')) +
   interior_annotation("a", sc_ltr_pos_mnist) +
@@
     alpha = 0.2) +
-  scale_color_manual(values = c("#000000","#e41a1c")) +
+  scale_color_manual(values = c("#000000","#984ea3")) +
   geom_segment(
@@
     alpha = 0.1) +
-  scale_color_manual(values = c("#000000", "#e41a1c")) +
+  scale_color_manual(values = c("#000000", "#984ea3")) +
   geom_segment(
@@
     alpha = 0.2) +
-  scale_color_manual(values = c("#000000", "#e41a1c")) +
+  scale_color_manual(values = c("#000000", "#984ea3")) +
   geom_segment(

### 2024-09-23 paper.qmd
Updated Figure 16, changed color of the small cluster

--- 2024-09-23 paper.qmd
@@
   geom_point(
-    size = 0.2,
+    #size = 0.2,
     alpha = 0.2) +
@@
   geom_point(
-    size = 0.2,
+    #size = 0.2,
     alpha = 0.1) +

### 2024-09-24 paper.qmd
Updated Figure 16 b) with more useful projection

@@
 ## First projection
+# projection <- cbind(
+#     c(0.71503,0.00502,-0.04370,-0.01811,-0.01103,0.09877,0.06199,0.05396,0.07771,-0.01713),
+#     c(0.00587,0.71879,0.03485,0.08284,-0.07534,-0.05420,0.04315,-0.00547,-0.02995,0.01810))
+
 projection <- cbind(
-    c(0.71503,0.00502,-0.04370,-0.01811,-0.01103,0.09877,0.06199,0.05396,0.07771,-0.01713),
-    c(0.00587,0.71879,0.03485,0.08284,-0.07534,-0.05420,0.04315,-0.00547,-0.02995,0.01810))
-
-projection_scaled <- projection * 1.3
+    c(-0.09936,0.36550,0.07658,0.29975,0.09085,-0.14009,-0.13417,0.48357,-0.00699,-0.13587),
+    c(0.23061,0.19332,-0.11264,0.14797,0.30351,0.14980,0.02785,-0.30209,-0.13540,-0.43284))
+
+projection_scaled <- projection * 2.5
 
@@
   proj = projection,
-  limits = 1.5,
-  axis_pos_x = -0.75,
-  axis_pos_y = -0.75,
+  limits = 1,
+  axis_pos_x = -0.83,
+  axis_pos_y = -0.83,
   axis_labels = names(scaled_mnist_data),
-  threshold = 0.084)
+  threshold = 0.06)
 

### 2024-09-24 paper.qmd
Figure 16 update the projections

@@
       yend = proj2_to), 
-    color = "#33a02c") +
+    color = "#33a02c",
+    linewidth = 1) +
   geom_point(
@@
   #coord_fixed() +
-  xlim(c(-1, 1)) +
-  ylim(c(-1, 1)) +
+  xlim(c(-1, 0.6)) +
+  ylim(c(-1, 0.6)) +
   interior_annotation("b") +
@@
       yend = proj2_to), 
-    color = "#33a02c") +
+    color = "#33a02c",
+    linewidth = 1) +
   geom_point(
     #size = 0.2,
-    alpha = 0.1) +
+    alpha = 0.2) +
   scale_color_manual(values = c("#000000", "#e41a1c")) +
@@
   coord_fixed() +
-  xlim(c(-1, 1)) +
-  ylim(c(-1, 1)) +
+  xlim(c(-1, 0.6)) +
+  ylim(c(-1, 0.6)) +
   interior_annotation("d") +

### 2024-09-24 paper.qmd
Added proj for twisted patterns

@@
 
+<!--proj for twisted pattern-->
+```{r}
+#| label: mnist-model-proj-twisted
+## fourth projection
+projection <- cbind(
+  c(-0.03265,-0.43663,0.02143,0.07384,0.26237,0.20909,0.32398,-0.34487,0.03723,-0.04418),
+  c(0.52862,-0.14177,0.20045,-0.37753,-0.09552,0.08851,0.07168,0.09255,-0.15653,-0.00030))
+
+projection_scaled <- projection * 2
+
+projected <- as.matrix(scaled_mnist_data) %*% projection_scaled
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  #dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number()) |>
+  mutate(cluster = tsne_minst_scaled_temp$cluster)
+
+projected_model <- as.matrix(scaled_mnist_data_model) %*% projection_scaled
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges_mnist |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+axes_obj <- gen_axes(  
+  proj = projection,
+  limits = 1,
+  axis_pos_x = -0.83,
+  axis_pos_y = -0.83,
+  axis_labels = names(scaled_mnist_data),
+  threshold = 0.06)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
+
+mnist_proj_tsne_model4 <- ggplot() +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#33a02c",
+    linewidth = 0.5) +
+  # geom_point(
+  #   #size = 0.2,
+  #   alpha = 0.2) +
+  # scale_color_manual(values = c("#000000", "#e41a1c")) +
+  geom_segment(
+    data=axes, 
+    aes(x=x1, y=y1, xend=x2, yend=y2), 
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
+    colour="grey50",
+    size = 2) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-1, 1)) +
+  ylim(c(-1, 1)) +
+  interior_annotation("d") +
+  theme(
+    legend.position = "none"
+  )
+
+## fifth projection
+projection <- cbind(
+  c(-0.36688,-0.10781,-0.15474,-0.23005,-0.20194,-0.10841,-0.32443,0.04531,-0.36514,-0.14546),
+  c(0.34256,-0.16215,-0.18747,-0.08515,-0.28689,-0.32807,0.28372,0.20894,-0.16990,0.09181))
+
+projection_scaled <- projection * 2.3
+
+projected <- as.matrix(scaled_mnist_data) %*% projection_scaled
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  #dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number()) |>
+  mutate(cluster = tsne_minst_scaled_temp$cluster)
+
+projected_model <- as.matrix(scaled_mnist_data_model) %*% projection_scaled
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges_mnist |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+axes_obj <- gen_axes(  
+  proj = projection,
+  limits = 1,
+  axis_pos_x = -0.83,
+  axis_pos_y = -0.83,
+  axis_labels = names(scaled_mnist_data),
+  threshold = 0.06)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
+
+mnist_proj_tsne_model5 <- ggplot() +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#33a02c",
+    linewidth = 0.5) +
+  # geom_point(
+  #   #size = 0.2,
+  #   alpha = 0.2) +
+  # scale_color_manual(values = c("#000000", "#e41a1c")) +
+  geom_segment(
+    data=axes, 
+    aes(x=x1, y=y1, xend=x2, yend=y2), 
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
+    colour="grey50",
+    size = 2) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-1, 1)) +
+  ylim(c(-1, 1)) +
+  interior_annotation("d") +
+  theme(
+    legend.position = "none"
+  )
+
+```
+
+```{r}
+#| label: fig-mnist-tri-proj-twisted
+#| fig-pos: H
+#| fig-height: 10
+#| fig-width: 10
+#| fig-cap: "twisted pattern"
+
+free(trimesh_removed_mnist) + free(trimesh_removed_mnist) + mnist_proj_tsne_model4 + mnist_proj_tsne_model5 +
+  plot_layout(ncol = 2)
+```
+
 <!-- ::: {#fig-mnist1-sc layout-ncol="4" fig-pos="H"} -->

### 2024-09-24 paper.qmd
Updated Figure 8

@@
   hex_grid_coloured_scurve3 +
-hex_grid_coloured_scurve1_dens +
-  hex_grid_coloured_scurve2_dens +
-  hex_grid_coloured_scurve3_dens +
   plot_layout(guides='collect', ncol = 3,

### 2024-09-24 paper.qmd
Updated Figure 6 (still need to work on that)

@@
 
+# error_df_scurve <- error_df_scurve |>
+#   mutate(type = case_when(
+#     row_wise_abs_error == 0 ~ "no error",
+#     row_wise_abs_error <= 0.2 ~ "error 0-0.2",
+#     row_wise_abs_error <= 0.4 ~ "error 0.2-0.4",
+#     row_wise_abs_error <= 0.6 ~ "error 0.4-0.6",
+#     .default = "error greter than 0.6"
+#   )) |>
+#   mutate(type = factor(type, levels = c(
+#     "no error", "error 0-0.2", "error 0.2-0.4", "error 0.4-0.6", "error greter than 0.6")))
+
 error_df_scurve <- error_df_scurve |>
-  mutate(type = case_when(
-    row_wise_abs_error == 0 ~ "no error",
-    row_wise_abs_error <= 0.2 ~ "error 0-0.2",
-    row_wise_abs_error <= 0.4 ~ "error 0.2-0.4",
-    row_wise_abs_error <= 0.6 ~ "error 0.4-0.6",
-    .default = "error greter than 0.6"
-  )) |>
-  mutate(type = factor(type, levels = c(
-    "no error", "error 0-0.2", "error 0.2-0.4", "error 0.4-0.6", "error greter than 0.6")))
+  mutate(sqrt_row_wise_abs_error = sqrt(row_wise_abs_error))
 
@@
              y = UMAP2,
-             color = type)) +
+             colour = sqrt_row_wise_abs_error)) +
   geom_point(alpha=0.5) +
-  scale_colour_brewer(
-    type = "qual",
-    palette = 2
-  ) +
-  theme(
-    legend.position = "right",
-    legend.text = element_text(size = 10)
-  ) +
+  scale_colour_continuous_sequential(palette = "Light Grays") +
   coord_equal()

### 2024-09-24 paper.qmd
Removed distance diag from Figure 9
Figure 10 initial work

@@
 The value of $b_2$ is determined analytically by $b_1, q, r_2$.
-Values of $b_1$ between $2$ and $b_1 = \sqrt{\frac{n}{2}}$ are allowed.
+Values of $b_1$ between $2$ and $b_1 = \sqrt{\frac{n}{r_2}}$ are allowed.
 @fig-param-scurve (b) shows the effect of different choices of $b_1$ on the MSE of the fitted model.
@@
 
+```{r}
+#| echo: false
+#| fig-cap: "Various plots to help assess best hexagon bin parameters, thresholds to remove low density bins and large edges.
+Both (b) and (c) show MSE, against number of bins along the x-axis and standardised count.
+A good benchmark value for these parameters is when the MSE drops and then flattens out.
+Plot (a) shows the distribution of stadardised counts of hexagons.
+Plot (d) shows the distribution of $2\\text{-}D$ Euclidean distances between bin centroids, with a good benchmark value for removing large edges would being the distance that shows the first large decrease."
+#| label: fig-param-scurve
+#| out-width: 100%
+#| fig-width: 12
+#| fig-height: 10
+#| fig-pos: H
+
+mse_scurve_b + a1_m_scurve +
+  mse_scurve_lwd + 
+  plot_layout(ncol=3)
+```
+
 #### Removing long edges
@@
   tri_object = tr1_object_scurve)
+
+trimesh_scurve <- ggplot() +
+  geom_segment(data = tr_from_to_df_scurve,
+               aes(
+                 x = x_from,
+                 y = y_from,
+                 xend = x_to,
+                 yend = y_to),
+               colour = "#33a02c") +
+  interior_annotation("a", position = c(0.92, 0.95))
 
@@
 
-benchmark_label_df <- tibble(
-  y = benchmark_scurve - 0.03, 
-  x = 1.3, 
-  label = paste0("benchmark is ", benchmark_scurve))
+## Benchmark 1
+
+distance_df_small_edges_scurve1 <- distance_scurve |>
+  filter(distance < benchmark_scurve)
+
+tr_from_to_df_scurve1 <- inner_join(
+  tr_from_to_df_scurve, distance_df_small_edges_scurve1,
+  by = c("from", "to"))
+
+trimesh_scurve_removed1 <- ggplot() +
+  geom_segment(data = tr_from_to_df_scurve1,
+               aes(
+                 x = x_from,
+                 y = y_from,
+                 xend = x_to,
+                 yend = y_to),
+               colour = "#33a02c") +
+  interior_annotation("c", position = c(0.92, 0.95))
+
+## Benchmark 2
+
+benchmark_scurve2 <- 0.6
+
+distance_df_small_edges_scurve2 <- distance_scurve |>
+  filter(distance < benchmark_scurve2)
+
+tr_from_to_df_scurve2 <- inner_join(
+  tr_from_to_df_scurve, distance_df_small_edges_scurve2,
+  by = c("from", "to"))
+
+trimesh_scurve_removed2 <- ggplot() +
+  geom_segment(data = tr_from_to_df_scurve2,
+               aes(
+                 x = x_from,
+                 y = y_from,
+                 xend = x_to,
+                 yend = y_to),
+               colour = "#33a02c") +
+  interior_annotation("b", position = c(0.92, 0.95))
 
@@
 ```{r}
-#| echo: false
-#| fig-cap: "Various plots to help assess best hexagon bin parameters, thresholds to remove low density bins and large edges.
-Both (b) and (c) show MSE, against number of bins along the x-axis and standardised count.
-A good benchmark value for these parameters is when the MSE drops and then flattens out.
-Plot (a) shows the distribution of stadardised counts of hexagons.
-Plot (d) shows the distribution of $2\\text{-}D$ Euclidean distances between bin centroids, with a good benchmark value for removing large edges would being the distance that shows the first large decrease."
-#| label: fig-param-scurve
-#| out-width: 100%
-#| fig-width: 12
-#| fig-height: 10
-#| fig-pos: H
-
-mse_scurve_b + a1_m_scurve +
-  mse_scurve_lwd + distance_hist + 
-  plot_layout(ncol=2)
+#| fig-width: 6
+#| fig-height: 4
+#| fig-cap: "long edge removal"
+
+trimesh_scurve + trimesh_scurve_removed2 + trimesh_scurve_removed1 +
+  plot_layout(ncol = 3)
 ```

### 2024-09-24 paper.tex
Updated the Figure 15 title
Added proj for twisted patterns
Removed distance diag from Figure 9
Figure 10 initial work

@@
 determined analytically by \(b_1, q, r_2\). Values of \(b_1\) between
-\(2\) and \(b_1 = \sqrt{\frac{n}{2}}\) are allowed.
+\(2\) and \(b_1 = \sqrt{\frac{n}{r_2}}\) are allowed.
 Figure~\ref{fig-param-scurve} (b) shows the effect of different choices
@@
 
-\subsubsection{Removing long edges}\label{removing-long-edges}
-
-Edges define the neighbourhood structure, in order to provide a smooth
-\gD{} representation of the fitted model.
-Figure~\ref{fig-scurve-true-sc} shows a wire frame of the true model
-that was used to generate the S-curve example data. The ideal is that
-the representation of the fitted model, at least for this example where
-we know the true model, should look similar to this.
-
-The Delaunay triangulation will ensure that all centroids are connected
-into a triangular mesh. For some structures, like clustered data, or
-highly non-linear shapes, breaks in the mesh are meaningful. When
-separated clusters are present the mesh should be broken across the
-gaps. For non-linear structures like the S-curve, the mesh should run
-unbroken along the S, but there should be no edges connecting the top of
-the S directly to the bottom of the S. For these reasons it is necessary
-to remove edges from the mesh in some applications.
-
-The decision on edge length removal is made based on the distribution of
-edge lengths. In particular, a gap between values, where there a
-concentration of small values and then a few larger values, likely
-suggests a cutoff for edge removal. Because the triangulation is
-typically done on the hexagon centroids, there are particular discrete
-edge lengths, based on bin widths. Figure~\ref{fig-NLDR-scurve} (d)
-illustrates edge length distributions.
-
-There is an additional step that is needed. When the model is lifted
-into \pD{}, if the fit is good all the edges should be relatively small
-in this space, too. If this is not the case, then there are several
-possible actions: (1) re-do the NLDR to get a more representative
-layout; (2) identify the edge and remove it from the model, in \gD{} and
-\pD{}; (3) consider different values for the model fit, number of bins,
-initial bin position or removing low density bins.
-
 \begin{figure}[H]
@@
 being the distance that shows the first large decrease.}
+
+\end{figure}%
+
+\subsubsection{Removing long edges}\label{removing-long-edges}
+
+Edges define the neighbourhood structure, in order to provide a smooth
+\gD{} representation of the fitted model.
+Figure~\ref{fig-scurve-true-sc} shows a wire frame of the true model
+that was used to generate the S-curve example data. The ideal is that
+the representation of the fitted model, at least for this example where
+we know the true model, should look similar to this.
+
+The Delaunay triangulation will ensure that all centroids are connected
+into a triangular mesh. For some structures, like clustered data, or
+highly non-linear shapes, breaks in the mesh are meaningful. When
+separated clusters are present the mesh should be broken across the
+gaps. For non-linear structures like the S-curve, the mesh should run
+unbroken along the S, but there should be no edges connecting the top of
+the S directly to the bottom of the S. For these reasons it is necessary
+to remove edges from the mesh in some applications.
+
+The decision on edge length removal is made based on the distribution of
+edge lengths. In particular, a gap between values, where there a
+concentration of small values and then a few larger values, likely
+suggests a cutoff for edge removal. Because the triangulation is
+typically done on the hexagon centroids, there are particular discrete
+edge lengths, based on bin widths. Figure~\ref{fig-NLDR-scurve} (d)
+illustrates edge length distributions.
+
+There is an additional step that is needed. When the model is lifted
+into \pD{}, if the fit is good all the edges should be relatively small
+in this space, too. If this is not the case, then there are several
+possible actions: (1) re-do the NLDR to get a more representative
+layout; (2) identify the edge and remove it from the model, in \gD{} and
+\pD{}; (3) consider different values for the model fit, number of bins,
+initial bin position or removing low density bins.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-40-1.pdf}
+
+}
+
+\caption{long edge removal}
 
@@
 varying binwidth (\(a_1\)). Colour used for the lines and points in the
-left plot and in the scatterplots represents NLDR layout (a-e). Layout c
-is universally poor. Layouts a that show two close clusters are
-universally suboptimal. Layout a is the best choice.}
+left plot and in the scatterplots represents NLDR layout (a-e). All the
+layouts appear to be very similar. Layout c is universally poor. Layouts
+a that show two close clusters are universally suboptimal. Layout a is
+the best choice.}
 
@@
 
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-mnist-tri-proj-twisted-1.pdf}
+
+}
+
+\caption{\label{fig-mnist-tri-proj-twisted}twisted pattern}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+\centering{
+
 \includegraphics[width=1\textwidth,height=0.8\textheight]{paper_files/figure-pdf/fig-model-error-mnist-1.pdf}

### 2024-09-24 paper.qmd
Removed twisted pattern in MNIST

@@
 
-<!--proj for twisted pattern-->
-```{r}
-#| label: mnist-model-proj-twisted
-## fourth projection
-projection <- cbind(
-  c(-0.03265,-0.43663,0.02143,0.07384,0.26237,0.20909,0.32398,-0.34487,0.03723,-0.04418),
-  c(0.52862,-0.14177,0.20045,-0.37753,-0.09552,0.08851,0.07168,0.09255,-0.15653,-0.00030))
-
-projection_scaled <- projection * 2
-
-projected <- as.matrix(scaled_mnist_data) %*% projection_scaled
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  #dplyr::mutate(type = df_exe$type) |>
-  dplyr::mutate(ID = dplyr::row_number()) |>
-  mutate(cluster = tsne_minst_scaled_temp$cluster)
-
-projected_model <- as.matrix(scaled_mnist_data_model) %*% projection_scaled
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  distance_df_small_edges_mnist |> select(-distance), 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-axes_obj <- gen_axes(  
-  proj = projection,
-  limits = 1,
-  axis_pos_x = -0.83,
-  axis_pos_y = -0.83,
-  axis_labels = names(scaled_mnist_data),
-  threshold = 0.06)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
-
-mnist_proj_tsne_model4 <- ggplot() +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#33a02c",
-    linewidth = 0.5) +
-  # geom_point(
-  #   #size = 0.2,
-  #   alpha = 0.2) +
-  # scale_color_manual(values = c("#000000", "#e41a1c")) +
-  geom_segment(
-    data=axes, 
-    aes(x=x1, y=y1, xend=x2, yend=y2), 
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 2) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-1, 1)) +
-  ylim(c(-1, 1)) +
-  interior_annotation("d") +
-  theme(
-    legend.position = "none"
-  )
-
-## fifth projection
-projection <- cbind(
-  c(-0.36688,-0.10781,-0.15474,-0.23005,-0.20194,-0.10841,-0.32443,0.04531,-0.36514,-0.14546),
-  c(0.34256,-0.16215,-0.18747,-0.08515,-0.28689,-0.32807,0.28372,0.20894,-0.16990,0.09181))
-
-projection_scaled <- projection * 2.3
-
-projected <- as.matrix(scaled_mnist_data) %*% projection_scaled
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  #dplyr::mutate(type = df_exe$type) |>
-  dplyr::mutate(ID = dplyr::row_number()) |>
-  mutate(cluster = tsne_minst_scaled_temp$cluster)
-
-projected_model <- as.matrix(scaled_mnist_data_model) %*% projection_scaled
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  distance_df_small_edges_mnist |> select(-distance), 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-axes_obj <- gen_axes(  
-  proj = projection,
-  limits = 1,
-  axis_pos_x = -0.83,
-  axis_pos_y = -0.83,
-  axis_labels = names(scaled_mnist_data),
-  threshold = 0.06)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
-
-mnist_proj_tsne_model5 <- ggplot() +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#33a02c",
-    linewidth = 0.5) +
-  # geom_point(
-  #   #size = 0.2,
-  #   alpha = 0.2) +
-  # scale_color_manual(values = c("#000000", "#e41a1c")) +
-  geom_segment(
-    data=axes, 
-    aes(x=x1, y=y1, xend=x2, yend=y2), 
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 2) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-1, 1)) +
-  ylim(c(-1, 1)) +
-  interior_annotation("d") +
-  theme(
-    legend.position = "none"
-  )
-
-```
-
-```{r}
-#| label: fig-mnist-tri-proj-twisted
-#| fig-pos: H
-#| fig-height: 10
-#| fig-width: 10
-#| fig-cap: "twisted pattern"
-
-free(trimesh_removed_mnist) + free(trimesh_removed_mnist) + mnist_proj_tsne_model4 + mnist_proj_tsne_model5 +
-  plot_layout(ncol = 2)
-```
-
 <!-- ::: {#fig-mnist1-sc layout-ncol="4" fig-pos="H"} -->

### 2024-09-24 paper.qmd
Updated Figure 9 (b)

@@
   r2 = r2,
-  q = 0.16,
+  q = 0.1,
   is_bin_centroid = TRUE,
@@
 ## To initialize number of bins along the x-axis
-bin1_vec_scurve <- 5:19 #sqrt(NROW(training_data_scurve)/2)
+bin1_vec_scurve <- 5:22 #sqrt(NROW(umap_scurve_scaled)/r2)
 #buffer_vec <- seq(0.05, 0.1, by = 0.01)
@@
 
+error_scurve <- error_scurve |>
+  mutate(prop_bins = b_non_empty/b)
+
 a1_m_scurve <- ggplot(error_scurve,
                      aes(x = a1,
-                         y = b_non_empty)) +
-  geom_vline(xintercept = 0.12,
-         linetype=2,
-         color = "#bdbdbd",
-         linewidth=1) +
-  geom_hline(yintercept = 52,
-           linetype=2,
-           color = "#bdbdbd",
-           linewidth=1) +
+                         y = prop_bins)) +
+  geom_vline(xintercept = 0.12, linetype="solid",
+             color = "#bdbdbd", linewidth=1) +
+  geom_vline(xintercept = 0.10, linetype=2,
+             color = "#bdbdbd", linewidth=1) +
+  geom_vline(xintercept = 0.18, linetype=2,
+             color = "#bdbdbd", linewidth=1) +
   geom_point(size = 1) +
   geom_line(linewidth = 0.05) +
-  # geom_vline(xintercept = 11, linetype="solid",
-  #            color = "#bdbdbd", linewidth=1) +
-  # geom_vline(xintercept = 9, linetype=2,
-  #            color = "#bdbdbd", linewidth=1) +
-  # geom_vline(xintercept = 13, linetype=2,
-  #            color = "#bdbdbd", linewidth=1) +
-  scale_y_continuous(breaks = sort(unique(error_scurve$b_non_empty))[-3]) +
+  # scale_y_continuous(breaks = sort(unique(error_scurve$b_non_empty))[-3]) +
   scale_x_continuous(breaks = sort(unique(round(error_scurve$a1, 2)))[c(1, 3, 5, 6, 8, 10:12)]) +
-  labs(x = expression(paste("binwidth (", a[1], ")")), y = expression(paste("non-empty bins (", m, ")"))) +
+  labs(x = expression(paste("binwidth (", a[1], ")")), y = expression(paste("proportion of non-empty bins ", bgroup("(", frac(m, b), ")")))) +
   ggtitle("(b)") +
@@
   mse_scurve_lwd + 
-  plot_layout(ncol=3)
+  plot_layout(ncol=2)
 ```

### 2024-09-24 paper.qmd
Choose the color palattes for Figure 8 and color the lines accordingly

@@
   scale_fill_viridis_c(direction = -1, 
-    na.value = "#ffffff", option = "C") +
+    na.value = "#ffffff", option = "E") +
   interior_annotation("a", position = c(0.92, 0.93)) +
@@
            size = 0.5) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "D") +
   xlim(c(-0.25, 1.35)) + ylim(c(-0.35, 2)) +
@@
   geom_vline(xintercept = 0.10, linetype=2,
-           color = "#bdbdbd", linewidth=1) +
+           color = "#54278f", linewidth=1) +
   geom_vline(xintercept = 0.12,
              linetype="solid",
-             color = "#bdbdbd",
+             color = "#006d2c",
              linewidth=1) +
   geom_vline(xintercept = 0.18, linetype=2,
-             color = "#bdbdbd", linewidth=1) +
+             color = "#08519c", linewidth=1) +
   geom_line(linewidth = 0.5) + 
@@
   geom_vline(xintercept = 0.12, linetype="solid",
-             color = "#bdbdbd", linewidth=1) +
+             color = "#006d2c", linewidth=1) +
   geom_vline(xintercept = 0.10, linetype=2,
-             color = "#bdbdbd", linewidth=1) +
+             color = "#54278f", linewidth=1) +
   geom_vline(xintercept = 0.18, linetype=2,
-             color = "#bdbdbd", linewidth=1) +
+             color = "#08519c", linewidth=1) +
   geom_point(size = 1) +

### 2024-09-24 paper.qmd
Updated Figure 9 (c)

@@
 ## To initialize benchmark values to remove low density hexagons
-benchmark_rm_hex_vec <- append(seq(0, 0.99, by=0.1), c(benchmark1, 0.99))
-
+benchmark_rm_hex_vec <- sort(append(seq(0, 0.6, by=0.15), benchmark1))
+
+## option 1
 error_rm_scurve <- data.frame(matrix(nrow = 0, ncol = 0))
@@
 
-benchmark_label_df <- tibble(x = 0.2, y = 0.1, 
-                          label = paste0("benchmark is ", benchmark1))
+# benchmark_label_df <- tibble(x = 0.2, y = 0.1, 
+#                           label = paste0("benchmark is ", benchmark1))
 
@@
 
+### Option 2
+
+scurve_model2 <- fit_highd_model(
+  training_data = training_data_scurve,
+  emb_df = umap_scurve_scaled,
+  bin1 = 7,
+  r2 = r2,
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = FALSE,
+  col_start_highd = "x"
+)
+
+df_bin_centroids_scurve2 <- scurve_model2$df_bin_centroids
+df_bin_scurve2 <- scurve_model2$df_bin
+
+error_rm_scurve2 <- data.frame(matrix(nrow = 0, ncol = 0))
+
+for (benchmark_rm_lwd in benchmark_rm_hex_vec) {
+
+  df_bin_centroids_scurve_high_dens <- df_bin_centroids_scurve2 |>
+    filter(std_counts > benchmark_rm_lwd)
+
+  df_bin_scurve_high_dens <- df_bin_scurve2 |>
+    filter(hb_id %in% df_bin_centroids_scurve_high_dens$hexID)
+
+  ## Compute error
+  error_df <- glance(
+    df_bin_centroids = df_bin_centroids_scurve_high_dens,
+    df_bin = df_bin_scurve_high_dens,
+    training_data = training_data_scurve,
+    newdata = NULL,
+    type_NLDR = "UMAP",
+    col_start = "x") |>
+    mutate(benchmark_rm_lwd = round(benchmark_rm_lwd, 2),
+           bin1 = 7,
+           bin2 = 13,
+           b = bin1 * bin2,
+           b_non_empty = NROW(df_bin_centroids_scurve_high_dens),
+           mean_counts = sum(df_bin_centroids_scurve_high_dens$bin_counts)/NROW(df_bin_centroids_scurve_high_dens))
+
+  error_rm_scurve2 <- bind_rows(error_rm_scurve2, error_df)
+
+}
+
+### Option 3
+
+scurve_model3 <- fit_highd_model(
+  training_data = training_data_scurve,
+  emb_df = umap_scurve_scaled,
+  bin1 = 13,
+  r2 = r2,
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = FALSE,
+  col_start_highd = "x"
+)
+
+df_bin_centroids_scurve3 <- scurve_model3$df_bin_centroids
+df_bin_scurve3 <- scurve_model3$df_bin
+
+error_rm_scurve3 <- data.frame(matrix(nrow = 0, ncol = 0))
+
+for (benchmark_rm_lwd in benchmark_rm_hex_vec) {
+
+  df_bin_centroids_scurve_high_dens <- df_bin_centroids_scurve3 |>
+    filter(std_counts > benchmark_rm_lwd)
+
+  df_bin_scurve_high_dens <- df_bin_scurve3 |>
+    filter(hb_id %in% df_bin_centroids_scurve_high_dens$hexID)
+
+  ## Compute error
+  error_df <- glance(
+    df_bin_centroids = df_bin_centroids_scurve_high_dens,
+    df_bin = df_bin_scurve_high_dens,
+    training_data = training_data_scurve,
+    newdata = NULL,
+    type_NLDR = "UMAP",
+    col_start = "x") |>
+    mutate(benchmark_rm_lwd = round(benchmark_rm_lwd, 2),
+           bin1 = 13,
+           bin2 = 24,
+           b = bin1 * bin2,
+           b_non_empty = NROW(df_bin_centroids_scurve_high_dens),
+           mean_counts = sum(df_bin_centroids_scurve_high_dens$bin_counts)/NROW(df_bin_centroids_scurve_high_dens))
+
+  error_rm_scurve3 <- bind_rows(error_rm_scurve3, error_df)
+
+}
+
+error_rm_scurve <- bind_rows(
+  error_rm_scurve,
+  error_rm_scurve2,
+  error_rm_scurve3
+) |>
+  mutate(bin1  = as.factor(bin1))
+
 mse_scurve_lwd <- ggplot(error_rm_scurve, 
                      aes(x = benchmark_rm_lwd, 
-                         y = MSE)) + 
+                         y = MSE,
+                         color = bin1)) + 
   geom_point(
@@
   scale_x_continuous(breaks = unique(error_rm_scurve$benchmark_rm_lwd)) +
+  scale_color_manual(values=c("#08519c", "#006d2c", "#54278f")) +
   xlab("threshold to remove low-density hexagons") +
@@
         axis.ticks.x = element_line(),
-        axis.ticks.y = element_line())
-
-# mse_scurve_lwd2 <- error_rm_scurve |>
-#   filter(b_non_empty >= 5) |>
-#   ggplot(aes(x = b_non_empty,y = MSE)) + 
-#   geom_point(
-#     size = 1
-#     ) +
-#   geom_line(
-#     linewidth = 0.3
-#     ) + 
-#   # geom_vline(xintercept = benchmark1, linetype="solid", 
-#   #            color = "#bdbdbd", linewidth=1, alpha = 0.5) +
-#   scale_x_continuous("m (b1 = 14)") + 
-#   ylab("MSE") +
-#   ggtitle("(e)") +
-#   theme(aspect.ratio = 0.75,
-#         axis.text.x = element_text(size = 12),
-#         axis.text.y = element_text(size = 12),
-#         axis.title.x = element_text(size = 12),
-#         axis.title.y = element_text(size = 12, 
-#             angle=90),
-#         plot.title = element_text(size = 12))
+        axis.ticks.y = element_line(),
+        legend.position = "none")
 ```

### 2024-09-24 paper.qmd
Adde prop density vs a1 plot to Figure 9 (d)

@@
 }
+```
+
+```{r}
+#| label: solve-quad-fun
+
+quad <- function(a = 3, b = 2 * a2, c = -(a2^2 + a1^2))
+{
+  a <- as.complex(a)
+  answer <- c((-b + sqrt(b^2 - 4 * a * c)) / (2 * a),
+              (-b - sqrt(b^2 - 4 * a * c)) / (2 * a))
+  if(all(Im(answer) == 0)) answer <- Re(answer)
+  if(answer[1] == answer[2]) return(answer[1])
+  answer[answer>0] ## only positive 
+}
+
 ```
@@
 
-  bin_obj <- calc_bins_y(bin1 = xbins, r2 = r2, q = 0.1)
+  bin_obj <- calc_bins_y(
+    bin1 = xbins, 
+    r2 = r2, 
+    q = 0.1)
 
@@
   a1 <- bin_obj$a1
+  a2 <- bin_obj$a2
 
@@
            b_non_empty = NROW(df_bin_centroids_scurve),
-           a1 = round(a1, 2))
+           a1 = round(a1, 2),
+           a2 = round(a2, 2),
+           side_length = quad(a=3, b = 2 * a2, c = -(a2^2 + a1^2)))
 
@@
   filter(MSE == min(MSE)) |>
-  ungroup()
+  ungroup() |>
+  mutate(prop_dens = b_non_empty/side_length)
+
+base_line_prop_dens <- error_scurve |>
+  filter(a1 == min(a1)) |>
+  pull(prop_dens)
+
+error_scurve <- error_scurve |>
+  mutate(prop_comp = prop_dens/base_line_prop_dens)
 
@@
   ggtitle("(a)") +
+  theme_minimal() +
+  theme(aspect.ratio = 0.75,
+        panel.border = element_rect(fill = 'transparent'),
+        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+        axis.ticks.x = element_line(),
+        axis.ticks.y = element_line())
+
+prop_dens_a1 <- ggplot(error_scurve,
+                     aes(x = a1,
+                         y = prop_comp)) +
+  geom_vline(xintercept = 0.12, linetype="solid",
+             color = "#006d2c", linewidth=1) +
+  geom_vline(xintercept = 0.10, linetype=2,
+             color = "#54278f", linewidth=1) +
+  geom_vline(xintercept = 0.18, linetype=2,
+             color = "#08519c", linewidth=1) +
+  geom_point(size = 1) +
+  geom_line(linewidth = 0.05) +
+  # scale_y_continuous(breaks = sort(unique(error_scurve$b_non_empty))[-3]) +
+  scale_x_continuous(breaks = sort(unique(round(error_scurve$a1, 2)))[c(1, 3, 5, 6, 8, 10:12)]) +
+  labs(x = expression(paste("binwidth (", a[1], ")")), y = paste("proportion of densities")) +
+  ggtitle("(d)") +
   theme_minimal() +
@@
 
+<!--base line -->
+
 <!-- Furthermore, selecting the benchmark value for removing low-density hexagons is important.
@@
 mse_scurve_b + a1_m_scurve +
-  mse_scurve_lwd + 
+  mse_scurve_lwd + prop_dens_a1 +
   plot_layout(ncol=2)

### 2024-09-24 paper.qmd
Updated Figure 11 (b)

@@
   geom_point(alpha=0.1, size=1, colour='#a65628') +
-  interior_annotation("b") +
+  interior_annotation("b", c(0.08, 0.93)) +
   theme(aspect.ratio = 1)

### 2024-09-25 paper.qmd
Figure 15 resize labels

@@
   )  +
-  interior_annotation("a1") +
+  interior_annotation("a1", cex = 2) +
   theme(
@@
   ylim(c(-0.6, 1)) +
-  interior_annotation("a2")
+  interior_annotation("a2", cex = 2)
 
@@
   ylim(c(-0.9, 0.5)) +
-  interior_annotation("a3")
+  interior_annotation("a3", cex = 2)
 
@@
   )  +
-  interior_annotation("b1") +
+  interior_annotation("b1", cex = 2) +
   theme(
@@
   ylim(c(-0.6, 1)) +
-  interior_annotation("b2")
+  interior_annotation("b2", cex = 2)
 
@@
   ylim(c(-0.9, 0.5)) +
-  interior_annotation("b3")
+  interior_annotation("b3", cex = 2)
 
@@
 #| echo: false
-#| fig-cap: "Model in \\gD{}, on the layouts, and two views of the fit in projections from $9\\text{-}D$, for the PBMC3k data.
-The \\gD{} model shows three well-separated clusters with large gaps in a1 ($a_1 = 0.05$, $b = 676 \\  (26, \ 26)$, $m = 92$, benchmark value to remove large edges is $0.08$), while a2 ($a_1 = 0.10$, $b = 195 \\  (13, \ 15)$, $m = 90$, benchmark value to remove large edges is $0.166$) shows small separation between the clusters.
-Because of the small separation, the three clusters are connected in \\gD{} and appear as long edges in $9\\text{-}D$.
-The full video for the model for layout a is available at <https://youtu.be/5Y1hE4i7N2k>.
-The full video for the model for layout c is available at <https://youtu.be/5Y1hE4i7N2k>."
+#| fig-cap: ""
 #| label: fig-model-pbmc-author-proj

### 2024-09-25 paper.tex
Removed twisted pattern in MNIST
Adde prop density vs a1 plot to Figure 9 (d)
Figure 15 resize labels

--- 2024-09-19 paper.tex
@@
 being the distance that shows the first large decrease.}
+
+\end{figure}%
+
+\subsubsection{Removing long edges}\label{removing-long-edges}
+
+Edges define the neighbourhood structure, in order to provide a smooth
+\gD{} representation of the fitted model.
+Figure~\ref{fig-scurve-true-sc} shows a wire frame of the true model
+that was used to generate the S-curve example data. The ideal is that
+the representation of the fitted model, at least for this example where
+we know the true model, should look similar to this.
+
+The Delaunay triangulation will ensure that all centroids are connected
+into a triangular mesh. For some structures, like clustered data, or
+highly non-linear shapes, breaks in the mesh are meaningful. When
+separated clusters are present the mesh should be broken across the
+gaps. For non-linear structures like the S-curve, the mesh should run
+unbroken along the S, but there should be no edges connecting the top of
+the S directly to the bottom of the S. For these reasons it is necessary
+to remove edges from the mesh in some applications.
+
+The decision on edge length removal is made based on the distribution of
+edge lengths. In particular, a gap between values, where there a
+concentration of small values and then a few larger values, likely
+suggests a cutoff for edge removal. Because the triangulation is
+typically done on the hexagon centroids, there are particular discrete
+edge lengths, based on bin widths. Figure~\ref{fig-NLDR-scurve} (d)
+illustrates edge length distributions.
+
+There is an additional step that is needed. When the model is lifted
+into \pD{}, if the fit is good all the edges should be relatively small
+in this space, too. If this is not the case, then there are several
+possible actions: (1) re-do the NLDR to get a more representative
+layout; (2) identify the edge and remove it from the model, in \gD{} and
+\pD{}; (3) consider different values for the model fit, number of bins,
+initial bin position or removing low density bins.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-41-1.pdf}
+
+}
+
+\caption{long edge removal}
 
@@
 
-\caption{\label{fig-model-pbmc-author-proj}Model in \gD{}, on the
-layouts, and two views of the fit in projections from \(9\text{-}D\),
-for the PBMC3k data. The \gD{} model shows three well-separated clusters
-with large gaps in a1 (\(a_1 = 0.05\), \(b = 676 \  (26,  26)\),
-\(m = 92\), benchmark value to remove large edges is \(0.08\)), while a2
-(\(a_1 = 0.10\), \(b = 195 \  (13,  15)\), \(m = 90\), benchmark value
-to remove large edges is \(0.166\)) shows small separation between the
-clusters. Because of the small separation, the three clusters are
-connected in \gD{} and appear as long edges in \(9\text{-}D\). The full
-video for the model for layout a is available at
-\url{https://youtu.be/5Y1hE4i7N2k}. The full video for the model for
-layout c is available at \url{https://youtu.be/5Y1hE4i7N2k}.}
+\caption{\label{fig-model-pbmc-author-proj}}
 

### 2024-09-25 paper.qmd
Updated Figure 15 with only one projection

@@
 projection <- cbind(
-    c(0.24156,0.85058,0.11564,-0.03255,-0.28106,0.27463,-0.10539,0.08429,0.13107),
-    c(-0.67212,0.02745,-0.26718,-0.33126,-0.34919,0.20839,-0.14498,-0.37708,0.15455))
+    c(-0.4545,-0.5533,-0.1235,0.3984,-0.2065,-0.3586,-0.0088,0.3572,-0.1368),
+    c(0.4576,-0.3157,0.4708,0.1410,-0.4479,-0.1685,0.3668,-0.2692,0.1332))
 
@@
   limits = 1,
-  axis_pos_x = -0.4,
-  axis_pos_y = -0.4,
+  axis_pos_x = -0.48,
+  axis_pos_y = -0.48,
   axis_labels = names(scaled_pbmc_data),
-  threshold = 0.057)
+  threshold = 0.082)
 
@@
 pbmc_proj_umap_model1 <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#33a02c",
-    linewidth = 1) +
-  geom_point(
-    #size = 0.5,
-    alpha = 0.2) +
-  scale_color_manual(values = c("#000000", "#33a02c")) +
-  geom_segment(
-    data=axes, 
-    aes(x=x1, y=y1, xend=x2, yend=y2), 
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 3) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.6, 1)) +
-  ylim(c(-0.6, 1)) +
-  interior_annotation("a2", cex = 2)
-
-## Second projection
-projection <- cbind(
-    c(0.00067,0.53963,-0.16241,-0.09777,0.32766,0.44543,-0.51359,-0.29884,0.00951),
-    c(0.62399,0.26016,0.06441,-0.34686,-0.40466,0.24643,0.23537,0.07972,0.34402))
-
-projection_scaled <- projection * 1.25
-
-projected <- as.matrix(scaled_pbmc_data) %*% projection_scaled
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  #dplyr::mutate(type = df_exe$type) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection_scaled
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  distance_df_small_edges_pbmc |> select(-distance), 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-axes_obj <- gen_axes(
-  proj = projection,
-  limits = 1,
-  axis_pos_x = -0.6,
-  axis_pos_y = -0.7,
-  axis_labels = names(scaled_pbmc_data),
-  threshold = 0.086)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
-
-pbmc_proj_umap_model2 <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#33a02c",
-    linewidth = 1) +
-  geom_point(
-    #size = 0.5,
-    alpha = 0.2) +
-  scale_color_manual(values = c("#000000", "#33a02c")) +
-  geom_segment(
-    data=axes, 
-    aes(x=x1, y=y1, xend=x2, yend=y2), 
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 3) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.9, 0.5)) +
-  ylim(c(-0.9, 0.5)) +
-  interior_annotation("a3", cex = 2)
-
-```
-
-<!-- ::: {#fig-pbmc1-sc layout-ncol="4" fig-pos="H"} -->
-<!-- ![](figures/pbmc3k/umap_trimesh_plot.png) -->
-
-<!-- ![](figures/pbmc3k/sc_1.png) -->
-
-<!-- ![](figures/pbmc3k/sc_2.png) -->
-
-<!-- ![](figures/pbmc3k/sc_3.png) -->
-
-<!-- Model in \gD{}, on the UMAP layout, and three views of the fit in projections from $9\text{-}D$, for the PBMC3k data ($(s_1, \ s_2) = (-0.050, \ -0.041)$, $b = 870 \  (30, \ 29)$, $m = 135$, benchmark value to remove large edges is $0.099$) (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/VqqWuE0Jj6A>).
--->
-<!-- ::: -->
-<!--compute absolute error for different parameter choices-->
-
-```{r}
-#| label: combine-data-pbmc
-
-error_pbmc_umap <- read_rds("data/pbmc3k/error_pbmc_umap_30_min_dist_0.3.rds")
-error_pbmc_umap2 <- read_rds("data/pbmc3k/error_pbmc_umap_5_min_dist_0.01.rds")
-# error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_15_min_dist_0.99.rds")
-error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_12_min_dist_0.99.rds")
-error_pbmc_tsne <- read_rds("data/pbmc3k/error_pbmc_tsne_5.rds")
-error_pbmc_tsne2 <- read_rds("data/pbmc3k/error_pbmc_tsne_30.rds")
-error_pbmc_phate <- read_rds("data/pbmc3k/error_pbmc_phate_5.rds")
-error_pbmc_trimap <- read_rds("data/pbmc3k/error_pbmc_trimap_12_4_3.rds")
-error_pbmc_pacmap <- read_rds("data/pbmc3k/error_pbmc_pacmap_30_random_0.9_5.rds")
-
-error_pbmc <- bind_rows(error_pbmc_umap, 
-                        error_pbmc_umap2,
-                        error_pbmc_umap3,
-                        error_pbmc_tsne,
-                        error_pbmc_tsne2,
-                        error_pbmc_phate,
-                        error_pbmc_trimap,
-                        error_pbmc_pacmap)
-
-error_pbmc <- error_pbmc |>
-  mutate(a1 = round(a1, 2)) |>
-  filter(bin1 >= 5) |>
-  group_by(method, a1) |>
-  filter(MSE == min(MSE)) |>
-  ungroup()
-```
-
-```{r}
-#| label: error-comp-pbmc
-# read the png file from device
-
-#pbmc_layouts <- readPNG("figures/pbmc3k/pbmc_nldr_layouts.png", native = TRUE)
-
-error_plot_pbmc <- ggplot(error_pbmc, 
-                          aes(x = a1, 
-                              y = MSE, 
-                              colour = method)) + 
-  geom_point(size = 0.8) +
-  geom_line(linewidth = 0.3) + 
-  # geom_vline(xintercept = 15, linetype="solid", 
-  #            color = "black", linewidth=0.8, alpha = 0.5) +
-  scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 13, 17, 21, 26)]) +
-  scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a','#ff7f00',
-                              '#984ea3','#999999','#a65628','#f781bf')) +
-  scale_y_log10() +
-  ylab("MSE") +
-  xlab(expression(paste("binwidth (", a[1], ")"))) +
-  theme_minimal() +
-  theme(panel.border = element_rect(fill = 'transparent'),
-        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
-        axis.ticks.x = element_line(),
-        axis.ticks.y = element_line(),
-        legend.position = "none",
-        axis.text.x = element_text(size = 7),
-        axis.text.y = element_text(size = 7),
-        axis.title.x = element_text(size = 7),
-        axis.title.y = element_text(size = 7))
-
-```
-
-```{r}
-#| fig-cap: "Assessing which of the 8 NLDR layouts on the PBMC3k data  (shown in @fig-NLDR-variety) is the better representation using MSE for varying binwidth ($a_1$).
-Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-h).
-Layout f is universally poor.
-Layouts a, b, g, h that show large separations between clusters are universally suboptimal.
-Layout d with little separation performs well at tiny binwidth (where most points are in their own bin) and poorly as binwidth increases.
-The choice of best is between layouts c and e, that have small separations between oddly shaped clusters.
-Layout e is the best choice."
-#| label: fig-pbmc-mse
-#| fig-pos: H
-#| out-height: 100%
-
-free(error_plot_pbmc) + wrap_plots(nldr1c, nldr2c, nldr3c, nldr4c,
-           nldr5c, nldr6c, nldr7c, nldr8c, ncol = 2)
-```
-
-<!-- The MSE for different binwidths ($a_1$) using tSNE, UMAP, PHATE, PaCMAP, and TriMAP with different (hyper-)parameter settings (@fig-pbmc-mse) were calculated.  -->
-
-<!-- It is found that tSNE (@fig-mnist-mse (a)) provide the most reasonable representation for the digit 1 dataset, showing universally best.
-However, \gD{} representation shows a big non-linear cluster and a small cluster with a small gap (@fig-tsne-best).
--->
-
-<!-- In order to find a reasonable NLDR representation for the PBMC3k dataset, the MSE for different $a_1$ using UMAP, tSNE, PHATE, PaCMAP, and TriMAP with different (hyper-)parameter settings (@fig-pbmc-mse) were calculated.
-tSNE with a perplexity value set to $30$, the default parameter setting, is considered as a reasonable representation for the PBMC3k dataset.
-As shown in @fig-NLDR-variety (e), there are three well-separated clusters, although the separation between the clusters are small.
-Additionally, non-linear structures can also be observed within the clusters (@fig-NLDR-variety (e)).
--->
-<!-- The visualization of the *model-in-the-data-space* provides evidence for this (@fig-pbmc2-sc).
-Therefore, tSNE with perplexity $30$ accurately captures the data structure of the PBMC3k dataset, which is better than UMAP.
--->
-
-<!--best choice-->
-
-```{r}
-#| label: umap-pbmc-read-best
-umap_pbmc2 <- read_rds("data/pbmc3k/pbmc_umap_12_min_dist_0.99.rds")
-pbmc_scaled_obj <- gen_scaled_data(
-  data = umap_pbmc2)
-umap_pbmc_scaled_best <- pbmc_scaled_obj$scaled_nldr
-
-# tsne_pbmc <- tsne_pbmc_scaled |>
-#   ggplot(aes(x = tSNE1,
-#              y = tSNE2)) +
-#   geom_point(alpha=0.3, color = "#000000")
-```
-
-<!-- We then fit the model for tSNE, and visualize the resultant model in the $9\text{-}D$ data space.
-The model shows a quirk, as shown in @fig-pbmc2-sc.
-All three clusters are connected by an edge except the small and large clusters.
-Because the clusters are so close in \gD{}, they attempt to maintain the structure in \pD{} as well.
-This is evident that tSNE with perplexity $30$ provides a reasonable representation of PBMC3k data.
--->
-
-<!--Fit the best model and compute error-->
-```{r}
-#| label: num-bins-pbmc
-## Compute hexbin parameters
-num_bins_x_pbmc <- 13
-lim1 <- pbmc_scaled_obj$lim1
-lim2 <- pbmc_scaled_obj$lim2
-r2_pbmc <- diff(lim2)/diff(lim1)
-
-pbmc_model <- fit_highd_model(
-  training_data = training_data_pbmc,
-  emb_df = umap_pbmc_scaled_best,
-  bin1 = num_bins_x_pbmc,
-  r2 = r2_pbmc,
-  q = 0.1,
-  is_bin_centroid = TRUE,
-  is_rm_lwd_hex = FALSE,
-  col_start_highd = "pc"
-)
-
-df_bin_centroids_pbmc <- pbmc_model$df_bin_centroids
-df_bin_pbmc <- pbmc_model$df_bin
-
-## Triangulate bin centroids
-tr1_object_pbmc <- tri_bin_centroids(
-  df_bin_centroids_pbmc, x = "c_x", y = "c_y")
-tr_from_to_df_pbmc <- gen_edges(
-  tri_object = tr1_object_pbmc)
-
-# tr_from_to_df_pbmc <- tr_from_to_df_pbmc |>
-#   filter(!(row_number() %in% c(65, 112, 130)))
-
-## Compute 2D distances
-distance_pbmc <- cal_2d_dist(
-  tr_coord_df = tr_from_to_df_pbmc,
-  start_x = "x_from",
-  start_y = "y_from",
-  end_x = "x_to",
-  end_y = "y_to",
-  select_vars = c("from", "to", "distance"))
-
-## To find the benchmark value
-benchmark_pbmc <- find_lg_benchmark(
-  distance_edges = distance_pbmc,
-  distance_col = "distance")
-
-#benchmark_pbmc <- 0.1
-
-## Remove an edge that exist as long edge in p-D
-# tr_from_to_df_pbmc |> mutate(ID = row_number()) |> filter(from == 37) |> filter(to == 43)
-
-tr_df <- distinct(tibble::tibble(
-  x = c(tr_from_to_df_pbmc[["x_from"]], tr_from_to_df_pbmc[["x_to"]]), 
-  y = c(tr_from_to_df_pbmc[["y_from"]], tr_from_to_df_pbmc[["y_to"]])))
-
-distance_df_small_edges_pbmc <- distance_pbmc |>
-  filter(distance < benchmark_pbmc)
-
-tr_from_to_df_pbmc <- inner_join(
-  tr_from_to_df_pbmc, distance_df_small_edges_pbmc, 
-  by = c("from", "to"))
-
-# tr_from_to_df_pbmc <- tr_from_to_df_pbmc |>
-#   filter(!(row_number() %in% c(155)))
-
-trimesh_removed_pbmc_best <- ggplot() + 
-  geom_segment(data = tr_from_to_df_pbmc, 
-               aes(
-                 x = x_from, 
-                 y = y_from, 
-                 xend = x_to, 
-                 yend = y_to),
-               colour = "#33a02c",
-               linewidth = 1) +
-  geom_point(
-    data = umap_pbmc_scaled_best,
-    aes(
-      x = UMAP1,
-      y = UMAP2
-    ),
-    alpha = 0.2,
-    color = "#000000"
-  )  +
-  interior_annotation("b1", cex = 2) +
-  theme(
-    aspect.ratio = 1
-  )
-
-# trimesh_removed_pbmc <- vis_rmlg_mesh(
-#   distance_edges = distance_pbmc,
-#   benchmark_value = benchmark_pbmc,
-#   tr_coord_df = tr_from_to_df_pbmc,
-#   distance_col = "distance") +
-#   geom_point(
-#     data = tsne_pbmc_scaled,
-#     aes(
-#       x = tSNE1,
-#       y = tSNE2
-#     ),
-#     alpha = 0.2,
-#     color = "#000000"
-#   ) 
-#   #xlim(sc_xlims) + ylim(sc_ylims) +
-#   #interior_annotation("a", sc_ltr_pos)
-
-## Compute error
-error_df <- augment(
-  df_bin_centroids = df_bin_centroids_pbmc,
-  df_bin = df_bin_pbmc,
-  training_data = training_data_pbmc,
-  newdata = NULL,
-  type_NLDR = "UMAP",
-  col_start = "pc")
-
-## Categorize error
-
-error_df <- error_df |>
-  mutate(type = case_when(
-    row_wise_abs_error <= 5 ~ "error 0-5",
-    row_wise_abs_error <= 10 ~ "error 5-10",
-    row_wise_abs_error <= 15 ~ "error 10-15",
-    row_wise_abs_error <= 20 ~ "error 15-20",
-    row_wise_abs_error <= 25 ~ "error 20-25",
-    .default = "error greter than 25"
-  )) |>
-  mutate(type = factor(type, levels = c(
-    "error 0-5", "error 5-10", "error 10-15", "error 15-20", "error 20-25", 
-    "error greter than 25")))
-
-## To join embedding
-error_df <- error_df |>
-  bind_cols(umap_pbmc_scaled_best |>
-              select(-ID))
-
-error_plot_pbmc_best <- error_df |>
-  ggplot(aes(x = UMAP1,
-             y = UMAP2,
-             color = type,
-             group = ID)) +
-  geom_point(alpha=0.5,
-             size = 0.1) +
-  interior_annotation("b", sc_ltr_pos)
-
-```
-
-```{r}
-#| label: pbmc-umap-model-proj-best
-
-df_b_pbmc <- df_bin_pbmc |>
-  dplyr::filter(hb_id %in% df_bin_centroids_pbmc$hexID) |>
-  dplyr::mutate(type = "model") ## Data with summarized mean
-
-## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
-df_b_pbmc <- df_b_pbmc[match(df_bin_centroids_pbmc$hexID, df_b_pbmc$hb_id),] |>
-  dplyr::select(-hb_id)
-
-# Apply the scaling
-df_model_data_pbmc <- bind_rows(data_pbmc, df_b_pbmc)
-scaled_pbmc <- scale_data_manual(df_model_data_pbmc, "type") |>
-  as_tibble()
-
-scaled_pbmc_data <- scaled_pbmc |>
-  filter(type == "data") |>
-  select(-type)
-
-scaled_pbmc_data_model <- scaled_pbmc |>
-  filter(type == "model") |>
-  select(-type)
-
-# Combine with the true model for visualization
-df <- dplyr::bind_rows(scaled_pbmc_data_model |> mutate(type = "model"),
-                       scaled_pbmc_data |> mutate(type = "data"))
-
-## Removed long edge in high-d
-# distance_df_small_edges_pbmc <- distance_df_small_edges_pbmc |>
-#   filter(!(row_number() %in% c(155)))
-
-## First projection
-projection <- cbind(
-    c(0.24156,0.85058,0.11564,-0.03255,-0.28106,0.27463,-0.10539,0.08429,0.13107),
-    c(-0.67212,0.02745,-0.26718,-0.33126,-0.34919,0.20839,-0.14498,-0.37708,0.15455))
-
-projection_scaled <- projection * 1.23
-
-projected <- as.matrix(scaled_pbmc_data) %*% projection_scaled
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  #dplyr::mutate(type = df_exe$type) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection_scaled
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  distance_df_small_edges_pbmc |> select(-distance), 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-axes_obj <- gen_axes(
-  proj = projection,
-  limits = 1,
-  axis_pos_x = -0.4,
-  axis_pos_y = -0.4,
-  axis_labels = names(scaled_pbmc_data),
-  threshold = 0.057)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
-
-pbmc_proj_umap_model1_best <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#33a02c",
-    linewidth = 1) +
-  geom_point(
-    #size = 0.5,
-    alpha = 0.2) +
-  scale_color_manual(values = c("#000000", "#33a02c")) +
-  geom_segment(
-    data=axes, 
-    aes(x=x1, y=y1, xend=x2, yend=y2), 
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 3) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.6, 1)) +
-  ylim(c(-0.6, 1)) +
-  interior_annotation("b2", cex = 2)
-
-## Second projection
-projection <- cbind(
-    c(0.00067,0.53963,-0.16241,-0.09777,0.32766,0.44543,-0.51359,-0.29884,0.00951),
-    c(0.62399,0.26016,0.06441,-0.34686,-0.40466,0.24643,0.23537,0.07972,0.34402))
-
-projection_scaled <- projection * 1.25
-
-projected <- as.matrix(scaled_pbmc_data) %*% projection_scaled
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  #dplyr::mutate(type = df_exe$type) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection_scaled
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  distance_df_small_edges_pbmc |> select(-distance), 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-axes_obj <- gen_axes(
-  proj = projection,
-  limits = 1,
-  axis_pos_x = -0.6,
-  axis_pos_y = -0.7,
-  axis_labels = names(scaled_pbmc_data),
-  threshold = 0.086)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
-
-pbmc_proj_umap_model2_best <- projected_df |>
   ggplot(
@@
   coord_fixed() +
-  xlim(c(-0.9, 0.5)) +
-  ylim(c(-0.9, 0.5)) +
+  xlim(c(-0.65, 0.65)) +
+  ylim(c(-0.65, 0.65)) +
+  interior_annotation("a2", cex = 2)
+
+## Second projection
+projection <- cbind(
+    c(-0.4545,-0.5533,-0.1235,0.3984,-0.2065,-0.3586,-0.0088,0.3572,-0.1368),
+    c(0.4576,-0.3157,0.4708,0.1410,-0.4479,-0.1685,0.3668,-0.2692,0.1332))
+
+projection_scaled <- projection * 1.23
+
+projected <- as.matrix(scaled_pbmc_data) %*% projection_scaled
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  #dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection_scaled
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges_pbmc |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+axes_obj <- gen_axes(
+  proj = projection,
+  limits = 1,
+  axis_pos_x = -0.48,
+  axis_pos_y = -0.48,
+  axis_labels = names(scaled_pbmc_data),
+  threshold = 0.082)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
+
+# pbmc_proj_umap_model2 <- projected_df |>
+#   ggplot(
+#     aes(
+#       x = proj1, 
+#       y = proj2)) +
+#   geom_segment(
+#     data = model_df, 
+#     aes(
+#       x = proj1_from, 
+#       y = proj2_from, 
+#       xend = proj1_to, 
+#       yend = proj2_to), 
+#     color = "#33a02c",
+#     linewidth = 1) +
+#   geom_point(
+#     #size = 0.5,
+#     alpha = 0.2) +
+#   scale_color_manual(values = c("#000000", "#33a02c")) +
+#   geom_segment(
+#     data=axes, 
+#     aes(x=x1, y=y1, xend=x2, yend=y2), 
+#     colour="grey70") +
+#   geom_text(
+#     data=axes,
+#     aes(x=x2, y=y2), 
+#     label=rownames(axes),
+#     colour="grey50",
+#     size = 3) +
+#   geom_path(
+#     data=circle, 
+#     aes(x=c1, y=c2), colour="grey70") +
+#   coord_fixed() +
+#   xlim(c(-0.9, 0.5)) +
+#   ylim(c(-0.9, 0.5)) +
+#   interior_annotation("a3", cex = 2)
+
+pbmc_proj_umap_model2 <- ggplot() +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#33a02c") +
+  scale_color_manual(values = c("#000000", "#33a02c")) +
+  geom_segment(
+    data=axes, 
+    aes(x=x1, y=y1, xend=x2, yend=y2), 
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
+    colour="grey50",
+    size = 3) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.65, 0.65)) +
+  ylim(c(-0.65, 0.65)) +
+  interior_annotation("a3", cex = 2)
+
+```
+
+<!-- ::: {#fig-pbmc1-sc layout-ncol="4" fig-pos="H"} -->
+<!-- ![](figures/pbmc3k/umap_trimesh_plot.png) -->
+
+<!-- ![](figures/pbmc3k/sc_1.png) -->
+
+<!-- ![](figures/pbmc3k/sc_2.png) -->
+
+<!-- ![](figures/pbmc3k/sc_3.png) -->
+
+<!-- Model in \gD{}, on the UMAP layout, and three views of the fit in projections from $9\text{-}D$, for the PBMC3k data ($(s_1, \ s_2) = (-0.050, \ -0.041)$, $b = 870 \  (30, \ 29)$, $m = 135$, benchmark value to remove large edges is $0.099$) (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/VqqWuE0Jj6A>).
+-->
+<!-- ::: -->
+<!--compute absolute error for different parameter choices-->
+
+```{r}
+#| label: combine-data-pbmc
+
+error_pbmc_umap <- read_rds("data/pbmc3k/error_pbmc_umap_30_min_dist_0.3.rds")
+error_pbmc_umap2 <- read_rds("data/pbmc3k/error_pbmc_umap_5_min_dist_0.01.rds")
+# error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_15_min_dist_0.99.rds")
+error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_12_min_dist_0.99.rds")
+error_pbmc_tsne <- read_rds("data/pbmc3k/error_pbmc_tsne_5.rds")
+error_pbmc_tsne2 <- read_rds("data/pbmc3k/error_pbmc_tsne_30.rds")
+error_pbmc_phate <- read_rds("data/pbmc3k/error_pbmc_phate_5.rds")
+error_pbmc_trimap <- read_rds("data/pbmc3k/error_pbmc_trimap_12_4_3.rds")
+error_pbmc_pacmap <- read_rds("data/pbmc3k/error_pbmc_pacmap_30_random_0.9_5.rds")
+
+error_pbmc <- bind_rows(error_pbmc_umap, 
+                        error_pbmc_umap2,
+                        error_pbmc_umap3,
+                        error_pbmc_tsne,
+                        error_pbmc_tsne2,
+                        error_pbmc_phate,
+                        error_pbmc_trimap,
+                        error_pbmc_pacmap)
+
+error_pbmc <- error_pbmc |>
+  mutate(a1 = round(a1, 2)) |>
+  filter(bin1 >= 5) |>
+  group_by(method, a1) |>
+  filter(MSE == min(MSE)) |>
+  ungroup()
+```
+
+```{r}
+#| label: error-comp-pbmc
+# read the png file from device
+
+#pbmc_layouts <- readPNG("figures/pbmc3k/pbmc_nldr_layouts.png", native = TRUE)
+
+error_plot_pbmc <- ggplot(error_pbmc, 
+                          aes(x = a1, 
+                              y = MSE, 
+                              colour = method)) + 
+  geom_point(size = 0.8) +
+  geom_line(linewidth = 0.3) + 
+  # geom_vline(xintercept = 15, linetype="solid", 
+  #            color = "black", linewidth=0.8, alpha = 0.5) +
+  scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 13, 17, 21, 26)]) +
+  scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a','#ff7f00',
+                              '#984ea3','#999999','#a65628','#f781bf')) +
+  scale_y_log10() +
+  ylab("MSE") +
+  xlab(expression(paste("binwidth (", a[1], ")"))) +
+  theme_minimal() +
+  theme(panel.border = element_rect(fill = 'transparent'),
+        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+        axis.ticks.x = element_line(),
+        axis.ticks.y = element_line(),
+        legend.position = "none",
+        axis.text.x = element_text(size = 7),
+        axis.text.y = element_text(size = 7),
+        axis.title.x = element_text(size = 7),
+        axis.title.y = element_text(size = 7))
+
+```
+
+```{r}
+#| fig-cap: "Assessing which of the 8 NLDR layouts on the PBMC3k data  (shown in @fig-NLDR-variety) is the better representation using MSE for varying binwidth ($a_1$).
+Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-h).
+Layout f is universally poor.
+Layouts a, b, g, h that show large separations between clusters are universally suboptimal.
+Layout d with little separation performs well at tiny binwidth (where most points are in their own bin) and poorly as binwidth increases.
+The choice of best is between layouts c and e, that have small separations between oddly shaped clusters.
+Layout e is the best choice."
+#| label: fig-pbmc-mse
+#| fig-pos: H
+#| out-height: 100%
+
+free(error_plot_pbmc) + wrap_plots(nldr1c, nldr2c, nldr3c, nldr4c,
+           nldr5c, nldr6c, nldr7c, nldr8c, ncol = 2)
+```
+
+<!-- The MSE for different binwidths ($a_1$) using tSNE, UMAP, PHATE, PaCMAP, and TriMAP with different (hyper-)parameter settings (@fig-pbmc-mse) were calculated.  -->
+
+<!-- It is found that tSNE (@fig-mnist-mse (a)) provide the most reasonable representation for the digit 1 dataset, showing universally best.
+However, \gD{} representation shows a big non-linear cluster and a small cluster with a small gap (@fig-tsne-best).
+-->
+
+<!-- In order to find a reasonable NLDR representation for the PBMC3k dataset, the MSE for different $a_1$ using UMAP, tSNE, PHATE, PaCMAP, and TriMAP with different (hyper-)parameter settings (@fig-pbmc-mse) were calculated.
+tSNE with a perplexity value set to $30$, the default parameter setting, is considered as a reasonable representation for the PBMC3k dataset.
+As shown in @fig-NLDR-variety (e), there are three well-separated clusters, although the separation between the clusters are small.
+Additionally, non-linear structures can also be observed within the clusters (@fig-NLDR-variety (e)).
+-->
+<!-- The visualization of the *model-in-the-data-space* provides evidence for this (@fig-pbmc2-sc).
+Therefore, tSNE with perplexity $30$ accurately captures the data structure of the PBMC3k dataset, which is better than UMAP.
+-->
+
+<!--best choice-->
+
+```{r}
+#| label: umap-pbmc-read-best
+umap_pbmc2 <- read_rds("data/pbmc3k/pbmc_umap_12_min_dist_0.99.rds")
+pbmc_scaled_obj <- gen_scaled_data(
+  data = umap_pbmc2)
+umap_pbmc_scaled_best <- pbmc_scaled_obj$scaled_nldr
+
+# tsne_pbmc <- tsne_pbmc_scaled |>
+#   ggplot(aes(x = tSNE1,
+#              y = tSNE2)) +
+#   geom_point(alpha=0.3, color = "#000000")
+```
+
+<!-- We then fit the model for tSNE, and visualize the resultant model in the $9\text{-}D$ data space.
+The model shows a quirk, as shown in @fig-pbmc2-sc.
+All three clusters are connected by an edge except the small and large clusters.
+Because the clusters are so close in \gD{}, they attempt to maintain the structure in \pD{} as well.
+This is evident that tSNE with perplexity $30$ provides a reasonable representation of PBMC3k data.
+-->
+
+<!--Fit the best model and compute error-->
+```{r}
+#| label: num-bins-pbmc
+## Compute hexbin parameters
+num_bins_x_pbmc <- 13
+lim1 <- pbmc_scaled_obj$lim1
+lim2 <- pbmc_scaled_obj$lim2
+r2_pbmc <- diff(lim2)/diff(lim1)
+
+pbmc_model <- fit_highd_model(
+  training_data = training_data_pbmc,
+  emb_df = umap_pbmc_scaled_best,
+  bin1 = num_bins_x_pbmc,
+  r2 = r2_pbmc,
+  q = 0.1,
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = FALSE,
+  col_start_highd = "pc"
+)
+
+df_bin_centroids_pbmc <- pbmc_model$df_bin_centroids
+df_bin_pbmc <- pbmc_model$df_bin
+
+## Triangulate bin centroids
+tr1_object_pbmc <- tri_bin_centroids(
+  df_bin_centroids_pbmc, x = "c_x", y = "c_y")
+tr_from_to_df_pbmc <- gen_edges(
+  tri_object = tr1_object_pbmc)
+
+# tr_from_to_df_pbmc <- tr_from_to_df_pbmc |>
+#   filter(!(row_number() %in% c(65, 112, 130)))
+
+## Compute 2D distances
+distance_pbmc <- cal_2d_dist(
+  tr_coord_df = tr_from_to_df_pbmc,
+  start_x = "x_from",
+  start_y = "y_from",
+  end_x = "x_to",
+  end_y = "y_to",
+  select_vars = c("from", "to", "distance"))
+
+## To find the benchmark value
+benchmark_pbmc <- find_lg_benchmark(
+  distance_edges = distance_pbmc,
+  distance_col = "distance")
+
+#benchmark_pbmc <- 0.1
+
+## Remove an edge that exist as long edge in p-D
+# tr_from_to_df_pbmc |> mutate(ID = row_number()) |> filter(from == 37) |> filter(to == 43)
+
+tr_df <- distinct(tibble::tibble(
+  x = c(tr_from_to_df_pbmc[["x_from"]], tr_from_to_df_pbmc[["x_to"]]), 
+  y = c(tr_from_to_df_pbmc[["y_from"]], tr_from_to_df_pbmc[["y_to"]])))
+
+distance_df_small_edges_pbmc <- distance_pbmc |>
+  filter(distance < benchmark_pbmc)
+
+tr_from_to_df_pbmc <- inner_join(
+  tr_from_to_df_pbmc, distance_df_small_edges_pbmc, 
+  by = c("from", "to"))
+
+# tr_from_to_df_pbmc <- tr_from_to_df_pbmc |>
+#   filter(!(row_number() %in% c(155)))
+
+trimesh_removed_pbmc_best <- ggplot() + 
+  geom_segment(data = tr_from_to_df_pbmc, 
+               aes(
+                 x = x_from, 
+                 y = y_from, 
+                 xend = x_to, 
+                 yend = y_to),
+               colour = "#33a02c",
+               linewidth = 1) +
+  geom_point(
+    data = umap_pbmc_scaled_best,
+    aes(
+      x = UMAP1,
+      y = UMAP2
+    ),
+    alpha = 0.2,
+    color = "#000000"
+  )  +
+  interior_annotation("b1", cex = 2) +
+  theme(
+    aspect.ratio = 1
+  )
+
+# trimesh_removed_pbmc <- vis_rmlg_mesh(
+#   distance_edges = distance_pbmc,
+#   benchmark_value = benchmark_pbmc,
+#   tr_coord_df = tr_from_to_df_pbmc,
+#   distance_col = "distance") +
+#   geom_point(
+#     data = tsne_pbmc_scaled,
+#     aes(
+#       x = tSNE1,
+#       y = tSNE2
+#     ),
+#     alpha = 0.2,
+#     color = "#000000"
+#   ) 
+#   #xlim(sc_xlims) + ylim(sc_ylims) +
+#   #interior_annotation("a", sc_ltr_pos)
+
+## Compute error
+error_df <- augment(
+  df_bin_centroids = df_bin_centroids_pbmc,
+  df_bin = df_bin_pbmc,
+  training_data = training_data_pbmc,
+  newdata = NULL,
+  type_NLDR = "UMAP",
+  col_start = "pc")
+
+## Categorize error
+
+error_df <- error_df |>
+  mutate(type = case_when(
+    row_wise_abs_error <= 5 ~ "error 0-5",
+    row_wise_abs_error <= 10 ~ "error 5-10",
+    row_wise_abs_error <= 15 ~ "error 10-15",
+    row_wise_abs_error <= 20 ~ "error 15-20",
+    row_wise_abs_error <= 25 ~ "error 20-25",
+    .default = "error greter than 25"
+  )) |>
+  mutate(type = factor(type, levels = c(
+    "error 0-5", "error 5-10", "error 10-15", "error 15-20", "error 20-25", 
+    "error greter than 25")))
+
+## To join embedding
+error_df <- error_df |>
+  bind_cols(umap_pbmc_scaled_best |>
+              select(-ID))
+
+error_plot_pbmc_best <- error_df |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2,
+             color = type,
+             group = ID)) +
+  geom_point(alpha=0.5,
+             size = 0.1) +
+  interior_annotation("b", sc_ltr_pos)
+
+```
+
+```{r}
+#| label: pbmc-umap-model-proj-best
+
+df_b_pbmc <- df_bin_pbmc |>
+  dplyr::filter(hb_id %in% df_bin_centroids_pbmc$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b_pbmc <- df_b_pbmc[match(df_bin_centroids_pbmc$hexID, df_b_pbmc$hb_id),] |>
+  dplyr::select(-hb_id)
+
+# Apply the scaling
+df_model_data_pbmc <- bind_rows(data_pbmc, df_b_pbmc)
+scaled_pbmc <- scale_data_manual(df_model_data_pbmc, "type") |>
+  as_tibble()
+
+scaled_pbmc_data <- scaled_pbmc |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_pbmc_data_model <- scaled_pbmc |>
+  filter(type == "model") |>
+  select(-type)
+
+# Combine with the true model for visualization
+df <- dplyr::bind_rows(scaled_pbmc_data_model |> mutate(type = "model"),
+                       scaled_pbmc_data |> mutate(type = "data"))
+
+## Removed long edge in high-d
+# distance_df_small_edges_pbmc <- distance_df_small_edges_pbmc |>
+#   filter(!(row_number() %in% c(155)))
+
+## First projection
+projection <- cbind(
+    c(-0.4545,-0.5533,-0.1235,0.3984,-0.2065,-0.3586,-0.0088,0.3572,-0.1368),
+    c(0.4576,-0.3157,0.4708,0.1410,-0.4479,-0.1685,0.3668,-0.2692,0.1332))
+
+projection_scaled <- projection * 1.23
+
+projected <- as.matrix(scaled_pbmc_data) %*% projection_scaled
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  #dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection_scaled
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges_pbmc |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+axes_obj <- gen_axes(
+  proj = projection,
+  limits = 1,
+  axis_pos_x = -0.48,
+  axis_pos_y = -0.48,
+  axis_labels = names(scaled_pbmc_data),
+  threshold = 0.082)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
+
+pbmc_proj_umap_model1_best <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#33a02c") +
+  geom_point(
+    #size = 0.5,
+    alpha = 0.2) +
+  scale_color_manual(values = c("#000000", "#33a02c")) +
+  geom_segment(
+    data=axes, 
+    aes(x=x1, y=y1, xend=x2, yend=y2), 
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
+    colour="grey50",
+    size = 3) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.65, 0.65)) +
+  ylim(c(-0.65, 0.65)) +
+  interior_annotation("b2", cex = 2)
+
+## Second projection
+projection <- cbind(
+    c(-0.4545,-0.5533,-0.1235,0.3984,-0.2065,-0.3586,-0.0088,0.3572,-0.1368),
+    c(0.4576,-0.3157,0.4708,0.1410,-0.4479,-0.1685,0.3668,-0.2692,0.1332))
+
+projection_scaled <- projection * 1.23
+
+projected <- as.matrix(scaled_pbmc_data) %*% projection_scaled
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  #dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection_scaled
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges_pbmc |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+axes_obj <- gen_axes(
+  proj = projection,
+  limits = 1,
+  axis_pos_x = -0.48,
+  axis_pos_y = -0.48,
+  axis_labels = names(scaled_pbmc_data),
+  threshold = 0.082)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
+
+# pbmc_proj_umap_model2_best <- projected_df |>
+#   ggplot(
+#     aes(
+#       x = proj1, 
+#       y = proj2)) +
+#   geom_segment(
+#     data = model_df, 
+#     aes(
+#       x = proj1_from, 
+#       y = proj2_from, 
+#       xend = proj1_to, 
+#       yend = proj2_to), 
+#     color = "#33a02c") +
+#   geom_point(
+#     #size = 0.5,
+#     alpha = 0.2) +
+#   scale_color_manual(values = c("#000000", "#33a02c")) +
+#   geom_segment(
+#     data=axes, 
+#     aes(x=x1, y=y1, xend=x2, yend=y2), 
+#     colour="grey70") +
+#   geom_text(
+#     data=axes,
+#     aes(x=x2, y=y2), 
+#     label=rownames(axes),
+#     colour="grey50",
+#     size = 3) +
+#   geom_path(
+#     data=circle, 
+#     aes(x=c1, y=c2), colour="grey70") +
+#   coord_fixed() +
+#   xlim(c(-0.9, 0.5)) +
+#   ylim(c(-0.9, 0.5)) +
+#   interior_annotation("b3", cex = 2)
+
+pbmc_proj_umap_model2_best <- ggplot() +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#33a02c") +
+  geom_segment(
+    data=axes, 
+    aes(x=x1, y=y1, xend=x2, yend=y2), 
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
+    colour="grey50",
+    size = 3) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.65, 0.65)) +
+  ylim(c(-0.65, 0.65)) +
   interior_annotation("b3", cex = 2)

### 2024-09-25 paper.qmd
Updated Figure 15

@@
                  yend = y_to),
-               colour = "#33a02c",
+               colour = "#000000",
                linewidth = 1) +
@@
     alpha = 0.2,
-    color = "#000000"
+    color = '#a65628'
   )  +
@@
       yend = proj2_to), 
-    color = "#33a02c") +
+    color = "#000000") +
   geom_point(
     #size = 0.5,
-    alpha = 0.2) +
-  scale_color_manual(values = c("#000000", "#33a02c")) +
+    alpha = 0.2,
+    color = '#a65628') +
   geom_segment(
@@
       yend = proj2_to), 
-    color = "#33a02c") +
-  scale_color_manual(values = c("#000000", "#33a02c")) +
+    color = "#000000") +
   geom_segment(
@@
     ),
-    alpha = 0.2,
-    color = "#000000"
+    alpha = 0.3,
+    color = "#999999"
   )  +
@@
       yend = proj2_to), 
-    color = "#33a02c") +
+    color = "#000000") +
   geom_point(
     #size = 0.5,
-    alpha = 0.2) +
-  scale_color_manual(values = c("#000000", "#33a02c")) +
+    alpha = 0.2,
+    color = "#999999") +
   geom_segment(
@@
       yend = proj2_to), 
-    color = "#33a02c") +
+    color = "#000000") +
   geom_segment(
@@
 #| echo: false
-#| fig-cap: ""
+#| fig-cap: "Model in \\gD{}, on the layout a and c, and two views of the fit in projections from $9\\text{-}D$, for the PBMC3k data ($n =  2622$ and $p = 9$).
+Layout a shows three-well separated clusters with big separation, while layout c shows close clusters.
+In $9\\text{-}D$, the data shows close clusters.
+This is the reason where b1, b2, and b3 shows connected edges between the clusters, and also proved by @fig-pbmc-mse as the best layout, which used the same NLDR method.
+By viewing the fitted model in $9\\text{-}D$, help to see some unobserved pattern of the data (i) densed points, and (ii) non-linear clusters."
 #| label: fig-model-pbmc-author-proj

### 2024-09-25 paper.tex
Updated Figure 15

@@
 
-\caption{\label{fig-model-pbmc-author-proj}}
+\caption{\label{fig-model-pbmc-author-proj}Model in \gD{}, on the layout
+a and c, and two views of the fit in projections from \(9\text{-}D\),
+for the PBMC3k data (\(n =  2622\) and \(p = 9\)). Layout a shows
+three-well separated clusters with big separation, while layout c shows
+close clusters. In \(9\text{-}D\), the data shows close clusters. This
+is the reason where b1, b2, and b3 shows connected edges between the
+clusters, and also proved by Figure~\ref{fig-pbmc-mse} as the best
+layout, which used the same NLDR method. By viewing the fitted model in
+\(9\text{-}D\), help to see some unobserved pattern of the data (i)
+densed points, and (ii) non-linear clusters.}
 

### 2024-09-25 paper.qmd
Updated PBMC3k layouts lables to have the consistency

@@
 
-umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_5_min_dist_0.01.rds")
-
+umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_12_min_dist_0.99.rds")
 nldr2 <- umap_pbmc |>
@@
              y = UMAP2))+
-  geom_point(alpha=0.1, size=1, colour='#f781bf') +
+  geom_point(alpha=0.1, size=1, colour='#999999') +
   interior_annotation("b")
 
-#umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_15_min_dist_0.99.rds")
-umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_12_min_dist_0.99.rds")
+umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_5_min_dist_0.01.rds")
+
 nldr3 <- umap_pbmc |>
@@
              y = UMAP2))+
-  geom_point(alpha=0.1, size=1, colour='#999999') +
+  geom_point(alpha=0.1, size=1, colour='#f781bf') +
   interior_annotation("c")
+
+#umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_15_min_dist_0.99.rds")
 
@@
 Layout f is universally poor.
-Layouts a, b, g, h that show large separations between clusters are universally suboptimal.
+Layouts a, c, g, h that show large separations between clusters are universally suboptimal.
 Layout d with little separation performs well at tiny binwidth (where most points are in their own bin) and poorly as binwidth increases.
-The choice of best is between layouts c and e, that have small separations between oddly shaped clusters.
+The choice of best is between layouts b and e, that have small separations between oddly shaped clusters.
 Layout e is the best choice."
@@
 #| echo: false
-#| fig-cap: "Model in \\gD{}, on the layout a and c, and two views of the fit in projections from $9\\text{-}D$, for the PBMC3k data ($n =  2622$ and $p = 9$).
-Layout a shows three-well separated clusters with big separation, while layout c shows close clusters.
+#| fig-cap: "Model in \\gD{}, on the layout a and b, and two views of the fit in projections from $9\\text{-}D$, for the PBMC3k data ($n =  2622$ and $p = 9$).
+Layout a shows three-well separated clusters with big separation, while layout b shows close clusters.
 In $9\\text{-}D$, the data shows close clusters.

### 2024-09-25 paper.tex
Updated PBMC3k layouts lables to have the consistency

@@
 the lines and points in the left plot and in the scatterplots represents
-NLDR layout (a-h). Layout f is universally poor. Layouts a, b, g, h that
+NLDR layout (a-h). Layout f is universally poor. Layouts a, c, g, h that
 show large separations between clusters are universally suboptimal.
@@
 most points are in their own bin) and poorly as binwidth increases. The
-choice of best is between layouts c and e, that have small separations
+choice of best is between layouts b and e, that have small separations
 between oddly shaped clusters. Layout e is the best choice.}
@@
 \caption{\label{fig-model-pbmc-author-proj}Model in \gD{}, on the layout
-a and c, and two views of the fit in projections from \(9\text{-}D\),
+a and b, and two views of the fit in projections from \(9\text{-}D\),
 for the PBMC3k data (\(n =  2622\) and \(p = 9\)). Layout a shows
-three-well separated clusters with big separation, while layout c shows
+three-well separated clusters with big separation, while layout b shows
 close clusters. In \(9\text{-}D\), the data shows close clusters. This

### 2024-09-25 paper.qmd
Updated Figure 15 with new projection

@@
 projection <- cbind(
-    c(-0.4545,-0.5533,-0.1235,0.3984,-0.2065,-0.3586,-0.0088,0.3572,-0.1368),
-    c(0.4576,-0.3157,0.4708,0.1410,-0.4479,-0.1685,0.3668,-0.2692,0.1332))
-
-projection_scaled <- projection * 1.23
+    c(0.0734,0.6138,-0.4671,0.2589,0.1189,0.0778,-0.0819,0.0308,-0.5561),
+    c(0.6625,0.0034,0.2412,0.3360,0.2463,0.1070,-0.2903,-0.4705,0.1293))
+
+projection_scaled <- projection * 1.25
 
@@
   proj = projection,
-  limits = 1,
-  axis_pos_x = -0.48,
-  axis_pos_y = -0.48,
+  limits = 1.35,
+  axis_pos_x = -0.75,
+  axis_pos_y = -0.75,
   axis_labels = names(scaled_pbmc_data),
-  threshold = 0.082)
+  threshold = 0.12)
 
@@
 
-# pbmc_proj_umap_model2 <- projected_df |>
-#   ggplot(
-#     aes(
-#       x = proj1, 
-#       y = proj2)) +
+pbmc_proj_umap_model2 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1,
+      y = proj2)) +
+  geom_segment(
+    data = model_df,
+    aes(
+      x = proj1_from,
+      y = proj2_from,
+      xend = proj1_to,
+      yend = proj2_to),
+    color = "#000000") +
+  geom_point(
+    #size = 0.5,
+    alpha = 0.2,
+    color = '#a65628') +
+  geom_segment(
+    data=axes,
+    aes(x=x1, y=y1, xend=x2, yend=y2),
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2),
+    label=rownames(axes),
+    colour="grey50",
+    size = 3) +
+  geom_path(
+    data=circle,
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-1, 0.75)) +
+  ylim(c(-1, 0.75)) +
+  interior_annotation("a3", cex = 2)
+
+# pbmc_proj_umap_model2 <- ggplot() +
 #   geom_segment(
@@
 #       yend = proj2_to), 
-#     color = "#33a02c",
-#     linewidth = 1) +
-#   geom_point(
-#     #size = 0.5,
-#     alpha = 0.2) +
-#   scale_color_manual(values = c("#000000", "#33a02c")) +
+#     color = "#000000") +
 #   geom_segment(
@@
 #   coord_fixed() +
-#   xlim(c(-0.9, 0.5)) +
-#   ylim(c(-0.9, 0.5)) +
+#   xlim(c(-0.65, 0.65)) +
+#   ylim(c(-0.65, 0.65)) +
 #   interior_annotation("a3", cex = 2)
-
-pbmc_proj_umap_model2 <- ggplot() +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#000000") +
-  geom_segment(
-    data=axes, 
-    aes(x=x1, y=y1, xend=x2, yend=y2), 
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 3) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.65, 0.65)) +
-  ylim(c(-0.65, 0.65)) +
-  interior_annotation("a3", cex = 2)
 
@@
 
-# pbmc_proj_umap_model2_best <- projected_df |>
-#   ggplot(
-#     aes(
-#       x = proj1, 
-#       y = proj2)) +
-#   geom_segment(
-#     data = model_df, 
-#     aes(
-#       x = proj1_from, 
-#       y = proj2_from, 
-#       xend = proj1_to, 
-#       yend = proj2_to), 
-#     color = "#33a02c") +
-#   geom_point(
-#     #size = 0.5,
-#     alpha = 0.2) +
-#   scale_color_manual(values = c("#000000", "#33a02c")) +
-#   geom_segment(
-#     data=axes, 
-#     aes(x=x1, y=y1, xend=x2, yend=y2), 
-#     colour="grey70") +
-#   geom_text(
-#     data=axes,
-#     aes(x=x2, y=y2), 
-#     label=rownames(axes),
-#     colour="grey50",
-#     size = 3) +
-#   geom_path(
-#     data=circle, 
-#     aes(x=c1, y=c2), colour="grey70") +
-#   coord_fixed() +
-#   xlim(c(-0.9, 0.5)) +
-#   ylim(c(-0.9, 0.5)) +
-#   interior_annotation("b3", cex = 2)
-
-pbmc_proj_umap_model2_best <- ggplot() +
+pbmc_proj_umap_model2_best <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1,
+      y = proj2)) +
   geom_segment(
@@
     color = "#000000") +
+  geom_point(
+    #size = 0.5,
+    alpha = 0.2,
+    color = '#999999') +
   geom_segment(
@@
   coord_fixed() +
-  xlim(c(-0.65, 0.65)) +
-  ylim(c(-0.65, 0.65)) +
+  xlim(c(-1, 0.75)) +
+  ylim(c(-1, 0.75)) +
   interior_annotation("b3", cex = 2)

### 2024-09-25 paper.qmd
Updated the colors of PBMC3k animations

@@
 In $9\\text{-}D$, the data shows close clusters.
-This is the reason where b1, b2, and b3 shows connected edges between the clusters, and also proved by @fig-pbmc-mse as the best layout, which used the same NLDR method.
-By viewing the fitted model in $9\\text{-}D$, help to see some unobserved pattern of the data (i) densed points, and (ii) non-linear clusters."
+This is why b1, b2, and b3 shows connected edges between the clusters, as proven by @fig-pbmc-mse as a best layout.
+Additionally, in the fitted model for layout b, the two big clusters are not connected, but two are connected with the small cluster.
+Viewing the fitted model in $9\\text{-}D$, helps to see some unobserved patterns of the data: (i) dense points, and (ii) non-linear clusters.
+Videos of the langevitour animations for the models are available at <https://youtu.be/swkOKdjnjq0> and <link2> respectively."
 #| label: fig-model-pbmc-author-proj

### 2024-09-25 paper.tex
Updated the colors of PBMC3k animations

@@
 close clusters. In \(9\text{-}D\), the data shows close clusters. This
-is the reason where b1, b2, and b3 shows connected edges between the
-clusters, and also proved by Figure~\ref{fig-pbmc-mse} as the best
-layout, which used the same NLDR method. By viewing the fitted model in
-\(9\text{-}D\), help to see some unobserved pattern of the data (i)
-densed points, and (ii) non-linear clusters.}
+is why b1, b2, and b3 shows connected edges between the clusters, as
+proven by Figure~\ref{fig-pbmc-mse} as a best layout. Additionally, in
+the fitted model for layout b, the two big clusters are not connected,
+but two are connected with the small cluster. Viewing the fitted model
+in \(9\text{-}D\), helps to see some unobserved patterns of the data:
+(i) dense points, and (ii) non-linear clusters. Videos of the
+langevitour animations for the models are available at
+\url{https://youtu.be/swkOKdjnjq0} and respectively.}
 

### 2024-09-25 paper.qmd
Changed Figure 15 title

@@
 
-#benchmark_pbmc <- 0.1
+benchmark_pbmc <- 0.1
 
@@
 Layout a shows three-well separated clusters with big separation, while layout b shows close clusters.
-In $9\\text{-}D$, the data shows close clusters.
-This is why b1, b2, and b3 shows connected edges between the clusters, as proven by @fig-pbmc-mse as a best layout.
-Additionally, in the fitted model for layout b, the two big clusters are not connected, but two are connected with the small cluster.
+In $9\\text{-}D$, the data shows three clusters with two of them close and one is separated from others.
+This is why b1, b2, and b3 shows a connected edge between the close clusters, as proven by @fig-pbmc-mse as a best layout.
 Viewing the fitted model in $9\\text{-}D$, helps to see some unobserved patterns of the data: (i) dense points, and (ii) non-linear clusters.
-Videos of the langevitour animations for the models are available at <https://youtu.be/swkOKdjnjq0> and <link2> respectively."
+Videos of the langevitour animations for the models are available at <https://youtu.be/0cKX_HG_n0k> and <link2> respectively."
 #| label: fig-model-pbmc-author-proj

### 2024-09-25 paper.tex
Changed Figure 15 title

--- 2024-09-25 paper.tex
@@
 three-well separated clusters with big separation, while layout b shows
-close clusters. In \(9\text{-}D\), the data shows close clusters. This
-is the reason where b1, b2, and b3 shows connected edges between the
-clusters, and also proved by Figure~\ref{fig-pbmc-mse} as the best
-layout, which used the same NLDR method. By viewing the fitted model in
-\(9\text{-}D\), help to see some unobserved pattern of the data (i)
-densed points, and (ii) non-linear clusters.}
+close clusters. In \(9\text{-}D\), the data shows three clusters with
+two of them close and one is separated from others. This is why b1, b2,
+and b3 shows a connected edge between the close clusters, as proven by
+Figure~\ref{fig-pbmc-mse} as a best layout. Viewing the fitted model in
+\(9\text{-}D\), helps to see some unobserved patterns of the data: (i)
+dense points, and (ii) non-linear clusters. Videos of the langevitour
+animations for the models are available at
+\url{https://youtu.be/0cKX_HG_n0k} and respectively.}
 

### 2024-09-25 paper.qmd
Added animation links for Figure 15

@@
 Layout a shows three-well separated clusters with big separation, while layout b shows close clusters.
-In $9\\text{-}D$, the data shows three clusters with two of them close and one is separated from others.
+In $9\\text{-}D$, the data shows three clusters with two of them being close and one separated from the others.
 This is why b1, b2, and b3 shows a connected edge between the close clusters, as proven by @fig-pbmc-mse as a best layout.
 Viewing the fitted model in $9\\text{-}D$, helps to see some unobserved patterns of the data: (i) dense points, and (ii) non-linear clusters.
-Videos of the langevitour animations for the models are available at <https://youtu.be/0cKX_HG_n0k> and <link2> respectively."
+Videos of the langevitour animations for the models are available at <https://youtu.be/0cKX_HG_n0k> and <https://youtu.be/KhJvsRtaX04> respectively."
 #| label: fig-model-pbmc-author-proj

### 2024-09-25 paper.tex
Added animation links for Figure 15

@@
 close clusters. In \(9\text{-}D\), the data shows three clusters with
-two of them close and one is separated from others. This is why b1, b2,
-and b3 shows a connected edge between the close clusters, as proven by
-Figure~\ref{fig-pbmc-mse} as a best layout. Viewing the fitted model in
-\(9\text{-}D\), helps to see some unobserved patterns of the data: (i)
-dense points, and (ii) non-linear clusters. Videos of the langevitour
-animations for the models are available at
-\url{https://youtu.be/0cKX_HG_n0k} and respectively.}
+two of them being close and one separated from the others. This is why
+b1, b2, and b3 shows a connected edge between the close clusters, as
+proven by Figure~\ref{fig-pbmc-mse} as a best layout. Viewing the fitted
+model in \(9\text{-}D\), helps to see some unobserved patterns of the
+data: (i) dense points, and (ii) non-linear clusters. Videos of the
+langevitour animations for the models are available at
+\url{https://youtu.be/0cKX_HG_n0k} and
+\url{https://youtu.be/KhJvsRtaX04} respectively.}
 

### 2024-09-25 paper.qmd
Updated Figure 15 title with more precise description
Updated title
Updated Figure 17 colors for the consistency

@@
 
-benchmark_pbmc <- 0.1
+#benchmark_pbmc <- 0.2
 
@@
 In $9\\text{-}D$, the data shows three clusters with two of them being close and one separated from the others.
-This is why b1, b2, and b3 shows a connected edge between the close clusters, as proven by @fig-pbmc-mse as a best layout.
+This is why b1, b2, and b3 shows a connected edge between the close clusters, as proven by @fig-pbmc-mse as a reasonable layout.
 Viewing the fitted model in $9\\text{-}D$, helps to see some unobserved patterns of the data: (i) dense points, and (ii) non-linear clusters.
@@
              y = TriMAP2))+
-  geom_point(alpha=0.1, size=1, colour='#4daf4a') +
+  geom_point(alpha=0.1, size=1, colour='#999999') +
   interior_annotation("d", position = c(0.86, 0.92)) +
@@
   scale_x_continuous(breaks = sort(unique(error_mnist$a1))[append(seq(1, length(unique(error_mnist$a1)), by = 5), 24)]) +
-  scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a', "#ff7f00",'#a65628')) +
+  scale_color_manual(values=c('#e41a1c','#377eb8','#999999', "#ff7f00",'#a65628')) +
   scale_y_log10() +
@@
                  yend = y_to),
-               colour = "#33a02c",
-               linewidth = 1) +
+               colour = "#000000") +
   geom_point(data = tsne_minst_scaled_temp,
@@
              ),
-             alpha=0.1)+
-  scale_color_manual(values=c('#000000', '#e41a1c')) +
+             alpha=0.2)+
+  scale_color_manual(values=c('#ff7f00', '#4daf4a')) +
   interior_annotation("a", sc_ltr_pos_mnist) +
@@
       yend = proj2_to), 
-    color = "#33a02c",
-    linewidth = 1) +
+    color = "#000000") +
   geom_point(
     #size = 0.2,
-    alpha = 0.2) +
-  scale_color_manual(values = c("#000000","#e41a1c")) +
+    alpha = 0.3) +
+  scale_color_manual(values = c('#ff7f00', '#4daf4a')) +
   geom_segment(
@@
       yend = proj2_to), 
-    color = "#33a02c",
-    linewidth = 1) +
+    color = "#000000") +
   geom_point(
     #size = 0.2,
-    alpha = 0.2) +
-  scale_color_manual(values = c("#000000", "#e41a1c")) +
+    alpha = 0.3) +
+  scale_color_manual(values = c('#ff7f00', '#4daf4a')) +
   geom_segment(

### 2024-09-25 scripts/mnist/4_gen_model_in_data_space.R
Try to find more aacurrate model

@@ -54,6 +54,8 @@ benchmark_mnist <- find_lg_benchmark(
   distance_edges = distance_mnist,
   distance_col = "distance")
 
+benchmark_mnist <- 0.1
+
 ## Hexagonal binning to have regular hexagons
 hb_obj_mnist <- hex_binning(
   data = tsne_minst_scaled,

### 2024-09-25 paper.qmd
Try to find more aacurrate model
Added Figure 18 title

@@
 
+benchmark_mnist <- 0.1
+
 sc_ltr_pos_mnist <- c(0.96, 0.96)
@@
 #| label: fig-model-error-mnist
-#| fig-cap: "(a) The $10\\text{-}D$ model error for tSNE layout in $\\gD{}$ of the MNIST digit 1 dataset.
-Most low model errors are distributed along the big non-linear clutser, while most large model errors are concentrated along the small cluster.
-(b) There are certain images that exhibit large errors due to their deviation from the non-linear cluster, which makes them anomalies.
+#| fig-cap: "The $10\\text{-}D$ model error in layout a of the MNIST digit 1 dataset show a pattern.
+Most low model errors are distributed along the big non-linear clutser, while most large model errors are distributed along the small cluster.
+Along the non-linear cluster, the angle of digit 1 changes.
+Some images have large errors due to their deviation from the non-linear cluster, which makes them anomalies.
 The images associated with large model errors shows different patterns of digit 1."

### 2024-09-25 paper.tex
Updated Figure 15 title with more precise description
Added Figure 18 title

@@
 b1, b2, and b3 shows a connected edge between the close clusters, as
-proven by Figure~\ref{fig-pbmc-mse} as a best layout. Viewing the fitted
-model in \(9\text{-}D\), helps to see some unobserved patterns of the
-data: (i) dense points, and (ii) non-linear clusters. Videos of the
+proven by Figure~\ref{fig-pbmc-mse} as a reasonable layout. Viewing the
+fitted model in \(9\text{-}D\), helps to see some unobserved patterns of
+the data: (i) dense points, and (ii) non-linear clusters. Videos of the
 langevitour animations for the models are available at
@@
 
-\caption{\label{fig-model-error-mnist}(a) The \(10\text{-}D\) model
-error for tSNE layout in \(\gD{}\) of the MNIST digit 1 dataset. Most
-low model errors are distributed along the big non-linear clutser, while
-most large model errors are concentrated along the small cluster. (b)
-There are certain images that exhibit large errors due to their
-deviation from the non-linear cluster, which makes them anomalies. The
-images associated with large model errors shows different patterns of
-digit 1.}
+\caption{\label{fig-model-error-mnist}The \(10\text{-}D\) model error in
+layout a of the MNIST digit 1 dataset show a pattern. Most low model
+errors are distributed along the big non-linear clutser, while most
+large model errors are distributed along the small cluster. Along the
+non-linear cluster, the angle of digit 1 changes. Some images have large
+errors due to their deviation from the non-linear cluster, which makes
+them anomalies. The images associated with large model errors shows
+different patterns of digit 1.}
 

### 2024-09-25 paper.qmd
Figure 17 initial figure title

@@
 #| fig-width: 10
-#| fig-cap: "Model in \\gD{}, on the tSNE layout (a), and three views (b, c, d) of the fit in projections from $10\\text{-}D$, for the digit 1 of MNIST data ($a_1 = 0.07$, $b = 456 \\  (19, \ 24)$, $m = 226$, and benchmark value to remove large edges is $0.114$).
-(b) The non-linear structure observed in the \\gD{} layout of tSNE (@fig-tsne-best) is also visible when visualizing the model overlaid on the data space.
-This indicates that tSNE accurately captures the non-linear structure of the $10\\text{-}D$ data.
-(c) the model shows a twisted pattern within the non-linear structure in $10\\text{-}D$ space, which is an additional pattern not visible in the \\gD{} representation (@fig-tsne-best).
-(d) some long edges exist in the $10\\text{-}D$ space that are not recognized as long edges in the \\gD{} representation.
-The full video is available at <https://youtu.be/zcg_GXBmqjA>."
+#| fig-cap: "Model in \\gD{}, on the layout a and three views of the fit in projections from $10\\text{-}D$, for the MNIST digit 1 data ($n =  7877$ and $p = 10$)."
 

### 2024-09-25 paper.tex
Figure 17 initial figure title

@@
 
-\caption{\label{fig-mnist-tri-proj}Model in \gD{}, on the tSNE layout
-(a), and three views (b, c, d) of the fit in projections from
-\(10\text{-}D\), for the digit 1 of MNIST data (\(a_1 = 0.07\),
-\(b = 456 \  (19,  24)\), \(m = 226\), and benchmark value to remove
-large edges is \(0.114\)). (b) The non-linear structure observed in the
-\gD{} layout of tSNE (\textbf{?@fig-tsne-best}) is also visible when
-visualizing the model overlaid on the data space. This indicates that
-tSNE accurately captures the non-linear structure of the \(10\text{-}D\)
-data. (c) the model shows a twisted pattern within the non-linear
-structure in \(10\text{-}D\) space, which is an additional pattern not
-visible in the \gD{} representation (\textbf{?@fig-tsne-best}). (d) some
-long edges exist in the \(10\text{-}D\) space that are not recognized as
-long edges in the \gD{} representation. The full video is available at
-\url{https://youtu.be/zcg_GXBmqjA}.}
+\caption{\label{fig-mnist-tri-proj}Model in \gD{}, on the layout a and
+three views of the fit in projections from \(10\text{-}D\), for the
+MNIST digit 1 data (\(n =  7877\) and \(p = 10\)).}
 

### 2024-09-25 paper.qmd
Added title for Figure 17

@@
 Viewing the fitted model in $9\\text{-}D$, helps to see some unobserved patterns of the data: (i) dense points, and (ii) non-linear clusters.
-Videos of the langevitour animations for the models are available at <https://youtu.be/0cKX_HG_n0k> and <https://youtu.be/KhJvsRtaX04> respectively."
+Videos of the langevitour animations are available at <https://youtu.be/0cKX_HG_n0k> and <https://youtu.be/KhJvsRtaX04> respectively."
 #| label: fig-model-pbmc-author-proj
@@
 #| fig-width: 10
-#| fig-cap: "Model in \\gD{}, on the layout a and three views of the fit in projections from $10\\text{-}D$, for the MNIST digit 1 data ($n =  7877$ and $p = 10$)."
+#| fig-cap: "Model in \\gD{}, on the layout a and three views of the fit in projections from $10\\text{-}D$, for the MNIST digit 1 data ($n =  7877$ and $p = 10$).
+There is a big non-linear cluster and a small clusterlocated very close to one corner of the big cluster.
+There is a twisted pattern that is hardly visible in the static plot.
+Video of the langevitour animation are available at <need to add the link>."
 

### 2024-09-25 paper.tex
Added title for Figure 17

@@
 the data: (i) dense points, and (ii) non-linear clusters. Videos of the
-langevitour animations for the models are available at
+langevitour animations are available at
 \url{https://youtu.be/0cKX_HG_n0k} and
@@
 three views of the fit in projections from \(10\text{-}D\), for the
-MNIST digit 1 data (\(n =  7877\) and \(p = 10\)).}
+MNIST digit 1 data (\(n =  7877\) and \(p = 10\)). There is a big
+non-linear cluster and a small clusterlocated very close to one corner
+of the big cluster. There is a twisted pattern that is hardly visible in
+the static plot. Video of the langevitour animation are available at .}
 

### 2024-09-25 paper.qmd
Updated Figure 13 NLDR layouts

@@
   )  +
-  coord_fixed() +
   interior_annotation("a1", 
                       position = c(0.08, 0.95),
-                      cex = 2)
+                      cex = 2) +
+  theme(
+    aspect.ratio = 1
+  )
 
@@
   )  +
-  coord_fixed() +
   interior_annotation("b1", 
                       position = c(0.06, 0.95),
-                      cex = 2)
+                      cex = 2) +
+  theme(
+    aspect.ratio = 1
+  )
 
@@
   )  +
-  coord_fixed() +
   interior_annotation("c1", 
                       position = c(0.08, 0.95),
-                      cex = 2)
+                      cex = 2) +
+  theme(
+    aspect.ratio = 1
+  )
 

### 2024-09-25 paper.qmd
Updated Figure 13 with layout a

@@
     ),
-    alpha = 0.3,
-    size = 0.1,
+    alpha = 0.2,
+    #size = 0.1,
     color = "#000000"
@@
 
-projected <- as.matrix(scaled_gau_data) %*% projection
+projection_scaled <- projection * 1
+
+projected <- as.matrix(scaled_gau_data) %*% projection_scaled
 
@@
 
-projected_model <- as.matrix(scaled_gau_data_model) %*% projection
+projected_model <- as.matrix(scaled_gau_data_model) %*% projection_scaled
 
@@
   proj = projection,
-  limits = 0.2,
-  axis_pos_x = 0.15,
+  limits = 0.3,
+  axis_pos_x = 0.05,
   axis_pos_y = -0.85,
@@
       yend = proj2_to), 
-    color = "#33a02c") +
+    color = "#33a02c",
+    linewidth = 1) +
   geom_point(
-    size = 0.5,
-    alpha = 0.3) +
+    #size = 0.5,
+    alpha = 0.2) +
   scale_color_manual(values = c("#000000", "#33a02c")) +
@@
     colour="grey50",
-    size = 3) +
+    size = 5) +
   geom_path(
@@
   coord_fixed() +
-  xlim(c(0.1, 0.38)) +
+  xlim(c(0, 0.4)) +
   ylim(c(-0.9, -0.5)) +
@@
   coord_fixed() +
-  xlim(c(-0.75, -0.35)) +
-  ylim(c(-0.75, -0.4)) +
+  xlim(c(-0.75, -0.36)) +
+  ylim(c(-0.75, -0.36)) +
   interior_annotation("a3", 

### 2024-09-26 paper.qmd
Updated Figure 13 layout b

@@
     ),
-    alpha = 0.3,
-    size = 0.1,
+    alpha = 0.05,
+    #size = 0.1,
     color = "#000000"
@@
   proj = projection,
-  limits = 0.2,
-  axis_pos_x = -0.55,
+  limits = 0.3,
+  axis_pos_x = -0.58,
   axis_pos_y = -0.05,
@@
       yend = proj2_to), 
-    color = "#33a02c") +
+    color = "#33a02c",
+    linewidth = 1) +
   geom_point(
-    size = 0.5,
+    #size = 0.5,
     alpha = 0.3) +
@@
   coord_fixed() +
-  xlim(c(-0.6, -0.3)) +
+  xlim(c(-0.64, -0.26)) +
   ylim(c(0, -0.38)) +
@@
   coord_fixed() +
-  xlim(c(-0.75, -0.35)) +
+  xlim(c(-0.75, -0.4)) +
   ylim(c(-0.75, -0.4)) +

### 2024-09-26 paper.qmd
Updated Figure 13 with layout c

@@
     alpha = 0.3,
-    size = 0.1,
+    #size = 0.1,
     color = "#000000"
@@
   coord_fixed() +
-  xlim(c(-0.41, -0.15)) +
+  xlim(c(-0.46, -0.12)) +
   ylim(c(0.18, 0.52)) +
@@
       yend = proj2_to), 
-    color = "#33a02c") +
+    color = "#33a02c",
+    linewidth = 1) +
   geom_point(
-    size = 0.8,
+    #size = 0.8,
     alpha = 0.15) +
@@
   xlim(c(-0.75, -0.35)) +
-  ylim(c(-0.75, -0.4))+
+  ylim(c(-0.75, -0.35))+
   interior_annotation("c3", 

### 2024-09-26 paper.qmd
Figure 13 resize axes circles and repositioned some

@@
   proj = projection,
-  limits = 0.3,
+  limits = 0.28,
   axis_pos_x = -0.58,
@@
   proj = projection,
-  limits = 0.3,
+  limits = 0.28,
   axis_pos_x = -0.7,
@@
   proj = projection,
-  limits = 0.2,
-  axis_pos_x = -0.37,
-  axis_pos_y = 0.22,
+  limits = 0.25,
+  axis_pos_x = -0.41,
+  axis_pos_y = 0.23,
   axis_labels = names(scaled_gau_data),

### 2024-09-26 paper.qmd
Figure 13 changed colors to specify the message clearly

@@
     #size = 0.1,
-    color = "#000000"
+    color = "#999999"
   )  +
@@
       yend = proj2_to), 
-    color = "#33a02c",
+    color = "#000000",
     linewidth = 1) +
@@
     #size = 0.5,
-    alpha = 0.2) +
-  scale_color_manual(values = c("#000000", "#33a02c")) +
+    alpha = 0.2,
+    color = "#999999") +
   geom_segment(
@@
     #size = 0.5,
-    alpha = 0.3) +
-  scale_color_manual(values = c("#000000", "#33a02c")) +
+    alpha = 0.3,
+    color = "#999999") +
   geom_segment(
@@
     #size = 0.8,
-    alpha = 0.15) +
-  scale_color_manual(values = c("#000000", "#33a02c")) +
+    alpha = 0.2,
+    color = "#999999") +
   geom_segment(

### 2024-09-26 paper.qmd
Fixed the issue with PaCMAP model in Figure 13

@@
   data = pacmap_data_gau)
-pacmap_gau_scaled <- gau1_scaled_obj$scaled_nldr
+pacmap_gau_scaled <- gau1_scaled_obj$scaled_nldr |>
+  select(PaCMAP1, PaCMAP2, ID)
 

### 2024-09-26 paper.qmd
Figure 13 removed long edge in PaCMAP
Figure 13, updated after removing long edges manually for UMAP
Updated Figure 13 tSNE layout by manually removing long edges

@@
 
-tr_from_to_df_gau1 <- tr_from_to_df_gau1 |>
-  filter(row_number() != 149)
+# tr_from_to_df_gau1 <- tr_from_to_df_gau1 |>
+#   filter(row_number() != 149)
 
@@
 distance_df_small_edges_gau1 <- distance_gau1 |>
-  filter(distance < benchmark_gau1)
+  filter(distance < benchmark_gau1) |>
+  filter(!(row_number() %in% c(33, 40, 48, 138, 143)))
 
@@
 
-distance_df_small_edges <- distance_df_small_edges |>
-  filter(row_number() != 108)
+# distance_df_small_edges <- distance_df_small_edges |>
+#   filter(row_number() != 108)
 
@@
 distance_df_small_edges_gau1 <- distance_gau1 |>
-  filter(distance < benchmark_gau1)
+  filter(distance < benchmark_gau1) |>
+  mutate(ID = row_number()) |>
+  filter(!(ID %in% c(112)))
 
@@
 distance_df_small_edges_gau1 <- distance_gau1 |>
-  filter(distance < benchmark_gau1)
+  filter(distance < benchmark_gau1) |>
+  mutate(ID = row_number()) |>
+  filter(!(ID %in% c(97)))
 

### 2024-09-26 paper.qmd
Added box around the selected cluster for layout a Figure13

@@
   ) + 
+  geom_rect(
+      aes(xmin = 0.3, 
+          xmax = 0.65, 
+          ymin = -0.05, 
+          ymax = 0.35),
+          fill = "transparent", 
+      color = "red", 
+      linewidth = 0.5) +
   interior_annotation("a1", 

### 2024-09-26 scripts/rm_lg_manually.R

@@ -1,0 +1,16 @@
+d <- tr_from_to_df_gau1 |>
+  mutate(ID = row_number())
+
+a <- ggplot(data = d,
+            aes(
+              x = x_from,
+              y = y_from,
+              xend = x_to,
+              yend = y_to,
+              group = ID)) +
+  geom_segment(colour = "#000000") +
+  geom_text(aes(label = from),
+            size = 2)
+
+ggplotly(a)
+

### 2024-09-26 paper.qmd
Figure 13 added rectangles for selected clusters

@@
   ) + 
+  geom_rect(
+    aes(xmin = 0.45, 
+        xmax = 0.57, 
+        ymin = 0, 
+        ymax = 0.1),
+    fill = "transparent", 
+    color = "red", 
+    linewidth = 0.5)  +
   interior_annotation("b1", 
@@
   ) +
+  geom_rect(
+    aes(xmin = -0.04, 
+        xmax = 0.22, 
+        ymin = 0.08, 
+        ymax = 0.34),
+    fill = "transparent", 
+    color = "red", 
+    linewidth = 0.5) +
   interior_annotation("c1", 

### 2024-09-26 paper.qmd
Updated Figure 13 title (Need to record videos)

@@
 #| fig-pos: H
-#| fig-cap: "Model in \\gD{}, on the layouts from tSNE, UMAP, and PaCMAP (a1, b1, c1).
-Two views of the fit in projections of each layout from $4\\text{-}D$, for the Gaussian cluster data.
+#| fig-cap: "Model in \\gD{}, on the layouts a1, b1, and c1, and two views of the fit in projections of each layout from $4\\text{-}D$, for the five Gaussian cluster data.
 All \\gD{} models show five well-separated clusters with different gaps.
@@
 On the other hand, layout b3 shows average gaps with exact positions as in $4\\text{-}D$.
-Assessing the different models fit for different layouts for a specific cluster shows filled-out shape in a1, a2, and b1, b2.
+Assessing the different models fit for different layouts for a specific cluster (highlighted in red) shows filled-out shape in a1, a2, and b1, b2.
 On the other hand, c1 and c2 shows pancake shape.
-The full video for the model for layout a, b, and c are available at <https://youtu.be/5Y1hE4i7N2k>, <https://youtu.be/5Y1hE4i7N2k>, and <https://youtu.be/5Y1hE4i7N2k> respectively."
+Videos of the langevitour animations are available at <https://youtu.be/5Y1hE4i7N2k>, <https://youtu.be/5Y1hE4i7N2k>, and <https://youtu.be/5Y1hE4i7N2k> respectively."
 

### 2024-09-26 paper.tex
Updated Figure 13 title (Need to record videos)

@@
 
-\caption{\label{fig-five-gau-projs}Model in \gD{}, on the layouts from
-tSNE, UMAP, and PaCMAP (a1, b1, c1). Two views of the fit in projections
-of each layout from \(4\text{-}D\), for the Gaussian cluster data. All
-\gD{} models show five well-separated clusters with different gaps.
-Layout a1 shows small separations, while layout b2 shows large gaps. On
-the other hand, layout b3 shows average gaps with exact positions as in
-\(4\text{-}D\). Assessing the different models fit for different layouts
-for a specific cluster shows filled-out shape in a1, a2, and b1, b2. On
-the other hand, c1 and c2 shows pancake shape. The full video for the
-model for layout a, b, and c are available at
+\caption{\label{fig-five-gau-projs}Model in \gD{}, on the layouts a1,
+b1, and c1, and two views of the fit in projections of each layout from
+\(4\text{-}D\), for the five Gaussian cluster data. All \gD{} models
+show five well-separated clusters with different gaps. Layout a1 shows
+small separations, while layout b2 shows large gaps. On the other hand,
+layout b3 shows average gaps with exact positions as in \(4\text{-}D\).
+Assessing the different models fit for different layouts for a specific
+cluster (highlighted in red) shows filled-out shape in a1, a2, and b1,
+b2. On the other hand, c1 and c2 shows pancake shape. Videos of the
+langevitour animations are available at
 \url{https://youtu.be/5Y1hE4i7N2k}, \url{https://youtu.be/5Y1hE4i7N2k},

### 2024-09-26 paper.qmd
Removed long edges manually for PBMC3k data
Changed colors og Figure 17

@@
 distance_df_small_edges_pbmc <- distance_pbmc |>
-  filter(distance < benchmark_pbmc)
+  filter(distance < benchmark_pbmc) |>
+  filter(!(row_number() %in% c(6, 12, 14, 16, 17, 19, 23, 29, 30, 143, 144, 146, 147, 157)))
 
@@
 distance_df_small_edges_mnist <- distance_mnist |>
-  filter(distance < benchmark_mnist)
+  filter(distance < benchmark_mnist) |>
+  filter(!(row_number() %in% c(143, 144, 162)))
 
@@
              alpha=0.2)+
-  scale_color_manual(values=c('#ff7f00', '#4daf4a')) +
+  scale_color_manual(values=c("#999999",'#ff7f00')) +
   interior_annotation("a", sc_ltr_pos_mnist) +
@@
     alpha = 0.3) +
-  scale_color_manual(values = c('#ff7f00', '#4daf4a')) +
+  scale_color_manual(values = c("#999999",'#ff7f00')) +
   geom_segment(

### 2024-09-26 paper.qmd
Updated data range for the small cluster for MNIST

@@
 tsne_minst_scaled_temp <- tsne_minst_scaled |>
-  mutate(cluster = if_else((tSNE1 <= 0.92) & (tSNE1 >= 0.75) & (tSNE2 <= 0.62) & (tSNE2 >= 0.44), "small_clust", "big_clust"))
+  mutate(cluster = if_else((tSNE1 <= 0.92) & (tSNE1 >= 0.76) & (tSNE2 <= 0.62) & (tSNE2 >= 0.44), "small_clust", "big_clust"))
 

### 2024-09-26 paper.qmd
Figure 18 changed the colors

@@
   geom_point(alpha=0.2) +
-  scale_colour_continuous_sequential(palette = "Light Grays") +
+  scale_colour_continuous_sequential(palette = "Blues") +
   geom_rect(aes(xmin = 0.8, xmax = 0.9, ymin = 0.45, ymax = 0.6),
+               fill = "transparent", color = "#000000", linewidth = 0.7) +
+  geom_rect(aes(xmin = 0.75, xmax = 0.85, ymin = 0.8, ymax = 0.95),
                fill = "transparent", color = "#e41a1c", linewidth = 0.7)  +
-  geom_rect(aes(xmin = 0.75, xmax = 0.85, ymin = 0.8, ymax = 0.95),
-               fill = "transparent", color = "#377eb8", linewidth = 0.7)  +
   geom_rect(aes(xmin = 0.45, xmax = 0.55, ymin = 0.45, ymax = 0.6),
@@
   coord_fixed() +
-  scale_fill_continuous_sequential(palette = "Blues 3") +
+  scale_fill_continuous_sequential(palette = "Reds 2") +
   theme(strip.background = element_blank(),
@@
   coord_fixed() +
-  scale_fill_continuous_sequential(palette = "Reds 2") +
+  scale_fill_continuous_sequential(palette = "Grays") +
   theme(strip.background = element_blank(),

### 2024-09-26 paper.qmd
Updated Figure 18

@@
              y = tSNE2))+
-  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
+  geom_point(alpha=0.1, size=1, colour='#999999') +
   interior_annotation("a", c(0.08, 0.93)) +
@@
              y = TriMAP2))+
-  geom_point(alpha=0.1, size=1, colour='#999999') +
+  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
   interior_annotation("d", position = c(0.86, 0.92)) +
@@
   scale_x_continuous(breaks = sort(unique(error_mnist$a1))[append(seq(1, length(unique(error_mnist$a1)), by = 5), 24)]) +
-  scale_color_manual(values=c('#e41a1c','#377eb8','#999999', "#ff7f00",'#a65628')) +
+  scale_color_manual(values=c('#e41a1c','#377eb8',"#ff7f00", '#999999','#a65628')) +
   scale_y_log10() +

### 2024-09-26 paper.qmd
Updated Figure 17

@@
   filter(distance < benchmark_mnist) |>
-  filter(!(row_number() %in% c(143, 144, 162)))
+  filter(!(row_number() %in% c(155, 554, 140)))
 
@@
   tr_from_to_df_mnist, distance_df_small_edges_mnist, 
-  by = c("from", "to"))
+  by = c("from", "to")) |>
+  mutate(ID = row_number())
 
@@
   plot_layout(ncol = 2)
+
 ```

### 2024-09-26 paper.qmd
Added initial work to replace the S-curve structure with another clustering structure

@@
 # and plot them on same aspect ratio
-tsne_two_curvilinear_clust <- read_rds("data/two_nonlinear_clusters/two_nonlinear_clusters_tsne_perplexity_30.rds")
+tsne_two_curvilinear_clust <- read_rds("data/two_curvy_clust/two_curvy_clust_tsne_perplexity_30.rds")
 
@@
 
-umap_two_curvilinear_clust <- read_rds("data/two_nonlinear_clusters/two_nonlinear_clusters_umap_n-neigbors_15_min-dist_0.1.rds")
+umap_two_curvilinear_clust <- read_rds("data/two_curvy_clust/two_curvy_clust_umap_n-neigbors_15_min-dist_0.1.rds")
 
@@
 
-phate_two_curvilinear_clust <- read_rds("data/two_nonlinear_clusters/two_nonlinear_clusters_phate_knn_5.rds")
+phate_two_curvilinear_clust <- read_rds("data/two_curvy_clust/two_curvy_clust_phate_knn_5.rds")
 
@@
 
-trimap_two_curvilinear_clust <- read_rds("data/two_nonlinear_clusters/two_nonlinear_clusters_trimap_n-inliers_12_n-outliers_4_n-random_3.rds")
+trimap_two_curvilinear_clust <- read_rds("data/two_curvy_clust/two_curvy_clust_trimap_n-inliers_12_n-outliers_4_n-random_3.rds")
 
@@
 
-pacmap_two_curvilinear_clust <- read_rds("data/two_nonlinear_clusters/two_nonlinear_clusters_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")
+pacmap_two_curvilinear_clust <- read_rds("data/two_curvy_clust/two_curvy_clust_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")
 
@@
 
-error_two_curvilinear_clust_umap <- read_rds("data/two_nonlinear_clusters/error_two_nonlinear_clusters_umap.rds")
-error_two_curvilinear_clust_tsne <- read_rds("data/two_nonlinear_clusters/error_two_nonlinear_clusters_tsne.rds")
-error_two_curvilinear_clust_phate <- read_rds("data/two_nonlinear_clusters/error_two_nonlinear_clusters_phate.rds")
-error_two_curvilinear_clust_trimap <- read_rds("data/two_nonlinear_clusters/error_two_nonlinear_clusters_trimap.rds")
-error_two_curvilinear_clust_pacmap <- read_rds("data/two_nonlinear_clusters/error_two_nonlinear_clusters_pacmap.rds")
+error_two_curvilinear_clust_umap <- read_rds("data/two_curvy_clust/error_two_curvy_clust_umap.rds")
+error_two_curvilinear_clust_tsne <- read_rds("data/two_curvy_clust/error_two_curvy_clust_tsne.rds")
+error_two_curvilinear_clust_phate <- read_rds("data/two_curvy_clust/error_two_curvy_clust_phate.rds")
+error_two_curvilinear_clust_trimap <- read_rds("data/two_curvy_clust/error_two_curvy_clust_trimap.rds")
+error_two_curvilinear_clust_pacmap <- read_rds("data/two_curvy_clust/error_two_curvy_clust_pacmap.rds")
 

### 2024-09-26 scripts/two_curvilinear_clusters/02_gen_true_model.R

@@ -1,82 +1,96 @@
+library(langevitour)
+library(rsample)
+library(scales)
 library(dplyr)
-library(tidyr)
+library(readr)
 
-# Function to generate a curvilinear grid pattern in 2D
-generate_curvilinear_grid_2d <- function(n_grid_x, n_grid_y) {
-  x <- seq(0, 2, length.out = n_grid_x)
-  y <- seq(-3, 0.5, length.out = n_grid_y)
+set.seed(20240110)
 
-  # Use expand.grid to create a grid of (x, y) pairs
-  curvilinear_grid <- expand.grid(x = x, y = y)
+# Number of points along the S-curve
+n_samples <- 50
 
-  # Adjust y values according to a curvilinear pattern based on x
-  curvilinear_grid$y <- -(curvilinear_grid$x^3 + curvilinear_grid$y)
+# Generate uniform parameter t
+a <- runif(n = n_samples, min = -0.5, max = 0)
 
-  return(curvilinear_grid)
+# Parametric equations for the S-curve in 3D
+x1 <- sin(3 * pi * a)
+x3 <- sign(a) * (cos(3 * pi * a) - 1)
+
+# Set band thickness for the S-curve in the third dimension (y)
+band_thickness <- 2.0
+num_y_points <- 5  # Number of points along the y direction for each (x, z) pair
+
+# Create a data frame for storing points
+true_model <- data.frame()
+
+# Loop through each point on the S-curve and generate thickness without jitter
+for (i in 1:n_samples) {
+  # Current (x, z) values
+  current_x <- x1[i]
+  current_z <- x3[i]
+
+  # Generate y values with a thickness band
+  y_values <- seq(0, band_thickness, length.out = num_y_points)
+
+  # Create points for the current (x, z) with different y values
+  new_points <- data.frame(x1 = rep(current_x, num_y_points),
+                           x2 = y_values,
+                           x3 = rep(current_z, num_y_points),
+                           ID = seq((i-1)*num_y_points + 1, i*num_y_points))
+
+  # Combine with existing points
+  true_model <- rbind(true_model, new_points)
 }
 
-# Number of grid points in each dimension
-n_grid_x <- 10  # Adjust this number for a finer or coarser grid along x
-n_grid_y <- 10  # Adjust this number for a finer or coarser grid along y
-
-# Generate the curvilinear grid of points in 2D
[…]

### 2024-09-26 scripts/two_curvilinear_clusters/04_gen_embeddings.R

@@ -20,12 +20,12 @@ use_condaenv("pcamp_env")
 
 #source(here::here("R/nldr_code.R"), local = TRUE)
 
-data <- read_rds(here::here("data/two_nonlinear_clusters/two_nonlinear_clusters_data.rds"))
+data <- read_rds(here::here("data/two_curvy_clust/two_curvy_clust_data.rds"))
 
 ## Scale the data
 
-# data <- data |>
-#   mutate(across(everything(), ~ (. - mean(.)) / sd(.)))
+data <- data |>
+  mutate(across(everything(), ~ (. - mean(.)) / sd(.)))
 
 ## tSNE
 perplexity <- 30
@@ -39,7 +39,7 @@ tSNE_data <- tSNE_fit$Y |>
   tibble::as_tibble(.name_repair = "unique")
 names(tSNE_data) <- c("tSNE1", "tSNE2")
 
-write_rds(tSNE_data, file = paste0("data/two_nonlinear_clusters/two_nonlinear_clusters_tsne_perplexity_", perplexity, ".rds"))
+write_rds(tSNE_data, file = paste0("data/two_curvy_clust/two_curvy_clust_tsne_perplexity_", perplexity, ".rds"))
 
 ## UMAP
 
@@ -58,9 +58,13 @@ min_dist <- 0.1
 # UMAP_data <- UMAP_fit$layout |>
 #   tibble::as_tibble(.name_repair = "unique")
 
-UMAP_fit <- umap(data,
-                 n_neighbors = n_neighbors,
-                 n_components =  min_dist)
+# Create a config list with the desired parameters
+umap_config <- umap.defaults
+umap_config$n_neighbors <- n_neighbors      # Set the number of neighbors
+umap_config$n_components <- 2    # Set the number of output dimensions (typically 2 or 3)
+umap_config$min_dist <- min_dist
+
+UMAP_fit <- umap(data, config = umap_config)
 
 UMAP_data <- UMAP_fit$layout |>
   as_tibble()
@@ -68,15 +72,15 @@ UMAP_data <- UMAP_fit$layout |>
 names(UMAP_data) <- c("UMAP1", "UMAP2")
 
 ## Run only once
[…]

### 2024-09-26 paper.qmd
Added removed S-curve codes to another script

@@
 ```{r}
-#| label: scurve-training
-training_data_scurve <- read_rds("data/s_curve/s_curve_training.rds")
-```
-
-<!--UMAP applied for Scurve data-->
-```{r}
-#| label: umap-scurve
-umap_scurve <- read_rds(file = "data/s_curve/s_curve_umap.rds")
-
-scurve_scaled_obj <- gen_scaled_data(
-  data = umap_scurve)
-
-umap_scurve_scaled <- scurve_scaled_obj$scaled_nldr
-lim1 <- scurve_scaled_obj$lim1
-lim2 <- scurve_scaled_obj$lim2
+#| label: two-curvy-training
+training_data_two_curvy <- read_rds("data/two_curvy_clust/two_curvy_clust_data.rds")
+```
+
+<!--UMAP applied for Two curvy data-->
+```{r}
+#| label: umap-two-curvy
+umap_two_curvy <- read_rds(file = "data/two_curvy_clust/two_curvy_clust_umap_n-neigbors_15_min-dist_0.1.rds")
+
+two_curvy_scaled_obj <- gen_scaled_data(
+  data = umap_two_curvy)
+
+umap_two_curvy_scaled <- two_curvy_scaled_obj$scaled_nldr
+lim1 <- two_curvy_scaled_obj$lim1
+lim2 <- two_curvy_scaled_obj$lim2
 r2 <- diff(lim2)/diff(lim1)

### 2024-09-26 paper.tex
Start to replace the methodology with a new example

@@
 resulting surface shown in \pD{}. The steps in this process are shown in
-Figure~\ref{fig-NLDR-scurve}, and documented below.
+Figure~\ref{fig-NLDR-two-curvy}, and documented below.
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-scurve-true-sc-1.pdf}
-
-}
-
-\caption{\label{fig-scurve-true-sc}Two views of the true model (blue
-lines) in \(2\text{-}D\) projections from \(7\text{-}D\), for the
-S-curve data (black points). The data is spread along the S shape, and
-does not vary much from this curve. (The \textbf{langevitour} software
-is used to view the data with a tour, and the full video is available at
-\url{https://youtu.be/I5GL23vLiw0}).}
-
-\end{figure}%
-
-\begin{figure}
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-NLDR-scurve-1.pdf}
-
-}
-
-\caption{\label{fig-NLDR-scurve}Key steps for constructing the model on
-the UMAP layout (\(k=2\)): (a) data, (b) hexagon bins, (c) bin
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-NLDR-two-curvy-1.pdf}
+
+}
+
+\caption{\label{fig-NLDR-two-curvy}Key steps for constructing the model
+on the UMAP layout (\(k=2\)): (a) data, (b) hexagon bins, (c) bin
 centroids, and (d) triangulated centroids. The S-curve data is shown.}
@@
 \(y_{1,\text{max}}=1\) and \(y_{2,\text{max}} = \frac{r_2}{r_1}\)
-(Figure~\ref{fig-NLDR-scurve}). If the NLDR aspect ratio is ignored then
-set \(y_ {2,\text{max}} = 1\).
+(Figure~\ref{fig-NLDR-two-curvy}). If the NLDR aspect ratio is ignored
+then set \(y_ {2,\text{max}} = 1\).
 
@@
 By default, the bin centroid is used for describing a hexagon (as done
-in Figure~\ref{fig-NLDR-scurve} (c)), but any measure of center, such as
-a mean or weighted mean of the points within each hexagon, could be
+in Figure~\ref{fig-NLDR-two-curvy} (c)), but any measure of center, such
+as a mean or weighted mean of the points within each hexagon, could be
 used. The bin centers, and the binned data, are the two important
@@
 points so that edges indicate neighbouring observations, in both the
-NLDR layout (Figure~\ref{fig-NLDR-scurve} (d)) and the \pD{} model
+NLDR layout (Figure~\ref{fig-NLDR-two-curvy} (d)) and the \pD{} model
 representation. When the data has been binned the triangulation connectd
@@
 
-\begin{figure}[H]
-
-\begin{minipage}{0.25\linewidth}
-\includegraphics{figures/scurve/umap_trimesh_layout.png}\end{minipage}%
-%
-\begin{minipage}{0.25\linewidth}
-\includegraphics{figures/scurve/sc_umap_best_1.png}\end{minipage}%
-%
-\begin{minipage}{0.25\linewidth}
-\includegraphics{figures/scurve/sc_umap_best_2.png}\end{minipage}%
-%
-\begin{minipage}{0.25\linewidth}
-\includegraphics{figures/scurve/sc_umap_best_3.png}\end{minipage}%
-
-\caption{\label{fig-scurve-sc}Model in \gD{}, on the UMAP layout, and
-three views of the fit in projections from \(7\text{-}D\), for the
-S-curve data (\((s_1, \ s_2) = (-0.160, \ -0.263)\),
-\(b = 405 \  (15, \ 27)\), \(m = 70\), benchmark value to remove low
-density hexagons is \(0.250\), and benchmark value to remove large edges
-is \(0.159\)). MSE is \(0.0462\). The model closely fits the shape, but
-it doesn't fully fill out the width of the S, which means that it does
-not adequately capture the surface (The \textbf{langevitour} software is
-used to view the data with a tour, and the full video is available at
-\url{https://youtu.be/i7F7xpN1Hz8}).}
-
-\end{figure}%
-
 \subsection{Measuring the fit}\label{sec-summary}
@@
 dimensional data of \(i^{th}\) observation in \(h^{th}\) hexagon.
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=0.3\textheight]{paper_files/figure-pdf/fig-p-d-error-in-2d-scurve-1.pdf}
-
-}
-
-\caption{\label{fig-p-d-error-in-2d-scurve}error}
-
-\end{figure}%
 
@@
 \(C_{h}^{(2)} = (c_{h1}, c_{h2})\).
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-predict-scurve-1.pdf}
-
-}
-
-\caption{\label{fig-predict-scurve}Comparison of prediction generated
-using the exiting \texttt{umap} prediction method and our method: (a) A
-view of the true model in projections from \(7\text{-}D\), (b) predicted
-data from the \texttt{umap} prediction method, (c) predicted data from
-our method, and (d) comparison of predictions from both methods. In plot
-(d), points representing the same data but predicted by different
-methods are connected by lines. Need to add concluded sentence after
-changing the data.}
-
-\end{figure}%
 
@@
 have some effect on the distribution of bin counts.
-Figure~\ref{fig-param-scurve} (a) illustrates this. The distribution of
-bin counts for \(s_1\) varying between \(-0.1-0.0\) is shown. Generally,
-a more uniform distribution among these possibilities would indicate
-that the bins are reliably capturing the underlying distribution of
-observations.
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-bins-scurve-1.pdf}
-
-}
-
-\caption{\label{fig-bins-scurve}Hexbin density plots of UMAP layout of
-the S-curve data, using three different bin inputs: (a)
-\(b = 91 \text{ } (7, \text{ }13)\), (b)
-\(b = 220 \text{ } (11, \text{ }20)\), and (c)
-\(b = 312 \text{ } (13, \text{ }24)\). Color indicates standardized
-counts, dark indicating high count and light indicates low count. At the
-smallest bin size the data segregates into two separate groups,
-suggesting this is too many bins. Using the MSE of the model fit in
-\(p\text{-}D\) helps decide on a useful choice of number of bins.}
-
-\end{figure}%
+\textbf{?@fig-param-two\_curvy} (a) illustrates this. The distribution
+of bin counts for \(s_1\) varying between \(-0.1-0.0\) is shown.
+Generally, a more uniform distribution among these possibilities would
+indicate that the bins are reliably capturing the underlying
+distribution of observations.
 
@@
 \(2\) and \(b_1 = \sqrt{\frac{n}{r_2}}\) are allowed.
-Figure~\ref{fig-param-scurve} (b) shows the effect of different choices
-of \(b_1\) on the MSE of the fitted model.
+\textbf{?@fig-param-two\_curvy} (b) shows the effect of different
+choices of \(b_1\) on the MSE of the fitted model.
 
@@
 the bins with density in the lowest 5\% of all bins, for example.
-Figure~\ref{fig-param-scurve} (c) illustrates the effect on the model
+\textbf{?@fig-param-two\_curvy} (c) illustrates the effect on the model
 representation of removing bins below different percentages. Generally,
@@
 model performance, it's essential to observe the change in MSE as the
-benchmark value increases (Figure~\ref{fig-param-scurve}). The MSE shows
-a gradual decrease as the benchmark value goes from \(1\) to \(0\).
-Evaluating this rate of increase is important. If the increment is not
-considerable, the decision might lean towards retaining low-density
-hexagons.
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-param-scurve-1.pdf}
-
-}
-
-\caption{\label{fig-param-scurve}Various plots to help assess best
-hexagon bin parameters, thresholds to remove low density bins and large
-edges. Both (b) and (c) show MSE, against number of bins along the
-x-axis and standardised count. A good benchmark value for these
-parameters is when the MSE drops and then flattens out. Plot (a) shows
-the distribution of stadardised counts of hexagons. Plot (d) shows the
-distribution of \(2\text{-}D\) Euclidean distances between bin
-centroids, with a good benchmark value for removing large edges would
-being the distance that shows the first large decrease.}
-
-\end{figure}%
+benchmark value increases (\textbf{?@fig-param-two\_curvy}). The MSE
+shows a gradual decrease as the benchmark value goes from \(1\) to
+\(0\). Evaluating this rate of increase is important. If the increment
+is not considerable, the decision might lean towards retaining
+low-density hexagons.
 
@@
 \gD{} representation of the fitted model.
-Figure~\ref{fig-scurve-true-sc} shows a wire frame of the true model
+\textbf{?@fig-two\_curvy-true-sc} shows a wire frame of the true model
 that was used to generate the S-curve example data. The ideal is that
@@
 typically done on the hexagon centroids, there are particular discrete
-edge lengths, based on bin widths. Figure~\ref{fig-NLDR-scurve} (d)
+edge lengths, based on bin widths. Figure~\ref{fig-NLDR-two-curvy} (d)
 illustrates edge length distributions.
@@
 initial bin position or removing low density bins.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-41-1.pdf}
-
-}
-
-\caption{long edge removal}
-
-\end{figure}%
 
@@
 
-\caption{\label{fig-scurve-sc-best}Model in \gD{}, on the tSNE layout,
-and three views of the fit in projections from \(7\text{-}D\), for the
-S-curve data (\((s_1, \ s_2) = (-0.170, \ -0.137)\),
+\caption{\label{fig-two_curvy-sc-best}Model in \gD{}, on the tSNE
+layout, and three views of the fit in projections from \(7\text{-}D\),
+for the S-curve data (\((s_1, \ s_2) = (-0.170, \ -0.137)\),
 \(b = 256 \  (16, \ 16)\), \(m = 70\), benchmark value to remove low

### 2024-09-26 paper.qmd
Updated Figure 3

@@
 This process serves to reduce some noisiness in the resulting surface shown in \pD{}.
-The steps in this process are shown in @fig-NLDR-scurve, and documented below.
+The steps in this process are shown in @fig-NLDR-two-curvy, and documented below.
 
@@
 
-umap_two_curvy_scaled <- two_curvy_scaled_obj$scaled_nldr
+umap_two_curvy_scaled <- two_curvy_scaled_obj$scaled_nldr |>
+  mutate(ID = row_number())
 lim1 <- two_curvy_scaled_obj$lim1
@@
 # sc_ylims <- c(-0.32, 2.1)
-sc_xlims <- c(-0.25, 1.25)
-sc_ylims <- c(-0.3, 2)
-
-nldr_scurve <- umap_scurve_scaled |> 
+sc_xlims <- c(-0.2, 1.2)
+sc_ylims <- c(-0.2, 0.8)
+
+nldr_two_curvy <- umap_two_curvy_scaled |> 
   ggplot(aes(x = UMAP1, y = UMAP2)) + 
@@
 ```{r}
-#| label: hexbin-scurve
+#| label: hexbin-two-curvy
 ## Compute hexbin parameters
-num_bins_x_scurve <- 7
+num_bins_x_two_curvy <- 10
 
 ## hexagon binning to have regular hexagons
-hb_obj_scurve <- hex_binning(
-  data = umap_scurve_scaled, 
-  bin1 = num_bins_x_scurve, 
+hb_obj_two_curvy <- hex_binning(
+  data = umap_two_curvy_scaled, 
+  bin1 = num_bins_x_two_curvy, 
   r2 = r2,
@@
 ## Data set with all centroids
-all_centroids_df <- hb_obj_scurve$centroids
+all_centroids_df <- hb_obj_two_curvy$centroids
 
 ## Generate all coordinates of hexagons
-hex_grid <- hb_obj_scurve$hex_poly
+hex_grid <- hb_obj_two_curvy$hex_poly
 
 ## To obtain the standardise counts within hexbins
-counts_df <- hb_obj_scurve$std_cts
+counts_df <- hb_obj_two_curvy$std_cts
 df_bin_centroids <- extract_hexbin_centroids(
@@
 
-hex_grid_scurve <- ggplot(
+hex_grid_two_curvy <- ggplot(
   data = hex_grid_with_counts, 
@@
                fill = "#ffffff") +
-  geom_point(data = umap_scurve_scaled, 
+  geom_point(data = umap_two_curvy_scaled, 
              aes(x = UMAP1, y = UMAP2), 
@@
 ```{r}
-#| label: empty-bin-scurve
+#| label: empty-bin-two-curvy
 hex_grid_nonempty <- hex_grid |>
@@
 
-hex_grid_nonempty_scurve <- ggplot(
+hex_grid_nonempty_two_curvy <- ggplot(
   data = hex_grid_nonempty, 
@@
 <!--2D model-->
-
-```{r}
-#| label: triangulate-scurve
+```{r}
+#| label: triangulate-two-curvy
 ## Triangulate bin centroids
-tr1_object_scurve <- tri_bin_centroids(
+tr1_object_two_curvy <- tri_bin_centroids(
   df_bin_centroids, x = "c_x", y = "c_y")
-tr_from_to_df_scurve <- gen_edges(
-  tri_object = tr1_object_scurve)
+tr_from_to_df_two_curvy <- gen_edges(
+  tri_object = tr1_object_two_curvy)
 
 ## Compute 2D distances
-distance_scurve <- cal_2d_dist(
-  tr_coord_df = tr_from_to_df_scurve, 
+distance_two_curvy <- cal_2d_dist(
+  tr_coord_df = tr_from_to_df_two_curvy, 
   start_x = "x_from", 
@@
 ## To find the benchmark value
-benchmark_scurve <- find_lg_benchmark(
-  distance_edges = distance_scurve, 
+benchmark_two_curvy <- find_lg_benchmark(
+  distance_edges = distance_two_curvy, 
   distance_col = "distance")
 
-trimesh_removed_scurve <- vis_rmlg_mesh(
-  distance_edges = distance_scurve, 
-  benchmark_value = benchmark_scurve, 
-  tr_coord_df = tr_from_to_df_scurve, 
+trimesh_removed_two_curvy <- vis_rmlg_mesh(
+  distance_edges = distance_two_curvy, 
+  benchmark_value = benchmark_two_curvy, 
+  tr_coord_df = tr_from_to_df_two_curvy, 
   distance_col = "distance") +
@@
 ```{r}
-#| label: s-curve-true-model-proj1
-
-true_model_df <- read_rds("data/s_curve/scurve_true_model.rds")
-wireframe_true_model <- read_rds("data/s_curve/scurve_true_model_wireframe.rds")
-
-data_scurve <- training_data_scurve |> 
-  select(-ID) |> 
-  mutate(type = "data")
-
-model_scurve <- true_model_df |> 
-    select(-ID) |> 
-  mutate(type = "model")
-
-# Apply the scaling
-df_model_data_scurve <- bind_rows(data_scurve, model_scurve)
-scaled_scurve <- scale_data_manual(df_model_data_scurve, "type") |>
-  as_tibble()
-
-scaled_s_curve_noise <- scaled_scurve |>
-  filter(type == "data") |>
-  select(-type)
-
-scaled_s_curve_true <- scaled_scurve |>
-  filter(type == "model") |>
-  select(-type)
-
-projection <- cbind(
-    c(0.45181,0.04747,0.02613,0.25735,0.20622,0.08755,-0.11488),
-    c(0.03231,0.23556,0.35166,-0.11741,0.12730,0.09544,0.34262))
-
-projected <- as.matrix(scaled_s_curve_noise) %*% projection
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-projected_true <- as.matrix(scaled_s_curve_true) %*% projection
-
-projected_true_df <- projected_true |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  wireframe_true_model, 
-  projected_true_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_true_df)[-NCOL(projected_true_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_true_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_true_df)):NCOL(model_df)] <- paste0(names(projected_true_df)[-NCOL(projected_true_df)], "_to")
-
-limits <- 1
-rng <- range(projected)
-projected <- projected/max(abs(rng))
-colnames(projected) <- c("P1", "P2")
-projected <- data.frame(projected)
-obs_labels <- as.character(1:nrow(scaled_s_curve_noise))
-
-axis_scale <- limits/6
-axis_pos <- -2/3 * limits
-
-# axis_scale <- limits/4
-# axis_pos <- -0.5 * limits
-
-adj <- function(x) axis_pos + x * axis_scale
-axes <- data.frame(x1 = adj(0), 
-                   y1 = adj(0), 
-                   x2 = adj(projection[, 1]), 
-                   y2 = adj(projection[, 2]))
-
-axis_labels <- colnames(scaled_s_curve_noise)
-rownames(axes) <- axis_labels
-
-# axis_scale <- 0.15 * limits
-# axis_pos <- -0.6 * limits
-# adj <- function(x) axis_pos + x * axis_scale
-theta <- seq(0, 2 * pi, length = 50)
-circle <- data.frame(c1 = adj(cos(theta)), 
-                     c2 = adj(sin(theta)))
-
-scurve_proj1_true_model <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#3182bd") + 
-  geom_point(
-        size = 0.8,
-        alpha = 0.5,
-        color = "#000000") +
-  geom_segment(
-    data=axes, 
-    aes(x=x1, y=y1, xend=x2, yend=y2), 
-    colour="grey70") +
-  geom_text(
-    data=axes, 
-    aes(x=x2, y=y2, label=rownames(axes)), 
-    colour="grey50",
-    size = 3) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed()
-
-```
-
-```{r}
-#| label: s-curve-true-model-proj2
-
-projection <- cbind(
-    c(0.03627,0.15991,0.39768,-0.21032,-0.14886,-0.17589,0.23305),
-    c(-0.34731,-0.08183,0.15819,0.31926,0.08105,0.10083,0.25627))
-
-projected <- as.matrix(scaled_s_curve_noise) %*% projection
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-projected_true <- as.matrix(scaled_s_curve_true) %*% projection
-
-projected_true_df <- projected_true |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  wireframe_true_model, 
-  projected_true_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_true_df)[-NCOL(projected_true_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_true_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_true_df)):NCOL(model_df)] <- paste0(names(projected_true_df)[-NCOL(projected_true_df)], "_to")
-
-limits <- 1
-rng <- range(projected)
-projected <- projected/max(abs(rng))
-colnames(projected) <- c("P1", "P2")
-projected <- data.frame(projected)
-obs_labels <- as.character(1:nrow(scaled_s_curve_noise))
-
-# axis_scale <- 2 * limits/3
-# axis_pos <- -0.6 * limits
-axis_scale <- limits/6
-axis_pos <- -2/3 * limits
-
-adj <- function(x) axis_pos + x * axis_scale
-axes <- data.frame(x1 = adj(0), 
-                   y1 = adj(0), 
-                   x2 = adj(projection[, 1]), 
-                   y2 = adj(projection[, 2]))
-
-axis_labels <- colnames(scaled_s_curve_noise)
-rownames(axes) <- axis_labels
-
-# axis_scale <- 0.15 * limits
-# axis_pos <- -0.6 * limits
-# adj <- function(x) axis_pos + x * axis_scale
-theta <- seq(0, 2 * pi, length = 50)
-circle <- data.frame(c1 = adj(cos(theta)), c2 = adj(sin(theta)))
-
-scurve_proj2_true_model <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#3182bd") +
-  geom_point(
-        size = 0.8,
-        alpha = 0.5,
-        color = "#000000") +
-  geom_segment(
-    data=axes, 
-    aes(x=x1, y=y1, xend=x2, yend=y2), 
-    colour="grey70") +
-  geom_text(
-    data=axes, 
-    aes(x=x2, y=y2, label=rownames(axes)), 
-    colour="grey50",
-    size = 3) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed()
-
-```
-
-```{r}
-#| label: fig-scurve-true-sc
-#| fig-cap: "Two views of the true model (blue lines) in $2\\text{-}D$ projections from $7\\text{-}D$, for the S-curve data (black points).
-The data is spread along the S shape, and does not vary much from this curve.
-(The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/I5GL23vLiw0>)."
-#| fig-pos: H
-#| fig-width: 10
-#| fig-height: 5
-
-scurve_proj1_true_model + scurve_proj2_true_model + 
-  plot_layout(ncol = 2)
+#| label: fig-NLDR-two-curvy
+#| echo: false
+#| fig-cap: "Key steps for constructing the model on the UMAP layout ($k=2$): (a) data, (b) hexagon bins, (c) bin centroids, and (d) triangulated centroids. The S-curve data is shown."
+#| fig-width: 8
+#| fig-height: 4
+#| out-width: 100%
+
+nldr_two_curvy + hex_grid_two_curvy + 
+  hex_grid_nonempty_two_curvy + 
+  trimesh_removed_two_curvy +
+  plot_layout(ncol = 4)
 ```
@@
 <!--langevitour with true model for S-curve--> 
-<!-- ::: {#fig-scurve-true-sc layout-ncol="3" fig-pos="H"} -->
-
-<!-- ![](figures/scurve/sc_true_1.png){width="150" fig-align="center"} -->
-
-<!-- ![](figures/scurve/sc_true_2.png){width="150" fig-align="center"} -->
-
-<!-- ![](figures/scurve/sc_true_3.png){width="150" fig-align="center"} -->
+<!-- ::: {#fig-two_curvy-true-sc layout-ncol="3" fig-pos="H"} -->
+
+<!-- ![](figures/two_curvy/sc_true_1.png){width="150" fig-align="center"} -->
+
+<!-- ![](figures/two_curvy/sc_true_2.png){width="150" fig-align="center"} -->
+
+<!-- ![](figures/two_curvy/sc_true_3.png){width="150" fig-align="center"} -->
 
@@
 
-```{r}
-#| label: fig-NLDR-scurve
-#| echo: false
-#| fig-cap: "Key steps for constructing the model on the UMAP layout ($k=2$): (a) data, (b) hexagon bins, (c) bin centroids, and (d) triangulated centroids. The S-curve data is shown."
-#| fig-width: 8
-#| fig-height: 4
-#| out-width: 100%
-
-nldr_scurve + hex_grid_scurve + 
-  hex_grid_nonempty_scurve + 
-  trimesh_removed_scurve +
-  plot_layout(ncol = 4)
-```
-
 ### Algorithm to represent the model in \gD{}
@@
 The default should take the aspect ratio produced by the NLDR $(r_1, r_2, ..., r_k)$ into account.
-When $k=2$, as in hexagon binning, the default range is $[0, y_{i,\text{max}}], i=1,2$, where $y_{1,\text{max}}=1$ and $y_{2,\text{max}} = \frac{r_2}{r_1}$ (@fig-NLDR-scurve).
+When $k=2$, as in hexagon binning, the default range is $[0, y_{i,\text{max}}], i=1,2$, where $y_{1,\text{max}}=1$ and $y_{2,\text{max}} = \frac{r_2}{r_1}$ (@fig-NLDR-two-curvy).
 If the NLDR aspect ratio is ignored then set $y_ {2,\text{max}} = 1$.
@@
       x = x1min  - rect_adj$x1, # 0 - 0.03, 
-      y = -0.35, 
+      y = -0.2, 
       xend = x1max - rect_adj$x1, #1 - 0.03, 
-      yend = -0.35), #-0.35),
+      yend = -0.2), #-0.35),
       arrow = arrow(length = unit(0.03, "npc"),
@@
                  color = "black")+
-    annotate("text", x=0.5, y=-0.45, 
+    annotate("text", x=0.5, y=-0.25, 
              label = expression(r[1]), color = "black") +
     geom_segment(data=d_rect, aes(
-      x = -0.25, 
+      x = -0.2, 
       y = x2min - rect_adj$x2, #0 - 0.05, 
-      xend = -0.25, 
+      xend = -0.2, 
       yend = x2max - rect_adj$x2), #r2 - 0.05),
@@
                  color = "black")+ 
-    annotate("text", x=-0.35, y=1, 
+    annotate("text", x=-0.25, y=0.25, 
              label = expression(r[2]), color = "black") +
@@
              color = "black", hjust=-0.2) +
-    annotate("text", x=-0.18, y=-0.25, 
+    annotate("text", x=-0.18, y=-0.2, 
       label = expression(group("(", list(s[1], s[2]), ")")),
@@
 
-When $k=2$, the number of bins along the $x$ and $y$ axes, $b_1$ and $b_2$, is computed by considering the scaling factors ${r_1, r_2}$, along with the offset along the axes ($q_1$, $q_2$), and the height ($a_1$) and width ($a_2$) of a regular hexagon (@fig-NLDR-scurve (b)).
+When $k=2$, the number of bins along the $x$ and $y$ axes, $b_1$ and $b_2$, is computed by considering the scaling factors ${r_1, r_2}$, along with the offset along the axes ($q_1$, $q_2$), and the height ($a_1$) and width ($a_2$) of a regular hexagon (@fig-NLDR-two-curvy (b)).
 ($b_1 = \frac{r_1 + q_1}{a_2}$, and $b_2 = \frac{r_2 + q_2}{0.75 \times a_1}$) where $0.75 \times a_1$ accommodate to have hexagons fill the entire 2D space without leaving any gaps between them.)-->
@@
 
-By default, the bin centroid is used for describing a hexagon (as done in @fig-NLDR-scurve (c)), but any measure of center, such as a mean or weighted mean of the points within each hexagon, could be used.
+By default, the bin centroid is used for describing a hexagon (as done in @fig-NLDR-two-curvy (c)), but any measure of center, such as a mean or weighted mean of the points within each hexagon, could be used.
 The bin centers, and the binned data, are the two important components needed to render the model representation in high dimensions.
@@
 
-Delaunay triangulation [@lee1980;@alb2024] is used to connect points so that edges indicate neighbouring observations, in both the NLDR layout (@fig-NLDR-scurve (d)) and the \pD{} model representation.
+Delaunay triangulation [@lee1980;@alb2024] is used to connect points so that edges indicate neighbouring observations, in both the NLDR layout (@fig-NLDR-two-curvy (d)) and the \pD{} model representation.
 When the data has been binned the triangulation connectd centroids.
@@
 
-<!-- When $k = 2$ Delaunay triangulation on $C^{(2)}$ generates the model in \gD{} space, which is a triangular mesh (@fig-NLDR-scurve (d)).
+<!-- When $k = 2$ Delaunay triangulation on $C^{(2)}$ generates the model in \gD{} space, which is a triangular mesh (@fig-NLDR-two-curvy (d)).
 It generates convex hulls of $C^{(2)}$ such that the circumcircle of every triangle in the triangulation contains no other points from $C^{(2)}$.
@@
 
-<!-- langevitour with pD model for S-curve -->
-::: {#fig-scurve-sc layout-ncol="4" fig-pos="H"}
-![](figures/scurve/umap_trimesh_layout.png)
-
-![](figures/scurve/sc_umap_best_1.png)
-
-![](figures/scurve/sc_umap_best_2.png)
-
-![](figures/scurve/sc_umap_best_3.png)
-
-Model in \gD{}, on the UMAP layout, and three views of the fit in projections from $7\text{-}D$, for the S-curve data ($(s_1, \ s_2) = (-0.160, \ -0.263)$, $b = 405 \  (15, \ 27)$, $m = 70$, benchmark value to remove low density hexagons is $0.250$, and benchmark value to remove large edges is $0.159$).
-MSE is $0.0462$.
-The model closely fits the shape, but it doesn't fully fill out the width of the S, which means that it does not adequately capture the surface (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/i7F7xpN1Hz8>).
-:::
-
 <!-- XXX You need to have consistent colouring. In langevitour the points corresponding to observations are purple. They should be purple in the \gD{} view too.  -->
@@
 
-```{r}
-#| label: best-umap-model-scurve
-
-# Initialize effective bins along x
-effective_bin1_scurve <- 15
-
-# Fit high-dimensional model
-scurve_model <- fit_highd_model(
-  training_data = training_data_scurve,
-  emb_df = umap_scurve_scaled,
-  bin1 = effective_bin1_scurve,
-  r2 = r2,
-  q = 0.1,
-  is_bin_centroid = TRUE,
-  is_rm_lwd_hex = FALSE,
-  col_start_highd = "x"
-)
-
-df_bin_centroids_scurve <- scurve_model$df_bin_centroids
-df_bin_scurve <- scurve_model$df_bin
-
-hex_grid_nonempty <- hex_grid |>
-  filter(hex_poly_id %in% df_bin_centroids_scurve$hexID)
-
-```
-
-```{r}
-#| label: p-d-error-in-2d-scurve
-
-## Compute error
-error_df_scurve <- augment(
-  df_bin_centroids = df_bin_centroids_scurve,
-  df_bin = df_bin_scurve,
-  training_data = training_data_scurve,
-  newdata = NULL,
-  type_NLDR = "UMAP",
-  col_start = "x")
-
-# error_df_scurve <- error_df_scurve |>
-#   mutate(type = case_when(
-#     row_wise_abs_error == 0 ~ "no error",
-#     row_wise_abs_error <= 0.2 ~ "error 0-0.2",
-#     row_wise_abs_error <= 0.4 ~ "error 0.2-0.4",
-#     row_wise_abs_error <= 0.6 ~ "error 0.4-0.6",
-#     .default = "error greter than 0.6"
-#   )) |>
-#   mutate(type = factor(type, levels = c(
-#     "no error", "error 0-0.2", "error 0.2-0.4", "error 0.4-0.6", "error greter than 0.6")))
-
-error_df_scurve <- error_df_scurve |>
-  mutate(sqrt_row_wise_abs_error = sqrt(row_wise_abs_error))
-
-error_df_scurve <- error_df_scurve |>
-  bind_cols(umap_scurve_scaled |>
-              select(-ID))
-
-error_plot_scurve <- error_df_scurve |>
-  ggplot(aes(x = UMAP1,
-             y = UMAP2,
-             colour = sqrt_row_wise_abs_error)) +
-  geom_point(alpha=0.5) +
-  scale_colour_continuous_sequential(palette = "Light Grays") +
-  coord_equal()
-```
-
-```{r}
-#| label: fig-p-d-error-in-2d-scurve
-#| fig-cap: "error"
-#| out-height: 30%
-#| fig-pos: H
-
-error_plot_scurve
-```
-
 ### Prediction into \gD{}
@@
 
-<!--prediction for true-model of S-curve-->
-```{r}
-#| label: model-prediction-umap-original
-
-predict_umap_scurve <- read_rds(file = "data/s_curve/s_curve_umap_predict_true.rds")
-
-predict_scurve_obj <- gen_scaled_data(
-    data = predict_umap_scurve)
-
-predict_umap_scurve_scaled <- predict_scurve_obj$scaled_nldr
-
-plot_predict_umap <- ggplot(data = predict_umap_scurve_scaled,
-                            aes(
-                              x = UMAP1,
-                              y = UMAP2
-                            )) +
-  geom_point(alpha = 0.5) +
-  interior_annotation("b", c(0.92, 0.96))
-```
-
-```{r}
-#| label: model-prediction-umap-quollr
-
-true_pred_df <- predict_emb(
-  test_data = true_model_df,
-  df_bin_centroids = df_bin_centroids_scurve,
-  df_bin = df_bin_scurve,
-  type_NLDR = "UMAP"
-)
-
-# Compute radius r
-a1 <- calc_bins_y(bin1 = 15, r2 = r2, q = 0.16)$a1
-r <- a1 / 2
-
-# Function to jitter points within a circumcircle of a hexagon
-jitter_within_circumcircle <- function(center_x, center_y, radius, num_points) {
-  theta <- runif(num_points, 0, 2 * pi)
-  r <- radius * sqrt(runif(num_points))
-  x <- center_x + r * cos(theta)
-  y <- center_y + r * sin(theta)
-  jittered_points <- data.frame(UMAP1 = x, UMAP2 = y, ID = points_in_hex$ID)
-  return(jittered_points)
-}
-
-# Jittered points data frame
-jittered_points_df <- data.frame()
-
-# Iterate through each hexagon and jitter points within the circumcircle
-for (hex_id in unique(true_pred_df$pred_hb_id)) {
-  # hex_points <- hex_grid_nonempty %>% filter(hex_poly_id == hex_id)
-  # center_x <- mean(hex_points$x)
-  # center_y <- mean(hex_points$y)
-  # points_in_hex <- true_pred_df %>% filter(pred_UMAP_1 >= min(hex_points$x) & pred_UMAP_1 <= max(hex_points$x) & pred_UMAP_2 >= min(hex_points$y) & pred_UMAP_2 <= max(hex_points$y))
-
-  points_in_hex <- true_pred_df |>
-    filter(pred_hb_id == hex_id)
-
-  center_x <- points_in_hex |> 
-    pull(pred_UMAP_1) |> 
-    unique()
-
-  center_y <- points_in_hex |> 
-    pull(pred_UMAP_2) |> 
-    unique()
-
-  if (nrow(points_in_hex) > 0) {
-    jittered_points <- jitter_within_circumcircle(center_x, center_y, r, nrow(points_in_hex))
-    jittered_points_df <- rbind(jittered_points_df, jittered_points)
-  }
-}
-
-plot_predict_umap_model <- ggplot(data = jittered_points_df,
-                            aes(
-                              x = UMAP1,
-                              y = UMAP2
-                            )) +
-  geom_point(alpha = 0.5) +
-  interior_annotation("c", c(0.92, 0.96))
-```
-
-```{r}
-#| label: true-model-projection-langevitour
-
-true_model_df_rm_id <- true_model_df |>
-  select(-ID)
-
-projection <- cbind(
-    c(0.18390,-0.18443,-0.06860,0.00654,-0.02921,0.02313,-0.08963),
-    c(-0.02902,0.03638,-0.16872,0.22418,0.01146,0.02777,0.01453))
-projected_df <- as.matrix(true_model_df_rm_id) %*% projection |> 
-  as_tibble(.name_repair = "unique") |> 
-  rename(c("proj1" = "...1", 
-           "proj2" = "...2"))
-
-proj_plot_scurve_true <- ggplot(
-  data = projected_df, 
-  aes(
-    x = proj1, 
-    y = proj2)) +
-    geom_point(alpha = 0.5) +
-  interior_annotation("a", c(0.92, 0.96))  
-```
-
-```{r}
-#| label: model-prediction-points-line
-
-predict_umap_scurve_scaled <- predict_umap_scurve_scaled |>
-  mutate(type = "predict_from_umap")
-
-jittered_points_df <- jittered_points_df |>
-  mutate(type = "predict_from_model")
-
-predict_point_df <- bind_rows(
-  predict_umap_scurve_scaled,
-  jittered_points_df
-)
-
-plot_predict_umap_model_connect <- ggplot(data = predict_point_df,
-                            aes(
-                              x = UMAP1,
-                              y = UMAP2, 
-                              group = ID
-                            )) +
-  geom_point(alpha = 0.5) +
-  geom_line(alpha = 0.3) +
-  interior_annotation("d", c(0.92, 0.96))
-```
-
-```{r}
-#| echo: false
-#| label: fig-predict-scurve
-#| fig-pos: H
-#| fig-width: 9
-#| fig-height: 5
-#| fig-cap: "Comparison of prediction generated using the exiting `umap` prediction method and our method: (a) A view of the true model in projections from $7\\text{-}D$, (b) predicted data from the `umap` prediction method, (c) predicted data from our method, and (d) comparison of predictions from both methods.
-In plot (d), points representing the same data but predicted by different methods are connected by lines.
-Need to add concluded sentence after changing the data."
-
-proj_plot_scurve_true + plot_predict_umap + 
-  plot_predict_umap_model + plot_predict_umap_model_connect +
-  plot_layout(ncol = 4)
-```
-
-<!-- ::: {#fig-scurve-pred-sc layout-ncol="2" fig-pos="H"} -->
-<!-- ![](figures/scurve/sc_true_only.png){width="150" fig-align="center"} -->
-
-<!-- ![](figures/scurve/pred_true_view.png){width="150" fig-align="center"} -->
+<!-- ::: {#fig-two_curvy-pred-sc layout-ncol="2" fig-pos="H"} -->
+<!-- ![](figures/two_curvy/sc_true_only.png){width="150" fig-align="center"} -->
+
+<!-- ![](figures/two_curvy/pred_true_view.png){width="150" fig-align="center"} -->
 
@@
 The choice of these values can have some effect on the distribution of bin counts.
-@fig-param-scurve (a) illustrates this.
+@fig-param-two_curvy (a) illustrates this.
 The distribution of bin counts for $s_1$ varying between $-0.1-0.0$ is shown.
@@
 
-<!--add three choice of bins 7, 11, 14-->
-
-```{r}
-#| label: hexbin-regular-scurve1
-## hexagon binning to have regular hexagons
-hb_obj_scurve1 <- hex_binning(
-  data = umap_scurve_scaled, 
-  bin1 = 7, 
-  r2 = r2)
-
-a1_1 <- calc_bins_y(
-  bin1 = 7, 
-  r2 = r2
-)$a1
-
-## Data set with all centroids
-all_centroids_df1 <- hb_obj_scurve1$centroids
-
-## Generate all coordinates of hexagons
-hex_grid1 <- hb_obj_scurve1$hex_poly
-
-## To obtain the standardise counts within hexbins
-counts_df1 <- hb_obj_scurve1$std_cts
-df_bin_centroids1 <- extract_hexbin_centroids(
-  centroids_df = all_centroids_df1, 
-  counts_df = counts_df1) |>
-  filter(drop_empty == FALSE) |>
-  mutate(b1 = "b1 = 7")
-
-hex_grid_with_counts_s_curve1 <- full_join(
-  hex_grid1, 
-  df_bin_centroids1 |> select(hexID, std_counts), 
-    by = c("hex_poly_id" = "hexID"))
-
-hex_grid_coloured_scurve1 <- ggplot() + 
-  geom_polygon(
-    data = hex_grid_with_counts_s_curve1, 
-    aes(x = x, y = y,
-      group = hex_poly_id, 
-      fill = std_counts), color = "grey70", linewidth=0.2) +
-  geom_point(data = umap_scurve_scaled,
-             aes(x = UMAP1, y = UMAP2),
-             alpha = 0.3,
-             size = 0.5) +
-  scale_fill_viridis_c(direction = -1, 
-    na.value = "#ffffff", option = "E") +
-  interior_annotation("a", position = c(0.92, 0.93)) +
-  xlim(c(-0.25, 1.35)) + ylim(c(-0.35, 2))
-
-hex_grid_coloured_scurve1_dens <- ggplot(hex_grid_with_counts_s_curve1) +
-  geom_histogram(aes(x=std_counts), 
-                 breaks=seq(0, 1, 0.05),
-                 fill="slategray4", colour="white")
-
-#hex_grid_coloured_scurve1_dens <- 
-#  hex_grid_with_counts_s_curve1 |>
-#  filter(!is.na(std_counts)) |>
-#  mutate(ord_std_counts = n() - rank(std_counts)) |>
-#  ggplot() +
-#  geom_line(aes(x=ord_std_counts, y=std_counts), 
-#                 colour="slategray4")
-```
-
-```{r}
-#| label: hexbin-regular-scurve2
-## hexagon binning to have regular hexagons
-hb_obj_scurve2 <- hex_binning(
-  data = umap_scurve_scaled, 
-  bin1 = 11, 
-  r2 = r2)
-
-a1_2 <- calc_bins_y(
-  bin1 = 11, 
-  r2 = r2
-  )$a1
-
-## Data set with all centroids
-all_centroids_df2 <- hb_obj_scurve2$centroids
-
-## Generate all coordinates of hexagons
-hex_grid2 <- hb_obj_scurve2$hex_poly
-
-## To obtain the standardise counts within hexbins
-counts_df2 <- hb_obj_scurve2$std_cts
-df_bin_centroids2 <- extract_hexbin_centroids(
-  centroids_df = all_centroids_df2, 
-  counts_df = counts_df2) |>
-  filter(drop_empty == FALSE) |>
-  mutate(b1 = "b1 = 11")
-
-hex_grid_with_counts_s_curve2 <- full_join(
-  hex_grid2, 
-  df_bin_centroids2 |> select(hexID, std_counts), 
-  by = c("hex_poly_id" = "hexID"))
-
-hex_grid_coloured_scurve2 <- ggplot() + 
-  geom_polygon(
-    data = hex_grid_with_counts_s_curve2, 
-    aes(x = x, y = y, 
-        group = hex_poly_id, 
-        fill = std_counts), 
-        color = "grey70", 
-        linewidth=0.2) +
-  geom_point(data = umap_scurve_scaled,
-           aes(x = UMAP1, y = UMAP2),
-           alpha = 0.3,
-           size = 0.5) +
-  scale_fill_viridis_c(direction = -1, 
-    na.value = "#ffffff", option = "C") +
-  xlim(c(-0.25, 1.35)) + ylim(c(-0.35, 2)) +
-  interior_annotation("b")
-
-hex_grid_coloured_scurve2_dens <- ggplot(hex_grid_with_counts_s_curve2) +
-  geom_histogram(aes(x=std_counts), 
-                 breaks=seq(0, 1, 0.05),
-                 fill="slategray4", colour="white")
-
-#hex_grid_coloured_scurve2_dens <- 
-#  hex_grid_with_counts_s_curve2 |>
-#  filter(!is.na(std_counts)) |>
-#  mutate(ord_std_counts = n() - rank(std_counts)) |>
-#  ggplot() +
-#  geom_line(aes(x=ord_std_counts, y=std_counts), 
-#                 colour="slategray4")
-
-```
-
-```{r}
-#| label: hexbin-regular-scurve3
-## hexagon binning to have regular hexagons
-hb_obj_scurve3 <- hex_binning(
-  data = umap_scurve_scaled, 
-  bin1 = 13, 
-  r2 = r2)
-
-a1_3 <- calc_bins_y(
-  bin1 = 13, 
-  r2 = r2
-)$a1
-
-## Data set with all centroids
-all_centroids_df3 <- hb_obj_scurve3$centroids
-
-## Generate all coordinates of hexagons
-hex_grid3 <- hb_obj_scurve3$hex_poly
-
-## To obtain the standardise counts within hexbins
-counts_df3 <- hb_obj_scurve3$std_cts
-df_bin_centroids3 <- extract_hexbin_centroids(
-  centroids_df = all_centroids_df3, 
-  counts_df = counts_df3) |>
-  filter(drop_empty == FALSE) |>
-  mutate(b1 = "b1 = 14")
-
-hex_grid_with_counts_s_curve3 <- full_join(hex_grid3, 
-                                          df_bin_centroids3 |> select(hexID, std_counts), 
-                                          by = c("hex_poly_id" = "hexID"))
-
-hex_grid_coloured_scurve3 <-  ggplot() + 
-  geom_polygon(
-    data = hex_grid_with_counts_s_curve3, 
-    aes(x = x, y = y, 
-        group = hex_poly_id, 
-        fill = std_counts), 
-        color = "grey70", linewidth=0.2) +
-  geom_point(data = umap_scurve_scaled,
-           aes(x = UMAP1, y = UMAP2),
-           alpha = 0.3,
-           size = 0.5) +
-  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "D") +
-  xlim(c(-0.25, 1.35)) + ylim(c(-0.35, 2)) +
-  interior_annotation("c")
-
-hex_grid_coloured_scurve3_dens <- ggplot(hex_grid_with_counts_s_curve3) +
-  geom_histogram(aes(x=std_counts), 
-                 breaks=seq(0, 1, 0.05),
-                 fill="slategray4", colour="white") 
-#hex_grid_coloured_scurve3_dens <- 
-#  hex_grid_with_counts_s_curve3 |>
-#  filter(!is.na(std_counts)) |>
-#  mutate(ord_std_counts = n() - rank(std_counts)) |>
-#  ggplot() +
-#  geom_line(aes(x=ord_std_counts, y=std_counts), 
-#                 colour="slategray4")
-```
-
-```{r}
-#| echo: false
-#| label: fig-bins-scurve
-#| fig-pos: H
-#| fig-cap: "Hexbin density plots of UMAP layout of the S-curve data, using three different bin inputs: (a) $b = 91 \\text{ } (7, \\text{ }13)$, (b) $b = 220 \\text{ } (11, \\text{ }20)$, and (c) $b = 312 \\text{ } (13, \\text{ }24)$.
-Color indicates standardized counts, dark indicating high count and light indicates low count.
-At the smallest bin size the data segregates into two separate groups, suggesting this is too many bins.
-Using the MSE of the model fit in $p\\text{-}D$ helps decide on a useful choice of number of bins."
-#| fig-width: 6
-#| fig-height: 4
-
-hex_grid_coloured_scurve1 + 
-  hex_grid_coloured_scurve2 + 
-  hex_grid_coloured_scurve3 +
-  plot_layout(guides='collect', ncol = 3,
-              heights = c(2,1)) &
-  theme(legend.position='none', plot.tag = element_text(size = 8))
-```
-
 The default number of bins $b=b_1\times b_2$ is computed based on the sample size, by setting $b_1=n^{1/3}$, consistent with the Diaconis-Freedman rule [@freedman1981].
@@
 Values of $b_1$ between $2$ and $b_1 = \sqrt{\frac{n}{r_2}}$ are allowed.
-@fig-param-scurve (b) shows the effect of different choices of $b_1$ on the MSE of the fitted model.
+@fig-param-two_curvy (b) shows the effect of different choices of $b_1$ on the MSE of the fitted model.
 
@@
 The minimum $b_1$ is set to $2$, while the maximum number is estimated by taking the square root of $\frac{n}{2}$.
-By evaluating MSE across varying $b$ within this range for different $q$, helps to determine an appropriate values for $b$ and $q$ (@fig-param-scurve (a)).-->
+By evaluating MSE across varying $b$ within this range for different $q$, helps to determine an appropriate values for $b$ and $q$ (@fig-param-two_curvy (a)).-->
 
@@
 <!--To generate errors for different total number of bins-->
-
-```{r}
-#| label: errors-scurve
-## To initialize number of bins along the x-axis
-bin1_vec_scurve <- 5:22 #sqrt(NROW(umap_scurve_scaled)/r2)
-#buffer_vec <- seq(0.05, 0.1, by = 0.01)
-
-error_scurve <- data.frame(matrix(nrow = 0, ncol = 0))
-
-for (xbins in bin1_vec_scurve) {
-  #for (buff in buffer_vec) {
-
-  bin_obj <- calc_bins_y(
-    bin1 = xbins, 
-    r2 = r2, 
-    q = 0.1)
-
-  bin2 <- bin_obj$bin2
-  a1 <- bin_obj$a1
-  a2 <- bin_obj$a2
-
-  scurve_model <- fit_highd_model(
-    training_data = training_data_scurve,
-    emb_df = umap_scurve_scaled,
-    bin1 = xbins,
-    r2 = r2,
-    q = 0.1, 
-    is_bin_centroid = TRUE,
-    is_rm_lwd_hex = FALSE,
-    col_start_highd = "x"
-  )
-
-  df_bin_centroids_scurve <- scurve_model$df_bin_centroids
-  df_bin_scurve <- scurve_model$df_bin
-
-  ## Compute error
-  error_df <- glance(
-    df_bin_centroids = df_bin_centroids_scurve,
-    df_bin = df_bin_scurve,
-    training_data = training_data_scurve,
-    newdata = NULL,
-    type_NLDR = "UMAP",
-    col_start = "x") |>
-    mutate(bin1 = xbins,
-           bin2 = bin2,
-           b = bin1 * bin2,
-           b_non_empty = NROW(df_bin_centroids_scurve),
-           a1 = round(a1, 2),
-           a2 = round(a2, 2),
-           side_length = quad(a=3, b = 2 * a2, c = -(a2^2 + a1^2)))
-
-  error_scurve <- bind_rows(error_scurve, error_df)
-
-  #}
-
-}
-
-# mse_scurve_b <- ggplot(error_scurve, 
-#                      aes(x = b, 
-#                          y = log(MSE),
-#                          color = buff,
-#                          group = buff)) + 
-#   geom_point(size = 1) +
-#   geom_line(linewidth = 0.05) + 
-#   geom_vline(xintercept = 220, linetype="solid", 
-#              color = "black", linewidth=0.1) +
-#   geom_vline(xintercept = 112, linetype=2, 
-#              color = "black", linewidth=0.1) +
-#   geom_vline(xintercept = 364, linetype=2, 
-#              color = "black", linewidth=0.1) +
-#   scale_colour_continuous_sequential(palette = "Magenta") +
-#   labs(x = expression(b), y = "log(MSE)", color = expression(q)) +
-#   ggtitle("(a)") +
-#   theme(aspect.ratio = 0.75,
-#         axis.text.x = element_text(size = 12),
-#         axis.text.y = element_text(size = 12),
-#         axis.title.x = element_text(size = 12),
-#         axis.title.y = element_text(size = 12, 
-#             angle=90),
-#         plot.title = element_text(size = 12),
-#         legend.title = element_text(size = 12),
-#         legend.text = element_text(size = 10),
-#         legend.key.height = unit(1, 'cm'),
-#         legend.key.width = unit(1, 'cm'),
-#         legend.key.size = unit(1, 'cm'))
-
-# diaconis_bin_width <- calc_bins_y(
-#     bin1 = ceiling(NROW(training_data_scurve)^(1/3)),
-#     r2 = r2,
-#     q = 0.1)$a1
-
-## Find the minimum MSE when have duplicate a1
-error_scurve <- error_scurve |>
-  group_by(a1) |>
-  filter(MSE == min(MSE)) |>
-  ungroup() |>
-  mutate(prop_dens = b_non_empty/side_length)
-
-base_line_prop_dens <- error_scurve |>
-  filter(a1 == min(a1)) |>
-  pull(prop_dens)
-
-error_scurve <- error_scurve |>
-  mutate(prop_comp = prop_dens/base_line_prop_dens)
-
-mse_scurve_b <- ggplot(error_scurve, 
-                     aes(x = a1, 
-                         y = MSE)) +
-  geom_vline(xintercept = 0.10, linetype=2,
-           color = "#54278f", linewidth=1) +
-  geom_vline(xintercept = 0.12,
-             linetype="solid",
-             color = "#006d2c",
-             linewidth=1) +
-  geom_vline(xintercept = 0.18, linetype=2,
-             color = "#08519c", linewidth=1) +
-  geom_line(linewidth = 0.5) + 
-  geom_point(size = 1) +
-  scale_x_continuous(breaks = sort(unique(round(error_scurve$a1, 2)))[c(1, 3, 5, 6, 8, 10:12)]) +
-  labs(x = expression(paste("binwidth (", a[1], ")")), y = "MSE") +
-  ggtitle("(a)") +
-  theme_minimal() +
-  theme(aspect.ratio = 0.75,
-        panel.border = element_rect(fill = 'transparent'),
-        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
-        axis.ticks.x = element_line(),
-        axis.ticks.y = element_line())
-
-prop_dens_a1 <- ggplot(error_scurve,
-                     aes(x = a1,
-                         y = prop_comp)) +
-  geom_vline(xintercept = 0.12, linetype="solid",
-             color = "#006d2c", linewidth=1) +
-  geom_vline(xintercept = 0.10, linetype=2,
-             color = "#54278f", linewidth=1) +
-  geom_vline(xintercept = 0.18, linetype=2,
-             color = "#08519c", linewidth=1) +
-  geom_point(size = 1) +
-  geom_line(linewidth = 0.05) +
-  # scale_y_continuous(breaks = sort(unique(error_scurve$b_non_empty))[-3]) +
-  scale_x_continuous(breaks = sort(unique(round(error_scurve$a1, 2)))[c(1, 3, 5, 6, 8, 10:12)]) +
-  labs(x = expression(paste("binwidth (", a[1], ")")), y = paste("proportion of densities")) +
-  ggtitle("(d)") +
-  theme_minimal() +
-  theme(aspect.ratio = 0.75,
-        panel.border = element_rect(fill = 'transparent'),
-        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
-        axis.ticks.x = element_line(),
-        axis.ticks.y = element_line())
-
-error_scurve <- error_scurve |>
-  mutate(prop_bins = b_non_empty/b)
-
-a1_m_scurve <- ggplot(error_scurve,
-                     aes(x = a1,
-                         y = prop_bins)) +
-  geom_vline(xintercept = 0.12, linetype="solid",
-             color = "#006d2c", linewidth=1) +
-  geom_vline(xintercept = 0.10, linetype=2,
-             color = "#54278f", linewidth=1) +
-  geom_vline(xintercept = 0.18, linetype=2,
-             color = "#08519c", linewidth=1) +
-  geom_point(size = 1) +
-  geom_line(linewidth = 0.05) +
-  # scale_y_continuous(breaks = sort(unique(error_scurve$b_non_empty))[-3]) +
-  scale_x_continuous(breaks = sort(unique(round(error_scurve$a1, 2)))[c(1, 3, 5, 6, 8, 10:12)]) +
-  labs(x = expression(paste("binwidth (", a[1], ")")), y = expression(paste("proportion of non-empty bins ", bgroup("(", frac(m, b), ")")))) +
-  ggtitle("(b)") +
-  theme_minimal() +
-  theme(aspect.ratio = 0.75,
-        panel.border = element_rect(fill = 'transparent'),
-        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
-        axis.ticks.x = element_line(),
-        axis.ticks.y = element_line())
-
-# 
-# b1_m_scurve2 <- ggplot(error_scurve, 
-#                      aes(x = bin1, 
-#                          y = b_non_empty)) + 
-#   geom_point(size = 1) +
-#   geom_line(linewidth = 0.05) + 
-#   # geom_vline(xintercept = 11, linetype="solid",
-#   #            color = "#bdbdbd", linewidth=1) +
-#   # geom_vline(xintercept = 9, linetype=2,
-#   #            color = "#bdbdbd", linewidth=1) +
-#   # geom_vline(xintercept = 13, linetype=2,
-#   #            color = "#bdbdbd", linewidth=1) +
-#   scale_x_continuous(breaks = 5:19) +
-#   labs(x = expression(b[1]), y = expression(m)) +
-#   ggtitle("(c)") +
-#   theme(aspect.ratio = 0.75,
-#         axis.text.x = element_text(size = 12),
-#         axis.text.y = element_text(size = 12),
-#         axis.title.x = element_text(size = 12),
-#         axis.title.y = element_text(size = 12, 
-#             angle=90),
-#         axis.ticks.x = element_line(linewidth = 0.5),
-#         axis.ticks.y = element_line(linewidth = 0.5),
-#         plot.title = element_text(size = 12),
-#         legend.title = element_text(size = 12),
-#         legend.text = element_text(size = 10),
-#         legend.key.height = unit(1, 'cm'),
-#         legend.key.width = unit(1, 'cm'),
-#         legend.key.size = unit(1, 'cm'))
-```
 
@@
 Alternatively, one could choose to remove based on a percentile, the bins with density in the lowest 5% of all bins, for example.
-@fig-param-scurve (c) illustrates the effect on the model representation of removing bins below different percentages.
+@fig-param-two_curvy (c) illustrates the effect on the model representation of removing bins below different percentages.
 Generally, we would urge caution in removing low count bins.
 
-<!-- Once setting up the hexagon grid with an appropriate number of bins, some hexagon bins may have few or no data points within them (@fig-bins-scurve (b)).
+<!-- Once setting up the hexagon grid with an appropriate number of bins, some hexagon bins may have few or no data points within them (@fig-bins-two_curvy (b)).
 To ensure comprehensive coverage of the NLDR data, it is necessary to select hexagon bins with a considerable number of data points.
@@
 Then, the standard count is computed by dividing the number of points within each hexagon by the maximum number of points in the grid.
-Next, bins with a standard count less than a benchmark value are removed (@fig-param-scurve (c)).
+Next, bins with a standard count less than a benchmark value are removed (@fig-param-two_curvy (c)).
 There is no specific rule for selecting a benchmark value.
@@
 
-1. Plot the distribution of the standardized counts (@fig-param-scurve (b)).
+1. Plot the distribution of the standardized counts (@fig-param-two_curvy (b)).
 2. Examine the distribution of counts.
@@
 
-<!--add distribution of density and add the first quantile as the benchmark-->
-
-```{r}
-#| label: distribution-scurve
-## To initialize effective bins along x
-effective_bin1_scurve <- 11
-
-effective_bin2_scurve <- calc_bins_y(
-  bin1 = effective_bin1_scurve, 
-  r2 = r2)$bin2
-
-scurve_model <- fit_highd_model(
-  training_data = training_data_scurve,
-  emb_df = umap_scurve_scaled,
-  bin1 = effective_bin1_scurve,
-  r2 = r2,
-  is_bin_centroid = TRUE,
-  is_rm_lwd_hex = FALSE,
-  col_start_highd = "x"
-)
-
-df_bin_centroids_scurve <- scurve_model$df_bin_centroids
-df_bin_scurve <- scurve_model$df_bin
-
-## To obtain the first quntile
-benchmark1 <- round(quantile(df_bin_centroids_scurve$std_counts, names = FALSE)[2], 3)
-```
-
-```{r}
-#| label: bin-counts-scurve
-#| eval: false
-
-df_bin_centroids_all <- bind_rows(
-  df_bin_centroids1,
-  df_bin_centroids2,
-  df_bin_centroids3
-)
-
-# cell_count_scurve <- ggplot(df_bin_centroids_all, 
-#                             aes(x = reorder(as.factor(hexID), 
-#                                             -bin_counts), 
-#                                 y = bin_counts,
-#                                 group = as.factor(b1),
-#                                 color = as.factor(b1))) + 
-#   geom_line() + 
-#   # geom_hline(yintercept = benchmark1, color = "black", linewidth=0.5, alpha = 0.5) +
-#   scale_colour_discrete_qualitative(palette = "Dark 3") +
-#   xlab("hexagon id (re-ordered)") + 
-#   ylab("bin count") +
-#   ggtitle("(a)") +
-#   theme(aspect.ratio = 0.75,
-#         axis.title.x = element_text(size = 12),
-#         axis.title.y = element_text(size = 12, angle = 90),
-#         axis.text.x = element_blank(),
-#         axis.text.y = element_text(size = 12),
-#         plot.title = element_text(size = 12),
-#         legend.text = element_text(size = 10),
-#         legend.key.size = unit(1, 'cm'),
-#         legend.key.height = unit(1, 'cm'))
-
-cell_count_scurve <- ggplot(df_bin_centroids_all, 
-                            aes(x = as.factor(hexID), 
-                                y = std_counts,
-                                group = as.factor(b1),
-                                color = as.factor(b1))) + 
-  geom_line() + 
-  # geom_hline(yintercept = benchmark1, color = "black", linewidth=0.5, alpha = 0.5) +
-  scale_colour_discrete_qualitative(palette = "Dark 3") +
-  xlab("hexagon id (re-ordered)") + 
-  ylab("standardised bin count") +
-  ggtitle("(a)") +
-  theme(aspect.ratio = 0.75,
-        axis.title.x = element_text(size = 12),
-        axis.title.y = element_text(size = 12, angle = 90),
-        axis.text.x = element_blank(),
-        axis.text.y = element_text(size = 12),
-        plot.title = element_text(size = 12),
-        legend.text = element_text(size = 10),
-        legend.key.size = unit(1, 'cm'),
-        legend.key.height = unit(1, 'cm'))
-```
-
 The benchmark value for removing low-density hexagons ranges between $0$ and $1$.
-When analyzing how these benchmark values influence model performance, it's essential to observe the change in MSE as the benchmark value increases (@fig-param-scurve).
+When analyzing how these benchmark values influence model performance, it's essential to observe the change in MSE as the benchmark value increases (@fig-param-two_curvy).
 The MSE shows a gradual decrease as the benchmark value goes from $1$ to $0$.
@@
 
-<!--To generate errors for different total number of bins-->
-```{r}
-#| label: bins-scurve
-
-## To initialize benchmark values to remove low density hexagons
-benchmark_rm_hex_vec <- sort(append(seq(0, 0.6, by=0.15), benchmark1))
-
-## option 1
-error_rm_scurve <- data.frame(matrix(nrow = 0, ncol = 0))
-
-for (benchmark_rm_lwd in benchmark_rm_hex_vec) {
-
-  df_bin_centroids_scurve_high_dens <- df_bin_centroids_scurve |>
-    filter(std_counts > benchmark_rm_lwd)
-
-  df_bin_scurve_high_dens <- df_bin_scurve |>
-    filter(hb_id %in% df_bin_centroids_scurve_high_dens$hexID)
-
-  ## Compute error
-  error_df <- glance(
-    df_bin_centroids = df_bin_centroids_scurve_high_dens,
-    df_bin = df_bin_scurve_high_dens,
-    training_data = training_data_scurve,
-    newdata = NULL,
-    type_NLDR = "UMAP",
-    col_start = "x") |>
-    mutate(benchmark_rm_lwd = round(benchmark_rm_lwd, 2),
-           bin1 = effective_bin1_scurve,
-           bin2 = effective_bin2_scurve,
-           b = bin1 * bin2,
-           b_non_empty = NROW(df_bin_centroids_scurve_high_dens),
-           mean_counts = sum(df_bin_centroids_scurve_high_dens$bin_counts)/NROW(df_bin_centroids_scurve_high_dens))
-
-  error_rm_scurve <- bind_rows(error_rm_scurve, error_df)
-
-}
-
-# benchmark_label_df <- tibble(x = 0.2, y = 0.1, 
-#                           label = paste0("benchmark is ", benchmark1))
-
-error_rm_scurve <- error_rm_scurve |>
-  mutate(mean_counts = round(mean_counts, 0)) |>
-  filter(b_non_empty >= 5)
-
-### Option 2
-
-scurve_model2 <- fit_highd_model(
-  training_data = training_data_scurve,
-  emb_df = umap_scurve_scaled,
-  bin1 = 7,
-  r2 = r2,
-  is_bin_centroid = TRUE,
-  is_rm_lwd_hex = FALSE,
-  col_start_highd = "x"
-)
-
-df_bin_centroids_scurve2 <- scurve_model2$df_bin_centroids
-df_bin_scurve2 <- scurve_model2$df_bin
-
-error_rm_scurve2 <- data.frame(matrix(nrow = 0, ncol = 0))
-
-for (benchmark_rm_lwd in benchmark_rm_hex_vec) {
-
-  df_bin_centroids_scurve_high_dens <- df_bin_centroids_scurve2 |>
-    filter(std_counts > benchmark_rm_lwd)
-
-  df_bin_scurve_high_dens <- df_bin_scurve2 |>
-    filter(hb_id %in% df_bin_centroids_scurve_high_dens$hexID)
-
-  ## Compute error
-  error_df <- glance(
-    df_bin_centroids = df_bin_centroids_scurve_high_dens,
-    df_bin = df_bin_scurve_high_dens,
-    training_data = training_data_scurve,
-    newdata = NULL,
-    type_NLDR = "UMAP",
-    col_start = "x") |>
-    mutate(benchmark_rm_lwd = round(benchmark_rm_lwd, 2),
-           bin1 = 7,
-           bin2 = 13,
-           b = bin1 * bin2,
-           b_non_empty = NROW(df_bin_centroids_scurve_high_dens),
-           mean_counts = sum(df_bin_centroids_scurve_high_dens$bin_counts)/NROW(df_bin_centroids_scurve_high_dens))
-
-  error_rm_scurve2 <- bind_rows(error_rm_scurve2, error_df)
-
-}
-
-### Option 3
-
-scurve_model3 <- fit_highd_model(
-  training_data = training_data_scurve,
-  emb_df = umap_scurve_scaled,
-  bin1 = 13,
-  r2 = r2,
-  is_bin_centroid = TRUE,
-  is_rm_lwd_hex = FALSE,
-  col_start_highd = "x"
-)
-
-df_bin_centroids_scurve3 <- scurve_model3$df_bin_centroids
-df_bin_scurve3 <- scurve_model3$df_bin
-
-error_rm_scurve3 <- data.frame(matrix(nrow = 0, ncol = 0))
-
-for (benchmark_rm_lwd in benchmark_rm_hex_vec) {
-
-  df_bin_centroids_scurve_high_dens <- df_bin_centroids_scurve3 |>
-    filter(std_counts > benchmark_rm_lwd)
-
-  df_bin_scurve_high_dens <- df_bin_scurve3 |>
-    filter(hb_id %in% df_bin_centroids_scurve_high_dens$hexID)
-
-  ## Compute error
-  error_df <- glance(
-    df_bin_centroids = df_bin_centroids_scurve_high_dens,
-    df_bin = df_bin_scurve_high_dens,
-    training_data = training_data_scurve,
-    newdata = NULL,
-    type_NLDR = "UMAP",
-    col_start = "x") |>
-    mutate(benchmark_rm_lwd = round(benchmark_rm_lwd, 2),
-           bin1 = 13,
-           bin2 = 24,
-           b = bin1 * bin2,
-           b_non_empty = NROW(df_bin_centroids_scurve_high_dens),
-           mean_counts = sum(df_bin_centroids_scurve_high_dens$bin_counts)/NROW(df_bin_centroids_scurve_high_dens))
-
-  error_rm_scurve3 <- bind_rows(error_rm_scurve3, error_df)
-
-}
-
-error_rm_scurve <- bind_rows(
-  error_rm_scurve,
-  error_rm_scurve2,
-  error_rm_scurve3
-) |>
-  mutate(bin1  = as.factor(bin1))
-
-mse_scurve_lwd <- ggplot(error_rm_scurve, 
-                     aes(x = benchmark_rm_lwd, 
-                         y = MSE,
-                         color = bin1)) + 
-  geom_point(
-    size = 1
-    ) +
-  geom_line(
-    linewidth = 0.3
-    ) + 
-  # geom_vline(xintercept = benchmark1, linetype="solid", 
-  #            color = "#bdbdbd", linewidth=1, alpha = 0.5) +
-  # scale_x_continuous("standardized bin count", 
-  #        transform = "reverse") + 
-  scale_x_continuous(breaks = unique(error_rm_scurve$benchmark_rm_lwd)) +
-  scale_color_manual(values=c("#08519c", "#006d2c", "#54278f")) +
-  xlab("threshold to remove low-density hexagons") +
-  ylab("MSE") +
-  ggtitle("(c)") +
-  theme_minimal() +
-  theme(aspect.ratio = 0.75,
-        panel.border = element_rect(fill = 'transparent'),
-        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
-        axis.ticks.x = element_line(),
-        axis.ticks.y = element_line(),
-        legend.position = "none")
-```
-
 <!--base line -->
@@
 Removing unnecessary bins may lead to the formation of long edges and an uneven \gD{} model.
-Hence, rather than solely relying on the benchmark value to identify hexagons for removal, it's essential to consider the standard number of points in the neighboring hexagons of the identified low-density bins (see @fig-lwd-scurve (b)).
+Hence, rather than solely relying on the benchmark value to identify hexagons for removal, it's essential to consider the standard number of points in the neighboring hexagons of the identified low-density bins (see @fig-lwd-two_curvy (b)).
 If neighboring bins also show low counts, only those bins will be removed.
@@
 
-<!--Observing neighboring bins as well-->
-```{r}
-#| label: bin-neighbours
-## First define low density bins using first quantile
-df_bin_centroids_low <- df_bin_centroids_scurve |> 
-  filter(std_counts <= benchmark1)
-
-## Check neighboring bins
-remove_id <- find_low_dens_hex(df_bin_centroids_all = df_bin_centroids_scurve, 
-                               bin1 = effective_bin1_scurve, 
-                               df_bin_centroids_low = df_bin_centroids_low)
-
-## Remove the identified bins
-df_bin_centroids_scurve_removed <- df_bin_centroids_scurve |>
-  filter(hexID %in%remove_id)
-
-df_bin_centroids_scurve_keep  <- df_bin_centroids_scurve |>
-  filter(!(hexID %in%remove_id))
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "Various plots to help assess best hexagon bin parameters, thresholds to remove low density bins and large edges.
-Both (b) and (c) show MSE, against number of bins along the x-axis and standardised count.
-A good benchmark value for these parameters is when the MSE drops and then flattens out.
-Plot (a) shows the distribution of stadardised counts of hexagons.
-Plot (d) shows the distribution of $2\\text{-}D$ Euclidean distances between bin centroids, with a good benchmark value for removing large edges would being the distance that shows the first large decrease."
-#| label: fig-param-scurve
-#| out-width: 100%
-#| fig-width: 12
-#| fig-height: 10
-#| fig-pos: H
-
-mse_scurve_b + a1_m_scurve +
-  mse_scurve_lwd + prop_dens_a1 +
-  plot_layout(ncol=2)
-```
-
 #### Removing long edges
@@
 Edges define the neighbourhood structure, in order to provide a smooth \gD{} representation of the fitted model.
-@fig-scurve-true-sc shows a wire frame of the true model that was used to generate the S-curve example data.
+@fig-two_curvy-true-sc shows a wire frame of the true model that was used to generate the S-curve example data.
 The ideal is that the representation of the fitted model, at least for this example where we know the true model, should look similar to this.
@@
 Because the triangulation is typically done on the hexagon centroids, there are particular discrete edge lengths, based on bin widths.
-@fig-NLDR-scurve (d) illustrates edge length distributions.
+@fig-NLDR-two-curvy (d) illustrates edge length distributions.
 
@@
 
-<!-- (@fig-NLDR-scurve (d)), it is necessary to remove edges that connect distant bin centroids in the triangular mesh.
+<!-- (@fig-NLDR-two-curvy (d)), it is necessary to remove edges that connect distant bin centroids in the triangular mesh.
 These edges only exist in the \gD{} model and do not extend into \pD{}, so their removal does not impact the model in \pD{}.
@@
 
-1. Plot the distribution of the \gD{} Euclidean distances (@fig-param-scurve (d)).
+1. Plot the distribution of the \gD{} Euclidean distances (@fig-param-two_curvy (d)).
 2. Identify the first largest difference between consecutive distance values.
@@
 <!--distribution of distance along with the default benchmark-->
-```{r}
-#| label: triangulate-scurve2
-
-## Triangulate bin centroids
-tr1_object_scurve <- tri_bin_centroids(
-  df_bin_centroids_scurve_keep, x = "c_x", y = "c_y")
-tr_from_to_df_scurve <- gen_edges(
-  tri_object = tr1_object_scurve)
-
-trimesh_scurve <- ggplot() +
-  geom_segment(data = tr_from_to_df_scurve,
-               aes(
-                 x = x_from,
-                 y = y_from,
-                 xend = x_to,
-                 yend = y_to),
-               colour = "#33a02c") +
-  interior_annotation("a", position = c(0.92, 0.95))
-
-## Compute 2D distances
-distance_scurve <- cal_2d_dist(
-  tr_coord_df = tr_from_to_df_scurve, 
-  start_x = "x_from", 
-  start_y = "y_from", 
-  end_x = "x_to", 
-  end_y = "y_to", 
-  select_vars = c("from", "to", "distance"))
-
-## To find the benchmark value
-benchmark_scurve <- find_lg_benchmark(
-  distance_edges = distance_scurve, 
-  distance_col = "distance")
-
-## Benchmark 1
-
-distance_df_small_edges_scurve1 <- distance_scurve |>
-  filter(distance < benchmark_scurve)
-
-tr_from_to_df_scurve1 <- inner_join(
-  tr_from_to_df_scurve, distance_df_small_edges_scurve1,
-  by = c("from", "to"))
-
-trimesh_scurve_removed1 <- ggplot() +
-  geom_segment(data = tr_from_to_df_scurve1,
-               aes(
-                 x = x_from,
-                 y = y_from,
-                 xend = x_to,
-                 yend = y_to),
-               colour = "#33a02c") +
-  interior_annotation("c", position = c(0.92, 0.95))
-
-## Benchmark 2
-
-benchmark_scurve2 <- 0.6
-
-distance_df_small_edges_scurve2 <- distance_scurve |>
-  filter(distance < benchmark_scurve2)
-
-tr_from_to_df_scurve2 <- inner_join(
-  tr_from_to_df_scurve, distance_df_small_edges_scurve2,
-  by = c("from", "to"))
-
-trimesh_scurve_removed2 <- ggplot() +
-  geom_segment(data = tr_from_to_df_scurve2,
-               aes(
-                 x = x_from,
-                 y = y_from,
-                 xend = x_to,
-                 yend = y_to),
-               colour = "#33a02c") +
-  interior_annotation("b", position = c(0.92, 0.95))
-
-## To draw the distance distribution
-# distance_scurve$group <- "1"
-# distance_scurve_plot <- ggplot(distance_scurve, 
-#                         aes(x = group, 
-#                             y = distance)) + 
-#   geom_quasirandom(
-#     size = 2, 
-#     alpha = 0.3
-#   ) + 
-#   ylim(0, max(unlist(distance_scurve$distance))+ 0.5) + 
-#   coord_flip() + 
-#   geom_hline(yintercept = benchmark_scurve,
-#      linetype="solid", 
-#      color = "black", linewidth=0.5, alpha = 0.5) +
-#   ylab(expression(d^{(2)})) +
-#   ggtitle("(d)") +
-#   theme(aspect.ratio = 0.75,
-#         axis.text.x = element_text(size = 12),
-#         axis.title.x = element_text(size = 12),
-#         plot.title = element_text(size = 12))
-
-# Define bin width and starting point
-bin_width <- hb_obj_scurve2$a1
-start_point <- bin_width/2
-
-# Create bins and calculate the mean value for each bin range
-bins <- seq(start_point, max(distance_scurve$distance) + bin_width, 
-            by = bin_width)
-bin_labels <- bins[-length(bins)] + bin_width / 2  # mean of each bin range
-
-distance_scurve <- distance_scurve |> 
-  mutate(bin = cut(distance, breaks = bins, include.lowest = TRUE, labels = bin_labels))
-
-distance_hist <- ggplot(
-  distance_scurve, aes(x = distance)) +
-    geom_histogram(
-      aes(y = after_stat(count) / sum(after_stat(count))), 
-      binwidth = bin_width) +
-  scale_x_continuous(breaks = bin_labels, 
-                     labels = c(
-                       expression(a[1]),
-                       expression(2*a[1]),
-                       expression(3*a[1]),
-                       expression(4*a[1]),
-                       expression(5*a[1]),
-                       expression(6*a[1]),
-                       expression(7*a[1]),
-                       expression(8*a[1]),
-                       expression(9*a[1]),
-                       expression(10*a[1]),
-                       expression(11*a[1]),
-                       expression(12*a[1]),
-                       expression(13*a[1])
-                     )) +
-  # geom_text(
-  #   data = tibble(x = 12*bin_width, y = 0.7,
-  #                 text = expression(a[1])),
-  #   aes(
-  #     x=x,
-  #     y=y,
-  #     label = text
-  #   )
-  # 
-  # ) +
-  geom_vline(xintercept = start_point + 2*bin_width,
-             linetype="solid",
-             color = "#bdbdbd",
-             linewidth=1) +
-  ylab("Proportion of the number of edges") +
-  xlab(expression(d^{(2)})) +
-  ggtitle("(d)") +
-  theme_minimal() +
-  theme(aspect.ratio = 0.75,
-        panel.border = element_rect(fill = 'transparent'),
-        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
-        panel.grid.major.x = element_blank(),
-        axis.ticks.x = element_line(),
-        axis.ticks.y = element_line())
-
-# distance_scurve_summary <- distance_scurve |>
-#   group_by(bin) |>
-#   summarise(count = n()) |>
-#   mutate(std_counts = count/sum(count))
-
-```
-
-```{r}
-#| fig-width: 6
-#| fig-height: 4
-#| fig-cap: "long edge removal"
-
-trimesh_scurve + trimesh_scurve_removed2 + trimesh_scurve_removed1 +
-  plot_layout(ncol = 3)
-```
 
@@
 
-<!-- @fig-scurve-sc-best shows one of the best fits for S-curve data.
+<!-- @fig-two_curvy-sc-best shows one of the best fits for S-curve data.
 TriMAP with $15$ nearest neighbors for forming the nearest neighbor triplet (`n_inliers`), $3$ outliers for forming the nearest neighbor triplets (`n_outliers`), $3$ random triplets per point (`n_random`) is used as the NLDR method.
@@
 <!-- langevitour with pD model for S-curve -->
-::: {#fig-scurve-sc-best layout-ncol="4" fig-pos="H"}
+::: {#fig-two_curvy-sc-best layout-ncol="4" fig-pos="H"}
 ![](figures/scurve/tsne_trimesh_layout.png)

### 2024-09-26 paper.qmd
Updated Figure 2 (need to work on the title)

@@
 
+umap_data_two_curvy_with_hb_id <- hb_obj_two_curvy$data_hb_id
+df_all_two_curvy <- dplyr::bind_cols(training_data_two_curvy, umap_data_two_curvy_with_hb_id)
+df_b_two_curvy <- avg_highd_data(data = df_all_two_curvy, col_start = "x")
+
 hex_grid_with_counts <- 
@@
 
+distance_df_small_edges_two_curvy <- distance_two_curvy |>
+    filter(distance < benchmark_two_curvy)
+
 trimesh_removed_two_curvy <- vis_rmlg_mesh(
@@
 We would consider $T=(X_1, X_2, X_3)$ to be the geometric structure (true model) that we hope to capture.
+
+```{r}
+#| label: two-curvy-true-proj
+
+true_model_df <- read_rds("data/two_curvy_clust/two_curvy_clust_true_model.rds")
+wireframe_true_model <- read_rds("data/two_curvy_clust/two_curvy_clust_true_model_connections.rds")
+
+data_two_curvy <- training_data_two_curvy |>
+  mutate(type = "data")
+
+true_model_two_curvy <- true_model_df |> 
+    select(-ID) |> 
+  mutate(type = "true model")
+
+# Apply the scaling
+df_model_data_two_curvy <- bind_rows(data_two_curvy, true_model_two_curvy)
+
+scaled_two_curvy <- scale_data_manual(df_model_data_two_curvy, "type") |>
+  as_tibble()
+
+scaled_two_curvy <- scale_data_manual(df_model_data_two_curvy, "type") |>
+  as_tibble()
+
+scaled_two_curvy_data <- scaled_two_curvy |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_two_curvy_data_model <- scaled_two_curvy |>
+  filter(type == "true model") |>
+  select(-type)
+
+## First projection
+projection <- cbind(
+    c(-0.08707,-0.06493,0.07017,-0.07349,-0.02955,-0.05286,-0.02765),
+    c(0.03398,0.03735,0.09272,0.01871,-0.09224,0.00451,0.08079))
+
+projection_scaled <- projection * 1
+
+projected <- as.matrix(scaled_two_curvy_data) %*% projection_scaled
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  #dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+projected_model <- as.matrix(scaled_two_curvy_data_model) %*% projection_scaled
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  wireframe_true_model, 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+axes_obj <- gen_axes(  
+  proj = projection * 5,
+  limits = 0.15,
+  axis_pos_x = -0.1,
+  axis_pos_y = -0.075,
+  axis_labels = names(scaled_two_curvy_data),
+  threshold = 0.01)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
+
+two_curvy_proj_tsne_model1 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#3182bd") +
+  geom_point(
+    #size = 0.2,
+    alpha = 0.3,
+    color = "#000000") +
+  geom_segment(
+    data=axes,
+    aes(x=x1, y=y1, xend=x2, yend=y2),
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
+    colour="grey50",
+    size = 3) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  interior_annotation("a") +
+  theme(
+    legend.position = "none"
+  )
+
+## Second projection
+projection <- cbind(
+    c(-0.02704,0.08281,-0.07671,-0.02996,0.08674,0.06497,-0.02090),
+    c(0.00095,0.08031,0.05721,0.10299,0.04816,-0.05837,-0.02226))
+
+projection_scaled <- projection * 1
+
+projected <- as.matrix(scaled_two_curvy_data) %*% projection_scaled
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  #dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+projected_model <- as.matrix(scaled_two_curvy_data_model) %*% projection_scaled
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  wireframe_true_model, 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+axes_obj <- gen_axes(  
+  proj = projection * 5,
+  limits = 0.15,
+  axis_pos_x = -0.1,
+  axis_pos_y = -0.075,
+  axis_labels = names(scaled_two_curvy_data),
+  threshold = 0.01)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
+
+two_curvy_proj_tsne_model2 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#3182bd") +
+  geom_point(
+    #size = 0.2,
+    alpha = 0.3,
+    color = "#000000") +
+  geom_segment(
+    data=axes,
+    aes(x=x1, y=y1, xend=x2, yend=y2),
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
+    colour="grey50",
+    size = 3) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  interior_annotation("b") +
+  theme(
+    legend.position = "none"
+  )
+```
+
+```{r}
+#| label: fig-two-curvy-true-proj
+#| fig-cap: "Two views of the true model (blue lines) in $2\\text{-}D$ projections from $7\\text{-}D$, for the S-curve data (black points).
+The data is spread along the S shape, and does not vary much from this curve.
+(The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/I5GL23vLiw0>)."
+#| fig-pos: H
+#| fig-width: 10
+#| fig-height: 5
+
+two_curvy_proj_tsne_model1 + two_curvy_proj_tsne_model2 + 
+  plot_layout(ncol = 2)
+```
 

### 2024-09-26 paper.tex
Updated Figure 2 (need to work on the title)

@@
 
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-two-curvy-true-proj-1.pdf}
+
+}
+
+\caption{\label{fig-two-curvy-true-proj}Two views of the true model
+(blue lines) in \(2\text{-}D\) projections from \(7\text{-}D\), for the
+S-curve data (black points). The data is spread along the S shape, and
+does not vary much from this curve. (The \textbf{langevitour} software
+is used to view the data with a tour, and the full video is available at
+\url{https://youtu.be/I5GL23vLiw0}).}
+
+\end{figure}%
+
 \begin{figure}

### 2024-09-26 paper.qmd
Updated Figure 2 labels

@@
   coord_fixed() +
-  interior_annotation("a") +
+  interior_annotation("a", cex = 2) +
   theme(
@@
   coord_fixed() +
-  interior_annotation("b") +
+  interior_annotation("b", c(0.08, 0.93), cex = 2) +
   theme(

### 2024-09-26 paper.qmd
Figure 2 fixed the aspect ratio

--- 2024-09-26 paper.qmd
@@
   axis_pos_x = -0.1,
-  axis_pos_y = -0.075,
+  axis_pos_y = -0.1,
   axis_labels = names(scaled_two_curvy_data),
@@
     colour="grey50",
-    size = 3) +
+    size = 4) +
   geom_path(
@@
   coord_fixed() +
-  interior_annotation("a") +
+  xlim(c(-0.126, 0.125)) + 
+  ylim(c(-0.126, 0.125)) +
+  interior_annotation("a", cex = 2) +
   theme(
@@
   proj = projection * 5,
-  limits = 0.15,
+  limits = 0.12,
   axis_pos_x = -0.1,
-  axis_pos_y = -0.075,
+  axis_pos_y = -0.1,
   axis_labels = names(scaled_two_curvy_data),
@@
   coord_fixed() +
-  interior_annotation("b") +
+  xlim(c(-0.125, 0.075)) + 
+  ylim(c(-0.125, 0.075)) +
+  interior_annotation("b", c(0.08, 0.93), cex = 2) +
   theme(
@@
 #| label: fig-two-curvy-true-proj
-#| fig-cap: "Two views of the true model (blue lines) in $2\\text{-}D$ projections from $7\\text{-}D$, for the S-curve data (black points).
+#| fig-cap: "Two views of the true model (blue lines) in $2\\text{-}D$ projections from $7\\text{-}D$, for the two-curvy data (black points).
 The data is spread along the S shape, and does not vary much from this curve.

### 2024-09-26 paper.tex
Figure 2 fixed the aspect ratio

@@
 (blue lines) in \(2\text{-}D\) projections from \(7\text{-}D\), for the
-S-curve data (black points). The data is spread along the S shape, and
+two-curvy data (black points). The data is spread along the S shape, and
 does not vary much from this curve. (The \textbf{langevitour} software

### 2024-09-27 paper.qmd
Updated true model with less sample..for better vis

@@
 ## Compute hexbin parameters
-num_bins_x_two_curvy <- 10
+num_bins_x_two_curvy <- 6
 
@@
       yend = proj2_to), 
-    color = "#3182bd") +
+    color = "#253494") +
   geom_point(
@@
     alpha = 0.3,
+    color = "#969696") +
+  geom_segment(
+    data=axes,
+    aes(x=x1, y=y1, xend=x2, yend=y2),
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
+    colour="grey50",
+    size = 4) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.126, 0.125)) + 
+  ylim(c(-0.126, 0.125)) +
+  interior_annotation("a", cex = 2) +
+  theme(
+    legend.position = "none"
+  )
+
+proj_plot_two_curvy_true <- model_df |>
+  ggplot(
+    aes(
+      x = proj1_from, 
+      y = proj2_from)) +
+  geom_point(
     color = "#000000") +
@@
     alpha = 0.3,
-    color = "#000000") +
+    color = "#969696") +
   geom_segment(
@@
 
+```{r}
+#| echo: false
+#| label: fig-predict-two-curvy
+#| fig-pos: H
+#| fig-width: 9
+#| fig-height: 5
+#| fig-cap: "Comparison of prediction generated using the exiting `umap` prediction method and our method: (a) A view of the true model in projections from $7\\text{-}D$, (b) predicted data from the `umap` prediction method, (c) predicted data from our method, and (d) comparison of predictions from both methods.
+In plot (d), points representing the same data but predicted by different methods are connected by lines.
+Need to add concluded sentence after changing the data."
+
+# proj_plot_two_curvy_true + plot_predict_umap_two_curvy + 
+#   plot_predict_umap_two_curvy + plot_predict_umap_connect_two_curvy +
+#   plot_layout(ncol = 4)
+proj_plot_two_curvy_true
+```
+
 <!--The prediction approach involves finding the nearest \kD{} model point for a new \pD{} point.

### 2024-09-27 paper.tex
Updated true model with less sample..for better vis

@@
 
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-predict-two-curvy-1.pdf}
+
+}
+
+\caption{\label{fig-predict-two-curvy}Comparison of prediction generated
+using the exiting \texttt{umap} prediction method and our method: (a) A
+view of the true model in projections from \(7\text{-}D\), (b) predicted
+data from the \texttt{umap} prediction method, (c) predicted data from
+our method, and (d) comparison of predictions from both methods. In plot
+(d), points representing the same data but predicted by different
+methods are connected by lines. Need to add concluded sentence after
+changing the data.}
+
+\end{figure}%
+
 \subsection{Tuning}\label{tuning}

### 2024-09-27 scripts/two_curvilinear_clusters/04_gen_embeddings.R

@@ -24,8 +24,8 @@ data <- read_rds(here::here("data/two_curvy_clust/two_curvy_clust_data.rds"))
 
 ## Scale the data
 
-# data <- data |>
-#   mutate(across(everything(), ~ (. - mean(.)) / sd(.)))
+data <- data |>
+  mutate(across(everything(), ~ (. - mean(.)) / sd(.)))
 
 ## tSNE
 perplexity <- 30

### 2024-09-27 paper.qmd
Added Figure 5

@@
 counts_df <- hb_obj_two_curvy$std_cts
-df_bin_centroids <- extract_hexbin_centroids(
+df_bin_centroids_two_curvy <- extract_hexbin_centroids(
   centroids_df = all_centroids_df, 
@@
 df_all_two_curvy <- dplyr::bind_cols(training_data_two_curvy, umap_data_two_curvy_with_hb_id)
-df_b_two_curvy <- avg_highd_data(data = df_all_two_curvy, col_start = "x")
+df_bin_two_curvy <- avg_highd_data(data = df_all_two_curvy, col_start = "x")
 
@@
 hex_grid_nonempty <- hex_grid |>
-  filter(hex_poly_id %in% df_bin_centroids$hexID)
+  filter(hex_poly_id %in% df_bin_centroids_two_curvy$hexID)
 
@@
                fill = "#ffffff") +
-  geom_point(data = df_bin_centroids, 
+  geom_point(data = df_bin_centroids_two_curvy, 
              aes(x = c_x, y = c_y), 
@@
 tr1_object_two_curvy <- tri_bin_centroids(
-  df_bin_centroids, x = "c_x", y = "c_y")
+  df_bin_centroids_two_curvy, x = "c_x", y = "c_y")
 tr_from_to_df_two_curvy <- gen_edges(
@@
 
+```{r}
+#| label: model-prediction-umap-quollr
+
+true_pred_df <- predict_emb(
+  test_data = true_model_df,
+  df_bin_centroids = df_bin_centroids_two_curvy,
+  df_bin = df_bin_two_curvy,
+  type_NLDR = "UMAP"
+)
+
+# Compute radius r
+a1 <- calc_bins_y(
+  bin1 = 15, 
+  r2 = r2, 
+  q = 0.1)$a1
+
+r <- a1/2
+
+# Function to jitter points within a circumcircle of a hexagon
+jitter_within_circumcircle <- function(center_x, center_y, radius, num_points) {
+  theta <- runif(num_points, 0, 2 * pi)
+  r <- radius * sqrt(runif(num_points))
+  x <- center_x + r * cos(theta)
+  y <- center_y + r * sin(theta)
+  jittered_points <- data.frame(UMAP1 = x, UMAP2 = y, ID = points_in_hex$ID)
+  return(jittered_points)
+}
+
+# Jittered points data frame
+jittered_points_df <- data.frame()
+
+# Iterate through each hexagon and jitter points within the circumcircle
+for (hex_id in unique(true_pred_df$pred_hb_id)) {
+  # hex_points <- hex_grid_nonempty %>% filter(hex_poly_id == hex_id)
+  # center_x <- mean(hex_points$x)
+  # center_y <- mean(hex_points$y)
+  # points_in_hex <- true_pred_df %>% filter(pred_UMAP_1 >= min(hex_points$x) & pred_UMAP_1 <= max(hex_points$x) & pred_UMAP_2 >= min(hex_points$y) & pred_UMAP_2 <= max(hex_points$y))
+
+  points_in_hex <- true_pred_df |>
+    filter(pred_hb_id == hex_id)
+
+  center_x <- points_in_hex |> 
+    pull(pred_UMAP_1) |> 
+    unique()
+
+  center_y <- points_in_hex |> 
+    pull(pred_UMAP_2) |> 
+    unique()
+
+  if (nrow(points_in_hex) > 0) {
+    jittered_points <- jitter_within_circumcircle(center_x, center_y, r, nrow(points_in_hex))
+    jittered_points_df <- rbind(jittered_points_df, jittered_points)
+  }
+}
+
+plot_predict_umap_two_curvy <- ggplot(
+  data = jittered_points_df, 
+  aes(x = UMAP1,
+      y = UMAP2)) +
+  geom_point(alpha = 0.5) +
+  interior_annotation("c", c(0.92, 0.96), cex = 2)
+```
+
 <!-- XXX You need to have consistent colouring. In langevitour the points corresponding to observations are purple. They should be purple in the \gD{} view too.  -->
@@
 
+<!--prediction for true-model of two-curvy clust-->
+```{r}
+#| label: model-prediction-umap-original
+
+predict_umap_two_curvy <- read_rds(file = "data/two_curvy_clust/two_curvy_clust_umap_predict_true.rds")
+
+predict_two_curvy_obj <- gen_scaled_data(
+    data = predict_umap_two_curvy)
+
+predict_umap_two_curvy_scaled <- predict_two_curvy_obj$scaled_nldr
+
+plot_predict_umap_fun_two_curvy <- ggplot(
+  data = predict_umap_two_curvy_scaled, 
+  aes(x = UMAP1, 
+      y = UMAP2)) +
+  geom_point(alpha = 0.5) +
+  interior_annotation("b", c(0.92, 0.96), cex = 2)
+```
+
+```{r}
+#| label: model-prediction-points-line
+
+predict_umap_two_curvy_scaled <- predict_umap_two_curvy_scaled |>
+  mutate(ID = row_number()) |>
+  mutate(type = "predict_from_umap")
+
+jittered_points_df <- jittered_points_df |>
+  mutate(type = "predict_from_model")
+
+predict_point_df <- bind_rows(
+  predict_umap_two_curvy_scaled,
+  jittered_points_df
+)
+
+plot_predict_umap_connect_two_curvy <- ggplot(
+  data = predict_point_df, 
+  aes(x = UMAP1, 
+      y = UMAP2, 
+      group = ID)) +
+  geom_point(alpha = 0.5) +
+  geom_line(alpha = 0.3) +
+  interior_annotation("d", c(0.92, 0.96), cex = 2)
+```
+
 ```{r}
@@
 #| fig-pos: H
-#| fig-width: 9
-#| fig-height: 5
+#| fig-width: 10
+#| fig-height: 10
 #| fig-cap: "Comparison of prediction generated using the exiting `umap` prediction method and our method: (a) A view of the true model in projections from $7\\text{-}D$, (b) predicted data from the `umap` prediction method, (c) predicted data from our method, and (d) comparison of predictions from both methods.
@@
 
-# proj_plot_two_curvy_true + plot_predict_umap_two_curvy + 
-#   plot_predict_umap_two_curvy + plot_predict_umap_connect_two_curvy +
-#   plot_layout(ncol = 4)
-proj_plot_two_curvy_true
+proj_plot_two_curvy_true + plot_predict_umap_fun_two_curvy +plot_predict_umap_two_curvy +
+  plot_predict_umap_connect_two_curvy +
+  plot_layout(ncol = 2)
 ```

### 2024-09-27 paper.qmd
Added Figure 7 to choose te number of bins

@@
 ## Compute hexbin parameters
-num_bins_x_two_curvy <- 6
+num_bins_x_two_curvy <- 7
 
@@
   interior_annotation("b", sc_ltr_pos)
+
+hex_grid_coloured_two_curvy1 <- ggplot() + 
+  geom_polygon(
+    data = hex_grid_with_counts, 
+    aes(x = x, y = y,
+      group = hex_poly_id, 
+      fill = std_counts), color = "grey70", linewidth=0.2) +
+  geom_point(data = umap_two_curvy_scaled,
+             aes(x = UMAP1, y = UMAP2),
+             alpha = 0.3,
+             size = 0.5) +
+  scale_fill_viridis_c(direction = -1, 
+    na.value = "#ffffff", option = "E") +
+  interior_annotation("a", position = c(0.92, 0.93)) +
+  xlim(c(-0.25, 1.35)) + ylim(c(-0.35, 2))
+
 ```
@@
 ```{r}
+#| label: hexbin-regular-two-curvy2
+## hexagon binning to have regular hexagons
+hb_obj_two_curvy2 <- hex_binning(
+  data = umap_two_curvy_scaled, 
+  bin1 = 10, 
+  r2 = r2)
+
+a1_2 <- calc_bins_y(
+  bin1 = 10, 
+  r2 = r2
+  )$a1
+
+## Data set with all centroids
+all_centroids_df2 <- hb_obj_two_curvy2$centroids
+
+## Generate all coordinates of hexagons
+hex_grid2 <- hb_obj_two_curvy2$hex_poly
+
+## To obtain the standardise counts within hexbins
+counts_df2 <- hb_obj_two_curvy2$std_cts
+df_bin_centroids2 <- extract_hexbin_centroids(
+  centroids_df = all_centroids_df2, 
+  counts_df = counts_df2) |>
+  filter(drop_empty == FALSE) |>
+  mutate(b1 = "b1 = 11")
+
+hex_grid_with_counts_s_curve2 <- full_join(
+  hex_grid2, 
+  df_bin_centroids2 |> select(hexID, std_counts), 
+  by = c("hex_poly_id" = "hexID"))
+
+hex_grid_coloured_two_curvy2 <- ggplot() + 
+  geom_polygon(
+    data = hex_grid_with_counts_s_curve2, 
+    aes(x = x, y = y, 
+        group = hex_poly_id, 
+        fill = std_counts), 
+        color = "grey70", 
+        linewidth=0.2) +
+  geom_point(data = umap_two_curvy_scaled,
+           aes(x = UMAP1, y = UMAP2),
+           alpha = 0.3,
+           size = 0.5) +
+  scale_fill_viridis_c(direction = -1, 
+    na.value = "#ffffff", option = "C") +
+  xlim(c(-0.25, 1.35)) + ylim(c(-0.35, 2)) +
+  interior_annotation("b")
+```
+
+```{r}
+#| label: hexbin-regular-two-curvy3
+## hexagon binning to have regular hexagons
+hb_obj_two_curvy3 <- hex_binning(
+  data = umap_two_curvy_scaled, 
+  bin1 = 13, 
+  r2 = r2)
+
+a1_3 <- calc_bins_y(
+  bin1 = 13, 
+  r2 = r2
+)$a1
+
+## Data set with all centroids
+all_centroids_df3 <- hb_obj_two_curvy3$centroids
+
+## Generate all coordinates of hexagons
+hex_grid3 <- hb_obj_two_curvy3$hex_poly
+
+## To obtain the standardise counts within hexbins
+counts_df3 <- hb_obj_two_curvy3$std_cts
+df_bin_centroids3 <- extract_hexbin_centroids(
+  centroids_df = all_centroids_df3, 
+  counts_df = counts_df3) |>
+  filter(drop_empty == FALSE) |>
+  mutate(b1 = "b1 = 14")
+
+hex_grid_with_counts_s_curve3 <- full_join(hex_grid3, 
+                                          df_bin_centroids3 |> select(hexID, std_counts), 
+                                          by = c("hex_poly_id" = "hexID"))
+
+hex_grid_coloured_two_curvy3 <-  ggplot() + 
+  geom_polygon(
+    data = hex_grid_with_counts_s_curve3, 
+    aes(x = x, y = y, 
+        group = hex_poly_id, 
+        fill = std_counts), 
+        color = "grey70", linewidth=0.2) +
+  geom_point(data = umap_two_curvy_scaled,
+           aes(x = UMAP1, y = UMAP2),
+           alpha = 0.3,
+           size = 0.5) +
+  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "D") +
+  xlim(c(-0.25, 1.35)) + ylim(c(-0.35, 2)) +
+  interior_annotation("c")
+```
+
+```{r}
 #| label: model-prediction-umap-quollr
@@
 
+```{r}
+#| label: error-two-curvy
+
+error_two_curvy_clust_umap <- read_rds("data/two_curvy_clust/error_two_curvy_clust_umap.rds")
+
+## Find the minimum MSE when have duplicate a1
+error_two_curvy_clust_umap <- error_two_curvy_clust_umap |>
+  group_by(a1) |>
+  filter(MSE == min(MSE)) |>
+  ungroup() |>
+  mutate(prop_dens = b_non_empty/side_length)
+
+base_line_prop_dens <- error_two_curvy_clust_umap |>
+  filter(a1 == min(a1)) |>
+  pull(prop_dens)
+
+error_two_curvy_clust_umap <- error_two_curvy_clust_umap |>
+  mutate(prop_comp = prop_dens/base_line_prop_dens)
+
+mse_two_curvy_b <- ggplot(error_two_curvy_clust_umap, 
+                     aes(x = a1, 
+                         y = MSE)) +
+  geom_vline(xintercept = 0.10, linetype=2,
+           color = "#54278f", linewidth=1) +
+  geom_vline(xintercept = 0.12,
+             linetype="solid",
+             color = "#006d2c",
+             linewidth=1) +
+  geom_vline(xintercept = 0.18, linetype=2,
+             color = "#08519c", linewidth=1) +
+  geom_line(linewidth = 0.5) + 
+  geom_point(size = 1) +
+  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_clust_umap$a1, 2)))[c(1, 3, 5, 6, 8, 10:12)]) +
+  labs(x = expression(paste("binwidth (", a[1], ")")), y = "MSE") +
+  ggtitle("(a)") +
+  theme_minimal() +
+  theme(aspect.ratio = 0.75,
+        panel.border = element_rect(fill = 'transparent'),
+        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+        axis.ticks.x = element_line(),
+        axis.ticks.y = element_line())
+
+prop_dens_a1 <- ggplot(error_two_curvy_clust_umap,
+                     aes(x = a1,
+                         y = prop_comp)) +
+  geom_vline(xintercept = 0.12, linetype="solid",
+             color = "#006d2c", linewidth=1) +
+  geom_vline(xintercept = 0.10, linetype=2,
+             color = "#54278f", linewidth=1) +
+  geom_vline(xintercept = 0.18, linetype=2,
+             color = "#08519c", linewidth=1) +
+  geom_point(size = 1) +
+  geom_line(linewidth = 0.05) +
+  # scale_y_continuous(breaks = sort(unique(error_two_curvy$b_non_empty))[-3]) +
+  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_clust_umap$a1, 2)))[c(1, 3, 5, 6, 8, 10:12)]) +
+  labs(x = expression(paste("binwidth (", a[1], ")")), y = paste("proportion of densities")) +
+  ggtitle("(d)") +
+  theme_minimal() +
+  theme(aspect.ratio = 0.75,
+        panel.border = element_rect(fill = 'transparent'),
+        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+        axis.ticks.x = element_line(),
+        axis.ticks.y = element_line())
+
+error_two_curvy_clust_umap <- error_two_curvy_clust_umap |>
+  mutate(prop_bins = b_non_empty/b)
+
+a1_m_two_curvy <- ggplot(error_two_curvy_clust_umap,
+                     aes(x = a1,
+                         y = prop_bins)) +
+  geom_vline(xintercept = 0.12, linetype="solid",
+             color = "#006d2c", linewidth=1) +
+  geom_vline(xintercept = 0.10, linetype=2,
+             color = "#54278f", linewidth=1) +
+  geom_vline(xintercept = 0.18, linetype=2,
+             color = "#08519c", linewidth=1) +
+  geom_point(size = 1) +
+  geom_line(linewidth = 0.05) +
+  # scale_y_continuous(breaks = sort(unique(error_two_curvy$b_non_empty))[-3]) +
+  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_clust_umap$a1, 2)))[c(1, 3, 5, 6, 8, 10:12)]) +
+  labs(x = expression(paste("binwidth (", a[1], ")")), y = expression(paste("proportion of non-empty bins ", bgroup("(", frac(m, b), ")")))) +
+  ggtitle("(b)") +
+  theme_minimal() +
+  theme(aspect.ratio = 0.75,
+        panel.border = element_rect(fill = 'transparent'),
+        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+        axis.ticks.x = element_line(),
+        axis.ticks.y = element_line())
+
+```
+
 ### Prediction into \gD{}
@@
 
+```{r}
+#| echo: false
+#| label: fig-bins-two-curvy
+#| fig-pos: H
+#| fig-cap: "Hexbin density plots of UMAP layout of the S-curve data, using three different bin inputs: (a) $b = 91 \\text{ } (7, \\text{ }13)$, (b) $b = 220 \\text{ } (11, \\text{ }20)$, and (c) $b = 312 \\text{ } (13, \\text{ }24)$.
+Color indicates standardized counts, dark indicating high count and light indicates low count.
+At the smallest bin size the data segregates into two separate groups, suggesting this is too many bins.
+Using the MSE of the model fit in $p\\text{-}D$ helps decide on a useful choice of number of bins."
+#| fig-width: 6
+#| fig-height: 4
+
+# hex_grid_coloured_two_curvy1 + 
+#   hex_grid_coloured_two_curvy2 + 
+#   hex_grid_coloured_two_curvy3 +
+#   plot_layout(guides='collect', ncol = 3,
+#               heights = c(2,1)) &
+#   theme(legend.position='none', plot.tag = element_text(size = 8))
+
+hex_grid_coloured_two_curvy1 +
+  hex_grid_coloured_two_curvy2 +
+  hex_grid_coloured_two_curvy3 +
+  plot_layout(guides='collect', ncol = 3) &
+  theme(legend.position='none', plot.tag = element_text(size = 8))
+```
+
 The default number of bins $b=b_1\times b_2$ is computed based on the sample size, by setting $b_1=n^{1/3}$, consistent with the Diaconis-Freedman rule [@freedman1981].
@@
 -->
+
+```{r}
+#| label: rm-lwd-bin-error
+
+error_rm_two_curvy <- read_rds("data/two_curvy_clust/error_rm_lwd_diff_bin.rds") |>
+  mutate(bin1  = as.factor(bin1))
+
+mse_two_curvy_lwd <- ggplot(error_rm_two_curvy, 
+                     aes(x = benchmark_rm_lwd, 
+                         y = MSE,
+                         color = bin1)) + 
+  geom_point(
+    size = 1
+    ) +
+  geom_line(
+    linewidth = 0.3
+    ) + 
+  # geom_vline(xintercept = benchmark1, linetype="solid", 
+  #            color = "#bdbdbd", linewidth=1, alpha = 0.5) +
+  # scale_x_continuous("standardized bin count", 
+  #        transform = "reverse") + 
+  scale_x_continuous(breaks = unique(error_rm_two_curvy$benchmark_rm_lwd)) +
+  scale_color_manual(values=c("#08519c", "#006d2c", "#54278f")) +
+  xlab("threshold to remove low-density hexagons") +
+  ylab("MSE") +
+  ggtitle("(c)") +
+  theme_minimal() +
+  theme(aspect.ratio = 0.75,
+        panel.border = element_rect(fill = 'transparent'),
+        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+        axis.ticks.x = element_line(),
+        axis.ticks.y = element_line(),
+        legend.position = "none")
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "Various plots to help assess best hexagon bin parameters, thresholds to remove low density bins and large edges.
+Both (b) and (c) show MSE, against number of bins along the x-axis and standardised count.
+A good benchmark value for these parameters is when the MSE drops and then flattens out.
+Plot (a) shows the distribution of stadardised counts of hexagons.
+Plot (d) shows the distribution of $2\\text{-}D$ Euclidean distances between bin centroids, with a good benchmark value for removing large edges would being the distance that shows the first large decrease."
+#| label: fig-param-two-curvy
+#| out-width: 100%
+#| fig-width: 12
+#| fig-height: 10
+#| fig-pos: H
+
+mse_two_curvy_b + a1_m_two_curvy +
+  mse_two_curvy_lwd + prop_dens_a1 +
+  plot_layout(ncol=2)
+```
 

### 2024-09-27 paper.tex
Added Figure 7 to choose te number of bins

@@
 
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-bins-two-curvy-1.pdf}
+
+}
+
+\caption{\label{fig-bins-two-curvy}Hexbin density plots of UMAP layout
+of the S-curve data, using three different bin inputs: (a)
+\(b = 91 \text{ } (7, \text{ }13)\), (b)
+\(b = 220 \text{ } (11, \text{ }20)\), and (c)
+\(b = 312 \text{ } (13, \text{ }24)\). Color indicates standardized
+counts, dark indicating high count and light indicates low count. At the
+smallest bin size the data segregates into two separate groups,
+suggesting this is too many bins. Using the MSE of the model fit in
+\(p\text{-}D\) helps decide on a useful choice of number of bins.}
+
+\end{figure}%
+
 The default number of bins \(b=b_1\times b_2\) is computed based on the
@@
 
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-param-two-curvy-1.pdf}
+
+}
+
+\caption{\label{fig-param-two-curvy}Various plots to help assess best
+hexagon bin parameters, thresholds to remove low density bins and large
+edges. Both (b) and (c) show MSE, against number of bins along the
+x-axis and standardised count. A good benchmark value for these
+parameters is when the MSE drops and then flattens out. Plot (a) shows
+the distribution of stadardised counts of hexagons. Plot (d) shows the
+distribution of \(2\text{-}D\) Euclidean distances between bin
+centroids, with a good benchmark value for removing large edges would
+being the distance that shows the first large decrease.}
+
+\end{figure}%
+
 \subsubsection{Removing long edges}\label{removing-long-edges}

### 2024-09-27 paper.qmd
Added error plot

@@
 
+umap_data_two_curvy2_with_hb_id <- hb_obj_two_curvy2$data_hb_id
+df_all_two_curvy2 <- dplyr::bind_cols(training_data_two_curvy, umap_data_two_curvy2_with_hb_id)
+df_bin_two_curvy2 <- avg_highd_data(data = df_all_two_curvy2, col_start = "x")
+
 hex_grid_coloured_two_curvy2 <- ggplot() + 
@@
 
+training_data_two_curvy <- training_data_two_curvy |>
+  mutate(ID = row_number())
+
+## Compute error
+error_df_two_curvy_abs <- augment(
+  df_bin_centroids = df_bin_centroids2,
+  df_bin = df_bin_two_curvy2,
+  training_data = training_data_two_curvy,
+  newdata = NULL,
+  type_NLDR = "UMAP",
+  col_start = "x")
+
+error_df_two_curvy_abs <- error_df_two_curvy_abs |>
+  mutate(sqrt_row_wise_abs_error = sqrt(row_wise_abs_error))
+
+error_df_two_curvy_abs <- error_df_two_curvy_abs |>
+  bind_cols(umap_two_curvy_scaled |>
+              select(-ID))
+
+error_plot_two_curvy <- error_df_two_curvy_abs |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2,
+             colour = sqrt_row_wise_abs_error)) +
+  geom_point(alpha=0.5) +
+  scale_colour_continuous_sequential(palette = "Light Grays") +
+  coord_equal()
+
+```
+
+```{r}
+#| label: fig-p-d-error-in-2d-two-curvy
+#| fig-cap: "error"
+#| out-height: 30%
+#| fig-pos: H
+
+error_plot_two_curvy
 ```

### 2024-09-27 paper.qmd
Added long edge plots

@@
   interior_annotation("b")
+
+## Triangulate bin centroids
+tr1_object_two_curvy2 <- tri_bin_centroids(
+  df_bin_centroids2, x = "c_x", y = "c_y")
+tr_from_to_df_two_curvy2 <- gen_edges(
+  tri_object = tr1_object_two_curvy2)
+
+trimesh_two_curvy2 <- ggplot() +
+  geom_segment(data = tr_from_to_df_two_curvy2,
+               aes(
+                 x = x_from,
+                 y = y_from,
+                 xend = x_to,
+                 yend = y_to),
+               colour = "#33a02c") +
+  interior_annotation("a", position = c(0.92, 0.95))
+
+## Compute 2D distances
+distance_two_curvy <- cal_2d_dist(
+  tr_coord_df = tr_from_to_df_two_curvy2, 
+  start_x = "x_from", 
+  start_y = "y_from", 
+  end_x = "x_to", 
+  end_y = "y_to", 
+  select_vars = c("from", "to", "distance"))
+
+## To find the benchmark value
+benchmark_two_curvy <- find_lg_benchmark(
+  distance_edges = distance_two_curvy, 
+  distance_col = "distance")
+
+## Benchmark 1
+
+distance_df_small_edges_two_curvy2 <- distance_two_curvy |>
+  filter(distance < benchmark_two_curvy) ## 0.231
+
+tr_from_to_df_two_curvy3 <- inner_join(
+  tr_from_to_df_two_curvy2, distance_df_small_edges_two_curvy2,
+  by = c("from", "to"))
+
+trimesh_two_curvy_removed1 <- ggplot() +
+  geom_segment(data = tr_from_to_df_two_curvy3,
+               aes(
+                 x = x_from,
+                 y = y_from,
+                 xend = x_to,
+                 yend = y_to),
+               colour = "#33a02c") +
+  interior_annotation("c", position = c(0.92, 0.95))
+
+## Benchmark 2
+
+benchmark_two_curvy <- 1.3
+
+distance_df_small_edges_two_curvy2 <- distance_two_curvy |>
+  filter(distance < benchmark_two_curvy)
+
+tr_from_to_df_two_curvy4 <- inner_join(
+  tr_from_to_df_two_curvy2, distance_df_small_edges_two_curvy2,
+  by = c("from", "to"))
+
+trimesh_two_curvy_removed2 <- ggplot() +
+  geom_segment(data = tr_from_to_df_two_curvy4,
+               aes(
+                 x = x_from,
+                 y = y_from,
+                 xend = x_to,
+                 yend = y_to),
+               colour = "#33a02c") +
+  interior_annotation("b", position = c(0.92, 0.95))
+
 ```
@@
   test_data = true_model_df,
-  df_bin_centroids = df_bin_centroids_two_curvy,
-  df_bin = df_bin_two_curvy,
+  df_bin_centroids = df_bin_centroids2,
+  df_bin = df_bin_two_curvy2,
   type_NLDR = "UMAP"
@@
 a1 <- calc_bins_y(
-  bin1 = 15, 
+  bin1 = 13, 
   r2 = r2, 
@@
 If this is not the case, then there are several possible actions: (1) re-do the NLDR to get a more representative layout; (2) identify the edge and remove it from the model, in \gD{} and \pD{}; (3) consider different values for the model fit, number of bins, initial bin position or removing low density bins.
+
+```{r}
+#| fig-width: 6
+#| fig-height: 4
+#| fig-cap: "long edge removal"
+
+trimesh_two_curvy2 + trimesh_two_curvy_removed2 + trimesh_two_curvy_removed1 +
+  plot_layout(ncol = 3)
+```
 

### 2024-09-27 paper.qmd
Added model initial projections (need to work on this)

@@
 df_model_data_two_curvy <- bind_rows(data_two_curvy, true_model_two_curvy)
-
-scaled_two_curvy <- scale_data_manual(df_model_data_two_curvy, "type") |>
-  as_tibble()
 
@@
 
+df_b_two_curvy2 <- df_bin_two_curvy2 |>
+  dplyr::filter(hb_id %in% df_bin_centroids2$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b_two_curvy2 <- df_b_two_curvy2[match(df_bin_centroids2$hexID, df_b_two_curvy2$hb_id),] |>
+  dplyr::select(-hb_id)
+
 hex_grid_coloured_two_curvy2 <- ggplot() + 
@@
 
+trimesh_two_curvy_removed1_with_data <- ggplot() +
+  geom_segment(data = tr_from_to_df_two_curvy3,
+               aes(
+                 x = x_from,
+                 y = y_from,
+                 xend = x_to,
+                 yend = y_to),
+               colour = "#33a02c") +
+  geom_point(data = umap_two_curvy_scaled,
+             aes(
+               x = UMAP1,
+               y = UMAP2
+             ),
+             color = "#000000",
+            alpha = 0.5) +
+  interior_annotation("b")
+
 ## Benchmark 2
@@
 
-distance_df_small_edges_two_curvy2 <- distance_two_curvy |>
+distance_df_small_edges_two_curvy3 <- distance_two_curvy |>
   filter(distance < benchmark_two_curvy)
@@
 tr_from_to_df_two_curvy4 <- inner_join(
-  tr_from_to_df_two_curvy2, distance_df_small_edges_two_curvy2,
+  tr_from_to_df_two_curvy2, distance_df_small_edges_two_curvy3,
   by = c("from", "to"))
@@
   interior_annotation("b", position = c(0.92, 0.95))
+
+```
+
+```{r}
+#| label: best-fit-two-curvy-proj
+
+# Apply the scaling
+df_model_data_two_curvy <- bind_rows(data_two_curvy, true_model_two_curvy, df_b_two_curvy2)
+
+scaled_two_curvy <- scale_data_manual(df_model_data_two_curvy, "type") |>
+  as_tibble()
+
+scaled_two_curvy_data <- scaled_two_curvy |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_two_curvy_data_true_model <- scaled_two_curvy |>
+  filter(type == "true model") |>
+  select(-type)
+
+scaled_two_curvy_data_model <- scaled_two_curvy |>
+  filter(type == "model") |>
+  select(-type)
+
+## First projection
+projection <- cbind(
+  c(-0.08707,-0.06493,0.07017,-0.07349,-0.02955,-0.05286,-0.02765),
+  c(0.03398,0.03735,0.09272,0.01871,-0.09224,0.00451,0.08079))
+
+projection_scaled <- projection * 1
+
+projected <- as.matrix(scaled_two_curvy_data) %*% projection_scaled
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+## For fitted model
+projected_model <- as.matrix(scaled_two_curvy_data_model) %*% projection_scaled
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges_two_curvy2 |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+## For true model
+
+projected_true_model <- as.matrix(scaled_two_curvy_data_true_model) %*% projection_scaled
+
+projected_true_model_df <- projected_true_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+true_model_df_proj <- dplyr::left_join(
+  wireframe_true_model, 
+  projected_true_model_df, 
+  by = c("from" = "ID"))
+
+names(true_model_df_proj)[3:NCOL(true_model_df_proj)] <- paste0(names(projected_true_model_df)[-NCOL(projected_true_model_df)], "_from")
+
+true_model_df_proj <- dplyr::left_join(true_model_df_proj, projected_true_model_df, by = c("to" = "ID"))
+names(true_model_df_proj)[(2 + NCOL(projected_true_model_df)):NCOL(true_model_df_proj)] <- paste0(names(projected_true_model_df)[-NCOL(projected_true_model_df)], "_to")
+
+axes_obj <- gen_axes(  
+  proj = projection * 5,
+  limits = 0.15,
+  axis_pos_x = -0.1,
+  axis_pos_y = -0.1,
+  axis_labels = names(scaled_two_curvy_data),
+  threshold = 0)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
+
+two_curvy_proj_umap_model <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#253494") +
+  geom_point(
+    #size = 0.2,
+    alpha = 0.3,
+    color = "#969696") +
+  geom_segment(
+    data=axes,
+    aes(x=x1, y=y1, xend=x2, yend=y2),
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
+    colour="grey50",
+    size = 4) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.126, 0.125)) + 
+  ylim(c(-0.126, 0.125)) +
+  interior_annotation("c") +
+  theme(
+    legend.position = "none"
+  )
+
+two_curvy_proj_umap_all_model <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = true_model_df_proj, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#253494") +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#253494") +
+  geom_point(
+    #size = 0.2,
+    alpha = 0.3,
+    color = "#969696") +
+  geom_segment(
+    data=axes,
+    aes(x=x1, y=y1, xend=x2, yend=y2),
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
+    colour="grey50",
+    size = 4) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.126, 0.125)) + 
+  ylim(c(-0.126, 0.125)) +
+  interior_annotation("d") +
+  theme(
+    legend.position = "none"
+  )
 
@@
 <!-- XXX Should there be a second long edge removal? You can compute the edges in \pD{} now, and if there are long edges maybe these should be pruned, and we re-plot the \gD{} view.  -->
+
+<!-- langevitour with pD model for two-curvy clusters -->
+```{r}
+#| label: best-fit
+#| fig-cap: "Best fit"
+#| fig-width: 10
+#| fig-height: 10
+
+nldr_two_curvy + 
+  trimesh_two_curvy_removed1_with_data +
+  two_curvy_proj_umap_model +
+  two_curvy_proj_umap_all_model +
+  plot_layout(ncol = 2)
+```
 
@@
 -->
-
-<!-- langevitour with pD model for S-curve -->
-::: {#fig-two_curvy-sc-best layout-ncol="4" fig-pos="H"}
-![](figures/scurve/tsne_trimesh_layout.png)
-
-![](figures/scurve/sc_tsne_best_1.png)
-
-![](figures/scurve/sc_tsne_best_2.png)
-
-![](figures/scurve/sc_tsne_best_3.png)
-
-Model in \gD{}, on the tSNE layout, and three views of the fit in projections from $7\text{-}D$, for the S-curve data ($(s_1, \ s_2) = (-0.170, \ -0.137)$, $b = 256 \  (16, \ 16)$, $m = 70$, benchmark value to remove low density hexagons is $0.250$, and benchmark value to remove large edges is $0.180$).
-MSE is $0.0432$.
-The model closely fits the shape, but it has some twists (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/ZuLdp89qJ6g>).
-:::
 

### 2024-09-27 paper.tex
Added error plot
Added long edge plots
Added model initial projections (need to work on this)

@@
 
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/best-fit-1.pdf}
+
+}
+
+\caption{Best fit}
+
+\end{figure}%
+
 \subsection{Measuring the fit}\label{sec-summary}
@@
 dimensional data of \(i^{th}\) observation in \(h^{th}\) hexagon.
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=0.3\textheight]{paper_files/figure-pdf/fig-p-d-error-in-2d-two-curvy-1.pdf}
+
+}
+
+\caption{\label{fig-p-d-error-in-2d-two-curvy}error}
+
+\end{figure}%
 
@@
 
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-34-1.pdf}
+
+}
+
+\caption{long edge removal}
+
+\end{figure}%
+
 \section{Best fit}\label{best-fit}
@@
 close clusters are universally suboptimal. Layout a is the best choice.}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\begin{minipage}{0.25\linewidth}
-\includegraphics{figures/scurve/tsne_trimesh_layout.png}\end{minipage}%
-%
-\begin{minipage}{0.25\linewidth}
-\includegraphics{figures/scurve/sc_tsne_best_1.png}\end{minipage}%
-%
-\begin{minipage}{0.25\linewidth}
-\includegraphics{figures/scurve/sc_tsne_best_2.png}\end{minipage}%
-%
-\begin{minipage}{0.25\linewidth}
-\includegraphics{figures/scurve/sc_tsne_best_3.png}\end{minipage}%
-
-\caption{\label{fig-two_curvy-sc-best}Model in \gD{}, on the tSNE
-layout, and three views of the fit in projections from \(7\text{-}D\),
-for the S-curve data (\((s_1, \ s_2) = (-0.170, \ -0.137)\),
-\(b = 256 \  (16, \ 16)\), \(m = 70\), benchmark value to remove low
-density hexagons is \(0.250\), and benchmark value to remove large edges
-is \(0.180\)). MSE is \(0.0432\). The model closely fits the shape, but
-it has some twists (The \textbf{langevitour} software is used to view
-the data with a tour, and the full video is available at
-\url{https://youtu.be/ZuLdp89qJ6g}).}
 

### 2024-09-27 paper.qmd
Updated some Figures with thir scaling

@@
 
-sc_ltr_pos <- c(0.08, 0.93)
+sc_ltr_pos <- c(0.08, 0.9)
 # sc_xlims <- c(-0.5, 1.47)
 # sc_ylims <- c(-0.32, 2.1)
-sc_xlims <- c(-0.2, 1.2)
-sc_ylims <- c(-0.2, 0.8)
+sc_xlims <- c(-0.3, 1.3)
+sc_ylims <- c(-0.3, 3.2)
 
@@
   interior_annotation("a", position = c(0.92, 0.93)) + 
-  xlim(c(-0.25, 1.35)) + ylim(c(-0.35, 2))
+  xlim(sc_xlims) + ylim(sc_ylims)
 
@@
     na.value = "#ffffff", option = "C") +
-  xlim(c(-0.25, 1.35)) + ylim(c(-0.35, 2)) +
+  xlim(sc_xlims) + ylim(sc_ylims) +
   interior_annotation("b")
@@
   scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "D") +
-  xlim(c(-0.25, 1.35)) + ylim(c(-0.35, 2)) +
+  xlim(sc_xlims) + ylim(sc_ylims) +
   interior_annotation("c")

### 2024-09-27 paper.qmd
Figure 4 fixed the grid, then according to the data this not going to change

@@
 # Code to draw illustration for notation
-start_pt <- all_centroids_df |> 
+## hexagon binning to have regular hexagons
+hb_obj_notation <- hex_binning(
+  data = umap_two_curvy_scaled, 
+  bin1 = 7, 
+  r2 = 1.6,
+  q = 0.1)
+
+## Data set with all centroids
+all_centroids_df_temp <- hb_obj_notation$centroids
+hex_grid_temp <- hb_obj_notation$hex_poly
+
+start_pt <- all_centroids_df_temp |> 
   filter(hexID == 1)
@@
                  x2min = 0,
-                 x2max = r2)
+                 x2max = 1.6)
 
@@
 
-a1 <- tibble(x = all_centroids_df$c_x[4],
-             xend = all_centroids_df$c_x[5],
-             y = all_centroids_df$c_y[21],
-             yend = all_centroids_df$c_y[21],
+a1 <- tibble(x = all_centroids_df_temp$c_x[4],
+             xend = all_centroids_df_temp$c_x[5],
+             y = all_centroids_df_temp$c_y[21],
+             yend = all_centroids_df_temp$c_y[21],
              label = expression(a[1]))
-a2 <- tibble(x = all_centroids_df$c_x[25],
-             xend = all_centroids_df$c_x[25],
-             y = all_centroids_df$c_y[25],
-             yend = all_centroids_df$c_y[33],
+a2 <- tibble(x = all_centroids_df_temp$c_x[25],
+             xend = all_centroids_df_temp$c_x[25],
+             y = all_centroids_df_temp$c_y[25],
+             yend = all_centroids_df_temp$c_y[33],
              label = expression(a[2]))
 hex_param_vis <- ggplot() + 
-    geom_polygon(data = hex_grid, 
+    geom_polygon(data = hex_grid_temp, 
                         aes(x = x, 
@@
                  color = "#bdbdbd") +
-    geom_point(data = all_centroids_df, aes(
+    geom_point(data = all_centroids_df_temp, aes(
       x = c_x, 
@@
             alpha = 0.5) +
+  #coord_equal() +
   interior_annotation("b")

### 2024-09-27 paper.qmd
Updated Figure 2 and 3 with their titles

@@
 # sc_ylims <- c(-0.32, 2.1)
-sc_xlims <- c(-0.3, 1.3)
-sc_ylims <- c(-0.3, 3.2)
+sc_xlims <- c(-0.27, 1.2)
+sc_ylims <- c(-0.2, 0.8)
 
@@
   interior_annotation("a", sc_ltr_pos)
+
+nldr_two_curvy_original <- umap_two_curvy_scaled |> 
+  ggplot(aes(x = UMAP1, y = UMAP2)) + 
+  geom_point(alpha=0.5, colour="#000000", size = 0.5) +
+  interior_annotation("a", sc_ltr_pos)
 ```
@@
 ## Compute hexbin parameters
-num_bins_x_two_curvy <- 7
+num_bins_x_two_curvy <- 9
 
@@
 
-two_curvy_proj_tsne_model1 <- projected_df |>
+two_curvy_proj_umap_model1 <- projected_df |>
   ggplot(
@@
       yend = proj2_to), 
-    color = "#253494") +
+    color = "#2b8cbe") +
   geom_point(
-    #size = 0.2,
-    alpha = 0.3,
-    color = "#969696") +
+    size = 0.5,
+    alpha = 0.5,
+    color = "#000000") +
   geom_segment(
@@
   ylim(c(-0.126, 0.125)) +
-  interior_annotation("a", cex = 2) +
+  interior_annotation("a", c(0.08, 0.93), cex = 2) +
   theme(
@@
 
-two_curvy_proj_tsne_model2 <- projected_df |>
+two_curvy_proj_umap_model2 <- projected_df |>
   ggplot(
@@
 #| fig-cap: "Two views of the true model (blue lines) in $2\\text{-}D$ projections from $7\\text{-}D$, for the two-curvy data (black points).
-The data is spread along the S shape, and does not vary much from this curve.
-(The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/I5GL23vLiw0>)."
+The data is spread along the non-linear shaped clusters, and does not vary much from the curves.
+The **langevitour** software is used to view the data with a tour, and the full video is available at <>."
 #| fig-pos: H
@@
 
-two_curvy_proj_tsne_model1 + two_curvy_proj_tsne_model2 + 
+two_curvy_proj_umap_model1 + two_curvy_proj_umap_model2 + 
   plot_layout(ncol = 2)
@@
 #| fig-cap: "Key steps for constructing the model on the UMAP layout ($k=2$): (a) data, (b) hexagon bins, (c) bin centroids, and (d) triangulated centroids. The S-curve data is shown."
-#| fig-width: 8
+#| fig-width: 12
 #| fig-height: 4
@@
   geom_point(alpha = 0.5) +
-  interior_annotation("c", c(0.92, 0.96), cex = 2)
+  interior_annotation("c", c(0.08, 0.93), cex = 2)
 ```
@@
 
-nldr_two_curvy + 
+nldr_two_curvy_original + 
   trimesh_two_curvy_removed1_with_data +
@@
   geom_point(alpha = 0.5) +
-  interior_annotation("b", c(0.92, 0.96), cex = 2)
+  interior_annotation("b", c(0.08, 0.93), cex = 2)
 ```
@@
   geom_line(alpha = 0.3) +
-  interior_annotation("d", c(0.92, 0.96), cex = 2)
+  interior_annotation("d", c(0.08, 0.93), cex = 2)
 ```

### 2024-09-27 paper.tex
Updated Figure 2 and 3 with their titles

@@
 (blue lines) in \(2\text{-}D\) projections from \(7\text{-}D\), for the
-two-curvy data (black points). The data is spread along the S shape, and
-does not vary much from this curve. (The \textbf{langevitour} software
-is used to view the data with a tour, and the full video is available at
-\url{https://youtu.be/I5GL23vLiw0}).}
+two-curvy data (black points). The data is spread along the non-linear
+shaped clusters, and does not vary much from the curves. The
+\textbf{langevitour} software is used to view the data with a tour, and
+the full video is available at \textless\textgreater.}
 

### 2024-09-27 paper.qmd
Updated Figure 4

@@
   ggplot(aes(x = UMAP1, y = UMAP2)) + 
-  geom_point(alpha=0.5, colour="#000000", size = 0.5) +
+  geom_point(alpha=0.5, colour="#000000", alpha = 0.5) +
   interior_annotation("a", sc_ltr_pos)
@@
   geom_point(
-    color = "#000000") +
+    color = "#000000",
+    alpha = 0.5) +
   geom_segment(
@@
       x = x1min  - rect_adj$x1, # 0 - 0.03, 
-      y = -0.2, 
+      y = -0.31, 
       xend = x1max - rect_adj$x1, #1 - 0.03, 
-      yend = -0.2), #-0.35),
+      yend = -0.31), #-0.35),
       arrow = arrow(length = unit(0.03, "npc"),
@@
                  color = "black")+
-    annotate("text", x=0.5, y=-0.25, 
+    annotate("text", x=0.5, y=-0.36, 
              label = expression(r[1]), color = "black") +
     geom_segment(data=d_rect, aes(
-      x = -0.2, 
+      x = -0.25, 
       y = x2min - rect_adj$x2, #0 - 0.05, 
-      xend = -0.2, 
+      xend = -0.25, 
       yend = x2max - rect_adj$x2), #r2 - 0.05),
@@
                  color = "black")+ 
-    annotate("text", x=-0.25, y=0.25, 
+    annotate("text", x=-0.3, y=0.4, 
              label = expression(r[2]), color = "black") +
@@
              color = "black", hjust=-0.2) +
-    annotate("text", x=-0.18, y=-0.2, 
+    annotate("text", x=-0.18, y=-0.22, 
       label = expression(group("(", list(s[1], s[2]), ")")),

### 2024-09-27 paper.qmd
Updated Figure 4 rectangle

@@
 #| label: fig-two-curvy-true-proj
-#| fig-cap: "Two views of the true model (blue lines) in $2\\text{-}D$ projections from $7\\text{-}D$, for the two-curvy data (black points).
+#| fig-cap: "Two views of the true model (blue lines) in $2\\text{-}D$ projections from $7\\text{-}D$, for the two-curvy clusters data (black points).
 The data is spread along the non-linear shaped clusters, and does not vary much from the curves.
@@
 #| echo: false
-#| fig-cap: "Key steps for constructing the model on the UMAP layout ($k=2$): (a) data, (b) hexagon bins, (c) bin centroids, and (d) triangulated centroids. The S-curve data is shown."
+#| fig-cap: "Key steps for constructing the model on the UMAP layout ($k=2$): (a) data, (b) hexagon bins, (c) bin centroids, and (d) triangulated centroids. The two-curvy clusters data is shown."
 #| fig-width: 12
@@
 # To move the rectangle to ignore the overlap with the centroids
-rect_adj <- tibble(x1 = 0.03, x2 = 0.03)
+# rect_adj <- tibble(x1 = 0.03, x2 = 0.03)
+rect_adj <- tibble(x1 = -0.03, x2 = 0.03)
 
@@
              color = "black", hjust=-0.2) +
-    annotate("text", x=-0.18, y=-0.22, 
+    annotate("text", x=-0.18, y=-0.24, 
       label = expression(group("(", list(s[1], s[2]), ")")),

### 2024-09-27 paper.tex
Updated Figure 4 rectangle

@@
 (blue lines) in \(2\text{-}D\) projections from \(7\text{-}D\), for the
-two-curvy data (black points). The data is spread along the non-linear
-shaped clusters, and does not vary much from the curves. The
+two-curvy clusters data (black points). The data is spread along the
+non-linear shaped clusters, and does not vary much from the curves. The
 \textbf{langevitour} software is used to view the data with a tour, and
@@
 on the UMAP layout (\(k=2\)): (a) data, (b) hexagon bins, (c) bin
-centroids, and (d) triangulated centroids. The S-curve data is shown.}
+centroids, and (d) triangulated centroids. The two-curvy clusters data
+is shown.}
 

### 2024-09-27 paper.qmd
Trying with different colors

@@
   ggplot(aes(x = UMAP1, y = UMAP2)) + 
-  geom_point(alpha=0.5, colour="#000000", alpha = 0.5) +
+  geom_point(alpha=0.5, color = "#636363", size = 0.5) +
+  xlim(sc_xlims) + ylim(sc_ylims) +
   interior_annotation("a", sc_ltr_pos)
@@
                  yend = y_to),
-               colour = "#33a02c") +
+               colour = "#000000") +
   interior_annotation("c", position = c(0.92, 0.95))
@@
                  yend = y_to),
-               colour = "#33a02c") +
+               colour = "#000000") +
   geom_point(data = umap_two_curvy_scaled,
@@
              ),
-             color = "#000000",
-            alpha = 0.5) +
-  #coord_equal() +
+             color = "#636363",
+            alpha = 0.5,
+            size = 0.5) +
+  coord_equal() +
   interior_annotation("b")
@@
   axis_labels = names(scaled_two_curvy_data),
-  threshold = 0)
+  threshold = 0.01)
 
@@
       yend = proj2_to), 
-    color = "#253494") +
+    color = "#000000") + #31a354
   geom_point(
-    #size = 0.2,
-    alpha = 0.3,
-    color = "#969696") +
+    size = 0.5,
+    alpha = 0.5,
+    color = "#636363") +
   geom_segment(
@@
       yend = proj2_to), 
-    color = "#253494") +
+    color = "#2b8cbe") +
   geom_segment(
@@
       yend = proj2_to), 
-    color = "#253494") +
+    color = "#000000") +
   geom_point(
-    #size = 0.2,
-    alpha = 0.3,
-    color = "#969696") +
+    size = 0.5,
+    alpha = 0.5,
+    color = "#636363") +
   geom_segment(

### 2024-09-28 paper.qmd
Initial work on ading best tSNE and UMAP layouts

@@
 
-hex_grid_with_counts_s_curve2 <- full_join(
+hex_grid_with_counts_two_curvy2 <- full_join(
   hex_grid2, 
@@
   geom_polygon(
-    data = hex_grid_with_counts_s_curve2, 
+    data = hex_grid_with_counts_two_curvy2, 
     aes(x = x, y = y, 
@@
 ```{r}
-#| label: best-fit
+#| label: best-fit-umap1
 #| fig-cap: "Best fit"
@@
 ```{r}
-#| label: read-two-curvilinear-clust-nldr
-# Read a variety of different NLDR representations of two_curvilinear_clust
+#| label: read-two-curvy-clust-nldr
+# Read a variety of different NLDR representations of two_curvy_clust
 # and plot them on same aspect ratio
-tsne_two_curvilinear_clust <- read_rds("data/two_curvy_clust/two_curvy_clust_tsne_perplexity_30.rds")
-
-nldr_two_curvilinear_clust1 <- tsne_two_curvilinear_clust |>
+tsne_two_curvy_clust <- read_rds("data/two_curvy_clust/two_curvy_clust_tsne_perplexity_30.rds")
+
+nldr_two_curvy_clust1 <- tsne_two_curvy_clust |>
   ggplot(aes(x = tSNE1,
@@
 
-umap_two_curvilinear_clust <- read_rds("data/two_curvy_clust/two_curvy_clust_umap_n-neigbors_15_min-dist_0.1.rds")
-
-nldr_two_curvilinear_clust2 <- umap_two_curvilinear_clust |>
+umap_two_curvy_clust <- read_rds("data/two_curvy_clust/two_curvy_clust_umap_n-neigbors_15_min-dist_0.1.rds")
+
+nldr_two_curvy_clust2 <- umap_two_curvy_clust |>
   ggplot(aes(x = UMAP1,
@@
 
-phate_two_curvilinear_clust <- read_rds("data/two_curvy_clust/two_curvy_clust_phate_knn_5.rds")
-
-nldr_two_curvilinear_clust3 <- phate_two_curvilinear_clust |>
+phate_two_curvy_clust <- read_rds("data/two_curvy_clust/two_curvy_clust_phate_knn_5.rds")
+
+nldr_two_curvy_clust3 <- phate_two_curvy_clust |>
   ggplot(aes(x = PHATE1,
@@
 
-trimap_two_curvilinear_clust <- read_rds("data/two_curvy_clust/two_curvy_clust_trimap_n-inliers_12_n-outliers_4_n-random_3.rds")
-
-nldr_two_curvilinear_clust4 <- trimap_two_curvilinear_clust |>
+trimap_two_curvy_clust <- read_rds("data/two_curvy_clust/two_curvy_clust_trimap_n-inliers_12_n-outliers_4_n-random_3.rds")
+
+nldr_two_curvy_clust4 <- trimap_two_curvy_clust |>
   ggplot(aes(x = TriMAP1,
@@
 
-pacmap_two_curvilinear_clust <- read_rds("data/two_curvy_clust/two_curvy_clust_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")
-
-nldr_two_curvilinear_clust5 <- pacmap_two_curvilinear_clust |>
+pacmap_two_curvy_clust <- read_rds("data/two_curvy_clust/two_curvy_clust_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")
+
+nldr_two_curvy_clust5 <- pacmap_two_curvy_clust |>
   ggplot(aes(x = PaCMAP1,
@@
 ```{r}
-#| label: combine-data-two_curvilinear_clust
-
-error_two_curvilinear_clust_umap <- read_rds("data/two_curvy_clust/error_two_curvy_clust_umap.rds")
-error_two_curvilinear_clust_tsne <- read_rds("data/two_curvy_clust/error_two_curvy_clust_tsne.rds")
-error_two_curvilinear_clust_phate <- read_rds("data/two_curvy_clust/error_two_curvy_clust_phate.rds")
-error_two_curvilinear_clust_trimap <- read_rds("data/two_curvy_clust/error_two_curvy_clust_trimap.rds")
-error_two_curvilinear_clust_pacmap <- read_rds("data/two_curvy_clust/error_two_curvy_clust_pacmap.rds")
-
-error_two_curvilinear_clust <- bind_rows(error_two_curvilinear_clust_umap, 
-                         error_two_curvilinear_clust_tsne,
-                         error_two_curvilinear_clust_phate,
-                         error_two_curvilinear_clust_trimap,
-                         error_two_curvilinear_clust_pacmap)
-
-error_two_curvilinear_clust <- error_two_curvilinear_clust |>
+#| label: combine-data-two_curvy_clust
+
+error_two_curvy_clust_umap <- read_rds("data/two_curvy_clust/error_two_curvy_clust_umap.rds")
+error_two_curvy_clust_tsne <- read_rds("data/two_curvy_clust/error_two_curvy_clust_tsne.rds")
+error_two_curvy_clust_phate <- read_rds("data/two_curvy_clust/error_two_curvy_clust_phate.rds")
+error_two_curvy_clust_trimap <- read_rds("data/two_curvy_clust/error_two_curvy_clust_trimap.rds")
+error_two_curvy_clust_pacmap <- read_rds("data/two_curvy_clust/error_two_curvy_clust_pacmap.rds")
+
+error_two_curvy_clust <- bind_rows(error_two_curvy_clust_umap, 
+                         error_two_curvy_clust_tsne,
+                         error_two_curvy_clust_phate,
+                         error_two_curvy_clust_trimap,
+                         error_two_curvy_clust_pacmap)
+
+error_two_curvy_clust <- error_two_curvy_clust |>
   mutate(a1 = round(a1, 2)) |>
@@
 ```{r}
-#| label: error-comp-two_curvilinear_clust
-
-error_plot_two_curvilinear_clust <- ggplot(error_two_curvilinear_clust, 
+#| label: error-comp-two_curvy_clust
+
+error_plot_two_curvy_clust <- ggplot(error_two_curvy_clust, 
                            aes(x = a1, 
@@
   #            color = "black", linewidth=0.8, alpha = 0.5) +
-  scale_x_continuous(breaks = sort(unique(error_two_curvilinear_clust$a1))[append(seq(1, length(unique(error_two_curvilinear_clust$a1)), by = 5), 24)]) +
+  scale_x_continuous(breaks = sort(unique(error_two_curvy_clust$a1))[append(seq(1, length(unique(error_two_curvy_clust$a1)), by = 5), 24)]) +
   scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a', "#ff7f00",'#a65628')) +
@@
 ```{r}
-#| fig-cap: "Assessing which of the 5 NLDR layouts on the two_curvilinear_clust digit 1 data is the better representation using MSE for varying binwidth ($a_1$).
+#| fig-cap: "Assessing which of the 5 NLDR layouts on the two_curvy_clust digit 1 data is the better representation using MSE for varying binwidth ($a_1$).
 Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-e).
@@
 Layout a is the best choice."
-#| label: fig-two_curvilinear_clust-mse
+#| label: fig-two_curvy_clust-mse
 #| fig-pos: H
@@
 
-free(error_plot_two_curvilinear_clust) + wrap_plots(nldr_two_curvilinear_clust1, nldr_two_curvilinear_clust2, nldr_two_curvilinear_clust3, 
-                                    nldr_two_curvilinear_clust4, nldr_two_curvilinear_clust5, ncol = 2)
+free(error_plot_two_curvy_clust) + wrap_plots(nldr_two_curvy_clust1, nldr_two_curvy_clust2, nldr_two_curvy_clust3, 
+                                    nldr_two_curvy_clust4, nldr_two_curvy_clust5, ncol = 2)
+```
+
+```{r}
+#| label: best-fit-tsne-two-curvy
+
+two_curvy_tsne_scaled_obj <- gen_scaled_data(
+  data = tsne_two_curvy_clust)
+
+tsne_two_curvy_scaled <- two_curvy_tsne_scaled_obj$scaled_nldr |>
+  mutate(ID = row_number())
+lim1 <- two_curvy_tsne_scaled_obj$lim1
+lim2 <- two_curvy_tsne_scaled_obj$lim2
+r2_tsne <- diff(lim2)/diff(lim1)
+
+## hexagon binning to have regular hexagons
+hb_obj_tsne_two_curvy2 <- hex_binning(
+  data = tsne_two_curvy_scaled, 
+  bin1 = 10, 
+  r2 = r2_tsne)
+
+a1_2_tsne <- calc_bins_y(
+  bin1 = 10, 
+  r2 = r2_tsne
+)$a1
+
+## Data set with all centroids
+all_centroids_df4 <- hb_obj_tsne_two_curvy2$centroids
+
+## Generate all coordinates of hexagons
+hex_grid4 <- hb_obj_tsne_two_curvy2$hex_poly
+
+## To obtain the standardise counts within hexbins
+counts_df4 <- hb_obj_tsne_two_curvy2$std_cts
+df_bin_centroids4 <- extract_hexbin_centroids(
+  centroids_df = all_centroids_df4, 
+  counts_df = counts_df4) |>
+  filter(drop_empty == FALSE) |>
+  mutate(b1 = "b1 = 11")
+
+hex_grid_with_counts_two_curvy4 <- full_join(
+  hex_grid4, 
+  df_bin_centroids4 |> select(hexID, std_counts), 
+  by = c("hex_poly_id" = "hexID"))
+
+tsne_data_two_curvy2_with_hb_id <- hb_obj_tsne_two_curvy2$data_hb_id
+df_all_two_curvy4 <- dplyr::bind_cols(training_data_two_curvy, tsne_data_two_curvy2_with_hb_id)
+df_bin_two_curvy4 <- avg_highd_data(data = df_all_two_curvy4, col_start = "x")
+
+df_b_two_curvy4 <- df_bin_two_curvy4 |>
+  dplyr::filter(hb_id %in% df_bin_centroids4$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b_two_curvy4 <- df_b_two_curvy4[match(df_bin_centroids4$hexID, df_b_two_curvy4$hb_id),] |>
+  dplyr::select(-hb_id)
+
+## Triangulate bin centroids
+tr1_object_two_curvy4 <- tri_bin_centroids(
+  df_bin_centroids4, x = "c_x", y = "c_y")
+tr_from_to_df_two_curvy4 <- gen_edges(
+  tri_object = tr1_object_two_curvy4)
+
+## Compute 2D distances
+distance_two_curvy4 <- cal_2d_dist(
+  tr_coord_df = tr_from_to_df_two_curvy4, 
+  start_x = "x_from", 
+  start_y = "y_from", 
+  end_x = "x_to", 
+  end_y = "y_to", 
+  select_vars = c("from", "to", "distance"))
+
+## To find the benchmark value
+benchmark_two_curvy4 <- find_lg_benchmark(
+  distance_edges = distance_two_curvy4, 
+  distance_col = "distance")
+
+## Benchmark 1
+
+distance_df_small_edges_two_curvy4 <- distance_two_curvy4 |>
+  filter(distance < benchmark_two_curvy4) ## 0.221
+
+tr_from_to_df_two_curvy4 <- inner_join(
+  tr_from_to_df_two_curvy4, distance_df_small_edges_two_curvy4,
+  by = c("from", "to"))
+
+trimesh_two_curvy_removed1_tsne_with_data <- ggplot() +
+  geom_segment(data = tr_from_to_df_two_curvy4,
+               aes(
+                 x = x_from,
+                 y = y_from,
+                 xend = x_to,
+                 yend = y_to),
+               colour = "#000000") +
+  geom_point(data = tsne_two_curvy_scaled,
+             aes(
+               x = tSNE1,
+               y = tSNE2
+             ),
+             color = "#636363",
+             alpha = 0.5,
+             size = 0.5) +
+  coord_equal() +
+  interior_annotation("b")
+
+```
+
+```{r}
+#| label: best-fit-tsne-two-curvy-proj
+
+# Apply the scaling
+df_model_data_two_curvy <- bind_rows(data_two_curvy, true_model_two_curvy, df_b_two_curvy4)
+
+scaled_two_curvy <- scale_data_manual(df_model_data_two_curvy, "type") |>
+  as_tibble()
+
+scaled_two_curvy_data <- scaled_two_curvy |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_two_curvy_data_true_model <- scaled_two_curvy |>
+  filter(type == "true model") |>
+  select(-type)
+
+scaled_two_curvy_data_model <- scaled_two_curvy |>
+  filter(type == "model") |>
+  select(-type)
+
+## First projection
+projection <- cbind(
+  c(-0.08707,-0.06493,0.07017,-0.07349,-0.02955,-0.05286,-0.02765),
+  c(0.03398,0.03735,0.09272,0.01871,-0.09224,0.00451,0.08079))
+
+projection_scaled <- projection * 1
+
+projected <- as.matrix(scaled_two_curvy_data) %*% projection_scaled
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+## For fitted model
+projected_model <- as.matrix(scaled_two_curvy_data_model) %*% projection_scaled
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges_two_curvy4 |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+## For true model
+
+projected_true_model <- as.matrix(scaled_two_curvy_data_true_model) %*% projection_scaled
+
+projected_true_model_df <- projected_true_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+true_model_df_proj <- dplyr::left_join(
+  wireframe_true_model, 
+  projected_true_model_df, 
+  by = c("from" = "ID"))
+
+names(true_model_df_proj)[3:NCOL(true_model_df_proj)] <- paste0(names(projected_true_model_df)[-NCOL(projected_true_model_df)], "_from")
+
+true_model_df_proj <- dplyr::left_join(true_model_df_proj, projected_true_model_df, by = c("to" = "ID"))
+names(true_model_df_proj)[(2 + NCOL(projected_true_model_df)):NCOL(true_model_df_proj)] <- paste0(names(projected_true_model_df)[-NCOL(projected_true_model_df)], "_to")
+
+axes_obj <- gen_axes(  
+  proj = projection * 5,
+  limits = 0.15,
+  axis_pos_x = -0.1,
+  axis_pos_y = -0.1,
+  axis_labels = names(scaled_two_curvy_data),
+  threshold = 0.01)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
+
+two_curvy_proj_tsne_model <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#000000") + #31a354
+  geom_point(
+    size = 0.5,
+    alpha = 0.5,
+    color = "#636363") +
+  geom_segment(
+    data=axes,
+    aes(x=x1, y=y1, xend=x2, yend=y2),
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
+    colour="grey50",
+    size = 4) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.126, 0.125)) + 
+  ylim(c(-0.126, 0.125)) +
+  interior_annotation("c") +
+  theme(
+    legend.position = "none"
+  )
+
+two_curvy_proj_tsne_all_model <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = true_model_df_proj, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#2b8cbe") +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#000000") +
+  geom_point(
+    size = 0.5,
+    alpha = 0.5,
+    color = "#636363") +
+  geom_segment(
+    data=axes,
+    aes(x=x1, y=y1, xend=x2, yend=y2),
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
+    colour="grey50",
+    size = 4) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.126, 0.125)) + 
+  ylim(c(-0.126, 0.125)) +
+  interior_annotation("d") +
+  theme(
+    legend.position = "none"
+  )
+```
+
+```{r}
+#| label: best-fit-umap-tsne
+#| fig-cap: "Best fit"
+#| fig-width: 10
+#| fig-height: 20
+
+nldr_two_curvy_original + nldr_two_curvy_original +
+  trimesh_two_curvy_removed1_with_data + trimesh_two_curvy_removed1_tsne_with_data +
+  two_curvy_proj_umap_model +
+  two_curvy_proj_tsne_model +
+  two_curvy_proj_umap_all_model +
+  two_curvy_proj_tsne_all_model +
+  plot_layout(ncol = 2)
 ```

### 2024-09-28 paper.tex
Initial work on ading best tSNE and UMAP layouts

@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/best-fit-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/best-fit-umap1-1.pdf}
 
@@
 
-\includegraphics[width=1\textwidth,height=1\textheight]{paper_files/figure-pdf/fig-two_curvilinear_clust-mse-1.pdf}
-
-}
-
-\caption{\label{fig-two_curvilinear_clust-mse}Assessing which of the 5
-NLDR layouts on the two\_curvilinear\_clust digit 1 data is the better
+\includegraphics[width=1\textwidth,height=1\textheight]{paper_files/figure-pdf/fig-two_curvy_clust-mse-1.pdf}
+
+}
+
+\caption{\label{fig-two_curvy_clust-mse}Assessing which of the 5 NLDR
+layouts on the two\_curvy\_clust digit 1 data is the better
 representation using MSE for varying binwidth (\(a_1\)). Colour used for
@@
 close clusters are universally suboptimal. Layout a is the best choice.}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/best-fit-umap-tsne-1.pdf}
+
+}
+
+\caption{Best fit}
 

### 2024-09-28 paper.qmd
Updated Figure 12

@@
 
+tsne_two_curvy_clust <- tsne_two_curvy_scaled |>
+  ggplot(aes(x = tSNE1,
+             y = tSNE2))+
+  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
+  interior_annotation("a", c(0.08, 0.93)) +
+  theme(aspect.ratio = 1)
+
 ## hexagon binning to have regular hexagons
@@
 
-nldr_two_curvy_original + nldr_two_curvy_original +
+nldr_two_curvy_original + tsne_two_curvy_clust +
   trimesh_two_curvy_removed1_with_data + trimesh_two_curvy_removed1_tsne_with_data +

### 2024-09-28 paper.qmd
Added UMAP first choose initially

@@
 
+```{r}
+#| label: two-curvy-true-proj-data
+
+true_model_df <- read_rds("data/two_curvy_clust/two_curvy_clust_true_model.rds")
+wireframe_true_model <- read_rds("data/two_curvy_clust/two_curvy_clust_true_model_connections.rds")
+
+data_two_curvy <- training_data_two_curvy |>
+  mutate(type = "data")
+
+true_model_two_curvy <- true_model_df |> 
+    select(-ID) |> 
+  mutate(type = "true model")
+```
+
 <!--UMAP applied for Two curvy data-->
@@
   geom_point(alpha=0.5, color = "#636363", size = 0.5) +
-  xlim(sc_xlims) + ylim(sc_ylims) +
+  xlim(c(-0.01, 1.1)) + ylim(c(-0.08, 0.57)) +
   interior_annotation("a", sc_ltr_pos)
@@
   interior_annotation("d", sc_ltr_pos)
+
+tr_from_to_df_two_curvy <- inner_join(
+  tr_from_to_df_two_curvy, distance_df_small_edges_two_curvy,
+  by = c("from", "to"))
+
+trimesh_two_curvy_removed1_umap_with_data <- ggplot() +
+  geom_segment(data = tr_from_to_df_two_curvy,
+               aes(
+                 x = x_from,
+                 y = y_from,
+                 xend = x_to,
+                 yend = y_to),
+               colour = "#000000") +
+  geom_point(data = umap_two_curvy_scaled,
+             aes(
+               x = UMAP1,
+               y = UMAP2
+             ),
+             color = "#636363",
+             alpha = 0.5,
+             size = 0.5) +
+  coord_equal() +
+  interior_annotation("b")
+```
+
+```{r}
+#| label: best-fit-two-curvy-proj-umap1
+
+df_bin_two_curvy <- df_bin_two_curvy |>
+  select(-hb_id) |>
+  mutate(type = "model")
+
+# Apply the scaling
+df_model_data_two_curvy <- bind_rows(data_two_curvy, true_model_two_curvy, df_bin_two_curvy)
+
+scaled_two_curvy <- scale_data_manual(df_model_data_two_curvy, "type") |>
+  as_tibble()
+
+scaled_two_curvy_data <- scaled_two_curvy |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_two_curvy_data_true_model <- scaled_two_curvy |>
+  filter(type == "true model") |>
+  select(-type)
+
+scaled_two_curvy_data_model <- scaled_two_curvy |>
+  filter(type == "model") |>
+  select(-type)
+
+## First projection
+projection <- cbind(
+  c(-0.08707,-0.06493,0.07017,-0.07349,-0.02955,-0.05286,-0.02765),
+  c(0.03398,0.03735,0.09272,0.01871,-0.09224,0.00451,0.08079))
+
+projection_scaled <- projection * 1
+
+projected <- as.matrix(scaled_two_curvy_data) %*% projection_scaled
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+## For fitted model
+projected_model <- as.matrix(scaled_two_curvy_data_model) %*% projection_scaled
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges_two_curvy |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+## For true model
+
+projected_true_model <- as.matrix(scaled_two_curvy_data_true_model) %*% projection_scaled
+
+projected_true_model_df <- projected_true_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+true_model_df_proj <- dplyr::left_join(
+  wireframe_true_model, 
+  projected_true_model_df, 
+  by = c("from" = "ID"))
+
+names(true_model_df_proj)[3:NCOL(true_model_df_proj)] <- paste0(names(projected_true_model_df)[-NCOL(projected_true_model_df)], "_from")
+
+true_model_df_proj <- dplyr::left_join(true_model_df_proj, projected_true_model_df, by = c("to" = "ID"))
+names(true_model_df_proj)[(2 + NCOL(projected_true_model_df)):NCOL(true_model_df_proj)] <- paste0(names(projected_true_model_df)[-NCOL(projected_true_model_df)], "_to")
+
+axes_obj <- gen_axes(  
+  proj = projection * 5,
+  limits = 0.15,
+  axis_pos_x = -0.1,
+  axis_pos_y = -0.1,
+  axis_labels = names(scaled_two_curvy_data),
+  threshold = 0.01)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
+
+two_curvy_proj_umap_first_model1 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#000000") + #31a354
+  geom_point(
+    size = 0.5,
+    alpha = 0.5,
+    color = "#636363") +
+  geom_segment(
+    data=axes,
+    aes(x=x1, y=y1, xend=x2, yend=y2),
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
+    colour="grey50",
+    size = 4) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.126, 0.125)) + 
+  ylim(c(-0.126, 0.125)) +
+  interior_annotation("c") +
+  theme(
+    legend.position = "none"
+  )
+
+two_curvy_proj_umap_all_model1 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = true_model_df_proj, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#2b8cbe") +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#000000") +
+  geom_point(
+    size = 0.5,
+    alpha = 0.5,
+    color = "#636363") +
+  geom_segment(
+    data=axes,
+    aes(x=x1, y=y1, xend=x2, yend=y2),
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
+    colour="grey50",
+    size = 4) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.126, 0.125)) + 
+  ylim(c(-0.126, 0.125)) +
+  interior_annotation("d") +
+  theme(
+    legend.position = "none"
+  )
+
 ```
@@
 #| label: two-curvy-true-proj
-
-true_model_df <- read_rds("data/two_curvy_clust/two_curvy_clust_true_model.rds")
-wireframe_true_model <- read_rds("data/two_curvy_clust/two_curvy_clust_true_model_connections.rds")
-
-data_two_curvy <- training_data_two_curvy |>
-  mutate(type = "data")
-
-true_model_two_curvy <- true_model_df |> 
-    select(-ID) |> 
-  mutate(type = "true model")
-
 # Apply the scaling
@@
             alpha = 0.5,
-            size = 0.5) +
+            size = 0.5
+            ) +
   coord_equal() +
+  xlim(c(-0.01, 1.1)) + ylim(c(-0.08, 0.57)) +
   interior_annotation("b")
@@
 nldr_two_curvy_original + 
-  trimesh_two_curvy_removed1_with_data +
-  two_curvy_proj_umap_model +
-  two_curvy_proj_umap_all_model +
+  trimesh_two_curvy_removed1_umap_with_data +
+  two_curvy_proj_umap_first_model1 +
+  two_curvy_proj_umap_all_model1 +
   plot_layout(ncol = 2)

### 2024-09-28 paper.qmd
Updated Figure 8

@@
     na.value = "#ffffff", option = "E") +
-  interior_annotation("a", position = c(0.92, 0.93)) + 
-  xlim(sc_xlims) + ylim(sc_ylims)
+  interior_annotation("a", sc_ltr_pos) + 
+  xlim(c(-0.3, 1.2)) + ylim(c(-0.17, 0.78))
 
@@
 #| label: best-fit-two-curvy-proj-umap1
+
+df_bin_two_curvy_temp <- df_bin_two_curvy
 
@@
   data = umap_two_curvy_scaled, 
-  bin1 = 10, 
+  bin1 = 12, 
   r2 = r2)
@@
 a1_2 <- calc_bins_y(
-  bin1 = 10, 
+  bin1 = 12, 
   r2 = r2
@@
     na.value = "#ffffff", option = "C") +
-  xlim(sc_xlims) + ylim(sc_ylims) +
-  interior_annotation("b")
+  xlim(c(-0.3, 1.2)) + ylim(c(-0.17, 0.78)) +
+  interior_annotation("b", sc_ltr_pos)
 
@@
   data = umap_two_curvy_scaled, 
-  bin1 = 13, 
+  bin1 = 14, 
   r2 = r2)
@@
 a1_3 <- calc_bins_y(
-  bin1 = 13, 
+  bin1 = 14, 
   r2 = r2
@@
   scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "D") +
-  xlim(sc_xlims) + ylim(sc_ylims) +
-  interior_annotation("c")
+  xlim(c(-0.3, 1.2)) + ylim(c(-0.17, 0.78)) +
+  interior_annotation("c", sc_ltr_pos)
 ```
@@
   test_data = true_model_df,
-  df_bin_centroids = df_bin_centroids2,
-  df_bin = df_bin_two_curvy2,
+  df_bin_centroids = df_bin_centroids_two_curvy,
+  df_bin = df_bin_two_curvy_temp,
   type_NLDR = "UMAP"
@@
 a1 <- calc_bins_y(
-  bin1 = 13, 
+  bin1 = 9, 
   r2 = r2, 
@@
 #| fig-pos: H
-#| fig-cap: "Hexbin density plots of UMAP layout of the S-curve data, using three different bin inputs: (a) $b = 91 \\text{ } (7, \\text{ }13)$, (b) $b = 220 \\text{ } (11, \\text{ }20)$, and (c) $b = 312 \\text{ } (13, \\text{ }24)$.
+#| fig-cap: "Hexbin density plots of UMAP layout of the two-curvy clusters data, using three different bin inputs: (a) $b = 63 \\text{ } (9, \\text{ }7)$, (b) $b = 108 \\text{ } (12, \\text{ }9)$, and (c) $b = 312 \\text{ } (14, \\text{ }10)$.
 Color indicates standardized counts, dark indicating high count and light indicates low count.
-At the smallest bin size the data segregates into two separate groups, suggesting this is too many bins.
-Using the MSE of the model fit in $p\\text{-}D$ helps decide on a useful choice of number of bins."
-#| fig-width: 6
-#| fig-height: 4
+Using the MSE of the model fit in $7\\text{-}D$ helps decide on a useful choice of number of bins."
+#| fig-width: 9
+#| fig-height: 2
 

### 2024-09-28 paper.tex
Added UMAP first choose initially
Updated Figure 8

@@
 \caption{\label{fig-bins-two-curvy}Hexbin density plots of UMAP layout
-of the S-curve data, using three different bin inputs: (a)
-\(b = 91 \text{ } (7, \text{ }13)\), (b)
-\(b = 220 \text{ } (11, \text{ }20)\), and (c)
-\(b = 312 \text{ } (13, \text{ }24)\). Color indicates standardized
-counts, dark indicating high count and light indicates low count. At the
-smallest bin size the data segregates into two separate groups,
-suggesting this is too many bins. Using the MSE of the model fit in
-\(p\text{-}D\) helps decide on a useful choice of number of bins.}
+of the two-curvy clusters data, using three different bin inputs: (a)
+\(b = 63 \text{ } (9, \text{ }7)\), (b)
+\(b = 108 \text{ } (12, \text{ }9)\), and (c)
+\(b = 312 \text{ } (14, \text{ }10)\). Color indicates standardized
+counts, dark indicating high count and light indicates low count. Using
+the MSE of the model fit in \(7\text{-}D\) helps decide on a useful
+choice of number of bins.}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-34-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-36-1.pdf}
 

### 2024-09-28 paper.qmd
Updated Figure 9

@@
 ## Compute hexbin parameters
-num_bins_x_two_curvy <- 9
+num_bins_x_two_curvy <- 10
 
@@
 all_centroids_df <- hb_obj_two_curvy$centroids
+
+a1_1 <- hb_obj_two_curvy$a1
 
@@
   data = umap_two_curvy_scaled, 
-  bin1 = 12, 
+  bin1 = 14, 
   r2 = r2)
@@
 a1_2 <- calc_bins_y(
-  bin1 = 12, 
+  bin1 = 14, 
   r2 = r2
@@
   data = umap_two_curvy_scaled, 
-  bin1 = 14, 
+  bin1 = 17, 
   r2 = r2)
@@
 a1_3 <- calc_bins_y(
-  bin1 = 14, 
+  bin1 = 17, 
   r2 = r2
@@
 a1 <- calc_bins_y(
-  bin1 = 9, 
+  bin1 = 10, 
   r2 = r2, 
@@
                          y = MSE)) +
-  geom_vline(xintercept = 0.10, linetype=2,
+  geom_vline(xintercept = 0.06,
+             linetype="solid",
            color = "#54278f", linewidth=1) +
-  geom_vline(xintercept = 0.12,
-             linetype="solid",
-             color = "#006d2c",
+  geom_vline(xintercept = 0.04,
+             color = "#006d2c", linetype=2,
              linewidth=1) +
-  geom_vline(xintercept = 0.18, linetype=2,
+  geom_vline(xintercept = 0.12, linetype=2,
              color = "#08519c", linewidth=1) +
@@
   geom_point(size = 1) +
-  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_clust_umap$a1, 2)))[c(1, 3, 5, 6, 8, 10:12)]) +
+  # scale_x_continuous(breaks = sort(unique(round(error_two_curvy_clust_umap$a1, 2)))[c(1, 3, 5, 6, 8, 10:12)]) +
   labs(x = expression(paste("binwidth (", a[1], ")")), y = "MSE") +
@@
                          y = prop_comp)) +
-  geom_vline(xintercept = 0.12, linetype="solid",
-             color = "#006d2c", linewidth=1) +
-  geom_vline(xintercept = 0.10, linetype=2,
+  geom_vline(xintercept = 0.06,
+             linetype="solid",
            color = "#54278f", linewidth=1) +
-  geom_vline(xintercept = 0.18, linetype=2,
+  geom_vline(xintercept = 0.04,
+             color = "#006d2c", linetype=2,
+             linewidth=1) +
+  geom_vline(xintercept = 0.12, linetype=2,
              color = "#08519c", linewidth=1) +

### 2024-09-29 paper.qmd
Updated Figure 9

@@
   data = umap_two_curvy_scaled, 
-  bin1 = 14, 
+  bin1 = 13, 
   r2 = r2)
@@
 a1_2 <- calc_bins_y(
-  bin1 = 14, 
+  bin1 = 13, 
   r2 = r2
@@
   data = umap_two_curvy_scaled, 
-  bin1 = 17, 
+  bin1 = 23, 
   r2 = r2)
@@
 a1_3 <- calc_bins_y(
-  bin1 = 17, 
+  bin1 = 23, 
   r2 = r2
@@
   group_by(a1) |>
-  filter(MSE == min(MSE)) |>
+  filter(bin1 == min(bin1)) |>
   ungroup() |>
@@
                          y = MSE)) +
+  geom_vline(xintercept = 0.09,
+             linetype="solid",
+           color = "#d95f02", linewidth=1) +
   geom_vline(xintercept = 0.06,
-             linetype="solid",
-           color = "#54278f", linewidth=1) +
-  geom_vline(xintercept = 0.04,
-             color = "#006d2c", linetype=2,
+             color = "#1b9e77", linetype=2,
              linewidth=1) +
   geom_vline(xintercept = 0.12, linetype=2,
-             color = "#08519c", linewidth=1) +
+             color = "#1f78b4", linewidth=1) +
   geom_line(linewidth = 0.5) + 
@@
                          y = prop_comp)) +
+  geom_vline(xintercept = 0.09,
+             linetype="solid",
+           color = "#d95f02", linewidth=1) +
   geom_vline(xintercept = 0.06,
-             linetype="solid",
-           color = "#54278f", linewidth=1) +
-  geom_vline(xintercept = 0.04,
-             color = "#006d2c", linetype=2,
+             color = "#1b9e77", linetype=2,
              linewidth=1) +
   geom_vline(xintercept = 0.12, linetype=2,
-             color = "#08519c", linewidth=1) +
+             color = "#1f78b4", linewidth=1) +
+  geom_line(linewidth = 0.5) +
   geom_point(size = 1) +
-  geom_line(linewidth = 0.05) +
   # scale_y_continuous(breaks = sort(unique(error_two_curvy$b_non_empty))[-3]) +
@@
                          y = prop_bins)) +
-  geom_vline(xintercept = 0.12, linetype="solid",
-             color = "#006d2c", linewidth=1) +
-  geom_vline(xintercept = 0.10, linetype=2,
-             color = "#54278f", linewidth=1) +
-  geom_vline(xintercept = 0.18, linetype=2,
-             color = "#08519c", linewidth=1) +
+  geom_vline(xintercept = 0.09,
+             linetype="solid",
+           color = "#d95f02", linewidth=1) +
+  geom_vline(xintercept = 0.06,
+             color = "#1b9e77", linetype=2,
+             linewidth=1) +
+  geom_vline(xintercept = 0.12, linetype=2,
+             color = "#1f78b4", linewidth=1) +
+  geom_line(linewidth = 0.5) +
   geom_point(size = 1) +
-  geom_line(linewidth = 0.05) +
   # scale_y_continuous(breaks = sort(unique(error_two_curvy$b_non_empty))[-3]) +
@@
   geom_line(
-    linewidth = 0.3
+    linewidth = 0.5
     ) + 
@@
   scale_x_continuous(breaks = unique(error_rm_two_curvy$benchmark_rm_lwd)) +
-  scale_color_manual(values=c("#08519c", "#006d2c", "#54278f")) +
+  scale_color_manual(values=c("#1f78b4", "#d95f02", "#1b9e77")) +
   xlab("threshold to remove low-density hexagons") +

### 2024-09-29 paper.qmd
Updated Figure 6

@@
   geom_point(size = 1) +
-  # scale_x_continuous(breaks = sort(unique(round(error_two_curvy_clust_umap$a1, 2)))[c(1, 3, 5, 6, 8, 10:12)]) +
+  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_clust_umap$a1, 2)))[c(1, 3, 5, 8, 10, 11, 12, 13, 14)]) +
   labs(x = expression(paste("binwidth (", a[1], ")")), y = "MSE") +
@@
   # scale_y_continuous(breaks = sort(unique(error_two_curvy$b_non_empty))[-3]) +
-  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_clust_umap$a1, 2)))[c(1, 3, 5, 6, 8, 10:12)]) +
+  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_clust_umap$a1, 2)))[c(1, 3, 5, 8, 10, 11, 12, 13, 14)]) +
   labs(x = expression(paste("binwidth (", a[1], ")")), y = paste("proportion of densities")) +
@@
   # scale_y_continuous(breaks = sort(unique(error_two_curvy$b_non_empty))[-3]) +
-  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_clust_umap$a1, 2)))[c(1, 3, 5, 6, 8, 10:12)]) +
+  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_clust_umap$a1, 2)))[c(1, 3, 5, 8, 10, 11, 12, 13, 14)]) +
   labs(x = expression(paste("binwidth (", a[1], ")")), y = expression(paste("proportion of non-empty bins ", bgroup("(", frac(m, b), ")")))) +
@@
   geom_point(alpha=0.5) +
-  scale_colour_continuous_sequential(palette = "Light Grays") +
+  scale_colour_continuous_sequential(palette = "Blues") +
   coord_equal()
@@
 #| label: fig-p-d-error-in-2d-two-curvy
-#| fig-cap: "error"
+#| fig-cap: "The $4\\text{-}D$ model error in \\gD{} layout.
+Color indicates square root of absolute error, dark blue indicating high error and light indicates low error.
+Most large errors are distributed near the edges of the clusters."
 #| out-height: 30%
@@
   #        transform = "reverse") + 
-  scale_x_continuous(breaks = unique(error_rm_two_curvy$benchmark_rm_lwd)) +
+  scale_x_continuous(breaks = sort(unique(error_rm_two_curvy$benchmark_rm_lwd))[c(1, 3, 5, 7, 9, 11, 13, 15, 17)]) +
   scale_color_manual(values=c("#1f78b4", "#d95f02", "#1b9e77")) +

### 2024-09-29 paper.tex
Updated Figure 6

@@
 
-\caption{\label{fig-p-d-error-in-2d-two-curvy}error}
+\caption{\label{fig-p-d-error-in-2d-two-curvy}The \(4\text{-}D\) model
+error in \gD{} layout. Color indicates square root of absolute error,
+dark blue indicating high error and light indicates low error. Most
+large errors are distributed near the edges of the clusters.}
 

### 2024-09-29 paper.qmd
Updated Figure 8

@@
   ggplot(aes(x = UMAP1, y = UMAP2)) + 
-  geom_point(alpha=0.5, colour="#000000", size = 0.5) +
+  geom_point(alpha=0.5, colour="#636363", size = 0.5) +
   xlim(sc_xlims) + ylim(sc_ylims) +
@@
   aes(x = x, y = y)) +
-  geom_polygon(color = "black", 
+  geom_polygon(color = "grey70", 
                aes(group = hex_poly_id), 
@@
              aes(x = UMAP1, y = UMAP2), 
-             alpha = 0.5, size = 0.5, color = "#000000") +
+             alpha = 0.5, size = 0.5, color = "#636363") +
   xlim(sc_xlims) + ylim(sc_ylims) +
@@
              aes(x = c_x, y = c_y), 
-             color = "#33a02c",
+             color = "#000000",
              size = 1) +
@@
 
-trimesh_removed_two_curvy <- vis_rmlg_mesh(
-  distance_edges = distance_two_curvy, 
-  benchmark_value = benchmark_two_curvy, 
-  tr_coord_df = tr_from_to_df_two_curvy, 
-  distance_col = "distance") +
-  xlim(sc_xlims) + ylim(sc_ylims) +
-  interior_annotation("d", sc_ltr_pos)
+# trimesh_removed_two_curvy <- vis_rmlg_mesh(
+#   distance_edges = distance_two_curvy, 
+#   benchmark_value = benchmark_two_curvy, 
+#   tr_coord_df = tr_from_to_df_two_curvy, 
+#   distance_col = "distance") +
+#   xlim(sc_xlims) + ylim(sc_ylims) +
+#   interior_annotation("d", sc_ltr_pos)
 
@@
   by = c("from", "to"))
+
+trimesh_removed_two_curvy <- ggplot() +
+  geom_segment(data = tr_from_to_df_two_curvy,
+               aes(
+                 x = x_from,
+                 y = y_from,
+                 xend = x_to,
+                 yend = y_to),
+               colour = "#000000") +
+  coord_equal() +
+  xlim(sc_xlims) + ylim(sc_ylims) +
+  interior_annotation("d", sc_ltr_pos)
 
@@
     alpha = 0.5,
-    color = "#000000") +
+    color = "#636363") +
   geom_segment(
@@
 #| label: fig-two-curvy-true-proj
-#| fig-cap: "Two views of the true model (blue lines) in $2\\text{-}D$ projections from $7\\text{-}D$, for the two-curvy clusters data (black points).
+#| fig-cap: "Two views of the true model (blue lines) in $2\\text{-}D$ projections from $7\\text{-}D$, for the two-curvy clusters data (grey points).
 The data is spread along the non-linear shaped clusters, and does not vary much from the curves.
@@
            color = "#d95f02", linewidth=1) +
-  geom_vline(xintercept = 0.06,
+  geom_vline(xintercept = 0.05,
              color = "#1b9e77", linetype=2,
@@
   geom_point(size = 1) +
-  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_clust_umap$a1, 2)))[c(1, 3, 5, 8, 10, 11, 12, 13, 14)]) +
+  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_clust_umap$a1, 2)))[c(1, 4, 8, 10, 11, 12, 13, 14)]) +
   labs(x = expression(paste("binwidth (", a[1], ")")), y = "MSE") +
@@
   # scale_y_continuous(breaks = sort(unique(error_two_curvy$b_non_empty))[-3]) +
-  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_clust_umap$a1, 2)))[c(1, 3, 5, 8, 10, 11, 12, 13, 14)]) +
+  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_clust_umap$a1, 2)))[c(1, 4, 8, 10, 11, 12, 13, 14)]) +
   labs(x = expression(paste("binwidth (", a[1], ")")), y = paste("proportion of densities")) +
@@
   # scale_y_continuous(breaks = sort(unique(error_two_curvy$b_non_empty))[-3]) +
-  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_clust_umap$a1, 2)))[c(1, 3, 5, 8, 10, 11, 12, 13, 14)]) +
+  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_clust_umap$a1, 2)))[c(1, 4, 8, 10, 11, 12, 13, 14)]) +
   labs(x = expression(paste("binwidth (", a[1], ")")), y = expression(paste("proportion of non-empty bins ", bgroup("(", frac(m, b), ")")))) +
@@
 #| label: fig-p-d-error-in-2d-two-curvy
-#| fig-cap: "The $4\\text{-}D$ model error in \\gD{} layout.
+#| fig-cap: "The $7\\text{-}D$ model error in \\gD{} layout.
 Color indicates square root of absolute error, dark blue indicating high error and light indicates low error.
@@
 #| fig-pos: H
-#| fig-cap: "Hexbin density plots of UMAP layout of the two-curvy clusters data, using three different bin inputs: (a) $b = 63 \\text{ } (9, \\text{ }7)$, (b) $b = 108 \\text{ } (12, \\text{ }9)$, and (c) $b = 312 \\text{ } (14, \\text{ }10)$.
+#| fig-cap: "Hexbin density plots of UMAP layout of the two-curvy clusters data, using three different bin inputs: (a) $b = 80 \\text{ } (10, \\text{ }8)$, (b) $b = 130 \\text{ } (13, \\text{ }10)$, and (c) $b = 391 \\text{ } (23, \\text{ }17)$.
 Color indicates standardized counts, dark indicating high count and light indicates low count.

### 2024-09-29 paper.tex
Updated Figure 8

@@
 (blue lines) in \(2\text{-}D\) projections from \(7\text{-}D\), for the
-two-curvy clusters data (black points). The data is spread along the
+two-curvy clusters data (grey points). The data is spread along the
 non-linear shaped clusters, and does not vary much from the curves. The
@@
 
-\caption{\label{fig-p-d-error-in-2d-two-curvy}The \(4\text{-}D\) model
+\caption{\label{fig-p-d-error-in-2d-two-curvy}The \(7\text{-}D\) model
 error in \gD{} layout. Color indicates square root of absolute error,
@@
 of the two-curvy clusters data, using three different bin inputs: (a)
-\(b = 63 \text{ } (9, \text{ }7)\), (b)
-\(b = 108 \text{ } (12, \text{ }9)\), and (c)
-\(b = 312 \text{ } (14, \text{ }10)\). Color indicates standardized
+\(b = 80 \text{ } (10, \text{ }8)\), (b)
+\(b = 130 \text{ } (13, \text{ }10)\), and (c)
+\(b = 391 \text{ } (23, \text{ }17)\). Color indicates standardized
 counts, dark indicating high count and light indicates low count. Using

### 2024-09-29 paper.qmd
Updated Figure 10

@@
                  yend = y_to),
-               colour = "#33a02c") +
+               colour = "#000000") +
   interior_annotation("a", position = c(0.92, 0.95))
@@
 
-benchmark_two_curvy <- 1.3
+benchmark_two_curvy <- 0.42
 
@@
                  yend = y_to),
-               colour = "#33a02c") +
+               colour = "#000000") +
   interior_annotation("b", position = c(0.92, 0.95))
@@
 ```{r}
-#| fig-width: 6
-#| fig-height: 4
+#| fig-width: 9
+#| fig-height: 2
 #| fig-cap: "long edge removal"

### 2024-09-29 paper.qmd
Changed the abstract (worked with Thiyanga's comments also)

@@
 abstract: |
-  Non-linear dimension reduction (NLDR) techniques such as tSNE, and UMAP provide a low-dimensional representation of high-dimensional ($p\text{-}D$) data using non-linear transformation.
-The methods and parameter choices can create wildly different representations, making it difficult to decide which is best, or whether any or all are accurate or misleading.
+  Non-linear dimension reduction (NLDR) techniques such as tSNE, and UMAP provide a low-dimensional representation of high-dimensional data ($p\text{-}D$) by applying a non-linear transformation.
 NLDR often exaggerates random patterns, sometimes due to the samples observed.
 But NLDR views have an important role in data analysis because, if done well, they provide a concise visual (and conceptual) summary of $p\text{-}D$ distributions.
-To help evaluate the NLDR we have developed an algorithm to show the $2\text{-}D$ NLDR model in the $p\text{-}D$ space, viewed with a tour.
-One can see if the model fits everywhere or better in some subspaces, or completely mismatches the data.
-It is used to evaluate which $2\text{-}D$ layout is the best representation of the $p\text{-}D$ distribution and see how different methods may have similar summaries or quirks.
-
-keywords: [high-dimensional data, dimension reduction, hexagon binning, low-dimensional representation, tour, data vizualization, model-in-the-data-space]
-keywords-formatted: [high-dimensional data, dimension reduction, hexagon binning, low-dimensional representation, tour, data vizualization, model-in-the-data-space]
+The NLDR methods and (hyper)parameter choices can create wildly different representations, making it difficult to decide which is best, or whether any or all are accurate or misleading.
+To help assess the NLDR and decide on which, if any, is the most reasonable representation of the structure(s) present in the $p\text{-}D$ data, we have developed an algorithm to show the \gD{} NLDR model in the $p\text{-}D$ space, viewed with a tour, a movie of linear projections.
+From this, one can see if the model fits everywhere, or better in some subspaces, or completely mismatches the data.
+Also, we can see how different methods may have similar summaries or quirks.
+
+keywords: [high-dimensional data vizualization, non-linear dimension reduction, tour]
+keywords-formatted: [high-dimensional data vizualization, non-linear dimension reduction, tour]
 
@@
                colour = "#000000") +
-  interior_annotation("a", position = c(0.92, 0.95))
+  interior_annotation("a")
 
@@
   coord_equal() +
-  xlim(c(-0.01, 1.1)) + ylim(c(-0.08, 0.57)) +
-  interior_annotation("b")
+  #xlim(c(-0.01, 1.1)) + ylim(c(-0.08, 0.57)) +
+  interior_annotation("a2")
 
@@
 #| label: best-fit-umap1
-#| fig-cap: "Best fit"
+#| fig-cap: "Model in \\gD{}, on the layout and two views of the fit in projections from $7\\text{-}D$, for the two-curvy clusters data ($n =  2000$ and $p = 7$). The fitted model"
 #| fig-width: 10
@@
 
-tsne_two_curvy_clust <- tsne_two_curvy_scaled |>
+tsne_two_curvy_clust_plt <- tsne_two_curvy_scaled |>
   ggplot(aes(x = tSNE1,
@@
   data = tsne_two_curvy_scaled, 
-  bin1 = 10, 
+  bin1 = 8, 
   r2 = r2_tsne)
@@
 a1_2_tsne <- calc_bins_y(
-  bin1 = 10, 
+  bin1 = 8, 
   r2 = r2_tsne
@@
   distance_col = "distance")
+
+benchmark_two_curvy4 <- 0.2
 
@@
 
-nldr_two_curvy_original + tsne_two_curvy_clust +
+nldr_two_curvy_original + tsne_two_curvy_clust_plt +
   trimesh_two_curvy_removed1_with_data + trimesh_two_curvy_removed1_tsne_with_data +

### 2024-09-29 paper.tex
Changed the abstract (worked with Thiyanga's comments also)

@@
   pdfauthor={Jayani P.G. Lakshika; Dianne Cook; Paul Harrison; Michael Lydeamore; Thiyanga S. Talagala},
-  pdfkeywords={high-dimensional data, dimension reduction, hexagon
-binning, low-dimensional representation, tour, data
-vizualization, model-in-the-data-space},
+  pdfkeywords={high-dimensional data vizualization, non-linear dimension
+reduction, tour},
   colorlinks=true,
@@
 Non-linear dimension reduction (NLDR) techniques such as tSNE, and UMAP
-provide a low-dimensional representation of high-dimensional
-(\(p\text{-}D\)) data using non-linear transformation. The methods and
-parameter choices can create wildly different representations, making it
-difficult to decide which is best, or whether any or all are accurate or
-misleading. NLDR often exaggerates random patterns, sometimes due to the
-samples observed. But NLDR views have an important role in data analysis
-because, if done well, they provide a concise visual (and conceptual)
-summary of \(p\text{-}D\) distributions. To help evaluate the NLDR we
-have developed an algorithm to show the \(2\text{-}D\) NLDR model in the
-\(p\text{-}D\) space, viewed with a tour. One can see if the model fits
-everywhere or better in some subspaces, or completely mismatches the
-data. It is used to evaluate which \(2\text{-}D\) layout is the best
-representation of the \(p\text{-}D\) distribution and see how different
-methods may have similar summaries or quirks.
+provide a low-dimensional representation of high-dimensional data
+(\(p\text{-}D\)) by applying a non-linear transformation. NLDR often
+exaggerates random patterns, sometimes due to the samples observed. But
+NLDR views have an important role in data analysis because, if done
+well, they provide a concise visual (and conceptual) summary of
+\(p\text{-}D\) distributions. The NLDR methods and (hyper)parameter
+choices can create wildly different representations, making it difficult
+to decide which is best, or whether any or all are accurate or
+misleading. To help assess the NLDR and decide on which, if any, is the
+most reasonable representation of the structure(s) present in the
+\(p\text{-}D\) data, we have developed an algorithm to show the \gD{}
+NLDR model in the \(p\text{-}D\) space, viewed with a tour, a movie of
+linear projections. From this, one can see if the model fits everywhere,
+or better in some subspaces, or completely mismatches the data. Also, we
+can see how different methods may have similar summaries or quirks.
 \end{abstract}
@@
 \noindent%
-{\it Keywords:} high-dimensional data, dimension reduction, hexagon
-binning, low-dimensional representation, tour, data
-vizualization, model-in-the-data-space
+{\it Keywords:} high-dimensional data vizualization, non-linear
+dimension reduction, tour
 \vfill
@@
 
-\caption{Best fit}
+\caption{Model in \gD{}, on the layout and two views of the fit in
+projections from \(7\text{-}D\), for the two-curvy clusters data
+(\(n =  2000\) and \(p = 7\)). The fitted model}
 

### 2024-09-29 paper.qmd
Added the Figure 5 title

@@
 abstract: |
-  Non-linear dimension reduction (NLDR) techniques such as tSNE, and UMAP provide a low-dimensional representation of high-dimensional data ($p\text{-}D$) by applying a non-linear transformation.
+  Non-linear dimension reduction (NLDR) techniques such as tSNE, and UMAP provide a low-dimensional representation of high-dimensional data (\pD{}) by applying a non-linear transformation.
 NLDR often exaggerates random patterns, sometimes due to the samples observed.
-But NLDR views have an important role in data analysis because, if done well, they provide a concise visual (and conceptual) summary of $p\text{-}D$ distributions.
+But NLDR views have an important role in data analysis because, if done well, they provide a concise visual (and conceptual) summary of \pD{} distributions.
 The NLDR methods and (hyper)parameter choices can create wildly different representations, making it difficult to decide which is best, or whether any or all are accurate or misleading.
-To help assess the NLDR and decide on which, if any, is the most reasonable representation of the structure(s) present in the $p\text{-}D$ data, we have developed an algorithm to show the \gD{} NLDR model in the $p\text{-}D$ space, viewed with a tour, a movie of linear projections.
+To help assess the NLDR and decide on which, if any, is the most reasonable representation of the structure(s) present in the \pD{} data, we have developed an algorithm to show the \gD{} NLDR model in the \pD{} space, viewed with a tour, a movie of linear projections.
 From this, one can see if the model fits everywhere, or better in some subspaces, or completely mismatches the data.
@@
 
-Non-linear dimension reduction (NLDR) is popular for making a convenient low-dimensional (\kD{}) representation of high-dimensional (\pD{}) data.
+Non-linear dimension reduction (NLDR) is popular for making a convenient low-dimensional (\kD{}) representation of high-dimensional (\pD{}) data ($k < p$).
 Recently developed methods include t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2022], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021].
@@
 
-The paper is organised as follows.
+The paper is organized as follows.
 @sec-background provides a summary of the literature on NLDR, and high-dimensional data visualization methods.
@@
 #| label: best-fit-umap1
-#| fig-cap: "Model in \\gD{}, on the layout and two views of the fit in projections from $7\\text{-}D$, for the two-curvy clusters data ($n =  2000$ and $p = 7$). The fitted model"
+#| fig-cap: "Model in \\gD{} ($a_1 = 0.12$), on the layout and two views of the fit in projections from $7\\text{-}D$, for the two-curvy clusters data ($n =  2000$ and $p = 7$).
+One view shows the fitted model, and the other shows the fitted and true models alongside the data.
+The fitted model accurately represents the shape of the clusters but does not capture the edges of the structure.
+This illustrates a characteristic of UMAP, which compresses the data when transforming the data into \\gD{}."
 #| fig-width: 10

### 2024-09-29 paper.tex
Added the Figure 5 title

@@
 provide a low-dimensional representation of high-dimensional data
-(\(p\text{-}D\)) by applying a non-linear transformation. NLDR often
-exaggerates random patterns, sometimes due to the samples observed. But
-NLDR views have an important role in data analysis because, if done
-well, they provide a concise visual (and conceptual) summary of
-\(p\text{-}D\) distributions. The NLDR methods and (hyper)parameter
-choices can create wildly different representations, making it difficult
-to decide which is best, or whether any or all are accurate or
-misleading. To help assess the NLDR and decide on which, if any, is the
-most reasonable representation of the structure(s) present in the
-\(p\text{-}D\) data, we have developed an algorithm to show the \gD{}
-NLDR model in the \(p\text{-}D\) space, viewed with a tour, a movie of
-linear projections. From this, one can see if the model fits everywhere,
-or better in some subspaces, or completely mismatches the data. Also, we
-can see how different methods may have similar summaries or quirks.
+(\pD{}) by applying a non-linear transformation. NLDR often exaggerates
+random patterns, sometimes due to the samples observed. But NLDR views
+have an important role in data analysis because, if done well, they
+provide a concise visual (and conceptual) summary of \pD{}
+distributions. The NLDR methods and (hyper)parameter choices can create
+wildly different representations, making it difficult to decide which is
+best, or whether any or all are accurate or misleading. To help assess
+the NLDR and decide on which, if any, is the most reasonable
+representation of the structure(s) present in the \pD{} data, we have
+developed an algorithm to show the \gD{} NLDR model in the \pD{} space,
+viewed with a tour, a movie of linear projections. From this, one can
+see if the model fits everywhere, or better in some subspaces, or
+completely mismatches the data. Also, we can see how different methods
+may have similar summaries or quirks.
 \end{abstract}
@@
 Non-linear dimension reduction (NLDR) is popular for making a convenient
-low-dimensional (\kD{}) representation of high-dimensional (\pD{}) data.
-Recently developed methods include t-distributed stochastic neighbor
-embedding (tSNE) \citep{laurens2008}, uniform manifold approximation and
-projection (UMAP) \citep{leland2018}, potential of heat-diffusion for
-affinity-based trajectory embedding (PHATE) algorithm \citep{moon2019},
-large-scale dimensionality reduction Using triplets (TriMAP)
-\citep{amid2022}, and pairwise controlled manifold approximation
-(PaCMAP) \citep{yingfan2021}. However, the representation generated can
-vary dramatically from method to method, and with different choices of
-parameters or random seeds made using the same method
-(Figure~\ref{fig-NLDR-variety}). The dilemma for the analyst is then,
-\textbf{which representation to use}. The choice might result in
+low-dimensional (\kD{}) representation of high-dimensional (\pD{}) data
+(\(k < p\)). Recently developed methods include t-distributed stochastic
+neighbor embedding (tSNE) \citep{laurens2008}, uniform manifold
+approximation and projection (UMAP) \citep{leland2018}, potential of
+heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm
+\citep{moon2019}, large-scale dimensionality reduction Using triplets
+(TriMAP) \citep{amid2022}, and pairwise controlled manifold
+approximation (PaCMAP) \citep{yingfan2021}. However, the representation
+generated can vary dramatically from method to method, and with
+different choices of parameters or random seeds made using the same
+method (Figure~\ref{fig-NLDR-variety}). The dilemma for the analyst is
+then, \textbf{which representation to use}. The choice might result in
 different procedures used in the downstream analysis, or different
@@
 
-The paper is organised as follows. Section~\ref{sec-background} provides
+The paper is organized as follows. Section~\ref{sec-background} provides
 a summary of the literature on NLDR, and high-dimensional data
@@
 
-\caption{Model in \gD{}, on the layout and two views of the fit in
-projections from \(7\text{-}D\), for the two-curvy clusters data
-(\(n =  2000\) and \(p = 7\)). The fitted model}
+\caption{Model in \gD{} (\(a_1 = 0.12\)), on the layout and two views of
+the fit in projections from \(7\text{-}D\), for the two-curvy clusters
+data (\(n =  2000\) and \(p = 7\)). One view shows the fitted model, and
+the other shows the fitted and true models alongside the data. The
+fitted model accurately represents the shape of the clusters but does
+not capture the edges of the structure. This illustrates a
+characteristic of UMAP, which compresses the data when transforming the
+data into \gD{}.}
 

### 2024-09-30 paper.qmd
Updated Figure 7 title

@@
 The fitted model accurately represents the shape of the clusters but does not capture the edges of the structure.
-This illustrates a characteristic of UMAP, which compresses the data when transforming the data into \\gD{}."
+This illustrates a characteristic of UMAP, which compresses the data when transforming the data into \\gD{}.
+Video of the langevitour animations is available at <>."
 #| fig-width: 10
@@
 #| fig-height: 10
-#| fig-cap: "Comparison of prediction generated using the exiting `umap` prediction method and our method: (a) A view of the true model in projections from $7\\text{-}D$, (b) predicted data from the `umap` prediction method, (c) predicted data from our method, and (d) comparison of predictions from both methods.
+#| fig-cap: "Comparison of prediction generated using the exiting `umap` R package's prediction method and our method: (a) A view of the true model in projection from $7\\text{-}D$, (b) predicted data from the `umap` prediction method, (c) predicted data from our method, and (d) comparison of predictions from both methods.
 In plot (d), points representing the same data but predicted by different methods are connected by lines.
-Need to add concluded sentence after changing the data."
+Both prediction methods work similarly within the clusters, as each method predicts the data within the cluster with slight differences."
 

### 2024-09-30 paper.tex
Updated Figure 7 title

@@
 characteristic of UMAP, which compresses the data when transforming the
-data into \gD{}.}
+data into \gD{}. Video of the langevitour animations is available at
+\textless\textgreater.}
 
@@
 \caption{\label{fig-predict-two-curvy}Comparison of prediction generated
-using the exiting \texttt{umap} prediction method and our method: (a) A
-view of the true model in projections from \(7\text{-}D\), (b) predicted
-data from the \texttt{umap} prediction method, (c) predicted data from
-our method, and (d) comparison of predictions from both methods. In plot
-(d), points representing the same data but predicted by different
-methods are connected by lines. Need to add concluded sentence after
-changing the data.}
+using the exiting \texttt{umap} R package's prediction method and our
+method: (a) A view of the true model in projection from \(7\text{-}D\),
+(b) predicted data from the \texttt{umap} prediction method, (c)
+predicted data from our method, and (d) comparison of predictions from
+both methods. In plot (d), points representing the same data but
+predicted by different methods are connected by lines. Both prediction
+methods work similarly within the clusters, as each method predicts the
+data within the cluster with slight differences.}
 

### 2024-09-30 paper.qmd
Updated Figure 10

@@
 
+## To plot the distribution of distances
+
+# Define bin width and starting point
+bin_width <- hb_obj_two_curvy2$a1
+start_point <- bin_width/2
+
+# Create bins and calculate the mean value for each bin range
+bins <- seq(start_point, max(distance_two_curvy$distance) + bin_width, 
+            by = bin_width)
+bin_labels <- bins[-length(bins)] + bin_width / 2  # mean of each bin range
+
+distance_two_curvy_dist <- distance_two_curvy |> 
+  mutate(bin = cut(distance, breaks = bins, include.lowest = TRUE, labels = bin_labels)) |>
+  dplyr::filter(distance > 0.1) ## To remove a1 distances, since always small
+
+distance_hist <- ggplot(
+  distance_two_curvy_dist, aes(x = distance)) +
+    geom_histogram(
+      aes(y = after_stat(count) / sum(after_stat(count))), 
+      binwidth = bin_width) +
+  scale_x_continuous(breaks = bin_labels, 
+                     labels = c(
+                       expression(a[1]),
+                       expression(2*a[1]),
+                       expression(3*a[1]),
+                       expression(4*a[1]),
+                       expression(5*a[1]),
+                       expression(6*a[1]),
+                       expression(7*a[1]),
+                       expression(8*a[1])
+                     )) +
+  geom_vline(xintercept = start_point + 2*bin_width,
+             linetype="solid",
+             color = "#bdbdbd",
+             linewidth=1) +
+  ylab("Proportion of the number of edges") +
+  xlab(expression(d^{(2)})) +
+  #ggtitle("(d)") +
+  theme_minimal() +
+  theme(aspect.ratio = 0.75,
+        panel.border = element_rect(fill = 'transparent'),
+        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+        panel.grid.major.x = element_blank(),
+        axis.ticks.x = element_line(),
+        axis.ticks.y = element_line())
+
 ## To find the benchmark value
@@
 ```{r}
-#| fig-width: 9
-#| fig-height: 2
-#| fig-cap: "long edge removal"
-
-trimesh_two_curvy2 + trimesh_two_curvy_removed2 + trimesh_two_curvy_removed1 +
-  plot_layout(ncol = 3)
+#| fig-cap: "Wireframes with Delaunay triangulation, and two choices of edge removal in \\gD{} and \\pD{}. "
+
+free(distance_hist) + wrap_plots(trimesh_two_curvy2, trimesh_two_curvy_removed2, trimesh_two_curvy_removed1, ncol = 2) 
 ```

### 2024-09-30 paper.qmd
Added initial writeup for the supplementary aand acknowledgments

@@
 This study makes several important contributions to the field of NLDR.
-We have developed an algorithm to evaluate the most useful NLDR method and (hyper-)parameter choices for creating a reasonable \gD{} layout of high-dimensional data.
+We have developed an algorithm to evaluate the most useful NLDR method and (hyper-)parameter choices for creating an accurate \gD{} layout of high-dimensional data.
 Our objective is to fit a model for the \gD{} layout that preserves the relationships between neighboring points and turns it into a high-dimensional wireframe, which can be overlaid on the data and visualized using a tour.
@@
 The effectiveness of this approach is illustrated through various examples.
-For instance, the S-curve example demonstrates how the model accurately fits the points, capturing both local and global structures in high-dimensional space.
+For instance, the two-curvy clusters example demonstrates how the model accurately fits the points, capturing both local and global structures in high-dimensional space.
 Our simulation case study further, five Gaussian cluster example shows that while all observed NLDR methods preserve the global structure, only tSNE effectively maintains the local structure, highlighting the specific strengths and quirks of different methods.
@@
 This emphasizes the importance of clear and distinct clusters.
-For example, in the UMAP layout of the **pbmc** dataset suggested by @chen2023, three distant, well-separated clusters are shown.
+For example, in the UMAP layout of the **PBMC3k** dataset suggested by @chen2023, three distant, well-separated clusters are shown.
 However, our model reveals that these clusters are actually close to each other in \pD{}.
@@
 
-Code, and data for reproducing this paper are available at [https://github.com/JayaniLakshika/paper-nldr-vis-algorithm](https://github.com/JayaniLakshika/paper-nldr-vis-algorithm).
+Appendix: The appendix includes more details about the hexagonal binning algorithm (appendix.pdf, Portable Document Format file).
+
+R package `quollr`: The R package `quollr` containing codes to fit, and visualize the model. (need to add quollr .zip file, GNU zipped tar file)
+
+## Acknowledgments
+
+These `R` packages were used for the work:
+The project's GitHub repository ([https://github.com/JayaniLakshika/paper-nldr-vis-algorithm](https://github.com/JayaniLakshika/paper-nldr-vis-algorithm)) contains all materials required to reproduce this article.
 

### 2024-09-30 paper.tex
Updated Figure 10
Added initial writeup for the supplementary aand acknowledgments

@@
 
-\caption{long edge removal}
+\caption{Wireframes with Delaunay triangulation, and two choices of edge
+removal in \gD{} and \pD{}.}
 
@@
 We have developed an algorithm to evaluate the most useful NLDR method
-and (hyper-)parameter choices for creating a reasonable \gD{} layout of
+and (hyper-)parameter choices for creating an accurate \gD{} layout of
 high-dimensional data. Our objective is to fit a model for the \gD{}
@@
 The effectiveness of this approach is illustrated through various
-examples. For instance, the S-curve example demonstrates how the model
-accurately fits the points, capturing both local and global structures
-in high-dimensional space. Our simulation case study further, five
-Gaussian cluster example shows that while all observed NLDR methods
+examples. For instance, the two-curvy clusters example demonstrates how
+the model accurately fits the points, capturing both local and global
+structures in high-dimensional space. Our simulation case study further,
+five Gaussian cluster example shows that while all observed NLDR methods
 preserve the global structure, only tSNE effectively maintains the local
@@
 prefer well-separated views. This emphasizes the importance of clear and
-distinct clusters. For example, in the UMAP layout of the \textbf{pbmc}
-dataset suggested by \citet{chen2023}, three distant, well-separated
-clusters are shown. However, our model reveals that these clusters are
-actually close to each other in \pD{}. Additionally, the model discovers
-non-uniform data distribution and non-linear structures within the
-clusters that are not visible in the UMAP layout, demonstrating the
-ability of our model in uncovering hidden data characteristics.
+distinct clusters. For example, in the UMAP layout of the
+\textbf{PBMC3k} dataset suggested by \citet{chen2023}, three distant,
+well-separated clusters are shown. However, our model reveals that these
+clusters are actually close to each other in \pD{}. Additionally, the
+model discovers non-uniform data distribution and non-linear structures
+within the clusters that are not visible in the UMAP layout,
+demonstrating the ability of our model in uncovering hidden data
+characteristics.
 
@@
 
-Code, and data for reproducing this paper are available at
-\url{https://github.com/JayaniLakshika/paper-nldr-vis-algorithm}.
+Appendix: The appendix includes more details about the hexagonal binning
+algorithm (appendix.pdf, Portable Document Format file).
+
+R package \texttt{quollr}: The R package \texttt{quollr} containing
+codes to fit, and visualize the model. (need to add quollr .zip file,
+GNU zipped tar file)
+
+\section{Acknowledgments}\label{acknowledgments}
+
+These \texttt{R} packages were used for the work: The project's GitHub
+repository
+(\url{https://github.com/JayaniLakshika/paper-nldr-vis-algorithm})
+contains all materials required to reproduce this article.
 

### 2024-09-30 paper.qmd
Updated acknowledgement

@@
 # library(DT)
-library(ggbeeswarm)
+#library(ggbeeswarm)
 library(ggplot2)
@@
 
-library(Rtsne)
-library(uwot)
-library(phateR)
+# library(Rtsne)
+# library(uwot)
+# library(phateR)
 library(patchwork)
 library(png)
-library(langevitour)
+#library(langevitour)
 library(colorspace)
 library(kableExtra)
-library(grid)
+#library(grid)
 library(conflicted)
@@
 
-These `R` packages were used for the work:
+These `R` packages were used for the work:.
+These `python` packages were used for the work:.
+The article was created with `R` packages `rticles`, `knitr`, and `rmarkdown`.
 The project's GitHub repository ([https://github.com/JayaniLakshika/paper-nldr-vis-algorithm](https://github.com/JayaniLakshika/paper-nldr-vis-algorithm)) contains all materials required to reproduce this article.

### 2024-09-30 paper.tex
Updated acknowledgement

@@
 
-These \texttt{R} packages were used for the work: The project's GitHub
-repository
+These \texttt{R} packages were used for the work:. These \texttt{python}
+packages were used for the work:. The article was created with
+\texttt{R} packages \texttt{rticles}, \texttt{knitr}, and
+\texttt{rmarkdown}. The project's GitHub repository
 (\url{https://github.com/JayaniLakshika/paper-nldr-vis-algorithm})

### 2024-09-30 paper.qmd
Added references for acknowledged works

@@
 
-These `R` packages were used for the work:.
+These `R` packages were used for the work: `tidyverse` (@hadley2019), `png` (@simon2022), `Rtsne` (@jesse2015), `umap` (@tomasz2023), `ggplot2` (@hadley2016), `patchwork` (@thomas2024), `colorspace` (@achim2020), `langevitour` (@harisson2024), `conflicted` (@hadley2023), `reticulate` (@kevin2024), `kableExtra` (@hao2024).
 These `python` packages were used for the work:.
-The article was created with `R` packages `rticles`, `knitr`, and `rmarkdown`.
+The article was created with `R` packages `rticles` (@jjallaire2024), `knitr` (@yihui2014), and `rmarkdown` (@yihui2020).
 The project's GitHub repository ([https://github.com/JayaniLakshika/paper-nldr-vis-algorithm](https://github.com/JayaniLakshika/paper-nldr-vis-algorithm)) contains all materials required to reproduce this article.

### 2024-09-30 paper.tex
Added references for acknowledged works

@@
 
-These \texttt{R} packages were used for the work:. These \texttt{python}
-packages were used for the work:. The article was created with
-\texttt{R} packages \texttt{rticles}, \texttt{knitr}, and
-\texttt{rmarkdown}. The project's GitHub repository
+These \texttt{R} packages were used for the work: \texttt{tidyverse}
+(\citet{hadley2019}), \texttt{png} (\citet{simon2022}), \texttt{Rtsne}
+(\citet{jesse2015}), \texttt{umap} (\citet{tomasz2023}),
+\texttt{ggplot2} (\citet{hadley2016}), \texttt{patchwork}
+(\citet{thomas2024}), \texttt{colorspace} (\citet{achim2020}),
+\texttt{langevitour} (\citet{harisson2024}), \texttt{conflicted}
+(\citet{hadley2023}), \texttt{reticulate} (\citet{kevin2024}),
+\texttt{kableExtra} (\citet{hao2024}). These \texttt{python} packages
+were used for the work:. The article was created with \texttt{R}
+packages \texttt{rticles} (\citet{jjallaire2024}), \texttt{knitr}
+(\citet{yihui2014}), and \texttt{rmarkdown} (\citet{yihui2020}). The
+project's GitHub repository
 (\url{https://github.com/JayaniLakshika/paper-nldr-vis-algorithm})

### 2024-09-30 paper.tex
Added python packages

@@
 \texttt{kableExtra} (\citet{hao2024}). These \texttt{python} packages
-were used for the work:. The article was created with \texttt{R}
-packages \texttt{rticles} (\citet{jjallaire2024}), \texttt{knitr}
-(\citet{yihui2014}), and \texttt{rmarkdown} (\citet{yihui2020}). The
-project's GitHub repository
+were used for the work: \texttt{trimap} and \texttt{pacmap}. The article
+was created with \texttt{R} packages \texttt{rticles}
+(\citet{jjallaire2024}), \texttt{knitr} (\citet{yihui2014}), and
+\texttt{rmarkdown} (\citet{yihui2020}). The project's GitHub repository
 (\url{https://github.com/JayaniLakshika/paper-nldr-vis-algorithm})

### 2024-09-30 paper.tex
Added python packages

--- 2024-09-30 paper.tex
@@
 \texttt{kableExtra} (\citet{hao2024}). These \texttt{python} packages
-were used for the work:. The article was created with \texttt{R}
-packages \texttt{rticles} (\citet{jjallaire2024}), \texttt{knitr}
-(\citet{yihui2014}), and \texttt{rmarkdown} (\citet{yihui2020}). The
-project's GitHub repository
+were used for the work: \texttt{trimap} (\citet{amid2022}) and
+\texttt{pacmap} (\citet{yingfan2021}). The article was created with
+\texttt{R} packages \texttt{rticles} (\citet{jjallaire2024}),
+\texttt{knitr} (\citet{yihui2014}), and \texttt{rmarkdown}
+(\citet{yihui2020}). The project's GitHub repository
 (\url{https://github.com/JayaniLakshika/paper-nldr-vis-algorithm})

### 2024-09-30 paper.qmd
Added python packages
Added Figure 9 title

@@
 #| echo: false
-#| fig-cap: "Various plots to help assess best hexagon bin parameters, thresholds to remove low density bins and large edges.
-Both (b) and (c) show MSE, against number of bins along the x-axis and standardised count.
+#| fig-cap: "Various plots to help assess best hexagon bin parameters and thresholds to remove low-density bins.
+Both (a) and (c) show MSE, against binwidth ($a_1$) and threshold.
 A good benchmark value for these parameters is when the MSE drops and then flattens out.
-Plot (a) shows the distribution of stadardised counts of hexagons.
-Plot (d) shows the distribution of $2\\text{-}D$ Euclidean distances between bin centroids, with a good benchmark value for removing large edges would being the distance that shows the first large decrease."
+Three binwidth choices were made: $0.05$, $0.09$, and $0.12$ to investigate.
+As the binwidth increases, the proportion of non-empty bins also increases.
+There are two peaks at binwidths $0.09$ and $0.12$.
+The proportion of densities, calculated as a density with the smallest possible binwidth, decreases and levels off.
+Binwidths $0.09$ and $0.12$ are favorable choices.
+Based on the MSE, $0.09$ was chosen as the initial best binwidth for further analysis.
+There is no need to remove the low-density hexagons because as shown in (b), there is no considerable drop in MSE."
 #| label: fig-param-two-curvy
@@
 These `R` packages were used for the work: `tidyverse` (@hadley2019), `png` (@simon2022), `Rtsne` (@jesse2015), `umap` (@tomasz2023), `ggplot2` (@hadley2016), `patchwork` (@thomas2024), `colorspace` (@achim2020), `langevitour` (@harisson2024), `conflicted` (@hadley2023), `reticulate` (@kevin2024), `kableExtra` (@hao2024).
-These `python` packages were used for the work:.
+These `python` packages were used for the work: `trimap` (@amid2022) and `pacmap` (@yingfan2021).
 The article was created with `R` packages `rticles` (@jjallaire2024), `knitr` (@yihui2014), and `rmarkdown` (@yihui2020).

### 2024-09-30 paper.tex
Added Figure 9 title

@@
 \caption{\label{fig-param-two-curvy}Various plots to help assess best
-hexagon bin parameters, thresholds to remove low density bins and large
-edges. Both (b) and (c) show MSE, against number of bins along the
-x-axis and standardised count. A good benchmark value for these
-parameters is when the MSE drops and then flattens out. Plot (a) shows
-the distribution of stadardised counts of hexagons. Plot (d) shows the
-distribution of \(2\text{-}D\) Euclidean distances between bin
-centroids, with a good benchmark value for removing large edges would
-being the distance that shows the first large decrease.}
+hexagon bin parameters and thresholds to remove low-density bins. Both
+(a) and (c) show MSE, against binwidth (\(a_1\)) and threshold. A good
+benchmark value for these parameters is when the MSE drops and then
+flattens out. Three binwidth choices were made: \(0.05\), \(0.09\), and
+\(0.12\) to investigate. As the binwidth increases, the proportion of
+non-empty bins also increases. There are two peaks at binwidths \(0.09\)
+and \(0.12\). The proportion of densities, calculated as a density with
+the smallest possible binwidth, decreases and levels off. Binwidths
+\(0.09\) and \(0.12\) are favorable choices. Based on the MSE, \(0.09\)
+was chosen as the initial best binwidth for further analysis. There is
+no need to remove the low-density hexagons because as shown in (b),
+there is no considerable drop in MSE.}
 

### 2024-09-30 paper.qmd
Updated Figure 10

@@
 distance_two_curvy_dist <- distance_two_curvy |> 
-  mutate(bin = cut(distance, breaks = bins, include.lowest = TRUE, labels = bin_labels)) |>
-  dplyr::filter(distance > 0.1) ## To remove a1 distances, since always small
+  mutate(bin = cut(distance, breaks = bins, include.lowest = TRUE, labels = bin_labels))
+
+distance_two_curvy_dist <- distance_two_curvy_dist |>
+  count(bin) |>
+  mutate(prop = n/sum(n))
+
+distance_two_curvy_dist <- distance_two_curvy_dist |>
+  filter(bin != unique(distance_two_curvy_dist$bin)[1]) ## To remove a1 distances, since always small
 
 distance_hist <- ggplot(
-  distance_two_curvy_dist, aes(x = distance)) +
-    geom_histogram(
-      aes(y = after_stat(count) / sum(after_stat(count))), 
-      binwidth = bin_width) +
-  scale_x_continuous(breaks = bin_labels, 
+  distance_two_curvy_dist, 
+  aes(x = as.numeric(bin),
+      y = prop)) +
+  geom_line() +
+  scale_x_continuous(breaks = as.numeric(distance_two_curvy_dist$bin),
                      labels = c(
-                       expression(a[1]),
                        expression(2*a[1]),
@@
                      )) +
-  geom_vline(xintercept = start_point + 2*bin_width,
+  geom_vline(xintercept = 3,
+             linetype="solid",
+             color = "#bdbdbd",
+             linewidth=1) +
+  geom_vline(xintercept = 5,
              linetype="solid",
@@
         axis.ticks.y = element_line())
+
+# distance_hist <- ggplot(
+#   distance_two_curvy_dist, aes(x = distance)) +
+#     geom_histogram(
+#       aes(y = after_stat(count) / sum(after_stat(count))), 
+#       binwidth = bin_width) +
+#   scale_x_continuous(breaks = bin_labels, 
+#                      labels = c(
+#                        expression(a[1]),
+#                        expression(2*a[1]),
+#                        expression(3*a[1]),
+#                        expression(4*a[1]),
+#                        expression(5*a[1]),
+#                        expression(6*a[1]),
+#                        expression(7*a[1]),
+#                        expression(8*a[1])
+#                      )) +
+#   geom_vline(xintercept = start_point + 2*bin_width,
+#              linetype="solid",
+#              color = "#bdbdbd",
+#              linewidth=1) +
+#   ylab("Proportion of the number of edges") +
+#   xlab(expression(d^{(2)})) +
+#   #ggtitle("(d)") +
+#   theme_minimal() +
+#   theme(aspect.ratio = 0.75,
+#         panel.border = element_rect(fill = 'transparent'),
+#         plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+#         panel.grid.major.x = element_blank(),
+#         axis.ticks.x = element_line(),
+#         axis.ticks.y = element_line())
 
@@
 ```{r}
-#| fig-cap: "Wireframes with Delaunay triangulation, and two choices of edge removal in \\gD{} and \\pD{}. "
-
-free(distance_hist) + wrap_plots(trimesh_two_curvy2, trimesh_two_curvy_removed2, trimesh_two_curvy_removed1, ncol = 2) 
+#| fig-cap: "Distribution of edge lengths and wireframes in \\gD{} and $7\\text{-}D$ with Delaunay triangulation, and two choices of long edge removal."
+
+free(distance_hist) / wrap_plots(trimesh_two_curvy2, trimesh_two_curvy_removed2, trimesh_two_curvy_removed1, ncol = 3) 
 ```

### 2024-10-01 paper.qmd
Updated long edge benchmark values

@@
                      )) +
-  geom_vline(xintercept = 3,
+  geom_vline(xintercept = 2,
              linetype="solid",
@@
 
+benchmark_two_curvy <- 2 * bin_width
+
 distance_df_small_edges_two_curvy2 <- distance_two_curvy |>
@@
 
-benchmark_two_curvy <- 0.42
+benchmark_two_curvy <- 5 * bin_width
 

### 2024-10-01 paper.qmd
Added Figure 12 title

@@
 #| label: best-fit-umap-tsne
-#| fig-cap: "Best fit"
+#| fig-cap: "Best fit using UMAP ($a_1$ = 0.09) and tSNE ($a_1$ = 0.16) with the same number of non-empty bins ($m$ = 39). Both models show twist in $7\\text{-}D$."
 #| fig-width: 10

### 2024-10-01 paper.tex
Updated Figure 10
Added Figure 12 title

@@
 
-\caption{Wireframes with Delaunay triangulation, and two choices of edge
-removal in \gD{} and \pD{}.}
+\caption{Distribution of edge lengths and wireframes in \gD{} and
+\(7\text{-}D\) with Delaunay triangulation, and two choices of long edge
+removal.}
 
@@
 
-\caption{Best fit}
+\caption{Best fit using UMAP (\(a_1\) = 0.09) and tSNE (\(a_1\) = 0.16)
+with the same number of non-empty bins (\(m\) = 39). Both models show
+twist in \(7\text{-}D\).}
 

### 2024-10-01 paper.qmd
Updated Figure 10 sizes

@@
                colour = "#000000") +
+  geom_point(data = umap_two_curvy_scaled,
+             aes(
+               x = UMAP1,
+               y = UMAP2
+             ),
+             color = "#636363",
+            alpha = 0.5,
+            size = 0.5
+            ) +
+  coord_equal() +
   interior_annotation("a")
@@
 
-trimesh_two_curvy_removed1 <- ggplot() +
-  geom_segment(data = tr_from_to_df_two_curvy3,
-               aes(
-                 x = x_from,
-                 y = y_from,
-                 xend = x_to,
-                 yend = y_to),
-               colour = "#000000") +
-  interior_annotation("c", position = c(0.92, 0.95))
+# trimesh_two_curvy_removed1 <- ggplot() +
+#   geom_segment(data = tr_from_to_df_two_curvy3,
+#                aes(
+#                  x = x_from,
+#                  y = y_from,
+#                  xend = x_to,
+#                  yend = y_to),
+#                colour = "#000000")+
+#   geom_point(data = umap_two_curvy_scaled,
+#              aes(
+#                x = UMAP1,
+#                y = UMAP2
+#              ),
+#              color = "#636363",
+#             alpha = 0.5,
+#             size = 0.5
+#             ) +
+#   coord_equal() +
+#   interior_annotation("c", position = c(0.92, 0.95))
 
@@
                  yend = y_to),
-               colour = "#000000") +
+               colour = "#000000")+
+  geom_point(data = umap_two_curvy_scaled,
+             aes(
+               x = UMAP1,
+               y = UMAP2
+             ),
+             color = "#636363",
+            alpha = 0.5,
+            size = 0.5
+            ) +
+  coord_equal() +
   interior_annotation("b", position = c(0.92, 0.95))
@@
 #| fig-cap: "Distribution of edge lengths and wireframes in \\gD{} and $7\\text{-}D$ with Delaunay triangulation, and two choices of long edge removal."
-
-free(distance_hist) / wrap_plots(trimesh_two_curvy2, trimesh_two_curvy_removed2, trimesh_two_curvy_removed1, ncol = 3) 
+#| fig-height: 10
+#| fig-width: 15
+
+free(distance_hist) / wrap_plots(trimesh_two_curvy2, trimesh_two_curvy_removed2, trimesh_two_curvy_removed1_with_data, ncol = 3) 
 ```

### 2024-10-01 paper.qmd
Updated Figure 10

@@
 
-# trimesh_two_curvy_removed1 <- ggplot() +
-#   geom_segment(data = tr_from_to_df_two_curvy3,
-#                aes(
-#                  x = x_from,
-#                  y = y_from,
-#                  xend = x_to,
-#                  yend = y_to),
-#                colour = "#000000")+
-#   geom_point(data = umap_two_curvy_scaled,
-#              aes(
-#                x = UMAP1,
-#                y = UMAP2
-#              ),
-#              color = "#636363",
-#             alpha = 0.5,
-#             size = 0.5
-#             ) +
-#   coord_equal() +
-#   interior_annotation("c", position = c(0.92, 0.95))
+trimesh_two_curvy_removed1 <- ggplot() +
+  geom_segment(data = tr_from_to_df_two_curvy3,
+               aes(
+                 x = x_from,
+                 y = y_from,
+                 xend = x_to,
+                 yend = y_to),
+               colour = "#000000")+
+  geom_point(data = umap_two_curvy_scaled,
+             aes(
+               x = UMAP1,
+               y = UMAP2
+             ),
+             color = "#636363",
+            alpha = 0.5,
+            size = 0.5
+            ) +
+  coord_equal() +
+  interior_annotation("c", position = c(0.92, 0.95))
 
@@
 #| fig-cap: "Distribution of edge lengths and wireframes in \\gD{} and $7\\text{-}D$ with Delaunay triangulation, and two choices of long edge removal."
-#| fig-height: 10
+#| fig-height: 15
 #| fig-width: 15
 
-free(distance_hist) / wrap_plots(trimesh_two_curvy2, trimesh_two_curvy_removed2, trimesh_two_curvy_removed1_with_data, ncol = 3) 
+free(distance_hist) / wrap_plots(trimesh_two_curvy2, trimesh_two_curvy_removed2, trimesh_two_curvy_removed1, two_curvy_proj_umap_model, ncol = 3) 
 ```

### 2024-10-01 paper.qmd
Added Figure 10 all projections

@@
 
+## Projection with Delauany triangulation
+
+model_df1 <- dplyr::left_join(
+  distance_two_curvy |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df1)[3:NCOL(model_df1)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df1 <- dplyr::left_join(model_df1, projected_model_df, by = c("to" = "ID"))
+names(model_df1)[(2 + NCOL(projected_model_df)):NCOL(model_df1)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+## With benchmark option 2
+
+model_df2 <- dplyr::left_join(
+  distance_df_small_edges_two_curvy3 |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df2)[3:NCOL(model_df2)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df2 <- dplyr::left_join(model_df2, projected_model_df, by = c("to" = "ID"))
+names(model_df2)[(2 + NCOL(projected_model_df)):NCOL(model_df2)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
 ## For true model
@@
 
+two_curvy_proj_umap_model_delaunay <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df1, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#000000") + #31a354
+  geom_point(
+    size = 0.5,
+    alpha = 0.5,
+    color = "#636363") +
+  geom_segment(
+    data=axes,
+    aes(x=x1, y=y1, xend=x2, yend=y2),
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
+    colour="grey50",
+    size = 4) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.126, 0.125)) + 
+  ylim(c(-0.126, 0.125)) +
+  interior_annotation("c") +
+  theme(
+    legend.position = "none"
+  )
+
+two_curvy_proj_umap_model_benchmark2 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df2, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#000000") + #31a354
+  geom_point(
+    size = 0.5,
+    alpha = 0.5,
+    color = "#636363") +
+  geom_segment(
+    data=axes,
+    aes(x=x1, y=y1, xend=x2, yend=y2),
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
+    colour="grey50",
+    size = 4) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.126, 0.125)) + 
+  ylim(c(-0.126, 0.125)) +
+  interior_annotation("c") +
+  theme(
+    legend.position = "none"
+  )
+
 two_curvy_proj_umap_all_model <- projected_df |>
@@
 ```{r}
-#| fig-cap: "Distribution of edge lengths and wireframes in \\gD{} and $7\\text{-}D$ with Delaunay triangulation, and two choices of long edge removal."
+#| fig-cap: "Distribution of edge lengths and wireframes in \\gD{} and $7\\text{-}D$ with Delaunay triangulation, and two choices of long edge removal: (b) benchmark = 2$a_1$ and (c) benchmark = 5$a_1$, where $a_1$ = 0.09."
 #| fig-height: 15
@@
 
-free(distance_hist) / wrap_plots(trimesh_two_curvy2, trimesh_two_curvy_removed2, trimesh_two_curvy_removed1, two_curvy_proj_umap_model, ncol = 3) 
+free(distance_hist) / wrap_plots(trimesh_two_curvy2, trimesh_two_curvy_removed2, trimesh_two_curvy_removed1,
+two_curvy_proj_umap_model_delaunay,
+two_curvy_proj_umap_model_benchmark2,
+two_curvy_proj_umap_model, ncol = 3) 
 ```

### 2024-10-01 paper.qmd
Changed colors for tSNE

@@
              y = tSNE2))+
-  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
+  geom_point(alpha=0.1, size=1, colour='#a65628') +
   interior_annotation("a", c(0.08, 0.93)) +
@@
              y = UMAP2)) +
-  geom_point(alpha=0.1, size=1, colour='#a65628') +
+  geom_point(alpha=0.1, size=1, colour='#636363') +
   interior_annotation("b", c(0.08, 0.93)) +
@@
              y = PHATE2))+
-  geom_point(alpha=0.1, size=1, colour='#377eb8') +
+  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
   interior_annotation("c") +
@@
   scale_x_continuous(breaks = sort(unique(error_two_curvy_clust$a1))[append(seq(1, length(unique(error_two_curvy_clust$a1)), by = 5), 24)]) +
-  scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a', "#ff7f00",'#a65628')) +
+  scale_color_manual(values=c('#e41a1c','#ff7f00','#4daf4a', "#a65628",'#636363')) +
   scale_y_log10() +
@@
              ),
-             color = "#636363",
+             color = "#a65628",
              alpha = 0.5,
@@
     alpha = 0.5,
-    color = "#636363") +
+    color = "#a65628") +
   geom_segment(

### 2024-10-01 paper.qmd
Updated Figure 11 error plot's x-axis

@@
   #            color = "black", linewidth=0.8, alpha = 0.5) +
-  scale_x_continuous(breaks = sort(unique(error_two_curvy_clust$a1))[append(seq(1, length(unique(error_two_curvy_clust$a1)), by = 5), 24)]) +
+  scale_x_continuous(breaks = sort(unique(error_two_curvy_clust$a1))[seq(1, length(unique(error_two_curvy_clust$a1)), by = 5)]) +
   scale_color_manual(values=c('#e41a1c','#ff7f00','#4daf4a', "#a65628",'#636363')) +

### 2024-10-01 paper.qmd
Try to remove the white spaces

@@
 #| fig-height: 10
+#| out-height: 80%
 #| fig-cap: "Comparison of prediction generated using the exiting `umap` R package's prediction method and our method: (a) A view of the true model in projection from $7\\text{-}D$, (b) predicted data from the `umap` prediction method, (c) predicted data from our method, and (d) comparison of predictions from both methods.
@@
 #| label: fig-param-two-curvy
-#| out-width: 100%
+#| out-width: 80%
 #| fig-width: 12
@@
 #| fig-width: 15
+#| out-height: 70%
 
@@
 #| fig-pos: H
-#| out-height: 100%
+#| out-height: 60%
 
@@
 #| fig-height: 20
+#| out-height: 90%
 

### 2024-10-01 paper.tex
Added Figure 10 all projections
Try to remove the white spaces

@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-predict-two-curvy-1.pdf}
+\includegraphics[width=1\textwidth,height=0.8\textheight]{paper_files/figure-pdf/fig-predict-two-curvy-1.pdf}
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-param-two-curvy-1.pdf}
+\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-param-two-curvy-1.pdf}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-36-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=0.7\textheight]{paper_files/figure-pdf/unnamed-chunk-36-1.pdf}
 
@@
 \(7\text{-}D\) with Delaunay triangulation, and two choices of long edge
-removal.}
+removal: (b) benchmark = 2\(a_1\) and (c) benchmark = 5\(a_1\), where
+\(a_1\) = 0.09.}
 
@@
 
-\includegraphics[width=1\textwidth,height=1\textheight]{paper_files/figure-pdf/fig-two_curvy_clust-mse-1.pdf}
+\includegraphics[width=1\textwidth,height=0.6\textheight]{paper_files/figure-pdf/fig-two_curvy_clust-mse-1.pdf}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/best-fit-umap-tsne-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=0.9\textheight]{paper_files/figure-pdf/best-fit-umap-tsne-1.pdf}
 

### 2024-10-01 paper.qmd
Updated Figure 12

@@
   geom_point(alpha=0.5, color = "#636363", size = 0.5) +
-  xlim(c(-0.01, 1.1)) + ylim(c(-0.08, 0.57)) +
+  theme(aspect.ratio = 1) +
+  #xlim(c(-0.01, 1.1)) + ylim(c(-0.08, 0.57)) +
   interior_annotation("a", sc_ltr_pos)
@@
 nldr_two_curvy_original + tsne_two_curvy_clust_plt +
-  trimesh_two_curvy_removed1_with_data + trimesh_two_curvy_removed1_tsne_with_data +
+  trimesh_two_curvy_removed1 + trimesh_two_curvy_removed1_tsne_with_data +
   two_curvy_proj_umap_model +

### 2024-10-01 paper.qmd
Updated the axes labels of MSE

@@
   scale_y_log10() +
-  ylab("MSE") +
+  ylab("log(MSE)") +
   xlab(expression(paste("binwidth (", a[1], ")"))) +

### 2024-10-01 paper.qmd
Updated Figure 10 by adding a1

@@
 
-distance_two_curvy_dist <- distance_two_curvy_dist |>
-  filter(bin != unique(distance_two_curvy_dist$bin)[1]) ## To remove a1 distances, since always small
+# distance_two_curvy_dist <- distance_two_curvy_dist |>
+#   filter(bin != unique(distance_two_curvy_dist$bin)[1]) ## To remove a1 distances, since always small
 
@@
                      labels = c(
+                       expression(a[1]),
                        expression(2*a[1]),

### 2024-10-01 paper.qmd
Updated Figure 10

@@
              linewidth=1) +
-  geom_vline(xintercept = 5,
+  geom_vline(xintercept = 4.5,
              linetype="solid",
@@
 
-benchmark_two_curvy <- 5 * bin_width
+benchmark_two_curvy <- 4.5 * bin_width
 
@@
 ```{r}
-#| fig-cap: "Distribution of edge lengths and wireframes in \\gD{} and $7\\text{-}D$ with Delaunay triangulation, and two choices of long edge removal: (b) benchmark = 2$a_1$ and (c) benchmark = 5$a_1$, where $a_1$ = 0.09."
+#| fig-cap: "Distribution of edge lengths and wireframes in \\gD{} and $7\\text{-}D$ with Delaunay triangulation, and two choices of long edge removal: (b) benchmark = 4.5$a_1$ and (c) benchmark = 2$a_1$, where $a_1$ = 0.09."
 #| fig-height: 15

### 2024-10-01 paper.qmd
Updated Figure 9 dens calculation with total number of bins (Didn't chnaged: Need to look)

@@
   ungroup() |>
-  mutate(prop_dens = b_non_empty/side_length)
+  mutate(prop_dens = b/side_length)
 

### 2024-10-01 paper.qmd
Added Figure 13 (Didn't see any pattern with error)

@@
 
+## Compute error
+
+error_df_gau_abs <- augment(
+  df_bin_centroids = df_bin_centroids_gau1,
+  df_bin = df_bin_gau1,
+  training_data = training_data_gau,
+  newdata = NULL,
+  type_NLDR = "tSNE",
+  col_start = "x")
+
+error_df_gau_abs <- error_df_gau_abs |>
+  mutate(sqrt_row_wise_abs_error = sqrt(row_wise_abs_error))
+
+error_df_gau_abs <- error_df_gau_abs |>
+  bind_cols(tsne_gau_scaled |>
+              select(-ID))
+
+error_df_gau_tsne <- error_df_gau_abs |>
+  ggplot(aes(x = tSNE1,
+             y = tSNE2,
+             colour = sqrt_row_wise_abs_error)) +
+  geom_point(alpha=0.5) + 
+  xlim(c(0.3, 0.66)) + ylim(c(-0.01, 0.35)) +
+  scale_colour_continuous_sequential(palette = "Blues") +
+  interior_annotation("a3", 
+                      position = c(0.08, 0.9),
+                      cex = 2)
+
 ## Triangulate bin centroids
@@
 
+## Compute error
+
+error_df_gau_abs <- augment(
+  df_bin_centroids = df_bin_centroids_gau1,
+  df_bin = df_bin_gau1,
+  training_data = training_data_gau,
+  newdata = NULL,
+  type_NLDR = "UMAP",
+  col_start = "x")
+
+error_df_gau_abs <- error_df_gau_abs |>
+  mutate(sqrt_row_wise_abs_error = sqrt(row_wise_abs_error))
+
+error_df_gau_abs <- error_df_gau_abs |>
+  bind_cols(umap_gau_scaled |>
+              select(-ID))
+
+error_df_gau_umap <- error_df_gau_abs |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2,
+             colour = sqrt_row_wise_abs_error)) +
+  geom_point(alpha=0.5) + 
+  xlim(c(0.47, 0.5475)) + ylim(c(0.0125, 0.09)) +
+  scale_colour_continuous_sequential(palette = "Blues") +
+  interior_annotation("b3", 
+                      position = c(0.08, 0.9),
+                      cex = 2)
+
 ## Triangulate bin centroids
@@
 
+## Compute error
+
+error_df_gau_abs <- augment(
+  df_bin_centroids = df_bin_centroids_gau1,
+  df_bin = df_bin_gau1,
+  training_data = training_data_gau,
+  newdata = NULL,
+  type_NLDR = "PaCMAP",
+  col_start = "x")
+
+error_df_gau_abs <- error_df_gau_abs |>
+  mutate(sqrt_row_wise_abs_error = sqrt(row_wise_abs_error))
+
+error_df_gau_abs <- error_df_gau_abs |>
+  bind_cols(pacmap_gau_scaled |>
+              select(-ID))
+
+error_df_gau_pacmap <- error_df_gau_abs |>
+  ggplot(aes(x = PaCMAP1,
+             y = PaCMAP2,
+             colour = sqrt_row_wise_abs_error)) +
+  geom_point(alpha=0.5) + 
+  xlim(c(-0.02, 0.18)) + ylim(c(0.12, 0.32)) +
+  scale_colour_continuous_sequential(palette = "Blues") +
+  interior_annotation("c3", 
+                      position = c(0.08, 0.9),
+                      cex = 2)
+
 ## Triangulate bin centroids
@@
 trimesh_removed_gau_tsne + trimesh_removed_gau_umap + trimesh_removed_gau_pacmap +
-five_gau_proj_tsne_model1 + five_gau_proj_umap_model1 + five_gau_proj_pacmap_model1 + five_gau_proj_tsne_model2 + five_gau_proj_umap_model2 + five_gau_proj_pacmap_model2 +
+five_gau_proj_tsne_model1 + five_gau_proj_umap_model1 + five_gau_proj_pacmap_model1 + error_df_gau_tsne + error_df_gau_umap + error_df_gau_pacmap +
   plot_layout(guides = "collect", nrow = 3) &

### 2024-10-01 paper.qmd
Updated Figure 10, nicely filtered out long edges outside the structure which showing in a small cluster, still need to work on how to remove long edges within the clutser (which is the most suitable benchmark)

@@
 
+### Compute highD distance
+dist_vec <- proxy::dist(x = df_bin_two_curvy2[, -1], method = "Euclidean") |> as.vector()
+
+from_vec <- c()
+to_vec <- c()
+num_obs <- 1:(NROW(df_bin_two_curvy2) - 1)
+
+for (obs in num_obs) {
+
+  from_val <- rep(obs, (NROW(df_bin_two_curvy2) - obs))
+  if ((obs + 1) <= NROW(df_bin_two_curvy2)) {
+    to_val <- (obs + 1):NROW(df_bin_two_curvy2)
+  }
+  from_vec <- append(from_vec, from_val)
+  to_vec <- append(to_vec, to_val)
+
+}
+
+dist_highd <- tibble::tibble(from = from_vec, to = to_vec, dist_highd = dist_vec)
+
 df_b_two_curvy2 <- df_bin_two_curvy2 |>
@@
 
-distance_two_curvy_dist <- distance_two_curvy_dist |>
-  count(bin) |>
-  mutate(prop = n/sum(n))
-
-# distance_two_curvy_dist <- distance_two_curvy_dist |>
-#   filter(bin != unique(distance_two_curvy_dist$bin)[1]) ## To remove a1 distances, since always small
+distance_two_curvy_dist <- left_join(distance_two_curvy_dist, dist_highd, by = c("from", "to"))
 
@@
   aes(x = as.numeric(bin),
-      y = prop)) +
-  geom_line() +
-  scale_x_continuous(breaks = as.numeric(distance_two_curvy_dist$bin),
+      y = dist_highd)) +
+  geom_point(alpha = 0.5) +
+  geom_jitter() +
+  scale_x_continuous(breaks = sort(unique(as.numeric(distance_two_curvy_dist$bin))),
                      labels = c(
@@
                      )) +
-  geom_vline(xintercept = 2,
-             linetype="solid",
-             color = "#bdbdbd",
-             linewidth=1) +
-  geom_vline(xintercept = 4.5,
-             linetype="solid",
-             color = "#bdbdbd",
-             linewidth=1) +
-  ylab("Proportion of the number of edges") +
+  # geom_vline(xintercept = 2,
+  #            linetype="solid",
+  #            color = "#bdbdbd",
+  #            linewidth=1) +
+  # geom_vline(xintercept = 4.5,
+  #            linetype="solid",
+  #            color = "#bdbdbd",
+  #            linewidth=1) +
+  ylab(expression(d^{(7)})) +
   xlab(expression(d^{(2)})) +
@@
         axis.ticks.y = element_line())
+
+# distance_two_curvy_dist <- distance_two_curvy_dist |>
+#   count(bin) |>
+#   mutate(prop = n/sum(n))
+
+# distance_two_curvy_dist <- distance_two_curvy_dist |>
+#   filter(bin != unique(distance_two_curvy_dist$bin)[1]) ## To remove a1 distances, since always small
+
+# distance_hist <- ggplot(
+#   distance_two_curvy_dist, 
+#   aes(x = as.numeric(bin),
+#       y = prop)) +
+#   geom_line() +
+#   scale_x_continuous(breaks = as.numeric(distance_two_curvy_dist$bin),
+#                      labels = c(
+#                        expression(a[1]),
+#                        expression(2*a[1]),
+#                        expression(3*a[1]),
+#                        expression(4*a[1]),
+#                        expression(5*a[1]),
+#                        expression(6*a[1]),
+#                        expression(7*a[1]),
+#                        expression(8*a[1])
+#                      )) +
+#   geom_vline(xintercept = 2,
+#              linetype="solid",
+#              color = "#bdbdbd",
+#              linewidth=1) +
+#   geom_vline(xintercept = 4.5,
+#              linetype="solid",
+#              color = "#bdbdbd",
+#              linewidth=1) +
+#   ylab("Proportion of the number of edges") +
+#   xlab(expression(d^{(2)})) +
+#   #ggtitle("(d)") +
+#   theme_minimal() +
+#   theme(aspect.ratio = 0.75,
+#         panel.border = element_rect(fill = 'transparent'),
+#         plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+#         panel.grid.major.x = element_blank(),
+#         axis.ticks.x = element_line(),
+#         axis.ticks.y = element_line())
 

### 2024-10-01 paper.qmd
Updated Figure 10

--- 2024-10-01 paper.qmd
@@
 
-distance_two_curvy_dist <- distance_two_curvy_dist |>
-  count(bin) |>
-  mutate(prop = n/sum(n))
-
-# distance_two_curvy_dist <- distance_two_curvy_dist |>
-#   filter(bin != unique(distance_two_curvy_dist$bin)[1]) ## To remove a1 distances, since always small
-
-distance_hist <- ggplot(
+distance_two_curvy_dist <- left_join(distance_two_curvy_dist, dist_highd, by = c("from", "to"))
+
+distance_points <- ggplot(
   distance_two_curvy_dist, 
   aes(x = as.numeric(bin),
-      y = prop)) +
-  geom_line() +
-  scale_x_continuous(breaks = as.numeric(distance_two_curvy_dist$bin),
+      y = dist_highd)) +
+  geom_point(alpha = 0.5) +
+  geom_jitter() +
+  scale_x_continuous(breaks = sort(unique(as.numeric(distance_two_curvy_dist$bin))),
                      labels = c(
@@
                      )) +
+  # geom_vline(xintercept = 2,
+  #            linetype="solid",
+  #            color = "#bdbdbd",
+  #            linewidth=1) +
+  # geom_vline(xintercept = 4.5,
+  #            linetype="solid",
+  #            color = "#bdbdbd",
+  #            linewidth=1) +
+  ylab(expression(d^{(7)})) +
+  xlab(expression(d^{(2)})) +
+  #ggtitle("(d)") +
+  theme_minimal() +
+  theme(aspect.ratio = 0.75,
+        panel.border = element_rect(fill = 'transparent'),
+        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+        panel.grid.major.x = element_blank(),
+        axis.ticks.x = element_line(),
+        axis.ticks.y = element_line())
+
+distance_two_curvy_dist_summary <- distance_two_curvy_dist |>
+  count(bin) |>
+  mutate(prop = n/sum(n)) |>
+  head(4)
+
+# distance_two_curvy_dist <- distance_two_curvy_dist |>
+#   filter(bin != unique(distance_two_curvy_dist$bin)[1]) ## To remove a1 distances, since always small
+
+distance_hist <- ggplot(
+  distance_two_curvy_dist_summary,
+  aes(x = as.numeric(bin),
+      y = prop)) +
+  geom_line() +
+  scale_x_continuous(breaks = sort(unique(as.numeric(distance_two_curvy_dist$bin)))[1:4],
+                     labels = c(
+                       expression(a[1]),
+                       expression(2*a[1]),
+                       expression(3*a[1]),
+                       expression(4*a[1])
+                     )) +
   geom_vline(xintercept = 2,
@@
              linewidth=1) +
-  geom_vline(xintercept = 4.5,
+  geom_vline(xintercept = 3.5,
              linetype="solid",
@@
 
-benchmark_two_curvy <- 4.5 * bin_width
+benchmark_two_curvy <- 3.5 * bin_width
 
@@
 
-free(distance_hist) / wrap_plots(trimesh_two_curvy2, trimesh_two_curvy_removed2, trimesh_two_curvy_removed1,
+wrap_plots(distance_points,distance_hist, ncol = 2) / wrap_plots(trimesh_two_curvy2, trimesh_two_curvy_removed2, trimesh_two_curvy_removed1,
 two_curvy_proj_umap_model_delaunay,

### 2024-10-01 paper.qmd
Figure 10 add labels on the line

@@
 
+text_df <- tibble(
+  x = c(1.9, 3.4),
+  y = c(0.5, 0.5),
+  text = c("c1", "b1")
+)
+
 distance_hist <- ggplot(
@@
              linewidth=1) +
+  geom_text(
+    data = text_df,
+    aes(
+      x = x,
+      y = y,
+      label = text
+    ),
+    size = 9
+  ) +
   ylab("Proportion of the number of edges") +

### 2024-10-07 paper.qmd
Checking continuous color palattes

@@
   geom_point(alpha=0.5) +
-  scale_colour_continuous_sequential(palette = "Blues") +
-  coord_equal()
+  scale_colour_continuous_sequential(palette = "YlOrRd", n_interp = 10) +
+  theme(
+    aspect.ratio = 1
+  )
 
@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 13
+num_bins_x_gau1 <- 21
 lim1 <- gau1_scaled_obj$lim1
@@
   xlim(c(0.3, 0.66)) + ylim(c(-0.01, 0.35)) +
-  scale_colour_continuous_sequential(palette = "Blues") +
+  scale_colour_continuous_sequential(palette = "YlOrRd", n_interp = 20) +
   interior_annotation("a3", 
@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 45
+num_bins_x_gau1 <- 76
 lim1 <- gau1_scaled_obj$lim1
@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 21
+num_bins_x_gau1 <- 33
 lim1 <- gau1_scaled_obj$lim1

### 2024-10-07 paper.qmd
Updated Figure 5 labels

@@
   coord_equal() +
-  interior_annotation("b")
+  # theme(
+  #   aspect.ratio = 1
+  #   ) +
+  interior_annotation("b", sc_ltr_pos)
 ```
@@
   ylim(c(-0.126, 0.125)) +
-  interior_annotation("c") +
+  interior_annotation("c", sc_ltr_pos) +
   theme(
@@
   ylim(c(-0.126, 0.125)) +
-  interior_annotation("d") +
+  interior_annotation("d", sc_ltr_pos) +
   theme(
@@
   coord_equal() +
-  interior_annotation("a")
+  interior_annotation("a1")
 
@@
   coord_equal() +
-  interior_annotation("c", position = c(0.92, 0.95))
+  interior_annotation("c1")
 
@@
   coord_equal() +
-  interior_annotation("b", position = c(0.92, 0.95))
+  interior_annotation("b1")
 
@@
   ylim(c(-0.126, 0.125)) +
-  interior_annotation("c") +
+  interior_annotation("c2") +
   theme(
@@
   ylim(c(-0.126, 0.125)) +
-  interior_annotation("c") +
+  interior_annotation("a2") +
   theme(
@@
   ylim(c(-0.126, 0.125)) +
-  interior_annotation("c") +
+  interior_annotation("b2") +
   theme(
@@
 #| fig-width: 10
-#| fig-height: 10
+#| fig-height: 8
 

### 2024-10-07 paper.qmd
Updated Figure 10 labels

@@
   coord_equal() +
-  interior_annotation("a1")
+  interior_annotation("a1", sc_ltr_pos)
 
@@
   coord_equal() +
-  interior_annotation("c1")
+  interior_annotation("c1", sc_ltr_pos)
 
@@
   coord_equal() +
-  interior_annotation("b1")
+  interior_annotation("b1", sc_ltr_pos)
 
@@
   ylim(c(-0.126, 0.125)) +
-  interior_annotation("c2") +
+  interior_annotation("c2", sc_ltr_pos) +
   theme(
@@
   ylim(c(-0.126, 0.125)) +
-  interior_annotation("a2") +
+  interior_annotation("a2", sc_ltr_pos) +
   theme(
@@
   ylim(c(-0.126, 0.125)) +
-  interior_annotation("b2") +
+  interior_annotation("b2", sc_ltr_pos) +
   theme(

### 2024-10-07 paper.qmd
Updated Figure 12

@@
   interior_annotation("a", sc_ltr_pos)
+
+nldr_two_curvy_original2 <- umap_two_curvy_scaled |> 
+  ggplot(aes(x = UMAP1, y = UMAP2)) + 
+  geom_point(alpha=0.5, color = "#636363", size = 0.5) +
+  theme(aspect.ratio = 1) +
+  #xlim(c(-0.01, 1.1)) + ylim(c(-0.08, 0.57)) +
+  interior_annotation("a1", sc_ltr_pos)
 ```
@@
   #xlim(c(-0.01, 1.1)) + ylim(c(-0.08, 0.57)) +
-  interior_annotation("a2")
+  interior_annotation("a2", sc_ltr_pos)
 
@@
 
+two_curvy_proj_umap_model2 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#000000") + #31a354
+  geom_point(
+    size = 0.5,
+    alpha = 0.5,
+    color = "#636363") +
+  geom_segment(
+    data=axes,
+    aes(x=x1, y=y1, xend=x2, yend=y2),
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
+    colour="grey50",
+    size = 4) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.126, 0.125)) + 
+  ylim(c(-0.126, 0.125)) +
+  interior_annotation("a3", sc_ltr_pos) +
+  theme(
+    legend.position = "none"
+  )
+
 two_curvy_proj_umap_model_delaunay <- projected_df |>
@@
   ylim(c(-0.126, 0.125)) +
-  interior_annotation("d") +
+  interior_annotation("a4", sc_ltr_pos) +
   theme(
@@
   geom_point(alpha=0.1, size=1, colour='#a65628') +
-  interior_annotation("a", c(0.08, 0.93)) +
+  interior_annotation("b1", sc_ltr_pos) +
   theme(aspect.ratio = 1)
@@
   coord_equal() +
-  interior_annotation("b")
+  interior_annotation("b2", sc_ltr_pos)
 
@@
   ylim(c(-0.126, 0.125)) +
-  interior_annotation("c") +
+  interior_annotation("b3", sc_ltr_pos) +
   theme(
@@
   ylim(c(-0.126, 0.125)) +
-  interior_annotation("d") +
+  interior_annotation("b4", sc_ltr_pos) +
   theme(
@@
 
-nldr_two_curvy_original + tsne_two_curvy_clust_plt +
+nldr_two_curvy_original2 + tsne_two_curvy_clust_plt +
   trimesh_two_curvy_removed1_with_data + trimesh_two_curvy_removed1_tsne_with_data +
-  two_curvy_proj_umap_model +
+  two_curvy_proj_umap_model2 +
   two_curvy_proj_tsne_model +

### 2024-10-07 paper.qmd
Updated Figure 5

@@
   two_curvy_proj_umap_all_model1 +
-  plot_layout(ncol = 2)
+  plot_layout(ncol = 2, heights = c(1, 2))
 ```

### 2024-10-07 paper.qmd
Added best fit with tNSE
Updated Figure 12 Caption

@@
 ```{r}
-#| fig-cap: "Assessing which of the 5 NLDR layouts on the two_curvy_clust digit 1 data is the better representation using MSE for varying binwidth ($a_1$).
+#| fig-cap: "Assessing which of the 5 NLDR layouts on the two-curvy clusters data is the better representation using MSE for varying binwidth ($a_1$).
 Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-e).
-Layout c is universally poor.
+Layout d is universally poor.
 Layouts a that show two close clusters are universally suboptimal.
@@
 distance_df_small_edges_two_curvy4 <- distance_two_curvy4 |>
-  filter(distance < benchmark_two_curvy4) ## 0.221
+  filter(distance < benchmark_two_curvy4) |> ## 0.221
+  filter(!(row_number() %in% c(35, 89, 83, 25, 80, 56, 13, 59, 16)))
 
@@
 #| label: best-fit-umap-tsne
-#| fig-cap: "Best fit using UMAP ($a_1$ = 0.09) and tSNE ($a_1$ = 0.16) with the same number of non-empty bins ($m$ = 39). Both models show twist in $7\\text{-}D$."
+#| fig-cap: "Best fit for two-curvy clusters data using UMAP (a1-a4) ($a_1$ = 0.09) and tSNE (b1-b4) ($a_1$ = 0.16) with the same number of non-empty bins ($m$ = 39).
+Both models show twist in $7\\text{-}D$ (a3, a4, b3, b4).
+Also, both NLDR methods are unable to accurately capture the width of the curvilinear clusters."
 #| fig-width: 10
@@
 It’s important to access the \gD{} layout and the generated model overlaid on data in \pD{} together to understand whether it fits the points everywhere, fits better in some places, or simply mismatches the pattern.
-Interactivity also helps in understanding the quirks that occur with different NLDR techniques (XXXXRefer the video after finalising the S-curve).
+Interactivity also helps in understanding the quirks that occur with different NLDR techniques (XXXXRefer the video after finalising the two-curvy clusters data).
 

### 2024-10-07 paper.tex
Updated Figure 10
Updated Figure 12 Caption

@@
 \(7\text{-}D\) with Delaunay triangulation, and two choices of long edge
-removal: (b) benchmark = 2\(a_1\) and (c) benchmark = 5\(a_1\), where
+removal: (b) benchmark = 4.5\(a_1\) and (c) benchmark = 2\(a_1\), where
 \(a_1\) = 0.09.}
@@
 \caption{\label{fig-two_curvy_clust-mse}Assessing which of the 5 NLDR
-layouts on the two\_curvy\_clust digit 1 data is the better
-representation using MSE for varying binwidth (\(a_1\)). Colour used for
-the lines and points in the left plot and in the scatterplots represents
-NLDR layout (a-e). Layout c is universally poor. Layouts a that show two
-close clusters are universally suboptimal. Layout a is the best choice.}
+layouts on the two-curvy clusters data is the better representation
+using MSE for varying binwidth (\(a_1\)). Colour used for the lines and
+points in the left plot and in the scatterplots represents NLDR layout
+(a-e). Layout d is universally poor. Layouts a that show two close
+clusters are universally suboptimal. Layout a is the best choice.}
 
@@
 
-\caption{Best fit using UMAP (\(a_1\) = 0.09) and tSNE (\(a_1\) = 0.16)
-with the same number of non-empty bins (\(m\) = 39). Both models show
-twist in \(7\text{-}D\).}
+\caption{Best fit for two-curvy clusters data using UMAP (a1-a4)
+(\(a_1\) = 0.09) and tSNE (b1-b4) (\(a_1\) = 0.16) with the same number
+of non-empty bins (\(m\) = 39). Both models show twist in \(7\text{-}D\)
+(a3, a4, b3, b4). Also, both NLDR methods are unable to accurately
+capture the width of the curvilinear clusters.}
 
@@
 with different NLDR techniques (XXXXRefer the video after finalising the
-S-curve).
+two-curvy clusters data).
 

### 2024-10-07 paper.qmd
Changed the sequqntial color palatte for Five_Gau clusters example

@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 21
+num_bins_x_gau1 <- 13
 lim1 <- gau1_scaled_obj$lim1
@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 76
+num_bins_x_gau1 <- 45
 lim1 <- gau1_scaled_obj$lim1
@@
   xlim(c(0.47, 0.5475)) + ylim(c(0.0125, 0.09)) +
-  scale_colour_continuous_sequential(palette = "Blues") +
+  scale_colour_continuous_sequential(palette = "YlOrRd", n_interp = 20) +
   interior_annotation("b3", 
@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 33
+num_bins_x_gau1 <- 21
 lim1 <- gau1_scaled_obj$lim1
@@
   xlim(c(-0.02, 0.18)) + ylim(c(0.12, 0.32)) +
-  scale_colour_continuous_sequential(palette = "Blues") +
+  scale_colour_continuous_sequential(palette = "YlOrRd", n_interp = 20) +
   interior_annotation("c3", 

### 2024-10-07 paper.qmd
Corrected the projection of Figure 13

@@
   geom_point(alpha=0.5) +
-  scale_colour_continuous_sequential(palette = "YlOrRd", n_interp = 10) +
+  scale_colour_continuous_sequential(palette = "YlOrRd", n_interp = 20) +
   theme(
@@
 #   c(-0.70709,0.42156,-0.02593,0.10965))
+# projection <- cbind(
+#   c(-0.00215,-0.68905,-0.04778,-0.54223),
+#   c(0.42558,-0.23854,-0.63659,0.35753))
+
 projection <- cbind(
-  c(-0.00215,-0.68905,-0.04778,-0.54223),
-  c(0.42558,-0.23854,-0.63659,0.35753))
+  c(-0.04778, -0.54223, -0.00215, -0.68905),
+  c(-0.63659, 0.35753, 0.42558, -0.23854))
 
@@
              colour = sqrt_row_wise_abs_error)) +
-  geom_point(alpha=0.5) + 
+  geom_point(alpha=0.6,
+             size = 1.5) + 
   xlim(c(-0.02, 0.18)) + ylim(c(0.12, 0.32)) +
-  scale_colour_continuous_sequential(palette = "YlOrRd", n_interp = 20) +
+  scale_colour_continuous_sequential(palette = "YlOrRd", breaks = 20) +
   interior_annotation("c3", 

### 2024-10-07 paper.qmd
Corrected Figure 13 projection

@@
 projection <- cbind(
-  c(-0.04778, -0.54223, -0.00215, -0.68905),
-  c(-0.63659, 0.35753, 0.42558, -0.23854))
+  c(-0.00215,-0.68905,-0.04778,-0.54223),
+  c(0.42558, -0.23854, -0.63659, 0.35753))
 
@@
   axis_pos_x = -0.58,
-  axis_pos_y = -0.05,
+  axis_pos_y = -0.35,
   axis_labels = names(scaled_gau_data),
@@
   coord_fixed() +
-  xlim(c(-0.64, -0.26)) +
-  ylim(c(0, -0.38)) +
+  xlim(c(-0.64, -0.29)) +
+  ylim(c(-0.4, -0.05)) +
   interior_annotation("b2", 

### 2024-10-07 paper.qmd
Changed the prop density calculation and added new notations to the table

@@
   ungroup() |>
-  mutate(prop_dens = b/side_length)
+  mutate(prop_dens = 1/(b*side_length^2))
 
@@
 #### Removal of low density bins
+
+##### Density of a bin
+
+Area of a hexagon is defined as $A = \frac{3\sqrt{3}}{2}l^2$ where $l$ is the side length of the hexagon.
+If we know $a_1$ and $a_2$, $l$ can be computed (see appendix).
+Therefore, the density of a hexagon grid is computed as $\frac{\sum^{h}_{i=1}n_h}{A}$ and the proportion is $\frac{\sum^{h}_{i=1}n_h}{A \times b}$.
+The base line proportion is defined as the
 

### 2024-10-07 paper.tex
Changed the prop density calculation and added new notations to the table

@@
 $h$ & hexagonal id\\
+$l$ & side length\\
+$A$ & area\\
 \bottomrule
@@
 bins}\label{removal-of-low-density-bins}
+
+\paragraph{Density of a bin}\label{density-of-a-bin}
+
+Area of a hexagon is defined as \(A = \frac{3\sqrt{3}}{2}l^2\) where
+\(l\) is the side length of the hexagon. If we know \(a_1\) and \(a_2\),
+\(l\) can be computed (see appendix). Therefore, the density of a
+hexagon grid is computed as \(\frac{\sum^{h}_{i=1}n_h}{A}\) and the
+proportion is \(\frac{\sum^{h}_{i=1}n_h}{A \times b}\). The base line
+proportion is defined as the
 

### 2024-10-07 paper.qmd
Added relative proportion density section to the diagnostics

@@
 
+### Diagnostics
+
+#### Relative proportion density
+
+The area of a hexagon is defined as $A = \frac{3\sqrt{3}}{2}l^2$ where $l$ is the side length of the hexagon.
+If we know $a_1$ and $a_2$, $l$ can be computed (see appendix).
+The density of a hexagon grid is calculated as $\frac{\sum^{h}_{i=1}n_h}{A}$ and the proportion is $\frac{\sum^{h}_{i=1}n_h}{A \times b}$.
+The baseline proportion is the proportion density at the smallest possible value of $a_1$.
+The relative proportion density is the ratio of the observed proportion density to the baseline proportion density.
+
 ### Tuning
@@
 #### Removal of low density bins
-
-##### Density of a bin
-
-Area of a hexagon is defined as $A = \frac{3\sqrt{3}}{2}l^2$ where $l$ is the side length of the hexagon.
-If we know $a_1$ and $a_2$, $l$ can be computed (see appendix).
-Therefore, the density of a hexagon grid is computed as $\frac{\sum^{h}_{i=1}n_h}{A}$ and the proportion is $\frac{\sum^{h}_{i=1}n_h}{A \times b}$.
-The base line proportion is defined as the
 
@@
 There are two peaks at binwidths $0.09$ and $0.12$.
-The proportion of densities, calculated as a density with the smallest possible binwidth, decreases and levels off.
-Binwidths $0.09$ and $0.12$ are favorable choices.
+The relative proportion density decreases and levels off.
+Binwidths $0.05$ is the favorable choices.
 Based on the MSE, $0.09$ was chosen as the initial best binwidth for further analysis.

### 2024-10-07 paper.tex
Added relative proportion density section to the diagnostics

--- 2024-10-07 paper.tex
@@
 \end{figure}%
+
+\subsection{Diagnostics}\label{diagnostics}
+
+\subsubsection{Relative proportion
+density}\label{relative-proportion-density}
+
+The area of a hexagon is defined as \(A = \frac{3\sqrt{3}}{2}l^2\) where
+\(l\) is the side length of the hexagon. If we know \(a_1\) and \(a_2\),
+\(l\) can be computed (see appendix). The density of a hexagon grid is
+calculated as \(\frac{\sum^{h}_{i=1}n_h}{A}\) and the proportion is
+\(\frac{\sum^{h}_{i=1}n_h}{A \times b}\). The baseline proportion is the
+proportion density at the smallest possible value of \(a_1\). The
+relative proportion density is the ratio of the observed proportion
+density to the baseline proportion density.
 
@@
 non-empty bins also increases. There are two peaks at binwidths \(0.09\)
-and \(0.12\). The proportion of densities, calculated as a density with
-the smallest possible binwidth, decreases and levels off. Binwidths
-\(0.09\) and \(0.12\) are favorable choices. Based on the MSE, \(0.09\)
+and \(0.12\). The relative proportion density decreases and levels off.
+Binwidths \(0.05\) is the favorable choices. Based on the MSE, \(0.09\)
 was chosen as the initial best binwidth for further analysis. There is

### 2024-10-07 paper.qmd
Changed the example dataset name

@@
 #| label: fig-two-curvy-true-proj
-#| fig-cap: "Two views of the true model (blue lines) in $2\\text{-}D$ projections from $7\\text{-}D$, for the two-curvy clusters data (grey points).
+#| fig-cap: "Two views of the true model (blue lines) in $2\\text{-}D$ projections from $7\\text{-}D$, for the two C-shaped clusters data (grey points).
 The data is spread along the non-linear shaped clusters, and does not vary much from the curves.
@@
 #| echo: false
-#| fig-cap: "Key steps for constructing the model on the UMAP layout ($k=2$): (a) data, (b) hexagon bins, (c) bin centroids, and (d) triangulated centroids. The two-curvy clusters data is shown."
+#| fig-cap: "Key steps for constructing the model on the UMAP layout ($k=2$): (a) data, (b) hexagon bins, (c) bin centroids, and (d) triangulated centroids. The two C-shaped clusters data is shown."
 #| fig-width: 12
@@
 #| label: best-fit-umap1
-#| fig-cap: "Model in \\gD{} ($a_1 = 0.12$), on the layout and two views of the fit in projections from $7\\text{-}D$, for the two-curvy clusters data ($n =  2000$ and $p = 7$).
+#| fig-cap: "Model in \\gD{} ($a_1 = 0.12$), on the layout and two views of the fit in projections from $7\\text{-}D$, for the two C-shaped clusters data ($n =  2000$ and $p = 7$).
 One view shows the fitted model, and the other shows the fitted and true models alongside the data.
@@
 #| fig-pos: H
-#| fig-cap: "Hexbin density plots of UMAP layout of the two-curvy clusters data, using three different bin inputs: (a) $b = 80 \\text{ } (10, \\text{ }8)$, (b) $b = 130 \\text{ } (13, \\text{ }10)$, and (c) $b = 391 \\text{ } (23, \\text{ }17)$.
+#| fig-cap: "Hexbin density plots of UMAP layout of the two C-shaped clusters data, using three different bin inputs: (a) $b = 80 \\text{ } (10, \\text{ }8)$, (b) $b = 130 \\text{ } (13, \\text{ }10)$, and (c) $b = 391 \\text{ } (23, \\text{ }17)$.
 Color indicates standardized counts, dark indicating high count and light indicates low count.
@@
 When separated clusters are present the mesh should be broken across the gaps.
-For non-linear structures like the S-curve, the mesh should run unbroken along the S, but there should be no edges connecting the top of the S directly to the bottom of the S.
+For non-linear structures like the C-shaped curvilinear, the mesh should run unbroken along the C, but there should be no edges connecting the top of the C directly to the bottom of the C.
 For these reasons it is necessary to remove edges from the mesh in some applications.
@@
 ```{r}
-#| fig-cap: "Assessing which of the 5 NLDR layouts on the two-curvy clusters data is the better representation using MSE for varying binwidth ($a_1$).
+#| fig-cap: "Assessing which of the 5 NLDR layouts on the two C-shaped clusters data is the better representation using MSE for varying binwidth ($a_1$).
 Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-e).
@@
 #| label: best-fit-umap-tsne
-#| fig-cap: "Best fit for two-curvy clusters data using UMAP (a1-a4) ($a_1$ = 0.09) and tSNE (b1-b4) ($a_1$ = 0.16) with the same number of non-empty bins ($m$ = 39).
+#| fig-cap: "Best fit for two C-shaped clusters data using UMAP (a1-a4) ($a_1$ = 0.09) and tSNE (b1-b4) ($a_1$ = 0.16) with the same number of non-empty bins ($m$ = 39).
 Both models show twist in $7\\text{-}D$ (a3, a4, b3, b4).
@@
 It’s important to access the \gD{} layout and the generated model overlaid on data in \pD{} together to understand whether it fits the points everywhere, fits better in some places, or simply mismatches the pattern.
-Interactivity also helps in understanding the quirks that occur with different NLDR techniques (XXXXRefer the video after finalising the two-curvy clusters data).
+Interactivity also helps in understanding the quirks that occur with different NLDR techniques (XXXXRefer the video after finalising the two C-shaped clusters data).
 

### 2024-10-07 paper.tex
Changed the example dataset name

@@
 (blue lines) in \(2\text{-}D\) projections from \(7\text{-}D\), for the
-two-curvy clusters data (grey points). The data is spread along the
+two C-shaped clusters data (grey points). The data is spread along the
 non-linear shaped clusters, and does not vary much from the curves. The
@@
 on the UMAP layout (\(k=2\)): (a) data, (b) hexagon bins, (c) bin
-centroids, and (d) triangulated centroids. The two-curvy clusters data
-is shown.}
+centroids, and (d) triangulated centroids. The two C-shaped clusters
+data is shown.}
 
@@
 \caption{Model in \gD{} (\(a_1 = 0.12\)), on the layout and two views of
-the fit in projections from \(7\text{-}D\), for the two-curvy clusters
-data (\(n =  2000\) and \(p = 7\)). One view shows the fitted model, and
-the other shows the fitted and true models alongside the data. The
-fitted model accurately represents the shape of the clusters but does
-not capture the edges of the structure. This illustrates a
+the fit in projections from \(7\text{-}D\), for the two C-shaped
+clusters data (\(n =  2000\) and \(p = 7\)). One view shows the fitted
+model, and the other shows the fitted and true models alongside the
+data. The fitted model accurately represents the shape of the clusters
+but does not capture the edges of the structure. This illustrates a
 characteristic of UMAP, which compresses the data when transforming the
@@
 \caption{\label{fig-bins-two-curvy}Hexbin density plots of UMAP layout
-of the two-curvy clusters data, using three different bin inputs: (a)
+of the two C-shaped clusters data, using three different bin inputs: (a)
 \(b = 80 \text{ } (10, \text{ }8)\), (b)
@@
 separated clusters are present the mesh should be broken across the
-gaps. For non-linear structures like the S-curve, the mesh should run
-unbroken along the S, but there should be no edges connecting the top of
-the S directly to the bottom of the S. For these reasons it is necessary
-to remove edges from the mesh in some applications.
+gaps. For non-linear structures like the C-shaped curvilinear, the mesh
+should run unbroken along the C, but there should be no edges connecting
+the top of the C directly to the bottom of the C. For these reasons it
+is necessary to remove edges from the mesh in some applications.
 
@@
 \caption{\label{fig-two_curvy_clust-mse}Assessing which of the 5 NLDR
-layouts on the two-curvy clusters data is the better representation
+layouts on the two C-shaped clusters data is the better representation
 using MSE for varying binwidth (\(a_1\)). Colour used for the lines and
@@
 
-\caption{Best fit for two-curvy clusters data using UMAP (a1-a4)
+\caption{Best fit for two C-shaped clusters data using UMAP (a1-a4)
 (\(a_1\) = 0.09) and tSNE (b1-b4) (\(a_1\) = 0.16) with the same number
@@
 with different NLDR techniques (XXXXRefer the video after finalising the
-two-curvy clusters data).
+two C-shaped clusters data).
 

### 2024-10-07 paper.qmd
Added some references

@@
 The values $(s_1, \ s_2)$ define the position of the centroid of the bottom left hexagon.
-By default, this is at $s_1 = -q, s_2 = -qr_2$, where $q$ is the buffer sound the data.
+By default, this is at $s_1 = -q, s_2 = -qr_2$, where $q$ is the buffer bound the data.
 The choice of these values can have some effect on the distribution of bin counts.
-@fig-param-two_curvy (a) illustrates this.
+@fig-bins-two-curvy illustrates this.
 The distribution of bin counts for $s_1$ varying between $-0.1-0.0$ is shown.
@@
 Values of $b_1$ between $2$ and $b_1 = \sqrt{\frac{n}{r_2}}$ are allowed.
-@fig-param-two_curvy (b) shows the effect of different choices of $b_1$ on the MSE of the fitted model.
+@fig-param-two-curvy (a) shows the effect of different choices of $b_1$ on the MSE of the fitted model.
 
@@
 Alternatively, one could choose to remove based on a percentile, the bins with density in the lowest 5% of all bins, for example.
-@fig-param-two_curvy (c) illustrates the effect on the model representation of removing bins below different percentages.
+@fig-param-two-curvy (c) illustrates the effect on the model representation of removing bins below different percentages.
 Generally, we would urge caution in removing low count bins.
@@
 The benchmark value for removing low-density hexagons ranges between $0$ and $1$.
-When analyzing how these benchmark values influence model performance, it's essential to observe the change in MSE as the benchmark value increases (@fig-param-two_curvy).
+When analyzing how these benchmark values influence model performance, it's essential to observe the change in MSE as the benchmark value increases (@fig-param-two-curvy (c)).
 The MSE shows a gradual decrease as the benchmark value goes from $1$ to $0$.
@@
 Edges define the neighbourhood structure, in order to provide a smooth \gD{} representation of the fitted model.
-@fig-two_curvy-true-sc shows a wire frame of the true model that was used to generate the S-curve example data.
+@fig-two-curvy-true-proj shows a wire frame of the true model that was used to generate the two C-shaped clusters example data.
 The ideal is that the representation of the fitted model, at least for this example where we know the true model, should look similar to this.
@@
 Because the triangulation is typically done on the hexagon centroids, there are particular discrete edge lengths, based on bin widths.
-@fig-NLDR-two-curvy (d) illustrates edge length distributions.
+@fig-rm-lg illustrates edge length distributions.
 
@@
 #| out-height: 70%
+#| fig-label: fig-rm-lg
 

### 2024-10-07 paper.tex
Added some references

@@
 bottom left hexagon. By default, this is at \(s_1 = -q, s_2 = -qr_2\),
-where \(q\) is the buffer sound the data. The choice of these values can
+where \(q\) is the buffer bound the data. The choice of these values can
 have some effect on the distribution of bin counts.
-\textbf{?@fig-param-two\_curvy} (a) illustrates this. The distribution
-of bin counts for \(s_1\) varying between \(-0.1-0.0\) is shown.
-Generally, a more uniform distribution among these possibilities would
-indicate that the bins are reliably capturing the underlying
-distribution of observations.
+Figure~\ref{fig-bins-two-curvy} illustrates this. The distribution of
+bin counts for \(s_1\) varying between \(-0.1-0.0\) is shown. Generally,
+a more uniform distribution among these possibilities would indicate
+that the bins are reliably capturing the underlying distribution of
+observations.
 
@@
 \(2\) and \(b_1 = \sqrt{\frac{n}{r_2}}\) are allowed.
-\textbf{?@fig-param-two\_curvy} (b) shows the effect of different
+Figure~\ref{fig-param-two-curvy} (a) shows the effect of different
 choices of \(b_1\) on the MSE of the fitted model.
@@
 the bins with density in the lowest 5\% of all bins, for example.
-\textbf{?@fig-param-two\_curvy} (c) illustrates the effect on the model
+Figure~\ref{fig-param-two-curvy} (c) illustrates the effect on the model
 representation of removing bins below different percentages. Generally,
@@
 model performance, it's essential to observe the change in MSE as the
-benchmark value increases (\textbf{?@fig-param-two\_curvy}). The MSE
-shows a gradual decrease as the benchmark value goes from \(1\) to
+benchmark value increases (Figure~\ref{fig-param-two-curvy} (c)). The
+MSE shows a gradual decrease as the benchmark value goes from \(1\) to
 \(0\). Evaluating this rate of increase is important. If the increment
@@
 \gD{} representation of the fitted model.
-\textbf{?@fig-two\_curvy-true-sc} shows a wire frame of the true model
-that was used to generate the S-curve example data. The ideal is that
-the representation of the fitted model, at least for this example where
-we know the true model, should look similar to this.
+Figure~\ref{fig-two-curvy-true-proj} shows a wire frame of the true
+model that was used to generate the two C-shaped clusters example data.
+The ideal is that the representation of the fitted model, at least for
+this example where we know the true model, should look similar to this.
 
@@
 typically done on the hexagon centroids, there are particular discrete
-edge lengths, based on bin widths. Figure~\ref{fig-NLDR-two-curvy} (d)
-illustrates edge length distributions.
+edge lengths, based on bin widths. \textbf{?@fig-rm-lg} illustrates edge
+length distributions.
 

### 2024-10-07 paper.qmd
Updated the fig reference

@@
 ```{r}
+#| label: fig-rm-lg
 #| fig-cap: "Distribution of edge lengths and wireframes in \\gD{} and $7\\text{-}D$ with Delaunay triangulation, and two choices of long edge removal: (b) benchmark = 4.5$a_1$ and (c) benchmark = 2$a_1$, where $a_1$ = 0.09."
@@
 #| out-height: 70%
-#| fig-label: fig-rm-lg
 

### 2024-10-07 paper.tex
Updated the fig reference

@@
 typically done on the hexagon centroids, there are particular discrete
-edge lengths, based on bin widths. \textbf{?@fig-rm-lg} illustrates edge
-length distributions.
+edge lengths, based on bin widths. Figure~\ref{fig-rm-lg} illustrates
+edge length distributions.
 
@@
 
-\begin{figure}[H]
-
-{\centering \includegraphics[width=1\textwidth,height=0.7\textheight]{paper_files/figure-pdf/unnamed-chunk-36-1.pdf}
-
-}
-
-\caption{Distribution of edge lengths and wireframes in \gD{} and
-\(7\text{-}D\) with Delaunay triangulation, and two choices of long edge
-removal: (b) benchmark = 4.5\(a_1\) and (c) benchmark = 2\(a_1\), where
-\(a_1\) = 0.09.}
+\begin{figure}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=0.7\textheight]{paper_files/figure-pdf/fig-rm-lg-1.pdf}
+
+}
+
+\caption{\label{fig-rm-lg}Distribution of edge lengths and wireframes in
+\gD{} and \(7\text{-}D\) with Delaunay triangulation, and two choices of
+long edge removal: (b) benchmark = 4.5\(a_1\) and (c) benchmark =
+2\(a_1\), where \(a_1\) = 0.09.}
 

### 2024-10-07 appendix.qmd
Added url

@@
     email: michael.lydeamore@monash.edu
-    url: 
+    url: https://www.michaellydeamore.com/
   - name: Thiyanga S. Talagala
@@
 library(dplyr)
-# remotes::install_github("jlmelville/snedata")
-library(snedata)
-library(ggflowchart)
-library(purrr) ## map function
-library(gridExtra) ## for grid.arrange
-library(rsample)
-library(DT)
-library(ggbeeswarm)
 library(ggplot2)
@@
 library(tidyr)
-
-library(Rtsne)
-library(uwot)
-library(phateR)
 library(patchwork)
@@
 
-source("nldr_code.R", local = TRUE)
+#source("nldr_code.R", local = TRUE)
 ```
@@
 #| message: false
+#| eval: false
 

### 2024-10-07 appendix.qmd
Restructured appendix

@@
 ---
-title: "Appendix: Visualising How Non-linear Dimension Reduction Warps Your Data"
+title: "Appendix: Looking at Non-Linear Dimension Reductions as Models in the Data Space"
+format: 
+    jasa-pdf:
+        keep-tex: true
+    jasa-html: default
 author:
@@
 
-```{r}
-#| warning: false
-#| echo: false
-library(quollr)
-library(dplyr)
-library(ggplot2)
-library(readr)
-library(tidyr)
-library(patchwork)
-library(langevitour)
-library(colorspace)
-library(kableExtra)
-library(grid)
-set.seed(20240110)
-
-#source("nldr_code.R", local = TRUE)
-```
-
-<!-- To find the number of bins should assign for the model in tSNE-->
-```{r}
-#| warning: false
-#| echo: false
-#| message: false
-#| eval: false
-
-training_data_gau <- read_rds("data/five_gau_clusters/data_five_gau_training.rds")
-tSNE_data_gau <- read_rds("data/five_gau_clusters/tsne_data_five_gau_61.rds")
-
-tsne_gau_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tSNE_data_gau, 
-                                    x = "tSNE1", y = "tSNE2"))) |>
-  dplyr::rename(c("tSNE1" = "scaled_tSNE1", 
-                  "tSNE2" = "scaled_tSNE2")) |>
-  dplyr::mutate(ID = 1:NROW(tSNE_data_gau))
-
-## tSNE
-hex_size_vec <- seq(0.02, 2, by = 0.01)
-
-vec <- stats::setNames(rep("", 6), c("num_bins", "aic", "mse", "num_bins_x", "num_bins_y", "hex_size"))  ## Define column names
-
-mse_df_gau <- dplyr::bind_rows(vec)[0, ]
-mse_df_gau <- mse_df_gau |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(hex_size_vec)) {
-
-  num_bin_list <- calc_bins(data = tsne_gau_scaled, 
-            x = "tSNE1", y = "tSNE2", 
-            hex_size = hex_size_vec[i], buffer_x = NA, buffer_y = NA)
-
-  num_bins_x <- num_bin_list$num_x
-  num_bins_y <- num_bin_list$num_y
-
-  model_object <- fit_highd_model( training_data = training_data_gau, 
-                                   nldr_df_with_id = tsne_gau_scaled, 
-                                   x = "tSNE1", y = "tSNE2", 
-                                   num_bins_x = num_bins_x, 
-                                   num_bins_y = num_bins_y, 
-                                   x_start = NA, y_start = NA, 
-                                   buffer_x = NA, buffer_y = NA, 
-                                   hex_size = hex_size_vec[i],
-                                   is_rm_lwd_hex = FALSE, 
-                                   benchmark_to_rm_lwd_hex = NA, 
-                                   col_start_2d = "tSNE", 
-                                   col_start_highd = "x")
-
-  centroid_df_training <- model_object$df_bin_centroids
-  avg_df_training <- model_object$df_bin
-
-  pred_emb_list <- predict_emb(test_data = training_data_gau, 
-                                  df_bin_centroids = centroid_df_training, 
-                                  df_bin = avg_df_training, type_NLDR = "tSNE")
-
-  pred_df_training <- as.data.frame(do.call(cbind, pred_emb_list))
-
-  eval_list <- gen_summary(test_data = training_data_gau, 
-                                  prediction_df = pred_df_training, 
-                                  df_bin = avg_df_training, col_start = "x")
-
-  mse_df_gau <- mse_df_gau |>
-    tibble::add_row(num_bins = num_bins_x * num_bins_y,
-                    aic = eval_list$aic,
-                    mse = eval_list$mse,
-                    num_bins_x = num_bins_x,
-                    num_bins_y = num_bins_y,
-                    hex_size = hex_size_vec[i])
-
-}
-
-## If same total number of bins occurred only select ones with minimum error
-### Obtain duplicate bins
-dupli_bins <- mse_df_gau |> 
-  dplyr::count(num_bins) |> 
-  dplyr::filter(n > 1) |> 
-  dplyr::pull(num_bins)
-
-### Group split by duplicated bins
-duplicate_df_list <- mse_df_gau |>
-  dplyr::filter(num_bins %in% dupli_bins) |>
-  dplyr::arrange(num_bins) |>
-  dplyr::group_split(num_bins)
-
-### Obtain one row from duplicates which have lowest error and hexsize
-duplicate_df <- data.frame(matrix(nrow = 0, ncol = 0))
-
-for (i in 1:length(duplicate_df_list)) {
-
-  dd <- duplicate_df_list[[i]] |>
-    dplyr::filter(mse == min(duplicate_df_list[[i]]$mse)) |>
-    dplyr::filter(hex_size == min(duplicate_df_list[[i]]$hex_size))
-
-  duplicate_df <- dplyr::bind_rows(duplicate_df, dd)
-
-}
-
-### Obtain the mse_df with not duplicated bins
-not_dupli_df <- mse_df_gau |>
-  dplyr::filter(!(num_bins %in% dupli_bins))
-
-### Combine duplicated and not duplicated(corrected) bins dfs
-mse_df_gau <- dplyr::bind_rows(not_dupli_df, duplicate_df) 
+```{r include=FALSE}
+# Set up chunk for for knitr
+knitr::opts_chunk$set(
+  fig.width = 5,
+  fig.height = 5,
+  fig.align = "center",
+  out.width = "100%",
+  code.line.numbers = FALSE,
+  fig.retina = 4,
+  echo = FALSE,
+  message = FALSE,
+  warning = FALSE,
+  cache = FALSE,
+  dev.args = list(pointsize = 11)
+)
 ```
@@
 ```{r}
-#| warning: false
-#| echo: false
-#| eval: false
+library(ggplot2)
+library(dplyr)
+library(quollr)
+library(readr)
+library(ggforce)
+library(patchwork)
 
-mse_plot_gau <- ggplot(mse_df_gau, aes(x = num_bins,
-                                       y = log(mse)
-)) +
-  geom_point(size = 0.8) +
-  geom_line() +
-   geom_vline(xintercept = 588, linetype="solid",
-                color = "red", size=0.8, alpha = 0.5) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  #scale_colour_manual(values = c("#377eb8")) +
-  ylab("MSE") +
-  xlab("total number of bins")
+theme_set(theme_linedraw() +
+   theme(
+     #aspect.ratio = 1,
+     plot.background = element_rect(fill = 'transparent', colour = NA),
+     plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
+     panel.background = element_rect(fill = 'transparent', 
+                                     colour = NA),
+     panel.grid.major = element_blank(), 
+     panel.grid.minor = element_blank(), 
+     axis.title.x = element_blank(), axis.title.y = element_blank(),
+     axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+     axis.text.y = element_blank(), axis.ticks.y = element_blank(),
+     legend.background = element_rect(fill = 'transparent', 
+                                      colour = NA),
+     legend.key = element_rect(fill = 'transparent', 
+                               colour = NA),
+     legend.position = "none", 
+     legend.title = element_text(size=5), 
+     legend.text = element_text(size=4),
+     legend.key.height = unit(0.25, 'cm'),
+     legend.key.width = unit(0.25, 'cm')
+   )
+)
+```
+
+## Computing hexagon grid configurations
+
+Given range of embedding component, $r_2$, number of bins along the x-axis, $b_1$, and buffer proportion, $q$, hexagonal starting point coordinates, $s_1 = -q$, and $s_2 = -q \times r_2$.
+The purpose is to find width of the hexagon.
+$a_1$, and number of bins along the y-axis, $b_2$.
+
+```{r}
+umap_scurve <- read_rds(file = here::here("data/s_curve/s_curve_umap.rds"))
+
+scurve_scaled_obj <- gen_scaled_data(
+  data = umap_scurve)
+
+umap_scurve_scaled <- scurve_scaled_obj$scaled_nldr
+lim1 <- scurve_scaled_obj$lim1
+lim2 <- scurve_scaled_obj$lim2
+r2 <- diff(lim2)/diff(lim1)
+
+sc_ltr_pos <- c(0.08, 0.96)
+sc_xlims <- c(-0.5, 1.47)
+sc_ylims <- c(-0.32, 2.1)
+```
+
+<!--Full hexagon grid with UMAP data-->
+
+```{r}
+## Compute hexbin parameters
+num_bins_x_scurve <- 7
+
+## hexagon binning to have regular hexagons
+hb_obj_scurve <- hex_binning(
+  data = umap_scurve_scaled, 
+  bin1 = num_bins_x_scurve, 
+  r2 = r2)
+
+## Data set with all centroids
+all_centroids_df <- hb_obj_scurve$centroids
+
+## Generate all coordinates of hexagons
+hex_grid <- hb_obj_scurve$hex_poly
+```
+
+```{r}
+# Code to draw illustration for notation
+start_pt <- all_centroids_df |> 
+  filter(hexID == 1)
+d_rect <- tibble(x1min = 0, 
+                 x1max = 1,
+                 x2min = 0,
+                 x2max = r2)
+
+c_start <- tibble(s1 = 0.03, s2 = -0.05)
+
+a1 <- tibble(x = all_centroids_df$c_x[4],
+             xend = all_centroids_df$c_x[5],
+             y = all_centroids_df$c_y[21],
+             yend = all_centroids_df$c_y[21],
+             label = expression(a[1]))
+a2 <- tibble(x = all_centroids_df$c_x[27],
+             xend = all_centroids_df$c_x[27],
+             y = all_centroids_df$c_y[27],
+             yend = all_centroids_df$c_y[33],
+             label = expression(a[2]))
+hex_param_vis <- ggplot() + 
+    geom_polygon(data = hex_grid, 
+                        aes(x = x, 
+                            y = y, 
+                            group = hex_poly_id),
+                 fill = "white", 
+                 color = "#bdbdbd") +
+    geom_point(data = all_centroids_df, aes(
+      x = c_x, 
+      y = c_y), 
+      color = "#31a354", size = 0.9) +
+    geom_point(data = start_pt, aes(x = c_x, 
+                                    y = c_y), 
+               color = "black") + 
+    geom_rect(data=d_rect, 
+              aes(xmin = x1min - c_start$s1,# - c_start$s1, 
+                  xmax = x1max - c_start$s1,# - c_start$s1, 
+                  ymin = x2min - c_start$s2,# - c_start$s2, 
+                  ymax = x2max - c_start$s2),# - c_start$s2), 
+              fill = "white", 
+              color = "black", 
+              alpha = 0, 
+              linewidth = 0.7) +
+    geom_point(data=d_rect, aes(x=x1min - c_start$s1, 
+                                y=x2min - c_start$s2)) + 
+    geom_point(data=d_rect, aes(x=x1max - c_start$s1, 
+                                y=x2min - c_start$s2)) + 
+    geom_point(data=d_rect, aes(x=x1min - c_start$s1, 
+                                y=x2max - c_start$s2)) + 
+    annotate("text", x=d_rect$x1min - c_start$s1, 
+                     y=d_rect$x2min - c_start$s2,
+                     label = "(0,0)", 
+             hjust=-0.1, vjust=-0.3) + 
+    annotate("text", x=d_rect$x1max - c_start$s1, 
+                     y=d_rect$x2min - c_start$s2,
+                     label = "(0,1)", 
+             hjust=1.1, vjust=-0.3) + 
+    annotate("text", x=d_rect$x1min - c_start$s1, 
+                     y=d_rect$x2max - c_start$s2,
+                     label = expression(group("(", 
+                        list(0, y[2][max]),")")), 
+            hjust=-0.1, vjust=1.2) + 
+    geom_segment(data=d_rect, aes(
+      x = x1min  - c_start$s1, # 0 - 0.03, 
+      y = -0.35, 
+      xend = x1max - c_start$s1, #1 - 0.03, 
+      yend = -0.35), #-0.35),
+      arrow = arrow(length = unit(0.03, "npc"),
+                               ends = "both"), 
+                 color = "black")+
+    annotate("text", x=0.5, y=-0.45, 
+             label = expression(r[1]), color = "black") +
+    geom_segment(data=d_rect, aes(
+      x = -0.25, 
+      y = x2min - c_start$s2, #0 - 0.05, 
+      xend = -0.25, 
+      yend = x2max - c_start$s2), #r2 - 0.05),
+      arrow = arrow(length = unit(0.03, "npc"),
+                       ends = "both"), 
+                 color = "black")+ 
+    annotate("text", x=-0.35, y=1, 
+             label = expression(r[2]), color = "black") +
+    geom_segment(data = a1, aes(
+      x = x, #-0.1 + 0.2087578, 
+      y = y, #-0.15, 
+      xend = xend, #-0.1 + 0.2087578*2, 
+      yend = yend), #-0.15),
+      arrow = arrow(length = unit(0.03, "npc"),
+        ends = "both"), 
+        color = "black")+ # a1 = 0.2087578
+    annotate("text", 
+             x=(a1$x+a1$xend)/2, 
+             y=a1$y, 
+             label = expression(a[1]), 
+             color = "black",
+             vjust = 1.2) +
+    geom_segment(data = a2, aes(
+      x = x, #-0.15, 
+      y = y, #-0.1*r2 + 0.1807896*2, 
+      xend = xend, #-0.15, 
+      yend = yend), #-0.1*r2 + 0.1807896*3),
+      arrow = arrow(length = unit(0.03, "npc"),
+                               ends = "both"), 
+      color = "black") + # a2 = 0.1807896
+    annotate("text", x=a2$x, y=(a2$y+a2$yend)/2, 
+             label = expression(a[2]), 
+             color = "black", hjust=-0.2) +
+    annotate("text", x=-0.18, y=-0.25, 
+      label = expression(group("(", list(s[1], s[2]), ")")),
+      color = "black") 
+```
+
+```{r}
+#| label: fig-hex-param
+#| fig-cap: "The components of the hexagon grid illustrating notation."
+#| out-height: 30%
+#| fig-width: 2.5
+#| fig-height: 3.3
+#| fig-pos: H
+
+hex_param_vis
+```
+
+Geometric arguments give rise to the following constraints.
+
+$\text{min }a_1 \text{ s.t.}$
+
+$$
+s_1 - \frac{a_1}{2} < 0,
+$$ {#eq-equation1}
+
+$$
+s_1 + (b_1 - 1) \times a_1 > 1,
+$$ {#eq-equation2}
+
+$$
+s_2 - \frac{a_2}{2} < 0,
+$$ {#eq-equation4}
+
+$$
+s_2 + (b_2 - 1) \times a_2 > r_2.
+$$ {#eq-equation5}
+
+Since $a_1$ and $a_2$ are distances,
+
+$$
+a_1, a_2 > 0.
+$$ Also, $(s_1, s_2) \in (-0.1, -0.05)$ as these are multiplicative offsets in the negative direction.
+
+@eq-equation1 can be rearranged as,
+
+$$
+a_1 > 2s_1
+$$
+
+which given $s_1 < 0$ and $a_1 > 0$ will *always* be true.
+The same logic follows for @eq-equation4 and substituting $a_2 = \frac{\sqrt{3}}{2}a_1$, and $s_2 = -q \times r_2$ to @eq-equation4 can be written as,
+
+$$
+a_1 > -\frac{4}{\sqrt{3}}qr_2
+$$
+
+Also, substituting $a_2 = \frac{\sqrt{3}}{2}a_1$, $s_2 = -q \times r_2$ and rearranging @eq-equation5 gives:
+
+$$
+a_1 > \frac{2(r_2 + qr_2)}{\sqrt{3}(b_2 - 1)}.
+$$ {#eq-equation6}
+
+Similarly, substituting $s_1 = -q$ @eq-equation2 becomes,
+
+$$
+a_1 > \frac{(1 + q)}{(b_1 - 1)}.
+$$ {#eq-equation7}
+
+This is a linear optimization problem. Therefore, the optimal solution must occur on a vertex. So, by setting @eq-equation6 equals to @eq-equation7 gives,
+
+$$
+\frac{2(r_2 + qr_2)}{\sqrt{3}(b_2 - 1)} = \frac{(1 + q)}{(b_1 - 1)}.
+$$ After rearranging this,
+
+$$
+b_2 = 1 + \frac{2r_2(b_1 - 1)}{\sqrt{3}}
+$$
+
+and since $b_2$ should be an integer,
+
+$$
+b_2 = \Big\lceil1 +\frac{2r_2(b_1 - 1)}{\sqrt{3}}\Big\rceil.
+$$ {#eq-equation8}
+
+Furthermore, with known $b_1$ and $b_2$, by considering @eq-equation2 or @eq-equation5 as the *binding* or *active constraint*, can compute $a_1$.
+
+If @eq-equation2 is active, then,
+
+$$
+\frac{(1 + q)}{(b_1 - 1)} < \frac{2(r_2 + qr_2)}{\sqrt{3}(b_2 - 1)}.
+$$
+
+Rearranging this gives,
+
+$$
+r_2 > \frac{\sqrt{3}(b_2 - 1)}{2(b_1 - 1)}.
+$$
+
+Therefore, if this equality is true, then $a_1 = \frac{(1+q)}{(b_1 - 1)}$, otherwise, $a_1 = \frac{2r_2(1+q)}{\sqrt{3}(b_2 - 1)}$.
+
+\newpage
+
+## Binning the data
+
+Points are assigned to the bin they fall into based on the nearest centroid. If a point is equidistant from multiple centroids, it is assigned to the centroid with the lowest hexagonal bin ID.
+
+```{r}
+all_centroids_df <- gen_centroids(bin1 = 4, r2 = r2, q = 0.1)
+
+hex_grid <- gen_hex_coord(all_centroids_df, a1 = 1.043789)
+
+## Data to generate circle
+circles <- data.frame(
+  x = c(0.422, 0.683),
+  y = c(0.740, 1.19),
+  r = rep(0.301316, 2)
+)
+
+emb_df1 <- tibble(emb1 = 0.58, emb2 = 0.93) |>
+  add_row(emb1 = 0.5, emb2 = 1.01)
+
+assign_data_vis1 <- ggplot(data = hex_grid, aes(x = x, y = y)) + 
+  geom_polygon(fill = "white", color = "#bdbdbd", aes(group = hex_poly_id)) +
+  geom_point(data = emb_df1, aes(x = emb1, y = emb2), color = "black", size = 0.7) +
+  geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID), size = 5, colour = "#bdbdbd") +
+  geom_circle(data = circles, aes(x0 = x, y0 = y, r = r), fill = NA) +
+  geom_segment(x = 0.58, y = 0.93, xend = 0.44, yend = 0.76, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "#969696") +
+  geom_segment(x = 0.5, y = 1.01, xend = 0.68, yend = 1.16, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "#969696")
+
+emb_df2 <- tibble(emb1 = 0.161, emb2 = 0.7395) |>
+  add_row(emb1 = 0.683, emb2 = 0.7395) |>
+  add_row(emb1 = 0.2915, emb2 = 0.965) |>
+  add_row(emb1 = 0.5525, emb2 = 0.965) |>
+  add_row(emb1 = 0.2915, emb2 = 0.5135) |>
+  add_row(emb1 = 0.5525, emb2 = 0.5135) |>
+  add_row(emb1 = 0.422, emb2 = 1.04) |>
+  add_row(emb1 = 0.161, emb2 = 0.890) |>
+  add_row(emb1 = 0.161, emb2 = 0.589) |>
+  add_row(emb1 = 0.422, emb2 = 0.438) |>
+  add_row(emb1 = 0.683, emb2 = 0.589) |>
+  add_row(emb1 = 0.683, emb2 = 0.890)
+
+assign_data_vis2 <- ggplot(data = hex_grid, aes(x = x, y = y)) + 
+  geom_polygon(fill = "white", color = "#bdbdbd", aes(group = hex_poly_id)) +
+  geom_point(data = emb_df2, aes(x = emb1, y = emb2), color = "black", size = 0.7) +
+  geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID), size = 5, colour = "#bdbdbd") +
+  geom_segment(x = 0.161, y = 0.7395, xend = -0.06, yend = 0.740, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  geom_segment(x = 0.683, y = 0.7395, xend = 0.45, yend = 0.740, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  geom_segment(x = 0.2915, y = 0.965, xend = 0.43, yend = 0.79, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  geom_segment(x = 0.5525, y = 0.965, xend = 0.44, yend = 0.8, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  geom_segment(x = 0.2915, y = 0.5135, xend = 0.18, yend = 0.32, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  geom_segment(x = 0.5525, y = 0.5135, xend = 0.67, yend = 0.32, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  geom_segment(x = 0.422, y = 1.04, xend = 0.42, yend = 0.85, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  geom_segment(x = 0.161, y = 0.890, xend = -0.02, yend = 0.8, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  geom_segment(x = 0.161, y = 0.589, xend = 0.161, yend = 0.4, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  geom_segment(x = 0.422, y = 0.438, xend = 0.25, yend = 0.3, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  geom_segment(x = 0.683, y = 0.589, xend = 0.683, yend = 0.4, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  geom_segment(x = 0.683, y = 0.890, xend = 0.5, yend = 0.82, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd")
 
@@
 ```{r}
-#| echo: false
-#| fig-cap: "Goodness of fit statistics from different NLDR techniques applied to training five spherical Gaussian cluster dataset.
-What is the best NLDR technique to represent the original data in 2D?"
-#| label: fig-diagnosticpltGau
-#| out-width: 100%
-#| fig-pos: H
-#| eval: false
-
-mse_plot_gau 
+#| label: fig-assign-data
+#| fig-cap: "Binning the data. Points are assigned to the nearest centroid. If a point is equidistant from multiple centroids, assigned to the lowest centroid."
+#| fig-width: 9
+#| fig-height: 6
+assign_data_vis1 + assign_data_vis2 +
+  plot_layout(ncol = 2)
 ```
 
-<!-- MSE plot to find which is the effective number of bins to use-->
-```{r}
-#| warning: false
-#| echo: false
-#| message: false
-#| eval: false
+## Density of a bin
 
-## UMAP
-## Prediction
-
-## Import data
-training_data_pbmc <- read_rds("data/pbmc3k/pbmc_pca_50.rds")
-training_data_pbmc <- training_data_pbmc[, 1:9] |>
-  mutate(ID = 1:NROW(training_data_pbmc))
-
-umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_30_min_dist_0.3.rds")
-umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc, 
-                                                                 x = "UMAP1", y = "UMAP2"))) |>
-  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
-                  "UMAP2" = "scaled_UMAP2")) |>
-  dplyr::mutate(ID = 1:NROW(umap_pbmc))
-
-#tot_num_bin_vec <- 2:51
-hex_size_vec <- seq(0.02, 2, by = 0.01)
-
-vec <- stats::setNames(rep("", 7), c("num_bins", "error", "mse", "num_bins_x", "num_bins_y", "hex_size", "num_non_empty_bins"))  ## Define column names
-
-mse_df_pbmc <- dplyr::bind_rows(vec)[0, ]
-mse_df_pbmc <- mse_df_pbmc |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(hex_size_vec)) {
-
-  num_bin_list <- calc_bins(data = umap_pbmc_scaled, 
-            x = "UMAP1", y = "UMAP2", 
-            hex_size = hex_size_vec[i], buffer_x = NA, buffer_y = NA)
-
-  num_bins_x <- num_bin_list$num_x
-  num_bins_y <- num_bin_list$num_y
-
-  model_object <- fit_highd_model( training_data = training_data_pbmc, 
-                                   nldr_df_with_id = umap_pbmc_scaled, 
-                                   x = "UMAP1", y = "UMAP2", 
-                                   num_bins_x = num_bins_x, 
-                                   num_bins_y = num_bins_y, 
-                                   x_start = NA, y_start = NA, 
-                                   buffer_x = NA, buffer_y = NA, 
-                                   hex_size = hex_size_vec[i],
-                                   is_rm_lwd_hex = FALSE, 
-                                   benchmark_to_rm_lwd_hex = NA, 
-                                   col_start_2d = "UMAP", 
-                                   col_start_highd = "PC_")
-
-  centroid_df_training <- model_object$df_bin_centroids
-  avg_df_training <- model_object$df_bin
-
-  pred_emb_list <- predict_emb(test_data = training_data_pbmc, 
-                                  df_bin_centroids = centroid_df_training, 
-                                  df_bin = avg_df_training, type_NLDR = "UMAP")
-
-  pred_df_training <- as.data.frame(do.call(cbind, pred_emb_list))
-
-  eval_list <- gen_summary(test_data = training_data_pbmc, 
-                                  prediction_df = pred_df_training, 
-                                  df_bin = avg_df_training, col_start = "PC_")
-
-  mse_df_pbmc <- mse_df_pbmc |>
-    tibble::add_row(num_bins = num_bins_x * num_bins_y,
-                    error = eval_list$error,
-                    mse = eval_list$mse,
-                    num_bins_x = num_bins_x,
-                    num_bins_y = num_bins_y,
-                    hex_size = hex_size_vec[i],
-                    num_non_empty_bins = NROW(centroid_df_training))
-
-}
-
-## If same total number of bins occurred only select ones with minimum error
-### Obtain duplicate bins
-dupli_bins <- mse_df_pbmc |> 
-  dplyr::count(num_bins) |> 
-  dplyr::filter(n > 1) |> 
-  dplyr::pull(num_bins)
-
-### Group split by duplicated bins
-duplicate_df_list <- mse_df_pbmc |>
-  dplyr::filter(num_bins %in% dupli_bins) |>
-  dplyr::arrange(num_bins) |>
-  dplyr::group_split(num_bins)
-
-### Obtain one row from duplicates which have lowest error and hexsize
-duplicate_df <- data.frame(matrix(nrow = 0, ncol = 0))
-
-for (i in 1:length(duplicate_df_list)) {
-
-  dd <- duplicate_df_list[[i]] |>
-    dplyr::filter(mse == min(duplicate_df_list[[i]]$mse)) |>
-    dplyr::filter(hex_size == min(duplicate_df_list[[i]]$hex_size))
-
-  duplicate_df <- dplyr::bind_rows(duplicate_df, dd)
-
-}
-
-### Obtain the mse_df with not duplicated bins
-not_dupli_df <- mse_df_pbmc |>
-  dplyr::filter(!(num_bins %in% dupli_bins))
-
-### Combine duplicated and not duplicated(corrected) bins dfs
-mse_df_pbmc <- dplyr::bind_rows(not_dupli_df, duplicate_df)
-```
-
-```{r}
-#| warning: false
-#| echo: false
-#| eval: false
-
-mse_plot_pbmc <- ggplot(mse_df_pbmc, aes(x = num_bins,
-                                       y = mse
-)) +
-  geom_point() +
-  geom_line() +
-   geom_vline(xintercept = 588, linetype="solid",
-                color = "red", size=0.8, alpha = 0.5) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  ylab("MSE") +
-  xlab("total number of bins")
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "Goodness of fit statistics from UMAP applied to training PBMC3k dataset.
-What is the effective number of bins in UMAP to create a 2D model?
-The MSE plot have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
-Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error.
-Using the elbow method, when the total number of bins is set to 588, the slope of the MSE plot experiences a sudden and noticeable change, resembling an elbow-like shape.
-This point indicates that adding less bins does not enough to capture the data structure."
-#| label: fig-msepltpbmc
-#| fig-pos: H
-#| eval: false
-
-mse_plot_pbmc
-```
-
-<!-- MSE plot to find which is the effective number of bins to use-->
-
-```{r}
-#| warning: false
-#| echo: false
-#| message: false
-#| eval: false
-
-## Import data
-training_data_mnist <- read_rds("data/mnist/mnist_10_pcs_of_digit_1.rds")
-training_data_mnist <- training_data_mnist |>
-  mutate(ID = 1:NROW(training_data_mnist))
-
-pacmap_minst <- read_rds("data/mnist/mnist_pacmap.rds")
-pacmap_minst_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = pacmap_minst, 
-                                    x = "PaCMAP1", y = "PaCMAP2"))) |>
-  dplyr::rename(c("PaCMAP1" = "scaled_PaCMAP1", 
-                  "PaCMAP2" = "scaled_PaCMAP2")) |>
-  dplyr::mutate(ID = 1:NROW(pacmap_minst))
-
-## UMAP
-## Prediction
-
-#tot_num_bin_vec <- 2:51
-hex_size_vec <- seq(0.02, 2, by = 0.01)
-
-vec <- stats::setNames(rep("", 6), c("num_bins", "aic", "mse", "num_bins_x", "num_bins_y", "hex_size"))  ## Define column names
-
-mse_df_mnist <- dplyr::bind_rows(vec)[0, ]
-mse_df_mnist <- mse_df_mnist |>
-  dplyr::mutate_if(is.character, as.numeric)
-
-for (i in 1:length(hex_size_vec)) {
-
-  num_bin_list <- calc_bins(data = pacmap_minst_scaled, 
-            x = "PaCMAP1", y = "PaCMAP2", 
-            hex_size = hex_size_vec[i], buffer_x = NA, buffer_y = NA)
-
-  num_bins_x <- num_bin_list$num_x
-  num_bins_y <- num_bin_list$num_y
-
-  model_object <- fit_highd_model( training_data = training_data_mnist, 
-                                   nldr_df_with_id = pacmap_minst_scaled, 
-                                   x = "PaCMAP1", y = "PaCMAP2", 
-                                   num_bins_x = num_bins_x, 
-                                   num_bins_y = num_bins_y, 
-                                   x_start = NA, y_start = NA, 
-                                   buffer_x = NA, buffer_y = NA, 
-                                   hex_size = hex_size_vec[i],
-                                   is_rm_lwd_hex = FALSE, 
-                                   benchmark_to_rm_lwd_hex = NA, 
-                                   col_start_2d = "PaCMAP", 
-                                   col_start_highd = "PC")
-
-  centroid_df_training <- model_object$df_bin_centroids
-  avg_df_training <- model_object$df_bin
-
-  pred_emb_list <- predict_emb(test_data = training_data_mnist, 
-                                  df_bin_centroids = centroid_df_training, 
-                                  df_bin = avg_df_training, type_NLDR = "PaCMAP")
-
-  pred_df_training <- as.data.frame(do.call(cbind, pred_emb_list))
-
-  eval_list <- gen_summary(test_data = training_data_mnist, 
-                                  prediction_df = pred_df_training, 
-                                  df_bin = avg_df_training, col_start = "PC")
-
-  mse_df_mnist <- mse_df_mnist |>
-    tibble::add_row(num_bins = num_bins_x * num_bins_y,
-                    aic = eval_list$aic,
-                    mse = eval_list$mse,
-                    num_bins_x = num_bins_x,
-                    num_bins_y = num_bins_y,
-                    hex_size = hex_size_vec[i])
-
-}
-
-## If same total number of bins occurred only select ones with minimum error
-### Obtain duplicate bins
-dupli_bins <- mse_df_mnist |> 
-  dplyr::count(num_bins) |> 
-  dplyr::filter(n > 1) |> 
-  dplyr::pull(num_bins)
-
-### Group split by duplicated bins
-duplicate_df_list <- mse_df_mnist |>
-  dplyr::filter(num_bins %in% dupli_bins) |>
-  dplyr::arrange(num_bins) |>
-  dplyr::group_split(num_bins)
-
-### Obtain one row from duplicates which have lowest error and hexsize
-duplicate_df <- data.frame(matrix(nrow = 0, ncol = 0))
-
-for (i in 1:length(duplicate_df_list)) {
-
-  dd <- duplicate_df_list[[i]] |>
-    dplyr::filter(mse == min(duplicate_df_list[[i]]$mse)) |>
-    dplyr::filter(hex_size == min(duplicate_df_list[[i]]$hex_size))
-
-  duplicate_df <- dplyr::bind_rows(duplicate_df, dd)
-
-}
-
-### Obtain the mse_df with not duplicated bins
-not_dupli_df <- mse_df_mnist |>
-  dplyr::filter(!(num_bins %in% dupli_bins))
-
-### Combine duplicated and not duplicated(corrected) bins dfs
-mse_df_mnist <- dplyr::bind_rows(not_dupli_df, duplicate_df)
-```
-
-```{r}
-#| warning: false
-#| echo: false
-#| eval: false
-
-mse_plot_mnist <- ggplot(mse_df_mnist, aes(x = num_bins,
-                                       y = mse
-)) +
-  geom_point() +
-  geom_line() +
-   geom_vline(xintercept = 180, linetype="solid",
-                color = "red", size=0.8, alpha = 0.5) +
-  theme_light() +
-  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
-        axis.title = element_text(size = 7),
-        axis.text = element_text(size = 7)) +
-  #scale_colour_manual(values = c("#377eb8")) +
-  ylab("MSE") +
-  xlab("total number of bins")
-
-```
-
-```{r}
-#| echo: false
-#| fig-cap: "Goodness of fit statistics from PaCMAP applied to training MNIST dataset.
-What is the effective number of bins in PaCMAP to create a 2D model?
-The MSE plot have a steep slope at the beginning, indicating that a smaller number of bins causes a larger amount of error.
-Then, the slope gradually declines or level off, indicating that a higher number of bins generates a smaller error.
-Using the elbow method, when the total number of bins is set to 180, the slope of the Mean Squared Error (MSE) plot experiences a sudden and noticeable change, resembling an elbow-like shape.
-This point indicates that adding less bins does not enough to capture the data structure."
-#| label: fig-diagnosticpltMNIST
-#| fig-pos: H
-#| eval: false
-
-mse_plot_mnist
-```
+hjisaghhasouho;uheew

### 2024-10-07 appendix.tex
Restructured appendix

@@
 \addtolength{\topmargin}{-1in}%
-
-\makeatletter
-\makeatother
-\makeatletter
-\makeatother
 \makeatletter
@@
 \makeatletter
+\makeatother
+\makeatletter
 \@ifpackageloaded{caption}{}{\usepackage{caption}}
@@
 \makeatother
-\makeatletter
-\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
-\makeatother
-\makeatletter
-\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
-\makeatother
-\makeatletter
-\makeatother
-\makeatletter
-\makeatother
+
 \ifLuaTeX
@@
 \hypersetup{
-  pdftitle={Appendix: Visualising How Non-linear Dimension Reduction Warps Your Data},
+  pdftitle={Appendix: Looking at Non-Linear Dimension Reductions as Models in the Data Space},
   pdfauthor={Jayani P.G. Lakshika; Dianne Cook; Paul Harrison; Michael Lydeamore; Thiyanga S. Talagala},
@@
 
-\title{\bf Appendix: Visualising How Non-linear Dimension Reduction
-Warps Your Data}
+\title{\bf Appendix: Looking at Non-Linear Dimension Reductions as
+Models in the Data Space}
 \author{
@@
 \newpage
-\spacingset{1} % DON'T change the spacing! (Default 1.9)
+\spacingset{1.9} % DON'T change the spacing!
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, boxrule=0pt, breakable, enhanced]}{\end{tcolorbox}}\fi
+\section{Computing hexagon grid
+configurations}\label{computing-hexagon-grid-configurations}
 
-Paramter tunning
+Given range of embedding component, \(r_2\), number of bins along the
+x-axis, \(b_1\), and buffer proportion, \(q\), hexagonal starting point
+coordinates, \(s_1 = -q\), and \(s_2 = -q \times r_2\). The purpose is
+to find width of the hexagon. \(a_1\), and number of bins along the
+y-axis, \(b_2\).
 
-\hypertarget{total-number-of-bins}{%
-\section{Total number of bins}\label{total-number-of-bins}}
+\begin{figure}[H]
 
-The total number of bins represents the overall count of hexagonal bins
-within the hexagonal grid. This count is determined by multiplying the
-number of bins along the x-axis (\(b_1\)) with the number of bins along
-the y-axis (\(b_2\)), according to the formula:
+\centering{
 
-\begin{equation}\protect\hypertarget{eq-equationp1}{}{
- b = b_1 \times b_2
-}\label{eq-equationp1}\end{equation}
+\includegraphics[width=1\textwidth,height=0.3\textheight]{appendix_files/figure-pdf/fig-hex-param-1.pdf}
 
-Here, \(b\) denotes the total number of bins. By adjusting the parameter
-\(b_1\), we have control over the total number of bins (\(b\)). Hence,
-fine-tuning \(b_1\) enables us to customize and optimize the total bin
-count based on the desired configuration along the x-axis.
+}
 
-\includegraphics[width=1\textwidth,height=\textheight]{appendix_files/figure-pdf/unnamed-chunk-9-1.pdf}
+\caption{\label{fig-hex-param}The components of the hexagon grid
+illustrating notation.}
 
-\hypertarget{benchmark-value-to-remove-low-density-hexagons}{%
-\section{Benchmark value to remove low-density
-hexagons}\label{benchmark-value-to-remove-low-density-hexagons}}
+\end{figure}%
 
-In the step of addressing low-density hexagons, which can arise from
-sparsely represented data in certain regions, we employ a systematic
-strategy. The goal is to ensure a more comprehensive coverage of the
-data by removing hexagons with low data density. To achieve this, we
-initiate the process by identifying, for each hex bin, the six nearest
-hex bins based on an equal 2D distance metric. Following this, we
-calculate the mean density (see Equation~\ref{eq-equationp3}), as
-outlined in the equation:
+Geometric arguments give rise to the following constraints.
 
-\begin{equation}\protect\hypertarget{eq-equationp2}{}{
- \text{standard count} = \frac{\text{count}}{\text{max count}}
-}\label{eq-equationp2}\end{equation}
+\(\text{min }a_1 \text{ s.t.}\)
 
-\begin{equation}\protect\hypertarget{eq-equationp3}{}{
- \text{mean density} = \frac{\text{standard count}}{6}
-}\label{eq-equationp3}\end{equation}
+\begin{equation}\phantomsection\label{eq-equation1}{
+s_1 - \frac{a_1}{2} < 0,
+}\end{equation}
 
-The standard count is derived from the number of observations in the hex
-bins (see Equation~\ref{eq-equationp2}). Next, we examine the
-distribution of mean densities across all hex bins and designate the
-first quartile as the benchmark value for removing low-density hexagons.
-Finally, hex bins with mean densities below this benchmark value are
-removed from consideration. This meticulous procedure ensures the
-elimination of regions with inadequate data density, allowing the focus
-to shift to areas with more significant data representation. The result
-is the preservation of the overall structure of the data in the
-low-dimensional space, as illustrated in \textbf{?@fig-bintorm} (c).
+\begin{equation}\phantomsection\label{eq-equation2}{
+s_1 + (b_1 - 1) \times a_1 > 1,
+}\end{equation}
 
-\includegraphics{appendix_files/figure-pdf/unnamed-chunk-14-1.pdf}
+\begin{equation}\phantomsection\label{eq-equation4}{
+s_2 - \frac{a_2}{2} < 0,
+}\end{equation}
 
-\begin{verbatim}
-Joining with `by = join_by(hexID)`
-\end{verbatim}
+\begin{equation}\phantomsection\label{eq-equation5}{
+s_2 + (b_2 - 1) \times a_2 > r_2.
+}\end{equation}
 
-\includegraphics{appendix_files/figure-pdf/unnamed-chunk-15-1.pdf}
+Since \(a_1\) and \(a_2\) are distances,
 
-\hypertarget{benchmark-value-to-remove-long-edges}{%
-\section{Benchmark value to remove long
-edges}\label{benchmark-value-to-remove-long-edges}}
+\[
+a_1, a_2 > 0.
+\] Also, \((s_1, s_2) \in (-0.1, -0.05)\) as these are multiplicative
+offsets in the negative direction.
 
-In this step, we aim to create a smoother surface in the low-dimensional
-space by removing long edges from the triangular mesh. This process
-helps eliminate outliers and noise while preserving important local
-relationships and intricate structures within the data. To achieve this,
-distances between vertices are sorted, and unique distance values are
-extracted. A data frame is created to calculate the differences between
-consecutive distance values, and the largest difference serves as a
-threshold to identify long edges. By removing edges that exceed this
-threshold, the algorithm refines the triangular mesh (see
-\textbf{?@fig-traingularmeshgr}).
+Equation~\ref{eq-equation1} can be rearranged as,
+
+\[
+a_1 > 2s_1
+\]
+
+which given \(s_1 < 0\) and \(a_1 > 0\) will \emph{always} be true. The
+same logic follows for Equation~\ref{eq-equation4} and substituting
+\(a_2 = \frac{\sqrt{3}}{2}a_1\), and \(s_2 = -q \times r_2\) to
+Equation~\ref{eq-equation4} can be written as,
+
+\[
+a_1 > -\frac{4}{\sqrt{3}}qr_2
+\]
+
+Also, substituting \(a_2 = \frac{\sqrt{3}}{2}a_1\),
+\(s_2 = -q \times r_2\) and rearranging Equation~\ref{eq-equation5}
+gives:
+
+\begin{equation}\phantomsection\label{eq-equation6}{
+a_1 > \frac{2(r_2 + qr_2)}{\sqrt{3}(b_2 - 1)}.
+}\end{equation}
+
+Similarly, substituting \(s_1 = -q\) Equation~\ref{eq-equation2}
+becomes,
+
+\begin{equation}\phantomsection\label{eq-equation7}{
+a_1 > \frac{(1 + q)}{(b_1 - 1)}.
+}\end{equation}
+
+This is a linear optimization problem. Therefore, the optimal solution
+must occur on a vertex. So, by setting Equation~\ref{eq-equation6}
+equals to Equation~\ref{eq-equation7} gives,
+
+\[
+\frac{2(r_2 + qr_2)}{\sqrt{3}(b_2 - 1)} = \frac{(1 + q)}{(b_1 - 1)}.
+\] After rearranging this,
+
+\[
+b_2 = 1 + \frac{2r_2(b_1 - 1)}{\sqrt{3}}
+\]
+
+and since \(b_2\) should be an integer,
+
+\begin{equation}\phantomsection\label{eq-equation8}{
+b_2 = \Big\lceil1 +\frac{2r_2(b_1 - 1)}{\sqrt{3}}\Big\rceil.
+}\end{equation}
+
+Furthermore, with known \(b_1\) and \(b_2\), by considering
+Equation~\ref{eq-equation2} or Equation~\ref{eq-equation5} as the
+\emph{binding} or \emph{active constraint}, can compute \(a_1\).
+
+If Equation~\ref{eq-equation2} is active, then,
+
+\[
+\frac{(1 + q)}{(b_1 - 1)} < \frac{2(r_2 + qr_2)}{\sqrt{3}(b_2 - 1)}.
+\]
+
+Rearranging this gives,
+
+\[
+r_2 > \frac{\sqrt{3}(b_2 - 1)}{2(b_1 - 1)}.
+\]
+
+Therefore, if this equality is true, then
+\(a_1 = \frac{(1+q)}{(b_1 - 1)}\), otherwise,
+\(a_1 = \frac{2r_2(1+q)}{\sqrt{3}(b_2 - 1)}\).
+
+\newpage
+
+\section{Binning the data}\label{binning-the-data}
+
+Points are assigned to the bin they fall into based on the nearest
+centroid. If a point is equidistant from multiple centroids, it is
+assigned to the centroid with the lowest hexagonal bin ID.
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=1\textheight]{figures/remove_long_edges_workflow.png}
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{appendix_files/figure-pdf/fig-assign-data-1.pdf}
 
@@
 
-\caption{\label{fig-rmlgmeth}A flow diagram detailing the steps taken to
-find the benchmark value to remove long edges.}
+\caption{\label{fig-assign-data}Binning the data. Points are assigned to
+the nearest centroid. If a point is equidistant from multiple centroids,
+assigned to the lowest centroid.}
 
-\end{figure}
+\end{figure}%
 
-\includegraphics{appendix_files/figure-pdf/unnamed-chunk-16-1.pdf}
+\section{Density of a bin}\label{density-of-a-bin}
 
-\includegraphics{appendix_files/figure-pdf/unnamed-chunk-17-1.pdf}
+hjisaghhasouho;uheew
 

### 2024-10-07 appendix.qmd
Added latex packages to appendix

@@
 tbl-cap-location: bottom
+header-includes: | 
+  \usepackage{amsmath}
+  \usepackage{float}
+  \usepackage{hyperref}
+  \usepackage[utf8]{inputenc}
+  \usepackage{bm}
+  \def\tightlist{}
+  \usepackage{setspace}
+  \newcommand\pD{$p\text{-}D$}
+  \newcommand\kD{$k\text{-}D$}
+  \newcommand\dD{$d\text{-}D$}
+  \newcommand\gD{$2\text{-}D$}
 ---
@@
 library(patchwork)
+library(kableExtra)
 
@@
 )
+```
+
+```{r}
+#| label: tbl-notation
+#| tbl-cap: "Summary of notation for describing new methodology."
+# Notation used in the paper
+
+notation_df <- read_csv("misc/notation.csv")
+
+# Create the table
+kable(notation_df, 
+      format = "latex", 
+      booktabs = TRUE, escape = FALSE) |>
+  kable_styling(position = "center", 
+                full_width = FALSE, 
+                font_size = 12) |>
+  row_spec(0, bold = TRUE) |>
+  column_spec(1:2, width = c("3cm", "12cm"))
 ```
@@
 
-## Density of a bin
+## Area of a hexagon
 
-hjisaghhasouho;uheew
+The area of a hexagon is defined as $A = \frac{3\sqrt{3}}{2}l^2$, where $l$ is the side length of the hexagon. $l$ can be computed using $a_1$ and $a_2$.

### 2024-10-07 appendix.tex
Added latex packages to appendix

@@
 \addtolength{\topmargin}{-1in}%
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
+\usepackage{amsmath}
+\usepackage{float}
+\usepackage{hyperref}
+\usepackage[utf8]{inputenc}
+\usepackage{bm}
+\def\tightlist{}
+\usepackage{setspace}
+\newcommand\pD{$p\text{-}D$}
+\newcommand\kD{$k\text{-}D$}
+\newcommand\dD{$d\text{-}D$}
+\newcommand\gD{$2\text{-}D$}
 \makeatletter
@@
 \spacingset{1.9} % DON'T change the spacing!
+
+\begin{table}
+
+\centering{
+
+\centering\begingroup\fontsize{12}{14}\selectfont
+
+\begin{tabular}{>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{12cm}}
+\toprule
+\textbf{Notation} & \textbf{Description}\\
+\midrule
+$n, p, k$ & number of observations, variables, embedding dimension, respectively\\
+$\mathbfit{X}, \mathbfit{x}$ & $p$-dimensional data (population, sample)\\
+$\mathbfit{y}$ & $k$-dimensional layout\\
+$P$ & orthonormal basis, generating a $d\text{-}dimensional$ linear projection of $p$-dimensional data\\
+$T$ & true  model\\
+\addlinespace
+$g$ & functional mapping from \pD{} to \kD{}, especially as prescribed by NLDR\\
+$\mathbfit{\theta}$ & (Hyper-) parameters for NLDR method\\
+$r$ & ranges of the embedding components\\
+$C^{(j)}$ & $j$-dimensional bin centers\\
+$(b_1, b_2)$ & number of bins in each direction\\
+\addlinespace
+$(a_1, a_2)$ & binwidths, distance between centroids in each direction\\
+$(s_1, \ s_2)$ & starting coordinates of the hexagonal grid\\
+$q$ & buffer to ensure hexgrid covers data, proportion of data range, 0-1\\
+$m$ & number of non-empty bins\\
+$b$ & number of  hexagons in the grid\\
+\addlinespace
+$h$ & hexagonal id\\
+$l$ & side length\\
+$A$ & area\\
+\bottomrule
+\end{tabular}
+\endgroup{}
+
+}
+
+\caption{\label{tbl-notation}Summary of notation for describing new
+methodology.}
+
+\end{table}%
 
@@
 
-\section{Density of a bin}\label{density-of-a-bin}
+\section{Area of a hexagon}\label{area-of-a-hexagon}
 
-hjisaghhasouho;uheew
+The area of a hexagon is defined as \(A = \frac{3\sqrt{3}}{2}l^2\),
+where \(l\) is the side length of the hexagon. \(l\) can be computed
+using \(a_1\) and \(a_2\).
 

### 2024-10-07 paper.qmd
Added side length to Figure 4

@@
 
+a1_temp <- hb_obj_notation$a1
+a2_temp <- hb_obj_notation$a2
+l_temp <- quad(a=3, b = 2 * a2_temp, c = -(a2_temp^2 + a1_temp^2))
+
 ## Data set with all centroids
@@
 hex_grid_temp <- hb_obj_notation$hex_poly
+
+hex_grid_temp45 <- hex_grid_temp |> 
+  filter(hex_poly_id == 45)
 
@@
              label = expression(a[2]))
+l <- tibble(x = hex_grid_temp45$x[2],
+            xend = hex_grid_temp45$x[3],
+            y = hex_grid_temp45$y[2],
+            yend = hex_grid_temp45$y[3],
+            label = expression(l))
+
 hex_param_vis <- ggplot() + 
@@
       color = "black") +
+  geom_segment(data = l, aes(
+      x = x, #-0.15, 
+      y = y, #-0.1*r2 + 0.1807896*2, 
+      xend = xend, #-0.15, 
+      yend = yend), #-0.1*r2 + 0.1807896*3),
+      arrow = arrow(length = unit(0.03, "npc"),
+                               ends = "both"), 
+      color = "black") + 
+    annotate("text", x=l$x + 0.03, y=(l$y+l$yend)/2, 
+             label = expression(l), 
+             color = "black", hjust=-0.2) +
   coord_equal()

### 2024-10-07 appendix.qmd
Added initial computation for the side length
Added intial triangle
Added notations for the triangles

@@
 The area of a hexagon is defined as $A = \frac{3\sqrt{3}}{2}l^2$, where $l$ is the side length of the hexagon. $l$ can be computed using $a_1$ and $a_2$.
+
+```{r}
+
+triangle_temp <- tibble(
+  x = c(0, 4, 0),
+  y = c(0, 0, 3)
+)
+
+triangle_line_temp <- tibble(
+  x = c(4, 0, 4),
+  xend = c(0, 0, 0),
+  y = c(0, 0, 0),
+  yend = c(0, 3, 3)
+)
+
+triangle_arrow_temp <- tibble(
+  x = c(4 + 0.04, 0 - 0.06, 4 + 0.06),
+  xend = c(0 + 0.04, 0 - 0.06, 0 + 0.06),
+  y = c(0 - 0.04, 0 - 0.06, 0 + 0.06),
+  yend = c(0 - 0.04, 3 - 0.06, 3 + 0.06)
+)
+
+triangle_plot <- ggplot() +
+  # geom_point(
+  #   data = triangle_temp,
+  #   aes(
+  #     x = x,
+  #     y = y
+  #   )
+  # ) +
+  geom_segment(
+    data = triangle_line_temp,
+    aes(
+      x = x,
+      xend = xend,
+      y = y,
+      yend = yend
+    )
+  ) + 
+  annotate("text", x=-1, y= 1.5, 
+           label = expression(frac(a[2] - a[1], 2)), 
+           color = "black", hjust=-0.2) + 
+  annotate("text", x=1.5, y= -0.3, 
+           label = expression(frac(a[1], 2)), 
+           color = "black", hjust=-0.2) + 
+  annotate("text", x=2.5, y= 1.5, 
+           label = expression(l), 
+           color = "black", hjust=-0.2)
+```
+
+```{r}
+#| out-height: 30%
+triangle_plot
+```
+
+By pythagorean theorem,
+
+$$
+l^2 = (\frac{a_1}{2})^2 + (\frac{a_2 - l}{2})^2
+$$
+$$
+l^2 - (\frac{a_2 - l}{2})^2 = (\frac{a_1}{2})^2
+$$
+$$
+[l - (\frac{a_2 - l}{2})][l + (\frac{a_2 - l}{2})] = (\frac{a_1}{2})^2
+$$
+
+$$
+3l^2 + 2la_2 - (a_1^2 + a_2^2) = 0
+$$
+
+$$
+l = \frac{-2a_2 \pm \sqrt{4a_2^2 - 24[-(a_1^2 + a_2^2)]}}{6}
+$$
+$$
+l = \frac{-a_2 \pm \sqrt{a_2^2 - 6[-(a_1^2 + a_2^2)]}}{3}
+$$

### 2024-10-07 appendix.tex
Added initial computation for the side length
Added intial triangle
Added the notations
Added the equations with a flow

@@
 
+\newpage
+
 \section{Area of a hexagon}\label{area-of-a-hexagon}
@@
 
+\begin{figure}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=0.3\textheight]{appendix_files/figure-pdf/fig-tri-param-1.pdf}
+
+}
+
+\caption{\label{fig-tri-param}The components of the right triangle
+illustrating notation.}
+
+\end{figure}%
+
+By applying the Pythagorean theorem, we obtain,
+
+\[
+l^2 = \left(\frac{a_1}{2}\right)^2 + \left(\frac{a_2 - l}{2}\right)^2.
+\] Next, rearranging the terms, we get,
+
+\[
+l^2 - \left(\frac{a_2 - l}{2}\right)^2 = \left(\frac{a_1}{2}\right)^2,
+\]
+
+\[
+\left[l - \left(\frac{a_2 - l}{2}\right)\right]\left[l + \left(\frac{a_2 - l}{2}\right)\right] = \left(\frac{a_1}{2}\right)^2,
+\]
+
+\[
+3l^2 + 2a_2l - (a_1^2 + a_2^2) = 0.
+\]
+
+Finally, by solving the quadratic equation, we compute,
+
+\[
+l = \frac{-2a_2 \pm \sqrt{4a_2^2 - 24[-(a_1^2 + a_2^2)]}}{6},
+\]
+
+\[
+l = \frac{-a_2 \pm \sqrt{a_2^2 - 6[-(a_1^2 + a_2^2)]}}{3},
+\]
+
+where \(l > 0\).
+
 \end{document}

### 2024-10-08 paper.qmd
Updated Figure 9 (d) y-axis title

@@
   scale_x_continuous(breaks = sort(unique(round(error_two_curvy_clust_umap$a1, 2)))[c(1, 4, 8, 10, 11, 12, 13, 14)]) +
-  labs(x = expression(paste("binwidth (", a[1], ")")), y = paste("proportion of densities")) +
+  labs(x = expression(paste("binwidth (", a[1], ")")), y = paste("Relative proportion density")) +
   ggtitle("(d)") +

### 2024-10-08 paper.qmd
Updated Figure 10

@@
   xlab(expression(d^{(2)})) +
-  #ggtitle("(d)") +
+  ggtitle("(a)") +
   theme_minimal() +
@@
   y = c(0.5, 0.5),
-  text = c("c1", "b1")
+  text = c("e1", "d1")
 )
@@
   xlab(expression(d^{(2)})) +
-  #ggtitle("(d)") +
+  ggtitle("(b)") +
   theme_minimal() +

### 2024-10-08 paper.qmd
Removed jitter from Fig 10

@@
   geom_point(alpha = 0.5) +
-  geom_jitter() +
+  #geom_jitter() +
   scale_x_continuous(breaks = sort(unique(as.numeric(distance_two_curvy_dist$bin))),

### 2024-10-08 paper.qmd
Update Fig10

--- 2024-10-08 paper.qmd
@@
   xlab(expression(d^{(2)})) +
-  ggtitle("(a)") +
+  #ggtitle("(a)") +
   theme_minimal() +
@@
 #| fig-cap: "Distribution of edge lengths and wireframes in \\gD{} and $7\\text{-}D$ with Delaunay triangulation, and two choices of long edge removal: (b) benchmark = 4.5$a_1$ and (c) benchmark = 2$a_1$, where $a_1$ = 0.09."
-#| fig-height: 15
-#| fig-width: 15
-#| out-height: 70%
-
-wrap_plots(distance_points,distance_hist, ncol = 2) / wrap_plots(trimesh_two_curvy2, trimesh_two_curvy_removed2, trimesh_two_curvy_removed1,
-two_curvy_proj_umap_model_delaunay,
-two_curvy_proj_umap_model_benchmark2,
-two_curvy_proj_umap_model, ncol = 3) 
+#| fig-height: 4
+#| fig-width: 20
+
+free(distance_points) + plot_annotation(tag_levels = 'A') +
+ (trimesh_two_curvy2 + trimesh_two_curvy_removed2 + trimesh_two_curvy_removed1 + two_curvy_proj_umap_model_delaunay + two_curvy_proj_umap_model_benchmark2 + two_curvy_proj_umap_model) +
+  plot_layout(widths = c(0.5, 1, 1, 1)) +
+  plot_annotation(tag_levels = list(c("(a)", "(b)"))) 
 ```

### 2024-10-08 paper.qmd
Updated Figure 7

@@
 plot_predict_umap_two_curvy <- ggplot(
-  data = jittered_points_df, 
-  aes(x = UMAP1,
-      y = UMAP2)) +
+  data = true_pred_df, 
+  aes(x = pred_UMAP_1,
+      y = pred_UMAP_2)) +
   geom_point(alpha = 0.5) +
@@
 #| out-height: 80%
-#| fig-cap: "Comparison of prediction generated using the exiting `umap` R package's prediction method and our method: (a) A view of the true model in projection from $7\\text{-}D$, (b) predicted data from the `umap` prediction method, (c) predicted data from our method, and (d) comparison of predictions from both methods.
-In plot (d), points representing the same data but predicted by different methods are connected by lines.
-Both prediction methods work similarly within the clusters, as each method predicts the data within the cluster with slight differences."
+#| fig-cap: "Comparison of prediction generated using the exiting `umap` R package's prediction method and our method: (a) A view of the true model in projection from $7\\text{-}D$, (b) predicted data from the `umap` prediction method, and (c) predicted data from our method."
 
 proj_plot_two_curvy_true + plot_predict_umap_fun_two_curvy +plot_predict_umap_two_curvy +
-  plot_predict_umap_connect_two_curvy +
-  plot_layout(ncol = 2)
+  plot_layout(ncol = 3)
 ```

### 2024-10-08 paper.tex
Update Fig10
Updated Figure 7

@@
 method: (a) A view of the true model in projection from \(7\text{-}D\),
-(b) predicted data from the \texttt{umap} prediction method, (c)
-predicted data from our method, and (d) comparison of predictions from
-both methods. In plot (d), points representing the same data but
-predicted by different methods are connected by lines. Both prediction
-methods work similarly within the clusters, as each method predicts the
-data within the cluster with slight differences.}
+(b) predicted data from the \texttt{umap} prediction method, and (c)
+predicted data from our method.}
 
@@
 
-\includegraphics[width=1\textwidth,height=0.7\textheight]{paper_files/figure-pdf/fig-rm-lg-1.pdf}
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-rm-lg-1.pdf}
 

### 2024-10-11 paper.qmd
Added relative proportion computation to the tuning section
Updated the example data structure..Need to work on the true model

@@
 #| label: two-curvy-training
-training_data_two_curvy <- read_rds("data/two_curvy_clust/two_curvy_clust_data.rds")
+training_data_two_curvy <- read_rds("data/two_curvy_diff_clust/two_curvy_diff_clust_data.rds")
 ```
@@
 
-true_model_df <- read_rds("data/two_curvy_clust/two_curvy_clust_true_model.rds")
-wireframe_true_model <- read_rds("data/two_curvy_clust/two_curvy_clust_true_model_connections.rds")
+true_model_df <- read_rds("data/two_curvy_diff_clust/two_curvy_diff_clust_true_model.rds")
+wireframe_true_model <- read_rds("data/two_curvy_diff_clust/two_curvy_diff_clust_true_model_connections.rds")
 
@@
 #| label: umap-two-curvy
-umap_two_curvy <- read_rds(file = "data/two_curvy_clust/two_curvy_clust_umap_n-neigbors_15_min-dist_0.1.rds")
+umap_two_curvy <- read_rds(file = "data/two_curvy_diff_clust/two_curvy_diff_clust_umap_n-neigbors_15_min-dist_0.1.rds")
 
@@
 
-error_two_curvy_clust_umap <- read_rds("data/two_curvy_clust/error_two_curvy_clust_umap.rds")
+error_two_curvy_diff_clust_umap <- read_rds("data/two_curvy_diff_clust/error_two_curvy_diff_clust_umap.rds")
 
 ## Find the minimum MSE when have duplicate a1
-error_two_curvy_clust_umap <- error_two_curvy_clust_umap |>
+error_two_curvy_diff_clust_umap <- error_two_curvy_diff_clust_umap |>
   group_by(a1) |>
@@
 
-base_line_prop_dens <- error_two_curvy_clust_umap |>
+base_line_prop_dens <- error_two_curvy_diff_clust_umap |>
   filter(a1 == min(a1)) |>
@@
 
-error_two_curvy_clust_umap <- error_two_curvy_clust_umap |>
+error_two_curvy_diff_clust_umap <- error_two_curvy_diff_clust_umap |>
   mutate(prop_comp = prop_dens/base_line_prop_dens)
 
-mse_two_curvy_b <- ggplot(error_two_curvy_clust_umap, 
+mse_two_curvy_b <- ggplot(error_two_curvy_diff_clust_umap, 
                      aes(x = a1, 
@@
   geom_point(size = 1) +
-  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_clust_umap$a1, 2)))[c(1, 4, 8, 10, 11, 12, 13, 14)]) +
+  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_diff_clust_umap$a1, 2)))[c(1, 4, 8, 10, 11, 12, 13, 14)]) +
   labs(x = expression(paste("binwidth (", a[1], ")")), y = "MSE") +
@@
 
-prop_dens_a1 <- ggplot(error_two_curvy_clust_umap,
+prop_dens_a1 <- ggplot(error_two_curvy_diff_clust_umap,
                      aes(x = a1,
@@
   # scale_y_continuous(breaks = sort(unique(error_two_curvy$b_non_empty))[-3]) +
-  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_clust_umap$a1, 2)))[c(1, 4, 8, 10, 11, 12, 13, 14)]) +
+  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_diff_clust_umap$a1, 2)))[c(1, 4, 8, 10, 11, 12, 13, 14)]) +
   labs(x = expression(paste("binwidth (", a[1], ")")), y = paste("Relative proportion density")) +
@@
 
-error_two_curvy_clust_umap <- error_two_curvy_clust_umap |>
+error_two_curvy_diff_clust_umap <- error_two_curvy_diff_clust_umap |>
   mutate(prop_bins = b_non_empty/b)
 
-a1_m_two_curvy <- ggplot(error_two_curvy_clust_umap,
+a1_m_two_curvy <- ggplot(error_two_curvy_diff_clust_umap,
                      aes(x = a1,
@@
   # scale_y_continuous(breaks = sort(unique(error_two_curvy$b_non_empty))[-3]) +
-  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_clust_umap$a1, 2)))[c(1, 4, 8, 10, 11, 12, 13, 14)]) +
+  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_diff_clust_umap$a1, 2)))[c(1, 4, 8, 10, 11, 12, 13, 14)]) +
   labs(x = expression(paste("binwidth (", a[1], ")")), y = expression(paste("proportion of non-empty bins ", bgroup("(", frac(m, b), ")")))) +
@@
 
-predict_umap_two_curvy <- read_rds(file = "data/two_curvy_clust/two_curvy_clust_umap_predict_true.rds")
+predict_umap_two_curvy <- read_rds(file = "data/two_curvy_diff_clust/two_curvy_diff_clust_umap_predict_true.rds")
 
@@
 
-### Diagnostics
-
-#### Relative proportion density
-
-The area of a hexagon is defined as $A = \frac{3\sqrt{3}}{2}l^2$ where $l$ is the side length of the hexagon.
-If we know $a_1$ and $a_2$, $l$ can be computed (see appendix).
-The density of a hexagon grid is calculated as $\frac{\sum^{h}_{i=1}n_h}{A}$ and the proportion is $\frac{\sum^{h}_{i=1}n_h}{A \times b}$.
-The baseline proportion is the proportion density at the smallest possible value of $a_1$.
-The relative proportion density is the ratio of the observed proportion density to the baseline proportion density.
-
 ### Tuning
@@
 
+#### Measurement of capturing the data shape in 2-D
+
+The area of a hexagon is defined as $A = \frac{3\sqrt{3}}{2}l^2$ where $l$ is the side length of the hexagon.
+If we know $a_1$ and $a_2$, $l$ can be computed (see appendix).
+The density of a hexagon grid is calculated as $\frac{\sum^{h}_{i=1}n_h}{A}$ and the proportion is $\frac{\sum^{h}_{i=1}n_h}{A \times b}$.
+The baseline proportion is the proportion density at the smallest possible value of $a_1$.
+The relative proportion density is the ratio of the observed proportion density to the baseline proportion density.
+
 #### Removal of low density bins
@@
 
-error_rm_two_curvy <- read_rds("data/two_curvy_clust/error_rm_lwd_diff_bin.rds") |>
+error_rm_two_curvy <- read_rds("data/two_curvy_diff_clust/error_rm_lwd_diff_bin.rds") |>
   mutate(bin1  = as.factor(bin1))
@@
 #| label: read-two-curvy-clust-nldr
-# Read a variety of different NLDR representations of two_curvy_clust
+# Read a variety of different NLDR representations of two_curvy_diff_clust
 # and plot them on same aspect ratio
-tsne_two_curvy_clust <- read_rds("data/two_curvy_clust/two_curvy_clust_tsne_perplexity_30.rds")
-
-nldr_two_curvy_clust1 <- tsne_two_curvy_clust |>
+tsne_two_curvy_diff_clust <- read_rds("data/two_curvy_diff_clust/two_curvy_diff_clust_tsne_perplexity_30.rds")
+
+nldr_two_curvy_diff_clust1 <- tsne_two_curvy_diff_clust |>
   ggplot(aes(x = tSNE1,
@@
 
-umap_two_curvy_clust <- read_rds("data/two_curvy_clust/two_curvy_clust_umap_n-neigbors_15_min-dist_0.1.rds")
-
-nldr_two_curvy_clust2 <- umap_two_curvy_clust |>
+umap_two_curvy_diff_clust <- read_rds("data/two_curvy_diff_clust/two_curvy_diff_clust_umap_n-neigbors_15_min-dist_0.1.rds")
+
+nldr_two_curvy_diff_clust2 <- umap_two_curvy_diff_clust |>
   ggplot(aes(x = UMAP1,
@@
 
-phate_two_curvy_clust <- read_rds("data/two_curvy_clust/two_curvy_clust_phate_knn_5.rds")
-
-nldr_two_curvy_clust3 <- phate_two_curvy_clust |>
+phate_two_curvy_diff_clust <- read_rds("data/two_curvy_diff_clust/two_curvy_diff_clust_phate_knn_5.rds")
+
+nldr_two_curvy_diff_clust3 <- phate_two_curvy_diff_clust |>
   ggplot(aes(x = PHATE1,
@@
 
-trimap_two_curvy_clust <- read_rds("data/two_curvy_clust/two_curvy_clust_trimap_n-inliers_12_n-outliers_4_n-random_3.rds")
-
-nldr_two_curvy_clust4 <- trimap_two_curvy_clust |>
+trimap_two_curvy_diff_clust <- read_rds("data/two_curvy_diff_clust/two_curvy_diff_clust_trimap_n-inliers_12_n-outliers_4_n-random_3.rds")
+
+nldr_two_curvy_diff_clust4 <- trimap_two_curvy_diff_clust |>
   ggplot(aes(x = TriMAP1,
@@
 
-pacmap_two_curvy_clust <- read_rds("data/two_curvy_clust/two_curvy_clust_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")
-
-nldr_two_curvy_clust5 <- pacmap_two_curvy_clust |>
+pacmap_two_curvy_diff_clust <- read_rds("data/two_curvy_diff_clust/two_curvy_diff_clust_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")
+
+nldr_two_curvy_diff_clust5 <- pacmap_two_curvy_diff_clust |>
   ggplot(aes(x = PaCMAP1,
@@
 ```{r}
-#| label: combine-data-two_curvy_clust
-
-error_two_curvy_clust_umap <- read_rds("data/two_curvy_clust/error_two_curvy_clust_umap.rds")
-error_two_curvy_clust_tsne <- read_rds("data/two_curvy_clust/error_two_curvy_clust_tsne.rds")
-error_two_curvy_clust_phate <- read_rds("data/two_curvy_clust/error_two_curvy_clust_phate.rds")
-error_two_curvy_clust_trimap <- read_rds("data/two_curvy_clust/error_two_curvy_clust_trimap.rds")
-error_two_curvy_clust_pacmap <- read_rds("data/two_curvy_clust/error_two_curvy_clust_pacmap.rds")
-
-error_two_curvy_clust <- bind_rows(error_two_curvy_clust_umap, 
-                         error_two_curvy_clust_tsne,
-                         error_two_curvy_clust_phate,
-                         error_two_curvy_clust_trimap,
-                         error_two_curvy_clust_pacmap)
-
-error_two_curvy_clust <- error_two_curvy_clust |>
+#| label: combine-data-two_curvy_diff_clust
+
+error_two_curvy_diff_clust_umap <- read_rds("data/two_curvy_diff_clust/error_two_curvy_diff_clust_umap.rds")
+error_two_curvy_diff_clust_tsne <- read_rds("data/two_curvy_diff_clust/error_two_curvy_diff_clust_tsne.rds")
+error_two_curvy_diff_clust_phate <- read_rds("data/two_curvy_diff_clust/error_two_curvy_diff_clust_phate.rds")
+error_two_curvy_diff_clust_trimap <- read_rds("data/two_curvy_diff_clust/error_two_curvy_diff_clust_trimap.rds")
+error_two_curvy_diff_clust_pacmap <- read_rds("data/two_curvy_diff_clust/error_two_curvy_diff_clust_pacmap.rds")
+
+error_two_curvy_diff_clust <- bind_rows(error_two_curvy_diff_clust_umap, 
+                         error_two_curvy_diff_clust_tsne,
+                         error_two_curvy_diff_clust_phate,
+                         error_two_curvy_diff_clust_trimap,
+                         error_two_curvy_diff_clust_pacmap)
+
+error_two_curvy_diff_clust <- error_two_curvy_diff_clust |>
   mutate(a1 = round(a1, 2)) |>
@@
 ```{r}
-#| label: error-comp-two_curvy_clust
-
-error_plot_two_curvy_clust <- ggplot(error_two_curvy_clust, 
+#| label: error-comp-two_curvy_diff_clust
+
+error_plot_two_curvy_diff_clust <- ggplot(error_two_curvy_diff_clust, 
                            aes(x = a1, 
@@
   #            color = "black", linewidth=0.8, alpha = 0.5) +
-  scale_x_continuous(breaks = sort(unique(error_two_curvy_clust$a1))[seq(1, length(unique(error_two_curvy_clust$a1)), by = 5)]) +
+  scale_x_continuous(breaks = sort(unique(error_two_curvy_diff_clust$a1))[seq(1, length(unique(error_two_curvy_diff_clust$a1)), by = 5)]) +
   scale_color_manual(values=c('#e41a1c','#ff7f00','#4daf4a', "#a65628",'#636363')) +
@@
 Layout a is the best choice."
-#| label: fig-two_curvy_clust-mse
+#| label: fig-two_curvy_diff_clust-mse
 #| fig-pos: H
@@
 
-free(error_plot_two_curvy_clust) + wrap_plots(nldr_two_curvy_clust1, nldr_two_curvy_clust2, nldr_two_curvy_clust3, 
-                                    nldr_two_curvy_clust4, nldr_two_curvy_clust5, ncol = 2)
+free(error_plot_two_curvy_diff_clust) + wrap_plots(nldr_two_curvy_diff_clust1, nldr_two_curvy_diff_clust2, nldr_two_curvy_diff_clust3, 
+                                    nldr_two_curvy_diff_clust4, nldr_two_curvy_diff_clust5, ncol = 2)
 ```
@@
 two_curvy_tsne_scaled_obj <- gen_scaled_data(
-  data = tsne_two_curvy_clust)
+  data = tsne_two_curvy_diff_clust)
 
@@
 
-tsne_two_curvy_clust_plt <- tsne_two_curvy_scaled |>
+tsne_two_curvy_diff_clust_plt <- tsne_two_curvy_scaled |>
   ggplot(aes(x = tSNE1,
@@
 
-nldr_two_curvy_original2 + tsne_two_curvy_clust_plt +
+nldr_two_curvy_original2 + tsne_two_curvy_diff_clust_plt +
   trimesh_two_curvy_removed1_with_data + trimesh_two_curvy_removed1_tsne_with_data +

### 2024-10-11 paper.tex
Added relative proportion computation to the tuning section
Updated the example data structure..Need to work on the true model

@@
 
-\subsection{Diagnostics}\label{diagnostics}
-
-\subsubsection{Relative proportion
-density}\label{relative-proportion-density}
-
-The area of a hexagon is defined as \(A = \frac{3\sqrt{3}}{2}l^2\) where
-\(l\) is the side length of the hexagon. If we know \(a_1\) and \(a_2\),
-\(l\) can be computed (see appendix). The density of a hexagon grid is
-calculated as \(\frac{\sum^{h}_{i=1}n_h}{A}\) and the proportion is
-\(\frac{\sum^{h}_{i=1}n_h}{A \times b}\). The baseline proportion is the
-proportion density at the smallest possible value of \(a_1\). The
-relative proportion density is the ratio of the observed proportion
-density to the baseline proportion density.
-
 \subsection{Tuning}\label{tuning}
@@
 choices of \(b_1\) on the MSE of the fitted model.
+
+\subsubsection{Measurement of capturing the data shape in
+2-D}\label{measurement-of-capturing-the-data-shape-in-2-d}
+
+The area of a hexagon is defined as \(A = \frac{3\sqrt{3}}{2}l^2\) where
+\(l\) is the side length of the hexagon. If we know \(a_1\) and \(a_2\),
+\(l\) can be computed (see appendix). The density of a hexagon grid is
+calculated as \(\frac{\sum^{h}_{i=1}n_h}{A}\) and the proportion is
+\(\frac{\sum^{h}_{i=1}n_h}{A \times b}\). The baseline proportion is the
+proportion density at the smallest possible value of \(a_1\). The
+relative proportion density is the ratio of the observed proportion
+density to the baseline proportion density.
 
@@
 
-\includegraphics[width=1\textwidth,height=0.6\textheight]{paper_files/figure-pdf/fig-two_curvy_clust-mse-1.pdf}
-
-}
-
-\caption{\label{fig-two_curvy_clust-mse}Assessing which of the 5 NLDR
-layouts on the two C-shaped clusters data is the better representation
-using MSE for varying binwidth (\(a_1\)). Colour used for the lines and
-points in the left plot and in the scatterplots represents NLDR layout
-(a-e). Layout d is universally poor. Layouts a that show two close
-clusters are universally suboptimal. Layout a is the best choice.}
+\includegraphics[width=1\textwidth,height=0.6\textheight]{paper_files/figure-pdf/fig-two_curvy_diff_clust-mse-1.pdf}
+
+}
+
+\caption{\label{fig-two_curvy_diff_clust-mse}Assessing which of the 5
+NLDR layouts on the two C-shaped clusters data is the better
+representation using MSE for varying binwidth (\(a_1\)). Colour used for
+the lines and points in the left plot and in the scatterplots represents
+NLDR layout (a-e). Layout d is universally poor. Layouts a that show two
+close clusters are universally suboptimal. Layout a is the best choice.}
 

### 2024-10-11 paper.tex
Updated Figure 8 title

@@
 \(b = 391 \text{ } (23, \text{ }17)\). Color indicates standardized
-counts, dark indicating high count and light indicates low count. Using
-the MSE of the model fit in \(7\text{-}D\) helps decide on a useful
-choice of number of bins.}
+counts, dark indicating high count and light indicates low count. At the
+smallest bin size, the data structure is discontinuous, suggesting that
+there are too many bins. Using the MSE of the model fit in
+\(7\text{-}D\) helps decide on a useful choice of number of bins.}
 

### 2024-10-11 paper.qmd
Updated Figure 8 title
Updated Figure 9

@@
              linewidth=1) +
-  geom_vline(xintercept = 0.12, linetype=2,
+  geom_vline(xintercept = 0.13, linetype=2,
              color = "#1f78b4", linewidth=1) +
@@
 Color indicates standardized counts, dark indicating high count and light indicates low count.
+At the smallest bin size, the data structure is discontinuous, suggesting that there are too many bins.
 Using the MSE of the model fit in $7\\text{-}D$ helps decide on a useful choice of number of bins."
@@
 A good benchmark value for these parameters is when the MSE drops and then flattens out.
-Three binwidth choices were made: $0.05$, $0.09$, and $0.12$ to investigate.
+Three binwidth choices were made: $0.05$, $0.09$, and $0.13$ to investigate.
 As the binwidth increases, the proportion of non-empty bins also increases.
-There are two peaks at binwidths $0.09$ and $0.12$.
+There are two peaks at binwidths $0.09$ and $0.13$.
 The relative proportion density decreases and levels off.

### 2024-10-11 paper.qmd
Updated Figure 10

@@
   distance_two_curvy_dist, 
-  aes(x = as.numeric(bin),
+  aes(x = distance,
       y = dist_highd)) +
@@
   #geom_jitter() +
-  scale_x_continuous(breaks = sort(unique(as.numeric(distance_two_curvy_dist$bin))),
-                     labels = c(
-                       expression(a[1]),
-                       expression(2*a[1]),
-                       expression(3*a[1]),
-                       expression(4*a[1]),
-                       expression(5*a[1]),
-                       expression(6*a[1]),
-                       expression(7*a[1]),
-                       expression(8*a[1])
-                     )) +
+  # scale_x_continuous(breaks = sort(unique(as.numeric(distance_two_curvy_dist$bin))),
+  #                    labels = c(
+  #                      expression(a[1]),
+  #                      expression(2*a[1]),
+  #                      expression(3*a[1]),
+  #                      expression(4*a[1]),
+  #                      expression(5*a[1]),
+  #                      expression(6*a[1]),
+  #                      expression(7*a[1]),
+  #                      expression(8*a[1])
+  #                    )) +
   # geom_vline(xintercept = 2,
@@
   geom_point(alpha=0.2) +
-  scale_colour_continuous_sequential(palette = "Blues") +
+  scale_colour_continuous_sequential(palette = "YlOrRd", n_interp = 20) +
   geom_rect(aes(xmin = 0.8, xmax = 0.9, ymin = 0.45, ymax = 0.6),

### 2024-10-16 paper.qmd
Updated the paper with new data structure

@@
 #| label: two-curvy-training
-training_data_two_curvy <- read_rds("data/two_curvy_diff_clust/two_curvy_diff_clust_data.rds")
+training_data_two_curvy <- read_rds("data/two_c_shaped_clust/two_c_shaped_clust_data.rds")
 ```
@@
 
-true_model_df <- read_rds("data/two_curvy_diff_clust/two_curvy_diff_clust_true_model.rds")
-wireframe_true_model <- read_rds("data/two_curvy_diff_clust/two_curvy_diff_clust_true_model_connections.rds")
+true_model_df <- read_rds("data/two_c_shaped_clust/two_c_shaped_clust_true_model.rds")
+wireframe_true_model <- read_rds("data/two_c_shaped_clust/two_c_shaped_clust_true_model_connections.rds")
 
@@
 #| label: umap-two-curvy
-umap_two_curvy <- read_rds(file = "data/two_curvy_diff_clust/two_curvy_diff_clust_umap_n-neigbors_15_min-dist_0.1.rds")
+umap_two_curvy <- read_rds(file = "data/two_c_shaped_clust/two_c_shaped_clust_umap_n-neigbors_15_min-dist_0.1.rds")
 
@@
 
-error_two_curvy_diff_clust_umap <- read_rds("data/two_curvy_diff_clust/error_two_curvy_diff_clust_umap.rds")
+error_two_c_shaped_clust_umap <- read_rds("data/two_c_shaped_clust/error_two_c_shaped_clust_umap.rds")
 
 ## Find the minimum MSE when have duplicate a1
-error_two_curvy_diff_clust_umap <- error_two_curvy_diff_clust_umap |>
+error_two_c_shaped_clust_umap <- error_two_c_shaped_clust_umap |>
   group_by(a1) |>
@@
 
-base_line_prop_dens <- error_two_curvy_diff_clust_umap |>
+base_line_prop_dens <- error_two_c_shaped_clust_umap |>
   filter(a1 == min(a1)) |>
@@
 
-error_two_curvy_diff_clust_umap <- error_two_curvy_diff_clust_umap |>
+error_two_c_shaped_clust_umap <- error_two_c_shaped_clust_umap |>
   mutate(prop_comp = prop_dens/base_line_prop_dens)
 
-mse_two_curvy_b <- ggplot(error_two_curvy_diff_clust_umap, 
+mse_two_curvy_b <- ggplot(error_two_c_shaped_clust_umap, 
                      aes(x = a1, 
@@
   geom_point(size = 1) +
-  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_diff_clust_umap$a1, 2)))[c(1, 4, 8, 10, 11, 12, 13, 14)]) +
+  scale_x_continuous(breaks = sort(unique(round(error_two_c_shaped_clust_umap$a1, 2)))[c(1, 4, 8, 10, 11, 12, 13, 14)]) +
   labs(x = expression(paste("binwidth (", a[1], ")")), y = "MSE") +
@@
 
-prop_dens_a1 <- ggplot(error_two_curvy_diff_clust_umap,
+prop_dens_a1 <- ggplot(error_two_c_shaped_clust_umap,
                      aes(x = a1,
@@
   # scale_y_continuous(breaks = sort(unique(error_two_curvy$b_non_empty))[-3]) +
-  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_diff_clust_umap$a1, 2)))[c(1, 4, 8, 10, 11, 12, 13, 14)]) +
+  scale_x_continuous(breaks = sort(unique(round(error_two_c_shaped_clust_umap$a1, 2)))[c(1, 4, 8, 10, 11, 12, 13, 14)]) +
   labs(x = expression(paste("binwidth (", a[1], ")")), y = paste("Relative proportion density")) +
@@
 
-error_two_curvy_diff_clust_umap <- error_two_curvy_diff_clust_umap |>
+error_two_c_shaped_clust_umap <- error_two_c_shaped_clust_umap |>
   mutate(prop_bins = b_non_empty/b)
 
-a1_m_two_curvy <- ggplot(error_two_curvy_diff_clust_umap,
+a1_m_two_curvy <- ggplot(error_two_c_shaped_clust_umap,
                      aes(x = a1,
@@
   # scale_y_continuous(breaks = sort(unique(error_two_curvy$b_non_empty))[-3]) +
-  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_diff_clust_umap$a1, 2)))[c(1, 4, 8, 10, 11, 12, 13, 14)]) +
+  scale_x_continuous(breaks = sort(unique(round(error_two_c_shaped_clust_umap$a1, 2)))[c(1, 4, 8, 10, 11, 12, 13, 14)]) +
   labs(x = expression(paste("binwidth (", a[1], ")")), y = expression(paste("proportion of non-empty bins ", bgroup("(", frac(m, b), ")")))) +
@@
 
-predict_umap_two_curvy <- read_rds(file = "data/two_curvy_diff_clust/two_curvy_diff_clust_umap_predict_true.rds")
+predict_umap_two_curvy <- read_rds(file = "data/two_c_shaped_clust/two_c_shaped_clust_umap_predict_true.rds")
 
@@
 
-error_rm_two_curvy <- read_rds("data/two_curvy_diff_clust/error_rm_lwd_diff_bin.rds") |>
+error_rm_two_curvy <- read_rds("data/two_c_shaped_clust/error_rm_lwd_diff_bin.rds") |>
   mutate(bin1  = as.factor(bin1))
@@
 #| label: read-two-curvy-clust-nldr
-# Read a variety of different NLDR representations of two_curvy_diff_clust
+# Read a variety of different NLDR representations of two_c_shaped_clust
 # and plot them on same aspect ratio
-tsne_two_curvy_diff_clust <- read_rds("data/two_curvy_diff_clust/two_curvy_diff_clust_tsne_perplexity_30.rds")
-
-nldr_two_curvy_diff_clust1 <- tsne_two_curvy_diff_clust |>
+tsne_two_c_shaped_clust <- read_rds("data/two_c_shaped_clust/two_c_shaped_clust_tsne_perplexity_30.rds")
+
+nldr_two_c_shaped_clust1 <- tsne_two_c_shaped_clust |>
   ggplot(aes(x = tSNE1,
@@
 
-umap_two_curvy_diff_clust <- read_rds("data/two_curvy_diff_clust/two_curvy_diff_clust_umap_n-neigbors_15_min-dist_0.1.rds")
-
-nldr_two_curvy_diff_clust2 <- umap_two_curvy_diff_clust |>
+umap_two_c_shaped_clust <- read_rds("data/two_c_shaped_clust/two_c_shaped_clust_umap_n-neigbors_15_min-dist_0.1.rds")
+
+nldr_two_c_shaped_clust2 <- umap_two_c_shaped_clust |>
   ggplot(aes(x = UMAP1,
@@
 
-phate_two_curvy_diff_clust <- read_rds("data/two_curvy_diff_clust/two_curvy_diff_clust_phate_knn_5.rds")
-
-nldr_two_curvy_diff_clust3 <- phate_two_curvy_diff_clust |>
+phate_two_c_shaped_clust <- read_rds("data/two_c_shaped_clust/two_c_shaped_clust_phate_knn_5.rds")
+
+nldr_two_c_shaped_clust3 <- phate_two_c_shaped_clust |>
   ggplot(aes(x = PHATE1,
@@
 
-trimap_two_curvy_diff_clust <- read_rds("data/two_curvy_diff_clust/two_curvy_diff_clust_trimap_n-inliers_12_n-outliers_4_n-random_3.rds")
-
-nldr_two_curvy_diff_clust4 <- trimap_two_curvy_diff_clust |>
+trimap_two_c_shaped_clust <- read_rds("data/two_c_shaped_clust/two_c_shaped_clust_trimap_n-inliers_12_n-outliers_4_n-random_3.rds")
+
+nldr_two_c_shaped_clust4 <- trimap_two_c_shaped_clust |>
   ggplot(aes(x = TriMAP1,
@@
 
-pacmap_two_curvy_diff_clust <- read_rds("data/two_curvy_diff_clust/two_curvy_diff_clust_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")
-
-nldr_two_curvy_diff_clust5 <- pacmap_two_curvy_diff_clust |>
+pacmap_two_c_shaped_clust <- read_rds("data/two_c_shaped_clust/two_c_shaped_clust_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")
+
+nldr_two_c_shaped_clust5 <- pacmap_two_c_shaped_clust |>
   ggplot(aes(x = PaCMAP1,
@@
 ```{r}
-#| label: combine-data-two_curvy_diff_clust
-
-error_two_curvy_diff_clust_umap <- read_rds("data/two_curvy_diff_clust/error_two_curvy_diff_clust_umap.rds")
-error_two_curvy_diff_clust_tsne <- read_rds("data/two_curvy_diff_clust/error_two_curvy_diff_clust_tsne.rds")
-error_two_curvy_diff_clust_phate <- read_rds("data/two_curvy_diff_clust/error_two_curvy_diff_clust_phate.rds")
-error_two_curvy_diff_clust_trimap <- read_rds("data/two_curvy_diff_clust/error_two_curvy_diff_clust_trimap.rds")
-error_two_curvy_diff_clust_pacmap <- read_rds("data/two_curvy_diff_clust/error_two_curvy_diff_clust_pacmap.rds")
-
-error_two_curvy_diff_clust <- bind_rows(error_two_curvy_diff_clust_umap, 
-                         error_two_curvy_diff_clust_tsne,
-                         error_two_curvy_diff_clust_phate,
-                         error_two_curvy_diff_clust_trimap,
-                         error_two_curvy_diff_clust_pacmap)
-
-error_two_curvy_diff_clust <- error_two_curvy_diff_clust |>
+#| label: combine-data-two_c_shaped_clust
+
+error_two_c_shaped_clust_umap <- read_rds("data/two_c_shaped_clust/error_two_c_shaped_clust_umap.rds")
+error_two_c_shaped_clust_tsne <- read_rds("data/two_c_shaped_clust/error_two_c_shaped_clust_tsne.rds")
+error_two_c_shaped_clust_phate <- read_rds("data/two_c_shaped_clust/error_two_c_shaped_clust_phate.rds")
+error_two_c_shaped_clust_trimap <- read_rds("data/two_c_shaped_clust/error_two_c_shaped_clust_trimap.rds")
+error_two_c_shaped_clust_pacmap <- read_rds("data/two_c_shaped_clust/error_two_c_shaped_clust_pacmap.rds")
+
+error_two_c_shaped_clust <- bind_rows(error_two_c_shaped_clust_umap, 
+                         error_two_c_shaped_clust_tsne,
+                         error_two_c_shaped_clust_phate,
+                         error_two_c_shaped_clust_trimap,
+                         error_two_c_shaped_clust_pacmap)
+
+error_two_c_shaped_clust <- error_two_c_shaped_clust |>
   mutate(a1 = round(a1, 2)) |>
@@
 ```{r}
-#| label: error-comp-two_curvy_diff_clust
-
-error_plot_two_curvy_diff_clust <- ggplot(error_two_curvy_diff_clust, 
+#| label: error-comp-two_c_shaped_clust
+
+error_plot_two_c_shaped_clust <- ggplot(error_two_c_shaped_clust, 
                            aes(x = a1, 
@@
   #            color = "black", linewidth=0.8, alpha = 0.5) +
-  scale_x_continuous(breaks = sort(unique(error_two_curvy_diff_clust$a1))[seq(1, length(unique(error_two_curvy_diff_clust$a1)), by = 5)]) +
+  scale_x_continuous(breaks = sort(unique(error_two_c_shaped_clust$a1))[seq(1, length(unique(error_two_c_shaped_clust$a1)), by = 5)]) +
   scale_color_manual(values=c('#e41a1c','#ff7f00','#4daf4a', "#a65628",'#636363')) +
@@
 Layout a is the best choice."
-#| label: fig-two_curvy_diff_clust-mse
+#| label: fig-two_c_shaped_clust-mse
 #| fig-pos: H
@@
 
-free(error_plot_two_curvy_diff_clust) + wrap_plots(nldr_two_curvy_diff_clust1, nldr_two_curvy_diff_clust2, nldr_two_curvy_diff_clust3, 
-                                    nldr_two_curvy_diff_clust4, nldr_two_curvy_diff_clust5, ncol = 2)
+free(error_plot_two_c_shaped_clust) + wrap_plots(nldr_two_c_shaped_clust1, nldr_two_c_shaped_clust2, nldr_two_c_shaped_clust3, 
+                                    nldr_two_c_shaped_clust4, nldr_two_c_shaped_clust5, ncol = 2)
 ```
@@
 two_curvy_tsne_scaled_obj <- gen_scaled_data(
-  data = tsne_two_curvy_diff_clust)
+  data = tsne_two_c_shaped_clust)
 
@@
 
-tsne_two_curvy_diff_clust_plt <- tsne_two_curvy_scaled |>
+tsne_two_c_shaped_clust_plt <- tsne_two_curvy_scaled |>
   ggplot(aes(x = tSNE1,
@@
 
-nldr_two_curvy_original2 + tsne_two_curvy_diff_clust_plt +
+nldr_two_curvy_original2 + tsne_two_c_shaped_clust_plt +
   trimesh_two_curvy_removed1_with_data + trimesh_two_curvy_removed1_tsne_with_data +

### 2024-10-16 paper.tex
Updated Figure 9
Updated the paper with new data structure

@@
 flattens out. Three binwidth choices were made: \(0.05\), \(0.09\), and
-\(0.12\) to investigate. As the binwidth increases, the proportion of
+\(0.13\) to investigate. As the binwidth increases, the proportion of
 non-empty bins also increases. There are two peaks at binwidths \(0.09\)
-and \(0.12\). The relative proportion density decreases and levels off.
+and \(0.13\). The relative proportion density decreases and levels off.
 Binwidths \(0.05\) is the favorable choices. Based on the MSE, \(0.09\)
@@
 
-\includegraphics[width=1\textwidth,height=0.6\textheight]{paper_files/figure-pdf/fig-two_curvy_diff_clust-mse-1.pdf}
-
-}
-
-\caption{\label{fig-two_curvy_diff_clust-mse}Assessing which of the 5
-NLDR layouts on the two C-shaped clusters data is the better
-representation using MSE for varying binwidth (\(a_1\)). Colour used for
-the lines and points in the left plot and in the scatterplots represents
-NLDR layout (a-e). Layout d is universally poor. Layouts a that show two
-close clusters are universally suboptimal. Layout a is the best choice.}
+\includegraphics[width=1\textwidth,height=0.6\textheight]{paper_files/figure-pdf/fig-two_c_shaped_clust-mse-1.pdf}
+
+}
+
+\caption{\label{fig-two_c_shaped_clust-mse}Assessing which of the 5 NLDR
+layouts on the two C-shaped clusters data is the better representation
+using MSE for varying binwidth (\(a_1\)). Colour used for the lines and
+points in the left plot and in the scatterplots represents NLDR layout
+(a-e). Layout d is universally poor. Layouts a that show two close
+clusters are universally suboptimal. Layout a is the best choice.}
 

### 2024-10-16 paper.qmd
Added PaCMAP model for two C-shaped data

@@
 ```{r}
-#| label: best-fit-umap-tsne
+#| label: best-fit-pacmap-two-curvy
+
+two_curvy_pacmap_scaled_obj <- gen_scaled_data(
+  data = pacmap_two_c_shaped_clust)
+
+pacmap_two_curvy_scaled <- two_curvy_pacmap_scaled_obj$scaled_nldr |>
+  mutate(ID = row_number())
+lim1 <- two_curvy_pacmap_scaled_obj$lim1
+lim2 <- two_curvy_pacmap_scaled_obj$lim2
+r2_pacmap <- diff(lim2)/diff(lim1)
+
+pacmap_two_c_shaped_clust_plt <- pacmap_two_curvy_scaled |>
+  ggplot(aes(x = PaCMAP1,
+             y = PaCMAP2))+
+  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
+  interior_annotation("b1", sc_ltr_pos) +
+  theme(aspect.ratio = 1)
+
+## hexagon binning to have regular hexagons
+hb_obj_pacmap_two_curvy2 <- hex_binning(
+  data = pacmap_two_curvy_scaled, 
+  bin1 = 8, 
+  r2 = r2_pacmap)
+
+a1_2_pacmap <- calc_bins_y(
+  bin1 = 8, 
+  r2 = r2_pacmap
+)$a1
+
+## Data set with all centroids
+all_centroids_df4 <- hb_obj_pacmap_two_curvy2$centroids
+
+## Generate all coordinates of hexagons
+hex_grid4 <- hb_obj_pacmap_two_curvy2$hex_poly
+
+## To obtain the standardise counts within hexbins
+counts_df4 <- hb_obj_pacmap_two_curvy2$std_cts
+df_bin_centroids4 <- extract_hexbin_centroids(
+  centroids_df = all_centroids_df4, 
+  counts_df = counts_df4) |>
+  filter(drop_empty == FALSE) |>
+  mutate(b1 = "b1 = 11")
+
+hex_grid_with_counts_two_curvy4 <- full_join(
+  hex_grid4, 
+  df_bin_centroids4 |> select(hexID, std_counts), 
+  by = c("hex_poly_id" = "hexID"))
+
+pacmap_data_two_curvy2_with_hb_id <- hb_obj_pacmap_two_curvy2$data_hb_id
+df_all_two_curvy4 <- dplyr::bind_cols(training_data_two_curvy, pacmap_data_two_curvy2_with_hb_id)
+df_bin_two_curvy4 <- avg_highd_data(data = df_all_two_curvy4, col_start = "x")
+
+df_b_two_curvy4 <- df_bin_two_curvy4 |>
+  dplyr::filter(hb_id %in% df_bin_centroids4$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b_two_curvy4 <- df_b_two_curvy4[match(df_bin_centroids4$hexID, df_b_two_curvy4$hb_id),] |>
+  dplyr::select(-hb_id)
+
+## Triangulate bin centroids
+tr1_object_two_curvy4 <- tri_bin_centroids(
+  df_bin_centroids4, x = "c_x", y = "c_y")
+tr_from_to_df_two_curvy4 <- gen_edges(
+  tri_object = tr1_object_two_curvy4)
+
+## Compute 2D distances
+distance_two_curvy4 <- cal_2d_dist(
+  tr_coord_df = tr_from_to_df_two_curvy4, 
+  start_x = "x_from", 
+  start_y = "y_from", 
+  end_x = "x_to", 
+  end_y = "y_to", 
+  select_vars = c("from", "to", "distance"))
+
+## To find the benchmark value
+benchmark_two_curvy4 <- find_lg_benchmark(
+  distance_edges = distance_two_curvy4, 
+  distance_col = "distance")
+
+benchmark_two_curvy4 <- 0.2
+
+## Benchmark 1
+
+distance_df_small_edges_two_curvy4 <- distance_two_curvy4 |>
+  filter(distance < benchmark_two_curvy4) |> ## 0.221
+  filter(!(row_number() %in% c(35, 89, 83, 25, 80, 56, 13, 59, 16)))
+
+tr_from_to_df_two_curvy4 <- inner_join(
+  tr_from_to_df_two_curvy4, distance_df_small_edges_two_curvy4,
+  by = c("from", "to"))
+
+trimesh_two_curvy_removed1_pacmap_with_data <- ggplot() +
+  geom_segment(data = tr_from_to_df_two_curvy4,
+               aes(
+                 x = x_from,
+                 y = y_from,
+                 xend = x_to,
+                 yend = y_to),
+               colour = "#000000") +
+  geom_point(data = pacmap_two_curvy_scaled,
+             aes(
+               x = PaCMAP1,
+               y = PaCMAP2
+             ),
+             color = "#a65628",
+             alpha = 0.5,
+             size = 0.5) +
+  coord_equal() +
+  interior_annotation("b2", sc_ltr_pos)
+
+```
+
+```{r}
+#| label: best-fit-pacmap-two-curvy-proj
+
+# Apply the scaling
+df_model_data_two_curvy <- bind_rows(data_two_curvy, true_model_two_curvy, df_b_two_curvy4)
+
+scaled_two_curvy <- scale_data_manual(df_model_data_two_curvy, "type") |>
+  as_tibble()
+
+scaled_two_curvy_data <- scaled_two_curvy |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_two_curvy_data_true_model <- scaled_two_curvy |>
+  filter(type == "true model") |>
+  select(-type)
+
+scaled_two_curvy_data_model <- scaled_two_curvy |>
+  filter(type == "model") |>
+  select(-type)
+
+## First projection
+projection <- cbind(
+  c(-0.08707,-0.06493,0.07017,-0.07349,-0.02955,-0.05286,-0.02765),
+  c(0.03398,0.03735,0.09272,0.01871,-0.09224,0.00451,0.08079))
+
+projection_scaled <- projection * 1
+
+projected <- as.matrix(scaled_two_curvy_data) %*% projection_scaled
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+## For fitted model
+projected_model <- as.matrix(scaled_two_curvy_data_model) %*% projection_scaled
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges_two_curvy4 |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+## For true model
+
+projected_true_model <- as.matrix(scaled_two_curvy_data_true_model) %*% projection_scaled
+
+projected_true_model_df <- projected_true_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+true_model_df_proj <- dplyr::left_join(
+  wireframe_true_model, 
+  projected_true_model_df, 
+  by = c("from" = "ID"))
+
+names(true_model_df_proj)[3:NCOL(true_model_df_proj)] <- paste0(names(projected_true_model_df)[-NCOL(projected_true_model_df)], "_from")
+
+true_model_df_proj <- dplyr::left_join(true_model_df_proj, projected_true_model_df, by = c("to" = "ID"))
+names(true_model_df_proj)[(2 + NCOL(projected_true_model_df)):NCOL(true_model_df_proj)] <- paste0(names(projected_true_model_df)[-NCOL(projected_true_model_df)], "_to")
+
+axes_obj <- gen_axes(  
+  proj = projection * 5,
+  limits = 0.15,
+  axis_pos_x = -0.1,
+  axis_pos_y = -0.1,
+  axis_labels = names(scaled_two_curvy_data),
+  threshold = 0.01)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
+
+two_curvy_proj_pacmap_model <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#000000") + #31a354
+  geom_point(
+    size = 0.5,
+    alpha = 0.5,
+    color = "#a65628") +
+  geom_segment(
+    data=axes,
+    aes(x=x1, y=y1, xend=x2, yend=y2),
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
+    colour="grey50",
+    size = 4) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.126, 0.125)) + 
+  ylim(c(-0.126, 0.125)) +
+  interior_annotation("b3", sc_ltr_pos) +
+  theme(
+    legend.position = "none"
+  )
+
+two_curvy_proj_pacmap_all_model <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = true_model_df_proj, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#2b8cbe") +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#000000") +
+  geom_point(
+    size = 0.5,
+    alpha = 0.5,
+    color = "#a65628") +
+  geom_segment(
+    data=axes,
+    aes(x=x1, y=y1, xend=x2, yend=y2),
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
+    colour="grey50",
+    size = 4) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.126, 0.125)) + 
+  ylim(c(-0.126, 0.125)) +
+  interior_annotation("b4", sc_ltr_pos) +
+  theme(
+    legend.position = "none"
+  )
+```
+
+```{r}
+#| label: best-fit-umap-tsne-pacamp
 #| fig-cap: "Best fit for two C-shaped clusters data using UMAP (a1-a4) ($a_1$ = 0.09) and tSNE (b1-b4) ($a_1$ = 0.16) with the same number of non-empty bins ($m$ = 39).
@@
 Also, both NLDR methods are unable to accurately capture the width of the curvilinear clusters."
-#| fig-width: 10
+#| fig-width: 15
 #| fig-height: 20
@@
 nldr_two_curvy_original2 + tsne_two_c_shaped_clust_plt +
+  pacmap_two_c_shaped_clust_plt +
   trimesh_two_curvy_removed1_with_data + trimesh_two_curvy_removed1_tsne_with_data +
+trimesh_two_curvy_removed1_pacmap_with_data +
   two_curvy_proj_umap_model2 +
   two_curvy_proj_tsne_model +
+  two_curvy_proj_pacmap_model +
   two_curvy_proj_umap_all_model +
   two_curvy_proj_tsne_all_model +
-  plot_layout(ncol = 2)
-```
-
-<!-- @fig-two_curvy-sc-best shows one of the best fits for S-curve data.
-TriMAP with $15$ nearest neighbors for forming the nearest neighbor triplet (`n_inliers`), $3$ outliers for forming the nearest neighbor triplets (`n_outliers`), $3$ random triplets per point (`n_random`) is used as the NLDR method.
-The model is generated with $103$ bins, which have a width of $0.154$ and a height of $0.133$.
-Edges with lengths greater than $0.266$ are removed.
--->
-
-<!-- The appropriate model for the S-curve using UMAP (n_neighbors: $15$) was created with $420$ bins.
-It was determined to be the best model after removing edges with a length greater than $0.134$.
-The decision was made considering that the model accurately captures the geometry of the S-curve.
-But better in some places only (@fig-scurve_sc_best).
--->
-
-<!-- In contrast, a poor model for the S-curve with PHATE was created using only $190$ bins and removing edges with a length greater than $0.3$.
-This model is inadequate as it has a higher MSE compared to the appropriate model.
-Visually, the model squeezed to the middle of the geometry of the S-curve (@fig-scurve-sc-bad).
--->
-
-<!-- XXX BEST FIT FOR S-CURVE MIGHT BE A DIFFERENT NLDR.  -->
-<!-- SHOW THE SEVERAL SUB-OPTIMAL FITS. -->
-
-<!--langevitour with best pD model for S-curve--> 
-<!-- ::: {#fig-scurve-sc-best layout-ncol="3" fig-pos="H"} -->
-<!-- ![](figures/scurve/sc_best_1.png) -->
-
-<!-- ![](figures/scurve/sc_best_2.png) -->
-
-<!-- ![](figures/scurve/sc_best_3.png) -->
-
-<!-- Screen shots of the **langevitour** of the S-curve, shows the model-in-data space, a video of the tour animation is available at (<https://youtu.be/4RFNENX_fXk>). MSE with $0.035$. -->
-<!-- ::: -->
-
-<!--langevitour with bad pD model for S-curve--> 
-<!-- ::: {#fig-scurve-sc-bad layout-ncol="3" fig-pos="H"} -->
-<!-- ![](figures/scurve/sc_bad_1.png) -->
-
-<!-- ![](figures/scurve/sc_bad_2.png) -->
-
-<!-- ![](figures/scurve/sc_bad_3.png) -->
-
-<!-- Screen shots of the **langevitour** of the S-curve, shows the model-in-data space, a video of the tour animation is available at (<https://youtu.be/DHPvhkZ0lUM>). MSE with $0.090$. -->
-<!-- ::: -->
-
-<!------>
-<!--add more details regarding to the how the change happening with different number of bins-->
-
-<!--best model-->
-<!--need to mention why it is the best-->
-
-<!--bad model-->
-<!--need to mention why it is the bad-->
-<!--Not represented the whole data structure because there are some discontinuity in the S-curve-->
-<!-- Why? bad
-
-<!-- add MSE values for the screenshot figure names-->
+  two_curvy_proj_pacmap_all_model +
+  plot_layout(ncol = 3)
+```
 

### 2024-10-17 paper.qmd
Updated Figure 10

@@
 #| fig-cap: "Distribution of edge lengths and wireframes in \\gD{} and $7\\text{-}D$ with Delaunay triangulation, and two choices of long edge removal: (b) benchmark = 4.5$a_1$ and (c) benchmark = 2$a_1$, where $a_1$ = 0.09."
-#| fig-height: 4
-#| fig-width: 20
-
-free(distance_points) + plot_annotation(tag_levels = 'A') +
- (trimesh_two_curvy2 + trimesh_two_curvy_removed2 + trimesh_two_curvy_removed1 + two_curvy_proj_umap_model_delaunay + two_curvy_proj_umap_model_benchmark2 + two_curvy_proj_umap_model) +
-  plot_layout(widths = c(0.5, 1, 1, 1)) +
-  plot_annotation(tag_levels = list(c("(a)", "(b)"))) 
-```
-
-<!-- (@fig-NLDR-two-curvy (d)), it is necessary to remove edges that connect distant bin centroids in the triangular mesh.
-These edges only exist in the \gD{} model and do not extend into \pD{}, so their removal does not impact the model in \pD{}.
-Although there are no specific criteria for determining the benchmark value to remove long edges, the following steps provide an approach to identifying a suitable threshold:
-
-1. Plot the distribution of the \gD{} Euclidean distances (@fig-param-two_curvy (d)).
-2. Identify the first largest difference between consecutive distance values.
-3. Take the distance value corresponding to this difference as the benchmark value.-->
-
-<!--distribution of distance along with the default benchmark-->
+#| fig-height: 15
+#| fig-width: 15
+
+free(distance_points) /
+  wrap_plots(trimesh_two_curvy2, trimesh_two_curvy_removed2, trimesh_two_curvy_removed1, two_curvy_proj_umap_model_delaunay, two_curvy_proj_umap_model_benchmark2, two_curvy_proj_umap_model, ncol = 3)
+
+# free(distance_points) + plot_annotation(tag_levels = 'A') +
+#  (trimesh_two_curvy2 + trimesh_two_curvy_removed2 + trimesh_two_curvy_removed1 + two_curvy_proj_umap_model_delaunay + two_curvy_proj_umap_model_benchmark2 + two_curvy_proj_umap_model) +
+#   plot_layout(widths = c(0.5, 1, 1, 1)) +
+#   plot_annotation(tag_levels = list(c("(a)", "(b)"))) 
+```
 

### 2024-10-17 paper.qmd
Added labels as well for Figure 10

@@
 free(distance_points) /
-  wrap_plots(trimesh_two_curvy2, trimesh_two_curvy_removed2, trimesh_two_curvy_removed1, two_curvy_proj_umap_model_delaunay, two_curvy_proj_umap_model_benchmark2, two_curvy_proj_umap_model, ncol = 3)
+  wrap_plots(trimesh_two_curvy2, trimesh_two_curvy_removed2, trimesh_two_curvy_removed1, two_curvy_proj_umap_model_delaunay, two_curvy_proj_umap_model_benchmark2, two_curvy_proj_umap_model, ncol = 3) +
+  plot_annotation(tag_levels = list(c("(a)", "(b)")))
 

### 2024-10-17 paper.qmd
Updated Figure 12

@@
   geom_point(alpha=0.1, size=1, colour='#e41a1c') +
-  interior_annotation("b1", sc_ltr_pos) +
+  interior_annotation("c1", sc_ltr_pos) +
   theme(aspect.ratio = 1)
@@
              ),
-             color = "#a65628",
+             color = "#e41a1c",
              alpha = 0.5,
@@
   coord_equal() +
-  interior_annotation("b2", sc_ltr_pos)
+  interior_annotation("c2", sc_ltr_pos)
 
@@
     alpha = 0.5,
-    color = "#a65628") +
+    color = "#e41a1c") +
   geom_segment(
@@
   ylim(c(-0.126, 0.125)) +
-  interior_annotation("b3", sc_ltr_pos) +
+  interior_annotation("c3", sc_ltr_pos) +
   theme(
@@
   ylim(c(-0.126, 0.125)) +
-  interior_annotation("b4", sc_ltr_pos) +
+  interior_annotation("c4", sc_ltr_pos) +
   theme(
@@
   two_curvy_proj_pacmap_all_model +
-  plot_layout(ncol = 3)
+  plot_layout(ncol = 3) +
+  plot_annotation(tag_levels = list(c("(a)", "(b)", "(c)")))
 ```

### 2024-10-17 paper.qmd
Updated Figure 2 projection

@@
 axes_obj <- gen_axes(  
-  proj = projection * 5,
-  limits = 0.12,
+  proj = projection * 6,
+  limits = 0.13,
   axis_pos_x = -0.1,
@@
   axis_labels = names(scaled_two_curvy_data),
-  threshold = 0.01)
+  threshold = 0.0113)
 
@@
   coord_fixed() +
-  xlim(c(-0.125, 0.075)) + 
-  ylim(c(-0.125, 0.075)) +
+  xlim(c(-0.125, 0.13)) + 
+  ylim(c(-0.125, 0.13)) +
   interior_annotation("b", c(0.08, 0.93), cex = 2) +

### 2024-10-17 paper.tex
Added PaCMAP model for two C-shaped data
Updated Figure 2 title

@@
 (blue lines) in \(2\text{-}D\) projections from \(7\text{-}D\), for the
-two C-shaped clusters data (grey points). The data is spread along the
-non-linear shaped clusters, and does not vary much from the curves. The
+two C-shaped clusters data (grey points). In one C-shaped cluster, the
+data is spread throughout the cluster, while in the other C-shaped
+cluster, the data is concentrated more in one corner of the cluster. The
 \textbf{langevitour} software is used to view the data with a tour, and
@@
 
-{\centering \includegraphics[width=1\textwidth,height=0.9\textheight]{paper_files/figure-pdf/best-fit-umap-tsne-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=0.9\textheight]{paper_files/figure-pdf/best-fit-umap-tsne-pacamp-1.pdf}
 

### 2024-10-17 paper.qmd
Updated Figure 2 title
Updated the distances between clusters

@@
 ## Compute hexbin parameters
-num_bins_x_two_curvy <- 10
+num_bins_x_two_curvy <- 15
 
@@
 #| fig-cap: "Two views of the true model (blue lines) in $2\\text{-}D$ projections from $7\\text{-}D$, for the two C-shaped clusters data (grey points).
-The data is spread along the non-linear shaped clusters, and does not vary much from the curves.
+In one C-shaped cluster, the data is spread throughout the cluster, while in the other C-shaped cluster, the data is concentrated more in one corner of the cluster.
 The **langevitour** software is used to view the data with a tour, and the full video is available at <>."

### 2024-11-18 paper.qmd
Regenerate MSE plots
Changed the script to view the true model along with the data

@@
 #| label: two-curvy-training
-training_data_two_curvy <- read_rds("data/two_c_shaped_clust/two_c_shaped_clust_data.rds")
+training_data_two_curvy <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_data.rds")
 ```
@@
 
-true_model_df <- read_rds("data/two_c_shaped_clust/two_c_shaped_clust_true_model.rds")
-wireframe_true_model <- read_rds("data/two_c_shaped_clust/two_c_shaped_clust_true_model_connections.rds")
+true_model_df <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_true_model.rds")
+wireframe_true_model <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_true_model_connections.rds")
 
@@
 #| label: umap-two-curvy
-umap_two_curvy <- read_rds(file = "data/two_c_shaped_clust/two_c_shaped_clust_umap_n-neigbors_15_min-dist_0.1.rds")
+umap_two_curvy <- read_rds(file = "data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_umap_n-neigbors_15_min-dist_0.1.rds")
 
@@
 ## Compute hexbin parameters
-num_bins_x_two_curvy <- 15
+num_bins_x_two_curvy <- 12
 
@@
 
-error_two_c_shaped_clust_umap <- read_rds("data/two_c_shaped_clust/error_two_c_shaped_clust_umap.rds")
+error_two_non_linear_diff_shaped_close_clusters_umap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_umap.rds")
 
 ## Find the minimum MSE when have duplicate a1
-error_two_c_shaped_clust_umap <- error_two_c_shaped_clust_umap |>
+error_two_non_linear_diff_shaped_close_clusters_umap <- error_two_non_linear_diff_shaped_close_clusters_umap |>
   group_by(a1) |>
@@
 
-base_line_prop_dens <- error_two_c_shaped_clust_umap |>
+base_line_prop_dens <- error_two_non_linear_diff_shaped_close_clusters_umap |>
   filter(a1 == min(a1)) |>
@@
 
-error_two_c_shaped_clust_umap <- error_two_c_shaped_clust_umap |>
+error_two_non_linear_diff_shaped_close_clusters_umap <- error_two_non_linear_diff_shaped_close_clusters_umap |>
   mutate(prop_comp = prop_dens/base_line_prop_dens)
 
-mse_two_curvy_b <- ggplot(error_two_c_shaped_clust_umap, 
+mse_two_curvy_b <- ggplot(error_two_non_linear_diff_shaped_close_clusters_umap, 
                      aes(x = a1, 
@@
   geom_point(size = 1) +
-  scale_x_continuous(breaks = sort(unique(round(error_two_c_shaped_clust_umap$a1, 2)))[c(1, 4, 8, 10, 11, 12, 13, 14)]) +
+  scale_x_continuous(breaks = sort(unique(round(error_two_non_linear_diff_shaped_close_clusters_umap$a1, 2)))[c(1, 4, 8, 10, 11, 12, 13, 14)]) +
   labs(x = expression(paste("binwidth (", a[1], ")")), y = "MSE") +
@@
 
-prop_dens_a1 <- ggplot(error_two_c_shaped_clust_umap,
+prop_dens_a1 <- ggplot(error_two_non_linear_diff_shaped_close_clusters_umap,
                      aes(x = a1,
@@
   # scale_y_continuous(breaks = sort(unique(error_two_curvy$b_non_empty))[-3]) +
-  scale_x_continuous(breaks = sort(unique(round(error_two_c_shaped_clust_umap$a1, 2)))[c(1, 4, 8, 10, 11, 12, 13, 14)]) +
+  scale_x_continuous(breaks = sort(unique(round(error_two_non_linear_diff_shaped_close_clusters_umap$a1, 2)))[c(1, 4, 8, 10, 11, 12, 13, 14)]) +
   labs(x = expression(paste("binwidth (", a[1], ")")), y = paste("Relative proportion density")) +
@@
 
-error_two_c_shaped_clust_umap <- error_two_c_shaped_clust_umap |>
+error_two_non_linear_diff_shaped_close_clusters_umap <- error_two_non_linear_diff_shaped_close_clusters_umap |>
   mutate(prop_bins = b_non_empty/b)
 
-a1_m_two_curvy <- ggplot(error_two_c_shaped_clust_umap,
+a1_m_two_curvy <- ggplot(error_two_non_linear_diff_shaped_close_clusters_umap,
                      aes(x = a1,
@@
   # scale_y_continuous(breaks = sort(unique(error_two_curvy$b_non_empty))[-3]) +
-  scale_x_continuous(breaks = sort(unique(round(error_two_c_shaped_clust_umap$a1, 2)))[c(1, 4, 8, 10, 11, 12, 13, 14)]) +
+  scale_x_continuous(breaks = sort(unique(round(error_two_non_linear_diff_shaped_close_clusters_umap$a1, 2)))[c(1, 4, 8, 10, 11, 12, 13, 14)]) +
   labs(x = expression(paste("binwidth (", a[1], ")")), y = expression(paste("proportion of non-empty bins ", bgroup("(", frac(m, b), ")")))) +
@@
 
-predict_umap_two_curvy <- read_rds(file = "data/two_c_shaped_clust/two_c_shaped_clust_umap_predict_true.rds")
+predict_umap_two_curvy <- read_rds(file = "data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_umap_predict_true.rds")
 
@@
 
-error_rm_two_curvy <- read_rds("data/two_c_shaped_clust/error_rm_lwd_diff_bin.rds") |>
+error_rm_two_curvy <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_rm_lwd_diff_bin.rds") |>
   mutate(bin1  = as.factor(bin1))
@@
 #| label: read-two-curvy-clust-nldr
-# Read a variety of different NLDR representations of two_c_shaped_clust
+# Read a variety of different NLDR representations of two_non_linear_diff_shaped_close_clusters
 # and plot them on same aspect ratio
-tsne_two_c_shaped_clust <- read_rds("data/two_c_shaped_clust/two_c_shaped_clust_tsne_perplexity_30.rds")
-
-nldr_two_c_shaped_clust1 <- tsne_two_c_shaped_clust |>
+tsne_two_non_linear_diff_shaped_close_clusters <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_30.rds")
+
+nldr_two_non_linear_diff_shaped_close_clusters1 <- tsne_two_non_linear_diff_shaped_close_clusters |>
   ggplot(aes(x = tSNE1,
@@
 
-umap_two_c_shaped_clust <- read_rds("data/two_c_shaped_clust/two_c_shaped_clust_umap_n-neigbors_15_min-dist_0.1.rds")
-
-nldr_two_c_shaped_clust2 <- umap_two_c_shaped_clust |>
+umap_two_non_linear_diff_shaped_close_clusters <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_umap_n-neigbors_15_min-dist_0.1.rds")
+
+nldr_two_non_linear_diff_shaped_close_clusters2 <- umap_two_non_linear_diff_shaped_close_clusters |>
   ggplot(aes(x = UMAP1,
@@
 
-phate_two_c_shaped_clust <- read_rds("data/two_c_shaped_clust/two_c_shaped_clust_phate_knn_5.rds")
-
-nldr_two_c_shaped_clust3 <- phate_two_c_shaped_clust |>
+phate_two_non_linear_diff_shaped_close_clusters <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_phate_knn_5.rds")
+
+nldr_two_non_linear_diff_shaped_close_clusters3 <- phate_two_non_linear_diff_shaped_close_clusters |>
   ggplot(aes(x = PHATE1,
@@
 
-trimap_two_c_shaped_clust <- read_rds("data/two_c_shaped_clust/two_c_shaped_clust_trimap_n-inliers_12_n-outliers_4_n-random_3.rds")
-
-nldr_two_c_shaped_clust4 <- trimap_two_c_shaped_clust |>
+trimap_two_non_linear_diff_shaped_close_clusters <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_trimap_n-inliers_12_n-outliers_4_n-random_3.rds")
+
+nldr_two_non_linear_diff_shaped_close_clusters4 <- trimap_two_non_linear_diff_shaped_close_clusters |>
   ggplot(aes(x = TriMAP1,
@@
 
-pacmap_two_c_shaped_clust <- read_rds("data/two_c_shaped_clust/two_c_shaped_clust_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")
-
-nldr_two_c_shaped_clust5 <- pacmap_two_c_shaped_clust |>
+pacmap_two_non_linear_diff_shaped_close_clusters <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")
+
+nldr_two_non_linear_diff_shaped_close_clusters5 <- pacmap_two_non_linear_diff_shaped_close_clusters |>
   ggplot(aes(x = PaCMAP1,
@@
 ```{r}
-#| label: combine-data-two_c_shaped_clust
-
-error_two_c_shaped_clust_umap <- read_rds("data/two_c_shaped_clust/error_two_c_shaped_clust_umap.rds")
-error_two_c_shaped_clust_tsne <- read_rds("data/two_c_shaped_clust/error_two_c_shaped_clust_tsne.rds")
-error_two_c_shaped_clust_phate <- read_rds("data/two_c_shaped_clust/error_two_c_shaped_clust_phate.rds")
-error_two_c_shaped_clust_trimap <- read_rds("data/two_c_shaped_clust/error_two_c_shaped_clust_trimap.rds")
-error_two_c_shaped_clust_pacmap <- read_rds("data/two_c_shaped_clust/error_two_c_shaped_clust_pacmap.rds")
-
-error_two_c_shaped_clust <- bind_rows(error_two_c_shaped_clust_umap, 
-                         error_two_c_shaped_clust_tsne,
-                         error_two_c_shaped_clust_phate,
-                         error_two_c_shaped_clust_trimap,
-                         error_two_c_shaped_clust_pacmap)
-
-error_two_c_shaped_clust <- error_two_c_shaped_clust |>
+#| label: combine-data-two_non_linear_diff_shaped_close_clusters
+
+error_two_non_linear_diff_shaped_close_clusters_umap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_umap.rds")
+error_two_non_linear_diff_shaped_close_clusters_tsne <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_tsne.rds")
+error_two_non_linear_diff_shaped_close_clusters_phate <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_phate.rds")
+error_two_non_linear_diff_shaped_close_clusters_trimap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_trimap.rds")
+error_two_non_linear_diff_shaped_close_clusters_pacmap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_pacmap.rds")
+
+error_two_non_linear_diff_shaped_close_clusters <- bind_rows(error_two_non_linear_diff_shaped_close_clusters_umap, 
+                         error_two_non_linear_diff_shaped_close_clusters_tsne,
+                         error_two_non_linear_diff_shaped_close_clusters_phate,
+                         error_two_non_linear_diff_shaped_close_clusters_trimap,
+                         error_two_non_linear_diff_shaped_close_clusters_pacmap)
+
+error_two_non_linear_diff_shaped_close_clusters <- error_two_non_linear_diff_shaped_close_clusters |>
   mutate(a1 = round(a1, 2)) |>
@@
 ```{r}
-#| label: error-comp-two_c_shaped_clust
-
-error_plot_two_c_shaped_clust <- ggplot(error_two_c_shaped_clust, 
+#| label: error-comp-two_non_linear_diff_shaped_close_clusters
+
+error_plot_two_non_linear_diff_shaped_close_clusters <- ggplot(error_two_non_linear_diff_shaped_close_clusters, 
                            aes(x = a1, 
@@
   #            color = "black", linewidth=0.8, alpha = 0.5) +
-  scale_x_continuous(breaks = sort(unique(error_two_c_shaped_clust$a1))[seq(1, length(unique(error_two_c_shaped_clust$a1)), by = 5)]) +
+  scale_x_continuous(breaks = sort(unique(error_two_non_linear_diff_shaped_close_clusters$a1))[seq(1, length(unique(error_two_non_linear_diff_shaped_close_clusters$a1)), by = 5)]) +
   scale_color_manual(values=c('#e41a1c','#ff7f00','#4daf4a', "#a65628",'#636363')) +
@@
 Layout a is the best choice."
-#| label: fig-two_c_shaped_clust-mse
+#| label: fig-two_non_linear_diff_shaped_close_clusters-mse
 #| fig-pos: H
@@
 
-free(error_plot_two_c_shaped_clust) + wrap_plots(nldr_two_c_shaped_clust1, nldr_two_c_shaped_clust2, nldr_two_c_shaped_clust3, 
-                                    nldr_two_c_shaped_clust4, nldr_two_c_shaped_clust5, ncol = 2)
+free(error_plot_two_non_linear_diff_shaped_close_clusters) + wrap_plots(nldr_two_non_linear_diff_shaped_close_clusters1, nldr_two_non_linear_diff_shaped_close_clusters2, nldr_two_non_linear_diff_shaped_close_clusters3,
+                                    nldr_two_non_linear_diff_shaped_close_clusters4, nldr_two_non_linear_diff_shaped_close_clusters5, ncol = 2)
 ```
@@
 two_curvy_tsne_scaled_obj <- gen_scaled_data(
-  data = tsne_two_c_shaped_clust)
+  data = tsne_two_non_linear_diff_shaped_close_clusters)
 
@@
 
-tsne_two_c_shaped_clust_plt <- tsne_two_curvy_scaled |>
+tsne_two_non_linear_diff_shaped_close_clusters_plt <- tsne_two_curvy_scaled |>
   ggplot(aes(x = tSNE1,
@@
 two_curvy_pacmap_scaled_obj <- gen_scaled_data(
-  data = pacmap_two_c_shaped_clust)
+  data = pacmap_two_non_linear_diff_shaped_close_clusters)
 
@@
 
-pacmap_two_c_shaped_clust_plt <- pacmap_two_curvy_scaled |>
+pacmap_two_non_linear_diff_shaped_close_clusters_plt <- pacmap_two_curvy_scaled |>
   ggplot(aes(x = PaCMAP1,
@@
 
-nldr_two_curvy_original2 + tsne_two_c_shaped_clust_plt +
-  pacmap_two_c_shaped_clust_plt +
+nldr_two_curvy_original2 + tsne_two_non_linear_diff_shaped_close_clusters_plt +
+  pacmap_two_non_linear_diff_shaped_close_clusters_plt +
   trimesh_two_curvy_removed1_with_data + trimesh_two_curvy_removed1_tsne_with_data +

### 2024-11-18 paper.tex
Changed the script to view the true model along with the data

@@
 
-\includegraphics[width=1\textwidth,height=0.6\textheight]{paper_files/figure-pdf/fig-two_c_shaped_clust-mse-1.pdf}
-
-}
-
-\caption{\label{fig-two_c_shaped_clust-mse}Assessing which of the 5 NLDR
-layouts on the two C-shaped clusters data is the better representation
-using MSE for varying binwidth (\(a_1\)). Colour used for the lines and
-points in the left plot and in the scatterplots represents NLDR layout
-(a-e). Layout d is universally poor. Layouts a that show two close
-clusters are universally suboptimal. Layout a is the best choice.}
+\includegraphics[width=1\textwidth,height=0.6\textheight]{paper_files/figure-pdf/fig-two_non_linear_diff_shaped_close_clusters-mse-1.pdf}
+
+}
+
+\caption{\label{fig-two_non_linear_diff_shaped_close_clusters-mse}Assessing
+which of the 5 NLDR layouts on the two C-shaped clusters data is the
+better representation using MSE for varying binwidth (\(a_1\)). Colour
+used for the lines and points in the left plot and in the scatterplots
+represents NLDR layout (a-e). Layout d is universally poor. Layouts a
+that show two close clusters are universally suboptimal. Layout a is the
+best choice.}
 

### 2024-11-18 paper.qmd
Updated Figure 5

@@
 
+nldr_two_curvy2 <- umap_two_curvy_scaled |> 
+  ggplot(aes(x = UMAP1, y = UMAP2)) + 
+  geom_point(alpha=0.5, colour="#636363", size = 0.5) +
+  interior_annotation("a", sc_ltr_pos)
+
 nldr_two_curvy_original <- umap_two_curvy_scaled |> 
@@
              size = 0.5) +
-  coord_equal() +
+  #coord_equal() +
   # theme(
@@
 
-nldr_two_curvy_original + 
+nldr_two_curvy2 + 
   trimesh_two_curvy_removed1_umap_with_data +

### 2024-11-19 paper.qmd
Added initial table for scDEED
Added 3 layouts for three UMAP hyperparameters

@@
 
+#### scDEED
+
+```{r}
+result_umap <- read_rds(here::here('data/pbmc3k/pbmc_scdeed_umap_results.rds'))
+
+result_umap |>
+  kableExtra::kable(format = "latex",
+                    booktabs = TRUE,
+                    longtable = TRUE,
+                    label = "scdeedumap") |>
+  kableExtra::kable_styling(latex_options = "scale_down")
+```
+
+```{r}
+#| label: read-pbmc-umap-layouts
+# Read a variety of different NLDR representations of PBMC
+# and plot them on same aspect ratio
+umap_pbmc1 <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_30_min_dist_0.3.rds")
+umap_pbmc1 <- as_tibble(umap_pbmc1)
+names(umap_pbmc1) <- c("UMAP1", "UMAP2")
+
+nldr1c <- umap_pbmc1 |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2)) +
+  geom_point(alpha=0.1, size=1, colour='#a65628') +
+  interior_annotation("a")
+
+umap_pbmc2 <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_5_min_dist_0.01.rds")
+umap_pbmc2 <- as_tibble(umap_pbmc2)
+names(umap_pbmc2) <- c("UMAP1", "UMAP2")
+
+nldr2c <- umap_pbmc2 |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2))+
+  geom_point(alpha=0.1, size=1, colour='#999999') +
+  interior_annotation("b")
+
+umap_pbmc3 <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_15_min_dist_0.99.rds")
+umap_pbmc3 <- as_tibble(umap_pbmc3)
+names(umap_pbmc3) <- c("UMAP1", "UMAP2")
+
+nldr3c <- umap_pbmc3 |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2))+
+  geom_point(alpha=0.1, size=1, colour='#f781bf') +
+  interior_annotation("c")
+```
+
+```{r}
+#| label: fig-pbmc-mse-umap
+#| fig-pos: H
+#| out-height: 100%
+
+wrap_plots(nldr1c, nldr2c, nldr3c, ncol = 3)
+```
+
 ### Hand-written digits

### 2024-11-20 paper.qmd
Added results for different UMAP parameter choices

@@
 
+result_umap <- result_umap |>
+  mutate(plot = c("b", "i", "h", "g", "f", "a", "e", "c", "d")) |>
+  arrange(number_dubious_cells)
+
 result_umap |>
@@
              y = UMAP2)) +
-  geom_point(alpha=0.1, size=1, colour='#a65628') +
+  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
   interior_annotation("a")
@@
              y = UMAP2))+
-  geom_point(alpha=0.1, size=1, colour='#999999') +
+  geom_point(alpha=0.1, size=1, colour='#a65628') +
   interior_annotation("b")
@@
              y = UMAP2))+
+  geom_point(alpha=0.1, size=1, colour='#4daf4a') +
+  interior_annotation("c")
+
+umap_pbmc4 <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_30_min_dist_0.99.rds")
+umap_pbmc4 <- as_tibble(umap_pbmc4)
+names(umap_pbmc4) <- c("UMAP1", "UMAP2")
+
+nldr4c <- umap_pbmc4 |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2))+
+  geom_point(alpha=0.1, size=1, colour='#8dd3c7') +
+  interior_annotation("d")
+
+umap_pbmc5 <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_5_min_dist_0.99.rds")
+umap_pbmc5 <- as_tibble(umap_pbmc5)
+names(umap_pbmc5) <- c("UMAP1", "UMAP2")
+
+nldr5c <- umap_pbmc5 |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2))+
+  geom_point(alpha=0.1, size=1, colour='#999999') +
+  interior_annotation("e")
+
+umap_pbmc6 <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_15_min_dist_0.3.rds")
+umap_pbmc6 <- as_tibble(umap_pbmc6)
+names(umap_pbmc6) <- c("UMAP1", "UMAP2")
+
+nldr6c <- umap_pbmc6 |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2))+
+  geom_point(alpha=0.1, size=1, colour='#377eb8') +
+  interior_annotation("f")
+
+umap_pbmc7 <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_5_min_dist_0.3.rds")
+umap_pbmc7 <- as_tibble(umap_pbmc7)
+names(umap_pbmc7) <- c("UMAP1", "UMAP2")
+
+nldr7c <- umap_pbmc7 |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2))+
   geom_point(alpha=0.1, size=1, colour='#f781bf') +
-  interior_annotation("c")
+  interior_annotation("g")
+
+umap_pbmc8 <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_30_min_dist_0.01.rds")
+umap_pbmc8 <- as_tibble(umap_pbmc8)
+names(umap_pbmc8) <- c("UMAP1", "UMAP2")
+
+nldr8c <- umap_pbmc8 |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2))+
+  geom_point(alpha=0.1, size=1, colour='#984ea3') +
+  interior_annotation("h")
+
+umap_pbmc9 <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_15_min_dist_0.01.rds")
+umap_pbmc9 <- as_tibble(umap_pbmc9)
+names(umap_pbmc9) <- c("UMAP1", "UMAP2")
+
+nldr9c <- umap_pbmc9 |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2))+
+  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
+  interior_annotation("i")
+```
+
+```{r}
+#| label: combine-data-pbmc-umap
+
+error_pbmc_umap <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_30_min_dist_0.3.rds")
+error_pbmc_umap2 <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_5_min_dist_0.01.rds")
+error_pbmc_umap3 <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_15_min_dist_0.99.rds")
+error_pbmc_umap4 <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_30_min_dist_0.99.rds")
+error_pbmc_umap5 <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_5_min_dist_0.99.rds")
+error_pbmc_umap6 <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_15_min_dist_0.3.rds")
+error_pbmc_umap7 <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_5_min_dist_0.3.rds")
+error_pbmc_umap8 <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_30_min_dist_0.01.rds")
+error_pbmc_umap9 <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_15_min_dist_0.01.rds")
+
+error_pbmc <- bind_rows(error_pbmc_umap, 
+                        error_pbmc_umap2,
+                        error_pbmc_umap3,
+                        error_pbmc_umap4,
+                        error_pbmc_umap5,
+                        error_pbmc_umap6,
+                        error_pbmc_umap7,
+                        error_pbmc_umap8,
+                        error_pbmc_umap9)
+
+error_pbmc <- error_pbmc |>
+  mutate(a1 = round(a1, 2)) |>
+  filter(bin1 >= 5) |>
+  group_by(method, a1) |>
+  filter(MSE == min(MSE)) |>
+  ungroup()
+```
+
+```{r}
+#| label: error-comp-pbmc-umap
+
+error_umap_plot_pbmc <- ggplot(error_pbmc, 
+                          aes(x = a1, 
+                              y = MSE, 
+                              colour = method)) + 
+  geom_point(size = 0.8) +
+  geom_line(linewidth = 0.3) + 
+  # geom_vline(xintercept = 15, linetype="solid", 
+  #            color = "black", linewidth=0.8, alpha = 0.5) +
+  scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 13, 17, 21, 26)]) +
+  scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#8dd3c7','#a65628','#f781bf','#999999')) +
+  scale_y_log10() +
+  ylab("log(MSE)") +
+  xlab(expression(paste("binwidth (", a[1], ")"))) +
+  theme_minimal() +
+  theme(panel.border = element_rect(fill = 'transparent'),
+        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+        axis.ticks.x = element_line(),
+        axis.ticks.y = element_line(),
+        legend.position = "none",
+        axis.text.x = element_text(size = 7),
+        axis.text.y = element_text(size = 7),
+        axis.title.x = element_text(size = 7),
+        axis.title.y = element_text(size = 7))
+
 ```
@@
 
-wrap_plots(nldr1c, nldr2c, nldr3c, ncol = 3)
+free(error_umap_plot_pbmc) + wrap_plots(nldr1c, nldr2c, nldr3c, nldr4c, nldr5c, nldr6c, nldr7c, nldr8c, nldr9c, ncol = 2)
 ```

### 2024-12-02 paper.qmd
Added initial set up for UMAP with PBMC..our method is working !!!!

--- 2024-11-18 paper.qmd
@@
 
+#### scDEED
+
+**tSNE**
+
+```{r}
+result_tsne <- read_rds(here::here('data/pbmc3k/pbmc_scdeed_tsne_results.rds'))
+```
+
+```{r}
+result_tsne |>
+  kableExtra::kable(format = "latex",
+                    booktabs = TRUE,
+                    longtable = TRUE,
+                    label = "scdeedtsne") |>
+  kableExtra::kable_styling(latex_options = "scale_down")
+```
+
+```{r}
+#| out-width: 50%
+
+ggplot(data = result_tsne,
+       aes(
+         x = as.numeric(perplexity),
+         y = number_dubious_cells
+       )) +
+  geom_line() +
+  geom_point() +
+  labs(x = "perplexity",
+       y = "# of dubious cells") +
+  theme_minimal()
+```
+
+**UMAP**
+
+```{r}
+result_umap <- read_rds(here::here('data/pbmc3k/pbmc_scdeed_umap_results.rds'))
+```
+
+```{r}
+result_umap |>
+  kableExtra::kable(format = "latex",
+                    booktabs = TRUE,
+                    longtable = TRUE,
+                    label = "scdeedumap") |>
+  kableExtra::kable_styling(latex_options = "scale_down")
+```
+
+```{r}
+#| out-width: 50%
+
+result_umap |>
+  filter(min.dist == 0.5) |>
+  ggplot(aes(
+           x = as.numeric(n_neighbors),
+           y = number_dubious_cells
+         )) +
+    geom_line() +
+    geom_point() +
+    labs(x = "n_neighbors",
+         y = "# of dubious cells") +
+    theme_minimal()
+```
+
+```{r}
+#| label: read-pbmc-umap-layouts
+# Read a variety of different NLDR representations of PBMC
+# and plot them on same aspect ratio
+umap_pbmc1 <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_30_min_dist_0.3.rds")
+umap_pbmc1 <- as_tibble(umap_pbmc1)
+names(umap_pbmc1) <- c("UMAP1", "UMAP2")
+
+nldr1c <- umap_pbmc1 |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2)) +
+  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
+  interior_annotation("a")
+
+umap_pbmc2 <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_80_min_dist_0.5.rds")
+umap_pbmc2 <- as_tibble(umap_pbmc2)
+names(umap_pbmc2) <- c("UMAP1", "UMAP2")
+
+nldr2c <- umap_pbmc2 |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2))+
+  geom_point(alpha=0.1, size=1, colour='#377eb8') +
+  interior_annotation("b")
+
+```
+
+```{r}
+#| label: combine-data-pbmc-umap
+
+error_pbmc_umap <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_30_min_dist_0.3.rds")
+error_pbmc_umap2 <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_80_min_dist_0.5.rds")
+
+error_pbmc <- bind_rows(error_pbmc_umap, 
+                        error_pbmc_umap2)
+
+error_pbmc <- error_pbmc |>
+  mutate(a1 = round(a1, 2)) |>
+  filter(bin1 >= 5) |>
+  group_by(method, a1) |>
+  filter(MSE == min(MSE)) |>
+  ungroup()
+```
+
+```{r}
+#| label: error-comp-pbmc-umap
+
+error_umap_plot_pbmc <- ggplot(error_pbmc, 
+                               aes(x = a1, 
+                                   y = MSE, 
+                                   colour = method)) + 
+  geom_point(size = 0.8) +
+  geom_line(linewidth = 0.3) + 
+  # geom_vline(xintercept = 15, linetype="solid", 
+  #            color = "black", linewidth=0.8, alpha = 0.5) +
+  scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 13, 17, 21, 26)]) +
+  scale_color_manual(values=c('#e41a1c','#377eb8')) +
+  scale_y_log10() +
+  ylab("log(MSE)") +
+  xlab(expression(paste("binwidth (", a[1], ")"))) +
+  theme_minimal() +
+  theme(panel.border = element_rect(fill = 'transparent'),
+        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+        axis.ticks.x = element_line(),
+        axis.ticks.y = element_line(),
+        legend.position = "none",
+        axis.text.x = element_text(size = 7),
+        axis.text.y = element_text(size = 7),
+        axis.title.x = element_text(size = 7),
+        axis.title.y = element_text(size = 7))
+
+```
+
+```{r}
+#| label: fig-pbmc-mse-umap
+#| fig-pos: H
+#| out-height: 100%
+
+free(error_umap_plot_pbmc) + wrap_plots(nldr1c, nldr2c, ncol = 2)
+```
+
 ### Hand-written digits

### 2024-12-02 paper.tex
Added initial table for scDEED
Added 3 layouts for three UMAP hyperparameters
Added results for different UMAP parameter choices
Set the seed to get the same results as before
Added initial set up for UMAP with PBMC..our method is working !!!!

@@
 
+\subsubsection{scDEED}\label{scdeed}
+
+\textbf{tSNE}
+
+\begin{longtable*}{lr}
+\toprule
+perplexity & number\_dubious\_cells\\
+\midrule
+5 & 87\\
+10 & 143\\
+15 & 146\\
+20 & 101\\
+25 & 93\\
+\addlinespace
+30 & 83\\
+35 & 82\\
+40 & 85\\
+45 & 71\\
+50 & 41\\
+\addlinespace
+55 & 48\\
+60 & 46\\
+65 & 35\\
+70 & 36\\
+75 & 49\\
+\addlinespace
+80 & 46\\
+85 & 75\\
+90 & 82\\
+95 & 62\\
+100 & 20\\
+\addlinespace
+105 & 19\\
+110 & 19\\
+115 & 21\\
+120 & 18\\
+125 & 17\\
+\addlinespace
+130 & 15\\
+135 & 17\\
+140 & 17\\
+145 & 18\\
+\bottomrule
+\end{longtable*}
+
+\begin{center}
+\includegraphics[width=0.5\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-62-1.pdf}
+\end{center}
+
+\textbf{UMAP}
+
+\begin{longtable*}{rrr}
+\toprule
+n\_neighbors & min.dist & number\_dubious\_cells\\
+\midrule
+5 & 0.0125 & 14\\
+6 & 0.0125 & 7\\
+7 & 0.0125 & 5\\
+8 & 0.0125 & 10\\
+9 & 0.0125 & 29\\
+\addlinespace
+10 & 0.0125 & 7\\
+20 & 0.0125 & 9\\
+30 & 0.0125 & 10\\
+40 & 0.0125 & 4\\
+50 & 0.0125 & 3\\
+\addlinespace
+80 & 0.0125 & 47\\
+160 & 0.0125 & 13\\
+240 & 0.0125 & 74\\
+5 & 0.0500 & 12\\
+6 & 0.0500 & 37\\
+\addlinespace
+7 & 0.0500 & 6\\
+8 & 0.0500 & 15\\
+9 & 0.0500 & 33\\
+10 & 0.0500 & 5\\
+20 & 0.0500 & 13\\
+\addlinespace
+30 & 0.0500 & 4\\
+40 & 0.0500 & 4\\
+50 & 0.0500 & 8\\
+80 & 0.0500 & 15\\
+160 & 0.0500 & 55\\
+\addlinespace
+240 & 0.0500 & 75\\
+5 & 0.1000 & 271\\
+6 & 0.1000 & 9\\
+7 & 0.1000 & 16\\
+8 & 0.1000 & 37\\
+\addlinespace
+9 & 0.1000 & 4\\
+10 & 0.1000 & 20\\
+20 & 0.1000 & 5\\
+30 & 0.1000 & 5\\
+40 & 0.1000 & 5\\
+\addlinespace
+50 & 0.1000 & 25\\
+80 & 0.1000 & 4\\
+160 & 0.1000 & 4\\
+240 & 0.1000 & 13\\
+5 & 0.3000 & 45\\
+\addlinespace
+6 & 0.3000 & 10\\
+7 & 0.3000 & 12\\
+8 & 0.3000 & 4\\
+9 & 0.3000 & 3\\
+10 & 0.3000 & 3\\
+\addlinespace
+20 & 0.3000 & 2\\
+30 & 0.3000 & 46\\
+40 & 0.3000 & 2\\
+50 & 0.3000 & 3\\
+80 & 0.3000 & 2\\
+\addlinespace
+160 & 0.3000 & 2\\
+240 & 0.3000 & 21\\
+5 & 0.5000 & 88\\
+6 & 0.5000 & 3\\
+7 & 0.5000 & 5\\
+\addlinespace
+8 & 0.5000 & 4\\
+9 & 0.5000 & 92\\
+10 & 0.5000 & 5\\
+20 & 0.5000 & 72\\
+30 & 0.5000 & 70\\
+\addlinespace
+40 & 0.5000 & 1\\
+50 & 0.5000 & 2\\
+80 & 0.5000 & 2\\
+160 & 0.5000 & 3\\
+240 & 0.5000 & 4\\
+\addlinespace
+5 & 0.7000 & 13\\
+6 & 0.7000 & 86\\
+7 & 0.7000 & 95\\
+8 & 0.7000 & 29\\
+9 & 0.7000 & 95\\
+\addlinespace
+10 & 0.7000 & 6\\
+20 & 0.7000 & 3\\
+30 & 0.7000 & 5\\
+40 & 0.7000 & 2\\
+50 & 0.7000 & 3\\
+\addlinespace
+80 & 0.7000 & 4\\
+160 & 0.7000 & 5\\
+240 & 0.7000 & 4\\
+5 & 0.9000 & 192\\
+6 & 0.9000 & 95\\
+\addlinespace
+7 & 0.9000 & 233\\
+8 & 0.9000 & 228\\
+9 & 0.9000 & 101\\
+10 & 0.9000 & 7\\
+20 & 0.9000 & 3\\
+\addlinespace
+30 & 0.9000 & 79\\
+40 & 0.9000 & 3\\
+50 & 0.9000 & 5\\
+80 & 0.9000 & 4\\
+160 & 0.9000 & 5\\
+\addlinespace
+240 & 0.9000 & 3\\
+\bottomrule
+\end{longtable*}
+
+\begin{center}
+\includegraphics[width=0.5\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-65-1.pdf}
+\end{center}
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=1\textheight]{paper_files/figure-pdf/fig-pbmc-mse-umap-1.pdf}
+
+}
+
+\caption{\label{fig-pbmc-mse-umap}}
+
+\end{figure}%
+
 \subsection{Hand-written digits}\label{hand-written-digits}

### 2024-12-02 paper.qmd
Added tSNE layouts as well

@@
 
+```{r}
+#| label: read-pbmc-tsne-layouts
+# Read a variety of different NLDR representations of PBMC
+# and plot them on same aspect ratio
+tsne_pbmc1 <- read_rds("data/pbmc3k/pbmc_scdeed_tsne_perplexity_30.rds")
+tsne_pbmc1 <- as_tibble(tsne_pbmc1)
+names(tsne_pbmc1) <- c("tSNE1", "tSNE2")
+
+nldr1c <- tsne_pbmc1 |>
+  ggplot(aes(x = tSNE1,
+             y = tSNE2)) +
+  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
+  interior_annotation("a")
+
+tsne_pbmc2 <- read_rds("data/pbmc3k/pbmc_scdeed_tsne_perplexity_320.rds")
+tsne_pbmc2 <- as_tibble(tsne_pbmc2)
+names(tsne_pbmc2) <- c("tSNE1", "tSNE2")
+
+nldr2c <- tsne_pbmc2 |>
+  ggplot(aes(x = tSNE1,
+             y = tSNE2))+
+  geom_point(alpha=0.1, size=1, colour='#377eb8') +
+  interior_annotation("b")
+
+```
+
+```{r}
+#| label: combine-data-pbmc-tsne
+
+error_pbmc_tsne <- read_rds("data/pbmc3k/error_scdeed_pbmc_tsne_perplexity_30.rds")
+error_pbmc_tsne2 <- read_rds("data/pbmc3k/error_scdeed_pbmc_tsne_perplexity_320.rds")
+
+error_pbmc <- bind_rows(error_pbmc_tsne, 
+                        error_pbmc_tsne2)
+
+error_pbmc <- error_pbmc |>
+  mutate(a1 = round(a1, 2)) |>
+  filter(bin1 >= 5) |>
+  group_by(method, a1) |>
+  filter(MSE == min(MSE)) |>
+  ungroup()
+```
+
+```{r}
+#| label: error-comp-pbmc-tsne
+
+error_tsne_plot_pbmc <- ggplot(error_pbmc, 
+                               aes(x = a1, 
+                                   y = MSE, 
+                                   colour = method)) + 
+  geom_point(size = 0.8) +
+  geom_line(linewidth = 0.3) + 
+  # geom_vline(xintercept = 15, linetype="solid", 
+  #            color = "black", linewidth=0.8, alpha = 0.5) +
+  scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 12, 15, 17, 19, 21)]) +
+  scale_color_manual(values=c('#e41a1c','#377eb8')) +
+  scale_y_log10() +
+  ylab("log(MSE)") +
+  xlab(expression(paste("binwidth (", a[1], ")"))) +
+  theme_minimal() +
+  theme(panel.border = element_rect(fill = 'transparent'),
+        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+        axis.ticks.x = element_line(),
+        axis.ticks.y = element_line(),
+        legend.position = "none",
+        axis.text.x = element_text(size = 7),
+        axis.text.y = element_text(size = 7),
+        axis.title.x = element_text(size = 7),
+        axis.title.y = element_text(size = 7))
+
+```
+
+```{r}
+#| label: fig-pbmc-mse-tsne
+#| fig-pos: H
+#| out-height: 100%
+
+free(error_tsne_plot_pbmc) + wrap_plots(nldr1c, nldr2c, ncol = 2)
+```
+
 **UMAP**
@@
   #            color = "black", linewidth=0.8, alpha = 0.5) +
-  scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 13, 17, 21, 26)]) +
+  scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 12, 15, 17, 19)]) +
   scale_color_manual(values=c('#e41a1c','#377eb8')) +

### 2024-12-02 paper.tex
Added tSNE layouts as well

@@
 
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=1\textheight]{paper_files/figure-pdf/fig-pbmc-mse-tsne-1.pdf}
+
+}
+
+\caption{\label{fig-pbmc-mse-tsne}}
+
+\end{figure}%
+
 \textbf{UMAP}
@@
 \begin{center}
-\includegraphics[width=0.5\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-65-1.pdf}
+\includegraphics[width=0.5\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-69-1.pdf}
 \end{center}

### 2024-12-02 paper.qmd
Finalize the aspect rations

--- 2024-12-02 paper.qmd
@@
 
-**tSNE**
-
-```{r}
-result_tsne <- read_rds(here::here('data/pbmc3k/pbmc_scdeed_tsne_results.rds'))
-```
-
-```{r}
-result_tsne |>
-  kableExtra::kable(format = "latex",
-                    booktabs = TRUE,
-                    longtable = TRUE,
-                    label = "scdeedtsne") |>
-  kableExtra::kable_styling(latex_options = "scale_down")
-```
-
-```{r}
-#| out-width: 50%
-
-ggplot(data = result_tsne,
-       aes(
-         x = as.numeric(perplexity),
-         y = number_dubious_cells
-       )) +
-  geom_line() +
-  geom_point() +
-  labs(x = "perplexity",
-       y = "# of dubious cells") +
-  theme_minimal()
-```
-
 **UMAP**
@@
 #| out-width: 50%
+#| eval: false
 
@@
 #| out-height: 100%
-
-free(error_umap_plot_pbmc) + wrap_plots(nldr1c, nldr2c, ncol = 2)
+#| fig-cap: "Assessing which of the 2 tSNE layouts with different perplexity ($30$ (red), $320$ (blue)) on the PBMC data is the better representation using MSE for varying binwidth ($a_1$).
+Colour  used for the lines and points in the left plot and in the scatterplots represents tSNE layout (a, b).
+Layout b is universally suboptimal.
+Layout b is the best choice."
+
+free(error_umap_plot_pbmc) + wrap_plots(nldr1c, nldr2c, ncol = 2) +
+  plot_layout(ncol = 1, heights = c(1, 1))
+```
+
+```{r}
+#| label: published-pbmc-scdeed
+
+training_data_pbmc <- read_rds("data/pbmc3k/pbmc_pca_50_scdeed.rds")
+training_data_pbmc <- training_data_pbmc[, 1:50] |>
+  mutate(ID = 1:NROW(training_data_pbmc))
+
+## For umap
+umap_pbmc <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_80_min_dist_0.5.rds")
+umap_pbmc <- as_tibble(umap_pbmc)
+names(umap_pbmc) <- c("UMAP1", "UMAP2")
+umap_pbmc <- umap_pbmc |>
+  mutate(ID = 1:NROW(umap_pbmc))
+
+pbmc_scaled_obj_umap <- gen_scaled_data(
+  data = umap_pbmc)
+umap_pbmc_scaled <- pbmc_scaled_obj_umap$scaled_nldr
+
+lim1 <- pbmc_scaled_obj_umap$lim1
+lim2 <- pbmc_scaled_obj_umap$lim2
+r2_umap <- diff(lim2)/diff(lim1)
+
+```
+
+```{r}
+#| label: hexbin-pbmc-scdeed
+## Compute hexbin parameters
+num_bins_x_pbmc <- 12
+lim1 <- pbmc_scaled_obj_umap$lim1
+lim2 <- pbmc_scaled_obj_umap$lim2
+r2_pbmc <- diff(lim2)/diff(lim1)
+
+pbmc_model <- fit_highd_model(
+  training_data = training_data_pbmc,
+  emb_df = umap_pbmc_scaled,
+  bin1 = num_bins_x_pbmc,
+  r2 = r2_pbmc,
+  q = 0.1,
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = FALSE,
+  col_start_highd = "pc"
+)
+
+df_bin_centroids_pbmc <- pbmc_model$df_bin_centroids
+df_bin_pbmc <- pbmc_model$df_bin
+
+## Triangulate bin centroids
+tr1_object_pbmc <- tri_bin_centroids(
+  df_bin_centroids_pbmc, x = "c_x", y = "c_y")
+tr_from_to_df_pbmc <- gen_edges(
+  tri_object = tr1_object_pbmc)
+
+## Compute 2D distances
+distance_pbmc <- cal_2d_dist(
+  tr_coord_df = tr_from_to_df_pbmc,
+  start_x = "x_from",
+  start_y = "y_from",
+  end_x = "x_to",
+  end_y = "y_to",
+  select_vars = c("from", "to", "distance"))
+
+## To find the benchmark value
+benchmark_pbmc <- find_lg_benchmark(
+  distance_edges = distance_pbmc,
+  distance_col = "distance")
+
+tr_df <- distinct(tibble(
+  x = c(tr_from_to_df_pbmc[["x_from"]], tr_from_to_df_pbmc[["x_to"]]), 
+  y = c(tr_from_to_df_pbmc[["y_from"]], tr_from_to_df_pbmc[["y_to"]])))
+
+distance_df_small_edges_pbmc <- distance_pbmc |>
+  filter(distance < benchmark_pbmc)
+
+tr_from_to_df_pbmc <- inner_join(
+  tr_from_to_df_pbmc, distance_df_small_edges_pbmc, 
+  by = c("from", "to"))
+
+trimesh_removed_pbmc <- ggplot() + 
+  geom_segment(data = tr_from_to_df_pbmc, 
+               aes(
+                 x = x_from, 
+                 y = y_from, 
+                 xend = x_to, 
+                 yend = y_to),
+               colour = "#000000",
+               linewidth = 1) +
+  geom_point(
+    data = umap_pbmc_scaled,
+    aes(
+      x = UMAP1,
+      y = UMAP2
+    ),
+    alpha = 0.2,
+    color = '#a65628'
+  )  +
+  interior_annotation("a1", cex = 2) +
+  theme(
+    aspect.ratio = 1
+  )
 ```

### 2024-12-02 paper.tex
Finalize the aspect rations

--- 2024-12-02 paper.tex
@@
 
-\textbf{tSNE}
-
-\begin{longtable*}{lr}
-\toprule
-perplexity & number\_dubious\_cells\\
-\midrule
-5 & 87\\
-10 & 143\\
-15 & 146\\
-20 & 101\\
-25 & 93\\
-\addlinespace
-30 & 83\\
-35 & 82\\
-40 & 85\\
-45 & 71\\
-50 & 41\\
-\addlinespace
-55 & 48\\
-60 & 46\\
-65 & 35\\
-70 & 36\\
-75 & 49\\
-\addlinespace
-80 & 46\\
-85 & 75\\
-90 & 82\\
-95 & 62\\
-100 & 20\\
-\addlinespace
-105 & 19\\
-110 & 19\\
-115 & 21\\
-120 & 18\\
-125 & 17\\
-\addlinespace
-130 & 15\\
-135 & 17\\
-140 & 17\\
-145 & 18\\
-\bottomrule
-\end{longtable*}
-
-\begin{center}
-\includegraphics[width=0.5\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-62-1.pdf}
-\end{center}
-
 \textbf{UMAP}
@@
 
-\begin{center}
-\includegraphics[width=0.5\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-65-1.pdf}
-\end{center}
-
 \begin{figure}[H]
@@
 
-\caption{\label{fig-pbmc-mse-umap}}
+\caption{\label{fig-pbmc-mse-umap}Assessing which of the 2 tSNE layouts
+with different perplexity (\(30\) (red), \(320\) (blue)) on the PBMC
+data is the better representation using MSE for varying binwidth
+(\(a_1\)). Colour used for the lines and points in the left plot and in
+the scatterplots represents tSNE layout (a, b). Layout b is universally
+suboptimal. Layout b is the best choice.}
 

### 2024-12-02 paper.qmd
Added initial projection layouts

@@
 
+```{r}
+#| label: pbmc-umap-model-proj-scdeed
+
+data_pbmc <- training_data_pbmc |> 
+  select(-ID) |>
+  mutate(type = "data")
+
+df_b_pbmc <- df_bin_pbmc |>
+  dplyr::filter(hb_id %in% df_bin_centroids_pbmc$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b_pbmc <- df_b_pbmc[match(df_bin_centroids_pbmc$hexID, df_b_pbmc$hb_id),] |>
+  dplyr::select(-hb_id)
+
+# Apply the scaling
+df_model_data_pbmc <- bind_rows(data_pbmc, df_b_pbmc)
+scaled_pbmc <- scale_data_manual(df_model_data_pbmc, "type") |>
+  as_tibble()
+
+scaled_pbmc_data <- scaled_pbmc |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_pbmc_data_model <- scaled_pbmc |>
+  filter(type == "model") |>
+  select(-type)
+
+# Combine with the true model for visualization
+df <- dplyr::bind_rows(scaled_pbmc_data_model |> mutate(type = "model"),
+                       scaled_pbmc_data |> mutate(type = "data"))
+
+## First projection
+projection <- cbind(
+  c(0.040856,0.000723,0.006000,0.007623,0.001986,-0.006599,0.005311,-0.000222,0.001648,0.008067,-0.001587,0.004423,0.004851,0.000638,0.002606,0.006616,0.006188,-0.000954,-0.004603,0.004998,-0.012220,0.008401,-0.007788,0.008972,-0.003903,0.003348,-0.001817,-0.000385,-0.002437,-0.004097,-0.001357,0.013660,0.001770,0.006982,0.002828,-0.009169,-0.003541,0.008582,-0.000032,-0.002375,0.005854,-0.005466,-0.000824,-0.001435,0.004148,0.007141,0.000883,-0.004159,0.005401,0.007882),
+  c(-0.000590,0.039365,-0.004481,-0.009041,-0.003805,0.003278,-0.009601,-0.000650,-0.002333,-0.005697,-0.001488,-0.009808,0.002557,0.004113,0.005705,-0.004665,-0.001170,-0.004727,-0.000845,0.005084,-0.018054,0.002538,0.006606,-0.007184,0.000238,0.001877,-0.002124,-0.005972,-0.009729,0.004958,-0.003181,-0.005185,-0.009598,-0.004243,0.007716,0.000396,-0.004258,0.005932,-0.003242,-0.004184,0.003493,-0.008405,0.000860,0.008304,0.004569,0.000798,0.004156,-0.005777,0.005278,-0.002531))
+
+projection_scaled <- projection * 1.23
+
+# projection_filtered <- projection
+# projection_filtered[apply(projection_filtered, 1, function(row) all(row < 0)), ] <- 0.75
+
+#rows_to_keep <- c(1, 2, 4, 5, 6)
+
+#projection_filtered <- projection[rows_to_keep, ]
+
+projected <- as.matrix(scaled_pbmc_data) %*% projection_scaled
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  #dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection_scaled
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges_pbmc |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+axes_obj <- gen_axes(
+  proj = projection,
+  limits = 1,
+  axis_pos_x = -0.48,
+  axis_pos_y = -0.48,
+  axis_labels = names(scaled_pbmc_data),
+  threshold = 0.082)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
+
+pbmc_proj_umap_model1 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#000000") +
+  geom_point(
+    #size = 0.5,
+    alpha = 0.2,
+    color = '#a65628') +
+  geom_segment(
+    data=axes, 
+    aes(x=x1, y=y1, xend=x2, yend=y2), 
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
+    colour="grey50",
+    size = 3) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.65, 0.65)) +
+  ylim(c(-0.65, 0.65)) +
+  interior_annotation("a2", cex = 2)
+
+## Second projection
+projection <- cbind(
+  c(-0.019825,-0.011101,0.011543,-0.006943,-0.004485,0.000062,0.002023,-0.005178,0.000800,-0.002903,-0.014182,0.008756,0.009595,0.007032,-0.007820,0.006143,0.004865,0.003352,0.007554,0.002265,0.012035,0.009961,-0.001825,-0.004942,-0.002237,0.002308,0.009531,-0.001146,0.013401,-0.001022,0.008434,0.000477,0.004239,-0.006772,-0.004702,-0.011432,-0.000431,0.008146,-0.009299,-0.004782,0.003457,-0.008214,-0.009604,0.011329,0.003805,-0.003004,-0.000800,0.015288,-0.013157,0.006984),
+  c(-0.001387,0.001643,0.010470,0.002930,-0.002335,-0.010569,0.013839,0.002917,-0.006825,0.010191,-0.006367,0.002220,0.002506,0.005139,-0.007213,-0.006235,-0.007123,-0.000222,-0.002225,0.011592,-0.003942,0.002788,-0.000055,0.011316,-0.009753,0.020785,0.004615,0.005172,0.011182,-0.003125,-0.010860,0.007820,-0.017510,0.007010,0.007267,0.002694,-0.003949,-0.009188,-0.005163,0.001898,-0.004744,0.013757,0.010630,-0.007688,0.004598,0.004658,-0.011154,0.003584,-0.005226,0.006650))
+
+projection_scaled <- projection * 1.25
+
+projected <- as.matrix(scaled_pbmc_data) %*% projection_scaled
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  #dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection_scaled
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges_pbmc |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+axes_obj <- gen_axes(
+  proj = projection,
+  limits = 1.35,
+  axis_pos_x = -0.75,
+  axis_pos_y = -0.75,
+  axis_labels = names(scaled_pbmc_data),
+  threshold = 0.12)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
+
+pbmc_proj_umap_model2 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1,
+      y = proj2)) +
+  geom_segment(
+    data = model_df,
+    aes(
+      x = proj1_from,
+      y = proj2_from,
+      xend = proj1_to,
+      yend = proj2_to),
+    color = "#000000") +
+  geom_point(
+    #size = 0.5,
+    alpha = 0.2,
+    color = '#a65628') +
+  geom_segment(
+    data=axes,
+    aes(x=x1, y=y1, xend=x2, yend=y2),
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2),
+    label=rownames(axes),
+    colour="grey50",
+    size = 3) +
+  geom_path(
+    data=circle,
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-1, 0.75)) +
+  ylim(c(-1, 0.75)) +
+  interior_annotation("a3", cex = 2)
+```
+
+```{r}
+#| echo: false
+#| fig-cap: "Model in \\gD{}, on the layout a, and two views of the fit in projections from $50\\text{-}D$, for the PBMC3k data ($n =  5858$ and $p = 50$).
+Layout shows close clusters.
+In $50\\text{-}D$, the data shows "
+#| label: fig-model-pbmc-author-proj-scdeed
+#| fig-pos: H
+#| fig-width: 10
+#| fig-height: 15
+#| out-height: 80%
+
+free(trimesh_removed_pbmc) + 
+  pbmc_proj_umap_model1 + 
+  pbmc_proj_umap_model2 + 
+  plot_layout(nrow=3) &
+  theme(legend.position='none')
+```
+
 ### Hand-written digits

### 2024-12-02 paper.tex
Added initial projection layouts

@@
 
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=0.8\textheight]{paper_files/figure-pdf/fig-model-pbmc-author-proj-scdeed-1.pdf}
+
+}
+
+\caption{\label{fig-model-pbmc-author-proj-scdeed}Model in \gD{}, on the
+layout a, and two views of the fit in projections from \(50\text{-}D\),
+for the PBMC3k data (\(n =  5858\) and \(p = 50\)). Layout shows close
+clusters. In \(50\text{-}D\), the data shows}
+
+\end{figure}%
+
 \subsection{Hand-written digits}\label{hand-written-digits}

### 2024-12-02 paper.qmd
initial scdeed intro

@@
 
+Motivated by this challenge, we present a statistical method, scDEED, for detecting dubious cell embeddings output by a 2D-embedding method.
+By calculating a reliability score for every cell embedding based on the similarity between the cell’s 2D-embedding neighbors and pre-embedding neighbors, scDEED identifies the cell embeddings with low reliability scores as dubious and those with high reliability scores as trustworthy.
+Moreover, by minimizing the number of dubious cell embeddings, scDEED provides intuitive guidance for optimizing the hyperparameters of an embedding method.
+
 **UMAP**

### 2024-12-02 paper.tex
initial scdeed intro

@@
 
+Motivated by this challenge, we present a statistical method, scDEED,
+for detecting dubious cell embeddings output by a 2D-embedding method.
+By calculating a reliability score for every cell embedding based on the
+similarity between the cell's 2D-embedding neighbors and pre-embedding
+neighbors, scDEED identifies the cell embeddings with low reliability
+scores as dubious and those with high reliability scores as trustworthy.
+Moreover, by minimizing the number of dubious cell embeddings, scDEED
+provides intuitive guidance for optimizing the hyperparameters of an
+embedding method.
+
 \textbf{UMAP}

### 2024-12-02 paper.qmd
Finalizing the scDEED section

--- 2024-12-02 paper.qmd
@@
 
-**UMAP**
+In the field of single-cell studies, clustering is a common analytical task used to identify groups of cells with similar expression profiles.
+Non-linear dimensional reduction (NLDR) methods are frequently employed to visualize these clusters and help validate the results.
+However, it is well known that the 2D embeddings produced by t-SNE and UMAP may not accurately reflect the similarities among cell clusters.
+
+To address this challenge, @xia2023 introduces a statistical method called scDEED, which detects unreliable cell embeddings generated by \gD{} embedding techniques.
+scDEED calculates a reliability score for each cell embedding based on the similarity between the cell’s \gD{} embedding neighbors and its neighbors prior to embedding.
+By identifying embeddings with low reliability scores as dubious and those with high scores as trustworthy, scDEED minimizes the number of questionable cell embeddings.
+This approach also provides clear guidance for optimizing the hyperparameters of an embedding method.
+
+To verify that the \gD{} cell embeddings became better aligned with the cell types after scDEED’s optimization, the Human Peripheral Blood Mononuclear Cells (PBMC) dataset was used.
+It contained $31,021$ cells with cell type labels, and the gene expression levels were in the unit of log-transformed UMI count per $10,000$.
+They focused on three sequencing methods (inDrops, DropSeq, and SeqWell) and four common cell types Cytotoxic T cell, CD4+T cell, CD14+ Monocyte, and B cell.
+
+For illustration purposes, we only selected cells generated with inDrops ($n=5858$ cells) and UMAP cell embeddings.
+Also, @xia2023 used first $50$ principal components to generate the UMAP.
+The objective is to assess the optimized layout by scDEED, and if it does not accurately represent the three clusters with small separations of the PBMC dataset, then select a reasonable \gD{} layout.
 
@@
 ```{r}
+#| eval: false
 result_umap |>
@@
 #| out-height: 100%
-#| fig-cap: "Assessing which of the 2 tSNE layouts with different perplexity ($30$ (red), $320$ (blue)) on the PBMC data is the better representation using MSE for varying binwidth ($a_1$).
-Colour  used for the lines and points in the left plot and in the scatterplots represents tSNE layout (a, b).
+#| fig-cap: "Assessing which of the 2 UMAP layouts with different hyperparameter settings (n_neighbors: $30$, min_dist: $0.3$ (red), n_neighbors: $80$, min_dist: $0.5$ (blue)) on the PBMC data is the better representation using MSE for varying binwidth ($a_1$).
+Colour  used for the lines and points in the left plot and in the scatterplots represents UMAP layout (a, b).
 Layout b is universally suboptimal.

### 2024-12-02 paper.tex
Finalizing the scDEED section

--- 2024-12-02 paper.tex
@@
 
-\textbf{UMAP}
-
-\begin{longtable*}{rrr}
-\toprule
-n\_neighbors & min.dist & number\_dubious\_cells\\
-\midrule
-5 & 0.0125 & 14\\
-6 & 0.0125 & 7\\
-7 & 0.0125 & 5\\
-8 & 0.0125 & 10\\
-9 & 0.0125 & 29\\
-\addlinespace
-10 & 0.0125 & 7\\
-20 & 0.0125 & 9\\
-30 & 0.0125 & 10\\
-40 & 0.0125 & 4\\
-50 & 0.0125 & 3\\
-\addlinespace
-80 & 0.0125 & 47\\
-160 & 0.0125 & 13\\
-240 & 0.0125 & 74\\
-5 & 0.0500 & 12\\
-6 & 0.0500 & 37\\
-\addlinespace
-7 & 0.0500 & 6\\
-8 & 0.0500 & 15\\
-9 & 0.0500 & 33\\
-10 & 0.0500 & 5\\
-20 & 0.0500 & 13\\
-\addlinespace
-30 & 0.0500 & 4\\
-40 & 0.0500 & 4\\
-50 & 0.0500 & 8\\
-80 & 0.0500 & 15\\
-160 & 0.0500 & 55\\
-\addlinespace
-240 & 0.0500 & 75\\
-5 & 0.1000 & 271\\
-6 & 0.1000 & 9\\
-7 & 0.1000 & 16\\
-8 & 0.1000 & 37\\
-\addlinespace
-9 & 0.1000 & 4\\
-10 & 0.1000 & 20\\
-20 & 0.1000 & 5\\
-30 & 0.1000 & 5\\
-40 & 0.1000 & 5\\
-\addlinespace
-50 & 0.1000 & 25\\
-80 & 0.1000 & 4\\
-160 & 0.1000 & 4\\
-240 & 0.1000 & 13\\
-5 & 0.3000 & 45\\
-\addlinespace
-6 & 0.3000 & 10\\
-7 & 0.3000 & 12\\
-8 & 0.3000 & 4\\
-9 & 0.3000 & 3\\
-10 & 0.3000 & 3\\
-\addlinespace
-20 & 0.3000 & 2\\
-30 & 0.3000 & 46\\
-40 & 0.3000 & 2\\
-50 & 0.3000 & 3\\
-80 & 0.3000 & 2\\
-\addlinespace
-160 & 0.3000 & 2\\
-240 & 0.3000 & 21\\
-5 & 0.5000 & 88\\
-6 & 0.5000 & 3\\
-7 & 0.5000 & 5\\
-\addlinespace
-8 & 0.5000 & 4\\
-9 & 0.5000 & 92\\
-10 & 0.5000 & 5\\
-20 & 0.5000 & 72\\
-30 & 0.5000 & 70\\
-\addlinespace
-40 & 0.5000 & 1\\
-50 & 0.5000 & 2\\
-80 & 0.5000 & 2\\
-160 & 0.5000 & 3\\
-240 & 0.5000 & 4\\
-\addlinespace
-5 & 0.7000 & 13\\
-6 & 0.7000 & 86\\
-7 & 0.7000 & 95\\
-8 & 0.7000 & 29\\
-9 & 0.7000 & 95\\
-\addlinespace
-10 & 0.7000 & 6\\
-20 & 0.7000 & 3\\
-30 & 0.7000 & 5\\
-40 & 0.7000 & 2\\
-50 & 0.7000 & 3\\
-\addlinespace
-80 & 0.7000 & 4\\
-160 & 0.7000 & 5\\
-240 & 0.7000 & 4\\
-5 & 0.9000 & 192\\
-6 & 0.9000 & 95\\
-\addlinespace
-7 & 0.9000 & 233\\
-8 & 0.9000 & 228\\
-9 & 0.9000 & 101\\
-10 & 0.9000 & 7\\
-20 & 0.9000 & 3\\
-\addlinespace
-30 & 0.9000 & 79\\
-40 & 0.9000 & 3\\
-50 & 0.9000 & 5\\
-80 & 0.9000 & 4\\
-160 & 0.9000 & 5\\
-\addlinespace
-240 & 0.9000 & 3\\
-\bottomrule
-\end{longtable*}
+In the field of single-cell studies, clustering is a common analytical
+task used to identify groups of cells with similar expression profiles.
+Non-linear dimensional reduction (NLDR) methods are frequently employed
+to visualize these clusters and help validate the results. However, it
+is well known that the 2D embeddings produced by t-SNE and UMAP may not
+accurately reflect the similarities among cell clusters.
+
+To address this challenge, \citet{xia2023} introduces a statistical
+method called scDEED, which detects unreliable cell embeddings generated
+by \gD{} embedding techniques. scDEED calculates a reliability score for
+each cell embedding based on the similarity between the cell's \gD{}
+embedding neighbors and its neighbors prior to embedding. By identifying
+embeddings with low reliability scores as dubious and those with high
+scores as trustworthy, scDEED minimizes the number of questionable cell
+embeddings. This approach also provides clear guidance for optimizing
+the hyperparameters of an embedding method.
+
+To verify that the \gD{} cell embeddings became better aligned with the
+cell types after scDEED's optimization, the Human Peripheral Blood
+Mononuclear Cells (PBMC) dataset was used. It contained \(31,021\) cells
+with cell type labels, and the gene expression levels were in the unit
+of log-transformed UMI count per \(10,000\). They focused on three
+sequencing methods (inDrops, DropSeq, and SeqWell) and four common cell
+types Cytotoxic T cell, CD4+T cell, CD14+ Monocyte, and B cell.
+
+For illustration purposes, we only selected cells generated with inDrops
+(\(n=5858\) cells) and UMAP cell embeddings. Also, \citet{xia2023} used
+first \(50\) principal components to generate the UMAP. The objective is
+to assess the optimized layout by scDEED, and if it does not accurately
+represent the three clusters with small separations of the PBMC dataset,
+then select a reasonable \gD{} layout.
 
@@
 
-\caption{\label{fig-pbmc-mse-umap}Assessing which of the 2 tSNE layouts
-with different perplexity (\(30\) (red), \(320\) (blue)) on the PBMC
-data is the better representation using MSE for varying binwidth
+\caption{\label{fig-pbmc-mse-umap}Assessing which of the 2 UMAP layouts
+with different hyperparameter settings (n\_neighbors: \(30\), min\_dist:
+\(0.3\) (red), n\_neighbors: \(80\), min\_dist: \(0.5\) (blue)) on the
+PBMC data is the better representation using MSE for varying binwidth
 (\(a_1\)). Colour used for the lines and points in the left plot and in
-the scatterplots represents tSNE layout (a, b). Layout b is universally
+the scatterplots represents UMAP layout (a, b). Layout b is universally
 suboptimal. Layout b is the best choice.}

### 2024-12-02 paper.qmd
Need to work on projections again

@@
 
-projection_scaled <- projection * 1.23
+projection_scaled <- projection * 3
 
@@
   proj = projection,
-  limits = 1,
-  axis_pos_x = -0.48,
-  axis_pos_y = -0.48,
+  limits = 0.04,
+  axis_pos_x = -0.025,
+  axis_pos_y = -0.025,
   axis_labels = names(scaled_pbmc_data),
-  threshold = 0.082)
+  threshold = 0.0001)
 
@@
   coord_fixed() +
-  xlim(c(-0.65, 0.65)) +
-  ylim(c(-0.65, 0.65)) +
+  xlim(c(-0.03, 0.06)) +
+  ylim(c(-0.03, 0.06)) +
   interior_annotation("a2", cex = 2)
@@
 
-projection_scaled <- projection * 1.25
+projection_scaled <- projection * 3
 
@@
   proj = projection,
-  limits = 1.35,
-  axis_pos_x = -0.75,
-  axis_pos_y = -0.75,
+  limits = 0.03,
+  axis_pos_x = -0.02,
+  axis_pos_y = -0.02,
   axis_labels = names(scaled_pbmc_data),
-  threshold = 0.12)
+  threshold = 0)
 
@@
   coord_fixed() +
-  xlim(c(-1, 0.75)) +
-  ylim(c(-1, 0.75)) +
+  xlim(c(-0.025, 0.02)) +
+  ylim(c(-0.025, 0.02)) +
   interior_annotation("a3", cex = 2)
@@
 #| echo: false
-#| fig-cap: "Model in \\gD{}, on the layout a, and two views of the fit in projections from $50\\text{-}D$, for the PBMC3k data ($n =  5858$ and $p = 50$).
+#| fig-cap: "Model in \\gD{}, on the layout a, and two views of the fit in projections from $50\\text{-}D$, for the PBMC data ($n =  5858$ and $p = 50$).
 Layout shows close clusters.
@@
 #| fig-pos: H
-#| fig-width: 10
-#| fig-height: 15
+#| fig-width: 15
+#| fig-height: 5
 #| out-height: 80%
@@
   pbmc_proj_umap_model2 + 
-  plot_layout(nrow=3) &
+  plot_layout(ncol=3) &
   theme(legend.position='none')

### 2024-12-02 paper.tex
Need to work on projections again

@@
 layout a, and two views of the fit in projections from \(50\text{-}D\),
-for the PBMC3k data (\(n =  5858\) and \(p = 50\)). Layout shows close
+for the PBMC data (\(n =  5858\) and \(p = 50\)). Layout shows close
 clusters. In \(50\text{-}D\), the data shows}

### 2024-12-03 paper.qmd
Initial work on replacing the tSNE with UMAP

@@
 #| label: umap-two-curvy
-umap_two_curvy <- read_rds(file = "data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_umap_n-neigbors_15_min-dist_0.1.rds")
+umap_two_curvy <- read_rds(file = "data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_30.rds")
 
@@
 nldr_two_curvy <- umap_two_curvy_scaled |> 
-  ggplot(aes(x = UMAP1, y = UMAP2)) + 
+  ggplot(aes(x = tSNE1, y = tSNE2)) + 
   geom_point(alpha=0.5, colour="#636363", size = 0.5) +
@@
 nldr_two_curvy2 <- umap_two_curvy_scaled |> 
-  ggplot(aes(x = UMAP1, y = UMAP2)) + 
+  ggplot(aes(x = tSNE1, y = tSNE2)) + 
   geom_point(alpha=0.5, colour="#636363", size = 0.5) +
@@
 nldr_two_curvy_original2 <- umap_two_curvy_scaled |> 
-  ggplot(aes(x = UMAP1, y = UMAP2)) + 
+  ggplot(aes(x = tSNE1, y = tSNE2)) + 
   geom_point(alpha=0.5, color = "#636363", size = 0.5) +
@@
   geom_point(data = umap_two_curvy_scaled, 
-             aes(x = UMAP1, y = UMAP2), 
+             aes(x = tSNE1, y = tSNE2), 
              alpha = 0.5, size = 0.5, color = "#636363") +
@@
   geom_point(data = umap_two_curvy_scaled,
-             aes(x = UMAP1, y = UMAP2),
+             aes(x = tSNE1, y = tSNE2),
              alpha = 0.3,
@@
              aes(
-               x = UMAP1,
-               y = UMAP2
+               x = tSNE1,
+               y = tSNE2
              ),
@@
   geom_point(data = umap_two_curvy_scaled,
-           aes(x = UMAP1, y = UMAP2),
+           aes(x = tSNE1, y = tSNE2),
            alpha = 0.3,
@@
 error_plot_two_curvy <- error_df_two_curvy_abs |>
-  ggplot(aes(x = UMAP1,
-             y = UMAP2,
+  ggplot(aes(x = tSNE1,
+             y = tSNE2,
              colour = sqrt_row_wise_abs_error)) +
@@
 #| fig-cap: "Comparison of prediction generated using the exiting `umap` R package's prediction method and our method: (a) A view of the true model in projection from $7\\text{-}D$, (b) predicted data from the `umap` prediction method, and (c) predicted data from our method."
+#| eval: false
 

### 2024-12-03 paper.tex
Initial work on replacing the tSNE with UMAP

@@
 
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=0.8\textheight]{paper_files/figure-pdf/fig-predict-two-curvy-1.pdf}
-
-}
-
-\caption{\label{fig-predict-two-curvy}Comparison of prediction generated
-using the exiting \texttt{umap} R package's prediction method and our
-method: (a) A view of the true model in projection from \(7\text{-}D\),
-(b) predicted data from the \texttt{umap} prediction method, and (c)
-predicted data from our method.}
-
-\end{figure}%
-
 \subsection{Tuning}\label{tuning}

### 2024-12-10 paper.qmd
working on curiosities section: look for DHC on things that need fixing

@@
 
-## A curious difference between tSNE, UMAP and PaCMAP revealer
+## Curiosities
 
@@
 
-In this section, the evaluation focuses on how effectively the tSNE, UMAP, and PaCMAP algorithms preserve the clustering structure.
+With the drawing of the model in the data, several interesting differences between NLDR methods can be observed.
+
+### Ordering of points
+
+DHC: Will need to have more bins in each of the methods to illustrate this properly.
+
+To illustrate a difference in how the methods organise points in the NLDR layout, simulated 4D data having five Gaussian clusters is used.
+@fig-five-gau-projs a1, b1, c1 show the 2D layouts for (a) t-SNE, (b) UMAP, and (c) PaCMAP, respectively.
+The default hyper-parameters are used.
+All three methods show the five clusters, with varying degrees of separation.
+
+The models are fitted to these layouts, but we focus on a single cluster to illustrate the curious detail.
+@fig-five-gau-projs a2, b2, c2 show the fitted models in a projection of the 4D space.
+In this projection the difference between methods can be seen.
+These clusters are fully 4D in nature, so we would expect the model to be a *crumpled 2D sheet* that stretches in all four dimensions.
+This is what is observed for t-SNE and UMAP.
+The curious detail is that the model for PaCMAP is closer to a *pancake* in shape!
+What this means is that there has to be some ordering of points in the 2D PaCMAP layout that induces the flat model, likely that the points are organised by the global principal components.
+So the layout of the model in 4D doesn't reflect the dimension of each cluster.
+DHC: is there something in the documentation that might help to explain what has happened?
+
+This pattern can also be observed by examining the error from the model fit (@fig-five-gau-projs a3, b3, c3).
+With t-SNE and UMAP the error is relatively uniformly distributed, but with PaCMAP there is more error in the center of the 2D layout, reflecting that there are more points in the middle that are further from the fitted model.
+
+<!-- In this section, the evaluation focuses on how effectively the tSNE, UMAP, and PaCMAP algorithms preserve the clustering structure.
 The data consists of five $4\text{-}D$ clusters, each with $1000$ points generated from a multivariate normal distribution with a small variance of $0.0025$.
-The first cluster is centered at $(0, 0, 0, 0)$, the second at $(1, 0, 0, 0)$, the third at $(0, 1, 0, 0)$, the fourth at $(0, 0, 1, 0)$, and the fifth at $(0, 0, 0, 1)$.
+The first cluster is centered at $(0, 0, 0, 0)$, the second at $(1, 0, 0, 0)$, the third at $(0, 1, 0, 0)$, the fourth at $(0, 0, 1, 0)$, and the fifth at $(0, 0, 0, 1)$.-->
 
@@
 
-In the tSNE layout (@fig-five-gau-projs (a1)), the distances between clusters appear smaller, leading to more tightly packed clusters, which is typical of tSNE's focus on preserving local structures.
+<!-- In the tSNE layout (@fig-five-gau-projs (a1)), the distances between clusters appear smaller, leading to more tightly packed clusters, which is typical of tSNE's focus on preserving local structures.
 This often results in less emphasis on the global distances between clusters.
 On the other hand, UMAP (@fig-five-gau-projs (b1)) tends to maintain larger separations between clusters, better reflecting the global structure and preserving the relative distances between clusters more effectively.
-In contrast, PaCMAP (@fig-five-gau-projs (c1)) aim to balance these aspects by maintaining the distance between clusters and accurately positioning them, effectively capturing both global and local structures.
-
-By visualizing the model of different layouts in $4\text{-}D$ reveal different characteristics of each NLDR method.
+In contrast, PaCMAP (@fig-five-gau-projs (c1)) aim to balance these aspects by maintaining the distance between clusters and accurately positioning them, effectively capturing both global and local structures.-->
+
+<!-- By visualizing the model of different layouts in $4\text{-}D$ reveal different characteristics of each NLDR method.
 When examining the model of a specific cluster, which appears more like a filled-out shape in tSNE (@fig-five-gau-projs (a2, a3)), and UMAP (@fig-five-gau-projs (b2, b3)), while the model has a pancake shape in PaCMAP (@fig-five-gau-projs (c2, c3)).
@@
 On the other hand, the reason for the pancake shape in PaCMAP is the fail to capture the local structure.
+-->
 
@@
 
-The differences in performance can be attributed to several factors.
+<!-- The differences in performance can be attributed to several factors.
 One key element is the choice of hyper-parameters, particularly the number of neighbors considered during the dimensionality reduction process.
@@
 Adjusting these parameters could lead to different results, potentially improving PaCMAP's ability to capture local structures or altering how tSNE and UMAP handle global and local structures.
-Thus, while PaCMAP generally balances global and local preservation, careful tuning of its parameters is important for achieving the best results in any given dataset.
+Thus, while PaCMAP generally balances global and local preservation, careful tuning of its parameters is important for achieving the best results in any given dataset.-->
 
@@
              y = tSNE2)) +
-  geom_point(alpha=0.3, color = "#000000") +
+  geom_point(alpha=0.3) +
   geom_rect(aes(xmin = -0.05, xmax = 0.35, ymin = 0.28, ymax = 0.6),
-               fill = "transparent", color = "red", linewidth = 0.8)
+               fill = "transparent", color = "grey50", linewidth = 0.8)
 
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 13
+num_bins_x_gau1 <- 20 #13
 lim1 <- gau1_scaled_obj$lim1
@@
              y = tSNE2,
-             colour = sqrt_row_wise_abs_error)) +
-  geom_point(alpha=0.5) + 
+             z = sqrt_row_wise_abs_error)) +
+  stat_summary_hex(bins=15) +  
   xlim(c(0.3, 0.66)) + ylim(c(-0.01, 0.35)) +
-  scale_colour_continuous_sequential(palette = "YlOrRd", n_interp = 20) +
+  scale_fill_continuous_sequential(palette = "YlOrRd", n_interp = 20) +
   interior_annotation("a3", 
@@
 trimesh_removed_gau_tsne <- ggplot() + 
-  geom_segment(data = tr_from_to_df_gau1, 
-               aes(
-                 x = x_from, 
-                 y = y_from, 
-                 xend = x_to, 
-                 yend = y_to),
-               colour = "#000000",
-               linewidth = 1) +
+  #geom_segment(data = tr_from_to_df_gau1, 
+  #             aes(
+  #               x = x_from, 
+  #               y = y_from, 
+  #               xend = x_to, 
+  #               yend = y_to),
+  #             colour = "#000000",
+  #             linewidth = 1) +
   geom_point(
@@
     ),
-    alpha = 0.2,
-    #size = 0.1,
-    color = "#999999"
-  ) + 
+    alpha = 0.2) + 
   geom_rect(
@@
           fill = "transparent", 
-      color = "red", 
+      color = "grey50", 
       linewidth = 0.5) +
@@
   dplyr::filter(distance < benchmark_gau1)
+
+# library(langevitour)
 
@@
              y = UMAP2,
-             colour = sqrt_row_wise_abs_error)) +
-  geom_point(alpha=0.5) + 
+             z = sqrt_row_wise_abs_error)) +
+  stat_summary_hex(bins=15) +  
   xlim(c(0.47, 0.5475)) + ylim(c(0.0125, 0.09)) +
-  scale_colour_continuous_sequential(palette = "YlOrRd", n_interp = 20) +
+  scale_fill_continuous_sequential(palette = "YlOrRd", n_interp = 20) +
   interior_annotation("b3", 
@@
 trimesh_removed_gau_umap <- ggplot() + 
-  geom_segment(data = tr_from_to_df_gau1, 
-               aes(
-                 x = x_from, 
-                 y = y_from, 
-                 xend = x_to, 
-                 yend = y_to),
-               colour = "#000000",
-               linewidth = 1) +
+  #geom_segment(data = tr_from_to_df_gau1, 
+  #             aes(
+  #               x = x_from, 
+  #               y = y_from, 
+  #               xend = x_to, 
+  #               yend = y_to),
+  #             colour = "#000000",
+  #             linewidth = 1) +
   geom_point(
@@
     ),
-    alpha = 0.05,
-    #size = 0.1,
-    color = "#999999"
-  ) + 
+    alpha = 0.05) + 
   geom_rect(
@@
     fill = "transparent", 
-    color = "red", 
+    color = "grey50", 
     linewidth = 0.5)  +
@@
              y = PaCMAP2,
-             colour = sqrt_row_wise_abs_error)) +
-  geom_point(alpha=0.6,
-             size = 1.5) + 
+             z = sqrt_row_wise_abs_error)) +
+  stat_summary_hex(bins=15) + 
   xlim(c(-0.02, 0.18)) + ylim(c(0.12, 0.32)) +
-  scale_colour_continuous_sequential(palette = "YlOrRd", breaks = 20) +
+  scale_fill_continuous_sequential(palette = "YlOrRd", breaks = 20) +
   interior_annotation("c3", 
@@
 trimesh_removed_gau_pacmap <- ggplot() + 
-  geom_segment(data = tr_from_to_df_gau1, 
-               aes(
-                 x = x_from, 
-                 y = y_from, 
-                 xend = x_to, 
-                 yend = y_to),
-               colour = "#000000",
-               linewidth = 1) +
+  #geom_segment(data = tr_from_to_df_gau1, 
+  #             aes(
+  #               x = x_from, 
+  #               y = y_from, 
+  #               xend = x_to, 
+  #               yend = y_to),
+  #             colour = "#000000",
+  #             linewidth = 1) +
   geom_point(
@@
     ),
-    alpha = 0.3,
-    #size = 0.1,
-    color = "#999999"
-  ) +
+    alpha = 0.3) +
   geom_rect(
@@
     fill = "transparent", 
-    color = "red", 
+    color = "grey50", 
     linewidth = 0.5) +
@@
 <!-- ::: -->
+
+### The effect of density
 

### 2024-12-10 paper.tex
working on curiosities section: look for DHC on things that need fixing

@@
 \makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\newsavebox\pandoc@box
+\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
+  \sbox\pandoc@box{#1}%
+  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
+  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
+  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
+  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
+  \else\usebox{\pandoc@box}%
+  \fi%
+}
 % Set default figure placement to htbp
-\makeatletter
 \def\fps@figure{htbp}
@@
 
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
 \usepackage[]{natbib}
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-NLDR-variety-1.pdf}
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-NLDR-variety-1.pdf}
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-two-curvy-true-proj-1.pdf}
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-two-curvy-true-proj-1.pdf}
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-NLDR-two-curvy-1.pdf}
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-NLDR-two-curvy-1.pdf}
 
@@
 
-\includegraphics[width=1\textwidth,height=0.3\textheight]{paper_files/figure-pdf/fig-hex-param-1.pdf}
+\includegraphics[width=1\linewidth,height=0.3\textheight]{paper_files/figure-pdf/fig-hex-param-1.pdf}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/best-fit-umap1-1.pdf}
+{\centering \includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/best-fit-umap1-1.pdf}
 
@@
 
-\includegraphics[width=1\textwidth,height=0.3\textheight]{paper_files/figure-pdf/fig-p-d-error-in-2d-two-curvy-1.pdf}
+\includegraphics[width=1\linewidth,height=0.3\textheight]{paper_files/figure-pdf/fig-p-d-error-in-2d-two-curvy-1.pdf}
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-bins-two-curvy-1.pdf}
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-bins-two-curvy-1.pdf}
 
@@
 
-\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-param-two-curvy-1.pdf}
+\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-param-two-curvy-1.pdf}
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-rm-lg-1.pdf}
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-rm-lg-1.pdf}
 
@@
 
-\includegraphics[width=1\textwidth,height=0.6\textheight]{paper_files/figure-pdf/fig-two_non_linear_diff_shaped_close_clusters-mse-1.pdf}
+\includegraphics[width=1\linewidth,height=0.6\textheight]{paper_files/figure-pdf/fig-two_non_linear_diff_shaped_close_clusters-mse-1.pdf}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=0.9\textheight]{paper_files/figure-pdf/best-fit-umap-tsne-pacamp-1.pdf}
+{\centering \includegraphics[width=1\linewidth,height=0.9\textheight]{paper_files/figure-pdf/best-fit-umap-tsne-pacamp-1.pdf}
 
@@
 
-\section{A curious difference between tSNE, UMAP and PaCMAP
-revealer}\label{a-curious-difference-between-tsne-umap-and-pacmap-revealer}
-
-In this section, the evaluation focuses on how effectively the tSNE,
-UMAP, and PaCMAP algorithms preserve the clustering structure. The data
-consists of five \(4\text{-}D\) clusters, each with \(1000\) points
-generated from a multivariate normal distribution with a small variance
-of \(0.0025\). The first cluster is centered at \((0, 0, 0, 0)\), the
-second at \((1, 0, 0, 0)\), the third at \((0, 1, 0, 0)\), the fourth at
-\((0, 0, 1, 0)\), and the fifth at \((0, 0, 0, 1)\).
-
-In the tSNE layout (Figure~\ref{fig-five-gau-projs} (a1)), the distances
-between clusters appear smaller, leading to more tightly packed
-clusters, which is typical of tSNE's focus on preserving local
-structures. This often results in less emphasis on the global distances
-between clusters. On the other hand, UMAP
-(Figure~\ref{fig-five-gau-projs} (b1)) tends to maintain larger
-separations between clusters, better reflecting the global structure and
-preserving the relative distances between clusters more effectively. In
-contrast, PaCMAP (Figure~\ref{fig-five-gau-projs} (c1)) aim to balance
-these aspects by maintaining the distance between clusters and
-accurately positioning them, effectively capturing both global and local
-structures.
-
-By visualizing the model of different layouts in \(4\text{-}D\) reveal
-different characteristics of each NLDR method. When examining the model
-of a specific cluster, which appears more like a filled-out shape in
-tSNE (Figure~\ref{fig-five-gau-projs} (a2, a3)), and UMAP
-(Figure~\ref{fig-five-gau-projs} (b2, b3)), while the model has a
-pancake shape in PaCMAP (Figure~\ref{fig-five-gau-projs} (c2, c3)). The
-reason to have filled-out shaped clusters in tSNE and UMAP is that the
-ability of capture the local structure. On the other hand, the reason
-for the pancake shape in PaCMAP is the fail to capture the local
-structure.
-
-The differences in performance can be attributed to several factors. One
-key element is the choice of hyper-parameters, particularly the number
-of neighbors considered during the dimensionality reduction process.
-This parameter significantly impacts how well local structures are
-maintained. Additionally, the balance between global and local
-structures, and issues related to optimization and initialization all
-contribute to the observed outcomes.
-
-It's important to note that these layouts are made with default
-hyper-parameter settings. Adjusting these parameters could lead to
-different results, potentially improving PaCMAP's ability to capture
-local structures or altering how tSNE and UMAP handle global and local
-structures. Thus, while PaCMAP generally balances global and local
-preservation, careful tuning of its parameters is important for
-achieving the best results in any given dataset.
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-five-gau-projs-1.pdf}
+\section{Curiosities}\label{curiosities}
+
+With the drawing of the model in the data, several interesting
+differences between NLDR methods can be observed.
+
+\subsection{Ordering of points}\label{ordering-of-points}
+
+DHC: Will need to have more bins in each of the methods to illustrate
+this properly.
+
+To illustrate a difference in how the methods organise points in the
+NLDR layout, simulated 4D data having five Gaussian clusters is used.
+Figure~\ref{fig-five-gau-projs} a1, b1, c1 show the 2D layouts for (a)
+t-SNE, (b) UMAP, and (c) PaCMAP, respectively. The default
+hyper-parameters are used. All three methods show the five clusters,
+with varying degrees of separation.
+
+The models are fitted to these layouts, but we focus on a single cluster
+to illustrate the curious detail. Figure~\ref{fig-five-gau-projs} a2,
+b2, c2 show the fitted models in a projection of the 4D space. In this
+projection the difference between methods can be seen. These clusters
+are fully 4D in nature, so we would expect the model to be a
+\emph{crumpled 2D sheet} that stretches in all four dimensions. This is
+what is observed for t-SNE and UMAP. The curious detail is that the
+model for PaCMAP is closer to a \emph{pancake} in shape! What this means
+is that there has to be some ordering of points in the 2D PaCMAP layout
+that induces the flat model, likely that the points are organised by the
+global principal components. So the layout of the model in 4D doesn't
+reflect the dimension of each cluster. DHC: is there something in the
+documentation that might help to explain what has happened?
+
+This pattern can also be observed by examining the error from the model
+fit (Figure~\ref{fig-five-gau-projs} a3, b3, c3). With t-SNE and UMAP
+the error is relatively uniformly distributed, but with PaCMAP there is
+more error in the center of the 2D layout, reflecting that there are
+more points in the middle that are further from the fitted model.
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-five-gau-projs-1.pdf}
 
@@
 
+\subsection{The effect of density}\label{the-effect-of-density}
+
 \section{Applications}\label{sec-applications}
@@
 
-\includegraphics[width=1\textwidth,height=1\textheight]{paper_files/figure-pdf/fig-pbmc-mse-1.pdf}
+\includegraphics[width=1\linewidth,height=1\textheight]{paper_files/figure-pdf/fig-pbmc-mse-1.pdf}
 
@@
 
-\includegraphics[width=1\textwidth,height=0.8\textheight]{paper_files/figure-pdf/fig-model-pbmc-author-proj-1.pdf}
+\includegraphics[width=1\linewidth,height=0.8\textheight]{paper_files/figure-pdf/fig-model-pbmc-author-proj-1.pdf}
 
@@
 
-\includegraphics[width=1\textwidth,height=1\textheight]{paper_files/figure-pdf/fig-pbmc-mse-umap-1.pdf}
+\includegraphics[width=1\linewidth,height=1\textheight]{paper_files/figure-pdf/fig-pbmc-mse-umap-1.pdf}
 
@@
 
-\includegraphics[width=1\textwidth,height=0.8\textheight]{paper_files/figure-pdf/fig-model-pbmc-author-proj-scdeed-1.pdf}
+\includegraphics[width=1\linewidth,height=0.8\textheight]{paper_files/figure-pdf/fig-model-pbmc-author-proj-scdeed-1.pdf}
 
@@
 
-\includegraphics[width=1\textwidth,height=1\textheight]{paper_files/figure-pdf/fig-mnist-mse-1.pdf}
+\includegraphics[width=1\linewidth,height=1\textheight]{paper_files/figure-pdf/fig-mnist-mse-1.pdf}
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-mnist-tri-proj-1.pdf}
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-mnist-tri-proj-1.pdf}
 
@@
 
-\includegraphics[width=1\textwidth,height=0.8\textheight]{paper_files/figure-pdf/fig-model-error-mnist-1.pdf}
+\includegraphics[width=1\linewidth,height=0.8\textheight]{paper_files/figure-pdf/fig-model-error-mnist-1.pdf}
 

### 2024-12-10 paper.tex
Initial work on dense cluster

--- 2024-12-03 paper.tex
@@
 
-\section{A curious difference between tSNE, UMAP and PaCMAP
-revealer}\label{a-curious-difference-between-tsne-umap-and-pacmap-revealer}
-
-In this section, the evaluation focuses on how effectively the tSNE,
-UMAP, and PaCMAP algorithms preserve the clustering structure. The data
-consists of five \(4\text{-}D\) clusters, each with \(1000\) points
-generated from a multivariate normal distribution with a small variance
-of \(0.0025\). The first cluster is centered at \((0, 0, 0, 0)\), the
-second at \((1, 0, 0, 0)\), the third at \((0, 1, 0, 0)\), the fourth at
-\((0, 0, 1, 0)\), and the fifth at \((0, 0, 0, 1)\).
-
-In the tSNE layout (Figure~\ref{fig-five-gau-projs} (a1)), the distances
-between clusters appear smaller, leading to more tightly packed
-clusters, which is typical of tSNE's focus on preserving local
-structures. This often results in less emphasis on the global distances
-between clusters. On the other hand, UMAP
-(Figure~\ref{fig-five-gau-projs} (b1)) tends to maintain larger
-separations between clusters, better reflecting the global structure and
-preserving the relative distances between clusters more effectively. In
-contrast, PaCMAP (Figure~\ref{fig-five-gau-projs} (c1)) aim to balance
-these aspects by maintaining the distance between clusters and
-accurately positioning them, effectively capturing both global and local
-structures.
-
-By visualizing the model of different layouts in \(4\text{-}D\) reveal
-different characteristics of each NLDR method. When examining the model
-of a specific cluster, which appears more like a filled-out shape in
-tSNE (Figure~\ref{fig-five-gau-projs} (a2, a3)), and UMAP
-(Figure~\ref{fig-five-gau-projs} (b2, b3)), while the model has a
-pancake shape in PaCMAP (Figure~\ref{fig-five-gau-projs} (c2, c3)). The
-reason to have filled-out shaped clusters in tSNE and UMAP is that the
-ability of capture the local structure. On the other hand, the reason
-for the pancake shape in PaCMAP is the fail to capture the local
-structure.
-
-The differences in performance can be attributed to several factors. One
-key element is the choice of hyper-parameters, particularly the number
-of neighbors considered during the dimensionality reduction process.
-This parameter significantly impacts how well local structures are
-maintained. Additionally, the balance between global and local
-structures, and issues related to optimization and initialization all
-contribute to the observed outcomes.
-
-It's important to note that these layouts are made with default
-hyper-parameter settings. Adjusting these parameters could lead to
-different results, potentially improving PaCMAP's ability to capture
-local structures or altering how tSNE and UMAP handle global and local
-structures. Thus, while PaCMAP generally balances global and local
-preservation, careful tuning of its parameters is important for
-achieving the best results in any given dataset.
+\section{Curiosities}\label{curiosities}
+
+With the drawing of the model in the data, several interesting
+differences between NLDR methods can be observed.
+
+\subsection{Ordering of points}\label{ordering-of-points}
+
+DHC: Will need to have more bins in each of the methods to illustrate
+this properly.
+
+To illustrate a difference in how the methods organise points in the
+NLDR layout, simulated 4D data having five Gaussian clusters is used.
+Figure~\ref{fig-five-gau-projs} a1, b1, c1 show the 2D layouts for (a)
+t-SNE, (b) UMAP, and (c) PaCMAP, respectively. The default
+hyper-parameters are used. All three methods show the five clusters,
+with varying degrees of separation.
+
+The models are fitted to these layouts, but we focus on a single cluster
+to illustrate the curious detail. Figure~\ref{fig-five-gau-projs} a2,
+b2, c2 show the fitted models in a projection of the 4D space. In this
+projection the difference between methods can be seen. These clusters
+are fully 4D in nature, so we would expect the model to be a
+\emph{crumpled 2D sheet} that stretches in all four dimensions. This is
+what is observed for t-SNE and UMAP. The curious detail is that the
+model for PaCMAP is closer to a \emph{pancake} in shape! What this means
+is that there has to be some ordering of points in the 2D PaCMAP layout
+that induces the flat model, likely that the points are organised by the
+global principal components. So the layout of the model in 4D doesn't
+reflect the dimension of each cluster. DHC: is there something in the
+documentation that might help to explain what has happened?
+
+This pattern can also be observed by examining the error from the model
+fit (Figure~\ref{fig-five-gau-projs} a3, b3, c3). With t-SNE and UMAP
+the error is relatively uniformly distributed, but with PaCMAP there is
+more error in the center of the 2D layout, reflecting that there are
+more points in the middle that are further from the fitted model.
 
@@
 \end{figure}%
+
+\subsection{The effect of density}\label{the-effect-of-density}
 

### 2024-12-10 scripts/one_c_shaped_dens_structure/01_gen_data.R

@@ -1,0 +1,31 @@
+### This script is to generate an example for the effect of density
+
+library(tibble)
+library(dplyr)
+library(langevitour)
+library(readr)
+
+set.seed(20240110)
+
+# Function to gen a curvilinear cluster in 4D space with an offset
+gen_curv1_3d <- function(n) {
+  # Bias sampling of 'a' but keep it within the same range (-0.5 to 0)
+  a <- 3 * pi * (stats::rbeta(n = n, shape1 = 1, shape2 = 10) * -0.9)
+  x2 <- sin(a)
+  x1 <- 2.0 * stats::runif(n = n)
+  x3 <- sign(a) * (cos(a) - 1)
+
+  df <- tibble(
+    x1 = x1,
+    x2 = x2,
+    x3 = x3
+  )
+
+  return(df)
+}
+
+sample_size <- 1000
+curve1 <- gen_curv1_3d(n = sample_size)
+langevitour(curve1)
+
+write_rds(curve1, here::here("data/one_c_shaped_dens_structure/one_c_shaped_dens_data.rds"))

### 2024-12-10 scripts/one_c_shaped_dens_structure/03_gen_model_with_tSNE.R

@@ -1,0 +1,176 @@
+## Import necessary libraries
+library(quollr)
+library(dplyr)
+library(readr)
+library(langevitour)
+library(ggplot2)
+
+## Import data
+training_data_one_nonlinear_structure <- read_rds("data/one_c_shaped_dens_structure/one_c_shaped_dens_data.rds")
+
+# training_data_one_nonlinear_structure <- training_data_one_nonlinear_structure |>
+#   mutate(across(everything(), ~ (. - mean(.)) / sd(.)))
+
+training_data_one_nonlinear_structure <- training_data_one_nonlinear_structure |>
+  mutate(ID = 1:NROW(training_data_one_nonlinear_structure))
+
+tSNE_one_nonlinear_structure <- read_rds("data/one_c_shaped_dens_structure/one_c_shaped_dens_structure_tsne_perplexity_30.rds")
+
+one_nonlinear_structure_scaled_obj <- gen_scaled_data(
+  data = tSNE_one_nonlinear_structure)
+tSNE_one_nonlinear_structure_scaled <- one_nonlinear_structure_scaled_obj$scaled_nldr |>
+  mutate(ID = 1:NROW(tSNE_one_nonlinear_structure))
+
+## Compute hexbin parameters
+num_bins_x_one_nonlinear_structure <- 12
+lim1 <- one_nonlinear_structure_scaled_obj$lim1
+lim2 <- one_nonlinear_structure_scaled_obj$lim2
+r2_one_nonlinear_structure <- diff(lim2)/diff(lim1)
+
+one_nonlinear_structure_model <- fit_highd_model(
+  training_data = training_data_one_nonlinear_structure,
+  emb_df = tSNE_one_nonlinear_structure_scaled,
+  bin1 = num_bins_x_one_nonlinear_structure,
+  r2 = r2_one_nonlinear_structure,
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = FALSE,
+  col_start_highd = "x"
+)
+
+df_bin_centroids_one_nonlinear_structure <- one_nonlinear_structure_model$df_bin_centroids
+df_bin_one_nonlinear_structure <- one_nonlinear_structure_model$df_bin
+
+## Triangulate bin centroids
+tr1_object_one_nonlinear_structure <- tri_bin_centroids(
+  df_bin_centroids_one_nonlinear_structure, x = "c_x", y = "c_y")
[…]

### 2025-01-06 paper.qmd
Updated Figure 3

@@
 #| label: umap-two-curvy
-umap_two_curvy <- read_rds(file = "data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_30.rds")
+tsne_two_curvy <- read_rds(file = "data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_30.rds")
 
 two_curvy_scaled_obj <- gen_scaled_data(
-  data = umap_two_curvy)
-
-umap_two_curvy_scaled <- two_curvy_scaled_obj$scaled_nldr |>
+  data = tsne_two_curvy)
+
+tsne_two_curvy_scaled <- two_curvy_scaled_obj$scaled_nldr |>
   mutate(ID = row_number())
@@
 sc_xlims <- c(-0.27, 1.2)
-sc_ylims <- c(-0.2, 0.8)
-
-nldr_two_curvy <- umap_two_curvy_scaled |> 
+sc_ylims <- c(-0.2, 1.1)
+
+nldr_two_curvy <- tsne_two_curvy_scaled |> 
   ggplot(aes(x = tSNE1, y = tSNE2)) + 
@@
 
-nldr_two_curvy2 <- umap_two_curvy_scaled |> 
+nldr_two_curvy2 <- tsne_two_curvy_scaled |> 
   ggplot(aes(x = tSNE1, y = tSNE2)) + 
@@
 
-nldr_two_curvy_original <- umap_two_curvy_scaled |> 
+nldr_two_curvy_original <- tsne_two_curvy_scaled |> 
   ggplot(aes(x = UMAP1, y = UMAP2)) + 
@@
 
-nldr_two_curvy_original2 <- umap_two_curvy_scaled |> 
+nldr_two_curvy_original2 <- tsne_two_curvy_scaled |> 
   ggplot(aes(x = tSNE1, y = tSNE2)) + 
@@
 ## Compute hexbin parameters
-num_bins_x_two_curvy <- 12
+num_bins_x_two_curvy <- 20
 
@@
 hb_obj_two_curvy <- hex_binning(
-  data = umap_two_curvy_scaled, 
+  data = tsne_two_curvy_scaled, 
   bin1 = num_bins_x_two_curvy, 
@@
                fill = "#ffffff") +
-  geom_point(data = umap_two_curvy_scaled, 
+  geom_point(data = tsne_two_curvy_scaled, 
              aes(x = tSNE1, y = tSNE2), 
@@
       fill = std_counts), color = "grey70", linewidth=0.2) +
-  geom_point(data = umap_two_curvy_scaled,
+  geom_point(data = tsne_two_curvy_scaled,
              aes(x = tSNE1, y = tSNE2),
@@
                colour = "#000000") +
-  geom_point(data = umap_two_curvy_scaled,
+  geom_point(data = tsne_two_curvy_scaled,
              aes(
@@
 hb_obj_notation <- hex_binning(
-  data = umap_two_curvy_scaled, 
+  data = tsne_two_curvy_scaled, 
   bin1 = 7, 
@@
 hb_obj_two_curvy2 <- hex_binning(
-  data = umap_two_curvy_scaled, 
+  data = tsne_two_curvy_scaled, 
   bin1 = 13, 
@@
         linewidth=0.2) +
-  geom_point(data = umap_two_curvy_scaled,
+  geom_point(data = tsne_two_curvy_scaled,
            aes(x = tSNE1, y = tSNE2),
@@
                colour = "#000000")+
-  geom_point(data = umap_two_curvy_scaled,
+  geom_point(data = tsne_two_curvy_scaled,
              aes(
@@
 hb_obj_two_curvy3 <- hex_binning(
-  data = umap_two_curvy_scaled, 
+  data = tsne_two_curvy_scaled, 
   bin1 = 23, 
@@
         color = "grey70", linewidth=0.2) +
-  geom_point(data = umap_two_curvy_scaled,
+  geom_point(data = tsne_two_curvy_scaled,
            aes(x = tSNE1, y = tSNE2),
@@
 
-plot_predict_umap_two_curvy <- ggplot(
+plot_predict_tsne_two_curvy <- ggplot(
   data = true_pred_df, 
@@
 error_df_two_curvy_abs <- error_df_two_curvy_abs |>
-  bind_cols(umap_two_curvy_scaled |>
+  bind_cols(tsne_two_curvy_scaled |>
               select(-ID))
@@
 
-predict_umap_two_curvy <- read_rds(file = "data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_umap_predict_true.rds")
+predict_tsne_two_curvy <- read_rds(file = "data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_umap_predict_true.rds")
 
 predict_two_curvy_obj <- gen_scaled_data(
-    data = predict_umap_two_curvy)
-
-predict_umap_two_curvy_scaled <- predict_two_curvy_obj$scaled_nldr
+    data = predict_tsne_two_curvy)
+
+predict_tsne_two_curvy_scaled <- predict_two_curvy_obj$scaled_nldr
 
 plot_predict_umap_fun_two_curvy <- ggplot(
-  data = predict_umap_two_curvy_scaled, 
+  data = predict_tsne_two_curvy_scaled, 
   aes(x = UMAP1, 
@@
 
-predict_umap_two_curvy_scaled <- predict_umap_two_curvy_scaled |>
+predict_tsne_two_curvy_scaled <- predict_tsne_two_curvy_scaled |>
   mutate(ID = row_number()) |>
@@
 predict_point_df <- bind_rows(
-  predict_umap_two_curvy_scaled,
+  predict_tsne_two_curvy_scaled,
   jittered_points_df
@@
 
-proj_plot_two_curvy_true + plot_predict_umap_fun_two_curvy +plot_predict_umap_two_curvy +
+proj_plot_two_curvy_true + plot_predict_umap_fun_two_curvy +plot_predict_tsne_two_curvy +
   plot_layout(ncol = 3)

### 2025-01-06 paper.qmd
Updated Figure 2

@@
 ## Compute hexbin parameters
-num_bins_x_two_curvy <- 20
+num_bins_x_two_curvy <- 23
 
@@
   coord_fixed() +
-  xlim(c(-0.126, 0.125)) + 
-  ylim(c(-0.126, 0.125)) +
+  xlim(c(-0.165, 0.165)) + 
+  ylim(c(-0.165, 0.165)) +
   interior_annotation("c", sc_ltr_pos) +
@@
   coord_fixed() +
-  xlim(c(-0.126, 0.125)) + 
-  ylim(c(-0.126, 0.125)) +
+  xlim(c(-0.165, 0.165)) + 
+  ylim(c(-0.165, 0.165)) +
   interior_annotation("d", sc_ltr_pos) +
@@
 projection <- cbind(
-    c(-0.08707,-0.06493,0.07017,-0.07349,-0.02955,-0.05286,-0.02765),
-    c(0.03398,0.03735,0.09272,0.01871,-0.09224,0.00451,0.08079))
+    c(0.11062,-0.02743,0.02786,-0.01930,0.02374,-0.03493,0.03159),
+    c(-0.01110,-0.06886,0.04730,0.09890,-0.00571,-0.00038,0.00169))
 
@@
   limits = 0.15,
-  axis_pos_x = -0.1,
-  axis_pos_y = -0.1,
+  axis_pos_x = -0.067,
+  axis_pos_y = -0.067,
   axis_labels = names(scaled_two_curvy_data),
-  threshold = 0.01)
+  threshold = 0.007)
 
@@
   coord_fixed() +
-  xlim(c(-0.126, 0.125)) + 
-  ylim(c(-0.126, 0.125)) +
+  xlim(c(-0.1, 0.1)) + 
+  ylim(c(-0.1, 0.1)) +
   interior_annotation("a", c(0.08, 0.93), cex = 2) +
@@
 projection <- cbind(
-    c(-0.02704,0.08281,-0.07671,-0.02996,0.08674,0.06497,-0.02090),
-    c(0.00095,0.08031,0.05721,0.10299,0.04816,-0.05837,-0.02226))
+    c(-0.03372,0.07726,0.03948,-0.00096,-0.02976,-0.07968,0.03194),
+    c(0.07080,0.00316,-0.00382,0.05069,-0.07334,-0.02506,-0.05749))
 
@@
   proj = projection * 6,
-  limits = 0.13,
-  axis_pos_x = -0.1,
-  axis_pos_y = -0.1,
+  limits = 0.1,
+  axis_pos_x = -0.05,
+  axis_pos_y = -0.05,
   axis_labels = names(scaled_two_curvy_data),
-  threshold = 0.0113)
+  threshold = 0.007)
 
@@
   coord_fixed() +
-  xlim(c(-0.125, 0.13)) + 
-  ylim(c(-0.125, 0.13)) +
+  xlim(c(-0.07, 0.07)) + 
+  ylim(c(-0.07, 0.07)) +
   interior_annotation("b", c(0.08, 0.93), cex = 2) +

### 2025-01-06 paper.qmd
Updated Figure 5

@@
 projection <- cbind(
-  c(-0.08707,-0.06493,0.07017,-0.07349,-0.02955,-0.05286,-0.02765),
-  c(0.03398,0.03735,0.09272,0.01871,-0.09224,0.00451,0.08079))
+    c(0.11062,-0.02743,0.02786,-0.01930,0.02374,-0.03493,0.03159),
+    c(-0.01110,-0.06886,0.04730,0.09890,-0.00571,-0.00038,0.00169))
 
@@
   coord_fixed() +
-  xlim(c(-0.165, 0.165)) + 
-  ylim(c(-0.165, 0.165)) +
+  xlim(c(-0.1, 0.1)) + 
+  ylim(c(-0.1, 0.1)) +
   interior_annotation("c", sc_ltr_pos) +
@@
   coord_fixed() +
-  xlim(c(-0.165, 0.165)) + 
-  ylim(c(-0.165, 0.165)) +
+  xlim(c(-0.1, 0.1)) + 
+  ylim(c(-0.1, 0.1)) +
   interior_annotation("d", sc_ltr_pos) +
@@
 #| fig-width: 10
-#| fig-height: 8
+#| fig-height: 10
 
@@
   two_curvy_proj_umap_all_model1 +
-  plot_layout(ncol = 2, heights = c(1, 2))
+  plot_layout(ncol = 2)
 ```

### 2025-01-06 paper.qmd
Updated Figure 7

@@
   interior_annotation("a", sc_ltr_pos) + 
-  xlim(c(-0.3, 1.2)) + ylim(c(-0.17, 0.78))
+  xlim(c(-0.3, 1.2)) + ylim(c(-0.17, 1.1))
 
@@
     na.value = "#ffffff", option = "C") +
-  xlim(c(-0.3, 1.2)) + ylim(c(-0.17, 0.78)) +
+  xlim(c(-0.3, 1.2)) + ylim(c(-0.17, 1.1)) +
   interior_annotation("b", sc_ltr_pos)
@@
   scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "D") +
-  xlim(c(-0.3, 1.2)) + ylim(c(-0.17, 0.78)) +
+  xlim(c(-0.3, 1.2)) + ylim(c(-0.17, 1.1)) +
   interior_annotation("c", sc_ltr_pos)
@@
 #| fig-pos: H
-#| fig-cap: "Hexbin density plots of UMAP layout of the two C-shaped clusters data, using three different bin inputs: (a) $b = 80 \\text{ } (10, \\text{ }8)$, (b) $b = 130 \\text{ } (13, \\text{ }10)$, and (c) $b = 391 \\text{ } (23, \\text{ }17)$.
+#| fig-cap: "Hexbin density plots of tSNE layout of the two C-shaped clusters data, using three different bin inputs: (a) $b = 80 \\text{ } (10, \\text{ }8)$, (b) $b = 130 \\text{ } (13, \\text{ }10)$, and (c) $b = 391 \\text{ } (23, \\text{ }17)$.
 Color indicates standardized counts, dark indicating high count and light indicates low count.
@@
 #| fig-width: 9
-#| fig-height: 2
+#| fig-height: 3
 

### 2025-01-06 paper.qmd
Updated benchmark value of 2nd figure 9

@@
   data = tsne_two_curvy_scaled, 
-  bin1 = 13, 
+  bin1 = 23, 
   r2 = r2)
@@
 a1_2 <- calc_bins_y(
-  bin1 = 13, 
+  bin1 = 23, 
   r2 = r2
@@
 
-benchmark_two_curvy <- 3.5 * bin_width
+benchmark_two_curvy <- 2.5 * bin_width
 
@@
   coord_fixed() +
-  xlim(c(-0.126, 0.125)) + 
-  ylim(c(-0.126, 0.125)) +
+  xlim(c(-0.1, 0.1)) + 
+  ylim(c(-0.1, 0.1)) +
   interior_annotation("c2", sc_ltr_pos) +
@@
 #| label: fig-rm-lg
-#| fig-cap: "Distribution of edge lengths and wireframes in \\gD{} and $7\\text{-}D$ with Delaunay triangulation, and two choices of long edge removal: (b) benchmark = 4.5$a_1$ and (c) benchmark = 2$a_1$, where $a_1$ = 0.09."
+#| fig-cap: "Distribution of edge lengths and wireframes in \\gD{} and $7\\text{-}D$ with Delaunay triangulation, and two choices of long edge removal: (b) benchmark = 2.5$a_1$ and (c) benchmark = 2$a_1$, where $a_1$ = 0.09."
 #| fig-height: 15

### 2025-01-06 paper.tex
Updated Figure 7
Updated benchmark value of 2nd figure 9

@@
 
-\caption{\label{fig-bins-two-curvy}Hexbin density plots of UMAP layout
+\caption{\label{fig-bins-two-curvy}Hexbin density plots of tSNE layout
 of the two C-shaped clusters data, using three different bin inputs: (a)
@@
 \gD{} and \(7\text{-}D\) with Delaunay triangulation, and two choices of
-long edge removal: (b) benchmark = 4.5\(a_1\) and (c) benchmark =
+long edge removal: (b) benchmark = 2.5\(a_1\) and (c) benchmark =
 2\(a_1\), where \(a_1\) = 0.09.}

### 2025-01-06 paper.qmd
Updated Figure9

@@
   #                    )) +
-  # geom_vline(xintercept = 2,
-  #            linetype="solid",
-  #            color = "#bdbdbd",
-  #            linewidth=1) +
+  geom_vline(xintercept = 2.5 * bin_width,
+             linetype="solid",
+             color = "#bdbdbd",
+             linewidth=1) +
+  geom_vline(xintercept = 2 * bin_width,
+           linetype="solid",
+           color = "#bdbdbd",
+           linewidth=1) +
   # geom_vline(xintercept = 4.5,
@@
     size = 0.5,
-    alpha = 0.5,
+    alpha = 0.2,
     color = "#636363") +
@@
   coord_fixed() +
-  xlim(c(-0.126, 0.125)) + 
-  ylim(c(-0.126, 0.125)) +
+  xlim(c(-0.1, 0.1)) + 
+  ylim(c(-0.1, 0.1)) +
   interior_annotation("a3", sc_ltr_pos) +
@@
   coord_fixed() +
-  xlim(c(-0.126, 0.125)) + 
-  ylim(c(-0.126, 0.125)) +
+  xlim(c(-0.1, 0.1)) + 
+  ylim(c(-0.1, 0.1)) +
   interior_annotation("a2", sc_ltr_pos) +
@@
   coord_fixed() +
-  xlim(c(-0.126, 0.125)) + 
-  ylim(c(-0.126, 0.125)) +
+  xlim(c(-0.1, 0.1)) + 
+  ylim(c(-0.1, 0.1)) +
   interior_annotation("b2", sc_ltr_pos) +
@@
   coord_fixed() +
-  xlim(c(-0.126, 0.125)) + 
-  ylim(c(-0.126, 0.125)) +
+  xlim(c(-0.1, 0.1)) + 
+  ylim(c(-0.1, 0.1)) +
   interior_annotation("a4", sc_ltr_pos) +

### 2025-01-06 paper.qmd
some fixes to the curiosities

@@
 #| fig-cap: "The components of the hexagon grid illustrating notation."
-#| out-height: 30%
+#| out-width: 30%
 #| fig-pos: H
@@
 Most large errors are distributed near the edges of the clusters."
-#| out-height: 30%
+#| out-width: 30%
 #| fig-pos: H
@@
 #| fig-height: 10
-#| out-height: 80%
+#| out-width: 80%
 #| fig-cap: "Comparison of prediction generated using the exiting `umap` R package's prediction method and our method: (a) A view of the true model in projection from $7\\text{-}D$, (b) predicted data from the `umap` prediction method, and (c) predicted data from our method."
@@
 At the smallest bin size, the data structure is discontinuous, suggesting that there are too many bins.
-Using the MSE of the model fit in $7\\text{-}D$ helps decide on a useful choice of number of bins."
+Using the MSE of the model fit in $7\\text{-}D$ helps decide on a useful choice of number of bins.
+XXX All look the same"
 #| fig-width: 9
@@
 #| fig-pos: H
-#| out-height: 60%
+#| out-width: 60%
 
@@
 #| fig-height: 20
-#| out-height: 90%
+#| out-width: 90%
 
@@
 #| label: fig-five-gau-projs
-#| fig-height: 15
+#| fig-height: 10
 #| fig-width: 15
 #| fig-pos: H
-#| fig-cap: "Model in \\gD{}, on the layouts a1, b1, and c1, and two views of the fit in projections of each layout from $4\\text{-}D$, for the five Gaussian cluster data.
-All \\gD{} models show five well-separated clusters with different gaps.
-Layout a1 shows small separations, while layout b2 shows large gaps.
-On the other hand, layout b3 shows average gaps with exact positions as in $4\\text{-}D$.
-Assessing the different models fit for different layouts for a specific cluster (highlighted in red) shows filled-out shape in a1, a2, and b1, b2.
-On the other hand, c1 and c2 shows pancake shape.
-Videos of the langevitour animations are available at <https://youtu.be/5Y1hE4i7N2k>, <https://youtu.be/5Y1hE4i7N2k>, and <https://youtu.be/5Y1hE4i7N2k> respectively."
+#| fig-cap: "NLDR's organise points in the \\gD{} layout in different ways, possibly misleadingly, illustrated using three layouts: (a) t-SNE, (b) UMAP, (c) PaCMap.
+The data has five Gaussian cluster in $4\\text{-}D$.
+The bottom row of plots shows a \\gD{} projection from a tour on $4\\text{-}D$ revealing the differences generated by the layouts on the model fits.
+We would expect the model fit to be like that in (a2) where it is distinctly is separate for each cluster but like a hairball in each.
+This would indicate the distinct clusters, each being fully $4\\text{-}D$.
+With (c2), the curiousity is that the model is a \\gD{} pancake shape in $4\\text{-}D$, indicating that there is some ordering of points done by PaCMAP, posisbly along some principal component axes.
+Videos of the langevitour animations are available at <https://youtu.be/oQxEb4wRdHI>, <https://youtu.be/JW49csPpDx4>, and XXX respectively."
 
 trimesh_removed_gau_tsne + trimesh_removed_gau_umap + trimesh_removed_gau_pacmap +
-five_gau_proj_tsne_model1 + five_gau_proj_umap_model1 + five_gau_proj_pacmap_model1 + error_df_gau_tsne + error_df_gau_umap + error_df_gau_pacmap +
-  plot_layout(guides = "collect", nrow = 3) &
+five_gau_proj_tsne_model1 + five_gau_proj_umap_model1 + five_gau_proj_pacmap_model1 + #error_df_gau_tsne + error_df_gau_umap + error_df_gau_pacmap +
+  plot_layout(guides = "collect", nrow = 2) &
   theme(legend.position='none')
@@
 #| fig-pos: H
-#| out-height: 100%
+#| out-width: 100%
 
@@
 #| fig-height: 15
-#| out-height: 80%
+#| out-width: 80%
 
@@
 
-#### scDEED
+#### Comparison with results of scDEED recommendations
 
@@
 #| fig-pos: H
-#| out-height: 100%
+#| out-width: 100%
 #| fig-cap: "Assessing which of the 2 UMAP layouts with different hyperparameter settings (n_neighbors: $30$, min_dist: $0.3$ (red), n_neighbors: $80$, min_dist: $0.5$ (blue)) on the PBMC data is the better representation using MSE for varying binwidth ($a_1$).
@@
 #| fig-height: 5
-#| out-height: 80%
+#| out-width: 80%
 
@@
 #| fig-pos: H
-#| out-height: 22%
+#| out-width: 22%
 #| eval: false

### 2025-01-06 paper.qmd
Commit

--- 2025-01-06 paper.qmd
@@
 ## Compute hexbin parameters
-num_bins_x_two_curvy <- 23
+num_bins_x_two_curvy <- 15
 
@@
   data = tsne_two_curvy_scaled, 
-  bin1 = 23, 
+  bin1 = 48, 
   r2 = r2)
@@
 a1_3 <- calc_bins_y(
-  bin1 = 23, 
+  bin1 = 48, 
   r2 = r2

### 2025-01-06 paper.tex

--- 2024-12-10 paper.tex
@@
 
-\includegraphics[width=1\linewidth,height=0.3\textheight]{paper_files/figure-pdf/fig-hex-param-1.pdf}
+\includegraphics[width=0.3\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-hex-param-1.pdf}
 
@@
 
-\includegraphics[width=1\linewidth,height=0.3\textheight]{paper_files/figure-pdf/fig-p-d-error-in-2d-two-curvy-1.pdf}
+\includegraphics[width=0.3\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-p-d-error-in-2d-two-curvy-1.pdf}
 
@@
 there are too many bins. Using the MSE of the model fit in
-\(7\text{-}D\) helps decide on a useful choice of number of bins.}
+\(7\text{-}D\) helps decide on a useful choice of number of bins. XXX
+All look the same}
 
@@
 
-\includegraphics[width=1\linewidth,height=0.6\textheight]{paper_files/figure-pdf/fig-two_non_linear_diff_shaped_close_clusters-mse-1.pdf}
+\includegraphics[width=0.6\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-two_non_linear_diff_shaped_close_clusters-mse-1.pdf}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.9\textheight]{paper_files/figure-pdf/best-fit-umap-tsne-pacamp-1.pdf}
+{\centering \includegraphics[width=0.9\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/best-fit-umap-tsne-pacamp-1.pdf}
 
@@
 
-\caption{\label{fig-five-gau-projs}Model in \gD{}, on the layouts a1,
-b1, and c1, and two views of the fit in projections of each layout from
-\(4\text{-}D\), for the five Gaussian cluster data. All \gD{} models
-show five well-separated clusters with different gaps. Layout a1 shows
-small separations, while layout b2 shows large gaps. On the other hand,
-layout b3 shows average gaps with exact positions as in \(4\text{-}D\).
-Assessing the different models fit for different layouts for a specific
-cluster (highlighted in red) shows filled-out shape in a1, a2, and b1,
-b2. On the other hand, c1 and c2 shows pancake shape. Videos of the
-langevitour animations are available at
-\url{https://youtu.be/5Y1hE4i7N2k}, \url{https://youtu.be/5Y1hE4i7N2k},
-and \url{https://youtu.be/5Y1hE4i7N2k} respectively.}
+\caption{\label{fig-five-gau-projs}NLDR's organise points in the \gD{}
+layout in different ways, possibly misleadingly, illustrated using three
+layouts: (a) t-SNE, (b) UMAP, (c) PaCMap. The data has five Gaussian
+cluster in \(4\text{-}D\). The bottom row of plots shows a \gD{}
+projection from a tour on \(4\text{-}D\) revealing the differences
+generated by the layouts on the model fits. We would expect the model
+fit to be like that in (a2) where it is distinctly is separate for each
+cluster but like a hairball in each. This would indicate the distinct
+clusters, each being fully \(4\text{-}D\). With (c2), the curiousity is
+that the model is a \gD{} pancake shape in \(4\text{-}D\), indicating
+that there is some ordering of points done by PaCMAP, posisbly along
+some principal component axes. Videos of the langevitour animations are
+available at \url{https://youtu.be/oQxEb4wRdHI},
+\url{https://youtu.be/JW49csPpDx4}, and XXX respectively.}
 
@@
 
-\includegraphics[width=1\linewidth,height=1\textheight]{paper_files/figure-pdf/fig-pbmc-mse-1.pdf}
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-pbmc-mse-1.pdf}
 
@@
 
-\includegraphics[width=1\linewidth,height=0.8\textheight]{paper_files/figure-pdf/fig-model-pbmc-author-proj-1.pdf}
+\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-model-pbmc-author-proj-1.pdf}
 
@@
 
-\subsubsection{scDEED}\label{scdeed}
+\subsubsection{Comparison with results of scDEED
+recommendations}\label{comparison-with-results-of-scdeed-recommendations}
 
@@
 
-\includegraphics[width=1\linewidth,height=1\textheight]{paper_files/figure-pdf/fig-pbmc-mse-umap-1.pdf}
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-pbmc-mse-umap-1.pdf}
 
@@
 
-\includegraphics[width=1\linewidth,height=0.8\textheight]{paper_files/figure-pdf/fig-model-pbmc-author-proj-scdeed-1.pdf}
+\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-model-pbmc-author-proj-scdeed-1.pdf}
 
@@
 
-\includegraphics[width=1\linewidth,height=1\textheight]{paper_files/figure-pdf/fig-mnist-mse-1.pdf}
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-mnist-mse-1.pdf}
 
@@
 
-\includegraphics[width=1\linewidth,height=0.8\textheight]{paper_files/figure-pdf/fig-model-error-mnist-1.pdf}
+\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-model-error-mnist-1.pdf}
 

### 2025-01-06 paper.tex
Fixed the URL issues in References

--- 2025-01-06 paper.tex
@@
 
-\includegraphics[width=1\textwidth,height=0.3\textheight]{paper_files/figure-pdf/fig-hex-param-1.pdf}
+\includegraphics[width=0.3\textwidth,height=\textheight]{paper_files/figure-pdf/fig-hex-param-1.pdf}
 
@@
 
-\includegraphics[width=1\textwidth,height=0.3\textheight]{paper_files/figure-pdf/fig-p-d-error-in-2d-two-curvy-1.pdf}
+\includegraphics[width=0.3\textwidth,height=\textheight]{paper_files/figure-pdf/fig-p-d-error-in-2d-two-curvy-1.pdf}
 
@@
 
-\includegraphics[width=1\textwidth,height=0.6\textheight]{paper_files/figure-pdf/fig-two_non_linear_diff_shaped_close_clusters-mse-1.pdf}
+\includegraphics[width=0.6\textwidth,height=\textheight]{paper_files/figure-pdf/fig-two_non_linear_diff_shaped_close_clusters-mse-1.pdf}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=0.9\textheight]{paper_files/figure-pdf/best-fit-umap-tsne-pacamp-1.pdf}
+{\centering \includegraphics[width=0.9\textwidth,height=\textheight]{paper_files/figure-pdf/best-fit-umap-tsne-pacamp-1.pdf}
 
@@
 
-\includegraphics[width=1\textwidth,height=1\textheight]{paper_files/figure-pdf/fig-pbmc-mse-1.pdf}
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-pbmc-mse-1.pdf}
 
@@
 
-\includegraphics[width=1\textwidth,height=0.8\textheight]{paper_files/figure-pdf/fig-model-pbmc-author-proj-1.pdf}
+\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-model-pbmc-author-proj-1.pdf}
 
@@
 
-\includegraphics[width=1\textwidth,height=1\textheight]{paper_files/figure-pdf/fig-pbmc-mse-umap-1.pdf}
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-pbmc-mse-umap-1.pdf}
 
@@
 
-\includegraphics[width=1\textwidth,height=0.8\textheight]{paper_files/figure-pdf/fig-model-pbmc-author-proj-scdeed-1.pdf}
+\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-model-pbmc-author-proj-scdeed-1.pdf}
 
@@
 
-\includegraphics[width=1\textwidth,height=1\textheight]{paper_files/figure-pdf/fig-mnist-mse-1.pdf}
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-mnist-mse-1.pdf}
 
@@
 
-\includegraphics[width=1\textwidth,height=0.8\textheight]{paper_files/figure-pdf/fig-model-error-mnist-1.pdf}
+\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-model-error-mnist-1.pdf}
 

### 2025-01-06 paper.qmd
Fixed Figure 7

@@
 ## Compute hexbin parameters
-num_bins_x_two_curvy <- 15
+num_bins_x_two_curvy <- 23
 
@@
   data = tsne_two_curvy_scaled, 
-  bin1 = 23, 
+  bin1 = 15, 
   r2 = r2)
@@
 a1_2 <- calc_bins_y(
-  bin1 = 23, 
+  bin1 = 15, 
   r2 = r2
@@
 
+hex_grid_coloured_two_curvy2 +
   hex_grid_coloured_two_curvy1 +
-  hex_grid_coloured_two_curvy2 +
   hex_grid_coloured_two_curvy3 +

### 2025-01-06 paper.qmd
Removed best fit layouts

@@
 #| out-width: 90%
+#| eval: false
 

### 2025-01-06 paper.tex
Removed best fit layouts

@@
 
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.9\textwidth,height=\textheight]{paper_files/figure-pdf/best-fit-umap-tsne-pacamp-1.pdf}
-
-}
-
-\caption{Best fit for two C-shaped clusters data using UMAP (a1-a4)
-(\(a_1\) = 0.09) and tSNE (b1-b4) (\(a_1\) = 0.16) with the same number
-of non-empty bins (\(m\) = 39). Both models show twist in \(7\text{-}D\)
-(a3, a4, b3, b4). Also, both NLDR methods are unable to accurately
-capture the width of the curvilinear clusters.}
-
-\end{figure}%
-
 \section{Linked plots}\label{linked-plots}

### 2025-01-06 paper.qmd
Updated Figure 8

@@
 
-error_two_non_linear_diff_shaped_close_clusters_umap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_umap.rds")
+error_two_non_linear_diff_shaped_close_clusters_umap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_tsne.rds")
 

### 2025-01-06 paper.qmd
Updated Figure 8

@@
                          y = MSE)) +
-  geom_vline(xintercept = 0.09,
+  geom_vline(xintercept = 0.08,
+           color = "#d95f02", linetype=2, 
+           linewidth=1) +
+  geom_vline(xintercept = 0.05,
+             color = "#1f78b4", 
              linetype="solid",
-           color = "#d95f02", linewidth=1) +
-  geom_vline(xintercept = 0.05,
-             color = "#1b9e77", linetype=2,
              linewidth=1) +
-  geom_vline(xintercept = 0.13, linetype=2,
-             color = "#1f78b4", linewidth=1) +
+  geom_vline(xintercept = 0.02, linetype=2,
+             color = "#1b9e77", linewidth=1) +
   geom_line(linewidth = 0.5) + 
@@
                          y = prop_comp)) +
-  geom_vline(xintercept = 0.09,
+  geom_vline(xintercept = 0.08,
+           color = "#d95f02", linetype=2, 
+           linewidth=1) +
+  geom_vline(xintercept = 0.05,
+             color = "#1f78b4", 
              linetype="solid",
-           color = "#d95f02", linewidth=1) +
-  geom_vline(xintercept = 0.05,
-             color = "#1b9e77", linetype=2,
              linewidth=1) +
-  geom_vline(xintercept = 0.13, linetype=2,
-             color = "#1f78b4", linewidth=1) +
+  geom_vline(xintercept = 0.02, linetype=2,
+             color = "#1b9e77", linewidth=1) +
   geom_line(linewidth = 0.5) +
@@
                          y = prop_bins)) +
-  geom_vline(xintercept = 0.09,
+  geom_vline(xintercept = 0.08,
+           color = "#d95f02", linetype=2, 
+           linewidth=1) +
+  geom_vline(xintercept = 0.05,
+             color = "#1f78b4", 
              linetype="solid",
-           color = "#d95f02", linewidth=1) +
-  geom_vline(xintercept = 0.05,
-             color = "#1b9e77", linetype=2,
              linewidth=1) +
-  geom_vline(xintercept = 0.13, linetype=2,
-             color = "#1f78b4", linewidth=1) +
+  geom_vline(xintercept = 0.02, linetype=2,
+             color = "#1b9e77", linewidth=1) +
   geom_line(linewidth = 0.5) +
@@
   scale_x_continuous(breaks = sort(unique(error_rm_two_curvy$benchmark_rm_lwd))[c(1, 3, 5, 7, 9, 11, 13, 15, 17)]) +
-  scale_color_manual(values=c("#1f78b4", "#d95f02", "#1b9e77")) +
+  scale_color_manual(values=c("#d95f02", "#1f78b4", "#1b9e77")) +
   xlab("threshold to remove low-density hexagons") +

### 2025-01-06 paper.qmd
Initial work on the effect of density

@@
   newdata = NULL,
-  type_NLDR = "UMAP",
+  type_NLDR = "tSNE",
   col_start = "x")
@@
 
+```{r}
+source("scripts/one_c_shaped_dens_structure/04_gen_error_plots.R")
+```
+
+```{r}
+#| fig-height: 5
+#| fig-width: 15
+#| fig-pos: H
+generate_error_plots_one_c_shaped()
+```
+
 ## Applications {#sec-applications}

### 2025-01-06 scripts/one_c_shaped_dens_structure/04_gen_error_plots.R

@@ -1,0 +1,89 @@
+### This script is to generate error plot for one_c_shaped_dens cluster with tSNE
+library(readr)
+library(dplyr)
+
+one_c_shaped_data <- read_rds(here::here("data/one_c_shaped_dens_structure/one_c_shaped_dens_data.rds"))
+one_c_shaped_data <- one_c_shaped_data |>
+  mutate(ID = row_number())
+
+tsne_one_c_shaped <- read_rds(file = "data/one_c_shaped_dens_structure/one_c_shaped_dens_structure_tsne_perplexity_30.rds")
+
+tsne_one_c_shaped_scaled_obj <- gen_scaled_data(
+  data = tsne_one_c_shaped)
+
+tsne_one_c_shaped_scaled <- tsne_one_c_shaped_scaled_obj$scaled_nldr |>
+  mutate(ID = row_number())
+lim1 <- tsne_one_c_shaped_scaled_obj$lim1
+lim2 <- tsne_one_c_shaped_scaled_obj$lim2
+r2 <- diff(lim2)/diff(lim1)
+
+num_bins <- 23
+
+## hexagon binning to have regular hexagons
+hb_obj_one_c_shaped <- hex_binning(
+  data = tsne_one_c_shaped_scaled,
+  bin1 = num_bins,
+  r2 = r2)
+
+a1_2 <- calc_bins_y(
+  bin1 = num_bins,
+  r2 = r2
+)$a1
+
+## Data set with all centroids
+all_centroids_df2 <- hb_obj_one_c_shaped$centroids
+
+## Generate all coordinates of hexagons
+hex_grid2 <- hb_obj_one_c_shaped$hex_poly
+
+## To obtain the standardise counts within hexbins
+counts_df2 <- hb_obj_one_c_shaped$std_cts
+df_bin_centroids2 <- extract_hexbin_centroids(
+  centroids_df = all_centroids_df2,
+  counts_df = counts_df2) |>
+  filter(drop_empty == FALSE)
+
+tsne_data_one_c_shaped_with_hb_id <- hb_obj_one_c_shaped$data_hb_id
+df_all_one_curvy2 <- dplyr::bind_cols(one_c_shaped_data,
+                                      tsne_data_one_c_shaped_with_hb_id)
+
+df_bin_one_curvy2 <- avg_highd_data(data = df_all_one_curvy2, col_start = "x")
+
+## Compute error
+error_df_one_curvy_abs <- augment(
[…]

### 2025-01-07 paper.qmd
Updated Figure 12

@@
 ```{r}
-#| fig-height: 5
-#| fig-width: 15
+#| label: fig-one-dens_clust-error
+#| fig-height: 3
+#| fig-width: 9
 #| fig-pos: H
+#| fig-cap: "All \\gD{} layouts show high error in the sparse end.
+Three \\gD{} layouts: (a) tSNE, (b) UMAP, (c) PaCMaP of C-shaped densed cluster coloured by high-dimensional absolute error.
+Darker color shows high error while light colors shows low errors.
+Irrespective of the NLDR method, dense end of the structure shows the low error while sparse end of the structure shows the high error, because."
 generate_error_plots_one_c_shaped()

### 2025-01-07 paper.tex
Initial work on the effect of density
Updated Figure 12

@@
 
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-one-dens_clust-error-1.pdf}
+
+}
+
+\caption{\label{fig-one-dens_clust-error}All \gD{} layouts show high
+error in the sparse end. Three \gD{} layouts: (a) tSNE, (b) UMAP, (c)
+PaCMaP of C-shaped densed cluster coloured by high-dimensional absolute
+error. Darker color shows high error while light colors shows low
+errors. Irrespective of the NLDR method, dense end of the structure
+shows the low error while sparse end of the structure shows the high
+error, because.}
+
+\end{figure}%
+
 \section{Applications}\label{sec-applications}

### 2025-01-07 scripts/one_c_shaped_dens_structure/04_gen_error_plots.R

@@ -1,10 +1,19 @@
-### This script is to generate error plot for one_c_shaped_dens cluster with tSNE
+### This script is to generate error plot for one_c_shaped_dens cluster with tSNE, UMAP, and PaCMAP
 library(readr)
 library(dplyr)
+set.seed(20240110)
+
+# creating Standardization function
+standardize = function(x){
+  z <- (x - mean(x)) / sd(x)
+  return( z)
+}
 
 one_c_shaped_data <- read_rds(here::here("data/one_c_shaped_dens_structure/one_c_shaped_dens_data.rds"))
 one_c_shaped_data <- one_c_shaped_data |>
   mutate(ID = row_number())
+
+##1. With tSNE
 
 tsne_one_c_shaped <- read_rds(file = "data/one_c_shaped_dens_structure/one_c_shaped_dens_structure_tsne_perplexity_30.rds")
 
@@ -65,6 +74,9 @@ error_df_one_curvy_abs <- error_df_one_curvy_abs |>
   bind_cols(tsne_one_c_shaped_scaled |>
               select(-ID))
 
+error_df_one_curvy_abs <- error_df_one_curvy_abs |>
+  mutate(sqrt_row_wise_abs_error = standardize(sqrt_row_wise_abs_error))
+
 error_plot_tsne <- error_df_one_curvy_abs |>
   ggplot(aes(x = tSNE1,
              y = tSNE2,
@@ -76,14 +88,172 @@ error_plot_tsne <- error_df_one_curvy_abs |>
   ) +
   interior_annotation("a",
                       position = c(0.08, 0.95),
-                      cex = 2)
+                      cex = 1.5)
+
+##2. With UMAP
+
+umap_one_c_shaped <- read_rds(file = "data/one_c_shaped_dens_structure/one_c_shaped_dens_structure_umap_n-neigbors_15_min-dist_0.1.rds")
+
+umap_one_c_shaped_scaled_obj <- gen_scaled_data(
+  data = umap_one_c_shaped)
+
+umap_one_c_shaped_scaled <- umap_one_c_shaped_scaled_obj$scaled_nldr |>
+  mutate(ID = row_number())
+lim1 <- umap_one_c_shaped_scaled_obj$lim1
+lim2 <- umap_one_c_shaped_scaled_obj$lim2
+r2 <- diff(lim2)/diff(lim1)
+
+num_bins <- 23
+
[…]

### 2025-01-07 paper.qmd
Figure 12 fixed the title

--- 2025-01-06 paper.qmd
@@
 ```{r}
-#| fig-height: 5
-#| fig-width: 15
+#| label: fig-one-dens_clust-error
+#| fig-height: 3
+#| fig-width: 9
 #| fig-pos: H
+#| fig-cap: "All \\gD{} layouts exhibit high error in the sparse end.
+The three \\gD{} layouts are: (a) tSNE, (b) UMAP, and (c) PaCMaP, which represent a C-shaped dense cluster colored according to the high-dimensional absolute error.
+Darker colors indicate high error, while lighter colors indicate low error.
+Regardless of the NLDR method used, the dense end of the structure shows low error, whereas the sparse end displays high error."
+
 generate_error_plots_one_c_shaped()

### 2025-01-07 paper.tex
Figure 12 fixed the title

--- 2025-01-06 paper.tex
@@
 
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-one-dens_clust-error-1.pdf}
+
+}
+
+\caption{\label{fig-one-dens_clust-error}All \gD{} layouts exhibit high
+error in the sparse end. The three \gD{} layouts are: (a) tSNE, (b)
+UMAP, and (c) PaCMaP, which represent a C-shaped dense cluster colored
+according to the high-dimensional absolute error. Darker colors indicate
+high error, while lighter colors indicate low error. Regardless of the
+NLDR method used, the dense end of the structure shows low error,
+whereas the sparse end displays high error.}
+
+\end{figure}%
+
 \section{Applications}\label{sec-applications}

### 2025-01-07 paper.qmd
Initial writeup
Added chunck to set up script

@@
 
+```
+
+```{r}
+#| label: import-scripts
+source("scripts/one_c_shaped_dens_structure/04_gen_error_plots.R")
 ```
@@
 
-```{r}
-source("scripts/one_c_shaped_dens_structure/04_gen_error_plots.R")
-```
+To illustrate how the methods organise difference number of points within the structure in the NLDR layout, simulated $7\text{-}D$ data having C-shaped structure is used.
 

### 2025-01-07 paper.qmd
modifying error plots

--- 2025-01-07 paper.qmd
@@
 where $n$ is the number of observations, $b$ is the number of bins, $n_h$ is the number of observations in $h^{th}$ bin, $p$ is the number of variables, $\mathbfit{x}_{hij}$ is the $j^{th}$ dimensional data of $i^{th}$ observation in $h^{th}$ hexagon.
+We can consider $e_{hj} = \sum_{j = 1}^{p} (\mathbfit{x}_{hij} - C^{(p)}_{hj})^2$ to be the residual for each observation.
 
@@
 
+error_plot_two_curvy_hist <- 
+  ggplot(error_df_two_curvy_abs) +
+  geom_histogram(aes(x=sqrt_row_wise_abs_error, y=..density..)) +
+  geom_density(aes(x=sqrt_row_wise_abs_error, y=..density..), colour="red") +
+  xlab(expression(group("|", e[hj], "|"))) +
+  ylab("") +
+  theme_bw() 
 ```
@@
 Most large errors are distributed near the edges of the clusters."
-#| out-width: 30%
+#| out-width: 60%
 #| fig-pos: H
 
-error_plot_two_curvy
+error_plot_two_curvy + error_plot_two_curvy_hist +
+  plot_layout(ncol=2)
 ```
@@
 At the smallest bin size, the data structure is discontinuous, suggesting that there are too many bins.
-Using the MSE of the model fit in $7\\text{-}D$ helps decide on a useful choice of number of bins.
-XXX All look the same"
+Using the MSE of the model fit in $7\\text{-}D$ helps decide on a useful choice of number of bins."
 #| fig-width: 9

### 2025-01-07 paper.tex
Initial writeup
modifying error plots

@@
 
-\includegraphics[width=0.3\textwidth,height=\textheight]{paper_files/figure-pdf/fig-hex-param-1.pdf}
+\includegraphics[width=0.3\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-hex-param-1.pdf}
 
@@
 number of variables, \(\mathbfit{x}_{hij}\) is the \(j^{th}\)
-dimensional data of \(i^{th}\) observation in \(h^{th}\) hexagon.
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=0.3\textwidth,height=\textheight]{paper_files/figure-pdf/fig-p-d-error-in-2d-two-curvy-1.pdf}
+dimensional data of \(i^{th}\) observation in \(h^{th}\) hexagon. We can
+consider
+\(e_{hj} = \sum_{j = 1}^{p} (\mathbfit{x}_{hij} - C^{(p)}_{hj})^2\) to
+be the residual for each observation.
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=0.6\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-p-d-error-in-2d-two-curvy-1.pdf}
 
@@
 there are too many bins. Using the MSE of the model fit in
-\(7\text{-}D\) helps decide on a useful choice of number of bins. XXX
-All look the same}
+\(7\text{-}D\) helps decide on a useful choice of number of bins.}
 
@@
 
-\includegraphics[width=0.6\textwidth,height=\textheight]{paper_files/figure-pdf/fig-two_non_linear_diff_shaped_close_clusters-mse-1.pdf}
+\includegraphics[width=0.6\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-two_non_linear_diff_shaped_close_clusters-mse-1.pdf}
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-five-gau-projs-1.pdf}
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-five-gau-projs-1.pdf}
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-one-dens_clust-error-1.pdf}
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-one-dens_clust-error-1.pdf}
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-pbmc-mse-1.pdf}
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-pbmc-mse-1.pdf}
 
@@
 
-\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-model-pbmc-author-proj-1.pdf}
+\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-model-pbmc-author-proj-1.pdf}
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-pbmc-mse-umap-1.pdf}
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-pbmc-mse-umap-1.pdf}
 
@@
 
-\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-model-pbmc-author-proj-scdeed-1.pdf}
+\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-model-pbmc-author-proj-scdeed-1.pdf}
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-mnist-mse-1.pdf}
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-mnist-mse-1.pdf}
 
@@
 
-\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-model-error-mnist-1.pdf}
+\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-model-error-mnist-1.pdf}
 

### 2025-01-07 paper.qmd
Added script to generate C-shaped uniform and dense clusters, generate tSNE for both

@@
 source("scripts/one_c_shaped_dens_structure/04_gen_error_plots.R")
+source("scripts/five_gau_clusters/5_gen_model_with_tsne.R")
 ```
@@
 
+```{r}
+gen_proj_five_clust()[[1]] + gen_proj_five_clust()[[2]] + plot_layout(nrow = 2)
+```
+
 ### The effect of density

### 2025-01-07 paper.tex
Added script to generate C-shaped uniform and dense clusters, generate tSNE for both

--- 2025-01-07 paper.tex
@@
 number of variables, \(\mathbfit{x}_{hij}\) is the \(j^{th}\)
-dimensional data of \(i^{th}\) observation in \(h^{th}\) hexagon.
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=0.3\textwidth,height=\textheight]{paper_files/figure-pdf/fig-p-d-error-in-2d-two-curvy-1.pdf}
+dimensional data of \(i^{th}\) observation in \(h^{th}\) hexagon. We can
+consider
+\(e_{hj} = \sum_{j = 1}^{p} (\mathbfit{x}_{hij} - C^{(p)}_{hj})^2\) to
+be the residual for each observation.
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=0.6\textwidth,height=\textheight]{paper_files/figure-pdf/fig-p-d-error-in-2d-two-curvy-1.pdf}
 
@@
 
+\begin{center}
+\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-51-1.pdf}
+\end{center}
+
 \subsection{The effect of density}\label{the-effect-of-density}
+
+To illustrate how the methods organise difference number of points
+within the structure in the NLDR layout, simulated \(7\text{-}D\) data
+having C-shaped structure is used.
 

### 2025-01-07 scripts/five_gau_clusters/5_gen_model_with_tsne.R

@@ -3,7 +3,6 @@ library(readr)
 library(dplyr)
 
 training_data_gau <- read_rds("data/five_gau_clusters/data_five_gau.rds")
-training_data_gau_with_label <- read_rds("data/five_gau_clusters/data_five_gau_with_labels.rds")
 
 data_gau <- training_data_gau |>
   select(-ID) |>
@@ -14,25 +13,10 @@ gau1_scaled_obj <- gen_scaled_data(
   data = tsne_data_gau)
 tsne_gau_scaled <- gau1_scaled_obj$scaled_nldr
 
-## To find the sepcific cluster
-df <- bind_cols(training_data_gau_with_label,
-                tsne_gau_scaled)
-
-df |>
-  ggplot(aes(x = tSNE1,
-             y = tSNE2,
-             color = cluster)) +
-  geom_point(alpha=0.3) +
-  geom_rect(aes(xmin = -0.05, xmax = 0.35, ymin = 0.28, ymax = 0.6),
-            fill = "transparent", color = "grey50", linewidth = 0.8)
-### Selected cluster 2
-
 tsne_gau <- tsne_gau_scaled |>
   ggplot(aes(x = tSNE1,
              y = tSNE2)) +
-  geom_point(alpha=0.3) +
-  geom_rect(aes(xmin = -0.05, xmax = 0.35, ymin = 0.28, ymax = 0.6),
-            fill = "transparent", color = "grey50", linewidth = 0.8)
+  geom_point(alpha=0.3)
 
 ## Compute hexbin parameters
 num_bins_x_gau1 <- 20 #13
@@ -130,14 +114,6 @@ tr_from_to_df_gau1 <- inner_join(
   by = c("from", "to"))
 
 trimesh_removed_gau_tsne <- ggplot() +
-  #geom_segment(data = tr_from_to_df_gau1,
-  #             aes(
-  #               x = x_from,
-  #               y = y_from,
-  #               xend = x_to,
-  #               yend = y_to),
-  #             colour = "#000000",
-  #             linewidth = 1) +
   geom_point(
     data = tsne_gau_scaled,
     aes(
@@ -203,15 +179,6 @@ scaled_gau_data_model <- scaled_gau |>
 distance_df_small_edges <- distance_gau1 |>
   dplyr::filter(distance < benchmark_gau1)
 
-# library(langevitour)
-
[…]

### 2025-01-07 paper.qmd
Updated Figure 12

@@
 #| fig-height: 3
-#| fig-width: 9
+#| fig-width: 6
 #| fig-pos: H

### 2025-01-07 paper.qmd
Updated Figure 10

@@
 
+tsne_two_non_linear_diff_shaped_close_clusters <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_62.rds")
+
+nldr_two_non_linear_diff_shaped_close_clusters6 <- tsne_two_non_linear_diff_shaped_close_clusters |>
+  ggplot(aes(x = tSNE1,
+             y = tSNE2))+
+  geom_point(alpha=0.1, size=1, colour='#984ea3') +
+  interior_annotation("f", c(0.08, 0.93)) +
+  theme(aspect.ratio = 1)
+
 umap_two_non_linear_diff_shaped_close_clusters <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_umap_n-neigbors_15_min-dist_0.1.rds")
@@
 error_two_non_linear_diff_shaped_close_clusters_pacmap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_pacmap.rds")
+error_two_non_linear_diff_shaped_close_clusters_tsne2 <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_tsne2.rds")
+
+error_two_non_linear_diff_shaped_close_clusters_tsne2 <- error_two_non_linear_diff_shaped_close_clusters_tsne2 |>
+  mutate(method = "tSNE2")
 
@@
                          error_two_non_linear_diff_shaped_close_clusters_trimap,
-                         error_two_non_linear_diff_shaped_close_clusters_pacmap)
+                         error_two_non_linear_diff_shaped_close_clusters_pacmap,
+                         error_two_non_linear_diff_shaped_close_clusters_tsne2)
 
@@
   scale_x_continuous(breaks = sort(unique(error_two_non_linear_diff_shaped_close_clusters$a1))[seq(1, length(unique(error_two_non_linear_diff_shaped_close_clusters$a1)), by = 5)]) +
-  scale_color_manual(values=c('#e41a1c','#ff7f00','#4daf4a', "#a65628",'#636363')) +
+  scale_color_manual(values=c('#e41a1c','#ff7f00','#4daf4a', "#a65628",'#636363', '#984ea3')) +
   scale_y_log10() +
@@
 free(error_plot_two_non_linear_diff_shaped_close_clusters) + wrap_plots(nldr_two_non_linear_diff_shaped_close_clusters1, nldr_two_non_linear_diff_shaped_close_clusters2, nldr_two_non_linear_diff_shaped_close_clusters3,
-                                    nldr_two_non_linear_diff_shaped_close_clusters4, nldr_two_non_linear_diff_shaped_close_clusters5, ncol = 2)
+                                    nldr_two_non_linear_diff_shaped_close_clusters4, nldr_two_non_linear_diff_shaped_close_clusters5,
+                                    nldr_two_non_linear_diff_shaped_close_clusters6, ncol = 2)
 ```

### 2025-01-07 scripts/two_non_linear_diff_shaped_close_clusters/05_validate_nldr_layouts.R

@@ -3,6 +3,7 @@ library(readr)
 library(ggplot2)
 
 tsne_df <- read_rds(here::here("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_30.rds"))
+tsne_df2 <- read_rds(here::here("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_62.rds"))
 umap_df <- read_rds(here::here("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_umap_n-neigbors_15_min-dist_0.1.rds"))
 phate_df <- read_rds(here::here("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_phate_knn_5.rds"))
 trimap_df <- read_rds(here::here("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_trimap_n-inliers_12_n-outliers_4_n-random_3.rds"))
@@ -12,6 +13,10 @@ pacmap_df <- read_rds(here::here("data/two_non_linear_diff_shaped_close_clusters
 tsne_df <- tsne_df |>
   mutate(method = "tSNE")
 names(tsne_df) <- c("emb_1", "emb_2", "method")
+
+tsne_df2 <- tsne_df2 |>
+  mutate(method = "tSNE2")
+names(tsne_df2) <- c("emb_1", "emb_2", "method")
 
 umap_df <- umap_df |>
   mutate(method = "UMAP")
@@ -30,6 +35,7 @@ pacmap_df <- pacmap_df |>
 names(pacmap_df) <- c("emb_1", "emb_2", "method")
 
 nldr_df <- bind_rows(tsne_df,
+                     tsne_df2,
                      umap_df,
                      phate_df,
                      pacmap_df,
@@ -44,5 +50,5 @@ ggplot(data = nldr_df,
   geom_point(
     alpha = 0.5
   ) +
-  facet_wrap(~factor(method, levels = c("tSNE", "UMAP", "PHATE", "TriMAP", "PaCMAP")),
+  facet_wrap(~factor(method, levels = c("tSNE", "tSNE2", "UMAP", "PHATE", "TriMAP", "PaCMAP")),
              scales = "free")

### 2025-01-07 paper.qmd
Updated error computation of Figure 6

@@
 error_df_two_curvy_abs <- error_df_two_curvy_abs |>
-  mutate(sqrt_row_wise_abs_error = sqrt(row_wise_abs_error))
+  mutate(sqrt_row_wise_total_error = sqrt(row_wise_total_error))
 
@@
              y = tSNE2,
-             colour = sqrt_row_wise_abs_error)) +
+             colour = sqrt_row_wise_total_error)) +
   geom_point(alpha=0.5) +
@@
   ggplot(error_df_two_curvy_abs) +
-  geom_histogram(aes(x=sqrt_row_wise_abs_error, y=..density..)) +
-  geom_density(aes(x=sqrt_row_wise_abs_error, y=..density..), colour="red") +
+  geom_histogram(aes(x=sqrt_row_wise_total_error, y=..density..)) +
+  geom_density(aes(x=sqrt_row_wise_total_error, y=..density..), colour="red") +
   xlab(expression(group("|", e[hj], "|"))) +
@@
 #| fig-cap: "The $7\\text{-}D$ model error in \\gD{} layout.
-Color indicates square root of absolute error, dark blue indicating high error and light indicates low error.
+Color indicates square root of total error, dark blue indicating high error and light indicates low error.
 Most large errors are distributed near the edges of the clusters."

### 2025-01-07 paper.qmd
Updated Figure 11

@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 20 #13
+num_bins_x_gau1 <- 35
 lim1 <- gau1_scaled_obj$lim1
@@
 
-benchmark_gau1 <- 0.1
+# benchmark_gau1 <- 0.1
 
@@
 distance_df_small_edges_gau1 <- distance_gau1 |>
-  filter(distance < benchmark_gau1) |>
-  filter(!(row_number() %in% c(33, 40, 48, 138, 143)))
+  filter(distance < benchmark_gau1)
 
@@
 trimesh_removed_gau_tsne <- ggplot() + 
-  #geom_segment(data = tr_from_to_df_gau1, 
-  #             aes(
-  #               x = x_from, 
-  #               y = y_from, 
-  #               xend = x_to, 
-  #               yend = y_to),
-  #             colour = "#000000",
-  #             linewidth = 1) +
   geom_point(
@@
     alpha = 0.2) + 
-  geom_rect(
-      aes(xmin = 0.3, 
-          xmax = 0.65, 
-          ymin = -0.05, 
-          ymax = 0.35),
-          fill = "transparent", 
-      color = "grey50", 
-      linewidth = 0.5) +
+  # geom_rect(
+  #     aes(xmin = 0.3, 
+  #         xmax = 0.65, 
+  #         ymin = -0.05, 
+  #         ymax = 0.35),
+  #         fill = "transparent", 
+  #     color = "grey50", 
+  #     linewidth = 0.5) +
   interior_annotation("a1", 
@@
   proj = projection,
-  limits = 0.3,
-  axis_pos_x = 0.05,
-  axis_pos_y = -0.85,
+  limits = 1,
+  axis_pos_x = -0.6,
+  axis_pos_y = -0.63,
   axis_labels = names(scaled_gau_data),
@@
       y = proj2)) +
+  geom_point(
+    #size = 0.5,
+    alpha = 0.3,
+    color = "#999999") +
   geom_segment(
@@
     color = "#000000",
-    linewidth = 1) +
-  geom_point(
-    #size = 0.5,
-    alpha = 0.2,
-    color = "#999999") +
+    alpha = 0.4,
+    linewidth = 0.1) +
   geom_segment(
@@
   coord_fixed() +
-  xlim(c(0, 0.4)) +
-  ylim(c(-0.9, -0.5)) +
+  xlim(c(-0.84, 0.84)) +
+  ylim(c(-0.84, 0.84)) +
   interior_annotation("a2", 
@@
                       cex = 2)
-
-## Second projection
-# projection <- cbind(
-#     c(-0.71895,0.29250,-0.29905,-0.01656),
-#     c(-0.16698,-0.62707,-0.23756,0.46328))
-projection <- cbind(
-  c(0.36030,0.00630,0.66545,-0.34307),
-  c(-0.01861,0.14409,-0.36798,-0.73066))
-
-projected <- as.matrix(scaled_gau_data) %*% projection
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  #dplyr::mutate(type = df_exe$type) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-projected_model <- as.matrix(scaled_gau_data_model) %*% projection
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  distance_df_small_edges |> select(-distance), 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-axes_obj <- gen_axes(  
-  proj = projection,
-  limits = 0.3,
-  axis_pos_x = -0.7,
-  axis_pos_y = -0.7,
-  axis_labels = names(scaled_gau_data),
-  threshold = 0)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
-
-five_gau_proj_tsne_model2 <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#000000",
-    linewidth = 1) +
-  geom_point(
-    #size = 0.5,
-    alpha = 0.2,
-    color = "#999999") +
-  geom_segment(
-    data=axes, 
-    aes(x=x1, y=y1, xend=x2, yend=y2), 
-    colour="grey70") +
-  geom_text(
-    data=axes, 
-    aes(x=x2, y=y2), 
-    label=rownames(axes), 
-    colour="grey50",
-    size = 5) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.75, -0.36)) +
-  ylim(c(-0.75, -0.36)) +
-  interior_annotation("a3", 
-                      position = c(0.08, 0.9),
-                      cex = 2)
-
-# five_gau_proj_tsne_model <- gen_proj_langevitour(
-#   points_df = df_exe,
-#   projection = projection,
-#   edge_df = distance_df_small_edges |> select(-distance)
-# ) +
-#   xlim(c(-0.18, 0.1)) +
-#   ylim(c(-0.78, -0.55))
 ```
@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 45
+num_bins_x_gau1 <- 127
 lim1 <- gau1_scaled_obj$lim1
@@
   filter(distance < benchmark_gau1) |>
-  mutate(ID = row_number()) |>
-  filter(!(ID %in% c(112)))
+  mutate(ID = row_number()) #|>
+  #filter(!(ID %in% c(112)))
 
@@
 trimesh_removed_gau_umap <- ggplot() + 
-  #geom_segment(data = tr_from_to_df_gau1, 
-  #             aes(
-  #               x = x_from, 
-  #               y = y_from, 
-  #               xend = x_to, 
-  #               yend = y_to),
-  #             colour = "#000000",
-  #             linewidth = 1) +
   geom_point(
@@
     alpha = 0.05) + 
-  geom_rect(
-    aes(xmin = 0.45, 
-        xmax = 0.57, 
-        ymin = 0, 
-        ymax = 0.1),
-    fill = "transparent", 
-    color = "grey50", 
-    linewidth = 0.5)  +
+  # geom_rect(
+  #   aes(xmin = 0.45, 
+  #       xmax = 0.57, 
+  #       ymin = 0, 
+  #       ymax = 0.1),
+  #   fill = "transparent", 
+  #   color = "grey50", 
+  #   linewidth = 0.5)  +
   interior_annotation("b1", 
@@
   proj = projection,
-  limits = 0.28,
-  axis_pos_x = -0.58,
-  axis_pos_y = -0.35,
+  limits = 1,
+  axis_pos_x = -0.6,
+  axis_pos_y = -0.63,
   axis_labels = names(scaled_gau_data),
@@
     color = "#000000",
-    linewidth = 1) +
-  geom_point(
-    #size = 0.5,
-    alpha = 0.3,
-    color = "#999999") +
+    alpha = 0.4,
+    linewidth = 0.1) +
   geom_segment(
@@
   coord_fixed() +
-  xlim(c(-0.64, -0.29)) +
-  ylim(c(-0.4, -0.05)) +
+  xlim(c(-0.83, 0.83)) +
+  ylim(c(-0.83, 0.83)) +
   interior_annotation("b2", 
@@
                       cex = 2)
-
-## Second projection
+```
+
+```{r}
+#| label: five-gau-proj-pacmap-model
+
+pacmap_data_gau <- read_rds("data/five_gau_clusters/pacmap_data_five_gau.rds")
+gau1_scaled_obj <- gen_scaled_data(
+  data = pacmap_data_gau)
+pacmap_gau_scaled <- gau1_scaled_obj$scaled_nldr |>
+  select(PaCMAP1, PaCMAP2, ID)
+
+pacmap_gau <- pacmap_gau_scaled |>
+  ggplot(aes(x = PaCMAP1,
+             y = PaCMAP2)) +
+  geom_point(alpha=0.3, color = "#000000")
+
+## Compute hexbin parameters
+num_bins_x_gau1 <- 56
+lim1 <- gau1_scaled_obj$lim1
+lim2 <- gau1_scaled_obj$lim2
+r2_gau1 <- diff(lim2)/diff(lim1)
+
+gau1_model <- fit_highd_model(
+  training_data = training_data_gau,
+  emb_df = pacmap_gau_scaled,
+  bin1 = num_bins_x_gau1,
+  r2 = r2_gau1,
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = FALSE,
+  col_start_highd = "x",
+  q = 0.1
+)
+
+df_bin_centroids_gau1 <- gau1_model$df_bin_centroids
+df_bin_gau1 <- gau1_model$df_bin
+
+## Compute error
+
+error_df_gau_abs <- augment(
+  df_bin_centroids = df_bin_centroids_gau1,
+  df_bin = df_bin_gau1,
+  training_data = training_data_gau,
+  newdata = NULL,
+  type_NLDR = "PaCMAP",
+  col_start = "x")
+
+error_df_gau_abs <- error_df_gau_abs |>
+  mutate(sqrt_row_wise_abs_error = sqrt(row_wise_abs_error))
+
+error_df_gau_abs <- error_df_gau_abs |>
+  bind_cols(pacmap_gau_scaled |>
+              select(-ID))
+
+error_df_gau_pacmap <- error_df_gau_abs |>
+  ggplot(aes(x = PaCMAP1,
+             y = PaCMAP2,
+             z = sqrt_row_wise_abs_error)) +
+  stat_summary_hex(bins=15) + 
+  xlim(c(-0.02, 0.18)) + ylim(c(0.12, 0.32)) +
+  scale_fill_continuous_sequential(palette = "YlOrRd", breaks = 20) +
+  interior_annotation("c3", 
+                      position = c(0.08, 0.9),
+                      cex = 2)
+
+## Triangulate bin centroids
+tr1_object_gau1 <- tri_bin_centroids(
+  df_bin_centroids_gau1, x = "c_x", y = "c_y")
+tr_from_to_df_gau1 <- gen_edges(
+  tri_object = tr1_object_gau1)
+
+## Compute 2D distances
+distance_gau1 <- cal_2d_dist(
+  tr_coord_df = tr_from_to_df_gau1,
+  start_x = "x_from",
+  start_y = "y_from",
+  end_x = "x_to",
+  end_y = "y_to",
+  select_vars = c("from", "to", "distance"))
+
+## To find the benchmark value
+benchmark_gau1 <- find_lg_benchmark(
+  distance_edges = distance_gau1,
+  distance_col = "distance")
+
+# trimesh_removed_gau1 <- vis_rmlg_mesh(
+#   distance_edges = distance_gau1,
+#   benchmark_value = benchmark_gau1,
+#   tr_coord_df = tr_from_to_df_gau1,
+#   distance_col = "distance")
+
+tr_df <- distinct(tibble(
+  x = c(tr_from_to_df_gau1[["x_from"]], tr_from_to_df_gau1[["x_to"]]), 
+  y = c(tr_from_to_df_gau1[["y_from"]], tr_from_to_df_gau1[["y_to"]])))
+
+distance_df_small_edges_gau1 <- distance_gau1 |>
+  filter(distance < benchmark_gau1) |>
+  mutate(ID = row_number()) |>
+  filter(!(ID %in% c(97)))
+
+tr_from_to_df_gau1 <- inner_join(
+  tr_from_to_df_gau1, distance_df_small_edges_gau1, 
+  by = c("from", "to"))
+
+trimesh_removed_gau_pacmap <- ggplot() + 
+  geom_point(
+    data = pacmap_gau_scaled,
+    aes(
+      x = PaCMAP1,
+      y = PaCMAP2
+    ),
+    alpha = 0.3) +
+  # geom_rect(
+  #   aes(xmin = -0.04, 
+  #       xmax = 0.22, 
+  #       ymin = 0.08, 
+  #       ymax = 0.34),
+  #   fill = "transparent", 
+  #   color = "grey50", 
+  #   linewidth = 0.5) +
+  interior_annotation("c1", 
+                      position = c(0.08, 0.95),
+                      cex = 2) +
+  theme(
+    aspect.ratio = 1
+  )
+
+## Hexagonal binning to have regular hexagons
+hb_obj_gau1 <- hex_binning(
+  data = pacmap_gau_scaled,
+  bin1 = num_bins_x_gau1,
+  r2 = r2_gau1,
+  q = 0.1)
+
+pacmap_data_with_hb_id <- hb_obj_gau1$data_hb_id
+
+df_all_gau1 <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID),
+                                pacmap_data_with_hb_id)
+
+### Define type column
+df <- df_all_gau1 |>
+  dplyr::select(tidyselect::starts_with("x")) |>
+  dplyr::mutate(type = "data") ## original dataset
+
+df_b <- df_bin_gau1 |>
+  dplyr::filter(hb_id %in% df_bin_centroids_gau1$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b <- df_b[match(df_bin_centroids_gau1$hexID, df_b$hb_id),] |>
+  dplyr::select(-hb_id)
+
+# Apply the scaling
+df_model_data <- bind_rows(data_gau, df_b)
+scaled_gau <- scale_data_manual(df_model_data, "type") |>
+  as_tibble()
+
+scaled_gau_data <- scaled_gau |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_gau_data_model <- scaled_gau |>
+  filter(type == "model") |>
+  select(-type)
+
+## Set the maximum difference as the criteria
+distance_df_small_edges <- distance_gau1 |>
+  dplyr::filter(distance < benchmark_gau1)
+
+## First projection
 projection <- cbind(
-  c(0.36030,0.00630,0.66545,-0.34307),
-  c(-0.01861,0.14409,-0.36798,-0.73066))
+  c(-0.00215,-0.68905,-0.04778,-0.54223),
+  c(0.42558,-0.23854,-0.63659,0.35753))
 
@@
   proj = projection,
-  limits = 0.28,
-  axis_pos_x = -0.7,
-  axis_pos_y = -0.7,
+  limits = 1,
+  axis_pos_x = -0.6,
+  axis_pos_y = -0.63,
   axis_labels = names(scaled_gau_data),
@@
 
-five_gau_proj_umap_model2 <- projected_df |>
+five_gau_proj_pacmap_model1 <- projected_df |>
   ggplot(
@@
   coord_fixed() +
-  xlim(c(-0.75, -0.4)) +
-  ylim(c(-0.75, -0.4)) +
-  interior_annotation("b3", 
-                      position = c(0.08, 0.9),
-                      cex = 2)
-
-# five_gau_proj_umap_model <- gen_proj_langevitour(
-#   points_df = df_exe,
-#   projection = projection,
-#   edge_df = distance_df_small_edges |> select(-distance)
-# ) +
-#   xlim(c(-0.18, 0.1)) +
-#   ylim(c(-0.78, -0.55))
-```
-
-```{r}
-#| label: five-gau-proj-pacmap-model
-
-pacmap_data_gau <- read_rds("data/five_gau_clusters/pacmap_data_five_gau.rds")
-gau1_scaled_obj <- gen_scaled_data(
-  data = pacmap_data_gau)
-pacmap_gau_scaled <- gau1_scaled_obj$scaled_nldr |>
-  select(PaCMAP1, PaCMAP2, ID)
-
-pacmap_gau <- pacmap_gau_scaled |>
-  ggplot(aes(x = PaCMAP1,
-             y = PaCMAP2)) +
-  geom_point(alpha=0.3, color = "#000000")
-
-## Compute hexbin parameters
-num_bins_x_gau1 <- 21
-lim1 <- gau1_scaled_obj$lim1
-lim2 <- gau1_scaled_obj$lim2
-r2_gau1 <- diff(lim2)/diff(lim1)
-
-gau1_model <- fit_highd_model(
-  training_data = training_data_gau,
-  emb_df = pacmap_gau_scaled,
-  bin1 = num_bins_x_gau1,
-  r2 = r2_gau1,
-  is_bin_centroid = TRUE,
-  is_rm_lwd_hex = FALSE,
-  col_start_highd = "x",
-  q = 0.1
-)
-
-df_bin_centroids_gau1 <- gau1_model$df_bin_centroids
-df_bin_gau1 <- gau1_model$df_bin
-
-## Compute error
-
-error_df_gau_abs <- augment(
-  df_bin_centroids = df_bin_centroids_gau1,
-  df_bin = df_bin_gau1,
-  training_data = training_data_gau,
-  newdata = NULL,
-  type_NLDR = "PaCMAP",
-  col_start = "x")
-
-error_df_gau_abs <- error_df_gau_abs |>
-  mutate(sqrt_row_wise_abs_error = sqrt(row_wise_abs_error))
-
-error_df_gau_abs <- error_df_gau_abs |>
-  bind_cols(pacmap_gau_scaled |>
-              select(-ID))
-
-error_df_gau_pacmap <- error_df_gau_abs |>
-  ggplot(aes(x = PaCMAP1,
-             y = PaCMAP2,
-             z = sqrt_row_wise_abs_error)) +
-  stat_summary_hex(bins=15) + 
-  xlim(c(-0.02, 0.18)) + ylim(c(0.12, 0.32)) +
-  scale_fill_continuous_sequential(palette = "YlOrRd", breaks = 20) +
-  interior_annotation("c3", 
-                      position = c(0.08, 0.9),
-                      cex = 2)
-
-## Triangulate bin centroids
-tr1_object_gau1 <- tri_bin_centroids(
-  df_bin_centroids_gau1, x = "c_x", y = "c_y")
-tr_from_to_df_gau1 <- gen_edges(
-  tri_object = tr1_object_gau1)
-
-## Compute 2D distances
-distance_gau1 <- cal_2d_dist(
-  tr_coord_df = tr_from_to_df_gau1,
-  start_x = "x_from",
-  start_y = "y_from",
-  end_x = "x_to",
-  end_y = "y_to",
-  select_vars = c("from", "to", "distance"))
-
-## To find the benchmark value
-benchmark_gau1 <- find_lg_benchmark(
-  distance_edges = distance_gau1,
-  distance_col = "distance")
-
-# trimesh_removed_gau1 <- vis_rmlg_mesh(
-#   distance_edges = distance_gau1,
-#   benchmark_value = benchmark_gau1,
-#   tr_coord_df = tr_from_to_df_gau1,
-#   distance_col = "distance")
-
-tr_df <- distinct(tibble(
-  x = c(tr_from_to_df_gau1[["x_from"]], tr_from_to_df_gau1[["x_to"]]), 
-  y = c(tr_from_to_df_gau1[["y_from"]], tr_from_to_df_gau1[["y_to"]])))
-
-distance_df_small_edges_gau1 <- distance_gau1 |>
-  filter(distance < benchmark_gau1) |>
-  mutate(ID = row_number()) |>
-  filter(!(ID %in% c(97)))
-
-tr_from_to_df_gau1 <- inner_join(
-  tr_from_to_df_gau1, distance_df_small_edges_gau1, 
-  by = c("from", "to"))
-
-trimesh_removed_gau_pacmap <- ggplot() + 
-  #geom_segment(data = tr_from_to_df_gau1, 
-  #             aes(
-  #               x = x_from, 
-  #               y = y_from, 
-  #               xend = x_to, 
-  #               yend = y_to),
-  #             colour = "#000000",
-  #             linewidth = 1) +
-  geom_point(
-    data = pacmap_gau_scaled,
-    aes(
-      x = PaCMAP1,
-      y = PaCMAP2
-    ),
-    alpha = 0.3) +
-  geom_rect(
-    aes(xmin = -0.04, 
-        xmax = 0.22, 
-        ymin = 0.08, 
-        ymax = 0.34),
-    fill = "transparent", 
-    color = "grey50", 
-    linewidth = 0.5) +
-  interior_annotation("c1", 
-                      position = c(0.08, 0.95),
-                      cex = 2) +
-  theme(
-    aspect.ratio = 1
-  )
-
-## Hexagonal binning to have regular hexagons
-hb_obj_gau1 <- hex_binning(
-  data = pacmap_gau_scaled,
-  bin1 = num_bins_x_gau1,
-  r2 = r2_gau1,
-  q = 0.1)
-
-pacmap_data_with_hb_id <- hb_obj_gau1$data_hb_id
-
-df_all_gau1 <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID),
-                                pacmap_data_with_hb_id)
-
-### Define type column
-df <- df_all_gau1 |>
-  dplyr::select(tidyselect::starts_with("x")) |>
-  dplyr::mutate(type = "data") ## original dataset
-
-df_b <- df_bin_gau1 |>
-  dplyr::filter(hb_id %in% df_bin_centroids_gau1$hexID) |>
-  dplyr::mutate(type = "model") ## Data with summarized mean
-
-## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
-df_b <- df_b[match(df_bin_centroids_gau1$hexID, df_b$hb_id),] |>
-  dplyr::select(-hb_id)
-
-# Apply the scaling
-df_model_data <- bind_rows(data_gau, df_b)
-scaled_gau <- scale_data_manual(df_model_data, "type") |>
-  as_tibble()
-
-scaled_gau_data <- scaled_gau |>
-  filter(type == "data") |>
-  select(-type)
-
-scaled_gau_data_model <- scaled_gau |>
-  filter(type == "model") |>
-  select(-type)
-
-## Set the maximum difference as the criteria
-distance_df_small_edges <- distance_gau1 |>
-  dplyr::filter(distance < benchmark_gau1)
-
-## First projection
-projection <- cbind(
-  c(-0.00215,-0.68905,-0.04778,-0.54223),
-  c(0.42558,-0.23854,-0.63659,0.35753))
-
-projected <- as.matrix(scaled_gau_data) %*% projection
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  #dplyr::mutate(type = df_exe$type) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-projected_model <- as.matrix(scaled_gau_data_model) %*% projection
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  distance_df_small_edges |> select(-distance), 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-axes_obj <- gen_axes(  
-  proj = projection,
-  limits = 0.25,
-  axis_pos_x = -0.41,
-  axis_pos_y = 0.23,
-  axis_labels = names(scaled_gau_data),
-  threshold = 0)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
-
-five_gau_proj_pacmap_model1 <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#000000",
-    linewidth = 1) +
-  geom_point(
-    #size = 0.5,
-    alpha = 0.3,
-    color = "#999999") +
-  geom_segment(
-    data=axes, 
-    aes(x=x1, y=y1, xend=x2, yend=y2), 
-    colour="grey70") +
-  geom_text(
-    data=axes, 
-    aes(x=x2, y=y2, label=rownames(axes)), 
-    colour="grey50",
-    size = 5) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.46, -0.12)) +
-  ylim(c(0.18, 0.52)) +
+  xlim(c(-0.83, 0.83)) +
+  ylim(c(-0.83, 0.83)) +
   interior_annotation("c2",
-                      position = c(0.08, 0.9),
-                      cex = 2)
-
-### Second projection
-# projection <- cbind(
-#     c(-0.60995,0.33361,0.32920,-0.31518),
-#     c(0.19525,-0.30554,0.74417,0.07600))
-projection <- cbind(
-  c(0.36030,0.00630,0.66545,-0.34307),
-  c(-0.01861,0.14409,-0.36798,-0.73066))
-
-projected <- as.matrix(scaled_gau_data) %*% projection
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  #dplyr::mutate(type = df_exe$type) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-projected_model <- as.matrix(scaled_gau_data_model) %*% projection
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  distance_df_small_edges |> select(-distance), 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-axes_obj <- gen_axes(  
-  proj = projection,
-  limits = 0.3,
-  axis_pos_x = -0.7,
-  axis_pos_y = -0.7,
-  axis_labels = names(scaled_gau_data),
-  threshold = 0)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
-
-five_gau_proj_pacmap_model2 <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#000000",
-    linewidth = 1) +
-  geom_point(
-    #size = 0.8,
-    alpha = 0.2,
-    color = "#999999") +
-  geom_segment(
-    data=axes, 
-    aes(x=x1, y=y1, xend=x2, yend=y2), 
-    colour="grey70") +
-  geom_text(
-    data=axes, 
-    aes(x=x2, y=y2, label=rownames(axes)), 
-    colour="grey50",
-    size = 5) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.75, -0.35)) +
-  ylim(c(-0.75, -0.35))+
-  interior_annotation("c3", 
                       position = c(0.08, 0.9),
@@
 #| fig-pos: H
-#| fig-cap: "NLDR's organise points in the \\gD{} layout in different ways, possibly misleadingly, illustrated using three layouts: (a) t-SNE, (b) UMAP, (c) PaCMap.
+#| fig-cap: "NLDR's organise points in the \\gD{} layout in different ways, possibly misleadingly, illustrated using three layouts: (a) tSNE, (b) UMAP, (c) PaCMAP.
 The data has five Gaussian cluster in $4\\text{-}D$.
@@
 ```{r}
-gen_proj_five_clust()[[1]] + gen_proj_five_clust()[[2]] + plot_layout(nrow = 2)
+# gen_proj_five_clust()[[1]] + gen_proj_five_clust()[[2]] + plot_layout(nrow = 2)
 ```

### 2025-01-07 paper.tex
Updated error computation of Figure 6
Updated Figure 11

@@
 \caption{\label{fig-p-d-error-in-2d-two-curvy}The \(7\text{-}D\) model
-error in \gD{} layout. Color indicates square root of absolute error,
-dark blue indicating high error and light indicates low error. Most
-large errors are distributed near the edges of the clusters.}
+error in \gD{} layout. Color indicates square root of total error, dark
+blue indicating high error and light indicates low error. Most large
+errors are distributed near the edges of the clusters.}
 
@@
 layout in different ways, possibly misleadingly, illustrated using three
-layouts: (a) t-SNE, (b) UMAP, (c) PaCMap. The data has five Gaussian
+layouts: (a) tSNE, (b) UMAP, (c) PaCMAP. The data has five Gaussian
 cluster in \(4\text{-}D\). The bottom row of plots shows a \gD{}
@@
 \end{figure}%
-
-\begin{center}
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-51-1.pdf}
-\end{center}
 

### 2025-01-08 paper.qmd
Updated Figure 11

@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 35
+num_bins_x_gau1 <- 26
 lim1 <- gau1_scaled_obj$lim1
@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 127
+num_bins_x_gau1 <- 90
 lim1 <- gau1_scaled_obj$lim1
@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 56
+num_bins_x_gau1 <- 40
 lim1 <- gau1_scaled_obj$lim1

### 2025-01-08 paper.qmd
Initial work on adding projections for Figure12

@@
 #| label: fig-one-dens_clust-error
-#| fig-height: 3
-#| fig-width: 6
+#| fig-height: 8
+#| fig-width: 8
 #| fig-pos: H

### 2025-01-08 paper.qmd
Updated error computation

@@
 #| label: fig-one-dens_clust-error
-#| fig-height: 8
-#| fig-width: 8
+#| fig-height: 10
+#| fig-width: 10
 #| fig-pos: H

### 2025-01-09 paper.qmd
Updated Figure 7

@@
     na.value = "#ffffff", option = "E") +
-  interior_annotation("a", sc_ltr_pos) + 
+  interior_annotation("b", sc_ltr_pos) + 
   xlim(c(-0.3, 1.2)) + ylim(c(-0.17, 1.1))
@@
   xlim(c(-0.3, 1.2)) + ylim(c(-0.17, 1.1)) +
-  interior_annotation("b", sc_ltr_pos)
+  interior_annotation("a", sc_ltr_pos)
 

### 2025-01-09 paper.qmd
Tried add hist..need to change

@@
   filter(drop_empty == FALSE)
+
+boxplot2 <- ggplot(df_bin_centroids_two_curvy, 
+       aes(x=bin_counts)) + 
+  geom_histogram()
 
@@
 
+boxplot1 <- ggplot(df_bin_centroids2, 
+                   aes(x=bin_counts)) + 
+    geom_histogram()
+
 hex_grid_with_counts_two_curvy2 <- full_join(
@@
 
+boxplot3 <- ggplot(df_bin_centroids3, 
+                   aes(x=bin_counts)) + 
+    geom_histogram()
+
 hex_grid_with_counts_s_curve3 <- full_join(hex_grid3, 
@@
   hex_grid_coloured_two_curvy3 +
+  boxplot1 + boxplot2 + boxplot3 +
   plot_layout(guides='collect', ncol = 3) &

### 2025-01-10 paper.qmd
Updated Figure 9 (a)

@@
 ```{r}
+#| label: dist-plot
+
+df_bin_two_curvy_temp <- df_bin_two_curvy
+
+### Compute highD distance
+dist_vec <- proxy::dist(x = df_bin_two_curvy_temp[, -1], method = "Euclidean") |> as.vector()
+
+from_vec <- c()
+to_vec <- c()
+num_obs <- 1:(NROW(df_bin_two_curvy_temp) - 1)
+
+for (obs in num_obs) {
+
+  from_val <- rep(obs, (NROW(df_bin_two_curvy_temp) - obs))
+  if ((obs + 1) <= NROW(df_bin_two_curvy_temp)) {
+    to_val <- (obs + 1):NROW(df_bin_two_curvy_temp)
+  }
+  from_vec <- append(from_vec, from_val)
+  to_vec <- append(to_vec, to_val)
+
+}
+
+dist_highd <- tibble::tibble(from = from_vec, to = to_vec, dist_highd = dist_vec)
+
+## To plot the distribution of distances
+
+# Define bin width and starting point
+bin_width <- hb_obj_two_curvy$a1
+start_point <- bin_width/2
+
+# Create bins and calculate the mean value for each bin range
+bins <- seq(start_point, max(distance_two_curvy$distance) + bin_width, 
+            by = bin_width)
+bin_labels <- bins[-length(bins)] + bin_width / 2  # mean of each bin range
+
+distance_two_curvy_dist <- distance_two_curvy |> 
+  mutate(bin = cut(distance, breaks = bins, include.lowest = TRUE, labels = bin_labels))
+
+distance_two_curvy_dist <- left_join(distance_two_curvy_dist, dist_highd, by = c("from", "to"))
+
+distance_points <- ggplot(
+  distance_two_curvy_dist, 
+  aes(x = distance,
+      y = dist_highd)) +
+  geom_point(alpha = 0.5) +
+  geom_vline(xintercept = 2.5 * bin_width,
+             linetype="solid",
+             color = "#bdbdbd",
+             linewidth=1) +
+  geom_vline(xintercept = 2 * bin_width,
+             linetype="solid",
+             color = "#bdbdbd",
+             linewidth=1) +
+  ylab(expression(d^{(7)})) +
+  xlab(expression(d^{(2)})) +
+  theme_minimal() +
+  theme(aspect.ratio = 0.75,
+        panel.border = element_rect(fill = 'transparent'),
+        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+        panel.grid.major.x = element_blank(),
+        axis.ticks.x = element_line(),
+        axis.ticks.y = element_line())
+
+```
+
+```{r}
 #| label: best-fit-two-curvy-proj-umap1
-
-df_bin_two_curvy_temp <- df_bin_two_curvy
 
@@
 
-### Compute highD distance
-dist_vec <- proxy::dist(x = df_bin_two_curvy2[, -1], method = "Euclidean") |> as.vector()
-
-from_vec <- c()
-to_vec <- c()
-num_obs <- 1:(NROW(df_bin_two_curvy2) - 1)
-
-for (obs in num_obs) {
-
-  from_val <- rep(obs, (NROW(df_bin_two_curvy2) - obs))
-  if ((obs + 1) <= NROW(df_bin_two_curvy2)) {
-    to_val <- (obs + 1):NROW(df_bin_two_curvy2)
-  }
-  from_vec <- append(from_vec, from_val)
-  to_vec <- append(to_vec, to_val)
-
-}
-
-dist_highd <- tibble::tibble(from = from_vec, to = to_vec, dist_highd = dist_vec)
-
 df_b_two_curvy2 <- df_bin_two_curvy2 |>
@@
   select_vars = c("from", "to", "distance"))
-
-## To plot the distribution of distances
-
-# Define bin width and starting point
-bin_width <- hb_obj_two_curvy2$a1
-start_point <- bin_width/2
-
-# Create bins and calculate the mean value for each bin range
-bins <- seq(start_point, max(distance_two_curvy$distance) + bin_width, 
-            by = bin_width)
-bin_labels <- bins[-length(bins)] + bin_width / 2  # mean of each bin range
-
-distance_two_curvy_dist <- distance_two_curvy |> 
-  mutate(bin = cut(distance, breaks = bins, include.lowest = TRUE, labels = bin_labels))
-
-distance_two_curvy_dist <- left_join(distance_two_curvy_dist, dist_highd, by = c("from", "to"))
-
-distance_points <- ggplot(
-  distance_two_curvy_dist, 
-  aes(x = distance,
-      y = dist_highd)) +
-  geom_point(alpha = 0.5) +
-  #geom_jitter() +
-  # scale_x_continuous(breaks = sort(unique(as.numeric(distance_two_curvy_dist$bin))),
-  #                    labels = c(
-  #                      expression(a[1]),
-  #                      expression(2*a[1]),
-  #                      expression(3*a[1]),
-  #                      expression(4*a[1]),
-  #                      expression(5*a[1]),
-  #                      expression(6*a[1]),
-  #                      expression(7*a[1]),
-  #                      expression(8*a[1])
-  #                    )) +
-  geom_vline(xintercept = 2.5 * bin_width,
-             linetype="solid",
-             color = "#bdbdbd",
-             linewidth=1) +
-  geom_vline(xintercept = 2 * bin_width,
-           linetype="solid",
-           color = "#bdbdbd",
-           linewidth=1) +
-  # geom_vline(xintercept = 4.5,
-  #            linetype="solid",
-  #            color = "#bdbdbd",
-  #            linewidth=1) +
-  ylab(expression(d^{(7)})) +
-  xlab(expression(d^{(2)})) +
-  #ggtitle("(a)") +
-  theme_minimal() +
-  theme(aspect.ratio = 0.75,
-        panel.border = element_rect(fill = 'transparent'),
-        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
-        panel.grid.major.x = element_blank(),
-        axis.ticks.x = element_line(),
-        axis.ticks.y = element_line())
-
-distance_two_curvy_dist_summary <- distance_two_curvy_dist |>
-  count(bin) |>
-  mutate(prop = n/sum(n)) |>
-  head(4)
-
-# distance_two_curvy_dist <- distance_two_curvy_dist |>
-#   filter(bin != unique(distance_two_curvy_dist$bin)[1]) ## To remove a1 distances, since always small
-
-text_df <- tibble(
-  x = c(1.9, 3.4),
-  y = c(0.5, 0.5),
-  text = c("e1", "d1")
-)
-
-distance_hist <- ggplot(
-  distance_two_curvy_dist_summary,
-  aes(x = as.numeric(bin),
-      y = prop)) +
-  geom_line() +
-  scale_x_continuous(breaks = sort(unique(as.numeric(distance_two_curvy_dist$bin)))[1:4],
-                     labels = c(
-                       expression(a[1]),
-                       expression(2*a[1]),
-                       expression(3*a[1]),
-                       expression(4*a[1])
-                     )) +
-  geom_vline(xintercept = 2,
-             linetype="solid",
-             color = "#bdbdbd",
-             linewidth=1) +
-  geom_vline(xintercept = 3.5,
-             linetype="solid",
-             color = "#bdbdbd",
-             linewidth=1) +
-  geom_text(
-    data = text_df,
-    aes(
-      x = x,
-      y = y,
-      label = text
-    ),
-    size = 9
-  ) +
-  ylab("Proportion of the number of edges") +
-  xlab(expression(d^{(2)})) +
-  ggtitle("(b)") +
-  theme_minimal() +
-  theme(aspect.ratio = 0.75,
-        panel.border = element_rect(fill = 'transparent'),
-        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
-        panel.grid.major.x = element_blank(),
-        axis.ticks.x = element_line(),
-        axis.ticks.y = element_line())
-
-# distance_hist <- ggplot(
-#   distance_two_curvy_dist, aes(x = distance)) +
-#     geom_histogram(
-#       aes(y = after_stat(count) / sum(after_stat(count))), 
-#       binwidth = bin_width) +
-#   scale_x_continuous(breaks = bin_labels, 
-#                      labels = c(
-#                        expression(a[1]),
-#                        expression(2*a[1]),
-#                        expression(3*a[1]),
-#                        expression(4*a[1]),
-#                        expression(5*a[1]),
-#                        expression(6*a[1]),
-#                        expression(7*a[1]),
-#                        expression(8*a[1])
-#                      )) +
-#   geom_vline(xintercept = start_point + 2*bin_width,
-#              linetype="solid",
-#              color = "#bdbdbd",
-#              linewidth=1) +
-#   ylab("Proportion of the number of edges") +
-#   xlab(expression(d^{(2)})) +
-#   #ggtitle("(d)") +
-#   theme_minimal() +
-#   theme(aspect.ratio = 0.75,
-#         panel.border = element_rect(fill = 'transparent'),
-#         plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
-#         panel.grid.major.x = element_blank(),
-#         axis.ticks.x = element_line(),
-#         axis.ticks.y = element_line())
 

### 2025-01-10 paper.qmd
Updated Figure 9b (a1)

@@
   tri_object = tr1_object_two_curvy)
+
+trimesh_two_curvy <- ggplot() +
+  geom_segment(data = tr_from_to_df_two_curvy,
+               aes(
+                 x = x_from,
+                 y = y_from,
+                 xend = x_to,
+                 yend = y_to),
+               colour = "#000000") +
+  geom_point(data = tsne_two_curvy_scaled,
+             aes(
+               x = tSNE1,
+               y = tSNE2
+             ),
+             color = "#636363",
+             alpha = 0.5,
+             size = 0.5
+  ) +
+  coord_equal() +
+  interior_annotation("a1", sc_ltr_pos)
 
@@
 free(distance_points) /
-  wrap_plots(trimesh_two_curvy2, trimesh_two_curvy_removed2, trimesh_two_curvy_removed1, two_curvy_proj_umap_model_delaunay, two_curvy_proj_umap_model_benchmark2, two_curvy_proj_umap_model, ncol = 3) +
+  wrap_plots(trimesh_two_curvy, trimesh_two_curvy_removed2, trimesh_two_curvy_removed1, two_curvy_proj_umap_model_delaunay, two_curvy_proj_umap_model_benchmark2, two_curvy_proj_umap_model, ncol = 3) +
   plot_annotation(tag_levels = list(c("(a)", "(b)")))

### 2025-01-10 paper.qmd
Updated Figure 9b (b1)

@@
 
+bin_width <- hb_obj_two_curvy$a1
+
+####With diff benchmark values
+
+## Benchmark 2
+
+benchmark_two_curvy <- 2.5 * bin_width
+
+distance_df_small_edges_two_curvy3 <- distance_two_curvy |>
+  filter(distance < benchmark_two_curvy)
+
+tr_from_to_df_two_curvy4 <- inner_join(
+  tr_from_to_df_two_curvy, distance_df_small_edges_two_curvy3,
+  by = c("from", "to"))
+
+trimesh_two_curvy_removed2 <- ggplot() +
+  geom_segment(data = tr_from_to_df_two_curvy4,
+               aes(
+                 x = x_from,
+                 y = y_from,
+                 xend = x_to,
+                 yend = y_to),
+               colour = "#000000")+
+  geom_point(data = tsne_two_curvy_scaled,
+             aes(
+               x = tSNE1,
+               y = tSNE2
+             ),
+             color = "#636363",
+             alpha = 0.5,
+             size = 0.5
+  ) +
+  coord_equal() +
+  interior_annotation("b1", sc_ltr_pos)
+
+## Benchmark 1
+
+benchmark_two_curvy <- 2 * bin_width
+
+distance_df_small_edges_two_curvy2 <- distance_two_curvy |>
+  filter(distance < benchmark_two_curvy) ## 0.231
+
+tr_from_to_df_two_curvy3 <- inner_join(
+  tr_from_to_df_two_curvy, distance_df_small_edges_two_curvy2,
+  by = c("from", "to"))
+
+trimesh_two_curvy_removed1 <- ggplot() +
+  geom_segment(data = tr_from_to_df_two_curvy3,
+               aes(
+                 x = x_from,
+                 y = y_from,
+                 xend = x_to,
+                 yend = y_to),
+               colour = "#000000")+
+  geom_point(data = tsne_two_curvy_scaled,
+             aes(
+               x = tSNE1,
+               y = tSNE2
+             ),
+             color = "#636363",
+             alpha = 0.5,
+             size = 0.5
+  ) +
+  coord_equal() +
+  interior_annotation("c1", sc_ltr_pos)
+
+trimesh_two_curvy_removed1_with_data <- ggplot() +
+  geom_segment(data = tr_from_to_df_two_curvy3,
+               aes(
+                 x = x_from,
+                 y = y_from,
+                 xend = x_to,
+                 yend = y_to),
+               colour = "#000000") +
+  geom_point(data = tsne_two_curvy_scaled,
+             aes(
+               x = tSNE1,
+               y = tSNE2
+             ),
+             color = "#636363",
+             alpha = 0.5,
+             size = 0.5
+  ) +
+  coord_equal() +
+  #xlim(c(-0.01, 1.1)) + ylim(c(-0.08, 0.57)) +
+  interior_annotation("a2", sc_ltr_pos)
+
+## Computed benchmark value
 tr_from_to_df_two_curvy <- inner_join(
@@
 # Define bin width and starting point
-bin_width <- hb_obj_two_curvy$a1
 start_point <- bin_width/2
@@
 
-## Benchmark 1
-
-benchmark_two_curvy <- 2 * bin_width
-
-distance_df_small_edges_two_curvy2 <- distance_two_curvy |>
-  filter(distance < benchmark_two_curvy) ## 0.231
-
-tr_from_to_df_two_curvy3 <- inner_join(
-  tr_from_to_df_two_curvy2, distance_df_small_edges_two_curvy2,
-  by = c("from", "to"))
-
-trimesh_two_curvy_removed1 <- ggplot() +
-  geom_segment(data = tr_from_to_df_two_curvy3,
-               aes(
-                 x = x_from,
-                 y = y_from,
-                 xend = x_to,
-                 yend = y_to),
-               colour = "#000000")+
-  geom_point(data = tsne_two_curvy_scaled,
-             aes(
-               x = tSNE1,
-               y = tSNE2
-             ),
-             color = "#636363",
-            alpha = 0.5,
-            size = 0.5
-            ) +
-  coord_equal() +
-  interior_annotation("c1", sc_ltr_pos)
-
-trimesh_two_curvy_removed1_with_data <- ggplot() +
-  geom_segment(data = tr_from_to_df_two_curvy3,
-               aes(
-                 x = x_from,
-                 y = y_from,
-                 xend = x_to,
-                 yend = y_to),
-               colour = "#000000") +
-  geom_point(data = tsne_two_curvy_scaled,
-             aes(
-               x = tSNE1,
-               y = tSNE2
-             ),
-             color = "#636363",
-            alpha = 0.5,
-            size = 0.5
-            ) +
-  coord_equal() +
-  #xlim(c(-0.01, 1.1)) + ylim(c(-0.08, 0.57)) +
-  interior_annotation("a2", sc_ltr_pos)
-
-## Benchmark 2
-
-benchmark_two_curvy <- 2.5 * bin_width
-
-distance_df_small_edges_two_curvy3 <- distance_two_curvy |>
-  filter(distance < benchmark_two_curvy)
-
-tr_from_to_df_two_curvy4 <- inner_join(
-  tr_from_to_df_two_curvy2, distance_df_small_edges_two_curvy3,
-  by = c("from", "to"))
-
-trimesh_two_curvy_removed2 <- ggplot() +
-  geom_segment(data = tr_from_to_df_two_curvy4,
-               aes(
-                 x = x_from,
-                 y = y_from,
-                 xend = x_to,
-                 yend = y_to),
-               colour = "#000000")+
-  geom_point(data = tsne_two_curvy_scaled,
-             aes(
-               x = tSNE1,
-               y = tSNE2
-             ),
-             color = "#636363",
-            alpha = 0.5,
-            size = 0.5
-            ) +
-  coord_equal() +
-  interior_annotation("b1", sc_ltr_pos)
-
 ```

### 2025-01-10 paper.qmd
Updated Figure 9b (a2)
Updated Figure 9b (b2)

@@
 
+## Projection with Delauany triangulation
+
+two_curvy_proj_umap_model_delaunay <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df1, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#000000") + #31a354
+  geom_point(
+    size = 0.5,
+    alpha = 0.2,
+    color = "#636363") +
+  geom_segment(
+    data=axes,
+    aes(x=x1, y=y1, xend=x2, yend=y2),
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
+    colour="grey50",
+    size = 4) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.1, 0.1)) + 
+  ylim(c(-0.1, 0.1)) +
+  interior_annotation("a2", sc_ltr_pos) +
+  theme(
+    legend.position = "none"
+  )
+
+two_curvy_proj_umap_model_benchmark2 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df2, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#000000") + #31a354
+  geom_point(
+    size = 0.5,
+    alpha = 0.2,
+    color = "#636363") +
+  geom_segment(
+    data=axes,
+    aes(x=x1, y=y1, xend=x2, yend=y2),
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
+    colour="grey50",
+    size = 4) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.1, 0.1)) + 
+  ylim(c(-0.1, 0.1)) +
+  interior_annotation("b2", sc_ltr_pos) +
+  theme(
+    legend.position = "none"
+  )
 ```
@@
 
-## Projection with Delauany triangulation
-
-model_df1 <- dplyr::left_join(
-  distance_two_curvy |> select(-distance), 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df1)[3:NCOL(model_df1)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df1 <- dplyr::left_join(model_df1, projected_model_df, by = c("to" = "ID"))
-names(model_df1)[(2 + NCOL(projected_model_df)):NCOL(model_df1)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-## With benchmark option 2
-
-model_df2 <- dplyr::left_join(
-  distance_df_small_edges_two_curvy3 |> select(-distance), 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df2)[3:NCOL(model_df2)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df2 <- dplyr::left_join(model_df2, projected_model_df, by = c("to" = "ID"))
-names(model_df2)[(2 + NCOL(projected_model_df)):NCOL(model_df2)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
 ## For true model
@@
 
-two_curvy_proj_umap_model_delaunay <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_segment(
-    data = model_df1, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#000000") + #31a354
-  geom_point(
-    size = 0.5,
-    alpha = 0.2,
-    color = "#636363") +
-  geom_segment(
-    data=axes,
-    aes(x=x1, y=y1, xend=x2, yend=y2),
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 4) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.1, 0.1)) + 
-  ylim(c(-0.1, 0.1)) +
-  interior_annotation("a2", sc_ltr_pos) +
-  theme(
-    legend.position = "none"
-  )
-
-two_curvy_proj_umap_model_benchmark2 <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_segment(
-    data = model_df2, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#000000") + #31a354
-  geom_point(
-    size = 0.5,
-    alpha = 0.2,
-    color = "#636363") +
-  geom_segment(
-    data=axes,
-    aes(x=x1, y=y1, xend=x2, yend=y2),
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 4) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.1, 0.1)) + 
-  ylim(c(-0.1, 0.1)) +
-  interior_annotation("b2", sc_ltr_pos) +
-  theme(
-    legend.position = "none"
-  )
-
 two_curvy_proj_umap_all_model <- projected_df |>

### 2025-01-10 paper.qmd
Updated Figure 9b (c1)

@@
 free(distance_points) /
-  wrap_plots(trimesh_two_curvy, trimesh_two_curvy_removed2, trimesh_two_curvy_removed1, two_curvy_proj_umap_model_delaunay, two_curvy_proj_umap_model_benchmark2, two_curvy_proj_umap_model, ncol = 3) +
+  wrap_plots(trimesh_two_curvy, trimesh_two_curvy_removed2, trimesh_two_curvy_removed1, two_curvy_proj_umap_model_delaunay, two_curvy_proj_umap_model_benchmark2, two_curvy_proj_umap_first_model1, ncol = 3) +
   plot_annotation(tag_levels = list(c("(a)", "(b)")))

### 2025-01-10 paper.qmd
Updated Figure7

@@
   hex_grid_coloured_two_curvy3 +
-  boxplot1 + boxplot2 + boxplot3 +
+  #boxplot1 + boxplot2 + boxplot3 +
   plot_layout(guides='collect', ncol = 3) &

### 2025-01-10 paper.qmd
Updated Figure 7 title

@@
 #| fig-pos: H
-#| fig-cap: "Hexbin density plots of tSNE layout of the two C-shaped clusters data, using three different bin inputs: (a) $b = 80 \\text{ } (10, \\text{ }8)$, (b) $b = 130 \\text{ } (13, \\text{ }10)$, and (c) $b = 391 \\text{ } (23, \\text{ }17)$.
+#| fig-cap: "Hexbin density plots of tSNE layout of the two C-shaped clusters data, using three different bin inputs: (a) $b = 255 \\text{ } (15, \\text{ }17)$, (b) $b = 598 \\text{ } (23, \\text{ }26)$, and (c) $b = 2544 \\text{ } (48, \\text{ }53)$.
 Color indicates standardized counts, dark indicating high count and light indicates low count.

### 2025-01-10 paper.qmd
Updated Figure 12

@@
 ```{r}
-#| fig-cap: "Assessing which of the 5 NLDR layouts on the two C-shaped clusters data is the better representation using MSE for varying binwidth ($a_1$).
+#| fig-cap: "Assessing which of the 6 NLDR layouts on the two C-shaped clusters data is the better representation using MSE for varying binwidth ($a_1$).
 Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-e).

### 2025-01-10 paper.tex
Updated Figure 7 title
Updated Figure 12

@@
 of the two C-shaped clusters data, using three different bin inputs: (a)
-\(b = 80 \text{ } (10, \text{ }8)\), (b)
-\(b = 130 \text{ } (13, \text{ }10)\), and (c)
-\(b = 391 \text{ } (23, \text{ }17)\). Color indicates standardized
+\(b = 255 \text{ } (15, \text{ }17)\), (b)
+\(b = 598 \text{ } (23, \text{ }26)\), and (c)
+\(b = 2544 \text{ } (48, \text{ }53)\). Color indicates standardized
 counts, dark indicating high count and light indicates low count. At the
@@
 \caption{\label{fig-two_non_linear_diff_shaped_close_clusters-mse}Assessing
-which of the 5 NLDR layouts on the two C-shaped clusters data is the
+which of the 6 NLDR layouts on the two C-shaped clusters data is the
 better representation using MSE for varying binwidth (\(a_1\)). Colour

### 2025-01-10 scripts/one_c_shaped_dens_structure/04_gen_error_plots.R

@@ -289,7 +289,8 @@ five_c_shaped_proj_tsne_model1 <- projected_df |>
       xend = proj1_to,
       yend = proj2_to),
     color = "#000000",
-    linewidth = 1) +
+    linewidth = 0.8,
+    alpha = 0.6) +
   geom_point(
     #size = 0.5,
     aes(
@@ -318,7 +319,7 @@ five_c_shaped_proj_tsne_model1 <- projected_df |>
   coord_fixed() +
   xlim(c(-0.6, 0.6)) +
   ylim(c(-0.6, 0.6)) +
-  interior_annotation("b2",
+  interior_annotation("a2",
                       position = c(0.08, 0.9),
                       cex = 1.5)
 
@@ -428,7 +429,7 @@ error_plot_tsne_uni <- error_df_one_curvy_abs |>
   theme(
     aspect.ratio = 1
   ) +
-  interior_annotation("a1",
+  interior_annotation("b1",
                       position = c(0.08, 0.95),
                       cex = 1.5)
 
@@ -590,7 +591,8 @@ five_c_shaped_proj_tsne_model2 <- projected_df |>
       xend = proj1_to,
       yend = proj2_to),
     color = "#000000",
-    linewidth = 1) +
+    linewidth = 0.8,
+    alpha = 0.6) +
   geom_point(
     #size = 0.5,
     aes(

### 2025-01-12 paper.qmd
Updated Figure 12 ordering

@@
 
-To illustrate how the methods organise difference number of points within the structure in the NLDR layout, simulated $7\text{-}D$ data having C-shaped structure is used.
+To illustrate how the methods organize difference number of points within the structure in the NLDR layout, simulated $7\text{-}D$ data having C-shaped structure is used.
 

### 2025-01-12 paper.tex
Updated Figure 12 ordering

@@
 
-To illustrate how the methods organise difference number of points
+To illustrate how the methods organize difference number of points
 within the structure in the NLDR layout, simulated \(7\text{-}D\) data

### 2025-01-12 paper.qmd
Added Figure 12 title

@@
 #| fig-pos: H
-#| fig-cap: "All \\gD{} layouts exhibit high error in the sparse end.
-The three \\gD{} layouts are: (a) tSNE, (b) UMAP, and (c) PaCMaP, which represent a C-shaped dense cluster colored according to the high-dimensional absolute error.
-Darker colors indicate high error, while lighter colors indicate low error.
-Regardless of the NLDR method used, the dense end of the structure shows low error, whereas the sparse end displays high error."
+#| fig-cap: "The sparse end of the structure exhibits a high model error, while the dense end shows a low model error.
+The tSNE layouts of the C-shaped cluster, with uniform points (a1) and dense points (b1), along with their selected projections (a2 and b2), are displayed.
+Darker colors represent a high error, while lighter colors indicate a low error.
+The uniformly distributed points demonstrate random model errors throughout the structure, whereas the dense structure shows large model error at the sparse end."
 

### 2025-01-12 paper.tex
Added Figure 12 title

@@
 
-\caption{\label{fig-one-dens_clust-error}All \gD{} layouts exhibit high
-error in the sparse end. The three \gD{} layouts are: (a) tSNE, (b)
-UMAP, and (c) PaCMaP, which represent a C-shaped dense cluster colored
-according to the high-dimensional absolute error. Darker colors indicate
-high error, while lighter colors indicate low error. Regardless of the
-NLDR method used, the dense end of the structure shows low error,
-whereas the sparse end displays high error.}
+\caption{\label{fig-one-dens_clust-error}The sparse end of the structure
+exhibits a high model error, while the dense end shows a low model
+error. The tSNE layouts of the C-shaped cluster, with uniform points
+(a1) and dense points (b1), along with their selected projections (a2
+and b2), are displayed. Darker colors represent a high error, while
+lighter colors indicate a low error. The uniformly distributed points
+demonstrate random model errors throughout the structure, whereas the
+dense structure shows large model error at the sparse end.}
 

### 2025-01-12 paper.qmd
changes by DC but not checked in early

--- 2025-01-10 paper.qmd
@@
 source("scripts/one_c_shaped_dens_structure/04_gen_error_plots.R")
-source("scripts/five_gau_clusters/5_gen_model_with_tsne.R")
 ```
@@
 
-boxplot2 <- ggplot(df_bin_centroids_two_curvy, 
-       aes(x=bin_counts)) + 
-  geom_histogram()
-
 umap_data_two_curvy_with_hb_id <- hb_obj_two_curvy$data_hb_id
@@
     na.value = "#ffffff", option = "E") +
-  interior_annotation("b", sc_ltr_pos) + 
+  interior_annotation("a", sc_ltr_pos) + 
   xlim(c(-0.3, 1.2)) + ylim(c(-0.17, 1.1))
@@
 
-trimesh_two_curvy <- ggplot() +
-  geom_segment(data = tr_from_to_df_two_curvy,
-               aes(
-                 x = x_from,
-                 y = y_from,
-                 xend = x_to,
-                 yend = y_to),
-               colour = "#000000") +
-  geom_point(data = tsne_two_curvy_scaled,
-             aes(
-               x = tSNE1,
-               y = tSNE2
-             ),
-             color = "#636363",
-             alpha = 0.5,
-             size = 0.5
-  ) +
-  coord_equal() +
-  interior_annotation("a1", sc_ltr_pos)
-
 ## Compute 2D distances
@@
 
-bin_width <- hb_obj_two_curvy$a1
-
-####With diff benchmark values
-
-## Benchmark 2
-
-benchmark_two_curvy <- 2.5 * bin_width
-
-distance_df_small_edges_two_curvy3 <- distance_two_curvy |>
-  filter(distance < benchmark_two_curvy)
-
-tr_from_to_df_two_curvy4 <- inner_join(
-  tr_from_to_df_two_curvy, distance_df_small_edges_two_curvy3,
-  by = c("from", "to"))
-
-trimesh_two_curvy_removed2 <- ggplot() +
-  geom_segment(data = tr_from_to_df_two_curvy4,
-               aes(
-                 x = x_from,
-                 y = y_from,
-                 xend = x_to,
-                 yend = y_to),
-               colour = "#000000")+
-  geom_point(data = tsne_two_curvy_scaled,
-             aes(
-               x = tSNE1,
-               y = tSNE2
-             ),
-             color = "#636363",
-             alpha = 0.5,
-             size = 0.5
-  ) +
-  coord_equal() +
-  interior_annotation("b1", sc_ltr_pos)
-
-## Benchmark 1
-
-benchmark_two_curvy <- 2 * bin_width
-
-distance_df_small_edges_two_curvy2 <- distance_two_curvy |>
-  filter(distance < benchmark_two_curvy) ## 0.231
-
-tr_from_to_df_two_curvy3 <- inner_join(
-  tr_from_to_df_two_curvy, distance_df_small_edges_two_curvy2,
-  by = c("from", "to"))
-
-trimesh_two_curvy_removed1 <- ggplot() +
-  geom_segment(data = tr_from_to_df_two_curvy3,
-               aes(
-                 x = x_from,
-                 y = y_from,
-                 xend = x_to,
-                 yend = y_to),
-               colour = "#000000")+
-  geom_point(data = tsne_two_curvy_scaled,
-             aes(
-               x = tSNE1,
-               y = tSNE2
-             ),
-             color = "#636363",
-             alpha = 0.5,
-             size = 0.5
-  ) +
-  coord_equal() +
-  interior_annotation("c1", sc_ltr_pos)
-
-trimesh_two_curvy_removed1_with_data <- ggplot() +
-  geom_segment(data = tr_from_to_df_two_curvy3,
-               aes(
-                 x = x_from,
-                 y = y_from,
-                 xend = x_to,
-                 yend = y_to),
-               colour = "#000000") +
-  geom_point(data = tsne_two_curvy_scaled,
-             aes(
-               x = tSNE1,
-               y = tSNE2
-             ),
-             color = "#636363",
-             alpha = 0.5,
-             size = 0.5
-  ) +
-  coord_equal() +
-  #xlim(c(-0.01, 1.1)) + ylim(c(-0.08, 0.57)) +
-  interior_annotation("a2", sc_ltr_pos)
-
-## Computed benchmark value
 tr_from_to_df_two_curvy <- inner_join(
@@
 ```{r}
-#| label: dist-plot
+#| label: best-fit-two-curvy-proj-umap1
 
 df_bin_two_curvy_temp <- df_bin_two_curvy
-
-### Compute highD distance
-dist_vec <- proxy::dist(x = df_bin_two_curvy_temp[, -1], method = "Euclidean") |> as.vector()
-
-from_vec <- c()
-to_vec <- c()
-num_obs <- 1:(NROW(df_bin_two_curvy_temp) - 1)
-
-for (obs in num_obs) {
-
-  from_val <- rep(obs, (NROW(df_bin_two_curvy_temp) - obs))
-  if ((obs + 1) <= NROW(df_bin_two_curvy_temp)) {
-    to_val <- (obs + 1):NROW(df_bin_two_curvy_temp)
-  }
-  from_vec <- append(from_vec, from_val)
-  to_vec <- append(to_vec, to_val)
-
-}
-
-dist_highd <- tibble::tibble(from = from_vec, to = to_vec, dist_highd = dist_vec)
-
-## To plot the distribution of distances
-
-# Define bin width and starting point
-start_point <- bin_width/2
-
-# Create bins and calculate the mean value for each bin range
-bins <- seq(start_point, max(distance_two_curvy$distance) + bin_width, 
-            by = bin_width)
-bin_labels <- bins[-length(bins)] + bin_width / 2  # mean of each bin range
-
-distance_two_curvy_dist <- distance_two_curvy |> 
-  mutate(bin = cut(distance, breaks = bins, include.lowest = TRUE, labels = bin_labels))
-
-distance_two_curvy_dist <- left_join(distance_two_curvy_dist, dist_highd, by = c("from", "to"))
-
-distance_points <- ggplot(
-  distance_two_curvy_dist, 
-  aes(x = distance,
-      y = dist_highd)) +
-  geom_point(alpha = 0.5) +
-  geom_vline(xintercept = 2.5 * bin_width,
-             linetype="solid",
-             color = "#bdbdbd",
-             linewidth=1) +
-  geom_vline(xintercept = 2 * bin_width,
-             linetype="solid",
-             color = "#bdbdbd",
-             linewidth=1) +
-  ylab(expression(d^{(7)})) +
-  xlab(expression(d^{(2)})) +
-  theme_minimal() +
-  theme(aspect.ratio = 0.75,
-        panel.border = element_rect(fill = 'transparent'),
-        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
-        panel.grid.major.x = element_blank(),
-        axis.ticks.x = element_line(),
-        axis.ticks.y = element_line())
-
-```
-
-```{r}
-#| label: best-fit-two-curvy-proj-umap1
 
@@
 names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-## Projection with Delauany triangulation
-
-model_df1 <- dplyr::left_join(
-  distance_two_curvy |> select(-distance), 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df1)[3:NCOL(model_df1)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df1 <- dplyr::left_join(model_df1, projected_model_df, by = c("to" = "ID"))
-names(model_df1)[(2 + NCOL(projected_model_df)):NCOL(model_df1)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-## With benchmark option 2
-
-model_df2 <- dplyr::left_join(
-  distance_df_small_edges_two_curvy3 |> select(-distance), 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df2)[3:NCOL(model_df2)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df2 <- dplyr::left_join(model_df2, projected_model_df, by = c("to" = "ID"))
-names(model_df2)[(2 + NCOL(projected_model_df)):NCOL(model_df2)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
 
@@
     size = 0.5,
-    alpha = 0.2,
+    alpha = 0.5,
     color = "#636363") +
@@
 
-## Projection with Delauany triangulation
-
-two_curvy_proj_umap_model_delaunay <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_segment(
-    data = model_df1, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#000000") + #31a354
-  geom_point(
-    size = 0.5,
-    alpha = 0.2,
-    color = "#636363") +
-  geom_segment(
-    data=axes,
-    aes(x=x1, y=y1, xend=x2, yend=y2),
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 4) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.1, 0.1)) + 
-  ylim(c(-0.1, 0.1)) +
-  interior_annotation("a2", sc_ltr_pos) +
-  theme(
-    legend.position = "none"
-  )
-
-two_curvy_proj_umap_model_benchmark2 <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_segment(
-    data = model_df2, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#000000") + #31a354
-  geom_point(
-    size = 0.5,
-    alpha = 0.2,
-    color = "#636363") +
-  geom_segment(
-    data=axes,
-    aes(x=x1, y=y1, xend=x2, yend=y2),
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 4) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.1, 0.1)) + 
-  ylim(c(-0.1, 0.1)) +
-  interior_annotation("b2", sc_ltr_pos) +
-  theme(
-    legend.position = "none"
-  )
 ```
@@
 
-boxplot1 <- ggplot(df_bin_centroids2, 
-                   aes(x=bin_counts)) + 
-    geom_histogram()
-
 hex_grid_with_counts_two_curvy2 <- full_join(
@@
 df_bin_two_curvy2 <- avg_highd_data(data = df_all_two_curvy2, col_start = "x")
+
+### Compute highD distance
+dist_vec <- proxy::dist(x = df_bin_two_curvy2[, -1], method = "Euclidean") |> as.vector()
+
+from_vec <- c()
+to_vec <- c()
+num_obs <- 1:(NROW(df_bin_two_curvy2) - 1)
+
+for (obs in num_obs) {
+
+  from_val <- rep(obs, (NROW(df_bin_two_curvy2) - obs))
+  if ((obs + 1) <= NROW(df_bin_two_curvy2)) {
+    to_val <- (obs + 1):NROW(df_bin_two_curvy2)
+  }
+  from_vec <- append(from_vec, from_val)
+  to_vec <- append(to_vec, to_val)
+
+}
+
+dist_highd <- tibble::tibble(from = from_vec, to = to_vec, dist_highd = dist_vec)
 
@@
   xlim(c(-0.3, 1.2)) + ylim(c(-0.17, 1.1)) +
-  interior_annotation("a", sc_ltr_pos)
+  interior_annotation("b", sc_ltr_pos)
 
@@
 
+## To plot the distribution of distances
+
+# Define bin width and starting point
+bin_width <- hb_obj_two_curvy2$a1
+start_point <- bin_width/2
+
+# Create bins and calculate the mean value for each bin range
+bins <- seq(start_point, max(distance_two_curvy$distance) + bin_width, 
+            by = bin_width)
+bin_labels <- bins[-length(bins)] + bin_width / 2  # mean of each bin range
+
+distance_two_curvy_dist <- distance_two_curvy |> 
+  mutate(bin = cut(distance, breaks = bins, include.lowest = TRUE, labels = bin_labels))
+
+distance_two_curvy_dist <- left_join(distance_two_curvy_dist, dist_highd, by = c("from", "to"))
+
+distance_points <- ggplot(
+  distance_two_curvy_dist, 
+  aes(x = distance,
+      y = dist_highd)) +
+  geom_point(alpha = 0.5) +
+  #geom_jitter() +
+  # scale_x_continuous(breaks = sort(unique(as.numeric(distance_two_curvy_dist$bin))),
+  #                    labels = c(
+  #                      expression(a[1]),
+  #                      expression(2*a[1]),
+  #                      expression(3*a[1]),
+  #                      expression(4*a[1]),
+  #                      expression(5*a[1]),
+  #                      expression(6*a[1]),
+  #                      expression(7*a[1]),
+  #                      expression(8*a[1])
+  #                    )) +
+  geom_vline(xintercept = 2.5 * bin_width,
+             linetype="solid",
+             color = "#bdbdbd",
+             linewidth=1) +
+  geom_vline(xintercept = 2 * bin_width,
+           linetype="solid",
+           color = "#bdbdbd",
+           linewidth=1) +
+  # geom_vline(xintercept = 4.5,
+  #            linetype="solid",
+  #            color = "#bdbdbd",
+  #            linewidth=1) +
+  ylab(expression(d^{(7)})) +
+  xlab(expression(d^{(2)})) +
+  #ggtitle("(a)") +
+  theme_minimal() +
+  theme(aspect.ratio = 0.75,
+        panel.border = element_rect(fill = 'transparent'),
+        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+        panel.grid.major.x = element_blank(),
+        axis.ticks.x = element_line(),
+        axis.ticks.y = element_line())
+
+distance_two_curvy_dist_summary <- distance_two_curvy_dist |>
+  count(bin) |>
+  mutate(prop = n/sum(n)) |>
+  head(4)
+
+# distance_two_curvy_dist <- distance_two_curvy_dist |>
+#   filter(bin != unique(distance_two_curvy_dist$bin)[1]) ## To remove a1 distances, since always small
+
+text_df <- tibble(
+  x = c(1.9, 3.4),
+  y = c(0.5, 0.5),
+  text = c("e1", "d1")
+)
+
+distance_hist <- ggplot(
+  distance_two_curvy_dist_summary,
+  aes(x = as.numeric(bin),
+      y = prop)) +
+  geom_line() +
+  scale_x_continuous(breaks = sort(unique(as.numeric(distance_two_curvy_dist$bin)))[1:4],
+                     labels = c(
+                       expression(a[1]),
+                       expression(2*a[1]),
+                       expression(3*a[1]),
+                       expression(4*a[1])
+                     )) +
+  geom_vline(xintercept = 2,
+             linetype="solid",
+             color = "#bdbdbd",
+             linewidth=1) +
+  geom_vline(xintercept = 3.5,
+             linetype="solid",
+             color = "#bdbdbd",
+             linewidth=1) +
+  geom_text(
+    data = text_df,
+    aes(
+      x = x,
+      y = y,
+      label = text
+    ),
+    size = 9
+  ) +
+  ylab("Proportion of the number of edges") +
+  xlab(expression(d^{(2)})) +
+  ggtitle("(b)") +
+  theme_minimal() +
+  theme(aspect.ratio = 0.75,
+        panel.border = element_rect(fill = 'transparent'),
+        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+        panel.grid.major.x = element_blank(),
+        axis.ticks.x = element_line(),
+        axis.ticks.y = element_line())
+
+# distance_hist <- ggplot(
+#   distance_two_curvy_dist, aes(x = distance)) +
+#     geom_histogram(
+#       aes(y = after_stat(count) / sum(after_stat(count))), 
+#       binwidth = bin_width) +
+#   scale_x_continuous(breaks = bin_labels, 
+#                      labels = c(
+#                        expression(a[1]),
+#                        expression(2*a[1]),
+#                        expression(3*a[1]),
+#                        expression(4*a[1]),
+#                        expression(5*a[1]),
+#                        expression(6*a[1]),
+#                        expression(7*a[1]),
+#                        expression(8*a[1])
+#                      )) +
+#   geom_vline(xintercept = start_point + 2*bin_width,
+#              linetype="solid",
+#              color = "#bdbdbd",
+#              linewidth=1) +
+#   ylab("Proportion of the number of edges") +
+#   xlab(expression(d^{(2)})) +
+#   #ggtitle("(d)") +
+#   theme_minimal() +
+#   theme(aspect.ratio = 0.75,
+#         panel.border = element_rect(fill = 'transparent'),
+#         plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+#         panel.grid.major.x = element_blank(),
+#         axis.ticks.x = element_line(),
+#         axis.ticks.y = element_line())
+
 ## To find the benchmark value
@@
   distance_col = "distance")
+
+## Benchmark 1
+
+benchmark_two_curvy <- 2 * bin_width
+
+distance_df_small_edges_two_curvy2 <- distance_two_curvy |>
+  filter(distance < benchmark_two_curvy) ## 0.231
+
+tr_from_to_df_two_curvy3 <- inner_join(
+  tr_from_to_df_two_curvy2, distance_df_small_edges_two_curvy2,
+  by = c("from", "to"))
+
+trimesh_two_curvy_removed1 <- ggplot() +
+  geom_segment(data = tr_from_to_df_two_curvy3,
+               aes(
+                 x = x_from,
+                 y = y_from,
+                 xend = x_to,
+                 yend = y_to),
+               colour = "#000000")+
+  geom_point(data = tsne_two_curvy_scaled,
+             aes(
+               x = tSNE1,
+               y = tSNE2
+             ),
+             color = "#636363",
+            alpha = 0.5,
+            size = 0.5
+            ) +
+  coord_equal() +
+  interior_annotation("c1", sc_ltr_pos)
+
+trimesh_two_curvy_removed1_with_data <- ggplot() +
+  geom_segment(data = tr_from_to_df_two_curvy3,
+               aes(
+                 x = x_from,
+                 y = y_from,
+                 xend = x_to,
+                 yend = y_to),
+               colour = "#000000") +
+  geom_point(data = tsne_two_curvy_scaled,
+             aes(
+               x = tSNE1,
+               y = tSNE2
+             ),
+             color = "#636363",
+            alpha = 0.5,
+            size = 0.5
+            ) +
+  coord_equal() +
+  #xlim(c(-0.01, 1.1)) + ylim(c(-0.08, 0.57)) +
+  interior_annotation("a2", sc_ltr_pos)
+
+## Benchmark 2
+
+benchmark_two_curvy <- 2.5 * bin_width
+
+distance_df_small_edges_two_curvy3 <- distance_two_curvy |>
+  filter(distance < benchmark_two_curvy)
+
+tr_from_to_df_two_curvy4 <- inner_join(
+  tr_from_to_df_two_curvy2, distance_df_small_edges_two_curvy3,
+  by = c("from", "to"))
+
+trimesh_two_curvy_removed2 <- ggplot() +
+  geom_segment(data = tr_from_to_df_two_curvy4,
+               aes(
+                 x = x_from,
+                 y = y_from,
+                 xend = x_to,
+                 yend = y_to),
+               colour = "#000000")+
+  geom_point(data = tsne_two_curvy_scaled,
+             aes(
+               x = tSNE1,
+               y = tSNE2
+             ),
+             color = "#636363",
+            alpha = 0.5,
+            size = 0.5
+            ) +
+  coord_equal() +
+  interior_annotation("b1", sc_ltr_pos)
 
@@
 names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+## Projection with Delauany triangulation
+
+model_df1 <- dplyr::left_join(
+  distance_two_curvy |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df1)[3:NCOL(model_df1)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df1 <- dplyr::left_join(model_df1, projected_model_df, by = c("to" = "ID"))
+names(model_df1)[(2 + NCOL(projected_model_df)):NCOL(model_df1)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+## With benchmark option 2
+
+model_df2 <- dplyr::left_join(
+  distance_df_small_edges_two_curvy3 |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df2)[3:NCOL(model_df2)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df2 <- dplyr::left_join(model_df2, projected_model_df, by = c("to" = "ID"))
+names(model_df2)[(2 + NCOL(projected_model_df)):NCOL(model_df2)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
 
@@
 
+two_curvy_proj_umap_model_delaunay <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df1, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#000000") + #31a354
+  geom_point(
+    size = 0.5,
+    alpha = 0.2,
+    color = "#636363") +
+  geom_segment(
+    data=axes,
+    aes(x=x1, y=y1, xend=x2, yend=y2),
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
+    colour="grey50",
+    size = 4) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.1, 0.1)) + 
+  ylim(c(-0.1, 0.1)) +
+  interior_annotation("a2", sc_ltr_pos) +
+  theme(
+    legend.position = "none"
+  )
+
+two_curvy_proj_umap_model_benchmark2 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df2, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#000000") + #31a354
+  geom_point(
+    size = 0.5,
+    alpha = 0.2,
+    color = "#636363") +
+  geom_segment(
+    data=axes,
+    aes(x=x1, y=y1, xend=x2, yend=y2),
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
+    colour="grey50",
+    size = 4) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.1, 0.1)) + 
+  ylim(c(-0.1, 0.1)) +
+  interior_annotation("b2", sc_ltr_pos) +
+  theme(
+    legend.position = "none"
+  )
+
 two_curvy_proj_umap_all_model <- projected_df |>
@@
 
-boxplot3 <- ggplot(df_bin_centroids3, 
-                   aes(x=bin_counts)) + 
-    geom_histogram()
-
 hex_grid_with_counts_s_curve3 <- full_join(hex_grid3, 
@@
 error_df_two_curvy_abs <- error_df_two_curvy_abs |>
-  mutate(sqrt_row_wise_total_error = sqrt(row_wise_total_error))
+  mutate(sqrt_row_wise_abs_error = sqrt(row_wise_abs_error))
 
@@
              y = tSNE2,
-             colour = sqrt_row_wise_total_error)) +
+             colour = sqrt_row_wise_abs_error)) +
   geom_point(alpha=0.5) +
@@
   ggplot(error_df_two_curvy_abs) +
-  geom_histogram(aes(x=sqrt_row_wise_total_error, y=..density..)) +
-  geom_density(aes(x=sqrt_row_wise_total_error, y=..density..), colour="red") +
+  geom_histogram(aes(x=sqrt_row_wise_abs_error^2, y=..density..)) +
+  geom_density(aes(x=sqrt_row_wise_abs_error^2, y=..density..), colour="red") +
   xlab(expression(group("|", e[hj], "|"))) +
@@
 #| fig-cap: "The $7\\text{-}D$ model error in \\gD{} layout.
-Color indicates square root of total error, dark blue indicating high error and light indicates low error.
+Color indicates square root of absolute error, dark blue indicating high error and light indicates low error.
 Most large errors are distributed near the edges of the clusters."
@@
 #| fig-pos: H
-#| fig-cap: "Hexbin density plots of tSNE layout of the two C-shaped clusters data, using three different bin inputs: (a) $b = 255 \\text{ } (15, \\text{ }17)$, (b) $b = 598 \\text{ } (23, \\text{ }26)$, and (c) $b = 2544 \\text{ } (48, \\text{ }53)$.
+#| fig-cap: "Hexbin density plots of tSNE layout of the two C-shaped clusters data, using three different bin inputs: (a) $b = 80 \\text{ } (10, \\text{ }8)$, (b) $b = 130 \\text{ } (13, \\text{ }10)$, and (c) $b = 391 \\text{ } (23, \\text{ }17)$.
 Color indicates standardized counts, dark indicating high count and light indicates low count.
@@
   hex_grid_coloured_two_curvy3 +
-  #boxplot1 + boxplot2 + boxplot3 +
   plot_layout(guides='collect', ncol = 3) &
@@
 free(distance_points) /
-  wrap_plots(trimesh_two_curvy, trimesh_two_curvy_removed2, trimesh_two_curvy_removed1, two_curvy_proj_umap_model_delaunay, two_curvy_proj_umap_model_benchmark2, two_curvy_proj_umap_first_model1, ncol = 3) +
+  wrap_plots(trimesh_two_curvy2, trimesh_two_curvy_removed2, trimesh_two_curvy_removed1, two_curvy_proj_umap_model_delaunay, two_curvy_proj_umap_model_benchmark2, two_curvy_proj_umap_model, ncol = 3) +
   plot_annotation(tag_levels = list(c("(a)", "(b)")))
@@
 
-tsne_two_non_linear_diff_shaped_close_clusters <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_62.rds")
-
-nldr_two_non_linear_diff_shaped_close_clusters6 <- tsne_two_non_linear_diff_shaped_close_clusters |>
-  ggplot(aes(x = tSNE1,
-             y = tSNE2))+
-  geom_point(alpha=0.1, size=1, colour='#984ea3') +
-  interior_annotation("f", c(0.08, 0.93)) +
-  theme(aspect.ratio = 1)
-
 umap_two_non_linear_diff_shaped_close_clusters <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_umap_n-neigbors_15_min-dist_0.1.rds")
@@
 error_two_non_linear_diff_shaped_close_clusters_pacmap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_pacmap.rds")
-error_two_non_linear_diff_shaped_close_clusters_tsne2 <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_tsne2.rds")
-
-error_two_non_linear_diff_shaped_close_clusters_tsne2 <- error_two_non_linear_diff_shaped_close_clusters_tsne2 |>
-  mutate(method = "tSNE2")
 
@@
                          error_two_non_linear_diff_shaped_close_clusters_trimap,
-                         error_two_non_linear_diff_shaped_close_clusters_pacmap,
-                         error_two_non_linear_diff_shaped_close_clusters_tsne2)
+                         error_two_non_linear_diff_shaped_close_clusters_pacmap)
 
@@
   scale_x_continuous(breaks = sort(unique(error_two_non_linear_diff_shaped_close_clusters$a1))[seq(1, length(unique(error_two_non_linear_diff_shaped_close_clusters$a1)), by = 5)]) +
-  scale_color_manual(values=c('#e41a1c','#ff7f00','#4daf4a', "#a65628",'#636363', '#984ea3')) +
+  scale_color_manual(values=c('#e41a1c','#ff7f00','#4daf4a', "#a65628",'#636363')) +
   scale_y_log10() +
@@
 ```{r}
-#| fig-cap: "Assessing which of the 6 NLDR layouts on the two C-shaped clusters data is the better representation using MSE for varying binwidth ($a_1$).
+#| fig-cap: "Assessing which of the 5 NLDR layouts on the two C-shaped clusters data is the better representation using MSE for varying binwidth ($a_1$).
 Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-e).
@@
 free(error_plot_two_non_linear_diff_shaped_close_clusters) + wrap_plots(nldr_two_non_linear_diff_shaped_close_clusters1, nldr_two_non_linear_diff_shaped_close_clusters2, nldr_two_non_linear_diff_shaped_close_clusters3,
-                                    nldr_two_non_linear_diff_shaped_close_clusters4, nldr_two_non_linear_diff_shaped_close_clusters5,
-                                    nldr_two_non_linear_diff_shaped_close_clusters6, ncol = 2)
+                                    nldr_two_non_linear_diff_shaped_close_clusters4, nldr_two_non_linear_diff_shaped_close_clusters5, ncol = 2)
 ```
@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 26
+num_bins_x_gau1 <- 20 #13
 lim1 <- gau1_scaled_obj$lim1
@@
 
-# benchmark_gau1 <- 0.1
+benchmark_gau1 <- 0.1
 
@@
 trimesh_removed_gau_tsne <- ggplot() + 
+  #geom_segment(data = tr_from_to_df_gau1, 
+  #             aes(
+  #               x = x_from, 
+  #               y = y_from, 
+  #               xend = x_to, 
+  #               yend = y_to),
+  #             colour = "#000000",
+  #             linewidth = 1) +
   geom_point(
@@
     alpha = 0.2) + 
-  # geom_rect(
-  #     aes(xmin = 0.3, 
-  #         xmax = 0.65, 
-  #         ymin = -0.05, 
-  #         ymax = 0.35),
-  #         fill = "transparent", 
-  #     color = "grey50", 
-  #     linewidth = 0.5) +
+  geom_rect(
+      aes(xmin = 0.3, 
+          xmax = 0.65, 
+          ymin = -0.05, 
+          ymax = 0.35),
+          fill = "transparent", 
+      color = "grey50", 
+      linewidth = 0.5) +
   interior_annotation("a1", 
@@
   proj = projection,
-  limits = 1,
-  axis_pos_x = -0.6,
-  axis_pos_y = -0.63,
+  limits = 0.3,
+  axis_pos_x = 0.05,
+  axis_pos_y = -0.85,
   axis_labels = names(scaled_gau_data),
@@
       y = proj2)) +
-  geom_point(
-    #size = 0.5,
-    alpha = 0.3,
-    color = "#999999") +
   geom_segment(
@@
     color = "#000000",
-    alpha = 0.4,
-    linewidth = 0.1) +
+    linewidth = 1) +
+  geom_point(
+    #size = 0.5,
+    alpha = 0.2,
+    color = "#999999") +
   geom_segment(
@@
   coord_fixed() +
-  xlim(c(-0.84, 0.84)) +
-  ylim(c(-0.84, 0.84)) +
+  xlim(c(0, 0.4)) +
+  ylim(c(-0.9, -0.5)) +
   interior_annotation("a2", 
@@
                       cex = 2)
+
+## Second projection
+# projection <- cbind(
+#     c(-0.71895,0.29250,-0.29905,-0.01656),
+#     c(-0.16698,-0.62707,-0.23756,0.46328))
+projection <- cbind(
+  c(0.36030,0.00630,0.66545,-0.34307),
+  c(-0.01861,0.14409,-0.36798,-0.73066))
+
+projected <- as.matrix(scaled_gau_data) %*% projection
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  #dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+projected_model <- as.matrix(scaled_gau_data_model) %*% projection
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+axes_obj <- gen_axes(  
+  proj = projection,
+  limits = 0.3,
+  axis_pos_x = -0.7,
+  axis_pos_y = -0.7,
+  axis_labels = names(scaled_gau_data),
+  threshold = 0)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
+
+five_gau_proj_tsne_model2 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#000000",
+    linewidth = 1) +
+  geom_point(
+    #size = 0.5,
+    alpha = 0.2,
+    color = "#999999") +
+  geom_segment(
+    data=axes, 
+    aes(x=x1, y=y1, xend=x2, yend=y2), 
+    colour="grey70") +
+  geom_text(
+    data=axes, 
+    aes(x=x2, y=y2), 
+    label=rownames(axes), 
+    colour="grey50",
+    size = 5) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.75, -0.36)) +
+  ylim(c(-0.75, -0.36)) +
+  interior_annotation("a3", 
+                      position = c(0.08, 0.9),
+                      cex = 2)
+
+# five_gau_proj_tsne_model <- gen_proj_langevitour(
+#   points_df = df_exe,
+#   projection = projection,
+#   edge_df = distance_df_small_edges |> select(-distance)
+# ) +
+#   xlim(c(-0.18, 0.1)) +
+#   ylim(c(-0.78, -0.55))
 ```
@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 90
+num_bins_x_gau1 <- 45
 lim1 <- gau1_scaled_obj$lim1
@@
   filter(distance < benchmark_gau1) |>
-  mutate(ID = row_number()) #|>
-  #filter(!(ID %in% c(112)))
+  mutate(ID = row_number()) |>
+  filter(!(ID %in% c(112)))
 
@@
 trimesh_removed_gau_umap <- ggplot() + 
+  #geom_segment(data = tr_from_to_df_gau1, 
+  #             aes(
+  #               x = x_from, 
+  #               y = y_from, 
+  #               xend = x_to, 
+  #               yend = y_to),
+  #             colour = "#000000",
+  #             linewidth = 1) +
   geom_point(
@@
     alpha = 0.05) + 
-  # geom_rect(
-  #   aes(xmin = 0.45, 
-  #       xmax = 0.57, 
-  #       ymin = 0, 
-  #       ymax = 0.1),
-  #   fill = "transparent", 
-  #   color = "grey50", 
-  #   linewidth = 0.5)  +
+  geom_rect(
+    aes(xmin = 0.45, 
+        xmax = 0.57, 
+        ymin = 0, 
+        ymax = 0.1),
+    fill = "transparent", 
+    color = "grey50", 
+    linewidth = 0.5)  +
   interior_annotation("b1", 
@@
   proj = projection,
-  limits = 1,
-  axis_pos_x = -0.6,
-  axis_pos_y = -0.63,
+  limits = 0.28,
+  axis_pos_x = -0.58,
+  axis_pos_y = -0.35,
   axis_labels = names(scaled_gau_data),
@@
     color = "#000000",
-    alpha = 0.4,
-    linewidth = 0.1) +
+    linewidth = 1) +
+  geom_point(
+    #size = 0.5,
+    alpha = 0.3,
+    color = "#999999") +
   geom_segment(
@@
   coord_fixed() +
-  xlim(c(-0.83, 0.83)) +
-  ylim(c(-0.83, 0.83)) +
+  xlim(c(-0.64, -0.29)) +
+  ylim(c(-0.4, -0.05)) +
   interior_annotation("b2", 
@@
                       cex = 2)
-```
-
-```{r}
-#| label: five-gau-proj-pacmap-model
-
-pacmap_data_gau <- read_rds("data/five_gau_clusters/pacmap_data_five_gau.rds")
-gau1_scaled_obj <- gen_scaled_data(
-  data = pacmap_data_gau)
-pacmap_gau_scaled <- gau1_scaled_obj$scaled_nldr |>
-  select(PaCMAP1, PaCMAP2, ID)
-
-pacmap_gau <- pacmap_gau_scaled |>
-  ggplot(aes(x = PaCMAP1,
-             y = PaCMAP2)) +
-  geom_point(alpha=0.3, color = "#000000")
-
-## Compute hexbin parameters
-num_bins_x_gau1 <- 40
-lim1 <- gau1_scaled_obj$lim1
-lim2 <- gau1_scaled_obj$lim2
-r2_gau1 <- diff(lim2)/diff(lim1)
-
-gau1_model <- fit_highd_model(
-  training_data = training_data_gau,
-  emb_df = pacmap_gau_scaled,
-  bin1 = num_bins_x_gau1,
-  r2 = r2_gau1,
-  is_bin_centroid = TRUE,
-  is_rm_lwd_hex = FALSE,
-  col_start_highd = "x",
-  q = 0.1
-)
-
-df_bin_centroids_gau1 <- gau1_model$df_bin_centroids
-df_bin_gau1 <- gau1_model$df_bin
-
-## Compute error
-
-error_df_gau_abs <- augment(
-  df_bin_centroids = df_bin_centroids_gau1,
-  df_bin = df_bin_gau1,
-  training_data = training_data_gau,
-  newdata = NULL,
-  type_NLDR = "PaCMAP",
-  col_start = "x")
-
-error_df_gau_abs <- error_df_gau_abs |>
-  mutate(sqrt_row_wise_abs_error = sqrt(row_wise_abs_error))
-
-error_df_gau_abs <- error_df_gau_abs |>
-  bind_cols(pacmap_gau_scaled |>
-              select(-ID))
-
-error_df_gau_pacmap <- error_df_gau_abs |>
-  ggplot(aes(x = PaCMAP1,
-             y = PaCMAP2,
-             z = sqrt_row_wise_abs_error)) +
-  stat_summary_hex(bins=15) + 
-  xlim(c(-0.02, 0.18)) + ylim(c(0.12, 0.32)) +
-  scale_fill_continuous_sequential(palette = "YlOrRd", breaks = 20) +
-  interior_annotation("c3", 
-                      position = c(0.08, 0.9),
-                      cex = 2)
-
-## Triangulate bin centroids
-tr1_object_gau1 <- tri_bin_centroids(
-  df_bin_centroids_gau1, x = "c_x", y = "c_y")
-tr_from_to_df_gau1 <- gen_edges(
-  tri_object = tr1_object_gau1)
-
-## Compute 2D distances
-distance_gau1 <- cal_2d_dist(
-  tr_coord_df = tr_from_to_df_gau1,
-  start_x = "x_from",
-  start_y = "y_from",
-  end_x = "x_to",
-  end_y = "y_to",
-  select_vars = c("from", "to", "distance"))
-
-## To find the benchmark value
-benchmark_gau1 <- find_lg_benchmark(
-  distance_edges = distance_gau1,
-  distance_col = "distance")
-
-# trimesh_removed_gau1 <- vis_rmlg_mesh(
-#   distance_edges = distance_gau1,
-#   benchmark_value = benchmark_gau1,
-#   tr_coord_df = tr_from_to_df_gau1,
-#   distance_col = "distance")
-
-tr_df <- distinct(tibble(
-  x = c(tr_from_to_df_gau1[["x_from"]], tr_from_to_df_gau1[["x_to"]]), 
-  y = c(tr_from_to_df_gau1[["y_from"]], tr_from_to_df_gau1[["y_to"]])))
-
-distance_df_small_edges_gau1 <- distance_gau1 |>
-  filter(distance < benchmark_gau1) |>
-  mutate(ID = row_number()) |>
-  filter(!(ID %in% c(97)))
-
-tr_from_to_df_gau1 <- inner_join(
-  tr_from_to_df_gau1, distance_df_small_edges_gau1, 
-  by = c("from", "to"))
-
-trimesh_removed_gau_pacmap <- ggplot() + 
-  geom_point(
-    data = pacmap_gau_scaled,
-    aes(
-      x = PaCMAP1,
-      y = PaCMAP2
-    ),
-    alpha = 0.3) +
-  # geom_rect(
-  #   aes(xmin = -0.04, 
-  #       xmax = 0.22, 
-  #       ymin = 0.08, 
-  #       ymax = 0.34),
-  #   fill = "transparent", 
-  #   color = "grey50", 
-  #   linewidth = 0.5) +
-  interior_annotation("c1", 
-                      position = c(0.08, 0.95),
-                      cex = 2) +
-  theme(
-    aspect.ratio = 1
-  )
-
-## Hexagonal binning to have regular hexagons
-hb_obj_gau1 <- hex_binning(
-  data = pacmap_gau_scaled,
-  bin1 = num_bins_x_gau1,
-  r2 = r2_gau1,
-  q = 0.1)
-
-pacmap_data_with_hb_id <- hb_obj_gau1$data_hb_id
-
-df_all_gau1 <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID),
-                                pacmap_data_with_hb_id)
-
-### Define type column
-df <- df_all_gau1 |>
-  dplyr::select(tidyselect::starts_with("x")) |>
-  dplyr::mutate(type = "data") ## original dataset
-
-df_b <- df_bin_gau1 |>
-  dplyr::filter(hb_id %in% df_bin_centroids_gau1$hexID) |>
-  dplyr::mutate(type = "model") ## Data with summarized mean
-
-## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
-df_b <- df_b[match(df_bin_centroids_gau1$hexID, df_b$hb_id),] |>
-  dplyr::select(-hb_id)
-
-# Apply the scaling
-df_model_data <- bind_rows(data_gau, df_b)
-scaled_gau <- scale_data_manual(df_model_data, "type") |>
-  as_tibble()
-
-scaled_gau_data <- scaled_gau |>
-  filter(type == "data") |>
-  select(-type)
-
-scaled_gau_data_model <- scaled_gau |>
-  filter(type == "model") |>
-  select(-type)
-
-## Set the maximum difference as the criteria
-distance_df_small_edges <- distance_gau1 |>
-  dplyr::filter(distance < benchmark_gau1)
-
-## First projection
+
+## Second projection
 projection <- cbind(
-  c(-0.00215,-0.68905,-0.04778,-0.54223),
-  c(0.42558,-0.23854,-0.63659,0.35753))
+  c(0.36030,0.00630,0.66545,-0.34307),
+  c(-0.01861,0.14409,-0.36798,-0.73066))
 
@@
   proj = projection,
-  limits = 1,
-  axis_pos_x = -0.6,
-  axis_pos_y = -0.63,
+  limits = 0.28,
+  axis_pos_x = -0.7,
+  axis_pos_y = -0.7,
   axis_labels = names(scaled_gau_data),
@@
 
-five_gau_proj_pacmap_model1 <- projected_df |>
+five_gau_proj_umap_model2 <- projected_df |>
   ggplot(
@@
   coord_fixed() +
-  xlim(c(-0.83, 0.83)) +
-  ylim(c(-0.83, 0.83)) +
+  xlim(c(-0.75, -0.4)) +
+  ylim(c(-0.75, -0.4)) +
+  interior_annotation("b3", 
+                      position = c(0.08, 0.9),
+                      cex = 2)
+
+# five_gau_proj_umap_model <- gen_proj_langevitour(
+#   points_df = df_exe,
+#   projection = projection,
+#   edge_df = distance_df_small_edges |> select(-distance)
+# ) +
+#   xlim(c(-0.18, 0.1)) +
+#   ylim(c(-0.78, -0.55))
+```
+
+```{r}
+#| label: five-gau-proj-pacmap-model
+
+pacmap_data_gau <- read_rds("data/five_gau_clusters/pacmap_data_five_gau.rds")
+gau1_scaled_obj <- gen_scaled_data(
+  data = pacmap_data_gau)
+pacmap_gau_scaled <- gau1_scaled_obj$scaled_nldr |>
+  select(PaCMAP1, PaCMAP2, ID)
+
+pacmap_gau <- pacmap_gau_scaled |>
+  ggplot(aes(x = PaCMAP1,
+             y = PaCMAP2)) +
+  geom_point(alpha=0.3, color = "#000000")
+
+## Compute hexbin parameters
+num_bins_x_gau1 <- 21
+lim1 <- gau1_scaled_obj$lim1
+lim2 <- gau1_scaled_obj$lim2
+r2_gau1 <- diff(lim2)/diff(lim1)
+
+gau1_model <- fit_highd_model(
+  training_data = training_data_gau,
+  emb_df = pacmap_gau_scaled,
+  bin1 = num_bins_x_gau1,
+  r2 = r2_gau1,
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = FALSE,
+  col_start_highd = "x",
+  q = 0.1
+)
+
+df_bin_centroids_gau1 <- gau1_model$df_bin_centroids
+df_bin_gau1 <- gau1_model$df_bin
+
+## Compute error
+
+error_df_gau_abs <- augment(
+  df_bin_centroids = df_bin_centroids_gau1,
+  df_bin = df_bin_gau1,
+  training_data = training_data_gau,
+  newdata = NULL,
+  type_NLDR = "PaCMAP",
+  col_start = "x")
+
+error_df_gau_abs <- error_df_gau_abs |>
+  mutate(sqrt_row_wise_abs_error = sqrt(row_wise_abs_error))
+
+error_df_gau_abs <- error_df_gau_abs |>
+  bind_cols(pacmap_gau_scaled |>
+              select(-ID))
+
+error_df_gau_pacmap <- error_df_gau_abs |>
+  ggplot(aes(x = PaCMAP1,
+             y = PaCMAP2,
+             z = sqrt_row_wise_abs_error)) +
+  stat_summary_hex(bins=15) + 
+  xlim(c(-0.02, 0.18)) + ylim(c(0.12, 0.32)) +
+  scale_fill_continuous_sequential(palette = "YlOrRd", breaks = 20) +
+  interior_annotation("c3", 
+                      position = c(0.08, 0.9),
+                      cex = 2)
+
+## Triangulate bin centroids
+tr1_object_gau1 <- tri_bin_centroids(
+  df_bin_centroids_gau1, x = "c_x", y = "c_y")
+tr_from_to_df_gau1 <- gen_edges(
+  tri_object = tr1_object_gau1)
+
+## Compute 2D distances
+distance_gau1 <- cal_2d_dist(
+  tr_coord_df = tr_from_to_df_gau1,
+  start_x = "x_from",
+  start_y = "y_from",
+  end_x = "x_to",
+  end_y = "y_to",
+  select_vars = c("from", "to", "distance"))
+
+## To find the benchmark value
+benchmark_gau1 <- find_lg_benchmark(
+  distance_edges = distance_gau1,
+  distance_col = "distance")
+
+# trimesh_removed_gau1 <- vis_rmlg_mesh(
+#   distance_edges = distance_gau1,
+#   benchmark_value = benchmark_gau1,
+#   tr_coord_df = tr_from_to_df_gau1,
+#   distance_col = "distance")
+
+tr_df <- distinct(tibble(
+  x = c(tr_from_to_df_gau1[["x_from"]], tr_from_to_df_gau1[["x_to"]]), 
+  y = c(tr_from_to_df_gau1[["y_from"]], tr_from_to_df_gau1[["y_to"]])))
+
+distance_df_small_edges_gau1 <- distance_gau1 |>
+  filter(distance < benchmark_gau1) |>
+  mutate(ID = row_number()) |>
+  filter(!(ID %in% c(97)))
+
+tr_from_to_df_gau1 <- inner_join(
+  tr_from_to_df_gau1, distance_df_small_edges_gau1, 
+  by = c("from", "to"))
+
+trimesh_removed_gau_pacmap <- ggplot() + 
+  #geom_segment(data = tr_from_to_df_gau1, 
+  #             aes(
+  #               x = x_from, 
+  #               y = y_from, 
+  #               xend = x_to, 
+  #               yend = y_to),
+  #             colour = "#000000",
+  #             linewidth = 1) +
+  geom_point(
+    data = pacmap_gau_scaled,
+    aes(
+      x = PaCMAP1,
+      y = PaCMAP2
+    ),
+    alpha = 0.3) +
+  geom_rect(
+    aes(xmin = -0.04, 
+        xmax = 0.22, 
+        ymin = 0.08, 
+        ymax = 0.34),
+    fill = "transparent", 
+    color = "grey50", 
+    linewidth = 0.5) +
+  interior_annotation("c1", 
+                      position = c(0.08, 0.95),
+                      cex = 2) +
+  theme(
+    aspect.ratio = 1
+  )
+
+## Hexagonal binning to have regular hexagons
+hb_obj_gau1 <- hex_binning(
+  data = pacmap_gau_scaled,
+  bin1 = num_bins_x_gau1,
+  r2 = r2_gau1,
+  q = 0.1)
+
+pacmap_data_with_hb_id <- hb_obj_gau1$data_hb_id
+
+df_all_gau1 <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID),
+                                pacmap_data_with_hb_id)
+
+### Define type column
+df <- df_all_gau1 |>
+  dplyr::select(tidyselect::starts_with("x")) |>
+  dplyr::mutate(type = "data") ## original dataset
+
+df_b <- df_bin_gau1 |>
+  dplyr::filter(hb_id %in% df_bin_centroids_gau1$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b <- df_b[match(df_bin_centroids_gau1$hexID, df_b$hb_id),] |>
+  dplyr::select(-hb_id)
+
+# Apply the scaling
+df_model_data <- bind_rows(data_gau, df_b)
+scaled_gau <- scale_data_manual(df_model_data, "type") |>
+  as_tibble()
+
+scaled_gau_data <- scaled_gau |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_gau_data_model <- scaled_gau |>
+  filter(type == "model") |>
+  select(-type)
+
+## Set the maximum difference as the criteria
+distance_df_small_edges <- distance_gau1 |>
+  dplyr::filter(distance < benchmark_gau1)
+
+## First projection
+projection <- cbind(
+  c(-0.00215,-0.68905,-0.04778,-0.54223),
+  c(0.42558,-0.23854,-0.63659,0.35753))
+
+projected <- as.matrix(scaled_gau_data) %*% projection
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  #dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+projected_model <- as.matrix(scaled_gau_data_model) %*% projection
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+axes_obj <- gen_axes(  
+  proj = projection,
+  limits = 0.25,
+  axis_pos_x = -0.41,
+  axis_pos_y = 0.23,
+  axis_labels = names(scaled_gau_data),
+  threshold = 0)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
+
+five_gau_proj_pacmap_model1 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#000000",
+    linewidth = 1) +
+  geom_point(
+    #size = 0.5,
+    alpha = 0.3,
+    color = "#999999") +
+  geom_segment(
+    data=axes, 
+    aes(x=x1, y=y1, xend=x2, yend=y2), 
+    colour="grey70") +
+  geom_text(
+    data=axes, 
+    aes(x=x2, y=y2, label=rownames(axes)), 
+    colour="grey50",
+    size = 5) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.46, -0.12)) +
+  ylim(c(0.18, 0.52)) +
   interior_annotation("c2", 
+                      position = c(0.08, 0.9),
+                      cex = 2)
+
+### Second projection
+# projection <- cbind(
+#     c(-0.60995,0.33361,0.32920,-0.31518),
+#     c(0.19525,-0.30554,0.74417,0.07600))
+projection <- cbind(
+  c(0.36030,0.00630,0.66545,-0.34307),
+  c(-0.01861,0.14409,-0.36798,-0.73066))
+
+projected <- as.matrix(scaled_gau_data) %*% projection
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  #dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+projected_model <- as.matrix(scaled_gau_data_model) %*% projection
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+axes_obj <- gen_axes(  
+  proj = projection,
+  limits = 0.3,
+  axis_pos_x = -0.7,
+  axis_pos_y = -0.7,
+  axis_labels = names(scaled_gau_data),
+  threshold = 0)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
+
+five_gau_proj_pacmap_model2 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#000000",
+    linewidth = 1) +
+  geom_point(
+    #size = 0.8,
+    alpha = 0.2,
+    color = "#999999") +
+  geom_segment(
+    data=axes, 
+    aes(x=x1, y=y1, xend=x2, yend=y2), 
+    colour="grey70") +
+  geom_text(
+    data=axes, 
+    aes(x=x2, y=y2, label=rownames(axes)), 
+    colour="grey50",
+    size = 5) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.75, -0.35)) +
+  ylim(c(-0.75, -0.35))+
+  interior_annotation("c3", 
                       position = c(0.08, 0.9),
@@
 #| fig-pos: H
-#| fig-cap: "NLDR's organise points in the \\gD{} layout in different ways, possibly misleadingly, illustrated using three layouts: (a) tSNE, (b) UMAP, (c) PaCMAP.
+#| fig-cap: "NLDR's organise points in the \\gD{} layout in different ways, possibly misleadingly, illustrated using three layouts: (a) t-SNE, (b) UMAP, (c) PaCMap.
 The data has five Gaussian cluster in $4\\text{-}D$.
@@
 
-```{r}
-# gen_proj_five_clust()[[1]] + gen_proj_five_clust()[[2]] + plot_layout(nrow = 2)
-```
-
 ### The effect of density
@@
 #| label: fig-one-dens_clust-error
-#| fig-height: 10
-#| fig-width: 10
+#| fig-height: 3
+#| fig-width: 9
 #| fig-pos: H

### 2025-01-12 paper.qmd
Updated ordering of points symbols

--- 2025-01-12 paper.qmd
@@
 
-DHC: Will need to have more bins in each of the methods to illustrate this properly.
-
-To illustrate a difference in how the methods organise points in the NLDR layout, simulated 4D data having five Gaussian clusters is used.
-@fig-five-gau-projs a1, b1, c1 show the 2D layouts for (a) t-SNE, (b) UMAP, and (c) PaCMAP, respectively.
+<!-- DHC: Will need to have more bins in each of the methods to illustrate this properly. -->
+
+To illustrate a difference in how the methods organise points in the NLDR layout, simulated $4\text{-}D$ data having five Gaussian clusters is used.
+@fig-five-gau-projs a1, b1, c1 show the \gD{} layouts for (a) tSNE, (b) UMAP, and (c) PaCMAP, respectively.
 The default hyper-parameters are used.
@@
 The models are fitted to these layouts, but we focus on a single cluster to illustrate the curious detail.
-@fig-five-gau-projs a2, b2, c2 show the fitted models in a projection of the 4D space.
+@fig-five-gau-projs a2, b2, c2 show the fitted models in a projection of the $4\text{-}D$ space.
 In this projection the difference between methods can be seen.
-These clusters are fully 4D in nature, so we would expect the model to be a *crumpled 2D sheet* that stretches in all four dimensions.
-This is what is observed for t-SNE and UMAP.
+These clusters are fully $4\text{-}D$ in nature, so we would expect the model to be a *crumpled \gD{} sheet* that stretches in all four dimensions.
+This is what is observed for tSNE and UMAP.
 The curious detail is that the model for PaCMAP is closer to a *pancake* in shape!
-What this means is that there has to be some ordering of points in the 2D PaCMAP layout that induces the flat model, likely that the points are organised by the global principal components.
-So the layout of the model in 4D doesn't reflect the dimension of each cluster.
+What this means is that there has to be some ordering of points in the \gD{} PaCMAP layout that induces the flat model, likely that the points are organised by the global principal components.
+So the layout of the model in $4\text{-}D$ doesn't reflect the dimension of each cluster.
 DHC: is there something in the documentation that might help to explain what has happened?
@@
 This pattern can also be observed by examining the error from the model fit (@fig-five-gau-projs a3, b3, c3).
-With t-SNE and UMAP the error is relatively uniformly distributed, but with PaCMAP there is more error in the center of the 2D layout, reflecting that there are more points in the middle that are further from the fitted model.
+With tSNE and UMAP the error is relatively uniformly distributed, but with PaCMAP there is more error in the center of the \gD{} layout, reflecting that there are more points in the middle that are further from the fitted model.
 
@@
 #| fig-cap: "The sparse end of the structure exhibits a high model error, while the dense end shows a low model error.
-The tSNE layouts of the C-shaped cluster, with uniform points (a1) and dense points (b1), along with their selected projections (a2 and b2), are displayed.
+The tSNE layouts of the C-shaped cluster, with uniform points (a1) and dense points (b1) are colored according to their high-dimensional error, along with their selected projections (a2 and b2), are displayed.
 Darker colors represent a high error, while lighter colors indicate a low error.

### 2025-01-12 paper.tex
Updated ordering of points symbols

@@
 
-DHC: Will need to have more bins in each of the methods to illustrate
-this properly.
-
 To illustrate a difference in how the methods organise points in the
-NLDR layout, simulated 4D data having five Gaussian clusters is used.
-Figure~\ref{fig-five-gau-projs} a1, b1, c1 show the 2D layouts for (a)
-t-SNE, (b) UMAP, and (c) PaCMAP, respectively. The default
-hyper-parameters are used. All three methods show the five clusters,
-with varying degrees of separation.
+NLDR layout, simulated \(4\text{-}D\) data having five Gaussian clusters
+is used. Figure~\ref{fig-five-gau-projs} a1, b1, c1 show the \gD{}
+layouts for (a) tSNE, (b) UMAP, and (c) PaCMAP, respectively. The
+default hyper-parameters are used. All three methods show the five
+clusters, with varying degrees of separation.
 
@@
 to illustrate the curious detail. Figure~\ref{fig-five-gau-projs} a2,
-b2, c2 show the fitted models in a projection of the 4D space. In this
-projection the difference between methods can be seen. These clusters
-are fully 4D in nature, so we would expect the model to be a
-\emph{crumpled 2D sheet} that stretches in all four dimensions. This is
-what is observed for t-SNE and UMAP. The curious detail is that the
-model for PaCMAP is closer to a \emph{pancake} in shape! What this means
-is that there has to be some ordering of points in the 2D PaCMAP layout
-that induces the flat model, likely that the points are organised by the
-global principal components. So the layout of the model in 4D doesn't
-reflect the dimension of each cluster. DHC: is there something in the
-documentation that might help to explain what has happened?
+b2, c2 show the fitted models in a projection of the \(4\text{-}D\)
+space. In this projection the difference between methods can be seen.
+These clusters are fully \(4\text{-}D\) in nature, so we would expect
+the model to be a \emph{crumpled \gD{} sheet} that stretches in all four
+dimensions. This is what is observed for tSNE and UMAP. The curious
+detail is that the model for PaCMAP is closer to a \emph{pancake} in
+shape! What this means is that there has to be some ordering of points
+in the \gD{} PaCMAP layout that induces the flat model, likely that the
+points are organised by the global principal components. So the layout
+of the model in \(4\text{-}D\) doesn't reflect the dimension of each
+cluster. DHC: is there something in the documentation that might help to
+explain what has happened?
 
 This pattern can also be observed by examining the error from the model
-fit (Figure~\ref{fig-five-gau-projs} a3, b3, c3). With t-SNE and UMAP
-the error is relatively uniformly distributed, but with PaCMAP there is
-more error in the center of the 2D layout, reflecting that there are
-more points in the middle that are further from the fitted model.
+fit (Figure~\ref{fig-five-gau-projs} a3, b3, c3). With tSNE and UMAP the
+error is relatively uniformly distributed, but with PaCMAP there is more
+error in the center of the \gD{} layout, reflecting that there are more
+points in the middle that are further from the fitted model.
 
@@
 error. The tSNE layouts of the C-shaped cluster, with uniform points
-(a1) and dense points (b1), along with their selected projections (a2
-and b2), are displayed. Darker colors represent a high error, while
-lighter colors indicate a low error. The uniformly distributed points
+(a1) and dense points (b1) are colored according to their
+high-dimensional error, along with their selected projections (a2 and
+b2), are displayed. Darker colors represent a high error, while lighter
+colors indicate a low error. The uniformly distributed points
 demonstrate random model errors throughout the structure, whereas the

### 2025-01-12 paper.qmd
Commented error description from 6.1

@@
 
-This pattern can also be observed by examining the error from the model fit (@fig-five-gau-projs a3, b3, c3).
+<!-- This pattern can also be observed by examining the error from the model fit (@fig-five-gau-projs a3, b3, c3).
 With tSNE and UMAP the error is relatively uniformly distributed, but with PaCMAP there is more error in the center of the \gD{} layout, reflecting that there are more points in the middle that are further from the fitted model.
-
-<!-- In this section, the evaluation focuses on how effectively the tSNE, UMAP, and PaCMAP algorithms preserve the clustering structure.
-The data consists of five $4\text{-}D$ clusters, each with $1000$ points generated from a multivariate normal distribution with a small variance of $0.0025$.
-The first cluster is centered at $(0, 0, 0, 0)$, the second at $(1, 0, 0, 0)$, the third at $(0, 1, 0, 0)$, the fourth at $(0, 0, 1, 0)$, and the fifth at $(0, 0, 0, 1)$.-->
-
-<!-- In this section, we will assess how tSNE, UMAP, and PaCMAP algorithms preserve the clustering structure.
-We use a simulated dataset consisting of five spherical Gaussian clusters where each cluster occupies a different corner of the $4\text{-}D$ space, with each cluster containing an equal number of points ($1000$) and the same within-cluster variation.
-The data simulated from $\Sigma = \begin{pmatrix} -->
-<!-- 0.0025 & 0 & 0 & 0 \\ -->
-<!-- 0 & 0.0025 & 0 & 0 \\ -->
-<!-- 0 & 0 & 0.0025 & 0 \\ -->
-<!-- 0 & 0 & 0 & 0.0025 \\ -->
-<!-- \end{pmatrix}$. -->
-
-<!-- In the tSNE layout (@fig-five-gau-projs (a1)), the distances between clusters appear smaller, leading to more tightly packed clusters, which is typical of tSNE's focus on preserving local structures.
-This often results in less emphasis on the global distances between clusters.
-On the other hand, UMAP (@fig-five-gau-projs (b1)) tends to maintain larger separations between clusters, better reflecting the global structure and preserving the relative distances between clusters more effectively.
-In contrast, PaCMAP (@fig-five-gau-projs (c1)) aim to balance these aspects by maintaining the distance between clusters and accurately positioning them, effectively capturing both global and local structures.-->
-
-<!-- By visualizing the model of different layouts in $4\text{-}D$ reveal different characteristics of each NLDR method.
-When examining the model of a specific cluster, which appears more like a filled-out shape in tSNE (@fig-five-gau-projs (a2, a3)), and UMAP (@fig-five-gau-projs (b2, b3)), while the model has a pancake shape in PaCMAP (@fig-five-gau-projs (c2, c3)).
-The reason to have filled-out shaped clusters in tSNE and UMAP is that the ability of capture the local structure.
-On the other hand, the reason for the pancake shape in PaCMAP is the fail to capture the local structure.
--->
-
-<!-- The tSNE, UMAP, and PaCMAP layouts each reveal different characteristics of the clustering structure within the dataset.  -->
-
-<!-- When examining the model of a specific cluster, which appears more like a filled-out pancake in $4\text{-}D$, we observe that tSNE, UMAP, and PaCMAP all effectively capture global structures.
-However, there is evidence suggesting that PaCMAP, with its default hyper-parameter settings, struggles to preserve the local structure as effectively as tSNE and UMAP.
--->
-
-<!-- The differences in performance can be attributed to several factors.
-One key element is the choice of hyper-parameters, particularly the number of neighbors considered during the dimensionality reduction process.
-This parameter significantly impacts how well local structures are maintained.
-Additionally, the balance between global and local structures, and issues related to optimization and initialization all contribute to the observed outcomes.
-
-It's important to note that these layouts are made with default hyper-parameter settings.
-Adjusting these parameters could lead to different results, potentially improving PaCMAP's ability to capture local structures or altering how tSNE and UMAP handle global and local structures.
-Thus, while PaCMAP generally balances global and local preservation, careful tuning of its parameters is important for achieving the best results in any given dataset.-->
-
-<!-- In this section, the effectiveness of the algorithm is described using a simulated dataset.
-The dataset consists of five spherical Gaussian clusters in $4\text{-}D$, with each cluster containing an equal number of points and the same within-cluster variation.
--->
-
-<!-- XXX Add tSNE, UMAP, and PaCMAP NLDR layouts and describe what is happening -->
-<!-- XXX Discuss the models of one specific cluster (filled out, pancakes): tSNE, UMAP, PaCMAP are useful to capture global structures, but this is an evidence where PaCMAP fail to capture the local structure with default hyper-parameter settings than tSNE and UMAP -->
-<!-- XXX Why?
-hyper-parameter choice specially number of neighbors, data characteristics (intrinsic dimensionality), imbalance between global and local structures, optimization and initialization issues -->
-<!-- XXX Add this is with specific hyper-parameter choices -->
-
-<!-- The \gD{} layouts generated by tSNE, UMAP, and PaCMAP show five well-separated clusters which evident these methods effectively preserve the global structure.
-In tSNE (@fig-gau-tsne-sc (a)), these clusters appear closely.
-UMAP arranges all clusters in a parallel manner, with three aligned in one line and the other two in a separate line (@fig-gau-umap-sc (a)).
-In contrast, PaCMAP shows one central cluster and the remaining four spread out in different directions (@fig-gau-pacmap-sc (a)).
--->
-
-<!-- The tSNE and UMAP shows *filled out* clusters which provide evidence that these methods preserve the local structure (@fig-gau-tsne-sc (c) and @fig-gau-umap-sc (c)).
-On the other hand, PaCMAP shows *flat* shapes clusters in the model and evident that PaCMAP fail to capture the within-cluster variation (@fig-gau-pacmap-sc (c)).
 -->
@@
 
-<!-- ::: {#fig-gau-tsne-sc layout-ncol="3" fig-pos="H"} -->
-<!-- ![](figures/five_gau_clusters/tsne_layout.png) -->
-
-<!-- ![](figures/five_gau_clusters/2d_model_tsne.png) -->
-
-<!-- ![](figures/five_gau_clusters/sc_tsne_3.png) -->
-
-<!-- The tSNE layout, model in \gD{}, and a view of the fit in projections from $4\text{-}D$, for the five Gaussian cluster data.
-The model fits the separation and tries to *filled out* the clusters.
-(The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/RASEE7N5MbM>).
--->
-<!-- ::: -->
-
-<!-- ::: {#fig-gau-umap-sc layout-ncol="3" fig-pos="H"} -->
-<!-- ![](figures/five_gau_clusters/umap_layout.png) -->
-
-<!-- ![](figures/five_gau_clusters/2d_model_umap.png) -->
-
-<!-- ![](figures/five_gau_clusters/sc_umap_3.png) -->
-
-<!-- The UMAP layout, model in \gD{}, and a view of the fit in projections from $4\text{-}D$, for the five Gaussian cluster data.
-The model fits the separation and tries to *filled out* the clusters, but not as much as tSNE.
-(The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/iG4bCPkJilw>).
--->
-<!-- ::: -->
-
-<!-- ::: {#fig-gau-pacmap-sc layout-ncol="3" fig-pos="H"} -->
-<!-- ![](figures/five_gau_clusters/pacmap_layout.png) -->
-
-<!-- ![](figures/five_gau_clusters/2d_model_pacmap.png) -->
-
-<!-- ![](figures/five_gau_clusters/sc_pacmap_2.png) -->
-
-<!-- The PaCMAP layout, model in \gD{}, and a view of the fit in projections from $4\text{-}D$, for the five Gaussian cluster data.
-The model fits the separation and shows *flat* shaped clusters.
-(The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/z07cKXi8EJQ>).
--->
-<!-- ::: -->
-
-```{r}
-# gen_proj_five_clust()[[1]] + gen_proj_five_clust()[[2]] + plot_layout(nrow = 2)
-```
-
 ### The effect of density
 
-To illustrate how the methods organize difference number of points within the structure in the NLDR layout, simulated $7\text{-}D$ data having C-shaped structure is used.
+To illustrate how tSNE organize different number of points within the structure in the tSNE layout, simulated $7\text{-}D$ data having C-shaped structure is used.
+@fig-one-dens_clust-error a1 and b1 shows \gD{} layouts for tSNE.
+The default hyper-parameters are used.
+
+The models are fitted to these layouts.
+@fig-one-dens_clust-error b1 and b2 show the fitted models in projections of the $7\text{-}D$ space.
+In these projection the difference between the distribution of high-dimensional model error can be seen.
+The uniform densed C-shaped cluster
 

### 2025-01-12 paper.tex
Commented error description from 6.1

@@
 
-This pattern can also be observed by examining the error from the model
-fit (Figure~\ref{fig-five-gau-projs} a3, b3, c3). With tSNE and UMAP the
-error is relatively uniformly distributed, but with PaCMAP there is more
-error in the center of the \gD{} layout, reflecting that there are more
-points in the middle that are further from the fitted model.
-
 \begin{figure}[H]
@@
 
-To illustrate how the methods organize difference number of points
-within the structure in the NLDR layout, simulated \(7\text{-}D\) data
-having C-shaped structure is used.
+To illustrate how tSNE organize different number of points within the
+structure in the tSNE layout, simulated \(7\text{-}D\) data having
+C-shaped structure is used. Figure~\ref{fig-one-dens_clust-error} a1 and
+b1 shows \gD{} layouts for tSNE. The default hyper-parameters are used.
+
+The models are fitted to these layouts.
+Figure~\ref{fig-one-dens_clust-error} b1 and b2 show the fitted models
+in projections of the \(7\text{-}D\) space. In these projection the
+difference between the distribution of high-dimensional model error can
+be seen. The uniform densed C-shaped cluster
 

### 2025-01-12 paper.qmd

--- 2025-01-12 paper.qmd
@@
 #| label: fig-rm-lg
-#| fig-cap: "Distribution of edge lengths and wireframes in \\gD{} and $7\\text{-}D$ with Delaunay triangulation, and two choices of long edge removal: (b) benchmark = 2.5$a_1$ and (c) benchmark = 2$a_1$, where $a_1$ = 0.09."
+#| fig-cap: "XXX Shouldn't the middle one have an edge connecting inside a C, so that it is clearly not as good as c?
+XXX Distribution of edge lengths and wireframes in \\gD{} and $7\\text{-}D$ with Delaunay triangulation, and two choices of long edge removal: (b) benchmark = 2.5$a_1$ and (c) benchmark = 2$a_1$, where $a_1$ = 0.09."
 #| fig-height: 15
@@
 
-<!-- DHC: Will need to have more bins in each of the methods to illustrate this properly. -->
-
-To illustrate a difference in how the methods organise points in the NLDR layout, simulated $4\text{-}D$ data having five Gaussian clusters is used.
-@fig-five-gau-projs a1, b1, c1 show the \gD{} layouts for (a) tSNE, (b) UMAP, and (c) PaCMAP, respectively.
+DHC: Will need to have more bins in each of the methods to illustrate this properly.
+
+To illustrate a difference in how the methods organise points in the NLDR layout, simulated 4D data having five Gaussian clusters is used.
+@fig-five-gau-projs a1, b1, c1 show the 2D layouts for (a) t-SNE, (b) UMAP, and (c) PaCMAP, respectively.
 The default hyper-parameters are used.
@@
 The models are fitted to these layouts, but we focus on a single cluster to illustrate the curious detail.
-@fig-five-gau-projs a2, b2, c2 show the fitted models in a projection of the $4\text{-}D$ space.
+@fig-five-gau-projs a2, b2, c2 show the fitted models in a projection of the 4D space.
 In this projection the difference between methods can be seen.
-These clusters are fully $4\text{-}D$ in nature, so we would expect the model to be a *crumpled \gD{} sheet* that stretches in all four dimensions.
-This is what is observed for tSNE and UMAP.
+These clusters are fully 4D in nature, so we would expect the model to be a *crumpled 2D sheet* that stretches in all four dimensions.
+This is what is observed for t-SNE and UMAP.
 The curious detail is that the model for PaCMAP is closer to a *pancake* in shape!
-What this means is that there has to be some ordering of points in the \gD{} PaCMAP layout that induces the flat model, likely that the points are organised by the global principal components.
-So the layout of the model in $4\text{-}D$ doesn't reflect the dimension of each cluster.
+What this means is that there has to be some ordering of points in the 2D PaCMAP layout that induces the flat model, likely that the points are organised by the global principal components.
+So the layout of the model in 4D doesn't reflect the dimension of each cluster.
 DHC: is there something in the documentation that might help to explain what has happened?
@@
 This pattern can also be observed by examining the error from the model fit (@fig-five-gau-projs a3, b3, c3).
-With tSNE and UMAP the error is relatively uniformly distributed, but with PaCMAP there is more error in the center of the \gD{} layout, reflecting that there are more points in the middle that are further from the fitted model.
+With t-SNE and UMAP the error is relatively uniformly distributed, but with PaCMAP there is more error in the center of the 2D layout, reflecting that there are more points in the middle that are further from the fitted model.
 
@@
 #| fig-cap: "The sparse end of the structure exhibits a high model error, while the dense end shows a low model error.
-The tSNE layouts of the C-shaped cluster, with uniform points (a1) and dense points (b1) are colored according to their high-dimensional error, along with their selected projections (a2 and b2), are displayed.
+The tSNE layouts of the C-shaped cluster, with uniform points (a1) and dense points (b1), along with their selected projections (a2 and b2), are displayed.
 Darker colors represent a high error, while lighter colors indicate a low error.
@@
 
-### Single-cell gene expression
+XXX WHY THESE EXAMPLES. NEEDS AN OVERVIEW PARAGRAPH
+
+### Single-cell gene expression {#sec-pbmc}
 <!--
@@
 
-In the field of single-cell studies, a common analytical task involves clustering to identify groups of cells with similar expression profiles.
-NLDR methods are commonly used to display clusters, and help to verify the results.
-For example, @chen2023 illustrates the use of UMAP to identify clusters in Human Peripheral Blood Mononuclear Cells (PBMC3k).
-There are $2622$ single cells.
-First 9 principal components are used to generate the UMAP.
-@fig-NLDR-variety (a) is the reproduction of the published plot.
-The objective is to assess the published layout, and if it does not accurately represent the three clusters with small separations of the PBMC3k dataset (@fig-model-pbmc-author-proj (a2)), then select a reasonable \gD{} layout.
+XXX What is single cell data????
+
+Clustering of single-cell data is used to identify groups of cells with similar expression profiles.
+NLDR often used to summarise the discovered clusters, and help to understand the results.
+The purpose of this example is to *illustrate how to use our method to help decide on an appropriate NLDR layout that accurately represents the data*.
+The cluster results of Human Peripheral Blood Mononuclear Cells (PBMC3k) in @chen2023 are examined.
+There are $2622$ single cells, with XXX gene expressions.
+Following their pre-processing, UMAP was performed using nine principal components.
+@fig-NLDR-variety (a) is the reproduction of the published plot, with colour indicating the clustering results overlaid on the NLDR layout.
+The question is whether this accurately represents the cluster structure in the data.
+Our method can help here, and also help to provide a better \gD{} layout, as needed.
 
@@
 As a result, it is necessary to find an appropriate layout for the dataset.
-MSE for different binwidths ($a_1$) using tSNE, UMAP, PHATE, PaCMAP, and TriMAP with various (hyper-)parameter settings were computed (@fig-pbmc-mse).
+MSE for different binwidths ($a_1$) using tSNE, UMAP, PHATE, PaCMAP, and TriMAP with various (hyper-)~parameter settings were computed (@fig-pbmc-mse).
 Layouts c, d, and e, which show small separations between clusters, are universally optimal.
@@
 #| echo: false
-#| fig-cap: "Model in \\gD{}, on the layout a and b, and two views of the fit in projections from $9\\text{-}D$, for the PBMC3k data ($n =  2622$ and $p = 9$).
+#| fig-cap: "XXX Number of bins for b is inadequate.
+Caption needs to start with what this figure is for XXX Model in \\gD{}, on the layout a and b, and two views of the fit in projections from $9\\text{-}D$, for the PBMC3k data ($n =  2622$ and $p = 9$).
 Layout a shows three-well separated clusters with big separation, while layout b shows close clusters.
@@
 
-In the field of single-cell studies, clustering is a common analytical task used to identify groups of cells with similar expression profiles.
+<!--
+XXX What this section needs:
+
+- connect this with previous section, not make it seem like a completely new. connections are: nldr used to show clustering, scDEED provides a different approach to decide which is best.
+- why these two? assume one is the published figure, and second is one recommended by scDEED, which our results confirm is better.
+- Refer to figure in text
+- Fig 16 not necessary
+
+CHANGES WERE MADE BASED ON THIS THINKING
+-->
+
+<!-- In the field of single-cell studies, clustering is a common analytical task used to identify groups of cells with similar expression profiles.
 Non-linear dimensional reduction (NLDR) methods are frequently employed to visualize these clusters and help validate the results.
 However, it is well known that the 2D embeddings produced by t-SNE and UMAP may not accurately reflect the similarities among cell clusters.
-
-To address this challenge, @xia2023 introduces a statistical method called scDEED, which detects unreliable cell embeddings generated by \gD{} embedding techniques.
+ALREADY STATED IN PREVIOUS SECTION -->
+
+As we were writing this paper @xia2023 appeared proposing a new method called scDEED helping to assess the validity of a \gD{} embedding.
 scDEED calculates a reliability score for each cell embedding based on the similarity between the cell’s \gD{} embedding neighbors and its neighbors prior to embedding.
-By identifying embeddings with low reliability scores as dubious and those with high scores as trustworthy, scDEED minimizes the number of questionable cell embeddings.
-This approach also provides clear guidance for optimizing the hyperparameters of an embedding method.
-
-To verify that the \gD{} cell embeddings became better aligned with the cell types after scDEED’s optimization, the Human Peripheral Blood Mononuclear Cells (PBMC) dataset was used.
-It contained $31,021$ cells with cell type labels, and the gene expression levels were in the unit of log-transformed UMI count per $10,000$.
+A low reliability score suggests a dubious embedding.
+It can help in the deciding on optimal hyperparameters.
+Here we illustrate how our method compares with the results from scDEED
+
+Following the process in @xia2023 again using the Human Peripheral Blood Mononuclear Cells (PBMC) data.
+Note that @xia2023 uses a different subset from @sec-pbmc, which contains $31,021$ cells including cell type labels, and the gene expression levels were in the unit of log-transformed UMI count per $10,000$.
 They focused on three sequencing methods (inDrops, DropSeq, and SeqWell) and four common cell types Cytotoxic T cell, CD4+T cell, CD14+ Monocyte, and B cell.
@@
 The objective is to assess the optimized layout by scDEED, and if it does not accurately represent the three clusters with small separations of the PBMC dataset, then select a reasonable \gD{} layout.
+
+@fig-model-pbmc-author-proj-scdeed shows layout from the original published article along with that suggested by scDEED to be more accurate.
+The log MSE vs binwidth plot illustrates that our approach would suggest that scDEED is correct here, that layout b is more accurately reflecting the cluster structure in the data.
 
@@
 Colour used for the lines and points in the left plot and in the scatterplots represents UMAP layout (a, b).
-Layout b is universally suboptimal.
-Layout b is the best choice."
+Of the two, layout b is optimal across all binwidths making it the best choice."
 
@@
 #| echo: false
+#| eval: false
 #| fig-cap: "Model in \\gD{}, on the layout a, and two views of the fit in projections from $50\\text{-}D$, for the PBMC data ($n =  5858$ and $p = 50$).

### 2025-01-12 paper.tex

--- 2025-01-12 paper.tex
@@
 
-\includegraphics[width=0.6\textwidth,height=\textheight]{paper_files/figure-pdf/fig-p-d-error-in-2d-two-curvy-1.pdf}
+\includegraphics[width=0.6\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-p-d-error-in-2d-two-curvy-1.pdf}
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-rm-lg-1.pdf}
-
-}
-
-\caption{\label{fig-rm-lg}Distribution of edge lengths and wireframes in
-\gD{} and \(7\text{-}D\) with Delaunay triangulation, and two choices of
-long edge removal: (b) benchmark = 2.5\(a_1\) and (c) benchmark =
-2\(a_1\), where \(a_1\) = 0.09.}
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-rm-lg-1.pdf}
+
+}
+
+\caption{\label{fig-rm-lg}XXX Shouldn't the middle one have an edge
+connecting inside a C, so that it is clearly not as good as c? XXX
+Distribution of edge lengths and wireframes in \gD{} and \(7\text{-}D\)
+with Delaunay triangulation, and two choices of long edge removal: (b)
+benchmark = 2.5\(a_1\) and (c) benchmark = 2\(a_1\), where \(a_1\) =
+0.09.}
 
@@
 
-\subsection{Single-cell gene
-expression}\label{single-cell-gene-expression}
-
-In the field of single-cell studies, a common analytical task involves
-clustering to identify groups of cells with similar expression profiles.
-NLDR methods are commonly used to display clusters, and help to verify
-the results. For example, \citet{chen2023} illustrates the use of UMAP
-to identify clusters in Human Peripheral Blood Mononuclear Cells
-(PBMC3k). There are \(2622\) single cells. First 9 principal components
-are used to generate the UMAP. Figure~\ref{fig-NLDR-variety} (a) is the
-reproduction of the published plot. The objective is to assess the
-published layout, and if it does not accurately represent the three
-clusters with small separations of the PBMC3k dataset
-(Figure~\ref{fig-model-pbmc-author-proj} (a2)), then select a reasonable
-\gD{} layout.
+XXX WHY THESE EXAMPLES. NEEDS AN OVERVIEW PARAGRAPH
+
+\subsection{Single-cell gene expression}\label{sec-pbmc}
+
+XXX What is single cell data????
+
+Clustering of single-cell data is used to identify groups of cells with
+similar expression profiles. NLDR often used to summarise the discovered
+clusters, and help to understand the results. The purpose of this
+example is to \emph{illustrate how to use our method to help decide on
+an appropriate NLDR layout that accurately represents the data}. The
+cluster results of Human Peripheral Blood Mononuclear Cells (PBMC3k) in
+\citet{chen2023} are examined. There are \(2622\) single cells, with XXX
+gene expressions. Following their pre-processing, UMAP was performed
+using nine principal components. Figure~\ref{fig-NLDR-variety} (a) is
+the reproduction of the published plot, with colour indicating the
+clustering results overlaid on the NLDR layout. The question is whether
+this accurately represents the cluster structure in the data. Our method
+can help here, and also help to provide a better \gD{} layout, as
+needed.
 
@@
 dataset. MSE for different binwidths (\(a_1\)) using tSNE, UMAP, PHATE,
-PaCMAP, and TriMAP with various (hyper-)parameter settings were computed
-(Figure~\ref{fig-pbmc-mse}). Layouts c, d, and e, which show small
-separations between clusters, are universally optimal. However, layout d
-performs well with smaller binwidths and poorly with larger binwidths.
-On the other hand, layout e performs well with larger binwidths. Layout
-c was selected for further analysis due to its better (hyper-)parameter
-selection for the same method of the published plot.
+PaCMAP, and TriMAP with various (hyper-)\textasciitilde parameter
+settings were computed (Figure~\ref{fig-pbmc-mse}). Layouts c, d, and e,
+which show small separations between clusters, are universally optimal.
+However, layout d performs well with smaller binwidths and poorly with
+larger binwidths. On the other hand, layout e performs well with larger
+binwidths. Layout c was selected for further analysis due to its better
+(hyper-)parameter selection for the same method of the published plot.
 
@@
 
-\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-model-pbmc-author-proj-1.pdf}
-
-}
-
-\caption{\label{fig-model-pbmc-author-proj}Model in \gD{}, on the layout
-a and b, and two views of the fit in projections from \(9\text{-}D\),
-for the PBMC3k data (\(n =  2622\) and \(p = 9\)). Layout a shows
-three-well separated clusters with big separation, while layout b shows
-close clusters. In \(9\text{-}D\), the data shows three clusters with
-two of them being close and one separated from the others. This is why
-b1, b2, and b3 shows a connected edge between the close clusters, as
-proven by Figure~\ref{fig-pbmc-mse} as a reasonable layout. Viewing the
-fitted model in \(9\text{-}D\), helps to see some unobserved patterns of
-the data: (i) dense points, and (ii) non-linear clusters. Videos of the
-langevitour animations are available at
-\url{https://youtu.be/0cKX_HG_n0k} and
+\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-model-pbmc-author-proj-1.pdf}
+
+}
+
+\caption{\label{fig-model-pbmc-author-proj}XXX Number of bins for b is
+inadequate. Caption needs to start with what this figure is for XXX
+Model in \gD{}, on the layout a and b, and two views of the fit in
+projections from \(9\text{-}D\), for the PBMC3k data (\(n =  2622\) and
+\(p = 9\)). Layout a shows three-well separated clusters with big
+separation, while layout b shows close clusters. In \(9\text{-}D\), the
+data shows three clusters with two of them being close and one separated
+from the others. This is why b1, b2, and b3 shows a connected edge
+between the close clusters, as proven by Figure~\ref{fig-pbmc-mse} as a
+reasonable layout. Viewing the fitted model in \(9\text{-}D\), helps to
+see some unobserved patterns of the data: (i) dense points, and (ii)
+non-linear clusters. Videos of the langevitour animations are available
+at \url{https://youtu.be/0cKX_HG_n0k} and
 \url{https://youtu.be/KhJvsRtaX04} respectively.}
@@
 
-In the field of single-cell studies, clustering is a common analytical
-task used to identify groups of cells with similar expression profiles.
-Non-linear dimensional reduction (NLDR) methods are frequently employed
-to visualize these clusters and help validate the results. However, it
-is well known that the 2D embeddings produced by t-SNE and UMAP may not
-accurately reflect the similarities among cell clusters.
-
-To address this challenge, \citet{xia2023} introduces a statistical
-method called scDEED, which detects unreliable cell embeddings generated
-by \gD{} embedding techniques. scDEED calculates a reliability score for
-each cell embedding based on the similarity between the cell's \gD{}
-embedding neighbors and its neighbors prior to embedding. By identifying
-embeddings with low reliability scores as dubious and those with high
-scores as trustworthy, scDEED minimizes the number of questionable cell
-embeddings. This approach also provides clear guidance for optimizing
-the hyperparameters of an embedding method.
-
-To verify that the \gD{} cell embeddings became better aligned with the
-cell types after scDEED's optimization, the Human Peripheral Blood
-Mononuclear Cells (PBMC) dataset was used. It contained \(31,021\) cells
-with cell type labels, and the gene expression levels were in the unit
-of log-transformed UMI count per \(10,000\). They focused on three
-sequencing methods (inDrops, DropSeq, and SeqWell) and four common cell
-types Cytotoxic T cell, CD4+T cell, CD14+ Monocyte, and B cell.
+As we were writing this paper \citet{xia2023} appeared proposing a new
+method called scDEED helping to assess the validity of a \gD{}
+embedding. scDEED calculates a reliability score for each cell embedding
+based on the similarity between the cell's \gD{} embedding neighbors and
+its neighbors prior to embedding. A low reliability score suggests a
+dubious embedding. It can help in the deciding on optimal
+hyperparameters. Here we illustrate how our method compares with the
+results from scDEED
+
+Following the process in \citet{xia2023} again using the Human
+Peripheral Blood Mononuclear Cells (PBMC) data. Note that
+\citet{xia2023} uses a different subset from Section~\ref{sec-pbmc},
+which contains \(31,021\) cells including cell type labels, and the gene
+expression levels were in the unit of log-transformed UMI count per
+\(10,000\). They focused on three sequencing methods (inDrops, DropSeq,
+and SeqWell) and four common cell types Cytotoxic T cell, CD4+T cell,
+CD14+ Monocyte, and B cell.
 
@@
 
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-pbmc-mse-umap-1.pdf}
+\textbf{?@fig-model-pbmc-author-proj-scdeed} shows layout from the
+original published article along with that suggested by scDEED to be
+more accurate. The log MSE vs binwidth plot illustrates that our
+approach would suggest that scDEED is correct here, that layout b is
+more accurately reflecting the cluster structure in the data.
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-pbmc-mse-umap-1.pdf}
 
@@
 (\(a_1\)). Colour used for the lines and points in the left plot and in
-the scatterplots represents UMAP layout (a, b). Layout b is universally
-suboptimal. Layout b is the best choice.}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-model-pbmc-author-proj-scdeed-1.pdf}
-
-}
-
-\caption{\label{fig-model-pbmc-author-proj-scdeed}Model in \gD{}, on the
-layout a, and two views of the fit in projections from \(50\text{-}D\),
-for the PBMC data (\(n =  5858\) and \(p = 50\)). Layout shows close
-clusters. In \(50\text{-}D\), the data shows}
+the scatterplots represents UMAP layout (a, b). Of the two, layout b is
+optimal across all binwidths making it the best choice.}
 

### 2025-01-12 paper.tex

@@
 to illustrate the curious detail. Figure~\ref{fig-five-gau-projs} a2,
-b2, c2 show the fitted models in a projection of the 4D space. In this
-projection the difference between methods can be seen. These clusters
-are fully 4D in nature, so we would expect the model to be a
-\emph{crumpled 2D sheet} that stretches in all four dimensions. This is
-what is observed for t-SNE and UMAP. The curious detail is that the
-model for PaCMAP is closer to a \emph{pancake} in shape! What this means
-is that there has to be some ordering of points in the 2D PaCMAP layout
-that induces the flat model, likely that the points are organised by the
-global principal components. So the layout of the model in 4D doesn't
-reflect the dimension of each cluster. DHC: is there something in the
-documentation that might help to explain what has happened?
-
-This pattern can also be observed by examining the error from the model
-fit (Figure~\ref{fig-five-gau-projs} a3, b3, c3). With t-SNE and UMAP
-the error is relatively uniformly distributed, but with PaCMAP there is
-more error in the center of the 2D layout, reflecting that there are
-more points in the middle that are further from the fitted model.
+b2, c2 show the fitted models in a projection of the \(4\text{-}D\)
+space. In this projection the difference between methods can be seen.
+These clusters are fully \(4\text{-}D\) in nature, so we would expect
+the model to be a \emph{crumpled \gD{} sheet} that stretches in all four
+dimensions. This is what is observed for tSNE and UMAP. The curious
+detail is that the model for PaCMAP is closer to a \emph{pancake} in
+shape! What this means is that there has to be some ordering of points
+in the \gD{} PaCMAP layout that induces the flat model, likely that the
+points are organised by the global principal components. So the layout
+of the model in \(4\text{-}D\) doesn't reflect the dimension of each
+cluster. DHC: is there something in the documentation that might help to
+explain what has happened?
 

### 2025-01-12 paper.qmd
no rigid page breaks allowed
Added description to the effect of density section

--- 2025-01-12 paper.qmd
@@
 The default hyper-parameters are used.
+The \gD{} layouts are colored according to the high-dimensional model error.
+In b1, the large model error points are clustered in the top left corner, forming a separate cluster with high model errors.
 
@@
 @fig-one-dens_clust-error b1 and b2 show the fitted models in projections of the $7\text{-}D$ space.
-In these projection the difference between the distribution of high-dimensional model error can be seen.
-The uniform densed C-shaped cluster
+In these projections, the differences in the distribution of high-dimensional model errors can be observed.
+The uniform C-shaped cluster displays randomly distributed model errors, while the dense C-shaped cluster shows high model errors at the sparse end.
+In the dense C-shaped cluster, the high model errors occur because there are fewer data points positioned within the bins at the sparse end of the structure.
 
@@
 #| fig-pos: H
+#| out-height: 50%
 #| fig-cap: "The sparse end of the structure exhibits a high model error, while the dense end shows a low model error.
-The tSNE layouts of the C-shaped cluster, with uniform points (a1) and dense points (b1) are colored according to their high-dimensional error, along with their selected projections (a2 and b2), are displayed.
+The tSNE layouts of the C-shaped structure, with uniform points (a1) and dense points (b1) are colored according to their high-dimensional error, along with their selected projections (a2 and b2), are displayed.
 Darker colors represent a high error, while lighter colors indicate a low error.
-The uniformly distributed points demonstrate random model errors throughout the structure, whereas the dense structure shows large model error at the sparse end."
+The uniformly distributed points demonstrate random model errors throughout the structure, whereas the dense structure shows large model error at the sparse end.
+Videos of the langevitour animations are available at."
 

### 2025-01-12 paper.tex
no rigid page breaks allowed
Added description to the effect of density section

--- 2025-01-12 paper.tex
@@
 b1 shows \gD{} layouts for tSNE. The default hyper-parameters are used.
+The \gD{} layouts are colored according to the high-dimensional model
+error. In b1, the large model error points are clustered in the top left
+corner, forming a separate cluster with high model errors.
 
@@
 Figure~\ref{fig-one-dens_clust-error} b1 and b2 show the fitted models
-in projections of the \(7\text{-}D\) space. In these projection the
-difference between the distribution of high-dimensional model error can
-be seen. The uniform densed C-shaped cluster
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-one-dens_clust-error-1.pdf}
+in projections of the \(7\text{-}D\) space. In these projections, the
+differences in the distribution of high-dimensional model errors can be
+observed. The uniform C-shaped cluster displays randomly distributed
+model errors, while the dense C-shaped cluster shows high model errors
+at the sparse end. In the dense C-shaped cluster, the high model errors
+occur because there are fewer data points positioned within the bins at
+the sparse end of the structure.
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=0.5\textheight]{paper_files/figure-pdf/fig-one-dens_clust-error-1.pdf}
 
@@
 exhibits a high model error, while the dense end shows a low model
-error. The tSNE layouts of the C-shaped cluster, with uniform points
+error. The tSNE layouts of the C-shaped structure, with uniform points
 (a1) and dense points (b1) are colored according to their
@@
 demonstrate random model errors throughout the structure, whereas the
-dense structure shows large model error at the sparse end.}
+dense structure shows large model error at the sparse end. Videos of the
+langevitour animations are available at.}
 

### 2025-01-13 paper.qmd
Added descriptions for effect of density

--- 2025-01-12 paper.qmd
@@
 :::
-
-{{< pagebreak >}}

### 2025-01-13 paper.qmd
Added what is meant by single cell data
Added number of gene experessions for PBMC3k

@@
 
-XXX What is single cell data????
+<!-- XXX What is single cell data???? -->
+
+Single-cell data refers to the characteristics of individual cells within a population (@haque2017).
+These data also provides insights at the level of individual cells, enabling deeper understanding of gene expression.
 
@@
 The cluster results of Human Peripheral Blood Mononuclear Cells (PBMC3k) in @chen2023 are examined.
-There are $2622$ single cells, with XXX gene expressions.
+There are $2622$ single cells, with $1000$ gene expressions.
 Following their pre-processing, UMAP was performed using nine principal components.

### 2025-01-13 paper.qmd
Resize figures

@@
 ```{r}
-#| fig-cap: "Assessing which of the 6 NLDR layouts on the two C-shaped clusters data is the better representation using MSE for varying binwidth ($a_1$).
+#| fig-cap: "Assessing which of the six NLDR layouts on the two C-shaped clusters data is the better representation using MSE for varying binwidth ($a_1$).
 Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-e).
@@
 #| fig-pos: H
-#| out-width: 100%
+#| out-width: 60%
 
@@
 #| fig-height: 15
-#| out-width: 80%
+#| out-width: 60%
 
@@
 #| fig-width: 10
+#| out-width: 60%
 #| fig-cap: "Model in \\gD{}, on the layout a and three views of the fit in projections from $10\\text{-}D$, for the MNIST digit 1 data ($n =  7877$ and $p = 10$).

### 2025-01-13 paper.qmd
Added script to make the model with tSNE

@@
 
-<!--UMAP applied for Two curvy data-->
-```{r}
-#| label: umap-two-curvy
+<!--tSNE applied for Two curvy data-->
+```{r}
+#| label: tsne-two-curvy
 tsne_two_curvy <- read_rds(file = "data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_30.rds")
@@
 ## Compute hexbin parameters
-num_bins_x_two_curvy <- 23
+num_bins_x_two_curvy <- 27
 

### 2025-01-13 paper.qmd
Tried to add long edges within C-shaped cluster. Found: With UMAP, the model generate long edges within the clusters, while with tSNE between the clusters. The resaon that I think is that the characteric of tSNE to prserve global distance. To keep the clusters close. With UMAP, if the clusters are close the global structure already preserved. But to preserve the local distances, and didn't make the edges of C-shaped cluster too large, it make the long edges to keep the distances.

@@
   interior_annotation("c", sc_ltr_pos) +
+  theme(
+    legend.position = "none"
+  )
+
+two_curvy_proj_umap_first_model1_temp <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#000000") + #31a354
+  geom_point(
+    size = 0.5,
+    alpha = 0.2,
+    color = "#636363") +
+  geom_segment(
+    data=axes,
+    aes(x=x1, y=y1, xend=x2, yend=y2),
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
+    colour="grey50",
+    size = 4) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.1, 0.1)) + 
+  ylim(c(-0.1, 0.1)) +
+  interior_annotation("c2", sc_ltr_pos) +
   theme(
@@
 free(distance_points) /
-  wrap_plots(trimesh_two_curvy, trimesh_two_curvy_removed2, trimesh_two_curvy_removed1, two_curvy_proj_umap_model_delaunay, two_curvy_proj_umap_model_benchmark2, two_curvy_proj_umap_first_model1, ncol = 3) +
+  wrap_plots(trimesh_two_curvy, trimesh_two_curvy_removed2, trimesh_two_curvy_removed1, two_curvy_proj_umap_model_delaunay, two_curvy_proj_umap_model_benchmark2, two_curvy_proj_umap_first_model1_temp, ncol = 3) +
   plot_annotation(tag_levels = list(c("(a)", "(b)")))

### 2025-01-13 scripts/two_non_linear_diff_shaped_close_clusters/02_gen_true_model.R

@@ -96,7 +96,7 @@ curvilinear_grid1 <- generate_curvilinear_grid_2d(n_grid_x, n_grid_y) |>
   as_tibble()
 
 # Apply an offset to one of the clusters to create a distance between them
-offset <- c(1.5, 2.3, 1)  # Adjust these values to set the desired distance
+offset <- c(2.5, 2.5, 2.5)  # Adjust these values to set the desired distance
 curvilinear_grid1 <- sweep(curvilinear_grid1, 2, offset, "+")
 
 # Add small noise to the grid data to match dimensions with curvilinear data

### 2025-01-13 paper.qmd
Updated Chen's paper citation
Initial writeup for the application intro

@@
 ## Compute hexbin parameters
-num_bins_x_two_curvy <- 27
+num_bins_x_two_curvy <- 23
 
@@
 
-benchmark_two_curvy <- 2.5 * bin_width
+benchmark_two_curvy <- 2.8 * bin_width
 
@@
 
-XXX WHY THESE EXAMPLES. NEEDS AN OVERVIEW PARAGRAPH
+<!-- XXX WHY THESE EXAMPLES. NEEDS AN OVERVIEW PARAGRAPH -->
+
+To showcase the robustness of our method, this section presents two case studies.
+These applications were chosen to highlight various types of data where NLDR techniques play a critical role.
+The first case study explores single-cell data, a domain characterized by high-dimensional, sparse, and complex data structures.
+The second case focuses on the classic hand-written digits dataset, which provides an interpretable and visually intuitive benchmark for comparing NLDR techniques.
 

### 2025-01-13 paper.qmd
Added intro to application section

--- 2025-01-13 paper.qmd
@@
 
-XXX WHY THESE EXAMPLES. NEEDS AN OVERVIEW PARAGRAPH
+<!-- XXX WHY THESE EXAMPLES. NEEDS AN OVERVIEW PARAGRAPH -->
+
+To showcase the robustness of our method when applied to data with different characteristics, this section presents two case studies.
+The first application focuses on single-cell data, a domain where the high dimensionality, sparsity, and heterogeneity of gene expression measurements pose significant analytical challenges.
+Our approach facilitates the identification of the most accurate NLDR layout, enabling the discovery of cell populations with similar expression profiles.
+The second case study explores the MNIST hand-written digits dataset, which serves as a benchmark for evaluating the preservation of local structures, specifically within a selected digit.
 

### 2025-01-13 paper.qmd
Updated Figure 9

@@
 source("scripts/five_gau_clusters/5_gen_model_with_tsne.R")
+source("scripts/two_non_linear_diff_shaped_close_clusters/10_gen_rm_lg_plots.R")
 ```
@@
 
-free(distance_points) /
-  wrap_plots(trimesh_two_curvy, trimesh_two_curvy_removed2, trimesh_two_curvy_removed1, two_curvy_proj_umap_model_delaunay, two_curvy_proj_umap_model_benchmark2, two_curvy_proj_umap_first_model1_temp, ncol = 3) +
-  plot_annotation(tag_levels = list(c("(a)", "(b)")))
-
-# free(distance_points) + plot_annotation(tag_levels = 'A') +
-#  (trimesh_two_curvy2 + trimesh_two_curvy_removed2 + trimesh_two_curvy_removed1 + two_curvy_proj_umap_model_delaunay + two_curvy_proj_umap_model_benchmark2 + two_curvy_proj_umap_model) +
-#   plot_layout(widths = c(0.5, 1, 1, 1)) +
-#   plot_annotation(tag_levels = list(c("(a)", "(b)"))) 
+gen_tuning_lg_plots()
 ```

### 2025-01-13 paper.qmd
Updated Figure 9

@@
 #| fig-cap: "XXX Shouldn't the middle one have an edge connecting inside a C, so that it is clearly not as good as c?
-XXX Distribution of edge lengths and wireframes in \\gD{} and $7\\text{-}D$ with Delaunay triangulation, and two choices of long edge removal: (b) benchmark = 2.5$a_1$ and (c) benchmark = 2$a_1$, where $a_1$ = 0.09."
+XXX Distribution of edge lengths and wireframes in \\gD{} and $7\\text{-}D$ with Delaunay triangulation, and two choices of long edge removal: (b) benchmark = 6.5$a_1$ and (c) benchmark = 2$a_1$, where $a_1$ = 0.09."
 #| fig-height: 15

### 2025-01-13 paper.qmd
Added label for distance plot in Figure 9

@@
 #| label: fig-rm-lg
-#| fig-cap: "XXX Shouldn't the middle one have an edge connecting inside a C, so that it is clearly not as good as c?
-XXX Distribution of edge lengths and wireframes in \\gD{} and $7\\text{-}D$ with Delaunay triangulation, and two choices of long edge removal: (b) benchmark = 6.5$a_1$ and (c) benchmark = 2$a_1$, where $a_1$ = 0.09."
+#| fig-cap: "\\gD{} Vs $7\\text{-}D$ distance between the bin centroids and wireframes in \\gD{} and $7\\text{-}D$ with Delaunay triangulation, and two choices of long edge removal: (b) benchmark = 6.5$a_1$ and (c) benchmark = 2$a_1$, where $a_1$ = 0.03."
 #| fig-height: 15

### 2025-01-13 paper.qmd
Added Figure 9 title

@@
 #| label: fig-rm-lg
-#| fig-cap: "\\gD{} Vs $7\\text{-}D$ distance between the bin centroids and wireframes in \\gD{} and $7\\text{-}D$ with Delaunay triangulation, and two choices of long edge removal: (b) benchmark = 6.5$a_1$ and (c) benchmark = 2$a_1$, where $a_1$ = 0.03."
+#| fig-cap: "The choice of long edge removal significantly impacts the model's accuracy.
+\\gD{} Vs $7\\text{-}D$ distances between the bin centroids, where the model is fitted with UMAP for the non-linear cluster data.
+The two benchmarks for long edge removal: (b) benchmark = 6.5$a_1$ and (c) benchmark = 2$a_1$, where $a_1$ = 0.03 are made.
+The wireframes in \\gD{} and a projection from $7\\text{-}D$ are shown with Delaunay triangulation (a1, a2) and the two benchmark values (b1, b2, c1, c2).
+With the benchmark set at 6.5$a_1$, long edges exist within the C-shaped cluster, which distrupts the model.
+In contrast, using 2$a_1$ is a better option, as it effectively captures the non-linear shapes of the clusters.
+To generate an accurate model that captures the data structure, the choice of long edge removal does matter."
 #| fig-height: 15

### 2025-01-13 appendix.qmd
Added the notations
Added the equations with a flow
Updated the appendix
Moved scDEED to appendix

@@
   \newcommand\gD{$2\text{-}D$}
+  \newcommand{\To}{\textbf{to}~}
+  \usepackage{abstract}
+  \renewcommand{\abstractname}{}    % clear the title
+  \renewcommand{\absnamepos}{empty} % originally center
 ---
@@
 
+\newpage
+
 ## Area of a hexagon
@@
 ```{r}
-
-triangle_temp <- tibble(
-  x = c(0, 4, 0),
-  y = c(0, 0, 3)
-)
+# triangle_temp <- tibble(
+#   x = c(0, 4, 0),
+#   y = c(0, 0, 3)
+# )
 
@@
 ```{r}
+#| label: fig-tri-param
+#| fig-cap: "The components of the right triangle illustrating notation."
 #| out-height: 30%
+
 triangle_plot
@@
 
-By pythagorean theorem,
+By applying the Pythagorean theorem, we obtain,
 
 $$
-l^2 = (\frac{a_1}{2})^2 + (\frac{a_2 - l}{2})^2
+l^2 = \left(\frac{a_1}{2}\right)^2 + \left(\frac{a_2 - l}{2}\right)^2.
 $$
+Next, rearranging the terms, we get,
+
 $$
-l^2 - (\frac{a_2 - l}{2})^2 = (\frac{a_1}{2})^2
-$$
-$$
-[l - (\frac{a_2 - l}{2})][l + (\frac{a_2 - l}{2})] = (\frac{a_1}{2})^2
+l^2 - \left(\frac{a_2 - l}{2}\right)^2 = \left(\frac{a_1}{2}\right)^2,
 $$
@@
 $$
-3l^2 + 2la_2 - (a_1^2 + a_2^2) = 0
+\left[l - \left(\frac{a_2 - l}{2}\right)\right]\left[l + \left(\frac{a_2 - l}{2}\right)\right] = \left(\frac{a_1}{2}\right)^2,
 $$
@@
 $$
-l = \frac{-2a_2 \pm \sqrt{4a_2^2 - 24[-(a_1^2 + a_2^2)]}}{6}
+3l^2 + 2a_2l - (a_1^2 + a_2^2) = 0.
 $$
+
+Finally, by solving the quadratic equation, we compute,
+
 $$
-l = \frac{-a_2 \pm \sqrt{a_2^2 - 6[-(a_1^2 + a_2^2)]}}{3}
+l = \frac{-2a_2 \pm \sqrt{4a_2^2 - 24[-(a_1^2 + a_2^2)]}}{6},
 $$
+
+$$
+l = \frac{-a_2 \pm \sqrt{a_2^2 - 6[-(a_1^2 + a_2^2)]}}{3},
+$$
+
+where $l > 0$.
+
+#### Comparison with results of scDEED recommendations
+
+<!--
+XXX What this section needs:
+
+- connect this with previous section, not make it seem like a completely new. connections are: nldr used to show clustering, scDEED provides a different approach to decide which is best.
+- why these two? assume one is the published figure, and second is one recommended by scDEED, which our results confirm is better.
+- Refer to figure in text
+- Fig 16 not necessary
+
+CHANGES WERE MADE BASED ON THIS THINKING
+-->
+
+<!-- In the field of single-cell studies, clustering is a common analytical task used to identify groups of cells with similar expression profiles.
+Non-linear dimensional reduction (NLDR) methods are frequently employed to visualize these clusters and help validate the results.
+However, it is well known that the 2D embeddings produced by t-SNE and UMAP may not accurately reflect the similarities among cell clusters.
+ALREADY STATED IN PREVIOUS SECTION -->
+
+As we were writing this paper @xia2023 appeared proposing a new method called scDEED helping to assess the validity of a \gD{} embedding.
+scDEED calculates a reliability score for each cell embedding based on the similarity between the cell’s \gD{} embedding neighbors and its neighbors prior to embedding.
+A low reliability score suggests a dubious embedding.
+It can help in the deciding on optimal hyperparameters.
+Here we illustrate how our method compares with the results from scDEED
+
+Following the process in @xia2023 again using the Human Peripheral Blood Mononuclear Cells (PBMC) data.
+Note that @xia2023 uses a different subset from @sec-pbmc, which contains $31,021$ cells including cell type labels, and the gene expression levels were in the unit of log-transformed UMI count per $10,000$.
+They focused on three sequencing methods (inDrops, DropSeq, and SeqWell) and four common cell types Cytotoxic T cell, CD4+T cell, CD14+ Monocyte, and B cell.
+
+For illustration purposes, we only selected cells generated with inDrops ($n=5858$ cells) and UMAP cell embeddings.
+Also, @xia2023 used first $50$ principal components to generate the UMAP.
+The objective is to assess the optimized layout by scDEED, and if it does not accurately represent the three clusters with small separations of the PBMC dataset, then select a reasonable \gD{} layout.
+
+@fig-model-pbmc-author-proj-scdeed shows layout from the original published article along with that suggested by scDEED to be more accurate.
+The log MSE vs binwidth plot illustrates that our approach would suggest that scDEED is correct here, that layout b is more accurately reflecting the cluster structure in the data.
+
+```{r}
+result_umap <- read_rds(here::here('data/pbmc3k/pbmc_scdeed_umap_results.rds'))
+```
+
+```{r}
+#| eval: false
+result_umap |>
+  kableExtra::kable(format = "latex",
+                    booktabs = TRUE,
+                    longtable = TRUE,
+                    label = "scdeedumap") |>
+  kableExtra::kable_styling(latex_options = "scale_down")
+```
+
+```{r}
+#| out-width: 50%
+#| eval: false
+
+result_umap |>
+  filter(min.dist == 0.5) |>
+  ggplot(aes(
+           x = as.numeric(n_neighbors),
+           y = number_dubious_cells
+         )) +
+    geom_line() +
+    geom_point() +
+    labs(x = "n_neighbors",
+         y = "# of dubious cells") +
+    theme_minimal()
+```
+
+```{r}
+#| label: read-pbmc-umap-layouts
+# Read a variety of different NLDR representations of PBMC
+# and plot them on same aspect ratio
+umap_pbmc1 <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_30_min_dist_0.3.rds")
+umap_pbmc1 <- as_tibble(umap_pbmc1)
+names(umap_pbmc1) <- c("UMAP1", "UMAP2")
+
+nldr1c <- umap_pbmc1 |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2)) +
+  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
+  interior_annotation("a")
+
+umap_pbmc2 <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_80_min_dist_0.5.rds")
+umap_pbmc2 <- as_tibble(umap_pbmc2)
+names(umap_pbmc2) <- c("UMAP1", "UMAP2")
+
+nldr2c <- umap_pbmc2 |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2))+
+  geom_point(alpha=0.1, size=1, colour='#377eb8') +
+  interior_annotation("b")
+
+```
+
+```{r}
+#| label: combine-data-pbmc-umap
+
+error_pbmc_umap <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_30_min_dist_0.3.rds")
+error_pbmc_umap2 <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_80_min_dist_0.5.rds")
+
+error_pbmc <- bind_rows(error_pbmc_umap, 
+                        error_pbmc_umap2)
+
+error_pbmc <- error_pbmc |>
+  mutate(a1 = round(a1, 2)) |>
+  filter(bin1 >= 5) |>
+  group_by(method, a1) |>
+  filter(MSE == min(MSE)) |>
+  ungroup()
+```
+
+```{r}
+#| label: error-comp-pbmc-umap
+
+error_umap_plot_pbmc <- ggplot(error_pbmc, 
+                               aes(x = a1, 
+                                   y = MSE, 
+                                   colour = method)) + 
+  geom_point(size = 0.8) +
+  geom_line(linewidth = 0.3) + 
+  # geom_vline(xintercept = 15, linetype="solid", 
+  #            color = "black", linewidth=0.8, alpha = 0.5) +
+  scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 12, 15, 17, 19)]) +
+  scale_color_manual(values=c('#e41a1c','#377eb8')) +
+  scale_y_log10() +
+  ylab("log(MSE)") +
+  xlab(expression(paste("binwidth (", a[1], ")"))) +
+  theme_minimal() +
+  theme(panel.border = element_rect(fill = 'transparent'),
+        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+        axis.ticks.x = element_line(),
+        axis.ticks.y = element_line(),
+        legend.position = "none",
+        axis.text.x = element_text(size = 7),
+        axis.text.y = element_text(size = 7),
+        axis.title.x = element_text(size = 7),
+        axis.title.y = element_text(size = 7))
+
+```
+
+```{r}
+#| label: fig-pbmc-mse-umap
+#| fig-pos: H
+#| out-width: 100%
+#| fig-cap: "Assessing which of the 2 UMAP layouts with different hyperparameter settings (n_neighbors: $30$, min_dist: $0.3$ (red), n_neighbors: $80$, min_dist: $0.5$ (blue)) on the PBMC data is the better representation using MSE for varying binwidth ($a_1$).
+Colour used for the lines and points in the left plot and in the scatterplots represents UMAP layout (a, b).
+Of the two, layout b is optimal across all binwidths making it the best choice."
+
+free(error_umap_plot_pbmc) + wrap_plots(nldr1c, nldr2c, ncol = 2) +
+  plot_layout(ncol = 1, heights = c(1, 1))
+```
+
+```{r}
+#| label: published-pbmc-scdeed
+
+training_data_pbmc <- read_rds("data/pbmc3k/pbmc_pca_50_scdeed.rds")
+training_data_pbmc <- training_data_pbmc[, 1:50] |>
+  mutate(ID = 1:NROW(training_data_pbmc))
+
+## For umap
+umap_pbmc <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_80_min_dist_0.5.rds")
+umap_pbmc <- as_tibble(umap_pbmc)
+names(umap_pbmc) <- c("UMAP1", "UMAP2")
+umap_pbmc <- umap_pbmc |>
+  mutate(ID = 1:NROW(umap_pbmc))
+
+pbmc_scaled_obj_umap <- gen_scaled_data(
+  data = umap_pbmc)
+umap_pbmc_scaled <- pbmc_scaled_obj_umap$scaled_nldr
+
+lim1 <- pbmc_scaled_obj_umap$lim1
+lim2 <- pbmc_scaled_obj_umap$lim2
+r2_umap <- diff(lim2)/diff(lim1)
+
+```
+
+```{r}
+#| label: hexbin-pbmc-scdeed
+## Compute hexbin parameters
+num_bins_x_pbmc <- 12
+lim1 <- pbmc_scaled_obj_umap$lim1
+lim2 <- pbmc_scaled_obj_umap$lim2
+r2_pbmc <- diff(lim2)/diff(lim1)
+
+pbmc_model <- fit_highd_model(
+  training_data = training_data_pbmc,
+  emb_df = umap_pbmc_scaled,
+  bin1 = num_bins_x_pbmc,
+  r2 = r2_pbmc,
+  q = 0.1,
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = FALSE,
+  col_start_highd = "pc"
+)
+
+df_bin_centroids_pbmc <- pbmc_model$df_bin_centroids
+df_bin_pbmc <- pbmc_model$df_bin
+
+## Triangulate bin centroids
+tr1_object_pbmc <- tri_bin_centroids(
+  df_bin_centroids_pbmc, x = "c_x", y = "c_y")
+tr_from_to_df_pbmc <- gen_edges(
+  tri_object = tr1_object_pbmc)
+
+## Compute 2D distances
+distance_pbmc <- cal_2d_dist(
+  tr_coord_df = tr_from_to_df_pbmc,
+  start_x = "x_from",
+  start_y = "y_from",
+  end_x = "x_to",
+  end_y = "y_to",
+  select_vars = c("from", "to", "distance"))
+
+## To find the benchmark value
+benchmark_pbmc <- find_lg_benchmark(
+  distance_edges = distance_pbmc,
+  distance_col = "distance")
+
+tr_df <- distinct(tibble(
+  x = c(tr_from_to_df_pbmc[["x_from"]], tr_from_to_df_pbmc[["x_to"]]), 
+  y = c(tr_from_to_df_pbmc[["y_from"]], tr_from_to_df_pbmc[["y_to"]])))
+
+distance_df_small_edges_pbmc <- distance_pbmc |>
+  filter(distance < benchmark_pbmc)
+
+tr_from_to_df_pbmc <- inner_join(
+  tr_from_to_df_pbmc, distance_df_small_edges_pbmc, 
+  by = c("from", "to"))
+
+trimesh_removed_pbmc <- ggplot() + 
+  geom_segment(data = tr_from_to_df_pbmc, 
+               aes(
+                 x = x_from, 
+                 y = y_from, 
+                 xend = x_to, 
+                 yend = y_to),
+               colour = "#000000",
+               linewidth = 1) +
+  geom_point(
+    data = umap_pbmc_scaled,
+    aes(
+      x = UMAP1,
+      y = UMAP2
+    ),
+    alpha = 0.2,
+    color = '#a65628'
+  )  +
+  interior_annotation("a1", cex = 2) +
+  theme(
+    aspect.ratio = 1
+  )
+```
+
+```{r}
+#| label: pbmc-umap-model-proj-scdeed
+
+data_pbmc <- training_data_pbmc |> 
+  select(-ID) |>
+  mutate(type = "data")
+
+df_b_pbmc <- df_bin_pbmc |>
+  dplyr::filter(hb_id %in% df_bin_centroids_pbmc$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b_pbmc <- df_b_pbmc[match(df_bin_centroids_pbmc$hexID, df_b_pbmc$hb_id),] |>
+  dplyr::select(-hb_id)
+
+# Apply the scaling
+df_model_data_pbmc <- bind_rows(data_pbmc, df_b_pbmc)
+scaled_pbmc <- scale_data_manual(df_model_data_pbmc, "type") |>
+  as_tibble()
+
+scaled_pbmc_data <- scaled_pbmc |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_pbmc_data_model <- scaled_pbmc |>
+  filter(type == "model") |>
+  select(-type)
+
+# Combine with the true model for visualization
+df <- dplyr::bind_rows(scaled_pbmc_data_model |> mutate(type = "model"),
+                       scaled_pbmc_data |> mutate(type = "data"))
+
+## First projection
+projection <- cbind(
+  c(0.040856,0.000723,0.006000,0.007623,0.001986,-0.006599,0.005311,-0.000222,0.001648,0.008067,-0.001587,0.004423,0.004851,0.000638,0.002606,0.006616,0.006188,-0.000954,-0.004603,0.004998,-0.012220,0.008401,-0.007788,0.008972,-0.003903,0.003348,-0.001817,-0.000385,-0.002437,-0.004097,-0.001357,0.013660,0.001770,0.006982,0.002828,-0.009169,-0.003541,0.008582,-0.000032,-0.002375,0.005854,-0.005466,-0.000824,-0.001435,0.004148,0.007141,0.000883,-0.004159,0.005401,0.007882),
+  c(-0.000590,0.039365,-0.004481,-0.009041,-0.003805,0.003278,-0.009601,-0.000650,-0.002333,-0.005697,-0.001488,-0.009808,0.002557,0.004113,0.005705,-0.004665,-0.001170,-0.004727,-0.000845,0.005084,-0.018054,0.002538,0.006606,-0.007184,0.000238,0.001877,-0.002124,-0.005972,-0.009729,0.004958,-0.003181,-0.005185,-0.009598,-0.004243,0.007716,0.000396,-0.004258,0.005932,-0.003242,-0.004184,0.003493,-0.008405,0.000860,0.008304,0.004569,0.000798,0.004156,-0.005777,0.005278,-0.002531))
+
+projection_scaled <- projection * 3
+
+# projection_filtered <- projection
+# projection_filtered[apply(projection_filtered, 1, function(row) all(row < 0)), ] <- 0.75
+
+#rows_to_keep <- c(1, 2, 4, 5, 6)
+
+#projection_filtered <- projection[rows_to_keep, ]
+
+projected <- as.matrix(scaled_pbmc_data) %*% projection_scaled
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  #dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection_scaled
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges_pbmc |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+axes_obj <- gen_axes(
+  proj = projection,
+  limits = 0.04,
+  axis_pos_x = -0.025,
+  axis_pos_y = -0.025,
+  axis_labels = names(scaled_pbmc_data),
+  threshold = 0.0001)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
+
+pbmc_proj_umap_model1 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#000000") +
+  geom_point(
+    #size = 0.5,
+    alpha = 0.2,
+    color = '#a65628') +
+  geom_segment(
+    data=axes, 
+    aes(x=x1, y=y1, xend=x2, yend=y2), 
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
+    colour="grey50",
+    size = 3) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.03, 0.06)) +
+  ylim(c(-0.03, 0.06)) +
+  interior_annotation("a2", cex = 2)
+
+## Second projection
+projection <- cbind(
+  c(-0.019825,-0.011101,0.011543,-0.006943,-0.004485,0.000062,0.002023,-0.005178,0.000800,-0.002903,-0.014182,0.008756,0.009595,0.007032,-0.007820,0.006143,0.004865,0.003352,0.007554,0.002265,0.012035,0.009961,-0.001825,-0.004942,-0.002237,0.002308,0.009531,-0.001146,0.013401,-0.001022,0.008434,0.000477,0.004239,-0.006772,-0.004702,-0.011432,-0.000431,0.008146,-0.009299,-0.004782,0.003457,-0.008214,-0.009604,0.011329,0.003805,-0.003004,-0.000800,0.015288,-0.013157,0.006984),
+  c(-0.001387,0.001643,0.010470,0.002930,-0.002335,-0.010569,0.013839,0.002917,-0.006825,0.010191,-0.006367,0.002220,0.002506,0.005139,-0.007213,-0.006235,-0.007123,-0.000222,-0.002225,0.011592,-0.003942,0.002788,-0.000055,0.011316,-0.009753,0.020785,0.004615,0.005172,0.011182,-0.003125,-0.010860,0.007820,-0.017510,0.007010,0.007267,0.002694,-0.003949,-0.009188,-0.005163,0.001898,-0.004744,0.013757,0.010630,-0.007688,0.004598,0.004658,-0.011154,0.003584,-0.005226,0.006650))
+
+projection_scaled <- projection * 3
+
+projected <- as.matrix(scaled_pbmc_data) %*% projection_scaled
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  #dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection_scaled
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges_pbmc |> select(-distance), 
+  projected_model_df, 
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+axes_obj <- gen_axes(
+  proj = projection,
+  limits = 0.03,
+  axis_pos_x = -0.02,
+  axis_pos_y = -0.02,
+  axis_labels = names(scaled_pbmc_data),
+  threshold = 0)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
+
+pbmc_proj_umap_model2 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1,
+      y = proj2)) +
+  geom_segment(
+    data = model_df,
+    aes(
+      x = proj1_from,
+      y = proj2_from,
+      xend = proj1_to,
+      yend = proj2_to),
+    color = "#000000") +
+  geom_point(
+    #size = 0.5,
+    alpha = 0.2,
+    color = '#a65628') +
+  geom_segment(
+    data=axes,
+    aes(x=x1, y=y1, xend=x2, yend=y2),
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2),
+    label=rownames(axes),
+    colour="grey50",
+    size = 3) +
+  geom_path(
+    data=circle,
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.025, 0.02)) +
+  ylim(c(-0.025, 0.02)) +
+  interior_annotation("a3", cex = 2)
+```
+
+```{r}
+#| echo: false
+#| eval: false
+#| fig-cap: "Model in \\gD{}, on the layout a, and two views of the fit in projections from $50\\text{-}D$, for the PBMC data ($n =  5858$ and $p = 50$).
+Layout shows close clusters.
+In $50\\text{-}D$, the data shows "
+#| label: fig-model-pbmc-author-proj-scdeed
+#| fig-pos: H
+#| fig-width: 15
+#| fig-height: 5
+#| out-width: 80%
+
+free(trimesh_removed_pbmc) + 
+  pbmc_proj_umap_model1 + 
+  pbmc_proj_umap_model2 + 
+  plot_layout(ncol=3) &
+  theme(legend.position='none')
+```

### 2025-01-13 paper.qmd
Moved scDEED to appendix

@@
 
-#### Comparison with results of scDEED recommendations
-
-<!--
-XXX What this section needs:
-
-- connect this with previous section, not make it seem like a completely new. connections are: nldr used to show clustering, scDEED provides a different approach to decide which is best.
-- why these two? assume one is the published figure, and second is one recommended by scDEED, which our results confirm is better.
-- Refer to figure in text
-- Fig 16 not necessary
-
-CHANGES WERE MADE BASED ON THIS THINKING
--->
-
-<!-- In the field of single-cell studies, clustering is a common analytical task used to identify groups of cells with similar expression profiles.
-Non-linear dimensional reduction (NLDR) methods are frequently employed to visualize these clusters and help validate the results.
-However, it is well known that the 2D embeddings produced by t-SNE and UMAP may not accurately reflect the similarities among cell clusters.
-ALREADY STATED IN PREVIOUS SECTION -->
-
-As we were writing this paper @xia2023 appeared proposing a new method called scDEED helping to assess the validity of a \gD{} embedding.
-scDEED calculates a reliability score for each cell embedding based on the similarity between the cell’s \gD{} embedding neighbors and its neighbors prior to embedding.
-A low reliability score suggests a dubious embedding.
-It can help in the deciding on optimal hyperparameters.
-Here we illustrate how our method compares with the results from scDEED
-
-Following the process in @xia2023 again using the Human Peripheral Blood Mononuclear Cells (PBMC) data.
-Note that @xia2023 uses a different subset from @sec-pbmc, which contains $31,021$ cells including cell type labels, and the gene expression levels were in the unit of log-transformed UMI count per $10,000$.
-They focused on three sequencing methods (inDrops, DropSeq, and SeqWell) and four common cell types Cytotoxic T cell, CD4+T cell, CD14+ Monocyte, and B cell.
-
-For illustration purposes, we only selected cells generated with inDrops ($n=5858$ cells) and UMAP cell embeddings.
-Also, @xia2023 used first $50$ principal components to generate the UMAP.
-The objective is to assess the optimized layout by scDEED, and if it does not accurately represent the three clusters with small separations of the PBMC dataset, then select a reasonable \gD{} layout.
-
-@fig-model-pbmc-author-proj-scdeed shows layout from the original published article along with that suggested by scDEED to be more accurate.
-The log MSE vs binwidth plot illustrates that our approach would suggest that scDEED is correct here, that layout b is more accurately reflecting the cluster structure in the data.
-
-```{r}
-result_umap <- read_rds(here::here('data/pbmc3k/pbmc_scdeed_umap_results.rds'))
-```
-
-```{r}
-#| eval: false
-result_umap |>
-  kableExtra::kable(format = "latex",
-                    booktabs = TRUE,
-                    longtable = TRUE,
-                    label = "scdeedumap") |>
-  kableExtra::kable_styling(latex_options = "scale_down")
-```
-
-```{r}
-#| out-width: 50%
-#| eval: false
-
-result_umap |>
-  filter(min.dist == 0.5) |>
-  ggplot(aes(
-           x = as.numeric(n_neighbors),
-           y = number_dubious_cells
-         )) +
-    geom_line() +
-    geom_point() +
-    labs(x = "n_neighbors",
-         y = "# of dubious cells") +
-    theme_minimal()
-```
-
-```{r}
-#| label: read-pbmc-umap-layouts
-# Read a variety of different NLDR representations of PBMC
-# and plot them on same aspect ratio
-umap_pbmc1 <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_30_min_dist_0.3.rds")
-umap_pbmc1 <- as_tibble(umap_pbmc1)
-names(umap_pbmc1) <- c("UMAP1", "UMAP2")
-
-nldr1c <- umap_pbmc1 |>
-  ggplot(aes(x = UMAP1,
-             y = UMAP2)) +
-  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
-  interior_annotation("a")
-
-umap_pbmc2 <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_80_min_dist_0.5.rds")
-umap_pbmc2 <- as_tibble(umap_pbmc2)
-names(umap_pbmc2) <- c("UMAP1", "UMAP2")
-
-nldr2c <- umap_pbmc2 |>
-  ggplot(aes(x = UMAP1,
-             y = UMAP2))+
-  geom_point(alpha=0.1, size=1, colour='#377eb8') +
-  interior_annotation("b")
-
-```
-
-```{r}
-#| label: combine-data-pbmc-umap
-
-error_pbmc_umap <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_30_min_dist_0.3.rds")
-error_pbmc_umap2 <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_80_min_dist_0.5.rds")
-
-error_pbmc <- bind_rows(error_pbmc_umap, 
-                        error_pbmc_umap2)
-
-error_pbmc <- error_pbmc |>
-  mutate(a1 = round(a1, 2)) |>
-  filter(bin1 >= 5) |>
-  group_by(method, a1) |>
-  filter(MSE == min(MSE)) |>
-  ungroup()
-```
-
-```{r}
-#| label: error-comp-pbmc-umap
-
-error_umap_plot_pbmc <- ggplot(error_pbmc, 
-                               aes(x = a1, 
-                                   y = MSE, 
-                                   colour = method)) + 
-  geom_point(size = 0.8) +
-  geom_line(linewidth = 0.3) + 
-  # geom_vline(xintercept = 15, linetype="solid", 
-  #            color = "black", linewidth=0.8, alpha = 0.5) +
-  scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 12, 15, 17, 19)]) +
-  scale_color_manual(values=c('#e41a1c','#377eb8')) +
-  scale_y_log10() +
-  ylab("log(MSE)") +
-  xlab(expression(paste("binwidth (", a[1], ")"))) +
-  theme_minimal() +
-  theme(panel.border = element_rect(fill = 'transparent'),
-        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
-        axis.ticks.x = element_line(),
-        axis.ticks.y = element_line(),
-        legend.position = "none",
-        axis.text.x = element_text(size = 7),
-        axis.text.y = element_text(size = 7),
-        axis.title.x = element_text(size = 7),
-        axis.title.y = element_text(size = 7))
-
-```
-
-```{r}
-#| label: fig-pbmc-mse-umap
-#| fig-pos: H
-#| out-width: 100%
-#| fig-cap: "Assessing which of the 2 UMAP layouts with different hyperparameter settings (n_neighbors: $30$, min_dist: $0.3$ (red), n_neighbors: $80$, min_dist: $0.5$ (blue)) on the PBMC data is the better representation using MSE for varying binwidth ($a_1$).
-Colour used for the lines and points in the left plot and in the scatterplots represents UMAP layout (a, b).
-Of the two, layout b is optimal across all binwidths making it the best choice."
-
-free(error_umap_plot_pbmc) + wrap_plots(nldr1c, nldr2c, ncol = 2) +
-  plot_layout(ncol = 1, heights = c(1, 1))
-```
-
-```{r}
-#| label: published-pbmc-scdeed
-
-training_data_pbmc <- read_rds("data/pbmc3k/pbmc_pca_50_scdeed.rds")
-training_data_pbmc <- training_data_pbmc[, 1:50] |>
-  mutate(ID = 1:NROW(training_data_pbmc))
-
-## For umap
-umap_pbmc <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_80_min_dist_0.5.rds")
-umap_pbmc <- as_tibble(umap_pbmc)
-names(umap_pbmc) <- c("UMAP1", "UMAP2")
-umap_pbmc <- umap_pbmc |>
-  mutate(ID = 1:NROW(umap_pbmc))
-
-pbmc_scaled_obj_umap <- gen_scaled_data(
-  data = umap_pbmc)
-umap_pbmc_scaled <- pbmc_scaled_obj_umap$scaled_nldr
-
-lim1 <- pbmc_scaled_obj_umap$lim1
-lim2 <- pbmc_scaled_obj_umap$lim2
-r2_umap <- diff(lim2)/diff(lim1)
-
-```
-
-```{r}
-#| label: hexbin-pbmc-scdeed
-## Compute hexbin parameters
-num_bins_x_pbmc <- 12
-lim1 <- pbmc_scaled_obj_umap$lim1
-lim2 <- pbmc_scaled_obj_umap$lim2
-r2_pbmc <- diff(lim2)/diff(lim1)
-
-pbmc_model <- fit_highd_model(
-  training_data = training_data_pbmc,
-  emb_df = umap_pbmc_scaled,
-  bin1 = num_bins_x_pbmc,
-  r2 = r2_pbmc,
-  q = 0.1,
-  is_bin_centroid = TRUE,
-  is_rm_lwd_hex = FALSE,
-  col_start_highd = "pc"
-)
-
-df_bin_centroids_pbmc <- pbmc_model$df_bin_centroids
-df_bin_pbmc <- pbmc_model$df_bin
-
-## Triangulate bin centroids
-tr1_object_pbmc <- tri_bin_centroids(
-  df_bin_centroids_pbmc, x = "c_x", y = "c_y")
-tr_from_to_df_pbmc <- gen_edges(
-  tri_object = tr1_object_pbmc)
-
-## Compute 2D distances
-distance_pbmc <- cal_2d_dist(
-  tr_coord_df = tr_from_to_df_pbmc,
-  start_x = "x_from",
-  start_y = "y_from",
-  end_x = "x_to",
-  end_y = "y_to",
-  select_vars = c("from", "to", "distance"))
-
-## To find the benchmark value
-benchmark_pbmc <- find_lg_benchmark(
-  distance_edges = distance_pbmc,
-  distance_col = "distance")
-
-tr_df <- distinct(tibble(
-  x = c(tr_from_to_df_pbmc[["x_from"]], tr_from_to_df_pbmc[["x_to"]]), 
-  y = c(tr_from_to_df_pbmc[["y_from"]], tr_from_to_df_pbmc[["y_to"]])))
-
-distance_df_small_edges_pbmc <- distance_pbmc |>
-  filter(distance < benchmark_pbmc)
-
-tr_from_to_df_pbmc <- inner_join(
-  tr_from_to_df_pbmc, distance_df_small_edges_pbmc, 
-  by = c("from", "to"))
-
-trimesh_removed_pbmc <- ggplot() + 
-  geom_segment(data = tr_from_to_df_pbmc, 
-               aes(
-                 x = x_from, 
-                 y = y_from, 
-                 xend = x_to, 
-                 yend = y_to),
-               colour = "#000000",
-               linewidth = 1) +
-  geom_point(
-    data = umap_pbmc_scaled,
-    aes(
-      x = UMAP1,
-      y = UMAP2
-    ),
-    alpha = 0.2,
-    color = '#a65628'
-  )  +
-  interior_annotation("a1", cex = 2) +
-  theme(
-    aspect.ratio = 1
-  )
-```
-
-```{r}
-#| label: pbmc-umap-model-proj-scdeed
-
-data_pbmc <- training_data_pbmc |> 
-  select(-ID) |>
-  mutate(type = "data")
-
-df_b_pbmc <- df_bin_pbmc |>
-  dplyr::filter(hb_id %in% df_bin_centroids_pbmc$hexID) |>
-  dplyr::mutate(type = "model") ## Data with summarized mean
-
-## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
-df_b_pbmc <- df_b_pbmc[match(df_bin_centroids_pbmc$hexID, df_b_pbmc$hb_id),] |>
-  dplyr::select(-hb_id)
-
-# Apply the scaling
-df_model_data_pbmc <- bind_rows(data_pbmc, df_b_pbmc)
-scaled_pbmc <- scale_data_manual(df_model_data_pbmc, "type") |>
-  as_tibble()
-
-scaled_pbmc_data <- scaled_pbmc |>
-  filter(type == "data") |>
-  select(-type)
-
-scaled_pbmc_data_model <- scaled_pbmc |>
-  filter(type == "model") |>
-  select(-type)
-
-# Combine with the true model for visualization
-df <- dplyr::bind_rows(scaled_pbmc_data_model |> mutate(type = "model"),
-                       scaled_pbmc_data |> mutate(type = "data"))
-
-## First projection
-projection <- cbind(
-  c(0.040856,0.000723,0.006000,0.007623,0.001986,-0.006599,0.005311,-0.000222,0.001648,0.008067,-0.001587,0.004423,0.004851,0.000638,0.002606,0.006616,0.006188,-0.000954,-0.004603,0.004998,-0.012220,0.008401,-0.007788,0.008972,-0.003903,0.003348,-0.001817,-0.000385,-0.002437,-0.004097,-0.001357,0.013660,0.001770,0.006982,0.002828,-0.009169,-0.003541,0.008582,-0.000032,-0.002375,0.005854,-0.005466,-0.000824,-0.001435,0.004148,0.007141,0.000883,-0.004159,0.005401,0.007882),
-  c(-0.000590,0.039365,-0.004481,-0.009041,-0.003805,0.003278,-0.009601,-0.000650,-0.002333,-0.005697,-0.001488,-0.009808,0.002557,0.004113,0.005705,-0.004665,-0.001170,-0.004727,-0.000845,0.005084,-0.018054,0.002538,0.006606,-0.007184,0.000238,0.001877,-0.002124,-0.005972,-0.009729,0.004958,-0.003181,-0.005185,-0.009598,-0.004243,0.007716,0.000396,-0.004258,0.005932,-0.003242,-0.004184,0.003493,-0.008405,0.000860,0.008304,0.004569,0.000798,0.004156,-0.005777,0.005278,-0.002531))
-
-projection_scaled <- projection * 3
-
-# projection_filtered <- projection
-# projection_filtered[apply(projection_filtered, 1, function(row) all(row < 0)), ] <- 0.75
-
-#rows_to_keep <- c(1, 2, 4, 5, 6)
-
-#projection_filtered <- projection[rows_to_keep, ]
-
-projected <- as.matrix(scaled_pbmc_data) %*% projection_scaled
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  #dplyr::mutate(type = df_exe$type) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection_scaled
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  distance_df_small_edges_pbmc |> select(-distance), 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-axes_obj <- gen_axes(
-  proj = projection,
-  limits = 0.04,
-  axis_pos_x = -0.025,
-  axis_pos_y = -0.025,
-  axis_labels = names(scaled_pbmc_data),
-  threshold = 0.0001)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
-
-pbmc_proj_umap_model1 <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#000000") +
-  geom_point(
-    #size = 0.5,
-    alpha = 0.2,
-    color = '#a65628') +
-  geom_segment(
-    data=axes, 
-    aes(x=x1, y=y1, xend=x2, yend=y2), 
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 3) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.03, 0.06)) +
-  ylim(c(-0.03, 0.06)) +
-  interior_annotation("a2", cex = 2)
-
-## Second projection
-projection <- cbind(
-  c(-0.019825,-0.011101,0.011543,-0.006943,-0.004485,0.000062,0.002023,-0.005178,0.000800,-0.002903,-0.014182,0.008756,0.009595,0.007032,-0.007820,0.006143,0.004865,0.003352,0.007554,0.002265,0.012035,0.009961,-0.001825,-0.004942,-0.002237,0.002308,0.009531,-0.001146,0.013401,-0.001022,0.008434,0.000477,0.004239,-0.006772,-0.004702,-0.011432,-0.000431,0.008146,-0.009299,-0.004782,0.003457,-0.008214,-0.009604,0.011329,0.003805,-0.003004,-0.000800,0.015288,-0.013157,0.006984),
-  c(-0.001387,0.001643,0.010470,0.002930,-0.002335,-0.010569,0.013839,0.002917,-0.006825,0.010191,-0.006367,0.002220,0.002506,0.005139,-0.007213,-0.006235,-0.007123,-0.000222,-0.002225,0.011592,-0.003942,0.002788,-0.000055,0.011316,-0.009753,0.020785,0.004615,0.005172,0.011182,-0.003125,-0.010860,0.007820,-0.017510,0.007010,0.007267,0.002694,-0.003949,-0.009188,-0.005163,0.001898,-0.004744,0.013757,0.010630,-0.007688,0.004598,0.004658,-0.011154,0.003584,-0.005226,0.006650))
-
-projection_scaled <- projection * 3
-
-projected <- as.matrix(scaled_pbmc_data) %*% projection_scaled
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  #dplyr::mutate(type = df_exe$type) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection_scaled
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  distance_df_small_edges_pbmc |> select(-distance), 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-axes_obj <- gen_axes(
-  proj = projection,
-  limits = 0.03,
-  axis_pos_x = -0.02,
-  axis_pos_y = -0.02,
-  axis_labels = names(scaled_pbmc_data),
-  threshold = 0)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
-
-pbmc_proj_umap_model2 <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1,
-      y = proj2)) +
-  geom_segment(
-    data = model_df,
-    aes(
-      x = proj1_from,
-      y = proj2_from,
-      xend = proj1_to,
-      yend = proj2_to),
-    color = "#000000") +
-  geom_point(
-    #size = 0.5,
-    alpha = 0.2,
-    color = '#a65628') +
-  geom_segment(
-    data=axes,
-    aes(x=x1, y=y1, xend=x2, yend=y2),
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2),
-    label=rownames(axes),
-    colour="grey50",
-    size = 3) +
-  geom_path(
-    data=circle,
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.025, 0.02)) +
-  ylim(c(-0.025, 0.02)) +
-  interior_annotation("a3", cex = 2)
-```
-
-```{r}
-#| echo: false
-#| eval: false
-#| fig-cap: "Model in \\gD{}, on the layout a, and two views of the fit in projections from $50\\text{-}D$, for the PBMC data ($n =  5858$ and $p = 50$).
-Layout shows close clusters.
-In $50\\text{-}D$, the data shows "
-#| label: fig-model-pbmc-author-proj-scdeed
-#| fig-pos: H
-#| fig-width: 15
-#| fig-height: 5
-#| out-width: 80%
-
-free(trimesh_removed_pbmc) + 
-  pbmc_proj_umap_model1 + 
-  pbmc_proj_umap_model2 + 
-  plot_layout(ncol=3) &
-  theme(legend.position='none')
-```
-
 ### Hand-written digits

### 2025-01-13 paper.qmd
Updated Figure 13 size

@@
 #| fig-pos: H
-#| out-width: 60%
+#| out-width: 80%
 

### 2025-01-13 paper.qmd
Added another layout for Figure 15..to have nice layout

@@
 
+tsne_mnist2 <- read_rds("data/mnist/mnist_tsne89.rds")
+
+nldr_mnist6 <- tsne_mnist2 |>
+  ggplot(aes(x = tSNE1,
+             y = tSNE2))+
+  geom_point(alpha=0.1, size=1, colour='#984ea3') +
+  interior_annotation("f", c(0.08, 0.93)) +
+  theme(aspect.ratio = 1)
+
 umap_mnist <- read_rds("data/mnist/mnist_umap.rds")
@@
 error_mnist_tsne <- read_rds("data/mnist/error_mnist_tsne.rds")
+error_mnist_tsne2 <- read_rds("data/mnist/error_mnist_tsne2.rds")
 error_mnist_phate <- read_rds("data/mnist/error_mnist_phate.rds")
@@
                         error_mnist_tsne,
+                        error_mnist_tsne2,
                         error_mnist_phate,
@@
   scale_x_continuous(breaks = sort(unique(error_mnist$a1))[append(seq(1, length(unique(error_mnist$a1)), by = 5), 24)]) +
-  scale_color_manual(values=c('#e41a1c','#377eb8',"#ff7f00", '#999999','#a65628')) +
+  scale_color_manual(values=c('#e41a1c','#377eb8',"#ff7f00", '#999999', '#984ea3', '#a65628')) +
   scale_y_log10() +
@@
 free(error_plot_mnist) + wrap_plots(nldr_mnist1, nldr_mnist2, nldr_mnist3, 
-                                    nldr_mnist4, nldr_mnist5, ncol = 2)
+                                    nldr_mnist4, nldr_mnist5, nldr_mnist6, ncol = 2)
 ```

### 2025-01-13 paper.qmd
Try to find a better model for Hand-written digit with different number of bins and new choice  with tSNE perplexity 89

@@
              y = tSNE2))+
-  geom_point(alpha=0.1, size=1, colour='#999999') +
+  geom_point(alpha=0.1, size=1, colour='#984ea3') +
   interior_annotation("a", c(0.08, 0.93)) +
@@
              y = tSNE2))+
-  geom_point(alpha=0.1, size=1, colour='#984ea3') +
+  geom_point(alpha=0.1, size=1, colour='#999999') +
   interior_annotation("f", c(0.08, 0.93)) +
@@
   scale_x_continuous(breaks = sort(unique(error_mnist$a1))[append(seq(1, length(unique(error_mnist$a1)), by = 5), 24)]) +
-  scale_color_manual(values=c('#e41a1c','#377eb8',"#ff7f00", '#999999', '#984ea3', '#a65628')) +
+  scale_color_manual(values=c('#e41a1c','#377eb8',"#ff7f00", '#984ea3', '#999999', '#a65628')) +
   scale_y_log10() +
@@
 Layout c is universally poor.
-Layouts a that show two close clusters are universally suboptimal.
-Layout a is the best choice."
+Layouts f that show two close clusters are universally suboptimal.
+Layout f is the best choice."
 #| label: fig-mnist-mse
@@
 mnist_scaled_obj <- gen_scaled_data(
-  data = tsne_mnist)
+  data = tsne_mnist2)
 tsne_minst_scaled <- mnist_scaled_obj$scaled_nldr
@@
 ## Compute hexbin parameters
-num_bins_x_mnist <- 19
+num_bins_x_mnist <- 30
 lim1 <- mnist_scaled_obj$lim1
@@
   is_bin_centroid = TRUE,
-  is_rm_lwd_hex = FALSE,
+  is_rm_lwd_hex = TRUE,
   col_start_highd = "PC"
@@
 
-benchmark_mnist <- 0.1
+#benchmark_mnist <- 0.1
 
@@
 error_df <- error_df |>
-  mutate(sqrt_row_wise_abs_error = sqrt(row_wise_abs_error))
+  mutate(sqrt_row_wise_tot_error = sqrt(row_wise_total_error))
 
@@
              y = tSNE2,
-             colour = sqrt_row_wise_abs_error)) +
+             colour = sqrt_row_wise_tot_error)) +
   geom_point(alpha=0.2) +
@@
     #size = 0.2,
-    alpha = 0.3) +
+    alpha = 0.1) +
   scale_color_manual(values = c("#999999",'#ff7f00')) +

### 2025-01-14 paper.qmd
Start working on Figure 14

@@
 
+<!-- XXX Number of bins for b is inadequate. Caption needs to start with what this figure is for XXX -->
+
 ```{r}
 #| echo: false
-#| fig-cap: "XXX Number of bins for b is inadequate.
-Caption needs to start with what this figure is for XXX Model in \\gD{}, on the layout a and b, and two views of the fit in projections from $9\\text{-}D$, for the PBMC3k data ($n =  2622$ and $p = 9$).
+#| fig-cap: "Comapre the author suggested layout and the layout selected by our approach.
+Model in \\gD{}, on the layout a and b, and two views of the fit in projections from $9\\text{-}D$, for the PBMC3k data ($n =  2622$ and $p = 9$).
 Layout a shows three-well separated clusters with big separation, while layout b shows close clusters.

### 2025-01-14 paper.qmd
Updated Figure 7 title

@@
 #| fig-pos: H
-#| fig-cap: "Hexbin density plots of tSNE layout of the two C-shaped clusters data, using three different bin inputs: (a) $b = 255 \\text{ } (15, \\text{ }17)$, (b) $b = 598 \\text{ } (23, \\text{ }26)$, and (c) $b = 2544 \\text{ } (48, \\text{ }53)$.
+#| fig-cap: "Hexbin density plots of tSNE layout of the non-linear cluster data, using three different bin inputs: (a) $b = 255/98 \\text{ } (15, \\text{ }17)$, (b) $b = 598/188 \\text{ } (23, \\text{ }26)$, and (c) $b = 2544/549 \\text{ } (48, \\text{ }53)$.
 Color indicates standardized counts, dark indicating high count and light indicates low count.

### 2025-01-15 paper.qmd
Updated Figure 2

--- 2025-01-14 paper.qmd
@@
 ## Compute hexbin parameters
-num_bins_x_two_curvy <- 23
+num_bins_x_two_curvy <- 22
 
@@
   limits = 0.15,
-  axis_pos_x = -0.067,
-  axis_pos_y = -0.067,
+  axis_pos_x = -0.09,
+  axis_pos_y = -0.09,
   axis_labels = names(scaled_two_curvy_data),
@@
   coord_fixed() +
-  xlim(c(-0.1, 0.1)) + 
-  ylim(c(-0.1, 0.1)) +
+  xlim(c(-0.12, 0.12)) + 
+  ylim(c(-0.12, 0.12)) +
   interior_annotation("a", c(0.08, 0.93), cex = 2) +
@@
   limits = 0.1,
-  axis_pos_x = -0.05,
-  axis_pos_y = -0.05,
+  axis_pos_x = -0.07,
+  axis_pos_y = -0.07,
   axis_labels = names(scaled_two_curvy_data),
@@
   coord_fixed() +
-  xlim(c(-0.07, 0.07)) + 
-  ylim(c(-0.07, 0.07)) +
+  xlim(c(-0.09, 0.09)) + 
+  ylim(c(-0.09, 0.09)) +
   interior_annotation("b", c(0.08, 0.93), cex = 2) +
@@
 #| fig-pos: H
-#| fig-cap: "Hexbin density plots of tSNE layout of the two C-shaped clusters data, using three different bin inputs: (a) $b = 255 \\text{ } (15, \\text{ }17)$, (b) $b = 598 \\text{ } (23, \\text{ }26)$, and (c) $b = 2544 \\text{ } (48, \\text{ }53)$.
+#| fig-cap: "Hexbin density plots of tSNE layout of the non-linear cluster data, using three different bin inputs: (a) $b = 240/98 \\text{ } (15, \\text{ }16)$, (b) $b = 528/173 \\text{ } (22, \\text{ }24)$, and (c) $b = 2496/549 \\text{ } (48, \\text{ }52)$.
 Color indicates standardized counts, dark indicating high count and light indicates low count.

### 2025-01-15 paper.qmd
Updated Figure 2 title

@@
 #| label: fig-two-curvy-true-proj
-#| fig-cap: "Two views of the true model (blue lines) in $2\\text{-}D$ projections from $7\\text{-}D$, for the two C-shaped clusters data (grey points).
-In one C-shaped cluster, the data is spread throughout the cluster, while in the other C-shaped cluster, the data is concentrated more in one corner of the cluster.
-The **langevitour** software is used to view the data with a tour, and the full video is available at <>."
+#| fig-cap: "Two projections from $7\\text{-}D$, for the two non-linear cluster data are shown.
+The sample obtained from the true model is shown, with the sample shown in grey and the true model in blue lines.
+In the C-shaped cluster, the data points are distributed randomly along the cluster, while the data is more concentrated more in one corner in the other non-linear cluster.
+Video of the langevitour animations are available at <>."
 #| fig-pos: H

### 2025-01-15 paper.qmd
Updated Figure 3 title

@@
 #| label: fig-two-curvy-true-proj
-#| fig-cap: "Two projections from $7\\text{-}D$, for the two non-linear cluster data are shown.
+#| fig-cap: "Two projections from $7\\text{-}D$, for the two non-linear clusters data are shown.
 The sample obtained from the true model is shown, with the sample shown in grey and the true model in blue lines.
@@
 #| echo: false
-#| fig-cap: "Key steps for constructing the model on the UMAP layout ($k=2$): (a) data, (b) hexagon bins, (c) bin centroids, and (d) triangulated centroids. The two C-shaped clusters data is shown."
+#| fig-cap: "Key steps for constructing the model on the tSNE layout ($k=2$): (a) data, (b) hexagon bins, (c) bin centroids, and (d) triangulated centroids. The two non-linear clusters data is shown."
 #| fig-width: 12

### 2025-01-15 paper.qmd
Updated Figure 5

@@
   coord_fixed() +
-  xlim(c(-0.1, 0.1)) + 
-  ylim(c(-0.1, 0.1)) +
+  xlim(c(-0.12, 0.12)) + 
+  ylim(c(-0.12, 0.12)) +
   interior_annotation("c", sc_ltr_pos) +
@@
   coord_fixed() +
-  xlim(c(-0.1, 0.1)) + 
-  ylim(c(-0.1, 0.1)) +
+  xlim(c(-0.12, 0.12)) + 
+  ylim(c(-0.12, 0.12)) +
   interior_annotation("d", sc_ltr_pos) +

### 2025-01-15 paper.qmd
Updated Figure 6 title
Updated Figure title 15

@@
 #| fig-cap: "The $7\\text{-}D$ model error in \\gD{} layout.
-Color indicates square root of total error, dark blue indicating high error and light indicates low error.
+Color indicates square root of total error, dark colour indicating high error and light indicates low error.
 Most large errors are distributed near the edges of the clusters."
@@
 ```{r}
-#| fig-cap: "Assessing which of the 5 NLDR layouts on the MNIST digit 1 data is the better representation using MSE for varying binwidth ($a_1$).
-Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-e).
+#| fig-cap: "Assessing which of the 6 NLDR layouts on the MNIST digit 1 data is the better representation using MSE for varying binwidth ($a_1$).
+Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f).
 All the layouts appear to be very similar.

### 2025-01-15 paper.qmd
Updated Figure 10 title

@@
 ```{r}
-#| fig-cap: "Assessing which of the six NLDR layouts on the two C-shaped clusters data is the better representation using MSE for varying binwidth ($a_1$).
-Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-e).
+#| fig-cap: "Assessing which of the 6 NLDR layouts on the two non-linear clusters data is the better representation using MSE for varying binwidth ($a_1$).
+Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f).
 Layout d is universally poor.

### 2025-01-15 paper.qmd
Updated Figure 8

@@
                          y = MSE)) +
-  geom_vline(xintercept = 0.08,
+  geom_vline(xintercept = 0.09,
            color = "#d95f02", linetype=2, 
            linewidth=1) +
-  geom_vline(xintercept = 0.05,
+  geom_vline(xintercept = 0.06,
              color = "#1f78b4", 
@@
              linewidth=1) +
-  geom_vline(xintercept = 0.02, linetype=2,
+  geom_vline(xintercept = 0.03, linetype=2,
              color = "#1b9e77", linewidth=1) +
@@
   geom_point(size = 1) +
-  scale_x_continuous(breaks = sort(unique(round(error_two_non_linear_diff_shaped_close_clusters_umap$a1, 2)))[c(1, 4, 8, 10, 11, 12, 13, 14)]) +
+  scale_x_continuous(breaks = sort(unique(round(error_two_non_linear_diff_shaped_close_clusters_umap$a1, 2)))[c(2, 5, 8, 10, 12, 14, 15, 16, 17)]) +
   labs(x = expression(paste("binwidth (", a[1], ")")), y = "MSE") +
@@
                          y = prop_comp)) +
-  geom_vline(xintercept = 0.08,
+  geom_vline(xintercept = 0.09,
            color = "#d95f02", linetype=2, 
            linewidth=1) +
-  geom_vline(xintercept = 0.05,
+  geom_vline(xintercept = 0.06,
              color = "#1f78b4", 
@@
   # scale_y_continuous(breaks = sort(unique(error_two_curvy$b_non_empty))[-3]) +
-  scale_x_continuous(breaks = sort(unique(round(error_two_non_linear_diff_shaped_close_clusters_umap$a1, 2)))[c(1, 4, 8, 10, 11, 12, 13, 14)]) +
+  scale_x_continuous(breaks = sort(unique(round(error_two_non_linear_diff_shaped_close_clusters_umap$a1, 2)))[c(2, 5, 8, 10, 12, 14, 15, 16, 17)]) +
   labs(x = expression(paste("binwidth (", a[1], ")")), y = paste("Relative proportion density")) +
@@
                          y = prop_bins)) +
-  geom_vline(xintercept = 0.08,
+  geom_vline(xintercept = 0.09,
            color = "#d95f02", linetype=2, 
            linewidth=1) +
-  geom_vline(xintercept = 0.05,
+  geom_vline(xintercept = 0.06,
              color = "#1f78b4", 
@@
   # scale_y_continuous(breaks = sort(unique(error_two_curvy$b_non_empty))[-3]) +
-  scale_x_continuous(breaks = sort(unique(round(error_two_non_linear_diff_shaped_close_clusters_umap$a1, 2)))[c(1, 4, 8, 10, 11, 12, 13, 14)]) +
+  scale_x_continuous(breaks = sort(unique(round(error_two_non_linear_diff_shaped_close_clusters_umap$a1, 2)))[c(2, 5, 8, 10, 12, 14, 15, 16, 17)]) +
   labs(x = expression(paste("binwidth (", a[1], ")")), y = expression(paste("proportion of non-empty bins ", bgroup("(", frac(m, b), ")")))) +

### 2025-01-15 paper.qmd
Updated Figure 16

@@
 #| fig-pos: H
-#| fig-height: 10
+#| fig-height: 30
 #| fig-width: 10
-#| out-width: 60%
+#| out-width: 100%
 #| fig-cap: "Model in \\gD{}, on the layout a and three views of the fit in projections from $10\\text{-}D$, for the MNIST digit 1 data ($n =  7877$ and $p = 10$).
@@
 free(trimesh_removed_mnist) + mnist_proj_tsne_model1 +
-  mnist_proj_tsne_model2 + mnist_proj_tsne_model3 +
-  plot_layout(ncol = 2)
+  mnist_proj_tsne_model2 + #mnist_proj_tsne_model3 +
+  plot_layout(ncol = 3)
 

### 2025-01-16 paper.qmd
Updated Figure 16 color range for high-d error cluster

@@
 distance_df_small_edges_mnist <- distance_mnist |>
-  filter(distance < benchmark_mnist) |>
-  filter(!(row_number() %in% c(155, 554, 140)))
+  filter(distance < benchmark_mnist) #|>
+  #filter(!(row_number() %in% c(155, 554, 140)))
 
@@
 tsne_minst_scaled_temp <- tsne_minst_scaled |>
-  mutate(cluster = if_else((tSNE1 <= 0.92) & (tSNE1 >= 0.76) & (tSNE2 <= 0.62) & (tSNE2 >= 0.44), "small_clust", "big_clust"))
+  mutate(cluster = if_else((tSNE1 <= 1) & (tSNE1 >= 0.78) & (tSNE2 <= 0.62) & (tSNE2 >= 0.39), "small_clust", "big_clust"))
 
@@
   interior_annotation("a", sc_ltr_pos_mnist) +
-  theme(aspect.ratio = 1,
-        legend.position = "none")
+  theme(legend.position = "none")
 

### 2025-01-16 paper.qmd
Updated Figure17 black box position

@@
   scale_color_manual(values=c("#999999",'#ff7f00')) +
-  interior_annotation("a", sc_ltr_pos_mnist) +
+  interior_annotation("a1", sc_ltr_pos_mnist) +
   theme(legend.position = "none")
@@
   scale_colour_continuous_sequential(palette = "YlOrRd", n_interp = 20) +
-  geom_rect(aes(xmin = 0.8, xmax = 0.9, ymin = 0.45, ymax = 0.6),
+  geom_rect(aes(xmin = 0.82, xmax = 0.92, ymin = 0.38, ymax = 0.58),
                fill = "transparent", color = "#000000", linewidth = 0.7) +
@@
 projection <- cbind(
-    c(-0.09936,0.36550,0.07658,0.29975,0.09085,-0.14009,-0.13417,0.48357,-0.00699,-0.13587),
-    c(0.23061,0.19332,-0.11264,0.14797,0.30351,0.14980,0.02785,-0.30209,-0.13540,-0.43284))
+  c(-0.04295,0.42361,-0.36702,0.13103,-0.19678,-0.29621,0.00471,0.09323,0.18822,-0.18484),
+  c(-0.36644,0.05058,0.18200,0.21860,0.13583,-0.03511,0.35339,0.15821,-0.26945,-0.27926))
 
@@
   limits = 1,
-  axis_pos_x = -0.83,
-  axis_pos_y = -0.83,
+  axis_pos_x = -0.9,
+  axis_pos_y = -0.9,
   axis_labels = names(scaled_mnist_data),
@@
       yend = proj2_to), 
-    color = "#000000") +
+    color = "#000000",
+    linewidth = 0.5,
+    alpha = 0.5) +
   geom_point(
-    #size = 0.2,
-    alpha = 0.1) +
+    size = 0.5,
+    alpha = 0.5) +
   scale_color_manual(values = c("#999999",'#ff7f00')) +
@@
   #coord_fixed() +
-  xlim(c(-1, 0.6)) +
-  ylim(c(-1, 0.6)) +
-  interior_annotation("b") +
+  xlim(c(-1.1, 0.9)) +
+  ylim(c(-1.1, 0.9)) +
+  interior_annotation("a2") +
   theme(
@@
       yend = proj2_to), 
-    color = "#000000") +
+    color = "#000000",
+    linewidth = 0.5) +
   geom_point(
-    #size = 0.2,
+    size = 0.5,
     alpha = 0.3) +
@@
   ylim(c(-1, 1)) +
-  interior_annotation("c") +
+  interior_annotation("a3") +
   theme(
@@
 #| fig-pos: H
-#| fig-height: 30
-#| fig-width: 10
-#| out-width: 100%
-#| fig-cap: "Model in \\gD{}, on the layout a and three views of the fit in projections from $10\\text{-}D$, for the MNIST digit 1 data ($n =  7877$ and $p = 10$).
-There is a big non-linear cluster and a small clusterlocated very close to one corner of the big cluster.
-There is a twisted pattern that is hardly visible in the static plot.
-Video of the langevitour animation are available at <need to add the link>."
+#| fig-height: 4
+#| fig-width: 12
+#| out-height: 100%
+#| fig-cap: "\\gD{} wireframe layout shows that there is a big non-linear cluster (grey) and a small cluster (orange) located very close to the one corner of the big cluster.
+The MNIST digit 1 data ($n =  7877$ and $p = 10$) has a non-linear structure in $10\\text{-}D$.
+Two \\gD{} projection from a tour on $10\\text{-}D$ reveal that the closeness of the clusters in $10\\text{-}D$ and the twisted pattern of the model fit with tSNE ($a_1 = 0.041$, $b = 659/585 \\text{ } (30, \\text{ }48)$).
+Video of the langevitour animation is available at <>."
 
 free(trimesh_removed_mnist) + mnist_proj_tsne_model1 +
-  mnist_proj_tsne_model2 + #mnist_proj_tsne_model3 +
+  #mnist_proj_tsne_model2 + mnist_proj_tsne_model3 +
+  mnist_proj_tsne_model2 +
   plot_layout(ncol = 3)

### 2025-01-16 paper.qmd
Updated Figure 17
Added handwritten digit 1 images with high error inside the non-linear cluster

@@
 #img_error_outside <- c(3554, 501, 4422, 6860, 3869, 1529, 4916, 6144)
-img_error_outside <- c(501, 1529, 6144)
+#img_error_outside <- c(501, 1529, 6144)
+img_error_outside <- c(999, 3650, 4916)
 
@@
 imge_error_sample_outside <- pixels_gathered_outside |>
+  ggplot(aes(x, y, fill = value)) +
+  geom_tile() +
+  facet_wrap(~ instance, ncol = 4) +
+  coord_fixed() +
+  scale_fill_continuous_sequential(palette = "Grays") +
+  theme(strip.background = element_blank(),
+        strip.text.x = element_blank(),
+        legend.position = "none")
+
+##### inside
+img_error_inside <- c(7228, 7363, 7650)
+
+pixels_gathered_inside <-  mnist_data |>
+  mutate(instance = row_number()) |>
+  gather(pixel, value, -Label, -instance) |>
+  extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
+  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
+  filter(instance %in% img_error_inside)
+
+imge_error_sample_inside <- pixels_gathered_inside |>
   ggplot(aes(x, y, fill = value)) +
@@
 
-highd_error_2d_mnist + (imge_error_sample_outside / right_top_img / middle_img / right_bottom_img)
+highd_error_2d_mnist + (imge_error_sample_outside/ imge_error_sample_inside / right_top_img / middle_img / right_bottom_img)
 ```

### 2025-01-16 paper.qmd
Updated red box of 2D layout

@@
                fill = "transparent", color = "#000000", linewidth = 0.7) +
-  geom_rect(aes(xmin = 0.75, xmax = 0.85, ymin = 0.8, ymax = 0.95),
+  geom_rect(aes(xmin = 0.9, xmax = 1, ymin = 1.25, ymax = 1.45),
                fill = "transparent", color = "#e41a1c", linewidth = 0.7)  +
@@
 
-img_right_top <- c(4795, 259, 6564)
+img_right_top <- c(868, 1465, 1787)
 

### 2025-01-16 paper.qmd
Updated Figure 17

@@
                fill = "transparent", color = "#e41a1c", linewidth = 0.7)  +
-  geom_rect(aes(xmin = 0.45, xmax = 0.55, ymin = 0.45, ymax = 0.6),
+  geom_rect(aes(xmin = 0.35, xmax = 0.45, ymin = 0.55, ymax = 0.7),
                fill = "transparent", color = "#4daf4a", linewidth = 0.7)  +
@@
 
-img_middle <- c(1069, 4604, 3579)
+#img_middle <- c(1069, 4604, 3579)
+img_middle <- c(130, 2091, 7795)
 
@@
 #img_right_bottom <- c(1695, 5338, 2805)
-img_right_bottom <- c(2035, 7710, 2517)
+img_right_bottom <- c(2035, 7708, 7710)
 

### 2025-01-16 paper.qmd
Updated Figure 17

@@
   facet_wrap(~ instance, ncol = 4) +
+  ggtitle("(c)") +
   coord_fixed() +
@@
         strip.text.x = element_blank(),
-        legend.position = "none")
+        legend.position = "none",
+        plot.title.position = "plot",
+        plot.title = element_text(hjust = 0.31))
 
@@
   facet_wrap(~ instance, ncol = 4) +
+  ggtitle("(d)") +
   coord_fixed() +
@@
   facet_wrap(~ instance, ncol = 4) +
+  ggtitle("(e)") +
   coord_fixed() +
@@
   facet_wrap(~ instance, ncol = 4) +
+  ggtitle("(a)") +
   coord_fixed() +
@@
   facet_wrap(~ instance, ncol = 4) +
+  ggtitle("(b)") +
   coord_fixed() +
@@
 #| label: fig-model-error-mnist
-#| fig-cap: "The $10\\text{-}D$ model error in layout a of the MNIST digit 1 dataset show a pattern.
-Most low model errors are distributed along the big non-linear clutser, while most large model errors are distributed along the small cluster.
-Along the non-linear cluster, the angle of digit 1 changes.
-Some images have large errors due to their deviation from the non-linear cluster, which makes them anomalies.
-The images associated with large model errors shows different patterns of digit 1."
+#| fig-cap: "The $10\\text{-}D$ model error in \\gD{} layout of the MNIST digit 1 data shows a pattern.
+Most low model errors are distributed along the big non-linear cluster, while most large model errors are distributed along the small cluster.
+The images associated with large model errors shows different patterns of digit 1, some inside (b) the non-linear structure and others outside (a).
+Along the non-linear cluster, the angle of digit 1 changes (c-e)."
 #| fig-pos: H
@@
 
-highd_error_2d_mnist + (imge_error_sample_outside/ imge_error_sample_inside / right_top_img / middle_img / right_bottom_img)
+highd_error_2d_mnist + (imge_error_sample_outside / imge_error_sample_inside / right_top_img / middle_img / right_bottom_img) 
 ```

### 2025-01-16 paper.qmd
Updated Section 7.2 text

@@
 The MSE for different binwidths ($a_1$) using tSNE, UMAP, PHATE, PaCMAP, and TriMAP with default (hyper-)parameter setting (@fig-mnist-mse) were calculated.
-It is found that tSNE (@fig-mnist-mse (a)) provide the most reasonable representation for the digit 1 dataset, showing universally best.
-However, \gD{} representation shows a big non-linear cluster and a small cluster with a small gap (@fig-tsne-best).
+It is found that tSNE (@fig-mnist-mse (f)) provide the most reasonable representation for the digit 1 dataset, showing universally best.
+However, \gD{} representation shows a big non-linear cluster and a small cluster with a small gap.
 
 In the case of digit 1 data, there should not be any clusters unless anomalies exist, indicating different digit 1 patterns.
-The angle of the digit 1 images varies along this non-linear clustering structure (@fig-tsne-best), while the small cluster contains the digit 1 images with different patterns of the digit 1, unlike the usual (@fig-model-error-mnist (c)).
+The angle of the digit 1 images varies along this non-linear clustering structure (@fig-model-error-mnist), while the small cluster contains the digit 1 images with different patterns of the digit 1, unlike the usual (@fig-model-error-mnist (c)).
 This provides the evidence for two close clusters.
@@
 By visualizing the model generated for tSNE in $10\text{-}D$ helps to assess the \gD{} layout.
-As shown in @fig-mnist-tri-proj (a), the model provides the evidence for the non-linear structure of the digit 1 data in $10\text{-}D$.
+As shown in @fig-mnist-tri-proj (a1), the model provides the evidence for the non-linear structure of the digit 1 data in $10\text{-}D$.
 The model shows some quirks.
-The model's twisted pattern provides evidence for the $10\text{-}D$ data structure, which is not observed by \gD{} layout (@fig-mnist-tri-proj (b)).
-Furthermore, the presence of long edges indicates the existence of the small cluster located closely together (@fig-mnist-tri-proj (c)).
+The model's twisted pattern provides evidence for the $10\text{-}D$ data structure, which is not observed by \gD{} layout (@fig-mnist-tri-proj (a2)).
 

### 2025-01-16 paper.qmd
Updated Figure 15
Updated Figure 10

@@
 Layout d is universally poor.
-Layouts a that show two close clusters are universally suboptimal.
+Layouts a, b, e that show two close clusters are universally suboptimal.
+Layout b with little separation performs well at tiny binwidth (where most points are in their own bin) and poorly as binwidth increases.
+Layout e has small separation with oddly shaped clusters.
 Layout a is the best choice."
@@
 The uniformly distributed points demonstrate random model errors throughout the structure, whereas the dense structure shows large model error at the sparse end.
-Videos of the langevitour animations are available at."
+Videos of the langevitour animations are available at <>."
 
@@
 Our approach facilitates the identification of the most accurate NLDR layout, enabling the discovery of cell populations with similar expression profiles.
-The second case study explores the MNIST hand-written digits dataset, which serves as a benchmark for evaluating the preservation of local structures, specifically within a selected digit.
+The second case study explores the MNIST hand-written digits dataset, which serves as a benchmark for evaluating the preservation of local structures, specifically with digit 1.
 
@@
 Layout c is universally poor.
-Layouts f that show two close clusters are universally suboptimal.
+Layouts a, f that show two close clusters are universally suboptimal.
+Layout a with little separation performs well at tiny binwidth (where most points are in their own bin) and poorly as binwidth increases.
 Layout f is the best choice."

### 2025-01-16 paper.qmd
Updated Figure 8 title

@@
 #| echo: false
-#| fig-cap: "Various plots to help assess best hexagon bin parameters and thresholds to remove low-density bins.
+#| fig-cap: "Various plots to help assess best hexagon bin parameters (a, b, d) and thresholds to remove low-density bins (c).
 Both (a) and (c) show MSE, against binwidth ($a_1$) and threshold.
 A good benchmark value for these parameters is when the MSE drops and then flattens out.
-Three binwidth choices were made: $0.05$, $0.09$, and $0.13$ to investigate.
-As the binwidth increases, the proportion of non-empty bins also increases.
-There are two peaks at binwidths $0.09$ and $0.13$.
-The relative proportion density decreases and levels off.
-Binwidths $0.05$ is the favorable choices.
-Based on the MSE, $0.09$ was chosen as the initial best binwidth for further analysis.
+Three binwidth choices were made: $0.03$, $0.06$, and $0.09$ to investigate.
+As the binwidth increases, the proportion of non-empty bins also increases (b).
+The relative proportion density decreases and levels off (d).
+Binwidth $0.06$ is chosen as the initial best binwidth for further analysis.
 There is no need to remove the low-density hexagons because as shown in (b), there is no considerable drop in MSE."

### 2025-01-16 paper.qmd
Updated Figure 6 title

@@
 #| label: best-fit-umap1
-#| fig-cap: "Model in \\gD{} ($a_1 = 0.12$), on the layout and two views of the fit in projections from $7\\text{-}D$, for the two C-shaped clusters data ($n =  2000$ and $p = 7$).
+#| fig-cap: "Model in \\gD{} ($a_1 = 0.12$), on the layout and two views of the fit in projections from $7\\text{-}D$, for the two non-linear clusters data ($n =  2000$ and $p = 7$).
 One view shows the fitted model, and the other shows the fitted and true models alongside the data.
 The fitted model accurately represents the shape of the clusters but does not capture the edges of the structure.
-This illustrates a characteristic of UMAP, which compresses the data when transforming the data into \\gD{}.
+This illustrates a characteristic of tSNE, which compresses the data when transforming the data into \\gD{}.
 Video of the langevitour animations is available at <>."
@@
   geom_density(aes(x=sqrt_row_wise_total_error, y=..density..), colour="red") +
-  xlab(expression(group("|", e[hj], "|"))) +
+  #xlab(expression(group("|", e[hj], "|"))) +
+  xlab(expression(e[hj])) +
   ylab("") +
@@
 Color indicates square root of total error, dark colour indicating high error and light indicates low error.
-Most large errors are distributed near the edges of the clusters."
+Most large errors are distributed near the sparse end of the non-linear cluster with dense corner."
 #| out-width: 60%

### 2025-01-16 paper.qmd
Updated Figure 5
Updated Figure 5 title

@@
 #| label: best-fit-umap1
-#| fig-cap: "Model in \\gD{} ($a_1 = 0.12$), on the layout and two views of the fit in projections from $7\\text{-}D$, for the two non-linear clusters data ($n =  2000$ and $p = 7$).
-One view shows the fitted model, and the other shows the fitted and true models alongside the data.
-The fitted model accurately represents the shape of the clusters but does not capture the edges of the structure.
-This illustrates a characteristic of tSNE, which compresses the data when transforming the data into \\gD{}.
-Video of the langevitour animations is available at <>."
+#| fig-cap: "We start with a tSNE layout (a) of non-linear clusters data ($n =  2000$ and $p = 7$), from which the \\gD{} wireframe is constructed (b) and then lifted to $7\\text{-}D$.
+A \\gD{} projection of the model fit from a tour through the $7\\text{-}D$ space is presented in c.
+Additionally, a \\gD{} projection of the fitted and true models (d) from the same tour demonstrates that the model accurately captures the non-linear structure of the data, although it does not fully represent the width of that structure.
+This highlights a key characteristic of tSNE, which compresses the data when applying tSNE.
+Furthermore, the sparse space observed in the middle of the tSNE layout is a result of (hyper-parameter) choice.
+The default is used.
+Video of the langevitour animation is available at <>."
 #| fig-width: 10

### 2025-01-16 paper.qmd
Updated Figure 14 with more bins

@@
 ## Compute hexbin parameters
-num_bins_x_pbmc <- 26
+num_bins_x_pbmc <- 60
 lim1 <- pbmc_scaled_obj$lim1
@@
 ## Compute hexbin parameters
-num_bins_x_pbmc <- 13
+num_bins_x_pbmc <- 30
 lim1 <- pbmc_scaled_obj$lim1
@@
 distance_df_small_edges_pbmc <- distance_pbmc |>
-  filter(distance < benchmark_pbmc) |>
-  filter(!(row_number() %in% c(6, 12, 14, 16, 17, 19, 23, 29, 30, 143, 144, 146, 147, 157)))
+  filter(distance < benchmark_pbmc) #|>
+  #filter(!(row_number() %in% c(6, 12, 14, 16, 17, 19, 23, 29, 30, 143, 144, 146, 147, 157)))
 

### 2025-01-17 paper.qmd
Updated the arrangments

@@
 #| fig-pos: H
-#| out-width: 80%
+#| out-width: 100%
 
@@
 #| echo: false
-#| fig-cap: "Comapre the author suggested layout and the layout selected by our approach.
-Model in \\gD{}, on the layout a and b, and two views of the fit in projections from $9\\text{-}D$, for the PBMC3k data ($n =  2622$ and $p = 9$).
-Layout a shows three-well separated clusters with big separation, while layout b shows close clusters.
-In $9\\text{-}D$, the data shows three clusters with two of them being close and one separated from the others.
-This is why b1, b2, and b3 shows a connected edge between the close clusters, as proven by @fig-pbmc-mse as a reasonable layout.
-Viewing the fitted model in $9\\text{-}D$, helps to see some unobserved patterns of the data: (i) dense points, and (ii) non-linear clusters.
+#| fig-cap: "Compare the author suggested \\gD{} layout (@fig-pbmc-mse a) and the \\gD{} layout selected (@fig-pbmc-mse b) by MSE plot (@fig-pbmc-mse) from the same NLDR with different hyper-parameter choice.
+The PBMC3k data ($n =  2622$) has three clusters in $9\\text{-}D$, where two of them being close and one separated from the others.
+Two \\gD{} projection from a tour on $9\\text{-}D$ of the model fit with @fig-pbmc-mse a ($a_1 = 0.02$, $b = 3480/349 (60, 58)$) shows three-well separated clusters with big separations.
+On the other hand, the model fit with @fig-pbmc-mse b ($a_1 = 0.04$, $b = 1020/347 (30, 34)$) shows two close clusters, and one separated cluster.
+The edge connected the two clusters proves the closeness of the two clusters.
+Therefore, @fig-pbmc-mse b is more reasonable than @fig-pbmc-mse a.
+Also, the fitted models in $9\\text{-}D$ help to see some unobserved patterns of the data: (i) dense points, and (ii) non-linear clusters.
 Videos of the langevitour animations are available at <https://youtu.be/0cKX_HG_n0k> and <https://youtu.be/KhJvsRtaX04> respectively."
@@
 #| fig-pos: H
-#| out-width: 80%
+#| out-width: 92%
 

### 2025-01-17 appendix.qmd
Update Figure 2 of appendix section 2

--- 2024-10-07 appendix.qmd
@@
 ```{r}
-all_centroids_df <- gen_centroids(bin1 = 4, r2 = r2, q = 0.1)
+# all_centroids_df <- gen_centroids(bin1 = 4, r2 = r2, q = 0.1)
+# 
+# hex_grid <- gen_hex_coord(all_centroids_df, a1 = 1.043789)
 
-hex_grid <- gen_hex_coord(all_centroids_df, a1 = 1.043789)
+## Compute hexbin parameters
+num_bins_x_scurve <- 3
+
+## hexagon binning to have regular hexagons
+hb_obj_scurve <- hex_binning(
+  data = umap_scurve_scaled, 
+  bin1 = num_bins_x_scurve, 
+  r2 = r2)
+
+## Data set with all centroids
+all_centroids_df <- hb_obj_scurve$centroids
+
+## Generate all coordinates of hexagons
+hex_grid <- hb_obj_scurve$hex_poly
 
@@
 $$
-l = \frac{-2a_2 \pm \sqrt{4a_2^2 - 24[-(a_1^2 + a_2^2)]}}{6}
+3l^2 + 2a_2l - (a_1^2 + a_2^2) = 0.
 $$
+
+Finally, by solving the quadratic equation, we compute,
+
 $$
-l = \frac{-a_2 \pm \sqrt{a_2^2 - 6[-(a_1^2 + a_2^2)]}}{3}
+l = \frac{-2a_2 \pm \sqrt{4a_2^2 - 24[-(a_1^2 + a_2^2)]}}{6},
 $$
+
+$$
+l = \frac{-a_2 \pm \sqrt{a_2^2 - 6[-(a_1^2 + a_2^2)]}}{3},
+$$
+
+where $l > 0$.
+
+<!-- #### Comparison with results of scDEED recommendations -->
+
+<!-- <!-- -->
+<!-- XXX What this section needs: -->
+
+<!-- - connect this with previous section, not make it seem like a completely new. connections are: nldr used to show clustering, scDEED provides a different approach to decide which is best. -->
+<!-- - why these two? assume one is the published figure, and second is one recommended by scDEED, which our results confirm is better. -->
+<!-- - Refer to figure in text -->
+<!-- - Fig 16 not necessary -->
+
+<!-- CHANGES WERE MADE BASED ON THIS THINKING -->
+<!-- --> -->
+
+<!-- <!-- In the field of single-cell studies, clustering is a common analytical task used to identify groups of cells with similar expression profiles.
+Non-linear dimensional reduction (NLDR) methods are frequently employed to visualize these clusters and help validate the results.
+However, it is well known that the 2D embeddings produced by t-SNE and UMAP may not accurately reflect the similarities among cell clusters.
+ALREADY STATED IN PREVIOUS SECTION -->  -->
+
+<!-- As we were writing this paper @xia2023 appeared proposing a new method called scDEED helping to assess the validity of a \gD{} embedding.
+scDEED calculates a reliability score for each cell embedding based on the similarity between the cell’s \gD{} embedding neighbors and its neighbors prior to embedding.
+A low reliability score suggests a dubious embedding.
+It can help in the deciding on optimal hyperparameters.
+Here we illustrate how our method compares with the results from scDEED -->
+
+<!-- Following the process in @xia2023 again using the Human Peripheral Blood Mononuclear Cells (PBMC) data.
+Note that @xia2023 uses a different subset from @sec-pbmc, which contains $31,021$ cells including cell type labels, and the gene expression levels were in the unit of log-transformed UMI count per $10,000$.
+They focused on three sequencing methods (inDrops, DropSeq, and SeqWell) and four common cell types Cytotoxic T cell, CD4+T cell, CD14+ Monocyte, and B cell.
+-->
+
+<!-- For illustration purposes, we only selected cells generated with inDrops ($n=5858$ cells) and UMAP cell embeddings.
+Also, @xia2023 used first $50$ principal components to generate the UMAP.
+The objective is to assess the optimized layout by scDEED, and if it does not accurately represent the three clusters with small separations of the PBMC dataset, then select a reasonable \gD{} layout.
+-->
+
+<!-- @fig-model-pbmc-author-proj-scdeed shows layout from the original published article along with that suggested by scDEED to be more accurate.
+The log MSE vs binwidth plot illustrates that our approach would suggest that scDEED is correct here, that layout b is more accurately reflecting the cluster structure in the data.
+-->
+
+<!-- ```{r} -->
+<!-- result_umap <- read_rds(here::here('data/pbmc3k/pbmc_scdeed_umap_results.rds')) -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- #| eval: false -->
+<!-- result_umap |> -->
+<!--   kableExtra::kable(format = "latex", -->
+<!--                     booktabs = TRUE, -->
+<!--                     longtable = TRUE, -->
+<!--                     label = "scdeedumap") |> -->
+<!--   kableExtra::kable_styling(latex_options = "scale_down") -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- #| out-width: 50% -->
+<!-- #| eval: false -->
+
+<!-- result_umap |> -->
+<!--   filter(min.dist == 0.5) |> -->
+<!--   ggplot(aes( -->
+<!--            x = as.numeric(n_neighbors), -->
+<!--            y = number_dubious_cells -->
+<!--          )) + -->
+<!--     geom_line() + -->
+<!--     geom_point() + -->
+<!--     labs(x = "n_neighbors", -->
+<!--          y = "# of dubious cells") + -->
+<!--     theme_minimal() -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- #| label: read-pbmc-umap-layouts -->
+<!-- # Read a variety of different NLDR representations of PBMC -->
+<!-- # and plot them on same aspect ratio -->
+<!-- umap_pbmc1 <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_30_min_dist_0.3.rds") -->
+<!-- umap_pbmc1 <- as_tibble(umap_pbmc1) -->
+<!-- names(umap_pbmc1) <- c("UMAP1", "UMAP2") -->
+
+<!-- nldr1c <- umap_pbmc1 |> -->
+<!--   ggplot(aes(x = UMAP1, -->
+<!--              y = UMAP2)) + -->
+<!--   geom_point(alpha=0.1, size=1, colour='#e41a1c') + -->
+<!--   interior_annotation("a") -->
+
+<!-- umap_pbmc2 <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_80_min_dist_0.5.rds") -->
+<!-- umap_pbmc2 <- as_tibble(umap_pbmc2) -->
+<!-- names(umap_pbmc2) <- c("UMAP1", "UMAP2") -->
+
+<!-- nldr2c <- umap_pbmc2 |> -->
+<!--   ggplot(aes(x = UMAP1, -->
+<!--              y = UMAP2))+ -->
+<!--   geom_point(alpha=0.1, size=1, colour='#377eb8') + -->
+<!--   interior_annotation("b") -->
+
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- #| label: combine-data-pbmc-umap -->
+
+<!-- error_pbmc_umap <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_30_min_dist_0.3.rds") -->
+<!-- error_pbmc_umap2 <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_80_min_dist_0.5.rds") -->
+
+<!-- error_pbmc <- bind_rows(error_pbmc_umap,  -->
+<!--                         error_pbmc_umap2) -->
+
+<!-- error_pbmc <- error_pbmc |> -->
+<!--   mutate(a1 = round(a1, 2)) |> -->
+<!--   filter(bin1 >= 5) |> -->
+<!--   group_by(method, a1) |> -->
+<!--   filter(MSE == min(MSE)) |> -->
+<!--   ungroup() -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- #| label: error-comp-pbmc-umap -->
+
+<!-- error_umap_plot_pbmc <- ggplot(error_pbmc,  -->
+<!--                                aes(x = a1,  -->
+<!--                                    y = MSE,  -->
+<!--                                    colour = method)) +  -->
+<!--   geom_point(size = 0.8) + -->
+<!--   geom_line(linewidth = 0.3) +  -->
+<!--   # geom_vline(xintercept = 15, linetype="solid",  -->
+<!--   #            color = "black", linewidth=0.8, alpha = 0.5) + -->
+<!--   scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 12, 15, 17, 19)]) + -->
+<!--   scale_color_manual(values=c('#e41a1c','#377eb8')) + -->
+<!--   scale_y_log10() + -->
+<!--   ylab("log(MSE)") + -->
+<!--   xlab(expression(paste("binwidth (", a[1], ")"))) + -->
+<!--   theme_minimal() + -->
+<!--   theme(panel.border = element_rect(fill = 'transparent'), -->
+<!--         plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5), -->
+<!--         axis.ticks.x = element_line(), -->
+<!--         axis.ticks.y = element_line(), -->
+<!--         legend.position = "none", -->
+<!--         axis.text.x = element_text(size = 7), -->
+<!--         axis.text.y = element_text(size = 7), -->
+<!--         axis.title.x = element_text(size = 7), -->
+<!--         axis.title.y = element_text(size = 7))  -->
+
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- #| label: fig-pbmc-mse-umap -->
+<!-- #| fig-pos: H -->
+<!-- #| out-width: 100% -->
+<!-- #| fig-cap: "Assessing which of the 2 UMAP layouts with different hyperparameter settings (n_neighbors: $30$, min_dist: $0.3$ (red), n_neighbors: $80$, min_dist: $0.5$ (blue)) on the PBMC data is the better representation using MSE for varying binwidth ($a_1$).
+Colour used for the lines and points in the left plot and in the scatterplots represents UMAP layout (a, b).
+Of the two, layout b is optimal across all binwidths making it the best choice." -->
+
+<!-- free(error_umap_plot_pbmc) + wrap_plots(nldr1c, nldr2c, ncol = 2) + -->
+<!--   plot_layout(ncol = 1, heights = c(1, 1)) -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- #| label: published-pbmc-scdeed -->
+
+<!-- training_data_pbmc <- read_rds("data/pbmc3k/pbmc_pca_50_scdeed.rds") -->
+<!-- training_data_pbmc <- training_data_pbmc[, 1:50] |> -->
+<!--   mutate(ID = 1:NROW(training_data_pbmc)) -->
+
+<!-- ## For umap -->
+<!-- umap_pbmc <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_80_min_dist_0.5.rds") -->
+<!-- umap_pbmc <- as_tibble(umap_pbmc) -->
+<!-- names(umap_pbmc) <- c("UMAP1", "UMAP2") -->
+<!-- umap_pbmc <- umap_pbmc |> -->
+<!--   mutate(ID = 1:NROW(umap_pbmc)) -->
+
+<!-- pbmc_scaled_obj_umap <- gen_scaled_data( -->
+<!--   data = umap_pbmc) -->
+<!-- umap_pbmc_scaled <- pbmc_scaled_obj_umap$scaled_nldr -->
+
+<!-- lim1 <- pbmc_scaled_obj_umap$lim1 -->
+<!-- lim2 <- pbmc_scaled_obj_umap$lim2 -->
+<!-- r2_umap <- diff(lim2)/diff(lim1) -->
+
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- #| label: hexbin-pbmc-scdeed -->
+<!-- ## Compute hexbin parameters -->
+<!-- num_bins_x_pbmc <- 12 -->
+<!-- lim1 <- pbmc_scaled_obj_umap$lim1 -->
+<!-- lim2 <- pbmc_scaled_obj_umap$lim2 -->
+<!-- r2_pbmc <- diff(lim2)/diff(lim1)  -->
+
+<!-- pbmc_model <- fit_highd_model( -->
+<!--   training_data = training_data_pbmc, -->
+<!--   emb_df = umap_pbmc_scaled, -->
+<!--   bin1 = num_bins_x_pbmc, -->
+<!--   r2 = r2_pbmc, -->
+<!--   q = 0.1, -->
+<!--   is_bin_centroid = TRUE, -->
+<!--   is_rm_lwd_hex = FALSE, -->
+<!--   col_start_highd = "pc" -->
+<!-- ) -->
+
+<!-- df_bin_centroids_pbmc <- pbmc_model$df_bin_centroids -->
+<!-- df_bin_pbmc <- pbmc_model$df_bin -->
+
+<!-- ## Triangulate bin centroids -->
+<!-- tr1_object_pbmc <- tri_bin_centroids( -->
+<!--   df_bin_centroids_pbmc, x = "c_x", y = "c_y") -->
+<!-- tr_from_to_df_pbmc <- gen_edges( -->
+<!--   tri_object = tr1_object_pbmc) -->
+
+<!-- ## Compute 2D distances -->
+<!-- distance_pbmc <- cal_2d_dist( -->
+<!--   tr_coord_df = tr_from_to_df_pbmc, -->
+<!--   start_x = "x_from", -->
+<!--   start_y = "y_from", -->
+<!--   end_x = "x_to", -->
+<!--   end_y = "y_to", -->
+<!--   select_vars = c("from", "to", "distance")) -->
+
+<!-- ## To find the benchmark value -->
+<!-- benchmark_pbmc <- find_lg_benchmark( -->
+<!--   distance_edges = distance_pbmc, -->
+<!--   distance_col = "distance") -->
+
+<!-- tr_df <- distinct(tibble( -->
+<!--   x = c(tr_from_to_df_pbmc[["x_from"]], tr_from_to_df_pbmc[["x_to"]]),  -->
+<!--   y = c(tr_from_to_df_pbmc[["y_from"]], tr_from_to_df_pbmc[["y_to"]]))) -->
+
+<!-- distance_df_small_edges_pbmc <- distance_pbmc |> -->
+<!--   filter(distance < benchmark_pbmc) -->
+
+<!-- tr_from_to_df_pbmc <- inner_join( -->
+<!--   tr_from_to_df_pbmc, distance_df_small_edges_pbmc,  -->
+<!--   by = c("from", "to")) -->
+
+<!-- trimesh_removed_pbmc <- ggplot() +  -->
+<!--   geom_segment(data = tr_from_to_df_pbmc,  -->
+<!--                aes( -->
+<!--                  x = x_from,  -->
+<!--                  y = y_from,  -->
+<!--                  xend = x_to,  -->
+<!--                  yend = y_to), -->
+<!--                colour = "#000000", -->
+<!--                linewidth = 1) + -->
+<!--   geom_point( -->
+<!--     data = umap_pbmc_scaled, -->
+<!--     aes( -->
+<!--       x = UMAP1, -->
+<!--       y = UMAP2 -->
+<!--     ), -->
+<!--     alpha = 0.2, -->
+<!--     color = '#a65628' -->
+<!--   )  + -->
+<!--   interior_annotation("a1", cex = 2) + -->
+<!--   theme( -->
+<!--     aspect.ratio = 1 -->
+<!--   ) -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- #| label: pbmc-umap-model-proj-scdeed -->
+
+<!-- data_pbmc <- training_data_pbmc |>  -->
+<!--   select(-ID) |> -->
+<!--   mutate(type = "data") -->
+
+<!-- df_b_pbmc <- df_bin_pbmc |> -->
+<!--   dplyr::filter(hb_id %in% df_bin_centroids_pbmc$hexID) |> -->
+<!--   dplyr::mutate(type = "model") ## Data with summarized mean -->
+
+<!-- ## Reorder the rows of df_b according to the hexID order in df_b_with_center_data -->
+<!-- df_b_pbmc <- df_b_pbmc[match(df_bin_centroids_pbmc$hexID, df_b_pbmc$hb_id),] |> -->
+<!--   dplyr::select(-hb_id)  -->
+
+<!-- # Apply the scaling -->
+<!-- df_model_data_pbmc <- bind_rows(data_pbmc, df_b_pbmc) -->
+<!-- scaled_pbmc <- scale_data_manual(df_model_data_pbmc, "type") |> -->
+<!--   as_tibble() -->
+
+<!-- scaled_pbmc_data <- scaled_pbmc |> -->
+<!--   filter(type == "data") |> -->
+<!--   select(-type) -->
+
+<!-- scaled_pbmc_data_model <- scaled_pbmc |> -->
+<!--   filter(type == "model") |> -->
+<!--   select(-type) -->
+
+<!-- # Combine with the true model for visualization -->
+<!-- df <- dplyr::bind_rows(scaled_pbmc_data_model |> mutate(type = "model"), -->
+<!--                        scaled_pbmc_data |> mutate(type = "data")) -->
+
+<!-- ## First projection -->
+<!-- projection <- cbind( -->
+<!--   c(0.040856,0.000723,0.006000,0.007623,0.001986,-0.006599,0.005311,-0.000222,0.001648,0.008067,-0.001587,0.004423,0.004851,0.000638,0.002606,0.006616,0.006188,-0.000954,-0.004603,0.004998,-0.012220,0.008401,-0.007788,0.008972,-0.003903,0.003348,-0.001817,-0.000385,-0.002437,-0.004097,-0.001357,0.013660,0.001770,0.006982,0.002828,-0.009169,-0.003541,0.008582,-0.000032,-0.002375,0.005854,-0.005466,-0.000824,-0.001435,0.004148,0.007141,0.000883,-0.004159,0.005401,0.007882), -->
+<!--   c(-0.000590,0.039365,-0.004481,-0.009041,-0.003805,0.003278,-0.009601,-0.000650,-0.002333,-0.005697,-0.001488,-0.009808,0.002557,0.004113,0.005705,-0.004665,-0.001170,-0.004727,-0.000845,0.005084,-0.018054,0.002538,0.006606,-0.007184,0.000238,0.001877,-0.002124,-0.005972,-0.009729,0.004958,-0.003181,-0.005185,-0.009598,-0.004243,0.007716,0.000396,-0.004258,0.005932,-0.003242,-0.004184,0.003493,-0.008405,0.000860,0.008304,0.004569,0.000798,0.004156,-0.005777,0.005278,-0.002531)) -->
+
+<!-- projection_scaled <- projection * 3 -->
+
+<!-- # projection_filtered <- projection -->
+<!-- # projection_filtered[apply(projection_filtered, 1, function(row) all(row < 0)), ] <- 0.75 -->
+
+<!-- #rows_to_keep <- c(1, 2, 4, 5, 6) -->
+
+<!-- #projection_filtered <- projection[rows_to_keep, ] -->
+
+<!-- projected <- as.matrix(scaled_pbmc_data) %*% projection_scaled -->
+
+<!-- projected_df <- projected |> -->
+<!--   tibble::as_tibble(.name_repair = "unique") |> -->
+<!--   dplyr::rename(c("proj1" = "...1", -->
+<!--                   "proj2" = "...2")) |> -->
+<!--   #dplyr::mutate(type = df_exe$type) |> -->
+<!--   dplyr::mutate(ID = dplyr::row_number()) -->
+
+<!-- projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection_scaled -->
+
+<!-- projected_model_df <- projected_model |> -->
+<!--   tibble::as_tibble(.name_repair = "unique") |> -->
+<!--   dplyr::rename(c("proj1" = "...1", -->
+<!--                   "proj2" = "...2")) |> -->
+<!--   dplyr::mutate(ID = dplyr::row_number()) -->
+
+<!-- model_df <- dplyr::left_join( -->
+<!--   distance_df_small_edges_pbmc |> select(-distance),  -->
+<!--   projected_model_df,  -->
+<!--   by = c("from" = "ID")) -->
+
+<!-- names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from") -->
+
+<!-- model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID")) -->
+<!-- names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to") -->
+
+<!-- axes_obj <- gen_axes( -->
+<!--   proj = projection, -->
+<!--   limits = 0.04, -->
+<!--   axis_pos_x = -0.025, -->
+<!--   axis_pos_y = -0.025, -->
+<!--   axis_labels = names(scaled_pbmc_data), -->
+<!--   threshold = 0.0001) -->
+
+<!-- axes <- axes_obj$axes -->
+<!-- circle <- axes_obj$circle -->
+
+<!-- pbmc_proj_umap_model1 <- projected_df |> -->
+<!--   ggplot( -->
+<!--     aes( -->
+<!--       x = proj1,  -->
+<!--       y = proj2)) + -->
+<!--   geom_segment( -->
+<!--     data = model_df,  -->
+<!--     aes( -->
+<!--       x = proj1_from,  -->
+<!--       y = proj2_from,  -->
+<!--       xend = proj1_to,  -->
+<!--       yend = proj2_to),  -->
+<!--     color = "#000000") + -->
+<!--   geom_point( -->
+<!--     #size = 0.5, -->
+<!--     alpha = 0.2, -->
+<!--     color = '#a65628') + -->
+<!--   geom_segment( -->
+<!--     data=axes,  -->
+<!--     aes(x=x1, y=y1, xend=x2, yend=y2),  -->
+<!--     colour="grey70") + -->
+<!--   geom_text( -->
+<!--     data=axes, -->
+<!--     aes(x=x2, y=y2),  -->
+<!--     label=rownames(axes), -->
+<!--     colour="grey50", -->
+<!--     size = 3) + -->
+<!--   geom_path( -->
+<!--     data=circle,  -->
+<!--     aes(x=c1, y=c2), colour="grey70") + -->
+<!--   coord_fixed() + -->
+<!--   xlim(c(-0.03, 0.06)) + -->
+<!--   ylim(c(-0.03, 0.06)) + -->
+<!--   interior_annotation("a2", cex = 2) -->
+
+<!-- ## Second projection -->
+<!-- projection <- cbind( -->
+<!--   c(-0.019825,-0.011101,0.011543,-0.006943,-0.004485,0.000062,0.002023,-0.005178,0.000800,-0.002903,-0.014182,0.008756,0.009595,0.007032,-0.007820,0.006143,0.004865,0.003352,0.007554,0.002265,0.012035,0.009961,-0.001825,-0.004942,-0.002237,0.002308,0.009531,-0.001146,0.013401,-0.001022,0.008434,0.000477,0.004239,-0.006772,-0.004702,-0.011432,-0.000431,0.008146,-0.009299,-0.004782,0.003457,-0.008214,-0.009604,0.011329,0.003805,-0.003004,-0.000800,0.015288,-0.013157,0.006984), -->
+<!--   c(-0.001387,0.001643,0.010470,0.002930,-0.002335,-0.010569,0.013839,0.002917,-0.006825,0.010191,-0.006367,0.002220,0.002506,0.005139,-0.007213,-0.006235,-0.007123,-0.000222,-0.002225,0.011592,-0.003942,0.002788,-0.000055,0.011316,-0.009753,0.020785,0.004615,0.005172,0.011182,-0.003125,-0.010860,0.007820,-0.017510,0.007010,0.007267,0.002694,-0.003949,-0.009188,-0.005163,0.001898,-0.004744,0.013757,0.010630,-0.007688,0.004598,0.004658,-0.011154,0.003584,-0.005226,0.006650)) -->
+
+<!-- projection_scaled <- projection * 3 -->
+
+<!-- projected <- as.matrix(scaled_pbmc_data) %*% projection_scaled -->
+
+<!-- projected_df <- projected |> -->
+<!--   tibble::as_tibble(.name_repair = "unique") |> -->
+<!--   dplyr::rename(c("proj1" = "...1", -->
+<!--                   "proj2" = "...2")) |> -->
+<!--   #dplyr::mutate(type = df_exe$type) |> -->
+<!--   dplyr::mutate(ID = dplyr::row_number()) -->
+
+<!-- projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection_scaled -->
+
+<!-- projected_model_df <- projected_model |> -->
+<!--   tibble::as_tibble(.name_repair = "unique") |> -->
+<!--   dplyr::rename(c("proj1" = "...1", -->
+<!--                   "proj2" = "...2")) |> -->
+<!--   dplyr::mutate(ID = dplyr::row_number()) -->
+
+<!-- model_df <- dplyr::left_join( -->
+<!--   distance_df_small_edges_pbmc |> select(-distance),  -->
+<!--   projected_model_df,  -->
+<!--   by = c("from" = "ID")) -->
+
+<!-- names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from") -->
+
+<!-- model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID")) -->
+<!-- names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to") -->
+
+<!-- axes_obj <- gen_axes( -->
+<!--   proj = projection, -->
+<!--   limits = 0.03, -->
+<!--   axis_pos_x = -0.02, -->
+<!--   axis_pos_y = -0.02, -->
+<!--   axis_labels = names(scaled_pbmc_data), -->
+<!--   threshold = 0) -->
+
+<!-- axes <- axes_obj$axes -->
+<!-- circle <- axes_obj$circle -->
+
+<!-- pbmc_proj_umap_model2 <- projected_df |> -->
+<!--   ggplot( -->
+<!--     aes( -->
+<!--       x = proj1, -->
+<!--       y = proj2)) + -->
+<!--   geom_segment( -->
+<!--     data = model_df, -->
+<!--     aes( -->
+<!--       x = proj1_from, -->
+<!--       y = proj2_from, -->
+<!--       xend = proj1_to, -->
+<!--       yend = proj2_to), -->
+<!--     color = "#000000") + -->
+<!--   geom_point( -->
+<!--     #size = 0.5, -->
+<!--     alpha = 0.2, -->
+<!--     color = '#a65628') + -->
+<!--   geom_segment( -->
+<!--     data=axes, -->
+<!--     aes(x=x1, y=y1, xend=x2, yend=y2), -->
+<!--     colour="grey70") + -->
+<!--   geom_text( -->
+<!--     data=axes, -->
+<!--     aes(x=x2, y=y2), -->
+<!--     label=rownames(axes), -->
+<!--     colour="grey50", -->
+<!--     size = 3) + -->
+<!--   geom_path( -->
+<!--     data=circle, -->
+<!--     aes(x=c1, y=c2), colour="grey70") + -->
+<!--   coord_fixed() + -->
+<!--   xlim(c(-0.025, 0.02)) + -->
+<!--   ylim(c(-0.025, 0.02)) + -->
+<!--   interior_annotation("a3", cex = 2) -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- #| echo: false -->
+<!-- #| eval: false -->
+<!-- #| fig-cap: "Model in \\gD{}, on the layout a, and two views of the fit in projections from $50\\text{-}D$, for the PBMC data ($n =  5858$ and $p = 50$).
+Layout shows close clusters.
+In $50\\text{-}D$, the data shows " -->
+<!-- #| label: fig-model-pbmc-author-proj-scdeed -->
+<!-- #| fig-pos: H -->
+<!-- #| fig-width: 15 -->
+<!-- #| fig-height: 5 -->
+<!-- #| out-width: 80% -->
+
+<!-- free(trimesh_removed_pbmc) +  -->
+<!--   pbmc_proj_umap_model1 +  -->
+<!--   pbmc_proj_umap_model2 +  -->
+<!--   plot_layout(ncol=3) & -->
+<!--   theme(legend.position='none') -->
+<!-- ``` -->

### 2025-01-17 appendix.qmd
Updated bin circles
Updated Figure 2

@@
 ```{r}
-# all_centroids_df <- gen_centroids(bin1 = 4, r2 = r2, q = 0.1)
-# 
-# hex_grid <- gen_hex_coord(all_centroids_df, a1 = 1.043789)
-
 ## Compute hexbin parameters
@@
 circles <- data.frame(
-  x = c(0.422, 0.683),
-  y = c(0.740, 1.19),
-  r = rep(0.301316, 2)
+  x = c(0.726, 0.451),
+  y = c(1.27, 0.790),
+  r = rep(0.32, 2)
 )
 
-emb_df1 <- tibble(emb1 = 0.58, emb2 = 0.93) |>
-  add_row(emb1 = 0.5, emb2 = 1.01)
+emb_df1 <- tibble(emb1 = 0.61, emb2 = 0.99) |>
+  add_row(emb1 = 0.54, emb2 = 1.08)
 
@@
   geom_circle(data = circles, aes(x0 = x, y0 = y, r = r), fill = NA) +
-  geom_segment(x = 0.58, y = 0.93, xend = 0.44, yend = 0.76, 
+  geom_segment(x = 0.61, y = 0.99, xend = 0.46, yend = 0.83, 
                arrow = arrow(length = unit(0.02, "npc")), color = "#969696") +
-  geom_segment(x = 0.5, y = 1.01, xend = 0.68, yend = 1.16, 
+  geom_segment(x = 0.54, y = 1.08, xend = 0.7, yend = 1.2, 
                arrow = arrow(length = unit(0.02, "npc")), color = "#969696")

### 2025-01-17 appendix.qmd
Updated Figure 2

@@
 library(kableExtra)
+library(tidyr)
 
@@
 
-emb_df2 <- tibble(emb1 = 0.161, emb2 = 0.7395) |>
-  add_row(emb1 = 0.683, emb2 = 0.7395) |>
-  add_row(emb1 = 0.2915, emb2 = 0.965) |>
-  add_row(emb1 = 0.5525, emb2 = 0.965) |>
-  add_row(emb1 = 0.2915, emb2 = 0.5135) |>
-  add_row(emb1 = 0.5525, emb2 = 0.5135) |>
-  add_row(emb1 = 0.422, emb2 = 1.04) |>
-  add_row(emb1 = 0.161, emb2 = 0.890) |>
-  add_row(emb1 = 0.161, emb2 = 0.589) |>
-  add_row(emb1 = 0.422, emb2 = 0.438) |>
-  add_row(emb1 = 0.683, emb2 = 0.589) |>
-  add_row(emb1 = 0.683, emb2 = 0.890)
+hex_data <- hex_grid |> 
+  dplyr::filter(hex_poly_id == 8)
+
+# Create all pairs of rows
+pairs <- expand_grid(row1 = 1:nrow(hex_data), row2 = 1:nrow(hex_data)) |>
+  filter(row1 < row2)
+
+# Compute middle points for each pair of rows
+middle_points <- pairs |>
+  rowwise() |>
+  mutate(
+    hex_poly_id = hex_data$hex_poly_id[row1],
+    x = (hex_data$x[row1] + hex_data$x[row2]) / 2,
+    y = (hex_data$y[row1] + hex_data$y[row2]) / 2
+  ) |>
+  ungroup() |>
+  select(hex_poly_id, x, y) |>
+  filter(row_number() %in% c(1, 5, 6, 10, 13, 15))
 
@@
   geom_polygon(fill = "white", color = "#bdbdbd", aes(group = hex_poly_id)) +
-  geom_point(data = emb_df2, aes(x = emb1, y = emb2), color = "black", size = 0.7) +
+  geom_point(data = middle_points, aes(x = x, y = y), color = "black", size = 0.7) +
   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID), size = 5, colour = "#bdbdbd") +

### 2025-01-17 appendix.qmd
Updated Figure 2

@@
 
+points_df <- bind_rows(hex_data, middle_points)
+
 assign_data_vis2 <- ggplot(data = hex_grid, aes(x = x, y = y)) + 
   geom_polygon(fill = "white", color = "#bdbdbd", aes(group = hex_poly_id)) +
-  geom_point(data = middle_points, aes(x = x, y = y), color = "black", size = 0.7) +
+  geom_point(data = points_df, aes(x = x, y = y), color = "black", size = 0.7) +
   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID), size = 5, colour = "#bdbdbd") +

### 2025-01-17 appendix.qmd
Updated Figure 2

@@
   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID), size = 5, colour = "#bdbdbd") +
-  geom_segment(x = 0.161, y = 0.7395, xend = -0.06, yend = 0.740, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.683, y = 0.7395, xend = 0.45, yend = 0.740, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.2915, y = 0.965, xend = 0.43, yend = 0.79, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.5525, y = 0.965, xend = 0.44, yend = 0.8, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.2915, y = 0.5135, xend = 0.18, yend = 0.32, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.5525, y = 0.5135, xend = 0.67, yend = 0.32, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.422, y = 1.04, xend = 0.42, yend = 0.85, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.161, y = 0.890, xend = -0.02, yend = 0.8, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.161, y = 0.589, xend = 0.161, yend = 0.4, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.422, y = 0.438, xend = 0.25, yend = 0.3, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.683, y = 0.589, xend = 0.683, yend = 0.4, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.683, y = 0.890, xend = 0.5, yend = 0.82, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd")
+  geom_segment(x = 0.175, y = 0.790, xend = -0.06, yend = 0.790, 
+               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") #+
+  # geom_segment(x = 0.683, y = 0.7395, xend = 0.45, yend = 0.740, 
+  #              arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  # geom_segment(x = 0.2915, y = 0.965, xend = 0.43, yend = 0.79, 
+  #              arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  # geom_segment(x = 0.5525, y = 0.965, xend = 0.44, yend = 0.8, 
+  #              arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  # geom_segment(x = 0.2915, y = 0.5135, xend = 0.18, yend = 0.32, 
+  #              arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  # geom_segment(x = 0.5525, y = 0.5135, xend = 0.67, yend = 0.32, 
+  #              arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  # geom_segment(x = 0.422, y = 1.04, xend = 0.42, yend = 0.85, 
+  #              arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  # geom_segment(x = 0.161, y = 0.890, xend = -0.02, yend = 0.8, 
+  #              arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  # geom_segment(x = 0.161, y = 0.589, xend = 0.161, yend = 0.4, 
+  #              arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  # geom_segment(x = 0.422, y = 0.438, xend = 0.25, yend = 0.3, 
+  #              arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  # geom_segment(x = 0.683, y = 0.589, xend = 0.683, yend = 0.4, 
+  #              arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  # geom_segment(x = 0.683, y = 0.890, xend = 0.5, yend = 0.82, 
+  #              arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd")
 

### 2025-01-17 appendix.qmd
Updated Figure 2

--- 2025-01-17 appendix.qmd
@@
   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID), size = 5, colour = "#bdbdbd") +
-  geom_segment(x = 0.161, y = 0.7395, xend = -0.06, yend = 0.740, 
+  geom_segment(x = 0.175, y = 0.790, xend = -0.06, yend = 0.790, 
                arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.683, y = 0.7395, xend = 0.45, yend = 0.740, 
+  geom_segment(x = 0.726, y = 0.790, xend = 0.50, yend = 0.790,
                arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.2915, y = 0.965, xend = 0.43, yend = 0.79, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.5525, y = 0.965, xend = 0.44, yend = 0.8, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.2915, y = 0.5135, xend = 0.18, yend = 0.32, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.5525, y = 0.5135, xend = 0.67, yend = 0.32, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.422, y = 1.04, xend = 0.42, yend = 0.85, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.161, y = 0.890, xend = -0.02, yend = 0.8, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.161, y = 0.589, xend = 0.161, yend = 0.4, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.422, y = 0.438, xend = 0.25, yend = 0.3, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.683, y = 0.589, xend = 0.683, yend = 0.4, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.683, y = 0.890, xend = 0.5, yend = 0.82, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd")
+  geom_segment(x = 0.313, y = 1.03, xend = 0.41, yend = 0.82,
+               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") #+
+  # geom_segment(x = 0.5525, y = 0.965, xend = 0.44, yend = 0.8, 
+  #              arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  # geom_segment(x = 0.2915, y = 0.5135, xend = 0.18, yend = 0.32, 
+  #              arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  # geom_segment(x = 0.5525, y = 0.5135, xend = 0.67, yend = 0.32, 
+  #              arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  # geom_segment(x = 0.422, y = 1.04, xend = 0.42, yend = 0.85, 
+  #              arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  # geom_segment(x = 0.161, y = 0.890, xend = -0.02, yend = 0.8, 
+  #              arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  # geom_segment(x = 0.161, y = 0.589, xend = 0.161, yend = 0.4, 
+  #              arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  # geom_segment(x = 0.422, y = 0.438, xend = 0.25, yend = 0.3, 
+  #              arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  # geom_segment(x = 0.683, y = 0.589, xend = 0.683, yend = 0.4, 
+  #              arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  # geom_segment(x = 0.683, y = 0.890, xend = 0.5, yend = 0.82, 
+  #              arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd")
 

### 2025-01-17 appendix.qmd
Updated Figure 2

--- 2025-01-17 appendix.qmd
@@
   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID), size = 5, colour = "#bdbdbd") +
-  geom_segment(x = 0.161, y = 0.7395, xend = -0.06, yend = 0.740, 
+  geom_segment(x = 0.175, y = 0.790, xend = -0.06, yend = 0.790, 
                arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.683, y = 0.7395, xend = 0.45, yend = 0.740, 
+  geom_segment(x = 0.726, y = 0.790, xend = 0.50, yend = 0.790,
                arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.2915, y = 0.965, xend = 0.43, yend = 0.79, 
+  geom_segment(x = 0.313, y = 1.03, xend = 0.41, yend = 0.82,
                arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.5525, y = 0.965, xend = 0.44, yend = 0.8, 
+  geom_segment(x = 0.588, y = 1.03, xend = 0.48, yend = 0.82,
                arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.2915, y = 0.5135, xend = 0.18, yend = 0.32, 
+  geom_segment(x = 0.313, y = 0.551, xend = 0.20, yend = 0.34,
                arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.5525, y = 0.5135, xend = 0.67, yend = 0.32, 
+  geom_segment(x = 0.588, y = 0.551, xend = 0.67, yend = 0.34,
                arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.422, y = 1.04, xend = 0.42, yend = 0.85, 
+  geom_segment(x = 0.451, y = 1.11, xend = 0.451, yend = 0.85,
                arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.161, y = 0.890, xend = -0.02, yend = 0.8, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.161, y = 0.589, xend = 0.161, yend = 0.4, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.422, y = 0.438, xend = 0.25, yend = 0.3, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.683, y = 0.589, xend = 0.683, yend = 0.4, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.683, y = 0.890, xend = 0.5, yend = 0.82, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd")
+  geom_segment(x = 0.175, y = 0.949, xend = -0.02, yend = 0.82,
+               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") #+
+  # geom_segment(x = 0.161, y = 0.589, xend = 0.161, yend = 0.4, 
+  #              arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  # geom_segment(x = 0.422, y = 0.438, xend = 0.25, yend = 0.3, 
+  #              arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  # geom_segment(x = 0.683, y = 0.589, xend = 0.683, yend = 0.4, 
+  #              arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  # geom_segment(x = 0.683, y = 0.890, xend = 0.5, yend = 0.82, 
+  #              arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd")
 

### 2025-01-17 appendix.qmd
Updated Figure 2

@@
   geom_segment(x = 0.175, y = 0.949, xend = -0.02, yend = 0.82,
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") #+
-  # geom_segment(x = 0.161, y = 0.589, xend = 0.161, yend = 0.4, 
-  #              arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  # geom_segment(x = 0.422, y = 0.438, xend = 0.25, yend = 0.3, 
-  #              arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  # geom_segment(x = 0.683, y = 0.589, xend = 0.683, yend = 0.4, 
-  #              arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  # geom_segment(x = 0.683, y = 0.890, xend = 0.5, yend = 0.82, 
-  #              arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd")
+               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  geom_segment(x = 0.175, y = 0.631, xend = 0.175, yend = 0.4,
+               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  geom_segment(x = 0.451, y = 0.472, xend = 0.25, yend = 0.3,
+               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  geom_segment(x = 0.726, y = 0.631, xend = 0.726, yend = 0.4,
+               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
+  geom_segment(x = 0.726, y = 0.949, xend = 0.5, yend = 0.82,
+               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd")
 

### 2025-01-17 appendix.qmd
Added scDEED section

@@
 )
+
+interior_annotation <- function(label, position = c(0.92, 0.92), cex = 1, col="grey70") {
+  annotation_custom(grid::textGrob(label = label,
+      x = unit(position[1], "npc"), y = unit(position[2], "npc"),
+      gp = grid::gpar(cex = cex, col=col)))
+}
+```
+
+```{r}
+#| label: scale-highd-data
+# Center the data by subtracting the mean of each column
+center_data <- function(data) {
+  apply(data, 2, function(col) col - mean(col))
+}
+
+# Function to scale data manually
+scale_data_manual <- function(data, type_col) {
+  # Step 1: Center the data (mean 0)
+  data_centered <- center_data(data |> select(-all_of(type_col)))
+
+  # Step 2: Calculate the standard deviation of each dimension
+  sds <- apply(data_centered, 2, sd)
+
+  # Step 3: Scale each dimension to have the range [0, 1]
+  data_scaled <- apply(data_centered, 2, function(col) col / max(abs(col)))
+
+  # Step 4: Scale dimensions according to their variation
+  # The dimension with the highest standard deviation is scaled to [-1, 1]
+  # Other dimensions are scaled to smaller ranges based on their standard deviations
+  max_sd <- max(sds)
+
+  # Normalize the standard deviations to get scaling factors
+  scaling_factors <- sds / max_sd
+
+  for (i in seq_along(scaling_factors)) {
+    data_scaled[, i] <- data_scaled[, i] * scaling_factors[i]
+  }
+
+  # Combine the scaled data with the 'type' column and return as a tibble
+  data_scaled <- as_tibble(data_scaled) %>% 
+    mutate(!!type_col := data[[type_col]])
+
+  return(data_scaled)
+}
+```
+
+```{r}
+#| label: gen-axes
+
+gen_axes <- function(proj, limits = 1, axis_pos_x = NULL, axis_pos_y = NULL, axis_labels, threshold) {
+
+  axis_scale <- limits/6
+
+  if (is.null(axis_pos_x)) {
+
+    axis_pos_x <- -2/3 * limits
+
+  }
+
+  if (is.null(axis_pos_y)) {
+
+    axis_pos_y <- -2/3 * limits
+
+  }
+
+  adj <- function(x, axis_pos) axis_pos + x * axis_scale
+  axes <- data.frame(x1 = adj(0, axis_pos_x), 
+                     y1 = adj(0, axis_pos_y), 
+                     x2 = adj(proj[, 1], axis_pos_x), 
+                     y2 = adj(proj[, 2], axis_pos_y))
+
+  rownames(axes) <- axis_labels
+
+  ## To remove axes
+  axes <- axes |>
+    mutate(distance = sqrt((x2 - x1)^2 + (y2 - y1)^2)) |>
+    filter(distance >= threshold)
+
+  theta <- seq(0, 2 * pi, length = 50)
+  circle <- data.frame(c1 = adj(cos(theta), axis_pos_x), 
+                       c2 = adj(sin(theta), axis_pos_y))
+
+  return(list(axes = axes, circle = circle))
+
+}
+```
+
+```{r}
+#| label: solve-quad-fun
+
+quad <- function(a = 3, b = 2 * a2, c = -(a2^2 + a1^2))
+{
+  a <- as.complex(a)
+  answer <- c((-b + sqrt(b^2 - 4 * a * c)) / (2 * a),
+              (-b - sqrt(b^2 - 4 * a * c)) / (2 * a))
+  if(all(Im(answer) == 0)) answer <- Re(answer)
+  if(answer[1] == answer[2]) return(answer[1])
+  answer[answer>0] ## only positive 
+}
+
 ```
@@
 
-<!-- #### Comparison with results of scDEED recommendations -->
-
-<!-- <!-- -->
+### Single-cell gene expression
+
+#### Comparison with results of scDEED recommendations
+
 <!-- XXX What this section needs: -->
@@
 <!-- CHANGES WERE MADE BASED ON THIS THINKING -->
-<!-- --> -->
 
@@
 However, it is well known that the 2D embeddings produced by t-SNE and UMAP may not accurately reflect the similarities among cell clusters.
-ALREADY STATED IN PREVIOUS SECTION -->  -->
-
-<!-- As we were writing this paper @xia2023 appeared proposing a new method called scDEED helping to assess the validity of a \gD{} embedding.
+ALREADY STATED IN PREVIOUS SECTION -->
+
+As we were writing this paper @xia2023 appeared proposing a new method called scDEED helping to assess the validity of a \gD{} embedding.
 scDEED calculates a reliability score for each cell embedding based on the similarity between the cell’s \gD{} embedding neighbors and its neighbors prior to embedding.
 A low reliability score suggests a dubious embedding.
-It can help in the deciding on optimal hyperparameters.
-Here we illustrate how our method compares with the results from scDEED -->
-
-<!-- Following the process in @xia2023 again using the Human Peripheral Blood Mononuclear Cells (PBMC) data.
+It can help in the deciding on optimal hyper-parameters.
+Here we illustrate how our method compares with the results from scDEED
+
+Following the process in @xia2023 again using the Human Peripheral Blood Mononuclear Cells (PBMC) data.
 Note that @xia2023 uses a different subset from @sec-pbmc, which contains $31,021$ cells including cell type labels, and the gene expression levels were in the unit of log-transformed UMI count per $10,000$.
 They focused on three sequencing methods (inDrops, DropSeq, and SeqWell) and four common cell types Cytotoxic T cell, CD4+T cell, CD14+ Monocyte, and B cell.
--->
-
-<!-- For illustration purposes, we only selected cells generated with inDrops ($n=5858$ cells) and UMAP cell embeddings.
+
+For illustration purposes, we only selected cells generated with inDrops ($n=5858$ cells) and UMAP cell embeddings.
 Also, @xia2023 used first $50$ principal components to generate the UMAP.
 The objective is to assess the optimized layout by scDEED, and if it does not accurately represent the three clusters with small separations of the PBMC dataset, then select a reasonable \gD{} layout.
--->
-
-<!-- @fig-model-pbmc-author-proj-scdeed shows layout from the original published article along with that suggested by scDEED to be more accurate.
+
+@fig-model-pbmc-author-proj-scdeed shows layout from the original published article along with that suggested by scDEED to be more accurate.
 The log MSE vs binwidth plot illustrates that our approach would suggest that scDEED is correct here, that layout b is more accurately reflecting the cluster structure in the data.
--->
-
-<!-- ```{r} -->
-<!-- result_umap <- read_rds(here::here('data/pbmc3k/pbmc_scdeed_umap_results.rds')) -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- #| eval: false -->
-<!-- result_umap |> -->
-<!--   kableExtra::kable(format = "latex", -->
-<!--                     booktabs = TRUE, -->
-<!--                     longtable = TRUE, -->
-<!--                     label = "scdeedumap") |> -->
-<!--   kableExtra::kable_styling(latex_options = "scale_down") -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- #| out-width: 50% -->
-<!-- #| eval: false -->
-
-<!-- result_umap |> -->
-<!--   filter(min.dist == 0.5) |> -->
-<!--   ggplot(aes( -->
-<!--            x = as.numeric(n_neighbors), -->
-<!--            y = number_dubious_cells -->
-<!--          )) + -->
-<!--     geom_line() + -->
-<!--     geom_point() + -->
-<!--     labs(x = "n_neighbors", -->
-<!--          y = "# of dubious cells") + -->
-<!--     theme_minimal() -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- #| label: read-pbmc-umap-layouts -->
-<!-- # Read a variety of different NLDR representations of PBMC -->
-<!-- # and plot them on same aspect ratio -->
-<!-- umap_pbmc1 <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_30_min_dist_0.3.rds") -->
-<!-- umap_pbmc1 <- as_tibble(umap_pbmc1) -->
-<!-- names(umap_pbmc1) <- c("UMAP1", "UMAP2") -->
-
-<!-- nldr1c <- umap_pbmc1 |> -->
-<!--   ggplot(aes(x = UMAP1, -->
-<!--              y = UMAP2)) + -->
-<!--   geom_point(alpha=0.1, size=1, colour='#e41a1c') + -->
-<!--   interior_annotation("a") -->
-
-<!-- umap_pbmc2 <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_80_min_dist_0.5.rds") -->
-<!-- umap_pbmc2 <- as_tibble(umap_pbmc2) -->
-<!-- names(umap_pbmc2) <- c("UMAP1", "UMAP2") -->
-
-<!-- nldr2c <- umap_pbmc2 |> -->
-<!--   ggplot(aes(x = UMAP1, -->
-<!--              y = UMAP2))+ -->
-<!--   geom_point(alpha=0.1, size=1, colour='#377eb8') + -->
-<!--   interior_annotation("b") -->
-
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- #| label: combine-data-pbmc-umap -->
-
-<!-- error_pbmc_umap <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_30_min_dist_0.3.rds") -->
-<!-- error_pbmc_umap2 <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_80_min_dist_0.5.rds") -->
-
-<!-- error_pbmc <- bind_rows(error_pbmc_umap,  -->
-<!--                         error_pbmc_umap2) -->
-
-<!-- error_pbmc <- error_pbmc |> -->
-<!--   mutate(a1 = round(a1, 2)) |> -->
-<!--   filter(bin1 >= 5) |> -->
-<!--   group_by(method, a1) |> -->
-<!--   filter(MSE == min(MSE)) |> -->
-<!--   ungroup() -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- #| label: error-comp-pbmc-umap -->
-
-<!-- error_umap_plot_pbmc <- ggplot(error_pbmc,  -->
-<!--                                aes(x = a1,  -->
-<!--                                    y = MSE,  -->
-<!--                                    colour = method)) +  -->
-<!--   geom_point(size = 0.8) + -->
-<!--   geom_line(linewidth = 0.3) +  -->
-<!--   # geom_vline(xintercept = 15, linetype="solid",  -->
-<!--   #            color = "black", linewidth=0.8, alpha = 0.5) + -->
-<!--   scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 12, 15, 17, 19)]) + -->
-<!--   scale_color_manual(values=c('#e41a1c','#377eb8')) + -->
-<!--   scale_y_log10() + -->
-<!--   ylab("log(MSE)") + -->
-<!--   xlab(expression(paste("binwidth (", a[1], ")"))) + -->
-<!--   theme_minimal() + -->
-<!--   theme(panel.border = element_rect(fill = 'transparent'), -->
-<!--         plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5), -->
-<!--         axis.ticks.x = element_line(), -->
-<!--         axis.ticks.y = element_line(), -->
-<!--         legend.position = "none", -->
-<!--         axis.text.x = element_text(size = 7), -->
-<!--         axis.text.y = element_text(size = 7), -->
-<!--         axis.title.x = element_text(size = 7), -->
-<!--         axis.title.y = element_text(size = 7))  -->
-
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- #| label: fig-pbmc-mse-umap -->
-<!-- #| fig-pos: H -->
-<!-- #| out-width: 100% -->
-<!-- #| fig-cap: "Assessing which of the 2 UMAP layouts with different hyperparameter settings (n_neighbors: $30$, min_dist: $0.3$ (red), n_neighbors: $80$, min_dist: $0.5$ (blue)) on the PBMC data is the better representation using MSE for varying binwidth ($a_1$).
+
+```{r}
+result_umap <- read_rds(here::here('data/pbmc3k/pbmc_scdeed_umap_results.rds'))
+```
+
+```{r}
+#| eval: false
+result_umap |>
+  kableExtra::kable(format = "latex",
+                    booktabs = TRUE,
+                    longtable = TRUE,
+                    label = "scdeedumap") |>
+  kableExtra::kable_styling(latex_options = "scale_down")
+```
+
+```{r}
+#| out-width: 50%
+#| eval: false
+
+result_umap |>
+  filter(min.dist == 0.5) |>
+  ggplot(aes(
+           x = as.numeric(n_neighbors),
+           y = number_dubious_cells
+         )) +
+    geom_line() +
+    geom_point() +
+    labs(x = "n_neighbors",
+         y = "# of dubious cells") +
+    theme_minimal()
+```
+
+```{r}
+#| label: read-pbmc-umap-layouts
+# Read a variety of different NLDR representations of PBMC
+# and plot them on same aspect ratio
+umap_pbmc1 <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_30_min_dist_0.3.rds")
+umap_pbmc1 <- as_tibble(umap_pbmc1)
+names(umap_pbmc1) <- c("UMAP1", "UMAP2")
+
+nldr1c <- umap_pbmc1 |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2)) +
+  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
+  interior_annotation("a")
+
+umap_pbmc2 <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_80_min_dist_0.5.rds")
+umap_pbmc2 <- as_tibble(umap_pbmc2)
+names(umap_pbmc2) <- c("UMAP1", "UMAP2")
+
+nldr2c <- umap_pbmc2 |>
+  ggplot(aes(x = UMAP1,
+             y = UMAP2))+
+  geom_point(alpha=0.1, size=1, colour='#377eb8') +
+  interior_annotation("b")
+
+```
+
+```{r}
+#| label: combine-data-pbmc-umap
+
+error_pbmc_umap <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_30_min_dist_0.3.rds")
+error_pbmc_umap2 <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_80_min_dist_0.5.rds")
+
+error_pbmc <- bind_rows(error_pbmc_umap,
+                        error_pbmc_umap2)
+
+error_pbmc <- error_pbmc |>
+  mutate(a1 = round(a1, 2)) |>
+  filter(bin1 >= 5) |>
+  group_by(method, a1) |>
+  filter(MSE == min(MSE)) |>
+  ungroup()
+```
+
+```{r}
+#| label: error-comp-pbmc-umap
+
+error_umap_plot_pbmc <- ggplot(error_pbmc,
+                               aes(x = a1,
+                                   y = MSE,
+                                   colour = method)) +
+  geom_point(size = 0.8) +
+  geom_line(linewidth = 0.3) +
+  # geom_vline(xintercept = 15, linetype="solid",
+  #            color = "black", linewidth=0.8, alpha = 0.5) +
+  scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 12, 15, 17, 19)]) +
+  scale_color_manual(values=c('#e41a1c','#377eb8')) +
+  scale_y_log10() +
+  ylab("log(MSE)") +
+  xlab(expression(paste("binwidth (", a[1], ")"))) +
+  theme_minimal() +
+  theme(panel.border = element_rect(fill = 'transparent'),
+        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+        axis.ticks.x = element_line(),
+        axis.ticks.y = element_line(),
+        legend.position = "none",
+        axis.text.x = element_text(size = 7),
+        axis.text.y = element_text(size = 7),
+        axis.title.x = element_text(size = 7),
+        axis.title.y = element_text(size = 7))
+
+```
+
+```{r}
+#| label: fig-pbmc-mse-umap
+#| fig-pos: H
+#| out-width: 100%
+#| fig-cap: "Assessing which of the 2 UMAP layouts with different hyperparameter settings (n_neighbors: $30$, min_dist: $0.3$ (red), n_neighbors: $80$, min_dist: $0.5$ (blue)) on the PBMC data is the better representation using MSE for varying binwidth ($a_1$).
 Colour used for the lines and points in the left plot and in the scatterplots represents UMAP layout (a, b).
-Of the two, layout b is optimal across all binwidths making it the best choice." -->
-
-<!-- free(error_umap_plot_pbmc) + wrap_plots(nldr1c, nldr2c, ncol = 2) + -->
-<!--   plot_layout(ncol = 1, heights = c(1, 1)) -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- #| label: published-pbmc-scdeed -->
-
-<!-- training_data_pbmc <- read_rds("data/pbmc3k/pbmc_pca_50_scdeed.rds") -->
-<!-- training_data_pbmc <- training_data_pbmc[, 1:50] |> -->
-<!--   mutate(ID = 1:NROW(training_data_pbmc)) -->
-
-<!-- ## For umap -->
-<!-- umap_pbmc <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_80_min_dist_0.5.rds") -->
-<!-- umap_pbmc <- as_tibble(umap_pbmc) -->
-<!-- names(umap_pbmc) <- c("UMAP1", "UMAP2") -->
-<!-- umap_pbmc <- umap_pbmc |> -->
-<!--   mutate(ID = 1:NROW(umap_pbmc)) -->
-
-<!-- pbmc_scaled_obj_umap <- gen_scaled_data( -->
-<!--   data = umap_pbmc) -->
-<!-- umap_pbmc_scaled <- pbmc_scaled_obj_umap$scaled_nldr -->
-
-<!-- lim1 <- pbmc_scaled_obj_umap$lim1 -->
-<!-- lim2 <- pbmc_scaled_obj_umap$lim2 -->
-<!-- r2_umap <- diff(lim2)/diff(lim1) -->
-
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- #| label: hexbin-pbmc-scdeed -->
-<!-- ## Compute hexbin parameters -->
-<!-- num_bins_x_pbmc <- 12 -->
-<!-- lim1 <- pbmc_scaled_obj_umap$lim1 -->
-<!-- lim2 <- pbmc_scaled_obj_umap$lim2 -->
-<!-- r2_pbmc <- diff(lim2)/diff(lim1)  -->
-
-<!-- pbmc_model <- fit_highd_model( -->
-<!--   training_data = training_data_pbmc, -->
-<!--   emb_df = umap_pbmc_scaled, -->
-<!--   bin1 = num_bins_x_pbmc, -->
-<!--   r2 = r2_pbmc, -->
-<!--   q = 0.1, -->
-<!--   is_bin_centroid = TRUE, -->
-<!--   is_rm_lwd_hex = FALSE, -->
-<!--   col_start_highd = "pc" -->
-<!-- ) -->
-
-<!-- df_bin_centroids_pbmc <- pbmc_model$df_bin_centroids -->
-<!-- df_bin_pbmc <- pbmc_model$df_bin -->
-
-<!-- ## Triangulate bin centroids -->
-<!-- tr1_object_pbmc <- tri_bin_centroids( -->
-<!--   df_bin_centroids_pbmc, x = "c_x", y = "c_y") -->
-<!-- tr_from_to_df_pbmc <- gen_edges( -->
-<!--   tri_object = tr1_object_pbmc) -->
-
-<!-- ## Compute 2D distances -->
-<!-- distance_pbmc <- cal_2d_dist( -->
-<!--   tr_coord_df = tr_from_to_df_pbmc, -->
-<!--   start_x = "x_from", -->
-<!--   start_y = "y_from", -->
-<!--   end_x = "x_to", -->
-<!--   end_y = "y_to", -->
-<!--   select_vars = c("from", "to", "distance")) -->
-
-<!-- ## To find the benchmark value -->
-<!-- benchmark_pbmc <- find_lg_benchmark( -->
-<!--   distance_edges = distance_pbmc, -->
-<!--   distance_col = "distance") -->
-
-<!-- tr_df <- distinct(tibble( -->
-<!--   x = c(tr_from_to_df_pbmc[["x_from"]], tr_from_to_df_pbmc[["x_to"]]),  -->
-<!--   y = c(tr_from_to_df_pbmc[["y_from"]], tr_from_to_df_pbmc[["y_to"]]))) -->
-
-<!-- distance_df_small_edges_pbmc <- distance_pbmc |> -->
-<!--   filter(distance < benchmark_pbmc) -->
-
-<!-- tr_from_to_df_pbmc <- inner_join( -->
-<!--   tr_from_to_df_pbmc, distance_df_small_edges_pbmc,  -->
-<!--   by = c("from", "to")) -->
-
-<!-- trimesh_removed_pbmc <- ggplot() +  -->
-<!--   geom_segment(data = tr_from_to_df_pbmc,  -->
-<!--                aes( -->
-<!--                  x = x_from,  -->
-<!--                  y = y_from,  -->
-<!--                  xend = x_to,  -->
-<!--                  yend = y_to), -->
-<!--                colour = "#000000", -->
-<!--                linewidth = 1) + -->
-<!--   geom_point( -->
-<!--     data = umap_pbmc_scaled, -->
-<!--     aes( -->
-<!--       x = UMAP1, -->
-<!--       y = UMAP2 -->
-<!--     ), -->
-<!--     alpha = 0.2, -->
-<!--     color = '#a65628' -->
-<!--   )  + -->
-<!--   interior_annotation("a1", cex = 2) + -->
-<!--   theme( -->
-<!--     aspect.ratio = 1 -->
-<!--   ) -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- #| label: pbmc-umap-model-proj-scdeed -->
-
-<!-- data_pbmc <- training_data_pbmc |>  -->
-<!--   select(-ID) |> -->
-<!--   mutate(type = "data") -->
-
-<!-- df_b_pbmc <- df_bin_pbmc |> -->
-<!--   dplyr::filter(hb_id %in% df_bin_centroids_pbmc$hexID) |> -->
-<!--   dplyr::mutate(type = "model") ## Data with summarized mean -->
-
-<!-- ## Reorder the rows of df_b according to the hexID order in df_b_with_center_data -->
-<!-- df_b_pbmc <- df_b_pbmc[match(df_bin_centroids_pbmc$hexID, df_b_pbmc$hb_id),] |> -->
-<!--   dplyr::select(-hb_id)  -->
-
-<!-- # Apply the scaling -->
-<!-- df_model_data_pbmc <- bind_rows(data_pbmc, df_b_pbmc) -->
-<!-- scaled_pbmc <- scale_data_manual(df_model_data_pbmc, "type") |> -->
-<!--   as_tibble() -->
-
-<!-- scaled_pbmc_data <- scaled_pbmc |> -->
-<!--   filter(type == "data") |> -->
-<!--   select(-type) -->
-
-<!-- scaled_pbmc_data_model <- scaled_pbmc |> -->
-<!--   filter(type == "model") |> -->
-<!--   select(-type) -->
-
-<!-- # Combine with the true model for visualization -->
-<!-- df <- dplyr::bind_rows(scaled_pbmc_data_model |> mutate(type = "model"), -->
-<!--                        scaled_pbmc_data |> mutate(type = "data")) -->
-
-<!-- ## First projection -->
-<!-- projection <- cbind( -->
-<!--   c(0.040856,0.000723,0.006000,0.007623,0.001986,-0.006599,0.005311,-0.000222,0.001648,0.008067,-0.001587,0.004423,0.004851,0.000638,0.002606,0.006616,0.006188,-0.000954,-0.004603,0.004998,-0.012220,0.008401,-0.007788,0.008972,-0.003903,0.003348,-0.001817,-0.000385,-0.002437,-0.004097,-0.001357,0.013660,0.001770,0.006982,0.002828,-0.009169,-0.003541,0.008582,-0.000032,-0.002375,0.005854,-0.005466,-0.000824,-0.001435,0.004148,0.007141,0.000883,-0.004159,0.005401,0.007882), -->
-<!--   c(-0.000590,0.039365,-0.004481,-0.009041,-0.003805,0.003278,-0.009601,-0.000650,-0.002333,-0.005697,-0.001488,-0.009808,0.002557,0.004113,0.005705,-0.004665,-0.001170,-0.004727,-0.000845,0.005084,-0.018054,0.002538,0.006606,-0.007184,0.000238,0.001877,-0.002124,-0.005972,-0.009729,0.004958,-0.003181,-0.005185,-0.009598,-0.004243,0.007716,0.000396,-0.004258,0.005932,-0.003242,-0.004184,0.003493,-0.008405,0.000860,0.008304,0.004569,0.000798,0.004156,-0.005777,0.005278,-0.002531)) -->
-
-<!-- projection_scaled <- projection * 3 -->
-
-<!-- # projection_filtered <- projection -->
-<!-- # projection_filtered[apply(projection_filtered, 1, function(row) all(row < 0)), ] <- 0.75 -->
-
-<!-- #rows_to_keep <- c(1, 2, 4, 5, 6) -->
-
-<!-- #projection_filtered <- projection[rows_to_keep, ] -->
-
-<!-- projected <- as.matrix(scaled_pbmc_data) %*% projection_scaled -->
-
-<!-- projected_df <- projected |> -->
-<!--   tibble::as_tibble(.name_repair = "unique") |> -->
-<!--   dplyr::rename(c("proj1" = "...1", -->
-<!--                   "proj2" = "...2")) |> -->
-<!--   #dplyr::mutate(type = df_exe$type) |> -->
-<!--   dplyr::mutate(ID = dplyr::row_number()) -->
-
-<!-- projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection_scaled -->
-
-<!-- projected_model_df <- projected_model |> -->
-<!--   tibble::as_tibble(.name_repair = "unique") |> -->
-<!--   dplyr::rename(c("proj1" = "...1", -->
-<!--                   "proj2" = "...2")) |> -->
-<!--   dplyr::mutate(ID = dplyr::row_number()) -->
-
-<!-- model_df <- dplyr::left_join( -->
-<!--   distance_df_small_edges_pbmc |> select(-distance),  -->
-<!--   projected_model_df,  -->
-<!--   by = c("from" = "ID")) -->
-
-<!-- names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from") -->
-
-<!-- model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID")) -->
-<!-- names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to") -->
-
-<!-- axes_obj <- gen_axes( -->
-<!--   proj = projection, -->
-<!--   limits = 0.04, -->
-<!--   axis_pos_x = -0.025, -->
-<!--   axis_pos_y = -0.025, -->
-<!--   axis_labels = names(scaled_pbmc_data), -->
-<!--   threshold = 0.0001) -->
-
-<!-- axes <- axes_obj$axes -->
-<!-- circle <- axes_obj$circle -->
-
-<!-- pbmc_proj_umap_model1 <- projected_df |> -->
-<!--   ggplot( -->
-<!--     aes( -->
-<!--       x = proj1,  -->
-<!--       y = proj2)) + -->
-<!--   geom_segment( -->
-<!--     data = model_df,  -->
-<!--     aes( -->
-<!--       x = proj1_from,  -->
-<!--       y = proj2_from,  -->
-<!--       xend = proj1_to,  -->
-<!--       yend = proj2_to),  -->
-<!--     color = "#000000") + -->
-<!--   geom_point( -->
-<!--     #size = 0.5, -->
-<!--     alpha = 0.2, -->
-<!--     color = '#a65628') + -->
-<!--   geom_segment( -->
-<!--     data=axes,  -->
-<!--     aes(x=x1, y=y1, xend=x2, yend=y2),  -->
-<!--     colour="grey70") + -->
-<!--   geom_text( -->
-<!--     data=axes, -->
-<!--     aes(x=x2, y=y2),  -->
-<!--     label=rownames(axes), -->
-<!--     colour="grey50", -->
-<!--     size = 3) + -->
-<!--   geom_path( -->
-<!--     data=circle,  -->
-<!--     aes(x=c1, y=c2), colour="grey70") + -->
-<!--   coord_fixed() + -->
-<!--   xlim(c(-0.03, 0.06)) + -->
-<!--   ylim(c(-0.03, 0.06)) + -->
-<!--   interior_annotation("a2", cex = 2) -->
-
-<!-- ## Second projection -->
-<!-- projection <- cbind( -->
-<!--   c(-0.019825,-0.011101,0.011543,-0.006943,-0.004485,0.000062,0.002023,-0.005178,0.000800,-0.002903,-0.014182,0.008756,0.009595,0.007032,-0.007820,0.006143,0.004865,0.003352,0.007554,0.002265,0.012035,0.009961,-0.001825,-0.004942,-0.002237,0.002308,0.009531,-0.001146,0.013401,-0.001022,0.008434,0.000477,0.004239,-0.006772,-0.004702,-0.011432,-0.000431,0.008146,-0.009299,-0.004782,0.003457,-0.008214,-0.009604,0.011329,0.003805,-0.003004,-0.000800,0.015288,-0.013157,0.006984), -->
-<!--   c(-0.001387,0.001643,0.010470,0.002930,-0.002335,-0.010569,0.013839,0.002917,-0.006825,0.010191,-0.006367,0.002220,0.002506,0.005139,-0.007213,-0.006235,-0.007123,-0.000222,-0.002225,0.011592,-0.003942,0.002788,-0.000055,0.011316,-0.009753,0.020785,0.004615,0.005172,0.011182,-0.003125,-0.010860,0.007820,-0.017510,0.007010,0.007267,0.002694,-0.003949,-0.009188,-0.005163,0.001898,-0.004744,0.013757,0.010630,-0.007688,0.004598,0.004658,-0.011154,0.003584,-0.005226,0.006650)) -->
-
-<!-- projection_scaled <- projection * 3 -->
-
-<!-- projected <- as.matrix(scaled_pbmc_data) %*% projection_scaled -->
-
-<!-- projected_df <- projected |> -->
-<!--   tibble::as_tibble(.name_repair = "unique") |> -->
-<!--   dplyr::rename(c("proj1" = "...1", -->
-<!--                   "proj2" = "...2")) |> -->
-<!--   #dplyr::mutate(type = df_exe$type) |> -->
-<!--   dplyr::mutate(ID = dplyr::row_number()) -->
-
-<!-- projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection_scaled -->
-
-<!-- projected_model_df <- projected_model |> -->
-<!--   tibble::as_tibble(.name_repair = "unique") |> -->
-<!--   dplyr::rename(c("proj1" = "...1", -->
-<!--                   "proj2" = "...2")) |> -->
-<!--   dplyr::mutate(ID = dplyr::row_number()) -->
-
-<!-- model_df <- dplyr::left_join( -->
-<!--   distance_df_small_edges_pbmc |> select(-distance),  -->
-<!--   projected_model_df,  -->
-<!--   by = c("from" = "ID")) -->
-
-<!-- names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from") -->
-
-<!-- model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID")) -->
-<!-- names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to") -->
-
-<!-- axes_obj <- gen_axes( -->
-<!--   proj = projection, -->
-<!--   limits = 0.03, -->
-<!--   axis_pos_x = -0.02, -->
-<!--   axis_pos_y = -0.02, -->
-<!--   axis_labels = names(scaled_pbmc_data), -->
-<!--   threshold = 0) -->
-
-<!-- axes <- axes_obj$axes -->
-<!-- circle <- axes_obj$circle -->
-
-<!-- pbmc_proj_umap_model2 <- projected_df |> -->
-<!--   ggplot( -->
-<!--     aes( -->
-<!--       x = proj1, -->
-<!--       y = proj2)) + -->
-<!--   geom_segment( -->
-<!--     data = model_df, -->
-<!--     aes( -->
-<!--       x = proj1_from, -->
-<!--       y = proj2_from, -->
-<!--       xend = proj1_to, -->
-<!--       yend = proj2_to), -->
-<!--     color = "#000000") + -->
-<!--   geom_point( -->
-<!--     #size = 0.5, -->
-<!--     alpha = 0.2, -->
-<!--     color = '#a65628') + -->
-<!--   geom_segment( -->
-<!--     data=axes, -->
-<!--     aes(x=x1, y=y1, xend=x2, yend=y2), -->
-<!--     colour="grey70") + -->
-<!--   geom_text( -->
-<!--     data=axes, -->
-<!--     aes(x=x2, y=y2), -->
-<!--     label=rownames(axes), -->
-<!--     colour="grey50", -->
-<!--     size = 3) + -->
-<!--   geom_path( -->
-<!--     data=circle, -->
-<!--     aes(x=c1, y=c2), colour="grey70") + -->
-<!--   coord_fixed() + -->
-<!--   xlim(c(-0.025, 0.02)) + -->
-<!--   ylim(c(-0.025, 0.02)) + -->
-<!--   interior_annotation("a3", cex = 2) -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- #| echo: false -->
-<!-- #| eval: false -->
-<!-- #| fig-cap: "Model in \\gD{}, on the layout a, and two views of the fit in projections from $50\\text{-}D$, for the PBMC data ($n =  5858$ and $p = 50$).
+Of the two, layout b is optimal across all binwidths making it the best choice."
+
+free(error_umap_plot_pbmc) + wrap_plots(nldr1c, nldr2c, ncol = 2) +
+  plot_layout(ncol = 1, heights = c(1, 1))
+```
+
+```{r}
+#| label: published-pbmc-scdeed
+
+training_data_pbmc <- read_rds("data/pbmc3k/pbmc_pca_50_scdeed.rds")
+training_data_pbmc <- training_data_pbmc[, 1:50] |>
+  mutate(ID = 1:NROW(training_data_pbmc))
+
+## For umap
+umap_pbmc <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_80_min_dist_0.5.rds")
+umap_pbmc <- as_tibble(umap_pbmc)
+names(umap_pbmc) <- c("UMAP1", "UMAP2")
+umap_pbmc <- umap_pbmc |>
+  mutate(ID = 1:NROW(umap_pbmc))
+
+pbmc_scaled_obj_umap <- gen_scaled_data(
+  data = umap_pbmc)
+umap_pbmc_scaled <- pbmc_scaled_obj_umap$scaled_nldr
+
+lim1 <- pbmc_scaled_obj_umap$lim1
+lim2 <- pbmc_scaled_obj_umap$lim2
+r2_umap <- diff(lim2)/diff(lim1)
+
+```
+
+```{r}
+#| label: hexbin-pbmc-scdeed
+## Compute hexbin parameters
+num_bins_x_pbmc <- 12
+lim1 <- pbmc_scaled_obj_umap$lim1
+lim2 <- pbmc_scaled_obj_umap$lim2
+r2_pbmc <- diff(lim2)/diff(lim1)
+
+pbmc_model <- fit_highd_model(
+  training_data = training_data_pbmc,
+  emb_df = umap_pbmc_scaled,
+  bin1 = num_bins_x_pbmc,
+  r2 = r2_pbmc,
+  q = 0.1,
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = FALSE,
+  col_start_highd = "pc"
+)
+
+df_bin_centroids_pbmc <- pbmc_model$df_bin_centroids
+df_bin_pbmc <- pbmc_model$df_bin
+
+## Triangulate bin centroids
+tr1_object_pbmc <- tri_bin_centroids(
+  df_bin_centroids_pbmc, x = "c_x", y = "c_y")
+tr_from_to_df_pbmc <- gen_edges(
+  tri_object = tr1_object_pbmc)
+
+## Compute 2D distances
+distance_pbmc <- cal_2d_dist(
+  tr_coord_df = tr_from_to_df_pbmc,
+  start_x = "x_from",
+  start_y = "y_from",
+  end_x = "x_to",
+  end_y = "y_to",
+  select_vars = c("from", "to", "distance"))
+
+## To find the benchmark value
+benchmark_pbmc <- find_lg_benchmark(
+  distance_edges = distance_pbmc,
+  distance_col = "distance")
+
+tr_df <- distinct(tibble(
+  x = c(tr_from_to_df_pbmc[["x_from"]], tr_from_to_df_pbmc[["x_to"]]),
+  y = c(tr_from_to_df_pbmc[["y_from"]], tr_from_to_df_pbmc[["y_to"]])))
+
+distance_df_small_edges_pbmc <- distance_pbmc |>
+  filter(distance < benchmark_pbmc)
+
+tr_from_to_df_pbmc <- inner_join(
+  tr_from_to_df_pbmc, distance_df_small_edges_pbmc,
+  by = c("from", "to"))
+
+trimesh_removed_pbmc <- ggplot() +
+  geom_segment(data = tr_from_to_df_pbmc,
+               aes(
+                 x = x_from,
+                 y = y_from,
+                 xend = x_to,
+                 yend = y_to),
+               colour = "#000000",
+               linewidth = 1) +
+  geom_point(
+    data = umap_pbmc_scaled,
+    aes(
+      x = UMAP1,
+      y = UMAP2
+    ),
+    alpha = 0.2,
+    color = '#a65628'
+  )  +
+  interior_annotation("a1", cex = 2) +
+  theme(
+    aspect.ratio = 1
+  )
+```
+
+```{r}
+#| label: pbmc-umap-model-proj-scdeed
+
+data_pbmc <- training_data_pbmc |>
+  select(-ID) |>
+  mutate(type = "data")
+
+df_b_pbmc <- df_bin_pbmc |>
+  dplyr::filter(hb_id %in% df_bin_centroids_pbmc$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b_pbmc <- df_b_pbmc[match(df_bin_centroids_pbmc$hexID, df_b_pbmc$hb_id),] |>
+  dplyr::select(-hb_id)
+
+# Apply the scaling
+df_model_data_pbmc <- bind_rows(data_pbmc, df_b_pbmc)
+scaled_pbmc <- scale_data_manual(df_model_data_pbmc, "type") |>
+  as_tibble()
+
+scaled_pbmc_data <- scaled_pbmc |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_pbmc_data_model <- scaled_pbmc |>
+  filter(type == "model") |>
+  select(-type)
+
+# Combine with the true model for visualization
+df <- dplyr::bind_rows(scaled_pbmc_data_model |> mutate(type = "model"),
+                       scaled_pbmc_data |> mutate(type = "data"))
+
+## First projection
+projection <- cbind(
+  c(0.040856,0.000723,0.006000,0.007623,0.001986,-0.006599,0.005311,-0.000222,0.001648,0.008067,-0.001587,0.004423,0.004851,0.000638,0.002606,0.006616,0.006188,-0.000954,-0.004603,0.004998,-0.012220,0.008401,-0.007788,0.008972,-0.003903,0.003348,-0.001817,-0.000385,-0.002437,-0.004097,-0.001357,0.013660,0.001770,0.006982,0.002828,-0.009169,-0.003541,0.008582,-0.000032,-0.002375,0.005854,-0.005466,-0.000824,-0.001435,0.004148,0.007141,0.000883,-0.004159,0.005401,0.007882),
+  c(-0.000590,0.039365,-0.004481,-0.009041,-0.003805,0.003278,-0.009601,-0.000650,-0.002333,-0.005697,-0.001488,-0.009808,0.002557,0.004113,0.005705,-0.004665,-0.001170,-0.004727,-0.000845,0.005084,-0.018054,0.002538,0.006606,-0.007184,0.000238,0.001877,-0.002124,-0.005972,-0.009729,0.004958,-0.003181,-0.005185,-0.009598,-0.004243,0.007716,0.000396,-0.004258,0.005932,-0.003242,-0.004184,0.003493,-0.008405,0.000860,0.008304,0.004569,0.000798,0.004156,-0.005777,0.005278,-0.002531))
+
+projection_scaled <- projection * 3
+
+# projection_filtered <- projection
+# projection_filtered[apply(projection_filtered, 1, function(row) all(row < 0)), ] <- 0.75
+
+#rows_to_keep <- c(1, 2, 4, 5, 6)
+
+#projection_filtered <- projection[rows_to_keep, ]
+
+projected <- as.matrix(scaled_pbmc_data) %*% projection_scaled
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  #dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection_scaled
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges_pbmc |> select(-distance),
+  projected_model_df,
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+axes_obj <- gen_axes(
+  proj = projection,
+  limits = 0.04,
+  axis_pos_x = -0.025,
+  axis_pos_y = -0.025,
+  axis_labels = names(scaled_pbmc_data),
+  threshold = 0.0001)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
+
+pbmc_proj_umap_model1 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1,
+      y = proj2)) +
+  geom_segment(
+    data = model_df,
+    aes(
+      x = proj1_from,
+      y = proj2_from,
+      xend = proj1_to,
+      yend = proj2_to),
+    color = "#000000") +
+  geom_point(
+    #size = 0.5,
+    alpha = 0.2,
+    color = '#a65628') +
+  geom_segment(
+    data=axes,
+    aes(x=x1, y=y1, xend=x2, yend=y2),
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2),
+    label=rownames(axes),
+    colour="grey50",
+    size = 3) +
+  geom_path(
+    data=circle,
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.03, 0.06)) +
+  ylim(c(-0.03, 0.06)) +
+  interior_annotation("a2", cex = 2)
+
+## Second projection
+projection <- cbind(
+  c(-0.019825,-0.011101,0.011543,-0.006943,-0.004485,0.000062,0.002023,-0.005178,0.000800,-0.002903,-0.014182,0.008756,0.009595,0.007032,-0.007820,0.006143,0.004865,0.003352,0.007554,0.002265,0.012035,0.009961,-0.001825,-0.004942,-0.002237,0.002308,0.009531,-0.001146,0.013401,-0.001022,0.008434,0.000477,0.004239,-0.006772,-0.004702,-0.011432,-0.000431,0.008146,-0.009299,-0.004782,0.003457,-0.008214,-0.009604,0.011329,0.003805,-0.003004,-0.000800,0.015288,-0.013157,0.006984),
+  c(-0.001387,0.001643,0.010470,0.002930,-0.002335,-0.010569,0.013839,0.002917,-0.006825,0.010191,-0.006367,0.002220,0.002506,0.005139,-0.007213,-0.006235,-0.007123,-0.000222,-0.002225,0.011592,-0.003942,0.002788,-0.000055,0.011316,-0.009753,0.020785,0.004615,0.005172,0.011182,-0.003125,-0.010860,0.007820,-0.017510,0.007010,0.007267,0.002694,-0.003949,-0.009188,-0.005163,0.001898,-0.004744,0.013757,0.010630,-0.007688,0.004598,0.004658,-0.011154,0.003584,-0.005226,0.006650))
+
+projection_scaled <- projection * 3
+
+projected <- as.matrix(scaled_pbmc_data) %*% projection_scaled
+
+projected_df <- projected |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  #dplyr::mutate(type = df_exe$type) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection_scaled
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+model_df <- dplyr::left_join(
+  distance_df_small_edges_pbmc |> select(-distance),
+  projected_model_df,
+  by = c("from" = "ID"))
+
+names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+axes_obj <- gen_axes(
+  proj = projection,
+  limits = 0.03,
+  axis_pos_x = -0.02,
+  axis_pos_y = -0.02,
+  axis_labels = names(scaled_pbmc_data),
+  threshold = 0)
+
+axes <- axes_obj$axes
+circle <- axes_obj$circle
+
+pbmc_proj_umap_model2 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1,
+      y = proj2)) +
+  geom_segment(
+    data = model_df,
+    aes(
+      x = proj1_from,
+      y = proj2_from,
+      xend = proj1_to,
+      yend = proj2_to),
+    color = "#000000") +
+  geom_point(
+    #size = 0.5,
+    alpha = 0.2,
+    color = '#a65628') +
+  geom_segment(
+    data=axes,
+    aes(x=x1, y=y1, xend=x2, yend=y2),
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2),
+    label=rownames(axes),
+    colour="grey50",
+    size = 3) +
+  geom_path(
+    data=circle,
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.025, 0.02)) +
+  ylim(c(-0.025, 0.02)) +
+  interior_annotation("a3", cex = 2)
+```
+
+```{r}
+#| echo: false
+#| eval: false
+#| fig-cap: "Model in \\gD{}, on the layout a, and two views of the fit in projections from $50\\text{-}D$, for the PBMC data ($n =  5858$ and $p = 50$).
 Layout shows close clusters.
-In $50\\text{-}D$, the data shows " -->
-<!-- #| label: fig-model-pbmc-author-proj-scdeed -->
-<!-- #| fig-pos: H -->
-<!-- #| fig-width: 15 -->
-<!-- #| fig-height: 5 -->
-<!-- #| out-width: 80% -->
-
-<!-- free(trimesh_removed_pbmc) +  -->
-<!--   pbmc_proj_umap_model1 +  -->
-<!--   pbmc_proj_umap_model2 +  -->
-<!--   plot_layout(ncol=3) & -->
-<!--   theme(legend.position='none') -->
-<!-- ``` -->
+In $50\\text{-}D$, the data shows "
+#| label: fig-model-pbmc-author-proj-scdeed
+#| fig-pos: H
+#| fig-width: 15
+#| fig-height: 5
+#| out-width: 80%
+
+free(trimesh_removed_pbmc) +
+  pbmc_proj_umap_model1 +
+  pbmc_proj_umap_model2 +
+  plot_layout(ncol=3) &
+  theme(legend.position='none')
+```

### 2025-01-17 appendix.tex
Updated the appendix
Update Figure 2 of appendix section 2
Added scDEED section

@@
 \newcommand\gD{$2\text{-}D$}
+\newcommand{\To}{\textbf{to}~}
+\usepackage{abstract}
+\renewcommand{\abstractname}{}    % clear the title
+\renewcommand{\absnamepos}{empty} % originally center
 \makeatletter
@@
 
+\subsection{Single-cell gene
+expression}\label{single-cell-gene-expression}
+
+\subsubsection{Comparison with results of scDEED
+recommendations}\label{comparison-with-results-of-scdeed-recommendations}
+
+As we were writing this paper \citet{xia2023} appeared proposing a new
+method called scDEED helping to assess the validity of a \gD{}
+embedding. scDEED calculates a reliability score for each cell embedding
+based on the similarity between the cell's \gD{} embedding neighbors and
+its neighbors prior to embedding. A low reliability score suggests a
+dubious embedding. It can help in the deciding on optimal
+hyper-parameters. Here we illustrate how our method compares with the
+results from scDEED
+
+Following the process in \citet{xia2023} again using the Human
+Peripheral Blood Mononuclear Cells (PBMC) data. Note that
+\citet{xia2023} uses a different subset from \textbf{?@sec-pbmc}, which
+contains \(31,021\) cells including cell type labels, and the gene
+expression levels were in the unit of log-transformed UMI count per
+\(10,000\). They focused on three sequencing methods (inDrops, DropSeq,
+and SeqWell) and four common cell types Cytotoxic T cell, CD4+T cell,
+CD14+ Monocyte, and B cell.
+
+For illustration purposes, we only selected cells generated with inDrops
+(\(n=5858\) cells) and UMAP cell embeddings. Also, \citet{xia2023} used
+first \(50\) principal components to generate the UMAP. The objective is
+to assess the optimized layout by scDEED, and if it does not accurately
+represent the three clusters with small separations of the PBMC dataset,
+then select a reasonable \gD{} layout.
+
+\textbf{?@fig-model-pbmc-author-proj-scdeed} shows layout from the
+original published article along with that suggested by scDEED to be
+more accurate. The log MSE vs binwidth plot illustrates that our
+approach would suggest that scDEED is correct here, that layout b is
+more accurately reflecting the cluster structure in the data.
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{appendix_files/figure-pdf/fig-pbmc-mse-umap-1.pdf}
+
+}
+
+\caption{\label{fig-pbmc-mse-umap}Assessing which of the 2 UMAP layouts
+with different hyperparameter settings (n\_neighbors: \(30\), min\_dist:
+\(0.3\) (red), n\_neighbors: \(80\), min\_dist: \(0.5\) (blue)) on the
+PBMC data is the better representation using MSE for varying binwidth
+(\(a_1\)). Colour used for the lines and points in the left plot and in
+the scatterplots represents UMAP layout (a, b). Of the two, layout b is
+optimal across all binwidths making it the best choice.}
+
+\end{figure}%
+
 \end{document}

### 2025-01-17 appendix.qmd
Updated Figure 4

@@
 tbl-cap-location: bottom
+bibliography: bibliography.bib 
 header-includes: | 
@@
 #| fig-pos: H
-#| out-width: 100%
+#| out-height: 50%
 #| fig-cap: "Assessing which of the 2 UMAP layouts with different hyperparameter settings (n_neighbors: $30$, min_dist: $0.3$ (red), n_neighbors: $80$, min_dist: $0.5$ (blue)) on the PBMC data is the better representation using MSE for varying binwidth ($a_1$).
@@
 #| echo: false
-#| eval: false
 #| fig-cap: "Model in \\gD{}, on the layout a, and two views of the fit in projections from $50\\text{-}D$, for the PBMC data ($n =  5858$ and $p = 50$).
@@
 #| fig-height: 5
-#| out-width: 80%
+#| out-width: 30%
 

### 2025-01-17 appendix.tex
Updated Figure 4

@@
 
-\textbf{?@fig-model-pbmc-author-proj-scdeed} shows layout from the
+Figure~\ref{fig-model-pbmc-author-proj-scdeed} shows layout from the
 original published article along with that suggested by scDEED to be
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{appendix_files/figure-pdf/fig-pbmc-mse-umap-1.pdf}
+\includegraphics[width=1\textwidth,height=0.5\textheight]{appendix_files/figure-pdf/fig-pbmc-mse-umap-1.pdf}
 
@@
 
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=0.3\textwidth,height=\textheight]{appendix_files/figure-pdf/fig-model-pbmc-author-proj-scdeed-1.pdf}
+
+}
+
+\caption{\label{fig-model-pbmc-author-proj-scdeed}Model in \gD{}, on the
+layout a, and two views of the fit in projections from \(50\text{-}D\),
+for the PBMC data (\(n =  5858\) and \(p = 50\)). Layout shows close
+clusters. In \(50\text{-}D\), the data shows}
+
+\end{figure}%
+
+  \bibliography{bibliography.bib}
+
 \end{document}

### 2025-01-17 appendix.tex
Updated PBMC3k data section
Updated PBMC3k data

--- 2025-01-17 appendix.tex
@@
 hyper-parameters. Here we illustrate how our method compares with the
-results from scDEED
+results from scDEED.
 
@@
 
+  \bibliography{bibliography.bib}
+
 \end{document}

### 2025-01-17 appendix.qmd
Updated PBMC3k data
Updated PBMC3k data figure link

@@
 It can help in the deciding on optimal hyper-parameters.
-Here we illustrate how our method compares with the results from scDEED
+Here we illustrate how our method compares with the results from scDEED.
 
@@
 
-@fig-model-pbmc-author-proj-scdeed shows layout from the original published article along with that suggested by scDEED to be more accurate.
+@fig-pbmc-mse-umap shows layout from the original published article along with that suggested by scDEED to be more accurate.
 The log MSE vs binwidth plot illustrates that our approach would suggest that scDEED is correct here, that layout b is more accurately reflecting the cluster structure in the data.

### 2025-01-17 appendix.qmd
Updated the PBMC3k section

@@
 Following the process in @xia2023 again using the Human Peripheral Blood Mononuclear Cells (PBMC) data.
-Note that @xia2023 uses a different subset from @sec-pbmc, which contains $31,021$ cells including cell type labels, and the gene expression levels were in the unit of log-transformed UMI count per $10,000$.
+Note that @xia2023 uses a different subset from the PBMC dataset used by @chen2023, which contains $31,021$ cells including cell type labels, and the gene expression levels were in the unit of log-transformed UMI count per $10,000$.
 They focused on three sequencing methods (inDrops, DropSeq, and SeqWell) and four common cell types Cytotoxic T cell, CD4+T cell, CD14+ Monocyte, and B cell.
@@
 #| fig-pos: H
-#| out-height: 50%
-#| fig-cap: "Assessing which of the 2 UMAP layouts with different hyperparameter settings (n_neighbors: $30$, min_dist: $0.3$ (red), n_neighbors: $80$, min_dist: $0.5$ (blue)) on the PBMC data is the better representation using MSE for varying binwidth ($a_1$).
-Colour used for the lines and points in the left plot and in the scatterplots represents UMAP layout (a, b).
+#| out-height: 53%
+#| fig-cap: "Assessing which of the 2 UMAP layouts with different hyper-parameter settings (n_neighbors: $30$, min_dist: $0.3$ (red); n_neighbors: $80$, min_dist: $0.5$ (blue)) on the PBMC data is the better representation using MSE for varying binwidth ($a_1$).
+Colour used for the lines and points in the top plot and in the scatterplots represents UMAP layout (a, b).
 Of the two, layout b is optimal across all binwidths making it the best choice."

### 2025-01-17 appendix.tex
Updated PBMC3k data figure link
Updated the PBMC3k section

@@
 Peripheral Blood Mononuclear Cells (PBMC) data. Note that
-\citet{xia2023} uses a different subset from \textbf{?@sec-pbmc}, which
-contains \(31,021\) cells including cell type labels, and the gene
-expression levels were in the unit of log-transformed UMI count per
-\(10,000\). They focused on three sequencing methods (inDrops, DropSeq,
-and SeqWell) and four common cell types Cytotoxic T cell, CD4+T cell,
-CD14+ Monocyte, and B cell.
+\citet{xia2023} uses a different subset from the PBMC dataset used by
+\citet{chen2023}, which contains \(31,021\) cells including cell type
+labels, and the gene expression levels were in the unit of
+log-transformed UMI count per \(10,000\). They focused on three
+sequencing methods (inDrops, DropSeq, and SeqWell) and four common cell
+types Cytotoxic T cell, CD4+T cell, CD14+ Monocyte, and B cell.
 
@@
 
-\textbf{?@fig-model-pbmc-author-proj-scdeed} shows layout from the
-original published article along with that suggested by scDEED to be
-more accurate. The log MSE vs binwidth plot illustrates that our
-approach would suggest that scDEED is correct here, that layout b is
-more accurately reflecting the cluster structure in the data.
+Figure~\ref{fig-pbmc-mse-umap} shows layout from the original published
+article along with that suggested by scDEED to be more accurate. The log
+MSE vs binwidth plot illustrates that our approach would suggest that
+scDEED is correct here, that layout b is more accurately reflecting the
+cluster structure in the data.
 
@@
 
-\includegraphics[width=1\textwidth,height=0.5\textheight]{appendix_files/figure-pdf/fig-pbmc-mse-umap-1.pdf}
+\includegraphics[width=1\textwidth,height=0.53\textheight]{appendix_files/figure-pdf/fig-pbmc-mse-umap-1.pdf}
 
@@
 \caption{\label{fig-pbmc-mse-umap}Assessing which of the 2 UMAP layouts
-with different hyperparameter settings (n\_neighbors: \(30\), min\_dist:
-\(0.3\) (red), n\_neighbors: \(80\), min\_dist: \(0.5\) (blue)) on the
-PBMC data is the better representation using MSE for varying binwidth
-(\(a_1\)). Colour used for the lines and points in the left plot and in
-the scatterplots represents UMAP layout (a, b). Of the two, layout b is
-optimal across all binwidths making it the best choice.}
+with different hyper-parameter settings (n\_neighbors: \(30\),
+min\_dist: \(0.3\) (red); n\_neighbors: \(80\), min\_dist: \(0.5\)
+(blue)) on the PBMC data is the better representation using MSE for
+varying binwidth (\(a_1\)). Colour used for the lines and points in the
+top plot and in the scatterplots represents UMAP layout (a, b). Of the
+two, layout b is optimal across all binwidths making it the best
+choice.}
 

### 2025-01-17 paper.qmd
Added text to supplementary section
Added description to the True model

@@
 
-To illustrate the method, we use $7\text{-}D$ simulated data, which we call the "S-curve".
-It is constructed by simulating $n=750$ observations from $\theta \sim U(-3\pi/2, 3\pi/2)$, $X_1 = \sin(\theta)$, $X_2 \sim U(0, 2)$ (adding thickness to the S), $X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)$.
+To illustrate the method, we use $7\text{-}D$ simulated data, which we call the "non-linear clusters".
+It is constructed by simulating $n=2000$ observations from $\theta \sim U(-3\pi/2, 0)$, $X_1 = \sin(\theta)$, $X_2 \sim U(0, 2)$ (adding thickness to the C), $X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)$.
 The remaining variables $X_4, X_5, X_6, X_7$ are all uniform error, with small variance.
@@
 
+Direct links to videos for viewing online are available in .
+
 ## Acknowledgments

### 2025-01-19 paper.qmd
Updated the true model

@@
 To illustrate the method, we use $7\text{-}D$ simulated data, which we call the "non-linear clusters".
-It is constructed by simulating $n=2000$ observations from $\theta \sim U(-3\pi/2, 0)$, $X_1 = \sin(\theta)$, $X_2 \sim U(0, 2)$ (adding thickness to the C), $X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)$.
-The remaining variables $X_4, X_5, X_6, X_7$ are all uniform error, with small variance.
-We would consider $T=(X_1, X_2, X_3)$ to be the geometric structure (true model) that we hope to capture.
+It is constructed by simulating two clusters, each consisting of $1000$ observations.
+The C-shaped cluster is generated from $\theta \sim U(-3\pi/2, 0)$, $X_1 = \sin(\theta)$, $X_2 \sim U(0, 2)$ (adding thickness to the C), $X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)$.
+The other cluster is from $X_1 \sim U(0, 2)$, $X_2 \sim U(0, 3)$, $\gamma \sim U(0, 0.5)$, and $X_3 = -(X_1^3 + X_2) + \gamma$.
+
+The remaining variables $X_4, X_5, X_6, X_7$ are all uniform error, with small variance. We would consider $T=(X_1, X_2, X_3)$ to be the geometric structure (true model) that we hope to capture.
 

### 2025-01-19 paper.qmd
Updated the paper
Discussion intro paragraph

@@
 The other cluster is from $X_1 \sim U(0, 2)$, $X_2 \sim U(0, 3)$, $\gamma \sim U(0, 0.5)$, and $X_3 = -(X_1^3 + X_2) + \gamma$.
-
-The remaining variables $X_4, X_5, X_6, X_7$ are all uniform error, with small variance. We would consider $T=(X_1, X_2, X_3)$ to be the geometric structure (true model) that we hope to capture.
+The remaining variables $X_4, X_5, X_6, X_7$ are all uniform error, with small variance.
+We would consider $T=(X_1, X_2, X_3)$ to be the geometric structure (true model) that we hope to capture.
 
@@
 #| fig-cap: "The choice of long edge removal significantly impacts the model's accuracy.
-\\gD{} Vs $7\\text{-}D$ distances between the bin centroids, where the model is fitted with UMAP for the non-linear cluster data.
+\\gD{} Vs $7\\text{-}D$ distances between the bin centroids, where the model is fitted with UMAP for the non-linear clusters data.
 The two benchmarks for long edge removal: (b) benchmark = 6.5$a_1$ and (c) benchmark = 2$a_1$, where $a_1$ = 0.03 are made.
@@
 This study makes several important contributions to the field of NLDR.
-We have developed an algorithm to evaluate the most useful NLDR method and (hyper-)parameter choices for creating an accurate \gD{} layout of high-dimensional data.
-Our objective is to fit a model for the \gD{} layout that preserves the relationships between neighboring points and turns it into a high-dimensional wireframe, which can be overlaid on the data and visualized using a tour.
+We have developed an algorithm to evaluate the most reasonable NLDR method and (hyper-)parameter choices for creating an accurate \gD{} layout of high-dimensional data.
+Our objective is to fit a model for the \gD{} layout as a \gD{} wireframe and lifts it into \pD{}, which can be overlaid on the data and visualized using a tour.
 This approach is defined as *model-in-data-space*.

### 2025-01-19 paper.qmd
Updated discussion

@@
 
-The effectiveness of this approach is illustrated through various examples.
-For instance, the two-curvy clusters example demonstrates how the model accurately fits the points, capturing both local and global structures in high-dimensional space.
-Our simulation case study further, five Gaussian cluster example shows that while all observed NLDR methods preserve the global structure, only tSNE effectively maintains the local structure, highlighting the specific strengths and quirks of different methods.
+With the model fit, how the NLDR methods organize points in \gD{} layout is examined using the simulated $4\text{-}D$ data having five Gaussian clusters.
+tSNE, UMAP, and PaCMAP preserved the global structure having well-separated clusters, while tSNE, and UMAP effectively maintains the local structure, showing the fitted model as *crumpled \gD{} sheet* that stretches in all four dimensions.
+The interesting result is that the model fit of PaCMAP show *pancake* shape.
 
@@
 
-In conclusion, while our method effectively captures and represents high-dimensional data structures, further enhancements could involve introducing approaches to bind the data, indicate line segments beyond \gD{}, and diagnose the fitted model.
+To extend layouts beyond $k$-D, when 2D is clearly inadequate, can use k-means instead of bin centroids.
+
+In conclusion, while our method effectively captures and represents high-dimensional data structures, further enhancements could involve introducing approaches to diagnose the fitted model.
 These improvements would help in creating a more accurate representation of the data when \gD{} layout is inadequate.

### 2025-01-20 paper.qmd
Add intro to discussion

@@
 
-The MSE for different binwidths ($a_1$) using tSNE, UMAP, PHATE, PaCMAP, and TriMAP with default (hyper-)parameter setting (@fig-mnist-mse) were calculated.
-It is found that tSNE (@fig-mnist-mse (f)) provide the most reasonable representation for the digit 1 dataset, showing universally best.
+The MSE for different binwidths ($a_1$) using UMAP, PHATE, PaCMAP, TriMAP with default (hyper-)parameter setting and tSNE with default (hyper-)parameter setting and perplexity $89$ (@fig-mnist-mse) were calculated.
+It is found that tSNE with perplexity $89$ (@fig-mnist-mse (f)) provide the most reasonable representation for the digit 1 dataset, showing universally best.
 However, \gD{} representation shows a big non-linear cluster and a small cluster with a small gap.
@@
 
-This study makes several important contributions to the field of NLDR.
-We have developed an algorithm to evaluate the most reasonable NLDR method and (hyper-)parameter choices for creating an accurate \gD{} layout of high-dimensional data.
-Our objective is to fit a model for the \gD{} layout as a \gD{} wireframe and lifts it into \pD{}, which can be overlaid on the data and visualized using a tour.
+We have developed an algorithm to assess the NLDR method and (hyper-)parameter choice(s) and choose the most accurate representation of the structure(s) present in the \pD{} data.
+Starting from \gD{} layout, the fitted model, as represented by the positions of points in \kD{}, and turn it into a \pD{} wireframe to overlay on the data, viewing it with a tour.
 This approach is defined as *model-in-data-space*.
@@
 
-To extend layouts beyond $k$-D, when 2D is clearly inadequate, can use k-means instead of bin centroids.
+To extend layouts beyond \kD{}, when \gD{} is clearly inadequate, can use k-means instead of bin centroids.
 

### 2025-01-20 paper.qmd
Updated supplementary material section

@@
 
-With the model fit, how the NLDR methods organize points in \gD{} layout is examined using the simulated $4\text{-}D$ data having five Gaussian clusters.
-tSNE, UMAP, and PaCMAP preserved the global structure having well-separated clusters, while tSNE, and UMAP effectively maintains the local structure, showing the fitted model as *crumpled \gD{} sheet* that stretches in all four dimensions.
-The interesting result is that the model fit of PaCMAP show *pancake* shape.
+With the model fit, some quirks can be observed.
+The simulated $4\text{-}D$ data having five Gaussian clusters is used.
+tSNE, and UMAP which showed the fitted model as *crumpled \gD{} sheet* that stretches in all four dimensions evident of preserving local structure.
+The flat model generated by PaCMAP evident the fail to capture the within cluster variation.
 
@@
 
-Appendix: The appendix includes more details about the hexagonal binning algorithm (appendix.pdf, Portable Document Format file).
+Appendix: The appendix includes more details about the hexagonal binning algorithm and scDEED method (appendix.pdf, Portable Document Format file).
 

### 2025-01-20 paper.qmd
Start updating the discussion

@@
 @fig-one-dens_clust-error b1 and b2 show the fitted models in projections of the $7\text{-}D$ space.
-In these projections, the differences in the distribution of high-dimensional model errors can be observed.
-The uniform C-shaped cluster displays randomly distributed model errors, while the dense C-shaped cluster shows high model errors at the sparse end.
-In the dense C-shaped cluster, the high model errors occur because there are fewer data points positioned within the bins at the sparse end of the structure.
+In these projections, the differences in the distribution of \pD{} model errors can be observed.
+The uniform C-shaped structure displays randomly distributed model errors, while the dense C-shaped structure shows high model errors at the sparse end.
+In the dense C-shaped structure, the high model errors occur because there are fewer data points positioned within the bins at the sparse end of the structure.
 
@@
 
-With the model fit, some quirks can be observed.
-The simulated $4\text{-}D$ data having five Gaussian clusters is used.
-tSNE, and UMAP which showed the fitted model as *crumpled \gD{} sheet* that stretches in all four dimensions evident of preserving local structure.
-The flat model generated by PaCMAP evident the fail to capture the within cluster variation.
+With the model fit, several observations can be made.
+The simulated $4\text{-}D$ data, which contains five Gaussian clusters, was analyzed.
+Both tSNE and UMAP demonstrated the fitted model as a *crumpled \gD{} sheet* that stretches across all four dimensions, highlighting its ability to preserve local structure.
+In contrast, the flat representation generated by PaCMAP indicates its failure to capture the variation within the clusters.
+This highlights how different NLDR methods organize the data points differently.
+
+Another usability of our approach is how NLDR methods behave with different density of the data structure.
+The tSNE layout of simulated $7\text{-}D$ C-shaped data with different densities are used.
+The high \pD{} model errors are scattered to a corner of the \gD{} layout of dense C-shaped structure.
+In the dense C-shaped structure, the high model errors occur because there are fewer data points positioned within the bins at the sparse end of the structure.
 
@@
 
-The **digit: 1** example further illustrates the model's ability to accurately capture non-linear structures and provide additional information.
+The digit 1 of **Hand-written digits** data shows how the model accurately capture the non-linear structure. Also, twisted pattern of the model evident that
+
+example further illustrates the model's ability to accurately capture non-linear structures and provide additional information.
 Key findings include a twisted pattern that compresses the structure in some projections and long line segments that detect anomalies.
@@
 
-To extend layouts beyond \kD{}, when \gD{} is clearly inadequate, can use k-means instead of bin centroids.
+Furthermore, to extend layouts beyond \kD{}, when \gD{} is clearly inadequate, k-means can be used as an alternative to bin centroids.
 

### 2025-01-20 paper.qmd
Explain why PaCMAP doesn't preserve local structure

@@
 So the layout of the model in $4\text{-}D$ doesn't reflect the dimension of each cluster.
-DHC: is there something in the documentation that might help to explain what has happened?
+This can be happen because PaCMAP balances three components (near, mid-range, and far) rather than focusing solely on near neighbors.
+Mid-range neighbors ensure that transitions between clusters or groups are preserved and less concern with local preservation.
+
+<!-- DHC: is there something in the documentation that might help to explain what has happened? -->
 

### 2025-01-21 paper.qmd
Added future work for the discussion

@@
 
-In conclusion, while our method effectively captures and represents high-dimensional data structures, further enhancements could involve introducing approaches to diagnose the fitted model.
+There are many avenues for future work. First, the development of
+evaluation metrics is important for NLDR.
+We introduced a qualitative approach.
+But having a quantitative measure also will be useful for various application domains.
+Second, one could use our approach to create an approach for tuning (hyper-)parameters.
+Another direction for future work is to introduce approaches to diagnose the fitted model.
+Finally, one can think of designing a new NLDR method by following the binning in high-dimensions.
+
+<!-- In conclusion, while our method effectively captures and represents high-dimensional data structures, further enhancements could involve introducing approaches to diagnose the fitted model.
 These improvements would help in creating a more accurate representation of the data when \gD{} layout is inadequate.
+-->
 

### 2025-01-21 paper.qmd
Designed the table for supplementary

@@
 Additionally, the model discovers non-uniform data distribution and non-linear structures within the clusters that are not visible in the UMAP layout, demonstrating the ability of our model in uncovering hidden data characteristics.
+(added about scDEED in Appendix as well)
 
@@
 
-Direct links to videos for viewing online are available in .
+Direct links to videos for viewing online are available in @tbl-links.
+
+```{r}
+#| label: tbl-links
+#| tbl-cap: "Example videos."
+
+links_df <- tibble(
+  data = c("PBMC3k"),
+  URL = c("link1")
+)
+
+links_df |>
+  kableExtra::kable(format = "latex", 
+                    booktabs = TRUE, 
+                    label = "tbllinks") |>
+  kableExtra::kable_styling(latex_options = "scale_down")
+
+```
 

### 2025-01-21 paper.qmd
Added table position for link table

@@
 
-The digit 1 of **Hand-written digits** data shows how the model accurately capture the non-linear structure. Also, twisted pattern of the model evident that
-
-example further illustrates the model's ability to accurately capture non-linear structures and provide additional information.
-Key findings include a twisted pattern that compresses the structure in some projections and long line segments that detect anomalies.
+The digit 1 of **Hand-written digits** data shows how the model accurately capture the non-linear structure.
+Not only that, the \pD{} model error help to find the anomalies of digit 1 images.
+Also, twisted pattern of the model evident that how tSNE tries to maintain the local structure by positioning the similar patterns of digit 1 close to each other.
 
@@
 #| tbl-cap: "Example videos."
+#| tbl-pos: H
 

### 2025-01-21 paper.qmd
Added PBMC3k results to discussion

@@
 
-Human behavior often shows a desire for more certainty and a tendency to prefer well-separated views.
-This emphasizes the importance of clear and distinct clusters.
-For example, in the UMAP layout of the **PBMC3k** dataset suggested by @chen2023, three distant, well-separated clusters are shown.
-However, our model reveals that these clusters are actually close to each other in \pD{}.
+Our algorithm can also be useful in practice to decide on a reasonable NLDR layout that accurately represents the data, as described with a single-cell gene expression data, **PBMC3k**.
+The published results of UMAP with default (hyper-)parameter setting by @chen2023 were evaluated.
+The data showed three clusters, two of which were close.
+But the published layout showed three well-separated clusters.
+Therefore, by changing the (hyper-)parameter, we found a better layout with UMAP, which is more representative the data.
 Additionally, the model discovers non-uniform data distribution and non-linear structures within the clusters that are not visible in the UMAP layout, demonstrating the ability of our model in uncovering hidden data characteristics.
-(added about scDEED in Appendix as well)
-
-Evaluating the error or unexplained variance is important for assessing how well the model fits the data.
-By examining the error for different numbers of bins, we found that tSNE with a perplexity of $30$ provides a reasonable representation for the **pbmc** dataset.
-Connecting the closest clusters with line segments in the fitted model further supports the preservation of neighborhood relationships.
+Also, as explained in Appendix, with recently introduced approach, scDEED also proved the layout suggested by our layout is better than the published one.
 

### 2025-01-21 paper.qmd
Added link plots intro
Start writing about link plots

@@
 
-## Linked plots
-
-It’s important to access the \gD{} layout and the generated model overlaid on data in \pD{} together to understand whether it fits the points everywhere, fits better in some places, or simply mismatches the pattern.
-Interactivity also helps in understanding the quirks that occur with different NLDR techniques (XXXXRefer the video after finalising the two C-shaped clusters data).
-
 ## Curiosities
@@
 
+## Linked plots
+
+It’s important to access the \gD{} layout and the generated model overlaid on data in \pD{} together to understand whether it fits the points everywhere, fits better in some places, or simply mismatches the pattern.
+Interactivity also helps in understanding the quirks.
+
+The embedding and tour views are linked together via rectangular brushes; when a brush is active, points will be highlighted in the adjacent view.
+Because the tour is dynamic, brush events that become active will pause the animation, so that a user can interrogate the current view.
+By default, brushing on the embedding view will produce a one-to-one linking with the tour view.
+For interpreting specific combinations of clusters, the multiple guides on the legend can be selected in order to see their relative orientations.
+
+The interface is constructed as a **shinygadget** specifically designed for interactive data analysis.
+
 ## Discussion {#sec-discussion}

### 2025-01-22 paper.qmd
Intro section for link plot

--- 2025-01-21 paper.qmd
@@
 
+## Linked plots
+
+It’s important to access the \gD{} layout and the generated model overlaid on data in \pD{} together to understand whether it fits the points everywhere, fits better in some places, or simply mismatches the pattern.
+The \gD{} layout and the langevitour view with the model are linked together via rectangular brushes; when a brush is active, points will be highlighted in the adjacent view.
+Because the langevitour is dynamic, brush events that become active will pause the animation, so that a user can interrogate the current view.
+The interface is constructed as a **browsable HTML widget** specifically designed for interactive data analysis.
+
 ## Discussion {#sec-discussion}

### 2025-01-22 paper.qmd
Added details of link plots

@@
 
-### The effect of density
+### The effect of density {#sec-effect-dens}
 
@@
 
-### Hand-written digits
+### Hand-written digits {#sec-mnist}
 <!--
@@
 
+Brushing the high model error in the \gD{} layout, with dense C-shaped data explained in @sec-effect-dens, helps in locating the positions of the related \pD{} points in langevitour view.
+This process helps in identifying the high model errors that occur due to the sparse end (add the video).
+
+Furthermore, when examining the digit 1 from the MNIST dataset (@sec-mnist), brushing the small cluster in the \gD{} layout helps in locating the positions of the related \pD{} points in langevitour view.
+This process helps in identifying how close the small cluster to the big cluster (add the video).
+We can also explore which data corresponds to the twisted pattern of the big cluster (add the video).
+
 ## Discussion {#sec-discussion}

### 2025-01-22 paper.qmd
Updated Figure 2 title

@@
 #| label: fig-two-curvy-true-proj
-#| fig-cap: "Two projections from $7\\text{-}D$, for the two non-linear clusters data are shown.
-The sample obtained from the true model is shown, with the sample shown in grey and the true model in blue lines.
-In the C-shaped cluster, the data points are distributed randomly along the cluster, while the data is more concentrated more in one corner in the other non-linear cluster.
+#| fig-cap: "Two \\gD{} projections from $7\\text{-}D$, for the two non-linear clusters data (grey points) and true model (blue lines) are shown.
+The top cluster in (a) shows uniformly distributed data along a C-shape, while in the bottom cluster, the data is more concentrated in one corner of the curvy shape.
 Video of the langevitour animations are available at <>."
@@
 The MNIST digit 1 data ($n =  7877$ and $p = 10$) has a non-linear structure in $10\\text{-}D$.
-Two \\gD{} projection from a tour on $10\\text{-}D$ reveal that the closeness of the clusters in $10\\text{-}D$ and the twisted pattern of the model fit with tSNE ($a_1 = 0.041$, $b = 659/585 \\text{ } (30, \\text{ }48)$).
+Two \\gD{} projections from a tour on $10\\text{-}D$ reveal that the closeness of the clusters in $10\\text{-}D$ and the twisted pattern of the model fit with tSNE ($a_1 = 0.041$, $b = 659/585 \\text{ } (30, \\text{ }48)$).
 Video of the langevitour animation is available at <>."

### 2025-02-03 paper.qmd
Added linked plot section to Methods
Updated the data

@@
 wireframe_true_model <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_true_model_connections.rds")
-
-data_two_curvy <- training_data_two_curvy |>
+training_data_two_curvy_not_std <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_data_without_std.rds")
+
+data_two_curvy <- training_data_two_curvy_not_std |>
   mutate(type = "data")
@@
 projection <- cbind(
-    c(0.11062,-0.02743,0.02786,-0.01930,0.02374,-0.03493,0.03159),
-    c(-0.01110,-0.06886,0.04730,0.09890,-0.00571,-0.00038,0.00169))
+    c(0.11062,-0.02743,0.02786,-0.01930),
+    c(-0.01110,-0.06886,0.04730,0.09890))
 
@@
 projection <- cbind(
-    c(-0.03372,0.07726,0.03948,-0.00096,-0.02976,-0.07968,0.03194),
-    c(0.07080,0.00316,-0.00382,0.05069,-0.07334,-0.02506,-0.05749))
+    c(-0.03372,0.07726,0.03948,-0.00096),
+    c(0.07080,0.00316,-0.00382,0.05069))
 
@@
 
+### Linked plots
+
+It’s important to access the \gD{} layout and the generated model overlaid on data in \pD{} together to understand whether it fits the points everywhere, fits better in some places, or simply mismatches the pattern.
+The \gD{} layout and the langevitour view with the model are linked together via rectangular brushes; when a brush is active, points will be highlighted in the adjacent view.
+Because the langevitour is dynamic, brush events that become active will pause the animation, so that a user can interrogate the current view.
+The interface is constructed as a **browsable HTML widget** specifically designed for interactive data analysis.
+
+Brushing the high model error in the \gD{} layout, with dense C-shaped data explained in @sec-effect-dens, helps in locating the positions of the related \pD{} points in langevitour view.
+This process helps in identifying the high model errors that occur due to the sparse end (add the video).
+
+Furthermore, when examining the digit 1 from the MNIST dataset (@sec-mnist), brushing the small cluster in the \gD{} layout helps in locating the positions of the related \pD{} points in langevitour view.
+This process helps in identifying how close the small cluster to the big cluster (add the video).
+We can also explore which data corresponds to the twisted pattern of the big cluster (add the video).
+
 ## Best fit
@@
 projection <- cbind(
-  c(-0.08707,-0.06493,0.07017,-0.07349,-0.02955,-0.05286,-0.02765),
-  c(0.03398,0.03735,0.09272,0.01871,-0.09224,0.00451,0.08079))
+  c(-0.08707,-0.06493,0.07017,-0.07349),
+  c(0.03398,0.03735,0.09272,0.01871))
 
@@
 
-## Linked plots
-
-It’s important to access the \gD{} layout and the generated model overlaid on data in \pD{} together to understand whether it fits the points everywhere, fits better in some places, or simply mismatches the pattern.
-The \gD{} layout and the langevitour view with the model are linked together via rectangular brushes; when a brush is active, points will be highlighted in the adjacent view.
-Because the langevitour is dynamic, brush events that become active will pause the animation, so that a user can interrogate the current view.
-The interface is constructed as a **browsable HTML widget** specifically designed for interactive data analysis.
-
-Brushing the high model error in the \gD{} layout, with dense C-shaped data explained in @sec-effect-dens, helps in locating the positions of the related \pD{} points in langevitour view.
-This process helps in identifying the high model errors that occur due to the sparse end (add the video).
-
-Furthermore, when examining the digit 1 from the MNIST dataset (@sec-mnist), brushing the small cluster in the \gD{} layout helps in locating the positions of the related \pD{} points in langevitour view.
-This process helps in identifying how close the small cluster to the big cluster (add the video).
-We can also explore which data corresponds to the twisted pattern of the big cluster (add the video).
-
 ## Discussion {#sec-discussion}

### 2025-02-04 paper.qmd
Updated Figure 3

@@
 # sc_ylims <- c(-0.32, 2.1)
-sc_xlims <- c(-0.27, 1.2)
-sc_ylims <- c(-0.2, 1.1)
+sc_xlims <- c(-0.3, 1.2)
+sc_ylims <- c(-0.17, 1.2)
 
@@
   interior_annotation("b", sc_ltr_pos) + 
-  xlim(c(-0.3, 1.2)) + ylim(c(-0.17, 1.1))
+  xlim(sc_xlims) + ylim(sc_ylims)
 
@@
 projection <- cbind(
-    c(0.11062,-0.02743,0.02786,-0.01930),
-    c(-0.01110,-0.06886,0.04730,0.09890))
+  c(0.37847,-0.06517,0.04231,0.01655),
+  c(-0.06466,-0.27052,0.22131,-0.15235))
 
@@
 axes_obj <- gen_axes(  
-  proj = projection * 5,
-  limits = 0.15,
-  axis_pos_x = -0.09,
-  axis_pos_y = -0.09,
+  proj = projection * 2,
+  limits = 0.6,
+  axis_pos_x = -0.4,
+  axis_pos_y = -0.4,
   axis_labels = names(scaled_two_curvy_data),
-  threshold = 0.007)
+  threshold = 0.04)
 
@@
   coord_fixed() +
-  xlim(c(-0.12, 0.12)) + 
-  ylim(c(-0.12, 0.12)) +
+  xlim(c(-0.52, 0.52)) + 
+  ylim(c(-0.52, 0.52)) +
   interior_annotation("a", c(0.08, 0.93), cex = 2) +
@@
 projection <- cbind(
-    c(-0.03372,0.07726,0.03948,-0.00096),
-    c(0.07080,0.00316,-0.00382,0.05069))
+  c(-0.38281,-0.04870,0.00619,-0.02451),
+  c(-0.01303,0.00634,-0.37421,0.09648))
 
@@
   proj = projection * 6,
-  limits = 0.1,
-  axis_pos_x = -0.07,
-  axis_pos_y = -0.07,
+  limits = 0.5,
+  axis_pos_x = -0.35,
+  axis_pos_y = -0.35,
   axis_labels = names(scaled_two_curvy_data),
-  threshold = 0.007)
+  threshold = 0)
 
@@
   coord_fixed() +
-  xlim(c(-0.09, 0.09)) + 
-  ylim(c(-0.09, 0.09)) +
+  xlim(c(-0.4, 0.4)) + 
+  ylim(c(-0.4, 0.4)) +
   interior_annotation("b", c(0.08, 0.93), cex = 2) +

### 2025-02-04 paper.qmd
Updated Figure 2

@@
   limits = 0.5,
-  axis_pos_x = -0.35,
-  axis_pos_y = -0.35,
+  axis_pos_x = -0.3,
+  axis_pos_y = -0.3,
   axis_labels = names(scaled_two_curvy_data),

### 2025-02-04 paper.qmd
Updated Figure 2

@@
 
-data_two_curvy <- training_data_two_curvy_not_std |>
+data_two_curvy <- training_data_two_curvy |>
+  mutate(type = "data")
+
+data_two_curvy_not_std <- training_data_two_curvy_not_std |>
   mutate(type = "data")
@@
   coord_fixed() +
-  xlim(c(-0.12, 0.12)) + 
-  ylim(c(-0.12, 0.12)) +
+  xlim(c(-0.25, 0.25)) + 
+  ylim(c(-0.25, 0.25)) +
   interior_annotation("c", sc_ltr_pos) +
@@
   coord_fixed() +
-  xlim(c(-0.1, 0.1)) + 
-  ylim(c(-0.1, 0.1)) +
+  xlim(c(-0.25, 0.25)) + 
+  ylim(c(-0.25, 0.25)) +
   interior_annotation("c2", sc_ltr_pos) +
@@
 # Apply the scaling
-df_model_data_two_curvy <- bind_rows(data_two_curvy, true_model_two_curvy)
+df_model_data_two_curvy <- bind_rows(data_two_curvy_not_std, true_model_two_curvy)
 

### 2025-02-04 paper.qmd
Updated Figure 2

@@
 axes_obj <- gen_axes(  
-  proj = projection * 5,
-  limits = 0.15,
-  axis_pos_x = -0.09,
-  axis_pos_y = -0.09,
+  proj = projection * 2,
+  limits = 0.3,
+  axis_pos_x = -0.2,
+  axis_pos_y = -0.2,
   axis_labels = names(scaled_two_curvy_data),
-  threshold = 0.007)
+  threshold = 0.02)
 
@@
   coord_fixed() +
-  xlim(c(-0.12, 0.12)) + 
-  ylim(c(-0.12, 0.12)) +
+  xlim(c(-0.25, 0.25)) + 
+  ylim(c(-0.25, 0.25)) +
   interior_annotation("d", sc_ltr_pos) +

### 2025-02-04 paper.qmd
Updated Figure 9

@@
 true_model_df <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_true_model.rds")
+true_model_df_std <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_true_model_std.rds")
 wireframe_true_model <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_true_model_connections.rds")
@@
   mutate(type = "data")
+
+true_model_two_curvy_std <- true_model_df_std |> 
+  select(-ID) |> 
+  mutate(type = "true model")
 
@@
 
-scaled_two_curvy_data_true_model <- scaled_two_curvy |>
+scaled_two_curvy_data_true_model <- df_model_data_two_curvy |>
   filter(type == "true model") |>

### 2025-02-04 scripts/two_non_linear_diff_shaped_close_clusters/10_gen_rm_lg_plots.R

@@ -309,10 +309,10 @@ scaled_two_curvy_data_model <- scaled_two_curvy |>
 
 ## First projection
 projection <- cbind(
-  c(-0.02291,-0.00834,0.06450,-0.09921),
-  c(-0.11183,-0.02761,-0.04394,-0.01066))
+  c(0.37847,-0.06517,0.04231,0.01655),
+  c(-0.06466,-0.27052,0.22131,-0.15235))
 
-projection_scaled <- projection * 5
+projection_scaled <- projection * 1
 
 projected <- as.matrix(scaled_two_curvy_data) %*% projection_scaled
 
@@ -366,12 +366,12 @@ model_df2 <- dplyr::left_join(model_df2, projected_model_df, by = c("to" = "ID")
 names(model_df2)[(2 + NCOL(projected_model_df)):NCOL(model_df2)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
 
 axes_obj <- gen_axes(
-  proj = projection * 5,
-  limits = 0.5,
-  axis_pos_x = -0.4,
-  axis_pos_y = -0.4,
+  proj = projection * 2,
+  limits = 0.6,
+  axis_pos_x = -0.5,
+  axis_pos_y = -0.5,
   axis_labels = names(scaled_two_curvy_data),
-  threshold = 0.03)
+  threshold = 0.04)
 
 axes <- axes_obj$axes
 circle <- axes_obj$circle
@@ -407,8 +407,8 @@ two_curvy_proj_first_model1_umap <- projected_df |>
     data=circle,
     aes(x=c1, y=c2), colour="grey70") +
   coord_fixed() +
-  xlim(c(-0.5, 0.5)) +
-  ylim(c(-0.5, 0.5)) +
+  xlim(c(-0.6, 0.6)) +
+  ylim(c(-0.6, 0.6)) +
   interior_annotation("c2", sc_ltr_pos) +
   theme(
     legend.position = "none"
@@ -447,8 +447,8 @@ two_curvy_proj_model_delaunay_umap <- projected_df |>
     data=circle,
     aes(x=c1, y=c2), colour="grey70") +
   coord_fixed() +
-  xlim(c(-0.5, 0.5)) +
-  ylim(c(-0.5, 0.5)) +
+  xlim(c(-0.6, 0.6)) +
+  ylim(c(-0.6, 0.6)) +
   interior_annotation("a2", sc_ltr_pos) +
[…]

### 2025-02-04 paper.qmd
Updated Figure 7

@@
 sc_xlims <- c(-0.3, 1.2)
-sc_ylims <- c(-0.17, 1.2)
+sc_ylims <- c(-0.17, 1.25)
 
@@
     na.value = "#ffffff", option = "C") +
-  xlim(c(-0.3, 1.2)) + ylim(c(-0.17, 1.1)) +
+  xlim(sc_xlims) + ylim(sc_ylims) +
   interior_annotation("a", sc_ltr_pos)
@@
   scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "D") +
-  xlim(c(-0.3, 1.2)) + ylim(c(-0.17, 1.1)) +
+  xlim(sc_xlims) + ylim(sc_ylims) +
   interior_annotation("c", sc_ltr_pos)

### 2025-02-04 paper.qmd
Updated Figure 11

@@
   limits = 1,
-  axis_pos_x = -0.6,
-  axis_pos_y = -0.63,
+  axis_pos_x = -0.72,
+  axis_pos_y = -0.72,
   axis_labels = names(scaled_gau_data),
@@
   coord_fixed() +
-  xlim(c(-0.84, 0.84)) +
-  ylim(c(-0.84, 0.84)) +
+  xlim(c(-0.89, 0.89)) +
+  ylim(c(-0.89, 0.89)) +
   interior_annotation("a2", 
@@
   coord_fixed() +
-  xlim(c(-0.83, 0.83)) +
-  ylim(c(-0.83, 0.83)) +
+  xlim(c(-0.89, 0.89)) +
+  ylim(c(-0.89, 0.89)) +
   interior_annotation("b2", 
@@
   coord_fixed() +
-  xlim(c(-0.83, 0.83)) +
-  ylim(c(-0.83, 0.83)) +
+  xlim(c(-0.89, 0.89)) +
+  ylim(c(-0.89, 0.89)) +
   interior_annotation("c2",

### 2025-02-04 paper.qmd
Added true model

@@
 It is constructed by simulating two clusters, each consisting of $1000$ observations.
-The C-shaped cluster is generated from $\theta \sim U(-3\pi/2, 0)$, $X_1 = \sin(\theta)$, $X_2 \sim U(0, 2)$ (adding thickness to the C), $X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)$.
-The other cluster is from $X_1 \sim U(0, 2)$, $X_2 \sim U(0, 3)$, $\gamma \sim U(0, 0.5)$, and $X_3 = -(X_1^3 + X_2) + \gamma$.
-The remaining variables $X_4, X_5, X_6, X_7$ are all uniform error, with small variance.
-We would consider $T=(X_1, X_2, X_3)$ to be the geometric structure (true model) that we hope to capture.
+The C-shaped cluster is generated from $\theta \sim U(-3\pi/2, 0)$, $X_1 = \sin(\theta)$, $X_2 \sim U(0, 2)$ (adding thickness to the C), $X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)$, $X_4 = \cos(\theta)$.
+The other cluster is from $X_1 \sim U(0, 2)$, $X_2 \sim U(0, 3)$, $\gamma \sim U(0, 0.5)$, $X_3 = -(X_1^3 + X_2) + \gamma$, and $X_4 \sim U(0, 2)$.
+We would consider $T=(X_1, X_2, X_3, X_4)$ to be the geometric structure (true model) that we hope to capture.
+
+<!-- The remaining variables $X_4, X_5, X_6, X_7$ are all uniform error, with small variance.  -->
 

### 2025-02-04 paper.qmd
Updated the paper

@@
 
-true_model_df <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_true_model.rds")
+true_model_df <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_true_model_std.rds")
 true_model_df_std <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_true_model_std.rds")
@@
 
-scaled_two_curvy_data_true_model <- df_model_data_two_curvy |>
+scaled_two_curvy_data_true_model <- scaled_two_curvy |>
   filter(type == "true model") |>
@@
 # Apply the scaling
-df_model_data_two_curvy <- bind_rows(data_two_curvy, true_model_two_curvy, df_b_two_curvy2)
+df_model_data_two_curvy <- bind_rows(
+  data_two_curvy, 
+  true_model_two_curvy, 
+  df_b_two_curvy2)
 

### 2025-02-05 paper.qmd
Updated Figure 17

@@
   facet_wrap(~ instance, ncol = 4) +
-  ggtitle("(c)") +
+  ggtitle("(a)") +
   coord_fixed() +
@@
   facet_wrap(~ instance, ncol = 4) +
-  ggtitle("(d)") +
+  ggtitle("(b)") +
   coord_fixed() +
@@
   facet_wrap(~ instance, ncol = 4) +
-  ggtitle("(e)") +
+  ggtitle("(c)") +
   coord_fixed() +
@@
   facet_wrap(~ instance, ncol = 4) +
-  ggtitle("(a)") +
+  ggtitle("(d)") +
   coord_fixed() +
@@
   facet_wrap(~ instance, ncol = 4) +
-  ggtitle("(b)") +
+  ggtitle("(e)") +
   coord_fixed() +
@@
 Most low model errors are distributed along the big non-linear cluster, while most large model errors are distributed along the small cluster.
-The images associated with large model errors shows different patterns of digit 1, some inside (b) the non-linear structure and others outside (a).
-Along the non-linear cluster, the angle of digit 1 changes (c-e)."
+The images associated with large model errors shows different patterns of digit 1, some inside (e) the non-linear structure and others outside (d).
+Along the non-linear cluster, the angle of digit 1 changes (a-c)."
 #| fig-pos: H
@@
 
-highd_error_2d_mnist + (imge_error_sample_outside / imge_error_sample_inside / right_top_img / middle_img / right_bottom_img) 
+highd_error_2d_mnist + (right_top_img / middle_img / right_bottom_img / imge_error_sample_outside) / imge_error_sample_inside
 ```

### 2025-02-05 paper.qmd
Updated Figure 17

@@
         plot.title.position = "plot",
-        plot.title = element_text(hjust = 0.31))
+        plot.title = element_text(hjust = 0))
 
@@
 #| fig-width: 10
-#| fig-height: 5
-#| out-width: 80%
-
-highd_error_2d_mnist + (right_top_img / middle_img / right_bottom_img / imge_error_sample_outside) / imge_error_sample_inside
+#| fig-height: 10
+#| out-width: 90%
+
+# Define your current layout
+top_layout <- right_top_img / middle_img / right_bottom_img / imge_error_sample_outside
+full_layout <- highd_error_2d_mnist + top_layout
+
+# Add the new full-width bottom row
+final_layout <- full_layout / imge_error_sample_inside
+
+# Display the final combined plot
+final_layout
 ```

### 2025-02-05 paper.qmd
Updated Figure 17

@@
 ##### inside
-img_error_inside <- c(7228, 7363, 7650)
+img_error_inside <- c(3179, 7363, 7650, 5241, 5173, 6981, 7228, 6673, 4636) #7228, 7363, 7650
 
@@
   geom_tile() +
-  facet_wrap(~ instance, ncol = 4) +
+  facet_wrap(~ instance, ncol = 9) +
   ggtitle("(e)") +
@@
 #| fig-width: 10
-#| fig-height: 10
+#| fig-height: 7
 #| out-width: 90%
@@
 # Add the new full-width bottom row
-final_layout <- full_layout / imge_error_sample_inside
+final_layout <- full_layout / imge_error_sample_inside + plot_layout(heights = c(1, 0.2))
 

### 2025-02-05 paper.qmd
Updated Figure 12

@@
 #| label: fig-one-dens_clust-error
-#| fig-height: 10
-#| fig-width: 10
+#| fig-height: 5
+#| fig-width: 15
 #| fig-pos: H

### 2025-02-06 paper.qmd
Updated Figure title

@@
 #| out-height: 50%
-#| fig-cap: "The sparse end of the structure exhibits a high model error, while the dense end shows a low model error.
-The tSNE layouts of the C-shaped structure, with uniform points (a1) and dense points (b1) are colored according to their high-dimensional error, along with their selected projections (a2 and b2), are displayed.
+#| fig-cap: "The sparse end of the structure shows a high model error, while the dense end shows a low model error.
+The tSNE layout of the C-shaped structure, with dense points (a) are colored according to their high-dimensional error.
 Darker colors represent a high error, while lighter colors indicate a low error.
-The uniformly distributed points demonstrate random model errors throughout the structure, whereas the dense structure shows large model error at the sparse end.
-Videos of the langevitour animations are available at <>."
+b and c present the results from linked plots that highlighted the high-dimensional error points in \\gD{} and $4\\text{-}D$.
+Videos of the linked plots are available at <>."
 

### 2025-02-06 paper.qmd
Updated Figure 12
Added linked plot section

@@
 
-It’s important to access the \gD{} layout and the generated model overlaid on data in \pD{} together to understand whether it fits the points everywhere, fits better in some places, or simply mismatches the pattern.
+Diagnosing the model while locating points in the \gD{} layout and displaying the generated model overlaid on data in the \pD{} is important.
+
 The \gD{} layout and the langevitour view with the model are linked together via rectangular brushes; when a brush is active, points will be highlighted in the adjacent view.
@@
 
-Brushing the high model error in the \gD{} layout, with dense C-shaped data explained in @sec-effect-dens, helps in locating the positions of the related \pD{} points in langevitour view.
+To understand how well the model fits the points—whether it fits consistently, works better in certain areas, or fails to match the overall pattern—it is important to link and brush the points with high model error in the \pD{} error plot, \gD{} layout, and the generated model overlay on the data in \pD{}.
+
+<!-- Brushing the high model error in the \gD{} layout, with dense C-shaped data explained in @sec-effect-dens, helps in locating the positions of the related \pD{} points in langevitour view.
 This process helps in identifying the high model errors that occur due to the sparse end (add the video).
+-->
 
@@
 #| fig-cap: "The sparse end of the structure shows a high model error, while the dense end shows a low model error.
-The tSNE layout of the C-shaped structure, with dense points (a) are colored according to their high-dimensional error.
+The tSNE layout of the C-shaped structure, with dense points (a) are colored according to their $4\\text{-}D$ error.
 Darker colors represent a high error, while lighter colors indicate a low error.
-b and c present the results from linked plots that highlighted the high-dimensional error points in \\gD{} and $4\\text{-}D$.
+b and c present the results from linked plots that brushed the high $4\\text{-}D$ error points in \\gD{} and $4\\text{-}D$.
+It helps in identifying the high $4\\text{-}D$ model errors that occur due to the sparse end.
 Videos of the linked plots are available at <>."

### 2025-02-06 paper.qmd
Updated long edge removal section

@@
 #| label: fig-two-curvy-true-proj
-#| fig-cap: "Two \\gD{} projections from $7\\text{-}D$, for the two non-linear clusters data (grey points) and true model (blue lines) are shown.
+#| fig-cap: "Two \\gD{} projections from $4\\text{-}D$, for the two non-linear clusters data (grey points) and true model (blue lines) are shown.
 The top cluster in (a) shows uniformly distributed data along a C-shape, while in the bottom cluster, the data is more concentrated in one corner of the curvy shape.
@@
 #| label: best-fit-umap1
-#| fig-cap: "We start with a tSNE layout (a) of non-linear clusters data ($n =  2000$ and $p = 7$), from which the \\gD{} wireframe is constructed (b) and then lifted to $7\\text{-}D$.
-A \\gD{} projection of the model fit from a tour through the $7\\text{-}D$ space is presented in c.
+#| fig-cap: "We start with a tSNE layout (a) of non-linear clusters data ($n =  2000$ and $p = 4$), from which the \\gD{} wireframe is constructed (b) and then lifted to $4\\text{-}D$.
+A \\gD{} projection of the model fit from a tour through the $4\\text{-}D$ space is presented in c.
 Additionally, a \\gD{} projection of the fitted and true models (d) from the same tour demonstrates that the model accurately captures the non-linear structure of the data, although it does not fully represent the width of that structure.
@@
 #| label: fig-p-d-error-in-2d-two-curvy
-#| fig-cap: "The $7\\text{-}D$ model error in \\gD{} layout.
+#| fig-cap: "The $4\\text{-}D$ model error in \\gD{} layout.
 Color indicates square root of total error, dark colour indicating high error and light indicates low error.
@@
 Edges define the neighbourhood structure, in order to provide a smooth \gD{} representation of the fitted model.
-@fig-two-curvy-true-proj shows a wire frame of the true model that was used to generate the two C-shaped clusters example data.
+@fig-two-curvy-true-proj shows a wire frame of the true model that was used to generate the non-linear clusters example data.
 The ideal is that the representation of the fitted model, at least for this example where we know the true model, should look similar to this.
@@
 Another usability of our approach is how NLDR methods behave with different density of the data structure.
-The tSNE layout of simulated $7\text{-}D$ C-shaped data with different densities are used.
+The tSNE layout of simulated $4\text{-}D$ C-shaped data with different densities are used.
 The high \pD{} model errors are scattered to a corner of the \gD{} layout of dense C-shaped structure.

### 2025-02-06 paper.qmd
Changed the text of effect density

@@
 
-To illustrate how tSNE organize different number of points within the structure in the tSNE layout, simulated $7\text{-}D$ data having C-shaped structure is used.
-@fig-one-dens_clust-error a1 and b1 shows \gD{} layouts for tSNE.
+To illustrate how tSNE organize different number of points within the structure in the tSNE layout, simulated $4\text{-}D$ data having C-shaped structure is used.
+@fig-one-dens_clust-error b shows \gD{} layouts for tSNE.
 The default hyper-parameters are used.
-The \gD{} layouts are colored according to the high-dimensional model error.
-In b1, the large model error points are clustered in the top left corner, forming a separate cluster with high model errors.
-
-The models are fitted to these layouts.
-@fig-one-dens_clust-error b1 and b2 show the fitted models in projections of the $7\text{-}D$ space.
-In these projections, the differences in the distribution of \pD{} model errors can be observed.
-The uniform C-shaped structure displays randomly distributed model errors, while the dense C-shaped structure shows high model errors at the sparse end.
+The \gD{} layout is colored according to the $4\text{-}D$ model error (@fig-one-dens_clust-error a).
+In @fig-one-dens_clust-error a, the large model error points are clustered in the top corner, forming a separate cluster with high model errors.
+
+The model is fitted to this layout.
+@fig-one-dens_clust-error c shows the $2\text{-}D$ projection of the fitted model.
+In this projection, the distribution of $4\text{-}D$ model errors can be observed.
+The dense C-shaped structure shows high model errors at the sparse end.
 In the dense C-shaped structure, the high model errors occur because there are fewer data points positioned within the bins at the sparse end of the structure.

### 2025-02-06 paper.qmd
Updated Measutring fit

@@
 
-The model here is similar to a confirmatory factor analysis model [@brown2015], $\widehat{T}(X_1, X_2, X_3) + \Epsilon$.
-The difference between the fitted model and observed values would be considered to be residuals, and for this problem are $7\text{-}D$.
+The model here is similar to a confirmatory factor analysis model [@brown2015], $\widehat{T}(X_1, X_2, X_3, X_4) + \Epsilon$.
+The difference between the fitted model and observed values would be considered to be residuals, and for this problem are $4\text{-}D$.
 

### 2025-02-06 paper.qmd
Updated with standardised error
added link brushing for MNIST

@@
 
+tsne_plot_mnist_clust <- ggplot() + 
+  geom_point(data = tsne_minst_scaled_temp,
+             aes(
+               x = tSNE1,
+               y = tSNE2,
+               colour = cluster
+             ),
+             alpha=0.2)+
+  scale_color_manual(values=c('#f0f0f0', '#ff7f00')) +
+  interior_annotation("a1", sc_ltr_pos_mnist) +
+  theme(legend.position = "none")
+
 # trimesh_removed_mnist <- vis_rmlg_mesh(
@@
 error_df <- error_df |>
-  mutate(sqrt_row_wise_tot_error = sqrt(row_wise_total_error))
+  mutate(sqrt_row_wise_tot_error = sqrt(row_wise_total_error)) |>
+  mutate(sqrt_row_wise_tot_error = standardize(sqrt_row_wise_tot_error))
 
@@
 
+mnist_proj_tsne_model1_link <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2,
+      colour = cluster)) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#000000",
+    linewidth = 0.5,
+    alpha = 0.5) +
+  geom_point(
+    size = 1.2,
+    alpha = 0.5) +
+  scale_color_manual(values = c("#d9d9d9",'#ff7f00')) +
+  geom_segment(
+    data=axes,
+    aes(x=x1, y=y1, xend=x2, yend=y2),
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
+    colour="grey50",
+    size = 2) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  #coord_fixed() +
+  xlim(c(-1.1, 0.9)) +
+  ylim(c(-1.1, 0.9)) +
+  interior_annotation("a2") +
+  theme(
+    legend.position = "none"
+  )
+
 ## Second projection
@@
 
+mnist_proj_tsne_model2_link <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2,
+      colour = cluster)) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#000000",
+    linewidth = 0.5) +
+  geom_point(
+    size = 1.2,
+    alpha = 0.3) +
+  scale_color_manual(values = c("#d9d9d9",'#ff7f00')) +
+  geom_segment(
+    data=axes, 
+    aes(x=x1, y=y1, xend=x2, yend=y2), 
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
+    colour="grey50",
+    size = 2) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-1, 1)) +
+  ylim(c(-1, 1)) +
+  interior_annotation("a3") +
+  theme(
+    legend.position = "none"
+  )
+
 ## Third projection
@@
 
+```{r}
+#| label: fig-clust-mnist
+#| fig-pos: H
+#| fig-height: 5
+#| fig-width: 15
+#| out-height: 100%
+
+tsne_plot_mnist_clust + mnist_proj_tsne_model1_link + mnist_proj_tsne_model2_link +
+  plot_layout(ncol = 3)
+```
+
 <!-- ::: {#fig-mnist1-sc layout-ncol="4" fig-pos="H"} -->

### 2025-02-06 scripts/one_c_shaped_dens_structure/04_gen_error_plots.R

@@ -134,7 +134,7 @@ plot_tsne_dens <- error_df_one_curvy_abs |>
              colour = factor(error_cat_n,
                              levels = c("high", "low")))) +
   geom_point(alpha=0.7) +
-  scale_color_manual(values=c('#800026', '#f0f0f0')) +
+  scale_color_manual(values=c('#800026', '#d9d9d9')) +
   theme(
     aspect.ratio = 1
   ) +
@@ -213,7 +213,7 @@ langevitour::langevitour(df_exe[1:(length(df_exe)-1)],
                          lineTo = distance_df_small_edges_c_shaped_structure$to,
                          group = factor(df_exe$type,
                                         c("low", "high","model")), pointSize = append(rep(1, NROW(df_b)), rep(2, NROW(df))),
-                         levelColors = c('#f0f0f0', '#800026', "#000000"))
+                         levelColors = c('#d9d9d9', '#800026', "#000000"))
 
 
 # Apply the scaling
@@ -317,7 +317,7 @@ five_c_shaped_proj_tsne_model1 <- projected_df |>
   geom_path(
     data=circle,
     aes(x=c1, y=c2), colour="grey70") +
-  scale_color_manual(values=c('#f0f0f0', '#800026')) +
+  scale_color_manual(values=c('#d9d9d9', '#800026')) +
   coord_fixed() +
   xlim(c(-0.35, 0.35)) +
   ylim(c(-0.35, 0.35)) +

### 2025-02-06 paper.qmd
Updated Figure 17

--- 2025-02-06 paper.qmd
@@
 
-Furthermore, when examining the digit 1 from the MNIST dataset (@sec-mnist), brushing the small cluster in the \gD{} layout helps in locating the positions of the related \pD{} points in langevitour view.
+<!-- Furthermore, when examining the digit 1 from the MNIST dataset (@sec-mnist), brushing the small cluster in the \gD{} layout helps in locating the positions of the related \pD{} points in langevitour view.
 This process helps in identifying how close the small cluster to the big cluster (add the video).
 We can also explore which data corresponds to the twisted pattern of the big cluster (add the video).
+-->
 
@@
   scale_color_manual(values=c("#999999",'#ff7f00')) +
-  interior_annotation("a1", sc_ltr_pos_mnist) +
+  interior_annotation("a", sc_ltr_pos_mnist) +
+  theme(legend.position = "none")
+
+tsne_plot_mnist_clust <- ggplot() + 
+  geom_point(data = tsne_minst_scaled_temp,
+             aes(
+               x = tSNE1,
+               y = tSNE2,
+               colour = cluster
+             ),
+             alpha=0.2)+
+  scale_color_manual(values=c('#f0f0f0', '#ff7f00')) +
+  interior_annotation("a", sc_ltr_pos_mnist) +
   theme(legend.position = "none")
@@
   ylim(c(-1.1, 0.9)) +
-  interior_annotation("a2") +
+  interior_annotation("b") +
+  theme(
+    legend.position = "none"
+  )
+
+mnist_proj_tsne_model1_link <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2,
+      colour = cluster)) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#000000",
+    linewidth = 0.5,
+    alpha = 0.5) +
+  geom_point(
+    size = 1.2,
+    alpha = 0.5) +
+  scale_color_manual(values = c("#d9d9d9",'#ff7f00')) +
+  geom_segment(
+    data=axes,
+    aes(x=x1, y=y1, xend=x2, yend=y2),
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
+    colour="grey50",
+    size = 2) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  #coord_fixed() +
+  xlim(c(-1.1, 0.9)) +
+  ylim(c(-1.1, 0.9)) +
+  interior_annotation("b") +
   theme(
@@
   ylim(c(-1, 1)) +
-  interior_annotation("a3") +
+  interior_annotation("c") +
+  theme(
+    legend.position = "none"
+  )
+
+mnist_proj_tsne_model2_link <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2,
+      colour = cluster)) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#000000",
+    linewidth = 0.5) +
+  geom_point(
+    size = 1.2,
+    alpha = 0.3) +
+  scale_color_manual(values = c("#d9d9d9",'#ff7f00')) +
+  geom_segment(
+    data=axes, 
+    aes(x=x1, y=y1, xend=x2, yend=y2), 
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
+    colour="grey50",
+    size = 2) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-1, 1)) +
+  ylim(c(-1, 1)) +
+  interior_annotation("c") +
   theme(
@@
 
+```{r}
+#| label: fig-clust-mnist
+#| fig-pos: H
+#| fig-height: 5
+#| fig-width: 15
+#| out-height: 100%
+#| fig-cap: "The tSNE layout of the MNIST digit 1 data shows a big non-linear cluster (grey) and a small cluster (orange) located very close to the one corner of the big cluster in \\gD{} (a).
+The brushing feature in the linked plots helps in visualizing the closeness of the small cluster to the big cluster.
+Videos of the linked plots are available at <>."
+
+tsne_plot_mnist_clust + mnist_proj_tsne_model1_link + mnist_proj_tsne_model2_link +
+  plot_layout(ncol = 3)
+```
+
 <!-- ::: {#fig-mnist1-sc layout-ncol="4" fig-pos="H"} -->

### 2025-02-06 paper.qmd
Updated Figure 17

@@
              alpha=0.2)+
-  scale_color_manual(values=c('#f0f0f0', '#ff7f00')) +
+  scale_color_manual(values=c('#d9d9d9', '#ff7f00')) +
   interior_annotation("a", sc_ltr_pos_mnist) +
@@
 The tSNE layout of simulated $4\text{-}D$ C-shaped data with different densities are used.
-The high \pD{} model errors are scattered to a corner of the \gD{} layout of dense C-shaped structure.
+The high $4\text{-}D$ model errors are scattered to a corner of the \gD{} layout of dense C-shaped structure.
 In the dense C-shaped structure, the high model errors occur because there are fewer data points positioned within the bins at the sparse end of the structure.
@@
 The digit 1 of **Hand-written digits** data shows how the model accurately capture the non-linear structure.
-Not only that, the \pD{} model error help to find the anomalies of digit 1 images.
+Not only that, the $10\text{-}D$ model error help to find the anomalies of digit 1 images.
 Also, twisted pattern of the model evident that how tSNE tries to maintain the local structure by positioning the similar patterns of digit 1 close to each other.
+
+Diagnose the fitted model is important. Interactivity helps to
 
@@
 Second, one could use our approach to create an approach for tuning (hyper-)parameters.
-Another direction for future work is to introduce approaches to diagnose the fitted model.
+Another direction for future work is to introduce more approaches to diagnose the fitted model.
 Finally, one can think of designing a new NLDR method by following the binning in high-dimensions.

### 2025-02-06 paper.qmd
Added linked plot for discussion

@@
 
-To understand how well the model fits the points—whether it fits consistently, works better in certain areas, or fails to match the overall pattern—it is important to link and brush the points with high model error in the \pD{} error plot, \gD{} layout, and the generated model overlay on the data in \pD{}.
+To understand how well the model fits the points—whether it fits well, works better in some positions, or fails to match the overall pattern—it is important to link and brush the points with high model error in the \pD{} error plot, \gD{} layout, and the generated model overlay on the data in \pD{}.
 
@@
 
-Diagnose the fitted model is important. Interactivity helps to
+Diagnose is important to find where the \gD{} points located in \pD{}.
+One approach is to use interactivity.
+By linking the \gD{} layout with the fitted model in \pD{} and employing brushing techniques, we can identify which points fit well and which do not.
+Additionally, linking the \gD{} error plot with the fitted model in \pD{} can help pinpoint where mismatches occur.
 

### 2025-02-06 paper.qmd
Found the solution for true model problem

@@
 true_model_df <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_true_model.rds")
-true_model_df_std <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_true_model_std.rds")
 wireframe_true_model <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_true_model_connections.rds")
-training_data_two_curvy_not_std <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_data_without_std.rds")
 
@@
 
-data_two_curvy_not_std <- training_data_two_curvy_not_std |>
-  mutate(type = "data")
-
-true_model_two_curvy_std <- true_model_df_std |> 
-  select(-ID) |> 
-  mutate(type = "true model")
-
 true_model_two_curvy <- true_model_df |> 
-  select(-ID) |> 
   mutate(type = "true model")
@@
 # Apply the scaling
-df_model_data_two_curvy <- bind_rows(data_two_curvy_not_std, true_model_two_curvy)
+df_model_data_two_curvy <- bind_rows(data_two_curvy, true_model_two_curvy)
 
@@
 #| label: model-prediction-umap-quollr
+#| eval: false
 
@@
 This highlights a key characteristic of tSNE, which compresses the data when applying tSNE.
-Furthermore, the sparse space observed in the middle of the tSNE layout is a result of (hyper-parameter) choice.
+Furthermore, the sparse space observed in the middle of the tSNE layout is a result of (hyper-)parameter choice.
 The default is used.
@@
 #| label: model-prediction-points-line
+#| eval: false
 

### 2025-02-06 paper.qmd
Updated Figure 2

@@
   coord_fixed() +
-  xlim(c(-0.52, 0.52)) + 
-  ylim(c(-0.52, 0.52)) +
+  xlim(c(-0.6, 0.6)) + 
+  ylim(c(-0.6, 0.6)) +
   interior_annotation("a", c(0.08, 0.93), cex = 2) +
@@
   coord_fixed() +
-  xlim(c(-0.1, 0.1)) + 
-  ylim(c(-0.1, 0.1)) +
+  xlim(c(-0.6, 0.6)) + 
+  ylim(c(-0.6, 0.6)) +
   interior_annotation("a", c(0.08, 0.93), cex = 2) +
@@
   coord_fixed() +
-  xlim(c(-0.4, 0.4)) + 
-  ylim(c(-0.4, 0.4)) +
-  interior_annotation("b", c(0.08, 0.93), cex = 2) +
+  xlim(c(-0.42, 0.42)) + 
+  ylim(c(-0.42, 0.42)) +
+  interior_annotation("b", cex = 2) +
   theme(

### 2025-02-06 paper.qmd
Updated Figure 5

@@
   proj = projection * 2,
-  limits = 0.3,
-  axis_pos_x = -0.2,
-  axis_pos_y = -0.2,
+  limits = 0.5,
+  axis_pos_x = -0.5,
+  axis_pos_y = -0.5,
   axis_labels = names(scaled_two_curvy_data),
-  threshold = 0.02)
+  threshold = 0.03)
 
@@
   coord_fixed() +
-  xlim(c(-0.25, 0.25)) + 
-  ylim(c(-0.25, 0.25)) +
+  xlim(c(-0.6, 0.6)) + 
+  ylim(c(-0.6, 0.6)) +
   interior_annotation("c", sc_ltr_pos) +
@@
   coord_fixed() +
-  xlim(c(-0.25, 0.25)) + 
-  ylim(c(-0.25, 0.25)) +
+  xlim(c(-0.6, 0.6)) + 
+  ylim(c(-0.6, 0.6)) +
   interior_annotation("c2", sc_ltr_pos) +
@@
   coord_fixed() +
-  xlim(c(-0.25, 0.25)) + 
-  ylim(c(-0.25, 0.25)) +
+  xlim(c(-0.6, 0.6)) + 
+  ylim(c(-0.6, 0.6)) +
   interior_annotation("d", sc_ltr_pos) +

### 2025-03-10 paper.qmd
Updated Figure 18 text

@@
 #| label: gen-axes
-
 gen_axes <- function(proj, limits = 1, axis_pos_x = NULL, axis_pos_y = NULL, axis_labels, threshold) {
@@
 set.seed(20240110)
-#source("nldr_code.R", local = TRUE)
 ```
@@
         plot.title.position = "plot",
-        plot.title = element_text(hjust = 0.31))
+        plot.title = element_text(hjust = 0.31,
+                                  size = 12))
 
@@
         plot.title.position = "plot",
-        plot.title = element_text(hjust = 0))
+        plot.title = element_text(hjust = 0,
+                                  size = 12))
 

### 2025-03-10 paper.qmd
Updated Figure 3

@@
   xlim(sc_xlims) + ylim(sc_ylims) +
-  interior_annotation("a", sc_ltr_pos)
+  interior_annotation("a", sc_ltr_pos, cex = 2)
 
@@
   xlim(sc_xlims) + ylim(sc_ylims) +
-  interior_annotation("b", sc_ltr_pos)
+  interior_annotation("b", sc_ltr_pos, cex = 2)
 
@@
   xlim(sc_xlims) + ylim(sc_ylims) +
-  interior_annotation("c", sc_ltr_pos) 
+  interior_annotation("c", sc_ltr_pos, cex = 2) 
 ```
@@
   xlim(sc_xlims) + ylim(sc_ylims) +
-  interior_annotation("d", sc_ltr_pos)
+  interior_annotation("d", sc_ltr_pos, cex = 2)
 
@@
   scale_color_manual(values=c("#999999",'#ff7f00')) +
-  interior_annotation("a", sc_ltr_pos_mnist) +
+  interior_annotation("a", sc_ltr_pos_mnist, cex = 2) +
   theme(legend.position = "none")
@@
   scale_color_manual(values=c('#d9d9d9', '#ff7f00')) +
-  interior_annotation("a", sc_ltr_pos_mnist) +
+  interior_annotation("a", sc_ltr_pos_mnist, cex = 2) +
   theme(legend.position = "none")
@@
   ylim(c(-1.1, 0.9)) +
-  interior_annotation("b") +
+  interior_annotation("b", cex = 2) +
   theme(
@@
   ylim(c(-1, 1)) +
-  interior_annotation("c") +
+  interior_annotation("c", cex = 2) +
   theme(
@@
   ylim(c(-1, 0.6)) +
-  interior_annotation("d") +
+  interior_annotation("d", cex = 2) +
   theme(

### 2025-03-10 paper.qmd
Updated Figure 5

@@
   geom_point(alpha=0.5, colour="#636363", size = 0.5) +
-  interior_annotation("a", sc_ltr_pos)
+  interior_annotation("a", sc_ltr_pos, cex = 2)
 
@@
   #   ) +
-  interior_annotation("b", sc_ltr_pos)
+  interior_annotation("b", sc_ltr_pos, cex = 2)
 ```
@@
   ylim(c(-0.6, 0.6)) +
-  interior_annotation("c", sc_ltr_pos) +
+  interior_annotation("c", sc_ltr_pos, cex = 2) +
   theme(
@@
   ylim(c(-0.6, 0.6)) +
-  interior_annotation("d", sc_ltr_pos) +
+  interior_annotation("d", sc_ltr_pos, cex = 2) +
   theme(

### 2025-03-10 paper.qmd
Updated Figure 6

@@
   theme(
-    aspect.ratio = 1
+    aspect.ratio = 1,
+    legend.position = "none"
   )

### 2025-03-10 paper.qmd
Updated Figure 7

@@
   xlim(sc_xlims) + ylim(sc_ylims) +
-  interior_annotation("a", sc_ltr_pos, cex = 2)
+  interior_annotation("a", sc_ltr_pos, cex = 1.5)
 
@@
   geom_point(alpha=0.5, colour="#636363", size = 0.5) +
-  interior_annotation("a", sc_ltr_pos, cex = 2)
+  interior_annotation("a", sc_ltr_pos, cex = 1.5)
 
@@
   xlim(sc_xlims) + ylim(sc_ylims) +
-  interior_annotation("b", sc_ltr_pos, cex = 2)
+  interior_annotation("b", sc_ltr_pos, cex = 1.5)
 
@@
     na.value = "#ffffff", option = "E") +
-  interior_annotation("b", sc_ltr_pos) + 
+  interior_annotation("b", sc_ltr_pos, cex = 1.5) + 
   xlim(sc_xlims) + ylim(sc_ylims)
@@
   xlim(sc_xlims) + ylim(sc_ylims) +
-  interior_annotation("c", sc_ltr_pos, cex = 2) 
+  interior_annotation("c", sc_ltr_pos, cex = 1.5) 
 ```
@@
   xlim(sc_xlims) + ylim(sc_ylims) +
-  interior_annotation("d", sc_ltr_pos, cex = 2)
+  interior_annotation("d", sc_ltr_pos, cex = 1.5)
 
@@
   #   ) +
-  interior_annotation("b", sc_ltr_pos, cex = 2)
+  interior_annotation("b", sc_ltr_pos, cex = 1.5)
 ```
@@
   ylim(c(-0.6, 0.6)) +
-  interior_annotation("c", sc_ltr_pos, cex = 2) +
+  interior_annotation("c", sc_ltr_pos, cex = 1.5) +
   theme(
@@
   ylim(c(-0.6, 0.6)) +
-  interior_annotation("d", sc_ltr_pos, cex = 2) +
+  interior_annotation("d", sc_ltr_pos, cex = 1.5) +
   theme(
@@
   ylim(c(-0.6, 0.6)) +
-  interior_annotation("a", c(0.08, 0.93), cex = 2) +
+  interior_annotation("a", c(0.08, 0.93), cex = 1.5) +
   theme(
@@
   ylim(c(-0.42, 0.42)) +
-  interior_annotation("b", cex = 2) +
+  interior_annotation("b", cex = 1.5) +
   theme(
@@
   xlim(sc_xlims) + ylim(sc_ylims) +
-  interior_annotation("a", sc_ltr_pos)
+  interior_annotation("a", sc_ltr_pos, cex = 1.5)
 
@@
   xlim(sc_xlims) + ylim(sc_ylims) +
-  interior_annotation("c", sc_ltr_pos)
+  interior_annotation("c", sc_ltr_pos, cex = 1.5)
 ```
@@
   geom_point(alpha = 0.5) +
-  interior_annotation("c", c(0.08, 0.93), cex = 2)
+  interior_annotation("c", c(0.08, 0.93), cex = 1.5)
 ```
@@
         panel.border = element_rect(fill = 'transparent'),
-        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+        plot.title = element_text(size = 15, hjust = 0.5, vjust = -0.5),
         axis.ticks.x = element_line(),
@@
   geom_point(alpha = 0.5) +
-  interior_annotation("b", c(0.08, 0.93), cex = 2)
+  interior_annotation("b", c(0.08, 0.93), cex = 1.5)
 ```
@@
   geom_line(alpha = 0.3) +
-  interior_annotation("d", c(0.08, 0.93), cex = 2)
+  interior_annotation("d", c(0.08, 0.93), cex = 1.5)
 ```
@@
   plot_layout(guides='collect', ncol = 3) &
-  theme(legend.position='none', plot.tag = element_text(size = 8))
+  theme(legend.position='none', plot.tag = element_text(size = 12))
 ```
@@
                       position = c(0.08, 0.9),
-                      cex = 2)
+                      cex = 1.5)
 
@@
                       position = c(0.08, 0.95),
-                      cex = 2) +
+                      cex = 1.5) +
   theme(
@@
                       position = c(0.08, 0.9),
-                      cex = 2)
+                      cex = 1.5)
 ```
@@
                       position = c(0.06, 0.95),
-                      cex = 2) +
+                      cex = 1.5) +
   theme(
@@
   )  +
-  interior_annotation("a1", cex = 2) +
+  interior_annotation("a1", cex = 1.5) +
   theme(
@@
   ylim(c(-0.65, 0.65)) +
-  interior_annotation("a2", cex = 2)
+  interior_annotation("a2", cex = 1.5)
 
@@
   ylim(c(-1, 0.75)) +
-  interior_annotation("a3", cex = 2)
+  interior_annotation("a3", cex = 1.5)
 
@@
 #   ylim(c(-0.65, 0.65)) +
-#   interior_annotation("a3", cex = 2)
+#   interior_annotation("a3", cex = 1.5)
 
@@
   )  +
-  interior_annotation("b1", cex = 2) +
+  interior_annotation("b1", cex = 1.5) +
   theme(
@@
   ylim(c(-0.65, 0.65)) +
-  interior_annotation("b2", cex = 2)
+  interior_annotation("b2", cex = 1.5)
 
@@
   ylim(c(-1, 0.75)) +
-  interior_annotation("b3", cex = 2)
+  interior_annotation("b3", cex = 1.5)
 
@@
   scale_color_manual(values=c("#999999",'#ff7f00')) +
-  interior_annotation("a", sc_ltr_pos_mnist, cex = 2) +
+  interior_annotation("a", sc_ltr_pos_mnist, cex = 1.5) +
   theme(legend.position = "none")
@@
   scale_color_manual(values=c('#d9d9d9', '#ff7f00')) +
-  interior_annotation("a", sc_ltr_pos_mnist, cex = 2) +
+  interior_annotation("a", sc_ltr_pos_mnist, cex = 1.5) +
   theme(legend.position = "none")
@@
   ylim(c(-1.1, 0.9)) +
-  interior_annotation("b", cex = 2) +
+  interior_annotation("b", cex = 1.5) +
   theme(
@@
   ylim(c(-1, 1)) +
-  interior_annotation("c", cex = 2) +
+  interior_annotation("c", cex = 1.5) +
   theme(
@@
   ylim(c(-1, 0.6)) +
-  interior_annotation("d", cex = 2) +
+  interior_annotation("d", cex = 1.5) +
   theme(

### 2025-03-10 paper.qmd
Updated Figure 18

@@
   geom_rect(aes(xmin = 0.7, xmax = 0.8, ymin = 0.25, ymax = 0.4),
-               fill = "transparent", color = "#984ea3", linewidth = 0.7) 
+               fill = "transparent", color = "#984ea3", linewidth = 0.7) +
+  theme(legend.position = "none")
   # scale_colour_brewer(

### 2025-03-10 paper.qmd
Updated Figure 13

@@
   )  +
-  interior_annotation("a1", cex = 1.5) +
+  interior_annotation("a1", cex = 2) +
   theme(
@@
   ylim(c(-0.65, 0.65)) +
-  interior_annotation("a2", cex = 1.5)
+  interior_annotation("a2", cex = 2)
 
@@
   ylim(c(-1, 0.75)) +
-  interior_annotation("a3", cex = 1.5)
+  interior_annotation("a3", cex = 2)
 
@@
   )  +
-  interior_annotation("b1", cex = 1.5) +
+  interior_annotation("b1", cex = 2) +
   theme(
@@
   ylim(c(-0.65, 0.65)) +
-  interior_annotation("b2", cex = 1.5)
+  interior_annotation("b2", cex = 2)
 
@@
   ylim(c(-1, 0.75)) +
-  interior_annotation("b3", cex = 1.5)
+  interior_annotation("b3", cex = 2)
 

### 2025-03-10 paper.qmd
Updated Figure 16 and 17

@@
   scale_color_manual(values=c("#999999",'#ff7f00')) +
-  interior_annotation("a", sc_ltr_pos_mnist, cex = 1.5) +
+  interior_annotation("a", c(0.08, 0.93), cex = 1.5) +
   theme(legend.position = "none")
@@
   scale_color_manual(values=c('#d9d9d9', '#ff7f00')) +
-  interior_annotation("a", sc_ltr_pos_mnist, cex = 1.5) +
+  interior_annotation("a", c(0.08, 0.93), cex = 2) +
   theme(legend.position = "none")
@@
   ylim(c(-1.1, 0.9)) +
-  interior_annotation("b", cex = 1.5) +
+  interior_annotation("b", cex = 2) +
   theme(
@@
   ylim(c(-1, 1)) +
-  interior_annotation("c", cex = 1.5) +
+  interior_annotation("c", cex = 2) +
   theme(

### 2025-03-10 paper.qmd
Updated Figure 11

@@
                       position = c(0.08, 0.95),
-                      cex = 1.5) +
+                      cex = 2) +
   theme(
@@
                       position = c(0.08, 0.9),
-                      cex = 1.5)
+                      cex = 2)
 ```
@@
   interior_annotation("b1", 
-                      position = c(0.06, 0.95),
-                      cex = 1.5) +
+                      position = c(0.95, 0.95),
+                      cex = 2) +
   theme(
@@
                       position = c(0.08, 0.9),
-                      cex = 1.5)
+                      cex = 2)
 

### 2025-03-10 paper.qmd
Added histograms for errors
Update Figure 12

@@
 Most large errors are distributed near the sparse end of the non-linear cluster with dense corner."
-#| out-width: 60%
+#| out-width: 100%
 #| fig-pos: H
@@
 #| label: fig-one-dens_clust-error
-#| fig-height: 5
-#| fig-width: 15
 #| fig-pos: H
-#| out-height: 50%
+#| out-height: 100%
 #| fig-cap: "The sparse end of the structure shows a high model error, while the dense end shows a low model error.

### 2025-03-10 paper.qmd
Removed MNIST tSNE plot code which is unnecessary

@@
 tsne_minst_scaled <- mnist_scaled_obj$scaled_nldr 
-
-tsne_plot_mnist <- tsne_minst_scaled |> 
-  ggplot(aes(x = tSNE1, 
-             y = tSNE2)) + 
-  geom_point(alpha=0.1, color = "#000000")
-
-# read the png file from device
-left_top_img <- readPNG("figures/mnist/mnist_right_top_img.png", native = TRUE)
-center_img <- readPNG("figures/mnist/mnist_middle_img.png", native = TRUE)
-right_bottom_img <- readPNG("figures/mnist/mnist_right_bottom_img.png", native = TRUE)
-
-tsne_plot_mnist <- tsne_plot_mnist +
-    inset_element(p = left_top_img,
-                  left = 1,
-                  bottom = 0.85,
-                  right = 0,
-                  top = 0.95) +
-    inset_element(p = center_img,
-                  left = 0.8,
-                  bottom = 0.5,
-                  right = 0,
-                  top = 0.6) +
-    inset_element(p = right_bottom_img,
-                  left = 1,
-                  bottom = 0.2,
-                  right = 0,
-                  top = 0.3)
 ```
@@
 <!--PaCMAP param: n_components=2, n_neighbors=10, init=random, MN_ratio=0.9, FP_ratio=2.0-->
-```{r}
-#| fig-cap: "The $\\gD{}$ layout from tSNE applied for the MNIST digit 1 dataset with defualt (hyper-)parameters.
-We use our model in the $10\\text{-}D$ space to assess whether this is an accurate representation of data structure present in $10\\text{-}D$, or if it is misleading.
-The angle of the digit 1 images varies along this structure.
-Images at the top-right of the $\\gD{}$ layout show the digit 1 angled more to the right, while those at the bottom-right show the digit 1 angled more to the left."
-#| label: fig-tsne-best
-#| fig-pos: H
-#| out-width: 22%
-#| eval: false
-
-tsne_plot_mnist
-```
 

### 2025-03-10 paper.qmd
Removed MNIST third projection

@@
 distance_df_small_edges_mnist <- distance_mnist |>
-  filter(distance < benchmark_mnist) #|>
-  #filter(!(row_number() %in% c(155, 554, 140)))
+  filter(distance < benchmark_mnist)
 
@@
 
-# trimesh_removed_mnist <- vis_rmlg_mesh(
-#   distance_edges = distance_mnist, 
-#   benchmark_value = benchmark_mnist, 
-#   tr_coord_df = tr_from_to_df_mnist, 
-#   distance_col = "distance") +
-#   geom_point(data = pacmap_minst_scaled,
-#              aes(
-#                x = PaCMAP1,
-#                y = PaCMAP2
-#              ),
-#              alpha=0.5, 
-#              size = 0.1) +
-#   interior_annotation("a", sc_ltr_pos_mnist)
-
 ## Compute error
@@
   col_start = "PC")
-
-## Categorize error
-
-# error_df <- error_df |>
-#   mutate(type = case_when(
-#     row_wise_abs_error <= 2 ~ "error 0-2",
-#     row_wise_abs_error <= 4 ~ "error 2-4",
-#     row_wise_abs_error <= 6 ~ "error 4-6",
-#     row_wise_abs_error <= 8 ~ "error 6-8",
-#     row_wise_abs_error <= 10 ~ "error 8-10",
-#     .default = "error greter than 10"
-#   )) |>
-#   mutate(type = factor(type, levels = c(
-#     "error 0-2", "error 2-4", "error 4-6", "error 6-8", "error 8-10", 
-#     "error greter than 10")))
 
@@
   theme(legend.position = "none")
-  # scale_colour_brewer(
-  #   type = "seq",
-  #   palette = 1
-  # ) #+ 
-  # theme(
-  #     legend.position = "bottom",
-  #     legend.text = element_text(size = 10),          # Increase legend text size
-  #     legend.key.size = unit(1, "lines")              # Increase size of legend keys
-  # ) +
-  # guides(
-  #     color = guide_legend(
-  #         override.aes = list(size = 2.5, alpha = 1)      # Increase point size and opacity in legend
-  #     )
-  # ) 
 ```
@@
   filter(type == "data") |>
-  # bind_cols(tsne_minst_scaled_temp) |>
-  # select(-type, -tSNE1, -tSNE2, -ID) |>
   select(-type)
-
-# scaled_mnist_data <- bind_cols(scaled_mnist_data,
-#                                tsne_minst_scaled |> select(-ID))
 
@@
   select(-type)
-
-# model_2d <- df_bin_centroids_mnist |>
-#   select(c_x, c_y) |>
-#   rename(c("tSNE1" = "c_x",
-#            "tSNE2" = "c_y"))
-# 
-# scaled_mnist_data_model <- bind_cols(scaled_mnist_data_model, model_2d)
-
-# Combine with the true model for visualization
-# df <- dplyr::bind_rows(scaled_mnist_data_model |> mutate(type = "model"),
-#                        scaled_mnist_data)
-
-# scaled_mnist_data <- scaled_mnist_data |>
-#   select(-type)
-
-## First projection
-# projection <- cbind(
-#     c(0.71503,0.00502,-0.04370,-0.01811,-0.01103,0.09877,0.06199,0.05396,0.07771,-0.01713),
-#     c(0.00587,0.71879,0.03485,0.08284,-0.07534,-0.05420,0.04315,-0.00547,-0.02995,0.01810))
 
@@
   )
-
-## Third projection
-projection <- cbind(
-    c(-0.12783,-0.50063,-0.02631,-0.00484,0.17733,0.37157,0.11310,0.09424,0.19844,0.19792),
-    c(-0.42542,0.29570,0.31106,-0.30307,0.00614,0.16714,0.21779,-0.00025,0.07020,-0.00689))
-
-projection_scaled <- projection * 1.5
-
-projected <- as.matrix(scaled_mnist_data) %*% projection_scaled
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  #dplyr::mutate(type = df_exe$type) |>
-  dplyr::mutate(ID = dplyr::row_number()) |>
-  mutate(cluster = tsne_minst_scaled_temp$cluster)
-
-projected_model <- as.matrix(scaled_mnist_data_model) %*% projection_scaled
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  distance_df_small_edges_mnist |> select(-distance), 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-axes_obj <- gen_axes(  
-  proj = projection,
-  limits = 1,
-  axis_pos_x = -0.83,
-  axis_pos_y = -0.83,
-  axis_labels = names(scaled_mnist_data),
-  threshold = 0.06)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
-
-mnist_proj_tsne_model3 <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2,
-      colour = cluster)) +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#000000") +
-  geom_point(
-    #size = 0.2,
-    alpha = 0.3) +
-  scale_color_manual(values = c("#999999",'#ff7f00')) +
-  geom_segment(
-    data=axes, 
-    aes(x=x1, y=y1, xend=x2, yend=y2), 
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 2) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-1, 0.6)) +
-  ylim(c(-1, 0.6)) +
-  interior_annotation("d", cex = 1.5) +
-  theme(
-    legend.position = "none"
-  )
-
 ```
@@
 free(trimesh_removed_mnist) + mnist_proj_tsne_model1 +
-  #mnist_proj_tsne_model2 + mnist_proj_tsne_model3 +
   mnist_proj_tsne_model2 +

### 2025-03-10 paper.qmd
Updated the code

@@
 #| echo: false
-# remotes::install_github("JayaniLakshika/quollr")
 library(quollr)
@@
 library(dplyr)
-# remotes::install_github("jlmelville/snedata")
-# library(snedata)
-# library(ggflowchart)
-#library(purrr) ## map function
-#library(gridExtra) ## for grid.arrange
-# library(rsample)
-# library(DT)
-#library(ggbeeswarm)
 library(ggplot2)
@@
 library(tidyr)
-
-# library(Rtsne)
-# library(uwot)
-# library(phateR)
 library(patchwork)
 library(png)
-#library(langevitour)
 library(colorspace)
 library(kableExtra)
-#library(grid)
 library(conflicted)
+
 conflicts_prefer(dplyr::filter)
+conflicts_prefer(dplyr::select)
 ```
@@
 )
+
 interior_annotation <- function(label, position = c(0.92, 0.92), cex = 1, col="grey70") {
@@
 sc_ltr_pos <- c(0.08, 0.9)
-# sc_xlims <- c(-0.5, 1.47)
-# sc_ylims <- c(-0.32, 2.1)
 sc_xlims <- c(-0.3, 1.2)
@@
 
-nldr_two_curvy <- tsne_two_curvy_scaled |> 
-  ggplot(aes(x = tSNE1, y = tSNE2)) + 
-  geom_point(alpha=0.5, colour="#636363", size = 0.5) +
-  xlim(sc_xlims) + ylim(sc_ylims) +
+nldr_two_curvy_original <- ggplot(tsne_two_curvy_scaled, aes(x = tSNE1, y = tSNE2)) +
+  geom_point(alpha = 0.5, color = "#636363", size = 0.5)
+
+# First plot with xlim and ylim
+nldr_two_curvy <- nldr_two_curvy_original+
+  interior_annotation("a", sc_ltr_pos, cex = 1.5) +
+  xlim(sc_xlims) +
+  ylim(sc_ylims)
+
+# Second plot (identical to base except for xlim/ylim)
+nldr_two_curvy2 <- nldr_two_curvy_original +
   interior_annotation("a", sc_ltr_pos, cex = 1.5)
 
-nldr_two_curvy2 <- tsne_two_curvy_scaled |> 
-  ggplot(aes(x = tSNE1, y = tSNE2)) + 
-  geom_point(alpha=0.5, colour="#636363", size = 0.5) +
-  interior_annotation("a", sc_ltr_pos, cex = 1.5)
-
-nldr_two_curvy_original <- tsne_two_curvy_scaled |> 
-  ggplot(aes(x = UMAP1, y = UMAP2)) + 
-  geom_point(alpha=0.5, color = "#636363", size = 0.5) +
+# Third plot with aspect ratio and different annotation
+nldr_two_curvy_original2 <- nldr_two_curvy_original +
   theme(aspect.ratio = 1) +
-  #xlim(c(-0.01, 1.1)) + ylim(c(-0.08, 0.57)) +
-  interior_annotation("a", sc_ltr_pos)
-
-nldr_two_curvy_original2 <- tsne_two_curvy_scaled |> 
-  ggplot(aes(x = tSNE1, y = tSNE2)) + 
-  geom_point(alpha=0.5, color = "#636363", size = 0.5) +
-  theme(aspect.ratio = 1) +
-  #xlim(c(-0.01, 1.1)) + ylim(c(-0.08, 0.57)) +
   interior_annotation("a1", sc_ltr_pos)

### 2025-03-10 paper.qmd
Removed unecessary code

@@
   xlim(sc_xlims) + ylim(sc_ylims) 
-
 ```
@@
     filter(distance < benchmark_two_curvy)
-
-# trimesh_removed_two_curvy <- vis_rmlg_mesh(
-#   distance_edges = distance_two_curvy, 
-#   benchmark_value = benchmark_two_curvy, 
-#   tr_coord_df = tr_from_to_df_two_curvy, 
-#   distance_col = "distance") +
-#   xlim(sc_xlims) + ylim(sc_ylims) +
-#   interior_annotation("d", sc_ltr_pos)
 
@@
 ```
-
-<!--langevitour with true model for S-curve--> 
-<!-- ::: {#fig-two_curvy-true-sc layout-ncol="3" fig-pos="H"} -->
-
-<!-- ![](figures/two_curvy/sc_true_1.png){width="150" fig-align="center"} -->
-
-<!-- ![](figures/two_curvy/sc_true_2.png){width="150" fig-align="center"} -->
-
-<!-- ![](figures/two_curvy/sc_true_3.png){width="150" fig-align="center"} -->
-
-<!-- Three views of the true model (grey points and lines) in $2\text{-}D$ projections from $7\text{-}D$, for the S-curve data (purple points).
-The data is spread along the S shape, and does not vary much from this curve.
-(The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/I5GL23vLiw0>).
--->
-<!-- ::: -->
-
-<!-- XXX The points on the "theoretical model" should be equidistant along the S. They are still to bunched in some spots and too far apart in others. -->
 
@@
 umap_pbmc_scaled <- pbmc_scaled_obj$scaled_nldr
-
-# umap_pbmc <- umap_pbmc_scaled |>
-#   ggplot(aes(x = UMAP1,
-#              y = UMAP2)) +
-#   geom_point(alpha=0.3, color = "#000000") 
 ```
@@
 
-# trimesh_removed_pbmc <- vis_rmlg_mesh(
-#   distance_edges = distance_pbmc,
-#   benchmark_value = benchmark_pbmc,
-#   tr_coord_df = tr_from_to_df_pbmc,
-#   distance_col = "distance") +
-#   geom_point(
-#     data = umap_pbmc_scaled,
-#     aes(
-#       x = UMAP1,
-#       y = UMAP2
-#     ),
-#     alpha = 0.3,
-#     size = 0.1,
-#     color = "#000000"
-#   )
-
 trimesh_removed_pbmc <- ggplot() + 
@@
   )
-
-# ## Compute error
-# error_df <- augment(
-#   df_bin_centroids = df_bin_centroids_pbmc,
-#   df_bin = df_bin_pbmc,
-#   training_data = training_data_pbmc,
-#   newdata = NULL,
-#   type_NLDR = "UMAP",
-#   col_start = "PC_")
-
-# ## Categorize error
-# 
-# error_df <- error_df |>
-#   mutate(type = case_when(
-#     row_wise_abs_error <= 5 ~ "error 0-5",
-#     row_wise_abs_error <= 10 ~ "error 5-10",
-#     row_wise_abs_error <= 15 ~ "error 10-15",
-#     row_wise_abs_error <= 20 ~ "error 15-20",
-#     .default = "error greter than 20"
-#   )) |>
-#   mutate(type = factor(type, levels = c(
-#     "error 0-5", "error 5-10", "error 10-15", "error 15-20", "error greter than 20")))
-# 
-# ## To join embedding
-# error_df <- error_df |>
-#   bind_cols(umap_pbmc_scaled |>
-#               select(-ID))
-# 
-# error_plot_pbmc <- error_df |>
-#   ggplot(aes(x = UMAP1,
-#              y = UMAP2,
-#              color = type,
-#              group = ID)) +
-#   geom_point(alpha=0.5,
-#              size = 0.1) +
-#   interior_annotation("b", sc_ltr_pos)
-
 ```
@@
 projection_scaled <- projection * 1.23
-
-# projection_filtered <- projection
-# projection_filtered[apply(projection_filtered, 1, function(row) all(row < 0)), ] <- 0.75
-
-#rows_to_keep <- c(1, 2, 4, 5, 6)
-
-#projection_filtered <- projection[rows_to_keep, ]
 
@@
 
-# pbmc_proj_umap_model2 <- ggplot() +
-#   geom_segment(
-#     data = model_df, 
-#     aes(
-#       x = proj1_from, 
-#       y = proj2_from, 
-#       xend = proj1_to, 
-#       yend = proj2_to), 
-#     color = "#000000") +
-#   geom_segment(
-#     data=axes, 
-#     aes(x=x1, y=y1, xend=x2, yend=y2), 
-#     colour="grey70") +
-#   geom_text(
-#     data=axes,
-#     aes(x=x2, y=y2), 
-#     label=rownames(axes),
-#     colour="grey50",
-#     size = 3) +
-#   geom_path(
-#     data=circle, 
-#     aes(x=c1, y=c2), colour="grey70") +
-#   coord_fixed() +
-#   xlim(c(-0.65, 0.65)) +
-#   ylim(c(-0.65, 0.65)) +
-#   interior_annotation("a3", cex = 1.5)
-
 ```
@@
 #| label: error-comp-pbmc
-# read the png file from device
-
-#pbmc_layouts <- readPNG("figures/pbmc3k/pbmc_nldr_layouts.png", native = TRUE)
 
@@
   geom_line(linewidth = 0.3) + 
-  # geom_vline(xintercept = 15, linetype="solid", 
-  #            color = "black", linewidth=0.8, alpha = 0.5) +
   scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 13, 17, 21, 26)]) +
@@
 distance_df_small_edges_pbmc <- distance_pbmc |>
-  filter(distance < benchmark_pbmc) #|>
-  #filter(!(row_number() %in% c(6, 12, 14, 16, 17, 19, 23, 29, 30, 143, 144, 146, 147, 157)))
+  filter(distance < benchmark_pbmc)
 
@@
   by = c("from", "to"))
-
-# tr_from_to_df_pbmc <- tr_from_to_df_pbmc |>
-#   filter(!(row_number() %in% c(155)))
 
@@
   )
-
-# trimesh_removed_pbmc <- vis_rmlg_mesh(
-#   distance_edges = distance_pbmc,
-#   benchmark_value = benchmark_pbmc,
-#   tr_coord_df = tr_from_to_df_pbmc,
-#   distance_col = "distance") +
-#   geom_point(
-#     data = tsne_pbmc_scaled,
-#     aes(
-#       x = tSNE1,
-#       y = tSNE2
-#     ),
-#     alpha = 0.2,
-#     color = "#000000"
-#   ) 
-#   #xlim(sc_xlims) + ylim(sc_ylims) +
-#   #interior_annotation("a", sc_ltr_pos)
 
@@
                        scaled_pbmc_data |> mutate(type = "data"))
-
-## Removed long edge in high-d
-# distance_df_small_edges_pbmc <- distance_df_small_edges_pbmc |>
-#   filter(!(row_number() %in% c(155)))
 

### 2025-03-10 paper.qmd
Removed unncessary code

@@
 
-img_right_top <- c(868, 1465, 1787)
-
-pixels_gathered_within <-  mnist_data |>
+mnist_data <- mnist_data |>
   mutate(instance = row_number()) |>
@@
   extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
-  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
+  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28)
+
+img_right_top <- c(868, 1465, 1787)
+
+pixels_gathered_within <-  mnist_data |>
   filter(instance %in% img_right_top)
@@
 pixels_gathered_within <-  mnist_data |>
-  mutate(instance = row_number()) |>
-  gather(pixel, value, -Label, -instance) |>
-  extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
-  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
   filter(instance %in% img_middle)
@@
 pixels_gathered_within <-  mnist_data |>
-  mutate(instance = row_number()) |>
-  gather(pixel, value, -Label, -instance) |>
-  extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
-  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
   filter(instance %in% img_right_bottom)
@@
 pixels_gathered_outside <-  mnist_data |>
-  mutate(instance = row_number()) |>
-  gather(pixel, value, -Label, -instance) |>
-  extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
-  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
   filter(instance %in% img_error_outside)
@@
 ##### inside
-img_error_inside <- c(3179, 7363, 7650, 5241, 5173, 6981, 7228, 6673, 4636) #7228, 7363, 7650
+img_error_inside <- c(3179, 7363, 7650, 5241, 5173, 6981, 7228, 6673, 4636)
 
 pixels_gathered_inside <-  mnist_data |>
-  mutate(instance = row_number()) |>
-  gather(pixel, value, -Label, -instance) |>
-  extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
-  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
   filter(instance %in% img_error_inside)

### 2025-03-10 paper.qmd
optimize the code of Figure 18

@@
 
+```
+
+```{r}
+#| label: plot-digit
+
+plot_digit_img <- function(digit_df, palette, title_text, hjust, ncol = 4) {
+
+  ggplot(data = digit_df, aes(x, y, fill = value)) +
+  geom_tile() +
+  facet_wrap(~ instance, ncol = ncol) +
+  ggtitle(title_text) +
+  coord_fixed() +
+  scale_fill_continuous_sequential(palette = palette) +
+  theme(strip.background = element_blank(),
+        strip.text.x = element_blank(),
+        legend.position = "none",
+        plot.title.position = "plot",
+        plot.title = element_text(hjust = hjust,
+                                  size = 12))
+
+}
 ```
@@
 
-right_top_img <- pixels_gathered_within |>
-  ggplot(aes(x, y, fill = value)) +
-  geom_tile() +
-  facet_wrap(~ instance, ncol = 4) +
-  ggtitle("(a)") +
-  coord_fixed() +
-  scale_fill_continuous_sequential(palette = "Reds 2") +
-  theme(strip.background = element_blank(),
-        strip.text.x = element_blank(),
-        legend.position = "none",
-        plot.title.position = "plot",
-        plot.title = element_text(hjust = 0.31,
-                                  size = 12))
-
-#img_middle <- c(1069, 4604, 3579)
+right_top_img <- plot_digit_img(pixels_gathered_within, palette = "Reds 2", title_text = "(a)", hjust = 0.31)
+
 img_middle <- c(130, 2091, 7795)
@@
 
-middle_img <- pixels_gathered_within |>
-  ggplot(aes(x, y, fill = value)) +
-  geom_tile() +
-  facet_wrap(~ instance, ncol = 4) +
-  ggtitle("(b)") +
-  coord_fixed() +
-  scale_fill_continuous_sequential(palette = "Greens 3") +
-  theme(strip.background = element_blank(),
-        strip.text.x = element_blank(),
-        legend.position = "none",
-        plot.title.position = "plot",
-        plot.title = element_text(hjust = 0.31,
-                                  size = 12))
-
-#img_right_bottom <- c(1695, 5338, 2805)
+middle_img <- plot_digit_img(pixels_gathered_within, palette = "Greens 3", title_text = "(b)", hjust = 0.31)
+
 img_right_bottom <- c(2035, 7708, 7710)
@@
 
-right_bottom_img <- pixels_gathered_within |>
-  ggplot(aes(x, y, fill = value)) +
-  geom_tile() +
-  facet_wrap(~ instance, ncol = 4) +
-  ggtitle("(c)") +
-  coord_fixed() +
-  scale_fill_continuous_sequential(palette = "Purples 2") +
-  theme(strip.background = element_blank(),
-        strip.text.x = element_blank(),
-        legend.position = "none",
-        plot.title.position = "plot",
-        plot.title = element_text(hjust = 0.31,
-                                  size = 12))
-
-#img_error_outside <- c(3554, 501, 4422, 6860, 3869, 1529, 4916, 6144)
-#img_error_outside <- c(501, 1529, 6144)
+right_bottom_img <- plot_digit_img(pixels_gathered_within, palette = "Purples 2", title_text = "(c)", hjust = 0.31)
+
 img_error_outside <- c(999, 3650, 4916)
@@
 
-imge_error_sample_outside <- pixels_gathered_outside |>
-  ggplot(aes(x, y, fill = value)) +
-  geom_tile() +
-  facet_wrap(~ instance, ncol = 4) +
-  ggtitle("(d)") +
-  coord_fixed() +
-  scale_fill_continuous_sequential(palette = "Grays") +
-  theme(strip.background = element_blank(),
-        strip.text.x = element_blank(),
-        legend.position = "none",
-        plot.title.position = "plot",
-        plot.title = element_text(hjust = 0.31,
-                                  size = 12))
+imge_error_sample_outside <- plot_digit_img(pixels_gathered_within, palette = "Grays", title_text = "(d)", hjust = 0.31)
 
@@
 
-imge_error_sample_inside <- pixels_gathered_inside |>
-  ggplot(aes(x, y, fill = value)) +
-  geom_tile() +
-  facet_wrap(~ instance, ncol = 9) +
-  ggtitle("(e)") +
-  coord_fixed() +
-  scale_fill_continuous_sequential(palette = "Grays") +
-  theme(strip.background = element_blank(),
-        strip.text.x = element_blank(),
-        legend.position = "none",
-        plot.title.position = "plot",
-        plot.title = element_text(hjust = 0,
-                                  size = 12))
+imge_error_sample_inside <- plot_digit_img(pixels_gathered_inside, palette = "Grays", title_text = "(e)", hjust = 0, ncol = 9)
 

### 2025-03-10 paper.qmd
Updated Figure 16 code

@@
   geom_line(linewidth = 0.3) + 
-  # geom_vline(xintercept = 15, linetype="solid", 
-  #            color = "black", linewidth=0.8, alpha = 0.5) +
   scale_x_continuous(breaks = sort(unique(error_mnist$a1))[append(seq(1, length(unique(error_mnist$a1)), by = 5), 24)]) +
@@
   distance_col = "distance")
-
-#benchmark_mnist <- 0.1
 

### 2025-03-10 paper.qmd
Updated the some code with new functions

@@
 
-#umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_15_min_dist_0.99.rds")
-
 tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_5.rds")
@@
   mutate(ID = row_number())
+
 lim1 <- two_curvy_scaled_obj$lim1
@@
   filter(drop_empty == FALSE)
-
-boxplot2 <- ggplot(df_bin_centroids_two_curvy, 
-       aes(x=bin_counts)) + 
-  geom_histogram()
 
@@
              size = 0.5) +
-  #coord_equal() +
-  # theme(
-  #   aspect.ratio = 1
-  #   ) +
   interior_annotation("b", sc_ltr_pos, cex = 1.5)
@@
 df_bin_two_curvy_temp <- df_bin_two_curvy
-
-### Compute highD distance
-dist_vec <- proxy::dist(x = df_bin_two_curvy_temp[, -1], method = "Euclidean") |> as.vector()
-
-from_vec <- c()
-to_vec <- c()
-num_obs <- 1:(NROW(df_bin_two_curvy_temp) - 1)
-
-for (obs in num_obs) {
-
-  from_val <- rep(obs, (NROW(df_bin_two_curvy_temp) - obs))
-  if ((obs + 1) <= NROW(df_bin_two_curvy_temp)) {
-    to_val <- (obs + 1):NROW(df_bin_two_curvy_temp)
-  }
-  from_vec <- append(from_vec, from_val)
-  to_vec <- append(to_vec, to_val)
-
-}
-
-dist_highd <- tibble::tibble(from = from_vec, to = to_vec, dist_highd = dist_vec)
-
-## To plot the distribution of distances
-
-# Define bin width and starting point
-start_point <- bin_width/2
-
-# Create bins and calculate the mean value for each bin range
-bins <- seq(start_point, max(distance_two_curvy$distance) + bin_width, 
-            by = bin_width)
-bin_labels <- bins[-length(bins)] + bin_width / 2  # mean of each bin range
-
-distance_two_curvy_dist <- distance_two_curvy |> 
-  mutate(bin = cut(distance, breaks = bins, include.lowest = TRUE, labels = bin_labels))
-
-distance_two_curvy_dist <- left_join(distance_two_curvy_dist, dist_highd, by = c("from", "to"))
-
-distance_points <- ggplot(
-  distance_two_curvy_dist, 
-  aes(x = distance,
-      y = dist_highd)) +
-  geom_point(alpha = 0.5) +
-  geom_vline(xintercept = 2.5 * bin_width,
-             linetype="solid",
-             color = "#bdbdbd",
-             linewidth=1) +
-  geom_vline(xintercept = 2 * bin_width,
-             linetype="solid",
-             color = "#bdbdbd",
-             linewidth=1) +
-  ylab(expression(d^{(7)})) +
-  xlab(expression(d^{(2)})) +
-  theme_minimal() +
-  theme(aspect.ratio = 0.75,
-        panel.border = element_rect(fill = 'transparent'),
-        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
-        panel.grid.major.x = element_blank(),
-        axis.ticks.x = element_line(),
-        axis.ticks.y = element_line())
-
 ```
@@
 
-# hex_grid_coloured_two_curvy1 + 
-#   hex_grid_coloured_two_curvy2 + 
-#   hex_grid_coloured_two_curvy3 +
-#   plot_layout(guides='collect', ncol = 3,
-#               heights = c(2,1)) &
-#   theme(legend.position='none', plot.tag = element_text(size = 8))
-
 hex_grid_coloured_two_curvy2 +

### 2025-03-10 paper.qmd
Updated the code for MNIST

@@
 
-# tr_from_to_df_gau1 <- tr_from_to_df_gau1 |>
-#   filter(row_number() != 76) |>
-#   filter(row_number() != 34)
-
-# tr_from_to_df_gau1 <- tr_from_to_df_gau1 |>
-#   filter(row_number() != 149)
-
 ## Compute 2D distances
@@
 
-# benchmark_gau1 <- 0.1
-
-# trimesh_removed_gau1 <- vis_rmlg_mesh(
-#   distance_edges = distance_gau1,
-#   benchmark_value = benchmark_gau1,
-#   tr_coord_df = tr_from_to_df_gau1,
-#   distance_col = "distance")
-
 tr_df <- distinct(tibble(
@@
     alpha = 0.2) + 
-  # geom_rect(
-  #     aes(xmin = 0.3, 
-  #         xmax = 0.65, 
-  #         ymin = -0.05, 
-  #         ymax = 0.35),
-  #         fill = "transparent", 
-  #     color = "grey50", 
-  #     linewidth = 0.5) +
   interior_annotation("a1", 
@@
 
-# library(langevitour)
-
-# distance_df_small_edges <- distance_df_small_edges |>
-#   filter(row_number() != 108)
-
-## First projection
-# projection <- cbind(
-#     c(0.46477,-0.02024,0.56378,0.39736),
-#     c(0.15607,-0.54876,-0.44090,0.41505))
 projection <- cbind(
@@
     alpha = 0.05) + 
-  # geom_rect(
-  #   aes(xmin = 0.45, 
-  #       xmax = 0.57, 
-  #       ymin = 0, 
-  #       ymax = 0.1),
-  #   fill = "transparent", 
-  #   color = "grey50", 
-  #   linewidth = 0.5)  +
   interior_annotation("b1", 
@@
 
-## First projection
-# projection <- cbind(
-#   c(0.02743,0.21427,0.62893,-0.49816),
-#   c(-0.70709,0.42156,-0.02593,0.10965))
-# projection <- cbind(
-#   c(-0.00215,-0.68905,-0.04778,-0.54223),
-#   c(0.42558,-0.23854,-0.63659,0.35753))
-
 projection <- cbind(
@@
                   "proj2" = "...2")) |>
-  #dplyr::mutate(type = df_exe$type) |>
   dplyr::mutate(ID = dplyr::row_number())
@@
   geom_point(
-    #size = 0.5,
     alpha = 0.3,
@@
 
-# trimesh_removed_gau1 <- vis_rmlg_mesh(
-#   distance_edges = distance_gau1,
-#   benchmark_value = benchmark_gau1,
-#   tr_coord_df = tr_from_to_df_gau1,
-#   distance_col = "distance")
-
 tr_df <- distinct(tibble(
@@
     alpha = 0.3) +
-  # geom_rect(
-  #   aes(xmin = -0.04, 
-  #       xmax = 0.22, 
-  #       ymin = 0.08, 
-  #       ymax = 0.34),
-  #   fill = "transparent", 
-  #   color = "grey50", 
-  #   linewidth = 0.5) +
   interior_annotation("c1", 
@@
 trimesh_removed_gau_tsne + trimesh_removed_gau_umap + trimesh_removed_gau_pacmap +
-five_gau_proj_tsne_model1 + five_gau_proj_umap_model1 + five_gau_proj_pacmap_model1 + #error_df_gau_tsne + error_df_gau_umap + error_df_gau_pacmap +
+five_gau_proj_tsne_model1 + five_gau_proj_umap_model1 + five_gau_proj_pacmap_model1 + 
   plot_layout(guides = "collect", nrow = 2) &

### 2025-03-11 paper.qmd
Remove unnecessary code

@@
 
-# tsne_pbmc <- tsne_pbmc_scaled |>
-#   ggplot(aes(x = tSNE1,
-#              y = tSNE2)) +
-#   geom_point(alpha=0.3, color = "#000000")
 ```
@@
 
-# tr_from_to_df_pbmc <- tr_from_to_df_pbmc |>
-#   filter(!(row_number() %in% c(65, 112, 130)))
-
 ## Compute 2D distances
@@
   distance_col = "distance")
-
-#benchmark_pbmc <- 0.2
-
-## Remove an edge that exist as long edge in p-D
-# tr_from_to_df_pbmc |> mutate(ID = row_number()) |> filter(from == 37) |> filter(to == 43)
 

### 2025-03-12 paper.qmd
Updated Figure 12

@@
 #| label: fig-one-dens_clust-error
+#| fig-width: 15
+#| fig-height: 10
 #| fig-pos: H

### 2025-03-12 scripts/one_c_shaped_dens_structure/04_gen_error_plots.R

@@ -728,7 +728,8 @@ generate_error_plots_one_c_shaped <- function(){
 
   error_plot_one_curvy_hist + plot_tsne_dens + five_c_shaped_proj_tsne_model1 +
   error_plot_one_curvy_hist_selected + plot_tsne_dens_selected + five_c_shaped_proj_tsne_model1_selected +
-    plot_layout(guides = "collect", ncol = 3, widths = c(1, 1.5, 1.5)) &
+    #plot_layout(guides = "collect", ncol = 3, widths = c(1, 1, 1)) &
+    plot_layout(guides = "collect", nrow = 2) &
     theme(legend.position='none')
 
 }

### 2025-03-12 paper.qmd
Updated Figure 12 title

@@
 #| fig-cap: "The sparse end of the structure shows a high model error, while the dense end shows a low model error.
-The tSNE layout of the C-shaped structure, with dense points (a) are colored according to their $4\\text{-}D$ error.
+The tSNE layout of the C-shaped structure, with dense points (a1-a3) are colored according to their $4\\text{-}D$ model error.
 Darker colors represent a high error, while lighter colors indicate a low error.
-b and c present the results from linked plots that brushed the high $4\\text{-}D$ error points in \\gD{} and $4\\text{-}D$.
+a2, a3 and b2, b3 present the results from linked plots that brushed the high $4\\text{-}D$ error points in \\gD{} and $4\\text{-}D$.
 It helps in identifying the high $4\\text{-}D$ model errors that occur due to the sparse end.

### 2025-03-13 paper.qmd
Updated the text size of Figure 5

@@
         panel.border = element_rect(fill = 'transparent'),
-        plot.title = element_text(size = 15, hjust = 0.5, vjust = -0.5),
+        plot.title = element_text(size = 20, hjust = 0.5, vjust = -0.5),
         axis.ticks.x = element_line(),
-        axis.ticks.y = element_line())
+        axis.ticks.y = element_line(),
+        axis.title = element_text(size = 18))
 

### 2025-03-13 paper.qmd
Updated Figure 8

--- 2025-03-12 paper.qmd
@@
         panel.border = element_rect(fill = 'transparent'),
-        plot.title = element_text(size = 15, hjust = 0.5, vjust = -0.5),
+        plot.title = element_text(size = 20, hjust = 0.5, vjust = -0.5),
         axis.ticks.x = element_line(),
-        axis.ticks.y = element_line())
+        axis.ticks.y = element_line(),
+        axis.title = element_text(size = 15),
+        axis.text = element_text(size = 12))
 

### 2025-03-13 paper.qmd
Updated Figure8

@@
         panel.border = element_rect(fill = 'transparent'),
-        plot.title = element_text(size = 15, hjust = 0.5, vjust = -0.5),
+        plot.title = element_text(size = 20, hjust = 0.5, vjust = -0.5),
         axis.ticks.x = element_line(),
-        axis.ticks.y = element_line())
+        axis.ticks.y = element_line(),
+        axis.title = element_text(size = 15),
+axis.text = element_text(size = 12))
 
@@
         panel.border = element_rect(fill = 'transparent'),
-        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+        plot.title = element_text(size = 20, hjust = 0.5, vjust = -0.5),
         axis.ticks.x = element_line(),
         axis.ticks.y = element_line(),
-        legend.position = "none")
+        legend.position = "none",
+        axis.title = element_text(size = 15),
+axis.text = element_text(size = 12))
 ```

### 2025-03-13 paper.qmd
Updated Figure 12

@@
         plot.title = element_text(hjust = hjust,
-                                  size = 12))
+                                  size = 15))
 
@@
   theme(panel.border = element_rect(fill = 'transparent'),
-        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+        plot.title = element_text(size = 15, hjust = 0.5, vjust = -0.5),
         axis.ticks.x = element_line(),
@@
         legend.position = "none",
-        axis.text.x = element_text(size = 7),
-        axis.text.y = element_text(size = 7),
-        axis.title.x = element_text(size = 7),
-        axis.title.y = element_text(size = 7))
+        axis.text = element_text(size = 7),
+        axis.title = element_text(size = 12))
 
@@
 #| fig-pos: H
-#| out-width: 60%
+#| out-width: 100%
 

### 2025-03-17 paper.qmd
Added function to get 2D projection data

@@
 ```{r}
+#| label: plot-projection
+
+get_projection <- function(projection, proj_scale, scaled_data, 
+                           scaled_data_model, distance_df_small_edges,
+                           axis_param) {
+
+  projection_scaled <- projection * proj_scale
+  projected <- as.matrix(scaled_data) %*% projection_scaled
+
+  projected_df <- projected |>
+    tibble::as_tibble(.name_repair = "unique") |>
+    dplyr::rename(c("proj1" = "...1",
+                    "proj2" = "...2")) |>
+    dplyr::mutate(ID = dplyr::row_number())
+
+  projected_model <- as.matrix(scaled_data_model) %*% projection_scaled
+
+  projected_model_df <- projected_model |>
+    tibble::as_tibble(.name_repair = "unique") |>
+    dplyr::rename(c("proj1" = "...1",
+                    "proj2" = "...2")) |>
+    dplyr::mutate(ID = dplyr::row_number())
+
+  model_df <- dplyr::left_join(
+    distance_df_small_edges |> select(-distance), 
+    projected_model_df, 
+    by = c("from" = "ID"))
+
+  names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+  model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+  names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+  limits <- axis_param$limits
+  axis_pos_x <- axis_param$axis_pos_x
+  axis_pos_y <- axis_param$axis_pos_y
+  threshold <- axis_param$threshold
+
+  axes_obj <- gen_axes(  
+    proj = projection,
+    limits = limits,
+    axis_pos_x = axis_pos_x,
+    axis_pos_y = axis_pos_y,
+    axis_labels = names(scaled_data),
+    threshold = threshold)
+
+  axes <- axes_obj$axes
+  circle <- axes_obj$circle
+
+  return(list(projected_df = projected_df, 
+              model_df = model_df, 
+              axes = axes, 
+              circle = circle))
+
+}
+```
+
+```{r}
 #| label: import-scripts
@@
     colour="grey50",
-    size = 4) +
+    size = 5) +
   geom_path(
@@
 
-projection_scaled <- projection * 2.5
-
-projected <- as.matrix(scaled_mnist_data) %*% projection_scaled
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  #dplyr::mutate(type = df_exe$type) |>
-  dplyr::mutate(ID = dplyr::row_number()) |>
-  mutate(cluster = tsne_minst_scaled_temp$cluster)
-
-projected_model <- as.matrix(scaled_mnist_data_model) %*% projection_scaled
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  distance_df_small_edges_mnist |> select(-distance), 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-axes_obj <- gen_axes(  
-  proj = projection,
-  limits = 1,
+proj_obj1 <- get_projection(projection = projection, 
+               proj_scale = 2.5, 
+               scaled_data = scaled_mnist_data, 
+               scaled_data_model = scaled_mnist_data_model, 
+               distance_df_small_edges = distance_df_small_edges_mnist, 
+               axis_param = list(limits = 1, 
                                  axis_pos_x = -0.9, 
                                  axis_pos_y = -0.9, 
-  axis_labels = names(scaled_mnist_data),
-  threshold = 0.06)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
+                                 threshold = 0.06))
+
+projected_df <- proj_obj1$projected_df |>
+  mutate(cluster = tsne_minst_scaled_temp$cluster)
+
+model_df <- proj_obj1$model_df
+axes <- proj_obj1$axes
+circle <- proj_obj1$circle
 
@@
 
-projection_scaled <- projection * 2
-
-projected <- as.matrix(scaled_mnist_data) %*% projection_scaled
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  #dplyr::mutate(type = df_exe$type) |>
-  dplyr::mutate(ID = dplyr::row_number()) |>
-  mutate(cluster = tsne_minst_scaled_temp$cluster)
-
-projected_model <- as.matrix(scaled_mnist_data_model) %*% projection_scaled
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  distance_df_small_edges_mnist |> select(-distance), 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-axes_obj <- gen_axes(  
-  proj = projection,
-  limits = 1,
+proj_obj2 <- get_projection(projection = projection, 
+               proj_scale = 2, 
+               scaled_data = scaled_mnist_data, 
+               scaled_data_model = scaled_mnist_data_model, 
+               distance_df_small_edges = distance_df_small_edges_mnist, 
+               axis_param = list(limits = 1, 
                                  axis_pos_x = -0.83, 
                                  axis_pos_y = -0.83, 
-  axis_labels = names(scaled_mnist_data),
-  threshold = 0.06)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
+                                 threshold = 0.06))
+
+projected_df <- proj_obj2$projected_df |>
+  mutate(cluster = tsne_minst_scaled_temp$cluster)
+
+model_df <- proj_obj2$model_df
+axes <- proj_obj2$axes
+circle <- proj_obj2$circle
 

### 2025-03-17 paper.qmd
Updated PBMC3k projections

@@
 
-projection_scaled <- projection * 1.23
-
-projected <- as.matrix(scaled_pbmc_data) %*% projection_scaled
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  #dplyr::mutate(type = df_exe$type) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection_scaled
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  distance_df_small_edges_pbmc |> select(-distance), 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-axes_obj <- gen_axes(
-  proj = projection,
-  limits = 1,
+proj_obj1 <- get_projection(projection = projection, 
+               proj_scale = 1.23, 
+               scaled_data = scaled_pbmc_data, 
+               scaled_data_model = scaled_pbmc_data_model, 
+               distance_df_small_edges = distance_df_small_edges_pbmc, 
+               axis_param = list(limits = 1, 
                                  axis_pos_x = -0.48, 
                                  axis_pos_y = -0.48, 
-  axis_labels = names(scaled_pbmc_data),
-  threshold = 0.082)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
+                                 threshold = 0.082))
+
+projected_df <- proj_obj1$projected_df
+model_df <- proj_obj1$model_df
+axes <- proj_obj1$axes
+circle <- proj_obj1$circle
 
@@
 
-projection_scaled <- projection * 1.25
-
-projected <- as.matrix(scaled_pbmc_data) %*% projection_scaled
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  #dplyr::mutate(type = df_exe$type) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection_scaled
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  distance_df_small_edges_pbmc |> select(-distance), 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-axes_obj <- gen_axes(
-  proj = projection,
-  limits = 1.35,
+proj_obj2 <- get_projection(projection = projection, 
+               proj_scale = 1.25, 
+               scaled_data = scaled_pbmc_data, 
+               scaled_data_model = scaled_pbmc_data_model, 
+               distance_df_small_edges = distance_df_small_edges_pbmc, 
+               axis_param = list(limits = 1.35, 
                                  axis_pos_x = -0.75, 
                                  axis_pos_y = -0.75, 
-  axis_labels = names(scaled_pbmc_data),
-  threshold = 0.12)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
+                                 threshold = 0.12))
+
+projected_df <- proj_obj2$projected_df
+model_df <- proj_obj2$model_df
+axes <- proj_obj2$axes
+circle <- proj_obj2$circle
 

### 2025-03-17 paper.qmd
Updated the Five Gaussian clusters section

--- 2025-03-17 paper.qmd
@@
   limits <- axis_param$limits
+  axis_scaled <- axis_param$axis_scaled
   axis_pos_x <- axis_param$axis_pos_x
@@
   axes_obj <- gen_axes(  
-    proj = projection,
+    proj = projection * axis_scaled,
     limits = limits,
@@
 
-projection_scaled <- projection * 1
-
-projected <- as.matrix(scaled_gau_data) %*% projection_scaled
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  #dplyr::mutate(type = df_exe$type) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-projected_model <- as.matrix(scaled_gau_data_model) %*% projection_scaled
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  distance_df_small_edges |> select(-distance), 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-axes_obj <- gen_axes(  
-  proj = projection,
-  limits = 1,
+proj_obj1 <- get_projection(projection = projection, 
+               proj_scale = 1, 
+               scaled_data = scaled_gau_data, 
+               scaled_data_model = scaled_gau_data_model, 
+               distance_df_small_edges = distance_df_small_edges, 
+               axis_param = list(limits = 1,
+                                 axis_scaled = 1,
                                  axis_pos_x = -0.72, 
                                  axis_pos_y = -0.72, 
-  axis_labels = names(scaled_gau_data),
-  threshold = 0)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
+                                 threshold = 0))
+
+projected_df <- proj_obj1$projected_df
+model_df <- proj_obj1$model_df
+axes <- proj_obj1$axes
+circle <- proj_obj1$circle
 
@@
 
-projected <- as.matrix(scaled_gau_data) %*% projection
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-projected_model <- as.matrix(scaled_gau_data_model) %*% projection
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  distance_df_small_edges |> select(-distance), 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-axes_obj <- gen_axes(  
-  proj = projection,
-  limits = 1,
+proj_obj1 <- get_projection(projection = projection, 
+               proj_scale = 1, 
+               scaled_data = scaled_gau_data, 
+               scaled_data_model = scaled_gau_data_model, 
+               distance_df_small_edges = distance_df_small_edges, 
+               axis_param = list(limits = 1, 
+                                 axis_scaled = 1,
                                  axis_pos_x = -0.72, 
                                  axis_pos_y = -0.72, 
-  axis_labels = names(scaled_gau_data),
-  threshold = 0)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
+                                 threshold = 0))
+
+projected_df <- proj_obj1$projected_df
+model_df <- proj_obj1$model_df
+axes <- proj_obj1$axes
+circle <- proj_obj1$circle
 
@@
 
-projected <- as.matrix(scaled_gau_data) %*% projection
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-projected_model <- as.matrix(scaled_gau_data_model) %*% projection
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  distance_df_small_edges |> select(-distance), 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-axes_obj <- gen_axes(  
-  proj = projection,
-  limits = 1,
+proj_obj1 <- get_projection(projection = projection, 
+               proj_scale = 1, 
+               scaled_data = scaled_gau_data, 
+               scaled_data_model = scaled_gau_data_model, 
+               distance_df_small_edges = distance_df_small_edges, 
+               axis_param = list(limits = 1,
+                                 axis_scaled = 1,
                                  axis_pos_x = -0.72, 
                                  axis_pos_y = -0.72, 
-  axis_labels = names(scaled_gau_data),
-  threshold = 0)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
+                                 threshold = 0))
+
+projected_df <- proj_obj1$projected_df
+model_df <- proj_obj1$model_df
+axes <- proj_obj1$axes
+circle <- proj_obj1$circle
 
@@
 
-projection_scaled <- projection * 1.23
-
-projected <- as.matrix(scaled_pbmc_data) %*% projection_scaled
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  #dplyr::mutate(type = df_exe$type) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection_scaled
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  distance_df_small_edges_pbmc |> select(-distance), 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-axes_obj <- gen_axes(
-  proj = projection,
-  limits = 1,
+proj_obj1 <- get_projection(projection = projection, 
+               proj_scale = 1.23, 
+               scaled_data = scaled_pbmc_data, 
+               scaled_data_model = scaled_pbmc_data_model, 
+               distance_df_small_edges = distance_df_small_edges_pbmc, 
+               axis_param = list(limits = 1,
+                                 axis_scaled = 1, 
                                  axis_pos_x = -0.48, 
                                  axis_pos_y = -0.48, 
-  axis_labels = names(scaled_pbmc_data),
-  threshold = 0.082)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
+                                 threshold = 0.082))
+
+projected_df <- proj_obj1$projected_df
+model_df <- proj_obj1$model_df
+axes <- proj_obj1$axes
+circle <- proj_obj1$circle
 
@@
 
-projection_scaled <- projection * 1.25
-
-projected <- as.matrix(scaled_pbmc_data) %*% projection_scaled
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  #dplyr::mutate(type = df_exe$type) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection_scaled
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  distance_df_small_edges_pbmc |> select(-distance), 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-axes_obj <- gen_axes(
-  proj = projection,
-  limits = 1.35,
+proj_obj2 <- get_projection(projection = projection, 
+               proj_scale = 1.25, 
+               scaled_data = scaled_pbmc_data, 
+               scaled_data_model = scaled_pbmc_data_model, 
+               distance_df_small_edges = distance_df_small_edges_pbmc, 
+               axis_param = list(limits = 1.35, 
+                                 axis_scaled = 1, 
                                  axis_pos_x = -0.75, 
                                  axis_pos_y = -0.75, 
-  axis_labels = names(scaled_pbmc_data),
-  threshold = 0.12)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
+                                 threshold = 0.12))
+
+projected_df <- proj_obj2$projected_df
+model_df <- proj_obj2$model_df
+axes <- proj_obj2$axes
+circle <- proj_obj2$circle
 
@@
                axis_param = list(limits = 1, 
+                                 axis_scaled = 1, 
                                  axis_pos_x = -0.9, 
@@
                axis_param = list(limits = 1, 
+                                 axis_scaled = 1, 
                                  axis_pos_x = -0.83, 

### 2025-03-17 paper.qmd
Updated with projection function

@@
 
-projection_scaled <- projection * 1
-
-projected <- as.matrix(scaled_two_curvy_data) %*% projection_scaled
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-## For fitted model
-projected_model <- as.matrix(scaled_two_curvy_data_model) %*% projection_scaled
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  distance_df_small_edges_two_curvy |> select(-distance), 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+proj_obj1 <- get_projection(projection = projection, 
+               proj_scale = 1, 
+               scaled_data = scaled_two_curvy_data, 
+               scaled_data_model = scaled_two_curvy_data_model, 
+               distance_df_small_edges = distance_df_small_edges_two_curvy, 
+               axis_param = list(limits = 0.5, 
+                                 axis_scaled = 2,
+                                 axis_pos_x = -0.5, 
+                                 axis_pos_y = -0.5, 
+                                 threshold = 0.03))
+
+projected_df <- proj_obj1$projected_df
+model_df <- proj_obj1$model_df
+axes <- proj_obj1$axes
+circle <- proj_obj1$circle
 
@@
 
-projection_scaled <- projection * 1
-
-projected <- as.matrix(scaled_two_curvy_data) %*% projection_scaled
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  #dplyr::mutate(type = df_exe$type) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-projected_model <- as.matrix(scaled_two_curvy_data_model) %*% projection_scaled
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  wireframe_true_model, 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-axes_obj <- gen_axes(  
-  proj = projection * 2,
-  limits = 0.6,
+proj_obj1 <- get_projection(projection = projection, 
+               proj_scale = 1, 
+               scaled_data = scaled_two_curvy_data, 
+               scaled_data_model = scaled_two_curvy_data_model, 
+               distance_df_small_edges = distance_df_small_edges_two_curvy, 
+               axis_param = list(limits = 0.6, 
+                                 axis_scaled = 2,
                                  axis_pos_x = -0.4, 
                                  axis_pos_y = -0.4, 
-  axis_labels = names(scaled_two_curvy_data),
-  threshold = 0.04)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
+                                 threshold = 0.04))
+
+projected_df <- proj_obj1$projected_df
+model_df <- proj_obj1$model_df
+axes <- proj_obj1$axes
+circle <- proj_obj1$circle
+
+# projection_scaled <- projection * 1
+# 
+# projected <- as.matrix(scaled_two_curvy_data) %*% projection_scaled
+# 
+# projected_df <- projected |>
+#   tibble::as_tibble(.name_repair = "unique") |>
+#   dplyr::rename(c("proj1" = "...1",
+#                   "proj2" = "...2")) |>
+#   #dplyr::mutate(type = df_exe$type) |>
+#   dplyr::mutate(ID = dplyr::row_number()) 
+# 
+# projected_model <- as.matrix(scaled_two_curvy_data_model) %*% projection_scaled
+# 
+# projected_model_df <- projected_model |>
+#   tibble::as_tibble(.name_repair = "unique") |>
+#   dplyr::rename(c("proj1" = "...1",
+#                   "proj2" = "...2")) |>
+#   dplyr::mutate(ID = dplyr::row_number())
+# 
+# model_df <- dplyr::left_join(
+#   wireframe_true_model, 
+#   projected_model_df, 
+#   by = c("from" = "ID"))
+# 
+# names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+# 
+# model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+# names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+# 
+# axes_obj <- gen_axes(  
+#   proj = projection * 2,
+#   limits = 0.6,
+#   axis_pos_x = -0.4,
+#   axis_pos_y = -0.4,
+#   axis_labels = names(scaled_two_curvy_data),
+#   threshold = 0.04)
+# 
+# axes <- axes_obj$axes
+# circle <- axes_obj$circle
 
@@
 
-projection_scaled <- projection * 1
-
-projected <- as.matrix(scaled_two_curvy_data) %*% projection_scaled
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  #dplyr::mutate(type = df_exe$type) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-projected_model <- as.matrix(scaled_two_curvy_data_model) %*% projection_scaled
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  wireframe_true_model, 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-axes_obj <- gen_axes(  
-  proj = projection * 6,
-  limits = 0.5,
+proj_obj2 <- get_projection(projection = projection, 
+               proj_scale = 1, 
+               scaled_data = scaled_two_curvy_data, 
+               scaled_data_model = scaled_two_curvy_data_model, 
+               distance_df_small_edges = distance_df_small_edges_two_curvy, 
+               axis_param = list(limits = 0.5, 
+                                 axis_scaled = 6,
                                  axis_pos_x = -0.3, 
                                  axis_pos_y = -0.3, 
-  axis_labels = names(scaled_two_curvy_data),
-  threshold = 0)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
+                                 threshold = 0))
+
+projected_df <- proj_obj2$projected_df
+model_df <- proj_obj2$model_df
+axes <- proj_obj2$axes
+circle <- proj_obj2$circle
 
@@
 
-projection_scaled <- projection * 1
-
-projected <- as.matrix(scaled_two_curvy_data) %*% projection_scaled
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-## For fitted model
-projected_model <- as.matrix(scaled_two_curvy_data_model) %*% projection_scaled
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  distance_df_small_edges_two_curvy2 |> select(-distance), 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1, 
+                            scaled_data = scaled_gau_data, 
+                            scaled_data_model = scaled_gau_data_model, 
+                            distance_df_small_edges = distance_df_small_edges, 
+                            axis_param = list(limits = 0.15, 
+                                              axis_scaled = 5,
+                                              axis_pos_x = -0.067, 
+                                              axis_pos_y = -0.067, 
+                                              threshold = 0.007))
+
+projected_df <- proj_obj1$projected_df
+model_df <- proj_obj1$model_df
+axes <- proj_obj1$axes
+circle <- proj_obj1$circle
+
+# projection_scaled <- projection * 1
+# 
+# projected <- as.matrix(scaled_two_curvy_data) %*% projection_scaled
+# 
+# projected_df <- projected |>
+#   tibble::as_tibble(.name_repair = "unique") |>
+#   dplyr::rename(c("proj1" = "...1",
+#                   "proj2" = "...2")) |>
+#   dplyr::mutate(ID = dplyr::row_number()) 
+# 
+# ## For fitted model
+# projected_model <- as.matrix(scaled_two_curvy_data_model) %*% projection_scaled
+# 
+# projected_model_df <- projected_model |>
+#   tibble::as_tibble(.name_repair = "unique") |>
+#   dplyr::rename(c("proj1" = "...1",
+#                   "proj2" = "...2")) |>
+#   dplyr::mutate(ID = dplyr::row_number())
+# 
+# model_df <- dplyr::left_join(
+#   distance_df_small_edges_two_curvy2 |> select(-distance), 
+#   projected_model_df, 
+#   by = c("from" = "ID"))
+# 
+# names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+# 
+# model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
+# names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
 
@@
 
-axes_obj <- gen_axes(  
-  proj = projection * 5,
-  limits = 0.15,
-  axis_pos_x = -0.067,
-  axis_pos_y = -0.067,
-  axis_labels = names(scaled_two_curvy_data),
-  threshold = 0.007)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
+# axes_obj <- gen_axes(  
+#   proj = projection * 5,
+#   limits = 0.15,
+#   axis_pos_x = -0.067,
+#   axis_pos_y = -0.067,
+#   axis_labels = names(scaled_two_curvy_data),
+#   threshold = 0.007)
+# 
+# axes <- axes_obj$axes
+# circle <- axes_obj$circle
 

### 2025-03-17 paper.qmd
Added function plot MSE

@@
               circle = circle))
+
+}
+```
+
+```{r}
+#| label: plot-mse
+
+plot_mse <- function(error_df) {
+
+  ggplot(error_df, 
+       aes(x = a1, 
+           y = MSE, 
+           colour = method)) + 
+  geom_point(size = 0.8) +
+  geom_line(linewidth = 0.3) + 
+  scale_y_log10() +
+  ylab("log(MSE)") +
+  xlab(expression(paste("binwidth (", a[1], ")"))) +
+  theme_minimal() +
+  theme(panel.border = element_rect(fill = 'transparent'),
+        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
+        axis.ticks.x = element_line(),
+        axis.ticks.y = element_line(),
+        legend.position = "none",
+        axis.text.x = element_text(size = 7),
+        axis.text.y = element_text(size = 7),
+        axis.title.x = element_text(size = 7),
+        axis.title.y = element_text(size = 7))
 
@@
 
-error_plot_two_non_linear_diff_shaped_close_clusters <- ggplot(error_two_non_linear_diff_shaped_close_clusters, 
-                           aes(x = a1, 
-                               y = MSE, 
-                               colour = method)) + 
-  geom_point(size = 0.8) +
-  geom_line(linewidth = 0.3) + 
-  # geom_vline(xintercept = 15, linetype="solid", 
-  #            color = "black", linewidth=0.8, alpha = 0.5) +
+error_plot_two_non_linear_diff_shaped_close_clusters <-
+  plot_mse(error_two_non_linear_diff_shaped_close_clusters) +
   scale_x_continuous(breaks = sort(unique(error_two_non_linear_diff_shaped_close_clusters$a1))[seq(1, length(unique(error_two_non_linear_diff_shaped_close_clusters$a1)), by = 5)]) +
   scale_color_manual(values=c('#e41a1c','#ff7f00','#4daf4a', "#a65628",'#636363', '#984ea3')) +
-  scale_y_log10() +
-  ylab("log(MSE)") +
-  xlab(expression(paste("binwidth (", a[1], ")"))) +
-  theme_minimal() +
-  theme(panel.border = element_rect(fill = 'transparent'),
-        plot.title = element_text(size = 15, hjust = 0.5, vjust = -0.5),
-        axis.ticks.x = element_line(),
-        axis.ticks.y = element_line(),
-        legend.position = "none",
-        axis.text = element_text(size = 7),
-        axis.title = element_text(size = 12))
+  scale_y_log10()
 
@@
 
-error_plot_pbmc <- ggplot(error_pbmc, 
-                          aes(x = a1, 
-                              y = MSE, 
-                              colour = method)) + 
-  geom_point(size = 0.8) +
-  geom_line(linewidth = 0.3) + 
+error_plot_pbmc <- plot_mse(error_pbmc) +
   scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 13, 17, 21, 26)]) +
   scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a','#ff7f00',
-                              '#984ea3','#999999','#a65628','#f781bf')) +
-  scale_y_log10() +
-  ylab("log(MSE)") +
-  xlab(expression(paste("binwidth (", a[1], ")"))) +
-  theme_minimal() +
-  theme(panel.border = element_rect(fill = 'transparent'),
-        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
-        axis.ticks.x = element_line(),
-        axis.ticks.y = element_line(),
-        legend.position = "none",
-        axis.text.x = element_text(size = 7),
-        axis.text.y = element_text(size = 7),
-        axis.title.x = element_text(size = 7),
-        axis.title.y = element_text(size = 7))
+                              '#984ea3','#999999','#a65628','#f781bf'))
 
@@
 
-error_plot_mnist <- ggplot(error_mnist, 
-                          aes(x = a1, 
-                              y = MSE, 
-                              colour = method)) + 
-  geom_point(size = 0.8) +
-  geom_line(linewidth = 0.3) + 
+error_plot_mnist <- plot_mse(error_mnist) +
   scale_x_continuous(breaks = sort(unique(error_mnist$a1))[append(seq(1, length(unique(error_mnist$a1)), by = 5), 24)]) +
-  scale_color_manual(values=c('#e41a1c','#377eb8',"#ff7f00", '#984ea3', '#999999', '#a65628')) +
-  scale_y_log10() +
-  ylab("log(MSE)") +
-  xlab(expression(paste("binwidth (", a[1], ")"))) +
-  theme_minimal() +
-  theme(panel.border = element_rect(fill = 'transparent'),
-        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
-        axis.ticks.x = element_line(),
-        axis.ticks.y = element_line(),
-        legend.position = "none",
-        axis.text.x = element_text(size = 7),
-        axis.text.y = element_text(size = 7),
-        axis.title.x = element_text(size = 7),
-        axis.title.y = element_text(size = 7))
+  scale_color_manual(values=c('#e41a1c','#377eb8',"#ff7f00", '#984ea3', '#999999', '#a65628'))
 

### 2025-03-17 paper.qmd
Removed prediction code

@@
   interior_annotation("b", sc_ltr_pos, cex = 1.5)
-```
-
-```{r}
-#| label: dist-plot
-
-df_bin_two_curvy_temp <- df_bin_two_curvy
 ```
@@
 
-```{r}
-#| label: model-prediction-umap-quollr
-#| eval: false
-
-true_pred_df <- predict_emb(
-  test_data = true_model_df,
-  df_bin_centroids = df_bin_centroids_two_curvy,
-  df_bin = df_bin_two_curvy_temp,
-  type_NLDR = "UMAP"
-)
-
-# Compute radius r
-a1 <- calc_bins_y(
-  bin1 = 10, 
-  r2 = r2, 
-  q = 0.1)$a1
-
-r <- a1/2
-
-# Function to jitter points within a circumcircle of a hexagon
-jitter_within_circumcircle <- function(center_x, center_y, radius, num_points) {
-  theta <- runif(num_points, 0, 2 * pi)
-  r <- radius * sqrt(runif(num_points))
-  x <- center_x + r * cos(theta)
-  y <- center_y + r * sin(theta)
-  jittered_points <- data.frame(UMAP1 = x, UMAP2 = y, ID = points_in_hex$ID)
-  return(jittered_points)
-}
-
-# Jittered points data frame
-jittered_points_df <- data.frame()
-
-# Iterate through each hexagon and jitter points within the circumcircle
-for (hex_id in unique(true_pred_df$pred_hb_id)) {
-  # hex_points <- hex_grid_nonempty %>% filter(hex_poly_id == hex_id)
-  # center_x <- mean(hex_points$x)
-  # center_y <- mean(hex_points$y)
-  # points_in_hex <- true_pred_df %>% filter(pred_UMAP_1 >= min(hex_points$x) & pred_UMAP_1 <= max(hex_points$x) & pred_UMAP_2 >= min(hex_points$y) & pred_UMAP_2 <= max(hex_points$y))
-
-  points_in_hex <- true_pred_df |>
-    filter(pred_hb_id == hex_id)
-
-  center_x <- points_in_hex |> 
-    pull(pred_UMAP_1) |> 
-    unique()
-
-  center_y <- points_in_hex |> 
-    pull(pred_UMAP_2) |> 
-    unique()
-
-  if (nrow(points_in_hex) > 0) {
-    jittered_points <- jitter_within_circumcircle(center_x, center_y, r, nrow(points_in_hex))
-    jittered_points_df <- rbind(jittered_points_df, jittered_points)
-  }
-}
-
-plot_predict_tsne_two_curvy <- ggplot(
-  data = true_pred_df, 
-  aes(x = pred_UMAP_1,
-      y = pred_UMAP_2)) +
-  geom_point(alpha = 0.5) +
-  interior_annotation("c", c(0.08, 0.93), cex = 1.5)
-```
-
 <!-- XXX You need to have consistent colouring. In langevitour the points corresponding to observations are purple. They should be purple in the \gD{} view too.  -->
@@
 
-error_two_non_linear_diff_shaped_close_clusters_umap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_tsne.rds")
+error_two_curvy_umap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_tsne.rds")
 
 ## Find the minimum MSE when have duplicate a1
-error_two_non_linear_diff_shaped_close_clusters_umap <- error_two_non_linear_diff_shaped_close_clusters_umap |>
+error_two_curvy_umap <- error_two_curvy_umap |>
   group_by(a1) |>
@@
 
-base_line_prop_dens <- error_two_non_linear_diff_shaped_close_clusters_umap |>
+base_line_prop_dens <- error_two_curvy_umap |>
   filter(a1 == min(a1)) |>
@@
 
-error_two_non_linear_diff_shaped_close_clusters_umap <- error_two_non_linear_diff_shaped_close_clusters_umap |>
+error_two_curvy_umap <- error_two_curvy_umap |>
   mutate(prop_comp = prop_dens/base_line_prop_dens)
 
-mse_two_curvy_b <- ggplot(error_two_non_linear_diff_shaped_close_clusters_umap, 
+mse_two_curvy_b <- ggplot(error_two_curvy_umap, 
                      aes(x = a1, 
@@
   geom_point(size = 1) +
-  scale_x_continuous(breaks = sort(unique(round(error_two_non_linear_diff_shaped_close_clusters_umap$a1, 2)))[c(2, 5, 8, 10, 12, 14, 15, 16, 17)]) +
+  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_umap$a1, 2)))[c(2, 5, 8, 10, 12, 14, 15, 16, 17)]) +
   labs(x = expression(paste("binwidth (", a[1], ")")), y = "MSE") +
@@
 
-prop_dens_a1 <- ggplot(error_two_non_linear_diff_shaped_close_clusters_umap,
+prop_dens_a1 <- ggplot(error_two_curvy_umap,
                      aes(x = a1,
@@
   # scale_y_continuous(breaks = sort(unique(error_two_curvy$b_non_empty))[-3]) +
-  scale_x_continuous(breaks = sort(unique(round(error_two_non_linear_diff_shaped_close_clusters_umap$a1, 2)))[c(2, 5, 8, 10, 12, 14, 15, 16, 17)]) +
+  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_umap$a1, 2)))[c(2, 5, 8, 10, 12, 14, 15, 16, 17)]) +
   labs(x = expression(paste("binwidth (", a[1], ")")), y = paste("Relative proportion density")) +
@@
 
-error_two_non_linear_diff_shaped_close_clusters_umap <- error_two_non_linear_diff_shaped_close_clusters_umap |>
+error_two_curvy_umap <- error_two_curvy_umap |>
   mutate(prop_bins = b_non_empty/b)
 
-a1_m_two_curvy <- ggplot(error_two_non_linear_diff_shaped_close_clusters_umap,
+a1_m_two_curvy <- ggplot(error_two_curvy_umap,
                      aes(x = a1,
@@
   # scale_y_continuous(breaks = sort(unique(error_two_curvy$b_non_empty))[-3]) +
-  scale_x_continuous(breaks = sort(unique(round(error_two_non_linear_diff_shaped_close_clusters_umap$a1, 2)))[c(2, 5, 8, 10, 12, 14, 15, 16, 17)]) +
+  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_umap$a1, 2)))[c(2, 5, 8, 10, 12, 14, 15, 16, 17)]) +
   labs(x = expression(paste("binwidth (", a[1], ")")), y = expression(paste("proportion of non-empty bins ", bgroup("(", frac(m, b), ")")))) +
@@
 
-error_two_non_linear_diff_shaped_close_clusters_umap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_umap.rds")
+error_two_curvy_umap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_umap.rds")
 error_two_non_linear_diff_shaped_close_clusters_tsne <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_tsne.rds")
@@
 
-error_two_non_linear_diff_shaped_close_clusters <- bind_rows(error_two_non_linear_diff_shaped_close_clusters_umap, 
+error_two_non_linear_diff_shaped_close_clusters <- bind_rows(error_two_curvy_umap, 
                          error_two_non_linear_diff_shaped_close_clusters_tsne,

### 2025-03-17 paper.qmd
Updated the code

@@
 
-axes_obj <- gen_axes(  
-  proj = projection * 2,
-  limits = 0.5,
-  axis_pos_x = -0.5,
-  axis_pos_y = -0.5,
-  axis_labels = names(scaled_two_curvy_data),
-  threshold = 0.03)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
+# axes_obj <- gen_axes(  
+#   proj = projection * 2,
+#   limits = 0.5,
+#   axis_pos_x = -0.5,
+#   axis_pos_y = -0.5,
+#   axis_labels = names(scaled_two_curvy_data),
+#   threshold = 0.03)
+# 
+# axes <- axes_obj$axes
+# circle <- axes_obj$circle
 

### 2025-03-17 paper.qmd
Updated the model and proj for pbmc best

@@
       gp = grid::gpar(cex = cex, col=col)))
+}
+```
+
+```{r}
+#| label: gen-vis-obj
+
+gen_nldr_vis_algo_obj <- function(high_d_data, nldr_data, num_x_bins, col_start = "x") {
+
+  nldr_scaled_obj <- gen_scaled_data(
+    data = nldr_data)
+
+  nldr_scaled <- nldr_scaled_obj$scaled_nldr |>
+    mutate(ID = row_number())
+
+  lim1 <- nldr_scaled_obj$lim1
+  lim2 <- nldr_scaled_obj$lim2
+  r2 <- diff(lim2)/diff(lim1)
+
+  ## hexagon binning to have regular hexagons
+  hb_obj <- hex_binning(
+    data = nldr_scaled, 
+    bin1 = num_x_bins, 
+    r2 = r2,
+    q = 0.1)
+
+  ## Data set with all centroids
+  all_centroids_df <- hb_obj$centroids
+
+  a1_1 <- hb_obj$a1
+
+  ## Generate all coordinates of hexagons
+  hex_grid <- hb_obj$hex_poly
+
+  ## To obtain the standardise counts within hexbins
+  counts_df <- hb_obj$std_cts
+  df_bin_centroids <- extract_hexbin_centroids(
+    centroids_df = all_centroids_df, 
+    counts_df = counts_df) |>
+    filter(drop_empty == FALSE)
+
+  nldr_with_hb_id <- hb_obj$data_hb_id
+  df_all <- dplyr::bind_cols(high_d_data, 
+                                       nldr_with_hb_id)
+  df_bin <- avg_highd_data(data = df_all, col_start = col_start)
+
+  hex_grid_with_counts <- 
+    left_join(hex_grid,
+              counts_df, 
+              by = c("hex_poly_id" = "hb_id"))
+
+  hex_grid_nonempty <- hex_grid |>
+    filter(hex_poly_id %in% df_bin_centroids$hexID)
+
+  tr1_object <- tri_bin_centroids(
+    df_bin_centroids, x = "c_x", y = "c_y")
+  tr_from_to_df <- gen_edges(
+    tri_object = tr1_object)
+
+  ## Compute 2D distances
+  distance_df <- cal_2d_dist(
+    tr_coord_df = tr_from_to_df, 
+    start_x = "x_from", 
+    start_y = "y_from", 
+    end_x = "x_to", 
+    end_y = "y_to", 
+    select_vars = c("from", "to", "distance"))
+
+  ## To find the benchmark value
+  benchmark <- find_lg_benchmark(
+    distance_edges = distance_df, 
+    distance_col = "distance")
+
+  return(list(nldr_scaled = nldr_scaled,
+              hex_grid_with_counts = hex_grid_with_counts,
+              hex_grid_nonempty = hex_grid_nonempty,
+              df_bin_centroids = df_bin_centroids,
+              tr_from_to_df = tr_from_to_df,
+              df_bin = df_bin,
+              distance_df = distance_df,
+              benchmark = benchmark))
+
 }
@@
 num_bins_x_pbmc <- 60
-lim1 <- pbmc_scaled_obj$lim1
-lim2 <- pbmc_scaled_obj$lim2
-r2_pbmc <- diff(lim2)/diff(lim1)
-
-pbmc_model <- fit_highd_model(
-  training_data = training_data_pbmc,
-  emb_df = umap_pbmc_scaled,
-  bin1 = num_bins_x_pbmc,
-  r2 = r2_pbmc,
-  q = 0.1,
-  is_bin_centroid = TRUE,
-  is_rm_lwd_hex = FALSE,
-  col_start_highd = "pc"
-)
-
-df_bin_centroids_pbmc <- pbmc_model$df_bin_centroids
-df_bin_pbmc <- pbmc_model$df_bin
-
-## Triangulate bin centroids
-tr1_object_pbmc <- tri_bin_centroids(
-  df_bin_centroids_pbmc, x = "c_x", y = "c_y")
-tr_from_to_df_pbmc <- gen_edges(
-  tri_object = tr1_object_pbmc)
-
-## Compute 2D distances
-distance_pbmc <- cal_2d_dist(
-  tr_coord_df = tr_from_to_df_pbmc,
-  start_x = "x_from",
-  start_y = "y_from",
-  end_x = "x_to",
-  end_y = "y_to",
-  select_vars = c("from", "to", "distance"))
-
-## To find the benchmark value
-benchmark_pbmc <- find_lg_benchmark(
-  distance_edges = distance_pbmc,
-  distance_col = "distance")
-
-tr_df <- distinct(tibble(
-  x = c(tr_from_to_df_pbmc[["x_from"]], tr_from_to_df_pbmc[["x_to"]]), 
-  y = c(tr_from_to_df_pbmc[["y_from"]], tr_from_to_df_pbmc[["y_to"]])))
+
+algo_obj_pbmc <- gen_nldr_vis_algo_obj(
+  high_d_data = training_data_pbmc, 
+  nldr_data = umap_pbmc, 
+  num_x_bins = num_bins_x_pbmc,
+  col_start = "pc")
+
+umap_pbmc_scaled <- algo_obj_pbmc$nldr_scaled
+distance_pbmc <- algo_obj_pbmc$distance_df
+tr_from_to_df_pbmc <- algo_obj_pbmc$tr_from_to_df
+benchmark_pbmc <- algo_obj_pbmc$benchmark
+df_bin_centroids_pbmc <- algo_obj_pbmc$df_bin_centroids
+df_bin_pbmc <- algo_obj_pbmc$df_bin
 
@@
 
-# Combine with the true model for visualization
-df <- dplyr::bind_rows(scaled_pbmc_data_model |> mutate(type = "model"),
-                       scaled_pbmc_data |> mutate(type = "data"))
+# # Combine with the true model for visualization
+# df <- dplyr::bind_rows(scaled_pbmc_data_model |> mutate(type = "model"),
+#                        scaled_pbmc_data |> mutate(type = "data"))
 
@@
 #| label: num-bins-pbmc
+
 ## Compute hexbin parameters
 num_bins_x_pbmc <- 30
-lim1 <- pbmc_scaled_obj$lim1
-lim2 <- pbmc_scaled_obj$lim2
-r2_pbmc <- diff(lim2)/diff(lim1)
-
-pbmc_model <- fit_highd_model(
-  training_data = training_data_pbmc,
-  emb_df = umap_pbmc_scaled_best,
-  bin1 = num_bins_x_pbmc,
-  r2 = r2_pbmc,
-  q = 0.1,
-  is_bin_centroid = TRUE,
-  is_rm_lwd_hex = FALSE,
-  col_start_highd = "pc"
-)
-
-df_bin_centroids_pbmc <- pbmc_model$df_bin_centroids
-df_bin_pbmc <- pbmc_model$df_bin
-
-## Triangulate bin centroids
-tr1_object_pbmc <- tri_bin_centroids(
-  df_bin_centroids_pbmc, x = "c_x", y = "c_y")
-tr_from_to_df_pbmc <- gen_edges(
-  tri_object = tr1_object_pbmc)
-
-## Compute 2D distances
-distance_pbmc <- cal_2d_dist(
-  tr_coord_df = tr_from_to_df_pbmc,
-  start_x = "x_from",
-  start_y = "y_from",
-  end_x = "x_to",
-  end_y = "y_to",
-  select_vars = c("from", "to", "distance"))
-
-## To find the benchmark value
-benchmark_pbmc <- find_lg_benchmark(
-  distance_edges = distance_pbmc,
-  distance_col = "distance")
-
-tr_df <- distinct(tibble::tibble(
-  x = c(tr_from_to_df_pbmc[["x_from"]], tr_from_to_df_pbmc[["x_to"]]),
-  y = c(tr_from_to_df_pbmc[["y_from"]], tr_from_to_df_pbmc[["y_to"]])))
+
+algo_obj_pbmc <- gen_nldr_vis_algo_obj(
+  high_d_data = training_data_pbmc, 
+  nldr_data = umap_pbmc2, 
+  num_x_bins = num_bins_x_pbmc,
+  col_start = "pc")
+
+umap_pbmc_scaled_best <- algo_obj_pbmc$nldr_scaled
+distance_pbmc <- algo_obj_pbmc$distance_df
+tr_from_to_df_pbmc <- algo_obj_pbmc$tr_from_to_df
+benchmark_pbmc <- algo_obj_pbmc$benchmark
+df_bin_centroids_pbmc <- algo_obj_pbmc$df_bin_centroids
+df_bin_pbmc <- algo_obj_pbmc$df_bin
 
@@
   )
-
-## Compute error
-error_df <- augment(
-  df_bin_centroids = df_bin_centroids_pbmc,
-  df_bin = df_bin_pbmc,
-  training_data = training_data_pbmc,
-  newdata = NULL,
-  type_NLDR = "UMAP",
-  col_start = "pc")
-
-## Categorize error
-
-error_df <- error_df |>
-  mutate(type = case_when(
-    row_wise_abs_error <= 5 ~ "error 0-5",
-    row_wise_abs_error <= 10 ~ "error 5-10",
-    row_wise_abs_error <= 15 ~ "error 10-15",
-    row_wise_abs_error <= 20 ~ "error 15-20",
-    row_wise_abs_error <= 25 ~ "error 20-25",
-    .default = "error greter than 25"
-  )) |>
-  mutate(type = factor(type, levels = c(
-    "error 0-5", "error 5-10", "error 10-15", "error 15-20", "error 20-25", 
-    "error greter than 25")))
-
-## To join embedding
-error_df <- error_df |>
-  bind_cols(umap_pbmc_scaled_best |>
-              select(-ID))
-
-error_plot_pbmc_best <- error_df |>
-  ggplot(aes(x = UMAP1,
-             y = UMAP2,
-             color = type,
-             group = ID)) +
-  geom_point(alpha=0.5,
-             size = 0.1) +
-  interior_annotation("b", sc_ltr_pos)
-
 ```
@@
 
-# Combine with the true model for visualization
-df <- dplyr::bind_rows(scaled_pbmc_data_model |> mutate(type = "model"),
-                       scaled_pbmc_data |> mutate(type = "data"))
-
 ## First projection
@@
 
-projection_scaled <- projection * 1.23
-
-projected <- as.matrix(scaled_pbmc_data) %*% projection_scaled
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  #dplyr::mutate(type = df_exe$type) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection_scaled
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  distance_df_small_edges_pbmc |> select(-distance), 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-axes_obj <- gen_axes(
-  proj = projection,
-  limits = 1,
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.23, 
+                            scaled_data = scaled_pbmc_data, 
+                            scaled_data_model = scaled_pbmc_data_model, 
+                            distance_df_small_edges = distance_df_small_edges_pbmc, 
+                            axis_param = list(limits = 1, 
+                                              axis_scaled = 1,
                                               axis_pos_x = -0.48, 
                                               axis_pos_y = -0.48, 
-  axis_labels = names(scaled_pbmc_data),
-  threshold = 0.082)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
+                                              threshold = 0.082))
+
+projected_df <- proj_obj1$projected_df
+model_df <- proj_obj1$model_df
+axes <- proj_obj1$axes
+circle <- proj_obj1$circle
 
@@
 
-projection_scaled <- projection * 1.25
-
-projected <- as.matrix(scaled_pbmc_data) %*% projection_scaled
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  #dplyr::mutate(type = df_exe$type) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection_scaled
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  distance_df_small_edges_pbmc |> select(-distance), 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-axes_obj <- gen_axes(
-  proj = projection,
-  limits = 1.35,
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.25, 
+                            scaled_data = scaled_pbmc_data, 
+                            scaled_data_model = scaled_pbmc_data_model, 
+                            distance_df_small_edges = distance_df_small_edges_pbmc, 
+                            axis_param = list(limits = 1.35, 
+                                              axis_scaled = 1,
                                               axis_pos_x = -0.75, 
                                               axis_pos_y = -0.75, 
-  axis_labels = names(scaled_pbmc_data),
-  threshold = 0.12)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
+                                              threshold = 0.12))
+
+projected_df <- proj_obj2$projected_df
+model_df <- proj_obj2$model_df
+axes <- proj_obj2$axes
+circle <- proj_obj2$circle
 
@@
   mutate(type = "data")
-
-mnist_scaled_obj <- gen_scaled_data(
-  data = tsne_mnist2)
-tsne_minst_scaled <- mnist_scaled_obj$scaled_nldr 
 ```
@@
 #| label: hexbin-mnist
-## Compute hexbin parameters
+
 num_bins_x_mnist <- 30
-lim1 <- mnist_scaled_obj$lim1
-lim2 <- mnist_scaled_obj$lim2
-r2_mnist <- diff(lim2)/diff(lim1)
-
-mnist_model <- fit_highd_model(
-  training_data = training_data_mnist,
-  emb_df = tsne_minst_scaled,
-  bin1 = num_bins_x_mnist,
-  r2 = r2_mnist,
-  is_bin_centroid = TRUE,
-  is_rm_lwd_hex = TRUE,
-  col_start_highd = "PC"
-)
-
-df_bin_centroids_mnist <- mnist_model$df_bin_centroids
-df_bin_mnist <- mnist_model$df_bin
-
-## Triangulate bin centroids
-tr1_object_mnist <- tri_bin_centroids(
-  df_bin_centroids_mnist, x = "c_x", y = "c_y")
-tr_from_to_df_mnist <- gen_edges(
-  tri_object = tr1_object_mnist)
-
-## Compute 2D distances
-distance_mnist <- cal_2d_dist(
-  tr_coord_df = tr_from_to_df_mnist, 
-  start_x = "x_from", 
-  start_y = "y_from", 
-  end_x = "x_to", 
-  end_y = "y_to", 
-  select_vars = c("from", "to", "distance"))
-
-## To find the benchmark value
-benchmark_mnist <- find_lg_benchmark(
-  distance_edges = distance_mnist, 
-  distance_col = "distance")
+
+algo_obj_mnist <- gen_nldr_vis_algo_obj(
+  high_d_data = training_data_mnist, 
+  nldr_data = tsne_mnist2, 
+  num_x_bins = num_bins_x_mnist,
+  col_start = "PC")
+
+tsne_minst_scaled <- algo_obj_mnist$nldr_scaled
+distance_mnist <- algo_obj_mnist$distance_df
+tr_from_to_df_mnist <- algo_obj_mnist$tr_from_to_df
+benchmark_mnist <- algo_obj_mnist$benchmark
+df_bin_centroids_mnist <- algo_obj_mnist$df_bin_centroids
+df_bin_mnist <- algo_obj_mnist$df_bin
 
 sc_ltr_pos_mnist <- c(0.96, 0.96)
-
-tr_df <- distinct(tibble::tibble(
-  x = c(tr_from_to_df_mnist[["x_from"]], tr_from_to_df_mnist[["x_to"]]), 
-  y = c(tr_from_to_df_mnist[["y_from"]], tr_from_to_df_mnist[["y_to"]])))
 
@@
 
-tr_from_to_df_mnist <- inner_join(
-  tr_from_to_df_mnist, distance_df_small_edges_mnist, 
-  by = c("from", "to")) |>
-  mutate(ID = row_number())
-
-tsne_minst_scaled_temp <- tsne_minst_scaled |>
+tsne_minst_scaled <- tsne_minst_scaled |>
   mutate(cluster = if_else((tSNE1 <= 1) & (tSNE1 >= 0.78) & (tSNE2 <= 0.62) & (tSNE2 >= 0.39), "small_clust", "big_clust"))
@@
                colour = "#000000") +
-  geom_point(data = tsne_minst_scaled_temp,
+  geom_point(data = tsne_minst_scaled,
              aes(
@@
 tsne_plot_mnist_clust <- ggplot() + 
-  geom_point(data = tsne_minst_scaled_temp,
+  geom_point(data = tsne_minst_scaled,
              aes(
@@
 projected_df <- proj_obj1$projected_df |>
-  mutate(cluster = tsne_minst_scaled_temp$cluster)
+  mutate(cluster = tsne_minst_scaled$cluster)
 
@@
 projected_df <- proj_obj2$projected_df |>
-  mutate(cluster = tsne_minst_scaled_temp$cluster)
+  mutate(cluster = tsne_minst_scaled$cluster)
 

### 2025-03-17 paper.qmd
Added function to plot the projection

@@
 ```{r}
-#| label: plot-projection
+#| label: get-projection
 
@@
               circle = circle))
+
+}
+```
+
+```{r}
+#| label: plot-projection
+
+plot_proj <- function(projected_df, model_df, axes, circle, 
+                      point_param = c(1, 0.5, "#000000"), # size, alpha, color
+                      plot_limits, title, cex = 2, 
+                      axis_text_size = 3,
+                      is_category = FALSE) {
+
+  if(is_category == FALSE) {
+
+    initial_plot <- projected_df |>
+      ggplot(
+        aes(
+          x = proj1, 
+          y = proj2))
+
+  } else {
+
+    initial_plot <- projected_df |>
+      ggplot(
+        aes(
+          x = proj1, 
+          y = proj2,
+          colour = cluster))
+
+  }
+
+  initial_plot + 
+    geom_segment(
+      data = model_df, 
+      aes(
+        x = proj1_from, 
+        y = proj2_from, 
+        xend = proj1_to, 
+        yend = proj2_to), 
+      color = "#000000") +
+    geom_point(
+      size = point_param[1],
+      alpha = point_param[2],
+      color = point_param[3]) +
+    geom_segment(
+      data=axes, 
+      aes(x=x1, y=y1, xend=x2, yend=y2), 
+      colour="grey70") +
+    geom_text(
+      data=axes,
+      aes(x=x2, y=y2), 
+      label=rownames(axes),
+      colour="grey50",
+      size = axis_text_size) +
+    geom_path(
+      data=circle, 
+      aes(x=c1, y=c2), colour="grey70") +
+    coord_fixed() +
+    xlim(plot_limits) +
+    ylim(plot_limits) +
+    interior_annotation(title, cex = cex)
 

### 2025-03-17 paper.qmd
Added function to plot the 2D projection

--- 2025-03-17 paper.qmd
@@
               circle = circle))
+
+}
+```
+
+```{r}
+#| label: plot-projection
+
+plot_proj <- function(projected_df, model_df, axes, circle, 
+                      point_param = c(1.5, 0.5, "#000000"), # size, alpha, color
+                      line_param = c(0.5, 0.5), #linewidth, alpha
+                      plot_limits, title, cex = 2, 
+                      axis_text_size = 3,
+                      is_category = FALSE) {
+
+  if(is_category == FALSE) {
+
+    initial_plot <- projected_df |>
+      ggplot(
+        aes(
+          x = proj1, 
+          y = proj2))
+
+  } else {
+
+    initial_plot <- projected_df |>
+      ggplot(
+        aes(
+          x = proj1, 
+          y = proj2,
+          colour = cluster))
+
+  }
+
+  initial_plot + 
+    geom_segment(
+      data = model_df, 
+      aes(
+        x = proj1_from, 
+        y = proj2_from, 
+        xend = proj1_to, 
+        yend = proj2_to), 
+      color = "#000000",
+      linewidth = line_param[1],
+      alpha = line_param[2]) +
+    geom_point(
+      size = as.numeric(point_param[1]),
+      alpha = as.numeric(point_param[2]),
+      color = point_param[3]) +
+    geom_segment(
+      data=axes, 
+      aes(x=x1, y=y1, xend=x2, yend=y2), 
+      colour="grey70") +
+    geom_text(
+      data=axes,
+      aes(x=x2, y=y2), 
+      label=rownames(axes),
+      colour="grey50",
+      size = axis_text_size) +
+    geom_path(
+      data=circle, 
+      aes(x=c1, y=c2), colour="grey70") +
+    xlim(plot_limits) +
+    ylim(plot_limits) +
+    interior_annotation(title, cex = cex)
 
@@
 
-pbmc_proj_umap_model1 <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#000000") +
-  geom_point(
-    #size = 0.5,
-    alpha = 0.2,
-    color = '#a65628') +
-  geom_segment(
-    data=axes, 
-    aes(x=x1, y=y1, xend=x2, yend=y2), 
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 3) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.65, 0.65)) +
-  ylim(c(-0.65, 0.65)) +
-  interior_annotation("a2", cex = 2)
+pbmc_proj_umap_model1 <- plot_proj(
+  projected_df = projected_df, 
+  model_df = model_df, 
+  axes = axes, 
+  circle = circle, 
+  point_param = c(1.5, 0.2, "#a65628"), # size, alpha, color
+  plot_limits = c(-0.65, 0.65), 
+  title = "a2", 
+  cex = 2, 
+  is_category = FALSE)
 
@@
 
-pbmc_proj_umap_model2 <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1,
-      y = proj2)) +
-  geom_segment(
-    data = model_df,
-    aes(
-      x = proj1_from,
-      y = proj2_from,
-      xend = proj1_to,
-      yend = proj2_to),
-    color = "#000000") +
-  geom_point(
-    #size = 0.5,
-    alpha = 0.2,
-    color = '#a65628') +
-  geom_segment(
-    data=axes,
-    aes(x=x1, y=y1, xend=x2, yend=y2),
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2),
-    label=rownames(axes),
-    colour="grey50",
-    size = 3) +
-  geom_path(
-    data=circle,
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-1, 0.75)) +
-  ylim(c(-1, 0.75)) +
-  interior_annotation("a3", cex = 2)
+pbmc_proj_umap_model2 <- plot_proj(
+    projected_df = projected_df, 
+    model_df = model_df, 
+    axes = axes, 
+    circle = circle, 
+    point_param = c(1.5, 0.2, "#a65628"), # size, alpha, color
+    plot_limits = c(-1, 0.75), 
+    title = "a3", 
+    cex = 2, 
+    is_category = FALSE)
 
@@
 
+tr_from_to_df_mnist <- inner_join(
+  tr_from_to_df_mnist, distance_df_small_edges_mnist, 
+  by = c("from", "to"))
+
 tsne_minst_scaled <- tsne_minst_scaled |>
@@
 
-mnist_proj_tsne_model1 <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2,
-      colour = cluster)) +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#000000",
-    linewidth = 0.5,
-    alpha = 0.5) +
-  geom_point(
-    size = 0.5,
-    alpha = 0.5) +
-  scale_color_manual(values = c("#999999",'#ff7f00')) +
-  geom_segment(
-    data=axes,
-    aes(x=x1, y=y1, xend=x2, yend=y2),
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 2) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  #coord_fixed() +
-  xlim(c(-1.1, 0.9)) +
-  ylim(c(-1.1, 0.9)) +
-  interior_annotation("b", cex = 1.5) +
+mnist_proj_tsne_model1 <- plot_proj(
+  projected_df = projected_df, 
+  model_df = model_df, 
+  axes = axes, 
+  circle = circle, 
+  point_param = c(0.5, 0.5, "#000000"), # size, alpha, color
+  line_param = c(0.5, 0.5), # linewidth, alpha
+  plot_limits = c(-1.1, 0.9), 
+  title = "b", 
+  cex = 2, 
+  axis_text_size = 3, 
+  is_category = TRUE) +
+  scale_color_manual(values = c("#d9d9d9",'#ff7f00')) +
   theme(

### 2025-03-17 paper.qmd
Updated the projectio plot with new function for PBMC3k

@@
 
-pbmc_proj_umap_model1_best <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#000000") +
-  geom_point(
-    #size = 0.5,
-    alpha = 0.2,
-    color = "#999999") +
-  geom_segment(
-    data=axes, 
-    aes(x=x1, y=y1, xend=x2, yend=y2), 
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 3) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.65, 0.65)) +
-  ylim(c(-0.65, 0.65)) +
-  interior_annotation("b2", cex = 2)
+pbmc_proj_umap_model1_best <- plot_proj(
+    projected_df = projected_df, 
+    model_df = model_df, 
+    axes = axes, 
+    circle = circle, 
+    point_param = c(1.5, 0.2, "#999999"), # size, alpha, color
+    plot_limits = c(-0.65, 0.65), 
+    title = "b2", 
+    cex = 2, 
+    is_category = FALSE)
 
@@
 
-pbmc_proj_umap_model2_best <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1,
-      y = proj2)) +
-  geom_segment(
-    data = model_df,
-    aes(
-      x = proj1_from,
-      y = proj2_from,
-      xend = proj1_to,
-      yend = proj2_to),
-    color = "#000000") +
-  geom_point(
-    #size = 0.5,
-    alpha = 0.2,
-    color = '#999999') +
-  geom_segment(
-    data=axes,
-    aes(x=x1, y=y1, xend=x2, yend=y2),
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2),
-    label=rownames(axes),
-    colour="grey50",
-    size = 3) +
-  geom_path(
-    data=circle,
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-1, 0.75)) +
-  ylim(c(-1, 0.75)) +
-  interior_annotation("b3", cex = 2)
-
+pbmc_proj_umap_model2_best <- plot_proj(
+    projected_df = projected_df, 
+    model_df = model_df, 
+    axes = axes, 
+    circle = circle, 
+    point_param = c(1.5, 0.2, "#999999"), # size, alpha, color
+    plot_limits = c(-1, 0.75), 
+    title = "b3", 
+    cex = 2, 
+    is_category = FALSE)
 ```

### 2025-03-17 paper.qmd
Try to update with new function

@@
 
-  a1_1 <- hb_obj$a1
+  a1 <- hb_obj$a1
 
@@
   return(list(nldr_scaled = nldr_scaled,
+              a1 = a1,
               hex_grid_with_counts = hex_grid_with_counts,
@@
 tsne_two_curvy <- read_rds(file = "data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_30.rds") 
-
-two_curvy_scaled_obj <- gen_scaled_data(
-  data = tsne_two_curvy)
-
-tsne_two_curvy_scaled <- two_curvy_scaled_obj$scaled_nldr |>
-  mutate(ID = row_number())
-
-lim1 <- two_curvy_scaled_obj$lim1
-lim2 <- two_curvy_scaled_obj$lim2
-r2 <- diff(lim2)/diff(lim1)
+```
+
+```{r}
+#| label: fit_model-two-curvy
+
+num_bins_x_two_curvy <- 22
+
+algo_obj_two_curvy <- gen_nldr_vis_algo_obj(
+  high_d_data = training_data_two_curvy, 
+  nldr_data = tsne_two_curvy, 
+  num_x_bins = num_bins_x_two_curvy,
+  col_start = "x")
+
+tsne_two_curvy_scaled <- algo_obj_two_curvy$nldr_scaled
+distance_two_curvy <- algo_obj_two_curvy$distance_df
+tr_from_to_df_two_curvy <- algo_obj_two_curvy$tr_from_to_df
+benchmark_two_curvy <- algo_obj_two_curvy$benchmark
+df_bin_centroids_two_curvy <- algo_obj_two_curvy$df_bin_centroids
+df_bin_two_curvy <- algo_obj_two_curvy$df_bin
+hex_grid_with_counts <- algo_obj_two_curvy$hex_grid_with_counts
+hex_grid_nonempty <- algo_obj_two_curvy$hex_grid_nonempty
+bin_width <- algo_obj_two_curvy$a1
 
@@
 sc_ylims <- c(-0.17, 1.25)
+```
+
+<!--Full hexagon grid with UMAP data-->
+
+```{r}
+#| label: hexbin-two-curvy
 
@@
   interior_annotation("a1", sc_ltr_pos)
-```
-
-<!--Full hexagon grid with UMAP data-->
-
-```{r}
-#| label: hexbin-two-curvy
-## Compute hexbin parameters
-num_bins_x_two_curvy <- 22
-
-## hexagon binning to have regular hexagons
-hb_obj_two_curvy <- hex_binning(
-  data = tsne_two_curvy_scaled, 
-  bin1 = num_bins_x_two_curvy, 
-  r2 = r2,
-  q = 0.1)
-
-## Data set with all centroids
-all_centroids_df <- hb_obj_two_curvy$centroids
-
-a1_1 <- hb_obj_two_curvy$a1
-
-## Generate all coordinates of hexagons
-hex_grid <- hb_obj_two_curvy$hex_poly
-
-## To obtain the standardise counts within hexbins
-counts_df <- hb_obj_two_curvy$std_cts
-df_bin_centroids_two_curvy <- extract_hexbin_centroids(
-  centroids_df = all_centroids_df, 
-  counts_df = counts_df) |>
-  filter(drop_empty == FALSE)
-
-umap_data_two_curvy_with_hb_id <- hb_obj_two_curvy$data_hb_id
-df_all_two_curvy <- dplyr::bind_cols(training_data_two_curvy, umap_data_two_curvy_with_hb_id)
-df_bin_two_curvy <- avg_highd_data(data = df_all_two_curvy, col_start = "x")
-
-hex_grid_with_counts <- 
-  left_join(hex_grid,
-            counts_df, 
-            by = c("hex_poly_id" = "hb_id"))
 
@@
 #| label: empty-bin-two-curvy
-hex_grid_nonempty <- hex_grid |>
-  filter(hex_poly_id %in% df_bin_centroids_two_curvy$hexID)
 
@@
 #| label: triangulate-two-curvy
-## Triangulate bin centroids
-tr1_object_two_curvy <- tri_bin_centroids(
-  df_bin_centroids_two_curvy, x = "c_x", y = "c_y")
-tr_from_to_df_two_curvy <- gen_edges(
-  tri_object = tr1_object_two_curvy)
 
@@
 
-## Compute 2D distances
-distance_two_curvy <- cal_2d_dist(
-  tr_coord_df = tr_from_to_df_two_curvy, 
-  start_x = "x_from", 
-  start_y = "y_from", 
-  end_x = "x_to", 
-  end_y = "y_to", 
-  select_vars = c("from", "to", "distance"))
-
-## To find the benchmark value
-benchmark_two_curvy <- find_lg_benchmark(
-  distance_edges = distance_two_curvy, 
-  distance_col = "distance")
-
 distance_df_small_edges_two_curvy <- distance_two_curvy |>
     filter(distance < benchmark_two_curvy)
-
-bin_width <- hb_obj_two_curvy$a1
-
-####With diff benchmark values
-
-## Benchmark 2
-
-benchmark_two_curvy <- 2.8 * bin_width
-
-distance_df_small_edges_two_curvy3 <- distance_two_curvy |>
-  filter(distance < benchmark_two_curvy)
-
-tr_from_to_df_two_curvy4 <- inner_join(
-  tr_from_to_df_two_curvy, distance_df_small_edges_two_curvy3,
-  by = c("from", "to"))
-
-trimesh_two_curvy_removed2 <- ggplot() +
-  geom_segment(data = tr_from_to_df_two_curvy4,
-               aes(
-                 x = x_from,
-                 y = y_from,
-                 xend = x_to,
-                 yend = y_to),
-               colour = "#000000")+
-  geom_point(data = tsne_two_curvy_scaled,
-             aes(
-               x = tSNE1,
-               y = tSNE2
-             ),
-             color = "#636363",
-             alpha = 0.5,
-             size = 0.5
-  ) +
-  coord_equal() +
-  interior_annotation("b1", sc_ltr_pos)
-
-## Benchmark 1
-
-benchmark_two_curvy <- 2 * bin_width
-
-distance_df_small_edges_two_curvy2 <- distance_two_curvy |>
-  filter(distance < benchmark_two_curvy) ## 0.231
-
-tr_from_to_df_two_curvy3 <- inner_join(
-  tr_from_to_df_two_curvy, distance_df_small_edges_two_curvy2,
-  by = c("from", "to"))
-
-trimesh_two_curvy_removed1 <- ggplot() +
-  geom_segment(data = tr_from_to_df_two_curvy3,
-               aes(
-                 x = x_from,
-                 y = y_from,
-                 xend = x_to,
-                 yend = y_to),
-               colour = "#000000")+
-  geom_point(data = tsne_two_curvy_scaled,
-             aes(
-               x = tSNE1,
-               y = tSNE2
-             ),
-             color = "#636363",
-             alpha = 0.5,
-             size = 0.5
-  ) +
-  coord_equal() +
-  interior_annotation("c1", sc_ltr_pos)
-
-trimesh_two_curvy_removed1_with_data <- ggplot() +
-  geom_segment(data = tr_from_to_df_two_curvy3,
-               aes(
-                 x = x_from,
-                 y = y_from,
-                 xend = x_to,
-                 yend = y_to),
-               colour = "#000000") +
-  geom_point(data = tsne_two_curvy_scaled,
-             aes(
-               x = tSNE1,
-               y = tSNE2
-             ),
-             color = "#636363",
-             alpha = 0.5,
-             size = 0.5
-  ) +
-  coord_equal() +
-  #xlim(c(-0.01, 1.1)) + ylim(c(-0.08, 0.57)) +
-  interior_annotation("a2", sc_ltr_pos)
 

### 2025-03-19 paper.qmd
Fixed Figure 2a

@@
 
+two_curvy_proj_true <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = true_model_df_proj, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#2b8cbe") +
+  geom_point(
+    size = 0.5,
+    alpha = 0.2,
+    color = "#636363") +
+  geom_segment(
+    data=axes,
+    aes(x=x1, y=y1, xend=x2, yend=y2),
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
+    colour="grey50",
+    size = 4) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.6, 0.6)) + 
+  ylim(c(-0.6, 0.6)) +
+  interior_annotation("a", sc_ltr_pos, cex = 1.5) +
+  theme(
+    legend.position = "none"
+  )
+
 two_curvy_proj_umap_all_model1 <- projected_df |>
@@
   )
+
+## Second projection
+projection <- cbind(
+  c(-0.38281,-0.04870,0.00619,-0.02451),
+  c(-0.01303,0.00634,-0.37421,0.09648))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1, 
+                            scaled_data = scaled_two_curvy_data, 
+                            scaled_data_model = scaled_two_curvy_data_model, 
+                            distance_df_small_edges = distance_df_small_edges_two_curvy, 
+                            axis_param = list(limits = 0.5, 
+                                              axis_scaled = 6,
+                                              axis_pos_x = -0.3, 
+                                              axis_pos_y = -0.3, 
+                                              threshold = 0))
+
+projected_df <- proj_obj2$projected_df
+model_df <- proj_obj2$model_df
+axes <- proj_obj2$axes
+circle <- proj_obj2$circle
+
+projected_true_model <- as.matrix(scaled_two_curvy_data_true_model) %*% projection_scaled
+
+projected_true_model_df <- projected_true_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+true_model_df_proj <- dplyr::left_join(
+  wireframe_true_model, 
+  projected_true_model_df, 
+  by = c("from" = "ID"))
+
+names(true_model_df_proj)[3:NCOL(true_model_df_proj)] <- paste0(names(projected_true_model_df)[-NCOL(projected_true_model_df)], "_from")
+
+true_model_df_proj <- dplyr::left_join(true_model_df_proj, projected_true_model_df, by = c("to" = "ID"))
+names(true_model_df_proj)[(2 + NCOL(projected_true_model_df)):NCOL(true_model_df_proj)] <- paste0(names(projected_true_model_df)[-NCOL(projected_true_model_df)], "_to")
+
+two_curvy_proj_true2 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_segment(
+    data = true_model_df_proj, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#2b8cbe") +
+  geom_point(
+    size = 0.5,
+    alpha = 0.2,
+    color = "#636363") +
+  geom_segment(
+    data=axes,
+    aes(x=x1, y=y1, xend=x2, yend=y2),
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2), 
+    label=rownames(axes),
+    colour="grey50",
+    size = 4) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.6, 0.6)) + 
+  ylim(c(-0.6, 0.6)) +
+  interior_annotation("b", sc_ltr_pos, cex = 1.5) +
+  theme(
+    legend.position = "none"
+  )
+
 ```
@@
 
-two_curvy_proj_umap_model1 + two_curvy_proj_umap_model2 + 
+two_curvy_proj_true + two_curvy_proj_true2 +
+#two_curvy_proj_umap_model1 + two_curvy_proj_umap_model2 + 
   plot_layout(ncol = 2)

### 2025-03-19 paper.qmd
Fixed Figure 2

@@
 
-projected_true_model <- as.matrix(scaled_two_curvy_data_true_model) %*% projection_scaled
+projected_true_model <- as.matrix(scaled_two_curvy_data_true_model) %*% projection
 
@@
   coord_fixed() +
-  xlim(c(-0.6, 0.6)) + 
-  ylim(c(-0.6, 0.6)) +
+  xlim(c(-0.42, 0.42)) +
+  ylim(c(-0.42, 0.42)) +
   interior_annotation("b", sc_ltr_pos, cex = 1.5) +
@@
 <!-- The remaining variables $X_4, X_5, X_6, X_7$ are all uniform error, with small variance.  -->
-
-```{r}
-#| label: two-curvy-true-proj
-# Apply the scaling
-df_model_data_two_curvy <- bind_rows(data_two_curvy, true_model_two_curvy)
-
-scaled_two_curvy <- scale_data_manual(df_model_data_two_curvy, "type") |>
-  as_tibble()
-
-scaled_two_curvy_data <- scaled_two_curvy |>
-  filter(type == "data") |>
-  select(-type)
-
-scaled_two_curvy_data_model <- scaled_two_curvy |>
-  filter(type == "true model") |>
-  select(-type)
-
-## First projection
-projection <- cbind(
-  c(0.37847,-0.06517,0.04231,0.01655),
-  c(-0.06466,-0.27052,0.22131,-0.15235))
-
-proj_obj1 <- get_projection(projection = projection, 
-               proj_scale = 1, 
-               scaled_data = scaled_two_curvy_data, 
-               scaled_data_model = scaled_two_curvy_data_model, 
-               distance_df_small_edges = distance_df_small_edges_two_curvy, 
-               axis_param = list(limits = 0.6, 
-                                 axis_scaled = 2,
-                                 axis_pos_x = -0.4, 
-                                 axis_pos_y = -0.4, 
-                                 threshold = 0.04))
-
-projected_df <- proj_obj1$projected_df
-model_df <- proj_obj1$model_df
-axes <- proj_obj1$axes
-circle <- proj_obj1$circle
-
-# projection_scaled <- projection * 1
-# 
-# projected <- as.matrix(scaled_two_curvy_data) %*% projection_scaled
-# 
-# projected_df <- projected |>
-#   tibble::as_tibble(.name_repair = "unique") |>
-#   dplyr::rename(c("proj1" = "...1",
-#                   "proj2" = "...2")) |>
-#   #dplyr::mutate(type = df_exe$type) |>
-#   dplyr::mutate(ID = dplyr::row_number()) 
-# 
-# projected_model <- as.matrix(scaled_two_curvy_data_model) %*% projection_scaled
-# 
-# projected_model_df <- projected_model |>
-#   tibble::as_tibble(.name_repair = "unique") |>
-#   dplyr::rename(c("proj1" = "...1",
-#                   "proj2" = "...2")) |>
-#   dplyr::mutate(ID = dplyr::row_number())
-# 
-# model_df <- dplyr::left_join(
-#   wireframe_true_model, 
-#   projected_model_df, 
-#   by = c("from" = "ID"))
-# 
-# names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-# 
-# model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-# names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-# 
-# axes_obj <- gen_axes(  
-#   proj = projection * 2,
-#   limits = 0.6,
-#   axis_pos_x = -0.4,
-#   axis_pos_y = -0.4,
-#   axis_labels = names(scaled_two_curvy_data),
-#   threshold = 0.04)
-# 
-# axes <- axes_obj$axes
-# circle <- axes_obj$circle
-
-two_curvy_proj_umap_model1 <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#2b8cbe") +
-  geom_point(
-    size = 0.5,
-    alpha = 0.5,
-    color = "#636363") +
-  geom_segment(
-    data=axes,
-    aes(x=x1, y=y1, xend=x2, yend=y2),
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 5) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.6, 0.6)) + 
-  ylim(c(-0.6, 0.6)) +
-  interior_annotation("a", c(0.08, 0.93), cex = 1.5) +
-  theme(
-    legend.position = "none"
-  )
-
-proj_plot_two_curvy_true <- model_df |>
-  ggplot(
-    aes(
-      x = proj1_from, 
-      y = proj2_from)) +
-  geom_point(
-    color = "#000000",
-    alpha = 0.5) +
-  geom_segment(
-    data=axes,
-    aes(x=x1, y=y1, xend=x2, yend=y2),
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 4) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.6, 0.6)) + 
-  ylim(c(-0.6, 0.6)) +
-  interior_annotation("a", c(0.08, 0.93), cex = 1.5) +
-  theme(
-    legend.position = "none"
-  )
-
-## Second projection
-projection <- cbind(
-  c(-0.38281,-0.04870,0.00619,-0.02451),
-  c(-0.01303,0.00634,-0.37421,0.09648))
-
-proj_obj2 <- get_projection(projection = projection, 
-               proj_scale = 1, 
-               scaled_data = scaled_two_curvy_data, 
-               scaled_data_model = scaled_two_curvy_data_model, 
-               distance_df_small_edges = distance_df_small_edges_two_curvy, 
-               axis_param = list(limits = 0.5, 
-                                 axis_scaled = 6,
-                                 axis_pos_x = -0.3, 
-                                 axis_pos_y = -0.3, 
-                                 threshold = 0))
-
-projected_df <- proj_obj2$projected_df
-model_df <- proj_obj2$model_df
-axes <- proj_obj2$axes
-circle <- proj_obj2$circle
-
-two_curvy_proj_umap_model2 <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#2b8cbe") +
-  geom_point(
-    size = 0.5,
-    alpha = 0.5,
-    color = "#636363") +
-  geom_segment(
-    data=axes,
-    aes(x=x1, y=y1, xend=x2, yend=y2),
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 4) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.42, 0.42)) + 
-  ylim(c(-0.42, 0.42)) +
-  interior_annotation("b", cex = 1.5) +
-  theme(
-    legend.position = "none"
-  )
-```
 
@@
 two_curvy_proj_true + two_curvy_proj_true2 +
-#two_curvy_proj_umap_model1 + two_curvy_proj_umap_model2 + 
   plot_layout(ncol = 2)

### 2025-03-19 paper.qmd
Removed unnecessry code

@@
 
-trimesh_two_curvy_removed1_umap_with_data <- ggplot() +
+trimesh_two_curvy_removed1_tsne_with_data <- ggplot() +
   geom_segment(data = tr_from_to_df_two_curvy,
@@
 ```{r}
-#| label: best-fit-two-curvy-proj-umap1
+#| label: best-fit-two-curvy-proj-tsne
 
@@
 
-# axes_obj <- gen_axes(  
-#   proj = projection * 2,
-#   limits = 0.5,
-#   axis_pos_x = -0.5,
-#   axis_pos_y = -0.5,
-#   axis_labels = names(scaled_two_curvy_data),
-#   threshold = 0.03)
-# 
-# axes <- axes_obj$axes
-# circle <- axes_obj$circle
-
-two_curvy_proj_umap_first_model1 <- projected_df |>
+two_curvy_proj_tsne_first_model1 <- projected_df |>
   ggplot(
@@
 
-two_curvy_proj_umap_first_model1_temp <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#000000") + #31a354
-  geom_point(
-    size = 0.5,
-    alpha = 0.2,
-    color = "#636363") +
-  geom_segment(
-    data=axes,
-    aes(x=x1, y=y1, xend=x2, yend=y2),
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 4) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.6, 0.6)) + 
-  ylim(c(-0.6, 0.6)) +
-  interior_annotation("c2", sc_ltr_pos) +
-  theme(
-    legend.position = "none"
-  )
-
 two_curvy_proj_true <- projected_df |>
@@
 
-two_curvy_proj_umap_all_model1 <- projected_df |>
+two_curvy_proj_tsne_all_model1 <- projected_df |>
   ggplot(
@@
   interior_annotation("d", sc_ltr_pos, cex = 1.5) +
-  theme(
-    legend.position = "none"
-  )
-
-## Projection with Delauany triangulation
-
-two_curvy_proj_umap_model_delaunay <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_segment(
-    data = model_df1, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#000000") + #31a354
-  geom_point(
-    size = 0.5,
-    alpha = 0.2,
-    color = "#636363") +
-  geom_segment(
-    data=axes,
-    aes(x=x1, y=y1, xend=x2, yend=y2),
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 4) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.1, 0.1)) + 
-  ylim(c(-0.1, 0.1)) +
-  interior_annotation("a2", sc_ltr_pos) +
-  theme(
-    legend.position = "none"
-  )
-
-two_curvy_proj_umap_model_benchmark2 <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_segment(
-    data = model_df2, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#000000") + #31a354
-  geom_point(
-    size = 0.5,
-    alpha = 0.2,
-    color = "#636363") +
-  geom_segment(
-    data=axes,
-    aes(x=x1, y=y1, xend=x2, yend=y2),
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 4) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.1, 0.1)) + 
-  ylim(c(-0.1, 0.1)) +
-  interior_annotation("b2", sc_ltr_pos) +
   theme(
@@
 ```{r}
-#| label: best-fit-umap1
+#| label: best-fit-tsne
 #| fig-cap: "We start with a tSNE layout (a) of non-linear clusters data ($n =  2000$ and $p = 4$), from which the \\gD{} wireframe is constructed (b) and then lifted to $4\\text{-}D$.
@@
 nldr_two_curvy2 + 
-  trimesh_two_curvy_removed1_umap_with_data +
-  two_curvy_proj_umap_first_model1 +
-  two_curvy_proj_umap_all_model1 +
+  trimesh_two_curvy_removed1_tsne_with_data +
+  two_curvy_proj_tsne_first_model1 +
+  two_curvy_proj_tsne_all_model1 +
   plot_layout(ncol = 2)

### 2025-03-19 paper.qmd
Updated the tru data projection with the new function

@@
                       point_param = c(1.5, 0.5, "#000000"), # size, alpha, color
-                      line_param = c(0.5, 0.5), #linewidth, alpha
+                      line_param = c(0.5, 0.5, "#000000"), #linewidth, alpha
                       plot_limits, title, cex = 2, 
@@
         yend = proj2_to), 
-      color = "#000000",
-      linewidth = line_param[1],
-      alpha = line_param[2]) +
+      color = line_param[3],
+      linewidth = as.numeric(line_param[1]),
+      alpha = as.numeric(line_param[2])) +
     geom_point(
@@
 
-two_curvy_proj_true <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_segment(
-    data = true_model_df_proj, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#2b8cbe") +
-  geom_point(
-    size = 0.5,
-    alpha = 0.2,
-    color = "#636363") +
-  geom_segment(
-    data=axes,
-    aes(x=x1, y=y1, xend=x2, yend=y2),
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 4) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.6, 0.6)) + 
-  ylim(c(-0.6, 0.6)) +
-  interior_annotation("a", sc_ltr_pos, cex = 1.5) +
+two_curvy_proj_true <- plot_proj(
+  projected_df = projected_df, 
+  model_df = true_model_df_proj, 
+  axes = axes, 
+  circle = circle, 
+  point_param = c(0.5, 0.5, "#636363"), # size, alpha, color
+  line_param = c(0.5, 0.5, "#2b8cbe"), # linewidth, alpha
+  plot_limits = c(-0.6, 0.6), 
+  title = "a", 
+  cex = 1.5, 
+  axis_text_size = 4, 
+  is_category = FALSE) +
   theme(
@@
 
-two_curvy_proj_true2 <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1,
-      y = proj2)) +
-  geom_segment(
-    data = true_model_df_proj,
-    aes(
-      x = proj1_from,
-      y = proj2_from,
-      xend = proj1_to,
-      yend = proj2_to),
-    color = "#2b8cbe") +
-  geom_point(
-    size = 0.5,
-    alpha = 0.2,
-    color = "#636363") +
-  geom_segment(
-    data=axes,
-    aes(x=x1, y=y1, xend=x2, yend=y2),
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2),
-    label=rownames(axes),
-    colour="grey50",
-    size = 4) +
-  geom_path(
-    data=circle,
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.42, 0.42)) +
-  ylim(c(-0.42, 0.42)) +
-  interior_annotation("b", sc_ltr_pos, cex = 1.5) +
+two_curvy_proj_true2 <- plot_proj(
+    projected_df = projected_df, 
+    model_df = true_model_df_proj, 
+    axes = axes, 
+    circle = circle, 
+    point_param = c(0.5, 0.5, "#636363"), # size, alpha, color
+    line_param = c(0.5, 0.5, "#2b8cbe"), # linewidth, alpha
+    plot_limits = c(-0.42, 0.42), 
+    title = "b", 
+    cex = 1.5, 
+    axis_text_size = 4, 
+    is_category = FALSE) +
     theme(

### 2025-03-19 paper.qmd
Updated Figure 5 with projection function

@@
                       plot_limits, title, cex = 2, 
+                      position = c(0.92, 0.92), 
                       axis_text_size = 3,
@@
 
-  initial_plot + 
+  initial_plot <- initial_plot + 
     geom_segment(
@@
     ylim(plot_limits) +
-    interior_annotation(title, cex = cex)
+    interior_annotation(title, position, cex = cex)
+
+  initial_plot
 
@@
 
-two_curvy_proj_tsne_first_model1 <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#000000") + #31a354
-  geom_point(
-    size = 0.5,
-    alpha = 0.2,
-    color = "#636363") +
-  geom_segment(
-    data=axes,
-    aes(x=x1, y=y1, xend=x2, yend=y2),
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 4) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.6, 0.6)) + 
-  ylim(c(-0.6, 0.6)) +
-  interior_annotation("c", sc_ltr_pos, cex = 1.5) +
+two_curvy_proj_tsne_first_model1 <- plot_proj(
+    projected_df = projected_df, 
+    model_df = model_df, 
+    axes = axes, 
+    circle = circle, 
+    point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
+    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
+    plot_limits = c(-0.6, 0.6), 
+    title = "c", 
+    cex = 1.5, 
+    position = sc_ltr_pos,
+    axis_text_size = 4, 
+    is_category = FALSE) +
     theme(
@@
 
-two_curvy_proj_tsne_all_model1 <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
+two_curvy_proj_tsne_all_model1 <- plot_proj(
+  projected_df = projected_df, 
+  model_df = true_model_df_proj, 
+  axes = axes, 
+  circle = circle, 
+  point_param = c(0.5, 0.5, "#636363"), # size, alpha, color
+  line_param = c(0.5, 0.5, "#2b8cbe"), # linewidth, alpha
+  plot_limits = c(-0.6, 0.6), 
+  title = "d", 
+  cex = 1.5, 
+  position = sc_ltr_pos,
+  axis_text_size = 4, 
+  is_category = FALSE) +
   geom_segment(
-    data = true_model_df_proj, 
+    data = model_df, 
     aes(
@@
       yend = proj2_to), 
-    color = "#2b8cbe") +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#000000") +
-  geom_point(
-    size = 0.5,
-    alpha = 0.2,
-    color = "#636363") +
-  geom_segment(
-    data=axes,
-    aes(x=x1, y=y1, xend=x2, yend=y2),
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 4) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.6, 0.6)) + 
-  ylim(c(-0.6, 0.6)) +
-  interior_annotation("d", sc_ltr_pos, cex = 1.5) +
-  theme(
-    legend.position = "none"
-  )
+    color = "#000000")
 

### 2025-03-19 paper.qmd
Fixed Figure 7a

@@
 #| label: hexbin-regular-two-curvy2
-## hexagon binning to have regular hexagons
-hb_obj_two_curvy2 <- hex_binning(
-  data = tsne_two_curvy_scaled, 
-  bin1 = 15, 
-  r2 = r2)
-
-a1_2 <- calc_bins_y(
-  bin1 = 15, 
-  r2 = r2
-  )$a1
-
-## Data set with all centroids
-all_centroids_df2 <- hb_obj_two_curvy2$centroids
-
-## Generate all coordinates of hexagons
-hex_grid2 <- hb_obj_two_curvy2$hex_poly
-
-## To obtain the standardise counts within hexbins
-counts_df2 <- hb_obj_two_curvy2$std_cts
-df_bin_centroids2 <- extract_hexbin_centroids(
-  centroids_df = all_centroids_df2, 
-  counts_df = counts_df2) |>
-  filter(drop_empty == FALSE) |>
-  mutate(b1 = "b1 = 11")
-
-hex_grid_with_counts_two_curvy2 <- full_join(
-  hex_grid2, 
-  df_bin_centroids2 |> select(hexID, std_counts), 
-  by = c("hex_poly_id" = "hexID"))
-
-umap_data_two_curvy2_with_hb_id <- hb_obj_two_curvy2$data_hb_id
-df_all_two_curvy2 <- dplyr::bind_cols(training_data_two_curvy, umap_data_two_curvy2_with_hb_id)
-df_bin_two_curvy2 <- avg_highd_data(data = df_all_two_curvy2, col_start = "x")
-
-df_b_two_curvy2 <- df_bin_two_curvy2 |>
-  dplyr::filter(hb_id %in% df_bin_centroids2$hexID) |>
-  dplyr::mutate(type = "model") ## Data with summarized mean
-
-## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
-df_b_two_curvy2 <- df_b_two_curvy2[match(df_bin_centroids2$hexID, df_b_two_curvy2$hb_id),] |>
-  dplyr::select(-hb_id)
+
+num_bins_x_two_curvy <- 15
+
+algo_obj_two_curvy2 <- gen_nldr_vis_algo_obj(
+  high_d_data = training_data_two_curvy, 
+  nldr_data = tsne_two_curvy, 
+  num_x_bins = num_bins_x_two_curvy,
+  col_start = "x")
+
+tsne_two_curvy_scaled <- algo_obj_two_curvy2$nldr_scaled
+distance_two_curvy2 <- algo_obj_two_curvy2$distance_df
+tr_from_to_df_two_curvy2 <- algo_obj_two_curvy2$tr_from_to_df
+benchmark_two_curvy2 <- algo_obj_two_curvy2$benchmark
+df_bin_centroids_two_curvy2 <- algo_obj_two_curvy2$df_bin_centroids
+df_bin_two_curvy2 <- algo_obj_two_curvy2$df_bin
+hex_grid_with_counts_two_curvy2 <- algo_obj_two_curvy2$hex_grid_with_counts
+
+distance_df_small_edges_two_curvy2 <- distance_two_curvy2 |>
+  filter(distance < benchmark_two_curvy2)
+
+tr_from_to_df_two_curvy2 <- inner_join(
+  tr_from_to_df_two_curvy2, distance_df_small_edges_two_curvy2, 
+  by = c("from", "to"))
 
@@
   interior_annotation("a", sc_ltr_pos, cex = 1.5)
-
-## Triangulate bin centroids
-tr1_object_two_curvy2 <- tri_bin_centroids(
-  df_bin_centroids2, x = "c_x", y = "c_y")
-tr_from_to_df_two_curvy2 <- gen_edges(
-  tri_object = tr1_object_two_curvy2)
 
@@
   interior_annotation("a1", sc_ltr_pos)
-
-## Compute 2D distances
-distance_two_curvy <- cal_2d_dist(
-  tr_coord_df = tr_from_to_df_two_curvy2, 
-  start_x = "x_from", 
-  start_y = "y_from", 
-  end_x = "x_to", 
-  end_y = "y_to", 
-  select_vars = c("from", "to", "distance"))
-
-## To find the benchmark value
-benchmark_two_curvy <- find_lg_benchmark(
-  distance_edges = distance_two_curvy, 
-  distance_col = "distance")
-
-```
-
-```{r}
-#| label: best-fit-two-curvy-proj
-
-# Apply the scaling
-df_model_data_two_curvy <- bind_rows(
-  data_two_curvy, 
-  true_model_two_curvy, 
-  df_b_two_curvy2)
-
-scaled_two_curvy <- scale_data_manual(df_model_data_two_curvy, "type") |>
-  as_tibble()
-
-scaled_two_curvy_data <- scaled_two_curvy |>
-  filter(type == "data") |>
-  select(-type)
-
-scaled_two_curvy_data_true_model <- scaled_two_curvy |>
-  filter(type == "true model") |>
-  select(-type)
-
-scaled_two_curvy_data_model <- scaled_two_curvy |>
-  filter(type == "model") |>
-  select(-type)
-
-## First projection
-projection <- cbind(
-    c(0.11062,-0.02743,0.02786,-0.01930),
-    c(-0.01110,-0.06886,0.04730,0.09890))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1, 
-                            scaled_data = scaled_gau_data, 
-                            scaled_data_model = scaled_gau_data_model, 
-                            distance_df_small_edges = distance_df_small_edges, 
-                            axis_param = list(limits = 0.15, 
-                                              axis_scaled = 5,
-                                              axis_pos_x = -0.067, 
-                                              axis_pos_y = -0.067, 
-                                              threshold = 0.007))
-
-projected_df <- proj_obj1$projected_df
-model_df <- proj_obj1$model_df
-axes <- proj_obj1$axes
-circle <- proj_obj1$circle
-
-# projection_scaled <- projection * 1
-# 
-# projected <- as.matrix(scaled_two_curvy_data) %*% projection_scaled
-# 
-# projected_df <- projected |>
-#   tibble::as_tibble(.name_repair = "unique") |>
-#   dplyr::rename(c("proj1" = "...1",
-#                   "proj2" = "...2")) |>
-#   dplyr::mutate(ID = dplyr::row_number()) 
-# 
-# ## For fitted model
-# projected_model <- as.matrix(scaled_two_curvy_data_model) %*% projection_scaled
-# 
-# projected_model_df <- projected_model |>
-#   tibble::as_tibble(.name_repair = "unique") |>
-#   dplyr::rename(c("proj1" = "...1",
-#                   "proj2" = "...2")) |>
-#   dplyr::mutate(ID = dplyr::row_number())
-# 
-# model_df <- dplyr::left_join(
-#   distance_df_small_edges_two_curvy2 |> select(-distance), 
-#   projected_model_df, 
-#   by = c("from" = "ID"))
-# 
-# names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-# 
-# model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-# names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-## For true model
-
-projected_true_model <- as.matrix(scaled_two_curvy_data_true_model) %*% projection_scaled
-
-projected_true_model_df <- projected_true_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-true_model_df_proj <- dplyr::left_join(
-  wireframe_true_model, 
-  projected_true_model_df, 
-  by = c("from" = "ID"))
-
-names(true_model_df_proj)[3:NCOL(true_model_df_proj)] <- paste0(names(projected_true_model_df)[-NCOL(projected_true_model_df)], "_from")
-
-true_model_df_proj <- dplyr::left_join(true_model_df_proj, projected_true_model_df, by = c("to" = "ID"))
-names(true_model_df_proj)[(2 + NCOL(projected_true_model_df)):NCOL(true_model_df_proj)] <- paste0(names(projected_true_model_df)[-NCOL(projected_true_model_df)], "_to")
-
-# axes_obj <- gen_axes(  
-#   proj = projection * 5,
-#   limits = 0.15,
-#   axis_pos_x = -0.067,
-#   axis_pos_y = -0.067,
-#   axis_labels = names(scaled_two_curvy_data),
-#   threshold = 0.007)
-# 
-# axes <- axes_obj$axes
-# circle <- axes_obj$circle
-
-two_curvy_proj_umap_model <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#000000") + #31a354
-  geom_point(
-    size = 0.5,
-    alpha = 0.2,
-    color = "#636363") +
-  geom_segment(
-    data=axes,
-    aes(x=x1, y=y1, xend=x2, yend=y2),
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 4) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.1, 0.1)) + 
-  ylim(c(-0.1, 0.1)) +
-  interior_annotation("c2", sc_ltr_pos) +
-  theme(
-    legend.position = "none"
-  )
-
-two_curvy_proj_umap_model2 <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#000000") + #31a354
-  geom_point(
-    size = 0.5,
-    alpha = 0.2,
-    color = "#636363") +
-  geom_segment(
-    data=axes,
-    aes(x=x1, y=y1, xend=x2, yend=y2),
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 5) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.1, 0.1)) + 
-  ylim(c(-0.1, 0.1)) +
-  interior_annotation("a3", sc_ltr_pos) +
-  theme(
-    legend.position = "none"
-  )
-
-two_curvy_proj_umap_all_model <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_segment(
-    data = true_model_df_proj, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#2b8cbe") +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#000000") +
-  geom_point(
-    size = 0.5,
-    alpha = 0.5,
-    color = "#636363") +
-  geom_segment(
-    data=axes,
-    aes(x=x1, y=y1, xend=x2, yend=y2),
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 4) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.1, 0.1)) + 
-  ylim(c(-0.1, 0.1)) +
-  interior_annotation("a4", sc_ltr_pos) +
-  theme(
-    legend.position = "none"
-  )
 

### 2025-03-19 paper.qmd
Fixed the issue of Figure 7

@@
 
-tsne_two_curvy_scaled <- algo_obj_two_curvy2$nldr_scaled
 distance_two_curvy2 <- algo_obj_two_curvy2$distance_df
@@
 #| label: hexbin-regular-two-curvy3
-## hexagon binning to have regular hexagons
-hb_obj_two_curvy3 <- hex_binning(
-  data = tsne_two_curvy_scaled, 
-  bin1 = 48, 
-  r2 = r2)
-
-a1_3 <- calc_bins_y(
-  bin1 = 48, 
-  r2 = r2
-)$a1
-
-## Data set with all centroids
-all_centroids_df3 <- hb_obj_two_curvy3$centroids
-
-## Generate all coordinates of hexagons
-hex_grid3 <- hb_obj_two_curvy3$hex_poly
-
-## To obtain the standardise counts within hexbins
-counts_df3 <- hb_obj_two_curvy3$std_cts
-df_bin_centroids3 <- extract_hexbin_centroids(
-  centroids_df = all_centroids_df3, 
-  counts_df = counts_df3) |>
-  filter(drop_empty == FALSE) |>
-  mutate(b1 = "b1 = 14")
-
-boxplot3 <- ggplot(df_bin_centroids3, 
-                   aes(x=bin_counts)) + 
-    geom_histogram()
-
-hex_grid_with_counts_s_curve3 <- full_join(hex_grid3, 
-                                          df_bin_centroids3 |> select(hexID, std_counts), 
-                                          by = c("hex_poly_id" = "hexID"))
+
+num_bins_x_two_curvy <- 48
+
+algo_obj_two_curvy3 <- gen_nldr_vis_algo_obj(
+  high_d_data = training_data_two_curvy, 
+  nldr_data = tsne_two_curvy, 
+  num_x_bins = num_bins_x_two_curvy,
+  col_start = "x")
+
+distance_two_curvy3 <- algo_obj_two_curvy3$distance_df
+tr_from_to_df_two_curvy3 <- algo_obj_two_curvy3$tr_from_to_df
+benchmark_two_curvy3 <- algo_obj_two_curvy3$benchmark
+df_bin_centroids_two_curvy3 <- algo_obj_two_curvy3$df_bin_centroids
+df_bin_two_curvy3 <- algo_obj_two_curvy3$df_bin
+hex_grid_with_counts_two_curvy3 <- algo_obj_two_curvy3$hex_grid_with_counts
+
+distance_df_small_edges_two_curvy3 <- distance_two_curvy3 |>
+  filter(distance < benchmark_two_curvy3)
+
+tr_from_to_df_two_curvy3 <- inner_join(
+  tr_from_to_df_two_curvy3, distance_df_small_edges_two_curvy3, 
+  by = c("from", "to"))
 
@@
   geom_polygon(
-    data = hex_grid_with_counts_s_curve3, 
+    data = hex_grid_with_counts_two_curvy3, 
     aes(x = x, y = y, 

### 2025-03-19 paper.qmd
Updated Figure 6

@@
   interior_annotation("b", sc_ltr_pos, cex = 1.5)
+```
+
+```{r}
+#| label: compute-error-tsne
+
+training_data_two_curvy <- training_data_two_curvy |>
+  mutate(ID = row_number())
+
+## Compute error
+error_df_two_curvy_abs <- augment(
+  df_bin_centroids = df_bin_centroids_two_curvy,
+  df_bin = df_bin_two_curvy,
+  training_data = training_data_two_curvy,
+  newdata = NULL,
+  type_NLDR = "tSNE",
+  col_start = "x")
+
+error_df_two_curvy_abs <- error_df_two_curvy_abs |>
+  mutate(sqrt_row_wise_total_error = sqrt(row_wise_total_error))
+
+error_df_two_curvy_abs <- error_df_two_curvy_abs |>
+  bind_cols(tsne_two_curvy_scaled |>
+              select(-ID))
+
+error_plot_two_curvy <- error_df_two_curvy_abs |>
+  ggplot(aes(x = tSNE1,
+             y = tSNE2,
+             colour = sqrt_row_wise_total_error)) +
+  geom_point(alpha=0.5) +
+  scale_colour_continuous_sequential(palette = "YlOrRd", n_interp = 20) +
+  theme(
+    aspect.ratio = 1,
+    legend.position = "none"
+  )
+
+error_plot_two_curvy_hist <- 
+  ggplot(error_df_two_curvy_abs) +
+  geom_histogram(aes(x=sqrt_row_wise_total_error, y=..density..)) +
+  geom_density(aes(x=sqrt_row_wise_total_error, y=..density..), colour="red") +
+  #xlab(expression(group("|", e[hj], "|"))) +
+  xlab(expression(e[hj])) +
+  ylab("") +
+  theme_bw() 
 ```
@@
 axis.text = element_text(size = 12))
-
-training_data_two_curvy <- training_data_two_curvy |>
-  mutate(ID = row_number())
-
-## Compute error
-error_df_two_curvy_abs <- augment(
-  df_bin_centroids = df_bin_centroids2,
-  df_bin = df_bin_two_curvy2,
-  training_data = training_data_two_curvy,
-  newdata = NULL,
-  type_NLDR = "tSNE",
-  col_start = "x")
-
-error_df_two_curvy_abs <- error_df_two_curvy_abs |>
-  mutate(sqrt_row_wise_total_error = sqrt(row_wise_total_error))
-
-error_df_two_curvy_abs <- error_df_two_curvy_abs |>
-  bind_cols(tsne_two_curvy_scaled |>
-              select(-ID))
-
-error_plot_two_curvy <- error_df_two_curvy_abs |>
-  ggplot(aes(x = tSNE1,
-             y = tSNE2,
-             colour = sqrt_row_wise_total_error)) +
-  geom_point(alpha=0.5) +
-  scale_colour_continuous_sequential(palette = "YlOrRd", n_interp = 20) +
-  theme(
-    aspect.ratio = 1,
-    legend.position = "none"
-  )
-
-error_plot_two_curvy_hist <- 
-  ggplot(error_df_two_curvy_abs) +
-  geom_histogram(aes(x=sqrt_row_wise_total_error, y=..density..)) +
-  geom_density(aes(x=sqrt_row_wise_total_error, y=..density..), colour="red") +
-  #xlab(expression(group("|", e[hj], "|"))) +
-  xlab(expression(e[hj])) +
-  ylab("") +
-  theme_bw() 
 ```

### 2025-03-19 paper.qmd
Updated Figure 14 axes label size

@@
   cex = 2, 
+  axis_text_size = 5,
   is_category = FALSE)
@@
     cex = 2, 
+    axis_text_size = 5,
     is_category = FALSE)

### 2025-03-19 paper.qmd
Fixed issues of figure 17

@@
           x = proj1, 
-          y = proj2))
-
-  } else {
-
-    initial_plot <- projected_df |>
-      ggplot(
-        aes(
-          x = proj1, 
-          y = proj2,
-          colour = cluster))
-
-  }
-
-  initial_plot <- initial_plot + 
+          y = proj2)) + 
       geom_segment(
@@
       alpha = as.numeric(point_param[2]),
-      color = point_param[3]) +
+      color = point_param[3])
+
+  } else {
+
+    initial_plot <- projected_df |>
+      ggplot(
+        aes(
+          x = proj1, 
+          y = proj2,
+          colour = cluster)) + 
+      geom_segment(
+        data = model_df, 
+        aes(
+          x = proj1_from, 
+          y = proj2_from, 
+          xend = proj1_to, 
+          yend = proj2_to), 
+        color = line_param[3],
+        linewidth = as.numeric(line_param[1]),
+        alpha = as.numeric(line_param[2])) +
+      geom_point(
+      size = as.numeric(point_param[1]),
+      alpha = as.numeric(point_param[2]))
+
+  }
+
+  initial_plot <- initial_plot + 
     geom_segment(
@@
   circle = circle, 
-  point_param = c(0.5, 0.5, "#000000"), # size, alpha, color
-  line_param = c(0.5, 0.5), # linewidth, alpha
+  point_param = c(0.5, 0.5), # size, alpha, color
+  line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
+  plot_limits = c(-1.1, 0.9), 
+  title = "b", 
+  cex = 1.5, 
+  axis_text_size = 3, 
+  is_category = TRUE) +
+  scale_color_manual(values = c("#999999",'#ff7f00')) +
+  theme(
+    legend.position = "none"
+  )
+
+mnist_proj_tsne_model1_link <- plot_proj(
+    projected_df = projected_df, 
+    model_df = model_df, 
+    axes = axes, 
+    circle = circle, 
+    point_param = c(1.2, 0.5), # size, alpha, color
+    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
     plot_limits = c(-1.1, 0.9), 
@@
     scale_color_manual(values = c("#d9d9d9",'#ff7f00')) +
-  theme(
-    legend.position = "none"
-  )
-
-mnist_proj_tsne_model1_link <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2,
-      colour = cluster)) +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#000000",
-    linewidth = 0.5,
-    alpha = 0.5) +
-  geom_point(
-    size = 1.2,
-    alpha = 0.5) +
-  scale_color_manual(values = c("#d9d9d9",'#ff7f00')) +
-  geom_segment(
-    data=axes,
-    aes(x=x1, y=y1, xend=x2, yend=y2),
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 2) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  #coord_fixed() +
-  xlim(c(-1.1, 0.9)) +
-  ylim(c(-1.1, 0.9)) +
-  interior_annotation("b", cex = 2) +
     theme(
@@
 
-mnist_proj_tsne_model2 <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2,
-      colour = cluster)) +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#000000",
-    linewidth = 0.5) +
-  geom_point(
-    size = 0.5,
-    alpha = 0.3) +
+mnist_proj_tsne_model2 <- plot_proj(
+    projected_df = projected_df, 
+    model_df = model_df, 
+    axes = axes, 
+    circle = circle, 
+    point_param = c(0.5, 0.3), # size, alpha, color
+    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
+    plot_limits = c(-1, 1), 
+    title = "c", 
+    cex = 1.5, 
+    axis_text_size = 3, 
+    is_category = TRUE) +
     scale_color_manual(values = c("#999999",'#ff7f00')) +
-  geom_segment(
-    data=axes, 
-    aes(x=x1, y=y1, xend=x2, yend=y2), 
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 2) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-1, 1)) +
-  ylim(c(-1, 1)) +
-  interior_annotation("c", cex = 1.5) +
     theme(
@@
 
-mnist_proj_tsne_model2_link <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2,
-      colour = cluster)) +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#000000",
-    linewidth = 0.5) +
-  geom_point(
-    size = 1.2,
-    alpha = 0.3) +
+mnist_proj_tsne_model2_link <- plot_proj(
+    projected_df = projected_df, 
+    model_df = model_df, 
+    axes = axes, 
+    circle = circle, 
+    point_param = c(1.2, 0.5), # size, alpha, color
+    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
+    plot_limits = c(-1, 1), 
+    title = "c", 
+    cex = 2, 
+    axis_text_size = 3, 
+    is_category = TRUE) +
     scale_color_manual(values = c("#d9d9d9",'#ff7f00')) +
-  geom_segment(
-    data=axes, 
-    aes(x=x1, y=y1, xend=x2, yend=y2), 
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 2) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-1, 1)) +
-  ylim(c(-1, 1)) +
-  interior_annotation("c", cex = 2) +
     theme(
@@
     )
+
 ```

### 2025-03-19 paper.qmd
Plan to update the source script

@@
 source("scripts/one_c_shaped_dens_structure/04_gen_error_plots.R")
-source("scripts/five_gau_clusters/5_gen_model_with_tsne.R")
 source("scripts/two_non_linear_diff_shaped_close_clusters/10_gen_rm_lg_plots.R")

### 2025-03-20 paper.qmd
Updated Figure 12 with the upated code

@@
 ```{r}
+#| label: model-one-dens-clust
+
+one_c_shaped_data <- read_rds(here::here("data/one_c_shaped_dens_structure/one_c_shaped_dens_data.rds"))
+one_c_shaped_data <- one_c_shaped_data |>
+  mutate(ID = row_number())
+
+tsne_one_c_shaped <- read_rds(file = "data/one_c_shaped_dens_structure/one_c_shaped_dens_structure_tsne_perplexity_52.rds")
+
+num_bins_x_one_c_shaped <- 20
+
+algo_obj_dens_clust <- gen_nldr_vis_algo_obj(
+  high_d_data = one_c_shaped_data, 
+  nldr_data = tsne_one_c_shaped, 
+  num_x_bins = num_bins_x_one_c_shaped,
+  col_start = "x")
+
+tsne_one_c_shaped_scaled <- algo_obj_dens_clust$nldr_scaled
+distance_dens_clust <- algo_obj_dens_clust$distance_df
+tr_from_to_df_dens_clust <- algo_obj_dens_clust$tr_from_to_df
+benchmark_dens_clust <- algo_obj_dens_clust$benchmark
+df_bin_centroids_dens_clust <- algo_obj_dens_clust$df_bin_centroids
+df_bin_dens_clust <- algo_obj_dens_clust$df_bin
+hex_grid_with_counts_dens_clust <- algo_obj_dens_clust$hex_grid_with_counts
+```
+
+```{r}
+#| label: error-model-one-dens-clust
+## Compute error
+error_df_one_curvy_abs <- augment(
+  df_bin_centroids = df_bin_centroids_dens_clust,
+  df_bin = df_bin_dens_clust,
+  training_data = one_c_shaped_data,
+  newdata = NULL,
+  type_NLDR = "tSNE",
+  col_start = "x")
+
+error_df_one_curvy_abs <- error_df_one_curvy_abs |>
+  bind_cols(tsne_one_c_shaped_scaled |>
+              select(-ID))
+
+error_df_one_curvy_abs <- error_df_one_curvy_abs |>
+  mutate(sqrt_row_wise_total_error = sqrt(row_wise_total_error))
+
+# Compute density
+density_data <- density(error_df_one_curvy_abs$sqrt_row_wise_total_error)
+density_df <- data.frame(x = density_data$x, y = density_data$y)
+
+# Add density values to the original dataset
+error_df_one_curvy_abs <- error_df_one_curvy_abs %>%
+  mutate(density = approx(density_df$x, density_df$y, xout = sqrt_row_wise_total_error)$y)
+
+error_df_one_curvy_abs <- error_df_one_curvy_abs |>
+  mutate(error_cat_n = if_else(tSNE1 <= 0.25 & tSNE2 <= 0.25, "high", "low")) |> ## high_error points
+  mutate(error_cat_n2 = if_else(tSNE2 >= 1.2, "high", "low")) ## corner points
+
+error_plot_one_curvy_hist_selected <- ggplot(error_df_one_curvy_abs, aes(x = sqrt_row_wise_total_error, y = density,
+                                                                         colour = factor(error_cat_n,
+                                                                                         levels = c("high", "low")))) +
+  geom_point() +
+  scale_color_manual(values=c('#800026', '#d9d9d9')) +
+  xlab(expression(e[hj])) +
+  ylab("") +
+  interior_annotation("b1",
+                      position = c(0.95, 0.9),
+                      cex = 1.5) +
+  theme_bw() +
+  theme(
+    axis.title = element_text(size = 18),
+    axis.text = element_text(size = 15)
+  )
+
+error_plot_one_curvy_hist <- ggplot(error_df_one_curvy_abs,
+                                    aes(x = sqrt_row_wise_total_error, y = density,
+                                        colour = factor(error_cat_n2,
+                                                        levels = c("high", "low")))) +
+  geom_point() +
+  scale_color_manual(values=c('#800026', '#d9d9d9')) +
+  xlab(expression(e[hj])) +
+  ylab("") +
+  interior_annotation("a1",
+                      position = c(0.95, 0.9),
+                      cex = 1.5) +
+  theme_bw() +
+  theme(
+    axis.title = element_text(size = 18),
+    axis.text = element_text(size = 15)
+  )
+```
+
+```{r}
+#| label: link-model-one-dens-clust
+plot_tsne_dens <- error_df_one_curvy_abs |>
+  ggplot(aes(x = tSNE1,
+             y = tSNE2,
+             colour = factor(error_cat_n2,
+                             levels = c("high", "low")))) +
+  geom_point(alpha=0.7) +
+  scale_color_manual(values=c('#800026', '#d9d9d9')) +
+  theme(
+    aspect.ratio = 1
+  ) +
+  interior_annotation("a2",
+                      position = c(0.08, 0.9),
+                      cex = 1.5)
+
+plot_tsne_dens_selected <- error_df_one_curvy_abs |>
+  ggplot(aes(x = tSNE1,
+             y = tSNE2,
+             colour = factor(error_cat_n,
+                             levels = c("high", "low")))) +
+  geom_point(alpha=0.7) +
+  scale_color_manual(values=c('#800026', '#d9d9d9')) +
+  theme(
+    aspect.ratio = 1
+  ) +
+  interior_annotation("b2",
+                      position = c(0.08, 0.9),
+                      cex = 1.5)
+
+distance_df_small_edges_c_shaped_structure <- distance_dens_clust |>
+  filter(distance < benchmark_dens_clust)
+
+tr_from_to_df_dens_clust <- inner_join(
+  tr_from_to_df_dens_clust, distance_df_small_edges_c_shaped_structure,
+  by = c("from", "to"))
+
+```
+
+```{r}
+#| label: proj-one-dens-clust
+
+# Apply the scaling
+
+data_c_shaped <- one_c_shaped_data |>
+  select(-ID) |>
+  mutate(type = "data")
+
+df_bin_dens_clust <- df_bin_dens_clust |>
+  select(-hb_id) |>
+  mutate(type = "model")
+
+df_model_data <- bind_rows(data_c_shaped, df_bin_dens_clust)
+scaled_c_shaped <- scale_data_manual(df_model_data, "type") |>
+  as_tibble()
+
+scaled_c_shaped_data <- scaled_c_shaped |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_c_shaped_data_model <- scaled_c_shaped |>
+  filter(type == "model") |>
+  select(-type)
+
+## First projection
+projection <- cbind(
+  c(0.05096,0.15399,0.19736,0.05110),
+  c(0.14608,-0.16929,0.11291,-0.07160))
+
+proj_obj1 <- get_projection(projection = projection,
+                            proj_scale = 1,
+                            scaled_data = scaled_c_shaped_data,
+                            scaled_data_model = scaled_c_shaped_data_model,
+                            distance_df_small_edges = distance_df_small_edges_c_shaped_structure,
+                            axis_param = list(limits = 0.35,
+                                              axis_scaled = 2,
+                                              axis_pos_x = -0.28,
+                                              axis_pos_y = -0.28,
+                                              threshold = 0.017))
+
+projected_df <- proj_obj1$projected_df
+model_df <- proj_obj1$model_df
+axes2 <- proj_obj1$axes
+circle <- proj_obj1$circle
+
+## To add error category
+projected_df <- projected_df |>
+  mutate(cluster = error_df_one_curvy_abs$error_cat_n2)
+
+five_c_shaped_proj_tsne_model1 <- plot_proj(
+    projected_df = projected_df, 
+    model_df = model_df, 
+    axes = axes2, 
+    circle = circle, 
+    point_param = c(1.5, 0.5), # size, alpha, color
+    line_param = c(0.8, 0.4, "#000000"), # linewidth, alpha
+    plot_limits = c(-0.35, 0.35), 
+    title = "a3", 
+    cex = 1.5,
+    position = c(0.08, 0.9),
+    axis_text_size = 4, 
+    is_category = TRUE) +
+    scale_color_manual(values = c('#800026', '#d9d9d9')) +
+    theme(
+        legend.position = "none"
+    )
+
+projected_df <- projected_df |>
+  select(-cluster) |>
+  mutate(cluster = error_df_one_curvy_abs$error_cat_n)
+
+five_c_shaped_proj_tsne_model1_selected <- plot_proj(
+    projected_df = projected_df, 
+    model_df = model_df, 
+    axes = axes2, 
+    circle = circle, 
+    point_param = c(1.5, 0.5), # size, alpha, color
+    line_param = c(0.8, 0.4, "#000000"), # linewidth, alpha
+    plot_limits = c(-0.35, 0.35), 
+    title = "a3", 
+    cex = 1.5,
+    position = c(0.08, 0.9),
+    axis_text_size = 4, 
+    is_category = TRUE) +
+    scale_color_manual(values = c('#800026', '#d9d9d9')) +
+    theme(
+        legend.position = "none"
+    )
+```
+
+```{r}
 #| label: fig-one-dens_clust-error
@@
 
-generate_error_plots_one_c_shaped()
+error_plot_one_curvy_hist + plot_tsne_dens + five_c_shaped_proj_tsne_model1 +
+  error_plot_one_curvy_hist_selected + plot_tsne_dens_selected + five_c_shaped_proj_tsne_model1_selected +
+    plot_layout(guides = "collect", nrow = 2) &
+    theme(legend.position='none')
 ```

### 2025-03-20 paper.qmd
Updated the source directory

@@
 ```{r}
+#| label: std-error
+
+# creating Standardization function
+standardize = function(x){
+  z <- (x - mean(x)) / sd(x)
+  return( z)
+}
+```
+
+```{r}
 #| label: import-scripts
-source("scripts/one_c_shaped_dens_structure/04_gen_error_plots.R")
 source("scripts/two_non_linear_diff_shaped_close_clusters/10_gen_rm_lg_plots.R")

### 2025-03-20 paper.qmd
Updated Figure 9

@@
 ```{r}
+#| label: model-two-curvy-umap
+
+umap_two_curvy <- read_rds(file = "data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_umap_n-neigbors_15_min-dist_0.1.rds")
+
+num_bins_x_two_curvy <- 42
+
+algo_obj_two_curvy3 <- gen_nldr_vis_algo_obj(
+  high_d_data = training_data_two_curvy, 
+  nldr_data = umap_two_curvy, 
+  num_x_bins = num_bins_x_two_curvy,
+  col_start = "x")
+
+umap_two_curvy_scaled <- algo_obj_two_curvy3$nldr_scaled
+distance_two_curvy <- algo_obj_two_curvy3$distance_df
+tr_from_to_df_two_curvy <- algo_obj_two_curvy3$tr_from_to_df
+benchmark_two_curvy <- algo_obj_two_curvy3$benchmark
+df_bin_centroids_two_curvy <- algo_obj_two_curvy3$df_bin_centroids
+df_bin_two_curvy <- algo_obj_two_curvy3$df_bin
+hex_grid_with_counts <- algo_obj_two_curvy3$hex_grid_with_counts
+a1 <- algo_obj_two_curvy3$a1
+
+trimesh_two_curvy_umap <- ggplot() +
+  geom_segment(data = tr_from_to_df_two_curvy,
+               aes(
+                 x = x_from,
+                 y = y_from,
+                 xend = x_to,
+                 yend = y_to),
+               colour = "#000000",
+               alpha = 0.5,
+               size = 0.5) +
+  geom_point(data = umap_two_curvy_scaled,
+             aes(
+               x = UMAP1,
+               y = UMAP2
+             ),
+             color = "#636363",
+             alpha = 0.5,
+             size = 0.5
+  ) +
+  coord_equal() +
+  xlim(c(0, 1)) +
+  ylim(c(0, 1)) +
+  interior_annotation("a1", sc_ltr_pos, cex = 2)
+
+```
+
+```{r}
+
+distance_df_small_edges_two_curvy <- distance_two_curvy |>
+  filter(distance < benchmark_two_curvy)
+
+####With diff benchmark values
+
+## Benchmark 2
+
+benchmark_two_curvy <- 7 * a1
+
+distance_df_small_edges_two_curvy3 <- distance_two_curvy |>
+  filter(distance < benchmark_two_curvy)
+
+tr_from_to_df_two_curvy4 <- inner_join(
+  tr_from_to_df_two_curvy, distance_df_small_edges_two_curvy3,
+  by = c("from", "to"))
+
+trimesh_two_curvy_removed2_umap <- ggplot() +
+  geom_segment(data = tr_from_to_df_two_curvy4,
+               aes(
+                 x = x_from,
+                 y = y_from,
+                 xend = x_to,
+                 yend = y_to),
+               colour = "#000000",
+               alpha = 0.5,
+               size = 0.5)+
+  geom_point(data = umap_two_curvy_scaled,
+             aes(
+               x = UMAP1,
+               y = UMAP2
+             ),
+             color = "#636363",
+             alpha = 0.5,
+             size = 0.5
+  ) +
+  coord_equal() +
+  xlim(c(0, 1)) +
+  ylim(c(0, 1)) +
+  interior_annotation("b1", sc_ltr_pos, cex = 2)
+
+## Benchmark 1
+
+benchmark_two_curvy <- 4 * bin_width
+
+distance_df_small_edges_two_curvy2 <- distance_two_curvy |>
+  filter(distance < benchmark_two_curvy) ## 0.231
+
+tr_from_to_df_two_curvy3 <- inner_join(
+  tr_from_to_df_two_curvy, distance_df_small_edges_two_curvy2,
+  by = c("from", "to"))
+
+trimesh_two_curvy_removed1_umap <- ggplot() +
+  geom_segment(data = tr_from_to_df_two_curvy3,
+               aes(
+                 x = x_from,
+                 y = y_from,
+                 xend = x_to,
+                 yend = y_to),
+               colour = "#000000",
+               alpha = 0.5,
+               size = 0.5)+
+  geom_point(data = umap_two_curvy_scaled,
+             aes(
+               x = UMAP1,
+               y = UMAP2
+             ),
+             color = "#636363",
+             alpha = 0.5,
+             size = 0.5
+  ) +
+  coord_equal() +
+  xlim(c(0, 1)) +
+  ylim(c(0, 1)) +
+  interior_annotation("c1", sc_ltr_pos, cex = 2)
+```
+
+```{r}
+#| label: compare-2d-and-4d-dist
+
+### Compute highD distance
+dist_vec <- proxy::dist(x = df_bin_two_curvy[, -1], method = "Euclidean") |> as.vector()
+
+from_vec <- c()
+to_vec <- c()
+num_obs <- 1:(NROW(df_bin_two_curvy) - 1)
+
+for (obs in num_obs) {
+
+  from_val <- rep(obs, (NROW(df_bin_two_curvy) - obs))
+  if ((obs + 1) <= NROW(df_bin_two_curvy)) {
+    to_val <- (obs + 1):NROW(df_bin_two_curvy)
+  }
+  from_vec <- append(from_vec, from_val)
+  to_vec <- append(to_vec, to_val)
+
+}
+
+dist_highd <- tibble::tibble(from = from_vec, to = to_vec, dist_highd = dist_vec)
+
+## To plot the distribution of distances
+
+# Define bin width and starting point
+start_point <- a1/2
+
+# Create bins and calculate the mean value for each bin range
+bins <- seq(start_point, max(distance_two_curvy$distance) + a1,
+            by = a1)
+bin_labels <- bins[-length(bins)] + a1 / 2  # mean of each bin range
+
+distance_two_curvy_dist <- distance_two_curvy |>
+  mutate(bin = cut(distance, breaks = bins, include.lowest = TRUE, labels = bin_labels))
+
+distance_two_curvy_dist <- left_join(distance_two_curvy_dist, dist_highd, by = c("from", "to"))
+
+text_df <- tibble::tibble(
+  x = c(0.035, 0.15),
+  y = c(10, 10),
+  text = c("c", "b")
+)
+
+distance_points_umap <- ggplot(
+  distance_two_curvy_dist,
+  aes(x = distance,
+      y = dist_highd)) +
+  geom_point(alpha = 0.5) +
+  geom_text(
+    data=text_df,
+    aes(x=x, y=y,
+        label=text),
+    colour="#bdbdbd",
+    size = 10) +
+  geom_vline(xintercept = 6.5 * bin_width,
+             linetype="solid",
+             color = "#bdbdbd",
+             linewidth=1) +
+  geom_vline(xintercept = 2 * bin_width,
+             linetype="solid",
+             color = "#bdbdbd",
+             linewidth=1) +
+  ylab(expression(d^{(4)})) +
+  xlab(expression(d^{(2)})) +
+  theme_minimal() +
+  theme(aspect.ratio = 0.75,
+        panel.border = element_rect(fill = 'transparent'),
+        plot.title = element_text(size = 20, hjust = 0.5, vjust = -0.5),
+        panel.grid.major.x = element_blank(),
+        axis.ticks.x = element_line(),
+        axis.ticks.y = element_line(),
+        axis.title = element_text(size = 15),
+        axis.text = element_text(size = 12))
+```
+
+```{r}
+#| label: proj-tsne-model
+
+df_bin_two_curvy <- df_bin_two_curvy |>
+  select(-hb_id) |>
+  mutate(type = "model")
+
+# Apply the scaling
+df_model_data_two_curvy <- bind_rows(data_two_curvy, df_bin_two_curvy)
+
+scaled_two_curvy <- scale_data_manual(df_model_data_two_curvy, "type") |>
+  as_tibble()
+
+scaled_two_curvy_data <- scaled_two_curvy |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_two_curvy_data_model <- scaled_two_curvy |>
+  filter(type == "model") |>
+  select(-type)
+
+## First projection
+projection <- cbind(
+  c(0.37847,-0.06517,0.04231,0.01655),
+  c(-0.06466,-0.27052,0.22131,-0.15235))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1, 
+                            scaled_data = scaled_two_curvy_data, 
+                            scaled_data_model = scaled_two_curvy_data_model, 
+                            distance_df_small_edges = distance_df_small_edges_two_curvy, 
+                            axis_param = list(limits = 0.6, 
+                                              axis_scaled = 2,
+                                              axis_pos_x = -0.5, 
+                                              axis_pos_y = -0.5, 
+                                              threshold = 0.04))
+
+projected_df <- proj_obj1$projected_df
+model_df <- proj_obj1$model_df
+axes <- proj_obj1$axes
+circle <- proj_obj1$circle
+
+## Projection with Delauany triangulation
+
+model_df1 <- dplyr::left_join(
+  distance_two_curvy |> select(-distance),
+  projected_model_df,
+  by = c("from" = "ID"))
+
+names(model_df1)[3:NCOL(model_df1)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df1 <- dplyr::left_join(model_df1, projected_model_df, by = c("to" = "ID"))
+names(model_df1)[(2 + NCOL(projected_model_df)):NCOL(model_df1)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+## With benchmark option 2
+
+model_df2 <- dplyr::left_join(
+  distance_df_small_edges_two_curvy3 |> select(-distance),
+  projected_model_df,
+  by = c("from" = "ID"))
+
+names(model_df2)[3:NCOL(model_df2)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+
+model_df2 <- dplyr::left_join(model_df2, projected_model_df, by = c("to" = "ID"))
+names(model_df2)[(2 + NCOL(projected_model_df)):NCOL(model_df2)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+
+two_curvy_proj_first_model1_umap <- plot_proj(
+    projected_df = projected_df, 
+    model_df = model_df, 
+    axes = axes, 
+    circle = circle, 
+    point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
+    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
+    plot_limits = c(-0.6, 0.6), 
+    title = "c2", 
+    cex = 2, 
+    position = sc_ltr_pos,
+    axis_text_size = 4, 
+    is_category = FALSE) +
+    theme(
+        legend.position = "none"
+    )
+
+## Projection with Delauany triangulation
+
+two_curvy_proj_model_delaunay_umap <- plot_proj(
+    projected_df = projected_df, 
+    model_df = model_df1, 
+    axes = axes, 
+    circle = circle, 
+    point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
+    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
+    plot_limits = c(-0.6, 0.6), 
+    title = "a2", 
+    cex = 2, 
+    position = sc_ltr_pos,
+    axis_text_size = 4, 
+    is_category = FALSE) +
+    theme(
+        legend.position = "none"
+    )
+
+two_curvy_proj_model_benchmark2_umap <- plot_proj(
+    projected_df = projected_df, 
+    model_df = model_df2, 
+    axes = axes, 
+    circle = circle, 
+    point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
+    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
+    plot_limits = c(-0.6, 0.6), 
+    title = "b2", 
+    cex = 2, 
+    position = sc_ltr_pos,
+    axis_text_size = 4, 
+    is_category = FALSE) +
+    theme(
+        legend.position = "none"
+    )
+
+```
+
+```{r}
 #| label: fig-rm-lg
@@
 
-gen_tuning_lg_plots()
+free(distance_points_umap) + 
+    wrap_plots(trimesh_two_curvy_umap, trimesh_two_curvy_removed2_umap,
+               trimesh_two_curvy_removed1_umap, two_curvy_proj_model_delaunay_umap,
+               two_curvy_proj_model_benchmark2_umap, two_curvy_proj_first_model1_umap,
+               ncol = 3) +
+    plot_layout(heights = c(0.8, 1))
 ```

### 2025-03-20 paper.qmd
Adde all the additional functions to another script

@@
 )
-
-interior_annotation <- function(label, position = c(0.92, 0.92), cex = 1, col="grey70") {
-  annotation_custom(grid::textGrob(label = label,
-      x = unit(position[1], "npc"), y = unit(position[2], "npc"),
-      gp = grid::gpar(cex = cex, col=col)))
-}
-```
-
-```{r}
-#| label: gen-vis-obj
-
-gen_nldr_vis_algo_obj <- function(high_d_data, nldr_data, num_x_bins, col_start = "x") {
-
-  nldr_scaled_obj <- gen_scaled_data(
-    data = nldr_data)
-
-  nldr_scaled <- nldr_scaled_obj$scaled_nldr |>
-    mutate(ID = row_number())
-
-  lim1 <- nldr_scaled_obj$lim1
-  lim2 <- nldr_scaled_obj$lim2
-  r2 <- diff(lim2)/diff(lim1)
-
-  ## hexagon binning to have regular hexagons
-  hb_obj <- hex_binning(
-    data = nldr_scaled, 
-    bin1 = num_x_bins, 
-    r2 = r2,
-    q = 0.1)
-
-  ## Data set with all centroids
-  all_centroids_df <- hb_obj$centroids
-
-  a1 <- hb_obj$a1
-
-  ## Generate all coordinates of hexagons
-  hex_grid <- hb_obj$hex_poly
-
-  ## To obtain the standardise counts within hexbins
-  counts_df <- hb_obj$std_cts
-  df_bin_centroids <- extract_hexbin_centroids(
-    centroids_df = all_centroids_df, 
-    counts_df = counts_df) |>
-    filter(drop_empty == FALSE)
-
-  nldr_with_hb_id <- hb_obj$data_hb_id
-  df_all <- dplyr::bind_cols(high_d_data, 
-                                       nldr_with_hb_id)
-  df_bin <- avg_highd_data(data = df_all, col_start = col_start)
-
-  hex_grid_with_counts <- 
-    left_join(hex_grid,
-              counts_df, 
-              by = c("hex_poly_id" = "hb_id"))
-
-  hex_grid_nonempty <- hex_grid |>
-    filter(hex_poly_id %in% df_bin_centroids$hexID)
-
-  tr1_object <- tri_bin_centroids(
-    df_bin_centroids, x = "c_x", y = "c_y")
-  tr_from_to_df <- gen_edges(
-    tri_object = tr1_object)
-
-  ## Compute 2D distances
-  distance_df <- cal_2d_dist(
-    tr_coord_df = tr_from_to_df, 
-    start_x = "x_from", 
-    start_y = "y_from", 
-    end_x = "x_to", 
-    end_y = "y_to", 
-    select_vars = c("from", "to", "distance"))
-
-  ## To find the benchmark value
-  benchmark <- find_lg_benchmark(
-    distance_edges = distance_df, 
-    distance_col = "distance")
-
-  return(list(nldr_scaled = nldr_scaled,
-              a1 = a1,
-              hex_grid_with_counts = hex_grid_with_counts,
-              hex_grid_nonempty = hex_grid_nonempty,
-              df_bin_centroids = df_bin_centroids,
-              tr_from_to_df = tr_from_to_df,
-              df_bin = df_bin,
-              distance_df = distance_df,
-              benchmark = benchmark))
-
-}
-```
-
-```{r}
-#| label: scale-highd-data
-# Center the data by subtracting the mean of each column
-center_data <- function(data) {
-  apply(data, 2, function(col) col - mean(col))
-}
-
-# Function to scale data manually
-scale_data_manual <- function(data, type_col) {
-  # Step 1: Center the data (mean 0)
-  data_centered <- center_data(data |> select(-all_of(type_col)))
-
-  # Step 2: Calculate the standard deviation of each dimension
-  sds <- apply(data_centered, 2, sd)
-
-  # Step 3: Scale each dimension to have the range [0, 1]
-  data_scaled <- apply(data_centered, 2, function(col) col / max(abs(col)))
-
-  # Step 4: Scale dimensions according to their variation
-  # The dimension with the highest standard deviation is scaled to [-1, 1]
-  # Other dimensions are scaled to smaller ranges based on their standard deviations
-  max_sd <- max(sds)
-
-  # Normalize the standard deviations to get scaling factors
-  scaling_factors <- sds / max_sd
-
-  for (i in seq_along(scaling_factors)) {
-    data_scaled[, i] <- data_scaled[, i] * scaling_factors[i]
-  }
-
-  # Combine the scaled data with the 'type' column and return as a tibble
-  data_scaled <- as_tibble(data_scaled) %>% 
-    mutate(!!type_col := data[[type_col]])
-
-  return(data_scaled)
-}
-```
-
-```{r}
-#| label: gen-axes
-gen_axes <- function(proj, limits = 1, axis_pos_x = NULL, axis_pos_y = NULL, axis_labels, threshold) {
-
-  axis_scale <- limits/6
-
-  if (is.null(axis_pos_x)) {
-
-    axis_pos_x <- -2/3 * limits
-
-  }
-
-  if (is.null(axis_pos_y)) {
-
-    axis_pos_y <- -2/3 * limits
-
-  }
-
-  adj <- function(x, axis_pos) axis_pos + x * axis_scale
-  axes <- data.frame(x1 = adj(0, axis_pos_x), 
-                     y1 = adj(0, axis_pos_y), 
-                     x2 = adj(proj[, 1], axis_pos_x), 
-                     y2 = adj(proj[, 2], axis_pos_y))
-
-  rownames(axes) <- axis_labels
-
-  ## To remove axes
-  axes <- axes |>
-    mutate(distance = sqrt((x2 - x1)^2 + (y2 - y1)^2)) |>
-    filter(distance >= threshold)
-
-  theta <- seq(0, 2 * pi, length = 50)
-  circle <- data.frame(c1 = adj(cos(theta), axis_pos_x), 
-                       c2 = adj(sin(theta), axis_pos_y))
-
-  return(list(axes = axes, circle = circle))
-
-}
-```
-
-```{r}
-#| label: solve-quad-fun
-
-quad <- function(a = 3, b = 2 * a2, c = -(a2^2 + a1^2))
-{
-  a <- as.complex(a)
-  answer <- c((-b + sqrt(b^2 - 4 * a * c)) / (2 * a),
-              (-b - sqrt(b^2 - 4 * a * c)) / (2 * a))
-  if(all(Im(answer) == 0)) answer <- Re(answer)
-  if(answer[1] == answer[2]) return(answer[1])
-  answer[answer>0] ## only positive 
-}
-
-```
-
-```{r}
-#| label: plot-digit
-
-plot_digit_img <- function(digit_df, palette, title_text, hjust, ncol = 4) {
-
-  ggplot(data = digit_df, aes(x, y, fill = value)) +
-  geom_tile() +
-  facet_wrap(~ instance, ncol = ncol) +
-  ggtitle(title_text) +
-  coord_fixed() +
-  scale_fill_continuous_sequential(palette = palette) +
-  theme(strip.background = element_blank(),
-        strip.text.x = element_blank(),
-        legend.position = "none",
-        plot.title.position = "plot",
-        plot.title = element_text(hjust = hjust,
-                                  size = 15))
-
-}
-```
-
-```{r}
-#| label: get-projection
-
-get_projection <- function(projection, proj_scale, scaled_data, 
-                           scaled_data_model, distance_df_small_edges,
-                           axis_param) {
-
-  projection_scaled <- projection * proj_scale
-  projected <- as.matrix(scaled_data) %*% projection_scaled
-
-  projected_df <- projected |>
-    tibble::as_tibble(.name_repair = "unique") |>
-    dplyr::rename(c("proj1" = "...1",
-                    "proj2" = "...2")) |>
-    dplyr::mutate(ID = dplyr::row_number())
-
-  projected_model <- as.matrix(scaled_data_model) %*% projection_scaled
-
-  projected_model_df <- projected_model |>
-    tibble::as_tibble(.name_repair = "unique") |>
-    dplyr::rename(c("proj1" = "...1",
-                    "proj2" = "...2")) |>
-    dplyr::mutate(ID = dplyr::row_number())
-
-  model_df <- dplyr::left_join(
-    distance_df_small_edges |> select(-distance), 
-    projected_model_df, 
-    by = c("from" = "ID"))
-
-  names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-  model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-  names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-  limits <- axis_param$limits
-  axis_scaled <- axis_param$axis_scaled
-  axis_pos_x <- axis_param$axis_pos_x
-  axis_pos_y <- axis_param$axis_pos_y
-  threshold <- axis_param$threshold
-
-  axes_obj <- gen_axes(  
-    proj = projection * axis_scaled,
-    limits = limits,
-    axis_pos_x = axis_pos_x,
-    axis_pos_y = axis_pos_y,
-    axis_labels = names(scaled_data),
-    threshold = threshold)
-
-  axes <- axes_obj$axes
-  circle <- axes_obj$circle
-
-  return(list(projected_df = projected_df, 
-              model_df = model_df, 
-              axes = axes, 
-              circle = circle))
-
-}
-```
-
-```{r}
-#| label: plot-projection
-
-plot_proj <- function(projected_df, model_df, axes, circle, 
-                      point_param = c(1.5, 0.5, "#000000"), # size, alpha, color
-                      line_param = c(0.5, 0.5, "#000000"), #linewidth, alpha
-                      plot_limits, title, cex = 2, 
-                      position = c(0.92, 0.92), 
-                      axis_text_size = 3,
-                      is_category = FALSE) {
-
-  if(is_category == FALSE) {
-
-    initial_plot <- projected_df |>
-      ggplot(
-        aes(
-          x = proj1, 
-          y = proj2)) + 
-      geom_segment(
-        data = model_df, 
-        aes(
-          x = proj1_from, 
-          y = proj2_from, 
-          xend = proj1_to, 
-          yend = proj2_to), 
-        color = line_param[3],
-        linewidth = as.numeric(line_param[1]),
-        alpha = as.numeric(line_param[2])) +
-      geom_point(
-      size = as.numeric(point_param[1]),
-      alpha = as.numeric(point_param[2]),
-      color = point_param[3])
-
-  } else {
-
-    initial_plot <- projected_df |>
-      ggplot(
-        aes(
-          x = proj1, 
-          y = proj2,
-          colour = cluster)) + 
-      geom_segment(
-        data = model_df, 
-        aes(
-          x = proj1_from, 
-          y = proj2_from, 
-          xend = proj1_to, 
-          yend = proj2_to), 
-        color = line_param[3],
-        linewidth = as.numeric(line_param[1]),
-        alpha = as.numeric(line_param[2])) +
-      geom_point(
-      size = as.numeric(point_param[1]),
-      alpha = as.numeric(point_param[2]))
-
-  }
-
-  initial_plot <- initial_plot + 
-    geom_segment(
-      data=axes, 
-      aes(x=x1, y=y1, xend=x2, yend=y2), 
-      colour="grey70") +
-    geom_text(
-      data=axes,
-      aes(x=x2, y=y2), 
-      label=rownames(axes),
-      colour="grey50",
-      size = axis_text_size) +
-    geom_path(
-      data=circle, 
-      aes(x=c1, y=c2), colour="grey70") +
-    xlim(plot_limits) +
-    ylim(plot_limits) +
-    interior_annotation(title, position, cex = cex)
-
-  initial_plot
-
-}
-```
-
-```{r}
-#| label: plot-mse
-
-plot_mse <- function(error_df) {
-
-  ggplot(error_df, 
-       aes(x = a1, 
-           y = MSE, 
-           colour = method)) + 
-  geom_point(size = 0.8) +
-  geom_line(linewidth = 0.3) + 
-  scale_y_log10() +
-  ylab("log(MSE)") +
-  xlab(expression(paste("binwidth (", a[1], ")"))) +
-  theme_minimal() +
-  theme(panel.border = element_rect(fill = 'transparent'),
-        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
-        axis.ticks.x = element_line(),
-        axis.ticks.y = element_line(),
-        legend.position = "none",
-        axis.text.x = element_text(size = 7),
-        axis.text.y = element_text(size = 7),
-        axis.title.x = element_text(size = 7),
-        axis.title.y = element_text(size = 7))
-
-}
-```
-
-```{r}
-#| label: std-error
-
-# creating Standardization function
-standardize = function(x){
-  z <- (x - mean(x)) / sd(x)
-  return( z)
-}
 ```
@@
 #| label: import-scripts
-source("scripts/two_non_linear_diff_shaped_close_clusters/10_gen_rm_lg_plots.R")
+source("scripts/additional_functions.R")
 ```
@@
 ## For true model
+
+projection_scaled <- projection * 1
 
@@
 
+projected_model <- as.matrix(scaled_two_curvy_data_model) %*% projection_scaled
+
+projected_model_df <- projected_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
 model_df1 <- dplyr::left_join(

### 2025-03-20 paper.qmd
Updated Figure 9

@@
                alpha = 0.5,
-               size = 0.5) +
+               linewidth = 0.5) +
   geom_point(data = umap_two_curvy_scaled,
@@
 
-distance_df_small_edges_two_curvy <- distance_two_curvy |>
-  filter(distance < benchmark_two_curvy)
+# With the computed benchmark value
+# distance_df_small_edges_two_curvy <- distance_two_curvy |>
+#   filter(distance < benchmark_two_curvy)
 
@@
 
-benchmark_two_curvy <- 7 * a1
+benchmark_two_curvy <- 8 * a1
 
@@
 
-benchmark_two_curvy <- 4 * bin_width
+benchmark_two_curvy <- 2.5 * bin_width
 
@@
 text_df <- tibble::tibble(
-  x = c(0.035, 0.15),
+  x = c(0.12, 0.43),
   y = c(10, 10),
@@
     size = 10) +
-  geom_vline(xintercept = 6.5 * bin_width,
+  geom_vline(xintercept = 8 * bin_width,
              linetype="solid",
@@
              linewidth=1) +
-  geom_vline(xintercept = 2 * bin_width,
+  geom_vline(xintercept = 2.5 * bin_width,
              linetype="solid",
@@
                             scaled_data_model = scaled_two_curvy_data_model, 
-                            distance_df_small_edges = distance_df_small_edges_two_curvy, 
+                            distance_df_small_edges = distance_df_small_edges_two_curvy2, 
                             axis_param = list(limits = 0.6, 
@@
 \\gD{} Vs $7\\text{-}D$ distances between the bin centroids, where the model is fitted with UMAP for the non-linear clusters data.
-The two benchmarks for long edge removal: (b) benchmark = 6.5$a_1$ and (c) benchmark = 2$a_1$, where $a_1$ = 0.03 are made.
+The two benchmarks for long edge removal: (b) benchmark = 8$a_1$ and (c) benchmark = 2.5$a_1$, where $a_1$ = 0.03 are made.
 The wireframes in \\gD{} and a projection from $7\\text{-}D$ are shown with Delaunay triangulation (a1, a2) and the two benchmark values (b1, b2, c1, c2).

### 2025-03-25 paper.qmd
Updated Figure 18 aspect ratio
changed to out-width

@@
   theme_bw() +
-  theme(
+  theme(aspect.ratio = 1,
         axis.title = element_text(size = 18),
-    axis.text = element_text(size = 15)
-  )
+        axis.text = element_text(size = 15))
 
@@
   theme_bw() +
-  theme(
+  theme(aspect.ratio = 1,
         axis.title = element_text(size = 18),
-    axis.text = element_text(size = 15)
-  )
+        axis.text = element_text(size = 15))
 ```
@@
   scale_color_manual(values=c('#800026', '#d9d9d9')) +
-  theme(
-    aspect.ratio = 1
-  ) +
+  theme(aspect.ratio = 1) +
   interior_annotation("a2",
@@
   scale_color_manual(values=c('#800026', '#d9d9d9')) +
-  theme(
-    aspect.ratio = 1
-  ) +
+  theme(aspect.ratio = 1) +
   interior_annotation("b2",
@@
     scale_color_manual(values = c('#800026', '#d9d9d9')) +
-    theme(
-        legend.position = "none"
-    )
+    theme(aspect.ratio = 1,
+          legend.position = "none")
 
@@
     scale_color_manual(values = c('#800026', '#d9d9d9')) +
-    theme(
-        legend.position = "none"
-    )
+    theme(aspect.ratio = 1,
+          legend.position = "none")
 ```
@@
 #| fig-pos: H
-#| out-height: 100%
+#| out-width: 100%
 #| fig-cap: "The sparse end of the structure shows a high model error, while the dense end shows a low model error.
@@
 
-error_plot_one_curvy_hist + plot_tsne_dens + five_c_shaped_proj_tsne_model1 +
-  error_plot_one_curvy_hist_selected + plot_tsne_dens_selected + five_c_shaped_proj_tsne_model1_selected +
+error_plot_one_curvy_hist + plot_tsne_dens + 
+  five_c_shaped_proj_tsne_model1 +
+  error_plot_one_curvy_hist_selected + plot_tsne_dens_selected +
+  five_c_shaped_proj_tsne_model1_selected +
     plot_layout(guides = "collect", nrow = 2) &
@@
 #| fig-width: 12
-#| out-height: 100%
+#| out-width: 100%
 #| fig-cap: "\\gD{} wireframe layout shows that there is a big non-linear cluster (grey) and a small cluster (orange) located very close to the one corner of the big cluster.
@@
 #| fig-width: 15
-#| out-height: 100%
+#| out-width: 100%
 #| fig-cap: "The tSNE layout of the MNIST digit 1 data shows a big non-linear cluster (grey) and a small cluster (orange) located very close to the one corner of the big cluster in \\gD{} (a).
@@
 #| fig-pos: H
-#| fig-width: 10
-#| fig-height: 7
+#| fig-width: 7
+#| fig-height: 5
 #| out-width: 90%

### 2025-03-26 paper.qmd
Add
Figure 18 updated
small typo

@@
     is_category = FALSE) +
-    theme(
-        legend.position = "none"
-    )
+    theme(aspect.ratio = 1,
+          legend.position = "none")
 
@@
   is_category = FALSE) +
-  theme(
-    legend.position = "none"
-  )
+  theme(aspect.ratio=1, 
+        legend.position = "none")
 
@@
     is_category = FALSE) +
-    theme(
-        legend.position = "none"
-    )
+    theme(aspect.ratio = 1, 
+          legend.position = "none")
 
@@
                alpha = 0.5,
-               size = 0.5)+
+               linewidth = 0.5)+
   geom_point(data = umap_two_curvy_scaled,
@@
 
-error_plot_one_curvy_hist_selected <- ggplot(error_df_one_curvy_abs, aes(x = sqrt_row_wise_total_error, y = density,
-                                                                         colour = factor(error_cat_n,
-                                                                                         levels = c("high", "low")))) +
-  geom_point() +
-  scale_color_manual(values=c('#800026', '#d9d9d9')) +
+error_plot_one_curvy_hist_selected <- ggplot(error_df_one_curvy_abs, 
+      aes(x = sqrt_row_wise_total_error, 
+          y = density,
+          colour = factor(error_cat_n, levels = c("low", "high")))) +
+  geom_point(alpha=0.7) +
+  scale_color_manual(values=c('#d9d9d9', '#800026')) +
   xlab(expression(e[hj])) +
@@
 error_plot_one_curvy_hist <- ggplot(error_df_one_curvy_abs,
-                                    aes(x = sqrt_row_wise_total_error, y = density,
-                                        colour = factor(error_cat_n2,
-                                                        levels = c("high", "low")))) +
-  geom_point() +
-  scale_color_manual(values=c('#800026', '#d9d9d9')) +
+        aes(x = sqrt_row_wise_total_error, 
+            y = density,
+            colour = factor(error_cat_n2, levels = c("low", "high")))) +
+  geom_point(alpha=0.7) +
+  scale_color_manual(values=c('#d9d9d9', '#800026')) +
   xlab(expression(e[hj])) +
@@
              colour = factor(error_cat_n2,
-                             levels = c("high", "low")))) +
+                             levels = c("low", "high")))) +
   geom_point(alpha=0.7) +
-  scale_color_manual(values=c('#800026', '#d9d9d9')) +
+  scale_color_manual(values=c('#d9d9d9', '#800026')) +
   theme(aspect.ratio = 1) +
@@
              colour = factor(error_cat_n,
-                             levels = c("high", "low")))) +
+                             levels = c("low", "high")))) +
   geom_point(alpha=0.7) +
-  scale_color_manual(values=c('#800026', '#d9d9d9')) +
+  scale_color_manual(values=c('#d9d9d9', '#800026')) +
   theme(aspect.ratio = 1) +
@@
 projected_df <- projected_df |>
-  mutate(cluster = error_df_one_curvy_abs$error_cat_n2)
+  mutate(cluster = factor(error_df_one_curvy_abs$error_cat_n2,
+                          levels=c("low", "high")))
 
@@
     is_category = TRUE) +
-    scale_color_manual(values = c('#800026', '#d9d9d9')) +
+    scale_color_manual(values = c('#d9d9d9', '#800026')) +
     theme(aspect.ratio = 1,

### 2025-03-30 paper.qmd
Changed augment arguments with new quollr version

--- 2025-03-25 paper.qmd
@@
 error_df_two_curvy_abs <- augment(
-  df_bin_centroids = df_bin_centroids_two_curvy,
-  df_bin = df_bin_two_curvy,
-  training_data = training_data_two_curvy,
-  newdata = NULL,
-  type_NLDR = "tSNE",
-  col_start = "x")
+  highd_data = training_data_two_curvy,
+  model_2d = df_bin_centroids_two_curvy,
+  model_highd = df_bin_two_curvy)
 
@@
   is_rm_lwd_hex = FALSE,
-  col_start_highd = "x",
   q = 0.1
@@
 error_df_gau_abs <- augment(
-  df_bin_centroids = df_bin_centroids_gau1,
-  df_bin = df_bin_gau1,
-  training_data = training_data_gau,
-  newdata = NULL,
-  type_NLDR = "tSNE",
-  col_start = "x")
+  highd_data = training_data_gau,
+  model_2d = df_bin_centroids_gau1,
+  model_highd = df_bin_gau1)
 
@@
   is_bin_centroid = TRUE,
-  is_rm_lwd_hex = FALSE,
-  col_start_highd = "x"
+  is_rm_lwd_hex = FALSE
 )
@@
 error_df_gau_abs <- augment(
-  df_bin_centroids = df_bin_centroids_gau1,
-  df_bin = df_bin_gau1,
-  training_data = training_data_gau,
-  newdata = NULL,
-  type_NLDR = "UMAP",
-  col_start = "x")
+  highd_data = training_data_gau,
+  model_2d = df_bin_centroids_gau1,
+  model_highd = df_bin_gau1)
 
@@
 error_df_gau_abs <- augment(
-  df_bin_centroids = df_bin_centroids_gau1,
-  df_bin = df_bin_gau1,
-  training_data = training_data_gau,
-  newdata = NULL,
-  type_NLDR = "PaCMAP",
-  col_start = "x")
+  highd_data = training_data_gau,
+  model_2d = df_bin_centroids_gau1,
+  model_highd = df_bin_gau1)
 
@@
 error_df_one_curvy_abs <- augment(
-  df_bin_centroids = df_bin_centroids_dens_clust,
-  df_bin = df_bin_dens_clust,
-  training_data = one_c_shaped_data,
-  newdata = NULL,
-  type_NLDR = "tSNE",
-  col_start = "x")
+  highd_data = one_c_shaped_data,
+  model_2d = df_bin_centroids_dens_clust,
+  model_highd = df_bin_dens_clust)
 
@@
 training_data_pbmc <- read_rds("data/pbmc3k/pbmc_pca_50.rds")
-names(training_data_pbmc) <- paste0("pc", 1:50)
+names(training_data_pbmc) <- paste0("x", 1:50)
 
@@
   num_x_bins = num_bins_x_pbmc,
-  col_start = "pc")
+  col_start = "x")
 
@@
 training_data_mnist <- read_rds("data/mnist/mnist_10_pcs_of_digit_1.rds")
+names(training_data_mnist) <- paste0("x", 1:NCOL(training_data_mnist))
+
 training_data_mnist <- training_data_mnist |>
@@
   num_x_bins = num_bins_x_mnist,
-  col_start = "PC")
+  col_start = "x")
 
@@
 error_df <- augment(
-  df_bin_centroids = df_bin_centroids_mnist, 
-  df_bin = df_bin_mnist, 
-  training_data = training_data_mnist, 
-  newdata = NULL, 
-  type_NLDR = "tSNE", 
-  col_start = "PC")
+  highd_data = training_data_mnist,
+  model_2d = df_bin_centroids_mnist,
+  model_highd = df_bin_mnist)
 
@@
 #| fig-pos: H
-#| fig-width: 7
-#| fig-height: 5
+#| fig-width: 10
+#| fig-height: 7
 #| out-width: 90%

### 2025-03-30 paper.qmd
Updated the paper.qmd

@@
   nldr_data = tsne_two_curvy, 
-  num_x_bins = num_bins_x_two_curvy,
-  col_start = "x")
+  num_x_bins = num_bins_x_two_curvy)
 
@@
   nldr_data = umap_two_curvy, 
-  num_x_bins = num_bins_x_two_curvy,
-  col_start = "x")
+  num_x_bins = num_bins_x_two_curvy)
 
@@
 df_all_two_curvy4 <- dplyr::bind_cols(training_data_two_curvy, tsne_data_two_curvy2_with_hb_id)
-df_bin_two_curvy4 <- avg_highd_data(data = df_all_two_curvy4, col_start = "x")
+df_bin_two_curvy4 <- avg_highd_data(data = df_all_two_curvy4)
 
@@
 df_all_two_curvy4 <- dplyr::bind_cols(training_data_two_curvy, pacmap_data_two_curvy2_with_hb_id)
-df_bin_two_curvy4 <- avg_highd_data(data = df_all_two_curvy4, col_start = "x")
+df_bin_two_curvy4 <- avg_highd_data(data = df_all_two_curvy4)
 
@@
 gau1_model <- fit_highd_model(
-  training_data = training_data_gau,
-  emb_df = tsne_gau_scaled,
+  highd_data = training_data_gau,
+  nldr_data = tsne_gau_scaled,
   bin1 = num_bins_x_gau1,
@@
   is_bin_centroid = TRUE,
-  is_rm_lwd_hex = FALSE,
   q = 0.1
@@
 gau1_model <- fit_highd_model(
-  training_data = training_data_gau,
-  emb_df = umap_gau_scaled,
+  highd_data = training_data_gau,
+  nldr_data = umap_gau_scaled,
   bin1 = num_bins_x_gau1,
@@
   is_bin_centroid = TRUE,
-  is_rm_lwd_hex = FALSE
+  q = 0.1
 )
@@
 gau1_model <- fit_highd_model(
-  training_data = training_data_gau,
-  emb_df = pacmap_gau_scaled,
+  highd_data = training_data_gau,
+  nldr_data = pacmap_gau_scaled,
   bin1 = num_bins_x_gau1,
@@
   nldr_data = tsne_one_c_shaped, 
-  num_x_bins = num_bins_x_one_c_shaped,
-  col_start = "x")
+  num_x_bins = num_bins_x_one_c_shaped)
 
@@
   nldr_data = umap_pbmc, 
-  num_x_bins = num_bins_x_pbmc,
-  col_start = "x")
+  num_x_bins = num_bins_x_pbmc)
 
@@
   nldr_data = umap_pbmc2, 
-  num_x_bins = num_bins_x_pbmc,
-  col_start = "pc")
+  num_x_bins = num_bins_x_pbmc)
 
@@
   nldr_data = tsne_mnist2, 
-  num_x_bins = num_bins_x_mnist,
-  col_start = "x")
+  num_x_bins = num_bins_x_mnist)
 

### 2025-03-30 paper.qmd

--- 2025-03-26 paper.qmd
@@
   is_bin_centroid = TRUE,
-  is_rm_lwd_hex = FALSE,
-  col_start_highd = "x",
   q = 0.1
@@
   is_bin_centroid = TRUE,
-  is_rm_lwd_hex = FALSE,
-  col_start_highd = "x"
+  q = 0.1
 )
@@
   nldr_data = umap_pbmc, 
-  num_x_bins = num_bins_x_pbmc,
-  col_start = "pc")
+  num_x_bins = num_bins_x_pbmc)
 
@@
   nldr_data = tsne_mnist2, 
-  num_x_bins = num_bins_x_mnist,
-  col_start = "PC")
+  num_x_bins = num_bins_x_mnist)
 

### 2025-03-31 paper.qmd
working on correct sizing of figures

@@
 
-## Best fit
+## Best fit, or at least avoiding an inaccurate representation
+
+@fig-toy-mse shows a summary of plots that can be used to help assess the strength of the fit, and compare the fits for different representations.
+What does it mean to be a best fit for this problem?
+There probably isn't a best fit, but there are wrong representations.
+The goal is to help users decide on a useful and appropriate low-dimensional representation of the high-dimensional data.
 
@@
   plot_mse(error_two_non_linear_diff_shaped_close_clusters) +
-  scale_x_continuous(breaks = sort(unique(error_two_non_linear_diff_shaped_close_clusters$a1))[seq(1, length(unique(error_two_non_linear_diff_shaped_close_clusters$a1)), by = 5)]) +
-  scale_color_manual(values=c('#e41a1c','#ff7f00','#4daf4a', "#a65628",'#636363', '#984ea3')) +
-  scale_y_log10()
-
+  scale_x_continuous(breaks =
+    sort(unique(error_two_non_linear_diff_shaped_close_clusters$a1))[
+      seq(1, length(
+        unique(error_two_non_linear_diff_shaped_close_clusters$a1)), 
+        by = 5)]) +
+  scale_color_manual(
+    values=c('#e41a1c','#ff7f00','#4daf4a', 
+             "#a65628",'#636363', '#984ea3')) +
+  scale_y_log10() + 
+  theme(aspect.ratio = 1.5)
 ```
@@
 Layout a is the best choice."
-#| label: fig-two_non_linear_diff_shaped_close_clusters-mse
+#| label: fig-toy-mse
 #| fig-pos: H
 #| out-width: 100%
-
-free(error_plot_two_non_linear_diff_shaped_close_clusters) + wrap_plots(nldr_two_non_linear_diff_shaped_close_clusters1, nldr_two_non_linear_diff_shaped_close_clusters2, nldr_two_non_linear_diff_shaped_close_clusters3,
+#| fig-height: 6
+#| fig-width: 6
+
+error_plot_two_non_linear_diff_shaped_close_clusters + wrap_plots(nldr_two_non_linear_diff_shaped_close_clusters1, nldr_two_non_linear_diff_shaped_close_clusters2, nldr_two_non_linear_diff_shaped_close_clusters3,
                                     nldr_two_non_linear_diff_shaped_close_clusters4, nldr_two_non_linear_diff_shaped_close_clusters5,
-                                    nldr_two_non_linear_diff_shaped_close_clusters6, ncol = 2)
+                                    nldr_two_non_linear_diff_shaped_close_clusters6, ncol = 2, 
+                                    widths = c(50, 50))
 ```

### 2025-04-01 paper.qmd
revised figure 11

@@
     ),
-    alpha = 0.2) + 
+    colour = "#0077A3",
+    alpha = 0.1) + 
+  geom_segment(
+    data = tr_from_to_df_gau1,
+    aes(
+      x = x_from,
+      y = y_from,
+      xend = x_to,
+      yend = y_to),
+    color = "#000000",
+    alpha = 0.4,
+    linewidth = 0.5) +
   interior_annotation("a1", 
@@
     #size = 0.5,
-    alpha = 0.3,
-    color = "#999999") +
+    alpha = 0.05,
+    color = "#0077A3") +
   geom_segment(
@@
     ),
+    colour = "#0077A3",
     alpha = 0.05) + 
+  geom_segment(
+    data = tr_from_to_df_gau1,
+    aes(
+      x = x_from,
+      y = y_from,
+      xend = x_to,
+      yend = y_to),
+    color = "#000000",
+    alpha = 0.4,
+    linewidth = 0.5) +
   interior_annotation("b1", 
@@
   geom_point(
-    alpha = 0.3,
-    color = "#999999") +
+    alpha = 0.05,
+    color = "#0077A3") +
   geom_segment(
@@
       x = PaCMAP1,
-      y = PaCMAP2
-    ),
-    alpha = 0.3) +
+      y = PaCMAP2),
+      colour = "#0077A3",
+    alpha = 0.1) +
+  geom_segment(
+    data = tr_from_to_df_gau1,
+    aes(
+      x = x_from,
+      y = y_from,
+      xend = x_to,
+      yend = y_to),
+    color = "#000000",
+    alpha = 0.4,
+    linewidth = 0.5) +
   interior_annotation("c1", 
@@
 #| echo: false
-#| fig-cap: "Compare the author suggested \\gD{} layout (@fig-pbmc-mse a) and the \\gD{} layout selected (@fig-pbmc-mse b) by MSE plot (@fig-pbmc-mse) from the same NLDR with different hyper-parameter choice.
+#| fig-cap: "Compare the published \\gD{} layout (@fig-pbmc-mse a) and the \\gD{} layout selected (@fig-pbmc-mse b) by MSE plot (@fig-pbmc-mse) from the same NLDR with different hyper-parameter choice.
 The PBMC3k data ($n =  2622$) has three clusters in $9\\text{-}D$, where two of them being close and one separated from the others.
@@
 #| fig-height: 15
-#| out-width: 60%
+#| out-width: 90%
 
@@
   pbmc_proj_umap_model2 + pbmc_proj_umap_model2_best +
-  plot_layout(nrow=3) &
+  plot_layout(nrow=2, byrow=FALSE) &
   theme(legend.position='none')
@@
 error_plot_mnist <- plot_mse(error_mnist) +
-  scale_x_continuous(breaks = sort(unique(error_mnist$a1))[append(seq(1, length(unique(error_mnist$a1)), by = 5), 24)]) +
-  scale_color_manual(values=c('#e41a1c','#377eb8',"#ff7f00", '#984ea3', '#999999', '#a65628'))
+  scale_x_continuous(breaks = sort(
+    unique(error_mnist$a1))[append(seq(1, length(unique(error_mnist$a1)), 
+                                       by = 5), 24)]) +
+  scale_color_manual(
+    values=c('#e41a1c','#377eb8',"#ff7f00", 
+             '#984ea3', '#999999', '#a65628')) +
+  theme(aspect.ratio=1.5)
 
@@
 
-free(error_plot_mnist) + wrap_plots(nldr_mnist1, nldr_mnist2, nldr_mnist3, 
+error_plot_mnist + wrap_plots(nldr_mnist1, nldr_mnist2, nldr_mnist3, 
                                     nldr_mnist4, nldr_mnist5, nldr_mnist6, ncol = 2)

### 2025-04-01 paper.qmd
correct citation for MDS

@@
 <!-- - Connection between NLDR and MDS-->
-Historically, \kD{}   representations of \pD{}   data have been computed using multidimensional scaling (MDS) [@borg2005], which includes principal components analysis (PCA) [@jolliffe2011] as a special case.
-The \kD{}   representation can be considered to be a layout of points in \kD{}   produced by an embedding procedure that maps the data from \pD{}.
+Historically, \kD{}   representations of \pD{} data have been computed using multidimensional scaling (MDS) [@kruskal1964], which includes principal components analysis (PCA) [@jolliffe2011] as a special case.
+(@borg2005 provides a comprehensive guide to MDS.) The \kD{}   representation can be considered to be a layout of points in \kD{}   produced by an embedding procedure that maps the data from \pD{}.
 In MDS, the \kD{}   layout is constructed by minimizing a stress function that differences distances between points in \pD{}   with potential distances between points in \kD{}.
@@
     alpha = 0.2,
+    size = 0.1,
     color = '#a65628'

### 2025-04-01 paper.qmd
name of bib file standardised
correct kruskal refence included

@@
 
-bibliography: bibliography.bib  
+bibliography: paper.bib  
 header-includes: | 
@@
 Historically, \kD{}   representations of \pD{} data have been computed using multidimensional scaling (MDS) [@kruskal1964], which includes principal components analysis (PCA) [@jolliffe2011] as a special case.
-(@borg2005 provides a comprehensive guide to MDS.) The \kD{}   representation can be considered to be a layout of points in \kD{}   produced by an embedding procedure that maps the data from \pD{}.
+(A contemporary comprehensive guide to MDS can be found in @borg2005.) The \kD{}   representation can be considered to be a layout of points in \kD{}   produced by an embedding procedure that maps the data from \pD{}.
 In MDS, the \kD{}   layout is constructed by minimizing a stress function that differences distances between points in \pD{}   with potential distances between points in \kD{}.
@@
 We can also imagine that the positioning of points in \gD{}    represent the fitted values, that will have some prescribed position in \pD{}   that can be compared with their observed values.
-This is the conceptual framework underlying the more formal versions of factor analysis [@cfa69] and multidimensional scaling (MDS) [@borg2005].
+This is the conceptual framework underlying the more formal versions of factor analysis [@cfa69] and MDS.
 (Note that, for this thinking the full \pD{}   data needs to be available, not just the interpoint distances.)

### 2025-04-01 paper.qmd
revised figure 5

@@
 
-nldr_two_curvy_original <- ggplot(tsne_two_curvy_scaled, aes(x = tSNE1, y = tSNE2)) +
-  geom_point(alpha = 0.5, color = "#636363", size = 0.5)
+nldr_two_curvy_original <- ggplot(tsne_two_curvy_scaled, 
+                                  aes(x = tSNE1, y = tSNE2)) +
+  geom_point(alpha = 0.5, color = clr_choice, size = 0.5)
 
 # First plot with xlim and ylim
-nldr_two_curvy <- nldr_two_curvy_original+
+nldr_two_curvy <- nldr_two_curvy_original +
   interior_annotation("a", sc_ltr_pos, cex = 1.5) +
@@
              aes(x = tSNE1, y = tSNE2), 
-             alpha = 0.5, size = 0.5, color = "#636363") +
+             alpha = 0.5, size = 0.5, color = clr_choice) +
   xlim(sc_xlims) + ylim(sc_ylims) +
@@
              aes(x = tSNE1, y = tSNE2),
+             colour = clr_choice,
              alpha = 0.3,
@@
              ),
-             color = "#636363",
+             color = clr_choice,
              alpha = 0.5,
@@
 ## First projection
-projection <- cbind(
+model_prj1 <- cbind(
   c(0.37847,-0.06517,0.04231,0.01655),
@@
 
-proj_obj1 <- get_projection(projection = projection, 
+proj_obj1 <- get_projection(projection = model_prj1, 
                proj_scale = 1, 
@@
 
-projection_scaled <- projection * 1
+projection_scaled <- model_prj1 * 1
 
@@
     circle = circle, 
-    point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
+    point_param = c(0.5, 0.2, clr_choice), # size, alpha, color
     line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
@@
   circle = circle, 
-  point_param = c(0.5, 0.5, "#636363"), # size, alpha, color
-  line_param = c(0.5, 0.5, "#2b8cbe"), # linewidth, alpha
+  point_param = c(0.5, 0.5, clr_choice), # size, alpha, color
+  line_param = c(0.5, 0.5, "black"), # linewidth, alpha
   plot_limits = c(-0.6, 0.6), 
@@
 
-two_curvy_proj_tsne_all_model1 <- plot_proj(
-  projected_df = projected_df, 
-  model_df = true_model_df_proj, 
-  axes = axes, 
-  circle = circle, 
-  point_param = c(0.5, 0.5, "#636363"), # size, alpha, color
-  line_param = c(0.5, 0.5, "#2b8cbe"), # linewidth, alpha
-  plot_limits = c(-0.6, 0.6), 
-  title = "d", 
-  cex = 1.5, 
-  position = sc_ltr_pos,
-  axis_text_size = 4, 
-  is_category = FALSE) +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#000000")
-
 ## Second projection
-projection <- cbind(
+model_prj2 <- cbind(
   c(-0.38281,-0.04870,0.00619,-0.02451),
@@
 
-proj_obj2 <- get_projection(projection = projection, 
+proj_obj2 <- get_projection(projection = model_prj2, 
                             proj_scale = 1, 
@@
 
-projected_true_model <- as.matrix(scaled_two_curvy_data_true_model) %*% projection
+projected_true_model <- as.matrix(scaled_two_curvy_data_true_model) %*% model_prj2
 
@@
     circle = circle, 
-    point_param = c(0.5, 0.5, "#636363"), # size, alpha, color
-    line_param = c(0.5, 0.5, "#2b8cbe"), # linewidth, alpha
+    point_param = c(0.5, 0.5, clr_choice), # size, alpha, color
+    line_param = c(0.5, 0.5, "black"), # linewidth, alpha
     plot_limits = c(-0.42, 0.42), 
@@
           legend.position = "none")
+
+two_curvy_proj_tsne_all_model1 <- plot_proj(
+  projected_df = projected_df, 
+  model_df = model_df, 
+  axes = axes, 
+  circle = circle, 
+  point_param = c(0.1, 0.5, clr_choice), # size, alpha, color
+  line_param = c(0.5, 0.5, "black"), # linewidth, alpha
+  plot_limits = c(-0.6, 0.6), 
+  title = "d", 
+  cex = 1.5, 
+  position = sc_ltr_pos,
+  axis_text_size = 4, 
+  is_category = FALSE) #+
+  #geom_segment(
+  #  data = model_df, 
+  #  aes(
+  #    x = proj1_from, 
+  #    y = proj2_from, 
+  #    xend = proj1_to, 
+  #    yend = proj2_to), 
+  #  color = "#000000")
 
@@
 #| label: best-fit-tsne
-#| fig-cap: "We start with a tSNE layout (a) of non-linear clusters data ($n =  2000$ and $p = 4$), from which the \\gD{} wireframe is constructed (b) and then lifted to $4\\text{-}D$.
-A \\gD{} projection of the model fit from a tour through the $4\\text{-}D$ space is presented in c.
-Additionally, a \\gD{} projection of the fitted and true models (d) from the same tour demonstrates that the model accurately captures the non-linear structure of the data, although it does not fully represent the width of that structure.
+#| fig-cap: "Lifting the fitted model into \\pD{}.
+Starting from a default tSNE layout (a) of non-linear clusters data ($n =  2000$ and $p = 4$), from which the \\gD{} wireframe is constructed (b) and then lifted to $4\\text{-}D$ (c, d).
+The fit is reasonbly tight with the data, although it does not fully spread the full width.
 This highlights a key characteristic of tSNE, which compresses the data when applying tSNE.
 Furthermore, the sparse space observed in the middle of the tSNE layout is a result of (hyper-)parameter choice.
-The default is used.
 Video of the langevitour animation is available at <>."
@@
   two_curvy_proj_tsne_all_model1 +
-  plot_layout(ncol = 2)
+  plot_layout(ncol = 4)
 ```
@@
     ),
-    colour = "#0077A3",
+    colour = clr_choice,
     alpha = 0.1) + 
@@
     alpha = 0.05,
-    color = "#0077A3") +
+    color = clr_choice) +
   geom_segment(
@@
     ),
-    colour = "#0077A3",
+    colour = clr_choice,
     alpha = 0.05) + 
@@
       y = PaCMAP2),
-      colour = "#0077A3",
+      colour = clr_choice,
     alpha = 0.1) +

### 2025-04-01 paper.qmd
Fixed Figure 5(c)

@@
 
-two_curvy_proj_tsne_first_model1 <- plot_proj(
-    projected_df = projected_df, 
-    model_df = model_df, 
-    axes = axes, 
-    circle = circle, 
-    point_param = c(0.5, 0.2, clr_choice), # size, alpha, color
-    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
-    plot_limits = c(-0.6, 0.6), 
-    title = "c", 
-    cex = 1.5, 
-    position = sc_ltr_pos,
-    axis_text_size = 4, 
-    is_category = FALSE) +
-    theme(aspect.ratio = 1,
-          legend.position = "none")
-
 two_curvy_proj_true <- plot_proj(
@@
 
+# Changed the axis parameters
+axis_obj <- gen_axes(
+    proj = model_prj1 * 2,
+    limits = 0.8,
+    axis_pos_x = -0.4,
+    axis_pos_y = -0.4,
+    axis_labels = names(scaled_two_curvy_data),
+    threshold = 0.05)
+
+axes <- axis_obj$axes
+circle <- axis_obj$circle
+
+two_curvy_proj_tsne_first_model1 <- plot_proj(
+    projected_df = projected_df, 
+    model_df = model_df, 
+    axes = axes, 
+    circle = circle, 
+    point_param = c(0.5, 0.2, clr_choice), # size, alpha, color
+    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
+    plot_limits = c(-0.6, 0.6), 
+    title = "c", 
+    cex = 1.5, 
+    position = sc_ltr_pos,
+    axis_text_size = 2, 
+    is_category = FALSE) +
+    theme(aspect.ratio = 1,
+          legend.position = "none")
+
 ## Second projection
@@
 
+# Changed the axis parameters
+axis_obj <- gen_axes(
+    proj = model_prj2 * 2,
+    limits = 0.8,
+    axis_pos_x = -0.3,
+    axis_pos_y = -0.3,
+    axis_labels = names(scaled_two_curvy_data),
+    threshold = 0)
+
+axes <- axis_obj$axes
+circle <- axis_obj$circle
+
 two_curvy_proj_tsne_all_model1 <- plot_proj(
@@
   line_param = c(0.5, 0.5, "black"), # linewidth, alpha
-  plot_limits = c(-0.6, 0.6), 
+  plot_limits = c(-0.5, 0.5), 
   title = "d", 
-  cex = 1.5, 
+  cex = 1, 
   position = sc_ltr_pos,
@@
 Video of the langevitour animation is available at <>."
-#| fig-width: 10
-#| fig-height: 10
+#| fig-width: 8
+#| fig-height: 2
 

### 2025-04-01 paper.qmd
Fixed Figure 5

@@
     proj = model_prj1 * 2,
-    limits = 0.8,
+    limits = 0.9,
     axis_pos_x = -0.4,
@@
                             axis_param = list(limits = 0.5, 
-                                              axis_scaled = 6,
+                                              axis_scaled = 2,
                                               axis_pos_x = -0.3, 
                                               axis_pos_y = -0.3, 
-                                              threshold = 0))
+                                              threshold = 0.01))
 
@@
 
-# Changed the axis parameters
+#Changed the axis parameters
 axis_obj <- gen_axes(
     proj = model_prj2 * 2,
-    limits = 0.8,
-    axis_pos_x = -0.3,
-    axis_pos_y = -0.3,
+    limits = 0.7,
+    axis_pos_x = -0.35,
+    axis_pos_y = -0.35,
     axis_labels = names(scaled_two_curvy_data),
-    threshold = 0)
+    threshold = 0.02)
 
@@
   position = sc_ltr_pos,
-  axis_text_size = 4, 
+  axis_text_size = 2, 
   is_category = FALSE) #+

### 2025-04-01 paper.qmd
Start working on adding langevitour

@@
 
+```
+
+```{r}
+#| eval: false
+langevitour::langevitour(df_model_data_two_curvy[1:(length(df_model_data_two_curvy)-1)],
+                         lineFrom = distance_df_small_edges_two_curvy$from,
+                         lineTo = distance_df_small_edges_two_curvy$to,
+                         group = factor(df_model_data_two_curvy$type,
+                                        c("data", "true mdeol","model")),
+                         levelColors = c(clr_choice, '#000000', "#000000"))
+```
+
+```{r}
 ## First projection

### 2025-04-01 paper.qmd
Added langevitour for true model of the toy example

--- 2025-04-01 paper.qmd
@@
 ```{r}
-#| label: best-fit-two-curvy-proj-tsne
+#| label: prep-data-for-proj-tsne-best
 
@@
 
+```
+
+```{r}
+#| eval: false
+df_model_data_two_curvy_filtered <- bind_rows(true_model_two_curvy, data_two_curvy)
+
+langevitour::langevitour(df_model_data_two_curvy_filtered[1:(length(df_model_data_two_curvy_filtered)-1)],
+                         lineFrom = wireframe_true_model$from,
+                         lineTo = wireframe_true_model$to,
+                         group = factor(df_model_data_two_curvy_filtered$type,
+                                        c("data", "true model")),
+                         levelColors = c(clr_choice, "#000000"))
+```
+
+```{r}
+#| label: best-fit-two-curvy-proj-tsne
 ## First projection

### 2025-04-01 paper.qmd
Added langevitour for model fit with tSNE

@@
 ```{r}
+#| label: langevitour-true-two-curvy-proj
 #| eval: false
@@
                                         c("data", "true model")),
+                         levelColors = c(clr_choice, "#000000"))
+```
+
+```{r}
+#| label: langevitour-best-fit-two-curvy-proj-tsne
+#| eval: false
+
+df_model_data_two_curvy_filtered <- bind_rows(df_bin_two_curvy, data_two_curvy)
+
+langevitour::langevitour(df_model_data_two_curvy_filtered[1:(length(df_model_data_two_curvy_filtered)-1)],
+                         lineFrom = distance_df_small_edges_two_curvy$from,
+                         lineTo = distance_df_small_edges_two_curvy$to,
+                         group = factor(df_model_data_two_curvy_filtered$type,
+                                        c("data", "model")),
                          levelColors = c(clr_choice, "#000000"))

### 2025-04-01 paper.qmd
Added langevitour for MNIST data

@@
 <!--bin1 = 22, bin2 = 17, b = 374, non_empty = 140-->
-```{r}
-#| label: mnist-model-proj
+
+```{r}
+#| label: data-prep-mnist-model-proj
 
@@
   select(-type)
+```
+
+```{r}
+#| label: langevitour-best-fit-two-curvy-proj-tsne
+#| eval: false
+
+data_mnist_cluster <- data_mnist |>
+  select(-type) |>
+  mutate(type = tsne_minst_scaled$cluster)
+
+df_model_data_mnist_n <- bind_rows(data_mnist_cluster, df_b_mnist)
+
+langevitour::langevitour(df_model_data_mnist_n[1:(length(df_model_data_mnist_n)-1)],
+                         lineFrom = distance_df_small_edges_mnist$from,
+                         lineTo = distance_df_small_edges_mnist$to,
+                         group = factor(df_model_data_mnist_n$type,
+                                        c("big_clust", "small_clust", "model")),
+                         levelColors = c("#999999",'#ff7f00', "#000000"))
+```
+
+```{r}
+#| label: mnist-model-proj
 

### 2025-04-02 paper.qmd
Added langevitour for Single cell data example

@@
 ```{r}
-#| label: pbmc-umap-model-proj
+#| label: prep-pbmc-umap-model-proj
 
@@
   select(-type)
-
-# # Combine with the true model for visualization
-# df <- dplyr::bind_rows(scaled_pbmc_data_model |> mutate(type = "model"),
-#                        scaled_pbmc_data |> mutate(type = "data"))
+```
+
+```{r}
+#| label: langevitour-pbmc-umap-author-proj
+#| eval: false
+
+df_model_data_pbmc_n <- bind_rows(df_b_pbmc, data_pbmc)
+
+langevitour::langevitour(df_model_data_pbmc_n[1:(length(df_model_data_pbmc_n)-1)],
+                         lineFrom = distance_df_small_edges_pbmc$from,
+                         lineTo = distance_df_small_edges_pbmc$to,
+                         group = factor(df_model_data_pbmc_n$type,
+                                        c("data", "model")),
+                         levelColors = c("#a65628", "#000000"))
+```
+
+```{r}
+#| label: pbmc-umap-model-proj
 
@@
 ```{r}
-#| label: pbmc-umap-model-proj-best
+#| label: prep-pbmc-umap-model-proj-best
 
@@
   select(-type)
+```
+
+```{r}
+#| label: langevitour-pbmc-umap-best-proj
+#| eval: false
+
+df_model_data_pbmc_n <- bind_rows(df_b_pbmc, data_pbmc)
+
+langevitour::langevitour(df_model_data_pbmc_n[1:(length(df_model_data_pbmc_n)-1)],
+                         lineFrom = distance_df_small_edges_pbmc$from,
+                         lineTo = distance_df_small_edges_pbmc$to,
+                         group = factor(df_model_data_pbmc_n$type,
+                                        c("data", "model")),
+                         levelColors = c("#999999", "#000000"))
+```
+
+```{r}
+#| label: pbmc-umap-model-proj-best
 

### 2025-04-02 paper.qmd
Added langevitour plot for dens cluster

@@
 ```{r}
-#| label: proj-one-dens-clust
+#| label: prep-proj-one-dens-clust
 
@@
   select(-type)
+```
+
+```{r}
+#| label: langevitour-pbmc-umap-author-proj
+#| eval: false
+
+data_c_shaped_n <- data_c_shaped |>
+  select(-type) |>
+  mutate(type = factor(error_df_one_curvy_abs$error_cat_n2,
+                          levels=c("low", "high")))
+
+df_model_data_n <- bind_rows(df_bin_dens_clust, data_c_shaped_n)
+
+langevitour::langevitour(df_model_data_n[1:(length(df_model_data_n)-1)],
+                         lineFrom = distance_df_small_edges_c_shaped_structure$from,
+                         lineTo = distance_df_small_edges_c_shaped_structure$to,
+                         group = factor(df_model_data_n$type,
+                                        c("low", "high", "model")),
+                         levelColors = c('#d9d9d9', '#800026', "#000000"))
+```
+
+```{r}
+#| label: proj-one-dens-clust
 

### 2025-04-02 paper.qmd
Added all the langevitour which produce the projections

@@
   dplyr::filter(distance < benchmark_gau1)
-
+```
+
+```{r}
+#| label: langevitour-five-gau-tsne-proj
+#| eval: false
+
+df_model_data_n <- bind_rows(data_gau, df_b)
+
+langevitour::langevitour(df_model_data_n[1:(length(df_model_data_n)-1)],
+                         lineFrom = distance_df_small_edges$from,
+                         lineTo = distance_df_small_edges$to,
+                         group = factor(df_model_data_n$type,
+                                        c("data", "model")),
+                         levelColors = c(clr_choice, "#000000"))
+```
+
+```{r}
+#| label: five-gau-tsne-projs
 projection <- cbind(
@@
   select(-type)
-
+```
+
+```{r}
+#| label: langevitour-five-gau-umap-proj
+#| eval: false
+
+df_model_data_n <- bind_rows(data_gau, df_b)
+
+langevitour::langevitour(df_model_data_n[1:(length(df_model_data_n)-1)],
+                         lineFrom = distance_df_small_edges$from,
+                         lineTo = distance_df_small_edges$to,
+                         group = factor(df_model_data_n$type,
+                                        c("data", "model")),
+                         levelColors = c(clr_choice, "#000000"))
+```
+
+```{r}
+#| label: five-gau-umap-projs
 projection <- cbind(
@@
   dplyr::filter(distance < benchmark_gau1)
-
+```
+
+```{r}
+#| label: langevitour-five-gau-pacmap-proj
+#| eval: false
+
+df_model_data_n <- bind_rows(data_gau, df_b)
+
+langevitour::langevitour(df_model_data_n[1:(length(df_model_data_n)-1)],
+                         lineFrom = distance_df_small_edges$from,
+                         lineTo = distance_df_small_edges$to,
+                         group = factor(df_model_data_n$type,
+                                        c("data", "model")),
+                         levelColors = c(clr_choice, "#000000"))
+```
+
+```{r}
+#| label: five-gau-pacmap-projs
 ## First projection
@@
 ```{r}
-#| label: langevitour-pbmc-umap-author-proj
+#| label: langevitour-c-shaped-tsne-proj
 #| eval: false
@@
 ```{r}
-#| label: langevitour-best-fit-two-curvy-proj-tsne
+#| label: langevitour-best-fit-mnist-proj-tsne
 #| eval: false

### 2025-04-04 paper.qmd
Updated Figure 12

@@
 error_df_one_curvy_abs <- error_df_one_curvy_abs |>
-  mutate(error_cat_n = if_else(tSNE1 <= 0.25 & tSNE2 <= 0.25, "high", "low")) |> ## high_error points
-  mutate(error_cat_n2 = if_else(tSNE2 >= 1.2, "high", "low")) ## corner points
+  mutate(error_cat_n = if_else(tSNE1 <= 0.25 & tSNE2 <= 0.25, "selected", "deselected")) |> ## high_error points
+  mutate(error_cat_n2 = if_else(tSNE2 >= 1.2, "selected", "deselected")) ## corner points
 
@@
           y = density,
-          colour = factor(error_cat_n, levels = c("low", "high")))) +
+          colour = factor(error_cat_n, levels = c("deselected", "selected")))) +
   geom_point(alpha=0.7) +
@@
             y = density,
-            colour = factor(error_cat_n2, levels = c("low", "high")))) +
+            colour = factor(error_cat_n2, levels = c("deselected", "selected")))) +
   geom_point(alpha=0.7) +
@@
              colour = factor(error_cat_n2,
-                             levels = c("low", "high")))) +
+                             levels = c("deselected", "selected")))) +
   geom_point(alpha=0.7) +
@@
              colour = factor(error_cat_n,
-                             levels = c("low", "high")))) +
+                             levels = c("deselected", "selected")))) +
   geom_point(alpha=0.7) +
@@
   mutate(type = factor(error_df_one_curvy_abs$error_cat_n2,
-                          levels=c("low", "high")))
+                          levels=c("deselected", "selected")))
 
@@
                          group = factor(df_model_data_n$type,
-                                        c("low", "high", "model")),
+                                        c("deselected", "selected", "model")),
                          levelColors = c('#d9d9d9', '#800026', "#000000"))
@@
   mutate(cluster = factor(error_df_one_curvy_abs$error_cat_n2,
-                          levels=c("low", "high")))
+                          levels=c("deselected", "selected")))
 
@@
   select(-cluster) |>
-  mutate(cluster = error_df_one_curvy_abs$error_cat_n)
+  mutate(cluster = factor(error_df_one_curvy_abs$error_cat_n,
+                          levels=c("deselected", "selected")))
 

### 2025-04-04 paper.qmd
Updated plot_proj

@@
   axis_text_size = 4, 
-  is_category = FALSE) +
+  is_color = FALSE) +
   theme(aspect.ratio=1, 
@@
     axis_text_size = 2, 
-    is_category = FALSE) +
+    is_color = FALSE) +
     theme(aspect.ratio = 1,
@@
     axis_text_size = 4, 
-    is_category = FALSE) +
+    is_color = FALSE) +
     theme(aspect.ratio = 1, 
@@
   axis_text_size = 2, 
-  is_category = FALSE) #+
+  is_color = FALSE) #+
   #geom_segment(
@@
     axis_text_size = 4, 
-    is_category = FALSE) +
+    is_color = FALSE) +
     theme(
@@
     axis_text_size = 4, 
-    is_category = TRUE) +
+    is_color = TRUE) +
     scale_color_manual(values = c('#d9d9d9', '#800026')) +
@@
   axis_text_size = 5,
-  is_category = FALSE)
+  is_color = FALSE)
 
@@
     axis_text_size = 5,
-    is_category = FALSE)
+    is_color = FALSE)
 
@@
     axis_text_size = 5,
-    is_category = FALSE)
+    is_color = FALSE)
 ```
@@
   axis_text_size = 3, 
-  is_category = TRUE) +
+  is_color = TRUE) +
   scale_color_manual(values = c("#999999",'#ff7f00')) +
@@
     axis_text_size = 3, 
-    is_category = TRUE) +
+    is_color = TRUE) +
     scale_color_manual(values = c("#d9d9d9",'#ff7f00')) +
@@
     axis_text_size = 3, 
-    is_category = TRUE) +
+    is_color = TRUE) +
     scale_color_manual(values = c("#999999",'#ff7f00')) +

### 2025-04-04 scripts/additional_functions.R

@@ -188,25 +188,27 @@ plot_proj <- function(projected_df, model_df, axes, circle,
                       plot_limits, title, cex = 2,
                       position = c(0.92, 0.92),
                       axis_text_size = 3,
-                      is_category = FALSE) {
+                      is_color = FALSE) {
 
-  if(is_category == FALSE) {
+  initial_plot <- projected_df |>
+    ggplot(
+      aes(
+        x = proj1,
+        y = proj2)) +
+    geom_segment(
+      data = model_df,
+      aes(
+        x = proj1_from,
+        y = proj2_from,
+        xend = proj1_to,
+        yend = proj2_to),
+      color = line_param[3],
+      linewidth = as.numeric(line_param[1]),
+      alpha = as.numeric(line_param[2]))
 
-    initial_plot <- projected_df |>
-      ggplot(
-        aes(
-          x = proj1,
-          y = proj2)) +
-      geom_segment(
-        data = model_df,
-        aes(
-          x = proj1_from,
-          y = proj2_from,
-          xend = proj1_to,
-          yend = proj2_to),
-        color = line_param[3],
-        linewidth = as.numeric(line_param[1]),
-        alpha = as.numeric(line_param[2])) +
+  if(is_color == FALSE) {
+
+    initial_plot <- initial_plot +
       geom_point(
         size = as.numeric(point_param[1]),
         alpha = as.numeric(point_param[2]),
@@ -214,22 +216,7 @@ plot_proj <- function(projected_df, model_df, axes, circle,
 
   } else {
 
-    initial_plot <- projected_df |>
-      ggplot(
-        aes(
-          x = proj1,
-          y = proj2,
-          colour = cluster)) +
-      geom_segment(
-        data = model_df,
-        aes(
-          x = proj1_from,
-          y = proj2_from,
-          xend = proj1_to,
-          yend = proj2_to),
-        color = line_param[3],
-        linewidth = as.numeric(line_param[1]),
-        alpha = as.numeric(line_param[2])) +
+    initial_plot <- initial_plot +
       geom_point(
         size = as.numeric(point_param[1]),
         alpha = as.numeric(point_param[2]))

### 2025-04-04 paper.qmd
Updated the plot project

@@
 
-projected_df <- proj_obj1$projected_df
-model_df <- proj_obj1$model_df
-axes <- proj_obj1$axes
-circle <- proj_obj1$circle
-
 ## For true model
@@
 
+proj_obj1_true <- proj_obj1
+proj_obj1_true[["model_df"]] <- true_model_df_proj
+
 two_curvy_proj_true <- plot_proj(
-  projected_df = projected_df, 
-  model_df = true_model_df_proj, 
-  axes = axes, 
-  circle = circle, 
+  proj_obj = proj_obj1_true, 
   point_param = c(0.5, 0.5, clr_choice), # size, alpha, color
@@
 
+proj_obj1[["axes"]] <- axes
+proj_obj1[["circle"]] <- circle
+
 two_curvy_proj_tsne_first_model1 <- plot_proj(
-    projected_df = projected_df, 
-    model_df = model_df, 
-    axes = axes, 
-    circle = circle, 
+    proj_obj = proj_obj1, 
     point_param = c(0.5, 0.2, clr_choice), # size, alpha, color
@@
 
-projected_df <- proj_obj2$projected_df
-model_df <- proj_obj2$model_df
-axes <- proj_obj2$axes
-circle <- proj_obj2$circle
-
 projected_true_model <- as.matrix(scaled_two_curvy_data_true_model) %*% model_prj2
@@
 
+proj_obj2_true <- proj_obj2
+proj_obj2_true[["model_df"]] <- true_model_df_proj
+
 two_curvy_proj_true2 <- plot_proj(
-    projected_df = projected_df, 
-    model_df = true_model_df_proj, 
-    axes = axes, 
-    circle = circle, 
+    proj_obj = proj_obj2_true, 
     point_param = c(0.5, 0.5, clr_choice), # size, alpha, color
@@
 
+proj_obj2[["axes"]] <- axes
+proj_obj2[["circle"]] <- circle
+
 two_curvy_proj_tsne_all_model1 <- plot_proj(
-  projected_df = projected_df, 
-  model_df = model_df, 
-  axes = axes, 
-  circle = circle, 
+  proj_obj = proj_obj2, 
   point_param = c(0.1, 0.5, clr_choice), # size, alpha, color
@@
 
-projected_df <- proj_obj1$projected_df
-model_df <- proj_obj1$model_df
-axes <- proj_obj1$axes
-circle <- proj_obj1$circle
-
 ## Projection with Delauany triangulation
@@
 two_curvy_proj_first_model1_umap <- plot_proj(
-    projected_df = projected_df, 
-    model_df = model_df, 
-    axes = axes, 
-    circle = circle, 
+    proj_obj = proj_obj1, 
     point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
@@
 
+proj_obj1_temp1 <- proj_obj1
+proj_obj1_temp2 <- proj_obj1
+proj_obj1_temp1[["model_df"]] <- model_df1
+
 two_curvy_proj_model_delaunay_umap <- plot_proj(
-    projected_df = projected_df, 
-    model_df = model_df1, 
-    axes = axes, 
-    circle = circle, 
+    proj_obj = proj_obj1_temp1, 
     point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
@@
 
+proj_obj1_temp2[["model_df"]] <- model_df2
+
 two_curvy_proj_model_benchmark2_umap <- plot_proj(
-    projected_df = projected_df, 
-    model_df = model_df2, 
-    axes = axes, 
-    circle = circle, 
+    proj_obj = proj_obj1_temp2, 
     point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
@@
 
-projected_df <- proj_obj1$projected_df
-model_df <- proj_obj1$model_df
-axes2 <- proj_obj1$axes
-circle <- proj_obj1$circle
-
-## To add error category
-projected_df <- projected_df |>
-  mutate(cluster = factor(error_df_one_curvy_abs$error_cat_n2,
-                          levels=c("deselected", "selected")))
+proj_obj1[["cluster"]] <- factor(error_df_one_curvy_abs$error_cat_n2,
+                          levels=c("deselected", "selected"))
 
 five_c_shaped_proj_tsne_model1 <- plot_proj(
-    projected_df = projected_df, 
-    model_df = model_df, 
-    axes = axes2, 
-    circle = circle, 
+    proj_obj = proj_obj1,
     point_param = c(1.5, 0.5), # size, alpha, color
@@
 
-projected_df <- projected_df |>
-  select(-cluster) |>
-  mutate(cluster = factor(error_df_one_curvy_abs$error_cat_n,
-                          levels=c("deselected", "selected")))
+proj_obj1[["cluster"]] <- factor(error_df_one_curvy_abs$error_cat_n,
+                          levels=c("deselected", "selected"))
 
 five_c_shaped_proj_tsne_model1_selected <- plot_proj(
-    projected_df = projected_df, 
-    model_df = model_df, 
-    axes = axes2, 
-    circle = circle, 
+    proj_obj = proj_obj1, 
     point_param = c(1.5, 0.5), # size, alpha, color
@@
 
-projected_df <- proj_obj1$projected_df
-model_df <- proj_obj1$model_df
-axes <- proj_obj1$axes
-circle <- proj_obj1$circle
-
 pbmc_proj_umap_model1 <- plot_proj(
-  projected_df = projected_df, 
-  model_df = model_df, 
-  axes = axes, 
-  circle = circle, 
+  proj_obj = proj_obj1, 
   point_param = c(1.5, 0.2, "#a65628"), # size, alpha, color
@@
 
-projected_df <- proj_obj2$projected_df
-model_df <- proj_obj2$model_df
-axes <- proj_obj2$axes
-circle <- proj_obj2$circle
-
 pbmc_proj_umap_model2 <- plot_proj(
-    projected_df = projected_df, 
-    model_df = model_df, 
-    axes = axes, 
-    circle = circle, 
+    proj_obj = proj_obj2, 
     point_param = c(1.5, 0.2, "#a65628"), # size, alpha, color
@@
 
-projected_df <- proj_obj1$projected_df
-model_df <- proj_obj1$model_df
-axes <- proj_obj1$axes
-circle <- proj_obj1$circle
-
 pbmc_proj_umap_model1_best <- plot_proj(
-    projected_df = projected_df, 
-    model_df = model_df, 
-    axes = axes, 
-    circle = circle, 
+    proj_obj = proj_obj1,
     point_param = c(1.5, 0.2, "#999999"), # size, alpha, color
@@
 
-projected_df <- proj_obj2$projected_df
-model_df <- proj_obj2$model_df
-axes <- proj_obj2$axes
-circle <- proj_obj2$circle
-
 pbmc_proj_umap_model2_best <- plot_proj(
-    projected_df = projected_df, 
-    model_df = model_df, 
-    axes = axes, 
-    circle = circle, 
+    proj_obj = proj_obj2,
     point_param = c(1.5, 0.2, "#999999"), # size, alpha, color
@@
 
-projected_df <- proj_obj1$projected_df |>
-  mutate(cluster = tsne_minst_scaled$cluster)
-
-model_df <- proj_obj1$model_df
-axes <- proj_obj1$axes
-circle <- proj_obj1$circle
+proj_obj1[["cluster"]] <- tsne_minst_scaled$cluster
 
 mnist_proj_tsne_model1 <- plot_proj(
-  projected_df = projected_df, 
-  model_df = model_df, 
-  axes = axes, 
-  circle = circle, 
+  proj_obj = proj_obj1, 
   point_param = c(0.5, 0.5), # size, alpha, color
@@
 mnist_proj_tsne_model1_link <- plot_proj(
-    projected_df = projected_df, 
-    model_df = model_df, 
-    axes = axes, 
-    circle = circle, 
+    proj_obj = proj_obj1,
     point_param = c(1.2, 0.5), # size, alpha, color
@@
 
-projected_df <- proj_obj2$projected_df |>
-  mutate(cluster = tsne_minst_scaled$cluster)
-
-model_df <- proj_obj2$model_df
-axes <- proj_obj2$axes
-circle <- proj_obj2$circle
+proj_obj2[["cluster"]] <- tsne_minst_scaled$cluster
 
 mnist_proj_tsne_model2 <- plot_proj(
-    projected_df = projected_df, 
-    model_df = model_df, 
-    axes = axes, 
-    circle = circle, 
+    proj_obj = proj_obj1,
     point_param = c(0.5, 0.3), # size, alpha, color
@@
 mnist_proj_tsne_model2_link <- plot_proj(
-    projected_df = projected_df, 
-    model_df = model_df, 
-    axes = axes, 
-    circle = circle, 
+    proj_obj = proj_obj1, 
     point_param = c(1.2, 0.5), # size, alpha, color

### 2025-04-04 scripts/additional_functions.R

@@ -182,13 +182,18 @@ get_projection <- function(projection, proj_scale, scaled_data,
 
 # Plot projection
 
-plot_proj <- function(projected_df, model_df, axes, circle,
+plot_proj <- function(proj_obj,
                       point_param = c(1.5, 0.5, "#000000"), # size, alpha, color
                       line_param = c(0.5, 0.5, "#000000"), #linewidth, alpha
                       plot_limits, title, cex = 2,
                       position = c(0.92, 0.92),
                       axis_text_size = 3,
                       is_color = FALSE) {
+
+  projected_df <- proj_obj$projected_df
+  model_df <- proj_obj$model_df
+  axes <- proj_obj$axes
+  circle <- proj_obj$circle
 
   if(is_color == FALSE) {
 
@@ -203,6 +208,9 @@ plot_proj <- function(projected_df, model_df, axes, circle,
         color = point_param[3])
 
   } else {
+
+    projected_df <- projected_df |>
+      dplyr::mutate(cluster = proj_obj$cluster)
 
     initial_plot <- ggplot() +
       geom_point(

### 2025-04-04 paper.qmd
Tried to find another good layout for PBMC3k

@@
 
-umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_12_min_dist_0.99.rds")
+# umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_12_min_dist_0.99.rds")
+umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_3_min_dist_0.6.rds")
 nldr2 <- umap_pbmc |>
-  ggplot(aes(x = UMAP1,
-             y = UMAP2))+
+  ggplot(aes(x = emb1,
+             y = emb2))+
   geom_point(alpha=0.1, size=1, colour=clr_choice) +
@@
 nldr2c <- umap_pbmc |>
-  ggplot(aes(x = UMAP1,
-             y = UMAP2))+
+  ggplot(aes(x = emb1,
+             y = emb2))+
   geom_point(alpha=0.1, size=1, colour='#999999') +
@@
              y = UMAP2))+
-  geom_point(alpha=0.1, size=1, colour='#f781bf') +
+  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
   interior_annotation("c")
@@
              y = tSNE2))+
-  geom_point(alpha=0.1, size=1, colour='#984ea3') +
+  geom_point(alpha=0.1, size=1, colour='#377eb8') +
   interior_annotation("d")
@@
              y = tSNE2))+
-  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
+  geom_point(alpha=0.1, size=1, colour='#4daf4a') +
   interior_annotation("e")
@@
              y = PHATE2))+
-  geom_point(alpha=0.1, size=1, colour='#377eb8') +
+  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
   interior_annotation("f")
@@
              y = TriMAP2))+
-  geom_point(alpha=0.1, size=1, colour='#4daf4a') +
+  geom_point(alpha=0.1, size=1, colour='#984ea3') +
   interior_annotation("g")
@@
              y = PaCMAP2))+
-  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
+  geom_point(alpha=0.1, size=1, colour='#f781bf') +
   interior_annotation("h")
@@
 error_pbmc_umap2 <- read_rds("data/pbmc3k/error_pbmc_umap_5_min_dist_0.01.rds")
-# error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_15_min_dist_0.99.rds")
-error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_12_min_dist_0.99.rds")
+# error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_12_min_dist_0.99.rds")
+error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_3_min_dist_0.6.rds")
 error_pbmc_tsne <- read_rds("data/pbmc3k/error_pbmc_tsne_5.rds")
@@
   ungroup()
+
+error_pbmc <- error_pbmc |>
+  mutate(method = factor(method,
+                         levels = c("UMAP_3_min_dist_0.6", "UMAP_30_min_dist_0.3", "UMAP_5_min_dist_0.01","tsne_5", "tsne_30", "phate_5", "trimap_12_4_3", "pacmap")))
 ```
@@
   scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 13, 17, 21, 26)]) +
-  scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a','#ff7f00',
-                              '#984ea3','#999999','#a65628','#f781bf'))
+  scale_color_manual(values=c('#999999','#a65628','#e41a1c','#377eb8','#4daf4a','#ff7f00','#984ea3','#f781bf'))
 
@@
 #| label: umap-pbmc-read-best
-umap_pbmc2 <- read_rds("data/pbmc3k/pbmc_umap_12_min_dist_0.99.rds")
+umap_pbmc2 <- read_rds("data/pbmc3k/pbmc_umap_3_min_dist_0.6.rds")
 pbmc_scaled_obj <- gen_scaled_data(
@@
 ## Compute hexbin parameters
-num_bins_x_pbmc <- 30
+num_bins_x_pbmc <- 29
 
@@
     aes(
-      x = UMAP1,
-      y = UMAP2
+      x = emb1,
+      y = emb2
     ),

### 2025-04-04 paper.qmd
Five Gaussian clusters with enough bins..earlier was too much

@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 40
+num_bins_x_gau1 <- 16
 lim1 <- gau1_scaled_obj$lim1
@@
   filter(distance < benchmark_gau1) |>
-  mutate(ID = row_number()) |>
-  filter(!(ID %in% c(97)))
+  mutate(ID = row_number()) #|>
+  #filter(!(ID %in% c(97)))
 

### 2025-04-04 paper.qmd
Updated Figure 11

@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 26
+num_bins_x_gau1 <- 22
 lim1 <- gau1_scaled_obj$lim1
@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 90
+num_bins_x_gau1 <- 62
 lim1 <- gau1_scaled_obj$lim1
@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 16
+num_bins_x_gau1 <- 25
 lim1 <- gau1_scaled_obj$lim1

### 2025-04-04 paper.qmd
Updated Figure 11

@@
     color = "#000000",
-    alpha = 0.4,
+    #alpha = 0.4,
     linewidth = 0.5) +
@@
     color = "#000000",
-    alpha = 0.4,
-    linewidth = 0.1) +
+    #alpha = 0.4,
+    linewidth = 0.5) +
   geom_segment(

### 2025-04-04 paper.qmd
Added initial cluster labels for PBMC

@@
 
+umap_pbmc_scaled_with_cluster <- umap_pbmc_scaled |>
+  mutate(cluster = if_else(UMAP1 <= 0.25, "cluster1", if_else(UMAP1 >= 0.625, "cluster2", "cluster3"))) |>
+  mutate(cluster = factor(cluster, levels = c("cluster1", "cluster2", "cluster3")))
+
 tr_from_to_df_pbmc <- inner_join(
@@
   geom_point(
-    data = umap_pbmc_scaled,
+    data = umap_pbmc_scaled_with_cluster,
     aes(
       x = UMAP1,
-      y = UMAP2
+      y = UMAP2,
+      color = cluster
     ),
     alpha = 0.2,
-    size = 0.1,
-    color = '#a65628'
+    size = 0.1#,
+    #color = '#a65628'
   )  +
+  scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada")) +
   interior_annotation("a1", cex = 2) +
@@
 
+umap_pbmc_scaled_best_with_cluster <- umap_pbmc_scaled_best |>
+  mutate(cluster = umap_pbmc_scaled_with_cluster$cluster)
+
 trimesh_removed_pbmc_best <- ggplot() + 
@@
   geom_point(
-    data = umap_pbmc_scaled_best,
+    data = umap_pbmc_scaled_best_with_cluster,
     aes(
       x = emb1,
-      y = emb2
+      y = emb2,
+      color = cluster
     ),
-    alpha = 0.3,
-    color = "#999999"
+    alpha = 0.3#,
+    #color = "#999999"
   )  +
+  scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada")) +
   interior_annotation("b1", cex = 2) +

### 2025-04-04 paper.qmd
Updated Figure 14

@@
 # umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_12_min_dist_0.99.rds")
-umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_3_min_dist_0.6.rds")
+umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_5_min_dist_0.8.rds")
 nldr2 <- umap_pbmc |>
@@
 ## Compute hexbin parameters
-num_bins_x_pbmc <- 60
+num_bins_x_pbmc <- 26
 
@@
 
-df_model_data_pbmc_n <- bind_rows(df_b_pbmc, data_pbmc)
+data_pbmc_n <- data_pbmc |>
+  select(-type) |>
+  mutate(type = as.character(umap_pbmc_scaled_with_cluster$cluster))
+
+df_model_data_pbmc_n <- bind_rows(df_b_pbmc, data_pbmc_n)
 
@@
                          group = factor(df_model_data_pbmc_n$type,
-                                        c("data", "model")),
-                         levelColors = c("#a65628", "#000000"))
+                                        c("cluster1", "cluster2", "cluster3", "model")),
+                         levelColors = c("#8dd3c7", "#fdb462", "#bebada", "#000000"))
 ```
@@
 
+proj_obj1[["cluster"]] <- as.character(umap_pbmc_scaled_with_cluster$cluster)
+
 pbmc_proj_umap_model1 <- plot_proj(
   proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2, "#a65628"), # size, alpha, color
+  point_param = c(1.5, 0.2), # size, alpha, color
   plot_limits = c(-0.65, 0.65), 
@@
   axis_text_size = 5,
-  is_color = FALSE)
+  is_color = TRUE) + scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))
 
@@
 
+proj_obj2[["cluster"]] <- as.character(umap_pbmc_scaled_with_cluster$cluster)
+
 pbmc_proj_umap_model2 <- plot_proj(
     proj_obj = proj_obj2, 
-    point_param = c(1.5, 0.2, "#a65628"), # size, alpha, color
+    point_param = c(1.5, 0.2), # size, alpha, color
     plot_limits = c(-1, 0.75), 
@@
     axis_text_size = 5,
-    is_color = FALSE)
+    is_color = TRUE) + scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))
 
@@
 # error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_12_min_dist_0.99.rds")
-error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_3_min_dist_0.6.rds")
+# error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_3_min_dist_0.6.rds")
+# error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_4_min_dist_0.6.rds")
+# error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_5_min_dist_0.8.rds")
+error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_5_min_dist_0.8.rds")
 error_pbmc_tsne <- read_rds("data/pbmc3k/error_pbmc_tsne_5.rds")
@@
   mutate(method = factor(method,
-                         levels = c("UMAP_3_min_dist_0.6", "UMAP_30_min_dist_0.3", "UMAP_5_min_dist_0.01","tsne_5", "tsne_30", "phate_5", "trimap_12_4_3", "pacmap")))
+                         levels = c("UMAP_5_min_dist_0.8", "UMAP_30_min_dist_0.3", "UMAP_5_min_dist_0.01","tsne_5", "tsne_30", "phate_5", "trimap_12_4_3", "pacmap")))
 ```
@@
 #| label: umap-pbmc-read-best
-umap_pbmc2 <- read_rds("data/pbmc3k/pbmc_umap_3_min_dist_0.6.rds")
+umap_pbmc2 <- read_rds("data/pbmc3k/pbmc_umap_5_min_dist_0.8.rds")
 pbmc_scaled_obj <- gen_scaled_data(
@@
 ## Compute hexbin parameters
-num_bins_x_pbmc <- 29
+num_bins_x_pbmc <- 18
 
@@
 distance_df_small_edges_pbmc <- distance_pbmc |>
-  filter(distance < benchmark_pbmc)
-
-tr_from_to_df_pbmc <- inner_join(
+  filter(distance < 0.07) #benchmark_pbmc
+
+tr_from_to_df_pbmc <- right_join(
   tr_from_to_df_pbmc, distance_df_small_edges_pbmc,
@@
 
-df_model_data_pbmc_n <- bind_rows(df_b_pbmc, data_pbmc)
+df_model_data_pbmc_n <- bind_rows(df_b_pbmc, data_pbmc_n)
 
@@
                          group = factor(df_model_data_pbmc_n$type,
-                                        c("data", "model")),
-                         levelColors = c("#999999", "#000000"))
+                                        c("cluster1", "cluster2", "cluster3", "model")),
+                         levelColors = c("#8dd3c7", "#fdb462", "#bebada", "#000000"))
 ```
@@
 
+proj_obj1[["cluster"]] <- as.character(umap_pbmc_scaled_with_cluster$cluster)
+
 pbmc_proj_umap_model1_best <- plot_proj(
     proj_obj = proj_obj1,
-    point_param = c(1.5, 0.2, "#999999"), # size, alpha, color
+    point_param = c(1.5, 0.2), # size, alpha, color
     plot_limits = c(-0.65, 0.65), 
@@
 
+proj_obj2[["cluster"]] <- as.character(umap_pbmc_scaled_with_cluster$cluster)
+
 pbmc_proj_umap_model2_best <- plot_proj(
     proj_obj = proj_obj2,
-    point_param = c(1.5, 0.2, "#999999"), # size, alpha, color
+    point_param = c(1.5, 0.2), # size, alpha, color
     plot_limits = c(-1, 0.75), 
@@
     axis_text_size = 5,
-    is_color = FALSE)
+    is_color = TRUE) + scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))
 ```

### 2025-04-05 paper.qmd
Updated with PBMC new layout (best)

@@
 ## Compute hexbin parameters
-num_bins_x_pbmc <- 26
+num_bins_x_pbmc <- 22
 
@@
 ## Compute hexbin parameters
-num_bins_x_pbmc <- 18
+num_bins_x_pbmc <- 13
 
@@
 distance_df_small_edges_pbmc <- distance_pbmc |>
-  filter(distance < 0.07) #benchmark_pbmc
+  filter(distance < 0.1) #benchmark_pbmc
 
@@
 The PBMC3k data ($n =  2622$) has three clusters in $9\\text{-}D$, where two of them being close and one separated from the others.
-Two \\gD{} projection from a tour on $9\\text{-}D$ of the model fit with @fig-pbmc-mse a ($a_1 = 0.02$, $b = 3480/349 (60, 58)$) shows three-well separated clusters with big separations.
-On the other hand, the model fit with @fig-pbmc-mse b ($a_1 = 0.04$, $b = 1020/347 (30, 34)$) shows two close clusters, and one separated cluster.
+Two \\gD{} projection from a tour on $9\\text{-}D$ of the model fit with @fig-pbmc-mse a ($a_1 = 0.06$, $b = 616/86 (22, 28)$) shows three-well separated clusters with big separations.
+On the other hand, the model fit with @fig-pbmc-mse b ($a_1 = 0.1$, $b = 234/98 (13, 18)$) shows two close clusters, and one separated cluster.
 The edge connected the two clusters proves the closeness of the two clusters.

### 2025-04-05 scripts/pbmc3k/3_gen_umap_diff_param.R

@@ -5,8 +5,8 @@ library(ggplot2)
 library(quollr)
 set.seed(20240110)
 
-n_neighbors <- 5
-min_dist <- 0.81
+n_neighbors <- 6
+min_dist <- 0.85
 
 ## Select PCs
 training_data_pbmc <- read_rds("data/pbmc3k/pbmc_pca_50.rds")

### 2025-04-05 paper.qmd
Removed long edges manually
Updated Figure 14

@@
 trimesh_removed_pbmc <- ggplot() + 
-  geom_segment(data = tr_from_to_df_pbmc, 
-               aes(
-                 x = x_from, 
-                 y = y_from, 
-                 xend = x_to, 
-                 yend = y_to),
-               colour = "#000000",
-               linewidth = 1) +
   geom_point(
@@
   )  +
+  geom_segment(data = tr_from_to_df_pbmc, 
+               aes(
+                 x = x_from, 
+                 y = y_from, 
+                 xend = x_to, 
+                 yend = y_to),
+               colour = "#000000",
+               linewidth = 1) +
   scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada")) +
@@
 
+tr_from_to_df_pbmc <- tr_from_to_df_pbmc |>
+  mutate(ID = row_number()) |>
+  filter(!(ID %in% c(19, 20, 23, 24, 29, 33, 48,
+                     183, 181, 178, 176))) #19
+
 tr_from_to_df_pbmc <- right_join(
@@
 trimesh_removed_pbmc_best <- ggplot() + 
-  geom_segment(data = tr_from_to_df_pbmc, 
-               aes(
-                 x = x_from, 
-                 y = y_from, 
-                 xend = x_to, 
-                 yend = y_to),
-               colour = "#000000",
-               linewidth = 1) +
   geom_point(
@@
   ) +
+    geom_segment(data = tr_from_to_df_pbmc, 
+               aes(
+                 x = x_from, 
+                 y = y_from, 
+                 xend = x_to, 
+                 yend = y_to),
+               colour = "#000000",
+               linewidth = 1) +
   scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada")) +

### 2025-04-05 paper.qmd
Tried to update figure 16

@@
   tr_from_to_df_pbmc, distance_df_small_edges_pbmc,
-  by = c("from", "to"))
+  by = c("from", "to")) |>
+  select(-ID)
 
@@
 
+# tr_from_to_df_mnist <- tr_from_to_df_mnist |>
+#   mutate(ID = row_number()) |>
+#   filter(!(ID %in% c(1205, 1629, 1170)))
+
 tsne_minst_scaled <- tsne_minst_scaled |>
@@
 trimesh_removed_mnist <- ggplot() + 
+  geom_point(data = tsne_minst_scaled,
+             aes(
+               x = tSNE1,
+               y = tSNE2,
+               colour = cluster
+             ),
+             alpha=0.2)+
     geom_segment(data = tr_from_to_df_mnist, 
@@
                colour = "#000000") +
-  geom_point(data = tsne_minst_scaled,
-             aes(
-               x = tSNE1,
-               y = tSNE2,
-               colour = cluster
-             ),
-             alpha=0.2)+
   scale_color_manual(values=c("#999999",'#ff7f00')) +

### 2025-04-06 paper.qmd
di changes

--- 2025-04-05 paper.qmd
@@
 
-# umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_12_min_dist_0.99.rds")
-umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_5_min_dist_0.8.rds")
+umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_12_min_dist_0.99.rds")
 nldr2 <- umap_pbmc |>
-  ggplot(aes(x = emb1,
-             y = emb2))+
+  ggplot(aes(x = UMAP1,
+             y = UMAP2))+
   geom_point(alpha=0.1, size=1, colour=clr_choice) +
@@
 nldr2c <- umap_pbmc |>
-  ggplot(aes(x = emb1,
-             y = emb2))+
+  ggplot(aes(x = UMAP1,
+             y = UMAP2))+
   geom_point(alpha=0.1, size=1, colour='#999999') +
@@
              y = UMAP2))+
-  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
+  geom_point(alpha=0.1, size=1, colour='#f781bf') +
   interior_annotation("c")
@@
              y = tSNE2))+
-  geom_point(alpha=0.1, size=1, colour='#377eb8') +
+  geom_point(alpha=0.1, size=1, colour='#984ea3') +
   interior_annotation("d")
@@
              y = tSNE2))+
-  geom_point(alpha=0.1, size=1, colour='#4daf4a') +
+  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
   interior_annotation("e")
@@
              y = PHATE2))+
-  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
+  geom_point(alpha=0.1, size=1, colour='#377eb8') +
   interior_annotation("f")
@@
              y = TriMAP2))+
-  geom_point(alpha=0.1, size=1, colour='#984ea3') +
+  geom_point(alpha=0.1, size=1, colour='#4daf4a') +
   interior_annotation("g")
@@
              y = PaCMAP2))+
-  geom_point(alpha=0.1, size=1, colour='#f781bf') +
+  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
   interior_annotation("h")
@@
 trimesh_two_curvy_removed1_tsne_with_data <- ggplot() +
-  geom_segment(data = tr_from_to_df_two_curvy,
-               aes(
-                 x = x_from,
-                 y = y_from,
-                 xend = x_to,
-                 yend = y_to),
-               colour = "#000000") +
   geom_point(data = tsne_two_curvy_scaled,
@@
              size = 0.5) +
-  interior_annotation("b", sc_ltr_pos, cex = 1.5)
+  geom_segment(data = tr_from_to_df_two_curvy,
+               aes(
+                 x = x_from,
+                 y = y_from,
+                 xend = x_to,
+                 yend = y_to),
+               colour = "#000000") +
+  interior_annotation("a", sc_ltr_pos, cex = 1.5) +
+  theme(aspect.ratio = 1)
 ```
@@
 
+projected_df <- proj_obj1$projected_df
+model_df <- proj_obj1$model_df
+axes <- proj_obj1$axes
+circle <- proj_obj1$circle
+
 ## For true model
@@
 
-proj_obj1_true <- proj_obj1
-proj_obj1_true[["model_df"]] <- true_model_df_proj
-
 two_curvy_proj_true <- plot_proj(
-  proj_obj = proj_obj1_true, 
+  projected_df = projected_df, 
+  model_df = true_model_df_proj, 
+  axes = axes, 
+  circle = circle, 
   point_param = c(0.5, 0.5, clr_choice), # size, alpha, color
@@
   axis_text_size = 4, 
-  is_color = FALSE) +
+  is_category = FALSE) +
   theme(aspect.ratio=1, 
@@
 
-proj_obj1[["axes"]] <- axes
-proj_obj1[["circle"]] <- circle
-
 two_curvy_proj_tsne_first_model1 <- plot_proj(
-    proj_obj = proj_obj1, 
+    projected_df = projected_df, 
+    model_df = model_df, 
+    axes = axes, 
+    circle = circle, 
     point_param = c(0.5, 0.2, clr_choice), # size, alpha, color
@@
     plot_limits = c(-0.6, 0.6), 
-    title = "c", 
-    cex = 1.5, 
+    title = "b", 
+    cex = 1, 
     position = sc_ltr_pos,
     axis_text_size = 2, 
-    is_color = FALSE) +
+    is_category = FALSE) +
     theme(aspect.ratio = 1,
@@
 
+projected_df <- proj_obj2$projected_df
+model_df <- proj_obj2$model_df
+axes <- proj_obj2$axes
+circle <- proj_obj2$circle
+
 projected_true_model <- as.matrix(scaled_two_curvy_data_true_model) %*% model_prj2
@@
 
-proj_obj2_true <- proj_obj2
-proj_obj2_true[["model_df"]] <- true_model_df_proj
-
 two_curvy_proj_true2 <- plot_proj(
-    proj_obj = proj_obj2_true, 
+    projected_df = projected_df, 
+    model_df = true_model_df_proj, 
+    axes = axes, 
+    circle = circle, 
     point_param = c(0.5, 0.5, clr_choice), # size, alpha, color
@@
     axis_text_size = 4, 
-    is_color = FALSE) +
+    is_category = FALSE) +
     theme(aspect.ratio = 1, 
@@
 
-proj_obj2[["axes"]] <- axes
-proj_obj2[["circle"]] <- circle
-
 two_curvy_proj_tsne_all_model1 <- plot_proj(
-  proj_obj = proj_obj2, 
+  projected_df = projected_df, 
+  model_df = model_df, 
+  axes = axes, 
+  circle = circle, 
   point_param = c(0.1, 0.5, clr_choice), # size, alpha, color
@@
   plot_limits = c(-0.5, 0.5), 
-  title = "d", 
+  title = "c", 
   cex = 1, 
@@
   axis_text_size = 2, 
-  is_color = FALSE) #+
+  is_category = FALSE) +
+  theme(aspect.ratio = 1)
+#+
   #geom_segment(
@@
 #| label: fig-two-curvy-true-proj
-#| fig-cap: "Two \\gD{} projections from $4\\text{-}D$, for the two non-linear clusters data (grey points) and true model (blue lines) are shown.
+#| fig-cap: "Two \\gD{} projections from $4\\text{-}D$, for the two non-linear clusters data (blue points) and true model (black lines) are shown.
 The top cluster in (a) shows uniformly distributed data along a C-shape, while in the bottom cluster, the data is more concentrated in one corner of the curvy shape.
@@
 #| fig-cap: "Lifting the fitted model into \\pD{}.
-Starting from a default tSNE layout (a) of non-linear clusters data ($n =  2000$ and $p = 4$), from which the \\gD{} wireframe is constructed (b) and then lifted to $4\\text{-}D$ (c, d).
+Starting from a default tSNE layout (a) of non-linear clusters data ($n =  2000$ and $p = 4$), from which the \\gD{} wireframe is constructed (a) and then lifted to $4\\text{-}D$ (b, c).
 The fit is reasonbly tight with the data, although it does not fully spread the full width.
@@
 
-nldr_two_curvy2 + 
+#nldr_two_curvy2 + 
   trimesh_two_curvy_removed1_tsne_with_data +
@@
   two_curvy_proj_tsne_all_model1 +
-  plot_layout(ncol = 4)
+  plot_layout(ncol = 3)
 ```
@@
 
+projected_df <- proj_obj1$projected_df
+model_df <- proj_obj1$model_df
+axes <- proj_obj1$axes
+circle <- proj_obj1$circle
+
 ## Projection with Delauany triangulation
@@
 two_curvy_proj_first_model1_umap <- plot_proj(
-    proj_obj = proj_obj1, 
+    projected_df = projected_df, 
+    model_df = model_df, 
+    axes = axes, 
+    circle = circle, 
     point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
@@
     axis_text_size = 4, 
-    is_color = FALSE) +
+    is_category = FALSE) +
     theme(
@@
 
-proj_obj1_temp1 <- proj_obj1
-proj_obj1_temp2 <- proj_obj1
-proj_obj1_temp1[["model_df"]] <- model_df1
-
 two_curvy_proj_model_delaunay_umap <- plot_proj(
-    proj_obj = proj_obj1_temp1, 
+    projected_df = projected_df, 
+    model_df = model_df1, 
+    axes = axes, 
+    circle = circle, 
     point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
@@
 
-proj_obj1_temp2[["model_df"]] <- model_df2
-
 two_curvy_proj_model_benchmark2_umap <- plot_proj(
-    proj_obj = proj_obj1_temp2, 
+    projected_df = projected_df, 
+    model_df = model_df2, 
+    axes = axes, 
+    circle = circle, 
     point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 22
+num_bins_x_gau1 <- 26
 lim1 <- gau1_scaled_obj$lim1
@@
     color = "#000000",
-    #alpha = 0.4,
+    alpha = 0.4,
     linewidth = 0.5) +
@@
     color = "#000000",
-    #alpha = 0.4,
-    linewidth = 0.5) +
+    alpha = 0.4,
+    linewidth = 0.1) +
   geom_segment(
@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 62
+num_bins_x_gau1 <- 90
 lim1 <- gau1_scaled_obj$lim1
@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 25
+num_bins_x_gau1 <- 40
 lim1 <- gau1_scaled_obj$lim1
@@
   filter(distance < benchmark_gau1) |>
-  mutate(ID = row_number()) #|>
-  #filter(!(ID %in% c(97)))
+  mutate(ID = row_number()) |>
+  filter(!(ID %in% c(97)))
 
@@
 error_df_one_curvy_abs <- error_df_one_curvy_abs |>
-  mutate(error_cat_n = if_else(tSNE1 <= 0.25 & tSNE2 <= 0.25, "selected", "deselected")) |> ## high_error points
-  mutate(error_cat_n2 = if_else(tSNE2 >= 1.2, "selected", "deselected")) ## corner points
+  mutate(error_cat_n = if_else(tSNE1 <= 0.25 & tSNE2 <= 0.25, "high", "low")) |> ## high_error points
+  mutate(error_cat_n2 = if_else(tSNE2 >= 1.2, "high", "low")) ## corner points
 
@@
           y = density,
-          colour = factor(error_cat_n, levels = c("deselected", "selected")))) +
+          colour = factor(error_cat_n, levels = c("low", "high")))) +
   geom_point(alpha=0.7) +
@@
             y = density,
-            colour = factor(error_cat_n2, levels = c("deselected", "selected")))) +
+            colour = factor(error_cat_n2, levels = c("low", "high")))) +
   geom_point(alpha=0.7) +
@@
              colour = factor(error_cat_n2,
-                             levels = c("deselected", "selected")))) +
+                             levels = c("low", "high")))) +
   geom_point(alpha=0.7) +
@@
              colour = factor(error_cat_n,
-                             levels = c("deselected", "selected")))) +
+                             levels = c("low", "high")))) +
   geom_point(alpha=0.7) +
@@
   mutate(type = factor(error_df_one_curvy_abs$error_cat_n2,
-                          levels=c("deselected", "selected")))
+                          levels=c("low", "high")))
 
@@
                          group = factor(df_model_data_n$type,
-                                        c("deselected", "selected", "model")),
+                                        c("low", "high", "model")),
                          levelColors = c('#d9d9d9', '#800026', "#000000"))
@@
 
-proj_obj1[["cluster"]] <- factor(error_df_one_curvy_abs$error_cat_n2,
-                          levels=c("deselected", "selected"))
+projected_df <- proj_obj1$projected_df
+model_df <- proj_obj1$model_df
+axes2 <- proj_obj1$axes
+circle <- proj_obj1$circle
+
+## To add error category
+projected_df <- projected_df |>
+  mutate(cluster = factor(error_df_one_curvy_abs$error_cat_n2,
+                          levels=c("low", "high")))
 
 five_c_shaped_proj_tsne_model1 <- plot_proj(
-    proj_obj = proj_obj1,
+    projected_df = projected_df, 
+    model_df = model_df, 
+    axes = axes2, 
+    circle = circle, 
     point_param = c(1.5, 0.5), # size, alpha, color
@@
     axis_text_size = 4, 
-    is_color = TRUE) +
+    is_category = TRUE) +
     scale_color_manual(values = c('#d9d9d9', '#800026')) +
@@
 
-proj_obj1[["cluster"]] <- factor(error_df_one_curvy_abs$error_cat_n,
-                          levels=c("deselected", "selected"))
+projected_df <- projected_df |>
+  select(-cluster) |>
+  mutate(cluster = error_df_one_curvy_abs$error_cat_n)
 
 five_c_shaped_proj_tsne_model1_selected <- plot_proj(
-    proj_obj = proj_obj1, 
+    projected_df = projected_df, 
+    model_df = model_df, 
+    axes = axes2, 
+    circle = circle, 
     point_param = c(1.5, 0.5), # size, alpha, color
@@
     axis_text_size = 4, 
-    is_color = TRUE) +
-    scale_color_manual(values = c('#d9d9d9', '#800026')) +
+    is_category = TRUE) +
+    scale_color_manual(values = c('#800026', '#d9d9d9')) +
     theme(aspect.ratio = 1,
@@
 ## Compute hexbin parameters
-num_bins_x_pbmc <- 22
+num_bins_x_pbmc <- 60
 
@@
   filter(distance < benchmark_pbmc)
-
-umap_pbmc_scaled_with_cluster <- umap_pbmc_scaled |>
-  mutate(cluster = if_else(UMAP1 <= 0.25, "cluster1", if_else(UMAP1 >= 0.625, "cluster2", "cluster3"))) |>
-  mutate(cluster = factor(cluster, levels = c("cluster1", "cluster2", "cluster3")))
 
@@
   geom_point(
-    data = umap_pbmc_scaled_with_cluster,
+    data = umap_pbmc_scaled,
     aes(
       x = UMAP1,
-      y = UMAP2,
-      color = cluster
+      y = UMAP2
     ),
     alpha = 0.2,
-    size = 0.1#,
-    #color = '#a65628'
+    size = 0.1,
+    color = '#a65628'
   )  +
-  scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada")) +
   interior_annotation("a1", cex = 2) +
@@
 
-data_pbmc_n <- data_pbmc |>
-  select(-type) |>
-  mutate(type = as.character(umap_pbmc_scaled_with_cluster$cluster))
-
-df_model_data_pbmc_n <- bind_rows(df_b_pbmc, data_pbmc_n)
+df_model_data_pbmc_n <- bind_rows(df_b_pbmc, data_pbmc)
 
@@
                          group = factor(df_model_data_pbmc_n$type,
-                                        c("cluster1", "cluster2", "cluster3", "model")),
-                         levelColors = c("#8dd3c7", "#fdb462", "#bebada", "#000000"))
+                                        c("data", "model")),
+                         levelColors = c("#a65628", "#000000"))
 ```
@@
 
-proj_obj1[["cluster"]] <- as.character(umap_pbmc_scaled_with_cluster$cluster)
+projected_df <- proj_obj1$projected_df
+model_df <- proj_obj1$model_df
+axes <- proj_obj1$axes
+circle <- proj_obj1$circle
 
 pbmc_proj_umap_model1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(1.5, 0.2), # size, alpha, color
+  projected_df = projected_df, 
+  model_df = model_df, 
+  axes = axes, 
+  circle = circle, 
+  point_param = c(1.5, 0.2, "#a65628"), # size, alpha, color
   plot_limits = c(-0.65, 0.65), 
@@
   axis_text_size = 5,
-  is_color = TRUE) + scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))
+  is_category = FALSE)
 
@@
 
-proj_obj2[["cluster"]] <- as.character(umap_pbmc_scaled_with_cluster$cluster)
+projected_df <- proj_obj2$projected_df
+model_df <- proj_obj2$model_df
+axes <- proj_obj2$axes
+circle <- proj_obj2$circle
 
 pbmc_proj_umap_model2 <- plot_proj(
-    proj_obj = proj_obj2, 
-    point_param = c(1.5, 0.2), # size, alpha, color
+    projected_df = projected_df, 
+    model_df = model_df, 
+    axes = axes, 
+    circle = circle, 
+    point_param = c(1.5, 0.2, "#a65628"), # size, alpha, color
     plot_limits = c(-1, 0.75), 
@@
     axis_text_size = 5,
-    is_color = TRUE) + scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))
+    is_category = FALSE)
 
@@
 error_pbmc_umap2 <- read_rds("data/pbmc3k/error_pbmc_umap_5_min_dist_0.01.rds")
-# error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_12_min_dist_0.99.rds")
-# error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_3_min_dist_0.6.rds")
-# error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_4_min_dist_0.6.rds")
-# error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_5_min_dist_0.8.rds")
-error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_5_min_dist_0.8.rds")
+# error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_15_min_dist_0.99.rds")
+error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_12_min_dist_0.99.rds")
 error_pbmc_tsne <- read_rds("data/pbmc3k/error_pbmc_tsne_5.rds")
@@
   ungroup()
-
-error_pbmc <- error_pbmc |>
-  mutate(method = factor(method,
-                         levels = c("UMAP_5_min_dist_0.8", "UMAP_30_min_dist_0.3", "UMAP_5_min_dist_0.01","tsne_5", "tsne_30", "phate_5", "trimap_12_4_3", "pacmap")))
 ```
@@
   scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 13, 17, 21, 26)]) +
-  scale_color_manual(values=c('#999999','#a65628','#e41a1c','#377eb8','#4daf4a','#ff7f00','#984ea3','#f781bf'))
+  scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a','#ff7f00',
+                              '#984ea3','#999999','#a65628','#f781bf'))
 
@@
 #| label: umap-pbmc-read-best
-umap_pbmc2 <- read_rds("data/pbmc3k/pbmc_umap_5_min_dist_0.8.rds")
+umap_pbmc2 <- read_rds("data/pbmc3k/pbmc_umap_12_min_dist_0.99.rds")
 pbmc_scaled_obj <- gen_scaled_data(
@@
 ## Compute hexbin parameters
-num_bins_x_pbmc <- 13
+num_bins_x_pbmc <- 30
 
@@
 distance_df_small_edges_pbmc <- distance_pbmc |>
-  filter(distance < 0.1) #benchmark_pbmc
-
-tr_from_to_df_pbmc <- right_join(
+  filter(distance < benchmark_pbmc)
+
+tr_from_to_df_pbmc <- inner_join(
   tr_from_to_df_pbmc, distance_df_small_edges_pbmc,
   by = c("from", "to"))
-
-umap_pbmc_scaled_best_with_cluster <- umap_pbmc_scaled_best |>
-  mutate(cluster = umap_pbmc_scaled_with_cluster$cluster)
 
@@
   geom_point(
-    data = umap_pbmc_scaled_best_with_cluster,
+    data = umap_pbmc_scaled_best,
     aes(
-      x = emb1,
-      y = emb2,
-      color = cluster
+      x = UMAP1,
+      y = UMAP2
     ),
-    alpha = 0.3#,
-    #color = "#999999"
+    alpha = 0.3,
+    color = "#999999"
   )  +
-  scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada")) +
   interior_annotation("b1", cex = 2) +
@@
 
-df_model_data_pbmc_n <- bind_rows(df_b_pbmc, data_pbmc_n)
+df_model_data_pbmc_n <- bind_rows(df_b_pbmc, data_pbmc)
 
@@
                          group = factor(df_model_data_pbmc_n$type,
-                                        c("cluster1", "cluster2", "cluster3", "model")),
-                         levelColors = c("#8dd3c7", "#fdb462", "#bebada", "#000000"))
+                                        c("data", "model")),
+                         levelColors = c("#999999", "#000000"))
 ```
@@
 
-proj_obj1[["cluster"]] <- as.character(umap_pbmc_scaled_with_cluster$cluster)
+projected_df <- proj_obj1$projected_df
+model_df <- proj_obj1$model_df
+axes <- proj_obj1$axes
+circle <- proj_obj1$circle
 
 pbmc_proj_umap_model1_best <- plot_proj(
-    proj_obj = proj_obj1,
-    point_param = c(1.5, 0.2), # size, alpha, color
+    projected_df = projected_df, 
+    model_df = model_df, 
+    axes = axes, 
+    circle = circle, 
+    point_param = c(1.5, 0.2, "#999999"), # size, alpha, color
     plot_limits = c(-0.65, 0.65), 
@@
 
-proj_obj2[["cluster"]] <- as.character(umap_pbmc_scaled_with_cluster$cluster)
+projected_df <- proj_obj2$projected_df
+model_df <- proj_obj2$model_df
+axes <- proj_obj2$axes
+circle <- proj_obj2$circle
 
 pbmc_proj_umap_model2_best <- plot_proj(
-    proj_obj = proj_obj2,
-    point_param = c(1.5, 0.2), # size, alpha, color
+    projected_df = projected_df, 
+    model_df = model_df, 
+    axes = axes, 
+    circle = circle, 
+    point_param = c(1.5, 0.2, "#999999"), # size, alpha, color
     plot_limits = c(-1, 0.75), 
@@
     axis_text_size = 5,
-    is_color = TRUE) + scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))
+    is_category = FALSE)
 ```
@@
 The PBMC3k data ($n =  2622$) has three clusters in $9\\text{-}D$, where two of them being close and one separated from the others.
-Two \\gD{} projection from a tour on $9\\text{-}D$ of the model fit with @fig-pbmc-mse a ($a_1 = 0.06$, $b = 616/86 (22, 28)$) shows three-well separated clusters with big separations.
-On the other hand, the model fit with @fig-pbmc-mse b ($a_1 = 0.1$, $b = 234/98 (13, 18)$) shows two close clusters, and one separated cluster.
+Two \\gD{} projection from a tour on $9\\text{-}D$ of the model fit with @fig-pbmc-mse a ($a_1 = 0.02$, $b = 3480/349 (60, 58)$) shows three-well separated clusters with big separations.
+On the other hand, the model fit with @fig-pbmc-mse b ($a_1 = 0.04$, $b = 1020/347 (30, 34)$) shows two close clusters, and one separated cluster.
 The edge connected the two clusters proves the closeness of the two clusters.
@@
 
-proj_obj1[["cluster"]] <- tsne_minst_scaled$cluster
+projected_df <- proj_obj1$projected_df |>
+  mutate(cluster = tsne_minst_scaled$cluster)
+
+model_df <- proj_obj1$model_df
+axes <- proj_obj1$axes
+circle <- proj_obj1$circle
 
 mnist_proj_tsne_model1 <- plot_proj(
-  proj_obj = proj_obj1, 
+  projected_df = projected_df, 
+  model_df = model_df, 
+  axes = axes, 
+  circle = circle, 
   point_param = c(0.5, 0.5), # size, alpha, color
@@
   axis_text_size = 3, 
-  is_color = TRUE) +
+  is_category = TRUE) +
   scale_color_manual(values = c("#999999",'#ff7f00')) +
@@
 mnist_proj_tsne_model1_link <- plot_proj(
-    proj_obj = proj_obj1,
+    projected_df = projected_df, 
+    model_df = model_df, 
+    axes = axes, 
+    circle = circle, 
     point_param = c(1.2, 0.5), # size, alpha, color
@@
     axis_text_size = 3, 
-    is_color = TRUE) +
+    is_category = TRUE) +
     scale_color_manual(values = c("#d9d9d9",'#ff7f00')) +
@@
 
-proj_obj2[["cluster"]] <- tsne_minst_scaled$cluster
+projected_df <- proj_obj2$projected_df |>
+  mutate(cluster = tsne_minst_scaled$cluster)
+
+model_df <- proj_obj2$model_df
+axes <- proj_obj2$axes
+circle <- proj_obj2$circle
 
 mnist_proj_tsne_model2 <- plot_proj(
-    proj_obj = proj_obj1,
+    projected_df = projected_df, 
+    model_df = model_df, 
+    axes = axes, 
+    circle = circle, 
     point_param = c(0.5, 0.3), # size, alpha, color
@@
     axis_text_size = 3, 
-    is_color = TRUE) +
+    is_category = TRUE) +
     scale_color_manual(values = c("#999999",'#ff7f00')) +
@@
 mnist_proj_tsne_model2_link <- plot_proj(
-    proj_obj = proj_obj1, 
+    projected_df = projected_df, 
+    model_df = model_df, 
+    axes = axes, 
+    circle = circle, 
     point_param = c(1.2, 0.5), # size, alpha, color

### 2025-04-06 scripts/additional_functions.R

@@ -181,7 +181,6 @@ get_projection <- function(projection, proj_scale, scaled_data,
 }
 
 # Plot projection
-
 plot_proj <- function(projected_df, model_df, axes, circle,
                       point_param = c(1.5, 0.5, "#000000"), # size, alpha, color
                       line_param = c(0.5, 0.5, "#000000"), #linewidth, alpha

### 2025-04-06 paper.qmd

--- 2025-04-05 paper.qmd
@@
     plot_limits = c(-0.6, 0.6), 
-    title = "c", 
-    cex = 1.5, 
+    title = "b", 
+    cex = 1, 
     position = sc_ltr_pos,
@@
   axis_text_size = 2, 
-  is_color = FALSE) #+
+  is_color = FALSE) +
+  theme(aspect.ratio = 1) #+
   #geom_segment(

### 2025-04-06 paper.qmd
Changed the best layout PBMC

--- 2025-04-05 paper.qmd
@@
 #| label: umap-pbmc-read-best
-umap_pbmc2 <- read_rds("data/pbmc3k/pbmc_umap_5_min_dist_0.8.rds")
+umap_pbmc2 <- read_rds("data/pbmc3k/pbmc_tsne_30.rds")
 pbmc_scaled_obj <- gen_scaled_data(
@@
     aes(
-      x = emb1,
-      y = emb2,
+      x = tSNE1,
+      y = tSNE2,
       color = cluster

### 2025-04-06 paper.qmd
Stash

@@
 ```{r}
-#| label: umap-pbmc-read-best
-umap_pbmc2 <- read_rds("data/pbmc3k/pbmc_tsne_30.rds")
+#| label: tsne-pbmc-read-best
+tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_30.rds")
 pbmc_scaled_obj <- gen_scaled_data(
-  data = umap_pbmc2)
-umap_pbmc_scaled_best <- pbmc_scaled_obj$scaled_nldr
+  data = tsne_pbmc)
+tsne_pbmc_scaled_best <- pbmc_scaled_obj$scaled_nldr
 
@@
 ```{r}
-#| label: num-bins-pbmc
+#| label: tsne-num-bins-pbmc
 
@@
   high_d_data = training_data_pbmc, 
-  nldr_data = umap_pbmc2, 
+  nldr_data = tsne_pbmc, 
   num_x_bins = num_bins_x_pbmc)
 
-umap_pbmc_scaled_best <- algo_obj_pbmc$nldr_scaled
+tsne_pbmc_scaled_best <- algo_obj_pbmc$nldr_scaled
 distance_pbmc <- algo_obj_pbmc$distance_df
@@
 
-umap_pbmc_scaled_best_with_cluster <- umap_pbmc_scaled_best |>
+tsne_pbmc_scaled_best_with_cluster <- tsne_pbmc_scaled_best |>
   mutate(cluster = umap_pbmc_scaled_with_cluster$cluster)
@@
   geom_point(
-    data = umap_pbmc_scaled_best_with_cluster,
+    data = tsne_pbmc_scaled_best_with_cluster,
     aes(
@@
 ```{r}
-#| label: prep-pbmc-umap-model-proj-best
+#| label: prep-pbmc-tsne-model-proj-best
 
@@
 ```{r}
-#| label: langevitour-pbmc-umap-best-proj
+#| label: langevitour-pbmc-tsne-best-proj
 #| eval: false
@@
 ```{r}
-#| label: pbmc-umap-model-proj-best
+#| label: pbmc-tsne-model-proj-best
 
@@
 
-pbmc_proj_umap_model1_best <- plot_proj(
+pbmc_proj_tsne_model1_best <- plot_proj(
   proj_obj = proj_obj1,
@@
 
-pbmc_proj_umap_model2_best <- plot_proj(
+pbmc_proj_tsne_model2_best <- plot_proj(
   proj_obj = proj_obj2,
@@
 free(trimesh_removed_pbmc) + free(trimesh_removed_pbmc_best) +
-  pbmc_proj_umap_model1 + pbmc_proj_umap_model1_best +
-  pbmc_proj_umap_model2 + pbmc_proj_umap_model2_best +
+  pbmc_proj_umap_model1 + pbmc_proj_tsne_model1_best +
+  pbmc_proj_umap_model2 + pbmc_proj_tsne_model2_best +
   plot_layout(nrow=2, byrow=FALSE) &

### 2025-04-06 paper.qmd
Updated Figure 14

@@
 ## Compute hexbin parameters
-num_bins_x_pbmc <- 22
+num_bins_x_pbmc <- 27
 
@@
 distance_df_small_edges_pbmc <- distance_pbmc |>
-  filter(distance < benchmark_pbmc)
+  filter(distance < 0.05)#benchmark_pbmc
 
@@
 ## Compute hexbin parameters
-num_bins_x_pbmc <- 13
+num_bins_x_tsne_pbmc <- 17
 
@@
   nldr_data = tsne_pbmc, 
-  num_x_bins = num_bins_x_pbmc)
+  num_x_bins = num_bins_x_tsne_pbmc)
 
@@
 
-tr_from_to_df_pbmc <- tr_from_to_df_pbmc |>
-  mutate(ID = row_number()) |>
-  filter(!(ID %in% c(19, 20, 23, 24, 29, 33, 48,
-                     183, 181, 178, 176))) #19
+# tr_from_to_df_pbmc <- tr_from_to_df_pbmc |>
+#   mutate(ID = row_number()) |>
+#   filter(!(ID %in% c(19, 20, 23, 24, 29, 33, 48,
+#                      183, 181, 178, 176))) #19
 
@@
   tr_from_to_df_pbmc, distance_df_small_edges_pbmc,
-  by = c("from", "to")) |>
-  select(-ID)
+  by = c("from", "to")) #|>
+  #select(-ID)
 

### 2025-04-06 paper.qmd
Updated Figure 14

@@
 ## Compute hexbin parameters
-num_bins_x_pbmc <- 27
+num_bins_x_pbmc <- 21
 
@@
 distance_df_small_edges_pbmc <- distance_pbmc |>
-  filter(distance < 0.05)#benchmark_pbmc
+  filter(distance < 0.06)#benchmark_pbmc
 
@@
 
-tr_from_to_df_pbmc <- inner_join(
+tr_from_to_df_pbmc <- right_join(
   tr_from_to_df_pbmc, distance_df_small_edges_pbmc, 
@@
 ## Compute hexbin parameters
-num_bins_x_tsne_pbmc <- 17
+num_bins_x_tsne_pbmc <- 22
 
@@
 distance_df_small_edges_pbmc <- distance_pbmc |>
-  filter(distance < 0.1) #benchmark_pbmc
+  filter(distance < benchmark_pbmc) #benchmark_pbmc
 

### 2025-04-06 paper.qmd
Updated Figure 14 with optimal bin widths

@@
 ## Compute hexbin parameters
-num_bins_x_pbmc <- 21
+num_bins_x_pbmc <- 22
 
@@
 distance_df_small_edges_pbmc <- distance_pbmc |>
-  filter(distance < 0.06)#benchmark_pbmc
+  filter(distance < benchmark_pbmc)#benchmark_pbmc
 

### 2025-04-06 paper.qmd
Updated Figure 14 title

@@
 #| echo: false
-#| fig-cap: "Compare the published \\gD{} layout (@fig-pbmc-mse a) and the \\gD{} layout selected (@fig-pbmc-mse b) by MSE plot (@fig-pbmc-mse) from the same NLDR with different hyper-parameter choice.
+#| fig-cap: "Compare the published \\gD{} layout (@fig-pbmc-mse a) and the \\gD{} layout selected (@fig-pbmc-mse e) by MSE plot (@fig-pbmc-mse) from the same NLDR with different hyper-parameter choice.
 The PBMC3k data ($n =  2622$) has three clusters in $9\\text{-}D$, where two of them being close and one separated from the others.
 Two \\gD{} projection from a tour on $9\\text{-}D$ of the model fit with @fig-pbmc-mse a ($a_1 = 0.06$, $b = 616/86 (22, 28)$) shows three-well separated clusters with big separations.
-On the other hand, the model fit with @fig-pbmc-mse b ($a_1 = 0.1$, $b = 234/98 (13, 18)$) shows two close clusters, and one separated cluster.
+On the other hand, the model fit with @fig-pbmc-mse e ($a_1 = 0.06$, $b = 704/227 (22, 32)$) shows two close clusters, and one separated cluster.
 The edge connected the two clusters proves the closeness of the two clusters.

### 2025-04-06 paper.qmd
Updated Figure 14 title

@@
 Two \\gD{} projection from a tour on $9\\text{-}D$ of the model fit with @fig-pbmc-mse a ($a_1 = 0.06$, $b = 616/86 (22, 28)$) shows three-well separated clusters with big separations.
-On the other hand, the model fit with @fig-pbmc-mse e ($a_1 = 0.06$, $b = 704/227 (22, 32)$) shows two close clusters, and one separated cluster.
-The edge connected the two clusters proves the closeness of the two clusters.
-Therefore, @fig-pbmc-mse b is more reasonable than @fig-pbmc-mse a.
+On the other hand, the model fit with @fig-pbmc-mse e ($a_1 = 0.06$, $b = 704/227 (22, 32)$) shows three-well separated clusters with small separations.
+Therefore, @fig-pbmc-mse e is more reasonable than @fig-pbmc-mse a.
 Also, the fitted models in $9\\text{-}D$ help to see some unobserved patterns of the data: (i) dense points, and (ii) non-linear clusters.

### 2025-04-06 paper.qmd
Updated langevitour for Five Gau clusters

@@
 
-df_model_data_n <- bind_rows(data_gau, df_b)
+df_model_data_n <- bind_rows(df_b, data_gau)
 
@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 25
+num_bins_x_gau1 <- 20
 lim1 <- gau1_scaled_obj$lim1
@@
 distance_df_small_edges_gau1 <- distance_gau1 |>
-  filter(distance < benchmark_gau1) |>
-  mutate(ID = row_number()) #|>
+  filter(distance < benchmark_gau1) #|>
+  #mutate(ID = row_number()) #|>
   #filter(!(ID %in% c(97)))

### 2025-04-06 paper.qmd
Updated Figure 11

@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 22
+num_bins_x_gau1 <- 17
 lim1 <- gau1_scaled_obj$lim1
@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 62
+num_bins_x_gau1 <- 52
 lim1 <- gau1_scaled_obj$lim1

### 2025-04-06 paper.qmd
Corrected the projection of Figure 16 and 17

@@
 mnist_proj_tsne_model2 <- plot_proj(
-    proj_obj = proj_obj1,
+    proj_obj = proj_obj2,
     point_param = c(0.5, 0.3), # size, alpha, color
@@
 mnist_proj_tsne_model2_link <- plot_proj(
-    proj_obj = proj_obj1, 
+    proj_obj = proj_obj2, 
     point_param = c(1.2, 0.5), # size, alpha, color

### 2025-04-06 paper.qmd
Tried to work on MNIST another layout

@@
 distance_df_small_edges_mnist <- distance_mnist |>
-  filter(distance < benchmark_mnist)
-
-tr_from_to_df_mnist <- inner_join(
+  filter(distance < benchmark_mnist) #benchmark_mnist
+
+tr_from_to_df_mnist <- right_join(
   tr_from_to_df_mnist, distance_df_small_edges_mnist, 
@@
 
-df_model_data_mnist_n <- bind_rows(data_mnist_cluster, df_b_mnist)
+df_model_data_mnist_n <- bind_rows(df_b_mnist, data_mnist_cluster)
 

### 2025-04-06 paper.qmd
Added another projection

@@
                       cex = 2)
+
+## Second projection
+projection <- cbind(
+  c(0.30943,-0.08452,-0.23005,-0.05284),
+  c(0.22463,-0.06682,0.32106,0.02452))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 3, 
+                            scaled_data = scaled_gau_data, 
+                            scaled_data_model = scaled_gau_data_model, 
+                            distance_df_small_edges = distance_df_small_edges, 
+                            axis_param = list(limits = 1.2,
+                                              axis_scaled = 2,
+                                              axis_pos_x = -1.1, 
+                                              axis_pos_y = -1.1, 
+                                              threshold = 0.05))
+
+five_gau_proj_tsne_model2 <- plot_proj(
+  proj_obj = proj_obj2,
+  point_param = c(1.5, 0.05, clr_choice), # size, alpha, color
+  line_param = c(0.8, 0.4, "#000000"), # linewidth, alpha
+  plot_limits = c(-1.3, 1.3), 
+  title = "a3", 
+  cex = 1.5,
+  position = c(0.08, 0.9),
+  axis_text_size = 4, 
+  is_color = FALSE) +
+  theme(aspect.ratio = 1,
+        legend.position = "none")
 ```
@@
                       cex = 2)
+
+## Second projection
+projection <- cbind(
+  c(0.30943,-0.08452,-0.23005,-0.05284),
+  c(0.22463,-0.06682,0.32106,0.02452))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 3, 
+                            scaled_data = scaled_gau_data, 
+                            scaled_data_model = scaled_gau_data_model, 
+                            distance_df_small_edges = distance_df_small_edges, 
+                            axis_param = list(limits = 1.2,
+                                              axis_scaled = 2,
+                                              axis_pos_x = -1.1, 
+                                              axis_pos_y = -1.1, 
+                                              threshold = 0.05))
+
+five_gau_proj_umap_model2 <- plot_proj(
+  proj_obj = proj_obj2,
+  point_param = c(1.5, 0.05, clr_choice), # size, alpha, color
+  line_param = c(0.8, 0.4, "#000000"), # linewidth, alpha
+  plot_limits = c(-1.3, 1.3), 
+  title = "b3", 
+  cex = 1.5,
+  position = c(0.08, 0.9),
+  axis_text_size = 4, 
+  is_color = FALSE) +
+  theme(aspect.ratio = 1,
+        legend.position = "none")
 ```
@@
 
+## Second projection
+projection <- cbind(
+  c(0.30943,-0.08452,-0.23005,-0.05284),
+  c(0.22463,-0.06682,0.32106,0.02452))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 3, 
+                            scaled_data = scaled_gau_data, 
+                            scaled_data_model = scaled_gau_data_model, 
+                            distance_df_small_edges = distance_df_small_edges, 
+                            axis_param = list(limits = 1.2,
+                                              axis_scaled = 2,
+                                              axis_pos_x = -1.1, 
+                                              axis_pos_y = -1.1, 
+                                              threshold = 0.05))
+
+five_gau_proj_pacmap_model2 <- plot_proj(
+  proj_obj = proj_obj2,
+  point_param = c(1.5, 0.05, clr_choice), # size, alpha, color
+  line_param = c(0.8, 0.4, "#000000"), # linewidth, alpha
+  plot_limits = c(-1.3, 1.3), 
+  title = "c3", 
+  cex = 1.5,
+  position = c(0.08, 0.9),
+  axis_text_size = 4, 
+  is_color = FALSE) +
+  theme(aspect.ratio = 1,
+        legend.position = "none")
+
 ```
@@
 five_gau_proj_tsne_model1 + five_gau_proj_umap_model1 + five_gau_proj_pacmap_model1 + 
-  plot_layout(guides = "collect", nrow = 2) &
+  five_gau_proj_tsne_model2 +
+  five_gau_proj_umap_model2 + five_gau_proj_pacmap_model2 +
+  plot_layout(guides = "collect", nrow = 3) &
   theme(legend.position='none')

### 2025-04-07 paper.qmd
Updated the Five Gaussian cluster example

@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 17
+num_bins_x_gau1 <- 14
 lim1 <- gau1_scaled_obj$lim1
@@
 distance_df_small_edges_gau1 <- distance_gau1 |>
-  filter(distance < benchmark_gau1)
-
-tr_from_to_df_gau1 <- inner_join(
+  filter(distance < 0.1) #benchmark_gau1
+
+distance_df_small_edges_gau1 <- distance_df_small_edges_gau1 |>
+  mutate(ID = row_number()) |>
+  dplyr::filter(!(ID %in% c(150, 110))) |>
+  dplyr::select(-ID)
+
+tr_from_to_df_gau1 <- right_join(
   tr_from_to_df_gau1, distance_df_small_edges_gau1, 
@@
 distance_df_small_edges <- distance_gau1 |>
-  dplyr::filter(distance < benchmark_gau1)
+  dplyr::filter(distance < 0.1) |>
+  mutate(ID = row_number()) |>
+  filter(!(ID %in% c(150, 110))) |>
+  dplyr::select(-ID) #benchmark_gau1
 ```
@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 52
+num_bins_x_gau1 <- 50
 lim1 <- gau1_scaled_obj$lim1
@@
 distance_df_small_edges_gau1 <- distance_gau1 |>
-  filter(distance < benchmark_gau1) |>
-  mutate(ID = row_number()) #|>
+  filter(distance < 0.05) #|> benchmark_gau1
+  #mutate(ID = row_number()) #|>
   #filter(!(ID %in% c(112)))
@@
 distance_df_small_edges <- distance_gau1 |>
-  dplyr::filter(distance < benchmark_gau1)
+  dplyr::filter(distance < 0.05) #benchmark_gau1
 
@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 20
+num_bins_x_gau1 <- 22
 lim1 <- gau1_scaled_obj$lim1
@@
 distance_df_small_edges_gau1 <- distance_gau1 |>
-  filter(distance < benchmark_gau1) #|>
+  filter(distance < 0.06) #|> #benchmark_gau1
   #mutate(ID = row_number()) #|>
@@
 
-tr_from_to_df_gau1 <- inner_join(
+tr_from_to_df_gau1 <- right_join(
   tr_from_to_df_gau1, distance_df_small_edges_gau1, 
@@
 distance_df_small_edges <- distance_gau1 |>
-  dplyr::filter(distance < benchmark_gau1)
+  dplyr::filter(distance < 0.06) #benchmark_gau1 #benchmark_gau1
 ```

### 2025-04-10 paper.qmd
Saved models for Gau

@@
   dplyr::select(-ID) #benchmark_gau1
+
+# write_rds(distance_df_small_edges, "data/five_gau_clusters/tsne_model.rds")
 ```
@@
   select(-type)
+
+# write_rds(distance_df_small_edges, "data/five_gau_clusters/umap_model.rds")
 ```
@@
   dplyr::filter(distance < 0.06) #benchmark_gau1 #benchmark_gau1
+
+# write_rds(distance_df_small_edges, "data/five_gau_clusters/pacmap_model.rds")
 ```

### 2025-04-10 paper.qmd
Found the way to projected the models
Saved data for other methods as well

@@
 
-# write_rds(distance_df_small_edges, "data/five_gau_clusters/tsne_model.rds")
+# write_rds(distance_df_small_edges, "data/five_gau_clusters/tsne_wireframe.rds")
+# write_rds(df_b, "data/five_gau_clusters/tsne_model.rds")
 ```
@@
 
-# write_rds(distance_df_small_edges, "data/five_gau_clusters/umap_model.rds")
+# write_rds(distance_df_small_edges, "data/five_gau_clusters/umap_wireframe.rds")
+# write_rds(df_b, "data/five_gau_clusters/umap_model.rds")
 ```
@@
 
-# write_rds(distance_df_small_edges, "data/five_gau_clusters/pacmap_model.rds")
+# write_rds(distance_df_small_edges, "data/five_gau_clusters/pacmap_wireframe.rds")
+# write_rds(df_b, "data/five_gau_clusters/pacmap_model.rds")
 ```

### 2025-04-10 paper.qmd
Saved hexID mappings

@@
 ```{r}
+
+## Give unique values to hex IDs
+hexid_unique <- tsne_data_with_hb_id |> 
+  arrange(hb_id) |> 
+  distinct(hb_id) |> 
+  mutate(ID = row_number())
+
+## Find points in each hex bin
+pts_df <- tsne_data_with_hb_id |>
+    group_by(hb_id) |>
+    summarize(pts_list = list(ID), .groups = "drop")
+
+unlisted_df <- pts_df |>
+  unnest(cols = pts_list) |>
+  rename(pts_ID = pts_list)
+
+pts_df <- left_join(pts_df, unlisted_df, by = c("hb_id")) |>
+  select(-pts_list)
+
+write_rds(pts_df, "data/five_gau_clusters/tsne_model_mapping_data.rds")
+```
+
+```{r}
 #| label: langevitour-five-gau-tsne-proj
@@
 ```{r}
+
+## Give unique values to hex IDs
+hexid_unique <- umap_data_with_hb_id |> 
+  arrange(hb_id) |> 
+  distinct(hb_id) |> 
+  mutate(ID = row_number())
+
+## Find points in each hex bin
+pts_df <- umap_data_with_hb_id |>
+    group_by(hb_id) |>
+    summarize(pts_list = list(ID), .groups = "drop")
+
+unlisted_df <- pts_df |>
+  unnest(cols = pts_list) |>
+  rename(pts_ID = pts_list)
+
+pts_df <- left_join(pts_df, unlisted_df, by = c("hb_id")) |>
+  select(-pts_list)
+
+write_rds(pts_df, "data/five_gau_clusters/umap_model_mapping_data.rds")
+```
+
+```{r}
 #| label: langevitour-five-gau-umap-proj
@@
 ```{r}
+
+## Give unique values to hex IDs
+hexid_unique <- pacmap_data_with_hb_id |> 
+  arrange(hb_id) |> 
+  distinct(hb_id) |> 
+  mutate(ID = row_number())
+
+## Find points in each hex bin
+pts_df <- pacmap_data_with_hb_id |>
+    group_by(hb_id) |>
+    summarize(pts_list = list(ID), .groups = "drop")
+
+unlisted_df <- pts_df |>
+  unnest(cols = pts_list) |>
+  rename(pts_ID = pts_list)
+
+pts_df <- left_join(pts_df, unlisted_df, by = c("hb_id")) |>
+  select(-pts_list)
+
+write_rds(pts_df, "data/five_gau_clusters/pacmap_model_mapping_data.rds")
+```
+
+```{r}
 #| label: langevitour-five-gau-pacmap-proj

### 2025-04-15 paper.qmd
Saved updated mapping data
Removed unncessary chunks
Added the models only

--- 2025-04-10 paper.qmd
@@
 
+# write_rds(df_b, "data/five_gau_clusters/pacmap_model.rds")
+
 ## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
@@
 
-# write_rds(distance_df_small_edges, "data/five_gau_clusters/pacmap_wireframe.rds")
+write_rds(distance_df_small_edges, "data/five_gau_clusters/pacmap_wireframe.rds")
 # write_rds(df_b, "data/five_gau_clusters/pacmap_model.rds")
+```
+
+```{r}
+
+ ## Give unique values to hex IDs
+ hexid_unique <- pacmap_data_with_hb_id |> 
+   arrange(hb_id) |> 
+   distinct(hb_id) |> 
+   mutate(ID = row_number())
+
+ ## Find points in each hex bin
+ pts_df <- pacmap_data_with_hb_id |>
+     group_by(hb_id) |>
+     summarize(pts_list = list(ID), .groups = "drop")
+
+ unlisted_df <- pts_df |>
+   unnest(cols = pts_list) |>
+   rename(pts_ID = pts_list)
+
+ pts_df <- left_join(hexid_unique, unlisted_df, by = c("hb_id"))
+
+ write_rds(pts_df, "data/five_gau_clusters/pacmap_model_mapping_data.rds")
 ```

### 2025-04-15 paper.qmd
Removed unnecessary code
Added install packages if it's already not installed

--- 2025-04-10 paper.qmd
@@
 
+```{r include=FALSE}
+#| label: install-libraries
+#| warning: false
+#| echo: false
+
+if (!requireNamespace("remotes", quietly = TRUE)) {
+  install.packages("remotes")
+}
+
+if (!requireNamespace("quollr", quietly = TRUE)) {
+  remotes::install_github("JayaniLakshika/quollr")
+}
+
+if (!requireNamespace("tibble", quietly = TRUE)) {
+  install.packages("tibble")
+}
+
+if (!requireNamespace("dplyr", quietly = TRUE)) {
+  install.packages("dplyr")
+}
+
+if (!requireNamespace("ggplot2", quietly = TRUE)) {
+  install.packages("ggplot2")
+}
+
+if (!requireNamespace("readr", quietly = TRUE)) {
+  install.packages("readr")
+}
+
+if (!requireNamespace("tidyr", quietly = TRUE)) {
+  install.packages("tidyr")
+}
+
+if (!requireNamespace("patchwork", quietly = TRUE)) {
+  install.packages("patchwork")
+}
+
+if (!requireNamespace("colorspace", quietly = TRUE)) {
+  install.packages("colorspace")
+}
+
+if (!requireNamespace("kableExtra", quietly = TRUE)) {
+  install.packages("kableExtra")
+}
+
+if (!requireNamespace("conflicted", quietly = TRUE)) {
+  install.packages("conflicted")
+}
+```
+
 ```{r}
@@
 library(patchwork)
-library(png)
 library(colorspace)

### 2025-04-15 paper.qmd
Added pkg dowloads with latest version

@@
 
-if (!requireNamespace("remotes", quietly = TRUE)) {
-  install.packages("remotes")
+options(repos = c(CRAN = "https://cran.rstudio.com")) # Setup mirror
+
+packages_to_check <- c("remotes", "tibble", "dplyr", "ggplot2", "tidyr", "readr", "patchwork", "colorspace", "kableExtra", "conflicted")
+
+for (pkg in packages_to_check) {
+  if (!requireNamespace(pkg, quietly = TRUE)) {
+    message(paste("Installing package:", pkg))
+    install.packages(pkg)
+  } else {
+    installed_version <- packageVersion(pkg)
+    available_version <- tryCatch({
+      utils::packageDescription(pkg)$Version
+    }, error = function(e) NA) # Handle cases where package info isn't readily available
+
+    if (!is.na(available_version) && installed_version < package_version(available_version)) {
+      message(paste("A newer version of package", pkg, "is available. Updating..."))
+      install.packages(pkg)
+    } else {
+      message(paste("Package", pkg, "is up-to-date (version", installed_version, ")."))
+    }
+  }
 }
@@
 
-if (!requireNamespace("tibble", quietly = TRUE)) {
-  install.packages("tibble")
-}
-
-if (!requireNamespace("dplyr", quietly = TRUE)) {
-  install.packages("dplyr")
-}
-
-if (!requireNamespace("ggplot2", quietly = TRUE)) {
-  install.packages("ggplot2")
-}
-
-if (!requireNamespace("readr", quietly = TRUE)) {
-  install.packages("readr")
-}
-
-if (!requireNamespace("tidyr", quietly = TRUE)) {
-  install.packages("tidyr")
-}
-
-if (!requireNamespace("patchwork", quietly = TRUE)) {
-  install.packages("patchwork")
-}
-
-if (!requireNamespace("colorspace", quietly = TRUE)) {
-  install.packages("colorspace")
-}
-
-if (!requireNamespace("kableExtra", quietly = TRUE)) {
-  install.packages("kableExtra")
-}
-
-if (!requireNamespace("conflicted", quietly = TRUE)) {
-  install.packages("conflicted")
-}
 ```

### 2025-04-15 paper.qmd
Updated the package installation by replacing with tidyverse
Try to remove the packages thatdon't want

@@
 
-packages_to_check <- c("remotes", "tibble", "dplyr", "ggplot2", "tidyr", "readr", "patchwork", "colorspace", "kableExtra", "conflicted")
+packages_to_check <- c("remotes", "tidyverse", "patchwork", "colorspace", "kableExtra", "conflicted")
 
@@
 
-These `R` packages were used for the work: `tidyverse` (@hadley2019), `png` (@simon2022), `Rtsne` (@jesse2015), `umap` (@tomasz2023), `ggplot2` (@hadley2016), `patchwork` (@thomas2024), `colorspace` (@achim2020), `langevitour` (@harisson2024), `conflicted` (@hadley2023), `reticulate` (@kevin2024), `kableExtra` (@hao2024).
+These `R` packages were used for the work: `tidyverse` (@hadley2019), `Rtsne` (@jesse2015), `umap` (@tomasz2023), `patchwork` (@thomas2024), `colorspace` (@achim2020), `langevitour` (@harisson2024), `conflicted` (@hadley2023), `reticulate` (@kevin2024), `kableExtra` (@hao2024).
 These `python` packages were used for the work: `trimap` (@amid2022) and `pacmap` (@yingfan2021).

### 2025-04-17 paper.qmd
Updated Figure1

@@
   geom_point(alpha=0.1, size=1, colour=clr_choice) +
-  interior_annotation("b")
+  interior_annotation("b", c(0.08, 0.9))
 
@@
   geom_point(alpha=0.1, size=1, colour='#999999') +
-  interior_annotation("b")
+  interior_annotation("b", c(0.08, 0.9))
 

### 2025-04-17 paper.qmd
Updated Figure 5 titles

@@
                colour = "#000000") +
-  interior_annotation("a", sc_ltr_pos, cex = 1.5) +
+  interior_annotation("a", sc_ltr_pos, cex = 1.2) +
   theme(aspect.ratio = 1)

### 2025-04-17 paper.qmd
Updated Figure 11

@@
                       position = c(0.95, 0.95),
-                      cex = 2) +
+                      cex = 1.5) +
   theme(

### 2025-04-17 paper.qmd
Updated Figure 12

@@
                       position = c(0.95, 0.9),
-                      cex = 1.5) +
+                      cex = 2) +
   theme_bw() +
@@
     title = "a3", 
-    cex = 1.5,
+    cex = 2,
     position = c(0.08, 0.9),

### 2025-04-17 paper.qmd
Updated Figure 12

@@
     plot_limits = c(-0.35, 0.35), 
-    title = "a3", 
+    title = "b3", 
     cex = 2,

### 2025-04-17 paper.qmd
Updated Figure 16

@@
   scale_color_manual(values=c("#999999",'#ff7f00')) +
-  interior_annotation("a", c(0.08, 0.93), cex = 1.5) +
+  interior_annotation("a", c(0.08, 0.93), cex = 2) +
   theme(legend.position = "none")
@@
   title = "b", 
-  cex = 1.5, 
+  cex = 2, 
   axis_text_size = 3, 
@@
     title = "c", 
-    cex = 1.5, 
+    cex = 2, 
     axis_text_size = 3, 

### 2025-04-17 paper.qmd
Updated Figure 17

@@
   scale_color_manual(values=c('#d9d9d9', '#ff7f00')) +
-  interior_annotation("a", c(0.08, 0.93), cex = 2) +
+  interior_annotation("a", c(0.08, 0.93), cex = 2.5) +
   theme(legend.position = "none")
@@
     title = "b", 
-    cex = 2, 
+    cex = 2.5, 
     axis_text_size = 3, 
@@
     title = "c", 
-    cex = 2, 
+    cex = 2.5, 
     axis_text_size = 3, 

### 2025-04-17 paper.qmd
Updated Figure 1

@@
   geom_point(alpha=0.1, size=1, colour=clr_choice) +
-  interior_annotation("a")
+  interior_annotation("a", cex = 1.3)
 
@@
   geom_point(alpha=0.1, size=1, colour=clr_choice) +
-  interior_annotation("b", c(0.08, 0.9))
+  interior_annotation("b", c(0.08, 0.9), cex = 1.3)
 
@@
   geom_point(alpha=0.1, size=1, colour=clr_choice) +
-  interior_annotation("c")
+  interior_annotation("c", cex = 1.3)
 
@@
   geom_point(alpha=0.1, size=1, colour=clr_choice) +
-  interior_annotation("d")
+  interior_annotation("d", cex = 1.3)
 
@@
   geom_point(alpha=0.1, size=1, colour=clr_choice) +
-  interior_annotation("e")
+  interior_annotation("e", cex = 1.3)
 
@@
   geom_point(alpha=0.1, size=1, colour=clr_choice) +
-  interior_annotation("f")
+  interior_annotation("f", cex = 1.3)
 
@@
   geom_point(alpha=0.1, size=1, colour=clr_choice) +
-  interior_annotation("g")
+  interior_annotation("g", cex = 1.3)
 
@@
   geom_point(alpha=0.1, size=1, colour=clr_choice) +
-  interior_annotation("h")
+  interior_annotation("h", cex = 1.3)
 

### 2025-04-17 paper.qmd
Updated Figure 2 axes

@@
                             distance_df_small_edges = distance_df_small_edges_two_curvy, 
-                            axis_param = list(limits = 0.5, 
+                            axis_param = list(limits = 0.35, 
                                               axis_scaled = 2,
-                                              axis_pos_x = -0.3, 
-                                              axis_pos_y = -0.3, 
+                                              axis_pos_x = -0.35, 
+                                              axis_pos_y = -0.35, 
                                               threshold = 0.01))

### 2025-04-17 paper.qmd
Updated Figure 11

@@
     colour="grey50",
-    size = 5) +
+    size = 3) +
   geom_path(

### 2025-04-17 paper.qmd
Updated Figure12

@@
                             distance_df_small_edges = distance_df_small_edges_c_shaped_structure,
-                            axis_param = list(limits = 0.35,
-                                              axis_scaled = 2,
+                            axis_param = list(limits = 0.4,
+                                              axis_scaled = 3,
                                               axis_pos_x = -0.28,
                                               axis_pos_y = -0.28,
-                                              threshold = 0.017))
+                                              threshold = 0.02))
 
@@
     position = c(0.08, 0.9),
-    axis_text_size = 4, 
+    axis_text_size = 4.5, 
     is_color = TRUE) +

### 2025-04-17 paper.qmd
Updated Figure 16 and 17 axes

@@
                                  axis_scaled = 1, 
-                                 axis_pos_x = -0.9, 
-                                 axis_pos_y = -0.9, 
+                                 axis_pos_x = -0.92, 
+                                 axis_pos_y = -0.92, 
                                  threshold = 0.06))
@@
   cex = 2, 
-  axis_text_size = 3, 
+  axis_text_size = 3.5, 
   is_color = TRUE) +
@@
     cex = 2.5, 
-    axis_text_size = 3, 
+    axis_text_size = 4.5, 
     is_color = TRUE) +
@@
     cex = 2, 
-    axis_text_size = 3, 
+    axis_text_size = 3.5, 
     is_color = TRUE) +

### 2025-04-17 paper.qmd
Updated Figure14

@@
   cex = 2, 
-  axis_text_size = 5,
+  axis_text_size = 4,
   is_color = TRUE) + scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))
@@
     cex = 2, 
-    axis_text_size = 5,
+    axis_text_size = 4,
     is_color = TRUE) + scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))
@@
   scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada")) +
-  interior_annotation("b1", cex = 2) +
+  interior_annotation("b1", cex = 2, c(0.08, 0.9)) +
   theme(

### 2025-04-17 paper.qmd
Initial update for .bib

@@
 Non-linear dimension reduction (NLDR) is popular for making a convenient low-dimensional (\kD{}) representation of high-dimensional (\pD{}) data ($k < p$).
-Recently developed methods include t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2022], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021].
+Recently developed methods include t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021].
 However, the representation generated can vary dramatically from method to method, and with different choices of parameters or random seeds made using the same method (@fig-NLDR-variety).
@@
 These `R` packages were used for the work: `tidyverse` (@hadley2019), `Rtsne` (@jesse2015), `umap` (@tomasz2023), `patchwork` (@thomas2024), `colorspace` (@achim2020), `langevitour` (@harisson2024), `conflicted` (@hadley2023), `reticulate` (@kevin2024), `kableExtra` (@hao2024).
-These `python` packages were used for the work: `trimap` (@amid2022) and `pacmap` (@yingfan2021).
+These `python` packages were used for the work: `trimap` (@amid2019) and `pacmap` (@yingfan2021).
 The article was created with `R` packages `rticles` (@jjallaire2024), `knitr` (@yihui2014), and `rmarkdown` (@yihui2020).

### 2025-04-17 paper.qmd
Added quarto citation
Figure 11 resized axes cricle..

@@
                             distance_df_small_edges = distance_df_small_edges, 
-                            axis_param = list(limits = 1.2,
+                            axis_param = list(limits = 1.5,
                                               axis_scaled = 2,
-                                              axis_pos_x = -1.1, 
-                                              axis_pos_y = -1.1, 
+                                              axis_pos_x = -1, 
+                                              axis_pos_y = -1, 
                                               threshold = 0.05))
@@
 These `python` packages were used for the work: `trimap` (@amid2019) and `pacmap` (@yingfan2021).
-The article was created with `R` packages `rticles` (@jjallaire2024), `knitr` (@yihui2014), and `rmarkdown` (@yihui2020).
+The article was created with `R` packages `quarto` (@jjallaire2024).
 The project's GitHub repository ([https://github.com/JayaniLakshika/paper-nldr-vis-algorithm](https://github.com/JayaniLakshika/paper-nldr-vis-algorithm)) contains all materials required to reproduce this article.

### 2025-04-20 appendix.qmd
Updated Figure 1

@@
 
-```{r}
+```{r include=FALSE}
+#| label: install-libraries
+#| warning: false
+#| echo: false
+
+options(repos = c(CRAN = "https://cran.rstudio.com")) # Setup mirror
+
+packages_to_check <- c("remotes", "tidyverse", "patchwork", "colorspace", "kableExtra", "conflicted", "ggforce")
+
+for (pkg in packages_to_check) {
+  if (!requireNamespace(pkg, quietly = TRUE)) {
+    message(paste("Installing package:", pkg))
+    install.packages(pkg)
+  } else {
+    installed_version <- packageVersion(pkg)
+    available_version <- tryCatch({
+      utils::packageDescription(pkg)$Version
+    }, error = function(e) NA) # Handle cases where package info isn't readily available
+
+    if (!is.na(available_version) && installed_version < package_version(available_version)) {
+      message(paste("A newer version of package", pkg, "is available. Updating..."))
+      install.packages(pkg)
+    } else {
+      message(paste("Package", pkg, "is up-to-date (version", installed_version, ")."))
+    }
+  }
+}
+
+if (!requireNamespace("quollr", quietly = TRUE)) {
+  remotes::install_github("JayaniLakshika/quollr")
+}
+
+```
+
+```{r}
+#| label: load-libraries
+#| warning: false
+#| echo: false
+library(quollr)
+library(tibble)
+library(dplyr)
 library(ggplot2)
-library(dplyr)
-library(quollr)
 library(readr)
+library(tidyr)
+library(patchwork)
+library(colorspace)
+library(kableExtra)
+library(conflicted)
 library(ggforce)
-library(patchwork)
-library(kableExtra)
-library(tidyr)
-
+
+conflicts_prefer(dplyr::filter)
+conflicts_prefer(dplyr::select)
+```
+
+```{r}
+#| label: plot-theme
 theme_set(theme_linedraw() +
@@
 )
-
-interior_annotation <- function(label, position = c(0.92, 0.92), cex = 1, col="grey70") {
-  annotation_custom(grid::textGrob(label = label,
-      x = unit(position[1], "npc"), y = unit(position[2], "npc"),
-      gp = grid::gpar(cex = cex, col=col)))
-}
-```
-
-```{r}
-#| label: scale-highd-data
-# Center the data by subtracting the mean of each column
-center_data <- function(data) {
-  apply(data, 2, function(col) col - mean(col))
-}
-
-# Function to scale data manually
-scale_data_manual <- function(data, type_col) {
-  # Step 1: Center the data (mean 0)
-  data_centered <- center_data(data |> select(-all_of(type_col)))
-
-  # Step 2: Calculate the standard deviation of each dimension
-  sds <- apply(data_centered, 2, sd)
-
-  # Step 3: Scale each dimension to have the range [0, 1]
-  data_scaled <- apply(data_centered, 2, function(col) col / max(abs(col)))
-
-  # Step 4: Scale dimensions according to their variation
-  # The dimension with the highest standard deviation is scaled to [-1, 1]
-  # Other dimensions are scaled to smaller ranges based on their standard deviations
-  max_sd <- max(sds)
-
-  # Normalize the standard deviations to get scaling factors
-  scaling_factors <- sds / max_sd
-
-  for (i in seq_along(scaling_factors)) {
-    data_scaled[, i] <- data_scaled[, i] * scaling_factors[i]
-  }
-
-  # Combine the scaled data with the 'type' column and return as a tibble
-  data_scaled <- as_tibble(data_scaled) %>% 
-    mutate(!!type_col := data[[type_col]])
-
-  return(data_scaled)
-}
-```
-
-```{r}
-#| label: gen-axes
-
-gen_axes <- function(proj, limits = 1, axis_pos_x = NULL, axis_pos_y = NULL, axis_labels, threshold) {
-
-  axis_scale <- limits/6
-
-  if (is.null(axis_pos_x)) {
-
-    axis_pos_x <- -2/3 * limits
-
-  }
-
-  if (is.null(axis_pos_y)) {
-
-    axis_pos_y <- -2/3 * limits
-
-  }
-
-  adj <- function(x, axis_pos) axis_pos + x * axis_scale
-  axes <- data.frame(x1 = adj(0, axis_pos_x), 
-                     y1 = adj(0, axis_pos_y), 
-                     x2 = adj(proj[, 1], axis_pos_x), 
-                     y2 = adj(proj[, 2], axis_pos_y))
-
-  rownames(axes) <- axis_labels
-
-  ## To remove axes
-  axes <- axes |>
-    mutate(distance = sqrt((x2 - x1)^2 + (y2 - y1)^2)) |>
-    filter(distance >= threshold)
-
-  theta <- seq(0, 2 * pi, length = 50)
-  circle <- data.frame(c1 = adj(cos(theta), axis_pos_x), 
-                       c2 = adj(sin(theta), axis_pos_y))
-
-  return(list(axes = axes, circle = circle))
-
-}
-```
-
-```{r}
-#| label: solve-quad-fun
-
-quad <- function(a = 3, b = 2 * a2, c = -(a2^2 + a1^2))
-{
-  a <- as.complex(a)
-  answer <- c((-b + sqrt(b^2 - 4 * a * c)) / (2 * a),
-              (-b - sqrt(b^2 - 4 * a * c)) / (2 * a))
-  if(all(Im(answer) == 0)) answer <- Re(answer)
-  if(answer[1] == answer[2]) return(answer[1])
-  answer[answer>0] ## only positive 
-}
-
+```
+
+```{r}
+#| label: import-scripts
+source("scripts/additional_functions.R")
+```
+
+```{r}
+#| label: code-setup
+set.seed(20240110)
 ```
@@
       label = expression(group("(", list(s[1], s[2]), ")")),
-      color = "black") 
+      color = "black") +
+  coord_equal()  
 ```
@@
 #| fig-cap: "The components of the hexagon grid illustrating notation."
-#| out-height: 30%
-#| fig-width: 2.5
-#| fig-height: 3.3
+#| out-width: 30%
 #| fig-pos: H
@@
 #| label: read-pbmc-umap-layouts
+#| eval: false
 # Read a variety of different NLDR representations of PBMC
@@
 #| label: combine-data-pbmc-umap
+#| eval: false
 
@@
 #| label: error-comp-pbmc-umap
+#| eval: false
 
@@
 #| label: fig-pbmc-mse-umap
+#| eval: false
 #| fig-pos: H
@@
 #| label: published-pbmc-scdeed
+#| eval: false
 
@@
 #| label: hexbin-pbmc-scdeed
+#| eval: false
 ## Compute hexbin parameters
@@
 #| label: pbmc-umap-model-proj-scdeed
+#| eval: false
 

### 2025-04-20 appendix.tex
Updated Figure 1

@@
 
-\includegraphics[width=1\textwidth,height=0.3\textheight]{appendix_files/figure-pdf/fig-hex-param-1.pdf}
+\includegraphics[width=0.3\linewidth,height=\textheight,keepaspectratio]{appendix_files/figure-pdf/fig-hex-param-1.pdf}
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{appendix_files/figure-pdf/fig-assign-data-1.pdf}
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{appendix_files/figure-pdf/fig-assign-data-1.pdf}
 
@@
 
-\includegraphics[width=1\textwidth,height=0.3\textheight]{appendix_files/figure-pdf/fig-tri-param-1.pdf}
+\includegraphics[width=1\linewidth,height=0.3\textheight]{appendix_files/figure-pdf/fig-tri-param-1.pdf}
 
@@
 
-Figure~\ref{fig-pbmc-mse-umap} shows layout from the original published
+\textbf{?@fig-pbmc-mse-umap} shows layout from the original published
 article along with that suggested by scDEED to be more accurate. The log
@@
 
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=0.53\textheight]{appendix_files/figure-pdf/fig-pbmc-mse-umap-1.pdf}
-
-}
-
-\caption{\label{fig-pbmc-mse-umap}Assessing which of the 2 UMAP layouts
-with different hyper-parameter settings (n\_neighbors: \(30\),
-min\_dist: \(0.3\) (red); n\_neighbors: \(80\), min\_dist: \(0.5\)
-(blue)) on the PBMC data is the better representation using MSE for
-varying binwidth (\(a_1\)). Colour used for the lines and points in the
-top plot and in the scatterplots represents UMAP layout (a, b). Of the
-two, layout b is optimal across all binwidths making it the best
-choice.}
-
-\end{figure}%
-
   \bibliography{bibliography.bib}

### 2025-04-20 appendix.qmd
Updated Figure 3

@@
            label = expression(frac(a[2] - a[1], 2)), 
-           color = "black", hjust=-0.2) + 
+           color = "black", hjust=-0.2, size = 5) + 
   annotate("text", x=1.5, y= -0.3, 
            label = expression(frac(a[1], 2)), 
-           color = "black", hjust=-0.2) + 
+           color = "black", hjust=-0.2, size = 5) + 
   annotate("text", x=2.5, y= 1.5, 
            label = expression(l), 
-           color = "black", hjust=-0.2)
+           color = "black", hjust=-0.2, size = 5)
 ```
@@
 #| fig-cap: "The components of the right triangle illustrating notation."
-#| out-height: 30%
+#| out-width: 30%
 

### 2025-04-20 appendix.qmd
Added the corrected .bib

@@
 tbl-cap-location: bottom
-bibliography: bibliography.bib 
+bibliography: paper.bib 
 header-includes: | 

### 2025-04-20 appendix.qmd
Updated Figure 4

@@
 ```
+
+\newpage
 
@@
 ```{r}
+#| label: fig-n_neighbors-dubious_cells
 #| out-width: 50%
+#| fig-pos: H
 #| eval: false
@@
   filter(min.dist == 0.5) |>
+  filter(n_neighbors < 240) |>
   ggplot(aes(
@@
     labs(x = "n_neighbors",
-         y = "# of dubious cells") +
+         y = "Number of dubious cells") +
     theme_minimal()
@@
 #| label: read-pbmc-umap-layouts
-#| eval: false
 # Read a variety of different NLDR representations of PBMC
@@
 #| label: combine-data-pbmc-umap
-#| eval: false
 
@@
 #| label: error-comp-pbmc-umap
-#| eval: false
 
@@
 #| label: fig-pbmc-mse-umap
-#| eval: false
 #| fig-pos: H
-#| out-height: 53%
+#| out-width: 100%
 #| fig-cap: "Assessing which of the 2 UMAP layouts with different hyper-parameter settings (n_neighbors: $30$, min_dist: $0.3$ (red); n_neighbors: $80$, min_dist: $0.5$ (blue)) on the PBMC data is the better representation using MSE for varying binwidth ($a_1$).

### 2025-04-20 appendix.tex
Updated Figure 3
Added the corrected .bib
Updated Figure 4

@@
 
+\newpage
+
 \section{Computing hexagon grid
@@
 
-\includegraphics[width=1\linewidth,height=0.3\textheight]{appendix_files/figure-pdf/fig-tri-param-1.pdf}
+\includegraphics[width=0.3\linewidth,height=\textheight,keepaspectratio]{appendix_files/figure-pdf/fig-tri-param-1.pdf}
 
@@
 
-\textbf{?@fig-pbmc-mse-umap} shows layout from the original published
+Figure~\ref{fig-pbmc-mse-umap} shows layout from the original published
 article along with that suggested by scDEED to be more accurate. The log
@@
 
-  \bibliography{bibliography.bib}
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{appendix_files/figure-pdf/fig-pbmc-mse-umap-1.pdf}
+
+}
+
+\caption{\label{fig-pbmc-mse-umap}Assessing which of the 2 UMAP layouts
+with different hyper-parameter settings (n\_neighbors: \(30\),
+min\_dist: \(0.3\) (red); n\_neighbors: \(80\), min\_dist: \(0.5\)
+(blue)) on the PBMC data is the better representation using MSE for
+varying binwidth (\(a_1\)). Colour used for the lines and points in the
+top plot and in the scatterplots represents UMAP layout (a, b). Of the
+two, layout b is optimal across all binwidths making it the best
+choice.}
+
+\end{figure}%
+
+  \bibliography{paper.bib}
 

### 2025-04-20 appendix.qmd
More described the approach

@@
 
-@fig-pbmc-mse-umap shows layout from the original published article along with that suggested by scDEED to be more accurate.
-The log MSE vs binwidth plot illustrates that our approach would suggest that scDEED is correct here, that layout b is more accurately reflecting the cluster structure in the data.
+The layout a (@fig-pbmc-mse-umap) is generated from the hyper-parameters suggested by @chen2023, and the layout b (@fig-pbmc-mse-umap) is with suggested hyper-parameters by scDEED to be more accurate.
+The layout a and b contain $46$ and $15$ dubious cells respectively.
+
+The log MSE vs binwidth ($a_1$) plot (@fig-pbmc-mse-umap) illustrates that our approach would suggest that scDEED is correct here, that layout b is more accurately reflecting the cluster structure in the PBMC data.
 
@@
   geom_line(linewidth = 0.3) +
-  # geom_vline(xintercept = 15, linetype="solid",
-  #            color = "black", linewidth=0.8, alpha = 0.5) +
   scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 12, 15, 17, 19)]) +
@@
 #| out-width: 100%
-#| fig-cap: "Assessing which of the 2 UMAP layouts with different hyper-parameter settings (n_neighbors: $30$, min_dist: $0.3$ (red); n_neighbors: $80$, min_dist: $0.5$ (blue)) on the PBMC data is the better representation using MSE for varying binwidth ($a_1$).
+#| fig-cap: "Assessing which of the two UMAP layouts with different hyper-parameter settings (n_neighbors: $30$, min_dist: $0.3$ (red); n_neighbors: $80$, min_dist: $0.5$ (blue)) on the PBMC data is the better representation using MSE for varying binwidth ($a_1$).
 Colour used for the lines and points in the top plot and in the scatterplots represents UMAP layout (a, b).

### 2025-04-20 appendix.tex
More described the approach

@@
 
-Figure~\ref{fig-pbmc-mse-umap} shows layout from the original published
-article along with that suggested by scDEED to be more accurate. The log
-MSE vs binwidth plot illustrates that our approach would suggest that
-scDEED is correct here, that layout b is more accurately reflecting the
-cluster structure in the data.
+The layout a (Figure~\ref{fig-pbmc-mse-umap}) is generated from the
+hyper-parameters suggested by \citet{chen2023}, and the layout b
+(Figure~\ref{fig-pbmc-mse-umap}) is with suggested hyper-parameters by
+scDEED to be more accurate. The layout a and b contain \(46\) and \(15\)
+dubious cells respectively.
+
+The log MSE vs binwidth (\(a_1\)) plot (Figure~\ref{fig-pbmc-mse-umap})
+illustrates that our approach would suggest that scDEED is correct here,
+that layout b is more accurately reflecting the cluster structure in the
+PBMC data.
 
@@
 
-\caption{\label{fig-pbmc-mse-umap}Assessing which of the 2 UMAP layouts
-with different hyper-parameter settings (n\_neighbors: \(30\),
+\caption{\label{fig-pbmc-mse-umap}Assessing which of the two UMAP
+layouts with different hyper-parameter settings (n\_neighbors: \(30\),
 min\_dist: \(0.3\) (red); n\_neighbors: \(80\), min\_dist: \(0.5\)

### 2025-04-20 appendix.qmd
Updated Figure 4

@@
 For illustration purposes, we only selected cells generated with inDrops ($n=5858$ cells) and UMAP cell embeddings.
-Also, @xia2023 used first $50$ principal components to generate the UMAP.
+Also, @xia2023 used first $9$ principal components to generate the UMAP.
 The objective is to assess the optimized layout by scDEED, and if it does not accurately represent the three clusters with small separations of the PBMC dataset, then select a reasonable \gD{} layout.
@@
 
-error_umap_plot_pbmc <- ggplot(error_pbmc,
-                               aes(x = a1,
-                                   y = MSE,
-                                   colour = method)) +
-  geom_point(size = 0.8) +
-  geom_line(linewidth = 0.3) +
-  scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 12, 15, 17, 19)]) +
-  scale_color_manual(values=c('#e41a1c','#377eb8')) +
-  scale_y_log10() +
-  ylab("log(MSE)") +
-  xlab(expression(paste("binwidth (", a[1], ")"))) +
-  theme_minimal() +
-  theme(panel.border = element_rect(fill = 'transparent'),
-        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
-        axis.ticks.x = element_line(),
-        axis.ticks.y = element_line(),
-        legend.position = "none",
-        axis.text.x = element_text(size = 7),
-        axis.text.y = element_text(size = 7),
-        axis.title.x = element_text(size = 7),
-        axis.title.y = element_text(size = 7))
-
+error_umap_plot_pbmc <-  plot_mse(error_pbmc) +
+  scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 12, 15, 17, 19, 20)]) +
+  scale_color_manual(values=c('#e41a1c','#377eb8'))
 ```
@@
 #| label: published-pbmc-scdeed
-#| eval: false
 
@@
 #| label: hexbin-pbmc-scdeed
-#| eval: false
+
 ## Compute hexbin parameters
@@
 pbmc_model <- fit_highd_model(
-  training_data = training_data_pbmc,
-  emb_df = umap_pbmc_scaled,
+    highd_data = training_data_pbmc,
+    nldr_data = umap_pbmc_scaled,
     bin1 = num_bins_x_pbmc,
@@
     q = 0.1,
-  is_bin_centroid = TRUE,
-  is_rm_lwd_hex = FALSE,
-  col_start_highd = "pc"
+    is_bin_centroid = TRUE
   )
@@
 #| label: pbmc-umap-model-proj-scdeed
-#| eval: false
 
@@
 projection_scaled <- projection * 3
-
-# projection_filtered <- projection
-# projection_filtered[apply(projection_filtered, 1, function(row) all(row < 0)), ] <- 0.75
-
-#rows_to_keep <- c(1, 2, 4, 5, 6)
-
-#projection_filtered <- projection[rows_to_keep, ]
 
@@
 ```{r}
-#| echo: false
-#| eval: false
 #| fig-cap: "Model in \\gD{}, on the layout a, and two views of the fit in projections from $50\\text{-}D$, for the PBMC data ($n =  5858$ and $p = 50$).

### 2025-04-20 appendix.tex
Updated Figure 4

@@
 (\(n=5858\) cells) and UMAP cell embeddings. Also, \citet{xia2023} used
-first \(50\) principal components to generate the UMAP. The objective is
+first \(9\) principal components to generate the UMAP. The objective is
 to assess the optimized layout by scDEED, and if it does not accurately
@@
 
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=0.3\linewidth,height=\textheight,keepaspectratio]{appendix_files/figure-pdf/fig-model-pbmc-author-proj-scdeed-1.pdf}
+
+}
+
+\caption{\label{fig-model-pbmc-author-proj-scdeed}Model in \gD{}, on the
+layout a, and two views of the fit in projections from \(50\text{-}D\),
+for the PBMC data (\(n =  5858\) and \(p = 50\)). Layout shows close
+clusters. In \(50\text{-}D\), the data shows}
+
+\end{figure}%
+
   \bibliography{paper.bib}

### 2025-04-20 scripts/pbmc3k/13_gen_mse_for_diff_umap_scdeed.R

@@ -1,3 +1,4 @@
+## This script is to generate MSE for hyper-parameter suggested by Chen and scDEED
 library(readr)
 library(quollr)
 library(dplyr)
@@ -36,14 +37,12 @@ for (xbins in bin1_vec_pbmc) {
   a1 <- hb_obj$a1
 
   pbmc_model <- fit_highd_model(
-    training_data = training_data_pbmc,
-    emb_df = umap_pbmc_scaled,
+    highd_data = training_data_pbmc,
+    nldr_data = umap_pbmc_scaled,
     bin1 = xbins,
     r2 = r2_umap,
     q = 0.1,
-    is_bin_centroid = TRUE,
-    is_rm_lwd_hex = FALSE,
-    col_start_highd = "PC_"
+    is_bin_centroid = TRUE
   )
 
   df_bin_centroids_pbmc <- pbmc_model$df_bin_centroids
@@ -51,12 +50,9 @@ for (xbins in bin1_vec_pbmc) {
 
   ## Compute error
   error_df <- glance(
-    df_bin_centroids = df_bin_centroids_pbmc,
-    df_bin = df_bin_pbmc,
-    training_data = training_data_pbmc,
-    newdata = NULL,
-    type_NLDR = "UMAP",
-    col_start = "PC_") |>
+    model_2d = df_bin_centroids_pbmc,
+    model_highd = df_bin_pbmc,
+    highd_data = training_data_pbmc) |>
     mutate(bin1 = xbins,
            bin2 = bin2,
            b = bin1 * bin2,
@@ -75,7 +71,7 @@ write_rds(error_pbmc_umap, "data/pbmc3k/error_scdeed_pbmc_umap_30_min_dist_0.3.r
 ## For umap
 umap_pbmc <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_80_min_dist_0.5.rds")
 umap_pbmc <- as_tibble(umap_pbmc)
-names(umap_pbmc) <- c("UMAP1", "UMAP2")
+names(umap_pbmc) <- c("emb1", "emb2")
 umap_pbmc <- umap_pbmc |>
   mutate(ID = 1:NROW(umap_pbmc))
 
@@ -99,14 +95,12 @@ for (xbins in bin1_vec_pbmc) {
   a1 <- hb_obj$a1
 
   pbmc_model <- fit_highd_model(
-    training_data = training_data_pbmc,
-    emb_df = umap_pbmc_scaled,
[…]

### 2025-04-20 appendix.qmd
Added initial image for Figure 5

@@
 # and plot them on same aspect ratio
-umap_pbmc1 <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_30_min_dist_0.3.rds")
-umap_pbmc1 <- as_tibble(umap_pbmc1)
-names(umap_pbmc1) <- c("UMAP1", "UMAP2")
-
-nldr1c <- umap_pbmc1 |>
-  ggplot(aes(x = UMAP1,
-             y = UMAP2)) +
+umap_pbmc <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_30_min_dist_0.3.rds")
+umap_pbmc <- as_tibble(umap_pbmc)
+names(umap_pbmc) <- c("emb1", "emb2")
+umap_pbmc <- umap_pbmc |>
+  dplyr::mutate(ID = row_number())
+
+nldr1c <- umap_pbmc |>
+  ggplot(aes(x = emb1,
+             y = emb2)) +
   geom_point(alpha=0.1, size=1, colour='#e41a1c') +
@@
 
-umap_pbmc2 <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_80_min_dist_0.5.rds")
-umap_pbmc2 <- as_tibble(umap_pbmc2)
-names(umap_pbmc2) <- c("UMAP1", "UMAP2")
-
-nldr2c <- umap_pbmc2 |>
-  ggplot(aes(x = UMAP1,
-             y = UMAP2))+
+tsne_pbmc <- read_rds("data/pbmc3k/pbmc_scdeed_tsne_perplexity_30.rds")
+tsne_pbmc <- as_tibble(tsne_pbmc)
+names(tsne_pbmc) <- c("emb1", "emb2")
+tsne_pbmc <- tsne_pbmc |>
+  dplyr::mutate(ID = row_number())
+
+nldr2c <- tsne_pbmc |>
+  ggplot(aes(x = emb1,
+             y = emb2))+
   geom_point(alpha=0.1, size=1, colour='#377eb8') +
@@
 ```{r}
-#| label: combine-data-pbmc-umap
+#| label: combine-data-pbmc-umap-tsne
 
 error_pbmc_umap <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_30_min_dist_0.3.rds")
-error_pbmc_umap2 <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_80_min_dist_0.5.rds")
+error_pbmc_tsne <- read_rds("data/pbmc3k/error_scdeed_pbmc_tsne_perplexity_30.rds")
 
 error_pbmc <- bind_rows(error_pbmc_umap,
-                        error_pbmc_umap2)
+                        error_pbmc_tsne)
 
@@
   scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 12, 15, 17, 19, 20)]) +
-  scale_color_manual(values=c('#e41a1c','#377eb8'))
+  scale_color_manual(values=c('#377eb8', '#e41a1c'))
 ```
@@
 ```{r}
-#| label: published-pbmc-scdeed
-
+#| label: hexbin-pbmc
 training_data_pbmc <- read_rds("data/pbmc3k/pbmc_pca_50_scdeed.rds")
-training_data_pbmc <- training_data_pbmc[, 1:50] |>
+training_data_pbmc <- training_data_pbmc[, 1:9] |>
   mutate(ID = 1:NROW(training_data_pbmc))
 
-## For umap
-umap_pbmc <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_80_min_dist_0.5.rds")
-umap_pbmc <- as_tibble(umap_pbmc)
-names(umap_pbmc) <- c("UMAP1", "UMAP2")
-umap_pbmc <- umap_pbmc |>
-  mutate(ID = 1:NROW(umap_pbmc))
-
-pbmc_scaled_obj_umap <- gen_scaled_data(
-  data = umap_pbmc)
-umap_pbmc_scaled <- pbmc_scaled_obj_umap$scaled_nldr
-
-lim1 <- pbmc_scaled_obj_umap$lim1
-lim2 <- pbmc_scaled_obj_umap$lim2
-r2_umap <- diff(lim2)/diff(lim1)
-
-```
-
-```{r}
-#| label: hexbin-pbmc-scdeed
+names(training_data_pbmc) <- append(paste0("x", 1:9), "ID")
 
 ## Compute hexbin parameters
-num_bins_x_pbmc <- 12
-lim1 <- pbmc_scaled_obj_umap$lim1
-lim2 <- pbmc_scaled_obj_umap$lim2
-r2_pbmc <- diff(lim2)/diff(lim1)
-
-pbmc_model <- fit_highd_model(
-    highd_data = training_data_pbmc,
-    nldr_data = umap_pbmc_scaled,
-    bin1 = num_bins_x_pbmc,
-    r2 = r2_pbmc,
-    q = 0.1,
-    is_bin_centroid = TRUE
-  )
-
-df_bin_centroids_pbmc <- pbmc_model$df_bin_centroids
-df_bin_pbmc <- pbmc_model$df_bin
-
-## Triangulate bin centroids
-tr1_object_pbmc <- tri_bin_centroids(
-  df_bin_centroids_pbmc, x = "c_x", y = "c_y")
-tr_from_to_df_pbmc <- gen_edges(
-  tri_object = tr1_object_pbmc)
-
-## Compute 2D distances
-distance_pbmc <- cal_2d_dist(
-  tr_coord_df = tr_from_to_df_pbmc,
-  start_x = "x_from",
-  start_y = "y_from",
-  end_x = "x_to",
-  end_y = "y_to",
-  select_vars = c("from", "to", "distance"))
-
-## To find the benchmark value
-benchmark_pbmc <- find_lg_benchmark(
-  distance_edges = distance_pbmc,
-  distance_col = "distance")
-
-tr_df <- distinct(tibble(
-  x = c(tr_from_to_df_pbmc[["x_from"]], tr_from_to_df_pbmc[["x_to"]]),
-  y = c(tr_from_to_df_pbmc[["y_from"]], tr_from_to_df_pbmc[["y_to"]])))
+num_bins_x_pbmc <- 22
+
+algo_obj_pbmc <- gen_nldr_vis_algo_obj(
+  high_d_data = training_data_pbmc, 
+  nldr_data = umap_pbmc, 
+  num_x_bins = num_bins_x_pbmc)
+
+umap_pbmc_scaled <- algo_obj_pbmc$nldr_scaled
+distance_pbmc <- algo_obj_pbmc$distance_df
+tr_from_to_df_pbmc <- algo_obj_pbmc$tr_from_to_df
+benchmark_pbmc <- algo_obj_pbmc$benchmark
+df_bin_centroids_pbmc <- algo_obj_pbmc$df_bin_centroids
+df_bin_pbmc <- algo_obj_pbmc$df_bin
 
 distance_df_small_edges_pbmc <- distance_pbmc |>
-  filter(distance < benchmark_pbmc)
-
-tr_from_to_df_pbmc <- inner_join(
+  filter(distance < benchmark_pbmc)#benchmark_pbmc
+
+umap_pbmc_scaled_with_cluster <- umap_pbmc_scaled |>
+  mutate(cluster = if_else(emb1 <= 0.3, "cluster1", if_else((emb1 >= 0.75) & (emb2 <= 0.25), "cluster3", "cluster2"))) |>
+  mutate(cluster = factor(cluster, levels = c("cluster1", "cluster2", "cluster3")))
+
+tr_from_to_df_pbmc <- right_join(
   tr_from_to_df_pbmc, distance_df_small_edges_pbmc, 
@@
 trimesh_removed_pbmc <- ggplot() + 
+  geom_point(
+    data = umap_pbmc_scaled_with_cluster,
+    aes(
+      x = emb1,
+      y = emb2,
+      color = cluster
+    ),
+    alpha = 0.2,
+    size = 0.1
+  )  +
   geom_segment(data = tr_from_to_df_pbmc, 
@@
                linewidth = 1) +
-  geom_point(
-    data = umap_pbmc_scaled,
-    aes(
-      x = UMAP1,
-      y = UMAP2
-    ),
-    alpha = 0.2,
-    color = '#a65628'
-  )  +
+  scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada")) +
   interior_annotation("a1", cex = 2) +
@@
 ```{r}
-#| label: pbmc-umap-model-proj-scdeed
+#| label: prep-pbmc-umap-model-proj
 
@@
   select(-type)
-
-# Combine with the true model for visualization
-df <- dplyr::bind_rows(scaled_pbmc_data_model |> mutate(type = "model"),
-                       scaled_pbmc_data |> mutate(type = "data"))
+```
+
+```{r}
+#| label: langevitour-pbmc-umap-author-proj
+#| eval: false
+
+data_pbmc_n <- data_pbmc |>
+  select(-type) |>
+  mutate(type = as.character(umap_pbmc_scaled_with_cluster$cluster))
+
+df_model_data_pbmc_n <- bind_rows(df_b_pbmc, data_pbmc_n)
+
+langevitour::langevitour(df_model_data_pbmc_n[1:(length(df_model_data_pbmc_n)-1)],
+                         lineFrom = distance_df_small_edges_pbmc$from,
+                         lineTo = distance_df_small_edges_pbmc$to,
+                         group = factor(df_model_data_pbmc_n$type,
+                                        c("cluster1", "cluster2", "cluster3", "model")),
+                         levelColors = c("#8dd3c7", "#fdb462", "#bebada", "#000000"))
+```
+
+```{r}
+#| label: pbmc-umap-model-proj
 
@@
 projection <- cbind(
-  c(0.040856,0.000723,0.006000,0.007623,0.001986,-0.006599,0.005311,-0.000222,0.001648,0.008067,-0.001587,0.004423,0.004851,0.000638,0.002606,0.006616,0.006188,-0.000954,-0.004603,0.004998,-0.012220,0.008401,-0.007788,0.008972,-0.003903,0.003348,-0.001817,-0.000385,-0.002437,-0.004097,-0.001357,0.013660,0.001770,0.006982,0.002828,-0.009169,-0.003541,0.008582,-0.000032,-0.002375,0.005854,-0.005466,-0.000824,-0.001435,0.004148,0.007141,0.000883,-0.004159,0.005401,0.007882),
-  c(-0.000590,0.039365,-0.004481,-0.009041,-0.003805,0.003278,-0.009601,-0.000650,-0.002333,-0.005697,-0.001488,-0.009808,0.002557,0.004113,0.005705,-0.004665,-0.001170,-0.004727,-0.000845,0.005084,-0.018054,0.002538,0.006606,-0.007184,0.000238,0.001877,-0.002124,-0.005972,-0.009729,0.004958,-0.003181,-0.005185,-0.009598,-0.004243,0.007716,0.000396,-0.004258,0.005932,-0.003242,-0.004184,0.003493,-0.008405,0.000860,0.008304,0.004569,0.000798,0.004156,-0.005777,0.005278,-0.002531))
-
-projection_scaled <- projection * 3
-
-projected <- as.matrix(scaled_pbmc_data) %*% projection_scaled
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  #dplyr::mutate(type = df_exe$type) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection_scaled
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  distance_df_small_edges_pbmc |> select(-distance),
-  projected_model_df,
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-axes_obj <- gen_axes(
-  proj = projection,
-  limits = 0.04,
-  axis_pos_x = -0.025,
-  axis_pos_y = -0.025,
-  axis_labels = names(scaled_pbmc_data),
-  threshold = 0.0001)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
-
-pbmc_proj_umap_model1 <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1,
-      y = proj2)) +
-  geom_segment(
-    data = model_df,
-    aes(
-      x = proj1_from,
-      y = proj2_from,
-      xend = proj1_to,
-      yend = proj2_to),
-    color = "#000000") +
-  geom_point(
-    #size = 0.5,
-    alpha = 0.2,
-    color = '#a65628') +
-  geom_segment(
-    data=axes,
-    aes(x=x1, y=y1, xend=x2, yend=y2),
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2),
-    label=rownames(axes),
-    colour="grey50",
-    size = 3) +
-  geom_path(
-    data=circle,
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.03, 0.06)) +
-  ylim(c(-0.03, 0.06)) +
-  interior_annotation("a2", cex = 2)
+    c(-0.4545,-0.5533,-0.1235,0.3984,-0.2065,-0.3586,-0.0088,0.3572,-0.1368),
+    c(0.4576,-0.3157,0.4708,0.1410,-0.4479,-0.1685,0.3668,-0.2692,0.1332))
+
+proj_obj1 <- get_projection(projection = projection, 
+               proj_scale = 1.23, 
+               scaled_data = scaled_pbmc_data, 
+               scaled_data_model = scaled_pbmc_data_model, 
+               distance_df_small_edges = distance_df_small_edges_pbmc, 
+               axis_param = list(limits = 1,
+                                 axis_scaled = 1, 
+                                 axis_pos_x = -0.48, 
+                                 axis_pos_y = -0.48, 
+                                 threshold = 0.082))
+
+proj_obj1[["cluster"]] <- as.character(umap_pbmc_scaled_with_cluster$cluster)
+
+pbmc_proj_umap_model1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2), # size, alpha, color
+  plot_limits = c(-0.65, 0.65), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 4,
+  is_color = TRUE) + scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))
 
@@
 projection <- cbind(
-  c(-0.019825,-0.011101,0.011543,-0.006943,-0.004485,0.000062,0.002023,-0.005178,0.000800,-0.002903,-0.014182,0.008756,0.009595,0.007032,-0.007820,0.006143,0.004865,0.003352,0.007554,0.002265,0.012035,0.009961,-0.001825,-0.004942,-0.002237,0.002308,0.009531,-0.001146,0.013401,-0.001022,0.008434,0.000477,0.004239,-0.006772,-0.004702,-0.011432,-0.000431,0.008146,-0.009299,-0.004782,0.003457,-0.008214,-0.009604,0.011329,0.003805,-0.003004,-0.000800,0.015288,-0.013157,0.006984),
-  c(-0.001387,0.001643,0.010470,0.002930,-0.002335,-0.010569,0.013839,0.002917,-0.006825,0.010191,-0.006367,0.002220,0.002506,0.005139,-0.007213,-0.006235,-0.007123,-0.000222,-0.002225,0.011592,-0.003942,0.002788,-0.000055,0.011316,-0.009753,0.020785,0.004615,0.005172,0.011182,-0.003125,-0.010860,0.007820,-0.017510,0.007010,0.007267,0.002694,-0.003949,-0.009188,-0.005163,0.001898,-0.004744,0.013757,0.010630,-0.007688,0.004598,0.004658,-0.011154,0.003584,-0.005226,0.006650))
-
-projection_scaled <- projection * 3
-
-projected <- as.matrix(scaled_pbmc_data) %*% projection_scaled
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  #dplyr::mutate(type = df_exe$type) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-projected_model <- as.matrix(scaled_pbmc_data_model) %*% projection_scaled
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  distance_df_small_edges_pbmc |> select(-distance),
-  projected_model_df,
-  by = c("from" = "ID"))
-
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-axes_obj <- gen_axes(
-  proj = projection,
-  limits = 0.03,
-  axis_pos_x = -0.02,
-  axis_pos_y = -0.02,
-  axis_labels = names(scaled_pbmc_data),
-  threshold = 0)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
-
-pbmc_proj_umap_model2 <- projected_df |>
-  ggplot(
+    c(0.0734,0.6138,-0.4671,0.2589,0.1189,0.0778,-0.0819,0.0308,-0.5561),
+    c(0.6625,0.0034,0.2412,0.3360,0.2463,0.1070,-0.2903,-0.4705,0.1293))
+
+proj_obj2 <- get_projection(projection = projection, 
+               proj_scale = 1.25, 
+               scaled_data = scaled_pbmc_data, 
+               scaled_data_model = scaled_pbmc_data_model, 
+               distance_df_small_edges = distance_df_small_edges_pbmc, 
+               axis_param = list(limits = 1.35, 
+                                 axis_scaled = 1, 
+                                 axis_pos_x = -0.75, 
+                                 axis_pos_y = -0.75, 
+                                 threshold = 0.12))
+
+proj_obj2[["cluster"]] <- as.character(umap_pbmc_scaled_with_cluster$cluster)
+
+pbmc_proj_umap_model2 <- plot_proj(
+    proj_obj = proj_obj2, 
+    point_param = c(1.5, 0.2), # size, alpha, color
+    plot_limits = c(-1, 0.75), 
+    title = "a3", 
+    cex = 2, 
+    axis_text_size = 4,
+    is_color = TRUE) + scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))
+
+```
+
+```{r}
+#| label: hexbin-pbmc2
+
+## Compute hexbin parameters
+num_bins_x_pbmc <- 22
+
+algo_obj_pbmc <- gen_nldr_vis_algo_obj(
+  high_d_data = training_data_pbmc, 
+  nldr_data = tsne_pbmc, 
+  num_x_bins = num_bins_x_pbmc)
+
+tsne_pbmc_scaled <- algo_obj_pbmc$nldr_scaled
+distance_pbmc <- algo_obj_pbmc$distance_df
+tr_from_to_df_pbmc <- algo_obj_pbmc$tr_from_to_df
+benchmark_pbmc <- algo_obj_pbmc$benchmark
+df_bin_centroids_pbmc <- algo_obj_pbmc$df_bin_centroids
+df_bin_pbmc <- algo_obj_pbmc$df_bin
+
+distance_df_small_edges_pbmc <- distance_pbmc |>
+  filter(distance < benchmark_pbmc)#benchmark_pbmc
+
+tsne_pbmc_scaled_with_cluster <- tsne_pbmc_scaled |>
+  mutate(cluster = if_else((emb1 <= 0.4) & (emb2 <= 0.6), "cluster1", if_else((emb1 >= 0.75) & (emb2 <= 0.25), "cluster3", "cluster2"))) |>
+  mutate(cluster = factor(cluster, levels = c("cluster1", "cluster2", "cluster3")))
+
+tr_from_to_df_pbmc <- right_join(
+  tr_from_to_df_pbmc, distance_df_small_edges_pbmc, 
+  by = c("from", "to"))
+
+trimesh_removed_pbmc_best <- ggplot() + 
+  geom_point(
+    data = tsne_pbmc_scaled_with_cluster,
     aes(
-      x = proj1,
-      y = proj2)) +
-  geom_segment(
-    data = model_df,
+      x = emb1,
+      y = emb2,
+      color = cluster
+    ),
+    alpha = 0.2,
+    size = 0.1
+  )  +
+  geom_segment(data = tr_from_to_df_pbmc, 
                aes(
-      x = proj1_from,
-      y = proj2_from,
-      xend = proj1_to,
-      yend = proj2_to),
-    color = "#000000") +
-  geom_point(
-    #size = 0.5,
-    alpha = 0.2,
-    color = '#a65628') +
-  geom_segment(
-    data=axes,
-    aes(x=x1, y=y1, xend=x2, yend=y2),
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2),
-    label=rownames(axes),
-    colour="grey50",
-    size = 3) +
-  geom_path(
-    data=circle,
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.025, 0.02)) +
-  ylim(c(-0.025, 0.02)) +
-  interior_annotation("a3", cex = 2)
-```
-
-```{r}
-#| fig-cap: "Model in \\gD{}, on the layout a, and two views of the fit in projections from $50\\text{-}D$, for the PBMC data ($n =  5858$ and $p = 50$).
-Layout shows close clusters.
-In $50\\text{-}D$, the data shows "
-#| label: fig-model-pbmc-author-proj-scdeed
+                 x = x_from, 
+                 y = y_from, 
+                 xend = x_to, 
+                 yend = y_to),
+               colour = "#000000",
+               linewidth = 1) +
+  scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada")) +
+  interior_annotation("a1", cex = 2) +
+  theme(
+    aspect.ratio = 1
+  )
+```
+
+```{r}
+#| label: prep-pbmc-tsne-model-proj-best
+
+data_pbmc <- training_data_pbmc |> 
+  select(-ID) |>
+  mutate(type = "data")
+
+df_b_pbmc <- df_bin_pbmc |>
+  dplyr::filter(hb_id %in% df_bin_centroids_pbmc$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b_pbmc <- df_b_pbmc[match(df_bin_centroids_pbmc$hexID, df_b_pbmc$hb_id),] |>
+  dplyr::select(-hb_id)
+
+# Apply the scaling
+df_model_data_pbmc <- bind_rows(data_pbmc, df_b_pbmc)
+scaled_pbmc <- scale_data_manual(df_model_data_pbmc, "type") |>
+  as_tibble()
+
+scaled_pbmc_data <- scaled_pbmc |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_pbmc_data_model <- scaled_pbmc |>
+  filter(type == "model") |>
+  select(-type)
+```
+
+```{r}
+#| label: langevitour-pbmc-tsne-author-proj-best
+#| eval: false
+
+data_pbmc_n <- data_pbmc |>
+  select(-type) |>
+  mutate(type = as.character(tsne_pbmc_scaled_with_cluster$cluster))
+
+df_model_data_pbmc_n <- bind_rows(df_b_pbmc, data_pbmc_n)
+
+langevitour::langevitour(df_model_data_pbmc_n[1:(length(df_model_data_pbmc_n)-1)],
+                         lineFrom = distance_df_small_edges_pbmc$from,
+                         lineTo = distance_df_small_edges_pbmc$to,
+                         group = factor(df_model_data_pbmc_n$type,
+                                        c("cluster1", "cluster2", "cluster3", "model")),
+                         levelColors = c("#8dd3c7", "#fdb462", "#bebada", "#000000"))
+```
+
+```{r}
+#| label: pbmc-tsne-model-proj-best
+
+## First projection
+projection <- cbind(
+  c(-0.4545,-0.5533,-0.1235,0.3984,-0.2065,-0.3586,-0.0088,0.3572,-0.1368),
+  c(0.4576,-0.3157,0.4708,0.1410,-0.4479,-0.1685,0.3668,-0.2692,0.1332))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.23, 
+                            scaled_data = scaled_pbmc_data, 
+                            scaled_data_model = scaled_pbmc_data_model, 
+                            distance_df_small_edges = distance_df_small_edges_pbmc, 
+                            axis_param = list(limits = 1,
+                                              axis_scaled = 1, 
+                                              axis_pos_x = -0.48, 
+                                              axis_pos_y = -0.48, 
+                                              threshold = 0.082))
+
+proj_obj1[["cluster"]] <- as.character(tsne_pbmc_scaled_with_cluster$cluster)
+
+pbmc_proj_tsne_model1_best <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(1.5, 0.2), # size, alpha, color
+  plot_limits = c(-0.65, 0.65), 
+  title = "a2", 
+  cex = 2, 
+  axis_text_size = 4,
+  is_color = TRUE) + scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))
+
+## Second projection
+projection <- cbind(
+  c(0.0734,0.6138,-0.4671,0.2589,0.1189,0.0778,-0.0819,0.0308,-0.5561),
+  c(0.6625,0.0034,0.2412,0.3360,0.2463,0.1070,-0.2903,-0.4705,0.1293))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.25, 
+                            scaled_data = scaled_pbmc_data, 
+                            scaled_data_model = scaled_pbmc_data_model, 
+                            distance_df_small_edges = distance_df_small_edges_pbmc, 
+                            axis_param = list(limits = 1.35, 
+                                              axis_scaled = 1, 
+                                              axis_pos_x = -0.75, 
+                                              axis_pos_y = -0.75, 
+                                              threshold = 0.12))
+
+proj_obj2[["cluster"]] <- as.character(tsne_pbmc_scaled_with_cluster$cluster)
+
+pbmc_proj_tsne_model2_best <- plot_proj(
+  proj_obj = proj_obj2, 
+  point_param = c(1.5, 0.2), # size, alpha, color
+  plot_limits = c(-1, 0.75), 
+  title = "a3", 
+  cex = 2, 
+  axis_text_size = 4,
+  is_color = TRUE) + scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))
+
+```
+
+```{r}
+#| echo: false
+#| fig-cap: ""
+#| label: fig-model-pbmc-author-proj
 #| fig-pos: H
-#| fig-width: 15
-#| fig-height: 5
-#| out-width: 30%
-
-free(trimesh_removed_pbmc) +
-  pbmc_proj_umap_model1 +
-  pbmc_proj_umap_model2 +
-  plot_layout(ncol=3) &
+#| fig-width: 10
+#| fig-height: 15
+#| out-width: 90%
+
+free(trimesh_removed_pbmc) + free(trimesh_removed_pbmc_best) +
+  pbmc_proj_umap_model1 + pbmc_proj_tsne_model1_best +
+  pbmc_proj_umap_model2 + pbmc_proj_tsne_model2_best +
+  plot_layout(nrow=2, byrow=FALSE) &
   theme(legend.position='none')

### 2025-04-20 appendix.qmd
Added cluster colors

@@
 distance_df_small_edges_pbmc <- distance_pbmc |>
-  filter(distance < benchmark_pbmc)#benchmark_pbmc
+  filter(distance < 0.1)#benchmark_pbmc
 
 tsne_pbmc_scaled_with_cluster <- tsne_pbmc_scaled |>
-  mutate(cluster = if_else((emb1 <= 0.4) & (emb2 <= 0.6), "cluster1", if_else((emb1 >= 0.75) & (emb2 <= 0.25), "cluster3", "cluster2"))) |>
+  mutate(cluster = if_else((emb2 <= -emb1 + 0.8) & (emb1 < 0.5), "cluster1", if_else(emb2 >= -2 * emb1 + 1.79, "cluster3", "cluster2"))) |>
   mutate(cluster = factor(cluster, levels = c("cluster1", "cluster2", "cluster3")))
@@
   scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada")) +
-  interior_annotation("a1", cex = 2) +
+  interior_annotation("b1", cex = 2) +
   theme(
@@
   plot_limits = c(-0.65, 0.65), 
-  title = "a2", 
+  title = "b2", 
   cex = 2, 
@@
   plot_limits = c(-1, 0.75), 
-  title = "a3", 
+  title = "b3", 
   cex = 2, 

### 2025-04-20 appendix.qmd
Added figure title

@@
 ## Compute hexbin parameters
-num_bins_x_pbmc <- 22
+num_bins_x_pbmc <- 30
 
@@
 distance_df_small_edges_pbmc <- distance_pbmc |>
-  filter(distance < 0.1)#benchmark_pbmc
+  filter(distance < benchmark_pbmc)#benchmark_pbmc
 
@@
 #| echo: false
-#| fig-cap: ""
+#| fig-cap: "Compare the published \\gD{} layout (@fig-pbmc-mse-umap a) and the \\gD{} layout selected (@fig-pbmc-mse-umap b) by MSE plot (@fig-pbmc-mse-umap) from tSNE and UMAP with default hyper-parameters.
+The PBMC3k data ($n =  5858$) has three clusters in $9\\text{-}D$, where three clusters are close.
+Two \\gD{} projection from a tour on $9\\text{-}D$ of the model fit with @fig-pbmc-mse-umap a ($a_1 = 0.06$, $b = 616/86 (22, 28)$) shows three-well separated clusters with big separations.
+On the other hand, the model fit with @fig-pbmc-mse-umap b ($a_1 = 0.06$, $b = 704/227 (22, 32)$) shows three-well separated clusters with small separations.
+Therefore, @fig-pbmc-mse-umap b is more reasonable than @fig-pbmc-mse-umap a.
+Videos of the langevitour animations are available at <https://youtu.be/0cKX_HG_n0k> and <https://youtu.be/KhJvsRtaX04> respectively."
 #| label: fig-model-pbmc-author-proj

### 2025-04-20 appendix.tex
Added initial image for Figure 5
Added figure title

--- 2025-04-20 appendix.tex
@@
 
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=0.9\linewidth,height=\textheight,keepaspectratio]{appendix_files/figure-pdf/fig-model-pbmc-author-proj-1.pdf}
+
+}
+
+\caption{\label{fig-model-pbmc-author-proj}Compare the published \gD{}
+layout (Figure~\ref{fig-pbmc-mse-umap} a) and the \gD{} layout selected
+(Figure~\ref{fig-pbmc-mse-umap} b) by MSE plot
+(Figure~\ref{fig-pbmc-mse-umap}) from tSNE and UMAP with default
+hyper-parameters. The PBMC3k data (\(n =  5858\)) has three clusters in
+\(9\text{-}D\), where three clusters are close. Two \gD{} projection
+from a tour on \(9\text{-}D\) of the model fit with
+Figure~\ref{fig-pbmc-mse-umap} a (\(a_1 = 0.06\),
+\(b = 616/86 (22, 28)\)) shows three-well separated clusters with big
+separations. On the other hand, the model fit with
+Figure~\ref{fig-pbmc-mse-umap} b (\(a_1 = 0.06\),
+\(b = 704/227 (22, 32)\)) shows three-well separated clusters with small
+separations. Therefore, Figure~\ref{fig-pbmc-mse-umap} b is more
+reasonable than Figure~\ref{fig-pbmc-mse-umap} a. Videos of the
+langevitour animations are available at
+\url{https://youtu.be/0cKX_HG_n0k} and
+\url{https://youtu.be/KhJvsRtaX04} respectively.}
+
+\end{figure}%
+
   \bibliography{paper.bib}

### 2025-04-20 paper.qmd
Added figure title

@@
 #| echo: false
-#| fig-cap: "Compare the published \\gD{} layout (@fig-pbmc-mse a) and the \\gD{} layout selected (@fig-pbmc-mse e) by MSE plot (@fig-pbmc-mse) from the same NLDR with different hyper-parameter choice.
+#| fig-cap: "Compare the published \\gD{} layout (@fig-pbmc-mse a) and the \\gD{} layout selected (@fig-pbmc-mse e) by MSE plot (@fig-pbmc-mse) from the UMAP and tSNE with default hyper-parameters.
 The PBMC3k data ($n =  2622$) has three clusters in $9\\text{-}D$, where two of them being close and one separated from the others.

### 2025-04-20 appendix.qmd
Added dobious cells

@@
 
-For illustration purposes, we only selected cells generated with inDrops ($n=5858$ cells) and UMAP cell embeddings.
-Also, @xia2023 used first $9$ principal components to generate the UMAP.
+For illustration purposes, we only selected cells generated with inDrops ($n=5858$ cells) and UMAP and tSNE cell embeddings.
+Also, @xia2023 used first $9$ principal components to generate the UMAP and tSNE.
 The objective is to assess the optimized layout by scDEED, and if it does not accurately represent the three clusters with small separations of the PBMC dataset, then select a reasonable \gD{} layout.
@@
 The layout a (@fig-pbmc-mse-umap) is generated from the hyper-parameters suggested by @chen2023, and the layout b (@fig-pbmc-mse-umap) is with suggested hyper-parameters by scDEED to be more accurate.
-The layout a and b contain $46$ and $15$ dubious cells respectively.
+The layout a and b contain $46$ and $83$ dubious cells respectively.
 
@@
 #| out-width: 100%
-#| fig-cap: "Assessing which of the two UMAP layouts with different hyper-parameter settings (n_neighbors: $30$, min_dist: $0.3$ (red); n_neighbors: $80$, min_dist: $0.5$ (blue)) on the PBMC data is the better representation using MSE for varying binwidth ($a_1$).
-Colour used for the lines and points in the top plot and in the scatterplots represents UMAP layout (a, b).
+#| fig-cap: "Assessing which of the two layouts with UMAP and tSNE with different hyper-parameter setting (n_neighbors: $30$, min_dist: $0.3$ (red); n_neighbors: $30$ (blue)) on the PBMC data is the better representation using MSE for varying binwidth ($a_1$).
+Colour used for the lines and points in the top plot and in the scatterplots represents UMAP and tSNE layouts (a, b).
 Of the two, layout b is optimal across all binwidths making it the best choice."

### 2025-04-20 appendix.tex
Added dobious cells

@@
 For illustration purposes, we only selected cells generated with inDrops
-(\(n=5858\) cells) and UMAP cell embeddings. Also, \citet{xia2023} used
-first \(9\) principal components to generate the UMAP. The objective is
-to assess the optimized layout by scDEED, and if it does not accurately
-represent the three clusters with small separations of the PBMC dataset,
-then select a reasonable \gD{} layout.
+(\(n=5858\) cells) and UMAP and tSNE cell embeddings. Also,
+\citet{xia2023} used first \(9\) principal components to generate the
+UMAP and tSNE. The objective is to assess the optimized layout by
+scDEED, and if it does not accurately represent the three clusters with
+small separations of the PBMC dataset, then select a reasonable \gD{}
+layout.
 
@@
 (Figure~\ref{fig-pbmc-mse-umap}) is with suggested hyper-parameters by
-scDEED to be more accurate. The layout a and b contain \(46\) and \(15\)
+scDEED to be more accurate. The layout a and b contain \(46\) and \(83\)
 dubious cells respectively.
@@
 
-\caption{\label{fig-pbmc-mse-umap}Assessing which of the two UMAP
-layouts with different hyper-parameter settings (n\_neighbors: \(30\),
-min\_dist: \(0.3\) (red); n\_neighbors: \(80\), min\_dist: \(0.5\)
-(blue)) on the PBMC data is the better representation using MSE for
-varying binwidth (\(a_1\)). Colour used for the lines and points in the
-top plot and in the scatterplots represents UMAP layout (a, b). Of the
-two, layout b is optimal across all binwidths making it the best
-choice.}
+\caption{\label{fig-pbmc-mse-umap}Assessing which of the two layouts
+with UMAP and tSNE with different hyper-parameter setting (n\_neighbors:
+\(30\), min\_dist: \(0.3\) (red); n\_neighbors: \(30\) (blue)) on the
+PBMC data is the better representation using MSE for varying binwidth
+(\(a_1\)). Colour used for the lines and points in the top plot and in
+the scatterplots represents UMAP and tSNE layouts (a, b). Of the two,
+layout b is optimal across all binwidths making it the best choice.}
 

### 2025-04-20 appendix.qmd
Updated my name

@@
 author:
-  - name: Jayani P.G. Lakshika
+  - name: Jayani P. Gamage
     affiliations:

### 2025-04-20 appendix.tex
Updated my name

@@
   pdftitle={Appendix: Looking at Non-Linear Dimension Reductions as Models in the Data Space},
-  pdfauthor={Jayani P.G. Lakshika; Dianne Cook; Paul Harrison; Michael Lydeamore; Thiyanga S. Talagala},
+  pdfauthor={Jayani P. Gamage; Dianne Cook; Paul Harrison; Michael Lydeamore; Thiyanga S. Talagala},
   colorlinks=true,
@@
 \author{
-Jayani P.G. Lakshika\\
+Jayani P. Gamage\\
 Econometrics \& Business Statistics, Monash University\\

### 2025-04-24 paper.qmd
Updated Figure 18
Added interp citation

@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 14
+num_bins_x_gau1 <- 25
 lim1 <- gau1_scaled_obj$lim1
@@
 The purpose of this example is to *illustrate how to use our method to help decide on an appropriate NLDR layout that accurately represents the data*.
-The cluster results of Human Peripheral Blood Mononuclear Cells (PBMC3k) in @chen2023 are examined.
+The cluster results of Human Peripheral Blood Mononuclear Cells (PBMC3k) in @chen2024 are examined.
 There are $2622$ single cells, with $1000$ gene expressions.
@@
 
-imge_error_sample_outside <- plot_digit_img(pixels_gathered_within, palette = "Grays", title_text = "(d)", hjust = 0.31)
+imge_error_sample_outside <- plot_digit_img(pixels_gathered_outside, palette = "Grays", title_text = "(d)", hjust = 0.31)
 
@@
 Our algorithm can also be useful in practice to decide on a reasonable NLDR layout that accurately represents the data, as described with a single-cell gene expression data, **PBMC3k**.
-The published results of UMAP with default (hyper-)parameter setting by @chen2023 were evaluated.
+The published results of UMAP with default (hyper-)parameter setting by @chen2024 were evaluated.
 The data showed three clusters, two of which were close.

### 2025-04-24 paper.qmd
Updated Figure 11

@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 25
+num_bins_x_gau1 <- 23
 lim1 <- gau1_scaled_obj$lim1
@@
 distance_df_small_edges_gau1 <- distance_gau1 |>
-  filter(distance < 0.1) #benchmark_gau1
-
-distance_df_small_edges_gau1 <- distance_df_small_edges_gau1 |>
-  mutate(ID = row_number()) |>
-  dplyr::filter(!(ID %in% c(150, 110))) |>
-  dplyr::select(-ID)
+  filter(distance < 0.06) #benchmark_gau1
+
+# distance_df_small_edges_gau1 <- distance_df_small_edges_gau1 |>
+#   mutate(ID = row_number()) |>
+#   dplyr::filter(!(ID %in% c(150, 110))) |>
+#   dplyr::select(-ID)
 
@@
 distance_df_small_edges <- distance_gau1 |>
-  dplyr::filter(distance < 0.1) |>
-  mutate(ID = row_number()) |>
-  filter(!(ID %in% c(150, 110))) |>
-  dplyr::select(-ID) #benchmark_gau1
+  dplyr::filter(distance < 0.06) #|>
+  #mutate(ID = row_number()) |>
+  #filter(!(ID %in% c(150, 110))) |>
+  #dplyr::select(-ID) #benchmark_gau1
 
@@
     color = "#000000",
-    #alpha = 0.4,
-    linewidth = 0.5) +
+    alpha = 0.4,
+    linewidth = 0.8) +
   geom_segment(
@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 50
+num_bins_x_gau1 <- 75
 lim1 <- gau1_scaled_obj$lim1
@@
 distance_df_small_edges_gau1 <- distance_gau1 |>
-  filter(distance < 0.05) #|> benchmark_gau1
+  filter(distance < benchmark_gau1) #|> benchmark_gau1
   #mutate(ID = row_number()) #|>
@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 22
+num_bins_x_gau1 <- 33
 lim1 <- gau1_scaled_obj$lim1
@@
 distance_df_small_edges_gau1 <- distance_gau1 |>
-  filter(distance < 0.06) #|> #benchmark_gau1
+  filter(distance < benchmark_gau1) #|> #benchmark_gau1
   #mutate(ID = row_number()) #|>

### 2025-05-05 paper.qmd
Added orange to high error
Started update the paper with the new version of quollr 0.2.6

@@
 
-distance_df_small_edges_two_curvy <- distance_two_curvy |>
-    filter(distance < benchmark_two_curvy)
-
-## Computed benchmark value
-tr_from_to_df_two_curvy <- inner_join(
-  tr_from_to_df_two_curvy, distance_df_small_edges_two_curvy,
-  by = c("from", "to"))
-
 trimesh_removed_two_curvy <- ggplot() +
@@
 #| label: langevitour-best-fit-two-curvy-proj-tsne
-#| eval: false
 
@@
 langevitour::langevitour(df_model_data_two_curvy_filtered[1:(length(df_model_data_two_curvy_filtered)-1)],
-                         lineFrom = distance_df_small_edges_two_curvy$from,
-                         lineTo = distance_df_small_edges_two_curvy$to,
+                         lineFrom = tr_from_to_df_two_curvy$from,
+                         lineTo = tr_from_to_df_two_curvy$to,
                          group = factor(df_model_data_two_curvy_filtered$type,
@@
 #| label: best-fit-two-curvy-proj-tsne
+
 ## First projection
@@
                scaled_data_model = scaled_two_curvy_data_model, 
-               distance_df_small_edges = distance_df_small_edges_two_curvy, 
+               distance_df_small_edges = tr_from_to_df_two_curvy, 
                axis_param = list(limits = 0.5, 
@@
                             scaled_data_model = scaled_two_curvy_data_model, 
-                            distance_df_small_edges = distance_df_small_edges_two_curvy, 
+                            distance_df_small_edges = tr_from_to_df_two_curvy, 
                             axis_param = list(limits = 0.35, 
@@
 #| fig-height: 5
+#| eval: false
 
@@
 
-distance_two_curvy2 <- algo_obj_two_curvy2$distance_df
 tr_from_to_df_two_curvy2 <- algo_obj_two_curvy2$tr_from_to_df
-benchmark_two_curvy2 <- algo_obj_two_curvy2$benchmark
 df_bin_centroids_two_curvy2 <- algo_obj_two_curvy2$df_bin_centroids
@@
 hex_grid_with_counts_two_curvy2 <- algo_obj_two_curvy2$hex_grid_with_counts
-
-distance_df_small_edges_two_curvy2 <- distance_two_curvy2 |>
-  filter(distance < benchmark_two_curvy2)
-
-tr_from_to_df_two_curvy2 <- inner_join(
-  tr_from_to_df_two_curvy2, distance_df_small_edges_two_curvy2, 
-  by = c("from", "to"))
 
@@
 
-distance_two_curvy3 <- algo_obj_two_curvy3$distance_df
 tr_from_to_df_two_curvy3 <- algo_obj_two_curvy3$tr_from_to_df
-benchmark_two_curvy3 <- algo_obj_two_curvy3$benchmark
 df_bin_centroids_two_curvy3 <- algo_obj_two_curvy3$df_bin_centroids
@@
 hex_grid_with_counts_two_curvy3 <- algo_obj_two_curvy3$hex_grid_with_counts
-
-distance_df_small_edges_two_curvy3 <- distance_two_curvy3 |>
-  filter(distance < benchmark_two_curvy3)
-
-tr_from_to_df_two_curvy3 <- inner_join(
-  tr_from_to_df_two_curvy3, distance_df_small_edges_two_curvy3, 
-  by = c("from", "to"))
 
@@
   trimesh_two_curvy_removed1_tsne_with_data +
-  two_curvy_proj_tsne_first_model1 +
-  two_curvy_proj_tsne_all_model1 +
+  #two_curvy_proj_tsne_first_model1 +
+  #two_curvy_proj_tsne_all_model1 +
   plot_layout(ncol = 3)
@@
 
-#### Removing long edges
-
-Edges define the neighbourhood structure, in order to provide a smooth \gD{} representation of the fitted model.
+<!-- #### Removing long edges -->
+
+<!-- Edges define the neighbourhood structure, in order to provide a smooth \gD{} representation of the fitted model.
 @fig-two-curvy-true-proj shows a wire frame of the true model that was used to generate the non-linear clusters example data.
 The ideal is that the representation of the fitted model, at least for this example where we know the true model, should look similar to this.
-
-The Delaunay triangulation will ensure that all centroids are connected into a triangular mesh.
+-->
+
+<!-- The Delaunay triangulation will ensure that all centroids are connected into a triangular mesh.
 For some structures, like clustered data, or highly non-linear shapes, breaks in the mesh are meaningful.
@@
 For these reasons it is necessary to remove edges from the mesh in some applications.
-
-The decision on edge length removal is made based on the distribution of edge lengths.
+-->
+
+<!-- The decision on edge length removal is made based on the distribution of edge lengths.
 In particular, a gap between values, where there a concentration of small values and then a few larger values, likely suggests a cutoff for edge removal.
@@
 @fig-rm-lg illustrates edge length distributions.
-
-There is an additional step that is needed.
+-->
+
+<!-- There is an additional step that is needed.
 When the model is lifted into \pD{}, if the fit is good all the edges should be relatively small in this space, too.
 If this is not the case, then there are several possible actions: (1) re-do the NLDR to get a more representative layout; (2) identify the edge and remove it from the model, in \gD{} and \pD{}; (3) consider different values for the model fit, number of bins, initial bin position or removing low density bins.
-
-```{r}
-#| label: model-two-curvy-umap
-
-umap_two_curvy <- read_rds(file = "data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_umap_n-neigbors_15_min-dist_0.1.rds")
-
-num_bins_x_two_curvy <- 42
-
-algo_obj_two_curvy3 <- gen_nldr_vis_algo_obj(
-  high_d_data = training_data_two_curvy, 
-  nldr_data = umap_two_curvy, 
-  num_x_bins = num_bins_x_two_curvy)
-
-umap_two_curvy_scaled <- algo_obj_two_curvy3$nldr_scaled
-distance_two_curvy <- algo_obj_two_curvy3$distance_df
-tr_from_to_df_two_curvy <- algo_obj_two_curvy3$tr_from_to_df
-benchmark_two_curvy <- algo_obj_two_curvy3$benchmark
-df_bin_centroids_two_curvy <- algo_obj_two_curvy3$df_bin_centroids
-df_bin_two_curvy <- algo_obj_two_curvy3$df_bin
-hex_grid_with_counts <- algo_obj_two_curvy3$hex_grid_with_counts
-a1 <- algo_obj_two_curvy3$a1
-
-trimesh_two_curvy_umap <- ggplot() +
-  geom_segment(data = tr_from_to_df_two_curvy,
-               aes(
-                 x = x_from,
-                 y = y_from,
-                 xend = x_to,
-                 yend = y_to),
-               colour = "#000000",
-               alpha = 0.5,
-               linewidth = 0.5) +
-  geom_point(data = umap_two_curvy_scaled,
-             aes(
-               x = UMAP1,
-               y = UMAP2
-             ),
-             color = "#636363",
-             alpha = 0.5,
-             size = 0.5
-  ) +
-  coord_equal() +
-  xlim(c(0, 1)) +
-  ylim(c(0, 1)) +
-  interior_annotation("a1", sc_ltr_pos, cex = 2)
-
-```
-
-```{r}
-
-# With the computed benchmark value
-# distance_df_small_edges_two_curvy <- distance_two_curvy |>
-#   filter(distance < benchmark_two_curvy)
-
-####With diff benchmark values
-
-## Benchmark 2
-
-benchmark_two_curvy <- 8 * a1
-
-distance_df_small_edges_two_curvy3 <- distance_two_curvy |>
-  filter(distance < benchmark_two_curvy)
-
-tr_from_to_df_two_curvy4 <- inner_join(
-  tr_from_to_df_two_curvy, distance_df_small_edges_two_curvy3,
-  by = c("from", "to"))
-
-trimesh_two_curvy_removed2_umap <- ggplot() +
-  geom_segment(data = tr_from_to_df_two_curvy4,
-               aes(
-                 x = x_from,
-                 y = y_from,
-                 xend = x_to,
-                 yend = y_to),
-               colour = "#000000",
-               alpha = 0.5,
-               linewidth = 0.5)+
-  geom_point(data = umap_two_curvy_scaled,
-             aes(
-               x = UMAP1,
-               y = UMAP2
-             ),
-             color = "#636363",
-             alpha = 0.5,
-             size = 0.5
-  ) +
-  coord_equal() +
-  xlim(c(0, 1)) +
-  ylim(c(0, 1)) +
-  interior_annotation("b1", sc_ltr_pos, cex = 2)
-
-## Benchmark 1
-
-benchmark_two_curvy <- 2.5 * bin_width
-
-distance_df_small_edges_two_curvy2 <- distance_two_curvy |>
-  filter(distance < benchmark_two_curvy) ## 0.231
-
-tr_from_to_df_two_curvy3 <- inner_join(
-  tr_from_to_df_two_curvy, distance_df_small_edges_two_curvy2,
-  by = c("from", "to"))
-
-trimesh_two_curvy_removed1_umap <- ggplot() +
-  geom_segment(data = tr_from_to_df_two_curvy3,
-               aes(
-                 x = x_from,
-                 y = y_from,
-                 xend = x_to,
-                 yend = y_to),
-               colour = "#000000",
-               alpha = 0.5,
-               linewidth = 0.5)+
-  geom_point(data = umap_two_curvy_scaled,
-             aes(
-               x = UMAP1,
-               y = UMAP2
-             ),
-             color = "#636363",
-             alpha = 0.5,
-             size = 0.5
-  ) +
-  coord_equal() +
-  xlim(c(0, 1)) +
-  ylim(c(0, 1)) +
-  interior_annotation("c1", sc_ltr_pos, cex = 2)
-```
-
-```{r}
-#| label: compare-2d-and-4d-dist
-
-### Compute highD distance
-dist_vec <- proxy::dist(x = df_bin_two_curvy[, -1], method = "Euclidean") |> as.vector()
-
-from_vec <- c()
-to_vec <- c()
-num_obs <- 1:(NROW(df_bin_two_curvy) - 1)
-
-for (obs in num_obs) {
-
-  from_val <- rep(obs, (NROW(df_bin_two_curvy) - obs))
-  if ((obs + 1) <= NROW(df_bin_two_curvy)) {
-    to_val <- (obs + 1):NROW(df_bin_two_curvy)
-  }
-  from_vec <- append(from_vec, from_val)
-  to_vec <- append(to_vec, to_val)
-
-}
-
-dist_highd <- tibble::tibble(from = from_vec, to = to_vec, dist_highd = dist_vec)
-
-## To plot the distribution of distances
-
-# Define bin width and starting point
-start_point <- a1/2
-
-# Create bins and calculate the mean value for each bin range
-bins <- seq(start_point, max(distance_two_curvy$distance) + a1,
-            by = a1)
-bin_labels <- bins[-length(bins)] + a1 / 2  # mean of each bin range
-
-distance_two_curvy_dist <- distance_two_curvy |>
-  mutate(bin = cut(distance, breaks = bins, include.lowest = TRUE, labels = bin_labels))
-
-distance_two_curvy_dist <- left_join(distance_two_curvy_dist, dist_highd, by = c("from", "to"))
-
-text_df <- tibble::tibble(
-  x = c(0.12, 0.43),
-  y = c(10, 10),
-  text = c("c", "b")
-)
-
-distance_points_umap <- ggplot(
-  distance_two_curvy_dist,
-  aes(x = distance,
-      y = dist_highd)) +
-  geom_point(alpha = 0.5) +
-  geom_text(
-    data=text_df,
-    aes(x=x, y=y,
-        label=text),
-    colour="#bdbdbd",
-    size = 10) +
-  geom_vline(xintercept = 8 * bin_width,
-             linetype="solid",
-             color = "#bdbdbd",
-             linewidth=1) +
-  geom_vline(xintercept = 2.5 * bin_width,
-             linetype="solid",
-             color = "#bdbdbd",
-             linewidth=1) +
-  ylab(expression(d^{(4)})) +
-  xlab(expression(d^{(2)})) +
-  theme_minimal() +
-  theme(aspect.ratio = 0.75,
-        panel.border = element_rect(fill = 'transparent'),
-        plot.title = element_text(size = 20, hjust = 0.5, vjust = -0.5),
-        panel.grid.major.x = element_blank(),
-        axis.ticks.x = element_line(),
-        axis.ticks.y = element_line(),
-        axis.title = element_text(size = 15),
-        axis.text = element_text(size = 12))
-```
-
-```{r}
-#| label: proj-tsne-model
-
-df_bin_two_curvy <- df_bin_two_curvy |>
-  select(-hb_id) |>
-  mutate(type = "model")
-
-# Apply the scaling
-df_model_data_two_curvy <- bind_rows(data_two_curvy, df_bin_two_curvy)
-
-scaled_two_curvy <- scale_data_manual(df_model_data_two_curvy, "type") |>
-  as_tibble()
-
-scaled_two_curvy_data <- scaled_two_curvy |>
-  filter(type == "data") |>
-  select(-type)
-
-scaled_two_curvy_data_model <- scaled_two_curvy |>
-  filter(type == "model") |>
-  select(-type)
-
-## First projection
-projection <- cbind(
-  c(0.37847,-0.06517,0.04231,0.01655),
-  c(-0.06466,-0.27052,0.22131,-0.15235))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1, 
-                            scaled_data = scaled_two_curvy_data, 
-                            scaled_data_model = scaled_two_curvy_data_model, 
-                            distance_df_small_edges = distance_df_small_edges_two_curvy2, 
-                            axis_param = list(limits = 0.6, 
-                                              axis_scaled = 2,
-                                              axis_pos_x = -0.5, 
-                                              axis_pos_y = -0.5, 
-                                              threshold = 0.04))
-
-## Projection with Delauany triangulation
-
-projected_model <- as.matrix(scaled_two_curvy_data_model) %*% projection_scaled
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df1 <- dplyr::left_join(
-  distance_two_curvy |> select(-distance),
-  projected_model_df,
-  by = c("from" = "ID"))
-
-names(model_df1)[3:NCOL(model_df1)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df1 <- dplyr::left_join(model_df1, projected_model_df, by = c("to" = "ID"))
-names(model_df1)[(2 + NCOL(projected_model_df)):NCOL(model_df1)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-## With benchmark option 2
-
-model_df2 <- dplyr::left_join(
-  distance_df_small_edges_two_curvy3 |> select(-distance),
-  projected_model_df,
-  by = c("from" = "ID"))
-
-names(model_df2)[3:NCOL(model_df2)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df2 <- dplyr::left_join(model_df2, projected_model_df, by = c("to" = "ID"))
-names(model_df2)[(2 + NCOL(projected_model_df)):NCOL(model_df2)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-two_curvy_proj_first_model1_umap <- plot_proj(
-    proj_obj = proj_obj1, 
-    point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
-    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
-    plot_limits = c(-0.6, 0.6), 
-    title = "c2", 
-    cex = 2, 
-    position = sc_ltr_pos,
-    axis_text_size = 4, 
-    is_color = FALSE) +
-    theme(
-        legend.position = "none"
-    )
-
-## Projection with Delauany triangulation
-
-proj_obj1_temp1 <- proj_obj1
-proj_obj1_temp2 <- proj_obj1
-proj_obj1_temp1[["model_df"]] <- model_df1
-
-two_curvy_proj_model_delaunay_umap <- plot_proj(
-    proj_obj = proj_obj1_temp1, 
-    point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
-    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
-    plot_limits = c(-0.6, 0.6), 
-    title = "a2", 
-    cex = 2, 
-    position = sc_ltr_pos,
-    axis_text_size = 4, 
-    is_color = FALSE) +
-    theme(aspect.ratio = 1, 
-          legend.position = "none")
-
-proj_obj1_temp2[["model_df"]] <- model_df2
-
-two_curvy_proj_model_benchmark2_umap <- plot_proj(
-    proj_obj = proj_obj1_temp2, 
-    point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
-    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
-    plot_limits = c(-0.6, 0.6), 
-    title = "b2", 
-    cex = 2, 
-    position = sc_ltr_pos,
-    axis_text_size = 4, 
-    is_color = FALSE) +
-    theme(aspect.ratio = 1, 
-          legend.position = "none")
-
-```
-
-```{r}
-#| label: fig-rm-lg
-#| fig-cap: "The choice of long edge removal significantly impacts the model's accuracy.
+-->
+
+<!-- ```{r} -->
+<!-- #| label: model-two-curvy-umap -->
+
+<!-- umap_two_curvy <- read_rds(file = "data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_umap_n-neigbors_15_min-dist_0.1.rds") -->
+
+<!-- num_bins_x_two_curvy <- 42 -->
+
+<!-- algo_obj_two_curvy3 <- gen_nldr_vis_algo_obj( -->
+<!--   high_d_data = training_data_two_curvy,  -->
+<!--   nldr_data = umap_two_curvy,  -->
+<!--   num_x_bins = num_bins_x_two_curvy) -->
+
+<!-- umap_two_curvy_scaled <- algo_obj_two_curvy3$nldr_scaled -->
+<!-- distance_two_curvy <- algo_obj_two_curvy3$distance_df -->
+<!-- tr_from_to_df_two_curvy <- algo_obj_two_curvy3$tr_from_to_df -->
+<!-- benchmark_two_curvy <- algo_obj_two_curvy3$benchmark -->
+<!-- df_bin_centroids_two_curvy <- algo_obj_two_curvy3$df_bin_centroids -->
+<!-- df_bin_two_curvy <- algo_obj_two_curvy3$df_bin -->
+<!-- hex_grid_with_counts <- algo_obj_two_curvy3$hex_grid_with_counts -->
+<!-- a1 <- algo_obj_two_curvy3$a1 -->
+
+<!-- trimesh_two_curvy_umap <- ggplot() + -->
+<!--   geom_segment(data = tr_from_to_df_two_curvy, -->
+<!--                aes( -->
+<!--                  x = x_from, -->
+<!--                  y = y_from, -->
+<!--                  xend = x_to, -->
+<!--                  yend = y_to), -->
+<!--                colour = "#000000", -->
+<!--                alpha = 0.5, -->
+<!--                linewidth = 0.5) + -->
+<!--   geom_point(data = umap_two_curvy_scaled, -->
+<!--              aes( -->
+<!--                x = UMAP1, -->
+<!--                y = UMAP2 -->
+<!--              ), -->
+<!--              color = "#636363", -->
+<!--              alpha = 0.5, -->
+<!--              size = 0.5 -->
+<!--   ) + -->
+<!--   coord_equal() + -->
+<!--   xlim(c(0, 1)) + -->
+<!--   ylim(c(0, 1)) + -->
+<!--   interior_annotation("a1", sc_ltr_pos, cex = 2) -->
+
+<!-- ``` -->
+
+<!-- ```{r} -->
+
+<!-- # With the computed benchmark value -->
+<!-- # distance_df_small_edges_two_curvy <- distance_two_curvy |> -->
+<!-- #   filter(distance < benchmark_two_curvy) -->
+
+<!-- ####With diff benchmark values -->
+
+<!-- ## Benchmark 2 -->
+
+<!-- benchmark_two_curvy <- 8 * a1 -->
+
+<!-- distance_df_small_edges_two_curvy3 <- distance_two_curvy |> -->
+<!--   filter(distance < benchmark_two_curvy) -->
+
+<!-- tr_from_to_df_two_curvy4 <- inner_join( -->
+<!--   tr_from_to_df_two_curvy, distance_df_small_edges_two_curvy3, -->
+<!--   by = c("from", "to")) -->
+
+<!-- trimesh_two_curvy_removed2_umap <- ggplot() + -->
+<!--   geom_segment(data = tr_from_to_df_two_curvy4, -->
+<!--                aes( -->
+<!--                  x = x_from, -->
+<!--                  y = y_from, -->
+<!--                  xend = x_to, -->
+<!--                  yend = y_to), -->
+<!--                colour = "#000000", -->
+<!--                alpha = 0.5, -->
+<!--                linewidth = 0.5)+ -->
+<!--   geom_point(data = umap_two_curvy_scaled, -->
+<!--              aes( -->
+<!--                x = UMAP1, -->
+<!--                y = UMAP2 -->
+<!--              ), -->
+<!--              color = "#636363", -->
+<!--              alpha = 0.5, -->
+<!--              size = 0.5 -->
+<!--   ) + -->
+<!--   coord_equal() + -->
+<!--   xlim(c(0, 1)) + -->
+<!--   ylim(c(0, 1)) + -->
+<!--   interior_annotation("b1", sc_ltr_pos, cex = 2) -->
+
+<!-- ## Benchmark 1 -->
+
+<!-- benchmark_two_curvy <- 2.5 * bin_width -->
+
+<!-- distance_df_small_edges_two_curvy2 <- distance_two_curvy |> -->
+<!--   filter(distance < benchmark_two_curvy) ## 0.231 -->
+
+<!-- tr_from_to_df_two_curvy3 <- inner_join( -->
+<!--   tr_from_to_df_two_curvy, distance_df_small_edges_two_curvy2, -->
+<!--   by = c("from", "to")) -->
+
+<!-- trimesh_two_curvy_removed1_umap <- ggplot() + -->
+<!--   geom_segment(data = tr_from_to_df_two_curvy3, -->
+<!--                aes( -->
+<!--                  x = x_from, -->
+<!--                  y = y_from, -->
+<!--                  xend = x_to, -->
+<!--                  yend = y_to), -->
+<!--                colour = "#000000", -->
+<!--                alpha = 0.5, -->
+<!--                linewidth = 0.5)+ -->
+<!--   geom_point(data = umap_two_curvy_scaled, -->
+<!--              aes( -->
+<!--                x = UMAP1, -->
+<!--                y = UMAP2 -->
+<!--              ), -->
+<!--              color = "#636363", -->
+<!--              alpha = 0.5, -->
+<!--              size = 0.5 -->
+<!--   ) + -->
+<!--   coord_equal() + -->
+<!--   xlim(c(0, 1)) + -->
+<!--   ylim(c(0, 1)) + -->
+<!--   interior_annotation("c1", sc_ltr_pos, cex = 2) -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- #| label: compare-2d-and-4d-dist -->
+
+<!-- ### Compute highD distance -->
+<!-- dist_vec <- proxy::dist(x = df_bin_two_curvy[, -1], method = "Euclidean") |> as.vector() -->
+
+<!-- from_vec <- c() -->
+<!-- to_vec <- c() -->
+<!-- num_obs <- 1:(NROW(df_bin_two_curvy) - 1) -->
+
+<!-- for (obs in num_obs) { -->
+
+<!--   from_val <- rep(obs, (NROW(df_bin_two_curvy) - obs)) -->
+<!--   if ((obs + 1) <= NROW(df_bin_two_curvy)) { -->
+<!--     to_val <- (obs + 1):NROW(df_bin_two_curvy) -->
+<!--   } -->
+<!--   from_vec <- append(from_vec, from_val) -->
+<!--   to_vec <- append(to_vec, to_val) -->
+
+<!-- } -->
+
+<!-- dist_highd <- tibble::tibble(from = from_vec, to = to_vec, dist_highd = dist_vec) -->
+
+<!-- ## To plot the distribution of distances -->
+
+<!-- # Define bin width and starting point -->
+<!-- start_point <- a1/2 -->
+
+<!-- # Create bins and calculate the mean value for each bin range -->
+<!-- bins <- seq(start_point, max(distance_two_curvy$distance) + a1, -->
+<!--             by = a1) -->
+<!-- bin_labels <- bins[-length(bins)] + a1 / 2  # mean of each bin range -->
+
+<!-- distance_two_curvy_dist <- distance_two_curvy |> -->
+<!--   mutate(bin = cut(distance, breaks = bins, include.lowest = TRUE, labels = bin_labels)) -->
+
+<!-- distance_two_curvy_dist <- left_join(distance_two_curvy_dist, dist_highd, by = c("from", "to")) -->
+
+<!-- text_df <- tibble::tibble( -->
+<!--   x = c(0.12, 0.43), -->
+<!--   y = c(10, 10), -->
+<!--   text = c("c", "b") -->
+<!-- ) -->
+
+<!-- distance_points_umap <- ggplot( -->
+<!--   distance_two_curvy_dist, -->
+<!--   aes(x = distance, -->
+<!--       y = dist_highd)) + -->
+<!--   geom_point(alpha = 0.5) + -->
+<!--   geom_text( -->
+<!--     data=text_df, -->
+<!--     aes(x=x, y=y, -->
+<!--         label=text), -->
+<!--     colour="#bdbdbd", -->
+<!--     size = 10) + -->
+<!--   geom_vline(xintercept = 8 * bin_width, -->
+<!--              linetype="solid", -->
+<!--              color = "#bdbdbd", -->
+<!--              linewidth=1) + -->
+<!--   geom_vline(xintercept = 2.5 * bin_width, -->
+<!--              linetype="solid", -->
+<!--              color = "#bdbdbd", -->
+<!--              linewidth=1) + -->
+<!--   ylab(expression(d^{(4)})) + -->
+<!--   xlab(expression(d^{(2)})) + -->
+<!--   theme_minimal() + -->
+<!--   theme(aspect.ratio = 0.75, -->
+<!--         panel.border = element_rect(fill = 'transparent'), -->
+<!--         plot.title = element_text(size = 20, hjust = 0.5, vjust = -0.5), -->
+<!--         panel.grid.major.x = element_blank(), -->
+<!--         axis.ticks.x = element_line(), -->
+<!--         axis.ticks.y = element_line(), -->
+<!--         axis.title = element_text(size = 15), -->
+<!--         axis.text = element_text(size = 12)) -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- #| label: proj-tsne-model -->
+
+<!-- df_bin_two_curvy <- df_bin_two_curvy |> -->
+<!--   select(-hb_id) |> -->
+<!--   mutate(type = "model") -->
+
+<!-- # Apply the scaling -->
+<!-- df_model_data_two_curvy <- bind_rows(data_two_curvy, df_bin_two_curvy) -->
+
+<!-- scaled_two_curvy <- scale_data_manual(df_model_data_two_curvy, "type") |> -->
+<!--   as_tibble() -->
+
+<!-- scaled_two_curvy_data <- scaled_two_curvy |> -->
+<!--   filter(type == "data") |> -->
+<!--   select(-type) -->
+
+<!-- scaled_two_curvy_data_model <- scaled_two_curvy |> -->
+<!--   filter(type == "model") |> -->
+<!--   select(-type) -->
+
+<!-- ## First projection -->
+<!-- projection <- cbind( -->
+<!--   c(0.37847,-0.06517,0.04231,0.01655), -->
+<!--   c(-0.06466,-0.27052,0.22131,-0.15235)) -->
+
+<!-- proj_obj1 <- get_projection(projection = projection,  -->
+<!--                             proj_scale = 1,  -->
+<!--                             scaled_data = scaled_two_curvy_data,  -->
+<!--                             scaled_data_model = scaled_two_curvy_data_model,  -->
+<!--                             distance_df_small_edges = distance_df_small_edges_two_curvy2,  -->
+<!--                             axis_param = list(limits = 0.6,  -->
+<!--                                               axis_scaled = 2, -->
+<!--                                               axis_pos_x = -0.5,  -->
+<!--                                               axis_pos_y = -0.5,  -->
+<!--                                               threshold = 0.04)) -->
+
+<!-- ## Projection with Delauany triangulation -->
+
+<!-- projected_model <- as.matrix(scaled_two_curvy_data_model) %*% projection_scaled -->
+
+<!-- projected_model_df <- projected_model |> -->
+<!--   tibble::as_tibble(.name_repair = "unique") |> -->
+<!--   dplyr::rename(c("proj1" = "...1", -->
+<!--                   "proj2" = "...2")) |> -->
+<!--   dplyr::mutate(ID = dplyr::row_number()) -->
+
+<!-- model_df1 <- dplyr::left_join( -->
+<!--   distance_two_curvy |> select(-distance), -->
+<!--   projected_model_df, -->
+<!--   by = c("from" = "ID")) -->
+
+<!-- names(model_df1)[3:NCOL(model_df1)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from") -->
+
+<!-- model_df1 <- dplyr::left_join(model_df1, projected_model_df, by = c("to" = "ID")) -->
+<!-- names(model_df1)[(2 + NCOL(projected_model_df)):NCOL(model_df1)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to") -->
+
+<!-- ## With benchmark option 2 -->
+
+<!-- model_df2 <- dplyr::left_join( -->
+<!--   distance_df_small_edges_two_curvy3 |> select(-distance), -->
+<!--   projected_model_df, -->
+<!--   by = c("from" = "ID")) -->
+
+<!-- names(model_df2)[3:NCOL(model_df2)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from") -->
+
+<!-- model_df2 <- dplyr::left_join(model_df2, projected_model_df, by = c("to" = "ID")) -->
+<!-- names(model_df2)[(2 + NCOL(projected_model_df)):NCOL(model_df2)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to") -->
+
+<!-- two_curvy_proj_first_model1_umap <- plot_proj( -->
+<!--     proj_obj = proj_obj1,  -->
+<!--     point_param = c(0.5, 0.2, "#636363"), # size, alpha, color -->
+<!--     line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha -->
+<!--     plot_limits = c(-0.6, 0.6),  -->
+<!--     title = "c2",  -->
+<!--     cex = 2,  -->
+<!--     position = sc_ltr_pos, -->
+<!--     axis_text_size = 4,  -->
+<!--     is_color = FALSE) + -->
+<!--     theme( -->
+<!--         legend.position = "none" -->
+<!--     ) -->
+
+<!-- ## Projection with Delauany triangulation -->
+
+<!-- proj_obj1_temp1 <- proj_obj1 -->
+<!-- proj_obj1_temp2 <- proj_obj1 -->
+<!-- proj_obj1_temp1[["model_df"]] <- model_df1 -->
+
+<!-- two_curvy_proj_model_delaunay_umap <- plot_proj( -->
+<!--     proj_obj = proj_obj1_temp1,  -->
+<!--     point_param = c(0.5, 0.2, "#636363"), # size, alpha, color -->
+<!--     line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha -->
+<!--     plot_limits = c(-0.6, 0.6),  -->
+<!--     title = "a2",  -->
+<!--     cex = 2,  -->
+<!--     position = sc_ltr_pos, -->
+<!--     axis_text_size = 4,  -->
+<!--     is_color = FALSE) + -->
+<!--     theme(aspect.ratio = 1,  -->
+<!--           legend.position = "none") -->
+
+<!-- proj_obj1_temp2[["model_df"]] <- model_df2 -->
+
+<!-- two_curvy_proj_model_benchmark2_umap <- plot_proj( -->
+<!--     proj_obj = proj_obj1_temp2,  -->
+<!--     point_param = c(0.5, 0.2, "#636363"), # size, alpha, color -->
+<!--     line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha -->
+<!--     plot_limits = c(-0.6, 0.6),  -->
+<!--     title = "b2",  -->
+<!--     cex = 2,  -->
+<!--     position = sc_ltr_pos, -->
+<!--     axis_text_size = 4,  -->
+<!--     is_color = FALSE) + -->
+<!--     theme(aspect.ratio = 1,  -->
+<!--           legend.position = "none") -->
+
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- #| label: fig-rm-lg -->
+<!-- #| fig-cap: "The choice of long edge removal significantly impacts the model's accuracy.
 \\gD{} Vs $7\\text{-}D$ distances between the bin centroids, where the model is fitted with UMAP for the non-linear clusters data.
@@
 In contrast, using 2$a_1$ is a better option, as it effectively captures the non-linear shapes of the clusters.
-To generate an accurate model that captures the data structure, the choice of long edge removal does matter."
-#| fig-height: 15
-#| fig-width: 15
-
-free(distance_points_umap) + 
-    wrap_plots(trimesh_two_curvy_umap, trimesh_two_curvy_removed2_umap,
-               trimesh_two_curvy_removed1_umap, two_curvy_proj_model_delaunay_umap,
-               two_curvy_proj_model_benchmark2_umap, two_curvy_proj_first_model1_umap,
-               ncol = 3) +
-    plot_layout(heights = c(0.8, 1))
-```
+To generate an accurate model that captures the data structure, the choice of long edge removal does matter." -->
+<!-- #| fig-height: 15 -->
+<!-- #| fig-width: 15 -->
+
+<!-- free(distance_points_umap) +  -->
+<!--     wrap_plots(trimesh_two_curvy_umap, trimesh_two_curvy_removed2_umap, -->
+<!--                trimesh_two_curvy_removed1_umap, two_curvy_proj_model_delaunay_umap, -->
+<!--                two_curvy_proj_model_benchmark2_umap, two_curvy_proj_first_model1_umap, -->
+<!--                ncol = 3) + -->
+<!--     plot_layout(heights = c(0.8, 1)) -->
+<!-- ``` -->
 
@@
   tri_object = tr1_object_two_curvy4)
-
-## Compute 2D distances
-distance_two_curvy4 <- cal_2d_dist(
-  tr_coord_df = tr_from_to_df_two_curvy4, 
-  start_x = "x_from", 
-  start_y = "y_from", 
-  end_x = "x_to", 
-  end_y = "y_to", 
-  select_vars = c("from", "to", "distance"))
-
-## To find the benchmark value
-benchmark_two_curvy4 <- find_lg_benchmark(
-  distance_edges = distance_two_curvy4, 
-  distance_col = "distance")
-
-benchmark_two_curvy4 <- 0.2
-
-## Benchmark 1
-
-distance_df_small_edges_two_curvy4 <- distance_two_curvy4 |>
-  filter(distance < benchmark_two_curvy4) |> ## 0.221
-  filter(!(row_number() %in% c(35, 89, 83, 25, 80, 56, 13, 59, 16)))
-
-tr_from_to_df_two_curvy4 <- inner_join(
-  tr_from_to_df_two_curvy4, distance_df_small_edges_two_curvy4,
-  by = c("from", "to"))
 
@@
 model_df <- dplyr::left_join(
-  distance_df_small_edges_two_curvy4 |> select(-distance), 
+  tr_from_to_df_two_curvy4, 
   projected_model_df, 
@@
 
-## Compute 2D distances
-distance_gau1 <- cal_2d_dist(
-  tr_coord_df = tr_from_to_df_gau1,
-  start_x = "x_from",
-  start_y = "y_from",
-  end_x = "x_to",
-  end_y = "y_to",
-  select_vars = c("from", "to", "distance"))
-
-## To find the benchmark value
-benchmark_gau1 <- find_lg_benchmark(
-  distance_edges = distance_gau1,
-  distance_col = "distance")
-
 tr_df <- distinct(tibble(
@@
   y = c(tr_from_to_df_gau1[["y_from"]], tr_from_to_df_gau1[["y_to"]])))
-
-distance_df_small_edges_gau1 <- distance_gau1 |>
-  filter(distance < 0.06) #benchmark_gau1
-
-# distance_df_small_edges_gau1 <- distance_df_small_edges_gau1 |>
-#   mutate(ID = row_number()) |>
-#   dplyr::filter(!(ID %in% c(150, 110))) |>
-#   dplyr::select(-ID)
-
-tr_from_to_df_gau1 <- right_join(
-  tr_from_to_df_gau1, distance_df_small_edges_gau1, 
-  by = c("from", "to"))
 
@@
   select(-type)
-
-## Set the maximum difference as the criteria
-distance_df_small_edges <- distance_gau1 |>
-  dplyr::filter(distance < 0.06) #|>
-  #mutate(ID = row_number()) |>
-  #filter(!(ID %in% c(150, 110))) |>
-  #dplyr::select(-ID) #benchmark_gau1
-
-# write_rds(distance_df_small_edges, "data/five_gau_clusters/tsne_wireframe.rds")
-# write_rds(df_b, "data/five_gau_clusters/tsne_model.rds")
 ```
@@
 langevitour::langevitour(df_model_data_n[1:(length(df_model_data_n)-1)],
-                         lineFrom = distance_df_small_edges$from,
-                         lineTo = distance_df_small_edges$to,
+                         lineFrom = tr_from_to_df_gau1$from,
+                         lineTo = tr_from_to_df_gau1$to,
                          group = factor(df_model_data_n$type,
@@
                scaled_data_model = scaled_gau_data_model, 
-               distance_df_small_edges = distance_df_small_edges, 
+               distance_df_small_edges = tr_from_to_df_gau1, 
                axis_param = list(limits = 1,
@@
                             scaled_data_model = scaled_gau_data_model, 
-                            distance_df_small_edges = distance_df_small_edges, 
+                            distance_df_small_edges = tr_from_to_df_gau1, 
                             axis_param = list(limits = 1.5,
@@
   y = c(tr_from_to_df_gau1[["y_from"]], tr_from_to_df_gau1[["y_to"]])))
-
-distance_df_small_edges_gau1 <- distance_gau1 |>
-  filter(distance < benchmark_gau1) #|> benchmark_gau1
-  #mutate(ID = row_number()) #|>
-  #filter(!(ID %in% c(112)))
-
-tr_from_to_df_gau1 <- inner_join(
-  tr_from_to_df_gau1, distance_df_small_edges_gau1, 
-  by = c("from", "to"))
 
@@
 
-## Set the maximum difference as the criteria
-distance_df_small_edges <- distance_gau1 |>
-  dplyr::filter(distance < 0.05) #benchmark_gau1
-
 # Apply the scaling
@@
 
-# write_rds(distance_df_small_edges, "data/five_gau_clusters/umap_wireframe.rds")
-# write_rds(df_b, "data/five_gau_clusters/umap_model.rds")
 ```
@@
                scaled_data_model = scaled_gau_data_model, 
-               distance_df_small_edges = distance_df_small_edges, 
+               distance_df_small_edges = tr_from_to_df_gau1, 
                axis_param = list(limits = 1, 
@@
   y = c(tr_from_to_df_gau1[["y_from"]], tr_from_to_df_gau1[["y_to"]])))
-
-distance_df_small_edges_gau1 <- distance_gau1 |>
-  filter(distance < benchmark_gau1) #|> #benchmark_gau1
-  #mutate(ID = row_number()) #|>
-  #filter(!(ID %in% c(97)))
-
-tr_from_to_df_gau1 <- right_join(
-  tr_from_to_df_gau1, distance_df_small_edges_gau1, 
-  by = c("from", "to"))
 
@@
   select(-type)
-
-## Set the maximum difference as the criteria
-distance_df_small_edges <- distance_gau1 |>
-  dplyr::filter(distance < 0.06) #benchmark_gau1 #benchmark_gau1
-
-# write_rds(distance_df_small_edges, "data/five_gau_clusters/pacmap_wireframe.rds")
-# write_rds(df_b, "data/five_gau_clusters/pacmap_model.rds")
 ```
@@
 tsne_one_c_shaped_scaled <- algo_obj_dens_clust$nldr_scaled
-distance_dens_clust <- algo_obj_dens_clust$distance_df
 tr_from_to_df_dens_clust <- algo_obj_dens_clust$tr_from_to_df
-benchmark_dens_clust <- algo_obj_dens_clust$benchmark
 df_bin_centroids_dens_clust <- algo_obj_dens_clust$df_bin_centroids
@@
                       cex = 2)
-
-distance_df_small_edges_c_shaped_structure <- distance_dens_clust |>
-  filter(distance < benchmark_dens_clust)
-
-tr_from_to_df_dens_clust <- inner_join(
-  tr_from_to_df_dens_clust, distance_df_small_edges_c_shaped_structure,
-  by = c("from", "to"))
-
 ```
@@
 langevitour::langevitour(df_model_data_n[1:(length(df_model_data_n)-1)],
-                         lineFrom = distance_df_small_edges_c_shaped_structure$from,
-                         lineTo = distance_df_small_edges_c_shaped_structure$to,
+                         lineFrom = tr_from_to_df_dens_clust$from,
+                         lineTo = tr_from_to_df_dens_clust$to,
                          group = factor(df_model_data_n$type,
@@
                             scaled_data_model = scaled_c_shaped_data_model,
-                            distance_df_small_edges = distance_df_small_edges_c_shaped_structure,
+                            distance_df_small_edges = tr_from_to_df_dens_clust,
                             axis_param = list(limits = 0.4,
@@
 umap_pbmc_scaled <- algo_obj_pbmc$nldr_scaled
-distance_pbmc <- algo_obj_pbmc$distance_df
 tr_from_to_df_pbmc <- algo_obj_pbmc$tr_from_to_df
-benchmark_pbmc <- algo_obj_pbmc$benchmark
 df_bin_centroids_pbmc <- algo_obj_pbmc$df_bin_centroids
 df_bin_pbmc <- algo_obj_pbmc$df_bin
-
-distance_df_small_edges_pbmc <- distance_pbmc |>
-  filter(distance < benchmark_pbmc)#benchmark_pbmc
 
@@
   mutate(cluster = factor(cluster, levels = c("cluster1", "cluster2", "cluster3")))
-
-tr_from_to_df_pbmc <- right_join(
-  tr_from_to_df_pbmc, distance_df_small_edges_pbmc, 
-  by = c("from", "to"))
 
@@
 langevitour::langevitour(df_model_data_pbmc_n[1:(length(df_model_data_pbmc_n)-1)],
-                         lineFrom = distance_df_small_edges_pbmc$from,
-                         lineTo = distance_df_small_edges_pbmc$to,
+                         lineFrom = tr_from_to_df_pbmc$from,
+                         lineTo = tr_from_to_df_pbmc$to,
                          group = factor(df_model_data_pbmc_n$type,
@@
                scaled_data_model = scaled_pbmc_data_model, 
-               distance_df_small_edges = distance_df_small_edges_pbmc, 
+               distance_df_small_edges = tr_from_to_df_pbmc, 
                axis_param = list(limits = 1,
@@
                scaled_data_model = scaled_pbmc_data_model, 
-               distance_df_small_edges = distance_df_small_edges_pbmc, 
+               distance_df_small_edges = tr_from_to_df_pbmc, 
                axis_param = list(limits = 1.35, 
@@
 tsne_minst_scaled <- algo_obj_mnist$nldr_scaled
-distance_mnist <- algo_obj_mnist$distance_df
 tr_from_to_df_mnist <- algo_obj_mnist$tr_from_to_df
-benchmark_mnist <- algo_obj_mnist$benchmark
 df_bin_centroids_mnist <- algo_obj_mnist$df_bin_centroids
@@
 sc_ltr_pos_mnist <- c(0.96, 0.96)
-
-distance_df_small_edges_mnist <- distance_mnist |>
-  filter(distance < benchmark_mnist) #benchmark_mnist
-
-tr_from_to_df_mnist <- right_join(
-  tr_from_to_df_mnist, distance_df_small_edges_mnist, 
-  by = c("from", "to"))
 
@@
 
+hexID_mnist <- df_bin_centroids_mnist |>
+  dplyr::filter(std_counts > 0) |>
+  dplyr::pull(hexID)
+
 df_b_mnist <- df_bin_mnist |>
-  dplyr::filter(hb_id %in% df_bin_centroids_mnist$hexID) |>
+  dplyr::filter(hb_id %in% hexID_mnist) |>
   dplyr::mutate(type = "model") ## Data with summarized mean
@@
 ## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
-df_b_mnist <- df_b_mnist[match(df_bin_centroids_mnist$hexID, df_b_mnist$hb_id),] |>
+df_b_mnist <- df_b_mnist[match(hexID_mnist, df_b_mnist$hb_id),] |>
   dplyr::select(-hb_id)
@@
 langevitour::langevitour(df_model_data_mnist_n[1:(length(df_model_data_mnist_n)-1)],
-                         lineFrom = distance_df_small_edges_mnist$from,
-                         lineTo = distance_df_small_edges_mnist$to,
+                         lineFrom = tr_from_to_df_mnist$from,
+                         lineTo = tr_from_to_df_mnist$to,
                          group = factor(df_model_data_mnist_n$type,
@@
                scaled_data_model = scaled_mnist_data_model, 
-               distance_df_small_edges = distance_df_small_edges_mnist, 
+               distance_df_small_edges = tr_from_to_df_mnist, 
                axis_param = list(limits = 1, 
@@
 
-imge_error_sample_inside <- plot_digit_img(pixels_gathered_inside, palette = "Grays", title_text = "(e)", hjust = 0, ncol = 9)
+imge_error_sample_inside <- plot_digit_img(pixels_gathered_inside, palette = "Oranges", title_text = "(e)", hjust = 0, ncol = 9)
 

### 2025-05-05 paper.qmd
Updated the projection function

@@
 #| label: langevitour-best-fit-two-curvy-proj-tsne
+#| eval: false
 
@@
   trimesh_two_curvy_removed1_tsne_with_data +
-  #two_curvy_proj_tsne_first_model1 +
-  #two_curvy_proj_tsne_all_model1 +
+  two_curvy_proj_tsne_first_model1 +
+  two_curvy_proj_tsne_all_model1 +
   plot_layout(ncol = 3)
@@
 
-names(model_df)[3:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+names(model_df)[7:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
 
 model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(2 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+names(model_df)[(6 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
 
@@
 
-## Compute 2D distances
-distance_two_curvy4 <- cal_2d_dist(
-  tr_coord_df = tr_from_to_df_two_curvy4, 
-  start_x = "x_from", 
-  start_y = "y_from", 
-  end_x = "x_to", 
-  end_y = "y_to", 
-  select_vars = c("from", "to", "distance"))
-
-## To find the benchmark value
-benchmark_two_curvy4 <- find_lg_benchmark(
-  distance_edges = distance_two_curvy4, 
-  distance_col = "distance")
-
-benchmark_two_curvy4 <- 0.2
-
-## Benchmark 1
-
-distance_df_small_edges_two_curvy4 <- distance_two_curvy4 |>
-  filter(distance < benchmark_two_curvy4) |> ## 0.221
-  filter(!(row_number() %in% c(35, 89, 83, 25, 80, 56, 13, 59, 16)))
-
-tr_from_to_df_two_curvy4 <- inner_join(
-  tr_from_to_df_two_curvy4, distance_df_small_edges_two_curvy4,
-  by = c("from", "to"))
-
 trimesh_two_curvy_removed1_pacmap_with_data <- ggplot() +
@@
 tr_from_to_df_pbmc <- algo_obj_pbmc$tr_from_to_df
-benchmark_pbmc <- algo_obj_pbmc$benchmark
 df_bin_centroids_pbmc <- algo_obj_pbmc$df_bin_centroids
 df_bin_pbmc <- algo_obj_pbmc$df_bin
-
-distance_df_small_edges_pbmc <- distance_pbmc |>
-  filter(distance < benchmark_pbmc) #benchmark_pbmc
-
-# tr_from_to_df_pbmc <- tr_from_to_df_pbmc |>
-#   mutate(ID = row_number()) |>
-#   filter(!(ID %in% c(19, 20, 23, 24, 29, 33, 48,
-#                      183, 181, 178, 176))) #19
-
-tr_from_to_df_pbmc <- right_join(
-  tr_from_to_df_pbmc, distance_df_small_edges_pbmc,
-  by = c("from", "to")) #|>
-  #select(-ID)
 
@@
                             scaled_data_model = scaled_pbmc_data_model, 
-                            distance_df_small_edges = distance_df_small_edges_pbmc, 
+                            distance_df_small_edges = tr_from_to_df_pbmc, 
                             axis_param = list(limits = 1, 
@@
                             scaled_data_model = scaled_pbmc_data_model, 
-                            distance_df_small_edges = distance_df_small_edges_pbmc, 
+                            distance_df_small_edges = tr_from_to_df_pbmc, 
                             axis_param = list(limits = 1.35, 

### 2025-05-05 paper.qmd
Removed long edge parameter

@@
     - the total number of bins ($b$), 
-- bin density cutoff, to remove low-density hexagons, and 
-- edge length maximum, remove long edges from \gD{} representation.
+- bin density cutoff, to remove low-density hexagons.
 

### 2025-05-05 paper.qmd
Updated Figure 1 (c)

@@
 tsne_two_curvy_scaled <- algo_obj_two_curvy$nldr_scaled
-distance_two_curvy <- algo_obj_two_curvy$distance_df
 tr_from_to_df_two_curvy <- algo_obj_two_curvy$tr_from_to_df
-benchmark_two_curvy <- algo_obj_two_curvy$benchmark
 df_bin_centroids_two_curvy <- algo_obj_two_curvy$df_bin_centroids
@@
 #| label: empty-bin-two-curvy
+
+df_bin_centroids_two_curvy <- df_bin_centroids_two_curvy |>
+  dplyr::filter(std_counts > 0)
 

### 2025-05-05 paper.qmd
Updated Figure 9

@@
 
-num_bins_x_two_curvy <- 22
+num_bins_x_two_curvy <- 24
 
@@
 tsne_data_gau <- read_rds("data/five_gau_clusters/tsne_data_five_gau_71.rds")
-gau1_scaled_obj <- gen_scaled_data(
-  data = tsne_data_gau)
-tsne_gau_scaled <- gau1_scaled_obj$scaled_nldr
-
-tsne_gau <- tsne_gau_scaled |>
-  ggplot(aes(x = tSNE1,
-             y = tSNE2)) +
-  geom_point(alpha=0.3) +
-  geom_rect(aes(xmin = -0.05, xmax = 0.35, ymin = 0.28, ymax = 0.6),
-               fill = "transparent", color = "grey50", linewidth = 0.8)
 
@@
 num_bins_x_gau1 <- 23
-lim1 <- gau1_scaled_obj$lim1
-lim2 <- gau1_scaled_obj$lim2
-r2_gau1 <- diff(lim2)/diff(lim1)
-
-gau1_model <- fit_highd_model(
-  highd_data = training_data_gau,
-  nldr_data = tsne_gau_scaled,
-  bin1 = num_bins_x_gau1,
-  r2 = r2_gau1,
-  is_bin_centroid = TRUE,
-  q = 0.1
-)
-
-df_bin_centroids_gau1 <- gau1_model$df_bin_centroids
-df_bin_gau1 <- gau1_model$df_bin
-
-## Compute error
-
-error_df_gau_abs <- augment(
-  highd_data = training_data_gau,
-  model_2d = df_bin_centroids_gau1,
-  model_highd = df_bin_gau1)
-
-error_df_gau_abs <- error_df_gau_abs |>
-  mutate(sqrt_row_wise_abs_error = sqrt(row_wise_abs_error))
-
-error_df_gau_abs <- error_df_gau_abs |>
-  bind_cols(tsne_gau_scaled |>
-              select(-ID))
-
-error_df_gau_tsne <- error_df_gau_abs |>
-  ggplot(aes(x = tSNE1,
-             y = tSNE2,
-             z = sqrt_row_wise_abs_error)) +
-  stat_summary_hex(bins=15) +  
-  xlim(c(0.3, 0.66)) + ylim(c(-0.01, 0.35)) +
-  scale_fill_continuous_sequential(palette = "YlOrRd", n_interp = 20) +
-  interior_annotation("a3", 
-                      position = c(0.08, 0.9),
-                      cex = 1.5)
-
-## Triangulate bin centroids
-tr1_object_gau1 <- tri_bin_centroids(
-  df_bin_centroids_gau1, x = "c_x", y = "c_y")
-tr_from_to_df_gau1 <- gen_edges(
-  tri_object = tr1_object_gau1)
-
-tr_df <- distinct(tibble(
-  x = c(tr_from_to_df_gau1[["x_from"]], tr_from_to_df_gau1[["x_to"]]), 
-  y = c(tr_from_to_df_gau1[["y_from"]], tr_from_to_df_gau1[["y_to"]])))
+
+algo_obj_gau1 <- gen_nldr_vis_algo_obj(
+  high_d_data = training_data_gau, 
+  nldr_data = tsne_data_gau, 
+  num_x_bins = num_bins_x_gau1)
+
+tsne_gau_scaled <- algo_obj_gau1$nldr_scaled
+tr_from_to_df_gau1 <- algo_obj_gau1$tr_from_to_df
+df_bin_centroids_gau1 <- algo_obj_gau1$df_bin_centroids
+df_bin_gau1 <- algo_obj_gau1$df_bin
 
@@
 
-## Hexagonal binning to have regular hexagons
-hb_obj_gau1 <- hex_binning(
-  data = tsne_gau_scaled,
-  bin1 = num_bins_x_gau1,
-  r2 = r2_gau1,
-  q = 0.1)
-
-tsne_data_with_hb_id <- hb_obj_gau1$data_hb_id
-
-df_all_gau1 <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID),
-                                  tsne_data_with_hb_id)
-
-### Define type column
-df <- df_all_gau1 |>
-  dplyr::select(tidyselect::starts_with("x")) |>
-  dplyr::mutate(type = "data") ## original dataset
+df_bin_centroids_gau1 <- df_bin_centroids_gau1 |>
+  dplyr::filter(std_counts > 0)
 
@@
   select(-type)
+
 ```
@@
 umap_data_gau <- read_rds("data/five_gau_clusters/umap_data_five_gau.rds")
-gau1_scaled_obj <- gen_scaled_data(
-  data = umap_data_gau)
-umap_gau_scaled <- gau1_scaled_obj$scaled_nldr
-
-umap_gau <- umap_gau_scaled |>
-  ggplot(aes(x = UMAP1,
-             y = UMAP2)) +
-  geom_point(alpha=0.3, color = "#000000")
 
@@
 num_bins_x_gau1 <- 75
-lim1 <- gau1_scaled_obj$lim1
-lim2 <- gau1_scaled_obj$lim2
-r2_gau1 <- diff(lim2)/diff(lim1)
-
-gau1_model <- fit_highd_model(
-  highd_data = training_data_gau,
-  nldr_data = umap_gau_scaled,
-  bin1 = num_bins_x_gau1,
-  r2 = r2_gau1,
-  is_bin_centroid = TRUE,
-  q = 0.1
-)
-
-df_bin_centroids_gau1 <- gau1_model$df_bin_centroids
-df_bin_gau1 <- gau1_model$df_bin
-
-## Compute error
-
-error_df_gau_abs <- augment(
-  highd_data = training_data_gau,
-  model_2d = df_bin_centroids_gau1,
-  model_highd = df_bin_gau1)
-
-error_df_gau_abs <- error_df_gau_abs |>
-  mutate(sqrt_row_wise_abs_error = sqrt(row_wise_abs_error))
-
-error_df_gau_abs <- error_df_gau_abs |>
-  bind_cols(umap_gau_scaled |>
-              select(-ID))
-
-error_df_gau_umap <- error_df_gau_abs |>
-  ggplot(aes(x = UMAP1,
-             y = UMAP2,
-             z = sqrt_row_wise_abs_error)) +
-  stat_summary_hex(bins=15) +  
-  xlim(c(0.47, 0.5475)) + ylim(c(0.0125, 0.09)) +
-  scale_fill_continuous_sequential(palette = "YlOrRd", n_interp = 20) +
-  interior_annotation("b3", 
-                      position = c(0.08, 0.9),
-                      cex = 1.5)
-
-## Triangulate bin centroids
-tr1_object_gau1 <- tri_bin_centroids(
-  df_bin_centroids_gau1, x = "c_x", y = "c_y")
-tr_from_to_df_gau1 <- gen_edges(
-  tri_object = tr1_object_gau1)
-
-tr_df <- distinct(tibble(
-  x = c(tr_from_to_df_gau1[["x_from"]], tr_from_to_df_gau1[["x_to"]]), 
-  y = c(tr_from_to_df_gau1[["y_from"]], tr_from_to_df_gau1[["y_to"]])))
+algo_obj_gau1 <- gen_nldr_vis_algo_obj(
+  high_d_data = training_data_gau, 
+  nldr_data = umap_data_gau, 
+  num_x_bins = num_bins_x_gau1)
+
+umap_gau_scaled <- algo_obj_gau1$nldr_scaled
+tr_from_to_df_gau1 <- algo_obj_gau1$tr_from_to_df
+df_bin_centroids_gau1 <- algo_obj_gau1$df_bin_centroids
+df_bin_gau1 <- algo_obj_gau1$df_bin
 
@@
 
-## Hexagonal binning to have regular hexagons
-hb_obj_gau1 <- hex_binning(
-  data = umap_gau_scaled,
-  bin1 = num_bins_x_gau1,
-  r2 = r2_gau1)
-
-umap_data_with_hb_id <- hb_obj_gau1$data_hb_id
-
-df_all_gau1 <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID),
-                                umap_data_with_hb_id)
-
-### Define type column
-df <- df_all_gau1 |>
-  dplyr::select(tidyselect::starts_with("x")) |>
-  dplyr::mutate(type = "data") ## original dataset
+df_bin_centroids_gau1 <- df_bin_centroids_gau1 |>
+  dplyr::filter(std_counts > 0)
 
@@
 pacmap_data_gau <- read_rds("data/five_gau_clusters/pacmap_data_five_gau.rds")
-gau1_scaled_obj <- gen_scaled_data(
-  data = pacmap_data_gau)
-pacmap_gau_scaled <- gau1_scaled_obj$scaled_nldr |>
-  select(PaCMAP1, PaCMAP2, ID)
-
-pacmap_gau <- pacmap_gau_scaled |>
-  ggplot(aes(x = PaCMAP1,
-             y = PaCMAP2)) +
-  geom_point(alpha=0.3, color = "#000000")
 
@@
 num_bins_x_gau1 <- 33
-lim1 <- gau1_scaled_obj$lim1
-lim2 <- gau1_scaled_obj$lim2
-r2_gau1 <- diff(lim2)/diff(lim1)
-
-gau1_model <- fit_highd_model(
-  highd_data = training_data_gau,
-  nldr_data = pacmap_gau_scaled,
-  bin1 = num_bins_x_gau1,
-  r2 = r2_gau1,
-  is_bin_centroid = TRUE,
-  q = 0.1
-)
-
-df_bin_centroids_gau1 <- gau1_model$df_bin_centroids
-df_bin_gau1 <- gau1_model$df_bin
-
-## Compute error
-
-error_df_gau_abs <- augment(
-  highd_data = training_data_gau,
-  model_2d = df_bin_centroids_gau1,
-  model_highd = df_bin_gau1)
-
-error_df_gau_abs <- error_df_gau_abs |>
-  mutate(sqrt_row_wise_abs_error = sqrt(row_wise_abs_error))
-
-error_df_gau_abs <- error_df_gau_abs |>
-  bind_cols(pacmap_gau_scaled |>
-              select(-ID))
-
-error_df_gau_pacmap <- error_df_gau_abs |>
-  ggplot(aes(x = PaCMAP1,
-             y = PaCMAP2,
-             z = sqrt_row_wise_abs_error)) +
-  stat_summary_hex(bins=15) + 
-  xlim(c(-0.02, 0.18)) + ylim(c(0.12, 0.32)) +
-  scale_fill_continuous_sequential(palette = "YlOrRd", breaks = 20) +
-  interior_annotation("c3", 
-                      position = c(0.08, 0.9),
-                      cex = 1.5)
-
-## Triangulate bin centroids
-tr1_object_gau1 <- tri_bin_centroids(
-  df_bin_centroids_gau1, x = "c_x", y = "c_y")
-tr_from_to_df_gau1 <- gen_edges(
-  tri_object = tr1_object_gau1)
-
-tr_df <- distinct(tibble(
-  x = c(tr_from_to_df_gau1[["x_from"]], tr_from_to_df_gau1[["x_to"]]), 
-  y = c(tr_from_to_df_gau1[["y_from"]], tr_from_to_df_gau1[["y_to"]])))
+
+algo_obj_gau1 <- gen_nldr_vis_algo_obj(
+  high_d_data = training_data_gau, 
+  nldr_data = pacmap_data_gau, 
+  num_x_bins = num_bins_x_gau1)
+
+pacmap_gau_scaled <- algo_obj_gau1$nldr_scaled
+tr_from_to_df_gau1 <- algo_obj_gau1$tr_from_to_df
+df_bin_centroids_gau1 <- algo_obj_gau1$df_bin_centroids
+df_bin_gau1 <- algo_obj_gau1$df_bin
 
@@
 
-## Hexagonal binning to have regular hexagons
-hb_obj_gau1 <- hex_binning(
-  data = pacmap_gau_scaled,
-  bin1 = num_bins_x_gau1,
-  r2 = r2_gau1,
-  q = 0.1)
-
-pacmap_data_with_hb_id <- hb_obj_gau1$data_hb_id
-
-df_all_gau1 <- dplyr::bind_cols(training_data_gau |> dplyr::select(-ID),
-                                pacmap_data_with_hb_id)
-
-### Define type column
-df <- df_all_gau1 |>
-  dplyr::select(tidyselect::starts_with("x")) |>
-  dplyr::mutate(type = "data") ## original dataset
+df_bin_centroids_gau1 <- df_bin_centroids_gau1 |>
+  dplyr::filter(std_counts > 0)
 
@@
   dplyr::mutate(type = "model") ## Data with summarized mean
-
-# write_rds(df_b, "data/five_gau_clusters/pacmap_model.rds")
 

### 2025-05-05 paper.qmd
Updated Figure 9

@@
 projection <- cbind(
-  c(-0.00215,-0.68905,-0.04778,-0.54223),
-  c(0.42558,-0.23854,-0.63659,0.35753))
+  c(0.07108,0.19751,-0.21229,0.26341),
+  c(-0.23494,0.27965,-0.01430,-0.15782))
 
 proj_obj1 <- get_projection(projection = projection, 
-               proj_scale = 1, 
+               proj_scale = 2, 
                scaled_data = scaled_gau_data, 
@@
                axis_param = list(limits = 1,
-                                 axis_scaled = 1,
+                                 axis_scaled = 2,
                                  axis_pos_x = -0.72, 
                                  axis_pos_y = -0.72, 
-                                 threshold = 0))
+                                 threshold = 0.1))
 
@@
 projection <- cbind(
-  c(-0.00215,-0.68905,-0.04778,-0.54223),
-  c(0.42558, -0.23854, -0.63659, 0.35753))
+  c(0.07108,0.19751,-0.21229,0.26341),
+  c(-0.23494,0.27965,-0.01430,-0.15782))
 
 proj_obj1 <- get_projection(projection = projection, 
-               proj_scale = 1, 
+               proj_scale = 2, 
                scaled_data = scaled_gau_data, 

### 2025-05-05 paper.qmd
Updated Figure 12

@@
 
+df_bin_centroids_pbmc <- df_bin_centroids_pbmc |>
+  dplyr::filter(std_counts > 0)
+
 df_b_pbmc <- df_bin_pbmc |>
@@
 tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_30.rds")
-pbmc_scaled_obj <- gen_scaled_data(
-  data = tsne_pbmc)
-tsne_pbmc_scaled_best <- pbmc_scaled_obj$scaled_nldr
-
 ```

### 2025-05-05 paper.qmd
Updated Figure 6

@@
 #| fig-pos: H
-#| fig-cap: "Hexbin density plots of tSNE layout of the non-linear cluster data, using three different bin inputs: (a) $b = 240/98 \\text{ } (15, \\text{ }16)$, (b) $b = 528/173 \\text{ } (22, \\text{ }24)$, and (c) $b = 2496/549 \\text{ } (48, \\text{ }52)$.
+#| fig-cap: "Hexbin density plots of tSNE layout of the non-linear cluster data, using three different bin inputs: (a) $b = 240/98 \\text{ } (15, \\text{ }16)$, (b) $b = 720/215 \\text{ } (24, \\text{ }30)$, and (c) $b = 2496/549 \\text{ } (48, \\text{ }52)$.
 Color indicates standardized counts, dark indicating high count and light indicates low count.
@@
 - the choice of NLDR method, and the parameters used to create it, and
-- model fit parameters: bin size, low density bin removal, long edge removal.
+- model fit parameters: bin size, low density bin removal.
 

### 2025-05-05 paper.qmd
Updated Figure 7

@@
            linewidth=1) +
-  geom_vline(xintercept = 0.06,
+  geom_vline(xintercept = 0.05,
              color = "#1f78b4", 
@@
   geom_point(size = 1) +
-  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_umap$a1, 2)))[c(2, 5, 8, 10, 12, 14, 15, 16, 17)]) +
+  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_umap$a1, 2)))[c(2, 4, 8, 11, 13, 14, 15, 16)]) +
   labs(x = expression(paste("binwidth (", a[1], ")")), y = "MSE") +
@@
   # scale_y_continuous(breaks = sort(unique(error_two_curvy$b_non_empty))[-3]) +
-  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_umap$a1, 2)))[c(2, 5, 8, 10, 12, 14, 15, 16, 17)]) +
+  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_umap$a1, 2)))[c(2, 4, 8, 11, 13, 14, 15, 16)]) +
   labs(x = expression(paste("binwidth (", a[1], ")")), y = paste("Relative proportion density")) +
@@
   # scale_y_continuous(breaks = sort(unique(error_two_curvy$b_non_empty))[-3]) +
-  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_umap$a1, 2)))[c(2, 5, 8, 10, 12, 14, 15, 16, 17)]) +
+  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_umap$a1, 2)))[c(2, 4, 8, 11, 13, 14, 15, 16)]) +
   labs(x = expression(paste("binwidth (", a[1], ")")), y = expression(paste("proportion of non-empty bins ", bgroup("(", frac(m, b), ")")))) +
@@
 A good benchmark value for these parameters is when the MSE drops and then flattens out.
-Three binwidth choices were made: $0.03$, $0.06$, and $0.09$ to investigate.
+Three binwidth choices were made: $0.03$, $0.05$, and $0.09$ to investigate.
 As the binwidth increases, the proportion of non-empty bins also increases (b).
 The relative proportion density decreases and levels off (d).
-Binwidth $0.06$ is chosen as the initial best binwidth for further analysis.
+Binwidth $0.05$ is chosen as the initial best binwidth for further analysis.
 There is no need to remove the low-density hexagons because as shown in (b), there is no considerable drop in MSE."
@@
   data = tsne_two_curvy_scaled, 
-  bin1 = 8, 
+  bin1 = 24, 
   r2 = r2_tsne)
@@
 a1_2_tsne <- calc_bins_y(
-  bin1 = 8, 
+  bin1 = 24, 
   r2 = r2_tsne

### 2025-05-06 paper.qmd
Updated spacing
Try to reduce the number of pages

@@
 
-\spacingset{1.0} <!--% command in JASA style, comment to go back to double spacing-->
+\spacingset{2.0} <!--% command in JASA style, comment to go back to double spacing-->
 
@@
 #| out-width: 100%
+#| fig-pos: H
 # (a) UMAP (n_neighbors = 30, min_dist = 0.3), (b) UMAP (n_neighbors = 5, min_dist = 0.01), (c) UMAP (n_neighbors = 15, min_dist = 0.99), (f) tSNE (perplexity = 5), (g) tSNE (perplexity = 30), (l) TriMAP (n_inliers = 12, n_outliers = 4, n_random = 3), (q) PaCMAP (n_neighbors = 30, init = random, MN_ratio = 0.9, FP_ratio = 5)

### 2025-05-12 paper.qmd
Start updating with new quollr version

@@
 #| label: two-curvy-training
-training_data_two_curvy <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_data.rds")
+training_data_two_curvy <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_data.rds") |>
+  mutate(ID = row_number())
 ```
@@
 data_two_curvy <- training_data_two_curvy |>
+  select(-ID) |>
   mutate(type = "data")
@@
 #| label: tsne-two-curvy
-tsne_two_curvy <- read_rds(file = "data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_30.rds") 
+tsne_two_curvy <- read_rds(file = "data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_30.rds") |>
+  mutate(ID = row_number()) |>
+  rename(c("emb1" = "tSNE1",
+           "emb2" = "tSNE2"))
 ```
@@
 
-algo_obj_two_curvy <- gen_nldr_vis_algo_obj(
-  high_d_data = training_data_two_curvy, 
+algo_obj_two_curvy <- fit_highd_model(
+  highd_data = training_data_two_curvy, 
   nldr_data = tsne_two_curvy, 
-  num_x_bins = num_bins_x_two_curvy)
-
-tsne_two_curvy_scaled <- algo_obj_two_curvy$nldr_scaled
-tr_from_to_df_two_curvy <- algo_obj_two_curvy$tr_from_to_df
-df_bin_centroids_two_curvy <- algo_obj_two_curvy$df_bin_centroids
-df_bin_two_curvy <- algo_obj_two_curvy$df_bin
-hex_grid_with_counts <- algo_obj_two_curvy$hex_grid_with_counts
-hex_grid_nonempty <- algo_obj_two_curvy$hex_grid_nonempty
-bin_width <- algo_obj_two_curvy$a1
+  bin1 = num_bins_x_two_curvy, 
+  q = 0.1, 
+  benchmark_highdens = 5)
+
+tsne_two_curvy_scaled <- algo_obj_two_curvy$nldr_obj$scaled_nldr
+tr_from_to_df_two_curvy <- algo_obj_two_curvy$trimesh_data
+df_bin_centroids_two_curvy <- algo_obj_two_curvy$model_2d
+df_bin_two_curvy <- algo_obj_two_curvy$model_highd
+hex_grid <- algo_obj_two_curvy$hb_obj$hex_poly
+counts_df <- algo_obj_two_curvy$hb_obj$std_cts
+
+hex_grid_with_counts <- left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hexID"))
+
+hex_grid_nonempty <- hex_grid |>
+  filter(hex_poly_id %in% df_bin_centroids_two_curvy$hexID)
+
+bin_width <- algo_obj_two_curvy$hb_obj$a1
 
@@
 nldr_two_curvy_original <- ggplot(tsne_two_curvy_scaled, 
-                                  aes(x = tSNE1, y = tSNE2)) +
+                                  aes(x = emb1, y = emb2)) +
   geom_point(alpha = 0.5, color = clr_choice, size = 0.5)
@@
   geom_point(data = tsne_two_curvy_scaled, 
-             aes(x = tSNE1, y = tSNE2), 
+             aes(x = emb1, y = emb2), 
              alpha = 0.5, size = 0.5, color = clr_choice) +
@@
   geom_point(data = tsne_two_curvy_scaled,
-             aes(x = tSNE1, y = tSNE2),
+             aes(x = emb1, y = emb2),
              colour = clr_choice,
@@
 
-df_bin_centroids_two_curvy <- df_bin_centroids_two_curvy |>
-  dplyr::filter(std_counts > 0)
+# df_bin_centroids_two_curvy <- df_bin_centroids_two_curvy |>
+#   dplyr::filter(std_counts > 0)
 
@@
              aes(
-               x = tSNE1,
-               y = tSNE2
+               x = emb1,
+               y = emb2
              ),
@@
 
-training_data_two_curvy <- training_data_two_curvy |>
-  mutate(ID = row_number())
-
 ## Compute error
@@
 error_plot_two_curvy <- error_df_two_curvy_abs |>
-  ggplot(aes(x = tSNE1,
-             y = tSNE2,
+  ggplot(aes(x = emb1,
+             y = emb2,
              colour = sqrt_row_wise_total_error)) +
@@
 df_bin_two_curvy <- df_bin_two_curvy |>
-  select(-hb_id) |>
+  select(-hexID) |>
   mutate(type = "model")
@@
 <!-- df_bin_two_curvy <- df_bin_two_curvy |> -->
-<!--   select(-hb_id) |> -->
+<!--   select(-hexID) |> -->
 <!--   mutate(type = "model") -->
@@
 df_b_two_curvy4 <- df_bin_two_curvy4 |>
-  dplyr::filter(hb_id %in% df_bin_centroids4$hexID) |>
+  dplyr::filter(hexID %in% df_bin_centroids4$hexID) |>
   dplyr::mutate(type = "model") ## Data with summarized mean
@@
 ## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
-df_b_two_curvy4 <- df_b_two_curvy4[match(df_bin_centroids4$hexID, df_b_two_curvy4$hb_id),] |>
-  dplyr::select(-hb_id)
+df_b_two_curvy4 <- df_b_two_curvy4[match(df_bin_centroids4$hexID, df_b_two_curvy4$hexID),] |>
+  dplyr::select(-hexID)
 
@@
 df_b <- df_bin_gau1 |>
-  dplyr::filter(hb_id %in% df_bin_centroids_gau1$hexID) |>
+  dplyr::filter(hexID %in% df_bin_centroids_gau1$hexID) |>
   dplyr::mutate(type = "model") ## Data with summarized mean
@@
 ## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
-df_b <- df_b[match(df_bin_centroids_gau1$hexID, df_b$hb_id),] |>
-  dplyr::select(-hb_id)
+df_b <- df_b[match(df_bin_centroids_gau1$hexID, df_b$hexID),] |>
+  dplyr::select(-hexID)
 
@@
 df_bin_dens_clust <- df_bin_dens_clust |>
-  select(-hb_id) |>
+  select(-hexID) |>
   mutate(type = "model")
@@
 df_b_pbmc <- df_bin_pbmc |>
-  dplyr::filter(hb_id %in% df_bin_centroids_pbmc$hexID) |>
+  dplyr::filter(hexID %in% df_bin_centroids_pbmc$hexID) |>
   dplyr::mutate(type = "model") ## Data with summarized mean
@@
 ## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
-df_b_pbmc <- df_b_pbmc[match(df_bin_centroids_pbmc$hexID, df_b_pbmc$hb_id),] |>
-  dplyr::select(-hb_id)
+df_b_pbmc <- df_b_pbmc[match(df_bin_centroids_pbmc$hexID, df_b_pbmc$hexID),] |>
+  dplyr::select(-hexID)
 
@@
 df_b_mnist <- df_bin_mnist |>
-  dplyr::filter(hb_id %in% hexID_mnist) |>
+  dplyr::filter(hexID %in% hexID_mnist) |>
   dplyr::mutate(type = "model") ## Data with summarized mean
@@
 ## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
-df_b_mnist <- df_b_mnist[match(hexID_mnist, df_b_mnist$hb_id),] |>
-  dplyr::select(-hb_id)
+df_b_mnist <- df_b_mnist[match(hexID_mnist, df_b_mnist$hexID),] |>
+  dplyr::select(-hexID)
 

### 2025-05-12 paper.qmd
Updated MNIST model building

@@
 
-algo_obj_mnist <- gen_nldr_vis_algo_obj(
-  high_d_data = training_data_mnist, 
+algo_obj_mnist <- fit_highd_model(
+  highd_data = training_data_mnist, 
   nldr_data = tsne_mnist2, 
-  num_x_bins = num_bins_x_mnist)
-
-tsne_minst_scaled <- algo_obj_mnist$nldr_scaled
-tr_from_to_df_mnist <- algo_obj_mnist$tr_from_to_df
-df_bin_centroids_mnist <- algo_obj_mnist$df_bin_centroids
-df_bin_mnist <- algo_obj_mnist$df_bin
+  bin1 = num_bins_x_mnist, 
+  q = 0.1, 
+  benchmark_highdens = 5)
+
+tsne_minst_scaled <- algo_obj_mnist$nldr_obj$scaled_nldr
+tr_from_to_df_mnist <- algo_obj_mnist$trimesh_data
+df_bin_centroids_mnist <- algo_obj_mnist$model_2d
+df_bin_mnist <- algo_obj_mnist$model_highd
 
 sc_ltr_pos_mnist <- c(0.96, 0.96)
-
-# tr_from_to_df_mnist <- tr_from_to_df_mnist |>
-#   mutate(ID = row_number()) |>
-#   filter(!(ID %in% c(1205, 1629, 1170)))
 
@@
 hexID_mnist <- df_bin_centroids_mnist |>
-  dplyr::filter(std_counts > 0) |>
+  #dplyr::filter(std_counts > 0) |>
   dplyr::pull(hexID)

### 2025-05-12 paper.qmd
Updated curosities

@@
 
-algo_obj_gau1 <- gen_nldr_vis_algo_obj(
-  high_d_data = training_data_gau, 
+algo_obj_gau1 <- fit_highd_model(
+  highd_data = training_data_gau, 
   nldr_data = tsne_data_gau, 
-  num_x_bins = num_bins_x_gau1)
-
-tsne_gau_scaled <- algo_obj_gau1$nldr_scaled
-tr_from_to_df_gau1 <- algo_obj_gau1$tr_from_to_df
-df_bin_centroids_gau1 <- algo_obj_gau1$df_bin_centroids
-df_bin_gau1 <- algo_obj_gau1$df_bin
+  bin1 = num_bins_x_gau1, 
+  q = 0.1, 
+  benchmark_highdens = 5)
+
+tsne_gau_scaled <- algo_obj_gau1$nldr_obj$scaled_nldr
+tr_from_to_df_gau1 <- algo_obj_gau1$trimesh_data
+df_bin_centroids_gau1 <- algo_obj_gau1$model_2d
+df_bin_gau1 <- algo_obj_gau1$model_highd
 
@@
 
-df_bin_centroids_gau1 <- df_bin_centroids_gau1 |>
-  dplyr::filter(std_counts > 0)
+# df_bin_centroids_gau1 <- df_bin_centroids_gau1 |>
+#   dplyr::filter(std_counts > 0)
 
@@
 num_bins_x_gau1 <- 75
-algo_obj_gau1 <- gen_nldr_vis_algo_obj(
-  high_d_data = training_data_gau, 
-  nldr_data = umap_data_gau, 
-  num_x_bins = num_bins_x_gau1)
-
-umap_gau_scaled <- algo_obj_gau1$nldr_scaled
-tr_from_to_df_gau1 <- algo_obj_gau1$tr_from_to_df
-df_bin_centroids_gau1 <- algo_obj_gau1$df_bin_centroids
-df_bin_gau1 <- algo_obj_gau1$df_bin
+
+algo_obj_gau1 <- fit_highd_model(
+  highd_data = training_data_gau, 
+  nldr_data = tsne_data_gau, 
+  bin1 = num_bins_x_gau1, 
+  q = 0.1, 
+  benchmark_highdens = 5)
+
+tsne_gau_scaled <- algo_obj_gau1$nldr_obj$scaled_nldr
+tr_from_to_df_gau1 <- algo_obj_gau1$trimesh_data
+df_bin_centroids_gau1 <- algo_obj_gau1$model_2d
+df_bin_gau1 <- algo_obj_gau1$model_highd
 
@@
 
-algo_obj_gau1 <- gen_nldr_vis_algo_obj(
-  high_d_data = training_data_gau, 
-  nldr_data = pacmap_data_gau, 
-  num_x_bins = num_bins_x_gau1)
-
-pacmap_gau_scaled <- algo_obj_gau1$nldr_scaled
-tr_from_to_df_gau1 <- algo_obj_gau1$tr_from_to_df
-df_bin_centroids_gau1 <- algo_obj_gau1$df_bin_centroids
-df_bin_gau1 <- algo_obj_gau1$df_bin
+algo_obj_gau1 <- fit_highd_model(
+  highd_data = training_data_gau, 
+  nldr_data = tsne_data_gau, 
+  bin1 = num_bins_x_gau1, 
+  q = 0.1, 
+  benchmark_highdens = 5)
+
+tsne_gau_scaled <- algo_obj_gau1$nldr_obj$scaled_nldr
+tr_from_to_df_gau1 <- algo_obj_gau1$trimesh_data
+df_bin_centroids_gau1 <- algo_obj_gau1$model_2d
+df_bin_gau1 <- algo_obj_gau1$model_highd
 
@@
 
-algo_obj_dens_clust <- gen_nldr_vis_algo_obj(
-  high_d_data = one_c_shaped_data, 
+algo_obj_dens_clust <- fit_highd_model(
+  highd_data = one_c_shaped_data, 
   nldr_data = tsne_one_c_shaped, 
-  num_x_bins = num_bins_x_one_c_shaped)
-
-tsne_one_c_shaped_scaled <- algo_obj_dens_clust$nldr_scaled
-tr_from_to_df_dens_clust <- algo_obj_dens_clust$tr_from_to_df
-df_bin_centroids_dens_clust <- algo_obj_dens_clust$df_bin_centroids
-df_bin_dens_clust <- algo_obj_dens_clust$df_bin
-hex_grid_with_counts_dens_clust <- algo_obj_dens_clust$hex_grid_with_counts
+  bin1 = num_bins_x_one_c_shaped, 
+  q = 0.1, 
+  benchmark_highdens = 5)
+
+tsne_one_c_shaped_scaled <- algo_obj_dens_clust$nldr_obj$scaled_nldr
+tr_from_to_df_dens_clust <- algo_obj_dens_clust$trimesh_data
+df_bin_centroids_dens_clust <- algo_obj_dens_clust$model_2d
+df_bin_dens_clust <- algo_obj_dens_clust$model_highd
 ```
@@
 
-algo_obj_pbmc <- gen_nldr_vis_algo_obj(
-  high_d_data = training_data_pbmc, 
+algo_obj_pbmc <- fit_highd_model(
+  highd_data = training_data_pbmc, 
   nldr_data = umap_pbmc, 
-  num_x_bins = num_bins_x_pbmc)
-
-umap_pbmc_scaled <- algo_obj_pbmc$nldr_scaled
-tr_from_to_df_pbmc <- algo_obj_pbmc$tr_from_to_df
-df_bin_centroids_pbmc <- algo_obj_pbmc$df_bin_centroids
-df_bin_pbmc <- algo_obj_pbmc$df_bin
+  bin1 = num_bins_x_pbmc, 
+  q = 0.1, 
+  benchmark_highdens = 5)
+
+umap_pbmc_scaled <- algo_obj_pbmc$nldr_obj$scaled_nldr
+tr_from_to_df_pbmc <- algo_obj_pbmc$trimesh_data
+df_bin_centroids_pbmc <- algo_obj_pbmc$model_2d
+df_bin_pbmc <- algo_obj_pbmc$model_highd
 
@@
 
-df_bin_centroids_pbmc <- df_bin_centroids_pbmc |>
-  dplyr::filter(std_counts > 0)
+# df_bin_centroids_pbmc <- df_bin_centroids_pbmc |>
+#   dplyr::filter(std_counts > 0)
 
@@
 
-algo_obj_pbmc <- gen_nldr_vis_algo_obj(
-  high_d_data = training_data_pbmc, 
+algo_obj_pbmc <- fit_highd_model(
+  highd_data = training_data_pbmc, 
   nldr_data = tsne_pbmc, 
-  num_x_bins = num_bins_x_tsne_pbmc)
-
-tsne_pbmc_scaled_best <- algo_obj_pbmc$nldr_scaled
-distance_pbmc <- algo_obj_pbmc$distance_df
-tr_from_to_df_pbmc <- algo_obj_pbmc$tr_from_to_df
-df_bin_centroids_pbmc <- algo_obj_pbmc$df_bin_centroids
-df_bin_pbmc <- algo_obj_pbmc$df_bin
+  bin1 = num_bins_x_tsne_pbmc, 
+  q = 0.1, 
+  benchmark_highdens = 5)
+
+tsne_pbmc_scaled_best <- algo_obj_pbmc$nldr_obj$scaled_nldr
+tr_from_to_df_pbmc <- algo_obj_pbmc$trimesh_data
+df_bin_centroids_pbmc <- algo_obj_pbmc$model_2d
+df_bin_pbmc <- algo_obj_pbmc$model_highd
 
@@
 
-df_bin_centroids_pbmc <- df_bin_centroids_pbmc |>
-  dplyr::filter(std_counts > 0)
-
 df_b_pbmc <- df_bin_pbmc |>

### 2025-05-12 paper.qmd
Updated plot_proj

@@
 
-proj_obj1 <- get_projection(projection = model_prj1, 
+proj_obj1 <- quollr::get_projection(projection = model_prj1, 
                proj_scale = 1, 
-               scaled_data = scaled_two_curvy_data, 
-               scaled_data_model = scaled_two_curvy_data_model, 
-               distance_df_small_edges = tr_from_to_df_two_curvy, 
+               highd_data = scaled_two_curvy_data, 
+               model_highd = scaled_two_curvy_data_model, 
+               trimesh_data = tr_from_to_df_two_curvy, 
                axis_param = list(limits = 0.5, 
@@
   plot_limits = c(-0.6, 0.6), 
-  title = "a", 
-  cex = 1.5, 
   axis_text_size = 4, 
-  is_color = FALSE) +
+  is_category = FALSE) +
+  interior_annotation(label = "a", cex = 1.5) +
   theme(aspect.ratio=1, 
@@
     plot_limits = c(-0.6, 0.6), 
-    title = "b", 
-    cex = 1, 
-    position = sc_ltr_pos,
     axis_text_size = 2, 
-    is_color = FALSE) +
+    is_category = FALSE) +
+  interior_annotation(label = "b", position = sc_ltr_pos, cex = 1) +
   theme(aspect.ratio = 1,
@@
                             proj_scale = 1, 
-                            scaled_data = scaled_two_curvy_data, 
-                            scaled_data_model = scaled_two_curvy_data_model, 
-                            distance_df_small_edges = tr_from_to_df_two_curvy, 
+                            highd_data = scaled_two_curvy_data, 
+                            model_highd = scaled_two_curvy_data_model, 
+                            trimesh_data = tr_from_to_df_two_curvy, 
                             axis_param = list(limits = 0.35, 
@@
     plot_limits = c(-0.42, 0.42), 
-    title = "b", 
-    cex = 1.5, 
     axis_text_size = 4, 
-    is_color = FALSE) +
+    is_category = FALSE) +
+  interior_annotation(label = "b", cex = 1.5) +
   theme(aspect.ratio = 1, 
@@
   plot_limits = c(-0.5, 0.5), 
-  title = "c", 
-  cex = 1, 
-  position = sc_ltr_pos,
   axis_text_size = 2, 
-  is_color = FALSE) +
+  is_category = FALSE) +
+  interior_annotation(label = "c", position = sc_ltr_pos, cex = 1) +
   theme(aspect.ratio = 1) #+
@@
 
-algo_obj_two_curvy2 <- gen_nldr_vis_algo_obj(
-  high_d_data = training_data_two_curvy, 
+algo_obj_two_curvy2 <- fit_highd_model(
+  highd_data = training_data_two_curvy, 
   nldr_data = tsne_two_curvy, 
-  num_x_bins = num_bins_x_two_curvy)
-
-tr_from_to_df_two_curvy2 <- algo_obj_two_curvy2$tr_from_to_df
-df_bin_centroids_two_curvy2 <- algo_obj_two_curvy2$df_bin_centroids
-df_bin_two_curvy2 <- algo_obj_two_curvy2$df_bin
-hex_grid_with_counts_two_curvy2 <- algo_obj_two_curvy2$hex_grid_with_counts
+  bin1 = num_bins_x_two_curvy, 
+  q = 0.1, 
+  benchmark_highdens = 5)
+
+tr_from_to_df_two_curvy2 <- algo_obj_two_curvy2$trimesh_data
+df_bin_centroids_two_curvy2 <- algo_obj_two_curvy2$model_2d
+df_bin_two_curvy2 <- algo_obj_two_curvy2$model_highd
+
+hex_grid_two_curvy2 <- algo_obj_two_curvy2$hb_obj$hex_poly
+counts_df_two_curvy2 <- algo_obj_two_curvy2$hb_obj$std_cts
+
+hex_grid_with_counts_two_curvy2 <- left_join(hex_grid_two_curvy2, counts_df_two_curvy2, by = c("hex_poly_id" = "hexID"))
 
@@
 df_bin_two_curvy3 <- algo_obj_two_curvy3$df_bin
-hex_grid_with_counts_two_curvy3 <- algo_obj_two_curvy3$hex_grid_with_counts
+
+hex_grid_two_curvy3 <- algo_obj_two_curvy3$hb_obj$hex_poly
+counts_df_two_curvy2 <- algo_obj_two_curvy3$hb_obj$std_cts
+
+hex_grid_with_counts_two_curvy3 <- left_join(hex_grid_two_curvy3, counts_df_two_curvy3, by = c("hex_poly_id" = "hexID"))
 
@@
 <!--     axis_text_size = 4,  -->
-<!--     is_color = FALSE) + -->
+<!--     is_category = FALSE) + -->
 <!--     theme( -->
@@
 <!--     axis_text_size = 4,  -->
-<!--     is_color = FALSE) + -->
+<!--     is_category = FALSE) + -->
 <!--     theme(aspect.ratio = 1,  -->
@@
                proj_scale = 2, 
-               scaled_data = scaled_gau_data, 
-               scaled_data_model = scaled_gau_data_model, 
-               distance_df_small_edges = tr_from_to_df_gau1, 
+               highd_data = scaled_gau_data, 
+               model_highd = scaled_gau_data_model, 
+               trimesh_data = tr_from_to_df_gau1, 
                axis_param = list(limits = 1,
@@
                             proj_scale = 3, 
-                            scaled_data = scaled_gau_data, 
-                            scaled_data_model = scaled_gau_data_model, 
-                            distance_df_small_edges = tr_from_to_df_gau1, 
+                            highd_data = scaled_gau_data, 
+                            model_highd = scaled_gau_data_model, 
+                            trimesh_data = tr_from_to_df_gau1, 
                             axis_param = list(limits = 1.5,
@@
   plot_limits = c(-1.3, 1.3), 
-  title = "a3", 
-  cex = 1.5,
-  position = c(0.08, 0.9),
   axis_text_size = 4, 
-  is_color = FALSE) +
+  is_category = FALSE) +
+  interior_annotation(label = "a3", position = c(0.08, 0.9), cex = 1.5) +
   theme(aspect.ratio = 1,
@@
   plot_limits = c(-1.3, 1.3), 
-  title = "b3", 
-  cex = 1.5,
-  position = c(0.08, 0.9),
   axis_text_size = 4, 
-  is_color = FALSE) +
+  is_category = FALSE) +
+  interior_annotation(label = "b3", position = c(0.08, 0.9), cex = 1.5) +
   theme(aspect.ratio = 1,
@@
   plot_limits = c(-1.3, 1.3), 
-  title = "c3", 
-  cex = 1.5,
-  position = c(0.08, 0.9),
   axis_text_size = 4, 
-  is_color = FALSE) +
+  is_category = FALSE) +
+  interior_annotation(label = "c3", position = c(0.08, 0.9), cex = 1.5) +
   theme(aspect.ratio = 1,
@@
                             proj_scale = 1,
-                            scaled_data = scaled_c_shaped_data,
-                            scaled_data_model = scaled_c_shaped_data_model,
-                            distance_df_small_edges = tr_from_to_df_dens_clust,
+                            highd_data = scaled_c_shaped_data,
+                            model_highd = scaled_c_shaped_data_model,
+                            trimesh_data = tr_from_to_df_dens_clust,
                             axis_param = list(limits = 0.4,
@@
     plot_limits = c(-0.35, 0.35), 
-    title = "a3", 
-    cex = 2,
-    position = c(0.08, 0.9),
     axis_text_size = 4.5, 
-    is_color = TRUE) +
+    is_category = TRUE) +
+  interior_annotation(label = "a3", position = c(0.08, 0.9), cex = 2) +
   scale_color_manual(values = c('#d9d9d9', '#800026')) +
@@
     plot_limits = c(-0.35, 0.35), 
-    title = "b3", 
-    cex = 2,
-    position = c(0.08, 0.9),
     axis_text_size = 4.5, 
-    is_color = TRUE) +
+    is_category = TRUE) +
+  interior_annotation(label = "b3", position = c(0.08, 0.9), cex = 2) +
   scale_color_manual(values = c('#d9d9d9', '#800026')) +
@@
                proj_scale = 1.23, 
-               scaled_data = scaled_pbmc_data, 
-               scaled_data_model = scaled_pbmc_data_model, 
-               distance_df_small_edges = tr_from_to_df_pbmc, 
+               highd_data = scaled_pbmc_data, 
+               model_highd = scaled_pbmc_data_model, 
+               trimesh_data = tr_from_to_df_pbmc, 
                axis_param = list(limits = 1,
@@
   plot_limits = c(-0.65, 0.65), 
-  title = "a2", 
-  cex = 2, 
   axis_text_size = 4,
-  is_color = TRUE) + scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))
+  is_category = TRUE) +
+  interior_annotation(label = "a2", cex = 2) + 
+  scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))
 
@@
                proj_scale = 1.25, 
-               scaled_data = scaled_pbmc_data, 
-               scaled_data_model = scaled_pbmc_data_model, 
-               distance_df_small_edges = tr_from_to_df_pbmc, 
+               highd_data = scaled_pbmc_data, 
+               model_highd = scaled_pbmc_data_model, 
+               trimesh_data = tr_from_to_df_pbmc, 
                axis_param = list(limits = 1.35, 
@@
     plot_limits = c(-1, 0.75), 
-    title = "a3", 
-    cex = 2, 
     axis_text_size = 4,
-    is_color = TRUE) + scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))
+    is_category = TRUE) +
+  interior_annotation(label = "a3", cex = 2) +
+  scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))
 
@@
                             proj_scale = 1.23, 
-                            scaled_data = scaled_pbmc_data, 
-                            scaled_data_model = scaled_pbmc_data_model, 
-                            distance_df_small_edges = tr_from_to_df_pbmc, 
+                            highd_data = scaled_pbmc_data, 
+                            model_highd = scaled_pbmc_data_model, 
+                            trimesh_data = tr_from_to_df_pbmc, 
                             axis_param = list(limits = 1, 
@@
   plot_limits = c(-0.65, 0.65), 
-  title = "b2", 
-  cex = 2, 
   axis_text_size = 4,
-  is_color = TRUE) + scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))
+  is_category = TRUE) +
+  interior_annotation(label = "b2", cex = 2) +
+  scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))
 
@@
                             proj_scale = 1.25, 
-                            scaled_data = scaled_pbmc_data, 
-                            scaled_data_model = scaled_pbmc_data_model, 
-                            distance_df_small_edges = tr_from_to_df_pbmc, 
+                            highd_data = scaled_pbmc_data, 
+                            model_highd = scaled_pbmc_data_model, 
+                            trimesh_data = tr_from_to_df_pbmc, 
                             axis_param = list(limits = 1.35, 
@@
   plot_limits = c(-1, 0.75), 
-  title = "b3", 
-  cex = 2, 
   axis_text_size = 4,
-  is_color = TRUE) + scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))
+  is_category = TRUE) +
+  interior_annotation(label = "b3", cex = 2) + 
+  scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))
 ```
@@
                proj_scale = 2.5, 
-               scaled_data = scaled_mnist_data, 
-               scaled_data_model = scaled_mnist_data_model, 
-               distance_df_small_edges = tr_from_to_df_mnist, 
+               highd_data = scaled_mnist_data, 
+               model_highd = scaled_mnist_data_model, 
+               trimesh_data = tr_from_to_df_mnist, 
                axis_param = list(limits = 1, 
@@
   plot_limits = c(-1.1, 0.9), 
-  title = "b", 
-  cex = 2, 
   axis_text_size = 3.5, 
-  is_color = TRUE) +
+  is_category = TRUE) +
+  interior_annotation(label = "b", cex = 2) +
   scale_color_manual(values = c("#999999",'#ff7f00')) +
@@
     plot_limits = c(-1.1, 0.9), 
-    title = "b", 
-    cex = 2.5, 
     axis_text_size = 4.5, 
-    is_color = TRUE) +
+    is_category = TRUE) +
+  interior_annotation(label = "b", cex = 2.5) +
   scale_color_manual(values = c("#d9d9d9",'#ff7f00')) +
@@
                proj_scale = 2, 
-               scaled_data = scaled_mnist_data, 
-               scaled_data_model = scaled_mnist_data_model, 
-               distance_df_small_edges = tr_from_to_df_mnist, 
+               highd_data = scaled_mnist_data, 
+               model_highd = scaled_mnist_data_model, 
+               trimesh_data = tr_from_to_df_mnist, 
                axis_param = list(limits = 1, 
@@
     plot_limits = c(-1, 1), 
-    title = "c", 
-    cex = 2, 
     axis_text_size = 3.5, 
-    is_color = TRUE) +
+    is_category = TRUE) +
+  interior_annotation(label = "c", cex = 2) +
   scale_color_manual(values = c("#999999",'#ff7f00')) +
@@
     plot_limits = c(-1, 1), 
-    title = "c", 
-    cex = 2.5, 
     axis_text_size = 4.5, 
-    is_color = TRUE) +
+    is_category = TRUE) +
+  interior_annotation(label = "c", cex = 2.5) +
   scale_color_manual(values = c("#d9d9d9",'#ff7f00')) +

### 2025-05-12 paper.qmd
Updated with new quollr version

@@
 hb_obj_notation <- hex_binning(
-  data = tsne_two_curvy_scaled, 
+  nldr_obj = algo_obj_two_curvy$nldr_obj, 
   bin1 = 7, 
-  r2 = 1.6,
   q = 0.1)
@@
   geom_point(data = tsne_two_curvy_scaled,
-           aes(x = tSNE1, y = tSNE2),
+           aes(x = emb1, y = emb2),
            alpha = 0.3,
@@
 
-algo_obj_two_curvy3 <- gen_nldr_vis_algo_obj(
-  high_d_data = training_data_two_curvy, 
+algo_obj_two_curvy3 <- fit_highd_model(
+    highd_data = training_data_two_curvy, 
     nldr_data = tsne_two_curvy, 
-  num_x_bins = num_bins_x_two_curvy)
-
-tr_from_to_df_two_curvy3 <- algo_obj_two_curvy3$tr_from_to_df
-df_bin_centroids_two_curvy3 <- algo_obj_two_curvy3$df_bin_centroids
-df_bin_two_curvy3 <- algo_obj_two_curvy3$df_bin
+    bin1 = num_bins_x_two_curvy, 
+    q = 0.1, 
+    benchmark_highdens = 5)
+
+tr_from_to_df_two_curvy3 <- algo_obj_two_curvy3$trimesh_data
+df_bin_centroids_two_curvy3 <- algo_obj_two_curvy3$model_2d
+df_bin_two_curvy3 <- algo_obj_two_curvy3$model_highd
 
 hex_grid_two_curvy3 <- algo_obj_two_curvy3$hb_obj$hex_poly
-counts_df_two_curvy2 <- algo_obj_two_curvy3$hb_obj$std_cts
+counts_df_two_curvy3 <- algo_obj_two_curvy3$hb_obj$std_cts
 
@@
 predict_two_curvy_obj <- gen_scaled_data(
-    data = predict_tsne_two_curvy)
+    nldr_data = predict_tsne_two_curvy)
 
@@
 # and plot them on same aspect ratio
-tsne_two_non_linear_diff_shaped_close_clusters <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_30.rds")
-
-nldr_two_non_linear_diff_shaped_close_clusters1 <- tsne_two_non_linear_diff_shaped_close_clusters |>
+tsne_two_curvy <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_30.rds")
+
+nldr_two_curvy1 <- tsne_two_curvy |>
   ggplot(aes(x = tSNE1,
@@
 
-tsne_two_non_linear_diff_shaped_close_clusters <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_62.rds")
-
-nldr_two_non_linear_diff_shaped_close_clusters6 <- tsne_two_non_linear_diff_shaped_close_clusters |>
+tsne_two_curvy <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_62.rds")
+
+nldr_two_curvy6 <- tsne_two_curvy |>
   ggplot(aes(x = tSNE1,
@@
 
-umap_two_non_linear_diff_shaped_close_clusters <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_umap_n-neigbors_15_min-dist_0.1.rds")
-
-nldr_two_non_linear_diff_shaped_close_clusters2 <- umap_two_non_linear_diff_shaped_close_clusters |>
+umap_two_curvy <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_umap_n-neigbors_15_min-dist_0.1.rds")
+
+nldr_two_curvy2 <- umap_two_curvy |>
   ggplot(aes(x = UMAP1,
@@
 
-phate_two_non_linear_diff_shaped_close_clusters <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_phate_knn_5.rds")
-
-nldr_two_non_linear_diff_shaped_close_clusters3 <- phate_two_non_linear_diff_shaped_close_clusters |>
+phate_two_curvy <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_phate_knn_5.rds")
+
+nldr_two_curvy3 <- phate_two_curvy |>
   ggplot(aes(x = PHATE1,
@@
 
-trimap_two_non_linear_diff_shaped_close_clusters <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_trimap_n-inliers_12_n-outliers_4_n-random_3.rds")
-
-nldr_two_non_linear_diff_shaped_close_clusters4 <- trimap_two_non_linear_diff_shaped_close_clusters |>
+trimap_two_curvy <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_trimap_n-inliers_12_n-outliers_4_n-random_3.rds")
+
+nldr_two_curvy4 <- trimap_two_curvy |>
   ggplot(aes(x = TriMAP1,
@@
 
-pacmap_two_non_linear_diff_shaped_close_clusters <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")
-
-nldr_two_non_linear_diff_shaped_close_clusters5 <- pacmap_two_non_linear_diff_shaped_close_clusters |>
+pacmap_two_curvy <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")
+
+nldr_two_curvy5 <- pacmap_two_curvy |>
   ggplot(aes(x = PaCMAP1,
@@
 ```{r}
-#| label: combine-data-two_non_linear_diff_shaped_close_clusters
+#| label: combine-data-two_curvy
 
 error_two_curvy_umap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_umap.rds")
-error_two_non_linear_diff_shaped_close_clusters_tsne <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_tsne.rds")
-error_two_non_linear_diff_shaped_close_clusters_phate <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_phate.rds")
-error_two_non_linear_diff_shaped_close_clusters_trimap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_trimap.rds")
-error_two_non_linear_diff_shaped_close_clusters_pacmap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_pacmap.rds")
-error_two_non_linear_diff_shaped_close_clusters_tsne2 <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_tsne2.rds")
-
-error_two_non_linear_diff_shaped_close_clusters_tsne2 <- error_two_non_linear_diff_shaped_close_clusters_tsne2 |>
+error_two_curvy_tsne <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_tsne.rds")
+error_two_curvy_phate <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_phate.rds")
+error_two_curvy_trimap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_trimap.rds")
+error_two_curvy_pacmap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_pacmap.rds")
+error_two_curvy_tsne2 <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_tsne2.rds")
+
+error_two_curvy_tsne2 <- error_two_curvy_tsne2 |>
   mutate(method = "tSNE2")
 
-error_two_non_linear_diff_shaped_close_clusters <- bind_rows(error_two_curvy_umap, 
-                         error_two_non_linear_diff_shaped_close_clusters_tsne,
-                         error_two_non_linear_diff_shaped_close_clusters_phate,
-                         error_two_non_linear_diff_shaped_close_clusters_trimap,
-                         error_two_non_linear_diff_shaped_close_clusters_pacmap,
-                         error_two_non_linear_diff_shaped_close_clusters_tsne2)
-
-error_two_non_linear_diff_shaped_close_clusters <- error_two_non_linear_diff_shaped_close_clusters |>
+error_two_curvy <- bind_rows(error_two_curvy_umap, 
+                         error_two_curvy_tsne,
+                         error_two_curvy_phate,
+                         error_two_curvy_trimap,
+                         error_two_curvy_pacmap,
+                         error_two_curvy_tsne2)
+
+error_two_curvy <- error_two_curvy |>
   mutate(a1 = round(a1, 2)) |>
@@
 ```{r}
-#| label: error-comp-two_non_linear_diff_shaped_close_clusters
-
-error_plot_two_non_linear_diff_shaped_close_clusters <-
-  plot_mse(error_two_non_linear_diff_shaped_close_clusters) +
+#| label: error-comp-two_curvy
+
+error_plot_two_curvy <-
+  plot_mse(error_two_curvy) +
   scale_x_continuous(breaks =
-    sort(unique(error_two_non_linear_diff_shaped_close_clusters$a1))[
+    sort(unique(error_two_curvy$a1))[
       seq(1, length(
-        unique(error_two_non_linear_diff_shaped_close_clusters$a1)), 
+        unique(error_two_curvy$a1)), 
         by = 5)]) +
@@
 
-error_plot_two_non_linear_diff_shaped_close_clusters + wrap_plots(nldr_two_non_linear_diff_shaped_close_clusters1, nldr_two_non_linear_diff_shaped_close_clusters2, nldr_two_non_linear_diff_shaped_close_clusters3,
-                                    nldr_two_non_linear_diff_shaped_close_clusters4, nldr_two_non_linear_diff_shaped_close_clusters5,
-                                    nldr_two_non_linear_diff_shaped_close_clusters6, ncol = 2, 
+error_plot_two_curvy + wrap_plots(nldr_two_curvy1, nldr_two_curvy2, nldr_two_curvy3, 
+                                    nldr_two_curvy4, nldr_two_curvy5,
+                                    nldr_two_curvy6, ncol = 2, 
                                     widths = c(50, 50))
@@
 two_curvy_tsne_scaled_obj <- gen_scaled_data(
-  data = tsne_two_non_linear_diff_shaped_close_clusters)
+  nldr_data = tsne_two_curvy)
 
@@
 
-tsne_two_non_linear_diff_shaped_close_clusters_plt <- tsne_two_curvy_scaled |>
+tsne_two_curvy_plt <- tsne_two_curvy_scaled |>
   ggplot(aes(x = tSNE1,
@@
 hb_obj_tsne_two_curvy2 <- hex_binning(
-  data = tsne_two_curvy_scaled, 
-  bin1 = 24, 
-  r2 = r2_tsne)
-
-a1_2_tsne <- calc_bins_y(
-  bin1 = 24, 
-  r2 = r2_tsne
-)$a1
+  nldr_obj = algo_obj_two_curvy$nldr_obj, 
+  bin1 = 24)
+
+a1_2_tsne <- hb_obj_tsne_two_curvy2$a1
 
@@
 df_bin_centroids4 <- extract_hexbin_centroids(
-  centroids_df = all_centroids_df4, 
-  counts_df = counts_df4) |>
-  filter(drop_empty == FALSE) |>
+  centroids_data = all_centroids_df4, 
+  counts_data = counts_df4) |>
   mutate(b1 = "b1 = 11")
@@
 tsne_data_two_curvy2_with_hb_id <- hb_obj_tsne_two_curvy2$data_hb_id
-df_all_two_curvy4 <- dplyr::bind_cols(training_data_two_curvy, tsne_data_two_curvy2_with_hb_id)
-df_bin_two_curvy4 <- avg_highd_data(data = df_all_two_curvy4)
+df_bin_two_curvy4 <- avg_highd_data(highd_data = training_data_two_curvy,
+                                    scaled_nldr_hexid = tsne_data_two_curvy2_with_hb_id)
 
@@
 tr1_object_two_curvy4 <- tri_bin_centroids(
-  df_bin_centroids4, x = "c_x", y = "c_y")
+  df_bin_centroids4)
 tr_from_to_df_two_curvy4 <- gen_edges(
-  tri_object = tr1_object_two_curvy4)
+  tri_object = tr1_object_two_curvy4,
+  a1 = a1_2_tsne)
 
@@
 
-names(model_df)[7:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
+names(model_df)[(length(names(model_df))-1):length(names(model_df))] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
 
 model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(6 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
+names(model_df)[(length(names(model_df))-1):length(names(model_df))] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
 
@@
 ```{r}
-#| label: best-fit-pacmap-two-curvy
-
-two_curvy_pacmap_scaled_obj <- gen_scaled_data(
-  data = pacmap_two_non_linear_diff_shaped_close_clusters)
-
-pacmap_two_curvy_scaled <- two_curvy_pacmap_scaled_obj$scaled_nldr |>
-  mutate(ID = row_number())
-lim1 <- two_curvy_pacmap_scaled_obj$lim1
-lim2 <- two_curvy_pacmap_scaled_obj$lim2
-r2_pacmap <- diff(lim2)/diff(lim1)
-
-pacmap_two_non_linear_diff_shaped_close_clusters_plt <- pacmap_two_curvy_scaled |>
-  ggplot(aes(x = PaCMAP1,
-             y = PaCMAP2))+
-  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
-  interior_annotation("c1", sc_ltr_pos) +
-  theme(aspect.ratio = 1)
-
-## hexagon binning to have regular hexagons
-hb_obj_pacmap_two_curvy2 <- hex_binning(
-  data = pacmap_two_curvy_scaled, 
-  bin1 = 8, 
-  r2 = r2_pacmap)
-
-a1_2_pacmap <- calc_bins_y(
-  bin1 = 8, 
-  r2 = r2_pacmap
-)$a1
-
-## Data set with all centroids
-all_centroids_df4 <- hb_obj_pacmap_two_curvy2$centroids
-
-## Generate all coordinates of hexagons
-hex_grid4 <- hb_obj_pacmap_two_curvy2$hex_poly
-
-## To obtain the standardise counts within hexbins
-counts_df4 <- hb_obj_pacmap_two_curvy2$std_cts
-df_bin_centroids4 <- extract_hexbin_centroids(
-  centroids_df = all_centroids_df4, 
-  counts_df = counts_df4) |>
-  filter(drop_empty == FALSE) |>
-  mutate(b1 = "b1 = 11")
-
-hex_grid_with_counts_two_curvy4 <- full_join(
-  hex_grid4, 
-  df_bin_centroids4 |> select(hexID, std_counts), 
-  by = c("hex_poly_id" = "hexID"))
-
-pacmap_data_two_curvy2_with_hb_id <- hb_obj_pacmap_two_curvy2$data_hb_id
-df_all_two_curvy4 <- dplyr::bind_cols(training_data_two_curvy, pacmap_data_two_curvy2_with_hb_id)
-df_bin_two_curvy4 <- avg_highd_data(data = df_all_two_curvy4)
-
-df_b_two_curvy4 <- df_bin_two_curvy4 |>
-  dplyr::filter(hexID %in% df_bin_centroids4$hexID) |>
-  dplyr::mutate(type = "model") ## Data with summarized mean
-
-## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
-df_b_two_curvy4 <- df_b_two_curvy4[match(df_bin_centroids4$hexID, df_b_two_curvy4$hexID),] |>
-  dplyr::select(-hexID)
-
-## Triangulate bin centroids
-tr1_object_two_curvy4 <- tri_bin_centroids(
-  df_bin_centroids4, x = "c_x", y = "c_y")
-tr_from_to_df_two_curvy4 <- gen_edges(
-  tri_object = tr1_object_two_curvy4)
-
-trimesh_two_curvy_removed1_pacmap_with_data <- ggplot() +
-  geom_segment(data = tr_from_to_df_two_curvy4,
-               aes(
-                 x = x_from,
-                 y = y_from,
-                 xend = x_to,
-                 yend = y_to),
-               colour = "#000000") +
-  geom_point(data = pacmap_two_curvy_scaled,
-             aes(
-               x = PaCMAP1,
-               y = PaCMAP2
-             ),
-             color = "#e41a1c",
-             alpha = 0.5,
-             size = 0.5) +
-  coord_equal() +
-  interior_annotation("c2", sc_ltr_pos)
-
-```
-
-```{r}
-#| label: best-fit-pacmap-two-curvy-proj
-
-# Apply the scaling
-df_model_data_two_curvy <- bind_rows(data_two_curvy, true_model_two_curvy, df_b_two_curvy4)
-
-scaled_two_curvy <- scale_data_manual(df_model_data_two_curvy, "type") |>
-  as_tibble()
-
-scaled_two_curvy_data <- scaled_two_curvy |>
-  filter(type == "data") |>
-  select(-type)
-
-scaled_two_curvy_data_true_model <- scaled_two_curvy |>
-  filter(type == "true model") |>
-  select(-type)
-
-scaled_two_curvy_data_model <- scaled_two_curvy |>
-  filter(type == "model") |>
-  select(-type)
-
-## First projection
-projection <- cbind(
-  c(-0.08707,-0.06493,0.07017,-0.07349),
-  c(0.03398,0.03735,0.09272,0.01871))
-
-projection_scaled <- projection * 1
-
-projected <- as.matrix(scaled_two_curvy_data) %*% projection_scaled
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-## For fitted model
-projected_model <- as.matrix(scaled_two_curvy_data_model) %*% projection_scaled
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  tr_from_to_df_two_curvy4, 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[7:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(6 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-## For true model
-
-projected_true_model <- as.matrix(scaled_two_curvy_data_true_model) %*% projection_scaled
-
-projected_true_model_df <- projected_true_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-true_model_df_proj <- dplyr::left_join(
-  wireframe_true_model, 
-  projected_true_model_df, 
-  by = c("from" = "ID"))
-
-names(true_model_df_proj)[3:NCOL(true_model_df_proj)] <- paste0(names(projected_true_model_df)[-NCOL(projected_true_model_df)], "_from")
-
-true_model_df_proj <- dplyr::left_join(true_model_df_proj, projected_true_model_df, by = c("to" = "ID"))
-names(true_model_df_proj)[(2 + NCOL(projected_true_model_df)):NCOL(true_model_df_proj)] <- paste0(names(projected_true_model_df)[-NCOL(projected_true_model_df)], "_to")
-
-axes_obj <- gen_axes(  
-  proj = projection * 5,
-  limits = 0.15,
-  axis_pos_x = -0.1,
-  axis_pos_y = -0.1,
-  axis_labels = names(scaled_two_curvy_data),
-  threshold = 0.01)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
-
-two_curvy_proj_pacmap_model <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#000000") + #31a354
-  geom_point(
-    size = 0.5,
-    alpha = 0.5,
-    color = "#e41a1c") +
-  geom_segment(
-    data=axes,
-    aes(x=x1, y=y1, xend=x2, yend=y2),
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 4) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.126, 0.125)) + 
-  ylim(c(-0.126, 0.125)) +
-  interior_annotation("c3", sc_ltr_pos) +
-  theme(
-    legend.position = "none"
-  )
-
-two_curvy_proj_pacmap_all_model <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_segment(
-    data = true_model_df_proj, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#2b8cbe") +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#000000") +
-  geom_point(
-    size = 0.5,
-    alpha = 0.5,
-    color = "#e41a1c") +
-  geom_segment(
-    data=axes,
-    aes(x=x1, y=y1, xend=x2, yend=y2),
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 4) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.126, 0.125)) + 
-  ylim(c(-0.126, 0.125)) +
-  interior_annotation("c4", sc_ltr_pos) +
-  theme(
-    legend.position = "none"
-  )
-```
-
-```{r}
 #| label: best-fit-umap-tsne-pacamp
@@
 
-nldr_two_curvy_original2 + tsne_two_non_linear_diff_shaped_close_clusters_plt +
-  pacmap_two_non_linear_diff_shaped_close_clusters_plt +
+nldr_two_curvy_original2 + tsne_two_curvy_plt +
+  pacmap_two_curvy_plt +
   trimesh_two_curvy_removed1_with_data + trimesh_two_curvy_removed1_tsne_with_data +
@@
   highd_data = training_data_gau, 
-  nldr_data = tsne_data_gau, 
+  nldr_data = umap_data_gau, 
   bin1 = num_bins_x_gau1, 
@@
 
-tsne_gau_scaled <- algo_obj_gau1$nldr_obj$scaled_nldr
+umap_gau_scaled <- algo_obj_gau1$nldr_obj$scaled_nldr
 tr_from_to_df_gau1 <- algo_obj_gau1$trimesh_data
@@
   highd_data = training_data_gau, 
-  nldr_data = tsne_data_gau, 
+  nldr_data = pacmap_data_gau, 
   bin1 = num_bins_x_gau1, 
@@
 
-tsne_gau_scaled <- algo_obj_gau1$nldr_obj$scaled_nldr
+pacmap_gau_scaled <- algo_obj_gau1$nldr_obj$scaled_nldr
 tr_from_to_df_gau1 <- algo_obj_gau1$trimesh_data
@@
 
-tsne_one_c_shaped <- read_rds(file = "data/one_c_shaped_dens_structure/one_c_shaped_dens_structure_tsne_perplexity_52.rds")
+tsne_one_c_shaped <- read_rds(file = "data/one_c_shaped_dens_structure/one_c_shaped_dens_structure_tsne_perplexity_52.rds") |>
+  mutate(ID = row_number())
 
@@
 pbmc_scaled_obj <- gen_scaled_data(
-  data = umap_pbmc)
+  nldr_data = umap_pbmc)
 umap_pbmc_scaled <- pbmc_scaled_obj$scaled_nldr

### 2025-05-12 scripts/additional_functions.R

@@ -214,56 +214,57 @@ scale_data_manual <- function(data, type_col) {
 #
 # }
 #
-# # Plot MSE
-#
-# plot_mse <- function(error_df) {
-#
-#   ggplot(error_df,
-#          aes(x = a1,
-#              y = sqrt(MSE),
-#              colour = method)) +
-#     geom_point(size = 0.8) +
-#     geom_line(linewidth = 0.3) +
-#     #scale_y_log10() +
-#     ylab("RMSE") +
-#     xlab(expression(paste("binwidth (", a[1], ")"))) +
-#     theme_minimal() +
-#     theme(panel.border = element_rect(fill = 'transparent'),
-#           plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
-#           axis.ticks.x = element_line(),
-#           axis.ticks.y = element_line(),
-#           legend.position = "none",
-#           axis.text.x = element_text(size = 7),
-#           axis.text.y = element_text(size = 7),
-#           axis.title.x = element_text(size = 7),
-#           axis.title.y = element_text(size = 7))
-#
-# }
-#
-# # creating Standardization function
-# standardize = function(x){
-#   z <- (x - mean(x)) / sd(x)
-#   return( z)
-# }
-#
-# # Plot digits in MNIST
-#
-# plot_digit_img <- function(digit_df, palette, title_text, hjust, ncol = 4) {
-#
-#   ggplot(data = digit_df, aes(x, y, fill = value)) +
-#     geom_tile() +
-#     facet_wrap(~ instance, ncol = ncol) +
-#     ggtitle(title_text) +
-#     coord_fixed() +
-#     scale_fill_continuous_sequential(palette = palette) +
-#     theme(strip.background = element_blank(),
-#           strip.text.x = element_blank(),
-#           legend.position = "none",
-#           plot.title.position = "plot",
-#           plot.title = element_text(hjust = hjust,
-#                                     size = 15))
-#
-# }
+
+# Plot MSE
+
+plot_mse <- function(error_df) {
+
+  ggplot(error_df,
+         aes(x = a1,
+             y = sqrt(MSE),
+             colour = method)) +
+    geom_point(size = 0.8) +
+    geom_line(linewidth = 0.3) +
+    #scale_y_log10() +
+    ylab("RMSE") +
+    xlab(expression(paste("binwidth (", a[1], ")"))) +
+    theme_minimal() +
+    theme(panel.border = element_rect(fill = 'transparent'),
[…]

### 2025-05-12 paper.qmd
Changed the bin density for two_curvy

@@
   q = 0.1, 
-  benchmark_highdens = 5)
+  benchmark_highdens = 1)
 

### 2025-05-12 paper.qmd
Updated Figure 3

@@
                  x2min = 0,
-                 x2max = 1.6)
+                 x2max = diff(algo_obj_two_curvy$nldr_obj$lim2)/diff(algo_obj_two_curvy$nldr_obj$lim1))
 

### 2025-05-12 paper.qmd
Updated Figure 10

@@
   filter(MSE == min(MSE)) |>
-  ungroup()
+  ungroup() |>
+  mutate(method = factor(method,
+                         levels = c("UMAP", "tSNE", "tSNE2", "PHATE", "TriMAP", "PaCMAP")))
 ```

### 2025-05-12 paper.qmd
Updated Figure 13

@@
 five_gau_proj_tsne_model1 + five_gau_proj_umap_model1 + five_gau_proj_pacmap_model1 + 
-  five_gau_proj_tsne_model2 +
-  five_gau_proj_umap_model2 + five_gau_proj_pacmap_model2 +
-  plot_layout(guides = "collect", nrow = 3) &
+  #five_gau_proj_tsne_model2 +
+  #five_gau_proj_umap_model2 + five_gau_proj_pacmap_model2 +
+  plot_layout(guides = "collect", nrow = 2) &
   theme(legend.position='none')
@@
   scale_color_manual(
-    values=c('#e41a1c','#377eb8',"#ff7f00", 
-             '#984ea3', '#999999', '#a65628')) +
+    values=c('#a65628', '#984ea3', '#999999', 
+             '#377eb8', '#ff7f00', '#e41a1c')) +
   theme(aspect.ratio=1.5)

### 2025-05-12 paper.qmd
Updated Projection

@@
 projection <- cbind(
-  c(0.07108,0.19751,-0.21229,0.26341),
-  c(-0.23494,0.27965,-0.01430,-0.15782))
+  c(0.00660,-0.34942,0.14548,-0.12289),
+  c(-0.18758,0.15398,0.30233,-0.08999))
 

### 2025-05-12 scripts/five_gau_clusters/1_five_gaussian_cluster_data_emb.R

@@ -34,19 +34,19 @@ vert2 <- simplex(p2)$points
 # df[2001:3000,] <- df[2001:3000,] + matrix(rep(vert2[3,], 1000), ncol=4, byrow=T)
 # df[3001:4000,] <- df[3001:4000,] + matrix(rep(vert2[4,], 1000), ncol=4, byrow=T)
 # df[4001:5000,] <- df[4001:5000,] + matrix(rep(vert2[5,], 1000), ncol=4, byrow=T)
-df1 <- rmvnorm(1000, mean = rep(0, 4), sigma = diag(4) * 0.05)/6
+df1 <- rmvnorm(1000, mean = rep(0, 4), sigma = diag(4) * 0.5)/6
 df1 <- df1 + matrix(rep(vert2[1,], 1000), ncol=4, byrow=T)
 
-df2 <- rmvnorm(1000, mean = rep(0, 4), sigma = diag(4) * 0.3)/6
+df2 <- rmvnorm(1000, mean = rep(0, 4), sigma = diag(4) * 0.5)/6
 df2 <- df2 + matrix(rep(vert2[2,], 1000), ncol=4, byrow=T)
 
 df3 <- rmvnorm(1000, mean = rep(0, 4), sigma = diag(4) * 0.5)/6
 df3 <- df3 + matrix(rep(vert2[3,], 1000), ncol=4, byrow=T)
 
-df4 <- rmvnorm(1000, mean = rep(0, 4), sigma = diag(4) * 0.02)/6
+df4 <- rmvnorm(1000, mean = rep(0, 4), sigma = diag(4) * 0.5)/6
 df4 <- df4 + matrix(rep(vert2[4,], 1000), ncol=4, byrow=T)
 
-df5 <- rmvnorm(1000, mean = rep(0, 4), sigma = diag(4) * 0.6)/6
+df5 <- rmvnorm(1000, mean = rep(0, 4), sigma = diag(4) * 0.5)/6
 df5 <- df5 + matrix(rep(vert2[5,], 1000), ncol=4, byrow=T)
 
 df1 <- df1 |>

### 2025-05-12 paper.qmd
Removed unnecessary code

@@
 
-```{r}
-#| label: best-fit-tsne-two-curvy
-
-two_curvy_tsne_scaled_obj <- gen_scaled_data(
-  nldr_data = tsne_two_curvy)
-
-tsne_two_curvy_scaled <- two_curvy_tsne_scaled_obj$scaled_nldr |>
-  mutate(ID = row_number())
-lim1 <- two_curvy_tsne_scaled_obj$lim1
-lim2 <- two_curvy_tsne_scaled_obj$lim2
-r2_tsne <- diff(lim2)/diff(lim1)
-
-tsne_two_curvy_plt <- tsne_two_curvy_scaled |>
-  ggplot(aes(x = tSNE1,
-             y = tSNE2))+
-  geom_point(alpha=0.1, size=1, colour='#a65628') +
-  interior_annotation("b1", sc_ltr_pos) +
-  theme(aspect.ratio = 1)
-
-## hexagon binning to have regular hexagons
-hb_obj_tsne_two_curvy2 <- hex_binning(
-  nldr_obj = algo_obj_two_curvy$nldr_obj, 
-  bin1 = 24)
-
-a1_2_tsne <- hb_obj_tsne_two_curvy2$a1
-
-## Data set with all centroids
-all_centroids_df4 <- hb_obj_tsne_two_curvy2$centroids
-
-## Generate all coordinates of hexagons
-hex_grid4 <- hb_obj_tsne_two_curvy2$hex_poly
-
-## To obtain the standardise counts within hexbins
-counts_df4 <- hb_obj_tsne_two_curvy2$std_cts
-df_bin_centroids4 <- extract_hexbin_centroids(
-  centroids_data = all_centroids_df4, 
-  counts_data = counts_df4) |>
-  mutate(b1 = "b1 = 11")
-
-hex_grid_with_counts_two_curvy4 <- full_join(
-  hex_grid4, 
-  df_bin_centroids4 |> select(hexID, std_counts), 
-  by = c("hex_poly_id" = "hexID"))
-
-tsne_data_two_curvy2_with_hb_id <- hb_obj_tsne_two_curvy2$data_hb_id
-df_bin_two_curvy4 <- avg_highd_data(highd_data = training_data_two_curvy,
-                                    scaled_nldr_hexid = tsne_data_two_curvy2_with_hb_id)
-
-df_b_two_curvy4 <- df_bin_two_curvy4 |>
-  dplyr::filter(hexID %in% df_bin_centroids4$hexID) |>
-  dplyr::mutate(type = "model") ## Data with summarized mean
-
-## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
-df_b_two_curvy4 <- df_b_two_curvy4[match(df_bin_centroids4$hexID, df_b_two_curvy4$hexID),] |>
-  dplyr::select(-hexID)
-
-## Triangulate bin centroids
-tr1_object_two_curvy4 <- tri_bin_centroids(
-  df_bin_centroids4)
-tr_from_to_df_two_curvy4 <- gen_edges(
-  tri_object = tr1_object_two_curvy4,
-  a1 = a1_2_tsne)
-
-trimesh_two_curvy_removed1_tsne_with_data <- ggplot() +
-  geom_segment(data = tr_from_to_df_two_curvy4,
-               aes(
-                 x = x_from,
-                 y = y_from,
-                 xend = x_to,
-                 yend = y_to),
-               colour = "#000000") +
-  geom_point(data = tsne_two_curvy_scaled,
-             aes(
-               x = tSNE1,
-               y = tSNE2
-             ),
-             color = "#a65628",
-             alpha = 0.5,
-             size = 0.5) +
-  coord_equal() +
-  interior_annotation("b2", sc_ltr_pos)
-
-```
-
-```{r}
-#| label: best-fit-tsne-two-curvy-proj
-
-# Apply the scaling
-df_model_data_two_curvy <- bind_rows(data_two_curvy, true_model_two_curvy, df_b_two_curvy4)
-
-scaled_two_curvy <- scale_data_manual(df_model_data_two_curvy, "type") |>
-  as_tibble()
-
-scaled_two_curvy_data <- scaled_two_curvy |>
-  filter(type == "data") |>
-  select(-type)
-
-scaled_two_curvy_data_true_model <- scaled_two_curvy |>
-  filter(type == "true model") |>
-  select(-type)
-
-scaled_two_curvy_data_model <- scaled_two_curvy |>
-  filter(type == "model") |>
-  select(-type)
-
-## First projection
-projection <- cbind(
-  c(-0.08707,-0.06493,0.07017,-0.07349),
-  c(0.03398,0.03735,0.09272,0.01871))
-
-projection_scaled <- projection * 1
-
-projected <- as.matrix(scaled_two_curvy_data) %*% projection_scaled
-
-projected_df <- projected |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-## For fitted model
-projected_model <- as.matrix(scaled_two_curvy_data_model) %*% projection_scaled
-
-projected_model_df <- projected_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-model_df <- dplyr::left_join(
-  tr_from_to_df_two_curvy4, 
-  projected_model_df, 
-  by = c("from" = "ID"))
-
-names(model_df)[(length(names(model_df))-1):length(names(model_df))] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")
-
-model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
-names(model_df)[(length(names(model_df))-1):length(names(model_df))] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")
-
-## For true model
-
-projected_true_model <- as.matrix(scaled_two_curvy_data_true_model) %*% projection_scaled
-
-projected_true_model_df <- projected_true_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-true_model_df_proj <- dplyr::left_join(
-  wireframe_true_model, 
-  projected_true_model_df, 
-  by = c("from" = "ID"))
-
-names(true_model_df_proj)[3:NCOL(true_model_df_proj)] <- paste0(names(projected_true_model_df)[-NCOL(projected_true_model_df)], "_from")
-
-true_model_df_proj <- dplyr::left_join(true_model_df_proj, projected_true_model_df, by = c("to" = "ID"))
-names(true_model_df_proj)[(2 + NCOL(projected_true_model_df)):NCOL(true_model_df_proj)] <- paste0(names(projected_true_model_df)[-NCOL(projected_true_model_df)], "_to")
-
-axes_obj <- gen_axes(  
-  proj = projection * 5,
-  limits = 0.15,
-  axis_pos_x = -0.1,
-  axis_pos_y = -0.1,
-  axis_labels = names(scaled_two_curvy_data),
-  threshold = 0.01)
-
-axes <- axes_obj$axes
-circle <- axes_obj$circle
-
-two_curvy_proj_tsne_model <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#000000") + #31a354
-  geom_point(
-    size = 0.5,
-    alpha = 0.5,
-    color = "#a65628") +
-  geom_segment(
-    data=axes,
-    aes(x=x1, y=y1, xend=x2, yend=y2),
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 4) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.126, 0.125)) + 
-  ylim(c(-0.126, 0.125)) +
-  interior_annotation("b3", sc_ltr_pos) +
-  theme(
-    legend.position = "none"
-  )
-
-two_curvy_proj_tsne_all_model <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_segment(
-    data = true_model_df_proj, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#2b8cbe") +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    color = "#000000") +
-  geom_point(
-    size = 0.5,
-    alpha = 0.5,
-    color = "#a65628") +
-  geom_segment(
-    data=axes,
-    aes(x=x1, y=y1, xend=x2, yend=y2),
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2), 
-    label=rownames(axes),
-    colour="grey50",
-    size = 4) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.126, 0.125)) + 
-  ylim(c(-0.126, 0.125)) +
-  interior_annotation("b4", sc_ltr_pos) +
-  theme(
-    legend.position = "none"
-  )
-```
-
-```{r}
-#| label: best-fit-umap-tsne-pacamp
-#| fig-cap: "Best fit for two C-shaped clusters data using UMAP (a1-a4) ($a_1$ = 0.09) and tSNE (b1-b4) ($a_1$ = 0.16) with the same number of non-empty bins ($m$ = 39).
-Both models show twist in $7\\text{-}D$ (a3, a4, b3, b4).
-Also, both NLDR methods are unable to accurately capture the width of the curvilinear clusters."
-#| fig-width: 15
-#| fig-height: 20
-#| out-width: 90%
-#| eval: false
-
-nldr_two_curvy_original2 + tsne_two_curvy_plt +
-  pacmap_two_curvy_plt +
-  trimesh_two_curvy_removed1_with_data + trimesh_two_curvy_removed1_tsne_with_data +
-trimesh_two_curvy_removed1_pacmap_with_data +
-  two_curvy_proj_umap_model2 +
-  two_curvy_proj_tsne_model +
-  two_curvy_proj_pacmap_model +
-  two_curvy_proj_umap_all_model +
-  two_curvy_proj_tsne_all_model +
-  two_curvy_proj_pacmap_all_model +
-  plot_layout(ncol = 3) +
-  plot_annotation(tag_levels = list(c("(a)", "(b)", "(c)")))
-```
-
 ## Curiosities

### 2025-05-12 paper.qmd
Fixed the cross reference

@@
 
-The visualization of the selected layout in the $9\text{-}D$ shows edges between the clusters (@fig-mnist-tri-proj (b2)).
+The visualization of the selected layout in the $9\text{-}D$ shows edges between the clusters (@fig-model-pbmc-author-proj (b2)).
 This supports the presence of small separations between clusters.
-Additionally, the data points are not uniformly distributed across the clusters, resulting in dense areas (@fig-mnist-tri-proj (b2)).
-Furthermore, the clusters shows non-linear shapes (@fig-mnist-tri-proj (b3)).
+Additionally, the data points are not uniformly distributed across the clusters, resulting in dense areas (@fig-model-pbmc-author-proj (b2)).
+Furthermore, the clusters shows non-linear shapes (@fig-model-pbmc-author-proj (b3)).
 
@@
 Before further analysis, PCA was used to preprocess the data, where the first $10$ principal components, explaining 83% of the total variation, were selected.
-The objective is to select a reasonable \gD{} layout, representing the non-linear structure of the digit 1 dataset in $10\text{-}D$ (@fig-mnist-tri-proj (a)).
+The objective is to select a reasonable \gD{} layout, representing the non-linear structure of the digit 1 dataset in $10\text{-}D$ (@fig-clust-mnist (a)).
 
@@
 By visualizing the model generated for tSNE in $10\text{-}D$ helps to assess the \gD{} layout.
-As shown in @fig-mnist-tri-proj (a1), the model provides the evidence for the non-linear structure of the digit 1 data in $10\text{-}D$.
+As shown in @fig-clust-mnist (a1), the model provides the evidence for the non-linear structure of the digit 1 data in $10\text{-}D$.
 The model shows some quirks.
-The model's twisted pattern provides evidence for the $10\text{-}D$ data structure, which is not observed by \gD{} layout (@fig-mnist-tri-proj (a2)).
+The model's twisted pattern provides evidence for the $10\text{-}D$ data structure, which is not observed by \gD{} layout (@fig-clust-mnist (a2)).
 
@@
 Video of the langevitour animation is available at <>."
+#| eval: false
 
@@
 #| fig-cap: "The tSNE layout of the MNIST digit 1 data shows a big non-linear cluster (grey) and a small cluster (orange) located very close to the one corner of the big cluster in \\gD{} (a).
+The MNIST digit 1 data ($n =  7877$ and $p = 10$) has a non-linear structure in $10\\text{-}D$.
+Two \\gD{} projections from a tour on $10\\text{-}D$ reveal that the closeness of the clusters in $10\\text{-}D$ and the twisted pattern of the model fit with tSNE ($a_1 = 0.041$, $b = 659/585 \\text{ } (30, \\text{ }48)$).
+Video of the langevitour animation is available at <>.
 The brushing feature in the linked plots helps in visualizing the closeness of the small cluster to the big cluster.

### 2025-05-12 paper.qmd
Updated discussion

@@
 The published results of UMAP with default (hyper-)parameter setting by @chen2024 were evaluated.
-The data showed three clusters, two of which were close.
+The data showed three close clusters.
 But the published layout showed three well-separated clusters.
-Therefore, by changing the (hyper-)parameter, we found a better layout with UMAP, which is more representative the data.
-Additionally, the model discovers non-uniform data distribution and non-linear structures within the clusters that are not visible in the UMAP layout, demonstrating the ability of our model in uncovering hidden data characteristics.
+Therefore, by changing the NLDR method to tSNE with default (hyper-)parameters, we found a better layout, which is more representative the data.
+Additionally, the model discovers non-uniform data distribution and non-linear structures within the clusters that are not visible in the \gD{} layout, demonstrating the ability of our model in uncovering hidden data characteristics.
 Also, as explained in Appendix, with recently introduced approach, scDEED also proved the layout suggested by our layout is better than the published one.

### 2025-05-12 paper.qmd
Updated some texts

@@
 
-To understand how well the model fits the points—whether it fits well, works better in some positions, or fails to match the overall pattern—it is important to link and brush the points with high model error in the \pD{} error plot, \gD{} layout, and the generated model overlay on the data in \pD{}.
+To understand how well the model fits the points whether it fits well, works better in some positions, or fails to match the overall pattern.
+It is important to link and brush the points with high model error in the \pD{} error plot, \gD{} layout, and the generated model overlay on the data in \pD{}.
 

### 2025-05-13 paper.qmd
Updated Figure 15 colors

@@
   geom_rect(aes(xmin = 0.82, xmax = 0.92, ymin = 0.38, ymax = 0.58),
-               fill = "transparent", color = "#000000", linewidth = 0.7) +
+               fill = "transparent", color = "#000000", linewidth = 1) +
   geom_rect(aes(xmin = 0.9, xmax = 1, ymin = 1.25, ymax = 1.45),
-               fill = "transparent", color = "#e41a1c", linewidth = 0.7)  +
+               fill = "transparent", color = "#830c15", linewidth = 1)  +
   geom_rect(aes(xmin = 0.35, xmax = 0.45, ymin = 0.55, ymax = 0.7),
-               fill = "transparent", color = "#4daf4a", linewidth = 0.7)  +
+               fill = "transparent", color = "#004616", linewidth = 1)  +
   geom_rect(aes(xmin = 0.7, xmax = 0.8, ymin = 0.25, ymax = 0.4),
-               fill = "transparent", color = "#984ea3", linewidth = 0.7) +
+               fill = "transparent", color = "#4c389a", linewidth = 1) +
   theme(legend.position = "none")

### 2025-05-13 paper.qmd
Updated Figure 9 text size

@@
   is_category = FALSE) +
-  interior_annotation(label = "a3", position = c(0.08, 0.9), cex = 1.5) +
+  interior_annotation(label = "a3", position = c(0.08, 0.9), cex = 2) +
   theme(aspect.ratio = 1,
@@
                       position = c(0.95, 0.95),
-                      cex = 1.5) +
+                      cex = 2) +
   theme(
@@
   is_category = FALSE) +
-  interior_annotation(label = "b3", position = c(0.08, 0.9), cex = 1.5) +
+  interior_annotation(label = "b3", position = c(0.08, 0.9), cex = 2) +
   theme(aspect.ratio = 1,
@@
   is_category = FALSE) +
-  interior_annotation(label = "c3", position = c(0.08, 0.9), cex = 1.5) +
+  interior_annotation(label = "c3", position = c(0.08, 0.9), cex = 2) +
   theme(aspect.ratio = 1,

### 2025-05-13 paper.qmd
Put the intro image to first page
Updated Figure 3 text

@@
 However, the representation generated can vary dramatically from method to method, and with different choices of parameters or random seeds made using the same method (@fig-NLDR-variety).
-The dilemma for the analyst is then, **which representation to use**.
-The choice might result in different procedures used in the downstream analysis, or different inferential conclusions.
-The research described here provides new visual tools to aid with this decision.
 
@@
 ```
+
+The dilemma for the analyst is then, **which representation to use**.
+The choice might result in different procedures used in the downstream analysis, or different inferential conclusions.
+The research described here provides new visual tools to aid with this decision.
 
@@
 
-hex_grid_temp45 <- hex_grid_temp |> 
-  filter(hex_poly_id == 45)
+hex_grid_temp40 <- hex_grid_temp |> 
+  filter(hex_poly_id == 40)
 
@@
              label = expression(a[2]))
-l <- tibble(x = hex_grid_temp45$x[2],
-            xend = hex_grid_temp45$x[3],
-            y = hex_grid_temp45$y[2],
-            yend = hex_grid_temp45$y[3],
+l <- tibble(x = hex_grid_temp40$x[2],
+            xend = hex_grid_temp40$x[3],
+            y = hex_grid_temp40$y[2],
+            yend = hex_grid_temp40$y[3],
             label = expression(l))
@@
                      label = "(0,0)", 
-             hjust=-0.1, vjust=-0.3) + 
+             hjust=-0.1, vjust=-0.3, size = 8) + 
     annotate("text", x=d_rect$x1max - rect_adj$x1, 
@@
                      label = "(0,1)", 
-             hjust=1.1, vjust=-0.3) + 
+             hjust=1.1, vjust=-0.3, size = 8) + 
     annotate("text", x=d_rect$x1min - rect_adj$x1, 
@@
                         list(0, y[2][max]),")")), 
-            hjust=-0.1, vjust=1.2) + 
+            hjust=-0.1, vjust=1.2, size = 8) + 
     geom_segment(data=d_rect, aes(
@@
     annotate("text", x=0.5, y=-0.36, 
-             label = expression(r[1]), color = "black") +
+             label = expression(r[1]), color = "black", size = 8) +
     geom_segment(data=d_rect, aes(
@@
     annotate("text", x=-0.3, y=0.4, 
-             label = expression(r[2]), color = "black") +
+             label = expression(r[2]), color = "black", size = 8) +
     geom_segment(data = a1, aes(
@@
              color = "black",
-             vjust = 1.2) +
+             vjust = 1.2, size = 8) +
     geom_segment(data = a2, aes(
@@
              label = expression(a[2]), 
-             color = "black", hjust=-0.2) +
+             color = "black", hjust=-0.2, size = 8) +
     annotate("text", x=-0.18, y=-0.24, 
       label = expression(group("(", list(s[1], s[2]), ")")),
-      color = "black") +
+      color = "black", size = 8) +
   geom_segment(data = l, aes(
@@
              label = expression(l), 
-             color = "black", hjust=-0.2) +
+             color = "black", hjust=-0.2, size = 8) +
   coord_equal()

### 2025-05-13 paper.qmd
Updated Figure labels

@@
 nldr_two_curvy <- nldr_two_curvy_original +
-  interior_annotation("a", sc_ltr_pos, cex = 1.5) +
+  interior_annotation("a", sc_ltr_pos, cex = 2) +
   xlim(sc_xlims) +
@@
   xlim(sc_xlims) + ylim(sc_ylims) +
-  interior_annotation("b", sc_ltr_pos, cex = 1.5)
+  interior_annotation("b", sc_ltr_pos, cex = 2)
 
@@
   xlim(sc_xlims) + ylim(sc_ylims) +
-  interior_annotation("c", sc_ltr_pos, cex = 1.5) 
+  interior_annotation("c", sc_ltr_pos, cex = 2) 
 ```
@@
   xlim(sc_xlims) + ylim(sc_ylims) +
-  interior_annotation("d", sc_ltr_pos, cex = 1.5)
+  interior_annotation("d", sc_ltr_pos, cex = 2)
 

### 2025-05-13 paper.qmd
Updated Figure 4

@@
     is_category = FALSE) +
-  interior_annotation(label = "b", position = sc_ltr_pos, cex = 1) +
+  interior_annotation(label = "b", position = sc_ltr_pos, cex = 1.2) +
   theme(aspect.ratio = 1,
@@
   is_category = FALSE) +
-  interior_annotation(label = "c", position = sc_ltr_pos, cex = 1) +
+  interior_annotation(label = "c", position = sc_ltr_pos, cex = 1.2) +
   theme(aspect.ratio = 1) #+

### 2025-05-13 paper.qmd
Updated Figure 6

@@
     na.value = "#ffffff", option = "E") +
-  interior_annotation("b", sc_ltr_pos, cex = 1.5) + 
+  interior_annotation("b", sc_ltr_pos, cex = 2) + 
   xlim(sc_xlims) + ylim(sc_ylims) 
@@
   xlim(sc_xlims) + ylim(sc_ylims) +
-  interior_annotation("a", sc_ltr_pos, cex = 1.5)
+  interior_annotation("a", sc_ltr_pos, cex = 2)
 
@@
   xlim(sc_xlims) + ylim(sc_ylims) +
-  interior_annotation("c", sc_ltr_pos, cex = 1.5)
+  interior_annotation("c", sc_ltr_pos, cex = 2)
 ```

### 2025-05-13 paper.qmd
Updated Figure 7

@@
   geom_point(size = 1) +
-  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_umap$a1, 2)))[c(2, 4, 8, 11, 13, 14, 15, 16)]) +
+  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_umap$a1, 2)))[c(1, 4, 7, 11, 13, 14, 15, 16)]) +
   labs(x = expression(paste("binwidth (", a[1], ")")), y = "MSE") +
@@
         axis.ticks.y = element_line(),
-        axis.title = element_text(size = 15),
-        axis.text = element_text(size = 12))
+        axis.title = element_text(size = 18),
+        axis.text = element_text(size = 15))
 
@@
   # scale_y_continuous(breaks = sort(unique(error_two_curvy$b_non_empty))[-3]) +
-  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_umap$a1, 2)))[c(2, 4, 8, 11, 13, 14, 15, 16)]) +
+  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_umap$a1, 2)))[c(1, 4, 7, 11, 13, 14, 15, 16)]) +
   labs(x = expression(paste("binwidth (", a[1], ")")), y = paste("Relative proportion density")) +
@@
         axis.ticks.y = element_line(),
-        axis.title = element_text(size = 15),
-axis.text = element_text(size = 12))
+        axis.title = element_text(size = 18),
+axis.text = element_text(size = 15))
 
@@
   # scale_y_continuous(breaks = sort(unique(error_two_curvy$b_non_empty))[-3]) +
-  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_umap$a1, 2)))[c(2, 4, 8, 11, 13, 14, 15, 16)]) +
+  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_umap$a1, 2)))[c(1, 4, 7, 11, 13, 14, 15, 16)]) +
   labs(x = expression(paste("binwidth (", a[1], ")")), y = expression(paste("proportion of non-empty bins ", bgroup("(", frac(m, b), ")")))) +
@@
         axis.ticks.y = element_line(),
-        axis.title = element_text(size = 15),
-axis.text = element_text(size = 12))
+        axis.title = element_text(size = 18),
+axis.text = element_text(size = 15))
 ```
@@
         legend.position = "none",
-        axis.title = element_text(size = 15),
-axis.text = element_text(size = 12))
+        axis.title = element_text(size = 18),
+axis.text = element_text(size = 15))
 ```

### 2025-05-13 paper.qmd
Updated Figure 14

@@
                axis_param = list(limits = 1, 
-                                 axis_scaled = 1, 
+                                 axis_scaled = 1.1, 
                                  axis_pos_x = -0.92, 
                                  axis_pos_y = -0.92, 
-                                 threshold = 0.06))
+                                 threshold = 0.065))
 
@@
   plot_limits = c(-1.1, 0.9), 
-  axis_text_size = 3.5, 
+  axis_text_size = 4, 
   is_category = TRUE) +
@@
     plot_limits = c(-1.1, 0.9), 
-    axis_text_size = 4.5, 
+    axis_text_size = 5, 
     is_category = TRUE) +
@@
                axis_param = list(limits = 1, 
-                                 axis_scaled = 1, 
+                                 axis_scaled = 1.1, 
                                  axis_pos_x = -0.83, 
                                  axis_pos_y = -0.83, 
-                                 threshold = 0.06))
+                                 threshold = 0.065))
 
@@
     plot_limits = c(-1, 1), 
-    axis_text_size = 4.5, 
+    axis_text_size = 5, 
     is_category = TRUE) +

### 2025-05-13 paper.qmd
Updated MSE with RMSE in asessing plots

@@
 ```{r}
-#| fig-cap: "Assessing which of the 6 NLDR layouts on the two non-linear clusters data is the better representation using MSE for varying binwidth ($a_1$).
+#| fig-cap: "Assessing which of the 6 NLDR layouts on the two non-linear clusters data is the better representation using RMSE for varying binwidth ($a_1$).
 Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f).
@@
 Following their pre-processing, UMAP was performed using nine principal components.
-@fig-NLDR-variety (a) is the reproduction of the published plot, with colour indicating the clustering results overlaid on the NLDR layout.
+@fig-NLDR-variety (a) is the reproduction of the published plot.
 The question is whether this accurately represents the cluster structure in the data.
@@
 ```{r}
-#| fig-cap: "Assessing which of the 8 NLDR layouts on the PBMC3k data  (shown in @fig-NLDR-variety) is the better representation using MSE for varying binwidth ($a_1$).
+#| fig-cap: "Assessing which of the 8 NLDR layouts on the PBMC3k data  (shown in @fig-NLDR-variety) is the better representation using RMSE for varying binwidth ($a_1$).
 Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-h).
@@
 #| fig-cap: "Compare the published \\gD{} layout (@fig-pbmc-mse a) and the \\gD{} layout selected (@fig-pbmc-mse e) by MSE plot (@fig-pbmc-mse) from the UMAP and tSNE with default hyper-parameters.
-The PBMC3k data ($n =  2622$) has three clusters in $9\\text{-}D$, where two of them being close and one separated from the others.
+The PBMC3k data ($n =  2622$) has three close clusters in $9\\text{-}D$.
 Two \\gD{} projection from a tour on $9\\text{-}D$ of the model fit with @fig-pbmc-mse a ($a_1 = 0.06$, $b = 616/86 (22, 28)$) shows three-well separated clusters with big separations.
@@
 ```{r}
-#| fig-cap: "Assessing which of the 6 NLDR layouts on the MNIST digit 1 data is the better representation using MSE for varying binwidth ($a_1$).
+#| fig-cap: "Assessing which of the 6 NLDR layouts on the MNIST digit 1 data is the better representation using RMSE for varying binwidth ($a_1$).
 Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f).

### 2025-05-13 paper.qmd
Updated the Five Gaussian cluster projection

@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 23
+num_bins_x_gau1 <- 18
 
@@
 projection <- cbind(
-  c(0.00660,-0.34942,0.14548,-0.12289),
-  c(-0.18758,0.15398,0.30233,-0.08999))
+  c(0.10394,0.08379,-0.17868,0.32960),
+  c(0.35111,-0.01137,-0.09657,-0.16018))
 
 proj_obj1 <- get_projection(projection = projection, 
-               proj_scale = 2, 
+               proj_scale = 2.5, 
                highd_data = scaled_gau_data, 
@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 75
+num_bins_x_gau1 <- 63
 
@@
 ## Compute hexbin parameters
-num_bins_x_gau1 <- 33
+num_bins_x_gau1 <- 28
 

### 2025-05-13 paper.qmd
Updated Figure 8

@@
              "#a65628",'#636363', '#984ea3')) +
-  scale_y_log10() + 
   theme(aspect.ratio = 1.5)

### 2025-05-13 paper.qmd
Changed MSE equation to RMSE
Updated Figure 5

@@
 library(conflicted)
+library(ggbeeswarm)
 
@@
   scale_colour_continuous_sequential(palette = "YlOrRd", n_interp = 20) +
+  xlab("tSNE1") +
+  ylab("tSNE2") +
+  theme_bw() +
   theme(
@@
 
-error_plot_two_curvy_hist <- 
-  ggplot(error_df_two_curvy_abs) +
-  geom_histogram(aes(x=sqrt_row_wise_total_error, y=..density..)) +
-  geom_density(aes(x=sqrt_row_wise_total_error, y=..density..), colour="red") +
-  #xlab(expression(group("|", e[hj], "|"))) +
+# error_plot_two_curvy_hist <- 
+#   ggplot(error_df_two_curvy_abs) +
+#   geom_histogram(aes(x=sqrt_row_wise_total_error, y=..density..)) +
+#   geom_density(aes(x=sqrt_row_wise_total_error, y=..density..), colour="red") +
+#   #xlab(expression(group("|", e[hj], "|"))) +
+#   xlab(expression(e[hj])) +
+#   ylab("") +
+#   theme_bw()
+
+error_plot_two_curvy_quasi <- ggplot(
+  error_df_two_curvy_abs, 
+  aes(x = sqrt_row_wise_total_error, 
+      y = 0, 
+      colour = sqrt_row_wise_total_error)) +
+  geom_quasirandom(alpha = 0.5) +
+  scale_colour_continuous_sequential(palette = "YlOrRd", n_interp = 20) +
   xlab(expression(e[hj])) +
   ylab("") +
-  theme_bw() 
+  theme_bw() +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank(),
+        legend.position = "none")
 ```
@@
 
-The error is computed by taking the squared \pD{} Euclidean distance, corresponding to computing the mean squared error (MSE) as:
-
-$$\frac{1}{n}\sum_{h = 1}^{b}\sum_{i = 1}^{n_h}\sum_{j = 1}^{p} (\mathbfit{x}_{hij} - C^{(p)}_{hj})^2$${#eq-equation1}
+The error is computed by taking the squared \pD{} Euclidean distance, corresponding to computing the root mean squared error (RMSE) as:
+
+$$\sqrt{\frac{1}{n}\sum_{h = 1}^{b}\sum_{i = 1}^{n_h}\sum_{j = 1}^{p} (\mathbfit{x}_{hij} - C^{(p)}_{hj})^2}$${#eq-equation1}
 
 where $n$ is the number of observations, $b$ is the number of bins, $n_h$ is the number of observations in $h^{th}$ bin, $p$ is the number of variables, $\mathbfit{x}_{hij}$ is the $j^{th}$ dimensional data of $i^{th}$ observation in $h^{th}$ hexagon.
-We can consider $e_{hj} = \sum_{j = 1}^{p} (\mathbfit{x}_{hij} - C^{(p)}_{hj})^2$ to be the residual for each observation.
+We can consider $e_{hj} = \sqrt{\sum_{j = 1}^{p} (\mathbfit{x}_{hij} - C^{(p)}_{hj})^2}$ to be the residual for each observation.
 
@@
 #| fig-cap: "The $4\\text{-}D$ model error in \\gD{} layout.
-Color indicates square root of total error, dark colour indicating high error and light indicates low error.
+Color indicates error ($e_{hj}$), dark colour indicating high error and light indicates low error.
 Most large errors are distributed near the sparse end of the non-linear cluster with dense corner."
@@
 
-error_plot_two_curvy + error_plot_two_curvy_hist +
+error_plot_two_curvy + error_plot_two_curvy_quasi +
   plot_layout(ncol=2)

### 2025-05-13 paper.qmd
Updated Figure 5
Updated Figure 7

@@
                      aes(x = a1, 
-                         y = MSE)) +
+                         y = sqrt(MSE))) +
   geom_vline(xintercept = 0.09,
-           color = "#d95f02", linetype=2, 
+           color = "#d95f02", linetype="dashed", 
            linewidth=1) +
@@
              linewidth=1) +
-  geom_vline(xintercept = 0.03, linetype=2,
+  geom_vline(xintercept = 0.03, linetype="dotted",
              color = "#1b9e77", linewidth=1) +
@@
   scale_x_continuous(breaks = sort(unique(round(error_two_curvy_umap$a1, 2)))[c(1, 4, 7, 11, 13, 14, 15, 16)]) +
-  labs(x = expression(paste("binwidth (", a[1], ")")), y = "MSE") +
-  ggtitle("(a)") +
+  labs(x = expression(paste("binwidth (", a[1], ")")), y = "RMSE") +
+  #ggtitle("(a)") +
+  interior_annotation("a", position = c(0.28, 0.15), cex = 2) +
   theme_minimal() +
@@
   scale_x_continuous(breaks = sort(unique(round(error_two_curvy_umap$a1, 2)))[c(1, 4, 7, 11, 13, 14, 15, 16)]) +
-  labs(x = expression(paste("binwidth (", a[1], ")")), y = expression(paste("proportion of non-empty bins ", bgroup("(", frac(m, b), ")")))) +
+  labs(x = expression(paste("binwidth (", a[1], ")")), y = expression("prop. non-empty bins " * bgroup("(", m / b, ")"))) +
   ggtitle("(b)") +
@@
 Most large errors are distributed near the sparse end of the non-linear cluster with dense corner."
-#| out-width: 100%
+#| out-width: 80%
 #| fig-pos: H

### 2025-05-13 paper.qmd
Added Figure labels

@@
   labs(x = expression(paste("binwidth (", a[1], ")")), y = "RMSE") +
-  #ggtitle("(a)") +
-  interior_annotation("a", position = c(0.28, 0.15), cex = 2) +
+  interior_annotation("a", position = c(0.9, 0.15), cex = 2) +
   theme_minimal() +
@@
   labs(x = expression(paste("binwidth (", a[1], ")")), y = paste("Relative proportion density")) +
-  ggtitle("(d)") +
+  interior_annotation("d", position = c(0.9, 0.9), cex = 2) +
   theme_minimal() +
@@
   labs(x = expression(paste("binwidth (", a[1], ")")), y = expression("prop. non-empty bins " * bgroup("(", m / b, ")"))) +
-  ggtitle("(b)") +
+  interior_annotation("b", position = c(0.9, 0.15), cex = 2) +
   theme_minimal() +
@@
                      aes(x = benchmark_rm_lwd, 
-                         y = MSE,
+                         y = sqrt(MSE),
                          color = bin1)) + 
@@
   xlab("threshold to remove low-density hexagons") +
-  ylab("MSE") +
-  ggtitle("(c)") +
+  ylab("RMSE") +
+  interior_annotation("c", position = c(0.1, 0.9), cex = 2) +
   theme_minimal() +

### 2025-05-13 paper.qmd
Changed the lines in a, b, and c

@@
   geom_vline(xintercept = 0.09,
-           color = "#d95f02", linetype=2, 
+           color = "#d95f02", linetype="dashed", 
            linewidth=1) +

### 2025-05-13 paper.qmd
Updated Figure 7

@@
                          y = sqrt(MSE),
-                         color = bin1)) + 
+                         color = bin1,
+                         linetype = bin1)) + 
   geom_point(
@@
     ) + 
-  # geom_vline(xintercept = benchmark1, linetype="solid", 
-  #            color = "#bdbdbd", linewidth=1, alpha = 0.5) +
-  # scale_x_continuous("standardized bin count", 
-  #        transform = "reverse") + 
   scale_x_continuous(breaks = sort(unique(error_rm_two_curvy$benchmark_rm_lwd))[c(1, 3, 5, 7, 9, 11, 13, 15, 17)]) +
   scale_color_manual(values=c("#d95f02", "#1f78b4", "#1b9e77")) +
+  scale_linetype_manual(values = c("dashed", "solid", "dotted")) +
   xlab("threshold to remove low-density hexagons") +
@@
         axis.ticks.y = element_line(),
-        legend.position = "none",
+        legend.position = "bottom",
         axis.title = element_text(size = 18),
@@
   mse_two_curvy_lwd + prop_dens_a1 +
-  plot_layout(ncol=2)
+  plot_layout(ncol=2, guides = "collect")
 ```

### 2025-05-13 paper.qmd
Updated Figure 7

@@
 library(ggbeeswarm)
+library(scales)
 
@@
                      aes(x = benchmark_rm_lwd, 
-                         y = sqrt(MSE),
-                         color = bin1,
-                         linetype = bin1)) + 
+                         y = RMSE,
+                         color = a1,
+                         linetype = a1)) + 
   geom_point(
@@
   scale_x_continuous(breaks = sort(unique(error_rm_two_curvy$benchmark_rm_lwd))[c(1, 3, 5, 7, 9, 11, 13, 15, 17)]) +
-  scale_color_manual(values=c("#d95f02", "#1f78b4", "#1b9e77")) +
-  scale_linetype_manual(values = c("dashed", "solid", "dotted")) +
+  scale_color_manual(values=c("#d95f02", "#1f78b4", "#1b9e77"),
+                     labels = parse_format()) +
+  scale_linetype_manual(values = c("dashed", "solid", "dotted"),
+                        labels = parse_format()) +
   xlab("threshold to remove low-density hexagons") +
@@
         legend.position = "bottom",
+        legend.title = element_blank(),
+        legend.text = element_text(size = 18),
         axis.title = element_text(size = 18),
@@
 #| fig-cap: "Various plots to help assess best hexagon bin parameters (a, b, d) and thresholds to remove low-density bins (c).
-Both (a) and (c) show MSE, against binwidth ($a_1$) and threshold.
-A good benchmark value for these parameters is when the MSE drops and then flattens out.
-Three binwidth choices were made: $0.03$, $0.05$, and $0.09$ to investigate.
+Both (a) and (c) show RMSE, against binwidth ($a_1$) and threshold.
+A good benchmark value for these parameters is when the RMSE drops and then flattens out.
+Three binwidth choices were made: $0.03$ (orange dashed), $0.05$ (blue solid), and $0.09$ (green dotted) to investigate.
 As the binwidth increases, the proportion of non-empty bins also increases (b).
@@
 Binwidth $0.05$ is chosen as the initial best binwidth for further analysis.
-There is no need to remove the low-density hexagons because as shown in (b), there is no considerable drop in MSE."
+There is no need to remove the low-density hexagons because as shown in (b), there is no considerable drop in RMSE."
 #| label: fig-param-two-curvy

### 2025-05-13 scripts/two_non_linear_diff_shaped_close_clusters/09_gen_rm_lwd_mse.R

@@ -42,6 +42,7 @@ for (benchmark_rm_lwd in benchmark_rm_hex_vec) {
     model_highd = df_bin_two_curvy_high_dens,
     highd_data = data_two_curvy) |>
     mutate(benchmark_rm_lwd = round(benchmark_rm_lwd, 2),
+           a1 = paste0("a[1] == ", round(two_curvy_model1$hb_obj$a1, 2)),
            bin1 = two_curvy_model1$hb_obj$bins[1],
            bin2 = two_curvy_model1$hb_obj$bins[2],
            b = bin1 * bin2,
@@ -83,6 +84,7 @@ for (benchmark_rm_lwd in benchmark_rm_hex_vec) {
     model_highd = df_bin_two_curvy_high_dens,
     highd_data = data_two_curvy) |>
     mutate(benchmark_rm_lwd = round(benchmark_rm_lwd, 2),
+           a1 = paste0("a[1] == ", round(two_curvy_model2$hb_obj$a1, 2)),
            bin1 = two_curvy_model2$hb_obj$bins[1],
            bin2 = two_curvy_model2$hb_obj$bins[2],
            b = bin1 * bin2,
@@ -124,8 +126,10 @@ for (benchmark_rm_lwd in benchmark_rm_hex_vec) {
     model_highd = df_bin_two_curvy_high_dens,
     highd_data = data_two_curvy) |>
     mutate(benchmark_rm_lwd = round(benchmark_rm_lwd, 2),
+           a1 = paste0("a[1] == ", round(two_curvy_model3$hb_obj$a1, 2)),
            bin1 = two_curvy_model3$hb_obj$bins[1],
            bin2 = two_curvy_model3$hb_obj$bins[2],
+           b = bin1 * bin2,
            b_non_empty = NROW(df_bin_centroids_two_curvy_high_dens),
            mean_counts = sum(df_bin_centroids_two_curvy_high_dens$bin_counts)/NROW(df_bin_centroids_two_curvy_high_dens))
 

### 2025-05-13 paper.qmd
Updated Figure 15

@@
   geom_rect(aes(xmin = 0.9, xmax = 1, ymin = 1.25, ymax = 1.45),
-               fill = "transparent", color = "#830c15", linewidth = 1)  +
+               fill = "transparent", color = "#375cb1", linewidth = 1)  +
   geom_rect(aes(xmin = 0.35, xmax = 0.45, ymin = 0.55, ymax = 0.7),
@@
 
-right_top_img <- plot_digit_img(pixels_gathered_within, palette = "Reds 2", title_text = "(a)", hjust = 0.31)
+right_top_img <- plot_digit_img(pixels_gathered_within, palette = "Blues 2", title_text = "(a)", hjust = 0.31)
 

### 2025-05-13 paper.qmd
Started updating Figure 15
Removed whitespace around the plot

--- 2025-05-13 paper.qmd
@@
                fill = "transparent", color = "#4c389a", linewidth = 1) +
-  theme(legend.position = "none")
+  theme(legend.position = "none",
+        plot.margin = margin(0, 0, 0, 0))
 ```
@@
 
-right_top_img <- plot_digit_img(pixels_gathered_within, palette = "Reds 2", title_text = "(a)", hjust = 0.31)
+right_top_img <- plot_digit_img(pixels_gathered_within, palette = "Blues 2", title_text = "(a)", hjust = 0.31) + theme(plot.margin = margin(0, 0, 0, 0))
 
@@
 
-middle_img <- plot_digit_img(pixels_gathered_within, palette = "Greens 3", title_text = "(b)", hjust = 0.31)
+middle_img <- plot_digit_img(pixels_gathered_within, palette = "Greens 3", title_text = "(b)", hjust = 0.31) + theme(plot.margin = margin(0, 0, 0, 0))
 
@@
 
-right_bottom_img <- plot_digit_img(pixels_gathered_within, palette = "Purples 2", title_text = "(c)", hjust = 0.31)
+right_bottom_img <- plot_digit_img(pixels_gathered_within, palette = "Purples 2", title_text = "(c)", hjust = 0.31) + theme(plot.margin = margin(0, 0, 0, 0))
 
@@
 
-imge_error_sample_outside <- plot_digit_img(pixels_gathered_outside, palette = "Grays", title_text = "(d)", hjust = 0.31)
+imge_error_sample_outside <- plot_digit_img(pixels_gathered_outside, palette = "Grays", title_text = "(d)", hjust = 0.31) + theme(plot.margin = margin(0, 0, 0, 0))
 
@@
 
-imge_error_sample_inside <- plot_digit_img(pixels_gathered_inside, palette = "Oranges", title_text = "(e)", hjust = 0, ncol = 9)
+imge_error_sample_inside <- plot_digit_img(pixels_gathered_inside, palette = "Oranges", title_text = "(e)", hjust = 0, ncol = 9) + theme(plot.margin = margin(0, 0, 0, 0))
 

### 2025-05-13 paper.qmd
Updated Figure 13

@@
   interior_annotation("a", c(0.08, 0.93)) +
-  theme(aspect.ratio = 1)
+  theme(aspect.ratio = 1,
+        plot.margin = margin(0, 0, 0, 0))
 
@@
   interior_annotation("f", c(0.08, 0.93)) +
-  theme(aspect.ratio = 1)
+  theme(aspect.ratio = 1,
+        plot.margin = margin(0, 0, 0, 0))
 
@@
   interior_annotation("b") +
-  theme(aspect.ratio = 1)
+  theme(aspect.ratio = 1,
+        plot.margin = margin(0, 0, 0, 0))
 
@@
   interior_annotation("c", position = c(0.94, 0.92)) +
-  theme(aspect.ratio = 1)
+  theme(aspect.ratio = 1,
+        plot.margin = margin(0, 0, 0, 0))
 
@@
   interior_annotation("d", position = c(0.86, 0.92)) +
-  theme(aspect.ratio = 1)
+  theme(aspect.ratio = 1,
+        plot.margin = margin(0, 0, 0, 0))
 
@@
   interior_annotation("e") +
-  theme(aspect.ratio = 1)
+  theme(aspect.ratio = 1,
+        plot.margin = margin(0, 0, 0, 0))
 ```
@@
              '#377eb8', '#ff7f00', '#e41a1c')) +
-  theme(aspect.ratio=1.5)
+  theme(aspect.ratio=1.5,
+        plot.margin = margin(0, 0, 0, 0))
 

### 2025-05-13 paper.qmd
Updated error figures

@@
   geom_point(alpha=0.1, size=1, colour='#a65628') +
-  interior_annotation("a")
+  interior_annotation("a") + 
+  theme(plot.margin = margin(0, 0, 0, 0))
 
@@
   geom_point(alpha=0.1, size=1, colour='#999999') +
-  interior_annotation("b", c(0.08, 0.9))
+  interior_annotation("b", c(0.08, 0.9)) + 
+  theme(plot.margin = margin(0, 0, 0, 0))
 
@@
   geom_point(alpha=0.1, size=1, colour='#e41a1c') +
-  interior_annotation("c")
+  interior_annotation("c") + 
+  theme(plot.margin = margin(0, 0, 0, 0))
 
@@
   geom_point(alpha=0.1, size=1, colour='#377eb8') +
-  interior_annotation("d")
+  interior_annotation("d") + 
+  theme(plot.margin = margin(0, 0, 0, 0))
 
@@
   geom_point(alpha=0.1, size=1, colour='#4daf4a') +
-  interior_annotation("e")
+  interior_annotation("e") + 
+  theme(plot.margin = margin(0, 0, 0, 0))
 
@@
   geom_point(alpha=0.1, size=1, colour='#ff7f00') +
-  interior_annotation("f")
+  interior_annotation("f") + 
+  theme(plot.margin = margin(0, 0, 0, 0))
 
@@
   geom_point(alpha=0.1, size=1, colour='#984ea3') +
-  interior_annotation("g")
+  interior_annotation("g") + 
+  theme(plot.margin = margin(0, 0, 0, 0))
 
@@
   geom_point(alpha=0.1, size=1, colour='#f781bf') +
-  interior_annotation("h")
+  interior_annotation("h") + 
+  theme(plot.margin = margin(0, 0, 0, 0))
 ```
@@
   interior_annotation("b", c(0.08, 0.93)) +
-  theme(aspect.ratio = 1)
+  theme(aspect.ratio = 1,
+        plot.margin = margin(0, 0, 0, 0))
 
@@
   interior_annotation("c") +
-  theme(aspect.ratio = 1)
+  theme(aspect.ratio = 1,
+        plot.margin = margin(0, 0, 0, 0))
 
@@
   interior_annotation("d") +
-  theme(aspect.ratio = 1)
+  theme(aspect.ratio = 1,
+        plot.margin = margin(0, 0, 0, 0))
 
@@
   interior_annotation("e") +
-  theme(aspect.ratio = 1)
+  theme(aspect.ratio = 1, 
+        plot.margin = margin(0, 0, 0, 0))
 ```
@@
              "#a65628",'#636363', '#984ea3')) +
-  theme(aspect.ratio = 1.5)
+  theme(aspect.ratio = 1.5,
+        plot.margin = margin(0, 0, 0, 0))
 ```
@@
   scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 13, 17, 21, 26)]) +
-  scale_color_manual(values=c('#999999','#a65628','#e41a1c','#377eb8','#4daf4a','#ff7f00','#984ea3','#f781bf'))
+  scale_color_manual(values=c('#999999','#a65628','#e41a1c','#377eb8','#4daf4a','#ff7f00','#984ea3','#f781bf')) + 
+  theme(plot.margin = margin(0, 0, 0, 0))
 

### 2025-05-14 paper.qmd
Updated Figure 15 label colour

@@
   theme(legend.position = "none",
-        plot.margin = margin(0, 0, 0, 0))
+        plot.margin = margin(0, 0, 0, 0),
+        aspect.ratio = 1)
 ```
@@
 
-right_top_img <- plot_digit_img(pixels_gathered_within, palette = "Blues 2", title_text = "(a)", hjust = 0.31) + theme(plot.margin = margin(0, 0, 0, 0))
+right_top_img <- plot_digit_img(pixels_gathered_within, palette = "Blues 2", title_text = "a", hjust = 0.31) + theme(plot.margin = margin(0, 0, 0, 0))
 
@@
 
-middle_img <- plot_digit_img(pixels_gathered_within, palette = "Greens 3", title_text = "(b)", hjust = 0.31) + theme(plot.margin = margin(0, 0, 0, 0))
+middle_img <- plot_digit_img(pixels_gathered_within, palette = "Greens 3", title_text = "b", hjust = 0.31) + theme(plot.margin = margin(0, 0, 0, 0))
 
@@
 
-right_bottom_img <- plot_digit_img(pixels_gathered_within, palette = "Purples 2", title_text = "(c)", hjust = 0.31) + theme(plot.margin = margin(0, 0, 0, 0))
+right_bottom_img <- plot_digit_img(pixels_gathered_within, palette = "Purples 2", title_text = "c", hjust = 0.31) + theme(plot.margin = margin(0, 0, 0, 0))
 
@@
 
-imge_error_sample_outside <- plot_digit_img(pixels_gathered_outside, palette = "Grays", title_text = "(d)", hjust = 0.31) + theme(plot.margin = margin(0, 0, 0, 0))
+imge_error_sample_outside <- plot_digit_img(pixels_gathered_outside, palette = "Grays", title_text = "d", hjust = 0.31) + theme(plot.margin = margin(0, 0, 0, 0))
 
 ##### inside
-img_error_inside <- c(3179, 7363, 7650, 5241, 5173, 6981, 7228, 6673, 4636)
+img_error_inside <- c(3179, 7363, 7650) #, 5241, 5173, 6981, 7228, 6673, 4636
 
@@
 
-imge_error_sample_inside <- plot_digit_img(pixels_gathered_inside, palette = "Oranges", title_text = "(e)", hjust = 0, ncol = 9) + theme(plot.margin = margin(0, 0, 0, 0))
+imge_error_sample_inside <- plot_digit_img(pixels_gathered_inside, palette = "Oranges", title_text = "e", hjust = 0, ncol = 9) + theme(plot.margin = margin(0, 0, 0, 0))
 
@@
 # Display the final combined plot
-final_layout
+#final_layout
+full_layout/imge_error_sample_inside + plot_layout(widths = c(1, 1), 
+                                                   heights = c(1, 0.2)) 
 ```

### 2025-05-14 paper.qmd
Added label inside the plot

@@
 
-right_top_img <- plot_digit_img(pixels_gathered_within, palette = "Blues 2", title_text = "a", hjust = 0.31) + theme(plot.margin = margin(0, 0, 0, 0))
+right_top_img <- plot_digit_img(pixels_gathered_within, palette = "Blues 2", title_text = "a", hjust = 0.17) + theme(plot.margin = margin(0, 0, 0, 0))
 
@@
 
-middle_img <- plot_digit_img(pixels_gathered_within, palette = "Greens 3", title_text = "b", hjust = 0.31) + theme(plot.margin = margin(0, 0, 0, 0))
+middle_img <- plot_digit_img(pixels_gathered_within, palette = "Greens 3", title_text = "b", hjust = 0.17) + theme(plot.margin = margin(0, 0, 0, 0))
 
@@
 
-right_bottom_img <- plot_digit_img(pixels_gathered_within, palette = "Purples 2", title_text = "c", hjust = 0.31) + theme(plot.margin = margin(0, 0, 0, 0))
+right_bottom_img <- plot_digit_img(pixels_gathered_within, palette = "Purples 2", title_text = "c", hjust = 0.17) + theme(plot.margin = margin(0, 0, 0, 0))
 
@@
 
-imge_error_sample_outside <- plot_digit_img(pixels_gathered_outside, palette = "Grays", title_text = "d", hjust = 0.31) + theme(plot.margin = margin(0, 0, 0, 0))
+imge_error_sample_outside <- plot_digit_img(pixels_gathered_outside, palette = "Grays", title_text = "d", hjust = 0.17) + theme(plot.margin = margin(0, 0, 0, 0))
 
@@
 
-imge_error_sample_inside <- plot_digit_img(pixels_gathered_inside, palette = "Oranges", title_text = "e", hjust = 0, ncol = 9) + theme(plot.margin = margin(0, 0, 0, 0))
+imge_error_sample_inside <- plot_digit_img(pixels_gathered_inside, palette = "Oranges", title_text = "e", hjust = 0.17, ncol = 3) + theme(plot.margin = margin(0, 0, 0, 0))
 

### 2025-05-14 paper.qmd
Trying to remove the whitespace

--- 2025-05-14 paper.qmd
@@
   theme(legend.position = "none",
-        plot.margin = margin(0, 0, 0, 0),
-        aspect.ratio = 1)
+        aspect.ratio = 1,
+        plot.margin = margin(0, 0, 0, 0))
 ```
@@
 
-right_top_img <- plot_digit_img(pixels_gathered_within, palette = "Blues 2", title_text = "a", hjust = 0.31) + theme(plot.margin = margin(0, 0, 0, 0))
+right_top_img <- plot_digit_img(
+  pixels_gathered_within, palette = "Blues 2", title_text = "a")
 
@@
 
-middle_img <- plot_digit_img(pixels_gathered_within, palette = "Greens 3", title_text = "b", hjust = 0.31) + theme(plot.margin = margin(0, 0, 0, 0))
+middle_img <- plot_digit_img(
+  pixels_gathered_within, palette = "Greens 3", title_text = "b")
 
@@
 
-right_bottom_img <- plot_digit_img(pixels_gathered_within, palette = "Purples 2", title_text = "c", hjust = 0.31) + theme(plot.margin = margin(0, 0, 0, 0))
+right_bottom_img <- plot_digit_img(
+  pixels_gathered_within, palette = "Purples 2", title_text = "c")
 
@@
 
-imge_error_sample_outside <- plot_digit_img(pixels_gathered_outside, palette = "Grays", title_text = "d", hjust = 0.31) + theme(plot.margin = margin(0, 0, 0, 0))
+imge_error_sample_outside <- plot_digit_img(
+  pixels_gathered_outside, palette = "Grays", title_text = "d")
 
@@
 
-imge_error_sample_inside <- plot_digit_img(pixels_gathered_inside, palette = "Oranges", title_text = "e", hjust = 0, ncol = 9) + theme(plot.margin = margin(0, 0, 0, 0))
+imge_error_sample_inside <- plot_digit_img(
+  pixels_gathered_inside, palette = "Oranges", title_text = "e", ncol = 3)
 
@@
 #| fig-pos: H
-#| fig-width: 10
-#| fig-height: 7
+#| fig-width: 12
+#| fig-height: 6
 #| out-width: 90%
@@
 # Define your current layout
-top_layout <- right_top_img / middle_img / right_bottom_img / imge_error_sample_outside
+# top_layout <- right_top_img / middle_img / right_bottom_img / imge_error_sample_outside
+
+top_layout <- wrap_plots(right_top_img, middle_img, right_bottom_img, imge_error_sample_outside, nrow = 4) + theme(plot.margin = margin(0, 0, 0, 0))
 full_layout <- highd_error_2d_mnist + top_layout
@@
 #final_layout
-full_layout/imge_error_sample_inside + plot_layout(widths = c(1, 1), 
-                                                   heights = c(1, 0.2)) 
+full_layout
+
+# /imge_error_sample_inside + plot_layout(widths = c(1, 1), 
+#                                                    heights = c(1, 0.2))
 ```

### 2025-05-14 scripts/additional_functions.R

@@ -77,7 +77,7 @@ standardize = function(x){
 
 # Plot digits in MNIST
 
-plot_digit_img <- function(digit_df, palette, title_text, hjust, ncol = 3) {
+plot_digit_img <- function(digit_df, palette, title_text, ncol = 3) {
 
   # Add a label column only to the first instance
   digit_df <- digit_df |>
@@ -95,7 +95,8 @@ plot_digit_img <- function(digit_df, palette, title_text, hjust, ncol = 3) {
     scale_fill_continuous_sequential(palette = palette) +
     theme(strip.background = element_blank(),
           strip.text.x = element_blank(),
-          legend.position = "none")
+          legend.position = "none",
+          plot.margin = margin(0, 0, 0, 0))
 
 }
 

### 2025-05-14 paper.qmd
Removed the spacing

@@
      legend.key.height = unit(0.25, 'cm'),
-     legend.key.width = unit(0.25, 'cm')
+     legend.key.width = unit(0.25, 'cm'),
+     plot.margin = margin(0, 0, 0, 0)
    )
@@
   geom_point(alpha=0.1, size=1, colour='#a65628') +
-  interior_annotation("a") + 
-  theme(plot.margin = margin(0, 0, 0, 0))
+  interior_annotation("a")
 
@@
   geom_point(alpha=0.1, size=1, colour='#999999') +
-  interior_annotation("b", c(0.08, 0.9)) + 
-  theme(plot.margin = margin(0, 0, 0, 0))
+  interior_annotation("b", c(0.08, 0.9))
 
@@
   geom_point(alpha=0.1, size=1, colour='#e41a1c') +
-  interior_annotation("c") + 
-  theme(plot.margin = margin(0, 0, 0, 0))
+  interior_annotation("c")
 
@@
   geom_point(alpha=0.1, size=1, colour='#377eb8') +
-  interior_annotation("d") + 
-  theme(plot.margin = margin(0, 0, 0, 0))
+  interior_annotation("d")
 
@@
   geom_point(alpha=0.1, size=1, colour='#4daf4a') +
-  interior_annotation("e") + 
-  theme(plot.margin = margin(0, 0, 0, 0))
+  interior_annotation("e")
 
@@
   geom_point(alpha=0.1, size=1, colour='#ff7f00') +
-  interior_annotation("f") + 
-  theme(plot.margin = margin(0, 0, 0, 0))
+  interior_annotation("f")
 
@@
   geom_point(alpha=0.1, size=1, colour='#984ea3') +
-  interior_annotation("g") + 
-  theme(plot.margin = margin(0, 0, 0, 0))
+  interior_annotation("g")
 
@@
   geom_point(alpha=0.1, size=1, colour='#f781bf') +
-  interior_annotation("h") + 
-  theme(plot.margin = margin(0, 0, 0, 0))
+  interior_annotation("h") 
 ```
@@
   interior_annotation("a", c(0.08, 0.93)) +
-  theme(aspect.ratio = 1,
-        plot.margin = margin(0, 0, 0, 0))
+  theme(aspect.ratio = 1)
 
@@
   interior_annotation("f", c(0.08, 0.93)) +
-  theme(aspect.ratio = 1,
-        plot.margin = margin(0, 0, 0, 0))
+  theme(aspect.ratio = 1)
 
@@
   interior_annotation("b", c(0.08, 0.93)) +
-  theme(aspect.ratio = 1,
-        plot.margin = margin(0, 0, 0, 0))
+  theme(aspect.ratio = 1)
 
@@
   interior_annotation("c") +
-  theme(aspect.ratio = 1,
-        plot.margin = margin(0, 0, 0, 0))
+  theme(aspect.ratio = 1)
 
@@
   interior_annotation("d") +
-  theme(aspect.ratio = 1,
-        plot.margin = margin(0, 0, 0, 0))
+  theme(aspect.ratio = 1)
 
@@
   interior_annotation("e") +
-  theme(aspect.ratio = 1, 
-        plot.margin = margin(0, 0, 0, 0))
+  theme(aspect.ratio = 1)
 ```
@@
              "#a65628",'#636363', '#984ea3')) +
-  theme(aspect.ratio = 1.5,
-        plot.margin = margin(0, 0, 0, 0))
+  theme(aspect.ratio = 1.5)
 ```
@@
   scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 13, 17, 21, 26)]) +
-  scale_color_manual(values=c('#999999','#a65628','#e41a1c','#377eb8','#4daf4a','#ff7f00','#984ea3','#f781bf')) + 
-  theme(plot.margin = margin(0, 0, 0, 0))
+  scale_color_manual(values=c('#999999','#a65628','#e41a1c','#377eb8','#4daf4a','#ff7f00','#984ea3','#f781bf'))
 
@@
   interior_annotation("b") +
-  theme(aspect.ratio = 1,
-        plot.margin = margin(0, 0, 0, 0))
+  theme(aspect.ratio = 1)
 
@@
   interior_annotation("c", position = c(0.94, 0.92)) +
-  theme(aspect.ratio = 1,
-        plot.margin = margin(0, 0, 0, 0))
+  theme(aspect.ratio = 1)
 
@@
   interior_annotation("d", position = c(0.86, 0.92)) +
-  theme(aspect.ratio = 1,
-        plot.margin = margin(0, 0, 0, 0))
+  theme(aspect.ratio = 1)
 
@@
   interior_annotation("e") +
-  theme(aspect.ratio = 1,
-        plot.margin = margin(0, 0, 0, 0))
+  theme(aspect.ratio = 1)
 ```
@@
              '#377eb8', '#ff7f00', '#e41a1c')) +
-  theme(aspect.ratio=1.5,
-        plot.margin = margin(0, 0, 0, 0))
+  theme(aspect.ratio=1.5)
 
@@
 
-top_layout <- wrap_plots(right_top_img, middle_img, right_bottom_img, imge_error_sample_outside, nrow = 4) + theme(plot.margin = margin(0, 0, 0, 0))
-full_layout <- highd_error_2d_mnist + top_layout
+top_layout <- wrap_plots(
+  right_top_img,
+  middle_img,
+  right_bottom_img,
+  imge_error_sample_outside,
+  ncol = 1,
+  heights = c(1, 1, 1, 1)
+) + plot_layout() + plot_annotation(theme = theme(plot.margin = margin(0, 0, 0, 0)))
+
+# Combine with main plot
+full_layout <- highd_error_2d_mnist + top_layout +
+  plot_layout(widths = c(1, 1)) +
+  plot_annotation(theme = theme(plot.margin = margin(0, 0, 0, 0)))
 

### 2025-05-14 paper.qmd
Found the way to arrange images

@@
         aspect.ratio = 1,
-        plot.margin = margin(0, 0, 0, 0))
+        plot.margin = margin(0, 0, 0, 0),
+        panel.spacing = unit(0, "pt"),
+        axis.ticks.length = unit(0, "pt"),
+        strip.text = element_blank(),   # Optional, if you're using facets
+        axis.title = element_blank()   # Optional
+        )
 ```
@@
 
-img_right_top <- c(868, 1465, 1787)
+img_right_top <- c(868, 1465, 1787, 1788, 1789)
 
@@
 
-img_middle <- c(130, 2091, 7795)
+img_middle <- c(130, 2091, 7795, 7796, 7797)
 
@@
 
-img_right_bottom <- c(2035, 7708, 7710)
+img_right_bottom <- c(2035, 7708, 7710, 7711, 7712)
 
@@
 
-img_error_outside <- c(999, 3650, 4916)
+img_error_outside <- c(999, 3650, 4916, 4917, 4918)
 
@@
 ##### inside
-img_error_inside <- c(3179, 7363, 7650) #, 5241, 5173, 6981, 7228, 6673, 4636
+img_error_inside <- c(5173, 6981, 7228, 6673, 3179) #, 6673, 3179, 7363, 7650, 5241, 4636
 
@@
 imge_error_sample_inside <- plot_digit_img(
-  pixels_gathered_inside, palette = "Oranges", title_text = "e", ncol = 3)
+  pixels_gathered_inside, palette = "Oranges", title_text = "e")
 
@@
 #| fig-pos: H
-#| fig-width: 12
-#| fig-height: 6
-#| out-width: 90%
+#| fig-width: 10
+#| fig-height: 10
+#| out-width: 100%
 
 # Define your current layout
-# top_layout <- right_top_img / middle_img / right_bottom_img / imge_error_sample_outside
-
-top_layout <- wrap_plots(
-  right_top_img,
-  middle_img,
-  right_bottom_img,
-  imge_error_sample_outside,
-  ncol = 1,
-  heights = c(1, 1, 1, 1)
-) + plot_layout() + plot_annotation(theme = theme(plot.margin = margin(0, 0, 0, 0)))
-
-# Combine with main plot
-full_layout <- highd_error_2d_mnist + top_layout +
-  plot_layout(widths = c(1, 1)) +
-  plot_annotation(theme = theme(plot.margin = margin(0, 0, 0, 0)))
-
-# Add the new full-width bottom row
-final_layout <- full_layout / imge_error_sample_inside + plot_layout(heights = c(1, 0.2))
-
-# Display the final combined plot
-#final_layout
-full_layout
-
-# /imge_error_sample_inside + plot_layout(widths = c(1, 1), 
-#                                                    heights = c(1, 0.2))
+top_layout <- right_top_img + middle_img + right_bottom_img + imge_error_sample_outside + imge_error_sample_inside + plot_layout(nrow = 5)
+
+final_layout <- wrap_plots(highd_error_2d_mnist, top_layout)
+
+final_layout
 ```

### 2025-05-14 paper.qmd
Updated Figure 15
Updated Figure 11

@@
 
-free(error_plot_pbmc) + wrap_plots(nldr1c, nldr2c, nldr3c, nldr4c,
-           nldr5c, nldr6c, nldr7c, nldr8c, ncol = 2)
+design <- "
+AABB
+AABB
+AABB
+"
+
+# Put error plot in area A, and wrap_plots in area B
+error_plot_pbmc + 
+  wrap_plots(nldr1c, nldr2c, nldr3c, nldr4c,
+             nldr5c, nldr6c, nldr7c, nldr8c, ncol = 2) +
+  plot_layout(design = design)
 ```
@@
 
-img_right_top <- c(868, 1465, 1787, 1788, 1789)
+img_right_top <- c(868, 1465, 1787, 2006, 4314)
 
@@
 
-img_middle <- c(130, 2091, 7795, 7796, 7797)
+img_middle <- c(130, 2091, 7795, 163, 77)
 
@@
 
-img_right_bottom <- c(2035, 7708, 7710, 7711, 7712)
+img_right_bottom <- c(2035, 7708, 7710, 4134, 5006)
 
@@
 
-img_error_outside <- c(999, 3650, 4916, 4917, 4918)
+img_error_outside <- c(999, 3650, 4916, 2964, 5622)
 

### 2025-05-14 paper.qmd
Updated Figure 10

@@
 
-error_plot_one_curvy_hist_selected <- ggplot(error_df_one_curvy_abs, 
+error_df_one_curvy_abs_selected1 <- error_df_one_curvy_abs |>
+  dplyr::filter(error_cat_n == "selected")
+
+error_df_one_curvy_abs_deselected1 <- error_df_one_curvy_abs |>
+  dplyr::filter(error_cat_n == "deselected")
+
+error_plot_one_curvy_hist_selected <- ggplot(error_df_one_curvy_abs_deselected1, 
       aes(x = sqrt_row_wise_total_error, 
-          y = density,
-          colour = factor(error_cat_n, levels = c("deselected", "selected")))) +
-  geom_point(alpha=0.7) +
-  scale_color_manual(values=c('#d9d9d9', '#800026')) +
+          y = density)) +
+  geom_point(alpha=0.7, color = "#d9d9d9") +
+  geom_point(data = error_df_one_curvy_abs_selected1, 
+      aes(x = sqrt_row_wise_total_error, 
+          y = density),
+      alpha=0.7, color = "#800026", size = 3) +
   xlab(expression(e[hj])) +
@@
 
-error_plot_one_curvy_hist <- ggplot(error_df_one_curvy_abs,
+error_df_one_curvy_abs_selected1 <- error_df_one_curvy_abs |>
+  dplyr::filter(error_cat_n2 == "selected")
+
+error_df_one_curvy_abs_deselected1 <- error_df_one_curvy_abs |>
+  dplyr::filter(error_cat_n2 == "deselected")
+
+error_plot_one_curvy_hist <- ggplot(error_df_one_curvy_abs_deselected1, 
       aes(x = sqrt_row_wise_total_error, 
-            y = density,
-            colour = factor(error_cat_n2, levels = c("deselected", "selected")))) +
-  geom_point(alpha=0.7) +
-  scale_color_manual(values=c('#d9d9d9', '#800026')) +
+          y = density)) +
+  geom_point(alpha=0.7, color = "#d9d9d9") +
+  geom_point(data = error_df_one_curvy_abs_selected1, 
+      aes(x = sqrt_row_wise_total_error, 
+          y = density),
+      alpha=0.7, color = "#800026", size = 3) +
   xlab(expression(e[hj])) +
@@
 "
-
-# Put error plot in area A, and wrap_plots in area B
 error_plot_pbmc + 

### 2025-05-14 paper.qmd
Updated Figure 10

--- 2025-05-14 paper.qmd
@@
 
-error_plot_one_curvy_hist_selected <- ggplot(error_df_one_curvy_abs, 
+error_df_one_curvy_abs_selected11 <- error_df_one_curvy_abs |>
+  dplyr::filter(error_cat_n == "selected")
+
+error_df_one_curvy_abs_deselected11 <- error_df_one_curvy_abs |>
+  dplyr::filter(error_cat_n == "deselected")
+
+error_plot_one_curvy_hist_selected <- ggplot(error_df_one_curvy_abs_deselected11, 
       aes(x = sqrt_row_wise_total_error, 
-          y = density,
-          colour = factor(error_cat_n, levels = c("deselected", "selected")))) +
-  geom_point(alpha=0.7) +
-  scale_color_manual(values=c('#d9d9d9', '#800026')) +
+          y = density)) +
+  geom_point(alpha=0.7, color = "#d9d9d9") +
+  geom_point(data = error_df_one_curvy_abs_selected11, 
+      aes(x = sqrt_row_wise_total_error, 
+          y = density),
+      alpha=0.7, color = "#800026", size = 3) +
   xlab(expression(e[hj])) +
@@
 
-error_plot_one_curvy_hist <- ggplot(error_df_one_curvy_abs,
+error_df_one_curvy_abs_selected12 <- error_df_one_curvy_abs |>
+  dplyr::filter(error_cat_n2 == "selected")
+
+error_df_one_curvy_abs_deselected12 <- error_df_one_curvy_abs |>
+  dplyr::filter(error_cat_n2 == "deselected")
+
+error_plot_one_curvy_hist <- ggplot(error_df_one_curvy_abs_deselected12, 
       aes(x = sqrt_row_wise_total_error, 
-            y = density,
-            colour = factor(error_cat_n2, levels = c("deselected", "selected")))) +
-  geom_point(alpha=0.7) +
-  scale_color_manual(values=c('#d9d9d9', '#800026')) +
+          y = density)) +
+  geom_point(alpha=0.7, color = "#d9d9d9") +
+  geom_point(data = error_df_one_curvy_abs_selected12, 
+      aes(x = sqrt_row_wise_total_error, 
+          y = density),
+      alpha=0.7, color = "#800026", size = 3) +
   xlab(expression(e[hj])) +
@@
 #| label: link-model-one-dens-clust
-plot_tsne_dens <- error_df_one_curvy_abs |>
+
+plot_tsne_dens <- error_df_one_curvy_abs_deselected12 |>
   ggplot(aes(x = tSNE1,
-             y = tSNE2,
-             colour = factor(error_cat_n2,
-                             levels = c("deselected", "selected")))) +
-  geom_point(alpha=0.7) +
-  scale_color_manual(values=c('#d9d9d9', '#800026')) +
+             y = tSNE2)) +
+  geom_point(alpha=0.7, color = "#d9d9d9") +
+  geom_point(data = error_df_one_curvy_abs_selected12,
+             aes(x = tSNE1, y = tSNE2),
+             alpha=0.7, color = "#800026", size = 3) +
   theme(aspect.ratio = 1) +
@@
 
-plot_tsne_dens_selected <- error_df_one_curvy_abs |>
+plot_tsne_dens_selected <- error_df_one_curvy_abs_deselected11 |>
   ggplot(aes(x = tSNE1,
-             y = tSNE2,
-             colour = factor(error_cat_n,
-                             levels = c("deselected", "selected")))) +
-  geom_point(alpha=0.7) +
-  scale_color_manual(values=c('#d9d9d9', '#800026')) +
+             y = tSNE2)) +
+  geom_point(alpha=0.7, color = "#d9d9d9") +
+  geom_point(data = error_df_one_curvy_abs_selected11,
+             aes(x = tSNE1, y = tSNE2),
+             alpha=0.7, color = "#800026", size = 3) +
   theme(aspect.ratio = 1) +
@@
                                         c("deselected", "selected", "model")),
-                         levelColors = c('#d9d9d9', '#800026', "#000000"))
+                         levelColors = c('#d9d9d9', '#800026', "#000000"),
+                         pointSize = c(1.5, 3, 0))
 ```
@@
 
-five_c_shaped_proj_tsne_model1 <- plot_proj(
-    proj_obj = proj_obj1,
-    point_param = c(1.5, 0.5), # size, alpha, color
-    line_param = c(0.8, 0.4, "#000000"), # linewidth, alpha
-    plot_limits = c(-0.35, 0.35), 
-    axis_text_size = 4.5, 
-    is_category = TRUE) +
+projected_df <- proj_obj1$projected_df
+model_df <- proj_obj1$model_df
+axes <- proj_obj1$axes
+circle <- proj_obj1$circle
+
+projected_df <- projected_df |>
+      dplyr::mutate(cluster = proj_obj1$cluster)
+
+five_c_shaped_proj_tsne_model1 <- ggplot() +
+      geom_point(
+        data = projected_df,
+        aes(
+          x = proj1,
+          y = proj2,
+          colour = factor(cluster, levels = c("deselected", "selected")),
+          size = factor(cluster, levels = c("deselected", "selected"))),
+        alpha = 0.5) +
+      geom_segment(
+        data = model_df,
+        aes(
+          x = proj1_from,
+          y = proj2_from,
+          xend = proj1_to,
+          yend = proj2_to),
+        color = "#000000",
+        linewidth = 0.8,
+        alpha = 0.4) +
+      geom_segment(
+          data=axes,
+          aes(x=x1, y=y1, xend=x2, yend=y2),
+          colour="grey70") +
+        geom_text(
+          data=axes,
+          aes(x=x2, y=y2),
+          label=rownames(axes),
+          colour="grey50",
+          size = 4.5) +
+        geom_path(
+          data=circle,
+          aes(x=c1, y=c2), colour="grey70") +
+        xlim(c(-0.35, 0.35)) +
+        ylim(c(-0.35, 0.35)) +
         interior_annotation(label = "a3", position = c(0.08, 0.9), cex = 2) +
         scale_color_manual(values = c('#d9d9d9', '#800026')) +
+        scale_size_manual(values = c('deselected' = 1.5, 'selected' = 3)) +
         theme(aspect.ratio = 1,
@@
 
-five_c_shaped_proj_tsne_model1_selected <- plot_proj(
-    proj_obj = proj_obj1, 
-    point_param = c(1.5, 0.5), # size, alpha, color
-    line_param = c(0.8, 0.4, "#000000"), # linewidth, alpha
-    plot_limits = c(-0.35, 0.35), 
-    axis_text_size = 4.5, 
-    is_category = TRUE) +
+projected_df <- proj_obj1$projected_df
+model_df <- proj_obj1$model_df
+axes <- proj_obj1$axes
+circle <- proj_obj1$circle
+
+projected_df <- projected_df |>
+      dplyr::mutate(cluster = proj_obj1$cluster)
+
+five_c_shaped_proj_tsne_model1_selected <- ggplot() +
+      geom_point(
+        data = projected_df,
+        aes(
+          x = proj1,
+          y = proj2,
+          colour = factor(cluster, levels = c("deselected", "selected")),
+          size = factor(cluster, levels = c("deselected", "selected"))),
+        alpha = 0.5) +
+      geom_segment(
+        data = model_df,
+        aes(
+          x = proj1_from,
+          y = proj2_from,
+          xend = proj1_to,
+          yend = proj2_to),
+        color = "#000000",
+        linewidth = 0.8,
+        alpha = 0.4) +
+      geom_segment(
+          data=axes,
+          aes(x=x1, y=y1, xend=x2, yend=y2),
+          colour="grey70") +
+        geom_text(
+          data=axes,
+          aes(x=x2, y=y2),
+          label=rownames(axes),
+          colour="grey50",
+          size = 4.5) +
+        geom_path(
+          data=circle,
+          aes(x=c1, y=c2), colour="grey70") +
+        xlim(c(-0.35, 0.35)) +
+        ylim(c(-0.35, 0.35)) +
   interior_annotation(label = "b3", position = c(0.08, 0.9), cex = 2) +
   scale_color_manual(values = c('#d9d9d9', '#800026')) +
+  scale_size_manual(values = c('deselected' = 1.5, 'selected' = 3)) +
   theme(aspect.ratio = 1,

### 2025-05-14 paper.qmd
Started updating the equations

@@
 The default should take the aspect ratio produced by the NLDR $(r_1, r_2, ..., r_k)$ into account.
-When $k=2$, as in hexagon binning, the default range is $[0, y_{i,\text{max}}], i=1,2$, where $y_{1,\text{max}}=1$ and $y_{2,\text{max}} = \frac{r_2}{r_1}$ (@fig-NLDR-two-curvy).
+When $k=2$, as in hexagon binning, the default range is $[0, y_{i,\text{max}}], i=1,2$, where $y_{1,\text{max}}=1$ and $y_{2,\text{max}} = r_2/r_1$ (@fig-NLDR-two-curvy).
 If the NLDR aspect ratio is ignored then set $y_ {2,\text{max}} = 1$.
@@
 The value for $b_2$ is computed by fixing $b_1$.
-Considering the upper bound of the first NLDR component, $a_1 > \frac{1+2q}{b_1 -1}$.
-Similarly, for the second NLDR component, $a_2 > \frac{r_2 + q(1 + r_2)}{(b_2 - 1)}$.
-Since $a_2 = \frac{\sqrt(3)}{2}a_1$ for regular hexagons, $a_1 > \frac{2[r_2 + q(1 + r_2)]}{\sqrt{3}(b_2 - 1)}$.
+Considering the upper bound of the first NLDR component, $a_1 > (1+2q)/(b_1 -1)$.
+Similarly, for the second NLDR component, $a_2 > (r_2 + q(1 + r_2))/(b_2 - 1)$.
+Since $a_2 = \sqrt(3)a_1/2$ for regular hexagons, $a_1 > \frac{2[r_2 + q(1 + r_2)]}{\sqrt{3}(b_2 - 1)}$.
 This is a linear optimization problem.

### 2025-05-14 paper.qmd
Updated equations to inline
Updated the remaining equations

@@
 When $k=2$, as in hexagon binning, the default range is $[0, y_{i,\text{max}}], i=1,2$, where $y_{1,\text{max}}=1$ and $y_{2,\text{max}} = r_2/r_1$ (@fig-NLDR-two-curvy).
-If the NLDR aspect ratio is ignored then set $y_ {2,\text{max}} = 1$.
-<!-- \times \frac{2}{\sqrt{3}}$.
+If the NLDR aspect ratio is ignored then set $y_ {2,\text{max}} = 1.$ <!-- \times \frac{2}{\sqrt{3}}$.
 (The $\frac{2}{\sqrt{3}}$ accounts for the different height ($a_1$) and width ($a_2$) of a regular hexagon.) The scaling of data should take the size of the hexagons into account, but choice of number of bins should.
@@
 
-The value for $b_2$ is computed by fixing $b_1$.
-Considering the upper bound of the first NLDR component, $a_1 > (1+2q)/(b_1 -1)$.
-Similarly, for the second NLDR component, $a_2 > (r_2 + q(1 + r_2))/(b_2 - 1)$.
-Since $a_2 = \sqrt(3)a_1/2$ for regular hexagons, $a_1 > \frac{2[r_2 + q(1 + r_2)]}{\sqrt{3}(b_2 - 1)}$.
-This is a linear optimization problem.
-Therefore, the optimal solution must occur on a vertex.
-Therefore, $b_2 = \Big\lceil1 +\frac{2[r_2 + q(1 + r_2)](b_1 - 1)}{\sqrt{3}(1 + 2q)}\Big\rceil$.
+The value for $b_2$ is computed by fixing $b_1$. Considering the upper bound of the first NLDR component, $a_1 > (1+2q)/(b_1 -1)$. Similarly, for the second NLDR component,
+
+$$
+a_2 > \frac{r_2 + q(1 + r_2)}{(b_2 - 1)}.
+$$
+
+Since $a_2 = \sqrt{3}a_1/2$ for regular hexagons,
+
+$$
+a_1 > \frac{2[r_2 + q(1 + r_2)]}{\sqrt{3}(b_2 - 1)}.
+$$
+
+This is a linear optimization problem. Therefore, the optimal solution must occur on a vertex. Therefore,
+
+$$
+b_2 = \Big\lceil1 +\frac{2[r_2 + q(1 + r_2)](b_1 - 1)}{\sqrt{3}(1 + 2q)}\Big\rceil.
+$$
 
@@
 The value of $b_2$ is determined analytically by $b_1, q, r_2$.
-Values of $b_1$ between $2$ and $b_1 = \sqrt{\frac{n}{r_2}}$ are allowed.
+Values of $b_1$ between $2$ and $b_1 = \sqrt{n/r_2}$ are allowed.
 @fig-param-two-curvy (a) shows the effect of different choices of $b_1$ on the MSE of the fitted model.
@@
 
-The area of a hexagon is defined as $A = \frac{3\sqrt{3}}{2}l^2$ where $l$ is the side length of the hexagon.
+The area of a hexagon is defined as $A = 3\sqrt{3}l^2/2$ where $l$ is the side length of the hexagon.
 If we know $a_1$ and $a_2$, $l$ can be computed (see appendix).
-The density of a hexagon grid is calculated as $\frac{\sum^{h}_{i=1}n_h}{A}$ and the proportion is $\frac{\sum^{h}_{i=1}n_h}{A \times b}$.
+The density of a hexagon grid is calculated as $\sum^{h}_{i=1}n_h/A$ and the proportion is $\sum^{h}_{i=1}n_h/Ab$.
 The baseline proportion is the proportion density at the smallest possible value of $a_1$.

### 2025-05-14 paper.qmd
Updated Figure1 label

@@
   geom_point(alpha=0.1, size=1, colour=clr_choice) +
-  interior_annotation("a", cex = 1.3)
+  interior_annotation("a", c(0.08, 0.9), cex = 1.3)
 
@@
   geom_point(alpha=0.1, size=1, colour='#a65628') +
-  interior_annotation("a")
+  interior_annotation("a", c(0.08, 0.9))
 
@@
   geom_point(alpha=0.1, size=1, colour=clr_choice) +
-  interior_annotation("e", cex = 1.3)
+  interior_annotation("e", c(0.08, 0.9), cex = 1.3)
 
@@
   geom_point(alpha=0.1, size=1, colour='#4daf4a') +
-  interior_annotation("e")
+  interior_annotation("e", c(0.08, 0.9))
 

### 2025-05-14 paper.qmd
Updated Figure8 labels

@@
   geom_point(alpha=0.1, size=1, colour='#a65628') +
-  interior_annotation("a", c(0.08, 0.93)) +
+  interior_annotation("a") +
   theme(aspect.ratio = 1)
@@
   geom_point(alpha=0.1, size=1, colour='#984ea3') +
-  interior_annotation("f", c(0.08, 0.93)) +
+  interior_annotation("f") +
   theme(aspect.ratio = 1)
@@
   geom_point(alpha=0.1, size=1, colour='#636363') +
-  interior_annotation("b", c(0.08, 0.93)) +
+  interior_annotation("b", c(0.08, 0.9)) +
   theme(aspect.ratio = 1)
@@
   geom_point(alpha=0.1, size=1, colour='#e41a1c') +
-  interior_annotation("e") +
+  interior_annotation("e", c(0.08, 0.9)) +
   theme(aspect.ratio = 1)

### 2025-05-14 paper.qmd
Updated Figure10 labels

@@
   interior_annotation("b1",
-                      position = c(0.95, 0.9),
                       cex = 2) +
@@
   interior_annotation("a1",
-                      position = c(0.95, 0.9),
                       cex = 2) +

### 2025-05-14 paper.qmd
Updated Figure15 labels

@@
   scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada")) +
-  interior_annotation("a1", cex = 2) +
+  interior_annotation("a1", c(0.08, 0.9), cex = 2) +
   theme(
@@
   geom_point(alpha=0.1, size=1, colour='#984ea3') +
-  interior_annotation("a", c(0.08, 0.93)) +
+  interior_annotation("a", c(0.08, 0.9)) +
   theme(aspect.ratio = 1)
@@
   geom_point(alpha=0.1, size=1, colour='#999999') +
-  interior_annotation("f", c(0.08, 0.93)) +
+  interior_annotation("f", c(0.08, 0.9)) +
   theme(aspect.ratio = 1)
@@
   geom_point(alpha=0.1, size=1, colour='#377eb8') +
-  interior_annotation("c", position = c(0.94, 0.92)) +
+  interior_annotation("c", c(0.94, 0.92)) +
   theme(aspect.ratio = 1)

### 2025-05-14 paper.qmd
Updated Figure7 labels

@@
   labs(x = expression(paste("binwidth (", a[1], ")")), y = "RMSE") +
-  interior_annotation("a", position = c(0.9, 0.15), cex = 2) +
+  interior_annotation("a", position = c(0.9, 0.1), cex = 2) +
   theme_minimal() +
@@
   labs(x = expression(paste("binwidth (", a[1], ")")), y = expression("prop. non-empty bins " * bgroup("(", m / b, ")"))) +
-  interior_annotation("b", position = c(0.9, 0.15), cex = 2) +
+  interior_annotation("b", position = c(0.9, 0.1), cex = 2) +
   theme_minimal() +
@@
   ylab("RMSE") +
-  interior_annotation("c", position = c(0.1, 0.9), cex = 2) +
+  interior_annotation("c", position = c(0.07, 0.9), cex = 2) +
   theme_minimal() +

### 2025-05-14 paper.qmd
Updated Figure 7c

@@
                         labels = parse_format()) +
-  xlab("threshold to remove low-density hexagons") +
+  xlab("bin density cutoff") +
   ylab("RMSE") +
@@
 #| fig-cap: "Various plots to help assess best hexagon bin parameters (a, b, d) and thresholds to remove low-density bins (c).
-Both (a) and (c) show RMSE, against binwidth ($a_1$) and threshold.
+Both (a) and (c) show RMSE, against binwidth ($a_1$) and bin density cutoff.
 A good benchmark value for these parameters is when the RMSE drops and then flattens out.

### 2025-05-14 paper.qmd
Updated Figure 11 with RMSE

@@
                      aes(x = a1, 
-                         y = sqrt(MSE))) +
+                         y = RMSE)) +
   geom_vline(xintercept = 0.09,
@@
   group_by(method, a1) |>
-  filter(MSE == min(MSE)) |>
+  filter(RMSE == min(RMSE)) |>
   ungroup()
@@
   group_by(method, a1) |>
-  filter(MSE == min(MSE)) |>
+  filter(RMSE == min(RMSE)) |>
   ungroup() |>

### 2025-05-14 appendix.qmd
Started updating the appendix

@@
 scurve_scaled_obj <- gen_scaled_data(
-  data = umap_scurve)
-
-umap_scurve_scaled <- scurve_scaled_obj$scaled_nldr
-lim1 <- scurve_scaled_obj$lim1
-lim2 <- scurve_scaled_obj$lim2
-r2 <- diff(lim2)/diff(lim1)
+  nldr_data = umap_scurve)
 
@@
 hb_obj_scurve <- hex_binning(
-  data = umap_scurve_scaled, 
-  bin1 = num_bins_x_scurve, 
-  r2 = r2)
+  nldr_obj = scurve_scaled_obj, 
+  bin1 = 8, 
+  q = 0.1)
 
@@
                  x2min = 0,
-                 x2max = r2)
+                 x2max = diff(scurve_scaled_obj$nldr_obj$lim2)/diff(scurve_scaled_obj$nldr_obj$lim1))
 
@@
 \frac{2(r_2 + qr_2)}{\sqrt{3}(b_2 - 1)} = \frac{(1 + q)}{(b_1 - 1)}.
-$$ After rearranging this,
+$$
+
+After rearranging this,
 
@@
 hb_obj_scurve <- hex_binning(
-  data = umap_scurve_scaled, 
+  nldr_obj = scurve_scaled_obj, 
   bin1 = num_bins_x_scurve, 
-  r2 = r2)
+  q = 0.1)
 
@@
 #| fig-height: 6
+
 assign_data_vis1 + assign_data_vis2 +
@@
 
-### Single-cell gene expression
-
-#### Comparison with results of scDEED recommendations
+## Single-cell gene expression: comparison with results of scDEED recommendations
+
+<!-- ### Comparison with results of scDEED recommendations -->
 
@@
 #| out-width: 100%
-#| fig-cap: "Assessing which of the two layouts with UMAP and tSNE with different hyper-parameter setting (n_neighbors: $30$, min_dist: $0.3$ (red); n_neighbors: $30$ (blue)) on the PBMC data is the better representation using MSE for varying binwidth ($a_1$).
+#| fig-cap: "Assessing which of the two layouts with UMAP and tSNE with different hyper-parameter setting (n_neighbors: $30$, min_dist: $0.3$ (red); n_neighbors: $30$ (blue)) on the PBMC data is the better representation using RMSE for varying binwidth ($a_1$).
 Colour used for the lines and points in the top plot and in the scatterplots represents UMAP and tSNE layouts (a, b).
@@
 
-algo_obj_pbmc <- gen_nldr_vis_algo_obj(
-  high_d_data = training_data_pbmc, 
+algo_obj_pbmc <- fit_highd_model(
+  highd_data = training_data_pbmc, 
   nldr_data = umap_pbmc, 
-  num_x_bins = num_bins_x_pbmc)
-
-umap_pbmc_scaled <- algo_obj_pbmc$nldr_scaled
-distance_pbmc <- algo_obj_pbmc$distance_df
-tr_from_to_df_pbmc <- algo_obj_pbmc$tr_from_to_df
-benchmark_pbmc <- algo_obj_pbmc$benchmark
-df_bin_centroids_pbmc <- algo_obj_pbmc$df_bin_centroids
-df_bin_pbmc <- algo_obj_pbmc$df_bin
-
-distance_df_small_edges_pbmc <- distance_pbmc |>
-  filter(distance < benchmark_pbmc)#benchmark_pbmc
+  bin1 = num_bins_x_pbmc, 
+  q = 0.1, 
+  benchmark_highdens = 5)
+
+umap_pbmc_scaled <- algo_obj_pbmc$nldr_obj$scaled_nldr
+tr_from_to_df_pbmc <- algo_obj_pbmc$trimesh_data
+df_bin_centroids_pbmc <- algo_obj_pbmc$model_2d
+df_bin_pbmc <- algo_obj_pbmc$model_highd
 
@@
                proj_scale = 1.23, 
-               scaled_data = scaled_pbmc_data, 
-               scaled_data_model = scaled_pbmc_data_model, 
-               distance_df_small_edges = distance_df_small_edges_pbmc, 
+               highd_data = scaled_pbmc_data, 
+               model_highd = scaled_pbmc_data_model, 
+               trimesh_data = tr_from_to_df_pbmc, 
                axis_param = list(limits = 1,
@@
   plot_limits = c(-0.65, 0.65), 
-  title = "a2", 
-  cex = 2, 
   axis_text_size = 4,
-  is_color = TRUE) + scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))
+  is_category = TRUE) + 
+  interior_annotation(label = "a2", cex = 2) +
+  scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))
 
@@
                proj_scale = 1.25, 
-               scaled_data = scaled_pbmc_data, 
-               scaled_data_model = scaled_pbmc_data_model, 
-               distance_df_small_edges = distance_df_small_edges_pbmc, 
+               highd_data = scaled_pbmc_data, 
+               model_highd = scaled_pbmc_data_model, 
+               trimesh_data = tr_from_to_df_pbmc, 
                axis_param = list(limits = 1.35, 
@@
 
-algo_obj_pbmc <- gen_nldr_vis_algo_obj(
-  high_d_data = training_data_pbmc, 
+algo_obj_pbmc <- fit_highd_model(
+  highd_data = training_data_pbmc, 
   nldr_data = tsne_pbmc, 
-  num_x_bins = num_bins_x_pbmc)
-
-tsne_pbmc_scaled <- algo_obj_pbmc$nldr_scaled
-distance_pbmc <- algo_obj_pbmc$distance_df
-tr_from_to_df_pbmc <- algo_obj_pbmc$tr_from_to_df
-benchmark_pbmc <- algo_obj_pbmc$benchmark
-df_bin_centroids_pbmc <- algo_obj_pbmc$df_bin_centroids
-df_bin_pbmc <- algo_obj_pbmc$df_bin
-
-distance_df_small_edges_pbmc <- distance_pbmc |>
-  filter(distance < benchmark_pbmc)#benchmark_pbmc
+  bin1 = num_bins_x_pbmc, 
+  q = 0.1, 
+  benchmark_highdens = 1)
+
+tsne_pbmc_scaled <- algo_obj_pbmc$nldr_obj$scaled_nldr
+tr_from_to_df_pbmc <- algo_obj_pbmc$trimesh_data
+df_bin_centroids_pbmc <- algo_obj_pbmc$model_2d
+df_bin_pbmc <- algo_obj_pbmc$model_highd
 
@@
                             proj_scale = 1.23, 
-                            scaled_data = scaled_pbmc_data, 
-                            scaled_data_model = scaled_pbmc_data_model, 
-                            distance_df_small_edges = distance_df_small_edges_pbmc, 
+                            highd_data = scaled_pbmc_data, 
+                            model_highd = scaled_pbmc_data_model, 
+                            trimesh_data = tr_from_to_df_pbmc, 
                             axis_param = list(limits = 1,
@@
   plot_limits = c(-0.65, 0.65), 
-  title = "b2", 
-  cex = 2, 
   axis_text_size = 4,
-  is_color = TRUE) + scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))
+  is_category = TRUE) + 
+  interior_annotation(label = "b2", cex = 2) +
+  scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))
 
@@
                             proj_scale = 1.25, 
-                            scaled_data = scaled_pbmc_data, 
-                            scaled_data_model = scaled_pbmc_data_model, 
-                            distance_df_small_edges = distance_df_small_edges_pbmc, 
+                            highd_data = scaled_pbmc_data, 
+                            model_highd = scaled_pbmc_data_model, 
+                            trimesh_data = tr_from_to_df_pbmc, 
                             axis_param = list(limits = 1.35, 
@@
   plot_limits = c(-1, 0.75), 
-  title = "b3", 
-  cex = 2, 
   axis_text_size = 4,
-  is_color = TRUE) + scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))
+  is_category = TRUE) + 
+  interior_annotation(label = "b3", cex = 2) +
+  scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))
 

### 2025-05-14 appendix.tex
Started updating the appendix

@@
 \frac{2(r_2 + qr_2)}{\sqrt{3}(b_2 - 1)} = \frac{(1 + q)}{(b_1 - 1)}.
-\] After rearranging this,
+\]
+
+After rearranging this,
 
@@
 
-\subsection{Single-cell gene
-expression}\label{single-cell-gene-expression}
-
-\subsubsection{Comparison with results of scDEED
-recommendations}\label{comparison-with-results-of-scdeed-recommendations}
+\section{Single-cell gene expression: comparison with results of scDEED
+recommendations}\label{single-cell-gene-expression-comparison-with-results-of-scdeed-recommendations}
 
@@
 \(30\), min\_dist: \(0.3\) (red); n\_neighbors: \(30\) (blue)) on the
-PBMC data is the better representation using MSE for varying binwidth
+PBMC data is the better representation using RMSE for varying binwidth
 (\(a_1\)). Colour used for the lines and points in the top plot and in

### 2025-05-14 appendix.qmd
Updated Figure 4

@@
 
-free(error_umap_plot_pbmc) + wrap_plots(nldr1c, nldr2c, ncol = 2) +
-  plot_layout(ncol = 1, heights = c(1, 1))
+# Define the layout design
+layout_design <- "
+AA
+BC
+"
+
+# Combine the plots using the layout
+(error_umap_plot_pbmc + nldr1c + nldr2c) +
+  plot_layout(design = layout_design)
 ```

### 2025-05-14 appendix.qmd
Fixed the reference

@@
 Following the process in @xia2023 again using the Human Peripheral Blood Mononuclear Cells (PBMC) data.
-Note that @xia2023 uses a different subset from the PBMC dataset used by @chen2023, which contains $31,021$ cells including cell type labels, and the gene expression levels were in the unit of log-transformed UMI count per $10,000$.
+Note that @xia2023 uses a different subset from the PBMC dataset used by @chen2024, which contains $31,021$ cells including cell type labels, and the gene expression levels were in the unit of log-transformed UMI count per $10,000$.
 They focused on three sequencing methods (inDrops, DropSeq, and SeqWell) and four common cell types Cytotoxic T cell, CD4+T cell, CD14+ Monocyte, and B cell.

### 2025-05-14 appendix.tex
Fixed the reference

@@
 \citet{xia2023} uses a different subset from the PBMC dataset used by
-\citet{chen2023}, which contains \(31,021\) cells including cell type
+\citet{chen2024}, which contains \(31,021\) cells including cell type
 labels, and the gene expression levels were in the unit of

### 2025-05-14 appendix.qmd
Updated Figure 1

@@
   filter(hexID == 1)
+
 d_rect <- tibble(x1min = 0, 
@@
                  x2min = 0,
-                 x2max = diff(scurve_scaled_obj$nldr_obj$lim2)/diff(scurve_scaled_obj$nldr_obj$lim1))
+                 x2max = diff(scurve_scaled_obj$lim2)/diff(scurve_scaled_obj$lim1))
 
@@
              label = expression(a[2]))
+
 hex_param_vis <- ggplot() + 
@@
                      label = "(0,0)", 
-             hjust=-0.1, vjust=-0.3) + 
+             hjust=-0.1, vjust=-0.3, size = 8) + 
     annotate("text", x=d_rect$x1max - c_start$s1, 
@@
                      label = "(0,1)", 
-             hjust=1.1, vjust=-0.3) + 
+             hjust=1.1, vjust=-0.3, size = 8) + 
     annotate("text", x=d_rect$x1min - c_start$s1, 
@@
     annotate("text", x=0.5, y=-0.45, 
-             label = expression(r[1]), color = "black") +
+             label = expression(r[1]), color = "black", size = 8) +
     geom_segment(data=d_rect, aes(
@@
     annotate("text", x=-0.35, y=1, 
-             label = expression(r[2]), color = "black") +
+             label = expression(r[2]), color = "black", size = 8) +
     geom_segment(data = a1, aes(
@@
              label = expression(a[2]), 
-             color = "black", hjust=-0.2) +
+             color = "black", hjust=-0.2, size = 8) +
     annotate("text", x=-0.18, y=-0.25, 
       label = expression(group("(", list(s[1], s[2]), ")")),
-      color = "black") +
+      color = "black", size = 8) +
   coord_equal()  

### 2025-05-15 appendix.qmd
Updated Figure 1

--- 2025-05-14 appendix.qmd
@@
   nldr_obj = scurve_scaled_obj, 
-  bin1 = 8, 
+  bin1 = num_bins_x_scurve, 
   q = 0.1)
@@
 hex_grid <- hb_obj_scurve$hex_poly
+
+a1_temp <- hb_obj_scurve$a1
+a2_temp <- hb_obj_scurve$a2
+l_temp <- quad(a=3, b = 2 * a2_temp, c = -(a2_temp^2 + a1_temp^2))
+
 ```
@@
                  x2min = 0,
-                 x2max = diff(scurve_scaled_obj$nldr_obj$lim2)/diff(scurve_scaled_obj$nldr_obj$lim1))
+                 x2max = diff(scurve_scaled_obj$lim2)/diff(scurve_scaled_obj$lim1))
+
+hex_grid_temp40 <- hex_grid |> 
+  filter(hex_poly_id == 53)
 
@@
              xend = all_centroids_df$c_x[5],
-             y = all_centroids_df$c_y[21],
-             yend = all_centroids_df$c_y[21],
+             y = all_centroids_df$c_y[29],
+             yend = all_centroids_df$c_y[29],
              label = expression(a[1]))
-a2 <- tibble(x = all_centroids_df$c_x[27],
-             xend = all_centroids_df$c_x[27],
-             y = all_centroids_df$c_y[27],
-             yend = all_centroids_df$c_y[33],
+a2 <- tibble(x = all_centroids_df$c_x[39],
+             xend = all_centroids_df$c_x[39],
+             y = all_centroids_df$c_y[39],
+             yend = all_centroids_df$c_y[47],
              label = expression(a[2]))
+
+l <- tibble(x = hex_grid_temp40$x[2],
+            xend = hex_grid_temp40$x[3],
+            y = hex_grid_temp40$y[2],
+            yend = hex_grid_temp40$y[3],
+            label = expression(l))
+
 hex_param_vis <- ggplot() + 
@@
              label = expression(a[2]), 
-             color = "black", hjust=-0.2) +
+             color = "black", hjust=-0.2, size = 8) +
     annotate("text", x=-0.18, y=-0.25, 
       label = expression(group("(", list(s[1], s[2]), ")")),
+      color = "black", size = 8) +
+  geom_segment(data = l, aes(
+      x = x, #-0.15, 
+      y = y, #-0.1*r2 + 0.1807896*2, 
+      xend = xend, #-0.15, 
+      yend = yend), #-0.1*r2 + 0.1807896*3),
+      arrow = arrow(length = unit(0.03, "npc"),
+                               ends = "both"), 
       color = "black") + 
+    annotate("text", x=l$x + 0.03, y=(l$y+l$yend)/2, 
+             label = expression(l), 
+             color = "black", hjust=-0.2, size = 8) +
   coord_equal()  

### 2025-05-15 appendix.qmd
Updated Figure 3

@@
   ) + 
-  annotate("text", x=-1, y= 1.5, 
+  annotate("text", x=-1.5, y= 1.5, 
            label = expression(frac(a[2] - a[1], 2)), 
-           color = "black", hjust=-0.2, size = 5) + 
-  annotate("text", x=1.5, y= -0.3, 
+           color = "black", hjust=-0.2, size = 8) + 
+  annotate("text", x=1.5, y= -0.5, 
            label = expression(frac(a[1], 2)), 
-           color = "black", hjust=-0.2, size = 5) + 
+           color = "black", hjust=-0.2, size = 8) + 
   annotate("text", x=2.5, y= 1.5, 
            label = expression(l), 
-           color = "black", hjust=-0.2, size = 5)
+           color = "black", hjust=-0.2, size = 8)
 ```

### 2025-05-15 appendix.qmd
Removed the whitespace

@@
 
-triangle_arrow_temp <- tibble(
-  x = c(4 + 0.04, 0 - 0.06, 4 + 0.06),
-  xend = c(0 + 0.04, 0 - 0.06, 0 + 0.06),
-  y = c(0 - 0.04, 0 - 0.06, 0 + 0.06),
-  yend = c(0 - 0.04, 3 - 0.06, 3 + 0.06)
-)
+# triangle_arrow_temp <- tibble(
+#   x = c(4 + 0.04, 0 - 0.06, 4 + 0.06),
+#   xend = c(0 + 0.04, 0 - 0.06, 0 + 0.06),
+#   y = c(0 - 0.04, 0 - 0.06, 0 + 0.06),
+#   yend = c(0 - 0.04, 3 - 0.06, 3 + 0.06)
+# )
 
@@
 
-The layout a (@fig-pbmc-mse-umap) is generated from the hyper-parameters suggested by @chen2023, and the layout b (@fig-pbmc-mse-umap) is with suggested hyper-parameters by scDEED to be more accurate.
-The layout a and b contain $46$ and $83$ dubious cells respectively.
-
-The log MSE vs binwidth ($a_1$) plot (@fig-pbmc-mse-umap) illustrates that our approach would suggest that scDEED is correct here, that layout b is more accurately reflecting the cluster structure in the PBMC data.
-
 ```{r}
@@
 ```
+The layout a (@fig-pbmc-mse-umap) is generated from the hyper-parameters suggested by @chen2024, and the layout b (@fig-pbmc-mse-umap) is with suggested hyper-parameters by scDEED to be more accurate.
+The layout a and b contain $46$ and $83$ dubious cells respectively.
+
+The RMSE vs binwidth ($a_1$) plot (@fig-pbmc-mse-umap) illustrates that our approach would suggest that scDEED is correct here, that layout b is more accurately reflecting the cluster structure in the PBMC data.
 

### 2025-05-15 appendix.tex
Removed the whitespace

@@
 
-The layout a (Figure~\ref{fig-pbmc-mse-umap}) is generated from the
-hyper-parameters suggested by \citet{chen2023}, and the layout b
-(Figure~\ref{fig-pbmc-mse-umap}) is with suggested hyper-parameters by
-scDEED to be more accurate. The layout a and b contain \(46\) and \(83\)
-dubious cells respectively.
-
-The log MSE vs binwidth (\(a_1\)) plot (Figure~\ref{fig-pbmc-mse-umap})
-illustrates that our approach would suggest that scDEED is correct here,
-that layout b is more accurately reflecting the cluster structure in the
-PBMC data.
-
 \begin{figure}[H]
@@
 \end{figure}%
+
+The layout a (Figure~\ref{fig-pbmc-mse-umap}) is generated from the
+hyper-parameters suggested by \citet{chen2024}, and the layout b
+(Figure~\ref{fig-pbmc-mse-umap}) is with suggested hyper-parameters by
+scDEED to be more accurate. The layout a and b contain \(46\) and \(83\)
+dubious cells respectively.
+
+The RMSE vs binwidth (\(a_1\)) plot (Figure~\ref{fig-pbmc-mse-umap})
+illustrates that our approach would suggest that scDEED is correct here,
+that layout b is more accurately reflecting the cluster structure in the
+PBMC data.
 

### 2025-05-15 appendix.qmd
Updated appendix
Updated point alpha

@@
 ```
-
-\newpage
 
@@
     ),
-    alpha = 0.2,
-    size = 0.1
+    alpha = 0.3#,
+    #size = 0.1
   )  +
@@
   q = 0.1, 
-  benchmark_highdens = 1)
+  benchmark_highdens = 5)
 
@@
 #| echo: false
-#| fig-cap: "Compare the published \\gD{} layout (@fig-pbmc-mse-umap a) and the \\gD{} layout selected (@fig-pbmc-mse-umap b) by MSE plot (@fig-pbmc-mse-umap) from tSNE and UMAP with default hyper-parameters.
-The PBMC3k data ($n =  5858$) has three clusters in $9\\text{-}D$, where three clusters are close.
+#| fig-cap: "Compare the published \\gD{} layout (@fig-pbmc-mse-umap a) and the \\gD{} layout selected (@fig-pbmc-mse-umap b) by RMSE plot (@fig-pbmc-mse-umap) from tSNE and UMAP with default hyper-parameters.
+The PBMC data ($n =  5858$) has three clusters in $9\\text{-}D$, where three clusters are close.
 Two \\gD{} projection from a tour on $9\\text{-}D$ of the model fit with @fig-pbmc-mse-umap a ($a_1 = 0.06$, $b = 616/86 (22, 28)$) shows three-well separated clusters with big separations.

### 2025-05-15 appendix.tex
Updated appendix
Updated point alpha

@@
 \end{table}%
-
-\newpage
 
@@
 layout (Figure~\ref{fig-pbmc-mse-umap} a) and the \gD{} layout selected
-(Figure~\ref{fig-pbmc-mse-umap} b) by MSE plot
+(Figure~\ref{fig-pbmc-mse-umap} b) by RMSE plot
 (Figure~\ref{fig-pbmc-mse-umap}) from tSNE and UMAP with default
-hyper-parameters. The PBMC3k data (\(n =  5858\)) has three clusters in
+hyper-parameters. The PBMC data (\(n =  5858\)) has three clusters in
 \(9\text{-}D\), where three clusters are close. Two \gD{} projection

### 2025-05-15 paper.qmd
Updated point alpha

@@
     ),
-    alpha = 0.2,
-    size = 0.1#,
+    alpha = 0.2#,
+    #size = 0.1#,
     #color = '#a65628'

### 2025-05-15 appendix.qmd
Updated Figure 5 title

@@
 tsne_pbmc_scaled <- algo_obj_pbmc$nldr_obj$scaled_nldr
-tr_from_to_df_pbmc <- algo_obj_pbmc$trimesh_data
+#tr_from_to_df_pbmc <- algo_obj_pbmc$trimesh_data
 df_bin_centroids_pbmc <- algo_obj_pbmc$model_2d
 df_bin_pbmc <- algo_obj_pbmc$model_highd
+
+## Removed some bins manually to fit a better model
+df_bin_centroids_pbmc <- df_bin_centroids_pbmc |>
+  filter(!(hexID %in% c(559, 617, 677)))
+
+tri_object <- tri_bin_centroids(
+  centroids_data = df_bin_centroids_pbmc)
+
+tr_from_to_df_pbmc <- gen_edges(tri_object = tri_object, a1 = algo_obj_pbmc$hb_obj$a1)
+tr_from_to_df_pbmc <- tr_from_to_df_pbmc <- update_trimesh_index(trimesh_data = tr_from_to_df_pbmc)
 
@@
 The PBMC data ($n =  5858$) has three clusters in $9\\text{-}D$, where three clusters are close.
-Two \\gD{} projection from a tour on $9\\text{-}D$ of the model fit with @fig-pbmc-mse-umap a ($a_1 = 0.06$, $b = 616/86 (22, 28)$) shows three-well separated clusters with big separations.
-On the other hand, the model fit with @fig-pbmc-mse-umap b ($a_1 = 0.06$, $b = 704/227 (22, 32)$) shows three-well separated clusters with small separations.
+Two \\gD{} projection from a tour on $9\\text{-}D$ of the model fit with @fig-pbmc-mse-umap a ($a_1 = 0.04$, $b = 900/179 (30, 30)$) shows three-well separated clusters with big separations.
+On the other hand, the model fit with @fig-pbmc-mse-umap b ($a_1 = 0.04$, $b = 1110/317 (30, 37)$) shows three-well separated clusters with small separations.
 Therefore, @fig-pbmc-mse-umap b is more reasonable than @fig-pbmc-mse-umap a.

### 2025-05-15 appendix.tex
Updated Figure 5 title

@@
 from a tour on \(9\text{-}D\) of the model fit with
-Figure~\ref{fig-pbmc-mse-umap} a (\(a_1 = 0.06\),
-\(b = 616/86 (22, 28)\)) shows three-well separated clusters with big
+Figure~\ref{fig-pbmc-mse-umap} a (\(a_1 = 0.04\),
+\(b = 900/179 (30, 30)\)) shows three-well separated clusters with big
 separations. On the other hand, the model fit with
-Figure~\ref{fig-pbmc-mse-umap} b (\(a_1 = 0.06\),
-\(b = 704/227 (22, 32)\)) shows three-well separated clusters with small
-separations. Therefore, Figure~\ref{fig-pbmc-mse-umap} b is more
+Figure~\ref{fig-pbmc-mse-umap} b (\(a_1 = 0.04\),
+\(b = 1110/317 (30, 37)\)) shows three-well separated clusters with
+small separations. Therefore, Figure~\ref{fig-pbmc-mse-umap} b is more
 reasonable than Figure~\ref{fig-pbmc-mse-umap} a. Videos of the

### 2025-05-15 appendix.qmd
Added new projections. Need to update the circle isse

@@
 projection <- cbind(
-    c(-0.4545,-0.5533,-0.1235,0.3984,-0.2065,-0.3586,-0.0088,0.3572,-0.1368),
-    c(0.4576,-0.3157,0.4708,0.1410,-0.4479,-0.1685,0.3668,-0.2692,0.1332))
+  c(0.025588,0.002516,0.002185,0.025284,0.004030,-0.026880,-0.015326,0.014851,0.026260),
+  c(0.026495,-0.024817,0.021623,-0.014036,-0.010220,0.016634,-0.006496,-0.022928,0.016045))
 
@@
                                  axis_scaled = 1, 
-                                 axis_pos_x = -0.48, 
-                                 axis_pos_y = -0.48, 
-                                 threshold = 0.082))
+                                 axis_pos_x = -0.03, 
+                                 axis_pos_y = -0.03, 
+                                 threshold = 0.0051))
 
@@
   point_param = c(1.5, 0.2), # size, alpha, color
-  plot_limits = c(-0.65, 0.65), 
-  axis_text_size = 4,
+  plot_limits = c(-0.04, 0.045), 
+  axis_text_size = 5,
   is_category = TRUE) + 
@@
 projection <- cbind(
-    c(0.0734,0.6138,-0.4671,0.2589,0.1189,0.0778,-0.0819,0.0308,-0.5561),
-    c(0.6625,0.0034,0.2412,0.3360,0.2463,0.1070,-0.2903,-0.4705,0.1293))
+  c(0.025013,0.029055,-0.003143,0.001821,-0.006208,0.019174,0.020630,0.021512,0.020281),
+  c(0.004528,0.006465,-0.042452,0.009837,0.010040,-0.023820,0.005236,0.014775,-0.017712))
 
@@
                                  axis_scaled = 1, 
-                                 axis_pos_x = -0.75, 
-                                 axis_pos_y = -0.75, 
-                                 threshold = 0.12))
+                                 axis_pos_x = -0.03, 
+                                 axis_pos_y = -0.02, 
+                                 threshold = 0.006))
 
@@
     point_param = c(1.5, 0.2), # size, alpha, color
-    plot_limits = c(-1, 0.75), 
+    plot_limits = c(-0.03, 0.03), 
     axis_text_size = 4,
@@
 projection <- cbind(
-  c(-0.4545,-0.5533,-0.1235,0.3984,-0.2065,-0.3586,-0.0088,0.3572,-0.1368),
-  c(0.4576,-0.3157,0.4708,0.1410,-0.4479,-0.1685,0.3668,-0.2692,0.1332))
+  c(0.025588,0.002516,0.002185,0.025284,0.004030,-0.026880,-0.015326,0.014851,0.026260),
+  c(0.026495,-0.024817,0.021623,-0.014036,-0.010220,0.016634,-0.006496,-0.022928,0.016045))
 
@@
                                               axis_scaled = 1, 
-                                              axis_pos_x = -0.48, 
-                                              axis_pos_y = -0.48, 
-                                              threshold = 0.082))
+                                              axis_pos_x = -0.03, 
+                                              axis_pos_y = -0.03, 
+                                              threshold = 0.0051))
 
@@
   point_param = c(1.5, 0.2), # size, alpha, color
-  plot_limits = c(-0.65, 0.65), 
+  plot_limits = c(-0.04, 0.045), 
   axis_text_size = 4,
@@
 projection <- cbind(
-  c(0.0734,0.6138,-0.4671,0.2589,0.1189,0.0778,-0.0819,0.0308,-0.5561),
-  c(0.6625,0.0034,0.2412,0.3360,0.2463,0.1070,-0.2903,-0.4705,0.1293))
+  c(0.025013,0.029055,-0.003143,0.001821,-0.006208,0.019174,0.020630,0.021512,0.020281),
+  c(0.004528,0.006465,-0.042452,0.009837,0.010040,-0.023820,0.005236,0.014775,-0.017712))
 
@@
                                               axis_scaled = 1, 
-                                              axis_pos_x = -0.75, 
-                                              axis_pos_y = -0.75, 
-                                              threshold = 0.12))
+                                              axis_pos_x = -0.03, 
+                                              axis_pos_y = -0.02, 
+                                              threshold = 0.006))
 
@@
   point_param = c(1.5, 0.2), # size, alpha, color
-  plot_limits = c(-1, 0.75), 
+  plot_limits = c(-0.03, 0.03), 
   axis_text_size = 4,

### 2025-05-15 appendix.qmd
Updated Figure 5

@@
                trimesh_data = tr_from_to_df_pbmc, 
-               axis_param = list(limits = 1,
-                                 axis_scaled = 1, 
+               axis_param = list(limits = 0.05,
+                                 axis_scaled = 17, 
                                  axis_pos_x = -0.03, 
                                  axis_pos_y = -0.03, 
-                                 threshold = 0.0051))
+                                 threshold = 0.0044))
 
@@
   plot_limits = c(-0.04, 0.045), 
-  axis_text_size = 5,
+  axis_text_size = 4,
   is_category = TRUE) + 
@@
                trimesh_data = tr_from_to_df_pbmc, 
-               axis_param = list(limits = 1.35, 
-                                 axis_scaled = 1, 
-                                 axis_pos_x = -0.03, 
-                                 axis_pos_y = -0.02, 
-                                 threshold = 0.006))
+               axis_param = list(limits = 0.035, 
+                                 axis_scaled = 17, 
+                                 axis_pos_x = -0.023, 
+                                 axis_pos_y = -0.023, 
+                                 threshold = 0.0027))
 
@@
                trimesh_data = tr_from_to_df_pbmc, 
-                            axis_param = list(limits = 1,
-                                              axis_scaled = 1, 
+               axis_param = list(limits = 0.05,
+                                 axis_scaled = 17, 
                                  axis_pos_x = -0.03, 
                                  axis_pos_y = -0.03, 
-                                              threshold = 0.0051))
+                                 threshold = 0.0044))
 
@@
                trimesh_data = tr_from_to_df_pbmc, 
-                            axis_param = list(limits = 1.35, 
-                                              axis_scaled = 1, 
-                                              axis_pos_x = -0.03, 
-                                              axis_pos_y = -0.02, 
-                                              threshold = 0.006))
+               axis_param = list(limits = 0.035, 
+                                 axis_scaled = 17, 
+                                 axis_pos_x = -0.023, 
+                                 axis_pos_y = -0.023, 
+                                 threshold = 0.0027))
 

### 2025-05-15 appendix.qmd
Updated the arrangement

@@
 $a_1$, and number of bins along the y-axis, $b_2$.
+
+Geometric arguments give rise to the following constraints.
+
+$\text{min }a_1 \text{ s.t.}$
+
+$$
+s_1 - \frac{a_1}{2} < 0,
+$$ {#eq-equation1}
+
+$$
+s_1 + (b_1 - 1) \times a_1 > 1,
+$$ {#eq-equation2}
+
+$$
+s_2 - \frac{a_2}{2} < 0,
+$$ {#eq-equation4}
+
+$$
+s_2 + (b_2 - 1) \times a_2 > r_2.
+$$ {#eq-equation5}
 
@@
 
-Geometric arguments give rise to the following constraints.
-
-$\text{min }a_1 \text{ s.t.}$
-
-$$
-s_1 - \frac{a_1}{2} < 0,
-$$ {#eq-equation1}
-
-$$
-s_1 + (b_1 - 1) \times a_1 > 1,
-$$ {#eq-equation2}
-
-$$
-s_2 - \frac{a_2}{2} < 0,
-$$ {#eq-equation4}
-
-$$
-s_2 + (b_2 - 1) \times a_2 > r_2.
-$$ {#eq-equation5}
-
 Since $a_1$ and $a_2$ are distances,
@@
 Therefore, if this equality is true, then $a_1 = \frac{(1+q)}{(b_1 - 1)}$, otherwise, $a_1 = \frac{2r_2(1+q)}{\sqrt{3}(b_2 - 1)}$.
-
-\newpage
 
@@
 #| fig-height: 6
+#| fig-pos: H
 
@@
 #| out-width: 30%
+#| fig-pos: H
 
@@
 The objective is to assess the optimized layout by scDEED, and if it does not accurately represent the three clusters with small separations of the PBMC dataset, then select a reasonable \gD{} layout.
+
+The layout a (@fig-pbmc-mse-umap) is generated from the hyper-parameters suggested by @chen2024, and the layout b (@fig-pbmc-mse-umap) is with suggested hyper-parameters by scDEED to be more accurate.
+The layout a and b contain $46$ and $83$ dubious cells respectively.
 
@@
 ```
-The layout a (@fig-pbmc-mse-umap) is generated from the hyper-parameters suggested by @chen2024, and the layout b (@fig-pbmc-mse-umap) is with suggested hyper-parameters by scDEED to be more accurate.
-The layout a and b contain $46$ and $83$ dubious cells respectively.
 

### 2025-05-15 appendix.qmd
Updated the arrangement of equations

@@
 which given $s_1 < 0$ and $a_1 > 0$ will *always* be true.
-The same logic follows for @eq-equation4 and substituting $a_2 = \frac{\sqrt{3}}{2}a_1$, and $s_2 = -q \times r_2$ to @eq-equation4 can be written as,
+The same logic follows for @eq-equation4 and substituting $a_2 = \sqrt{3}a_1/{2}$, and $s_2 = -q \times r_2$ to @eq-equation4 can be written as,
 
@@
 
-Also, substituting $a_2 = \frac{\sqrt{3}}{2}a_1$, $s_2 = -q \times r_2$ and rearranging @eq-equation5 gives:
+Also, substituting $a_2 = \sqrt{3}a_1/{2}$, $s_2 = -q \times r_2$ and rearranging @eq-equation5 gives:
 
@@
 
-Therefore, if this equality is true, then $a_1 = \frac{(1+q)}{(b_1 - 1)}$, otherwise, $a_1 = \frac{2r_2(1+q)}{\sqrt{3}(b_2 - 1)}$.
+Therefore, if this equality is true, then 
+$$
+a_1 = \frac{(1+q)}{(b_1 - 1)},
+$$
+otherwise, 
+$$
+a_1 = \frac{2r_2(1+q)}{\sqrt{3}(b_2 - 1)}.
+$$
 

### 2025-05-15 appendix.tex
Updated the arrangement
Updated the arrangement of equations

@@
 
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=0.3\linewidth,height=\textheight,keepaspectratio]{appendix_files/figure-pdf/fig-hex-param-1.pdf}
-
-}
-
-\caption{\label{fig-hex-param}The components of the hexagon grid
-illustrating notation.}
-
-\end{figure}%
-
 Geometric arguments give rise to the following constraints.
@@
 
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=0.3\linewidth,height=\textheight,keepaspectratio]{appendix_files/figure-pdf/fig-hex-param-1.pdf}
+
+}
+
+\caption{\label{fig-hex-param}The components of the hexagon grid
+illustrating notation.}
+
+\end{figure}%
+
 Since \(a_1\) and \(a_2\) are distances,
@@
 same logic follows for Equation~\ref{eq-equation4} and substituting
-\(a_2 = \frac{\sqrt{3}}{2}a_1\), and \(s_2 = -q \times r_2\) to
+\(a_2 = \sqrt{3}a_1/{2}\), and \(s_2 = -q \times r_2\) to
 Equation~\ref{eq-equation4} can be written as,
@@
 
-Also, substituting \(a_2 = \frac{\sqrt{3}}{2}a_1\),
-\(s_2 = -q \times r_2\) and rearranging Equation~\ref{eq-equation5}
-gives:
+Also, substituting \(a_2 = \sqrt{3}a_1/{2}\), \(s_2 = -q \times r_2\)
+and rearranging Equation~\ref{eq-equation5} gives:
 
@@
 
-Therefore, if this equality is true, then
-\(a_1 = \frac{(1+q)}{(b_1 - 1)}\), otherwise,
-\(a_1 = \frac{2r_2(1+q)}{\sqrt{3}(b_2 - 1)}\).
-
-\newpage
+Therefore, if this equality is true, then \[
+a_1 = \frac{(1+q)}{(b_1 - 1)},
+\] otherwise, \[
+a_1 = \frac{2r_2(1+q)}{\sqrt{3}(b_2 - 1)}.
+\]
 
@@
 
-\newpage
-
 \section{Area of a hexagon}\label{area-of-a-hexagon}
@@
 
+The layout a (Figure~\ref{fig-pbmc-mse-umap}) is generated from the
+hyper-parameters suggested by \citet{chen2024}, and the layout b
+(Figure~\ref{fig-pbmc-mse-umap}) is with suggested hyper-parameters by
+scDEED to be more accurate. The layout a and b contain \(46\) and \(83\)
+dubious cells respectively.
+
 \begin{figure}[H]
@@
 \end{figure}%
-
-The layout a (Figure~\ref{fig-pbmc-mse-umap}) is generated from the
-hyper-parameters suggested by \citet{chen2024}, and the layout b
-(Figure~\ref{fig-pbmc-mse-umap}) is with suggested hyper-parameters by
-scDEED to be more accurate. The layout a and b contain \(46\) and \(83\)
-dubious cells respectively.
 

### 2025-06-09 paper.qmd
some changes

@@
 ---
+
+XXX TO DO:
+
+- Make sure langevitour is explained before using it
+- Too much code needs to be streamlined
+- Remove all the extraneous commented text
+- Ensure that the SCdeed paper is mentioned in main part of this paper
 
@@
 
-## Best fit, or at least avoiding an inaccurate representation
-
-@fig-toy-mse shows a summary of plots that can be used to help assess the strength of the fit, and compare the fits for different representations.
+## Assessing the best fit, and hence most accurate 2D layout
+
+@fig-toy-mse illustrates the approach to compare the fits for different representations and assess the strength of any fit.
 What does it mean to be a best fit for this problem?
-There probably isn't a best fit, but there are wrong representations.
-The goal is to help users decide on a useful and appropriate low-dimensional representation of the high-dimensional data.
-
-Deciding on the best fit relies on several elements:
-
-- the choice of NLDR method, and the parameters used to create it, and
-- model fit parameters: bin size, low density bin removal.
-
-Comparing the MSE to obtain the best fit is suitable if one starts from the same NLDR representation.
-In theory, because the MSE is computed on \pD{} measuring the fit between model and data it might still be useful to compare different NLDR representations.
-A good NLDR representation should produce a good fit, producing a low MSE if the model fits the data well.
-However, it technically might be quite variable.
+Analysts use an NLDR layout to display the structure present in high-dimensional data in a convenient 2D display.
+It is a competitor to linear dimension reduction that can better represent non-linear associations such as clusters.
+However, these methods can hallucinate, suggesting patterns that don't exist, and grossly exaggerate other patterns.
+Having a layout that best fits the high-dimensional structure is desirable but more important is to identify bad representations so they can be avoided.
+The goal is to help users decide on a the most useful and appropriate low-dimensional representation of the high-dimensional data.
+
+A particular pattern that we commonly see is that analysts tend to pick layouts with clusters that have big separations between them.
+When you examine their data in a tour, it is almost always that we see there are no big separations, and actually the suggested clusters are not even present.
+While we don't expect that analysts include animated gifs of tours in their papers, we should expect that any 2D representation adequately indicates the clustering that is present, and honestly show lack of separation or lack of clustering when it doesn't exist.
+It is important for analysts to have tools to select the accurate representation not the pretty but wrong representation.
+
+To compare and assess a range of representations needs these items:
+
+- a selection of NLDR representations made with a range of parameter choices and possibly different methods (tSNE, UMAP, ...).
+- a range of model fits made by varying bin size and low density bin removal.
+- calculated RMSE for each layout, when it is transformed into the \pD{} space.
+
+Comparing the RMSE to obtain the best fit is appropraite if the same NLDR method is used.
+However, because the RMSE is computed on \pD{} data it measures the fit between model and data so it can also be used to compare the fit of different NLDR methods.
+A lower RMSE indicates a better NLDR representation.
 
@@
 
-## Curiosities
-
-<!-- XXX THIS IS NOT INTERESTING IN ITS CURRENT FORM. THIS IS ONLY ANY INTERESTING EXAMPLE WHEN THE FILLED OUT VS FLAT SHAPES ARE DISCUSSED. -->
+## Curiosities discovered by examining the model in the data space
 
@@
 
-### Ordering of points
-
-<!-- DHC: Will need to have more bins in each of the methods to illustrate this properly. -->
-
-To illustrate a difference in how the methods organise points in the NLDR layout, simulated $4\text{-}D$ data having five Gaussian clusters is used.
+### Some methods appear to order points in the layout
+
+When examining the \gD{} model representations it appears to be flatter or like a pancake with some methods, especially PACMAP, when the data is structure is higher than two dimensional.
+A simple example of this can be seen with data simulated to contain five $4\text{-}D$ Gaussian clusters.
+Each cluster is essentially a ball in $4\text{-}D$, so there is no \2D{} representation, rather the model in each cluster should resemble a crumpled sheet of paper that fills out $4\text{-}D$.
+
 @fig-five-gau-projs a1, b1, c1 show the \gD{} layouts for (a) tSNE, (b) UMAP, and (c) PaCMAP, respectively.
-The default hyper-parameters are used.
-All three methods show the five clusters, with varying degrees of separation.
-
-The models are fitted to these layouts, but we focus on a single cluster to illustrate the curious detail.
-@fig-five-gau-projs a2, b2, c2 show the fitted models in a projection of the $4\text{-}D$ space.
-In this projection the difference between methods can be seen.
-These clusters are fully $4\text{-}D$ in nature, so we would expect the model to be a *crumpled \gD{} sheet* that stretches in all four dimensions.
-This is what is observed for tSNE and UMAP.
-The curious detail is that the model for PaCMAP is closer to a *pancake* in shape!
-What this means is that there has to be some ordering of points in the \gD{} PaCMAP layout that induces the flat model, likely that the points are organised by the global principal components.
+The default hyper-parameters for each method are used.
+In each layout we can see an accurate representation where all five clusters are visible, although with varying degrees of separation.
+
+The models are fitted to each these layouts.
+@fig-five-gau-projs a2, b2, c2 show the fitted models in a projection of the $4\text{-}D$ space, taken from a tour.
+These clusters are fully $4\text{-}D$ in nature, so we would expect the model to be a *crumpled sheet* that stretches in all four dimensions.
+This is what is mostly observed for tSNE and UMAP.
+The curious detail is that the model for PaCMAP is closer to a *pancake* in shape in every cluster!
+This seingle projection only shows this in three of the five clusters but if we examine a different projection the other clusters exhibit the pancake also.
+While we don't know what exactly causes this it is likely due to some ordering of points in the \gD{} PaCMAP layout that induces the flat model.
+One could imagine that if the method used principal components on all the data, that it might induce some ordering that would produce the flat model.
+If this were the reason, the pancaking would be the same in all clusters, but it is not: The pancake is visible in some clusters in some projections but in other clusters it is visible in different projections.
+It might be due to some ordering by nearest neighbours in a cluster.
+The PaCMAP documentation doesn't provide any helpful clues.
+That this happens, though, makes the PaCMAP layout inadequate for representing the high-dimensional data.
+<!-- likely that the points are organised by the global principal components.
 So the layout of the model in $4\text{-}D$ doesn't reflect the dimension of each cluster.
 This can be happen because PaCMAP balances three components (near, mid-range, and far) rather than focusing solely on near neighbors.
-Mid-range neighbors ensure that transitions between clusters or groups are preserved and less concern with local preservation.
-
-<!-- DHC: is there something in the documentation that might help to explain what has happened? -->
-
-<!-- This pattern can also be observed by examining the error from the model fit (@fig-five-gau-projs a3, b3, c3).
-With tSNE and UMAP the error is relatively uniformly distributed, but with PaCMAP there is more error in the center of the \gD{} layout, reflecting that there are more points in the middle that are further from the fitted model.
--->
+Mid-range neighbors ensure that transitions between clusters or groups are preserved and less concern with local preservation.-->
 
@@
 
-### The effect of density {#sec-effect-dens}
+### Low density results in a squishing of the 2D representation in all methods {#sec-effect-dens}
+
+Differences in density can be due to be sampling at different rates in different subspaces of \gD{}.
+For example, the data shown in @fig-one-dens_clust-error all lies on a \2D{} curved sheet in $4\text{-}D$, but one end of the sheet is sampled densely and the other very sparsely.
+It was simulated to illustrate the effect on layout generated by tSNE.
 
@@
 
+Other tour software instead of langevitour.
+
 <!-- In conclusion, while our method effectively captures and represents high-dimensional data structures, further enhancements could involve introducing approaches to diagnose the fitted model.

### 2025-06-09 paper.qmd
using residual once model is fitted, and made a short cut for 4D

@@
   \newcommand\gD{$2\text{-}D$}
+  \newcommand\fD{$4\text{-}D$}
 ---
@@
 The model here is similar to a confirmatory factor analysis model [@brown2015], $\widehat{T}(X_1, X_2, X_3, X_4) + \Epsilon$.
-The difference between the fitted model and observed values would be considered to be residuals, and for this problem are $4\text{-}D$.
+The difference between the fitted model and observed values would be considered to be residuals, and for this problem are \fD{}.
 
@@
 
-## Curiosities discovered by examining the model in the data space
+## Curiosities about NLDR results discovered by examining the model in the data space
 
@@
 When examining the \gD{} model representations it appears to be flatter or like a pancake with some methods, especially PACMAP, when the data is structure is higher than two dimensional.
-A simple example of this can be seen with data simulated to contain five $4\text{-}D$ Gaussian clusters.
-Each cluster is essentially a ball in $4\text{-}D$, so there is no \2D{} representation, rather the model in each cluster should resemble a crumpled sheet of paper that fills out $4\text{-}D$.
+A simple example of this can be seen with data simulated to contain five \fD{} Gaussian clusters.
+Each cluster is essentially a ball in \fD{}, so there is no \2D{} representation, rather the model in each cluster should resemble a crumpled sheet of paper that fills out \fD{}.
 
@@
 The models are fitted to each these layouts.
-@fig-five-gau-projs a2, b2, c2 show the fitted models in a projection of the $4\text{-}D$ space, taken from a tour.
-These clusters are fully $4\text{-}D$ in nature, so we would expect the model to be a *crumpled sheet* that stretches in all four dimensions.
+@fig-five-gau-projs a2, b2, c2 show the fitted models in a projection of the \fD{} space, taken from a tour.
+These clusters are fully \fD{} in nature, so we would expect the model to be a *crumpled sheet* that stretches in all four dimensions.
 This is what is mostly observed for tSNE and UMAP.
@@
 <!-- likely that the points are organised by the global principal components.
-So the layout of the model in $4\text{-}D$ doesn't reflect the dimension of each cluster.
+So the layout of the model in \fD{} doesn't reflect the dimension of each cluster.
 This can be happen because PaCMAP balances three components (near, mid-range, and far) rather than focusing solely on near neighbors.
@@
 
-Differences in density can be due to be sampling at different rates in different subspaces of \gD{}.
-For example, the data shown in @fig-one-dens_clust-error all lies on a \2D{} curved sheet in $4\text{-}D$, but one end of the sheet is sampled densely and the other very sparsely.
-It was simulated to illustrate the effect on layout generated by tSNE.
-
-To illustrate how tSNE organize different number of points within the structure in the tSNE layout, simulated $4\text{-}D$ data having C-shaped structure is used.
-@fig-one-dens_clust-error b shows \gD{} layouts for tSNE.
-The default hyper-parameters are used.
-The \gD{} layout is colored according to the $4\text{-}D$ model error (@fig-one-dens_clust-error a).
+Differences in density can arise by sampling at different rates in different subspaces of \gD{}.
+For example, the data shown in @fig-one-dens_clust-error all lies on a \2D{} curved sheet in \fD{}, but one end of the sheet is sampled densely and the other very sparsely.
+It was simulated to illustrate the effect of the density difference on layout generated by an NLDR, illustrated using the tSNE results.
+
+@fig-one-dens_clust-error (a2, b2) shows \gD{} layout for tSNE created using the default hyper-parameters.
+One would expect to see a rectangular shape if the curved sheet is flattened, but the layout is triangular.
+The other two displays show the residuals as a dot density plot (a1, b1), and a \2D{} projection of the data and the model from \fD{} (a3, b3).
+Using the linked brushing tool, we can highlight points in the residuals
+
+The \gD{} layout is colored according to the \fD{} model error (@fig-one-dens_clust-error a).
 In @fig-one-dens_clust-error a, the large model error points are clustered in the top corner, forming a separate cluster with high model errors.
@@
 The model is fitted to this layout.
-@fig-one-dens_clust-error c shows the $2\text{-}D$ projection of the fitted model.
-In this projection, the distribution of $4\text{-}D$ model errors can be observed.
+@fig-one-dens_clust-error c shows the \gD{} projection of the fitted model.
+In this projection, the distribution of \fD{} model errors can be observed.
 The dense C-shaped structure shows high model errors at the sparse end.
@@
 With the model fit, several observations can be made.
-The simulated $4\text{-}D$ data, which contains five Gaussian clusters, was analyzed.
+The simulated \fD{} data, which contains five Gaussian clusters, was analyzed.
 Both tSNE and UMAP demonstrated the fitted model as a *crumpled \gD{} sheet* that stretches across all four dimensions, highlighting its ability to preserve local structure.
@@
 Another usability of our approach is how NLDR methods behave with different density of the data structure.
-The tSNE layout of simulated $4\text{-}D$ C-shaped data with different densities are used.
-The high $4\text{-}D$ model errors are scattered to a corner of the \gD{} layout of dense C-shaped structure.
+The tSNE layout of simulated \fD{} C-shaped data with different densities are used.
+The high \fD{} model errors are scattered to a corner of the \gD{} layout of dense C-shaped structure.
 In the dense C-shaped structure, the high model errors occur because there are fewer data points positioned within the bins at the sparse end of the structure.

### 2025-06-09 paper.qmd
finished with Section 5, Curiosities

@@
 
-Differences in density can arise by sampling at different rates in different subspaces of \gD{}.
+Differences in density can arise by sampling at different rates in different subspaces of \pD{}.
 For example, the data shown in @fig-one-dens_clust-error all lies on a \2D{} curved sheet in \fD{}, but one end of the sheet is sampled densely and the other very sparsely.
@@
 One would expect to see a rectangular shape if the curved sheet is flattened, but the layout is triangular.
-The other two displays show the residuals as a dot density plot (a1, b1), and a \2D{} projection of the data and the model from \fD{} (a3, b3).
-Using the linked brushing tool, we can highlight points in the residuals
-
-The \gD{} layout is colored according to the \fD{} model error (@fig-one-dens_clust-error a).
-In @fig-one-dens_clust-error a, the large model error points are clustered in the top corner, forming a separate cluster with high model errors.
-
-The model is fitted to this layout.
-@fig-one-dens_clust-error c shows the \gD{} projection of the fitted model.
-In this projection, the distribution of \fD{} model errors can be observed.
-The dense C-shaped structure shows high model errors at the sparse end.
-In the dense C-shaped structure, the high model errors occur because there are fewer data points positioned within the bins at the sparse end of the structure.
+The other two displays show the residuals as a dot density plot (a1, b1), and a \gD{} projection of the data and the model from \fD{} (a3, b3).
+Using linked brushing between the plots, we can highlight points in the tSNE layout, and examine where they fall in the original \fD{}.
+The darker (maroon) points indicate points that have been highlighted by linking.
+In row a, the points at the top of the triangle are highlighted, and we can see these correspond to higher residuals, and also to all points at the low density end of the curved sheet.
+In row b, points at the lower left side of the triangle are highlighted which corresponds to smaller residuals and one corner of the sheet at the high density end of the curved sheet.
+
+The tSNE behaviour is to squeeze the low density area of the data together into the layout.
+This is common in other NLDR methods also, which means analysts needs to be aware the if their data is not sampled relatively uniformly apparent closeness in the \gD{} may correspond to sparseness in \pD{}.
 

### 2025-06-09 paper.qmd
revised discussion and supp materials and change round brackets

@@
 - Remove all the extraneous commented text
-- Ensure that the SCdeed paper is mentioned in main part of this paper
+- Check how \kD{} and \gD{} are used, and make sure it's sensible
 
@@
 
-With our goal being to make a representation of this \gD{} layout that can be lifted into high-dimensional space, the layout needs to be augmented to include neighbour information.
+With our goal being to make a representation of this \gD{} layout that can be lifted into high-dimensional space, the layout needs to be augmented to include neighbor information.
 A simple approach would be to triangulate the points and add edges.
@@
 
-Delaunay triangulation [@lee1980;@alb2024] is used to connect points so that edges indicate neighbouring observations, in both the NLDR layout (@fig-NLDR-two-curvy (d)) and the \pD{} model representation.
+Delaunay triangulation [@lee1980;@alb2024] is used to connect points so that edges indicate neighboring observations, in both the NLDR layout (@fig-NLDR-two-curvy (d)) and the \pD{} model representation.
 When the data has been binned the triangulation connectd centroids.
@@
 If this were the reason, the pancaking would be the same in all clusters, but it is not: The pancake is visible in some clusters in some projections but in other clusters it is visible in different projections.
-It might be due to some ordering by nearest neighbours in a cluster.
+It might be due to some ordering by nearest neighbors in a cluster.
 The PaCMAP documentation doesn't provide any helpful clues.
@@
 
-Single-cell data refers to the characteristics of individual cells within a population (@haque2017).
+Single-cell data refers to the characteristics of individual cells within a population [@haque2017].
 These data also provides insights at the level of individual cells, enabling deeper understanding of gene expression.
@@
 
-We have developed an algorithm to assess the NLDR method and (hyper-)parameter choice(s) and choose the most accurate representation of the structure(s) present in the \pD{} data.
+We have developed an approach to help assess and compare NLDR layouts, generated by different methods and (hyper-)parameter choice(s).
+It depends on conceptualising the \gD{} layout as a model, creating a wireframe representation of it that can be lifted into \pD{}.
+The fit is assessed by viewing the model in the data space, computing residuals and RMSE.
+Different layouts can be compared using the RMSE, and provides quantitative and objective methods for deciding on the most suitable NLDR layout to represent the \pD{} data.
+It also provides a way to predict the values of new \pD{} observations in the \gD{}, which could be useful for implementing uncertainty checks such as using training and testing samples.
+
+Two examples illustrating usage are provided: the PBMC3k data where the NLDR is summarising clustering in \pD{} and hand-written digits illustrating how NLDR represents an intrinsically lower dimensional nonlinear manifold.
+We demonstrated that the published layout of the PBMC3k is inaccurate, because it grossly exaggerates separation between clusters, and even suggests separation when there is none.
+This is common when layouts are chosen subjectively -- often a preference for the "prettiest".
+Our approach provides a way to objectively choose the layout and hopefully avoids the use of misleading layouts in the future.
+In the hand-written digits we illustrate how our model fit statistics show that a flat disc layout is superior to the curved shaped layouts, and how to identify oddly written "1"s using the residuals of the fitted model.
+
+Additional exploration of metrics to summarise the fit could be a new direction for the work.
+The difficulty is capturing nonlinear fits, for which Euclidean distance can be sub-optimal.
+We have used a very simple approach based on clustering methods, Euclidean distances to nearest centroid, which can approximate nonlinear patterns.
+Other cluster metrics would be natural choices to explore.
+
+This new method also reveals some interesting curiosities about NLDR procedures.
+The fitted model appears as a "pancake" in some data where clusters are regularly shaped and high-dimensional, for some methods but not others, which is odd.
+One can imagine that if algorithms are initiated using principal components then some ordering of points along the major axes might generate this pattern.
+Alternatively, if local distances dominate the algorithm then is might be possible to see this pattern with well-separated regular clusters.
+We also demonstrated that there is a tendency for NLDR algorithms to be confused by different density in the data space, and some patterns in the layout are due to density differences rather than nonlinear associations between variables.
+
+Most NLDR methods only provide a \gD{} but if a \kD{} ($k>2$) layout is provided the approach developed here could be extended.
+Binning into cubes could be done in $3\text{-}D$ or higher, relatively easily, and used asa the basis for a wireframe of the fitted model.
+@barber1996 and the software @cxhull have algorithms for convex hulls, which \pD{} which serve as an inspiration.
+A simpler approach using $k$-means clustering to provide centroids could also be possible, but the complication would be to determine how to connect the centroids into an appropriate wireframe.
+
+The new methodology is accompanied by an R package called quollr, so that it is readily usable and broadly accessible.
+The package has methods to fit the model, compute diagnostics and also visualise the results, with interactivity.
+We have primarily used the langevitour software [@harisson2024] to view the model in the data space, but other tour software such as tourr [@tourr] and [@detourr] could be also used.
+
+<!-- and choose the most accurate representation of the structure(s) present in the \pD{} data.
 Starting from \gD{} layout, the fitted model, as represented by the positions of points in \kD{}, and turn it into a \pD{} wireframe to overlay on the data, viewing it with a tour.
@@
 Additionally, the model discovers non-uniform data distribution and non-linear structures within the clusters that are not visible in the \gD{} layout, demonstrating the ability of our model in uncovering hidden data characteristics.
-Also, as explained in Appendix, with recently introduced approach, scDEED also proved the layout suggested by our layout is better than the published one.
+Also, as explained in Appendix, with recently introduced approach, scDEED [@xia2023] also proved the layout suggested by our layout is better than the published one.
 
@@
 Other tour software instead of langevitour. 
+-->
 
@@
 
-Appendix: The appendix includes more details about the hexagonal binning algorithm and scDEED method (appendix.pdf, Portable Document Format file).
-
-R package `quollr`: The R package `quollr` containing codes to fit, and visualize the model. (need to add quollr .zip file, GNU zipped tar file)
+The GitHub repository for the paper can be found at https://github.com/JayaniLakshika/paper-nldr-vis-algorithm.
+The Appendix includes more details about the hexagonal binning algorithm and a comparison to the results of the newly reported scDEED [@xia2023] statistic.
+
+The R package `quollr`, available on CRAN and at https://jayanilakshika.github.io/quollr/, provides software accompaying this paper to fit the wireframe model representation, compute diagnostics, visualize the model in the data with langevitour and link multiple plots interactively.
 
@@
 
-These `R` packages were used for the work: `tidyverse` (@hadley2019), `Rtsne` (@jesse2015), `umap` (@tomasz2023), `patchwork` (@thomas2024), `colorspace` (@achim2020), `langevitour` (@harisson2024), `conflicted` (@hadley2023), `reticulate` (@kevin2024), `kableExtra` (@hao2024).
-These `python` packages were used for the work: `trimap` (@amid2019) and `pacmap` (@yingfan2021).
-The article was created with `R` packages `quarto` (@jjallaire2024).
-The project's GitHub repository ([https://github.com/JayaniLakshika/paper-nldr-vis-algorithm](https://github.com/JayaniLakshika/paper-nldr-vis-algorithm)) contains all materials required to reproduce this article.
+These `R` packages were used for the work: `tidyverse` [@hadley2019], `Rtsne` [@jesse2015], `umap` [@tomasz2023], `patchwork` [@thomas2024], `colorspace` [@achim2020], `langevitour` [@harisson2024], `conflicted` [@hadley2023], `reticulate` [@kevin2024], `kableExtra` [@hao2024].
+These `python` packages were used for the work: `trimap` [@amid2019] and `pacmap` [@yingfan2021].
+The article was created with `R` packages `quarto` [@jjallaire2024].
 

### 2025-06-09 paper.qmd
Added tourr and detourr references

@@
 The package has methods to fit the model, compute diagnostics and also visualise the results, with interactivity.
-We have primarily used the langevitour software [@harisson2024] to view the model in the data space, but other tour software such as tourr [@tourr] and [@detourr] could be also used.
+We have primarily used the langevitour software [@harisson2024] to view the model in the data space, but other tour software such as tourr [@wickham2011] and detourr [@hart2022] could be also used.
 

### 2025-06-09 paper.qmd
Removed long comments

@@
 
-<!-- - What's the problem:
-
-  Non-linear dimension reduction being used to summarise high-dimensional data.
-
-  - Summary of literature
-
-  Relevant high-d vis, NLDR history
--->
-
 ```{r}
@@
 
-<!-- Linked brushing as done by @article21
-
-- Model-in-the-data-space: how can we represent the model, eg plane for PCA, grid of values for classification boundaries, ellipses for LDA and mclust, nets for SOM.-->
-
 ## Method {#sec-method}
@@
 (Note that, for this thinking the full \pD{}   data needs to be available, not just the interpoint distances.)
-<!--### Notation -->
-
-<!-- @tbl-notation summarises the notation used to explain the new methodology.
-The observed data is denoted as $\mathbfit{x}_{n \times p}$ where $x_{ij}$ would indicate the $i^{th}$ observation on the $j^{th}$ variable sampled from a population $\mathbfit{X}$.
-To refer to variable $j$, we would use $X_j$.-->
-
-<!--
-$X_{n \times p} = \begin{bmatrix} \textbf{x} _{1} & \textbf{x}_ {2} & \cdots & \textbf{x}_{n} \\  \end{bmatrix}^\top$
-
-$Y_{n \times d} = \begin{bmatrix} \textbf{y} _{1} & \textbf{y}_ {2} & \cdots & \textbf{y}_{n} \\  \end{bmatrix}^\top$
-
-$C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2})$
-
-$C_k^{(p)} \equiv (C_{kx_1}, ..., C_{kx_p})$ $p$-D mappings of 2D hexagon bin centroids of the $k^{th}$ hexagon
-
--->
 
@@
 When $k=2$, as in hexagon binning, the default range is $[0, y_{i,\text{max}}], i=1,2$, where $y_{1,\text{max}}=1$ and $y_{2,\text{max}} = r_2/r_1$ (@fig-NLDR-two-curvy).
-If the NLDR aspect ratio is ignored then set $y_ {2,\text{max}} = 1.$ <!-- \times \frac{2}{\sqrt{3}}$.
-(The $\frac{2}{\sqrt{3}}$ accounts for the different height ($a_1$) and width ($a_2$) of a regular hexagon.) The scaling of data should take the size of the hexagons into account, but choice of number of bins should.
--->
+If the NLDR aspect ratio is ignored then set $y_ {2,\text{max}} = 1.$
 
@@
 
-<!-- Number of bins is set by fixing $b_1$, which determines the binwidth accounting the offset $q_1$ and breaks on $x_1$, is calculated as  -->
-
-<!-- $$ -->
-<!-- a_1 = \frac{r_1 + q_1}{b_1}. -->
-<!-- $$ -->
-
-<!-- The computed binwidth then determines the number of vertical bins, $b_2$, and vertical binwidth, which are computed based on $r_2$, the offset $q_2$, and the height/width ratio of a regular hexagon, $\frac{2}{\sqrt{3}}$ as -->
-
-<!-- $$ -->
-<!-- a_2 = \frac{r_2 + q_2}{0.75 \times b_2}. -->
-<!-- $$ -->
-
-<!-- To define a hexagon grid across the \kD{} space, with hexagons of fixed height ($a_1$), width ($a_2$),  it is necessary to determine how many hexagons should be represented along each axis in the \kD{} space.
-
-When $k=2$, the number of bins along the $x$ and $y$ axes, $b_1$ and $b_2$, is computed by considering the scaling factors ${r_1, r_2}$, along with the offset along the axes ($q_1$, $q_2$), and the height ($a_1$) and width ($a_2$) of a regular hexagon (@fig-NLDR-two-curvy (b)).
-($b_1 = \frac{r_1 + q_1}{a_2}$, and $b_2 = \frac{r_2 + q_2}{0.75 \times a_1}$) where $0.75 \times a_1$ accommodate to have hexagons fill the entire 2D space without leaving any gaps between them.)-->
-
-<!-- XXX We don't need parameters for the regular hexagon, these are fixed constants. -->
-
 #### Binning the data
-
-<!-- Points are allocated to the bin they fall into based on the nearest centroid.
-In situations where a point is equidistant from multiple centroids, tie-breaking rules are applied.
-If multiple centroids are in the same row, the point is assigned to the leftmost centroid.
-If multiple centroids are in different rows, the point is assigned to the bottom centroid.
--->
-
-<!-- $\{ i \in H_h, h = 1, \dots, b, \text{ and } i = 1, \dots, n\}$ -->
 
@@
 
-<!--
-Define a hexagon grid across the \kD{} space, using hexagons with fixed height ($a_2$), width ($a_1$).
-Each of the \kD{} points will belong to a hexagon bin.
-That is, for each $y \in \mathbfit{Y}$, we can (uniquely) identify the hexagon that the point belongs to.
-This identification is done finding the nearest bin centroid for the \kD{} points by considering the \kD{} Euclidean distance.
-
-When $k=2$, the starting coordinates $(s_1, s_2)$ mark the lower left of the grid.
-This is the bottom left bin centroid.
-By starting from there, points are generated to fill the grid accounting $b_1$, $b_2$, $a_1$, and $a_2$.
--->
-
-<!--We deliberately separate out the creation of the hexagon grid from the mapping of points on the grid.-->
-
 #### Indicating neighborhood
@@
 
-<!-- When $k = 2$ Delaunay triangulation on $C^{(2)}$ generates the model in \gD{} space, which is a triangular mesh (@fig-NLDR-two-curvy (d)).
-It generates convex hulls of $C^{(2)}$ such that the circumcircle of every triangle in the triangulation contains no other points from $C^{(2)}$.
--->
-
 When shapes are non-linear in the NLDR layout, some edges could be long.
@@
 In order to generate a smooth surface in \gD{}, these long line segments should be removed when tuning the model fit.
-
-<!--need to add what is meant by a long edge-->
 
@@
 
-<!-- XXX You need to have consistent colouring. In langevitour the points corresponding to observations are purple. They should be purple in the \gD{} view too.  -->
-
-<!-- XXX Should there be a second long edge removal? You can compute the edges in \pD{} now, and if there are long edges maybe these should be pruned, and we re-plot the \gD{} view.  -->
-
-<!-- langevitour with pD model for two-curvy clusters -->
 ```{r}
@@
 
-<!--The prediction approach involves finding the nearest \kD{} model point for a new \pD{} point.
-We define the function $z: \mathbb{R}^{n\times p} \rightarrow \mathbb{R}^{m\times p}$, where $z(i) = \arg\min_{j = 1, \dots, b} \sqrt{\sum_{v=1}^{p}(x_{iv} - C^{(p)}_{jv})^2}$ maps each \pD{} point to its nearest \pD{} mapping of the model.
-Therefore, the new observation $i$ falls in the hexagon, $H_h = \{i| z(i) = h\}$ and the corresponding \kD{} bin centroids, $C_{h}^{(2)} = (c_{h1}, c_{h2})$ be the predicted values.-->
-
-<!-- ::: {#fig-two_curvy-pred-sc layout-ncol="2" fig-pos="H"} -->
-<!-- ![](figures/two_curvy/sc_true_only.png){width="150" fig-align="center"} -->
-
-<!-- ![](figures/two_curvy/pred_true_view.png){width="150" fig-align="center"} -->
-
-<!-- A view of the true model in projections from $7\text{-}D$, and predictions of the true model in \gD{}, for the S-curve data.
-The predictions fits the UMAP layout which means that it capture the geometry of S-curve with UMAP.
--->
-<!-- ::: -->
-
-<!-- XXX Prediction should have a jitter option, to spread the predicted points out fully in the hexagons. -->
-
 ### Tuning
-<!-- removal of low density bins, removing long edges, choice of bins-->
 
@@
 Generally, a more uniform distribution among these possibilities would indicate that the bins are reliably capturing the underlying distribution of observations.
-
-<!--The starting position of the hexagonal grid is important because different starting points result in different distributions of data across bins, even with the same total number of bins.
-This variation affects the model due to the differing number of non-empty bins.
-Therefore, it is necessary to evaluate various starting points with different total numbers of bins to determine which configurations are more effective at capturing the structure and fitting the model.-->
 
@@
 
-<!-- To determine the effective $b$, candidate values are selected based on the range between the minimum and approximate maximum $b_1$, because $b_2$ is computed from $b_1$.
-The minimum $b_1$ is set to $2$, while the maximum number is estimated by taking the square root of $\frac{n}{2}$.
-By evaluating MSE across varying $b$ within this range for different $q$, helps to determine an appropriate values for $b$ and $q$ (@fig-param-two_curvy (a)).-->
-
-<!--add MSE vs total number of error plot-->
-
-<!--To generate errors for different total number of bins-->
-
 #### Measurement of capturing the data shape in 2-D 
-
 The area of a hexagon is defined as $A = 3\sqrt{3}l^2/2$ where $l$ is the side length of the hexagon.
@@
 
-<!-- Once setting up the hexagon grid with an appropriate number of bins, some hexagon bins may have few or no data points within them (@fig-bins-two_curvy (b)).
-To ensure comprehensive coverage of the NLDR data, it is necessary to select hexagon bins with a considerable number of data points.
-This involves calculating the number of points within each hexagon.
-Then, the standard count is computed by dividing the number of points within each hexagon by the maximum number of points in the grid.
-Next, bins with a standard count less than a benchmark value are removed (@fig-param-two_curvy (c)).
-There is no specific rule for selecting a benchmark value.
-However, the following steps can help determine a suitable value for removing low-density hexagons:
-
-1. Plot the distribution of the standardized counts (@fig-param-two_curvy (b)).
-2. Examine the distribution of counts.
-3. Select the first quantile value if the distribution is skewed.
--->
-
 The benchmark value for removing low-density hexagons ranges between $0$ and $1$.
@@
 If the increment is not considerable, the decision might lean towards retaining low-density hexagons.
-
-<!--add MSE vs density (0 to 1)-->
-
-<!--base line -->
-
-<!-- Furthermore, selecting the benchmark value for removing low-density hexagons is important.
-Removing unnecessary bins may lead to the formation of long edges and an uneven \gD{} model.
-Hence, rather than solely relying on the benchmark value to identify hexagons for removal, it's essential to consider the standard number of points in the neighboring hexagons of the identified low-density bins (see @fig-lwd-two_curvy (b)).
-If neighboring bins also show low counts, only those bins will be removed.
-The remaining bins are used to construct the \gD{} model.
--->
 
@@
 
-<!-- #### Removing long edges -->
-
-<!-- Edges define the neighbourhood structure, in order to provide a smooth \gD{} representation of the fitted model.
-@fig-two-curvy-true-proj shows a wire frame of the true model that was used to generate the non-linear clusters example data.
-The ideal is that the representation of the fitted model, at least for this example where we know the true model, should look similar to this.
--->
-
-<!-- The Delaunay triangulation will ensure that all centroids are connected into a triangular mesh.
-For some structures, like clustered data, or highly non-linear shapes, breaks in the mesh are meaningful.
-When separated clusters are present the mesh should be broken across the gaps.
-For non-linear structures like the C-shaped curvilinear, the mesh should run unbroken along the C, but there should be no edges connecting the top of the C directly to the bottom of the C.
-For these reasons it is necessary to remove edges from the mesh in some applications.
--->
-
-<!-- The decision on edge length removal is made based on the distribution of edge lengths.
-In particular, a gap between values, where there a concentration of small values and then a few larger values, likely suggests a cutoff for edge removal.
-Because the triangulation is typically done on the hexagon centroids, there are particular discrete edge lengths, based on bin widths.
-@fig-rm-lg illustrates edge length distributions.
--->
-
-<!-- There is an additional step that is needed.
-When the model is lifted into \pD{}, if the fit is good all the edges should be relatively small in this space, too.
-If this is not the case, then there are several possible actions: (1) re-do the NLDR to get a more representative layout; (2) identify the edge and remove it from the model, in \gD{} and \pD{}; (3) consider different values for the model fit, number of bins, initial bin position or removing low density bins.
--->
-
-<!-- ```{r} -->
-<!-- #| label: model-two-curvy-umap -->
-
-<!-- umap_two_curvy <- read_rds(file = "data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_umap_n-neigbors_15_min-dist_0.1.rds") -->
-
-<!-- num_bins_x_two_curvy <- 42 -->
-
-<!-- algo_obj_two_curvy3 <- gen_nldr_vis_algo_obj( -->
-<!--   high_d_data = training_data_two_curvy,  -->
-<!--   nldr_data = umap_two_curvy,  -->
-<!--   num_x_bins = num_bins_x_two_curvy) -->
-
-<!-- umap_two_curvy_scaled <- algo_obj_two_curvy3$nldr_scaled -->
-<!-- distance_two_curvy <- algo_obj_two_curvy3$distance_df -->
-<!-- tr_from_to_df_two_curvy <- algo_obj_two_curvy3$tr_from_to_df -->
-<!-- benchmark_two_curvy <- algo_obj_two_curvy3$benchmark -->
-<!-- df_bin_centroids_two_curvy <- algo_obj_two_curvy3$df_bin_centroids -->
-<!-- df_bin_two_curvy <- algo_obj_two_curvy3$df_bin -->
-<!-- hex_grid_with_counts <- algo_obj_two_curvy3$hex_grid_with_counts -->
-<!-- a1 <- algo_obj_two_curvy3$a1 -->
-
-<!-- trimesh_two_curvy_umap <- ggplot() + -->
-<!--   geom_segment(data = tr_from_to_df_two_curvy, -->
-<!--                aes( -->
-<!--                  x = x_from, -->
-<!--                  y = y_from, -->
-<!--                  xend = x_to, -->
-<!--                  yend = y_to), -->
-<!--                colour = "#000000", -->
-<!--                alpha = 0.5, -->
-<!--                linewidth = 0.5) + -->
-<!--   geom_point(data = umap_two_curvy_scaled, -->
-<!--              aes( -->
-<!--                x = UMAP1, -->
-<!--                y = UMAP2 -->
-<!--              ), -->
-<!--              color = "#636363", -->
-<!--              alpha = 0.5, -->
-<!--              size = 0.5 -->
-<!--   ) + -->
-<!--   coord_equal() + -->
-<!--   xlim(c(0, 1)) + -->
-<!--   ylim(c(0, 1)) + -->
-<!--   interior_annotation("a1", sc_ltr_pos, cex = 2) -->
-
-<!-- ``` -->
-
-<!-- ```{r} -->
-
-<!-- # With the computed benchmark value -->
-<!-- # distance_df_small_edges_two_curvy <- distance_two_curvy |> -->
-<!-- #   filter(distance < benchmark_two_curvy) -->
-
-<!-- ####With diff benchmark values -->
-
-<!-- ## Benchmark 2 -->
-
-<!-- benchmark_two_curvy <- 8 * a1 -->
-
-<!-- distance_df_small_edges_two_curvy3 <- distance_two_curvy |> -->
-<!--   filter(distance < benchmark_two_curvy) -->
-
-<!-- tr_from_to_df_two_curvy4 <- inner_join( -->
-<!--   tr_from_to_df_two_curvy, distance_df_small_edges_two_curvy3, -->
-<!--   by = c("from", "to")) -->
-
-<!-- trimesh_two_curvy_removed2_umap <- ggplot() + -->
-<!--   geom_segment(data = tr_from_to_df_two_curvy4, -->
-<!--                aes( -->
-<!--                  x = x_from, -->
-<!--                  y = y_from, -->
-<!--                  xend = x_to, -->
-<!--                  yend = y_to), -->
-<!--                colour = "#000000", -->
-<!--                alpha = 0.5, -->
-<!--                linewidth = 0.5)+ -->
-<!--   geom_point(data = umap_two_curvy_scaled, -->
-<!--              aes( -->
-<!--                x = UMAP1, -->
-<!--                y = UMAP2 -->
-<!--              ), -->
-<!--              color = "#636363", -->
-<!--              alpha = 0.5, -->
-<!--              size = 0.5 -->
-<!--   ) + -->
-<!--   coord_equal() + -->
-<!--   xlim(c(0, 1)) + -->
-<!--   ylim(c(0, 1)) + -->
-<!--   interior_annotation("b1", sc_ltr_pos, cex = 2) -->
-
-<!-- ## Benchmark 1 -->
-
-<!-- benchmark_two_curvy <- 2.5 * bin_width -->
-
-<!-- distance_df_small_edges_two_curvy2 <- distance_two_curvy |> -->
-<!--   filter(distance < benchmark_two_curvy) ## 0.231 -->
-
-<!-- tr_from_to_df_two_curvy3 <- inner_join( -->
-<!--   tr_from_to_df_two_curvy, distance_df_small_edges_two_curvy2, -->
-<!--   by = c("from", "to")) -->
-
-<!-- trimesh_two_curvy_removed1_umap <- ggplot() + -->
-<!--   geom_segment(data = tr_from_to_df_two_curvy3, -->
-<!--                aes( -->
-<!--                  x = x_from, -->
-<!--                  y = y_from, -->
-<!--                  xend = x_to, -->
-<!--                  yend = y_to), -->
-<!--                colour = "#000000", -->
-<!--                alpha = 0.5, -->
-<!--                linewidth = 0.5)+ -->
-<!--   geom_point(data = umap_two_curvy_scaled, -->
-<!--              aes( -->
-<!--                x = UMAP1, -->
-<!--                y = UMAP2 -->
-<!--              ), -->
-<!--              color = "#636363", -->
-<!--              alpha = 0.5, -->
-<!--              size = 0.5 -->
-<!--   ) + -->
-<!--   coord_equal() + -->
-<!--   xlim(c(0, 1)) + -->
-<!--   ylim(c(0, 1)) + -->
-<!--   interior_annotation("c1", sc_ltr_pos, cex = 2) -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- #| label: compare-2d-and-4d-dist -->
-
-<!-- ### Compute highD distance -->
-<!-- dist_vec <- proxy::dist(x = df_bin_two_curvy[, -1], method = "Euclidean") |> as.vector() -->
-
-<!-- from_vec <- c() -->
-<!-- to_vec <- c() -->
-<!-- num_obs <- 1:(NROW(df_bin_two_curvy) - 1) -->
-
-<!-- for (obs in num_obs) { -->
-
-<!--   from_val <- rep(obs, (NROW(df_bin_two_curvy) - obs)) -->
-<!--   if ((obs + 1) <= NROW(df_bin_two_curvy)) { -->
-<!--     to_val <- (obs + 1):NROW(df_bin_two_curvy) -->
-<!--   } -->
-<!--   from_vec <- append(from_vec, from_val) -->
-<!--   to_vec <- append(to_vec, to_val) -->
-
-<!-- } -->
-
-<!-- dist_highd <- tibble::tibble(from = from_vec, to = to_vec, dist_highd = dist_vec) -->
-
-<!-- ## To plot the distribution of distances -->
-
-<!-- # Define bin width and starting point -->
-<!-- start_point <- a1/2 -->
-
-<!-- # Create bins and calculate the mean value for each bin range -->
-<!-- bins <- seq(start_point, max(distance_two_curvy$distance) + a1, -->
-<!--             by = a1) -->
-<!-- bin_labels <- bins[-length(bins)] + a1 / 2  # mean of each bin range -->
-
-<!-- distance_two_curvy_dist <- distance_two_curvy |> -->
-<!--   mutate(bin = cut(distance, breaks = bins, include.lowest = TRUE, labels = bin_labels)) -->
-
-<!-- distance_two_curvy_dist <- left_join(distance_two_curvy_dist, dist_highd, by = c("from", "to")) -->
-
-<!-- text_df <- tibble::tibble( -->
-<!--   x = c(0.12, 0.43), -->
-<!--   y = c(10, 10), -->
-<!--   text = c("c", "b") -->
-<!-- ) -->
-
-<!-- distance_points_umap <- ggplot( -->
-<!--   distance_two_curvy_dist, -->
-<!--   aes(x = distance, -->
-<!--       y = dist_highd)) + -->
-<!--   geom_point(alpha = 0.5) + -->
-<!--   geom_text( -->
-<!--     data=text_df, -->
-<!--     aes(x=x, y=y, -->
-<!--         label=text), -->
-<!--     colour="#bdbdbd", -->
-<!--     size = 10) + -->
-<!--   geom_vline(xintercept = 8 * bin_width, -->
-<!--              linetype="solid", -->
-<!--              color = "#bdbdbd", -->
-<!--              linewidth=1) + -->
-<!--   geom_vline(xintercept = 2.5 * bin_width, -->
-<!--              linetype="solid", -->
-<!--              color = "#bdbdbd", -->
-<!--              linewidth=1) + -->
-<!--   ylab(expression(d^{(4)})) + -->
-<!--   xlab(expression(d^{(2)})) + -->
-<!--   theme_minimal() + -->
-<!--   theme(aspect.ratio = 0.75, -->
-<!--         panel.border = element_rect(fill = 'transparent'), -->
-<!--         plot.title = element_text(size = 20, hjust = 0.5, vjust = -0.5), -->
-<!--         panel.grid.major.x = element_blank(), -->
-<!--         axis.ticks.x = element_line(), -->
-<!--         axis.ticks.y = element_line(), -->
-<!--         axis.title = element_text(size = 15), -->
-<!--         axis.text = element_text(size = 12)) -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- #| label: proj-tsne-model -->
-
-<!-- df_bin_two_curvy <- df_bin_two_curvy |> -->
-<!--   select(-hexID) |> -->
-<!--   mutate(type = "model") -->
-
-<!-- # Apply the scaling -->
-<!-- df_model_data_two_curvy <- bind_rows(data_two_curvy, df_bin_two_curvy) -->
-
-<!-- scaled_two_curvy <- scale_data_manual(df_model_data_two_curvy, "type") |> -->
-<!--   as_tibble() -->
-
-<!-- scaled_two_curvy_data <- scaled_two_curvy |> -->
-<!--   filter(type == "data") |> -->
-<!--   select(-type) -->
-
-<!-- scaled_two_curvy_data_model <- scaled_two_curvy |> -->
-<!--   filter(type == "model") |> -->
-<!--   select(-type) -->
-
-<!-- ## First projection -->
-<!-- projection <- cbind( -->
-<!--   c(0.37847,-0.06517,0.04231,0.01655), -->
-<!--   c(-0.06466,-0.27052,0.22131,-0.15235)) -->
-
-<!-- proj_obj1 <- get_projection(projection = projection,  -->
-<!--                             proj_scale = 1,  -->
-<!--                             scaled_data = scaled_two_curvy_data,  -->
-<!--                             scaled_data_model = scaled_two_curvy_data_model,  -->
-<!--                             distance_df_small_edges = distance_df_small_edges_two_curvy2,  -->
-<!--                             axis_param = list(limits = 0.6,  -->
-<!--                                               axis_scaled = 2, -->
-<!--                                               axis_pos_x = -0.5,  -->
-<!--                                               axis_pos_y = -0.5,  -->
-<!--                                               threshold = 0.04)) -->
-
-<!-- ## Projection with Delauany triangulation -->
-
-<!-- projected_model <- as.matrix(scaled_two_curvy_data_model) %*% projection_scaled -->
-
-<!-- projected_model_df <- projected_model |> -->
-<!--   tibble::as_tibble(.name_repair = "unique") |> -->
-<!--   dplyr::rename(c("proj1" = "...1", -->
-<!--                   "proj2" = "...2")) |> -->
-<!--   dplyr::mutate(ID = dplyr::row_number()) -->
-
-<!-- model_df1 <- dplyr::left_join( -->
-<!--   distance_two_curvy |> select(-distance), -->
-<!--   projected_model_df, -->
-<!--   by = c("from" = "ID")) -->
-
-<!-- names(model_df1)[3:NCOL(model_df1)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from") -->
-
-<!-- model_df1 <- dplyr::left_join(model_df1, projected_model_df, by = c("to" = "ID")) -->
-<!-- names(model_df1)[(2 + NCOL(projected_model_df)):NCOL(model_df1)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to") -->
-
-<!-- ## With benchmark option 2 -->
-
-<!-- model_df2 <- dplyr::left_join( -->
-<!--   distance_df_small_edges_two_curvy3 |> select(-distance), -->
-<!--   projected_model_df, -->
-<!--   by = c("from" = "ID")) -->
-
-<!-- names(model_df2)[3:NCOL(model_df2)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from") -->
-
-<!-- model_df2 <- dplyr::left_join(model_df2, projected_model_df, by = c("to" = "ID")) -->
-<!-- names(model_df2)[(2 + NCOL(projected_model_df)):NCOL(model_df2)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to") -->
-
-<!-- two_curvy_proj_first_model1_umap <- plot_proj( -->
-<!--     proj_obj = proj_obj1,  -->
-<!--     point_param = c(0.5, 0.2, "#636363"), # size, alpha, color -->
-<!--     line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha -->
-<!--     plot_limits = c(-0.6, 0.6),  -->
-<!--     title = "c2",  -->
-<!--     cex = 2,  -->
-<!--     position = sc_ltr_pos, -->
-<!--     axis_text_size = 4,  -->
-<!--     is_category = FALSE) + -->
-<!--     theme( -->
-<!--         legend.position = "none" -->
-<!--     ) -->
-
-<!-- ## Projection with Delauany triangulation -->
-
-<!-- proj_obj1_temp1 <- proj_obj1 -->
-<!-- proj_obj1_temp2 <- proj_obj1 -->
-<!-- proj_obj1_temp1[["model_df"]] <- model_df1 -->
-
-<!-- two_curvy_proj_model_delaunay_umap <- plot_proj( -->
-<!--     proj_obj = proj_obj1_temp1,  -->
-<!--     point_param = c(0.5, 0.2, "#636363"), # size, alpha, color -->
-<!--     line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha -->
-<!--     plot_limits = c(-0.6, 0.6),  -->
-<!--     title = "a2",  -->
-<!--     cex = 2,  -->
-<!--     position = sc_ltr_pos, -->
-<!--     axis_text_size = 4,  -->
-<!--     is_category = FALSE) + -->
-<!--     theme(aspect.ratio = 1,  -->
-<!--           legend.position = "none") -->
-
-<!-- proj_obj1_temp2[["model_df"]] <- model_df2 -->
-
-<!-- two_curvy_proj_model_benchmark2_umap <- plot_proj( -->
-<!--     proj_obj = proj_obj1_temp2,  -->
-<!--     point_param = c(0.5, 0.2, "#636363"), # size, alpha, color -->
-<!--     line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha -->
-<!--     plot_limits = c(-0.6, 0.6),  -->
-<!--     title = "b2",  -->
-<!--     cex = 2,  -->
-<!--     position = sc_ltr_pos, -->
-<!--     axis_text_size = 4,  -->
-<!--     is_category = FALSE) + -->
-<!--     theme(aspect.ratio = 1,  -->
-<!--           legend.position = "none") -->
-
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- #| label: fig-rm-lg -->
-<!-- #| fig-cap: "The choice of long edge removal significantly impacts the model's accuracy.
-\\gD{} Vs $7\\text{-}D$ distances between the bin centroids, where the model is fitted with UMAP for the non-linear clusters data.
-The two benchmarks for long edge removal: (b) benchmark = 8$a_1$ and (c) benchmark = 2.5$a_1$, where $a_1$ = 0.03 are made.
-The wireframes in \\gD{} and a projection from $7\\text{-}D$ are shown with Delaunay triangulation (a1, a2) and the two benchmark values (b1, b2, c1, c2).
-With the benchmark set at 6.5$a_1$, long edges exist within the C-shaped cluster, which distrupts the model.
-In contrast, using 2$a_1$ is a better option, as it effectively captures the non-linear shapes of the clusters.
-To generate an accurate model that captures the data structure, the choice of long edge removal does matter." -->
-<!-- #| fig-height: 15 -->
-<!-- #| fig-width: 15 -->
-
-<!-- free(distance_points_umap) +  -->
-<!--     wrap_plots(trimesh_two_curvy_umap, trimesh_two_curvy_removed2_umap, -->
-<!--                trimesh_two_curvy_removed1_umap, two_curvy_proj_model_delaunay_umap, -->
-<!--                two_curvy_proj_model_benchmark2_umap, two_curvy_proj_first_model1_umap, -->
-<!--                ncol = 3) + -->
-<!--     plot_layout(heights = c(0.8, 1)) -->
-<!-- ``` -->
-
 ### Linked plots
@@
 It is important to link and brush the points with high model error in the \pD{} error plot, \gD{} layout, and the generated model overlay on the data in \pD{}.
-
-<!-- Brushing the high model error in the \gD{} layout, with dense C-shaped data explained in @sec-effect-dens, helps in locating the positions of the related \pD{} points in langevitour view.
-This process helps in identifying the high model errors that occur due to the sparse end (add the video).
--->
-
-<!-- Furthermore, when examining the digit 1 from the MNIST dataset (@sec-mnist), brushing the small cluster in the \gD{} layout helps in locating the positions of the related \pD{} points in langevitour view.
-This process helps in identifying how close the small cluster to the big cluster (add the video).
-We can also explore which data corresponds to the twisted pattern of the big cluster (add the video).
--->
 
@@
 That this happens, though, makes the PaCMAP layout inadequate for representing the high-dimensional data.
-<!-- likely that the points are organised by the global principal components.
-So the layout of the model in \fD{} doesn't reflect the dimension of each cluster.
-This can be happen because PaCMAP balances three components (near, mid-range, and far) rather than focusing solely on near neighbors.
-Mid-range neighbors ensure that transitions between clusters or groups are preserved and less concern with local preservation.-->
 
@@
 
-<!-- XXX WHY THESE EXAMPLES. NEEDS AN OVERVIEW PARAGRAPH -->
-
 To showcase the robustness of our method when applied to data with different characteristics, this section presents two case studies.
@@
 ### Single-cell gene expression {#sec-pbmc}
-<!--
-- NLDR view used to illustrate clusters
-- Use our method to assess is it a reasonable representation
-- Demonstrate that it is not
-- Illustrate how to use our method to get a better representation
--->
-
-<!-- XXX What is single cell data???? -->
 
@@
 Furthermore, the clusters shows non-linear shapes (@fig-model-pbmc-author-proj (b3)).
-
-<!-- As shown in @fig-NLDR-variety, layout c, d, and e that show small separations between clusters are universally optimal.
-But layout d with little separations perform well at tiny binwidth and poorly as binwidth increases.
-Also, layout e performs well at large binwidths.
-The layout c is chosen as the best, since it used the UMAP with a different (hyper-) parameter settings.
--->
-
-<!-- The visualization of the model generated for the selected layout in $9\text{-}D$ shows edges between the clusters (@fig-mnist-tri-proj (b2)).
-This provides evidence for the small separations between clusters.
-Furthermore, the data does not uniformally distributed along the clusters and therefore shows densed points (@fig-mnist-tri-proj (b2).
-Also, the clusters are non-linear shaped (@fig-mnist-tri-proj (b3).
--->
-
-<!-- To determine whether the UMAP representation with the specific (hyper-)parameter choice suggested by @chen2023 preserves the data structure present in $9\text{-}D$ (@fig-model-pbmc-author-proj (b)), we visualize the model constructed with UMAP overlaid on the $9\text{-}D$ data.
--->
-
-<!-- But, when visualizing the *model-in-the-data-space* some unobserved structures can be seen.
-Some clusters have non-linear continuity patterns and high-density patches (@fig-model-pbmc-author-proj).
--->
-
-<!-- As shown in @fig-NLDR-variety (e), there are three well-separated clusters, although the separation between the clusters are small.
-Additionally, non-linear structures can also be observed within the clusters (@fig-NLDR-variety (e)).
-This demonstrates that tSNE accurately captures the data structure of the PBMC3k dataset, which UMAP did not achieve.
--->
-
-<!-- In order to find a reasonable NLDR representation for the PBMC3k dataset, the MSE for different $b_1$ using UMAP with different (hyper-)parameter settings and tSNE with default (hyper-)parameter setting (@fig-pbmc-mse) were calculated.
-After analyzing the results, it was found that tSNE with default (hyper-)parameter setting (perplexity: $30$) achieved the lowest error.
-Therefore, tSNE with a perplexity value set to $30$, the default parameter setting, is considered as a reasonable representation for the PBMC3k dataset.
--->
 
@@
 
-<!-- ::: {#fig-pbmc1-sc layout-ncol="4" fig-pos="H"} -->
-<!-- ![](figures/pbmc3k/umap_trimesh_plot.png) -->
-
-<!-- ![](figures/pbmc3k/sc_1.png) -->
-
-<!-- ![](figures/pbmc3k/sc_2.png) -->
-
-<!-- ![](figures/pbmc3k/sc_3.png) -->
-
-<!-- Model in \gD{}, on the UMAP layout, and three views of the fit in projections from $9\text{-}D$, for the PBMC3k data ($(s_1, \ s_2) = (-0.050, \ -0.041)$, $b = 870 \  (30, \ 29)$, $m = 135$, benchmark value to remove large edges is $0.099$) (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/VqqWuE0Jj6A>).
--->
-<!-- ::: -->
-<!--compute absolute error for different parameter choices-->
-
 ```{r}
@@
 
-<!-- The MSE for different binwidths ($a_1$) using tSNE, UMAP, PHATE, PaCMAP, and TriMAP with different (hyper-)parameter settings (@fig-pbmc-mse) were calculated.  -->
-
-<!-- It is found that tSNE (@fig-mnist-mse (a)) provide the most reasonable representation for the digit 1 dataset, showing universally best.
-However, \gD{} representation shows a big non-linear cluster and a small cluster with a small gap (@fig-tsne-best).
--->
-
-<!-- In order to find a reasonable NLDR representation for the PBMC3k dataset, the MSE for different $a_1$ using UMAP, tSNE, PHATE, PaCMAP, and TriMAP with different (hyper-)parameter settings (@fig-pbmc-mse) were calculated.
-tSNE with a perplexity value set to $30$, the default parameter setting, is considered as a reasonable representation for the PBMC3k dataset.
-As shown in @fig-NLDR-variety (e), there are three well-separated clusters, although the separation between the clusters are small.
-Additionally, non-linear structures can also be observed within the clusters (@fig-NLDR-variety (e)).
--->
-<!-- The visualization of the *model-in-the-data-space* provides evidence for this (@fig-pbmc2-sc).
-Therefore, tSNE with perplexity $30$ accurately captures the data structure of the PBMC3k dataset, which is better than UMAP.
--->
-
 <!--best choice-->
@@
 ```
-
-<!-- We then fit the model for tSNE, and visualize the resultant model in the $9\text{-}D$ data space.
-The model shows a quirk, as shown in @fig-pbmc2-sc.
-All three clusters are connected by an edge except the small and large clusters.
-Because the clusters are so close in \gD{}, they attempt to maintain the structure in \pD{} as well.
-This is evident that tSNE with perplexity $30$ provides a reasonable representation of PBMC3k data.
--->
 
@@
 
-<!-- <!--bin1 = 15, bin2 = 20, b = 300, non_empty = 136--> 
-<!-- ```{r} -->
-<!-- #| echo: false -->
-<!-- #| fig-cap: "(a) Model generated in $2\\text{-}D$ with tSNE, and (b) $p\\text{-}D$ model error in $2\\text{-}D$.
-The $2\\text{-}D$ model shows three well-separated distant clusters.
-The $p\\text{-}D$ model errors are distributed along clusters, but most low $p\\text{-}D$ model errors present in the large cluster." -->
-<!-- #| label: fig-model-pbmc -->
-<!-- #| fig-pos: H -->
-<!-- #| out-height: 30% -->
-
-<!-- trimesh_removed_pbmc + error_plot_pbmc +  -->
-<!--   plot_layout(guides='collect', ncol=2) & -->
-<!--   theme(legend.position='bottom') -->
-<!-- ``` -->
-
-<!-- XXX Number of bins for b is inadequate. Caption needs to start with what this figure is for XXX -->
-
 ```{r}
@@
 
-<!-- ::: {#fig-pbmc2-sc layout-ncol="2" fig-pos="H"} -->
-<!-- ![](figures/pbmc3k/tsne_trimesh_plot.png) -->
-
-<!-- ![](figures/pbmc3k/sc_4.png) -->
-
-<!-- ![](figures/pbmc3k/sc_5.png) -->
-
-<!-- ![](figures/pbmc3k/sc_6.png) -->
-
-<!-- Model in \gD{}, on the tSNE layout, and three views of the fit in projections from $9\text{-}D$, for the PBMC3k data ($(s_1, \ s_2) = (-0.050, \ -0.058)$, $b = 300 \  (15, \ 20)$, $m = 136$, benchmark value to remove large edges is $0.133$).
-(The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/5Y1hE4i7N2k>).
--->
-<!-- ::: -->
-
 ### Hand-written digits {#sec-mnist}
-<!--
-- NLDR is used to illustrate different ways 1's are drawn
-- Use our method to assess is it a reasonable representation
-- Demonstrate that it is, except for the anomalies 
--->
-
-<!--add different NLDR layouts-->
 
@@
 The objective is to select a reasonable \gD{} layout, representing the non-linear structure of the digit 1 dataset in $10\text{-}D$ (@fig-clust-mnist (a)).
-
-<!-- @yingfan2021 used this dataset to evaluate how PaCMAP preserves local structures in \gD{}.
-Because of the large number of images, the focus was specifically on the handwritten digit 1, which exhibits a non-linear structure in \gD{} (@fig-pacmap-author).
-There are $7877$ images of the digit 1 in the dataset.
-Additionally, PCA was applied as a preprocessing step and the first $10$ principal components were selected.
-The objective is to assess whether PaCMAP effectively preserves the non-linear structure in \gD{}.
--->
 
@@
 ```
-
-<!-- PaCMAP preserves non-linear structure in $784\text{-}D$.
-To evaluate whether PaCMAP provides a reasonable representation of the data, the \gD{} embedding of the handwritten digit 1 was selected.
-As shown in @fig-pacmap-author, the angle of the digit 1 images varies along the \gD{} structure.
--->
 
@@
 
-<!-- As a result of visualizing the *model-in-the-data-space*, PaCMAP provides a reasonable \gD{} representation of MNIST digit 1 data, as it preserves the non-linear structure present in the $10\text{-}D$ data (@fig-mnist-tri-proj).
-It also shows some quirks like twisted patterns and long edges (@fig-mnist-tri-proj).
-Furthermore, the results help to identify anomalies (with large model errors) within and the outside the non-linear structure, displaying different patterns of the digit 1 (@fig-model-error-mnist).
--->
-
-<!-- According to @fig-mnist-tri-proj (b), the non-linear structure observed in the \gD{} layout of PaCMAP (@fig-pacmap-author) is also visible when visualizing the model overlaid on the data space.
-This indicates that PaCMAP accurately captures the non-linear structure of the $10\text{-}D$ data.
-Additionally, the model shows a twisted pattern within the non-linear structure in $10\text{-}D$ space (@fig-mnist-tri-proj (c)), which is an additional pattern not visible in the \gD{} representation (@fig-pacmap-author).
-Furthermore, as shown in @fig-mnist-tri-proj (d), some long edges exist in the $10\text{-}D$ space that are not recognized as long edges in the \gD{} representation.
-However, PaCMAP provides a reasonable \gD{} representation of MNIST digit 1 data, as it preserves the non-linear structure present in the $10\text{-}D$ data.
--->
-
 <!--bin1 = 22, bin2 = 17, b = 374, non_empty = 140-->
@@
 
-<!-- ::: {#fig-mnist1-sc layout-ncol="4" fig-pos="H"} -->
-<!-- ![](figures/mnist/mnist_model_2d.png){#fig-mnist1-model} -->
-
-<!-- ![](figures/mnist/sc_1.png){#fig-mnist1-sc1} -->
-
-<!-- ![](figures/mnist/sc_2.png){#fig-mnist1-sc2} -->
-
-<!-- ![](figures/mnist/sc_3.png){#fig-mnist1-sc3} -->
-
-<!-- Model in \gD{}, on the PaCMAP layout, and three views of the fit in projections from $10\text{-}D$, for the digit 1 of MNIST data ($(s_1, \ s_2) = (-0.100, \ -0.059)$, $b = 374 \  (22, \ 17)$, $m = 140$, benchmark value to remove large edges is $0.094$).
-(The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/zcg_GXBmqjA>).
--->
-
-<!-- ::: -->
-<!-- need to update the youtube recording-->
-
-<!--images that occur large error-->
-<!-- within the nonlinear structure-->
-<!--outside the nonlinear structure-->
-
 ```{r}
@@
 
-<!-- These anomalies can be classified into two types: those that are anomalies within the non-linear cluster and those that lie in the small cluster.
-The images associated with large model error points within the non-linear cluster display different patterns of the digit 1.
--->
-
-<!-- There are certain data points that exhibit high error rates due to their deviation from the usual $10\text{-}D$ data structure, which makes them anomalies (@fig-model-mnist).
-These anomalies can be classified into two types: those that are anomalies within the non-linear structure and those that lie outside of it.
-The images associated with high model error points within the non-linear structure display different patterns of the digit 1, as shown in @fig-mnist1-within.
-However, when comparing these images to the ones found outside of the non-linear structure, it becomes evident that the latter display different patterns of the digit 1 (@fig-mnist1-out).
--->
-
-<!--images that occur large error-->
-<!-- within the nonlinear structure-->
-<!--outside the nonlinear structure-->
-
-<!-- ::: {#fig-mnist-anomalies layout-ncol="2" fig-pos="H"} -->
-<!-- ![](figures/mnist/img_error_sample_within.png){#fig-mnist1-within} -->
-
-<!-- ![](figures/mnist/img_error_sample_out.png){#fig-mnist1-out} -->
-
-<!-- Some images of handwritten digit 1 which occur high model error (a) within the non-linear structure, and (b) outside the non-linear structure. The images shows different patterns of digit 1. -->
-<!-- ::: -->
-
 ## Discussion {#sec-discussion}
-
-<!-- - Summarise contributions -->
-<!-- - Explain where it is expected or not expected to work, eg higher dimensional relationships -->
-<!-- - Human behaviour, the desire to have more certainty, and a tendency to prefer the well-separated views (need to add) -->
-<!-- - Predicting new observations in $k$-D -->
-<!-- - Extending layouts beyond $k$-D, when 2D is clearly inadequate. -->
-<!-- - Diagnostic app to explore differences in distances (need to add) -->
-<!-- - What might be useful enhancements -->
 
@@
 
-<!-- and choose the most accurate representation of the structure(s) present in the \pD{} data.
-Starting from \gD{} layout, the fitted model, as represented by the positions of points in \kD{}, and turn it into a \pD{} wireframe to overlay on the data, viewing it with a tour.
-This approach is defined as *model-in-data-space*.
-Viewing a model in the data space is an ideal way to examine the fit.
-
-With the model fit, several observations can be made.
-The simulated \fD{} data, which contains five Gaussian clusters, was analyzed.
-Both tSNE and UMAP demonstrated the fitted model as a *crumpled \gD{} sheet* that stretches across all four dimensions, highlighting its ability to preserve local structure.
-In contrast, the flat representation generated by PaCMAP indicates its failure to capture the variation within the clusters.
-This highlights how different NLDR methods organize the data points differently.
-
-Another usability of our approach is how NLDR methods behave with different density of the data structure.
-The tSNE layout of simulated \fD{} C-shaped data with different densities are used.
-The high \fD{} model errors are scattered to a corner of the \gD{} layout of dense C-shaped structure.
-In the dense C-shaped structure, the high model errors occur because there are fewer data points positioned within the bins at the sparse end of the structure.
-
-Our algorithm can also be useful in practice to decide on a reasonable NLDR layout that accurately represents the data, as described with a single-cell gene expression data, **PBMC3k**.
-The published results of UMAP with default (hyper-)parameter setting by @chen2024 were evaluated.
-The data showed three close clusters.
-But the published layout showed three well-separated clusters.
-Therefore, by changing the NLDR method to tSNE with default (hyper-)parameters, we found a better layout, which is more representative the data.
-Additionally, the model discovers non-uniform data distribution and non-linear structures within the clusters that are not visible in the \gD{} layout, demonstrating the ability of our model in uncovering hidden data characteristics.
-Also, as explained in Appendix, with recently introduced approach, scDEED [@xia2023] also proved the layout suggested by our layout is better than the published one.
-
-The digit 1 of **Hand-written digits** data shows how the model accurately capture the non-linear structure.
-Not only that, the $10\text{-}D$ model error help to find the anomalies of digit 1 images.
-Also, twisted pattern of the model evident that how tSNE tries to maintain the local structure by positioning the similar patterns of digit 1 close to each other.
-
-Diagnose is important to find where the \gD{} points located in \pD{}.
-One approach is to use interactivity.
-By linking the \gD{} layout with the fitted model in \pD{} and employing brushing techniques, we can identify which points fit well and which do not.
-Additionally, linking the \gD{} error plot with the fitted model in \pD{} can help pinpoint where mismatches occur.
-
-Predicting new observations in \kD{} is particularly valuable due to the limitations of some NLDR methods, like tSNE, which don't provide a straightforward method for prediction.
-As a result, our approach offers a solution that capable of generating predicted \kD{} embedding regardless of the NLDR method employed, effectively addressing this functional gap.
-
-Furthermore, to extend layouts beyond \kD{}, when \gD{} is clearly inadequate, k-means can be used as an alternative to bin centroids.
-
-There are many avenues for future work. First, the development of
-evaluation metrics is important for NLDR.
-We introduced a qualitative approach.
-But having a quantitative measure also will be useful for various application domains.
-Second, one could use our approach to create an approach for tuning (hyper-)parameters.
-Another direction for future work is to introduce more approaches to diagnose the fitted model.
-Finally, one can think of designing a new NLDR method by following the binning in high-dimensions.
-
-Other tour software instead of langevitour. 
--->
-
-<!-- In conclusion, while our method effectively captures and represents high-dimensional data structures, further enhancements could involve introducing approaches to diagnose the fitted model.
-These improvements would help in creating a more accurate representation of the data when \gD{} layout is inadequate.
--->
-
 ## Supplementary Materials

### 2025-06-09 paper.qmd
Checked 2D notation

@@
 - Too much code needs to be streamlined
-- Remove all the extraneous commented text
+- Remove all the extraneous commented text (Done)
 - Check how \kD{} and \gD{} are used, and make sure it's sensible
@@
 
-## Assessing the best fit, and hence most accurate 2D layout
+## Assessing the best fit, and hence most accurate \gD{} layout
 
@@
 A simple example of this can be seen with data simulated to contain five \fD{} Gaussian clusters.
-Each cluster is essentially a ball in \fD{}, so there is no \2D{} representation, rather the model in each cluster should resemble a crumpled sheet of paper that fills out \fD{}.
+Each cluster is essentially a ball in \fD{}, so there is no \gD{} representation, rather the model in each cluster should resemble a crumpled sheet of paper that fills out \fD{}.
 
@@
 
-### Low density results in a squishing of the 2D representation in all methods {#sec-effect-dens}
+### Low density results in a squishing of the \gD{} representation in all methods {#sec-effect-dens}
 
 Differences in density can arise by sampling at different rates in different subspaces of \pD{}.
-For example, the data shown in @fig-one-dens_clust-error all lies on a \2D{} curved sheet in \fD{}, but one end of the sheet is sampled densely and the other very sparsely.
+For example, the data shown in @fig-one-dens_clust-error all lies on a \gD{} curved sheet in \fD{}, but one end of the sheet is sampled densely and the other very sparsely.
 It was simulated to illustrate the effect of the density difference on layout generated by an NLDR, illustrated using the tSNE results.

### 2025-06-09 paper.qmd
Langevitour is introduced in page 5

@@
 
-- Make sure langevitour is explained before using it
+- Make sure langevitour is explained before using it (introduced in page 5)
 - Too much code needs to be streamlined

### 2025-06-10 paper.qmd
finished with PBMC3k example

--- 2025-06-09 paper.qmd
@@
 - Check how \kD{} and \gD{} are used, and make sure it's sensible
+- Check spelling is consistently American
 
@@
 
-To showcase the robustness of our method when applied to data with different characteristics, this section presents two case studies.
+To illustrate the approach we use two examples: PBMC3k data (single cell gene expression) where an NLDR layout is used to represent cluster structure present in the \pD{} data, and MNIST hand-written digits where NLDR is used to represent essentially a low-dimensional nonlinear manifold in \pD{}.
+
+<!-- To showcase the robustness of our method when applied to data with different characteristics, this section presents two case studies.
 The first application focuses on single-cell data, a domain where the high dimensionality, sparsity, and heterogeneity of gene expression measurements pose significant analytical challenges.
 Our approach facilitates the identification of the most accurate NLDR layout, enabling the discovery of cell populations with similar expression profiles.
-The second case study explores the MNIST hand-written digits dataset, which serves as a benchmark for evaluating the preservation of local structures, specifically with digit 1.
-
-### Single-cell gene expression {#sec-pbmc}
-
-Single-cell data refers to the characteristics of individual cells within a population [@haque2017].
-These data also provides insights at the level of individual cells, enabling deeper understanding of gene expression.
-
+The second case study explores the MNIST hand-written digits dataset, which serves as a benchmark for evaluating the preservation of local structures, specifically with digit 1.-->
+
+### PBMC3k {#sec-pbmc}
+
+This is a benchmark single-cell data set collected on Human Peripheral Blood Mononuclear Cells (PBMC3k) as used in @chen2024.
+Single-cell data measures the gene expression of individual cells in a sample of tissue (see  for example, @haque2017).
+This type of data is used to obtain an understanding of cellular level behavior and heterogeneity in their activity.
 Clustering of single-cell data is used to identify groups of cells with similar expression profiles.
-NLDR often used to summarise the discovered clusters, and help to understand the results.
-The purpose of this example is to *illustrate how to use our method to help decide on an appropriate NLDR layout that accurately represents the data*.
-The cluster results of Human Peripheral Blood Mononuclear Cells (PBMC3k) in @chen2024 are examined.
-There are $2622$ single cells, with $1000$ gene expressions.
-Following their pre-processing, UMAP was performed using nine principal components.
-@fig-NLDR-variety (a) is the reproduction of the published plot.
-The question is whether this accurately represents the cluster structure in the data.
-Our method can help here, and also help to provide a better \gD{} layout, as needed.
+NLDR often used to summarise the cluster structure.
+Usually, NLDR does not use the cluster labels to compute the layout, but uses color to represent the cluster labels when it is plotted.
+<!-- The purpose of this example is to *illustrate how to use our method to help decide on an appropriate NLDR layout that accurately represents the data*.
+The cluster results of Human Peripheral Blood Mononuclear Cells (PBMC3k) in @chen2024 are examined.-->
+
+In this data there are $2622$ single cells and $1000$ gene expressions (variables).
+Following their pre-processing, UMAP was performed on the first nine principal components.
+@fig-NLDR-variety shows this data using a variety of methods, and different (hyper-)~parameters.
+You can see that the result is wildly different depending on the choices.
+Layout a is a reproduction of the layout published in @chen2024.
+This layout suggests that the data has three very well separated clusters, each with an odd shape.
+The question is whether this accurately represents the cluster structure in the data, or whether they should have chosen b or c or d or e or f or g.
+This is what our new method can help with -- to decide which is the more accurate \gD{} representation of the cluster structure in the \pD{} data.
 
@@
 
-The @fig-NLDR-variety (a) shows three well-separated clusters with big separations.
+<!-- The @fig-NLDR-variety (a) shows three well-separated clusters with big separations.
 However, as shown in @fig-model-pbmc-author-proj (a2), there is no big separation between three clusters in $9\text{-}D$.
-Therefore, the suggested UMAP representation (@fig-NLDR-variety (a)) does not accurately represent the structure of PBMC3k dataset.
-
-As a result, it is necessary to find an appropriate layout for the dataset.
-MSE for different binwidths ($a_1$) using tSNE, UMAP, PHATE, PaCMAP, and TriMAP with various (hyper-)~parameter settings were computed (@fig-pbmc-mse).
-Layouts c, d, and e, which show small separations between clusters, are universally optimal.
+Therefore, the suggested UMAP representation (@fig-NLDR-variety (a)) does not accurately represent the structure of PBMC3k dataset.-->
+
+@fig-pbmc-mse shows RMSE across a range of binwidths ($a_1$) for each of the layouts in @fig-NLDR-variety.
+The layouts were made using tSNE, UMAP, PHATE, PaCMAP, and TriMAP with various (hyper-)~parameter settings.
+Lines are color coded to match the color of the layouts shown on the right.
+Lower RMSE indicates the better fit.
+Using a range of binwidths shows how the model changes, with possibly the best model being one that is universally low RMSE across all binwidths.
+It can be seen that layout f is sub-optimal with universally higher RMSE.
+Layout a, the published one, is better but it is not as good as layouts b, d, or e.
+With some imagination layout d perhaps shows three barely distinguishable clusters.
+Layout e shows three, possibly four, clusters that are more separated.
+The choice reduces from eight to these two.
+Layout d has slightly better RMSE when the $a_1$ is small, but layout e beats it at larger values.
+Thus we could argue that layout e is the most accurate representation of the cluster structure, of these eight.
+
+<!-- Layouts c, d, and e, which show small separations between clusters, are universally optimal.
 However, layout d performs well with smaller binwidths and poorly with larger binwidths.
 On the other hand, layout e performs well with larger binwidths.
-Layout c was selected for further analysis due to its better (hyper-)parameter selection for the same method of the published plot.
-
-The visualization of the selected layout in the $9\text{-}D$ shows edges between the clusters (@fig-model-pbmc-author-proj (b2)).
+Layout c was selected for further analysis due to its better (hyper-)parameter selection for the same method of the published plot.-->
+
+XXX can we use a and e in these figures instead of a and b, so it is easier to describe relative to the first figure, and previous figure?
+
+To further assess the choices, we need to look at the model in the data space, by using a tour to show the wireframe model overlaid on the data in the $9\text{-}D$ space (@fig-model-pbmc-author-proj).
+Here we compare the published layout (a) versus what we argue is the best layout (e).
+The top row (a1, a2, a3) correspond to the published layout and the bottom row (e1, e2, e3) correspond to the optimal choice according to our procedure.
+The middle and right plots show two projections.
+The primary difference between the two models is that, the model of layout e does not fill out to the extent of the data but concentrates in the center of each point cloud.
+Both suggest that three clusters is a reasonable interpretation of the structure, but layout e more accurately reflects the separation between them, which is small.
+
+<!-- The visualization of the selected layout in the $9\text{-}D$ shows edges between the clusters (@fig-model-pbmc-author-proj (b2)).
 This supports the presence of small separations between clusters.
 Additionally, the data points are not uniformly distributed across the clusters, resulting in dense areas (@fig-model-pbmc-author-proj (b2)).
-Furthermore, the clusters shows non-linear shapes (@fig-model-pbmc-author-proj (b3)).
+Furthermore, the clusters shows non-linear shapes (@fig-model-pbmc-author-proj (b3)).-->
 
@@
 
-### Hand-written digits {#sec-mnist}
-
-The digit 1 of the MNIST dataset consists of $7877$ grayscale images of handwritten digits [@lecun2010].
+### MNIST hand-written digits {#sec-mnist}
+
+The digit "1" of the MNIST dataset [@lecun2010] consists of $7877$ grayscale images of handwritten 1s.
 Before further analysis, PCA was used to preprocess the data, where the first $10$ principal components, explaining 83% of the total variation, were selected.
@@
 
-The GitHub repository for the paper can be found at https://github.com/JayaniLakshika/paper-nldr-vis-algorithm.
+All the materials to reproduce the paper can be found at [https://github.com/JayaniLakshika/paper-nldr-vis-algorithm](https://github.com/JayaniLakshika/paper-nldr-vis-algorithm).
 The Appendix includes more details about the hexagonal binning algorithm and a comparison to the results of the newly reported scDEED [@xia2023] statistic.
 
-The R package `quollr`, available on CRAN and at https://jayanilakshika.github.io/quollr/, provides software accompaying this paper to fit the wireframe model representation, compute diagnostics, visualize the model in the data with langevitour and link multiple plots interactively.
+The R package `quollr`, available on CRAN and at [https://jayanilakshika.github.io/quollr/](https://jayanilakshika.github.io/quollr/), provides software accompaying this paper to fit the wireframe model representation, compute diagnostics, visualize the model in the data with langevitour and link multiple plots interactively.
 

### 2025-06-10 paper.qmd
Updated Figure 12

@@
 
-XXX can we use a and e in these figures instead of a and b, so it is easier to describe relative to the first figure, and previous figure?
+<!-- XXX can we use a and e in these figures instead of a and b, so it is easier to describe relative to the first figure, and previous figure? -->
 
@@
   scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada")) +
-  interior_annotation("b1", cex = 2, c(0.08, 0.9)) +
+  interior_annotation("e1", cex = 2, c(0.08, 0.9)) +
   theme(
@@
   is_category = TRUE) +
-  interior_annotation(label = "b2", cex = 2) +
+  interior_annotation(label = "e2", cex = 2) +
   scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))
@@
   is_category = TRUE) +
-  interior_annotation(label = "b3", cex = 2) + 
+  interior_annotation(label = "e3", cex = 2) + 
   scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))

### 2025-06-10 paper.qmd
Updated PBMC3k reference

@@
 
-This is a benchmark single-cell data set collected on Human Peripheral Blood Mononuclear Cells (PBMC3k) as used in @chen2024.
+This is a benchmark single-cell data set collected on Human Peripheral Blood Mononuclear Cells (PBMC3k) as used in @pbmc2019.
 Single-cell data measures the gene expression of individual cells in a sample of tissue (see  for example, @haque2017).
@@
 
-Direct links to videos for viewing online are available in @tbl-links.
-
-```{r}
-#| label: tbl-links
-#| tbl-cap: "Example videos."
-#| tbl-pos: H
-
-links_df <- tibble(
-  data = c("PBMC3k"),
-  URL = c("link1")
-)
-
-links_df |>
-  kableExtra::kable(format = "latex", 
-                    booktabs = TRUE, 
-                    label = "tbllinks") |>
-  kableExtra::kable_styling(latex_options = "scale_down")
-
-```
-
 ## Acknowledgments

### 2025-06-10 paper.qmd
Added Langevitour links for Gaussian data

@@
                                         c("data", "model")),
-                         levelColors = c(clr_choice, "#000000"))
+                         levelColors = c(clr_choice, "#000000"),
+                         pointSize = 0.7)
 ```
@@
                                         c("data", "model")),
-                         levelColors = c(clr_choice, "#000000"))
+                         levelColors = c(clr_choice, "#000000"), pointSize = 0.7)
 ```
@@
 With (c2), the curiousity is that the model is a \\gD{} pancake shape in $4\\text{-}D$, indicating that there is some ordering of points done by PaCMAP, posisbly along some principal component axes.
-Videos of the langevitour animations are available at <https://youtu.be/oQxEb4wRdHI>, <https://youtu.be/JW49csPpDx4>, and XXX respectively."
+Videos of the langevitour animations are available at <https://youtu.be/I-kxCwVfqiQ>, <https://youtu.be/gD1P01FUPyU>, and <https://youtu.be/MxJ_srOFQNk> respectively."
 

### 2025-06-10 paper.qmd
Updated layout a link in PBMC3k

@@
 - Check spelling is consistently American
+- Check references
+- Add videos
 
@@
 error_pbmc_umap2 <- read_rds("data/pbmc3k/error_pbmc_umap_5_min_dist_0.01.rds")
-# error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_12_min_dist_0.99.rds")
-# error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_3_min_dist_0.6.rds")
-# error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_4_min_dist_0.6.rds")
-# error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_5_min_dist_0.8.rds")
 error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_5_min_dist_0.8.rds")
@@
 Also, the fitted models in $9\\text{-}D$ help to see some unobserved patterns of the data: (i) dense points, and (ii) non-linear clusters.
-Videos of the langevitour animations are available at <https://youtu.be/0cKX_HG_n0k> and <https://youtu.be/KhJvsRtaX04> respectively."
+Videos of the langevitour animations are available at <https://youtu.be/3VfK3M2gnZM> and <> respectively."
 #| label: fig-model-pbmc-author-proj

### 2025-06-10 paper.qmd
done with MNIST example

@@
 
-The digit "1" of the MNIST dataset [@lecun2010] consists of $7877$ grayscale images of handwritten 1s.
-Before further analysis, PCA was used to preprocess the data, where the first $10$ principal components, explaining 83% of the total variation, were selected.
-The objective is to select a reasonable \gD{} layout, representing the non-linear structure of the digit 1 dataset in $10\text{-}D$ (@fig-clust-mnist (a)).
+The digit "1" of the MNIST dataset [@lecun2010] consists of $7877$ grayscale images of handwritten "1"s.
+Each image is $28 \times 28$ pixels which corresponds to 784 variables.
+The first 10 principal components, explaining 83% of the total variation, are used.
+This data is essentially lies on a nonlinear manifold in the high dimensions, defined by the shapes that "1"s make in when sketched, menaing that some pixels are almost always dark and some almost always light.
+We expect that the best layout captures this type of structure and does not exhibit distinct clusters.
+<!-- The objective is to select a reasonable \gD{} layout, representing the non-linear structure of the digit 1 dataset in $10\text{-}D$ (@fig-clust-mnist (a)).
+-->
 
@@
 
-The MSE for different binwidths ($a_1$) using UMAP, PHATE, PaCMAP, TriMAP with default (hyper-)parameter setting and tSNE with default (hyper-)parameter setting and perplexity $89$ (@fig-mnist-mse) were calculated.
+@fig-mnist-mse compares the fit of six layouts computed using UMAP (a), PHATE (b), PaCMAP (c), TriMAP (d) with default (hyper-)parameter setting and two tSNE runs, one with default (hyper-)parameter setting (e) and the other changing perplexity to $89$ (f).
+The layouts are reasonably similar in that they all have the observations in a single blob.
+Some (b, c) have a more curved shape than others.
+Layout e is the most different having a linear shape, and a single very large outlier.
+Both a and f have a small clump of points perhaps slightly disconnected from the other points, in the lower to middle right.
+XXX are the labels I put for each method above correct?
+please fix if not.
+
+The layout plot are colored to match the lines in the RMSE vs binwidth ($a_1$) plot.
+Layouts a, b and f fit the data better than c, d, e, and layout f appears to be the bets fit.
+@fig-clust-mnist shows this model in the data space in two projections from a tour.
+The data is curved in the $10\text{-}D$ space, and the fitted model captures this curve.
+The small clump of points in the \gD{} layout is highlighted in both displays.
+These are almost all inside the curve of the bulk of points and are sparsely located.
+The fact that they are packed together in the \gD{} layout is likely due to the handling of density differences by the NLDR.
+
+The next step is to investigate the \gD{} to understand what information is learned from this representation.
+@fig-model-error-mnist summarizes this investigation.
+Plot a shows the layout with points colored by their residual value - darker color indicates larger residual and poor fit.
+The plots b, c, d, e show samples of hand-written digits taken from inside the colored boxes.
+Going from top to bottom around the curve shape we can see that the "1"s are drawn with from right slant to a left slant.
+The "1"s in d (black box) tend to have the extra up stroke but are quite varied in appearance.
+The "1"s shown in the plots labelled e correspond to points with big residuals.
+They can be seen to be more strangely drawn than the others.
+Overall, this \gD{} layout shows a useful way to summarise the variation in way "1"s are drawn.
+
+<!-- The MSE for different binwidths ($a_1$) using UMAP, PHATE, PaCMAP, TriMAP with default (hyper-)parameter setting and tSNE with default (hyper-)parameter setting and perplexity $89$ (@fig-mnist-mse) were calculated.
 It is found that tSNE with perplexity $89$ (@fig-mnist-mse (f)) provide the most reasonable representation for the digit 1 dataset, showing universally best.
-However, \gD{} representation shows a big non-linear cluster and a small cluster with a small gap.
-
-In the case of digit 1 data, there should not be any clusters unless anomalies exist, indicating different digit 1 patterns.
+However, \gD{} representation shows a big non-linear cluster and a small cluster with a small gap.-->
+
+<!-- In the case of digit 1 data, there should not be any clusters unless anomalies exist, indicating different digit 1 patterns.
 The angle of the digit 1 images varies along this non-linear clustering structure (@fig-model-error-mnist), while the small cluster contains the digit 1 images with different patterns of the digit 1, unlike the usual (@fig-model-error-mnist (c)).
@@
 The model's twisted pattern provides evidence for the $10\text{-}D$ data structure, which is not observed by \gD{} layout (@fig-clust-mnist (a2)).
+-->
 
@@
   geom_rect(aes(xmin = 0.35, xmax = 0.45, ymin = 0.55, ymax = 0.7),
-               fill = "transparent", color = "#004616", linewidth = 1)  +
+               fill = "transparent", color = "#008000", linewidth = 1)  +
   geom_rect(aes(xmin = 0.7, xmax = 0.8, ymin = 0.25, ymax = 0.4),
                fill = "transparent", color = "#4c389a", linewidth = 1) +
+  interior_annotation(label = "a", c(0.1, 1), cex = 2.5) +
   theme(legend.position = "none",
@@
 right_top_img <- plot_digit_img(
-  pixels_gathered_within, palette = "Blues 2", title_text = "a")
+  pixels_gathered_within, palette = "Blues 2", title_text = "b")
 
@@
 middle_img <- plot_digit_img(
-  pixels_gathered_within, palette = "Greens 3", title_text = "b")
+  pixels_gathered_within, palette = "Greens 3", title_text = "c")
 
@@
 right_bottom_img <- plot_digit_img(
-  pixels_gathered_within, palette = "Purples 2", title_text = "c")
+  pixels_gathered_within, palette = "Purples 2", title_text = "d")
 
@@
 imge_error_sample_outside <- plot_digit_img(
-  pixels_gathered_outside, palette = "Grays", title_text = "d")
+  pixels_gathered_outside, palette = "Grays", title_text = "e")
 
@@
 imge_error_sample_inside <- plot_digit_img(
-  pixels_gathered_inside, palette = "Oranges", title_text = "e")
+  pixels_gathered_inside, palette = "Oranges", title_text = "f")
 

### 2025-06-10 paper.qmd
TO DO list modified)

@@
 
-- Make sure langevitour is explained before using it (introduced in page 5)
-- Too much code needs to be streamlined
-- Remove all the extraneous commented text (Done)
+- Remove all the extraneous commented text (NOT Done)
 - Check how \kD{} and \gD{} are used, and make sure it's sensible
@@
 - Check references
-- Add videos
+- Add videos - make a list of the videos in the supplementary section, with the number of the figure that it corresponds too.
+This avoids having captions that are too long and awkward with raw URLs.
+And it keeps all of this supplementary material in one place.
+- Move the notation table to appendix
+- Code clean up: recommend that all of the calculations done with quollr are done with a separate script(s) and put in a scripts directory.
+Save the resulting data need to make plots and tables.
+Only code for the plots and tables is directly in the qmd in the final version.
 

### 2025-06-11 paper.qmd
Initial work on the table

@@
 
+Direct links to videos for viewing online are available in @tbl-links.
+
+```{r}
+#| label: tbl-links
+#| tbl-cap: "Example videos."
+#| tbl-pos: H
+
+links_df <- tibble(
+  data = c("PBMC3k"),
+  URL = c("link1")
+)
+
+links_df |>
+  kableExtra::kable(format = "latex", 
+                    booktabs = TRUE, 
+                    label = "tbllinks") |>
+  kableExtra::kable_styling(latex_options = "scale_down")
+
+```
+
 ## Acknowledgments

### 2025-06-11 paper.qmd
Added the notation table to appendix

@@
 
-- Remove all the extraneous commented text (NOT Done)
+- Remove all the extraneous commented text (Done)
 - Check how \kD{} and \gD{} are used, and make sure it's sensible
@@
 
-```{r}
-#| label: tbl-notation
-#| tbl-cap: "Summary of notation for describing new methodology."
-# Notation used in the paper
-
-notation_df <- read_csv("misc/notation.csv")
-
-# Create the table
-kable(notation_df, 
-      format = "latex", 
-      booktabs = TRUE, escape = FALSE) |>
-  kable_styling(position = "center", 
-                full_width = FALSE, 
-                font_size = 12) |>
-  row_spec(0, bold = TRUE) |>
-  column_spec(1:2, width = c("3cm", "12cm"))
-```
-
 We define the NLDR as a function $g\text{:}~ \mathbb{R}^{n\times p} \rightarrow \mathbb{R}^{n\times k}$, with (hyper-)parameters $\mathbfit{\theta}$.
@@
 Common choices for $g$ include functions used in tSNE, UMAP, PHATE, TriMAP, PaCMAP, or MDS, although in theory any function that does this mapping is suitable.
-<!--Any input requirements for the data (such as normalization, or preprocessing through the use of PCA or similar) is considered part of the function $g$.-->
 
@@
 
-# error_plot_two_curvy_hist <- 
-#   ggplot(error_df_two_curvy_abs) +
-#   geom_histogram(aes(x=sqrt_row_wise_total_error, y=..density..)) +
-#   geom_density(aes(x=sqrt_row_wise_total_error, y=..density..), colour="red") +
-#   #xlab(expression(group("|", e[hj], "|"))) +
-#   xlab(expression(e[hj])) +
-#   ylab("") +
-#   theme_bw()
-
 error_plot_two_curvy_quasi <- ggplot(
@@
   interior_annotation(label = "c", position = sc_ltr_pos, cex = 1.2) +
-  theme(aspect.ratio = 1) #+
-  #geom_segment(
-  #  data = model_df, 
-  #  aes(
-  #    x = proj1_from, 
-  #    y = proj2_from, 
-  #    xend = proj1_to, 
-  #    yend = proj2_to), 
-  #  color = "#000000")
+  theme(aspect.ratio = 1)
 
@@
 
-<!-- XXX How are you doing this? Do you check the bounds of the hexagon? Or do you use distance to centroid? -->
-
 #### Indicating neighborhood
@@
 
-Observations are associated with their bin center, $C_{h}^{(p)}$, which are also considered to be the *fitted values*.
-These can also be denoted as $\widehat{X}$.
-<!--The fitted values of the points in $H_h$ refers to the \pD{} mapping $C_{h}^{(p)}$ of the corresponding \kD{} model point $C_{h}^{(2)}$.-->
-
+Observations are associated with their bin center, $C_{h}^{(p)}$, which are also considered to be the *fitted values*. These can also be denoted as $\widehat{X}$. 
 <!--#### Error-->
@@
   )
-
-# df_bin_centroids_gau1 <- df_bin_centroids_gau1 |>
-#   dplyr::filter(std_counts > 0)
 
@@
 
-# df_bin_centroids_gau1 <- df_bin_centroids_gau1 |>
-#   dplyr::filter(std_counts > 0)
-
 df_b <- df_bin_gau1 |>
@@
 five_gau_proj_tsne_model1 + five_gau_proj_umap_model1 + five_gau_proj_pacmap_model1 + 
-  #five_gau_proj_tsne_model2 +
-  #five_gau_proj_umap_model2 + five_gau_proj_pacmap_model2 +
   plot_layout(guides = "collect", nrow = 2) &
@@
 
-<!-- To showcase the robustness of our method when applied to data with different characteristics, this section presents two case studies.
-The first application focuses on single-cell data, a domain where the high dimensionality, sparsity, and heterogeneity of gene expression measurements pose significant analytical challenges.
-Our approach facilitates the identification of the most accurate NLDR layout, enabling the discovery of cell populations with similar expression profiles.
-The second case study explores the MNIST hand-written digits dataset, which serves as a benchmark for evaluating the preservation of local structures, specifically with digit 1.-->
-
 ### PBMC3k {#sec-pbmc}
@@
 Usually, NLDR does not use the cluster labels to compute the layout, but uses color to represent the cluster labels when it is plotted.
-<!-- The purpose of this example is to *illustrate how to use our method to help decide on an appropriate NLDR layout that accurately represents the data*.
-The cluster results of Human Peripheral Blood Mononuclear Cells (PBMC3k) in @chen2024 are examined.-->
 
@@
 ```
-
-<!-- The @fig-NLDR-variety (a) shows three well-separated clusters with big separations.
-However, as shown in @fig-model-pbmc-author-proj (a2), there is no big separation between three clusters in $9\text{-}D$.
-Therefore, the suggested UMAP representation (@fig-NLDR-variety (a)) does not accurately represent the structure of PBMC3k dataset.-->
 
@@
 
-<!-- Layouts c, d, and e, which show small separations between clusters, are universally optimal.
-However, layout d performs well with smaller binwidths and poorly with larger binwidths.
-On the other hand, layout e performs well with larger binwidths.
-Layout c was selected for further analysis due to its better (hyper-)parameter selection for the same method of the published plot.-->
-
-<!-- XXX can we use a and e in these figures instead of a and b, so it is easier to describe relative to the first figure, and previous figure? -->
-
 To further assess the choices, we need to look at the model in the data space, by using a tour to show the wireframe model overlaid on the data in the $9\text{-}D$ space (@fig-model-pbmc-author-proj).
@@
 Both suggest that three clusters is a reasonable interpretation of the structure, but layout e more accurately reflects the separation between them, which is small.
-
-<!-- The visualization of the selected layout in the $9\text{-}D$ shows edges between the clusters (@fig-model-pbmc-author-proj (b2)).
-This supports the presence of small separations between clusters.
-Additionally, the data points are not uniformly distributed across the clusters, resulting in dense areas (@fig-model-pbmc-author-proj (b2)).
-Furthermore, the clusters shows non-linear shapes (@fig-model-pbmc-author-proj (b3)).-->
 
@@
 We expect that the best layout captures this type of structure and does not exhibit distinct clusters.
-<!-- The objective is to select a reasonable \gD{} layout, representing the non-linear structure of the digit 1 dataset in $10\text{-}D$ (@fig-clust-mnist (a)).
--->
 
@@
 Overall, this \gD{} layout shows a useful way to summarise the variation in way "1"s are drawn.
-
-<!-- The MSE for different binwidths ($a_1$) using UMAP, PHATE, PaCMAP, TriMAP with default (hyper-)parameter setting and tSNE with default (hyper-)parameter setting and perplexity $89$ (@fig-mnist-mse) were calculated.
-It is found that tSNE with perplexity $89$ (@fig-mnist-mse (f)) provide the most reasonable representation for the digit 1 dataset, showing universally best.
-However, \gD{} representation shows a big non-linear cluster and a small cluster with a small gap.-->
-
-<!-- In the case of digit 1 data, there should not be any clusters unless anomalies exist, indicating different digit 1 patterns.
-The angle of the digit 1 images varies along this non-linear clustering structure (@fig-model-error-mnist), while the small cluster contains the digit 1 images with different patterns of the digit 1, unlike the usual (@fig-model-error-mnist (c)).
-This provides the evidence for two close clusters.
-
-By visualizing the model generated for tSNE in $10\text{-}D$ helps to assess the \gD{} layout.
-As shown in @fig-clust-mnist (a1), the model provides the evidence for the non-linear structure of the digit 1 data in $10\text{-}D$.
-The model shows some quirks.
-The model's twisted pattern provides evidence for the $10\text{-}D$ data structure, which is not observed by \gD{} layout (@fig-clust-mnist (a2)).
--->
 

### 2025-06-11 paper.qmd
Match the layouts in MNIST

@@
 And it keeps all of this supplementary material in one place.
-- Move the notation table to appendix
+- Move the notation table to appendix (Done)
 - Code clean up: recommend that all of the calculations done with quollr are done with a separate script(s) and put in a scripts directory.
@@
 
-@fig-mnist-mse compares the fit of six layouts computed using UMAP (a), PHATE (b), PaCMAP (c), TriMAP (d) with default (hyper-)parameter setting and two tSNE runs, one with default (hyper-)parameter setting (e) and the other changing perplexity to $89$ (f).
+@fig-mnist-mse compares the fit of six layouts computed using UMAP (b), PHATE (c), TriMAP (d), PaCMAP (e) with default (hyper-)parameter setting and two tSNE runs, one with default (hyper-)parameter setting (a) and the other changing perplexity to $89$ (f).
 The layouts are reasonably similar in that they all have the observations in a single blob.
@@
 Both a and f have a small clump of points perhaps slightly disconnected from the other points, in the lower to middle right.
-XXX are the labels I put for each method above correct?
-please fix if not.
 

### 2025-06-11 paper.qmd
Updated Figure 15 labels and figure title

@@
                fill = "transparent", color = "#4c389a", linewidth = 1) +
-  interior_annotation(label = "a", c(0.1, 1), cex = 2.5) +
+  interior_annotation(label = "a", c(0.08, 0.93), cex = 2) +
   theme(legend.position = "none",
@@
 Most low model errors are distributed along the big non-linear cluster, while most large model errors are distributed along the small cluster.
-The images associated with large model errors shows different patterns of digit 1, some inside (e) the non-linear structure and others outside (d).
-Along the non-linear cluster, the angle of digit 1 changes (a-c)."
+The images associated with large model errors shows different patterns of digit 1, some inside (f) the non-linear structure and others outside (e).
+Along the non-linear cluster, the angle of digit 1 changes (b-d)."
 #| fig-pos: H

### 2025-06-11 paper.qmd
Updated .bib
Removed unnecessary references from .bib

@@
 We can also imagine that the positioning of points in \gD{}    represent the fitted values, that will have some prescribed position in \pD{}   that can be compared with their observed values.
-This is the conceptual framework underlying the more formal versions of factor analysis [@cfa69] and MDS.
+This is the conceptual framework underlying the more formal versions of factor analysis [@joreskog1969] and MDS.
 (Note that, for this thinking the full \pD{}   data needs to be available, not just the interpoint distances.)
@@
 Binning into cubes could be done in $3\text{-}D$ or higher, relatively easily, and used asa the basis for a wireframe of the fitted model.
-@barber1996 and the software @cxhull have algorithms for convex hulls, which \pD{} which serve as an inspiration.
+@barber1996 and the software @stephane2023 have algorithms for convex hulls, which \pD{} which serve as an inspiration.
 A simpler approach using $k$-means clustering to provide centroids could also be possible, but the complication would be to determine how to connect the centroids into an appropriate wireframe.

### 2025-06-11 paper.qmd
Finalized the references
Initial work on video link table

@@
 - Check spelling is consistently American
-- Check references
+- Check references (Done)
 - Add videos - make a list of the videos in the supplementary section, with the number of the figure that it corresponds too.
@@
 Also, the fitted models in $9\\text{-}D$ help to see some unobserved patterns of the data: (i) dense points, and (ii) non-linear clusters.
-Videos of the langevitour animations are available at <https://youtu.be/3VfK3M2gnZM> and <> respectively."
+Videos of the langevitour animations are available at <https://youtu.be/3VfK3M2gnZM> and <https://youtu.be/Es84bwQcndU> respectively."
 #| label: fig-model-pbmc-author-proj
@@
 
-The digit "1" of the MNIST dataset [@lecun2010] consists of $7877$ grayscale images of handwritten "1"s.
+The digit "1" of the MNIST dataset [@lecun1998] consists of $7877$ grayscale images of handwritten "1"s.
 Each image is $28 \times 28$ pixels which corresponds to 784 variables.
@@
 Two \\gD{} projections from a tour on $10\\text{-}D$ reveal that the closeness of the clusters in $10\\text{-}D$ and the twisted pattern of the model fit with tSNE ($a_1 = 0.041$, $b = 659/585 \\text{ } (30, \\text{ }48)$).
-Video of the langevitour animation is available at <>.
+Video of the langevitour animation is available at <https://youtu.be/sUcGd57Swdg>.
 The brushing feature in the linked plots helps in visualizing the closeness of the small cluster to the big cluster.
@@
 #| label: tbl-links
-#| tbl-cap: "Example videos."
 #| tbl-pos: H
-
-links_df <- tibble(
-  data = c("PBMC3k"),
-  URL = c("link1")
-)
-
-links_df |>
-  kableExtra::kable(format = "latex", 
-                    booktabs = TRUE, 
-                    label = "tbllinks") |>
-  kableExtra::kable_styling(latex_options = "scale_down")
-
+#| tbl-cap: "Videos of the langevitour animations and linked plots."
+# Video links in the paper
+
+links_df <- read_csv("misc/video_links.csv")
+
+# Create the table
+kable(links_df, 
+      format = "latex", 
+      booktabs = TRUE, escape = FALSE) |>
+  kable_styling(position = "center", 
+                full_width = FALSE, 
+                font_size = 12) |>
+  row_spec(0, bold = TRUE) |>
+  column_spec(1:2, width = c("3cm", "12cm"))
 ```

### 2025-06-11 paper.qmd
Initial set up for the video link table

@@
 
-## Supplementary Materials
+## Supplementary Materials {#sec-supplementary}
 
@@
 #| tbl-pos: H
-#| tbl-cap: "Videos of the langevitour animations and linked plots."
+#| tbl-cap: "Videos of the langevitour animations and the linked plots."
 # Video links in the paper
@@
   row_spec(0, bold = TRUE) |>
-  column_spec(1:2, width = c("3cm", "12cm"))
+  column_spec(1:3, width = c("3cm", "5cm", "7cm"))
 ```

### 2025-06-11 paper.qmd
Added descriptions to supplementary table

@@
   row_spec(0, bold = TRUE) |>
-  column_spec(1:3, width = c("3cm", "5cm", "7cm"))
+  column_spec(1:3, width = c("1cm", "5cm", "9cm"))
 ```

### 2025-06-11 paper.qmd
Removed unnecessary code chunk in MNIST

@@
 
-<!--add langevitour screenshots and youtube animation link-->
-```{r}
-#| label: fig-mnist-tri-proj
-#| fig-pos: H
-#| fig-height: 4
-#| fig-width: 12
-#| out-width: 100%
-#| fig-cap: "\\gD{} wireframe layout shows that there is a big non-linear cluster (grey) and a small cluster (orange) located very close to the one corner of the big cluster.
-The MNIST digit 1 data ($n =  7877$ and $p = 10$) has a non-linear structure in $10\\text{-}D$.
-Two \\gD{} projections from a tour on $10\\text{-}D$ reveal that the closeness of the clusters in $10\\text{-}D$ and the twisted pattern of the model fit with tSNE ($a_1 = 0.041$, $b = 659/585 \\text{ } (30, \\text{ }48)$).
-Video of the langevitour animation is available at <>."
-#| eval: false
-
-free(trimesh_removed_mnist) + mnist_proj_tsne_model1 +
-  mnist_proj_tsne_model2 +
-  plot_layout(ncol = 3)
-
-```
-
 ```{r}
@@
   row_spec(0, bold = TRUE) |>
-  column_spec(1:3, width = c("1cm", "5cm", "9cm"))
+  column_spec(1:3, width = c("1.5cm", "4.5cm", "9cm"))
 ```

### 2025-06-11 paper.qmd
Added initial link plots

@@
 
+df_bin_dens_clust_temp <- df_bin_dens_clust
+
+error_df_one_curvy_abs_temp <- error_df_one_curvy_abs
+
 error_df_one_curvy_abs <- error_df_one_curvy_abs |>
@@
 error_df_one_curvy_abs <- error_df_one_curvy_abs |>
-  mutate(error_cat_n = if_else(tSNE1 <= 0.25 & tSNE2 <= 0.25, "selected", "deselected")) |> ## high_error points
-  mutate(error_cat_n2 = if_else(tSNE2 >= 1.2, "selected", "deselected")) ## corner points
+  mutate(error_cat_n = if_else(emb1 <= 0.25 & emb2 <= 0.25, "selected", "deselected")) |> ## high_error points
+  mutate(error_cat_n2 = if_else(emb2 >= 1.2, "selected", "deselected")) ## corner points
 
@@
 plot_tsne_dens <- error_df_one_curvy_abs_deselected12 |>
-  ggplot(aes(x = tSNE1,
-             y = tSNE2)) +
+  ggplot(aes(x = emb1,
+             y = emb2)) +
   geom_point(alpha=0.7, color = "#d9d9d9") +
   geom_point(data = error_df_one_curvy_abs_selected12,
-             aes(x = tSNE1, y = tSNE2),
+             aes(x = emb1, y = emb2),
              alpha=0.7, color = "#800026", size = 3) +
@@
 plot_tsne_dens_selected <- error_df_one_curvy_abs_deselected11 |>
-  ggplot(aes(x = tSNE1,
-             y = tSNE2)) +
+  ggplot(aes(x = emb1,
+             y = emb2)) +
   geom_point(alpha=0.7, color = "#d9d9d9") +
   geom_point(data = error_df_one_curvy_abs_selected11,
-             aes(x = tSNE1, y = tSNE2),
+             aes(x = emb1, y = emb2),
              alpha=0.7, color = "#800026", size = 3) +
@@
 ```{r}
-#| label: langevitour-c-shaped-tsne-proj
+#| label: link-plts-c-shaped-data
 #| eval: false
 
-data_c_shaped_n <- data_c_shaped |>
-  select(-type) |>
-  mutate(type = factor(error_df_one_curvy_abs$error_cat_n2,
-                          levels=c("deselected", "selected")))
-
-df_model_data_n <- bind_rows(df_bin_dens_clust, data_c_shaped_n)
-
-langevitour::langevitour(df_model_data_n[1:(length(df_model_data_n)-1)],
-                         lineFrom = tr_from_to_df_dens_clust$from,
-                         lineTo = tr_from_to_df_dens_clust$to,
-                         group = factor(df_model_data_n$type,
-                                        c("deselected", "selected", "model")),
-                         levelColors = c('#d9d9d9', '#800026', "#000000"),
-                         pointSize = c(1.5, 3, 0))
+df_exe_c_shaped_data <- comb_all_data_model_error(
+  highd_data = one_c_shaped_data,
+  nldr_data = tsne_one_c_shaped,
+  model_highd = df_bin_dens_clust_temp,
+  model_2d = df_bin_centroids_dens_clust,
+  error_data = error_df_one_curvy_abs_temp)
+
+show_error_link_plots(point_data = df_exe_c_shaped_data,
+                      edge_data = tr_from_to_df_dens_clust)
 ```
@@
 
+tsne_minst_scaled_temp <- tsne_minst_scaled |>
+  rename(c("emb1" = "tSNE1",
+           "emb2" = "tSNE2"))
+
 tsne_minst_scaled <- tsne_minst_scaled |>
@@
 ```{r}
+#| label: link-plts-mnist
+#| eval: false
+
+df_exe_mnist <- comb_all_data_model(
+  highd_data = training_data_mnist,
+  nldr_data = tsne_minst_scaled_temp,
+  model_highd = df_bin_mnist,
+  model_2d = df_bin_centroids_dens_clust)
+
+show_link_plots(point_data = df_exe_mnist,
+                edge_data = tr_from_to_df_dens_clust)
+```
+
+```{r}
 #| label: mnist-model-proj

### 2025-06-11 paper.qmd
Consistent with the word hyper-parameter

@@
 But NLDR views have an important role in data analysis because, if done well, they provide a concise visual (and conceptual) summary of \pD{} distributions.
-The NLDR methods and (hyper)parameter choices can create wildly different representations, making it difficult to decide which is best, or whether any or all are accurate or misleading.
+The NLDR methods and (hyper-)~parameter choices can create wildly different representations, making it difficult to decide which is best, or whether any or all are accurate or misleading.
 To help assess the NLDR and decide on which, if any, is the most reasonable representation of the structure(s) present in the \pD{} data, we have developed an algorithm to show the \gD{} NLDR model in the \pD{} space, viewed with a tour, a movie of linear projections.
@@
 
-We define the NLDR as a function $g\text{:}~ \mathbb{R}^{n\times p} \rightarrow \mathbb{R}^{n\times k}$, with (hyper-)parameters $\mathbfit{\theta}$.
+We define the NLDR as a function $g\text{:}~ \mathbb{R}^{n\times p} \rightarrow \mathbb{R}^{n\times k}$, with (hyper-)~parameters $\mathbfit{\theta}$.
 The parameters, $\mathbfit{\theta}$, depend on the choice of $g$, and can be considered part of model fitting in the traditional sense.
@@
 This highlights a key characteristic of tSNE, which compresses the data when applying tSNE.
-Furthermore, the sparse space observed in the middle of the tSNE layout is a result of (hyper-)parameter choice.
+Furthermore, the sparse space observed in the middle of the tSNE layout is a result of (hyper-)~parameter choice.
 Video of the langevitour animation is available at <>."
@@
 @fig-five-gau-projs a1, b1, c1 show the \gD{} layouts for (a) tSNE, (b) UMAP, and (c) PaCMAP, respectively.
-The default hyper-parameters for each method are used.
+The default (hyper-)~parameters for each method are used.
 In each layout we can see an accurate representation where all five clusters are visible, although with varying degrees of separation.
@@
 
-@fig-one-dens_clust-error (a2, b2) shows \gD{} layout for tSNE created using the default hyper-parameters.
+@fig-one-dens_clust-error (a2, b2) shows \gD{} layout for tSNE created using the default (hyper-)~parameters.
 One would expect to see a rectangular shape if the curved sheet is flattened, but the layout is triangular.
@@
 #| echo: false
-#| fig-cap: "Compare the published \\gD{} layout (@fig-pbmc-mse a) and the \\gD{} layout selected (@fig-pbmc-mse e) by MSE plot (@fig-pbmc-mse) from the UMAP and tSNE with default hyper-parameters.
+#| fig-cap: "Compare the published \\gD{} layout (@fig-pbmc-mse a) and the \\gD{} layout selected (@fig-pbmc-mse e) by MSE plot (@fig-pbmc-mse) from the UMAP and tSNE with default (hyper-)~parameters.
 The PBMC3k data ($n =  2622$) has three close clusters in $9\\text{-}D$.
@@
 
-@fig-mnist-mse compares the fit of six layouts computed using UMAP (b), PHATE (c), TriMAP (d), PaCMAP (e) with default (hyper-)parameter setting and two tSNE runs, one with default (hyper-)parameter setting (a) and the other changing perplexity to $89$ (f).
+@fig-mnist-mse compares the fit of six layouts computed using UMAP (b), PHATE (c), TriMAP (d), PaCMAP (e) with default (hyper-)~parameter setting and two tSNE runs, one with default (hyper-)~parameter setting (a) and the other changing perplexity to $89$ (f).
 The layouts are reasonably similar in that they all have the observations in a single blob.
@@
 
-We have developed an approach to help assess and compare NLDR layouts, generated by different methods and (hyper-)parameter choice(s).
+We have developed an approach to help assess and compare NLDR layouts, generated by different methods and (hyper-)~parameter choice(s).
 It depends on conceptualising the \gD{} layout as a model, creating a wireframe representation of it that can be lifted into \pD{}.

### 2025-06-11 paper.qmd
Changed some kD to gD

@@
 - Remove all the extraneous commented text (Done)
-- Check how \kD{} and \gD{} are used, and make sure it's sensible
+- Check how \kD{} and \gD{} are used, and make sure it's sensible (changed in Rendering the model)
 - Check spelling is consistently American
@@
 
-The last step is to lift the \kD{} model into \pD{} by computing \pD{} vectors that represent bin centroids.
+The last step is to lift the \gD{} model into \pD{} by computing \pD{} vectors that represent bin centroids.
 We use the \pD{} mean of the points in $H_h$ to map the centroid $C_{h}^{(2)} = (c_{h1}, c_{h2})$ to a point in \pD{}.
@@
 $$C_{h}^{(p)} = \frac{1}{n_h}\sum_{i =1}^{n_h} x_i, h = {1, \dots, b; n_h > 0}.$$
-Furthermore, line segments that exist in the \kD{} model generate line segments in \pD{} by connecting the \pD{} means of the corresponding \kD{} bin centroids.
+Furthermore, line segments that exist in the \gD{} model generate line segments in \pD{} by connecting the \pD{} means of the corresponding \gD{} bin centroids.
 If additional long edges need to be removed, compute the edges in \pD{} and pruned any detected long edges to improve the accuracy.

### 2025-06-11 paper.qmd
Updated -0.1-0 range mathematical notations
Updated word colour to color

@@
                                   aes(x = emb1, y = emb2)) +
-  geom_point(alpha = 0.5, color = clr_choice, size = 0.5)
+  geom_point(alpha = 0.5, colour = clr_choice, size = 0.5)
 
@@
   aes(x = x, y = y)) +
-  geom_polygon(color = "grey70", 
+  geom_polygon(colour = "grey70", 
                aes(group = hex_poly_id), 
@@
              aes(x = emb1, y = emb2), 
-             alpha = 0.5, size = 0.5, color = clr_choice) +
+             alpha = 0.5, size = 0.5, colour = clr_choice) +
   xlim(sc_xlims) + ylim(sc_ylims) +
@@
       group = hex_poly_id, 
-      fill = std_counts), color = "grey70", linewidth=0.2) +
+      fill = std_counts), colour = "grey70", linewidth=0.2) +
   geom_point(data = tsne_two_curvy_scaled,
@@
              aes(x = c_x, y = c_y), 
-             color = "#000000",
+             colour = "#000000",
              size = 1) +
@@
              ),
-             color = "#636363",
+             colour = "#636363",
              alpha = 0.5,
@@
              ),
-             color = clr_choice,
+             colour = clr_choice,
              alpha = 0.5,
@@
                  fill = "white", 
-                 color = "#bdbdbd") +
+                 colour = "#bdbdbd") +
     geom_point(data = all_centroids_df_temp, aes(
@@
       y = c_y), 
-      color = "#31a354", size = 0.9) +
+      colour = "#31a354", size = 0.9) +
     geom_point(data = start_pt, aes(x = c_x, 
                                     y = c_y), 
-               color = "black") + 
+               colour = "black") + 
     geom_rect(data=d_rect, 
@@
               fill = "white", 
-              color = "black", 
+              colour = "black", 
               alpha = 0, 
@@
                                ends = "both"), 
-                 color = "black")+
+                 colour = "black")+
     annotate("text", x=0.5, y=-0.36, 
-             label = expression(r[1]), color = "black", size = 8) +
+             label = expression(r[1]), colour = "black", size = 8) +
     geom_segment(data=d_rect, aes(
@@
                        ends = "both"), 
-                 color = "black")+ 
+                 colour = "black")+ 
     annotate("text", x=-0.3, y=0.4, 
-             label = expression(r[2]), color = "black", size = 8) +
+             label = expression(r[2]), colour = "black", size = 8) +
     geom_segment(data = a1, aes(
@@
         ends = "both"), 
-        color = "black")+ # a1 = 0.2087578
+        colour = "black")+ # a1 = 0.2087578
     annotate("text", 
@@
              label = expression(a[1]), 
-             color = "black",
+             colour = "black",
              vjust = 1.2, size = 8) +
@@
                                ends = "both"), 
-      color = "black") + # a2 = 0.1807896
+      colour = "black") + # a2 = 0.1807896
     annotate("text", x=a2$x, y=(a2$y+a2$yend)/2, 
              label = expression(a[2]), 
-             color = "black", hjust=-0.2, size = 8) +
+             colour = "black", hjust=-0.2, size = 8) +
     annotate("text", x=-0.18, y=-0.24, 
       label = expression(group("(", list(s[1], s[2]), ")")),
-      color = "black", size = 8) +
+      colour = "black", size = 8) +
   geom_segment(data = l, aes(
@@
                                ends = "both"), 
-      color = "black") + 
+      colour = "black") + 
     annotate("text", x=l$x + 0.03, y=(l$y+l$yend)/2, 
              label = expression(l), 
-             color = "black", hjust=-0.2, size = 8) +
+             colour = "black", hjust=-0.2, size = 8) +
   coord_equal()
@@
         fill = std_counts), 
-        color = "grey70", 
+        colour = "grey70", 
         linewidth=0.2) +
@@
              ),
-             color = "#636363",
+             colour = "#636363",
             alpha = 0.5,
@@
         fill = std_counts), 
-        color = "grey70", linewidth=0.2) +
+        colour = "grey70", linewidth=0.2) +
   geom_point(data = tsne_two_curvy_scaled,
@@
   geom_vline(xintercept = 0.09,
-           color = "#d95f02", linetype="dashed", 
+           colour = "#d95f02", linetype="dashed", 
            linewidth=1) +
   geom_vline(xintercept = 0.05,
-             color = "#1f78b4", 
+             colour = "#1f78b4", 
              linetype="solid",
@@
   geom_vline(xintercept = 0.03, linetype="dotted",
-             color = "#1b9e77", linewidth=1) +
+             colour = "#1b9e77", linewidth=1) +
   geom_line(linewidth = 0.5) + 
@@
   geom_vline(xintercept = 0.03, linetype="dotted",
-             color = "#1b9e77", linewidth=1) +
+             colour = "#1b9e77", linewidth=1) +
   geom_line(linewidth = 0.5) +
@@
 #| fig-cap: "The $4\\text{-}D$ model error in \\gD{} layout.
-Color indicates error ($e_{hj}$), dark colour indicating high error and light indicates low error.
+Color indicates error ($e_{hj}$), dark color indicating high error and light indicates low error.
 Most large errors are distributed near the sparse end of the non-linear cluster with dense corner."
@@
 The values $(s_1, \ s_2)$ define the position of the centroid of the bottom left hexagon.
-By default, this is at $s_1 = -q, s_2 = -qr_2$, where $q$ is the buffer bound the data.
+By default, this is at $s_1 = \text{-}q, s_2 = \text{-}qr_2$, where $q$ is the buffer bound the data.
 The choice of these values can have some effect on the distribution of bin counts.
 @fig-bins-two-curvy illustrates this.
-The distribution of bin counts for $s_1$ varying between $-0.1-0.0$ is shown.
+The distribution of bin counts for $s_1$ varying between $\text{-}0.1\,\text{--}\,0.0$ is shown.
 Generally, a more uniform distribution among these possibilities would indicate that the bins are reliably capturing the underlying distribution of observations.
@@
                          y = RMSE,
-                         color = a1,
+                         colour = a1,
                          linetype = a1)) + 
@@
 What does it mean to be a best fit for this problem?
-Analysts use an NLDR layout to display the structure present in high-dimensional data in a convenient 2D display.
+Analysts use an NLDR layout to display the structure present in high-dimensional data in a convenient \gD{} display.
 It is a competitor to linear dimension reduction that can better represent non-linear associations such as clusters.
@@
 When you examine their data in a tour, it is almost always that we see there are no big separations, and actually the suggested clusters are not even present.
-While we don't expect that analysts include animated gifs of tours in their papers, we should expect that any 2D representation adequately indicates the clustering that is present, and honestly show lack of separation or lack of clustering when it doesn't exist.
+While we don't expect that analysts include animated gifs of tours in their papers, we should expect that any \gD{} representation adequately indicates the clustering that is present, and honestly show lack of separation or lack of clustering when it doesn't exist.
 It is important for analysts to have tools to select the accurate representation not the pretty but wrong representation.
@@
 #| fig-cap: "Assessing which of the 6 NLDR layouts on the two non-linear clusters data is the better representation using RMSE for varying binwidth ($a_1$).
-Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f).
+Color used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f).
 Layout d is universally poor.
@@
       yend = y_to),
-    color = "#000000",
+    colour = "#000000",
     #alpha = 0.4,
@@
     alpha = 0.05,
-    color = clr_choice) +
+    colour = clr_choice) +
   geom_segment(
@@
       yend = proj2_to), 
-    color = "#000000",
+    colour = "#000000",
     alpha = 0.4,
@@
       yend = proj2_to),
-    color = "#000000",
+    colour = "#000000",
     alpha = 0.4,
@@
           y = density)) +
-  geom_point(alpha=0.7, color = "#d9d9d9") +
+  geom_point(alpha=0.7, colour = "#d9d9d9") +
   geom_point(data = error_df_one_curvy_abs_selected11, 
@@
           y = density),
-      alpha=0.7, color = "#800026", size = 3) +
+      alpha=0.7, colour = "#800026", size = 3) +
   xlab(expression(e[hj])) +
@@
           y = density)) +
-  geom_point(alpha=0.7, color = "#d9d9d9") +
+  geom_point(alpha=0.7, colour = "#d9d9d9") +
   geom_point(data = error_df_one_curvy_abs_selected12, 
@@
              y = emb2)) +
-  geom_point(alpha=0.7, color = "#d9d9d9") +
+  geom_point(alpha=0.7, colour = "#d9d9d9") +
   geom_point(data = error_df_one_curvy_abs_selected12,
              aes(x = emb1, y = emb2),
-             alpha=0.7, color = "#800026", size = 3) +
+             alpha=0.7, colour = "#800026", size = 3) +
   theme(aspect.ratio = 1) +
@@
              y = emb2)) +
-  geom_point(alpha=0.7, color = "#d9d9d9") +
+  geom_point(alpha=0.7, colour = "#d9d9d9") +
   geom_point(data = error_df_one_curvy_abs_selected11,
              aes(x = emb1, y = emb2),
-             alpha=0.7, color = "#800026", size = 3) +
+             alpha=0.7, colour = "#800026", size = 3) +
   theme(aspect.ratio = 1) +
@@
           yend = proj2_to),
-        color = "#000000",
+        colour = "#000000",
         linewidth = 0.8,
@@
       y = UMAP2,
-      color = cluster
+      colour = cluster
     ),
@@
     #size = 0.1#,
-    #color = '#a65628'
+    #colour = '#a65628'
   )  +
@@
 #| fig-cap: "Assessing which of the 8 NLDR layouts on the PBMC3k data  (shown in @fig-NLDR-variety) is the better representation using RMSE for varying binwidth ($a_1$).
-Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-h).
+Color used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-h).
 Layout f is universally poor.
@@
       y = tSNE2,
-      color = cluster
+      colour = cluster
     ),
     alpha = 0.3#,
-    #color = "#999999"
+    #colour = "#999999"
   ) +
@@
 #| fig-cap: "Assessing which of the 6 NLDR layouts on the MNIST digit 1 data is the better representation using RMSE for varying binwidth ($a_1$).
-Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f).
+Color used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f).
 All the layouts appear to be very similar.
@@
   geom_rect(aes(xmin = 0.82, xmax = 0.92, ymin = 0.38, ymax = 0.58),
-               fill = "transparent", color = "#000000", linewidth = 1) +
+               fill = "transparent", colour = "#000000", linewidth = 1) +
   geom_rect(aes(xmin = 0.9, xmax = 1, ymin = 1.25, ymax = 1.45),
-               fill = "transparent", color = "#375cb1", linewidth = 1)  +
+               fill = "transparent", colour = "#375cb1", linewidth = 1)  +
   geom_rect(aes(xmin = 0.35, xmax = 0.45, ymin = 0.55, ymax = 0.7),
-               fill = "transparent", color = "#008000", linewidth = 1)  +
+               fill = "transparent", colour = "#008000", linewidth = 1)  +
   geom_rect(aes(xmin = 0.7, xmax = 0.8, ymin = 0.25, ymax = 0.4),
-               fill = "transparent", color = "#4c389a", linewidth = 1) +
+               fill = "transparent", colour = "#4c389a", linewidth = 1) +
   interior_annotation(label = "a", c(0.08, 0.93), cex = 2) +

### 2025-06-11 paper.qmd
Corrected the spelling

@@
 The curious detail is that the model for PaCMAP is closer to a *pancake* in shape in every cluster!
-This seingle projection only shows this in three of the five clusters but if we examine a different projection the other clusters exhibit the pancake also.
+This single projection only shows this in three of the five clusters but if we examine a different projection the other clusters exhibit the pancake also.
 While we don't know what exactly causes this it is likely due to some ordering of points in the \gD{} PaCMAP layout that induces the flat model.

### 2025-06-11 paper.qmd
replaced non-linear with nonlinear

@@
 
-To illustrate the method, we use $7\text{-}D$ simulated data, which we call the "non-linear clusters".
+To illustrate the method, we use $7\text{-}D$ simulated data, which we call the "nonlinear clusters".
 It is constructed by simulating two clusters, each consisting of $1000$ observations.
@@
 #| label: fig-two-curvy-true-proj
-#| fig-cap: "Two \\gD{} projections from $4\\text{-}D$, for the two non-linear clusters data (blue points) and true model (black lines) are shown.
+#| fig-cap: "Two \\gD{} projections from $4\\text{-}D$, for the two nonlinear clusters data (blue points) and true model (black lines) are shown.
 The top cluster in (a) shows uniformly distributed data along a C-shape, while in the bottom cluster, the data is more concentrated in one corner of the curvy shape.
@@
 #| echo: false
-#| fig-cap: "Key steps for constructing the model on the tSNE layout ($k=2$): (a) data, (b) hexagon bins, (c) bin centroids, and (d) triangulated centroids. The two non-linear clusters data is shown."
+#| fig-cap: "Key steps for constructing the model on the tSNE layout ($k=2$): (a) data, (b) hexagon bins, (c) bin centroids, and (d) triangulated centroids. The two nonlinear clusters data is shown."
 #| fig-width: 12
@@
 
-When shapes are non-linear in the NLDR layout, some edges could be long.
+When shapes are nonlinear in the NLDR layout, some edges could be long.
 It can also happen that distant centroids can be connected, particularly if clustering is present, which can result in long line segments.
@@
 #| fig-cap: "Lifting the fitted model into \\pD{}.
-Starting from a default tSNE layout (a) of non-linear clusters data ($n =  2000$ and $p = 4$), from which the \\gD{} wireframe is constructed (a) and then lifted to $4\\text{-}D$ (b, c).
+Starting from a default tSNE layout (a) of nonlinear clusters data ($n =  2000$ and $p = 4$), from which the \\gD{} wireframe is constructed (a) and then lifted to $4\\text{-}D$ (b, c).
 The fit is reasonbly tight with the data, although it does not fully spread the full width.
@@
 Color indicates error ($e_{hj}$), dark color indicating high error and light indicates low error.
-Most large errors are distributed near the sparse end of the non-linear cluster with dense corner."
+Most large errors are distributed near the sparse end of the nonlinear cluster with dense corner."
 #| out-width: 80%
@@
 #| fig-pos: H
-#| fig-cap: "Hexbin density plots of tSNE layout of the non-linear cluster data, using three different bin inputs: (a) $b = 240/98 \\text{ } (15, \\text{ }16)$, (b) $b = 720/215 \\text{ } (24, \\text{ }30)$, and (c) $b = 2496/549 \\text{ } (48, \\text{ }52)$.
+#| fig-cap: "Hexbin density plots of tSNE layout of the nonlinear cluster data, using three different bin inputs: (a) $b = 240/98 \\text{ } (15, \\text{ }16)$, (b) $b = 720/215 \\text{ } (24, \\text{ }30)$, and (c) $b = 2496/549 \\text{ } (48, \\text{ }52)$.
 Color indicates standardized counts, dark indicating high count and light indicates low count.
@@
 Analysts use an NLDR layout to display the structure present in high-dimensional data in a convenient \gD{} display.
-It is a competitor to linear dimension reduction that can better represent non-linear associations such as clusters.
+It is a competitor to linear dimension reduction that can better represent nonlinear associations such as clusters.
 However, these methods can hallucinate, suggesting patterns that don't exist, and grossly exaggerate other patterns.
@@
 ```{r}
-#| fig-cap: "Assessing which of the 6 NLDR layouts on the two non-linear clusters data is the better representation using RMSE for varying binwidth ($a_1$).
+#| fig-cap: "Assessing which of the 6 NLDR layouts on the two nonlinear clusters data is the better representation using RMSE for varying binwidth ($a_1$).
 Color used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f).
@@
 Therefore, @fig-pbmc-mse e is more reasonable than @fig-pbmc-mse a.
-Also, the fitted models in $9\\text{-}D$ help to see some unobserved patterns of the data: (i) dense points, and (ii) non-linear clusters.
+Also, the fitted models in $9\\text{-}D$ help to see some unobserved patterns of the data: (i) dense points, and (ii) nonlinear clusters.
 Videos of the langevitour animations are available at <https://youtu.be/3VfK3M2gnZM> and <https://youtu.be/Es84bwQcndU> respectively."
@@
 #| out-width: 100%
-#| fig-cap: "The tSNE layout of the MNIST digit 1 data shows a big non-linear cluster (grey) and a small cluster (orange) located very close to the one corner of the big cluster in \\gD{} (a).
-The MNIST digit 1 data ($n =  7877$ and $p = 10$) has a non-linear structure in $10\\text{-}D$.
+#| fig-cap: "The tSNE layout of the MNIST digit 1 data shows a big nonlinear cluster (grey) and a small cluster (orange) located very close to the one corner of the big cluster in \\gD{} (a).
+The MNIST digit 1 data ($n =  7877$ and $p = 10$) has a nonlinear structure in $10\\text{-}D$.
 Two \\gD{} projections from a tour on $10\\text{-}D$ reveal that the closeness of the clusters in $10\\text{-}D$ and the twisted pattern of the model fit with tSNE ($a_1 = 0.041$, $b = 659/585 \\text{ } (30, \\text{ }48)$).
@@
 #| fig-cap: "The $10\\text{-}D$ model error in \\gD{} layout of the MNIST digit 1 data shows a pattern.
-Most low model errors are distributed along the big non-linear cluster, while most large model errors are distributed along the small cluster.
-The images associated with large model errors shows different patterns of digit 1, some inside (f) the non-linear structure and others outside (e).
-Along the non-linear cluster, the angle of digit 1 changes (b-d)."
+Most low model errors are distributed along the big nonlinear cluster, while most large model errors are distributed along the small cluster.
+The images associated with large model errors shows different patterns of digit 1, some inside (f) the nonlinear structure and others outside (e).
+Along the nonlinear cluster, the angle of digit 1 changes (b-d)."
 #| fig-pos: H

### 2025-06-11 paper.qmd
Corrected grammar

@@
 The tSNE behaviour is to squeeze the low density area of the data together into the layout.
-This is common in other NLDR methods also, which means analysts needs to be aware the if their data is not sampled relatively uniformly apparent closeness in the \gD{} may correspond to sparseness in \pD{}.
+This is common in other NLDR methods also, which means analysts need to be aware that if their data is not sampled relatively uniformly, apparent closeness in the \gD{} may correspond to sparseness in \pD{}.
 

### 2025-06-11 paper.qmd
Updated summarise to summarize

@@
 Clustering of single-cell data is used to identify groups of cells with similar expression profiles.
-NLDR often used to summarise the cluster structure.
+NLDR often used to summarize the cluster structure.
 Usually, NLDR does not use the cluster labels to compute the layout, but uses color to represent the cluster labels when it is plotted.
@@
 They can be seen to be more strangely drawn than the others.
-Overall, this \gD{} layout shows a useful way to summarise the variation in way "1"s are drawn.
+Overall, this \gD{} layout shows a useful way to summarize the variation in way "1"s are drawn.
 
@@
 
-Additional exploration of metrics to summarise the fit could be a new direction for the work.
+Additional exploration of metrics to summarize the fit could be a new direction for the work.
 The difficulty is capturing nonlinear fits, for which Euclidean distance can be sub-optimal.
@@
 Most NLDR methods only provide a \gD{} but if a \kD{} ($k>2$) layout is provided the approach developed here could be extended.
-Binning into cubes could be done in $3\text{-}D$ or higher, relatively easily, and used asa the basis for a wireframe of the fitted model.
+Binning into cubes could be done in $3\text{-}D$ or higher, relatively easily, and used as a the basis for a wireframe of the fitted model.
 @barber1996 and the software @stephane2023 have algorithms for convex hulls, which \pD{} which serve as an inspiration.
@@
 The new methodology is accompanied by an R package called quollr, so that it is readily usable and broadly accessible.
-The package has methods to fit the model, compute diagnostics and also visualise the results, with interactivity.
+The package has methods to fit the model, compute diagnostics and also visualize the results, with interactivity.
 We have primarily used the langevitour software [@harisson2024] to view the model in the data space, but other tour software such as tourr [@wickham2011] and detourr [@hart2022] could be also used.

### 2025-06-11 paper.qmd
Finalize American English spellings in the content

@@
 - Remove all the extraneous commented text (Done)
-- Check how \kD{} and \gD{} are used, and make sure it's sensible (changed in Rendering the model)
-- Check spelling is consistently American
+- Check how \kD{} and \gD{} are used, and make sure it's sensible (changed in Rendering the model: Done)
+- Check spelling is consistently American (Done)
 - Check references (Done)
@@
 The digit "1" of the MNIST dataset [@lecun1998] consists of $7877$ grayscale images of handwritten "1"s.
-Each image is $28 \times 28$ pixels which corresponds to 784 variables.
-The first 10 principal components, explaining 83% of the total variation, are used.
-This data is essentially lies on a nonlinear manifold in the high dimensions, defined by the shapes that "1"s make in when sketched, menaing that some pixels are almost always dark and some almost always light.
+Each image is $28 \times 28$ pixels which corresponds to $784$ variables.
+The first $10$ principal components, explaining $83\%$ of the total variation, are used.
+This data is essentially lies on a nonlinear manifold in the high dimensions, defined by the shapes that "1"s make in when sketched, meaning that some pixels are almost always dark and some almost always light.
 We expect that the best layout captures this type of structure and does not exhibit distinct clusters.
@@
 We have developed an approach to help assess and compare NLDR layouts, generated by different methods and (hyper-)~parameter choice(s).
-It depends on conceptualising the \gD{} layout as a model, creating a wireframe representation of it that can be lifted into \pD{}.
+It depends on conceptualizing the \gD{} layout as a model, creating a wireframe representation of it that can be lifted into \pD{}.
 The fit is assessed by viewing the model in the data space, computing residuals and RMSE.
@@
 
-Two examples illustrating usage are provided: the PBMC3k data where the NLDR is summarising clustering in \pD{} and hand-written digits illustrating how NLDR represents an intrinsically lower dimensional nonlinear manifold.
+Two examples illustrating usage are provided: the PBMC3k data where the NLDR is summarizing clustering in \pD{} and hand-written digits illustrating how NLDR represents an intrinsically lower dimensional nonlinear manifold.
 We demonstrated that the published layout of the PBMC3k is inaccurate, because it grossly exaggerates separation between clusters, and even suggests separation when there is none.

### 2025-06-11 paper.qmd
Updated - values with proper notations

@@
 It is constructed by simulating two clusters, each consisting of $1000$ observations.
-The C-shaped cluster is generated from $\theta \sim U(-3\pi/2, 0)$, $X_1 = \sin(\theta)$, $X_2 \sim U(0, 2)$ (adding thickness to the C), $X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)$, $X_4 = \cos(\theta)$.
-The other cluster is from $X_1 \sim U(0, 2)$, $X_2 \sim U(0, 3)$, $\gamma \sim U(0, 0.5)$, $X_3 = -(X_1^3 + X_2) + \gamma$, and $X_4 \sim U(0, 2)$.
+The C-shaped cluster is generated from $\theta \sim U(\text{-}3\pi/2, 0)$, $X_1 = \sin(\theta)$, $X_2 \sim U(0, 2)$ (adding thickness to the C), $X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)$, $X_4 = \cos(\theta)$.
+The other cluster is from $X_1 \sim U(0, 2)$, $X_2 \sim U(0, 3)$, $\gamma \sim U(0, 0.5)$, $X_3 = \text{-}(X_1^3 + X_2) + \gamma$, and $X_4 \sim U(0, 2)$.
 We would consider $T=(X_1, X_2, X_3, X_4)$ to be the geometric structure (true model) that we hope to capture.
@@
 This must be smaller than the minimum data value.
-Because it is one buffer unit, $q$ below the minimum data values, $s_1 = -q$ and $s_2 = -qr_2$.
+Because it is one buffer unit, $q$ below the minimum data values, $s_1 = \text{-}q$ and $s_2 = \text{-}qr_2$.
 

### 2025-06-11 appendix.qmd
Updated - sign

@@
 
-Given range of embedding component, $r_2$, number of bins along the x-axis, $b_1$, and buffer proportion, $q$, hexagonal starting point coordinates, $s_1 = -q$, and $s_2 = -q \times r_2$.
+Given range of embedding component, $r_2$, number of bins along the x-axis, $b_1$, and buffer proportion, $q$, hexagonal starting point coordinates, $s_1 = \text{-}q$, and $s_2 = \text{-}qr_2$.
 The purpose is to find width of the hexagon.
@@
 a_1, a_2 > 0.
-$$ Also, $(s_1, s_2) \in (-0.1, -0.05)$ as these are multiplicative offsets in the negative direction.
+$$ Also, $(s_1, s_2) \in (\text{-}0.1, \text{-}0.05)$ as these are multiplicative offsets in the negative direction.
 
@@
 which given $s_1 < 0$ and $a_1 > 0$ will *always* be true.
-The same logic follows for @eq-equation4 and substituting $a_2 = \sqrt{3}a_1/{2}$, and $s_2 = -q \times r_2$ to @eq-equation4 can be written as,
+The same logic follows for @eq-equation4 and substituting $a_2 = \sqrt{3}a_1/{2}$, and $s_2 = \text{-}qr_2$ to @eq-equation4 can be written as,
 
@@
 
-Also, substituting $a_2 = \sqrt{3}a_1/{2}$, $s_2 = -q \times r_2$ and rearranging @eq-equation5 gives:
+Also, substituting $a_2 = \sqrt{3}a_1/{2}$, $s_2 = \text{-}qr_2$ and rearranging @eq-equation5 gives:
 
@@
 
-Similarly, substituting $s_1 = -q$ @eq-equation2 becomes,
+Similarly, substituting $s_1 = \text{-}q$ @eq-equation2 becomes,
 
@@
 
-The area of a hexagon is defined as $A = \frac{3\sqrt{3}}{2}l^2$, where $l$ is the side length of the hexagon. $l$ can be computed using $a_1$ and $a_2$.
+The area of a hexagon is defined as $A = 3\sqrt{3}l^2/2$, where $l$ is the side length of the hexagon. $l$ can be computed using $a_1$ and $a_2$.
 

### 2025-06-11 appendix.tex
Added the notation table to appendix
Updated - sign

@@
+% Options for packages loaded elsewhere
 % Options for packages loaded elsewhere
@@
   12pt]{article}
-
+\usepackage{xcolor}
 \usepackage{amsmath,amssymb}
+\setcounter{secnumdepth}{5}
 \usepackage{iftex}
@@
 \else % if luatex or xetex
-  \usepackage{unicode-math}
+  \usepackage{unicode-math} % this also loads fontspec
   \defaultfontfeatures{Scale=MatchLowercase}
@@
 \makeatother
-\usepackage{xcolor}
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\setcounter{secnumdepth}{5}
 % Make \paragraph and \subparagraph free-standing
@@
 
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
+\usepackage{longtable,booktabs,array}
 \usepackage{calc} % for calculating minipage widths
@@
 \makeatother
+
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+\usepackage[]{natbib}
+\bibliographystyle{agsm}
 
@@
 \makeatother
-
-\usepackage[]{natbib}
-\bibliographystyle{agsm}
 \usepackage{bookmark}
-
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same} % disable monospaced font for URLs
+\urlstyle{same}
 \hypersetup{
@@
 x-axis, \(b_1\), and buffer proportion, \(q\), hexagonal starting point
-coordinates, \(s_1 = -q\), and \(s_2 = -q \times r_2\). The purpose is
-to find width of the hexagon. \(a_1\), and number of bins along the
-y-axis, \(b_2\).
+coordinates, \(s_1 = \text{-}q\), and \(s_2 = \text{-}qr_2\). The
+purpose is to find width of the hexagon. \(a_1\), and number of bins
+along the y-axis, \(b_2\).
 
@@
 a_1, a_2 > 0.
-\] Also, \((s_1, s_2) \in (-0.1, -0.05)\) as these are multiplicative
-offsets in the negative direction.
+\] Also, \((s_1, s_2) \in (\text{-}0.1, \text{-}0.05)\) as these are
+multiplicative offsets in the negative direction.
 
@@
 same logic follows for Equation~\ref{eq-equation4} and substituting
-\(a_2 = \sqrt{3}a_1/{2}\), and \(s_2 = -q \times r_2\) to
+\(a_2 = \sqrt{3}a_1/{2}\), and \(s_2 = \text{-}qr_2\) to
 Equation~\ref{eq-equation4} can be written as,
@@
 
-Also, substituting \(a_2 = \sqrt{3}a_1/{2}\), \(s_2 = -q \times r_2\)
-and rearranging Equation~\ref{eq-equation5} gives:
+Also, substituting \(a_2 = \sqrt{3}a_1/{2}\), \(s_2 = \text{-}qr_2\) and
+rearranging Equation~\ref{eq-equation5} gives:
 
@@
 
-Similarly, substituting \(s_1 = -q\) Equation~\ref{eq-equation2}
+Similarly, substituting \(s_1 = \text{-}q\) Equation~\ref{eq-equation2}
 becomes,
@@
 
-The area of a hexagon is defined as \(A = \frac{3\sqrt{3}}{2}l^2\),
-where \(l\) is the side length of the hexagon. \(l\) can be computed
-using \(a_1\) and \(a_2\).
+The area of a hexagon is defined as \(A = 3\sqrt{3}l^2/2\), where \(l\)
+is the side length of the hexagon. \(l\) can be computed using \(a_1\)
+and \(a_2\).
 

### 2025-06-15 paper.qmd
grammar fixes and some larger overall comments

@@
 #| fig-cap: "Eight different NLDR representations of the same data.
-Different techniques and different parameter choices are used.
+Different methods and different parameter choices are used.
 Researchers may have seen any of these in their analysis of this data, depending on their choice of method, or typical parameter choice.
@@
 
+<!-- Where are sections 4 & 5 in this description? -->
+
 ## Background {#sec-background}
@@
 <!-- - Connection between NLDR and MDS-->
-Historically, \kD{} representations of \pD{} data have been computed using multidimensional scaling (MDS) [@kruskal1964], which includes principal components analysis (PCA) [@jolliffe2011] as a special case.
-(A contemporary comprehensive guide to MDS can be found in @borg2005.) The \kD{} representation can be considered to be a layout of points in \kD{}   produced by an embedding procedure that maps the data from \pD{}.
+Historically, low-dimensional (\kD{}) representations of high-dimensional (\pD{}) data have been computed using multidimensional scaling (MDS) [@kruskal1964], which includes principal components analysis (PCA) [@jolliffe2011] as a special case.
+(A contemporary comprehensive guide to MDS can be found in @borg2005.)
+<!-- Does JASA allow footnotes?
+If so this is a footnote, otherwise the punctuation has gone a bit rogue - can't have an entire sentence in brackets surely --> The \kD{} representation can be considered to be a layout of points in \kD{}   produced by an embedding procedure that maps the data from \pD{}.
 In MDS, the \kD{} layout is constructed by minimizing a stress function that differences distances between points in \pD{} with potential distances between points in \kD{}.
@@
 
+<!-- Non-expert comment: This to me is a leap from MDS to NLDR. Is NLDR an MDS? I don't think so, but I'm missing the link -->
+<!-- Suggested rewording: "To better capture specific structures potentially existing in \pD{}, many new methods for NLDR have emerged in recent years." -->
 Many new methods for NLDR have emerged in recent years, all designed to better capture specific structures potentially existing in \pD{}.
-Here we focus on five currently popular techniques, tSNE, UMAP, PHATE, TriMAP and PaCMAP.
-tNSE and UMAP can be considered to produce the \kD{}   minimizing the divergence between two distributions, where the distributions are modeling the inter-point distances.
+Here we focus on five currently popular techniques: tSNE, UMAP, PHATE, TriMAP and PaCMAP.
+Both tNSE and UMAP can be considered to produce the \kD{} representation by minimizing the divergence between two distributions, where the distributions are modeling the inter-point distances.
 PHATE, TriMAP and PaCMAP are examples of diffusion processes [@coifman2005] spreading to capture geometric shapes, that include both global and local structure.
@@
 
-We define the NLDR as a function $g\text{:}~ \mathbb{R}^{n\times p} \rightarrow \mathbb{R}^{n\times k}$, with (hyper-)~parameters $\mathbfit{\theta}$.
-The parameters, $\mathbfit{\theta}$, depend on the choice of $g$, and can be considered part of model fitting in the traditional sense.
+We define the NLDR as a function $g\text{:}~ \mathbb{R}^{n\times p} \rightarrow \mathbb{R}^{n\times k}$, with (hyper-)parameters $\mathbfit{\theta}$.
+These parameters, $\mathbfit{\theta}$, depend on the choice of $g$, and can be considered part of model fitting in the traditional sense.
 Common choices for $g$ include functions used in tSNE, UMAP, PHATE, TriMAP, PaCMAP, or MDS, although in theory any function that does this mapping is suitable.
@@
 This must be smaller than the minimum data value.
-Because it is one buffer unit, $q$ below the minimum data values, $s_1 = \text{-}q$ and $s_2 = \text{-}qr_2$.
+Because it is one buffer unit, $q$ below the minimum data values, $s_1 = -q$ and $s_2 = -qr_2$.
 
@@
 This can be defined using the function $u: \mathbb{R}^{n\times 2} \rightarrow \mathbb{R}^{m\times 2}$, where
-$u(i) = \arg\min_{j = 1, \dots, b} \sqrt{(y_{i1} - C^{(2)}_{j1})^2 + (y_{i2} - C^{(2)}_{j2})^2}$, mapping observation $i$ into $H_h = \{i| u(i) = h\}$.
+$$u(i) = \arg\min_{j = 1, \dots, b} \sqrt{(y_{i1} - C^{(2)}_{j1})^2 + (y_{i2} - C^{(2)}_{j2})^2}$$, maps observation $i$ into $H_h = \{i| u(i) = h\}$.
+
+<!-- I'm not thrilled about H_h but I guess it can stay because changing it everywhere is a pain -->
 
@@
 Delaunay triangulation [@lee1980;@alb2024] is used to connect points so that edges indicate neighboring observations, in both the NLDR layout (@fig-NLDR-two-curvy (d)) and the \pD{} model representation.
-When the data has been binned the triangulation connectd centroids.
-The edges preserve the neighborhood information when the model is lifted into \pD{}.
+When the data has been binned the triangulation connects centroids.
+The edges preserve the neighborhood information from the \kD{} representation when the model is lifted into \pD{}.
 
@@
 In order to generate a smooth surface in \gD{}, these long line segments should be removed when tuning the model fit.
+<!-- I don't think this is a thing anymore -->
 
@@
 The last step is to lift the \gD{} model into \pD{} by computing \pD{} vectors that represent bin centroids.
-We use the \pD{} mean of the points in $H_h$ to map the centroid $C_{h}^{(2)} = (c_{h1}, c_{h2})$ to a point in \pD{}.
+We use the \pD{} mean of the points in a given hexagon, $H_h$, denoted $C_{h}^{(p)}$, to map the centroid $C_{h}^{(2)} = (c_{h1}, c_{h2})$ to a point in \pD{}.
 Let the \pD{} mean be
@@
 $$C_{h}^{(p)} = \frac{1}{n_h}\sum_{i =1}^{n_h} x_i, h = {1, \dots, b; n_h > 0}.$$
-Furthermore, line segments that exist in the \gD{} model generate line segments in \pD{} by connecting the \pD{} means of the corresponding \gD{} bin centroids.
-If additional long edges need to be removed, compute the edges in \pD{} and pruned any detected long edges to improve the accuracy.
+
+Line segments that exist in the \gD{} model generate line segments in \pD{} by connecting the \pD{} means of the corresponding \gD{} bin centroids.
+If additional long edges need to be removed, compute the edges in \pD{} and prune any detected long edges to improve the accuracy.
 Once pruned, re-plot the \gD{} view to ensure it accurately captures the data.
+<!-- Do we still do this after the new binning system?
+-->
 
@@
 
+<!-- In Figure 4, is the "sparse space" in the middle of each cluster (i.e.
+the holes) or is it the gap between clusters?
+If it's between clusters, I'd argue tSNE is making this gap smaller than in reality, which I don't think the caption currently conveys.
+-->
+
 ```{r}
@@
 
-where $n$ is the number of observations, $b$ is the number of bins, $n_h$ is the number of observations in $h^{th}$ bin, $p$ is the number of variables, $\mathbfit{x}_{hij}$ is the $j^{th}$ dimensional data of $i^{th}$ observation in $h^{th}$ hexagon.
+where $n$ is the number of observations, $b$ is the number of bins, $n_h$ is the number of observations in $h^{th}$ bin, $p$ is the number of variables and $\mathbfit{x}_{hij}$ is the $j^{th}$ dimensional data of $i^{th}$ observation in $h^{th}$ hexagon.
 We can consider $e_{hj} = \sqrt{\sum_{j = 1}^{p} (\mathbfit{x}_{hij} - C^{(p)}_{hj})^2}$ to be the residual for each observation.
@@
 
+<!-- Sparse end with dense corner to me is very hard to understand - sparse and dense are pretty much antonyms -->
 ```{r}
@@
 
-A new benefit of this fitted model is that it allows us to now predict a new observation's value in the NLDR, for any method.
-The steps are to determine the closest bin centroid in \pD{}, $C^{(p)}_{h}$ and predict it to be the centroid of this bin in \gD{}, $C^{(2)}_{h}$.
-This can be written as, let $z(i) = \arg\min_{j = 1, \dots, b} \sqrt{\sum_{v=1}^{p}(x_{iv} - C^{(p)}_{jv})^2}$, then the new observation $i$ falls in the hexagon, $H_h = \{i| z(i) = h\}$ and the corresponding \kD{} bin centroids, $C_{h}^{(2)} = (c_{h1}, c_{h2})$.
+<!-- Does this really need an entire subsection? It's only 3 sentences long. -->
+
+In this paradigm of model fitting, it is now possible to predict a new observation's value in the NLDR representation, independent of the choice of NLDR method.
+To do this, we first need to determine the closest bin centroid in \pD{}, $C^{(p)}_{h}$, to the new observation, which is,
+
+$$z = \arg\min_{j = 1, \dots, b} \sqrt{\sum_{v=1}^{p}(x_{iv} - C^{(p)}_{jv})^2}.$$
+
+Then, the predicted value is is the centroid of this bin in \gD{}, $C^{(2)}_{z} = (c_{z1}, c_{z2})$.
 
@@
 The values $(s_1, \ s_2)$ define the position of the centroid of the bottom left hexagon.
-By default, this is at $s_1 = \text{-}q, s_2 = \text{-}qr_2$, where $q$ is the buffer bound the data.
-The choice of these values can have some effect on the distribution of bin counts.
-@fig-bins-two-curvy illustrates this.
-The distribution of bin counts for $s_1$ varying between $\text{-}0.1\,\text{--}\,0.0$ is shown.
+By default, this is at $s_1 = -q, s_2 = -qr_2$, where $q$ is the buffer bound the data.
+The choice of these values can have some effect on the distribution of bin counts which is seen in @fig-bins-two-curvy.
+<!-- Figure 6 doesn't seem to have any varying s_1 unless I'm missing something --> The distribution of bin counts for $s_1$ varying between $\text{-}0.1\,\text{--}\,0.0$ is shown.
 Generally, a more uniform distribution among these possibilities would indicate that the bins are reliably capturing the underlying distribution of observations.
@@
 The default number of bins $b=b_1\times b_2$ is computed based on the sample size, by setting $b_1=n^{1/3}$, consistent with the Diaconis-Freedman rule [@freedman1981].
-The value of $b_2$ is determined analytically by $b_1, q, r_2$.
+The value of $b_2$ is determined analytically by $b_1, q, r_2$ <!-- What is the formula for this?
+-->.
 Values of $b_1$ between $2$ and $b_1 = \sqrt{n/r_2}$ are allowed.
@@
 #### Measurement of capturing the data shape in 2-D
+
+<!-- I'll be honest I've no idea what this section is -->
 The area of a hexagon is defined as $A = 3\sqrt{3}l^2/2$ where $l$ is the side length of the hexagon.
@@
 This more accurately reflects the hexagon grid relative the MSE than the full number of bins in the grid.
-It may also be beneficial to remove low count bins also, in the situation where data is clustered or stringy, where the observed data is sparse.
+It may be beneficial to remove low count bins also, such as when the data is clustered or stringy, or where the observed data is sparse.
 In order to decide if this is necessary, you would examine the distribution of bin counts, or the density which puts the counts on a standard scale.
@@
 
-The benchmark value for removing low-density hexagons ranges between $0$ and $1$.
+The benchmark value for removing low-density hexagons ranges between $0$ and $1$, where $0$ corresponds to no bins being removed, and $1$ to 100% of bins being removed.
 When analyzing how these benchmark values influence model performance, it's essential to observe the change in MSE as the benchmark value increases (@fig-param-two-curvy (c)).
@@
 
-To understand how well the model fits the points whether it fits well, works better in some positions, or fails to match the overall pattern.
-It is important to link and brush the points with high model error in the \pD{} error plot, \gD{} layout, and the generated model overlay on the data in \pD{}.
+To understand how well the model fits the points whether it fits well, works better in some positions, or fails to match the overall pattern, it is important to link and brush the points with high model error in the \pD{} error plot, \gD{} layout, and the generated model overlay on the data in \pD{}.
 
@@
 A particular pattern that we commonly see is that analysts tend to pick layouts with clusters that have big separations between them.
-When you examine their data in a tour, it is almost always that we see there are no big separations, and actually the suggested clusters are not even present.
+When you examine their data in a tour, it is almost always that we see there are no big separations, and actually often the suggested clusters are not even present.
 While we don't expect that analysts include animated gifs of tours in their papers, we should expect that any \gD{} representation adequately indicates the clustering that is present, and honestly show lack of separation or lack of clustering when it doesn't exist.
@@
 
-To compare and assess a range of representations needs these items:
+To compare and assess a range of representations an analyst needs:
 
@@
 This single projection only shows this in three of the five clusters but if we examine a different projection the other clusters exhibit the pancake also.
-While we don't know what exactly causes this it is likely due to some ordering of points in the \gD{} PaCMAP layout that induces the flat model.
+While we don't know what exactly causes this, it is likely due to some ordering of points in the \gD{} PaCMAP layout that induces the flat model.
 One could imagine that if the method used principal components on all the data, that it might induce some ordering that would produce the flat model.
@@
 
-@fig-one-dens_clust-error (a2, b2) shows \gD{} layout for tSNE created using the default (hyper-)~parameters.
+@fig-one-dens_clust-error (a2, b2) shows a \gD{} layout for tSNE created using the default (hyper-)~parameters.
 One would expect to see a rectangular shape if the curved sheet is flattened, but the layout is triangular.
@@
 @fig-NLDR-variety shows this data using a variety of methods, and different (hyper-)~parameters.
-You can see that the result is wildly different depending on the choices.
+<!-- I think some of these panels aren't umap - if so saying "umap was performed on the first nine principal components" is not necessarily correct.
+I think its what the publication did but not what is in our figure --> You can see that the result is wildly different depending on the choices.
 Layout a is a reproduction of the layout published in @chen2024.
@@
 
-The layout plot are colored to match the lines in the RMSE vs binwidth ($a_1$) plot.
+The layout plots are colored to match the lines in the RMSE vs binwidth ($a_1$) plot.
 Layouts a, b and f fit the data better than c, d, e, and layout f appears to be the bets fit.
@@
 
-The next step is to investigate the \gD{} to understand what information is learned from this representation.
+The next step is to investigate the \gD{} layout to understand what information is learned from this representation.
 @fig-model-error-mnist summarizes this investigation.
-Plot a shows the layout with points colored by their residual value - darker color indicates larger residual and poor fit.
+<!-- I understand the convenience of b-f all lined up but I did go looking for a red box.
+With enough reading I know why it's not there but I wonder if this can be improved --> Plot a shows the layout with points colored by their residual value - darker color indicates larger residual and poor fit.
 The plots b, c, d, e show samples of hand-written digits taken from inside the colored boxes.
@@
 We have developed an approach to help assess and compare NLDR layouts, generated by different methods and (hyper-)~parameter choice(s).
-It depends on conceptualizing the \gD{} layout as a model, creating a wireframe representation of it that can be lifted into \pD{}.
+It depends on conceptualizing the \gD{} layout as a model, allowing for the creation of a wireframe representation of the model that can be lifted into \pD{}.
 The fit is assessed by viewing the model in the data space, computing residuals and RMSE.

### 2025-06-16 paper.qmd
Added section descriptions

@@
 @sec-method contains the details of the new methodology, including simulated data examples.
+In @sec-bestfit, we describe how to assess the best fit and identify the most accurate \gD{} layout based on the proposed model diagnostics.
+Curiosities and unexpected patterns discovered in NLDR results by examining the model in the data space are discussed in @sec-curiosities.
 Two applications illustrating the use of the new methodology for bioinformatics and image classification are in @sec-applications.
 Limitations and future directions are provided in @sec-discussion.
-
-<!-- Where are sections 4 & 5 in this description? -->
 
@@
 
-## Assessing the best fit, and hence most accurate \gD{} layout
+## Assessing the best fit, and hence most accurate \gD{} layout {#sec-bestfit}
 
@@
 
-## Curiosities about NLDR results discovered by examining the model in the data space
+## Curiosities about NLDR results discovered by examining the model in the data space {#sec-curiosities}
 

### 2025-06-16 paper.qmd
Improved some comments

@@
 
-When shapes are nonlinear in the NLDR layout, some edges could be long.
-It can also happen that distant centroids can be connected, particularly if clustering is present, which can result in long line segments.
-In order to generate a smooth surface in \gD{}, these long line segments should be removed when tuning the model fit.
-<!-- I don't think this is a thing anymore -->
-
 ### Rendering the model in \pD{}
@@
 
-#### Measurement of capturing the data shape in 2-D
+#### Measurement of capturing the data shape in \gD{}
 
@@
 In this data there are $2622$ single cells and $1000$ gene expressions (variables).
-Following their pre-processing, UMAP was performed on the first nine principal components.
+Following their pre-processing, different NLDR techniques were performed on the first nine principal components.
 @fig-NLDR-variety shows this data using a variety of methods, and different (hyper-)~parameters.
-<!-- I think some of these panels aren't umap - if so saying "umap was performed on the first nine principal components" is not necessarily correct.
-I think its what the publication did but not what is in our figure --> You can see that the result is wildly different depending on the choices.
+You can see that the result is wildly different depending on the choices.
 Layout a is a reproduction of the layout published in @chen2024.

### 2025-06-16 paper.qmd
updated with comments

@@
 
-Line segments that exist in the \gD{} model generate line segments in \pD{} by connecting the \pD{} means of the corresponding \gD{} bin centroids.
-If additional long edges need to be removed, compute the edges in \pD{} and prune any detected long edges to improve the accuracy.
-Once pruned, re-plot the \gD{} view to ensure it accurately captures the data.
-<!-- Do we still do this after the new binning system?
--->
-
 ```{r}

### 2025-06-16 paper.qmd
Updated the comments

@@
 This can be defined using the function $u: \mathbb{R}^{n\times 2} \rightarrow \mathbb{R}^{m\times 2}$, where
-$$u(i) = \arg\min_{j = 1, \dots, b} \sqrt{(y_{i1} - C^{(2)}_{j1})^2 + (y_{i2} - C^{(2)}_{j2})^2}$$, maps observation $i$ into $H_h = \{i| u(i) = h\}$.
+$$u(i) = \arg\min_{j = 1, \dots, b} \sqrt{(y_{i1} - C^{(2)}_{j1})^2 + (y_{i2} - C^{(2)}_{j2})^2},$$ maps observation $i$ into $H_h = \{i| u(i) = h\}$.
 
@@
 Color indicates error ($e_{hj}$), dark color indicating high error and light indicates low error.
-Most large errors are distributed near the sparse end of the nonlinear cluster with dense corner."
+Most large errors are distributed near the sparse end of the nonlinear cluster and most small errors are distributed near the dense corners."
 #| out-width: 80%
@@
 The choice of these values can have some effect on the distribution of bin counts which is seen in @fig-bins-two-curvy.
-<!-- Figure 6 doesn't seem to have any varying s_1 unless I'm missing something --> The distribution of bin counts for $s_1$ varying between $\text{-}0.1\,\text{--}\,0.0$ is shown.
+The distribution of bin counts for $s_1$ varying between $\text{-}0.1\,\text{--}\,0.0$.
 Generally, a more uniform distribution among these possibilities would indicate that the bins are reliably capturing the underlying distribution of observations.

### 2025-06-16 paper.qmd
Changed MSE to RMSE

@@
 
-## Find the minimum MSE when have duplicate a1
+## Find the minimum RMSE when have duplicate a1
 error_two_curvy_umap <- error_two_curvy_umap |>
@@
 
-Default values are provided for each of these, but it is expected that the user will examine the MSE for a range of choices.
-Choosing these parameters according to MSE can be automated but it is recommended that the user examine the resulting model representation by overlaying it on the data in \pD{}.
+Default values are provided for each of these, but it is expected that the user will examine the RMSE for a range of choices.
+Choosing these parameters according to RMSE can be automated but it is recommended that the user examine the resulting model representation by overlaying it on the data in \pD{}.
 The next few subsections describe the calculation of default values, and the effect that different choices have on the model fit.
@@
 At the smallest bin size, the data structure is discontinuous, suggesting that there are too many bins.
-Using the MSE of the model fit in $7\\text{-}D$ helps decide on a useful choice of number of bins."
+Using the RMSE of the model fit in $7\\text{-}D$ helps decide on a useful choice of number of bins."
 #| fig-width: 9
@@
 Values of $b_1$ between $2$ and $b_1 = \sqrt{n/r_2}$ are allowed.
-@fig-param-two-curvy (a) shows the effect of different choices of $b_1$ on the MSE of the fitted model.
+@fig-param-two-curvy (a) shows the effect of different choices of $b_1$ on the RMSE of the fitted model.
 
@@
 By default, when assessing the choice of $b_1$, the total number of bins is measured by the number of **non-empty** bins.
-This more accurately reflects the hexagon grid relative the MSE than the full number of bins in the grid.
+This more accurately reflects the hexagon grid relative the RMSE than the full number of bins in the grid.
 It may be beneficial to remove low count bins also, such as when the data is clustered or stringy, or where the observed data is sparse.
@@
 The benchmark value for removing low-density hexagons ranges between $0$ and $1$, where $0$ corresponds to no bins being removed, and $1$ to 100% of bins being removed.
-When analyzing how these benchmark values influence model performance, it's essential to observe the change in MSE as the benchmark value increases (@fig-param-two-curvy (c)).
-The MSE shows a gradual decrease as the benchmark value goes from $1$ to $0$.
+When analyzing how these benchmark values influence model performance, it's essential to observe the change in RMSE as the benchmark value increases (@fig-param-two-curvy (c)).
+The RMSE shows a gradual decrease as the benchmark value goes from $1$ to $0$.
 Evaluating this rate of increase is important.
@@
 
-@fig-toy-mse illustrates the approach to compare the fits for different representations and assess the strength of any fit.
+@fig-toy-rmse illustrates the approach to compare the fits for different representations and assess the strength of any fit.
 What does it mean to be a best fit for this problem?
@@
 Layout a is the best choice."
-#| label: fig-toy-mse
+#| label: fig-toy-rmse
 #| fig-pos: H
@@
 
-@fig-pbmc-mse shows RMSE across a range of binwidths ($a_1$) for each of the layouts in @fig-NLDR-variety.
+@fig-pbmc-rmse shows RMSE across a range of binwidths ($a_1$) for each of the layouts in @fig-NLDR-variety.
 The layouts were made using tSNE, UMAP, PHATE, PaCMAP, and TriMAP with various (hyper-)~parameter settings.
@@
 Layout e is the best choice."
-#| label: fig-pbmc-mse
+#| label: fig-pbmc-rmse
 #| fig-pos: H
@@
 #| echo: false
-#| fig-cap: "Compare the published \\gD{} layout (@fig-pbmc-mse a) and the \\gD{} layout selected (@fig-pbmc-mse e) by MSE plot (@fig-pbmc-mse) from the UMAP and tSNE with default (hyper-)~parameters.
+#| fig-cap: "Compare the published \\gD{} layout (@fig-pbmc-rmse a) and the \\gD{} layout selected (@fig-pbmc-rmse e) by RMSE plot (@fig-pbmc-rmse) from the UMAP and tSNE with default (hyper-)~parameters.
 The PBMC3k data ($n =  2622$) has three close clusters in $9\\text{-}D$.
-Two \\gD{} projection from a tour on $9\\text{-}D$ of the model fit with @fig-pbmc-mse a ($a_1 = 0.06$, $b = 616/86 (22, 28)$) shows three-well separated clusters with big separations.
-On the other hand, the model fit with @fig-pbmc-mse e ($a_1 = 0.06$, $b = 704/227 (22, 32)$) shows three-well separated clusters with small separations.
-Therefore, @fig-pbmc-mse e is more reasonable than @fig-pbmc-mse a.
+Two \\gD{} projection from a tour on $9\\text{-}D$ of the model fit with @fig-pbmc-rmse a ($a_1 = 0.06$, $b = 616/86 (22, 28)$) shows three-well separated clusters with big separations.
+On the other hand, the model fit with @fig-pbmc-rmse e ($a_1 = 0.06$, $b = 704/227 (22, 32)$) shows three-well separated clusters with small separations.
+Therefore, @fig-pbmc-rmse e is more reasonable than @fig-pbmc-rmse a.
 Also, the fitted models in $9\\text{-}D$ help to see some unobserved patterns of the data: (i) dense points, and (ii) nonlinear clusters.
@@
 Layout f is the best choice."
-#| label: fig-mnist-mse
+#| label: fig-mnist-rmse
 #| fig-pos: H
@@
 
-@fig-mnist-mse compares the fit of six layouts computed using UMAP (b), PHATE (c), TriMAP (d), PaCMAP (e) with default (hyper-)~parameter setting and two tSNE runs, one with default (hyper-)~parameter setting (a) and the other changing perplexity to $89$ (f).
+@fig-mnist-rmse compares the fit of six layouts computed using UMAP (b), PHATE (c), TriMAP (d), PaCMAP (e) with default (hyper-)~parameter setting and two tSNE runs, one with default (hyper-)~parameter setting (a) and the other changing perplexity to $89$ (f).
 The layouts are reasonably similar in that they all have the observations in a single blob.
@@
 @fig-model-error-mnist summarizes this investigation.
-<!-- I understand the convenience of b-f all lined up but I did go looking for a red box.
-With enough reading I know why it's not there but I wonder if this can be improved --> Plot a shows the layout with points colored by their residual value - darker color indicates larger residual and poor fit.
+Plot a shows the layout with points colored by their residual value - darker color indicates larger residual and poor fit.
 The plots b, c, d, e show samples of hand-written digits taken from inside the colored boxes.

### 2025-06-16 paper.qmd
Added bin2 equation with cross reference
Added linked plots video for C-shaped data

@@
 b_2 = \Big\lceil1 +\frac{2[r_2 + q(1 + r_2)](b_1 - 1)}{\sqrt{3}(1 + 2q)}\Big\rceil.
-$$
+$${#eq-bin2}
 
@@
 The default number of bins $b=b_1\times b_2$ is computed based on the sample size, by setting $b_1=n^{1/3}$, consistent with the Diaconis-Freedman rule [@freedman1981].
-The value of $b_2$ is determined analytically by $b_1, q, r_2$ <!-- What is the formula for this?
--->.
+The value of $b_2$ is determined analytically by $b_1, q, r_2$ (@eq-bin2).
 Values of $b_1$ between $2$ and $b_1 = \sqrt{n/r_2}$ are allowed.
@@
 ## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
-df_b_mnist <- df_b_mnist[match(hexID_mnist, df_b_mnist$hexID),] |>
+df_b_mnist <- df_b_mnist[match(hexID_mnist, df_b_mnist$hexID),]
+
+df_b_mnist_temp <- df_b_mnist
+
+df_b_mnist <- df_b_mnist |>
   dplyr::select(-hexID)
@@
   nldr_data = tsne_minst_scaled_temp,
-  model_highd = df_bin_mnist,
+  model_highd = df_b_mnist_temp,
   model_2d = df_bin_centroids_dens_clust)

### 2025-06-16 paper.qmd
Updated naming with convention

@@
 nldr_mnist1 <- tsne_mnist |>
-  ggplot(aes(x = tSNE1,
-             y = tSNE2))+
+  ggplot(aes(x = emb1,
+             y = emb2))+
   geom_point(alpha=0.1, size=1, colour='#984ea3') +
@@
 nldr_mnist6 <- tsne_mnist2 |>
-  ggplot(aes(x = tSNE1,
-             y = tSNE2))+
+  ggplot(aes(x = emb1,
+             y = emb2))+
   geom_point(alpha=0.1, size=1, colour='#999999') +
@@
 nldr_mnist2 <- umap_mnist |>
-  ggplot(aes(x = UMAP1,
-             y = UMAP2)) +
+  ggplot(aes(x = emb1,
+             y = emb2)) +
   geom_point(alpha=0.1, size=1, colour='#a65628') +
@@
 nldr_mnist3 <- phate_mnist |>
-  ggplot(aes(x = PHATE1,
-             y = PHATE2))+
+  ggplot(aes(x = emb1,
+             y = emb2))+
   geom_point(alpha=0.1, size=1, colour='#377eb8') +
@@
 nldr_mnist4 <- trimap_mnist |>
-  ggplot(aes(x = TriMAP1,
-             y = TriMAP2))+
+  ggplot(aes(x = emb1,
+             y = emb2))+
   geom_point(alpha=0.1, size=1, colour='#ff7f00') +
@@
 nldr_mnist5 <- pacmap_mnist |>
-  ggplot(aes(x = PaCMAP1,
-             y = PaCMAP2))+
+  ggplot(aes(x = emb1,
+             y = emb2))+
   geom_point(alpha=0.1, size=1, colour='#e41a1c') +
@@
 
-tsne_minst_scaled_temp <- tsne_minst_scaled |>
-  rename(c("emb1" = "tSNE1",
-           "emb2" = "tSNE2"))
-
 tsne_minst_scaled <- tsne_minst_scaled |>
-  mutate(cluster = if_else((tSNE1 <= 1) & (tSNE1 >= 0.78) & (tSNE2 <= 0.62) & (tSNE2 >= 0.39), "small_clust", "big_clust"))
+  mutate(cluster = if_else((emb1 <= 1) & (emb1 >= 0.78) & (emb2 <= 0.62) & (emb2 >= 0.39), "small_clust", "big_clust"))
 
@@
              aes(
-               x = tSNE1,
-               y = tSNE2,
+               x = emb1,
+               y = emb2,
                colour = cluster
@@
 highd_error_2d_mnist <- error_df |>
-  ggplot(aes(x = tSNE1,
-             y = tSNE2,
+  ggplot(aes(x = emb1,
+             y = emb2,
              colour = sqrt_row_wise_tot_error)) +

### 2025-06-16 paper.qmd
Finalize the video table

@@
 And it keeps all of this supplementary material in one place.
+(Done)
 - Move the notation table to appendix (Done)
@@
 
-trimesh_removed_mnist <- ggplot() + 
-  geom_point(data = tsne_minst_scaled,
-             aes(
-               x = emb1,
-               y = emb2,
-               colour = cluster
-             ),
-             alpha=0.2)+
-    geom_segment(data = tr_from_to_df_mnist, 
-               aes(
-                 x = x_from, 
-                 y = y_from, 
-                 xend = x_to, 
-                 yend = y_to),
-               colour = "#000000") +
-  scale_color_manual(values=c("#999999",'#ff7f00')) +
-  interior_annotation("a", c(0.08, 0.93), cex = 2) +
-  theme(legend.position = "none")
-
 tsne_plot_mnist_clust <- ggplot() + 
@@
 
-mnist_proj_tsne_model1 <- plot_proj(
-  proj_obj = proj_obj1, 
-  point_param = c(0.5, 0.5), # size, alpha, color
-  line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
-  plot_limits = c(-1.1, 0.9), 
-  axis_text_size = 4, 
-  is_category = TRUE) +
-  interior_annotation(label = "b", cex = 2) +
-  scale_color_manual(values = c("#999999",'#ff7f00')) +
-  theme(
-    legend.position = "none"
-  )
-
 mnist_proj_tsne_model1_link <- plot_proj(
@@
 
-mnist_proj_tsne_model2 <- plot_proj(
-    proj_obj = proj_obj2,
-    point_param = c(0.5, 0.3), # size, alpha, color
-    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
-    plot_limits = c(-1, 1), 
-    axis_text_size = 3.5, 
-    is_category = TRUE) +
-  interior_annotation(label = "c", cex = 2) +
-  scale_color_manual(values = c("#999999",'#ff7f00')) +
-  theme(
-      legend.position = "none"
-  )
-
 mnist_proj_tsne_model2_link <- plot_proj(

### 2025-06-16 paper.qmd
Added MNISt code to a script

@@
 
-<!-- PaCMAP layout with author's suggested parameter choice-->
-```{r}
-#| label: read-mnist
-## Import data
-training_data_mnist <- read_rds("data/mnist/mnist_10_pcs_of_digit_1.rds")
-names(training_data_mnist) <- paste0("x", 1:NCOL(training_data_mnist))
-
-training_data_mnist <- training_data_mnist |>
-  mutate(ID = 1:NROW(training_data_mnist))
-
-data_mnist <- training_data_mnist |>
-  select(-ID) |>
-  mutate(type = "data")
-```
-
 <!--PaCMAP param: n_components=2, n_neighbors=10, init=random, MN_ratio=0.9, FP_ratio=2.0-->
@@
 
-num_bins_x_mnist <- 30
-
-algo_obj_mnist <- fit_highd_model(
-  highd_data = training_data_mnist, 
-  nldr_data = tsne_mnist2, 
-  bin1 = num_bins_x_mnist, 
-  q = 0.1, 
-  benchmark_highdens = 1)
-
-tsne_minst_scaled <- algo_obj_mnist$nldr_obj$scaled_nldr
-tr_from_to_df_mnist <- algo_obj_mnist$trimesh_data
-df_bin_centroids_mnist <- algo_obj_mnist$model_2d
-df_bin_mnist <- algo_obj_mnist$model_highd
-
 sc_ltr_pos_mnist <- c(0.96, 0.96)
 
-tsne_minst_scaled <- tsne_minst_scaled |>
-  mutate(cluster = if_else((emb1 <= 1) & (emb1 >= 0.78) & (emb2 <= 0.62) & (emb2 >= 0.39), "small_clust", "big_clust"))
+tsne_minst_scaled <- read_rds("data/mnist/mnist_tsne_minst_scaled.rds")
 
@@
 ## Compute error
-error_df <- augment(
-  highd_data = training_data_mnist,
-  model_2d = df_bin_centroids_mnist,
-  model_highd = df_bin_mnist)
-
-## To join embedding
-error_df <- error_df |>
-  bind_cols(tsne_minst_scaled |> 
-              select(-ID))
-
-## To transform the data
-## ggplot(data = error_df, aes(x=row_wise_abs_error)) + geom_density() ## right skewed
-
-error_df <- error_df |>
-  mutate(sqrt_row_wise_tot_error = sqrt(row_wise_total_error)) |>
-  mutate(sqrt_row_wise_tot_error = standardize(sqrt_row_wise_tot_error))
+error_df <- read_rds("data/mnist/mnist_error_df.rds")
 
@@
 
-<!--bin1 = 22, bin2 = 17, b = 374, non_empty = 140-->
-
-```{r}
-#| label: data-prep-mnist-model-proj
-
-hexID_mnist <- df_bin_centroids_mnist |>
-  #dplyr::filter(std_counts > 0) |>
-  dplyr::pull(hexID)
-
-df_b_mnist <- df_bin_mnist |>
-  dplyr::filter(hexID %in% hexID_mnist) |>
-  dplyr::mutate(type = "model") ## Data with summarized mean
-
-## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
-df_b_mnist <- df_b_mnist[match(hexID_mnist, df_b_mnist$hexID),]
-
-df_b_mnist_temp <- df_b_mnist
-
-df_b_mnist <- df_b_mnist |>
-  dplyr::select(-hexID)
-
-# Apply the scaling
-df_model_data_mnist <- bind_rows(data_mnist, df_b_mnist)
-scaled_mnist <- scale_data_manual(df_model_data_mnist, "type") |>
-  as_tibble()
-
-scaled_mnist_data <- scaled_mnist |>
-  filter(type == "data") |>
-  select(-type)
-
-scaled_mnist_data_model <- scaled_mnist |>
-  filter(type == "model") |>
-  select(-type)
-```
-
-```{r}
-#| label: langevitour-best-fit-mnist-proj-tsne
-#| eval: false
-
-data_mnist_cluster <- data_mnist |>
-  select(-type) |>
-  mutate(type = tsne_minst_scaled$cluster)
-
-df_model_data_mnist_n <- bind_rows(df_b_mnist, data_mnist_cluster)
-
-langevitour::langevitour(df_model_data_mnist_n[1:(length(df_model_data_mnist_n)-1)],
-                         lineFrom = tr_from_to_df_mnist$from,
-                         lineTo = tr_from_to_df_mnist$to,
-                         group = factor(df_model_data_mnist_n$type,
-                                        c("big_clust", "small_clust", "model")),
-                         levelColors = c("#999999",'#ff7f00', "#000000"))
-```
-
-```{r}
-#| label: link-plts-mnist
-#| eval: false
-
-df_exe_mnist <- comb_all_data_model(
-  highd_data = training_data_mnist,
-  nldr_data = tsne_minst_scaled_temp,
-  model_highd = df_b_mnist_temp,
-  model_2d = df_bin_centroids_dens_clust)
-
-show_link_plots(point_data = df_exe_mnist,
-                edge_data = tr_from_to_df_dens_clust)
-```
-
 ```{r}
@@
 
-projection <- cbind(
-  c(-0.04295,0.42361,-0.36702,0.13103,-0.19678,-0.29621,0.00471,0.09323,0.18822,-0.18484),
-  c(-0.36644,0.05058,0.18200,0.21860,0.13583,-0.03511,0.35339,0.15821,-0.26945,-0.27926))
-
-proj_obj1 <- get_projection(projection = projection, 
-               proj_scale = 2.5, 
-               highd_data = scaled_mnist_data, 
-               model_highd = scaled_mnist_data_model, 
-               trimesh_data = tr_from_to_df_mnist, 
-               axis_param = list(limits = 1, 
-                                 axis_scaled = 1.1, 
-                                 axis_pos_x = -0.92, 
-                                 axis_pos_y = -0.92, 
-                                 threshold = 0.065))
-
-proj_obj1[["cluster"]] <- tsne_minst_scaled$cluster
+proj_obj1 <- read_rds("data/mnist/mnist_proj_obj1.rds")
 
@@
 ## Second projection
-projection <- cbind(
-    c(-0.33305,0.20223,0.06303,0.11124,0.13299,-0.30093,0.34082,0.15581,-0.08864,0.33573),
-    c(0.53615,0.14969,-0.09393,-0.20053,0.07992,-0.34791,0.13057,-0.06814,-0.16737,0.03716))
-
-proj_obj2 <- get_projection(projection = projection, 
-               proj_scale = 2, 
-               highd_data = scaled_mnist_data, 
-               model_highd = scaled_mnist_data_model, 
-               trimesh_data = tr_from_to_df_mnist, 
-               axis_param = list(limits = 1, 
-                                 axis_scaled = 1.1, 
-                                 axis_pos_x = -0.83, 
-                                 axis_pos_y = -0.83, 
-                                 threshold = 0.065))
-
-proj_obj2[["cluster"]] <- tsne_minst_scaled$cluster
+proj_obj2 <- read_rds("data/mnist/mnist_proj_obj2.rds")
 
@@
 Two \\gD{} projections from a tour on $10\\text{-}D$ reveal that the closeness of the clusters in $10\\text{-}D$ and the twisted pattern of the model fit with tSNE ($a_1 = 0.041$, $b = 659/585 \\text{ } (30, \\text{ }48)$).
-Video of the langevitour animation is available at <https://youtu.be/sUcGd57Swdg>.
+The link to the video of the langevitour animation is available in @tbl-links.
 The brushing feature in the linked plots helps in visualizing the closeness of the small cluster to the big cluster.
-Videos of the linked plots are available at <>."
+The link to the video of the linked plots is available in @tbl-links."
 
@@
 ## Data with pixel values
-mnist_data <- read_rds("data/mnist/mnist_digit_1.rds")
-
-mnist_data <- mnist_data |>
-  mutate(instance = row_number()) |>
-  gather(pixel, value, -Label, -instance) |>
-  extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
-  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28)
-
-img_right_top <- c(868, 1465, 1787, 2006, 4314)
-
-pixels_gathered_within <-  mnist_data |>
-  filter(instance %in% img_right_top)
+
+img_right_top <- read_rds("data/mnist/mnist_img_right_top.rds")
 
 right_top_img <- plot_digit_img(
-  pixels_gathered_within, palette = "Blues 2", title_text = "b")
-
-img_middle <- c(130, 2091, 7795, 163, 77)
-
-pixels_gathered_within <-  mnist_data |>
-  filter(instance %in% img_middle)
+  img_right_top, palette = "Blues 2", title_text = "b")
+
+img_middle <- read_rds("data/mnist/mnist_img_middle.rds")
 
 middle_img <- plot_digit_img(
-  pixels_gathered_within, palette = "Greens 3", title_text = "c")
-
-img_right_bottom <- c(2035, 7708, 7710, 4134, 5006)
-
-pixels_gathered_within <-  mnist_data |>
-  filter(instance %in% img_right_bottom)
+  img_middle, palette = "Greens 3", title_text = "c")
+
+img_right_bottom <- read_rds("data/mnist/mnist_img_right_bottom.rds")
 
 right_bottom_img <- plot_digit_img(
-  pixels_gathered_within, palette = "Purples 2", title_text = "d")
-
-img_error_outside <- c(999, 3650, 4916, 2964, 5622)
-
-pixels_gathered_outside <-  mnist_data |>
-  filter(instance %in% img_error_outside)
+  img_right_bottom, palette = "Purples 2", title_text = "d")
+
+img_error_outside <- read_rds("data/mnist/mnist_img_error_outside.rds")
 
 imge_error_sample_outside <- plot_digit_img(
-  pixels_gathered_outside, palette = "Grays", title_text = "e")
+  img_error_outside, palette = "Grays", title_text = "e")
 
 ##### inside
-img_error_inside <- c(5173, 6981, 7228, 6673, 3179) #, 6673, 3179, 7363, 7650, 5241, 4636
-
-pixels_gathered_inside <-  mnist_data |>
-  filter(instance %in% img_error_inside)
+img_error_inside <- read_rds("data/mnist/mnist_img_error_inside.rds")
 
 imge_error_sample_inside <- plot_digit_img(
-  pixels_gathered_inside, palette = "Oranges", title_text = "f")
+  img_error_inside, palette = "Oranges", title_text = "f")
 

### 2025-06-16 paper.qmd
Moved PBMC3k codes to scripts

@@
 
-<!-- UMAP layout with author's suggested parameter choice-->
-```{r}
-#| label: published-pbmc
-## Import data
-training_data_pbmc <- read_rds("data/pbmc3k/pbmc_pca_50.rds")
-names(training_data_pbmc) <- paste0("x", 1:50)
-
-training_data_pbmc <- training_data_pbmc[, 1:9] |>
-  mutate(ID = 1:NROW(training_data_pbmc))
-
-umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_30_min_dist_0.3.rds")
-pbmc_scaled_obj <- gen_scaled_data(
-  nldr_data = umap_pbmc)
-umap_pbmc_scaled <- pbmc_scaled_obj$scaled_nldr
-```
-
 @fig-pbmc-rmse shows RMSE across a range of binwidths ($a_1$) for each of the layouts in @fig-NLDR-variety.
@@
 #| label: hexbin-pbmc
-## Compute hexbin parameters
-num_bins_x_pbmc <- 22
-
-algo_obj_pbmc <- fit_highd_model(
-  highd_data = training_data_pbmc, 
-  nldr_data = umap_pbmc, 
-  bin1 = num_bins_x_pbmc, 
-  q = 0.1, 
-  benchmark_highdens = 5)
-
-umap_pbmc_scaled <- algo_obj_pbmc$nldr_obj$scaled_nldr
-tr_from_to_df_pbmc <- algo_obj_pbmc$trimesh_data
-df_bin_centroids_pbmc <- algo_obj_pbmc$model_2d
-df_bin_pbmc <- algo_obj_pbmc$model_highd
-
-umap_pbmc_scaled_with_cluster <- umap_pbmc_scaled |>
-  mutate(cluster = if_else(UMAP1 <= 0.25, "cluster1", if_else(UMAP1 >= 0.625, "cluster2", "cluster3"))) |>
-  mutate(cluster = factor(cluster, levels = c("cluster1", "cluster2", "cluster3")))
+#| 
+umap_pbmc_scaled_with_cluster <- read_rds("data/pbmc3k/umap_pbmc_scaled_with_cluster.rds")
+
+umap_tr_from_to_df_pbmc <- read_rds("data/pbmc3k/umap_tr_from_to_df_pbmc.rds")
 
@@
   )  +
-  geom_segment(data = tr_from_to_df_pbmc, 
+  geom_segment(data = umap_tr_from_to_df_pbmc, 
                aes(
@@
 ```{r}
-#| label: prep-pbmc-umap-model-proj
-
-data_pbmc <- training_data_pbmc |> 
-  select(-ID) |>
-  mutate(type = "data")
-
-# df_bin_centroids_pbmc <- df_bin_centroids_pbmc |>
-#   dplyr::filter(std_counts > 0)
-
-df_b_pbmc <- df_bin_pbmc |>
-  dplyr::filter(hexID %in% df_bin_centroids_pbmc$hexID) |>
-  dplyr::mutate(type = "model") ## Data with summarized mean
-
-## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
-df_b_pbmc <- df_b_pbmc[match(df_bin_centroids_pbmc$hexID, df_b_pbmc$hexID),] |>
-  dplyr::select(-hexID)
-
-# Apply the scaling
-df_model_data_pbmc <- bind_rows(data_pbmc, df_b_pbmc)
-scaled_pbmc <- scale_data_manual(df_model_data_pbmc, "type") |>
-  as_tibble()
-
-scaled_pbmc_data <- scaled_pbmc |>
-  filter(type == "data") |>
-  select(-type)
-
-scaled_pbmc_data_model <- scaled_pbmc |>
-  filter(type == "model") |>
-  select(-type)
-```
-
-```{r}
-#| label: langevitour-pbmc-umap-author-proj
-#| eval: false
-
-data_pbmc_n <- data_pbmc |>
-  select(-type) |>
-  mutate(type = as.character(umap_pbmc_scaled_with_cluster$cluster))
-
-df_model_data_pbmc_n <- bind_rows(df_b_pbmc, data_pbmc_n)
-
-langevitour::langevitour(df_model_data_pbmc_n[1:(length(df_model_data_pbmc_n)-1)],
-                         lineFrom = tr_from_to_df_pbmc$from,
-                         lineTo = tr_from_to_df_pbmc$to,
-                         group = factor(df_model_data_pbmc_n$type,
-                                        c("cluster1", "cluster2", "cluster3", "model")),
-                         levelColors = c("#8dd3c7", "#fdb462", "#bebada", "#000000"))
-```
-
-```{r}
 #| label: pbmc-umap-model-proj
@@
 ## First projection
-projection <- cbind(
-    c(-0.4545,-0.5533,-0.1235,0.3984,-0.2065,-0.3586,-0.0088,0.3572,-0.1368),
-    c(0.4576,-0.3157,0.4708,0.1410,-0.4479,-0.1685,0.3668,-0.2692,0.1332))
-
-proj_obj1 <- get_projection(projection = projection, 
-               proj_scale = 1.23, 
-               highd_data = scaled_pbmc_data, 
-               model_highd = scaled_pbmc_data_model, 
-               trimesh_data = tr_from_to_df_pbmc, 
-               axis_param = list(limits = 1,
-                                 axis_scaled = 1, 
-                                 axis_pos_x = -0.48, 
-                                 axis_pos_y = -0.48, 
-                                 threshold = 0.082))
-
-proj_obj1[["cluster"]] <- as.character(umap_pbmc_scaled_with_cluster$cluster)
+proj_obj1 <- read_rds("data/pbmc3k/pbmc_umap_proj_obj1.rds")
 
@@
 ## Second projection
-projection <- cbind(
-    c(0.0734,0.6138,-0.4671,0.2589,0.1189,0.0778,-0.0819,0.0308,-0.5561),
-    c(0.6625,0.0034,0.2412,0.3360,0.2463,0.1070,-0.2903,-0.4705,0.1293))
-
-proj_obj2 <- get_projection(projection = projection, 
-               proj_scale = 1.25, 
-               highd_data = scaled_pbmc_data, 
-               model_highd = scaled_pbmc_data_model, 
-               trimesh_data = tr_from_to_df_pbmc, 
-               axis_param = list(limits = 1.35, 
-                                 axis_scaled = 1, 
-                                 axis_pos_x = -0.75, 
-                                 axis_pos_y = -0.75, 
-                                 threshold = 0.12))
-
-proj_obj2[["cluster"]] <- as.character(umap_pbmc_scaled_with_cluster$cluster)
+proj_obj2 <- read_rds("data/pbmc3k/pbmc_umap_proj_obj2.rds")
 
@@
 <!--best choice-->
-
-```{r}
-#| label: tsne-pbmc-read-best
-tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_30.rds")
-```
-
 <!--Fit the best model and compute error-->
@@
 #| label: tsne-num-bins-pbmc
-
-## Compute hexbin parameters
-num_bins_x_tsne_pbmc <- 22
-
-algo_obj_pbmc <- fit_highd_model(
-  highd_data = training_data_pbmc, 
-  nldr_data = tsne_pbmc, 
-  bin1 = num_bins_x_tsne_pbmc, 
-  q = 0.1, 
-  benchmark_highdens = 1)
-
-tsne_pbmc_scaled_best <- algo_obj_pbmc$nldr_obj$scaled_nldr
-tr_from_to_df_pbmc <- algo_obj_pbmc$trimesh_data
-df_bin_centroids_pbmc <- algo_obj_pbmc$model_2d
-df_bin_pbmc <- algo_obj_pbmc$model_highd
-
-tsne_pbmc_scaled_best_with_cluster <- tsne_pbmc_scaled_best |>
-  mutate(cluster = umap_pbmc_scaled_with_cluster$cluster)
+tsne_pbmc_scaled_best_with_cluster <- read_rds("data/pbmc3k/tsne_pbmc_scaled_best_with_cluster.rds")
+
+tsne_tr_from_to_df_pbmc <- read_rds("data/pbmc3k/tsne_tr_from_to_df_pbmc.rds")
 
@@
   ) +
-  geom_segment(data = tr_from_to_df_pbmc, 
+  geom_segment(data = tsne_tr_from_to_df_pbmc, 
                aes(
@@
 ```{r}
-#| label: prep-pbmc-tsne-model-proj-best
-
-df_b_pbmc <- df_bin_pbmc |>
-  dplyr::filter(hexID %in% df_bin_centroids_pbmc$hexID) |>
-  dplyr::mutate(type = "model") ## Data with summarized mean
-
-## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
-df_b_pbmc <- df_b_pbmc[match(df_bin_centroids_pbmc$hexID, df_b_pbmc$hexID),] |>
-  dplyr::select(-hexID)
-
-# Apply the scaling
-df_model_data_pbmc <- bind_rows(data_pbmc, df_b_pbmc)
-scaled_pbmc <- scale_data_manual(df_model_data_pbmc, "type") |>
-  as_tibble()
-
-scaled_pbmc_data <- scaled_pbmc |>
-  filter(type == "data") |>
-  select(-type)
-
-scaled_pbmc_data_model <- scaled_pbmc |>
-  filter(type == "model") |>
-  select(-type)
-```
-
-```{r}
-#| label: langevitour-pbmc-tsne-best-proj
-#| eval: false
-
-df_model_data_pbmc_n <- bind_rows(df_b_pbmc, data_pbmc_n)
-
-langevitour::langevitour(df_model_data_pbmc_n[1:(length(df_model_data_pbmc_n)-1)],
-                         lineFrom = tr_from_to_df_pbmc$from,
-                         lineTo = tr_from_to_df_pbmc$to,
-                         group = factor(df_model_data_pbmc_n$type,
-                                        c("cluster1", "cluster2", "cluster3", "model")),
-                         levelColors = c("#8dd3c7", "#fdb462", "#bebada", "#000000"))
-```
-
-```{r}
 #| label: pbmc-tsne-model-proj-best
@@
 ## First projection
-projection <- cbind(
-  c(-0.4545,-0.5533,-0.1235,0.3984,-0.2065,-0.3586,-0.0088,0.3572,-0.1368),
-  c(0.4576,-0.3157,0.4708,0.1410,-0.4479,-0.1685,0.3668,-0.2692,0.1332))
-
-proj_obj1 <- get_projection(projection = projection, 
-                            proj_scale = 1.23, 
-                            highd_data = scaled_pbmc_data, 
-                            model_highd = scaled_pbmc_data_model, 
-                            trimesh_data = tr_from_to_df_pbmc, 
-                            axis_param = list(limits = 1, 
-                                              axis_scaled = 1,
-                                              axis_pos_x = -0.48, 
-                                              axis_pos_y = -0.48, 
-                                              threshold = 0.082))
-
-proj_obj1[["cluster"]] <- as.character(umap_pbmc_scaled_with_cluster$cluster)
+proj_obj1 <- read_rds("data/pbmc3k/pbmc_tsne_proj_obj1.rds")
 
@@
 ## Second projection
-projection <- cbind(
-  c(0.0734,0.6138,-0.4671,0.2589,0.1189,0.0778,-0.0819,0.0308,-0.5561),
-  c(0.6625,0.0034,0.2412,0.3360,0.2463,0.1070,-0.2903,-0.4705,0.1293))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 1.25, 
-                            highd_data = scaled_pbmc_data, 
-                            model_highd = scaled_pbmc_data_model, 
-                            trimesh_data = tr_from_to_df_pbmc, 
-                            axis_param = list(limits = 1.35, 
-                                              axis_scaled = 1,
-                                              axis_pos_x = -0.75, 
-                                              axis_pos_y = -0.75, 
-                                              threshold = 0.12))
-
-proj_obj2[["cluster"]] <- as.character(umap_pbmc_scaled_with_cluster$cluster)
+proj_obj2 <- read_rds("data/pbmc3k/pbmc_tsne_proj_obj2.rds")
 
@@
 Also, the fitted models in $9\\text{-}D$ help to see some unobserved patterns of the data: (i) dense points, and (ii) nonlinear clusters.
-Videos of the langevitour animations are available at <https://youtu.be/3VfK3M2gnZM> and <https://youtu.be/Es84bwQcndU> respectively."
+The links to the videos of langevitour animations are available in @tbl-links."
 #| label: fig-model-pbmc-author-proj

### 2025-06-16 paper.qmd
Moved C-shaped data code to a script

@@
 ```{r}
-#| label: model-one-dens-clust
-
-one_c_shaped_data <- read_rds(here::here("data/one_c_shaped_dens_structure/one_c_shaped_dens_data.rds"))
-one_c_shaped_data <- one_c_shaped_data |>
-  mutate(ID = row_number())
-
-tsne_one_c_shaped <- read_rds(file = "data/one_c_shaped_dens_structure/one_c_shaped_dens_structure_tsne_perplexity_52.rds") |>
-  mutate(ID = row_number())
-
-num_bins_x_one_c_shaped <- 20
-
-algo_obj_dens_clust <- fit_highd_model(
-  highd_data = one_c_shaped_data, 
-  nldr_data = tsne_one_c_shaped, 
-  bin1 = num_bins_x_one_c_shaped, 
-  q = 0.1, 
-  benchmark_highdens = 1)
-
-tsne_one_c_shaped_scaled <- algo_obj_dens_clust$nldr_obj$scaled_nldr
-tr_from_to_df_dens_clust <- algo_obj_dens_clust$trimesh_data
-df_bin_centroids_dens_clust <- algo_obj_dens_clust$model_2d
-df_bin_dens_clust <- algo_obj_dens_clust$model_highd
-```
-
-```{r}
 #| label: error-model-one-dens-clust
-## Compute error
-error_df_one_curvy_abs <- augment(
-  highd_data = one_c_shaped_data,
-  model_2d = df_bin_centroids_dens_clust,
-  model_highd = df_bin_dens_clust)
-
-df_bin_dens_clust_temp <- df_bin_dens_clust
-
-error_df_one_curvy_abs_temp <- error_df_one_curvy_abs
-
-error_df_one_curvy_abs <- error_df_one_curvy_abs |>
-  bind_cols(tsne_one_c_shaped_scaled |>
-              select(-ID))
-
-error_df_one_curvy_abs <- error_df_one_curvy_abs |>
-  mutate(sqrt_row_wise_total_error = sqrt(row_wise_total_error))
-
-# Compute density
-density_data <- density(error_df_one_curvy_abs$sqrt_row_wise_total_error)
-density_df <- data.frame(x = density_data$x, y = density_data$y)
-
-# Add density values to the original dataset
-error_df_one_curvy_abs <- error_df_one_curvy_abs %>%
-  mutate(density = approx(density_df$x, density_df$y, xout = sqrt_row_wise_total_error)$y)
-
-error_df_one_curvy_abs <- error_df_one_curvy_abs |>
-  mutate(error_cat_n = if_else(emb1 <= 0.25 & emb2 <= 0.25, "selected", "deselected")) |> ## high_error points
-  mutate(error_cat_n2 = if_else(emb2 >= 1.2, "selected", "deselected")) ## corner points
-
-error_df_one_curvy_abs_selected11 <- error_df_one_curvy_abs |>
-  dplyr::filter(error_cat_n == "selected")
-
-error_df_one_curvy_abs_deselected11 <- error_df_one_curvy_abs |>
-  dplyr::filter(error_cat_n == "deselected")
+
+error_df_one_curvy_abs_selected11 <- read_rds("data/one_c_shaped_dens_structure/error_df_one_curvy_abs_selected11.rds")
+
+error_df_one_curvy_abs_deselected11 <- read_rds("data/one_c_shaped_dens_structure/error_df_one_curvy_abs_deselected11.rds")
 
@@
 
-error_df_one_curvy_abs_selected12 <- error_df_one_curvy_abs |>
-  dplyr::filter(error_cat_n2 == "selected")
-
-error_df_one_curvy_abs_deselected12 <- error_df_one_curvy_abs |>
-  dplyr::filter(error_cat_n2 == "deselected")
+error_df_one_curvy_abs_selected12 <- read_rds("data/one_c_shaped_dens_structure/error_df_one_curvy_abs_selected12.rds")
+
+error_df_one_curvy_abs_deselected12 <- read_rds("data/one_c_shaped_dens_structure/error_df_one_curvy_abs_deselected12.rds")
 
@@
 ```{r}
-#| label: prep-proj-one-dens-clust
-
-# Apply the scaling
-
-data_c_shaped <- one_c_shaped_data |>
-  select(-ID) |>
-  mutate(type = "data")
-
-df_bin_dens_clust <- df_bin_dens_clust |>
-  select(-hexID) |>
-  mutate(type = "model")
-
-df_model_data <- bind_rows(data_c_shaped, df_bin_dens_clust)
-scaled_c_shaped <- scale_data_manual(df_model_data, "type") |>
-  as_tibble()
-
-scaled_c_shaped_data <- scaled_c_shaped |>
-  filter(type == "data") |>
-  select(-type)
-
-scaled_c_shaped_data_model <- scaled_c_shaped |>
-  filter(type == "model") |>
-  select(-type)
-```
-
-```{r}
-#| label: link-plts-c-shaped-data
-#| eval: false
-
-df_exe_c_shaped_data <- comb_all_data_model_error(
-  highd_data = one_c_shaped_data,
-  nldr_data = tsne_one_c_shaped,
-  model_highd = df_bin_dens_clust_temp,
-  model_2d = df_bin_centroids_dens_clust,
-  error_data = error_df_one_curvy_abs_temp)
-
-show_error_link_plots(point_data = df_exe_c_shaped_data,
-                      edge_data = tr_from_to_df_dens_clust)
-```
-
-```{r}
 #| label: proj-one-dens-clust
@@
 ## First projection
-projection <- cbind(
-  c(0.05096,0.15399,0.19736,0.05110),
-  c(0.14608,-0.16929,0.11291,-0.07160))
-
-proj_obj1 <- get_projection(projection = projection,
-                            proj_scale = 1,
-                            highd_data = scaled_c_shaped_data,
-                            model_highd = scaled_c_shaped_data_model,
-                            trimesh_data = tr_from_to_df_dens_clust,
-                            axis_param = list(limits = 0.4,
-                                              axis_scaled = 3,
-                                              axis_pos_x = -0.28,
-                                              axis_pos_y = -0.28,
-                                              threshold = 0.02))
-
-proj_obj1[["cluster"]] <- factor(error_df_one_curvy_abs$error_cat_n2,
-                          levels=c("deselected", "selected"))
-
-projected_df <- proj_obj1$projected_df
-model_df <- proj_obj1$model_df
-axes <- proj_obj1$axes
-circle <- proj_obj1$circle
-
-projected_df <- projected_df |>
-      dplyr::mutate(cluster = proj_obj1$cluster)
+
+projected_df <- read_rds("data/one_c_shaped_dens_structure/projected_df1.rds")
+model_df <- read_rds("data/one_c_shaped_dens_structure/model_df1.rds")
+axes <- read_rds("data/one_c_shaped_dens_structure/axes1.rds")
+circle <- read_rds("data/one_c_shaped_dens_structure/circle1.rds")
 
@@
 
-proj_obj1[["cluster"]] <- factor(error_df_one_curvy_abs$error_cat_n,
-                          levels=c("deselected", "selected"))
-
-projected_df <- proj_obj1$projected_df
-model_df <- proj_obj1$model_df
-axes <- proj_obj1$axes
-circle <- proj_obj1$circle
-
-projected_df <- projected_df |>
-      dplyr::mutate(cluster = proj_obj1$cluster)
+projected_df <- read_rds("data/one_c_shaped_dens_structure/projected_df2.rds")
+model_df <- read_rds("data/one_c_shaped_dens_structure/model_df2.rds")
+axes <- read_rds("data/one_c_shaped_dens_structure/axes2.rds")
+circle <- read_rds("data/one_c_shaped_dens_structure/circle2.rds")
 
@@
     ),
-    alpha = 0.2#,
-    #size = 0.1#,
-    #colour = '#a65628'
+    alpha = 0.2
   )  +
@@
     ),
-    alpha = 0.3#,
-    #colour = "#999999"
+    alpha = 0.3
   ) +

### 2025-06-16 scripts/one_c_shaped_dens_structure/06_gen_model_with_tSNE.R

@@ -1,0 +1,166 @@
+## This script is to generate the model with tSNE for C-shaped data
+library(quollr)
+library(tidyverse)
+
+source("scripts/additional_functions.R")
+set.seed(20240110)
+
+## Data
+one_c_shaped_data <- read_rds(here::here("data/one_c_shaped_dens_structure/one_c_shaped_dens_data.rds"))
+one_c_shaped_data <- one_c_shaped_data |>
+  mutate(ID = row_number())
+
+## NLSR
+tsne_one_c_shaped <- read_rds(file = "data/one_c_shaped_dens_structure/one_c_shaped_dens_structure_tsne_perplexity_52.rds") |>
+  mutate(ID = row_number())
+
+## To fit the model
+num_bins_x_one_c_shaped <- 20
+
+algo_obj_dens_clust <- fit_highd_model(
+  highd_data = one_c_shaped_data,
+  nldr_data = tsne_one_c_shaped,
+  bin1 = num_bins_x_one_c_shaped,
+  q = 0.1,
+  benchmark_highdens = 1)
+
+tsne_one_c_shaped_scaled <- algo_obj_dens_clust$nldr_obj$scaled_nldr
+tr_from_to_df_dens_clust <- algo_obj_dens_clust$trimesh_data
+df_bin_centroids_dens_clust <- algo_obj_dens_clust$model_2d
+df_bin_dens_clust <- algo_obj_dens_clust$model_highd
+
+## Compute error
+error_df_one_curvy_abs <- augment(
+  highd_data = one_c_shaped_data,
+  model_2d = df_bin_centroids_dens_clust,
+  model_highd = df_bin_dens_clust)
+
+df_bin_dens_clust_temp <- df_bin_dens_clust
+
+error_df_one_curvy_abs_temp <- error_df_one_curvy_abs
+
+error_df_one_curvy_abs <- error_df_one_curvy_abs |>
+  bind_cols(tsne_one_c_shaped_scaled |>
+              select(-ID))
+
+error_df_one_curvy_abs <- error_df_one_curvy_abs |>
+  mutate(sqrt_row_wise_total_error = sqrt(row_wise_total_error))
+
+# Compute density
+density_data <- density(error_df_one_curvy_abs$sqrt_row_wise_total_error)
+density_df <- data.frame(x = density_data$x, y = density_data$y)
+
+# Add density values to the original dataset
+error_df_one_curvy_abs <- error_df_one_curvy_abs %>%
[…]

### 2025-06-16 paper.qmd
Replaced 4-D with notation

@@
 #| label: fig-two-curvy-true-proj
-#| fig-cap: "Two \\gD{} projections from $4\\text{-}D$, for the two nonlinear clusters data (blue points) and true model (black lines) are shown.
+#| fig-cap: "Two \\gD{} projections from \\fD{}, for the two nonlinear clusters data (blue points) and true model (black lines) are shown.
 The top cluster in (a) shows uniformly distributed data along a C-shape, while in the bottom cluster, the data is more concentrated in one corner of the curvy shape.
@@
 #| fig-cap: "Lifting the fitted model into \\pD{}.
-Starting from a default tSNE layout (a) of nonlinear clusters data ($n =  2000$ and $p = 4$), from which the \\gD{} wireframe is constructed (a) and then lifted to $4\\text{-}D$ (b, c).
+Starting from a default tSNE layout (a) of nonlinear clusters data ($n =  2000$ and $p = 4$), from which the \\gD{} wireframe is constructed (a) and then lifted to \\fD{} (b, c).
 The fit is reasonbly tight with the data, although it does not fully spread the full width.
@@
 #| label: fig-p-d-error-in-2d-two-curvy
-#| fig-cap: "The $4\\text{-}D$ model error in \\gD{} layout.
+#| fig-cap: "The \\fD{} model error in \\gD{} layout.
 Color indicates error ($e_{hj}$), dark color indicating high error and light indicates low error.
@@
 #| fig-cap: "NLDR's organise points in the \\gD{} layout in different ways, possibly misleadingly, illustrated using three layouts: (a) tSNE, (b) UMAP, (c) PaCMAP.
-The data has five Gaussian cluster in $4\\text{-}D$.
-The bottom row of plots shows a \\gD{} projection from a tour on $4\\text{-}D$ revealing the differences generated by the layouts on the model fits.
+The data has five Gaussian cluster in \\fD{}.
+The bottom row of plots shows a \\gD{} projection from a tour on \\fD{} revealing the differences generated by the layouts on the model fits.
 We would expect the model fit to be like that in (a2) where it is distinctly is separate for each cluster but like a hairball in each.
-This would indicate the distinct clusters, each being fully $4\\text{-}D$.
-With (c2), the curiousity is that the model is a \\gD{} pancake shape in $4\\text{-}D$, indicating that there is some ordering of points done by PaCMAP, posisbly along some principal component axes.
+This would indicate the distinct clusters, each being fully \\fD{}.
+With (c2), the curiousity is that the model is a \\gD{} pancake shape in \\fD{}, indicating that there is some ordering of points done by PaCMAP, posisbly along some principal component axes.
 Videos of the langevitour animations are available at <https://youtu.be/I-kxCwVfqiQ>, <https://youtu.be/gD1P01FUPyU>, and <https://youtu.be/MxJ_srOFQNk> respectively."
@@
 #| fig-cap: "The sparse end of the structure shows a high model error, while the dense end shows a low model error.
-The tSNE layout of the C-shaped structure, with dense points (a1-a3) are colored according to their $4\\text{-}D$ model error.
+The tSNE layout of the C-shaped structure, with dense points (a1-a3) are colored according to their \\fD{} model error.
 Darker colors represent a high error, while lighter colors indicate a low error.
-a2, a3 and b2, b3 present the results from linked plots that brushed the high $4\\text{-}D$ error points in \\gD{} and $4\\text{-}D$.
-It helps in identifying the high $4\\text{-}D$ model errors that occur due to the sparse end.
-Videos of the linked plots are available at <>."
+a2, a3 and b2, b3 present the results from linked plots that brushed the high \\fD{} error points in \\gD{} and \\fD{}.
+It helps in identifying the high \\fD{} model errors that occur due to the sparse end.
+The link to the video of the linked plots is available in @tbl-links."
 

### 2025-06-16 paper.qmd
Added video links to the figure titles

@@
 Furthermore, the sparse space observed in the middle of the tSNE layout is a result of (hyper-)~parameter choice.
-Video of the langevitour animation is available at <>."
+The link to the video of the langevitour animation is available in @tbl-links."
 #| fig-width: 8
@@
 With (c2), the curiousity is that the model is a \\gD{} pancake shape in \\fD{}, indicating that there is some ordering of points done by PaCMAP, posisbly along some principal component axes.
-Videos of the langevitour animations are available at <https://youtu.be/I-kxCwVfqiQ>, <https://youtu.be/gD1P01FUPyU>, and <https://youtu.be/MxJ_srOFQNk> respectively."
+The links to the videos of the langevitour animations are available in @tbl-links."
 

### 2025-06-16 paper.qmd
Moved Five Gaussian clusters code to script

@@
 
-training_data_gau <- read_rds("data/five_gau_clusters/data_five_gau.rds")
-
-data_gau <- training_data_gau |> 
-  select(-ID) |>
-  mutate(type = "data")
-
-tsne_data_gau <- read_rds("data/five_gau_clusters/tsne_data_five_gau_71.rds")
-
-## Compute hexbin parameters
-num_bins_x_gau1 <- 18
-
-algo_obj_gau1 <- fit_highd_model(
-  highd_data = training_data_gau, 
-  nldr_data = tsne_data_gau, 
-  bin1 = num_bins_x_gau1, 
-  q = 0.1, 
-  benchmark_highdens = 5)
-
-tsne_gau_scaled <- algo_obj_gau1$nldr_obj$scaled_nldr
-tr_from_to_df_gau1 <- algo_obj_gau1$trimesh_data
-df_bin_centroids_gau1 <- algo_obj_gau1$model_2d
-df_bin_gau1 <- algo_obj_gau1$model_highd
+tsne_gau_scaled <- read_rds("data/five_gau_clusters/tsne_gau_scaled.rds")
+tr_from_to_df_gau1 <- read_rds("data/five_gau_clusters/tr_from_to_df_gau1.rds")
 
@@
     colour = "#000000",
-    #alpha = 0.4,
     linewidth = 0.5) +
@@
 
-df_b <- df_bin_gau1 |>
-  dplyr::filter(hexID %in% df_bin_centroids_gau1$hexID) |>
-  dplyr::mutate(type = "model") ## Data with summarized mean
-
-## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
-df_b <- df_b[match(df_bin_centroids_gau1$hexID, df_b$hexID),] |>
-  dplyr::select(-hexID)
-
-# Apply the scaling
-df_model_data <- bind_rows(data_gau, df_b)
-scaled_gau <- scale_data_manual(df_model_data, "type") |>
-  as_tibble()
-
-scaled_gau_data <- scaled_gau |>
-  filter(type == "data") |>
-  select(-type)
-
-scaled_gau_data_model <- scaled_gau |>
-  filter(type == "model") |>
-  select(-type)
-
-```
-
-```{r}
-#| label: langevitour-five-gau-tsne-proj
-#| eval: false
-
-df_model_data_n <- bind_rows(df_b, data_gau)
-
-langevitour::langevitour(df_model_data_n[1:(length(df_model_data_n)-1)],
-                         lineFrom = tr_from_to_df_gau1$from,
-                         lineTo = tr_from_to_df_gau1$to,
-                         group = factor(df_model_data_n$type,
-                                        c("data", "model")),
-                         levelColors = c(clr_choice, "#000000"),
-                         pointSize = 0.7)
 ```
@@
 #| label: five-gau-tsne-projs
-projection <- cbind(
-  c(0.10394,0.08379,-0.17868,0.32960),
-  c(0.35111,-0.01137,-0.09657,-0.16018))
-
-proj_obj1 <- get_projection(projection = projection, 
-               proj_scale = 2.5, 
-               highd_data = scaled_gau_data, 
-               model_highd = scaled_gau_data_model, 
-               trimesh_data = tr_from_to_df_gau1, 
-               axis_param = list(limits = 1,
-                                 axis_scaled = 2,
-                                 axis_pos_x = -0.72, 
-                                 axis_pos_y = -0.72, 
-                                 threshold = 0.1))
-
-projected_df <- proj_obj1$projected_df
-model_df <- proj_obj1$model_df
-axes <- proj_obj1$axes
-circle <- proj_obj1$circle
-
-five_gau_proj_tsne_model1 <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_point(
-    #size = 0.5,
-    alpha = 0.05,
-    colour = clr_choice) +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    colour = "#000000",
-    alpha = 0.4,
-    linewidth = 0.8) +
-  geom_segment(
-    data=axes, 
-    aes(x=x1, y=y1, xend=x2, yend=y2), 
-    colour="grey70") +
-  geom_text(
-    data=axes, 
-    aes(x=x2, y=y2),
-    label=rownames(axes), 
-    colour="grey50",
-    size = 5) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.89, 0.89)) +
-  ylim(c(-0.89, 0.89)) +
+## First projection
+proj_obj1 <- read_rds("data/five_gau_clusters/tsne_gau_proj_obj1.rds")
+
+five_gau_proj_tsne_model1 <- plot_proj(
+  proj_obj = proj_obj1,
+  point_param = c(1.5, 0.05, clr_choice),
+  line_param = c(0.8, 0.4, "#000000"),
+  plot_limits = c(-0.89, 0.89),
+  axis_text_size = 5,
+  is_category = FALSE
+) +
   interior_annotation("a2", 
@@
 ## Second projection
-projection <- cbind(
-  c(0.30943,-0.08452,-0.23005,-0.05284),
-  c(0.22463,-0.06682,0.32106,0.02452))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 3, 
-                            highd_data = scaled_gau_data, 
-                            model_highd = scaled_gau_data_model, 
-                            trimesh_data = tr_from_to_df_gau1, 
-                            axis_param = list(limits = 1.5,
-                                              axis_scaled = 2,
-                                              axis_pos_x = -1, 
-                                              axis_pos_y = -1, 
-                                              threshold = 0.05))
+
+proj_obj2 <- read_rds("data/five_gau_clusters/tsne_gau_proj_obj2.rds")
 
@@
 
-umap_data_gau <- read_rds("data/five_gau_clusters/umap_data_five_gau.rds")
-
-## Compute hexbin parameters
-num_bins_x_gau1 <- 63
-
-algo_obj_gau1 <- fit_highd_model(
-  highd_data = training_data_gau, 
-  nldr_data = umap_data_gau, 
-  bin1 = num_bins_x_gau1, 
-  q = 0.1, 
-  benchmark_highdens = 5)
-
-umap_gau_scaled <- algo_obj_gau1$nldr_obj$scaled_nldr
-tr_from_to_df_gau1 <- algo_obj_gau1$trimesh_data
-df_bin_centroids_gau1 <- algo_obj_gau1$model_2d
-df_bin_gau1 <- algo_obj_gau1$model_highd
+umap_gau_scaled <- read_rds("data/five_gau_clusters/umap_gau_scaled.rds")
+tr_from_to_df_gau1_umap <- read_rds("data/five_gau_clusters/tr_from_to_df_gau1_umap.rds")
 
@@
   geom_segment(
-    data = tr_from_to_df_gau1,
+    data = tr_from_to_df_gau1_umap,
     aes(
@@
   )
-
-df_b <- df_bin_gau1 |>
-  dplyr::filter(hexID %in% df_bin_centroids_gau1$hexID) |>
-  dplyr::mutate(type = "model") ## Data with summarized mean
-
-## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
-df_b <- df_b[match(df_bin_centroids_gau1$hexID, df_b$hexID),] |>
-  dplyr::select(-hexID)
-
-# Apply the scaling
-df_model_data <- bind_rows(data_gau, df_b)
-scaled_gau <- scale_data_manual(df_model_data, "type") |>
-  as_tibble()
-
-scaled_gau_data <- scaled_gau |>
-  filter(type == "data") |>
-  select(-type)
-
-scaled_gau_data_model <- scaled_gau |>
-  filter(type == "model") |>
-  select(-type)
-
-```
-
-```{r}
-#| label: langevitour-five-gau-umap-proj
-#| eval: false
-
-df_model_data_n <- bind_rows(df_b, data_gau)
-
-langevitour::langevitour(df_model_data_n[1:(length(df_model_data_n)-1)],
-                         lineFrom = tr_from_to_df_gau1$from,
-                         lineTo = tr_from_to_df_gau1$to,
-                         group = factor(df_model_data_n$type,
-                                        c("data", "model")),
-                         levelColors = c(clr_choice, "#000000"), pointSize = 0.7)
 ```
@@
 #| label: five-gau-umap-projs
-projection <- cbind(
-  c(0.10394,0.08379,-0.17868,0.32960),
-  c(0.35111,-0.01137,-0.09657,-0.16018))
-
-proj_obj1 <- get_projection(projection = projection, 
-               proj_scale = 2.5, 
-               highd_data = scaled_gau_data, 
-               model_highd = scaled_gau_data_model, 
-               trimesh_data = tr_from_to_df_gau1, 
-               axis_param = list(limits = 1,
-                                 axis_scaled = 2,
-                                 axis_pos_x = -0.72, 
-                                 axis_pos_y = -0.72, 
-                                 threshold = 0.1))
-
-projected_df <- proj_obj1$projected_df
-model_df <- proj_obj1$model_df
-axes <- proj_obj1$axes
-circle <- proj_obj1$circle
-
-five_gau_proj_umap_model1 <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1, 
-      y = proj2)) +
-  geom_point(
-    alpha = 0.05,
-    colour = clr_choice) +
-  geom_segment(
-    data = model_df, 
-    aes(
-      x = proj1_from, 
-      y = proj2_from, 
-      xend = proj1_to, 
-      yend = proj2_to), 
-    colour = "#000000",
-    alpha = 0.4,
-    linewidth = 0.8) +
-  geom_segment(
-    data=axes, 
-    aes(x=x1, y=y1, xend=x2, yend=y2), 
-    colour="grey70") +
-  geom_text(
-    data=axes, 
-    aes(x=x2, y=y2, label=rownames(axes)), 
-    colour="grey50",
-    size = 5) +
-  geom_path(
-    data=circle, 
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.89, 0.89)) +
-  ylim(c(-0.89, 0.89)) +
+
+## First projection
+proj_obj1 <- read_rds("data/five_gau_clusters/umap_gau_proj_obj1.rds")
+
+five_gau_proj_umap_model1 <- plot_proj(
+  proj_obj = proj_obj1,
+  point_param = c(1.5, 0.05, clr_choice),
+  line_param = c(0.8, 0.4, "#000000"),
+  plot_limits = c(-0.89, 0.89),
+  axis_text_size = 5,
+  is_category = FALSE
+) +
   interior_annotation("b2", 
@@
 ## Second projection
-projection <- cbind(
-  c(0.30943,-0.08452,-0.23005,-0.05284),
-  c(0.22463,-0.06682,0.32106,0.02452))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 3, 
-                            highd_data = scaled_gau_data, 
-                            model_highd = scaled_gau_data_model, 
-                            trimesh_data = tr_from_to_df_gau1, 
-                            axis_param = list(limits = 1.5,
-                                              axis_scaled = 2,
-                                              axis_pos_x = -1, 
-                                              axis_pos_y = -1, 
-                                              threshold = 0.05))
+proj_obj2 <- read_rds("data/five_gau_clusters/umap_gau_proj_obj2.rds")
 
@@
 
-pacmap_data_gau <- read_rds("data/five_gau_clusters/pacmap_data_five_gau.rds")
-
-## Compute hexbin parameters
-num_bins_x_gau1 <- 28
-
-algo_obj_gau1 <- fit_highd_model(
-  highd_data = training_data_gau, 
-  nldr_data = pacmap_data_gau, 
-  bin1 = num_bins_x_gau1, 
-  q = 0.1, 
-  benchmark_highdens = 5)
-
-pacmap_gau_scaled <- algo_obj_gau1$nldr_obj$scaled_nldr
-tr_from_to_df_gau1 <- algo_obj_gau1$trimesh_data
-df_bin_centroids_gau1 <- algo_obj_gau1$model_2d
-df_bin_gau1 <- algo_obj_gau1$model_highd
+pacmap_gau_scaled <- read_rds("data/five_gau_clusters/pacmap_gau_scaled.rds")
+tr_from_to_df_gau1_pacmap <- read_rds("data/five_gau_clusters/tr_from_to_df_gau1_pacmap.rds")
 
@@
   geom_segment(
-    data = tr_from_to_df_gau1,
+    data = tr_from_to_df_gau1_pacmap,
     aes(
@@
   )
-
-df_b <- df_bin_gau1 |>
-  dplyr::filter(hexID %in% df_bin_centroids_gau1$hexID) |>
-  dplyr::mutate(type = "model") ## Data with summarized mean
-
-## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
-df_b <- df_b[match(df_bin_centroids_gau1$hexID, df_b$hexID),] |>
-  dplyr::select(-hexID)
-
-# Apply the scaling
-df_model_data <- bind_rows(data_gau, df_b)
-scaled_gau <- scale_data_manual(df_model_data, "type") |>
-  as_tibble()
-
-scaled_gau_data <- scaled_gau |>
-  filter(type == "data") |>
-  select(-type)
-
-scaled_gau_data_model <- scaled_gau |>
-  filter(type == "model") |>
-  select(-type)
-```
-
-```{r}
-#| label: langevitour-five-gau-pacmap-proj
-#| eval: false
-
-df_model_data_n <- bind_rows(df_b, data_gau)
-
-langevitour::langevitour(df_model_data_n[1:(length(df_model_data_n)-1)],
-                         lineFrom = tr_from_to_df_gau1$from,
-                         lineTo = tr_from_to_df_gau1$to,
-                         group = factor(df_model_data_n$type,
-                                        c("data", "model")),
-                         levelColors = c(clr_choice, "#000000"), pointSize = 0.7)
 ```
@@
 ## First projection
-projection <- cbind(
-  c(0.10394,0.08379,-0.17868,0.32960),
-  c(0.35111,-0.01137,-0.09657,-0.16018))
-
-proj_obj1 <- get_projection(projection = projection, 
-               proj_scale = 2.5, 
-               highd_data = scaled_gau_data, 
-               model_highd = scaled_gau_data_model, 
-               trimesh_data = tr_from_to_df_gau1, 
-               axis_param = list(limits = 1,
-                                 axis_scaled = 2,
-                                 axis_pos_x = -0.72, 
-                                 axis_pos_y = -0.72, 
-                                 threshold = 0.1))
-
-projected_df <- proj_obj1$projected_df
-model_df <- proj_obj1$model_df
-axes <- proj_obj1$axes
-circle <- proj_obj1$circle
-
-five_gau_proj_pacmap_model1 <- projected_df |>
-  ggplot(
-    aes(
-      x = proj1,
-      y = proj2)) +
-  geom_point(
-    alpha = 0.05,
-    colour = clr_choice) +
-  geom_segment(
-    data = model_df,
-    aes(
-      x = proj1_from,
-      y = proj2_from,
-      xend = proj1_to,
-      yend = proj2_to),
-    colour = "#000000",
-    alpha = 0.4,
-    linewidth = 0.8) +
-  geom_segment(
-    data=axes,
-    aes(x=x1, y=y1, xend=x2, yend=y2),
-    colour="grey70") +
-  geom_text(
-    data=axes,
-    aes(x=x2, y=y2, label=rownames(axes)),
-    colour="grey50",
-    size = 5) +
-  geom_path(
-    data=circle,
-    aes(x=c1, y=c2), colour="grey70") +
-  coord_fixed() +
-  xlim(c(-0.89, 0.89)) +
-  ylim(c(-0.89, 0.89)) +
+proj_obj1 <- read_rds("data/five_gau_clusters/pacmap_gau_proj_obj1.rds")
+
+five_gau_proj_pacmap_model1 <- plot_proj(
+  proj_obj = proj_obj1,
+  point_param = c(1.5, 0.05, clr_choice),
+  line_param = c(0.8, 0.4, "#000000"),
+  plot_limits = c(-0.89, 0.89),
+  axis_text_size = 5,
+  is_category = FALSE
+) +
   interior_annotation("c2",
@@
 ## Second projection
-projection <- cbind(
-  c(0.30943,-0.08452,-0.23005,-0.05284),
-  c(0.22463,-0.06682,0.32106,0.02452))
-
-proj_obj2 <- get_projection(projection = projection, 
-                            proj_scale = 3, 
-                            highd_data = scaled_gau_data, 
-                            model_highd = scaled_gau_data_model, 
-                            trimesh_data = tr_from_to_df_gau1, 
-                            axis_param = list(limits = 1.5,
-                                              axis_scaled = 2,
-                                              axis_pos_x = -1, 
-                                              axis_pos_y = -1, 
-                                              threshold = 0.05))
+proj_obj2 <- read_rds("data/five_gau_clusters/pacmap_gau_proj_obj2.rds")
 

### 2025-06-16 scripts/five_gau_clusters/8_gen_model_with_UMAP.R

@@ -1,0 +1,104 @@
+## This script is to generate the model with UMAP for Five Gaussian clusters data
+library(quollr)
+library(tidyverse)
+
+source("scripts/additional_functions.R")
+set.seed(20240110)
+
+clr_choice <- "#0077A3"
+
+## Data
+training_data_gau <- read_rds("data/five_gau_clusters/data_five_gau.rds")
+
+data_gau <- training_data_gau |>
+  select(-ID) |>
+  mutate(type = "data")
+
+## NLDR
+umap_data_gau <- read_rds("data/five_gau_clusters/umap_data_five_gau.rds")
+
+## Compute hexbin parameters
+num_bins_x_gau1 <- 63
+
+algo_obj_gau1 <- fit_highd_model(
+  highd_data = training_data_gau,
+  nldr_data = umap_data_gau,
+  bin1 = num_bins_x_gau1,
+  q = 0.1,
+  benchmark_highdens = 5)
+
+umap_gau_scaled <- algo_obj_gau1$nldr_obj$scaled_nldr
+tr_from_to_df_gau1 <- algo_obj_gau1$trimesh_data
+df_bin_centroids_gau1 <- algo_obj_gau1$model_2d
+df_bin_gau1 <- algo_obj_gau1$model_highd
+
+write_rds(umap_gau_scaled, "data/five_gau_clusters/umap_gau_scaled.rds")
+write_rds(tr_from_to_df_gau1, "data/five_gau_clusters/tr_from_to_df_gau1_umap.rds")
+
+df_b <- df_bin_gau1 |>
+  dplyr::filter(hexID %in% df_bin_centroids_gau1$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b <- df_b[match(df_bin_centroids_gau1$hexID, df_b$hexID),] |>
+  dplyr::select(-hexID)
+
+# Apply the scaling
+df_model_data <- bind_rows(data_gau, df_b)
+scaled_gau <- scale_data_manual(df_model_data, "type") |>
+  as_tibble()
+
+scaled_gau_data <- scaled_gau |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_gau_data_model <- scaled_gau |>
+  filter(type == "model") |>
+  select(-type)
+
+df_model_data_n <- bind_rows(df_b, data_gau)
+
+langevitour::langevitour(df_model_data_n[1:(length(df_model_data_n)-1)],
[…]

### 2025-06-16 paper.qmd
Removed prediction codes

@@
 
-# df_bin_centroids_two_curvy <- df_bin_centroids_two_curvy |>
-#   dplyr::filter(std_counts > 0)
-
 hex_grid_nonempty_two_curvy <- ggplot(
@@
 
-proj_obj1 <- quollr::get_projection(projection = model_prj1, 
+proj_obj1 <- get_projection(projection = model_prj1, 
                proj_scale = 1, 
@@
 <!-- The remaining variables $X_4, X_5, X_6, X_7$ are all uniform error, with small variance.  -->
-
-```{r}
-#| label: fig-two-curvy-true-proj
-#| fig-cap: "Two \\gD{} projections from \\fD{}, for the two nonlinear clusters data (blue points) and true model (black lines) are shown.
-The top cluster in (a) shows uniformly distributed data along a C-shape, while in the bottom cluster, the data is more concentrated in one corner of the curvy shape.
-Video of the langevitour animations are available at <>."
-#| fig-pos: H
-#| fig-width: 10
-#| fig-height: 5
-#| eval: false
-
-two_curvy_proj_true + two_curvy_proj_true2 +
-  plot_layout(ncol = 2)
-```
 
@@
 
-<!--prediction for true-model of two-curvy clust-->
-```{r}
-#| label: model-prediction-umap-original
-
-predict_tsne_two_curvy <- read_rds(file = "data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_umap_predict_true.rds")
-
-predict_two_curvy_obj <- gen_scaled_data(
-    nldr_data = predict_tsne_two_curvy)
-
-predict_tsne_two_curvy_scaled <- predict_two_curvy_obj$scaled_nldr
-
-plot_predict_umap_fun_two_curvy <- ggplot(
-  data = predict_tsne_two_curvy_scaled, 
-  aes(x = UMAP1, 
-      y = UMAP2)) +
-  geom_point(alpha = 0.5) +
-  interior_annotation("b", c(0.08, 0.93), cex = 1.5)
-```
-
-```{r}
-#| label: model-prediction-points-line
-#| eval: false
-
-predict_tsne_two_curvy_scaled <- predict_tsne_two_curvy_scaled |>
-  mutate(ID = row_number()) |>
-  mutate(type = "predict_from_umap")
-
-jittered_points_df <- jittered_points_df |>
-  mutate(type = "predict_from_model")
-
-predict_point_df <- bind_rows(
-  predict_tsne_two_curvy_scaled,
-  jittered_points_df
-)
-
-plot_predict_umap_connect_two_curvy <- ggplot(
-  data = predict_point_df, 
-  aes(x = UMAP1, 
-      y = UMAP2, 
-      group = ID)) +
-  geom_point(alpha = 0.5) +
-  geom_line(alpha = 0.3) +
-  interior_annotation("d", c(0.08, 0.93), cex = 1.5)
-```
-
-```{r}
-#| echo: false
-#| label: fig-predict-two-curvy
-#| fig-pos: H
-#| fig-width: 10
-#| fig-height: 10
-#| out-width: 80%
-#| fig-cap: "Comparison of prediction generated using the exiting `umap` R package's prediction method and our method: (a) A view of the true model in projection from $7\\text{-}D$, (b) predicted data from the `umap` prediction method, and (c) predicted data from our method."
-#| eval: false
-
-proj_plot_two_curvy_true + plot_predict_umap_fun_two_curvy +plot_predict_tsne_two_curvy +
-  plot_layout(ncol = 3)
-```
-
 ### Tuning

### 2025-06-17 paper.qmd
Moved two nonlinear data code to script

@@
 ```{r}
-#| label: two-curvy-training
-training_data_two_curvy <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_data.rds") |>
-  mutate(ID = row_number())
-```
-
-```{r}
-#| label: two-curvy-true-proj-data
-
-true_model_df <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_true_model.rds")
-wireframe_true_model <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_true_model_connections.rds")
-
-data_two_curvy <- training_data_two_curvy |>
-  select(-ID) |>
-  mutate(type = "data")
-
-true_model_two_curvy <- true_model_df |> 
-  mutate(type = "true model")
-```
-
-<!--tSNE applied for Two curvy data-->
-```{r}
-#| label: tsne-two-curvy
-tsne_two_curvy <- read_rds(file = "data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_30.rds") |>
-  mutate(ID = row_number()) |>
-  rename(c("emb1" = "tSNE1",
-           "emb2" = "tSNE2"))
-```
-
-```{r}
 #| label: fit_model-two-curvy
 
-num_bins_x_two_curvy <- 24
-
-algo_obj_two_curvy <- fit_highd_model(
-  highd_data = training_data_two_curvy, 
-  nldr_data = tsne_two_curvy, 
-  bin1 = num_bins_x_two_curvy, 
-  q = 0.1, 
-  benchmark_highdens = 1)
-
-tsne_two_curvy_scaled <- algo_obj_two_curvy$nldr_obj$scaled_nldr
-tr_from_to_df_two_curvy <- algo_obj_two_curvy$trimesh_data
-df_bin_centroids_two_curvy <- algo_obj_two_curvy$model_2d
-df_bin_two_curvy <- algo_obj_two_curvy$model_highd
-hex_grid <- algo_obj_two_curvy$hb_obj$hex_poly
-counts_df <- algo_obj_two_curvy$hb_obj$std_cts
-
-hex_grid_with_counts <- left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hexID"))
-
-hex_grid_nonempty <- hex_grid |>
-  filter(hex_poly_id %in% df_bin_centroids_two_curvy$hexID)
-
-bin_width <- algo_obj_two_curvy$hb_obj$a1
+tsne_two_curvy_scaled <- read_rds("data/two_nonlinear/tsne_two_curvy_scaled.rds")
+hex_grid_with_counts <- read_rds("data/two_nonlinear/hex_grid_with_counts.rds")
+hex_grid_nonempty <- read_rds("data/two_nonlinear/hex_grid_nonempty.rds")
+tr_from_to_df_two_curvy <- read_rds("data/two_nonlinear/tr_from_to_df_two_curvy.rds")
+df_bin_centroids_two_curvy <- read_rds("data/two_nonlinear/df_bin_centroids_two_curvy.rds")
 
@@
   ylim(sc_ylims)
-
-# Second plot (identical to base except for xlim/ylim)
-nldr_two_curvy2 <- nldr_two_curvy_original +
-  interior_annotation("a", sc_ltr_pos, cex = 1.5)
-
-# Third plot with aspect ratio and different annotation
-nldr_two_curvy_original2 <- nldr_two_curvy_original +
-  theme(aspect.ratio = 1) +
-  interior_annotation("a1", sc_ltr_pos)
 
@@
 ## Compute error
-error_df_two_curvy_abs <- augment(
-  highd_data = training_data_two_curvy,
-  model_2d = df_bin_centroids_two_curvy,
-  model_highd = df_bin_two_curvy)
-
-error_df_two_curvy_abs <- error_df_two_curvy_abs |>
-  mutate(sqrt_row_wise_total_error = sqrt(row_wise_total_error))
-
-error_df_two_curvy_abs <- error_df_two_curvy_abs |>
-  bind_cols(tsne_two_curvy_scaled |>
-              select(-ID))
+error_df_two_curvy_abs <- read_rds("data/two_nonlinear/error_df_two_curvy_abs.rds")
 
@@
 ```{r}
-#| label: prep-data-for-proj-tsne-best
-
-df_bin_two_curvy <- df_bin_two_curvy |>
-  select(-hexID) |>
-  mutate(type = "model")
-
-# Apply the scaling
-df_model_data_two_curvy <- bind_rows(data_two_curvy, true_model_two_curvy, df_bin_two_curvy)
-
-scaled_two_curvy <- scale_data_manual(df_model_data_two_curvy, "type") |>
-  as_tibble()
-
-scaled_two_curvy_data <- scaled_two_curvy |>
-  filter(type == "data") |>
-  select(-type)
-
-scaled_two_curvy_data_true_model <- scaled_two_curvy |>
-  filter(type == "true model") |>
-  select(-type)
-
-scaled_two_curvy_data_model <- scaled_two_curvy |>
-  filter(type == "model") |>
-  select(-type)
-
-```
-
-```{r}
-#| label: langevitour-true-two-curvy-proj
-#| eval: false
-df_model_data_two_curvy_filtered <- bind_rows(true_model_two_curvy, data_two_curvy)
-
-langevitour::langevitour(df_model_data_two_curvy_filtered[1:(length(df_model_data_two_curvy_filtered)-1)],
-                         lineFrom = wireframe_true_model$from,
-                         lineTo = wireframe_true_model$to,
-                         group = factor(df_model_data_two_curvy_filtered$type,
-                                        c("data", "true model")),
-                         levelColors = c(clr_choice, "#000000"))
-```
-
-```{r}
-#| label: langevitour-best-fit-two-curvy-proj-tsne
-#| eval: false
-
-df_model_data_two_curvy_filtered <- bind_rows(df_bin_two_curvy, data_two_curvy)
-
-langevitour::langevitour(df_model_data_two_curvy_filtered[1:(length(df_model_data_two_curvy_filtered)-1)],
-                         lineFrom = tr_from_to_df_two_curvy$from,
-                         lineTo = tr_from_to_df_two_curvy$to,
-                         group = factor(df_model_data_two_curvy_filtered$type,
-                                        c("data", "model")),
-                         levelColors = c(clr_choice, "#000000"))
-```
-
-```{r}
 #| label: best-fit-two-curvy-proj-tsne
@@
 ## First projection
-model_prj1 <- cbind(
-  c(0.37847,-0.06517,0.04231,0.01655),
-  c(-0.06466,-0.27052,0.22131,-0.15235))
-
-proj_obj1 <- get_projection(projection = model_prj1, 
-               proj_scale = 1, 
-               highd_data = scaled_two_curvy_data, 
-               model_highd = scaled_two_curvy_data_model, 
-               trimesh_data = tr_from_to_df_two_curvy, 
-               axis_param = list(limits = 0.5, 
-                                 axis_scaled = 2,
-                                 axis_pos_x = -0.5, 
-                                 axis_pos_y = -0.5, 
-                                 threshold = 0.03))
-
-## For true model
-
-projection_scaled <- model_prj1 * 1
-
-projected_true_model <- as.matrix(scaled_two_curvy_data_true_model) %*% projection_scaled
-
-projected_true_model_df <- projected_true_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-true_model_df_proj <- dplyr::left_join(
-  wireframe_true_model, 
-  projected_true_model_df, 
-  by = c("from" = "ID"))
-
-names(true_model_df_proj)[3:NCOL(true_model_df_proj)] <- paste0(names(projected_true_model_df)[-NCOL(projected_true_model_df)], "_from")
-
-true_model_df_proj <- dplyr::left_join(true_model_df_proj, projected_true_model_df, by = c("to" = "ID"))
-names(true_model_df_proj)[(2 + NCOL(projected_true_model_df)):NCOL(true_model_df_proj)] <- paste0(names(projected_true_model_df)[-NCOL(projected_true_model_df)], "_to")
-
-proj_obj1_true <- proj_obj1
-proj_obj1_true[["model_df"]] <- true_model_df_proj
-
-two_curvy_proj_true <- plot_proj(
-  proj_obj = proj_obj1_true, 
-  point_param = c(0.5, 0.5, clr_choice), # size, alpha, color
-  line_param = c(0.5, 0.5, "black"), # linewidth, alpha
-  plot_limits = c(-0.6, 0.6), 
-  axis_text_size = 4, 
-  is_category = FALSE) +
-  interior_annotation(label = "a", cex = 1.5) +
-  theme(aspect.ratio=1, 
-        legend.position = "none")
-
-# Changed the axis parameters
-axis_obj <- gen_axes(
-    proj = model_prj1 * 2,
-    limits = 0.9,
-    axis_pos_x = -0.4,
-    axis_pos_y = -0.4,
-    axis_labels = names(scaled_two_curvy_data),
-    threshold = 0.05)
-
-axes <- axis_obj$axes
-circle <- axis_obj$circle
-
-proj_obj1[["axes"]] <- axes
-proj_obj1[["circle"]] <- circle
+proj_obj1 <- read_rds("data/two_nonlinear/two_nonlinear_proj_obj1.rds")
 
@@
 ## Second projection
-model_prj2 <- cbind(
-  c(-0.38281,-0.04870,0.00619,-0.02451),
-  c(-0.01303,0.00634,-0.37421,0.09648))
-
-proj_obj2 <- get_projection(projection = model_prj2, 
-                            proj_scale = 1, 
-                            highd_data = scaled_two_curvy_data, 
-                            model_highd = scaled_two_curvy_data_model, 
-                            trimesh_data = tr_from_to_df_two_curvy, 
-                            axis_param = list(limits = 0.35, 
-                                              axis_scaled = 2,
-                                              axis_pos_x = -0.35, 
-                                              axis_pos_y = -0.35, 
-                                              threshold = 0.01))
-
-projected_true_model <- as.matrix(scaled_two_curvy_data_true_model) %*% model_prj2
-
-projected_true_model_df <- projected_true_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-true_model_df_proj <- dplyr::left_join(
-  wireframe_true_model, 
-  projected_true_model_df, 
-  by = c("from" = "ID"))
-
-names(true_model_df_proj)[3:NCOL(true_model_df_proj)] <- paste0(names(projected_true_model_df)[-NCOL(projected_true_model_df)], "_from")
-
-true_model_df_proj <- dplyr::left_join(true_model_df_proj, projected_true_model_df, by = c("to" = "ID"))
-names(true_model_df_proj)[(2 + NCOL(projected_true_model_df)):NCOL(true_model_df_proj)] <- paste0(names(projected_true_model_df)[-NCOL(projected_true_model_df)], "_to")
-
-proj_obj2_true <- proj_obj2
-proj_obj2_true[["model_df"]] <- true_model_df_proj
-
-two_curvy_proj_true2 <- plot_proj(
-    proj_obj = proj_obj2_true, 
-    point_param = c(0.5, 0.5, clr_choice), # size, alpha, color
-    line_param = c(0.5, 0.5, "black"), # linewidth, alpha
-    plot_limits = c(-0.42, 0.42), 
-    axis_text_size = 4, 
-    is_category = FALSE) +
-  interior_annotation(label = "b", cex = 1.5) +
-  theme(aspect.ratio = 1, 
-        legend.position = "none")
-
-#Changed the axis parameters
-axis_obj <- gen_axes(
-    proj = model_prj2 * 2,
-    limits = 0.7,
-    axis_pos_x = -0.35,
-    axis_pos_y = -0.35,
-    axis_labels = names(scaled_two_curvy_data),
-    threshold = 0.02)
-
-axes <- axis_obj$axes
-circle <- axis_obj$circle
-
-proj_obj2[["axes"]] <- axes
-proj_obj2[["circle"]] <- circle
+
+proj_obj2 <- read_rds("data/two_nonlinear/two_nonlinear_proj_obj2.rds")
 
@@
 ## hexagon binning to have regular hexagons
-hb_obj_notation <- hex_binning(
-  nldr_obj = algo_obj_two_curvy$nldr_obj, 
-  bin1 = 7, 
-  q = 0.1)
-
-a1_temp <- hb_obj_notation$a1
-a2_temp <- hb_obj_notation$a2
-l_temp <- quad(a=3, b = 2 * a2_temp, c = -(a2_temp^2 + a1_temp^2))
-
-## Data set with all centroids
-all_centroids_df_temp <- hb_obj_notation$centroids
-hex_grid_temp <- hb_obj_notation$hex_poly
-
-hex_grid_temp40 <- hex_grid_temp |> 
-  filter(hex_poly_id == 40)
-
-start_pt <- all_centroids_df_temp |> 
-  filter(hexID == 1)
-d_rect <- tibble(x1min = 0, 
-                 x1max = 1,
-                 x2min = 0,
-                 x2max = diff(algo_obj_two_curvy$nldr_obj$lim2)/diff(algo_obj_two_curvy$nldr_obj$lim1))
-
-# To move the rectangle to ignore the overlap with the centroids
-# rect_adj <- tibble(x1 = 0.03, x2 = 0.03)
-rect_adj <- tibble(x1 = -0.03, x2 = 0.03)
-
-a1 <- tibble(x = all_centroids_df_temp$c_x[4],
-             xend = all_centroids_df_temp$c_x[5],
-             y = all_centroids_df_temp$c_y[21],
-             yend = all_centroids_df_temp$c_y[21],
-             label = expression(a[1]))
-a2 <- tibble(x = all_centroids_df_temp$c_x[25],
-             xend = all_centroids_df_temp$c_x[25],
-             y = all_centroids_df_temp$c_y[25],
-             yend = all_centroids_df_temp$c_y[33],
-             label = expression(a[2]))
-l <- tibble(x = hex_grid_temp40$x[2],
-            xend = hex_grid_temp40$x[3],
-            y = hex_grid_temp40$y[2],
-            yend = hex_grid_temp40$y[3],
-            label = expression(l))
+
+hex_grid_temp <- read_rds("data/two_nonlinear/hex_grid_temp.rds")
+all_centroids_df_temp <- read_rds("data/two_nonlinear/all_centroids_df_tempp.rds")
+start_pt <- read_rds("data/two_nonlinear/start_pt.rds")
+d_rect <- read_rds("data/two_nonlinear/d_rect.rds")
+a1 <- read_rds("data/two_nonlinear/a2_data.rds")
+a2 <- read_rds("data/two_nonlinear/a1_data.rds")
+l <- read_rds("data/two_nonlinear/l_data.rds")
+rect_adj <- read_rds("data/two_nonlinear/rect_adj.rds")
 
@@
 
-num_bins_x_two_curvy <- 15
-
-algo_obj_two_curvy2 <- fit_highd_model(
-  highd_data = training_data_two_curvy, 
-  nldr_data = tsne_two_curvy, 
-  bin1 = num_bins_x_two_curvy, 
-  q = 0.1, 
-  benchmark_highdens = 5)
-
-tr_from_to_df_two_curvy2 <- algo_obj_two_curvy2$trimesh_data
-df_bin_centroids_two_curvy2 <- algo_obj_two_curvy2$model_2d
-df_bin_two_curvy2 <- algo_obj_two_curvy2$model_highd
-
-hex_grid_two_curvy2 <- algo_obj_two_curvy2$hb_obj$hex_poly
-counts_df_two_curvy2 <- algo_obj_two_curvy2$hb_obj$std_cts
-
-hex_grid_with_counts_two_curvy2 <- left_join(hex_grid_two_curvy2, counts_df_two_curvy2, by = c("hex_poly_id" = "hexID"))
+hex_grid_with_counts_two_curvy2 <- read_rds("data/two_nonlinear/two_nonlinear_hex_grid_with_counts_two_curvy2.rds")
+tr_from_to_df_two_curvy2 <- read_rds("data/two_nonlinear/two_nonlinear_tr_from_to_df_two_curvy2.rds")
 
@@
 
-num_bins_x_two_curvy <- 48
-
-algo_obj_two_curvy3 <- fit_highd_model(
-    highd_data = training_data_two_curvy, 
-    nldr_data = tsne_two_curvy, 
-    bin1 = num_bins_x_two_curvy, 
-    q = 0.1, 
-    benchmark_highdens = 5)
-
-tr_from_to_df_two_curvy3 <- algo_obj_two_curvy3$trimesh_data
-df_bin_centroids_two_curvy3 <- algo_obj_two_curvy3$model_2d
-df_bin_two_curvy3 <- algo_obj_two_curvy3$model_highd
-
-hex_grid_two_curvy3 <- algo_obj_two_curvy3$hb_obj$hex_poly
-counts_df_two_curvy3 <- algo_obj_two_curvy3$hb_obj$std_cts
-
-hex_grid_with_counts_two_curvy3 <- left_join(hex_grid_two_curvy3, counts_df_two_curvy3, by = c("hex_poly_id" = "hexID"))
+hex_grid_with_counts_two_curvy3 <- read_rds("data/two_nonlinear/two_nonlinear_hex_grid_with_counts_two_curvy3.rds")
+tr_from_to_df_two_curvy3 <- read_rds("data/two_nonlinear/two_nonlinear_tr_from_to_df_two_curvy3.rds")
 
@@
 
-#nldr_two_curvy2 + 
 trimesh_two_curvy_removed1_tsne_with_data +
@@
 
-error_two_curvy_umap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_tsne.rds")
-
-## Find the minimum RMSE when have duplicate a1
-error_two_curvy_umap <- error_two_curvy_umap |>
-  group_by(a1) |>
-  filter(bin1 == min(bin1)) |>
-  ungroup() |>
-  mutate(prop_dens = 1/(b*side_length^2))
-
-base_line_prop_dens <- error_two_curvy_umap |>
-  filter(a1 == min(a1)) |>
-  pull(prop_dens)
-
-error_two_curvy_umap <- error_two_curvy_umap |>
-  mutate(prop_comp = prop_dens/base_line_prop_dens)
+error_two_curvy_umap <- read_rds("data/two_nonlinear/error_two_curvy_umap.rds")
 
@@
 
-error_two_curvy_umap <- error_two_curvy_umap |>
-  mutate(prop_bins = b_non_empty/b)
-
 a1_m_two_curvy <- ggplot(error_two_curvy_umap,
@@
 
-error_rm_two_curvy <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_rm_lwd_diff_bin.rds") |>
+error_rm_two_curvy <- read_rds("data/two_nonlinear/error_rm_lwd_diff_bin.rds") |>
   mutate(bin1  = as.factor(bin1))
@@
 # and plot them on same aspect ratio
-tsne_two_curvy <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_30.rds")
+tsne_two_curvy <- read_rds("data/two_nonlinear/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_30.rds")
 
@@
 
-tsne_two_curvy <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_62.rds")
+tsne_two_curvy <- read_rds("data/two_nonlinear/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_62.rds")
 
@@
 
-umap_two_curvy <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_umap_n-neigbors_15_min-dist_0.1.rds")
+umap_two_curvy <- read_rds("data/two_nonlinear/two_non_linear_diff_shaped_close_clusters_umap_n-neigbors_15_min-dist_0.1.rds")
 
@@
 
-phate_two_curvy <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_phate_knn_5.rds")
+phate_two_curvy <- read_rds("data/two_nonlinear/two_non_linear_diff_shaped_close_clusters_phate_knn_5.rds")
 
@@
 
-trimap_two_curvy <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_trimap_n-inliers_12_n-outliers_4_n-random_3.rds")
+trimap_two_curvy <- read_rds("data/two_nonlinear/two_non_linear_diff_shaped_close_clusters_trimap_n-inliers_12_n-outliers_4_n-random_3.rds")
 
@@
 
-pacmap_two_curvy <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")
+pacmap_two_curvy <- read_rds("data/two_nonlinear/two_non_linear_diff_shaped_close_clusters_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")
 
@@
 
-error_two_curvy_umap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_umap.rds")
-error_two_curvy_tsne <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_tsne.rds")
-error_two_curvy_phate <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_phate.rds")
-error_two_curvy_trimap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_trimap.rds")
-error_two_curvy_pacmap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_pacmap.rds")
-error_two_curvy_tsne2 <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_tsne2.rds")
+error_two_curvy_umap <- read_rds("data/two_nonlinear/error_two_non_linear_diff_shaped_close_clusters_umap.rds")
+error_two_curvy_tsne <- read_rds("data/two_nonlinear/error_two_non_linear_diff_shaped_close_clusters_tsne.rds")
+error_two_curvy_phate <- read_rds("data/two_nonlinear/error_two_non_linear_diff_shaped_close_clusters_phate.rds")
+error_two_curvy_trimap <- read_rds("data/two_nonlinear/error_two_non_linear_diff_shaped_close_clusters_trimap.rds")
+error_two_curvy_pacmap <- read_rds("data/two_nonlinear/error_two_non_linear_diff_shaped_close_clusters_pacmap.rds")
+error_two_curvy_tsne2 <- read_rds("data/two_nonlinear/error_two_non_linear_diff_shaped_close_clusters_tsne2.rds")
 

### 2025-06-17 scripts/two_nonlinear/01_gen_data.R

@@ -1,0 +1,77 @@
+library(tibble)
+library(dplyr)
+library(langevitour)
+library(readr)
+
+set.seed(20240110)
+
+gen_curv1_3d <- function(n = 100) {
+
+  x1 <- runif(n, 0, 2)
+  x2 <- runif(n, 0, 3)
+  x3 <- -(x1^3 + x2)
+  x4 <- runif(n, 0, 1)
+
+  df2 <- tibble(
+    x1 = x1,
+    x2 = x2,
+    x3 = x3,
+    x4 = x4
+  )
+
+  df2
+
+}
+
+gen_curv2_3d <- function(n) {
+  a <- 3 * pi * stats::runif(n = n, min = -0.5, max = 0)
+  x1 <- sin(a)
+  x2 <- 2.0 * stats::runif(n = n)
+  x3 <- sign(a) * (cos(a) - 1)
+  x4 <- cos(a)
+
+  df <- tibble(
+    x1 = x1,
+    x2 = x2,
+    x3 = x3,
+    x4 = x4
+  )
+
+  df
+}
+
+# Simulate some s_curve_noise
+
+sample_size <- 1000
+curve2 <- gen_curv1_3d(n = sample_size)
+langevitour(curve2)
+
+curve1 <- gen_curv2_3d(n = sample_size)
+langevitour(curve1)
+
+
+# Apply an offset to one of the clusters to create a distance between them
+offset <- c(1.5, 3.3, 2, 1.5)  # Adjust these values to set the desired distance
+curve2 <- sweep(curve2, 2, offset, "+")
+
+df <- bind_rows(
+  curve1,
+  curve2
+)
+
+write_rds(df, here::here("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_data_without_std.rds"))
+
+# Compute and store column means and standard deviations
+stats <- df |>
+  summarise(across(everything(), list(mean = mean, sd = sd), .names = "{.col}_{.fn}"))
+
+write_rds(stats, here::here("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_data_stats.rds"))
+
+
+df <- df |>
+  mutate(across(everything(), ~ (. - mean(.)) / sd(.)))
+
+langevitour(df)
+
+write_rds(df, here::here("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_data.rds"))
+

### 2025-06-17 scripts/two_nonlinear/09_gen_rm_lwd_mse.R

@@ -1,0 +1,146 @@
+### This script is to generate error three bin choices
+library(readr)
+library(quollr)
+library(dplyr)
+
+data_two_curvy <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_data.rds")
+data_two_curvy <- data_two_curvy |>
+  mutate(ID = row_number())
+
+tsne_two_curvy <- read_rds(file = "data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_30.rds") |>
+  mutate(ID = row_number())
+
+
+### Option 1
+
+two_curvy_model1 <- fit_highd_model(
+  highd_data = data_two_curvy,
+  nldr_data = tsne_two_curvy,
+  bin1 = 15,
+  q = 0.1,
+  benchmark_highdens = 0)
+
+df_bin_centroids_two_curvy1 <- two_curvy_model1$model_2d
+df_bin_two_curvy1 <- two_curvy_model1$model_highd
+
+error_rm_two_curvy1 <- data.frame(matrix(nrow = 0, ncol = 0))
+
+## To initialize benchmark values to remove low density hexagons
+benchmark_rm_hex_vec <- seq(0, 0.47, by=0.03)
+
+for (benchmark_rm_lwd in benchmark_rm_hex_vec) {
+
+  df_bin_centroids_two_curvy_high_dens <- df_bin_centroids_two_curvy1 |>
+    filter(std_counts > benchmark_rm_lwd)
+
+  df_bin_two_curvy_high_dens <- df_bin_two_curvy1 |>
+    filter(hexID %in% df_bin_centroids_two_curvy_high_dens$hexID)
+
+  ## Compute error
+  error_df <- glance(
+    model_2d = df_bin_centroids_two_curvy_high_dens,
+    model_highd = df_bin_two_curvy_high_dens,
+    highd_data = data_two_curvy) |>
+    mutate(benchmark_rm_lwd = round(benchmark_rm_lwd, 2),
+           a1 = paste0("a[1] == ", round(two_curvy_model1$hb_obj$a1, 2)),
+           bin1 = two_curvy_model1$hb_obj$bins[1],
+           bin2 = two_curvy_model1$hb_obj$bins[2],
+           b = bin1 * bin2,
[…]

### 2025-06-17 paper.qmd
Fixed two curvy projection dara in the script
Updated Figure plot title to right

@@
 
-We define the NLDR as a function $g\text{:}~ \mathbb{R}^{n\times p} \rightarrow \mathbb{R}^{n\times k}$, with (hyper-)parameters $\mathbfit{\theta}$.
+We define the NLDR as a function $g\text{:}~ \mathbb{R}^{n\times p} \rightarrow \mathbb{R}^{n\times k}$, with (hyper-)~parameters $\mathbfit{\theta}$.
 These parameters, $\mathbfit{\theta}$, depend on the choice of $g$, and can be considered part of model fitting in the traditional sense.
@@
                colour = "#000000") +
-  interior_annotation("a", sc_ltr_pos, cex = 1.2) +
+  interior_annotation("a", cex = 1.2) +
   theme(aspect.ratio = 1)
@@
     is_category = FALSE) +
-  interior_annotation(label = "b", position = sc_ltr_pos, cex = 1.2) +
+  interior_annotation(label = "b", cex = 1.2) +
   theme(aspect.ratio = 1,
@@
   is_category = FALSE) +
-  interior_annotation(label = "c", position = sc_ltr_pos, cex = 1.2) +
+  interior_annotation(label = "c", cex = 1.2) +
   theme(aspect.ratio = 1)

### 2025-06-17 paper.qmd
Removed TO DO

@@
 ---
-
-XXX TO DO:
-
-- Remove all the extraneous commented text (Done)
-- Check how \kD{} and \gD{} are used, and make sure it's sensible (changed in Rendering the model: Done)
-- Check spelling is consistently American (Done)
-- Check references (Done)
-- Add videos - make a list of the videos in the supplementary section, with the number of the figure that it corresponds too.
-This avoids having captions that are too long and awkward with raw URLs.
-And it keeps all of this supplementary material in one place.
-(Done)
-- Move the notation table to appendix (Done)
-- Code clean up: recommend that all of the calculations done with quollr are done with a separate script(s) and put in a scripts directory.
-Save the resulting data need to make plots and tables.
-Only code for the plots and tables is directly in the qmd in the final version.
 

### 2025-06-17 paper.qmd
Removed white space
figure captions in the two applications

--- 2025-06-17 paper.qmd
@@
 But NLDR views have an important role in data analysis because, if done well, they provide a concise visual (and conceptual) summary of \pD{} distributions.
-The NLDR methods and (hyper-)~parameter choices can create wildly different representations, making it difficult to decide which is best, or whether any or all are accurate or misleading.
+The NLDR methods and (hyper)parameter choices can create wildly different representations, making it difficult to decide which is best, or whether any or all are accurate or misleading.
 To help assess the NLDR and decide on which, if any, is the most reasonable representation of the structure(s) present in the \pD{} data, we have developed an algorithm to show the \gD{} NLDR model in the \pD{} space, viewed with a tour, a movie of linear projections.
@@
 
-- Remove all the extraneous commented text (Done)
-- Check how \kD{} and \gD{} are used, and make sure it's sensible (changed in Rendering the model: Done)
-- Check spelling is consistently American (Done)
-- Check references (Done)
+- Remove all the extraneous commented text (NOT Done)
+- Check how \kD{} and \gD{} are used, and make sure it's sensible
+- Check spelling is consistently American
+- Check references
 - Add videos - make a list of the videos in the supplementary section, with the number of the figure that it corresponds too.
@@
 And it keeps all of this supplementary material in one place.
-(Done)
-- Move the notation table to appendix (Done)
+- Move the notation table to appendix
 - Code clean up: recommend that all of the calculations done with quollr are done with a separate script(s) and put in a scripts directory.
@@
 #| fig-cap: "Eight different NLDR representations of the same data.
-Different methods and different parameter choices are used.
+Different techniques and different parameter choices are used.
 Researchers may have seen any of these in their analysis of this data, depending on their choice of method, or typical parameter choice.
@@
 @sec-method contains the details of the new methodology, including simulated data examples.
-In @sec-bestfit, we describe how to assess the best fit and identify the most accurate \gD{} layout based on the proposed model diagnostics.
-Curiosities and unexpected patterns discovered in NLDR results by examining the model in the data space are discussed in @sec-curiosities.
 Two applications illustrating the use of the new methodology for bioinformatics and image classification are in @sec-applications.
@@
 <!-- - Connection between NLDR and MDS-->
-Historically, low-dimensional (\kD{}) representations of high-dimensional (\pD{}) data have been computed using multidimensional scaling (MDS) [@kruskal1964], which includes principal components analysis (PCA) [@jolliffe2011] as a special case.
-(A contemporary comprehensive guide to MDS can be found in @borg2005.)
-<!-- Does JASA allow footnotes?
-If so this is a footnote, otherwise the punctuation has gone a bit rogue - can't have an entire sentence in brackets surely --> The \kD{} representation can be considered to be a layout of points in \kD{}   produced by an embedding procedure that maps the data from \pD{}.
+Historically, \kD{} representations of \pD{} data have been computed using multidimensional scaling (MDS) [@kruskal1964], which includes principal components analysis (PCA) [@jolliffe2011] as a special case.
+(A contemporary comprehensive guide to MDS can be found in @borg2005.) The \kD{} representation can be considered to be a layout of points in \kD{}   produced by an embedding procedure that maps the data from \pD{}.
 In MDS, the \kD{} layout is constructed by minimizing a stress function that differences distances between points in \pD{} with potential distances between points in \kD{}.
@@
 
-<!-- Non-expert comment: This to me is a leap from MDS to NLDR. Is NLDR an MDS? I don't think so, but I'm missing the link -->
-<!-- Suggested rewording: "To better capture specific structures potentially existing in \pD{}, many new methods for NLDR have emerged in recent years." -->
 Many new methods for NLDR have emerged in recent years, all designed to better capture specific structures potentially existing in \pD{}.
-Here we focus on five currently popular techniques: tSNE, UMAP, PHATE, TriMAP and PaCMAP.
-Both tNSE and UMAP can be considered to produce the \kD{} representation by minimizing the divergence between two distributions, where the distributions are modeling the inter-point distances.
+Here we focus on five currently popular techniques, tSNE, UMAP, PHATE, TriMAP and PaCMAP.
+tNSE and UMAP can be considered to produce the \kD{}   minimizing the divergence between two distributions, where the distributions are modeling the inter-point distances.
 PHATE, TriMAP and PaCMAP are examples of diffusion processes [@coifman2005] spreading to capture geometric shapes, that include both global and local structure.
@@
 We can also imagine that the positioning of points in \gD{}    represent the fitted values, that will have some prescribed position in \pD{}   that can be compared with their observed values.
-This is the conceptual framework underlying the more formal versions of factor analysis [@joreskog1969] and MDS.
+This is the conceptual framework underlying the more formal versions of factor analysis [@cfa69] and MDS.
 (Note that, for this thinking the full \pD{}   data needs to be available, not just the interpoint distances.)
 
+```{r}
+#| label: tbl-notation
+#| tbl-cap: "Summary of notation for describing new methodology."
+# Notation used in the paper
+
+notation_df <- read_csv("misc/notation.csv")
+
+# Create the table
+kable(notation_df, 
+      format = "latex", 
+      booktabs = TRUE, escape = FALSE) |>
+  kable_styling(position = "center", 
+                full_width = FALSE, 
+                font_size = 12) |>
+  row_spec(0, bold = TRUE) |>
+  column_spec(1:2, width = c("3cm", "12cm"))
+```
+
 We define the NLDR as a function $g\text{:}~ \mathbb{R}^{n\times p} \rightarrow \mathbb{R}^{n\times k}$, with (hyper-)parameters $\mathbfit{\theta}$.
-These parameters, $\mathbfit{\theta}$, depend on the choice of $g$, and can be considered part of model fitting in the traditional sense.
+The parameters, $\mathbfit{\theta}$, depend on the choice of $g$, and can be considered part of model fitting in the traditional sense.
 Common choices for $g$ include functions used in tSNE, UMAP, PHATE, TriMAP, PaCMAP, or MDS, although in theory any function that does this mapping is suitable.
+<!--Any input requirements for the data (such as normalization, or preprocessing through the use of PCA or similar) is considered part of the function $g$.-->
 
@@
 ```{r}
+#| label: two-curvy-training
+training_data_two_curvy <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_data.rds") |>
+  mutate(ID = row_number())
+```
+
+```{r}
+#| label: two-curvy-true-proj-data
+
+true_model_df <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_true_model.rds")
+wireframe_true_model <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_true_model_connections.rds")
+
+data_two_curvy <- training_data_two_curvy |>
+  select(-ID) |>
+  mutate(type = "data")
+
+true_model_two_curvy <- true_model_df |> 
+  mutate(type = "true model")
+```
+
+<!--tSNE applied for Two curvy data-->
+```{r}
+#| label: tsne-two-curvy
+tsne_two_curvy <- read_rds(file = "data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_30.rds") |>
+  mutate(ID = row_number()) |>
+  rename(c("emb1" = "tSNE1",
+           "emb2" = "tSNE2"))
+```
+
+```{r}
 #| label: fit_model-two-curvy
 
-tsne_two_curvy_scaled <- read_rds("data/two_nonlinear/tsne_two_curvy_scaled.rds")
-hex_grid_with_counts <- read_rds("data/two_nonlinear/hex_grid_with_counts.rds")
-hex_grid_nonempty <- read_rds("data/two_nonlinear/hex_grid_nonempty.rds")
-tr_from_to_df_two_curvy <- read_rds("data/two_nonlinear/tr_from_to_df_two_curvy.rds")
-df_bin_centroids_two_curvy <- read_rds("data/two_nonlinear/df_bin_centroids_two_curvy.rds")
+num_bins_x_two_curvy <- 24
+
+algo_obj_two_curvy <- fit_highd_model(
+  highd_data = training_data_two_curvy, 
+  nldr_data = tsne_two_curvy, 
+  bin1 = num_bins_x_two_curvy, 
+  q = 0.1, 
+  benchmark_highdens = 1)
+
+tsne_two_curvy_scaled <- algo_obj_two_curvy$nldr_obj$scaled_nldr
+tr_from_to_df_two_curvy <- algo_obj_two_curvy$trimesh_data
+df_bin_centroids_two_curvy <- algo_obj_two_curvy$model_2d
+df_bin_two_curvy <- algo_obj_two_curvy$model_highd
+hex_grid <- algo_obj_two_curvy$hb_obj$hex_poly
+counts_df <- algo_obj_two_curvy$hb_obj$std_cts
+
+hex_grid_with_counts <- left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hexID"))
+
+hex_grid_nonempty <- hex_grid |>
+  filter(hex_poly_id %in% df_bin_centroids_two_curvy$hexID)
+
+bin_width <- algo_obj_two_curvy$hb_obj$a1
 
@@
                                   aes(x = emb1, y = emb2)) +
-  geom_point(alpha = 0.5, colour = clr_choice, size = 0.5)
+  geom_point(alpha = 0.5, color = clr_choice, size = 0.5)
 
@@
 
+# Second plot (identical to base except for xlim/ylim)
+nldr_two_curvy2 <- nldr_two_curvy_original +
+  interior_annotation("a", sc_ltr_pos, cex = 1.5)
+
+# Third plot with aspect ratio and different annotation
+nldr_two_curvy_original2 <- nldr_two_curvy_original +
+  theme(aspect.ratio = 1) +
+  interior_annotation("a1", sc_ltr_pos)
+
 hex_grid_two_curvy <- ggplot(
@@
   aes(x = x, y = y)) +
-  geom_polygon(colour = "grey70", 
+  geom_polygon(color = "grey70", 
                aes(group = hex_poly_id), 
@@
              aes(x = emb1, y = emb2), 
-             alpha = 0.5, size = 0.5, colour = clr_choice) +
+             alpha = 0.5, size = 0.5, color = clr_choice) +
   xlim(sc_xlims) + ylim(sc_ylims) +
@@
       group = hex_poly_id, 
-      fill = std_counts), colour = "grey70", linewidth=0.2) +
+      fill = std_counts), color = "grey70", linewidth=0.2) +
   geom_point(data = tsne_two_curvy_scaled,
@@
 
+# df_bin_centroids_two_curvy <- df_bin_centroids_two_curvy |>
+#   dplyr::filter(std_counts > 0)
+
 hex_grid_nonempty_two_curvy <- ggplot(
@@
              aes(x = c_x, y = c_y), 
-             colour = "#000000",
+             color = "#000000",
              size = 1) +
@@
              ),
-             colour = "#636363",
+             color = "#636363",
              alpha = 0.5,
@@
              ),
-             colour = clr_choice,
+             color = clr_choice,
              alpha = 0.5,
@@
 ## Compute error
-error_df_two_curvy_abs <- read_rds("data/two_nonlinear/error_df_two_curvy_abs.rds")
+error_df_two_curvy_abs <- augment(
+  highd_data = training_data_two_curvy,
+  model_2d = df_bin_centroids_two_curvy,
+  model_highd = df_bin_two_curvy)
+
+error_df_two_curvy_abs <- error_df_two_curvy_abs |>
+  mutate(sqrt_row_wise_total_error = sqrt(row_wise_total_error))
+
+error_df_two_curvy_abs <- error_df_two_curvy_abs |>
+  bind_cols(tsne_two_curvy_scaled |>
+              select(-ID))
 
@@
   )
+
+# error_plot_two_curvy_hist <- 
+#   ggplot(error_df_two_curvy_abs) +
+#   geom_histogram(aes(x=sqrt_row_wise_total_error, y=..density..)) +
+#   geom_density(aes(x=sqrt_row_wise_total_error, y=..density..), colour="red") +
+#   #xlab(expression(group("|", e[hj], "|"))) +
+#   xlab(expression(e[hj])) +
+#   ylab("") +
+#   theme_bw()
 
@@
 ```{r}
+#| label: prep-data-for-proj-tsne-best
+
+df_bin_two_curvy <- df_bin_two_curvy |>
+  select(-hexID) |>
+  mutate(type = "model")
+
+# Apply the scaling
+df_model_data_two_curvy <- bind_rows(data_two_curvy, true_model_two_curvy, df_bin_two_curvy)
+
+scaled_two_curvy <- scale_data_manual(df_model_data_two_curvy, "type") |>
+  as_tibble()
+
+scaled_two_curvy_data <- scaled_two_curvy |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_two_curvy_data_true_model <- scaled_two_curvy |>
+  filter(type == "true model") |>
+  select(-type)
+
+scaled_two_curvy_data_model <- scaled_two_curvy |>
+  filter(type == "model") |>
+  select(-type)
+
+```
+
+```{r}
+#| label: langevitour-true-two-curvy-proj
+#| eval: false
+df_model_data_two_curvy_filtered <- bind_rows(true_model_two_curvy, data_two_curvy)
+
+langevitour::langevitour(df_model_data_two_curvy_filtered[1:(length(df_model_data_two_curvy_filtered)-1)],
+                         lineFrom = wireframe_true_model$from,
+                         lineTo = wireframe_true_model$to,
+                         group = factor(df_model_data_two_curvy_filtered$type,
+                                        c("data", "true model")),
+                         levelColors = c(clr_choice, "#000000"))
+```
+
+```{r}
+#| label: langevitour-best-fit-two-curvy-proj-tsne
+#| eval: false
+
+df_model_data_two_curvy_filtered <- bind_rows(df_bin_two_curvy, data_two_curvy)
+
+langevitour::langevitour(df_model_data_two_curvy_filtered[1:(length(df_model_data_two_curvy_filtered)-1)],
+                         lineFrom = tr_from_to_df_two_curvy$from,
+                         lineTo = tr_from_to_df_two_curvy$to,
+                         group = factor(df_model_data_two_curvy_filtered$type,
+                                        c("data", "model")),
+                         levelColors = c(clr_choice, "#000000"))
+```
+
+```{r}
 #| label: best-fit-two-curvy-proj-tsne
@@
 ## First projection
-proj_obj1 <- read_rds("data/two_nonlinear/two_nonlinear_proj_obj1.rds")
+model_prj1 <- cbind(
+  c(0.37847,-0.06517,0.04231,0.01655),
+  c(-0.06466,-0.27052,0.22131,-0.15235))
+
+proj_obj1 <- quollr::get_projection(projection = model_prj1, 
+               proj_scale = 1, 
+               highd_data = scaled_two_curvy_data, 
+               model_highd = scaled_two_curvy_data_model, 
+               trimesh_data = tr_from_to_df_two_curvy, 
+               axis_param = list(limits = 0.5, 
+                                 axis_scaled = 2,
+                                 axis_pos_x = -0.5, 
+                                 axis_pos_y = -0.5, 
+                                 threshold = 0.03))
+
+## For true model
+
+projection_scaled <- model_prj1 * 1
+
+projected_true_model <- as.matrix(scaled_two_curvy_data_true_model) %*% projection_scaled
+
+projected_true_model_df <- projected_true_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+true_model_df_proj <- dplyr::left_join(
+  wireframe_true_model, 
+  projected_true_model_df, 
+  by = c("from" = "ID"))
+
+names(true_model_df_proj)[3:NCOL(true_model_df_proj)] <- paste0(names(projected_true_model_df)[-NCOL(projected_true_model_df)], "_from")
+
+true_model_df_proj <- dplyr::left_join(true_model_df_proj, projected_true_model_df, by = c("to" = "ID"))
+names(true_model_df_proj)[(2 + NCOL(projected_true_model_df)):NCOL(true_model_df_proj)] <- paste0(names(projected_true_model_df)[-NCOL(projected_true_model_df)], "_to")
+
+proj_obj1_true <- proj_obj1
+proj_obj1_true[["model_df"]] <- true_model_df_proj
+
+two_curvy_proj_true <- plot_proj(
+  proj_obj = proj_obj1_true, 
+  point_param = c(0.5, 0.5, clr_choice), # size, alpha, color
+  line_param = c(0.5, 0.5, "black"), # linewidth, alpha
+  plot_limits = c(-0.6, 0.6), 
+  axis_text_size = 4, 
+  is_category = FALSE) +
+  interior_annotation(label = "a", cex = 1.5) +
+  theme(aspect.ratio=1, 
+        legend.position = "none")
+
+# Changed the axis parameters
+axis_obj <- gen_axes(
+    proj = model_prj1 * 2,
+    limits = 0.9,
+    axis_pos_x = -0.4,
+    axis_pos_y = -0.4,
+    axis_labels = names(scaled_two_curvy_data),
+    threshold = 0.05)
+
+axes <- axis_obj$axes
+circle <- axis_obj$circle
+
+proj_obj1[["axes"]] <- axes
+proj_obj1[["circle"]] <- circle
 
@@
 ## Second projection
-
-proj_obj2 <- read_rds("data/two_nonlinear/two_nonlinear_proj_obj2.rds")
+model_prj2 <- cbind(
+  c(-0.38281,-0.04870,0.00619,-0.02451),
+  c(-0.01303,0.00634,-0.37421,0.09648))
+
+proj_obj2 <- get_projection(projection = model_prj2, 
+                            proj_scale = 1, 
+                            highd_data = scaled_two_curvy_data, 
+                            model_highd = scaled_two_curvy_data_model, 
+                            trimesh_data = tr_from_to_df_two_curvy, 
+                            axis_param = list(limits = 0.35, 
+                                              axis_scaled = 2,
+                                              axis_pos_x = -0.35, 
+                                              axis_pos_y = -0.35, 
+                                              threshold = 0.01))
+
+projected_true_model <- as.matrix(scaled_two_curvy_data_true_model) %*% model_prj2
+
+projected_true_model_df <- projected_true_model |>
+  tibble::as_tibble(.name_repair = "unique") |>
+  dplyr::rename(c("proj1" = "...1",
+                  "proj2" = "...2")) |>
+  dplyr::mutate(ID = dplyr::row_number())
+
+true_model_df_proj <- dplyr::left_join(
+  wireframe_true_model, 
+  projected_true_model_df, 
+  by = c("from" = "ID"))
+
+names(true_model_df_proj)[3:NCOL(true_model_df_proj)] <- paste0(names(projected_true_model_df)[-NCOL(projected_true_model_df)], "_from")
+
+true_model_df_proj <- dplyr::left_join(true_model_df_proj, projected_true_model_df, by = c("to" = "ID"))
+names(true_model_df_proj)[(2 + NCOL(projected_true_model_df)):NCOL(true_model_df_proj)] <- paste0(names(projected_true_model_df)[-NCOL(projected_true_model_df)], "_to")
+
+proj_obj2_true <- proj_obj2
+proj_obj2_true[["model_df"]] <- true_model_df_proj
+
+two_curvy_proj_true2 <- plot_proj(
+    proj_obj = proj_obj2_true, 
+    point_param = c(0.5, 0.5, clr_choice), # size, alpha, color
+    line_param = c(0.5, 0.5, "black"), # linewidth, alpha
+    plot_limits = c(-0.42, 0.42), 
+    axis_text_size = 4, 
+    is_category = FALSE) +
+  interior_annotation(label = "b", cex = 1.5) +
+  theme(aspect.ratio = 1, 
+        legend.position = "none")
+
+#Changed the axis parameters
+axis_obj <- gen_axes(
+    proj = model_prj2 * 2,
+    limits = 0.7,
+    axis_pos_x = -0.35,
+    axis_pos_y = -0.35,
+    axis_labels = names(scaled_two_curvy_data),
+    threshold = 0.02)
+
+axes <- axis_obj$axes
+circle <- axis_obj$circle
+
+proj_obj2[["axes"]] <- axes
+proj_obj2[["circle"]] <- circle
 
@@
   interior_annotation(label = "c", position = sc_ltr_pos, cex = 1.2) +
-  theme(aspect.ratio = 1)
-
-```
-
-To illustrate the method, we use $7\text{-}D$ simulated data, which we call the "nonlinear clusters".
+  theme(aspect.ratio = 1) #+
+  #geom_segment(
+  #  data = model_df, 
+  #  aes(
+  #    x = proj1_from, 
+  #    y = proj2_from, 
+  #    xend = proj1_to, 
+  #    yend = proj2_to), 
+  #  color = "#000000")
+
+```
+
+To illustrate the method, we use $7\text{-}D$ simulated data, which we call the "non-linear clusters".
 It is constructed by simulating two clusters, each consisting of $1000$ observations.
-The C-shaped cluster is generated from $\theta \sim U(\text{-}3\pi/2, 0)$, $X_1 = \sin(\theta)$, $X_2 \sim U(0, 2)$ (adding thickness to the C), $X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)$, $X_4 = \cos(\theta)$.
-The other cluster is from $X_1 \sim U(0, 2)$, $X_2 \sim U(0, 3)$, $\gamma \sim U(0, 0.5)$, $X_3 = \text{-}(X_1^3 + X_2) + \gamma$, and $X_4 \sim U(0, 2)$.
+The C-shaped cluster is generated from $\theta \sim U(-3\pi/2, 0)$, $X_1 = \sin(\theta)$, $X_2 \sim U(0, 2)$ (adding thickness to the C), $X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)$, $X_4 = \cos(\theta)$.
+The other cluster is from $X_1 \sim U(0, 2)$, $X_2 \sim U(0, 3)$, $\gamma \sim U(0, 0.5)$, $X_3 = -(X_1^3 + X_2) + \gamma$, and $X_4 \sim U(0, 2)$.
 We would consider $T=(X_1, X_2, X_3, X_4)$ to be the geometric structure (true model) that we hope to capture.
@@
 <!-- The remaining variables $X_4, X_5, X_6, X_7$ are all uniform error, with small variance.  -->
+
+```{r}
+#| label: fig-two-curvy-true-proj
+#| fig-cap: "Two \\gD{} projections from $4\\text{-}D$, for the two non-linear clusters data (blue points) and true model (black lines) are shown.
+The top cluster in (a) shows uniformly distributed data along a C-shape, while in the bottom cluster, the data is more concentrated in one corner of the curvy shape.
+Video of the langevitour animations are available at <>."
+#| fig-pos: H
+#| fig-width: 10
+#| fig-height: 5
+#| eval: false
+
+two_curvy_proj_true + two_curvy_proj_true2 +
+  plot_layout(ncol = 2)
+```
 
@@
 #| echo: false
-#| fig-cap: "Key steps for constructing the model on the tSNE layout ($k=2$): (a) data, (b) hexagon bins, (c) bin centroids, and (d) triangulated centroids. The two nonlinear clusters data is shown."
+#| fig-cap: "Key steps for constructing the model on the tSNE layout ($k=2$): (a) data, (b) hexagon bins, (c) bin centroids, and (d) triangulated centroids. The two non-linear clusters data is shown."
 #| fig-width: 12
@@
 b_2 = \Big\lceil1 +\frac{2[r_2 + q(1 + r_2)](b_1 - 1)}{\sqrt{3}(1 + 2q)}\Big\rceil.
-$${#eq-bin2}
+$$
 
@@
 ## hexagon binning to have regular hexagons
-
-hex_grid_temp <- read_rds("data/two_nonlinear/hex_grid_temp.rds")
-all_centroids_df_temp <- read_rds("data/two_nonlinear/all_centroids_df_tempp.rds")
-start_pt <- read_rds("data/two_nonlinear/start_pt.rds")
-d_rect <- read_rds("data/two_nonlinear/d_rect.rds")
-a1 <- read_rds("data/two_nonlinear/a2_data.rds")
-a2 <- read_rds("data/two_nonlinear/a1_data.rds")
-l <- read_rds("data/two_nonlinear/l_data.rds")
-rect_adj <- read_rds("data/two_nonlinear/rect_adj.rds")
+hb_obj_notation <- hex_binning(
+  nldr_obj = algo_obj_two_curvy$nldr_obj, 
+  bin1 = 7, 
+  q = 0.1)
+
+a1_temp <- hb_obj_notation$a1
+a2_temp <- hb_obj_notation$a2
+l_temp <- quad(a=3, b = 2 * a2_temp, c = -(a2_temp^2 + a1_temp^2))
+
+## Data set with all centroids
+all_centroids_df_temp <- hb_obj_notation$centroids
+hex_grid_temp <- hb_obj_notation$hex_poly
+
+hex_grid_temp40 <- hex_grid_temp |> 
+  filter(hex_poly_id == 40)
+
+start_pt <- all_centroids_df_temp |> 
+  filter(hexID == 1)
+d_rect <- tibble(x1min = 0, 
+                 x1max = 1,
+                 x2min = 0,
+                 x2max = diff(algo_obj_two_curvy$nldr_obj$lim2)/diff(algo_obj_two_curvy$nldr_obj$lim1))
+
+# To move the rectangle to ignore the overlap with the centroids
+# rect_adj <- tibble(x1 = 0.03, x2 = 0.03)
+rect_adj <- tibble(x1 = -0.03, x2 = 0.03)
+
+a1 <- tibble(x = all_centroids_df_temp$c_x[4],
+             xend = all_centroids_df_temp$c_x[5],
+             y = all_centroids_df_temp$c_y[21],
+             yend = all_centroids_df_temp$c_y[21],
+             label = expression(a[1]))
+a2 <- tibble(x = all_centroids_df_temp$c_x[25],
+             xend = all_centroids_df_temp$c_x[25],
+             y = all_centroids_df_temp$c_y[25],
+             yend = all_centroids_df_temp$c_y[33],
+             label = expression(a[2]))
+l <- tibble(x = hex_grid_temp40$x[2],
+            xend = hex_grid_temp40$x[3],
+            y = hex_grid_temp40$y[2],
+            yend = hex_grid_temp40$y[3],
+            label = expression(l))
 
@@
                  fill = "white", 
-                 colour = "#bdbdbd") +
+                 color = "#bdbdbd") +
     geom_point(data = all_centroids_df_temp, aes(
@@
       y = c_y), 
-      colour = "#31a354", size = 0.9) +
+      color = "#31a354", size = 0.9) +
     geom_point(data = start_pt, aes(x = c_x, 
                                     y = c_y), 
-               colour = "black") + 
+               color = "black") + 
     geom_rect(data=d_rect, 
@@
               fill = "white", 
-              colour = "black", 
+              color = "black", 
               alpha = 0, 
@@
                                ends = "both"), 
-                 colour = "black")+
+                 color = "black")+
     annotate("text", x=0.5, y=-0.36, 
-             label = expression(r[1]), colour = "black", size = 8) +
+             label = expression(r[1]), color = "black", size = 8) +
     geom_segment(data=d_rect, aes(
@@
                        ends = "both"), 
-                 colour = "black")+ 
+                 color = "black")+ 
     annotate("text", x=-0.3, y=0.4, 
-             label = expression(r[2]), colour = "black", size = 8) +
+             label = expression(r[2]), color = "black", size = 8) +
     geom_segment(data = a1, aes(
@@
         ends = "both"), 
-        colour = "black")+ # a1 = 0.2087578
+        color = "black")+ # a1 = 0.2087578
     annotate("text", 
@@
              label = expression(a[1]), 
-             colour = "black",
+             color = "black",
              vjust = 1.2, size = 8) +
@@
                                ends = "both"), 
-      colour = "black") + # a2 = 0.1807896
+      color = "black") + # a2 = 0.1807896
     annotate("text", x=a2$x, y=(a2$y+a2$yend)/2, 
              label = expression(a[2]), 
-             colour = "black", hjust=-0.2, size = 8) +
+             color = "black", hjust=-0.2, size = 8) +
     annotate("text", x=-0.18, y=-0.24, 
       label = expression(group("(", list(s[1], s[2]), ")")),
-      colour = "black", size = 8) +
+      color = "black", size = 8) +
   geom_segment(data = l, aes(
@@
                                ends = "both"), 
-      colour = "black") + 
+      color = "black") + 
     annotate("text", x=l$x + 0.03, y=(l$y+l$yend)/2, 
              label = expression(l), 
-             colour = "black", hjust=-0.2, size = 8) +
+             color = "black", hjust=-0.2, size = 8) +
   coord_equal()
@@
 This can be defined using the function $u: \mathbb{R}^{n\times 2} \rightarrow \mathbb{R}^{m\times 2}$, where
-$$u(i) = \arg\min_{j = 1, \dots, b} \sqrt{(y_{i1} - C^{(2)}_{j1})^2 + (y_{i2} - C^{(2)}_{j2})^2},$$ maps observation $i$ into $H_h = \{i| u(i) = h\}$.
-
-<!-- I'm not thrilled about H_h but I guess it can stay because changing it everywhere is a pain -->
+$u(i) = \arg\min_{j = 1, \dots, b} \sqrt{(y_{i1} - C^{(2)}_{j1})^2 + (y_{i2} - C^{(2)}_{j2})^2}$, mapping observation $i$ into $H_h = \{i| u(i) = h\}$.
 
@@
 
+<!-- XXX How are you doing this? Do you check the bounds of the hexagon? Or do you use distance to centroid? -->
+
 #### Indicating neighborhood
@@
 Delaunay triangulation [@lee1980;@alb2024] is used to connect points so that edges indicate neighboring observations, in both the NLDR layout (@fig-NLDR-two-curvy (d)) and the \pD{} model representation.
-When the data has been binned the triangulation connects centroids.
-The edges preserve the neighborhood information from the \kD{} representation when the model is lifted into \pD{}.
+When the data has been binned the triangulation connectd centroids.
+The edges preserve the neighborhood information when the model is lifted into \pD{}.
+
+When shapes are non-linear in the NLDR layout, some edges could be long.
+It can also happen that distant centroids can be connected, particularly if clustering is present, which can result in long line segments.
+In order to generate a smooth surface in \gD{}, these long line segments should be removed when tuning the model fit.
 
@@
 
-The last step is to lift the \gD{} model into \pD{} by computing \pD{} vectors that represent bin centroids.
-We use the \pD{} mean of the points in a given hexagon, $H_h$, denoted $C_{h}^{(p)}$, to map the centroid $C_{h}^{(2)} = (c_{h1}, c_{h2})$ to a point in \pD{}.
+The last step is to lift the \kD{} model into \pD{} by computing \pD{} vectors that represent bin centroids.
+We use the \pD{} mean of the points in $H_h$ to map the centroid $C_{h}^{(2)} = (c_{h1}, c_{h2})$ to a point in \pD{}.
 Let the \pD{} mean be
@@
 $$C_{h}^{(p)} = \frac{1}{n_h}\sum_{i =1}^{n_h} x_i, h = {1, \dots, b; n_h > 0}.$$
+Furthermore, line segments that exist in the \kD{} model generate line segments in \pD{} by connecting the \pD{} means of the corresponding \kD{} bin centroids.
+If additional long edges need to be removed, compute the edges in \pD{} and pruned any detected long edges to improve the accuracy.
+Once pruned, re-plot the \gD{} view to ensure it accurately captures the data.
 
@@
 
-hex_grid_with_counts_two_curvy2 <- read_rds("data/two_nonlinear/two_nonlinear_hex_grid_with_counts_two_curvy2.rds")
-tr_from_to_df_two_curvy2 <- read_rds("data/two_nonlinear/two_nonlinear_tr_from_to_df_two_curvy2.rds")
+num_bins_x_two_curvy <- 15
+
+algo_obj_two_curvy2 <- fit_highd_model(
+  highd_data = training_data_two_curvy, 
+  nldr_data = tsne_two_curvy, 
+  bin1 = num_bins_x_two_curvy, 
+  q = 0.1, 
+  benchmark_highdens = 5)
+
+tr_from_to_df_two_curvy2 <- algo_obj_two_curvy2$trimesh_data
+df_bin_centroids_two_curvy2 <- algo_obj_two_curvy2$model_2d
+df_bin_two_curvy2 <- algo_obj_two_curvy2$model_highd
+
+hex_grid_two_curvy2 <- algo_obj_two_curvy2$hb_obj$hex_poly
+counts_df_two_curvy2 <- algo_obj_two_curvy2$hb_obj$std_cts
+
+hex_grid_with_counts_two_curvy2 <- left_join(hex_grid_two_curvy2, counts_df_two_curvy2, by = c("hex_poly_id" = "hexID"))
 
@@
         fill = std_counts), 
-        colour = "grey70", 
+        color = "grey70", 
         linewidth=0.2) +
@@
              ),
-             colour = "#636363",
+             color = "#636363",
             alpha = 0.5,
@@
 
-hex_grid_with_counts_two_curvy3 <- read_rds("data/two_nonlinear/two_nonlinear_hex_grid_with_counts_two_curvy3.rds")
-tr_from_to_df_two_curvy3 <- read_rds("data/two_nonlinear/two_nonlinear_tr_from_to_df_two_curvy3.rds")
+num_bins_x_two_curvy <- 48
+
+algo_obj_two_curvy3 <- fit_highd_model(
+    highd_data = training_data_two_curvy, 
+    nldr_data = tsne_two_curvy, 
+    bin1 = num_bins_x_two_curvy, 
+    q = 0.1, 
+    benchmark_highdens = 5)
+
+tr_from_to_df_two_curvy3 <- algo_obj_two_curvy3$trimesh_data
+df_bin_centroids_two_curvy3 <- algo_obj_two_curvy3$model_2d
+df_bin_two_curvy3 <- algo_obj_two_curvy3$model_highd
+
+hex_grid_two_curvy3 <- algo_obj_two_curvy3$hb_obj$hex_poly
+counts_df_two_curvy3 <- algo_obj_two_curvy3$hb_obj$std_cts
+
+hex_grid_with_counts_two_curvy3 <- left_join(hex_grid_two_curvy3, counts_df_two_curvy3, by = c("hex_poly_id" = "hexID"))
 
@@
         fill = std_counts), 
-        colour = "grey70", linewidth=0.2) +
+        color = "grey70", linewidth=0.2) +
   geom_point(data = tsne_two_curvy_scaled,
@@
 
-<!-- In Figure 4, is the "sparse space" in the middle of each cluster (i.e.
-the holes) or is it the gap between clusters?
-If it's between clusters, I'd argue tSNE is making this gap smaller than in reality, which I don't think the caption currently conveys.
--->
-
 ```{r}
@@
 #| fig-cap: "Lifting the fitted model into \\pD{}.
-Starting from a default tSNE layout (a) of nonlinear clusters data ($n =  2000$ and $p = 4$), from which the \\gD{} wireframe is constructed (a) and then lifted to \\fD{} (b, c).
+Starting from a default tSNE layout (a) of non-linear clusters data ($n =  2000$ and $p = 4$), from which the \\gD{} wireframe is constructed (a) and then lifted to $4\\text{-}D$ (b, c).
 The fit is reasonbly tight with the data, although it does not fully spread the full width.
 This highlights a key characteristic of tSNE, which compresses the data when applying tSNE.
-Furthermore, the sparse space observed in the middle of the tSNE layout is a result of (hyper-)~parameter choice.
-The link to the video of the langevitour animation is available in @tbl-links."
+Furthermore, the sparse space observed in the middle of the tSNE layout is a result of (hyper-)parameter choice.
+Video of the langevitour animation is available at <>."
 #| fig-width: 8
@@
 
+#nldr_two_curvy2 + 
   trimesh_two_curvy_removed1_tsne_with_data +
@@
 
-Observations are associated with their bin center, $C_{h}^{(p)}$, which are also considered to be the *fitted values*. These can also be denoted as $\widehat{X}$. 
+Observations are associated with their bin center, $C_{h}^{(p)}$, which are also considered to be the *fitted values*.
+These can also be denoted as $\widehat{X}$.
+<!--The fitted values of the points in $H_h$ refers to the \pD{} mapping $C_{h}^{(p)}$ of the corresponding \kD{} model point $C_{h}^{(2)}$.-->
+
 <!--#### Error-->
@@
 
-where $n$ is the number of observations, $b$ is the number of bins, $n_h$ is the number of observations in $h^{th}$ bin, $p$ is the number of variables and $\mathbfit{x}_{hij}$ is the $j^{th}$ dimensional data of $i^{th}$ observation in $h^{th}$ hexagon.
+where $n$ is the number of observations, $b$ is the number of bins, $n_h$ is the number of observations in $h^{th}$ bin, $p$ is the number of variables, $\mathbfit{x}_{hij}$ is the $j^{th}$ dimensional data of $i^{th}$ observation in $h^{th}$ hexagon.
 We can consider $e_{hj} = \sqrt{\sum_{j = 1}^{p} (\mathbfit{x}_{hij} - C^{(p)}_{hj})^2}$ to be the residual for each observation.
@@
 
-error_two_curvy_umap <- read_rds("data/two_nonlinear/error_two_curvy_umap.rds")
+error_two_curvy_umap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_tsne.rds")
+
+## Find the minimum MSE when have duplicate a1
+error_two_curvy_umap <- error_two_curvy_umap |>
+  group_by(a1) |>
+  filter(bin1 == min(bin1)) |>
+  ungroup() |>
+  mutate(prop_dens = 1/(b*side_length^2))
+
+base_line_prop_dens <- error_two_curvy_umap |>
+  filter(a1 == min(a1)) |>
+  pull(prop_dens)
+
+error_two_curvy_umap <- error_two_curvy_umap |>
+  mutate(prop_comp = prop_dens/base_line_prop_dens)
 
@@
   geom_vline(xintercept = 0.09,
-           colour = "#d95f02", linetype="dashed", 
+           color = "#d95f02", linetype="dashed", 
            linewidth=1) +
   geom_vline(xintercept = 0.05,
-             colour = "#1f78b4", 
+             color = "#1f78b4", 
              linetype="solid",
@@
   geom_vline(xintercept = 0.03, linetype="dotted",
-             colour = "#1b9e77", linewidth=1) +
+             color = "#1b9e77", linewidth=1) +
   geom_line(linewidth = 0.5) + 
@@
   geom_vline(xintercept = 0.03, linetype="dotted",
-             colour = "#1b9e77", linewidth=1) +
+             color = "#1b9e77", linewidth=1) +
   geom_line(linewidth = 0.5) +
@@
 
+error_two_curvy_umap <- error_two_curvy_umap |>
+  mutate(prop_bins = b_non_empty/b)
+
 a1_m_two_curvy <- ggplot(error_two_curvy_umap,
@@
 
-<!-- Sparse end with dense corner to me is very hard to understand - sparse and dense are pretty much antonyms -->
 ```{r}
 #| label: fig-p-d-error-in-2d-two-curvy
-#| fig-cap: "The \\fD{} model error in \\gD{} layout.
-Color indicates error ($e_{hj}$), dark color indicating high error and light indicates low error.
-Most large errors are distributed near the sparse end of the nonlinear cluster and most small errors are distributed near the dense corners."
+#| fig-cap: "The $4\\text{-}D$ model error in \\gD{} layout.
+Color indicates error ($e_{hj}$), dark colour indicating high error and light indicates low error.
+Most large errors are distributed near the sparse end of the non-linear cluster with dense corner."
 #| out-width: 80%
@@
 
-<!-- Does this really need an entire subsection? It's only 3 sentences long. -->
-
-In this paradigm of model fitting, it is now possible to predict a new observation's value in the NLDR representation, independent of the choice of NLDR method.
-To do this, we first need to determine the closest bin centroid in \pD{}, $C^{(p)}_{h}$, to the new observation, which is,
-
-$$z = \arg\min_{j = 1, \dots, b} \sqrt{\sum_{v=1}^{p}(x_{iv} - C^{(p)}_{jv})^2}.$$
-
-Then, the predicted value is is the centroid of this bin in \gD{}, $C^{(2)}_{z} = (c_{z1}, c_{z2})$.
+A new benefit of this fitted model is that it allows us to now predict a new observation's value in the NLDR, for any method.
+The steps are to determine the closest bin centroid in \pD{}, $C^{(p)}_{h}$ and predict it to be the centroid of this bin in \gD{}, $C^{(2)}_{h}$.
+This can be written as, let $z(i) = \arg\min_{j = 1, \dots, b} \sqrt{\sum_{v=1}^{p}(x_{iv} - C^{(p)}_{jv})^2}$, then the new observation $i$ falls in the hexagon, $H_h = \{i| z(i) = h\}$ and the corresponding \kD{} bin centroids, $C_{h}^{(2)} = (c_{h1}, c_{h2})$.
+
+<!--prediction for true-model of two-curvy clust-->
+```{r}
+#| label: model-prediction-umap-original
+
+predict_tsne_two_curvy <- read_rds(file = "data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_umap_predict_true.rds")
+
+predict_two_curvy_obj <- gen_scaled_data(
+    nldr_data = predict_tsne_two_curvy)
+
+predict_tsne_two_curvy_scaled <- predict_two_curvy_obj$scaled_nldr
+
+plot_predict_umap_fun_two_curvy <- ggplot(
+  data = predict_tsne_two_curvy_scaled, 
+  aes(x = UMAP1, 
+      y = UMAP2)) +
+  geom_point(alpha = 0.5) +
+  interior_annotation("b", c(0.08, 0.93), cex = 1.5)
+```
+
+```{r}
+#| label: model-prediction-points-line
+#| eval: false
+
+predict_tsne_two_curvy_scaled <- predict_tsne_two_curvy_scaled |>
+  mutate(ID = row_number()) |>
+  mutate(type = "predict_from_umap")
+
+jittered_points_df <- jittered_points_df |>
+  mutate(type = "predict_from_model")
+
+predict_point_df <- bind_rows(
+  predict_tsne_two_curvy_scaled,
+  jittered_points_df
+)
+
+plot_predict_umap_connect_two_curvy <- ggplot(
+  data = predict_point_df, 
+  aes(x = UMAP1, 
+      y = UMAP2, 
+      group = ID)) +
+  geom_point(alpha = 0.5) +
+  geom_line(alpha = 0.3) +
+  interior_annotation("d", c(0.08, 0.93), cex = 1.5)
+```
+
+```{r}
+#| echo: false
+#| label: fig-predict-two-curvy
+#| fig-pos: H
+#| fig-width: 10
+#| fig-height: 10
+#| out-width: 80%
+#| fig-cap: "Comparison of prediction generated using the exiting `umap` R package's prediction method and our method: (a) A view of the true model in projection from $7\\text{-}D$, (b) predicted data from the `umap` prediction method, and (c) predicted data from our method."
+#| eval: false
+
+proj_plot_two_curvy_true + plot_predict_umap_fun_two_curvy +plot_predict_tsne_two_curvy +
+  plot_layout(ncol = 3)
+```
 
@@
 
-Default values are provided for each of these, but it is expected that the user will examine the RMSE for a range of choices.
-Choosing these parameters according to RMSE can be automated but it is recommended that the user examine the resulting model representation by overlaying it on the data in \pD{}.
+Default values are provided for each of these, but it is expected that the user will examine the MSE for a range of choices.
+Choosing these parameters according to MSE can be automated but it is recommended that the user examine the resulting model representation by overlaying it on the data in \pD{}.
 The next few subsections describe the calculation of default values, and the effect that different choices have on the model fit.
@@
 By default, this is at $s_1 = -q, s_2 = -qr_2$, where $q$ is the buffer bound the data.
-The choice of these values can have some effect on the distribution of bin counts which is seen in @fig-bins-two-curvy.
-The distribution of bin counts for $s_1$ varying between $\text{-}0.1\,\text{--}\,0.0$.
+The choice of these values can have some effect on the distribution of bin counts.
+@fig-bins-two-curvy illustrates this.
+The distribution of bin counts for $s_1$ varying between $-0.1-0.0$ is shown.
 Generally, a more uniform distribution among these possibilities would indicate that the bins are reliably capturing the underlying distribution of observations.
@@
 #| fig-pos: H
-#| fig-cap: "Hexbin density plots of tSNE layout of the nonlinear cluster data, using three different bin inputs: (a) $b = 240/98 \\text{ } (15, \\text{ }16)$, (b) $b = 720/215 \\text{ } (24, \\text{ }30)$, and (c) $b = 2496/549 \\text{ } (48, \\text{ }52)$.
+#| fig-cap: "Hexbin density plots of tSNE layout of the non-linear cluster data, using three different bin inputs: (a) $b = 240/98 \\text{ } (15, \\text{ }16)$, (b) $b = 720/215 \\text{ } (24, \\text{ }30)$, and (c) $b = 2496/549 \\text{ } (48, \\text{ }52)$.
 Color indicates standardized counts, dark indicating high count and light indicates low count.
 At the smallest bin size, the data structure is discontinuous, suggesting that there are too many bins.
-Using the RMSE of the model fit in $7\\text{-}D$ helps decide on a useful choice of number of bins."
+Using the MSE of the model fit in $7\\text{-}D$ helps decide on a useful choice of number of bins."
 #| fig-width: 9
@@
 The default number of bins $b=b_1\times b_2$ is computed based on the sample size, by setting $b_1=n^{1/3}$, consistent with the Diaconis-Freedman rule [@freedman1981].
-The value of $b_2$ is determined analytically by $b_1, q, r_2$ (@eq-bin2).
+The value of $b_2$ is determined analytically by $b_1, q, r_2$.
 Values of $b_1$ between $2$ and $b_1 = \sqrt{n/r_2}$ are allowed.
-@fig-param-two-curvy (a) shows the effect of different choices of $b_1$ on the RMSE of the fitted model.
-
-#### Measurement of capturing the data shape in \gD{}
-
-<!-- I'll be honest I've no idea what this section is -->
+@fig-param-two-curvy (a) shows the effect of different choices of $b_1$ on the MSE of the fitted model.
+
+#### Measurement of capturing the data shape in 2-D 
 The area of a hexagon is defined as $A = 3\sqrt{3}l^2/2$ where $l$ is the side length of the hexagon.
@@
 By default, when assessing the choice of $b_1$, the total number of bins is measured by the number of **non-empty** bins.
-This more accurately reflects the hexagon grid relative the RMSE than the full number of bins in the grid.
-It may be beneficial to remove low count bins also, such as when the data is clustered or stringy, or where the observed data is sparse.
+This more accurately reflects the hexagon grid relative the MSE than the full number of bins in the grid.
+It may also be beneficial to remove low count bins also, in the situation where data is clustered or stringy, where the observed data is sparse.
 In order to decide if this is necessary, you would examine the distribution of bin counts, or the density which puts the counts on a standard scale.
@@
 
-The benchmark value for removing low-density hexagons ranges between $0$ and $1$, where $0$ corresponds to no bins being removed, and $1$ to 100% of bins being removed.
-When analyzing how these benchmark values influence model performance, it's essential to observe the change in RMSE as the benchmark value increases (@fig-param-two-curvy (c)).
-The RMSE shows a gradual decrease as the benchmark value goes from $1$ to $0$.
+The benchmark value for removing low-density hexagons ranges between $0$ and $1$.
+When analyzing how these benchmark values influence model performance, it's essential to observe the change in MSE as the benchmark value increases (@fig-param-two-curvy (c)).
+The MSE shows a gradual decrease as the benchmark value goes from $1$ to $0$.
 Evaluating this rate of increase is important.
@@
 
-error_rm_two_curvy <- read_rds("data/two_nonlinear/error_rm_lwd_diff_bin.rds") |>
+error_rm_two_curvy <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_rm_lwd_diff_bin.rds") |>
   mutate(bin1  = as.factor(bin1))
@@
                          y = RMSE,
-                         colour = a1,
+                         color = a1,
                          linetype = a1)) + 
@@
 
-To understand how well the model fits the points whether it fits well, works better in some positions, or fails to match the overall pattern, it is important to link and brush the points with high model error in the \pD{} error plot, \gD{} layout, and the generated model overlay on the data in \pD{}.
-
-## Assessing the best fit, and hence most accurate \gD{} layout {#sec-bestfit}
-
-@fig-toy-rmse illustrates the approach to compare the fits for different representations and assess the strength of any fit.
+To understand how well the model fits the points whether it fits well, works better in some positions, or fails to match the overall pattern.
+It is important to link and brush the points with high model error in the \pD{} error plot, \gD{} layout, and the generated model overlay on the data in \pD{}.
+
+## Assessing the best fit, and hence most accurate \gD{} layout
+
+@fig-toy-mse illustrates the approach to compare the fits for different representations and assess the strength of any fit.
 What does it mean to be a best fit for this problem?
-Analysts use an NLDR layout to display the structure present in high-dimensional data in a convenient \gD{} display.
-It is a competitor to linear dimension reduction that can better represent nonlinear associations such as clusters.
+Analysts use an NLDR layout to display the structure present in high-dimensional data in a convenient 2D display.
+It is a competitor to linear dimension reduction that can better represent non-linear associations such as clusters.
 However, these methods can hallucinate, suggesting patterns that don't exist, and grossly exaggerate other patterns.
@@
 A particular pattern that we commonly see is that analysts tend to pick layouts with clusters that have big separations between them.
-When you examine their data in a tour, it is almost always that we see there are no big separations, and actually often the suggested clusters are not even present.
-While we don't expect that analysts include animated gifs of tours in their papers, we should expect that any \gD{} representation adequately indicates the clustering that is present, and honestly show lack of separation or lack of clustering when it doesn't exist.
+When you examine their data in a tour, it is almost always that we see there are no big separations, and actually the suggested clusters are not even present.
+While we don't expect that analysts include animated gifs of tours in their papers, we should expect that any 2D representation adequately indicates the clustering that is present, and honestly show lack of separation or lack of clustering when it doesn't exist.
 It is important for analysts to have tools to select the accurate representation not the pretty but wrong representation.
 
-To compare and assess a range of representations an analyst needs:
+To compare and assess a range of representations needs these items:
 
@@
 # and plot them on same aspect ratio
-tsne_two_curvy <- read_rds("data/two_nonlinear/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_30.rds")
+tsne_two_curvy <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_30.rds")
 
@@
 
-tsne_two_curvy <- read_rds("data/two_nonlinear/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_62.rds")
+tsne_two_curvy <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_62.rds")
 
@@
 
-umap_two_curvy <- read_rds("data/two_nonlinear/two_non_linear_diff_shaped_close_clusters_umap_n-neigbors_15_min-dist_0.1.rds")
+umap_two_curvy <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_umap_n-neigbors_15_min-dist_0.1.rds")
 
@@
 
-phate_two_curvy <- read_rds("data/two_nonlinear/two_non_linear_diff_shaped_close_clusters_phate_knn_5.rds")
+phate_two_curvy <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_phate_knn_5.rds")
 
@@
 
-trimap_two_curvy <- read_rds("data/two_nonlinear/two_non_linear_diff_shaped_close_clusters_trimap_n-inliers_12_n-outliers_4_n-random_3.rds")
+trimap_two_curvy <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_trimap_n-inliers_12_n-outliers_4_n-random_3.rds")
 
@@
 
-pacmap_two_curvy <- read_rds("data/two_nonlinear/two_non_linear_diff_shaped_close_clusters_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")
+pacmap_two_curvy <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")
 
@@
 
-error_two_curvy_umap <- read_rds("data/two_nonlinear/error_two_non_linear_diff_shaped_close_clusters_umap.rds")
-error_two_curvy_tsne <- read_rds("data/two_nonlinear/error_two_non_linear_diff_shaped_close_clusters_tsne.rds")
-error_two_curvy_phate <- read_rds("data/two_nonlinear/error_two_non_linear_diff_shaped_close_clusters_phate.rds")
-error_two_curvy_trimap <- read_rds("data/two_nonlinear/error_two_non_linear_diff_shaped_close_clusters_trimap.rds")
-error_two_curvy_pacmap <- read_rds("data/two_nonlinear/error_two_non_linear_diff_shaped_close_clusters_pacmap.rds")
-error_two_curvy_tsne2 <- read_rds("data/two_nonlinear/error_two_non_linear_diff_shaped_close_clusters_tsne2.rds")
+error_two_curvy_umap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_umap.rds")
+error_two_curvy_tsne <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_tsne.rds")
+error_two_curvy_phate <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_phate.rds")
+error_two_curvy_trimap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_trimap.rds")
+error_two_curvy_pacmap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_pacmap.rds")
+error_two_curvy_tsne2 <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_tsne2.rds")
 
@@
 ```{r}
-#| fig-cap: "Assessing which of the 6 NLDR layouts on the two nonlinear clusters data is the better representation using RMSE for varying binwidth ($a_1$).
-Color used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f).
+#| fig-cap: "Assessing which of the 6 NLDR layouts on the two non-linear clusters data is the better representation using RMSE for varying binwidth ($a_1$).
+Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f).
 Layout d is universally poor.
@@
 Layout a is the best choice."
-#| label: fig-toy-rmse
+#| label: fig-toy-mse
 #| fig-pos: H
@@
 
-## Curiosities about NLDR results discovered by examining the model in the data space {#sec-curiosities}
+## Curiosities about NLDR results discovered by examining the model in the data space
 
@@
 @fig-five-gau-projs a1, b1, c1 show the \gD{} layouts for (a) tSNE, (b) UMAP, and (c) PaCMAP, respectively.
-The default (hyper-)~parameters for each method are used.
+The default hyper-parameters for each method are used.
 In each layout we can see an accurate representation where all five clusters are visible, although with varying degrees of separation.
@@
 The curious detail is that the model for PaCMAP is closer to a *pancake* in shape in every cluster!
-This single projection only shows this in three of the five clusters but if we examine a different projection the other clusters exhibit the pancake also.
-While we don't know what exactly causes this, it is likely due to some ordering of points in the \gD{} PaCMAP layout that induces the flat model.
+This seingle projection only shows this in three of the five clusters but if we examine a different projection the other clusters exhibit the pancake also.
+While we don't know what exactly causes this it is likely due to some ordering of points in the \gD{} PaCMAP layout that induces the flat model.
 One could imagine that if the method used principal components on all the data, that it might induce some ordering that would produce the flat model.
@@
 
-tsne_gau_scaled <- read_rds("data/five_gau_clusters/tsne_gau_scaled.rds")
-tr_from_to_df_gau1 <- read_rds("data/five_gau_clusters/tr_from_to_df_gau1.rds")
+training_data_gau <- read_rds("data/five_gau_clusters/data_five_gau.rds")
+
+data_gau <- training_data_gau |> 
+  select(-ID) |>
+  mutate(type = "data")
+
+tsne_data_gau <- read_rds("data/five_gau_clusters/tsne_data_five_gau_71.rds")
+
+## Compute hexbin parameters
+num_bins_x_gau1 <- 18
+
+algo_obj_gau1 <- fit_highd_model(
+  highd_data = training_data_gau, 
+  nldr_data = tsne_data_gau, 
+  bin1 = num_bins_x_gau1, 
+  q = 0.1, 
+  benchmark_highdens = 5)
+
+tsne_gau_scaled <- algo_obj_gau1$nldr_obj$scaled_nldr
+tr_from_to_df_gau1 <- algo_obj_gau1$trimesh_data
+df_bin_centroids_gau1 <- algo_obj_gau1$model_2d
+df_bin_gau1 <- algo_obj_gau1$model_highd
 
@@
       yend = y_to),
-    colour = "#000000",
+    color = "#000000",
+    #alpha = 0.4,
     linewidth = 0.5) +
@@
 
+# df_bin_centroids_gau1 <- df_bin_centroids_gau1 |>
+#   dplyr::filter(std_counts > 0)
+
+df_b <- df_bin_gau1 |>
+  dplyr::filter(hexID %in% df_bin_centroids_gau1$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b <- df_b[match(df_bin_centroids_gau1$hexID, df_b$hexID),] |>
+  dplyr::select(-hexID)
+
+# Apply the scaling
+df_model_data <- bind_rows(data_gau, df_b)
+scaled_gau <- scale_data_manual(df_model_data, "type") |>
+  as_tibble()
+
+scaled_gau_data <- scaled_gau |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_gau_data_model <- scaled_gau |>
+  filter(type == "model") |>
+  select(-type)
+
+```
+
+```{r}
+#| label: langevitour-five-gau-tsne-proj
+#| eval: false
+
+df_model_data_n <- bind_rows(df_b, data_gau)
+
+langevitour::langevitour(df_model_data_n[1:(length(df_model_data_n)-1)],
+                         lineFrom = tr_from_to_df_gau1$from,
+                         lineTo = tr_from_to_df_gau1$to,
+                         group = factor(df_model_data_n$type,
+                                        c("data", "model")),
+                         levelColors = c(clr_choice, "#000000"),
+                         pointSize = 0.7)
 ```
@@
 #| label: five-gau-tsne-projs
-## First projection
-proj_obj1 <- read_rds("data/five_gau_clusters/tsne_gau_proj_obj1.rds")
-
-five_gau_proj_tsne_model1 <- plot_proj(
-  proj_obj = proj_obj1,
-  point_param = c(1.5, 0.05, clr_choice),
-  line_param = c(0.8, 0.4, "#000000"),
-  plot_limits = c(-0.89, 0.89),
-  axis_text_size = 5,
-  is_category = FALSE
-) +
+projection <- cbind(
+  c(0.10394,0.08379,-0.17868,0.32960),
+  c(0.35111,-0.01137,-0.09657,-0.16018))
+
+proj_obj1 <- get_projection(projection = projection, 
+               proj_scale = 2.5, 
+               highd_data = scaled_gau_data, 
+               model_highd = scaled_gau_data_model, 
+               trimesh_data = tr_from_to_df_gau1, 
+               axis_param = list(limits = 1,
+                                 axis_scaled = 2,
+                                 axis_pos_x = -0.72, 
+                                 axis_pos_y = -0.72, 
+                                 threshold = 0.1))
+
+projected_df <- proj_obj1$projected_df
+model_df <- proj_obj1$model_df
+axes <- proj_obj1$axes
+circle <- proj_obj1$circle
+
+five_gau_proj_tsne_model1 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_point(
+    #size = 0.5,
+    alpha = 0.05,
+    color = clr_choice) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#000000",
+    alpha = 0.4,
+    linewidth = 0.8) +
+  geom_segment(
+    data=axes, 
+    aes(x=x1, y=y1, xend=x2, yend=y2), 
+    colour="grey70") +
+  geom_text(
+    data=axes, 
+    aes(x=x2, y=y2),
+    label=rownames(axes), 
+    colour="grey50",
+    size = 5) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.89, 0.89)) +
+  ylim(c(-0.89, 0.89)) +
   interior_annotation("a2", 
@@
 ## Second projection
-
-proj_obj2 <- read_rds("data/five_gau_clusters/tsne_gau_proj_obj2.rds")
+projection <- cbind(
+  c(0.30943,-0.08452,-0.23005,-0.05284),
+  c(0.22463,-0.06682,0.32106,0.02452))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 3, 
+                            highd_data = scaled_gau_data, 
+                            model_highd = scaled_gau_data_model, 
+                            trimesh_data = tr_from_to_df_gau1, 
+                            axis_param = list(limits = 1.5,
+                                              axis_scaled = 2,
+                                              axis_pos_x = -1, 
+                                              axis_pos_y = -1, 
+                                              threshold = 0.05))
 
@@
 
-umap_gau_scaled <- read_rds("data/five_gau_clusters/umap_gau_scaled.rds")
-tr_from_to_df_gau1_umap <- read_rds("data/five_gau_clusters/tr_from_to_df_gau1_umap.rds")
+umap_data_gau <- read_rds("data/five_gau_clusters/umap_data_five_gau.rds")
+
+## Compute hexbin parameters
+num_bins_x_gau1 <- 63
+
+algo_obj_gau1 <- fit_highd_model(
+  highd_data = training_data_gau, 
+  nldr_data = umap_data_gau, 
+  bin1 = num_bins_x_gau1, 
+  q = 0.1, 
+  benchmark_highdens = 5)
+
+umap_gau_scaled <- algo_obj_gau1$nldr_obj$scaled_nldr
+tr_from_to_df_gau1 <- algo_obj_gau1$trimesh_data
+df_bin_centroids_gau1 <- algo_obj_gau1$model_2d
+df_bin_gau1 <- algo_obj_gau1$model_highd
 
@@
   geom_segment(
-    data = tr_from_to_df_gau1_umap,
+    data = tr_from_to_df_gau1,
     aes(
@@
       yend = y_to),
-    colour = "#000000",
+    color = "#000000",
     #alpha = 0.4,
@@
   )
+
+# df_bin_centroids_gau1 <- df_bin_centroids_gau1 |>
+#   dplyr::filter(std_counts > 0)
+
+df_b <- df_bin_gau1 |>
+  dplyr::filter(hexID %in% df_bin_centroids_gau1$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b <- df_b[match(df_bin_centroids_gau1$hexID, df_b$hexID),] |>
+  dplyr::select(-hexID)
+
+# Apply the scaling
+df_model_data <- bind_rows(data_gau, df_b)
+scaled_gau <- scale_data_manual(df_model_data, "type") |>
+  as_tibble()
+
+scaled_gau_data <- scaled_gau |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_gau_data_model <- scaled_gau |>
+  filter(type == "model") |>
+  select(-type)
+
+```
+
+```{r}
+#| label: langevitour-five-gau-umap-proj
+#| eval: false
+
+df_model_data_n <- bind_rows(df_b, data_gau)
+
+langevitour::langevitour(df_model_data_n[1:(length(df_model_data_n)-1)],
+                         lineFrom = tr_from_to_df_gau1$from,
+                         lineTo = tr_from_to_df_gau1$to,
+                         group = factor(df_model_data_n$type,
+                                        c("data", "model")),
+                         levelColors = c(clr_choice, "#000000"), pointSize = 0.7)
 ```
@@
 #| label: five-gau-umap-projs
-
-## First projection
-proj_obj1 <- read_rds("data/five_gau_clusters/umap_gau_proj_obj1.rds")
-
-five_gau_proj_umap_model1 <- plot_proj(
-  proj_obj = proj_obj1,
-  point_param = c(1.5, 0.05, clr_choice),
-  line_param = c(0.8, 0.4, "#000000"),
-  plot_limits = c(-0.89, 0.89),
-  axis_text_size = 5,
-  is_category = FALSE
-) +
+projection <- cbind(
+  c(0.10394,0.08379,-0.17868,0.32960),
+  c(0.35111,-0.01137,-0.09657,-0.16018))
+
+proj_obj1 <- get_projection(projection = projection, 
+               proj_scale = 2.5, 
+               highd_data = scaled_gau_data, 
+               model_highd = scaled_gau_data_model, 
+               trimesh_data = tr_from_to_df_gau1, 
+               axis_param = list(limits = 1,
+                                 axis_scaled = 2,
+                                 axis_pos_x = -0.72, 
+                                 axis_pos_y = -0.72, 
+                                 threshold = 0.1))
+
+projected_df <- proj_obj1$projected_df
+model_df <- proj_obj1$model_df
+axes <- proj_obj1$axes
+circle <- proj_obj1$circle
+
+five_gau_proj_umap_model1 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1, 
+      y = proj2)) +
+  geom_point(
+    alpha = 0.05,
+    color = clr_choice) +
+  geom_segment(
+    data = model_df, 
+    aes(
+      x = proj1_from, 
+      y = proj2_from, 
+      xend = proj1_to, 
+      yend = proj2_to), 
+    color = "#000000",
+    alpha = 0.4,
+    linewidth = 0.8) +
+  geom_segment(
+    data=axes, 
+    aes(x=x1, y=y1, xend=x2, yend=y2), 
+    colour="grey70") +
+  geom_text(
+    data=axes, 
+    aes(x=x2, y=y2, label=rownames(axes)), 
+    colour="grey50",
+    size = 5) +
+  geom_path(
+    data=circle, 
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.89, 0.89)) +
+  ylim(c(-0.89, 0.89)) +
   interior_annotation("b2", 
@@
 ## Second projection
-proj_obj2 <- read_rds("data/five_gau_clusters/umap_gau_proj_obj2.rds")
+projection <- cbind(
+  c(0.30943,-0.08452,-0.23005,-0.05284),
+  c(0.22463,-0.06682,0.32106,0.02452))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 3, 
+                            highd_data = scaled_gau_data, 
+                            model_highd = scaled_gau_data_model, 
+                            trimesh_data = tr_from_to_df_gau1, 
+                            axis_param = list(limits = 1.5,
+                                              axis_scaled = 2,
+                                              axis_pos_x = -1, 
+                                              axis_pos_y = -1, 
+                                              threshold = 0.05))
 
@@
 
-pacmap_gau_scaled <- read_rds("data/five_gau_clusters/pacmap_gau_scaled.rds")
-tr_from_to_df_gau1_pacmap <- read_rds("data/five_gau_clusters/tr_from_to_df_gau1_pacmap.rds")
+pacmap_data_gau <- read_rds("data/five_gau_clusters/pacmap_data_five_gau.rds")
+
+## Compute hexbin parameters
+num_bins_x_gau1 <- 28
+
+algo_obj_gau1 <- fit_highd_model(
+  highd_data = training_data_gau, 
+  nldr_data = pacmap_data_gau, 
+  bin1 = num_bins_x_gau1, 
+  q = 0.1, 
+  benchmark_highdens = 5)
+
+pacmap_gau_scaled <- algo_obj_gau1$nldr_obj$scaled_nldr
+tr_from_to_df_gau1 <- algo_obj_gau1$trimesh_data
+df_bin_centroids_gau1 <- algo_obj_gau1$model_2d
+df_bin_gau1 <- algo_obj_gau1$model_highd
 
@@
   geom_segment(
-    data = tr_from_to_df_gau1_pacmap,
+    data = tr_from_to_df_gau1,
     aes(
@@
   )
+
+# df_bin_centroids_gau1 <- df_bin_centroids_gau1 |>
+#   dplyr::filter(std_counts > 0)
+
+df_b <- df_bin_gau1 |>
+  dplyr::filter(hexID %in% df_bin_centroids_gau1$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b <- df_b[match(df_bin_centroids_gau1$hexID, df_b$hexID),] |>
+  dplyr::select(-hexID)
+
+# Apply the scaling
+df_model_data <- bind_rows(data_gau, df_b)
+scaled_gau <- scale_data_manual(df_model_data, "type") |>
+  as_tibble()
+
+scaled_gau_data <- scaled_gau |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_gau_data_model <- scaled_gau |>
+  filter(type == "model") |>
+  select(-type)
+```
+
+```{r}
+#| label: langevitour-five-gau-pacmap-proj
+#| eval: false
+
+df_model_data_n <- bind_rows(df_b, data_gau)
+
+langevitour::langevitour(df_model_data_n[1:(length(df_model_data_n)-1)],
+                         lineFrom = tr_from_to_df_gau1$from,
+                         lineTo = tr_from_to_df_gau1$to,
+                         group = factor(df_model_data_n$type,
+                                        c("data", "model")),
+                         levelColors = c(clr_choice, "#000000"), pointSize = 0.7)
 ```
@@
 ## First projection
-proj_obj1 <- read_rds("data/five_gau_clusters/pacmap_gau_proj_obj1.rds")
-
-five_gau_proj_pacmap_model1 <- plot_proj(
-  proj_obj = proj_obj1,
-  point_param = c(1.5, 0.05, clr_choice),
-  line_param = c(0.8, 0.4, "#000000"),
-  plot_limits = c(-0.89, 0.89),
-  axis_text_size = 5,
-  is_category = FALSE
-) +
+projection <- cbind(
+  c(0.10394,0.08379,-0.17868,0.32960),
+  c(0.35111,-0.01137,-0.09657,-0.16018))
+
+proj_obj1 <- get_projection(projection = projection, 
+               proj_scale = 2.5, 
+               highd_data = scaled_gau_data, 
+               model_highd = scaled_gau_data_model, 
+               trimesh_data = tr_from_to_df_gau1, 
+               axis_param = list(limits = 1,
+                                 axis_scaled = 2,
+                                 axis_pos_x = -0.72, 
+                                 axis_pos_y = -0.72, 
+                                 threshold = 0.1))
+
+projected_df <- proj_obj1$projected_df
+model_df <- proj_obj1$model_df
+axes <- proj_obj1$axes
+circle <- proj_obj1$circle
+
+five_gau_proj_pacmap_model1 <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1,
+      y = proj2)) +
+  geom_point(
+    alpha = 0.05,
+    color = clr_choice) +
+  geom_segment(
+    data = model_df,
+    aes(
+      x = proj1_from,
+      y = proj2_from,
+      xend = proj1_to,
+      yend = proj2_to),
+    color = "#000000",
+    alpha = 0.4,
+    linewidth = 0.8) +
+  geom_segment(
+    data=axes,
+    aes(x=x1, y=y1, xend=x2, yend=y2),
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2, label=rownames(axes)),
+    colour="grey50",
+    size = 5) +
+  geom_path(
+    data=circle,
+    aes(x=c1, y=c2), colour="grey70") +
+  coord_fixed() +
+  xlim(c(-0.89, 0.89)) +
+  ylim(c(-0.89, 0.89)) +
   interior_annotation("c2",
@@
 ## Second projection
-proj_obj2 <- read_rds("data/five_gau_clusters/pacmap_gau_proj_obj2.rds")
+projection <- cbind(
+  c(0.30943,-0.08452,-0.23005,-0.05284),
+  c(0.22463,-0.06682,0.32106,0.02452))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 3, 
+                            highd_data = scaled_gau_data, 
+                            model_highd = scaled_gau_data_model, 
+                            trimesh_data = tr_from_to_df_gau1, 
+                            axis_param = list(limits = 1.5,
+                                              axis_scaled = 2,
+                                              axis_pos_x = -1, 
+                                              axis_pos_y = -1, 
+                                              threshold = 0.05))
 
@@
 #| fig-cap: "NLDR's organise points in the \\gD{} layout in different ways, possibly misleadingly, illustrated using three layouts: (a) tSNE, (b) UMAP, (c) PaCMAP.
-The data has five Gaussian cluster in \\fD{}.
-The bottom row of plots shows a \\gD{} projection from a tour on \\fD{} revealing the differences generated by the layouts on the model fits.
+The data has five Gaussian cluster in $4\\text{-}D$.
+The bottom row of plots shows a \\gD{} projection from a tour on $4\\text{-}D$ revealing the differences generated by the layouts on the model fits.
 We would expect the model fit to be like that in (a2) where it is distinctly is separate for each cluster but like a hairball in each.
-This would indicate the distinct clusters, each being fully \\fD{}.
-With (c2), the curiousity is that the model is a \\gD{} pancake shape in \\fD{}, indicating that there is some ordering of points done by PaCMAP, posisbly along some principal component axes.
-The links to the videos of the langevitour animations are available in @tbl-links."
+This would indicate the distinct clusters, each being fully $4\\text{-}D$.
+With (c2), the curiousity is that the model is a \\gD{} pancake shape in $4\\text{-}D$, indicating that there is some ordering of points done by PaCMAP, posisbly along some principal component axes.
+Videos of the langevitour animations are available at <https://youtu.be/I-kxCwVfqiQ>, <https://youtu.be/gD1P01FUPyU>, and <https://youtu.be/MxJ_srOFQNk> respectively."
 
@@
 five_gau_proj_tsne_model1 + five_gau_proj_umap_model1 + five_gau_proj_pacmap_model1 + 
+  #five_gau_proj_tsne_model2 +
+  #five_gau_proj_umap_model2 + five_gau_proj_pacmap_model2 +
   plot_layout(guides = "collect", nrow = 2) &
@@
 
-@fig-one-dens_clust-error (a2, b2) shows a \gD{} layout for tSNE created using the default (hyper-)~parameters.
+@fig-one-dens_clust-error (a2, b2) shows \gD{} layout for tSNE created using the default hyper-parameters.
 One would expect to see a rectangular shape if the curved sheet is flattened, but the layout is triangular.
@@
 The tSNE behaviour is to squeeze the low density area of the data together into the layout.
-This is common in other NLDR methods also, which means analysts need to be aware that if their data is not sampled relatively uniformly, apparent closeness in the \gD{} may correspond to sparseness in \pD{}.
+This is common in other NLDR methods also, which means analysts needs to be aware the if their data is not sampled relatively uniformly apparent closeness in the \gD{} may correspond to sparseness in \pD{}.
+
+```{r}
+#| label: model-one-dens-clust
+
+one_c_shaped_data <- read_rds(here::here("data/one_c_shaped_dens_structure/one_c_shaped_dens_data.rds"))
+one_c_shaped_data <- one_c_shaped_data |>
+  mutate(ID = row_number())
+
+tsne_one_c_shaped <- read_rds(file = "data/one_c_shaped_dens_structure/one_c_shaped_dens_structure_tsne_perplexity_52.rds") |>
+  mutate(ID = row_number())
+
+num_bins_x_one_c_shaped <- 20
+
+algo_obj_dens_clust <- fit_highd_model(
+  highd_data = one_c_shaped_data, 
+  nldr_data = tsne_one_c_shaped, 
+  bin1 = num_bins_x_one_c_shaped, 
+  q = 0.1, 
+  benchmark_highdens = 1)
+
+tsne_one_c_shaped_scaled <- algo_obj_dens_clust$nldr_obj$scaled_nldr
+tr_from_to_df_dens_clust <- algo_obj_dens_clust$trimesh_data
+df_bin_centroids_dens_clust <- algo_obj_dens_clust$model_2d
+df_bin_dens_clust <- algo_obj_dens_clust$model_highd
+```
 
@@
 #| label: error-model-one-dens-clust
-
-error_df_one_curvy_abs_selected11 <- read_rds("data/one_c_shaped_dens_structure/error_df_one_curvy_abs_selected11.rds")
-
-error_df_one_curvy_abs_deselected11 <- read_rds("data/one_c_shaped_dens_structure/error_df_one_curvy_abs_deselected11.rds")
+## Compute error
+error_df_one_curvy_abs <- augment(
+  highd_data = one_c_shaped_data,
+  model_2d = df_bin_centroids_dens_clust,
+  model_highd = df_bin_dens_clust)
+
+error_df_one_curvy_abs <- error_df_one_curvy_abs |>
+  bind_cols(tsne_one_c_shaped_scaled |>
+              select(-ID))
+
+error_df_one_curvy_abs <- error_df_one_curvy_abs |>
+  mutate(sqrt_row_wise_total_error = sqrt(row_wise_total_error))
+
+# Compute density
+density_data <- density(error_df_one_curvy_abs$sqrt_row_wise_total_error)
+density_df <- data.frame(x = density_data$x, y = density_data$y)
+
+# Add density values to the original dataset
+error_df_one_curvy_abs <- error_df_one_curvy_abs %>%
+  mutate(density = approx(density_df$x, density_df$y, xout = sqrt_row_wise_total_error)$y)
+
+error_df_one_curvy_abs <- error_df_one_curvy_abs |>
+  mutate(error_cat_n = if_else(tSNE1 <= 0.25 & tSNE2 <= 0.25, "selected", "deselected")) |> ## high_error points
+  mutate(error_cat_n2 = if_else(tSNE2 >= 1.2, "selected", "deselected")) ## corner points
+
+error_df_one_curvy_abs_selected11 <- error_df_one_curvy_abs |>
+  dplyr::filter(error_cat_n == "selected")
+
+error_df_one_curvy_abs_deselected11 <- error_df_one_curvy_abs |>
+  dplyr::filter(error_cat_n == "deselected")
 
@@
           y = density)) +
-  geom_point(alpha=0.7, colour = "#d9d9d9") +
+  geom_point(alpha=0.7, color = "#d9d9d9") +
   geom_point(data = error_df_one_curvy_abs_selected11, 
@@
           y = density),
-      alpha=0.7, colour = "#800026", size = 3) +
+      alpha=0.7, color = "#800026", size = 3) +
   xlab(expression(e[hj])) +
@@
 
-error_df_one_curvy_abs_selected12 <- read_rds("data/one_c_shaped_dens_structure/error_df_one_curvy_abs_selected12.rds")
-
-error_df_one_curvy_abs_deselected12 <- read_rds("data/one_c_shaped_dens_structure/error_df_one_curvy_abs_deselected12.rds")
+error_df_one_curvy_abs_selected12 <- error_df_one_curvy_abs |>
+  dplyr::filter(error_cat_n2 == "selected")
+
+error_df_one_curvy_abs_deselected12 <- error_df_one_curvy_abs |>
+  dplyr::filter(error_cat_n2 == "deselected")
 
@@
           y = density)) +
-  geom_point(alpha=0.7, colour = "#d9d9d9") +
+  geom_point(alpha=0.7, color = "#d9d9d9") +
   geom_point(data = error_df_one_curvy_abs_selected12, 
@@
 plot_tsne_dens <- error_df_one_curvy_abs_deselected12 |>
-  ggplot(aes(x = emb1,
-             y = emb2)) +
-  geom_point(alpha=0.7, colour = "#d9d9d9") +
+  ggplot(aes(x = tSNE1,
+             y = tSNE2)) +
+  geom_point(alpha=0.7, color = "#d9d9d9") +
   geom_point(data = error_df_one_curvy_abs_selected12,
-             aes(x = emb1, y = emb2),
-             alpha=0.7, colour = "#800026", size = 3) +
+             aes(x = tSNE1, y = tSNE2),
+             alpha=0.7, color = "#800026", size = 3) +
   theme(aspect.ratio = 1) +
@@
 plot_tsne_dens_selected <- error_df_one_curvy_abs_deselected11 |>
-  ggplot(aes(x = emb1,
-             y = emb2)) +
-  geom_point(alpha=0.7, colour = "#d9d9d9") +
+  ggplot(aes(x = tSNE1,
+             y = tSNE2)) +
+  geom_point(alpha=0.7, color = "#d9d9d9") +
   geom_point(data = error_df_one_curvy_abs_selected11,
-             aes(x = emb1, y = emb2),
-             alpha=0.7, colour = "#800026", size = 3) +
+             aes(x = tSNE1, y = tSNE2),
+             alpha=0.7, color = "#800026", size = 3) +
   theme(aspect.ratio = 1) +
@@
 ```{r}
+#| label: prep-proj-one-dens-clust
+
+# Apply the scaling
+
+data_c_shaped <- one_c_shaped_data |>
+  select(-ID) |>
+  mutate(type = "data")
+
+df_bin_dens_clust <- df_bin_dens_clust |>
+  select(-hexID) |>
+  mutate(type = "model")
+
+df_model_data <- bind_rows(data_c_shaped, df_bin_dens_clust)
+scaled_c_shaped <- scale_data_manual(df_model_data, "type") |>
+  as_tibble()
+
+scaled_c_shaped_data <- scaled_c_shaped |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_c_shaped_data_model <- scaled_c_shaped |>
+  filter(type == "model") |>
+  select(-type)
+```
+
+```{r}
+#| label: langevitour-c-shaped-tsne-proj
+#| eval: false
+
+data_c_shaped_n <- data_c_shaped |>
+  select(-type) |>
+  mutate(type = factor(error_df_one_curvy_abs$error_cat_n2,
+                          levels=c("deselected", "selected")))
+
+df_model_data_n <- bind_rows(df_bin_dens_clust, data_c_shaped_n)
+
+langevitour::langevitour(df_model_data_n[1:(length(df_model_data_n)-1)],
+                         lineFrom = tr_from_to_df_dens_clust$from,
+                         lineTo = tr_from_to_df_dens_clust$to,
+                         group = factor(df_model_data_n$type,
+                                        c("deselected", "selected", "model")),
+                         levelColors = c('#d9d9d9', '#800026', "#000000"),
+                         pointSize = c(1.5, 3, 0))
+```
+
+```{r}
 #| label: proj-one-dens-clust
@@
 ## First projection
-
-projected_df <- read_rds("data/one_c_shaped_dens_structure/projected_df1.rds")
-model_df <- read_rds("data/one_c_shaped_dens_structure/model_df1.rds")
-axes <- read_rds("data/one_c_shaped_dens_structure/axes1.rds")
-circle <- read_rds("data/one_c_shaped_dens_structure/circle1.rds")
+projection <- cbind(
+  c(0.05096,0.15399,0.19736,0.05110),
+  c(0.14608,-0.16929,0.11291,-0.07160))
+
+proj_obj1 <- get_projection(projection = projection,
+                            proj_scale = 1,
+                            highd_data = scaled_c_shaped_data,
+                            model_highd = scaled_c_shaped_data_model,
+                            trimesh_data = tr_from_to_df_dens_clust,
+                            axis_param = list(limits = 0.4,
+                                              axis_scaled = 3,
+                                              axis_pos_x = -0.28,
+                                              axis_pos_y = -0.28,
+                                              threshold = 0.02))
+
+proj_obj1[["cluster"]] <- factor(error_df_one_curvy_abs$error_cat_n2,
+                          levels=c("deselected", "selected"))
+
+projected_df <- proj_obj1$projected_df
+model_df <- proj_obj1$model_df
+axes <- proj_obj1$axes
+circle <- proj_obj1$circle
+
+projected_df <- projected_df |>
+      dplyr::mutate(cluster = proj_obj1$cluster)
 
@@
           yend = proj2_to),
-        colour = "#000000",
+        color = "#000000",
         linewidth = 0.8,
@@
 
-projected_df <- read_rds("data/one_c_shaped_dens_structure/projected_df2.rds")
-model_df <- read_rds("data/one_c_shaped_dens_structure/model_df2.rds")
-axes <- read_rds("data/one_c_shaped_dens_structure/axes2.rds")
-circle <- read_rds("data/one_c_shaped_dens_structure/circle2.rds")
+proj_obj1[["cluster"]] <- factor(error_df_one_curvy_abs$error_cat_n,
+                          levels=c("deselected", "selected"))
+
+projected_df <- proj_obj1$projected_df
+model_df <- proj_obj1$model_df
+axes <- proj_obj1$axes
+circle <- proj_obj1$circle
+
+projected_df <- projected_df |>
+      dplyr::mutate(cluster = proj_obj1$cluster)
 
@@
 #| fig-cap: "The sparse end of the structure shows a high model error, while the dense end shows a low model error.
-The tSNE layout of the C-shaped structure, with dense points (a1-a3) are colored according to their \\fD{} model error.
+The tSNE layout of the C-shaped structure, with dense points (a1-a3) are colored according to their $4\\text{-}D$ model error.
 Darker colors represent a high error, while lighter colors indicate a low error.
-a2, a3 and b2, b3 present the results from linked plots that brushed the high \\fD{} error points in \\gD{} and \\fD{}.
-It helps in identifying the high \\fD{} model errors that occur due to the sparse end.
-The link to the video of the linked plots is available in @tbl-links."
+a2, a3 and b2, b3 present the results from linked plots that brushed the high $4\\text{-}D$ error points in \\gD{} and $4\\text{-}D$.
+It helps in identifying the high $4\\text{-}D$ model errors that occur due to the sparse end.
+Videos of the linked plots are available at <>."
 
@@
 
+<!-- To showcase the robustness of our method when applied to data with different characteristics, this section presents two case studies.
+The first application focuses on single-cell data, a domain where the high dimensionality, sparsity, and heterogeneity of gene expression measurements pose significant analytical challenges.
+Our approach facilitates the identification of the most accurate NLDR layout, enabling the discovery of cell populations with similar expression profiles.
+The second case study explores the MNIST hand-written digits dataset, which serves as a benchmark for evaluating the preservation of local structures, specifically with digit 1.-->
+
 ### PBMC3k {#sec-pbmc}
@@
 Clustering of single-cell data is used to identify groups of cells with similar expression profiles.
-NLDR often used to summarize the cluster structure.
+NLDR often used to summarise the cluster structure.
 Usually, NLDR does not use the cluster labels to compute the layout, but uses color to represent the cluster labels when it is plotted.
+<!-- The purpose of this example is to *illustrate how to use our method to help decide on an appropriate NLDR layout that accurately represents the data*.
+The cluster results of Human Peripheral Blood Mononuclear Cells (PBMC3k) in @chen2024 are examined.-->
 
 In this data there are $2622$ single cells and $1000$ gene expressions (variables).
-Following their pre-processing, different NLDR techniques were performed on the first nine principal components.
-@fig-NLDR-variety shows this data using a variety of methods, and different (hyper-)~parameters.
+Following their pre-processing, UMAP was performed on the first nine principal components.
+@fig-NLDR-variety shows this data using a variety of methods, and different (hyper-)parameters.
 You can see that the result is wildly different depending on the choices.
@@
 
-@fig-pbmc-rmse shows RMSE across a range of binwidths ($a_1$) for each of the layouts in @fig-NLDR-variety.
-The layouts were made using tSNE, UMAP, PHATE, PaCMAP, and TriMAP with various (hyper-)~parameter settings.
+<!-- UMAP layout with author's suggested parameter choice-->
+```{r}
+#| label: published-pbmc
+## Import data
+training_data_pbmc <- read_rds("data/pbmc3k/pbmc_pca_50.rds")
+names(training_data_pbmc) <- paste0("x", 1:50)
+
+training_data_pbmc <- training_data_pbmc[, 1:9] |>
+  mutate(ID = 1:NROW(training_data_pbmc))
+
+umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_30_min_dist_0.3.rds")
+pbmc_scaled_obj <- gen_scaled_data(
+  nldr_data = umap_pbmc)
+umap_pbmc_scaled <- pbmc_scaled_obj$scaled_nldr
+```
+
+<!-- The @fig-NLDR-variety (a) shows three well-separated clusters with big separations.
+However, as shown in @fig-model-pbmc-author-proj (a2), there is no big separation between three clusters in $9\text{-}D$.
+Therefore, the suggested UMAP representation (@fig-NLDR-variety (a)) does not accurately represent the structure of PBMC3k dataset.-->
+
+@fig-pbmc-mse shows RMSE across a range of binwidths ($a_1$) for each of the layouts in @fig-NLDR-variety.
+The layouts were made using tSNE, UMAP, PHATE, PaCMAP, and TriMAP with various (hyper-)parameter settings.
 Lines are color coded to match the color of the layouts shown on the right.
@@
 
+<!-- Layouts c, d, and e, which show small separations between clusters, are universally optimal.
+However, layout d performs well with smaller binwidths and poorly with larger binwidths.
+On the other hand, layout e performs well with larger binwidths.
+Layout c was selected for further analysis due to its better (hyper-)parameter selection for the same method of the published plot.-->
+
+<!-- XXX can we use a and e in these figures instead of a and b, so it is easier to describe relative to the first figure, and previous figure? -->
+
 To further assess the choices, we need to look at the model in the data space, by using a tour to show the wireframe model overlaid on the data in the $9\text{-}D$ space (@fig-model-pbmc-author-proj).
@@
 
+<!-- The visualization of the selected layout in the $9\text{-}D$ shows edges between the clusters (@fig-model-pbmc-author-proj (b2)).
+This supports the presence of small separations between clusters.
+Additionally, the data points are not uniformly distributed across the clusters, resulting in dense areas (@fig-model-pbmc-author-proj (b2)).
+Furthermore, the clusters shows non-linear shapes (@fig-model-pbmc-author-proj (b3)).-->
+
 <!--Fit the best model for author suggestion and compute error-->
@@
 #| label: hexbin-pbmc
-#| 
-umap_pbmc_scaled_with_cluster <- read_rds("data/pbmc3k/umap_pbmc_scaled_with_cluster.rds")
-
-umap_tr_from_to_df_pbmc <- read_rds("data/pbmc3k/umap_tr_from_to_df_pbmc.rds")
+## Compute hexbin parameters
+num_bins_x_pbmc <- 22
+
+algo_obj_pbmc <- fit_highd_model(
+  highd_data = training_data_pbmc, 
+  nldr_data = umap_pbmc, 
+  bin1 = num_bins_x_pbmc, 
+  q = 0.1, 
+  benchmark_highdens = 5)
+
+umap_pbmc_scaled <- algo_obj_pbmc$nldr_obj$scaled_nldr
+tr_from_to_df_pbmc <- algo_obj_pbmc$trimesh_data
+df_bin_centroids_pbmc <- algo_obj_pbmc$model_2d
+df_bin_pbmc <- algo_obj_pbmc$model_highd
+
+umap_pbmc_scaled_with_cluster <- umap_pbmc_scaled |>
+  mutate(cluster = if_else(UMAP1 <= 0.25, "cluster1", if_else(UMAP1 >= 0.625, "cluster2", "cluster3"))) |>
+  mutate(cluster = factor(cluster, levels = c("cluster1", "cluster2", "cluster3")))
 
@@
       y = UMAP2,
-      colour = cluster
+      color = cluster
     ),
-    alpha = 0.2
+    alpha = 0.2#,
+    #size = 0.1#,
+    #color = '#a65628'
   )  +
-  geom_segment(data = umap_tr_from_to_df_pbmc, 
+  geom_segment(data = tr_from_to_df_pbmc, 
                aes(
@@
 ```{r}
+#| label: prep-pbmc-umap-model-proj
+
+data_pbmc <- training_data_pbmc |> 
+  select(-ID) |>
+  mutate(type = "data")
+
+# df_bin_centroids_pbmc <- df_bin_centroids_pbmc |>
+#   dplyr::filter(std_counts > 0)
+
+df_b_pbmc <- df_bin_pbmc |>
+  dplyr::filter(hexID %in% df_bin_centroids_pbmc$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b_pbmc <- df_b_pbmc[match(df_bin_centroids_pbmc$hexID, df_b_pbmc$hexID),] |>
+  dplyr::select(-hexID)
+
+# Apply the scaling
+df_model_data_pbmc <- bind_rows(data_pbmc, df_b_pbmc)
+scaled_pbmc <- scale_data_manual(df_model_data_pbmc, "type") |>
+  as_tibble()
+
+scaled_pbmc_data <- scaled_pbmc |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_pbmc_data_model <- scaled_pbmc |>
+  filter(type == "model") |>
+  select(-type)
+```
+
+```{r}
+#| label: langevitour-pbmc-umap-author-proj
+#| eval: false
+
+data_pbmc_n <- data_pbmc |>
+  select(-type) |>
+  mutate(type = as.character(umap_pbmc_scaled_with_cluster$cluster))
+
+df_model_data_pbmc_n <- bind_rows(df_b_pbmc, data_pbmc_n)
+
+langevitour::langevitour(df_model_data_pbmc_n[1:(length(df_model_data_pbmc_n)-1)],
+                         lineFrom = tr_from_to_df_pbmc$from,
+                         lineTo = tr_from_to_df_pbmc$to,
+                         group = factor(df_model_data_pbmc_n$type,
+                                        c("cluster1", "cluster2", "cluster3", "model")),
+                         levelColors = c("#8dd3c7", "#fdb462", "#bebada", "#000000"))
+```
+
+```{r}
 #| label: pbmc-umap-model-proj
@@
 ## First projection
-proj_obj1 <- read_rds("data/pbmc3k/pbmc_umap_proj_obj1.rds")
+projection <- cbind(
+    c(-0.4545,-0.5533,-0.1235,0.3984,-0.2065,-0.3586,-0.0088,0.3572,-0.1368),
+    c(0.4576,-0.3157,0.4708,0.1410,-0.4479,-0.1685,0.3668,-0.2692,0.1332))
+
+proj_obj1 <- get_projection(projection = projection, 
+               proj_scale = 1.23, 
+               highd_data = scaled_pbmc_data, 
+               model_highd = scaled_pbmc_data_model, 
+               trimesh_data = tr_from_to_df_pbmc, 
+               axis_param = list(limits = 1,
+                                 axis_scaled = 1, 
+                                 axis_pos_x = -0.48, 
+                                 axis_pos_y = -0.48, 
+                                 threshold = 0.082))
+
+proj_obj1[["cluster"]] <- as.character(umap_pbmc_scaled_with_cluster$cluster)
 
@@
 ## Second projection
-proj_obj2 <- read_rds("data/pbmc3k/pbmc_umap_proj_obj2.rds")
+projection <- cbind(
+    c(0.0734,0.6138,-0.4671,0.2589,0.1189,0.0778,-0.0819,0.0308,-0.5561),
+    c(0.6625,0.0034,0.2412,0.3360,0.2463,0.1070,-0.2903,-0.4705,0.1293))
+
+proj_obj2 <- get_projection(projection = projection, 
+               proj_scale = 1.25, 
+               highd_data = scaled_pbmc_data, 
+               model_highd = scaled_pbmc_data_model, 
+               trimesh_data = tr_from_to_df_pbmc, 
+               axis_param = list(limits = 1.35, 
+                                 axis_scaled = 1, 
+                                 axis_pos_x = -0.75, 
+                                 axis_pos_y = -0.75, 
+                                 threshold = 0.12))
+
+proj_obj2[["cluster"]] <- as.character(umap_pbmc_scaled_with_cluster$cluster)
 
@@
 #| fig-cap: "Assessing which of the 8 NLDR layouts on the PBMC3k data  (shown in @fig-NLDR-variety) is the better representation using RMSE for varying binwidth ($a_1$).
-Color used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-h).
+Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-h).
 Layout f is universally poor.
@@
 Layout e is the best choice."
-#| label: fig-pbmc-rmse
+#| label: fig-pbmc-mse
 #| fig-pos: H
@@
 <!--best choice-->
+
+```{r}
+#| label: tsne-pbmc-read-best
+tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_30.rds")
+```
+
 <!--Fit the best model and compute error-->
@@
 #| label: tsne-num-bins-pbmc
-tsne_pbmc_scaled_best_with_cluster <- read_rds("data/pbmc3k/tsne_pbmc_scaled_best_with_cluster.rds")
-
-tsne_tr_from_to_df_pbmc <- read_rds("data/pbmc3k/tsne_tr_from_to_df_pbmc.rds")
+
+## Compute hexbin parameters
+num_bins_x_tsne_pbmc <- 22
+
+algo_obj_pbmc <- fit_highd_model(
+  highd_data = training_data_pbmc, 
+  nldr_data = tsne_pbmc, 
+  bin1 = num_bins_x_tsne_pbmc, 
+  q = 0.1, 
+  benchmark_highdens = 1)
+
+tsne_pbmc_scaled_best <- algo_obj_pbmc$nldr_obj$scaled_nldr
+tr_from_to_df_pbmc <- algo_obj_pbmc$trimesh_data
+df_bin_centroids_pbmc <- algo_obj_pbmc$model_2d
+df_bin_pbmc <- algo_obj_pbmc$model_highd
+
+tsne_pbmc_scaled_best_with_cluster <- tsne_pbmc_scaled_best |>
+  mutate(cluster = umap_pbmc_scaled_with_cluster$cluster)
 
@@
       y = tSNE2,
-      colour = cluster
+      color = cluster
     ),
-    alpha = 0.3
+    alpha = 0.3#,
+    #color = "#999999"
   ) +
-  geom_segment(data = tsne_tr_from_to_df_pbmc, 
+  geom_segment(data = tr_from_to_df_pbmc, 
                aes(
@@
 ```{r}
+#| label: prep-pbmc-tsne-model-proj-best
+
+df_b_pbmc <- df_bin_pbmc |>
+  dplyr::filter(hexID %in% df_bin_centroids_pbmc$hexID) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b_pbmc <- df_b_pbmc[match(df_bin_centroids_pbmc$hexID, df_b_pbmc$hexID),] |>
+  dplyr::select(-hexID)
+
+# Apply the scaling
+df_model_data_pbmc <- bind_rows(data_pbmc, df_b_pbmc)
+scaled_pbmc <- scale_data_manual(df_model_data_pbmc, "type") |>
+  as_tibble()
+
+scaled_pbmc_data <- scaled_pbmc |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_pbmc_data_model <- scaled_pbmc |>
+  filter(type == "model") |>
+  select(-type)
+```
+
+```{r}
+#| label: langevitour-pbmc-tsne-best-proj
+#| eval: false
+
+df_model_data_pbmc_n <- bind_rows(df_b_pbmc, data_pbmc_n)
+
+langevitour::langevitour(df_model_data_pbmc_n[1:(length(df_model_data_pbmc_n)-1)],
+                         lineFrom = tr_from_to_df_pbmc$from,
+                         lineTo = tr_from_to_df_pbmc$to,
+                         group = factor(df_model_data_pbmc_n$type,
+                                        c("cluster1", "cluster2", "cluster3", "model")),
+                         levelColors = c("#8dd3c7", "#fdb462", "#bebada", "#000000"))
+```
+
+```{r}
 #| label: pbmc-tsne-model-proj-best
@@
 ## First projection
-proj_obj1 <- read_rds("data/pbmc3k/pbmc_tsne_proj_obj1.rds")
+projection <- cbind(
+  c(-0.4545,-0.5533,-0.1235,0.3984,-0.2065,-0.3586,-0.0088,0.3572,-0.1368),
+  c(0.4576,-0.3157,0.4708,0.1410,-0.4479,-0.1685,0.3668,-0.2692,0.1332))
+
+proj_obj1 <- get_projection(projection = projection, 
+                            proj_scale = 1.23, 
+                            highd_data = scaled_pbmc_data, 
+                            model_highd = scaled_pbmc_data_model, 
+                            trimesh_data = tr_from_to_df_pbmc, 
+                            axis_param = list(limits = 1, 
+                                              axis_scaled = 1,
+                                              axis_pos_x = -0.48, 
+                                              axis_pos_y = -0.48, 
+                                              threshold = 0.082))
+
+proj_obj1[["cluster"]] <- as.character(umap_pbmc_scaled_with_cluster$cluster)
 
@@
 ## Second projection
-proj_obj2 <- read_rds("data/pbmc3k/pbmc_tsne_proj_obj2.rds")
+projection <- cbind(
+  c(0.0734,0.6138,-0.4671,0.2589,0.1189,0.0778,-0.0819,0.0308,-0.5561),
+  c(0.6625,0.0034,0.2412,0.3360,0.2463,0.1070,-0.2903,-0.4705,0.1293))
+
+proj_obj2 <- get_projection(projection = projection, 
+                            proj_scale = 1.25, 
+                            highd_data = scaled_pbmc_data, 
+                            model_highd = scaled_pbmc_data_model, 
+                            trimesh_data = tr_from_to_df_pbmc, 
+                            axis_param = list(limits = 1.35, 
+                                              axis_scaled = 1,
+                                              axis_pos_x = -0.75, 
+                                              axis_pos_y = -0.75, 
+                                              threshold = 0.12))
+
+proj_obj2[["cluster"]] <- as.character(umap_pbmc_scaled_with_cluster$cluster)
 
@@
 #| echo: false
-#| fig-cap: "Compare the published \\gD{} layout (@fig-pbmc-rmse a) and the \\gD{} layout selected (@fig-pbmc-rmse e) by RMSE plot (@fig-pbmc-rmse) from the UMAP and tSNE with default (hyper-)~parameters.
-The PBMC3k data ($n =  2622$) has three close clusters in $9\\text{-}D$.
-Two \\gD{} projection from a tour on $9\\text{-}D$ of the model fit with @fig-pbmc-rmse a ($a_1 = 0.06$, $b = 616/86 (22, 28)$) shows three-well separated clusters with big separations.
-On the other hand, the model fit with @fig-pbmc-rmse e ($a_1 = 0.06$, $b = 704/227 (22, 32)$) shows three-well separated clusters with small separations.
-Therefore, @fig-pbmc-rmse e is more reasonable than @fig-pbmc-rmse a.
-Also, the fitted models in $9\\text{-}D$ help to see some unobserved patterns of the data: (i) dense points, and (ii) nonlinear clusters.
-The links to the videos of langevitour animations are available in @tbl-links."
+#| fig-cap: "Compare the published \\gD{} layout (a) made with UMAP and the \\gD{} layout selected by RMSE plot (e) made by tSNE, both with default (hyper-)parameters.
+The two plots on the right show  projections from a tour, with the models overlaid.
+The published layout suggested three very separated clusters, but this is not present in the data.
+While there may be three clusters they are not well-separated.
+The difference in model fit also indicates this: the published layout a does not spread out fully into the point cloud like the model generated from layout e.
+This supports the choice that layout e is the better representation of the data, because it does not exaggerate separation between clusters."
 #| label: fig-model-pbmc-author-proj
@@
 
-The digit "1" of the MNIST dataset [@lecun1998] consists of $7877$ grayscale images of handwritten "1"s.
-Each image is $28 \times 28$ pixels which corresponds to $784$ variables.
-The first $10$ principal components, explaining $83\%$ of the total variation, are used.
-This data is essentially lies on a nonlinear manifold in the high dimensions, defined by the shapes that "1"s make in when sketched, meaning that some pixels are almost always dark and some almost always light.
+The digit "1" of the MNIST dataset [@lecun2010] consists of $7877$ grayscale images of handwritten "1"s.
+Each image is $28 \times 28$ pixels which corresponds to 784 variables.
+The first 10 principal components, explaining 83% of the total variation, are used.
+This data is essentially lies on a nonlinear manifold in the high dimensions, defined by the shapes that "1"s make in when sketched, menaing that some pixels are almost always dark and some almost always light.
 We expect that the best layout captures this type of structure and does not exhibit distinct clusters.
+<!-- The objective is to select a reasonable \gD{} layout, representing the non-linear structure of the digit 1 dataset in $10\text{-}D$ (@fig-clust-mnist (a)).
+-->
 
@@
 nldr_mnist1 <- tsne_mnist |>
-  ggplot(aes(x = emb1,
-             y = emb2))+
+  ggplot(aes(x = tSNE1,
+             y = tSNE2))+
   geom_point(alpha=0.1, size=1, colour='#984ea3') +
@@
 nldr_mnist6 <- tsne_mnist2 |>
-  ggplot(aes(x = emb1,
-             y = emb2))+
+  ggplot(aes(x = tSNE1,
+             y = tSNE2))+
   geom_point(alpha=0.1, size=1, colour='#999999') +
@@
 nldr_mnist2 <- umap_mnist |>
-  ggplot(aes(x = emb1,
-             y = emb2)) +
+  ggplot(aes(x = UMAP1,
+             y = UMAP2)) +
   geom_point(alpha=0.1, size=1, colour='#a65628') +
@@
 nldr_mnist3 <- phate_mnist |>
-  ggplot(aes(x = emb1,
-             y = emb2))+
+  ggplot(aes(x = PHATE1,
+             y = PHATE2))+
   geom_point(alpha=0.1, size=1, colour='#377eb8') +
@@
 nldr_mnist4 <- trimap_mnist |>
-  ggplot(aes(x = emb1,
-             y = emb2))+
+  ggplot(aes(x = TriMAP1,
+             y = TriMAP2))+
   geom_point(alpha=0.1, size=1, colour='#ff7f00') +
@@
 nldr_mnist5 <- pacmap_mnist |>
-  ggplot(aes(x = emb1,
-             y = emb2))+
+  ggplot(aes(x = PaCMAP1,
+             y = PaCMAP2))+
   geom_point(alpha=0.1, size=1, colour='#e41a1c') +
@@
 ```{r}
-#| fig-cap: "Assessing which of the 6 NLDR layouts on the MNIST digit 1 data is the better representation using RMSE for varying binwidth ($a_1$).
-Color used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f).
-All the layouts appear to be very similar.
+#| fig-cap: "Assessing which of the 6 NLDR layouts of the MNIST digit 1 data is the better representation using RMSE for varying binwidth ($a_1$).
+Colour is used for the lines and points in the left plot to match the scatterplots of the NLDR layouts (a-f).
 Layout c is universally poor.
-Layouts a, f that show two close clusters are universally suboptimal.
-Layout a with little separation performs well at tiny binwidth (where most points are in their own bin) and poorly as binwidth increases.
-Layout f is the best choice."
-#| label: fig-mnist-rmse
+Layouts a, f that show a big cluster and a small circular cluster are universally optimal.
+Layout a performs well at tiny binwidth (where most points are in their own bin) and not as well as f with larger binwidth, thus layout f is the best choice."
+#| label: fig-mnist-mse
 #| fig-pos: H
@@
 
-@fig-mnist-rmse compares the fit of six layouts computed using UMAP (b), PHATE (c), TriMAP (d), PaCMAP (e) with default (hyper-)~parameter setting and two tSNE runs, one with default (hyper-)~parameter setting (a) and the other changing perplexity to $89$ (f).
+@fig-mnist-mse compares the fit of six layouts computed using UMAP (a), PHATE (b), PaCMAP (c), TriMAP (d) with default (hyper-)parameter setting and two tSNE runs, one with default (hyper-)parameter setting (e) and the other changing perplexity to $89$ (f).
 The layouts are reasonably similar in that they all have the observations in a single blob.
@@
 Both a and f have a small clump of points perhaps slightly disconnected from the other points, in the lower to middle right.
-
-The layout plots are colored to match the lines in the RMSE vs binwidth ($a_1$) plot.
+XXX are the labels I put for each method above correct?
+please fix if not.
+
+The layout plot are colored to match the lines in the RMSE vs binwidth ($a_1$) plot.
 Layouts a, b and f fit the data better than c, d, e, and layout f appears to be the bets fit.
@@
 
-The next step is to investigate the \gD{} layout to understand what information is learned from this representation.
+The next step is to investigate the \gD{} to understand what information is learned from this representation.
 @fig-model-error-mnist summarizes this investigation.
@@
 They can be seen to be more strangely drawn than the others.
-Overall, this \gD{} layout shows a useful way to summarize the variation in way "1"s are drawn.
+Overall, this \gD{} layout shows a useful way to summarise the variation in way "1"s are drawn.
+
+<!-- The MSE for different binwidths ($a_1$) using UMAP, PHATE, PaCMAP, TriMAP with default (hyper-)parameter setting and tSNE with default (hyper-)parameter setting and perplexity $89$ (@fig-mnist-mse) were calculated.
+It is found that tSNE with perplexity $89$ (@fig-mnist-mse (f)) provide the most reasonable representation for the digit 1 dataset, showing universally best.
+However, \gD{} representation shows a big non-linear cluster and a small cluster with a small gap.-->
+
+<!-- In the case of digit 1 data, there should not be any clusters unless anomalies exist, indicating different digit 1 patterns.
+The angle of the digit 1 images varies along this non-linear clustering structure (@fig-model-error-mnist), while the small cluster contains the digit 1 images with different patterns of the digit 1, unlike the usual (@fig-model-error-mnist (c)).
+This provides the evidence for two close clusters.
+
+By visualizing the model generated for tSNE in $10\text{-}D$ helps to assess the \gD{} layout.
+As shown in @fig-clust-mnist (a1), the model provides the evidence for the non-linear structure of the digit 1 data in $10\text{-}D$.
+The model shows some quirks.
+The model's twisted pattern provides evidence for the $10\text{-}D$ data structure, which is not observed by \gD{} layout (@fig-clust-mnist (a2)).
+-->
+
+<!-- PaCMAP layout with author's suggested parameter choice-->
+```{r}
+#| label: read-mnist
+## Import data
+training_data_mnist <- read_rds("data/mnist/mnist_10_pcs_of_digit_1.rds")
+names(training_data_mnist) <- paste0("x", 1:NCOL(training_data_mnist))
+
+training_data_mnist <- training_data_mnist |>
+  mutate(ID = 1:NROW(training_data_mnist))
+
+data_mnist <- training_data_mnist |>
+  select(-ID) |>
+  mutate(type = "data")
+```
 
@@
 
+num_bins_x_mnist <- 30
+
+algo_obj_mnist <- fit_highd_model(
+  highd_data = training_data_mnist, 
+  nldr_data = tsne_mnist2, 
+  bin1 = num_bins_x_mnist, 
+  q = 0.1, 
+  benchmark_highdens = 1)
+
+tsne_minst_scaled <- algo_obj_mnist$nldr_obj$scaled_nldr
+tr_from_to_df_mnist <- algo_obj_mnist$trimesh_data
+df_bin_centroids_mnist <- algo_obj_mnist$model_2d
+df_bin_mnist <- algo_obj_mnist$model_highd
+
 sc_ltr_pos_mnist <- c(0.96, 0.96)
 
-tsne_minst_scaled <- read_rds("data/mnist/mnist_tsne_minst_scaled.rds")
+tsne_minst_scaled <- tsne_minst_scaled |>
+  mutate(cluster = if_else((tSNE1 <= 1) & (tSNE1 >= 0.78) & (tSNE2 <= 0.62) & (tSNE2 >= 0.39), "small_clust", "big_clust"))
+
+trimesh_removed_mnist <- ggplot() + 
+  geom_point(data = tsne_minst_scaled,
+             aes(
+               x = tSNE1,
+               y = tSNE2,
+               colour = cluster
+             ),
+             alpha=0.2)+
+    geom_segment(data = tr_from_to_df_mnist, 
+               aes(
+                 x = x_from, 
+                 y = y_from, 
+                 xend = x_to, 
+                 yend = y_to),
+               colour = "#000000") +
+  scale_color_manual(values=c("#999999",'#ff7f00')) +
+  interior_annotation("a", c(0.08, 0.93), cex = 2) +
+  theme(legend.position = "none")
 
@@
              aes(
-               x = emb1,
-               y = emb2,
+               x = tSNE1,
+               y = tSNE2,
                colour = cluster
@@
 ## Compute error
-error_df <- read_rds("data/mnist/mnist_error_df.rds")
+error_df <- augment(
+  highd_data = training_data_mnist,
+  model_2d = df_bin_centroids_mnist,
+  model_highd = df_bin_mnist)
+
+## To join embedding
+error_df <- error_df |>
+  bind_cols(tsne_minst_scaled |> 
+              select(-ID))
+
+## To transform the data
+## ggplot(data = error_df, aes(x=row_wise_abs_error)) + geom_density() ## right skewed
+
+error_df <- error_df |>
+  mutate(sqrt_row_wise_tot_error = sqrt(row_wise_total_error)) |>
+  mutate(sqrt_row_wise_tot_error = standardize(sqrt_row_wise_tot_error))
 
 highd_error_2d_mnist <- error_df |>
-  ggplot(aes(x = emb1,
-             y = emb2,
+  ggplot(aes(x = tSNE1,
+             y = tSNE2,
              colour = sqrt_row_wise_tot_error)) +
@@
   geom_rect(aes(xmin = 0.82, xmax = 0.92, ymin = 0.38, ymax = 0.58),
-               fill = "transparent", colour = "#000000", linewidth = 1) +
+               fill = "transparent", color = "#000000", linewidth = 1) +
   geom_rect(aes(xmin = 0.9, xmax = 1, ymin = 1.25, ymax = 1.45),
-               fill = "transparent", colour = "#375cb1", linewidth = 1)  +
+               fill = "transparent", color = "#375cb1", linewidth = 1)  +
   geom_rect(aes(xmin = 0.35, xmax = 0.45, ymin = 0.55, ymax = 0.7),
-               fill = "transparent", colour = "#008000", linewidth = 1)  +
+               fill = "transparent", color = "#008000", linewidth = 1)  +
   geom_rect(aes(xmin = 0.7, xmax = 0.8, ymin = 0.25, ymax = 0.4),
-               fill = "transparent", colour = "#4c389a", linewidth = 1) +
-  interior_annotation(label = "a", c(0.08, 0.93), cex = 2) +
+               fill = "transparent", color = "#4c389a", linewidth = 1) +
+  interior_annotation(label = "a", c(0.1, 1), cex = 2.5) +
   theme(legend.position = "none",
@@
 
+<!--bin1 = 22, bin2 = 17, b = 374, non_empty = 140-->
+
+```{r}
+#| label: data-prep-mnist-model-proj
+
+hexID_mnist <- df_bin_centroids_mnist |>
+  #dplyr::filter(std_counts > 0) |>
+  dplyr::pull(hexID)
+
+df_b_mnist <- df_bin_mnist |>
+  dplyr::filter(hexID %in% hexID_mnist) |>
+  dplyr::mutate(type = "model") ## Data with summarized mean
+
+## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
+df_b_mnist <- df_b_mnist[match(hexID_mnist, df_b_mnist$hexID),] |>
+  dplyr::select(-hexID)
+
+# Apply the scaling
+df_model_data_mnist <- bind_rows(data_mnist, df_b_mnist)
+scaled_mnist <- scale_data_manual(df_model_data_mnist, "type") |>
+  as_tibble()
+
+scaled_mnist_data <- scaled_mnist |>
+  filter(type == "data") |>
+  select(-type)
+
+scaled_mnist_data_model <- scaled_mnist |>
+  filter(type == "model") |>
+  select(-type)
+```
+
+```{r}
+#| label: langevitour-best-fit-mnist-proj-tsne
+#| eval: false
+
+data_mnist_cluster <- data_mnist |>
+  select(-type) |>
+  mutate(type = tsne_minst_scaled$cluster)
+
+df_model_data_mnist_n <- bind_rows(df_b_mnist, data_mnist_cluster)
+
+langevitour::langevitour(df_model_data_mnist_n[1:(length(df_model_data_mnist_n)-1)],
+                         lineFrom = tr_from_to_df_mnist$from,
+                         lineTo = tr_from_to_df_mnist$to,
+                         group = factor(df_model_data_mnist_n$type,
+                                        c("big_clust", "small_clust", "model")),
+                         levelColors = c("#999999",'#ff7f00', "#000000"))
+```
+
 ```{r}
@@
 
-proj_obj1 <- read_rds("data/mnist/mnist_proj_obj1.rds")
+projection <- cbind(
+  c(-0.04295,0.42361,-0.36702,0.13103,-0.19678,-0.29621,0.00471,0.09323,0.18822,-0.18484),
+  c(-0.36644,0.05058,0.18200,0.21860,0.13583,-0.03511,0.35339,0.15821,-0.26945,-0.27926))
+
+proj_obj1 <- get_projection(projection = projection, 
+               proj_scale = 2.5, 
+               highd_data = scaled_mnist_data, 
+               model_highd = scaled_mnist_data_model, 
+               trimesh_data = tr_from_to_df_mnist, 
+               axis_param = list(limits = 1, 
+                                 axis_scaled = 1.1, 
+                                 axis_pos_x = -0.92, 
+                                 axis_pos_y = -0.92, 
+                                 threshold = 0.065))
+
+proj_obj1[["cluster"]] <- tsne_minst_scaled$cluster
+
+mnist_proj_tsne_model1 <- plot_proj(
+  proj_obj = proj_obj1, 
+  point_param = c(0.5, 0.5), # size, alpha, color
+  line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
+  plot_limits = c(-1.1, 0.9), 
+  axis_text_size = 4, 
+  is_category = TRUE) +
+  interior_annotation(label = "b", cex = 2) +
+  scale_color_manual(values = c("#999999",'#ff7f00')) +
+  theme(
+    legend.position = "none"
+  )
 
@@
 ## Second projection
-proj_obj2 <- read_rds("data/mnist/mnist_proj_obj2.rds")
+projection <- cbind(
+    c(-0.33305,0.20223,0.06303,0.11124,0.13299,-0.30093,0.34082,0.15581,-0.08864,0.33573),
+    c(0.53615,0.14969,-0.09393,-0.20053,0.07992,-0.34791,0.13057,-0.06814,-0.16737,0.03716))
+
+proj_obj2 <- get_projection(projection = projection, 
+               proj_scale = 2, 
+               highd_data = scaled_mnist_data, 
+               model_highd = scaled_mnist_data_model, 
+               trimesh_data = tr_from_to_df_mnist, 
+               axis_param = list(limits = 1, 
+                                 axis_scaled = 1.1, 
+                                 axis_pos_x = -0.83, 
+                                 axis_pos_y = -0.83, 
+                                 threshold = 0.065))
+
+proj_obj2[["cluster"]] <- tsne_minst_scaled$cluster
+
+mnist_proj_tsne_model2 <- plot_proj(
+    proj_obj = proj_obj2,
+    point_param = c(0.5, 0.3), # size, alpha, color
+    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
+    plot_limits = c(-1, 1), 
+    axis_text_size = 3.5, 
+    is_category = TRUE) +
+  interior_annotation(label = "c", cex = 2) +
+  scale_color_manual(values = c("#999999",'#ff7f00')) +
+  theme(
+      legend.position = "none"
+  )
 
@@
 
+<!--add langevitour screenshots and youtube animation link-->
+```{r}
+#| label: fig-mnist-tri-proj
+#| fig-pos: H
+#| fig-height: 4
+#| fig-width: 12
+#| out-width: 100%
+#| fig-cap: "\\gD{} wireframe layout shows that there is a big non-linear cluster (grey) and a small cluster (orange) located very close to the one corner of the big cluster.
+The MNIST digit 1 data ($n =  7877$ and $p = 10$) has a non-linear structure in $10\\text{-}D$.
+Two \\gD{} projections from a tour on $10\\text{-}D$ reveal that the closeness of the clusters in $10\\text{-}D$ and the twisted pattern of the model fit with tSNE ($a_1 = 0.041$, $b = 659/585 \\text{ } (30, \\text{ }48)$).
+Video of the langevitour animation is available at <>."
+#| eval: false
+
+free(trimesh_removed_mnist) + mnist_proj_tsne_model1 +
+  mnist_proj_tsne_model2 +
+  plot_layout(ncol = 3)
+
+```
+
 ```{r}
@@
 #| out-width: 100%
-#| fig-cap: "The tSNE layout of the MNIST digit 1 data shows a big nonlinear cluster (grey) and a small cluster (orange) located very close to the one corner of the big cluster in \\gD{} (a).
-The MNIST digit 1 data ($n =  7877$ and $p = 10$) has a nonlinear structure in $10\\text{-}D$.
+#| fig-cap: "The tSNE layout of the MNIST digit 1 data shows a big non-linear cluster (grey) and a small cluster (orange) located very close to the one corner of the big cluster in \\gD{} (a).
+The MNIST digit 1 data ($n =  7877$ and $p = 10$) has a non-linear structure in $10\\text{-}D$.
 Two \\gD{} projections from a tour on $10\\text{-}D$ reveal that the closeness of the clusters in $10\\text{-}D$ and the twisted pattern of the model fit with tSNE ($a_1 = 0.041$, $b = 659/585 \\text{ } (30, \\text{ }48)$).
-The link to the video of the langevitour animation is available in @tbl-links.
+Video of the langevitour animation is available at <>.
 The brushing feature in the linked plots helps in visualizing the closeness of the small cluster to the big cluster.
-The link to the video of the linked plots is available in @tbl-links."
+Videos of the linked plots are available at <>."
 
@@
 ## Data with pixel values
-
-img_right_top <- read_rds("data/mnist/mnist_img_right_top.rds")
+mnist_data <- read_rds("data/mnist/mnist_digit_1.rds")
+
+mnist_data <- mnist_data |>
+  mutate(instance = row_number()) |>
+  gather(pixel, value, -Label, -instance) |>
+  extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
+  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28)
+
+img_right_top <- c(868, 1465, 1787, 2006, 4314)
+
+pixels_gathered_within <-  mnist_data |>
+  filter(instance %in% img_right_top)
 
 right_top_img <- plot_digit_img(
-  img_right_top, palette = "Blues 2", title_text = "b")
-
-img_middle <- read_rds("data/mnist/mnist_img_middle.rds")
+  pixels_gathered_within, palette = "Blues 2", title_text = "b")
+
+img_middle <- c(130, 2091, 7795, 163, 77)
+
+pixels_gathered_within <-  mnist_data |>
+  filter(instance %in% img_middle)
 
 middle_img <- plot_digit_img(
-  img_middle, palette = "Greens 3", title_text = "c")
-
-img_right_bottom <- read_rds("data/mnist/mnist_img_right_bottom.rds")
+  pixels_gathered_within, palette = "Greens 3", title_text = "c")
+
+img_right_bottom <- c(2035, 7708, 7710, 4134, 5006)
+
+pixels_gathered_within <-  mnist_data |>
+  filter(instance %in% img_right_bottom)
 
 right_bottom_img <- plot_digit_img(
-  img_right_bottom, palette = "Purples 2", title_text = "d")
-
-img_error_outside <- read_rds("data/mnist/mnist_img_error_outside.rds")
+  pixels_gathered_within, palette = "Purples 2", title_text = "d")
+
+img_error_outside <- c(999, 3650, 4916, 2964, 5622)
+
+pixels_gathered_outside <-  mnist_data |>
+  filter(instance %in% img_error_outside)
 
 imge_error_sample_outside <- plot_digit_img(
-  img_error_outside, palette = "Grays", title_text = "e")
+  pixels_gathered_outside, palette = "Grays", title_text = "e")
 
 ##### inside
-img_error_inside <- read_rds("data/mnist/mnist_img_error_inside.rds")
+img_error_inside <- c(5173, 6981, 7228, 6673, 3179) #, 6673, 3179, 7363, 7650, 5241, 4636
+
+pixels_gathered_inside <-  mnist_data |>
+  filter(instance %in% img_error_inside)
 
 imge_error_sample_inside <- plot_digit_img(
-  img_error_inside, palette = "Oranges", title_text = "f")
+  pixels_gathered_inside, palette = "Oranges", title_text = "f")
 
@@
 #| fig-cap: "The $10\\text{-}D$ model error in \\gD{} layout of the MNIST digit 1 data shows a pattern.
-Most low model errors are distributed along the big nonlinear cluster, while most large model errors are distributed along the small cluster.
-The images associated with large model errors shows different patterns of digit 1, some inside (f) the nonlinear structure and others outside (e).
-Along the nonlinear cluster, the angle of digit 1 changes (b-d)."
+Most low model errors are distributed along the big non-linear cluster, while most large model errors are distributed along the small cluster.
+The images associated with large model errors shows different patterns of digit 1, some inside (e) the non-linear structure and others outside (d).
+Along the non-linear cluster, the angle of digit 1 changes (a-c)."
 #| fig-pos: H
@@
 
-We have developed an approach to help assess and compare NLDR layouts, generated by different methods and (hyper-)~parameter choice(s).
-It depends on conceptualizing the \gD{} layout as a model, allowing for the creation of a wireframe representation of the model that can be lifted into \pD{}.
+We have developed an approach to help assess and compare NLDR layouts, generated by different methods and (hyper-)parameter choice(s).
+It depends on conceptualising the \gD{} layout as a model, creating a wireframe representation of it that can be lifted into \pD{}.
 The fit is assessed by viewing the model in the data space, computing residuals and RMSE.
@@
 
-Two examples illustrating usage are provided: the PBMC3k data where the NLDR is summarizing clustering in \pD{} and hand-written digits illustrating how NLDR represents an intrinsically lower dimensional nonlinear manifold.
+Two examples illustrating usage are provided: the PBMC3k data where the NLDR is summarising clustering in \pD{} and hand-written digits illustrating how NLDR represents an intrinsically lower dimensional nonlinear manifold.
 We demonstrated that the published layout of the PBMC3k is inaccurate, because it grossly exaggerates separation between clusters, and even suggests separation when there is none.
@@
 
-Additional exploration of metrics to summarize the fit could be a new direction for the work.
+Additional exploration of metrics to summarise the fit could be a new direction for the work.
 The difficulty is capturing nonlinear fits, for which Euclidean distance can be sub-optimal.
@@
 Most NLDR methods only provide a \gD{} but if a \kD{} ($k>2$) layout is provided the approach developed here could be extended.
-Binning into cubes could be done in $3\text{-}D$ or higher, relatively easily, and used as a the basis for a wireframe of the fitted model.
-@barber1996 and the software @stephane2023 have algorithms for convex hulls, which \pD{} which serve as an inspiration.
+Binning into cubes could be done in $3\text{-}D$ or higher, relatively easily, and used asa the basis for a wireframe of the fitted model.
+@barber1996 and the software @cxhull have algorithms for convex hulls, which \pD{} which serve as an inspiration.
 A simpler approach using $k$-means clustering to provide centroids could also be possible, but the complication would be to determine how to connect the centroids into an appropriate wireframe.
@@
 The new methodology is accompanied by an R package called quollr, so that it is readily usable and broadly accessible.
-The package has methods to fit the model, compute diagnostics and also visualize the results, with interactivity.
+The package has methods to fit the model, compute diagnostics and also visualise the results, with interactivity.
 We have primarily used the langevitour software [@harisson2024] to view the model in the data space, but other tour software such as tourr [@wickham2011] and detourr [@hart2022] could be also used.
 
-## Supplementary Materials {#sec-supplementary}
+## Supplementary Materials
 
@@
 The R package `quollr`, available on CRAN and at [https://jayanilakshika.github.io/quollr/](https://jayanilakshika.github.io/quollr/), provides software accompaying this paper to fit the wireframe model representation, compute diagnostics, visualize the model in the data with langevitour and link multiple plots interactively.
-
-Direct links to videos for viewing online are available in @tbl-links.
-
-```{r}
-#| label: tbl-links
-#| tbl-pos: H
-#| tbl-cap: "Videos of the langevitour animations and the linked plots."
-# Video links in the paper
-
-links_df <- read_csv("misc/video_links.csv")
-
-# Create the table
-kable(links_df, 
-      format = "latex", 
-      booktabs = TRUE, escape = FALSE) |>
-  kable_styling(position = "center", 
-                full_width = FALSE, 
-                font_size = 12) |>
-  row_spec(0, bold = TRUE) |>
-  column_spec(1:3, width = c("1.5cm", "4.5cm", "9cm"))
-```
 

### 2025-06-17 paper.qmd

@@
 
+<!-- 
 XXX TO DO:
@@
 Only code for the plots and tables is directly in the qmd in the final version.
+-->
 
@@
 
-```{r}
-#| label: tbl-notation
-#| tbl-cap: "Summary of notation for describing new methodology."
-# Notation used in the paper
-
-notation_df <- read_csv("misc/notation.csv")
-
-# Create the table
-kable(notation_df, 
-      format = "latex", 
-      booktabs = TRUE, escape = FALSE) |>
-  kable_styling(position = "center", 
-                full_width = FALSE, 
-                font_size = 12) |>
-  row_spec(0, bold = TRUE) |>
-  column_spec(1:2, width = c("3cm", "12cm"))
-```
-
-We define the NLDR as a function $g\text{:}~ \mathbb{R}^{n\times p} \rightarrow \mathbb{R}^{n\times k}$, with (hyper-)parameters $\mathbfit{\theta}$.
-The parameters, $\mathbfit{\theta}$, depend on the choice of $g$, and can be considered part of model fitting in the traditional sense.
+We define the NLDR as a function $g\text{:}~ \mathbb{R}^{n\times p} \rightarrow \mathbb{R}^{n\times k}$, with (hyper-)~parameters $\mathbfit{\theta}$.
+These parameters, $\mathbfit{\theta}$, depend on the choice of $g$, and can be considered part of model fitting in the traditional sense.
 Common choices for $g$ include functions used in tSNE, UMAP, PHATE, TriMAP, PaCMAP, or MDS, although in theory any function that does this mapping is suitable.
-<!--Any input requirements for the data (such as normalization, or preprocessing through the use of PCA or similar) is considered part of the function $g$.-->
 
@@
 #| label: empty-bin-two-curvy
-
-# df_bin_centroids_two_curvy <- df_bin_centroids_two_curvy |>
-#   dplyr::filter(std_counts > 0)
 
@@
 ## First projection
-model_prj1 <- cbind(
-  c(0.37847,-0.06517,0.04231,0.01655),
-  c(-0.06466,-0.27052,0.22131,-0.15235))
-
-proj_obj1 <- quollr::get_projection(projection = model_prj1, 
-               proj_scale = 1, 
-               highd_data = scaled_two_curvy_data, 
-               model_highd = scaled_two_curvy_data_model, 
-               trimesh_data = tr_from_to_df_two_curvy, 
-               axis_param = list(limits = 0.5, 
-                                 axis_scaled = 2,
-                                 axis_pos_x = -0.5, 
-                                 axis_pos_y = -0.5, 
-                                 threshold = 0.03))
-
-## For true model
-
-projection_scaled <- model_prj1 * 1
-
-projected_true_model <- as.matrix(scaled_two_curvy_data_true_model) %*% projection_scaled
-
-projected_true_model_df <- projected_true_model |>
-  tibble::as_tibble(.name_repair = "unique") |>
-  dplyr::rename(c("proj1" = "...1",
-                  "proj2" = "...2")) |>
-  dplyr::mutate(ID = dplyr::row_number())
-
-true_model_df_proj <- dplyr::left_join(
-  wireframe_true_model, 
-  projected_true_model_df, 
-  by = c("from" = "ID"))
-
-names(true_model_df_proj)[3:NCOL(true_model_df_proj)] <- paste0(names(projected_true_model_df)[-NCOL(projected_true_model_df)], "_from")
-
-true_model_df_proj <- dplyr::left_join(true_model_df_proj, projected_true_model_df, by = c("to" = "ID"))
-names(true_model_df_proj)[(2 + NCOL(projected_true_model_df)):NCOL(true_model_df_proj)] <- paste0(names(projected_true_model_df)[-NCOL(projected_true_model_df)], "_to")
-
-proj_obj1_true <- proj_obj1
-proj_obj1_true[["model_df"]] <- true_model_df_proj
-
-two_curvy_proj_true <- plot_proj(
-  proj_obj = proj_obj1_true, 
-  point_param = c(0.5, 0.5, clr_choice), # size, alpha, color
-  line_param = c(0.5, 0.5, "black"), # linewidth, alpha
-  plot_limits = c(-0.6, 0.6), 
-  axis_text_size = 4, 
-  is_category = FALSE) +
-  interior_annotation(label = "a", cex = 1.5) +
-  theme(aspect.ratio=1, 
-        legend.position = "none")
-
-# Changed the axis parameters
-axis_obj <- gen_axes(
-    proj = model_prj1 * 2,
-    limits = 0.9,
-    axis_pos_x = -0.4,
-    axis_pos_y = -0.4,
-    axis_labels = names(scaled_two_curvy_data),
-    threshold = 0.05)
-
-axes <- axis_obj$axes
-circle <- axis_obj$circle
-
-proj_obj1[["axes"]] <- axes
-proj_obj1[["circle"]] <- circle
+proj_obj1 <- read_rds("data/two_nonlinear/two_nonlinear_proj_obj1.rds")
 
@@
   is_category = FALSE) +
-  interior_annotation(label = "c", position = sc_ltr_pos, cex = 1.2) +
-  theme(aspect.ratio = 1) #+
-  #geom_segment(
-  #  data = model_df, 
-  #  aes(
-  #    x = proj1_from, 
-  #    y = proj2_from, 
-  #    xend = proj1_to, 
-  #    yend = proj2_to), 
-  #  color = "#000000")
+  interior_annotation(label = "c", cex = 1.2) +
+  theme(aspect.ratio = 1)
 
@@
 <!-- The remaining variables $X_4, X_5, X_6, X_7$ are all uniform error, with small variance.  -->
-
-```{r}
-#| label: fig-two-curvy-true-proj
-#| fig-cap: "Two \\gD{} projections from $4\\text{-}D$, for the two non-linear clusters data (blue points) and true model (black lines) are shown.
-The top cluster in (a) shows uniformly distributed data along a C-shape, while in the bottom cluster, the data is more concentrated in one corner of the curvy shape.
-Video of the langevitour animations are available at <>."
-#| fig-pos: H
-#| fig-width: 10
-#| fig-height: 5
-#| eval: false
-
-two_curvy_proj_true + two_curvy_proj_true2 +
-  plot_layout(ncol = 2)
-```
 
@@
 #| label: error-two-curvy
-
-error_two_curvy_umap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_tsne.rds")
-
-## Find the minimum MSE when have duplicate a1
-error_two_curvy_umap <- error_two_curvy_umap |>
-  group_by(a1) |>
-  filter(bin1 == min(bin1)) |>
-  ungroup() |>
-  mutate(prop_dens = 1/(b*side_length^2))
-
-base_line_prop_dens <- error_two_curvy_umap |>
-  filter(a1 == min(a1)) |>
-  pull(prop_dens)
-
-error_two_curvy_umap <- error_two_curvy_umap |>
-  mutate(prop_comp = prop_dens/base_line_prop_dens)
+error_two_curvy_umap <- read_rds("data/two_nonlinear/error_two_curvy_umap.rds")
 
@@
 
-A new benefit of this fitted model is that it allows us to now predict a new observation's value in the NLDR, for any method.
-The steps are to determine the closest bin centroid in \pD{}, $C^{(p)}_{h}$ and predict it to be the centroid of this bin in \gD{}, $C^{(2)}_{h}$.
-This can be written as, let $z(i) = \arg\min_{j = 1, \dots, b} \sqrt{\sum_{v=1}^{p}(x_{iv} - C^{(p)}_{jv})^2}$, then the new observation $i$ falls in the hexagon, $H_h = \{i| z(i) = h\}$ and the corresponding \kD{} bin centroids, $C_{h}^{(2)} = (c_{h1}, c_{h2})$.
-
-<!--prediction for true-model of two-curvy clust-->
-```{r}
-#| label: model-prediction-umap-original
-
-predict_tsne_two_curvy <- read_rds(file = "data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_umap_predict_true.rds")
-
-predict_two_curvy_obj <- gen_scaled_data(
-    nldr_data = predict_tsne_two_curvy)
-
-predict_tsne_two_curvy_scaled <- predict_two_curvy_obj$scaled_nldr
-
-plot_predict_umap_fun_two_curvy <- ggplot(
-  data = predict_tsne_two_curvy_scaled, 
-  aes(x = UMAP1, 
-      y = UMAP2)) +
-  geom_point(alpha = 0.5) +
-  interior_annotation("b", c(0.08, 0.93), cex = 1.5)
-```
-
-```{r}
-#| label: model-prediction-points-line
-#| eval: false
-
-predict_tsne_two_curvy_scaled <- predict_tsne_two_curvy_scaled |>
-  mutate(ID = row_number()) |>
-  mutate(type = "predict_from_umap")
-
-jittered_points_df <- jittered_points_df |>
-  mutate(type = "predict_from_model")
-
-predict_point_df <- bind_rows(
-  predict_tsne_two_curvy_scaled,
-  jittered_points_df
-)
-
-plot_predict_umap_connect_two_curvy <- ggplot(
-  data = predict_point_df, 
-  aes(x = UMAP1, 
-      y = UMAP2, 
-      group = ID)) +
-  geom_point(alpha = 0.5) +
-  geom_line(alpha = 0.3) +
-  interior_annotation("d", c(0.08, 0.93), cex = 1.5)
-```
-
-```{r}
-#| echo: false
-#| label: fig-predict-two-curvy
-#| fig-pos: H
-#| fig-width: 10
-#| fig-height: 10
-#| out-width: 80%
-#| fig-cap: "Comparison of prediction generated using the exiting `umap` R package's prediction method and our method: (a) A view of the true model in projection from $7\\text{-}D$, (b) predicted data from the `umap` prediction method, and (c) predicted data from our method."
-#| eval: false
-
-proj_plot_two_curvy_true + plot_predict_umap_fun_two_curvy +plot_predict_tsne_two_curvy +
-  plot_layout(ncol = 3)
-```
+<!-- Does this really need an entire subsection? It's only 3 sentences long. -->
+
+In this paradigm of model fitting, it is now possible to predict a new observation's value in the NLDR representation, independent of the choice of NLDR method.
+To do this, we first need to determine the closest bin centroid in \pD{}, $C^{(p)}_{h}$, to the new observation, which is,
+
+$$z = \arg\min_{j = 1, \dots, b} \sqrt{\sum_{v=1}^{p}(x_{iv} - C^{(p)}_{jv})^2}.$$
+
+Then, the predicted value is is the centroid of this bin in \gD{}, $C^{(2)}_{z} = (c_{z1}, c_{z2})$.
 
@@
 
+Direct links to videos for viewing online are available in @tbl-links.
+
+```{r}
+#| label: tbl-links
+#| tbl-cap: "Videos of the langevitour animations and the linked plots."
+# Video links in the paper
+
+links_df <- read_csv("misc/video_links.csv")
+
+# Create the table
+kable(links_df, 
+      format = "latex", 
+      booktabs = TRUE, escape = FALSE) |>
+  kable_styling(position = "center", 
+                full_width = FALSE, 
+                font_size = 12) |>
+  row_spec(0, bold = TRUE) |>
+  column_spec(1:3, width = c("1.5cm", "4.5cm", "9cm"))
+```
+
 ## Acknowledgments

### 2025-06-17 paper.qmd
Do not need quollr to run the paper

--- 2025-06-17 paper.qmd
@@
 
-if (!requireNamespace("quollr", quietly = TRUE)) {
-  remotes::install_github("JayaniLakshika/quollr")
-}
-
 ```
@@
 #| echo: false
-library(quollr)
 library(tibble)
@@
 #| label: tbl-links
-#| tbl-pos: H
 #| tbl-cap: "Videos of the langevitour animations and the linked plots."

### 2025-06-17 paper.qmd
Updated prediction section with Di's comments

@@
 
-In this paradigm of model fitting, it is now possible to predict a new observation's value in the NLDR representation, independent of the choice of NLDR method.
-To do this, we first need to determine the closest bin centroid in \pD{}, $C^{(p)}_{h}$, to the new observation, which is,
-
-$$z = \arg\min_{j = 1, \dots, b} \sqrt{\sum_{v=1}^{p}(x_{iv} - C^{(p)}_{jv})^2}.$$
-
-Then, the predicted value is is the centroid of this bin in \gD{}, $C^{(2)}_{z} = (c_{z1}, c_{z2})$.
+A new benefit of this fitted model is that it allows us to now predict a new observation's value in the NLDR, for any method.
+The steps are to determine the closest bin centroid in \pD{}, $C^{(p)}_{h}$ and predict it to be the centroid of this bin in \gD{}, $C^{(2)}_{h}$.
+This can be written as, let $z(i) = \arg\min_{j = 1, \dots, b} \sqrt{\sum_{v=1}^{p}(x_{iv} - C^{(p)}_{jv})^2}$, then the new observation $i$ falls in the hexagon, $H_h = \{i| z(i) = h\}$ and the corresponding \kD{} bin centroids, $C_{h}^{(2)} = (c_{h1}, c_{h2})$.
 

### 2025-06-17 paper.qmd
Updated some texts

@@
 This more accurately reflects the hexagon grid relative the RMSE than the full number of bins in the grid.
-It may be beneficial to remove low count bins also, such as when the data is clustered or stringy, or where the observed data is sparse.
+It may also be beneficial to remove low count bins also, in the situation where data is clustered or stringy, where the observed data is sparse.
 In order to decide if this is necessary, you would examine the distribution of bin counts, or the density which puts the counts on a standard scale.
 If there is something of a gap at low values, this would suggest a potential value to use as a cutoff.
-Alternatively, one could choose to remove based on a percentile, the bins with density in the lowest 5% of all bins, for example.
+Alternatively, one could choose to remove based on a percentile, the bins with density in the lowest $5\%$ of all bins, for example.
 @fig-param-two-curvy (c) illustrates the effect on the model representation of removing bins below different percentages.
@@
 
-The benchmark value for removing low-density hexagons ranges between $0$ and $1$, where $0$ corresponds to no bins being removed, and $1$ to 100% of bins being removed.
+The benchmark value for removing low-density hexagons ranges between $0$ and $1$.
 When analyzing how these benchmark values influence model performance, it's essential to observe the change in RMSE as the benchmark value increases (@fig-param-two-curvy (c)).
@@
 
-To understand how well the model fits the points whether it fits well, works better in some positions, or fails to match the overall pattern, it is important to link and brush the points with high model error in the \pD{} error plot, \gD{} layout, and the generated model overlay on the data in \pD{}.
+To understand how well the model fits the points whether it fits well, works better in some positions, or fails to match the overall pattern.
+It is important to link and brush the points with high model error in the \pD{} error plot, \gD{} layout, and the generated model overlay on the data in \pD{}.
 

### 2025-06-17 paper.qmd
captions put back in

@@
 ```{r}
-#| fig-cap: "Assessing which of the 6 NLDR layouts on the MNIST digit 1 data is the better representation using RMSE for varying binwidth ($a_1$).
-Color used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f).
-All the layouts appear to be very similar.
+#| fig-cap: "Assessing which of the 6 NLDR layouts of the MNIST digit 1 data is the better representation using RMSE for varying binwidth ($a_1$).
+Colour is used for the lines and points in the left plot to match the scatterplots of the NLDR layouts (a-f).
 Layout c is universally poor.
-Layouts a, f that show two close clusters are universally suboptimal.
-Layout a with little separation performs well at tiny binwidth (where most points are in their own bin) and poorly as binwidth increases.
-Layout f is the best choice."
+Layouts a, f that show a big cluster and a small circular cluster are universally optimal.
+Layout a performs well at tiny binwidth (where most points are in their own bin) and not as well as f with larger binwidth, thus layout f is the best choice."
 #| label: fig-mnist-rmse

### 2025-06-17 paper.qmd
Removed text with "links" in captions

@@
 This highlights a key characteristic of tSNE, which compresses the data when applying tSNE.
-Furthermore, the sparse space observed in the middle of the tSNE layout is a result of (hyper-)~parameter choice.
-The link to the video of the langevitour animation is available in @tbl-links."
+Furthermore, the sparse space observed in the middle of the tSNE layout is a result of (hyper-)~parameter choice."
 #| fig-width: 8
@@
 This would indicate the distinct clusters, each being fully \\fD{}.
-With (c2), the curiousity is that the model is a \\gD{} pancake shape in \\fD{}, indicating that there is some ordering of points done by PaCMAP, posisbly along some principal component axes.
-The links to the videos of the langevitour animations are available in @tbl-links."
+With (c2), the curiousity is that the model is a \\gD{} pancake shape in \\fD{}, indicating that there is some ordering of points done by PaCMAP, posisbly along some principal component axes."
 
@@
 a2, a3 and b2, b3 present the results from linked plots that brushed the high \\fD{} error points in \\gD{} and \\fD{}.
-It helps in identifying the high \\fD{} model errors that occur due to the sparse end.
-The link to the video of the linked plots is available in @tbl-links."
+It helps in identifying the high \\fD{} model errors that occur due to the sparse end."
 
@@
 #| echo: false
-#| fig-cap: "Compare the published \\gD{} layout (a) made with UMAP and the \\gD{} layout selected by RMSE plot (e) made by tSNE, both with default (hyper-)parameters.
+#| fig-cap: "Compare the published \\gD{} layout (a) made with UMAP and the \\gD{} layout selected by RMSE plot (e) made by tSNE.
 The two plots on the right show  projections from a tour, with the models overlaid.
@@
 #| fig-cap: "The tSNE layout of the MNIST digit 1 data shows a big nonlinear cluster (grey) and a small cluster (orange) located very close to the one corner of the big cluster in \\gD{} (a).
-The MNIST digit 1 data ($n =  7877$ and $p = 10$) has a nonlinear structure in $10\\text{-}D$.
-Two \\gD{} projections from a tour on $10\\text{-}D$ reveal that the closeness of the clusters in $10\\text{-}D$ and the twisted pattern of the model fit with tSNE ($a_1 = 0.041$, $b = 659/585 \\text{ } (30, \\text{ }48)$).
-The link to the video of the langevitour animation is available in @tbl-links.
-The brushing feature in the linked plots helps in visualizing the closeness of the small cluster to the big cluster.
-The link to the video of the linked plots is available in @tbl-links."
+The MNIST digit 1 data has a nonlinear structure in $10\\text{-}D$.
+Two \\gD{} projections from a tour on $10\\text{-}D$ reveal that the closeness of the clusters in $10\\text{-}D$ and the twisted pattern of the model fit with tSNE.
+The brushing feature in the linked plots helps in visualizing the closeness of the small cluster to the big cluster."
 

### 2025-06-17 paper.qmd
Restructured video links table

@@
 But NLDR views have an important role in data analysis because, if done well, they provide a concise visual (and conceptual) summary of \pD{} distributions.
-The NLDR methods and (hyper-)~parameter choices can create wildly different representations, making it difficult to decide which is best, or whether any or all are accurate or misleading.
+The NLDR methods and hyper-parameter choices can create wildly different representations, making it difficult to decide which is best, or whether any or all are accurate or misleading.
 To help assess the NLDR and decide on which, if any, is the most reasonable representation of the structure(s) present in the \pD{} data, we have developed an algorithm to show the \gD{} NLDR model in the \pD{} space, viewed with a tour, a movie of linear projections.
@@
 Historically, low-dimensional (\kD{}) representations of high-dimensional (\pD{}) data have been computed using multidimensional scaling (MDS) [@kruskal1964], which includes principal components analysis (PCA) [@jolliffe2011] as a special case.
-(A contemporary comprehensive guide to MDS can be found in @borg2005.)
-<!-- Does JASA allow footnotes?
-If so this is a footnote, otherwise the punctuation has gone a bit rogue - can't have an entire sentence in brackets surely --> The \kD{} representation can be considered to be a layout of points in \kD{}   produced by an embedding procedure that maps the data from \pD{}.
+(A contemporary comprehensive guide to MDS can be found in @borg2005.) The \kD{} representation can be considered to be a layout of points in \kD{}   produced by an embedding procedure that maps the data from \pD{}.
 In MDS, the \kD{} layout is constructed by minimizing a stress function that differences distances between points in \pD{} with potential distances between points in \kD{}.
@@
 
-<!-- Non-expert comment: This to me is a leap from MDS to NLDR. Is NLDR an MDS? I don't think so, but I'm missing the link -->
-<!-- Suggested rewording: "To better capture specific structures potentially existing in \pD{}, many new methods for NLDR have emerged in recent years." -->
 Many new methods for NLDR have emerged in recent years, all designed to better capture specific structures potentially existing in \pD{}.
@@
 
-We define the NLDR as a function $g\text{:}~ \mathbb{R}^{n\times p} \rightarrow \mathbb{R}^{n\times k}$, with (hyper-)~parameters $\mathbfit{\theta}$.
+We define the NLDR as a function $g\text{:}~ \mathbb{R}^{n\times p} \rightarrow \mathbb{R}^{n\times k}$, with hyper-parameters $\mathbfit{\theta}$.
 These parameters, $\mathbfit{\theta}$, depend on the choice of $g$, and can be considered part of model fitting in the traditional sense.
@@
 
-<!-- I'm not thrilled about H_h but I guess it can stay because changing it everywhere is a pain -->
-
 By default, the bin centroid is used for describing a hexagon (as done in @fig-NLDR-two-curvy (c)), but any measure of center, such as a mean or weighted mean of the points within each hexagon, could be used.
@@
 
-<!-- In Figure 4, is the "sparse space" in the middle of each cluster (i.e.
-the holes) or is it the gap between clusters?
-If it's between clusters, I'd argue tSNE is making this gap smaller than in reality, which I don't think the caption currently conveys.
--->
-
 ```{r}
 #| label: best-fit-tsne
-#| fig-cap: "Lifting the fitted model into \\pD{}.
-Starting from a default tSNE layout (a) of nonlinear clusters data ($n =  2000$ and $p = 4$), from which the \\gD{} wireframe is constructed (a) and then lifted to \\fD{} (b, c).
-The fit is reasonbly tight with the data, although it does not fully spread the full width.
-This highlights a key characteristic of tSNE, which compresses the data when applying tSNE.
-Furthermore, the sparse space observed in the middle of the tSNE layout is a result of (hyper-)~parameter choice."
+#| fig-cap: "Lifting the \\gD{} fitted model into \\pD{}.
+Two projections of the \\pD{} fitted model overlaying the data are shown in b, c.
+The fit is reasonbly tight with the data in one cluster (top one in b), but slightly less so in the other cluster probably because it is in $3-\\text{D}$.
+Notice also that, in the \\gD{} layout the two clusters have internal gaps which creates a model with some holes.
+This lacy pattern happens regardless of the hyper-parameter choice, but this doen't severely impact the \\pD{} model representation."
 #| fig-width: 8
@@
 
-@fig-one-dens_clust-error (a2, b2) shows a \gD{} layout for tSNE created using the default (hyper-)~parameters.
+@fig-one-dens_clust-error (a2, b2) shows a \gD{} layout for tSNE created using the default hyper-parameters.
 One would expect to see a rectangular shape if the curved sheet is flattened, but the layout is triangular.
@@
 #| out-width: 100%
-#| fig-cap: "The sparse end of the structure shows a high model error, while the dense end shows a low model error.
-The tSNE layout of the C-shaped structure, with dense points (a1-a3) are colored according to their \\fD{} model error.
+#| fig-cap: "Exploring the effect of density on the NLDR layout using a C-shaped structure with different density in each end.
+The sparse end of the structure shows a high model error, while the dense end shows a low model error.
+The top part of the triangular shaped with linked brushing.
+The tSNE layout, with dense points (a1-a3) are colored according to their \\fD{} model error.
 Darker colors represent a high error, while lighter colors indicate a low error.
@@
 Following their pre-processing, different NLDR techniques were performed on the first nine principal components.
-@fig-NLDR-variety shows this data using a variety of methods, and different (hyper-)~parameters.
+@fig-NLDR-variety shows this data using a variety of methods, and different hyper-parameters.
 You can see that the result is wildly different depending on the choices.
@@
 @fig-pbmc-rmse shows RMSE across a range of binwidths ($a_1$) for each of the layouts in @fig-NLDR-variety.
-The layouts were made using tSNE, UMAP, PHATE, PaCMAP, and TriMAP with various (hyper-)~parameter settings.
+The layouts were made using tSNE, UMAP, PHATE, PaCMAP, and TriMAP with various hyper-parameter settings.
 Lines are color coded to match the color of the layouts shown on the right.
@@
 
-@fig-mnist-rmse compares the fit of six layouts computed using UMAP (b), PHATE (c), TriMAP (d), PaCMAP (e) with default (hyper-)~parameter setting and two tSNE runs, one with default (hyper-)~parameter setting (a) and the other changing perplexity to $89$ (f).
+@fig-mnist-rmse compares the fit of six layouts computed using UMAP (b), PHATE (c), TriMAP (d), PaCMAP (e) with default hyper-parameter setting and two tSNE runs, one with default hyper-parameter setting (a) and the other changing perplexity to $89$ (f).
 The layouts are reasonably similar in that they all have the observations in a single blob.
@@
 
-We have developed an approach to help assess and compare NLDR layouts, generated by different methods and (hyper-)~parameter choice(s).
+We have developed an approach to help assess and compare NLDR layouts, generated by different methods and hyper-parameter choice(s).
 It depends on conceptualizing the \gD{} layout as a model, allowing for the creation of a wireframe representation of the model that can be lifted into \pD{}.
@@
 #| label: tbl-links
+#| tbl-pos: H
 #| tbl-cap: "Videos of the langevitour animations and the linked plots."
@@
   row_spec(0, bold = TRUE) |>
-  column_spec(1:3, width = c("1.5cm", "4.5cm", "9cm"))
+  column_spec(1:2, width = c("1.5cm", "4.5cm"))
 ```

### 2025-06-17 paper.qmd
Updated 10X Genometic bib

@@
 Two projections of the \\pD{} fitted model overlaying the data are shown in b, c.
-The fit is reasonbly tight with the data in one cluster (top one in b), but slightly less so in the other cluster probably because it is in $3-\\text{D}$.
+The fit is reasonbly tight with the data in one cluster (top one in b), but slightly less so in the other cluster probably because it is in $3\\text{-}D$.
 Notice also that, in the \\gD{} layout the two clusters have internal gaps which creates a model with some holes.

### 2025-06-17 paper.qmd
Added high-d projection

@@
              y = emb2,
-             colour = sqrt_row_wise_total_error)) +
+             colour = error_cat)) +
   geom_point(alpha=0.5) +
-  scale_colour_continuous_sequential(palette = "YlOrRd", n_interp = 20) +
+  scale_colour_manual(values = c('#ffffcc','#ffeda0','#fed976',
+                                 '#feb24c','#fd8d3c','#fc4e2a',
+                                 '#e31a1c','#bd0026','#800026')) +
   xlab("tSNE1") +
@@
       y = 0, 
-      colour = sqrt_row_wise_total_error)) +
+      colour = error_cat)) +
   geom_quasirandom(alpha = 0.5) +
-  scale_colour_continuous_sequential(palette = "YlOrRd", n_interp = 20) +
+  scale_colour_manual(values = c('#ffffcc','#ffeda0','#fed976',
+                                 '#feb24c','#fd8d3c','#fc4e2a',
+                                 '#e31a1c','#bd0026','#800026')) +
   xlab(expression(e[hj])) +
@@
         panel.grid.minor.y = element_blank(),
-        legend.position = "none")
+        legend.position = "none",
+        plot.margin = margin(0, 0, 0, 0))
 ```
@@
   interior_annotation(label = "b", cex = 1.2) +
+  theme(aspect.ratio = 1,
+        legend.position = "none")
+
+projected_df <- proj_obj1$projected_df
+axes <- proj_obj1$axes
+circle <- proj_obj1$circle
+cluster <- proj_obj1$cluster
+
+projected_df <- projected_df |>
+      dplyr::mutate(cluster = proj_obj1$cluster)
+
+two_curvy_proj_tsne_first_model1_coloured <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1,
+      y = proj2,
+      colour = cluster)) +
+  geom_point(
+    size = 0.5,
+    alpha = 0.2) +
+  geom_segment(
+    data=axes,
+    aes(x=x1, y=y1, xend=x2, yend=y2),
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2),
+    label=rownames(axes),
+    colour="grey50",
+    size = 2) +
+  geom_path(
+    data=circle,
+    aes(x=c1, y=c2), colour="grey70") +
+  scale_colour_manual(values = c('#ffffcc','#ffeda0','#fed976',
+                                 '#feb24c','#fd8d3c','#fc4e2a',
+                                 '#e31a1c','#bd0026','#800026')) +
+  xlim(c(-0.6, 0.6)) +
+  ylim(c(-0.6, 0.6)) +
   theme(aspect.ratio = 1,
@@
 
-<!-- Sparse end with dense corner to me is very hard to understand - sparse and dense are pretty much antonyms -->
 ```{r}
@@
 
-error_plot_two_curvy + error_plot_two_curvy_quasi +
-  plot_layout(ncol=2)
+ error_plot_two_curvy_quasi + error_plot_two_curvy + two_curvy_proj_tsne_first_model1_coloured +
+  plot_layout(ncol=3)
 ```

### 2025-06-17 paper.qmd
Added error categories

--- 2025-06-17 paper.qmd
@@
              y = emb2,
-             colour = sqrt_row_wise_total_error)) +
+             colour = error_cat)) +
   geom_point(alpha=0.5) +
-  scale_colour_continuous_sequential(palette = "YlOrRd", n_interp = 20) +
+  scale_colour_manual(values = c("#FFFFC8", "#FFFAC0", "#FEF2B3", "#FCE9A3",
+                                 "#FADE8F", "#F9D378", "#F8C65D", "#F6B938",
+                                 "#F5AA00", "#F49B00", "#F28A00", "#F07800",
+                                 "#ED6400", "#EA4D00", "#DE3900", "#CD2A00",
+                                 "#BB1B00", "#A70C00", "#93001C", "#7D0025")) +
   xlab("tSNE1") +
@@
       y = 0, 
-      colour = sqrt_row_wise_total_error)) +
+      colour = error_cat)) +
   geom_quasirandom(alpha = 0.5) +
-  scale_colour_continuous_sequential(palette = "YlOrRd", n_interp = 20) +
+  scale_colour_manual(values = c("#FFFFC8", "#FFFAC0", "#FEF2B3", "#FCE9A3",
+                                 "#FADE8F", "#F9D378", "#F8C65D", "#F6B938",
+                                 "#F5AA00", "#F49B00", "#F28A00", "#F07800",
+                                 "#ED6400", "#EA4D00", "#DE3900", "#CD2A00",
+                                 "#BB1B00", "#A70C00", "#93001C", "#7D0025")) +
   xlab(expression(e[hj])) +
@@
   interior_annotation(label = "b", cex = 1.2) +
+  theme(aspect.ratio = 1,
+        legend.position = "none")
+
+projected_df <- proj_obj1$projected_df
+axes <- proj_obj1$axes
+circle <- proj_obj1$circle
+cluster <- proj_obj1$cluster
+
+projected_df <- projected_df |>
+      dplyr::mutate(cluster = proj_obj1$cluster)
+
+two_curvy_proj_tsne_first_model1_coloured <- projected_df |>
+  ggplot(
+    aes(
+      x = proj1,
+      y = proj2,
+      colour = cluster)) +
+  geom_point(
+    size = 0.5,
+    alpha = 0.2) +
+  geom_segment(
+    data=axes,
+    aes(x=x1, y=y1, xend=x2, yend=y2),
+    colour="grey70") +
+  geom_text(
+    data=axes,
+    aes(x=x2, y=y2),
+    label=rownames(axes),
+    colour="grey50",
+    size = 2) +
+  geom_path(
+    data=circle,
+    aes(x=c1, y=c2), colour="grey70") +
+  scale_colour_manual(values = c("#FFFFC8", "#FFFAC0", "#FEF2B3", "#FCE9A3",
+                                 "#FADE8F", "#F9D378", "#F8C65D", "#F6B938",
+                                 "#F5AA00", "#F49B00", "#F28A00", "#F07800",
+                                 "#ED6400", "#EA4D00", "#DE3900", "#CD2A00",
+                                 "#BB1B00", "#A70C00", "#93001C", "#7D0025")) +
+  xlim(c(-0.6, 0.6)) +
+  ylim(c(-0.6, 0.6)) +
   theme(aspect.ratio = 1,

### 2025-06-17 appendix.qmd
Reorganized video link table

@@
       format = "latex", 
-      booktabs = TRUE, escape = FALSE) |>
+      booktabs = TRUE, escape = FALSE, linesep = "") |>
   kable_styling(position = "center", 

### 2025-06-17 paper.qmd
Reorganized video link table

@@
 The R package `quollr`, available on CRAN and at [https://jayanilakshika.github.io/quollr/](https://jayanilakshika.github.io/quollr/), provides software accompaying this paper to fit the wireframe model representation, compute diagnostics, visualize the model in the data with langevitour and link multiple plots interactively.
-
 Direct links to videos for viewing online are available in @tbl-links.
@@
   row_spec(0, bold = TRUE) |>
-  column_spec(1:2, width = c("1.5cm", "4.5cm"))
+  column_spec(1:2, width = c("1.5cm", "14cm"))
 ```

### 2025-06-20 paper.qmd
Add sentence to mention about video links
Updated Figure 10 title

@@
 Limitations and future directions are provided in @sec-discussion.
+Links to the langevitour animation videos showing the \gD{} projections are provided in @tbl-links.
 
@@
 #| out-width: 100%
-#| fig-cap: "Exploring the effect of density on the NLDR layout using a C-shaped structure with different density in each end.
-The sparse end of the structure shows a high model error, while the dense end shows a low model error.
-The top part of the triangular shaped with linked brushing.
-The tSNE layout, with dense points (a1-a3) are colored according to their \\fD{} model error.
-Darker colors represent a high error, while lighter colors indicate a low error.
-a2, a3 and b2, b3 present the results from linked plots that brushed the high \\fD{} error points in \\gD{} and \\fD{}.
+#| fig-cap: "Exploring the effect of density on the NLDR layout using a C-shaped structure with different density at each end.
+The top part of the triangular shape, shown with linked brushing, identify high model error corresponding to the sparse end of the structure (a1-a3).
+On the other hand, one of the dense corners of the structure shows low model error and is highlighted in the right part of the triangular shape (b1-b3).
+The tSNE layouts (a1, b1) are colored according to selected points.
+a2, a3 and b2, b3 present the results from linked plots that brushed the \\fD{} error points in \\gD{} and \\fD{}.
 It helps in identifying the high \\fD{} model errors that occur due to the sparse end."

### 2025-06-20 paper.qmd
Updated Figure 5 title

@@
   geom_point(alpha=0.5) +
+  interior_annotation(label = "b", cex = 1.2) +
   scale_colour_manual(values = c("#FFFFC8", "#FFFAC0", "#FEF2B3", "#FCE9A3",
@@
   geom_quasirandom(alpha = 0.5) +
+  interior_annotation(label = "a", cex = 1.2) +
   scale_colour_manual(values = c("#FFFFC8", "#FFFAC0", "#FEF2B3", "#FCE9A3",
@@
   ylim(c(-0.6, 0.6)) +
+  interior_annotation(label = "c", cex = 1.2) +
   theme(aspect.ratio = 1,
@@
 #| label: fig-p-d-error-in-2d-two-curvy
-#| fig-cap: "The \\fD{} model error in \\gD{} layout.
+#| fig-cap: "Explore the distribution of residuals in \\gD{} and \\fD{}.
 Color indicates error ($e_{hj}$), dark color indicating high error and light indicates low error.
-Most large errors are distributed near the sparse end of the nonlinear cluster and most small errors are distributed near the dense corners."
+Most large errors are distributed in one cluster (bottom one in c) and most small errors are distributed in the other cluster."
 #| out-width: 80%

### 2025-06-20 paper.qmd
Updated the function name

@@
 error_plot_two_curvy <-
-  plot_mse(error_two_curvy) +
+  plot_rmse(error_two_curvy) +
   scale_x_continuous(breaks =
@@
 
-error_plot_pbmc <- plot_mse(error_pbmc) +
+error_plot_pbmc <- plot_rmse(error_pbmc) +
   scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 13, 17, 21, 26)]) +
@@
 
-error_plot_mnist <- plot_mse(error_mnist) +
+error_plot_mnist <- plot_rmse(error_mnist) +
   scale_x_continuous(breaks = sort(

### 2025-06-23 paper.qmd
keywords fixed and fig 5 tidied

@@
 
-keywords: [high-dimensional data vizualization, non-linear dimension reduction, tour]
-keywords-formatted: [high-dimensional data vizualization, non-linear dimension reduction, tour]
+keywords: [high-dimensional data, data visualization, tour, statistical graphics, exploratory data analysis, unsupervised learning]
+keywords-formatted: [high-dimensional data, data visualization, tour, statistical graphics, exploratory data analysis, unsupervised learning]
 
@@
   geom_point(alpha=0.5) +
-  interior_annotation(label = "b", cex = 1.2) +
+  interior_annotation(label = "b", cex = 1.5) +
   scale_colour_manual(values = c("#FFFFC8", "#FFFAC0", "#FEF2B3", "#FCE9A3",
@@
     aspect.ratio = 1,
-    legend.position = "none"
+    legend.position = "none",
+    axis.text = element_blank(),
+    axis.ticks = element_blank(),
+    panel.grid = element_blank()
   )
@@
   geom_quasirandom(alpha = 0.5) +
-  interior_annotation(label = "a", cex = 1.2) +
+  interior_annotation(label = "a", cex = 1.5) +
   scale_colour_manual(values = c("#FFFFC8", "#FFFAC0", "#FEF2B3", "#FCE9A3",
@@
   ylim(c(-0.6, 0.6)) +
-  interior_annotation(label = "c", cex = 1.2) +
+  interior_annotation(label = "c", cex = 1.5) +
   theme(aspect.ratio = 1,
@@
 #| label: fig-p-d-error-in-2d-two-curvy
-#| fig-cap: "Explore the distribution of residuals in \\gD{} and \\fD{}.
+#| fig-cap: "Examining the distribution of residuals in a jittered dotplot (a), \\gD{} NLDR layout (b) and a tour of \\fD{} data space (c).
 Color indicates error ($e_{hj}$), dark color indicating high error and light indicates low error.
 Most large errors are distributed in one cluster (bottom one in c) and most small errors are distributed in the other cluster."
-#| out-width: 80%
+#| out-width: 100%
 #| fig-pos: H
+#| fig-width: 8
+#| fig-height: 3
 

### 2025-06-24 paper.qmd
abstract - PH comment
background updated - PH comments

@@
   Non-linear dimension reduction (NLDR) techniques such as tSNE, and UMAP provide a low-dimensional representation of high-dimensional data (\pD{}) by applying a non-linear transformation.
-NLDR often exaggerates random patterns, sometimes due to the samples observed.
+NLDR often exaggerates random patterns.
 But NLDR views have an important role in data analysis because, if done well, they provide a concise visual (and conceptual) summary of \pD{} distributions.
@@
 <!-- - Connection between NLDR and MDS-->
-Historically, low-dimensional (\kD{}) representations of high-dimensional (\pD{}) data have been computed using multidimensional scaling (MDS) [@kruskal1964], which includes principal components analysis (PCA) [@jolliffe2011] as a special case.
+Historically, low-dimensional (\kD{}) representations of high-dimensional (\pD{}) data have been computed using multidimensional scaling (MDS) [@kruskal1964], which includes principal components analysis (PCA) (for an overview see @jolliffe2011).
 (A contemporary comprehensive guide to MDS can be found in @borg2005.) The \kD{} representation can be considered to be a layout of points in \kD{}   produced by an embedding procedure that maps the data from \pD{}.
@@
 Key structures interpreted from these views suggest: (1) highly **separated clusters** (a, b, e, g, h) with the number ranging from 3-6; (2) **stringy branches** (f), and (3) **barely separated clusters** (c, d) which would **contradict** the other representations.
-
-It happens because these methods and parameter choices provide different lenses on the interpoint distances in the data.
+These contradictions arise because these methods and parameter choices provide different lenses on the interpoint distances in the data.
 
@@
 PCA is the classical approach, resulting in a set of new variables which are linear combinations of the original variables.
-Tours, defined by @lee2021, broaden the scope by providing movies of linear projections, that provide views the data from all directions.
-@lee2021 provides an review of the main developments in tours.
-There are many tour algorithms implemented, with many available in the R package `tourr` [@wickham2011], and versions enabling better interactivity in `langevitour` [@harisson2024] and `detourr` [@hart2022].
+Tours, defined by @As85, broaden the scope by providing movies of linear projections, that provide views the data from all directions.
+(See @lee2021 for a review of tour methods.) There are many tour algorithms implemented, with many available in the R package `tourr` [@wickham2011], and versions enabling better interactivity in `langevitour` [@harisson2024] and `detourr` [@hart2022].
 Linear projections are a safe way to view high-dimensional data, because they do not warp the space, so they are more faithful representations of the structure.
@@
 
-The solution is to use the tour to examine how the NLDR is warping the space.
-This approach follows what @wickham2015 describes as *model-in-the-data-space*.
+Our proposed solution is to use the tour to examine how the NLDR is warping the space.
+It follows what @wickham2015 describes as *model-in-the-data-space*.
 The fitted model should be overlaid on the data, to examine the fit relative the spread of the observations.

### 2025-06-24 extra/paper_original.tex
Move data files to extra

--- 2024-04-07 paper_original.tex
@@
 
-In terms of model error, some data points show high error (see
-Figure~\ref{fig-nldervisMNISTPACMAP}) due to their deviation from the
-typical high-D data structure, acting as outliers. This is evident when
-comparing the images associated with these high model error points (see
-Figure~\ref{fig-sampleimgerror}) to others (see
-Figure~\ref{fig-sampleimg}).
+There are certain data points that exhibit high error rates due to their
+deviation from the usual high-dimensional data structure, which makes
+them outliers. These outliers can be classified into two types: those
+that are outliers within the nonlinear structure and those that lie
+outside of it. The images associated with high model error points within
+the nonlinear structure display different patterns of the digit 1, as
+shown in Figure~\ref{fig-sampleimgerrorwithin}. However, when comparing
+these images to the ones found outside of the nonlinear structure, it
+becomes evident that the latter display different patterns of the digit
+1 (see Figure~\ref{fig-sampleimgerror}).
+Figure~\ref{fig-nldervisMNISTPACMAP} shows some data points that have
+high error rates due to their outlier status.
 

### 2025-06-24 extra/one_curvy_and_one_gau/08_gen_model_for_best_PaCMAP_layout.R

@@ -1,0 +1,209 @@
+## Import necessary libraries
+library(quollr)
+library(dplyr)
+library(readr)
+library(langevitour)
+library(munsell)
+
+## Import data
+training_data_two_nonlinear_clusters <- read_rds("data/one_curvy_one_gau_clust/one_curvy_one_gau_clust_data.rds")
+
+# training_data_two_nonlinear_clusters <- training_data_two_nonlinear_clusters |>
+#   mutate(across(everything(), ~ (. - mean(.)) / sd(.)))
+
+training_data_two_nonlinear_clusters <- training_data_two_nonlinear_clusters |>
+  mutate(ID = 1:NROW(training_data_two_nonlinear_clusters))
+
+PaCMAP_two_nonlinear_clusters <- read_rds("data/one_curvy_one_gau_clust/one_curvy_one_gau_clust_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")
+
+two_nonlinear_clusters_scaled_obj <- gen_scaled_data(
+  data = PaCMAP_two_nonlinear_clusters)
+PaCMAP_two_nonlinear_clusters_scaled <- two_nonlinear_clusters_scaled_obj$scaled_nldr |>
+  mutate(ID = 1:NROW(PaCMAP_two_nonlinear_clusters))
+
+## Compute hexbin parameters
+num_bins_x_two_nonlinear_clusters <- 15
+lim1 <- two_nonlinear_clusters_scaled_obj$lim1
+lim2 <- two_nonlinear_clusters_scaled_obj$lim2
+r2_two_nonlinear_clusters <- diff(lim2)/diff(lim1)
+
+two_nonlinear_clusters_model <- fit_highd_model(
+  training_data = training_data_two_nonlinear_clusters,
+  emb_df = PaCMAP_two_nonlinear_clusters_scaled,
+  bin1 = num_bins_x_two_nonlinear_clusters,
+  r2 = r2_two_nonlinear_clusters,
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = FALSE,
+  col_start_highd = "x"
+)
+
+df_bin_centroids_two_nonlinear_clusters <- two_nonlinear_clusters_model$df_bin_centroids
+df_bin_two_nonlinear_clusters <- two_nonlinear_clusters_model$df_bin
+
+## Triangulate bin centroids
[…]

### 2025-06-24 extra/one_gau_cluster/3_gen_model_in_data_space_with_pacmap.R

@@ -1,0 +1,90 @@
+## Import necessary libraries
+library(quollr)
+library(dplyr)
+library(reader)
+library(langevitour)
+
+training_data_gau <- read_rds("data/one_gau_cluster/data_one_gau_training.rds")
+
+pacmap_data_gau <- read_rds("data/one_gau_cluster/pacmap_data_one_gau.rds")
+gau1_scaled_obj <- gen_scaled_data(
+  data = pacmap_data_gau)
+pacmap_gau_scaled <- gau1_scaled_obj$scaled_nldr
+
+## Compute hexbin parameters
+num_bins_x_gau1 <- 21
+lim1 <- gau1_scaled_obj$lim1
+lim2 <- gau1_scaled_obj$lim2
+r2_gau1 <- diff(lim2)/diff(lim1)
+
+gau1_model <- fit_highd_model(
+  training_data = training_data_gau,
+  emb_df = pacmap_gau_scaled,
+  bin1 = num_bins_x_gau1,
+  r2 = r2_gau1,
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = FALSE,
+  col_start_highd = "x",
+  q = 0.18
+)
+
+df_bin_centroids_gau1 <- gau1_model$df_bin_centroids
+df_bin_gau1 <- gau1_model$df_bin
+
+## Triangulate bin centroids
+tr1_object_gau1 <- tri_bin_centroids(
+  df_bin_centroids_gau1, x = "c_x", y = "c_y")
+tr_from_to_df_gau1 <- gen_edges(
+  tri_object = tr1_object_gau1)
+
+## Compute 2D distances
+distance_gau1 <- cal_2d_dist(
+  tr_coord_df = tr_from_to_df_gau1,
+  start_x = "x_from",
+  start_y = "y_from",
+  end_x = "x_to",
+  end_y = "y_to",
+  select_vars = c("from", "to", "distance"))
+
+## To find the benchmark value
+benchmark_gau1 <- find_lg_benchmark(
+  distance_edges = distance_gau1,
+  distance_col = "distance")
+
+## Hexagonal binning to have regular hexagons
+hb_obj_gau1 <- hex_binning(
+  data = pacmap_gau_scaled,
+  bin1 = num_bins_x_gau1,
+  r2 = r2_gau1,
+  q = 0.18)
+
+pacmap_data_with_hb_id <- hb_obj_gau1$data_hb_id
+
[…]

### 2025-06-24 extra/scurve/1_gen_embeddings.R

@@ -1,0 +1,121 @@
+library(dplyr)
+library(rsample)
+library(DT)
+library(ggbeeswarm)
+library(ggplot2)
+library(readr)
+
+library(Rtsne)
+library(umap)
+library(phateR)
+library(reticulate)
+library(patchwork)
+
+library(grid)
+
+set.seed(20240110)
+
+use_python("~/miniforge3/envs/pcamp_env/bin/python")
+use_condaenv("pcamp_env")
+
+reticulate::source_python(paste0(here::here(), "/examples/function_scripts/Fit_PacMAP_code.py"))
+reticulate::source_python(paste0(here::here(), "/examples/function_scripts/Fit_TriMAP_code.py"))
+
+source("nldr_code.R", local = TRUE)
+
+training_data <- read_rds("data/s_curve/s_curve_training.rds")
+
+# training_data <- training_data |>
+#   mutate(ID = row_number())
+
+## tSNE
+
+tSNE_data <- Fit_tSNE(training_data |> select(-ID), opt_perplexity = calculate_effective_perplexity(training_data), with_seed = 20240110)
+
+tSNE_data <- tSNE_data |>
+  select(-ID) |>
+  mutate(ID = training_data$ID)
+
+write_rds(tSNE_data, file = paste0("data/s_curve/s_curve_tsne_", calculate_effective_perplexity(training_data), ".rds"))
+
+## UMAP
+
+UMAP_fit <- umap(training_data |> dplyr::select(-ID), n_neighbors = 50, n_components =  2)
+
+UMAP_data <- UMAP_fit$layout |>
+  as.data.frame()
+names(UMAP_data)[1:(ncol(UMAP_data))] <- paste0(rep("UMAP",(ncol(UMAP_data))), 1:(ncol(UMAP_data)))
+
+UMAP_data <- UMAP_data |>
+  mutate(ID = training_data$ID)
+
+## Run only once
+write_rds(UMAP_data, file = "data/s_curve/s_curve_umap.rds")
+
+predict_UMAP_df <- predict(UMAP_fit, test_data |> dplyr::select(-ID)) |>
+  as.data.frame()
+
+names(predict_UMAP_df)[1:(ncol(predict_UMAP_df))] <- paste0(rep("UMAP",(ncol(predict_UMAP_df))), 1:(ncol(predict_UMAP_df)))
+
[…]

### 2025-06-24 extra/scurve/s_curve_noise.R

@@ -1,0 +1,70 @@
+library(umap)
+library(rsample)
+library(scales)
+
+set.seed(20230531)
+
+## To generate S-curve data
+s_curve <- function(n_samples = 100) {
+  tt <- 3 * pi * stats::runif(n = n_samples, min = -0.5, max = 0.5)
+  x <- sin(tt)
+  y <- 2.0 * stats::runif(n = n_samples)
+  z <- sign(tt) * (cos(tt) - 1)
+  X <- cbind(x, y, z)
+
+  data.frame(X, stringsAsFactors = FALSE)
+}
+
+# Simulate some s_curve_noise
+
+sample_size <- 100
+s_curve_noise <- s_curve(n_samples = sample_size)
+names(s_curve_noise) <- c("x1", "x2", "x3")
+
+s_curve_noise$x4 <- runif(sample_size, -0.02, 0.02)
+s_curve_noise$x5 <- runif(sample_size, -0.02, 0.02)
+s_curve_noise$x6 <- runif(sample_size, -0.1, 0.1)
+s_curve_noise$x7 <- runif(sample_size, -0.01, 0.01)
+
+## Add the ID to the s_curve_noise
+s_curve_noise <- s_curve_noise |>
+  dplyr::mutate(ID = dplyr::row_number()) |>
+  tibble::as_tibble()
+
+## Split the s_curve_noise as training and test
+data_split <- initial_split(s_curve_noise)
+s_curve_noise_training <- training(data_split) |>
+  dplyr::arrange(ID) |>
+  tibble::as_tibble()
+
+s_curve_noise_test <- testing(data_split) |>
+  dplyr::arrange(ID) |>
+  tibble::as_tibble()
+
+## Fit umap
+UMAP_fit <- umap(s_curve_noise_training |> dplyr::select(-ID), n_neighbors = 15, n_components =  2)
+
+s_curve_noise_umap <- UMAP_fit$layout |>
+  as.data.frame() |>
+  tibble::as_tibble()
+
+names(s_curve_noise_umap)[1:(ncol(s_curve_noise_umap))] <- paste0(rep("UMAP",(ncol(s_curve_noise_umap))), 1:(ncol(s_curve_noise_umap)))
+
+s_curve_noise_umap <- s_curve_noise_umap |>
+  dplyr::mutate(ID = s_curve_noise_training$ID)
+
+## scaled 2D embeddings
+
[…]

### 2025-06-24 extra/true_model_s_curve/1_gen_true_model_s_curve.R

@@ -1,0 +1,597 @@
+training_data <- read_rds("data/s_curve/s_curve_training.rds") |>
+  dplyr::select(-ID) |>
+  dplyr::mutate(type = "data")
+
+corner_points <- tibble::tibble(x1 = c(min(training_data$x1), min(training_data$x1), ## added
+                                       max(training_data$x1), max(training_data$x1), ## added
+                                       mean(training_data$x1), mean(training_data$x1),
+                                       mean(training_data$x1), mean(training_data$x1),
+                                       mean(training_data$x1), mean(training_data$x1),
+                                       sort(training_data$x1)[210], sort(training_data$x1)[210],
+                                       sort(training_data$x1)[537], sort(training_data$x1)[537],
+                                       sort(training_data$x1)[665], sort(training_data$x1)[665],
+                                       sort(training_data$x1)[98], sort(training_data$x1)[98],
+                                       sort(training_data$x1)[280], sort(training_data$x1)[280],
+                                       sort(training_data$x1)[665], sort(training_data$x1)[537],
+                                       sort(training_data$x1)[665], sort(training_data$x1)[537],
+                                       sort(training_data$x1)[665], sort(training_data$x1)[537]),
+                                x2 = c(min(training_data$x2), max(training_data$x2),
+                                       min(training_data$x2), max(training_data$x2),
+                                       min(training_data$x2), max(training_data$x2),
+                                       min(training_data$x2), max(training_data$x2),
+                                       min(training_data$x2), max(training_data$x2),
+                                       min(training_data$x2), max(training_data$x2),
+                                       min(training_data$x2), max(training_data$x2),
+                                       min(training_data$x2), max(training_data$x2),
+                                       min(training_data$x2), max(training_data$x2),
+                                       min(training_data$x2), max(training_data$x2),
+                                       mean(training_data$x2), mean(training_data$x2),
+                                       min(training_data$x2), min(training_data$x2),
+                                       max(training_data$x2), max(training_data$x2)),
[…]

### 2025-06-24 extra/two_c_shaped_clust/05_validate_nldr_layouts.R

@@ -1,0 +1,48 @@
+library(dplyr)
+library(readr)
+library(ggplot2)
+
+tsne_df <- read_rds(here::here("data/two_c_shaped_clust/two_c_shaped_clust_tsne_perplexity_30.rds"))
+umap_df <- read_rds(here::here("data/two_c_shaped_clust/two_c_shaped_clust_umap_n-neigbors_15_min-dist_0.1.rds"))
+phate_df <- read_rds(here::here("data/two_c_shaped_clust/two_c_shaped_clust_phate_knn_5.rds"))
+trimap_df <- read_rds(here::here("data/two_c_shaped_clust/two_c_shaped_clust_trimap_n-inliers_12_n-outliers_4_n-random_3.rds"))
+pacmap_df <- read_rds(here::here("data/two_c_shaped_clust/two_c_shaped_clust_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds"))
+
+
+tsne_df <- tsne_df |>
+  mutate(method = "tSNE")
+names(tsne_df) <- c("emb_1", "emb_2", "method")
+
+umap_df <- umap_df |>
+  mutate(method = "UMAP")
+names(umap_df) <- c("emb_1", "emb_2", "method")
+
+phate_df <- phate_df |>
+  mutate(method = "PHATE")
+names(phate_df) <- c("emb_1", "emb_2", "method")
+
+trimap_df <- trimap_df |>
+  mutate(method = "TriMAP")
+names(trimap_df) <- c("emb_1", "emb_2", "method")
+
+pacmap_df <- pacmap_df |>
+  mutate(method = "PaCMAP")
+names(pacmap_df) <- c("emb_1", "emb_2", "method")
+
+nldr_df <- bind_rows(tsne_df,
+                     umap_df,
+                     phate_df,
+                     pacmap_df,
+                     trimap_df)
+
+
+ggplot(data = nldr_df,
+       aes(
+         x = emb_1,
+         y = emb_2
+       )) +
+  geom_point(
+    alpha = 0.5
+  ) +
+  facet_wrap(~factor(method, levels = c("tSNE", "UMAP", "PHATE", "TriMAP", "PaCMAP")),
+             scales = "free")

### 2025-06-24 extra/two_c_shaped_clust/07_find_the_best_layout.R

@@ -1,0 +1,52 @@
+library(ggplot2)
+library(dplyr)
+library(readr)
+
+conflicted::conflicts_prefer(dplyr::filter)
+
+error_two_curvy_diff_clust_umap <- read_rds("data/two_curvy_diff_clust/error_two_curvy_diff_clust_umap.rds")
+error_two_curvy_diff_clust_tsne <- read_rds("data/two_curvy_diff_clust/error_two_curvy_diff_clust_tsne.rds")
+error_two_curvy_diff_clust_phate <- read_rds("data/two_curvy_diff_clust/error_two_curvy_diff_clust_phate.rds")
+error_two_curvy_diff_clust_trimap <- read_rds("data/two_curvy_diff_clust/error_two_curvy_diff_clust_trimap.rds")
+error_two_curvy_diff_clust_pacmap <- read_rds("data/two_curvy_diff_clust/error_two_curvy_diff_clust_pacmap.rds")
+
+error_two_curvy_diff_clust <- bind_rows(error_two_curvy_diff_clust_umap,
+                                   error_two_curvy_diff_clust_tsne,
+                                   error_two_curvy_diff_clust_phate,
+                                   error_two_curvy_diff_clust_trimap,
+                                   error_two_curvy_diff_clust_pacmap)
+
+error_two_curvy_diff_clust <- error_two_curvy_diff_clust |>
+  mutate(a1 = round(a1, 2)) |>
+  filter(bin1 >= 5) |>
+  group_by(method, a1) |>
+  filter(MSE == min(MSE)) |>
+  ungroup()
+
+error_plot_two_curvy_diff_clust <- ggplot(error_two_curvy_diff_clust,
+                                     aes(x = a1,
+                                         y = MSE,
+                                         colour = method)) +
+  geom_point(size = 0.8) +
+  geom_line(linewidth = 0.3) +
+  # geom_vline(xintercept = 15, linetype="solid",
+  #            color = "black", linewidth=0.8, alpha = 0.5) +
+  scale_x_continuous(breaks = sort(unique(error_two_curvy_diff_clust$a1))[seq(1, length(unique(error_two_curvy_diff_clust$a1)), by = 5)]) +
+  scale_color_manual(values=c('#e41a1c','#ff7f00','#4daf4a', "#a65628",'#636363')) +
+  scale_y_log10() +
+  ylab("log(MSE)") +
+  xlab(expression(paste("binwidth (", a[1], ")"))) +
+  theme_minimal() +
[…]

### 2025-06-24 extra/two_c_shaped_clust/08_gen_model_for_best_UMAP_layout.R

@@ -1,0 +1,212 @@
+## Import necessary libraries
+library(quollr)
+library(dplyr)
+library(readr)
+library(langevitour)
+library(conflicted)
+library(ggplot2)
+library(colorspace)
+
+conflicts_prefer(dplyr::filter)
+
+## Import data
+training_data_two_nonlinear_clusters <- read_rds("data/two_c_shaped_clust/two_c_shaped_clust_data.rds")
+
+# training_data_two_nonlinear_clusters <- training_data_two_nonlinear_clusters |>
+#   mutate(across(everything(), ~ (. - mean(.)) / sd(.)))
+
+training_data_two_nonlinear_clusters <- training_data_two_nonlinear_clusters |>
+  mutate(ID = 1:NROW(training_data_two_nonlinear_clusters))
+
+UMAP_two_nonlinear_clusters <- read_rds("data/two_c_shaped_clust/two_c_shaped_clust_umap_n-neigbors_15_min-dist_0.1.rds")
+
+two_nonlinear_clusters_scaled_obj <- gen_scaled_data(
+  data = UMAP_two_nonlinear_clusters)
+UMAP_two_nonlinear_clusters_scaled <- two_nonlinear_clusters_scaled_obj$scaled_nldr |>
+  mutate(ID = 1:NROW(UMAP_two_nonlinear_clusters))
+
+## Compute hexbin parameters
+num_bins_x_two_nonlinear_clusters <- 15
+lim1 <- two_nonlinear_clusters_scaled_obj$lim1
+lim2 <- two_nonlinear_clusters_scaled_obj$lim2
+r2_two_nonlinear_clusters <- diff(lim2)/diff(lim1)
+
+two_nonlinear_clusters_model <- fit_highd_model(
+  training_data = training_data_two_nonlinear_clusters,
+  emb_df = UMAP_two_nonlinear_clusters_scaled,
+  bin1 = num_bins_x_two_nonlinear_clusters,
+  r2 = r2_two_nonlinear_clusters,
+  is_bin_centroid = TRUE,
+  is_rm_lwd_hex = FALSE,
+  col_start_highd = "x"
+)
+
+df_bin_centroids_two_nonlinear_clusters <- two_nonlinear_clusters_model$df_bin_centroids
+df_bin_two_nonlinear_clusters <- two_nonlinear_clusters_model$df_bin
+
+## Triangulate bin centroids
[…]

### 2025-06-24 motivationletter/motivationletter.qmd
Initial work on the motivation letter

---
author: Jayani P. Gamage
qualifications: B.Sc. (Hons) in Statistics
position: PhD student
www: jayanilakshika.netlify.app/
email: jayani.piyadigamage@monash.edu
phone: +61 4 5229 8443
#signature: sigfile.png
address:
  - Editors
  - Journal of Computational and Graphical Statistics
opening: "Dear Professor Yuguo Chen and Professor Laura M. Sangalli,"
closing: "Sincerely"
linestretch: 1.4
format: letter-pdf
---

\bigskip
Please consider our manuscript "Looking at Non-Linear Dimension Reductions as Models in the Data Space" for publication in the Journal of Computational and Graphical Statistics.

Thank you for the consideration of this manuscript. We believe that it is a good fit for the Journal
of Computational and Graphical Statistics.

### 2025-06-24 paper.qmd
Added another parameter choice with tSNE

@@
 # and plot them on same aspect ratio
-tsne_two_curvy <- read_rds("data/two_nonlinear/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_30.rds")
+tsne_two_curvy <- read_rds("data/two_nonlinear/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_47.rds")
 

### 2025-06-24 paper.qmd
Arranged the scripts

@@
   geom_point(alpha=0.1, size=1, colour='#636363') +
-  interior_annotation("b", c(0.08, 0.9)) +
+  interior_annotation("b") +
   theme(aspect.ratio = 1)

### 2025-06-24 scripts/two_nonlinear/02_gen_true_model.R

@@ -164,7 +164,7 @@ stats <- read_rds(here::here("data/two_nonlinear/two_non_linear_diff_shaped_clos
 write_rds(model_data, "data/two_nonlinear/two_non_linear_diff_shaped_close_clusters_true_model.rds")
 write_rds(connections_all, "data/two_nonlinear/two_non_linear_diff_shaped_close_clusters_true_model_connections.rds")
 
-data <- read_rds(here::here("data/two_nonlinear/two_non_linear_diff_shaped_close_clusters_data_without_std.rds"))
+data <- read_rds(here::here("data/two_nonlinear/two_non_linear_diff_shaped_close_clusters_data.rds"))
 
 df <- bind_rows(model_data |> dplyr::mutate(type = "model"),
                 data |> dplyr::mutate(type = "data"))

### 2025-06-24 paper.qmd
Updated Figure 4b

@@
     line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
-    plot_limits = c(-0.6, 0.6), 
+    plot_limits = c(-0.75, 0.75), 
     axis_text_size = 2, 
@@
                                  "#BB1B00", "#A70C00", "#93001C", "#7D0025")) +
-  xlim(c(-0.6, 0.6)) +
-  ylim(c(-0.6, 0.6)) +
+  xlim(c(-0.75, 0.75)) +
+  ylim(c(-0.75, 0.75)) +
   interior_annotation(label = "c", cex = 1.5) +

### 2025-06-25 appendix.qmd
working on fig 7

@@
 ```
+
+## Generating the nonlinear clusters data
 
@@
 
-Points are assigned to the bin they fall into based on the nearest centroid. If a point is equidistant from multiple centroids, it is assigned to the centroid with the lowest hexagonal bin ID.
+Points are assigned to the bin they fall into based on the nearest centroid. If a point is equidistant from multiple centroids, it is assigned to the centroid with the smallest bin ID.
 
@@
 #| label: fig-assign-data
-#| fig-cap: "Binning the data. Points are assigned to the nearest centroid. If a point is equidistant from multiple centroids, assigned to the lowest centroid."
+#| fig-cap: "Binning the data. Points are assigned to the nearest centroid. If a point is equidistant from multiple centroids, assigned to the centroid with the smallest bin ID."
 #| fig-width: 9
@@
 
-Following the process in @xia2023 again using the Human Peripheral Blood Mononuclear Cells (PBMC) data.
-Note that @xia2023 uses a different subset from the PBMC dataset used by @chen2024, which contains $31,021$ cells including cell type labels, and the gene expression levels were in the unit of log-transformed UMI count per $10,000$.
+Note that @xia2023 uses a different subset from the PBMC dataset used by @chen2024, shown by us in the main paper example, which is why this comparison is shown here and not in the main paper.
+Their data contains $31,021$ cells including cell type labels, and the gene expression levels were in the unit of log-transformed UMI count per $10,000$.
 They focused on three sequencing methods (inDrops, DropSeq, and SeqWell) and four common cell types Cytotoxic T cell, CD4+T cell, CD14+ Monocyte, and B cell.
+Pre-processing follows the process in @xia2023 again using the Human Peripheral Blood Mononuclear Cells (PBMC) data.
 

### 2025-06-25 paper.qmd
working on fig 7

--- 2025-06-24 paper.qmd
@@
 Both tNSE and UMAP can be considered to produce the \kD{} representation by minimizing the divergence between two distributions, where the distributions are modeling the inter-point distances.
-PHATE, TriMAP and PaCMAP are examples of diffusion processes [@coifman2005] spreading to capture geometric shapes, that include both global and local structure.
+PHATE are examples of diffusion processes spreading to capture geometric shapes, that include both global and local structure.
+(See @coifman2005 for an explanation of diffusion processes.) TriMAP and PaCMAP
 
@@
 It is possible to see whether the elliptical shapes appropriately matches the variance of the relevant clusters, and to compare and contrast different fits.
-For PCA, one can display the \kD{} plane of the reduced dimension using wireframes of transformed cubes.
+For PCA, one can display the model (a \kD{} plane of the reduced dimension) using wireframes of transformed cubes.
 Using a wireframe is the approach we take here, to represent the NLDR model in \pD{}.
@@
 The \pD{}   observations are the realization of the phenomenon, and the \kD{} NLDR layout is the simplified representation.
-From a statistical perspective we can consider the distances between points in the \kD{}   layout to be variance that the model explains, and the (relative) difference with their distances in \pD{}   is the error, or unexplained variance.
+Typically, $k=2$, which is used for the rest of this paper.
+From a statistical perspective we can consider the distances between points in the \gD{}   layout to be variance that the model explains, and the (relative) difference with their distances in \pD{}   is the error, or unexplained variance.
 We can also imagine that the positioning of points in \gD{}    represent the fitted values, that will have some prescribed position in \pD{}   that can be compared with their observed values.
@@
 
-We define the NLDR as a function $g\text{:}~ \mathbb{R}^{n\times p} \rightarrow \mathbb{R}^{n\times k}$, with hyper-parameters $\mathbfit{\theta}$.
+We define the NLDR as a function $g\text{:}~ \mathbb{R}^{n\times p} \rightarrow \mathbb{R}^{n\times 2}$, with hyper-parameters $\mathbfit{\theta}$.
 These parameters, $\mathbfit{\theta}$, depend on the choice of $g$, and can be considered part of model fitting in the traditional sense.
@@
                                  "#BB1B00", "#A70C00", "#93001C", "#7D0025")) +
-  xlab(expression(e[hj])) +
+  xlab(expression(e[hi])) +
   ylab("") +
@@
 
-#### Computing hexagon grid configuration
+#### Hexagon grid configuration
 
@@
 When the data has been binned the triangulation connects centroids.
-The edges preserve the neighborhood information from the \kD{} representation when the model is lifted into \pD{}.
+The edges preserve the neighborhood information from the \gD{} representation when the model is lifted into \pD{}.
 
@@
 We use the \pD{} mean of the points in a given hexagon, $H_h$, denoted $C_{h}^{(p)}$, to map the centroid $C_{h}^{(2)} = (c_{h1}, c_{h2})$ to a point in \pD{}.
-Let the \pD{} mean be
-
-$$C_{h}^{(p)} = \frac{1}{n_h}\sum_{i =1}^{n_h} x_i, h = {1, \dots, b; n_h > 0}.$$
+Let the $j^{th}$ component of the \pD{} mean be
+
+$$C_{hj}^{(p)} = \frac{1}{n_h}\sum_{i =1}^{n_h} x_{hij}, ~~~h = 1, \dots, b;~ j=1, \dots, p; ~ n_h > 0.$$
 
@@
 
-The model here is similar to a confirmatory factor analysis model [@brown2015], $\widehat{T}(X_1, X_2, X_3, X_4) + \Epsilon$.
-The difference between the fitted model and observed values would be considered to be residuals, and for this problem are \fD{}.
+The model here is similar to a confirmatory factor analysis model [@brown2015], $\widehat{T} + \Epsilon$.
+The difference between the fitted model and observed values would be considered to be residuals, and are \pD{}.
 
@@
 
-$$\sqrt{\frac{1}{n}\sum_{h = 1}^{b}\sum_{i = 1}^{n_h}\sum_{j = 1}^{p} (\mathbfit{x}_{hij} - C^{(p)}_{hj})^2}$${#eq-equation1}
-
-where $n$ is the number of observations, $b$ is the number of bins, $n_h$ is the number of observations in $h^{th}$ bin, $p$ is the number of variables and $\mathbfit{x}_{hij}$ is the $j^{th}$ dimensional data of $i^{th}$ observation in $h^{th}$ hexagon.
-We can consider $e_{hj} = \sqrt{\sum_{j = 1}^{p} (\mathbfit{x}_{hij} - C^{(p)}_{hj})^2}$ to be the residual for each observation.
+$$\sqrt{\frac{1}{n}\sum_{h = 1}^{m}\sum_{i = 1}^{n_h}\sum_{j = 1}^{p} (\mathbfit{x}_{hij} - C^{(p)}_{hj})^2}$${#eq-equation1}
+
+where $n$ is the number of observations, $m$ is the number of non-empty bins, $n_h$ is the number of observations in $h^{th}$ bin, $p$ is the number of variables and $\mathbfit{x}_{hij}$ is the $j^{th}$ dimensional data of $i^{th}$ observation in $h^{th}$ hexagon.
+We can consider $e_{hi} = \sqrt{\sum_{j = 1}^{p} (\mathbfit{x}_{hij} - C^{(p)}_{hj})^2}$ to be the residual for each observation.
 
@@
 #| fig-cap: "Examining the distribution of residuals in a jittered dotplot (a), \\gD{} NLDR layout (b) and a tour of \\fD{} data space (c).
-Color indicates error ($e_{hj}$), dark color indicating high error and light indicates low error.
+Color indicates error ($e_{hi}$), dark color indicating high error and light indicates low error.
 Most large errors are distributed in one cluster (bottom one in c) and most small errors are distributed in the other cluster."
@@
 
-A new benefit of this fitted model is that it allows us to now predict a new observation's value in the NLDR, for any method.
+A new benefit of this fitted model is that it allows us to now predict the NLDR value of a new observation, $x'$, for any method.
 The steps are to determine the closest bin centroid in \pD{}, $C^{(p)}_{h}$ and predict it to be the centroid of this bin in \gD{}, $C^{(2)}_{h}$.
-This can be written as, let $z(i) = \arg\min_{j = 1, \dots, b} \sqrt{\sum_{v=1}^{p}(x_{iv} - C^{(p)}_{jv})^2}$, then the new observation $i$ falls in the hexagon, $H_h = \{i| z(i) = h\}$ and the corresponding \kD{} bin centroids, $C_{h}^{(2)} = (c_{h1}, c_{h2})$.
 
@@
 The model fitting can be adjusted using these parameters: 
-
 - hexagon bin parameters
     - bottom left bin position $(s_1, \ s_2)$, 
-    - the total number of bins ($b$), 
-- bin density cutoff, to remove low-density hexagons.
+    - the number of bins in the horizontal direction ($b_1$), which controls the number of bins the vertical direction ($b_2$), total number of bins ($b$), and total number of non-empty bins ($m$).
+- bin density cutoff, to possibly remove low-density hexagons.
 
@@
 #| fig-pos: H
-#| fig-cap: "Hexbin density plots of tSNE layout of the nonlinear cluster data, using three different bin inputs: (a) $b = 240/98 \\text{ } (15, \\text{ }16)$, (b) $b = 720/215 \\text{ } (24, \\text{ }30)$, and (c) $b = 2496/549 \\text{ } (48, \\text{ }52)$.
+#| fig-cap: "Hexbin density plots of tSNE layout of the nonlinear cluster data, using three different bin specifications ($b_1, b_2, b, m$): (a) 15, 16, 240, 98, (b) 24, 30, 720, 215, and (c) 48, 52, 2496, 549.
 Color indicates standardized counts, dark indicating high count and light indicates low count.
@@
 
-#### Measurement of capturing the data shape in \gD{}
+#### Handling of low density bins
+Bin density is computed as $n_h/n$.
+To make comparisons between hexagon configurations made from different parameter choices we use the number of **non-empty** bins ($m$) rather than the total number of bins.
+
+By default, when assessing the choice of horizontal bins, $b_1$, the total number of bins is measured by the number of **non-empty** bins ($m$).
+This more accurately reflects the hexagon grid relative to the RMSE than the full number of bins in the grid.
+It may also be beneficial to remove low count bins also, in the situation where data is clustered or stringy, where the observed data is sparse.
+In order to decide if this is necessary, you would examine the distribution of bin counts, or the density which puts the counts on a standard scale.
+If there is something of a gap at low values, this would suggest a potential value to use as a cutoff.
+Alternatively, one could choose to remove based on a percentile, the bins with density in the lowest $5\%$ of all bins, for example.
+@fig-param-two-curvy (c) illustrates the effect on the model representation of removing bins below different percentages.
+
+The benchmark value for removing low-density hexagons ranges between $0$ and $1$.
+When analyzing how these benchmark values influence model performance, it's essential to observe the change in RMSE as the benchmark value increases (@fig-param-two-curvy (c)).
+The RMSE shows a gradual decrease as the benchmark value goes from $1$ to $0$.
+Evaluating this rate of increase is important.
+If the increment is not considerable, the decision might lean towards retaining low-density hexagons.
 
@@
 The relative proportion density is the ratio of the observed proportion density to the baseline proportion density.
-
-#### Removal of low density bins
-
-By default, when assessing the choice of $b_1$, the total number of bins is measured by the number of **non-empty** bins.
-This more accurately reflects the hexagon grid relative the RMSE than the full number of bins in the grid.
-It may also be beneficial to remove low count bins also, in the situation where data is clustered or stringy, where the observed data is sparse.
-In order to decide if this is necessary, you would examine the distribution of bin counts, or the density which puts the counts on a standard scale.
-If there is something of a gap at low values, this would suggest a potential value to use as a cutoff.
-Alternatively, one could choose to remove based on a percentile, the bins with density in the lowest $5\%$ of all bins, for example.
-@fig-param-two-curvy (c) illustrates the effect on the model representation of removing bins below different percentages.
-Generally, we would urge caution in removing low count bins.
-
-The benchmark value for removing low-density hexagons ranges between $0$ and $1$.
-When analyzing how these benchmark values influence model performance, it's essential to observe the change in RMSE as the benchmark value increases (@fig-param-two-curvy (c)).
-The RMSE shows a gradual decrease as the benchmark value goes from $1$ to $0$.
-Evaluating this rate of increase is important.
-If the increment is not considerable, the decision might lean towards retaining low-density hexagons.
-
 ```{r}
@@
 # and plot them on same aspect ratio
-tsne_two_curvy <- read_rds("data/two_nonlinear/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_47.rds")
+tsne_two_curvy <- read_rds("data/two_nonlinear/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_30.rds")
 

### 2025-06-25 extra/paper_original.qmd

--- 2024-04-07 paper_original.qmd
@@
 
-In terms of model error, some data points show high error (see @fig-nldervisMNISTPACMAP) due to their deviation from the typical high-D data structure, acting as outliers.
-This is evident when comparing the images associated with these high model error points (see @fig-sampleimgerror) to others (see @fig-sampleimg).
+There are certain data points that exhibit high error rates due to their deviation from the usual high-dimensional data structure, which makes them outliers.
+These outliers can be classified into two types: those that are outliers within the nonlinear structure and those that lie outside of it.
+The images associated with high model error points within the nonlinear structure display different patterns of the digit 1, as shown in @fig-sampleimgerrorwithin.
+However, when comparing these images to the ones found outside of the nonlinear structure, it becomes evident that the latter display different patterns of the digit 1 (see @fig-sampleimgerror).
+@fig-nldervisMNISTPACMAP shows some data points that have high error rates due to their outlier status.
 
@@
     4370, 1227, 4042, 2444, 836,
-    1654)), ncol = 6) +
+    1654)), ncol = 12) +
   theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),

### 2025-06-25 motivationletter/motivationletter.qmd

@@
 position: PhD student
-www: jayanilakshika.netlify.app/
+www: jayanilakshika.netlify.app
 email: jayani.piyadigamage@monash.edu
@@
 
-\bigskip
-Please consider our manuscript "Looking at Non-Linear Dimension Reductions as Models in the Data Space" for publication in the Journal of Computational and Graphical Statistics.
+Please consider our manuscript entitled *“Looking at Non-Linear Dimension Reductions as Models in the Data Space”* for publication in the *Journal of Computational and Graphical Statistics*.
 
-Thank you for the consideration of this manuscript. We believe that it is a good fit for the Journal
-of Computational and Graphical Statistics.
+Non-linear dimension reduction (NLDR) techniques, such as tSNE and UMAP, are widely used for producing low-dimensional representations of high-dimensional data.
+While these methods can provide concise and powerful summaries, they also have the potential to exaggerate random patterns or create misleading structures.
+Different NLDR methods and different choices of hyper-parameters often produce wildly varying representations from the same data, leaving analysts with difficult decisions about which, if any, of these views are accurate.
+
+Our manuscript introduces an approach that treats NLDR outputs as models and visualizes them directly in the original high-dimensional space using tours (movies of linear projections).
+This allows analysts to assess how well the NLDR model fits across all subspaces, identify mismatches, and compare the summaries provided by different methods.
+The approach we describe offer a new, visually interpretable way to diagnose NLDR results and support more trustworthy use of these techniques in high-dimensional data analysis.
+
+We believe that this work is a strong fit for *JCGS*, given the journal’s focus on innovative statistical graphics, computational methodology, and tools that help analysts better understand high-dimensional data.
+Our approach combines statistical modeling, dynamic visualization, and interactive diagnostics, contributing to the journal’s mission of advancing the use of computation and graphics in statistical practice.
+
+Thank you for the consideration of this manuscript. We believe that it is a good fit for the Journal of Computational and Graphical Statistics.

### 2025-06-25 paper.qmd

@@
 The model fitting can be adjusted using these parameters:
+
 - hexagon bin parameters
@@
 #### Handling of low density bins
-Bin density is computed as $n_h/n$.
+
+Standardised bin counts is computed as $w_h = n_h/n, ~~h=1, \dots m$, and $n$ is the number of observations.
+Density is computed as $d_h = w_h/A$ where $A$ is the area of the hexagon.
+As a measure of the denseness of a single grid, we can compute the average bin density, $\bar{d_h}$.
+This can be used to compare grids.
+
+These quantities can be used to assess and compare the models resulting from different binwidths:
+
+- RMSE should decrease when binwidth decreases, as the binning gets closer to observations being in their own bin.
+But a big drop in RMSE would indicate the lower binwidth is substantially better than the larger one.
+- Bins with no observations or a small number might be dropped from the model.
+This will create the wireframe not extending into sparse areas, or allowing for holes in the data to be better captured.
+Note that, all observations are used for the RMSE calculation, though.
+RMSE can be examined relative to dropping a fraction of the low density bins.
+- The proportion of non-empty bins is interesting to examine across different binwidths.
+A good binning should have just the right amount of bins to neatly cover the shape of the data, and no more or less.
+- Lastly, an ideal distribution of the density of a grid is uniform, in the sense that if each bin captures a similar number of observations, then it has just the right number of bins to neatly cover the shape of the data.
+To examine this, the average bin density is compared against the range of binwidths.
+
+<!-- 
 To make comparisons between hexagon configurations made from different parameter choices we use the number of **non-empty** bins ($m$) rather than the total number of bins.
@@
 If the increment is not considerable, the decision might lean towards retaining low-density hexagons.
-
-<!-- I'll be honest I've no idea what this section is -->
+-->
+
+<!-- I'll be honest I've no idea what this section is 
 The area of a hexagon is defined as $A = 3\sqrt{3}l^2/2$ where $l$ is the side length of the hexagon.
@@
 The baseline proportion is the proportion density at the smallest possible value of $a_1$.
-The relative proportion density is the ratio of the observed proportion density to the baseline proportion density.
+The relative proportion density is the ratio of the observed proportion density to the baseline proportion density.-->
+
 ```{r}

### 2025-06-25 paper.qmd
Corrected some spelling mistakes

@@
 Two projections of the \\pD{} fitted model overlaying the data are shown in b, c.
-The fit is reasonbly tight with the data in one cluster (top one in b), but slightly less so in the other cluster probably because it is in $3\\text{-}D$.
+The fit is reasonably tight with the data in one cluster (top one in b), but slightly less so in the other cluster probably because it is in $3\\text{-}D$.
 Notice also that, in the \\gD{} layout the two clusters have internal gaps which creates a model with some holes.
@@
 
-When examining the \gD{} model representations it appears to be flatter or like a pancake with some methods, especially PACMAP, when the data is structure is higher than two dimensional.
+When examining the \gD{} model representations it appears to be flatter or like a pancake with some methods, especially PaCMAP, when the data is structure is higher than two dimensional.
 A simple example of this can be seen with data simulated to contain five \fD{} Gaussian clusters.
@@
 #| fig-cap: "NLDR's organise points in the \\gD{} layout in different ways, possibly misleadingly, illustrated using three layouts: (a) tSNE, (b) UMAP, (c) PaCMAP.
-The data has five Gaussian cluster in \\fD{}.
+The data has five Gaussian clusters in \\fD{}.
 The bottom row of plots shows a \\gD{} projection from a tour on \\fD{} revealing the differences generated by the layouts on the model fits.
@@
 On the other hand, one of the dense corners of the structure shows low model error and is highlighted in the right part of the triangular shape (b1-b3).
-The tSNE layouts (a1, b1) are colored according to selected points.
+The tSNE layouts (a2, b2) are colored according to selected points.
 a2, a3 and b2, b3 present the results from linked plots that brushed the \\fD{} error points in \\gD{} and \\fD{}.
@@
 Clustering of single-cell data is used to identify groups of cells with similar expression profiles.
-NLDR often used to summarize the cluster structure.
+NLDR is often used to summarize the cluster structure.
 Usually, NLDR does not use the cluster labels to compute the layout, but uses color to represent the cluster labels when it is plotted.

### 2025-06-25 paper.qmd
Add and updated some words with Paul's comment

@@
 
-This is a benchmark single-cell data set collected on Human Peripheral Blood Mononuclear Cells (PBMC3k) as used in @pbmc2019.
+This is a benchmark single-cell RNA-Seq data set collected on Human Peripheral Blood Mononuclear Cells (PBMC3k) as used in @pbmc2019.
 Single-cell data measures the gene expression of individual cells in a sample of tissue (see  for example, @haque2017).
@@
 This layout suggests that the data has three very well separated clusters, each with an odd shape.
-The question is whether this accurately represents the cluster structure in the data, or whether they should have chosen b or c or d or e or f or g.
+The question is whether this accurately represents the cluster structure in the data, or whether they should have chosen b or c or d or e or f or g or h.
 This is what our new method can help with -- to decide which is the more accurate \gD{} representation of the cluster structure in the \pD{} data.

### 2025-06-25 paper.qmd
Updated error_two_curvy_umap to error_two_curvy_tsne

@@
 
-error_two_curvy_umap <- read_rds("data/two_nonlinear/error_two_curvy_umap.rds")
-
-mse_two_curvy_b <- ggplot(error_two_curvy_umap, 
+error_two_curvy_tsne <- read_rds("data/two_nonlinear/error_two_curvy_tsne.rds")
+
+mse_two_curvy_b <- ggplot(error_two_curvy_tsne, 
                      aes(x = a1, 
@@
   geom_point(size = 1) +
-  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_umap$a1, 2)))[c(1, 4, 7, 11, 13, 14, 15, 16)]) +
+  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_tsne$a1, 2)))[c(1, 4, 7, 11, 13, 14, 15, 16)]) +
   labs(x = expression(paste("binwidth (", a[1], ")")), y = "RMSE") +
@@
 
-prop_dens_a1 <- ggplot(error_two_curvy_umap,
+prop_dens_a1 <- ggplot(error_two_curvy_tsne,
                      aes(x = a1,
@@
   # scale_y_continuous(breaks = sort(unique(error_two_curvy$b_non_empty))[-3]) +
-  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_umap$a1, 2)))[c(1, 4, 7, 11, 13, 14, 15, 16)]) +
+  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_tsne$a1, 2)))[c(1, 4, 7, 11, 13, 14, 15, 16)]) +
   labs(x = expression(paste("binwidth (", a[1], ")")), y = paste("Relative proportion density")) +
@@
 
-a1_m_two_curvy <- ggplot(error_two_curvy_umap,
+a1_m_two_curvy <- ggplot(error_two_curvy_tsne,
                      aes(x = a1,
@@
   # scale_y_continuous(breaks = sort(unique(error_two_curvy$b_non_empty))[-3]) +
-  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_umap$a1, 2)))[c(1, 4, 7, 11, 13, 14, 15, 16)]) +
+  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_tsne$a1, 2)))[c(1, 4, 7, 11, 13, 14, 15, 16)]) +
   labs(x = expression(paste("binwidth (", a[1], ")")), y = expression("prop. non-empty bins " * bgroup("(", m / b, ")"))) +

### 2025-06-25 paper.qmd
True model equation (removed gamma)
Updated Figure 7 text

@@
 The C-shaped cluster is generated from $\theta \sim U(\text{-}3\pi/2, 0)$, $X_1 = \sin(\theta)$, $X_2 \sim U(0, 2)$ (adding thickness to the C), $X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)$, $X_4 = \cos(\theta)$.
-The other cluster is from $X_1 \sim U(0, 2)$, $X_2 \sim U(0, 3)$, $\gamma \sim U(0, 0.5)$, $X_3 = \text{-}(X_1^3 + X_2) + \gamma$, and $X_4 \sim U(0, 2)$.
+The other cluster is from $X_1 \sim U(0, 2)$, $X_2 \sim U(0, 3)$, $X_3 = \text{-}(X_1^3 + X_2)$, and $X_4 \sim U(0, 2)$.
 We would consider $T=(X_1, X_2, X_3, X_4)$ to be the geometric structure (true model) that we hope to capture.
@@
                          y = RMSE)) +
-  geom_vline(xintercept = 0.09,
+  geom_vline(xintercept = 0.03,
            colour = "#d95f02", linetype="dashed", 
@@
              linewidth=1) +
-  geom_vline(xintercept = 0.03, linetype="dotted",
+  geom_vline(xintercept = 0.08, linetype="dotted",
              colour = "#1b9e77", linewidth=1) +
@@
                          y = prop_comp)) +
-  geom_vline(xintercept = 0.09,
+  geom_vline(xintercept = 0.03,
            colour = "#d95f02", linetype="dashed", 
@@
   scale_x_continuous(breaks = sort(unique(round(error_two_curvy_tsne$a1, 2)))[c(1, 4, 7, 11, 13, 14, 15, 16)]) +
-  labs(x = expression(paste("binwidth (", a[1], ")")), y = paste("Relative proportion density")) +
+  labs(x = expression(paste("binwidth (", a[1], ")")), y = expression(paste("relative ", bar(d[h])))) +
   interior_annotation("d", position = c(0.9, 0.9), cex = 2) +
@@
                          y = prop_bins)) +
-  geom_vline(xintercept = 0.09,
+  geom_vline(xintercept = 0.03,
            colour = "#d95f02", linetype="dashed", 
@@
         axis.ticks.y = element_line(),
-        legend.position = "bottom",
+        legend.position = "none",
         legend.title = element_blank(),
@@
 A good benchmark value for these parameters is when the RMSE drops and then flattens out.
-Three binwidth choices were made: $0.03$ (orange dashed), $0.05$ (blue solid), and $0.09$ (green dotted) to investigate.
+Three binwidth choices were made: $0.03$ (orange dashed), $0.05$ (blue solid), and $0.08$ (green dotted) to investigate.
 As the binwidth increases, the proportion of non-empty bins also increases (b).
-The relative proportion density decreases and levels off (d).
+The relative $\\bar{d_h}$ decreases and levels off (d).
 Binwidth $0.05$ is chosen as the initial best binwidth for further analysis.
@@
   mse_two_curvy_lwd + prop_dens_a1 +
-  plot_layout(ncol=2, guides = "collect")
+  plot_layout(ncol=2)
 ```

### 2025-06-25 scripts/two_nonlinear/06_gen_model_with_tSNE.R

@@ -411,14 +411,14 @@ error_two_curvy_tsne <- error_two_curvy_tsne |>
   group_by(a1) |>
   filter(bin1 == min(bin1)) |>
   ungroup() |>
-  mutate(prop_dens = 1/(b*side_length^2))
+  mutate(dens = 1/(b_non_empty*side_length^2)) #m = b_non_empty
 
-base_line_prop_dens <- error_two_curvy_tsne |>
+base_line_dens <- error_two_curvy_tsne |>
   filter(a1 == min(a1)) |>
-  pull(prop_dens)
+  pull(dens)
 
 error_two_curvy_tsne <- error_two_curvy_tsne |>
-  mutate(prop_comp = prop_dens/base_line_prop_dens)
+  mutate(prop_comp = dens/base_line_dens)
 
 error_two_curvy_tsne <- error_two_curvy_tsne |>
   mutate(prop_bins = b_non_empty/b)

### 2025-06-25 paper.qmd
Updated 7c

@@
 
-mse_two_curvy_b <- ggplot(error_two_curvy_tsne, 
+rmse_two_curvy_b <- ggplot(error_two_curvy_tsne, 
                      aes(x = a1, 
@@
 
-mse_two_curvy_lwd <- ggplot(error_rm_two_curvy, 
+rmse_two_curvy_lwd <- ggplot(error_rm_two_curvy, 
                      aes(x = benchmark_rm_lwd, 
@@
     ) + 
-  scale_x_continuous(breaks = sort(unique(error_rm_two_curvy$benchmark_rm_lwd))[c(1, 3, 5, 7, 9, 11, 13, 15, 17)]) +
+  scale_x_continuous(breaks = sort(unique(error_rm_two_curvy$benchmark_rm_lwd))[c(1, 4, 7, 11, 17, 21)]) +
   scale_color_manual(values=c("#d95f02", "#1f78b4", "#1b9e77"),
@@
 
-mse_two_curvy_b + a1_m_two_curvy +
-  mse_two_curvy_lwd + prop_dens_a1 +
+rmse_two_curvy_b + a1_m_two_curvy +
+  rmse_two_curvy_lwd + prop_dens_a1 +
   plot_layout(ncol=2)

### 2025-06-25 paper.qmd
Added scripts for PBMC3k

@@
 <!--Fit the best model for author suggestion and compute error-->
+<!--pbmc3k/14_gen_model_with_UMAP.R-->
 ```{r}
@@
 
+<!--pbmc3k/11_gen_mse_for_diff_methods.R-->
 ```{r}
@@
 <!--Fit the best model and compute error-->
+<!--pbmc3k/15_gen_model_with_tSNE.R-->
 ```{r}

### 2025-06-25 scripts/pbmc3k/4_gen_tsne_diff_param.R

@@ -26,65 +26,65 @@ tSNE_data <- tSNE_data |>
 write_rds(tSNE_data, file = paste0("data/pbmc3k/pbmc_tsne_", perplexity, ".rds"))
 
 
-## Scaled data
-pbmc_scaled_obj <- gen_scaled_data(
-  data = tSNE_data)
-tsne_pbmc_scaled <- pbmc_scaled_obj$scaled_nldr
-
-
-## To initialise number of bins along the x-axis
-bin1_vec <- 2:40
-
-lim1 <- pbmc_scaled_obj$lim1
-lim2 <- pbmc_scaled_obj$lim2
-r2_pbmc <- diff(lim2)/diff(lim1)
-
-error_pbmc <- data.frame(matrix(nrow = 0, ncol = 0))
-
-for (xbins in bin1_vec) {
-
-  bin2 <- calc_bins_y(bin1 = xbins, q = 0.1, r2 = r2_pbmc)$bin2
-
-  pbmc_model <- fit_highd_model(
-    training_data = training_data_pbmc,
-    emb_df = tsne_pbmc_scaled,
-    bin1 = xbins,
-    q = 0.1,
-    r2 = r2_pbmc,
-    is_bin_centroid = TRUE,
-    is_rm_lwd_hex = FALSE,
-    col_start_highd = "PC_"
-  )
-
-  df_bin_centroids_pbmc <- pbmc_model$df_bin_centroids
-  df_bin_pbmc <- pbmc_model$df_bin
-
-  ## Compute error
-  error_df <- glance(
-    df_bin_centroids = df_bin_centroids_pbmc,
-    df_bin = df_bin_pbmc,
-    training_data = training_data_pbmc,
-    newdata = NULL,
-    type_NLDR = "tSNE",
-    col_start = "PC_") |>
-    mutate(bin1 = xbins,
-           bin2 = bin2,
-           b = bin1 * bin2,
-           b_non_empty = NROW(df_bin_centroids_pbmc))
-
-  error_pbmc <- bind_rows(error_pbmc, error_df)
-
-}
-
-
-error_pbmc <- error_pbmc |>
-  mutate(method = "tsne",
-         perplexity = perplexity)
-
-
-ggplot(error_pbmc, aes(x = b_non_empty,
-                       y = log(MSE))) +
-  geom_point() +
-  geom_line()
-
-write_csv(error_pbmc, "data/pbmc3k/error_df_tsne.csv", append = TRUE)
[…]

### 2025-06-25 paper.qmd
Added MNIST scripts

@@
 
+<!--mnist/6_gen_mse_for_diff_methods.R-->
 ```{r}
@@
 <!-- Fit the model and compute error-->
+<!--mnist/9_gen_model_with_tSNE.R-->
 ```{r}
@@
 
+<!--mnist/5_gen_images.R-->
 ```{r}

### 2025-06-25 paper.qmd
Arrange scripts

--- 2025-06-25 paper.qmd
@@
 
+<!--two_nonlinear/06_gen_model_with_tSNE.R-->
 ```{r}
@@
 
+<!--two_nonlinear/05_gen_rm_lwd_mse.R-->
 ```{r}
@@
 
+<!--two_nonlinear/04_gen_mse_for_diff_methods.R-->
 ```{r}
@@
 <!--Projections-->
+<!--five_gau_clusters/02_gen_model_with_tSNE.R-->
 ```{r}
@@
 
+<!--five_gau_clusters/03_gen_model_with_UMAP.R-->
 ```{r}
@@
 
+<!--five_gau_clusters/04_gen_model_with_PaCMAP.R-->
 ```{r}
@@
 
+<!--one_c_shaped_dens_structure/03_gen_model_with_tSNE.R-->
 ```{r}
@@
 
+<!--mnist/04_gen_mse_for_diff_methods.R-->
 ```{r}
@@
 <!-- Fit the model and compute error-->
+<!--mnist/05_gen_model_with_tSNE.R-->
 ```{r}
@@
 
+<!--mnist/03_gen_images.R-->
 ```{r}

### 2025-06-25 appendix.qmd
Moved appendix computations with quollr to scripts

@@
 ```{r}
-result_umap <- read_rds(here::here('data/pbmc3k/pbmc_scdeed_umap_results.rds'))
-```
-
-```{r}
-#| eval: false
-result_umap |>
-  kableExtra::kable(format = "latex",
-                    booktabs = TRUE,
-                    longtable = TRUE,
-                    label = "scdeedumap") |>
-  kableExtra::kable_styling(latex_options = "scale_down")
-```
-
-```{r}
-#| label: fig-n_neighbors-dubious_cells
-#| out-width: 50%
-#| fig-pos: H
-#| eval: false
-
-result_umap |>
-  filter(min.dist == 0.5) |>
-  filter(n_neighbors < 240) |>
-  ggplot(aes(
-           x = as.numeric(n_neighbors),
-           y = number_dubious_cells
-         )) +
-    geom_line() +
-    geom_point() +
-    labs(x = "n_neighbors",
-         y = "Number of dubious cells") +
-    theme_minimal()
-```
-
-```{r}
 #| label: read-pbmc-umap-layouts
@@
 names(umap_pbmc) <- c("emb1", "emb2")
-umap_pbmc <- umap_pbmc |>
-  dplyr::mutate(ID = row_number())
 
@@
 names(tsne_pbmc) <- c("emb1", "emb2")
-tsne_pbmc <- tsne_pbmc |>
-  dplyr::mutate(ID = row_number())
 
@@
 
+<!--pbmc3k/11_gen_mse_for_diff_tsne_scDEED.R-->
+<!--pbmc3k/11_gen_mse_for_diff_umap_scDEED.R-->
 ```{r}
@@
 
+<!--pbmc3k/15_gen_model_with_UMAP_scDEED.R-->
 ```{r}
 #| label: hexbin-pbmc
-training_data_pbmc <- read_rds("data/pbmc3k/pbmc_pca_50_scdeed.rds")
-training_data_pbmc <- training_data_pbmc[, 1:9] |>
-  mutate(ID = 1:NROW(training_data_pbmc))
-
-names(training_data_pbmc) <- append(paste0("x", 1:9), "ID")
-
-## Compute hexbin parameters
-num_bins_x_pbmc <- 30
-
-algo_obj_pbmc <- fit_highd_model(
-  highd_data = training_data_pbmc, 
-  nldr_data = umap_pbmc, 
-  bin1 = num_bins_x_pbmc, 
-  q = 0.1, 
-  benchmark_highdens = 5)
-
-umap_pbmc_scaled <- algo_obj_pbmc$nldr_obj$scaled_nldr
-tr_from_to_df_pbmc <- algo_obj_pbmc$trimesh_data
-df_bin_centroids_pbmc <- algo_obj_pbmc$model_2d
-df_bin_pbmc <- algo_obj_pbmc$model_highd
-
-umap_pbmc_scaled_with_cluster <- umap_pbmc_scaled |>
-  mutate(cluster = if_else(emb1 <= 0.3, "cluster1", if_else((emb1 >= 0.75) & (emb2 <= 0.25), "cluster3", "cluster2"))) |>
-  mutate(cluster = factor(cluster, levels = c("cluster1", "cluster2", "cluster3")))
+
+umap_pbmc_scaled_with_cluster <- read_rds("data/pbmc3k/umap_pbmc_scaled_with_cluster_scDEED.rds")
+
+tr_from_to_df_pbmc <- read_rds("data/pbmc3k/tr_from_to_df_pbmc_scDEED.rds")
 
@@
 ```{r}
-#| label: prep-pbmc-umap-model-proj
-
-data_pbmc <- training_data_pbmc |> 
-  select(-ID) |>
-  mutate(type = "data")
-
-df_b_pbmc <- df_bin_pbmc |>
-  dplyr::filter(hexID %in% df_bin_centroids_pbmc$hexID) |>
-  dplyr::mutate(type = "model") ## Data with summarized mean
-
-## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
-df_b_pbmc <- df_b_pbmc[match(df_bin_centroids_pbmc$hexID, df_b_pbmc$hexID),] |>
-  dplyr::select(-hexID)
-
-# Apply the scaling
-df_model_data_pbmc <- bind_rows(data_pbmc, df_b_pbmc)
-scaled_pbmc <- scale_data_manual(df_model_data_pbmc, "type") |>
-  as_tibble()
-
-scaled_pbmc_data <- scaled_pbmc |>
-  filter(type == "data") |>
-  select(-type)
-
-scaled_pbmc_data_model <- scaled_pbmc |>
-  filter(type == "model") |>
-  select(-type)
-```
-
-```{r}
-#| label: langevitour-pbmc-umap-author-proj
-#| eval: false
-
-data_pbmc_n <- data_pbmc |>
-  select(-type) |>
-  mutate(type = as.character(umap_pbmc_scaled_with_cluster$cluster))
-
-df_model_data_pbmc_n <- bind_rows(df_b_pbmc, data_pbmc_n)
-
-langevitour::langevitour(df_model_data_pbmc_n[1:(length(df_model_data_pbmc_n)-1)],
-                         lineFrom = tr_from_to_df_pbmc$from,
-                         lineTo = tr_from_to_df_pbmc$to,
-                         group = factor(df_model_data_pbmc_n$type,
-                                        c("cluster1", "cluster2", "cluster3", "model")),
-                         levelColors = c("#8dd3c7", "#fdb462", "#bebada", "#000000"))
-```
-
-```{r}
 #| label: pbmc-umap-model-proj
 
-## First projection
-projection <- cbind(
-  c(0.025588,0.002516,0.002185,0.025284,0.004030,-0.026880,-0.015326,0.014851,0.026260),
-  c(0.026495,-0.024817,0.021623,-0.014036,-0.010220,0.016634,-0.006496,-0.022928,0.016045))
-
-proj_obj1 <- get_projection(projection = projection, 
-               proj_scale = 1.23, 
-               highd_data = scaled_pbmc_data, 
-               model_highd = scaled_pbmc_data_model, 
-               trimesh_data = tr_from_to_df_pbmc, 
-               axis_param = list(limits = 0.05,
-                                 axis_scaled = 17, 
-                                 axis_pos_x = -0.03, 
-                                 axis_pos_y = -0.03, 
-                                 threshold = 0.0044))
-
-proj_obj1[["cluster"]] <- as.character(umap_pbmc_scaled_with_cluster$cluster)
+proj_obj1 <- read_rds("data/pbmc3k/proj_obj1_umap_scDEED.rds")
 
@@
 
-## Second projection
-projection <- cbind(
-  c(0.025013,0.029055,-0.003143,0.001821,-0.006208,0.019174,0.020630,0.021512,0.020281),
-  c(0.004528,0.006465,-0.042452,0.009837,0.010040,-0.023820,0.005236,0.014775,-0.017712))
-
-proj_obj2 <- get_projection(projection = projection, 
-               proj_scale = 1.25, 
-               highd_data = scaled_pbmc_data, 
-               model_highd = scaled_pbmc_data_model, 
-               trimesh_data = tr_from_to_df_pbmc, 
-               axis_param = list(limits = 0.035, 
-                                 axis_scaled = 17, 
-                                 axis_pos_x = -0.023, 
-                                 axis_pos_y = -0.023, 
-                                 threshold = 0.0027))
-
-proj_obj2[["cluster"]] <- as.character(umap_pbmc_scaled_with_cluster$cluster)
+proj_obj2 <- read_rds("data/pbmc3k/proj_obj2_umap_scDEED.rds")
 
@@
 
+<!--pbmc3k/16_gen_model_with_tSNE_scDEED.R-->
 ```{r}
@@
 
-## Compute hexbin parameters
-num_bins_x_pbmc <- 30
-
-algo_obj_pbmc <- fit_highd_model(
-  highd_data = training_data_pbmc, 
-  nldr_data = tsne_pbmc, 
-  bin1 = num_bins_x_pbmc, 
-  q = 0.1, 
-  benchmark_highdens = 5)
-
-tsne_pbmc_scaled <- algo_obj_pbmc$nldr_obj$scaled_nldr
-#tr_from_to_df_pbmc <- algo_obj_pbmc$trimesh_data
-df_bin_centroids_pbmc <- algo_obj_pbmc$model_2d
-df_bin_pbmc <- algo_obj_pbmc$model_highd
-
-## Removed some bins manually to fit a better model
-df_bin_centroids_pbmc <- df_bin_centroids_pbmc |>
-  filter(!(hexID %in% c(559, 617, 677)))
-
-tri_object <- tri_bin_centroids(
-  centroids_data = df_bin_centroids_pbmc)
-
-tr_from_to_df_pbmc <- gen_edges(tri_object = tri_object, a1 = algo_obj_pbmc$hb_obj$a1)
-tr_from_to_df_pbmc <- tr_from_to_df_pbmc <- update_trimesh_index(trimesh_data = tr_from_to_df_pbmc)
-
-tsne_pbmc_scaled_with_cluster <- tsne_pbmc_scaled |>
-  mutate(cluster = if_else((emb2 <= -emb1 + 0.8) & (emb1 < 0.5), "cluster1", if_else(emb2 >= -2 * emb1 + 1.79, "cluster3", "cluster2"))) |>
-  mutate(cluster = factor(cluster, levels = c("cluster1", "cluster2", "cluster3")))
+tsne_pbmc_scaled_with_cluster <- read_rds("data/pbmc3k/tsne_pbmc_scaled_with_cluster_scDEED.rds")
+
+tr_from_to_df_pbmc <- read_rds("data/pbmc3k/tr_from_to_df_pbmc_tsne_scDEED.rds")
 
@@
 ```{r}
-#| label: prep-pbmc-tsne-model-proj-best
-
-data_pbmc <- training_data_pbmc |> 
-  select(-ID) |>
-  mutate(type = "data")
-
-df_b_pbmc <- df_bin_pbmc |>
-  dplyr::filter(hexID %in% df_bin_centroids_pbmc$hexID) |>
-  dplyr::mutate(type = "model") ## Data with summarized mean
-
-## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
-df_b_pbmc <- df_b_pbmc[match(df_bin_centroids_pbmc$hexID, df_b_pbmc$hexID),] |>
-  dplyr::select(-hexID)
-
-# Apply the scaling
-df_model_data_pbmc <- bind_rows(data_pbmc, df_b_pbmc)
-scaled_pbmc <- scale_data_manual(df_model_data_pbmc, "type") |>
-  as_tibble()
-
-scaled_pbmc_data <- scaled_pbmc |>
-  filter(type == "data") |>
-  select(-type)
-
-scaled_pbmc_data_model <- scaled_pbmc |>
-  filter(type == "model") |>
-  select(-type)
-```
-
-```{r}
-#| label: langevitour-pbmc-tsne-author-proj-best
-#| eval: false
-
-data_pbmc_n <- data_pbmc |>
-  select(-type) |>
-  mutate(type = as.character(tsne_pbmc_scaled_with_cluster$cluster))
-
-df_model_data_pbmc_n <- bind_rows(df_b_pbmc, data_pbmc_n)
-
-langevitour::langevitour(df_model_data_pbmc_n[1:(length(df_model_data_pbmc_n)-1)],
-                         lineFrom = tr_from_to_df_pbmc$from,
-                         lineTo = tr_from_to_df_pbmc$to,
-                         group = factor(df_model_data_pbmc_n$type,
-                                        c("cluster1", "cluster2", "cluster3", "model")),
-                         levelColors = c("#8dd3c7", "#fdb462", "#bebada", "#000000"))
-```
-
-```{r}
 #| label: pbmc-tsne-model-proj-best
 
-## First projection
-projection <- cbind(
-  c(0.025588,0.002516,0.002185,0.025284,0.004030,-0.026880,-0.015326,0.014851,0.026260),
-  c(0.026495,-0.024817,0.021623,-0.014036,-0.010220,0.016634,-0.006496,-0.022928,0.016045))
-
-proj_obj1 <- get_projection(projection = projection, 
-               proj_scale = 1.23, 
-               highd_data = scaled_pbmc_data, 
-               model_highd = scaled_pbmc_data_model, 
-               trimesh_data = tr_from_to_df_pbmc, 
-               axis_param = list(limits = 0.05,
-                                 axis_scaled = 17, 
-                                 axis_pos_x = -0.03, 
-                                 axis_pos_y = -0.03, 
-                                 threshold = 0.0044))
-
-proj_obj1[["cluster"]] <- as.character(tsne_pbmc_scaled_with_cluster$cluster)
+proj_obj1 <- read_rds("data/pbmc3k/proj_obj1_tsne_scDEED.rds")
 
@@
 
-## Second projection
-projection <- cbind(
-  c(0.025013,0.029055,-0.003143,0.001821,-0.006208,0.019174,0.020630,0.021512,0.020281),
-  c(0.004528,0.006465,-0.042452,0.009837,0.010040,-0.023820,0.005236,0.014775,-0.017712))
-
-proj_obj2 <- get_projection(projection = projection, 
-               proj_scale = 1.25, 
-               highd_data = scaled_pbmc_data, 
-               model_highd = scaled_pbmc_data_model, 
-               trimesh_data = tr_from_to_df_pbmc, 
-               axis_param = list(limits = 0.035, 
-                                 axis_scaled = 17, 
-                                 axis_pos_x = -0.023, 
-                                 axis_pos_y = -0.023, 
-                                 threshold = 0.0027))
-
-proj_obj2[["cluster"]] <- as.character(tsne_pbmc_scaled_with_cluster$cluster)
+proj_obj2 <- read_rds("data/pbmc3k/proj_obj2_tsne_scDEED.rds")
 

### 2025-06-25 appendix.qmd
Updated Figure 1 with reproduable codes

@@
 ```{r}
-umap_scurve <- read_rds(file = here::here("data/s_curve/s_curve_umap.rds"))
+#| label: code-illustration
+# Code to draw illustration for notation
+## hexagon binning to have regular hexagons
 
-scurve_scaled_obj <- gen_scaled_data(
-  nldr_data = umap_scurve)
-
-sc_ltr_pos <- c(0.08, 0.96)
-sc_xlims <- c(-0.5, 1.47)
-sc_ylims <- c(-0.32, 2.1)
-```
-
-<!--Full hexagon grid with UMAP data-->
-
-```{r}
-## Compute hexbin parameters
-num_bins_x_scurve <- 7
-
-## hexagon binning to have regular hexagons
-hb_obj_scurve <- hex_binning(
-  nldr_obj = scurve_scaled_obj, 
-  bin1 = num_bins_x_scurve, 
-  q = 0.1)
-
-## Data set with all centroids
-all_centroids_df <- hb_obj_scurve$centroids
-
-## Generate all coordinates of hexagons
-hex_grid <- hb_obj_scurve$hex_poly
-
-a1_temp <- hb_obj_scurve$a1
-a2_temp <- hb_obj_scurve$a2
-l_temp <- quad(a=3, b = 2 * a2_temp, c = -(a2_temp^2 + a1_temp^2))
-
-```
-
-```{r}
-# Code to draw illustration for notation
-start_pt <- all_centroids_df |> 
-  filter(hexID == 1)
-
-d_rect <- tibble(x1min = 0, 
-                 x1max = 1,
-                 x2min = 0,
-                 x2max = diff(scurve_scaled_obj$lim2)/diff(scurve_scaled_obj$lim1))
-
-hex_grid_temp40 <- hex_grid |> 
-  filter(hex_poly_id == 53)
-
-c_start <- tibble(s1 = 0.03, s2 = -0.05)
-
-a1 <- tibble(x = all_centroids_df$c_x[4],
-             xend = all_centroids_df$c_x[5],
-             y = all_centroids_df$c_y[29],
-             yend = all_centroids_df$c_y[29],
-             label = expression(a[1]))
-a2 <- tibble(x = all_centroids_df$c_x[39],
-             xend = all_centroids_df$c_x[39],
-             y = all_centroids_df$c_y[39],
-             yend = all_centroids_df$c_y[47],
-             label = expression(a[2]))
-
-l <- tibble(x = hex_grid_temp40$x[2],
-            xend = hex_grid_temp40$x[3],
-            y = hex_grid_temp40$y[2],
-            yend = hex_grid_temp40$y[3],
-            label = expression(l))
+hex_grid_temp <- read_rds("data/two_nonlinear/hex_grid_temp.rds")
+all_centroids_df_temp <- read_rds("data/two_nonlinear/all_centroids_df_tempp.rds")
+start_pt <- read_rds("data/two_nonlinear/start_pt.rds")
+d_rect <- read_rds("data/two_nonlinear/d_rect.rds")
+a1 <- read_rds("data/two_nonlinear/a2_data.rds")
+a2 <- read_rds("data/two_nonlinear/a1_data.rds")
+l <- read_rds("data/two_nonlinear/l_data.rds")
+rect_adj <- read_rds("data/two_nonlinear/rect_adj.rds")
 
 hex_param_vis <- ggplot() + 
-    geom_polygon(data = hex_grid, 
+    geom_polygon(data = hex_grid_temp, 
                         aes(x = x, 
@@
                  fill = "white", 
-                 color = "#bdbdbd") +
-    geom_point(data = all_centroids_df, aes(
+                 colour = "#bdbdbd") +
+    geom_point(data = all_centroids_df_temp, aes(
       x = c_x, 
       y = c_y), 
-      color = "#31a354", size = 0.9) +
+      colour = "#31a354", size = 0.9) +
     geom_point(data = start_pt, aes(x = c_x, 
                                     y = c_y), 
-               color = "black") + 
+               colour = "black") + 
     geom_rect(data=d_rect, 
-              aes(xmin = x1min - c_start$s1,# - c_start$s1, 
-                  xmax = x1max - c_start$s1,# - c_start$s1, 
-                  ymin = x2min - c_start$s2,# - c_start$s2, 
-                  ymax = x2max - c_start$s2),# - c_start$s2), 
+              aes(xmin = x1min - rect_adj$x1,# - rect_adj$s1, 
+                  xmax = x1max - rect_adj$x1,# - rect_adj$s1, 
+                  ymin = x2min - rect_adj$x2,# - rect_adj$s2, 
+                  ymax = x2max - rect_adj$x2),# - rect_adj$s2), 
               fill = "white", 
-              color = "black", 
+              colour = "black", 
               alpha = 0, 
               linewidth = 0.7) +
-    geom_point(data=d_rect, aes(x=x1min - c_start$s1, 
-                                y=x2min - c_start$s2)) + 
-    geom_point(data=d_rect, aes(x=x1max - c_start$s1, 
-                                y=x2min - c_start$s2)) + 
-    geom_point(data=d_rect, aes(x=x1min - c_start$s1, 
-                                y=x2max - c_start$s2)) + 
-    annotate("text", x=d_rect$x1min - c_start$s1, 
-                     y=d_rect$x2min - c_start$s2,
+    geom_point(data=d_rect, aes(x=x1min - rect_adj$x1, 
+                                y=x2min - rect_adj$x2)) + 
+    geom_point(data=d_rect, aes(x=x1max - rect_adj$x1, 
+                                y=x2min - rect_adj$x2)) + 
+    geom_point(data=d_rect, aes(x=x1min - rect_adj$x1, 
+                                y=x2max - rect_adj$x2)) + 
+    annotate("text", x=d_rect$x1min - rect_adj$x1, 
+                     y=d_rect$x2min - rect_adj$x2,
                      label = "(0,0)", 
              hjust=-0.1, vjust=-0.3, size = 8) + 
-    annotate("text", x=d_rect$x1max - c_start$s1, 
-                     y=d_rect$x2min - c_start$s2,
+    annotate("text", x=d_rect$x1max - rect_adj$x1, 
+                     y=d_rect$x2min - rect_adj$x2,
                      label = "(0,1)", 
              hjust=1.1, vjust=-0.3, size = 8) + 
-    annotate("text", x=d_rect$x1min - c_start$s1, 
-                     y=d_rect$x2max - c_start$s2,
+    annotate("text", x=d_rect$x1min - rect_adj$x1, 
+                     y=d_rect$x2max - rect_adj$x2,
                      label = expression(group("(", 
@@
     geom_segment(data=d_rect, aes(
-      x = x1min  - c_start$s1, # 0 - 0.03, 
-      y = -0.35, 
-      xend = x1max - c_start$s1, #1 - 0.03, 
-      yend = -0.35), #-0.35),
+      x = x1min  - rect_adj$x1, # 0 - 0.03, 
+      y = -0.31, 
+      xend = x1max - rect_adj$x1, #1 - 0.03, 
+      yend = -0.31), #-0.35),
       arrow = arrow(length = unit(0.03, "npc"),
                                ends = "both"), 
-                 color = "black")+
-    annotate("text", x=0.5, y=-0.45, 
-             label = expression(r[1]), color = "black", size = 8) +
+                 colour = "black")+
+    annotate("text", x=0.5, y=-0.36, 
+             label = expression(r[1]), colour = "black", size = 8) +
     geom_segment(data=d_rect, aes(
       x = -0.25, 
-      y = x2min - c_start$s2, #0 - 0.05, 
+      y = x2min - rect_adj$x2, #0 - 0.05, 
       xend = -0.25, 
-      yend = x2max - c_start$s2), #r2 - 0.05),
+      yend = x2max - rect_adj$x2), #r2 - 0.05),
       arrow = arrow(length = unit(0.03, "npc"),
                        ends = "both"), 
-                 color = "black")+ 
-    annotate("text", x=-0.35, y=1, 
-             label = expression(r[2]), color = "black", size = 8) +
+                 colour = "black")+ 
+    annotate("text", x=-0.3, y=0.4, 
+             label = expression(r[2]), colour = "black", size = 8) +
     geom_segment(data = a1, aes(
@@
                                ends = "both"), 
-      color = "black") + # a2 = 0.1807896
+      colour = "black") + # a2 = 0.1807896
     annotate("text", x=a2$x, y=(a2$y+a2$yend)/2, 
              label = expression(a[2]), 
-             color = "black", hjust=-0.2, size = 8) +
-    annotate("text", x=-0.18, y=-0.25, 
+             colour = "black", hjust=-0.2, size = 8) +
+    annotate("text", x=-0.18, y=-0.24, 
       label = expression(group("(", list(s[1], s[2]), ")")),
-      color = "black", size = 8) +
+      colour = "black", size = 8) +
   geom_segment(data = l, aes(
@@
 ```{r}
-## Compute hexbin parameters
-num_bins_x_scurve <- 3
-
-## hexagon binning to have regular hexagons
-hb_obj_scurve <- hex_binning(
-  nldr_obj = scurve_scaled_obj, 
-  bin1 = num_bins_x_scurve, 
-  q = 0.1)
 
 ## Data set with all centroids
-all_centroids_df <- hb_obj_scurve$centroids
+all_centroids_df <- read_rds("data/two_nonlinear/two_nonlinear_all_centroids_two_curvy4.rds")
 
 ## Generate all coordinates of hexagons
-hex_grid <- hb_obj_scurve$hex_poly
+hex_grid <- read_rds("data/two_nonlinear/two_nonlinear_hex_grid_two_curvy4.rds")
 
@@
 circles <- data.frame(
-  x = c(0.726, 0.451),
-  y = c(1.27, 0.790),
-  r = rep(0.32, 2)
+  x = c(0.608, 0.372),
+  y = c(0.306, 0.715),
+  r = rep(0.27, 2)
 )
 
-emb_df1 <- tibble(emb1 = 0.61, emb2 = 0.99) |>
-  add_row(emb1 = 0.54, emb2 = 1.08)
+emb_df1 <- tibble(emb1 = 0.45, emb2 = 0.47) |>
+  add_row(emb1 = 0.54, emb2 = 0.56)
 
@@
   geom_circle(data = circles, aes(x0 = x, y0 = y, r = r), fill = NA) +
-  geom_segment(x = 0.61, y = 0.99, xend = 0.46, yend = 0.83, 
+  geom_segment(x = 0.45, y = 0.47, xend = 0.58, yend = 0.3, 
                arrow = arrow(length = unit(0.02, "npc")), color = "#969696") +
-  geom_segment(x = 0.54, y = 1.08, xend = 0.7, yend = 1.2, 
+  geom_segment(x = 0.54, y = 0.56, xend = 0.4, yend = 0.715, 
                arrow = arrow(length = unit(0.02, "npc")), color = "#969696")
+```
 
+```{r}
 hex_data <- hex_grid |> 
@@
 
+arrow_df <- tibble(
+  x = points_df$x, 
+  y = points_df$y,
+  xend = c(0.37, points_df$x[2] - 0.1, 0.136, points_df$x[4] - 0.1, 0.61, points_df$x[6] - 0.1, 0.33, points_df$x[8] - 0.1, 0,  points_df$x[10] - 0.1, 0.58, 0.45),
+  yend = c(0.82, points_df$y[2] - 0.1, 0.4, points_df$y[4] - 0.1, 0.4, points_df$y[6] - 0.1, 0.82, points_df$y[8] - 0.1, 0.715, points_df$y[10] - 0.1, 0.4, 0.71)
+)
+
 assign_data_vis2 <- ggplot(data = hex_grid, aes(x = x, y = y)) + 
@@
   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID), size = 5, colour = "#bdbdbd") +
-  geom_segment(x = 0.175, y = 0.790, xend = -0.06, yend = 0.790, 
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.726, y = 0.790, xend = 0.50, yend = 0.790,
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.313, y = 1.03, xend = 0.41, yend = 0.82,
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.588, y = 1.03, xend = 0.48, yend = 0.82,
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.313, y = 0.551, xend = 0.20, yend = 0.34,
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.588, y = 0.551, xend = 0.67, yend = 0.34,
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.451, y = 1.11, xend = 0.451, yend = 0.85,
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.175, y = 0.949, xend = -0.02, yend = 0.82,
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.175, y = 0.631, xend = 0.175, yend = 0.4,
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.451, y = 0.472, xend = 0.25, yend = 0.3,
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.726, y = 0.631, xend = 0.726, yend = 0.4,
-               arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") +
-  geom_segment(x = 0.726, y = 0.949, xend = 0.5, yend = 0.82,
+  geom_segment(data = arrow_df, aes(x = x, y = y, xend = xend, yend = yend), 
                arrow = arrow(length = unit(0.02, "npc")), color = "#bdbdbd") 
-
 ```
@@
 #| fig-cap: "Binning the data. Points are assigned to the nearest centroid. If a point is equidistant from multiple centroids, assigned to the centroid with the smallest bin ID."
-#| fig-width: 9
+#| fig-width: 12
 #| fig-height: 6
@@
 
-error_umap_plot_pbmc <-  plot_mse(error_pbmc) +
+error_umap_plot_pbmc <-  plot_rmse(error_pbmc) +
   scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 12, 15, 17, 19, 20)]) +

### 2025-06-25 appendix.tex
Updated Figure 1 with reproduable codes

@@
 $T$ & true  model\\
-\addlinespace
 $g$ & functional mapping from \pD{} to \kD{}, especially as prescribed by NLDR\\
@@
 $(b_1, b_2)$ & number of bins in each direction\\
-\addlinespace
 $(a_1, a_2)$ & binwidths, distance between centroids in each direction\\
@@
 $b$ & number of  hexagons in the grid\\
-\addlinespace
 $h$ & hexagonal id\\
@@
 $A$ & area\\
+$n_h$ & bin count\\
+$w_h$ & standardised bin counts\\
 \bottomrule
@@
 \end{table}%
+
+\section{Generating the nonlinear clusters
+data}\label{generating-the-nonlinear-clusters-data}
 
@@
 }\end{equation}
-
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=0.3\linewidth,height=\textheight,keepaspectratio]{appendix_files/figure-pdf/fig-hex-param-1.pdf}
-
-}
-
-\caption{\label{fig-hex-param}The components of the hexagon grid
-illustrating notation.}
-
-\end{figure}%
 
@@
 centroid. If a point is equidistant from multiple centroids, it is
-assigned to the centroid with the lowest hexagonal bin ID.
+assigned to the centroid with the smallest bin ID.
 
@@
 the nearest centroid. If a point is equidistant from multiple centroids,
-assigned to the lowest centroid.}
+assigned to the centroid with the smallest bin ID.}
 
@@
 
-Following the process in \citet{xia2023} again using the Human
-Peripheral Blood Mononuclear Cells (PBMC) data. Note that
-\citet{xia2023} uses a different subset from the PBMC dataset used by
-\citet{chen2024}, which contains \(31,021\) cells including cell type
-labels, and the gene expression levels were in the unit of
-log-transformed UMI count per \(10,000\). They focused on three
-sequencing methods (inDrops, DropSeq, and SeqWell) and four common cell
-types Cytotoxic T cell, CD4+T cell, CD14+ Monocyte, and B cell.
+Note that \citet{xia2023} uses a different subset from the PBMC dataset
+used by \citet{chen2024}, shown by us in the main paper example, which
+is why this comparison is shown here and not in the main paper. Their
+data contains \(31,021\) cells including cell type labels, and the gene
+expression levels were in the unit of log-transformed UMI count per
+\(10,000\). They focused on three sequencing methods (inDrops, DropSeq,
+and SeqWell) and four common cell types Cytotoxic T cell, CD4+T cell,
+CD14+ Monocyte, and B cell. Pre-processing follows the process in
+\citet{xia2023} again using the Human Peripheral Blood Mononuclear Cells
+(PBMC) data.
 

### 2025-06-25 paper.qmd
Updated Figure 1 with reproduable codes

@@
 <!--Fit the best model for author suggestion and compute error-->
-<!--pbmc3k/14_gen_model_with_UMAP.R-->
+<!--pbmc3k/13_gen_model_with_UMAP.R-->
 ```{r}
@@
 
-<!--pbmc3k/11_gen_mse_for_diff_methods.R-->
+<!--pbmc3k/08_gen_mse_for_diff_methods.R-->
 ```{r}
@@
 <!--Fit the best model and compute error-->
-<!--pbmc3k/15_gen_model_with_tSNE.R-->
+<!--pbmc3k/14_gen_model_with_tSNE.R-->
 ```{r}

### 2025-06-25 appendix.qmd
Added True model projections to appendix

@@
 ```{r}
+clr_choice <- "#0077A3"
+```
+
+```{r}
 #| label: tbl-notation
@@
 ## Generating the nonlinear clusters data
+
+```{r}
+proj_obj1_true <- read_rds("data/two_nonlinear/two_nonlinear_true_proj_obj1.rds")
+
+two_curvy_proj_true1 <- plot_proj(
+  proj_obj = proj_obj1_true,
+  point_param = c(0.5, 0.2, clr_choice), # size, alpha, color
+  line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
+  plot_limits = c(-0.75, 0.75),
+  axis_text_size = 2,
+  is_category = FALSE) +
+  interior_annotation(label = "a", cex = 1.2) +
+  theme(aspect.ratio = 1,
+        legend.position = "none")
+
+proj_obj2_true <- read_rds("data/two_nonlinear/two_nonlinear_true_proj_obj2.rds")
+
+two_curvy_proj_true2 <- plot_proj(
+  proj_obj = proj_obj2_true,
+  point_param = c(0.1, 0.5, clr_choice), # size, alpha, color
+  line_param = c(0.5, 0.5, "black"), # linewidth, alpha
+  plot_limits = c(-0.5, 0.5),
+  axis_text_size = 2,
+  is_category = FALSE) +
+  interior_annotation(label = "b", cex = 1.2) +
+  theme(aspect.ratio = 1)
+```
+
+```{r}
+#| fig-width: 6
+#| fig-height: 3
+#| fig-cap: "Two projections of the \\pD{} true model overlaying the data are shown in a, b."
+
+two_curvy_proj_true1 + two_curvy_proj_true2
+```
 

### 2025-06-26 paper.qmd
tuing section revised

@@
   labs(x = expression(paste("binwidth (", a[1], ")")), y = "RMSE") +
+  ylim(c(0, 0.75)) +
   interior_annotation("a", position = c(0.9, 0.1), cex = 2) +
@@
   scale_x_continuous(breaks = sort(unique(round(error_two_curvy_tsne$a1, 2)))[c(1, 4, 7, 11, 13, 14, 15, 16)]) +
-  labs(x = expression(paste("binwidth (", a[1], ")")), y = expression(paste("relative ", bar(d[h])))) +
+  labs(x = expression(paste("binwidth (", a[1], ")")), y = expression(paste("ave bin density (", bar(d), ")"))) +
   interior_annotation("d", position = c(0.9, 0.9), cex = 2) +
@@
 
-Default values are provided for each of these, but it is expected that the user will examine the RMSE for a range of choices.
-Choosing these parameters according to RMSE can be automated but it is recommended that the user examine the resulting model representation by overlaying it on the data in \pD{}.
-The next few subsections describe the calculation of default values, and the effect that different choices have on the model fit.
+Default values are provided for each of these, but deciding on the best model fit is assisted by examining the RMSE for a range of choices.
+<!--Choosing these parameters according to RMSE can be automated but it is recommended that the user examine the resulting model representation by overlaying it on the data in \pD{}.
+The next few subsections describe the calculation of default values, and the effect that different choices have on the model fit.-->
 
@@
 Density is computed as $d_h = w_h/A$ where $A$ is the area of the hexagon.
-As a measure of the denseness of a single grid, we can compute the average bin density, $\bar{d_h}$.
-This can be used to compare grids.
+As a measure of the denseness of a single grid, we can compute the average bin density, $\bar{d}$.
+This can be used to compare the model fit for different grids.
 
@@
 - RMSE should decrease when binwidth decreases, as the binning gets closer to observations being in their own bin.
-But a big drop in RMSE would indicate the lower binwidth is substantially better than the larger one.
+But a big drop in RMSE would indicate the lower binwidth is substantially better than the larger one (@fig-param-two-curvy a).
+- The proportion of non-empty bins is interesting to examine across different binwidths.
+A good binning should have just the right amount of bins to neatly cover the shape of the data, and no more or less (@fig-param-two-curvy b).
 - Bins with no observations or a small number might be dropped from the model.
@@
 Note that, all observations are used for the RMSE calculation, though.
-RMSE can be examined relative to dropping a fraction of the low density bins.
-- The proportion of non-empty bins is interesting to examine across different binwidths.
-A good binning should have just the right amount of bins to neatly cover the shape of the data, and no more or less.
+RMSE can be examined relative to dropping a fraction of the low density bins (@fig-param-two-curvy c).
 - Lastly, an ideal distribution of the density of a grid is uniform, in the sense that if each bin captures a similar number of observations, then it has just the right number of bins to neatly cover the shape of the data.
-To examine this, the average bin density is compared against the range of binwidths.
+To examine this, the average bin density, $\bar{d}$, is compared against the range of binwidths (@fig-param-two-curvy d).
 
@@
                      labels = parse_format()) +
-  scale_linetype_manual(values = c("dashed", "solid", "dotted"),
+  scale_linetype_manual(values = c("longdash", "solid", "dotted"),
                         labels = parse_format()) +
-  xlab("bin density cutoff") +
+  xlab(expression(paste("std bin count (", w[h], ")"))) +
   ylab("RMSE") +
+  ylim(c(0, 0.75)) +
   interior_annotation("c", position = c(0.07, 0.9), cex = 2) +
@@
 #| echo: false
-#| fig-cap: "Various plots to help assess best hexagon bin parameters (a, b, d) and thresholds to remove low-density bins (c).
-Both (a) and (c) show RMSE, against binwidth ($a_1$) and bin density cutoff.
-A good benchmark value for these parameters is when the RMSE drops and then flattens out.
-Three binwidth choices were made: $0.03$ (orange dashed), $0.05$ (blue solid), and $0.08$ (green dotted) to investigate.
-As the binwidth increases, the proportion of non-empty bins also increases (b).
-The relative $\\bar{d_h}$ decreases and levels off (d).
-Binwidth $0.05$ is chosen as the initial best binwidth for further analysis.
-There is no need to remove the low-density hexagons because as shown in (b), there is no considerable drop in RMSE."
+#| fig-cap: "Various plots to help tune the model fit, using the binwidth $a_1$ and low density bin removal.
+In (a) RMSE against $a_1$, RMSE is going to increase as $a_1$ increases, but just before a big increase is a good choice.
+Here RMSE steadily increases so there is no clear choice.
+For comparison in the other displays we have chosen $0.03$ (orange dashed), $0.05$ (blue solid), and $0.08$ (green dotted).
+Proportion of non-empty bins tends to increase with $a_1$ (b).
+Removing low count bins (up to bins with 20% of observations) doesn't change RMSE substantially (c).
+The average bin density ($\\bar{d}$) tends to decrease with $a_1$ (d)."
 #| label: fig-param-two-curvy
-#| out-width: 80%
+#| out-width: 100%
 #| fig-width: 12

### 2025-06-26 paper.qmd
tuning small change

@@
   scale_x_continuous(breaks = sort(unique(round(error_two_curvy_tsne$a1, 2)))[c(1, 4, 7, 11, 13, 14, 15, 16)]) +
-  labs(x = expression(paste("binwidth (", a[1], ")")), y = expression(paste("ave bin density (", bar(d), ")"))) +
+  labs(x = expression(paste("binwidth (", a[1], ")")), y = expression(paste("relative ave bin density (", bar(d), ")"))) +
   interior_annotation("d", position = c(0.9, 0.9), cex = 2) +
@@
 - Lastly, an ideal distribution of the density of a grid is uniform, in the sense that if each bin captures a similar number of observations, then it has just the right number of bins to neatly cover the shape of the data.
-To examine this, the average bin density, $\bar{d}$, is compared against the range of binwidths (@fig-param-two-curvy d).
+To examine this, the average bin density, $\bar{d}$, is compared against the range of binwidths (@fig-param-two-curvy d), relative to that of the hexgrid with the smallest binwidth.
 

### 2025-06-26 paper.qmd
section 3.7 and 4 retitled

@@
 
-### Linked plots
-
-Diagnosing the model while locating points in the \gD{} layout and displaying the generated model overlaid on data in the \pD{} is important.
-
-The \gD{} layout and the langevitour view with the model are linked together via rectangular brushes; when a brush is active, points will be highlighted in the adjacent view.
+### Interactive graphics
+
+Matching points in the \gD{} layout with their positions in \pD{} is useful for assessing the fit.
+This can be used to examine the fitted model in some subspaces in \pD{}, in particular in association with residual plots.
+
+The \gD{} layout and the langevitour view with the fitted model overlaid can be linked using a **browsable HTML widget**.
+A rectangular "brush" is used to select points in one plot, which will highlight the corresponding points in the other plot(s).
 Because the langevitour is dynamic, brush events that become active will pause the animation, so that a user can interrogate the current view.
-The interface is constructed as a **browsable HTML widget** specifically designed for interactive data analysis.
-
-To understand how well the model fits the points whether it fits well, works better in some positions, or fails to match the overall pattern.
-It is important to link and brush the points with high model error in the \pD{} error plot, \gD{} layout, and the generated model overlay on the data in \pD{}.
-
-## Assessing the best fit, and hence most accurate \gD{} layout {#sec-bestfit}
+This capability will be illustrated on the examples, to show how it can help to understand how the NLDR has organised the observations, and learn where it does not do well.
+
+## Choosing the best \gD{} layout {#sec-bestfit}
 

### 2025-06-26 paper.qmd
Updated levels in Figure 8

@@
   filter(RMSE == min(RMSE)) |>
-  ungroup()
+  ungroup() |>
+  mutate(method = factor(method, levels = c("PaCMAP", "PHATE", "TriMAP", "tSNE", "UMAP", "tSNE2")))
 ```

### 2025-06-26 paper.qmd
tinkering with title

--- 2025-06-26 paper.qmd
@@
 ---
-title: "Looking at Non-Linear Dimension Reductions as Models in the Data Space"
+title: "Beyond Subjectivity: New Visual Tools to Choose the Best Nonlinear Dimension Reduction Layout"
 format: 
@@
 ---
+<!-- Looking at Non-Linear Dimension Reductions as Models in the Data Space -->
 
@@
 
-### Low density results in a squishing of the \gD{} representation in all methods {#sec-effect-dens}
+### Sparseness creates a contracted \gD{} layout {#sec-effect-dens}
 
@@
 For example, the data shown in @fig-one-dens_clust-error all lies on a \gD{} curved sheet in \fD{}, but one end of the sheet is sampled densely and the other very sparsely.
-It was simulated to illustrate the effect of the density difference on layout generated by an NLDR, illustrated using the tSNE results.
+It was simulated to illustrate the effect of the density difference on layout generated by an NLDR, illustrated using the tSNE results, but it happens with all methods.
 

### 2025-06-26 paper.qmd
Updated Figure 7 (c)

--- 2025-06-26 paper.qmd
@@
   ylab("RMSE") +
-  ylim(c(0, 0.75)) +
+  #ylim(c(0, 0.75)) +
   interior_annotation("c", position = c(0.07, 0.9), cex = 2) +

### 2025-06-26 paper.qmd
Updated 7 (c)

--- 2025-06-26 paper.qmd
@@
     ) + 
-  scale_x_continuous(breaks = sort(unique(error_rm_two_curvy$benchmark_rm_lwd))[c(1, 4, 7, 11, 17, 21)]) +
+  scale_x_continuous(breaks = sort(unique(error_rm_two_curvy$benchmark_rm_lwd))[c(1, 4, 7, 11, 15, 17, 21)]) +
   scale_color_manual(values=c("#d95f02", "#1f78b4", "#1b9e77"),

### 2025-06-26 paper.qmd

--- 2025-06-26 paper.qmd
@@
 ---
-title: "Beyond Subjectivity: New Visual Tools to Choose the Best Nonlinear Dimension Reduction Layout"
+title: "Beyond Subjectivity: New Visual Tools to Choose a Nonlinear Dimension Reduction"
 format: 
@@
 ---
-<!-- Looking at Non-Linear Dimension Reductions as Models in the Data Space -->
+<!-- Looking at Non-Linear Dimension Reductions as Models in the Data Space
+
+Quantitative Visual Tools to Honestly Select Nonlinear Dimension Reductions
+
+Stop Lying to Me: Visual Evidence for Honest Dimension Reduction
+
+When Looks Deceive: Quantifying the Honesty of Nonlinear Layouts
+
+Not Just Pretty Pictures: Quantitative Visual Evaluation of Nonlinear DR Methods
+
+Beauty Isn’t Truth: Choosing Nonlinear Embeddings that Stay True to the Data
+
+Truth in the Tangle: Choosing Nonlinear Reductions with Quantitative Visual Evidence
+-->
 
@@
   geom_point(alpha=0.1, size=1, colour='#e41a1c') +
-  interior_annotation("e", c(0.08, 0.9)) +
+  interior_annotation("e") +
   theme(aspect.ratio = 1)

### 2025-06-26 paper.qmd

@@
   ylab("RMSE") +
+  xlim(c(0, 0.14)) +
   ylim(c(0, 0.75)) +
@@
         panel.border = element_rect(fill = 'transparent'),
-        plot.title = element_text(size = 20, hjust = 0.5, vjust = -0.5),
+        plot.title = element_text(size = 20, 
+                                  hjust = 0.5, 
+                                  vjust = -0.5),
         axis.ticks.x = element_line(),

### 2025-06-26 paper.qmd
fixed caption of fig 7
Try to remove bins with less number of points first

--- 2025-06-26 paper.qmd
@@
     ) + 
-  scale_x_continuous(breaks = sort(unique(error_rm_two_curvy$benchmark_rm_lwd))[c(1, 4, 7, 11, 17, 21)]) +
+  scale_x_continuous(breaks = sort(unique(error_rm_two_curvy$benchmark_rm_lwd))[c(1, 4, 7, 11, 15)]) +
   scale_color_manual(values=c("#d95f02", "#1f78b4", "#1b9e77"),
@@
 Proportion of non-empty bins tends to increase with $a_1$ (b).
-Removing low count bins (up to bins with 20% of observations) doesn't change RMSE substantially (c).
+Removing low count bins changes RMSE substantially (c).
 The average bin density ($\\bar{d}$) tends to decrease with $a_1$ (d)."

### 2025-06-26 paper.qmd
title change

--- 2025-06-26 paper.qmd
@@
 ---
-title: "Beyond Subjectivity: New Visual Tools to Choose a Nonlinear Dimension Reduction"
+title: "Stop Lying to Me: New Visual Tools to Choose the Most Honest Nonlinear Dimension Reduction"
 format: 
@@
 Truth in the Tangle: Choosing Nonlinear Reductions with Quantitative Visual Evidence
+
+Beyond Subjectivity: New Visual Tools to Choose a Nonlinear Dimension Reduction
 -->

### 2025-06-26 paper.qmd

--- 2025-06-26 paper.qmd
@@
 In (a) RMSE against $a_1$, RMSE is going to increase as $a_1$ increases, but just before a big increase is a good choice.
-Here RMSE steadily increases so there is no clear choice.
-For comparison in the other displays we have chosen $0.03$ (orange dashed), $0.05$ (blue solid), and $0.08$ (green dotted).
+Here RMSE steadily increases so there is no clear choice - nevertheless, we have chosen three binwidths, $0.03$ (orange dashed), $0.05$ (blue solid), and $0.08$ (green dotted), to compare in the other plots.
 Proportion of non-empty bins tends to increase with $a_1$ (b).

### 2025-06-26 appendix.qmd
moved data description to appendix

@@
   \newcommand\gD{$2\text{-}D$}
+  \newcommand\tD{$3\text{-}D$}
+  \newcommand\fD{$4\text{-}D$}
+  \newcommand\sD{$7\text{-}D$}
   \newcommand{\To}{\textbf{to}~}
@@
 
-## Generating the nonlinear clusters data
+## Generating the 2NC7 data
+
+This data is constructed by simulating two clusters, each consisting of $1000$ observations.
+The C-shaped cluster is generated from $\theta \sim U(\text{-}3\pi/2, 0)$, $X_1 = \sin(\theta)$, $X_2 \sim U(0, 2)$ (adding thickness to the C), $X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)$, $X_4 = \cos(\theta)$.
+Observations lie on a \gD{} manifold in \sD{}.
+The other cluster is from $X_1 \sim U(0, 2)$, $X_2 \sim U(0, 3)$, $X_3 = \text{-}(X_1^3 + X_2)$, and $X_4 \sim U(0, 2)$.
+It is also curved, but observations lie on a \tD{} manifold in\sD{}.
+Three more variables, $X_5, X_6, X_7$, that are small amounts of pure noise are added.
+We would consider $T=(X_1, X_2, X_3, X_4)$ to be the geometric structure (true model) that we hope to capture.
 

### 2025-06-26 paper.qmd
moved data description to appendix

@@
   \newcommand\gD{$2\text{-}D$}
+  \newcommand\tD{$3\text{-}D$}
   \newcommand\fD{$4\text{-}D$}
+  \newcommand\sD{$7\text{-}D$}
 ---
@@
 
-To illustrate the method, we use $7\text{-}D$ simulated data, which we call the "nonlinear clusters".
-It is constructed by simulating two clusters, each consisting of $1000$ observations.
-The C-shaped cluster is generated from $\theta \sim U(\text{-}3\pi/2, 0)$, $X_1 = \sin(\theta)$, $X_2 \sim U(0, 2)$ (adding thickness to the C), $X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)$, $X_4 = \cos(\theta)$.
-The other cluster is from $X_1 \sim U(0, 2)$, $X_2 \sim U(0, 3)$, $X_3 = \text{-}(X_1^3 + X_2)$, and $X_4 \sim U(0, 2)$.
+To illustrate the method, we use \sD{} simulated data, which we call the "2NC7" data.
+It is has two separated nonlinear clusters, one forming a \gD{} curved shape, and the other a \tD{} curved shape, each consisting of $1000$ observations.
+The first four variables hold this cluster structure, and the remaining three are purely noise.
 We would consider $T=(X_1, X_2, X_3, X_4)$ to be the geometric structure (true model) that we hope to capture.
@@
 #| echo: false
-#| fig-cap: "Key steps for constructing the model on the tSNE layout ($k=2$): (a) data, (b) hexagon bins, (c) bin centroids, and (d) triangulated centroids. The two nonlinear clusters data is shown."
+#| fig-cap: "Key steps for constructing the model on the tSNE layout ($k=2$) of 2NC7: (a) data, (b) hexagon bins, (c) bin centroids, and (d) triangulated centroids. The 2NC7 data is shown."
 #| fig-width: 12
@@
 #| fig-pos: H
-#| fig-cap: "Hexbin density plots of tSNE layout of the nonlinear cluster data, using three different bin specifications ($b_1, b_2, b, m$): (a) 15, 16, 240, 98, (b) 24, 30, 720, 215, and (c) 48, 52, 2496, 549.
+#| fig-cap: "Hexbin density plots of tSNE layout of the 2NC7 data, using three different bin specifications ($b_1, b_2, b, m$): (a) 15, 16, 240, 98, (b) 24, 30, 720, 215, and (c) 48, 52, 2496, 549.
 Color indicates standardized counts, dark indicating high count and light indicates low count.
@@
 ```{r}
-#| fig-cap: "Assessing which of the 6 NLDR layouts on the two nonlinear clusters data is the better representation using RMSE for varying binwidth ($a_1$).
+#| fig-cap: "Assessing which of the 6 NLDR layouts on the 2NC7 data is the better representation using RMSE for varying binwidth ($a_1$).
 Color used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-f).

### 2025-06-26 paper.qmd
Updated Figure 6 title with new Layout details

@@
 #| fig-pos: H
-#| fig-cap: "Hexbin density plots of tSNE layout of the 2NC7 data, using three different bin specifications ($b_1, b_2, b, m$): (a) 15, 16, 240, 98, (b) 24, 30, 720, 215, and (c) 48, 52, 2496, 549.
+#| fig-cap: "Hexbin density plots of tSNE layout of the 2NC7 data, using three different bin specifications ($b_1, b_2, b, m$): (a) 15, 18, 270, 98, (b) 24, 29, 696, 209, and (c) 35, 42, 1470, 386.
 Color indicates standardized counts, dark indicating high count and light indicates low count.

### 2025-06-26 paper.qmd
Updated Figure 7 text

@@
              linewidth=1) +
-  geom_vline(xintercept = 0.08, linetype="dotted",
+  geom_vline(xintercept = 0.07, linetype="dotted",
              colour = "#1b9e77", linewidth=1) +
@@
 In (a) RMSE against $a_1$, RMSE is going to increase as $a_1$ increases, but just before a big increase is a good choice.
-Here RMSE steadily increases so there is no clear choice - nevertheless, we have chosen three binwidths, $0.03$ (orange dashed), $0.05$ (blue solid), and $0.08$ (green dotted), to compare in the other plots.
+Here RMSE steadily increases so there is no clear choice - nevertheless, we have chosen three binwidths, $0.03$ (orange dashed), $0.05$ (blue solid), and $0.07$ (green dotted), to compare in the other plots.
 Proportion of non-empty bins tends to increase with $a_1$ (b).

### 2025-06-26 paper.qmd
curiosties section revised relative to PH comments

--- 2025-06-26 paper.qmd
@@
 
-When examining the \gD{} model representations it appears to be flatter or like a pancake with some methods, especially PaCMAP, when the data is structure is higher than two dimensional.
+The \gD{} model representations generated from some NLDR methods, especially PaCMAP, are unreasonably flat or like a pancake.
 A simple example of this can be seen with data simulated to contain five \fD{} Gaussian clusters.
@@
 #| out-width: 100%
-#| fig-cap: "Exploring the effect of density on the NLDR layout using a C-shaped structure with different density at each end.
-The top part of the triangular shape, shown with linked brushing, identify high model error corresponding to the sparse end of the structure (a1-a3).
-On the other hand, one of the dense corners of the structure shows low model error and is highlighted in the right part of the triangular shape (b1-b3).
-The tSNE layouts (a2, b2) are colored according to selected points.
-a2, a3 and b2, b3 present the results from linked plots that brushed the \\fD{} error points in \\gD{} and \\fD{}.
-It helps in identifying the high \\fD{} model errors that occur due to the sparse end."
+#| fig-cap: "Exploring the effect of density on the NLDR layout using a \\gD{} curved sheet in \\fD{} with different density at each end.
+Three plots are linked: density plot of residuals (a1, b1), NLDR layout (a2, b2), projection of \\fD{} model and data (a3, b3).
+The brown points indicate the selected set, which are different in each row.
+In (a2), the top part of the triangular shape is selected which corresponds to higher residuals (a1) and the sparse end of the structure (a3).
+In (b2) one of other corners is highlighted, which can be seen to correspond to low residuals (b1) and one side of the dense end of the data (b3).
+While the tSNE layout represents the dense end of the sheet correctly as two corners in the layout, it contracts the sparse end of the sheet into a single corner.
+"
 

### 2025-06-26 paper.qmd
axes of fig 10 - PH

@@
       alpha=0.7, colour = "#800026", size = 3) +
-  xlab(expression(e[hj])) +
+  xlab(expression(paste("residual (",e[hi],")"))) +
   ylab("") +
@@
         axis.title = element_text(size = 18),
-        axis.text = element_text(size = 15))
+        axis.text.x = element_text(size = 15),
+        axis.text.y = element_blank())
 
@@
         axis.title = element_text(size = 18),
-        axis.text = element_text(size = 15))
+        axis.text = element_text(size = 15),
+        axis.text.y = element_blank())
 ```

### 2025-06-26 appendix.qmd
PBMC example revised - PH

@@
 
-Note that @xia2023 uses a different subset from the PBMC dataset used by @chen2024, shown by us in the main paper example, which is why this comparison is shown here and not in the main paper.
+Note that @xia2023 uses a different PBMC dataset than that used by @chen2024, shown by us in the main paper example, which is why this comparison is shown here and not in the main paper.
 Their data contains $31,021$ cells including cell type labels, and the gene expression levels were in the unit of log-transformed UMI count per $10,000$.

### 2025-06-26 appendix.tex
Added True model projections to appendix
PBMC example revised - PH

@@
 \newcommand\gD{$2\text{-}D$}
+\newcommand\tD{$3\text{-}D$}
+\newcommand\fD{$4\text{-}D$}
+\newcommand\sD{$7\text{-}D$}
 \newcommand{\To}{\textbf{to}~}
@@
 
-\section{Generating the nonlinear clusters
-data}\label{generating-the-nonlinear-clusters-data}
+\section{Generating the 2NC7 data}\label{generating-the-2nc7-data}
+
+This data is constructed by simulating two clusters, each consisting of
+\(1000\) observations. The C-shaped cluster is generated from
+\(\theta \sim U(\text{-}3\pi/2, 0)\), \(X_1 = \sin(\theta)\),
+\(X_2 \sim U(0, 2)\) (adding thickness to the C),
+\(X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)\),
+\(X_4 = \cos(\theta)\). Observations lie on a \gD{} manifold in \sD{}.
+The other cluster is from \(X_1 \sim U(0, 2)\), \(X_2 \sim U(0, 3)\),
+\(X_3 = \text{-}(X_1^3 + X_2)\), and \(X_4 \sim U(0, 2)\). It is also
+curved, but observations lie on a \tD{} manifold in\sD{}. Three more
+variables, \(X_5, X_6, X_7\), that are small amounts of pure noise are
+added. We would consider \(T=(X_1, X_2, X_3, X_4)\) to be the geometric
+structure (true model) that we hope to capture.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{appendix_files/figure-pdf/unnamed-chunk-4-1.pdf}
+
+}
+
+\caption{Two projections of the \pD{} true model overlaying the data are
+shown in a, b.}
+
+\end{figure}%
 
@@
 
-Note that \citet{xia2023} uses a different subset from the PBMC dataset
-used by \citet{chen2024}, shown by us in the main paper example, which
-is why this comparison is shown here and not in the main paper. Their
-data contains \(31,021\) cells including cell type labels, and the gene
+Note that \citet{xia2023} uses a different PBMC dataset than that used
+by \citet{chen2024}, shown by us in the main paper example, which is why
+this comparison is shown here and not in the main paper. Their data
+contains \(31,021\) cells including cell type labels, and the gene
 expression levels were in the unit of log-transformed UMI count per

### 2025-06-26 paper.qmd
PBMC example revised - PH

@@
 In this data there are $2622$ single cells and $1000$ gene expressions (variables).
-Following their pre-processing, different NLDR techniques were performed on the first nine principal components.
+Following the same pre-processing as @chen2024, different NLDR techniques were performed on the first nine principal components.
 @fig-NLDR-variety shows this data using a variety of methods, and different hyper-parameters.
 You can see that the result is wildly different depending on the choices.
-Layout a is a reproduction of the layout published in @chen2024.
+Layout a is a reproduction of the layout that was published in @chen2024.
 This layout suggests that the data has three very well separated clusters, each with an odd shape.
@@
 The middle and right plots show two projections.
-The primary difference between the two models is that, the model of layout e does not fill out to the extent of the data but concentrates in the center of each point cloud.
+The primary difference between the two models is that the model of layout e does not fill out to the extent of the data but concentrates in the center of each point cloud.
 Both suggest that three clusters is a reasonable interpretation of the structure, but layout e more accurately reflects the separation between them, which is small.
@@
 The choice of best is between layouts b and e, that have small separations between oddly shaped clusters.
-Layout e is the best choice."
+Layout e is chosen as the best."
 #| label: fig-pbmc-rmse

### 2025-06-26 paper.qmd
MNIST example - PH

@@
 The first $10$ principal components, explaining $83\%$ of the total variation, are used.
-This data is essentially lies on a nonlinear manifold in the high dimensions, defined by the shapes that "1"s make in when sketched, meaning that some pixels are almost always dark and some almost always light.
+This data essentially lies on a nonlinear manifold in the high dimensions, defined by the shapes that "1"s make when sketched.
 We expect that the best layout captures this type of structure and does not exhibit distinct clusters.

### 2025-06-26 paper.qmd
discussion - PH

@@
 Two examples illustrating usage are provided: the PBMC3k data where the NLDR is summarizing clustering in \pD{} and hand-written digits illustrating how NLDR represents an intrinsically lower dimensional nonlinear manifold.
-We demonstrated that the published layout of the PBMC3k is inaccurate, because it grossly exaggerates separation between clusters, and even suggests separation when there is none.
-This is common when layouts are chosen subjectively -- often a preference for the "prettiest".
-Our approach provides a way to objectively choose the layout and hopefully avoids the use of misleading layouts in the future.
+We examined a typical published usage of UMAP with the PBMC3k dataset (Chen, 2024).
+As is typical of UMAP layout with default settings, the separation between clusters is grossly exaggerated.
+The layout even suggests separation where there is none.
+<!--This is common when layouts are chosen subjectively -- often a preference for the "prettiest".--> Our approach provides a way to objectively choose the layout and hopefully avoids the use of misleading layouts in the future.
 In the hand-written digits we illustrate how our model fit statistics show that a flat disc layout is superior to the curved shaped layouts, and how to identify oddly written "1"s using the residuals of the fitted model.
@@
 Binning into cubes could be done in $3\text{-}D$ or higher, relatively easily, and used as a the basis for a wireframe of the fitted model.
-@barber1996 and the software @stephane2023 have algorithms for convex hulls, which \pD{} which serve as an inspiration.
+@barber1996 (and the associated software @stephane2023) has an algorithm for a convex hull in \pD{}m which serves as an inspiration.
 A simpler approach using $k$-means clustering to provide centroids could also be possible, but the complication would be to determine how to connect the centroids into an appropriate wireframe.

### 2025-06-26 motivationletter/motivationletter.qmd
motivation letter revised

@@
 
-Please consider our manuscript entitled *“Looking at Non-Linear Dimension Reductions as Models in the Data Space”* for publication in the *Journal of Computational and Graphical Statistics*.
+Please consider our manuscript entitled *“Stop Lying to Me: New Visual Tools to Choose the Most Honest Nonlinear Dimension Reduction”* for publication in the *Journal of Computational and Graphical Statistics*.
 
@@
 
-Our manuscript introduces an approach that treats NLDR outputs as models and visualizes them directly in the original high-dimensional space using tours (movies of linear projections).
-This allows analysts to assess how well the NLDR model fits across all subspaces, identify mismatches, and compare the summaries provided by different methods.
-The approach we describe offer a new, visually interpretable way to diagnose NLDR results and support more trustworthy use of these techniques in high-dimensional data analysis.
+Our manuscript introduces an approach that treats NLDR outputs as models and visualizes the models with the data in the original high-dimensional space using tours (movies of linear projections).
+Numerical diagnostics are also generated from the fitted model.
+These allow analysts to have support to help them choose an appropriate NLDR representation of their data.
 
 We believe that this work is a strong fit for *JCGS*, given the journal’s focus on innovative statistical graphics, computational methodology, and tools that help analysts better understand high-dimensional data.
-Our approach combines statistical modeling, dynamic visualization, and interactive diagnostics, contributing to the journal’s mission of advancing the use of computation and graphics in statistical practice.
+Our approach combines statistical modeling, interactive and dynamic graphics, and numerical diagnostics, contributing to the journal’s mission of advancing the use of computation and graphics in statistical practice.
 

### 2025-06-26 appendix.qmd
Changed Apendix names

@@
 ---
-title: "Appendix: Looking at Non-Linear Dimension Reductions as Models in the Data Space"
+title: "Appendix: Stop Lying to Me: New Visual Tools to Choose the Most Honest Nonlinear Dimension Reduction"
 format: 

### 2025-06-26 appendix.tex
Changed Apendix names

@@
 \hypersetup{
-  pdftitle={Appendix: Looking at Non-Linear Dimension Reductions as Models in the Data Space},
+  pdftitle={Appendix: Stop Lying to Me: New Visual Tools to Choose the Most Honest Nonlinear Dimension Reduction},
   pdfauthor={Jayani P. Gamage; Dianne Cook; Paul Harrison; Michael Lydeamore; Thiyanga S. Talagala},
@@
 
-\title{\bf Appendix: Looking at Non-Linear Dimension Reductions as
-Models in the Data Space}
+\title{\bf Appendix: Stop Lying to Me: New Visual Tools to Choose the
+Most Honest Nonlinear Dimension Reduction}
 \author{

### 2025-06-26 paper.qmd
Updated non-linear to nonlinear

@@
 abstract: |
-  Non-linear dimension reduction (NLDR) techniques such as tSNE, and UMAP provide a low-dimensional representation of high-dimensional data (\pD{}) by applying a non-linear transformation.
+  Nonlinear dimension reduction (NLDR) techniques such as tSNE, and UMAP provide a low-dimensional representation of high-dimensional data (\pD{}) by applying a nonlinear transformation.
 NLDR often exaggerates random patterns.
@@
 ---
-<!-- Looking at Non-Linear Dimension Reductions as Models in the Data Space
+<!-- Looking at nonlinear Dimension Reductions as Models in the Data Space
 
@@
 
-Non-linear dimension reduction (NLDR) is popular for making a convenient low-dimensional (\kD{}) representation of high-dimensional (\pD{}) data ($k < p$).
+Nonlinear dimension reduction (NLDR) is popular for making a convenient low-dimensional (\kD{}) representation of high-dimensional (\pD{}) data ($k < p$).
 Recently developed methods include t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021].

### 2025-06-26 paper.qmd
Updated Figure 7 (c)

@@
   ylab("RMSE") +
-  xlim(c(0, 0.14)) +
-  ylim(c(0, 0.75)) +
+  #xlim(c(0, 0.14)) +
+  #ylim(c(0, 0.75)) +
   interior_annotation("c", position = c(0.07, 0.9), cex = 2) +

### 2025-06-26 scripts/two_nonlinear/05_gen_rm_lwd_mse.R

@@ -26,14 +26,12 @@ df_bin_two_curvy1 <- two_curvy_model1$model_highd
 error_rm_two_curvy1 <- data.frame(matrix(nrow = 0, ncol = 0))
 
 ## To initialize benchmark values to remove low density hexagons
-benchmark_rm_hex_vec <- seq(0, 0.14, by=0.01)
+benchmark_rm_hex_vec <- seq(0, 0.014, by=0.001)
 
 for (benchmark_rm_lwd in benchmark_rm_hex_vec) {
 
   df_bin_centroids_two_curvy_high_dens <- df_bin_centroids_two_curvy1 |>
-    arrange(std_counts) |>
-    mutate(cum_std_counts = cumsum(std_counts)) |>
-    filter(cum_std_counts >= benchmark_rm_lwd)
+    dplyr::filter(std_counts >= benchmark_rm_lwd)
 
   df_bin_two_curvy_high_dens <- df_bin_two_curvy1 |>
     filter(hexID %in% df_bin_centroids_two_curvy_high_dens$hexID)
@@ -43,7 +41,7 @@ for (benchmark_rm_lwd in benchmark_rm_hex_vec) {
     model_2d = df_bin_centroids_two_curvy_high_dens,
     model_highd = df_bin_two_curvy_high_dens,
     highd_data = data_two_curvy) |>
-    mutate(benchmark_rm_lwd = round(benchmark_rm_lwd, 2),
+    mutate(benchmark_rm_lwd = round(benchmark_rm_lwd, 3),
            a1 = paste0("a[1] == ", round(two_curvy_model1$hb_obj$a1, 2)),
            bin1 = two_curvy_model1$hb_obj$bins[1],
            bin2 = two_curvy_model1$hb_obj$bins[2],
@@ -70,14 +68,12 @@ df_bin_two_curvy2 <- two_curvy_model2$model_highd
 error_rm_two_curvy2 <- data.frame(matrix(nrow = 0, ncol = 0))
 
 ## To initialize benchmark values to remove low density hexagons
-benchmark_rm_hex_vec <- seq(0, 0.14, by=0.01)
+benchmark_rm_hex_vec <- seq(0, 0.008, by=0.001)
 
 for (benchmark_rm_lwd in benchmark_rm_hex_vec) {
 
   df_bin_centroids_two_curvy_high_dens <- df_bin_centroids_two_curvy2 |>
-    arrange(std_counts) |>
[…]

### 2025-06-26 paper.qmd
Figure 7 y rage fix with 7 (a)

--- 2025-06-26 paper.qmd
@@
   ylab("RMSE") +
-  xlim(c(0, 0.14)) +
+  #xlim(c(0, 0.14)) +
   ylim(c(0, 0.75)) +

### 2025-06-26 paper.qmd
Updated 7 (c, d)

@@
       group = hex_poly_id, 
-      fill = std_counts), colour = "grey70", linewidth=0.2) +
+      fill = w_h), colour = "grey70", linewidth=0.2) +
   geom_point(data = tsne_two_curvy_scaled,
@@
         group = hex_poly_id, 
-        fill = std_counts), 
+        fill = w_h), 
         colour = "grey70", 
@@
         group = hex_poly_id, 
-        fill = std_counts), 
+        fill = w_h), 
         colour = "grey70", linewidth=0.2) +
@@
 error_rm_two_curvy <- read_rds("data/two_nonlinear/error_rm_lwd_diff_bin.rds") |>
-  mutate(bin1  = as.factor(bin1))
+  mutate(b1  = as.factor(b1))
 
@@
   mutate(a1 = round(a1, 2)) |>
-  filter(bin1 >= 5) |>
+  filter(b1 >= 5) |>
   group_by(method, a1) |>

### 2025-06-26 paper.qmd
Labels done with Figure 7

@@
   scale_x_continuous(breaks = sort(unique(round(error_two_curvy_tsne$a1, 2)))[c(1, 4, 7, 11, 13, 14, 15, 16)]) +
-  labs(x = expression(paste("binwidth (", a[1], ")")), y = expression(paste("relative ave bin density (", bar(d), ")"))) +
-  interior_annotation("d", position = c(0.9, 0.9), cex = 2) +
+  labs(x = expression(paste("binwidth (", a[1], ")")), y = paste("relative ave bin density")) +
+  interior_annotation("d", position = c(0.9, 0.1), cex = 2) +
   theme_minimal() +
@@
 Removing low count bins changes RMSE substantially (c).
-The average bin density ($\\bar{d}$) tends to decrease with $a_1$ (d)."
+The relative average bin density tends to increase with $a_1$ (d)."
 #| label: fig-param-two-curvy

### 2025-06-26 paper.qmd
Changed PBMC3k text
Updated Figure 15

@@
 @fig-pbmc-rmse shows RMSE across a range of binwidths ($a_1$) for each of the layouts in @fig-NLDR-variety.
-The layouts were made using tSNE, UMAP, PHATE, PaCMAP, and TriMAP with various hyper-parameter settings.
+The layouts were generated using tSNE and UMAP with various hyper-parameter settings, while PHATE, PaCMAP, and TriMAP were applied using their default settings.
 Lines are color coded to match the color of the layouts shown on the right.
@@
 
-<!--mnist/04_gen_mse_for_diff_methods.R-->
+<!--mnist/03_gen_mse_for_diff_methods.R-->
 ```{r}
@@
 <!-- Fit the model and compute error-->
-<!--mnist/05_gen_model_with_tSNE.R-->
+<!--mnist/04_gen_model_with_tSNE.R-->
 ```{r}
@@
 
-<!--mnist/03_gen_images.R-->
 ```{r}
@@
 right_top_img <- plot_digit_img(
-  img_right_top, palette = "Blues 2", title_text = "b")
+  img_right_top, color_on = "#375cb1", title_text = "b")
 
@@
 middle_img <- plot_digit_img(
-  img_middle, palette = "Greens 3", title_text = "c")
+  img_middle, color_on = "#008000", title_text = "c")
 
@@
 right_bottom_img <- plot_digit_img(
-  img_right_bottom, palette = "Purples 2", title_text = "d")
+  img_right_bottom, color_on = "#4c389a", title_text = "d")
 
@@
 imge_error_sample_outside <- plot_digit_img(
-  img_error_outside, palette = "Grays", title_text = "e")
+  img_error_outside, color_on = "#000000", title_text = "e")
 
@@
 imge_error_sample_inside <- plot_digit_img(
-  img_error_inside, palette = "Oranges", title_text = "f")
+  img_error_inside, color_on = "#ff7f00", title_text = "f")
 

### 2025-06-26 scripts/additional_functions.R

@@ -37,7 +37,7 @@ scale_data_manual <- function(data, type_col) {
   }
 
   # Combine the scaled data with the 'type' column and return as a tibble
-  data_scaled <- as_tibble(data_scaled) %>%
+  data_scaled <- as_tibble(data_scaled) |>
     mutate(!!type_col := data[[type_col]])
 
   return(data_scaled)
@@ -77,26 +77,49 @@ standardize = function(x){
 
 # Plot digits in MNIST
 
-plot_digit_img <- function(digit_df, palette, title_text, ncol = 5) {
+plot_digit_img <- function(
+    digit_df,
+    title_text,
+    ncol = 5,
+    threshold = 100, # Added as a parameter
+    color_off = "white", # Added as a parameter
+    color_on = "darkblue" # Added as a parameter
+) {
 
-  # Add a label column only to the first instance
-  digit_df <- digit_df |>
-    mutate(label = ifelse(instance == unique(instance)[1], title_text, NA))
+  # --- Data Processing: Binarize and prepare for plotting ---
+  digit_df_processed <- digit_df |>
+    # Add a label column only to the first instance if title_text is provided
+    # (The original code had `mutate(label = ifelse(instance == unique(instance)[1], title_text, NA))`)
+    # If digit_df already has a 'label' column, you might want to adjust this.
+    # For now, I'll keep your original label logic.
+    mutate(label = ifelse(instance == unique(instance)[1], title_text, NA_character_)) |>
 
-  ggplot(data = digit_df, aes(x, y, fill = value)) +
+    # Binarize the 'value' column based on the threshold
+    mutate(
+      binary_value = ifelse(value <= threshold, 0, 1),
+      # Convert to factor for discrete manual scaling and controlled order
+      binary_value = factor(binary_value, levels = c(0, 1))
+    )
+
+  ggplot(data = digit_df_processed, aes(x = x, y = y, fill = binary_value)) +
     geom_tile() +
-    geom_text(data = digit_df |> filter(!is.na(label)),
+    geom_text(data = digit_df_processed |> filter(!is.na(label)),
               aes(label = label),
-              x = min(digit_df$x), y = max(digit_df$y), inherit.aes = FALSE,
+              x = min(digit_df_processed$x, na.rm = TRUE), # Use na.rm=TRUE for robustness
+              y = max(digit_df_processed$y, na.rm = TRUE), # Use na.rm=TRUE for robustness
+              inherit.aes = FALSE,
[…]

### 2025-06-26 appendix.qmd
Updated appendix with new notations

@@
   geom_point(data = emb_df1, aes(x = emb1, y = emb2), color = "black", size = 0.7) +
-  geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID), size = 5, colour = "#bdbdbd") +
+  geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = h), size = 5, colour = "#bdbdbd") +
   geom_circle(data = circles, aes(x0 = x, y0 = y, r = r), fill = NA) +
@@
   geom_point(data = points_df, aes(x = x, y = y), color = "black", size = 0.7) +
-  geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID), size = 5, colour = "#bdbdbd") +
+  geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = h), size = 5, colour = "#bdbdbd") +
   geom_segment(data = arrow_df, aes(x = x, y = y, xend = xend, yend = yend), 

### 2025-06-26 appendix.qmd
Updated appendix Figure 1

@@
   point_param = c(0.5, 0.2, clr_choice), # size, alpha, color
-  line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
+  line_param = c(0.5, 0.3, "#000000"), # linewidth, alpha
   plot_limits = c(-0.75, 0.75),
-  axis_text_size = 2,
+  axis_text_size = 4,
   is_category = FALSE) +
-  interior_annotation(label = "a", cex = 1.2) +
+  interior_annotation(label = "a", cex = 2) +
   theme(aspect.ratio = 1,
@@
   point_param = c(0.1, 0.5, clr_choice), # size, alpha, color
-  line_param = c(0.5, 0.5, "black"), # linewidth, alpha
+  line_param = c(0.5, 0.3, "black"), # linewidth, alpha
   plot_limits = c(-0.5, 0.5),
-  axis_text_size = 2,
+  axis_text_size = 4,
   is_category = FALSE) +
-  interior_annotation(label = "b", cex = 1.2) +
+  interior_annotation(label = "b", cex = 2) +
   theme(aspect.ratio = 1)
@@
 ```{r}
-#| fig-width: 6
-#| fig-height: 3
+#| fig-width: 10
+#| fig-height: 5
 #| fig-cap: "Two projections of the \\pD{} true model overlaying the data are shown in a, b."
 
-two_curvy_proj_true1 + two_curvy_proj_true2
+two_curvy_proj_true1 + two_curvy_proj_true2 +
+  plot_layout(ncol = 2)
 ```

### 2025-06-26 appendix.qmd
Figure 1 link video

@@
 #| fig-height: 5
-#| fig-cap: "Two projections of the \\pD{} true model overlaying the data are shown in a, b."
+#| fig-cap: "Two projections of the \\pD{} true model overlaying the data are shown in a, b. Video of the langevitour animations is available at <https://youtu.be/35TrnYJsUUI>."
 

### 2025-06-26 appendix.qmd
Updated appendix Figure 5 title

@@
 #| out-width: 100%
-#| fig-cap: "Assessing which of the two layouts with UMAP and tSNE with different hyper-parameter setting (n_neighbors: $30$, min_dist: $0.3$ (red); n_neighbors: $30$ (blue)) on the PBMC data is the better representation using RMSE for varying binwidth ($a_1$).
-Colour used for the lines and points in the top plot and in the scatterplots represents UMAP and tSNE layouts (a, b).
+#| fig-cap: "Assessing which of the 2 NLDR layouts on the PBMC3k data is the better representation using RMSE for varying binwidth ($a_1$).
+Color used for the lines and points in the left plot and in the scatterplots represents NLDR layouts (a, b).
 Of the two, layout b is optimal across all binwidths making it the best choice."
@@
 #| echo: false
-#| fig-cap: "Compare the published \\gD{} layout (@fig-pbmc-mse-umap a) and the \\gD{} layout selected (@fig-pbmc-mse-umap b) by RMSE plot (@fig-pbmc-mse-umap) from tSNE and UMAP with default hyper-parameters.
-The PBMC data ($n =  5858$) has three clusters in $9\\text{-}D$, where three clusters are close.
-Two \\gD{} projection from a tour on $9\\text{-}D$ of the model fit with @fig-pbmc-mse-umap a ($a_1 = 0.04$, $b = 900/179 (30, 30)$) shows three-well separated clusters with big separations.
-On the other hand, the model fit with @fig-pbmc-mse-umap b ($a_1 = 0.04$, $b = 1110/317 (30, 37)$) shows three-well separated clusters with small separations.
-Therefore, @fig-pbmc-mse-umap b is more reasonable than @fig-pbmc-mse-umap a.
-Videos of the langevitour animations are available at <https://youtu.be/0cKX_HG_n0k> and <https://youtu.be/KhJvsRtaX04> respectively."
+#| fig-cap: "Compare the published \\gD{} layout (@fig-pbmc-mse-umap a) made with UMAP and the \\gD{} layout selected (@fig-pbmc-mse-umap b) made with tSNE by RMSE plot (@fig-pbmc-mse-umap).
+The two plots on the right show  projections from a tour, with the models overlaid.
+The published layout a suggested three separated clusters with two of them are close, but this is not present in the data.
+While there may be three clusters they are not well-separated.
+The difference in model fit also indicates this: the published layout a does not capture the nonlinear structure of the clusters like the model generated from layout b.
+This supports the choice that layout b is the better representation of the data, because it shows close clusters.
+Videos of the langevitour animations are available at <https://youtu.be/ffiB4MGWyn8> and <https://youtu.be/e7XNL18co1c> respectively."
 #| label: fig-model-pbmc-author-proj

### 2025-06-26 appendix.tex
Figure 1 link video
Updated appendix Figure 5 title

@@
 \caption{Two projections of the \pD{} true model overlaying the data are
-shown in a, b.}
+shown in a, b. Video of the langevitour animations is available at
+\url{https://youtu.be/35TrnYJsUUI}.}
 
@@
 
-\caption{\label{fig-pbmc-mse-umap}Assessing which of the two layouts
-with UMAP and tSNE with different hyper-parameter setting (n\_neighbors:
-\(30\), min\_dist: \(0.3\) (red); n\_neighbors: \(30\) (blue)) on the
-PBMC data is the better representation using RMSE for varying binwidth
-(\(a_1\)). Colour used for the lines and points in the top plot and in
-the scatterplots represents UMAP and tSNE layouts (a, b). Of the two,
+\caption{\label{fig-pbmc-mse-umap}Assessing which of the 2 NLDR layouts
+on the PBMC3k data is the better representation using RMSE for varying
+binwidth (\(a_1\)). Color used for the lines and points in the left plot
+and in the scatterplots represents NLDR layouts (a, b). Of the two,
 layout b is optimal across all binwidths making it the best choice.}
@@
 \caption{\label{fig-model-pbmc-author-proj}Compare the published \gD{}
-layout (Figure~\ref{fig-pbmc-mse-umap} a) and the \gD{} layout selected
-(Figure~\ref{fig-pbmc-mse-umap} b) by RMSE plot
-(Figure~\ref{fig-pbmc-mse-umap}) from tSNE and UMAP with default
-hyper-parameters. The PBMC data (\(n =  5858\)) has three clusters in
-\(9\text{-}D\), where three clusters are close. Two \gD{} projection
-from a tour on \(9\text{-}D\) of the model fit with
-Figure~\ref{fig-pbmc-mse-umap} a (\(a_1 = 0.04\),
-\(b = 900/179 (30, 30)\)) shows three-well separated clusters with big
-separations. On the other hand, the model fit with
-Figure~\ref{fig-pbmc-mse-umap} b (\(a_1 = 0.04\),
-\(b = 1110/317 (30, 37)\)) shows three-well separated clusters with
-small separations. Therefore, Figure~\ref{fig-pbmc-mse-umap} b is more
-reasonable than Figure~\ref{fig-pbmc-mse-umap} a. Videos of the
-langevitour animations are available at
-\url{https://youtu.be/0cKX_HG_n0k} and
-\url{https://youtu.be/KhJvsRtaX04} respectively.}
+layout (Figure~\ref{fig-pbmc-mse-umap} a) made with UMAP and the \gD{}
+layout selected (Figure~\ref{fig-pbmc-mse-umap} b) made with tSNE by
+RMSE plot (Figure~\ref{fig-pbmc-mse-umap}). The two plots on the right
+show projections from a tour, with the models overlaid. The published
+layout a suggested three separated clusters with two of them are close,
+but this is not present in the data. While there may be three clusters
+they are not well-separated. The difference in model fit also indicates
+this: the published layout a does not capture the nonlinear structure of
+the clusters like the model generated from layout b. This supports the
+choice that layout b is the better representation of the data, because
+it shows close clusters. Videos of the langevitour animations are
+available at \url{https://youtu.be/ffiB4MGWyn8} and
+\url{https://youtu.be/e7XNL18co1c} respectively.}
 

### 2025-06-26 paper.qmd
Updated Figure 15
Updated appendix Figure 5 title

@@
 The two plots on the right show  projections from a tour, with the models overlaid.
-The published layout suggested three very separated clusters, but this is not present in the data.
+The published layout a suggested three very separated clusters, but this is not present in the data.
 While there may be three clusters they are not well-separated.
@@
 imge_error_sample_inside <- plot_digit_img(
-  img_error_inside, color_on = "#ff7f00", title_text = "f")
+  img_error_inside, color_on = "#d95f0e", title_text = "f")
 

### 2025-06-26 appendix.qmd
Added parameter choices

@@
 
-For illustration purposes, we only selected cells generated with inDrops ($n=5858$ cells) and UMAP and tSNE cell embeddings.
-Also, @xia2023 used first $9$ principal components to generate the UMAP and tSNE.
+For illustration purposes, we only selected cells generated with inDrops ($n=5858$ cells).
+Also, @xia2023 used first $9$ principal components to generate the UMAP and tSNE with default hyper-parameters.
 The objective is to assess the optimized layout by scDEED, and if it does not accurately represent the three clusters with small separations of the PBMC dataset, then select a reasonable \gD{} layout.

### 2025-06-26 appendix.tex
Added parameter choices

@@
 For illustration purposes, we only selected cells generated with inDrops
-(\(n=5858\) cells) and UMAP and tSNE cell embeddings. Also,
-\citet{xia2023} used first \(9\) principal components to generate the
-UMAP and tSNE. The objective is to assess the optimized layout by
-scDEED, and if it does not accurately represent the three clusters with
-small separations of the PBMC dataset, then select a reasonable \gD{}
-layout.
+(\(n=5858\) cells). Also, \citet{xia2023} used first \(9\) principal
+components to generate the UMAP and tSNE with default hyper-parameters.
+The objective is to assess the optimized layout by scDEED, and if it
+does not accurately represent the three clusters with small separations
+of the PBMC dataset, then select a reasonable \gD{} layout.
 

### 2025-06-26 motivationletter/motivationletter.qmd
Updated the word non-linear

@@
 
-Non-linear dimension reduction (NLDR) techniques, such as tSNE and UMAP, are widely used for producing low-dimensional representations of high-dimensional data.
+Nonlinear dimension reduction (NLDR) techniques, such as tSNE and UMAP, are widely used for producing low-dimensional representations of high-dimensional data.
 While these methods can provide concise and powerful summaries, they also have the potential to exaggerate random patterns or create misleading structures.

### 2025-06-26 paper.qmd
fixed use of residual vs error

@@
 Two projections of the \\pD{} fitted model overlaying the data are shown in b, c.
-The fit is reasonably tight with the data in one cluster (top one in b), but slightly less so in the other cluster probably because it is in $3\\text{-}D$.
+The fit is reasonably tight with the data in one cluster (top one in b), but slightly less so in the other cluster probably because it is $3\\text{-}D$.
 Notice also that, in the \\gD{} layout the two clusters have internal gaps which creates a model with some holes.
@@
 #| fig-cap: "Examining the distribution of residuals in a jittered dotplot (a), \\gD{} NLDR layout (b) and a tour of \\fD{} data space (c).
-Color indicates error ($e_{hi}$), dark color indicating high error and light indicates low error.
-Most large errors are distributed in one cluster (bottom one in c) and most small errors are distributed in the other cluster."
+Color indicates residual ($e_{hi}$), dark color indicating high value.
+Most large residuals are distributed in one cluster (bottom one in c) and most small residuals are distributed in the other cluster."
 #| out-width: 100%
@@
 #| label: fig-model-error-mnist
-#| fig-cap: "The $10\\text{-}D$ model error in \\gD{} layout of the MNIST digit 1 data shows a pattern.
-Most low model errors are distributed along the big nonlinear cluster, while most large model errors are distributed along the small cluster.
-The images associated with large model errors shows different patterns of digit 1, some inside (f) the nonlinear structure and others outside (e).
-Along the nonlinear cluster, the angle of digit 1 changes (b-d)."
+#| fig-cap: "There is a pattern to error of the fit of the model for \\gD{} layout of the MNIST digit 1 data.
+On the left is the NLDR layout (a), and at right (b-e) are images of samples of observations taken at locations along the big cluster, and the small cluster, showing how the 1's were drawn.
+Set (f) are images corresponding to large residuals in the big cluster.
+Along the big cluster, the angle of digit 1 changes (b-d).
+The small cluster has larger residuals, and the images show that these tend to be European style with a flag at the top, and a base at the bottom.
+The set in (f) show various poorly written digits.
+"
 #| fig-pos: H

### 2025-06-30 appendix.qmd
removed html version and made changes pointed out by PH

@@
         keep-tex: true
-    jasa-html: default
 author:

### 2025-06-30 paper.qmd
removed html version and made changes pointed out by PH

@@
 ---
+
 <!-- Looking at nonlinear Dimension Reductions as Models in the Data Space
@@
 Nonlinear dimension reduction (NLDR) is popular for making a convenient low-dimensional (\kD{}) representation of high-dimensional (\pD{}) data ($k < p$).
-Recently developed methods include t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021].
+Recently developed methods include t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021].
 However, the representation generated can vary dramatically from method to method, and with different choices of parameters or random seeds made using the same method (@fig-NLDR-variety).
@@
 Here we focus on five currently popular techniques: tSNE, UMAP, PHATE, TriMAP and PaCMAP.
-Both tNSE and UMAP can be considered to produce the \kD{} representation by minimizing the divergence between two distributions, where the distributions are modeling the inter-point distances.
-PHATE are examples of diffusion processes spreading to capture geometric shapes, that include both global and local structure.
-(See @coifman2005 for an explanation of diffusion processes.) TriMAP and PaCMAP
+The methods tNSE, UMAP, TriMAP and PaCMAP can be considered to produce the \kD{} representation by minimizing the divergence between two inter-point distance distributions.
+PHATE is an example of a diffusion process spreading to capture geometric shapes, that include both global and local structure.
+(See @coifman2005 for an explanation of diffusion processes.)
 
@@
 Binning into cubes could be done in $3\text{-}D$ or higher, relatively easily, and used as a the basis for a wireframe of the fitted model.
-@barber1996 (and the associated software @stephane2023) has an algorithm for a convex hull in \pD{}m which serves as an inspiration.
+@barber1996 (and the associated software @stephane2023) has an algorithm for a convex hull in \pD{} which serves as an inspiration.
 A simpler approach using $k$-means clustering to provide centroids could also be possible, but the complication would be to determine how to connect the centroids into an appropriate wireframe.

### 2025-08-09 paper.qmd
Updated a1, a2 equality signs
Fixed Gammars and typos suggested by reviewers

@@
 Here we focus on five currently popular techniques: tSNE, UMAP, PHATE, TriMAP and PaCMAP.
-The methods tNSE, UMAP, TriMAP and PaCMAP can be considered to produce the \kD{} representation by minimizing the divergence between two inter-point distance distributions.
+The methods tNSE, UMAP, TriMAP and PaCMAP can be considered for producing the \kD{} representation by minimizing the divergence between two inter-point distance distributions.
 PHATE is an example of a diffusion process spreading to capture geometric shapes, that include both global and local structure.
@@
   geom_polygon(colour = "grey70", 
-               aes(group = hex_poly_id), 
+               aes(group = h), 
                fill = "#ffffff") +
@@
     aes(x = x, y = y,
-      group = hex_poly_id, 
+      group = h, 
       fill = w_h), colour = "grey70", linewidth=0.2) +
@@
 To illustrate the method, we use \sD{} simulated data, which we call the "2NC7" data.
-It is has two separated nonlinear clusters, one forming a \gD{} curved shape, and the other a \tD{} curved shape, each consisting of $1000$ observations.
+It has two separated nonlinear clusters, one forming a \gD{} curved shape, and the other a \tD{} curved shape, each consisting of $1000$ observations.
 The first four variables hold this cluster structure, and the remaining three are purely noise.
@@
 $$
-a_2 > \frac{r_2 + q(1 + r_2)}{(b_2 - 1)}.
+a_2 \geq \frac{r_2 + q(1 + r_2)}{(b_2 - 1)}.
 $$
@@
 $$
-a_1 > \frac{2[r_2 + q(1 + r_2)]}{\sqrt{3}(b_2 - 1)}.
+a_1 \geq \frac{2[r_2 + q(1 + r_2)]}{\sqrt{3}(b_2 - 1)}.
 $$
@@
                             y = y, 
-                            group = hex_poly_id),
+                            group = h),
                  fill = "white", 
@@
     aes(x = x, y = y, 
-        group = hex_poly_id, 
+        group = h, 
         fill = w_h), 
@@
 Notice also that, in the \\gD{} layout the two clusters have internal gaps which creates a model with some holes.
-This lacy pattern happens regardless of the hyper-parameter choice, but this doen't severely impact the \\pD{} model representation."
+This lacy pattern happens regardless of the hyper-parameter choice, but this doesn't severely impact the \\pD{} model representation."
 #| fig-width: 8
@@
 
-Comparing the RMSE to obtain the best fit is appropraite if the same NLDR method is used.
+Comparing the RMSE to obtain the best fit is appropriate if the same NLDR method is used.
 However, because the RMSE is computed on \pD{} data it measures the fit between model and data so it can also be used to compare the fit of different NLDR methods.

### 2025-08-09 scripts/two_nonlinear/06_gen_model_with_tSNE.R

@@ -379,7 +379,7 @@ df_bin_two_curvy2 <- algo_obj_two_curvy2$model_highd
 hex_grid_two_curvy2 <- algo_obj_two_curvy2$hb_obj$hex_poly
 counts_df_two_curvy2 <- algo_obj_two_curvy2$hb_obj$std_cts
 
-hex_grid_with_counts_two_curvy2 <- left_join(hex_grid_two_curvy2, counts_df_two_curvy2, by = c("hex_poly_id" = "h"))
+hex_grid_with_counts_two_curvy2 <- left_join(hex_grid_two_curvy2, counts_df_two_curvy2, by = c("h" = "h"))
 
 write_rds(hex_grid_with_counts_two_curvy2, "data/two_nonlinear/two_nonlinear_hex_grid_with_counts_two_curvy2.rds")
 write_rds(tr_from_to_df_two_curvy2, "data/two_nonlinear/two_nonlinear_tr_from_to_df_two_curvy2.rds")
@@ -403,7 +403,7 @@ df_bin_two_curvy3 <- algo_obj_two_curvy3$model_highd
 hex_grid_two_curvy3 <- algo_obj_two_curvy3$hb_obj$hex_poly
 counts_df_two_curvy3 <- algo_obj_two_curvy3$hb_obj$std_cts
 
-hex_grid_with_counts_two_curvy3 <- left_join(hex_grid_two_curvy3, counts_df_two_curvy3, by = c("hex_poly_id" = "h"))
+hex_grid_with_counts_two_curvy3 <- left_join(hex_grid_two_curvy3, counts_df_two_curvy3, by = c("h" = "h"))
 
 write_rds(hex_grid_with_counts_two_curvy3, "data/two_nonlinear/two_nonlinear_hex_grid_with_counts_two_curvy3.rds")
 write_rds(tr_from_to_df_two_curvy3, "data/two_nonlinear/two_nonlinear_tr_from_to_df_two_curvy3.rds")

### 2025-08-19 JCGS_response_to_reviews/response_to_reviews.qmd
Start working on the reviews

---
title: "Response to Reviews: JCGS-25-420"
author: Jayani P. Gamage, Dianne Cook, Paul Harrison, Michael Lydeamore
format: pdf
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 2025-08-19 JCGS_response_to_reviews/response_to_reviews.qmd
Added Reviewer 1 comments

@@
 ```
+
+# Reviewer 1
+
+## Summary:
+
+This manuscript proposes an algorithm to evaluate nonlinear dimension reduction (NLDR) techniques.
+The authors identify some characteristics and limitations of NLDR and develop an algorithm to analyze NLDR results' quality of how well they match the original high-dimensional data.
+
+## Strengths:
+
+1. This manuscript is well-organized, providing sufficient background knowledge and examples which makes this paper easy to follow.
+
+2. This paper contains sufficient technical details in the introduction of the algorithm.
+
+## Weakness:
+
+1.
+The main concern is whether the proposed algorithm can properly evaluate the NLDR's performance.
+Using the RMSE plot with different binwidth values (e.g., in Figure 8 and Figure 11) to evaluate NLDR's performance is problematic for not considering the density within each cluster.
+For NLDR results that have denser clusters, for example layout a,c,g,h in Figure 11, a smaller binwidth needs be used to achieve the same average count in each bin, which will cause their lines in Figure 11 to be more upper left.
+However, denser clusters are not worse NLDR results.
+
+2.
+In Figure 7, as RMSE steadily increases with the binwidth, it's not clear why the three binwidths 0.03, 0.05, and 0.07 are chosen.
+In addition, we shouldn't use a single simulated dataset to determine the default hyperparameters without any real analysis.
+
+3.
+For several figures that contains results from multiple NLDR algorithms, it would be nice to add the information of which algorithm/ what hyperparameter values for each of the subfigure.
+Many readers may have tried several NLDR algorithms themselves, and they will benefit from those figures to gain insights about behaviors of different NLDRs.
+Also it's nice to mention which data is used to generate each figure (e.g., in Figure 7).
+
+4. In section 3.6.1, it says "Values of b1 between 2 and b1=sqrt(n/r2) are allowed", why the choice of b1 should consider r2?
+
+# Reviewer 2
+
+## Minor Issues
+
+1.
+First part of the title seems inappropriate for a scientific paper to me, as does the implied (and not substantiated, IMO) claim that the proposal is (primarily) useful to reliably select one of many NLDRs (as opposed to primarily useful for visually investigating details about a specific 2D NLDR).
+
+2. For marketing and clarity reasons, I would strongly suggest to not name the metric “RMSE”, but something more distinctive/descriptive.
+
+3.
+Flow from 3.1 to 3.2 (with too many sub-subsections) is difficult.
+Sections 3.2.1 (Scale the data) and 3.2.2 (Construct hexagon grid) are just short paragraphs, would be clearer if combined into a single numbered list of steps for the model construction.
+Much of 3 might be clearer and more compact if simply replaced by a direct step-by-step description of the algorithm in algorithmic pseudocode form with suitable short explanations.
+
+4.
+Section 3.5 is a claim without supporting evidence – many out-of-sample embedding methods exist (some based on rather similar ideas, which should be referenced), you would need to show that this particular one actually works reasonably well for this to deserve space in the paper.
+I would strongly suggest to simply mention this in the discussion section as a possible avenue for further development instead.
+
+5. Section 3.6 should be drastically shortened and details moved into an appendix.
+
+6.
+p.
+18: “A particular pattern that we commonly see is that analysts tend to pick lay- outs with clusters that have big separations between them.” The literature/internet contains many lively discussions of the perils of mis-interpreting and post-hoc’ing NL-DRs (e.g.
+(Izarry, 2024), (Chari & Pachter, 2023)), it might be good to cite/summarize these instead of/in addition to basing your motivation on personal anecdotal evidence?
+The tone here seems overly conversational to me (also: “we almost always see there are no big separations in the data”) and should be made more formal.
+
+7. Section 4, esp. p.19: IMO, “to compare and assess a range of representations” an analyst really should look at their respective RNX curves, stress values, and Shepard diagrams,
+think about whether they care more about local or global structure preservation, and then decide based on those factors.
+The method proposed here seems to me to be mostly useful to then investigate specific regions / slices of high-D space to see for which sets of data points the high-D structure
+is (not) preserved well in the NLDR.
+I find this whole section to be vastly overselling the proposal – it ignores and implicitly discards most of the prior work in this field.
+Its intro paragraphs are also highly redundant with other content in the paper and should be cut.
+
+8.
+Section 5 should be moved into an appendix to shorten the paper somewhat.
+Figure in Section 5.1.
+could benefit from coloring the clusters in different colors so global structure preservation is also somewhat legible from the figures for the NLDRs (e.g.
+“is orange cluster between green and blue as in high-D or do they switch positions”).
+Section 5.2 might also have benefited from a corresponding ISOMAP or similar embedding of this data suitable for truthfully (isometrically!) unrolling the high-D manifold, and then showcasing that such a successful “unrolling” can be discerned from the way the projections look?
+tSNE just isn’t isometric, so the distortions visible here are entirely expected, IMO?
+
+### Typos and grammar:
+
+Revised version should be much more carefully proofread: “appropraite”, “doen’t”, “it is has two separated nonlinear clusters”, “the methods tNSE”

### 2025-08-19 JCGS_response_to_reviews/response_to_reviews.qmd
Added Reviewer 2 comments

@@
 
+## Summary
+
+The manuscript proposes a novel, mostly graphical diagnostic to evaluate nonlinear dimension reduction (NLDR) layouts by effectively “lifting” the 2-D embedding back into the original high-dimensional space and inspecting the fit visually via tours (i.e., sequences of 2D-linear projections of the original high-D space).
+The goal is to help analysts discern which 2-D projection is the most “honest” (faithful) representation of the true structure in the data.
+The paper’s contributions are:
+
+1.
+an algorithm that overlays a geometric model on an NLDR layout using hexagonal binning and Delaunay triangulation in the embedding, then maps this model into the original large-D data space for visual assessment;
+
+2.
+a new goodness-of-fit metric, the RMSE in original space, computed as the root mean squared distance between each point and its bin centroid in high-D (Eq 2) (3) interactive visualization tools (implemented in published software) to compare/investigate different NLDRs
+
+It is mostly well written, adresses an important problem with critical real-world relevance, and provides a very well done, extensively documented R implementation.
+However, this reviewer is skeptical about the suitability of the newly proposed GOF metric to choose the most “reasonable” or “faithful” NLDR, finds the empirical evaluation of it to be severely lacking without comparison to alternative measures from the literature, and would request some restructuring and shortening of the paper.
+
+## Strengths
+
+### Relevance, Accessibility, and Intuitiveness
+
+The way this method visualizes the spatial distortions induced by NLDR seems uniquely acces sible to me, and, coupled with the modern interactive/animated visualizations implemented for it, seems likely to become a very useful tool to investigate and understand high-D data structures while avoiding over-reliance on / misinterpretation of misleading 2D patterns so prevalent in many areas at the moment (e.g.
+Izarry ( 2024), Chari & Pachter (2023)).
+
+### Background Material & Examples
+
+The paper is fully fleshed out with illuminating examples, and I commend the authors for the effort they put into documenting their package and providing very informative videos etc of the method in action.
+
+### Reproducibility and Code Availability
+
+The paper is exemplary in this regard - the underlying software is a CRAN compliant R package, and the Github repository itself allows to recompile the entire paper and appendix with all figures etc from scratch.
+
+## Major Issues
+
+### No Standard Embedding Quality Metrics or Comparisons
+
+My primary concern is the absence of established quantitative metrics
+
+1. to evaluate to what extent the newly proposed “RMSE” metric corresponds to established notion of embedding faithfulness and
+
+2. to validate the different embeddings under consideration.
+
+The paper’s only evaluation of “honesty” is done via the proposed custom “RMSE” in high dimensional space (defined in Section 3.4) and visual heuristics, but it does not report any conventional measures like trustworthiness/continuity scores, $k$-Neighborhood preservation rates (as derived from the co-ranking matrix - e.g.
+$R_{NX}$-curves for different neighborhood sizes $k$, c.f.
+(Lee et al, 2015), (Lueks et al, 2011)), or Shepard diagrams of original space vs.
+embedding space distances.
+This is a significant omission because NLDR almost always faces a trade-off between optimizing preservation of local structures vs preservation of global structures, which these metrics are (mostly) designed to resolve / analyze.
+For example, the authors could have computed such measures of trustworthiness for each NLDR layout to see if the one with lowest “RMSE” indeed had the highest (local and or global) neighborhood structure preservation but no such analysis was provided.
+The manuscript doesn’t even mention or discuss these metrics, which is not acceptable in a study focused on faithful representation, in my opinion.
+
+Questions a revised version should aim to resolve:
+
+1.
+does “RMSE” measure local or global structure preservation (or a mixture of both)?
+From the definition, it seems to be focused on how well small neighborhoods in embed ding space (the bins)) correspond to small neighborhoods in high-D space, with, IIUC, particularly high values if there are lots of “hard intrusions” (i.e., points that are embed ded in the same 2D bin despite being far apart in high-D space)?
+
+2. how strongly are “RMSE” values correlated to other established measures of (local or global) structure preservation? and (how) are these correlations affected by the binning hyperparameters?
+
+3.
+are the NLDRs selected by lowest “RMSE” the ones that would also be selected by other, more established measures of embedding faithfulness whose characteristics are well understood?
+at the very least, such measures should be provided, compared and discussed for all the embeddings under consideration in the case studies.
+
+The reliance of “RMSE” on euclidean distances in high-D space also seems limiting:
+
+1. at least for coarser bins and/or highly non-linear data structures, geodesic distances along the data manifold à la ISOMAP are likely to measure a more relevant kind of dissimilarity
+
+2.
+a definition based on a more general notion of distance would make this applicable to in vestigating 2D embeddings of data types for which the euclidean distance of their numeri cal data representations is not a suitable/applicable measure of dissimilarity (e.g.
+images, text).
+
+### Insufficient Validation of the Proposed Diagnostic (RMSE) Across Methods
+
+Related to the above points - the paper has no theoretical or empirical demonstration that minimizing this “RMSE” aligns with minimizing other distortion measures like stress or max imizing neighborhoood preservation (as measured by the co-ranking matrix).
+Furthermore, the claim that RMSE can be used to compare different NLDR methods on equal footing de serves scrutiny.
+Different methods optimize different criteria; it’s plausible that one method might (tends to) achieve lower “RMSE” simply by construction, especially if my reading that it mostly rewards embedings that manage to avoid hard intrusions is correct.
+Overall, much more critical validation of the RMSE diagnostic is needed to convince readers that “honesty” as defined here aligns with meaningful goodness-of-fit in embedding.
+
+### Limitations of the Evaluation (Simulation Study Design):
+
+The manuscript includes a simulation to illustrate the approach (the “2NC7” dataset with two nonlinear clusters in 7-D) and two real-data applications (single-cell RNA-seq and an image dataset of handwritten digits).
+While these examples are appropriate, the scope of simulation scenarios is narrow, and the analysis lacks any notion of variability.
+The simulation seems to be a single realization of a specific two-cluster scenario.
+The authors did not explore, for instance, varying the overlap between clusters, adding more clusters or different manifold shapes, or introducing different noise levels.
+Ideally, the authors would systematically vary factors like: number of clusters or pattern types, intrinsic dimensionality of structures, sample size, etc., and show that their approach reliably selects the truth (or at least, the most interpretable view) in each case, and how well alternative measures would do so.
+Because only one run is shown, there is also no information how stable the RMSE curves in Figures 8 and 13 are across replications.
+I encourage the authors to perform more systematic, replicated simulations, report measures of variability for their performance evaluations, compare against relevant alternatives, and take much more care not to overstate their conclusions beyond the tested scenarios (e.g.
+the claim that it “identify bad representations so they can be avoided” is based on a very limited set of examples).
+
 ## Minor Issues

### 2025-08-19 JCGS_response_to_reviews/response_to_reviews.qmd
Add the plan for writing the responses

@@
 ```
+
+We thank the anonymous associate editor and reviewers for their careful reviews. What follows is our point-by-point response to reviewer comments.
+
+Note that the original reviewer comments are in normal and *our response is in italic* text.
+
+<!-- Text in the paper corresponding to the changes as a response to reviewers' comments are highlighted in red.
+We have also made a lot of minor changes smoothing the language and fixing small issues in the paper which are not highlighted.
+-->
 
